### 2021-04-11 Supporting_old.tex
new files

\documentclass[draft,tec]{agutexSI2019}

\usepackage{graphicx}

\setkeys{Gin}{draft=false}

\begin{document}

\begin{article}

\section{Introduction}\label{sec:intro}

This supplement includes a description of the code, with a presentation of the features implemented followed by results of extensive benchmarks, implemented to verify that the code solves correctly a large range of problems (Table \ref{tab:list}).

The code is written in Fortran90 and uses the finite element method (FEM) with quadrilateral $Q_1P_0$ elements (continuous bilinear velocity and discontinuous constant pressure), associated with the direct solver MUMPS (see Section \ref{sec:solver}). Since $Q_1P_0$ elements do not satisfy Ladyzhenskaya, Babuska and Brezzi (LBB) stability condition \cite{Donea2003}, the elements are stabilised in postprocessing to avoid spurious pressures by means of a triple interpolation to smooth the pressure. Following this procedure, the elemental pressure is interpolated onto nodes, back onto elements and back again onto nodes (see Section \ref{sec:stokes}).

Elemental properties (density, viscosity, thermal conductivity, specific heat and thermal expansion) are related to the composition of each elemenet ,determined by means of Lagrangian markers. Each marker is characteristic of different materials in the domain. Therefore, elemental properties (with exception of the viscosity) is calculated using an arithmetic mean, as \[P_e=\frac{1}{n}\sum_{i=1}^n P_i\]
where $P_e$ is the elemental property, $P_i$ is the property characteristic of the material of each marker and $n$ is the number of markers of the element. Differently, the average scheme for the elemental viscosity can be chosen between harmonic, geometric and arithmetic mean. Markers can be placed regularly or randomly at the beginning of the simulation and their advection is performed by a 4th-order Runge-Kutta, interpolating the velocity field on each markers by means of the shape functions (see Section \ref{sec:runge}). During the simulation, each marker carries memory of temperature, pressure and accumulated strain, which are determined interpolating the nodal parameters, as for the velocities. The number of markers contained in each element is mantained between $n_{min}$ and $n_{max}$, which can be chosen by the user at the begininng of the simulation. When in an element there are less markers than $n_{min}$ the code adds random markers to reach the $n_{min}$, while if the number is higher than $n_{max}$ some of them are deleted. When new markers are added, they assume the properties of the nearest marker. In this way elements are never empty and mantain a number of markers inside a prefixed range.

The thermo-mechanics of crust-mantle systems can be described by means of the conservation of mass, momentum and energy equations, expressed as follows:
\begin{equation}\label{eq:mass}
\nabla \cdot \vec{u}=0
\end{equation}
\begin{equation}\label{eq:momentum}
-\nabla p + \nabla \cdot \vec{\tau} + \rho \vec{g} = 0
\end{equation}
\begin{equation}\label{eq:energy}
\rho C_p \left(\frac{\partial T}{\partial t} + \vec{u} \cdot \nabla T \right) = \nabla \cdot \left(K\nabla T\right) + H
\end{equation}

where $\vec{u}$ is the velocity, \textit{p} is the pressure, $\vec{\tau}$ is the deviatoric stress, $\rho$ is the density, $\vec{g}$ is the gravity acceleration, $C_p$ is the specific heat at constant pressure, \textit{T} is the temperature, \textit{K} is the thermal conductivity and \textit{H} is the total internal heat production. The deviatoric stress tensor can be write in terms of the strain rate tensor as $\vec{\tau}=2\mu\dot{\epsilon}$, with $\dot{\epsilon}=\frac{1}{2}\nabla\vec{v}$. Therefore, eq. \ref{eq:momentum} can be rewrite as:
\begin{equation}\label{eq:momentum_pressure}
-\nabla p + \nabla \cdot \left(\mu \nabla \vec{v}\right) + \rho \vec{g} = 0
\end{equation}
The penalty method has been implemented in order to enforce incompressibility, so that eq. \ref{eq:mass} can be write as:
\begin{equation}\label{eq:mass_penalty}
\nabla \cdot \vec{u} + \frac{p}{\lambda}=0
\end{equation}
where $\lambda$ is the so-called penalty parameter, which should be 6-7 orders of magnitude larger than the shear viscosity to ensure that mass conservation is satisfied \cite{Donea2003,Thieulot2014}. To support large viscosities variations the penalty parameter has been related to the effective elemental viscosity by means of a dimensionless coefficient, so that $\lambda=\lambda_e(e)\mu_{eff}(e)$ \cite{Marotta2006,Dabrowski2008,Thieulot2014}. Finally, using pressure from eq. \ref{eq:mass_penalty}, eq. \ref{eq:momentum_pressure} can be rewrite as:
\begin{equation}\label{eq:momentum_penalty}
\lambda \nabla \left(\nabla \cdot \vec{v} \right) + \nabla \cdot \left(\mu \nabla \vec{v}\right) + \rho \vec{g} = 0
\end{equation}
Results of benchmarks performed to verify the correctness in the immplementation of eqs. \ref{eq:momentum_penalty} and \ref{eq:energy} are shown in Section \ref{sec:momentum}. To strengthen the penalty method, it has been associated to the iterative Uzawa method, as described in detail in \citeA{Dabrowski2008} and \citeA{Thieulot2014} (see Section \ref{sec:poiseuille}). The stabilisation of the advection term of the energy equation (eq. \ref{eq:energy}) needed to avoid possible oscillations in the thermal solution in case where advection dominates over diffusion, has been implemented by means of a streamline-upwind Petrov–Galerkin (SUPG) method, for which the shape functions in the advection term are modified to have
\[(\textbf{N}^*)^T=\textbf{N}^T + \tau \vec{v} \textbf{B}\]
The choice of parameter $\tau$ follows the discussion in \citeA{Thieulot2011} and \citeA{Thieulot2014} (see Section \ref{sec:advection}).

\!!!!!!! Marcatori Lagrangiani e accennare a tutto  After the last interpolation, nodal pressure is used to determine pressure onto each Lagrangian marker.

\section{Solver}\label{sec:solver}
MUMPS (MUltifrontal Massively Parallel Solver) is a software package for solving systems of linear equations of the form $A \cdot x = b$, where \textit{A} is a square sparse matrix, by means of a direct method. Correctness of the solution and performances of the solver in terms of time and memory usage have been tested solving the Stokes flow with the analytical solution proposed by \citeA{Donea2003}.

\subsection{Stokes flow}\label{sec:stokes}
The problem consists of determining velocity field $(u,v)$ and pressure $p$ in case of prescribed body forces:
\[b_1=(12-24y)x^4+(-24+48y)x^3+(-48y+72y^2-48y^3+12)x^2+(-2+24y-72y^2+48y^3)x+1-4y+12y^2-8y^3\]
\[b_2=(8-48y+48y^2)x^3+(-12y+72y-72y^2)x^2+(4-24y+48y^2-48y^3+24y^4)x-12y^2+24y^3-12y^4\]
for which the exact solution is:
\[u(x,y)=x^2(1-x)^2(2y-6y^2+4y^3)\]
\[v(x,y)=-y^2(1-y)^2(2x-6x^2+4x^3)\]
\[p(x,y)=x(1-x)\]
The domain is a square with $L_x=L_y=1$ and a constant viscosity ($\mu=1$). Velocities are set to no slip conditions $(\vec{v}=0)$ on all boundaries. The problem has been performed for different grid resolution (between 8x8 and 1024x1024 elements).

Fig. \ref{fig:errors} shows that both velocity field and pressure converge to the exact solution with the decrease of the element size, following the theoretical convergence rate. The convergence of pressure, in contrast with the observation by \citeA{Donea2003} for $Q_1P_0$ elements, indicate the effectiveness of the smoothing procedure. Solve times and memory usage needed to generate the solution are shown in Fig. \ref{fig:MUMPS}.

\section{Momentum and energy equations}\label{sec:momentum}
Benchmarks in this Section show the accomplishment of the mass conservation equation with the Uzawa method (Poiseuille flow benchmark), the correctness of the solution of the momentum and energy equations (Instantaneous 2D Stokes sphere, Rayleigh-Taylor, mantle convection and thin-layer experiments), and the functionality of the stabilisation of the advection term in the energy equation.

\subsection{Poiseuille flow}\label{sec:poiseuille}
The domain is a rectangle with $L_x=2$ and $L_y=1$ and constant density and viscosity ($\rho=1$ and $\mu=1$), gravity acceleration $\vec{g}=0$ and penalty parameter $\lambda=10^8$. The grid is composed by 40×20 elements. Velocities are set to no slip conditions $(\vec{v}=0)$ at the top and the bottom, and a parabolic profile is imposed on the sides, with $u=y(1-y)$ and $v=0$.

The velocity field predicted by the model follows the expected parabolic profile (Fig. \ref{fig:poiseuille}a) and, in case of the classic penalty method with no iterations, the pressure is clearly related to the divergence by means of the penalty parameter (Fig. \ref{fig:poiseuille}b and c). Fig. \ref{fig:poiseuille}d shows that one iteration is sufficient to bring the divergence down to $10^{15}$, with no correlation with the pressure field.

\subsection{Instantaneous 2D sphere}\label{sec:ist_sphere}
The domain is a square with $L_x=L_y=1$ and gravity $g=(0,-1)$. The fluid has constant density and viscosity ($\rho_f=1$ and $\mu_f=1$). The sphere is in the middle of the domain with a radius R = 0.123456798 and has constant density and viscosity ($\rho_s=1.01$ and $\mu_s=10^3 \cdot \mu_f$).
Velocities are set in order to distinguish three cases:
\begin{enumerate}
\item FS: free slip conditions on all sides.
\item NS: no slip conditions on all sides.
\item OT: free slip conditions on the sides and the bottom, and open boundary at the top.
\end{enumerate}
All the three cases have been performed for different frid resolutions (between 16x16 and 512x512 elements) and with harmonic, geometric and arithmetic averages for the viscosity. The solutions generated in terms of the velocity ($v_{rms}$, $u_{min}$, $u_{max}$, $v_{min}$, $v_{max}$, $\vec{v}(0.5;0.5)$) and pressure ($p_{min}$, $p_{max}$, $p_{avg}$) fields have been compared with solutions generated by ASPECT.

\subsection{Rayleigh-Taylor instability}\label{sec:rayleigh}
This problem has been performed as the isoviscous case (Case 1a) originally presented by \citeA{vanKeken1997}. The domain has $L_x=0.9142$ and $L_y=1$ and gravity $g=(0,-1)$. Two fluids with same constant viscosities ($\mu_1=\mu_2=1$) and different densities ($\rho_1=1000$ and $\rho_2=1010$), with the buoyant fluid at the bottom. The interface between the fluids is given by $y(x)=0.2+0.02 cos \left(\frac{\pi x}{L_x}\right)$.

The experiment has been performed with different grid sizes (50x50, 80x80, 100x100, 256x256). Velocities are set to no slip conditions at the top and the bottom, and to free slip conditions at the sides of the domain. The root-mean-square velocity ($v_{rms}$) as function of time has been reported in \ref{fig:RT}, matching well with results shown by \citeA{vanKeken1997}, \citeA{Tackley2003} and \citeA{Thieulot2014}. \ref{fig:rayleigh} shows the evolution of the experiment at different time steps.

\subsection{Mantle convection}\label{sec:mantle}
This problem has been performed as the constant viscosity cases (Case 1a, 1b and 1c) presented by \citeA{Blankenbach1989} and \citeA{Thieulot2014}. Two fluids are separated in a square domain with $L_x=L_y=1$ and gravity acceleration $g=10^{10}Ra$. The experiment has been performed with three different Rayleigh numbers ($Ra=10^4$, $10^5$ and $10^6$) and with different grid resolution (between 32x32 and 128x128 elements). Both fluids have constant viscosity, initial density, heat capacity, thermal conductivity ($\mu= \rho_0= C_p = K=1$), reference temperature ($T_0=0$) and thermal expansion coefficient ($\alpha=10^{-10}$). Temperatures are set to 0 on top and 1 on bottom of the domain. Velocities are set to free slip conditions on all sides. The initial temperature field is given by \[T(x,y)=(1-y)-0.01 cos (\pi x) sin (\pi x)\]
Perturbation in the temperature field causes density perturbations, with \[\rho(T)=\rho_0(1-\alpha(T-T_0))\]

The solution generated by the code in terms of $v_{rms}$ and Nusselt number ($Nu$) as function of time are reported for all the simulations in Fig. \ref{fig:mantle}. At the steady state, $v_{rms}$ and $Nu$ of all simulations converge well toward the values from \citeA{Blankenbach1989}, with lower errors for higher resolution grids (Table \ref{tab:mantle}).

\subsection{Thin layer}\label{sec:thin}
This experiment has been originally proposed by \citeA{vanKeken1997}. Two fluids are separated in a rectangular domain with $L_x=2$ and $L_y=1$ and gravity acceleration $g=10^{10}Ra$. The Rayleigh number and the compositional Rayleigh number are fixed to $Ra=300000$ and $Rc=450000$, respectively. Both fluids have constant viscosity, thermal conductivity, specific heat ($\mu=\rho=C_p=1$) and thermal expansion coefficient ($\alpha=10^{-10}$). Fluid 1 has a density $\rho_1=1$, while fluid 2 is denser ($\rho_2=\rho_1+1.5 \cdot 10^{-10}$) and is placed at the bottom of the domain, for $y \leq 0.025$. Temperature are set to 0 on top and 1 on bottom of the domain. Velocities are set to free slip conditions on all sides of the domain. The initial temperature field is given by
\[T(x,y)=T_u(x,y)+T_l(x,y)+T_r(x,y)+T_s(x,y)-\frac{3}{2}\]
with
\[T_u(x,y)=\frac{1}{2}erf\left(\frac{1-y}{2}\sqrt{\frac{u_0}{x}}\right)\]
\[T_l(x,y)=1-\frac{1}{2}erf\left(\frac{y}{2}\sqrt{\frac{u_0}{L_x-x}}\right)\]
\[T_r(x,y)=\frac{1}{2}+\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{y+1}} exp\left(-\frac{x^2u_0}{4y+4}\right)\]
\[T_s(x,y)=\frac{1}{2}-\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{2-y}} exp\left(-\frac{(L_x-x)^2u_0}{8-4y}\right)\]
and
\[u_0=\frac{L_x^{7/3}}{(1+L_x^4)^{2/3}}\left(\frac{Ra}{2\sqrt{\pi}}\right)^{\frac{2}{3}}\]
\[Q=2\sqrt{\frac{L_x}{\pi u_0}}\]

The experiment has been performed with two grid resolution (125x40 and 200x80 elements), with 100 markers per element and Courant number $Cn=0.25$. $v_{rms}$ calculated for both the simulations match well the results from \citeA{vanKeken1997} and \citeA{Thieulot2014}, obtained with same grid resolutions (\ref{fig:thin}). In addition, a variety of simulations with different resolutions and aspect ratios have been performed to verify that the sensitivity of the initial velocity field agreed with the curve extrapolated by \citeA{Thieulot2014} (\ref{fig:thin_initial}).

\subsection{Advection stabilisation}\label{sec:advection}
The 1D advection problem proposed by \citeA{Thieulot2011} has been performed to verify the effectiveness of the SUPG method to stabilise the advection term of the energy equation. The domain is a 1D segment with $L_x=1$ composed by 50 elements and a discontinuity in the thermal field placed at $x=0.25$. Temperature is set to 1 for $x \leq 0.25$ and to 0 for $x > 0.25$. Velocity is set to $\vec{v}=1$ in the entire domain. The simulation has been performed for 250 time steps, with $dt=0.002$, so the thermal discontinuity should be at $x=0.75$ at the end of the simulation.

Temperature profiles at the end of the simulation are shown in \ref{fig:advection} as function of the dimensionless coefficient $\gamma=\tau \vec{v}/h$. In case of the classic Galerkin method ($\gamma=0$, blue line in \ref{fig:advection}) the final thermal profile is characterised by strong oscillations, which are strongly attenuated in case of the SUPG method ($\gamma=0.045$, orange line in \ref{fig:advection}).

\section{Markers advection}\label{sec:markers}

\bibliography{export}

\end{article}
\clearpage

\begin{figure}
\noindent\includegraphics[width=\textwidth]{./Figures/errors.pdf}
\caption{Velocity and pressure error between the generated and the analytical solution as a function of element size.}
\label{fig:errors}
\end{figure}

\begin{figure}
\noindent\includegraphics[width=\textwidth]{./Figures/MUMPS.pdf}
\caption{Panel a) Analysis, factorisation, solution and total solve times (panel a) and factorisation memory usage (panel b) as a function of the total number of degrees of freedom.}
\label{fig:MUMPS}
\end{figure}

\begin{figure}
\noindent\includegraphics[width=\textwidth]{./Figures/Poiseuille.pdf}
\caption{Velocity field (panel a), pressure (panel b) and divergence velocity in case of the classic penalty method (no iterations) and after one Uzawa iteration (panel c and d, respectively).}
\label{fig:poiseuille}
\end{figure}

\begin{figure}
\noindent\includegraphics[width=\textwidth]{./Figures/RT.pdf}
\caption{Root-mean-square velocity ($v_{rms}$) in function of time for different resolution of the grid.}
\label{fig:RT}
\end{figure}

\begin{figure}
\noindent\includegraphics[width=\textwidth]{./Figures/Rayleigh.pdf}
\caption{Evolution of the Rayleigh-Taylor experiment for a grid of 256x256 elements at $t=0, 500, 1000$ and $2000$ (panels a, b, c and d, respectively).}
\label{fig:rayleigh}
\end{figure}

\begin{figure}
\noindent\includegraphics[width=\textwidth]{./Figures/Convection.pdf}
\caption{$v_{rms}$ (panels a, c and e) and $Nu$ (panels b, d and f) as function of time for different grid resolution. Panels a and b show the results for $Ra=10^4$; Panels c and d show the results for $Ra=10^5$; panels e and f show the results for $Ra=10^6$. Purple lines indicate the convergence values for the $v_{rms}$ and the $Nu$ from \citeA{Blankenbach1989} at the steady state.}
\label{fig:mantle}
\end{figure}

\begin{figure}
\noindent\includegraphics[width=\textwidth]{./Figures/Thin_Layer.pdf}
\caption{$v_{rms}$ as function of time for different grid resolution.}
\label{fig:thin}
\end{figure}

\begin{figure}
\noindent\includegraphics[width=\textwidth]{./Figures/Thin_Layer_Initial.pdf}
\caption{$v_{rms}$ at $t=0$ as function of the element size for different aspect ratios of the elements.}
\label{fig:thin_initial}
\end{figure}

\begin{figure}
\noindent\includegraphics[width=\textwidth]{./Figures/Advection.pdf}
\caption{Temperature profile as function of \textit{x}. Purple line indicate the initial temperature profile; the green line inidicate the analytical temperature profile after 250 time steps; blue line indicate the temperature profile after 250 time steps in case of the classic Galerkin method ($\gamma=0$); orange line indicate the temperature profile after 250 time steps in case of the SUPG method ($\gamma=0.045$).}
\label{fig:advection}
\end{figure}

\begin{table}
\caption{List of benchmarks implemented for each feature of the code.}
\centering
\begin{tabular}{| l | l |}
\hline
 \textbf{Feature}  & \textbf{Benchmark}  \\
\hline
\hline
  Solver  & Stokes flow   \\
\hline
  Momentum and energy equations  & Poiseuille flow   \\
    & Instantaneous 2D sphere   \\
    & Rayleigh-Taylor experiment  \\
    & Mantle convection   \\
    & Thin layer   \\
    & Advection stabilisation   \\
\hline
  4th-order Runge-Kutta advection  & Zalesak disk   \\
\hline
  Large viscosity contrast  & Indenter   \\
    & Falling block   \\
\hline
  Non-linear visco-plasticity  & Indenter   \\
    & Brick   \\
    & Slab detachment   \\
\hline
  Shear and adiabatic heating  & ex. 9.4 in \citeA{Gerya2010b}   \\
\hline
  Free surface/sticky air  & Topography relaxation   \\
\hline
  Free surface stabilisation  & \citeA{Kaus2010a}, \citeA{Thieulot2014}   \\
\hline
  Markers chain  & Sinking ball   \\
  \hline
  Hydration  & Hydrated sinking ball   \\
\hline
\end{tabular}
\label{tab:list}
\end{table}

\begin{table}
\caption{Comparison between $v_{rms}$ and $Nu$ predicted by the code and same values reported in literature.}
\centering
\begin{tabular}{| l | l | l | l | l l l |}
\hline
  & & \citeA{Blankenbach1989} & \citeA{Thieulot2014} & 32x32 & 64x64 & 128x128  \\
\hline
\hline
 $Ra=10^4$ & $v_{rms}$ & $42.864947 \pm 0.000020$ & 42.867 & 42.83226 & 42.852793 & 42.861394  \\
           & $Nu$      & $4.884409 \pm 0.000010$  & 4.882  & 4.781297 & 4.857475  & 4.877573   \\
\hline
 $Ra=10^5$ & $v_{rms}$ & $193.21454 \pm 0.00010$  & 193.255& 193.872643 & 193.377472 & 193.252290  \\
           & $Nu$      & $10.534095 \pm 0.000010$ & 10.507 & 9.602514 & 10.270735  & 10.465629   \\
\hline
 $Ra=10^6$ & $v_{rms}$ & $833.98977 \pm 0.00020$  & 834.712& 848.091176 & 837.767911 & 834.945793  \\
           & $Nu$      & $21.972465 \pm 0.000020$ & 21.695 & 15.999266 & 19.703682  & 21.306939   \\
\hline
\end{tabular}
\label{tab:mantle}
\end{table}

%
%
% EXAMPLE FIGURES
% ---------------
% If you get an error about an unknown bounding box, try specifying the width and height of the figure with the natwidth and natheight options.
% \begin{figure}
%\setfigurenum{S1} %%You can change number for each figure if you want, not required. "S" prepended automatically.
% \noindent\includegraphics[natwidth=800px,natheight=600px]{samplefigure.eps}
%\caption{caption}
%\label{epsfiguresample}
%\end{figure}
%
%
% Giving latex a width will help it to scale the figure properly. A simple trick is to use \textwidth. Try this if large figures run off the side of the page.
% \begin{figure}
% \noindent\includegraphics[width=\textwidth]{anothersample.png}
%\caption{caption}
%\label{pngfiguresample}
%\end{figure}
%
%
%\begin{figure}
%\noindent\includegraphics[width=\textwidth]{athirdsample.pdf}
%\caption{A pdf test figure}
%\label{pdffiguresample}
%\end{figure}
%
% PDFLatex does not seem to be able to process EPS figures. You may want to try the epstopdf package.
%
%
% ---------------
% EXAMPLE TABLE
%
%\begin{table}
%\settablenum{S1} %%Change number for each table
%\caption{Time of the Transition Between Phase 1 and Phase 2\tablenotemark{a}}
%\centering
%\begin{tabular}{l c}
%\hline
% Run  & Time (min)  \\
%\hline
%  $l1$  & 260   \\
%  $l2$  & 300   \\
%  $l3$  & 340   \\
%  $h1$  & 270   \\
%  $h2$  & 250   \\
%  $h3$  & 380   \\
%  $r1$  & 370   \\
%  $r2$  & 390   \\
%\hline
%\end{tabular}
%\tablenotetext{a}{Footnote text here.}
%\end{table}
% ---------------
%
% EXAMPLE LARGE TABLE (UPLOADED SEPARATELY)
%\begin{table}
%\settablenum{S1} %%Change number for each table
%\caption{Time of the Transition Between Phase 1 and Phase 2\tablenotemark{a}}
%\end{table}

\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

More Information and Advice:

%% ------------------------------------------------------------------------ %%
%
%  SECTION HEADS
%
%% ------------------------------------------------------------------------ %%

% Capitalize the first letter of each word (except for
% prepositions, conjunctions, and articles that are
% three or fewer letters).

% AGU follows standard outline style; therefore, there cannot be a section 1 without
% a section 2, or a section 2.3.1 without a section 2.3.2.
% Please make sure your section numbers are balanced.
% ---------------
% Level 1 head
%
% Use the \section{} command to identify level 1 heads;
% type the appropriate head wording between the curly
% brackets, as shown below.
%
%An example:
%\section{Level 1 Head: Introduction}
%
% ---------------
% Level 2 head
%
% Use the \subsection{} command to identify level 2 heads.
%An example:
%\subsection{Level 2 Head}
%
% ---------------
% Level 3 head
%
% Use the \subsubsection{} command to identify level 3 heads
%An example:
%\subsubsection{Level 3 Head}
%
%---------------
% Level 4 head
%
% Use the \subsubsubsection{} command to identify level 3 heads
% An example:
%\subsubsubsection{Level 4 Head} An example.
%
%% ------------------------------------------------------------------------ %%
%
%  IN-TEXT LISTS
%
%% ------------------------------------------------------------------------ %%
%
% Do not use bulleted lists; enumerated lists are okay.
% \begin{enumerate}
% \item
% \item
% \item
% \end{enumerate}
%
%% ------------------------------------------------------------------------ %%
%
%  EQUATIONS
%
%% ------------------------------------------------------------------------ %%

% Single-line equations are centered.
% Equation arrays will appear left-aligned.

Math coded inside display math mode \[ ...\]
 will not be numbered, e.g.,:
 \[ x^2=y^2 + z^2\]

 Math coded inside \begin{equation} and \end{equation} will
 be automatically numbered, e.g.,:
 \begin{equation}
 x^2=y^2 + z^2
 \end{equation}

% IF YOU HAVE MULTI-LINE EQUATIONS, PLEASE
% BREAK THE EQUATIONS INTO TWO OR MORE LINES
% OF SINGLE COLUMN WIDTH (20 pc, 8.3 cm)
% using double backslashes (\\).

% To create multiline equations, use the
% \begin{eqnarray} and \end{eqnarray} environment
% as demonstrated below.
\begin{eqnarray}
  x_{1} & = & (x - x_{0}) \cos \Theta \nonumber \\
        && + (y - y_{0}) \sin \Theta  \nonumber \\
  y_{1} & = & -(x - x_{0}) \sin \Theta \nonumber \\
        && + (y - y_{0}) \cos \Theta.
\end{eqnarray}

%If you don't want an equation number, use the star form:
%\begin{eqnarray*}...\end{eqnarray*}

% Break each line at a sign of operation
% (+, -, etc.) if possible, with the sign of operation
% on the new line.

% Indent second and subsequent lines to align with
% the first character following the equal sign on the
% first line.

% Use an \hspace{} command to insert horizontal space
% into your equation if necessary. Place an appropriate
% unit of measure between the curly braces, e.g.
% \hspace{1in}; you may have to experiment to achieve
% the correct amount of space.

%% ------------------------------------------------------------------------ %%
%
%  EQUATION NUMBERING: COUNTER
%
%% ------------------------------------------------------------------------ %%

% You may change equation numbering by resetting
% the equation counter or by explicitly numbering
% an equation.

% To explicitly number an equation, type \eqnum{}
% (with the desired number between the brackets)
% after the \begin{equation} or \begin{eqnarray}
% command.  The \eqnum{} command will affect only
% the equation it appears with; LaTeX will number
% any equations appearing later in the manuscript
% according to the equation counter.
%

% If you have a multiline equation that needs only
% one equation number, use a \nonumber command in
% front of the double backslashes (\\) as shown in
% the multiline equation above.

%% ------------------------------------------------------------------------ %%
%
%  SIDEWAYS FIGURE AND TABLE EXAMPLES
%
%% ------------------------------------------------------------------------ %%
%
% For tables and figures, add \usepackage{rotating} to the paper and add the rotating.sty file to the folder.
% AGU prefers the use of {sidewaystable} over {landscapetable} as it causes fewer problems.
%
% \begin{sidewaysfigure}
% \includegraphics[width=20pc]{samplefigure.eps}
% \caption{caption here}
% \label{label_here}
% \end{sidewaysfigure}
%
%
%
% \begin{sidewaystable}
% \caption{}
% \begin{tabular}
% Table layout here.
% \end{tabular}
% \end{sidewaystable}
%
%
\grid

### 2021-04-11 Paper.tex
modified

\documentclass[hidelinks,11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage[cm]{fullpage}
\usepackage[round]{natbib}
\bibliographystyle{abbrvnat}
\usepackage{doi}
\usepackage{bm}
\usepackage{xcolor}
\hypersetup{
    colorlinks,
    linkcolor={blue},
    citecolor={blue},
    urlcolor={blue}
}
\usepackage{float}
\restylefloat{table}
\usepackage{siunitx}
\DeclareSIUnit\year{yr}

\begin{document}

\section{Introduction}\label{sec:intro}
PROVAThis supplement includes a description of the code, with a presentation of the features implemented followed by results of extensive benchmarks listed in Table \ref{tab:list}, implemented to verify that the code solves correctly a large range of problems. Description and results of the benchmarks are shown in Section \ref{sec:benchmark}.

\begin{table}[H]
\caption{List of benchmarks implemented for each feature of the code.}
\centering
\begin{tabular}{| l | l |}
\hline
 \textbf{Feature}  & \textbf{Benchmark}  \\
\hline
\hline
  Solver  & Stokes flow   \\
\hline
  4th-order Runge-Kutta  &  \\
  markers advection  & Zalesak disk   \\
\hline
  Momentum equation  & Poiseuille flow   \\
    & Instantaneous 2D sphere   \\
    & Rayleigh-Taylor experiment  \\
    & Falling block  \\
\hline
  Sticky air and  & 2D Stokes sphere \\
  free surface  & Stabilisation algorithm   \\
  & Topography relaxation   \\
  & Spontaneous subduction   \\
\hline
  Erosion and sedimentation  & -  \\
\hline
   Non-linear  & Slab detachment   \\
   visco-plasticity & Indenter   \\
    & Brick   \\
\hline
  Energy equation  & Advection stabilisation   \\
    & Simple shear heating \\
    & Shear and adiabatic heating \\
    & Mantle convection   \\
    & Thin layer entrainment   \\
\hline
  Phase changes  &  \\
  and hydration  & Hydrated sinking cylinder   \\
\hline
  Melting & Experimental melting curves \\
\hline
\end{tabular}
\label{tab:list}
\end{table}

The code is written in Fortran90 and uses the finite element method (FEM) with quadrilateral $Q_1 \times P_0$ elements (continuous bilinear velocity and discontinuous constant pressure), associated with MUMPS\footnote{\url{http://mumps.enseeiht.fr/}} \citep{Amestoy2001,Amestoy2019} that is a software package for solving systems of linear equations of the form $A \cdot x = b$, where \textit{A} is a square sparse matrix, by means of a direct method. Since $Q_1 \times P_0$ elements do not satisfy Ladyzhenskaya, Babuska and Brezzi (LBB) stability condition \citep{Donea2003}, elemental pressure is elaborated in postprocessing to avoid spurious pressures by means of a double interpolation to smooth the pressure. Following this procedure, the elemental pressure is interpolated onto nodes and then back onto elements. After the smoothing procedure, the pressure field is calculated again on the nodes to be used in combination with the lithostatic pressure, which is calculated onto nodes. Correctness of the solution and performances of the solver in terms of time and memory usage are tested solving the Stokes flow with the analytical solution proposed by \citet{Donea2003} (Section \ref{sec:stokes}).

The thermo-mechanics of crust-mantle systems is described by means of the conservation of mass, momentum and energy equations, expressed as follows:
\begin{eqnarray}
\label{eq:mass}\frac{\partial \rho}{\partial t}+\bm{\nabla} \cdot (\rho \bm{u})&=&0\\
\label{eq:momentum}-\bm{\nabla} p + \bm{\nabla} \cdot \bm{\tau} + \rho \bm{g}&=&0\\
\label{eq:energy}\rho C_p \left(\frac{\partial T}{\partial t} + \bm{u} \cdot \bm{\nabla} T \right)&=&\bm{\nabla} \cdot \left(k\bm{\nabla} T\right) + H_{tot}
\end{eqnarray}
where $\rho$ is the density, $\bm{u}$ is the velocity, \textit{p} is the pressure, $\bm{\tau}$ is the deviatoric stress, $\bm{g}$ is the gravity acceleration, $C_p$ is the specific heat at constant pressure, \textit{T} is the temperature, \textit{k} is the thermal conductivity and $H_{tot}$ is the total internal heat production. Density variations due to temperature are generally small enough to assume the density as constant ($\rho=\rho_0$) in Eq. \ref{eq:mass} and in Eq. \ref{eq:energy}, while it must be treated as a variable in the buoyancy term of Eq. \ref{eq:momentum}, such that \[\rho=\rho_0(1-\alpha(T-T_0))\]
where $\rho_0$ is the density at a reference temperature $T_0$ and $\alpha$ is the coefficient of thermal expansion. Eq. \ref{eq:mass} then can be rewritten as
\begin{equation}\label{eq:mass_0}
\bm{\nabla} \cdot \bm{u}=0
\end{equation}
This simplification is known as the extended Boussinesq approximation.

The time step $dt$ is chosen in according to Courant-Friedrichs-Lewy condition \citep{Anderson}, such that
\[dt=C_n\; \min\left(\frac{\Delta x}{\max|\bm{u}|},\frac{(\Delta x)^2}{\kappa}\right)\]
where $0<C_n<1$ is the Courant (CFL) number, $\Delta x$ is the minimum dimension of smallest element, $\max|\bm{u}|$ is the maximum velocity calculated on the entire domain and $\kappa=\frac{k}{\rho C_p}$ is the thermal diffusivity \citep{Thieulot2014}.

\section{Lagrangian markers}\label{sec:markers}
Elemental properties (density, viscosity, thermal conductivity, specific heat and thermal expansion) needed to solve Eqs. \ref{eq:mass_0}, \ref{eq:momentum} and \ref{eq:energy} are related to the composition of each element, determined by means of Lagrangian markers that are characteristic of different materials in the domain. Elemental properties (with the sole exception of the viscosity) are calculated using an arithmetic mean, as \[P_e=\frac{1}{n}\sum_{i=1}^n P_i\]
where $P_e$ is the elemental property, $P_i$ is the property characteristic of the material of each marker and $n$ is the number of markers of the element. Differently, the average scheme for the elemental viscosity can be chosen between harmonic, geometric and arithmetic mean. Markers can be placed regularly or randomly at the beginning of the simulation and their advection is performed by a 4th-order Runge-Kutta in space, interpolating the velocity field on each marker by means of the shape functions (see Section \ref{sec:runge}). The initial distribution of the markers is created using the open-source code library Geodynamic World Builder \citep{Fraters2019}. During the simulation, each marker carries memory of temperature, pressure and accumulated strain, which are determined interpolating the relative nodal parameter. In particular, pressure is interpolated onto each marker after the smoothing procedure. The number of markers contained in each element is maintained between $n_{min}$ and $n_{max}$. When in an element there are less markers than $n_{min}$ the code adds random markers to reach the $n_{min}$, while if the number is higher than $n_{max}$ some of them are randomly deleted. When new markers are added, they assume the properties of the nearest marker. In this way elements are never empty and maintain a number of markers inside a prefixed range.

\section{The momentum equation}\label{sec:mom_eq}
The penalty method is implemented so that Eq. \ref{eq:mass_0} can be written as:
\begin{equation}\label{eq:mass_penalty}
\bm{\nabla} \cdot \bm{u} + \frac{p}{\lambda}=0
\end{equation}
where $\lambda$ is the so-called penalty parameter, which should be 6-7 orders of magnitude larger than the shear viscosity to ensure that mass conservation is satisfied \citep{Donea2003,Thieulot2014}.

The deviatoric stress tensor in Eq. \ref{eq:momentum} can be written in terms of the strain rate tensor as $\bm{\tau}=2\eta\dot{\bm{\epsilon}}$, with $\dot{\bm{\epsilon}}=\frac{1}{2}\left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)$. Therefore, Eq. \ref{eq:momentum} can be rewritten as:
\begin{equation}\label{eq:momentum_pressure}
-\bm{\nabla} p + \bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right) + \rho \bm{g} = 0
\end{equation}
Finally, using pressure from Eq. \ref{eq:mass_penalty}, Eq. \ref{eq:momentum_pressure} can be rewritten as:
\begin{equation}\label{eq:momentum_penalty}
\lambda \bm{\nabla} \left(\bm{\nabla} \cdot \bm{v} \right) + \bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right) + \rho \bm{g} = 0
\end{equation}
The penalty method is been associated to the iterative Uzawa method, as described in detail in \citet{Dabrowski2008} and \citet{Thieulot2014} (see Section \ref{sec:poiseuille}). Results of benchmarks performed to verify the correctness in the implementation of Eq. \ref{eq:momentum_penalty} are shown in Sections \ref{sec:ist_sphere} and \ref{sec:rayleigh}.

To support large viscosities variations the penalty parameter is related to the effective elemental viscosity by means of a dimensionless coefficient, so that $\lambda=\lambda_e(e)\eta_{eff}(e)$ \citep{Marotta2006,Dabrowski2008,Thieulot2014}. Benchmark of the falling block \citep{Gerya2003a,Gerya2010b,Thieulot2011} is performed to verify that the code can correctly deal buoyancy driven flows with strong viscosity contrasts (Section \ref{sec:block}).

\section{Sticky air and free surface}\label{sec:sticky}
The Earth's surface can be treated by means of either the so-called sticky-air or a true free surface method, both of them implemented in the code. In the sticky-air method the surface is approximated with the introduction of a buoyant layer with a viscosity at least four orders of magnitude lower than the crust \citep{Schmeling2008,Crameri2012} and the interface between lithosphere and air is defined using a chain of passive markers that are advected as the Lagrangian markers. The correctness of the evolution of the markers chain is tested with the experiment of a 2D time-dependent Stokes sphere below a free surface and compared with results from ASPECT \citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} (Section \ref{sec:sphere}). In the true free surface case the top boundary is assumed stress-free and velocities are not fixed. In this case topography variations are described by vertical deformations of the mesh that depend on the velocity field of the nodes that identify the free surface, while horizontal deformations are not taken into account. This procedure is known as the Arbitrary Lagrangian-Eulerian (ALE) method and its implementation follows the technique described in \citet{Thieulot2011}. However, although the implementation of a true free surface better reproduces laboratory experiment, extremely small time steps can be necessary to maintain stability \citep{Kaus2010a,Quinquis2011,Thieulot2014}. Therefore, the stability algorithm proposed by \citet{Kaus2010a} is implemented to avoid instabilities due to high density differences at the free surface when using too large time steps. The implementation of this algorithm are tested by performing the experiment described by \citet{Kaus2010a} (Section \ref{sec:stab}).

The topography relaxation benchmark proposed by \citet{Crameri2012} is performed to verify that the code correctly recovers topography variations in case of both the sticky-air and the true free surface method (Section \ref{sec:crameri}). Finally, Section \ref{sec:subduction} show the results of the spontaneous subduction experiment described by \citet{Schmeling2008}.

\section{Erosion and sedimentation}\label{sec:erosion}
Surface processes at the lithosphere-air interface are implemented taking into account fluvial erosion (long-range processes), hillslope processes (short-range processes) and failure of slopes exceeding a threshold angle $\theta$ \citep{Burov1997,Tucker1999,Sacek2012,Croissant2014,Ueda2015,Beucher2020}. Erosion rates $\dot{E}_{lr}$ due to long-range processes can be described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, in according to
\begin{equation}
\dot{E}_{lr}=K_f A^{m_e} \bm{\nabla} h^{n_e}
\end{equation}
where $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $\bm{\nabla} h$ is the slope and $m_e$ and $n_e$ are parameters chosen in to constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}. Erosion rates $\dot{E}_{sr}$ due to short-range processes are implemented by assuming that transport is linearly proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}, in according to
\begin{equation}
\dot{E}_{sr}=K_d \bm{\nabla}^2 h
\end{equation}
where $K_d$ is the effective diffusivity. In addition, the slope is modified if it still exceeds a threshold angle after long- and short-range processes \citep{Ueda2015}.

Offshore sedimentation processes are implemented similarly to hillslope processes but with a depth-dependent coefficient \citep{Sacek2012,Beucher2020}. The sedimentation rate $\dot{S}$ is calculated as
\begin{equation}
\dot{S}=K_s \bm{\nabla}^2 h
\end{equation}
with
\begin{equation}
K_s=K^*_s \exp\left(\frac{h-h_{sl}}{h_0}\right)
\end{equation}
where $K^*_s$ is a constant, $h_{sl}$ is the sea level and $h_0$ is a characteristic length. In case of sticky air, slope is determined by means of the markers chain, which is vertically corrected considering erosion and sedimentation rates. After the correction, continental markers above the markers chain are transformed in air markers and air markers below the markers chain become sediments. Similarly, slope is determined considering the nodes on the top boundary in case of true free surface and they are vertically corrected as for the markers chain. In this case, continental markers above the top boundary are simply deleted.

\section{Non-linear rheologies}\label{sec:nonlinear}
Non-linear rheologies are implemented combining viscous creep (dislocation and diffusion) and plastic yielding. For each marker, diffusion and dislocation viscosity ($\eta_{df|ds}$) can be determined as follows \citep{Karato1993,Wang2016,Glerum2018}:
\begin{equation}\label{eq:viscous}
\eta_{df|ds}=\left(\frac{d^m}{A}\right)^{\frac{1}{n}}I_2^{\frac{(1-n)}{n}}\exp\left(\frac{Q+pV}{nRT}\right)
\end{equation}
where $d$ is the grain size, $m$ is the grain size exponent, $n$ is the stress exponent, $A$ is the pre-exponensional factor, $I_2$ is the square root of the second invariant of the strain rate tensor, $Q$ is the activation energy, $p$ is the pressure, $V$ is the activation volume, $R$ is the gas constant and $T$ is the temperature. Pressures and temperatures are determined by the interpolation of the nodal parameters, while strain rates are calculated by means of the derivatve of the velocities on the nodes. In case of diffusion creep $n=1$ and $m>0$, in case of dislocation creep $n>1$ and $m=0$. The implementation of the non-linear viscous creep viscosity depending on the strain rate is tested by means of the slab detachment benchmark (Section \ref{sec:slab}).

Plastic yielding is implemented rescaling $\eta_{vc}$ in order to limit the stress below the yield stress $\sigma_y$ \citep{Thieulot2008,Thieulot2014,Glerum2018}, obtaining
\begin{equation}\label{eq:plastic}
\eta_{pl}=\frac{\sigma_y}{2I_2}
\end{equation}
where $\eta_{pl}$ is the plastic viscosity and the yield stress is determined following the Drucker-Prager criterion, such as
\begin{equation}\label{eq:yield}
\sigma_y=C\;\cos(\phi)+p\;\sin(\phi)
\end{equation}
where $C$ is the cohesion and $\phi$ is the internal friction angle. In case of negative pressure it is imposed equal to 0, so that negative yield stress are excluded. The correctness of the non-linear solution in case of plastic viscosity with variable internal friction angle is verified performing indenter and brick experiments (Sections \ref{sec:indenter} and \ref{sec:brick}, respectively).

Strain softening is taken into account for both viscous creep and plastic viscosity \citep{Huismans2003,Babeyko2005,Huismans2005,Sobolev2005,Warren2008a} by means of the accumulated strain $\epsilon$ memorised by each marker. Softening in the viscous creep determines a linear decrease of $\eta_{vc}$ by means of a viscous strain softening factor $W_S$ that increases linearly from 1 to 10 for $5<\epsilon<10$ \citep{Warren2008a}. This viscous softening can be related to strain-induced grain size reduction \citep{Warren2008a}. Differently, plastic softening is simulated with a linear decrease of internal friction angle $\phi(\epsilon)$ and cohesion $C(\epsilon)$ in according to
\begin{eqnarray}
\label{eq:friction1}\phi(\epsilon)&=&\phi_0+(\phi_{\infty}-\phi_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}\\
\label{eq:friction2}C(\epsilon)&=&C_0+(C_{\infty}-C_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}
\end{eqnarray}
where $\phi_0$, $C_0$ and $\phi_{\infty}$, $C_{\infty}$ are internal friction angle and cohesion for $\epsilon=0.5$ and $\epsilon=1.5$, respectively \citep{Huismans2003,Huismans2005,Warren2008a,Thieulot2014}. Plastic softening approximates deformation-induced softening of faults and brittle shear zones \citep{Warren2008a}.

Viscous creep and plastic yielding are then combined to obtain a viscoplastic viscosity $\eta_{vp}$, which is computed with an harmonic average \citep{Tackley2010,Glerum2018} as follows
\begin{equation}\label{eq:viscoplastic}
\eta_{vp}=\left(\frac{1}{\eta_{df}}+\frac{1}{\eta_{ds}}+\frac{1}{\eta_{pl}}\right)^{-1}
\end{equation}
Finally, effective viscosity $\eta_{eff}$ is capped by the minimum and the maximum viscosity ($\eta_{min}$ and $\eta_{max}$, respectively) to avoid extremely low or high viscosity \citep{Glerum2018} as follows
\begin{equation}\label{eq:effective}
\eta_{eff}=\min(\max(\eta_{vp},\eta_{min}),\eta_{max})
\end{equation}
Viscous creep and plastic yielding are non-linear rheologies because of their dependence on the velocity field through pressure and strain rates. Therefore, the solution is determined by means of Picard-type iterations, until convergence of the velocity field \citep{Thieulot2011}. The correlation between two consecutive non-linear iterations is given by
\begin{equation}\label{eq:convergence}
\chi=1-\frac{\langle (\bm{v}^i-\langle\bm{v}^i\rangle)\cdot (\bm{v}^{i+1}-\langle\bm{v}^{i+1}\rangle) \rangle}{\sqrt{\sigma^i\sigma^{i+1}}}
\end{equation}
Iterations are performed until either convergence ($\chi<tol$) or maximal number of non-linear iterations ($it_{max}$) is reached.

\section{The energy equation}\label{sec:en_eq}
The stabilisation of the advection term of the energy equation (Eq. \ref{eq:energy}) needed to avoid possible oscillations in the thermal solution in the case when advection dominates over diffusion, is implemented by means of a streamline-upwind Petrov–Galerkin (SUPG) method, for which the advection term is modified follows the discussion in \citet{Thieulot2011} and \citet{Thieulot2014} (see Section \ref{sec:advection} for benchmark). The total internal heat production $H_{tot}$ in Eq. \ref{eq:energy} includes radiogenic heating $H_r$, shear heating $H_s$ and adiabatic heating $H_a$, where
\begin{equation}\label{eq:shear}
H_s=2\eta \dot{\bm{\epsilon}} : \dot{\bm{\epsilon}} = 2\eta(\dot{\epsilon}_{xx}^2+\dot{\epsilon}_{yy}^2+2\dot{\epsilon}_{xy}^2)
\end{equation}
and
\begin{equation}\label{eq:adiabatic}
H_a=T \alpha \frac{Dp}{Dt} \approx -\alpha T \rho g_y v
\end{equation}
The correctness of the implementation and the computation of shear and adiabatic heating is verified performing an exercise with an analytical solution (Section \ref{sec:simple_shear}) and Exercise 9.4 in \citet{Gerya2010b} (Section \ref{sec:shear}). Furthermore, two benchmarks are performed to verity that the code solve correctly Eqs. \ref{eq:momentum_penalty} and \ref{eq:energy} in case of a temperature field characterised by an initial perturbation and with a temperature-dependent density.

\section{Phase transitions and hydration}\label{sec:phase}
Variations of effective density, specific heat and coefficient of thermal expansion during the evolution of crustal and mantle materials at different pressure-temperature ($p$-$T$) conditions are computed by means of version 6.8.6 of Perple\_X software package \citep{Connolly2005}, similarly to the implementation described in \citet{Marotta2020}. By default, mineral assemblages and properties of each lithology (oceanic crust, lower and upper continental crust, sediments and dry or hydrated mantle) are calculated for temperatures from \SIrange{330}{1600}{\kelvin}, with increments of approximately \SI{4}{\kelvin}, and pressures from \SIrange{0.1}{30}{\giga\pascal}, with increments of approximately \SI{0.03}{\giga\pascal}, for a total of almost 100000 points.

The introduction of phase transitions produces two main effects that must be taken into account in numerical models: variations in density and release/absorption of latent heat \citep{Tackley2010}. Effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into account considering effective coefficient of thermal expansion \citep{Christensen1985,Zhong2015}. Similarly, effects of latent heat in the energy equation (eq \ref{eq:energy}) are taken into account by considering effective specific heat and coefficient of thermal expansion \citep{Christensen1985,Tackley2010,Zhong2015}. Effective density, specific heat and coefficient of thermal expansion are included in Perple\_X files and are assumed by each Lagrangian marker in according to its $p$-$T$ conditions.

Hydration processes related to dehydration of subducting lithosphere \citep{Schmidt1998,Liu2007,Faccenda2009,Faccenda2010,Faccenda2014,Rosas2016} strongly influence the thermo-mechanics inside the mantle wedge, mainly because of both weakening effects on the mantle rhology and density variations in case of mantle serpentinization \citep{Gerya2002,Honda2003,Arcay2005,Roda2010,Regorda2017}. In the case that hydration is switched on, the amount of bound and free water is memorised by each marker, following the implementation of \citet{Quinquis2014}; hydration and dehydration processes are related to the amount of bound water of each marker and to the maximum amount of water it can transport, i.e. if the amount of bound water exceeds the maximum amount of water, the marker dehydrates and releases free water that can hydrates undersaturated markers. Maximum water content of each marker is determined as function of lithology and $p$-$T$ conditions and it is calculated using Perple\_X, in the same way than effective density, specific heat and coefficient of thermal expansion. Bound water is advected together with the markers, neglecting the effect of bound water diffusion, while free water simply migrates vertically and is not coupled to the solid-phase flow of the mantle wedge \citep{Arcay2005,Quinquis2014}. The correctness of the migration scheme of the free water and its absorption into bound water by undersaturated markers are verified performing the experiment described in \citet{Quinquis2014} (Sec. \ref{sec:quinquis}).

The mantle viscosity weakening related to the amount of water has been extensively studied \citep{Chopra1981,Hirth2003} and it is implemented in according to \citet{Arcay2005} and \citet{Horiuchi2016}, as follows
\begin{equation}\label{eq:wet_visc}
\eta_{wet}=\eta_{dry}\left(\left[1-\frac{1}{f_v}\right]\exp\left(-\frac{[OH^-]}{[OH^-]_0}\right)+\frac{1}{f_v}\right)
\end{equation}
where $f_v$ is the viscosity reduction factor between dry and wet conditions, $[OH^-]$ is the water content and $[OH^-]_0$ is a reference water content set to 620 ppm (0.062 wt.\%) \citep{Arcay2005}. By default, $f_v$ is set to 100, in according to \citet{Arcay2005} and \citet{Horiuchi2016}.

\section{Mantle melting}\label{sec:melt}
Melt fraction in the mantle depends on temperature ($T$, in °C), pressure ($p$, in \si{\giga\pascal}) and the water content in the melt ($X_{H_2O}$, in wt.\%). The determination of the melt fraction $M$ is implemented as explained by \citet{Katz2003}, according to the successive modification by \citet{Langmuir2006} and \citet{Kelley2010}, as follows
\begin{eqnarray}\label{eq:melt_fr}
M(p,T,X_{H_2O})&=&-T+T_{sol}+(x_t\cdot ln(p+y_t))\cdot X_{H_2O}-\nonumber\\
&&+K_m\left(\frac{X_{H_2O}^{bulk}}{D_{H_2O}(1-X_{H_2O})+X_{H_2O}}\right)^{\gamma}
\end{eqnarray}
where $T_{sol}$ is the temperature od the dry solidus, $D_{H_2O}$ is the partition coefficient and $x_t\cdot log(p+y_t)$ indicates the pressure dependence of the melting curve \citep{Kelley2010}. $x_t$, $y_t$, $K_m$ and $\gamma$ are parameters chosen in according to \citet{Kelley2010}.
However, $X_{H_2O}$ depends on the melt fraction as
\begin{equation}\label{eq:melt_wt}
X_{H_2O}(M)=\frac{X_{H_2O}^{bulk}}{D_{H_2O}+M(1-D_{H_2O})}
\end{equation}
and a numerical solution can be found using a root-finder method \citep{Katz2003,Wang2016}. Furthermore, the content of water in the melt is limited by the pressure-dependent saturation concentration of water in the melt, determined as
\begin{equation}\label{eq:melt_max}
X_{H_2O}^{sat}=\chi_1 p^{\lambda_m}+\chi_2 p
\end{equation}
where $\chi_1$, $\chi_2$ and $\lambda_m$ are parameters chosen according to \citet{Katz2003}, \citet{Langmuir2006} and \citet{Kelley2010}. The correctness in the determination of melt fractions and water contents in the melt are tested comparing model results with experimental curves obtained by \citet{Katz2003} (Sec. \ref{sec:katz}).

In case of melting, viscous creep viscosity ($\eta_{vp}$) is modified according to \citet{Wang2016}, as
\begin{equation}\label{eq:melt_visc}
\eta_{melt}=\eta_{vp}\cdot \exp(\alpha_m M)
\end{equation}
where $\alpha_m$ are melt fraction factors for dislocation and diffusion creep, chosen in according to \citet{Wang2016}.

Effective density ($\rho_{eff}$) for partially molten rocks is calculated as
\begin{equation}\label{eq:melt_dens}
\rho_{eff}=\rho_s(1-M)+\rho_m M
\end{equation}
where $\rho_s$ and $\rho_m$ are the densities of the solid and the molten rock, respectively \citep{Gerya2003a,Wang2016}. The density of the solid rock is extracted by Perple\_X, while the density of the molten portion is calculated as
\begin{equation}\label{eq:melt_dens_M}
\rho_m(p,T)=\rho_0[1-\alpha(T-T_0)][1+\beta(p-p_0)]
\end{equation}
where $\rho_0$ is the density at the reference temperature ($T_0$) and pressure($p_0$), and $\alpha$ and $\beta$ are the thermal expansion and compressibility coefficients, respectively \citep{Gerya2003a,Wang2016}.

As for the phase transitions, effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into account considering effective coefficient of thermal expansion. Similarly, latent heat due to melting/crystallisation can be implicitily included in the energy equation (Eq. \ref{eq:energy}) by considering effective specific heat ($C_{p_{eff}}$) and thermal expansion ($\alpha_{eff}$) of the partially molten rocks, in according to
\begin{eqnarray}
\label{eq:cp_effective}C_{p_{eff}}&=&C_p+H_L\left(\frac{\partial M}{\partial T}\right)_p \\
\label{eq:alpha_effective}\alpha_{eff}&=&\alpha+\rho\frac{H_L}{T}\left(\frac{\partial M}{\partial p}\right)_T
\end{eqnarray}
where $C_p$ and $\alpha$ are the specific heat and the coefficient of thermal expansion of the solid rock, respectively, and $H_L$ is the latent heat of the molten rock \citep{Gerya2003a,Gerya2007,Tackley2010}.

\section{Benchmarks}\label{sec:benchmark}
\subsection{Error measurements}\label{sec:error}
In order to determine the accuracy of velocity and pressure field of the benchmarks, the $L_2$ norm is computed by numerical integration on the Gauss-Legendre quadrature points. $L_2$ norm for pressure and velocity errors can be evaluated as
\begin{eqnarray}
err_p&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|p^n_i(r_q)-p^a_i(r_q)|^2\right)w_q|J_q|}\nonumber \\
err_v&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|u^n_i(r_q)-u^a_i(r_q)|^2+|v^n_i(r_q)-v^a_i(r_q)|^2\right)w_q|J_q|}\nonumber
\end{eqnarray}
respectively, where $n_e$ is the number of elements, $n_q$ is the number of quadrature points per element, $p^n_i(r_q)$ and $p^a_i(r_q)$ are the numerical and analytical pressure, respectively, in each quadrature point $q$, $w_q$  and $J_q$ are the weight and the Jacobian at the quadrature point $q$, $u^n_i(r_q)$, $v^n_i(r_q)$, $u^a_i(r_q)$ and $v^a_i(r_q)$ are the numerical and analytical velocities, respectively, in each quadrature point $q$.

Other important quantities calculated in some benchmarks are the root-mean-square velocity and the Nusselt number, calculated as
\[v_{rms}=\sqrt{\frac{1}{V_{\Omega}}\int_{\Omega}|\bm{v}|^2 d\Omega}\]
and
\[Nu=\frac{\int_{top}\bm{\nabla} T \cdot \bm{n} ds}{\int_{bottom}T ds}\]
respectively.

\subsection{Stokes flow}\label{sec:stokes}
The problem consists of determining velocity field $(u,v)$ and pressure $p$ in case of a manufactured solution with prescribed body forces such as
\begin{eqnarray}
b_1&=&(12-24y)x^4+(-24+48y)x^3+(-48y+72y^2-48y^3+12)x^2+\nonumber \\
&&+(-2+24y-72y^2+48y^3)x+1-4y+12y^2-8y^3\nonumber \\
b_2&=&(8-48y+48y^2)x^3+(-12y+72y-72y^2)x^2+(4-24y+48y^2-48y^3+\nonumber \\
&&+24y^4)x-12y^2+24y^3-12y^4\nonumber
\end{eqnarray}
for which the exact solution is:
\begin{eqnarray}
u(x,y)&=&x^2(1-x)^2(2y-6y^2+4y^3)\nonumber \\
v(x,y)&=&-y^2(1-y)^2(2x-6x^2+4x^3)\nonumber \\
p(x,y)&=&x(1-x)-\frac{1}{6}\nonumber
\end{eqnarray}
The domain is a unit square a constant viscosity ($\eta=1$) and the penalty parameter is set to $\lambda=10^7$. Velocity boundary conditions are set to no slip $(\bm{v}=\bm{0})$ on all boundaries. The problem is performed for different grid resolution between \num{8x8} and \num{1024x1024} elements.

The errors between the analytical solution and the numerical prediction of the velocity and the pressure field are calculated as in Section \ref{sec:error}. Fig. \ref{fig:errors}a shows that both velocity and pressure field converge to the exact solution with the decrease of the element size, following the theoretical convergence rate. The convergence of pressure, in contrast with the observation by \citet{Donea2003} for $Q_1 \times P_0$ elements, indicate the effectiveness of the smoothing procedure. Smoothed pressure field for grid resolution of \num{128x128} elements is shown in Fig. \ref{fig:errors}b (continuous red line), in comparison with the analytical solution (dashed blue line). Solve times and memory usage needed to generate the solution are shown in Fig. \ref{fig:MUMPS}.

\begin{figure}
\centering
\noindent\includegraphics[width=\textwidth]{./Figures/errors.pdf}
\caption{Velocity and pressure error for the Stokes flow experiment between generated and analytical solution as a function of element size (panel a), and comparison between smoothed pressure and analytical solution as function of $x$ coordinate for a grid resolution of \num{128x128} elements (panel b).}
\label{fig:errors}
\end{figure}

\begin{figure}
\noindent\includegraphics[width=\textwidth]{./Figures/MUMPS.pdf}
\caption{Panel a) Analysis, factorisation, solution and total solve times (panel a) and factorisation memory usage (panel b) as a function of the total number of degrees of freedom for the Stokes flow experiment.}
\label{fig:MUMPS}
\end{figure}

\subsection{Markers advection}\label{sec:runge}
The 4th-order Runge-Kutta advection scheme in space is tested by means of the Zalesak disk test \citep{Zalesak1979,Thieulot2014}. The benchmark is performed in a unit square domain with different grid resolution (\num{100x100} and \num{200x200} elements) and 1000000 markers. The Courant number varies between 0.1 and 0.5 and the velocity field is prescribed in the entire domain as
\begin{eqnarray}
u(x,y)&=&2\pi\left(y-\frac{L_x}{2}\right)\nonumber \\
v(x,y)&=&-2\pi\left(x-\frac{L_x}{2}\right)\nonumber
\end{eqnarray}
At $t=0$, the disk is centered at position $(0.5;0.75)$ with a radius $R=0.15$ and has a vertical fissure 0.05 wide and 0.2 high (orange in Fig. \ref{fig:runge}). As we expect, markers are back to their initial location after a $2\pi$ rotation ($t=1$, red, green and blue in Fig. \ref{fig:runge}), with an error of the final position that increase moving away from the centre, more evident in case of higher time steps or lower grid resolution (Fig. \ref{fig:runge_err}).

\begin{figure}
\centering
\noindent\includegraphics[width=200px]{./Figures/Zalesak.pdf}
\caption{Comparison between the initial position of markers of the Zalesak disk experiment (orange) and their final position for different CFL in case of grid resolution of \num{100x100} elements. Blue, green and red markers indicate the final position of the disk in case of Courant number of 0.1, 0.25 and 0.5, respectively.}
\label{fig:runge}
\end{figure}

\begin{figure}
\centering
\noindent\includegraphics[width=400px]{./Figures/zal_error.pdf}
\caption{Error between final and initial position of markers of the disk as function of their distance from the centre of the domain. Results for different time steps are shown in blue, red and green for a grid resolution of \num{100x100} elements (for CFL of 0.1, 0.25 and 0.5, respectively) and in light blue, dark green and purple in case of \num{200x200} elements (for CFL of 0.1, 0.25 and 0.5, respectively).}
\label{fig:runge_err}
\end{figure}

\subsection{Poiseuille flow}\label{sec:poiseuille}
The domain is a rectangle with $L_x=2$ and $L_y=1$ and constant density and viscosity ($\rho=1$ and $\eta=1$), gravity acceleration $\bm{g}=\bm{0}$ and penalty parameter $\lambda=10^8$. The grid is composed by \num{40x20} elements. Velocity boundary conditions are set to no slip $(\bm{v}=\bm{0})$ at the top and the bottom, and a parabolic profile is imposed on the sides, with $u=y(1-y)$ and $v=0$. The analytical solution is then given by:
\begin{eqnarray}
u(x,y)&=&y(1-y)\nonumber \\
v(x,y,)&=&0\nonumber \\
p(x,y)&=&2\eta\left(\frac{L_x}{2}-x\right)\nonumber
\end{eqnarray}

The velocity field predicted by the model follows the expected parabolic profile (Fig. \ref{fig:poiseuille}a and Fig. \ref{fig:poi_plot}a) and, in case of the classic penalty method with no iterations, the pressure is clearly related to the divergence by means of the penalty parameter (Fig. \ref{fig:poiseuille}b and c and Fig. \ref{fig:poi_plot}b). Fig. \ref{fig:poiseuille}d shows that one Uzawa iteration is sufficient to bring the divergence down to \num{e-15}, with no correlation with the pressure field.

\begin{figure}
\noindent\includegraphics[width=\textwidth]{./Figures/Poiseuille.pdf}
\caption{Velocity field (panel a), pressure (panel b) and divergence velocity of a Poiseuille flow in case of the classic penalty method (no iterations) and after one Uzawa iteration (panel c and d, respectively).}
\label{fig:poiseuille}
\end{figure}

\begin{figure}
\noindent\includegraphics[width=\textwidth]{./Figures/Analytical.pdf}
\caption{Velocity (panel a) and pressure (panel b) field predicted by the model for a Poiseuille flow with respect with their analytical solutions. The velocity field is plotted as function of the vertical coordinate in $L_x/2$ and the pressure is plotted as function of the x coordinate in $L_y/2$.}
\label{fig:poi_plot}
\end{figure}

\subsection{Instantaneous 2D sphere}\label{sec:ist_sphere}
The domain is a unit square with gravity $\bm{g}=(0,-1)$. The fluid has constant density and viscosity ($\rho_f=1$ and $\eta_f=1$). The sphere is in the middle of the domain with a radius R = 0.123456798 and has constant density and viscosity ($\rho_s=10^{-2}+\rho_f$ and $\eta_s=10^3 \cdot \eta_f$).
We distinguish three different types of velocity boundary conditions:
\begin{enumerate}
\item FS: free slip conditions on all sides.
\item NS: no slip conditions on all sides.
\item OT: free slip conditions on the sides and the bottom, and open boundary at the top.
\end{enumerate}
All three cases are performed for different grid resolutions, between \num{16x16} and \num{512x512} elements with 50 randomly distributed markers per element, and different average schemes for the viscosity (harmonic, geometric and arithmetic). The velocity in the centre of the sphere $\bm{v}(0.5;0.5)$), minimum and maximum velocities ($u_{min}$, $u_{max}$, $v_{min}$, $v_{max}$) and pressures ($p_{min}$, $p_{max}$) on the entire domain, $v_{rms}$ and average pressure on the entire domain ($p_{avg}$) are compared with solutions generated by ASPECT \citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} and all results can be found at \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere2D}. Results in terms of the $v_{rms}$ for NS model are also shown in Fig. \ref{fig:inst_sphere}.

\begin{figure}
\centering
\noindent\includegraphics[width=400px]{./Figures/vrms_NS.pdf}
\caption{$v_{rms}$ as function of element size for the instantaneous 2D sphere experiment in case of no slip boundary conditions and with different average schemes for the viscosity. Results are compared with results obtained by other numerical codes.}
\label{fig:inst_sphere}
\end{figure}

\subsection{Rayleigh-Taylor instability}\label{sec:rayleigh}
This problem has been originally presented by \citet{vanKeken1997} and here it is performed as the isoviscous case (Case 1a). The domain has $L_x=0.9142$ and $L_y=1$ and gravity $\bm{g}=(0,-1)$. Two fluids with same constant viscosities ($\eta_1=\eta_2=1$) and different densities ($\rho_1=1000$ and $\rho_2=1010$), with the lighter fluid at the bottom. The initial interface between the fluids is given by $y(x)=0.2+0.02 \cos \left(\frac{\pi x}{L_x}\right)$. The experiment is performed with different grid sizes (\num{50x50}, \num{80x80}, \num{100x100} and \num{256x256}). A total of 1960000 markers are randomly distributed at the begininning of the simulation. Velocity boundary conditions are set to no slip at the top and the bottom, and to free slip at the sides of the domain.

$v_{rms}$ as function of time is reported in Fig. \ref{fig:RT}, matching well with results shown by \citet{vanKeken1997}, \citet{Tackley2003} and \citet{Thieulot2014}. Fig. \ref{fig:rayleigh} shows the evolution of the experiment at different time steps.

\begin{figure}
\centering
\noindent\includegraphics[width=400px]{./Figures/RT.pdf}
\caption{$v_{rms}$ of the Rayleigh-Taylor experiment as function of time for different resolution of the grid.}
\label{fig:RT}
\end{figure}

\begin{figure}
\centering
\noindent\includegraphics[width=300px]{./Figures/Rayleigh.pdf}
\caption{Evolution of the Rayleigh-Taylor experiment for a grid of \num{256x256} elements at $t=0, 500, 1000$ and $2000$ (panels a, b, c and d, respectively).}
\label{fig:rayleigh}
\end{figure}

\subsection{Falling block}\label{sec:block}
This benchmark is described as proposed by \citet{Gerya2003a}, \citet{Gerya2010b} and \citet{Thieulot2011}. The domain is square with $L_x=L_y=$ \SI{500}{\km} and the grid is composed by \num{50x50} elements with 25 markers in each element. The block is initially centered at ($x=$ \SI{250}{\km}; $y=$ \SI{400}{\km}) and has a size of \num{100x100} km (Fig. \ref{fig:block}a). The fluid surrounding the block has $\eta_f=$ \SI{e21}{\pascal\per\s} and $\rho_f=$ \SI{3200}{\kg\per\cubic\m}. The benchmark tests with different viscosities of the block, with $\eta_b$ from \SIrange{e15}{e27}{\pascal\per\s}. In each experiment also the density $\rho_b$ of the block varies from \SIrange{3220}{9900}{\kg\per\cubic\m}. Velocity boundary conditions are set to free slip conditions on all sides of the domain. Density distributions at $t=$ \SI{20}{\mega\year} for $\rho_b=$ \SI{3300}{\kg\per\cubic\m} and different $\eta_b$ are plotted in Fig. \ref{fig:block}b-f, showing that the code correctly preserve the block geometry in case of large viscosity contrast, with a block stiffer than the surrounding fluid (Fig. \ref{fig:block}f).

The velocity in the centre of the falling block at $t=0$ is measured for all experiments and, since the velocity should increase with density contrast, the quantity $\bm{v}/(\rho_b-\rho_f)$ is plotted as function of the viscosity contrast. All results perfectly match with those from \citet{Gerya2010b} and they line up on a single curve, demonstrating that the code can correctly deal with large viscosity and density contrasts (Fig. \ref{fig:falling}).

\begin{figure}
\noindent\includegraphics[width=300px]{./Figures/Block.pdf}
\centering
\caption{Density evolution of the falling block experiment at $t=$ \SI{0}{\mega\year} (panel a) and $t=$ \SI{20}{\mega\year} for different viscosities of the block (panels b-f).}
\label{fig:block}
\end{figure}

\begin{figure}
\centering
\noindent\includegraphics[width=400px]{./Figures/Falling.pdf}
\caption{Initial velocity relative to the density contrast at the centre of the falling block as function of the viscosity contrast between the block and the surrounding fluid.}
\label{fig:falling}
\end{figure}

\subsection{2D time-dependent Stokes sphere with deformable free surface}\label{sec:sphere}
This experiment is perfomed in a unit square domain with the gravity acceleration fixed to $g_y=-1$. The sphere has $\rho_s=2$ and $\eta_s=10^3$ and is initially centered at ($0.5;0.6$) with radius $R=0.123456789$. The fluid surrounding the sphere has $\rho_f=1$ and $\eta_f=1$ and occupies the domain for $y\leq 0.75$, while for $y>0.75$ the air has $\rho_a=0$ and $\eta_a=10^{-3}$. Velocity boundary conditions are set to free slip on all sides. The Courant number is set to 0.25. The experiments run for 200 s using grid resolutions from \num{150x150} to \num{512x512} elements, with 25 randomly distributed markers per element, and for different average schemes for the viscosity. The interface between the fluid and the air is tracked by means of the markers chain. Results in terms of velocity and pressure fieldd and topography variations are compared with results obtained with ASPECT \citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} and they can be found at \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere_fs2D}. Results in terms of the $v_{rms}$ in case of an arithmetic average are also shown in Fig. \ref{fig:inst_spherefs}.

\begin{figure}
\centering
\noindent\includegraphics[width=400px]{./Figures/vrms_arithm.pdf}
\caption{$v_{rms}$ for different grid resolutions as function of time for the 2D time-dependent Stokes sphere experiment in case of an arithmetic average. Results are compared with results obtained by ASPECT.}
\label{fig:inst_spherefs}
\end{figure}

\subsection{Free surface stabilisation}\label{sec:stab}
The benchmark is performed as discussed in \citet{Kaus2010a} and \citet{Thieulot2014}. The domain is square with $L_x=L_y=$ \SI{500}{\km} and a grid resolution of \num{200x200} elements, each of them containing 16 markers. A buoyant fluid with $\rho_1=$ \SI{3200}{\kg\per\cubic\metre} and $\eta_1=$ \SI{e20}{\pascal\s} is overlain by a denser fluid with $\rho_2=$ \SI{3300}{\kg\per\cubic\metre} and $\eta_2=$ \SI{e21}{\pascal\s}. The initial interface between the fluids has an initial sinusoidal shape of 5 km amplitude. Velocity boundary conditions are set to free slip conditions at the sides and no slip at the bottom, while the top is free surface. The experiment is performed with various fixed time steps, without the use of the Courant number. The vertical position of the free surface at $x=L_x$ is tracked for each simulation, with or without the stabilisation algorithm.

The results show an instability (drunken sailor effect) increasing the time step in case that the stabilisation algorithm is not activated ($dt=$ \SI{4200}{\year}, green line in Fig. \ref{fig:stab}), as already observed by \citet{Kaus2010a} and \citet{Thieulot2014}. Activating the algorithm the instability is fixed and the simulation remains stable using time step up to \SI{50000}{\year} (red line in Fig. \ref{fig:stab}).

\begin{figure}
\centering
\noindent\includegraphics[width=400px]{./Figures/Kaus_random.pdf}
\caption{Evolution of the y coordinate in $x=L_x$ as function of time for different time steps, with and without the stabilisation algorithm.}
\label{fig:stab}
\end{figure}

\subsection{Topography relaxation}\label{sec:crameri}
This experiment is performed as in \citet{Crameri2012} using free surface and sticky air (Case 1). The domain is rectangular with $L_x=$ \SI{2800}{\km} and $L_y=$ \SI{700}{\km} in case of free surface and $L_y=$ \SI{800}{\km} in case of sticky air ($h_{st}=$ \SI{100}{\km}). The grid is composed by \num{256x64} elements for the free surface \num{560x320} elements for the sticky air. For both cases, there is a mantle of 600 km thickness with $\eta_m=$ \SI{e21}{\pascal\s} overlain by a highly viscous cosine-shaped lithosphere with thickness between 93 and 107 km and $\eta_l=$ \SI{e23}{\pascal\s}. Both mediums have $\rho=$ \SI{3300}{\kg\per\cubic\metre}. In case of sticky air experiment, the air has $\rho_a=$ \SI{0}{\kg\per\cubic\metre} and $\eta_a=$ \SI{e18}{\pascal\s}. The gravity acceleration is set to $g_y=$ \SI{-10}{\metre\per\square\s}. Velocity boundary conditions are set to no slip at the bottom and free slip at the sides of the domain. In case of sticky air, velocities on top are set to free slip conditions.

The maximum topography as function of time ($h(t)$) can be analytically derived in according to \[h(t)=h_0 \exp(\gamma t)\]
where $h_0=7$ km is the initial topography, $\gamma=\num{-0.2139e-11}$ is the characteristic relaxation rate and $t$ is the time (black line in Fig. \ref{fig:crameri}). Maximum topography at the characteristic relaxation time $t_{rlx}=$ \SI{14.825}{\kilo\year} can be found to be $h_{rlx}=$ \SI{2576}{\m}. The results show that a viscosity of the air of \SI{e18}{\pascal\s} is needed in case of $h_{st}=$ \SI{100}{\km} to correctly describe the topography relaxation for this problem, as pointed out by \citet{Crameri2012} (purple line in Fig. \ref{fig:crameri}). The case with a true free surface well follows the analytical solution (orange line in Fig. \ref{fig:crameri}). In particular, at the relaxation time in case of the true free surface predicts a maximum topography of \SI{2570}{\m}, with an error of \SI{6}{\m} with respect to the analytical solution. Results are also compared with results obtained with a free surface in UNDERWORLD (\num{256x64} elements) and SULEC (\num{401x201} elements) (with courtesy from \citet{Crameri2012}).

\begin{figure}
\centering
\noindent\includegraphics[width=400px]{./Figures/Crameri.pdf}
\caption{Maximum topography as function of time for the Crameri benchmark (coloured lines) in comparison with the analytical solution (black line) and results obtained with UNDERWORLD (yellow line) and SULEC (blue line).}
\label{fig:crameri}
\end{figure}

\subsection{Spontaneous subduction}\label{sec:subduction}
This experiment is performed as presented by \citet{Schmeling2008}, in case of both a sticky air and a true free surface and using different average schemes for the viscosities. The domain is rectangular with $L_x=$ \SI{3000}{\km} and $L_y=$ \SI{700}{\km}, a grid resolution of \num{256x64} elements with 50 markers per element and Courant number of 0.01. At the beginning of the simulation, a 100 km-thick lithosperic layer with $\rho_m=$ \SI{3300}{\kg\per\cubic\metre} and $\eta_m=$ \SI{e23}{\pascal\s} is located at the top of the domain for $x$ between 1000 and 3000 km, while the asthenospheric mantle has $\rho_m=$ \SI{3200}{\pascal\s} and $\eta_m=$ \SI{e21}{\pascal\s}. In addition, a 200 km-depth lithosperic slab is already subducted in the mantle in order to have a spontaneous subduction. In case of sticky air, $L_y=750$ km and a 50 km-thick air layer with $\eta_a$ from \SIrange{e19}{e21}{\pascal\s} overlie the mantle. Velocity boundary conditions are set to free slip on the sides and on the bottom of the domain. In case of sticky air, velocities on the top boundary are set to free slip conditions as well.

The maximum depth of the slab is tracked for all the simulations (Fig. \ref{fig:subduction}), showing a good match with results from \citet{Schmeling2008}. Low viscous sticky air models (i.e., \SI{e19}{\pascal\s}) enlighten a strong dependence of the sinking velocity with the chosen average scheme (dashed lines in Fig. \ref{fig:subduction}), while high viscous sticky air models (i.e., \SI{e21}{\pascal\s}) show a higher resistance at the trench, underestimating the correct solution (dotted lines in Fig. \ref{fig:subduction}). A higher resistance can be also observed in case of a true free surface for the low resolution grid used (continuous lines in Fig. \ref{fig:subduction}), as pointed out by \citet{Schmeling2008}.

\begin{figure}
\centering
\noindent\includegraphics[width=400px]{./Figures/Subduction.pdf}
\caption{Maximum depth of the slab as function of time for all the simulations of the \citet{Schmeling2008} benchmark, in case of harmonic, geometric and arithmetic means (red, green and blue lines, respectively) using a sticky air or a true free surface (dashed and continuous lines, respectively).}
\label{fig:subduction}
\end{figure}

\subsection{The slab detachment}\label{sec:slab}
The slab detachment benchmark is performed as described by \citet{Schmalholz2011} and \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{1000}{\km} and $L_y=$ \SI{660}{\km} and a grid resolution of \num{256x256} elements. A non-linear viscous T-shaped layer with $\rho_l=$ \SI{3300}{\kg\per\cubic\metre} is placed at the top of the domain and surrounded by a linear viscous fluid with $\rho_f=$ \SI{3150}{\kg\per\cubic\m}. The top layer is 80 km-thick and a 250 km-long and 80 km-wide slab is placed at $x=L_x/2$ (Fig. \ref{fig:slab}a). The effective viscosity of the top layer is given by
\[\eta_{eff}=\eta_0 I_2^{\frac{(1-n)}{n}}\]
with $\eta_0=$ \SI{4.75e11}{\pascal\s} and $n=4$, while the surrounding fluid has $\eta_f=$ \SI{e21}{\pascal\s}. The effective viscosity is capped between $\eta_{min}=$ \SI{e21}{\pascal\s} and $\eta_{max}=$ \SI{e25}{\pascal\s}. Velocity bounadry conditions are set to free slip at the top and the bottom, and to no slip at the sides of the domain.

The sinking of the slab determines a decrease of the effective viscosity in the T-shaped layer during the first part of the evolution (Fig. \ref{fig:slab}a), as consequence of local strain rates, while effective viscosities increase up to $\eta_{max}$ in the top layer after approximately \SI{20}{\mega\year}, when necking is complete (Fig. \ref{fig:slab}b).

The width of necking is tracked by means of markers position at the side of the slab. Necking width and time are normalised in function of the initial slab (80 km) and the characteristic time ($t_c=$ \SI{7.1158e14}{\s}) \citep{Schmalholz2011,Glerum2018}. The results in case of arithmetic, geometric and harmonic mean of the viscosity are shown in Fig. \ref{fig:necking}, compared with results from \citet{Glerum2018} for which an infinity norm mean was used. The influence of the average schemes agrees with results shown in Section \ref{sec:subduction}.

\begin{figure}
\noindent\includegraphics[width=\textwidth]{./Figures/Slab.pdf}
\caption{Effective viscosity fot the slab detachment benchmark at the beginning of the evolution (panel a) and when necking is complete (panel b).}
\label{fig:slab}
\end{figure}

\begin{figure}
\centering
\noindent\includegraphics[width=400px]{./Figures/Necking.pdf}
\caption{Normalised width of the necking of the slab detachment benchmark as function of normalised time.}
\label{fig:necking}
\end{figure}

\subsection{The indenter experiment}\label{sec:indenter}
The indenter benchmark simulate a rigid punch on a purely plastic von Mises material, for which exists an analytical solution \citep{Thieulot2008,Thieulot2014,Glerum2018}. The domain is a unit square with a grid resolution of \num{256x256} elements. The gravity acceleration is set to $\bm{g}=\bm{0}$ and effective viscosity is capped using $\eta_{min}=10^{-4}$ and $\eta_{max}=10^3$. The tolerance for the convergence of the non-linear solution is $tol=10^{-9}$, with a maximum number of non-linear iterations set to 500. The medium in the domain has $\rho=0.01$, an initial viscosity $\eta_0=10$, cohesion $C=1$ and an angle of internal friction $\phi=0$°. Velocity boundary conditions are set to no slip at the bottom and free slip at the sides of the domain. The top of the domain is open with the exception of the central portion (punch area) with width $w_p=0.125$, where $v=-1.05$ and $u$ is fixed either to 0 (rough punch experiment) or free (smooth punch experiment).

Fig. \ref{fig:indenter} show the results in terms of viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure (panels g and h) for both experiments. The analytical solution indicate that pressure at ($0.5;1$) and ($0.5 \pm w_p;1$)is $p=\pi +1$ and $p=1$, respectively (black and red lines in Fig. \ref{fig:smooth_p}). Fig. \ref{fig:smooth_p} shows a clear improvement in the pressure solution when passing from a rough to a smooth punch (green and blue lines in Fig. \ref{fig:smooth_p}, respectively), as previously observed by \citet{Thieulot2014} and \citet{Glerum2018}.

\begin{figure}
\centering
\noindent\includegraphics[width=400px]{./Figures/Indenter.pdf}
\caption{Viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure (panels g and h) fields for rough (left column) and smooth (right column) punch experiments.}
\label{fig:indenter}
\end{figure}

\begin{figure}
\centering
\noindent\includegraphics[width=400px]{./Figures/Smooth_Rough.pdf}
\caption{Nodal pressure in function of the x coordinate at $y=1$ for rough (green line) and smooth (blue line) punch experiment. Black and red lines indicate the analytical solution at $x=0.5$ and $x=0.5 \pm w_p$, respectively.}
\label{fig:smooth_p}
\end{figure}

\subsection{The brick experiment}\label{sec:brick}
An instantaneous version of the brick benchmark is performed to verify the correctness of the pressure-dependent plasticity for different angle of internal friction, as proposed by \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{40}{\km} and $L_y=$ \SI{10}{\km} and a grid resolution of \num{512x128} elements. The gravity acceleration is set to $g_y=$ \SI{-10}{\m\square\s} and effective viscosity is capped using $\eta_{min}=$ \SI{e19}{\pascal\s} and $\eta_{max}=$ \SI{e26}{\pascal\s}. The tolerance for the convergence of the non-linear solution is $tol=10^{-7}$, with a maximum number of non-linear iterations set to 1000. A 800 m-wide and 400 m-height inclusion with $\eta_b=$ \SI{e20}{\pascal\s} and $\rho_b=$ \SI{2700}{\kg\per\cubic\m} is placed at the bottom of the domain at $x=L_x/2$. The inclusion is surrounded by a non-linear viscous medium with $\rho_m=$ \SI{2700}{\kg\per\cubic\m}, an initial viscosity of \SI{e23}{\pascal\s}, a linear viscous viscosity of \SI{e25}{\pascal\s} and a cohesion of \SI{40}{\mega\pascal}. Velocities are set to free slip conditions at the bottom of the domain and the top is open. Velocities on sides of the domain are fixed to $u=$ \SI{+-2e-11}{\m\per\s} and $v=0$. The experiment is performed in compressional and extensional contexts with an angle of internal friction $\phi$ of the non-linear medium variable between 0° and 20°.

As expected, two shear bands stem from the inclusion with variable angles in relation to both the dynamics context and the internal friction angle. Shear band angles formed at 45° for $\phi=0$° in both extensional and compressional contexts (Fig. \ref{fig:brick_beam}a), while they are different in case of $\phi\neq0$°. Shear band angles in case of $\phi=20$° are shown for compression and extension in Fig. \ref{fig:brick_beam}b  and c, respectively. Results for internal friction angles from 0° to 20° are plotted in Fig. \ref{fig:brick}, in comparison with theoretical Roscoe, Arthur and Coulomb shear band angles.

\begin{figure}
\centering
\noindent\includegraphics[width=400px]{./Figures/Brick_Beam.pdf}
\caption{Shear band angles predicted for the brick experiment for $\phi=0$° (panel a) and for $\phi=20$° in case of compression and extension (panels b and c, respectively).}
\label{fig:brick_beam}
\end{figure}

\begin{figure}
\centering
\noindent\includegraphics[width=400px]{./Figures/Brick.pdf}
\caption{Shear band angles predicted for the brick experiment in case of compressional and extensional contexts as function of different internal angle of friction (purple line and dots), compared with theoretical Roscoe, Artur and Coulomb angles (orange, blue and green lines, respectively).}
\label{fig:brick}
\end{figure}

\subsection{Advection stabilisation}\label{sec:advection}
The 1D advection problem proposed by \citet{Donea2003} and \citet{Thieulot2011} is performed to verify the effectiveness of the SUPG method to stabilise the advection term of the energy equation. The domain is a 1D segment with $L_x=1$ composed by 50 elements and a discontinuity in the thermal field placed at $x=0.25$. Temperature is set to 1 for $x \leq 0.25$ and to 0 for $x > 0.25$. Velocity is set to ${u}=1$ in the entire domain. The simulation is performed for 250 time steps, with $dt=0.002$, so the thermal discontinuity should be at $x=0.75$ at the end of the simulation.

Temperature profiles at the end of the simulation are shown in \ref{fig:advection} as function of the dimensionless coefficient $\gamma=\tau \bm{v}/h$. In case of the classic Galerkin method ($\gamma=0$, blue line in \ref{fig:advection}) the final thermal profile is characterised by strong oscillations, which are eliminated in case of the SUPG method ($\gamma=0.045$, orange line in \ref{fig:advection}).

\begin{figure}
\centering
\noindent\includegraphics[width=400px]{./Figures/Advection.pdf}
\caption{Temperature profile as function of \textit{x} for the advection stabilisation benchmark. Purple line indicates the initial temperature profile; the green line inidicates the analytical temperature profile after 250 time steps; blue line indicates the temperature profile after 250 time steps in case of the classic Galerkin method ($\gamma=0$); orange line indicates the temperature profile after 250 time steps in case of the SUPG method ($\gamma=0.045$).}
\label{fig:advection}
\end{figure}

\subsection{Shear and adiabatic heating}\label{sec:simple_shear}
This exercise is performed in an unit square domain composed by \num{128x128} elements. The velocity field is prescribed on the entire domain with $\bm{v}=(L_y-y)y\bm{e_x}$; viscosity, density and specific heat are fixed to 1, while thermal conductivity and radiogenic energy are fixed to 0. Therefore, the energy equation (Eq. \ref{eq:energy}) can be simplified as \[\frac{\partial T}{\partial t}=H_s\] and fixing $T(t=0)=0$, the temperature can be find as \[T(t)=H_s t\]
In this case we have
\begin{eqnarray}
\dot{\epsilon_{xx}}&=&\frac{\partial u}{\partial x}=0 \nonumber \\
\dot{\epsilon_{yy}}&=&\frac{\partial v}{\partial y}=0 \nonumber \\
\dot{\epsilon_{xy}}&=&\frac{1}{2}\left(\frac{\partial v}{\partial x}+\frac{\partial u}{\partial y}\right)=\frac{1}{2}(L_y-2y)\nonumber
\end{eqnarray}
and, simplifying Eq. \ref{eq:shear}, shear heating can be calculated as \[H_s(x,y)=(1-2y)^2\]

The solution predicted by the model in terms of velocity, temperature and shear heating match well with the analytical solutions (Fig. \ref{fig:analytical_en}).

%\begin{figure}
%\centering
%\noindent\includegraphics[width=400px]{./Figures/Energy_Thieulot.pdf}
%\caption{Temperature and shear heating predicted by the model after 1 time step.}
%\label{fig:en_thieulot}
%\end{figure}

\begin{figure}
\centering
\noindent\includegraphics[width=500px]{./Figures/Analytical_Shear.pdf}
\caption{Velocity (panel a), temperature (panel b) and shear heating (panel c) as function of the \textit{y} coordinate for the simple shear experiment. The solutions predicted by the model (red lines) are compared with the analytical solutions (dashed blue lines).}
\label{fig:analytical_en}
\end{figure}

\subsection{Shear and adiabatic heating}\label{sec:shear}
This problem is performed as presented in Exercise 9.4 in \citet{Gerya2010b}. Two materials are vertically separated in a rectangular domain with $L_x=$ \SI{1000}{\km}, $L_y=$ \SI{1500}{\km} and a grid resolution of \num{30x20} elements. Constant thermal coefficient expansion ($\alpha=$ \SI{3e-5}{\per\kelvin}), temperature ($T=$ \SI{1300}{\kelvin}) and gravity acceleration ($g_y=$ \SI{-10}{\m\square\s}) are assumed in the whole domain. Fluid 1 (on the left side of the domain) has $\rho_1=$ \SI{3200}{\kg\per\cubic\m} and $\eta_1=$ \SI{e20}{\pascal\s}; fluid 2 (on the right side of the domain) has $\rho_2=$ \SI{3300}{\kg\per\cubic\m} and $\eta_2=$ \SI{e22}{\pascal\s}. Velocity boundary conditions are set to free slip on all sides of the domain.

As shown in Fig. \ref{fig:energy}, both shear and adiabatic heating precited by the code (first row in Fig. \ref{fig:energy}) well recreate the results obtained by means of example ${Shear\_adiabatic\_heating.m}$ from \citet{Gerya2010b} (second row in Fig. \ref{fig:energy}).

\begin{figure}
\centering
\noindent\includegraphics[width=400px]{./Figures/Energy.pdf}
\caption{Comparison between shear (first column), adiabatic (second column) and total (third column) energy predicted by the code (panels a, b and c) and those created using example ${Shear\_adiabatic\_heating.m}$ from Exercise 9.4 in \citet{Gerya2010b} (panels d, e and f).}
\label{fig:energy}
\end{figure}

\subsection{Mantle convection}\label{sec:mantle}
This problem is performed as presented by \citet{Blankenbach1989} (constant viscosity cases) and \citet{Thieulot2014}, in a 2D unit square domain with gravity acceleration $g_y=10^{10}Ra$. The experiment is performed with three different Rayleigh numbers ($Ra=10^4$, $10^5$ and $10^6$) and with different grid resolution (between \num{32x32} and \num{128x128} elements). Both fluids have constant viscosity, initial density, heat capacity, thermal conductivity ($\eta= \rho_0= C_p = k=1$), reference temperature ($T_0=0$) and thermal expansion coefficient ($\alpha=10^{-10}$). Temperatures are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by \[T(x,y)=(1-y)-0.01 \cos (\pi x) \sin (\pi x)\]

The solution generated by the code in terms of $v_{rms}$ (calculated as in Sec. \ref{sec:error}) and $Nu$ as function of time are reported for all the simulations in Fig. \ref{fig:mantle}. At the steady state, $v_{rms}$ and $Nu$ of all simulations converge well toward the values from \citet{Blankenbach1989}, with lower errors for higher resolution grids (Table \ref{tab:mantle}).

\begin{figure}
\centering
\noindent\includegraphics[width=400px]{./Figures/Convection.pdf}
\caption{$v_{rms}$ (panels a, c and e) and $Nu$ (panels b, d and f) for the mantle convection benchmark as function of time for different grid resolution. Panels a and b show the results for $Ra=10^4$; Panels c and d show the results for $Ra=10^5$; panels e and f show the results for $Ra=10^6$. Purple lines indicate the convergence values for the $v_{rms}$ and the $Nu$ from \citet{Blankenbach1989} at the steady state.}
\label{fig:mantle}
\end{figure}

\begin{table}[H]
\caption{Comparison between $v_{rms}$ and $Nu$ predicted by the code for the mantle convection experiment and same values as reported in literature.}
\centering
\small
\begin{tabular}{|p{0.09\textwidth}|p{0.05\textwidth}|p{0.24\textwidth}|p{0.15\textwidth}|p{0.1\textwidth}p{0.1\textwidth}p{0.1\textwidth}|}
\hline
  & & \citet{Blankenbach1989} & \citet{Thieulot2014} & & &   \\
 & & & \num{200x200} & \num{32x32} & \num{64x64} & \num{128x128}  \\
\hline
\hline
 $Ra=10^4$ & $v_{rms}$ & $42.864947 \pm 0.000020$ & 42.867 & 42.83226 & 42.852793 & 42.861394  \\
           & $Nu$      & $4.884409 \pm 0.000010$  & 4.882  & 4.781297 & 4.857475  & 4.877573   \\
\hline
 $Ra=10^5$ & $v_{rms}$ & $193.21454 \pm 0.00010$  & 193.255& 193.872643 & 193.377472 & 193.252290  \\
           & $Nu$      & $10.534095 \pm 0.000010$ & 10.507 & 9.602514 & 10.270735  & 10.465629   \\
\hline
 $Ra=10^6$ & $v_{rms}$ & $833.98977 \pm 0.00020$  & 834.712& 848.091176 & 837.767911 & 834.945793  \\
           & $Nu$      & $21.972465 \pm 0.000020$ & 21.695 & 15.999266 & 19.703682  & 21.306939   \\
\hline
\end{tabular}
\label{tab:mantle}
\end{table}

\subsection{Thin layer entrainment}\label{sec:thin}
This experiment is performed as originally proposed by \citet{vanKeken1997} with two incompressible fluids in a rectangular domain with $L_x=2$ and $L_y=1$ and gravity acceleration $g_y=10^{10}Ra$. The Rayleigh number and the compositional Rayleigh number are fixed to $Ra=300000$ and $Rc=450000$, respectively. Both fluids have constant viscosity, thermal conductivity, specific heat ($\eta=\rho=C_p=1$) and thermal expansion coefficient ($\alpha=10^{-10}$). Fluid 1 has a density $\rho_1=1$, while fluid 2 is denser ($\rho_2=\rho_1+1.5 \cdot 10^{-10}$) and is placed at the bottom of the domain, for $y \leq 0.025$. Temperature are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides of the domain. The initial temperature field is given by
\[T(x,y)=T_u(x,y)+T_l(x,y)+T_r(x,y)+T_s(x,y)-\frac{3}{2}\]
with
\begin{eqnarray}
T_u(x,y)&=&\frac{1}{2}erf\left(\frac{1-y}{2}\sqrt{\frac{u_0}{x}}\right)\nonumber \\
T_l(x,y)&=&1-\frac{1}{2}erf\left(\frac{y}{2}\sqrt{\frac{u_0}{L_x-x}}\right)\nonumber \\
T_r(x,y)&=&\frac{1}{2}+\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{y+1}} \exp\left(-\frac{x^2u_0}{4y+4}\right)\nonumber \\
T_s(x,y)&=&\frac{1}{2}-\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{2-y}} \exp\left(-\frac{(L_x-x)^2u_0}{8-4y}\right)\nonumber
\end{eqnarray}
and
\begin{eqnarray}
u_0&=&\frac{L_x^{7/3}}{(1+L_x^4)^{2/3}}\left(\frac{Ra}{2\sqrt{\pi}}\right)^{\frac{2}{3}}\nonumber \\
Q&=&2\sqrt{\frac{L_x}{\pi u_0}}\nonumber
\end{eqnarray}

The experiment is performed with two grid resolution (\num{125x40} and \num{200x80} elements), with 100 markers per element and CFL of 0.25. $v_{rms}$ (see Sec. \ref{sec:error}) calculated for both the simulations match well with results from \citet{vanKeken1997} and obtained with ASPECT \citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} and ELEFANT \citep{Thieulot2014}, obtained with similar grid resolutions (\ref{fig:thin}). In addition, a variety of simulations with different resolutions and aspect ratios are performed to verify that the sensitivity of the initial velocity field. Results match well with results from \citet{vanKeken1997} (in the grey area) and \citet{Thieulot2014} (\ref{fig:thin_initial}).

\begin{figure}
\centering
\noindent\includegraphics[width=400px]{./Figures/vrms_thin.pdf}
\caption{$v_{rms}$ for the thin layer experiment as function of time for different grid resolution. Results are compared with results obtained by \citet{vanKeken1997} (black lines) and with ASPECT and ELEFANT.}
\label{fig:thin}
\end{figure}

\begin{figure}
\centering
\noindent\includegraphics[width=400px]{./Figures/vrmszero.pdf}
\caption{$v_{rms}$ at $t=0$ for the thin layer experiment as function of the element size for different aspect ratios of the elements. Results are compared with results obtained by \citet{vanKeken1997} (black dots in the grey area) and with ASPECT and ELEFANT.}
\label{fig:thin_initial}
\end{figure}

\subsection{Sinking hydrated cylinder}\label{sec:quinquis}
This experiment is performed as explained in \citet{Quinquis2014} and simulates the sinking of a cold, hydrated cylinder into a warm, dry mantle. The domain is square with $L_x=L_y=$ \SI{300}{\km}, a grid resolution of \num{300x300} elements and 25 markers per element. The sinking cylinder is located at $x=$ \SI{150}{\km} and $y=$ \SI{170}{\km} with a radius of \SI{20}{\km} and has $\rho_c=$ \SI{3250}{\kg\per\cubic\m}, $\eta_c=$ \SI{e23}{\pascal\s}, a thermal conductivity $k_c=$ \SI{4.5}{\watt\per\m\per\kelvin},a specific heat $C_{p_c}=$ \SI{1250}{\joule\per\kg\per\kelvin}; the surrounding mantle has $\rho_m=$ \SI{3200}{\kg\per\cubic\m}, $\eta_m=$ \SI{e20}{\pascal\s}, a thermal conductivity $k_m=$ \SI{105}{\watt\per\m\per\kelvin} and specific heat $C_{p_m}=$ \SI{1250}{\joule\per\kg\per\kelvin}; a 58 km-thick lithosphere overlies the mantle and has $\rho_l=$ \SI{3200}{\kg\per\cubic\m}, $\eta_l=$ \SI{e23}{\pascal\s}, a thermal conductivity $k_l=$ \SI{4.5}{\watt\per\m\per\kelvin}, specific heat $C_{p_l}=$ \SI{750}{\joule\per\kg\per\kelvin} and a thermal diffusivity of \SI{e-6}{\square\m\per\s}. Velocity boundary conditions are set to free slip on all sides of the domain. Temperature increase linearly in the lithosphere, from 0° to 1300°C, and in the mantle, from 1300° to 1360.5°C. The initial temperature of the cylider is fixed to 400°C. Throughout the evolution, temperature is fixed to 0° and 1360.5°C at the top and bottom of the domain, respectively. The initial bound water content is imposed at 2 wt.\% in the cylinder and at 0 wt.\% in both the mantle and the lithosphere. The maximum amount of water in the mantle is fixed to 0.2 wt.\%, while maximum water content in the cylinder and in the lithosphere is function of pressure and temperature and is calculated using a serpentinized harzburgite with Perple\_X, as in \citet{Quinquis2014}.

As shown in \citet{Quinquis2014}, the model is characterised by a progressive dehydration in the external portion of the cylinder, due to the increase of temperature, with a consequent vertical migration of free water that hydrates the mantle above the cylinder, up to the lithosphere. The distribution of bound and free water during the evolution of the experiment is shown in Fig. \ref{fig:quinquis} (left and right columns, respectively).

\begin{figure}
\centering
\noindent\includegraphics[width=400px]{./Figures/Hydration.pdf}
\caption{Distribution of bound and free water (left and right columns, respectively) at $t=0.5$ and $t=$ \SI{1.25}{\mega\year} (first and second rows, respectively) for the sinking hydrated cylinder experiment. Only markers with bound or free water not equal to 0 are plotted.}
\label{fig:quinquis}
\end{figure}

\subsection{Melting curves}\label{sec:katz}
This instantanous experiment is performed in order to recreate isobaric melting curves, as function of temperature, bulk water and content of water in the melt, to compare with melting curves obtained by \citet{Katz2003}. The domain is square with $L_x=L_y=100$ km, a grid resolution of \num{100x100} elements and 25 markers per element. Density is set in the entire domain to \SI{3300}{\kg\per\cubic\m} and temperature increases from the sides to the centre from 1000° to 1500°C. The experiment is performed for bulk water contents from 0 to 0.3 wt.\%.

Fig. \ref{fig:fraction} shows isobaric melting curves for \SI{1}{\giga\pascal} (panel a) and \SI{3}{\giga\pascal} (panel b) calculated for different bulk water contents. The curves well match with those obtained by \citet{Katz2003}, showing the saturation at the solidus for bulk water of 0.2 and 0.3 wt.\%, indicating that the code correctly determines melt fraction as function of pressure, temperature and bulk water content. Same results can be observed in Fig. \ref{fig:water}a, where isobaric and isothermal melting curves are shown as function of bulk water content. In addition, Fig. \ref{fig:water}b shows that the code correctly determines also the percentage of water in the melt as function of pressure, temperature and melt fraction.

\begin{figure}
\noindent\includegraphics[width=\textwidth]{./Figures/Fraction.pdf}
\caption{Isobaric melting curves for 1 GPa (panel a) and 3 GPa (panel b) as function of temperature with different bulk water contents.}
\label{fig:fraction}
\end{figure}

\begin{figure}
\noindent\includegraphics[width=\textwidth]{./Figures/Water.pdf}
\caption{Isobaric melting curves for 1.5 GPa as function of bulk water (panel a) and water in the melt (panel b) calculated for different temperarures.}
\label{fig:water}
\end{figure}

\clearpage

\bibliography{export}

\end{document}

### 2021-04-11 Paper.tex
modified

@@
 \section{Introduction}\label{sec:intro}
-PROVAThis supplement includes a description of the code, with a presentation of the features implemented followed by results of extensive benchmarks listed in Table \ref{tab:list}, implemented to verify that the code solves correctly a large range of problems.
+This supplement includes a description of the code, with a presentation of the features implemented followed by results of extensive benchmarks listed in Table \ref{tab:list}, implemented to verify that the code solves correctly a large range of problems.
 Description and results of the benchmarks are shown in Section \ref{sec:benchmark}.

### 2021-04-11 Paper.tex
modify

@@
 The maximum depth of the slab is tracked for all the simulations (Fig.
-\ref{fig:subduction}), showing a good match with results from \citet{Schmeling2008}.
+\ref{fig:subduction}) and they are compared with results obtained by \citet{Schmeling2008} with models with regular grids and comparable grid resolutions (rectangles in Fig.
+\ref{fig:subduction}).
 Low viscous sticky air models (i.e., \SI{e19}{\pascal\s}) enlighten a strong dependence of the sinking velocity with the chosen average scheme (dashed lines in Fig.
@@
 \ref{fig:subduction}).
-A higher resistance can be also observed in case of a true free surface for the low resolution grid used (continuous lines in Fig.
+A high resistance at the trench can be also observed in case of a true free surface for the low resolution grid used (continuous lines in Fig.
 \ref{fig:subduction}), as pointed out by \citet{Schmeling2008}.
@@
 \noindent\includegraphics[width=400px]{./Figures/Subduction.pdf}
-\caption{Maximum depth of the slab as function of time for all the simulations of the \citet{Schmeling2008} benchmark, in case of harmonic, geometric and arithmetic means (red, green and blue lines, respectively) using a sticky air or a true free surface (dashed and continuous lines, respectively).}
+\caption{Maximum depth of the slab as function of time for all the simulations of the \citet{Schmeling2008} benchmark, in case of harmonic, geometric and arithmetic means (red, green and blue lines, respectively) using a sticky air or a true free surface (discontinuous and continuous lines, respectively).
+Red, green and blue rectangular areas indicate the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{500}{\km} in case of sticky air, while the grey rectangular area indicates the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{600}{\km} in case of true free surface.}
 \label{fig:subduction}

### 2021-04-13 Paper.tex
modified

@@
 The code is written in Fortran90 and uses the finite element method (FEM) with quadrilateral $Q_1 \times P_0$ elements (continuous bilinear velocity and discontinuous constant pressure), associated with MUMPS\footnote{\url{http://mumps.enseeiht.fr/}} \citep{Amestoy2001,Amestoy2019} that is a software package for solving systems of linear equations of the form $A \cdot x = b$, where \textit{A} is a square sparse matrix, by means of a direct method.
-Since $Q_1 \times P_0$ elements do not satisfy Ladyzhenskaya, Babuska and Brezzi (LBB) stability condition \citep{Donea2003}, elemental pressure is elaborated in postprocessing to avoid spurious pressures by means of a double interpolation to smooth the pressure.
+Since $Q_1 \times P_0$ elements do not satisfy Ladyzhenskaya, Babuska and Brezzi (LBB) stability condition \citep{Donea2003}, elemental pressure is elaborated in post-processing to avoid spurious pressures by means of a double interpolation to smooth the pressure.
 Following this procedure, the elemental pressure is interpolated onto nodes and then back onto elements.
@@
 \end{equation}
-where $d$ is the grain size, $m$ is the grain size exponent, $n$ is the stress exponent, $A$ is the pre-exponensional factor, $I_2$ is the square root of the second invariant of the strain rate tensor, $Q$ is the activation energy, $p$ is the pressure, $V$ is the activation volume, $R$ is the gas constant and $T$ is the temperature.
-Pressures and temperatures are determined by the interpolation of the nodal parameters, while strain rates are calculated by means of the derivatve of the velocities on the nodes.
+where $d$ is the grain size, $m$ is the grain size exponent, $n$ is the stress exponent, $A$ is the pre-exponential factor, $I_2$ is the square root of the second invariant of the strain rate tensor, $Q$ is the activation energy, $p$ is the pressure, $V$ is the activation volume, $R$ is the gas constant and $T$ is the temperature.
+Pressures and temperatures are determined by the interpolation of the nodal parameters, while strain rates are calculated by means of the derivative of the velocities on the nodes.
 In case of diffusion creep $n=1$ and $m>0$, in case of dislocation creep $n>1$ and $m=0$.
@@
 
-Hydration processes related to dehydration of subducting lithosphere \citep{Schmidt1998,Liu2007,Faccenda2009,Faccenda2010,Faccenda2014,Rosas2016} strongly influence the thermo-mechanics inside the mantle wedge, mainly because of both weakening effects on the mantle rhology and density variations in case of mantle serpentinization \citep{Gerya2002,Honda2003,Arcay2005,Roda2010,Regorda2017}.
+Hydration processes related to dehydration of subducting lithosphere \citep{Schmidt1998,Liu2007,Faccenda2009,Faccenda2010,Faccenda2014,Rosas2016} strongly influence the thermo-mechanics inside the mantle wedge, mainly because of both weakening effects on the mantle rheology and density variations in case of mantle serpentinization \citep{Gerya2002,Honda2003,Arcay2005,Roda2010,Regorda2017}.
 In the case that hydration is switched on, the amount of bound and free water is memorised by each marker, following the implementation of \citet{Quinquis2014}; hydration and dehydration processes are related to the amount of bound water of each marker and to the maximum amount of water it can transport, i.e.
-if the amount of bound water exceeds the maximum amount of water, the marker dehydrates and releases free water that can hydrates undersaturated markers.
+if the amount of bound water exceeds the maximum amount of water, the marker dehydrates and releases free water that can hydrates under-saturated markers.
 Maximum water content of each marker is determined as function of lithology and $p$-$T$ conditions and it is calculated using Perple\_X, in the same way than effective density, specific heat and coefficient of thermal expansion.
 Bound water is advected together with the markers, neglecting the effect of bound water diffusion, while free water simply migrates vertically and is not coupled to the solid-phase flow of the mantle wedge \citep{Arcay2005,Quinquis2014}.
-The correctness of the migration scheme of the free water and its absorption into bound water by undersaturated markers are verified performing the experiment described in \citet{Quinquis2014} (Sec.
+The correctness of the migration scheme of the free water and its absorption into bound water by under-saturated markers are verified performing the experiment described in \citet{Quinquis2014} (Sec.
 \ref{sec:quinquis}).
@@
 \end{eqnarray}
-where $T_{sol}$ is the temperature od the dry solidus, $D_{H_2O}$ is the partition coefficient and $x_t\cdot log(p+y_t)$ indicates the pressure dependence of the melting curve \citep{Kelley2010}.
+where $T_{sol}$ is the temperature of the dry solidus, $D_{H_2O}$ is the partition coefficient and $x_t\cdot log(p+y_t)$ indicates the pressure dependence of the melting curve \citep{Kelley2010}.
 $x_t$, $y_t$, $K_m$ and $\gamma$ are parameters chosen in according to \citet{Kelley2010}.
@@
 \ref{eq:momentum_penalty}) are taken into account considering effective coefficient of thermal expansion.
-Similarly, latent heat due to melting/crystallisation can be implicitily included in the energy equation (Eq.
+Similarly, latent heat due to melting/crystallisation can be implicitly included in the energy equation (Eq.
 \ref{eq:energy}) by considering effective specific heat ($C_{p_{eff}}$) and thermal expansion ($\alpha_{eff}$) of the partially molten rocks, in according to
@@
 \end{eqnarray}
-At $t=0$, the disk is centered at position $(0.5;0.75)$ with a radius $R=0.15$ and has a vertical fissure 0.05 wide and 0.2 high (orange in Fig.
+At $t=0$, the disk is centred at position $(0.5;0.75)$ with a radius $R=0.15$ and has a vertical fissure 0.05 wide and 0.2 high (orange in Fig.
 \ref{fig:runge}).
@@
 The experiment is performed with different grid sizes (\num{50x50}, \num{80x80}, \num{100x100} and \num{256x256}).
-A total of 1960000 markers are randomly distributed at the begininning of the simulation.
+A total of 1960000 markers are randomly distributed at the beginning of the simulation.
 Velocity boundary conditions are set to no slip at the top and the bottom, and to free slip at the sides of the domain.
@@
 The domain is square with $L_x=L_y=$ \SI{500}{\km} and the grid is composed by \num{50x50} elements with 25 markers in each element.
-The block is initially centered at ($x=$ \SI{250}{\km}; $y=$ \SI{400}{\km}) and has a size of \num{100x100} km (Fig.
+The block is initially centred at ($x=$ \SI{250}{\km}; $y=$ \SI{400}{\km}) and has a size of \num{100x100} km (Fig.
 \ref{fig:block}a).
@@
 \subsection{2D time-dependent Stokes sphere with deformable free surface}\label{sec:sphere}
-This experiment is perfomed in a unit square domain with the gravity acceleration fixed to $g_y=-1$.
-The sphere has $\rho_s=2$ and $\eta_s=10^3$ and is initially centered at ($0.5;0.6$) with radius $R=0.123456789$.
+This experiment is performed in a unit square domain with the gravity acceleration fixed to $g_y=-1$.
+The sphere has $\rho_s=2$ and $\eta_s=10^3$ and is initially centred at ($0.5;0.6$) with radius $R=0.123456789$.
 The fluid surrounding the sphere has $\rho_f=1$ and $\eta_f=1$ and occupies the domain for $y\leq 0.75$, while for $y>0.75$ the air has $\rho_a=0$ and $\eta_a=10^{-3}$.
@@
 The interface between the fluid and the air is tracked by means of the markers chain.
-Results in terms of velocity and pressure fieldd and topography variations are compared with results obtained with ASPECT \citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} and they can be found at \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere_fs2D}.
+Results in terms of velocity and pressure fields and topography variations are compared with results obtained with ASPECT \citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} and they can be found at \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere_fs2D}.
 Results in terms of the $v_{rms}$ in case of an arithmetic average are also shown in Fig.
@@
 The domain is rectangular with $L_x=$ \SI{3000}{\km} and $L_y=$ \SI{700}{\km}, a grid resolution of \num{256x64} elements with 50 markers per element and Courant number of 0.01.
-At the beginning of the simulation, a 100 km-thick lithosperic layer with $\rho_m=$ \SI{3300}{\kg\per\cubic\metre} and $\eta_m=$ \SI{e23}{\pascal\s} is located at the top of the domain for $x$ between 1000 and 3000 km, while the asthenospheric mantle has $\rho_m=$ \SI{3200}{\pascal\s} and $\eta_m=$ \SI{e21}{\pascal\s}.
-In addition, a 200 km-depth lithosperic slab is already subducted in the mantle in order to have a spontaneous subduction.
+At the beginning of the simulation, a 100 km-thick lithospheric layer with $\rho_m=$ \SI{3300}{\kg\per\cubic\metre} and $\eta_m=$ \SI{e23}{\pascal\s} is located at the top of the domain for $x$ between 1000 and 3000 km, while the asthenospheric mantle has $\rho_m=$ \SI{3200}{\pascal\s} and $\eta_m=$ \SI{e21}{\pascal\s}.
+In addition, a 200 km-depth lithospheric slab is already subducted in the mantle in order to have a spontaneous subduction.
 In case of sticky air, $L_y=750$ km and a 50 km-thick air layer with $\eta_a$ from \SIrange{e19}{e21}{\pascal\s} overlie the mantle.
@@
 The effective viscosity is capped between $\eta_{min}=$ \SI{e21}{\pascal\s} and $\eta_{max}=$ \SI{e25}{\pascal\s}.
-Velocity bounadry conditions are set to free slip at the top and the bottom, and to no slip at the sides of the domain.
+Velocity boundary conditions are set to free slip at the top and the bottom, and to no slip at the sides of the domain.
 
@@
 \noindent\includegraphics[width=\textwidth]{./Figures/Slab.pdf}
-\caption{Effective viscosity fot the slab detachment benchmark at the beginning of the evolution (panel a) and when necking is complete (panel b).}
+\caption{Effective viscosity for the slab detachment benchmark at the beginning of the evolution (panel a) and when necking is complete (panel b).}
 \label{fig:slab}
@@
 \caption{Temperature profile as function of \textit{x} for the advection stabilisation benchmark.
-Purple line indicates the initial temperature profile; the green line inidicates the analytical temperature profile after 250 time steps; blue line indicates the temperature profile after 250 time steps in case of the classic Galerkin method ($\gamma=0$); orange line indicates the temperature profile after 250 time steps in case of the SUPG method ($\gamma=0.045$).}
+Purple line indicates the initial temperature profile; the green line indicates the analytical temperature profile after 250 time steps; blue line indicates the temperature profile after 250 time steps in case of the classic Galerkin method ($\gamma=0$); orange line indicates the temperature profile after 250 time steps in case of the SUPG method ($\gamma=0.045$).}
 \label{fig:advection}
@@
 As shown in Fig.
-\ref{fig:energy}, both shear and adiabatic heating precited by the code (first row in Fig.
+\ref{fig:energy}, both shear and adiabatic heating predicted by the code (first row in Fig.
 \ref{fig:energy}) well recreate the results obtained by means of example ${Shear\_adiabatic\_heating.m}$ from \citet{Gerya2010b} (second row in Fig.
@@
 Temperature increase linearly in the lithosphere, from 0° to 1300°C, and in the mantle, from 1300° to 1360.5°C.
-The initial temperature of the cylider is fixed to 400°C.
+The initial temperature of the cylinder is fixed to 400°C.
 Throughout the evolution, temperature is fixed to 0° and 1360.5°C at the top and bottom of the domain, respectively.
@@
 \subsection{Melting curves}\label{sec:katz}
-This instantanous experiment is performed in order to recreate isobaric melting curves, as function of temperature, bulk water and content of water in the melt, to compare with melting curves obtained by \citet{Katz2003}.
+This instantaneous experiment is performed in order to recreate isobaric melting curves, as function of temperature, bulk water and content of water in the melt, to compare with melting curves obtained by \citet{Katz2003}.
 The domain is square with $L_x=L_y=100$ km, a grid resolution of \num{100x100} elements and 25 markers per element.
@@
 \noindent\includegraphics[width=\textwidth]{./Figures/Water.pdf}
-\caption{Isobaric melting curves for 1.5 GPa as function of bulk water (panel a) and water in the melt (panel b) calculated for different temperarures.}
+\caption{Isobaric melting curves for 1.5 GPa as function of bulk water (panel a) and water in the melt (panel b) calculated for different temperatures.}
 \label{fig:water}

### 2021-04-13 Paper.tex
modified

@@
 Viscous creep and plastic yielding are non-linear rheologies because of their dependence on the velocity field through pressure and strain rates.
-Therefore, the solution is determined by means of Picard-type iterations, until convergence of the velocity field \citep{Thieulot2011}.
-The correlation between two consecutive non-linear iterations is given by
+Therefore, the solution is determined by means of Picard-type iterations, until convergence of the velocity field \citep{Glerum2018}.
+The convergence is verified at each iteration $i$ via the nonlinear residual $\bm{\mathcal{R}}^i$ that can be determined as
 \begin{equation}\label{eq:convergence}
-\chi=1-\frac{\langle (\bm{v}^i-\langle\bm{v}^i\rangle)\cdot (\bm{v}^{i+1}-\langle\bm{v}^{i+1}\rangle) \rangle}{\sqrt{\sigma^i\sigma^{i+1}}}
+\bm{\mathcal{R}}^i= \mathbb{K} (\eta_{eff}(\dot{\bm{\epsilon}}^{i-1},p^{i-1}))\cdot \bm{v}^{i-1} - \bm{f}^i
 \end{equation}
-Iterations are performed until either convergence ($\chi<tol$) or maximal number of non-linear iterations ($it_{max}$) is reached.
+%\begin{equation}\label{eq:convergence}
+%\chi=1-\frac{\langle (\bm{v}^i-\langle\bm{v}^i\rangle)\cdot (\bm{v}^{i+1}-\langle\bm{v}^{i+1}\rangle) \rangle}{\sqrt{\sigma^i\sigma^{i+1}}}
+%\end{equation}
+%Iterations are performed until either convergence ($\chi<tol$) or maximal number of non-linear iterations ($it_{max}$) is reached.
+where $\mathbb{K}$ is the Stiffness matrix and $\bm{f}$ is the right hand side vector \citep{Spiegelman2016,Glerum2018}.
+The $L_2$-norm is extracted from $\bm{\mathcal{R}}^i$ and it is normalised using the $L_2$-norm extracted from $\bm{f}^0$.
+Iterations are performed until either convergence ($\frac{||\bm{\mathcal{R}}^i||_2}{||\bm{f}^0||_2}<tol$) or maximal number of non-linear iterations ($it_{max}$) is reached.
 

### 2021-04-13 Paper.tex
Modified

@@
 \section{Introduction}\label{sec:intro}
-This supplement includes a description of the code, with a presentation of the features implemented followed by results of extensive benchmarks listed in Table \ref{tab:list}, implemented to verify that the code solves correctly a large range of problems.
-Description and results of the benchmarks are shown in Section \ref{sec:benchmark}.
+This supplement includes a description of the code, with a presentation of the features implemented followed by results of extensive benchmarks listed in Table \ref{tab:list}, implemented 
+to verify that the code solves correctly a large range of problems. Description and results of the benchmarks are shown in Section \ref{sec:benchmark}.
 
@@
 
-The code is written in Fortran90 and uses the finite element method (FEM) with quadrilateral $Q_1 \times P_0$ elements (continuous bilinear velocity and discontinuous constant pressure), associated with MUMPS\footnote{\url{http://mumps.enseeiht.fr/}} \citep{Amestoy2001,Amestoy2019} that is a software package for solving systems of linear equations of the form $A \cdot x = b$, where \textit{A} is a square sparse matrix, by means of a direct method.
-Since $Q_1 \times P_0$ elements do not satisfy Ladyzhenskaya, Babuska and Brezzi (LBB) stability condition \citep{Donea2003}, elemental pressure is elaborated in post-processing to avoid spurious pressures by means of a double interpolation to smooth the pressure.
-Following this procedure, the elemental pressure is interpolated onto nodes and then back onto elements.
-After the smoothing procedure, the pressure field is calculated again on the nodes to be used in combination with the lithostatic pressure, which is calculated onto nodes.
-Correctness of the solution and performances of the solver in terms of time and memory usage are tested solving the Stokes flow with the analytical solution proposed by \citet{Donea2003} (Section \ref{sec:stokes}).
+The code is written in Fortran90 and uses the finite element method (FEM) with quadrilateral $Q_1 \times P_0$ elements (continuous bilinear velocity and
+discontinuous constant pressure), associated with MUMPS\footnote{\url{http://mumps.enseeiht.fr/}} \citep{Amestoy2001,Amestoy2019} that is a software package
+for solving systems of linear equations of the form $A \cdot x = b$, where \textit{A} is a square sparse matrix, by means of a direct method. Since
+$Q_1 \times P_0$ elements do not satisfy Ladyzhenskaya, Babuska and Brezzi (LBB) stability condition \citep{Donea2003}, elemental pressure is elaborated
+in post-processing to avoid spurious pressures by means of a double interpolation to smooth the pressure. Following this procedure, the elemental pressure
+is interpolated onto nodes and then back onto elements. After the smoothing procedure, the pressure field is calculated again on the nodes to be used in
+combination with the lithostatic pressure, which is calculated onto nodes. Correctness of the solution and performances of the solver in terms of time and
+memory usage are tested solving the Stokes flow with the analytical solution proposed by \citet{Donea2003} (Section \ref{sec:stokes}).
 
@@
 \end{eqnarray}
-where $\rho$ is the density, $\bm{u}$ is the velocity, \textit{p} is the pressure, $\bm{\tau}$ is the deviatoric stress, $\bm{g}$ is the gravity acceleration, $C_p$ is the specific heat at constant pressure, \textit{T} is the temperature, \textit{k} is the thermal conductivity and $H_{tot}$ is the total internal heat production.
-Density variations due to temperature are generally small enough to assume the density as constant ($\rho=\rho_0$) in Eq.
-\ref{eq:mass} and in Eq.
-\ref{eq:energy}, while it must be treated as a variable in the buoyancy term of Eq.
-\ref{eq:momentum}, such that \[\rho=\rho_0(1-\alpha(T-T_0))\]
+where $\rho$ is the density, $\bm{u}$ is the velocity, \textit{p} is the pressure, $\bm{\tau}$ is the deviatoric stress, $\bm{g}$ is the gravity acceleration,
+$C_p$ is the specific heat at constant pressure, \textit{T} is the temperature, \textit{k} is the thermal conductivity and $H_{tot}$ is the total internal heat
+production. Density variations due to temperature are generally small enough to assume the density as constant ($\rho=\rho_0$) in Eq. \ref{eq:mass} and in
+Eq. \ref{eq:energy}, while it must be treated as a variable in the buoyancy term of Eq. \ref{eq:momentum}, such that \[\rho=\rho_0(1-\alpha(T-T_0))\]
 where $\rho_0$ is the density at a reference temperature $T_0$ and $\alpha$ is the coefficient of thermal expansion. Eq. \ref{eq:mass} then can be rewritten as
@@
 \[dt=C_n\; \min\left(\frac{\Delta x}{\max|\bm{u}|},\frac{(\Delta x)^2}{\kappa}\right)\]
-where $0<C_n<1$ is the Courant (CFL) number, $\Delta x$ is the minimum dimension of smallest element, $\max|\bm{u}|$ is the maximum velocity calculated on the entire domain and $\kappa=\frac{k}{\rho C_p}$ is the thermal diffusivity \citep{Thieulot2014}.
+where $0<C_n<1$ is the Courant (CFL) number, $\Delta x$ is the minimum dimension of smallest element, $\max|\bm{u}|$ is the maximum velocity calculated on the
+entire domain and
+$\kappa=\frac{k}{\rho C_p}$ is the thermal diffusivity \citep{Thieulot2014}.
 
 \section{Lagrangian markers}\label{sec:markers}
-Elemental properties (density, viscosity, thermal conductivity, specific heat and thermal expansion) needed to solve Eqs.
-\ref{eq:mass_0}, \ref{eq:momentum} and \ref{eq:energy} are related to the composition of each element, determined by means of Lagrangian markers that are characteristic of different materials in the domain.
-Elemental properties (with the sole exception of the viscosity) are calculated using an arithmetic mean, as \[P_e=\frac{1}{n}\sum_{i=1}^n P_i\]
+Elemental properties (density, viscosity, thermal conductivity, specific heat and thermal expansion) needed to solve Eqs. \ref{eq:mass_0}, \ref{eq:momentum}
+and \ref{eq:energy} are related to the composition of each element, determined by means of Lagrangian markers that are characteristic of different materials
+in the domain. Elemental properties (with the sole exception of the viscosity) are calculated using an arithmetic mean, as \[P_e=\frac{1}{n}\sum_{i=1}^n P_i\]
 where $P_e$ is the elemental property, $P_i$ is the property characteristic of the material of each marker and $n$ is the number of markers of the element.
-Differently, the average scheme for the elemental viscosity can be chosen between harmonic, geometric and arithmetic mean.
-Markers can be placed regularly or randomly at the beginning of the simulation and their advection is performed by a 4th-order Runge-Kutta in space, interpolating the velocity field on each marker by means of the shape functions (see Section \ref{sec:runge}).
-The initial distribution of the markers is created using the open-source code library Geodynamic World Builder \citep{Fraters2019}.
-During the simulation, each marker carries memory of temperature, pressure and accumulated strain, which are determined interpolating the relative nodal parameter.
-In particular, pressure is interpolated onto each marker after the smoothing procedure.
-The number of markers contained in each element is maintained between $n_{min}$ and $n_{max}$.
-When in an element there are less markers than $n_{min}$ the code adds random markers to reach the $n_{min}$, while if the number is higher than $n_{max}$ some of them are randomly deleted.
-When new markers are added, they assume the properties of the nearest marker.
-In this way elements are never empty and maintain a number of markers inside a prefixed range.
+Differently, the average scheme for the elemental viscosity can be chosen between harmonic, geometric and arithmetic mean. Markers can be placed regularly or
+randomly at the beginning of the simulation and their advection is performed by a 4th-order Runge-Kutta in space, interpolating the velocity field on each
+marker by means of the shape functions (see Section \ref{sec:runge}). The initial distribution of the markers is created using the open-source code library
+Geodynamic World Builder \citep{Fraters2019}. During the simulation, each marker carries memory of temperature, pressure and accumulated strain, which are
+determined interpolating the relative nodal parameter. In particular, pressure is interpolated onto each marker after the smoothing procedure. The number
+of markers contained in each element is maintained between $n_{min}$ and $n_{max}$. When in an element there are less markers than $n_{min}$ the code adds
+random markers to reach the $n_{min}$, while if the number is higher than $n_{max}$ some of them are randomly deleted. When new markers are added, they assume
+the properties of the nearest marker. In this way elements are never empty and maintain a number of markers inside a prefixed range.
 
@@
 \end{equation}
-where $\lambda$ is the so-called penalty parameter, which should be 6-7 orders of magnitude larger than the shear viscosity to ensure that mass conservation is satisfied \citep{Donea2003,Thieulot2014}.
+where $\lambda$ is the so-called penalty parameter, which should be 6-7 orders of magnitude larger than the shear viscosity to ensure that mass conservation
+is satisfied \citep{Donea2003,Thieulot2014}.
 
-The deviatoric stress tensor in Eq.
-\ref{eq:momentum} can be written in terms of the strain rate tensor as $\bm{\tau}=2\eta\dot{\bm{\epsilon}}$, with $\dot{\bm{\epsilon}}=\frac{1}{2}\left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)$.
-Therefore, Eq.
-\ref{eq:momentum} can be rewritten as:
+The deviatoric stress tensor in Eq. \ref{eq:momentum} can be written in terms of the strain rate tensor as $\bm{\tau}=2\eta\dot{\bm{\epsilon}}$, with
+$\dot{\bm{\epsilon}}=\frac{1}{2}\left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)$. Therefore, Eq. \ref{eq:momentum} can be rewritten as:
 \begin{equation}\label{eq:momentum_pressure}
@@
 \begin{equation}\label{eq:momentum_penalty}
-\lambda \bm{\nabla} \left(\bm{\nabla} \cdot \bm{v} \right) + \bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right) + \rho \bm{g} = 0
+\lambda \bm{\nabla} \left(\bm{\nabla} \cdot \bm{v}\right)+\bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right)+\rho\bm{g}=0
 \end{equation}
-The penalty method is been associated to the iterative Uzawa method, as described in detail in \citet{Dabrowski2008} and \citet{Thieulot2014} (see Section \ref{sec:poiseuille}).
-Results of benchmarks performed to verify the correctness in the implementation of Eq.
-\ref{eq:momentum_penalty} are shown in Sections \ref{sec:ist_sphere} and \ref{sec:rayleigh}.
+The penalty method is been associated to the iterative Uzawa method, as described in detail in \citet{Dabrowski2008} and \citet{Thieulot2014}
+(see Section \ref{sec:poiseuille}). Results of benchmarks performed to verify the correctness in the implementation of Eq. \ref{eq:momentum_penalty} are
+shown in Sections \ref{sec:ist_sphere} and \ref{sec:rayleigh}.
 
-To support large viscosities variations the penalty parameter is related to the effective elemental viscosity by means of a dimensionless coefficient, so that $\lambda=\lambda_e(e)\eta_{eff}(e)$ \citep{Marotta2006,Dabrowski2008,Thieulot2014}.
-Benchmark of the falling block \citep{Gerya2003a,Gerya2010b,Thieulot2011} is performed to verify that the code can correctly deal buoyancy driven flows with strong viscosity contrasts (Section \ref{sec:block}).
+To support large viscosities variations the penalty parameter is related to the effective elemental viscosity by means of a dimensionless coefficient, so that
+$\lambda=\lambda_e(e)\eta_{eff}(e)$ \citep{Marotta2006,Dabrowski2008,Thieulot2014}. Benchmark of the falling block \citep{Gerya2003a,Gerya2010b,Thieulot2011}
+is performed to verify that the code can correctly deal buoyancy driven flows with strong viscosity contrasts (Section \ref{sec:block}).
 
 \section{Sticky air and free surface}\label{sec:sticky}
-The Earth's surface can be treated by means of either the so-called sticky-air or a true free surface method, both of them implemented in the code.
-In the sticky-air method the surface is approximated with the introduction of a buoyant layer with a viscosity at least four orders of magnitude lower than the crust \citep{Schmeling2008,Crameri2012} and the interface between lithosphere and air is defined using a chain of passive markers that are advected as the Lagrangian markers.
-The correctness of the evolution of the markers chain is tested with the experiment of a 2D time-dependent Stokes sphere below a free surface and compared with results from ASPECT \citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} (Section \ref{sec:sphere}).
-In the true free surface case the top boundary is assumed stress-free and velocities are not fixed.
-In this case topography variations are described by vertical deformations of the mesh that depend on the velocity field of the nodes that identify the free surface, while horizontal deformations are not taken into account.
-This procedure is known as the Arbitrary Lagrangian-Eulerian (ALE) method and its implementation follows the technique described in \citet{Thieulot2011}.
-However, although the implementation of a true free surface better reproduces laboratory experiment, extremely small time steps can be necessary to maintain stability \citep{Kaus2010a,Quinquis2011,Thieulot2014}.
-Therefore, the stability algorithm proposed by \citet{Kaus2010a} is implemented to avoid instabilities due to high density differences at the free surface when using too large time steps.
-The implementation of this algorithm are tested by performing the experiment described by \citet{Kaus2010a} (Section \ref{sec:stab}).
+The Earth's surface can be treated by means of either the so-called sticky-air or a true free surface method, both of them implemented in the code. In the
+sticky-air method the surface is approximated with the introduction of a buoyant layer with a viscosity at least four orders of magnitude lower than the crust
+\citep{Schmeling2008,Crameri2012} and the interface between lithosphere and air is defined using a chain of passive markers that are advected as the Lagrangian
+markers. The correctness of the evolution of the markers chain is tested with the experiment of a 2D time-dependent Stokes sphere below a free surface and
+compared with results from ASPECT \citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} (Section \ref{sec:sphere}). In the true free surface
+case the top boundary is assumed stress-free and velocities are not fixed. In this case topography variations are described by vertical deformations of the
+mesh that depend on the velocity field of the nodes that identify the free surface, while horizontal deformations are not taken into account. This procedure
+is known as the Arbitrary Lagrangian-Eulerian (ALE) method and its implementation follows the technique described in \citet{Thieulot2011}. However, although
+the implementation of a true free surface better reproduces laboratory experiment, extremely small time steps can be necessary to maintain stability
+\citep{Kaus2010a,Quinquis2011,Thieulot2014}. Therefore, the stability algorithm proposed by \citet{Kaus2010a} is implemented to avoid instabilities due to
+high density differences at the free surface when using too large time steps. The implementation of this algorithm are tested by performing the experiment
+described by \citet{Kaus2010a} (Section \ref{sec:stab}).
 
-The topography relaxation benchmark proposed by \citet{Crameri2012} is performed to verify that the code correctly recovers topography variations in case of both the sticky-air and the true free surface method (Section \ref{sec:crameri}).
-Finally, Section \ref{sec:subduction} show the results of the spontaneous subduction experiment described by \citet{Schmeling2008}.
+The topography relaxation benchmark proposed by \citet{Crameri2012} is performed to verify that the code correctly recovers topography variations in case of
+both the sticky-air and the true free surface method (Section \ref{sec:crameri}). Finally, Section \ref{sec:subduction} show the results of the spontaneous
+subduction experiment described by \citet{Schmeling2008}.
 
 \section{Erosion and sedimentation}\label{sec:erosion}
-Surface processes at the lithosphere-air interface are implemented taking into account fluvial erosion (long-range processes), hillslope processes (short-range processes) and failure of slopes exceeding a threshold angle $\theta$ \citep{Burov1997,Tucker1999,Sacek2012,Croissant2014,Ueda2015,Beucher2020}.
-Erosion rates $\dot{E}_{lr}$ due to long-range processes can be described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, in according to
+Surface processes at the lithosphere-air interface are implemented taking into account fluvial erosion (long-range processes), hillslope processes (short-range
+processes) and failure of slopes exceeding a threshold angle $\theta$ \citep{Burov1997,Tucker1999,Sacek2012,Croissant2014,Ueda2015,Beucher2020}. Erosion rates
+$\dot{E}_{lr}$ due to long-range processes can be described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, in according to
 \begin{equation}
@@
 \end{equation}
-where $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $\bm{\nabla} h$ is the slope and $m_e$ and $n_e$ are parameters chosen in to constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}.
-Erosion rates $\dot{E}_{sr}$ due to short-range processes are implemented by assuming that transport is linearly proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}, in according to
+where $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $\bm{\nabla} h$ is the slope and $m_e$ and $n_e$ are parameters chosen in to
+constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}. Erosion rates $\dot{E}_{sr}$ due to short-range processes are
+implemented by assuming that transport is linearly proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}, in according to
 \begin{equation}
@@
 \end{equation}
-where $K_d$ is the effective diffusivity. In addition, the slope is modified if it still exceeds a threshold angle after long- and short-range processes \citep{Ueda2015}.
+where $K_d$ is the effective diffusivity. In addition, the slope is modified if it still exceeds a threshold angle after long- and short-range processes
+\citep{Ueda2015}.
 
@@
 \end{equation}
-where $K^*_s$ is a constant, $h_{sl}$ is the sea level and $h_0$ is a characteristic length.
-In case of sticky air, slope is determined by means of the markers chain, which is vertically corrected considering erosion and sedimentation rates.
-After the correction, continental markers above the markers chain are transformed in air markers and air markers below the markers chain become sediments.
-Similarly, slope is determined considering the nodes on the top boundary in case of true free surface and they are vertically corrected as for the markers chain.
-In this case, continental markers above the top boundary are simply deleted.
+where $K^*_s$ is a constant, $h_{sl}$ is the sea level and $h_0$ is a characteristic length. In case of sticky air, slope is determined by means of the markers
+chain, which is vertically corrected considering erosion and sedimentation rates. After the correction, continental markers above the markers chain are
+transformed in air markers and air markers below the markers chain become sediments. Similarly, slope is determined considering the nodes on the top boundary
+in case of true free surface and they are vertically corrected as for the markers chain. In this case, continental markers above the top boundary are simply
+deleted.
 
 \section{Non-linear rheologies}\label{sec:nonlinear}
-Non-linear rheologies are implemented combining viscous creep (dislocation and diffusion) and plastic yielding.
-For each marker, diffusion and dislocation viscosity ($\eta_{df|ds}$) can be determined as follows \citep{Karato1993,Wang2016,Glerum2018}:
+Non-linear rheologies are implemented combining viscous creep (dislocation and diffusion) and plastic yielding. For each marker, diffusion and dislocation
+viscosity ($\eta_{df|ds}$) can be determined as follows \citep{Karato1993,Wang2016,Glerum2018}:
 \begin{equation}\label{eq:viscous}
@@
 \end{equation}
-where $d$ is the grain size, $m$ is the grain size exponent, $n$ is the stress exponent, $A$ is the pre-exponential factor, $I_2$ is the square root of the second invariant of the strain rate tensor, $Q$ is the activation energy, $p$ is the pressure, $V$ is the activation volume, $R$ is the gas constant and $T$ is the temperature.
-Pressures and temperatures are determined by the interpolation of the nodal parameters, while strain rates are calculated by means of the derivative of the velocities on the nodes.
-In case of diffusion creep $n=1$ and $m>0$, in case of dislocation creep $n>1$ and $m=0$.
-The implementation of the non-linear viscous creep viscosity depending on the strain rate is tested by means of the slab detachment benchmark (Section \ref{sec:slab}).
+where $d$ is the grain size, $m$ is the grain size exponent, $n$ is the stress exponent, $A$ is the pre-exponential factor, $I_2$ is the square root of the
+second invariant of the strain rate tensor, $Q$ is the activation energy, $p$ is the pressure, $V$ is the activation volume, $R$ is the gas constant and $T$
+is the temperature. Pressures and temperatures are determined by the interpolation of the nodal parameters, while strain rates are calculated by means of the
+derivative of the velocities on the nodes. In case of diffusion creep $n=1$ and $m>0$, in case of dislocation creep $n>1$ and $m=0$. The implementation of the
+non-linear viscous creep viscosity depending on the strain rate is tested by means of the slab detachment benchmark (Section \ref{sec:slab}).
 
-Plastic yielding is implemented rescaling $\eta_{vc}$ in order to limit the stress below the yield stress $\sigma_y$ \citep{Thieulot2008,Thieulot2014,Glerum2018}, obtaining
+Plastic yielding is implemented rescaling $\eta_{vc}$ in order to limit the stress below the yield stress $\sigma_y$ \citep{Thieulot2008,Thieulot2014,Glerum2018},
+obtaining
 \begin{equation}\label{eq:plastic}
@@
 \end{equation}
-where $C$ is the cohesion and $\phi$ is the internal friction angle.
-In case of negative pressure it is imposed equal to 0, so that negative yield stress are excluded.
-The correctness of the non-linear solution in case of plastic viscosity with variable internal friction angle is verified performing indenter and brick experiments (Sections \ref{sec:indenter} and \ref{sec:brick}, respectively).
+where $C$ is the cohesion and $\phi$ is the internal friction angle. In case of negative pressure it is imposed equal to 0, so that negative yield stress are
+excluded. The correctness of the non-linear solution in case of plastic viscosity with variable internal friction angle is verified performing indenter and
+brick experiments (Sections \ref{sec:indenter} and \ref{sec:brick}, respectively).
 
-Strain softening is taken into account for both viscous creep and plastic viscosity \citep{Huismans2003,Babeyko2005,Huismans2005,Sobolev2005,Warren2008a} by means of the accumulated strain $\epsilon$ memorised by each marker.
-Softening in the viscous creep determines a linear decrease of $\eta_{vc}$ by means of a viscous strain softening factor $W_S$ that increases linearly from 1 to 10 for $5<\epsilon<10$ \citep{Warren2008a}.
-This viscous softening can be related to strain-induced grain size reduction \citep{Warren2008a}.
-Differently, plastic softening is simulated with a linear decrease of internal friction angle $\phi(\epsilon)$ and cohesion $C(\epsilon)$ in according to
+Strain softening is taken into account for both viscous creep and plastic viscosity \citep{Huismans2003,Babeyko2005,Huismans2005,Sobolev2005,Warren2008a} by
+means of the accumulated strain $\epsilon$ memorised by each marker. Softening in the viscous creep determines a linear decrease of $\eta_{vc}$ by means of a
+viscous strain softening factor $W_S$ that increases linearly from 1 to 10 for $5<\epsilon<10$ \citep{Warren2008a}. This viscous softening can be related to
+strain-induced grain size reduction \citep{Warren2008a}. Differently, plastic softening is simulated with a linear decrease of internal friction angle
+$\phi(\epsilon)$ and cohesion $C(\epsilon)$ in according to
 \begin{eqnarray}
@@
 \end{eqnarray}
-where $\phi_0$, $C_0$ and $\phi_{\infty}$, $C_{\infty}$ are internal friction angle and cohesion for $\epsilon=0.5$ and $\epsilon=1.5$, respectively \citep{Huismans2003,Huismans2005,Warren2008a,Thieulot2014}.
-Plastic softening approximates deformation-induced softening of faults and brittle shear zones \citep{Warren2008a}.
+where $\phi_0$, $C_0$ and $\phi_{\infty}$, $C_{\infty}$ are internal friction angle and cohesion for $\epsilon=0.5$ and $\epsilon=1.5$, respectively
+\citep{Huismans2003,Huismans2005,Warren2008a,Thieulot2014}. Plastic softening approximates deformation-induced softening of faults and brittle shear zones
+\citep{Warren2008a}.
 
-Viscous creep and plastic yielding are then combined to obtain a viscoplastic viscosity $\eta_{vp}$, which is computed with an harmonic average \citep{Tackley2010,Glerum2018} as follows
+Viscous creep and plastic yielding are then combined to obtain a viscoplastic viscosity $\eta_{vp}$, which is computed with an harmonic average
+\citep{Tackley2010,Glerum2018} as follows
 \begin{equation}\label{eq:viscoplastic}
@@
 \end{equation}
-Finally, effective viscosity $\eta_{eff}$ is capped by the minimum and the maximum viscosity ($\eta_{min}$ and $\eta_{max}$, respectively) to avoid extremely low or high viscosity \citep{Glerum2018} as follows
+Finally, effective viscosity $\eta_{eff}$ is capped by the minimum and the maximum viscosity ($\eta_{min}$ and $\eta_{max}$, respectively) to avoid extremely
+low or high viscosity \citep{Glerum2018} as follows
 \begin{equation}\label{eq:effective}
@@
 \end{equation}
-Viscous creep and plastic yielding are non-linear rheologies because of their dependence on the velocity field through pressure and strain rates.
-Therefore, the solution is determined by means of Picard-type iterations, until convergence of the velocity field \citep{Glerum2018}.
-The convergence is verified at each iteration $i$ via the nonlinear residual $\bm{\mathcal{R}}^i$ that can be determined as
+Viscous creep and plastic yielding are non-linear rheologies because of their dependence on the velocity field through pressure and strain rates. Therefore,
+the solution is determined by means of Picard-type iterations, until convergence of the velocity field \citep{Glerum2018}. The convergence is verified at each
+iteration $i$ via the nonlinear residual $\bm{\mathcal{R}}^i$ that can be determined as
 \begin{equation}\label{eq:convergence}
@@
 %Iterations are performed until either convergence ($\chi<tol$) or maximal number of non-linear iterations ($it_{max}$) is reached.
-where $\mathbb{K}$ is the Stiffness matrix and $\bm{f}$ is the right hand side vector \citep{Spiegelman2016,Glerum2018}.
-The $L_2$-norm is extracted from $\bm{\mathcal{R}}^i$ and it is normalised using the $L_2$-norm extracted from $\bm{f}^0$.
-Iterations are performed until either convergence ($\frac{||\bm{\mathcal{R}}^i||_2}{||\bm{f}^0||_2}<tol$) or maximal number of non-linear iterations ($it_{max}$) is reached.
+where $\mathbb{K}$ is the Stiffness matrix and $\bm{f}$ is the right hand side vector \citep{Spiegelman2016,Glerum2018}.  The $L_2$-norm is extracted from
+$\bm{\mathcal{R}}^i$ and it is normalised using the $L_2$-norm extracted from $\bm{f}^0$. Iterations are performed until either convergence
+($\frac{||\bm{\mathcal{R}}^i||_2}{||\bm{f}^0||_2}<tol$) or maximal number of non-linear iterations ($it_{max}$) is reached.
 
 \section{The energy equation}\label{sec:en_eq}
-The stabilisation of the advection term of the energy equation (Eq.
-\ref{eq:energy}) needed to avoid possible oscillations in the thermal solution in the case when advection dominates over diffusion, is implemented by means of a streamline-upwind Petrov–Galerkin (SUPG) method, for which the advection term is modified follows the discussion in \citet{Thieulot2011} and \citet{Thieulot2014} (see Section \ref{sec:advection} for benchmark).
-The total internal heat production $H_{tot}$ in Eq.
-\ref{eq:energy} includes radiogenic heating $H_r$, shear heating $H_s$ and adiabatic heating $H_a$, where
+The stabilisation of the advection term of the energy equation (Eq. \ref{eq:energy}) needed to avoid possible oscillations in the thermal solution in the case
+when advection dominates over diffusion, is implemented by means of a streamline-upwind Petrov–Galerkin (SUPG) method, for which the advection term is modified
+follows the discussion in \citet{Thieulot2011} and \citet{Thieulot2014} (see Section \ref{sec:advection} for benchmark). The total internal heat production
+$H_{tot}$ in Eq. \ref{eq:energy} includes radiogenic heating $H_r$, shear heating $H_s$ and adiabatic heating $H_a$, where
 \begin{equation}\label{eq:shear}
@@
 \end{equation}
-The correctness of the implementation and the computation of shear and adiabatic heating is verified performing an exercise with an analytical solution (Section \ref{sec:simple_shear}) and Exercise 9.4 in \citet{Gerya2010b} (Section \ref{sec:shear}).
-Furthermore, two benchmarks are performed to verity that the code solve correctly Eqs.
-\ref{eq:momentum_penalty} and \ref{eq:energy} in case of a temperature field characterised by an initial perturbation and with a temperature-dependent density.
+The correctness of the implementation and the computation of shear and adiabatic heating is verified performing an exercise with an analytical solution
+(Section \ref{sec:simple_shear}) and Exercise 9.4 in \citet{Gerya2010b} (Section \ref{sec:shear}). Furthermore, two benchmarks are performed to verity that
+the code solve correctly Eqs. \ref{eq:momentum_penalty} and \ref{eq:energy} in case of a temperature field characterised by an initial perturbation and with
+a temperature-dependent density.
 
 \section{Phase transitions and hydration}\label{sec:phase}
-Variations of effective density, specific heat and coefficient of thermal expansion during the evolution of crustal and mantle materials at different pressure-temperature ($p$-$T$) conditions are computed by means of version 6.8.6 of Perple\_X software package \citep{Connolly2005}, similarly to the implementation described in \citet{Marotta2020}.
-By default, mineral assemblages and properties of each lithology (oceanic crust, lower and upper continental crust, sediments and dry or hydrated mantle) are calculated for temperatures from \SIrange{330}{1600}{\kelvin}, with increments of approximately \SI{4}{\kelvin}, and pressures from \SIrange{0.1}{30}{\giga\pascal}, with increments of approximately \SI{0.03}{\giga\pascal}, for a total of almost 100000 points.
+Variations of effective density, specific heat and coefficient of thermal expansion during the evolution of crustal and mantle materials at different
+pressure-temperature ($p$-$T$) conditions are computed by means of version 6.8.6 of Perple\_X software package \citep{Connolly2005}, similarly to the
+implementation described in \citet{Marotta2020}. By default, mineral assemblages and properties of each lithology (oceanic crust, lower and upper continental
+crust, sediments and dry or hydrated mantle) are calculated for temperatures from \SIrange{330}{1600}{\kelvin}, with increments of approximately \SI{4}{\kelvin},
+and pressures from \SIrange{0.1}{30}{\giga\pascal}, with increments of approximately \SI{0.03}{\giga\pascal}, for a total of almost 100000 points.
 
-The introduction of phase transitions produces two main effects that must be taken into account in numerical models: variations in density and release/absorption of latent heat \citep{Tackley2010}.
-Effects of density variations on buoyancy force in the momentum equation (Eq.
-\ref{eq:momentum_penalty}) are taken into account considering effective coefficient of thermal expansion \citep{Christensen1985,Zhong2015}.
-Similarly, effects of latent heat in the energy equation (eq \ref{eq:energy}) are taken into account by considering effective specific heat and coefficient of thermal expansion \citep{Christensen1985,Tackley2010,Zhong2015}.
-Effective density, specific heat and coefficient of thermal expansion are included in Perple\_X files and are assumed by each Lagrangian marker in according to its $p$-$T$ conditions.
+The introduction of phase transitions produces two main effects that must be taken into account in numerical models: variations in density and release/absorption
+of latent heat \citep{Tackley2010}. Effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into
+account considering effective coefficient of thermal expansion \citep{Christensen1985,Zhong2015}. Similarly, effects of latent heat in the energy equation
+(eq \ref{eq:energy}) are taken into account by considering effective specific heat and coefficient of thermal expansion
+\citep{Christensen1985,Tackley2010,Zhong2015}. Effective density, specific heat and coefficient of thermal expansion are included in Perple\_X files and are
+assumed by each Lagrangian marker in according to its $p$-$T$ conditions.
 
-Hydration processes related to dehydration of subducting lithosphere \citep{Schmidt1998,Liu2007,Faccenda2009,Faccenda2010,Faccenda2014,Rosas2016} strongly influence the thermo-mechanics inside the mantle wedge, mainly because of both weakening effects on the mantle rheology and density variations in case of mantle serpentinization \citep{Gerya2002,Honda2003,Arcay2005,Roda2010,Regorda2017}.
-In the case that hydration is switched on, the amount of bound and free water is memorised by each marker, following the implementation of \citet{Quinquis2014}; hydration and dehydration processes are related to the amount of bound water of each marker and to the maximum amount of water it can transport, i.e.
-if the amount of bound water exceeds the maximum amount of water, the marker dehydrates and releases free water that can hydrates under-saturated markers.
-Maximum water content of each marker is determined as function of lithology and $p$-$T$ conditions and it is calculated using Perple\_X, in the same way than effective density, specific heat and coefficient of thermal expansion.
-Bound water is advected together with the markers, neglecting the effect of bound water diffusion, while free water simply migrates vertically and is not coupled to the solid-phase flow of the mantle wedge \citep{Arcay2005,Quinquis2014}.
-The correctness of the migration scheme of the free water and its absorption into bound water by under-saturated markers are verified performing the experiment described in \citet{Quinquis2014} (Sec.
-\ref{sec:quinquis}).
+Hydration processes related to dehydration of subducting lithosphere \citep{Schmidt1998,Liu2007,Faccenda2009,Faccenda2010,Faccenda2014,Rosas2016} strongly
+influence the thermo-mechanics inside the mantle wedge, mainly because of both weakening effects on the mantle rheology and density variations in case of
+mantle serpentinization \citep{Gerya2002,Honda2003,Arcay2005,Roda2010,Regorda2017}. In the case that hydration is switched on, the amount of bound and free
+water is memorised by each marker, following the implementation of \citet{Quinquis2014}; hydration and dehydration processes are related to the amount of bound
+water of each marker and to the maximum amount of water it can transport, i.e. if the amount of bound water exceeds the maximum amount of water, the marker
+dehydrates and releases free water that can hydrates under-saturated markers. Maximum water content of each marker is determined as function of lithology and
+$p$-$T$ conditions and it is calculated using Perple\_X, in the same way than effective density, specific heat and coefficient of thermal expansion. Bound water
+is advected together with the markers, neglecting the effect of bound water diffusion, while free water simply migrates vertically and is not coupled to the
+solid-phase flow of the mantle wedge \citep{Arcay2005,Quinquis2014}. The correctness of the migration scheme of the free water and its absorption into bound
+water by under-saturated markers are verified performing the experiment described in \citet{Quinquis2014} (Sec. \ref{sec:quinquis}).
 
-The mantle viscosity weakening related to the amount of water has been extensively studied \citep{Chopra1981,Hirth2003} and it is implemented in according to \citet{Arcay2005} and \citet{Horiuchi2016}, as follows
+The mantle viscosity weakening related to the amount of water has been extensively studied \citep{Chopra1981,Hirth2003} and it is implemented in according to
+\citet{Arcay2005} and \citet{Horiuchi2016}, as follows
 \begin{equation}\label{eq:wet_visc}
@@
 \end{equation}
-where $f_v$ is the viscosity reduction factor between dry and wet conditions, $[OH^-]$ is the water content and $[OH^-]_0$ is a reference water content set to 620 ppm (0.062 wt.\%) \citep{Arcay2005}.
-By default, $f_v$ is set to 100, in according to \citet{Arcay2005} and \citet{Horiuchi2016}.
+where $f_v$ is the viscosity reduction factor between dry and wet conditions, $[OH^-]$ is the water content and $[OH^-]_0$ is a reference water content set to
+620 ppm (0.062 wt.\%) \citep{Arcay2005}. By default, $f_v$ is set to 100, in according to \citet{Arcay2005} and \citet{Horiuchi2016}.
 
@@
 Melt fraction in the mantle depends on temperature ($T$, in °C), pressure ($p$, in \si{\giga\pascal}) and the water content in the melt ($X_{H_2O}$, in wt.\%).
-The determination of the melt fraction $M$ is implemented as explained by \citet{Katz2003}, according to the successive modification by \citet{Langmuir2006} and \citet{Kelley2010}, as follows
+The determination of the melt fraction $M$ is implemented as explained by \citet{Katz2003}, according to the successive modification by \citet{Langmuir2006} and
+\citet{Kelley2010}, as follows
 \begin{eqnarray}\label{eq:melt_fr}
@@
 \end{eqnarray}
-where $T_{sol}$ is the temperature of the dry solidus, $D_{H_2O}$ is the partition coefficient and $x_t\cdot log(p+y_t)$ indicates the pressure dependence of the melting curve \citep{Kelley2010}.
-$x_t$, $y_t$, $K_m$ and $\gamma$ are parameters chosen in according to \citet{Kelley2010}.
+where $T_{sol}$ is the temperature of the dry solidus, $D_{H_2O}$ is the partition coefficient and $x_t\cdot log(p+y_t)$ indicates the pressure dependence of
+the melting curve \citep{Kelley2010}. $x_t$, $y_t$, $K_m$ and $\gamma$ are parameters chosen in according to \citet{Kelley2010}.
 However, $X_{H_2O}$ depends on the melt fraction as
@@
 \end{equation}
-and a numerical solution can be found using a root-finder method \citep{Katz2003,Wang2016}.
-Furthermore, the content of water in the melt is limited by the pressure-dependent saturation concentration of water in the melt, determined as
+and a numerical solution can be found using a root-finder method \citep{Katz2003,Wang2016}. Furthermore, the content of water in the melt is limited by the
+pressure-dependent saturation concentration of water in the melt, determined as
 \begin{equation}\label{eq:melt_max}
@@
 \end{equation}
-where $\chi_1$, $\chi_2$ and $\lambda_m$ are parameters chosen according to \citet{Katz2003}, \citet{Langmuir2006} and \citet{Kelley2010}.
-The correctness in the determination of melt fractions and water contents in the melt are tested comparing model results with experimental curves obtained by \citet{Katz2003} (Sec.
-\ref{sec:katz}).
+where $\chi_1$, $\chi_2$ and $\lambda_m$ are parameters chosen according to \citet{Katz2003}, \citet{Langmuir2006} and \citet{Kelley2010}. The correctness in
+the determination of melt fractions and water contents in the melt are tested comparing model results with experimental curves obtained by \citet{Katz2003}
+(Sec. \ref{sec:katz}).
 
@@
 \end{equation}
-where $\rho_s$ and $\rho_m$ are the densities of the solid and the molten rock, respectively \citep{Gerya2003a,Wang2016}.
-The density of the solid rock is extracted by Perple\_X, while the density of the molten portion is calculated as
+where $\rho_s$ and $\rho_m$ are the densities of the solid and the molten rock, respectively \citep{Gerya2003a,Wang2016}. The density of the solid rock is
+extracted by Perple\_X, while the density of the molten portion is calculated as
 \begin{equation}\label{eq:melt_dens_M}
@@
 \end{equation}
-where $\rho_0$ is the density at the reference temperature ($T_0$) and pressure($p_0$), and $\alpha$ and $\beta$ are the thermal expansion and compressibility coefficients, respectively \citep{Gerya2003a,Wang2016}.
+where $\rho_0$ is the density at the reference temperature ($T_0$) and pressure($p_0$), and $\alpha$ and $\beta$ are the thermal expansion and compressibility
+coefficients, respectively \citep{Gerya2003a,Wang2016}.
 
-As for the phase transitions, effects of density variations on buoyancy force in the momentum equation (Eq.
-\ref{eq:momentum_penalty}) are taken into account considering effective coefficient of thermal expansion.
-Similarly, latent heat due to melting/crystallisation can be implicitly included in the energy equation (Eq.
-\ref{eq:energy}) by considering effective specific heat ($C_{p_{eff}}$) and thermal expansion ($\alpha_{eff}$) of the partially molten rocks, in according to
+As for the phase transitions, effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into account
+considering effective coefficient of thermal expansion. Similarly, latent heat due to melting/crystallisation can be implicitly included in the energy equation
+(Eq. \ref{eq:energy}) by considering effective specific heat ($C_{p_{eff}}$) and thermal expansion ($\alpha_{eff}$) of the partially molten rocks, in according to
 \begin{eqnarray}
@@
 \end{eqnarray}
-where $C_p$ and $\alpha$ are the specific heat and the coefficient of thermal expansion of the solid rock, respectively, and $H_L$ is the latent heat of the molten rock \citep{Gerya2003a,Gerya2007,Tackley2010}.
+where $C_p$ and $\alpha$ are the specific heat and the coefficient of thermal expansion of the solid rock, respectively, and $H_L$ is the latent heat of the
+molten rock \citep{Gerya2003a,Gerya2007,Tackley2010}.
 
@@
 \subsection{Error measurements}\label{sec:error}
-In order to determine the accuracy of velocity and pressure field of the benchmarks, the $L_2$ norm is computed by numerical integration on the Gauss-Legendre quadrature points.
-$L_2$ norm for pressure and velocity errors can be evaluated as
+In order to determine the accuracy of velocity and pressure field of the benchmarks, the $L_2$-norm is computed by numerical integration on the Gauss-Legendre
+quadrature points. $L_2$-norm for pressure and velocity errors can be evaluated as
 \begin{eqnarray}
@@
 \end{eqnarray}
-respectively, where $n_e$ is the number of elements, $n_q$ is the number of quadrature points per element, $p^n_i(r_q)$ and $p^a_i(r_q)$ are the numerical and analytical pressure, respectively, in each quadrature point $q$, $w_q$  and $J_q$ are the weight and the Jacobian at the quadrature point $q$, $u^n_i(r_q)$, $v^n_i(r_q)$, $u^a_i(r_q)$ and $v^a_i(r_q)$ are the numerical and analytical velocities, respectively, in each quadrature point $q$.
+respectively, where $n_e$ is the number of elements, $n_q$ is the number of quadrature points per element, $p^n_i(r_q)$ and $p^a_i(r_q)$ are the numerical and
+analytical pressure, respectively, in each quadrature point $q$, $w_q$  and $J_q$ are the weight and the Jacobian at the quadrature point
+$q$, $u^n_i(r_q)$, $v^n_i(r_q)$, $u^a_i(r_q)$ and $v^a_i(r_q)$ are the numerical and analytical velocities, respectively, in each quadrature point $q$.
 
@@
 \end{eqnarray}
-The domain is a unit square a constant viscosity ($\eta=1$) and the penalty parameter is set to $\lambda=10^7$.
-Velocity boundary conditions are set to no slip $(\bm{v}=\bm{0})$ on all boundaries.
-The problem is performed for different grid resolution between \num{8x8} and \num{1024x1024} elements.
+The domain is a unit square a constant viscosity ($\eta=1$) and the penalty parameter is set to $\lambda=10^7$. Velocity boundary conditions are set to no slip
+$(\bm{v}=\bm{0})$ on all boundaries. The problem is performed for different grid resolution between \num{8x8} and \num{1024x1024} elements.
 
 The errors between the analytical solution and the numerical prediction of the velocity and the pressure field are calculated as in Section \ref{sec:error}.
-Fig.
-\ref{fig:errors}a shows that both velocity and pressure field converge to the exact solution with the decrease of the element size, following the theoretical convergence rate.
-The convergence of pressure, in contrast with the observation by \citet{Donea2003} for $Q_1 \times P_0$ elements, indicate the effectiveness of the smoothing procedure.
-Smoothed pressure field for grid resolution of \num{128x128} elements is shown in Fig.
-\ref{fig:errors}b (continuous red line), in comparison with the analytical solution (dashed blue line).
-Solve times and memory usage needed to generate the solution are shown in Fig.
-\ref{fig:MUMPS}.
+Fig. \ref{fig:errors}a shows that both velocity and pressure field converge to the exact solution with the decrease of the element size, following the
+theoretical convergence rate. The convergence of pressure, in contrast with the observation by \citet{Donea2003} for $Q_1 \times P_0$ elements, indicate the
+effectiveness of the smoothing procedure. Smoothed pressure field for grid resolution of \num{128x128} elements is shown in Fig. \ref{fig:errors}b (continuous
+red line), in comparison with the analytical solution (dashed blue line). Solve times and memory usage needed to generate the solution are shown in
+Fig. \ref{fig:MUMPS}.
 
@@
 \noindent\includegraphics[width=\textwidth]{./Figures/errors.pdf}
-\caption{Velocity and pressure error for the Stokes flow experiment between generated and analytical solution as a function of element size (panel a), and comparison between smoothed pressure and analytical solution as function of $x$ coordinate for a grid resolution of \num{128x128} elements (panel b).}
+\caption{Velocity and pressure error for the Stokes flow experiment between generated and analytical solution as a function of element size (panel a),
+and comparison between smoothed pressure and analytical solution as function of $x$ coordinate for a grid resolution of \num{128x128} elements (panel b).}
 \label{fig:errors}
@@
 \noindent\includegraphics[width=\textwidth]{./Figures/MUMPS.pdf}
-\caption{Panel a) Analysis, factorisation, solution and total solve times (panel a) and factorisation memory usage (panel b) as a function of the total number of degrees of freedom for the Stokes flow experiment.}
+\caption{Panel a) Analysis, factorisation, solution and total solve times (panel a) and factorisation memory usage (panel b) as a function of the total number
+of degrees of freedom for the Stokes flow experiment.}
 \label{fig:MUMPS}
@@
 \subsection{Markers advection}\label{sec:runge}
-The 4th-order Runge-Kutta advection scheme in space is tested by means of the Zalesak disk test \citep{Zalesak1979,Thieulot2014}.
-The benchmark is performed in a unit square domain with different grid resolution (\num{100x100} and \num{200x200} elements) and 1000000 markers.
-The Courant number varies between 0.1 and 0.5 and the velocity field is prescribed in the entire domain as
+The 4th-order Runge-Kutta advection scheme in space is tested by means of the Zalesak disk test \citep{Zalesak1979,Thieulot2014}. The benchmark is performed
+in a unit square domain with different grid resolution (\num{100x100} and \num{200x200} elements) and 1000000 markers. The Courant number varies between 0.1
+and 0.5 and the velocity field is prescribed in the entire domain as
 \begin{eqnarray}
@@
 \end{eqnarray}
-At $t=0$, the disk is centred at position $(0.5;0.75)$ with a radius $R=0.15$ and has a vertical fissure 0.05 wide and 0.2 high (orange in Fig.
-\ref{fig:runge}).
-As we expect, markers are back to their initial location after a $2\pi$ rotation ($t=1$, red, green and blue in Fig.
-\ref{fig:runge}), with an error of the final position that increase moving away from the centre, more evident in case of higher time steps or lower grid resolution (Fig.
-\ref{fig:runge_err}).
+At $t=0$, the disk is centred at position $(0.5;0.75)$ with a radius $R=0.15$ and has a vertical fissure 0.05 wide and 0.2 high (orange in Fig. \ref{fig:runge}).
+As we expect, markers are back to their initial location after a $2\pi$ rotation ($t=1$, red, green and blue in Fig. \ref{fig:runge}), with an error of the
+final position that increase moving away from the centre, more evident in case of higher time steps or lower grid resolution (Fig. \ref{fig:runge_err}).
 
@@
 \noindent\includegraphics[width=200px]{./Figures/Zalesak.pdf}
-\caption{Comparison between the initial position of markers of the Zalesak disk experiment (orange) and their final position for different CFL in case of grid resolution of \num{100x100} elements.
-Blue, green and red markers indicate the final position of the disk in case of Courant number of 0.1, 0.25 and 0.5, respectively.}
+\caption{Comparison between the initial position of markers of the Zalesak disk experiment (orange) and their final position for different CFL in case of grid
+resolution of \num{100x100} elements. Blue, green and red markers indicate the final position of the disk in case of Courant number of 0.1, 0.25 and 0.5,
+respectively.}
 \label{fig:runge}
@@
 \noindent\includegraphics[width=400px]{./Figures/zal_error.pdf}
-\caption{Error between final and initial position of markers of the disk as function of their distance from the centre of the domain.
-Results for different time steps are shown in blue, red and green for a grid resolution of \num{100x100} elements (for CFL of 0.1, 0.25 and 0.5, respectively) and in light blue, dark green and purple in case of \num{200x200} elements (for CFL of 0.1, 0.25 and 0.5, respectively).}
+\caption{Error between final and initial position of markers of the disk as function of their distance from the centre of the domain. Results for different
+time steps are shown in blue, red and green for a grid resolution of \num{100x100} elements (for CFL of 0.1, 0.25 and 0.5, respectively) and in light blue,
+dark green and purple in case of \num{200x200} elements (for CFL of 0.1, 0.25 and 0.5, respectively).}
 \label{fig:runge_err}
@@
 \subsection{Poiseuille flow}\label{sec:poiseuille}
-The domain is a rectangle with $L_x=2$ and $L_y=1$ and constant density and viscosity ($\rho=1$ and $\eta=1$), gravity acceleration $\bm{g}=\bm{0}$ and penalty parameter $\lambda=10^8$.
-The grid is composed by \num{40x20} elements.
-Velocity boundary conditions are set to no slip $(\bm{v}=\bm{0})$ at the top and the bottom, and a parabolic profile is imposed on the sides, with $u=y(1-y)$ and $v=0$.
-The analytical solution is then given by:
+The domain is a rectangle with $L_x=2$ and $L_y=1$ and constant density and viscosity ($\rho=1$ and $\eta=1$), gravity acceleration $\bm{g}=\bm{0}$ and penalty
+parameter $\lambda=10^8$. The grid is composed by \num{40x20} elements. Velocity boundary conditions are set to no slip $(\bm{v}=\bm{0})$ at the top and the
+bottom, and a parabolic profile is imposed on the sides, with $u=y(1-y)$ and $v=0$. The analytical solution is then given by:
 \begin{eqnarray}
@@
 
-The velocity field predicted by the model follows the expected parabolic profile (Fig.
-\ref{fig:poiseuille}a and Fig.
-\ref{fig:poi_plot}a) and, in case of the classic penalty method with no iterations, the pressure is clearly related to the divergence by means of the penalty parameter (Fig.
-\ref{fig:poiseuille}b and c and Fig.
-\ref{fig:poi_plot}b).
-Fig.
-\ref{fig:poiseuille}d shows that one Uzawa iteration is sufficient to bring the divergence down to \num{e-15}, with no correlation with the pressure field.
+The velocity field predicted by the model follows the expected parabolic profile (Fig. \ref{fig:poiseuille}a and Fig. \ref{fig:poi_plot}a) and, in case of the
+classic penalty method with no iterations, the pressure is clearly related to the divergence by means of the penalty parameter (Fig. \ref{fig:poiseuille}b and
+c and Fig. \ref{fig:poi_plot}b). Fig. \ref{fig:poiseuille}d shows that one Uzawa iteration is sufficient to bring the divergence down to \num{e-15}, with no
+correlation with the pressure field.
 
@@
 \noindent\includegraphics[width=\textwidth]{./Figures/Poiseuille.pdf}
-\caption{Velocity field (panel a), pressure (panel b) and divergence velocity of a Poiseuille flow in case of the classic penalty method (no iterations) and after one Uzawa iteration (panel c and d, respectively).}
+\caption{Velocity field (panel a), pressure (panel b) and divergence velocity of a Poiseuille flow in case of the classic penalty method (no iterations) and
+after one Uzawa iteration (panel c and d, respectively).}
 \label{fig:poiseuille}
@@
 \noindent\includegraphics[width=\textwidth]{./Figures/Analytical.pdf}
-\caption{Velocity (panel a) and pressure (panel b) field predicted by the model for a Poiseuille flow with respect with their analytical solutions.
-The velocity field is plotted as function of the vertical coordinate in $L_x/2$ and the pressure is plotted as function of the x coordinate in $L_y/2$.}
+\caption{Velocity (panel a) and pressure (panel b) field predicted by the model for a Poiseuille flow with respect with their analytical solutions. The
+velocity field is plotted as function of the vertical coordinate in $L_x/2$ and the pressure is plotted as function of the x coordinate in $L_y/2$.}
 \label{fig:poi_plot}
@@
 \subsection{Instantaneous 2D sphere}\label{sec:ist_sphere}
-The domain is a unit square with gravity $\bm{g}=(0,-1)$.
-The fluid has constant density and viscosity ($\rho_f=1$ and $\eta_f=1$).
-The sphere is in the middle of the domain with a radius R = 0.123456798 and has constant density and viscosity ($\rho_s=10^{-2}+\rho_f$ and $\eta_s=10^3 \cdot \eta_f$).
+The domain is a unit square with gravity $\bm{g}=(0,-1)$. The fluid has constant density and viscosity ($\rho_f=1$ and $\eta_f=1$). The sphere is in the middle
+of the domain with a radius R = 0.123456798 and has constant density and viscosity ($\rho_s=10^{-2}+\rho_f$ and $\eta_s=10^3 \cdot \eta_f$).
 We distinguish three different types of velocity boundary conditions:
@@
 \end{enumerate}
-All three cases are performed for different grid resolutions, between \num{16x16} and \num{512x512} elements with 50 randomly distributed markers per element, and different average schemes for the viscosity (harmonic, geometric and arithmetic).
-The velocity in the centre of the sphere $\bm{v}(0.5;0.5)$), minimum and maximum velocities ($u_{min}$, $u_{max}$, $v_{min}$, $v_{max}$) and pressures ($p_{min}$, $p_{max}$) on the entire domain, $v_{rms}$ and average pressure on the entire domain ($p_{avg}$) are compared with solutions generated by ASPECT \citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} and all results can be found at \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere2D}.
-Results in terms of the $v_{rms}$ for NS model are also shown in Fig.
-\ref{fig:inst_sphere}.
+All three cases are performed for different grid resolutions, between \num{16x16} and \num{512x512} elements with 50 randomly distributed markers per element,
+and different average schemes for the viscosity (harmonic, geometric and arithmetic). The velocity in the centre of the sphere $\bm{v}(0.5;0.5)$), minimum and
+maximum velocities ($u_{min}$, $u_{max}$, $v_{min}$, $v_{max}$) and pressures ($p_{min}$, $p_{max}$) on the entire domain, $v_{rms}$ and average pressure on the
+entire domain ($p_{avg}$) are compared with solutions generated by ASPECT \citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} and all results
+can be found at \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere2D}. Results in terms of the $v_{rms}$ for NS model are also shown
+in Fig. \ref{fig:inst_sphere}.
 
@@
 \noindent\includegraphics[width=400px]{./Figures/vrms_NS.pdf}
-\caption{$v_{rms}$ as function of element size for the instantaneous 2D sphere experiment in case of no slip boundary conditions and with different average schemes for the viscosity.
-Results are compared with results obtained by other numerical codes.}
+\caption{$v_{rms}$ as function of element size for the instantaneous 2D sphere experiment in case of no slip boundary conditions and with different average
+schemes for the viscosity. Results are compared with results obtained by other numerical codes.}
 \label{fig:inst_sphere}
@@
 \subsection{Rayleigh-Taylor instability}\label{sec:rayleigh}
-This problem has been originally presented by \citet{vanKeken1997} and here it is performed as the isoviscous case (Case 1a).
-The domain has $L_x=0.9142$ and $L_y=1$ and gravity $\bm{g}=(0,-1)$.
-Two fluids with same constant viscosities ($\eta_1=\eta_2=1$) and different densities ($\rho_1=1000$ and $\rho_2=1010$), with the lighter fluid at the bottom.
-The initial interface between the fluids is given by $y(x)=0.2+0.02 \cos \left(\frac{\pi x}{L_x}\right)$.
-The experiment is performed with different grid sizes (\num{50x50}, \num{80x80}, \num{100x100} and \num{256x256}).
-A total of 1960000 markers are randomly distributed at the beginning of the simulation.
-Velocity boundary conditions are set to no slip at the top and the bottom, and to free slip at the sides of the domain.
+This problem has been originally presented by \citet{vanKeken1997} and here it is performed as the isoviscous case (Case 1a). The domain has $L_x=0.9142$ and
+$L_y=1$ and gravity $\bm{g}=(0,-1)$. Two fluids with same constant viscosities ($\eta_1=\eta_2=1$) and different densities ($\rho_1=1000$ and $\rho_2=1010$),
+with the lighter fluid at the bottom. The initial interface between the fluids is given by $y(x)=0.2+0.02 \cos \left(\frac{\pi x}{L_x}\right)$. The experiment
+is performed with different grid sizes (\num{50x50}, \num{80x80}, \num{100x100} and \num{256x256}). A total of 1960000 markers are randomly distributed at the
+beginning of the simulation. Velocity boundary conditions are set to no slip at the top and the bottom, and to free slip at the sides of the domain.
 
-$v_{rms}$ as function of time is reported in Fig.
-\ref{fig:RT}, matching well with results shown by \citet{vanKeken1997}, \citet{Tackley2003} and \citet{Thieulot2014}.
-Fig.
-\ref{fig:rayleigh} shows the evolution of the experiment at different time steps.
+$v_{rms}$ as function of time is reported in Fig. \ref{fig:RT}, matching well with results shown by \citet{vanKeken1997}, \citet{Tackley2003} and
+\citet{Thieulot2014}. Fig. \ref{fig:rayleigh} shows the evolution of the experiment at different time steps.
 
@@
 
-\subsection{Falling block}\label{sec:block}
-This benchmark is described as proposed by \citet{Gerya2003a}, \citet{Gerya2010b} and \citet{Thieulot2011}.
-The domain is square with $L_x=L_y=$ \SI{500}{\km} and the grid is composed by \num{50x50} elements with 25 markers in each element.
-The block is initially centred at ($x=$ \SI{250}{\km}; $y=$ \SI{400}{\km}) and has a size of \num{100x100} km (Fig.
-\ref{fig:block}a).
-The fluid surrounding the block has $\eta_f=$ \SI{e21}{\pascal\per\s} and $\rho_f=$ \SI{3200}{\kg\per\cubic\m}.
-The benchmark tests with different viscosities of the block, with $\eta_b$ from \SIrange{e15}{e27}{\pascal\per\s}.
-In each experiment also the density $\rho_b$ of the block varies from \SIrange{3220}{9900}{\kg\per\cubic\m}.
-Velocity boundary conditions are set to free slip conditions on all sides of the domain.
-Density distributions at $t=$ \SI{20}{\mega\year} for $\rho_b=$ \SI{3300}{\kg\per\cubic\m} and different $\eta_b$ are plotted in Fig.
-\ref{fig:block}b-f, showing that the code correctly preserve the block geometry in case of large viscosity contrast, with a block stiffer than the surrounding fluid (Fig.
-\ref{fig:block}f).
+\subsection{Falling blocks}\label{sec:block}
+This benchmark is described as proposed by \citet{Gerya2003a}, \citet{Gerya2010b} and \citet{Thieulot2011}. The domain is square with $L_x=L_y=$ \SI{500}{\km}
+and the grid is composed by \num{50x50} elements with 25 markers in each element. The block is initially centred at ($x=$ \SI{250}{\km}; $y=$ \SI{400}{\km})
+and has a size of \num{100x100} km (Fig. \ref{fig:block}a). The fluid surrounding the block has $\eta_f=$ \SI{e21}{\pascal\per\s} and $\rho_f=$
+\SI{3200}{\kg\per\cubic\m}. The benchmark tests with different viscosities of the block, with $\eta_b$ from \SIrange{e15}{e27}{\pascal\per\s}. In each
+experiment also the density $\rho_b$ of the block varies from \SIrange{3220}{9900}{\kg\per\cubic\m}. Velocity boundary conditions are set to free slip
+conditions on all sides of the domain. Density distributions at $t=$ \SI{20}{\mega\year} for $\rho_b=$ \SI{3300}{\kg\per\cubic\m} and different $\eta_b$
+are plotted in Fig. \ref{fig:block}b-f, showing that the code correctly preserve the block geometry in case of large viscosity contrast, with a block stiffer
+than the surrounding fluid (Fig. \ref{fig:block}f).
 
-The velocity in the centre of the falling block at $t=0$ is measured for all experiments and, since the velocity should increase with density contrast, the quantity $\bm{v}/(\rho_b-\rho_f)$ is plotted as function of the viscosity contrast.
-All results perfectly match with those from \citet{Gerya2010b} and they line up on a single curve, demonstrating that the code can correctly deal with large viscosity and density contrasts (Fig.
-\ref{fig:falling}).
+The velocity in the centre of the falling block at $t=0$ is measured for all experiments and, since the velocity should increase with density contrast, the
+quantity $\bm{v}/(\rho_b-\rho_f)$ is plotted as function of the viscosity contrast. All results perfectly match with those from \citet{Gerya2010b} and they
+line up on a single curve, demonstrating that the code can correctly deal with large viscosity and density contrasts (Fig. \ref{fig:falling}).
 
@@
 \centering
-\caption{Density evolution of the falling block experiment at $t=$ \SI{0}{\mega\year} (panel a) and $t=$ \SI{20}{\mega\year} for different viscosities of the block (panels b-f).}
+\caption{Density evolution of the falling block experiment at $t=$ \SI{0}{\mega\year} (panel a) and $t=$ \SI{20}{\mega\year} for different viscosities of the
+block (panels b-f).}
 \label{fig:block}
@@
 \noindent\includegraphics[width=400px]{./Figures/Falling.pdf}
-\caption{Initial velocity relative to the density contrast at the centre of the falling block as function of the viscosity contrast between the block and the surrounding fluid.}
+\caption{Initial velocity relative to the density contrast at the centre of the falling block as function of the viscosity contrast between the block and the
+surrounding fluid.}
 \label{fig:falling}
@@
 \subsection{2D time-dependent Stokes sphere with deformable free surface}\label{sec:sphere}
-This experiment is performed in a unit square domain with the gravity acceleration fixed to $g_y=-1$.
-The sphere has $\rho_s=2$ and $\eta_s=10^3$ and is initially centred at ($0.5;0.6$) with radius $R=0.123456789$.
-The fluid surrounding the sphere has $\rho_f=1$ and $\eta_f=1$ and occupies the domain for $y\leq 0.75$, while for $y>0.75$ the air has $\rho_a=0$ and $\eta_a=10^{-3}$.
-Velocity boundary conditions are set to free slip on all sides.
-The Courant number is set to 0.25.
-The experiments run for 200 s using grid resolutions from \num{150x150} to \num{512x512} elements, with 25 randomly distributed markers per element, and for different average schemes for the viscosity.
-The interface between the fluid and the air is tracked by means of the markers chain.
-Results in terms of velocity and pressure fields and topography variations are compared with results obtained with ASPECT \citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} and they can be found at \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere_fs2D}.
-Results in terms of the $v_{rms}$ in case of an arithmetic average are also shown in Fig.
-\ref{fig:inst_spherefs}.
+This experiment is performed in a unit square domain with the gravity acceleration fixed to $g_y=-1$. The sphere has $\rho_s=2$ and $\eta_s=10^3$ and is
+initially centred at ($0.5;0.6$) with radius $R=0.123456789$. The fluid surrounding the sphere has $\rho_f=1$ and $\eta_f=1$ and occupies the domain for
+$y\leq 0.75$, while for $y>0.75$ the air has $\rho_a=0$ and $\eta_a=10^{-3}$. Velocity boundary conditions are set to free slip on all sides. The Courant
+number is set to 0.25. The experiments run for 200 s using grid resolutions from \num{150x150} to \num{512x512} elements, with 25 randomly distributed markers
+per element, and for different average schemes for the viscosity. The interface between the fluid and the air is tracked by means of the markers chain. Results
+in terms of velocity and pressure fields and topography variations are compared with results obtained with ASPECT
+\citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} and they can be found at
+\url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere_fs2D}. Results in terms of the $v_{rms}$ in case of an arithmetic average are also
+shown in Fig. \ref{fig:inst_spherefs}.
 
@@
 \subsection{Free surface stabilisation}\label{sec:stab}
-The benchmark is performed as discussed in \citet{Kaus2010a} and \citet{Thieulot2014}.
-The domain is square with $L_x=L_y=$ \SI{500}{\km} and a grid resolution of \num{200x200} elements, each of them containing 16 markers.
-A buoyant fluid with $\rho_1=$ \SI{3200}{\kg\per\cubic\metre} and $\eta_1=$ \SI{e20}{\pascal\s} is overlain by a denser fluid with $\rho_2=$ \SI{3300}{\kg\per\cubic\metre} and $\eta_2=$ \SI{e21}{\pascal\s}.
-The initial interface between the fluids has an initial sinusoidal shape of 5 km amplitude.
-Velocity boundary conditions are set to free slip conditions at the sides and no slip at the bottom, while the top is free surface.
-The experiment is performed with various fixed time steps, without the use of the Courant number.
-The vertical position of the free surface at $x=L_x$ is tracked for each simulation, with or without the stabilisation algorithm.
+The benchmark is performed as discussed in \citet{Kaus2010a} and \citet{Thieulot2014}. The domain is square with $L_x=L_y=$ \SI{500}{\km} and a grid resolution
+of \num{200x200} elements, each of them containing 16 markers. A buoyant fluid with $\rho_1=$ \SI{3200}{\kg\per\cubic\metre} and $\eta_1=$ \SI{e20}{\pascal\s}
+is overlain by a denser fluid with $\rho_2=$ \SI{3300}{\kg\per\cubic\metre} and $\eta_2=$ \SI{e21}{\pascal\s}. The initial interface between the fluids has an
+initial sinusoidal shape of 5 km amplitude. Velocity boundary conditions are set to free slip conditions at the sides and no slip at the bottom, while the top
+is free surface. The experiment is performed with various fixed time steps, without the use of the Courant number. The vertical position of the free surface at
+$x=L_x$ is tracked for each simulation, with or without the stabilisation algorithm.
 
-The results show an instability (drunken sailor effect) increasing the time step in case that the stabilisation algorithm is not activated ($dt=$ \SI{4200}{\year}, green line in Fig.
-\ref{fig:stab}), as already observed by \citet{Kaus2010a} and \citet{Thieulot2014}.
-Activating the algorithm the instability is fixed and the simulation remains stable using time step up to \SI{50000}{\year} (red line in Fig.
-\ref{fig:stab}).
+The results show an instability (drunken sailor effect) increasing the time step in case that the stabilisation algorithm is not activated
+($dt=$ \SI{4200}{\year}, green line in Fig. \ref{fig:stab}), as already observed by \citet{Kaus2010a} and \citet{Thieulot2014}. Activating the algorithm the
+instability is fixed and the simulation remains stable using time step up to \SI{50000}{\year} (red line in Fig. \ref{fig:stab}).
 
@@
 \subsection{Topography relaxation}\label{sec:crameri}
-This experiment is performed as in \citet{Crameri2012} using free surface and sticky air (Case 1).
-The domain is rectangular with $L_x=$ \SI{2800}{\km} and $L_y=$ \SI{700}{\km} in case of free surface and $L_y=$ \SI{800}{\km} in case of sticky air ($h_{st}=$ \SI{100}{\km}).
-The grid is composed by \num{256x64} elements for the free surface \num{560x320} elements for the sticky air.
-For both cases, there is a mantle of 600 km thickness with $\eta_m=$ \SI{e21}{\pascal\s} overlain by a highly viscous cosine-shaped lithosphere with thickness between 93 and 107 km and $\eta_l=$ \SI{e23}{\pascal\s}.
-Both mediums have $\rho=$ \SI{3300}{\kg\per\cubic\metre}.
-In case of sticky air experiment, the air has $\rho_a=$ \SI{0}{\kg\per\cubic\metre} and $\eta_a=$ \SI{e18}{\pascal\s}.
-The gravity acceleration is set to $g_y=$ \SI{-10}{\metre\per\square\s}.
-Velocity boundary conditions are set to no slip at the bottom and free slip at the sides of the domain.
-In case of sticky air, velocities on top are set to free slip conditions.
+This experiment is performed as in \citet{Crameri2012} using free surface and sticky air (Case 1). The domain is rectangular with $L_x=$ \SI{2800}{\km} and
+$L_y=$ \SI{700}{\km} in case of free surface and $L_y=$ \SI{800}{\km} in case of sticky air ($h_{st}=$ \SI{100}{\km}). The grid is composed by \num{256x64}
+elements for the free surface \num{560x320} elements for the sticky air. For both cases, there is a mantle of 600 km thickness with $\eta_m=$ \SI{e21}{\pascal\s}
+overlain by a highly viscous cosine-shaped lithosphere with thickness between 93 and 107 km and $\eta_l=$ \SI{e23}{\pascal\s}. Both mediums have $\rho=$
+\SI{3300}{\kg\per\cubic\metre}. In case of sticky air experiment, the air has $\rho_a=$ \SI{0}{\kg\per\cubic\metre} and $\eta_a=$ \SI{e18}{\pascal\s}. The
+gravity acceleration is set to $g_y=$ \SI{-10}{\metre\per\square\s}. Velocity boundary conditions are set to no slip at the bottom and free slip at the sides
+of the domain. In case of sticky air, velocities on top are set to free slip conditions.
 
@@
 where $h_0=7$ km is the initial topography, $\gamma=\num{-0.2139e-11}$ is the characteristic relaxation rate and $t$ is the time (black line in Fig.
-\ref{fig:crameri}).
-Maximum topography at the characteristic relaxation time $t_{rlx}=$ \SI{14.825}{\kilo\year} can be found to be $h_{rlx}=$ \SI{2576}{\m}.
-The results show that a viscosity of the air of \SI{e18}{\pascal\s} is needed in case of $h_{st}=$ \SI{100}{\km} to correctly describe the topography relaxation for this problem, as pointed out by \citet{Crameri2012} (purple line in Fig.
-\ref{fig:crameri}).
-The case with a true free surface well follows the analytical solution (orange line in Fig.
-\ref{fig:crameri}).
-In particular, at the relaxation time in case of the true free surface predicts a maximum topography of \SI{2570}{\m}, with an error of \SI{6}{\m} with respect to the analytical solution.
-Results are also compared with results obtained with a free surface in UNDERWORLD (\num{256x64} elements) and SULEC (\num{401x201} elements) (with courtesy from \citet{Crameri2012}).
+\ref{fig:crameri}). Maximum topography at the characteristic relaxation time $t_{rlx}=$ \SI{14.825}{\kilo\year} can be found to be $h_{rlx}=$ \SI{2576}{\m}.
+The results show that a viscosity of the air of \SI{e18}{\pascal\s} is needed in case of $h_{st}=$ \SI{100}{\km} to correctly describe the topography relaxation
+for this problem, as pointed out by \citet{Crameri2012} (purple line in Fig. \ref{fig:crameri}). The case with a true free surface well follows the analytical
+solution (orange line in Fig. \ref{fig:crameri}). In particular, at the relaxation time in case of the true free surface predicts a maximum topography of
+\SI{2570}{\m}, with an error of \SI{6}{\m} with respect to the analytical solution. Results are also compared with results obtained with a free surface in
+UNDERWORLD (\num{256x64} elements) and SULEC (\num{401x201} elements) (with courtesy from \citealp{Crameri2012}).
 
@@
 \noindent\includegraphics[width=400px]{./Figures/Crameri.pdf}
-\caption{Maximum topography as function of time for the Crameri benchmark (coloured lines) in comparison with the analytical solution (black line) and results obtained with UNDERWORLD (yellow line) and SULEC (blue line).}
+\caption{Maximum topography as function of time for the Crameri benchmark (coloured lines) in comparison with the analytical solution (black line) and results
+obtained with UNDERWORLD (yellow line) and SULEC (blue line).}
 \label{fig:crameri}
@@
 \subsection{Spontaneous subduction}\label{sec:subduction}
-This experiment is performed as presented by \citet{Schmeling2008}, in case of both a sticky air and a true free surface and using different average schemes for the viscosities.
-The domain is rectangular with $L_x=$ \SI{3000}{\km} and $L_y=$ \SI{700}{\km}, a grid resolution of \num{256x64} elements with 50 markers per element and Courant number of 0.01.
-At the beginning of the simulation, a 100 km-thick lithospheric layer with $\rho_m=$ \SI{3300}{\kg\per\cubic\metre} and $\eta_m=$ \SI{e23}{\pascal\s} is located at the top of the domain for $x$ between 1000 and 3000 km, while the asthenospheric mantle has $\rho_m=$ \SI{3200}{\pascal\s} and $\eta_m=$ \SI{e21}{\pascal\s}.
-In addition, a 200 km-depth lithospheric slab is already subducted in the mantle in order to have a spontaneous subduction.
-In case of sticky air, $L_y=750$ km and a 50 km-thick air layer with $\eta_a$ from \SIrange{e19}{e21}{\pascal\s} overlie the mantle.
-Velocity boundary conditions are set to free slip on the sides and on the bottom of the domain.
-In case of sticky air, velocities on the top boundary are set to free slip conditions as well.
+This experiment is performed as presented by \citet{Schmeling2008}, in case of both a sticky air and a true free surface and using different average schemes
+for the viscosities. The domain is rectangular with $L_x=$ \SI{3000}{\km} and $L_y=$ \SI{700}{\km}, a grid resolution of \num{256x64} elements with 50 markers
+per element and Courant number of 0.01. At the beginning of the simulation, a 100 km-thick lithospheric layer with $\rho_m=$ \SI{3300}{\kg\per\cubic\metre}
+and $\eta_m=$ \SI{e23}{\pascal\s} is located at the top of the domain for $x$ between 1000 and 3000 km, while the asthenospheric mantle has $\rho_m=$
+\SI{3200}{\pascal\s} and $\eta_m=$ \SI{e21}{\pascal\s}. In addition, a 200 km-depth lithospheric slab is already subducted in the mantle in order to have
+a spontaneous subduction. In case of sticky air, $L_y=750$ km and a 50 km-thick air layer with $\eta_a$ from \SIrange{e19}{e21}{\pascal\s} overlie the mantle.
+Velocity boundary conditions are set to free slip on the sides and on the bottom of the domain. In case of sticky air, velocities on the top boundary are set
+to free slip conditions as well.
 
-The maximum depth of the slab is tracked for all the simulations (Fig.
-\ref{fig:subduction}) and they are compared with results obtained by \citet{Schmeling2008} with models with regular grids and comparable grid resolutions (rectangles in Fig.
-\ref{fig:subduction}).
-Low viscous sticky air models (i.e., \SI{e19}{\pascal\s}) enlighten a strong dependence of the sinking velocity with the chosen average scheme (dashed lines in Fig.
-\ref{fig:subduction}), while high viscous sticky air models (i.e., \SI{e21}{\pascal\s}) show a higher resistance at the trench, underestimating the correct solution (dotted lines in Fig.
-\ref{fig:subduction}).
-A high resistance at the trench can be also observed in case of a true free surface for the low resolution grid used (continuous lines in Fig.
-\ref{fig:subduction}), as pointed out by \citet{Schmeling2008}.
+The maximum depth of the slab is tracked for all the simulations (Fig. \ref{fig:subduction}) and they are compared with results obtained by
+\citet{Schmeling2008} with models with regular grids and comparable grid resolutions (rectangles in Fig. \ref{fig:subduction}). Low viscous sticky air models
+(i.e., \SI{e19}{\pascal\s}) enlighten a strong dependence of the sinking velocity with the chosen average scheme (dashed lines in Fig. \ref{fig:subduction}),
+while high viscous sticky air models (i.e., \SI{e21}{\pascal\s}) show a higher resistance at the trench, underestimating the correct solution (dotted lines in
+Fig. \ref{fig:subduction}). A high resistance at the trench can be also observed in case of a true free surface for the low resolution grid used (continuous
+lines in Fig. \ref{fig:subduction}), as pointed out by \citet{Schmeling2008}.
 
@@
 \noindent\includegraphics[width=400px]{./Figures/Subduction.pdf}
-\caption{Maximum depth of the slab as function of time for all the simulations of the \citet{Schmeling2008} benchmark, in case of harmonic, geometric and arithmetic means (red, green and blue lines, respectively) using a sticky air or a true free surface (discontinuous and continuous lines, respectively).
-Red, green and blue rectangular areas indicate the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{500}{\km} in case of sticky air, while the grey rectangular area indicates the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{600}{\km} in case of true free surface.}
+\caption{Maximum depth of the slab as function of time for all the simulations of the \citet{Schmeling2008} benchmark, in case of harmonic, geometric and
+arithmetic means (red, green and blue lines, respectively) using a sticky air or a true free surface (discontinuous and continuous lines, respectively).
+Red, green and blue rectangular areas indicate the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{500}{\km} in case of sticky air,
+while the grey rectangular area indicates the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{600}{\km} in case of true free surface.}
 \label{fig:subduction}
@@
 \subsection{The slab detachment}\label{sec:slab}
-The slab detachment benchmark is performed as described by \citet{Schmalholz2011} and \citet{Glerum2018}.
-The domain is rectangular with $L_x=$ \SI{1000}{\km} and $L_y=$ \SI{660}{\km} and a grid resolution of \num{256x256} elements.
-A non-linear viscous T-shaped layer with $\rho_l=$ \SI{3300}{\kg\per\cubic\metre} is placed at the top of the domain and surrounded by a linear viscous fluid with $\rho_f=$ \SI{3150}{\kg\per\cubic\m}.
-The top layer is 80 km-thick and a 250 km-long and 80 km-wide slab is placed at $x=L_x/2$ (Fig.
-\ref{fig:slab}a).
-The effective viscosity of the top layer is given by
+The slab detachment benchmark is performed as described by \citet{Schmalholz2011} and \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{1000}{\km}
+and $L_y=$ \SI{660}{\km} and a grid resolution of \num{256x256} elements. A non-linear viscous T-shaped layer with $\rho_l=$ \SI{3300}{\kg\per\cubic\metre} is
+placed at the top of the domain and surrounded by a linear viscous fluid with $\rho_f=$ \SI{3150}{\kg\per\cubic\m}. The top layer is 80 km-thick and a 250
+km-long and 80 km-wide slab is placed at $x=L_x/2$ (Fig. \ref{fig:slab}a). The effective viscosity of the top layer is given by
 \[\eta_{eff}=\eta_0 I_2^{\frac{(1-n)}{n}}\]
-with $\eta_0=$ \SI{4.75e11}{\pascal\s} and $n=4$, while the surrounding fluid has $\eta_f=$ \SI{e21}{\pascal\s}.
-The effective viscosity is capped between $\eta_{min}=$ \SI{e21}{\pascal\s} and $\eta_{max}=$ \SI{e25}{\pascal\s}.
-Velocity boundary conditions are set to free slip at the top and the bottom, and to no slip at the sides of the domain.
+with $\eta_0=$ \SI{4.75e11}{\pascal\s} and $n=4$, while the surrounding fluid has $\eta_f=$ \SI{e21}{\pascal\s}. The effective viscosity is capped between
+$\eta_{min}=$ \SI{e21}{\pascal\s} and $\eta_{max}=$ \SI{e25}{\pascal\s}. Velocity boundary conditions are set to free slip at the top and the bottom, and to
+no slip at the sides of the domain.
 
-The sinking of the slab determines a decrease of the effective viscosity in the T-shaped layer during the first part of the evolution (Fig.
-\ref{fig:slab}a), as consequence of local strain rates, while effective viscosities increase up to $\eta_{max}$ in the top layer after approximately \SI{20}{\mega\year}, when necking is complete (Fig.
-\ref{fig:slab}b).
+The sinking of the slab determines a decrease of the effective viscosity in the T-shaped layer during the first part of the evolution (Fig. \ref{fig:slab}a),
+as consequence of local strain rates, while effective viscosities increase up to $\eta_{max}$ in the top layer after approximately \SI{20}{\mega\year}, when
+necking is complete (Fig. \ref{fig:slab}b).
 
-The width of necking is tracked by means of markers position at the side of the slab.
-Necking width and time are normalised in function of the initial slab (80 km) and the characteristic time ($t_c=$ \SI{7.1158e14}{\s}) \citep{Schmalholz2011,Glerum2018}.
-The results in case of arithmetic, geometric and harmonic mean of the viscosity are shown in Fig.
-\ref{fig:necking}, compared with results from \citet{Glerum2018} for which an infinity norm mean was used.
+The width of necking is tracked by means of markers position at the side of the slab. Necking width and time are normalised in function of the initial slab
+(\SI{80}{\km}) and the characteristic time ($t_c=$ \SI{7.1158e14}{\s}) \citep{Schmalholz2011,Glerum2018}. The results in case of arithmetic, geometric and
+harmonic mean of the viscosity are shown in Fig. \ref{fig:necking}, compared with results from \citet{Glerum2018} for which an infinity norm mean was used.
 The influence of the average schemes agrees with results shown in Section \ref{sec:subduction}.
@@
 \subsection{The indenter experiment}\label{sec:indenter}
-The indenter benchmark simulate a rigid punch on a purely plastic von Mises material, for which exists an analytical solution \citep{Thieulot2008,Thieulot2014,Glerum2018}.
-The domain is a unit square with a grid resolution of \num{256x256} elements.
-The gravity acceleration is set to $\bm{g}=\bm{0}$ and effective viscosity is capped using $\eta_{min}=10^{-4}$ and $\eta_{max}=10^3$.
-The tolerance for the convergence of the non-linear solution is $tol=10^{-9}$, with a maximum number of non-linear iterations set to 500.
-The medium in the domain has $\rho=0.01$, an initial viscosity $\eta_0=10$, cohesion $C=1$ and an angle of internal friction $\phi=0$°.
-Velocity boundary conditions are set to no slip at the bottom and free slip at the sides of the domain.
-The top of the domain is open with the exception of the central portion (punch area) with width $w_p=0.125$, where $v=-1.05$ and $u$ is fixed either to 0 (rough punch experiment) or free (smooth punch experiment).
+The indenter benchmark simulate a rigid punch on a purely plastic von Mises material, for which exists an analytical solution
+\citep{Thieulot2008,Thieulot2014,Glerum2018}. The domain is a unit square with a grid resolution of \num{256x256} elements. The gravity acceleration is set
+to $\bm{g}=\bm{0}$ and effective viscosity is capped using $\eta_{min}=10^{-4}$ and $\eta_{max}=10^3$. The tolerance for the convergence of the non-linear
+solution is $tol=10^{-9}$, with a maximum number of non-linear iterations set to 500. The medium in the domain has $\rho=0.01$, an initial viscosity
+$\eta_0=10$, cohesion $C=1$ and an angle of internal friction $\phi=0$°. Velocity boundary conditions are set to no slip at the bottom and free slip at the
+sides of the domain. The top of the domain is open with the exception of the central portion (punch area) with width $w_p=0.125$, where $v=-1.05$ and $u$ is
+fixed either to 0 (rough punch experiment) or free (smooth punch experiment).
 
-Fig.
-\ref{fig:indenter} show the results in terms of viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure (panels g and h) for both experiments.
-The analytical solution indicate that pressure at ($0.5;1$) and ($0.5 \pm w_p;1$)is $p=\pi +1$ and $p=1$, respectively (black and red lines in Fig.
-\ref{fig:smooth_p}).
-Fig.
-\ref{fig:smooth_p} shows a clear improvement in the pressure solution when passing from a rough to a smooth punch (green and blue lines in Fig.
-\ref{fig:smooth_p}, respectively), as previously observed by \citet{Thieulot2014} and \citet{Glerum2018}.
+Fig. \ref{fig:indenter} show the results in terms of viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure
+(panels g and h) for both experiments. The analytical solution indicate that pressure at ($0.5;1$) and ($0.5 \pm w_p;1$)is $p=\pi +1$ and $p=1$, respectively
+(black and red lines in Fig. \ref{fig:smooth_p}). Fig. \ref{fig:smooth_p} shows a clear improvement in the pressure solution when passing from a rough to a
+smooth punch (green and blue lines in Fig. \ref{fig:smooth_p}, respectively), as previously observed by \citet{Thieulot2014} and \citet{Glerum2018}.
 
@@
 \noindent\includegraphics[width=400px]{./Figures/Indenter.pdf}
-\caption{Viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure (panels g and h) fields for rough (left column) and smooth (right column) punch experiments.}
+\caption{Viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure (panels g and h) fields for rough (left column) and
+smooth (right column) punch experiments.}
 \label{fig:indenter}
@@
 \noindent\includegraphics[width=400px]{./Figures/Smooth_Rough.pdf}
-\caption{Nodal pressure in function of the x coordinate at $y=1$ for rough (green line) and smooth (blue line) punch experiment.
-Black and red lines indicate the analytical solution at $x=0.5$ and $x=0.5 \pm w_p$, respectively.}
+\caption{Nodal pressure in function of the x coordinate at $y=1$ for rough (green line) and smooth (blue line) punch experiment. Black and red lines indicate
+the analytical solution at $x=0.5$ and $x=0.5 \pm w_p$, respectively.}
 \label{fig:smooth_p}
@@
 \subsection{The brick experiment}\label{sec:brick}
-An instantaneous version of the brick benchmark is performed to verify the correctness of the pressure-dependent plasticity for different angle of internal friction, as proposed by \citet{Glerum2018}.
-The domain is rectangular with $L_x=$ \SI{40}{\km} and $L_y=$ \SI{10}{\km} and a grid resolution of \num{512x128} elements.
-The gravity acceleration is set to $g_y=$ \SI{-10}{\m\square\s} and effective viscosity is capped using $\eta_{min}=$ \SI{e19}{\pascal\s} and $\eta_{max}=$ \SI{e26}{\pascal\s}.
-The tolerance for the convergence of the non-linear solution is $tol=10^{-7}$, with a maximum number of non-linear iterations set to 1000.
-A 800 m-wide and 400 m-height inclusion with $\eta_b=$ \SI{e20}{\pascal\s} and $\rho_b=$ \SI{2700}{\kg\per\cubic\m} is placed at the bottom of the domain at $x=L_x/2$.
-The inclusion is surrounded by a non-linear viscous medium with $\rho_m=$ \SI{2700}{\kg\per\cubic\m}, an initial viscosity of \SI{e23}{\pascal\s}, a linear viscous viscosity of \SI{e25}{\pascal\s} and a cohesion of \SI{40}{\mega\pascal}.
-Velocities are set to free slip conditions at the bottom of the domain and the top is open.
-Velocities on sides of the domain are fixed to $u=$ \SI{+-2e-11}{\m\per\s} and $v=0$.
-The experiment is performed in compressional and extensional contexts with an angle of internal friction $\phi$ of the non-linear medium variable between 0° and 20°.
+An instantaneous version of the brick benchmark is performed to verify the correctness of the pressure-dependent plasticity for different angle of internal
+friction, as proposed by \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{40}{\km} and $L_y=$ \SI{10}{\km} and a grid resolution of \num{512x128}
+elements. The gravity acceleration is set to $g_y=$ \SI{-10}{\m\square\s} and effective viscosity is capped using $\eta_{min}=$ \SI{e19}{\pascal\s} and
+$\eta_{max}=$ \SI{e26}{\pascal\s}. The tolerance for the convergence of the non-linear solution is $tol=10^{-7}$, with a maximum number of non-linear iterations
+set to 1000. A 800 m-wide and 400 m-height inclusion with $\eta_b=$ \SI{e20}{\pascal\s} and $\rho_b=$ \SI{2700}{\kg\per\cubic\m} is placed at the bottom of the
+domain at $x=L_x/2$. The inclusion is surrounded by a non-linear viscous medium with $\rho_m=$ \SI{2700}{\kg\per\cubic\m}, an initial viscosity of
+\SI{e23}{\pascal\s}, a linear viscous viscosity of \SI{e25}{\pascal\s} and a cohesion of \SI{40}{\mega\pascal}. Velocities are set to free slip conditions at
+the bottom of the domain and the top is open. Velocities on sides of the domain are fixed to $u=$ \SI{+-2e-11}{\m\per\s} and $v=0$. The experiment is performed
+in compressional and extensional contexts with an angle of internal friction $\phi$ of the non-linear medium variable between 0° and 20°.
 
-As expected, two shear bands stem from the inclusion with variable angles in relation to both the dynamics context and the internal friction angle.
-Shear band angles formed at 45° for $\phi=0$° in both extensional and compressional contexts (Fig.
-\ref{fig:brick_beam}a), while they are different in case of $\phi\neq0$°.
-Shear band angles in case of $\phi=20$° are shown for compression and extension in Fig.
-\ref{fig:brick_beam}b  and c, respectively.
-Results for internal friction angles from 0° to 20° are plotted in Fig.
-\ref{fig:brick}, in comparison with theoretical Roscoe, Arthur and Coulomb shear band angles.
+As expected, two shear bands stem from the inclusion with variable angles in relation to both the dynamics context and the internal friction angle. Shear band
+angles formed at 45° for $\phi=0$° in both extensional and compressional contexts (Fig. \ref{fig:brick_beam}a), while they are different in case of $\phi\neq0$°.
+Shear band angles in case of $\phi=20$° are shown for compression and extension in Fig. \ref{fig:brick_beam}b  and c, respectively. Results for internal friction
+angles from 0° to 20° are plotted in Fig. \ref{fig:brick}, in comparison with theoretical Roscoe, Arthur and Coulomb shear band angles.
 
@@
 \noindent\includegraphics[width=400px]{./Figures/Brick_Beam.pdf}
-\caption{Shear band angles predicted for the brick experiment for $\phi=0$° (panel a) and for $\phi=20$° in case of compression and extension (panels b and c, respectively).}
+\caption{Shear band angles predicted for the brick experiment for $\phi=0$° (panel a) and for $\phi=20$° in case of compression and extension (panels b and c,
+respectively).}
 \label{fig:brick_beam}
@@
 \noindent\includegraphics[width=400px]{./Figures/Brick.pdf}
-\caption{Shear band angles predicted for the brick experiment in case of compressional and extensional contexts as function of different internal angle of friction (purple line and dots), compared with theoretical Roscoe, Artur and Coulomb angles (orange, blue and green lines, respectively).}
+\caption{Shear band angles predicted for the brick experiment in case of compressional and extensional contexts as function of different internal angle of
+friction (purple line and dots), compared with theoretical Roscoe, Artur and Coulomb angles (orange, blue and green lines, respectively).}
 \label{fig:brick}
@@
 \subsection{Advection stabilisation}\label{sec:advection}
-The 1D advection problem proposed by \citet{Donea2003} and \citet{Thieulot2011} is performed to verify the effectiveness of the SUPG method to stabilise the advection term of the energy equation.
-The domain is a 1D segment with $L_x=1$ composed by 50 elements and a discontinuity in the thermal field placed at $x=0.25$.
-Temperature is set to 1 for $x \leq 0.25$ and to 0 for $x > 0.25$.
-Velocity is set to ${u}=1$ in the entire domain.
-The simulation is performed for 250 time steps, with $dt=0.002$, so the thermal discontinuity should be at $x=0.75$ at the end of the simulation.
+The 1D advection problem proposed by \citet{Donea2003} and \citet{Thieulot2011} is performed to verify the effectiveness of the SUPG method to stabilise the
+advection term of the energy equation. The domain is a 1D segment with $L_x=1$ composed by 50 elements and a discontinuity in the thermal field placed at
+$x=0.25$. Temperature is set to 1 for $x \leq 0.25$ and to 0 for $x > 0.25$. Velocity is set to ${u}=1$ in the entire domain. The simulation is performed for
+250 time steps, with $dt=0.002$, so the thermal discontinuity should be at $x=0.75$ at the end of the simulation.
 
-Temperature profiles at the end of the simulation are shown in \ref{fig:advection} as function of the dimensionless coefficient $\gamma=\tau \bm{v}/h$.
-In case of the classic Galerkin method ($\gamma=0$, blue line in \ref{fig:advection}) the final thermal profile is characterised by strong oscillations, which are eliminated in case of the SUPG method ($\gamma=0.045$, orange line in \ref{fig:advection}).
+Temperature profiles at the end of the simulation are shown in \ref{fig:advection} as function of the dimensionless coefficient $\gamma=\tau \bm{v}/h$. In case
+of the classic Galerkin method ($\gamma=0$, blue line in \ref{fig:advection}) the final thermal profile is characterised by strong oscillations, which are
+eliminated in case of the SUPG method ($\gamma=0.045$, orange line in \ref{fig:advection}).
 
@@
 \noindent\includegraphics[width=400px]{./Figures/Advection.pdf}
-\caption{Temperature profile as function of \textit{x} for the advection stabilisation benchmark.
-Purple line indicates the initial temperature profile; the green line indicates the analytical temperature profile after 250 time steps; blue line indicates the temperature profile after 250 time steps in case of the classic Galerkin method ($\gamma=0$); orange line indicates the temperature profile after 250 time steps in case of the SUPG method ($\gamma=0.045$).}
+\caption{Temperature profile as function of \textit{x} for the advection stabilisation benchmark. Purple line indicates the initial temperature profile; the
+green line indicates the analytical temperature profile after 250 time steps; blue line indicates the temperature profile after 250 time steps in case of the
+classic Galerkin method ($\gamma=0$); orange line indicates the temperature profile after 250 time steps in case of the SUPG method ($\gamma=0.045$).}
 \label{fig:advection}
@@
 \subsection{Shear and adiabatic heating}\label{sec:simple_shear}
-This exercise is performed in an unit square domain composed by \num{128x128} elements.
-The velocity field is prescribed on the entire domain with $\bm{v}=(L_y-y)y\bm{e_x}$; viscosity, density and specific heat are fixed to 1, while thermal conductivity and radiogenic energy are fixed to 0.
-Therefore, the energy equation (Eq.
-\ref{eq:energy}) can be simplified as \[\frac{\partial T}{\partial t}=H_s\] and fixing $T(t=0)=0$, the temperature can be find as \[T(t)=H_s t\]
+This exercise is performed in an unit square domain composed by \num{128x128} elements. The velocity field is prescribed on the entire domain with
+$\bm{v}=(L_y-y)y\bm{e_x}$; viscosity, density and specific heat are fixed to 1, while thermal conductivity and radiogenic energy are fixed to 0. Therefore,
+the energy equation (Eq. \ref{eq:energy}) can be simplified as \[\frac{\partial T}{\partial t}=H_s\] and fixing $T(t=0)=0$, the temperature can be find as
+\[T(t)=H_s t\]
 In this case we have
@@
 \noindent\includegraphics[width=500px]{./Figures/Analytical_Shear.pdf}
-\caption{Velocity (panel a), temperature (panel b) and shear heating (panel c) as function of the \textit{y} coordinate for the simple shear experiment.
-The solutions predicted by the model (red lines) are compared with the analytical solutions (dashed blue lines).}
+\caption{Velocity (panel a), temperature (panel b) and shear heating (panel c) as function of the \textit{y} coordinate for the simple shear experiment. The
+solutions predicted by the model (red lines) are compared with the analytical solutions (dashed blue lines).}
 \label{fig:analytical_en}
@@
 \subsection{Shear and adiabatic heating}\label{sec:shear}
-This problem is performed as presented in Exercise 9.4 in \citet{Gerya2010b}.
-Two materials are vertically separated in a rectangular domain with $L_x=$ \SI{1000}{\km}, $L_y=$ \SI{1500}{\km} and a grid resolution of \num{30x20} elements.
-Constant thermal coefficient expansion ($\alpha=$ \SI{3e-5}{\per\kelvin}), temperature ($T=$ \SI{1300}{\kelvin}) and gravity acceleration ($g_y=$ \SI{-10}{\m\square\s}) are assumed in the whole domain.
-Fluid 1 (on the left side of the domain) has $\rho_1=$ \SI{3200}{\kg\per\cubic\m} and $\eta_1=$ \SI{e20}{\pascal\s}; fluid 2 (on the right side of the domain) has $\rho_2=$ \SI{3300}{\kg\per\cubic\m} and $\eta_2=$ \SI{e22}{\pascal\s}.
-Velocity boundary conditions are set to free slip on all sides of the domain.
+This problem is performed as presented in Exercise 9.4 in \citet{Gerya2010b}. Two materials are vertically separated in a rectangular domain with $L_x=$
+\SI{1000}{\km}, $L_y=$ \SI{1500}{\km} and a grid resolution of \num{30x20} elements. Constant thermal coefficient expansion ($\alpha=$ \SI{3e-5}{\per\kelvin}),
+temperature ($T=$ \SI{1300}{\kelvin}) and gravity acceleration ($g_y=$ \SI{-10}{\m\square\s}) are assumed in the whole domain. Fluid 1 (on the left side of the
+domain) has $\rho_1=$ \SI{3200}{\kg\per\cubic\m} and $\eta_1=$ \SI{e20}{\pascal\s}; fluid 2 (on the right side of the domain) has $\rho_2=$
+\SI{3300}{\kg\per\cubic\m} and $\eta_2=$ \SI{e22}{\pascal\s}. Velocity boundary conditions are set to free slip on all sides of the domain.
 
-As shown in Fig.
-\ref{fig:energy}, both shear and adiabatic heating predicted by the code (first row in Fig.
-\ref{fig:energy}) well recreate the results obtained by means of example ${Shear\_adiabatic\_heating.m}$ from \citet{Gerya2010b} (second row in Fig.
-\ref{fig:energy}).
+As shown in Fig. \ref{fig:energy}, both shear and adiabatic heating predicted by the code (first row in Fig. \ref{fig:energy}) well recreate the results
+obtained by means of example ${Shear\_adiabatic\_heating.m}$ from \citet{Gerya2010b} (second row in Fig. \ref{fig:energy}).
 
@@
 \noindent\includegraphics[width=400px]{./Figures/Energy.pdf}
-\caption{Comparison between shear (first column), adiabatic (second column) and total (third column) energy predicted by the code (panels a, b and c) and those created using example ${Shear\_adiabatic\_heating.m}$ from Exercise 9.4 in \citet{Gerya2010b} (panels d, e and f).}
+\caption{Comparison between shear (first column), adiabatic (second column) and total (third column) energy predicted by the code (panels a, b and c) and those
+created using example ${Shear\_adiabatic\_heating.m}$ from Exercise 9.4 in \citet{Gerya2010b} (panels d, e and f).}
 \label{fig:energy}
@@
 \subsection{Mantle convection}\label{sec:mantle}
-This problem is performed as presented by \citet{Blankenbach1989} (constant viscosity cases) and \citet{Thieulot2014}, in a 2D unit square domain with gravity acceleration $g_y=10^{10}Ra$.
-The experiment is performed with three different Rayleigh numbers ($Ra=10^4$, $10^5$ and $10^6$) and with different grid resolution (between \num{32x32} and \num{128x128} elements).
-Both fluids have constant viscosity, initial density, heat capacity, thermal conductivity ($\eta= \rho_0= C_p = k=1$), reference temperature ($T_0=0$) and thermal expansion coefficient ($\alpha=10^{-10}$).
-Temperatures are set to 0 on top and 1 on bottom of the domain.
-Velocity boundary conditions are set to free slip on all sides.
-The initial temperature field is given by \[T(x,y)=(1-y)-0.01 \cos (\pi x) \sin (\pi x)\]
+This problem is performed as presented by \citet{Blankenbach1989} (constant viscosity cases) and \citet{Thieulot2014}, in a 2D unit square domain with gravity
+acceleration $g_y=10^{10}Ra$. The experiment is performed with three different Rayleigh numbers ($Ra=10^4$, $10^5$ and $10^6$) and with different grid
+resolution (between \num{32x32} and \num{128x128} elements). Both fluids have constant viscosity, initial density, heat capacity, thermal conductivity
+($\eta= \rho_0= C_p = k=1$), reference temperature ($T_0=0$) and thermal expansion coefficient ($\alpha=10^{-10}$). Temperatures are set to 0 on top and 1 on
+bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
+\[T(x,y)=(1-y)-0.01 \cos (\pi x) \sin (\pi x)\]
 
-The solution generated by the code in terms of $v_{rms}$ (calculated as in Sec.
-\ref{sec:error}) and $Nu$ as function of time are reported for all the simulations in Fig.
-\ref{fig:mantle}.
-At the steady state, $v_{rms}$ and $Nu$ of all simulations converge well toward the values from \citet{Blankenbach1989}, with lower errors for higher resolution grids (Table \ref{tab:mantle}).
+The solution generated by the code in terms of $v_{rms}$ (calculated as in Sec. \ref{sec:error}) and $Nu$ as function of time are reported for all the
+simulations in Fig. \ref{fig:mantle}. At the steady state, $v_{rms}$ and $Nu$ of all simulations converge well toward the values from \citet{Blankenbach1989},
+with lower errors for higher resolution grids (Table \ref{tab:mantle}).
 
@@
 \caption{$v_{rms}$ (panels a, c and e) and $Nu$ (panels b, d and f) for the mantle convection benchmark as function of time for different grid resolution.
-Panels a and b show the results for $Ra=10^4$; Panels c and d show the results for $Ra=10^5$; panels e and f show the results for $Ra=10^6$.
-Purple lines indicate the convergence values for the $v_{rms}$ and the $Nu$ from \citet{Blankenbach1989} at the steady state.}
+Panels a and b show the results for $Ra=10^4$; Panels c and d show the results for $Ra=10^5$; panels e and f show the results for $Ra=10^6$. Purple lines
+indicate the convergence values for the $v_{rms}$ and the $Nu$ from \citet{Blankenbach1989} at the steady state.}
 \label{fig:mantle}
@@
 \subsection{Thin layer entrainment}\label{sec:thin}
-This experiment is performed as originally proposed by \citet{vanKeken1997} with two incompressible fluids in a rectangular domain with $L_x=2$ and $L_y=1$ and gravity acceleration $g_y=10^{10}Ra$.
-The Rayleigh number and the compositional Rayleigh number are fixed to $Ra=300000$ and $Rc=450000$, respectively.
-Both fluids have constant viscosity, thermal conductivity, specific heat ($\eta=\rho=C_p=1$) and thermal expansion coefficient ($\alpha=10^{-10}$).
-Fluid 1 has a density $\rho_1=1$, while fluid 2 is denser ($\rho_2=\rho_1+1.5 \cdot 10^{-10}$) and is placed at the bottom of the domain, for $y \leq 0.025$.
-Temperature are set to 0 on top and 1 on bottom of the domain.
-Velocity boundary conditions are set to free slip on all sides of the domain.
-The initial temperature field is given by
+This experiment is performed as originally proposed by \citet{vanKeken1997} with two incompressible fluids in a rectangular domain with $L_x=2$ and $L_y=1$
+and gravity acceleration $g_y=10^{10}Ra$. The Rayleigh number and the compositional Rayleigh number are fixed to $Ra=300000$ and $Rc=450000$, respectively.
+Both fluids have constant viscosity, thermal conductivity, specific heat ($\eta=\rho=C_p=1$) and thermal expansion coefficient ($\alpha=10^{-10}$). Fluid 1
+has a density $\rho_1=1$, while fluid 2 is denser ($\rho_2=\rho_1+1.5 \cdot 10^{-10}$) and is placed at the bottom of the domain, for $y \leq 0.025$.
+Temperature are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides of the domain. The initial
+temperature field is given by
 \[T(x,y)=T_u(x,y)+T_l(x,y)+T_r(x,y)+T_s(x,y)-\frac{3}{2}\]
@@
 
-The experiment is performed with two grid resolution (\num{125x40} and \num{200x80} elements), with 100 markers per element and CFL of 0.25.
-$v_{rms}$ (see Sec.
-\ref{sec:error}) calculated for both the simulations match well with results from \citet{vanKeken1997} and obtained with ASPECT \citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} and ELEFANT \citep{Thieulot2014}, obtained with similar grid resolutions (\ref{fig:thin}).
-In addition, a variety of simulations with different resolutions and aspect ratios are performed to verify that the sensitivity of the initial velocity field.
-Results match well with results from \citet{vanKeken1997} (in the grey area) and \citet{Thieulot2014} (\ref{fig:thin_initial}).
+The experiment is performed with two grid resolution (\num{125x40} and \num{200x80} elements), with 100 markers per element and CFL of 0.25. $v_{rms}$
+(see Sec. \ref{sec:error}) calculated for both the simulations match well with results from \citet{vanKeken1997} and obtained with ASPECT
+\citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} and ELEFANT \citep{Thieulot2014}, obtained with similar grid resolutions
+(Fig. \ref{fig:thin}). In addition, a variety of simulations with different resolutions and aspect ratios are performed to verify that the sensitivity of the
+initial velocity field. Results match well with results from \citet{vanKeken1997} (in the grey area) and \citet{Thieulot2014} (Fig. \ref{fig:thin_initial}).
 
@@
 \noindent\includegraphics[width=400px]{./Figures/vrms_thin.pdf}
-\caption{$v_{rms}$ for the thin layer experiment as function of time for different grid resolution.
-Results are compared with results obtained by \citet{vanKeken1997} (black lines) and with ASPECT and ELEFANT.}
+\caption{$v_{rms}$ for the thin layer experiment as function of time for different grid resolution. Results are compared with results obtained by
+\citet{vanKeken1997} (black lines) and with ASPECT and ELEFANT.}
 \label{fig:thin}
@@
 \noindent\includegraphics[width=400px]{./Figures/vrmszero.pdf}
-\caption{$v_{rms}$ at $t=0$ for the thin layer experiment as function of the element size for different aspect ratios of the elements.
-Results are compared with results obtained by \citet{vanKeken1997} (black dots in the grey area) and with ASPECT and ELEFANT.}
+\caption{$v_{rms}$ at $t=0$ for the thin layer experiment as function of the element size for different aspect ratios of the elements. Results are compared
+with results obtained by \citet{vanKeken1997} (black dots in the grey area) and with ASPECT and ELEFANT.}
 \label{fig:thin_initial}
@@
 \subsection{Sinking hydrated cylinder}\label{sec:quinquis}
-This experiment is performed as explained in \citet{Quinquis2014} and simulates the sinking of a cold, hydrated cylinder into a warm, dry mantle.
-The domain is square with $L_x=L_y=$ \SI{300}{\km}, a grid resolution of \num{300x300} elements and 25 markers per element.
-The sinking cylinder is located at $x=$ \SI{150}{\km} and $y=$ \SI{170}{\km} with a radius of \SI{20}{\km} and has $\rho_c=$ \SI{3250}{\kg\per\cubic\m}, $\eta_c=$ \SI{e23}{\pascal\s}, a thermal conductivity $k_c=$ \SI{4.5}{\watt\per\m\per\kelvin},a specific heat $C_{p_c}=$ \SI{1250}{\joule\per\kg\per\kelvin}; the surrounding mantle has $\rho_m=$ \SI{3200}{\kg\per\cubic\m}, $\eta_m=$ \SI{e20}{\pascal\s}, a thermal conductivity $k_m=$ \SI{105}{\watt\per\m\per\kelvin} and specific heat $C_{p_m}=$ \SI{1250}{\joule\per\kg\per\kelvin}; a 58 km-thick lithosphere overlies the mantle and has $\rho_l=$ \SI{3200}{\kg\per\cubic\m}, $\eta_l=$ \SI{e23}{\pascal\s}, a thermal conductivity $k_l=$ \SI{4.5}{\watt\per\m\per\kelvin}, specific heat $C_{p_l}=$ \SI{750}{\joule\per\kg\per\kelvin} and a thermal diffusivity of \SI{e-6}{\square\m\per\s}.
-Velocity boundary conditions are set to free slip on all sides of the domain.
-Temperature increase linearly in the lithosphere, from 0° to 1300°C, and in the mantle, from 1300° to 1360.5°C.
-The initial temperature of the cylinder is fixed to 400°C.
-Throughout the evolution, temperature is fixed to 0° and 1360.5°C at the top and bottom of the domain, respectively.
-The initial bound water content is imposed at 2 wt.\% in the cylinder and at 0 wt.\% in both the mantle and the lithosphere.
-The maximum amount of water in the mantle is fixed to 0.2 wt.\%, while maximum water content in the cylinder and in the lithosphere is function of pressure and temperature and is calculated using a serpentinized harzburgite with Perple\_X, as in \citet{Quinquis2014}.
+This experiment is performed as explained in \citet{Quinquis2014} and simulates the sinking of a cold, hydrated cylinder into a warm, dry mantle. The domain
+is square with $L_x=L_y=$ \SI{300}{\km}, a grid resolution of \num{300x300} elements and 25 markers per element. The sinking cylinder is located at $x=$
+\SI{150}{\km} and $y=$ \SI{170}{\km} with a radius of \SI{20}{\km} and has $\rho_c=$ \SI{3250}{\kg\per\cubic\m}, $\eta_c=$ \SI{e23}{\pascal\s}, a thermal
+conductivity $k_c=$ \SI{4.5}{\watt\per\m\per\kelvin},a specific heat $C_{p_c}=$ \SI{1250}{\joule\per\kg\per\kelvin}; the surrounding mantle has $\rho_m=$
+\SI{3200}{\kg\per\cubic\m}, $\eta_m=$ \SI{e20}{\pascal\s}, a thermal conductivity $k_m=$ \SI{105}{\watt\per\m\per\kelvin} and specific heat $C_{p_m}=$
+\SI{1250}{\joule\per\kg\per\kelvin}; a 58 km-thick lithosphere overlies the mantle and has $\rho_l=$ \SI{3200}{\kg\per\cubic\m}, $\eta_l=$ \SI{e23}{\pascal\s},
+a thermal conductivity $k_l=$ \SI{4.5}{\watt\per\m\per\kelvin}, specific heat $C_{p_l}=$ \SI{750}{\joule\per\kg\per\kelvin} and a thermal diffusivity of
+\SI{e-6}{\square\m\per\s}. Velocity boundary conditions are set to free slip on all sides of the domain. Temperature increase linearly in the lithosphere,
+from 0° to 1300°C, and in the mantle, from 1300° to 1360.5°C. The initial temperature of the cylinder is fixed to 400°C. Throughout the evolution, temperature
+is fixed to 0° and 1360.5°C at the top and bottom of the domain, respectively. The initial bound water content is imposed at 2 wt.\% in the cylinder and at
+0 wt.\% in both the mantle and the lithosphere. The maximum amount of water in the mantle is fixed to 0.2 wt.\%, while maximum water content in the cylinder
+and in the lithosphere is function of pressure and temperature and is calculated using a serpentinized harzburgite with Perple\_X, as in \citet{Quinquis2014}.
 
-As shown in \citet{Quinquis2014}, the model is characterised by a progressive dehydration in the external portion of the cylinder, due to the increase of temperature, with a consequent vertical migration of free water that hydrates the mantle above the cylinder, up to the lithosphere.
-The distribution of bound and free water during the evolution of the experiment is shown in Fig.
-\ref{fig:quinquis} (left and right columns, respectively).
+As shown in \citet{Quinquis2014}, the model is characterised by a progressive dehydration in the external portion of the cylinder, due to the increase of
+temperature, with a consequent vertical migration of free water that hydrates the mantle above the cylinder, up to the lithosphere. The distribution of
+bound and free water during the evolution of the experiment is shown in Fig. \ref{fig:quinquis} (left and right columns, respectively).
 
@@
 \noindent\includegraphics[width=400px]{./Figures/Hydration.pdf}
-\caption{Distribution of bound and free water (left and right columns, respectively) at $t=0.5$ and $t=$ \SI{1.25}{\mega\year} (first and second rows, respectively) for the sinking hydrated cylinder experiment.
-Only markers with bound or free water not equal to 0 are plotted.}
+\caption{Distribution of bound and free water (left and right columns, respectively) at $t=0.5$ and $t=$ \SI{1.25}{\mega\year} (first and second rows,
+respectively) for the sinking hydrated cylinder experiment. Only markers with bound or free water not equal to 0 are plotted.}
 \label{fig:quinquis}
@@
 \subsection{Melting curves}\label{sec:katz}
-This instantaneous experiment is performed in order to recreate isobaric melting curves, as function of temperature, bulk water and content of water in the melt, to compare with melting curves obtained by \citet{Katz2003}.
-The domain is square with $L_x=L_y=100$ km, a grid resolution of \num{100x100} elements and 25 markers per element.
-Density is set in the entire domain to \SI{3300}{\kg\per\cubic\m} and temperature increases from the sides to the centre from 1000° to 1500°C.
-The experiment is performed for bulk water contents from 0 to 0.3 wt.\%.
+This instantaneous experiment is performed in order to recreate isobaric melting curves, as function of temperature, bulk water and content of water in the
+melt, to compare with melting curves obtained by \citet{Katz2003}. The domain is square with $L_x=L_y=100$ km, a grid resolution of \num{100x100} elements and
+25 markers per element. Density is set in the entire domain to \SI{3300}{\kg\per\cubic\m} and temperature increases from the sides to the centre from 1000° to
+1500°C. The experiment is performed for bulk water contents from 0 to 0.3 wt.\%.
 
-Fig.
-\ref{fig:fraction} shows isobaric melting curves for \SI{1}{\giga\pascal} (panel a) and \SI{3}{\giga\pascal} (panel b) calculated for different bulk water contents.
-The curves well match with those obtained by \citet{Katz2003}, showing the saturation at the solidus for bulk water of 0.2 and 0.3 wt.\%, indicating that the code correctly determines melt fraction as function of pressure, temperature and bulk water content.
-Same results can be observed in Fig.
-\ref{fig:water}a, where isobaric and isothermal melting curves are shown as function of bulk water content.
-In addition, Fig.
-\ref{fig:water}b shows that the code correctly determines also the percentage of water in the melt as function of pressure, temperature and melt fraction.
+Fig. \ref{fig:fraction} shows isobaric melting curves for \SI{1}{\giga\pascal} (panel a) and \SI{3}{\giga\pascal} (panel b) calculated for different bulk water
+contents. The curves well match with those obtained by \citet{Katz2003}, showing the saturation at the solidus for bulk water of 0.2 and 0.3 wt.\%, indicating
+that the code correctly determines melt fraction as function of pressure, temperature and bulk water content. Same results can be observed in Fig.
+\ref{fig:water}a, where isobaric and isothermal melting curves are shown as function of bulk water content. In addition, Fig. \ref{fig:water}b shows that the
+code correctly determines also the percentage of water in the melt as function of pressure, temperature and melt fraction.
 

### 2021-04-13 Paper.tex
modified

@@
 markers. The correctness of the evolution of the markers chain is tested with the experiment of a 2D time-dependent Stokes sphere below a free surface and
-compared with results from ASPECT \citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} (Section \ref{sec:sphere}). In the true free surface
+compared with results from ASPECT\footnote{\url{https://aspect.geodynamics.org/}} \citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} (Section \ref{sec:sphere}).
+In the true free surface
 case the top boundary is assumed stress-free and velocities are not fixed. In this case topography variations are described by vertical deformations of the

### 2021-04-15 Paper.tex
modified

--- 2021-04-13 Paper.tex
@@
 \section{Introduction}\label{sec:intro}
-This supplement includes a description of the code, with a presentation of the features implemented followed by results of extensive benchmarks listed in Table \ref{tab:list}, implemented 
-to verify that the code solves correctly a large range of problems. Description and results of the benchmarks are shown in Section \ref{sec:benchmark}.
+This supplement includes a description of the code, with a presentation of the features implemented followed by results of extensive benchmarks listed in
+Table \ref{tab:list}, implemented to verify that the code solves correctly a large range of problems. Description and results of the benchmarks are shown
+in Section \ref{sec:benchmark}.
 
@@
 markers. The correctness of the evolution of the markers chain is tested with the experiment of a 2D time-dependent Stokes sphere below a free surface and
-compared with results from ASPECT \citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} (Section \ref{sec:sphere}). In the true free surface
-case the top boundary is assumed stress-free and velocities are not fixed. In this case topography variations are described by vertical deformations of the
-mesh that depend on the velocity field of the nodes that identify the free surface, while horizontal deformations are not taken into account. This procedure
-is known as the Arbitrary Lagrangian-Eulerian (ALE) method and its implementation follows the technique described in \citet{Thieulot2011}. However, although
-the implementation of a true free surface better reproduces laboratory experiment, extremely small time steps can be necessary to maintain stability
-\citep{Kaus2010a,Quinquis2011,Thieulot2014}. Therefore, the stability algorithm proposed by \citet{Kaus2010a} is implemented to avoid instabilities due to
-high density differences at the free surface when using too large time steps. The implementation of this algorithm are tested by performing the experiment
-described by \citet{Kaus2010a} (Section \ref{sec:stab}).
+compared with results from ASPECT\footnote{\url{https://aspect.geodynamics.org/}} \citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} (Section
+\ref{sec:sphere}). In the true free surface case the top boundary is assumed stress-free and velocities are not fixed. In this case topography variations are
+described by vertical deformations of the mesh that depend on the velocity field of the nodes that identify the free surface, while horizontal deformations are
+not taken into account. This procedure is known as the Arbitrary Lagrangian-Eulerian (ALE) method and its implementation follows the technique described in
+\citet{Thieulot2011}. However, although the implementation of a true free surface better reproduces laboratory experiment, extremely small time steps can be
+necessary to maintain stability \citep{Kaus2010a,Quinquis2011,Thieulot2014}. Therefore, the stability algorithm proposed by \citet{Kaus2010a} is implemented
+to avoid instabilities due to high density differences at the free surface when using too large time steps. The implementation of this algorithm are tested by
+performing the experiment described by \citet{Kaus2010a} (Section \ref{sec:stab}).
 
@@
 where $\mathbb{K}$ is the Stiffness matrix and $\bm{f}$ is the right hand side vector \citep{Spiegelman2016,Glerum2018}.  The $L_2$-norm is extracted from
-$\bm{\mathcal{R}}^i$ and it is normalised using the $L_2$-norm extracted from $\bm{f}^0$. Iterations are performed until either convergence
-($\frac{||\bm{\mathcal{R}}^i||_2}{||\bm{f}^0||_2}<tol$) or maximal number of non-linear iterations ($it_{max}$) is reached.
+$\bm{\mathcal{R}}^i$ and it is normalised as follows
+\begin{equation}\label{eq:convergence}
+\frac{||\bm{\mathcal{R}}^i||_2}{||\bm{\mathcal{R}}^0||_2}
+\end{equation}
+where $||\bm{\mathcal{R}}^0||_2$ is the $L_2$-norm extracted at the first nonlinear iteration.
+Since often the normalised nonlinear residual drop very quickly over the first 2-3 iterations, also the $L_2$-norm of the velocity residual is calculated as
+\begin{equation}\label{eq:convergence_vel}
+\frac{||\bm{v}_1-\bm{v}_{1-1}||_2}{||\bm{v}_{i-1}||_2}
+\end{equation}
+Iterations are performed until either maximal number of non-linear iterations ($it_{max}$) is reached or both the normalised nonlinear residual and the
+velocity residual converge under a defined tolerance $tol$.
 
@@
 the bottom of the domain and the top is open. Velocities on sides of the domain are fixed to $u=$ \SI{+-2e-11}{\m\per\s} and $v=0$. The experiment is performed
-in compressional and extensional contexts with an angle of internal friction $\phi$ of the non-linear medium variable between 0° and 20°.
+in compressional and extensional contexts with an angle of internal friction $\phi$ of the non-linear medium variable between 0° and 30°.
 
@@
 angles formed at 45° for $\phi=0$° in both extensional and compressional contexts (Fig. \ref{fig:brick_beam}a), while they are different in case of $\phi\neq0$°.
-Shear band angles in case of $\phi=20$° are shown for compression and extension in Fig. \ref{fig:brick_beam}b  and c, respectively. Results for internal friction
-angles from 0° to 20° are plotted in Fig. \ref{fig:brick}, in comparison with theoretical Roscoe, Arthur and Coulomb shear band angles.
+Shear band angles in case of $\phi=20$° are shown for compression and extension in Fig. \ref{fig:brick_beam}b  and c, respectively. Values of shear band angles
+as function of internal friction angles from 0° to 30° are extracted at different depths and minimum and maximum values are plotted in Fig. \ref{fig:brick},
+in comparison with theoretical Roscoe, Arthur and Coulomb shear band angles. For all tests the tolerance for nonlinear convergence is defined $tol=10^{-7}$
+and none converges before the maximum number of iterations ($it_{max}=1000$) is reached. However, tests with internal friction angles up to 15° show a constant
+decrease of the velocity residuals, which can not be observed for for tests with higher internal friction angles (Fig. \ref{fig:convergence}).
 
@@
 \caption{Shear band angles predicted for the brick experiment in case of compressional and extensional contexts as function of different internal angle of
-friction (purple line and dots), compared with theoretical Roscoe, Artur and Coulomb angles (orange, blue and green lines, respectively).}
+friction (continuous black line and red dots), compared with theoretical Roscoe, Artur and Coulomb angles (discontinuous black lines). Red lines indicate the
+range of angles calculated at different depths.}
 \label{fig:brick}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \noindent\includegraphics[width=500px]{./Figures/convergence.pdf}
+  \caption{Convergence rates for all tests. Normalised residuals (left column) are compared with velocity residuals (right column) for extensional (first row)
+  and compressional (second row) settings.}
+  \label{fig:convergence}
   \end{figure}

### 2021-04-28 Paper.tex
modified

@@
 \end{equation}
-where $||\bm{\mathcal{R}}^0||_2$ is the $L_2$-norm extracted at the first nonlinear iteration.
+where $||\bm{\mathcal{R}}^0||_2$ is the $L_2$-norm of the first nonlinear iteration.
 Since often the normalised nonlinear residual drop very quickly over the first 2-3 iterations, also the $L_2$-norm of the velocity residual is calculated as
@@
 The correctness of the implementation and the computation of shear and adiabatic heating is verified performing an exercise with an analytical solution
-(Section \ref{sec:simple_shear}) and Exercise 9.4 in \citet{Gerya2010b} (Section \ref{sec:shear}). Furthermore, two benchmarks are performed to verity that
-the code solve correctly Eqs. \ref{eq:momentum_penalty} and \ref{eq:energy} in case of a temperature field characterised by an initial perturbation and with
-a temperature-dependent density.
+(Section \ref{sec:simple_shear}) and Exercise 9.4 in \citet{Gerya2010b} (Section \ref{sec:shear}). Furthermore, mantle convection and thin layer entrainment
+benchmarks (Sections \ref{sec:mantle} and \ref{sec:thin}, respectively) are performed to verify that the code solve correctly Eqs. \ref{eq:momentum_penalty}
+and \ref{eq:energy} in case of a temperature field characterised by an initial perturbation and with a temperature-dependent density. At last, the viscoplastic
+mantle convection benchmark is performed to check also the viscoplasticity in case of a initial perturbed temperature field (Section \ref{sec:visco_mantle}).
 
@@
 
+\subsection{Viscoplastic mantle convection}\label{sec:visco_mantle}
+This benchmark is performed as described in \citet{Tosi2015}.
+
 \subsection{Sinking hydrated cylinder}\label{sec:quinquis}

### 2021-05-05 Paper.tex
modified

@@
-\documentclass[hidelinks,11pt,a4paper]{article}
+\documentclass[hidelinks,10pt,a4paper]{article}
 \usepackage[utf8]{inputenc}
@@
 \usepackage{hyperref}
+\usepackage{booktabs}
+\usepackage{multirow}
 \usepackage[cm]{fullpage}
@@
 \usepackage{bm}
-\usepackage{xcolor}
+\usepackage[table]{xcolor}
 \hypersetup{
@@
 \begin{tabular}{| l | l |}
-\hline
+\toprule
  \textbf{Feature}  & \textbf{Benchmark}  \\
-\hline
-\hline
+\midrule
   Solver  & Stokes flow   \\
 \hline
-  4th-order Runge-Kutta  &  \\
-  markers advection  & Zalesak disk   \\
+  4th-order Runge-Kutta  & \multirow{2}{*}{Zalesak disk} \\
+  markers advection  &  \\
 \hline
-  Momentum equation  & Poiseuille flow   \\
+  \multirow{4}{*}{Momentum equation}  & Poiseuille flow   \\
     & Instantaneous 2D sphere   \\
@@
 \hline
-  Sticky air and  & 2D Stokes sphere \\
-  free surface  & Stabilisation algorithm   \\
-  & Topography relaxation   \\
+    & 2D Stokes sphere \\
+    Sticky air and  & Stabilisation algorithm   \\
+    free surface & Topography relaxation   \\
   & Spontaneous subduction   \\
@@
 \hline
-  Energy equation  & Advection stabilisation   \\
+  \multirow{5}{*}{Energy equation}  & Advection stabilisation   \\
     & Simple shear heating \\
@@
 \hline
-  Phase changes  &  \\
-  and hydration  & Hydrated sinking cylinder   \\
+  Phase changes  & \multirow{2}{*}{Hydrated sinking cylinder} \\
+  and hydration  &  \\
 \hline
   Melting & Experimental melting curves \\
-\hline
+\bottomrule
 \end{tabular}
@@
 $\bm{\mathcal{R}}^i$ and it is normalised as follows
-\begin{equation}\label{eq:convergence}
+\begin{equation}\label{eq:convergence_res}
 \frac{||\bm{\mathcal{R}}^i||_2}{||\bm{\mathcal{R}}^0||_2}
@@
 \begin{eqnarray}
-err_p&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|p^n_i(r_q)-p^a_i(r_q)|^2\right)w_q|J_q|}\nonumber \\
-err_v&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|u^n_i(r_q)-u^a_i(r_q)|^2+|v^n_i(r_q)-v^a_i(r_q)|^2\right)w_q|J_q|}\nonumber
+\textrm{err}_p&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|p^n_i(r_q)-p^a_i(r_q)|^2\right)w_q|J_q|}\nonumber \\
+\textrm{err}_v&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|u^n_i(r_q)-u^a_i(r_q)|^2+|v^n_i(r_q)-v^a_i(r_q)|^2\right)w_q|J_q|}\nonumber
 \end{eqnarray}
@@
 Other important quantities calculated in some benchmarks are the root-mean-square velocity and the Nusselt number, calculated as
-\[v_{rms}=\sqrt{\frac{1}{V_{\Omega}}\int_{\Omega}|\bm{v}|^2 d\Omega}\]
+\[v_{\textrm{rms}}=\sqrt{\frac{1}{V_{\Omega}}\int_{\Omega}|\bm{v}|^2 d\Omega}\]
 and
-\[Nu=\frac{\int_{top}\bm{\nabla} T \cdot \bm{n} ds}{\int_{bottom}T ds}\]
+\[\textrm{Nu}=\frac{\int_{top}\bm{\nabla} T \cdot \bm{n} ds}{\int_{bottom}T ds}\]
 respectively.
@@
 and different average schemes for the viscosity (harmonic, geometric and arithmetic). The velocity in the centre of the sphere $\bm{v}(0.5;0.5)$), minimum and
-maximum velocities ($u_{min}$, $u_{max}$, $v_{min}$, $v_{max}$) and pressures ($p_{min}$, $p_{max}$) on the entire domain, $v_{rms}$ and average pressure on the
+maximum velocities ($u_{min}$, $u_{max}$, $v_{min}$, $v_{max}$) and pressures ($p_{min}$, $p_{max}$) on the entire domain, $v_{\textrm{rms}}$ and average pressure on the
 entire domain ($p_{avg}$) are compared with solutions generated by ASPECT \citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} and all results
-can be found at \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere2D}. Results in terms of the $v_{rms}$ for NS model are also shown
+can be found at \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere2D}. Results in terms of the $v_{\textrm{rms}}$ for NS model are also shown
 in Fig. \ref{fig:inst_sphere}.
@@
 \noindent\includegraphics[width=400px]{./Figures/vrms_NS.pdf}
-\caption{$v_{rms}$ as function of element size for the instantaneous 2D sphere experiment in case of no slip boundary conditions and with different average
+\caption{$v_{\textrm{rms}}$ as function of element size for the instantaneous 2D sphere experiment in case of no slip boundary conditions and with different average
 schemes for the viscosity. Results are compared with results obtained by other numerical codes.}
@@
 
-$v_{rms}$ as function of time is reported in Fig. \ref{fig:RT}, matching well with results shown by \citet{vanKeken1997}, \citet{Tackley2003} and
+$v_{\textrm{rms}}$ as function of time is reported in Fig. \ref{fig:RT}, matching well with results shown by \citet{vanKeken1997}, \citet{Tackley2003} and
 \citet{Thieulot2014}. Fig. \ref{fig:rayleigh} shows the evolution of the experiment at different time steps.
@@
 \noindent\includegraphics[width=400px]{./Figures/RT.pdf}
-\caption{$v_{rms}$ of the Rayleigh-Taylor experiment as function of time for different resolution of the grid.}
+\caption{$v_{\textrm{rms}}$ of the Rayleigh-Taylor experiment as function of time for different resolution of the grid.}
 \label{fig:RT}
@@
 \citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} and they can be found at
-\url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere_fs2D}. Results in terms of the $v_{rms}$ in case of an arithmetic average are also
+\url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere_fs2D}. Results in terms of the $v_{\textrm{rms}}$ in case of an arithmetic average are also
 shown in Fig. \ref{fig:inst_spherefs}.
@@
 \noindent\includegraphics[width=400px]{./Figures/vrms_arithm.pdf}
-\caption{$v_{rms}$ for different grid resolutions as function of time for the 2D time-dependent Stokes sphere experiment in case of an arithmetic average.
+\caption{$v_{\textrm{rms}}$ for different grid resolutions as function of time for the 2D time-dependent Stokes sphere experiment in case of an arithmetic average.
 Results are compared with results obtained by ASPECT.}
@@
 bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)-0.01 \cos (\pi x) \sin (\pi x)\]
+\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\]
 
-The solution generated by the code in terms of $v_{rms}$ (calculated as in Sec. \ref{sec:error}) and $Nu$ as function of time are reported for all the
-simulations in Fig. \ref{fig:mantle}. At the steady state, $v_{rms}$ and $Nu$ of all simulations converge well toward the values from \citet{Blankenbach1989},
+The solution generated by the code in terms of $v_{\textrm{rms}}$ (calculated as in Sec. \ref{sec:error}) and Nu as function of time are reported for all the
+simulations in Fig. \ref{fig:mantle}. At the steady state, $v_{\textrm{rms}}$ and Nu of all simulations converge well toward the values from \citet{Blankenbach1989},
 with lower errors for higher resolution grids (Table \ref{tab:mantle}).
@@
 \noindent\includegraphics[width=400px]{./Figures/Convection.pdf}
-\caption{$v_{rms}$ (panels a, c and e) and $Nu$ (panels b, d and f) for the mantle convection benchmark as function of time for different grid resolution.
+\caption{$v_{\textrm{rms}}$ (panels a, c and e) and Nu (panels b, d and f) for the mantle convection benchmark as function of time for different grid resolution.
 Panels a and b show the results for $Ra=10^4$; Panels c and d show the results for $Ra=10^5$; panels e and f show the results for $Ra=10^6$. Purple lines
-indicate the convergence values for the $v_{rms}$ and the $Nu$ from \citet{Blankenbach1989} at the steady state.}
+indicate the convergence values for the $v_{\textrm{rms}}$ and the Nu from \citet{Blankenbach1989} at the steady state.}
 \label{fig:mantle}
@@
 \begin{table}[H]
-\caption{Comparison between $v_{rms}$ and $Nu$ predicted by the code for the mantle convection experiment and same values as reported in literature.}
+\caption{Comparison between $v_{\textrm{rms}}$ and Nu predicted by the code for the mantle convection experiment and same values as reported in literature.}
 \centering
 \small
-\begin{tabular}{|p{0.09\textwidth}|p{0.05\textwidth}|p{0.24\textwidth}|p{0.15\textwidth}|p{0.1\textwidth}p{0.1\textwidth}p{0.1\textwidth}|}
+\begin{tabular}{|>{\centering}m{0.09\textwidth}|>{\centering}m{0.05\textwidth}|>{\centering}m{0.24\textwidth}|>{\centering}m{0.15\textwidth}|
+  >{\centering}m{0.1\textwidth} >{\centering}m{0.1\textwidth} >{\centering\arraybackslash}m{0.1\textwidth}|}
+\toprule
+& & \multirow{2}{*}{\citet{Blankenbach1989}} & \citet{Thieulot2014} & \multicolumn{3}{c|}{FALCON}  \\
+ & & & \num{200x200} & \num{32x32} & \num{64x64} & \num{128x128}  \\
+\midrule
+\multirow{2}{*}{$Ra=10^4$} & $v_{\textrm{rms}}$ & $42.864947 \pm 0.000020$ & 42.867 & 42.83226 & 42.852793 & 42.861394  \\
+           & Nu      & $4.884409 \pm 0.000010$  & 4.882  & 4.781297 & 4.857475  & 4.877573   \\
 \hline
-  & & \citet{Blankenbach1989} & \citet{Thieulot2014} & & &   \\
- & & & \num{200x200} & \num{32x32} & \num{64x64} & \num{128x128}  \\
+\multirow{2}{*}{$Ra=10^5$} & $v_{\textrm{rms}}$ & $193.21454 \pm 0.00010$  & 193.255& 193.872643 & 193.377472 & 193.252290  \\
+           & Nu      & $10.534095 \pm 0.000010$ & 10.507 & 9.602514 & 10.270735  & 10.465629   \\
 \hline
-\hline
- $Ra=10^4$ & $v_{rms}$ & $42.864947 \pm 0.000020$ & 42.867 & 42.83226 & 42.852793 & 42.861394  \\
-           & $Nu$      & $4.884409 \pm 0.000010$  & 4.882  & 4.781297 & 4.857475  & 4.877573   \\
-\hline
- $Ra=10^5$ & $v_{rms}$ & $193.21454 \pm 0.00010$  & 193.255& 193.872643 & 193.377472 & 193.252290  \\
-           & $Nu$      & $10.534095 \pm 0.000010$ & 10.507 & 9.602514 & 10.270735  & 10.465629   \\
-\hline
- $Ra=10^6$ & $v_{rms}$ & $833.98977 \pm 0.00020$  & 834.712& 848.091176 & 837.767911 & 834.945793  \\
-           & $Nu$      & $21.972465 \pm 0.000020$ & 21.695 & 15.999266 & 19.703682  & 21.306939   \\
-\hline
+\multirow{2}{*}{$Ra=10^6$} & $v_{\textrm{rms}}$ & $833.98977 \pm 0.00020$  & 834.712& 848.091176 & 837.767911 & 834.945793  \\
+           & Nu      & $21.972465 \pm 0.000020$ & 21.695 & 15.999266 & 19.703682  & 21.306939   \\
+\bottomrule
 \end{tabular}
@@
 
-The experiment is performed with two grid resolution (\num{125x40} and \num{200x80} elements), with 100 markers per element and CFL of 0.25. $v_{rms}$
+The experiment is performed with two grid resolution (\num{125x40} and \num{200x80} elements), with 100 markers per element and CFL of 0.25. $v_{\textrm{rms}}$
 (see Sec. \ref{sec:error}) calculated for both the simulations match well with results from \citet{vanKeken1997} and obtained with ASPECT
@@
 \noindent\includegraphics[width=400px]{./Figures/vrms_thin.pdf}
-\caption{$v_{rms}$ for the thin layer experiment as function of time for different grid resolution. Results are compared with results obtained by
+\caption{$v_{\textrm{rms}}$ for the thin layer experiment as function of time for different grid resolution. Results are compared with results obtained by
 \citet{vanKeken1997} (black lines) and with ASPECT and ELEFANT.}
@@
 \noindent\includegraphics[width=400px]{./Figures/vrmszero.pdf}
-\caption{$v_{rms}$ at $t=0$ for the thin layer experiment as function of the element size for different aspect ratios of the elements. Results are compared
+\caption{$v_{\textrm{rms}}$ at $t=0$ for the thin layer experiment as function of the element size for different aspect ratios of the elements. Results are compared
 with results obtained by \citet{vanKeken1997} (black dots in the grey area) and with ASPECT and ELEFANT.}
@@
 \subsection{Viscoplastic mantle convection}\label{sec:visco_mantle}
-This benchmark is performed as described in \citet{Tosi2015}.
+This benchmark is performed as described in \citet{Tosi2015} for Cases 1-5a. Table \ref{tab:visco_mantle}.WORK IN PROGRESS.
+
+\renewcommand{\arraystretch}{1.5}
+{\rowcolors{3}{white!100!}{gray!25!}
+\begin{table}[H]
+  \caption{Results of the viscoplastic mantle convection benchmark for Cases 1-4 compared with results obtained with ELEFANT and YACC (from \citealp{Tosi2015}).}
+  \centering
+  \resizebox{\textwidth}{!}{
+\begin{tabular}{|c|ccc|ccc|ccc|ccc|}
+  \toprule
+    & \multicolumn{3}{c|}{\textbf{Case 1}} & \multicolumn{3}{c|}{\textbf{Case 2}} & \multicolumn{3}{c|}{\textbf{Case 3}} &
+    \multicolumn{3}{c|}{\textbf{Case 4}} \\
+   & FALCON & ELEFANT& YACC & FALCON& ELEFANT& YACC & FALCON & ELEFANT & YACC & FALCON & ELEFANT & YACC \\
+  \midrule
+   $\left\langle T \right\rangle$ & 0.7759 & 0.7758 & 0.7767 & & 0.6036 & 0.6058 & 0.7275 & 0.7275 & 0.7286 & & 0.5277 & 0.5289\\
+    Nu$^{top}$ & 3.4280 & 3.4214 & 3.4298 & & 8.5115 & 8.5278 & 3.0368 & 3.0347 & 3.0374 && 6.5912 & 6.5572 \\
+    Nu$^{bottom}$ & 3.3168 & 3.313 & 3.3143 & & 8.4465 & 8.3990 & 3.0086 & 2.9908 & 2.9628 && 6.5834 & 6.5243 \\
+    $v_{\textrm{rms}}$ & 249.901 & 249.134 & 251.7997 & & 140.7087 & 142.2020 & 100.2776 & 100.1208 & 100.9467 && 79.1105 & 79.6202 \\
+    $v_{\textrm{rms}}^{top}$ & 1.8907 & 1.8642 & 1.8298 & & 104.1228 & 105.1767 & 2.0832 & 2.0652 & 2.0374 && 74.7596 & 75.8414 \\
+    $u_{max}^{top}$ & 2.6340 & 2.6119 & 2.5516 & & 121.766 & 122.3238 & 2.9086 & 2.9019 & 2.8458 && 88.9146 & 89.2940 \\
+    $\eta_{min}$ & - & - & - & & $2.0600\times10^{-5}$ & $1.9853\times10^{-5}$ & $4.8129\times10^{-5}$ & $4.8080\times10^{-5}$ & $4.7907\times10^{-5}$ &
+    & $1.9860\times10^{-4}$ & $1.9174\times10^{-4}$ \\
+    $\eta_{max}$ & - & - & - & & 1.85 & 1.6095 & 0.9360 & 0.9023 & 1 && 1.52 & 1.6773 \\
+    $\left\langle W \right\rangle$ & 2.4299 & 2.4316 & 2.4583 & & 7.5665 & 7.6222 & 2.0387 & 2.0384 & 2.0400 && 5.6216 & 5.6512 \\
+    $\left\langle \Phi \right\rangle/Ra$ & 2.4311 & 2.4276 & 2.4333 & & 7.5585 & 7.6194 & 2.0386 & 2.0370 & 2.0335 && 5.6182 & 5.6463 \\
+    $\delta$ & 0.05\% & 0.16\% & 1.02\% & & 0.11\% & 0.04\% & $<$0.01\% & 0.07\% & 0.32\% && 0.06\% & 0.09\% \\
+  \bottomrule
+  \end{tabular}}
+  \label{tab:visco_mantle}
+  \end{table}
 

### 2021-05-13 Paper.tex
modified

@@
 \midrule
-  Solver  & Stokes flow   \\
+  Solver  & Stokes flow (Section \ref{sec:stokes})  \\
 \hline
-  4th-order Runge-Kutta  & \multirow{2}{*}{Zalesak disk} \\
+  4th-order Runge-Kutta  & \multirow{2}{*}{Zalesak disk (Section \ref{sec:runge})} \\
   markers advection  &  \\
 \hline
-  \multirow{4}{*}{Momentum equation}  & Poiseuille flow   \\
-    & Instantaneous 2D sphere   \\
-    & Rayleigh-Taylor experiment  \\
-    & Falling block  \\
+  \multirow{4}{*}{Momentum equation}  & Poiseuille flow (Section \ref{sec:poiseuille}) \\
+    & Instantaneous 2D sphere (Section \ref{sec:ist_sphere})   \\
+    & Rayleigh-Taylor experiment (Section \ref{sec:rayleigh})  \\
+    & Falling block (Section \ref{sec:block})  \\
 \hline
-    & 2D Stokes sphere \\
-    Sticky air and  & Stabilisation algorithm   \\
-    free surface & Topography relaxation   \\
-  & Spontaneous subduction   \\
+    & 2D Stokes sphere (Section \ref{sec:sphere}) \\
+    Sticky air and  & Stabilisation algorithm (Section \ref{sec:stab})   \\
+    free surface & Topography relaxation (Section \ref{sec:crameri})   \\
+  & Spontaneous subduction (Section \ref{sec:subduction})   \\
 \hline
@@
 \hline
-   Non-linear  & Slab detachment   \\
-   visco-plasticity & Indenter   \\
-    & Brick   \\
+   Non-linear  & Slab detachment (Section \ref{sec:slab})   \\
+   visco-plasticity & Indenter (Section \ref{sec:indenter})   \\
+    & Brick (Section \ref{sec:brick})   \\
 \hline
-  \multirow{5}{*}{Energy equation}  & Advection stabilisation   \\
-    & Simple shear heating \\
-    & Shear and adiabatic heating \\
-    & Mantle convection   \\
-    & Thin layer entrainment   \\
+  \multirow{5}{*}{Energy equation}  & Advection stabilisation (Section \ref{sec:advection})   \\
+    & Simple shear heating (Section \ref{sec:simple_shear}) \\
+    & Shear and adiabatic heating (Section \ref{sec:shear}) \\
+    & Mantle convection (Section \ref{sec:mantle})   \\
+    & Thin layer entrainment (Section \ref{sec:thin})   \\
+    & Viscoplastic mantle convection (Section \ref{sec:visco_mantle})   \\
 \hline
-  Phase changes  & \multirow{2}{*}{Hydrated sinking cylinder} \\
+  Phase changes  & \multirow{2}{*}{Hydrated sinking cylinder (Section \ref{sec:quinquis})} \\
   and hydration  &  \\
 \hline
-  Melting & Experimental melting curves \\
+  Melting & Experimental melting curves (Section \ref{sec:katz}) \\
 \bottomrule

### 2021-05-18 Paper.tex
modified

@@
 \subsection{Viscoplastic mantle convection}\label{sec:visco_mantle}
-This benchmark is performed as described in \citet{Tosi2015} for Cases 1-5a. Table \ref{tab:visco_mantle}.WORK IN PROGRESS.
+This benchmark is performed as Cases 1-5a described in \citet{Tosi2015}. Table \ref{tab:tosi_setup}. Table \ref{tab:visco_mantle}.WORK IN PROGRESS.
+
+%\renewcommand{\arraystretch}{1.5}
+%{\rowcolors{3}{white!100!}{gray!25!}
+%\begin{table}[H]
+%  \caption{Results of the viscoplastic mantle convection benchmark for Cases 1-4 compared with results obtained with ELEFANT and YACC (from \citealp{Tosi2015}).}
+%  \centering
+%  \resizebox{\textwidth}{!}{
+%\begin{tabular}{|c|ccc|ccc|ccc|ccc|}
+%  \toprule
+%    & \multicolumn{3}{c|}{\textbf{Case 1}} & \multicolumn{3}{c|}{\textbf{Case 2}} & \multicolumn{3}{c|}{\textbf{Case 3}} &
+%    \multicolumn{3}{c|}{\textbf{Case 4}} \\
+%   & FALCON & ELEFANT& YACC & FALCON& ELEFANT& YACC & FALCON & ELEFANT & YACC & FALCON & ELEFANT & YACC \\
+%  \midrule
+%   $\left\langle T \right\rangle$ & 0.7759 & 0.7758 & 0.7767 & 0.6043 & 0.6036 & 0.6058 & 0.7275 & 0.7275 & 0.7286 & & 0.5277 & 0.5289\\
+%    Nu$^{top}$ & 3.4280 & 3.4214 & 3.4298 & 8.5242 & 8.5115 & 8.5278 & 3.0368 & 3.0347 & 3.0374 && 6.5912 & 6.5572 \\
+%    Nu$^{bottom}$ & 3.3168 & 3.313 & 3.3143 & 8.4599 & 8.4465 & 8.3990 & 3.0086 & 2.9908 & 2.9628 && 6.5834 & 6.5243 \\
+%    $v_{\textrm{rms}}$ & 249.901 & 249.134 & 251.7997 & 141.2451 & 140.7087 & 142.2020 & 100.2776 & 100.1208 & 100.9467 && 79.1105 & 79.6202 \\
+%    $v_{\textrm{rms}}^{top}$ & 1.8907 & 1.8642 & 1.8298 & 104.7928 & 104.1228 & 105.1767 & 2.0832 & 2.0652 & 2.0374 && 74.7596 & 75.8414 \\
+%    $u_{max}^{top}$ & 2.6340 & 2.6119 & 2.5516 & 121.9212& 121.766 & 122.3238 & 2.9086 & 2.9019 & 2.8458 && 88.9146 & 89.2940 \\
+%    $\eta_{min}$ & - & - & - & $2.0303\times10^{-5}$ & $2.0600\times10^{-5}$ & $1.9853\times10^{-5}$ & $4.8129\times10^{-5}$ & $4.8080\times10^{-5}$ & $4.7907\times10^{-5}$ &
+%    & $1.9860\times10^{-4}$ & $1.9174\times10^{-4}$ \\
+%    $\eta_{max}$ & - & - & - & 1.3325 & 1.85 & 1.6095 & 0.9360 & 0.9023 & 1 && 1.52 & 1.6773 \\
+%    $\left\langle W \right\rangle$ & 2.4299 & 2.4316 & 2.4583 & 7.5648 & 7.5665 & 7.6222 & 2.0387 & 2.0384 & 2.0400 && 5.6216 & 5.6512 \\
+%    $\left\langle \Phi \right\rangle/Ra$ & 2.4311 & 2.4276 & 2.4333 & 7.5661 & 7.5585 & 7.6194 & 2.0386 & 2.0370 & 2.0335 && 5.6182 & 5.6463 \\
+%    $\delta$ & 0.05\% & 0.16\% & 1.02\% & 0.02\% & 0.11\% & 0.04\% & $<$0.01\% & 0.07\% & 0.32\% && 0.06\% & 0.09\% \\
+%  \bottomrule
+%  \end{tabular}}
+%  \label{tab:visco_mantle}
+%  \end{table}
+%\renewcommand{\arraystretch}{1.5}
+%{\rowcolors{2}{white!100!}{gray!25!}
+\begin{table}[H]
+  \caption{Benchmark cases and corresponding parameters.}
+  \centering
+  \small
+\begin{tabular}{|>{\centering}m{0.08\textwidth}|>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}
+  >{\centering}m{0.05\textwidth} >{\centering}m{0.05\textwidth} |>{\centering\arraybackslash}m{0.2\textwidth}|}
+  \toprule
+    Case & $Ra$ & $\Delta\eta_T$ & $\Delta\eta_y$ & $\eta^*$ & $\sigma_Y$ & Convective regime \\
+  \midrule
+  \rowcolor{gray!25!}
+    1 & $10^2$ & $10^5$ & 1 & - & - & Stagnant lid \\
+    2 & $10^2$ & $10^5$ & 1 & $10^{-3}$ & 1 & Mobile lid  \\
+    \rowcolor{gray!25!}
+    3 & $10^2$ & $10^5$ & 10 & - & - & Stagnant lid  \\
+    4 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 1 & Mobile lid \\
+    \rowcolor{gray!25!}
+    5 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 4 & Periodic \\
+  \bottomrule
+  \end{tabular}
+  \label{tab:tosi_setup}
+  \end{table}
 
 \renewcommand{\arraystretch}{1.5}
-{\rowcolors{3}{white!100!}{gray!25!}
+%{\rowcolors{3}{gray!25!}{white!100!}
 \begin{table}[H]
-  \caption{Results of the viscoplastic mantle convection benchmark for Cases 1-4 compared with results obtained with ELEFANT and YACC (from \citealp{Tosi2015}).}
+  \caption{Results of the viscoplastic mantle convection benchmark for different grid resolutions in Cases 1-4 compared with results obtained with ELEFANT
+  and YACC ($100\times100$ elements, from \citealp{Tosi2015}) and in Stone 28 ($32\times32$ elements, from
+  \url{https://github.com/cedrict/fieldstone/tree/master/python_codes/fieldstone_28}).}
   \centering
   \resizebox{\textwidth}{!}{
-\begin{tabular}{|c|ccc|ccc|ccc|ccc|}
+\begin{tabular}{|cc|ccccccccccc|}
   \toprule
-    & \multicolumn{3}{c|}{\textbf{Case 1}} & \multicolumn{3}{c|}{\textbf{Case 2}} & \multicolumn{3}{c|}{\textbf{Case 3}} &
-    \multicolumn{3}{c|}{\textbf{Case 4}} \\
-   & FALCON & ELEFANT& YACC & FALCON& ELEFANT& YACC & FALCON & ELEFANT & YACC & FALCON & ELEFANT & YACC \\
+  \multicolumn{13}{|c|}{\textbf{Case 1}} \\
   \midrule
-   $\left\langle T \right\rangle$ & 0.7759 & 0.7758 & 0.7767 & & 0.6036 & 0.6058 & 0.7275 & 0.7275 & 0.7286 & & 0.5277 & 0.5289\\
-    Nu$^{top}$ & 3.4280 & 3.4214 & 3.4298 & & 8.5115 & 8.5278 & 3.0368 & 3.0347 & 3.0374 && 6.5912 & 6.5572 \\
-    Nu$^{bottom}$ & 3.3168 & 3.313 & 3.3143 & & 8.4465 & 8.3990 & 3.0086 & 2.9908 & 2.9628 && 6.5834 & 6.5243 \\
-    $v_{\textrm{rms}}$ & 249.901 & 249.134 & 251.7997 & & 140.7087 & 142.2020 & 100.2776 & 100.1208 & 100.9467 && 79.1105 & 79.6202 \\
-    $v_{\textrm{rms}}^{top}$ & 1.8907 & 1.8642 & 1.8298 & & 104.1228 & 105.1767 & 2.0832 & 2.0652 & 2.0374 && 74.7596 & 75.8414 \\
-    $u_{max}^{top}$ & 2.6340 & 2.6119 & 2.5516 & & 121.766 & 122.3238 & 2.9086 & 2.9019 & 2.8458 && 88.9146 & 89.2940 \\
-    $\eta_{min}$ & - & - & - & & $2.0600\times10^{-5}$ & $1.9853\times10^{-5}$ & $4.8129\times10^{-5}$ & $4.8080\times10^{-5}$ & $4.7907\times10^{-5}$ &
-    & $1.9860\times10^{-4}$ & $1.9174\times10^{-4}$ \\
-    $\eta_{max}$ & - & - & - & & 1.85 & 1.6095 & 0.9360 & 0.9023 & 1 && 1.52 & 1.6773 \\
-    $\left\langle W \right\rangle$ & 2.4299 & 2.4316 & 2.4583 & & 7.5665 & 7.6222 & 2.0387 & 2.0384 & 2.0400 && 5.6216 & 5.6512 \\
-    $\left\langle \Phi \right\rangle/Ra$ & 2.4311 & 2.4276 & 2.4333 & & 7.5585 & 7.6194 & 2.0386 & 2.0370 & 2.0335 && 5.6182 & 5.6463 \\
-    $\delta$ & 0.05\% & 0.16\% & 1.02\% & & 0.11\% & 0.04\% & $<$0.01\% & 0.07\% & 0.32\% && 0.06\% & 0.09\% \\
+  Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
+  & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
+  \hline
+  \rowcolor{gray!25!}
+    YACC & $100\times100$ & 0.7767 & 3.4298 & 3.3143 & 251.7997 & 1.8298 & 2.5516 & - & - & 2.4583 & 2.4333 & 1.02\% \\
+    ELEFANT & $100\times100$ & 0.7758 & 3.4214 & 3.313 & 249.134 & 1.8642 & 2.6119 & - & - & 2.4316 & 2.4276 & 0.16\% \\
+    \rowcolor{gray!25!}
+    Stone 28 & $32\times32$ & 0.7737 & 3.3987 & - & 243.872 & - & - & - & - & - & - & - \\
+%    FALCON & $32\times32$ & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & - & - & 2.4745 & 2.4864 & 0.48\% \\
+%    \rowcolor{gray!25!}
+    FALCON & $100\times100$ & 0.7759 & 3.4280 & 3.3168 & 249.901 & 1.8907 & 2.6340 & - & - & 2.4299 & 2.4311 & 0.05\% \\
+    \midrule
+    \multicolumn{13}{|c|}{\textbf{Case 2}} \\
+    \midrule
+    Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
+    & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
+    \hline
+    \rowcolor{gray!25!}
+     YACC & $100\times100$ & 0.6058 & 8.5278 & 8.3990 & 142.2020 & 105.1767 & 122.3238 & $1.9853\times10^{-5}$ & 1.6095 & 7.6222 & 7.6194 & 0.04\% \\
+     ELEFANT & $100\times100$ & 0.6034 & 8.5115 & 8.4465 & 140.8067 & 104.1228 & 121.766 & $2.0600\times10^{-5}$ & 1.85 & 7.5665 & 7.5585 & 0.11\% \\
+     \rowcolor{gray!25!}
+     Stone 28 & $32\times32$ & 0.6052 & 8.1715 & - & 141.518 & - & - & - & - & - & - & - \\
+%     FALCON & $32\times32$ & 0.6454 & 8.1422 & 7.4020 & 156.6908 & 95.7226 & 116.9302 & $2.2572\times10^{-5}$ & 0.7194 & 7.4289 & 7.4452 & 0.22\% \\
+%     \rowcolor{gray!25!}
+     FALCON & $100\times100$ & 0.6044 & 8.5266 & 8.4590 & 141.3684 & 104.8867 & 122.0470 & $2.0333\times10^{-5}$ & 1.2914 & 7.5707 & 7.5718 & 0.01\% \\
+     \midrule
+     \multicolumn{13}{|c|}{\textbf{Case 3}} \\
+     \midrule
+     Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
+     & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
+     \hline
+     \rowcolor{gray!25!} 
+     YACC & $100\times100$ & 0.7286 & 3.0374 & 2.9628 & 100.9467 & 2.0374 & 2.8458 & $4.7907\times10^{-5}$ & 1 & 2.0400 & 2.0335 & 0.32\% \\
+      ELEFANT & $100\times100$ & 0.7275 & 3.0347 & 2.9908 & 100.1208 & 2.0652 & 2.9019 & $4.8080\times10^{-5}$ & 0.9023 & 2.0384 & 2.037 & 0.07\% \\
+      \rowcolor{gray!25!}
+      Stone 28 & $32\times32$ & 0.7277 & 3.0347 & - & 100.018 & - & - & - & - & - & - & - \\
+%      FALCON & $32\times32$ & 0.7281 & 3.0618 & 2.8132 & 101.0092 & 2.1630 & 3.0178 & $4.8500\times10^{-5}$ & 0.7845 & 2.0699 & 2.0695 & 0.02\% \\
+%      \rowcolor{gray!25!}
+      FALCON & $100\times100$ & 0.7275 & 3.0368 & 3.0086 & 100.2776 & 2.0832 & 2.9086 & $4.8129\times10^{-5}$ & 0.9360 & 2.0387 & 2.0386 & $<$0.01\% \\
+      \midrule
+      \multicolumn{13}{|c|}{\textbf{Case 4}} \\
+      \midrule
+      Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
+      & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
+      \hline
+      \rowcolor{gray!25!} 
+      YACC & $100\times100$ & 0.5289 & 6.5572 & 6.5243 & 79.6202 & 75.4814 & 89.2940 & $1.9174\times10^{-4}$ & 1.6773 & 5.6512 & 5.6463 & 0.09\% \\
+       ELEFANT & $100\times100$ & 0.5277 & 6.5912 & 6.5834 & 79.1105 & 74.7596 & 88.9146 & $1.9860\times10^{-4}$ & 1.5200 & 5.6216 & 5.6182 & 0.06\% \\
+       \rowcolor{gray!25!}
+       Stone 28 & $32\times32$ & 0.5286 & 6.4036 & - & 79.4746 & - & - & - & - & - & - & - \\
+%       FALCON & $32\times32$ & 0.5334 & 6.4679 & 6.3435 & 80.6153 & 76.0511 & 90.2058 & $2.0393\times10^{-4}$ & 0.6453 & 5.6676 & 5.6698 & 0.04\% \\
+%       \rowcolor{gray!25!}
+       FALCON & $100\times100$ & 0.5279 & 6.5897 & 6.5830 & 79.1220 & 75.0715 & 88.8484 & $1.9693\times10^{-4}$ & 1.4133 & 5.6127 & 5.6113 & 0.03\% \\
   \bottomrule

### 2021-05-21 Paper.tex
modified

@@
 This supplement includes a description of the code, with a presentation of the features implemented followed by results of extensive benchmarks listed in
-Table \ref{tab:list}, implemented to verify that the code solves correctly a large range of problems. Description and results of the benchmarks are shown
+Table \ref{tab:list}, implemented to verify that the code solves correctly a large range of problems. Description and results of all benchmarks are shown
 in Section \ref{sec:benchmark}.
@@
 \centering
-\begin{tabular}{| l | l |}
+\begin{tabular}{| l | l | l |}
 \toprule
- \textbf{Feature}  & \textbf{Benchmark}  \\
+ \textbf{Feature}  & \textbf{Benchmark} & \textbf{Section} \\
 \midrule
-  Solver  & Stokes flow (Section \ref{sec:stokes})  \\
-\hline
-  4th-order Runge-Kutta  & \multirow{2}{*}{Zalesak disk (Section \ref{sec:runge})} \\
-  markers advection  &  \\
-\hline
-  \multirow{4}{*}{Momentum equation}  & Poiseuille flow (Section \ref{sec:poiseuille}) \\
-    & Instantaneous 2D sphere (Section \ref{sec:ist_sphere})   \\
-    & Rayleigh-Taylor experiment (Section \ref{sec:rayleigh})  \\
-    & Falling block (Section \ref{sec:block})  \\
-\hline
-    & 2D Stokes sphere (Section \ref{sec:sphere}) \\
-    Sticky air and  & Stabilisation algorithm (Section \ref{sec:stab})   \\
-    free surface & Topography relaxation (Section \ref{sec:crameri})   \\
-  & Spontaneous subduction (Section \ref{sec:subduction})   \\
-\hline
-  Erosion and sedimentation  & -  \\
-\hline
-   Non-linear  & Slab detachment (Section \ref{sec:slab})   \\
-   visco-plasticity & Indenter (Section \ref{sec:indenter})   \\
-    & Brick (Section \ref{sec:brick})   \\
-\hline
-  \multirow{5}{*}{Energy equation}  & Advection stabilisation (Section \ref{sec:advection})   \\
-    & Simple shear heating (Section \ref{sec:simple_shear}) \\
-    & Shear and adiabatic heating (Section \ref{sec:shear}) \\
-    & Mantle convection (Section \ref{sec:mantle})   \\
-    & Thin layer entrainment (Section \ref{sec:thin})   \\
-    & Viscoplastic mantle convection (Section \ref{sec:visco_mantle})   \\
-\hline
-  Phase changes  & \multirow{2}{*}{Hydrated sinking cylinder (Section \ref{sec:quinquis})} \\
-  and hydration  &  \\
-\hline
-  Melting & Experimental melting curves (Section \ref{sec:katz}) \\
+  Solver  & Stokes flow & \ref{sec:stokes} \\
+\hline
+  4th-order Runge-Kutta  & \multirow{2}{*}{Zalesak disk} & \multirow{2}{*}{\ref{sec:runge}} \\
+  markers advection  &  & \\
+\hline
+  \multirow{4}{*}{Momentum equation}  & Poiseuille flow & \ref{sec:poiseuille}\\
+    & Instantaneous 2D sphere & \ref{sec:ist_sphere} \\
+    & Rayleigh-Taylor experiment & \ref{sec:rayleigh} \\
+    & Falling block & \ref{sec:block} \\
+\hline
+    & 2D Stokes sphere & \ref{sec:sphere} \\
+    Sticky air and  & Stabilisation algorithm & \ref{sec:stab} \\
+    free surface & Topography relaxation & \ref{sec:crameri} \\
+  & Spontaneous subduction & \ref{sec:subduction} \\
+\hline
+  Erosion and sedimentation  & - & - \\
+\hline
+   Non-linear  & Slab detachment & \ref{sec:slab} \\
+   visco-plasticity & Indenter & \ref{sec:indenter} \\
+    & Brick & \ref{sec:brick} \\
+\hline
+  \multirow{3}{*}{Energy equation}  & Advection stabilisation & \ref{sec:advection} \\
+    & Simple shear heating & \ref{sec:simple_shear} \\
+    & Shear and adiabatic heating & \ref{sec:shear} \\
+\hline
+\multirow{3}{*}{Energy + momentum} & Mantle convection & \ref{sec:mantle} \\
+& Thin layer entrainment & \ref{sec:thin} \\
+& Viscoplastic mantle convection & \ref{sec:visco_mantle} \\
+\hline
+  Phase changes  & \multirow{2}{*}{Hydrated sinking cylinder} & \multirow{2}{*}{\ref{sec:quinquis}} \\
+  and hydration  &  & \\
+\hline
+  Melting & Experimental melting curves & \ref{sec:katz} \\
 \bottomrule

### 2021-05-26 Paper.tex
modified

@@
 \multirow{3}{*}{Energy + momentum} & Mantle convection & \ref{sec:mantle} \\
+& Viscoplastic mantle convection & \ref{sec:visco_mantle} \\
 & Thin layer entrainment & \ref{sec:thin} \\
-& Viscoplastic mantle convection & \ref{sec:visco_mantle} \\
 \hline
@@
 acceleration $g_y=10^{10}Ra$. The experiment is performed with three different Rayleigh numbers ($Ra=10^4$, $10^5$ and $10^6$) and with different grid
-resolution (between \num{32x32} and \num{128x128} elements). Both fluids have constant viscosity, initial density, heat capacity, thermal conductivity
+resolution (between \num{32x32} and \num{128x128} elements). The fluid has constant viscosity, initial density, heat capacity, thermal conductivity
 ($\eta= \rho_0= C_p = k=1$), reference temperature ($T_0=0$) and thermal expansion coefficient ($\alpha=10^{-10}$). Temperatures are set to 0 on top and 1 on
@@
 
-\subsection{Thin layer entrainment}\label{sec:thin}
-This experiment is performed as originally proposed by \citet{vanKeken1997} with two incompressible fluids in a rectangular domain with $L_x=2$ and $L_y=1$
-and gravity acceleration $g_y=10^{10}Ra$. The Rayleigh number and the compositional Rayleigh number are fixed to $Ra=300000$ and $Rc=450000$, respectively.
-Both fluids have constant viscosity, thermal conductivity, specific heat ($\eta=\rho=C_p=1$) and thermal expansion coefficient ($\alpha=10^{-10}$). Fluid 1
-has a density $\rho_1=1$, while fluid 2 is denser ($\rho_2=\rho_1+1.5 \cdot 10^{-10}$) and is placed at the bottom of the domain, for $y \leq 0.025$.
-Temperature are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides of the domain. The initial
-temperature field is given by
-\[T(x,y)=T_u(x,y)+T_l(x,y)+T_r(x,y)+T_s(x,y)-\frac{3}{2}\]
-with
-\begin{eqnarray}
-T_u(x,y)&=&\frac{1}{2}erf\left(\frac{1-y}{2}\sqrt{\frac{u_0}{x}}\right)\nonumber \\
-T_l(x,y)&=&1-\frac{1}{2}erf\left(\frac{y}{2}\sqrt{\frac{u_0}{L_x-x}}\right)\nonumber \\
-T_r(x,y)&=&\frac{1}{2}+\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{y+1}} \exp\left(-\frac{x^2u_0}{4y+4}\right)\nonumber \\
-T_s(x,y)&=&\frac{1}{2}-\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{2-y}} \exp\left(-\frac{(L_x-x)^2u_0}{8-4y}\right)\nonumber
-\end{eqnarray}
-and
-\begin{eqnarray}
-u_0&=&\frac{L_x^{7/3}}{(1+L_x^4)^{2/3}}\left(\frac{Ra}{2\sqrt{\pi}}\right)^{\frac{2}{3}}\nonumber \\
-Q&=&2\sqrt{\frac{L_x}{\pi u_0}}\nonumber
-\end{eqnarray}
-
-The experiment is performed with two grid resolution (\num{125x40} and \num{200x80} elements), with 100 markers per element and CFL of 0.25. $v_{\textrm{rms}}$
-(see Sec. \ref{sec:error}) calculated for both the simulations match well with results from \citet{vanKeken1997} and obtained with ASPECT
-\citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} and ELEFANT \citep{Thieulot2014}, obtained with similar grid resolutions
-(Fig. \ref{fig:thin}). In addition, a variety of simulations with different resolutions and aspect ratios are performed to verify that the sensitivity of the
-initial velocity field. Results match well with results from \citet{vanKeken1997} (in the grey area) and \citet{Thieulot2014} (Fig. \ref{fig:thin_initial}).
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_thin.pdf}
-\caption{$v_{\textrm{rms}}$ for the thin layer experiment as function of time for different grid resolution. Results are compared with results obtained by
-\citet{vanKeken1997} (black lines) and with ASPECT and ELEFANT.}
-\label{fig:thin}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrmszero.pdf}
-\caption{$v_{\textrm{rms}}$ at $t=0$ for the thin layer experiment as function of the element size for different aspect ratios of the elements. Results are compared
-with results obtained by \citet{vanKeken1997} (black dots in the grey area) and with ASPECT and ELEFANT.}
-\label{fig:thin_initial}
-\end{figure}
-
 \subsection{Viscoplastic mantle convection}\label{sec:visco_mantle}
-This benchmark is performed as Cases 1-5a described in \citet{Tosi2015}. Table \ref{tab:tosi_setup}. Table \ref{tab:visco_mantle}.WORK IN PROGRESS.
+This benchmark is performed as Cases 1-5a described in \citet{Tosi2015} in a 2D unit square domain with two different grid resolutions of \num{32x32} 
+and \num{100x100} elements. Reference viscosity, reference density, heat capacity, thermal conductivity and thermal expansion coefficient are fixed to a constant
+value of 1 ($\eta_0= \rho_0= C_p = k=\alpha=1$), while the gravity has been chosen as $g_y=10^2$, in order to obtain a Rayleigh number $Ra=10^2$. Temperatures
+are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
+\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\].
+The viscosity ﬁeld $\eta$ is calculated as the harmonic average between a linear part $\eta_{lin}$ that depends on temperature only or on temperature and 
+depth $y$ and a nonlinear-plastic part $\eta_{pl}$ that depends on the strain rate $\dot{\bm{\epsilon}}$, as follows
+\[\eta(T,y,\dot{\bm{\epsilon}})=2\left(\frac{1}{\eta_{lin}(T,y)}+\frac{1}{\eta_{pl}(\dot{\bm{\epsilon}})}\right)^{-1}\]
+The linear and the nonlinear-plastic parts of the viscosity are calculated following \citet{Tosi2015} as follows
+\[\eta_{lin}(T,y)=\exp(-\ln(\Delta\eta_T)T+\ln(\Delta\eta_y)y)\]
+\[\eta_{pl}(\dot{\bm{\epsilon}})=\eta^*+\frac{\sigma_Y}{\sqrt{\dot{\bm{\epsilon}}:\dot{\bm{\epsilon}}}}\]
+where $\Delta\eta_T$, $\Delta\eta_y$, $\eta^*$ and $\sigma_Y$ are parameters chosen as listed in Table \ref{tab:tosi_setup}.
+
+For each case, some diagnostic quantities measured at the steady state are reported in Table \ref{tab:visco_mantle} (see Section \ref{sec:error} for 
+the computation of all the quantities), showing a well fit with results described in \citet{Tosi2015}.
 
@@
 \begin{table}[H]
-  \caption{Results of the viscoplastic mantle convection benchmark for different grid resolutions in Cases 1-4 compared with results obtained with ELEFANT
+  \caption{Results of the viscoplastic mantle convection benchmark for different grid resolutions in Cases 1-5 compared with results obtained with ELEFANT
   and YACC ($100\times100$ elements, from \citealp{Tosi2015}) and in Stone 28 ($32\times32$ elements, from
@@
        FALCON & $100\times100$ & 0.5279 & 6.5897 & 6.5830 & 79.1220 & 75.0715 & 88.8484 & $1.9693\times10^{-4}$ & 1.4133 & 5.6127 & 5.6113 & 0.03\% \\
+       \midrule
+       \multicolumn{13}{|c|}{\textbf{Case 5}} \\
+       \midrule
+       Code & Resolution & $P$ & $\left\langle T \right\rangle_{min}$ & $\left\langle T \right\rangle_{max}$ & Nu$^{top}_{min}$ & Nu$^{top}_{max}$ &
+       $v_{\textrm{rms}}^{min}$ & $v_{\textrm{rms}}^{max}$ & \multicolumn{2}{c}{$(\left\langle \Phi \right\rangle/Ra)_{min}$} & 
+       \multicolumn{2}{c|}{$(\left\langle \Phi \right\rangle/Ra)_{max}$} \\
+       \hline
+       \rowcolor{gray!25!}
+         YACC & $100\times100$ & 0.0792 & 0.6543 & 0.6722 & 2.6950 & 7.2792 & 41.9813 & 98.8230 & \multicolumn{2}{c}{1.4347} & 
+         \multicolumn{2}{c|}{9.2748} \\
+         ELEFANT & $100\times100$ & 0.076 & 0.6514 & 0.6693 & 2.6808 & 7.3506 & 41.2578 & 100.089 & \multicolumn{2}{c}{1.3324} & 
+         \multicolumn{2}{c|}{9.5289} \\
+         \rowcolor{gray!25!}
+         Stone 28 & $32\times32$ & 0.0764 & 0.6521 & 0.6697 & 2.6912 & 7.0763 & 41.61 & 98.183 & \multicolumn{2}{c}{-} & 
+         \multicolumn{2}{c|}{-} \\
+     %    FALCON & $32\times32$ & & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & \multicolumn{2}{c}{-} & 
+     %     \multicolumn{2}{c|}{-} \\
+     %    \rowcolor{gray!25!}
+         FALCON & $100\times100$ & 0.0775 & 0.6519 & 0.6697 & 2.6847 & 7.3629 & 41.5256 & 99.5466 & \multicolumn{2}{c}{1.3399} & 
+         \multicolumn{2}{c|}{9.3993} \\
   \bottomrule
@@
   \end{table}
+
+\subsection{Thin layer entrainment}\label{sec:thin}
+This experiment is performed as originally proposed by \citet{vanKeken1997} with two incompressible fluids in a rectangular domain with $L_x=2$ and $L_y=1$
+and gravity acceleration $g_y=10^{10}Ra$. The Rayleigh number and the compositional Rayleigh number are fixed to $Ra=300000$ and $Rc=450000$, respectively.
+Both fluids have constant viscosity, thermal conductivity, specific heat ($\eta=\rho=C_p=1$) and thermal expansion coefficient ($\alpha=10^{-10}$). Fluid 1
+has a density $\rho_1=1$, while fluid 2 is denser ($\rho_2=\rho_1+1.5 \cdot 10^{-10}$) and is placed at the bottom of the domain, for $y \leq 0.025$.
+Temperature are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides of the domain. The initial
+temperature field is given by
+\[T(x,y)=T_u(x,y)+T_l(x,y)+T_r(x,y)+T_s(x,y)-\frac{3}{2}\]
+with
+\begin{eqnarray}
+T_u(x,y)&=&\frac{1}{2}erf\left(\frac{1-y}{2}\sqrt{\frac{u_0}{x}}\right)\nonumber \\
+T_l(x,y)&=&1-\frac{1}{2}erf\left(\frac{y}{2}\sqrt{\frac{u_0}{L_x-x}}\right)\nonumber \\
+T_r(x,y)&=&\frac{1}{2}+\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{y+1}} \exp\left(-\frac{x^2u_0}{4y+4}\right)\nonumber \\
+T_s(x,y)&=&\frac{1}{2}-\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{2-y}} \exp\left(-\frac{(L_x-x)^2u_0}{8-4y}\right)\nonumber
+\end{eqnarray}
+and
+\begin{eqnarray}
+u_0&=&\frac{L_x^{7/3}}{(1+L_x^4)^{2/3}}\left(\frac{Ra}{2\sqrt{\pi}}\right)^{\frac{2}{3}}\nonumber \\
+Q&=&2\sqrt{\frac{L_x}{\pi u_0}}\nonumber
+\end{eqnarray}
+
+The experiment is performed with two grid resolution (\num{125x40} and \num{200x80} elements), with 100 markers per element and CFL of 0.25. $v_{\textrm{rms}}$
+(see Sec. \ref{sec:error}) calculated for both the simulations match well with results from \citet{vanKeken1997} and obtained with ASPECT
+\citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} and ELEFANT \citep{Thieulot2014}, obtained with similar grid resolutions
+(Fig. \ref{fig:thin}). In addition, a variety of simulations with different resolutions and aspect ratios are performed to verify that the sensitivity of the
+initial velocity field. Results match well with results from \citet{vanKeken1997} (in the grey area) and \citet{Thieulot2014} (Fig. \ref{fig:thin_initial}).
+
+\begin{figure}
+\centering
+\noindent\includegraphics[width=400px]{./Figures/vrms_thin.pdf}
+\caption{$v_{\textrm{rms}}$ for the thin layer experiment as function of time for different grid resolution. Results are compared with results obtained by
+\citet{vanKeken1997} (black lines) and with ASPECT and ELEFANT.}
+\label{fig:thin}
+\end{figure}
+
+\begin{figure}
+\centering
+\noindent\includegraphics[width=400px]{./Figures/vrmszero.pdf}
+\caption{$v_{\textrm{rms}}$ at $t=0$ for the thin layer experiment as function of the element size for different aspect ratios of the elements. Results are compared
+with results obtained by \citet{vanKeken1997} (black dots in the grey area) and with ASPECT and ELEFANT.}
+\label{fig:thin_initial}
+\end{figure}
 
@@
 
-\subsection{Melting curves}\label{sec:katz}
+\subsection{Mantle melting curves}\label{sec:katz}
 This instantaneous experiment is performed in order to recreate isobaric melting curves, as function of temperature, bulk water and content of water in the
@@
 Fig. \ref{fig:fraction} shows isobaric melting curves for \SI{1}{\giga\pascal} (panel a) and \SI{3}{\giga\pascal} (panel b) calculated for different bulk water
-contents. The curves well match with those obtained by \citet{Katz2003}, showing the saturation at the solidus for bulk water of 0.2 and 0.3 wt.\%, indicating
+contents. The curves well match with those obtained for the mantle by \citet{Katz2003}, showing the saturation at the solidus for bulk water of 0.2 and 0.3 wt.\%, indicating
 that the code correctly determines melt fraction as function of pressure, temperature and bulk water content. Same results can be observed in Fig.

### 2021-05-26 Paper.tex
modified

@@
 \begin{table}[H]
-\caption{List of benchmarks implemented for each feature of the code.}
+\caption{List of benchmarks implemented for each feature of the code. Data of all benchmarks can be found at 
+\url{https://github.com/aleregorda/Benchmarks}}
 \centering
@@
 
-Other important quantities calculated in some benchmarks are the root-mean-square velocity and the Nusselt number, calculated as
-\[v_{\textrm{rms}}=\sqrt{\frac{1}{V_{\Omega}}\int_{\Omega}|\bm{v}|^2 d\Omega}\]
-and
-\[\textrm{Nu}=\frac{\int_{top}\bm{\nabla} T \cdot \bm{n} ds}{\int_{bottom}T ds}\]
-respectively.
+Other quantities used as comparison with original benchmarks are the root-mean-square velocity over the whole domain and over the surface
+\begin{eqnarray}
+v_{\textrm{rms}}&=&\sqrt{\int_{\Omega}|\bm{v}|^2 d\Omega}\nonumber \\
+v_{\textrm{rms}}^{top}&=&\sqrt{\int_0^1 u^2 \bigg|_{y=1} dx}\nonumber
+\end{eqnarray}
+respectively, top and bottom Nusselt numbers
+\[\textrm{Nu}^{top/bottom}=-\int_0^1\frac{\partial T}{\partial y} \bigg|_{y=1/y=0} dx\]
+the average rate of work done against gravity
+\[\left\langle W \right\rangle=\int_{\Omega}T u_y d\Omega\]
+the average rate of viscous dissipation
+\[\left\langle \Phi \right\rangle=\int_{\Omega}\tau_{ij}\dot{\epsilon}_{ij} d\Omega\]
+and the percentage error between $\left\langle W \right\rangle$ and $\left\langle \Phi \right\rangle/Ra$
+\[\delta=\frac{\left|\left\langle W \right\rangle -\frac{\left\langle \Phi \right\rangle}{Ra} \right|}{\max \left(\left\langle W \right\rangle,
+\frac{\left\langle \Phi \right\rangle}{Ra}\right)}\times100\]
 
@@
 red line), in comparison with the analytical solution (dashed blue line). Solve times and memory usage needed to generate the solution are shown in
-Fig. \ref{fig:MUMPS}.
+Fig. \ref{fig:MUMPS}. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Solver/Stokes_Flow}.
 
@@
 final position that increase moving away from the centre, more evident in case of higher time steps or lower grid resolution (Fig. \ref{fig:runge_err}). 
+All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Zalesak_Disk}.
 
@@
 c and Fig. \ref{fig:poi_plot}b). Fig. \ref{fig:poiseuille}d shows that one Uzawa iteration is sufficient to bring the divergence down to \num{e-15}, with no
-correlation with the pressure field.
+correlation with the pressure field. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Poiseuille%20Flow}.
 
@@
 entire domain ($p_{avg}$) are compared with solutions generated by ASPECT \citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} and all results
-can be found at \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere2D}. Results in terms of the $v_{\textrm{rms}}$ for NS model are also shown
-in Fig. \ref{fig:inst_sphere}.
+can be found at \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere2D} and 
+\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Instantaneous_2D_Sphere}. Results in terms of the $v_{\textrm{rms}}$ for NS model 
+are also shown in Fig. \ref{fig:inst_sphere}.
 
@@
 $v_{\textrm{rms}}$ as function of time is reported in Fig. \ref{fig:RT}, matching well with results shown by \citet{vanKeken1997}, \citet{Tackley2003} and
-\citet{Thieulot2014}. Fig. \ref{fig:rayleigh} shows the evolution of the experiment at different time steps.
+\citet{Thieulot2014}. Fig. \ref{fig:rayleigh} shows the evolution of the experiment at different time steps. All data can be found at 
+\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Rayleigh_Taylor_experiment/ISOVISCOUS}.
 
@@
 quantity $\bm{v}/(\rho_b-\rho_f)$ is plotted as function of the viscosity contrast. All results perfectly match with those from \citet{Gerya2010b} and they
-line up on a single curve, demonstrating that the code can correctly deal with large viscosity and density contrasts (Fig. \ref{fig:falling}).
+line up on a single curve, demonstrating that the code can correctly deal with large viscosity and density contrasts (Fig. \ref{fig:falling}). All data can be 
+found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Falling%20blocks}.
 
@@
 \citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} and they can be found at
-\url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere_fs2D}. Results in terms of the $v_{\textrm{rms}}$ in case of an arithmetic average are also
-shown in Fig. \ref{fig:inst_spherefs}.
+\url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere_fs2D} and
+\url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Time_dependent_sphere}. Results in terms of the $v_{\textrm{rms}}$ in case of 
+an arithmetic average are also shown in Fig. \ref{fig:inst_spherefs}.
 
@@
 ($dt=$ \SI{4200}{\year}, green line in Fig. \ref{fig:stab}), as already observed by \citet{Kaus2010a} and \citet{Thieulot2014}. Activating the algorithm the
-instability is fixed and the simulation remains stable using time step up to \SI{50000}{\year} (red line in Fig. \ref{fig:stab}).
+instability is fixed and the simulation remains stable using time step up to \SI{50000}{\year} (red line in Fig. \ref{fig:stab}). All data can be 
+found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Stabilisation_algorithm}.
 
@@
 \SI{2570}{\m}, with an error of \SI{6}{\m} with respect to the analytical solution. Results are also compared with results obtained with a free surface in
-UNDERWORLD (\num{256x64} elements) and SULEC (\num{401x201} elements) (with courtesy from \citealp{Crameri2012}).
+UNDERWORLD (\num{256x64} elements) and SULEC (\num{401x201} elements) (with courtesy from \citealp{Crameri2012}). All data can be 
+found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Topography_relaxation}.
 
@@
 Fig. \ref{fig:subduction}). A high resistance at the trench can be also observed in case of a true free surface for the low resolution grid used (continuous
-lines in Fig. \ref{fig:subduction}), as pointed out by \citet{Schmeling2008}.
+lines in Fig. \ref{fig:subduction}), as pointed out by \citet{Schmeling2008}. All data can be 
+found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Spontaneous_subduction}.
 
@@
 harmonic mean of the viscosity are shown in Fig. \ref{fig:necking}, compared with results from \citet{Glerum2018} for which an infinity norm mean was used.
-The influence of the average schemes agrees with results shown in Section \ref{sec:subduction}.
+The influence of the average schemes agrees with results shown in Section \ref{sec:subduction}. All data can be 
+found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Slab_detachment}.
 
@@
 smooth punch (green and blue lines in Fig. \ref{fig:smooth_p}, respectively), as previously observed by \citet{Thieulot2014} and \citet{Glerum2018}. 
+All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Indenter}.
 
@@
 decrease of the velocity residuals, which can not be observed for for tests with higher internal friction angles (Fig. \ref{fig:convergence}). 
+All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Brick_experiment}.
 
@@
 eliminated in case of the SUPG method ($\gamma=0.045$, orange line in \ref{fig:advection}). 
+All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Advection_stabilisation}.
 
@@
 
-\subsection{Shear and adiabatic heating}\label{sec:simple_shear}
+\subsection{Shear heating}\label{sec:simple_shear}
 This exercise is performed in an unit square domain composed by \num{128x128} elements. The velocity field is prescribed on the entire domain with
@@
 The solution predicted by the model in terms of velocity, temperature and shear heating match well with the analytical solutions (Fig. \ref{fig:analytical_en}). 
+All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Simple_shear}.
 
@@
 obtained by means of example ${Shear\_adiabatic\_heating.m}$ from \citet{Gerya2010b} (second row in Fig. \ref{fig:energy}). 
+All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Adiabatic%2BShear}.
 
@@
 simulations in Fig. \ref{fig:mantle}. At the steady state, $v_{\textrm{rms}}$ and Nu of all simulations converge well toward the values from \citet{Blankenbach1989},
-with lower errors for higher resolution grids (Table \ref{tab:mantle}).
+with lower errors for higher resolution grids (Table \ref{tab:mantle}). All data can be found at 
+\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Mantle_convection}.
 
@@
 For each case, some diagnostic quantities measured at the steady state are reported in Table \ref{tab:visco_mantle} (see Section \ref{sec:error} for 
-the computation of all the quantities), showing a well fit with results described in \citet{Tosi2015}.
+the computation of all the quantities), showing a well fit with results described in \citet{Tosi2015}. Data and charts for all cases can be found at 
+\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Viscoplastic_mantle_convection}.
 
@@
 %{\rowcolors{3}{gray!25!}{white!100!}
-\begin{table}[H]
+\begin{table}
   \caption{Results of the viscoplastic mantle convection benchmark for different grid resolutions in Cases 1-5 compared with results obtained with ELEFANT
@@
 initial velocity field. Results match well with results from \citet{vanKeken1997} (in the grey area) and \citet{Thieulot2014} (Fig. \ref{fig:thin_initial}). 
+All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Thin_layer}.
 
@@
 bound and free water during the evolution of the experiment is shown in Fig. \ref{fig:quinquis} (left and right columns, respectively). 
+All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Hydration}.
 
@@
 code correctly determines also the percentage of water in the melt as function of pressure, temperature and melt fraction. 
+All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Melting}.
 

### 2021-10-18 Paper.tex
modified

@@
 Differently, the average scheme for the elemental viscosity can be chosen between harmonic, geometric and arithmetic mean. Markers can be placed regularly or
-randomly at the beginning of the simulation and their advection is performed by a 4th-order Runge-Kutta in space, interpolating the velocity field on each
-marker by means of the shape functions (see Section \ref{sec:runge}). The initial distribution of the markers is created using the open-source code library
+randomly at the beginning of the simulation and their advection is performed by either a 2nd-order or a 4th-order Runge-Kutta in space, interpolating the velocity 
+field on each marker by means of the shape functions (see Section \ref{sec:runge}). The interpolated velocity is then corrected by means of the Conservative
+Velocity Interpolation (CVI), which introduces a corrective term to reduce dispersion and clustering of particles in both steady state and time-dependent 
+\citep{Wang2015}.
+
+The initial distribution of the markers is created using the open-source code library
 Geodynamic World Builder \citep{Fraters2019}. During the simulation, each marker carries memory of temperature, pressure and accumulated strain, which are
@@
 Non-linear rheologies are implemented combining viscous creep (dislocation and diffusion) and plastic yielding. For each marker, diffusion and dislocation
-viscosity ($\eta_{df|ds}$) can be determined as follows \citep{Karato1993,Wang2016,Glerum2018}:
+viscosity ($\eta_{df|ds}$) can be determined as follows \citep{Karato1993,Wang2016}:
 \begin{equation}\label{eq:viscous}
@@
 &&+(-2+24y-72y^2+48y^3)x+1-4y+12y^2-8y^3\nonumber \\
-b_2&=&(8-48y+48y^2)x^3+(-12y+72y-72y^2)x^2+(4-24y+48y^2-48y^3+\nonumber \\
+b_2&=&(8-48y+48y^2)x^3+(-12+72y-72y^2)x^2+(4-24y+48y^2-48y^3+\nonumber \\
 &&+24y^4)x-12y^2+24y^3-12y^4\nonumber

### 2022-01-28 FALCON_description.tex
modified

--- 2021-10-18 Paper.tex
@@
 \usepackage{amsfonts}
+\usepackage{textcomp}
 \usepackage{amssymb}
@@
 \section{Introduction}\label{sec:intro}
-This supplement includes a description of the code, with a presentation of the features implemented followed by results of extensive benchmarks listed in
-Table \ref{tab:list}, implemented to verify that the code solves correctly a large range of problems. Description and results of all benchmarks are shown
+This document includes a description of the code FALCON, with a presentation of the features implemented followed by results of extensive benchmarks listed in
+Table \ref{tab:list}, performed to verify that the code solves correctly a wide range of problems. Description and results of all benchmarks are shown
 in Section \ref{sec:benchmark}.
@@
 \begin{table}[H]
-\caption{List of benchmarks implemented for each feature of the code. Data of all benchmarks can be found at 
+\caption{List of benchmarks performed for each feature of the code. Data of all benchmarks can be found at 
 \url{https://github.com/aleregorda/Benchmarks}}
@@
 \hline
-  4th-order Runge-Kutta  & \multirow{2}{*}{Zalesak disk} & \multirow{2}{*}{\ref{sec:runge}} \\
-  markers advection  &  & \\
+  \multirow{2}{*}{Markers advection} & Zalesak disk & \ref{sec:runge} \\
+    & Conservative Interpolation Velocity & \ref{sec:CVI} \\
 \hline
@@
 
-The time step $dt$ is chosen in according to Courant-Friedrichs-Lewy condition \citep{Anderson}, such that
+The time step $dt$ is chosen in according to Courant-Friedrichs-Lewy condition \citep{Anderson1995}, such that
 \[dt=C_n\; \min\left(\frac{\Delta x}{\max|\bm{u}|},\frac{(\Delta x)^2}{\kappa}\right)\]
@@
 Velocity Interpolation (CVI), which introduces a corrective term to reduce dispersion and clustering of particles in both steady state and time-dependent 
-\citep{Wang2015}.
+\citep{Wang2015} (see Section \ref{sec:CVI}).
 
@@
 markers. The correctness of the evolution of the markers chain is tested with the experiment of a 2D time-dependent Stokes sphere below a free surface and
-compared with results from ASPECT\footnote{\url{https://aspect.geodynamics.org/}} \citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} (Section
+compared with results from ASPECT\footnote{\url{https://aspect.geodynamics.org/}} \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} (Section
 \ref{sec:sphere}). In the true free surface case the top boundary is assumed stress-free and velocities are not fixed. In this case topography variations are
@@
 \section{Erosion and sedimentation}\label{sec:erosion}
-Surface processes at the lithosphere-air interface are implemented taking into account fluvial erosion (long-range processes), hillslope processes (short-range
-processes) and failure of slopes exceeding a threshold angle $\theta$ \citep{Burov1997,Tucker1999,Sacek2012,Croissant2014,Ueda2015,Beucher2020}. Erosion rates
-$\dot{E}_{lr}$ due to long-range processes can be described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, in according to
+Surface processes at either the lithosphere-air interface or in correspondence of the free surface have been implemented by means of the software
+Fastscape\footnote{\url{https://fastscape.org/}} \citep{Braun2013,Cordonnier2019,Yuan2019a,Yuan2019b}. FastScapeLib is a set of routines that solve the stream
+power law (enriched by a sediment transport/deposition term) \citep{Yuan2019a}, hillslope diffusion and marine transport and deposition \citep{Yuan2019b}.
+The partial differential equation solved by FastScapeLib is:
 \begin{equation}
-\dot{E}_{lr}=K_f A^{m_e} \bm{\nabla} h^{n_e}
-\end{equation}
-where $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $\bm{\nabla} h$ is the slope and $m_e$ and $n_e$ are parameters chosen in to
-constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}. Erosion rates $\dot{E}_{sr}$ due to short-range processes are
-implemented by assuming that transport is linearly proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}, in according to
-\begin{equation}
-\dot{E}_{sr}=K_d \bm{\nabla}^2 h
-\end{equation}
-where $K_d$ is the effective diffusivity. In addition, the slope is modified if it still exceeds a threshold angle after long- and short-range processes
-\citep{Ueda2015}.
-
-Offshore sedimentation processes are implemented similarly to hillslope processes but with a depth-dependent coefficient \citep{Sacek2012,Beucher2020}.
-The sedimentation rate $\dot{S}$ is calculated as
-\begin{equation}
-\dot{S}=K_s \bm{\nabla}^2 h
-\end{equation}
-with
-\begin{equation}
-K_s=K^*_s \exp\left(\frac{h-h_{sl}}{h_0}\right)
-\end{equation}
-where $K^*_s$ is a constant, $h_{sl}$ is the sea level and $h_0$ is a characteristic length. In case of sticky air, slope is determined by means of the markers
-chain, which is vertically corrected considering erosion and sedimentation rates. After the correction, continental markers above the markers chain are
-transformed in air markers and air markers below the markers chain become sediments. Similarly, slope is determined considering the nodes on the top boundary
-in case of true free surface and they are vertically corrected as for the markers chain. In this case, continental markers above the top boundary are simply
-deleted.
+\frac{\partial h}{ \partial t}=U-K_f A^{m_e}S^{n_e} + \frac{G}{A}\int_A \left( U-\frac{\partial h}{\partial t} \right) dA + K_d \bm{\nabla}^2 h 
+\end{equation}
+where $h$ is the topography, $U$ is the uplift, $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $S$ is the slope, $m_e$ and $n_e$ are
+parameters that should be chosen to constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}, $G$ is the dimensionless
+deposition/transport coefficient for the enriched stream power law and $K_d$ is the effective diffusivity. Terms $K_f A^{m_e}S^{n_e}$ and $K_d \bm{\nabla}^2 h$
+consider long-range processes, described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, and short-range processes,
+proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}.
+
+In case of sticky air, slope is determined by means of the markers chain, which is vertically corrected considering erosion and sedimentation rates. After the
+correction, continental markers above the markers chain are transformed in air markers and air markers below the markers chain become sediments. Similarly,
+slope is determined considering the nodes on the top boundary in case of true free surface and they are vertically corrected as for the markers chain. In this
+case, continental markers above the top boundary are simply deleted.
+%Surface processes at the lithosphere-air interface are implemented taking into account fluvial erosion (long-range processes), hillslope processes (short-range
+%processes) and failure of slopes exceeding a threshold angle $\theta$ \citep{Burov1997,Tucker1999,Sacek2012,Croissant2014,Ueda2015,Beucher2020}. Erosion rates
+%$\dot{E}_{lr}$ due to long-range processes can be described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, in according to
+%\begin{equation}
+%\dot{E}_{lr}=K_f A^{m_e} \bm{\nabla} h^{n_e}
+%\end{equation}
+%where $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $\bm{\nabla} h$ is the slope and $m_e$ and $n_e$ are parameters chosen in to
+%constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}. Erosion rates $\dot{E}_{sr}$ due to short-range processes are
+%implemented by assuming that transport is linearly proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}, in according to
+%\begin{equation}
+%\dot{E}_{sr}=K_d \bm{\nabla}^2 h
+%\end{equation}
+%where $K_d$ is the effective diffusivity. In addition, the slope is modified if it still exceeds a threshold angle after long- and short-range processes
+%\citep{Ueda2015}.
+
+%Offshore sedimentation processes are implemented similarly to hillslope processes but with a depth-dependent coefficient \citep{Sacek2012,Beucher2020}.
+%The sedimentation rate $\dot{S}$ is calculated as
+%\begin{equation}
+%\dot{S}=K_s \bm{\nabla}^2 h
+%\end{equation}
+%with
+%\begin{equation}
+%K_s=K^*_s \exp\left(\frac{h-h_{sl}}{h_0}\right)
+%\end{equation}
+%where $K^*_s$ is a constant, $h_{sl}$ is the sea level and $h_0$ is a characteristic length. In case of sticky air, slope is determined by means of the markers
+%chain, which is vertically corrected considering erosion and sedimentation rates. After the correction, continental markers above the markers chain are
+%transformed in air markers and air markers below the markers chain become sediments. Similarly, slope is determined considering the nodes on the top boundary
+%in case of true free surface and they are vertically corrected as for the markers chain. In this case, continental markers above the top boundary are simply
+%deleted.
 
@@
 Non-linear rheologies are implemented combining viscous creep (dislocation and diffusion) and plastic yielding. For each marker, diffusion and dislocation
-viscosity ($\eta_{df|ds}$) can be determined as follows \citep{Karato1993,Wang2016}:
+viscosity ($\eta_{df|ds}$) can be determined as follows \citep{Karato1993,Warren2008b,Wang2016}:
 \begin{equation}\label{eq:viscous}
-\eta_{df|ds}=\left(\frac{d^m}{A}\right)^{\frac{1}{n}}I_2^{\frac{(1-n)}{n}}\exp\left(\frac{Q+pV}{nRT}\right)
-\end{equation}
-where $d$ is the grain size, $m$ is the grain size exponent, $n$ is the stress exponent, $A$ is the pre-exponential factor, $I_2$ is the square root of the
-second invariant of the strain rate tensor, $Q$ is the activation energy, $p$ is the pressure, $V$ is the activation volume, $R$ is the gas constant and $T$
-is the temperature. Pressures and temperatures are determined by the interpolation of the nodal parameters, while strain rates are calculated by means of the
-derivative of the velocities on the nodes. In case of diffusion creep $n=1$ and $m>0$, in case of dislocation creep $n>1$ and $m=0$. The implementation of the
-non-linear viscous creep viscosity depending on the strain rate is tested by means of the slab detachment benchmark (Section \ref{sec:slab}).
-
-Plastic yielding is implemented rescaling $\eta_{vc}$ in order to limit the stress below the yield stress $\sigma_y$ \citep{Thieulot2008,Thieulot2014,Glerum2018},
+\eta_{df|ds}=f_s \left(\frac{d^m}{A}\right)^{\frac{1}{n}} \left (\frac{2^{\frac{1-n}{n}}}{3^{\frac{n+1}{2n}}} \right)
+I_2^{\frac{(1-n)}{n}}\exp\left(\frac{Q+pV}{nRT}\right)
+\end{equation}
+where $f_s$ is a scaling factor used to represent lithologies that are stronger or weaker than the base set, $d$ is the grain size, $m$ is the grain size exponent,
+$n$ is the stress exponent, $A$ is the uniaxial pre-exponential factor, $I_2$ is the square root of the second invariant of the strain rate tensor, $Q$ is the
+activation energy, $p$ is the pressure, $V$ is the activation volume, $R$ is the gas constant and $T$ is the temperature.
+Pressures and temperatures are determined by the interpolation of the nodal parameters, while strain rates are calculated by means of the
+derivative of the velocities on the nodes. In case of diffusion creep $n=1$ and $m>0$, in case of dislocation creep $n>1$ and $m=0$. The viscous creep $\eta_{cp}$
+is then calculated as the harmonic average between $\eta_{df}$ and $\eta_{ds}$:
+\begin{equation}\label{eq:visco_creep}
+\eta_{cp}=\left(\frac{1}{\eta_{df}}+\frac{1}{\eta_{ds}}\right)^{-1}
+\end{equation}
+The implementation of the non-linear viscous creep viscosity depending on the strain rate is tested by means of the slab detachment benchmark
+(Section \ref{sec:slab}).
+
+Plastic yielding is implemented rescaling $\eta_{cp}$ in order to limit the stress below the yield stress $\sigma_y$ \citep{Thieulot2008,Thieulot2014,Glerum2018},
 obtaining
@@
 means of the accumulated strain $\epsilon$ memorised by each marker. Softening in the viscous creep determines a linear decrease of $\eta_{vc}$ by means of a
-viscous strain softening factor $W_S$ that increases linearly from 1 to 10 for $5<\epsilon<10$ \citep{Warren2008a}. This viscous softening can be related to
-strain-induced grain size reduction \citep{Warren2008a}. Differently, plastic softening is simulated with a linear decrease of internal friction angle
-$\phi(\epsilon)$ and cohesion $C(\epsilon)$ in according to
+viscous strain softening factor $W_S$ that increases linearly from $W_{S_0}$ to $W_{S_{\infty}}$ for $\epsilon_{S_0}<\epsilon_S<\epsilon_{S_{\infty}}$ 
+\citep{Warren2008a}. This viscous softening can be related to strain-induced grain size reduction \citep{Warren2008a}. Differently, plastic softening is 
+simulated with a linear decrease of internal friction angle $\phi(\epsilon)$ and cohesion $C(\epsilon)$ in according to
 \begin{eqnarray}
@@
 \end{eqnarray}
-where $\phi_0$, $C_0$ and $\phi_{\infty}$, $C_{\infty}$ are internal friction angle and cohesion for $\epsilon=0.5$ and $\epsilon=1.5$, respectively
+where $\phi_0$, $C_0$ and $\phi_{\infty}$, $C_{\infty}$ are internal friction angle and cohesion for $\epsilon_0$ and $\epsilon_{\infty}$, respectively
 \citep{Huismans2003,Huismans2005,Warren2008a,Thieulot2014}. Plastic softening approximates deformation-induced softening of faults and brittle shear zones
@@
 
-Viscous creep and plastic yielding are then combined to obtain a viscoplastic viscosity $\eta_{vp}$, which is computed with an harmonic average
-\citep{Tackley2010,Glerum2018} as follows
+Plastic yielding and viscous creep are then combined to obtain a viscoplastic viscosity $\eta_{vp}$ as follows, assuming that they are independent processes
+\citep{Karato2008,Glerum2018}:
 \begin{equation}\label{eq:viscoplastic}
-\eta_{vp}=\left(\frac{1}{\eta_{df}}+\frac{1}{\eta_{ds}}+\frac{1}{\eta_{pl}}\right)^{-1}
+\eta_{vp}=\min(\eta_{cp},\eta_{pl})
 \end{equation}
@@
 where $||\bm{\mathcal{R}}^0||_2$ is the $L_2$-norm of the first nonlinear iteration.
-Since often the normalised nonlinear residual drop very quickly over the first 2-3 iterations, also the $L_2$-norm of the velocity residual is calculated as
+Since often the normalised nonlinear residual drop very quickly over the first 2-3 iterations, also the $L_2$-norm of velocity and pressure residuals are
+calculated as
 \begin{equation}\label{eq:convergence_vel}
-\frac{||\bm{v}_1-\bm{v}_{1-1}||_2}{||\bm{v}_{i-1}||_2}
-\end{equation}
-Iterations are performed until either maximal number of non-linear iterations ($it_{max}$) is reached or both the normalised nonlinear residual and the
-velocity residual converge under a defined tolerance $tol$.
+\frac{||\Delta\bm{v}||_2}{||\bm{v}||_2}=\frac{||\bm{v}_i-\bm{v}_{i-1}||_2}{||\bm{v}_{i}||_2}
+\end{equation}
+and
+\begin{equation}\label{eq:convergence_press}
+  \frac{||\Delta p||_2}{||p||_2}=\frac{||p_i-p_{i-1}||_2}{||p_{i}||_2}
+\end{equation}
+respectively. Iterations are performed until either maximal number of non-linear iterations ($it_{max}$) is reached or all the normalised nonlinear residual, 
+the velocity residual and the pressure residual converge under a defined tolerance $tol$.
 
@@
 \section{Mantle melting}\label{sec:melt}
-Melt fraction in the mantle depends on temperature ($T$, in °C), pressure ($p$, in \si{\giga\pascal}) and the water content in the melt ($X_{H_2O}$, in wt.\%).
+Melt fraction in the mantle depends on temperature ($T$, in \textdegree C), pressure ($p$, in \si{\giga\pascal}) and the water content in the melt ($X_{H_2O}$, in wt.\%).
 The determination of the melt fraction $M$ is implemented as explained by \citet{Katz2003}, according to the successive modification by \citet{Langmuir2006} and
@@
 \subsection{Markers advection}\label{sec:runge}
-The 4th-order Runge-Kutta advection scheme in space is tested by means of the Zalesak disk test \citep{Zalesak1979,Thieulot2014}. The benchmark is performed
-in a unit square domain with different grid resolution (\num{100x100} and \num{200x200} elements) and 1000000 markers. The Courant number varies between 0.1
-and 0.5 and the velocity field is prescribed in the entire domain as
+The 2nd-order Runge-Kutta advection scheme in space is tested by means of the Zalesak disk test \citep{Zalesak1979,Thieulot2014}. The benchmark is performed
+in a unit square domain with a grid resolution of \num{32x32} elements and 1000000 markers and values of Courant number between 0.25 and 1.
+At $t=0$, the disk is centred at position $(0.5;0.75)$ with a radius $R=0.15$ and has a vertical fissure 0.05 wide and 0.2 high (orange in Fig. \ref{fig:runge})
+and with the velocity field prescribed in the entire domain as
 \begin{eqnarray}
@@
 \end{eqnarray}
-At $t=0$, the disk is centred at position $(0.5;0.75)$ with a radius $R=0.15$ and has a vertical fissure 0.05 wide and 0.2 high (orange in Fig. \ref{fig:runge}).
-As we expect, markers are back to their initial location after a $2\pi$ rotation ($t=1$, red, green and blue in Fig. \ref{fig:runge}), with an error of the
-final position that increase moving away from the centre, more evident in case of higher time steps or lower grid resolution (Fig. \ref{fig:runge_err}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Zalesak_Disk}.
+the markers are back to their initial location after a $2\pi$ rotation ($t=1$). The clockwise rotation of the disk at $t=0.25$, 0.5 and 0.75 
+is shown in Fig. \ref{fig:runge} (red, blue and green, respectively). The distance from the centre of 1 marker is calculated throughout the rotation to
+evaluate the error for different values of Courant number. As expected, the error increase with the increasing of the time step, as shown in 
+Fig. \ref{fig:runge_err}. 
+All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/Zalesak_Disk}.
 
@@
 \noindent\includegraphics[width=200px]{./Figures/Zalesak.pdf}
-\caption{Comparison between the initial position of markers of the Zalesak disk experiment (orange) and their final position for different CFL in case of grid
-resolution of \num{100x100} elements. Blue, green and red markers indicate the final position of the disk in case of Courant number of 0.1, 0.25 and 0.5,
-respectively.}
+\caption{Position of the Zalesak disk at $t=0$, 0.25, 0.5 and 0.75 (orange, red, blue and green, respectively) for the case with Courant number of 1.}
 \label{fig:runge}
@@
 \centering
-\noindent\includegraphics[width=400px]{./Figures/zal_error.pdf}
-\caption{Error between final and initial position of markers of the disk as function of their distance from the centre of the domain. Results for different
-time steps are shown in blue, red and green for a grid resolution of \num{100x100} elements (for CFL of 0.1, 0.25 and 0.5, respectively) and in light blue,
-dark green and purple in case of \num{200x200} elements (for CFL of 0.1, 0.25 and 0.5, respectively).}
+\noindent\includegraphics[width=400px]{./Figures/zal_marker.pdf}
+\caption{Distance from the centre as function of time for values of Courant number of  of 0.25, 0.3, 0.5, 0.75 and 1 (green, orange, blue, black and red,
+respectively). R indicates the distance from the centre considering a perfect circle trajectory.}
 \label{fig:runge_err}
+\end{figure}
+
+\subsection{Conservative Velocity Interpolation (CVI)}\label{sec:CVI}
+The CVI correction is checked by means of the Stokes flow experiment by \citet{Donea2003}, as presented in \ref{sec:stokes}. The advection of Lagrangian
+markers is performed using either a 2nd-order or a 4th-order Runge-Kutta scheme with CFL=0.5 and an initial random distribution of 25 markers per element.
+Fig. \ref{fig:CVI} shows the minimum and maximum values of markers per element throughout the experiment with and without the CVI correction and using
+different Runge-Kutta schemes. The introduction of the CVI correction allows the
+tests to run for times longer than $t=2000$ without show empty elements, while tests without the CVI correction stop at approximately $t=600$, when at least
+one element goes empty. Fig. \ref{fig:CVI_time} shows the different number of markers per element and the markers distribution at $t=600$ obtained without and
+with the CVI correction (panels b and c, respectively) in case of a 2nd-order Runge-Kutta scheme. All data can be found at
+\url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/CVI}.
+\begin{figure}
+\centering
+\noindent\includegraphics[width=\textwidth]{./Figures/CVI.pdf}
+\caption{Maximum and minimum number of markers per element throughout the experiment, with (green and orange lines) or without (purple and blue lines)
+the CVI correction and using either a 2nd-order (blue and orange lines) or a 4th-order (purple and green lines) Runge-Kutta scheme.}
+\label{fig:CVI}
+\end{figure}
+\begin{figure}
+\centering
+\noindent\includegraphics[width=\textwidth]{./Figures/CVI_time.pdf}
+\caption{Number of markers per element and markers distribution through time using 2nd.order Runge-Kutta scheme.
+Panels a: number of markers per element (panel a1) and distribution of the markers (panel a2) after the first
+iteration; panels b and c: comparison between number of markers per element (panels b1 and c1) and markers distribution (panels b2 and c2) without and with
+the CVI correction at $t=600$; panels d: number of markers per element (panel d1) and markers distribution (panel d2) in case of CVI correction at $t=2000$.}
+\label{fig:CVI_time}
 \end{figure}
@@
 maximum velocities ($u_{min}$, $u_{max}$, $v_{min}$, $v_{max}$) and pressures ($p_{min}$, $p_{max}$) on the entire domain, $v_{\textrm{rms}}$ and average pressure on the
-entire domain ($p_{avg}$) are compared with solutions generated by ASPECT \citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} and all results
+entire domain ($p_{avg}$) are compared with solutions generated by ASPECT \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and all results
 can be found at \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere2D} and 
@@
 in terms of velocity and pressure fields and topography variations are compared with results obtained with ASPECT
-\citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} and they can be found at
+\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and they can be found at
 \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere_fs2D} and
@@
 \[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\].
-The viscosity ﬁeld $\eta$ is calculated as the harmonic average between a linear part $\eta_{lin}$ that depends on temperature only or on temperature and 
+The viscosity field $\eta$ is calculated as the harmonic average between a linear part $\eta_{lin}$ that depends on temperature only or on temperature and 
 depth $y$ and a nonlinear-plastic part $\eta_{pl}$ that depends on the strain rate $\dot{\bm{\epsilon}}$, as follows
@@
 (see Sec. \ref{sec:error}) calculated for both the simulations match well with results from \citet{vanKeken1997} and obtained with ASPECT
-\citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} and ELEFANT \citep{Thieulot2014}, obtained with similar grid resolutions
+\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and ELEFANT \citep{Thieulot2014}, obtained with similar grid resolutions
 (Fig. \ref{fig:thin}). In addition, a variety of simulations with different resolutions and aspect ratios are performed to verify that the sensitivity of the

### 2022-01-28 Paper.tex
modified

@@
 \begin{equation}\label{eq:viscous}
-\eta_{df|ds}=\left(\frac{d^m}{A}\right)^{\frac{1}{n}}I_2^{\frac{(1-n)}{n}}\exp\left(\frac{Q+pV}{nRT}\right)
-\end{equation}
-where $d$ is the grain size, $m$ is the grain size exponent, $n$ is the stress exponent, $A$ is the pre-exponential factor, $I_2$ is the square root of the
-second invariant of the strain rate tensor, $Q$ is the activation energy, $p$ is the pressure, $V$ is the activation volume, $R$ is the gas constant and $T$
-is the temperature. Pressures and temperatures are determined by the interpolation of the nodal parameters, while strain rates are calculated by means of the
-derivative of the velocities on the nodes. In case of diffusion creep $n=1$ and $m>0$, in case of dislocation creep $n>1$ and $m=0$. The implementation of the
-non-linear viscous creep viscosity depending on the strain rate is tested by means of the slab detachment benchmark (Section \ref{sec:slab}).
-
-Plastic yielding is implemented rescaling $\eta_{vc}$ in order to limit the stress below the yield stress $\sigma_y$ \citep{Thieulot2008,Thieulot2014,Glerum2018},
+\eta_{df|ds}=f_s \left(\frac{d^m}{A}\right)^{\frac{1}{n}} \left (\frac{2^{\frac{1-n}{n}}}{3^{\frac{n+1}{2n}}} \right)
+I_2^{\frac{(1-n)}{n}}\exp\left(\frac{Q+pV}{nRT}\right)
+\end{equation}
+where $f_s$ is a scaling factor used to represent lithologies that are stronger or weaker than the base set, $d$ is the grain size, $m$ is the grain size exponent,
+$n$ is the stress exponent, $A$ is the uniaxial pre-exponential factor, $I_2$ is the square root of the second invariant of the strain rate tensor, $Q$ is the
+activation energy, $p$ is the pressure, $V$ is the activation volume, $R$ is the gas constant and $T$ is the temperature.
+Pressures and temperatures are determined by the interpolation of the nodal parameters, while strain rates are calculated by means of the
+derivative of the velocities on the nodes. In case of diffusion creep $n=1$ and $m>0$, in case of dislocation creep $n>1$ and $m=0$. The viscous creep $\eta_{cp}$
+is then calculated as the harmonic average between $\eta_{df}$ and $\eta_{ds}$:
+\begin{equation}\label{eq:visco_creep}
+\eta_{cp}=\left(\frac{1}{\eta_{df}}+\frac{1}{\eta_{ds}}\right)^{-1}
+\end{equation}
+The implementation of the non-linear viscous creep viscosity depending on the strain rate is tested by means of the slab detachment benchmark
+(Section \ref{sec:slab}).
+
+Plastic yielding is implemented rescaling $\eta_{cp}$ in order to limit the stress below the yield stress $\sigma_y$ \citep{Thieulot2008,Thieulot2014,Glerum2018},
 obtaining

### 2022-01-28 .debris/2022-01-27 11.00.19.00/Paper.tex
modified

--- 2022-01-28 Paper.tex
@@
 \usepackage{amsfonts}
-\usepackage{textcomp}
 \usepackage{amssymb}
@@
 \section{Introduction}\label{sec:intro}
-This document includes a description of the code FALCON, with a presentation of the features implemented followed by results of extensive benchmarks listed in
-Table \ref{tab:list}, performed to verify that the code solves correctly a wide range of problems. Description and results of all benchmarks are shown
+This supplement includes a description of the code, with a presentation of the features implemented followed by results of extensive benchmarks listed in
+Table \ref{tab:list}, implemented to verify that the code solves correctly a large range of problems. Description and results of all benchmarks are shown
 in Section \ref{sec:benchmark}.
@@
 \begin{table}[H]
-\caption{List of benchmarks performed for each feature of the code. Data of all benchmarks can be found at 
+\caption{List of benchmarks implemented for each feature of the code. Data of all benchmarks can be found at 
 \url{https://github.com/aleregorda/Benchmarks}}
@@
 
-The time step $dt$ is chosen in according to Courant-Friedrichs-Lewy condition \citep{Anderson1995}, such that
+The time step $dt$ is chosen in according to Courant-Friedrichs-Lewy condition \citep{Anderson}, such that
 \[dt=C_n\; \min\left(\frac{\Delta x}{\max|\bm{u}|},\frac{(\Delta x)^2}{\kappa}\right)\]
@@
 markers. The correctness of the evolution of the markers chain is tested with the experiment of a 2D time-dependent Stokes sphere below a free surface and
-compared with results from ASPECT\footnote{\url{https://aspect.geodynamics.org/}} \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} (Section
+compared with results from ASPECT\footnote{\url{https://aspect.geodynamics.org/}} \citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} (Section
 \ref{sec:sphere}). In the true free surface case the top boundary is assumed stress-free and velocities are not fixed. In this case topography variations are
@@
 \section{Erosion and sedimentation}\label{sec:erosion}
-Surface processes at either the lithosphere-air interface or in correspondence of the free surface have been implemented by means of the software
-Fastscape\footnote{\url{https://fastscape.org/}} \citep{Braun2013,Cordonnier2019,Yuan2019a,Yuan2019b}. FastScapeLib is a set of routines that solve the stream
-power law (enriched by a sediment transport/deposition term) \citep{Yuan2019a}, hillslope diffusion and marine transport and deposition \citep{Yuan2019b}.
-The partial differential equation solved by FastScapeLib is:
-\begin{equation}
-\frac{\partial h}{ \partial t}=U-K_f A^{m_e}S^{n_e} + \frac{G}{A}\int_A \left( U-\frac{\partial h}{\partial t} \right) dA + K_d \bm{\nabla}^2 h 
-\end{equation}
-where $h$ is the topography, $U$ is the uplift, $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $S$ is the slope, $m_e$ and $n_e$ are
-parameters that should be chosen to constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}, $G$ is the dimensionless
-deposition/transport coefficient for the enriched stream power law and $K_d$ is the effective diffusivity. Terms $K_f A^{m_e}S^{n_e}$ and $K_d \bm{\nabla}^2 h$
-consider long-range processes, described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, and short-range processes,
-proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}.
-
-In case of sticky air, slope is determined by means of the markers chain, which is vertically corrected considering erosion and sedimentation rates. After the
-correction, continental markers above the markers chain are transformed in air markers and air markers below the markers chain become sediments. Similarly,
-slope is determined considering the nodes on the top boundary in case of true free surface and they are vertically corrected as for the markers chain. In this
-case, continental markers above the top boundary are simply deleted.
 %Surface processes at the lithosphere-air interface are implemented taking into account fluvial erosion (long-range processes), hillslope processes (short-range
@@
 \begin{equation}\label{eq:viscous}
-\eta_{df|ds}=f_s \left(\frac{d^m}{A}\right)^{\frac{1}{n}} \left (\frac{2^{\frac{1-n}{n}}}{3^{\frac{n+1}{2n}}} \right)
-I_2^{\frac{(1-n)}{n}}\exp\left(\frac{Q+pV}{nRT}\right)
-\end{equation}
-where $f_s$ is a scaling factor used to represent lithologies that are stronger or weaker than the base set, $d$ is the grain size, $m$ is the grain size exponent,
-$n$ is the stress exponent, $A$ is the uniaxial pre-exponential factor, $I_2$ is the square root of the second invariant of the strain rate tensor, $Q$ is the
-activation energy, $p$ is the pressure, $V$ is the activation volume, $R$ is the gas constant and $T$ is the temperature.
-Pressures and temperatures are determined by the interpolation of the nodal parameters, while strain rates are calculated by means of the
+\eta_{df|ds}=\left(\frac{d^m}{A}\right)^{\frac{1}{n}}I_2^{\frac{(1-n)}{n}}\exp\left(\frac{Q+pV}{nRT}\right)
+\end{equation}
+where $d$ is the grain size, $m$ is the grain size exponent, $n$ is the stress exponent, $A$ is the pre-exponential factor, $I_2$ is the square root of the
+second invariant of the strain rate tensor, $Q$ is the activation energy, $p$ is the pressure, $V$ is the activation volume, $R$ is the gas constant and $T$
+is the temperature. Pressures and temperatures are determined by the interpolation of the nodal parameters, while strain rates are calculated by means of the
 derivative of the velocities on the nodes. In case of diffusion creep $n=1$ and $m>0$, in case of dislocation creep $n>1$ and $m=0$. The viscous creep $\eta_{cp}$
@@
 \section{Mantle melting}\label{sec:melt}
-Melt fraction in the mantle depends on temperature ($T$, in \textdegree C), pressure ($p$, in \si{\giga\pascal}) and the water content in the melt ($X_{H_2O}$, in wt.\%).
+Melt fraction in the mantle depends on temperature ($T$, in °C), pressure ($p$, in \si{\giga\pascal}) and the water content in the melt ($X_{H_2O}$, in wt.\%).
 The determination of the melt fraction $M$ is implemented as explained by \citet{Katz2003}, according to the successive modification by \citet{Langmuir2006} and
@@
 maximum velocities ($u_{min}$, $u_{max}$, $v_{min}$, $v_{max}$) and pressures ($p_{min}$, $p_{max}$) on the entire domain, $v_{\textrm{rms}}$ and average pressure on the
-entire domain ($p_{avg}$) are compared with solutions generated by ASPECT \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and all results
+entire domain ($p_{avg}$) are compared with solutions generated by ASPECT \citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} and all results
 can be found at \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere2D} and 
@@
 in terms of velocity and pressure fields and topography variations are compared with results obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and they can be found at
+\citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} and they can be found at
 \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere_fs2D} and
@@
 \[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\].
-The viscosity field $\eta$ is calculated as the harmonic average between a linear part $\eta_{lin}$ that depends on temperature only or on temperature and 
+The viscosity ﬁeld $\eta$ is calculated as the harmonic average between a linear part $\eta_{lin}$ that depends on temperature only or on temperature and 
 depth $y$ and a nonlinear-plastic part $\eta_{pl}$ that depends on the strain rate $\dot{\bm{\epsilon}}$, as follows
@@
 (see Sec. \ref{sec:error}) calculated for both the simulations match well with results from \citet{vanKeken1997} and obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and ELEFANT \citep{Thieulot2014}, obtained with similar grid resolutions
+\citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} and ELEFANT \citep{Thieulot2014}, obtained with similar grid resolutions
 (Fig. \ref{fig:thin}). In addition, a variety of simulations with different resolutions and aspect ratios are performed to verify that the sensitivity of the

### 2022-01-28 .debris/2022-01-27 11.21.09.00/Paper.tex
modified

--- 2022-01-28 .debris/2022-01-27 11.00.19.00/Paper.tex
@@
 \usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
+%\usepackage[english]{babel}
 \usepackage{amsmath}
@@
 \section{Erosion and sedimentation}\label{sec:erosion}
+Surface processes at either the lithosphere-air interface or in correspondence of the free surface have been implemented by means of the software
+Fastscape\footnote{\url{https://fastscape.org/}} \citep{Braun2013,Cordonnier2019,Yuan2019a,Yuan2019b}
 %Surface processes at the lithosphere-air interface are implemented taking into account fluvial erosion (long-range processes), hillslope processes (short-range

### 2022-01-28 .debris/2022-01-27 11.27.05.00/Paper.tex
modified

--- 2022-01-28 .debris/2022-01-27 11.00.19.00/Paper.tex
@@
 \section{Erosion and sedimentation}\label{sec:erosion}
+Surface processes at either the lithosphere-air interface or in correspondence of the free surface have been implemented by means of the software
+Fastscape\footnote{\url{https://fastscape.org/}} \citep{Braun2013,Cordonnier2019,Yuan2019a,Yuan2019b}. Fastscape is an interface to model landscape evolution
+by river incision, sediment transport and deposition in continental and marine environments.
+
+FastScapeLib is a set of routines that solve the stream power law (enriched by a sediment transport/deposition term) (see Yuan et al, 2019a),
+hillslope diffusion and marine transport and deposition (see Yuan et al, 2019b).
 %Surface processes at the lithosphere-air interface are implemented taking into account fluvial erosion (long-range processes), hillslope processes (short-range

### 2022-01-28 .debris/2022-01-27 11.27.36.00/Paper.tex
modified

--- 2022-01-28 .debris/2022-01-27 11.27.05.00/Paper.tex
@@
 
-FastScapeLib is a set of routines that solve the stream power law (enriched by a sediment transport/deposition term) (see Yuan et al, 2019a),
-hillslope diffusion and marine transport and deposition (see Yuan et al, 2019b).
+FastScapeLib is a set of routines that solve the stream power law (enriched by a sediment transport/deposition term) \citep{Yuan2019a},
+hillslope diffusion and marine transport and deposition \citep{Yuan2019b}.
 %Surface processes at the lithosphere-air interface are implemented taking into account fluvial erosion (long-range processes), hillslope processes (short-range

### 2022-01-28 .debris/2022-01-27 11.28.18.00/Paper.tex
modified

--- 2022-01-28 .debris/2022-01-27 11.21.09.00/Paper.tex
@@
 \usepackage[utf8]{inputenc}
-%\usepackage[english]{babel}
+\usepackage[english]{babel}
 \usepackage{amsmath}
@@
 Surface processes at either the lithosphere-air interface or in correspondence of the free surface have been implemented by means of the software
-Fastscape\footnote{\url{https://fastscape.org/}} \citep{Braun2013,Cordonnier2019,Yuan2019a,Yuan2019b}
+Fastscape\footnote{\url{https://fastscape.org/}} \citep{Braun2013,Cordonnier2019,Yuan2019a,Yuan2019b}.FastScapeLib is a set of routines that solve the stream
+power law (enriched by a sediment transport/deposition term) \citep{Yuan2019a}, hillslope diffusion and marine transport and deposition \citep{Yuan2019b}.
 %Surface processes at the lithosphere-air interface are implemented taking into account fluvial erosion (long-range processes), hillslope processes (short-range

### 2022-01-28 .debris/2022-01-27 11.40.55.00/Paper.tex
modified

--- 2022-01-28 .debris/2022-01-27 11.28.18.00/Paper.tex
@@
 power law (enriched by a sediment transport/deposition term) \citep{Yuan2019a}, hillslope diffusion and marine transport and deposition \citep{Yuan2019b}.
+The partial differential equation solved by FastScapeLib is:
+\begin{equation}
+\frac{\partial h}{ \partial t}=U-K_f A^{m_e}S^{n_e}+\frac{G}{A}\int_A \left( U-\frac{\partial h}{\partial t} \right) dA + K_d \bm{\nabla}^2 h
+\end{equation}
+where $h$ is the topography, $U$ is the uplift, $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $S$ is the slope, $m_e$ and $n_e$ are
+parameters that should be chosen to constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020} and $K_d$ is the effective
+diffusivity.
 %Surface processes at the lithosphere-air interface are implemented taking into account fluvial erosion (long-range processes), hillslope processes (short-range

### 2022-01-28 .debris/2022-01-27 11.41.55.00/Paper.tex
modified

--- 2022-01-28 .debris/2022-01-27 11.28.18.00/Paper.tex
@@
 power law (enriched by a sediment transport/deposition term) \citep{Yuan2019a}, hillslope diffusion and marine transport and deposition \citep{Yuan2019b}.
+The partial differential equation solved by FastScapeLib is:
+\begin{equation}
+\frac{\partial h}{ \partial t}=U-K_f A^{m_e}S^{n_e}+\frac{G}{A}\int_A \left( U-\frac{\partial h}{\partial t} \right) dA + K_d \bm{\nabla}^2 h
+\end{equation}
+where $h$ is the topography, $U$ is the uplift, $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $S$ is the slope, $m_e$ and $n_e$ are
+parameters that should be chosen to constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}, $G$ is the dimensionless
+deposition/transport coefficient for the enriched stream power law and $K_d$ is the effective diffusivity.
 %Surface processes at the lithosphere-air interface are implemented taking into account fluvial erosion (long-range processes), hillslope processes (short-range

### 2022-01-28 .debris/2022-01-27 11.42.38.00/Paper.tex
modified

--- 2022-01-28 .debris/2022-01-27 11.41.55.00/Paper.tex
@@
 deposition/transport coefficient for the enriched stream power law and $K_d$ is the effective diffusivity.
+In case of sticky air, slope is determined by means of the markers chain, which is vertically corrected considering erosion and sedimentation rates. After the
+correction, continental markers above the markers chain are transformed in air markers and air markers below the markers chain become sediments. Similarly,
+slope is determined considering the nodes on the top boundary in case of true free surface and they are vertically corrected as for the markers chain. In this
+case, continental markers above the top boundary are simply deleted.
 %Surface processes at the lithosphere-air interface are implemented taking into account fluvial erosion (long-range processes), hillslope processes (short-range

### 2022-01-28 .debris/2022-01-27 11.54.33.00/Paper.tex
modified

--- 2022-01-28 .debris/2022-01-27 11.42.38.00/Paper.tex
@@
 \begin{equation}
-\frac{\partial h}{ \partial t}=U-K_f A^{m_e}S^{n_e}+\frac{G}{A}\int_A \left( U-\frac{\partial h}{\partial t} \right) dA + K_d \bm{\nabla}^2 h
+\frac{\partial h}{ \partial t}=U-K_f A^{m_e}S^{n_e} + \frac{G}{A}\int_A \left( U-\frac{\partial h}{\partial t} \right) dA + K_d \bm{\nabla}^2 h 
 \end{equation}
@@
 parameters that should be chosen to constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}, $G$ is the dimensionless
-deposition/transport coefficient for the enriched stream power law and $K_d$ is the effective diffusivity.
+deposition/transport coefficient for the enriched stream power law and $K_d$ is the effective diffusivity. Terms $$K_f A^{m_e}S^{n_e}$$ and $$K_d \bm{\nabla}^2 h$$
+consider long-range processes described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020} and short-range processes are
+proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}.
+
 In case of sticky air, slope is determined by means of the markers chain, which is vertically corrected considering erosion and sedimentation rates. After the
@@
 \end{equation}
-where $d$ is the grain size, $m$ is the grain size exponent, $n$ is the stress exponent, $A$ is the pre-exponential factor, $I_2$ is the square root of the
-second invariant of the strain rate tensor, $Q$ is the activation energy, $p$ is the pressure, $V$ is the activation volume, $R$ is the gas constant and $T$
+where $d$ is the grain size, $m$ is the grain size exponent, $n$ is the stress exponent, $A$ is the un9i-axial pre-exponential factor, $I_2$ is the square root of
+the second invariant of the strain rate tensor, $Q$ is the activation energy, $p$ is the pressure, $V$ is the activation volume, $R$ is the gas constant and $T$
 is the temperature. Pressures and temperatures are determined by the interpolation of the nodal parameters, while strain rates are calculated by means of the

### 2022-01-28 .debris/2022-01-27 11.57.22.00/Paper.tex
modified

--- 2022-01-28 .debris/2022-01-27 11.54.33.00/Paper.tex
@@
 \begin{equation}\label{eq:viscous}
-\eta_{df|ds}=\left(\frac{d^m}{A}\right)^{\frac{1}{n}}I_2^{\frac{(1-n)}{n}}\exp\left(\frac{Q+pV}{nRT}\right)
-\end{equation}
-where $d$ is the grain size, $m$ is the grain size exponent, $n$ is the stress exponent, $A$ is the un9i-axial pre-exponential factor, $I_2$ is the square root of
+\eta_{df|ds}=\left(\frac{d^m}{A}\right)^{\frac{1}{n}} 2^{\frac{1-n}{2n}} I_2^{\frac{(1-n)}{n}}\exp\left(\frac{Q+pV}{nRT}\right)
+\end{equation}
+where $d$ is the grain size, $m$ is the grain size exponent, $n$ is the stress exponent, $A$ is the uniaxial pre-exponential factor, $I_2$ is the square root of
 the second invariant of the strain rate tensor, $Q$ is the activation energy, $p$ is the pressure, $V$ is the activation volume, $R$ is the gas constant and $T$

### 2022-01-28 .debris/2022-01-27 12.02.23.00/Paper.tex
modified

--- 2022-01-28 .debris/2022-01-27 11.54.33.00/Paper.tex
@@
 \begin{equation}\label{eq:viscous}
-\eta_{df|ds}=\left(\frac{d^m}{A}\right)^{\frac{1}{n}}I_2^{\frac{(1-n)}{n}}\exp\left(\frac{Q+pV}{nRT}\right)
-\end{equation}
-where $d$ is the grain size, $m$ is the grain size exponent, $n$ is the stress exponent, $A$ is the un9i-axial pre-exponential factor, $I_2$ is the square root of
-the second invariant of the strain rate tensor, $Q$ is the activation energy, $p$ is the pressure, $V$ is the activation volume, $R$ is the gas constant and $T$
-is the temperature. Pressures and temperatures are determined by the interpolation of the nodal parameters, while strain rates are calculated by means of the
+\eta_{df|ds}=f_s \left(\frac{d^m}{A}\right)^{\frac{1}{n}} \frac{2^{\frac{1-n}{n}}}{3^{\frac{n+1}{2n}}} I_2^{\frac{(1-n)}{n}}\exp\left(\frac{Q+pV}{nRT}\right)
+\end{equation}
+where $f_s$ is a scaling factor used to represent lithologies that are stronger or weaker than the base set, $d$ is the grain size, $m$ is the grain size exponent,
+$n$ is the stress exponent, $A$ is the uniaxial pre-exponential factor, $I_2$ is the square root of the second invariant of the strain rate tensor, $Q$ is the
+activation energy, $p$ is the pressure, $V$ is the activation volume, $R$ is the gas constant and $T$ is the temperature.
+Pressures and temperatures are determined by the interpolation of the nodal parameters, while strain rates are calculated by means of the
 derivative of the velocities on the nodes. In case of diffusion creep $n=1$ and $m>0$, in case of dislocation creep $n>1$ and $m=0$. The viscous creep $\eta_{cp}$

### 2022-01-28 .debris/2022-01-27 12.30.55.00/Paper.tex
modified

--- 2022-01-28 .debris/2022-01-27 12.02.23.00/Paper.tex
@@
 \section{Introduction}\label{sec:intro}
-This supplement includes a description of the code, with a presentation of the features implemented followed by results of extensive benchmarks listed in
+This document includes a description of the code, with a presentation of the features implemented followed by results of extensive benchmarks listed in
 Table \ref{tab:list}, implemented to verify that the code solves correctly a large range of problems. Description and results of all benchmarks are shown

### 2022-01-28 .debris/2022-01-27 13.15.54.00/Paper.tex
modified

--- 2022-01-28 .debris/2022-01-27 12.30.55.00/Paper.tex
@@
 \section{Introduction}\label{sec:intro}
-This document includes a description of the code, with a presentation of the features implemented followed by results of extensive benchmarks listed in
+This document includes a description of the code FALCON, with a presentation of the features implemented followed by results of extensive benchmarks listed in
 Table \ref{tab:list}, implemented to verify that the code solves correctly a large range of problems. Description and results of all benchmarks are shown

### 2022-01-28 .debris/2022-01-27 13.19.32.00/Paper.tex
modified

--- 2022-01-28 .debris/2022-01-27 13.15.54.00/Paper.tex
@@
 Surface processes at either the lithosphere-air interface or in correspondence of the free surface have been implemented by means of the software
-Fastscape\footnote{\url{https://fastscape.org/}} \citep{Braun2013,Cordonnier2019,Yuan2019a,Yuan2019b}.FastScapeLib is a set of routines that solve the stream
+Fastscape\footnote{\url{https://fastscape.org/}} \citep{Braun2013,Cordonnier2019,Yuan2019a,Yuan2019b}. FastScapeLib is a set of routines that solve the stream
 power law (enriched by a sediment transport/deposition term) \citep{Yuan2019a}, hillslope diffusion and marine transport and deposition \citep{Yuan2019b}.

### 2022-01-28 .debris/2022-01-27 13.27.16.00/Paper.tex
modified

--- 2022-01-28 .debris/2022-01-27 13.23.17.00/Paper.tex
@@
 \begin{equation}\label{eq:viscous}
-\eta_{df|ds}=f_s \left(\frac{d^m}{A}\right)^{\frac{1}{n}} \frac{2^{\frac{1-n}{n}}}{3^{\frac{n+1}{2n}}} I_2^{\frac{(1-n)}{n}}\exp\left(\frac{Q+pV}{nRT}\right)
+\eta_{df|ds}=f_s \left(\frac{d^m}{A}\right)^{\frac{1}{n}} \left (\frac{2^{\frac{1-n}{n}}}{3^{\frac{n+1}{2n}}} \right)
+I_2^{\frac{(1-n)}{n}}\exp\left(\frac{Q+pV}{nRT}\right)
 \end{equation}

### 2022-01-28 .debris/2022-01-27 14.29.51.00/Paper.tex
modified

--- 2022-01-28 .debris/2022-01-27 13.27.16.00/Paper.tex
@@
 parameters that should be chosen to constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}, $G$ is the dimensionless
-deposition/transport coefficient for the enriched stream power law and $K_d$ is the effective diffusivity. Terms $$K_f A^{m_e}S^{n_e}$$ and $$K_d \bm{\nabla}^2 h$$
+deposition/transport coefficient for the enriched stream power law and $K_d$ is the effective diffusivity. Terms $K_f A^{m_e}S^{n_e}$ and $K_d \bm{\nabla}^2 h$
 consider long-range processes described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020} and short-range processes are

### 2022-01-28 .debris/2022-01-27 14.30.56.00/Paper.tex
modified

--- 2022-01-28 .debris/2022-01-27 14.29.51.00/Paper.tex
@@
 This document includes a description of the code FALCON, with a presentation of the features implemented followed by results of extensive benchmarks listed in
-Table \ref{tab:list}, implemented to verify that the code solves correctly a large range of problems. Description and results of all benchmarks are shown
+Table \ref{tab:list}, performed to verify that the code solves correctly a large range of problems. Description and results of all benchmarks are shown
 in Section \ref{sec:benchmark}.

### 2022-01-28 .debris/2022-01-27 14.32.17.00/Paper.tex
modified

--- 2022-01-28 .debris/2022-01-27 14.30.56.00/Paper.tex
@@
 This document includes a description of the code FALCON, with a presentation of the features implemented followed by results of extensive benchmarks listed in
-Table \ref{tab:list}, performed to verify that the code solves correctly a large range of problems. Description and results of all benchmarks are shown
+Table \ref{tab:list}, performed to verify that the code solves correctly a wide range of problems. Description and results of all benchmarks are shown
 in Section \ref{sec:benchmark}.

### 2022-01-28 .debris/2022-01-27 14.33.22.00/Paper.tex
modified

--- 2022-01-28 .debris/2022-01-27 14.32.17.00/Paper.tex
@@
 Surface processes at either the lithosphere-air interface or in correspondence of the free surface have been implemented by means of the software
-Fastscape\footnote{\url{https://fastscape.org/}} \citep{Braun2013,Cordonnier2019,Yuan2019a,Yuan2019b}. FastScapeLib is a set of routines that solve the stream
-power law (enriched by a sediment transport/deposition term) \citep{Yuan2019a}, hillslope diffusion and marine transport and deposition \citep{Yuan2019b}.
-The partial differential equation solved by FastScapeLib is:
+Fastscape\footnote{\url{https://fastscape.org/}} \citep{Braun2013,Cordonnier2019,Yuan2019a,Yuan2019b}. FastScapeLib\footnote{\url{https://fastscape.org/}}
+is a set of routines that solve the stream power law (enriched by a sediment transport/deposition term) \citep{Yuan2019a}, hillslope diffusion and marine
+transport and deposition \citep{Yuan2019b}. The partial differential equation solved by FastScapeLib\footnote{\url{https://fastscape.org/}} is:
 \begin{equation}

### 2022-01-28 .debris/2022-01-27 15.12.43.00/Paper.tex
modified

--- 2022-01-28 .debris/2022-01-27 14.32.17.00/Paper.tex
@@
 deposition/transport coefficient for the enriched stream power law and $K_d$ is the effective diffusivity. Terms $K_f A^{m_e}S^{n_e}$ and $K_d \bm{\nabla}^2 h$
-consider long-range processes described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020} and short-range processes are
+consider long-range processes, described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, and short-range processes,
 proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}.

### 2022-01-28 .debris/2022-01-27 15.13.53.00/Paper.tex
modified

--- 2022-01-28 .debris/2022-01-27 15.12.43.00/Paper.tex
@@
 markers. The correctness of the evolution of the markers chain is tested with the experiment of a 2D time-dependent Stokes sphere below a free surface and
-compared with results from ASPECT\footnote{\url{https://aspect.geodynamics.org/}} \citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} (Section
+compared with results from ASPECT\footnote{\url{https://aspect.geodynamics.org/}} \citep{Kronbichler2012,Heister2017,aspect,aspectmanual} (Section
 \ref{sec:sphere}). In the true free surface case the top boundary is assumed stress-free and velocities are not fixed. In this case topography variations are

### 2022-01-28 .debris/2022-01-27 15.42.10.00/Paper.tex
modified

--- 2022-01-28 .debris/2022-01-27 15.13.53.00/Paper.tex
@@
 maximum velocities ($u_{min}$, $u_{max}$, $v_{min}$, $v_{max}$) and pressures ($p_{min}$, $p_{max}$) on the entire domain, $v_{\textrm{rms}}$ and average pressure on the
-entire domain ($p_{avg}$) are compared with solutions generated by ASPECT \citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} and all results
+entire domain ($p_{avg}$) are compared with solutions generated by ASPECT \citep{Kronbichler2012,Heister2017,aspect,aspectmanual} and all results
 can be found at \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere2D} and 
@@
 in terms of velocity and pressure fields and topography variations are compared with results obtained with ASPECT
-\citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} and they can be found at
+\citep{Kronbichler2012,Heister2017,aspect,aspectmanual} and they can be found at
 \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere_fs2D} and
@@
 (see Sec. \ref{sec:error}) calculated for both the simulations match well with results from \citet{vanKeken1997} and obtained with ASPECT
-\citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} and ELEFANT \citep{Thieulot2014}, obtained with similar grid resolutions
+\citep{Kronbichler2012,Heister2017,aspect,aspectmanual} and ELEFANT \citep{Thieulot2014}, obtained with similar grid resolutions
 (Fig. \ref{fig:thin}). In addition, a variety of simulations with different resolutions and aspect ratios are performed to verify that the sensitivity of the

### 2022-01-28 .debris/2022-01-27/Paper.tex
modified

--- 2022-01-28 .debris/2022-01-27 11.42.38.00/Paper.tex
@@
 \section{Erosion and sedimentation}\label{sec:erosion}
-Surface processes at either the lithosphere-air interface or in correspondence of the free surface have been implemented by means of the software
-Fastscape\footnote{\url{https://fastscape.org/}} \citep{Braun2013,Cordonnier2019,Yuan2019a,Yuan2019b}.FastScapeLib is a set of routines that solve the stream
-power law (enriched by a sediment transport/deposition term) \citep{Yuan2019a}, hillslope diffusion and marine transport and deposition \citep{Yuan2019b}.
-The partial differential equation solved by FastScapeLib is:
+Surface processes at the lithosphere-air interface are implemented taking into account fluvial erosion (long-range processes), hillslope processes (short-range
+processes) and failure of slopes exceeding a threshold angle $\theta$ \citep{Burov1997,Tucker1999,Sacek2012,Croissant2014,Ueda2015,Beucher2020}. Erosion rates
+$\dot{E}_{lr}$ due to long-range processes can be described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, in according to
 \begin{equation}
-\frac{\partial h}{ \partial t}=U-K_f A^{m_e}S^{n_e}+\frac{G}{A}\int_A \left( U-\frac{\partial h}{\partial t} \right) dA + K_d \bm{\nabla}^2 h
-\end{equation}
-where $h$ is the topography, $U$ is the uplift, $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $S$ is the slope, $m_e$ and $n_e$ are
-parameters that should be chosen to constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}, $G$ is the dimensionless
-deposition/transport coefficient for the enriched stream power law and $K_d$ is the effective diffusivity.
-In case of sticky air, slope is determined by means of the markers chain, which is vertically corrected considering erosion and sedimentation rates. After the
-correction, continental markers above the markers chain are transformed in air markers and air markers below the markers chain become sediments. Similarly,
-slope is determined considering the nodes on the top boundary in case of true free surface and they are vertically corrected as for the markers chain. In this
-case, continental markers above the top boundary are simply deleted.
-%Surface processes at the lithosphere-air interface are implemented taking into account fluvial erosion (long-range processes), hillslope processes (short-range
-%processes) and failure of slopes exceeding a threshold angle $\theta$ \citep{Burov1997,Tucker1999,Sacek2012,Croissant2014,Ueda2015,Beucher2020}. Erosion rates
-%$\dot{E}_{lr}$ due to long-range processes can be described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{lr}=K_f A^{m_e} \bm{\nabla} h^{n_e}
-%\end{equation}
-%where $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $\bm{\nabla} h$ is the slope and $m_e$ and $n_e$ are parameters chosen in to
-%constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}. Erosion rates $\dot{E}_{sr}$ due to short-range processes are
-%implemented by assuming that transport is linearly proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{sr}=K_d \bm{\nabla}^2 h
-%\end{equation}
-%where $K_d$ is the effective diffusivity. In addition, the slope is modified if it still exceeds a threshold angle after long- and short-range processes
-%\citep{Ueda2015}.
-
-%Offshore sedimentation processes are implemented similarly to hillslope processes but with a depth-dependent coefficient \citep{Sacek2012,Beucher2020}.
-%The sedimentation rate $\dot{S}$ is calculated as
-%\begin{equation}
-%\dot{S}=K_s \bm{\nabla}^2 h
-%\end{equation}
-%with
-%\begin{equation}
-%K_s=K^*_s \exp\left(\frac{h-h_{sl}}{h_0}\right)
-%\end{equation}
-%where $K^*_s$ is a constant, $h_{sl}$ is the sea level and $h_0$ is a characteristic length. In case of sticky air, slope is determined by means of the markers
-%chain, which is vertically corrected considering erosion and sedimentation rates. After the correction, continental markers above the markers chain are
-%transformed in air markers and air markers below the markers chain become sediments. Similarly, slope is determined considering the nodes on the top boundary
-%in case of true free surface and they are vertically corrected as for the markers chain. In this case, continental markers above the top boundary are simply
-%deleted.
+\dot{E}_{lr}=K_f A^{m_e} \bm{\nabla} h^{n_e}
+\end{equation}
+where $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $\bm{\nabla} h$ is the slope and $m_e$ and $n_e$ are parameters chosen in to
+constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}. Erosion rates $\dot{E}_{sr}$ due to short-range processes are
+implemented by assuming that transport is linearly proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}, in according to
+\begin{equation}
+\dot{E}_{sr}=K_d \bm{\nabla}^2 h
+\end{equation}
+where $K_d$ is the effective diffusivity. In addition, the slope is modified if it still exceeds a threshold angle after long- and short-range processes
+\citep{Ueda2015}.
+
+Offshore sedimentation processes are implemented similarly to hillslope processes but with a depth-dependent coefficient \citep{Sacek2012,Beucher2020}.
+The sedimentation rate $\dot{S}$ is calculated as
+\begin{equation}
+\dot{S}=K_s \bm{\nabla}^2 h
+\end{equation}
+with
+\begin{equation}
+K_s=K^*_s \exp\left(\frac{h-h_{sl}}{h_0}\right)
+\end{equation}
+where $K^*_s$ is a constant, $h_{sl}$ is the sea level and $h_0$ is a characteristic length. In case of sticky air, slope is determined by means of the markers
+chain, which is vertically corrected considering erosion and sedimentation rates. After the correction, continental markers above the markers chain are
+transformed in air markers and air markers below the markers chain become sediments. Similarly, slope is determined considering the nodes on the top boundary
+in case of true free surface and they are vertically corrected as for the markers chain. In this case, continental markers above the top boundary are simply
+deleted.
 

### 2022-03-31 FALCON_description.tex
mod

@@
 \end{equation}
-where $\rho_s$ and $\rho_m$ are the densities of the solid and the molten rock, respectively \citep{Gerya2003a,Wang2016}. The density of the solid rock is
-extracted by Perple\_X, while the density of the molten portion is calculated as
+where $\rho_s$ and $\rho_m$ are the densities of the solid and the molten rock, respectively \citep{Gerya2004b,Gerya2007,Wang2016}. The density of the
+solid rock is extracted by Perple\_X, while the density of the molten portion is calculated as
 \begin{equation}\label{eq:melt_dens_M}
@@
 where $\rho_0$ is the density at the reference temperature ($T_0$) and pressure($p_0$), and $\alpha$ and $\beta$ are the thermal expansion and compressibility
-coefficients, respectively \citep{Gerya2003a,Wang2016}.
+coefficients, respectively \citep{Gerya2004b,Gerya2007,Wang2016}.
 
@@
 where $C_p$ and $\alpha$ are the specific heat and the coefficient of thermal expansion of the solid rock, respectively, and $H_L$ is the latent heat of the
-molten rock \citep{Gerya2003a,Gerya2007,Tackley2010}.
+molten rock \citep{Gerya2004b,Gerya2007,Tackley2010}.
 

### 2022-04-15 FALCON_description.tex
healing

@@
 To support large viscosities variations the penalty parameter is related to the effective elemental viscosity by means of a dimensionless coefficient, so that
-$\lambda=\lambda_e(e)\eta_{eff}(e)$ \citep{Marotta2006,Dabrowski2008,Thieulot2014}. Benchmark of the falling block \citep{Gerya2003a,Gerya2010b,Thieulot2011}
+$\lambda=\lambda_e(e)\eta_{eff}(e)$ \citep{Marotta2006,Dabrowski2008,Thieulot2014}. Benchmark of the falling block \citep{Gerya2003a,Gerya2010,Thieulot2011}
 is performed to verify that the code can correctly deal buoyancy driven flows with strong viscosity contrasts (Section \ref{sec:block}).
@@
 
+Plastic yielding and viscous creep are then combined to obtain a viscoplastic viscosity $\eta_{vp}$ as follows, assuming that they are independent processes
+\citep{Karato2008,Glerum2018}:
+\begin{equation}\label{eq:viscoplastic}
+\eta_{vp}=\min(\eta_{cp},\eta_{pl})
+\end{equation}
+Finally, effective viscosity $\eta_{eff}$ is capped by the minimum and the maximum viscosity ($\eta_{min}$ and $\eta_{max}$, respectively) to avoid extremely
+low or high viscosity \citep{Glerum2018} as follows
+\begin{equation}\label{eq:effective}
+\eta_{eff}=\min(\max(\eta_{vp},\eta_{min}),\eta_{max})
+\end{equation}
+Viscous creep and plastic yielding are non-linear rheologies because of their dependence on the velocity field through pressure and strain rates. Therefore,
+the solution is determined by means of Picard-type iterations, until convergence of the velocity field \citep{Glerum2018}. The convergence is verified at each
+iteration $i$ via the nonlinear residual $\bm{\mathcal{R}}^i$ that can be determined as
+\begin{equation}\label{eq:convergence}
+\bm{\mathcal{R}}^i= \mathbb{K} (\eta_{eff}(\dot{\bm{\epsilon}}^{i-1},p^{i-1}))\cdot \bm{v}^{i-1} - \bm{f}^i
+\end{equation}
+%\begin{equation}\label{eq:convergence}
+%\chi=1-\frac{\langle (\bm{v}^i-\langle\bm{v}^i\rangle)\cdot (\bm{v}^{i+1}-\langle\bm{v}^{i+1}\rangle) \rangle}{\sqrt{\sigma^i\sigma^{i+1}}}
+%\end{equation}
+%Iterations are performed until either convergence ($\chi<tol$) or maximal number of non-linear iterations ($it_{max}$) is reached.
+where $\mathbb{K}$ is the Stiffness matrix and $\bm{f}$ is the right hand side vector \citep{Spiegelman2016,Glerum2018}.  The $L_2$-norm is extracted from
+$\bm{\mathcal{R}}^i$ and it is normalised as follows
+\begin{equation}\label{eq:convergence_res}
+\frac{||\bm{\mathcal{R}}^i||_2}{||\bm{\mathcal{R}}^0||_2}
+\end{equation}
+where $||\bm{\mathcal{R}}^0||_2$ is the $L_2$-norm of the first nonlinear iteration.
+Since often the normalised nonlinear residual drop very quickly over the first 2-3 iterations, also the $L_2$-norm of velocity and pressure residuals are
+calculated as
+\begin{equation}\label{eq:convergence_vel}
+\frac{||\Delta\bm{v}||_2}{||\bm{v}||_2}=\frac{||\bm{v}_i-\bm{v}_{i-1}||_2}{||\bm{v}_{i}||_2}
+\end{equation}
+and
+\begin{equation}\label{eq:convergence_press}
+  \frac{||\Delta p||_2}{||p||_2}=\frac{||p_i-p_{i-1}||_2}{||p_{i}||_2}
+\end{equation}
+respectively. Iterations are performed until either maximal number of non-linear iterations ($it_{max}$) is reached or all the normalised nonlinear residual, 
+the velocity residual and the pressure residual converge under a defined tolerance $tol$.
+
+\section{Healing and weakening}
+The temporal evolution of the accumulated strain $\epsilon$ has been implemented as in \citet{Fuchs2019,Fuchs2021}, as following
+\begin{equation}\label{eq:healing}
+  \frac{d\epsilon}{dt}=I_2 - \epsilon H_{rate}(T)
+\end{equation}
+where the first term on the right-hand side is a source term given by the second invariant of the strain rate $I_2$ and the second term a
+temperature-dependent healing factor $H_{rate}$, calculated as
+\begin{equation}\label{eq:healing_rate}
+  H_{rate}(T)=B \exp \left[-\frac{\mu}{2}\left(\frac{1}{T+1}-\frac{1}{2}\right)\right]
+\end{equation}
+where $B$ is a constant describing the time scale of healing, while $\mu$ and $T$ are the non-dimensional temperature activation and temperature,
+respectively \citep{Fuchs2019,Fuchs2021}.
+
 Strain softening is taken into account for both viscous creep and plastic viscosity \citep{Huismans2003,Babeyko2005,Huismans2005,Sobolev2005,Warren2008a} by
@@
 
-Plastic yielding and viscous creep are then combined to obtain a viscoplastic viscosity $\eta_{vp}$ as follows, assuming that they are independent processes
-\citep{Karato2008,Glerum2018}:
-\begin{equation}\label{eq:viscoplastic}
-\eta_{vp}=\min(\eta_{cp},\eta_{pl})
-\end{equation}
-Finally, effective viscosity $\eta_{eff}$ is capped by the minimum and the maximum viscosity ($\eta_{min}$ and $\eta_{max}$, respectively) to avoid extremely
-low or high viscosity \citep{Glerum2018} as follows
-\begin{equation}\label{eq:effective}
-\eta_{eff}=\min(\max(\eta_{vp},\eta_{min}),\eta_{max})
-\end{equation}
-Viscous creep and plastic yielding are non-linear rheologies because of their dependence on the velocity field through pressure and strain rates. Therefore,
-the solution is determined by means of Picard-type iterations, until convergence of the velocity field \citep{Glerum2018}. The convergence is verified at each
-iteration $i$ via the nonlinear residual $\bm{\mathcal{R}}^i$ that can be determined as
-\begin{equation}\label{eq:convergence}
-\bm{\mathcal{R}}^i= \mathbb{K} (\eta_{eff}(\dot{\bm{\epsilon}}^{i-1},p^{i-1}))\cdot \bm{v}^{i-1} - \bm{f}^i
-\end{equation}
-%\begin{equation}\label{eq:convergence}
-%\chi=1-\frac{\langle (\bm{v}^i-\langle\bm{v}^i\rangle)\cdot (\bm{v}^{i+1}-\langle\bm{v}^{i+1}\rangle) \rangle}{\sqrt{\sigma^i\sigma^{i+1}}}
-%\end{equation}
-%Iterations are performed until either convergence ($\chi<tol$) or maximal number of non-linear iterations ($it_{max}$) is reached.
-where $\mathbb{K}$ is the Stiffness matrix and $\bm{f}$ is the right hand side vector \citep{Spiegelman2016,Glerum2018}.  The $L_2$-norm is extracted from
-$\bm{\mathcal{R}}^i$ and it is normalised as follows
-\begin{equation}\label{eq:convergence_res}
-\frac{||\bm{\mathcal{R}}^i||_2}{||\bm{\mathcal{R}}^0||_2}
-\end{equation}
-where $||\bm{\mathcal{R}}^0||_2$ is the $L_2$-norm of the first nonlinear iteration.
-Since often the normalised nonlinear residual drop very quickly over the first 2-3 iterations, also the $L_2$-norm of velocity and pressure residuals are
-calculated as
-\begin{equation}\label{eq:convergence_vel}
-\frac{||\Delta\bm{v}||_2}{||\bm{v}||_2}=\frac{||\bm{v}_i-\bm{v}_{i-1}||_2}{||\bm{v}_{i}||_2}
-\end{equation}
-and
-\begin{equation}\label{eq:convergence_press}
-  \frac{||\Delta p||_2}{||p||_2}=\frac{||p_i-p_{i-1}||_2}{||p_{i}||_2}
-\end{equation}
-respectively. Iterations are performed until either maximal number of non-linear iterations ($it_{max}$) is reached or all the normalised nonlinear residual, 
-the velocity residual and the pressure residual converge under a defined tolerance $tol$.
-
 \section{The energy equation}\label{sec:en_eq}
@@
 The correctness of the implementation and the computation of shear and adiabatic heating is verified performing an exercise with an analytical solution
-(Section \ref{sec:simple_shear}) and Exercise 9.4 in \citet{Gerya2010b} (Section \ref{sec:shear}). Furthermore, mantle convection and thin layer entrainment
+(Section \ref{sec:simple_shear}) and Exercise 9.4 in \citet{Gerya2010} (Section \ref{sec:shear}). Furthermore, mantle convection and thin layer entrainment
 benchmarks (Sections \ref{sec:mantle} and \ref{sec:thin}, respectively) are performed to verify that the code solve correctly Eqs. \ref{eq:momentum_penalty}
@@
 \subsection{Falling blocks}\label{sec:block}
-This benchmark is described as proposed by \citet{Gerya2003a}, \citet{Gerya2010b} and \citet{Thieulot2011}. The domain is square with $L_x=L_y=$ \SI{500}{\km}
+This benchmark is described as proposed by \citet{Gerya2003a}, \citet{Gerya2010} and \citet{Thieulot2011}. The domain is square with $L_x=L_y=$ \SI{500}{\km}
 and the grid is composed by \num{50x50} elements with 25 markers in each element. The block is initially centred at ($x=$ \SI{250}{\km}; $y=$ \SI{400}{\km})
@@
 The velocity in the centre of the falling block at $t=0$ is measured for all experiments and, since the velocity should increase with density contrast, the
-quantity $\bm{v}/(\rho_b-\rho_f)$ is plotted as function of the viscosity contrast. All results perfectly match with those from \citet{Gerya2010b} and they
+quantity $\bm{v}/(\rho_b-\rho_f)$ is plotted as function of the viscosity contrast. All results perfectly match with those from \citet{Gerya2010} and they
 line up on a single curve, demonstrating that the code can correctly deal with large viscosity and density contrasts (Fig. \ref{fig:falling}). All data can be 
@@
 \subsection{Shear and adiabatic heating}\label{sec:shear}
-This problem is performed as presented in Exercise 9.4 in \citet{Gerya2010b}. Two materials are vertically separated in a rectangular domain with $L_x=$
+This problem is performed as presented in Exercise 9.4 in \citet{Gerya2010}. Two materials are vertically separated in a rectangular domain with $L_x=$
 \SI{1000}{\km}, $L_y=$ \SI{1500}{\km} and a grid resolution of \num{30x20} elements. Constant thermal coefficient expansion ($\alpha=$ \SI{3e-5}{\per\kelvin}),
@@
 As shown in Fig. \ref{fig:energy}, both shear and adiabatic heating predicted by the code (first row in Fig. \ref{fig:energy}) well recreate the results
-obtained by means of example ${Shear\_adiabatic\_heating.m}$ from \citet{Gerya2010b} (second row in Fig. \ref{fig:energy}). 
+obtained by means of example ${Shear\_adiabatic\_heating.m}$ from \citet{Gerya2010} (second row in Fig. \ref{fig:energy}). 
 All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Adiabatic%2BShear}.
@@
 \caption{Comparison between shear (first column), adiabatic (second column) and total (third column) energy predicted by the code (panels a, b and c) and those
-created using example ${Shear\_adiabatic\_heating.m}$ from Exercise 9.4 in \citet{Gerya2010b} (panels d, e and f).}
+created using example ${Shear\_adiabatic\_heating.m}$ from Exercise 9.4 in \citet{Gerya2010} (panels d, e and f).}
 \label{fig:energy}

### 2022-07-20 FALCON_description.tex
fixed

@@
 \centering
-\noindent\includegraphics[width=\textwidth]{./Figures/errors.pdf}
+\includegraphics[width=\linewidth]{./Figures/errors.pdf}
 \caption{Velocity and pressure error for the Stokes flow experiment between generated and analytical solution as a function of element size (panel a),
@@
 \begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/MUMPS.pdf}
+\includegraphics[width=\linewidth]{./Figures/MUMPS.pdf}
 \caption{Panel a) Analysis, factorisation, solution and total solve times (panel a) and factorisation memory usage (panel b) as a function of the total number
@@
 \centering
-\noindent\includegraphics[width=200px]{./Figures/Zalesak.pdf}
+\includegraphics[width=200px]{./Figures/Zalesak.pdf}
 \caption{Position of the Zalesak disk at $t=0$, 0.25, 0.5 and 0.75 (orange, red, blue and green, respectively) for the case with Courant number of 1.}
@@
 \centering
-\noindent\includegraphics[width=400px]{./Figures/zal_marker.pdf}
+\includegraphics[width=400px]{./Figures/zal_marker.pdf}
 \caption{Distance from the centre as function of time for values of Courant number of  of 0.25, 0.3, 0.5, 0.75 and 1 (green, orange, blue, black and red,
@@
 \centering
-\noindent\includegraphics[width=\textwidth]{./Figures/CVI.pdf}
+\includegraphics[width=\linewidth]{./Figures/CVI.pdf}
 \caption{Maximum and minimum number of markers per element throughout the experiment, with (green and orange lines) or without (purple and blue lines)
@@
 \centering
-\noindent\includegraphics[width=\textwidth]{./Figures/CVI_time.pdf}
-\caption{Number of markers per element and markers distribution through time using 2nd.order Runge-Kutta scheme.
+\includegraphics[width=\linewidth]{./Figures/CVI_time.pdf}
+\caption{Number of markers per element and markers distribution through time using 2nd-order Runge-Kutta scheme.
 Panels a: number of markers per element (panel a1) and distribution of the markers (panel a2) after the first
@@
 \begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Poiseuille.pdf}
+\includegraphics[width=\linewidth]{./Figures/Poiseuille.pdf}
 \caption{Velocity field (panel a), pressure (panel b) and divergence velocity of a Poiseuille flow in case of the classic penalty method (no iterations) and
@@
 \begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Analytical.pdf}
+\includegraphics[width=\linewidth]{./Figures/Analytical.pdf}
 \caption{Velocity (panel a) and pressure (panel b) field predicted by the model for a Poiseuille flow with respect with their analytical solutions. The
@@
 \centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_NS.pdf}
+\includegraphics[width=400px]{./Figures/vrms_NS.pdf}
 \caption{$v_{\textrm{rms}}$ as function of element size for the instantaneous 2D sphere experiment in case of no slip boundary conditions and with different average
@@
 \centering
-\noindent\includegraphics[width=400px]{./Figures/RT.pdf}
+\includegraphics[width=400px]{./Figures/RT.pdf}
 \caption{$v_{\textrm{rms}}$ of the Rayleigh-Taylor experiment as function of time for different resolution of the grid.}
@@
 \centering
-\noindent\includegraphics[width=300px]{./Figures/Rayleigh.pdf}
+\includegraphics[width=300px]{./Figures/Rayleigh.pdf}
 \caption{Evolution of the Rayleigh-Taylor experiment for a grid of \num{256x256} elements at $t=0, 500, 1000$ and $2000$ (panels a, b, c and d, respectively).}
@@
 \begin{figure}
-\noindent\includegraphics[width=300px]{./Figures/Block.pdf}
+\includegraphics[width=300px]{./Figures/Block.pdf}
 \centering
@@
 \centering
-\noindent\includegraphics[width=400px]{./Figures/Falling.pdf}
+\includegraphics[width=400px]{./Figures/Falling.pdf}
 \caption{Initial velocity relative to the density contrast at the centre of the falling block as function of the viscosity contrast between the block and the
@@
 \centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_arithm.pdf}
+\includegraphics[width=400px]{./Figures/vrms_arithm.pdf}
 \caption{$v_{\textrm{rms}}$ for different grid resolutions as function of time for the 2D time-dependent Stokes sphere experiment in case of an arithmetic average.
@@
 \centering
-\noindent\includegraphics[width=400px]{./Figures/Kaus_random.pdf}
+\includegraphics[width=400px]{./Figures/Kaus_random.pdf}
 \caption{Evolution of the y coordinate in $x=L_x$ as function of time for different time steps, with and without the stabilisation algorithm.}
@@
 \centering
-\noindent\includegraphics[width=400px]{./Figures/Crameri.pdf}
+\includegraphics[width=400px]{./Figures/Crameri.pdf}
 \caption{Maximum topography as function of time for the Crameri benchmark (coloured lines) in comparison with the analytical solution (black line) and results
@@
 \centering
-\noindent\includegraphics[width=400px]{./Figures/Subduction.pdf}
+\includegraphics[width=400px]{./Figures/Subduction.pdf}
 \caption{Maximum depth of the slab as function of time for all the simulations of the \citet{Schmeling2008} benchmark, in case of harmonic, geometric and
@@
 \begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Slab.pdf}
+\includegraphics[width=\linewidth]{./Figures/Slab.pdf}
 \caption{Effective viscosity for the slab detachment benchmark at the beginning of the evolution (panel a) and when necking is complete (panel b).}
@@
 \centering
-\noindent\includegraphics[width=400px]{./Figures/Necking.pdf}
+\includegraphics[width=400px]{./Figures/Necking.pdf}
 \caption{Normalised width of the necking of the slab detachment benchmark as function of normalised time.}
@@
 \centering
-\noindent\includegraphics[width=400px]{./Figures/Indenter.pdf}
+\includegraphics[width=400px]{./Figures/Indenter.pdf}
 \caption{Viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure (panels g and h) fields for rough (left column) and
@@
 \centering
-\noindent\includegraphics[width=400px]{./Figures/Smooth_Rough.pdf}
+\includegraphics[width=400px]{./Figures/Smooth_Rough.pdf}
 \caption{Nodal pressure in function of the x coordinate at $y=1$ for rough (green line) and smooth (blue line) punch experiment. Black and red lines indicate
@@
 \centering
-\noindent\includegraphics[width=400px]{./Figures/Brick_Beam.pdf}
+\includegraphics[width=400px]{./Figures/Brick_Beam.pdf}
 \caption{Shear band angles predicted for the brick experiment for $\phi=0$° (panel a) and for $\phi=20$° in case of compression and extension (panels b and c,
@@
 \centering
-\noindent\includegraphics[width=400px]{./Figures/Brick.pdf}
+\includegraphics[width=400px]{./Figures/Brick.pdf}
 \caption{Shear band angles predicted for the brick experiment in case of compressional and extensional contexts as function of different internal angle of
@@
   \centering
-  \noindent\includegraphics[width=500px]{./Figures/convergence.pdf}
+  \includegraphics[width=500px]{./Figures/convergence.pdf}
   \caption{Convergence rates for all tests. Normalised residuals (left column) are compared with velocity residuals (right column) for extensional (first row)
@@
 \centering
-\noindent\includegraphics[width=400px]{./Figures/Advection.pdf}
+\includegraphics[width=400px]{./Figures/Advection.pdf}
 \caption{Temperature profile as function of \textit{x} for the advection stabilisation benchmark. Purple line indicates the initial temperature profile; the
@@
 %\centering
-%\noindent\includegraphics[width=400px]{./Figures/Energy_Thieulot.pdf}
+%\includegraphics[width=400px]{./Figures/Energy_Thieulot.pdf}
 %\caption{Temperature and shear heating predicted by the model after 1 time step.}
@@
 \centering
-\noindent\includegraphics[width=500px]{./Figures/Analytical_Shear.pdf}
+\includegraphics[width=500px]{./Figures/Analytical_Shear.pdf}
 \caption{Velocity (panel a), temperature (panel b) and shear heating (panel c) as function of the \textit{y} coordinate for the simple shear experiment. The
@@
 \centering
-\noindent\includegraphics[width=400px]{./Figures/Energy.pdf}
+\includegraphics[width=400px]{./Figures/Energy.pdf}
 \caption{Comparison between shear (first column), adiabatic (second column) and total (third column) energy predicted by the code (panels a, b and c) and those
@@
 \centering
-\noindent\includegraphics[width=400px]{./Figures/Convection.pdf}
+\includegraphics[width=400px]{./Figures/Convection.pdf}
 \caption{$v_{\textrm{rms}}$ (panels a, c and e) and Nu (panels b, d and f) for the mantle convection benchmark as function of time for different grid resolution.
@@
 \small
-\begin{tabular}{|>{\centering}m{0.09\textwidth}|>{\centering}m{0.05\textwidth}|>{\centering}m{0.24\textwidth}|>{\centering}m{0.15\textwidth}|
-  >{\centering}m{0.1\textwidth} >{\centering}m{0.1\textwidth} >{\centering\arraybackslash}m{0.1\textwidth}|}
+\begin{tabular}{|>{\centering}m{0.09\linewidth}|>{\centering}m{0.05\linewidth}|>{\centering}m{0.24\linewidth}|>{\centering}m{0.15\linewidth}|
+  >{\centering}m{0.1\linewidth} >{\centering}m{0.1\linewidth} >{\centering\arraybackslash}m{0.1\linewidth}|}
 \toprule
@@
 %  \centering
-%  \resizebox{\textwidth}{!}{
+%  \resizebox{\linewidth}{!}{
 %\begin{tabular}{|c|ccc|ccc|ccc|ccc|}
@@
   \small
-\begin{tabular}{|>{\centering}m{0.08\textwidth}|>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}
-  >{\centering}m{0.05\textwidth} >{\centering}m{0.05\textwidth} |>{\centering\arraybackslash}m{0.2\textwidth}|}
+\begin{tabular}{|>{\centering}m{0.08\linewidth}|>{\centering}m{0.05\linewidth}>{\centering}m{0.05\linewidth}>{\centering}m{0.05\linewidth}
+  >{\centering}m{0.05\linewidth} >{\centering}m{0.05\linewidth} |>{\centering\arraybackslash}m{0.2\linewidth}|}
   \toprule
@@
   \centering
-  \resizebox{\textwidth}{!}{
+  \resizebox{\linewidth}{!}{
 \begin{tabular}{|cc|ccccccccccc|}
@@
 \centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_thin.pdf}
+\includegraphics[width=400px]{./Figures/vrms_thin.pdf}
 \caption{$v_{\textrm{rms}}$ for the thin layer experiment as function of time for different grid resolution. Results are compared with results obtained by
@@
 \centering
-\noindent\includegraphics[width=400px]{./Figures/vrmszero.pdf}
+\includegraphics[width=400px]{./Figures/vrmszero.pdf}
 \caption{$v_{\textrm{rms}}$ at $t=0$ for the thin layer experiment as function of the element size for different aspect ratios of the elements. Results are compared
@@
 \centering
-\noindent\includegraphics[width=400px]{./Figures/Hydration.pdf}
+\includegraphics[width=400px]{./Figures/Hydration.pdf}
 \caption{Distribution of bound and free water (left and right columns, respectively) at $t=0.5$ and $t=$ \SI{1.25}{\mega\year} (first and second rows,
@@
 \begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Fraction.pdf}
+\includegraphics[width=\linewidth]{./Figures/Fraction.pdf}
 \caption{Isobaric melting curves for 1 GPa (panel a) and 3 GPa (panel b) as function of temperature with different bulk water contents.}
@@
 \begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Water.pdf}
+\includegraphics[width=\linewidth]{./Figures/Water.pdf}
 \caption{Isobaric melting curves for 1.5 GPa as function of bulk water (panel a) and water in the melt (panel b) calculated for different temperatures.}

### 2022-09-07 tex/Benchmark/CVI.tex
update main

--- 2022-03-31 FALCON_description.tex
@@
-\documentclass[hidelinks,10pt,a4paper]{article}
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{textcomp}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage{booktabs}
-\usepackage{multirow}
-\usepackage[cm]{fullpage}
-\usepackage[round]{natbib}
-\bibliographystyle{abbrvnat}
-\usepackage{doi}
-\usepackage{bm}
-\usepackage[table]{xcolor}
-\hypersetup{
-    colorlinks,
-    linkcolor={blue},
-    citecolor={blue},
-    urlcolor={blue}
-}
-\usepackage{float}
-\restylefloat{table}
-\usepackage{siunitx}
-\DeclareSIUnit\year{yr}
-
-\begin{document}
-
-\section{Introduction}\label{sec:intro}
-This document includes a description of the code FALCON, with a presentation of the features implemented followed by results of extensive benchmarks listed in
-Table \ref{tab:list}, performed to verify that the code solves correctly a wide range of problems. Description and results of all benchmarks are shown
-in Section \ref{sec:benchmark}.
-
-\begin{table}[H]
-\caption{List of benchmarks performed for each feature of the code. Data of all benchmarks can be found at 
-\url{https://github.com/aleregorda/Benchmarks}}
-\centering
-\begin{tabular}{| l | l | l |}
-\toprule
- \textbf{Feature}  & \textbf{Benchmark} & \textbf{Section} \\
-\midrule
-  Solver  & Stokes flow & \ref{sec:stokes} \\
-\hline
-  \multirow{2}{*}{Markers advection} & Zalesak disk & \ref{sec:runge} \\
-    & Conservative Interpolation Velocity & \ref{sec:CVI} \\
-\hline
-  \multirow{4}{*}{Momentum equation}  & Poiseuille flow & \ref{sec:poiseuille}\\
-    & Instantaneous 2D sphere & \ref{sec:ist_sphere} \\
-    & Rayleigh-Taylor experiment & \ref{sec:rayleigh} \\
-    & Falling block & \ref{sec:block} \\
-\hline
-    & 2D Stokes sphere & \ref{sec:sphere} \\
-    Sticky air and  & Stabilisation algorithm & \ref{sec:stab} \\
-    free surface & Topography relaxation & \ref{sec:crameri} \\
-  & Spontaneous subduction & \ref{sec:subduction} \\
-\hline
-  Erosion and sedimentation  & - & - \\
-\hline
-   Non-linear  & Slab detachment & \ref{sec:slab} \\
-   visco-plasticity & Indenter & \ref{sec:indenter} \\
-    & Brick & \ref{sec:brick} \\
-\hline
-  \multirow{3}{*}{Energy equation}  & Advection stabilisation & \ref{sec:advection} \\
-    & Simple shear heating & \ref{sec:simple_shear} \\
-    & Shear and adiabatic heating & \ref{sec:shear} \\
-\hline
-\multirow{3}{*}{Energy + momentum} & Mantle convection & \ref{sec:mantle} \\
-& Viscoplastic mantle convection & \ref{sec:visco_mantle} \\
-& Thin layer entrainment & \ref{sec:thin} \\
-\hline
-  Phase changes  & \multirow{2}{*}{Hydrated sinking cylinder} & \multirow{2}{*}{\ref{sec:quinquis}} \\
-  and hydration  &  & \\
-\hline
-  Melting & Experimental melting curves & \ref{sec:katz} \\
-\bottomrule
-\end{tabular}
-\label{tab:list}
-\end{table}
-
-The code is written in Fortran90 and uses the finite element method (FEM) with quadrilateral $Q_1 \times P_0$ elements (continuous bilinear velocity and
-discontinuous constant pressure), associated with MUMPS\footnote{\url{http://mumps.enseeiht.fr/}} \citep{Amestoy2001,Amestoy2019} that is a software package
-for solving systems of linear equations of the form $A \cdot x = b$, where \textit{A} is a square sparse matrix, by means of a direct method. Since
-$Q_1 \times P_0$ elements do not satisfy Ladyzhenskaya, Babuska and Brezzi (LBB) stability condition \citep{Donea2003}, elemental pressure is elaborated
-in post-processing to avoid spurious pressures by means of a double interpolation to smooth the pressure. Following this procedure, the elemental pressure
-is interpolated onto nodes and then back onto elements. After the smoothing procedure, the pressure field is calculated again on the nodes to be used in
-combination with the lithostatic pressure, which is calculated onto nodes. Correctness of the solution and performances of the solver in terms of time and
-memory usage are tested solving the Stokes flow with the analytical solution proposed by \citet{Donea2003} (Section \ref{sec:stokes}).
-
-The thermo-mechanics of crust-mantle systems is described by means of the conservation of mass, momentum and energy equations, expressed as follows:
-\begin{eqnarray}
-\label{eq:mass}\frac{\partial \rho}{\partial t}+\bm{\nabla} \cdot (\rho \bm{u})&=&0\\
-\label{eq:momentum}-\bm{\nabla} p + \bm{\nabla} \cdot \bm{\tau} + \rho \bm{g}&=&0\\
-\label{eq:energy}\rho C_p \left(\frac{\partial T}{\partial t} + \bm{u} \cdot \bm{\nabla} T \right)&=&\bm{\nabla} \cdot \left(k\bm{\nabla} T\right) + H_{tot}
-\end{eqnarray}
-where $\rho$ is the density, $\bm{u}$ is the velocity, \textit{p} is the pressure, $\bm{\tau}$ is the deviatoric stress, $\bm{g}$ is the gravity acceleration,
-$C_p$ is the specific heat at constant pressure, \textit{T} is the temperature, \textit{k} is the thermal conductivity and $H_{tot}$ is the total internal heat
-production. Density variations due to temperature are generally small enough to assume the density as constant ($\rho=\rho_0$) in Eq. \ref{eq:mass} and in
-Eq. \ref{eq:energy}, while it must be treated as a variable in the buoyancy term of Eq. \ref{eq:momentum}, such that \[\rho=\rho_0(1-\alpha(T-T_0))\]
-where $\rho_0$ is the density at a reference temperature $T_0$ and $\alpha$ is the coefficient of thermal expansion. Eq. \ref{eq:mass} then can be rewritten as
-\begin{equation}\label{eq:mass_0}
-\bm{\nabla} \cdot \bm{u}=0
-\end{equation}
-This simplification is known as the extended Boussinesq approximation.
-
-The time step $dt$ is chosen in according to Courant-Friedrichs-Lewy condition \citep{Anderson1995}, such that
-\[dt=C_n\; \min\left(\frac{\Delta x}{\max|\bm{u}|},\frac{(\Delta x)^2}{\kappa}\right)\]
-where $0<C_n<1$ is the Courant (CFL) number, $\Delta x$ is the minimum dimension of smallest element, $\max|\bm{u}|$ is the maximum velocity calculated on the
-entire domain and
-$\kappa=\frac{k}{\rho C_p}$ is the thermal diffusivity \citep{Thieulot2014}.
-
-\section{Lagrangian markers}\label{sec:markers}
-Elemental properties (density, viscosity, thermal conductivity, specific heat and thermal expansion) needed to solve Eqs. \ref{eq:mass_0}, \ref{eq:momentum}
-and \ref{eq:energy} are related to the composition of each element, determined by means of Lagrangian markers that are characteristic of different materials
-in the domain. Elemental properties (with the sole exception of the viscosity) are calculated using an arithmetic mean, as \[P_e=\frac{1}{n}\sum_{i=1}^n P_i\]
-where $P_e$ is the elemental property, $P_i$ is the property characteristic of the material of each marker and $n$ is the number of markers of the element.
-Differently, the average scheme for the elemental viscosity can be chosen between harmonic, geometric and arithmetic mean. Markers can be placed regularly or
-randomly at the beginning of the simulation and their advection is performed by either a 2nd-order or a 4th-order Runge-Kutta in space, interpolating the velocity 
-field on each marker by means of the shape functions (see Section \ref{sec:runge}). The interpolated velocity is then corrected by means of the Conservative
-Velocity Interpolation (CVI), which introduces a corrective term to reduce dispersion and clustering of particles in both steady state and time-dependent 
-\citep{Wang2015} (see Section \ref{sec:CVI}).
-
-The initial distribution of the markers is created using the open-source code library
-Geodynamic World Builder \citep{Fraters2019}. During the simulation, each marker carries memory of temperature, pressure and accumulated strain, which are
-determined interpolating the relative nodal parameter. In particular, pressure is interpolated onto each marker after the smoothing procedure. The number
-of markers contained in each element is maintained between $n_{min}$ and $n_{max}$. When in an element there are less markers than $n_{min}$ the code adds
-random markers to reach the $n_{min}$, while if the number is higher than $n_{max}$ some of them are randomly deleted. When new markers are added, they assume
-the properties of the nearest marker. In this way elements are never empty and maintain a number of markers inside a prefixed range.
-
-\section{The momentum equation}\label{sec:mom_eq}
-The penalty method is implemented so that Eq. \ref{eq:mass_0} can be written as:
-\begin{equation}\label{eq:mass_penalty}
-\bm{\nabla} \cdot \bm{u} + \frac{p}{\lambda}=0
-\end{equation}
-where $\lambda$ is the so-called penalty parameter, which should be 6-7 orders of magnitude larger than the shear viscosity to ensure that mass conservation
-is satisfied \citep{Donea2003,Thieulot2014}.
-
-The deviatoric stress tensor in Eq. \ref{eq:momentum} can be written in terms of the strain rate tensor as $\bm{\tau}=2\eta\dot{\bm{\epsilon}}$, with
-$\dot{\bm{\epsilon}}=\frac{1}{2}\left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)$. Therefore, Eq. \ref{eq:momentum} can be rewritten as:
-\begin{equation}\label{eq:momentum_pressure}
--\bm{\nabla} p + \bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right) + \rho \bm{g} = 0
-\end{equation}
-Finally, using pressure from Eq. \ref{eq:mass_penalty}, Eq. \ref{eq:momentum_pressure} can be rewritten as:
-\begin{equation}\label{eq:momentum_penalty}
-\lambda \bm{\nabla} \left(\bm{\nabla} \cdot \bm{v}\right)+\bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right)+\rho\bm{g}=0
-\end{equation}
-The penalty method is been associated to the iterative Uzawa method, as described in detail in \citet{Dabrowski2008} and \citet{Thieulot2014}
-(see Section \ref{sec:poiseuille}). Results of benchmarks performed to verify the correctness in the implementation of Eq. \ref{eq:momentum_penalty} are
-shown in Sections \ref{sec:ist_sphere} and \ref{sec:rayleigh}.
-
-To support large viscosities variations the penalty parameter is related to the effective elemental viscosity by means of a dimensionless coefficient, so that
-$\lambda=\lambda_e(e)\eta_{eff}(e)$ \citep{Marotta2006,Dabrowski2008,Thieulot2014}. Benchmark of the falling block \citep{Gerya2003a,Gerya2010b,Thieulot2011}
-is performed to verify that the code can correctly deal buoyancy driven flows with strong viscosity contrasts (Section \ref{sec:block}).
-
-\section{Sticky air and free surface}\label{sec:sticky}
-The Earth's surface can be treated by means of either the so-called sticky-air or a true free surface method, both of them implemented in the code. In the
-sticky-air method the surface is approximated with the introduction of a buoyant layer with a viscosity at least four orders of magnitude lower than the crust
-\citep{Schmeling2008,Crameri2012} and the interface between lithosphere and air is defined using a chain of passive markers that are advected as the Lagrangian
-markers. The correctness of the evolution of the markers chain is tested with the experiment of a 2D time-dependent Stokes sphere below a free surface and
-compared with results from ASPECT\footnote{\url{https://aspect.geodynamics.org/}} \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} (Section
-\ref{sec:sphere}). In the true free surface case the top boundary is assumed stress-free and velocities are not fixed. In this case topography variations are
-described by vertical deformations of the mesh that depend on the velocity field of the nodes that identify the free surface, while horizontal deformations are
-not taken into account. This procedure is known as the Arbitrary Lagrangian-Eulerian (ALE) method and its implementation follows the technique described in
-\citet{Thieulot2011}. However, although the implementation of a true free surface better reproduces laboratory experiment, extremely small time steps can be
-necessary to maintain stability \citep{Kaus2010a,Quinquis2011,Thieulot2014}. Therefore, the stability algorithm proposed by \citet{Kaus2010a} is implemented
-to avoid instabilities due to high density differences at the free surface when using too large time steps. The implementation of this algorithm are tested by
-performing the experiment described by \citet{Kaus2010a} (Section \ref{sec:stab}).
-
-The topography relaxation benchmark proposed by \citet{Crameri2012} is performed to verify that the code correctly recovers topography variations in case of
-both the sticky-air and the true free surface method (Section \ref{sec:crameri}). Finally, Section \ref{sec:subduction} show the results of the spontaneous
-subduction experiment described by \citet{Schmeling2008}.
-
-\section{Erosion and sedimentation}\label{sec:erosion}
-Surface processes at either the lithosphere-air interface or in correspondence of the free surface have been implemented by means of the software
-Fastscape\footnote{\url{https://fastscape.org/}} \citep{Braun2013,Cordonnier2019,Yuan2019a,Yuan2019b}. FastScapeLib is a set of routines that solve the stream
-power law (enriched by a sediment transport/deposition term) \citep{Yuan2019a}, hillslope diffusion and marine transport and deposition \citep{Yuan2019b}.
-The partial differential equation solved by FastScapeLib is:
-\begin{equation}
-\frac{\partial h}{ \partial t}=U-K_f A^{m_e}S^{n_e} + \frac{G}{A}\int_A \left( U-\frac{\partial h}{\partial t} \right) dA + K_d \bm{\nabla}^2 h 
-\end{equation}
-where $h$ is the topography, $U$ is the uplift, $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $S$ is the slope, $m_e$ and $n_e$ are
-parameters that should be chosen to constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}, $G$ is the dimensionless
-deposition/transport coefficient for the enriched stream power law and $K_d$ is the effective diffusivity. Terms $K_f A^{m_e}S^{n_e}$ and $K_d \bm{\nabla}^2 h$
-consider long-range processes, described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, and short-range processes,
-proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}.
-
-In case of sticky air, slope is determined by means of the markers chain, which is vertically corrected considering erosion and sedimentation rates. After the
-correction, continental markers above the markers chain are transformed in air markers and air markers below the markers chain become sediments. Similarly,
-slope is determined considering the nodes on the top boundary in case of true free surface and they are vertically corrected as for the markers chain. In this
-case, continental markers above the top boundary are simply deleted.
-%Surface processes at the lithosphere-air interface are implemented taking into account fluvial erosion (long-range processes), hillslope processes (short-range
-%processes) and failure of slopes exceeding a threshold angle $\theta$ \citep{Burov1997,Tucker1999,Sacek2012,Croissant2014,Ueda2015,Beucher2020}. Erosion rates
-%$\dot{E}_{lr}$ due to long-range processes can be described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{lr}=K_f A^{m_e} \bm{\nabla} h^{n_e}
-%\end{equation}
-%where $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $\bm{\nabla} h$ is the slope and $m_e$ and $n_e$ are parameters chosen in to
-%constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}. Erosion rates $\dot{E}_{sr}$ due to short-range processes are
-%implemented by assuming that transport is linearly proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{sr}=K_d \bm{\nabla}^2 h
-%\end{equation}
-%where $K_d$ is the effective diffusivity. In addition, the slope is modified if it still exceeds a threshold angle after long- and short-range processes
-%\citep{Ueda2015}.
-
-%Offshore sedimentation processes are implemented similarly to hillslope processes but with a depth-dependent coefficient \citep{Sacek2012,Beucher2020}.
-%The sedimentation rate $\dot{S}$ is calculated as
-%\begin{equation}
-%\dot{S}=K_s \bm{\nabla}^2 h
-%\end{equation}
-%with
-%\begin{equation}
-%K_s=K^*_s \exp\left(\frac{h-h_{sl}}{h_0}\right)
-%\end{equation}
-%where $K^*_s$ is a constant, $h_{sl}$ is the sea level and $h_0$ is a characteristic length. In case of sticky air, slope is determined by means of the markers
-%chain, which is vertically corrected considering erosion and sedimentation rates. After the correction, continental markers above the markers chain are
-%transformed in air markers and air markers below the markers chain become sediments. Similarly, slope is determined considering the nodes on the top boundary
-%in case of true free surface and they are vertically corrected as for the markers chain. In this case, continental markers above the top boundary are simply
-%deleted.
-
-\section{Non-linear rheologies}\label{sec:nonlinear}
-Non-linear rheologies are implemented combining viscous creep (dislocation and diffusion) and plastic yielding. For each marker, diffusion and dislocation
-viscosity ($\eta_{df|ds}$) can be determined as follows \citep{Karato1993,Warren2008b,Wang2016}:
-\begin{equation}\label{eq:viscous}
-\eta_{df|ds}=f_s \left(\frac{d^m}{A}\right)^{\frac{1}{n}} \left (\frac{2^{\frac{1-n}{n}}}{3^{\frac{n+1}{2n}}} \right)
-I_2^{\frac{(1-n)}{n}}\exp\left(\frac{Q+pV}{nRT}\right)
-\end{equation}
-where $f_s$ is a scaling factor used to represent lithologies that are stronger or weaker than the base set, $d$ is the grain size, $m$ is the grain size exponent,
-$n$ is the stress exponent, $A$ is the uniaxial pre-exponential factor, $I_2$ is the square root of the second invariant of the strain rate tensor, $Q$ is the
-activation energy, $p$ is the pressure, $V$ is the activation volume, $R$ is the gas constant and $T$ is the temperature.
-Pressures and temperatures are determined by the interpolation of the nodal parameters, while strain rates are calculated by means of the
-derivative of the velocities on the nodes. In case of diffusion creep $n=1$ and $m>0$, in case of dislocation creep $n>1$ and $m=0$. The viscous creep $\eta_{cp}$
-is then calculated as the harmonic average between $\eta_{df}$ and $\eta_{ds}$:
-\begin{equation}\label{eq:visco_creep}
-\eta_{cp}=\left(\frac{1}{\eta_{df}}+\frac{1}{\eta_{ds}}\right)^{-1}
-\end{equation}
-The implementation of the non-linear viscous creep viscosity depending on the strain rate is tested by means of the slab detachment benchmark
-(Section \ref{sec:slab}).
-
-Plastic yielding is implemented rescaling $\eta_{cp}$ in order to limit the stress below the yield stress $\sigma_y$ \citep{Thieulot2008,Thieulot2014,Glerum2018},
-obtaining
-\begin{equation}\label{eq:plastic}
-\eta_{pl}=\frac{\sigma_y}{2I_2}
-\end{equation}
-where $\eta_{pl}$ is the plastic viscosity and the yield stress is determined following the Drucker-Prager criterion, such as
-\begin{equation}\label{eq:yield}
-\sigma_y=C\;\cos(\phi)+p\;\sin(\phi)
-\end{equation}
-where $C$ is the cohesion and $\phi$ is the internal friction angle. In case of negative pressure it is imposed equal to 0, so that negative yield stress are
-excluded. The correctness of the non-linear solution in case of plastic viscosity with variable internal friction angle is verified performing indenter and
-brick experiments (Sections \ref{sec:indenter} and \ref{sec:brick}, respectively).
-
-Strain softening is taken into account for both viscous creep and plastic viscosity \citep{Huismans2003,Babeyko2005,Huismans2005,Sobolev2005,Warren2008a} by
-means of the accumulated strain $\epsilon$ memorised by each marker. Softening in the viscous creep determines a linear decrease of $\eta_{vc}$ by means of a
-viscous strain softening factor $W_S$ that increases linearly from $W_{S_0}$ to $W_{S_{\infty}}$ for $\epsilon_{S_0}<\epsilon_S<\epsilon_{S_{\infty}}$ 
-\citep{Warren2008a}. This viscous softening can be related to strain-induced grain size reduction \citep{Warren2008a}. Differently, plastic softening is 
-simulated with a linear decrease of internal friction angle $\phi(\epsilon)$ and cohesion $C(\epsilon)$ in according to
-\begin{eqnarray}
-\label{eq:friction1}\phi(\epsilon)&=&\phi_0+(\phi_{\infty}-\phi_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}\\
-\label{eq:friction2}C(\epsilon)&=&C_0+(C_{\infty}-C_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}
-\end{eqnarray}
-where $\phi_0$, $C_0$ and $\phi_{\infty}$, $C_{\infty}$ are internal friction angle and cohesion for $\epsilon_0$ and $\epsilon_{\infty}$, respectively
-\citep{Huismans2003,Huismans2005,Warren2008a,Thieulot2014}. Plastic softening approximates deformation-induced softening of faults and brittle shear zones
-\citep{Warren2008a}.
-
-Plastic yielding and viscous creep are then combined to obtain a viscoplastic viscosity $\eta_{vp}$ as follows, assuming that they are independent processes
-\citep{Karato2008,Glerum2018}:
-\begin{equation}\label{eq:viscoplastic}
-\eta_{vp}=\min(\eta_{cp},\eta_{pl})
-\end{equation}
-Finally, effective viscosity $\eta_{eff}$ is capped by the minimum and the maximum viscosity ($\eta_{min}$ and $\eta_{max}$, respectively) to avoid extremely
-low or high viscosity \citep{Glerum2018} as follows
-\begin{equation}\label{eq:effective}
-\eta_{eff}=\min(\max(\eta_{vp},\eta_{min}),\eta_{max})
-\end{equation}
-Viscous creep and plastic yielding are non-linear rheologies because of their dependence on the velocity field through pressure and strain rates. Therefore,
-the solution is determined by means of Picard-type iterations, until convergence of the velocity field \citep{Glerum2018}. The convergence is verified at each
-iteration $i$ via the nonlinear residual $\bm{\mathcal{R}}^i$ that can be determined as
-\begin{equation}\label{eq:convergence}
-\bm{\mathcal{R}}^i= \mathbb{K} (\eta_{eff}(\dot{\bm{\epsilon}}^{i-1},p^{i-1}))\cdot \bm{v}^{i-1} - \bm{f}^i
-\end{equation}
-%\begin{equation}\label{eq:convergence}
-%\chi=1-\frac{\langle (\bm{v}^i-\langle\bm{v}^i\rangle)\cdot (\bm{v}^{i+1}-\langle\bm{v}^{i+1}\rangle) \rangle}{\sqrt{\sigma^i\sigma^{i+1}}}
-%\end{equation}
-%Iterations are performed until either convergence ($\chi<tol$) or maximal number of non-linear iterations ($it_{max}$) is reached.
-where $\mathbb{K}$ is the Stiffness matrix and $\bm{f}$ is the right hand side vector \citep{Spiegelman2016,Glerum2018}.  The $L_2$-norm is extracted from
-$\bm{\mathcal{R}}^i$ and it is normalised as follows
-\begin{equation}\label{eq:convergence_res}
-\frac{||\bm{\mathcal{R}}^i||_2}{||\bm{\mathcal{R}}^0||_2}
-\end{equation}
-where $||\bm{\mathcal{R}}^0||_2$ is the $L_2$-norm of the first nonlinear iteration.
-Since often the normalised nonlinear residual drop very quickly over the first 2-3 iterations, also the $L_2$-norm of velocity and pressure residuals are
-calculated as
-\begin{equation}\label{eq:convergence_vel}
-\frac{||\Delta\bm{v}||_2}{||\bm{v}||_2}=\frac{||\bm{v}_i-\bm{v}_{i-1}||_2}{||\bm{v}_{i}||_2}
-\end{equation}
-and
-\begin{equation}\label{eq:convergence_press}
-  \frac{||\Delta p||_2}{||p||_2}=\frac{||p_i-p_{i-1}||_2}{||p_{i}||_2}
-\end{equation}
-respectively. Iterations are performed until either maximal number of non-linear iterations ($it_{max}$) is reached or all the normalised nonlinear residual, 
-the velocity residual and the pressure residual converge under a defined tolerance $tol$.
-
-\section{The energy equation}\label{sec:en_eq}
-The stabilisation of the advection term of the energy equation (Eq. \ref{eq:energy}) needed to avoid possible oscillations in the thermal solution in the case
-when advection dominates over diffusion, is implemented by means of a streamline-upwind Petrov–Galerkin (SUPG) method, for which the advection term is modified
-follows the discussion in \citet{Thieulot2011} and \citet{Thieulot2014} (see Section \ref{sec:advection} for benchmark). The total internal heat production
-$H_{tot}$ in Eq. \ref{eq:energy} includes radiogenic heating $H_r$, shear heating $H_s$ and adiabatic heating $H_a$, where
-\begin{equation}\label{eq:shear}
-H_s=2\eta \dot{\bm{\epsilon}} : \dot{\bm{\epsilon}} = 2\eta(\dot{\epsilon}_{xx}^2+\dot{\epsilon}_{yy}^2+2\dot{\epsilon}_{xy}^2)
-\end{equation}
-and
-\begin{equation}\label{eq:adiabatic}
-H_a=T \alpha \frac{Dp}{Dt} \approx -\alpha T \rho g_y v
-\end{equation}
-The correctness of the implementation and the computation of shear and adiabatic heating is verified performing an exercise with an analytical solution
-(Section \ref{sec:simple_shear}) and Exercise 9.4 in \citet{Gerya2010b} (Section \ref{sec:shear}). Furthermore, mantle convection and thin layer entrainment
-benchmarks (Sections \ref{sec:mantle} and \ref{sec:thin}, respectively) are performed to verify that the code solve correctly Eqs. \ref{eq:momentum_penalty}
-and \ref{eq:energy} in case of a temperature field characterised by an initial perturbation and with a temperature-dependent density. At last, the viscoplastic
-mantle convection benchmark is performed to check also the viscoplasticity in case of a initial perturbed temperature field (Section \ref{sec:visco_mantle}).
-
-\section{Phase transitions and hydration}\label{sec:phase}
-Variations of effective density, specific heat and coefficient of thermal expansion during the evolution of crustal and mantle materials at different
-pressure-temperature ($p$-$T$) conditions are computed by means of version 6.8.6 of Perple\_X software package \citep{Connolly2005}, similarly to the
-implementation described in \citet{Marotta2020}. By default, mineral assemblages and properties of each lithology (oceanic crust, lower and upper continental
-crust, sediments and dry or hydrated mantle) are calculated for temperatures from \SIrange{330}{1600}{\kelvin}, with increments of approximately \SI{4}{\kelvin},
-and pressures from \SIrange{0.1}{30}{\giga\pascal}, with increments of approximately \SI{0.03}{\giga\pascal}, for a total of almost 100000 points.
-
-The introduction of phase transitions produces two main effects that must be taken into account in numerical models: variations in density and release/absorption
-of latent heat \citep{Tackley2010}. Effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into
-account considering effective coefficient of thermal expansion \citep{Christensen1985,Zhong2015}. Similarly, effects of latent heat in the energy equation
-(eq \ref{eq:energy}) are taken into account by considering effective specific heat and coefficient of thermal expansion
-\citep{Christensen1985,Tackley2010,Zhong2015}. Effective density, specific heat and coefficient of thermal expansion are included in Perple\_X files and are
-assumed by each Lagrangian marker in according to its $p$-$T$ conditions.
-
-Hydration processes related to dehydration of subducting lithosphere \citep{Schmidt1998,Liu2007,Faccenda2009,Faccenda2010,Faccenda2014,Rosas2016} strongly
-influence the thermo-mechanics inside the mantle wedge, mainly because of both weakening effects on the mantle rheology and density variations in case of
-mantle serpentinization \citep{Gerya2002,Honda2003,Arcay2005,Roda2010,Regorda2017}. In the case that hydration is switched on, the amount of bound and free
-water is memorised by each marker, following the implementation of \citet{Quinquis2014}; hydration and dehydration processes are related to the amount of bound
-water of each marker and to the maximum amount of water it can transport, i.e. if the amount of bound water exceeds the maximum amount of water, the marker
-dehydrates and releases free water that can hydrates under-saturated markers. Maximum water content of each marker is determined as function of lithology and
-$p$-$T$ conditions and it is calculated using Perple\_X, in the same way than effective density, specific heat and coefficient of thermal expansion. Bound water
-is advected together with the markers, neglecting the effect of bound water diffusion, while free water simply migrates vertically and is not coupled to the
-solid-phase flow of the mantle wedge \citep{Arcay2005,Quinquis2014}. The correctness of the migration scheme of the free water and its absorption into bound
-water by under-saturated markers are verified performing the experiment described in \citet{Quinquis2014} (Sec. \ref{sec:quinquis}).
-
-The mantle viscosity weakening related to the amount of water has been extensively studied \citep{Chopra1981,Hirth2003} and it is implemented in according to
-\citet{Arcay2005} and \citet{Horiuchi2016}, as follows
-\begin{equation}\label{eq:wet_visc}
-\eta_{wet}=\eta_{dry}\left(\left[1-\frac{1}{f_v}\right]\exp\left(-\frac{[OH^-]}{[OH^-]_0}\right)+\frac{1}{f_v}\right)
-\end{equation}
-where $f_v$ is the viscosity reduction factor between dry and wet conditions, $[OH^-]$ is the water content and $[OH^-]_0$ is a reference water content set to
-620 ppm (0.062 wt.\%) \citep{Arcay2005}. By default, $f_v$ is set to 100, in according to \citet{Arcay2005} and \citet{Horiuchi2016}.
-
-\section{Mantle melting}\label{sec:melt}
-Melt fraction in the mantle depends on temperature ($T$, in \textdegree C), pressure ($p$, in \si{\giga\pascal}) and the water content in the melt ($X_{H_2O}$, in wt.\%).
-The determination of the melt fraction $M$ is implemented as explained by \citet{Katz2003}, according to the successive modification by \citet{Langmuir2006} and
-\citet{Kelley2010}, as follows
-\begin{eqnarray}\label{eq:melt_fr}
-M(p,T,X_{H_2O})&=&-T+T_{sol}+(x_t\cdot ln(p+y_t))\cdot X_{H_2O}-\nonumber\\
-&&+K_m\left(\frac{X_{H_2O}^{bulk}}{D_{H_2O}(1-X_{H_2O})+X_{H_2O}}\right)^{\gamma}
-\end{eqnarray}
-where $T_{sol}$ is the temperature of the dry solidus, $D_{H_2O}$ is the partition coefficient and $x_t\cdot log(p+y_t)$ indicates the pressure dependence of
-the melting curve \citep{Kelley2010}. $x_t$, $y_t$, $K_m$ and $\gamma$ are parameters chosen in according to \citet{Kelley2010}.
-However, $X_{H_2O}$ depends on the melt fraction as
-\begin{equation}\label{eq:melt_wt}
-X_{H_2O}(M)=\frac{X_{H_2O}^{bulk}}{D_{H_2O}+M(1-D_{H_2O})}
-\end{equation}
-and a numerical solution can be found using a root-finder method \citep{Katz2003,Wang2016}. Furthermore, the content of water in the melt is limited by the
-pressure-dependent saturation concentration of water in the melt, determined as
-\begin{equation}\label{eq:melt_max}
-X_{H_2O}^{sat}=\chi_1 p^{\lambda_m}+\chi_2 p
-\end{equation}
-where $\chi_1$, $\chi_2$ and $\lambda_m$ are parameters chosen according to \citet{Katz2003}, \citet{Langmuir2006} and \citet{Kelley2010}. The correctness in
-the determination of melt fractions and water contents in the melt are tested comparing model results with experimental curves obtained by \citet{Katz2003}
-(Sec. \ref{sec:katz}).
-
-In case of melting, viscous creep viscosity ($\eta_{vp}$) is modified according to \citet{Wang2016}, as
-\begin{equation}\label{eq:melt_visc}
-\eta_{melt}=\eta_{vp}\cdot \exp(\alpha_m M)
-\end{equation}
-where $\alpha_m$ are melt fraction factors for dislocation and diffusion creep, chosen in according to \citet{Wang2016}.
-
-Effective density ($\rho_{eff}$) for partially molten rocks is calculated as
-\begin{equation}\label{eq:melt_dens}
-\rho_{eff}=\rho_s(1-M)+\rho_m M
-\end{equation}
-where $\rho_s$ and $\rho_m$ are the densities of the solid and the molten rock, respectively \citep{Gerya2004b,Gerya2007,Wang2016}. The density of the
-solid rock is extracted by Perple\_X, while the density of the molten portion is calculated as
-\begin{equation}\label{eq:melt_dens_M}
-\rho_m(p,T)=\rho_0[1-\alpha(T-T_0)][1+\beta(p-p_0)]
-\end{equation}
-where $\rho_0$ is the density at the reference temperature ($T_0$) and pressure($p_0$), and $\alpha$ and $\beta$ are the thermal expansion and compressibility
-coefficients, respectively \citep{Gerya2004b,Gerya2007,Wang2016}.
-
-As for the phase transitions, effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into account
-considering effective coefficient of thermal expansion. Similarly, latent heat due to melting/crystallisation can be implicitly included in the energy equation
-(Eq. \ref{eq:energy}) by considering effective specific heat ($C_{p_{eff}}$) and thermal expansion ($\alpha_{eff}$) of the partially molten rocks, in according to
-\begin{eqnarray}
-\label{eq:cp_effective}C_{p_{eff}}&=&C_p+H_L\left(\frac{\partial M}{\partial T}\right)_p \\
-\label{eq:alpha_effective}\alpha_{eff}&=&\alpha+\rho\frac{H_L}{T}\left(\frac{\partial M}{\partial p}\right)_T
-\end{eqnarray}
-where $C_p$ and $\alpha$ are the specific heat and the coefficient of thermal expansion of the solid rock, respectively, and $H_L$ is the latent heat of the
-molten rock \citep{Gerya2004b,Gerya2007,Tackley2010}.
-
-\section{Benchmarks}\label{sec:benchmark}
-\subsection{Error measurements}\label{sec:error}
-In order to determine the accuracy of velocity and pressure field of the benchmarks, the $L_2$-norm is computed by numerical integration on the Gauss-Legendre
-quadrature points. $L_2$-norm for pressure and velocity errors can be evaluated as
-\begin{eqnarray}
-\textrm{err}_p&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|p^n_i(r_q)-p^a_i(r_q)|^2\right)w_q|J_q|}\nonumber \\
-\textrm{err}_v&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|u^n_i(r_q)-u^a_i(r_q)|^2+|v^n_i(r_q)-v^a_i(r_q)|^2\right)w_q|J_q|}\nonumber
-\end{eqnarray}
-respectively, where $n_e$ is the number of elements, $n_q$ is the number of quadrature points per element, $p^n_i(r_q)$ and $p^a_i(r_q)$ are the numerical and
-analytical pressure, respectively, in each quadrature point $q$, $w_q$  and $J_q$ are the weight and the Jacobian at the quadrature point
-$q$, $u^n_i(r_q)$, $v^n_i(r_q)$, $u^a_i(r_q)$ and $v^a_i(r_q)$ are the numerical and analytical velocities, respectively, in each quadrature point $q$.
-
-Other quantities used as comparison with original benchmarks are the root-mean-square velocity over the whole domain and over the surface
-\begin{eqnarray}
-v_{\textrm{rms}}&=&\sqrt{\int_{\Omega}|\bm{v}|^2 d\Omega}\nonumber \\
-v_{\textrm{rms}}^{top}&=&\sqrt{\int_0^1 u^2 \bigg|_{y=1} dx}\nonumber
-\end{eqnarray}
-respectively, top and bottom Nusselt numbers
-\[\textrm{Nu}^{top/bottom}=-\int_0^1\frac{\partial T}{\partial y} \bigg|_{y=1/y=0} dx\]
-the average rate of work done against gravity
-\[\left\langle W \right\rangle=\int_{\Omega}T u_y d\Omega\]
-the average rate of viscous dissipation
-\[\left\langle \Phi \right\rangle=\int_{\Omega}\tau_{ij}\dot{\epsilon}_{ij} d\Omega\]
-and the percentage error between $\left\langle W \right\rangle$ and $\left\langle \Phi \right\rangle/Ra$
-\[\delta=\frac{\left|\left\langle W \right\rangle -\frac{\left\langle \Phi \right\rangle}{Ra} \right|}{\max \left(\left\langle W \right\rangle,
-\frac{\left\langle \Phi \right\rangle}{Ra}\right)}\times100\]
-
-\subsection{Stokes flow}\label{sec:stokes}
-The problem consists of determining velocity field $(u,v)$ and pressure $p$ in case of a manufactured solution with prescribed body forces such as
-\begin{eqnarray}
-b_1&=&(12-24y)x^4+(-24+48y)x^3+(-48y+72y^2-48y^3+12)x^2+\nonumber \\
-&&+(-2+24y-72y^2+48y^3)x+1-4y+12y^2-8y^3\nonumber \\
-b_2&=&(8-48y+48y^2)x^3+(-12+72y-72y^2)x^2+(4-24y+48y^2-48y^3+\nonumber \\
-&&+24y^4)x-12y^2+24y^3-12y^4\nonumber
-\end{eqnarray}
-for which the exact solution is:
-\begin{eqnarray}
-u(x,y)&=&x^2(1-x)^2(2y-6y^2+4y^3)\nonumber \\
-v(x,y)&=&-y^2(1-y)^2(2x-6x^2+4x^3)\nonumber \\
-p(x,y)&=&x(1-x)-\frac{1}{6}\nonumber
-\end{eqnarray}
-The domain is a unit square a constant viscosity ($\eta=1$) and the penalty parameter is set to $\lambda=10^7$. Velocity boundary conditions are set to no slip
-$(\bm{v}=\bm{0})$ on all boundaries. The problem is performed for different grid resolution between \num{8x8} and \num{1024x1024} elements.
-
-The errors between the analytical solution and the numerical prediction of the velocity and the pressure field are calculated as in Section \ref{sec:error}.
-Fig. \ref{fig:errors}a shows that both velocity and pressure field converge to the exact solution with the decrease of the element size, following the
-theoretical convergence rate. The convergence of pressure, in contrast with the observation by \citet{Donea2003} for $Q_1 \times P_0$ elements, indicate the
-effectiveness of the smoothing procedure. Smoothed pressure field for grid resolution of \num{128x128} elements is shown in Fig. \ref{fig:errors}b (continuous
-red line), in comparison with the analytical solution (dashed blue line). Solve times and memory usage needed to generate the solution are shown in
-Fig. \ref{fig:MUMPS}. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Solver/Stokes_Flow}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/errors.pdf}
-\caption{Velocity and pressure error for the Stokes flow experiment between generated and analytical solution as a function of element size (panel a),
-and comparison between smoothed pressure and analytical solution as function of $x$ coordinate for a grid resolution of \num{128x128} elements (panel b).}
-\label{fig:errors}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/MUMPS.pdf}
-\caption{Panel a) Analysis, factorisation, solution and total solve times (panel a) and factorisation memory usage (panel b) as a function of the total number
-of degrees of freedom for the Stokes flow experiment.}
-\label{fig:MUMPS}
-\end{figure}
-
-\subsection{Markers advection}\label{sec:runge}
-The 2nd-order Runge-Kutta advection scheme in space is tested by means of the Zalesak disk test \citep{Zalesak1979,Thieulot2014}. The benchmark is performed
-in a unit square domain with a grid resolution of \num{32x32} elements and 1000000 markers and values of Courant number between 0.25 and 1.
-At $t=0$, the disk is centred at position $(0.5;0.75)$ with a radius $R=0.15$ and has a vertical fissure 0.05 wide and 0.2 high (orange in Fig. \ref{fig:runge})
-and with the velocity field prescribed in the entire domain as
-\begin{eqnarray}
-u(x,y)&=&2\pi\left(y-\frac{L_x}{2}\right)\nonumber \\
-v(x,y)&=&-2\pi\left(x-\frac{L_x}{2}\right)\nonumber
-\end{eqnarray}
-the markers are back to their initial location after a $2\pi$ rotation ($t=1$). The clockwise rotation of the disk at $t=0.25$, 0.5 and 0.75 
-is shown in Fig. \ref{fig:runge} (red, blue and green, respectively). The distance from the centre of 1 marker is calculated throughout the rotation to
-evaluate the error for different values of Courant number. As expected, the error increase with the increasing of the time step, as shown in 
-Fig. \ref{fig:runge_err}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/Zalesak_Disk}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=200px]{./Figures/Zalesak.pdf}
-\caption{Position of the Zalesak disk at $t=0$, 0.25, 0.5 and 0.75 (orange, red, blue and green, respectively) for the case with Courant number of 1.}
-\label{fig:runge}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/zal_marker.pdf}
-\caption{Distance from the centre as function of time for values of Courant number of  of 0.25, 0.3, 0.5, 0.75 and 1 (green, orange, blue, black and red,
-respectively). R indicates the distance from the centre considering a perfect circle trajectory.}
-\label{fig:runge_err}
-\end{figure}
-
 \subsection{Conservative Velocity Interpolation (CVI)}\label{sec:CVI}
@@
 \end{figure}
-
-\subsection{Poiseuille flow}\label{sec:poiseuille}
-The domain is a rectangle with $L_x=2$ and $L_y=1$ and constant density and viscosity ($\rho=1$ and $\eta=1$), gravity acceleration $\bm{g}=\bm{0}$ and penalty
-parameter $\lambda=10^8$. The grid is composed by \num{40x20} elements. Velocity boundary conditions are set to no slip $(\bm{v}=\bm{0})$ at the top and the
-bottom, and a parabolic profile is imposed on the sides, with $u=y(1-y)$ and $v=0$. The analytical solution is then given by:
-\begin{eqnarray}
-u(x,y)&=&y(1-y)\nonumber \\
-v(x,y,)&=&0\nonumber \\
-p(x,y)&=&2\eta\left(\frac{L_x}{2}-x\right)\nonumber
-\end{eqnarray}
-
-The velocity field predicted by the model follows the expected parabolic profile (Fig. \ref{fig:poiseuille}a and Fig. \ref{fig:poi_plot}a) and, in case of the
-classic penalty method with no iterations, the pressure is clearly related to the divergence by means of the penalty parameter (Fig. \ref{fig:poiseuille}b and
-c and Fig. \ref{fig:poi_plot}b). Fig. \ref{fig:poiseuille}d shows that one Uzawa iteration is sufficient to bring the divergence down to \num{e-15}, with no
-correlation with the pressure field. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Poiseuille%20Flow}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Poiseuille.pdf}
-\caption{Velocity field (panel a), pressure (panel b) and divergence velocity of a Poiseuille flow in case of the classic penalty method (no iterations) and
-after one Uzawa iteration (panel c and d, respectively).}
-\label{fig:poiseuille}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Analytical.pdf}
-\caption{Velocity (panel a) and pressure (panel b) field predicted by the model for a Poiseuille flow with respect with their analytical solutions. The
-velocity field is plotted as function of the vertical coordinate in $L_x/2$ and the pressure is plotted as function of the x coordinate in $L_y/2$.}
-\label{fig:poi_plot}
-\end{figure}
-
-\subsection{Instantaneous 2D sphere}\label{sec:ist_sphere}
-The domain is a unit square with gravity $\bm{g}=(0,-1)$. The fluid has constant density and viscosity ($\rho_f=1$ and $\eta_f=1$). The sphere is in the middle
-of the domain with a radius R = 0.123456798 and has constant density and viscosity ($\rho_s=10^{-2}+\rho_f$ and $\eta_s=10^3 \cdot \eta_f$).
-We distinguish three different types of velocity boundary conditions:
-\begin{enumerate}
-\item FS: free slip conditions on all sides.
-\item NS: no slip conditions on all sides.
-\item OT: free slip conditions on the sides and the bottom, and open boundary at the top.
-\end{enumerate}
-All three cases are performed for different grid resolutions, between \num{16x16} and \num{512x512} elements with 50 randomly distributed markers per element,
-and different average schemes for the viscosity (harmonic, geometric and arithmetic). The velocity in the centre of the sphere $\bm{v}(0.5;0.5)$), minimum and
-maximum velocities ($u_{min}$, $u_{max}$, $v_{min}$, $v_{max}$) and pressures ($p_{min}$, $p_{max}$) on the entire domain, $v_{\textrm{rms}}$ and average pressure on the
-entire domain ($p_{avg}$) are compared with solutions generated by ASPECT \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and all results
-can be found at \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere2D} and 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Instantaneous_2D_Sphere}. Results in terms of the $v_{\textrm{rms}}$ for NS model 
-are also shown in Fig. \ref{fig:inst_sphere}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_NS.pdf}
-\caption{$v_{\textrm{rms}}$ as function of element size for the instantaneous 2D sphere experiment in case of no slip boundary conditions and with different average
-schemes for the viscosity. Results are compared with results obtained by other numerical codes.}
-\label{fig:inst_sphere}
-\end{figure}
-
-\subsection{Rayleigh-Taylor instability}\label{sec:rayleigh}
-This problem has been originally presented by \citet{vanKeken1997} and here it is performed as the isoviscous case (Case 1a). The domain has $L_x=0.9142$ and
-$L_y=1$ and gravity $\bm{g}=(0,-1)$. Two fluids with same constant viscosities ($\eta_1=\eta_2=1$) and different densities ($\rho_1=1000$ and $\rho_2=1010$),
-with the lighter fluid at the bottom. The initial interface between the fluids is given by $y(x)=0.2+0.02 \cos \left(\frac{\pi x}{L_x}\right)$. The experiment
-is performed with different grid sizes (\num{50x50}, \num{80x80}, \num{100x100} and \num{256x256}). A total of 1960000 markers are randomly distributed at the
-beginning of the simulation. Velocity boundary conditions are set to no slip at the top and the bottom, and to free slip at the sides of the domain.
-
-$v_{\textrm{rms}}$ as function of time is reported in Fig. \ref{fig:RT}, matching well with results shown by \citet{vanKeken1997}, \citet{Tackley2003} and
-\citet{Thieulot2014}. Fig. \ref{fig:rayleigh} shows the evolution of the experiment at different time steps. All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Rayleigh_Taylor_experiment/ISOVISCOUS}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/RT.pdf}
-\caption{$v_{\textrm{rms}}$ of the Rayleigh-Taylor experiment as function of time for different resolution of the grid.}
-\label{fig:RT}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=300px]{./Figures/Rayleigh.pdf}
-\caption{Evolution of the Rayleigh-Taylor experiment for a grid of \num{256x256} elements at $t=0, 500, 1000$ and $2000$ (panels a, b, c and d, respectively).}
-\label{fig:rayleigh}
-\end{figure}
-
-\subsection{Falling blocks}\label{sec:block}
-This benchmark is described as proposed by \citet{Gerya2003a}, \citet{Gerya2010b} and \citet{Thieulot2011}. The domain is square with $L_x=L_y=$ \SI{500}{\km}
-and the grid is composed by \num{50x50} elements with 25 markers in each element. The block is initially centred at ($x=$ \SI{250}{\km}; $y=$ \SI{400}{\km})
-and has a size of \num{100x100} km (Fig. \ref{fig:block}a). The fluid surrounding the block has $\eta_f=$ \SI{e21}{\pascal\per\s} and $\rho_f=$
-\SI{3200}{\kg\per\cubic\m}. The benchmark tests with different viscosities of the block, with $\eta_b$ from \SIrange{e15}{e27}{\pascal\per\s}. In each
-experiment also the density $\rho_b$ of the block varies from \SIrange{3220}{9900}{\kg\per\cubic\m}. Velocity boundary conditions are set to free slip
-conditions on all sides of the domain. Density distributions at $t=$ \SI{20}{\mega\year} for $\rho_b=$ \SI{3300}{\kg\per\cubic\m} and different $\eta_b$
-are plotted in Fig. \ref{fig:block}b-f, showing that the code correctly preserve the block geometry in case of large viscosity contrast, with a block stiffer
-than the surrounding fluid (Fig. \ref{fig:block}f).
-
-The velocity in the centre of the falling block at $t=0$ is measured for all experiments and, since the velocity should increase with density contrast, the
-quantity $\bm{v}/(\rho_b-\rho_f)$ is plotted as function of the viscosity contrast. All results perfectly match with those from \citet{Gerya2010b} and they
-line up on a single curve, demonstrating that the code can correctly deal with large viscosity and density contrasts (Fig. \ref{fig:falling}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Falling%20blocks}.
-
-\begin{figure}
-\noindent\includegraphics[width=300px]{./Figures/Block.pdf}
-\centering
-\caption{Density evolution of the falling block experiment at $t=$ \SI{0}{\mega\year} (panel a) and $t=$ \SI{20}{\mega\year} for different viscosities of the
-block (panels b-f).}
-\label{fig:block}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Falling.pdf}
-\caption{Initial velocity relative to the density contrast at the centre of the falling block as function of the viscosity contrast between the block and the
-surrounding fluid.}
-\label{fig:falling}
-\end{figure}
-
-\subsection{2D time-dependent Stokes sphere with deformable free surface}\label{sec:sphere}
-This experiment is performed in a unit square domain with the gravity acceleration fixed to $g_y=-1$. The sphere has $\rho_s=2$ and $\eta_s=10^3$ and is
-initially centred at ($0.5;0.6$) with radius $R=0.123456789$. The fluid surrounding the sphere has $\rho_f=1$ and $\eta_f=1$ and occupies the domain for
-$y\leq 0.75$, while for $y>0.75$ the air has $\rho_a=0$ and $\eta_a=10^{-3}$. Velocity boundary conditions are set to free slip on all sides. The Courant
-number is set to 0.25. The experiments run for 200 s using grid resolutions from \num{150x150} to \num{512x512} elements, with 25 randomly distributed markers
-per element, and for different average schemes for the viscosity. The interface between the fluid and the air is tracked by means of the markers chain. Results
-in terms of velocity and pressure fields and topography variations are compared with results obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and they can be found at
-\url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere_fs2D} and
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Time_dependent_sphere}. Results in terms of the $v_{\textrm{rms}}$ in case of 
-an arithmetic average are also shown in Fig. \ref{fig:inst_spherefs}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_arithm.pdf}
-\caption{$v_{\textrm{rms}}$ for different grid resolutions as function of time for the 2D time-dependent Stokes sphere experiment in case of an arithmetic average.
-Results are compared with results obtained by ASPECT.}
-\label{fig:inst_spherefs}
-\end{figure}
-
-\subsection{Free surface stabilisation}\label{sec:stab}
-The benchmark is performed as discussed in \citet{Kaus2010a} and \citet{Thieulot2014}. The domain is square with $L_x=L_y=$ \SI{500}{\km} and a grid resolution
-of \num{200x200} elements, each of them containing 16 markers. A buoyant fluid with $\rho_1=$ \SI{3200}{\kg\per\cubic\metre} and $\eta_1=$ \SI{e20}{\pascal\s}
-is overlain by a denser fluid with $\rho_2=$ \SI{3300}{\kg\per\cubic\metre} and $\eta_2=$ \SI{e21}{\pascal\s}. The initial interface between the fluids has an
-initial sinusoidal shape of 5 km amplitude. Velocity boundary conditions are set to free slip conditions at the sides and no slip at the bottom, while the top
-is free surface. The experiment is performed with various fixed time steps, without the use of the Courant number. The vertical position of the free surface at
-$x=L_x$ is tracked for each simulation, with or without the stabilisation algorithm.
-
-The results show an instability (drunken sailor effect) increasing the time step in case that the stabilisation algorithm is not activated
-($dt=$ \SI{4200}{\year}, green line in Fig. \ref{fig:stab}), as already observed by \citet{Kaus2010a} and \citet{Thieulot2014}. Activating the algorithm the
-instability is fixed and the simulation remains stable using time step up to \SI{50000}{\year} (red line in Fig. \ref{fig:stab}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Stabilisation_algorithm}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Kaus_random.pdf}
-\caption{Evolution of the y coordinate in $x=L_x$ as function of time for different time steps, with and without the stabilisation algorithm.}
-\label{fig:stab}
-\end{figure}
-
-\subsection{Topography relaxation}\label{sec:crameri}
-This experiment is performed as in \citet{Crameri2012} using free surface and sticky air (Case 1). The domain is rectangular with $L_x=$ \SI{2800}{\km} and
-$L_y=$ \SI{700}{\km} in case of free surface and $L_y=$ \SI{800}{\km} in case of sticky air ($h_{st}=$ \SI{100}{\km}). The grid is composed by \num{256x64}
-elements for the free surface \num{560x320} elements for the sticky air. For both cases, there is a mantle of 600 km thickness with $\eta_m=$ \SI{e21}{\pascal\s}
-overlain by a highly viscous cosine-shaped lithosphere with thickness between 93 and 107 km and $\eta_l=$ \SI{e23}{\pascal\s}. Both mediums have $\rho=$
-\SI{3300}{\kg\per\cubic\metre}. In case of sticky air experiment, the air has $\rho_a=$ \SI{0}{\kg\per\cubic\metre} and $\eta_a=$ \SI{e18}{\pascal\s}. The
-gravity acceleration is set to $g_y=$ \SI{-10}{\metre\per\square\s}. Velocity boundary conditions are set to no slip at the bottom and free slip at the sides
-of the domain. In case of sticky air, velocities on top are set to free slip conditions.
-
-The maximum topography as function of time ($h(t)$) can be analytically derived in according to \[h(t)=h_0 \exp(\gamma t)\]
-where $h_0=7$ km is the initial topography, $\gamma=\num{-0.2139e-11}$ is the characteristic relaxation rate and $t$ is the time (black line in Fig.
-\ref{fig:crameri}). Maximum topography at the characteristic relaxation time $t_{rlx}=$ \SI{14.825}{\kilo\year} can be found to be $h_{rlx}=$ \SI{2576}{\m}.
-The results show that a viscosity of the air of \SI{e18}{\pascal\s} is needed in case of $h_{st}=$ \SI{100}{\km} to correctly describe the topography relaxation
-for this problem, as pointed out by \citet{Crameri2012} (purple line in Fig. \ref{fig:crameri}). The case with a true free surface well follows the analytical
-solution (orange line in Fig. \ref{fig:crameri}). In particular, at the relaxation time in case of the true free surface predicts a maximum topography of
-\SI{2570}{\m}, with an error of \SI{6}{\m} with respect to the analytical solution. Results are also compared with results obtained with a free surface in
-UNDERWORLD (\num{256x64} elements) and SULEC (\num{401x201} elements) (with courtesy from \citealp{Crameri2012}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Topography_relaxation}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Crameri.pdf}
-\caption{Maximum topography as function of time for the Crameri benchmark (coloured lines) in comparison with the analytical solution (black line) and results
-obtained with UNDERWORLD (yellow line) and SULEC (blue line).}
-\label{fig:crameri}
-\end{figure}
-
-\subsection{Spontaneous subduction}\label{sec:subduction}
-This experiment is performed as presented by \citet{Schmeling2008}, in case of both a sticky air and a true free surface and using different average schemes
-for the viscosities. The domain is rectangular with $L_x=$ \SI{3000}{\km} and $L_y=$ \SI{700}{\km}, a grid resolution of \num{256x64} elements with 50 markers
-per element and Courant number of 0.01. At the beginning of the simulation, a 100 km-thick lithospheric layer with $\rho_m=$ \SI{3300}{\kg\per\cubic\metre}
-and $\eta_m=$ \SI{e23}{\pascal\s} is located at the top of the domain for $x$ between 1000 and 3000 km, while the asthenospheric mantle has $\rho_m=$
-\SI{3200}{\pascal\s} and $\eta_m=$ \SI{e21}{\pascal\s}. In addition, a 200 km-depth lithospheric slab is already subducted in the mantle in order to have
-a spontaneous subduction. In case of sticky air, $L_y=750$ km and a 50 km-thick air layer with $\eta_a$ from \SIrange{e19}{e21}{\pascal\s} overlie the mantle.
-Velocity boundary conditions are set to free slip on the sides and on the bottom of the domain. In case of sticky air, velocities on the top boundary are set
-to free slip conditions as well.
-
-The maximum depth of the slab is tracked for all the simulations (Fig. \ref{fig:subduction}) and they are compared with results obtained by
-\citet{Schmeling2008} with models with regular grids and comparable grid resolutions (rectangles in Fig. \ref{fig:subduction}). Low viscous sticky air models
-(i.e., \SI{e19}{\pascal\s}) enlighten a strong dependence of the sinking velocity with the chosen average scheme (dashed lines in Fig. \ref{fig:subduction}),
-while high viscous sticky air models (i.e., \SI{e21}{\pascal\s}) show a higher resistance at the trench, underestimating the correct solution (dotted lines in
-Fig. \ref{fig:subduction}). A high resistance at the trench can be also observed in case of a true free surface for the low resolution grid used (continuous
-lines in Fig. \ref{fig:subduction}), as pointed out by \citet{Schmeling2008}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Spontaneous_subduction}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Subduction.pdf}
-\caption{Maximum depth of the slab as function of time for all the simulations of the \citet{Schmeling2008} benchmark, in case of harmonic, geometric and
-arithmetic means (red, green and blue lines, respectively) using a sticky air or a true free surface (discontinuous and continuous lines, respectively).
-Red, green and blue rectangular areas indicate the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{500}{\km} in case of sticky air,
-while the grey rectangular area indicates the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{600}{\km} in case of true free surface.}
-\label{fig:subduction}
-\end{figure}
-
-\subsection{The slab detachment}\label{sec:slab}
-The slab detachment benchmark is performed as described by \citet{Schmalholz2011} and \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{1000}{\km}
-and $L_y=$ \SI{660}{\km} and a grid resolution of \num{256x256} elements. A non-linear viscous T-shaped layer with $\rho_l=$ \SI{3300}{\kg\per\cubic\metre} is
-placed at the top of the domain and surrounded by a linear viscous fluid with $\rho_f=$ \SI{3150}{\kg\per\cubic\m}. The top layer is 80 km-thick and a 250
-km-long and 80 km-wide slab is placed at $x=L_x/2$ (Fig. \ref{fig:slab}a). The effective viscosity of the top layer is given by
-\[\eta_{eff}=\eta_0 I_2^{\frac{(1-n)}{n}}\]
-with $\eta_0=$ \SI{4.75e11}{\pascal\s} and $n=4$, while the surrounding fluid has $\eta_f=$ \SI{e21}{\pascal\s}. The effective viscosity is capped between
-$\eta_{min}=$ \SI{e21}{\pascal\s} and $\eta_{max}=$ \SI{e25}{\pascal\s}. Velocity boundary conditions are set to free slip at the top and the bottom, and to
-no slip at the sides of the domain.
-
-The sinking of the slab determines a decrease of the effective viscosity in the T-shaped layer during the first part of the evolution (Fig. \ref{fig:slab}a),
-as consequence of local strain rates, while effective viscosities increase up to $\eta_{max}$ in the top layer after approximately \SI{20}{\mega\year}, when
-necking is complete (Fig. \ref{fig:slab}b).
-
-The width of necking is tracked by means of markers position at the side of the slab. Necking width and time are normalised in function of the initial slab
-(\SI{80}{\km}) and the characteristic time ($t_c=$ \SI{7.1158e14}{\s}) \citep{Schmalholz2011,Glerum2018}. The results in case of arithmetic, geometric and
-harmonic mean of the viscosity are shown in Fig. \ref{fig:necking}, compared with results from \citet{Glerum2018} for which an infinity norm mean was used.
-The influence of the average schemes agrees with results shown in Section \ref{sec:subduction}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Slab_detachment}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Slab.pdf}
-\caption{Effective viscosity for the slab detachment benchmark at the beginning of the evolution (panel a) and when necking is complete (panel b).}
-\label{fig:slab}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Necking.pdf}
-\caption{Normalised width of the necking of the slab detachment benchmark as function of normalised time.}
-\label{fig:necking}
-\end{figure}
-
-\subsection{The indenter experiment}\label{sec:indenter}
-The indenter benchmark simulate a rigid punch on a purely plastic von Mises material, for which exists an analytical solution
-\citep{Thieulot2008,Thieulot2014,Glerum2018}. The domain is a unit square with a grid resolution of \num{256x256} elements. The gravity acceleration is set
-to $\bm{g}=\bm{0}$ and effective viscosity is capped using $\eta_{min}=10^{-4}$ and $\eta_{max}=10^3$. The tolerance for the convergence of the non-linear
-solution is $tol=10^{-9}$, with a maximum number of non-linear iterations set to 500. The medium in the domain has $\rho=0.01$, an initial viscosity
-$\eta_0=10$, cohesion $C=1$ and an angle of internal friction $\phi=0$°. Velocity boundary conditions are set to no slip at the bottom and free slip at the
-sides of the domain. The top of the domain is open with the exception of the central portion (punch area) with width $w_p=0.125$, where $v=-1.05$ and $u$ is
-fixed either to 0 (rough punch experiment) or free (smooth punch experiment).
-
-Fig. \ref{fig:indenter} show the results in terms of viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure
-(panels g and h) for both experiments. The analytical solution indicate that pressure at ($0.5;1$) and ($0.5 \pm w_p;1$)is $p=\pi +1$ and $p=1$, respectively
-(black and red lines in Fig. \ref{fig:smooth_p}). Fig. \ref{fig:smooth_p} shows a clear improvement in the pressure solution when passing from a rough to a
-smooth punch (green and blue lines in Fig. \ref{fig:smooth_p}, respectively), as previously observed by \citet{Thieulot2014} and \citet{Glerum2018}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Indenter}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Indenter.pdf}
-\caption{Viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure (panels g and h) fields for rough (left column) and
-smooth (right column) punch experiments.}
-\label{fig:indenter}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Smooth_Rough.pdf}
-\caption{Nodal pressure in function of the x coordinate at $y=1$ for rough (green line) and smooth (blue line) punch experiment. Black and red lines indicate
-the analytical solution at $x=0.5$ and $x=0.5 \pm w_p$, respectively.}
-\label{fig:smooth_p}
-\end{figure}
-
-\subsection{The brick experiment}\label{sec:brick}
-An instantaneous version of the brick benchmark is performed to verify the correctness of the pressure-dependent plasticity for different angle of internal
-friction, as proposed by \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{40}{\km} and $L_y=$ \SI{10}{\km} and a grid resolution of \num{512x128}
-elements. The gravity acceleration is set to $g_y=$ \SI{-10}{\m\square\s} and effective viscosity is capped using $\eta_{min}=$ \SI{e19}{\pascal\s} and
-$\eta_{max}=$ \SI{e26}{\pascal\s}. The tolerance for the convergence of the non-linear solution is $tol=10^{-7}$, with a maximum number of non-linear iterations
-set to 1000. A 800 m-wide and 400 m-height inclusion with $\eta_b=$ \SI{e20}{\pascal\s} and $\rho_b=$ \SI{2700}{\kg\per\cubic\m} is placed at the bottom of the
-domain at $x=L_x/2$. The inclusion is surrounded by a non-linear viscous medium with $\rho_m=$ \SI{2700}{\kg\per\cubic\m}, an initial viscosity of
-\SI{e23}{\pascal\s}, a linear viscous viscosity of \SI{e25}{\pascal\s} and a cohesion of \SI{40}{\mega\pascal}. Velocities are set to free slip conditions at
-the bottom of the domain and the top is open. Velocities on sides of the domain are fixed to $u=$ \SI{+-2e-11}{\m\per\s} and $v=0$. The experiment is performed
-in compressional and extensional contexts with an angle of internal friction $\phi$ of the non-linear medium variable between 0° and 30°.
-
-As expected, two shear bands stem from the inclusion with variable angles in relation to both the dynamics context and the internal friction angle. Shear band
-angles formed at 45° for $\phi=0$° in both extensional and compressional contexts (Fig. \ref{fig:brick_beam}a), while they are different in case of $\phi\neq0$°.
-Shear band angles in case of $\phi=20$° are shown for compression and extension in Fig. \ref{fig:brick_beam}b  and c, respectively. Values of shear band angles
-as function of internal friction angles from 0° to 30° are extracted at different depths and minimum and maximum values are plotted in Fig. \ref{fig:brick},
-in comparison with theoretical Roscoe, Arthur and Coulomb shear band angles. For all tests the tolerance for nonlinear convergence is defined $tol=10^{-7}$
-and none converges before the maximum number of iterations ($it_{max}=1000$) is reached. However, tests with internal friction angles up to 15° show a constant
-decrease of the velocity residuals, which can not be observed for for tests with higher internal friction angles (Fig. \ref{fig:convergence}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Brick_experiment}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Brick_Beam.pdf}
-\caption{Shear band angles predicted for the brick experiment for $\phi=0$° (panel a) and for $\phi=20$° in case of compression and extension (panels b and c,
-respectively).}
-\label{fig:brick_beam}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Brick.pdf}
-\caption{Shear band angles predicted for the brick experiment in case of compressional and extensional contexts as function of different internal angle of
-friction (continuous black line and red dots), compared with theoretical Roscoe, Artur and Coulomb angles (discontinuous black lines). Red lines indicate the
-range of angles calculated at different depths.}
-\label{fig:brick}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \noindent\includegraphics[width=500px]{./Figures/convergence.pdf}
-  \caption{Convergence rates for all tests. Normalised residuals (left column) are compared with velocity residuals (right column) for extensional (first row)
-  and compressional (second row) settings.}
-  \label{fig:convergence}
-  \end{figure}
-
-\subsection{Advection stabilisation}\label{sec:advection}
-The 1D advection problem proposed by \citet{Donea2003} and \citet{Thieulot2011} is performed to verify the effectiveness of the SUPG method to stabilise the
-advection term of the energy equation. The domain is a 1D segment with $L_x=1$ composed by 50 elements and a discontinuity in the thermal field placed at
-$x=0.25$. Temperature is set to 1 for $x \leq 0.25$ and to 0 for $x > 0.25$. Velocity is set to ${u}=1$ in the entire domain. The simulation is performed for
-250 time steps, with $dt=0.002$, so the thermal discontinuity should be at $x=0.75$ at the end of the simulation.
-
-Temperature profiles at the end of the simulation are shown in \ref{fig:advection} as function of the dimensionless coefficient $\gamma=\tau \bm{v}/h$. In case
-of the classic Galerkin method ($\gamma=0$, blue line in \ref{fig:advection}) the final thermal profile is characterised by strong oscillations, which are
-eliminated in case of the SUPG method ($\gamma=0.045$, orange line in \ref{fig:advection}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Advection_stabilisation}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Advection.pdf}
-\caption{Temperature profile as function of \textit{x} for the advection stabilisation benchmark. Purple line indicates the initial temperature profile; the
-green line indicates the analytical temperature profile after 250 time steps; blue line indicates the temperature profile after 250 time steps in case of the
-classic Galerkin method ($\gamma=0$); orange line indicates the temperature profile after 250 time steps in case of the SUPG method ($\gamma=0.045$).}
-\label{fig:advection}
-\end{figure}
-
-\subsection{Shear heating}\label{sec:simple_shear}
-This exercise is performed in an unit square domain composed by \num{128x128} elements. The velocity field is prescribed on the entire domain with
-$\bm{v}=(L_y-y)y\bm{e_x}$; viscosity, density and specific heat are fixed to 1, while thermal conductivity and radiogenic energy are fixed to 0. Therefore,
-the energy equation (Eq. \ref{eq:energy}) can be simplified as \[\frac{\partial T}{\partial t}=H_s\] and fixing $T(t=0)=0$, the temperature can be find as
-\[T(t)=H_s t\]
-In this case we have
-\begin{eqnarray}
-\dot{\epsilon_{xx}}&=&\frac{\partial u}{\partial x}=0 \nonumber \\
-\dot{\epsilon_{yy}}&=&\frac{\partial v}{\partial y}=0 \nonumber \\
-\dot{\epsilon_{xy}}&=&\frac{1}{2}\left(\frac{\partial v}{\partial x}+\frac{\partial u}{\partial y}\right)=\frac{1}{2}(L_y-2y)\nonumber
-\end{eqnarray}
-and, simplifying Eq. \ref{eq:shear}, shear heating can be calculated as \[H_s(x,y)=(1-2y)^2\]
-
-The solution predicted by the model in terms of velocity, temperature and shear heating match well with the analytical solutions (Fig. \ref{fig:analytical_en}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Simple_shear}.
-
-%\begin{figure}
-%\centering
-%\noindent\includegraphics[width=400px]{./Figures/Energy_Thieulot.pdf}
-%\caption{Temperature and shear heating predicted by the model after 1 time step.}
-%\label{fig:en_thieulot}
-%\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=500px]{./Figures/Analytical_Shear.pdf}
-\caption{Velocity (panel a), temperature (panel b) and shear heating (panel c) as function of the \textit{y} coordinate for the simple shear experiment. The
-solutions predicted by the model (red lines) are compared with the analytical solutions (dashed blue lines).}
-\label{fig:analytical_en}
-\end{figure}
-
-\subsection{Shear and adiabatic heating}\label{sec:shear}
-This problem is performed as presented in Exercise 9.4 in \citet{Gerya2010b}. Two materials are vertically separated in a rectangular domain with $L_x=$
-\SI{1000}{\km}, $L_y=$ \SI{1500}{\km} and a grid resolution of \num{30x20} elements. Constant thermal coefficient expansion ($\alpha=$ \SI{3e-5}{\per\kelvin}),
-temperature ($T=$ \SI{1300}{\kelvin}) and gravity acceleration ($g_y=$ \SI{-10}{\m\square\s}) are assumed in the whole domain. Fluid 1 (on the left side of the
-domain) has $\rho_1=$ \SI{3200}{\kg\per\cubic\m} and $\eta_1=$ \SI{e20}{\pascal\s}; fluid 2 (on the right side of the domain) has $\rho_2=$
-\SI{3300}{\kg\per\cubic\m} and $\eta_2=$ \SI{e22}{\pascal\s}. Velocity boundary conditions are set to free slip on all sides of the domain.
-
-As shown in Fig. \ref{fig:energy}, both shear and adiabatic heating predicted by the code (first row in Fig. \ref{fig:energy}) well recreate the results
-obtained by means of example ${Shear\_adiabatic\_heating.m}$ from \citet{Gerya2010b} (second row in Fig. \ref{fig:energy}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Adiabatic%2BShear}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Energy.pdf}
-\caption{Comparison between shear (first column), adiabatic (second column) and total (third column) energy predicted by the code (panels a, b and c) and those
-created using example ${Shear\_adiabatic\_heating.m}$ from Exercise 9.4 in \citet{Gerya2010b} (panels d, e and f).}
-\label{fig:energy}
-\end{figure}
-
-\subsection{Mantle convection}\label{sec:mantle}
-This problem is performed as presented by \citet{Blankenbach1989} (constant viscosity cases) and \citet{Thieulot2014}, in a 2D unit square domain with gravity
-acceleration $g_y=10^{10}Ra$. The experiment is performed with three different Rayleigh numbers ($Ra=10^4$, $10^5$ and $10^6$) and with different grid
-resolution (between \num{32x32} and \num{128x128} elements). The fluid has constant viscosity, initial density, heat capacity, thermal conductivity
-($\eta= \rho_0= C_p = k=1$), reference temperature ($T_0=0$) and thermal expansion coefficient ($\alpha=10^{-10}$). Temperatures are set to 0 on top and 1 on
-bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\]
-
-The solution generated by the code in terms of $v_{\textrm{rms}}$ (calculated as in Sec. \ref{sec:error}) and Nu as function of time are reported for all the
-simulations in Fig. \ref{fig:mantle}. At the steady state, $v_{\textrm{rms}}$ and Nu of all simulations converge well toward the values from \citet{Blankenbach1989},
-with lower errors for higher resolution grids (Table \ref{tab:mantle}). All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Mantle_convection}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Convection.pdf}
-\caption{$v_{\textrm{rms}}$ (panels a, c and e) and Nu (panels b, d and f) for the mantle convection benchmark as function of time for different grid resolution.
-Panels a and b show the results for $Ra=10^4$; Panels c and d show the results for $Ra=10^5$; panels e and f show the results for $Ra=10^6$. Purple lines
-indicate the convergence values for the $v_{\textrm{rms}}$ and the Nu from \citet{Blankenbach1989} at the steady state.}
-\label{fig:mantle}
-\end{figure}
-
-\begin{table}[H]
-\caption{Comparison between $v_{\textrm{rms}}$ and Nu predicted by the code for the mantle convection experiment and same values as reported in literature.}
-\centering
-\small
-\begin{tabular}{|>{\centering}m{0.09\textwidth}|>{\centering}m{0.05\textwidth}|>{\centering}m{0.24\textwidth}|>{\centering}m{0.15\textwidth}|
-  >{\centering}m{0.1\textwidth} >{\centering}m{0.1\textwidth} >{\centering\arraybackslash}m{0.1\textwidth}|}
-\toprule
-& & \multirow{2}{*}{\citet{Blankenbach1989}} & \citet{Thieulot2014} & \multicolumn{3}{c|}{FALCON}  \\
- & & & \num{200x200} & \num{32x32} & \num{64x64} & \num{128x128}  \\
-\midrule
-\multirow{2}{*}{$Ra=10^4$} & $v_{\textrm{rms}}$ & $42.864947 \pm 0.000020$ & 42.867 & 42.83226 & 42.852793 & 42.861394  \\
-           & Nu      & $4.884409 \pm 0.000010$  & 4.882  & 4.781297 & 4.857475  & 4.877573   \\
-\hline
-\multirow{2}{*}{$Ra=10^5$} & $v_{\textrm{rms}}$ & $193.21454 \pm 0.00010$  & 193.255& 193.872643 & 193.377472 & 193.252290  \\
-           & Nu      & $10.534095 \pm 0.000010$ & 10.507 & 9.602514 & 10.270735  & 10.465629   \\
-\hline
-\multirow{2}{*}{$Ra=10^6$} & $v_{\textrm{rms}}$ & $833.98977 \pm 0.00020$  & 834.712& 848.091176 & 837.767911 & 834.945793  \\
-           & Nu      & $21.972465 \pm 0.000020$ & 21.695 & 15.999266 & 19.703682  & 21.306939   \\
-\bottomrule
-\end{tabular}
-\label{tab:mantle}
-\end{table}
-
-\subsection{Viscoplastic mantle convection}\label{sec:visco_mantle}
-This benchmark is performed as Cases 1-5a described in \citet{Tosi2015} in a 2D unit square domain with two different grid resolutions of \num{32x32} 
-and \num{100x100} elements. Reference viscosity, reference density, heat capacity, thermal conductivity and thermal expansion coefficient are fixed to a constant
-value of 1 ($\eta_0= \rho_0= C_p = k=\alpha=1$), while the gravity has been chosen as $g_y=10^2$, in order to obtain a Rayleigh number $Ra=10^2$. Temperatures
-are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\].
-The viscosity field $\eta$ is calculated as the harmonic average between a linear part $\eta_{lin}$ that depends on temperature only or on temperature and 
-depth $y$ and a nonlinear-plastic part $\eta_{pl}$ that depends on the strain rate $\dot{\bm{\epsilon}}$, as follows
-\[\eta(T,y,\dot{\bm{\epsilon}})=2\left(\frac{1}{\eta_{lin}(T,y)}+\frac{1}{\eta_{pl}(\dot{\bm{\epsilon}})}\right)^{-1}\]
-The linear and the nonlinear-plastic parts of the viscosity are calculated following \citet{Tosi2015} as follows
-\[\eta_{lin}(T,y)=\exp(-\ln(\Delta\eta_T)T+\ln(\Delta\eta_y)y)\]
-\[\eta_{pl}(\dot{\bm{\epsilon}})=\eta^*+\frac{\sigma_Y}{\sqrt{\dot{\bm{\epsilon}}:\dot{\bm{\epsilon}}}}\]
-where $\Delta\eta_T$, $\Delta\eta_y$, $\eta^*$ and $\sigma_Y$ are parameters chosen as listed in Table \ref{tab:tosi_setup}.
-
-For each case, some diagnostic quantities measured at the steady state are reported in Table \ref{tab:visco_mantle} (see Section \ref{sec:error} for 
-the computation of all the quantities), showing a well fit with results described in \citet{Tosi2015}. Data and charts for all cases can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Viscoplastic_mantle_convection}.
-
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{white!100!}{gray!25!}
-%\begin{table}[H]
-%  \caption{Results of the viscoplastic mantle convection benchmark for Cases 1-4 compared with results obtained with ELEFANT and YACC (from \citealp{Tosi2015}).}
-%  \centering
-%  \resizebox{\textwidth}{!}{
-%\begin{tabular}{|c|ccc|ccc|ccc|ccc|}
-%  \toprule
-%    & \multicolumn{3}{c|}{\textbf{Case 1}} & \multicolumn{3}{c|}{\textbf{Case 2}} & \multicolumn{3}{c|}{\textbf{Case 3}} &
-%    \multicolumn{3}{c|}{\textbf{Case 4}} \\
-%   & FALCON & ELEFANT& YACC & FALCON& ELEFANT& YACC & FALCON & ELEFANT & YACC & FALCON & ELEFANT & YACC \\
-%  \midrule
-%   $\left\langle T \right\rangle$ & 0.7759 & 0.7758 & 0.7767 & 0.6043 & 0.6036 & 0.6058 & 0.7275 & 0.7275 & 0.7286 & & 0.5277 & 0.5289\\
-%    Nu$^{top}$ & 3.4280 & 3.4214 & 3.4298 & 8.5242 & 8.5115 & 8.5278 & 3.0368 & 3.0347 & 3.0374 && 6.5912 & 6.5572 \\
-%    Nu$^{bottom}$ & 3.3168 & 3.313 & 3.3143 & 8.4599 & 8.4465 & 8.3990 & 3.0086 & 2.9908 & 2.9628 && 6.5834 & 6.5243 \\
-%    $v_{\textrm{rms}}$ & 249.901 & 249.134 & 251.7997 & 141.2451 & 140.7087 & 142.2020 & 100.2776 & 100.1208 & 100.9467 && 79.1105 & 79.6202 \\
-%    $v_{\textrm{rms}}^{top}$ & 1.8907 & 1.8642 & 1.8298 & 104.7928 & 104.1228 & 105.1767 & 2.0832 & 2.0652 & 2.0374 && 74.7596 & 75.8414 \\
-%    $u_{max}^{top}$ & 2.6340 & 2.6119 & 2.5516 & 121.9212& 121.766 & 122.3238 & 2.9086 & 2.9019 & 2.8458 && 88.9146 & 89.2940 \\
-%    $\eta_{min}$ & - & - & - & $2.0303\times10^{-5}$ & $2.0600\times10^{-5}$ & $1.9853\times10^{-5}$ & $4.8129\times10^{-5}$ & $4.8080\times10^{-5}$ & $4.7907\times10^{-5}$ &
-%    & $1.9860\times10^{-4}$ & $1.9174\times10^{-4}$ \\
-%    $\eta_{max}$ & - & - & - & 1.3325 & 1.85 & 1.6095 & 0.9360 & 0.9023 & 1 && 1.52 & 1.6773 \\
-%    $\left\langle W \right\rangle$ & 2.4299 & 2.4316 & 2.4583 & 7.5648 & 7.5665 & 7.6222 & 2.0387 & 2.0384 & 2.0400 && 5.6216 & 5.6512 \\
-%    $\left\langle \Phi \right\rangle/Ra$ & 2.4311 & 2.4276 & 2.4333 & 7.5661 & 7.5585 & 7.6194 & 2.0386 & 2.0370 & 2.0335 && 5.6182 & 5.6463 \\
-%    $\delta$ & 0.05\% & 0.16\% & 1.02\% & 0.02\% & 0.11\% & 0.04\% & $<$0.01\% & 0.07\% & 0.32\% && 0.06\% & 0.09\% \\
-%  \bottomrule
-%  \end{tabular}}
-%  \label{tab:visco_mantle}
-%  \end{table}
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{2}{white!100!}{gray!25!}
-\begin{table}[H]
-  \caption{Benchmark cases and corresponding parameters.}
-  \centering
-  \small
-\begin{tabular}{|>{\centering}m{0.08\textwidth}|>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}
-  >{\centering}m{0.05\textwidth} >{\centering}m{0.05\textwidth} |>{\centering\arraybackslash}m{0.2\textwidth}|}
-  \toprule
-    Case & $Ra$ & $\Delta\eta_T$ & $\Delta\eta_y$ & $\eta^*$ & $\sigma_Y$ & Convective regime \\
-  \midrule
-  \rowcolor{gray!25!}
-    1 & $10^2$ & $10^5$ & 1 & - & - & Stagnant lid \\
-    2 & $10^2$ & $10^5$ & 1 & $10^{-3}$ & 1 & Mobile lid  \\
-    \rowcolor{gray!25!}
-    3 & $10^2$ & $10^5$ & 10 & - & - & Stagnant lid  \\
-    4 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 1 & Mobile lid \\
-    \rowcolor{gray!25!}
-    5 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 4 & Periodic \\
-  \bottomrule
-  \end{tabular}
-  \label{tab:tosi_setup}
-  \end{table}
-
-\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{gray!25!}{white!100!}
-\begin{table}
-  \caption{Results of the viscoplastic mantle convection benchmark for different grid resolutions in Cases 1-5 compared with results obtained with ELEFANT
-  and YACC ($100\times100$ elements, from \citealp{Tosi2015}) and in Stone 28 ($32\times32$ elements, from
-  \url{https://github.com/cedrict/fieldstone/tree/master/python_codes/fieldstone_28}).}
-  \centering
-  \resizebox{\textwidth}{!}{
-\begin{tabular}{|cc|ccccccccccc|}
-  \toprule
-  \multicolumn{13}{|c|}{\textbf{Case 1}} \\
-  \midrule
-  Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-  & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-  \hline
-  \rowcolor{gray!25!}
-    YACC & $100\times100$ & 0.7767 & 3.4298 & 3.3143 & 251.7997 & 1.8298 & 2.5516 & - & - & 2.4583 & 2.4333 & 1.02\% \\
-    ELEFANT & $100\times100$ & 0.7758 & 3.4214 & 3.313 & 249.134 & 1.8642 & 2.6119 & - & - & 2.4316 & 2.4276 & 0.16\% \\
-    \rowcolor{gray!25!}
-    Stone 28 & $32\times32$ & 0.7737 & 3.3987 & - & 243.872 & - & - & - & - & - & - & - \\
-%    FALCON & $32\times32$ & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & - & - & 2.4745 & 2.4864 & 0.48\% \\
-%    \rowcolor{gray!25!}
-    FALCON & $100\times100$ & 0.7759 & 3.4280 & 3.3168 & 249.901 & 1.8907 & 2.6340 & - & - & 2.4299 & 2.4311 & 0.05\% \\
-    \midrule
-    \multicolumn{13}{|c|}{\textbf{Case 2}} \\
-    \midrule
-    Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-    & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-    \hline
-    \rowcolor{gray!25!}
-     YACC & $100\times100$ & 0.6058 & 8.5278 & 8.3990 & 142.2020 & 105.1767 & 122.3238 & $1.9853\times10^{-5}$ & 1.6095 & 7.6222 & 7.6194 & 0.04\% \\
-     ELEFANT & $100\times100$ & 0.6034 & 8.5115 & 8.4465 & 140.8067 & 104.1228 & 121.766 & $2.0600\times10^{-5}$ & 1.85 & 7.5665 & 7.5585 & 0.11\% \\
-     \rowcolor{gray!25!}
-     Stone 28 & $32\times32$ & 0.6052 & 8.1715 & - & 141.518 & - & - & - & - & - & - & - \\
-%     FALCON & $32\times32$ & 0.6454 & 8.1422 & 7.4020 & 156.6908 & 95.7226 & 116.9302 & $2.2572\times10^{-5}$ & 0.7194 & 7.4289 & 7.4452 & 0.22\% \\
-%     \rowcolor{gray!25!}
-     FALCON & $100\times100$ & 0.6044 & 8.5266 & 8.4590 & 141.3684 & 104.8867 & 122.0470 & $2.0333\times10^{-5}$ & 1.2914 & 7.5707 & 7.5718 & 0.01\% \\
-     \midrule
-     \multicolumn{13}{|c|}{\textbf{Case 3}} \\
-     \midrule
-     Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-     & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-     \hline
-     \rowcolor{gray!25!} 
-     YACC & $100\times100$ & 0.7286 & 3.0374 & 2.9628 & 100.9467 & 2.0374 & 2.8458 & $4.7907\times10^{-5}$ & 1 & 2.0400 & 2.0335 & 0.32\% \\
-      ELEFANT & $100\times100$ & 0.7275 & 3.0347 & 2.9908 & 100.1208 & 2.0652 & 2.9019 & $4.8080\times10^{-5}$ & 0.9023 & 2.0384 & 2.037 & 0.07\% \\
-      \rowcolor{gray!25!}
-      Stone 28 & $32\times32$ & 0.7277 & 3.0347 & - & 100.018 & - & - & - & - & - & - & - \\
-%      FALCON & $32\times32$ & 0.7281 & 3.0618 & 2.8132 & 101.0092 & 2.1630 & 3.0178 & $4.8500\times10^{-5}$ & 0.7845 & 2.0699 & 2.0695 & 0.02\% \\
-%      \rowcolor{gray!25!}
-      FALCON & $100\times100$ & 0.7275 & 3.0368 & 3.0086 & 100.2776 & 2.0832 & 2.9086 & $4.8129\times10^{-5}$ & 0.9360 & 2.0387 & 2.0386 & $<$0.01\% \\
-      \midrule
-      \multicolumn{13}{|c|}{\textbf{Case 4}} \\
-      \midrule
-      Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-      & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-      \hline
-      \rowcolor{gray!25!} 
-      YACC & $100\times100$ & 0.5289 & 6.5572 & 6.5243 & 79.6202 & 75.4814 & 89.2940 & $1.9174\times10^{-4}$ & 1.6773 & 5.6512 & 5.6463 & 0.09\% \\
-       ELEFANT & $100\times100$ & 0.5277 & 6.5912 & 6.5834 & 79.1105 & 74.7596 & 88.9146 & $1.9860\times10^{-4}$ & 1.5200 & 5.6216 & 5.6182 & 0.06\% \\
-       \rowcolor{gray!25!}
-       Stone 28 & $32\times32$ & 0.5286 & 6.4036 & - & 79.4746 & - & - & - & - & - & - & - \\
-%       FALCON & $32\times32$ & 0.5334 & 6.4679 & 6.3435 & 80.6153 & 76.0511 & 90.2058 & $2.0393\times10^{-4}$ & 0.6453 & 5.6676 & 5.6698 & 0.04\% \\
-%       \rowcolor{gray!25!}
-       FALCON & $100\times100$ & 0.5279 & 6.5897 & 6.5830 & 79.1220 & 75.0715 & 88.8484 & $1.9693\times10^{-4}$ & 1.4133 & 5.6127 & 5.6113 & 0.03\% \\
-       \midrule
-       \multicolumn{13}{|c|}{\textbf{Case 5}} \\
-       \midrule
-       Code & Resolution & $P$ & $\left\langle T \right\rangle_{min}$ & $\left\langle T \right\rangle_{max}$ & Nu$^{top}_{min}$ & Nu$^{top}_{max}$ &
-       $v_{\textrm{rms}}^{min}$ & $v_{\textrm{rms}}^{max}$ & \multicolumn{2}{c}{$(\left\langle \Phi \right\rangle/Ra)_{min}$} & 
-       \multicolumn{2}{c|}{$(\left\langle \Phi \right\rangle/Ra)_{max}$} \\
-       \hline
-       \rowcolor{gray!25!}
-         YACC & $100\times100$ & 0.0792 & 0.6543 & 0.6722 & 2.6950 & 7.2792 & 41.9813 & 98.8230 & \multicolumn{2}{c}{1.4347} & 
-         \multicolumn{2}{c|}{9.2748} \\
-         ELEFANT & $100\times100$ & 0.076 & 0.6514 & 0.6693 & 2.6808 & 7.3506 & 41.2578 & 100.089 & \multicolumn{2}{c}{1.3324} & 
-         \multicolumn{2}{c|}{9.5289} \\
-         \rowcolor{gray!25!}
-         Stone 28 & $32\times32$ & 0.0764 & 0.6521 & 0.6697 & 2.6912 & 7.0763 & 41.61 & 98.183 & \multicolumn{2}{c}{-} & 
-         \multicolumn{2}{c|}{-} \\
-     %    FALCON & $32\times32$ & & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & \multicolumn{2}{c}{-} & 
-     %     \multicolumn{2}{c|}{-} \\
-     %    \rowcolor{gray!25!}
-         FALCON & $100\times100$ & 0.0775 & 0.6519 & 0.6697 & 2.6847 & 7.3629 & 41.5256 & 99.5466 & \multicolumn{2}{c}{1.3399} & 
-         \multicolumn{2}{c|}{9.3993} \\
-  \bottomrule
-  \end{tabular}}
-  \label{tab:visco_mantle}
-  \end{table}
-
-\subsection{Thin layer entrainment}\label{sec:thin}
-This experiment is performed as originally proposed by \citet{vanKeken1997} with two incompressible fluids in a rectangular domain with $L_x=2$ and $L_y=1$
-and gravity acceleration $g_y=10^{10}Ra$. The Rayleigh number and the compositional Rayleigh number are fixed to $Ra=300000$ and $Rc=450000$, respectively.
-Both fluids have constant viscosity, thermal conductivity, specific heat ($\eta=\rho=C_p=1$) and thermal expansion coefficient ($\alpha=10^{-10}$). Fluid 1
-has a density $\rho_1=1$, while fluid 2 is denser ($\rho_2=\rho_1+1.5 \cdot 10^{-10}$) and is placed at the bottom of the domain, for $y \leq 0.025$.
-Temperature are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides of the domain. The initial
-temperature field is given by
-\[T(x,y)=T_u(x,y)+T_l(x,y)+T_r(x,y)+T_s(x,y)-\frac{3}{2}\]
-with
-\begin{eqnarray}
-T_u(x,y)&=&\frac{1}{2}erf\left(\frac{1-y}{2}\sqrt{\frac{u_0}{x}}\right)\nonumber \\
-T_l(x,y)&=&1-\frac{1}{2}erf\left(\frac{y}{2}\sqrt{\frac{u_0}{L_x-x}}\right)\nonumber \\
-T_r(x,y)&=&\frac{1}{2}+\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{y+1}} \exp\left(-\frac{x^2u_0}{4y+4}\right)\nonumber \\
-T_s(x,y)&=&\frac{1}{2}-\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{2-y}} \exp\left(-\frac{(L_x-x)^2u_0}{8-4y}\right)\nonumber
-\end{eqnarray}
-and
-\begin{eqnarray}
-u_0&=&\frac{L_x^{7/3}}{(1+L_x^4)^{2/3}}\left(\frac{Ra}{2\sqrt{\pi}}\right)^{\frac{2}{3}}\nonumber \\
-Q&=&2\sqrt{\frac{L_x}{\pi u_0}}\nonumber
-\end{eqnarray}
-
-The experiment is performed with two grid resolution (\num{125x40} and \num{200x80} elements), with 100 markers per element and CFL of 0.25. $v_{\textrm{rms}}$
-(see Sec. \ref{sec:error}) calculated for both the simulations match well with results from \citet{vanKeken1997} and obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and ELEFANT \citep{Thieulot2014}, obtained with similar grid resolutions
-(Fig. \ref{fig:thin}). In addition, a variety of simulations with different resolutions and aspect ratios are performed to verify that the sensitivity of the
-initial velocity field. Results match well with results from \citet{vanKeken1997} (in the grey area) and \citet{Thieulot2014} (Fig. \ref{fig:thin_initial}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Thin_layer}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_thin.pdf}
-\caption{$v_{\textrm{rms}}$ for the thin layer experiment as function of time for different grid resolution. Results are compared with results obtained by
-\citet{vanKeken1997} (black lines) and with ASPECT and ELEFANT.}
-\label{fig:thin}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrmszero.pdf}
-\caption{$v_{\textrm{rms}}$ at $t=0$ for the thin layer experiment as function of the element size for different aspect ratios of the elements. Results are compared
-with results obtained by \citet{vanKeken1997} (black dots in the grey area) and with ASPECT and ELEFANT.}
-\label{fig:thin_initial}
-\end{figure}
-
-\subsection{Sinking hydrated cylinder}\label{sec:quinquis}
-This experiment is performed as explained in \citet{Quinquis2014} and simulates the sinking of a cold, hydrated cylinder into a warm, dry mantle. The domain
-is square with $L_x=L_y=$ \SI{300}{\km}, a grid resolution of \num{300x300} elements and 25 markers per element. The sinking cylinder is located at $x=$
-\SI{150}{\km} and $y=$ \SI{170}{\km} with a radius of \SI{20}{\km} and has $\rho_c=$ \SI{3250}{\kg\per\cubic\m}, $\eta_c=$ \SI{e23}{\pascal\s}, a thermal
-conductivity $k_c=$ \SI{4.5}{\watt\per\m\per\kelvin},a specific heat $C_{p_c}=$ \SI{1250}{\joule\per\kg\per\kelvin}; the surrounding mantle has $\rho_m=$
-\SI{3200}{\kg\per\cubic\m}, $\eta_m=$ \SI{e20}{\pascal\s}, a thermal conductivity $k_m=$ \SI{105}{\watt\per\m\per\kelvin} and specific heat $C_{p_m}=$
-\SI{1250}{\joule\per\kg\per\kelvin}; a 58 km-thick lithosphere overlies the mantle and has $\rho_l=$ \SI{3200}{\kg\per\cubic\m}, $\eta_l=$ \SI{e23}{\pascal\s},
-a thermal conductivity $k_l=$ \SI{4.5}{\watt\per\m\per\kelvin}, specific heat $C_{p_l}=$ \SI{750}{\joule\per\kg\per\kelvin} and a thermal diffusivity of
-\SI{e-6}{\square\m\per\s}. Velocity boundary conditions are set to free slip on all sides of the domain. Temperature increase linearly in the lithosphere,
-from 0° to 1300°C, and in the mantle, from 1300° to 1360.5°C. The initial temperature of the cylinder is fixed to 400°C. Throughout the evolution, temperature
-is fixed to 0° and 1360.5°C at the top and bottom of the domain, respectively. The initial bound water content is imposed at 2 wt.\% in the cylinder and at
-0 wt.\% in both the mantle and the lithosphere. The maximum amount of water in the mantle is fixed to 0.2 wt.\%, while maximum water content in the cylinder
-and in the lithosphere is function of pressure and temperature and is calculated using a serpentinized harzburgite with Perple\_X, as in \citet{Quinquis2014}.
-
-As shown in \citet{Quinquis2014}, the model is characterised by a progressive dehydration in the external portion of the cylinder, due to the increase of
-temperature, with a consequent vertical migration of free water that hydrates the mantle above the cylinder, up to the lithosphere. The distribution of
-bound and free water during the evolution of the experiment is shown in Fig. \ref{fig:quinquis} (left and right columns, respectively). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Hydration}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Hydration.pdf}
-\caption{Distribution of bound and free water (left and right columns, respectively) at $t=0.5$ and $t=$ \SI{1.25}{\mega\year} (first and second rows,
-respectively) for the sinking hydrated cylinder experiment. Only markers with bound or free water not equal to 0 are plotted.}
-\label{fig:quinquis}
-\end{figure}
-
-\subsection{Mantle melting curves}\label{sec:katz}
-This instantaneous experiment is performed in order to recreate isobaric melting curves, as function of temperature, bulk water and content of water in the
-melt, to compare with melting curves obtained by \citet{Katz2003}. The domain is square with $L_x=L_y=100$ km, a grid resolution of \num{100x100} elements and
-25 markers per element. Density is set in the entire domain to \SI{3300}{\kg\per\cubic\m} and temperature increases from the sides to the centre from 1000° to
-1500°C. The experiment is performed for bulk water contents from 0 to 0.3 wt.\%.
-
-Fig. \ref{fig:fraction} shows isobaric melting curves for \SI{1}{\giga\pascal} (panel a) and \SI{3}{\giga\pascal} (panel b) calculated for different bulk water
-contents. The curves well match with those obtained for the mantle by \citet{Katz2003}, showing the saturation at the solidus for bulk water of 0.2 and 0.3 wt.\%, indicating
-that the code correctly determines melt fraction as function of pressure, temperature and bulk water content. Same results can be observed in Fig.
-\ref{fig:water}a, where isobaric and isothermal melting curves are shown as function of bulk water content. In addition, Fig. \ref{fig:water}b shows that the
-code correctly determines also the percentage of water in the melt as function of pressure, temperature and melt fraction. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Melting}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Fraction.pdf}
-\caption{Isobaric melting curves for 1 GPa (panel a) and 3 GPa (panel b) as function of temperature with different bulk water contents.}
-\label{fig:fraction}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Water.pdf}
-\caption{Isobaric melting curves for 1.5 GPa as function of bulk water (panel a) and water in the melt (panel b) calculated for different temperatures.}
-\label{fig:water}
-\end{figure}
-
-\clearpage
-
-\bibliography{export}
-
-\end{document}

### 2022-09-07 tex/Benchmark/advection.tex
update main

--- 2022-03-31 FALCON_description.tex
@@
-\documentclass[hidelinks,10pt,a4paper]{article}
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{textcomp}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage{booktabs}
-\usepackage{multirow}
-\usepackage[cm]{fullpage}
-\usepackage[round]{natbib}
-\bibliographystyle{abbrvnat}
-\usepackage{doi}
-\usepackage{bm}
-\usepackage[table]{xcolor}
-\hypersetup{
-    colorlinks,
-    linkcolor={blue},
-    citecolor={blue},
-    urlcolor={blue}
-}
-\usepackage{float}
-\restylefloat{table}
-\usepackage{siunitx}
-\DeclareSIUnit\year{yr}
-
-\begin{document}
-
-\section{Introduction}\label{sec:intro}
-This document includes a description of the code FALCON, with a presentation of the features implemented followed by results of extensive benchmarks listed in
-Table \ref{tab:list}, performed to verify that the code solves correctly a wide range of problems. Description and results of all benchmarks are shown
-in Section \ref{sec:benchmark}.
-
-\begin{table}[H]
-\caption{List of benchmarks performed for each feature of the code. Data of all benchmarks can be found at 
-\url{https://github.com/aleregorda/Benchmarks}}
-\centering
-\begin{tabular}{| l | l | l |}
-\toprule
- \textbf{Feature}  & \textbf{Benchmark} & \textbf{Section} \\
-\midrule
-  Solver  & Stokes flow & \ref{sec:stokes} \\
-\hline
-  \multirow{2}{*}{Markers advection} & Zalesak disk & \ref{sec:runge} \\
-    & Conservative Interpolation Velocity & \ref{sec:CVI} \\
-\hline
-  \multirow{4}{*}{Momentum equation}  & Poiseuille flow & \ref{sec:poiseuille}\\
-    & Instantaneous 2D sphere & \ref{sec:ist_sphere} \\
-    & Rayleigh-Taylor experiment & \ref{sec:rayleigh} \\
-    & Falling block & \ref{sec:block} \\
-\hline
-    & 2D Stokes sphere & \ref{sec:sphere} \\
-    Sticky air and  & Stabilisation algorithm & \ref{sec:stab} \\
-    free surface & Topography relaxation & \ref{sec:crameri} \\
-  & Spontaneous subduction & \ref{sec:subduction} \\
-\hline
-  Erosion and sedimentation  & - & - \\
-\hline
-   Non-linear  & Slab detachment & \ref{sec:slab} \\
-   visco-plasticity & Indenter & \ref{sec:indenter} \\
-    & Brick & \ref{sec:brick} \\
-\hline
-  \multirow{3}{*}{Energy equation}  & Advection stabilisation & \ref{sec:advection} \\
-    & Simple shear heating & \ref{sec:simple_shear} \\
-    & Shear and adiabatic heating & \ref{sec:shear} \\
-\hline
-\multirow{3}{*}{Energy + momentum} & Mantle convection & \ref{sec:mantle} \\
-& Viscoplastic mantle convection & \ref{sec:visco_mantle} \\
-& Thin layer entrainment & \ref{sec:thin} \\
-\hline
-  Phase changes  & \multirow{2}{*}{Hydrated sinking cylinder} & \multirow{2}{*}{\ref{sec:quinquis}} \\
-  and hydration  &  & \\
-\hline
-  Melting & Experimental melting curves & \ref{sec:katz} \\
-\bottomrule
-\end{tabular}
-\label{tab:list}
-\end{table}
-
-The code is written in Fortran90 and uses the finite element method (FEM) with quadrilateral $Q_1 \times P_0$ elements (continuous bilinear velocity and
-discontinuous constant pressure), associated with MUMPS\footnote{\url{http://mumps.enseeiht.fr/}} \citep{Amestoy2001,Amestoy2019} that is a software package
-for solving systems of linear equations of the form $A \cdot x = b$, where \textit{A} is a square sparse matrix, by means of a direct method. Since
-$Q_1 \times P_0$ elements do not satisfy Ladyzhenskaya, Babuska and Brezzi (LBB) stability condition \citep{Donea2003}, elemental pressure is elaborated
-in post-processing to avoid spurious pressures by means of a double interpolation to smooth the pressure. Following this procedure, the elemental pressure
-is interpolated onto nodes and then back onto elements. After the smoothing procedure, the pressure field is calculated again on the nodes to be used in
-combination with the lithostatic pressure, which is calculated onto nodes. Correctness of the solution and performances of the solver in terms of time and
-memory usage are tested solving the Stokes flow with the analytical solution proposed by \citet{Donea2003} (Section \ref{sec:stokes}).
-
-The thermo-mechanics of crust-mantle systems is described by means of the conservation of mass, momentum and energy equations, expressed as follows:
-\begin{eqnarray}
-\label{eq:mass}\frac{\partial \rho}{\partial t}+\bm{\nabla} \cdot (\rho \bm{u})&=&0\\
-\label{eq:momentum}-\bm{\nabla} p + \bm{\nabla} \cdot \bm{\tau} + \rho \bm{g}&=&0\\
-\label{eq:energy}\rho C_p \left(\frac{\partial T}{\partial t} + \bm{u} \cdot \bm{\nabla} T \right)&=&\bm{\nabla} \cdot \left(k\bm{\nabla} T\right) + H_{tot}
-\end{eqnarray}
-where $\rho$ is the density, $\bm{u}$ is the velocity, \textit{p} is the pressure, $\bm{\tau}$ is the deviatoric stress, $\bm{g}$ is the gravity acceleration,
-$C_p$ is the specific heat at constant pressure, \textit{T} is the temperature, \textit{k} is the thermal conductivity and $H_{tot}$ is the total internal heat
-production. Density variations due to temperature are generally small enough to assume the density as constant ($\rho=\rho_0$) in Eq. \ref{eq:mass} and in
-Eq. \ref{eq:energy}, while it must be treated as a variable in the buoyancy term of Eq. \ref{eq:momentum}, such that \[\rho=\rho_0(1-\alpha(T-T_0))\]
-where $\rho_0$ is the density at a reference temperature $T_0$ and $\alpha$ is the coefficient of thermal expansion. Eq. \ref{eq:mass} then can be rewritten as
-\begin{equation}\label{eq:mass_0}
-\bm{\nabla} \cdot \bm{u}=0
-\end{equation}
-This simplification is known as the extended Boussinesq approximation.
-
-The time step $dt$ is chosen in according to Courant-Friedrichs-Lewy condition \citep{Anderson1995}, such that
-\[dt=C_n\; \min\left(\frac{\Delta x}{\max|\bm{u}|},\frac{(\Delta x)^2}{\kappa}\right)\]
-where $0<C_n<1$ is the Courant (CFL) number, $\Delta x$ is the minimum dimension of smallest element, $\max|\bm{u}|$ is the maximum velocity calculated on the
-entire domain and
-$\kappa=\frac{k}{\rho C_p}$ is the thermal diffusivity \citep{Thieulot2014}.
-
-\section{Lagrangian markers}\label{sec:markers}
-Elemental properties (density, viscosity, thermal conductivity, specific heat and thermal expansion) needed to solve Eqs. \ref{eq:mass_0}, \ref{eq:momentum}
-and \ref{eq:energy} are related to the composition of each element, determined by means of Lagrangian markers that are characteristic of different materials
-in the domain. Elemental properties (with the sole exception of the viscosity) are calculated using an arithmetic mean, as \[P_e=\frac{1}{n}\sum_{i=1}^n P_i\]
-where $P_e$ is the elemental property, $P_i$ is the property characteristic of the material of each marker and $n$ is the number of markers of the element.
-Differently, the average scheme for the elemental viscosity can be chosen between harmonic, geometric and arithmetic mean. Markers can be placed regularly or
-randomly at the beginning of the simulation and their advection is performed by either a 2nd-order or a 4th-order Runge-Kutta in space, interpolating the velocity 
-field on each marker by means of the shape functions (see Section \ref{sec:runge}). The interpolated velocity is then corrected by means of the Conservative
-Velocity Interpolation (CVI), which introduces a corrective term to reduce dispersion and clustering of particles in both steady state and time-dependent 
-\citep{Wang2015} (see Section \ref{sec:CVI}).
-
-The initial distribution of the markers is created using the open-source code library
-Geodynamic World Builder \citep{Fraters2019}. During the simulation, each marker carries memory of temperature, pressure and accumulated strain, which are
-determined interpolating the relative nodal parameter. In particular, pressure is interpolated onto each marker after the smoothing procedure. The number
-of markers contained in each element is maintained between $n_{min}$ and $n_{max}$. When in an element there are less markers than $n_{min}$ the code adds
-random markers to reach the $n_{min}$, while if the number is higher than $n_{max}$ some of them are randomly deleted. When new markers are added, they assume
-the properties of the nearest marker. In this way elements are never empty and maintain a number of markers inside a prefixed range.
-
-\section{The momentum equation}\label{sec:mom_eq}
-The penalty method is implemented so that Eq. \ref{eq:mass_0} can be written as:
-\begin{equation}\label{eq:mass_penalty}
-\bm{\nabla} \cdot \bm{u} + \frac{p}{\lambda}=0
-\end{equation}
-where $\lambda$ is the so-called penalty parameter, which should be 6-7 orders of magnitude larger than the shear viscosity to ensure that mass conservation
-is satisfied \citep{Donea2003,Thieulot2014}.
-
-The deviatoric stress tensor in Eq. \ref{eq:momentum} can be written in terms of the strain rate tensor as $\bm{\tau}=2\eta\dot{\bm{\epsilon}}$, with
-$\dot{\bm{\epsilon}}=\frac{1}{2}\left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)$. Therefore, Eq. \ref{eq:momentum} can be rewritten as:
-\begin{equation}\label{eq:momentum_pressure}
--\bm{\nabla} p + \bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right) + \rho \bm{g} = 0
-\end{equation}
-Finally, using pressure from Eq. \ref{eq:mass_penalty}, Eq. \ref{eq:momentum_pressure} can be rewritten as:
-\begin{equation}\label{eq:momentum_penalty}
-\lambda \bm{\nabla} \left(\bm{\nabla} \cdot \bm{v}\right)+\bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right)+\rho\bm{g}=0
-\end{equation}
-The penalty method is been associated to the iterative Uzawa method, as described in detail in \citet{Dabrowski2008} and \citet{Thieulot2014}
-(see Section \ref{sec:poiseuille}). Results of benchmarks performed to verify the correctness in the implementation of Eq. \ref{eq:momentum_penalty} are
-shown in Sections \ref{sec:ist_sphere} and \ref{sec:rayleigh}.
-
-To support large viscosities variations the penalty parameter is related to the effective elemental viscosity by means of a dimensionless coefficient, so that
-$\lambda=\lambda_e(e)\eta_{eff}(e)$ \citep{Marotta2006,Dabrowski2008,Thieulot2014}. Benchmark of the falling block \citep{Gerya2003a,Gerya2010b,Thieulot2011}
-is performed to verify that the code can correctly deal buoyancy driven flows with strong viscosity contrasts (Section \ref{sec:block}).
-
-\section{Sticky air and free surface}\label{sec:sticky}
-The Earth's surface can be treated by means of either the so-called sticky-air or a true free surface method, both of them implemented in the code. In the
-sticky-air method the surface is approximated with the introduction of a buoyant layer with a viscosity at least four orders of magnitude lower than the crust
-\citep{Schmeling2008,Crameri2012} and the interface between lithosphere and air is defined using a chain of passive markers that are advected as the Lagrangian
-markers. The correctness of the evolution of the markers chain is tested with the experiment of a 2D time-dependent Stokes sphere below a free surface and
-compared with results from ASPECT\footnote{\url{https://aspect.geodynamics.org/}} \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} (Section
-\ref{sec:sphere}). In the true free surface case the top boundary is assumed stress-free and velocities are not fixed. In this case topography variations are
-described by vertical deformations of the mesh that depend on the velocity field of the nodes that identify the free surface, while horizontal deformations are
-not taken into account. This procedure is known as the Arbitrary Lagrangian-Eulerian (ALE) method and its implementation follows the technique described in
-\citet{Thieulot2011}. However, although the implementation of a true free surface better reproduces laboratory experiment, extremely small time steps can be
-necessary to maintain stability \citep{Kaus2010a,Quinquis2011,Thieulot2014}. Therefore, the stability algorithm proposed by \citet{Kaus2010a} is implemented
-to avoid instabilities due to high density differences at the free surface when using too large time steps. The implementation of this algorithm are tested by
-performing the experiment described by \citet{Kaus2010a} (Section \ref{sec:stab}).
-
-The topography relaxation benchmark proposed by \citet{Crameri2012} is performed to verify that the code correctly recovers topography variations in case of
-both the sticky-air and the true free surface method (Section \ref{sec:crameri}). Finally, Section \ref{sec:subduction} show the results of the spontaneous
-subduction experiment described by \citet{Schmeling2008}.
-
-\section{Erosion and sedimentation}\label{sec:erosion}
-Surface processes at either the lithosphere-air interface or in correspondence of the free surface have been implemented by means of the software
-Fastscape\footnote{\url{https://fastscape.org/}} \citep{Braun2013,Cordonnier2019,Yuan2019a,Yuan2019b}. FastScapeLib is a set of routines that solve the stream
-power law (enriched by a sediment transport/deposition term) \citep{Yuan2019a}, hillslope diffusion and marine transport and deposition \citep{Yuan2019b}.
-The partial differential equation solved by FastScapeLib is:
-\begin{equation}
-\frac{\partial h}{ \partial t}=U-K_f A^{m_e}S^{n_e} + \frac{G}{A}\int_A \left( U-\frac{\partial h}{\partial t} \right) dA + K_d \bm{\nabla}^2 h 
-\end{equation}
-where $h$ is the topography, $U$ is the uplift, $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $S$ is the slope, $m_e$ and $n_e$ are
-parameters that should be chosen to constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}, $G$ is the dimensionless
-deposition/transport coefficient for the enriched stream power law and $K_d$ is the effective diffusivity. Terms $K_f A^{m_e}S^{n_e}$ and $K_d \bm{\nabla}^2 h$
-consider long-range processes, described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, and short-range processes,
-proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}.
-
-In case of sticky air, slope is determined by means of the markers chain, which is vertically corrected considering erosion and sedimentation rates. After the
-correction, continental markers above the markers chain are transformed in air markers and air markers below the markers chain become sediments. Similarly,
-slope is determined considering the nodes on the top boundary in case of true free surface and they are vertically corrected as for the markers chain. In this
-case, continental markers above the top boundary are simply deleted.
-%Surface processes at the lithosphere-air interface are implemented taking into account fluvial erosion (long-range processes), hillslope processes (short-range
-%processes) and failure of slopes exceeding a threshold angle $\theta$ \citep{Burov1997,Tucker1999,Sacek2012,Croissant2014,Ueda2015,Beucher2020}. Erosion rates
-%$\dot{E}_{lr}$ due to long-range processes can be described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{lr}=K_f A^{m_e} \bm{\nabla} h^{n_e}
-%\end{equation}
-%where $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $\bm{\nabla} h$ is the slope and $m_e$ and $n_e$ are parameters chosen in to
-%constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}. Erosion rates $\dot{E}_{sr}$ due to short-range processes are
-%implemented by assuming that transport is linearly proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{sr}=K_d \bm{\nabla}^2 h
-%\end{equation}
-%where $K_d$ is the effective diffusivity. In addition, the slope is modified if it still exceeds a threshold angle after long- and short-range processes
-%\citep{Ueda2015}.
-
-%Offshore sedimentation processes are implemented similarly to hillslope processes but with a depth-dependent coefficient \citep{Sacek2012,Beucher2020}.
-%The sedimentation rate $\dot{S}$ is calculated as
-%\begin{equation}
-%\dot{S}=K_s \bm{\nabla}^2 h
-%\end{equation}
-%with
-%\begin{equation}
-%K_s=K^*_s \exp\left(\frac{h-h_{sl}}{h_0}\right)
-%\end{equation}
-%where $K^*_s$ is a constant, $h_{sl}$ is the sea level and $h_0$ is a characteristic length. In case of sticky air, slope is determined by means of the markers
-%chain, which is vertically corrected considering erosion and sedimentation rates. After the correction, continental markers above the markers chain are
-%transformed in air markers and air markers below the markers chain become sediments. Similarly, slope is determined considering the nodes on the top boundary
-%in case of true free surface and they are vertically corrected as for the markers chain. In this case, continental markers above the top boundary are simply
-%deleted.
-
-\section{Non-linear rheologies}\label{sec:nonlinear}
-Non-linear rheologies are implemented combining viscous creep (dislocation and diffusion) and plastic yielding. For each marker, diffusion and dislocation
-viscosity ($\eta_{df|ds}$) can be determined as follows \citep{Karato1993,Warren2008b,Wang2016}:
-\begin{equation}\label{eq:viscous}
-\eta_{df|ds}=f_s \left(\frac{d^m}{A}\right)^{\frac{1}{n}} \left (\frac{2^{\frac{1-n}{n}}}{3^{\frac{n+1}{2n}}} \right)
-I_2^{\frac{(1-n)}{n}}\exp\left(\frac{Q+pV}{nRT}\right)
-\end{equation}
-where $f_s$ is a scaling factor used to represent lithologies that are stronger or weaker than the base set, $d$ is the grain size, $m$ is the grain size exponent,
-$n$ is the stress exponent, $A$ is the uniaxial pre-exponential factor, $I_2$ is the square root of the second invariant of the strain rate tensor, $Q$ is the
-activation energy, $p$ is the pressure, $V$ is the activation volume, $R$ is the gas constant and $T$ is the temperature.
-Pressures and temperatures are determined by the interpolation of the nodal parameters, while strain rates are calculated by means of the
-derivative of the velocities on the nodes. In case of diffusion creep $n=1$ and $m>0$, in case of dislocation creep $n>1$ and $m=0$. The viscous creep $\eta_{cp}$
-is then calculated as the harmonic average between $\eta_{df}$ and $\eta_{ds}$:
-\begin{equation}\label{eq:visco_creep}
-\eta_{cp}=\left(\frac{1}{\eta_{df}}+\frac{1}{\eta_{ds}}\right)^{-1}
-\end{equation}
-The implementation of the non-linear viscous creep viscosity depending on the strain rate is tested by means of the slab detachment benchmark
-(Section \ref{sec:slab}).
-
-Plastic yielding is implemented rescaling $\eta_{cp}$ in order to limit the stress below the yield stress $\sigma_y$ \citep{Thieulot2008,Thieulot2014,Glerum2018},
-obtaining
-\begin{equation}\label{eq:plastic}
-\eta_{pl}=\frac{\sigma_y}{2I_2}
-\end{equation}
-where $\eta_{pl}$ is the plastic viscosity and the yield stress is determined following the Drucker-Prager criterion, such as
-\begin{equation}\label{eq:yield}
-\sigma_y=C\;\cos(\phi)+p\;\sin(\phi)
-\end{equation}
-where $C$ is the cohesion and $\phi$ is the internal friction angle. In case of negative pressure it is imposed equal to 0, so that negative yield stress are
-excluded. The correctness of the non-linear solution in case of plastic viscosity with variable internal friction angle is verified performing indenter and
-brick experiments (Sections \ref{sec:indenter} and \ref{sec:brick}, respectively).
-
-Strain softening is taken into account for both viscous creep and plastic viscosity \citep{Huismans2003,Babeyko2005,Huismans2005,Sobolev2005,Warren2008a} by
-means of the accumulated strain $\epsilon$ memorised by each marker. Softening in the viscous creep determines a linear decrease of $\eta_{vc}$ by means of a
-viscous strain softening factor $W_S$ that increases linearly from $W_{S_0}$ to $W_{S_{\infty}}$ for $\epsilon_{S_0}<\epsilon_S<\epsilon_{S_{\infty}}$ 
-\citep{Warren2008a}. This viscous softening can be related to strain-induced grain size reduction \citep{Warren2008a}. Differently, plastic softening is 
-simulated with a linear decrease of internal friction angle $\phi(\epsilon)$ and cohesion $C(\epsilon)$ in according to
-\begin{eqnarray}
-\label{eq:friction1}\phi(\epsilon)&=&\phi_0+(\phi_{\infty}-\phi_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}\\
-\label{eq:friction2}C(\epsilon)&=&C_0+(C_{\infty}-C_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}
-\end{eqnarray}
-where $\phi_0$, $C_0$ and $\phi_{\infty}$, $C_{\infty}$ are internal friction angle and cohesion for $\epsilon_0$ and $\epsilon_{\infty}$, respectively
-\citep{Huismans2003,Huismans2005,Warren2008a,Thieulot2014}. Plastic softening approximates deformation-induced softening of faults and brittle shear zones
-\citep{Warren2008a}.
-
-Plastic yielding and viscous creep are then combined to obtain a viscoplastic viscosity $\eta_{vp}$ as follows, assuming that they are independent processes
-\citep{Karato2008,Glerum2018}:
-\begin{equation}\label{eq:viscoplastic}
-\eta_{vp}=\min(\eta_{cp},\eta_{pl})
-\end{equation}
-Finally, effective viscosity $\eta_{eff}$ is capped by the minimum and the maximum viscosity ($\eta_{min}$ and $\eta_{max}$, respectively) to avoid extremely
-low or high viscosity \citep{Glerum2018} as follows
-\begin{equation}\label{eq:effective}
-\eta_{eff}=\min(\max(\eta_{vp},\eta_{min}),\eta_{max})
-\end{equation}
-Viscous creep and plastic yielding are non-linear rheologies because of their dependence on the velocity field through pressure and strain rates. Therefore,
-the solution is determined by means of Picard-type iterations, until convergence of the velocity field \citep{Glerum2018}. The convergence is verified at each
-iteration $i$ via the nonlinear residual $\bm{\mathcal{R}}^i$ that can be determined as
-\begin{equation}\label{eq:convergence}
-\bm{\mathcal{R}}^i= \mathbb{K} (\eta_{eff}(\dot{\bm{\epsilon}}^{i-1},p^{i-1}))\cdot \bm{v}^{i-1} - \bm{f}^i
-\end{equation}
-%\begin{equation}\label{eq:convergence}
-%\chi=1-\frac{\langle (\bm{v}^i-\langle\bm{v}^i\rangle)\cdot (\bm{v}^{i+1}-\langle\bm{v}^{i+1}\rangle) \rangle}{\sqrt{\sigma^i\sigma^{i+1}}}
-%\end{equation}
-%Iterations are performed until either convergence ($\chi<tol$) or maximal number of non-linear iterations ($it_{max}$) is reached.
-where $\mathbb{K}$ is the Stiffness matrix and $\bm{f}$ is the right hand side vector \citep{Spiegelman2016,Glerum2018}.  The $L_2$-norm is extracted from
-$\bm{\mathcal{R}}^i$ and it is normalised as follows
-\begin{equation}\label{eq:convergence_res}
-\frac{||\bm{\mathcal{R}}^i||_2}{||\bm{\mathcal{R}}^0||_2}
-\end{equation}
-where $||\bm{\mathcal{R}}^0||_2$ is the $L_2$-norm of the first nonlinear iteration.
-Since often the normalised nonlinear residual drop very quickly over the first 2-3 iterations, also the $L_2$-norm of velocity and pressure residuals are
-calculated as
-\begin{equation}\label{eq:convergence_vel}
-\frac{||\Delta\bm{v}||_2}{||\bm{v}||_2}=\frac{||\bm{v}_i-\bm{v}_{i-1}||_2}{||\bm{v}_{i}||_2}
-\end{equation}
-and
-\begin{equation}\label{eq:convergence_press}
-  \frac{||\Delta p||_2}{||p||_2}=\frac{||p_i-p_{i-1}||_2}{||p_{i}||_2}
-\end{equation}
-respectively. Iterations are performed until either maximal number of non-linear iterations ($it_{max}$) is reached or all the normalised nonlinear residual, 
-the velocity residual and the pressure residual converge under a defined tolerance $tol$.
-
-\section{The energy equation}\label{sec:en_eq}
-The stabilisation of the advection term of the energy equation (Eq. \ref{eq:energy}) needed to avoid possible oscillations in the thermal solution in the case
-when advection dominates over diffusion, is implemented by means of a streamline-upwind Petrov–Galerkin (SUPG) method, for which the advection term is modified
-follows the discussion in \citet{Thieulot2011} and \citet{Thieulot2014} (see Section \ref{sec:advection} for benchmark). The total internal heat production
-$H_{tot}$ in Eq. \ref{eq:energy} includes radiogenic heating $H_r$, shear heating $H_s$ and adiabatic heating $H_a$, where
-\begin{equation}\label{eq:shear}
-H_s=2\eta \dot{\bm{\epsilon}} : \dot{\bm{\epsilon}} = 2\eta(\dot{\epsilon}_{xx}^2+\dot{\epsilon}_{yy}^2+2\dot{\epsilon}_{xy}^2)
-\end{equation}
-and
-\begin{equation}\label{eq:adiabatic}
-H_a=T \alpha \frac{Dp}{Dt} \approx -\alpha T \rho g_y v
-\end{equation}
-The correctness of the implementation and the computation of shear and adiabatic heating is verified performing an exercise with an analytical solution
-(Section \ref{sec:simple_shear}) and Exercise 9.4 in \citet{Gerya2010b} (Section \ref{sec:shear}). Furthermore, mantle convection and thin layer entrainment
-benchmarks (Sections \ref{sec:mantle} and \ref{sec:thin}, respectively) are performed to verify that the code solve correctly Eqs. \ref{eq:momentum_penalty}
-and \ref{eq:energy} in case of a temperature field characterised by an initial perturbation and with a temperature-dependent density. At last, the viscoplastic
-mantle convection benchmark is performed to check also the viscoplasticity in case of a initial perturbed temperature field (Section \ref{sec:visco_mantle}).
-
-\section{Phase transitions and hydration}\label{sec:phase}
-Variations of effective density, specific heat and coefficient of thermal expansion during the evolution of crustal and mantle materials at different
-pressure-temperature ($p$-$T$) conditions are computed by means of version 6.8.6 of Perple\_X software package \citep{Connolly2005}, similarly to the
-implementation described in \citet{Marotta2020}. By default, mineral assemblages and properties of each lithology (oceanic crust, lower and upper continental
-crust, sediments and dry or hydrated mantle) are calculated for temperatures from \SIrange{330}{1600}{\kelvin}, with increments of approximately \SI{4}{\kelvin},
-and pressures from \SIrange{0.1}{30}{\giga\pascal}, with increments of approximately \SI{0.03}{\giga\pascal}, for a total of almost 100000 points.
-
-The introduction of phase transitions produces two main effects that must be taken into account in numerical models: variations in density and release/absorption
-of latent heat \citep{Tackley2010}. Effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into
-account considering effective coefficient of thermal expansion \citep{Christensen1985,Zhong2015}. Similarly, effects of latent heat in the energy equation
-(eq \ref{eq:energy}) are taken into account by considering effective specific heat and coefficient of thermal expansion
-\citep{Christensen1985,Tackley2010,Zhong2015}. Effective density, specific heat and coefficient of thermal expansion are included in Perple\_X files and are
-assumed by each Lagrangian marker in according to its $p$-$T$ conditions.
-
-Hydration processes related to dehydration of subducting lithosphere \citep{Schmidt1998,Liu2007,Faccenda2009,Faccenda2010,Faccenda2014,Rosas2016} strongly
-influence the thermo-mechanics inside the mantle wedge, mainly because of both weakening effects on the mantle rheology and density variations in case of
-mantle serpentinization \citep{Gerya2002,Honda2003,Arcay2005,Roda2010,Regorda2017}. In the case that hydration is switched on, the amount of bound and free
-water is memorised by each marker, following the implementation of \citet{Quinquis2014}; hydration and dehydration processes are related to the amount of bound
-water of each marker and to the maximum amount of water it can transport, i.e. if the amount of bound water exceeds the maximum amount of water, the marker
-dehydrates and releases free water that can hydrates under-saturated markers. Maximum water content of each marker is determined as function of lithology and
-$p$-$T$ conditions and it is calculated using Perple\_X, in the same way than effective density, specific heat and coefficient of thermal expansion. Bound water
-is advected together with the markers, neglecting the effect of bound water diffusion, while free water simply migrates vertically and is not coupled to the
-solid-phase flow of the mantle wedge \citep{Arcay2005,Quinquis2014}. The correctness of the migration scheme of the free water and its absorption into bound
-water by under-saturated markers are verified performing the experiment described in \citet{Quinquis2014} (Sec. \ref{sec:quinquis}).
-
-The mantle viscosity weakening related to the amount of water has been extensively studied \citep{Chopra1981,Hirth2003} and it is implemented in according to
-\citet{Arcay2005} and \citet{Horiuchi2016}, as follows
-\begin{equation}\label{eq:wet_visc}
-\eta_{wet}=\eta_{dry}\left(\left[1-\frac{1}{f_v}\right]\exp\left(-\frac{[OH^-]}{[OH^-]_0}\right)+\frac{1}{f_v}\right)
-\end{equation}
-where $f_v$ is the viscosity reduction factor between dry and wet conditions, $[OH^-]$ is the water content and $[OH^-]_0$ is a reference water content set to
-620 ppm (0.062 wt.\%) \citep{Arcay2005}. By default, $f_v$ is set to 100, in according to \citet{Arcay2005} and \citet{Horiuchi2016}.
-
-\section{Mantle melting}\label{sec:melt}
-Melt fraction in the mantle depends on temperature ($T$, in \textdegree C), pressure ($p$, in \si{\giga\pascal}) and the water content in the melt ($X_{H_2O}$, in wt.\%).
-The determination of the melt fraction $M$ is implemented as explained by \citet{Katz2003}, according to the successive modification by \citet{Langmuir2006} and
-\citet{Kelley2010}, as follows
-\begin{eqnarray}\label{eq:melt_fr}
-M(p,T,X_{H_2O})&=&-T+T_{sol}+(x_t\cdot ln(p+y_t))\cdot X_{H_2O}-\nonumber\\
-&&+K_m\left(\frac{X_{H_2O}^{bulk}}{D_{H_2O}(1-X_{H_2O})+X_{H_2O}}\right)^{\gamma}
-\end{eqnarray}
-where $T_{sol}$ is the temperature of the dry solidus, $D_{H_2O}$ is the partition coefficient and $x_t\cdot log(p+y_t)$ indicates the pressure dependence of
-the melting curve \citep{Kelley2010}. $x_t$, $y_t$, $K_m$ and $\gamma$ are parameters chosen in according to \citet{Kelley2010}.
-However, $X_{H_2O}$ depends on the melt fraction as
-\begin{equation}\label{eq:melt_wt}
-X_{H_2O}(M)=\frac{X_{H_2O}^{bulk}}{D_{H_2O}+M(1-D_{H_2O})}
-\end{equation}
-and a numerical solution can be found using a root-finder method \citep{Katz2003,Wang2016}. Furthermore, the content of water in the melt is limited by the
-pressure-dependent saturation concentration of water in the melt, determined as
-\begin{equation}\label{eq:melt_max}
-X_{H_2O}^{sat}=\chi_1 p^{\lambda_m}+\chi_2 p
-\end{equation}
-where $\chi_1$, $\chi_2$ and $\lambda_m$ are parameters chosen according to \citet{Katz2003}, \citet{Langmuir2006} and \citet{Kelley2010}. The correctness in
-the determination of melt fractions and water contents in the melt are tested comparing model results with experimental curves obtained by \citet{Katz2003}
-(Sec. \ref{sec:katz}).
-
-In case of melting, viscous creep viscosity ($\eta_{vp}$) is modified according to \citet{Wang2016}, as
-\begin{equation}\label{eq:melt_visc}
-\eta_{melt}=\eta_{vp}\cdot \exp(\alpha_m M)
-\end{equation}
-where $\alpha_m$ are melt fraction factors for dislocation and diffusion creep, chosen in according to \citet{Wang2016}.
-
-Effective density ($\rho_{eff}$) for partially molten rocks is calculated as
-\begin{equation}\label{eq:melt_dens}
-\rho_{eff}=\rho_s(1-M)+\rho_m M
-\end{equation}
-where $\rho_s$ and $\rho_m$ are the densities of the solid and the molten rock, respectively \citep{Gerya2004b,Gerya2007,Wang2016}. The density of the
-solid rock is extracted by Perple\_X, while the density of the molten portion is calculated as
-\begin{equation}\label{eq:melt_dens_M}
-\rho_m(p,T)=\rho_0[1-\alpha(T-T_0)][1+\beta(p-p_0)]
-\end{equation}
-where $\rho_0$ is the density at the reference temperature ($T_0$) and pressure($p_0$), and $\alpha$ and $\beta$ are the thermal expansion and compressibility
-coefficients, respectively \citep{Gerya2004b,Gerya2007,Wang2016}.
-
-As for the phase transitions, effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into account
-considering effective coefficient of thermal expansion. Similarly, latent heat due to melting/crystallisation can be implicitly included in the energy equation
-(Eq. \ref{eq:energy}) by considering effective specific heat ($C_{p_{eff}}$) and thermal expansion ($\alpha_{eff}$) of the partially molten rocks, in according to
-\begin{eqnarray}
-\label{eq:cp_effective}C_{p_{eff}}&=&C_p+H_L\left(\frac{\partial M}{\partial T}\right)_p \\
-\label{eq:alpha_effective}\alpha_{eff}&=&\alpha+\rho\frac{H_L}{T}\left(\frac{\partial M}{\partial p}\right)_T
-\end{eqnarray}
-where $C_p$ and $\alpha$ are the specific heat and the coefficient of thermal expansion of the solid rock, respectively, and $H_L$ is the latent heat of the
-molten rock \citep{Gerya2004b,Gerya2007,Tackley2010}.
-
-\section{Benchmarks}\label{sec:benchmark}
-\subsection{Error measurements}\label{sec:error}
-In order to determine the accuracy of velocity and pressure field of the benchmarks, the $L_2$-norm is computed by numerical integration on the Gauss-Legendre
-quadrature points. $L_2$-norm for pressure and velocity errors can be evaluated as
-\begin{eqnarray}
-\textrm{err}_p&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|p^n_i(r_q)-p^a_i(r_q)|^2\right)w_q|J_q|}\nonumber \\
-\textrm{err}_v&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|u^n_i(r_q)-u^a_i(r_q)|^2+|v^n_i(r_q)-v^a_i(r_q)|^2\right)w_q|J_q|}\nonumber
-\end{eqnarray}
-respectively, where $n_e$ is the number of elements, $n_q$ is the number of quadrature points per element, $p^n_i(r_q)$ and $p^a_i(r_q)$ are the numerical and
-analytical pressure, respectively, in each quadrature point $q$, $w_q$  and $J_q$ are the weight and the Jacobian at the quadrature point
-$q$, $u^n_i(r_q)$, $v^n_i(r_q)$, $u^a_i(r_q)$ and $v^a_i(r_q)$ are the numerical and analytical velocities, respectively, in each quadrature point $q$.
-
-Other quantities used as comparison with original benchmarks are the root-mean-square velocity over the whole domain and over the surface
-\begin{eqnarray}
-v_{\textrm{rms}}&=&\sqrt{\int_{\Omega}|\bm{v}|^2 d\Omega}\nonumber \\
-v_{\textrm{rms}}^{top}&=&\sqrt{\int_0^1 u^2 \bigg|_{y=1} dx}\nonumber
-\end{eqnarray}
-respectively, top and bottom Nusselt numbers
-\[\textrm{Nu}^{top/bottom}=-\int_0^1\frac{\partial T}{\partial y} \bigg|_{y=1/y=0} dx\]
-the average rate of work done against gravity
-\[\left\langle W \right\rangle=\int_{\Omega}T u_y d\Omega\]
-the average rate of viscous dissipation
-\[\left\langle \Phi \right\rangle=\int_{\Omega}\tau_{ij}\dot{\epsilon}_{ij} d\Omega\]
-and the percentage error between $\left\langle W \right\rangle$ and $\left\langle \Phi \right\rangle/Ra$
-\[\delta=\frac{\left|\left\langle W \right\rangle -\frac{\left\langle \Phi \right\rangle}{Ra} \right|}{\max \left(\left\langle W \right\rangle,
-\frac{\left\langle \Phi \right\rangle}{Ra}\right)}\times100\]
-
-\subsection{Stokes flow}\label{sec:stokes}
-The problem consists of determining velocity field $(u,v)$ and pressure $p$ in case of a manufactured solution with prescribed body forces such as
-\begin{eqnarray}
-b_1&=&(12-24y)x^4+(-24+48y)x^3+(-48y+72y^2-48y^3+12)x^2+\nonumber \\
-&&+(-2+24y-72y^2+48y^3)x+1-4y+12y^2-8y^3\nonumber \\
-b_2&=&(8-48y+48y^2)x^3+(-12+72y-72y^2)x^2+(4-24y+48y^2-48y^3+\nonumber \\
-&&+24y^4)x-12y^2+24y^3-12y^4\nonumber
-\end{eqnarray}
-for which the exact solution is:
-\begin{eqnarray}
-u(x,y)&=&x^2(1-x)^2(2y-6y^2+4y^3)\nonumber \\
-v(x,y)&=&-y^2(1-y)^2(2x-6x^2+4x^3)\nonumber \\
-p(x,y)&=&x(1-x)-\frac{1}{6}\nonumber
-\end{eqnarray}
-The domain is a unit square a constant viscosity ($\eta=1$) and the penalty parameter is set to $\lambda=10^7$. Velocity boundary conditions are set to no slip
-$(\bm{v}=\bm{0})$ on all boundaries. The problem is performed for different grid resolution between \num{8x8} and \num{1024x1024} elements.
-
-The errors between the analytical solution and the numerical prediction of the velocity and the pressure field are calculated as in Section \ref{sec:error}.
-Fig. \ref{fig:errors}a shows that both velocity and pressure field converge to the exact solution with the decrease of the element size, following the
-theoretical convergence rate. The convergence of pressure, in contrast with the observation by \citet{Donea2003} for $Q_1 \times P_0$ elements, indicate the
-effectiveness of the smoothing procedure. Smoothed pressure field for grid resolution of \num{128x128} elements is shown in Fig. \ref{fig:errors}b (continuous
-red line), in comparison with the analytical solution (dashed blue line). Solve times and memory usage needed to generate the solution are shown in
-Fig. \ref{fig:MUMPS}. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Solver/Stokes_Flow}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/errors.pdf}
-\caption{Velocity and pressure error for the Stokes flow experiment between generated and analytical solution as a function of element size (panel a),
-and comparison between smoothed pressure and analytical solution as function of $x$ coordinate for a grid resolution of \num{128x128} elements (panel b).}
-\label{fig:errors}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/MUMPS.pdf}
-\caption{Panel a) Analysis, factorisation, solution and total solve times (panel a) and factorisation memory usage (panel b) as a function of the total number
-of degrees of freedom for the Stokes flow experiment.}
-\label{fig:MUMPS}
-\end{figure}
-
-\subsection{Markers advection}\label{sec:runge}
-The 2nd-order Runge-Kutta advection scheme in space is tested by means of the Zalesak disk test \citep{Zalesak1979,Thieulot2014}. The benchmark is performed
-in a unit square domain with a grid resolution of \num{32x32} elements and 1000000 markers and values of Courant number between 0.25 and 1.
-At $t=0$, the disk is centred at position $(0.5;0.75)$ with a radius $R=0.15$ and has a vertical fissure 0.05 wide and 0.2 high (orange in Fig. \ref{fig:runge})
-and with the velocity field prescribed in the entire domain as
-\begin{eqnarray}
-u(x,y)&=&2\pi\left(y-\frac{L_x}{2}\right)\nonumber \\
-v(x,y)&=&-2\pi\left(x-\frac{L_x}{2}\right)\nonumber
-\end{eqnarray}
-the markers are back to their initial location after a $2\pi$ rotation ($t=1$). The clockwise rotation of the disk at $t=0.25$, 0.5 and 0.75 
-is shown in Fig. \ref{fig:runge} (red, blue and green, respectively). The distance from the centre of 1 marker is calculated throughout the rotation to
-evaluate the error for different values of Courant number. As expected, the error increase with the increasing of the time step, as shown in 
-Fig. \ref{fig:runge_err}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/Zalesak_Disk}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=200px]{./Figures/Zalesak.pdf}
-\caption{Position of the Zalesak disk at $t=0$, 0.25, 0.5 and 0.75 (orange, red, blue and green, respectively) for the case with Courant number of 1.}
-\label{fig:runge}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/zal_marker.pdf}
-\caption{Distance from the centre as function of time for values of Courant number of  of 0.25, 0.3, 0.5, 0.75 and 1 (green, orange, blue, black and red,
-respectively). R indicates the distance from the centre considering a perfect circle trajectory.}
-\label{fig:runge_err}
-\end{figure}
-
-\subsection{Conservative Velocity Interpolation (CVI)}\label{sec:CVI}
-The CVI correction is checked by means of the Stokes flow experiment by \citet{Donea2003}, as presented in \ref{sec:stokes}. The advection of Lagrangian
-markers is performed using either a 2nd-order or a 4th-order Runge-Kutta scheme with CFL=0.5 and an initial random distribution of 25 markers per element.
-Fig. \ref{fig:CVI} shows the minimum and maximum values of markers per element throughout the experiment with and without the CVI correction and using
-different Runge-Kutta schemes. The introduction of the CVI correction allows the
-tests to run for times longer than $t=2000$ without show empty elements, while tests without the CVI correction stop at approximately $t=600$, when at least
-one element goes empty. Fig. \ref{fig:CVI_time} shows the different number of markers per element and the markers distribution at $t=600$ obtained without and
-with the CVI correction (panels b and c, respectively) in case of a 2nd-order Runge-Kutta scheme. All data can be found at
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/CVI}.
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/CVI.pdf}
-\caption{Maximum and minimum number of markers per element throughout the experiment, with (green and orange lines) or without (purple and blue lines)
-the CVI correction and using either a 2nd-order (blue and orange lines) or a 4th-order (purple and green lines) Runge-Kutta scheme.}
-\label{fig:CVI}
-\end{figure}
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/CVI_time.pdf}
-\caption{Number of markers per element and markers distribution through time using 2nd.order Runge-Kutta scheme.
-Panels a: number of markers per element (panel a1) and distribution of the markers (panel a2) after the first
-iteration; panels b and c: comparison between number of markers per element (panels b1 and c1) and markers distribution (panels b2 and c2) without and with
-the CVI correction at $t=600$; panels d: number of markers per element (panel d1) and markers distribution (panel d2) in case of CVI correction at $t=2000$.}
-\label{fig:CVI_time}
-\end{figure}
-
-\subsection{Poiseuille flow}\label{sec:poiseuille}
-The domain is a rectangle with $L_x=2$ and $L_y=1$ and constant density and viscosity ($\rho=1$ and $\eta=1$), gravity acceleration $\bm{g}=\bm{0}$ and penalty
-parameter $\lambda=10^8$. The grid is composed by \num{40x20} elements. Velocity boundary conditions are set to no slip $(\bm{v}=\bm{0})$ at the top and the
-bottom, and a parabolic profile is imposed on the sides, with $u=y(1-y)$ and $v=0$. The analytical solution is then given by:
-\begin{eqnarray}
-u(x,y)&=&y(1-y)\nonumber \\
-v(x,y,)&=&0\nonumber \\
-p(x,y)&=&2\eta\left(\frac{L_x}{2}-x\right)\nonumber
-\end{eqnarray}
-
-The velocity field predicted by the model follows the expected parabolic profile (Fig. \ref{fig:poiseuille}a and Fig. \ref{fig:poi_plot}a) and, in case of the
-classic penalty method with no iterations, the pressure is clearly related to the divergence by means of the penalty parameter (Fig. \ref{fig:poiseuille}b and
-c and Fig. \ref{fig:poi_plot}b). Fig. \ref{fig:poiseuille}d shows that one Uzawa iteration is sufficient to bring the divergence down to \num{e-15}, with no
-correlation with the pressure field. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Poiseuille%20Flow}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Poiseuille.pdf}
-\caption{Velocity field (panel a), pressure (panel b) and divergence velocity of a Poiseuille flow in case of the classic penalty method (no iterations) and
-after one Uzawa iteration (panel c and d, respectively).}
-\label{fig:poiseuille}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Analytical.pdf}
-\caption{Velocity (panel a) and pressure (panel b) field predicted by the model for a Poiseuille flow with respect with their analytical solutions. The
-velocity field is plotted as function of the vertical coordinate in $L_x/2$ and the pressure is plotted as function of the x coordinate in $L_y/2$.}
-\label{fig:poi_plot}
-\end{figure}
-
-\subsection{Instantaneous 2D sphere}\label{sec:ist_sphere}
-The domain is a unit square with gravity $\bm{g}=(0,-1)$. The fluid has constant density and viscosity ($\rho_f=1$ and $\eta_f=1$). The sphere is in the middle
-of the domain with a radius R = 0.123456798 and has constant density and viscosity ($\rho_s=10^{-2}+\rho_f$ and $\eta_s=10^3 \cdot \eta_f$).
-We distinguish three different types of velocity boundary conditions:
-\begin{enumerate}
-\item FS: free slip conditions on all sides.
-\item NS: no slip conditions on all sides.
-\item OT: free slip conditions on the sides and the bottom, and open boundary at the top.
-\end{enumerate}
-All three cases are performed for different grid resolutions, between \num{16x16} and \num{512x512} elements with 50 randomly distributed markers per element,
-and different average schemes for the viscosity (harmonic, geometric and arithmetic). The velocity in the centre of the sphere $\bm{v}(0.5;0.5)$), minimum and
-maximum velocities ($u_{min}$, $u_{max}$, $v_{min}$, $v_{max}$) and pressures ($p_{min}$, $p_{max}$) on the entire domain, $v_{\textrm{rms}}$ and average pressure on the
-entire domain ($p_{avg}$) are compared with solutions generated by ASPECT \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and all results
-can be found at \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere2D} and 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Instantaneous_2D_Sphere}. Results in terms of the $v_{\textrm{rms}}$ for NS model 
-are also shown in Fig. \ref{fig:inst_sphere}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_NS.pdf}
-\caption{$v_{\textrm{rms}}$ as function of element size for the instantaneous 2D sphere experiment in case of no slip boundary conditions and with different average
-schemes for the viscosity. Results are compared with results obtained by other numerical codes.}
-\label{fig:inst_sphere}
-\end{figure}
-
-\subsection{Rayleigh-Taylor instability}\label{sec:rayleigh}
-This problem has been originally presented by \citet{vanKeken1997} and here it is performed as the isoviscous case (Case 1a). The domain has $L_x=0.9142$ and
-$L_y=1$ and gravity $\bm{g}=(0,-1)$. Two fluids with same constant viscosities ($\eta_1=\eta_2=1$) and different densities ($\rho_1=1000$ and $\rho_2=1010$),
-with the lighter fluid at the bottom. The initial interface between the fluids is given by $y(x)=0.2+0.02 \cos \left(\frac{\pi x}{L_x}\right)$. The experiment
-is performed with different grid sizes (\num{50x50}, \num{80x80}, \num{100x100} and \num{256x256}). A total of 1960000 markers are randomly distributed at the
-beginning of the simulation. Velocity boundary conditions are set to no slip at the top and the bottom, and to free slip at the sides of the domain.
-
-$v_{\textrm{rms}}$ as function of time is reported in Fig. \ref{fig:RT}, matching well with results shown by \citet{vanKeken1997}, \citet{Tackley2003} and
-\citet{Thieulot2014}. Fig. \ref{fig:rayleigh} shows the evolution of the experiment at different time steps. All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Rayleigh_Taylor_experiment/ISOVISCOUS}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/RT.pdf}
-\caption{$v_{\textrm{rms}}$ of the Rayleigh-Taylor experiment as function of time for different resolution of the grid.}
-\label{fig:RT}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=300px]{./Figures/Rayleigh.pdf}
-\caption{Evolution of the Rayleigh-Taylor experiment for a grid of \num{256x256} elements at $t=0, 500, 1000$ and $2000$ (panels a, b, c and d, respectively).}
-\label{fig:rayleigh}
-\end{figure}
-
-\subsection{Falling blocks}\label{sec:block}
-This benchmark is described as proposed by \citet{Gerya2003a}, \citet{Gerya2010b} and \citet{Thieulot2011}. The domain is square with $L_x=L_y=$ \SI{500}{\km}
-and the grid is composed by \num{50x50} elements with 25 markers in each element. The block is initially centred at ($x=$ \SI{250}{\km}; $y=$ \SI{400}{\km})
-and has a size of \num{100x100} km (Fig. \ref{fig:block}a). The fluid surrounding the block has $\eta_f=$ \SI{e21}{\pascal\per\s} and $\rho_f=$
-\SI{3200}{\kg\per\cubic\m}. The benchmark tests with different viscosities of the block, with $\eta_b$ from \SIrange{e15}{e27}{\pascal\per\s}. In each
-experiment also the density $\rho_b$ of the block varies from \SIrange{3220}{9900}{\kg\per\cubic\m}. Velocity boundary conditions are set to free slip
-conditions on all sides of the domain. Density distributions at $t=$ \SI{20}{\mega\year} for $\rho_b=$ \SI{3300}{\kg\per\cubic\m} and different $\eta_b$
-are plotted in Fig. \ref{fig:block}b-f, showing that the code correctly preserve the block geometry in case of large viscosity contrast, with a block stiffer
-than the surrounding fluid (Fig. \ref{fig:block}f).
-
-The velocity in the centre of the falling block at $t=0$ is measured for all experiments and, since the velocity should increase with density contrast, the
-quantity $\bm{v}/(\rho_b-\rho_f)$ is plotted as function of the viscosity contrast. All results perfectly match with those from \citet{Gerya2010b} and they
-line up on a single curve, demonstrating that the code can correctly deal with large viscosity and density contrasts (Fig. \ref{fig:falling}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Falling%20blocks}.
-
-\begin{figure}
-\noindent\includegraphics[width=300px]{./Figures/Block.pdf}
-\centering
-\caption{Density evolution of the falling block experiment at $t=$ \SI{0}{\mega\year} (panel a) and $t=$ \SI{20}{\mega\year} for different viscosities of the
-block (panels b-f).}
-\label{fig:block}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Falling.pdf}
-\caption{Initial velocity relative to the density contrast at the centre of the falling block as function of the viscosity contrast between the block and the
-surrounding fluid.}
-\label{fig:falling}
-\end{figure}
-
-\subsection{2D time-dependent Stokes sphere with deformable free surface}\label{sec:sphere}
-This experiment is performed in a unit square domain with the gravity acceleration fixed to $g_y=-1$. The sphere has $\rho_s=2$ and $\eta_s=10^3$ and is
-initially centred at ($0.5;0.6$) with radius $R=0.123456789$. The fluid surrounding the sphere has $\rho_f=1$ and $\eta_f=1$ and occupies the domain for
-$y\leq 0.75$, while for $y>0.75$ the air has $\rho_a=0$ and $\eta_a=10^{-3}$. Velocity boundary conditions are set to free slip on all sides. The Courant
-number is set to 0.25. The experiments run for 200 s using grid resolutions from \num{150x150} to \num{512x512} elements, with 25 randomly distributed markers
-per element, and for different average schemes for the viscosity. The interface between the fluid and the air is tracked by means of the markers chain. Results
-in terms of velocity and pressure fields and topography variations are compared with results obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and they can be found at
-\url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere_fs2D} and
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Time_dependent_sphere}. Results in terms of the $v_{\textrm{rms}}$ in case of 
-an arithmetic average are also shown in Fig. \ref{fig:inst_spherefs}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_arithm.pdf}
-\caption{$v_{\textrm{rms}}$ for different grid resolutions as function of time for the 2D time-dependent Stokes sphere experiment in case of an arithmetic average.
-Results are compared with results obtained by ASPECT.}
-\label{fig:inst_spherefs}
-\end{figure}
-
-\subsection{Free surface stabilisation}\label{sec:stab}
-The benchmark is performed as discussed in \citet{Kaus2010a} and \citet{Thieulot2014}. The domain is square with $L_x=L_y=$ \SI{500}{\km} and a grid resolution
-of \num{200x200} elements, each of them containing 16 markers. A buoyant fluid with $\rho_1=$ \SI{3200}{\kg\per\cubic\metre} and $\eta_1=$ \SI{e20}{\pascal\s}
-is overlain by a denser fluid with $\rho_2=$ \SI{3300}{\kg\per\cubic\metre} and $\eta_2=$ \SI{e21}{\pascal\s}. The initial interface between the fluids has an
-initial sinusoidal shape of 5 km amplitude. Velocity boundary conditions are set to free slip conditions at the sides and no slip at the bottom, while the top
-is free surface. The experiment is performed with various fixed time steps, without the use of the Courant number. The vertical position of the free surface at
-$x=L_x$ is tracked for each simulation, with or without the stabilisation algorithm.
-
-The results show an instability (drunken sailor effect) increasing the time step in case that the stabilisation algorithm is not activated
-($dt=$ \SI{4200}{\year}, green line in Fig. \ref{fig:stab}), as already observed by \citet{Kaus2010a} and \citet{Thieulot2014}. Activating the algorithm the
-instability is fixed and the simulation remains stable using time step up to \SI{50000}{\year} (red line in Fig. \ref{fig:stab}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Stabilisation_algorithm}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Kaus_random.pdf}
-\caption{Evolution of the y coordinate in $x=L_x$ as function of time for different time steps, with and without the stabilisation algorithm.}
-\label{fig:stab}
-\end{figure}
-
-\subsection{Topography relaxation}\label{sec:crameri}
-This experiment is performed as in \citet{Crameri2012} using free surface and sticky air (Case 1). The domain is rectangular with $L_x=$ \SI{2800}{\km} and
-$L_y=$ \SI{700}{\km} in case of free surface and $L_y=$ \SI{800}{\km} in case of sticky air ($h_{st}=$ \SI{100}{\km}). The grid is composed by \num{256x64}
-elements for the free surface \num{560x320} elements for the sticky air. For both cases, there is a mantle of 600 km thickness with $\eta_m=$ \SI{e21}{\pascal\s}
-overlain by a highly viscous cosine-shaped lithosphere with thickness between 93 and 107 km and $\eta_l=$ \SI{e23}{\pascal\s}. Both mediums have $\rho=$
-\SI{3300}{\kg\per\cubic\metre}. In case of sticky air experiment, the air has $\rho_a=$ \SI{0}{\kg\per\cubic\metre} and $\eta_a=$ \SI{e18}{\pascal\s}. The
-gravity acceleration is set to $g_y=$ \SI{-10}{\metre\per\square\s}. Velocity boundary conditions are set to no slip at the bottom and free slip at the sides
-of the domain. In case of sticky air, velocities on top are set to free slip conditions.
-
-The maximum topography as function of time ($h(t)$) can be analytically derived in according to \[h(t)=h_0 \exp(\gamma t)\]
-where $h_0=7$ km is the initial topography, $\gamma=\num{-0.2139e-11}$ is the characteristic relaxation rate and $t$ is the time (black line in Fig.
-\ref{fig:crameri}). Maximum topography at the characteristic relaxation time $t_{rlx}=$ \SI{14.825}{\kilo\year} can be found to be $h_{rlx}=$ \SI{2576}{\m}.
-The results show that a viscosity of the air of \SI{e18}{\pascal\s} is needed in case of $h_{st}=$ \SI{100}{\km} to correctly describe the topography relaxation
-for this problem, as pointed out by \citet{Crameri2012} (purple line in Fig. \ref{fig:crameri}). The case with a true free surface well follows the analytical
-solution (orange line in Fig. \ref{fig:crameri}). In particular, at the relaxation time in case of the true free surface predicts a maximum topography of
-\SI{2570}{\m}, with an error of \SI{6}{\m} with respect to the analytical solution. Results are also compared with results obtained with a free surface in
-UNDERWORLD (\num{256x64} elements) and SULEC (\num{401x201} elements) (with courtesy from \citealp{Crameri2012}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Topography_relaxation}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Crameri.pdf}
-\caption{Maximum topography as function of time for the Crameri benchmark (coloured lines) in comparison with the analytical solution (black line) and results
-obtained with UNDERWORLD (yellow line) and SULEC (blue line).}
-\label{fig:crameri}
-\end{figure}
-
-\subsection{Spontaneous subduction}\label{sec:subduction}
-This experiment is performed as presented by \citet{Schmeling2008}, in case of both a sticky air and a true free surface and using different average schemes
-for the viscosities. The domain is rectangular with $L_x=$ \SI{3000}{\km} and $L_y=$ \SI{700}{\km}, a grid resolution of \num{256x64} elements with 50 markers
-per element and Courant number of 0.01. At the beginning of the simulation, a 100 km-thick lithospheric layer with $\rho_m=$ \SI{3300}{\kg\per\cubic\metre}
-and $\eta_m=$ \SI{e23}{\pascal\s} is located at the top of the domain for $x$ between 1000 and 3000 km, while the asthenospheric mantle has $\rho_m=$
-\SI{3200}{\pascal\s} and $\eta_m=$ \SI{e21}{\pascal\s}. In addition, a 200 km-depth lithospheric slab is already subducted in the mantle in order to have
-a spontaneous subduction. In case of sticky air, $L_y=750$ km and a 50 km-thick air layer with $\eta_a$ from \SIrange{e19}{e21}{\pascal\s} overlie the mantle.
-Velocity boundary conditions are set to free slip on the sides and on the bottom of the domain. In case of sticky air, velocities on the top boundary are set
-to free slip conditions as well.
-
-The maximum depth of the slab is tracked for all the simulations (Fig. \ref{fig:subduction}) and they are compared with results obtained by
-\citet{Schmeling2008} with models with regular grids and comparable grid resolutions (rectangles in Fig. \ref{fig:subduction}). Low viscous sticky air models
-(i.e., \SI{e19}{\pascal\s}) enlighten a strong dependence of the sinking velocity with the chosen average scheme (dashed lines in Fig. \ref{fig:subduction}),
-while high viscous sticky air models (i.e., \SI{e21}{\pascal\s}) show a higher resistance at the trench, underestimating the correct solution (dotted lines in
-Fig. \ref{fig:subduction}). A high resistance at the trench can be also observed in case of a true free surface for the low resolution grid used (continuous
-lines in Fig. \ref{fig:subduction}), as pointed out by \citet{Schmeling2008}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Spontaneous_subduction}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Subduction.pdf}
-\caption{Maximum depth of the slab as function of time for all the simulations of the \citet{Schmeling2008} benchmark, in case of harmonic, geometric and
-arithmetic means (red, green and blue lines, respectively) using a sticky air or a true free surface (discontinuous and continuous lines, respectively).
-Red, green and blue rectangular areas indicate the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{500}{\km} in case of sticky air,
-while the grey rectangular area indicates the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{600}{\km} in case of true free surface.}
-\label{fig:subduction}
-\end{figure}
-
-\subsection{The slab detachment}\label{sec:slab}
-The slab detachment benchmark is performed as described by \citet{Schmalholz2011} and \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{1000}{\km}
-and $L_y=$ \SI{660}{\km} and a grid resolution of \num{256x256} elements. A non-linear viscous T-shaped layer with $\rho_l=$ \SI{3300}{\kg\per\cubic\metre} is
-placed at the top of the domain and surrounded by a linear viscous fluid with $\rho_f=$ \SI{3150}{\kg\per\cubic\m}. The top layer is 80 km-thick and a 250
-km-long and 80 km-wide slab is placed at $x=L_x/2$ (Fig. \ref{fig:slab}a). The effective viscosity of the top layer is given by
-\[\eta_{eff}=\eta_0 I_2^{\frac{(1-n)}{n}}\]
-with $\eta_0=$ \SI{4.75e11}{\pascal\s} and $n=4$, while the surrounding fluid has $\eta_f=$ \SI{e21}{\pascal\s}. The effective viscosity is capped between
-$\eta_{min}=$ \SI{e21}{\pascal\s} and $\eta_{max}=$ \SI{e25}{\pascal\s}. Velocity boundary conditions are set to free slip at the top and the bottom, and to
-no slip at the sides of the domain.
-
-The sinking of the slab determines a decrease of the effective viscosity in the T-shaped layer during the first part of the evolution (Fig. \ref{fig:slab}a),
-as consequence of local strain rates, while effective viscosities increase up to $\eta_{max}$ in the top layer after approximately \SI{20}{\mega\year}, when
-necking is complete (Fig. \ref{fig:slab}b).
-
-The width of necking is tracked by means of markers position at the side of the slab. Necking width and time are normalised in function of the initial slab
-(\SI{80}{\km}) and the characteristic time ($t_c=$ \SI{7.1158e14}{\s}) \citep{Schmalholz2011,Glerum2018}. The results in case of arithmetic, geometric and
-harmonic mean of the viscosity are shown in Fig. \ref{fig:necking}, compared with results from \citet{Glerum2018} for which an infinity norm mean was used.
-The influence of the average schemes agrees with results shown in Section \ref{sec:subduction}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Slab_detachment}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Slab.pdf}
-\caption{Effective viscosity for the slab detachment benchmark at the beginning of the evolution (panel a) and when necking is complete (panel b).}
-\label{fig:slab}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Necking.pdf}
-\caption{Normalised width of the necking of the slab detachment benchmark as function of normalised time.}
-\label{fig:necking}
-\end{figure}
-
-\subsection{The indenter experiment}\label{sec:indenter}
-The indenter benchmark simulate a rigid punch on a purely plastic von Mises material, for which exists an analytical solution
-\citep{Thieulot2008,Thieulot2014,Glerum2018}. The domain is a unit square with a grid resolution of \num{256x256} elements. The gravity acceleration is set
-to $\bm{g}=\bm{0}$ and effective viscosity is capped using $\eta_{min}=10^{-4}$ and $\eta_{max}=10^3$. The tolerance for the convergence of the non-linear
-solution is $tol=10^{-9}$, with a maximum number of non-linear iterations set to 500. The medium in the domain has $\rho=0.01$, an initial viscosity
-$\eta_0=10$, cohesion $C=1$ and an angle of internal friction $\phi=0$°. Velocity boundary conditions are set to no slip at the bottom and free slip at the
-sides of the domain. The top of the domain is open with the exception of the central portion (punch area) with width $w_p=0.125$, where $v=-1.05$ and $u$ is
-fixed either to 0 (rough punch experiment) or free (smooth punch experiment).
-
-Fig. \ref{fig:indenter} show the results in terms of viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure
-(panels g and h) for both experiments. The analytical solution indicate that pressure at ($0.5;1$) and ($0.5 \pm w_p;1$)is $p=\pi +1$ and $p=1$, respectively
-(black and red lines in Fig. \ref{fig:smooth_p}). Fig. \ref{fig:smooth_p} shows a clear improvement in the pressure solution when passing from a rough to a
-smooth punch (green and blue lines in Fig. \ref{fig:smooth_p}, respectively), as previously observed by \citet{Thieulot2014} and \citet{Glerum2018}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Indenter}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Indenter.pdf}
-\caption{Viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure (panels g and h) fields for rough (left column) and
-smooth (right column) punch experiments.}
-\label{fig:indenter}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Smooth_Rough.pdf}
-\caption{Nodal pressure in function of the x coordinate at $y=1$ for rough (green line) and smooth (blue line) punch experiment. Black and red lines indicate
-the analytical solution at $x=0.5$ and $x=0.5 \pm w_p$, respectively.}
-\label{fig:smooth_p}
-\end{figure}
-
-\subsection{The brick experiment}\label{sec:brick}
-An instantaneous version of the brick benchmark is performed to verify the correctness of the pressure-dependent plasticity for different angle of internal
-friction, as proposed by \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{40}{\km} and $L_y=$ \SI{10}{\km} and a grid resolution of \num{512x128}
-elements. The gravity acceleration is set to $g_y=$ \SI{-10}{\m\square\s} and effective viscosity is capped using $\eta_{min}=$ \SI{e19}{\pascal\s} and
-$\eta_{max}=$ \SI{e26}{\pascal\s}. The tolerance for the convergence of the non-linear solution is $tol=10^{-7}$, with a maximum number of non-linear iterations
-set to 1000. A 800 m-wide and 400 m-height inclusion with $\eta_b=$ \SI{e20}{\pascal\s} and $\rho_b=$ \SI{2700}{\kg\per\cubic\m} is placed at the bottom of the
-domain at $x=L_x/2$. The inclusion is surrounded by a non-linear viscous medium with $\rho_m=$ \SI{2700}{\kg\per\cubic\m}, an initial viscosity of
-\SI{e23}{\pascal\s}, a linear viscous viscosity of \SI{e25}{\pascal\s} and a cohesion of \SI{40}{\mega\pascal}. Velocities are set to free slip conditions at
-the bottom of the domain and the top is open. Velocities on sides of the domain are fixed to $u=$ \SI{+-2e-11}{\m\per\s} and $v=0$. The experiment is performed
-in compressional and extensional contexts with an angle of internal friction $\phi$ of the non-linear medium variable between 0° and 30°.
-
-As expected, two shear bands stem from the inclusion with variable angles in relation to both the dynamics context and the internal friction angle. Shear band
-angles formed at 45° for $\phi=0$° in both extensional and compressional contexts (Fig. \ref{fig:brick_beam}a), while they are different in case of $\phi\neq0$°.
-Shear band angles in case of $\phi=20$° are shown for compression and extension in Fig. \ref{fig:brick_beam}b  and c, respectively. Values of shear band angles
-as function of internal friction angles from 0° to 30° are extracted at different depths and minimum and maximum values are plotted in Fig. \ref{fig:brick},
-in comparison with theoretical Roscoe, Arthur and Coulomb shear band angles. For all tests the tolerance for nonlinear convergence is defined $tol=10^{-7}$
-and none converges before the maximum number of iterations ($it_{max}=1000$) is reached. However, tests with internal friction angles up to 15° show a constant
-decrease of the velocity residuals, which can not be observed for for tests with higher internal friction angles (Fig. \ref{fig:convergence}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Brick_experiment}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Brick_Beam.pdf}
-\caption{Shear band angles predicted for the brick experiment for $\phi=0$° (panel a) and for $\phi=20$° in case of compression and extension (panels b and c,
-respectively).}
-\label{fig:brick_beam}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Brick.pdf}
-\caption{Shear band angles predicted for the brick experiment in case of compressional and extensional contexts as function of different internal angle of
-friction (continuous black line and red dots), compared with theoretical Roscoe, Artur and Coulomb angles (discontinuous black lines). Red lines indicate the
-range of angles calculated at different depths.}
-\label{fig:brick}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \noindent\includegraphics[width=500px]{./Figures/convergence.pdf}
-  \caption{Convergence rates for all tests. Normalised residuals (left column) are compared with velocity residuals (right column) for extensional (first row)
-  and compressional (second row) settings.}
-  \label{fig:convergence}
-  \end{figure}
-
 \subsection{Advection stabilisation}\label{sec:advection}
@@
 \end{figure}
-
-\subsection{Shear heating}\label{sec:simple_shear}
-This exercise is performed in an unit square domain composed by \num{128x128} elements. The velocity field is prescribed on the entire domain with
-$\bm{v}=(L_y-y)y\bm{e_x}$; viscosity, density and specific heat are fixed to 1, while thermal conductivity and radiogenic energy are fixed to 0. Therefore,
-the energy equation (Eq. \ref{eq:energy}) can be simplified as \[\frac{\partial T}{\partial t}=H_s\] and fixing $T(t=0)=0$, the temperature can be find as
-\[T(t)=H_s t\]
-In this case we have
-\begin{eqnarray}
-\dot{\epsilon_{xx}}&=&\frac{\partial u}{\partial x}=0 \nonumber \\
-\dot{\epsilon_{yy}}&=&\frac{\partial v}{\partial y}=0 \nonumber \\
-\dot{\epsilon_{xy}}&=&\frac{1}{2}\left(\frac{\partial v}{\partial x}+\frac{\partial u}{\partial y}\right)=\frac{1}{2}(L_y-2y)\nonumber
-\end{eqnarray}
-and, simplifying Eq. \ref{eq:shear}, shear heating can be calculated as \[H_s(x,y)=(1-2y)^2\]
-
-The solution predicted by the model in terms of velocity, temperature and shear heating match well with the analytical solutions (Fig. \ref{fig:analytical_en}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Simple_shear}.
-
-%\begin{figure}
-%\centering
-%\noindent\includegraphics[width=400px]{./Figures/Energy_Thieulot.pdf}
-%\caption{Temperature and shear heating predicted by the model after 1 time step.}
-%\label{fig:en_thieulot}
-%\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=500px]{./Figures/Analytical_Shear.pdf}
-\caption{Velocity (panel a), temperature (panel b) and shear heating (panel c) as function of the \textit{y} coordinate for the simple shear experiment. The
-solutions predicted by the model (red lines) are compared with the analytical solutions (dashed blue lines).}
-\label{fig:analytical_en}
-\end{figure}
-
-\subsection{Shear and adiabatic heating}\label{sec:shear}
-This problem is performed as presented in Exercise 9.4 in \citet{Gerya2010b}. Two materials are vertically separated in a rectangular domain with $L_x=$
-\SI{1000}{\km}, $L_y=$ \SI{1500}{\km} and a grid resolution of \num{30x20} elements. Constant thermal coefficient expansion ($\alpha=$ \SI{3e-5}{\per\kelvin}),
-temperature ($T=$ \SI{1300}{\kelvin}) and gravity acceleration ($g_y=$ \SI{-10}{\m\square\s}) are assumed in the whole domain. Fluid 1 (on the left side of the
-domain) has $\rho_1=$ \SI{3200}{\kg\per\cubic\m} and $\eta_1=$ \SI{e20}{\pascal\s}; fluid 2 (on the right side of the domain) has $\rho_2=$
-\SI{3300}{\kg\per\cubic\m} and $\eta_2=$ \SI{e22}{\pascal\s}. Velocity boundary conditions are set to free slip on all sides of the domain.
-
-As shown in Fig. \ref{fig:energy}, both shear and adiabatic heating predicted by the code (first row in Fig. \ref{fig:energy}) well recreate the results
-obtained by means of example ${Shear\_adiabatic\_heating.m}$ from \citet{Gerya2010b} (second row in Fig. \ref{fig:energy}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Adiabatic%2BShear}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Energy.pdf}
-\caption{Comparison between shear (first column), adiabatic (second column) and total (third column) energy predicted by the code (panels a, b and c) and those
-created using example ${Shear\_adiabatic\_heating.m}$ from Exercise 9.4 in \citet{Gerya2010b} (panels d, e and f).}
-\label{fig:energy}
-\end{figure}
-
-\subsection{Mantle convection}\label{sec:mantle}
-This problem is performed as presented by \citet{Blankenbach1989} (constant viscosity cases) and \citet{Thieulot2014}, in a 2D unit square domain with gravity
-acceleration $g_y=10^{10}Ra$. The experiment is performed with three different Rayleigh numbers ($Ra=10^4$, $10^5$ and $10^6$) and with different grid
-resolution (between \num{32x32} and \num{128x128} elements). The fluid has constant viscosity, initial density, heat capacity, thermal conductivity
-($\eta= \rho_0= C_p = k=1$), reference temperature ($T_0=0$) and thermal expansion coefficient ($\alpha=10^{-10}$). Temperatures are set to 0 on top and 1 on
-bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\]
-
-The solution generated by the code in terms of $v_{\textrm{rms}}$ (calculated as in Sec. \ref{sec:error}) and Nu as function of time are reported for all the
-simulations in Fig. \ref{fig:mantle}. At the steady state, $v_{\textrm{rms}}$ and Nu of all simulations converge well toward the values from \citet{Blankenbach1989},
-with lower errors for higher resolution grids (Table \ref{tab:mantle}). All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Mantle_convection}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Convection.pdf}
-\caption{$v_{\textrm{rms}}$ (panels a, c and e) and Nu (panels b, d and f) for the mantle convection benchmark as function of time for different grid resolution.
-Panels a and b show the results for $Ra=10^4$; Panels c and d show the results for $Ra=10^5$; panels e and f show the results for $Ra=10^6$. Purple lines
-indicate the convergence values for the $v_{\textrm{rms}}$ and the Nu from \citet{Blankenbach1989} at the steady state.}
-\label{fig:mantle}
-\end{figure}
-
-\begin{table}[H]
-\caption{Comparison between $v_{\textrm{rms}}$ and Nu predicted by the code for the mantle convection experiment and same values as reported in literature.}
-\centering
-\small
-\begin{tabular}{|>{\centering}m{0.09\textwidth}|>{\centering}m{0.05\textwidth}|>{\centering}m{0.24\textwidth}|>{\centering}m{0.15\textwidth}|
-  >{\centering}m{0.1\textwidth} >{\centering}m{0.1\textwidth} >{\centering\arraybackslash}m{0.1\textwidth}|}
-\toprule
-& & \multirow{2}{*}{\citet{Blankenbach1989}} & \citet{Thieulot2014} & \multicolumn{3}{c|}{FALCON}  \\
- & & & \num{200x200} & \num{32x32} & \num{64x64} & \num{128x128}  \\
-\midrule
-\multirow{2}{*}{$Ra=10^4$} & $v_{\textrm{rms}}$ & $42.864947 \pm 0.000020$ & 42.867 & 42.83226 & 42.852793 & 42.861394  \\
-           & Nu      & $4.884409 \pm 0.000010$  & 4.882  & 4.781297 & 4.857475  & 4.877573   \\
-\hline
-\multirow{2}{*}{$Ra=10^5$} & $v_{\textrm{rms}}$ & $193.21454 \pm 0.00010$  & 193.255& 193.872643 & 193.377472 & 193.252290  \\
-           & Nu      & $10.534095 \pm 0.000010$ & 10.507 & 9.602514 & 10.270735  & 10.465629   \\
-\hline
-\multirow{2}{*}{$Ra=10^6$} & $v_{\textrm{rms}}$ & $833.98977 \pm 0.00020$  & 834.712& 848.091176 & 837.767911 & 834.945793  \\
-           & Nu      & $21.972465 \pm 0.000020$ & 21.695 & 15.999266 & 19.703682  & 21.306939   \\
-\bottomrule
-\end{tabular}
-\label{tab:mantle}
-\end{table}
-
-\subsection{Viscoplastic mantle convection}\label{sec:visco_mantle}
-This benchmark is performed as Cases 1-5a described in \citet{Tosi2015} in a 2D unit square domain with two different grid resolutions of \num{32x32} 
-and \num{100x100} elements. Reference viscosity, reference density, heat capacity, thermal conductivity and thermal expansion coefficient are fixed to a constant
-value of 1 ($\eta_0= \rho_0= C_p = k=\alpha=1$), while the gravity has been chosen as $g_y=10^2$, in order to obtain a Rayleigh number $Ra=10^2$. Temperatures
-are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\].
-The viscosity field $\eta$ is calculated as the harmonic average between a linear part $\eta_{lin}$ that depends on temperature only or on temperature and 
-depth $y$ and a nonlinear-plastic part $\eta_{pl}$ that depends on the strain rate $\dot{\bm{\epsilon}}$, as follows
-\[\eta(T,y,\dot{\bm{\epsilon}})=2\left(\frac{1}{\eta_{lin}(T,y)}+\frac{1}{\eta_{pl}(\dot{\bm{\epsilon}})}\right)^{-1}\]
-The linear and the nonlinear-plastic parts of the viscosity are calculated following \citet{Tosi2015} as follows
-\[\eta_{lin}(T,y)=\exp(-\ln(\Delta\eta_T)T+\ln(\Delta\eta_y)y)\]
-\[\eta_{pl}(\dot{\bm{\epsilon}})=\eta^*+\frac{\sigma_Y}{\sqrt{\dot{\bm{\epsilon}}:\dot{\bm{\epsilon}}}}\]
-where $\Delta\eta_T$, $\Delta\eta_y$, $\eta^*$ and $\sigma_Y$ are parameters chosen as listed in Table \ref{tab:tosi_setup}.
-
-For each case, some diagnostic quantities measured at the steady state are reported in Table \ref{tab:visco_mantle} (see Section \ref{sec:error} for 
-the computation of all the quantities), showing a well fit with results described in \citet{Tosi2015}. Data and charts for all cases can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Viscoplastic_mantle_convection}.
-
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{white!100!}{gray!25!}
-%\begin{table}[H]
-%  \caption{Results of the viscoplastic mantle convection benchmark for Cases 1-4 compared with results obtained with ELEFANT and YACC (from \citealp{Tosi2015}).}
-%  \centering
-%  \resizebox{\textwidth}{!}{
-%\begin{tabular}{|c|ccc|ccc|ccc|ccc|}
-%  \toprule
-%    & \multicolumn{3}{c|}{\textbf{Case 1}} & \multicolumn{3}{c|}{\textbf{Case 2}} & \multicolumn{3}{c|}{\textbf{Case 3}} &
-%    \multicolumn{3}{c|}{\textbf{Case 4}} \\
-%   & FALCON & ELEFANT& YACC & FALCON& ELEFANT& YACC & FALCON & ELEFANT & YACC & FALCON & ELEFANT & YACC \\
-%  \midrule
-%   $\left\langle T \right\rangle$ & 0.7759 & 0.7758 & 0.7767 & 0.6043 & 0.6036 & 0.6058 & 0.7275 & 0.7275 & 0.7286 & & 0.5277 & 0.5289\\
-%    Nu$^{top}$ & 3.4280 & 3.4214 & 3.4298 & 8.5242 & 8.5115 & 8.5278 & 3.0368 & 3.0347 & 3.0374 && 6.5912 & 6.5572 \\
-%    Nu$^{bottom}$ & 3.3168 & 3.313 & 3.3143 & 8.4599 & 8.4465 & 8.3990 & 3.0086 & 2.9908 & 2.9628 && 6.5834 & 6.5243 \\
-%    $v_{\textrm{rms}}$ & 249.901 & 249.134 & 251.7997 & 141.2451 & 140.7087 & 142.2020 & 100.2776 & 100.1208 & 100.9467 && 79.1105 & 79.6202 \\
-%    $v_{\textrm{rms}}^{top}$ & 1.8907 & 1.8642 & 1.8298 & 104.7928 & 104.1228 & 105.1767 & 2.0832 & 2.0652 & 2.0374 && 74.7596 & 75.8414 \\
-%    $u_{max}^{top}$ & 2.6340 & 2.6119 & 2.5516 & 121.9212& 121.766 & 122.3238 & 2.9086 & 2.9019 & 2.8458 && 88.9146 & 89.2940 \\
-%    $\eta_{min}$ & - & - & - & $2.0303\times10^{-5}$ & $2.0600\times10^{-5}$ & $1.9853\times10^{-5}$ & $4.8129\times10^{-5}$ & $4.8080\times10^{-5}$ & $4.7907\times10^{-5}$ &
-%    & $1.9860\times10^{-4}$ & $1.9174\times10^{-4}$ \\
-%    $\eta_{max}$ & - & - & - & 1.3325 & 1.85 & 1.6095 & 0.9360 & 0.9023 & 1 && 1.52 & 1.6773 \\
-%    $\left\langle W \right\rangle$ & 2.4299 & 2.4316 & 2.4583 & 7.5648 & 7.5665 & 7.6222 & 2.0387 & 2.0384 & 2.0400 && 5.6216 & 5.6512 \\
-%    $\left\langle \Phi \right\rangle/Ra$ & 2.4311 & 2.4276 & 2.4333 & 7.5661 & 7.5585 & 7.6194 & 2.0386 & 2.0370 & 2.0335 && 5.6182 & 5.6463 \\
-%    $\delta$ & 0.05\% & 0.16\% & 1.02\% & 0.02\% & 0.11\% & 0.04\% & $<$0.01\% & 0.07\% & 0.32\% && 0.06\% & 0.09\% \\
-%  \bottomrule
-%  \end{tabular}}
-%  \label{tab:visco_mantle}
-%  \end{table}
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{2}{white!100!}{gray!25!}
-\begin{table}[H]
-  \caption{Benchmark cases and corresponding parameters.}
-  \centering
-  \small
-\begin{tabular}{|>{\centering}m{0.08\textwidth}|>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}
-  >{\centering}m{0.05\textwidth} >{\centering}m{0.05\textwidth} |>{\centering\arraybackslash}m{0.2\textwidth}|}
-  \toprule
-    Case & $Ra$ & $\Delta\eta_T$ & $\Delta\eta_y$ & $\eta^*$ & $\sigma_Y$ & Convective regime \\
-  \midrule
-  \rowcolor{gray!25!}
-    1 & $10^2$ & $10^5$ & 1 & - & - & Stagnant lid \\
-    2 & $10^2$ & $10^5$ & 1 & $10^{-3}$ & 1 & Mobile lid  \\
-    \rowcolor{gray!25!}
-    3 & $10^2$ & $10^5$ & 10 & - & - & Stagnant lid  \\
-    4 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 1 & Mobile lid \\
-    \rowcolor{gray!25!}
-    5 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 4 & Periodic \\
-  \bottomrule
-  \end{tabular}
-  \label{tab:tosi_setup}
-  \end{table}
-
-\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{gray!25!}{white!100!}
-\begin{table}
-  \caption{Results of the viscoplastic mantle convection benchmark for different grid resolutions in Cases 1-5 compared with results obtained with ELEFANT
-  and YACC ($100\times100$ elements, from \citealp{Tosi2015}) and in Stone 28 ($32\times32$ elements, from
-  \url{https://github.com/cedrict/fieldstone/tree/master/python_codes/fieldstone_28}).}
-  \centering
-  \resizebox{\textwidth}{!}{
-\begin{tabular}{|cc|ccccccccccc|}
-  \toprule
-  \multicolumn{13}{|c|}{\textbf{Case 1}} \\
-  \midrule
-  Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-  & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-  \hline
-  \rowcolor{gray!25!}
-    YACC & $100\times100$ & 0.7767 & 3.4298 & 3.3143 & 251.7997 & 1.8298 & 2.5516 & - & - & 2.4583 & 2.4333 & 1.02\% \\
-    ELEFANT & $100\times100$ & 0.7758 & 3.4214 & 3.313 & 249.134 & 1.8642 & 2.6119 & - & - & 2.4316 & 2.4276 & 0.16\% \\
-    \rowcolor{gray!25!}
-    Stone 28 & $32\times32$ & 0.7737 & 3.3987 & - & 243.872 & - & - & - & - & - & - & - \\
-%    FALCON & $32\times32$ & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & - & - & 2.4745 & 2.4864 & 0.48\% \\
-%    \rowcolor{gray!25!}
-    FALCON & $100\times100$ & 0.7759 & 3.4280 & 3.3168 & 249.901 & 1.8907 & 2.6340 & - & - & 2.4299 & 2.4311 & 0.05\% \\
-    \midrule
-    \multicolumn{13}{|c|}{\textbf{Case 2}} \\
-    \midrule
-    Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-    & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-    \hline
-    \rowcolor{gray!25!}
-     YACC & $100\times100$ & 0.6058 & 8.5278 & 8.3990 & 142.2020 & 105.1767 & 122.3238 & $1.9853\times10^{-5}$ & 1.6095 & 7.6222 & 7.6194 & 0.04\% \\
-     ELEFANT & $100\times100$ & 0.6034 & 8.5115 & 8.4465 & 140.8067 & 104.1228 & 121.766 & $2.0600\times10^{-5}$ & 1.85 & 7.5665 & 7.5585 & 0.11\% \\
-     \rowcolor{gray!25!}
-     Stone 28 & $32\times32$ & 0.6052 & 8.1715 & - & 141.518 & - & - & - & - & - & - & - \\
-%     FALCON & $32\times32$ & 0.6454 & 8.1422 & 7.4020 & 156.6908 & 95.7226 & 116.9302 & $2.2572\times10^{-5}$ & 0.7194 & 7.4289 & 7.4452 & 0.22\% \\
-%     \rowcolor{gray!25!}
-     FALCON & $100\times100$ & 0.6044 & 8.5266 & 8.4590 & 141.3684 & 104.8867 & 122.0470 & $2.0333\times10^{-5}$ & 1.2914 & 7.5707 & 7.5718 & 0.01\% \\
-     \midrule
-     \multicolumn{13}{|c|}{\textbf{Case 3}} \\
-     \midrule
-     Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-     & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-     \hline
-     \rowcolor{gray!25!} 
-     YACC & $100\times100$ & 0.7286 & 3.0374 & 2.9628 & 100.9467 & 2.0374 & 2.8458 & $4.7907\times10^{-5}$ & 1 & 2.0400 & 2.0335 & 0.32\% \\
-      ELEFANT & $100\times100$ & 0.7275 & 3.0347 & 2.9908 & 100.1208 & 2.0652 & 2.9019 & $4.8080\times10^{-5}$ & 0.9023 & 2.0384 & 2.037 & 0.07\% \\
-      \rowcolor{gray!25!}
-      Stone 28 & $32\times32$ & 0.7277 & 3.0347 & - & 100.018 & - & - & - & - & - & - & - \\
-%      FALCON & $32\times32$ & 0.7281 & 3.0618 & 2.8132 & 101.0092 & 2.1630 & 3.0178 & $4.8500\times10^{-5}$ & 0.7845 & 2.0699 & 2.0695 & 0.02\% \\
-%      \rowcolor{gray!25!}
-      FALCON & $100\times100$ & 0.7275 & 3.0368 & 3.0086 & 100.2776 & 2.0832 & 2.9086 & $4.8129\times10^{-5}$ & 0.9360 & 2.0387 & 2.0386 & $<$0.01\% \\
-      \midrule
-      \multicolumn{13}{|c|}{\textbf{Case 4}} \\
-      \midrule
-      Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-      & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-      \hline
-      \rowcolor{gray!25!} 
-      YACC & $100\times100$ & 0.5289 & 6.5572 & 6.5243 & 79.6202 & 75.4814 & 89.2940 & $1.9174\times10^{-4}$ & 1.6773 & 5.6512 & 5.6463 & 0.09\% \\
-       ELEFANT & $100\times100$ & 0.5277 & 6.5912 & 6.5834 & 79.1105 & 74.7596 & 88.9146 & $1.9860\times10^{-4}$ & 1.5200 & 5.6216 & 5.6182 & 0.06\% \\
-       \rowcolor{gray!25!}
-       Stone 28 & $32\times32$ & 0.5286 & 6.4036 & - & 79.4746 & - & - & - & - & - & - & - \\
-%       FALCON & $32\times32$ & 0.5334 & 6.4679 & 6.3435 & 80.6153 & 76.0511 & 90.2058 & $2.0393\times10^{-4}$ & 0.6453 & 5.6676 & 5.6698 & 0.04\% \\
-%       \rowcolor{gray!25!}
-       FALCON & $100\times100$ & 0.5279 & 6.5897 & 6.5830 & 79.1220 & 75.0715 & 88.8484 & $1.9693\times10^{-4}$ & 1.4133 & 5.6127 & 5.6113 & 0.03\% \\
-       \midrule
-       \multicolumn{13}{|c|}{\textbf{Case 5}} \\
-       \midrule
-       Code & Resolution & $P$ & $\left\langle T \right\rangle_{min}$ & $\left\langle T \right\rangle_{max}$ & Nu$^{top}_{min}$ & Nu$^{top}_{max}$ &
-       $v_{\textrm{rms}}^{min}$ & $v_{\textrm{rms}}^{max}$ & \multicolumn{2}{c}{$(\left\langle \Phi \right\rangle/Ra)_{min}$} & 
-       \multicolumn{2}{c|}{$(\left\langle \Phi \right\rangle/Ra)_{max}$} \\
-       \hline
-       \rowcolor{gray!25!}
-         YACC & $100\times100$ & 0.0792 & 0.6543 & 0.6722 & 2.6950 & 7.2792 & 41.9813 & 98.8230 & \multicolumn{2}{c}{1.4347} & 
-         \multicolumn{2}{c|}{9.2748} \\
-         ELEFANT & $100\times100$ & 0.076 & 0.6514 & 0.6693 & 2.6808 & 7.3506 & 41.2578 & 100.089 & \multicolumn{2}{c}{1.3324} & 
-         \multicolumn{2}{c|}{9.5289} \\
-         \rowcolor{gray!25!}
-         Stone 28 & $32\times32$ & 0.0764 & 0.6521 & 0.6697 & 2.6912 & 7.0763 & 41.61 & 98.183 & \multicolumn{2}{c}{-} & 
-         \multicolumn{2}{c|}{-} \\
-     %    FALCON & $32\times32$ & & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & \multicolumn{2}{c}{-} & 
-     %     \multicolumn{2}{c|}{-} \\
-     %    \rowcolor{gray!25!}
-         FALCON & $100\times100$ & 0.0775 & 0.6519 & 0.6697 & 2.6847 & 7.3629 & 41.5256 & 99.5466 & \multicolumn{2}{c}{1.3399} & 
-         \multicolumn{2}{c|}{9.3993} \\
-  \bottomrule
-  \end{tabular}}
-  \label{tab:visco_mantle}
-  \end{table}
-
-\subsection{Thin layer entrainment}\label{sec:thin}
-This experiment is performed as originally proposed by \citet{vanKeken1997} with two incompressible fluids in a rectangular domain with $L_x=2$ and $L_y=1$
-and gravity acceleration $g_y=10^{10}Ra$. The Rayleigh number and the compositional Rayleigh number are fixed to $Ra=300000$ and $Rc=450000$, respectively.
-Both fluids have constant viscosity, thermal conductivity, specific heat ($\eta=\rho=C_p=1$) and thermal expansion coefficient ($\alpha=10^{-10}$). Fluid 1
-has a density $\rho_1=1$, while fluid 2 is denser ($\rho_2=\rho_1+1.5 \cdot 10^{-10}$) and is placed at the bottom of the domain, for $y \leq 0.025$.
-Temperature are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides of the domain. The initial
-temperature field is given by
-\[T(x,y)=T_u(x,y)+T_l(x,y)+T_r(x,y)+T_s(x,y)-\frac{3}{2}\]
-with
-\begin{eqnarray}
-T_u(x,y)&=&\frac{1}{2}erf\left(\frac{1-y}{2}\sqrt{\frac{u_0}{x}}\right)\nonumber \\
-T_l(x,y)&=&1-\frac{1}{2}erf\left(\frac{y}{2}\sqrt{\frac{u_0}{L_x-x}}\right)\nonumber \\
-T_r(x,y)&=&\frac{1}{2}+\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{y+1}} \exp\left(-\frac{x^2u_0}{4y+4}\right)\nonumber \\
-T_s(x,y)&=&\frac{1}{2}-\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{2-y}} \exp\left(-\frac{(L_x-x)^2u_0}{8-4y}\right)\nonumber
-\end{eqnarray}
-and
-\begin{eqnarray}
-u_0&=&\frac{L_x^{7/3}}{(1+L_x^4)^{2/3}}\left(\frac{Ra}{2\sqrt{\pi}}\right)^{\frac{2}{3}}\nonumber \\
-Q&=&2\sqrt{\frac{L_x}{\pi u_0}}\nonumber
-\end{eqnarray}
-
-The experiment is performed with two grid resolution (\num{125x40} and \num{200x80} elements), with 100 markers per element and CFL of 0.25. $v_{\textrm{rms}}$
-(see Sec. \ref{sec:error}) calculated for both the simulations match well with results from \citet{vanKeken1997} and obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and ELEFANT \citep{Thieulot2014}, obtained with similar grid resolutions
-(Fig. \ref{fig:thin}). In addition, a variety of simulations with different resolutions and aspect ratios are performed to verify that the sensitivity of the
-initial velocity field. Results match well with results from \citet{vanKeken1997} (in the grey area) and \citet{Thieulot2014} (Fig. \ref{fig:thin_initial}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Thin_layer}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_thin.pdf}
-\caption{$v_{\textrm{rms}}$ for the thin layer experiment as function of time for different grid resolution. Results are compared with results obtained by
-\citet{vanKeken1997} (black lines) and with ASPECT and ELEFANT.}
-\label{fig:thin}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrmszero.pdf}
-\caption{$v_{\textrm{rms}}$ at $t=0$ for the thin layer experiment as function of the element size for different aspect ratios of the elements. Results are compared
-with results obtained by \citet{vanKeken1997} (black dots in the grey area) and with ASPECT and ELEFANT.}
-\label{fig:thin_initial}
-\end{figure}
-
-\subsection{Sinking hydrated cylinder}\label{sec:quinquis}
-This experiment is performed as explained in \citet{Quinquis2014} and simulates the sinking of a cold, hydrated cylinder into a warm, dry mantle. The domain
-is square with $L_x=L_y=$ \SI{300}{\km}, a grid resolution of \num{300x300} elements and 25 markers per element. The sinking cylinder is located at $x=$
-\SI{150}{\km} and $y=$ \SI{170}{\km} with a radius of \SI{20}{\km} and has $\rho_c=$ \SI{3250}{\kg\per\cubic\m}, $\eta_c=$ \SI{e23}{\pascal\s}, a thermal
-conductivity $k_c=$ \SI{4.5}{\watt\per\m\per\kelvin},a specific heat $C_{p_c}=$ \SI{1250}{\joule\per\kg\per\kelvin}; the surrounding mantle has $\rho_m=$
-\SI{3200}{\kg\per\cubic\m}, $\eta_m=$ \SI{e20}{\pascal\s}, a thermal conductivity $k_m=$ \SI{105}{\watt\per\m\per\kelvin} and specific heat $C_{p_m}=$
-\SI{1250}{\joule\per\kg\per\kelvin}; a 58 km-thick lithosphere overlies the mantle and has $\rho_l=$ \SI{3200}{\kg\per\cubic\m}, $\eta_l=$ \SI{e23}{\pascal\s},
-a thermal conductivity $k_l=$ \SI{4.5}{\watt\per\m\per\kelvin}, specific heat $C_{p_l}=$ \SI{750}{\joule\per\kg\per\kelvin} and a thermal diffusivity of
-\SI{e-6}{\square\m\per\s}. Velocity boundary conditions are set to free slip on all sides of the domain. Temperature increase linearly in the lithosphere,
-from 0° to 1300°C, and in the mantle, from 1300° to 1360.5°C. The initial temperature of the cylinder is fixed to 400°C. Throughout the evolution, temperature
-is fixed to 0° and 1360.5°C at the top and bottom of the domain, respectively. The initial bound water content is imposed at 2 wt.\% in the cylinder and at
-0 wt.\% in both the mantle and the lithosphere. The maximum amount of water in the mantle is fixed to 0.2 wt.\%, while maximum water content in the cylinder
-and in the lithosphere is function of pressure and temperature and is calculated using a serpentinized harzburgite with Perple\_X, as in \citet{Quinquis2014}.
-
-As shown in \citet{Quinquis2014}, the model is characterised by a progressive dehydration in the external portion of the cylinder, due to the increase of
-temperature, with a consequent vertical migration of free water that hydrates the mantle above the cylinder, up to the lithosphere. The distribution of
-bound and free water during the evolution of the experiment is shown in Fig. \ref{fig:quinquis} (left and right columns, respectively). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Hydration}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Hydration.pdf}
-\caption{Distribution of bound and free water (left and right columns, respectively) at $t=0.5$ and $t=$ \SI{1.25}{\mega\year} (first and second rows,
-respectively) for the sinking hydrated cylinder experiment. Only markers with bound or free water not equal to 0 are plotted.}
-\label{fig:quinquis}
-\end{figure}
-
-\subsection{Mantle melting curves}\label{sec:katz}
-This instantaneous experiment is performed in order to recreate isobaric melting curves, as function of temperature, bulk water and content of water in the
-melt, to compare with melting curves obtained by \citet{Katz2003}. The domain is square with $L_x=L_y=100$ km, a grid resolution of \num{100x100} elements and
-25 markers per element. Density is set in the entire domain to \SI{3300}{\kg\per\cubic\m} and temperature increases from the sides to the centre from 1000° to
-1500°C. The experiment is performed for bulk water contents from 0 to 0.3 wt.\%.
-
-Fig. \ref{fig:fraction} shows isobaric melting curves for \SI{1}{\giga\pascal} (panel a) and \SI{3}{\giga\pascal} (panel b) calculated for different bulk water
-contents. The curves well match with those obtained for the mantle by \citet{Katz2003}, showing the saturation at the solidus for bulk water of 0.2 and 0.3 wt.\%, indicating
-that the code correctly determines melt fraction as function of pressure, temperature and bulk water content. Same results can be observed in Fig.
-\ref{fig:water}a, where isobaric and isothermal melting curves are shown as function of bulk water content. In addition, Fig. \ref{fig:water}b shows that the
-code correctly determines also the percentage of water in the melt as function of pressure, temperature and melt fraction. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Melting}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Fraction.pdf}
-\caption{Isobaric melting curves for 1 GPa (panel a) and 3 GPa (panel b) as function of temperature with different bulk water contents.}
-\label{fig:fraction}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Water.pdf}
-\caption{Isobaric melting curves for 1.5 GPa as function of bulk water (panel a) and water in the melt (panel b) calculated for different temperatures.}
-\label{fig:water}
-\end{figure}
-
-\clearpage
-
-\bibliography{export}
-
-\end{document}

### 2022-09-07 tex/Benchmark/block.tex
update main

--- 2022-07-20 FALCON_description.tex
@@
-\documentclass[hidelinks,10pt,a4paper]{article}
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{textcomp}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage{booktabs}
-\usepackage{multirow}
-\usepackage[cm]{fullpage}
-\usepackage[round]{natbib}
-\bibliographystyle{abbrvnat}
-\usepackage{doi}
-\usepackage{bm}
-\usepackage[table]{xcolor}
-\hypersetup{
-    colorlinks,
-    linkcolor={blue},
-    citecolor={blue},
-    urlcolor={blue}
-}
-\usepackage{float}
-\restylefloat{table}
-\usepackage{siunitx}
-\DeclareSIUnit\year{yr}
-
-\begin{document}
-
-\section{Introduction}\label{sec:intro}
-This document includes a description of the code FALCON, with a presentation of the features implemented followed by results of extensive benchmarks listed in
-Table \ref{tab:list}, performed to verify that the code solves correctly a wide range of problems. Description and results of all benchmarks are shown
-in Section \ref{sec:benchmark}.
-
-\begin{table}[H]
-\caption{List of benchmarks performed for each feature of the code. Data of all benchmarks can be found at 
-\url{https://github.com/aleregorda/Benchmarks}}
-\centering
-\begin{tabular}{| l | l | l |}
-\toprule
- \textbf{Feature}  & \textbf{Benchmark} & \textbf{Section} \\
-\midrule
-  Solver  & Stokes flow & \ref{sec:stokes} \\
-\hline
-  \multirow{2}{*}{Markers advection} & Zalesak disk & \ref{sec:runge} \\
-    & Conservative Interpolation Velocity & \ref{sec:CVI} \\
-\hline
-  \multirow{4}{*}{Momentum equation}  & Poiseuille flow & \ref{sec:poiseuille}\\
-    & Instantaneous 2D sphere & \ref{sec:ist_sphere} \\
-    & Rayleigh-Taylor experiment & \ref{sec:rayleigh} \\
-    & Falling block & \ref{sec:block} \\
-\hline
-    & 2D Stokes sphere & \ref{sec:sphere} \\
-    Sticky air and  & Stabilisation algorithm & \ref{sec:stab} \\
-    free surface & Topography relaxation & \ref{sec:crameri} \\
-  & Spontaneous subduction & \ref{sec:subduction} \\
-\hline
-  Erosion and sedimentation  & - & - \\
-\hline
-   Non-linear  & Slab detachment & \ref{sec:slab} \\
-   visco-plasticity & Indenter & \ref{sec:indenter} \\
-    & Brick & \ref{sec:brick} \\
-\hline
-  \multirow{3}{*}{Energy equation}  & Advection stabilisation & \ref{sec:advection} \\
-    & Simple shear heating & \ref{sec:simple_shear} \\
-    & Shear and adiabatic heating & \ref{sec:shear} \\
-\hline
-\multirow{3}{*}{Energy + momentum} & Mantle convection & \ref{sec:mantle} \\
-& Viscoplastic mantle convection & \ref{sec:visco_mantle} \\
-& Thin layer entrainment & \ref{sec:thin} \\
-\hline
-  Phase changes  & \multirow{2}{*}{Hydrated sinking cylinder} & \multirow{2}{*}{\ref{sec:quinquis}} \\
-  and hydration  &  & \\
-\hline
-  Melting & Experimental melting curves & \ref{sec:katz} \\
-\bottomrule
-\end{tabular}
-\label{tab:list}
-\end{table}
-
-The code is written in Fortran90 and uses the finite element method (FEM) with quadrilateral $Q_1 \times P_0$ elements (continuous bilinear velocity and
-discontinuous constant pressure), associated with MUMPS\footnote{\url{http://mumps.enseeiht.fr/}} \citep{Amestoy2001,Amestoy2019} that is a software package
-for solving systems of linear equations of the form $A \cdot x = b$, where \textit{A} is a square sparse matrix, by means of a direct method. Since
-$Q_1 \times P_0$ elements do not satisfy Ladyzhenskaya, Babuska and Brezzi (LBB) stability condition \citep{Donea2003}, elemental pressure is elaborated
-in post-processing to avoid spurious pressures by means of a double interpolation to smooth the pressure. Following this procedure, the elemental pressure
-is interpolated onto nodes and then back onto elements. After the smoothing procedure, the pressure field is calculated again on the nodes to be used in
-combination with the lithostatic pressure, which is calculated onto nodes. Correctness of the solution and performances of the solver in terms of time and
-memory usage are tested solving the Stokes flow with the analytical solution proposed by \citet{Donea2003} (Section \ref{sec:stokes}).
-
-The thermo-mechanics of crust-mantle systems is described by means of the conservation of mass, momentum and energy equations, expressed as follows:
-\begin{eqnarray}
-\label{eq:mass}\frac{\partial \rho}{\partial t}+\bm{\nabla} \cdot (\rho \bm{u})&=&0\\
-\label{eq:momentum}-\bm{\nabla} p + \bm{\nabla} \cdot \bm{\tau} + \rho \bm{g}&=&0\\
-\label{eq:energy}\rho C_p \left(\frac{\partial T}{\partial t} + \bm{u} \cdot \bm{\nabla} T \right)&=&\bm{\nabla} \cdot \left(k\bm{\nabla} T\right) + H_{tot}
-\end{eqnarray}
-where $\rho$ is the density, $\bm{u}$ is the velocity, \textit{p} is the pressure, $\bm{\tau}$ is the deviatoric stress, $\bm{g}$ is the gravity acceleration,
-$C_p$ is the specific heat at constant pressure, \textit{T} is the temperature, \textit{k} is the thermal conductivity and $H_{tot}$ is the total internal heat
-production. Density variations due to temperature are generally small enough to assume the density as constant ($\rho=\rho_0$) in Eq. \ref{eq:mass} and in
-Eq. \ref{eq:energy}, while it must be treated as a variable in the buoyancy term of Eq. \ref{eq:momentum}, such that \[\rho=\rho_0(1-\alpha(T-T_0))\]
-where $\rho_0$ is the density at a reference temperature $T_0$ and $\alpha$ is the coefficient of thermal expansion. Eq. \ref{eq:mass} then can be rewritten as
-\begin{equation}\label{eq:mass_0}
-\bm{\nabla} \cdot \bm{u}=0
-\end{equation}
-This simplification is known as the extended Boussinesq approximation.
-
-The time step $dt$ is chosen in according to Courant-Friedrichs-Lewy condition \citep{Anderson1995}, such that
-\[dt=C_n\; \min\left(\frac{\Delta x}{\max|\bm{u}|},\frac{(\Delta x)^2}{\kappa}\right)\]
-where $0<C_n<1$ is the Courant (CFL) number, $\Delta x$ is the minimum dimension of smallest element, $\max|\bm{u}|$ is the maximum velocity calculated on the
-entire domain and
-$\kappa=\frac{k}{\rho C_p}$ is the thermal diffusivity \citep{Thieulot2014}.
-
-\section{Lagrangian markers}\label{sec:markers}
-Elemental properties (density, viscosity, thermal conductivity, specific heat and thermal expansion) needed to solve Eqs. \ref{eq:mass_0}, \ref{eq:momentum}
-and \ref{eq:energy} are related to the composition of each element, determined by means of Lagrangian markers that are characteristic of different materials
-in the domain. Elemental properties (with the sole exception of the viscosity) are calculated using an arithmetic mean, as \[P_e=\frac{1}{n}\sum_{i=1}^n P_i\]
-where $P_e$ is the elemental property, $P_i$ is the property characteristic of the material of each marker and $n$ is the number of markers of the element.
-Differently, the average scheme for the elemental viscosity can be chosen between harmonic, geometric and arithmetic mean. Markers can be placed regularly or
-randomly at the beginning of the simulation and their advection is performed by either a 2nd-order or a 4th-order Runge-Kutta in space, interpolating the velocity 
-field on each marker by means of the shape functions (see Section \ref{sec:runge}). The interpolated velocity is then corrected by means of the Conservative
-Velocity Interpolation (CVI), which introduces a corrective term to reduce dispersion and clustering of particles in both steady state and time-dependent 
-\citep{Wang2015} (see Section \ref{sec:CVI}).
-
-The initial distribution of the markers is created using the open-source code library
-Geodynamic World Builder \citep{Fraters2019}. During the simulation, each marker carries memory of temperature, pressure and accumulated strain, which are
-determined interpolating the relative nodal parameter. In particular, pressure is interpolated onto each marker after the smoothing procedure. The number
-of markers contained in each element is maintained between $n_{min}$ and $n_{max}$. When in an element there are less markers than $n_{min}$ the code adds
-random markers to reach the $n_{min}$, while if the number is higher than $n_{max}$ some of them are randomly deleted. When new markers are added, they assume
-the properties of the nearest marker. In this way elements are never empty and maintain a number of markers inside a prefixed range.
-
-\section{The momentum equation}\label{sec:mom_eq}
-The penalty method is implemented so that Eq. \ref{eq:mass_0} can be written as:
-\begin{equation}\label{eq:mass_penalty}
-\bm{\nabla} \cdot \bm{u} + \frac{p}{\lambda}=0
-\end{equation}
-where $\lambda$ is the so-called penalty parameter, which should be 6-7 orders of magnitude larger than the shear viscosity to ensure that mass conservation
-is satisfied \citep{Donea2003,Thieulot2014}.
-
-The deviatoric stress tensor in Eq. \ref{eq:momentum} can be written in terms of the strain rate tensor as $\bm{\tau}=2\eta\dot{\bm{\epsilon}}$, with
-$\dot{\bm{\epsilon}}=\frac{1}{2}\left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)$. Therefore, Eq. \ref{eq:momentum} can be rewritten as:
-\begin{equation}\label{eq:momentum_pressure}
--\bm{\nabla} p + \bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right) + \rho \bm{g} = 0
-\end{equation}
-Finally, using pressure from Eq. \ref{eq:mass_penalty}, Eq. \ref{eq:momentum_pressure} can be rewritten as:
-\begin{equation}\label{eq:momentum_penalty}
-\lambda \bm{\nabla} \left(\bm{\nabla} \cdot \bm{v}\right)+\bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right)+\rho\bm{g}=0
-\end{equation}
-The penalty method is been associated to the iterative Uzawa method, as described in detail in \citet{Dabrowski2008} and \citet{Thieulot2014}
-(see Section \ref{sec:poiseuille}). Results of benchmarks performed to verify the correctness in the implementation of Eq. \ref{eq:momentum_penalty} are
-shown in Sections \ref{sec:ist_sphere} and \ref{sec:rayleigh}.
-
-To support large viscosities variations the penalty parameter is related to the effective elemental viscosity by means of a dimensionless coefficient, so that
-$\lambda=\lambda_e(e)\eta_{eff}(e)$ \citep{Marotta2006,Dabrowski2008,Thieulot2014}. Benchmark of the falling block \citep{Gerya2003a,Gerya2010,Thieulot2011}
-is performed to verify that the code can correctly deal buoyancy driven flows with strong viscosity contrasts (Section \ref{sec:block}).
-
-\section{Sticky air and free surface}\label{sec:sticky}
-The Earth's surface can be treated by means of either the so-called sticky-air or a true free surface method, both of them implemented in the code. In the
-sticky-air method the surface is approximated with the introduction of a buoyant layer with a viscosity at least four orders of magnitude lower than the crust
-\citep{Schmeling2008,Crameri2012} and the interface between lithosphere and air is defined using a chain of passive markers that are advected as the Lagrangian
-markers. The correctness of the evolution of the markers chain is tested with the experiment of a 2D time-dependent Stokes sphere below a free surface and
-compared with results from ASPECT\footnote{\url{https://aspect.geodynamics.org/}} \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} (Section
-\ref{sec:sphere}). In the true free surface case the top boundary is assumed stress-free and velocities are not fixed. In this case topography variations are
-described by vertical deformations of the mesh that depend on the velocity field of the nodes that identify the free surface, while horizontal deformations are
-not taken into account. This procedure is known as the Arbitrary Lagrangian-Eulerian (ALE) method and its implementation follows the technique described in
-\citet{Thieulot2011}. However, although the implementation of a true free surface better reproduces laboratory experiment, extremely small time steps can be
-necessary to maintain stability \citep{Kaus2010a,Quinquis2011,Thieulot2014}. Therefore, the stability algorithm proposed by \citet{Kaus2010a} is implemented
-to avoid instabilities due to high density differences at the free surface when using too large time steps. The implementation of this algorithm are tested by
-performing the experiment described by \citet{Kaus2010a} (Section \ref{sec:stab}).
-
-The topography relaxation benchmark proposed by \citet{Crameri2012} is performed to verify that the code correctly recovers topography variations in case of
-both the sticky-air and the true free surface method (Section \ref{sec:crameri}). Finally, Section \ref{sec:subduction} show the results of the spontaneous
-subduction experiment described by \citet{Schmeling2008}.
-
-\section{Erosion and sedimentation}\label{sec:erosion}
-Surface processes at either the lithosphere-air interface or in correspondence of the free surface have been implemented by means of the software
-Fastscape\footnote{\url{https://fastscape.org/}} \citep{Braun2013,Cordonnier2019,Yuan2019a,Yuan2019b}. FastScapeLib is a set of routines that solve the stream
-power law (enriched by a sediment transport/deposition term) \citep{Yuan2019a}, hillslope diffusion and marine transport and deposition \citep{Yuan2019b}.
-The partial differential equation solved by FastScapeLib is:
-\begin{equation}
-\frac{\partial h}{ \partial t}=U-K_f A^{m_e}S^{n_e} + \frac{G}{A}\int_A \left( U-\frac{\partial h}{\partial t} \right) dA + K_d \bm{\nabla}^2 h 
-\end{equation}
-where $h$ is the topography, $U$ is the uplift, $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $S$ is the slope, $m_e$ and $n_e$ are
-parameters that should be chosen to constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}, $G$ is the dimensionless
-deposition/transport coefficient for the enriched stream power law and $K_d$ is the effective diffusivity. Terms $K_f A^{m_e}S^{n_e}$ and $K_d \bm{\nabla}^2 h$
-consider long-range processes, described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, and short-range processes,
-proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}.
-
-In case of sticky air, slope is determined by means of the markers chain, which is vertically corrected considering erosion and sedimentation rates. After the
-correction, continental markers above the markers chain are transformed in air markers and air markers below the markers chain become sediments. Similarly,
-slope is determined considering the nodes on the top boundary in case of true free surface and they are vertically corrected as for the markers chain. In this
-case, continental markers above the top boundary are simply deleted.
-%Surface processes at the lithosphere-air interface are implemented taking into account fluvial erosion (long-range processes), hillslope processes (short-range
-%processes) and failure of slopes exceeding a threshold angle $\theta$ \citep{Burov1997,Tucker1999,Sacek2012,Croissant2014,Ueda2015,Beucher2020}. Erosion rates
-%$\dot{E}_{lr}$ due to long-range processes can be described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{lr}=K_f A^{m_e} \bm{\nabla} h^{n_e}
-%\end{equation}
-%where $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $\bm{\nabla} h$ is the slope and $m_e$ and $n_e$ are parameters chosen in to
-%constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}. Erosion rates $\dot{E}_{sr}$ due to short-range processes are
-%implemented by assuming that transport is linearly proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{sr}=K_d \bm{\nabla}^2 h
-%\end{equation}
-%where $K_d$ is the effective diffusivity. In addition, the slope is modified if it still exceeds a threshold angle after long- and short-range processes
-%\citep{Ueda2015}.
-
-%Offshore sedimentation processes are implemented similarly to hillslope processes but with a depth-dependent coefficient \citep{Sacek2012,Beucher2020}.
-%The sedimentation rate $\dot{S}$ is calculated as
-%\begin{equation}
-%\dot{S}=K_s \bm{\nabla}^2 h
-%\end{equation}
-%with
-%\begin{equation}
-%K_s=K^*_s \exp\left(\frac{h-h_{sl}}{h_0}\right)
-%\end{equation}
-%where $K^*_s$ is a constant, $h_{sl}$ is the sea level and $h_0$ is a characteristic length. In case of sticky air, slope is determined by means of the markers
-%chain, which is vertically corrected considering erosion and sedimentation rates. After the correction, continental markers above the markers chain are
-%transformed in air markers and air markers below the markers chain become sediments. Similarly, slope is determined considering the nodes on the top boundary
-%in case of true free surface and they are vertically corrected as for the markers chain. In this case, continental markers above the top boundary are simply
-%deleted.
-
-\section{Non-linear rheologies}\label{sec:nonlinear}
-Non-linear rheologies are implemented combining viscous creep (dislocation and diffusion) and plastic yielding. For each marker, diffusion and dislocation
-viscosity ($\eta_{df|ds}$) can be determined as follows \citep{Karato1993,Warren2008b,Wang2016}:
-\begin{equation}\label{eq:viscous}
-\eta_{df|ds}=f_s \left(\frac{d^m}{A}\right)^{\frac{1}{n}} \left (\frac{2^{\frac{1-n}{n}}}{3^{\frac{n+1}{2n}}} \right)
-I_2^{\frac{(1-n)}{n}}\exp\left(\frac{Q+pV}{nRT}\right)
-\end{equation}
-where $f_s$ is a scaling factor used to represent lithologies that are stronger or weaker than the base set, $d$ is the grain size, $m$ is the grain size exponent,
-$n$ is the stress exponent, $A$ is the uniaxial pre-exponential factor, $I_2$ is the square root of the second invariant of the strain rate tensor, $Q$ is the
-activation energy, $p$ is the pressure, $V$ is the activation volume, $R$ is the gas constant and $T$ is the temperature.
-Pressures and temperatures are determined by the interpolation of the nodal parameters, while strain rates are calculated by means of the
-derivative of the velocities on the nodes. In case of diffusion creep $n=1$ and $m>0$, in case of dislocation creep $n>1$ and $m=0$. The viscous creep $\eta_{cp}$
-is then calculated as the harmonic average between $\eta_{df}$ and $\eta_{ds}$:
-\begin{equation}\label{eq:visco_creep}
-\eta_{cp}=\left(\frac{1}{\eta_{df}}+\frac{1}{\eta_{ds}}\right)^{-1}
-\end{equation}
-The implementation of the non-linear viscous creep viscosity depending on the strain rate is tested by means of the slab detachment benchmark
-(Section \ref{sec:slab}).
-
-Plastic yielding is implemented rescaling $\eta_{cp}$ in order to limit the stress below the yield stress $\sigma_y$ \citep{Thieulot2008,Thieulot2014,Glerum2018},
-obtaining
-\begin{equation}\label{eq:plastic}
-\eta_{pl}=\frac{\sigma_y}{2I_2}
-\end{equation}
-where $\eta_{pl}$ is the plastic viscosity and the yield stress is determined following the Drucker-Prager criterion, such as
-\begin{equation}\label{eq:yield}
-\sigma_y=C\;\cos(\phi)+p\;\sin(\phi)
-\end{equation}
-where $C$ is the cohesion and $\phi$ is the internal friction angle. In case of negative pressure it is imposed equal to 0, so that negative yield stress are
-excluded. The correctness of the non-linear solution in case of plastic viscosity with variable internal friction angle is verified performing indenter and
-brick experiments (Sections \ref{sec:indenter} and \ref{sec:brick}, respectively).
-
-Plastic yielding and viscous creep are then combined to obtain a viscoplastic viscosity $\eta_{vp}$ as follows, assuming that they are independent processes
-\citep{Karato2008,Glerum2018}:
-\begin{equation}\label{eq:viscoplastic}
-\eta_{vp}=\min(\eta_{cp},\eta_{pl})
-\end{equation}
-Finally, effective viscosity $\eta_{eff}$ is capped by the minimum and the maximum viscosity ($\eta_{min}$ and $\eta_{max}$, respectively) to avoid extremely
-low or high viscosity \citep{Glerum2018} as follows
-\begin{equation}\label{eq:effective}
-\eta_{eff}=\min(\max(\eta_{vp},\eta_{min}),\eta_{max})
-\end{equation}
-Viscous creep and plastic yielding are non-linear rheologies because of their dependence on the velocity field through pressure and strain rates. Therefore,
-the solution is determined by means of Picard-type iterations, until convergence of the velocity field \citep{Glerum2018}. The convergence is verified at each
-iteration $i$ via the nonlinear residual $\bm{\mathcal{R}}^i$ that can be determined as
-\begin{equation}\label{eq:convergence}
-\bm{\mathcal{R}}^i= \mathbb{K} (\eta_{eff}(\dot{\bm{\epsilon}}^{i-1},p^{i-1}))\cdot \bm{v}^{i-1} - \bm{f}^i
-\end{equation}
-%\begin{equation}\label{eq:convergence}
-%\chi=1-\frac{\langle (\bm{v}^i-\langle\bm{v}^i\rangle)\cdot (\bm{v}^{i+1}-\langle\bm{v}^{i+1}\rangle) \rangle}{\sqrt{\sigma^i\sigma^{i+1}}}
-%\end{equation}
-%Iterations are performed until either convergence ($\chi<tol$) or maximal number of non-linear iterations ($it_{max}$) is reached.
-where $\mathbb{K}$ is the Stiffness matrix and $\bm{f}$ is the right hand side vector \citep{Spiegelman2016,Glerum2018}.  The $L_2$-norm is extracted from
-$\bm{\mathcal{R}}^i$ and it is normalised as follows
-\begin{equation}\label{eq:convergence_res}
-\frac{||\bm{\mathcal{R}}^i||_2}{||\bm{\mathcal{R}}^0||_2}
-\end{equation}
-where $||\bm{\mathcal{R}}^0||_2$ is the $L_2$-norm of the first nonlinear iteration.
-Since often the normalised nonlinear residual drop very quickly over the first 2-3 iterations, also the $L_2$-norm of velocity and pressure residuals are
-calculated as
-\begin{equation}\label{eq:convergence_vel}
-\frac{||\Delta\bm{v}||_2}{||\bm{v}||_2}=\frac{||\bm{v}_i-\bm{v}_{i-1}||_2}{||\bm{v}_{i}||_2}
-\end{equation}
-and
-\begin{equation}\label{eq:convergence_press}
-  \frac{||\Delta p||_2}{||p||_2}=\frac{||p_i-p_{i-1}||_2}{||p_{i}||_2}
-\end{equation}
-respectively. Iterations are performed until either maximal number of non-linear iterations ($it_{max}$) is reached or all the normalised nonlinear residual, 
-the velocity residual and the pressure residual converge under a defined tolerance $tol$.
-
-\section{Healing and weakening}
-The temporal evolution of the accumulated strain $\epsilon$ has been implemented as in \citet{Fuchs2019,Fuchs2021}, as following
-\begin{equation}\label{eq:healing}
-  \frac{d\epsilon}{dt}=I_2 - \epsilon H_{rate}(T)
-\end{equation}
-where the first term on the right-hand side is a source term given by the second invariant of the strain rate $I_2$ and the second term a
-temperature-dependent healing factor $H_{rate}$, calculated as
-\begin{equation}\label{eq:healing_rate}
-  H_{rate}(T)=B \exp \left[-\frac{\mu}{2}\left(\frac{1}{T+1}-\frac{1}{2}\right)\right]
-\end{equation}
-where $B$ is a constant describing the time scale of healing, while $\mu$ and $T$ are the non-dimensional temperature activation and temperature,
-respectively \citep{Fuchs2019,Fuchs2021}.
-
-Strain softening is taken into account for both viscous creep and plastic viscosity \citep{Huismans2003,Babeyko2005,Huismans2005,Sobolev2005,Warren2008a} by
-means of the accumulated strain $\epsilon$ memorised by each marker. Softening in the viscous creep determines a linear decrease of $\eta_{vc}$ by means of a
-viscous strain softening factor $W_S$ that increases linearly from $W_{S_0}$ to $W_{S_{\infty}}$ for $\epsilon_{S_0}<\epsilon_S<\epsilon_{S_{\infty}}$ 
-\citep{Warren2008a}. This viscous softening can be related to strain-induced grain size reduction \citep{Warren2008a}. Differently, plastic softening is 
-simulated with a linear decrease of internal friction angle $\phi(\epsilon)$ and cohesion $C(\epsilon)$ in according to
-\begin{eqnarray}
-\label{eq:friction1}\phi(\epsilon)&=&\phi_0+(\phi_{\infty}-\phi_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}\\
-\label{eq:friction2}C(\epsilon)&=&C_0+(C_{\infty}-C_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}
-\end{eqnarray}
-where $\phi_0$, $C_0$ and $\phi_{\infty}$, $C_{\infty}$ are internal friction angle and cohesion for $\epsilon_0$ and $\epsilon_{\infty}$, respectively
-\citep{Huismans2003,Huismans2005,Warren2008a,Thieulot2014}. Plastic softening approximates deformation-induced softening of faults and brittle shear zones
-\citep{Warren2008a}.
-
-\section{The energy equation}\label{sec:en_eq}
-The stabilisation of the advection term of the energy equation (Eq. \ref{eq:energy}) needed to avoid possible oscillations in the thermal solution in the case
-when advection dominates over diffusion, is implemented by means of a streamline-upwind Petrov–Galerkin (SUPG) method, for which the advection term is modified
-follows the discussion in \citet{Thieulot2011} and \citet{Thieulot2014} (see Section \ref{sec:advection} for benchmark). The total internal heat production
-$H_{tot}$ in Eq. \ref{eq:energy} includes radiogenic heating $H_r$, shear heating $H_s$ and adiabatic heating $H_a$, where
-\begin{equation}\label{eq:shear}
-H_s=2\eta \dot{\bm{\epsilon}} : \dot{\bm{\epsilon}} = 2\eta(\dot{\epsilon}_{xx}^2+\dot{\epsilon}_{yy}^2+2\dot{\epsilon}_{xy}^2)
-\end{equation}
-and
-\begin{equation}\label{eq:adiabatic}
-H_a=T \alpha \frac{Dp}{Dt} \approx -\alpha T \rho g_y v
-\end{equation}
-The correctness of the implementation and the computation of shear and adiabatic heating is verified performing an exercise with an analytical solution
-(Section \ref{sec:simple_shear}) and Exercise 9.4 in \citet{Gerya2010} (Section \ref{sec:shear}). Furthermore, mantle convection and thin layer entrainment
-benchmarks (Sections \ref{sec:mantle} and \ref{sec:thin}, respectively) are performed to verify that the code solve correctly Eqs. \ref{eq:momentum_penalty}
-and \ref{eq:energy} in case of a temperature field characterised by an initial perturbation and with a temperature-dependent density. At last, the viscoplastic
-mantle convection benchmark is performed to check also the viscoplasticity in case of a initial perturbed temperature field (Section \ref{sec:visco_mantle}).
-
-\section{Phase transitions and hydration}\label{sec:phase}
-Variations of effective density, specific heat and coefficient of thermal expansion during the evolution of crustal and mantle materials at different
-pressure-temperature ($p$-$T$) conditions are computed by means of version 6.8.6 of Perple\_X software package \citep{Connolly2005}, similarly to the
-implementation described in \citet{Marotta2020}. By default, mineral assemblages and properties of each lithology (oceanic crust, lower and upper continental
-crust, sediments and dry or hydrated mantle) are calculated for temperatures from \SIrange{330}{1600}{\kelvin}, with increments of approximately \SI{4}{\kelvin},
-and pressures from \SIrange{0.1}{30}{\giga\pascal}, with increments of approximately \SI{0.03}{\giga\pascal}, for a total of almost 100000 points.
-
-The introduction of phase transitions produces two main effects that must be taken into account in numerical models: variations in density and release/absorption
-of latent heat \citep{Tackley2010}. Effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into
-account considering effective coefficient of thermal expansion \citep{Christensen1985,Zhong2015}. Similarly, effects of latent heat in the energy equation
-(eq \ref{eq:energy}) are taken into account by considering effective specific heat and coefficient of thermal expansion
-\citep{Christensen1985,Tackley2010,Zhong2015}. Effective density, specific heat and coefficient of thermal expansion are included in Perple\_X files and are
-assumed by each Lagrangian marker in according to its $p$-$T$ conditions.
-
-Hydration processes related to dehydration of subducting lithosphere \citep{Schmidt1998,Liu2007,Faccenda2009,Faccenda2010,Faccenda2014,Rosas2016} strongly
-influence the thermo-mechanics inside the mantle wedge, mainly because of both weakening effects on the mantle rheology and density variations in case of
-mantle serpentinization \citep{Gerya2002,Honda2003,Arcay2005,Roda2010,Regorda2017}. In the case that hydration is switched on, the amount of bound and free
-water is memorised by each marker, following the implementation of \citet{Quinquis2014}; hydration and dehydration processes are related to the amount of bound
-water of each marker and to the maximum amount of water it can transport, i.e. if the amount of bound water exceeds the maximum amount of water, the marker
-dehydrates and releases free water that can hydrates under-saturated markers. Maximum water content of each marker is determined as function of lithology and
-$p$-$T$ conditions and it is calculated using Perple\_X, in the same way than effective density, specific heat and coefficient of thermal expansion. Bound water
-is advected together with the markers, neglecting the effect of bound water diffusion, while free water simply migrates vertically and is not coupled to the
-solid-phase flow of the mantle wedge \citep{Arcay2005,Quinquis2014}. The correctness of the migration scheme of the free water and its absorption into bound
-water by under-saturated markers are verified performing the experiment described in \citet{Quinquis2014} (Sec. \ref{sec:quinquis}).
-
-The mantle viscosity weakening related to the amount of water has been extensively studied \citep{Chopra1981,Hirth2003} and it is implemented in according to
-\citet{Arcay2005} and \citet{Horiuchi2016}, as follows
-\begin{equation}\label{eq:wet_visc}
-\eta_{wet}=\eta_{dry}\left(\left[1-\frac{1}{f_v}\right]\exp\left(-\frac{[OH^-]}{[OH^-]_0}\right)+\frac{1}{f_v}\right)
-\end{equation}
-where $f_v$ is the viscosity reduction factor between dry and wet conditions, $[OH^-]$ is the water content and $[OH^-]_0$ is a reference water content set to
-620 ppm (0.062 wt.\%) \citep{Arcay2005}. By default, $f_v$ is set to 100, in according to \citet{Arcay2005} and \citet{Horiuchi2016}.
-
-\section{Mantle melting}\label{sec:melt}
-Melt fraction in the mantle depends on temperature ($T$, in \textdegree C), pressure ($p$, in \si{\giga\pascal}) and the water content in the melt ($X_{H_2O}$, in wt.\%).
-The determination of the melt fraction $M$ is implemented as explained by \citet{Katz2003}, according to the successive modification by \citet{Langmuir2006} and
-\citet{Kelley2010}, as follows
-\begin{eqnarray}\label{eq:melt_fr}
-M(p,T,X_{H_2O})&=&-T+T_{sol}+(x_t\cdot ln(p+y_t))\cdot X_{H_2O}-\nonumber\\
-&&+K_m\left(\frac{X_{H_2O}^{bulk}}{D_{H_2O}(1-X_{H_2O})+X_{H_2O}}\right)^{\gamma}
-\end{eqnarray}
-where $T_{sol}$ is the temperature of the dry solidus, $D_{H_2O}$ is the partition coefficient and $x_t\cdot log(p+y_t)$ indicates the pressure dependence of
-the melting curve \citep{Kelley2010}. $x_t$, $y_t$, $K_m$ and $\gamma$ are parameters chosen in according to \citet{Kelley2010}.
-However, $X_{H_2O}$ depends on the melt fraction as
-\begin{equation}\label{eq:melt_wt}
-X_{H_2O}(M)=\frac{X_{H_2O}^{bulk}}{D_{H_2O}+M(1-D_{H_2O})}
-\end{equation}
-and a numerical solution can be found using a root-finder method \citep{Katz2003,Wang2016}. Furthermore, the content of water in the melt is limited by the
-pressure-dependent saturation concentration of water in the melt, determined as
-\begin{equation}\label{eq:melt_max}
-X_{H_2O}^{sat}=\chi_1 p^{\lambda_m}+\chi_2 p
-\end{equation}
-where $\chi_1$, $\chi_2$ and $\lambda_m$ are parameters chosen according to \citet{Katz2003}, \citet{Langmuir2006} and \citet{Kelley2010}. The correctness in
-the determination of melt fractions and water contents in the melt are tested comparing model results with experimental curves obtained by \citet{Katz2003}
-(Sec. \ref{sec:katz}).
-
-In case of melting, viscous creep viscosity ($\eta_{vp}$) is modified according to \citet{Wang2016}, as
-\begin{equation}\label{eq:melt_visc}
-\eta_{melt}=\eta_{vp}\cdot \exp(\alpha_m M)
-\end{equation}
-where $\alpha_m$ are melt fraction factors for dislocation and diffusion creep, chosen in according to \citet{Wang2016}.
-
-Effective density ($\rho_{eff}$) for partially molten rocks is calculated as
-\begin{equation}\label{eq:melt_dens}
-\rho_{eff}=\rho_s(1-M)+\rho_m M
-\end{equation}
-where $\rho_s$ and $\rho_m$ are the densities of the solid and the molten rock, respectively \citep{Gerya2004b,Gerya2007,Wang2016}. The density of the
-solid rock is extracted by Perple\_X, while the density of the molten portion is calculated as
-\begin{equation}\label{eq:melt_dens_M}
-\rho_m(p,T)=\rho_0[1-\alpha(T-T_0)][1+\beta(p-p_0)]
-\end{equation}
-where $\rho_0$ is the density at the reference temperature ($T_0$) and pressure($p_0$), and $\alpha$ and $\beta$ are the thermal expansion and compressibility
-coefficients, respectively \citep{Gerya2004b,Gerya2007,Wang2016}.
-
-As for the phase transitions, effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into account
-considering effective coefficient of thermal expansion. Similarly, latent heat due to melting/crystallisation can be implicitly included in the energy equation
-(Eq. \ref{eq:energy}) by considering effective specific heat ($C_{p_{eff}}$) and thermal expansion ($\alpha_{eff}$) of the partially molten rocks, in according to
-\begin{eqnarray}
-\label{eq:cp_effective}C_{p_{eff}}&=&C_p+H_L\left(\frac{\partial M}{\partial T}\right)_p \\
-\label{eq:alpha_effective}\alpha_{eff}&=&\alpha+\rho\frac{H_L}{T}\left(\frac{\partial M}{\partial p}\right)_T
-\end{eqnarray}
-where $C_p$ and $\alpha$ are the specific heat and the coefficient of thermal expansion of the solid rock, respectively, and $H_L$ is the latent heat of the
-molten rock \citep{Gerya2004b,Gerya2007,Tackley2010}.
-
-\section{Benchmarks}\label{sec:benchmark}
-\subsection{Error measurements}\label{sec:error}
-In order to determine the accuracy of velocity and pressure field of the benchmarks, the $L_2$-norm is computed by numerical integration on the Gauss-Legendre
-quadrature points. $L_2$-norm for pressure and velocity errors can be evaluated as
-\begin{eqnarray}
-\textrm{err}_p&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|p^n_i(r_q)-p^a_i(r_q)|^2\right)w_q|J_q|}\nonumber \\
-\textrm{err}_v&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|u^n_i(r_q)-u^a_i(r_q)|^2+|v^n_i(r_q)-v^a_i(r_q)|^2\right)w_q|J_q|}\nonumber
-\end{eqnarray}
-respectively, where $n_e$ is the number of elements, $n_q$ is the number of quadrature points per element, $p^n_i(r_q)$ and $p^a_i(r_q)$ are the numerical and
-analytical pressure, respectively, in each quadrature point $q$, $w_q$  and $J_q$ are the weight and the Jacobian at the quadrature point
-$q$, $u^n_i(r_q)$, $v^n_i(r_q)$, $u^a_i(r_q)$ and $v^a_i(r_q)$ are the numerical and analytical velocities, respectively, in each quadrature point $q$.
-
-Other quantities used as comparison with original benchmarks are the root-mean-square velocity over the whole domain and over the surface
-\begin{eqnarray}
-v_{\textrm{rms}}&=&\sqrt{\int_{\Omega}|\bm{v}|^2 d\Omega}\nonumber \\
-v_{\textrm{rms}}^{top}&=&\sqrt{\int_0^1 u^2 \bigg|_{y=1} dx}\nonumber
-\end{eqnarray}
-respectively, top and bottom Nusselt numbers
-\[\textrm{Nu}^{top/bottom}=-\int_0^1\frac{\partial T}{\partial y} \bigg|_{y=1/y=0} dx\]
-the average rate of work done against gravity
-\[\left\langle W \right\rangle=\int_{\Omega}T u_y d\Omega\]
-the average rate of viscous dissipation
-\[\left\langle \Phi \right\rangle=\int_{\Omega}\tau_{ij}\dot{\epsilon}_{ij} d\Omega\]
-and the percentage error between $\left\langle W \right\rangle$ and $\left\langle \Phi \right\rangle/Ra$
-\[\delta=\frac{\left|\left\langle W \right\rangle -\frac{\left\langle \Phi \right\rangle}{Ra} \right|}{\max \left(\left\langle W \right\rangle,
-\frac{\left\langle \Phi \right\rangle}{Ra}\right)}\times100\]
-
-\subsection{Stokes flow}\label{sec:stokes}
-The problem consists of determining velocity field $(u,v)$ and pressure $p$ in case of a manufactured solution with prescribed body forces such as
-\begin{eqnarray}
-b_1&=&(12-24y)x^4+(-24+48y)x^3+(-48y+72y^2-48y^3+12)x^2+\nonumber \\
-&&+(-2+24y-72y^2+48y^3)x+1-4y+12y^2-8y^3\nonumber \\
-b_2&=&(8-48y+48y^2)x^3+(-12+72y-72y^2)x^2+(4-24y+48y^2-48y^3+\nonumber \\
-&&+24y^4)x-12y^2+24y^3-12y^4\nonumber
-\end{eqnarray}
-for which the exact solution is:
-\begin{eqnarray}
-u(x,y)&=&x^2(1-x)^2(2y-6y^2+4y^3)\nonumber \\
-v(x,y)&=&-y^2(1-y)^2(2x-6x^2+4x^3)\nonumber \\
-p(x,y)&=&x(1-x)-\frac{1}{6}\nonumber
-\end{eqnarray}
-The domain is a unit square a constant viscosity ($\eta=1$) and the penalty parameter is set to $\lambda=10^7$. Velocity boundary conditions are set to no slip
-$(\bm{v}=\bm{0})$ on all boundaries. The problem is performed for different grid resolution between \num{8x8} and \num{1024x1024} elements.
-
-The errors between the analytical solution and the numerical prediction of the velocity and the pressure field are calculated as in Section \ref{sec:error}.
-Fig. \ref{fig:errors}a shows that both velocity and pressure field converge to the exact solution with the decrease of the element size, following the
-theoretical convergence rate. The convergence of pressure, in contrast with the observation by \citet{Donea2003} for $Q_1 \times P_0$ elements, indicate the
-effectiveness of the smoothing procedure. Smoothed pressure field for grid resolution of \num{128x128} elements is shown in Fig. \ref{fig:errors}b (continuous
-red line), in comparison with the analytical solution (dashed blue line). Solve times and memory usage needed to generate the solution are shown in
-Fig. \ref{fig:MUMPS}. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Solver/Stokes_Flow}.
-
-\begin{figure}
-\centering
-\includegraphics[width=\linewidth]{./Figures/errors.pdf}
-\caption{Velocity and pressure error for the Stokes flow experiment between generated and analytical solution as a function of element size (panel a),
-and comparison between smoothed pressure and analytical solution as function of $x$ coordinate for a grid resolution of \num{128x128} elements (panel b).}
-\label{fig:errors}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\linewidth]{./Figures/MUMPS.pdf}
-\caption{Panel a) Analysis, factorisation, solution and total solve times (panel a) and factorisation memory usage (panel b) as a function of the total number
-of degrees of freedom for the Stokes flow experiment.}
-\label{fig:MUMPS}
-\end{figure}
-
-\subsection{Markers advection}\label{sec:runge}
-The 2nd-order Runge-Kutta advection scheme in space is tested by means of the Zalesak disk test \citep{Zalesak1979,Thieulot2014}. The benchmark is performed
-in a unit square domain with a grid resolution of \num{32x32} elements and 1000000 markers and values of Courant number between 0.25 and 1.
-At $t=0$, the disk is centred at position $(0.5;0.75)$ with a radius $R=0.15$ and has a vertical fissure 0.05 wide and 0.2 high (orange in Fig. \ref{fig:runge})
-and with the velocity field prescribed in the entire domain as
-\begin{eqnarray}
-u(x,y)&=&2\pi\left(y-\frac{L_x}{2}\right)\nonumber \\
-v(x,y)&=&-2\pi\left(x-\frac{L_x}{2}\right)\nonumber
-\end{eqnarray}
-the markers are back to their initial location after a $2\pi$ rotation ($t=1$). The clockwise rotation of the disk at $t=0.25$, 0.5 and 0.75 
-is shown in Fig. \ref{fig:runge} (red, blue and green, respectively). The distance from the centre of 1 marker is calculated throughout the rotation to
-evaluate the error for different values of Courant number. As expected, the error increase with the increasing of the time step, as shown in 
-Fig. \ref{fig:runge_err}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/Zalesak_Disk}.
-
-\begin{figure}
-\centering
-\includegraphics[width=200px]{./Figures/Zalesak.pdf}
-\caption{Position of the Zalesak disk at $t=0$, 0.25, 0.5 and 0.75 (orange, red, blue and green, respectively) for the case with Courant number of 1.}
-\label{fig:runge}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/zal_marker.pdf}
-\caption{Distance from the centre as function of time for values of Courant number of  of 0.25, 0.3, 0.5, 0.75 and 1 (green, orange, blue, black and red,
-respectively). R indicates the distance from the centre considering a perfect circle trajectory.}
-\label{fig:runge_err}
-\end{figure}
-
-\subsection{Conservative Velocity Interpolation (CVI)}\label{sec:CVI}
-The CVI correction is checked by means of the Stokes flow experiment by \citet{Donea2003}, as presented in \ref{sec:stokes}. The advection of Lagrangian
-markers is performed using either a 2nd-order or a 4th-order Runge-Kutta scheme with CFL=0.5 and an initial random distribution of 25 markers per element.
-Fig. \ref{fig:CVI} shows the minimum and maximum values of markers per element throughout the experiment with and without the CVI correction and using
-different Runge-Kutta schemes. The introduction of the CVI correction allows the
-tests to run for times longer than $t=2000$ without show empty elements, while tests without the CVI correction stop at approximately $t=600$, when at least
-one element goes empty. Fig. \ref{fig:CVI_time} shows the different number of markers per element and the markers distribution at $t=600$ obtained without and
-with the CVI correction (panels b and c, respectively) in case of a 2nd-order Runge-Kutta scheme. All data can be found at
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/CVI}.
-\begin{figure}
-\centering
-\includegraphics[width=\linewidth]{./Figures/CVI.pdf}
-\caption{Maximum and minimum number of markers per element throughout the experiment, with (green and orange lines) or without (purple and blue lines)
-the CVI correction and using either a 2nd-order (blue and orange lines) or a 4th-order (purple and green lines) Runge-Kutta scheme.}
-\label{fig:CVI}
-\end{figure}
-\begin{figure}
-\centering
-\includegraphics[width=\linewidth]{./Figures/CVI_time.pdf}
-\caption{Number of markers per element and markers distribution through time using 2nd-order Runge-Kutta scheme.
-Panels a: number of markers per element (panel a1) and distribution of the markers (panel a2) after the first
-iteration; panels b and c: comparison between number of markers per element (panels b1 and c1) and markers distribution (panels b2 and c2) without and with
-the CVI correction at $t=600$; panels d: number of markers per element (panel d1) and markers distribution (panel d2) in case of CVI correction at $t=2000$.}
-\label{fig:CVI_time}
-\end{figure}
-
-\subsection{Poiseuille flow}\label{sec:poiseuille}
-The domain is a rectangle with $L_x=2$ and $L_y=1$ and constant density and viscosity ($\rho=1$ and $\eta=1$), gravity acceleration $\bm{g}=\bm{0}$ and penalty
-parameter $\lambda=10^8$. The grid is composed by \num{40x20} elements. Velocity boundary conditions are set to no slip $(\bm{v}=\bm{0})$ at the top and the
-bottom, and a parabolic profile is imposed on the sides, with $u=y(1-y)$ and $v=0$. The analytical solution is then given by:
-\begin{eqnarray}
-u(x,y)&=&y(1-y)\nonumber \\
-v(x,y,)&=&0\nonumber \\
-p(x,y)&=&2\eta\left(\frac{L_x}{2}-x\right)\nonumber
-\end{eqnarray}
-
-The velocity field predicted by the model follows the expected parabolic profile (Fig. \ref{fig:poiseuille}a and Fig. \ref{fig:poi_plot}a) and, in case of the
-classic penalty method with no iterations, the pressure is clearly related to the divergence by means of the penalty parameter (Fig. \ref{fig:poiseuille}b and
-c and Fig. \ref{fig:poi_plot}b). Fig. \ref{fig:poiseuille}d shows that one Uzawa iteration is sufficient to bring the divergence down to \num{e-15}, with no
-correlation with the pressure field. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Poiseuille%20Flow}.
-
-\begin{figure}
-\includegraphics[width=\linewidth]{./Figures/Poiseuille.pdf}
-\caption{Velocity field (panel a), pressure (panel b) and divergence velocity of a Poiseuille flow in case of the classic penalty method (no iterations) and
-after one Uzawa iteration (panel c and d, respectively).}
-\label{fig:poiseuille}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\linewidth]{./Figures/Analytical.pdf}
-\caption{Velocity (panel a) and pressure (panel b) field predicted by the model for a Poiseuille flow with respect with their analytical solutions. The
-velocity field is plotted as function of the vertical coordinate in $L_x/2$ and the pressure is plotted as function of the x coordinate in $L_y/2$.}
-\label{fig:poi_plot}
-\end{figure}
-
-\subsection{Instantaneous 2D sphere}\label{sec:ist_sphere}
-The domain is a unit square with gravity $\bm{g}=(0,-1)$. The fluid has constant density and viscosity ($\rho_f=1$ and $\eta_f=1$). The sphere is in the middle
-of the domain with a radius R = 0.123456798 and has constant density and viscosity ($\rho_s=10^{-2}+\rho_f$ and $\eta_s=10^3 \cdot \eta_f$).
-We distinguish three different types of velocity boundary conditions:
-\begin{enumerate}
-\item FS: free slip conditions on all sides.
-\item NS: no slip conditions on all sides.
-\item OT: free slip conditions on the sides and the bottom, and open boundary at the top.
-\end{enumerate}
-All three cases are performed for different grid resolutions, between \num{16x16} and \num{512x512} elements with 50 randomly distributed markers per element,
-and different average schemes for the viscosity (harmonic, geometric and arithmetic). The velocity in the centre of the sphere $\bm{v}(0.5;0.5)$), minimum and
-maximum velocities ($u_{min}$, $u_{max}$, $v_{min}$, $v_{max}$) and pressures ($p_{min}$, $p_{max}$) on the entire domain, $v_{\textrm{rms}}$ and average pressure on the
-entire domain ($p_{avg}$) are compared with solutions generated by ASPECT \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and all results
-can be found at \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere2D} and 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Instantaneous_2D_Sphere}. Results in terms of the $v_{\textrm{rms}}$ for NS model 
-are also shown in Fig. \ref{fig:inst_sphere}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/vrms_NS.pdf}
-\caption{$v_{\textrm{rms}}$ as function of element size for the instantaneous 2D sphere experiment in case of no slip boundary conditions and with different average
-schemes for the viscosity. Results are compared with results obtained by other numerical codes.}
-\label{fig:inst_sphere}
-\end{figure}
-
-\subsection{Rayleigh-Taylor instability}\label{sec:rayleigh}
-This problem has been originally presented by \citet{vanKeken1997} and here it is performed as the isoviscous case (Case 1a). The domain has $L_x=0.9142$ and
-$L_y=1$ and gravity $\bm{g}=(0,-1)$. Two fluids with same constant viscosities ($\eta_1=\eta_2=1$) and different densities ($\rho_1=1000$ and $\rho_2=1010$),
-with the lighter fluid at the bottom. The initial interface between the fluids is given by $y(x)=0.2+0.02 \cos \left(\frac{\pi x}{L_x}\right)$. The experiment
-is performed with different grid sizes (\num{50x50}, \num{80x80}, \num{100x100} and \num{256x256}). A total of 1960000 markers are randomly distributed at the
-beginning of the simulation. Velocity boundary conditions are set to no slip at the top and the bottom, and to free slip at the sides of the domain.
-
-$v_{\textrm{rms}}$ as function of time is reported in Fig. \ref{fig:RT}, matching well with results shown by \citet{vanKeken1997}, \citet{Tackley2003} and
-\citet{Thieulot2014}. Fig. \ref{fig:rayleigh} shows the evolution of the experiment at different time steps. All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Rayleigh_Taylor_experiment/ISOVISCOUS}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/RT.pdf}
-\caption{$v_{\textrm{rms}}$ of the Rayleigh-Taylor experiment as function of time for different resolution of the grid.}
-\label{fig:RT}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=300px]{./Figures/Rayleigh.pdf}
-\caption{Evolution of the Rayleigh-Taylor experiment for a grid of \num{256x256} elements at $t=0, 500, 1000$ and $2000$ (panels a, b, c and d, respectively).}
-\label{fig:rayleigh}
-\end{figure}
-
 \subsection{Falling blocks}\label{sec:block}
 This benchmark is described as proposed by \citet{Gerya2003a}, \citet{Gerya2010} and \citet{Thieulot2011}. The domain is square with $L_x=L_y=$ \SI{500}{\km}
-and the grid is composed by \num{50x50} elements with 25 markers in each element. The block is initially centred at ($x=$ \SI{250}{\km}; $y=$ \SI{400}{\km})
-and has a size of \num{100x100} km (Fig. \ref{fig:block}a). The fluid surrounding the block has $\eta_f=$ \SI{e21}{\pascal\per\s} and $\rho_f=$
+and the grid is composed by $50\times50$ elements with 25 markers in each element. The block is initially centred at ($x=$ \SI{250}{\km}; $y=$ \SI{400}{\km})
+and has a size of $100\times100$ km (Fig. \ref{fig:block}a). The fluid surrounding the block has $\eta_f=$ \SI{e21}{\pascal\per\s} and $\rho_f=$
 \SI{3200}{\kg\per\cubic\m}. The benchmark tests with different viscosities of the block, with $\eta_b$ from \SIrange{e15}{e27}{\pascal\per\s}. In each
@@
 \end{figure}
-
-\subsection{2D time-dependent Stokes sphere with deformable free surface}\label{sec:sphere}
-This experiment is performed in a unit square domain with the gravity acceleration fixed to $g_y=-1$. The sphere has $\rho_s=2$ and $\eta_s=10^3$ and is
-initially centred at ($0.5;0.6$) with radius $R=0.123456789$. The fluid surrounding the sphere has $\rho_f=1$ and $\eta_f=1$ and occupies the domain for
-$y\leq 0.75$, while for $y>0.75$ the air has $\rho_a=0$ and $\eta_a=10^{-3}$. Velocity boundary conditions are set to free slip on all sides. The Courant
-number is set to 0.25. The experiments run for 200 s using grid resolutions from \num{150x150} to \num{512x512} elements, with 25 randomly distributed markers
-per element, and for different average schemes for the viscosity. The interface between the fluid and the air is tracked by means of the markers chain. Results
-in terms of velocity and pressure fields and topography variations are compared with results obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and they can be found at
-\url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere_fs2D} and
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Time_dependent_sphere}. Results in terms of the $v_{\textrm{rms}}$ in case of 
-an arithmetic average are also shown in Fig. \ref{fig:inst_spherefs}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/vrms_arithm.pdf}
-\caption{$v_{\textrm{rms}}$ for different grid resolutions as function of time for the 2D time-dependent Stokes sphere experiment in case of an arithmetic average.
-Results are compared with results obtained by ASPECT.}
-\label{fig:inst_spherefs}
-\end{figure}
-
-\subsection{Free surface stabilisation}\label{sec:stab}
-The benchmark is performed as discussed in \citet{Kaus2010a} and \citet{Thieulot2014}. The domain is square with $L_x=L_y=$ \SI{500}{\km} and a grid resolution
-of \num{200x200} elements, each of them containing 16 markers. A buoyant fluid with $\rho_1=$ \SI{3200}{\kg\per\cubic\metre} and $\eta_1=$ \SI{e20}{\pascal\s}
-is overlain by a denser fluid with $\rho_2=$ \SI{3300}{\kg\per\cubic\metre} and $\eta_2=$ \SI{e21}{\pascal\s}. The initial interface between the fluids has an
-initial sinusoidal shape of 5 km amplitude. Velocity boundary conditions are set to free slip conditions at the sides and no slip at the bottom, while the top
-is free surface. The experiment is performed with various fixed time steps, without the use of the Courant number. The vertical position of the free surface at
-$x=L_x$ is tracked for each simulation, with or without the stabilisation algorithm.
-
-The results show an instability (drunken sailor effect) increasing the time step in case that the stabilisation algorithm is not activated
-($dt=$ \SI{4200}{\year}, green line in Fig. \ref{fig:stab}), as already observed by \citet{Kaus2010a} and \citet{Thieulot2014}. Activating the algorithm the
-instability is fixed and the simulation remains stable using time step up to \SI{50000}{\year} (red line in Fig. \ref{fig:stab}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Stabilisation_algorithm}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Kaus_random.pdf}
-\caption{Evolution of the y coordinate in $x=L_x$ as function of time for different time steps, with and without the stabilisation algorithm.}
-\label{fig:stab}
-\end{figure}
-
-\subsection{Topography relaxation}\label{sec:crameri}
-This experiment is performed as in \citet{Crameri2012} using free surface and sticky air (Case 1). The domain is rectangular with $L_x=$ \SI{2800}{\km} and
-$L_y=$ \SI{700}{\km} in case of free surface and $L_y=$ \SI{800}{\km} in case of sticky air ($h_{st}=$ \SI{100}{\km}). The grid is composed by \num{256x64}
-elements for the free surface \num{560x320} elements for the sticky air. For both cases, there is a mantle of 600 km thickness with $\eta_m=$ \SI{e21}{\pascal\s}
-overlain by a highly viscous cosine-shaped lithosphere with thickness between 93 and 107 km and $\eta_l=$ \SI{e23}{\pascal\s}. Both mediums have $\rho=$
-\SI{3300}{\kg\per\cubic\metre}. In case of sticky air experiment, the air has $\rho_a=$ \SI{0}{\kg\per\cubic\metre} and $\eta_a=$ \SI{e18}{\pascal\s}. The
-gravity acceleration is set to $g_y=$ \SI{-10}{\metre\per\square\s}. Velocity boundary conditions are set to no slip at the bottom and free slip at the sides
-of the domain. In case of sticky air, velocities on top are set to free slip conditions.
-
-The maximum topography as function of time ($h(t)$) can be analytically derived in according to \[h(t)=h_0 \exp(\gamma t)\]
-where $h_0=7$ km is the initial topography, $\gamma=\num{-0.2139e-11}$ is the characteristic relaxation rate and $t$ is the time (black line in Fig.
-\ref{fig:crameri}). Maximum topography at the characteristic relaxation time $t_{rlx}=$ \SI{14.825}{\kilo\year} can be found to be $h_{rlx}=$ \SI{2576}{\m}.
-The results show that a viscosity of the air of \SI{e18}{\pascal\s} is needed in case of $h_{st}=$ \SI{100}{\km} to correctly describe the topography relaxation
-for this problem, as pointed out by \citet{Crameri2012} (purple line in Fig. \ref{fig:crameri}). The case with a true free surface well follows the analytical
-solution (orange line in Fig. \ref{fig:crameri}). In particular, at the relaxation time in case of the true free surface predicts a maximum topography of
-\SI{2570}{\m}, with an error of \SI{6}{\m} with respect to the analytical solution. Results are also compared with results obtained with a free surface in
-UNDERWORLD (\num{256x64} elements) and SULEC (\num{401x201} elements) (with courtesy from \citealp{Crameri2012}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Topography_relaxation}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Crameri.pdf}
-\caption{Maximum topography as function of time for the Crameri benchmark (coloured lines) in comparison with the analytical solution (black line) and results
-obtained with UNDERWORLD (yellow line) and SULEC (blue line).}
-\label{fig:crameri}
-\end{figure}
-
-\subsection{Spontaneous subduction}\label{sec:subduction}
-This experiment is performed as presented by \citet{Schmeling2008}, in case of both a sticky air and a true free surface and using different average schemes
-for the viscosities. The domain is rectangular with $L_x=$ \SI{3000}{\km} and $L_y=$ \SI{700}{\km}, a grid resolution of \num{256x64} elements with 50 markers
-per element and Courant number of 0.01. At the beginning of the simulation, a 100 km-thick lithospheric layer with $\rho_m=$ \SI{3300}{\kg\per\cubic\metre}
-and $\eta_m=$ \SI{e23}{\pascal\s} is located at the top of the domain for $x$ between 1000 and 3000 km, while the asthenospheric mantle has $\rho_m=$
-\SI{3200}{\pascal\s} and $\eta_m=$ \SI{e21}{\pascal\s}. In addition, a 200 km-depth lithospheric slab is already subducted in the mantle in order to have
-a spontaneous subduction. In case of sticky air, $L_y=750$ km and a 50 km-thick air layer with $\eta_a$ from \SIrange{e19}{e21}{\pascal\s} overlie the mantle.
-Velocity boundary conditions are set to free slip on the sides and on the bottom of the domain. In case of sticky air, velocities on the top boundary are set
-to free slip conditions as well.
-
-The maximum depth of the slab is tracked for all the simulations (Fig. \ref{fig:subduction}) and they are compared with results obtained by
-\citet{Schmeling2008} with models with regular grids and comparable grid resolutions (rectangles in Fig. \ref{fig:subduction}). Low viscous sticky air models
-(i.e., \SI{e19}{\pascal\s}) enlighten a strong dependence of the sinking velocity with the chosen average scheme (dashed lines in Fig. \ref{fig:subduction}),
-while high viscous sticky air models (i.e., \SI{e21}{\pascal\s}) show a higher resistance at the trench, underestimating the correct solution (dotted lines in
-Fig. \ref{fig:subduction}). A high resistance at the trench can be also observed in case of a true free surface for the low resolution grid used (continuous
-lines in Fig. \ref{fig:subduction}), as pointed out by \citet{Schmeling2008}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Spontaneous_subduction}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Subduction.pdf}
-\caption{Maximum depth of the slab as function of time for all the simulations of the \citet{Schmeling2008} benchmark, in case of harmonic, geometric and
-arithmetic means (red, green and blue lines, respectively) using a sticky air or a true free surface (discontinuous and continuous lines, respectively).
-Red, green and blue rectangular areas indicate the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{500}{\km} in case of sticky air,
-while the grey rectangular area indicates the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{600}{\km} in case of true free surface.}
-\label{fig:subduction}
-\end{figure}
-
-\subsection{The slab detachment}\label{sec:slab}
-The slab detachment benchmark is performed as described by \citet{Schmalholz2011} and \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{1000}{\km}
-and $L_y=$ \SI{660}{\km} and a grid resolution of \num{256x256} elements. A non-linear viscous T-shaped layer with $\rho_l=$ \SI{3300}{\kg\per\cubic\metre} is
-placed at the top of the domain and surrounded by a linear viscous fluid with $\rho_f=$ \SI{3150}{\kg\per\cubic\m}. The top layer is 80 km-thick and a 250
-km-long and 80 km-wide slab is placed at $x=L_x/2$ (Fig. \ref{fig:slab}a). The effective viscosity of the top layer is given by
-\[\eta_{eff}=\eta_0 I_2^{\frac{(1-n)}{n}}\]
-with $\eta_0=$ \SI{4.75e11}{\pascal\s} and $n=4$, while the surrounding fluid has $\eta_f=$ \SI{e21}{\pascal\s}. The effective viscosity is capped between
-$\eta_{min}=$ \SI{e21}{\pascal\s} and $\eta_{max}=$ \SI{e25}{\pascal\s}. Velocity boundary conditions are set to free slip at the top and the bottom, and to
-no slip at the sides of the domain.
-
-The sinking of the slab determines a decrease of the effective viscosity in the T-shaped layer during the first part of the evolution (Fig. \ref{fig:slab}a),
-as consequence of local strain rates, while effective viscosities increase up to $\eta_{max}$ in the top layer after approximately \SI{20}{\mega\year}, when
-necking is complete (Fig. \ref{fig:slab}b).
-
-The width of necking is tracked by means of markers position at the side of the slab. Necking width and time are normalised in function of the initial slab
-(\SI{80}{\km}) and the characteristic time ($t_c=$ \SI{7.1158e14}{\s}) \citep{Schmalholz2011,Glerum2018}. The results in case of arithmetic, geometric and
-harmonic mean of the viscosity are shown in Fig. \ref{fig:necking}, compared with results from \citet{Glerum2018} for which an infinity norm mean was used.
-The influence of the average schemes agrees with results shown in Section \ref{sec:subduction}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Slab_detachment}.
-
-\begin{figure}
-\includegraphics[width=\linewidth]{./Figures/Slab.pdf}
-\caption{Effective viscosity for the slab detachment benchmark at the beginning of the evolution (panel a) and when necking is complete (panel b).}
-\label{fig:slab}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Necking.pdf}
-\caption{Normalised width of the necking of the slab detachment benchmark as function of normalised time.}
-\label{fig:necking}
-\end{figure}
-
-\subsection{The indenter experiment}\label{sec:indenter}
-The indenter benchmark simulate a rigid punch on a purely plastic von Mises material, for which exists an analytical solution
-\citep{Thieulot2008,Thieulot2014,Glerum2018}. The domain is a unit square with a grid resolution of \num{256x256} elements. The gravity acceleration is set
-to $\bm{g}=\bm{0}$ and effective viscosity is capped using $\eta_{min}=10^{-4}$ and $\eta_{max}=10^3$. The tolerance for the convergence of the non-linear
-solution is $tol=10^{-9}$, with a maximum number of non-linear iterations set to 500. The medium in the domain has $\rho=0.01$, an initial viscosity
-$\eta_0=10$, cohesion $C=1$ and an angle of internal friction $\phi=0$°. Velocity boundary conditions are set to no slip at the bottom and free slip at the
-sides of the domain. The top of the domain is open with the exception of the central portion (punch area) with width $w_p=0.125$, where $v=-1.05$ and $u$ is
-fixed either to 0 (rough punch experiment) or free (smooth punch experiment).
-
-Fig. \ref{fig:indenter} show the results in terms of viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure
-(panels g and h) for both experiments. The analytical solution indicate that pressure at ($0.5;1$) and ($0.5 \pm w_p;1$)is $p=\pi +1$ and $p=1$, respectively
-(black and red lines in Fig. \ref{fig:smooth_p}). Fig. \ref{fig:smooth_p} shows a clear improvement in the pressure solution when passing from a rough to a
-smooth punch (green and blue lines in Fig. \ref{fig:smooth_p}, respectively), as previously observed by \citet{Thieulot2014} and \citet{Glerum2018}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Indenter}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Indenter.pdf}
-\caption{Viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure (panels g and h) fields for rough (left column) and
-smooth (right column) punch experiments.}
-\label{fig:indenter}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Smooth_Rough.pdf}
-\caption{Nodal pressure in function of the x coordinate at $y=1$ for rough (green line) and smooth (blue line) punch experiment. Black and red lines indicate
-the analytical solution at $x=0.5$ and $x=0.5 \pm w_p$, respectively.}
-\label{fig:smooth_p}
-\end{figure}
-
-\subsection{The brick experiment}\label{sec:brick}
-An instantaneous version of the brick benchmark is performed to verify the correctness of the pressure-dependent plasticity for different angle of internal
-friction, as proposed by \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{40}{\km} and $L_y=$ \SI{10}{\km} and a grid resolution of \num{512x128}
-elements. The gravity acceleration is set to $g_y=$ \SI{-10}{\m\square\s} and effective viscosity is capped using $\eta_{min}=$ \SI{e19}{\pascal\s} and
-$\eta_{max}=$ \SI{e26}{\pascal\s}. The tolerance for the convergence of the non-linear solution is $tol=10^{-7}$, with a maximum number of non-linear iterations
-set to 1000. A 800 m-wide and 400 m-height inclusion with $\eta_b=$ \SI{e20}{\pascal\s} and $\rho_b=$ \SI{2700}{\kg\per\cubic\m} is placed at the bottom of the
-domain at $x=L_x/2$. The inclusion is surrounded by a non-linear viscous medium with $\rho_m=$ \SI{2700}{\kg\per\cubic\m}, an initial viscosity of
-\SI{e23}{\pascal\s}, a linear viscous viscosity of \SI{e25}{\pascal\s} and a cohesion of \SI{40}{\mega\pascal}. Velocities are set to free slip conditions at
-the bottom of the domain and the top is open. Velocities on sides of the domain are fixed to $u=$ \SI{+-2e-11}{\m\per\s} and $v=0$. The experiment is performed
-in compressional and extensional contexts with an angle of internal friction $\phi$ of the non-linear medium variable between 0° and 30°.
-
-As expected, two shear bands stem from the inclusion with variable angles in relation to both the dynamics context and the internal friction angle. Shear band
-angles formed at 45° for $\phi=0$° in both extensional and compressional contexts (Fig. \ref{fig:brick_beam}a), while they are different in case of $\phi\neq0$°.
-Shear band angles in case of $\phi=20$° are shown for compression and extension in Fig. \ref{fig:brick_beam}b  and c, respectively. Values of shear band angles
-as function of internal friction angles from 0° to 30° are extracted at different depths and minimum and maximum values are plotted in Fig. \ref{fig:brick},
-in comparison with theoretical Roscoe, Arthur and Coulomb shear band angles. For all tests the tolerance for nonlinear convergence is defined $tol=10^{-7}$
-and none converges before the maximum number of iterations ($it_{max}=1000$) is reached. However, tests with internal friction angles up to 15° show a constant
-decrease of the velocity residuals, which can not be observed for for tests with higher internal friction angles (Fig. \ref{fig:convergence}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Brick_experiment}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Brick_Beam.pdf}
-\caption{Shear band angles predicted for the brick experiment for $\phi=0$° (panel a) and for $\phi=20$° in case of compression and extension (panels b and c,
-respectively).}
-\label{fig:brick_beam}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Brick.pdf}
-\caption{Shear band angles predicted for the brick experiment in case of compressional and extensional contexts as function of different internal angle of
-friction (continuous black line and red dots), compared with theoretical Roscoe, Artur and Coulomb angles (discontinuous black lines). Red lines indicate the
-range of angles calculated at different depths.}
-\label{fig:brick}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=500px]{./Figures/convergence.pdf}
-  \caption{Convergence rates for all tests. Normalised residuals (left column) are compared with velocity residuals (right column) for extensional (first row)
-  and compressional (second row) settings.}
-  \label{fig:convergence}
-  \end{figure}
-
-\subsection{Advection stabilisation}\label{sec:advection}
-The 1D advection problem proposed by \citet{Donea2003} and \citet{Thieulot2011} is performed to verify the effectiveness of the SUPG method to stabilise the
-advection term of the energy equation. The domain is a 1D segment with $L_x=1$ composed by 50 elements and a discontinuity in the thermal field placed at
-$x=0.25$. Temperature is set to 1 for $x \leq 0.25$ and to 0 for $x > 0.25$. Velocity is set to ${u}=1$ in the entire domain. The simulation is performed for
-250 time steps, with $dt=0.002$, so the thermal discontinuity should be at $x=0.75$ at the end of the simulation.
-
-Temperature profiles at the end of the simulation are shown in \ref{fig:advection} as function of the dimensionless coefficient $\gamma=\tau \bm{v}/h$. In case
-of the classic Galerkin method ($\gamma=0$, blue line in \ref{fig:advection}) the final thermal profile is characterised by strong oscillations, which are
-eliminated in case of the SUPG method ($\gamma=0.045$, orange line in \ref{fig:advection}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Advection_stabilisation}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Advection.pdf}
-\caption{Temperature profile as function of \textit{x} for the advection stabilisation benchmark. Purple line indicates the initial temperature profile; the
-green line indicates the analytical temperature profile after 250 time steps; blue line indicates the temperature profile after 250 time steps in case of the
-classic Galerkin method ($\gamma=0$); orange line indicates the temperature profile after 250 time steps in case of the SUPG method ($\gamma=0.045$).}
-\label{fig:advection}
-\end{figure}
-
-\subsection{Shear heating}\label{sec:simple_shear}
-This exercise is performed in an unit square domain composed by \num{128x128} elements. The velocity field is prescribed on the entire domain with
-$\bm{v}=(L_y-y)y\bm{e_x}$; viscosity, density and specific heat are fixed to 1, while thermal conductivity and radiogenic energy are fixed to 0. Therefore,
-the energy equation (Eq. \ref{eq:energy}) can be simplified as \[\frac{\partial T}{\partial t}=H_s\] and fixing $T(t=0)=0$, the temperature can be find as
-\[T(t)=H_s t\]
-In this case we have
-\begin{eqnarray}
-\dot{\epsilon_{xx}}&=&\frac{\partial u}{\partial x}=0 \nonumber \\
-\dot{\epsilon_{yy}}&=&\frac{\partial v}{\partial y}=0 \nonumber \\
-\dot{\epsilon_{xy}}&=&\frac{1}{2}\left(\frac{\partial v}{\partial x}+\frac{\partial u}{\partial y}\right)=\frac{1}{2}(L_y-2y)\nonumber
-\end{eqnarray}
-and, simplifying Eq. \ref{eq:shear}, shear heating can be calculated as \[H_s(x,y)=(1-2y)^2\]
-
-The solution predicted by the model in terms of velocity, temperature and shear heating match well with the analytical solutions (Fig. \ref{fig:analytical_en}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Simple_shear}.
-
-%\begin{figure}
-%\centering
-%\includegraphics[width=400px]{./Figures/Energy_Thieulot.pdf}
-%\caption{Temperature and shear heating predicted by the model after 1 time step.}
-%\label{fig:en_thieulot}
-%\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=500px]{./Figures/Analytical_Shear.pdf}
-\caption{Velocity (panel a), temperature (panel b) and shear heating (panel c) as function of the \textit{y} coordinate for the simple shear experiment. The
-solutions predicted by the model (red lines) are compared with the analytical solutions (dashed blue lines).}
-\label{fig:analytical_en}
-\end{figure}
-
-\subsection{Shear and adiabatic heating}\label{sec:shear}
-This problem is performed as presented in Exercise 9.4 in \citet{Gerya2010}. Two materials are vertically separated in a rectangular domain with $L_x=$
-\SI{1000}{\km}, $L_y=$ \SI{1500}{\km} and a grid resolution of \num{30x20} elements. Constant thermal coefficient expansion ($\alpha=$ \SI{3e-5}{\per\kelvin}),
-temperature ($T=$ \SI{1300}{\kelvin}) and gravity acceleration ($g_y=$ \SI{-10}{\m\square\s}) are assumed in the whole domain. Fluid 1 (on the left side of the
-domain) has $\rho_1=$ \SI{3200}{\kg\per\cubic\m} and $\eta_1=$ \SI{e20}{\pascal\s}; fluid 2 (on the right side of the domain) has $\rho_2=$
-\SI{3300}{\kg\per\cubic\m} and $\eta_2=$ \SI{e22}{\pascal\s}. Velocity boundary conditions are set to free slip on all sides of the domain.
-
-As shown in Fig. \ref{fig:energy}, both shear and adiabatic heating predicted by the code (first row in Fig. \ref{fig:energy}) well recreate the results
-obtained by means of example ${Shear\_adiabatic\_heating.m}$ from \citet{Gerya2010} (second row in Fig. \ref{fig:energy}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Adiabatic%2BShear}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Energy.pdf}
-\caption{Comparison between shear (first column), adiabatic (second column) and total (third column) energy predicted by the code (panels a, b and c) and those
-created using example ${Shear\_adiabatic\_heating.m}$ from Exercise 9.4 in \citet{Gerya2010} (panels d, e and f).}
-\label{fig:energy}
-\end{figure}
-
-\subsection{Mantle convection}\label{sec:mantle}
-This problem is performed as presented by \citet{Blankenbach1989} (constant viscosity cases) and \citet{Thieulot2014}, in a 2D unit square domain with gravity
-acceleration $g_y=10^{10}Ra$. The experiment is performed with three different Rayleigh numbers ($Ra=10^4$, $10^5$ and $10^6$) and with different grid
-resolution (between \num{32x32} and \num{128x128} elements). The fluid has constant viscosity, initial density, heat capacity, thermal conductivity
-($\eta= \rho_0= C_p = k=1$), reference temperature ($T_0=0$) and thermal expansion coefficient ($\alpha=10^{-10}$). Temperatures are set to 0 on top and 1 on
-bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\]
-
-The solution generated by the code in terms of $v_{\textrm{rms}}$ (calculated as in Sec. \ref{sec:error}) and Nu as function of time are reported for all the
-simulations in Fig. \ref{fig:mantle}. At the steady state, $v_{\textrm{rms}}$ and Nu of all simulations converge well toward the values from \citet{Blankenbach1989},
-with lower errors for higher resolution grids (Table \ref{tab:mantle}). All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Mantle_convection}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Convection.pdf}
-\caption{$v_{\textrm{rms}}$ (panels a, c and e) and Nu (panels b, d and f) for the mantle convection benchmark as function of time for different grid resolution.
-Panels a and b show the results for $Ra=10^4$; Panels c and d show the results for $Ra=10^5$; panels e and f show the results for $Ra=10^6$. Purple lines
-indicate the convergence values for the $v_{\textrm{rms}}$ and the Nu from \citet{Blankenbach1989} at the steady state.}
-\label{fig:mantle}
-\end{figure}
-
-\begin{table}[H]
-\caption{Comparison between $v_{\textrm{rms}}$ and Nu predicted by the code for the mantle convection experiment and same values as reported in literature.}
-\centering
-\small
-\begin{tabular}{|>{\centering}m{0.09\linewidth}|>{\centering}m{0.05\linewidth}|>{\centering}m{0.24\linewidth}|>{\centering}m{0.15\linewidth}|
-  >{\centering}m{0.1\linewidth} >{\centering}m{0.1\linewidth} >{\centering\arraybackslash}m{0.1\linewidth}|}
-\toprule
-& & \multirow{2}{*}{\citet{Blankenbach1989}} & \citet{Thieulot2014} & \multicolumn{3}{c|}{FALCON}  \\
- & & & \num{200x200} & \num{32x32} & \num{64x64} & \num{128x128}  \\
-\midrule
-\multirow{2}{*}{$Ra=10^4$} & $v_{\textrm{rms}}$ & $42.864947 \pm 0.000020$ & 42.867 & 42.83226 & 42.852793 & 42.861394  \\
-           & Nu      & $4.884409 \pm 0.000010$  & 4.882  & 4.781297 & 4.857475  & 4.877573   \\
-\hline
-\multirow{2}{*}{$Ra=10^5$} & $v_{\textrm{rms}}$ & $193.21454 \pm 0.00010$  & 193.255& 193.872643 & 193.377472 & 193.252290  \\
-           & Nu      & $10.534095 \pm 0.000010$ & 10.507 & 9.602514 & 10.270735  & 10.465629   \\
-\hline
-\multirow{2}{*}{$Ra=10^6$} & $v_{\textrm{rms}}$ & $833.98977 \pm 0.00020$  & 834.712& 848.091176 & 837.767911 & 834.945793  \\
-           & Nu      & $21.972465 \pm 0.000020$ & 21.695 & 15.999266 & 19.703682  & 21.306939   \\
-\bottomrule
-\end{tabular}
-\label{tab:mantle}
-\end{table}
-
-\subsection{Viscoplastic mantle convection}\label{sec:visco_mantle}
-This benchmark is performed as Cases 1-5a described in \citet{Tosi2015} in a 2D unit square domain with two different grid resolutions of \num{32x32} 
-and \num{100x100} elements. Reference viscosity, reference density, heat capacity, thermal conductivity and thermal expansion coefficient are fixed to a constant
-value of 1 ($\eta_0= \rho_0= C_p = k=\alpha=1$), while the gravity has been chosen as $g_y=10^2$, in order to obtain a Rayleigh number $Ra=10^2$. Temperatures
-are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\].
-The viscosity field $\eta$ is calculated as the harmonic average between a linear part $\eta_{lin}$ that depends on temperature only or on temperature and 
-depth $y$ and a nonlinear-plastic part $\eta_{pl}$ that depends on the strain rate $\dot{\bm{\epsilon}}$, as follows
-\[\eta(T,y,\dot{\bm{\epsilon}})=2\left(\frac{1}{\eta_{lin}(T,y)}+\frac{1}{\eta_{pl}(\dot{\bm{\epsilon}})}\right)^{-1}\]
-The linear and the nonlinear-plastic parts of the viscosity are calculated following \citet{Tosi2015} as follows
-\[\eta_{lin}(T,y)=\exp(-\ln(\Delta\eta_T)T+\ln(\Delta\eta_y)y)\]
-\[\eta_{pl}(\dot{\bm{\epsilon}})=\eta^*+\frac{\sigma_Y}{\sqrt{\dot{\bm{\epsilon}}:\dot{\bm{\epsilon}}}}\]
-where $\Delta\eta_T$, $\Delta\eta_y$, $\eta^*$ and $\sigma_Y$ are parameters chosen as listed in Table \ref{tab:tosi_setup}.
-
-For each case, some diagnostic quantities measured at the steady state are reported in Table \ref{tab:visco_mantle} (see Section \ref{sec:error} for 
-the computation of all the quantities), showing a well fit with results described in \citet{Tosi2015}. Data and charts for all cases can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Viscoplastic_mantle_convection}.
-
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{white!100!}{gray!25!}
-%\begin{table}[H]
-%  \caption{Results of the viscoplastic mantle convection benchmark for Cases 1-4 compared with results obtained with ELEFANT and YACC (from \citealp{Tosi2015}).}
-%  \centering
-%  \resizebox{\linewidth}{!}{
-%\begin{tabular}{|c|ccc|ccc|ccc|ccc|}
-%  \toprule
-%    & \multicolumn{3}{c|}{\textbf{Case 1}} & \multicolumn{3}{c|}{\textbf{Case 2}} & \multicolumn{3}{c|}{\textbf{Case 3}} &
-%    \multicolumn{3}{c|}{\textbf{Case 4}} \\
-%   & FALCON & ELEFANT& YACC & FALCON& ELEFANT& YACC & FALCON & ELEFANT & YACC & FALCON & ELEFANT & YACC \\
-%  \midrule
-%   $\left\langle T \right\rangle$ & 0.7759 & 0.7758 & 0.7767 & 0.6043 & 0.6036 & 0.6058 & 0.7275 & 0.7275 & 0.7286 & & 0.5277 & 0.5289\\
-%    Nu$^{top}$ & 3.4280 & 3.4214 & 3.4298 & 8.5242 & 8.5115 & 8.5278 & 3.0368 & 3.0347 & 3.0374 && 6.5912 & 6.5572 \\
-%    Nu$^{bottom}$ & 3.3168 & 3.313 & 3.3143 & 8.4599 & 8.4465 & 8.3990 & 3.0086 & 2.9908 & 2.9628 && 6.5834 & 6.5243 \\
-%    $v_{\textrm{rms}}$ & 249.901 & 249.134 & 251.7997 & 141.2451 & 140.7087 & 142.2020 & 100.2776 & 100.1208 & 100.9467 && 79.1105 & 79.6202 \\
-%    $v_{\textrm{rms}}^{top}$ & 1.8907 & 1.8642 & 1.8298 & 104.7928 & 104.1228 & 105.1767 & 2.0832 & 2.0652 & 2.0374 && 74.7596 & 75.8414 \\
-%    $u_{max}^{top}$ & 2.6340 & 2.6119 & 2.5516 & 121.9212& 121.766 & 122.3238 & 2.9086 & 2.9019 & 2.8458 && 88.9146 & 89.2940 \\
-%    $\eta_{min}$ & - & - & - & $2.0303\times10^{-5}$ & $2.0600\times10^{-5}$ & $1.9853\times10^{-5}$ & $4.8129\times10^{-5}$ & $4.8080\times10^{-5}$ & $4.7907\times10^{-5}$ &
-%    & $1.9860\times10^{-4}$ & $1.9174\times10^{-4}$ \\
-%    $\eta_{max}$ & - & - & - & 1.3325 & 1.85 & 1.6095 & 0.9360 & 0.9023 & 1 && 1.52 & 1.6773 \\
-%    $\left\langle W \right\rangle$ & 2.4299 & 2.4316 & 2.4583 & 7.5648 & 7.5665 & 7.6222 & 2.0387 & 2.0384 & 2.0400 && 5.6216 & 5.6512 \\
-%    $\left\langle \Phi \right\rangle/Ra$ & 2.4311 & 2.4276 & 2.4333 & 7.5661 & 7.5585 & 7.6194 & 2.0386 & 2.0370 & 2.0335 && 5.6182 & 5.6463 \\
-%    $\delta$ & 0.05\% & 0.16\% & 1.02\% & 0.02\% & 0.11\% & 0.04\% & $<$0.01\% & 0.07\% & 0.32\% && 0.06\% & 0.09\% \\
-%  \bottomrule
-%  \end{tabular}}
-%  \label{tab:visco_mantle}
-%  \end{table}
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{2}{white!100!}{gray!25!}
-\begin{table}[H]
-  \caption{Benchmark cases and corresponding parameters.}
-  \centering
-  \small
-\begin{tabular}{|>{\centering}m{0.08\linewidth}|>{\centering}m{0.05\linewidth}>{\centering}m{0.05\linewidth}>{\centering}m{0.05\linewidth}
-  >{\centering}m{0.05\linewidth} >{\centering}m{0.05\linewidth} |>{\centering\arraybackslash}m{0.2\linewidth}|}
-  \toprule
-    Case & $Ra$ & $\Delta\eta_T$ & $\Delta\eta_y$ & $\eta^*$ & $\sigma_Y$ & Convective regime \\
-  \midrule
-  \rowcolor{gray!25!}
-    1 & $10^2$ & $10^5$ & 1 & - & - & Stagnant lid \\
-    2 & $10^2$ & $10^5$ & 1 & $10^{-3}$ & 1 & Mobile lid  \\
-    \rowcolor{gray!25!}
-    3 & $10^2$ & $10^5$ & 10 & - & - & Stagnant lid  \\
-    4 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 1 & Mobile lid \\
-    \rowcolor{gray!25!}
-    5 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 4 & Periodic \\
-  \bottomrule
-  \end{tabular}
-  \label{tab:tosi_setup}
-  \end{table}
-
-\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{gray!25!}{white!100!}
-\begin{table}
-  \caption{Results of the viscoplastic mantle convection benchmark for different grid resolutions in Cases 1-5 compared with results obtained with ELEFANT
-  and YACC ($100\times100$ elements, from \citealp{Tosi2015}) and in Stone 28 ($32\times32$ elements, from
-  \url{https://github.com/cedrict/fieldstone/tree/master/python_codes/fieldstone_28}).}
-  \centering
-  \resizebox{\linewidth}{!}{
-\begin{tabular}{|cc|ccccccccccc|}
-  \toprule
-  \multicolumn{13}{|c|}{\textbf{Case 1}} \\
-  \midrule
-  Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-  & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-  \hline
-  \rowcolor{gray!25!}
-    YACC & $100\times100$ & 0.7767 & 3.4298 & 3.3143 & 251.7997 & 1.8298 & 2.5516 & - & - & 2.4583 & 2.4333 & 1.02\% \\
-    ELEFANT & $100\times100$ & 0.7758 & 3.4214 & 3.313 & 249.134 & 1.8642 & 2.6119 & - & - & 2.4316 & 2.4276 & 0.16\% \\
-    \rowcolor{gray!25!}
-    Stone 28 & $32\times32$ & 0.7737 & 3.3987 & - & 243.872 & - & - & - & - & - & - & - \\
-%    FALCON & $32\times32$ & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & - & - & 2.4745 & 2.4864 & 0.48\% \\
-%    \rowcolor{gray!25!}
-    FALCON & $100\times100$ & 0.7759 & 3.4280 & 3.3168 & 249.901 & 1.8907 & 2.6340 & - & - & 2.4299 & 2.4311 & 0.05\% \\
-    \midrule
-    \multicolumn{13}{|c|}{\textbf{Case 2}} \\
-    \midrule
-    Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-    & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-    \hline
-    \rowcolor{gray!25!}
-     YACC & $100\times100$ & 0.6058 & 8.5278 & 8.3990 & 142.2020 & 105.1767 & 122.3238 & $1.9853\times10^{-5}$ & 1.6095 & 7.6222 & 7.6194 & 0.04\% \\
-     ELEFANT & $100\times100$ & 0.6034 & 8.5115 & 8.4465 & 140.8067 & 104.1228 & 121.766 & $2.0600\times10^{-5}$ & 1.85 & 7.5665 & 7.5585 & 0.11\% \\
-     \rowcolor{gray!25!}
-     Stone 28 & $32\times32$ & 0.6052 & 8.1715 & - & 141.518 & - & - & - & - & - & - & - \\
-%     FALCON & $32\times32$ & 0.6454 & 8.1422 & 7.4020 & 156.6908 & 95.7226 & 116.9302 & $2.2572\times10^{-5}$ & 0.7194 & 7.4289 & 7.4452 & 0.22\% \\
-%     \rowcolor{gray!25!}
-     FALCON & $100\times100$ & 0.6044 & 8.5266 & 8.4590 & 141.3684 & 104.8867 & 122.0470 & $2.0333\times10^{-5}$ & 1.2914 & 7.5707 & 7.5718 & 0.01\% \\
-     \midrule
-     \multicolumn{13}{|c|}{\textbf{Case 3}} \\
-     \midrule
-     Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-     & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-     \hline
-     \rowcolor{gray!25!} 
-     YACC & $100\times100$ & 0.7286 & 3.0374 & 2.9628 & 100.9467 & 2.0374 & 2.8458 & $4.7907\times10^{-5}$ & 1 & 2.0400 & 2.0335 & 0.32\% \\
-      ELEFANT & $100\times100$ & 0.7275 & 3.0347 & 2.9908 & 100.1208 & 2.0652 & 2.9019 & $4.8080\times10^{-5}$ & 0.9023 & 2.0384 & 2.037 & 0.07\% \\
-      \rowcolor{gray!25!}
-      Stone 28 & $32\times32$ & 0.7277 & 3.0347 & - & 100.018 & - & - & - & - & - & - & - \\
-%      FALCON & $32\times32$ & 0.7281 & 3.0618 & 2.8132 & 101.0092 & 2.1630 & 3.0178 & $4.8500\times10^{-5}$ & 0.7845 & 2.0699 & 2.0695 & 0.02\% \\
-%      \rowcolor{gray!25!}
-      FALCON & $100\times100$ & 0.7275 & 3.0368 & 3.0086 & 100.2776 & 2.0832 & 2.9086 & $4.8129\times10^{-5}$ & 0.9360 & 2.0387 & 2.0386 & $<$0.01\% \\
-      \midrule
-      \multicolumn{13}{|c|}{\textbf{Case 4}} \\
-      \midrule
-      Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-      & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-      \hline
-      \rowcolor{gray!25!} 
-      YACC & $100\times100$ & 0.5289 & 6.5572 & 6.5243 & 79.6202 & 75.4814 & 89.2940 & $1.9174\times10^{-4}$ & 1.6773 & 5.6512 & 5.6463 & 0.09\% \\
-       ELEFANT & $100\times100$ & 0.5277 & 6.5912 & 6.5834 & 79.1105 & 74.7596 & 88.9146 & $1.9860\times10^{-4}$ & 1.5200 & 5.6216 & 5.6182 & 0.06\% \\
-       \rowcolor{gray!25!}
-       Stone 28 & $32\times32$ & 0.5286 & 6.4036 & - & 79.4746 & - & - & - & - & - & - & - \\
-%       FALCON & $32\times32$ & 0.5334 & 6.4679 & 6.3435 & 80.6153 & 76.0511 & 90.2058 & $2.0393\times10^{-4}$ & 0.6453 & 5.6676 & 5.6698 & 0.04\% \\
-%       \rowcolor{gray!25!}
-       FALCON & $100\times100$ & 0.5279 & 6.5897 & 6.5830 & 79.1220 & 75.0715 & 88.8484 & $1.9693\times10^{-4}$ & 1.4133 & 5.6127 & 5.6113 & 0.03\% \\
-       \midrule
-       \multicolumn{13}{|c|}{\textbf{Case 5}} \\
-       \midrule
-       Code & Resolution & $P$ & $\left\langle T \right\rangle_{min}$ & $\left\langle T \right\rangle_{max}$ & Nu$^{top}_{min}$ & Nu$^{top}_{max}$ &
-       $v_{\textrm{rms}}^{min}$ & $v_{\textrm{rms}}^{max}$ & \multicolumn{2}{c}{$(\left\langle \Phi \right\rangle/Ra)_{min}$} & 
-       \multicolumn{2}{c|}{$(\left\langle \Phi \right\rangle/Ra)_{max}$} \\
-       \hline
-       \rowcolor{gray!25!}
-         YACC & $100\times100$ & 0.0792 & 0.6543 & 0.6722 & 2.6950 & 7.2792 & 41.9813 & 98.8230 & \multicolumn{2}{c}{1.4347} & 
-         \multicolumn{2}{c|}{9.2748} \\
-         ELEFANT & $100\times100$ & 0.076 & 0.6514 & 0.6693 & 2.6808 & 7.3506 & 41.2578 & 100.089 & \multicolumn{2}{c}{1.3324} & 
-         \multicolumn{2}{c|}{9.5289} \\
-         \rowcolor{gray!25!}
-         Stone 28 & $32\times32$ & 0.0764 & 0.6521 & 0.6697 & 2.6912 & 7.0763 & 41.61 & 98.183 & \multicolumn{2}{c}{-} & 
-         \multicolumn{2}{c|}{-} \\
-     %    FALCON & $32\times32$ & & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & \multicolumn{2}{c}{-} & 
-     %     \multicolumn{2}{c|}{-} \\
-     %    \rowcolor{gray!25!}
-         FALCON & $100\times100$ & 0.0775 & 0.6519 & 0.6697 & 2.6847 & 7.3629 & 41.5256 & 99.5466 & \multicolumn{2}{c}{1.3399} & 
-         \multicolumn{2}{c|}{9.3993} \\
-  \bottomrule
-  \end{tabular}}
-  \label{tab:visco_mantle}
-  \end{table}
-
-\subsection{Thin layer entrainment}\label{sec:thin}
-This experiment is performed as originally proposed by \citet{vanKeken1997} with two incompressible fluids in a rectangular domain with $L_x=2$ and $L_y=1$
-and gravity acceleration $g_y=10^{10}Ra$. The Rayleigh number and the compositional Rayleigh number are fixed to $Ra=300000$ and $Rc=450000$, respectively.
-Both fluids have constant viscosity, thermal conductivity, specific heat ($\eta=\rho=C_p=1$) and thermal expansion coefficient ($\alpha=10^{-10}$). Fluid 1
-has a density $\rho_1=1$, while fluid 2 is denser ($\rho_2=\rho_1+1.5 \cdot 10^{-10}$) and is placed at the bottom of the domain, for $y \leq 0.025$.
-Temperature are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides of the domain. The initial
-temperature field is given by
-\[T(x,y)=T_u(x,y)+T_l(x,y)+T_r(x,y)+T_s(x,y)-\frac{3}{2}\]
-with
-\begin{eqnarray}
-T_u(x,y)&=&\frac{1}{2}erf\left(\frac{1-y}{2}\sqrt{\frac{u_0}{x}}\right)\nonumber \\
-T_l(x,y)&=&1-\frac{1}{2}erf\left(\frac{y}{2}\sqrt{\frac{u_0}{L_x-x}}\right)\nonumber \\
-T_r(x,y)&=&\frac{1}{2}+\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{y+1}} \exp\left(-\frac{x^2u_0}{4y+4}\right)\nonumber \\
-T_s(x,y)&=&\frac{1}{2}-\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{2-y}} \exp\left(-\frac{(L_x-x)^2u_0}{8-4y}\right)\nonumber
-\end{eqnarray}
-and
-\begin{eqnarray}
-u_0&=&\frac{L_x^{7/3}}{(1+L_x^4)^{2/3}}\left(\frac{Ra}{2\sqrt{\pi}}\right)^{\frac{2}{3}}\nonumber \\
-Q&=&2\sqrt{\frac{L_x}{\pi u_0}}\nonumber
-\end{eqnarray}
-
-The experiment is performed with two grid resolution (\num{125x40} and \num{200x80} elements), with 100 markers per element and CFL of 0.25. $v_{\textrm{rms}}$
-(see Sec. \ref{sec:error}) calculated for both the simulations match well with results from \citet{vanKeken1997} and obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and ELEFANT \citep{Thieulot2014}, obtained with similar grid resolutions
-(Fig. \ref{fig:thin}). In addition, a variety of simulations with different resolutions and aspect ratios are performed to verify that the sensitivity of the
-initial velocity field. Results match well with results from \citet{vanKeken1997} (in the grey area) and \citet{Thieulot2014} (Fig. \ref{fig:thin_initial}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Thin_layer}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/vrms_thin.pdf}
-\caption{$v_{\textrm{rms}}$ for the thin layer experiment as function of time for different grid resolution. Results are compared with results obtained by
-\citet{vanKeken1997} (black lines) and with ASPECT and ELEFANT.}
-\label{fig:thin}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/vrmszero.pdf}
-\caption{$v_{\textrm{rms}}$ at $t=0$ for the thin layer experiment as function of the element size for different aspect ratios of the elements. Results are compared
-with results obtained by \citet{vanKeken1997} (black dots in the grey area) and with ASPECT and ELEFANT.}
-\label{fig:thin_initial}
-\end{figure}
-
-\subsection{Sinking hydrated cylinder}\label{sec:quinquis}
-This experiment is performed as explained in \citet{Quinquis2014} and simulates the sinking of a cold, hydrated cylinder into a warm, dry mantle. The domain
-is square with $L_x=L_y=$ \SI{300}{\km}, a grid resolution of \num{300x300} elements and 25 markers per element. The sinking cylinder is located at $x=$
-\SI{150}{\km} and $y=$ \SI{170}{\km} with a radius of \SI{20}{\km} and has $\rho_c=$ \SI{3250}{\kg\per\cubic\m}, $\eta_c=$ \SI{e23}{\pascal\s}, a thermal
-conductivity $k_c=$ \SI{4.5}{\watt\per\m\per\kelvin},a specific heat $C_{p_c}=$ \SI{1250}{\joule\per\kg\per\kelvin}; the surrounding mantle has $\rho_m=$
-\SI{3200}{\kg\per\cubic\m}, $\eta_m=$ \SI{e20}{\pascal\s}, a thermal conductivity $k_m=$ \SI{105}{\watt\per\m\per\kelvin} and specific heat $C_{p_m}=$
-\SI{1250}{\joule\per\kg\per\kelvin}; a 58 km-thick lithosphere overlies the mantle and has $\rho_l=$ \SI{3200}{\kg\per\cubic\m}, $\eta_l=$ \SI{e23}{\pascal\s},
-a thermal conductivity $k_l=$ \SI{4.5}{\watt\per\m\per\kelvin}, specific heat $C_{p_l}=$ \SI{750}{\joule\per\kg\per\kelvin} and a thermal diffusivity of
-\SI{e-6}{\square\m\per\s}. Velocity boundary conditions are set to free slip on all sides of the domain. Temperature increase linearly in the lithosphere,
-from 0° to 1300°C, and in the mantle, from 1300° to 1360.5°C. The initial temperature of the cylinder is fixed to 400°C. Throughout the evolution, temperature
-is fixed to 0° and 1360.5°C at the top and bottom of the domain, respectively. The initial bound water content is imposed at 2 wt.\% in the cylinder and at
-0 wt.\% in both the mantle and the lithosphere. The maximum amount of water in the mantle is fixed to 0.2 wt.\%, while maximum water content in the cylinder
-and in the lithosphere is function of pressure and temperature and is calculated using a serpentinized harzburgite with Perple\_X, as in \citet{Quinquis2014}.
-
-As shown in \citet{Quinquis2014}, the model is characterised by a progressive dehydration in the external portion of the cylinder, due to the increase of
-temperature, with a consequent vertical migration of free water that hydrates the mantle above the cylinder, up to the lithosphere. The distribution of
-bound and free water during the evolution of the experiment is shown in Fig. \ref{fig:quinquis} (left and right columns, respectively). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Hydration}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Hydration.pdf}
-\caption{Distribution of bound and free water (left and right columns, respectively) at $t=0.5$ and $t=$ \SI{1.25}{\mega\year} (first and second rows,
-respectively) for the sinking hydrated cylinder experiment. Only markers with bound or free water not equal to 0 are plotted.}
-\label{fig:quinquis}
-\end{figure}
-
-\subsection{Mantle melting curves}\label{sec:katz}
-This instantaneous experiment is performed in order to recreate isobaric melting curves, as function of temperature, bulk water and content of water in the
-melt, to compare with melting curves obtained by \citet{Katz2003}. The domain is square with $L_x=L_y=100$ km, a grid resolution of \num{100x100} elements and
-25 markers per element. Density is set in the entire domain to \SI{3300}{\kg\per\cubic\m} and temperature increases from the sides to the centre from 1000° to
-1500°C. The experiment is performed for bulk water contents from 0 to 0.3 wt.\%.
-
-Fig. \ref{fig:fraction} shows isobaric melting curves for \SI{1}{\giga\pascal} (panel a) and \SI{3}{\giga\pascal} (panel b) calculated for different bulk water
-contents. The curves well match with those obtained for the mantle by \citet{Katz2003}, showing the saturation at the solidus for bulk water of 0.2 and 0.3 wt.\%, indicating
-that the code correctly determines melt fraction as function of pressure, temperature and bulk water content. Same results can be observed in Fig.
-\ref{fig:water}a, where isobaric and isothermal melting curves are shown as function of bulk water content. In addition, Fig. \ref{fig:water}b shows that the
-code correctly determines also the percentage of water in the melt as function of pressure, temperature and melt fraction. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Melting}.
-
-\begin{figure}
-\includegraphics[width=\linewidth]{./Figures/Fraction.pdf}
-\caption{Isobaric melting curves for 1 GPa (panel a) and 3 GPa (panel b) as function of temperature with different bulk water contents.}
-\label{fig:fraction}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\linewidth]{./Figures/Water.pdf}
-\caption{Isobaric melting curves for 1.5 GPa as function of bulk water (panel a) and water in the melt (panel b) calculated for different temperatures.}
-\label{fig:water}
-\end{figure}
-
-\clearpage
-
-\bibliography{export}
-
-\end{document}

### 2022-09-07 tex/Benchmark/brick.tex
update main

--- 2022-03-31 FALCON_description.tex
@@
-\documentclass[hidelinks,10pt,a4paper]{article}
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{textcomp}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage{booktabs}
-\usepackage{multirow}
-\usepackage[cm]{fullpage}
-\usepackage[round]{natbib}
-\bibliographystyle{abbrvnat}
-\usepackage{doi}
-\usepackage{bm}
-\usepackage[table]{xcolor}
-\hypersetup{
-    colorlinks,
-    linkcolor={blue},
-    citecolor={blue},
-    urlcolor={blue}
-}
-\usepackage{float}
-\restylefloat{table}
-\usepackage{siunitx}
-\DeclareSIUnit\year{yr}
-
-\begin{document}
-
-\section{Introduction}\label{sec:intro}
-This document includes a description of the code FALCON, with a presentation of the features implemented followed by results of extensive benchmarks listed in
-Table \ref{tab:list}, performed to verify that the code solves correctly a wide range of problems. Description and results of all benchmarks are shown
-in Section \ref{sec:benchmark}.
-
-\begin{table}[H]
-\caption{List of benchmarks performed for each feature of the code. Data of all benchmarks can be found at 
-\url{https://github.com/aleregorda/Benchmarks}}
-\centering
-\begin{tabular}{| l | l | l |}
-\toprule
- \textbf{Feature}  & \textbf{Benchmark} & \textbf{Section} \\
-\midrule
-  Solver  & Stokes flow & \ref{sec:stokes} \\
-\hline
-  \multirow{2}{*}{Markers advection} & Zalesak disk & \ref{sec:runge} \\
-    & Conservative Interpolation Velocity & \ref{sec:CVI} \\
-\hline
-  \multirow{4}{*}{Momentum equation}  & Poiseuille flow & \ref{sec:poiseuille}\\
-    & Instantaneous 2D sphere & \ref{sec:ist_sphere} \\
-    & Rayleigh-Taylor experiment & \ref{sec:rayleigh} \\
-    & Falling block & \ref{sec:block} \\
-\hline
-    & 2D Stokes sphere & \ref{sec:sphere} \\
-    Sticky air and  & Stabilisation algorithm & \ref{sec:stab} \\
-    free surface & Topography relaxation & \ref{sec:crameri} \\
-  & Spontaneous subduction & \ref{sec:subduction} \\
-\hline
-  Erosion and sedimentation  & - & - \\
-\hline
-   Non-linear  & Slab detachment & \ref{sec:slab} \\
-   visco-plasticity & Indenter & \ref{sec:indenter} \\
-    & Brick & \ref{sec:brick} \\
-\hline
-  \multirow{3}{*}{Energy equation}  & Advection stabilisation & \ref{sec:advection} \\
-    & Simple shear heating & \ref{sec:simple_shear} \\
-    & Shear and adiabatic heating & \ref{sec:shear} \\
-\hline
-\multirow{3}{*}{Energy + momentum} & Mantle convection & \ref{sec:mantle} \\
-& Viscoplastic mantle convection & \ref{sec:visco_mantle} \\
-& Thin layer entrainment & \ref{sec:thin} \\
-\hline
-  Phase changes  & \multirow{2}{*}{Hydrated sinking cylinder} & \multirow{2}{*}{\ref{sec:quinquis}} \\
-  and hydration  &  & \\
-\hline
-  Melting & Experimental melting curves & \ref{sec:katz} \\
-\bottomrule
-\end{tabular}
-\label{tab:list}
-\end{table}
-
-The code is written in Fortran90 and uses the finite element method (FEM) with quadrilateral $Q_1 \times P_0$ elements (continuous bilinear velocity and
-discontinuous constant pressure), associated with MUMPS\footnote{\url{http://mumps.enseeiht.fr/}} \citep{Amestoy2001,Amestoy2019} that is a software package
-for solving systems of linear equations of the form $A \cdot x = b$, where \textit{A} is a square sparse matrix, by means of a direct method. Since
-$Q_1 \times P_0$ elements do not satisfy Ladyzhenskaya, Babuska and Brezzi (LBB) stability condition \citep{Donea2003}, elemental pressure is elaborated
-in post-processing to avoid spurious pressures by means of a double interpolation to smooth the pressure. Following this procedure, the elemental pressure
-is interpolated onto nodes and then back onto elements. After the smoothing procedure, the pressure field is calculated again on the nodes to be used in
-combination with the lithostatic pressure, which is calculated onto nodes. Correctness of the solution and performances of the solver in terms of time and
-memory usage are tested solving the Stokes flow with the analytical solution proposed by \citet{Donea2003} (Section \ref{sec:stokes}).
-
-The thermo-mechanics of crust-mantle systems is described by means of the conservation of mass, momentum and energy equations, expressed as follows:
-\begin{eqnarray}
-\label{eq:mass}\frac{\partial \rho}{\partial t}+\bm{\nabla} \cdot (\rho \bm{u})&=&0\\
-\label{eq:momentum}-\bm{\nabla} p + \bm{\nabla} \cdot \bm{\tau} + \rho \bm{g}&=&0\\
-\label{eq:energy}\rho C_p \left(\frac{\partial T}{\partial t} + \bm{u} \cdot \bm{\nabla} T \right)&=&\bm{\nabla} \cdot \left(k\bm{\nabla} T\right) + H_{tot}
-\end{eqnarray}
-where $\rho$ is the density, $\bm{u}$ is the velocity, \textit{p} is the pressure, $\bm{\tau}$ is the deviatoric stress, $\bm{g}$ is the gravity acceleration,
-$C_p$ is the specific heat at constant pressure, \textit{T} is the temperature, \textit{k} is the thermal conductivity and $H_{tot}$ is the total internal heat
-production. Density variations due to temperature are generally small enough to assume the density as constant ($\rho=\rho_0$) in Eq. \ref{eq:mass} and in
-Eq. \ref{eq:energy}, while it must be treated as a variable in the buoyancy term of Eq. \ref{eq:momentum}, such that \[\rho=\rho_0(1-\alpha(T-T_0))\]
-where $\rho_0$ is the density at a reference temperature $T_0$ and $\alpha$ is the coefficient of thermal expansion. Eq. \ref{eq:mass} then can be rewritten as
-\begin{equation}\label{eq:mass_0}
-\bm{\nabla} \cdot \bm{u}=0
-\end{equation}
-This simplification is known as the extended Boussinesq approximation.
-
-The time step $dt$ is chosen in according to Courant-Friedrichs-Lewy condition \citep{Anderson1995}, such that
-\[dt=C_n\; \min\left(\frac{\Delta x}{\max|\bm{u}|},\frac{(\Delta x)^2}{\kappa}\right)\]
-where $0<C_n<1$ is the Courant (CFL) number, $\Delta x$ is the minimum dimension of smallest element, $\max|\bm{u}|$ is the maximum velocity calculated on the
-entire domain and
-$\kappa=\frac{k}{\rho C_p}$ is the thermal diffusivity \citep{Thieulot2014}.
-
-\section{Lagrangian markers}\label{sec:markers}
-Elemental properties (density, viscosity, thermal conductivity, specific heat and thermal expansion) needed to solve Eqs. \ref{eq:mass_0}, \ref{eq:momentum}
-and \ref{eq:energy} are related to the composition of each element, determined by means of Lagrangian markers that are characteristic of different materials
-in the domain. Elemental properties (with the sole exception of the viscosity) are calculated using an arithmetic mean, as \[P_e=\frac{1}{n}\sum_{i=1}^n P_i\]
-where $P_e$ is the elemental property, $P_i$ is the property characteristic of the material of each marker and $n$ is the number of markers of the element.
-Differently, the average scheme for the elemental viscosity can be chosen between harmonic, geometric and arithmetic mean. Markers can be placed regularly or
-randomly at the beginning of the simulation and their advection is performed by either a 2nd-order or a 4th-order Runge-Kutta in space, interpolating the velocity 
-field on each marker by means of the shape functions (see Section \ref{sec:runge}). The interpolated velocity is then corrected by means of the Conservative
-Velocity Interpolation (CVI), which introduces a corrective term to reduce dispersion and clustering of particles in both steady state and time-dependent 
-\citep{Wang2015} (see Section \ref{sec:CVI}).
-
-The initial distribution of the markers is created using the open-source code library
-Geodynamic World Builder \citep{Fraters2019}. During the simulation, each marker carries memory of temperature, pressure and accumulated strain, which are
-determined interpolating the relative nodal parameter. In particular, pressure is interpolated onto each marker after the smoothing procedure. The number
-of markers contained in each element is maintained between $n_{min}$ and $n_{max}$. When in an element there are less markers than $n_{min}$ the code adds
-random markers to reach the $n_{min}$, while if the number is higher than $n_{max}$ some of them are randomly deleted. When new markers are added, they assume
-the properties of the nearest marker. In this way elements are never empty and maintain a number of markers inside a prefixed range.
-
-\section{The momentum equation}\label{sec:mom_eq}
-The penalty method is implemented so that Eq. \ref{eq:mass_0} can be written as:
-\begin{equation}\label{eq:mass_penalty}
-\bm{\nabla} \cdot \bm{u} + \frac{p}{\lambda}=0
-\end{equation}
-where $\lambda$ is the so-called penalty parameter, which should be 6-7 orders of magnitude larger than the shear viscosity to ensure that mass conservation
-is satisfied \citep{Donea2003,Thieulot2014}.
-
-The deviatoric stress tensor in Eq. \ref{eq:momentum} can be written in terms of the strain rate tensor as $\bm{\tau}=2\eta\dot{\bm{\epsilon}}$, with
-$\dot{\bm{\epsilon}}=\frac{1}{2}\left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)$. Therefore, Eq. \ref{eq:momentum} can be rewritten as:
-\begin{equation}\label{eq:momentum_pressure}
--\bm{\nabla} p + \bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right) + \rho \bm{g} = 0
-\end{equation}
-Finally, using pressure from Eq. \ref{eq:mass_penalty}, Eq. \ref{eq:momentum_pressure} can be rewritten as:
-\begin{equation}\label{eq:momentum_penalty}
-\lambda \bm{\nabla} \left(\bm{\nabla} \cdot \bm{v}\right)+\bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right)+\rho\bm{g}=0
-\end{equation}
-The penalty method is been associated to the iterative Uzawa method, as described in detail in \citet{Dabrowski2008} and \citet{Thieulot2014}
-(see Section \ref{sec:poiseuille}). Results of benchmarks performed to verify the correctness in the implementation of Eq. \ref{eq:momentum_penalty} are
-shown in Sections \ref{sec:ist_sphere} and \ref{sec:rayleigh}.
-
-To support large viscosities variations the penalty parameter is related to the effective elemental viscosity by means of a dimensionless coefficient, so that
-$\lambda=\lambda_e(e)\eta_{eff}(e)$ \citep{Marotta2006,Dabrowski2008,Thieulot2014}. Benchmark of the falling block \citep{Gerya2003a,Gerya2010b,Thieulot2011}
-is performed to verify that the code can correctly deal buoyancy driven flows with strong viscosity contrasts (Section \ref{sec:block}).
-
-\section{Sticky air and free surface}\label{sec:sticky}
-The Earth's surface can be treated by means of either the so-called sticky-air or a true free surface method, both of them implemented in the code. In the
-sticky-air method the surface is approximated with the introduction of a buoyant layer with a viscosity at least four orders of magnitude lower than the crust
-\citep{Schmeling2008,Crameri2012} and the interface between lithosphere and air is defined using a chain of passive markers that are advected as the Lagrangian
-markers. The correctness of the evolution of the markers chain is tested with the experiment of a 2D time-dependent Stokes sphere below a free surface and
-compared with results from ASPECT\footnote{\url{https://aspect.geodynamics.org/}} \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} (Section
-\ref{sec:sphere}). In the true free surface case the top boundary is assumed stress-free and velocities are not fixed. In this case topography variations are
-described by vertical deformations of the mesh that depend on the velocity field of the nodes that identify the free surface, while horizontal deformations are
-not taken into account. This procedure is known as the Arbitrary Lagrangian-Eulerian (ALE) method and its implementation follows the technique described in
-\citet{Thieulot2011}. However, although the implementation of a true free surface better reproduces laboratory experiment, extremely small time steps can be
-necessary to maintain stability \citep{Kaus2010a,Quinquis2011,Thieulot2014}. Therefore, the stability algorithm proposed by \citet{Kaus2010a} is implemented
-to avoid instabilities due to high density differences at the free surface when using too large time steps. The implementation of this algorithm are tested by
-performing the experiment described by \citet{Kaus2010a} (Section \ref{sec:stab}).
-
-The topography relaxation benchmark proposed by \citet{Crameri2012} is performed to verify that the code correctly recovers topography variations in case of
-both the sticky-air and the true free surface method (Section \ref{sec:crameri}). Finally, Section \ref{sec:subduction} show the results of the spontaneous
-subduction experiment described by \citet{Schmeling2008}.
-
-\section{Erosion and sedimentation}\label{sec:erosion}
-Surface processes at either the lithosphere-air interface or in correspondence of the free surface have been implemented by means of the software
-Fastscape\footnote{\url{https://fastscape.org/}} \citep{Braun2013,Cordonnier2019,Yuan2019a,Yuan2019b}. FastScapeLib is a set of routines that solve the stream
-power law (enriched by a sediment transport/deposition term) \citep{Yuan2019a}, hillslope diffusion and marine transport and deposition \citep{Yuan2019b}.
-The partial differential equation solved by FastScapeLib is:
-\begin{equation}
-\frac{\partial h}{ \partial t}=U-K_f A^{m_e}S^{n_e} + \frac{G}{A}\int_A \left( U-\frac{\partial h}{\partial t} \right) dA + K_d \bm{\nabla}^2 h 
-\end{equation}
-where $h$ is the topography, $U$ is the uplift, $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $S$ is the slope, $m_e$ and $n_e$ are
-parameters that should be chosen to constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}, $G$ is the dimensionless
-deposition/transport coefficient for the enriched stream power law and $K_d$ is the effective diffusivity. Terms $K_f A^{m_e}S^{n_e}$ and $K_d \bm{\nabla}^2 h$
-consider long-range processes, described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, and short-range processes,
-proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}.
-
-In case of sticky air, slope is determined by means of the markers chain, which is vertically corrected considering erosion and sedimentation rates. After the
-correction, continental markers above the markers chain are transformed in air markers and air markers below the markers chain become sediments. Similarly,
-slope is determined considering the nodes on the top boundary in case of true free surface and they are vertically corrected as for the markers chain. In this
-case, continental markers above the top boundary are simply deleted.
-%Surface processes at the lithosphere-air interface are implemented taking into account fluvial erosion (long-range processes), hillslope processes (short-range
-%processes) and failure of slopes exceeding a threshold angle $\theta$ \citep{Burov1997,Tucker1999,Sacek2012,Croissant2014,Ueda2015,Beucher2020}. Erosion rates
-%$\dot{E}_{lr}$ due to long-range processes can be described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{lr}=K_f A^{m_e} \bm{\nabla} h^{n_e}
-%\end{equation}
-%where $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $\bm{\nabla} h$ is the slope and $m_e$ and $n_e$ are parameters chosen in to
-%constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}. Erosion rates $\dot{E}_{sr}$ due to short-range processes are
-%implemented by assuming that transport is linearly proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{sr}=K_d \bm{\nabla}^2 h
-%\end{equation}
-%where $K_d$ is the effective diffusivity. In addition, the slope is modified if it still exceeds a threshold angle after long- and short-range processes
-%\citep{Ueda2015}.
-
-%Offshore sedimentation processes are implemented similarly to hillslope processes but with a depth-dependent coefficient \citep{Sacek2012,Beucher2020}.
-%The sedimentation rate $\dot{S}$ is calculated as
-%\begin{equation}
-%\dot{S}=K_s \bm{\nabla}^2 h
-%\end{equation}
-%with
-%\begin{equation}
-%K_s=K^*_s \exp\left(\frac{h-h_{sl}}{h_0}\right)
-%\end{equation}
-%where $K^*_s$ is a constant, $h_{sl}$ is the sea level and $h_0$ is a characteristic length. In case of sticky air, slope is determined by means of the markers
-%chain, which is vertically corrected considering erosion and sedimentation rates. After the correction, continental markers above the markers chain are
-%transformed in air markers and air markers below the markers chain become sediments. Similarly, slope is determined considering the nodes on the top boundary
-%in case of true free surface and they are vertically corrected as for the markers chain. In this case, continental markers above the top boundary are simply
-%deleted.
-
-\section{Non-linear rheologies}\label{sec:nonlinear}
-Non-linear rheologies are implemented combining viscous creep (dislocation and diffusion) and plastic yielding. For each marker, diffusion and dislocation
-viscosity ($\eta_{df|ds}$) can be determined as follows \citep{Karato1993,Warren2008b,Wang2016}:
-\begin{equation}\label{eq:viscous}
-\eta_{df|ds}=f_s \left(\frac{d^m}{A}\right)^{\frac{1}{n}} \left (\frac{2^{\frac{1-n}{n}}}{3^{\frac{n+1}{2n}}} \right)
-I_2^{\frac{(1-n)}{n}}\exp\left(\frac{Q+pV}{nRT}\right)
-\end{equation}
-where $f_s$ is a scaling factor used to represent lithologies that are stronger or weaker than the base set, $d$ is the grain size, $m$ is the grain size exponent,
-$n$ is the stress exponent, $A$ is the uniaxial pre-exponential factor, $I_2$ is the square root of the second invariant of the strain rate tensor, $Q$ is the
-activation energy, $p$ is the pressure, $V$ is the activation volume, $R$ is the gas constant and $T$ is the temperature.
-Pressures and temperatures are determined by the interpolation of the nodal parameters, while strain rates are calculated by means of the
-derivative of the velocities on the nodes. In case of diffusion creep $n=1$ and $m>0$, in case of dislocation creep $n>1$ and $m=0$. The viscous creep $\eta_{cp}$
-is then calculated as the harmonic average between $\eta_{df}$ and $\eta_{ds}$:
-\begin{equation}\label{eq:visco_creep}
-\eta_{cp}=\left(\frac{1}{\eta_{df}}+\frac{1}{\eta_{ds}}\right)^{-1}
-\end{equation}
-The implementation of the non-linear viscous creep viscosity depending on the strain rate is tested by means of the slab detachment benchmark
-(Section \ref{sec:slab}).
-
-Plastic yielding is implemented rescaling $\eta_{cp}$ in order to limit the stress below the yield stress $\sigma_y$ \citep{Thieulot2008,Thieulot2014,Glerum2018},
-obtaining
-\begin{equation}\label{eq:plastic}
-\eta_{pl}=\frac{\sigma_y}{2I_2}
-\end{equation}
-where $\eta_{pl}$ is the plastic viscosity and the yield stress is determined following the Drucker-Prager criterion, such as
-\begin{equation}\label{eq:yield}
-\sigma_y=C\;\cos(\phi)+p\;\sin(\phi)
-\end{equation}
-where $C$ is the cohesion and $\phi$ is the internal friction angle. In case of negative pressure it is imposed equal to 0, so that negative yield stress are
-excluded. The correctness of the non-linear solution in case of plastic viscosity with variable internal friction angle is verified performing indenter and
-brick experiments (Sections \ref{sec:indenter} and \ref{sec:brick}, respectively).
-
-Strain softening is taken into account for both viscous creep and plastic viscosity \citep{Huismans2003,Babeyko2005,Huismans2005,Sobolev2005,Warren2008a} by
-means of the accumulated strain $\epsilon$ memorised by each marker. Softening in the viscous creep determines a linear decrease of $\eta_{vc}$ by means of a
-viscous strain softening factor $W_S$ that increases linearly from $W_{S_0}$ to $W_{S_{\infty}}$ for $\epsilon_{S_0}<\epsilon_S<\epsilon_{S_{\infty}}$ 
-\citep{Warren2008a}. This viscous softening can be related to strain-induced grain size reduction \citep{Warren2008a}. Differently, plastic softening is 
-simulated with a linear decrease of internal friction angle $\phi(\epsilon)$ and cohesion $C(\epsilon)$ in according to
-\begin{eqnarray}
-\label{eq:friction1}\phi(\epsilon)&=&\phi_0+(\phi_{\infty}-\phi_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}\\
-\label{eq:friction2}C(\epsilon)&=&C_0+(C_{\infty}-C_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}
-\end{eqnarray}
-where $\phi_0$, $C_0$ and $\phi_{\infty}$, $C_{\infty}$ are internal friction angle and cohesion for $\epsilon_0$ and $\epsilon_{\infty}$, respectively
-\citep{Huismans2003,Huismans2005,Warren2008a,Thieulot2014}. Plastic softening approximates deformation-induced softening of faults and brittle shear zones
-\citep{Warren2008a}.
-
-Plastic yielding and viscous creep are then combined to obtain a viscoplastic viscosity $\eta_{vp}$ as follows, assuming that they are independent processes
-\citep{Karato2008,Glerum2018}:
-\begin{equation}\label{eq:viscoplastic}
-\eta_{vp}=\min(\eta_{cp},\eta_{pl})
-\end{equation}
-Finally, effective viscosity $\eta_{eff}$ is capped by the minimum and the maximum viscosity ($\eta_{min}$ and $\eta_{max}$, respectively) to avoid extremely
-low or high viscosity \citep{Glerum2018} as follows
-\begin{equation}\label{eq:effective}
-\eta_{eff}=\min(\max(\eta_{vp},\eta_{min}),\eta_{max})
-\end{equation}
-Viscous creep and plastic yielding are non-linear rheologies because of their dependence on the velocity field through pressure and strain rates. Therefore,
-the solution is determined by means of Picard-type iterations, until convergence of the velocity field \citep{Glerum2018}. The convergence is verified at each
-iteration $i$ via the nonlinear residual $\bm{\mathcal{R}}^i$ that can be determined as
-\begin{equation}\label{eq:convergence}
-\bm{\mathcal{R}}^i= \mathbb{K} (\eta_{eff}(\dot{\bm{\epsilon}}^{i-1},p^{i-1}))\cdot \bm{v}^{i-1} - \bm{f}^i
-\end{equation}
-%\begin{equation}\label{eq:convergence}
-%\chi=1-\frac{\langle (\bm{v}^i-\langle\bm{v}^i\rangle)\cdot (\bm{v}^{i+1}-\langle\bm{v}^{i+1}\rangle) \rangle}{\sqrt{\sigma^i\sigma^{i+1}}}
-%\end{equation}
-%Iterations are performed until either convergence ($\chi<tol$) or maximal number of non-linear iterations ($it_{max}$) is reached.
-where $\mathbb{K}$ is the Stiffness matrix and $\bm{f}$ is the right hand side vector \citep{Spiegelman2016,Glerum2018}.  The $L_2$-norm is extracted from
-$\bm{\mathcal{R}}^i$ and it is normalised as follows
-\begin{equation}\label{eq:convergence_res}
-\frac{||\bm{\mathcal{R}}^i||_2}{||\bm{\mathcal{R}}^0||_2}
-\end{equation}
-where $||\bm{\mathcal{R}}^0||_2$ is the $L_2$-norm of the first nonlinear iteration.
-Since often the normalised nonlinear residual drop very quickly over the first 2-3 iterations, also the $L_2$-norm of velocity and pressure residuals are
-calculated as
-\begin{equation}\label{eq:convergence_vel}
-\frac{||\Delta\bm{v}||_2}{||\bm{v}||_2}=\frac{||\bm{v}_i-\bm{v}_{i-1}||_2}{||\bm{v}_{i}||_2}
-\end{equation}
-and
-\begin{equation}\label{eq:convergence_press}
-  \frac{||\Delta p||_2}{||p||_2}=\frac{||p_i-p_{i-1}||_2}{||p_{i}||_2}
-\end{equation}
-respectively. Iterations are performed until either maximal number of non-linear iterations ($it_{max}$) is reached or all the normalised nonlinear residual, 
-the velocity residual and the pressure residual converge under a defined tolerance $tol$.
-
-\section{The energy equation}\label{sec:en_eq}
-The stabilisation of the advection term of the energy equation (Eq. \ref{eq:energy}) needed to avoid possible oscillations in the thermal solution in the case
-when advection dominates over diffusion, is implemented by means of a streamline-upwind Petrov–Galerkin (SUPG) method, for which the advection term is modified
-follows the discussion in \citet{Thieulot2011} and \citet{Thieulot2014} (see Section \ref{sec:advection} for benchmark). The total internal heat production
-$H_{tot}$ in Eq. \ref{eq:energy} includes radiogenic heating $H_r$, shear heating $H_s$ and adiabatic heating $H_a$, where
-\begin{equation}\label{eq:shear}
-H_s=2\eta \dot{\bm{\epsilon}} : \dot{\bm{\epsilon}} = 2\eta(\dot{\epsilon}_{xx}^2+\dot{\epsilon}_{yy}^2+2\dot{\epsilon}_{xy}^2)
-\end{equation}
-and
-\begin{equation}\label{eq:adiabatic}
-H_a=T \alpha \frac{Dp}{Dt} \approx -\alpha T \rho g_y v
-\end{equation}
-The correctness of the implementation and the computation of shear and adiabatic heating is verified performing an exercise with an analytical solution
-(Section \ref{sec:simple_shear}) and Exercise 9.4 in \citet{Gerya2010b} (Section \ref{sec:shear}). Furthermore, mantle convection and thin layer entrainment
-benchmarks (Sections \ref{sec:mantle} and \ref{sec:thin}, respectively) are performed to verify that the code solve correctly Eqs. \ref{eq:momentum_penalty}
-and \ref{eq:energy} in case of a temperature field characterised by an initial perturbation and with a temperature-dependent density. At last, the viscoplastic
-mantle convection benchmark is performed to check also the viscoplasticity in case of a initial perturbed temperature field (Section \ref{sec:visco_mantle}).
-
-\section{Phase transitions and hydration}\label{sec:phase}
-Variations of effective density, specific heat and coefficient of thermal expansion during the evolution of crustal and mantle materials at different
-pressure-temperature ($p$-$T$) conditions are computed by means of version 6.8.6 of Perple\_X software package \citep{Connolly2005}, similarly to the
-implementation described in \citet{Marotta2020}. By default, mineral assemblages and properties of each lithology (oceanic crust, lower and upper continental
-crust, sediments and dry or hydrated mantle) are calculated for temperatures from \SIrange{330}{1600}{\kelvin}, with increments of approximately \SI{4}{\kelvin},
-and pressures from \SIrange{0.1}{30}{\giga\pascal}, with increments of approximately \SI{0.03}{\giga\pascal}, for a total of almost 100000 points.
-
-The introduction of phase transitions produces two main effects that must be taken into account in numerical models: variations in density and release/absorption
-of latent heat \citep{Tackley2010}. Effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into
-account considering effective coefficient of thermal expansion \citep{Christensen1985,Zhong2015}. Similarly, effects of latent heat in the energy equation
-(eq \ref{eq:energy}) are taken into account by considering effective specific heat and coefficient of thermal expansion
-\citep{Christensen1985,Tackley2010,Zhong2015}. Effective density, specific heat and coefficient of thermal expansion are included in Perple\_X files and are
-assumed by each Lagrangian marker in according to its $p$-$T$ conditions.
-
-Hydration processes related to dehydration of subducting lithosphere \citep{Schmidt1998,Liu2007,Faccenda2009,Faccenda2010,Faccenda2014,Rosas2016} strongly
-influence the thermo-mechanics inside the mantle wedge, mainly because of both weakening effects on the mantle rheology and density variations in case of
-mantle serpentinization \citep{Gerya2002,Honda2003,Arcay2005,Roda2010,Regorda2017}. In the case that hydration is switched on, the amount of bound and free
-water is memorised by each marker, following the implementation of \citet{Quinquis2014}; hydration and dehydration processes are related to the amount of bound
-water of each marker and to the maximum amount of water it can transport, i.e. if the amount of bound water exceeds the maximum amount of water, the marker
-dehydrates and releases free water that can hydrates under-saturated markers. Maximum water content of each marker is determined as function of lithology and
-$p$-$T$ conditions and it is calculated using Perple\_X, in the same way than effective density, specific heat and coefficient of thermal expansion. Bound water
-is advected together with the markers, neglecting the effect of bound water diffusion, while free water simply migrates vertically and is not coupled to the
-solid-phase flow of the mantle wedge \citep{Arcay2005,Quinquis2014}. The correctness of the migration scheme of the free water and its absorption into bound
-water by under-saturated markers are verified performing the experiment described in \citet{Quinquis2014} (Sec. \ref{sec:quinquis}).
-
-The mantle viscosity weakening related to the amount of water has been extensively studied \citep{Chopra1981,Hirth2003} and it is implemented in according to
-\citet{Arcay2005} and \citet{Horiuchi2016}, as follows
-\begin{equation}\label{eq:wet_visc}
-\eta_{wet}=\eta_{dry}\left(\left[1-\frac{1}{f_v}\right]\exp\left(-\frac{[OH^-]}{[OH^-]_0}\right)+\frac{1}{f_v}\right)
-\end{equation}
-where $f_v$ is the viscosity reduction factor between dry and wet conditions, $[OH^-]$ is the water content and $[OH^-]_0$ is a reference water content set to
-620 ppm (0.062 wt.\%) \citep{Arcay2005}. By default, $f_v$ is set to 100, in according to \citet{Arcay2005} and \citet{Horiuchi2016}.
-
-\section{Mantle melting}\label{sec:melt}
-Melt fraction in the mantle depends on temperature ($T$, in \textdegree C), pressure ($p$, in \si{\giga\pascal}) and the water content in the melt ($X_{H_2O}$, in wt.\%).
-The determination of the melt fraction $M$ is implemented as explained by \citet{Katz2003}, according to the successive modification by \citet{Langmuir2006} and
-\citet{Kelley2010}, as follows
-\begin{eqnarray}\label{eq:melt_fr}
-M(p,T,X_{H_2O})&=&-T+T_{sol}+(x_t\cdot ln(p+y_t))\cdot X_{H_2O}-\nonumber\\
-&&+K_m\left(\frac{X_{H_2O}^{bulk}}{D_{H_2O}(1-X_{H_2O})+X_{H_2O}}\right)^{\gamma}
-\end{eqnarray}
-where $T_{sol}$ is the temperature of the dry solidus, $D_{H_2O}$ is the partition coefficient and $x_t\cdot log(p+y_t)$ indicates the pressure dependence of
-the melting curve \citep{Kelley2010}. $x_t$, $y_t$, $K_m$ and $\gamma$ are parameters chosen in according to \citet{Kelley2010}.
-However, $X_{H_2O}$ depends on the melt fraction as
-\begin{equation}\label{eq:melt_wt}
-X_{H_2O}(M)=\frac{X_{H_2O}^{bulk}}{D_{H_2O}+M(1-D_{H_2O})}
-\end{equation}
-and a numerical solution can be found using a root-finder method \citep{Katz2003,Wang2016}. Furthermore, the content of water in the melt is limited by the
-pressure-dependent saturation concentration of water in the melt, determined as
-\begin{equation}\label{eq:melt_max}
-X_{H_2O}^{sat}=\chi_1 p^{\lambda_m}+\chi_2 p
-\end{equation}
-where $\chi_1$, $\chi_2$ and $\lambda_m$ are parameters chosen according to \citet{Katz2003}, \citet{Langmuir2006} and \citet{Kelley2010}. The correctness in
-the determination of melt fractions and water contents in the melt are tested comparing model results with experimental curves obtained by \citet{Katz2003}
-(Sec. \ref{sec:katz}).
-
-In case of melting, viscous creep viscosity ($\eta_{vp}$) is modified according to \citet{Wang2016}, as
-\begin{equation}\label{eq:melt_visc}
-\eta_{melt}=\eta_{vp}\cdot \exp(\alpha_m M)
-\end{equation}
-where $\alpha_m$ are melt fraction factors for dislocation and diffusion creep, chosen in according to \citet{Wang2016}.
-
-Effective density ($\rho_{eff}$) for partially molten rocks is calculated as
-\begin{equation}\label{eq:melt_dens}
-\rho_{eff}=\rho_s(1-M)+\rho_m M
-\end{equation}
-where $\rho_s$ and $\rho_m$ are the densities of the solid and the molten rock, respectively \citep{Gerya2004b,Gerya2007,Wang2016}. The density of the
-solid rock is extracted by Perple\_X, while the density of the molten portion is calculated as
-\begin{equation}\label{eq:melt_dens_M}
-\rho_m(p,T)=\rho_0[1-\alpha(T-T_0)][1+\beta(p-p_0)]
-\end{equation}
-where $\rho_0$ is the density at the reference temperature ($T_0$) and pressure($p_0$), and $\alpha$ and $\beta$ are the thermal expansion and compressibility
-coefficients, respectively \citep{Gerya2004b,Gerya2007,Wang2016}.
-
-As for the phase transitions, effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into account
-considering effective coefficient of thermal expansion. Similarly, latent heat due to melting/crystallisation can be implicitly included in the energy equation
-(Eq. \ref{eq:energy}) by considering effective specific heat ($C_{p_{eff}}$) and thermal expansion ($\alpha_{eff}$) of the partially molten rocks, in according to
-\begin{eqnarray}
-\label{eq:cp_effective}C_{p_{eff}}&=&C_p+H_L\left(\frac{\partial M}{\partial T}\right)_p \\
-\label{eq:alpha_effective}\alpha_{eff}&=&\alpha+\rho\frac{H_L}{T}\left(\frac{\partial M}{\partial p}\right)_T
-\end{eqnarray}
-where $C_p$ and $\alpha$ are the specific heat and the coefficient of thermal expansion of the solid rock, respectively, and $H_L$ is the latent heat of the
-molten rock \citep{Gerya2004b,Gerya2007,Tackley2010}.
-
-\section{Benchmarks}\label{sec:benchmark}
-\subsection{Error measurements}\label{sec:error}
-In order to determine the accuracy of velocity and pressure field of the benchmarks, the $L_2$-norm is computed by numerical integration on the Gauss-Legendre
-quadrature points. $L_2$-norm for pressure and velocity errors can be evaluated as
-\begin{eqnarray}
-\textrm{err}_p&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|p^n_i(r_q)-p^a_i(r_q)|^2\right)w_q|J_q|}\nonumber \\
-\textrm{err}_v&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|u^n_i(r_q)-u^a_i(r_q)|^2+|v^n_i(r_q)-v^a_i(r_q)|^2\right)w_q|J_q|}\nonumber
-\end{eqnarray}
-respectively, where $n_e$ is the number of elements, $n_q$ is the number of quadrature points per element, $p^n_i(r_q)$ and $p^a_i(r_q)$ are the numerical and
-analytical pressure, respectively, in each quadrature point $q$, $w_q$  and $J_q$ are the weight and the Jacobian at the quadrature point
-$q$, $u^n_i(r_q)$, $v^n_i(r_q)$, $u^a_i(r_q)$ and $v^a_i(r_q)$ are the numerical and analytical velocities, respectively, in each quadrature point $q$.
-
-Other quantities used as comparison with original benchmarks are the root-mean-square velocity over the whole domain and over the surface
-\begin{eqnarray}
-v_{\textrm{rms}}&=&\sqrt{\int_{\Omega}|\bm{v}|^2 d\Omega}\nonumber \\
-v_{\textrm{rms}}^{top}&=&\sqrt{\int_0^1 u^2 \bigg|_{y=1} dx}\nonumber
-\end{eqnarray}
-respectively, top and bottom Nusselt numbers
-\[\textrm{Nu}^{top/bottom}=-\int_0^1\frac{\partial T}{\partial y} \bigg|_{y=1/y=0} dx\]
-the average rate of work done against gravity
-\[\left\langle W \right\rangle=\int_{\Omega}T u_y d\Omega\]
-the average rate of viscous dissipation
-\[\left\langle \Phi \right\rangle=\int_{\Omega}\tau_{ij}\dot{\epsilon}_{ij} d\Omega\]
-and the percentage error between $\left\langle W \right\rangle$ and $\left\langle \Phi \right\rangle/Ra$
-\[\delta=\frac{\left|\left\langle W \right\rangle -\frac{\left\langle \Phi \right\rangle}{Ra} \right|}{\max \left(\left\langle W \right\rangle,
-\frac{\left\langle \Phi \right\rangle}{Ra}\right)}\times100\]
-
-\subsection{Stokes flow}\label{sec:stokes}
-The problem consists of determining velocity field $(u,v)$ and pressure $p$ in case of a manufactured solution with prescribed body forces such as
-\begin{eqnarray}
-b_1&=&(12-24y)x^4+(-24+48y)x^3+(-48y+72y^2-48y^3+12)x^2+\nonumber \\
-&&+(-2+24y-72y^2+48y^3)x+1-4y+12y^2-8y^3\nonumber \\
-b_2&=&(8-48y+48y^2)x^3+(-12+72y-72y^2)x^2+(4-24y+48y^2-48y^3+\nonumber \\
-&&+24y^4)x-12y^2+24y^3-12y^4\nonumber
-\end{eqnarray}
-for which the exact solution is:
-\begin{eqnarray}
-u(x,y)&=&x^2(1-x)^2(2y-6y^2+4y^3)\nonumber \\
-v(x,y)&=&-y^2(1-y)^2(2x-6x^2+4x^3)\nonumber \\
-p(x,y)&=&x(1-x)-\frac{1}{6}\nonumber
-\end{eqnarray}
-The domain is a unit square a constant viscosity ($\eta=1$) and the penalty parameter is set to $\lambda=10^7$. Velocity boundary conditions are set to no slip
-$(\bm{v}=\bm{0})$ on all boundaries. The problem is performed for different grid resolution between \num{8x8} and \num{1024x1024} elements.
-
-The errors between the analytical solution and the numerical prediction of the velocity and the pressure field are calculated as in Section \ref{sec:error}.
-Fig. \ref{fig:errors}a shows that both velocity and pressure field converge to the exact solution with the decrease of the element size, following the
-theoretical convergence rate. The convergence of pressure, in contrast with the observation by \citet{Donea2003} for $Q_1 \times P_0$ elements, indicate the
-effectiveness of the smoothing procedure. Smoothed pressure field for grid resolution of \num{128x128} elements is shown in Fig. \ref{fig:errors}b (continuous
-red line), in comparison with the analytical solution (dashed blue line). Solve times and memory usage needed to generate the solution are shown in
-Fig. \ref{fig:MUMPS}. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Solver/Stokes_Flow}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/errors.pdf}
-\caption{Velocity and pressure error for the Stokes flow experiment between generated and analytical solution as a function of element size (panel a),
-and comparison between smoothed pressure and analytical solution as function of $x$ coordinate for a grid resolution of \num{128x128} elements (panel b).}
-\label{fig:errors}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/MUMPS.pdf}
-\caption{Panel a) Analysis, factorisation, solution and total solve times (panel a) and factorisation memory usage (panel b) as a function of the total number
-of degrees of freedom for the Stokes flow experiment.}
-\label{fig:MUMPS}
-\end{figure}
-
-\subsection{Markers advection}\label{sec:runge}
-The 2nd-order Runge-Kutta advection scheme in space is tested by means of the Zalesak disk test \citep{Zalesak1979,Thieulot2014}. The benchmark is performed
-in a unit square domain with a grid resolution of \num{32x32} elements and 1000000 markers and values of Courant number between 0.25 and 1.
-At $t=0$, the disk is centred at position $(0.5;0.75)$ with a radius $R=0.15$ and has a vertical fissure 0.05 wide and 0.2 high (orange in Fig. \ref{fig:runge})
-and with the velocity field prescribed in the entire domain as
-\begin{eqnarray}
-u(x,y)&=&2\pi\left(y-\frac{L_x}{2}\right)\nonumber \\
-v(x,y)&=&-2\pi\left(x-\frac{L_x}{2}\right)\nonumber
-\end{eqnarray}
-the markers are back to their initial location after a $2\pi$ rotation ($t=1$). The clockwise rotation of the disk at $t=0.25$, 0.5 and 0.75 
-is shown in Fig. \ref{fig:runge} (red, blue and green, respectively). The distance from the centre of 1 marker is calculated throughout the rotation to
-evaluate the error for different values of Courant number. As expected, the error increase with the increasing of the time step, as shown in 
-Fig. \ref{fig:runge_err}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/Zalesak_Disk}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=200px]{./Figures/Zalesak.pdf}
-\caption{Position of the Zalesak disk at $t=0$, 0.25, 0.5 and 0.75 (orange, red, blue and green, respectively) for the case with Courant number of 1.}
-\label{fig:runge}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/zal_marker.pdf}
-\caption{Distance from the centre as function of time for values of Courant number of  of 0.25, 0.3, 0.5, 0.75 and 1 (green, orange, blue, black and red,
-respectively). R indicates the distance from the centre considering a perfect circle trajectory.}
-\label{fig:runge_err}
-\end{figure}
-
-\subsection{Conservative Velocity Interpolation (CVI)}\label{sec:CVI}
-The CVI correction is checked by means of the Stokes flow experiment by \citet{Donea2003}, as presented in \ref{sec:stokes}. The advection of Lagrangian
-markers is performed using either a 2nd-order or a 4th-order Runge-Kutta scheme with CFL=0.5 and an initial random distribution of 25 markers per element.
-Fig. \ref{fig:CVI} shows the minimum and maximum values of markers per element throughout the experiment with and without the CVI correction and using
-different Runge-Kutta schemes. The introduction of the CVI correction allows the
-tests to run for times longer than $t=2000$ without show empty elements, while tests without the CVI correction stop at approximately $t=600$, when at least
-one element goes empty. Fig. \ref{fig:CVI_time} shows the different number of markers per element and the markers distribution at $t=600$ obtained without and
-with the CVI correction (panels b and c, respectively) in case of a 2nd-order Runge-Kutta scheme. All data can be found at
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/CVI}.
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/CVI.pdf}
-\caption{Maximum and minimum number of markers per element throughout the experiment, with (green and orange lines) or without (purple and blue lines)
-the CVI correction and using either a 2nd-order (blue and orange lines) or a 4th-order (purple and green lines) Runge-Kutta scheme.}
-\label{fig:CVI}
-\end{figure}
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/CVI_time.pdf}
-\caption{Number of markers per element and markers distribution through time using 2nd.order Runge-Kutta scheme.
-Panels a: number of markers per element (panel a1) and distribution of the markers (panel a2) after the first
-iteration; panels b and c: comparison between number of markers per element (panels b1 and c1) and markers distribution (panels b2 and c2) without and with
-the CVI correction at $t=600$; panels d: number of markers per element (panel d1) and markers distribution (panel d2) in case of CVI correction at $t=2000$.}
-\label{fig:CVI_time}
-\end{figure}
-
-\subsection{Poiseuille flow}\label{sec:poiseuille}
-The domain is a rectangle with $L_x=2$ and $L_y=1$ and constant density and viscosity ($\rho=1$ and $\eta=1$), gravity acceleration $\bm{g}=\bm{0}$ and penalty
-parameter $\lambda=10^8$. The grid is composed by \num{40x20} elements. Velocity boundary conditions are set to no slip $(\bm{v}=\bm{0})$ at the top and the
-bottom, and a parabolic profile is imposed on the sides, with $u=y(1-y)$ and $v=0$. The analytical solution is then given by:
-\begin{eqnarray}
-u(x,y)&=&y(1-y)\nonumber \\
-v(x,y,)&=&0\nonumber \\
-p(x,y)&=&2\eta\left(\frac{L_x}{2}-x\right)\nonumber
-\end{eqnarray}
-
-The velocity field predicted by the model follows the expected parabolic profile (Fig. \ref{fig:poiseuille}a and Fig. \ref{fig:poi_plot}a) and, in case of the
-classic penalty method with no iterations, the pressure is clearly related to the divergence by means of the penalty parameter (Fig. \ref{fig:poiseuille}b and
-c and Fig. \ref{fig:poi_plot}b). Fig. \ref{fig:poiseuille}d shows that one Uzawa iteration is sufficient to bring the divergence down to \num{e-15}, with no
-correlation with the pressure field. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Poiseuille%20Flow}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Poiseuille.pdf}
-\caption{Velocity field (panel a), pressure (panel b) and divergence velocity of a Poiseuille flow in case of the classic penalty method (no iterations) and
-after one Uzawa iteration (panel c and d, respectively).}
-\label{fig:poiseuille}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Analytical.pdf}
-\caption{Velocity (panel a) and pressure (panel b) field predicted by the model for a Poiseuille flow with respect with their analytical solutions. The
-velocity field is plotted as function of the vertical coordinate in $L_x/2$ and the pressure is plotted as function of the x coordinate in $L_y/2$.}
-\label{fig:poi_plot}
-\end{figure}
-
-\subsection{Instantaneous 2D sphere}\label{sec:ist_sphere}
-The domain is a unit square with gravity $\bm{g}=(0,-1)$. The fluid has constant density and viscosity ($\rho_f=1$ and $\eta_f=1$). The sphere is in the middle
-of the domain with a radius R = 0.123456798 and has constant density and viscosity ($\rho_s=10^{-2}+\rho_f$ and $\eta_s=10^3 \cdot \eta_f$).
-We distinguish three different types of velocity boundary conditions:
-\begin{enumerate}
-\item FS: free slip conditions on all sides.
-\item NS: no slip conditions on all sides.
-\item OT: free slip conditions on the sides and the bottom, and open boundary at the top.
-\end{enumerate}
-All three cases are performed for different grid resolutions, between \num{16x16} and \num{512x512} elements with 50 randomly distributed markers per element,
-and different average schemes for the viscosity (harmonic, geometric and arithmetic). The velocity in the centre of the sphere $\bm{v}(0.5;0.5)$), minimum and
-maximum velocities ($u_{min}$, $u_{max}$, $v_{min}$, $v_{max}$) and pressures ($p_{min}$, $p_{max}$) on the entire domain, $v_{\textrm{rms}}$ and average pressure on the
-entire domain ($p_{avg}$) are compared with solutions generated by ASPECT \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and all results
-can be found at \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere2D} and 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Instantaneous_2D_Sphere}. Results in terms of the $v_{\textrm{rms}}$ for NS model 
-are also shown in Fig. \ref{fig:inst_sphere}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_NS.pdf}
-\caption{$v_{\textrm{rms}}$ as function of element size for the instantaneous 2D sphere experiment in case of no slip boundary conditions and with different average
-schemes for the viscosity. Results are compared with results obtained by other numerical codes.}
-\label{fig:inst_sphere}
-\end{figure}
-
-\subsection{Rayleigh-Taylor instability}\label{sec:rayleigh}
-This problem has been originally presented by \citet{vanKeken1997} and here it is performed as the isoviscous case (Case 1a). The domain has $L_x=0.9142$ and
-$L_y=1$ and gravity $\bm{g}=(0,-1)$. Two fluids with same constant viscosities ($\eta_1=\eta_2=1$) and different densities ($\rho_1=1000$ and $\rho_2=1010$),
-with the lighter fluid at the bottom. The initial interface between the fluids is given by $y(x)=0.2+0.02 \cos \left(\frac{\pi x}{L_x}\right)$. The experiment
-is performed with different grid sizes (\num{50x50}, \num{80x80}, \num{100x100} and \num{256x256}). A total of 1960000 markers are randomly distributed at the
-beginning of the simulation. Velocity boundary conditions are set to no slip at the top and the bottom, and to free slip at the sides of the domain.
-
-$v_{\textrm{rms}}$ as function of time is reported in Fig. \ref{fig:RT}, matching well with results shown by \citet{vanKeken1997}, \citet{Tackley2003} and
-\citet{Thieulot2014}. Fig. \ref{fig:rayleigh} shows the evolution of the experiment at different time steps. All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Rayleigh_Taylor_experiment/ISOVISCOUS}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/RT.pdf}
-\caption{$v_{\textrm{rms}}$ of the Rayleigh-Taylor experiment as function of time for different resolution of the grid.}
-\label{fig:RT}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=300px]{./Figures/Rayleigh.pdf}
-\caption{Evolution of the Rayleigh-Taylor experiment for a grid of \num{256x256} elements at $t=0, 500, 1000$ and $2000$ (panels a, b, c and d, respectively).}
-\label{fig:rayleigh}
-\end{figure}
-
-\subsection{Falling blocks}\label{sec:block}
-This benchmark is described as proposed by \citet{Gerya2003a}, \citet{Gerya2010b} and \citet{Thieulot2011}. The domain is square with $L_x=L_y=$ \SI{500}{\km}
-and the grid is composed by \num{50x50} elements with 25 markers in each element. The block is initially centred at ($x=$ \SI{250}{\km}; $y=$ \SI{400}{\km})
-and has a size of \num{100x100} km (Fig. \ref{fig:block}a). The fluid surrounding the block has $\eta_f=$ \SI{e21}{\pascal\per\s} and $\rho_f=$
-\SI{3200}{\kg\per\cubic\m}. The benchmark tests with different viscosities of the block, with $\eta_b$ from \SIrange{e15}{e27}{\pascal\per\s}. In each
-experiment also the density $\rho_b$ of the block varies from \SIrange{3220}{9900}{\kg\per\cubic\m}. Velocity boundary conditions are set to free slip
-conditions on all sides of the domain. Density distributions at $t=$ \SI{20}{\mega\year} for $\rho_b=$ \SI{3300}{\kg\per\cubic\m} and different $\eta_b$
-are plotted in Fig. \ref{fig:block}b-f, showing that the code correctly preserve the block geometry in case of large viscosity contrast, with a block stiffer
-than the surrounding fluid (Fig. \ref{fig:block}f).
-
-The velocity in the centre of the falling block at $t=0$ is measured for all experiments and, since the velocity should increase with density contrast, the
-quantity $\bm{v}/(\rho_b-\rho_f)$ is plotted as function of the viscosity contrast. All results perfectly match with those from \citet{Gerya2010b} and they
-line up on a single curve, demonstrating that the code can correctly deal with large viscosity and density contrasts (Fig. \ref{fig:falling}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Falling%20blocks}.
-
-\begin{figure}
-\noindent\includegraphics[width=300px]{./Figures/Block.pdf}
-\centering
-\caption{Density evolution of the falling block experiment at $t=$ \SI{0}{\mega\year} (panel a) and $t=$ \SI{20}{\mega\year} for different viscosities of the
-block (panels b-f).}
-\label{fig:block}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Falling.pdf}
-\caption{Initial velocity relative to the density contrast at the centre of the falling block as function of the viscosity contrast between the block and the
-surrounding fluid.}
-\label{fig:falling}
-\end{figure}
-
-\subsection{2D time-dependent Stokes sphere with deformable free surface}\label{sec:sphere}
-This experiment is performed in a unit square domain with the gravity acceleration fixed to $g_y=-1$. The sphere has $\rho_s=2$ and $\eta_s=10^3$ and is
-initially centred at ($0.5;0.6$) with radius $R=0.123456789$. The fluid surrounding the sphere has $\rho_f=1$ and $\eta_f=1$ and occupies the domain for
-$y\leq 0.75$, while for $y>0.75$ the air has $\rho_a=0$ and $\eta_a=10^{-3}$. Velocity boundary conditions are set to free slip on all sides. The Courant
-number is set to 0.25. The experiments run for 200 s using grid resolutions from \num{150x150} to \num{512x512} elements, with 25 randomly distributed markers
-per element, and for different average schemes for the viscosity. The interface between the fluid and the air is tracked by means of the markers chain. Results
-in terms of velocity and pressure fields and topography variations are compared with results obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and they can be found at
-\url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere_fs2D} and
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Time_dependent_sphere}. Results in terms of the $v_{\textrm{rms}}$ in case of 
-an arithmetic average are also shown in Fig. \ref{fig:inst_spherefs}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_arithm.pdf}
-\caption{$v_{\textrm{rms}}$ for different grid resolutions as function of time for the 2D time-dependent Stokes sphere experiment in case of an arithmetic average.
-Results are compared with results obtained by ASPECT.}
-\label{fig:inst_spherefs}
-\end{figure}
-
-\subsection{Free surface stabilisation}\label{sec:stab}
-The benchmark is performed as discussed in \citet{Kaus2010a} and \citet{Thieulot2014}. The domain is square with $L_x=L_y=$ \SI{500}{\km} and a grid resolution
-of \num{200x200} elements, each of them containing 16 markers. A buoyant fluid with $\rho_1=$ \SI{3200}{\kg\per\cubic\metre} and $\eta_1=$ \SI{e20}{\pascal\s}
-is overlain by a denser fluid with $\rho_2=$ \SI{3300}{\kg\per\cubic\metre} and $\eta_2=$ \SI{e21}{\pascal\s}. The initial interface between the fluids has an
-initial sinusoidal shape of 5 km amplitude. Velocity boundary conditions are set to free slip conditions at the sides and no slip at the bottom, while the top
-is free surface. The experiment is performed with various fixed time steps, without the use of the Courant number. The vertical position of the free surface at
-$x=L_x$ is tracked for each simulation, with or without the stabilisation algorithm.
-
-The results show an instability (drunken sailor effect) increasing the time step in case that the stabilisation algorithm is not activated
-($dt=$ \SI{4200}{\year}, green line in Fig. \ref{fig:stab}), as already observed by \citet{Kaus2010a} and \citet{Thieulot2014}. Activating the algorithm the
-instability is fixed and the simulation remains stable using time step up to \SI{50000}{\year} (red line in Fig. \ref{fig:stab}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Stabilisation_algorithm}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Kaus_random.pdf}
-\caption{Evolution of the y coordinate in $x=L_x$ as function of time for different time steps, with and without the stabilisation algorithm.}
-\label{fig:stab}
-\end{figure}
-
-\subsection{Topography relaxation}\label{sec:crameri}
-This experiment is performed as in \citet{Crameri2012} using free surface and sticky air (Case 1). The domain is rectangular with $L_x=$ \SI{2800}{\km} and
-$L_y=$ \SI{700}{\km} in case of free surface and $L_y=$ \SI{800}{\km} in case of sticky air ($h_{st}=$ \SI{100}{\km}). The grid is composed by \num{256x64}
-elements for the free surface \num{560x320} elements for the sticky air. For both cases, there is a mantle of 600 km thickness with $\eta_m=$ \SI{e21}{\pascal\s}
-overlain by a highly viscous cosine-shaped lithosphere with thickness between 93 and 107 km and $\eta_l=$ \SI{e23}{\pascal\s}. Both mediums have $\rho=$
-\SI{3300}{\kg\per\cubic\metre}. In case of sticky air experiment, the air has $\rho_a=$ \SI{0}{\kg\per\cubic\metre} and $\eta_a=$ \SI{e18}{\pascal\s}. The
-gravity acceleration is set to $g_y=$ \SI{-10}{\metre\per\square\s}. Velocity boundary conditions are set to no slip at the bottom and free slip at the sides
-of the domain. In case of sticky air, velocities on top are set to free slip conditions.
-
-The maximum topography as function of time ($h(t)$) can be analytically derived in according to \[h(t)=h_0 \exp(\gamma t)\]
-where $h_0=7$ km is the initial topography, $\gamma=\num{-0.2139e-11}$ is the characteristic relaxation rate and $t$ is the time (black line in Fig.
-\ref{fig:crameri}). Maximum topography at the characteristic relaxation time $t_{rlx}=$ \SI{14.825}{\kilo\year} can be found to be $h_{rlx}=$ \SI{2576}{\m}.
-The results show that a viscosity of the air of \SI{e18}{\pascal\s} is needed in case of $h_{st}=$ \SI{100}{\km} to correctly describe the topography relaxation
-for this problem, as pointed out by \citet{Crameri2012} (purple line in Fig. \ref{fig:crameri}). The case with a true free surface well follows the analytical
-solution (orange line in Fig. \ref{fig:crameri}). In particular, at the relaxation time in case of the true free surface predicts a maximum topography of
-\SI{2570}{\m}, with an error of \SI{6}{\m} with respect to the analytical solution. Results are also compared with results obtained with a free surface in
-UNDERWORLD (\num{256x64} elements) and SULEC (\num{401x201} elements) (with courtesy from \citealp{Crameri2012}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Topography_relaxation}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Crameri.pdf}
-\caption{Maximum topography as function of time for the Crameri benchmark (coloured lines) in comparison with the analytical solution (black line) and results
-obtained with UNDERWORLD (yellow line) and SULEC (blue line).}
-\label{fig:crameri}
-\end{figure}
-
-\subsection{Spontaneous subduction}\label{sec:subduction}
-This experiment is performed as presented by \citet{Schmeling2008}, in case of both a sticky air and a true free surface and using different average schemes
-for the viscosities. The domain is rectangular with $L_x=$ \SI{3000}{\km} and $L_y=$ \SI{700}{\km}, a grid resolution of \num{256x64} elements with 50 markers
-per element and Courant number of 0.01. At the beginning of the simulation, a 100 km-thick lithospheric layer with $\rho_m=$ \SI{3300}{\kg\per\cubic\metre}
-and $\eta_m=$ \SI{e23}{\pascal\s} is located at the top of the domain for $x$ between 1000 and 3000 km, while the asthenospheric mantle has $\rho_m=$
-\SI{3200}{\pascal\s} and $\eta_m=$ \SI{e21}{\pascal\s}. In addition, a 200 km-depth lithospheric slab is already subducted in the mantle in order to have
-a spontaneous subduction. In case of sticky air, $L_y=750$ km and a 50 km-thick air layer with $\eta_a$ from \SIrange{e19}{e21}{\pascal\s} overlie the mantle.
-Velocity boundary conditions are set to free slip on the sides and on the bottom of the domain. In case of sticky air, velocities on the top boundary are set
-to free slip conditions as well.
-
-The maximum depth of the slab is tracked for all the simulations (Fig. \ref{fig:subduction}) and they are compared with results obtained by
-\citet{Schmeling2008} with models with regular grids and comparable grid resolutions (rectangles in Fig. \ref{fig:subduction}). Low viscous sticky air models
-(i.e., \SI{e19}{\pascal\s}) enlighten a strong dependence of the sinking velocity with the chosen average scheme (dashed lines in Fig. \ref{fig:subduction}),
-while high viscous sticky air models (i.e., \SI{e21}{\pascal\s}) show a higher resistance at the trench, underestimating the correct solution (dotted lines in
-Fig. \ref{fig:subduction}). A high resistance at the trench can be also observed in case of a true free surface for the low resolution grid used (continuous
-lines in Fig. \ref{fig:subduction}), as pointed out by \citet{Schmeling2008}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Spontaneous_subduction}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Subduction.pdf}
-\caption{Maximum depth of the slab as function of time for all the simulations of the \citet{Schmeling2008} benchmark, in case of harmonic, geometric and
-arithmetic means (red, green and blue lines, respectively) using a sticky air or a true free surface (discontinuous and continuous lines, respectively).
-Red, green and blue rectangular areas indicate the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{500}{\km} in case of sticky air,
-while the grey rectangular area indicates the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{600}{\km} in case of true free surface.}
-\label{fig:subduction}
-\end{figure}
-
-\subsection{The slab detachment}\label{sec:slab}
-The slab detachment benchmark is performed as described by \citet{Schmalholz2011} and \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{1000}{\km}
-and $L_y=$ \SI{660}{\km} and a grid resolution of \num{256x256} elements. A non-linear viscous T-shaped layer with $\rho_l=$ \SI{3300}{\kg\per\cubic\metre} is
-placed at the top of the domain and surrounded by a linear viscous fluid with $\rho_f=$ \SI{3150}{\kg\per\cubic\m}. The top layer is 80 km-thick and a 250
-km-long and 80 km-wide slab is placed at $x=L_x/2$ (Fig. \ref{fig:slab}a). The effective viscosity of the top layer is given by
-\[\eta_{eff}=\eta_0 I_2^{\frac{(1-n)}{n}}\]
-with $\eta_0=$ \SI{4.75e11}{\pascal\s} and $n=4$, while the surrounding fluid has $\eta_f=$ \SI{e21}{\pascal\s}. The effective viscosity is capped between
-$\eta_{min}=$ \SI{e21}{\pascal\s} and $\eta_{max}=$ \SI{e25}{\pascal\s}. Velocity boundary conditions are set to free slip at the top and the bottom, and to
-no slip at the sides of the domain.
-
-The sinking of the slab determines a decrease of the effective viscosity in the T-shaped layer during the first part of the evolution (Fig. \ref{fig:slab}a),
-as consequence of local strain rates, while effective viscosities increase up to $\eta_{max}$ in the top layer after approximately \SI{20}{\mega\year}, when
-necking is complete (Fig. \ref{fig:slab}b).
-
-The width of necking is tracked by means of markers position at the side of the slab. Necking width and time are normalised in function of the initial slab
-(\SI{80}{\km}) and the characteristic time ($t_c=$ \SI{7.1158e14}{\s}) \citep{Schmalholz2011,Glerum2018}. The results in case of arithmetic, geometric and
-harmonic mean of the viscosity are shown in Fig. \ref{fig:necking}, compared with results from \citet{Glerum2018} for which an infinity norm mean was used.
-The influence of the average schemes agrees with results shown in Section \ref{sec:subduction}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Slab_detachment}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Slab.pdf}
-\caption{Effective viscosity for the slab detachment benchmark at the beginning of the evolution (panel a) and when necking is complete (panel b).}
-\label{fig:slab}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Necking.pdf}
-\caption{Normalised width of the necking of the slab detachment benchmark as function of normalised time.}
-\label{fig:necking}
-\end{figure}
-
-\subsection{The indenter experiment}\label{sec:indenter}
-The indenter benchmark simulate a rigid punch on a purely plastic von Mises material, for which exists an analytical solution
-\citep{Thieulot2008,Thieulot2014,Glerum2018}. The domain is a unit square with a grid resolution of \num{256x256} elements. The gravity acceleration is set
-to $\bm{g}=\bm{0}$ and effective viscosity is capped using $\eta_{min}=10^{-4}$ and $\eta_{max}=10^3$. The tolerance for the convergence of the non-linear
-solution is $tol=10^{-9}$, with a maximum number of non-linear iterations set to 500. The medium in the domain has $\rho=0.01$, an initial viscosity
-$\eta_0=10$, cohesion $C=1$ and an angle of internal friction $\phi=0$°. Velocity boundary conditions are set to no slip at the bottom and free slip at the
-sides of the domain. The top of the domain is open with the exception of the central portion (punch area) with width $w_p=0.125$, where $v=-1.05$ and $u$ is
-fixed either to 0 (rough punch experiment) or free (smooth punch experiment).
-
-Fig. \ref{fig:indenter} show the results in terms of viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure
-(panels g and h) for both experiments. The analytical solution indicate that pressure at ($0.5;1$) and ($0.5 \pm w_p;1$)is $p=\pi +1$ and $p=1$, respectively
-(black and red lines in Fig. \ref{fig:smooth_p}). Fig. \ref{fig:smooth_p} shows a clear improvement in the pressure solution when passing from a rough to a
-smooth punch (green and blue lines in Fig. \ref{fig:smooth_p}, respectively), as previously observed by \citet{Thieulot2014} and \citet{Glerum2018}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Indenter}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Indenter.pdf}
-\caption{Viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure (panels g and h) fields for rough (left column) and
-smooth (right column) punch experiments.}
-\label{fig:indenter}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Smooth_Rough.pdf}
-\caption{Nodal pressure in function of the x coordinate at $y=1$ for rough (green line) and smooth (blue line) punch experiment. Black and red lines indicate
-the analytical solution at $x=0.5$ and $x=0.5 \pm w_p$, respectively.}
-\label{fig:smooth_p}
-\end{figure}
-
 \subsection{The brick experiment}\label{sec:brick}
 An instantaneous version of the brick benchmark is performed to verify the correctness of the pressure-dependent plasticity for different angle of internal
-friction, as proposed by \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{40}{\km} and $L_y=$ \SI{10}{\km} and a grid resolution of \num{512x128}
+friction, as proposed by \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{40}{\km} and $L_y=$ \SI{10}{\km} and a grid resolution of $512\times128$
 elements. The gravity acceleration is set to $g_y=$ \SI{-10}{\m\square\s} and effective viscosity is capped using $\eta_{min}=$ \SI{e19}{\pascal\s} and
@@
   \end{figure}
-
-\subsection{Advection stabilisation}\label{sec:advection}
-The 1D advection problem proposed by \citet{Donea2003} and \citet{Thieulot2011} is performed to verify the effectiveness of the SUPG method to stabilise the
-advection term of the energy equation. The domain is a 1D segment with $L_x=1$ composed by 50 elements and a discontinuity in the thermal field placed at
-$x=0.25$. Temperature is set to 1 for $x \leq 0.25$ and to 0 for $x > 0.25$. Velocity is set to ${u}=1$ in the entire domain. The simulation is performed for
-250 time steps, with $dt=0.002$, so the thermal discontinuity should be at $x=0.75$ at the end of the simulation.
-
-Temperature profiles at the end of the simulation are shown in \ref{fig:advection} as function of the dimensionless coefficient $\gamma=\tau \bm{v}/h$. In case
-of the classic Galerkin method ($\gamma=0$, blue line in \ref{fig:advection}) the final thermal profile is characterised by strong oscillations, which are
-eliminated in case of the SUPG method ($\gamma=0.045$, orange line in \ref{fig:advection}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Advection_stabilisation}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Advection.pdf}
-\caption{Temperature profile as function of \textit{x} for the advection stabilisation benchmark. Purple line indicates the initial temperature profile; the
-green line indicates the analytical temperature profile after 250 time steps; blue line indicates the temperature profile after 250 time steps in case of the
-classic Galerkin method ($\gamma=0$); orange line indicates the temperature profile after 250 time steps in case of the SUPG method ($\gamma=0.045$).}
-\label{fig:advection}
-\end{figure}
-
-\subsection{Shear heating}\label{sec:simple_shear}
-This exercise is performed in an unit square domain composed by \num{128x128} elements. The velocity field is prescribed on the entire domain with
-$\bm{v}=(L_y-y)y\bm{e_x}$; viscosity, density and specific heat are fixed to 1, while thermal conductivity and radiogenic energy are fixed to 0. Therefore,
-the energy equation (Eq. \ref{eq:energy}) can be simplified as \[\frac{\partial T}{\partial t}=H_s\] and fixing $T(t=0)=0$, the temperature can be find as
-\[T(t)=H_s t\]
-In this case we have
-\begin{eqnarray}
-\dot{\epsilon_{xx}}&=&\frac{\partial u}{\partial x}=0 \nonumber \\
-\dot{\epsilon_{yy}}&=&\frac{\partial v}{\partial y}=0 \nonumber \\
-\dot{\epsilon_{xy}}&=&\frac{1}{2}\left(\frac{\partial v}{\partial x}+\frac{\partial u}{\partial y}\right)=\frac{1}{2}(L_y-2y)\nonumber
-\end{eqnarray}
-and, simplifying Eq. \ref{eq:shear}, shear heating can be calculated as \[H_s(x,y)=(1-2y)^2\]
-
-The solution predicted by the model in terms of velocity, temperature and shear heating match well with the analytical solutions (Fig. \ref{fig:analytical_en}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Simple_shear}.
-
-%\begin{figure}
-%\centering
-%\noindent\includegraphics[width=400px]{./Figures/Energy_Thieulot.pdf}
-%\caption{Temperature and shear heating predicted by the model after 1 time step.}
-%\label{fig:en_thieulot}
-%\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=500px]{./Figures/Analytical_Shear.pdf}
-\caption{Velocity (panel a), temperature (panel b) and shear heating (panel c) as function of the \textit{y} coordinate for the simple shear experiment. The
-solutions predicted by the model (red lines) are compared with the analytical solutions (dashed blue lines).}
-\label{fig:analytical_en}
-\end{figure}
-
-\subsection{Shear and adiabatic heating}\label{sec:shear}
-This problem is performed as presented in Exercise 9.4 in \citet{Gerya2010b}. Two materials are vertically separated in a rectangular domain with $L_x=$
-\SI{1000}{\km}, $L_y=$ \SI{1500}{\km} and a grid resolution of \num{30x20} elements. Constant thermal coefficient expansion ($\alpha=$ \SI{3e-5}{\per\kelvin}),
-temperature ($T=$ \SI{1300}{\kelvin}) and gravity acceleration ($g_y=$ \SI{-10}{\m\square\s}) are assumed in the whole domain. Fluid 1 (on the left side of the
-domain) has $\rho_1=$ \SI{3200}{\kg\per\cubic\m} and $\eta_1=$ \SI{e20}{\pascal\s}; fluid 2 (on the right side of the domain) has $\rho_2=$
-\SI{3300}{\kg\per\cubic\m} and $\eta_2=$ \SI{e22}{\pascal\s}. Velocity boundary conditions are set to free slip on all sides of the domain.
-
-As shown in Fig. \ref{fig:energy}, both shear and adiabatic heating predicted by the code (first row in Fig. \ref{fig:energy}) well recreate the results
-obtained by means of example ${Shear\_adiabatic\_heating.m}$ from \citet{Gerya2010b} (second row in Fig. \ref{fig:energy}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Adiabatic%2BShear}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Energy.pdf}
-\caption{Comparison between shear (first column), adiabatic (second column) and total (third column) energy predicted by the code (panels a, b and c) and those
-created using example ${Shear\_adiabatic\_heating.m}$ from Exercise 9.4 in \citet{Gerya2010b} (panels d, e and f).}
-\label{fig:energy}
-\end{figure}
-
-\subsection{Mantle convection}\label{sec:mantle}
-This problem is performed as presented by \citet{Blankenbach1989} (constant viscosity cases) and \citet{Thieulot2014}, in a 2D unit square domain with gravity
-acceleration $g_y=10^{10}Ra$. The experiment is performed with three different Rayleigh numbers ($Ra=10^4$, $10^5$ and $10^6$) and with different grid
-resolution (between \num{32x32} and \num{128x128} elements). The fluid has constant viscosity, initial density, heat capacity, thermal conductivity
-($\eta= \rho_0= C_p = k=1$), reference temperature ($T_0=0$) and thermal expansion coefficient ($\alpha=10^{-10}$). Temperatures are set to 0 on top and 1 on
-bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\]
-
-The solution generated by the code in terms of $v_{\textrm{rms}}$ (calculated as in Sec. \ref{sec:error}) and Nu as function of time are reported for all the
-simulations in Fig. \ref{fig:mantle}. At the steady state, $v_{\textrm{rms}}$ and Nu of all simulations converge well toward the values from \citet{Blankenbach1989},
-with lower errors for higher resolution grids (Table \ref{tab:mantle}). All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Mantle_convection}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Convection.pdf}
-\caption{$v_{\textrm{rms}}$ (panels a, c and e) and Nu (panels b, d and f) for the mantle convection benchmark as function of time for different grid resolution.
-Panels a and b show the results for $Ra=10^4$; Panels c and d show the results for $Ra=10^5$; panels e and f show the results for $Ra=10^6$. Purple lines
-indicate the convergence values for the $v_{\textrm{rms}}$ and the Nu from \citet{Blankenbach1989} at the steady state.}
-\label{fig:mantle}
-\end{figure}
-
-\begin{table}[H]
-\caption{Comparison between $v_{\textrm{rms}}$ and Nu predicted by the code for the mantle convection experiment and same values as reported in literature.}
-\centering
-\small
-\begin{tabular}{|>{\centering}m{0.09\textwidth}|>{\centering}m{0.05\textwidth}|>{\centering}m{0.24\textwidth}|>{\centering}m{0.15\textwidth}|
-  >{\centering}m{0.1\textwidth} >{\centering}m{0.1\textwidth} >{\centering\arraybackslash}m{0.1\textwidth}|}
-\toprule
-& & \multirow{2}{*}{\citet{Blankenbach1989}} & \citet{Thieulot2014} & \multicolumn{3}{c|}{FALCON}  \\
- & & & \num{200x200} & \num{32x32} & \num{64x64} & \num{128x128}  \\
-\midrule
-\multirow{2}{*}{$Ra=10^4$} & $v_{\textrm{rms}}$ & $42.864947 \pm 0.000020$ & 42.867 & 42.83226 & 42.852793 & 42.861394  \\
-           & Nu      & $4.884409 \pm 0.000010$  & 4.882  & 4.781297 & 4.857475  & 4.877573   \\
-\hline
-\multirow{2}{*}{$Ra=10^5$} & $v_{\textrm{rms}}$ & $193.21454 \pm 0.00010$  & 193.255& 193.872643 & 193.377472 & 193.252290  \\
-           & Nu      & $10.534095 \pm 0.000010$ & 10.507 & 9.602514 & 10.270735  & 10.465629   \\
-\hline
-\multirow{2}{*}{$Ra=10^6$} & $v_{\textrm{rms}}$ & $833.98977 \pm 0.00020$  & 834.712& 848.091176 & 837.767911 & 834.945793  \\
-           & Nu      & $21.972465 \pm 0.000020$ & 21.695 & 15.999266 & 19.703682  & 21.306939   \\
-\bottomrule
-\end{tabular}
-\label{tab:mantle}
-\end{table}
-
-\subsection{Viscoplastic mantle convection}\label{sec:visco_mantle}
-This benchmark is performed as Cases 1-5a described in \citet{Tosi2015} in a 2D unit square domain with two different grid resolutions of \num{32x32} 
-and \num{100x100} elements. Reference viscosity, reference density, heat capacity, thermal conductivity and thermal expansion coefficient are fixed to a constant
-value of 1 ($\eta_0= \rho_0= C_p = k=\alpha=1$), while the gravity has been chosen as $g_y=10^2$, in order to obtain a Rayleigh number $Ra=10^2$. Temperatures
-are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\].
-The viscosity field $\eta$ is calculated as the harmonic average between a linear part $\eta_{lin}$ that depends on temperature only or on temperature and 
-depth $y$ and a nonlinear-plastic part $\eta_{pl}$ that depends on the strain rate $\dot{\bm{\epsilon}}$, as follows
-\[\eta(T,y,\dot{\bm{\epsilon}})=2\left(\frac{1}{\eta_{lin}(T,y)}+\frac{1}{\eta_{pl}(\dot{\bm{\epsilon}})}\right)^{-1}\]
-The linear and the nonlinear-plastic parts of the viscosity are calculated following \citet{Tosi2015} as follows
-\[\eta_{lin}(T,y)=\exp(-\ln(\Delta\eta_T)T+\ln(\Delta\eta_y)y)\]
-\[\eta_{pl}(\dot{\bm{\epsilon}})=\eta^*+\frac{\sigma_Y}{\sqrt{\dot{\bm{\epsilon}}:\dot{\bm{\epsilon}}}}\]
-where $\Delta\eta_T$, $\Delta\eta_y$, $\eta^*$ and $\sigma_Y$ are parameters chosen as listed in Table \ref{tab:tosi_setup}.
-
-For each case, some diagnostic quantities measured at the steady state are reported in Table \ref{tab:visco_mantle} (see Section \ref{sec:error} for 
-the computation of all the quantities), showing a well fit with results described in \citet{Tosi2015}. Data and charts for all cases can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Viscoplastic_mantle_convection}.
-
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{white!100!}{gray!25!}
-%\begin{table}[H]
-%  \caption{Results of the viscoplastic mantle convection benchmark for Cases 1-4 compared with results obtained with ELEFANT and YACC (from \citealp{Tosi2015}).}
-%  \centering
-%  \resizebox{\textwidth}{!}{
-%\begin{tabular}{|c|ccc|ccc|ccc|ccc|}
-%  \toprule
-%    & \multicolumn{3}{c|}{\textbf{Case 1}} & \multicolumn{3}{c|}{\textbf{Case 2}} & \multicolumn{3}{c|}{\textbf{Case 3}} &
-%    \multicolumn{3}{c|}{\textbf{Case 4}} \\
-%   & FALCON & ELEFANT& YACC & FALCON& ELEFANT& YACC & FALCON & ELEFANT & YACC & FALCON & ELEFANT & YACC \\
-%  \midrule
-%   $\left\langle T \right\rangle$ & 0.7759 & 0.7758 & 0.7767 & 0.6043 & 0.6036 & 0.6058 & 0.7275 & 0.7275 & 0.7286 & & 0.5277 & 0.5289\\
-%    Nu$^{top}$ & 3.4280 & 3.4214 & 3.4298 & 8.5242 & 8.5115 & 8.5278 & 3.0368 & 3.0347 & 3.0374 && 6.5912 & 6.5572 \\
-%    Nu$^{bottom}$ & 3.3168 & 3.313 & 3.3143 & 8.4599 & 8.4465 & 8.3990 & 3.0086 & 2.9908 & 2.9628 && 6.5834 & 6.5243 \\
-%    $v_{\textrm{rms}}$ & 249.901 & 249.134 & 251.7997 & 141.2451 & 140.7087 & 142.2020 & 100.2776 & 100.1208 & 100.9467 && 79.1105 & 79.6202 \\
-%    $v_{\textrm{rms}}^{top}$ & 1.8907 & 1.8642 & 1.8298 & 104.7928 & 104.1228 & 105.1767 & 2.0832 & 2.0652 & 2.0374 && 74.7596 & 75.8414 \\
-%    $u_{max}^{top}$ & 2.6340 & 2.6119 & 2.5516 & 121.9212& 121.766 & 122.3238 & 2.9086 & 2.9019 & 2.8458 && 88.9146 & 89.2940 \\
-%    $\eta_{min}$ & - & - & - & $2.0303\times10^{-5}$ & $2.0600\times10^{-5}$ & $1.9853\times10^{-5}$ & $4.8129\times10^{-5}$ & $4.8080\times10^{-5}$ & $4.7907\times10^{-5}$ &
-%    & $1.9860\times10^{-4}$ & $1.9174\times10^{-4}$ \\
-%    $\eta_{max}$ & - & - & - & 1.3325 & 1.85 & 1.6095 & 0.9360 & 0.9023 & 1 && 1.52 & 1.6773 \\
-%    $\left\langle W \right\rangle$ & 2.4299 & 2.4316 & 2.4583 & 7.5648 & 7.5665 & 7.6222 & 2.0387 & 2.0384 & 2.0400 && 5.6216 & 5.6512 \\
-%    $\left\langle \Phi \right\rangle/Ra$ & 2.4311 & 2.4276 & 2.4333 & 7.5661 & 7.5585 & 7.6194 & 2.0386 & 2.0370 & 2.0335 && 5.6182 & 5.6463 \\
-%    $\delta$ & 0.05\% & 0.16\% & 1.02\% & 0.02\% & 0.11\% & 0.04\% & $<$0.01\% & 0.07\% & 0.32\% && 0.06\% & 0.09\% \\
-%  \bottomrule
-%  \end{tabular}}
-%  \label{tab:visco_mantle}
-%  \end{table}
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{2}{white!100!}{gray!25!}
-\begin{table}[H]
-  \caption{Benchmark cases and corresponding parameters.}
-  \centering
-  \small
-\begin{tabular}{|>{\centering}m{0.08\textwidth}|>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}
-  >{\centering}m{0.05\textwidth} >{\centering}m{0.05\textwidth} |>{\centering\arraybackslash}m{0.2\textwidth}|}
-  \toprule
-    Case & $Ra$ & $\Delta\eta_T$ & $\Delta\eta_y$ & $\eta^*$ & $\sigma_Y$ & Convective regime \\
-  \midrule
-  \rowcolor{gray!25!}
-    1 & $10^2$ & $10^5$ & 1 & - & - & Stagnant lid \\
-    2 & $10^2$ & $10^5$ & 1 & $10^{-3}$ & 1 & Mobile lid  \\
-    \rowcolor{gray!25!}
-    3 & $10^2$ & $10^5$ & 10 & - & - & Stagnant lid  \\
-    4 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 1 & Mobile lid \\
-    \rowcolor{gray!25!}
-    5 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 4 & Periodic \\
-  \bottomrule
-  \end{tabular}
-  \label{tab:tosi_setup}
-  \end{table}
-
-\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{gray!25!}{white!100!}
-\begin{table}
-  \caption{Results of the viscoplastic mantle convection benchmark for different grid resolutions in Cases 1-5 compared with results obtained with ELEFANT
-  and YACC ($100\times100$ elements, from \citealp{Tosi2015}) and in Stone 28 ($32\times32$ elements, from
-  \url{https://github.com/cedrict/fieldstone/tree/master/python_codes/fieldstone_28}).}
-  \centering
-  \resizebox{\textwidth}{!}{
-\begin{tabular}{|cc|ccccccccccc|}
-  \toprule
-  \multicolumn{13}{|c|}{\textbf{Case 1}} \\
-  \midrule
-  Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-  & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-  \hline
-  \rowcolor{gray!25!}
-    YACC & $100\times100$ & 0.7767 & 3.4298 & 3.3143 & 251.7997 & 1.8298 & 2.5516 & - & - & 2.4583 & 2.4333 & 1.02\% \\
-    ELEFANT & $100\times100$ & 0.7758 & 3.4214 & 3.313 & 249.134 & 1.8642 & 2.6119 & - & - & 2.4316 & 2.4276 & 0.16\% \\
-    \rowcolor{gray!25!}
-    Stone 28 & $32\times32$ & 0.7737 & 3.3987 & - & 243.872 & - & - & - & - & - & - & - \\
-%    FALCON & $32\times32$ & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & - & - & 2.4745 & 2.4864 & 0.48\% \\
-%    \rowcolor{gray!25!}
-    FALCON & $100\times100$ & 0.7759 & 3.4280 & 3.3168 & 249.901 & 1.8907 & 2.6340 & - & - & 2.4299 & 2.4311 & 0.05\% \\
-    \midrule
-    \multicolumn{13}{|c|}{\textbf{Case 2}} \\
-    \midrule
-    Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-    & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-    \hline
-    \rowcolor{gray!25!}
-     YACC & $100\times100$ & 0.6058 & 8.5278 & 8.3990 & 142.2020 & 105.1767 & 122.3238 & $1.9853\times10^{-5}$ & 1.6095 & 7.6222 & 7.6194 & 0.04\% \\
-     ELEFANT & $100\times100$ & 0.6034 & 8.5115 & 8.4465 & 140.8067 & 104.1228 & 121.766 & $2.0600\times10^{-5}$ & 1.85 & 7.5665 & 7.5585 & 0.11\% \\
-     \rowcolor{gray!25!}
-     Stone 28 & $32\times32$ & 0.6052 & 8.1715 & - & 141.518 & - & - & - & - & - & - & - \\
-%     FALCON & $32\times32$ & 0.6454 & 8.1422 & 7.4020 & 156.6908 & 95.7226 & 116.9302 & $2.2572\times10^{-5}$ & 0.7194 & 7.4289 & 7.4452 & 0.22\% \\
-%     \rowcolor{gray!25!}
-     FALCON & $100\times100$ & 0.6044 & 8.5266 & 8.4590 & 141.3684 & 104.8867 & 122.0470 & $2.0333\times10^{-5}$ & 1.2914 & 7.5707 & 7.5718 & 0.01\% \\
-     \midrule
-     \multicolumn{13}{|c|}{\textbf{Case 3}} \\
-     \midrule
-     Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-     & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-     \hline
-     \rowcolor{gray!25!} 
-     YACC & $100\times100$ & 0.7286 & 3.0374 & 2.9628 & 100.9467 & 2.0374 & 2.8458 & $4.7907\times10^{-5}$ & 1 & 2.0400 & 2.0335 & 0.32\% \\
-      ELEFANT & $100\times100$ & 0.7275 & 3.0347 & 2.9908 & 100.1208 & 2.0652 & 2.9019 & $4.8080\times10^{-5}$ & 0.9023 & 2.0384 & 2.037 & 0.07\% \\
-      \rowcolor{gray!25!}
-      Stone 28 & $32\times32$ & 0.7277 & 3.0347 & - & 100.018 & - & - & - & - & - & - & - \\
-%      FALCON & $32\times32$ & 0.7281 & 3.0618 & 2.8132 & 101.0092 & 2.1630 & 3.0178 & $4.8500\times10^{-5}$ & 0.7845 & 2.0699 & 2.0695 & 0.02\% \\
-%      \rowcolor{gray!25!}
-      FALCON & $100\times100$ & 0.7275 & 3.0368 & 3.0086 & 100.2776 & 2.0832 & 2.9086 & $4.8129\times10^{-5}$ & 0.9360 & 2.0387 & 2.0386 & $<$0.01\% \\
-      \midrule
-      \multicolumn{13}{|c|}{\textbf{Case 4}} \\
-      \midrule
-      Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-      & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-      \hline
-      \rowcolor{gray!25!} 
-      YACC & $100\times100$ & 0.5289 & 6.5572 & 6.5243 & 79.6202 & 75.4814 & 89.2940 & $1.9174\times10^{-4}$ & 1.6773 & 5.6512 & 5.6463 & 0.09\% \\
-       ELEFANT & $100\times100$ & 0.5277 & 6.5912 & 6.5834 & 79.1105 & 74.7596 & 88.9146 & $1.9860\times10^{-4}$ & 1.5200 & 5.6216 & 5.6182 & 0.06\% \\
-       \rowcolor{gray!25!}
-       Stone 28 & $32\times32$ & 0.5286 & 6.4036 & - & 79.4746 & - & - & - & - & - & - & - \\
-%       FALCON & $32\times32$ & 0.5334 & 6.4679 & 6.3435 & 80.6153 & 76.0511 & 90.2058 & $2.0393\times10^{-4}$ & 0.6453 & 5.6676 & 5.6698 & 0.04\% \\
-%       \rowcolor{gray!25!}
-       FALCON & $100\times100$ & 0.5279 & 6.5897 & 6.5830 & 79.1220 & 75.0715 & 88.8484 & $1.9693\times10^{-4}$ & 1.4133 & 5.6127 & 5.6113 & 0.03\% \\
-       \midrule
-       \multicolumn{13}{|c|}{\textbf{Case 5}} \\
-       \midrule
-       Code & Resolution & $P$ & $\left\langle T \right\rangle_{min}$ & $\left\langle T \right\rangle_{max}$ & Nu$^{top}_{min}$ & Nu$^{top}_{max}$ &
-       $v_{\textrm{rms}}^{min}$ & $v_{\textrm{rms}}^{max}$ & \multicolumn{2}{c}{$(\left\langle \Phi \right\rangle/Ra)_{min}$} & 
-       \multicolumn{2}{c|}{$(\left\langle \Phi \right\rangle/Ra)_{max}$} \\
-       \hline
-       \rowcolor{gray!25!}
-         YACC & $100\times100$ & 0.0792 & 0.6543 & 0.6722 & 2.6950 & 7.2792 & 41.9813 & 98.8230 & \multicolumn{2}{c}{1.4347} & 
-         \multicolumn{2}{c|}{9.2748} \\
-         ELEFANT & $100\times100$ & 0.076 & 0.6514 & 0.6693 & 2.6808 & 7.3506 & 41.2578 & 100.089 & \multicolumn{2}{c}{1.3324} & 
-         \multicolumn{2}{c|}{9.5289} \\
-         \rowcolor{gray!25!}
-         Stone 28 & $32\times32$ & 0.0764 & 0.6521 & 0.6697 & 2.6912 & 7.0763 & 41.61 & 98.183 & \multicolumn{2}{c}{-} & 
-         \multicolumn{2}{c|}{-} \\
-     %    FALCON & $32\times32$ & & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & \multicolumn{2}{c}{-} & 
-     %     \multicolumn{2}{c|}{-} \\
-     %    \rowcolor{gray!25!}
-         FALCON & $100\times100$ & 0.0775 & 0.6519 & 0.6697 & 2.6847 & 7.3629 & 41.5256 & 99.5466 & \multicolumn{2}{c}{1.3399} & 
-         \multicolumn{2}{c|}{9.3993} \\
-  \bottomrule
-  \end{tabular}}
-  \label{tab:visco_mantle}
-  \end{table}
-
-\subsection{Thin layer entrainment}\label{sec:thin}
-This experiment is performed as originally proposed by \citet{vanKeken1997} with two incompressible fluids in a rectangular domain with $L_x=2$ and $L_y=1$
-and gravity acceleration $g_y=10^{10}Ra$. The Rayleigh number and the compositional Rayleigh number are fixed to $Ra=300000$ and $Rc=450000$, respectively.
-Both fluids have constant viscosity, thermal conductivity, specific heat ($\eta=\rho=C_p=1$) and thermal expansion coefficient ($\alpha=10^{-10}$). Fluid 1
-has a density $\rho_1=1$, while fluid 2 is denser ($\rho_2=\rho_1+1.5 \cdot 10^{-10}$) and is placed at the bottom of the domain, for $y \leq 0.025$.
-Temperature are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides of the domain. The initial
-temperature field is given by
-\[T(x,y)=T_u(x,y)+T_l(x,y)+T_r(x,y)+T_s(x,y)-\frac{3}{2}\]
-with
-\begin{eqnarray}
-T_u(x,y)&=&\frac{1}{2}erf\left(\frac{1-y}{2}\sqrt{\frac{u_0}{x}}\right)\nonumber \\
-T_l(x,y)&=&1-\frac{1}{2}erf\left(\frac{y}{2}\sqrt{\frac{u_0}{L_x-x}}\right)\nonumber \\
-T_r(x,y)&=&\frac{1}{2}+\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{y+1}} \exp\left(-\frac{x^2u_0}{4y+4}\right)\nonumber \\
-T_s(x,y)&=&\frac{1}{2}-\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{2-y}} \exp\left(-\frac{(L_x-x)^2u_0}{8-4y}\right)\nonumber
-\end{eqnarray}
-and
-\begin{eqnarray}
-u_0&=&\frac{L_x^{7/3}}{(1+L_x^4)^{2/3}}\left(\frac{Ra}{2\sqrt{\pi}}\right)^{\frac{2}{3}}\nonumber \\
-Q&=&2\sqrt{\frac{L_x}{\pi u_0}}\nonumber
-\end{eqnarray}
-
-The experiment is performed with two grid resolution (\num{125x40} and \num{200x80} elements), with 100 markers per element and CFL of 0.25. $v_{\textrm{rms}}$
-(see Sec. \ref{sec:error}) calculated for both the simulations match well with results from \citet{vanKeken1997} and obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and ELEFANT \citep{Thieulot2014}, obtained with similar grid resolutions
-(Fig. \ref{fig:thin}). In addition, a variety of simulations with different resolutions and aspect ratios are performed to verify that the sensitivity of the
-initial velocity field. Results match well with results from \citet{vanKeken1997} (in the grey area) and \citet{Thieulot2014} (Fig. \ref{fig:thin_initial}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Thin_layer}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_thin.pdf}
-\caption{$v_{\textrm{rms}}$ for the thin layer experiment as function of time for different grid resolution. Results are compared with results obtained by
-\citet{vanKeken1997} (black lines) and with ASPECT and ELEFANT.}
-\label{fig:thin}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrmszero.pdf}
-\caption{$v_{\textrm{rms}}$ at $t=0$ for the thin layer experiment as function of the element size for different aspect ratios of the elements. Results are compared
-with results obtained by \citet{vanKeken1997} (black dots in the grey area) and with ASPECT and ELEFANT.}
-\label{fig:thin_initial}
-\end{figure}
-
-\subsection{Sinking hydrated cylinder}\label{sec:quinquis}
-This experiment is performed as explained in \citet{Quinquis2014} and simulates the sinking of a cold, hydrated cylinder into a warm, dry mantle. The domain
-is square with $L_x=L_y=$ \SI{300}{\km}, a grid resolution of \num{300x300} elements and 25 markers per element. The sinking cylinder is located at $x=$
-\SI{150}{\km} and $y=$ \SI{170}{\km} with a radius of \SI{20}{\km} and has $\rho_c=$ \SI{3250}{\kg\per\cubic\m}, $\eta_c=$ \SI{e23}{\pascal\s}, a thermal
-conductivity $k_c=$ \SI{4.5}{\watt\per\m\per\kelvin},a specific heat $C_{p_c}=$ \SI{1250}{\joule\per\kg\per\kelvin}; the surrounding mantle has $\rho_m=$
-\SI{3200}{\kg\per\cubic\m}, $\eta_m=$ \SI{e20}{\pascal\s}, a thermal conductivity $k_m=$ \SI{105}{\watt\per\m\per\kelvin} and specific heat $C_{p_m}=$
-\SI{1250}{\joule\per\kg\per\kelvin}; a 58 km-thick lithosphere overlies the mantle and has $\rho_l=$ \SI{3200}{\kg\per\cubic\m}, $\eta_l=$ \SI{e23}{\pascal\s},
-a thermal conductivity $k_l=$ \SI{4.5}{\watt\per\m\per\kelvin}, specific heat $C_{p_l}=$ \SI{750}{\joule\per\kg\per\kelvin} and a thermal diffusivity of
-\SI{e-6}{\square\m\per\s}. Velocity boundary conditions are set to free slip on all sides of the domain. Temperature increase linearly in the lithosphere,
-from 0° to 1300°C, and in the mantle, from 1300° to 1360.5°C. The initial temperature of the cylinder is fixed to 400°C. Throughout the evolution, temperature
-is fixed to 0° and 1360.5°C at the top and bottom of the domain, respectively. The initial bound water content is imposed at 2 wt.\% in the cylinder and at
-0 wt.\% in both the mantle and the lithosphere. The maximum amount of water in the mantle is fixed to 0.2 wt.\%, while maximum water content in the cylinder
-and in the lithosphere is function of pressure and temperature and is calculated using a serpentinized harzburgite with Perple\_X, as in \citet{Quinquis2014}.
-
-As shown in \citet{Quinquis2014}, the model is characterised by a progressive dehydration in the external portion of the cylinder, due to the increase of
-temperature, with a consequent vertical migration of free water that hydrates the mantle above the cylinder, up to the lithosphere. The distribution of
-bound and free water during the evolution of the experiment is shown in Fig. \ref{fig:quinquis} (left and right columns, respectively). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Hydration}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Hydration.pdf}
-\caption{Distribution of bound and free water (left and right columns, respectively) at $t=0.5$ and $t=$ \SI{1.25}{\mega\year} (first and second rows,
-respectively) for the sinking hydrated cylinder experiment. Only markers with bound or free water not equal to 0 are plotted.}
-\label{fig:quinquis}
-\end{figure}
-
-\subsection{Mantle melting curves}\label{sec:katz}
-This instantaneous experiment is performed in order to recreate isobaric melting curves, as function of temperature, bulk water and content of water in the
-melt, to compare with melting curves obtained by \citet{Katz2003}. The domain is square with $L_x=L_y=100$ km, a grid resolution of \num{100x100} elements and
-25 markers per element. Density is set in the entire domain to \SI{3300}{\kg\per\cubic\m} and temperature increases from the sides to the centre from 1000° to
-1500°C. The experiment is performed for bulk water contents from 0 to 0.3 wt.\%.
-
-Fig. \ref{fig:fraction} shows isobaric melting curves for \SI{1}{\giga\pascal} (panel a) and \SI{3}{\giga\pascal} (panel b) calculated for different bulk water
-contents. The curves well match with those obtained for the mantle by \citet{Katz2003}, showing the saturation at the solidus for bulk water of 0.2 and 0.3 wt.\%, indicating
-that the code correctly determines melt fraction as function of pressure, temperature and bulk water content. Same results can be observed in Fig.
-\ref{fig:water}a, where isobaric and isothermal melting curves are shown as function of bulk water content. In addition, Fig. \ref{fig:water}b shows that the
-code correctly determines also the percentage of water in the melt as function of pressure, temperature and melt fraction. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Melting}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Fraction.pdf}
-\caption{Isobaric melting curves for 1 GPa (panel a) and 3 GPa (panel b) as function of temperature with different bulk water contents.}
-\label{fig:fraction}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Water.pdf}
-\caption{Isobaric melting curves for 1.5 GPa as function of bulk water (panel a) and water in the melt (panel b) calculated for different temperatures.}
-\label{fig:water}
-\end{figure}
-
-\clearpage
-
-\bibliography{export}
-
-\end{document}

### 2022-09-07 tex/Benchmark/crameri.tex
update main

--- 2022-03-31 FALCON_description.tex
@@
-\documentclass[hidelinks,10pt,a4paper]{article}
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{textcomp}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage{booktabs}
-\usepackage{multirow}
-\usepackage[cm]{fullpage}
-\usepackage[round]{natbib}
-\bibliographystyle{abbrvnat}
-\usepackage{doi}
-\usepackage{bm}
-\usepackage[table]{xcolor}
-\hypersetup{
-    colorlinks,
-    linkcolor={blue},
-    citecolor={blue},
-    urlcolor={blue}
-}
-\usepackage{float}
-\restylefloat{table}
-\usepackage{siunitx}
-\DeclareSIUnit\year{yr}
-
-\begin{document}
-
-\section{Introduction}\label{sec:intro}
-This document includes a description of the code FALCON, with a presentation of the features implemented followed by results of extensive benchmarks listed in
-Table \ref{tab:list}, performed to verify that the code solves correctly a wide range of problems. Description and results of all benchmarks are shown
-in Section \ref{sec:benchmark}.
-
-\begin{table}[H]
-\caption{List of benchmarks performed for each feature of the code. Data of all benchmarks can be found at 
-\url{https://github.com/aleregorda/Benchmarks}}
-\centering
-\begin{tabular}{| l | l | l |}
-\toprule
- \textbf{Feature}  & \textbf{Benchmark} & \textbf{Section} \\
-\midrule
-  Solver  & Stokes flow & \ref{sec:stokes} \\
-\hline
-  \multirow{2}{*}{Markers advection} & Zalesak disk & \ref{sec:runge} \\
-    & Conservative Interpolation Velocity & \ref{sec:CVI} \\
-\hline
-  \multirow{4}{*}{Momentum equation}  & Poiseuille flow & \ref{sec:poiseuille}\\
-    & Instantaneous 2D sphere & \ref{sec:ist_sphere} \\
-    & Rayleigh-Taylor experiment & \ref{sec:rayleigh} \\
-    & Falling block & \ref{sec:block} \\
-\hline
-    & 2D Stokes sphere & \ref{sec:sphere} \\
-    Sticky air and  & Stabilisation algorithm & \ref{sec:stab} \\
-    free surface & Topography relaxation & \ref{sec:crameri} \\
-  & Spontaneous subduction & \ref{sec:subduction} \\
-\hline
-  Erosion and sedimentation  & - & - \\
-\hline
-   Non-linear  & Slab detachment & \ref{sec:slab} \\
-   visco-plasticity & Indenter & \ref{sec:indenter} \\
-    & Brick & \ref{sec:brick} \\
-\hline
-  \multirow{3}{*}{Energy equation}  & Advection stabilisation & \ref{sec:advection} \\
-    & Simple shear heating & \ref{sec:simple_shear} \\
-    & Shear and adiabatic heating & \ref{sec:shear} \\
-\hline
-\multirow{3}{*}{Energy + momentum} & Mantle convection & \ref{sec:mantle} \\
-& Viscoplastic mantle convection & \ref{sec:visco_mantle} \\
-& Thin layer entrainment & \ref{sec:thin} \\
-\hline
-  Phase changes  & \multirow{2}{*}{Hydrated sinking cylinder} & \multirow{2}{*}{\ref{sec:quinquis}} \\
-  and hydration  &  & \\
-\hline
-  Melting & Experimental melting curves & \ref{sec:katz} \\
-\bottomrule
-\end{tabular}
-\label{tab:list}
-\end{table}
-
-The code is written in Fortran90 and uses the finite element method (FEM) with quadrilateral $Q_1 \times P_0$ elements (continuous bilinear velocity and
-discontinuous constant pressure), associated with MUMPS\footnote{\url{http://mumps.enseeiht.fr/}} \citep{Amestoy2001,Amestoy2019} that is a software package
-for solving systems of linear equations of the form $A \cdot x = b$, where \textit{A} is a square sparse matrix, by means of a direct method. Since
-$Q_1 \times P_0$ elements do not satisfy Ladyzhenskaya, Babuska and Brezzi (LBB) stability condition \citep{Donea2003}, elemental pressure is elaborated
-in post-processing to avoid spurious pressures by means of a double interpolation to smooth the pressure. Following this procedure, the elemental pressure
-is interpolated onto nodes and then back onto elements. After the smoothing procedure, the pressure field is calculated again on the nodes to be used in
-combination with the lithostatic pressure, which is calculated onto nodes. Correctness of the solution and performances of the solver in terms of time and
-memory usage are tested solving the Stokes flow with the analytical solution proposed by \citet{Donea2003} (Section \ref{sec:stokes}).
-
-The thermo-mechanics of crust-mantle systems is described by means of the conservation of mass, momentum and energy equations, expressed as follows:
-\begin{eqnarray}
-\label{eq:mass}\frac{\partial \rho}{\partial t}+\bm{\nabla} \cdot (\rho \bm{u})&=&0\\
-\label{eq:momentum}-\bm{\nabla} p + \bm{\nabla} \cdot \bm{\tau} + \rho \bm{g}&=&0\\
-\label{eq:energy}\rho C_p \left(\frac{\partial T}{\partial t} + \bm{u} \cdot \bm{\nabla} T \right)&=&\bm{\nabla} \cdot \left(k\bm{\nabla} T\right) + H_{tot}
-\end{eqnarray}
-where $\rho$ is the density, $\bm{u}$ is the velocity, \textit{p} is the pressure, $\bm{\tau}$ is the deviatoric stress, $\bm{g}$ is the gravity acceleration,
-$C_p$ is the specific heat at constant pressure, \textit{T} is the temperature, \textit{k} is the thermal conductivity and $H_{tot}$ is the total internal heat
-production. Density variations due to temperature are generally small enough to assume the density as constant ($\rho=\rho_0$) in Eq. \ref{eq:mass} and in
-Eq. \ref{eq:energy}, while it must be treated as a variable in the buoyancy term of Eq. \ref{eq:momentum}, such that \[\rho=\rho_0(1-\alpha(T-T_0))\]
-where $\rho_0$ is the density at a reference temperature $T_0$ and $\alpha$ is the coefficient of thermal expansion. Eq. \ref{eq:mass} then can be rewritten as
-\begin{equation}\label{eq:mass_0}
-\bm{\nabla} \cdot \bm{u}=0
-\end{equation}
-This simplification is known as the extended Boussinesq approximation.
-
-The time step $dt$ is chosen in according to Courant-Friedrichs-Lewy condition \citep{Anderson1995}, such that
-\[dt=C_n\; \min\left(\frac{\Delta x}{\max|\bm{u}|},\frac{(\Delta x)^2}{\kappa}\right)\]
-where $0<C_n<1$ is the Courant (CFL) number, $\Delta x$ is the minimum dimension of smallest element, $\max|\bm{u}|$ is the maximum velocity calculated on the
-entire domain and
-$\kappa=\frac{k}{\rho C_p}$ is the thermal diffusivity \citep{Thieulot2014}.
-
-\section{Lagrangian markers}\label{sec:markers}
-Elemental properties (density, viscosity, thermal conductivity, specific heat and thermal expansion) needed to solve Eqs. \ref{eq:mass_0}, \ref{eq:momentum}
-and \ref{eq:energy} are related to the composition of each element, determined by means of Lagrangian markers that are characteristic of different materials
-in the domain. Elemental properties (with the sole exception of the viscosity) are calculated using an arithmetic mean, as \[P_e=\frac{1}{n}\sum_{i=1}^n P_i\]
-where $P_e$ is the elemental property, $P_i$ is the property characteristic of the material of each marker and $n$ is the number of markers of the element.
-Differently, the average scheme for the elemental viscosity can be chosen between harmonic, geometric and arithmetic mean. Markers can be placed regularly or
-randomly at the beginning of the simulation and their advection is performed by either a 2nd-order or a 4th-order Runge-Kutta in space, interpolating the velocity 
-field on each marker by means of the shape functions (see Section \ref{sec:runge}). The interpolated velocity is then corrected by means of the Conservative
-Velocity Interpolation (CVI), which introduces a corrective term to reduce dispersion and clustering of particles in both steady state and time-dependent 
-\citep{Wang2015} (see Section \ref{sec:CVI}).
-
-The initial distribution of the markers is created using the open-source code library
-Geodynamic World Builder \citep{Fraters2019}. During the simulation, each marker carries memory of temperature, pressure and accumulated strain, which are
-determined interpolating the relative nodal parameter. In particular, pressure is interpolated onto each marker after the smoothing procedure. The number
-of markers contained in each element is maintained between $n_{min}$ and $n_{max}$. When in an element there are less markers than $n_{min}$ the code adds
-random markers to reach the $n_{min}$, while if the number is higher than $n_{max}$ some of them are randomly deleted. When new markers are added, they assume
-the properties of the nearest marker. In this way elements are never empty and maintain a number of markers inside a prefixed range.
-
-\section{The momentum equation}\label{sec:mom_eq}
-The penalty method is implemented so that Eq. \ref{eq:mass_0} can be written as:
-\begin{equation}\label{eq:mass_penalty}
-\bm{\nabla} \cdot \bm{u} + \frac{p}{\lambda}=0
-\end{equation}
-where $\lambda$ is the so-called penalty parameter, which should be 6-7 orders of magnitude larger than the shear viscosity to ensure that mass conservation
-is satisfied \citep{Donea2003,Thieulot2014}.
-
-The deviatoric stress tensor in Eq. \ref{eq:momentum} can be written in terms of the strain rate tensor as $\bm{\tau}=2\eta\dot{\bm{\epsilon}}$, with
-$\dot{\bm{\epsilon}}=\frac{1}{2}\left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)$. Therefore, Eq. \ref{eq:momentum} can be rewritten as:
-\begin{equation}\label{eq:momentum_pressure}
--\bm{\nabla} p + \bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right) + \rho \bm{g} = 0
-\end{equation}
-Finally, using pressure from Eq. \ref{eq:mass_penalty}, Eq. \ref{eq:momentum_pressure} can be rewritten as:
-\begin{equation}\label{eq:momentum_penalty}
-\lambda \bm{\nabla} \left(\bm{\nabla} \cdot \bm{v}\right)+\bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right)+\rho\bm{g}=0
-\end{equation}
-The penalty method is been associated to the iterative Uzawa method, as described in detail in \citet{Dabrowski2008} and \citet{Thieulot2014}
-(see Section \ref{sec:poiseuille}). Results of benchmarks performed to verify the correctness in the implementation of Eq. \ref{eq:momentum_penalty} are
-shown in Sections \ref{sec:ist_sphere} and \ref{sec:rayleigh}.
-
-To support large viscosities variations the penalty parameter is related to the effective elemental viscosity by means of a dimensionless coefficient, so that
-$\lambda=\lambda_e(e)\eta_{eff}(e)$ \citep{Marotta2006,Dabrowski2008,Thieulot2014}. Benchmark of the falling block \citep{Gerya2003a,Gerya2010b,Thieulot2011}
-is performed to verify that the code can correctly deal buoyancy driven flows with strong viscosity contrasts (Section \ref{sec:block}).
-
-\section{Sticky air and free surface}\label{sec:sticky}
-The Earth's surface can be treated by means of either the so-called sticky-air or a true free surface method, both of them implemented in the code. In the
-sticky-air method the surface is approximated with the introduction of a buoyant layer with a viscosity at least four orders of magnitude lower than the crust
-\citep{Schmeling2008,Crameri2012} and the interface between lithosphere and air is defined using a chain of passive markers that are advected as the Lagrangian
-markers. The correctness of the evolution of the markers chain is tested with the experiment of a 2D time-dependent Stokes sphere below a free surface and
-compared with results from ASPECT\footnote{\url{https://aspect.geodynamics.org/}} \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} (Section
-\ref{sec:sphere}). In the true free surface case the top boundary is assumed stress-free and velocities are not fixed. In this case topography variations are
-described by vertical deformations of the mesh that depend on the velocity field of the nodes that identify the free surface, while horizontal deformations are
-not taken into account. This procedure is known as the Arbitrary Lagrangian-Eulerian (ALE) method and its implementation follows the technique described in
-\citet{Thieulot2011}. However, although the implementation of a true free surface better reproduces laboratory experiment, extremely small time steps can be
-necessary to maintain stability \citep{Kaus2010a,Quinquis2011,Thieulot2014}. Therefore, the stability algorithm proposed by \citet{Kaus2010a} is implemented
-to avoid instabilities due to high density differences at the free surface when using too large time steps. The implementation of this algorithm are tested by
-performing the experiment described by \citet{Kaus2010a} (Section \ref{sec:stab}).
-
-The topography relaxation benchmark proposed by \citet{Crameri2012} is performed to verify that the code correctly recovers topography variations in case of
-both the sticky-air and the true free surface method (Section \ref{sec:crameri}). Finally, Section \ref{sec:subduction} show the results of the spontaneous
-subduction experiment described by \citet{Schmeling2008}.
-
-\section{Erosion and sedimentation}\label{sec:erosion}
-Surface processes at either the lithosphere-air interface or in correspondence of the free surface have been implemented by means of the software
-Fastscape\footnote{\url{https://fastscape.org/}} \citep{Braun2013,Cordonnier2019,Yuan2019a,Yuan2019b}. FastScapeLib is a set of routines that solve the stream
-power law (enriched by a sediment transport/deposition term) \citep{Yuan2019a}, hillslope diffusion and marine transport and deposition \citep{Yuan2019b}.
-The partial differential equation solved by FastScapeLib is:
-\begin{equation}
-\frac{\partial h}{ \partial t}=U-K_f A^{m_e}S^{n_e} + \frac{G}{A}\int_A \left( U-\frac{\partial h}{\partial t} \right) dA + K_d \bm{\nabla}^2 h 
-\end{equation}
-where $h$ is the topography, $U$ is the uplift, $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $S$ is the slope, $m_e$ and $n_e$ are
-parameters that should be chosen to constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}, $G$ is the dimensionless
-deposition/transport coefficient for the enriched stream power law and $K_d$ is the effective diffusivity. Terms $K_f A^{m_e}S^{n_e}$ and $K_d \bm{\nabla}^2 h$
-consider long-range processes, described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, and short-range processes,
-proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}.
-
-In case of sticky air, slope is determined by means of the markers chain, which is vertically corrected considering erosion and sedimentation rates. After the
-correction, continental markers above the markers chain are transformed in air markers and air markers below the markers chain become sediments. Similarly,
-slope is determined considering the nodes on the top boundary in case of true free surface and they are vertically corrected as for the markers chain. In this
-case, continental markers above the top boundary are simply deleted.
-%Surface processes at the lithosphere-air interface are implemented taking into account fluvial erosion (long-range processes), hillslope processes (short-range
-%processes) and failure of slopes exceeding a threshold angle $\theta$ \citep{Burov1997,Tucker1999,Sacek2012,Croissant2014,Ueda2015,Beucher2020}. Erosion rates
-%$\dot{E}_{lr}$ due to long-range processes can be described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{lr}=K_f A^{m_e} \bm{\nabla} h^{n_e}
-%\end{equation}
-%where $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $\bm{\nabla} h$ is the slope and $m_e$ and $n_e$ are parameters chosen in to
-%constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}. Erosion rates $\dot{E}_{sr}$ due to short-range processes are
-%implemented by assuming that transport is linearly proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{sr}=K_d \bm{\nabla}^2 h
-%\end{equation}
-%where $K_d$ is the effective diffusivity. In addition, the slope is modified if it still exceeds a threshold angle after long- and short-range processes
-%\citep{Ueda2015}.
-
-%Offshore sedimentation processes are implemented similarly to hillslope processes but with a depth-dependent coefficient \citep{Sacek2012,Beucher2020}.
-%The sedimentation rate $\dot{S}$ is calculated as
-%\begin{equation}
-%\dot{S}=K_s \bm{\nabla}^2 h
-%\end{equation}
-%with
-%\begin{equation}
-%K_s=K^*_s \exp\left(\frac{h-h_{sl}}{h_0}\right)
-%\end{equation}
-%where $K^*_s$ is a constant, $h_{sl}$ is the sea level and $h_0$ is a characteristic length. In case of sticky air, slope is determined by means of the markers
-%chain, which is vertically corrected considering erosion and sedimentation rates. After the correction, continental markers above the markers chain are
-%transformed in air markers and air markers below the markers chain become sediments. Similarly, slope is determined considering the nodes on the top boundary
-%in case of true free surface and they are vertically corrected as for the markers chain. In this case, continental markers above the top boundary are simply
-%deleted.
-
-\section{Non-linear rheologies}\label{sec:nonlinear}
-Non-linear rheologies are implemented combining viscous creep (dislocation and diffusion) and plastic yielding. For each marker, diffusion and dislocation
-viscosity ($\eta_{df|ds}$) can be determined as follows \citep{Karato1993,Warren2008b,Wang2016}:
-\begin{equation}\label{eq:viscous}
-\eta_{df|ds}=f_s \left(\frac{d^m}{A}\right)^{\frac{1}{n}} \left (\frac{2^{\frac{1-n}{n}}}{3^{\frac{n+1}{2n}}} \right)
-I_2^{\frac{(1-n)}{n}}\exp\left(\frac{Q+pV}{nRT}\right)
-\end{equation}
-where $f_s$ is a scaling factor used to represent lithologies that are stronger or weaker than the base set, $d$ is the grain size, $m$ is the grain size exponent,
-$n$ is the stress exponent, $A$ is the uniaxial pre-exponential factor, $I_2$ is the square root of the second invariant of the strain rate tensor, $Q$ is the
-activation energy, $p$ is the pressure, $V$ is the activation volume, $R$ is the gas constant and $T$ is the temperature.
-Pressures and temperatures are determined by the interpolation of the nodal parameters, while strain rates are calculated by means of the
-derivative of the velocities on the nodes. In case of diffusion creep $n=1$ and $m>0$, in case of dislocation creep $n>1$ and $m=0$. The viscous creep $\eta_{cp}$
-is then calculated as the harmonic average between $\eta_{df}$ and $\eta_{ds}$:
-\begin{equation}\label{eq:visco_creep}
-\eta_{cp}=\left(\frac{1}{\eta_{df}}+\frac{1}{\eta_{ds}}\right)^{-1}
-\end{equation}
-The implementation of the non-linear viscous creep viscosity depending on the strain rate is tested by means of the slab detachment benchmark
-(Section \ref{sec:slab}).
-
-Plastic yielding is implemented rescaling $\eta_{cp}$ in order to limit the stress below the yield stress $\sigma_y$ \citep{Thieulot2008,Thieulot2014,Glerum2018},
-obtaining
-\begin{equation}\label{eq:plastic}
-\eta_{pl}=\frac{\sigma_y}{2I_2}
-\end{equation}
-where $\eta_{pl}$ is the plastic viscosity and the yield stress is determined following the Drucker-Prager criterion, such as
-\begin{equation}\label{eq:yield}
-\sigma_y=C\;\cos(\phi)+p\;\sin(\phi)
-\end{equation}
-where $C$ is the cohesion and $\phi$ is the internal friction angle. In case of negative pressure it is imposed equal to 0, so that negative yield stress are
-excluded. The correctness of the non-linear solution in case of plastic viscosity with variable internal friction angle is verified performing indenter and
-brick experiments (Sections \ref{sec:indenter} and \ref{sec:brick}, respectively).
-
-Strain softening is taken into account for both viscous creep and plastic viscosity \citep{Huismans2003,Babeyko2005,Huismans2005,Sobolev2005,Warren2008a} by
-means of the accumulated strain $\epsilon$ memorised by each marker. Softening in the viscous creep determines a linear decrease of $\eta_{vc}$ by means of a
-viscous strain softening factor $W_S$ that increases linearly from $W_{S_0}$ to $W_{S_{\infty}}$ for $\epsilon_{S_0}<\epsilon_S<\epsilon_{S_{\infty}}$ 
-\citep{Warren2008a}. This viscous softening can be related to strain-induced grain size reduction \citep{Warren2008a}. Differently, plastic softening is 
-simulated with a linear decrease of internal friction angle $\phi(\epsilon)$ and cohesion $C(\epsilon)$ in according to
-\begin{eqnarray}
-\label{eq:friction1}\phi(\epsilon)&=&\phi_0+(\phi_{\infty}-\phi_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}\\
-\label{eq:friction2}C(\epsilon)&=&C_0+(C_{\infty}-C_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}
-\end{eqnarray}
-where $\phi_0$, $C_0$ and $\phi_{\infty}$, $C_{\infty}$ are internal friction angle and cohesion for $\epsilon_0$ and $\epsilon_{\infty}$, respectively
-\citep{Huismans2003,Huismans2005,Warren2008a,Thieulot2014}. Plastic softening approximates deformation-induced softening of faults and brittle shear zones
-\citep{Warren2008a}.
-
-Plastic yielding and viscous creep are then combined to obtain a viscoplastic viscosity $\eta_{vp}$ as follows, assuming that they are independent processes
-\citep{Karato2008,Glerum2018}:
-\begin{equation}\label{eq:viscoplastic}
-\eta_{vp}=\min(\eta_{cp},\eta_{pl})
-\end{equation}
-Finally, effective viscosity $\eta_{eff}$ is capped by the minimum and the maximum viscosity ($\eta_{min}$ and $\eta_{max}$, respectively) to avoid extremely
-low or high viscosity \citep{Glerum2018} as follows
-\begin{equation}\label{eq:effective}
-\eta_{eff}=\min(\max(\eta_{vp},\eta_{min}),\eta_{max})
-\end{equation}
-Viscous creep and plastic yielding are non-linear rheologies because of their dependence on the velocity field through pressure and strain rates. Therefore,
-the solution is determined by means of Picard-type iterations, until convergence of the velocity field \citep{Glerum2018}. The convergence is verified at each
-iteration $i$ via the nonlinear residual $\bm{\mathcal{R}}^i$ that can be determined as
-\begin{equation}\label{eq:convergence}
-\bm{\mathcal{R}}^i= \mathbb{K} (\eta_{eff}(\dot{\bm{\epsilon}}^{i-1},p^{i-1}))\cdot \bm{v}^{i-1} - \bm{f}^i
-\end{equation}
-%\begin{equation}\label{eq:convergence}
-%\chi=1-\frac{\langle (\bm{v}^i-\langle\bm{v}^i\rangle)\cdot (\bm{v}^{i+1}-\langle\bm{v}^{i+1}\rangle) \rangle}{\sqrt{\sigma^i\sigma^{i+1}}}
-%\end{equation}
-%Iterations are performed until either convergence ($\chi<tol$) or maximal number of non-linear iterations ($it_{max}$) is reached.
-where $\mathbb{K}$ is the Stiffness matrix and $\bm{f}$ is the right hand side vector \citep{Spiegelman2016,Glerum2018}.  The $L_2$-norm is extracted from
-$\bm{\mathcal{R}}^i$ and it is normalised as follows
-\begin{equation}\label{eq:convergence_res}
-\frac{||\bm{\mathcal{R}}^i||_2}{||\bm{\mathcal{R}}^0||_2}
-\end{equation}
-where $||\bm{\mathcal{R}}^0||_2$ is the $L_2$-norm of the first nonlinear iteration.
-Since often the normalised nonlinear residual drop very quickly over the first 2-3 iterations, also the $L_2$-norm of velocity and pressure residuals are
-calculated as
-\begin{equation}\label{eq:convergence_vel}
-\frac{||\Delta\bm{v}||_2}{||\bm{v}||_2}=\frac{||\bm{v}_i-\bm{v}_{i-1}||_2}{||\bm{v}_{i}||_2}
-\end{equation}
-and
-\begin{equation}\label{eq:convergence_press}
-  \frac{||\Delta p||_2}{||p||_2}=\frac{||p_i-p_{i-1}||_2}{||p_{i}||_2}
-\end{equation}
-respectively. Iterations are performed until either maximal number of non-linear iterations ($it_{max}$) is reached or all the normalised nonlinear residual, 
-the velocity residual and the pressure residual converge under a defined tolerance $tol$.
-
-\section{The energy equation}\label{sec:en_eq}
-The stabilisation of the advection term of the energy equation (Eq. \ref{eq:energy}) needed to avoid possible oscillations in the thermal solution in the case
-when advection dominates over diffusion, is implemented by means of a streamline-upwind Petrov–Galerkin (SUPG) method, for which the advection term is modified
-follows the discussion in \citet{Thieulot2011} and \citet{Thieulot2014} (see Section \ref{sec:advection} for benchmark). The total internal heat production
-$H_{tot}$ in Eq. \ref{eq:energy} includes radiogenic heating $H_r$, shear heating $H_s$ and adiabatic heating $H_a$, where
-\begin{equation}\label{eq:shear}
-H_s=2\eta \dot{\bm{\epsilon}} : \dot{\bm{\epsilon}} = 2\eta(\dot{\epsilon}_{xx}^2+\dot{\epsilon}_{yy}^2+2\dot{\epsilon}_{xy}^2)
-\end{equation}
-and
-\begin{equation}\label{eq:adiabatic}
-H_a=T \alpha \frac{Dp}{Dt} \approx -\alpha T \rho g_y v
-\end{equation}
-The correctness of the implementation and the computation of shear and adiabatic heating is verified performing an exercise with an analytical solution
-(Section \ref{sec:simple_shear}) and Exercise 9.4 in \citet{Gerya2010b} (Section \ref{sec:shear}). Furthermore, mantle convection and thin layer entrainment
-benchmarks (Sections \ref{sec:mantle} and \ref{sec:thin}, respectively) are performed to verify that the code solve correctly Eqs. \ref{eq:momentum_penalty}
-and \ref{eq:energy} in case of a temperature field characterised by an initial perturbation and with a temperature-dependent density. At last, the viscoplastic
-mantle convection benchmark is performed to check also the viscoplasticity in case of a initial perturbed temperature field (Section \ref{sec:visco_mantle}).
-
-\section{Phase transitions and hydration}\label{sec:phase}
-Variations of effective density, specific heat and coefficient of thermal expansion during the evolution of crustal and mantle materials at different
-pressure-temperature ($p$-$T$) conditions are computed by means of version 6.8.6 of Perple\_X software package \citep{Connolly2005}, similarly to the
-implementation described in \citet{Marotta2020}. By default, mineral assemblages and properties of each lithology (oceanic crust, lower and upper continental
-crust, sediments and dry or hydrated mantle) are calculated for temperatures from \SIrange{330}{1600}{\kelvin}, with increments of approximately \SI{4}{\kelvin},
-and pressures from \SIrange{0.1}{30}{\giga\pascal}, with increments of approximately \SI{0.03}{\giga\pascal}, for a total of almost 100000 points.
-
-The introduction of phase transitions produces two main effects that must be taken into account in numerical models: variations in density and release/absorption
-of latent heat \citep{Tackley2010}. Effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into
-account considering effective coefficient of thermal expansion \citep{Christensen1985,Zhong2015}. Similarly, effects of latent heat in the energy equation
-(eq \ref{eq:energy}) are taken into account by considering effective specific heat and coefficient of thermal expansion
-\citep{Christensen1985,Tackley2010,Zhong2015}. Effective density, specific heat and coefficient of thermal expansion are included in Perple\_X files and are
-assumed by each Lagrangian marker in according to its $p$-$T$ conditions.
-
-Hydration processes related to dehydration of subducting lithosphere \citep{Schmidt1998,Liu2007,Faccenda2009,Faccenda2010,Faccenda2014,Rosas2016} strongly
-influence the thermo-mechanics inside the mantle wedge, mainly because of both weakening effects on the mantle rheology and density variations in case of
-mantle serpentinization \citep{Gerya2002,Honda2003,Arcay2005,Roda2010,Regorda2017}. In the case that hydration is switched on, the amount of bound and free
-water is memorised by each marker, following the implementation of \citet{Quinquis2014}; hydration and dehydration processes are related to the amount of bound
-water of each marker and to the maximum amount of water it can transport, i.e. if the amount of bound water exceeds the maximum amount of water, the marker
-dehydrates and releases free water that can hydrates under-saturated markers. Maximum water content of each marker is determined as function of lithology and
-$p$-$T$ conditions and it is calculated using Perple\_X, in the same way than effective density, specific heat and coefficient of thermal expansion. Bound water
-is advected together with the markers, neglecting the effect of bound water diffusion, while free water simply migrates vertically and is not coupled to the
-solid-phase flow of the mantle wedge \citep{Arcay2005,Quinquis2014}. The correctness of the migration scheme of the free water and its absorption into bound
-water by under-saturated markers are verified performing the experiment described in \citet{Quinquis2014} (Sec. \ref{sec:quinquis}).
-
-The mantle viscosity weakening related to the amount of water has been extensively studied \citep{Chopra1981,Hirth2003} and it is implemented in according to
-\citet{Arcay2005} and \citet{Horiuchi2016}, as follows
-\begin{equation}\label{eq:wet_visc}
-\eta_{wet}=\eta_{dry}\left(\left[1-\frac{1}{f_v}\right]\exp\left(-\frac{[OH^-]}{[OH^-]_0}\right)+\frac{1}{f_v}\right)
-\end{equation}
-where $f_v$ is the viscosity reduction factor between dry and wet conditions, $[OH^-]$ is the water content and $[OH^-]_0$ is a reference water content set to
-620 ppm (0.062 wt.\%) \citep{Arcay2005}. By default, $f_v$ is set to 100, in according to \citet{Arcay2005} and \citet{Horiuchi2016}.
-
-\section{Mantle melting}\label{sec:melt}
-Melt fraction in the mantle depends on temperature ($T$, in \textdegree C), pressure ($p$, in \si{\giga\pascal}) and the water content in the melt ($X_{H_2O}$, in wt.\%).
-The determination of the melt fraction $M$ is implemented as explained by \citet{Katz2003}, according to the successive modification by \citet{Langmuir2006} and
-\citet{Kelley2010}, as follows
-\begin{eqnarray}\label{eq:melt_fr}
-M(p,T,X_{H_2O})&=&-T+T_{sol}+(x_t\cdot ln(p+y_t))\cdot X_{H_2O}-\nonumber\\
-&&+K_m\left(\frac{X_{H_2O}^{bulk}}{D_{H_2O}(1-X_{H_2O})+X_{H_2O}}\right)^{\gamma}
-\end{eqnarray}
-where $T_{sol}$ is the temperature of the dry solidus, $D_{H_2O}$ is the partition coefficient and $x_t\cdot log(p+y_t)$ indicates the pressure dependence of
-the melting curve \citep{Kelley2010}. $x_t$, $y_t$, $K_m$ and $\gamma$ are parameters chosen in according to \citet{Kelley2010}.
-However, $X_{H_2O}$ depends on the melt fraction as
-\begin{equation}\label{eq:melt_wt}
-X_{H_2O}(M)=\frac{X_{H_2O}^{bulk}}{D_{H_2O}+M(1-D_{H_2O})}
-\end{equation}
-and a numerical solution can be found using a root-finder method \citep{Katz2003,Wang2016}. Furthermore, the content of water in the melt is limited by the
-pressure-dependent saturation concentration of water in the melt, determined as
-\begin{equation}\label{eq:melt_max}
-X_{H_2O}^{sat}=\chi_1 p^{\lambda_m}+\chi_2 p
-\end{equation}
-where $\chi_1$, $\chi_2$ and $\lambda_m$ are parameters chosen according to \citet{Katz2003}, \citet{Langmuir2006} and \citet{Kelley2010}. The correctness in
-the determination of melt fractions and water contents in the melt are tested comparing model results with experimental curves obtained by \citet{Katz2003}
-(Sec. \ref{sec:katz}).
-
-In case of melting, viscous creep viscosity ($\eta_{vp}$) is modified according to \citet{Wang2016}, as
-\begin{equation}\label{eq:melt_visc}
-\eta_{melt}=\eta_{vp}\cdot \exp(\alpha_m M)
-\end{equation}
-where $\alpha_m$ are melt fraction factors for dislocation and diffusion creep, chosen in according to \citet{Wang2016}.
-
-Effective density ($\rho_{eff}$) for partially molten rocks is calculated as
-\begin{equation}\label{eq:melt_dens}
-\rho_{eff}=\rho_s(1-M)+\rho_m M
-\end{equation}
-where $\rho_s$ and $\rho_m$ are the densities of the solid and the molten rock, respectively \citep{Gerya2004b,Gerya2007,Wang2016}. The density of the
-solid rock is extracted by Perple\_X, while the density of the molten portion is calculated as
-\begin{equation}\label{eq:melt_dens_M}
-\rho_m(p,T)=\rho_0[1-\alpha(T-T_0)][1+\beta(p-p_0)]
-\end{equation}
-where $\rho_0$ is the density at the reference temperature ($T_0$) and pressure($p_0$), and $\alpha$ and $\beta$ are the thermal expansion and compressibility
-coefficients, respectively \citep{Gerya2004b,Gerya2007,Wang2016}.
-
-As for the phase transitions, effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into account
-considering effective coefficient of thermal expansion. Similarly, latent heat due to melting/crystallisation can be implicitly included in the energy equation
-(Eq. \ref{eq:energy}) by considering effective specific heat ($C_{p_{eff}}$) and thermal expansion ($\alpha_{eff}$) of the partially molten rocks, in according to
-\begin{eqnarray}
-\label{eq:cp_effective}C_{p_{eff}}&=&C_p+H_L\left(\frac{\partial M}{\partial T}\right)_p \\
-\label{eq:alpha_effective}\alpha_{eff}&=&\alpha+\rho\frac{H_L}{T}\left(\frac{\partial M}{\partial p}\right)_T
-\end{eqnarray}
-where $C_p$ and $\alpha$ are the specific heat and the coefficient of thermal expansion of the solid rock, respectively, and $H_L$ is the latent heat of the
-molten rock \citep{Gerya2004b,Gerya2007,Tackley2010}.
-
-\section{Benchmarks}\label{sec:benchmark}
-\subsection{Error measurements}\label{sec:error}
-In order to determine the accuracy of velocity and pressure field of the benchmarks, the $L_2$-norm is computed by numerical integration on the Gauss-Legendre
-quadrature points. $L_2$-norm for pressure and velocity errors can be evaluated as
-\begin{eqnarray}
-\textrm{err}_p&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|p^n_i(r_q)-p^a_i(r_q)|^2\right)w_q|J_q|}\nonumber \\
-\textrm{err}_v&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|u^n_i(r_q)-u^a_i(r_q)|^2+|v^n_i(r_q)-v^a_i(r_q)|^2\right)w_q|J_q|}\nonumber
-\end{eqnarray}
-respectively, where $n_e$ is the number of elements, $n_q$ is the number of quadrature points per element, $p^n_i(r_q)$ and $p^a_i(r_q)$ are the numerical and
-analytical pressure, respectively, in each quadrature point $q$, $w_q$  and $J_q$ are the weight and the Jacobian at the quadrature point
-$q$, $u^n_i(r_q)$, $v^n_i(r_q)$, $u^a_i(r_q)$ and $v^a_i(r_q)$ are the numerical and analytical velocities, respectively, in each quadrature point $q$.
-
-Other quantities used as comparison with original benchmarks are the root-mean-square velocity over the whole domain and over the surface
-\begin{eqnarray}
-v_{\textrm{rms}}&=&\sqrt{\int_{\Omega}|\bm{v}|^2 d\Omega}\nonumber \\
-v_{\textrm{rms}}^{top}&=&\sqrt{\int_0^1 u^2 \bigg|_{y=1} dx}\nonumber
-\end{eqnarray}
-respectively, top and bottom Nusselt numbers
-\[\textrm{Nu}^{top/bottom}=-\int_0^1\frac{\partial T}{\partial y} \bigg|_{y=1/y=0} dx\]
-the average rate of work done against gravity
-\[\left\langle W \right\rangle=\int_{\Omega}T u_y d\Omega\]
-the average rate of viscous dissipation
-\[\left\langle \Phi \right\rangle=\int_{\Omega}\tau_{ij}\dot{\epsilon}_{ij} d\Omega\]
-and the percentage error between $\left\langle W \right\rangle$ and $\left\langle \Phi \right\rangle/Ra$
-\[\delta=\frac{\left|\left\langle W \right\rangle -\frac{\left\langle \Phi \right\rangle}{Ra} \right|}{\max \left(\left\langle W \right\rangle,
-\frac{\left\langle \Phi \right\rangle}{Ra}\right)}\times100\]
-
-\subsection{Stokes flow}\label{sec:stokes}
-The problem consists of determining velocity field $(u,v)$ and pressure $p$ in case of a manufactured solution with prescribed body forces such as
-\begin{eqnarray}
-b_1&=&(12-24y)x^4+(-24+48y)x^3+(-48y+72y^2-48y^3+12)x^2+\nonumber \\
-&&+(-2+24y-72y^2+48y^3)x+1-4y+12y^2-8y^3\nonumber \\
-b_2&=&(8-48y+48y^2)x^3+(-12+72y-72y^2)x^2+(4-24y+48y^2-48y^3+\nonumber \\
-&&+24y^4)x-12y^2+24y^3-12y^4\nonumber
-\end{eqnarray}
-for which the exact solution is:
-\begin{eqnarray}
-u(x,y)&=&x^2(1-x)^2(2y-6y^2+4y^3)\nonumber \\
-v(x,y)&=&-y^2(1-y)^2(2x-6x^2+4x^3)\nonumber \\
-p(x,y)&=&x(1-x)-\frac{1}{6}\nonumber
-\end{eqnarray}
-The domain is a unit square a constant viscosity ($\eta=1$) and the penalty parameter is set to $\lambda=10^7$. Velocity boundary conditions are set to no slip
-$(\bm{v}=\bm{0})$ on all boundaries. The problem is performed for different grid resolution between \num{8x8} and \num{1024x1024} elements.
-
-The errors between the analytical solution and the numerical prediction of the velocity and the pressure field are calculated as in Section \ref{sec:error}.
-Fig. \ref{fig:errors}a shows that both velocity and pressure field converge to the exact solution with the decrease of the element size, following the
-theoretical convergence rate. The convergence of pressure, in contrast with the observation by \citet{Donea2003} for $Q_1 \times P_0$ elements, indicate the
-effectiveness of the smoothing procedure. Smoothed pressure field for grid resolution of \num{128x128} elements is shown in Fig. \ref{fig:errors}b (continuous
-red line), in comparison with the analytical solution (dashed blue line). Solve times and memory usage needed to generate the solution are shown in
-Fig. \ref{fig:MUMPS}. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Solver/Stokes_Flow}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/errors.pdf}
-\caption{Velocity and pressure error for the Stokes flow experiment between generated and analytical solution as a function of element size (panel a),
-and comparison between smoothed pressure and analytical solution as function of $x$ coordinate for a grid resolution of \num{128x128} elements (panel b).}
-\label{fig:errors}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/MUMPS.pdf}
-\caption{Panel a) Analysis, factorisation, solution and total solve times (panel a) and factorisation memory usage (panel b) as a function of the total number
-of degrees of freedom for the Stokes flow experiment.}
-\label{fig:MUMPS}
-\end{figure}
-
-\subsection{Markers advection}\label{sec:runge}
-The 2nd-order Runge-Kutta advection scheme in space is tested by means of the Zalesak disk test \citep{Zalesak1979,Thieulot2014}. The benchmark is performed
-in a unit square domain with a grid resolution of \num{32x32} elements and 1000000 markers and values of Courant number between 0.25 and 1.
-At $t=0$, the disk is centred at position $(0.5;0.75)$ with a radius $R=0.15$ and has a vertical fissure 0.05 wide and 0.2 high (orange in Fig. \ref{fig:runge})
-and with the velocity field prescribed in the entire domain as
-\begin{eqnarray}
-u(x,y)&=&2\pi\left(y-\frac{L_x}{2}\right)\nonumber \\
-v(x,y)&=&-2\pi\left(x-\frac{L_x}{2}\right)\nonumber
-\end{eqnarray}
-the markers are back to their initial location after a $2\pi$ rotation ($t=1$). The clockwise rotation of the disk at $t=0.25$, 0.5 and 0.75 
-is shown in Fig. \ref{fig:runge} (red, blue and green, respectively). The distance from the centre of 1 marker is calculated throughout the rotation to
-evaluate the error for different values of Courant number. As expected, the error increase with the increasing of the time step, as shown in 
-Fig. \ref{fig:runge_err}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/Zalesak_Disk}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=200px]{./Figures/Zalesak.pdf}
-\caption{Position of the Zalesak disk at $t=0$, 0.25, 0.5 and 0.75 (orange, red, blue and green, respectively) for the case with Courant number of 1.}
-\label{fig:runge}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/zal_marker.pdf}
-\caption{Distance from the centre as function of time for values of Courant number of  of 0.25, 0.3, 0.5, 0.75 and 1 (green, orange, blue, black and red,
-respectively). R indicates the distance from the centre considering a perfect circle trajectory.}
-\label{fig:runge_err}
-\end{figure}
-
-\subsection{Conservative Velocity Interpolation (CVI)}\label{sec:CVI}
-The CVI correction is checked by means of the Stokes flow experiment by \citet{Donea2003}, as presented in \ref{sec:stokes}. The advection of Lagrangian
-markers is performed using either a 2nd-order or a 4th-order Runge-Kutta scheme with CFL=0.5 and an initial random distribution of 25 markers per element.
-Fig. \ref{fig:CVI} shows the minimum and maximum values of markers per element throughout the experiment with and without the CVI correction and using
-different Runge-Kutta schemes. The introduction of the CVI correction allows the
-tests to run for times longer than $t=2000$ without show empty elements, while tests without the CVI correction stop at approximately $t=600$, when at least
-one element goes empty. Fig. \ref{fig:CVI_time} shows the different number of markers per element and the markers distribution at $t=600$ obtained without and
-with the CVI correction (panels b and c, respectively) in case of a 2nd-order Runge-Kutta scheme. All data can be found at
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/CVI}.
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/CVI.pdf}
-\caption{Maximum and minimum number of markers per element throughout the experiment, with (green and orange lines) or without (purple and blue lines)
-the CVI correction and using either a 2nd-order (blue and orange lines) or a 4th-order (purple and green lines) Runge-Kutta scheme.}
-\label{fig:CVI}
-\end{figure}
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/CVI_time.pdf}
-\caption{Number of markers per element and markers distribution through time using 2nd.order Runge-Kutta scheme.
-Panels a: number of markers per element (panel a1) and distribution of the markers (panel a2) after the first
-iteration; panels b and c: comparison between number of markers per element (panels b1 and c1) and markers distribution (panels b2 and c2) without and with
-the CVI correction at $t=600$; panels d: number of markers per element (panel d1) and markers distribution (panel d2) in case of CVI correction at $t=2000$.}
-\label{fig:CVI_time}
-\end{figure}
-
-\subsection{Poiseuille flow}\label{sec:poiseuille}
-The domain is a rectangle with $L_x=2$ and $L_y=1$ and constant density and viscosity ($\rho=1$ and $\eta=1$), gravity acceleration $\bm{g}=\bm{0}$ and penalty
-parameter $\lambda=10^8$. The grid is composed by \num{40x20} elements. Velocity boundary conditions are set to no slip $(\bm{v}=\bm{0})$ at the top and the
-bottom, and a parabolic profile is imposed on the sides, with $u=y(1-y)$ and $v=0$. The analytical solution is then given by:
-\begin{eqnarray}
-u(x,y)&=&y(1-y)\nonumber \\
-v(x,y,)&=&0\nonumber \\
-p(x,y)&=&2\eta\left(\frac{L_x}{2}-x\right)\nonumber
-\end{eqnarray}
-
-The velocity field predicted by the model follows the expected parabolic profile (Fig. \ref{fig:poiseuille}a and Fig. \ref{fig:poi_plot}a) and, in case of the
-classic penalty method with no iterations, the pressure is clearly related to the divergence by means of the penalty parameter (Fig. \ref{fig:poiseuille}b and
-c and Fig. \ref{fig:poi_plot}b). Fig. \ref{fig:poiseuille}d shows that one Uzawa iteration is sufficient to bring the divergence down to \num{e-15}, with no
-correlation with the pressure field. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Poiseuille%20Flow}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Poiseuille.pdf}
-\caption{Velocity field (panel a), pressure (panel b) and divergence velocity of a Poiseuille flow in case of the classic penalty method (no iterations) and
-after one Uzawa iteration (panel c and d, respectively).}
-\label{fig:poiseuille}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Analytical.pdf}
-\caption{Velocity (panel a) and pressure (panel b) field predicted by the model for a Poiseuille flow with respect with their analytical solutions. The
-velocity field is plotted as function of the vertical coordinate in $L_x/2$ and the pressure is plotted as function of the x coordinate in $L_y/2$.}
-\label{fig:poi_plot}
-\end{figure}
-
-\subsection{Instantaneous 2D sphere}\label{sec:ist_sphere}
-The domain is a unit square with gravity $\bm{g}=(0,-1)$. The fluid has constant density and viscosity ($\rho_f=1$ and $\eta_f=1$). The sphere is in the middle
-of the domain with a radius R = 0.123456798 and has constant density and viscosity ($\rho_s=10^{-2}+\rho_f$ and $\eta_s=10^3 \cdot \eta_f$).
-We distinguish three different types of velocity boundary conditions:
-\begin{enumerate}
-\item FS: free slip conditions on all sides.
-\item NS: no slip conditions on all sides.
-\item OT: free slip conditions on the sides and the bottom, and open boundary at the top.
-\end{enumerate}
-All three cases are performed for different grid resolutions, between \num{16x16} and \num{512x512} elements with 50 randomly distributed markers per element,
-and different average schemes for the viscosity (harmonic, geometric and arithmetic). The velocity in the centre of the sphere $\bm{v}(0.5;0.5)$), minimum and
-maximum velocities ($u_{min}$, $u_{max}$, $v_{min}$, $v_{max}$) and pressures ($p_{min}$, $p_{max}$) on the entire domain, $v_{\textrm{rms}}$ and average pressure on the
-entire domain ($p_{avg}$) are compared with solutions generated by ASPECT \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and all results
-can be found at \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere2D} and 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Instantaneous_2D_Sphere}. Results in terms of the $v_{\textrm{rms}}$ for NS model 
-are also shown in Fig. \ref{fig:inst_sphere}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_NS.pdf}
-\caption{$v_{\textrm{rms}}$ as function of element size for the instantaneous 2D sphere experiment in case of no slip boundary conditions and with different average
-schemes for the viscosity. Results are compared with results obtained by other numerical codes.}
-\label{fig:inst_sphere}
-\end{figure}
-
-\subsection{Rayleigh-Taylor instability}\label{sec:rayleigh}
-This problem has been originally presented by \citet{vanKeken1997} and here it is performed as the isoviscous case (Case 1a). The domain has $L_x=0.9142$ and
-$L_y=1$ and gravity $\bm{g}=(0,-1)$. Two fluids with same constant viscosities ($\eta_1=\eta_2=1$) and different densities ($\rho_1=1000$ and $\rho_2=1010$),
-with the lighter fluid at the bottom. The initial interface between the fluids is given by $y(x)=0.2+0.02 \cos \left(\frac{\pi x}{L_x}\right)$. The experiment
-is performed with different grid sizes (\num{50x50}, \num{80x80}, \num{100x100} and \num{256x256}). A total of 1960000 markers are randomly distributed at the
-beginning of the simulation. Velocity boundary conditions are set to no slip at the top and the bottom, and to free slip at the sides of the domain.
-
-$v_{\textrm{rms}}$ as function of time is reported in Fig. \ref{fig:RT}, matching well with results shown by \citet{vanKeken1997}, \citet{Tackley2003} and
-\citet{Thieulot2014}. Fig. \ref{fig:rayleigh} shows the evolution of the experiment at different time steps. All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Rayleigh_Taylor_experiment/ISOVISCOUS}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/RT.pdf}
-\caption{$v_{\textrm{rms}}$ of the Rayleigh-Taylor experiment as function of time for different resolution of the grid.}
-\label{fig:RT}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=300px]{./Figures/Rayleigh.pdf}
-\caption{Evolution of the Rayleigh-Taylor experiment for a grid of \num{256x256} elements at $t=0, 500, 1000$ and $2000$ (panels a, b, c and d, respectively).}
-\label{fig:rayleigh}
-\end{figure}
-
-\subsection{Falling blocks}\label{sec:block}
-This benchmark is described as proposed by \citet{Gerya2003a}, \citet{Gerya2010b} and \citet{Thieulot2011}. The domain is square with $L_x=L_y=$ \SI{500}{\km}
-and the grid is composed by \num{50x50} elements with 25 markers in each element. The block is initially centred at ($x=$ \SI{250}{\km}; $y=$ \SI{400}{\km})
-and has a size of \num{100x100} km (Fig. \ref{fig:block}a). The fluid surrounding the block has $\eta_f=$ \SI{e21}{\pascal\per\s} and $\rho_f=$
-\SI{3200}{\kg\per\cubic\m}. The benchmark tests with different viscosities of the block, with $\eta_b$ from \SIrange{e15}{e27}{\pascal\per\s}. In each
-experiment also the density $\rho_b$ of the block varies from \SIrange{3220}{9900}{\kg\per\cubic\m}. Velocity boundary conditions are set to free slip
-conditions on all sides of the domain. Density distributions at $t=$ \SI{20}{\mega\year} for $\rho_b=$ \SI{3300}{\kg\per\cubic\m} and different $\eta_b$
-are plotted in Fig. \ref{fig:block}b-f, showing that the code correctly preserve the block geometry in case of large viscosity contrast, with a block stiffer
-than the surrounding fluid (Fig. \ref{fig:block}f).
-
-The velocity in the centre of the falling block at $t=0$ is measured for all experiments and, since the velocity should increase with density contrast, the
-quantity $\bm{v}/(\rho_b-\rho_f)$ is plotted as function of the viscosity contrast. All results perfectly match with those from \citet{Gerya2010b} and they
-line up on a single curve, demonstrating that the code can correctly deal with large viscosity and density contrasts (Fig. \ref{fig:falling}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Falling%20blocks}.
-
-\begin{figure}
-\noindent\includegraphics[width=300px]{./Figures/Block.pdf}
-\centering
-\caption{Density evolution of the falling block experiment at $t=$ \SI{0}{\mega\year} (panel a) and $t=$ \SI{20}{\mega\year} for different viscosities of the
-block (panels b-f).}
-\label{fig:block}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Falling.pdf}
-\caption{Initial velocity relative to the density contrast at the centre of the falling block as function of the viscosity contrast between the block and the
-surrounding fluid.}
-\label{fig:falling}
-\end{figure}
-
-\subsection{2D time-dependent Stokes sphere with deformable free surface}\label{sec:sphere}
-This experiment is performed in a unit square domain with the gravity acceleration fixed to $g_y=-1$. The sphere has $\rho_s=2$ and $\eta_s=10^3$ and is
-initially centred at ($0.5;0.6$) with radius $R=0.123456789$. The fluid surrounding the sphere has $\rho_f=1$ and $\eta_f=1$ and occupies the domain for
-$y\leq 0.75$, while for $y>0.75$ the air has $\rho_a=0$ and $\eta_a=10^{-3}$. Velocity boundary conditions are set to free slip on all sides. The Courant
-number is set to 0.25. The experiments run for 200 s using grid resolutions from \num{150x150} to \num{512x512} elements, with 25 randomly distributed markers
-per element, and for different average schemes for the viscosity. The interface between the fluid and the air is tracked by means of the markers chain. Results
-in terms of velocity and pressure fields and topography variations are compared with results obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and they can be found at
-\url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere_fs2D} and
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Time_dependent_sphere}. Results in terms of the $v_{\textrm{rms}}$ in case of 
-an arithmetic average are also shown in Fig. \ref{fig:inst_spherefs}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_arithm.pdf}
-\caption{$v_{\textrm{rms}}$ for different grid resolutions as function of time for the 2D time-dependent Stokes sphere experiment in case of an arithmetic average.
-Results are compared with results obtained by ASPECT.}
-\label{fig:inst_spherefs}
-\end{figure}
-
-\subsection{Free surface stabilisation}\label{sec:stab}
-The benchmark is performed as discussed in \citet{Kaus2010a} and \citet{Thieulot2014}. The domain is square with $L_x=L_y=$ \SI{500}{\km} and a grid resolution
-of \num{200x200} elements, each of them containing 16 markers. A buoyant fluid with $\rho_1=$ \SI{3200}{\kg\per\cubic\metre} and $\eta_1=$ \SI{e20}{\pascal\s}
-is overlain by a denser fluid with $\rho_2=$ \SI{3300}{\kg\per\cubic\metre} and $\eta_2=$ \SI{e21}{\pascal\s}. The initial interface between the fluids has an
-initial sinusoidal shape of 5 km amplitude. Velocity boundary conditions are set to free slip conditions at the sides and no slip at the bottom, while the top
-is free surface. The experiment is performed with various fixed time steps, without the use of the Courant number. The vertical position of the free surface at
-$x=L_x$ is tracked for each simulation, with or without the stabilisation algorithm.
-
-The results show an instability (drunken sailor effect) increasing the time step in case that the stabilisation algorithm is not activated
-($dt=$ \SI{4200}{\year}, green line in Fig. \ref{fig:stab}), as already observed by \citet{Kaus2010a} and \citet{Thieulot2014}. Activating the algorithm the
-instability is fixed and the simulation remains stable using time step up to \SI{50000}{\year} (red line in Fig. \ref{fig:stab}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Stabilisation_algorithm}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Kaus_random.pdf}
-\caption{Evolution of the y coordinate in $x=L_x$ as function of time for different time steps, with and without the stabilisation algorithm.}
-\label{fig:stab}
-\end{figure}
-
 \subsection{Topography relaxation}\label{sec:crameri}
 This experiment is performed as in \citet{Crameri2012} using free surface and sticky air (Case 1). The domain is rectangular with $L_x=$ \SI{2800}{\km} and
-$L_y=$ \SI{700}{\km} in case of free surface and $L_y=$ \SI{800}{\km} in case of sticky air ($h_{st}=$ \SI{100}{\km}). The grid is composed by \num{256x64}
-elements for the free surface \num{560x320} elements for the sticky air. For both cases, there is a mantle of 600 km thickness with $\eta_m=$ \SI{e21}{\pascal\s}
+$L_y=$ \SI{700}{\km} in case of free surface and $L_y=$ \SI{800}{\km} in case of sticky air ($h_{st}=$ \SI{100}{\km}). The grid is composed by $256\times64$
+elements for the free surface $560\times320$ elements for the sticky air. For both cases, there is a mantle of 600 km thickness with $\eta_m=$ \SI{e21}{\pascal\s}
 overlain by a highly viscous cosine-shaped lithosphere with thickness between 93 and 107 km and $\eta_l=$ \SI{e23}{\pascal\s}. Both mediums have $\rho=$
@@
 \SI{2570}{\m}, with an error of \SI{6}{\m} with respect to the analytical solution. Results are also compared with results obtained with a free surface in
-UNDERWORLD (\num{256x64} elements) and SULEC (\num{401x201} elements) (with courtesy from \citealp{Crameri2012}). All data can be 
+UNDERWORLD $256\times64$ elements) and SULEC ($401\times201$ elements) (with courtesy from \citealp{Crameri2012}). All data can be 
 found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Topography_relaxation}.
@@
 \end{figure}
-
-\subsection{Spontaneous subduction}\label{sec:subduction}
-This experiment is performed as presented by \citet{Schmeling2008}, in case of both a sticky air and a true free surface and using different average schemes
-for the viscosities. The domain is rectangular with $L_x=$ \SI{3000}{\km} and $L_y=$ \SI{700}{\km}, a grid resolution of \num{256x64} elements with 50 markers
-per element and Courant number of 0.01. At the beginning of the simulation, a 100 km-thick lithospheric layer with $\rho_m=$ \SI{3300}{\kg\per\cubic\metre}
-and $\eta_m=$ \SI{e23}{\pascal\s} is located at the top of the domain for $x$ between 1000 and 3000 km, while the asthenospheric mantle has $\rho_m=$
-\SI{3200}{\pascal\s} and $\eta_m=$ \SI{e21}{\pascal\s}. In addition, a 200 km-depth lithospheric slab is already subducted in the mantle in order to have
-a spontaneous subduction. In case of sticky air, $L_y=750$ km and a 50 km-thick air layer with $\eta_a$ from \SIrange{e19}{e21}{\pascal\s} overlie the mantle.
-Velocity boundary conditions are set to free slip on the sides and on the bottom of the domain. In case of sticky air, velocities on the top boundary are set
-to free slip conditions as well.
-
-The maximum depth of the slab is tracked for all the simulations (Fig. \ref{fig:subduction}) and they are compared with results obtained by
-\citet{Schmeling2008} with models with regular grids and comparable grid resolutions (rectangles in Fig. \ref{fig:subduction}). Low viscous sticky air models
-(i.e., \SI{e19}{\pascal\s}) enlighten a strong dependence of the sinking velocity with the chosen average scheme (dashed lines in Fig. \ref{fig:subduction}),
-while high viscous sticky air models (i.e., \SI{e21}{\pascal\s}) show a higher resistance at the trench, underestimating the correct solution (dotted lines in
-Fig. \ref{fig:subduction}). A high resistance at the trench can be also observed in case of a true free surface for the low resolution grid used (continuous
-lines in Fig. \ref{fig:subduction}), as pointed out by \citet{Schmeling2008}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Spontaneous_subduction}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Subduction.pdf}
-\caption{Maximum depth of the slab as function of time for all the simulations of the \citet{Schmeling2008} benchmark, in case of harmonic, geometric and
-arithmetic means (red, green and blue lines, respectively) using a sticky air or a true free surface (discontinuous and continuous lines, respectively).
-Red, green and blue rectangular areas indicate the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{500}{\km} in case of sticky air,
-while the grey rectangular area indicates the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{600}{\km} in case of true free surface.}
-\label{fig:subduction}
-\end{figure}
-
-\subsection{The slab detachment}\label{sec:slab}
-The slab detachment benchmark is performed as described by \citet{Schmalholz2011} and \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{1000}{\km}
-and $L_y=$ \SI{660}{\km} and a grid resolution of \num{256x256} elements. A non-linear viscous T-shaped layer with $\rho_l=$ \SI{3300}{\kg\per\cubic\metre} is
-placed at the top of the domain and surrounded by a linear viscous fluid with $\rho_f=$ \SI{3150}{\kg\per\cubic\m}. The top layer is 80 km-thick and a 250
-km-long and 80 km-wide slab is placed at $x=L_x/2$ (Fig. \ref{fig:slab}a). The effective viscosity of the top layer is given by
-\[\eta_{eff}=\eta_0 I_2^{\frac{(1-n)}{n}}\]
-with $\eta_0=$ \SI{4.75e11}{\pascal\s} and $n=4$, while the surrounding fluid has $\eta_f=$ \SI{e21}{\pascal\s}. The effective viscosity is capped between
-$\eta_{min}=$ \SI{e21}{\pascal\s} and $\eta_{max}=$ \SI{e25}{\pascal\s}. Velocity boundary conditions are set to free slip at the top and the bottom, and to
-no slip at the sides of the domain.
-
-The sinking of the slab determines a decrease of the effective viscosity in the T-shaped layer during the first part of the evolution (Fig. \ref{fig:slab}a),
-as consequence of local strain rates, while effective viscosities increase up to $\eta_{max}$ in the top layer after approximately \SI{20}{\mega\year}, when
-necking is complete (Fig. \ref{fig:slab}b).
-
-The width of necking is tracked by means of markers position at the side of the slab. Necking width and time are normalised in function of the initial slab
-(\SI{80}{\km}) and the characteristic time ($t_c=$ \SI{7.1158e14}{\s}) \citep{Schmalholz2011,Glerum2018}. The results in case of arithmetic, geometric and
-harmonic mean of the viscosity are shown in Fig. \ref{fig:necking}, compared with results from \citet{Glerum2018} for which an infinity norm mean was used.
-The influence of the average schemes agrees with results shown in Section \ref{sec:subduction}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Slab_detachment}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Slab.pdf}
-\caption{Effective viscosity for the slab detachment benchmark at the beginning of the evolution (panel a) and when necking is complete (panel b).}
-\label{fig:slab}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Necking.pdf}
-\caption{Normalised width of the necking of the slab detachment benchmark as function of normalised time.}
-\label{fig:necking}
-\end{figure}
-
-\subsection{The indenter experiment}\label{sec:indenter}
-The indenter benchmark simulate a rigid punch on a purely plastic von Mises material, for which exists an analytical solution
-\citep{Thieulot2008,Thieulot2014,Glerum2018}. The domain is a unit square with a grid resolution of \num{256x256} elements. The gravity acceleration is set
-to $\bm{g}=\bm{0}$ and effective viscosity is capped using $\eta_{min}=10^{-4}$ and $\eta_{max}=10^3$. The tolerance for the convergence of the non-linear
-solution is $tol=10^{-9}$, with a maximum number of non-linear iterations set to 500. The medium in the domain has $\rho=0.01$, an initial viscosity
-$\eta_0=10$, cohesion $C=1$ and an angle of internal friction $\phi=0$°. Velocity boundary conditions are set to no slip at the bottom and free slip at the
-sides of the domain. The top of the domain is open with the exception of the central portion (punch area) with width $w_p=0.125$, where $v=-1.05$ and $u$ is
-fixed either to 0 (rough punch experiment) or free (smooth punch experiment).
-
-Fig. \ref{fig:indenter} show the results in terms of viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure
-(panels g and h) for both experiments. The analytical solution indicate that pressure at ($0.5;1$) and ($0.5 \pm w_p;1$)is $p=\pi +1$ and $p=1$, respectively
-(black and red lines in Fig. \ref{fig:smooth_p}). Fig. \ref{fig:smooth_p} shows a clear improvement in the pressure solution when passing from a rough to a
-smooth punch (green and blue lines in Fig. \ref{fig:smooth_p}, respectively), as previously observed by \citet{Thieulot2014} and \citet{Glerum2018}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Indenter}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Indenter.pdf}
-\caption{Viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure (panels g and h) fields for rough (left column) and
-smooth (right column) punch experiments.}
-\label{fig:indenter}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Smooth_Rough.pdf}
-\caption{Nodal pressure in function of the x coordinate at $y=1$ for rough (green line) and smooth (blue line) punch experiment. Black and red lines indicate
-the analytical solution at $x=0.5$ and $x=0.5 \pm w_p$, respectively.}
-\label{fig:smooth_p}
-\end{figure}
-
-\subsection{The brick experiment}\label{sec:brick}
-An instantaneous version of the brick benchmark is performed to verify the correctness of the pressure-dependent plasticity for different angle of internal
-friction, as proposed by \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{40}{\km} and $L_y=$ \SI{10}{\km} and a grid resolution of \num{512x128}
-elements. The gravity acceleration is set to $g_y=$ \SI{-10}{\m\square\s} and effective viscosity is capped using $\eta_{min}=$ \SI{e19}{\pascal\s} and
-$\eta_{max}=$ \SI{e26}{\pascal\s}. The tolerance for the convergence of the non-linear solution is $tol=10^{-7}$, with a maximum number of non-linear iterations
-set to 1000. A 800 m-wide and 400 m-height inclusion with $\eta_b=$ \SI{e20}{\pascal\s} and $\rho_b=$ \SI{2700}{\kg\per\cubic\m} is placed at the bottom of the
-domain at $x=L_x/2$. The inclusion is surrounded by a non-linear viscous medium with $\rho_m=$ \SI{2700}{\kg\per\cubic\m}, an initial viscosity of
-\SI{e23}{\pascal\s}, a linear viscous viscosity of \SI{e25}{\pascal\s} and a cohesion of \SI{40}{\mega\pascal}. Velocities are set to free slip conditions at
-the bottom of the domain and the top is open. Velocities on sides of the domain are fixed to $u=$ \SI{+-2e-11}{\m\per\s} and $v=0$. The experiment is performed
-in compressional and extensional contexts with an angle of internal friction $\phi$ of the non-linear medium variable between 0° and 30°.
-
-As expected, two shear bands stem from the inclusion with variable angles in relation to both the dynamics context and the internal friction angle. Shear band
-angles formed at 45° for $\phi=0$° in both extensional and compressional contexts (Fig. \ref{fig:brick_beam}a), while they are different in case of $\phi\neq0$°.
-Shear band angles in case of $\phi=20$° are shown for compression and extension in Fig. \ref{fig:brick_beam}b  and c, respectively. Values of shear band angles
-as function of internal friction angles from 0° to 30° are extracted at different depths and minimum and maximum values are plotted in Fig. \ref{fig:brick},
-in comparison with theoretical Roscoe, Arthur and Coulomb shear band angles. For all tests the tolerance for nonlinear convergence is defined $tol=10^{-7}$
-and none converges before the maximum number of iterations ($it_{max}=1000$) is reached. However, tests with internal friction angles up to 15° show a constant
-decrease of the velocity residuals, which can not be observed for for tests with higher internal friction angles (Fig. \ref{fig:convergence}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Brick_experiment}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Brick_Beam.pdf}
-\caption{Shear band angles predicted for the brick experiment for $\phi=0$° (panel a) and for $\phi=20$° in case of compression and extension (panels b and c,
-respectively).}
-\label{fig:brick_beam}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Brick.pdf}
-\caption{Shear band angles predicted for the brick experiment in case of compressional and extensional contexts as function of different internal angle of
-friction (continuous black line and red dots), compared with theoretical Roscoe, Artur and Coulomb angles (discontinuous black lines). Red lines indicate the
-range of angles calculated at different depths.}
-\label{fig:brick}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \noindent\includegraphics[width=500px]{./Figures/convergence.pdf}
-  \caption{Convergence rates for all tests. Normalised residuals (left column) are compared with velocity residuals (right column) for extensional (first row)
-  and compressional (second row) settings.}
-  \label{fig:convergence}
-  \end{figure}
-
-\subsection{Advection stabilisation}\label{sec:advection}
-The 1D advection problem proposed by \citet{Donea2003} and \citet{Thieulot2011} is performed to verify the effectiveness of the SUPG method to stabilise the
-advection term of the energy equation. The domain is a 1D segment with $L_x=1$ composed by 50 elements and a discontinuity in the thermal field placed at
-$x=0.25$. Temperature is set to 1 for $x \leq 0.25$ and to 0 for $x > 0.25$. Velocity is set to ${u}=1$ in the entire domain. The simulation is performed for
-250 time steps, with $dt=0.002$, so the thermal discontinuity should be at $x=0.75$ at the end of the simulation.
-
-Temperature profiles at the end of the simulation are shown in \ref{fig:advection} as function of the dimensionless coefficient $\gamma=\tau \bm{v}/h$. In case
-of the classic Galerkin method ($\gamma=0$, blue line in \ref{fig:advection}) the final thermal profile is characterised by strong oscillations, which are
-eliminated in case of the SUPG method ($\gamma=0.045$, orange line in \ref{fig:advection}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Advection_stabilisation}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Advection.pdf}
-\caption{Temperature profile as function of \textit{x} for the advection stabilisation benchmark. Purple line indicates the initial temperature profile; the
-green line indicates the analytical temperature profile after 250 time steps; blue line indicates the temperature profile after 250 time steps in case of the
-classic Galerkin method ($\gamma=0$); orange line indicates the temperature profile after 250 time steps in case of the SUPG method ($\gamma=0.045$).}
-\label{fig:advection}
-\end{figure}
-
-\subsection{Shear heating}\label{sec:simple_shear}
-This exercise is performed in an unit square domain composed by \num{128x128} elements. The velocity field is prescribed on the entire domain with
-$\bm{v}=(L_y-y)y\bm{e_x}$; viscosity, density and specific heat are fixed to 1, while thermal conductivity and radiogenic energy are fixed to 0. Therefore,
-the energy equation (Eq. \ref{eq:energy}) can be simplified as \[\frac{\partial T}{\partial t}=H_s\] and fixing $T(t=0)=0$, the temperature can be find as
-\[T(t)=H_s t\]
-In this case we have
-\begin{eqnarray}
-\dot{\epsilon_{xx}}&=&\frac{\partial u}{\partial x}=0 \nonumber \\
-\dot{\epsilon_{yy}}&=&\frac{\partial v}{\partial y}=0 \nonumber \\
-\dot{\epsilon_{xy}}&=&\frac{1}{2}\left(\frac{\partial v}{\partial x}+\frac{\partial u}{\partial y}\right)=\frac{1}{2}(L_y-2y)\nonumber
-\end{eqnarray}
-and, simplifying Eq. \ref{eq:shear}, shear heating can be calculated as \[H_s(x,y)=(1-2y)^2\]
-
-The solution predicted by the model in terms of velocity, temperature and shear heating match well with the analytical solutions (Fig. \ref{fig:analytical_en}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Simple_shear}.
-
-%\begin{figure}
-%\centering
-%\noindent\includegraphics[width=400px]{./Figures/Energy_Thieulot.pdf}
-%\caption{Temperature and shear heating predicted by the model after 1 time step.}
-%\label{fig:en_thieulot}
-%\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=500px]{./Figures/Analytical_Shear.pdf}
-\caption{Velocity (panel a), temperature (panel b) and shear heating (panel c) as function of the \textit{y} coordinate for the simple shear experiment. The
-solutions predicted by the model (red lines) are compared with the analytical solutions (dashed blue lines).}
-\label{fig:analytical_en}
-\end{figure}
-
-\subsection{Shear and adiabatic heating}\label{sec:shear}
-This problem is performed as presented in Exercise 9.4 in \citet{Gerya2010b}. Two materials are vertically separated in a rectangular domain with $L_x=$
-\SI{1000}{\km}, $L_y=$ \SI{1500}{\km} and a grid resolution of \num{30x20} elements. Constant thermal coefficient expansion ($\alpha=$ \SI{3e-5}{\per\kelvin}),
-temperature ($T=$ \SI{1300}{\kelvin}) and gravity acceleration ($g_y=$ \SI{-10}{\m\square\s}) are assumed in the whole domain. Fluid 1 (on the left side of the
-domain) has $\rho_1=$ \SI{3200}{\kg\per\cubic\m} and $\eta_1=$ \SI{e20}{\pascal\s}; fluid 2 (on the right side of the domain) has $\rho_2=$
-\SI{3300}{\kg\per\cubic\m} and $\eta_2=$ \SI{e22}{\pascal\s}. Velocity boundary conditions are set to free slip on all sides of the domain.
-
-As shown in Fig. \ref{fig:energy}, both shear and adiabatic heating predicted by the code (first row in Fig. \ref{fig:energy}) well recreate the results
-obtained by means of example ${Shear\_adiabatic\_heating.m}$ from \citet{Gerya2010b} (second row in Fig. \ref{fig:energy}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Adiabatic%2BShear}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Energy.pdf}
-\caption{Comparison between shear (first column), adiabatic (second column) and total (third column) energy predicted by the code (panels a, b and c) and those
-created using example ${Shear\_adiabatic\_heating.m}$ from Exercise 9.4 in \citet{Gerya2010b} (panels d, e and f).}
-\label{fig:energy}
-\end{figure}
-
-\subsection{Mantle convection}\label{sec:mantle}
-This problem is performed as presented by \citet{Blankenbach1989} (constant viscosity cases) and \citet{Thieulot2014}, in a 2D unit square domain with gravity
-acceleration $g_y=10^{10}Ra$. The experiment is performed with three different Rayleigh numbers ($Ra=10^4$, $10^5$ and $10^6$) and with different grid
-resolution (between \num{32x32} and \num{128x128} elements). The fluid has constant viscosity, initial density, heat capacity, thermal conductivity
-($\eta= \rho_0= C_p = k=1$), reference temperature ($T_0=0$) and thermal expansion coefficient ($\alpha=10^{-10}$). Temperatures are set to 0 on top and 1 on
-bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\]
-
-The solution generated by the code in terms of $v_{\textrm{rms}}$ (calculated as in Sec. \ref{sec:error}) and Nu as function of time are reported for all the
-simulations in Fig. \ref{fig:mantle}. At the steady state, $v_{\textrm{rms}}$ and Nu of all simulations converge well toward the values from \citet{Blankenbach1989},
-with lower errors for higher resolution grids (Table \ref{tab:mantle}). All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Mantle_convection}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Convection.pdf}
-\caption{$v_{\textrm{rms}}$ (panels a, c and e) and Nu (panels b, d and f) for the mantle convection benchmark as function of time for different grid resolution.
-Panels a and b show the results for $Ra=10^4$; Panels c and d show the results for $Ra=10^5$; panels e and f show the results for $Ra=10^6$. Purple lines
-indicate the convergence values for the $v_{\textrm{rms}}$ and the Nu from \citet{Blankenbach1989} at the steady state.}
-\label{fig:mantle}
-\end{figure}
-
-\begin{table}[H]
-\caption{Comparison between $v_{\textrm{rms}}$ and Nu predicted by the code for the mantle convection experiment and same values as reported in literature.}
-\centering
-\small
-\begin{tabular}{|>{\centering}m{0.09\textwidth}|>{\centering}m{0.05\textwidth}|>{\centering}m{0.24\textwidth}|>{\centering}m{0.15\textwidth}|
-  >{\centering}m{0.1\textwidth} >{\centering}m{0.1\textwidth} >{\centering\arraybackslash}m{0.1\textwidth}|}
-\toprule
-& & \multirow{2}{*}{\citet{Blankenbach1989}} & \citet{Thieulot2014} & \multicolumn{3}{c|}{FALCON}  \\
- & & & \num{200x200} & \num{32x32} & \num{64x64} & \num{128x128}  \\
-\midrule
-\multirow{2}{*}{$Ra=10^4$} & $v_{\textrm{rms}}$ & $42.864947 \pm 0.000020$ & 42.867 & 42.83226 & 42.852793 & 42.861394  \\
-           & Nu      & $4.884409 \pm 0.000010$  & 4.882  & 4.781297 & 4.857475  & 4.877573   \\
-\hline
-\multirow{2}{*}{$Ra=10^5$} & $v_{\textrm{rms}}$ & $193.21454 \pm 0.00010$  & 193.255& 193.872643 & 193.377472 & 193.252290  \\
-           & Nu      & $10.534095 \pm 0.000010$ & 10.507 & 9.602514 & 10.270735  & 10.465629   \\
-\hline
-\multirow{2}{*}{$Ra=10^6$} & $v_{\textrm{rms}}$ & $833.98977 \pm 0.00020$  & 834.712& 848.091176 & 837.767911 & 834.945793  \\
-           & Nu      & $21.972465 \pm 0.000020$ & 21.695 & 15.999266 & 19.703682  & 21.306939   \\
-\bottomrule
-\end{tabular}
-\label{tab:mantle}
-\end{table}
-
-\subsection{Viscoplastic mantle convection}\label{sec:visco_mantle}
-This benchmark is performed as Cases 1-5a described in \citet{Tosi2015} in a 2D unit square domain with two different grid resolutions of \num{32x32} 
-and \num{100x100} elements. Reference viscosity, reference density, heat capacity, thermal conductivity and thermal expansion coefficient are fixed to a constant
-value of 1 ($\eta_0= \rho_0= C_p = k=\alpha=1$), while the gravity has been chosen as $g_y=10^2$, in order to obtain a Rayleigh number $Ra=10^2$. Temperatures
-are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\].
-The viscosity field $\eta$ is calculated as the harmonic average between a linear part $\eta_{lin}$ that depends on temperature only or on temperature and 
-depth $y$ and a nonlinear-plastic part $\eta_{pl}$ that depends on the strain rate $\dot{\bm{\epsilon}}$, as follows
-\[\eta(T,y,\dot{\bm{\epsilon}})=2\left(\frac{1}{\eta_{lin}(T,y)}+\frac{1}{\eta_{pl}(\dot{\bm{\epsilon}})}\right)^{-1}\]
-The linear and the nonlinear-plastic parts of the viscosity are calculated following \citet{Tosi2015} as follows
-\[\eta_{lin}(T,y)=\exp(-\ln(\Delta\eta_T)T+\ln(\Delta\eta_y)y)\]
-\[\eta_{pl}(\dot{\bm{\epsilon}})=\eta^*+\frac{\sigma_Y}{\sqrt{\dot{\bm{\epsilon}}:\dot{\bm{\epsilon}}}}\]
-where $\Delta\eta_T$, $\Delta\eta_y$, $\eta^*$ and $\sigma_Y$ are parameters chosen as listed in Table \ref{tab:tosi_setup}.
-
-For each case, some diagnostic quantities measured at the steady state are reported in Table \ref{tab:visco_mantle} (see Section \ref{sec:error} for 
-the computation of all the quantities), showing a well fit with results described in \citet{Tosi2015}. Data and charts for all cases can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Viscoplastic_mantle_convection}.
-
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{white!100!}{gray!25!}
-%\begin{table}[H]
-%  \caption{Results of the viscoplastic mantle convection benchmark for Cases 1-4 compared with results obtained with ELEFANT and YACC (from \citealp{Tosi2015}).}
-%  \centering
-%  \resizebox{\textwidth}{!}{
-%\begin{tabular}{|c|ccc|ccc|ccc|ccc|}
-%  \toprule
-%    & \multicolumn{3}{c|}{\textbf{Case 1}} & \multicolumn{3}{c|}{\textbf{Case 2}} & \multicolumn{3}{c|}{\textbf{Case 3}} &
-%    \multicolumn{3}{c|}{\textbf{Case 4}} \\
-%   & FALCON & ELEFANT& YACC & FALCON& ELEFANT& YACC & FALCON & ELEFANT & YACC & FALCON & ELEFANT & YACC \\
-%  \midrule
-%   $\left\langle T \right\rangle$ & 0.7759 & 0.7758 & 0.7767 & 0.6043 & 0.6036 & 0.6058 & 0.7275 & 0.7275 & 0.7286 & & 0.5277 & 0.5289\\
-%    Nu$^{top}$ & 3.4280 & 3.4214 & 3.4298 & 8.5242 & 8.5115 & 8.5278 & 3.0368 & 3.0347 & 3.0374 && 6.5912 & 6.5572 \\
-%    Nu$^{bottom}$ & 3.3168 & 3.313 & 3.3143 & 8.4599 & 8.4465 & 8.3990 & 3.0086 & 2.9908 & 2.9628 && 6.5834 & 6.5243 \\
-%    $v_{\textrm{rms}}$ & 249.901 & 249.134 & 251.7997 & 141.2451 & 140.7087 & 142.2020 & 100.2776 & 100.1208 & 100.9467 && 79.1105 & 79.6202 \\
-%    $v_{\textrm{rms}}^{top}$ & 1.8907 & 1.8642 & 1.8298 & 104.7928 & 104.1228 & 105.1767 & 2.0832 & 2.0652 & 2.0374 && 74.7596 & 75.8414 \\
-%    $u_{max}^{top}$ & 2.6340 & 2.6119 & 2.5516 & 121.9212& 121.766 & 122.3238 & 2.9086 & 2.9019 & 2.8458 && 88.9146 & 89.2940 \\
-%    $\eta_{min}$ & - & - & - & $2.0303\times10^{-5}$ & $2.0600\times10^{-5}$ & $1.9853\times10^{-5}$ & $4.8129\times10^{-5}$ & $4.8080\times10^{-5}$ & $4.7907\times10^{-5}$ &
-%    & $1.9860\times10^{-4}$ & $1.9174\times10^{-4}$ \\
-%    $\eta_{max}$ & - & - & - & 1.3325 & 1.85 & 1.6095 & 0.9360 & 0.9023 & 1 && 1.52 & 1.6773 \\
-%    $\left\langle W \right\rangle$ & 2.4299 & 2.4316 & 2.4583 & 7.5648 & 7.5665 & 7.6222 & 2.0387 & 2.0384 & 2.0400 && 5.6216 & 5.6512 \\
-%    $\left\langle \Phi \right\rangle/Ra$ & 2.4311 & 2.4276 & 2.4333 & 7.5661 & 7.5585 & 7.6194 & 2.0386 & 2.0370 & 2.0335 && 5.6182 & 5.6463 \\
-%    $\delta$ & 0.05\% & 0.16\% & 1.02\% & 0.02\% & 0.11\% & 0.04\% & $<$0.01\% & 0.07\% & 0.32\% && 0.06\% & 0.09\% \\
-%  \bottomrule
-%  \end{tabular}}
-%  \label{tab:visco_mantle}
-%  \end{table}
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{2}{white!100!}{gray!25!}
-\begin{table}[H]
-  \caption{Benchmark cases and corresponding parameters.}
-  \centering
-  \small
-\begin{tabular}{|>{\centering}m{0.08\textwidth}|>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}
-  >{\centering}m{0.05\textwidth} >{\centering}m{0.05\textwidth} |>{\centering\arraybackslash}m{0.2\textwidth}|}
-  \toprule
-    Case & $Ra$ & $\Delta\eta_T$ & $\Delta\eta_y$ & $\eta^*$ & $\sigma_Y$ & Convective regime \\
-  \midrule
-  \rowcolor{gray!25!}
-    1 & $10^2$ & $10^5$ & 1 & - & - & Stagnant lid \\
-    2 & $10^2$ & $10^5$ & 1 & $10^{-3}$ & 1 & Mobile lid  \\
-    \rowcolor{gray!25!}
-    3 & $10^2$ & $10^5$ & 10 & - & - & Stagnant lid  \\
-    4 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 1 & Mobile lid \\
-    \rowcolor{gray!25!}
-    5 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 4 & Periodic \\
-  \bottomrule
-  \end{tabular}
-  \label{tab:tosi_setup}
-  \end{table}
-
-\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{gray!25!}{white!100!}
-\begin{table}
-  \caption{Results of the viscoplastic mantle convection benchmark for different grid resolutions in Cases 1-5 compared with results obtained with ELEFANT
-  and YACC ($100\times100$ elements, from \citealp{Tosi2015}) and in Stone 28 ($32\times32$ elements, from
-  \url{https://github.com/cedrict/fieldstone/tree/master/python_codes/fieldstone_28}).}
-  \centering
-  \resizebox{\textwidth}{!}{
-\begin{tabular}{|cc|ccccccccccc|}
-  \toprule
-  \multicolumn{13}{|c|}{\textbf{Case 1}} \\
-  \midrule
-  Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-  & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-  \hline
-  \rowcolor{gray!25!}
-    YACC & $100\times100$ & 0.7767 & 3.4298 & 3.3143 & 251.7997 & 1.8298 & 2.5516 & - & - & 2.4583 & 2.4333 & 1.02\% \\
-    ELEFANT & $100\times100$ & 0.7758 & 3.4214 & 3.313 & 249.134 & 1.8642 & 2.6119 & - & - & 2.4316 & 2.4276 & 0.16\% \\
-    \rowcolor{gray!25!}
-    Stone 28 & $32\times32$ & 0.7737 & 3.3987 & - & 243.872 & - & - & - & - & - & - & - \\
-%    FALCON & $32\times32$ & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & - & - & 2.4745 & 2.4864 & 0.48\% \\
-%    \rowcolor{gray!25!}
-    FALCON & $100\times100$ & 0.7759 & 3.4280 & 3.3168 & 249.901 & 1.8907 & 2.6340 & - & - & 2.4299 & 2.4311 & 0.05\% \\
-    \midrule
-    \multicolumn{13}{|c|}{\textbf{Case 2}} \\
-    \midrule
-    Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-    & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-    \hline
-    \rowcolor{gray!25!}
-     YACC & $100\times100$ & 0.6058 & 8.5278 & 8.3990 & 142.2020 & 105.1767 & 122.3238 & $1.9853\times10^{-5}$ & 1.6095 & 7.6222 & 7.6194 & 0.04\% \\
-     ELEFANT & $100\times100$ & 0.6034 & 8.5115 & 8.4465 & 140.8067 & 104.1228 & 121.766 & $2.0600\times10^{-5}$ & 1.85 & 7.5665 & 7.5585 & 0.11\% \\
-     \rowcolor{gray!25!}
-     Stone 28 & $32\times32$ & 0.6052 & 8.1715 & - & 141.518 & - & - & - & - & - & - & - \\
-%     FALCON & $32\times32$ & 0.6454 & 8.1422 & 7.4020 & 156.6908 & 95.7226 & 116.9302 & $2.2572\times10^{-5}$ & 0.7194 & 7.4289 & 7.4452 & 0.22\% \\
-%     \rowcolor{gray!25!}
-     FALCON & $100\times100$ & 0.6044 & 8.5266 & 8.4590 & 141.3684 & 104.8867 & 122.0470 & $2.0333\times10^{-5}$ & 1.2914 & 7.5707 & 7.5718 & 0.01\% \\
-     \midrule
-     \multicolumn{13}{|c|}{\textbf{Case 3}} \\
-     \midrule
-     Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-     & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-     \hline
-     \rowcolor{gray!25!} 
-     YACC & $100\times100$ & 0.7286 & 3.0374 & 2.9628 & 100.9467 & 2.0374 & 2.8458 & $4.7907\times10^{-5}$ & 1 & 2.0400 & 2.0335 & 0.32\% \\
-      ELEFANT & $100\times100$ & 0.7275 & 3.0347 & 2.9908 & 100.1208 & 2.0652 & 2.9019 & $4.8080\times10^{-5}$ & 0.9023 & 2.0384 & 2.037 & 0.07\% \\
-      \rowcolor{gray!25!}
-      Stone 28 & $32\times32$ & 0.7277 & 3.0347 & - & 100.018 & - & - & - & - & - & - & - \\
-%      FALCON & $32\times32$ & 0.7281 & 3.0618 & 2.8132 & 101.0092 & 2.1630 & 3.0178 & $4.8500\times10^{-5}$ & 0.7845 & 2.0699 & 2.0695 & 0.02\% \\
-%      \rowcolor{gray!25!}
-      FALCON & $100\times100$ & 0.7275 & 3.0368 & 3.0086 & 100.2776 & 2.0832 & 2.9086 & $4.8129\times10^{-5}$ & 0.9360 & 2.0387 & 2.0386 & $<$0.01\% \\
-      \midrule
-      \multicolumn{13}{|c|}{\textbf{Case 4}} \\
-      \midrule
-      Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-      & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-      \hline
-      \rowcolor{gray!25!} 
-      YACC & $100\times100$ & 0.5289 & 6.5572 & 6.5243 & 79.6202 & 75.4814 & 89.2940 & $1.9174\times10^{-4}$ & 1.6773 & 5.6512 & 5.6463 & 0.09\% \\
-       ELEFANT & $100\times100$ & 0.5277 & 6.5912 & 6.5834 & 79.1105 & 74.7596 & 88.9146 & $1.9860\times10^{-4}$ & 1.5200 & 5.6216 & 5.6182 & 0.06\% \\
-       \rowcolor{gray!25!}
-       Stone 28 & $32\times32$ & 0.5286 & 6.4036 & - & 79.4746 & - & - & - & - & - & - & - \\
-%       FALCON & $32\times32$ & 0.5334 & 6.4679 & 6.3435 & 80.6153 & 76.0511 & 90.2058 & $2.0393\times10^{-4}$ & 0.6453 & 5.6676 & 5.6698 & 0.04\% \\
-%       \rowcolor{gray!25!}
-       FALCON & $100\times100$ & 0.5279 & 6.5897 & 6.5830 & 79.1220 & 75.0715 & 88.8484 & $1.9693\times10^{-4}$ & 1.4133 & 5.6127 & 5.6113 & 0.03\% \\
-       \midrule
-       \multicolumn{13}{|c|}{\textbf{Case 5}} \\
-       \midrule
-       Code & Resolution & $P$ & $\left\langle T \right\rangle_{min}$ & $\left\langle T \right\rangle_{max}$ & Nu$^{top}_{min}$ & Nu$^{top}_{max}$ &
-       $v_{\textrm{rms}}^{min}$ & $v_{\textrm{rms}}^{max}$ & \multicolumn{2}{c}{$(\left\langle \Phi \right\rangle/Ra)_{min}$} & 
-       \multicolumn{2}{c|}{$(\left\langle \Phi \right\rangle/Ra)_{max}$} \\
-       \hline
-       \rowcolor{gray!25!}
-         YACC & $100\times100$ & 0.0792 & 0.6543 & 0.6722 & 2.6950 & 7.2792 & 41.9813 & 98.8230 & \multicolumn{2}{c}{1.4347} & 
-         \multicolumn{2}{c|}{9.2748} \\
-         ELEFANT & $100\times100$ & 0.076 & 0.6514 & 0.6693 & 2.6808 & 7.3506 & 41.2578 & 100.089 & \multicolumn{2}{c}{1.3324} & 
-         \multicolumn{2}{c|}{9.5289} \\
-         \rowcolor{gray!25!}
-         Stone 28 & $32\times32$ & 0.0764 & 0.6521 & 0.6697 & 2.6912 & 7.0763 & 41.61 & 98.183 & \multicolumn{2}{c}{-} & 
-         \multicolumn{2}{c|}{-} \\
-     %    FALCON & $32\times32$ & & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & \multicolumn{2}{c}{-} & 
-     %     \multicolumn{2}{c|}{-} \\
-     %    \rowcolor{gray!25!}
-         FALCON & $100\times100$ & 0.0775 & 0.6519 & 0.6697 & 2.6847 & 7.3629 & 41.5256 & 99.5466 & \multicolumn{2}{c}{1.3399} & 
-         \multicolumn{2}{c|}{9.3993} \\
-  \bottomrule
-  \end{tabular}}
-  \label{tab:visco_mantle}
-  \end{table}
-
-\subsection{Thin layer entrainment}\label{sec:thin}
-This experiment is performed as originally proposed by \citet{vanKeken1997} with two incompressible fluids in a rectangular domain with $L_x=2$ and $L_y=1$
-and gravity acceleration $g_y=10^{10}Ra$. The Rayleigh number and the compositional Rayleigh number are fixed to $Ra=300000$ and $Rc=450000$, respectively.
-Both fluids have constant viscosity, thermal conductivity, specific heat ($\eta=\rho=C_p=1$) and thermal expansion coefficient ($\alpha=10^{-10}$). Fluid 1
-has a density $\rho_1=1$, while fluid 2 is denser ($\rho_2=\rho_1+1.5 \cdot 10^{-10}$) and is placed at the bottom of the domain, for $y \leq 0.025$.
-Temperature are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides of the domain. The initial
-temperature field is given by
-\[T(x,y)=T_u(x,y)+T_l(x,y)+T_r(x,y)+T_s(x,y)-\frac{3}{2}\]
-with
-\begin{eqnarray}
-T_u(x,y)&=&\frac{1}{2}erf\left(\frac{1-y}{2}\sqrt{\frac{u_0}{x}}\right)\nonumber \\
-T_l(x,y)&=&1-\frac{1}{2}erf\left(\frac{y}{2}\sqrt{\frac{u_0}{L_x-x}}\right)\nonumber \\
-T_r(x,y)&=&\frac{1}{2}+\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{y+1}} \exp\left(-\frac{x^2u_0}{4y+4}\right)\nonumber \\
-T_s(x,y)&=&\frac{1}{2}-\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{2-y}} \exp\left(-\frac{(L_x-x)^2u_0}{8-4y}\right)\nonumber
-\end{eqnarray}
-and
-\begin{eqnarray}
-u_0&=&\frac{L_x^{7/3}}{(1+L_x^4)^{2/3}}\left(\frac{Ra}{2\sqrt{\pi}}\right)^{\frac{2}{3}}\nonumber \\
-Q&=&2\sqrt{\frac{L_x}{\pi u_0}}\nonumber
-\end{eqnarray}
-
-The experiment is performed with two grid resolution (\num{125x40} and \num{200x80} elements), with 100 markers per element and CFL of 0.25. $v_{\textrm{rms}}$
-(see Sec. \ref{sec:error}) calculated for both the simulations match well with results from \citet{vanKeken1997} and obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and ELEFANT \citep{Thieulot2014}, obtained with similar grid resolutions
-(Fig. \ref{fig:thin}). In addition, a variety of simulations with different resolutions and aspect ratios are performed to verify that the sensitivity of the
-initial velocity field. Results match well with results from \citet{vanKeken1997} (in the grey area) and \citet{Thieulot2014} (Fig. \ref{fig:thin_initial}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Thin_layer}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_thin.pdf}
-\caption{$v_{\textrm{rms}}$ for the thin layer experiment as function of time for different grid resolution. Results are compared with results obtained by
-\citet{vanKeken1997} (black lines) and with ASPECT and ELEFANT.}
-\label{fig:thin}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrmszero.pdf}
-\caption{$v_{\textrm{rms}}$ at $t=0$ for the thin layer experiment as function of the element size for different aspect ratios of the elements. Results are compared
-with results obtained by \citet{vanKeken1997} (black dots in the grey area) and with ASPECT and ELEFANT.}
-\label{fig:thin_initial}
-\end{figure}
-
-\subsection{Sinking hydrated cylinder}\label{sec:quinquis}
-This experiment is performed as explained in \citet{Quinquis2014} and simulates the sinking of a cold, hydrated cylinder into a warm, dry mantle. The domain
-is square with $L_x=L_y=$ \SI{300}{\km}, a grid resolution of \num{300x300} elements and 25 markers per element. The sinking cylinder is located at $x=$
-\SI{150}{\km} and $y=$ \SI{170}{\km} with a radius of \SI{20}{\km} and has $\rho_c=$ \SI{3250}{\kg\per\cubic\m}, $\eta_c=$ \SI{e23}{\pascal\s}, a thermal
-conductivity $k_c=$ \SI{4.5}{\watt\per\m\per\kelvin},a specific heat $C_{p_c}=$ \SI{1250}{\joule\per\kg\per\kelvin}; the surrounding mantle has $\rho_m=$
-\SI{3200}{\kg\per\cubic\m}, $\eta_m=$ \SI{e20}{\pascal\s}, a thermal conductivity $k_m=$ \SI{105}{\watt\per\m\per\kelvin} and specific heat $C_{p_m}=$
-\SI{1250}{\joule\per\kg\per\kelvin}; a 58 km-thick lithosphere overlies the mantle and has $\rho_l=$ \SI{3200}{\kg\per\cubic\m}, $\eta_l=$ \SI{e23}{\pascal\s},
-a thermal conductivity $k_l=$ \SI{4.5}{\watt\per\m\per\kelvin}, specific heat $C_{p_l}=$ \SI{750}{\joule\per\kg\per\kelvin} and a thermal diffusivity of
-\SI{e-6}{\square\m\per\s}. Velocity boundary conditions are set to free slip on all sides of the domain. Temperature increase linearly in the lithosphere,
-from 0° to 1300°C, and in the mantle, from 1300° to 1360.5°C. The initial temperature of the cylinder is fixed to 400°C. Throughout the evolution, temperature
-is fixed to 0° and 1360.5°C at the top and bottom of the domain, respectively. The initial bound water content is imposed at 2 wt.\% in the cylinder and at
-0 wt.\% in both the mantle and the lithosphere. The maximum amount of water in the mantle is fixed to 0.2 wt.\%, while maximum water content in the cylinder
-and in the lithosphere is function of pressure and temperature and is calculated using a serpentinized harzburgite with Perple\_X, as in \citet{Quinquis2014}.
-
-As shown in \citet{Quinquis2014}, the model is characterised by a progressive dehydration in the external portion of the cylinder, due to the increase of
-temperature, with a consequent vertical migration of free water that hydrates the mantle above the cylinder, up to the lithosphere. The distribution of
-bound and free water during the evolution of the experiment is shown in Fig. \ref{fig:quinquis} (left and right columns, respectively). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Hydration}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Hydration.pdf}
-\caption{Distribution of bound and free water (left and right columns, respectively) at $t=0.5$ and $t=$ \SI{1.25}{\mega\year} (first and second rows,
-respectively) for the sinking hydrated cylinder experiment. Only markers with bound or free water not equal to 0 are plotted.}
-\label{fig:quinquis}
-\end{figure}
-
-\subsection{Mantle melting curves}\label{sec:katz}
-This instantaneous experiment is performed in order to recreate isobaric melting curves, as function of temperature, bulk water and content of water in the
-melt, to compare with melting curves obtained by \citet{Katz2003}. The domain is square with $L_x=L_y=100$ km, a grid resolution of \num{100x100} elements and
-25 markers per element. Density is set in the entire domain to \SI{3300}{\kg\per\cubic\m} and temperature increases from the sides to the centre from 1000° to
-1500°C. The experiment is performed for bulk water contents from 0 to 0.3 wt.\%.
-
-Fig. \ref{fig:fraction} shows isobaric melting curves for \SI{1}{\giga\pascal} (panel a) and \SI{3}{\giga\pascal} (panel b) calculated for different bulk water
-contents. The curves well match with those obtained for the mantle by \citet{Katz2003}, showing the saturation at the solidus for bulk water of 0.2 and 0.3 wt.\%, indicating
-that the code correctly determines melt fraction as function of pressure, temperature and bulk water content. Same results can be observed in Fig.
-\ref{fig:water}a, where isobaric and isothermal melting curves are shown as function of bulk water content. In addition, Fig. \ref{fig:water}b shows that the
-code correctly determines also the percentage of water in the melt as function of pressure, temperature and melt fraction. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Melting}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Fraction.pdf}
-\caption{Isobaric melting curves for 1 GPa (panel a) and 3 GPa (panel b) as function of temperature with different bulk water contents.}
-\label{fig:fraction}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Water.pdf}
-\caption{Isobaric melting curves for 1.5 GPa as function of bulk water (panel a) and water in the melt (panel b) calculated for different temperatures.}
-\label{fig:water}
-\end{figure}
-
-\clearpage
-
-\bibliography{export}
-
-\end{document}

### 2022-09-07 tex/Benchmark/error.tex
update main

--- 2022-01-28 .debris/2022-01-27/Paper.tex
@@
-\documentclass[hidelinks,10pt,a4paper]{article}
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage{booktabs}
-\usepackage{multirow}
-\usepackage[cm]{fullpage}
-\usepackage[round]{natbib}
-\bibliographystyle{abbrvnat}
-\usepackage{doi}
-\usepackage{bm}
-\usepackage[table]{xcolor}
-\hypersetup{
-    colorlinks,
-    linkcolor={blue},
-    citecolor={blue},
-    urlcolor={blue}
-}
-\usepackage{float}
-\restylefloat{table}
-\usepackage{siunitx}
-\DeclareSIUnit\year{yr}
-
-\begin{document}
-
-\section{Introduction}\label{sec:intro}
-This supplement includes a description of the code, with a presentation of the features implemented followed by results of extensive benchmarks listed in
-Table \ref{tab:list}, implemented to verify that the code solves correctly a large range of problems. Description and results of all benchmarks are shown
-in Section \ref{sec:benchmark}.
-
-\begin{table}[H]
-\caption{List of benchmarks implemented for each feature of the code. Data of all benchmarks can be found at 
-\url{https://github.com/aleregorda/Benchmarks}}
-\centering
-\begin{tabular}{| l | l | l |}
-\toprule
- \textbf{Feature}  & \textbf{Benchmark} & \textbf{Section} \\
-\midrule
-  Solver  & Stokes flow & \ref{sec:stokes} \\
-\hline
-  \multirow{2}{*}{Markers advection} & Zalesak disk & \ref{sec:runge} \\
-    & Conservative Interpolation Velocity & \ref{sec:CVI} \\
-\hline
-  \multirow{4}{*}{Momentum equation}  & Poiseuille flow & \ref{sec:poiseuille}\\
-    & Instantaneous 2D sphere & \ref{sec:ist_sphere} \\
-    & Rayleigh-Taylor experiment & \ref{sec:rayleigh} \\
-    & Falling block & \ref{sec:block} \\
-\hline
-    & 2D Stokes sphere & \ref{sec:sphere} \\
-    Sticky air and  & Stabilisation algorithm & \ref{sec:stab} \\
-    free surface & Topography relaxation & \ref{sec:crameri} \\
-  & Spontaneous subduction & \ref{sec:subduction} \\
-\hline
-  Erosion and sedimentation  & - & - \\
-\hline
-   Non-linear  & Slab detachment & \ref{sec:slab} \\
-   visco-plasticity & Indenter & \ref{sec:indenter} \\
-    & Brick & \ref{sec:brick} \\
-\hline
-  \multirow{3}{*}{Energy equation}  & Advection stabilisation & \ref{sec:advection} \\
-    & Simple shear heating & \ref{sec:simple_shear} \\
-    & Shear and adiabatic heating & \ref{sec:shear} \\
-\hline
-\multirow{3}{*}{Energy + momentum} & Mantle convection & \ref{sec:mantle} \\
-& Viscoplastic mantle convection & \ref{sec:visco_mantle} \\
-& Thin layer entrainment & \ref{sec:thin} \\
-\hline
-  Phase changes  & \multirow{2}{*}{Hydrated sinking cylinder} & \multirow{2}{*}{\ref{sec:quinquis}} \\
-  and hydration  &  & \\
-\hline
-  Melting & Experimental melting curves & \ref{sec:katz} \\
-\bottomrule
-\end{tabular}
-\label{tab:list}
-\end{table}
-
-The code is written in Fortran90 and uses the finite element method (FEM) with quadrilateral $Q_1 \times P_0$ elements (continuous bilinear velocity and
-discontinuous constant pressure), associated with MUMPS\footnote{\url{http://mumps.enseeiht.fr/}} \citep{Amestoy2001,Amestoy2019} that is a software package
-for solving systems of linear equations of the form $A \cdot x = b$, where \textit{A} is a square sparse matrix, by means of a direct method. Since
-$Q_1 \times P_0$ elements do not satisfy Ladyzhenskaya, Babuska and Brezzi (LBB) stability condition \citep{Donea2003}, elemental pressure is elaborated
-in post-processing to avoid spurious pressures by means of a double interpolation to smooth the pressure. Following this procedure, the elemental pressure
-is interpolated onto nodes and then back onto elements. After the smoothing procedure, the pressure field is calculated again on the nodes to be used in
-combination with the lithostatic pressure, which is calculated onto nodes. Correctness of the solution and performances of the solver in terms of time and
-memory usage are tested solving the Stokes flow with the analytical solution proposed by \citet{Donea2003} (Section \ref{sec:stokes}).
-
-The thermo-mechanics of crust-mantle systems is described by means of the conservation of mass, momentum and energy equations, expressed as follows:
-\begin{eqnarray}
-\label{eq:mass}\frac{\partial \rho}{\partial t}+\bm{\nabla} \cdot (\rho \bm{u})&=&0\\
-\label{eq:momentum}-\bm{\nabla} p + \bm{\nabla} \cdot \bm{\tau} + \rho \bm{g}&=&0\\
-\label{eq:energy}\rho C_p \left(\frac{\partial T}{\partial t} + \bm{u} \cdot \bm{\nabla} T \right)&=&\bm{\nabla} \cdot \left(k\bm{\nabla} T\right) + H_{tot}
-\end{eqnarray}
-where $\rho$ is the density, $\bm{u}$ is the velocity, \textit{p} is the pressure, $\bm{\tau}$ is the deviatoric stress, $\bm{g}$ is the gravity acceleration,
-$C_p$ is the specific heat at constant pressure, \textit{T} is the temperature, \textit{k} is the thermal conductivity and $H_{tot}$ is the total internal heat
-production. Density variations due to temperature are generally small enough to assume the density as constant ($\rho=\rho_0$) in Eq. \ref{eq:mass} and in
-Eq. \ref{eq:energy}, while it must be treated as a variable in the buoyancy term of Eq. \ref{eq:momentum}, such that \[\rho=\rho_0(1-\alpha(T-T_0))\]
-where $\rho_0$ is the density at a reference temperature $T_0$ and $\alpha$ is the coefficient of thermal expansion. Eq. \ref{eq:mass} then can be rewritten as
-\begin{equation}\label{eq:mass_0}
-\bm{\nabla} \cdot \bm{u}=0
-\end{equation}
-This simplification is known as the extended Boussinesq approximation.
-
-The time step $dt$ is chosen in according to Courant-Friedrichs-Lewy condition \citep{Anderson}, such that
-\[dt=C_n\; \min\left(\frac{\Delta x}{\max|\bm{u}|},\frac{(\Delta x)^2}{\kappa}\right)\]
-where $0<C_n<1$ is the Courant (CFL) number, $\Delta x$ is the minimum dimension of smallest element, $\max|\bm{u}|$ is the maximum velocity calculated on the
-entire domain and
-$\kappa=\frac{k}{\rho C_p}$ is the thermal diffusivity \citep{Thieulot2014}.
-
-\section{Lagrangian markers}\label{sec:markers}
-Elemental properties (density, viscosity, thermal conductivity, specific heat and thermal expansion) needed to solve Eqs. \ref{eq:mass_0}, \ref{eq:momentum}
-and \ref{eq:energy} are related to the composition of each element, determined by means of Lagrangian markers that are characteristic of different materials
-in the domain. Elemental properties (with the sole exception of the viscosity) are calculated using an arithmetic mean, as \[P_e=\frac{1}{n}\sum_{i=1}^n P_i\]
-where $P_e$ is the elemental property, $P_i$ is the property characteristic of the material of each marker and $n$ is the number of markers of the element.
-Differently, the average scheme for the elemental viscosity can be chosen between harmonic, geometric and arithmetic mean. Markers can be placed regularly or
-randomly at the beginning of the simulation and their advection is performed by either a 2nd-order or a 4th-order Runge-Kutta in space, interpolating the velocity 
-field on each marker by means of the shape functions (see Section \ref{sec:runge}). The interpolated velocity is then corrected by means of the Conservative
-Velocity Interpolation (CVI), which introduces a corrective term to reduce dispersion and clustering of particles in both steady state and time-dependent 
-\citep{Wang2015} (see Section \ref{sec:CVI}).
-
-The initial distribution of the markers is created using the open-source code library
-Geodynamic World Builder \citep{Fraters2019}. During the simulation, each marker carries memory of temperature, pressure and accumulated strain, which are
-determined interpolating the relative nodal parameter. In particular, pressure is interpolated onto each marker after the smoothing procedure. The number
-of markers contained in each element is maintained between $n_{min}$ and $n_{max}$. When in an element there are less markers than $n_{min}$ the code adds
-random markers to reach the $n_{min}$, while if the number is higher than $n_{max}$ some of them are randomly deleted. When new markers are added, they assume
-the properties of the nearest marker. In this way elements are never empty and maintain a number of markers inside a prefixed range.
-
-\section{The momentum equation}\label{sec:mom_eq}
-The penalty method is implemented so that Eq. \ref{eq:mass_0} can be written as:
-\begin{equation}\label{eq:mass_penalty}
-\bm{\nabla} \cdot \bm{u} + \frac{p}{\lambda}=0
-\end{equation}
-where $\lambda$ is the so-called penalty parameter, which should be 6-7 orders of magnitude larger than the shear viscosity to ensure that mass conservation
-is satisfied \citep{Donea2003,Thieulot2014}.
-
-The deviatoric stress tensor in Eq. \ref{eq:momentum} can be written in terms of the strain rate tensor as $\bm{\tau}=2\eta\dot{\bm{\epsilon}}$, with
-$\dot{\bm{\epsilon}}=\frac{1}{2}\left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)$. Therefore, Eq. \ref{eq:momentum} can be rewritten as:
-\begin{equation}\label{eq:momentum_pressure}
--\bm{\nabla} p + \bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right) + \rho \bm{g} = 0
-\end{equation}
-Finally, using pressure from Eq. \ref{eq:mass_penalty}, Eq. \ref{eq:momentum_pressure} can be rewritten as:
-\begin{equation}\label{eq:momentum_penalty}
-\lambda \bm{\nabla} \left(\bm{\nabla} \cdot \bm{v}\right)+\bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right)+\rho\bm{g}=0
-\end{equation}
-The penalty method is been associated to the iterative Uzawa method, as described in detail in \citet{Dabrowski2008} and \citet{Thieulot2014}
-(see Section \ref{sec:poiseuille}). Results of benchmarks performed to verify the correctness in the implementation of Eq. \ref{eq:momentum_penalty} are
-shown in Sections \ref{sec:ist_sphere} and \ref{sec:rayleigh}.
-
-To support large viscosities variations the penalty parameter is related to the effective elemental viscosity by means of a dimensionless coefficient, so that
-$\lambda=\lambda_e(e)\eta_{eff}(e)$ \citep{Marotta2006,Dabrowski2008,Thieulot2014}. Benchmark of the falling block \citep{Gerya2003a,Gerya2010b,Thieulot2011}
-is performed to verify that the code can correctly deal buoyancy driven flows with strong viscosity contrasts (Section \ref{sec:block}).
-
-\section{Sticky air and free surface}\label{sec:sticky}
-The Earth's surface can be treated by means of either the so-called sticky-air or a true free surface method, both of them implemented in the code. In the
-sticky-air method the surface is approximated with the introduction of a buoyant layer with a viscosity at least four orders of magnitude lower than the crust
-\citep{Schmeling2008,Crameri2012} and the interface between lithosphere and air is defined using a chain of passive markers that are advected as the Lagrangian
-markers. The correctness of the evolution of the markers chain is tested with the experiment of a 2D time-dependent Stokes sphere below a free surface and
-compared with results from ASPECT\footnote{\url{https://aspect.geodynamics.org/}} \citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} (Section
-\ref{sec:sphere}). In the true free surface case the top boundary is assumed stress-free and velocities are not fixed. In this case topography variations are
-described by vertical deformations of the mesh that depend on the velocity field of the nodes that identify the free surface, while horizontal deformations are
-not taken into account. This procedure is known as the Arbitrary Lagrangian-Eulerian (ALE) method and its implementation follows the technique described in
-\citet{Thieulot2011}. However, although the implementation of a true free surface better reproduces laboratory experiment, extremely small time steps can be
-necessary to maintain stability \citep{Kaus2010a,Quinquis2011,Thieulot2014}. Therefore, the stability algorithm proposed by \citet{Kaus2010a} is implemented
-to avoid instabilities due to high density differences at the free surface when using too large time steps. The implementation of this algorithm are tested by
-performing the experiment described by \citet{Kaus2010a} (Section \ref{sec:stab}).
-
-The topography relaxation benchmark proposed by \citet{Crameri2012} is performed to verify that the code correctly recovers topography variations in case of
-both the sticky-air and the true free surface method (Section \ref{sec:crameri}). Finally, Section \ref{sec:subduction} show the results of the spontaneous
-subduction experiment described by \citet{Schmeling2008}.
-
-\section{Erosion and sedimentation}\label{sec:erosion}
-Surface processes at the lithosphere-air interface are implemented taking into account fluvial erosion (long-range processes), hillslope processes (short-range
-processes) and failure of slopes exceeding a threshold angle $\theta$ \citep{Burov1997,Tucker1999,Sacek2012,Croissant2014,Ueda2015,Beucher2020}. Erosion rates
-$\dot{E}_{lr}$ due to long-range processes can be described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, in according to
-\begin{equation}
-\dot{E}_{lr}=K_f A^{m_e} \bm{\nabla} h^{n_e}
-\end{equation}
-where $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $\bm{\nabla} h$ is the slope and $m_e$ and $n_e$ are parameters chosen in to
-constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}. Erosion rates $\dot{E}_{sr}$ due to short-range processes are
-implemented by assuming that transport is linearly proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}, in according to
-\begin{equation}
-\dot{E}_{sr}=K_d \bm{\nabla}^2 h
-\end{equation}
-where $K_d$ is the effective diffusivity. In addition, the slope is modified if it still exceeds a threshold angle after long- and short-range processes
-\citep{Ueda2015}.
-
-Offshore sedimentation processes are implemented similarly to hillslope processes but with a depth-dependent coefficient \citep{Sacek2012,Beucher2020}.
-The sedimentation rate $\dot{S}$ is calculated as
-\begin{equation}
-\dot{S}=K_s \bm{\nabla}^2 h
-\end{equation}
-with
-\begin{equation}
-K_s=K^*_s \exp\left(\frac{h-h_{sl}}{h_0}\right)
-\end{equation}
-where $K^*_s$ is a constant, $h_{sl}$ is the sea level and $h_0$ is a characteristic length. In case of sticky air, slope is determined by means of the markers
-chain, which is vertically corrected considering erosion and sedimentation rates. After the correction, continental markers above the markers chain are
-transformed in air markers and air markers below the markers chain become sediments. Similarly, slope is determined considering the nodes on the top boundary
-in case of true free surface and they are vertically corrected as for the markers chain. In this case, continental markers above the top boundary are simply
-deleted.
-
-\section{Non-linear rheologies}\label{sec:nonlinear}
-Non-linear rheologies are implemented combining viscous creep (dislocation and diffusion) and plastic yielding. For each marker, diffusion and dislocation
-viscosity ($\eta_{df|ds}$) can be determined as follows \citep{Karato1993,Wang2016}:
-\begin{equation}\label{eq:viscous}
-\eta_{df|ds}=\left(\frac{d^m}{A}\right)^{\frac{1}{n}}I_2^{\frac{(1-n)}{n}}\exp\left(\frac{Q+pV}{nRT}\right)
-\end{equation}
-where $d$ is the grain size, $m$ is the grain size exponent, $n$ is the stress exponent, $A$ is the pre-exponential factor, $I_2$ is the square root of the
-second invariant of the strain rate tensor, $Q$ is the activation energy, $p$ is the pressure, $V$ is the activation volume, $R$ is the gas constant and $T$
-is the temperature. Pressures and temperatures are determined by the interpolation of the nodal parameters, while strain rates are calculated by means of the
-derivative of the velocities on the nodes. In case of diffusion creep $n=1$ and $m>0$, in case of dislocation creep $n>1$ and $m=0$. The viscous creep $\eta_{cp}$
-is then calculated as the harmonic average between $\eta_{df}$ and $\eta_{ds}$:
-\begin{equation}\label{eq:visco_creep}
-\eta_{cp}=\left(\frac{1}{\eta_{df}}+\frac{1}{\eta_{ds}}\right)^{-1}
-\end{equation}
-The implementation of the non-linear viscous creep viscosity depending on the strain rate is tested by means of the slab detachment benchmark
-(Section \ref{sec:slab}).
-
-Plastic yielding is implemented rescaling $\eta_{cp}$ in order to limit the stress below the yield stress $\sigma_y$ \citep{Thieulot2008,Thieulot2014,Glerum2018},
-obtaining
-\begin{equation}\label{eq:plastic}
-\eta_{pl}=\frac{\sigma_y}{2I_2}
-\end{equation}
-where $\eta_{pl}$ is the plastic viscosity and the yield stress is determined following the Drucker-Prager criterion, such as
-\begin{equation}\label{eq:yield}
-\sigma_y=C\;\cos(\phi)+p\;\sin(\phi)
-\end{equation}
-where $C$ is the cohesion and $\phi$ is the internal friction angle. In case of negative pressure it is imposed equal to 0, so that negative yield stress are
-excluded. The correctness of the non-linear solution in case of plastic viscosity with variable internal friction angle is verified performing indenter and
-brick experiments (Sections \ref{sec:indenter} and \ref{sec:brick}, respectively).
-
-Strain softening is taken into account for both viscous creep and plastic viscosity \citep{Huismans2003,Babeyko2005,Huismans2005,Sobolev2005,Warren2008a} by
-means of the accumulated strain $\epsilon$ memorised by each marker. Softening in the viscous creep determines a linear decrease of $\eta_{vc}$ by means of a
-viscous strain softening factor $W_S$ that increases linearly from $W_{S_0}$ to $W_{S_{\infty}}$ for $\epsilon_{S_0}<\epsilon_S<\epsilon_{S_{\infty}}$ 
-\citep{Warren2008a}. This viscous softening can be related to strain-induced grain size reduction \citep{Warren2008a}. Differently, plastic softening is 
-simulated with a linear decrease of internal friction angle $\phi(\epsilon)$ and cohesion $C(\epsilon)$ in according to
-\begin{eqnarray}
-\label{eq:friction1}\phi(\epsilon)&=&\phi_0+(\phi_{\infty}-\phi_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}\\
-\label{eq:friction2}C(\epsilon)&=&C_0+(C_{\infty}-C_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}
-\end{eqnarray}
-where $\phi_0$, $C_0$ and $\phi_{\infty}$, $C_{\infty}$ are internal friction angle and cohesion for $\epsilon_0$ and $\epsilon_{\infty}$, respectively
-\citep{Huismans2003,Huismans2005,Warren2008a,Thieulot2014}. Plastic softening approximates deformation-induced softening of faults and brittle shear zones
-\citep{Warren2008a}.
-
-Plastic yielding and viscous creep are then combined to obtain a viscoplastic viscosity $\eta_{vp}$ as follows, assuming that they are independent processes
-\citep{Karato2008,Glerum2018}:
-\begin{equation}\label{eq:viscoplastic}
-\eta_{vp}=\min(\eta_{cp},\eta_{pl})
-\end{equation}
-Finally, effective viscosity $\eta_{eff}$ is capped by the minimum and the maximum viscosity ($\eta_{min}$ and $\eta_{max}$, respectively) to avoid extremely
-low or high viscosity \citep{Glerum2018} as follows
-\begin{equation}\label{eq:effective}
-\eta_{eff}=\min(\max(\eta_{vp},\eta_{min}),\eta_{max})
-\end{equation}
-Viscous creep and plastic yielding are non-linear rheologies because of their dependence on the velocity field through pressure and strain rates. Therefore,
-the solution is determined by means of Picard-type iterations, until convergence of the velocity field \citep{Glerum2018}. The convergence is verified at each
-iteration $i$ via the nonlinear residual $\bm{\mathcal{R}}^i$ that can be determined as
-\begin{equation}\label{eq:convergence}
-\bm{\mathcal{R}}^i= \mathbb{K} (\eta_{eff}(\dot{\bm{\epsilon}}^{i-1},p^{i-1}))\cdot \bm{v}^{i-1} - \bm{f}^i
-\end{equation}
-%\begin{equation}\label{eq:convergence}
-%\chi=1-\frac{\langle (\bm{v}^i-\langle\bm{v}^i\rangle)\cdot (\bm{v}^{i+1}-\langle\bm{v}^{i+1}\rangle) \rangle}{\sqrt{\sigma^i\sigma^{i+1}}}
-%\end{equation}
-%Iterations are performed until either convergence ($\chi<tol$) or maximal number of non-linear iterations ($it_{max}$) is reached.
-where $\mathbb{K}$ is the Stiffness matrix and $\bm{f}$ is the right hand side vector \citep{Spiegelman2016,Glerum2018}.  The $L_2$-norm is extracted from
-$\bm{\mathcal{R}}^i$ and it is normalised as follows
-\begin{equation}\label{eq:convergence_res}
-\frac{||\bm{\mathcal{R}}^i||_2}{||\bm{\mathcal{R}}^0||_2}
-\end{equation}
-where $||\bm{\mathcal{R}}^0||_2$ is the $L_2$-norm of the first nonlinear iteration.
-Since often the normalised nonlinear residual drop very quickly over the first 2-3 iterations, also the $L_2$-norm of velocity and pressure residuals are
-calculated as
-\begin{equation}\label{eq:convergence_vel}
-\frac{||\Delta\bm{v}||_2}{||\bm{v}||_2}=\frac{||\bm{v}_i-\bm{v}_{i-1}||_2}{||\bm{v}_{i}||_2}
-\end{equation}
-and
-\begin{equation}\label{eq:convergence_press}
-  \frac{||\Delta p||_2}{||p||_2}=\frac{||p_i-p_{i-1}||_2}{||p_{i}||_2}
-\end{equation}
-respectively. Iterations are performed until either maximal number of non-linear iterations ($it_{max}$) is reached or all the normalised nonlinear residual, 
-the velocity residual and the pressure residual converge under a defined tolerance $tol$.
-
-\section{The energy equation}\label{sec:en_eq}
-The stabilisation of the advection term of the energy equation (Eq. \ref{eq:energy}) needed to avoid possible oscillations in the thermal solution in the case
-when advection dominates over diffusion, is implemented by means of a streamline-upwind Petrov–Galerkin (SUPG) method, for which the advection term is modified
-follows the discussion in \citet{Thieulot2011} and \citet{Thieulot2014} (see Section \ref{sec:advection} for benchmark). The total internal heat production
-$H_{tot}$ in Eq. \ref{eq:energy} includes radiogenic heating $H_r$, shear heating $H_s$ and adiabatic heating $H_a$, where
-\begin{equation}\label{eq:shear}
-H_s=2\eta \dot{\bm{\epsilon}} : \dot{\bm{\epsilon}} = 2\eta(\dot{\epsilon}_{xx}^2+\dot{\epsilon}_{yy}^2+2\dot{\epsilon}_{xy}^2)
-\end{equation}
-and
-\begin{equation}\label{eq:adiabatic}
-H_a=T \alpha \frac{Dp}{Dt} \approx -\alpha T \rho g_y v
-\end{equation}
-The correctness of the implementation and the computation of shear and adiabatic heating is verified performing an exercise with an analytical solution
-(Section \ref{sec:simple_shear}) and Exercise 9.4 in \citet{Gerya2010b} (Section \ref{sec:shear}). Furthermore, mantle convection and thin layer entrainment
-benchmarks (Sections \ref{sec:mantle} and \ref{sec:thin}, respectively) are performed to verify that the code solve correctly Eqs. \ref{eq:momentum_penalty}
-and \ref{eq:energy} in case of a temperature field characterised by an initial perturbation and with a temperature-dependent density. At last, the viscoplastic
-mantle convection benchmark is performed to check also the viscoplasticity in case of a initial perturbed temperature field (Section \ref{sec:visco_mantle}).
-
-\section{Phase transitions and hydration}\label{sec:phase}
-Variations of effective density, specific heat and coefficient of thermal expansion during the evolution of crustal and mantle materials at different
-pressure-temperature ($p$-$T$) conditions are computed by means of version 6.8.6 of Perple\_X software package \citep{Connolly2005}, similarly to the
-implementation described in \citet{Marotta2020}. By default, mineral assemblages and properties of each lithology (oceanic crust, lower and upper continental
-crust, sediments and dry or hydrated mantle) are calculated for temperatures from \SIrange{330}{1600}{\kelvin}, with increments of approximately \SI{4}{\kelvin},
-and pressures from \SIrange{0.1}{30}{\giga\pascal}, with increments of approximately \SI{0.03}{\giga\pascal}, for a total of almost 100000 points.
-
-The introduction of phase transitions produces two main effects that must be taken into account in numerical models: variations in density and release/absorption
-of latent heat \citep{Tackley2010}. Effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into
-account considering effective coefficient of thermal expansion \citep{Christensen1985,Zhong2015}. Similarly, effects of latent heat in the energy equation
-(eq \ref{eq:energy}) are taken into account by considering effective specific heat and coefficient of thermal expansion
-\citep{Christensen1985,Tackley2010,Zhong2015}. Effective density, specific heat and coefficient of thermal expansion are included in Perple\_X files and are
-assumed by each Lagrangian marker in according to its $p$-$T$ conditions.
-
-Hydration processes related to dehydration of subducting lithosphere \citep{Schmidt1998,Liu2007,Faccenda2009,Faccenda2010,Faccenda2014,Rosas2016} strongly
-influence the thermo-mechanics inside the mantle wedge, mainly because of both weakening effects on the mantle rheology and density variations in case of
-mantle serpentinization \citep{Gerya2002,Honda2003,Arcay2005,Roda2010,Regorda2017}. In the case that hydration is switched on, the amount of bound and free
-water is memorised by each marker, following the implementation of \citet{Quinquis2014}; hydration and dehydration processes are related to the amount of bound
-water of each marker and to the maximum amount of water it can transport, i.e. if the amount of bound water exceeds the maximum amount of water, the marker
-dehydrates and releases free water that can hydrates under-saturated markers. Maximum water content of each marker is determined as function of lithology and
-$p$-$T$ conditions and it is calculated using Perple\_X, in the same way than effective density, specific heat and coefficient of thermal expansion. Bound water
-is advected together with the markers, neglecting the effect of bound water diffusion, while free water simply migrates vertically and is not coupled to the
-solid-phase flow of the mantle wedge \citep{Arcay2005,Quinquis2014}. The correctness of the migration scheme of the free water and its absorption into bound
-water by under-saturated markers are verified performing the experiment described in \citet{Quinquis2014} (Sec. \ref{sec:quinquis}).
-
-The mantle viscosity weakening related to the amount of water has been extensively studied \citep{Chopra1981,Hirth2003} and it is implemented in according to
-\citet{Arcay2005} and \citet{Horiuchi2016}, as follows
-\begin{equation}\label{eq:wet_visc}
-\eta_{wet}=\eta_{dry}\left(\left[1-\frac{1}{f_v}\right]\exp\left(-\frac{[OH^-]}{[OH^-]_0}\right)+\frac{1}{f_v}\right)
-\end{equation}
-where $f_v$ is the viscosity reduction factor between dry and wet conditions, $[OH^-]$ is the water content and $[OH^-]_0$ is a reference water content set to
-620 ppm (0.062 wt.\%) \citep{Arcay2005}. By default, $f_v$ is set to 100, in according to \citet{Arcay2005} and \citet{Horiuchi2016}.
-
-\section{Mantle melting}\label{sec:melt}
-Melt fraction in the mantle depends on temperature ($T$, in °C), pressure ($p$, in \si{\giga\pascal}) and the water content in the melt ($X_{H_2O}$, in wt.\%).
-The determination of the melt fraction $M$ is implemented as explained by \citet{Katz2003}, according to the successive modification by \citet{Langmuir2006} and
-\citet{Kelley2010}, as follows
-\begin{eqnarray}\label{eq:melt_fr}
-M(p,T,X_{H_2O})&=&-T+T_{sol}+(x_t\cdot ln(p+y_t))\cdot X_{H_2O}-\nonumber\\
-&&+K_m\left(\frac{X_{H_2O}^{bulk}}{D_{H_2O}(1-X_{H_2O})+X_{H_2O}}\right)^{\gamma}
-\end{eqnarray}
-where $T_{sol}$ is the temperature of the dry solidus, $D_{H_2O}$ is the partition coefficient and $x_t\cdot log(p+y_t)$ indicates the pressure dependence of
-the melting curve \citep{Kelley2010}. $x_t$, $y_t$, $K_m$ and $\gamma$ are parameters chosen in according to \citet{Kelley2010}.
-However, $X_{H_2O}$ depends on the melt fraction as
-\begin{equation}\label{eq:melt_wt}
-X_{H_2O}(M)=\frac{X_{H_2O}^{bulk}}{D_{H_2O}+M(1-D_{H_2O})}
-\end{equation}
-and a numerical solution can be found using a root-finder method \citep{Katz2003,Wang2016}. Furthermore, the content of water in the melt is limited by the
-pressure-dependent saturation concentration of water in the melt, determined as
-\begin{equation}\label{eq:melt_max}
-X_{H_2O}^{sat}=\chi_1 p^{\lambda_m}+\chi_2 p
-\end{equation}
-where $\chi_1$, $\chi_2$ and $\lambda_m$ are parameters chosen according to \citet{Katz2003}, \citet{Langmuir2006} and \citet{Kelley2010}. The correctness in
-the determination of melt fractions and water contents in the melt are tested comparing model results with experimental curves obtained by \citet{Katz2003}
-(Sec. \ref{sec:katz}).
-
-In case of melting, viscous creep viscosity ($\eta_{vp}$) is modified according to \citet{Wang2016}, as
-\begin{equation}\label{eq:melt_visc}
-\eta_{melt}=\eta_{vp}\cdot \exp(\alpha_m M)
-\end{equation}
-where $\alpha_m$ are melt fraction factors for dislocation and diffusion creep, chosen in according to \citet{Wang2016}.
-
-Effective density ($\rho_{eff}$) for partially molten rocks is calculated as
-\begin{equation}\label{eq:melt_dens}
-\rho_{eff}=\rho_s(1-M)+\rho_m M
-\end{equation}
-where $\rho_s$ and $\rho_m$ are the densities of the solid and the molten rock, respectively \citep{Gerya2003a,Wang2016}. The density of the solid rock is
-extracted by Perple\_X, while the density of the molten portion is calculated as
-\begin{equation}\label{eq:melt_dens_M}
-\rho_m(p,T)=\rho_0[1-\alpha(T-T_0)][1+\beta(p-p_0)]
-\end{equation}
-where $\rho_0$ is the density at the reference temperature ($T_0$) and pressure($p_0$), and $\alpha$ and $\beta$ are the thermal expansion and compressibility
-coefficients, respectively \citep{Gerya2003a,Wang2016}.
-
-As for the phase transitions, effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into account
-considering effective coefficient of thermal expansion. Similarly, latent heat due to melting/crystallisation can be implicitly included in the energy equation
-(Eq. \ref{eq:energy}) by considering effective specific heat ($C_{p_{eff}}$) and thermal expansion ($\alpha_{eff}$) of the partially molten rocks, in according to
-\begin{eqnarray}
-\label{eq:cp_effective}C_{p_{eff}}&=&C_p+H_L\left(\frac{\partial M}{\partial T}\right)_p \\
-\label{eq:alpha_effective}\alpha_{eff}&=&\alpha+\rho\frac{H_L}{T}\left(\frac{\partial M}{\partial p}\right)_T
-\end{eqnarray}
-where $C_p$ and $\alpha$ are the specific heat and the coefficient of thermal expansion of the solid rock, respectively, and $H_L$ is the latent heat of the
-molten rock \citep{Gerya2003a,Gerya2007,Tackley2010}.
-
-\section{Benchmarks}\label{sec:benchmark}
 \subsection{Error measurements}\label{sec:error}
@@
 \frac{\left\langle \Phi \right\rangle}{Ra}\right)}\times100\]
-
-\subsection{Stokes flow}\label{sec:stokes}
-The problem consists of determining velocity field $(u,v)$ and pressure $p$ in case of a manufactured solution with prescribed body forces such as
-\begin{eqnarray}
-b_1&=&(12-24y)x^4+(-24+48y)x^3+(-48y+72y^2-48y^3+12)x^2+\nonumber \\
-&&+(-2+24y-72y^2+48y^3)x+1-4y+12y^2-8y^3\nonumber \\
-b_2&=&(8-48y+48y^2)x^3+(-12+72y-72y^2)x^2+(4-24y+48y^2-48y^3+\nonumber \\
-&&+24y^4)x-12y^2+24y^3-12y^4\nonumber
-\end{eqnarray}
-for which the exact solution is:
-\begin{eqnarray}
-u(x,y)&=&x^2(1-x)^2(2y-6y^2+4y^3)\nonumber \\
-v(x,y)&=&-y^2(1-y)^2(2x-6x^2+4x^3)\nonumber \\
-p(x,y)&=&x(1-x)-\frac{1}{6}\nonumber
-\end{eqnarray}
-The domain is a unit square a constant viscosity ($\eta=1$) and the penalty parameter is set to $\lambda=10^7$. Velocity boundary conditions are set to no slip
-$(\bm{v}=\bm{0})$ on all boundaries. The problem is performed for different grid resolution between \num{8x8} and \num{1024x1024} elements.
-
-The errors between the analytical solution and the numerical prediction of the velocity and the pressure field are calculated as in Section \ref{sec:error}.
-Fig. \ref{fig:errors}a shows that both velocity and pressure field converge to the exact solution with the decrease of the element size, following the
-theoretical convergence rate. The convergence of pressure, in contrast with the observation by \citet{Donea2003} for $Q_1 \times P_0$ elements, indicate the
-effectiveness of the smoothing procedure. Smoothed pressure field for grid resolution of \num{128x128} elements is shown in Fig. \ref{fig:errors}b (continuous
-red line), in comparison with the analytical solution (dashed blue line). Solve times and memory usage needed to generate the solution are shown in
-Fig. \ref{fig:MUMPS}. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Solver/Stokes_Flow}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/errors.pdf}
-\caption{Velocity and pressure error for the Stokes flow experiment between generated and analytical solution as a function of element size (panel a),
-and comparison between smoothed pressure and analytical solution as function of $x$ coordinate for a grid resolution of \num{128x128} elements (panel b).}
-\label{fig:errors}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/MUMPS.pdf}
-\caption{Panel a) Analysis, factorisation, solution and total solve times (panel a) and factorisation memory usage (panel b) as a function of the total number
-of degrees of freedom for the Stokes flow experiment.}
-\label{fig:MUMPS}
-\end{figure}
-
-\subsection{Markers advection}\label{sec:runge}
-The 2nd-order Runge-Kutta advection scheme in space is tested by means of the Zalesak disk test \citep{Zalesak1979,Thieulot2014}. The benchmark is performed
-in a unit square domain with a grid resolution of \num{32x32} elements and 1000000 markers and values of Courant number between 0.25 and 1.
-At $t=0$, the disk is centred at position $(0.5;0.75)$ with a radius $R=0.15$ and has a vertical fissure 0.05 wide and 0.2 high (orange in Fig. \ref{fig:runge})
-and with the velocity field prescribed in the entire domain as
-\begin{eqnarray}
-u(x,y)&=&2\pi\left(y-\frac{L_x}{2}\right)\nonumber \\
-v(x,y)&=&-2\pi\left(x-\frac{L_x}{2}\right)\nonumber
-\end{eqnarray}
-the markers are back to their initial location after a $2\pi$ rotation ($t=1$). The clockwise rotation of the disk at $t=0.25$, 0.5 and 0.75 
-is shown in Fig. \ref{fig:runge} (red, blue and green, respectively). The distance from the centre of 1 marker is calculated throughout the rotation to
-evaluate the error for different values of Courant number. As expected, the error increase with the increasing of the time step, as shown in 
-Fig. \ref{fig:runge_err}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/Zalesak_Disk}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=200px]{./Figures/Zalesak.pdf}
-\caption{Position of the Zalesak disk at $t=0$, 0.25, 0.5 and 0.75 (orange, red, blue and green, respectively) for the case with Courant number of 1.}
-\label{fig:runge}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/zal_marker.pdf}
-\caption{Distance from the centre as function of time for values of Courant number of  of 0.25, 0.3, 0.5, 0.75 and 1 (green, orange, blue, black and red,
-respectively). R indicates the distance from the centre considering a perfect circle trajectory.}
-\label{fig:runge_err}
-\end{figure}
-
-\subsection{Conservative Velocity Interpolation (CVI)}\label{sec:CVI}
-The CVI correction is checked by means of the Stokes flow experiment by \citet{Donea2003}, as presented in \ref{sec:stokes}. The advection of Lagrangian
-markers is performed using either a 2nd-order or a 4th-order Runge-Kutta scheme with CFL=0.5 and an initial random distribution of 25 markers per element.
-Fig. \ref{fig:CVI} shows the minimum and maximum values of markers per element throughout the experiment with and without the CVI correction and using
-different Runge-Kutta schemes. The introduction of the CVI correction allows the
-tests to run for times longer than $t=2000$ without show empty elements, while tests without the CVI correction stop at approximately $t=600$, when at least
-one element goes empty. Fig. \ref{fig:CVI_time} shows the different number of markers per element and the markers distribution at $t=600$ obtained without and
-with the CVI correction (panels b and c, respectively) in case of a 2nd-order Runge-Kutta scheme. All data can be found at
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/CVI}.
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/CVI.pdf}
-\caption{Maximum and minimum number of markers per element throughout the experiment, with (green and orange lines) or without (purple and blue lines)
-the CVI correction and using either a 2nd-order (blue and orange lines) or a 4th-order (purple and green lines) Runge-Kutta scheme.}
-\label{fig:CVI}
-\end{figure}
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/CVI_time.pdf}
-\caption{Number of markers per element and markers distribution through time using 2nd.order Runge-Kutta scheme.
-Panels a: number of markers per element (panel a1) and distribution of the markers (panel a2) after the first
-iteration; panels b and c: comparison between number of markers per element (panels b1 and c1) and markers distribution (panels b2 and c2) without and with
-the CVI correction at $t=600$; panels d: number of markers per element (panel d1) and markers distribution (panel d2) in case of CVI correction at $t=2000$.}
-\label{fig:CVI_time}
-\end{figure}
-
-\subsection{Poiseuille flow}\label{sec:poiseuille}
-The domain is a rectangle with $L_x=2$ and $L_y=1$ and constant density and viscosity ($\rho=1$ and $\eta=1$), gravity acceleration $\bm{g}=\bm{0}$ and penalty
-parameter $\lambda=10^8$. The grid is composed by \num{40x20} elements. Velocity boundary conditions are set to no slip $(\bm{v}=\bm{0})$ at the top and the
-bottom, and a parabolic profile is imposed on the sides, with $u=y(1-y)$ and $v=0$. The analytical solution is then given by:
-\begin{eqnarray}
-u(x,y)&=&y(1-y)\nonumber \\
-v(x,y,)&=&0\nonumber \\
-p(x,y)&=&2\eta\left(\frac{L_x}{2}-x\right)\nonumber
-\end{eqnarray}
-
-The velocity field predicted by the model follows the expected parabolic profile (Fig. \ref{fig:poiseuille}a and Fig. \ref{fig:poi_plot}a) and, in case of the
-classic penalty method with no iterations, the pressure is clearly related to the divergence by means of the penalty parameter (Fig. \ref{fig:poiseuille}b and
-c and Fig. \ref{fig:poi_plot}b). Fig. \ref{fig:poiseuille}d shows that one Uzawa iteration is sufficient to bring the divergence down to \num{e-15}, with no
-correlation with the pressure field. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Poiseuille%20Flow}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Poiseuille.pdf}
-\caption{Velocity field (panel a), pressure (panel b) and divergence velocity of a Poiseuille flow in case of the classic penalty method (no iterations) and
-after one Uzawa iteration (panel c and d, respectively).}
-\label{fig:poiseuille}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Analytical.pdf}
-\caption{Velocity (panel a) and pressure (panel b) field predicted by the model for a Poiseuille flow with respect with their analytical solutions. The
-velocity field is plotted as function of the vertical coordinate in $L_x/2$ and the pressure is plotted as function of the x coordinate in $L_y/2$.}
-\label{fig:poi_plot}
-\end{figure}
-
-\subsection{Instantaneous 2D sphere}\label{sec:ist_sphere}
-The domain is a unit square with gravity $\bm{g}=(0,-1)$. The fluid has constant density and viscosity ($\rho_f=1$ and $\eta_f=1$). The sphere is in the middle
-of the domain with a radius R = 0.123456798 and has constant density and viscosity ($\rho_s=10^{-2}+\rho_f$ and $\eta_s=10^3 \cdot \eta_f$).
-We distinguish three different types of velocity boundary conditions:
-\begin{enumerate}
-\item FS: free slip conditions on all sides.
-\item NS: no slip conditions on all sides.
-\item OT: free slip conditions on the sides and the bottom, and open boundary at the top.
-\end{enumerate}
-All three cases are performed for different grid resolutions, between \num{16x16} and \num{512x512} elements with 50 randomly distributed markers per element,
-and different average schemes for the viscosity (harmonic, geometric and arithmetic). The velocity in the centre of the sphere $\bm{v}(0.5;0.5)$), minimum and
-maximum velocities ($u_{min}$, $u_{max}$, $v_{min}$, $v_{max}$) and pressures ($p_{min}$, $p_{max}$) on the entire domain, $v_{\textrm{rms}}$ and average pressure on the
-entire domain ($p_{avg}$) are compared with solutions generated by ASPECT \citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} and all results
-can be found at \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere2D} and 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Instantaneous_2D_Sphere}. Results in terms of the $v_{\textrm{rms}}$ for NS model 
-are also shown in Fig. \ref{fig:inst_sphere}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_NS.pdf}
-\caption{$v_{\textrm{rms}}$ as function of element size for the instantaneous 2D sphere experiment in case of no slip boundary conditions and with different average
-schemes for the viscosity. Results are compared with results obtained by other numerical codes.}
-\label{fig:inst_sphere}
-\end{figure}
-
-\subsection{Rayleigh-Taylor instability}\label{sec:rayleigh}
-This problem has been originally presented by \citet{vanKeken1997} and here it is performed as the isoviscous case (Case 1a). The domain has $L_x=0.9142$ and
-$L_y=1$ and gravity $\bm{g}=(0,-1)$. Two fluids with same constant viscosities ($\eta_1=\eta_2=1$) and different densities ($\rho_1=1000$ and $\rho_2=1010$),
-with the lighter fluid at the bottom. The initial interface between the fluids is given by $y(x)=0.2+0.02 \cos \left(\frac{\pi x}{L_x}\right)$. The experiment
-is performed with different grid sizes (\num{50x50}, \num{80x80}, \num{100x100} and \num{256x256}). A total of 1960000 markers are randomly distributed at the
-beginning of the simulation. Velocity boundary conditions are set to no slip at the top and the bottom, and to free slip at the sides of the domain.
-
-$v_{\textrm{rms}}$ as function of time is reported in Fig. \ref{fig:RT}, matching well with results shown by \citet{vanKeken1997}, \citet{Tackley2003} and
-\citet{Thieulot2014}. Fig. \ref{fig:rayleigh} shows the evolution of the experiment at different time steps. All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Rayleigh_Taylor_experiment/ISOVISCOUS}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/RT.pdf}
-\caption{$v_{\textrm{rms}}$ of the Rayleigh-Taylor experiment as function of time for different resolution of the grid.}
-\label{fig:RT}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=300px]{./Figures/Rayleigh.pdf}
-\caption{Evolution of the Rayleigh-Taylor experiment for a grid of \num{256x256} elements at $t=0, 500, 1000$ and $2000$ (panels a, b, c and d, respectively).}
-\label{fig:rayleigh}
-\end{figure}
-
-\subsection{Falling blocks}\label{sec:block}
-This benchmark is described as proposed by \citet{Gerya2003a}, \citet{Gerya2010b} and \citet{Thieulot2011}. The domain is square with $L_x=L_y=$ \SI{500}{\km}
-and the grid is composed by \num{50x50} elements with 25 markers in each element. The block is initially centred at ($x=$ \SI{250}{\km}; $y=$ \SI{400}{\km})
-and has a size of \num{100x100} km (Fig. \ref{fig:block}a). The fluid surrounding the block has $\eta_f=$ \SI{e21}{\pascal\per\s} and $\rho_f=$
-\SI{3200}{\kg\per\cubic\m}. The benchmark tests with different viscosities of the block, with $\eta_b$ from \SIrange{e15}{e27}{\pascal\per\s}. In each
-experiment also the density $\rho_b$ of the block varies from \SIrange{3220}{9900}{\kg\per\cubic\m}. Velocity boundary conditions are set to free slip
-conditions on all sides of the domain. Density distributions at $t=$ \SI{20}{\mega\year} for $\rho_b=$ \SI{3300}{\kg\per\cubic\m} and different $\eta_b$
-are plotted in Fig. \ref{fig:block}b-f, showing that the code correctly preserve the block geometry in case of large viscosity contrast, with a block stiffer
-than the surrounding fluid (Fig. \ref{fig:block}f).
-
-The velocity in the centre of the falling block at $t=0$ is measured for all experiments and, since the velocity should increase with density contrast, the
-quantity $\bm{v}/(\rho_b-\rho_f)$ is plotted as function of the viscosity contrast. All results perfectly match with those from \citet{Gerya2010b} and they
-line up on a single curve, demonstrating that the code can correctly deal with large viscosity and density contrasts (Fig. \ref{fig:falling}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Falling%20blocks}.
-
-\begin{figure}
-\noindent\includegraphics[width=300px]{./Figures/Block.pdf}
-\centering
-\caption{Density evolution of the falling block experiment at $t=$ \SI{0}{\mega\year} (panel a) and $t=$ \SI{20}{\mega\year} for different viscosities of the
-block (panels b-f).}
-\label{fig:block}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Falling.pdf}
-\caption{Initial velocity relative to the density contrast at the centre of the falling block as function of the viscosity contrast between the block and the
-surrounding fluid.}
-\label{fig:falling}
-\end{figure}
-
-\subsection{2D time-dependent Stokes sphere with deformable free surface}\label{sec:sphere}
-This experiment is performed in a unit square domain with the gravity acceleration fixed to $g_y=-1$. The sphere has $\rho_s=2$ and $\eta_s=10^3$ and is
-initially centred at ($0.5;0.6$) with radius $R=0.123456789$. The fluid surrounding the sphere has $\rho_f=1$ and $\eta_f=1$ and occupies the domain for
-$y\leq 0.75$, while for $y>0.75$ the air has $\rho_a=0$ and $\eta_a=10^{-3}$. Velocity boundary conditions are set to free slip on all sides. The Courant
-number is set to 0.25. The experiments run for 200 s using grid resolutions from \num{150x150} to \num{512x512} elements, with 25 randomly distributed markers
-per element, and for different average schemes for the viscosity. The interface between the fluid and the air is tracked by means of the markers chain. Results
-in terms of velocity and pressure fields and topography variations are compared with results obtained with ASPECT
-\citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} and they can be found at
-\url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere_fs2D} and
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Time_dependent_sphere}. Results in terms of the $v_{\textrm{rms}}$ in case of 
-an arithmetic average are also shown in Fig. \ref{fig:inst_spherefs}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_arithm.pdf}
-\caption{$v_{\textrm{rms}}$ for different grid resolutions as function of time for the 2D time-dependent Stokes sphere experiment in case of an arithmetic average.
-Results are compared with results obtained by ASPECT.}
-\label{fig:inst_spherefs}
-\end{figure}
-
-\subsection{Free surface stabilisation}\label{sec:stab}
-The benchmark is performed as discussed in \citet{Kaus2010a} and \citet{Thieulot2014}. The domain is square with $L_x=L_y=$ \SI{500}{\km} and a grid resolution
-of \num{200x200} elements, each of them containing 16 markers. A buoyant fluid with $\rho_1=$ \SI{3200}{\kg\per\cubic\metre} and $\eta_1=$ \SI{e20}{\pascal\s}
-is overlain by a denser fluid with $\rho_2=$ \SI{3300}{\kg\per\cubic\metre} and $\eta_2=$ \SI{e21}{\pascal\s}. The initial interface between the fluids has an
-initial sinusoidal shape of 5 km amplitude. Velocity boundary conditions are set to free slip conditions at the sides and no slip at the bottom, while the top
-is free surface. The experiment is performed with various fixed time steps, without the use of the Courant number. The vertical position of the free surface at
-$x=L_x$ is tracked for each simulation, with or without the stabilisation algorithm.
-
-The results show an instability (drunken sailor effect) increasing the time step in case that the stabilisation algorithm is not activated
-($dt=$ \SI{4200}{\year}, green line in Fig. \ref{fig:stab}), as already observed by \citet{Kaus2010a} and \citet{Thieulot2014}. Activating the algorithm the
-instability is fixed and the simulation remains stable using time step up to \SI{50000}{\year} (red line in Fig. \ref{fig:stab}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Stabilisation_algorithm}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Kaus_random.pdf}
-\caption{Evolution of the y coordinate in $x=L_x$ as function of time for different time steps, with and without the stabilisation algorithm.}
-\label{fig:stab}
-\end{figure}
-
-\subsection{Topography relaxation}\label{sec:crameri}
-This experiment is performed as in \citet{Crameri2012} using free surface and sticky air (Case 1). The domain is rectangular with $L_x=$ \SI{2800}{\km} and
-$L_y=$ \SI{700}{\km} in case of free surface and $L_y=$ \SI{800}{\km} in case of sticky air ($h_{st}=$ \SI{100}{\km}). The grid is composed by \num{256x64}
-elements for the free surface \num{560x320} elements for the sticky air. For both cases, there is a mantle of 600 km thickness with $\eta_m=$ \SI{e21}{\pascal\s}
-overlain by a highly viscous cosine-shaped lithosphere with thickness between 93 and 107 km and $\eta_l=$ \SI{e23}{\pascal\s}. Both mediums have $\rho=$
-\SI{3300}{\kg\per\cubic\metre}. In case of sticky air experiment, the air has $\rho_a=$ \SI{0}{\kg\per\cubic\metre} and $\eta_a=$ \SI{e18}{\pascal\s}. The
-gravity acceleration is set to $g_y=$ \SI{-10}{\metre\per\square\s}. Velocity boundary conditions are set to no slip at the bottom and free slip at the sides
-of the domain. In case of sticky air, velocities on top are set to free slip conditions.
-
-The maximum topography as function of time ($h(t)$) can be analytically derived in according to \[h(t)=h_0 \exp(\gamma t)\]
-where $h_0=7$ km is the initial topography, $\gamma=\num{-0.2139e-11}$ is the characteristic relaxation rate and $t$ is the time (black line in Fig.
-\ref{fig:crameri}). Maximum topography at the characteristic relaxation time $t_{rlx}=$ \SI{14.825}{\kilo\year} can be found to be $h_{rlx}=$ \SI{2576}{\m}.
-The results show that a viscosity of the air of \SI{e18}{\pascal\s} is needed in case of $h_{st}=$ \SI{100}{\km} to correctly describe the topography relaxation
-for this problem, as pointed out by \citet{Crameri2012} (purple line in Fig. \ref{fig:crameri}). The case with a true free surface well follows the analytical
-solution (orange line in Fig. \ref{fig:crameri}). In particular, at the relaxation time in case of the true free surface predicts a maximum topography of
-\SI{2570}{\m}, with an error of \SI{6}{\m} with respect to the analytical solution. Results are also compared with results obtained with a free surface in
-UNDERWORLD (\num{256x64} elements) and SULEC (\num{401x201} elements) (with courtesy from \citealp{Crameri2012}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Topography_relaxation}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Crameri.pdf}
-\caption{Maximum topography as function of time for the Crameri benchmark (coloured lines) in comparison with the analytical solution (black line) and results
-obtained with UNDERWORLD (yellow line) and SULEC (blue line).}
-\label{fig:crameri}
-\end{figure}
-
-\subsection{Spontaneous subduction}\label{sec:subduction}
-This experiment is performed as presented by \citet{Schmeling2008}, in case of both a sticky air and a true free surface and using different average schemes
-for the viscosities. The domain is rectangular with $L_x=$ \SI{3000}{\km} and $L_y=$ \SI{700}{\km}, a grid resolution of \num{256x64} elements with 50 markers
-per element and Courant number of 0.01. At the beginning of the simulation, a 100 km-thick lithospheric layer with $\rho_m=$ \SI{3300}{\kg\per\cubic\metre}
-and $\eta_m=$ \SI{e23}{\pascal\s} is located at the top of the domain for $x$ between 1000 and 3000 km, while the asthenospheric mantle has $\rho_m=$
-\SI{3200}{\pascal\s} and $\eta_m=$ \SI{e21}{\pascal\s}. In addition, a 200 km-depth lithospheric slab is already subducted in the mantle in order to have
-a spontaneous subduction. In case of sticky air, $L_y=750$ km and a 50 km-thick air layer with $\eta_a$ from \SIrange{e19}{e21}{\pascal\s} overlie the mantle.
-Velocity boundary conditions are set to free slip on the sides and on the bottom of the domain. In case of sticky air, velocities on the top boundary are set
-to free slip conditions as well.
-
-The maximum depth of the slab is tracked for all the simulations (Fig. \ref{fig:subduction}) and they are compared with results obtained by
-\citet{Schmeling2008} with models with regular grids and comparable grid resolutions (rectangles in Fig. \ref{fig:subduction}). Low viscous sticky air models
-(i.e., \SI{e19}{\pascal\s}) enlighten a strong dependence of the sinking velocity with the chosen average scheme (dashed lines in Fig. \ref{fig:subduction}),
-while high viscous sticky air models (i.e., \SI{e21}{\pascal\s}) show a higher resistance at the trench, underestimating the correct solution (dotted lines in
-Fig. \ref{fig:subduction}). A high resistance at the trench can be also observed in case of a true free surface for the low resolution grid used (continuous
-lines in Fig. \ref{fig:subduction}), as pointed out by \citet{Schmeling2008}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Spontaneous_subduction}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Subduction.pdf}
-\caption{Maximum depth of the slab as function of time for all the simulations of the \citet{Schmeling2008} benchmark, in case of harmonic, geometric and
-arithmetic means (red, green and blue lines, respectively) using a sticky air or a true free surface (discontinuous and continuous lines, respectively).
-Red, green and blue rectangular areas indicate the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{500}{\km} in case of sticky air,
-while the grey rectangular area indicates the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{600}{\km} in case of true free surface.}
-\label{fig:subduction}
-\end{figure}
-
-\subsection{The slab detachment}\label{sec:slab}
-The slab detachment benchmark is performed as described by \citet{Schmalholz2011} and \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{1000}{\km}
-and $L_y=$ \SI{660}{\km} and a grid resolution of \num{256x256} elements. A non-linear viscous T-shaped layer with $\rho_l=$ \SI{3300}{\kg\per\cubic\metre} is
-placed at the top of the domain and surrounded by a linear viscous fluid with $\rho_f=$ \SI{3150}{\kg\per\cubic\m}. The top layer is 80 km-thick and a 250
-km-long and 80 km-wide slab is placed at $x=L_x/2$ (Fig. \ref{fig:slab}a). The effective viscosity of the top layer is given by
-\[\eta_{eff}=\eta_0 I_2^{\frac{(1-n)}{n}}\]
-with $\eta_0=$ \SI{4.75e11}{\pascal\s} and $n=4$, while the surrounding fluid has $\eta_f=$ \SI{e21}{\pascal\s}. The effective viscosity is capped between
-$\eta_{min}=$ \SI{e21}{\pascal\s} and $\eta_{max}=$ \SI{e25}{\pascal\s}. Velocity boundary conditions are set to free slip at the top and the bottom, and to
-no slip at the sides of the domain.
-
-The sinking of the slab determines a decrease of the effective viscosity in the T-shaped layer during the first part of the evolution (Fig. \ref{fig:slab}a),
-as consequence of local strain rates, while effective viscosities increase up to $\eta_{max}$ in the top layer after approximately \SI{20}{\mega\year}, when
-necking is complete (Fig. \ref{fig:slab}b).
-
-The width of necking is tracked by means of markers position at the side of the slab. Necking width and time are normalised in function of the initial slab
-(\SI{80}{\km}) and the characteristic time ($t_c=$ \SI{7.1158e14}{\s}) \citep{Schmalholz2011,Glerum2018}. The results in case of arithmetic, geometric and
-harmonic mean of the viscosity are shown in Fig. \ref{fig:necking}, compared with results from \citet{Glerum2018} for which an infinity norm mean was used.
-The influence of the average schemes agrees with results shown in Section \ref{sec:subduction}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Slab_detachment}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Slab.pdf}
-\caption{Effective viscosity for the slab detachment benchmark at the beginning of the evolution (panel a) and when necking is complete (panel b).}
-\label{fig:slab}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Necking.pdf}
-\caption{Normalised width of the necking of the slab detachment benchmark as function of normalised time.}
-\label{fig:necking}
-\end{figure}
-
-\subsection{The indenter experiment}\label{sec:indenter}
-The indenter benchmark simulate a rigid punch on a purely plastic von Mises material, for which exists an analytical solution
-\citep{Thieulot2008,Thieulot2014,Glerum2018}. The domain is a unit square with a grid resolution of \num{256x256} elements. The gravity acceleration is set
-to $\bm{g}=\bm{0}$ and effective viscosity is capped using $\eta_{min}=10^{-4}$ and $\eta_{max}=10^3$. The tolerance for the convergence of the non-linear
-solution is $tol=10^{-9}$, with a maximum number of non-linear iterations set to 500. The medium in the domain has $\rho=0.01$, an initial viscosity
-$\eta_0=10$, cohesion $C=1$ and an angle of internal friction $\phi=0$°. Velocity boundary conditions are set to no slip at the bottom and free slip at the
-sides of the domain. The top of the domain is open with the exception of the central portion (punch area) with width $w_p=0.125$, where $v=-1.05$ and $u$ is
-fixed either to 0 (rough punch experiment) or free (smooth punch experiment).
-
-Fig. \ref{fig:indenter} show the results in terms of viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure
-(panels g and h) for both experiments. The analytical solution indicate that pressure at ($0.5;1$) and ($0.5 \pm w_p;1$)is $p=\pi +1$ and $p=1$, respectively
-(black and red lines in Fig. \ref{fig:smooth_p}). Fig. \ref{fig:smooth_p} shows a clear improvement in the pressure solution when passing from a rough to a
-smooth punch (green and blue lines in Fig. \ref{fig:smooth_p}, respectively), as previously observed by \citet{Thieulot2014} and \citet{Glerum2018}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Indenter}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Indenter.pdf}
-\caption{Viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure (panels g and h) fields for rough (left column) and
-smooth (right column) punch experiments.}
-\label{fig:indenter}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Smooth_Rough.pdf}
-\caption{Nodal pressure in function of the x coordinate at $y=1$ for rough (green line) and smooth (blue line) punch experiment. Black and red lines indicate
-the analytical solution at $x=0.5$ and $x=0.5 \pm w_p$, respectively.}
-\label{fig:smooth_p}
-\end{figure}
-
-\subsection{The brick experiment}\label{sec:brick}
-An instantaneous version of the brick benchmark is performed to verify the correctness of the pressure-dependent plasticity for different angle of internal
-friction, as proposed by \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{40}{\km} and $L_y=$ \SI{10}{\km} and a grid resolution of \num{512x128}
-elements. The gravity acceleration is set to $g_y=$ \SI{-10}{\m\square\s} and effective viscosity is capped using $\eta_{min}=$ \SI{e19}{\pascal\s} and
-$\eta_{max}=$ \SI{e26}{\pascal\s}. The tolerance for the convergence of the non-linear solution is $tol=10^{-7}$, with a maximum number of non-linear iterations
-set to 1000. A 800 m-wide and 400 m-height inclusion with $\eta_b=$ \SI{e20}{\pascal\s} and $\rho_b=$ \SI{2700}{\kg\per\cubic\m} is placed at the bottom of the
-domain at $x=L_x/2$. The inclusion is surrounded by a non-linear viscous medium with $\rho_m=$ \SI{2700}{\kg\per\cubic\m}, an initial viscosity of
-\SI{e23}{\pascal\s}, a linear viscous viscosity of \SI{e25}{\pascal\s} and a cohesion of \SI{40}{\mega\pascal}. Velocities are set to free slip conditions at
-the bottom of the domain and the top is open. Velocities on sides of the domain are fixed to $u=$ \SI{+-2e-11}{\m\per\s} and $v=0$. The experiment is performed
-in compressional and extensional contexts with an angle of internal friction $\phi$ of the non-linear medium variable between 0° and 30°.
-
-As expected, two shear bands stem from the inclusion with variable angles in relation to both the dynamics context and the internal friction angle. Shear band
-angles formed at 45° for $\phi=0$° in both extensional and compressional contexts (Fig. \ref{fig:brick_beam}a), while they are different in case of $\phi\neq0$°.
-Shear band angles in case of $\phi=20$° are shown for compression and extension in Fig. \ref{fig:brick_beam}b  and c, respectively. Values of shear band angles
-as function of internal friction angles from 0° to 30° are extracted at different depths and minimum and maximum values are plotted in Fig. \ref{fig:brick},
-in comparison with theoretical Roscoe, Arthur and Coulomb shear band angles. For all tests the tolerance for nonlinear convergence is defined $tol=10^{-7}$
-and none converges before the maximum number of iterations ($it_{max}=1000$) is reached. However, tests with internal friction angles up to 15° show a constant
-decrease of the velocity residuals, which can not be observed for for tests with higher internal friction angles (Fig. \ref{fig:convergence}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Brick_experiment}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Brick_Beam.pdf}
-\caption{Shear band angles predicted for the brick experiment for $\phi=0$° (panel a) and for $\phi=20$° in case of compression and extension (panels b and c,
-respectively).}
-\label{fig:brick_beam}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Brick.pdf}
-\caption{Shear band angles predicted for the brick experiment in case of compressional and extensional contexts as function of different internal angle of
-friction (continuous black line and red dots), compared with theoretical Roscoe, Artur and Coulomb angles (discontinuous black lines). Red lines indicate the
-range of angles calculated at different depths.}
-\label{fig:brick}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \noindent\includegraphics[width=500px]{./Figures/convergence.pdf}
-  \caption{Convergence rates for all tests. Normalised residuals (left column) are compared with velocity residuals (right column) for extensional (first row)
-  and compressional (second row) settings.}
-  \label{fig:convergence}
-  \end{figure}
-
-\subsection{Advection stabilisation}\label{sec:advection}
-The 1D advection problem proposed by \citet{Donea2003} and \citet{Thieulot2011} is performed to verify the effectiveness of the SUPG method to stabilise the
-advection term of the energy equation. The domain is a 1D segment with $L_x=1$ composed by 50 elements and a discontinuity in the thermal field placed at
-$x=0.25$. Temperature is set to 1 for $x \leq 0.25$ and to 0 for $x > 0.25$. Velocity is set to ${u}=1$ in the entire domain. The simulation is performed for
-250 time steps, with $dt=0.002$, so the thermal discontinuity should be at $x=0.75$ at the end of the simulation.
-
-Temperature profiles at the end of the simulation are shown in \ref{fig:advection} as function of the dimensionless coefficient $\gamma=\tau \bm{v}/h$. In case
-of the classic Galerkin method ($\gamma=0$, blue line in \ref{fig:advection}) the final thermal profile is characterised by strong oscillations, which are
-eliminated in case of the SUPG method ($\gamma=0.045$, orange line in \ref{fig:advection}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Advection_stabilisation}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Advection.pdf}
-\caption{Temperature profile as function of \textit{x} for the advection stabilisation benchmark. Purple line indicates the initial temperature profile; the
-green line indicates the analytical temperature profile after 250 time steps; blue line indicates the temperature profile after 250 time steps in case of the
-classic Galerkin method ($\gamma=0$); orange line indicates the temperature profile after 250 time steps in case of the SUPG method ($\gamma=0.045$).}
-\label{fig:advection}
-\end{figure}
-
-\subsection{Shear heating}\label{sec:simple_shear}
-This exercise is performed in an unit square domain composed by \num{128x128} elements. The velocity field is prescribed on the entire domain with
-$\bm{v}=(L_y-y)y\bm{e_x}$; viscosity, density and specific heat are fixed to 1, while thermal conductivity and radiogenic energy are fixed to 0. Therefore,
-the energy equation (Eq. \ref{eq:energy}) can be simplified as \[\frac{\partial T}{\partial t}=H_s\] and fixing $T(t=0)=0$, the temperature can be find as
-\[T(t)=H_s t\]
-In this case we have
-\begin{eqnarray}
-\dot{\epsilon_{xx}}&=&\frac{\partial u}{\partial x}=0 \nonumber \\
-\dot{\epsilon_{yy}}&=&\frac{\partial v}{\partial y}=0 \nonumber \\
-\dot{\epsilon_{xy}}&=&\frac{1}{2}\left(\frac{\partial v}{\partial x}+\frac{\partial u}{\partial y}\right)=\frac{1}{2}(L_y-2y)\nonumber
-\end{eqnarray}
-and, simplifying Eq. \ref{eq:shear}, shear heating can be calculated as \[H_s(x,y)=(1-2y)^2\]
-
-The solution predicted by the model in terms of velocity, temperature and shear heating match well with the analytical solutions (Fig. \ref{fig:analytical_en}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Simple_shear}.
-
-%\begin{figure}
-%\centering
-%\noindent\includegraphics[width=400px]{./Figures/Energy_Thieulot.pdf}
-%\caption{Temperature and shear heating predicted by the model after 1 time step.}
-%\label{fig:en_thieulot}
-%\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=500px]{./Figures/Analytical_Shear.pdf}
-\caption{Velocity (panel a), temperature (panel b) and shear heating (panel c) as function of the \textit{y} coordinate for the simple shear experiment. The
-solutions predicted by the model (red lines) are compared with the analytical solutions (dashed blue lines).}
-\label{fig:analytical_en}
-\end{figure}
-
-\subsection{Shear and adiabatic heating}\label{sec:shear}
-This problem is performed as presented in Exercise 9.4 in \citet{Gerya2010b}. Two materials are vertically separated in a rectangular domain with $L_x=$
-\SI{1000}{\km}, $L_y=$ \SI{1500}{\km} and a grid resolution of \num{30x20} elements. Constant thermal coefficient expansion ($\alpha=$ \SI{3e-5}{\per\kelvin}),
-temperature ($T=$ \SI{1300}{\kelvin}) and gravity acceleration ($g_y=$ \SI{-10}{\m\square\s}) are assumed in the whole domain. Fluid 1 (on the left side of the
-domain) has $\rho_1=$ \SI{3200}{\kg\per\cubic\m} and $\eta_1=$ \SI{e20}{\pascal\s}; fluid 2 (on the right side of the domain) has $\rho_2=$
-\SI{3300}{\kg\per\cubic\m} and $\eta_2=$ \SI{e22}{\pascal\s}. Velocity boundary conditions are set to free slip on all sides of the domain.
-
-As shown in Fig. \ref{fig:energy}, both shear and adiabatic heating predicted by the code (first row in Fig. \ref{fig:energy}) well recreate the results
-obtained by means of example ${Shear\_adiabatic\_heating.m}$ from \citet{Gerya2010b} (second row in Fig. \ref{fig:energy}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Adiabatic%2BShear}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Energy.pdf}
-\caption{Comparison between shear (first column), adiabatic (second column) and total (third column) energy predicted by the code (panels a, b and c) and those
-created using example ${Shear\_adiabatic\_heating.m}$ from Exercise 9.4 in \citet{Gerya2010b} (panels d, e and f).}
-\label{fig:energy}
-\end{figure}
-
-\subsection{Mantle convection}\label{sec:mantle}
-This problem is performed as presented by \citet{Blankenbach1989} (constant viscosity cases) and \citet{Thieulot2014}, in a 2D unit square domain with gravity
-acceleration $g_y=10^{10}Ra$. The experiment is performed with three different Rayleigh numbers ($Ra=10^4$, $10^5$ and $10^6$) and with different grid
-resolution (between \num{32x32} and \num{128x128} elements). The fluid has constant viscosity, initial density, heat capacity, thermal conductivity
-($\eta= \rho_0= C_p = k=1$), reference temperature ($T_0=0$) and thermal expansion coefficient ($\alpha=10^{-10}$). Temperatures are set to 0 on top and 1 on
-bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\]
-
-The solution generated by the code in terms of $v_{\textrm{rms}}$ (calculated as in Sec. \ref{sec:error}) and Nu as function of time are reported for all the
-simulations in Fig. \ref{fig:mantle}. At the steady state, $v_{\textrm{rms}}$ and Nu of all simulations converge well toward the values from \citet{Blankenbach1989},
-with lower errors for higher resolution grids (Table \ref{tab:mantle}). All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Mantle_convection}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Convection.pdf}
-\caption{$v_{\textrm{rms}}$ (panels a, c and e) and Nu (panels b, d and f) for the mantle convection benchmark as function of time for different grid resolution.
-Panels a and b show the results for $Ra=10^4$; Panels c and d show the results for $Ra=10^5$; panels e and f show the results for $Ra=10^6$. Purple lines
-indicate the convergence values for the $v_{\textrm{rms}}$ and the Nu from \citet{Blankenbach1989} at the steady state.}
-\label{fig:mantle}
-\end{figure}
-
-\begin{table}[H]
-\caption{Comparison between $v_{\textrm{rms}}$ and Nu predicted by the code for the mantle convection experiment and same values as reported in literature.}
-\centering
-\small
-\begin{tabular}{|>{\centering}m{0.09\textwidth}|>{\centering}m{0.05\textwidth}|>{\centering}m{0.24\textwidth}|>{\centering}m{0.15\textwidth}|
-  >{\centering}m{0.1\textwidth} >{\centering}m{0.1\textwidth} >{\centering\arraybackslash}m{0.1\textwidth}|}
-\toprule
-& & \multirow{2}{*}{\citet{Blankenbach1989}} & \citet{Thieulot2014} & \multicolumn{3}{c|}{FALCON}  \\
- & & & \num{200x200} & \num{32x32} & \num{64x64} & \num{128x128}  \\
-\midrule
-\multirow{2}{*}{$Ra=10^4$} & $v_{\textrm{rms}}$ & $42.864947 \pm 0.000020$ & 42.867 & 42.83226 & 42.852793 & 42.861394  \\
-           & Nu      & $4.884409 \pm 0.000010$  & 4.882  & 4.781297 & 4.857475  & 4.877573   \\
-\hline
-\multirow{2}{*}{$Ra=10^5$} & $v_{\textrm{rms}}$ & $193.21454 \pm 0.00010$  & 193.255& 193.872643 & 193.377472 & 193.252290  \\
-           & Nu      & $10.534095 \pm 0.000010$ & 10.507 & 9.602514 & 10.270735  & 10.465629   \\
-\hline
-\multirow{2}{*}{$Ra=10^6$} & $v_{\textrm{rms}}$ & $833.98977 \pm 0.00020$  & 834.712& 848.091176 & 837.767911 & 834.945793  \\
-           & Nu      & $21.972465 \pm 0.000020$ & 21.695 & 15.999266 & 19.703682  & 21.306939   \\
-\bottomrule
-\end{tabular}
-\label{tab:mantle}
-\end{table}
-
-\subsection{Viscoplastic mantle convection}\label{sec:visco_mantle}
-This benchmark is performed as Cases 1-5a described in \citet{Tosi2015} in a 2D unit square domain with two different grid resolutions of \num{32x32} 
-and \num{100x100} elements. Reference viscosity, reference density, heat capacity, thermal conductivity and thermal expansion coefficient are fixed to a constant
-value of 1 ($\eta_0= \rho_0= C_p = k=\alpha=1$), while the gravity has been chosen as $g_y=10^2$, in order to obtain a Rayleigh number $Ra=10^2$. Temperatures
-are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\].
-The viscosity ﬁeld $\eta$ is calculated as the harmonic average between a linear part $\eta_{lin}$ that depends on temperature only or on temperature and 
-depth $y$ and a nonlinear-plastic part $\eta_{pl}$ that depends on the strain rate $\dot{\bm{\epsilon}}$, as follows
-\[\eta(T,y,\dot{\bm{\epsilon}})=2\left(\frac{1}{\eta_{lin}(T,y)}+\frac{1}{\eta_{pl}(\dot{\bm{\epsilon}})}\right)^{-1}\]
-The linear and the nonlinear-plastic parts of the viscosity are calculated following \citet{Tosi2015} as follows
-\[\eta_{lin}(T,y)=\exp(-\ln(\Delta\eta_T)T+\ln(\Delta\eta_y)y)\]
-\[\eta_{pl}(\dot{\bm{\epsilon}})=\eta^*+\frac{\sigma_Y}{\sqrt{\dot{\bm{\epsilon}}:\dot{\bm{\epsilon}}}}\]
-where $\Delta\eta_T$, $\Delta\eta_y$, $\eta^*$ and $\sigma_Y$ are parameters chosen as listed in Table \ref{tab:tosi_setup}.
-
-For each case, some diagnostic quantities measured at the steady state are reported in Table \ref{tab:visco_mantle} (see Section \ref{sec:error} for 
-the computation of all the quantities), showing a well fit with results described in \citet{Tosi2015}. Data and charts for all cases can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Viscoplastic_mantle_convection}.
-
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{white!100!}{gray!25!}
-%\begin{table}[H]
-%  \caption{Results of the viscoplastic mantle convection benchmark for Cases 1-4 compared with results obtained with ELEFANT and YACC (from \citealp{Tosi2015}).}
-%  \centering
-%  \resizebox{\textwidth}{!}{
-%\begin{tabular}{|c|ccc|ccc|ccc|ccc|}
-%  \toprule
-%    & \multicolumn{3}{c|}{\textbf{Case 1}} & \multicolumn{3}{c|}{\textbf{Case 2}} & \multicolumn{3}{c|}{\textbf{Case 3}} &
-%    \multicolumn{3}{c|}{\textbf{Case 4}} \\
-%   & FALCON & ELEFANT& YACC & FALCON& ELEFANT& YACC & FALCON & ELEFANT & YACC & FALCON & ELEFANT & YACC \\
-%  \midrule
-%   $\left\langle T \right\rangle$ & 0.7759 & 0.7758 & 0.7767 & 0.6043 & 0.6036 & 0.6058 & 0.7275 & 0.7275 & 0.7286 & & 0.5277 & 0.5289\\
-%    Nu$^{top}$ & 3.4280 & 3.4214 & 3.4298 & 8.5242 & 8.5115 & 8.5278 & 3.0368 & 3.0347 & 3.0374 && 6.5912 & 6.5572 \\
-%    Nu$^{bottom}$ & 3.3168 & 3.313 & 3.3143 & 8.4599 & 8.4465 & 8.3990 & 3.0086 & 2.9908 & 2.9628 && 6.5834 & 6.5243 \\
-%    $v_{\textrm{rms}}$ & 249.901 & 249.134 & 251.7997 & 141.2451 & 140.7087 & 142.2020 & 100.2776 & 100.1208 & 100.9467 && 79.1105 & 79.6202 \\
-%    $v_{\textrm{rms}}^{top}$ & 1.8907 & 1.8642 & 1.8298 & 104.7928 & 104.1228 & 105.1767 & 2.0832 & 2.0652 & 2.0374 && 74.7596 & 75.8414 \\
-%    $u_{max}^{top}$ & 2.6340 & 2.6119 & 2.5516 & 121.9212& 121.766 & 122.3238 & 2.9086 & 2.9019 & 2.8458 && 88.9146 & 89.2940 \\
-%    $\eta_{min}$ & - & - & - & $2.0303\times10^{-5}$ & $2.0600\times10^{-5}$ & $1.9853\times10^{-5}$ & $4.8129\times10^{-5}$ & $4.8080\times10^{-5}$ & $4.7907\times10^{-5}$ &
-%    & $1.9860\times10^{-4}$ & $1.9174\times10^{-4}$ \\
-%    $\eta_{max}$ & - & - & - & 1.3325 & 1.85 & 1.6095 & 0.9360 & 0.9023 & 1 && 1.52 & 1.6773 \\
-%    $\left\langle W \right\rangle$ & 2.4299 & 2.4316 & 2.4583 & 7.5648 & 7.5665 & 7.6222 & 2.0387 & 2.0384 & 2.0400 && 5.6216 & 5.6512 \\
-%    $\left\langle \Phi \right\rangle/Ra$ & 2.4311 & 2.4276 & 2.4333 & 7.5661 & 7.5585 & 7.6194 & 2.0386 & 2.0370 & 2.0335 && 5.6182 & 5.6463 \\
-%    $\delta$ & 0.05\% & 0.16\% & 1.02\% & 0.02\% & 0.11\% & 0.04\% & $<$0.01\% & 0.07\% & 0.32\% && 0.06\% & 0.09\% \\
-%  \bottomrule
-%  \end{tabular}}
-%  \label{tab:visco_mantle}
-%  \end{table}
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{2}{white!100!}{gray!25!}
-\begin{table}[H]
-  \caption{Benchmark cases and corresponding parameters.}
-  \centering
-  \small
-\begin{tabular}{|>{\centering}m{0.08\textwidth}|>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}
-  >{\centering}m{0.05\textwidth} >{\centering}m{0.05\textwidth} |>{\centering\arraybackslash}m{0.2\textwidth}|}
-  \toprule
-    Case & $Ra$ & $\Delta\eta_T$ & $\Delta\eta_y$ & $\eta^*$ & $\sigma_Y$ & Convective regime \\
-  \midrule
-  \rowcolor{gray!25!}
-    1 & $10^2$ & $10^5$ & 1 & - & - & Stagnant lid \\
-    2 & $10^2$ & $10^5$ & 1 & $10^{-3}$ & 1 & Mobile lid  \\
-    \rowcolor{gray!25!}
-    3 & $10^2$ & $10^5$ & 10 & - & - & Stagnant lid  \\
-    4 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 1 & Mobile lid \\
-    \rowcolor{gray!25!}
-    5 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 4 & Periodic \\
-  \bottomrule
-  \end{tabular}
-  \label{tab:tosi_setup}
-  \end{table}
-
-\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{gray!25!}{white!100!}
-\begin{table}
-  \caption{Results of the viscoplastic mantle convection benchmark for different grid resolutions in Cases 1-5 compared with results obtained with ELEFANT
-  and YACC ($100\times100$ elements, from \citealp{Tosi2015}) and in Stone 28 ($32\times32$ elements, from
-  \url{https://github.com/cedrict/fieldstone/tree/master/python_codes/fieldstone_28}).}
-  \centering
-  \resizebox{\textwidth}{!}{
-\begin{tabular}{|cc|ccccccccccc|}
-  \toprule
-  \multicolumn{13}{|c|}{\textbf{Case 1}} \\
-  \midrule
-  Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-  & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-  \hline
-  \rowcolor{gray!25!}
-    YACC & $100\times100$ & 0.7767 & 3.4298 & 3.3143 & 251.7997 & 1.8298 & 2.5516 & - & - & 2.4583 & 2.4333 & 1.02\% \\
-    ELEFANT & $100\times100$ & 0.7758 & 3.4214 & 3.313 & 249.134 & 1.8642 & 2.6119 & - & - & 2.4316 & 2.4276 & 0.16\% \\
-    \rowcolor{gray!25!}
-    Stone 28 & $32\times32$ & 0.7737 & 3.3987 & - & 243.872 & - & - & - & - & - & - & - \\
-%    FALCON & $32\times32$ & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & - & - & 2.4745 & 2.4864 & 0.48\% \\
-%    \rowcolor{gray!25!}
-    FALCON & $100\times100$ & 0.7759 & 3.4280 & 3.3168 & 249.901 & 1.8907 & 2.6340 & - & - & 2.4299 & 2.4311 & 0.05\% \\
-    \midrule
-    \multicolumn{13}{|c|}{\textbf{Case 2}} \\
-    \midrule
-    Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-    & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-    \hline
-    \rowcolor{gray!25!}
-     YACC & $100\times100$ & 0.6058 & 8.5278 & 8.3990 & 142.2020 & 105.1767 & 122.3238 & $1.9853\times10^{-5}$ & 1.6095 & 7.6222 & 7.6194 & 0.04\% \\
-     ELEFANT & $100\times100$ & 0.6034 & 8.5115 & 8.4465 & 140.8067 & 104.1228 & 121.766 & $2.0600\times10^{-5}$ & 1.85 & 7.5665 & 7.5585 & 0.11\% \\
-     \rowcolor{gray!25!}
-     Stone 28 & $32\times32$ & 0.6052 & 8.1715 & - & 141.518 & - & - & - & - & - & - & - \\
-%     FALCON & $32\times32$ & 0.6454 & 8.1422 & 7.4020 & 156.6908 & 95.7226 & 116.9302 & $2.2572\times10^{-5}$ & 0.7194 & 7.4289 & 7.4452 & 0.22\% \\
-%     \rowcolor{gray!25!}
-     FALCON & $100\times100$ & 0.6044 & 8.5266 & 8.4590 & 141.3684 & 104.8867 & 122.0470 & $2.0333\times10^{-5}$ & 1.2914 & 7.5707 & 7.5718 & 0.01\% \\
-     \midrule
-     \multicolumn{13}{|c|}{\textbf{Case 3}} \\
-     \midrule
-     Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-     & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-     \hline
-     \rowcolor{gray!25!} 
-     YACC & $100\times100$ & 0.7286 & 3.0374 & 2.9628 & 100.9467 & 2.0374 & 2.8458 & $4.7907\times10^{-5}$ & 1 & 2.0400 & 2.0335 & 0.32\% \\
-      ELEFANT & $100\times100$ & 0.7275 & 3.0347 & 2.9908 & 100.1208 & 2.0652 & 2.9019 & $4.8080\times10^{-5}$ & 0.9023 & 2.0384 & 2.037 & 0.07\% \\
-      \rowcolor{gray!25!}
-      Stone 28 & $32\times32$ & 0.7277 & 3.0347 & - & 100.018 & - & - & - & - & - & - & - \\
-%      FALCON & $32\times32$ & 0.7281 & 3.0618 & 2.8132 & 101.0092 & 2.1630 & 3.0178 & $4.8500\times10^{-5}$ & 0.7845 & 2.0699 & 2.0695 & 0.02\% \\
-%      \rowcolor{gray!25!}
-      FALCON & $100\times100$ & 0.7275 & 3.0368 & 3.0086 & 100.2776 & 2.0832 & 2.9086 & $4.8129\times10^{-5}$ & 0.9360 & 2.0387 & 2.0386 & $<$0.01\% \\
-      \midrule
-      \multicolumn{13}{|c|}{\textbf{Case 4}} \\
-      \midrule
-      Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-      & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-      \hline
-      \rowcolor{gray!25!} 
-      YACC & $100\times100$ & 0.5289 & 6.5572 & 6.5243 & 79.6202 & 75.4814 & 89.2940 & $1.9174\times10^{-4}$ & 1.6773 & 5.6512 & 5.6463 & 0.09\% \\
-       ELEFANT & $100\times100$ & 0.5277 & 6.5912 & 6.5834 & 79.1105 & 74.7596 & 88.9146 & $1.9860\times10^{-4}$ & 1.5200 & 5.6216 & 5.6182 & 0.06\% \\
-       \rowcolor{gray!25!}
-       Stone 28 & $32\times32$ & 0.5286 & 6.4036 & - & 79.4746 & - & - & - & - & - & - & - \\
-%       FALCON & $32\times32$ & 0.5334 & 6.4679 & 6.3435 & 80.6153 & 76.0511 & 90.2058 & $2.0393\times10^{-4}$ & 0.6453 & 5.6676 & 5.6698 & 0.04\% \\
-%       \rowcolor{gray!25!}
-       FALCON & $100\times100$ & 0.5279 & 6.5897 & 6.5830 & 79.1220 & 75.0715 & 88.8484 & $1.9693\times10^{-4}$ & 1.4133 & 5.6127 & 5.6113 & 0.03\% \\
-       \midrule
-       \multicolumn{13}{|c|}{\textbf{Case 5}} \\
-       \midrule
-       Code & Resolution & $P$ & $\left\langle T \right\rangle_{min}$ & $\left\langle T \right\rangle_{max}$ & Nu$^{top}_{min}$ & Nu$^{top}_{max}$ &
-       $v_{\textrm{rms}}^{min}$ & $v_{\textrm{rms}}^{max}$ & \multicolumn{2}{c}{$(\left\langle \Phi \right\rangle/Ra)_{min}$} & 
-       \multicolumn{2}{c|}{$(\left\langle \Phi \right\rangle/Ra)_{max}$} \\
-       \hline
-       \rowcolor{gray!25!}
-         YACC & $100\times100$ & 0.0792 & 0.6543 & 0.6722 & 2.6950 & 7.2792 & 41.9813 & 98.8230 & \multicolumn{2}{c}{1.4347} & 
-         \multicolumn{2}{c|}{9.2748} \\
-         ELEFANT & $100\times100$ & 0.076 & 0.6514 & 0.6693 & 2.6808 & 7.3506 & 41.2578 & 100.089 & \multicolumn{2}{c}{1.3324} & 
-         \multicolumn{2}{c|}{9.5289} \\
-         \rowcolor{gray!25!}
-         Stone 28 & $32\times32$ & 0.0764 & 0.6521 & 0.6697 & 2.6912 & 7.0763 & 41.61 & 98.183 & \multicolumn{2}{c}{-} & 
-         \multicolumn{2}{c|}{-} \\
-     %    FALCON & $32\times32$ & & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & \multicolumn{2}{c}{-} & 
-     %     \multicolumn{2}{c|}{-} \\
-     %    \rowcolor{gray!25!}
-         FALCON & $100\times100$ & 0.0775 & 0.6519 & 0.6697 & 2.6847 & 7.3629 & 41.5256 & 99.5466 & \multicolumn{2}{c}{1.3399} & 
-         \multicolumn{2}{c|}{9.3993} \\
-  \bottomrule
-  \end{tabular}}
-  \label{tab:visco_mantle}
-  \end{table}
-
-\subsection{Thin layer entrainment}\label{sec:thin}
-This experiment is performed as originally proposed by \citet{vanKeken1997} with two incompressible fluids in a rectangular domain with $L_x=2$ and $L_y=1$
-and gravity acceleration $g_y=10^{10}Ra$. The Rayleigh number and the compositional Rayleigh number are fixed to $Ra=300000$ and $Rc=450000$, respectively.
-Both fluids have constant viscosity, thermal conductivity, specific heat ($\eta=\rho=C_p=1$) and thermal expansion coefficient ($\alpha=10^{-10}$). Fluid 1
-has a density $\rho_1=1$, while fluid 2 is denser ($\rho_2=\rho_1+1.5 \cdot 10^{-10}$) and is placed at the bottom of the domain, for $y \leq 0.025$.
-Temperature are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides of the domain. The initial
-temperature field is given by
-\[T(x,y)=T_u(x,y)+T_l(x,y)+T_r(x,y)+T_s(x,y)-\frac{3}{2}\]
-with
-\begin{eqnarray}
-T_u(x,y)&=&\frac{1}{2}erf\left(\frac{1-y}{2}\sqrt{\frac{u_0}{x}}\right)\nonumber \\
-T_l(x,y)&=&1-\frac{1}{2}erf\left(\frac{y}{2}\sqrt{\frac{u_0}{L_x-x}}\right)\nonumber \\
-T_r(x,y)&=&\frac{1}{2}+\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{y+1}} \exp\left(-\frac{x^2u_0}{4y+4}\right)\nonumber \\
-T_s(x,y)&=&\frac{1}{2}-\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{2-y}} \exp\left(-\frac{(L_x-x)^2u_0}{8-4y}\right)\nonumber
-\end{eqnarray}
-and
-\begin{eqnarray}
-u_0&=&\frac{L_x^{7/3}}{(1+L_x^4)^{2/3}}\left(\frac{Ra}{2\sqrt{\pi}}\right)^{\frac{2}{3}}\nonumber \\
-Q&=&2\sqrt{\frac{L_x}{\pi u_0}}\nonumber
-\end{eqnarray}
-
-The experiment is performed with two grid resolution (\num{125x40} and \num{200x80} elements), with 100 markers per element and CFL of 0.25. $v_{\textrm{rms}}$
-(see Sec. \ref{sec:error}) calculated for both the simulations match well with results from \citet{vanKeken1997} and obtained with ASPECT
-\citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} and ELEFANT \citep{Thieulot2014}, obtained with similar grid resolutions
-(Fig. \ref{fig:thin}). In addition, a variety of simulations with different resolutions and aspect ratios are performed to verify that the sensitivity of the
-initial velocity field. Results match well with results from \citet{vanKeken1997} (in the grey area) and \citet{Thieulot2014} (Fig. \ref{fig:thin_initial}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Thin_layer}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_thin.pdf}
-\caption{$v_{\textrm{rms}}$ for the thin layer experiment as function of time for different grid resolution. Results are compared with results obtained by
-\citet{vanKeken1997} (black lines) and with ASPECT and ELEFANT.}
-\label{fig:thin}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrmszero.pdf}
-\caption{$v_{\textrm{rms}}$ at $t=0$ for the thin layer experiment as function of the element size for different aspect ratios of the elements. Results are compared
-with results obtained by \citet{vanKeken1997} (black dots in the grey area) and with ASPECT and ELEFANT.}
-\label{fig:thin_initial}
-\end{figure}
-
-\subsection{Sinking hydrated cylinder}\label{sec:quinquis}
-This experiment is performed as explained in \citet{Quinquis2014} and simulates the sinking of a cold, hydrated cylinder into a warm, dry mantle. The domain
-is square with $L_x=L_y=$ \SI{300}{\km}, a grid resolution of \num{300x300} elements and 25 markers per element. The sinking cylinder is located at $x=$
-\SI{150}{\km} and $y=$ \SI{170}{\km} with a radius of \SI{20}{\km} and has $\rho_c=$ \SI{3250}{\kg\per\cubic\m}, $\eta_c=$ \SI{e23}{\pascal\s}, a thermal
-conductivity $k_c=$ \SI{4.5}{\watt\per\m\per\kelvin},a specific heat $C_{p_c}=$ \SI{1250}{\joule\per\kg\per\kelvin}; the surrounding mantle has $\rho_m=$
-\SI{3200}{\kg\per\cubic\m}, $\eta_m=$ \SI{e20}{\pascal\s}, a thermal conductivity $k_m=$ \SI{105}{\watt\per\m\per\kelvin} and specific heat $C_{p_m}=$
-\SI{1250}{\joule\per\kg\per\kelvin}; a 58 km-thick lithosphere overlies the mantle and has $\rho_l=$ \SI{3200}{\kg\per\cubic\m}, $\eta_l=$ \SI{e23}{\pascal\s},
-a thermal conductivity $k_l=$ \SI{4.5}{\watt\per\m\per\kelvin}, specific heat $C_{p_l}=$ \SI{750}{\joule\per\kg\per\kelvin} and a thermal diffusivity of
-\SI{e-6}{\square\m\per\s}. Velocity boundary conditions are set to free slip on all sides of the domain. Temperature increase linearly in the lithosphere,
-from 0° to 1300°C, and in the mantle, from 1300° to 1360.5°C. The initial temperature of the cylinder is fixed to 400°C. Throughout the evolution, temperature
-is fixed to 0° and 1360.5°C at the top and bottom of the domain, respectively. The initial bound water content is imposed at 2 wt.\% in the cylinder and at
-0 wt.\% in both the mantle and the lithosphere. The maximum amount of water in the mantle is fixed to 0.2 wt.\%, while maximum water content in the cylinder
-and in the lithosphere is function of pressure and temperature and is calculated using a serpentinized harzburgite with Perple\_X, as in \citet{Quinquis2014}.
-
-As shown in \citet{Quinquis2014}, the model is characterised by a progressive dehydration in the external portion of the cylinder, due to the increase of
-temperature, with a consequent vertical migration of free water that hydrates the mantle above the cylinder, up to the lithosphere. The distribution of
-bound and free water during the evolution of the experiment is shown in Fig. \ref{fig:quinquis} (left and right columns, respectively). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Hydration}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Hydration.pdf}
-\caption{Distribution of bound and free water (left and right columns, respectively) at $t=0.5$ and $t=$ \SI{1.25}{\mega\year} (first and second rows,
-respectively) for the sinking hydrated cylinder experiment. Only markers with bound or free water not equal to 0 are plotted.}
-\label{fig:quinquis}
-\end{figure}
-
-\subsection{Mantle melting curves}\label{sec:katz}
-This instantaneous experiment is performed in order to recreate isobaric melting curves, as function of temperature, bulk water and content of water in the
-melt, to compare with melting curves obtained by \citet{Katz2003}. The domain is square with $L_x=L_y=100$ km, a grid resolution of \num{100x100} elements and
-25 markers per element. Density is set in the entire domain to \SI{3300}{\kg\per\cubic\m} and temperature increases from the sides to the centre from 1000° to
-1500°C. The experiment is performed for bulk water contents from 0 to 0.3 wt.\%.
-
-Fig. \ref{fig:fraction} shows isobaric melting curves for \SI{1}{\giga\pascal} (panel a) and \SI{3}{\giga\pascal} (panel b) calculated for different bulk water
-contents. The curves well match with those obtained for the mantle by \citet{Katz2003}, showing the saturation at the solidus for bulk water of 0.2 and 0.3 wt.\%, indicating
-that the code correctly determines melt fraction as function of pressure, temperature and bulk water content. Same results can be observed in Fig.
-\ref{fig:water}a, where isobaric and isothermal melting curves are shown as function of bulk water content. In addition, Fig. \ref{fig:water}b shows that the
-code correctly determines also the percentage of water in the melt as function of pressure, temperature and melt fraction. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Melting}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Fraction.pdf}
-\caption{Isobaric melting curves for 1 GPa (panel a) and 3 GPa (panel b) as function of temperature with different bulk water contents.}
-\label{fig:fraction}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Water.pdf}
-\caption{Isobaric melting curves for 1.5 GPa as function of bulk water (panel a) and water in the melt (panel b) calculated for different temperatures.}
-\label{fig:water}
-\end{figure}
-
-\clearpage
-
-\bibliography{export}
-
-\end{document}

### 2022-09-07 tex/Benchmark/indenter.tex
update main

--- 2022-03-31 FALCON_description.tex
@@
-\documentclass[hidelinks,10pt,a4paper]{article}
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{textcomp}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage{booktabs}
-\usepackage{multirow}
-\usepackage[cm]{fullpage}
-\usepackage[round]{natbib}
-\bibliographystyle{abbrvnat}
-\usepackage{doi}
-\usepackage{bm}
-\usepackage[table]{xcolor}
-\hypersetup{
-    colorlinks,
-    linkcolor={blue},
-    citecolor={blue},
-    urlcolor={blue}
-}
-\usepackage{float}
-\restylefloat{table}
-\usepackage{siunitx}
-\DeclareSIUnit\year{yr}
-
-\begin{document}
-
-\section{Introduction}\label{sec:intro}
-This document includes a description of the code FALCON, with a presentation of the features implemented followed by results of extensive benchmarks listed in
-Table \ref{tab:list}, performed to verify that the code solves correctly a wide range of problems. Description and results of all benchmarks are shown
-in Section \ref{sec:benchmark}.
-
-\begin{table}[H]
-\caption{List of benchmarks performed for each feature of the code. Data of all benchmarks can be found at 
-\url{https://github.com/aleregorda/Benchmarks}}
-\centering
-\begin{tabular}{| l | l | l |}
-\toprule
- \textbf{Feature}  & \textbf{Benchmark} & \textbf{Section} \\
-\midrule
-  Solver  & Stokes flow & \ref{sec:stokes} \\
-\hline
-  \multirow{2}{*}{Markers advection} & Zalesak disk & \ref{sec:runge} \\
-    & Conservative Interpolation Velocity & \ref{sec:CVI} \\
-\hline
-  \multirow{4}{*}{Momentum equation}  & Poiseuille flow & \ref{sec:poiseuille}\\
-    & Instantaneous 2D sphere & \ref{sec:ist_sphere} \\
-    & Rayleigh-Taylor experiment & \ref{sec:rayleigh} \\
-    & Falling block & \ref{sec:block} \\
-\hline
-    & 2D Stokes sphere & \ref{sec:sphere} \\
-    Sticky air and  & Stabilisation algorithm & \ref{sec:stab} \\
-    free surface & Topography relaxation & \ref{sec:crameri} \\
-  & Spontaneous subduction & \ref{sec:subduction} \\
-\hline
-  Erosion and sedimentation  & - & - \\
-\hline
-   Non-linear  & Slab detachment & \ref{sec:slab} \\
-   visco-plasticity & Indenter & \ref{sec:indenter} \\
-    & Brick & \ref{sec:brick} \\
-\hline
-  \multirow{3}{*}{Energy equation}  & Advection stabilisation & \ref{sec:advection} \\
-    & Simple shear heating & \ref{sec:simple_shear} \\
-    & Shear and adiabatic heating & \ref{sec:shear} \\
-\hline
-\multirow{3}{*}{Energy + momentum} & Mantle convection & \ref{sec:mantle} \\
-& Viscoplastic mantle convection & \ref{sec:visco_mantle} \\
-& Thin layer entrainment & \ref{sec:thin} \\
-\hline
-  Phase changes  & \multirow{2}{*}{Hydrated sinking cylinder} & \multirow{2}{*}{\ref{sec:quinquis}} \\
-  and hydration  &  & \\
-\hline
-  Melting & Experimental melting curves & \ref{sec:katz} \\
-\bottomrule
-\end{tabular}
-\label{tab:list}
-\end{table}
-
-The code is written in Fortran90 and uses the finite element method (FEM) with quadrilateral $Q_1 \times P_0$ elements (continuous bilinear velocity and
-discontinuous constant pressure), associated with MUMPS\footnote{\url{http://mumps.enseeiht.fr/}} \citep{Amestoy2001,Amestoy2019} that is a software package
-for solving systems of linear equations of the form $A \cdot x = b$, where \textit{A} is a square sparse matrix, by means of a direct method. Since
-$Q_1 \times P_0$ elements do not satisfy Ladyzhenskaya, Babuska and Brezzi (LBB) stability condition \citep{Donea2003}, elemental pressure is elaborated
-in post-processing to avoid spurious pressures by means of a double interpolation to smooth the pressure. Following this procedure, the elemental pressure
-is interpolated onto nodes and then back onto elements. After the smoothing procedure, the pressure field is calculated again on the nodes to be used in
-combination with the lithostatic pressure, which is calculated onto nodes. Correctness of the solution and performances of the solver in terms of time and
-memory usage are tested solving the Stokes flow with the analytical solution proposed by \citet{Donea2003} (Section \ref{sec:stokes}).
-
-The thermo-mechanics of crust-mantle systems is described by means of the conservation of mass, momentum and energy equations, expressed as follows:
-\begin{eqnarray}
-\label{eq:mass}\frac{\partial \rho}{\partial t}+\bm{\nabla} \cdot (\rho \bm{u})&=&0\\
-\label{eq:momentum}-\bm{\nabla} p + \bm{\nabla} \cdot \bm{\tau} + \rho \bm{g}&=&0\\
-\label{eq:energy}\rho C_p \left(\frac{\partial T}{\partial t} + \bm{u} \cdot \bm{\nabla} T \right)&=&\bm{\nabla} \cdot \left(k\bm{\nabla} T\right) + H_{tot}
-\end{eqnarray}
-where $\rho$ is the density, $\bm{u}$ is the velocity, \textit{p} is the pressure, $\bm{\tau}$ is the deviatoric stress, $\bm{g}$ is the gravity acceleration,
-$C_p$ is the specific heat at constant pressure, \textit{T} is the temperature, \textit{k} is the thermal conductivity and $H_{tot}$ is the total internal heat
-production. Density variations due to temperature are generally small enough to assume the density as constant ($\rho=\rho_0$) in Eq. \ref{eq:mass} and in
-Eq. \ref{eq:energy}, while it must be treated as a variable in the buoyancy term of Eq. \ref{eq:momentum}, such that \[\rho=\rho_0(1-\alpha(T-T_0))\]
-where $\rho_0$ is the density at a reference temperature $T_0$ and $\alpha$ is the coefficient of thermal expansion. Eq. \ref{eq:mass} then can be rewritten as
-\begin{equation}\label{eq:mass_0}
-\bm{\nabla} \cdot \bm{u}=0
-\end{equation}
-This simplification is known as the extended Boussinesq approximation.
-
-The time step $dt$ is chosen in according to Courant-Friedrichs-Lewy condition \citep{Anderson1995}, such that
-\[dt=C_n\; \min\left(\frac{\Delta x}{\max|\bm{u}|},\frac{(\Delta x)^2}{\kappa}\right)\]
-where $0<C_n<1$ is the Courant (CFL) number, $\Delta x$ is the minimum dimension of smallest element, $\max|\bm{u}|$ is the maximum velocity calculated on the
-entire domain and
-$\kappa=\frac{k}{\rho C_p}$ is the thermal diffusivity \citep{Thieulot2014}.
-
-\section{Lagrangian markers}\label{sec:markers}
-Elemental properties (density, viscosity, thermal conductivity, specific heat and thermal expansion) needed to solve Eqs. \ref{eq:mass_0}, \ref{eq:momentum}
-and \ref{eq:energy} are related to the composition of each element, determined by means of Lagrangian markers that are characteristic of different materials
-in the domain. Elemental properties (with the sole exception of the viscosity) are calculated using an arithmetic mean, as \[P_e=\frac{1}{n}\sum_{i=1}^n P_i\]
-where $P_e$ is the elemental property, $P_i$ is the property characteristic of the material of each marker and $n$ is the number of markers of the element.
-Differently, the average scheme for the elemental viscosity can be chosen between harmonic, geometric and arithmetic mean. Markers can be placed regularly or
-randomly at the beginning of the simulation and their advection is performed by either a 2nd-order or a 4th-order Runge-Kutta in space, interpolating the velocity 
-field on each marker by means of the shape functions (see Section \ref{sec:runge}). The interpolated velocity is then corrected by means of the Conservative
-Velocity Interpolation (CVI), which introduces a corrective term to reduce dispersion and clustering of particles in both steady state and time-dependent 
-\citep{Wang2015} (see Section \ref{sec:CVI}).
-
-The initial distribution of the markers is created using the open-source code library
-Geodynamic World Builder \citep{Fraters2019}. During the simulation, each marker carries memory of temperature, pressure and accumulated strain, which are
-determined interpolating the relative nodal parameter. In particular, pressure is interpolated onto each marker after the smoothing procedure. The number
-of markers contained in each element is maintained between $n_{min}$ and $n_{max}$. When in an element there are less markers than $n_{min}$ the code adds
-random markers to reach the $n_{min}$, while if the number is higher than $n_{max}$ some of them are randomly deleted. When new markers are added, they assume
-the properties of the nearest marker. In this way elements are never empty and maintain a number of markers inside a prefixed range.
-
-\section{The momentum equation}\label{sec:mom_eq}
-The penalty method is implemented so that Eq. \ref{eq:mass_0} can be written as:
-\begin{equation}\label{eq:mass_penalty}
-\bm{\nabla} \cdot \bm{u} + \frac{p}{\lambda}=0
-\end{equation}
-where $\lambda$ is the so-called penalty parameter, which should be 6-7 orders of magnitude larger than the shear viscosity to ensure that mass conservation
-is satisfied \citep{Donea2003,Thieulot2014}.
-
-The deviatoric stress tensor in Eq. \ref{eq:momentum} can be written in terms of the strain rate tensor as $\bm{\tau}=2\eta\dot{\bm{\epsilon}}$, with
-$\dot{\bm{\epsilon}}=\frac{1}{2}\left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)$. Therefore, Eq. \ref{eq:momentum} can be rewritten as:
-\begin{equation}\label{eq:momentum_pressure}
--\bm{\nabla} p + \bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right) + \rho \bm{g} = 0
-\end{equation}
-Finally, using pressure from Eq. \ref{eq:mass_penalty}, Eq. \ref{eq:momentum_pressure} can be rewritten as:
-\begin{equation}\label{eq:momentum_penalty}
-\lambda \bm{\nabla} \left(\bm{\nabla} \cdot \bm{v}\right)+\bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right)+\rho\bm{g}=0
-\end{equation}
-The penalty method is been associated to the iterative Uzawa method, as described in detail in \citet{Dabrowski2008} and \citet{Thieulot2014}
-(see Section \ref{sec:poiseuille}). Results of benchmarks performed to verify the correctness in the implementation of Eq. \ref{eq:momentum_penalty} are
-shown in Sections \ref{sec:ist_sphere} and \ref{sec:rayleigh}.
-
-To support large viscosities variations the penalty parameter is related to the effective elemental viscosity by means of a dimensionless coefficient, so that
-$\lambda=\lambda_e(e)\eta_{eff}(e)$ \citep{Marotta2006,Dabrowski2008,Thieulot2014}. Benchmark of the falling block \citep{Gerya2003a,Gerya2010b,Thieulot2011}
-is performed to verify that the code can correctly deal buoyancy driven flows with strong viscosity contrasts (Section \ref{sec:block}).
-
-\section{Sticky air and free surface}\label{sec:sticky}
-The Earth's surface can be treated by means of either the so-called sticky-air or a true free surface method, both of them implemented in the code. In the
-sticky-air method the surface is approximated with the introduction of a buoyant layer with a viscosity at least four orders of magnitude lower than the crust
-\citep{Schmeling2008,Crameri2012} and the interface between lithosphere and air is defined using a chain of passive markers that are advected as the Lagrangian
-markers. The correctness of the evolution of the markers chain is tested with the experiment of a 2D time-dependent Stokes sphere below a free surface and
-compared with results from ASPECT\footnote{\url{https://aspect.geodynamics.org/}} \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} (Section
-\ref{sec:sphere}). In the true free surface case the top boundary is assumed stress-free and velocities are not fixed. In this case topography variations are
-described by vertical deformations of the mesh that depend on the velocity field of the nodes that identify the free surface, while horizontal deformations are
-not taken into account. This procedure is known as the Arbitrary Lagrangian-Eulerian (ALE) method and its implementation follows the technique described in
-\citet{Thieulot2011}. However, although the implementation of a true free surface better reproduces laboratory experiment, extremely small time steps can be
-necessary to maintain stability \citep{Kaus2010a,Quinquis2011,Thieulot2014}. Therefore, the stability algorithm proposed by \citet{Kaus2010a} is implemented
-to avoid instabilities due to high density differences at the free surface when using too large time steps. The implementation of this algorithm are tested by
-performing the experiment described by \citet{Kaus2010a} (Section \ref{sec:stab}).
-
-The topography relaxation benchmark proposed by \citet{Crameri2012} is performed to verify that the code correctly recovers topography variations in case of
-both the sticky-air and the true free surface method (Section \ref{sec:crameri}). Finally, Section \ref{sec:subduction} show the results of the spontaneous
-subduction experiment described by \citet{Schmeling2008}.
-
-\section{Erosion and sedimentation}\label{sec:erosion}
-Surface processes at either the lithosphere-air interface or in correspondence of the free surface have been implemented by means of the software
-Fastscape\footnote{\url{https://fastscape.org/}} \citep{Braun2013,Cordonnier2019,Yuan2019a,Yuan2019b}. FastScapeLib is a set of routines that solve the stream
-power law (enriched by a sediment transport/deposition term) \citep{Yuan2019a}, hillslope diffusion and marine transport and deposition \citep{Yuan2019b}.
-The partial differential equation solved by FastScapeLib is:
-\begin{equation}
-\frac{\partial h}{ \partial t}=U-K_f A^{m_e}S^{n_e} + \frac{G}{A}\int_A \left( U-\frac{\partial h}{\partial t} \right) dA + K_d \bm{\nabla}^2 h 
-\end{equation}
-where $h$ is the topography, $U$ is the uplift, $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $S$ is the slope, $m_e$ and $n_e$ are
-parameters that should be chosen to constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}, $G$ is the dimensionless
-deposition/transport coefficient for the enriched stream power law and $K_d$ is the effective diffusivity. Terms $K_f A^{m_e}S^{n_e}$ and $K_d \bm{\nabla}^2 h$
-consider long-range processes, described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, and short-range processes,
-proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}.
-
-In case of sticky air, slope is determined by means of the markers chain, which is vertically corrected considering erosion and sedimentation rates. After the
-correction, continental markers above the markers chain are transformed in air markers and air markers below the markers chain become sediments. Similarly,
-slope is determined considering the nodes on the top boundary in case of true free surface and they are vertically corrected as for the markers chain. In this
-case, continental markers above the top boundary are simply deleted.
-%Surface processes at the lithosphere-air interface are implemented taking into account fluvial erosion (long-range processes), hillslope processes (short-range
-%processes) and failure of slopes exceeding a threshold angle $\theta$ \citep{Burov1997,Tucker1999,Sacek2012,Croissant2014,Ueda2015,Beucher2020}. Erosion rates
-%$\dot{E}_{lr}$ due to long-range processes can be described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{lr}=K_f A^{m_e} \bm{\nabla} h^{n_e}
-%\end{equation}
-%where $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $\bm{\nabla} h$ is the slope and $m_e$ and $n_e$ are parameters chosen in to
-%constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}. Erosion rates $\dot{E}_{sr}$ due to short-range processes are
-%implemented by assuming that transport is linearly proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{sr}=K_d \bm{\nabla}^2 h
-%\end{equation}
-%where $K_d$ is the effective diffusivity. In addition, the slope is modified if it still exceeds a threshold angle after long- and short-range processes
-%\citep{Ueda2015}.
-
-%Offshore sedimentation processes are implemented similarly to hillslope processes but with a depth-dependent coefficient \citep{Sacek2012,Beucher2020}.
-%The sedimentation rate $\dot{S}$ is calculated as
-%\begin{equation}
-%\dot{S}=K_s \bm{\nabla}^2 h
-%\end{equation}
-%with
-%\begin{equation}
-%K_s=K^*_s \exp\left(\frac{h-h_{sl}}{h_0}\right)
-%\end{equation}
-%where $K^*_s$ is a constant, $h_{sl}$ is the sea level and $h_0$ is a characteristic length. In case of sticky air, slope is determined by means of the markers
-%chain, which is vertically corrected considering erosion and sedimentation rates. After the correction, continental markers above the markers chain are
-%transformed in air markers and air markers below the markers chain become sediments. Similarly, slope is determined considering the nodes on the top boundary
-%in case of true free surface and they are vertically corrected as for the markers chain. In this case, continental markers above the top boundary are simply
-%deleted.
-
-\section{Non-linear rheologies}\label{sec:nonlinear}
-Non-linear rheologies are implemented combining viscous creep (dislocation and diffusion) and plastic yielding. For each marker, diffusion and dislocation
-viscosity ($\eta_{df|ds}$) can be determined as follows \citep{Karato1993,Warren2008b,Wang2016}:
-\begin{equation}\label{eq:viscous}
-\eta_{df|ds}=f_s \left(\frac{d^m}{A}\right)^{\frac{1}{n}} \left (\frac{2^{\frac{1-n}{n}}}{3^{\frac{n+1}{2n}}} \right)
-I_2^{\frac{(1-n)}{n}}\exp\left(\frac{Q+pV}{nRT}\right)
-\end{equation}
-where $f_s$ is a scaling factor used to represent lithologies that are stronger or weaker than the base set, $d$ is the grain size, $m$ is the grain size exponent,
-$n$ is the stress exponent, $A$ is the uniaxial pre-exponential factor, $I_2$ is the square root of the second invariant of the strain rate tensor, $Q$ is the
-activation energy, $p$ is the pressure, $V$ is the activation volume, $R$ is the gas constant and $T$ is the temperature.
-Pressures and temperatures are determined by the interpolation of the nodal parameters, while strain rates are calculated by means of the
-derivative of the velocities on the nodes. In case of diffusion creep $n=1$ and $m>0$, in case of dislocation creep $n>1$ and $m=0$. The viscous creep $\eta_{cp}$
-is then calculated as the harmonic average between $\eta_{df}$ and $\eta_{ds}$:
-\begin{equation}\label{eq:visco_creep}
-\eta_{cp}=\left(\frac{1}{\eta_{df}}+\frac{1}{\eta_{ds}}\right)^{-1}
-\end{equation}
-The implementation of the non-linear viscous creep viscosity depending on the strain rate is tested by means of the slab detachment benchmark
-(Section \ref{sec:slab}).
-
-Plastic yielding is implemented rescaling $\eta_{cp}$ in order to limit the stress below the yield stress $\sigma_y$ \citep{Thieulot2008,Thieulot2014,Glerum2018},
-obtaining
-\begin{equation}\label{eq:plastic}
-\eta_{pl}=\frac{\sigma_y}{2I_2}
-\end{equation}
-where $\eta_{pl}$ is the plastic viscosity and the yield stress is determined following the Drucker-Prager criterion, such as
-\begin{equation}\label{eq:yield}
-\sigma_y=C\;\cos(\phi)+p\;\sin(\phi)
-\end{equation}
-where $C$ is the cohesion and $\phi$ is the internal friction angle. In case of negative pressure it is imposed equal to 0, so that negative yield stress are
-excluded. The correctness of the non-linear solution in case of plastic viscosity with variable internal friction angle is verified performing indenter and
-brick experiments (Sections \ref{sec:indenter} and \ref{sec:brick}, respectively).
-
-Strain softening is taken into account for both viscous creep and plastic viscosity \citep{Huismans2003,Babeyko2005,Huismans2005,Sobolev2005,Warren2008a} by
-means of the accumulated strain $\epsilon$ memorised by each marker. Softening in the viscous creep determines a linear decrease of $\eta_{vc}$ by means of a
-viscous strain softening factor $W_S$ that increases linearly from $W_{S_0}$ to $W_{S_{\infty}}$ for $\epsilon_{S_0}<\epsilon_S<\epsilon_{S_{\infty}}$ 
-\citep{Warren2008a}. This viscous softening can be related to strain-induced grain size reduction \citep{Warren2008a}. Differently, plastic softening is 
-simulated with a linear decrease of internal friction angle $\phi(\epsilon)$ and cohesion $C(\epsilon)$ in according to
-\begin{eqnarray}
-\label{eq:friction1}\phi(\epsilon)&=&\phi_0+(\phi_{\infty}-\phi_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}\\
-\label{eq:friction2}C(\epsilon)&=&C_0+(C_{\infty}-C_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}
-\end{eqnarray}
-where $\phi_0$, $C_0$ and $\phi_{\infty}$, $C_{\infty}$ are internal friction angle and cohesion for $\epsilon_0$ and $\epsilon_{\infty}$, respectively
-\citep{Huismans2003,Huismans2005,Warren2008a,Thieulot2014}. Plastic softening approximates deformation-induced softening of faults and brittle shear zones
-\citep{Warren2008a}.
-
-Plastic yielding and viscous creep are then combined to obtain a viscoplastic viscosity $\eta_{vp}$ as follows, assuming that they are independent processes
-\citep{Karato2008,Glerum2018}:
-\begin{equation}\label{eq:viscoplastic}
-\eta_{vp}=\min(\eta_{cp},\eta_{pl})
-\end{equation}
-Finally, effective viscosity $\eta_{eff}$ is capped by the minimum and the maximum viscosity ($\eta_{min}$ and $\eta_{max}$, respectively) to avoid extremely
-low or high viscosity \citep{Glerum2018} as follows
-\begin{equation}\label{eq:effective}
-\eta_{eff}=\min(\max(\eta_{vp},\eta_{min}),\eta_{max})
-\end{equation}
-Viscous creep and plastic yielding are non-linear rheologies because of their dependence on the velocity field through pressure and strain rates. Therefore,
-the solution is determined by means of Picard-type iterations, until convergence of the velocity field \citep{Glerum2018}. The convergence is verified at each
-iteration $i$ via the nonlinear residual $\bm{\mathcal{R}}^i$ that can be determined as
-\begin{equation}\label{eq:convergence}
-\bm{\mathcal{R}}^i= \mathbb{K} (\eta_{eff}(\dot{\bm{\epsilon}}^{i-1},p^{i-1}))\cdot \bm{v}^{i-1} - \bm{f}^i
-\end{equation}
-%\begin{equation}\label{eq:convergence}
-%\chi=1-\frac{\langle (\bm{v}^i-\langle\bm{v}^i\rangle)\cdot (\bm{v}^{i+1}-\langle\bm{v}^{i+1}\rangle) \rangle}{\sqrt{\sigma^i\sigma^{i+1}}}
-%\end{equation}
-%Iterations are performed until either convergence ($\chi<tol$) or maximal number of non-linear iterations ($it_{max}$) is reached.
-where $\mathbb{K}$ is the Stiffness matrix and $\bm{f}$ is the right hand side vector \citep{Spiegelman2016,Glerum2018}.  The $L_2$-norm is extracted from
-$\bm{\mathcal{R}}^i$ and it is normalised as follows
-\begin{equation}\label{eq:convergence_res}
-\frac{||\bm{\mathcal{R}}^i||_2}{||\bm{\mathcal{R}}^0||_2}
-\end{equation}
-where $||\bm{\mathcal{R}}^0||_2$ is the $L_2$-norm of the first nonlinear iteration.
-Since often the normalised nonlinear residual drop very quickly over the first 2-3 iterations, also the $L_2$-norm of velocity and pressure residuals are
-calculated as
-\begin{equation}\label{eq:convergence_vel}
-\frac{||\Delta\bm{v}||_2}{||\bm{v}||_2}=\frac{||\bm{v}_i-\bm{v}_{i-1}||_2}{||\bm{v}_{i}||_2}
-\end{equation}
-and
-\begin{equation}\label{eq:convergence_press}
-  \frac{||\Delta p||_2}{||p||_2}=\frac{||p_i-p_{i-1}||_2}{||p_{i}||_2}
-\end{equation}
-respectively. Iterations are performed until either maximal number of non-linear iterations ($it_{max}$) is reached or all the normalised nonlinear residual, 
-the velocity residual and the pressure residual converge under a defined tolerance $tol$.
-
-\section{The energy equation}\label{sec:en_eq}
-The stabilisation of the advection term of the energy equation (Eq. \ref{eq:energy}) needed to avoid possible oscillations in the thermal solution in the case
-when advection dominates over diffusion, is implemented by means of a streamline-upwind Petrov–Galerkin (SUPG) method, for which the advection term is modified
-follows the discussion in \citet{Thieulot2011} and \citet{Thieulot2014} (see Section \ref{sec:advection} for benchmark). The total internal heat production
-$H_{tot}$ in Eq. \ref{eq:energy} includes radiogenic heating $H_r$, shear heating $H_s$ and adiabatic heating $H_a$, where
-\begin{equation}\label{eq:shear}
-H_s=2\eta \dot{\bm{\epsilon}} : \dot{\bm{\epsilon}} = 2\eta(\dot{\epsilon}_{xx}^2+\dot{\epsilon}_{yy}^2+2\dot{\epsilon}_{xy}^2)
-\end{equation}
-and
-\begin{equation}\label{eq:adiabatic}
-H_a=T \alpha \frac{Dp}{Dt} \approx -\alpha T \rho g_y v
-\end{equation}
-The correctness of the implementation and the computation of shear and adiabatic heating is verified performing an exercise with an analytical solution
-(Section \ref{sec:simple_shear}) and Exercise 9.4 in \citet{Gerya2010b} (Section \ref{sec:shear}). Furthermore, mantle convection and thin layer entrainment
-benchmarks (Sections \ref{sec:mantle} and \ref{sec:thin}, respectively) are performed to verify that the code solve correctly Eqs. \ref{eq:momentum_penalty}
-and \ref{eq:energy} in case of a temperature field characterised by an initial perturbation and with a temperature-dependent density. At last, the viscoplastic
-mantle convection benchmark is performed to check also the viscoplasticity in case of a initial perturbed temperature field (Section \ref{sec:visco_mantle}).
-
-\section{Phase transitions and hydration}\label{sec:phase}
-Variations of effective density, specific heat and coefficient of thermal expansion during the evolution of crustal and mantle materials at different
-pressure-temperature ($p$-$T$) conditions are computed by means of version 6.8.6 of Perple\_X software package \citep{Connolly2005}, similarly to the
-implementation described in \citet{Marotta2020}. By default, mineral assemblages and properties of each lithology (oceanic crust, lower and upper continental
-crust, sediments and dry or hydrated mantle) are calculated for temperatures from \SIrange{330}{1600}{\kelvin}, with increments of approximately \SI{4}{\kelvin},
-and pressures from \SIrange{0.1}{30}{\giga\pascal}, with increments of approximately \SI{0.03}{\giga\pascal}, for a total of almost 100000 points.
-
-The introduction of phase transitions produces two main effects that must be taken into account in numerical models: variations in density and release/absorption
-of latent heat \citep{Tackley2010}. Effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into
-account considering effective coefficient of thermal expansion \citep{Christensen1985,Zhong2015}. Similarly, effects of latent heat in the energy equation
-(eq \ref{eq:energy}) are taken into account by considering effective specific heat and coefficient of thermal expansion
-\citep{Christensen1985,Tackley2010,Zhong2015}. Effective density, specific heat and coefficient of thermal expansion are included in Perple\_X files and are
-assumed by each Lagrangian marker in according to its $p$-$T$ conditions.
-
-Hydration processes related to dehydration of subducting lithosphere \citep{Schmidt1998,Liu2007,Faccenda2009,Faccenda2010,Faccenda2014,Rosas2016} strongly
-influence the thermo-mechanics inside the mantle wedge, mainly because of both weakening effects on the mantle rheology and density variations in case of
-mantle serpentinization \citep{Gerya2002,Honda2003,Arcay2005,Roda2010,Regorda2017}. In the case that hydration is switched on, the amount of bound and free
-water is memorised by each marker, following the implementation of \citet{Quinquis2014}; hydration and dehydration processes are related to the amount of bound
-water of each marker and to the maximum amount of water it can transport, i.e. if the amount of bound water exceeds the maximum amount of water, the marker
-dehydrates and releases free water that can hydrates under-saturated markers. Maximum water content of each marker is determined as function of lithology and
-$p$-$T$ conditions and it is calculated using Perple\_X, in the same way than effective density, specific heat and coefficient of thermal expansion. Bound water
-is advected together with the markers, neglecting the effect of bound water diffusion, while free water simply migrates vertically and is not coupled to the
-solid-phase flow of the mantle wedge \citep{Arcay2005,Quinquis2014}. The correctness of the migration scheme of the free water and its absorption into bound
-water by under-saturated markers are verified performing the experiment described in \citet{Quinquis2014} (Sec. \ref{sec:quinquis}).
-
-The mantle viscosity weakening related to the amount of water has been extensively studied \citep{Chopra1981,Hirth2003} and it is implemented in according to
-\citet{Arcay2005} and \citet{Horiuchi2016}, as follows
-\begin{equation}\label{eq:wet_visc}
-\eta_{wet}=\eta_{dry}\left(\left[1-\frac{1}{f_v}\right]\exp\left(-\frac{[OH^-]}{[OH^-]_0}\right)+\frac{1}{f_v}\right)
-\end{equation}
-where $f_v$ is the viscosity reduction factor between dry and wet conditions, $[OH^-]$ is the water content and $[OH^-]_0$ is a reference water content set to
-620 ppm (0.062 wt.\%) \citep{Arcay2005}. By default, $f_v$ is set to 100, in according to \citet{Arcay2005} and \citet{Horiuchi2016}.
-
-\section{Mantle melting}\label{sec:melt}
-Melt fraction in the mantle depends on temperature ($T$, in \textdegree C), pressure ($p$, in \si{\giga\pascal}) and the water content in the melt ($X_{H_2O}$, in wt.\%).
-The determination of the melt fraction $M$ is implemented as explained by \citet{Katz2003}, according to the successive modification by \citet{Langmuir2006} and
-\citet{Kelley2010}, as follows
-\begin{eqnarray}\label{eq:melt_fr}
-M(p,T,X_{H_2O})&=&-T+T_{sol}+(x_t\cdot ln(p+y_t))\cdot X_{H_2O}-\nonumber\\
-&&+K_m\left(\frac{X_{H_2O}^{bulk}}{D_{H_2O}(1-X_{H_2O})+X_{H_2O}}\right)^{\gamma}
-\end{eqnarray}
-where $T_{sol}$ is the temperature of the dry solidus, $D_{H_2O}$ is the partition coefficient and $x_t\cdot log(p+y_t)$ indicates the pressure dependence of
-the melting curve \citep{Kelley2010}. $x_t$, $y_t$, $K_m$ and $\gamma$ are parameters chosen in according to \citet{Kelley2010}.
-However, $X_{H_2O}$ depends on the melt fraction as
-\begin{equation}\label{eq:melt_wt}
-X_{H_2O}(M)=\frac{X_{H_2O}^{bulk}}{D_{H_2O}+M(1-D_{H_2O})}
-\end{equation}
-and a numerical solution can be found using a root-finder method \citep{Katz2003,Wang2016}. Furthermore, the content of water in the melt is limited by the
-pressure-dependent saturation concentration of water in the melt, determined as
-\begin{equation}\label{eq:melt_max}
-X_{H_2O}^{sat}=\chi_1 p^{\lambda_m}+\chi_2 p
-\end{equation}
-where $\chi_1$, $\chi_2$ and $\lambda_m$ are parameters chosen according to \citet{Katz2003}, \citet{Langmuir2006} and \citet{Kelley2010}. The correctness in
-the determination of melt fractions and water contents in the melt are tested comparing model results with experimental curves obtained by \citet{Katz2003}
-(Sec. \ref{sec:katz}).
-
-In case of melting, viscous creep viscosity ($\eta_{vp}$) is modified according to \citet{Wang2016}, as
-\begin{equation}\label{eq:melt_visc}
-\eta_{melt}=\eta_{vp}\cdot \exp(\alpha_m M)
-\end{equation}
-where $\alpha_m$ are melt fraction factors for dislocation and diffusion creep, chosen in according to \citet{Wang2016}.
-
-Effective density ($\rho_{eff}$) for partially molten rocks is calculated as
-\begin{equation}\label{eq:melt_dens}
-\rho_{eff}=\rho_s(1-M)+\rho_m M
-\end{equation}
-where $\rho_s$ and $\rho_m$ are the densities of the solid and the molten rock, respectively \citep{Gerya2004b,Gerya2007,Wang2016}. The density of the
-solid rock is extracted by Perple\_X, while the density of the molten portion is calculated as
-\begin{equation}\label{eq:melt_dens_M}
-\rho_m(p,T)=\rho_0[1-\alpha(T-T_0)][1+\beta(p-p_0)]
-\end{equation}
-where $\rho_0$ is the density at the reference temperature ($T_0$) and pressure($p_0$), and $\alpha$ and $\beta$ are the thermal expansion and compressibility
-coefficients, respectively \citep{Gerya2004b,Gerya2007,Wang2016}.
-
-As for the phase transitions, effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into account
-considering effective coefficient of thermal expansion. Similarly, latent heat due to melting/crystallisation can be implicitly included in the energy equation
-(Eq. \ref{eq:energy}) by considering effective specific heat ($C_{p_{eff}}$) and thermal expansion ($\alpha_{eff}$) of the partially molten rocks, in according to
-\begin{eqnarray}
-\label{eq:cp_effective}C_{p_{eff}}&=&C_p+H_L\left(\frac{\partial M}{\partial T}\right)_p \\
-\label{eq:alpha_effective}\alpha_{eff}&=&\alpha+\rho\frac{H_L}{T}\left(\frac{\partial M}{\partial p}\right)_T
-\end{eqnarray}
-where $C_p$ and $\alpha$ are the specific heat and the coefficient of thermal expansion of the solid rock, respectively, and $H_L$ is the latent heat of the
-molten rock \citep{Gerya2004b,Gerya2007,Tackley2010}.
-
-\section{Benchmarks}\label{sec:benchmark}
-\subsection{Error measurements}\label{sec:error}
-In order to determine the accuracy of velocity and pressure field of the benchmarks, the $L_2$-norm is computed by numerical integration on the Gauss-Legendre
-quadrature points. $L_2$-norm for pressure and velocity errors can be evaluated as
-\begin{eqnarray}
-\textrm{err}_p&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|p^n_i(r_q)-p^a_i(r_q)|^2\right)w_q|J_q|}\nonumber \\
-\textrm{err}_v&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|u^n_i(r_q)-u^a_i(r_q)|^2+|v^n_i(r_q)-v^a_i(r_q)|^2\right)w_q|J_q|}\nonumber
-\end{eqnarray}
-respectively, where $n_e$ is the number of elements, $n_q$ is the number of quadrature points per element, $p^n_i(r_q)$ and $p^a_i(r_q)$ are the numerical and
-analytical pressure, respectively, in each quadrature point $q$, $w_q$  and $J_q$ are the weight and the Jacobian at the quadrature point
-$q$, $u^n_i(r_q)$, $v^n_i(r_q)$, $u^a_i(r_q)$ and $v^a_i(r_q)$ are the numerical and analytical velocities, respectively, in each quadrature point $q$.
-
-Other quantities used as comparison with original benchmarks are the root-mean-square velocity over the whole domain and over the surface
-\begin{eqnarray}
-v_{\textrm{rms}}&=&\sqrt{\int_{\Omega}|\bm{v}|^2 d\Omega}\nonumber \\
-v_{\textrm{rms}}^{top}&=&\sqrt{\int_0^1 u^2 \bigg|_{y=1} dx}\nonumber
-\end{eqnarray}
-respectively, top and bottom Nusselt numbers
-\[\textrm{Nu}^{top/bottom}=-\int_0^1\frac{\partial T}{\partial y} \bigg|_{y=1/y=0} dx\]
-the average rate of work done against gravity
-\[\left\langle W \right\rangle=\int_{\Omega}T u_y d\Omega\]
-the average rate of viscous dissipation
-\[\left\langle \Phi \right\rangle=\int_{\Omega}\tau_{ij}\dot{\epsilon}_{ij} d\Omega\]
-and the percentage error between $\left\langle W \right\rangle$ and $\left\langle \Phi \right\rangle/Ra$
-\[\delta=\frac{\left|\left\langle W \right\rangle -\frac{\left\langle \Phi \right\rangle}{Ra} \right|}{\max \left(\left\langle W \right\rangle,
-\frac{\left\langle \Phi \right\rangle}{Ra}\right)}\times100\]
-
-\subsection{Stokes flow}\label{sec:stokes}
-The problem consists of determining velocity field $(u,v)$ and pressure $p$ in case of a manufactured solution with prescribed body forces such as
-\begin{eqnarray}
-b_1&=&(12-24y)x^4+(-24+48y)x^3+(-48y+72y^2-48y^3+12)x^2+\nonumber \\
-&&+(-2+24y-72y^2+48y^3)x+1-4y+12y^2-8y^3\nonumber \\
-b_2&=&(8-48y+48y^2)x^3+(-12+72y-72y^2)x^2+(4-24y+48y^2-48y^3+\nonumber \\
-&&+24y^4)x-12y^2+24y^3-12y^4\nonumber
-\end{eqnarray}
-for which the exact solution is:
-\begin{eqnarray}
-u(x,y)&=&x^2(1-x)^2(2y-6y^2+4y^3)\nonumber \\
-v(x,y)&=&-y^2(1-y)^2(2x-6x^2+4x^3)\nonumber \\
-p(x,y)&=&x(1-x)-\frac{1}{6}\nonumber
-\end{eqnarray}
-The domain is a unit square a constant viscosity ($\eta=1$) and the penalty parameter is set to $\lambda=10^7$. Velocity boundary conditions are set to no slip
-$(\bm{v}=\bm{0})$ on all boundaries. The problem is performed for different grid resolution between \num{8x8} and \num{1024x1024} elements.
-
-The errors between the analytical solution and the numerical prediction of the velocity and the pressure field are calculated as in Section \ref{sec:error}.
-Fig. \ref{fig:errors}a shows that both velocity and pressure field converge to the exact solution with the decrease of the element size, following the
-theoretical convergence rate. The convergence of pressure, in contrast with the observation by \citet{Donea2003} for $Q_1 \times P_0$ elements, indicate the
-effectiveness of the smoothing procedure. Smoothed pressure field for grid resolution of \num{128x128} elements is shown in Fig. \ref{fig:errors}b (continuous
-red line), in comparison with the analytical solution (dashed blue line). Solve times and memory usage needed to generate the solution are shown in
-Fig. \ref{fig:MUMPS}. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Solver/Stokes_Flow}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/errors.pdf}
-\caption{Velocity and pressure error for the Stokes flow experiment between generated and analytical solution as a function of element size (panel a),
-and comparison between smoothed pressure and analytical solution as function of $x$ coordinate for a grid resolution of \num{128x128} elements (panel b).}
-\label{fig:errors}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/MUMPS.pdf}
-\caption{Panel a) Analysis, factorisation, solution and total solve times (panel a) and factorisation memory usage (panel b) as a function of the total number
-of degrees of freedom for the Stokes flow experiment.}
-\label{fig:MUMPS}
-\end{figure}
-
-\subsection{Markers advection}\label{sec:runge}
-The 2nd-order Runge-Kutta advection scheme in space is tested by means of the Zalesak disk test \citep{Zalesak1979,Thieulot2014}. The benchmark is performed
-in a unit square domain with a grid resolution of \num{32x32} elements and 1000000 markers and values of Courant number between 0.25 and 1.
-At $t=0$, the disk is centred at position $(0.5;0.75)$ with a radius $R=0.15$ and has a vertical fissure 0.05 wide and 0.2 high (orange in Fig. \ref{fig:runge})
-and with the velocity field prescribed in the entire domain as
-\begin{eqnarray}
-u(x,y)&=&2\pi\left(y-\frac{L_x}{2}\right)\nonumber \\
-v(x,y)&=&-2\pi\left(x-\frac{L_x}{2}\right)\nonumber
-\end{eqnarray}
-the markers are back to their initial location after a $2\pi$ rotation ($t=1$). The clockwise rotation of the disk at $t=0.25$, 0.5 and 0.75 
-is shown in Fig. \ref{fig:runge} (red, blue and green, respectively). The distance from the centre of 1 marker is calculated throughout the rotation to
-evaluate the error for different values of Courant number. As expected, the error increase with the increasing of the time step, as shown in 
-Fig. \ref{fig:runge_err}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/Zalesak_Disk}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=200px]{./Figures/Zalesak.pdf}
-\caption{Position of the Zalesak disk at $t=0$, 0.25, 0.5 and 0.75 (orange, red, blue and green, respectively) for the case with Courant number of 1.}
-\label{fig:runge}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/zal_marker.pdf}
-\caption{Distance from the centre as function of time for values of Courant number of  of 0.25, 0.3, 0.5, 0.75 and 1 (green, orange, blue, black and red,
-respectively). R indicates the distance from the centre considering a perfect circle trajectory.}
-\label{fig:runge_err}
-\end{figure}
-
-\subsection{Conservative Velocity Interpolation (CVI)}\label{sec:CVI}
-The CVI correction is checked by means of the Stokes flow experiment by \citet{Donea2003}, as presented in \ref{sec:stokes}. The advection of Lagrangian
-markers is performed using either a 2nd-order or a 4th-order Runge-Kutta scheme with CFL=0.5 and an initial random distribution of 25 markers per element.
-Fig. \ref{fig:CVI} shows the minimum and maximum values of markers per element throughout the experiment with and without the CVI correction and using
-different Runge-Kutta schemes. The introduction of the CVI correction allows the
-tests to run for times longer than $t=2000$ without show empty elements, while tests without the CVI correction stop at approximately $t=600$, when at least
-one element goes empty. Fig. \ref{fig:CVI_time} shows the different number of markers per element and the markers distribution at $t=600$ obtained without and
-with the CVI correction (panels b and c, respectively) in case of a 2nd-order Runge-Kutta scheme. All data can be found at
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/CVI}.
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/CVI.pdf}
-\caption{Maximum and minimum number of markers per element throughout the experiment, with (green and orange lines) or without (purple and blue lines)
-the CVI correction and using either a 2nd-order (blue and orange lines) or a 4th-order (purple and green lines) Runge-Kutta scheme.}
-\label{fig:CVI}
-\end{figure}
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/CVI_time.pdf}
-\caption{Number of markers per element and markers distribution through time using 2nd.order Runge-Kutta scheme.
-Panels a: number of markers per element (panel a1) and distribution of the markers (panel a2) after the first
-iteration; panels b and c: comparison between number of markers per element (panels b1 and c1) and markers distribution (panels b2 and c2) without and with
-the CVI correction at $t=600$; panels d: number of markers per element (panel d1) and markers distribution (panel d2) in case of CVI correction at $t=2000$.}
-\label{fig:CVI_time}
-\end{figure}
-
-\subsection{Poiseuille flow}\label{sec:poiseuille}
-The domain is a rectangle with $L_x=2$ and $L_y=1$ and constant density and viscosity ($\rho=1$ and $\eta=1$), gravity acceleration $\bm{g}=\bm{0}$ and penalty
-parameter $\lambda=10^8$. The grid is composed by \num{40x20} elements. Velocity boundary conditions are set to no slip $(\bm{v}=\bm{0})$ at the top and the
-bottom, and a parabolic profile is imposed on the sides, with $u=y(1-y)$ and $v=0$. The analytical solution is then given by:
-\begin{eqnarray}
-u(x,y)&=&y(1-y)\nonumber \\
-v(x,y,)&=&0\nonumber \\
-p(x,y)&=&2\eta\left(\frac{L_x}{2}-x\right)\nonumber
-\end{eqnarray}
-
-The velocity field predicted by the model follows the expected parabolic profile (Fig. \ref{fig:poiseuille}a and Fig. \ref{fig:poi_plot}a) and, in case of the
-classic penalty method with no iterations, the pressure is clearly related to the divergence by means of the penalty parameter (Fig. \ref{fig:poiseuille}b and
-c and Fig. \ref{fig:poi_plot}b). Fig. \ref{fig:poiseuille}d shows that one Uzawa iteration is sufficient to bring the divergence down to \num{e-15}, with no
-correlation with the pressure field. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Poiseuille%20Flow}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Poiseuille.pdf}
-\caption{Velocity field (panel a), pressure (panel b) and divergence velocity of a Poiseuille flow in case of the classic penalty method (no iterations) and
-after one Uzawa iteration (panel c and d, respectively).}
-\label{fig:poiseuille}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Analytical.pdf}
-\caption{Velocity (panel a) and pressure (panel b) field predicted by the model for a Poiseuille flow with respect with their analytical solutions. The
-velocity field is plotted as function of the vertical coordinate in $L_x/2$ and the pressure is plotted as function of the x coordinate in $L_y/2$.}
-\label{fig:poi_plot}
-\end{figure}
-
-\subsection{Instantaneous 2D sphere}\label{sec:ist_sphere}
-The domain is a unit square with gravity $\bm{g}=(0,-1)$. The fluid has constant density and viscosity ($\rho_f=1$ and $\eta_f=1$). The sphere is in the middle
-of the domain with a radius R = 0.123456798 and has constant density and viscosity ($\rho_s=10^{-2}+\rho_f$ and $\eta_s=10^3 \cdot \eta_f$).
-We distinguish three different types of velocity boundary conditions:
-\begin{enumerate}
-\item FS: free slip conditions on all sides.
-\item NS: no slip conditions on all sides.
-\item OT: free slip conditions on the sides and the bottom, and open boundary at the top.
-\end{enumerate}
-All three cases are performed for different grid resolutions, between \num{16x16} and \num{512x512} elements with 50 randomly distributed markers per element,
-and different average schemes for the viscosity (harmonic, geometric and arithmetic). The velocity in the centre of the sphere $\bm{v}(0.5;0.5)$), minimum and
-maximum velocities ($u_{min}$, $u_{max}$, $v_{min}$, $v_{max}$) and pressures ($p_{min}$, $p_{max}$) on the entire domain, $v_{\textrm{rms}}$ and average pressure on the
-entire domain ($p_{avg}$) are compared with solutions generated by ASPECT \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and all results
-can be found at \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere2D} and 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Instantaneous_2D_Sphere}. Results in terms of the $v_{\textrm{rms}}$ for NS model 
-are also shown in Fig. \ref{fig:inst_sphere}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_NS.pdf}
-\caption{$v_{\textrm{rms}}$ as function of element size for the instantaneous 2D sphere experiment in case of no slip boundary conditions and with different average
-schemes for the viscosity. Results are compared with results obtained by other numerical codes.}
-\label{fig:inst_sphere}
-\end{figure}
-
-\subsection{Rayleigh-Taylor instability}\label{sec:rayleigh}
-This problem has been originally presented by \citet{vanKeken1997} and here it is performed as the isoviscous case (Case 1a). The domain has $L_x=0.9142$ and
-$L_y=1$ and gravity $\bm{g}=(0,-1)$. Two fluids with same constant viscosities ($\eta_1=\eta_2=1$) and different densities ($\rho_1=1000$ and $\rho_2=1010$),
-with the lighter fluid at the bottom. The initial interface between the fluids is given by $y(x)=0.2+0.02 \cos \left(\frac{\pi x}{L_x}\right)$. The experiment
-is performed with different grid sizes (\num{50x50}, \num{80x80}, \num{100x100} and \num{256x256}). A total of 1960000 markers are randomly distributed at the
-beginning of the simulation. Velocity boundary conditions are set to no slip at the top and the bottom, and to free slip at the sides of the domain.
-
-$v_{\textrm{rms}}$ as function of time is reported in Fig. \ref{fig:RT}, matching well with results shown by \citet{vanKeken1997}, \citet{Tackley2003} and
-\citet{Thieulot2014}. Fig. \ref{fig:rayleigh} shows the evolution of the experiment at different time steps. All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Rayleigh_Taylor_experiment/ISOVISCOUS}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/RT.pdf}
-\caption{$v_{\textrm{rms}}$ of the Rayleigh-Taylor experiment as function of time for different resolution of the grid.}
-\label{fig:RT}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=300px]{./Figures/Rayleigh.pdf}
-\caption{Evolution of the Rayleigh-Taylor experiment for a grid of \num{256x256} elements at $t=0, 500, 1000$ and $2000$ (panels a, b, c and d, respectively).}
-\label{fig:rayleigh}
-\end{figure}
-
-\subsection{Falling blocks}\label{sec:block}
-This benchmark is described as proposed by \citet{Gerya2003a}, \citet{Gerya2010b} and \citet{Thieulot2011}. The domain is square with $L_x=L_y=$ \SI{500}{\km}
-and the grid is composed by \num{50x50} elements with 25 markers in each element. The block is initially centred at ($x=$ \SI{250}{\km}; $y=$ \SI{400}{\km})
-and has a size of \num{100x100} km (Fig. \ref{fig:block}a). The fluid surrounding the block has $\eta_f=$ \SI{e21}{\pascal\per\s} and $\rho_f=$
-\SI{3200}{\kg\per\cubic\m}. The benchmark tests with different viscosities of the block, with $\eta_b$ from \SIrange{e15}{e27}{\pascal\per\s}. In each
-experiment also the density $\rho_b$ of the block varies from \SIrange{3220}{9900}{\kg\per\cubic\m}. Velocity boundary conditions are set to free slip
-conditions on all sides of the domain. Density distributions at $t=$ \SI{20}{\mega\year} for $\rho_b=$ \SI{3300}{\kg\per\cubic\m} and different $\eta_b$
-are plotted in Fig. \ref{fig:block}b-f, showing that the code correctly preserve the block geometry in case of large viscosity contrast, with a block stiffer
-than the surrounding fluid (Fig. \ref{fig:block}f).
-
-The velocity in the centre of the falling block at $t=0$ is measured for all experiments and, since the velocity should increase with density contrast, the
-quantity $\bm{v}/(\rho_b-\rho_f)$ is plotted as function of the viscosity contrast. All results perfectly match with those from \citet{Gerya2010b} and they
-line up on a single curve, demonstrating that the code can correctly deal with large viscosity and density contrasts (Fig. \ref{fig:falling}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Falling%20blocks}.
-
-\begin{figure}
-\noindent\includegraphics[width=300px]{./Figures/Block.pdf}
-\centering
-\caption{Density evolution of the falling block experiment at $t=$ \SI{0}{\mega\year} (panel a) and $t=$ \SI{20}{\mega\year} for different viscosities of the
-block (panels b-f).}
-\label{fig:block}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Falling.pdf}
-\caption{Initial velocity relative to the density contrast at the centre of the falling block as function of the viscosity contrast between the block and the
-surrounding fluid.}
-\label{fig:falling}
-\end{figure}
-
-\subsection{2D time-dependent Stokes sphere with deformable free surface}\label{sec:sphere}
-This experiment is performed in a unit square domain with the gravity acceleration fixed to $g_y=-1$. The sphere has $\rho_s=2$ and $\eta_s=10^3$ and is
-initially centred at ($0.5;0.6$) with radius $R=0.123456789$. The fluid surrounding the sphere has $\rho_f=1$ and $\eta_f=1$ and occupies the domain for
-$y\leq 0.75$, while for $y>0.75$ the air has $\rho_a=0$ and $\eta_a=10^{-3}$. Velocity boundary conditions are set to free slip on all sides. The Courant
-number is set to 0.25. The experiments run for 200 s using grid resolutions from \num{150x150} to \num{512x512} elements, with 25 randomly distributed markers
-per element, and for different average schemes for the viscosity. The interface between the fluid and the air is tracked by means of the markers chain. Results
-in terms of velocity and pressure fields and topography variations are compared with results obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and they can be found at
-\url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere_fs2D} and
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Time_dependent_sphere}. Results in terms of the $v_{\textrm{rms}}$ in case of 
-an arithmetic average are also shown in Fig. \ref{fig:inst_spherefs}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_arithm.pdf}
-\caption{$v_{\textrm{rms}}$ for different grid resolutions as function of time for the 2D time-dependent Stokes sphere experiment in case of an arithmetic average.
-Results are compared with results obtained by ASPECT.}
-\label{fig:inst_spherefs}
-\end{figure}
-
-\subsection{Free surface stabilisation}\label{sec:stab}
-The benchmark is performed as discussed in \citet{Kaus2010a} and \citet{Thieulot2014}. The domain is square with $L_x=L_y=$ \SI{500}{\km} and a grid resolution
-of \num{200x200} elements, each of them containing 16 markers. A buoyant fluid with $\rho_1=$ \SI{3200}{\kg\per\cubic\metre} and $\eta_1=$ \SI{e20}{\pascal\s}
-is overlain by a denser fluid with $\rho_2=$ \SI{3300}{\kg\per\cubic\metre} and $\eta_2=$ \SI{e21}{\pascal\s}. The initial interface between the fluids has an
-initial sinusoidal shape of 5 km amplitude. Velocity boundary conditions are set to free slip conditions at the sides and no slip at the bottom, while the top
-is free surface. The experiment is performed with various fixed time steps, without the use of the Courant number. The vertical position of the free surface at
-$x=L_x$ is tracked for each simulation, with or without the stabilisation algorithm.
-
-The results show an instability (drunken sailor effect) increasing the time step in case that the stabilisation algorithm is not activated
-($dt=$ \SI{4200}{\year}, green line in Fig. \ref{fig:stab}), as already observed by \citet{Kaus2010a} and \citet{Thieulot2014}. Activating the algorithm the
-instability is fixed and the simulation remains stable using time step up to \SI{50000}{\year} (red line in Fig. \ref{fig:stab}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Stabilisation_algorithm}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Kaus_random.pdf}
-\caption{Evolution of the y coordinate in $x=L_x$ as function of time for different time steps, with and without the stabilisation algorithm.}
-\label{fig:stab}
-\end{figure}
-
-\subsection{Topography relaxation}\label{sec:crameri}
-This experiment is performed as in \citet{Crameri2012} using free surface and sticky air (Case 1). The domain is rectangular with $L_x=$ \SI{2800}{\km} and
-$L_y=$ \SI{700}{\km} in case of free surface and $L_y=$ \SI{800}{\km} in case of sticky air ($h_{st}=$ \SI{100}{\km}). The grid is composed by \num{256x64}
-elements for the free surface \num{560x320} elements for the sticky air. For both cases, there is a mantle of 600 km thickness with $\eta_m=$ \SI{e21}{\pascal\s}
-overlain by a highly viscous cosine-shaped lithosphere with thickness between 93 and 107 km and $\eta_l=$ \SI{e23}{\pascal\s}. Both mediums have $\rho=$
-\SI{3300}{\kg\per\cubic\metre}. In case of sticky air experiment, the air has $\rho_a=$ \SI{0}{\kg\per\cubic\metre} and $\eta_a=$ \SI{e18}{\pascal\s}. The
-gravity acceleration is set to $g_y=$ \SI{-10}{\metre\per\square\s}. Velocity boundary conditions are set to no slip at the bottom and free slip at the sides
-of the domain. In case of sticky air, velocities on top are set to free slip conditions.
-
-The maximum topography as function of time ($h(t)$) can be analytically derived in according to \[h(t)=h_0 \exp(\gamma t)\]
-where $h_0=7$ km is the initial topography, $\gamma=\num{-0.2139e-11}$ is the characteristic relaxation rate and $t$ is the time (black line in Fig.
-\ref{fig:crameri}). Maximum topography at the characteristic relaxation time $t_{rlx}=$ \SI{14.825}{\kilo\year} can be found to be $h_{rlx}=$ \SI{2576}{\m}.
-The results show that a viscosity of the air of \SI{e18}{\pascal\s} is needed in case of $h_{st}=$ \SI{100}{\km} to correctly describe the topography relaxation
-for this problem, as pointed out by \citet{Crameri2012} (purple line in Fig. \ref{fig:crameri}). The case with a true free surface well follows the analytical
-solution (orange line in Fig. \ref{fig:crameri}). In particular, at the relaxation time in case of the true free surface predicts a maximum topography of
-\SI{2570}{\m}, with an error of \SI{6}{\m} with respect to the analytical solution. Results are also compared with results obtained with a free surface in
-UNDERWORLD (\num{256x64} elements) and SULEC (\num{401x201} elements) (with courtesy from \citealp{Crameri2012}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Topography_relaxation}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Crameri.pdf}
-\caption{Maximum topography as function of time for the Crameri benchmark (coloured lines) in comparison with the analytical solution (black line) and results
-obtained with UNDERWORLD (yellow line) and SULEC (blue line).}
-\label{fig:crameri}
-\end{figure}
-
-\subsection{Spontaneous subduction}\label{sec:subduction}
-This experiment is performed as presented by \citet{Schmeling2008}, in case of both a sticky air and a true free surface and using different average schemes
-for the viscosities. The domain is rectangular with $L_x=$ \SI{3000}{\km} and $L_y=$ \SI{700}{\km}, a grid resolution of \num{256x64} elements with 50 markers
-per element and Courant number of 0.01. At the beginning of the simulation, a 100 km-thick lithospheric layer with $\rho_m=$ \SI{3300}{\kg\per\cubic\metre}
-and $\eta_m=$ \SI{e23}{\pascal\s} is located at the top of the domain for $x$ between 1000 and 3000 km, while the asthenospheric mantle has $\rho_m=$
-\SI{3200}{\pascal\s} and $\eta_m=$ \SI{e21}{\pascal\s}. In addition, a 200 km-depth lithospheric slab is already subducted in the mantle in order to have
-a spontaneous subduction. In case of sticky air, $L_y=750$ km and a 50 km-thick air layer with $\eta_a$ from \SIrange{e19}{e21}{\pascal\s} overlie the mantle.
-Velocity boundary conditions are set to free slip on the sides and on the bottom of the domain. In case of sticky air, velocities on the top boundary are set
-to free slip conditions as well.
-
-The maximum depth of the slab is tracked for all the simulations (Fig. \ref{fig:subduction}) and they are compared with results obtained by
-\citet{Schmeling2008} with models with regular grids and comparable grid resolutions (rectangles in Fig. \ref{fig:subduction}). Low viscous sticky air models
-(i.e., \SI{e19}{\pascal\s}) enlighten a strong dependence of the sinking velocity with the chosen average scheme (dashed lines in Fig. \ref{fig:subduction}),
-while high viscous sticky air models (i.e., \SI{e21}{\pascal\s}) show a higher resistance at the trench, underestimating the correct solution (dotted lines in
-Fig. \ref{fig:subduction}). A high resistance at the trench can be also observed in case of a true free surface for the low resolution grid used (continuous
-lines in Fig. \ref{fig:subduction}), as pointed out by \citet{Schmeling2008}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Spontaneous_subduction}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Subduction.pdf}
-\caption{Maximum depth of the slab as function of time for all the simulations of the \citet{Schmeling2008} benchmark, in case of harmonic, geometric and
-arithmetic means (red, green and blue lines, respectively) using a sticky air or a true free surface (discontinuous and continuous lines, respectively).
-Red, green and blue rectangular areas indicate the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{500}{\km} in case of sticky air,
-while the grey rectangular area indicates the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{600}{\km} in case of true free surface.}
-\label{fig:subduction}
-\end{figure}
-
-\subsection{The slab detachment}\label{sec:slab}
-The slab detachment benchmark is performed as described by \citet{Schmalholz2011} and \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{1000}{\km}
-and $L_y=$ \SI{660}{\km} and a grid resolution of \num{256x256} elements. A non-linear viscous T-shaped layer with $\rho_l=$ \SI{3300}{\kg\per\cubic\metre} is
-placed at the top of the domain and surrounded by a linear viscous fluid with $\rho_f=$ \SI{3150}{\kg\per\cubic\m}. The top layer is 80 km-thick and a 250
-km-long and 80 km-wide slab is placed at $x=L_x/2$ (Fig. \ref{fig:slab}a). The effective viscosity of the top layer is given by
-\[\eta_{eff}=\eta_0 I_2^{\frac{(1-n)}{n}}\]
-with $\eta_0=$ \SI{4.75e11}{\pascal\s} and $n=4$, while the surrounding fluid has $\eta_f=$ \SI{e21}{\pascal\s}. The effective viscosity is capped between
-$\eta_{min}=$ \SI{e21}{\pascal\s} and $\eta_{max}=$ \SI{e25}{\pascal\s}. Velocity boundary conditions are set to free slip at the top and the bottom, and to
-no slip at the sides of the domain.
-
-The sinking of the slab determines a decrease of the effective viscosity in the T-shaped layer during the first part of the evolution (Fig. \ref{fig:slab}a),
-as consequence of local strain rates, while effective viscosities increase up to $\eta_{max}$ in the top layer after approximately \SI{20}{\mega\year}, when
-necking is complete (Fig. \ref{fig:slab}b).
-
-The width of necking is tracked by means of markers position at the side of the slab. Necking width and time are normalised in function of the initial slab
-(\SI{80}{\km}) and the characteristic time ($t_c=$ \SI{7.1158e14}{\s}) \citep{Schmalholz2011,Glerum2018}. The results in case of arithmetic, geometric and
-harmonic mean of the viscosity are shown in Fig. \ref{fig:necking}, compared with results from \citet{Glerum2018} for which an infinity norm mean was used.
-The influence of the average schemes agrees with results shown in Section \ref{sec:subduction}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Slab_detachment}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Slab.pdf}
-\caption{Effective viscosity for the slab detachment benchmark at the beginning of the evolution (panel a) and when necking is complete (panel b).}
-\label{fig:slab}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Necking.pdf}
-\caption{Normalised width of the necking of the slab detachment benchmark as function of normalised time.}
-\label{fig:necking}
-\end{figure}
-
 \subsection{The indenter experiment}\label{sec:indenter}
 The indenter benchmark simulate a rigid punch on a purely plastic von Mises material, for which exists an analytical solution
-\citep{Thieulot2008,Thieulot2014,Glerum2018}. The domain is a unit square with a grid resolution of \num{256x256} elements. The gravity acceleration is set
+\citep{Thieulot2008,Thieulot2014,Glerum2018}. The domain is a unit square with a grid resolution of $256\times256$ elements. The gravity acceleration is set
 to $\bm{g}=\bm{0}$ and effective viscosity is capped using $\eta_{min}=10^{-4}$ and $\eta_{max}=10^3$. The tolerance for the convergence of the non-linear
@@
 \end{figure}
-
-\subsection{The brick experiment}\label{sec:brick}
-An instantaneous version of the brick benchmark is performed to verify the correctness of the pressure-dependent plasticity for different angle of internal
-friction, as proposed by \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{40}{\km} and $L_y=$ \SI{10}{\km} and a grid resolution of \num{512x128}
-elements. The gravity acceleration is set to $g_y=$ \SI{-10}{\m\square\s} and effective viscosity is capped using $\eta_{min}=$ \SI{e19}{\pascal\s} and
-$\eta_{max}=$ \SI{e26}{\pascal\s}. The tolerance for the convergence of the non-linear solution is $tol=10^{-7}$, with a maximum number of non-linear iterations
-set to 1000. A 800 m-wide and 400 m-height inclusion with $\eta_b=$ \SI{e20}{\pascal\s} and $\rho_b=$ \SI{2700}{\kg\per\cubic\m} is placed at the bottom of the
-domain at $x=L_x/2$. The inclusion is surrounded by a non-linear viscous medium with $\rho_m=$ \SI{2700}{\kg\per\cubic\m}, an initial viscosity of
-\SI{e23}{\pascal\s}, a linear viscous viscosity of \SI{e25}{\pascal\s} and a cohesion of \SI{40}{\mega\pascal}. Velocities are set to free slip conditions at
-the bottom of the domain and the top is open. Velocities on sides of the domain are fixed to $u=$ \SI{+-2e-11}{\m\per\s} and $v=0$. The experiment is performed
-in compressional and extensional contexts with an angle of internal friction $\phi$ of the non-linear medium variable between 0° and 30°.
-
-As expected, two shear bands stem from the inclusion with variable angles in relation to both the dynamics context and the internal friction angle. Shear band
-angles formed at 45° for $\phi=0$° in both extensional and compressional contexts (Fig. \ref{fig:brick_beam}a), while they are different in case of $\phi\neq0$°.
-Shear band angles in case of $\phi=20$° are shown for compression and extension in Fig. \ref{fig:brick_beam}b  and c, respectively. Values of shear band angles
-as function of internal friction angles from 0° to 30° are extracted at different depths and minimum and maximum values are plotted in Fig. \ref{fig:brick},
-in comparison with theoretical Roscoe, Arthur and Coulomb shear band angles. For all tests the tolerance for nonlinear convergence is defined $tol=10^{-7}$
-and none converges before the maximum number of iterations ($it_{max}=1000$) is reached. However, tests with internal friction angles up to 15° show a constant
-decrease of the velocity residuals, which can not be observed for for tests with higher internal friction angles (Fig. \ref{fig:convergence}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Brick_experiment}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Brick_Beam.pdf}
-\caption{Shear band angles predicted for the brick experiment for $\phi=0$° (panel a) and for $\phi=20$° in case of compression and extension (panels b and c,
-respectively).}
-\label{fig:brick_beam}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Brick.pdf}
-\caption{Shear band angles predicted for the brick experiment in case of compressional and extensional contexts as function of different internal angle of
-friction (continuous black line and red dots), compared with theoretical Roscoe, Artur and Coulomb angles (discontinuous black lines). Red lines indicate the
-range of angles calculated at different depths.}
-\label{fig:brick}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \noindent\includegraphics[width=500px]{./Figures/convergence.pdf}
-  \caption{Convergence rates for all tests. Normalised residuals (left column) are compared with velocity residuals (right column) for extensional (first row)
-  and compressional (second row) settings.}
-  \label{fig:convergence}
-  \end{figure}
-
-\subsection{Advection stabilisation}\label{sec:advection}
-The 1D advection problem proposed by \citet{Donea2003} and \citet{Thieulot2011} is performed to verify the effectiveness of the SUPG method to stabilise the
-advection term of the energy equation. The domain is a 1D segment with $L_x=1$ composed by 50 elements and a discontinuity in the thermal field placed at
-$x=0.25$. Temperature is set to 1 for $x \leq 0.25$ and to 0 for $x > 0.25$. Velocity is set to ${u}=1$ in the entire domain. The simulation is performed for
-250 time steps, with $dt=0.002$, so the thermal discontinuity should be at $x=0.75$ at the end of the simulation.
-
-Temperature profiles at the end of the simulation are shown in \ref{fig:advection} as function of the dimensionless coefficient $\gamma=\tau \bm{v}/h$. In case
-of the classic Galerkin method ($\gamma=0$, blue line in \ref{fig:advection}) the final thermal profile is characterised by strong oscillations, which are
-eliminated in case of the SUPG method ($\gamma=0.045$, orange line in \ref{fig:advection}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Advection_stabilisation}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Advection.pdf}
-\caption{Temperature profile as function of \textit{x} for the advection stabilisation benchmark. Purple line indicates the initial temperature profile; the
-green line indicates the analytical temperature profile after 250 time steps; blue line indicates the temperature profile after 250 time steps in case of the
-classic Galerkin method ($\gamma=0$); orange line indicates the temperature profile after 250 time steps in case of the SUPG method ($\gamma=0.045$).}
-\label{fig:advection}
-\end{figure}
-
-\subsection{Shear heating}\label{sec:simple_shear}
-This exercise is performed in an unit square domain composed by \num{128x128} elements. The velocity field is prescribed on the entire domain with
-$\bm{v}=(L_y-y)y\bm{e_x}$; viscosity, density and specific heat are fixed to 1, while thermal conductivity and radiogenic energy are fixed to 0. Therefore,
-the energy equation (Eq. \ref{eq:energy}) can be simplified as \[\frac{\partial T}{\partial t}=H_s\] and fixing $T(t=0)=0$, the temperature can be find as
-\[T(t)=H_s t\]
-In this case we have
-\begin{eqnarray}
-\dot{\epsilon_{xx}}&=&\frac{\partial u}{\partial x}=0 \nonumber \\
-\dot{\epsilon_{yy}}&=&\frac{\partial v}{\partial y}=0 \nonumber \\
-\dot{\epsilon_{xy}}&=&\frac{1}{2}\left(\frac{\partial v}{\partial x}+\frac{\partial u}{\partial y}\right)=\frac{1}{2}(L_y-2y)\nonumber
-\end{eqnarray}
-and, simplifying Eq. \ref{eq:shear}, shear heating can be calculated as \[H_s(x,y)=(1-2y)^2\]
-
-The solution predicted by the model in terms of velocity, temperature and shear heating match well with the analytical solutions (Fig. \ref{fig:analytical_en}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Simple_shear}.
-
-%\begin{figure}
-%\centering
-%\noindent\includegraphics[width=400px]{./Figures/Energy_Thieulot.pdf}
-%\caption{Temperature and shear heating predicted by the model after 1 time step.}
-%\label{fig:en_thieulot}
-%\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=500px]{./Figures/Analytical_Shear.pdf}
-\caption{Velocity (panel a), temperature (panel b) and shear heating (panel c) as function of the \textit{y} coordinate for the simple shear experiment. The
-solutions predicted by the model (red lines) are compared with the analytical solutions (dashed blue lines).}
-\label{fig:analytical_en}
-\end{figure}
-
-\subsection{Shear and adiabatic heating}\label{sec:shear}
-This problem is performed as presented in Exercise 9.4 in \citet{Gerya2010b}. Two materials are vertically separated in a rectangular domain with $L_x=$
-\SI{1000}{\km}, $L_y=$ \SI{1500}{\km} and a grid resolution of \num{30x20} elements. Constant thermal coefficient expansion ($\alpha=$ \SI{3e-5}{\per\kelvin}),
-temperature ($T=$ \SI{1300}{\kelvin}) and gravity acceleration ($g_y=$ \SI{-10}{\m\square\s}) are assumed in the whole domain. Fluid 1 (on the left side of the
-domain) has $\rho_1=$ \SI{3200}{\kg\per\cubic\m} and $\eta_1=$ \SI{e20}{\pascal\s}; fluid 2 (on the right side of the domain) has $\rho_2=$
-\SI{3300}{\kg\per\cubic\m} and $\eta_2=$ \SI{e22}{\pascal\s}. Velocity boundary conditions are set to free slip on all sides of the domain.
-
-As shown in Fig. \ref{fig:energy}, both shear and adiabatic heating predicted by the code (first row in Fig. \ref{fig:energy}) well recreate the results
-obtained by means of example ${Shear\_adiabatic\_heating.m}$ from \citet{Gerya2010b} (second row in Fig. \ref{fig:energy}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Adiabatic%2BShear}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Energy.pdf}
-\caption{Comparison between shear (first column), adiabatic (second column) and total (third column) energy predicted by the code (panels a, b and c) and those
-created using example ${Shear\_adiabatic\_heating.m}$ from Exercise 9.4 in \citet{Gerya2010b} (panels d, e and f).}
-\label{fig:energy}
-\end{figure}
-
-\subsection{Mantle convection}\label{sec:mantle}
-This problem is performed as presented by \citet{Blankenbach1989} (constant viscosity cases) and \citet{Thieulot2014}, in a 2D unit square domain with gravity
-acceleration $g_y=10^{10}Ra$. The experiment is performed with three different Rayleigh numbers ($Ra=10^4$, $10^5$ and $10^6$) and with different grid
-resolution (between \num{32x32} and \num{128x128} elements). The fluid has constant viscosity, initial density, heat capacity, thermal conductivity
-($\eta= \rho_0= C_p = k=1$), reference temperature ($T_0=0$) and thermal expansion coefficient ($\alpha=10^{-10}$). Temperatures are set to 0 on top and 1 on
-bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\]
-
-The solution generated by the code in terms of $v_{\textrm{rms}}$ (calculated as in Sec. \ref{sec:error}) and Nu as function of time are reported for all the
-simulations in Fig. \ref{fig:mantle}. At the steady state, $v_{\textrm{rms}}$ and Nu of all simulations converge well toward the values from \citet{Blankenbach1989},
-with lower errors for higher resolution grids (Table \ref{tab:mantle}). All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Mantle_convection}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Convection.pdf}
-\caption{$v_{\textrm{rms}}$ (panels a, c and e) and Nu (panels b, d and f) for the mantle convection benchmark as function of time for different grid resolution.
-Panels a and b show the results for $Ra=10^4$; Panels c and d show the results for $Ra=10^5$; panels e and f show the results for $Ra=10^6$. Purple lines
-indicate the convergence values for the $v_{\textrm{rms}}$ and the Nu from \citet{Blankenbach1989} at the steady state.}
-\label{fig:mantle}
-\end{figure}
-
-\begin{table}[H]
-\caption{Comparison between $v_{\textrm{rms}}$ and Nu predicted by the code for the mantle convection experiment and same values as reported in literature.}
-\centering
-\small
-\begin{tabular}{|>{\centering}m{0.09\textwidth}|>{\centering}m{0.05\textwidth}|>{\centering}m{0.24\textwidth}|>{\centering}m{0.15\textwidth}|
-  >{\centering}m{0.1\textwidth} >{\centering}m{0.1\textwidth} >{\centering\arraybackslash}m{0.1\textwidth}|}
-\toprule
-& & \multirow{2}{*}{\citet{Blankenbach1989}} & \citet{Thieulot2014} & \multicolumn{3}{c|}{FALCON}  \\
- & & & \num{200x200} & \num{32x32} & \num{64x64} & \num{128x128}  \\
-\midrule
-\multirow{2}{*}{$Ra=10^4$} & $v_{\textrm{rms}}$ & $42.864947 \pm 0.000020$ & 42.867 & 42.83226 & 42.852793 & 42.861394  \\
-           & Nu      & $4.884409 \pm 0.000010$  & 4.882  & 4.781297 & 4.857475  & 4.877573   \\
-\hline
-\multirow{2}{*}{$Ra=10^5$} & $v_{\textrm{rms}}$ & $193.21454 \pm 0.00010$  & 193.255& 193.872643 & 193.377472 & 193.252290  \\
-           & Nu      & $10.534095 \pm 0.000010$ & 10.507 & 9.602514 & 10.270735  & 10.465629   \\
-\hline
-\multirow{2}{*}{$Ra=10^6$} & $v_{\textrm{rms}}$ & $833.98977 \pm 0.00020$  & 834.712& 848.091176 & 837.767911 & 834.945793  \\
-           & Nu      & $21.972465 \pm 0.000020$ & 21.695 & 15.999266 & 19.703682  & 21.306939   \\
-\bottomrule
-\end{tabular}
-\label{tab:mantle}
-\end{table}
-
-\subsection{Viscoplastic mantle convection}\label{sec:visco_mantle}
-This benchmark is performed as Cases 1-5a described in \citet{Tosi2015} in a 2D unit square domain with two different grid resolutions of \num{32x32} 
-and \num{100x100} elements. Reference viscosity, reference density, heat capacity, thermal conductivity and thermal expansion coefficient are fixed to a constant
-value of 1 ($\eta_0= \rho_0= C_p = k=\alpha=1$), while the gravity has been chosen as $g_y=10^2$, in order to obtain a Rayleigh number $Ra=10^2$. Temperatures
-are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\].
-The viscosity field $\eta$ is calculated as the harmonic average between a linear part $\eta_{lin}$ that depends on temperature only or on temperature and 
-depth $y$ and a nonlinear-plastic part $\eta_{pl}$ that depends on the strain rate $\dot{\bm{\epsilon}}$, as follows
-\[\eta(T,y,\dot{\bm{\epsilon}})=2\left(\frac{1}{\eta_{lin}(T,y)}+\frac{1}{\eta_{pl}(\dot{\bm{\epsilon}})}\right)^{-1}\]
-The linear and the nonlinear-plastic parts of the viscosity are calculated following \citet{Tosi2015} as follows
-\[\eta_{lin}(T,y)=\exp(-\ln(\Delta\eta_T)T+\ln(\Delta\eta_y)y)\]
-\[\eta_{pl}(\dot{\bm{\epsilon}})=\eta^*+\frac{\sigma_Y}{\sqrt{\dot{\bm{\epsilon}}:\dot{\bm{\epsilon}}}}\]
-where $\Delta\eta_T$, $\Delta\eta_y$, $\eta^*$ and $\sigma_Y$ are parameters chosen as listed in Table \ref{tab:tosi_setup}.
-
-For each case, some diagnostic quantities measured at the steady state are reported in Table \ref{tab:visco_mantle} (see Section \ref{sec:error} for 
-the computation of all the quantities), showing a well fit with results described in \citet{Tosi2015}. Data and charts for all cases can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Viscoplastic_mantle_convection}.
-
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{white!100!}{gray!25!}
-%\begin{table}[H]
-%  \caption{Results of the viscoplastic mantle convection benchmark for Cases 1-4 compared with results obtained with ELEFANT and YACC (from \citealp{Tosi2015}).}
-%  \centering
-%  \resizebox{\textwidth}{!}{
-%\begin{tabular}{|c|ccc|ccc|ccc|ccc|}
-%  \toprule
-%    & \multicolumn{3}{c|}{\textbf{Case 1}} & \multicolumn{3}{c|}{\textbf{Case 2}} & \multicolumn{3}{c|}{\textbf{Case 3}} &
-%    \multicolumn{3}{c|}{\textbf{Case 4}} \\
-%   & FALCON & ELEFANT& YACC & FALCON& ELEFANT& YACC & FALCON & ELEFANT & YACC & FALCON & ELEFANT & YACC \\
-%  \midrule
-%   $\left\langle T \right\rangle$ & 0.7759 & 0.7758 & 0.7767 & 0.6043 & 0.6036 & 0.6058 & 0.7275 & 0.7275 & 0.7286 & & 0.5277 & 0.5289\\
-%    Nu$^{top}$ & 3.4280 & 3.4214 & 3.4298 & 8.5242 & 8.5115 & 8.5278 & 3.0368 & 3.0347 & 3.0374 && 6.5912 & 6.5572 \\
-%    Nu$^{bottom}$ & 3.3168 & 3.313 & 3.3143 & 8.4599 & 8.4465 & 8.3990 & 3.0086 & 2.9908 & 2.9628 && 6.5834 & 6.5243 \\
-%    $v_{\textrm{rms}}$ & 249.901 & 249.134 & 251.7997 & 141.2451 & 140.7087 & 142.2020 & 100.2776 & 100.1208 & 100.9467 && 79.1105 & 79.6202 \\
-%    $v_{\textrm{rms}}^{top}$ & 1.8907 & 1.8642 & 1.8298 & 104.7928 & 104.1228 & 105.1767 & 2.0832 & 2.0652 & 2.0374 && 74.7596 & 75.8414 \\
-%    $u_{max}^{top}$ & 2.6340 & 2.6119 & 2.5516 & 121.9212& 121.766 & 122.3238 & 2.9086 & 2.9019 & 2.8458 && 88.9146 & 89.2940 \\
-%    $\eta_{min}$ & - & - & - & $2.0303\times10^{-5}$ & $2.0600\times10^{-5}$ & $1.9853\times10^{-5}$ & $4.8129\times10^{-5}$ & $4.8080\times10^{-5}$ & $4.7907\times10^{-5}$ &
-%    & $1.9860\times10^{-4}$ & $1.9174\times10^{-4}$ \\
-%    $\eta_{max}$ & - & - & - & 1.3325 & 1.85 & 1.6095 & 0.9360 & 0.9023 & 1 && 1.52 & 1.6773 \\
-%    $\left\langle W \right\rangle$ & 2.4299 & 2.4316 & 2.4583 & 7.5648 & 7.5665 & 7.6222 & 2.0387 & 2.0384 & 2.0400 && 5.6216 & 5.6512 \\
-%    $\left\langle \Phi \right\rangle/Ra$ & 2.4311 & 2.4276 & 2.4333 & 7.5661 & 7.5585 & 7.6194 & 2.0386 & 2.0370 & 2.0335 && 5.6182 & 5.6463 \\
-%    $\delta$ & 0.05\% & 0.16\% & 1.02\% & 0.02\% & 0.11\% & 0.04\% & $<$0.01\% & 0.07\% & 0.32\% && 0.06\% & 0.09\% \\
-%  \bottomrule
-%  \end{tabular}}
-%  \label{tab:visco_mantle}
-%  \end{table}
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{2}{white!100!}{gray!25!}
-\begin{table}[H]
-  \caption{Benchmark cases and corresponding parameters.}
-  \centering
-  \small
-\begin{tabular}{|>{\centering}m{0.08\textwidth}|>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}
-  >{\centering}m{0.05\textwidth} >{\centering}m{0.05\textwidth} |>{\centering\arraybackslash}m{0.2\textwidth}|}
-  \toprule
-    Case & $Ra$ & $\Delta\eta_T$ & $\Delta\eta_y$ & $\eta^*$ & $\sigma_Y$ & Convective regime \\
-  \midrule
-  \rowcolor{gray!25!}
-    1 & $10^2$ & $10^5$ & 1 & - & - & Stagnant lid \\
-    2 & $10^2$ & $10^5$ & 1 & $10^{-3}$ & 1 & Mobile lid  \\
-    \rowcolor{gray!25!}
-    3 & $10^2$ & $10^5$ & 10 & - & - & Stagnant lid  \\
-    4 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 1 & Mobile lid \\
-    \rowcolor{gray!25!}
-    5 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 4 & Periodic \\
-  \bottomrule
-  \end{tabular}
-  \label{tab:tosi_setup}
-  \end{table}
-
-\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{gray!25!}{white!100!}
-\begin{table}
-  \caption{Results of the viscoplastic mantle convection benchmark for different grid resolutions in Cases 1-5 compared with results obtained with ELEFANT
-  and YACC ($100\times100$ elements, from \citealp{Tosi2015}) and in Stone 28 ($32\times32$ elements, from
-  \url{https://github.com/cedrict/fieldstone/tree/master/python_codes/fieldstone_28}).}
-  \centering
-  \resizebox{\textwidth}{!}{
-\begin{tabular}{|cc|ccccccccccc|}
-  \toprule
-  \multicolumn{13}{|c|}{\textbf{Case 1}} \\
-  \midrule
-  Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-  & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-  \hline
-  \rowcolor{gray!25!}
-    YACC & $100\times100$ & 0.7767 & 3.4298 & 3.3143 & 251.7997 & 1.8298 & 2.5516 & - & - & 2.4583 & 2.4333 & 1.02\% \\
-    ELEFANT & $100\times100$ & 0.7758 & 3.4214 & 3.313 & 249.134 & 1.8642 & 2.6119 & - & - & 2.4316 & 2.4276 & 0.16\% \\
-    \rowcolor{gray!25!}
-    Stone 28 & $32\times32$ & 0.7737 & 3.3987 & - & 243.872 & - & - & - & - & - & - & - \\
-%    FALCON & $32\times32$ & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & - & - & 2.4745 & 2.4864 & 0.48\% \\
-%    \rowcolor{gray!25!}
-    FALCON & $100\times100$ & 0.7759 & 3.4280 & 3.3168 & 249.901 & 1.8907 & 2.6340 & - & - & 2.4299 & 2.4311 & 0.05\% \\
-    \midrule
-    \multicolumn{13}{|c|}{\textbf{Case 2}} \\
-    \midrule
-    Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-    & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-    \hline
-    \rowcolor{gray!25!}
-     YACC & $100\times100$ & 0.6058 & 8.5278 & 8.3990 & 142.2020 & 105.1767 & 122.3238 & $1.9853\times10^{-5}$ & 1.6095 & 7.6222 & 7.6194 & 0.04\% \\
-     ELEFANT & $100\times100$ & 0.6034 & 8.5115 & 8.4465 & 140.8067 & 104.1228 & 121.766 & $2.0600\times10^{-5}$ & 1.85 & 7.5665 & 7.5585 & 0.11\% \\
-     \rowcolor{gray!25!}
-     Stone 28 & $32\times32$ & 0.6052 & 8.1715 & - & 141.518 & - & - & - & - & - & - & - \\
-%     FALCON & $32\times32$ & 0.6454 & 8.1422 & 7.4020 & 156.6908 & 95.7226 & 116.9302 & $2.2572\times10^{-5}$ & 0.7194 & 7.4289 & 7.4452 & 0.22\% \\
-%     \rowcolor{gray!25!}
-     FALCON & $100\times100$ & 0.6044 & 8.5266 & 8.4590 & 141.3684 & 104.8867 & 122.0470 & $2.0333\times10^{-5}$ & 1.2914 & 7.5707 & 7.5718 & 0.01\% \\
-     \midrule
-     \multicolumn{13}{|c|}{\textbf{Case 3}} \\
-     \midrule
-     Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-     & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-     \hline
-     \rowcolor{gray!25!} 
-     YACC & $100\times100$ & 0.7286 & 3.0374 & 2.9628 & 100.9467 & 2.0374 & 2.8458 & $4.7907\times10^{-5}$ & 1 & 2.0400 & 2.0335 & 0.32\% \\
-      ELEFANT & $100\times100$ & 0.7275 & 3.0347 & 2.9908 & 100.1208 & 2.0652 & 2.9019 & $4.8080\times10^{-5}$ & 0.9023 & 2.0384 & 2.037 & 0.07\% \\
-      \rowcolor{gray!25!}
-      Stone 28 & $32\times32$ & 0.7277 & 3.0347 & - & 100.018 & - & - & - & - & - & - & - \\
-%      FALCON & $32\times32$ & 0.7281 & 3.0618 & 2.8132 & 101.0092 & 2.1630 & 3.0178 & $4.8500\times10^{-5}$ & 0.7845 & 2.0699 & 2.0695 & 0.02\% \\
-%      \rowcolor{gray!25!}
-      FALCON & $100\times100$ & 0.7275 & 3.0368 & 3.0086 & 100.2776 & 2.0832 & 2.9086 & $4.8129\times10^{-5}$ & 0.9360 & 2.0387 & 2.0386 & $<$0.01\% \\
-      \midrule
-      \multicolumn{13}{|c|}{\textbf{Case 4}} \\
-      \midrule
-      Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-      & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-      \hline
-      \rowcolor{gray!25!} 
-      YACC & $100\times100$ & 0.5289 & 6.5572 & 6.5243 & 79.6202 & 75.4814 & 89.2940 & $1.9174\times10^{-4}$ & 1.6773 & 5.6512 & 5.6463 & 0.09\% \\
-       ELEFANT & $100\times100$ & 0.5277 & 6.5912 & 6.5834 & 79.1105 & 74.7596 & 88.9146 & $1.9860\times10^{-4}$ & 1.5200 & 5.6216 & 5.6182 & 0.06\% \\
-       \rowcolor{gray!25!}
-       Stone 28 & $32\times32$ & 0.5286 & 6.4036 & - & 79.4746 & - & - & - & - & - & - & - \\
-%       FALCON & $32\times32$ & 0.5334 & 6.4679 & 6.3435 & 80.6153 & 76.0511 & 90.2058 & $2.0393\times10^{-4}$ & 0.6453 & 5.6676 & 5.6698 & 0.04\% \\
-%       \rowcolor{gray!25!}
-       FALCON & $100\times100$ & 0.5279 & 6.5897 & 6.5830 & 79.1220 & 75.0715 & 88.8484 & $1.9693\times10^{-4}$ & 1.4133 & 5.6127 & 5.6113 & 0.03\% \\
-       \midrule
-       \multicolumn{13}{|c|}{\textbf{Case 5}} \\
-       \midrule
-       Code & Resolution & $P$ & $\left\langle T \right\rangle_{min}$ & $\left\langle T \right\rangle_{max}$ & Nu$^{top}_{min}$ & Nu$^{top}_{max}$ &
-       $v_{\textrm{rms}}^{min}$ & $v_{\textrm{rms}}^{max}$ & \multicolumn{2}{c}{$(\left\langle \Phi \right\rangle/Ra)_{min}$} & 
-       \multicolumn{2}{c|}{$(\left\langle \Phi \right\rangle/Ra)_{max}$} \\
-       \hline
-       \rowcolor{gray!25!}
-         YACC & $100\times100$ & 0.0792 & 0.6543 & 0.6722 & 2.6950 & 7.2792 & 41.9813 & 98.8230 & \multicolumn{2}{c}{1.4347} & 
-         \multicolumn{2}{c|}{9.2748} \\
-         ELEFANT & $100\times100$ & 0.076 & 0.6514 & 0.6693 & 2.6808 & 7.3506 & 41.2578 & 100.089 & \multicolumn{2}{c}{1.3324} & 
-         \multicolumn{2}{c|}{9.5289} \\
-         \rowcolor{gray!25!}
-         Stone 28 & $32\times32$ & 0.0764 & 0.6521 & 0.6697 & 2.6912 & 7.0763 & 41.61 & 98.183 & \multicolumn{2}{c}{-} & 
-         \multicolumn{2}{c|}{-} \\
-     %    FALCON & $32\times32$ & & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & \multicolumn{2}{c}{-} & 
-     %     \multicolumn{2}{c|}{-} \\
-     %    \rowcolor{gray!25!}
-         FALCON & $100\times100$ & 0.0775 & 0.6519 & 0.6697 & 2.6847 & 7.3629 & 41.5256 & 99.5466 & \multicolumn{2}{c}{1.3399} & 
-         \multicolumn{2}{c|}{9.3993} \\
-  \bottomrule
-  \end{tabular}}
-  \label{tab:visco_mantle}
-  \end{table}
-
-\subsection{Thin layer entrainment}\label{sec:thin}
-This experiment is performed as originally proposed by \citet{vanKeken1997} with two incompressible fluids in a rectangular domain with $L_x=2$ and $L_y=1$
-and gravity acceleration $g_y=10^{10}Ra$. The Rayleigh number and the compositional Rayleigh number are fixed to $Ra=300000$ and $Rc=450000$, respectively.
-Both fluids have constant viscosity, thermal conductivity, specific heat ($\eta=\rho=C_p=1$) and thermal expansion coefficient ($\alpha=10^{-10}$). Fluid 1
-has a density $\rho_1=1$, while fluid 2 is denser ($\rho_2=\rho_1+1.5 \cdot 10^{-10}$) and is placed at the bottom of the domain, for $y \leq 0.025$.
-Temperature are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides of the domain. The initial
-temperature field is given by
-\[T(x,y)=T_u(x,y)+T_l(x,y)+T_r(x,y)+T_s(x,y)-\frac{3}{2}\]
-with
-\begin{eqnarray}
-T_u(x,y)&=&\frac{1}{2}erf\left(\frac{1-y}{2}\sqrt{\frac{u_0}{x}}\right)\nonumber \\
-T_l(x,y)&=&1-\frac{1}{2}erf\left(\frac{y}{2}\sqrt{\frac{u_0}{L_x-x}}\right)\nonumber \\
-T_r(x,y)&=&\frac{1}{2}+\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{y+1}} \exp\left(-\frac{x^2u_0}{4y+4}\right)\nonumber \\
-T_s(x,y)&=&\frac{1}{2}-\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{2-y}} \exp\left(-\frac{(L_x-x)^2u_0}{8-4y}\right)\nonumber
-\end{eqnarray}
-and
-\begin{eqnarray}
-u_0&=&\frac{L_x^{7/3}}{(1+L_x^4)^{2/3}}\left(\frac{Ra}{2\sqrt{\pi}}\right)^{\frac{2}{3}}\nonumber \\
-Q&=&2\sqrt{\frac{L_x}{\pi u_0}}\nonumber
-\end{eqnarray}
-
-The experiment is performed with two grid resolution (\num{125x40} and \num{200x80} elements), with 100 markers per element and CFL of 0.25. $v_{\textrm{rms}}$
-(see Sec. \ref{sec:error}) calculated for both the simulations match well with results from \citet{vanKeken1997} and obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and ELEFANT \citep{Thieulot2014}, obtained with similar grid resolutions
-(Fig. \ref{fig:thin}). In addition, a variety of simulations with different resolutions and aspect ratios are performed to verify that the sensitivity of the
-initial velocity field. Results match well with results from \citet{vanKeken1997} (in the grey area) and \citet{Thieulot2014} (Fig. \ref{fig:thin_initial}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Thin_layer}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_thin.pdf}
-\caption{$v_{\textrm{rms}}$ for the thin layer experiment as function of time for different grid resolution. Results are compared with results obtained by
-\citet{vanKeken1997} (black lines) and with ASPECT and ELEFANT.}
-\label{fig:thin}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrmszero.pdf}
-\caption{$v_{\textrm{rms}}$ at $t=0$ for the thin layer experiment as function of the element size for different aspect ratios of the elements. Results are compared
-with results obtained by \citet{vanKeken1997} (black dots in the grey area) and with ASPECT and ELEFANT.}
-\label{fig:thin_initial}
-\end{figure}
-
-\subsection{Sinking hydrated cylinder}\label{sec:quinquis}
-This experiment is performed as explained in \citet{Quinquis2014} and simulates the sinking of a cold, hydrated cylinder into a warm, dry mantle. The domain
-is square with $L_x=L_y=$ \SI{300}{\km}, a grid resolution of \num{300x300} elements and 25 markers per element. The sinking cylinder is located at $x=$
-\SI{150}{\km} and $y=$ \SI{170}{\km} with a radius of \SI{20}{\km} and has $\rho_c=$ \SI{3250}{\kg\per\cubic\m}, $\eta_c=$ \SI{e23}{\pascal\s}, a thermal
-conductivity $k_c=$ \SI{4.5}{\watt\per\m\per\kelvin},a specific heat $C_{p_c}=$ \SI{1250}{\joule\per\kg\per\kelvin}; the surrounding mantle has $\rho_m=$
-\SI{3200}{\kg\per\cubic\m}, $\eta_m=$ \SI{e20}{\pascal\s}, a thermal conductivity $k_m=$ \SI{105}{\watt\per\m\per\kelvin} and specific heat $C_{p_m}=$
-\SI{1250}{\joule\per\kg\per\kelvin}; a 58 km-thick lithosphere overlies the mantle and has $\rho_l=$ \SI{3200}{\kg\per\cubic\m}, $\eta_l=$ \SI{e23}{\pascal\s},
-a thermal conductivity $k_l=$ \SI{4.5}{\watt\per\m\per\kelvin}, specific heat $C_{p_l}=$ \SI{750}{\joule\per\kg\per\kelvin} and a thermal diffusivity of
-\SI{e-6}{\square\m\per\s}. Velocity boundary conditions are set to free slip on all sides of the domain. Temperature increase linearly in the lithosphere,
-from 0° to 1300°C, and in the mantle, from 1300° to 1360.5°C. The initial temperature of the cylinder is fixed to 400°C. Throughout the evolution, temperature
-is fixed to 0° and 1360.5°C at the top and bottom of the domain, respectively. The initial bound water content is imposed at 2 wt.\% in the cylinder and at
-0 wt.\% in both the mantle and the lithosphere. The maximum amount of water in the mantle is fixed to 0.2 wt.\%, while maximum water content in the cylinder
-and in the lithosphere is function of pressure and temperature and is calculated using a serpentinized harzburgite with Perple\_X, as in \citet{Quinquis2014}.
-
-As shown in \citet{Quinquis2014}, the model is characterised by a progressive dehydration in the external portion of the cylinder, due to the increase of
-temperature, with a consequent vertical migration of free water that hydrates the mantle above the cylinder, up to the lithosphere. The distribution of
-bound and free water during the evolution of the experiment is shown in Fig. \ref{fig:quinquis} (left and right columns, respectively). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Hydration}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Hydration.pdf}
-\caption{Distribution of bound and free water (left and right columns, respectively) at $t=0.5$ and $t=$ \SI{1.25}{\mega\year} (first and second rows,
-respectively) for the sinking hydrated cylinder experiment. Only markers with bound or free water not equal to 0 are plotted.}
-\label{fig:quinquis}
-\end{figure}
-
-\subsection{Mantle melting curves}\label{sec:katz}
-This instantaneous experiment is performed in order to recreate isobaric melting curves, as function of temperature, bulk water and content of water in the
-melt, to compare with melting curves obtained by \citet{Katz2003}. The domain is square with $L_x=L_y=100$ km, a grid resolution of \num{100x100} elements and
-25 markers per element. Density is set in the entire domain to \SI{3300}{\kg\per\cubic\m} and temperature increases from the sides to the centre from 1000° to
-1500°C. The experiment is performed for bulk water contents from 0 to 0.3 wt.\%.
-
-Fig. \ref{fig:fraction} shows isobaric melting curves for \SI{1}{\giga\pascal} (panel a) and \SI{3}{\giga\pascal} (panel b) calculated for different bulk water
-contents. The curves well match with those obtained for the mantle by \citet{Katz2003}, showing the saturation at the solidus for bulk water of 0.2 and 0.3 wt.\%, indicating
-that the code correctly determines melt fraction as function of pressure, temperature and bulk water content. Same results can be observed in Fig.
-\ref{fig:water}a, where isobaric and isothermal melting curves are shown as function of bulk water content. In addition, Fig. \ref{fig:water}b shows that the
-code correctly determines also the percentage of water in the melt as function of pressure, temperature and melt fraction. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Melting}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Fraction.pdf}
-\caption{Isobaric melting curves for 1 GPa (panel a) and 3 GPa (panel b) as function of temperature with different bulk water contents.}
-\label{fig:fraction}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Water.pdf}
-\caption{Isobaric melting curves for 1.5 GPa as function of bulk water (panel a) and water in the melt (panel b) calculated for different temperatures.}
-\label{fig:water}
-\end{figure}
-
-\clearpage
-
-\bibliography{export}
-
-\end{document}

### 2022-09-07 tex/Benchmark/ist_sphere.tex
update main

--- 2022-03-31 FALCON_description.tex
@@
-\documentclass[hidelinks,10pt,a4paper]{article}
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{textcomp}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage{booktabs}
-\usepackage{multirow}
-\usepackage[cm]{fullpage}
-\usepackage[round]{natbib}
-\bibliographystyle{abbrvnat}
-\usepackage{doi}
-\usepackage{bm}
-\usepackage[table]{xcolor}
-\hypersetup{
-    colorlinks,
-    linkcolor={blue},
-    citecolor={blue},
-    urlcolor={blue}
-}
-\usepackage{float}
-\restylefloat{table}
-\usepackage{siunitx}
-\DeclareSIUnit\year{yr}
-
-\begin{document}
-
-\section{Introduction}\label{sec:intro}
-This document includes a description of the code FALCON, with a presentation of the features implemented followed by results of extensive benchmarks listed in
-Table \ref{tab:list}, performed to verify that the code solves correctly a wide range of problems. Description and results of all benchmarks are shown
-in Section \ref{sec:benchmark}.
-
-\begin{table}[H]
-\caption{List of benchmarks performed for each feature of the code. Data of all benchmarks can be found at 
-\url{https://github.com/aleregorda/Benchmarks}}
-\centering
-\begin{tabular}{| l | l | l |}
-\toprule
- \textbf{Feature}  & \textbf{Benchmark} & \textbf{Section} \\
-\midrule
-  Solver  & Stokes flow & \ref{sec:stokes} \\
-\hline
-  \multirow{2}{*}{Markers advection} & Zalesak disk & \ref{sec:runge} \\
-    & Conservative Interpolation Velocity & \ref{sec:CVI} \\
-\hline
-  \multirow{4}{*}{Momentum equation}  & Poiseuille flow & \ref{sec:poiseuille}\\
-    & Instantaneous 2D sphere & \ref{sec:ist_sphere} \\
-    & Rayleigh-Taylor experiment & \ref{sec:rayleigh} \\
-    & Falling block & \ref{sec:block} \\
-\hline
-    & 2D Stokes sphere & \ref{sec:sphere} \\
-    Sticky air and  & Stabilisation algorithm & \ref{sec:stab} \\
-    free surface & Topography relaxation & \ref{sec:crameri} \\
-  & Spontaneous subduction & \ref{sec:subduction} \\
-\hline
-  Erosion and sedimentation  & - & - \\
-\hline
-   Non-linear  & Slab detachment & \ref{sec:slab} \\
-   visco-plasticity & Indenter & \ref{sec:indenter} \\
-    & Brick & \ref{sec:brick} \\
-\hline
-  \multirow{3}{*}{Energy equation}  & Advection stabilisation & \ref{sec:advection} \\
-    & Simple shear heating & \ref{sec:simple_shear} \\
-    & Shear and adiabatic heating & \ref{sec:shear} \\
-\hline
-\multirow{3}{*}{Energy + momentum} & Mantle convection & \ref{sec:mantle} \\
-& Viscoplastic mantle convection & \ref{sec:visco_mantle} \\
-& Thin layer entrainment & \ref{sec:thin} \\
-\hline
-  Phase changes  & \multirow{2}{*}{Hydrated sinking cylinder} & \multirow{2}{*}{\ref{sec:quinquis}} \\
-  and hydration  &  & \\
-\hline
-  Melting & Experimental melting curves & \ref{sec:katz} \\
-\bottomrule
-\end{tabular}
-\label{tab:list}
-\end{table}
-
-The code is written in Fortran90 and uses the finite element method (FEM) with quadrilateral $Q_1 \times P_0$ elements (continuous bilinear velocity and
-discontinuous constant pressure), associated with MUMPS\footnote{\url{http://mumps.enseeiht.fr/}} \citep{Amestoy2001,Amestoy2019} that is a software package
-for solving systems of linear equations of the form $A \cdot x = b$, where \textit{A} is a square sparse matrix, by means of a direct method. Since
-$Q_1 \times P_0$ elements do not satisfy Ladyzhenskaya, Babuska and Brezzi (LBB) stability condition \citep{Donea2003}, elemental pressure is elaborated
-in post-processing to avoid spurious pressures by means of a double interpolation to smooth the pressure. Following this procedure, the elemental pressure
-is interpolated onto nodes and then back onto elements. After the smoothing procedure, the pressure field is calculated again on the nodes to be used in
-combination with the lithostatic pressure, which is calculated onto nodes. Correctness of the solution and performances of the solver in terms of time and
-memory usage are tested solving the Stokes flow with the analytical solution proposed by \citet{Donea2003} (Section \ref{sec:stokes}).
-
-The thermo-mechanics of crust-mantle systems is described by means of the conservation of mass, momentum and energy equations, expressed as follows:
-\begin{eqnarray}
-\label{eq:mass}\frac{\partial \rho}{\partial t}+\bm{\nabla} \cdot (\rho \bm{u})&=&0\\
-\label{eq:momentum}-\bm{\nabla} p + \bm{\nabla} \cdot \bm{\tau} + \rho \bm{g}&=&0\\
-\label{eq:energy}\rho C_p \left(\frac{\partial T}{\partial t} + \bm{u} \cdot \bm{\nabla} T \right)&=&\bm{\nabla} \cdot \left(k\bm{\nabla} T\right) + H_{tot}
-\end{eqnarray}
-where $\rho$ is the density, $\bm{u}$ is the velocity, \textit{p} is the pressure, $\bm{\tau}$ is the deviatoric stress, $\bm{g}$ is the gravity acceleration,
-$C_p$ is the specific heat at constant pressure, \textit{T} is the temperature, \textit{k} is the thermal conductivity and $H_{tot}$ is the total internal heat
-production. Density variations due to temperature are generally small enough to assume the density as constant ($\rho=\rho_0$) in Eq. \ref{eq:mass} and in
-Eq. \ref{eq:energy}, while it must be treated as a variable in the buoyancy term of Eq. \ref{eq:momentum}, such that \[\rho=\rho_0(1-\alpha(T-T_0))\]
-where $\rho_0$ is the density at a reference temperature $T_0$ and $\alpha$ is the coefficient of thermal expansion. Eq. \ref{eq:mass} then can be rewritten as
-\begin{equation}\label{eq:mass_0}
-\bm{\nabla} \cdot \bm{u}=0
-\end{equation}
-This simplification is known as the extended Boussinesq approximation.
-
-The time step $dt$ is chosen in according to Courant-Friedrichs-Lewy condition \citep{Anderson1995}, such that
-\[dt=C_n\; \min\left(\frac{\Delta x}{\max|\bm{u}|},\frac{(\Delta x)^2}{\kappa}\right)\]
-where $0<C_n<1$ is the Courant (CFL) number, $\Delta x$ is the minimum dimension of smallest element, $\max|\bm{u}|$ is the maximum velocity calculated on the
-entire domain and
-$\kappa=\frac{k}{\rho C_p}$ is the thermal diffusivity \citep{Thieulot2014}.
-
-\section{Lagrangian markers}\label{sec:markers}
-Elemental properties (density, viscosity, thermal conductivity, specific heat and thermal expansion) needed to solve Eqs. \ref{eq:mass_0}, \ref{eq:momentum}
-and \ref{eq:energy} are related to the composition of each element, determined by means of Lagrangian markers that are characteristic of different materials
-in the domain. Elemental properties (with the sole exception of the viscosity) are calculated using an arithmetic mean, as \[P_e=\frac{1}{n}\sum_{i=1}^n P_i\]
-where $P_e$ is the elemental property, $P_i$ is the property characteristic of the material of each marker and $n$ is the number of markers of the element.
-Differently, the average scheme for the elemental viscosity can be chosen between harmonic, geometric and arithmetic mean. Markers can be placed regularly or
-randomly at the beginning of the simulation and their advection is performed by either a 2nd-order or a 4th-order Runge-Kutta in space, interpolating the velocity 
-field on each marker by means of the shape functions (see Section \ref{sec:runge}). The interpolated velocity is then corrected by means of the Conservative
-Velocity Interpolation (CVI), which introduces a corrective term to reduce dispersion and clustering of particles in both steady state and time-dependent 
-\citep{Wang2015} (see Section \ref{sec:CVI}).
-
-The initial distribution of the markers is created using the open-source code library
-Geodynamic World Builder \citep{Fraters2019}. During the simulation, each marker carries memory of temperature, pressure and accumulated strain, which are
-determined interpolating the relative nodal parameter. In particular, pressure is interpolated onto each marker after the smoothing procedure. The number
-of markers contained in each element is maintained between $n_{min}$ and $n_{max}$. When in an element there are less markers than $n_{min}$ the code adds
-random markers to reach the $n_{min}$, while if the number is higher than $n_{max}$ some of them are randomly deleted. When new markers are added, they assume
-the properties of the nearest marker. In this way elements are never empty and maintain a number of markers inside a prefixed range.
-
-\section{The momentum equation}\label{sec:mom_eq}
-The penalty method is implemented so that Eq. \ref{eq:mass_0} can be written as:
-\begin{equation}\label{eq:mass_penalty}
-\bm{\nabla} \cdot \bm{u} + \frac{p}{\lambda}=0
-\end{equation}
-where $\lambda$ is the so-called penalty parameter, which should be 6-7 orders of magnitude larger than the shear viscosity to ensure that mass conservation
-is satisfied \citep{Donea2003,Thieulot2014}.
-
-The deviatoric stress tensor in Eq. \ref{eq:momentum} can be written in terms of the strain rate tensor as $\bm{\tau}=2\eta\dot{\bm{\epsilon}}$, with
-$\dot{\bm{\epsilon}}=\frac{1}{2}\left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)$. Therefore, Eq. \ref{eq:momentum} can be rewritten as:
-\begin{equation}\label{eq:momentum_pressure}
--\bm{\nabla} p + \bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right) + \rho \bm{g} = 0
-\end{equation}
-Finally, using pressure from Eq. \ref{eq:mass_penalty}, Eq. \ref{eq:momentum_pressure} can be rewritten as:
-\begin{equation}\label{eq:momentum_penalty}
-\lambda \bm{\nabla} \left(\bm{\nabla} \cdot \bm{v}\right)+\bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right)+\rho\bm{g}=0
-\end{equation}
-The penalty method is been associated to the iterative Uzawa method, as described in detail in \citet{Dabrowski2008} and \citet{Thieulot2014}
-(see Section \ref{sec:poiseuille}). Results of benchmarks performed to verify the correctness in the implementation of Eq. \ref{eq:momentum_penalty} are
-shown in Sections \ref{sec:ist_sphere} and \ref{sec:rayleigh}.
-
-To support large viscosities variations the penalty parameter is related to the effective elemental viscosity by means of a dimensionless coefficient, so that
-$\lambda=\lambda_e(e)\eta_{eff}(e)$ \citep{Marotta2006,Dabrowski2008,Thieulot2014}. Benchmark of the falling block \citep{Gerya2003a,Gerya2010b,Thieulot2011}
-is performed to verify that the code can correctly deal buoyancy driven flows with strong viscosity contrasts (Section \ref{sec:block}).
-
-\section{Sticky air and free surface}\label{sec:sticky}
-The Earth's surface can be treated by means of either the so-called sticky-air or a true free surface method, both of them implemented in the code. In the
-sticky-air method the surface is approximated with the introduction of a buoyant layer with a viscosity at least four orders of magnitude lower than the crust
-\citep{Schmeling2008,Crameri2012} and the interface between lithosphere and air is defined using a chain of passive markers that are advected as the Lagrangian
-markers. The correctness of the evolution of the markers chain is tested with the experiment of a 2D time-dependent Stokes sphere below a free surface and
-compared with results from ASPECT\footnote{\url{https://aspect.geodynamics.org/}} \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} (Section
-\ref{sec:sphere}). In the true free surface case the top boundary is assumed stress-free and velocities are not fixed. In this case topography variations are
-described by vertical deformations of the mesh that depend on the velocity field of the nodes that identify the free surface, while horizontal deformations are
-not taken into account. This procedure is known as the Arbitrary Lagrangian-Eulerian (ALE) method and its implementation follows the technique described in
-\citet{Thieulot2011}. However, although the implementation of a true free surface better reproduces laboratory experiment, extremely small time steps can be
-necessary to maintain stability \citep{Kaus2010a,Quinquis2011,Thieulot2014}. Therefore, the stability algorithm proposed by \citet{Kaus2010a} is implemented
-to avoid instabilities due to high density differences at the free surface when using too large time steps. The implementation of this algorithm are tested by
-performing the experiment described by \citet{Kaus2010a} (Section \ref{sec:stab}).
-
-The topography relaxation benchmark proposed by \citet{Crameri2012} is performed to verify that the code correctly recovers topography variations in case of
-both the sticky-air and the true free surface method (Section \ref{sec:crameri}). Finally, Section \ref{sec:subduction} show the results of the spontaneous
-subduction experiment described by \citet{Schmeling2008}.
-
-\section{Erosion and sedimentation}\label{sec:erosion}
-Surface processes at either the lithosphere-air interface or in correspondence of the free surface have been implemented by means of the software
-Fastscape\footnote{\url{https://fastscape.org/}} \citep{Braun2013,Cordonnier2019,Yuan2019a,Yuan2019b}. FastScapeLib is a set of routines that solve the stream
-power law (enriched by a sediment transport/deposition term) \citep{Yuan2019a}, hillslope diffusion and marine transport and deposition \citep{Yuan2019b}.
-The partial differential equation solved by FastScapeLib is:
-\begin{equation}
-\frac{\partial h}{ \partial t}=U-K_f A^{m_e}S^{n_e} + \frac{G}{A}\int_A \left( U-\frac{\partial h}{\partial t} \right) dA + K_d \bm{\nabla}^2 h 
-\end{equation}
-where $h$ is the topography, $U$ is the uplift, $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $S$ is the slope, $m_e$ and $n_e$ are
-parameters that should be chosen to constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}, $G$ is the dimensionless
-deposition/transport coefficient for the enriched stream power law and $K_d$ is the effective diffusivity. Terms $K_f A^{m_e}S^{n_e}$ and $K_d \bm{\nabla}^2 h$
-consider long-range processes, described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, and short-range processes,
-proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}.
-
-In case of sticky air, slope is determined by means of the markers chain, which is vertically corrected considering erosion and sedimentation rates. After the
-correction, continental markers above the markers chain are transformed in air markers and air markers below the markers chain become sediments. Similarly,
-slope is determined considering the nodes on the top boundary in case of true free surface and they are vertically corrected as for the markers chain. In this
-case, continental markers above the top boundary are simply deleted.
-%Surface processes at the lithosphere-air interface are implemented taking into account fluvial erosion (long-range processes), hillslope processes (short-range
-%processes) and failure of slopes exceeding a threshold angle $\theta$ \citep{Burov1997,Tucker1999,Sacek2012,Croissant2014,Ueda2015,Beucher2020}. Erosion rates
-%$\dot{E}_{lr}$ due to long-range processes can be described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{lr}=K_f A^{m_e} \bm{\nabla} h^{n_e}
-%\end{equation}
-%where $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $\bm{\nabla} h$ is the slope and $m_e$ and $n_e$ are parameters chosen in to
-%constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}. Erosion rates $\dot{E}_{sr}$ due to short-range processes are
-%implemented by assuming that transport is linearly proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{sr}=K_d \bm{\nabla}^2 h
-%\end{equation}
-%where $K_d$ is the effective diffusivity. In addition, the slope is modified if it still exceeds a threshold angle after long- and short-range processes
-%\citep{Ueda2015}.
-
-%Offshore sedimentation processes are implemented similarly to hillslope processes but with a depth-dependent coefficient \citep{Sacek2012,Beucher2020}.
-%The sedimentation rate $\dot{S}$ is calculated as
-%\begin{equation}
-%\dot{S}=K_s \bm{\nabla}^2 h
-%\end{equation}
-%with
-%\begin{equation}
-%K_s=K^*_s \exp\left(\frac{h-h_{sl}}{h_0}\right)
-%\end{equation}
-%where $K^*_s$ is a constant, $h_{sl}$ is the sea level and $h_0$ is a characteristic length. In case of sticky air, slope is determined by means of the markers
-%chain, which is vertically corrected considering erosion and sedimentation rates. After the correction, continental markers above the markers chain are
-%transformed in air markers and air markers below the markers chain become sediments. Similarly, slope is determined considering the nodes on the top boundary
-%in case of true free surface and they are vertically corrected as for the markers chain. In this case, continental markers above the top boundary are simply
-%deleted.
-
-\section{Non-linear rheologies}\label{sec:nonlinear}
-Non-linear rheologies are implemented combining viscous creep (dislocation and diffusion) and plastic yielding. For each marker, diffusion and dislocation
-viscosity ($\eta_{df|ds}$) can be determined as follows \citep{Karato1993,Warren2008b,Wang2016}:
-\begin{equation}\label{eq:viscous}
-\eta_{df|ds}=f_s \left(\frac{d^m}{A}\right)^{\frac{1}{n}} \left (\frac{2^{\frac{1-n}{n}}}{3^{\frac{n+1}{2n}}} \right)
-I_2^{\frac{(1-n)}{n}}\exp\left(\frac{Q+pV}{nRT}\right)
-\end{equation}
-where $f_s$ is a scaling factor used to represent lithologies that are stronger or weaker than the base set, $d$ is the grain size, $m$ is the grain size exponent,
-$n$ is the stress exponent, $A$ is the uniaxial pre-exponential factor, $I_2$ is the square root of the second invariant of the strain rate tensor, $Q$ is the
-activation energy, $p$ is the pressure, $V$ is the activation volume, $R$ is the gas constant and $T$ is the temperature.
-Pressures and temperatures are determined by the interpolation of the nodal parameters, while strain rates are calculated by means of the
-derivative of the velocities on the nodes. In case of diffusion creep $n=1$ and $m>0$, in case of dislocation creep $n>1$ and $m=0$. The viscous creep $\eta_{cp}$
-is then calculated as the harmonic average between $\eta_{df}$ and $\eta_{ds}$:
-\begin{equation}\label{eq:visco_creep}
-\eta_{cp}=\left(\frac{1}{\eta_{df}}+\frac{1}{\eta_{ds}}\right)^{-1}
-\end{equation}
-The implementation of the non-linear viscous creep viscosity depending on the strain rate is tested by means of the slab detachment benchmark
-(Section \ref{sec:slab}).
-
-Plastic yielding is implemented rescaling $\eta_{cp}$ in order to limit the stress below the yield stress $\sigma_y$ \citep{Thieulot2008,Thieulot2014,Glerum2018},
-obtaining
-\begin{equation}\label{eq:plastic}
-\eta_{pl}=\frac{\sigma_y}{2I_2}
-\end{equation}
-where $\eta_{pl}$ is the plastic viscosity and the yield stress is determined following the Drucker-Prager criterion, such as
-\begin{equation}\label{eq:yield}
-\sigma_y=C\;\cos(\phi)+p\;\sin(\phi)
-\end{equation}
-where $C$ is the cohesion and $\phi$ is the internal friction angle. In case of negative pressure it is imposed equal to 0, so that negative yield stress are
-excluded. The correctness of the non-linear solution in case of plastic viscosity with variable internal friction angle is verified performing indenter and
-brick experiments (Sections \ref{sec:indenter} and \ref{sec:brick}, respectively).
-
-Strain softening is taken into account for both viscous creep and plastic viscosity \citep{Huismans2003,Babeyko2005,Huismans2005,Sobolev2005,Warren2008a} by
-means of the accumulated strain $\epsilon$ memorised by each marker. Softening in the viscous creep determines a linear decrease of $\eta_{vc}$ by means of a
-viscous strain softening factor $W_S$ that increases linearly from $W_{S_0}$ to $W_{S_{\infty}}$ for $\epsilon_{S_0}<\epsilon_S<\epsilon_{S_{\infty}}$ 
-\citep{Warren2008a}. This viscous softening can be related to strain-induced grain size reduction \citep{Warren2008a}. Differently, plastic softening is 
-simulated with a linear decrease of internal friction angle $\phi(\epsilon)$ and cohesion $C(\epsilon)$ in according to
-\begin{eqnarray}
-\label{eq:friction1}\phi(\epsilon)&=&\phi_0+(\phi_{\infty}-\phi_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}\\
-\label{eq:friction2}C(\epsilon)&=&C_0+(C_{\infty}-C_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}
-\end{eqnarray}
-where $\phi_0$, $C_0$ and $\phi_{\infty}$, $C_{\infty}$ are internal friction angle and cohesion for $\epsilon_0$ and $\epsilon_{\infty}$, respectively
-\citep{Huismans2003,Huismans2005,Warren2008a,Thieulot2014}. Plastic softening approximates deformation-induced softening of faults and brittle shear zones
-\citep{Warren2008a}.
-
-Plastic yielding and viscous creep are then combined to obtain a viscoplastic viscosity $\eta_{vp}$ as follows, assuming that they are independent processes
-\citep{Karato2008,Glerum2018}:
-\begin{equation}\label{eq:viscoplastic}
-\eta_{vp}=\min(\eta_{cp},\eta_{pl})
-\end{equation}
-Finally, effective viscosity $\eta_{eff}$ is capped by the minimum and the maximum viscosity ($\eta_{min}$ and $\eta_{max}$, respectively) to avoid extremely
-low or high viscosity \citep{Glerum2018} as follows
-\begin{equation}\label{eq:effective}
-\eta_{eff}=\min(\max(\eta_{vp},\eta_{min}),\eta_{max})
-\end{equation}
-Viscous creep and plastic yielding are non-linear rheologies because of their dependence on the velocity field through pressure and strain rates. Therefore,
-the solution is determined by means of Picard-type iterations, until convergence of the velocity field \citep{Glerum2018}. The convergence is verified at each
-iteration $i$ via the nonlinear residual $\bm{\mathcal{R}}^i$ that can be determined as
-\begin{equation}\label{eq:convergence}
-\bm{\mathcal{R}}^i= \mathbb{K} (\eta_{eff}(\dot{\bm{\epsilon}}^{i-1},p^{i-1}))\cdot \bm{v}^{i-1} - \bm{f}^i
-\end{equation}
-%\begin{equation}\label{eq:convergence}
-%\chi=1-\frac{\langle (\bm{v}^i-\langle\bm{v}^i\rangle)\cdot (\bm{v}^{i+1}-\langle\bm{v}^{i+1}\rangle) \rangle}{\sqrt{\sigma^i\sigma^{i+1}}}
-%\end{equation}
-%Iterations are performed until either convergence ($\chi<tol$) or maximal number of non-linear iterations ($it_{max}$) is reached.
-where $\mathbb{K}$ is the Stiffness matrix and $\bm{f}$ is the right hand side vector \citep{Spiegelman2016,Glerum2018}.  The $L_2$-norm is extracted from
-$\bm{\mathcal{R}}^i$ and it is normalised as follows
-\begin{equation}\label{eq:convergence_res}
-\frac{||\bm{\mathcal{R}}^i||_2}{||\bm{\mathcal{R}}^0||_2}
-\end{equation}
-where $||\bm{\mathcal{R}}^0||_2$ is the $L_2$-norm of the first nonlinear iteration.
-Since often the normalised nonlinear residual drop very quickly over the first 2-3 iterations, also the $L_2$-norm of velocity and pressure residuals are
-calculated as
-\begin{equation}\label{eq:convergence_vel}
-\frac{||\Delta\bm{v}||_2}{||\bm{v}||_2}=\frac{||\bm{v}_i-\bm{v}_{i-1}||_2}{||\bm{v}_{i}||_2}
-\end{equation}
-and
-\begin{equation}\label{eq:convergence_press}
-  \frac{||\Delta p||_2}{||p||_2}=\frac{||p_i-p_{i-1}||_2}{||p_{i}||_2}
-\end{equation}
-respectively. Iterations are performed until either maximal number of non-linear iterations ($it_{max}$) is reached or all the normalised nonlinear residual, 
-the velocity residual and the pressure residual converge under a defined tolerance $tol$.
-
-\section{The energy equation}\label{sec:en_eq}
-The stabilisation of the advection term of the energy equation (Eq. \ref{eq:energy}) needed to avoid possible oscillations in the thermal solution in the case
-when advection dominates over diffusion, is implemented by means of a streamline-upwind Petrov–Galerkin (SUPG) method, for which the advection term is modified
-follows the discussion in \citet{Thieulot2011} and \citet{Thieulot2014} (see Section \ref{sec:advection} for benchmark). The total internal heat production
-$H_{tot}$ in Eq. \ref{eq:energy} includes radiogenic heating $H_r$, shear heating $H_s$ and adiabatic heating $H_a$, where
-\begin{equation}\label{eq:shear}
-H_s=2\eta \dot{\bm{\epsilon}} : \dot{\bm{\epsilon}} = 2\eta(\dot{\epsilon}_{xx}^2+\dot{\epsilon}_{yy}^2+2\dot{\epsilon}_{xy}^2)
-\end{equation}
-and
-\begin{equation}\label{eq:adiabatic}
-H_a=T \alpha \frac{Dp}{Dt} \approx -\alpha T \rho g_y v
-\end{equation}
-The correctness of the implementation and the computation of shear and adiabatic heating is verified performing an exercise with an analytical solution
-(Section \ref{sec:simple_shear}) and Exercise 9.4 in \citet{Gerya2010b} (Section \ref{sec:shear}). Furthermore, mantle convection and thin layer entrainment
-benchmarks (Sections \ref{sec:mantle} and \ref{sec:thin}, respectively) are performed to verify that the code solve correctly Eqs. \ref{eq:momentum_penalty}
-and \ref{eq:energy} in case of a temperature field characterised by an initial perturbation and with a temperature-dependent density. At last, the viscoplastic
-mantle convection benchmark is performed to check also the viscoplasticity in case of a initial perturbed temperature field (Section \ref{sec:visco_mantle}).
-
-\section{Phase transitions and hydration}\label{sec:phase}
-Variations of effective density, specific heat and coefficient of thermal expansion during the evolution of crustal and mantle materials at different
-pressure-temperature ($p$-$T$) conditions are computed by means of version 6.8.6 of Perple\_X software package \citep{Connolly2005}, similarly to the
-implementation described in \citet{Marotta2020}. By default, mineral assemblages and properties of each lithology (oceanic crust, lower and upper continental
-crust, sediments and dry or hydrated mantle) are calculated for temperatures from \SIrange{330}{1600}{\kelvin}, with increments of approximately \SI{4}{\kelvin},
-and pressures from \SIrange{0.1}{30}{\giga\pascal}, with increments of approximately \SI{0.03}{\giga\pascal}, for a total of almost 100000 points.
-
-The introduction of phase transitions produces two main effects that must be taken into account in numerical models: variations in density and release/absorption
-of latent heat \citep{Tackley2010}. Effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into
-account considering effective coefficient of thermal expansion \citep{Christensen1985,Zhong2015}. Similarly, effects of latent heat in the energy equation
-(eq \ref{eq:energy}) are taken into account by considering effective specific heat and coefficient of thermal expansion
-\citep{Christensen1985,Tackley2010,Zhong2015}. Effective density, specific heat and coefficient of thermal expansion are included in Perple\_X files and are
-assumed by each Lagrangian marker in according to its $p$-$T$ conditions.
-
-Hydration processes related to dehydration of subducting lithosphere \citep{Schmidt1998,Liu2007,Faccenda2009,Faccenda2010,Faccenda2014,Rosas2016} strongly
-influence the thermo-mechanics inside the mantle wedge, mainly because of both weakening effects on the mantle rheology and density variations in case of
-mantle serpentinization \citep{Gerya2002,Honda2003,Arcay2005,Roda2010,Regorda2017}. In the case that hydration is switched on, the amount of bound and free
-water is memorised by each marker, following the implementation of \citet{Quinquis2014}; hydration and dehydration processes are related to the amount of bound
-water of each marker and to the maximum amount of water it can transport, i.e. if the amount of bound water exceeds the maximum amount of water, the marker
-dehydrates and releases free water that can hydrates under-saturated markers. Maximum water content of each marker is determined as function of lithology and
-$p$-$T$ conditions and it is calculated using Perple\_X, in the same way than effective density, specific heat and coefficient of thermal expansion. Bound water
-is advected together with the markers, neglecting the effect of bound water diffusion, while free water simply migrates vertically and is not coupled to the
-solid-phase flow of the mantle wedge \citep{Arcay2005,Quinquis2014}. The correctness of the migration scheme of the free water and its absorption into bound
-water by under-saturated markers are verified performing the experiment described in \citet{Quinquis2014} (Sec. \ref{sec:quinquis}).
-
-The mantle viscosity weakening related to the amount of water has been extensively studied \citep{Chopra1981,Hirth2003} and it is implemented in according to
-\citet{Arcay2005} and \citet{Horiuchi2016}, as follows
-\begin{equation}\label{eq:wet_visc}
-\eta_{wet}=\eta_{dry}\left(\left[1-\frac{1}{f_v}\right]\exp\left(-\frac{[OH^-]}{[OH^-]_0}\right)+\frac{1}{f_v}\right)
-\end{equation}
-where $f_v$ is the viscosity reduction factor between dry and wet conditions, $[OH^-]$ is the water content and $[OH^-]_0$ is a reference water content set to
-620 ppm (0.062 wt.\%) \citep{Arcay2005}. By default, $f_v$ is set to 100, in according to \citet{Arcay2005} and \citet{Horiuchi2016}.
-
-\section{Mantle melting}\label{sec:melt}
-Melt fraction in the mantle depends on temperature ($T$, in \textdegree C), pressure ($p$, in \si{\giga\pascal}) and the water content in the melt ($X_{H_2O}$, in wt.\%).
-The determination of the melt fraction $M$ is implemented as explained by \citet{Katz2003}, according to the successive modification by \citet{Langmuir2006} and
-\citet{Kelley2010}, as follows
-\begin{eqnarray}\label{eq:melt_fr}
-M(p,T,X_{H_2O})&=&-T+T_{sol}+(x_t\cdot ln(p+y_t))\cdot X_{H_2O}-\nonumber\\
-&&+K_m\left(\frac{X_{H_2O}^{bulk}}{D_{H_2O}(1-X_{H_2O})+X_{H_2O}}\right)^{\gamma}
-\end{eqnarray}
-where $T_{sol}$ is the temperature of the dry solidus, $D_{H_2O}$ is the partition coefficient and $x_t\cdot log(p+y_t)$ indicates the pressure dependence of
-the melting curve \citep{Kelley2010}. $x_t$, $y_t$, $K_m$ and $\gamma$ are parameters chosen in according to \citet{Kelley2010}.
-However, $X_{H_2O}$ depends on the melt fraction as
-\begin{equation}\label{eq:melt_wt}
-X_{H_2O}(M)=\frac{X_{H_2O}^{bulk}}{D_{H_2O}+M(1-D_{H_2O})}
-\end{equation}
-and a numerical solution can be found using a root-finder method \citep{Katz2003,Wang2016}. Furthermore, the content of water in the melt is limited by the
-pressure-dependent saturation concentration of water in the melt, determined as
-\begin{equation}\label{eq:melt_max}
-X_{H_2O}^{sat}=\chi_1 p^{\lambda_m}+\chi_2 p
-\end{equation}
-where $\chi_1$, $\chi_2$ and $\lambda_m$ are parameters chosen according to \citet{Katz2003}, \citet{Langmuir2006} and \citet{Kelley2010}. The correctness in
-the determination of melt fractions and water contents in the melt are tested comparing model results with experimental curves obtained by \citet{Katz2003}
-(Sec. \ref{sec:katz}).
-
-In case of melting, viscous creep viscosity ($\eta_{vp}$) is modified according to \citet{Wang2016}, as
-\begin{equation}\label{eq:melt_visc}
-\eta_{melt}=\eta_{vp}\cdot \exp(\alpha_m M)
-\end{equation}
-where $\alpha_m$ are melt fraction factors for dislocation and diffusion creep, chosen in according to \citet{Wang2016}.
-
-Effective density ($\rho_{eff}$) for partially molten rocks is calculated as
-\begin{equation}\label{eq:melt_dens}
-\rho_{eff}=\rho_s(1-M)+\rho_m M
-\end{equation}
-where $\rho_s$ and $\rho_m$ are the densities of the solid and the molten rock, respectively \citep{Gerya2004b,Gerya2007,Wang2016}. The density of the
-solid rock is extracted by Perple\_X, while the density of the molten portion is calculated as
-\begin{equation}\label{eq:melt_dens_M}
-\rho_m(p,T)=\rho_0[1-\alpha(T-T_0)][1+\beta(p-p_0)]
-\end{equation}
-where $\rho_0$ is the density at the reference temperature ($T_0$) and pressure($p_0$), and $\alpha$ and $\beta$ are the thermal expansion and compressibility
-coefficients, respectively \citep{Gerya2004b,Gerya2007,Wang2016}.
-
-As for the phase transitions, effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into account
-considering effective coefficient of thermal expansion. Similarly, latent heat due to melting/crystallisation can be implicitly included in the energy equation
-(Eq. \ref{eq:energy}) by considering effective specific heat ($C_{p_{eff}}$) and thermal expansion ($\alpha_{eff}$) of the partially molten rocks, in according to
-\begin{eqnarray}
-\label{eq:cp_effective}C_{p_{eff}}&=&C_p+H_L\left(\frac{\partial M}{\partial T}\right)_p \\
-\label{eq:alpha_effective}\alpha_{eff}&=&\alpha+\rho\frac{H_L}{T}\left(\frac{\partial M}{\partial p}\right)_T
-\end{eqnarray}
-where $C_p$ and $\alpha$ are the specific heat and the coefficient of thermal expansion of the solid rock, respectively, and $H_L$ is the latent heat of the
-molten rock \citep{Gerya2004b,Gerya2007,Tackley2010}.
-
-\section{Benchmarks}\label{sec:benchmark}
-\subsection{Error measurements}\label{sec:error}
-In order to determine the accuracy of velocity and pressure field of the benchmarks, the $L_2$-norm is computed by numerical integration on the Gauss-Legendre
-quadrature points. $L_2$-norm for pressure and velocity errors can be evaluated as
-\begin{eqnarray}
-\textrm{err}_p&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|p^n_i(r_q)-p^a_i(r_q)|^2\right)w_q|J_q|}\nonumber \\
-\textrm{err}_v&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|u^n_i(r_q)-u^a_i(r_q)|^2+|v^n_i(r_q)-v^a_i(r_q)|^2\right)w_q|J_q|}\nonumber
-\end{eqnarray}
-respectively, where $n_e$ is the number of elements, $n_q$ is the number of quadrature points per element, $p^n_i(r_q)$ and $p^a_i(r_q)$ are the numerical and
-analytical pressure, respectively, in each quadrature point $q$, $w_q$  and $J_q$ are the weight and the Jacobian at the quadrature point
-$q$, $u^n_i(r_q)$, $v^n_i(r_q)$, $u^a_i(r_q)$ and $v^a_i(r_q)$ are the numerical and analytical velocities, respectively, in each quadrature point $q$.
-
-Other quantities used as comparison with original benchmarks are the root-mean-square velocity over the whole domain and over the surface
-\begin{eqnarray}
-v_{\textrm{rms}}&=&\sqrt{\int_{\Omega}|\bm{v}|^2 d\Omega}\nonumber \\
-v_{\textrm{rms}}^{top}&=&\sqrt{\int_0^1 u^2 \bigg|_{y=1} dx}\nonumber
-\end{eqnarray}
-respectively, top and bottom Nusselt numbers
-\[\textrm{Nu}^{top/bottom}=-\int_0^1\frac{\partial T}{\partial y} \bigg|_{y=1/y=0} dx\]
-the average rate of work done against gravity
-\[\left\langle W \right\rangle=\int_{\Omega}T u_y d\Omega\]
-the average rate of viscous dissipation
-\[\left\langle \Phi \right\rangle=\int_{\Omega}\tau_{ij}\dot{\epsilon}_{ij} d\Omega\]
-and the percentage error between $\left\langle W \right\rangle$ and $\left\langle \Phi \right\rangle/Ra$
-\[\delta=\frac{\left|\left\langle W \right\rangle -\frac{\left\langle \Phi \right\rangle}{Ra} \right|}{\max \left(\left\langle W \right\rangle,
-\frac{\left\langle \Phi \right\rangle}{Ra}\right)}\times100\]
-
-\subsection{Stokes flow}\label{sec:stokes}
-The problem consists of determining velocity field $(u,v)$ and pressure $p$ in case of a manufactured solution with prescribed body forces such as
-\begin{eqnarray}
-b_1&=&(12-24y)x^4+(-24+48y)x^3+(-48y+72y^2-48y^3+12)x^2+\nonumber \\
-&&+(-2+24y-72y^2+48y^3)x+1-4y+12y^2-8y^3\nonumber \\
-b_2&=&(8-48y+48y^2)x^3+(-12+72y-72y^2)x^2+(4-24y+48y^2-48y^3+\nonumber \\
-&&+24y^4)x-12y^2+24y^3-12y^4\nonumber
-\end{eqnarray}
-for which the exact solution is:
-\begin{eqnarray}
-u(x,y)&=&x^2(1-x)^2(2y-6y^2+4y^3)\nonumber \\
-v(x,y)&=&-y^2(1-y)^2(2x-6x^2+4x^3)\nonumber \\
-p(x,y)&=&x(1-x)-\frac{1}{6}\nonumber
-\end{eqnarray}
-The domain is a unit square a constant viscosity ($\eta=1$) and the penalty parameter is set to $\lambda=10^7$. Velocity boundary conditions are set to no slip
-$(\bm{v}=\bm{0})$ on all boundaries. The problem is performed for different grid resolution between \num{8x8} and \num{1024x1024} elements.
-
-The errors between the analytical solution and the numerical prediction of the velocity and the pressure field are calculated as in Section \ref{sec:error}.
-Fig. \ref{fig:errors}a shows that both velocity and pressure field converge to the exact solution with the decrease of the element size, following the
-theoretical convergence rate. The convergence of pressure, in contrast with the observation by \citet{Donea2003} for $Q_1 \times P_0$ elements, indicate the
-effectiveness of the smoothing procedure. Smoothed pressure field for grid resolution of \num{128x128} elements is shown in Fig. \ref{fig:errors}b (continuous
-red line), in comparison with the analytical solution (dashed blue line). Solve times and memory usage needed to generate the solution are shown in
-Fig. \ref{fig:MUMPS}. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Solver/Stokes_Flow}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/errors.pdf}
-\caption{Velocity and pressure error for the Stokes flow experiment between generated and analytical solution as a function of element size (panel a),
-and comparison between smoothed pressure and analytical solution as function of $x$ coordinate for a grid resolution of \num{128x128} elements (panel b).}
-\label{fig:errors}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/MUMPS.pdf}
-\caption{Panel a) Analysis, factorisation, solution and total solve times (panel a) and factorisation memory usage (panel b) as a function of the total number
-of degrees of freedom for the Stokes flow experiment.}
-\label{fig:MUMPS}
-\end{figure}
-
-\subsection{Markers advection}\label{sec:runge}
-The 2nd-order Runge-Kutta advection scheme in space is tested by means of the Zalesak disk test \citep{Zalesak1979,Thieulot2014}. The benchmark is performed
-in a unit square domain with a grid resolution of \num{32x32} elements and 1000000 markers and values of Courant number between 0.25 and 1.
-At $t=0$, the disk is centred at position $(0.5;0.75)$ with a radius $R=0.15$ and has a vertical fissure 0.05 wide and 0.2 high (orange in Fig. \ref{fig:runge})
-and with the velocity field prescribed in the entire domain as
-\begin{eqnarray}
-u(x,y)&=&2\pi\left(y-\frac{L_x}{2}\right)\nonumber \\
-v(x,y)&=&-2\pi\left(x-\frac{L_x}{2}\right)\nonumber
-\end{eqnarray}
-the markers are back to their initial location after a $2\pi$ rotation ($t=1$). The clockwise rotation of the disk at $t=0.25$, 0.5 and 0.75 
-is shown in Fig. \ref{fig:runge} (red, blue and green, respectively). The distance from the centre of 1 marker is calculated throughout the rotation to
-evaluate the error for different values of Courant number. As expected, the error increase with the increasing of the time step, as shown in 
-Fig. \ref{fig:runge_err}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/Zalesak_Disk}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=200px]{./Figures/Zalesak.pdf}
-\caption{Position of the Zalesak disk at $t=0$, 0.25, 0.5 and 0.75 (orange, red, blue and green, respectively) for the case with Courant number of 1.}
-\label{fig:runge}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/zal_marker.pdf}
-\caption{Distance from the centre as function of time for values of Courant number of  of 0.25, 0.3, 0.5, 0.75 and 1 (green, orange, blue, black and red,
-respectively). R indicates the distance from the centre considering a perfect circle trajectory.}
-\label{fig:runge_err}
-\end{figure}
-
-\subsection{Conservative Velocity Interpolation (CVI)}\label{sec:CVI}
-The CVI correction is checked by means of the Stokes flow experiment by \citet{Donea2003}, as presented in \ref{sec:stokes}. The advection of Lagrangian
-markers is performed using either a 2nd-order or a 4th-order Runge-Kutta scheme with CFL=0.5 and an initial random distribution of 25 markers per element.
-Fig. \ref{fig:CVI} shows the minimum and maximum values of markers per element throughout the experiment with and without the CVI correction and using
-different Runge-Kutta schemes. The introduction of the CVI correction allows the
-tests to run for times longer than $t=2000$ without show empty elements, while tests without the CVI correction stop at approximately $t=600$, when at least
-one element goes empty. Fig. \ref{fig:CVI_time} shows the different number of markers per element and the markers distribution at $t=600$ obtained without and
-with the CVI correction (panels b and c, respectively) in case of a 2nd-order Runge-Kutta scheme. All data can be found at
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/CVI}.
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/CVI.pdf}
-\caption{Maximum and minimum number of markers per element throughout the experiment, with (green and orange lines) or without (purple and blue lines)
-the CVI correction and using either a 2nd-order (blue and orange lines) or a 4th-order (purple and green lines) Runge-Kutta scheme.}
-\label{fig:CVI}
-\end{figure}
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/CVI_time.pdf}
-\caption{Number of markers per element and markers distribution through time using 2nd.order Runge-Kutta scheme.
-Panels a: number of markers per element (panel a1) and distribution of the markers (panel a2) after the first
-iteration; panels b and c: comparison between number of markers per element (panels b1 and c1) and markers distribution (panels b2 and c2) without and with
-the CVI correction at $t=600$; panels d: number of markers per element (panel d1) and markers distribution (panel d2) in case of CVI correction at $t=2000$.}
-\label{fig:CVI_time}
-\end{figure}
-
-\subsection{Poiseuille flow}\label{sec:poiseuille}
-The domain is a rectangle with $L_x=2$ and $L_y=1$ and constant density and viscosity ($\rho=1$ and $\eta=1$), gravity acceleration $\bm{g}=\bm{0}$ and penalty
-parameter $\lambda=10^8$. The grid is composed by \num{40x20} elements. Velocity boundary conditions are set to no slip $(\bm{v}=\bm{0})$ at the top and the
-bottom, and a parabolic profile is imposed on the sides, with $u=y(1-y)$ and $v=0$. The analytical solution is then given by:
-\begin{eqnarray}
-u(x,y)&=&y(1-y)\nonumber \\
-v(x,y,)&=&0\nonumber \\
-p(x,y)&=&2\eta\left(\frac{L_x}{2}-x\right)\nonumber
-\end{eqnarray}
-
-The velocity field predicted by the model follows the expected parabolic profile (Fig. \ref{fig:poiseuille}a and Fig. \ref{fig:poi_plot}a) and, in case of the
-classic penalty method with no iterations, the pressure is clearly related to the divergence by means of the penalty parameter (Fig. \ref{fig:poiseuille}b and
-c and Fig. \ref{fig:poi_plot}b). Fig. \ref{fig:poiseuille}d shows that one Uzawa iteration is sufficient to bring the divergence down to \num{e-15}, with no
-correlation with the pressure field. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Poiseuille%20Flow}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Poiseuille.pdf}
-\caption{Velocity field (panel a), pressure (panel b) and divergence velocity of a Poiseuille flow in case of the classic penalty method (no iterations) and
-after one Uzawa iteration (panel c and d, respectively).}
-\label{fig:poiseuille}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Analytical.pdf}
-\caption{Velocity (panel a) and pressure (panel b) field predicted by the model for a Poiseuille flow with respect with their analytical solutions. The
-velocity field is plotted as function of the vertical coordinate in $L_x/2$ and the pressure is plotted as function of the x coordinate in $L_y/2$.}
-\label{fig:poi_plot}
-\end{figure}
-
 \subsection{Instantaneous 2D sphere}\label{sec:ist_sphere}
@@
 \end{enumerate}
-All three cases are performed for different grid resolutions, between \num{16x16} and \num{512x512} elements with 50 randomly distributed markers per element,
+All three cases are performed for different grid resolutions, between $16 \times16$ and $512\times 512$ elements with 50 randomly distributed markers per element,
 and different average schemes for the viscosity (harmonic, geometric and arithmetic). The velocity in the centre of the sphere $\bm{v}(0.5;0.5)$), minimum and
@@
 \end{figure}
-
-\subsection{Rayleigh-Taylor instability}\label{sec:rayleigh}
-This problem has been originally presented by \citet{vanKeken1997} and here it is performed as the isoviscous case (Case 1a). The domain has $L_x=0.9142$ and
-$L_y=1$ and gravity $\bm{g}=(0,-1)$. Two fluids with same constant viscosities ($\eta_1=\eta_2=1$) and different densities ($\rho_1=1000$ and $\rho_2=1010$),
-with the lighter fluid at the bottom. The initial interface between the fluids is given by $y(x)=0.2+0.02 \cos \left(\frac{\pi x}{L_x}\right)$. The experiment
-is performed with different grid sizes (\num{50x50}, \num{80x80}, \num{100x100} and \num{256x256}). A total of 1960000 markers are randomly distributed at the
-beginning of the simulation. Velocity boundary conditions are set to no slip at the top and the bottom, and to free slip at the sides of the domain.
-
-$v_{\textrm{rms}}$ as function of time is reported in Fig. \ref{fig:RT}, matching well with results shown by \citet{vanKeken1997}, \citet{Tackley2003} and
-\citet{Thieulot2014}. Fig. \ref{fig:rayleigh} shows the evolution of the experiment at different time steps. All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Rayleigh_Taylor_experiment/ISOVISCOUS}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/RT.pdf}
-\caption{$v_{\textrm{rms}}$ of the Rayleigh-Taylor experiment as function of time for different resolution of the grid.}
-\label{fig:RT}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=300px]{./Figures/Rayleigh.pdf}
-\caption{Evolution of the Rayleigh-Taylor experiment for a grid of \num{256x256} elements at $t=0, 500, 1000$ and $2000$ (panels a, b, c and d, respectively).}
-\label{fig:rayleigh}
-\end{figure}
-
-\subsection{Falling blocks}\label{sec:block}
-This benchmark is described as proposed by \citet{Gerya2003a}, \citet{Gerya2010b} and \citet{Thieulot2011}. The domain is square with $L_x=L_y=$ \SI{500}{\km}
-and the grid is composed by \num{50x50} elements with 25 markers in each element. The block is initially centred at ($x=$ \SI{250}{\km}; $y=$ \SI{400}{\km})
-and has a size of \num{100x100} km (Fig. \ref{fig:block}a). The fluid surrounding the block has $\eta_f=$ \SI{e21}{\pascal\per\s} and $\rho_f=$
-\SI{3200}{\kg\per\cubic\m}. The benchmark tests with different viscosities of the block, with $\eta_b$ from \SIrange{e15}{e27}{\pascal\per\s}. In each
-experiment also the density $\rho_b$ of the block varies from \SIrange{3220}{9900}{\kg\per\cubic\m}. Velocity boundary conditions are set to free slip
-conditions on all sides of the domain. Density distributions at $t=$ \SI{20}{\mega\year} for $\rho_b=$ \SI{3300}{\kg\per\cubic\m} and different $\eta_b$
-are plotted in Fig. \ref{fig:block}b-f, showing that the code correctly preserve the block geometry in case of large viscosity contrast, with a block stiffer
-than the surrounding fluid (Fig. \ref{fig:block}f).
-
-The velocity in the centre of the falling block at $t=0$ is measured for all experiments and, since the velocity should increase with density contrast, the
-quantity $\bm{v}/(\rho_b-\rho_f)$ is plotted as function of the viscosity contrast. All results perfectly match with those from \citet{Gerya2010b} and they
-line up on a single curve, demonstrating that the code can correctly deal with large viscosity and density contrasts (Fig. \ref{fig:falling}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Falling%20blocks}.
-
-\begin{figure}
-\noindent\includegraphics[width=300px]{./Figures/Block.pdf}
-\centering
-\caption{Density evolution of the falling block experiment at $t=$ \SI{0}{\mega\year} (panel a) and $t=$ \SI{20}{\mega\year} for different viscosities of the
-block (panels b-f).}
-\label{fig:block}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Falling.pdf}
-\caption{Initial velocity relative to the density contrast at the centre of the falling block as function of the viscosity contrast between the block and the
-surrounding fluid.}
-\label{fig:falling}
-\end{figure}
-
-\subsection{2D time-dependent Stokes sphere with deformable free surface}\label{sec:sphere}
-This experiment is performed in a unit square domain with the gravity acceleration fixed to $g_y=-1$. The sphere has $\rho_s=2$ and $\eta_s=10^3$ and is
-initially centred at ($0.5;0.6$) with radius $R=0.123456789$. The fluid surrounding the sphere has $\rho_f=1$ and $\eta_f=1$ and occupies the domain for
-$y\leq 0.75$, while for $y>0.75$ the air has $\rho_a=0$ and $\eta_a=10^{-3}$. Velocity boundary conditions are set to free slip on all sides. The Courant
-number is set to 0.25. The experiments run for 200 s using grid resolutions from \num{150x150} to \num{512x512} elements, with 25 randomly distributed markers
-per element, and for different average schemes for the viscosity. The interface between the fluid and the air is tracked by means of the markers chain. Results
-in terms of velocity and pressure fields and topography variations are compared with results obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and they can be found at
-\url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere_fs2D} and
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Time_dependent_sphere}. Results in terms of the $v_{\textrm{rms}}$ in case of 
-an arithmetic average are also shown in Fig. \ref{fig:inst_spherefs}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_arithm.pdf}
-\caption{$v_{\textrm{rms}}$ for different grid resolutions as function of time for the 2D time-dependent Stokes sphere experiment in case of an arithmetic average.
-Results are compared with results obtained by ASPECT.}
-\label{fig:inst_spherefs}
-\end{figure}
-
-\subsection{Free surface stabilisation}\label{sec:stab}
-The benchmark is performed as discussed in \citet{Kaus2010a} and \citet{Thieulot2014}. The domain is square with $L_x=L_y=$ \SI{500}{\km} and a grid resolution
-of \num{200x200} elements, each of them containing 16 markers. A buoyant fluid with $\rho_1=$ \SI{3200}{\kg\per\cubic\metre} and $\eta_1=$ \SI{e20}{\pascal\s}
-is overlain by a denser fluid with $\rho_2=$ \SI{3300}{\kg\per\cubic\metre} and $\eta_2=$ \SI{e21}{\pascal\s}. The initial interface between the fluids has an
-initial sinusoidal shape of 5 km amplitude. Velocity boundary conditions are set to free slip conditions at the sides and no slip at the bottom, while the top
-is free surface. The experiment is performed with various fixed time steps, without the use of the Courant number. The vertical position of the free surface at
-$x=L_x$ is tracked for each simulation, with or without the stabilisation algorithm.
-
-The results show an instability (drunken sailor effect) increasing the time step in case that the stabilisation algorithm is not activated
-($dt=$ \SI{4200}{\year}, green line in Fig. \ref{fig:stab}), as already observed by \citet{Kaus2010a} and \citet{Thieulot2014}. Activating the algorithm the
-instability is fixed and the simulation remains stable using time step up to \SI{50000}{\year} (red line in Fig. \ref{fig:stab}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Stabilisation_algorithm}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Kaus_random.pdf}
-\caption{Evolution of the y coordinate in $x=L_x$ as function of time for different time steps, with and without the stabilisation algorithm.}
-\label{fig:stab}
-\end{figure}
-
-\subsection{Topography relaxation}\label{sec:crameri}
-This experiment is performed as in \citet{Crameri2012} using free surface and sticky air (Case 1). The domain is rectangular with $L_x=$ \SI{2800}{\km} and
-$L_y=$ \SI{700}{\km} in case of free surface and $L_y=$ \SI{800}{\km} in case of sticky air ($h_{st}=$ \SI{100}{\km}). The grid is composed by \num{256x64}
-elements for the free surface \num{560x320} elements for the sticky air. For both cases, there is a mantle of 600 km thickness with $\eta_m=$ \SI{e21}{\pascal\s}
-overlain by a highly viscous cosine-shaped lithosphere with thickness between 93 and 107 km and $\eta_l=$ \SI{e23}{\pascal\s}. Both mediums have $\rho=$
-\SI{3300}{\kg\per\cubic\metre}. In case of sticky air experiment, the air has $\rho_a=$ \SI{0}{\kg\per\cubic\metre} and $\eta_a=$ \SI{e18}{\pascal\s}. The
-gravity acceleration is set to $g_y=$ \SI{-10}{\metre\per\square\s}. Velocity boundary conditions are set to no slip at the bottom and free slip at the sides
-of the domain. In case of sticky air, velocities on top are set to free slip conditions.
-
-The maximum topography as function of time ($h(t)$) can be analytically derived in according to \[h(t)=h_0 \exp(\gamma t)\]
-where $h_0=7$ km is the initial topography, $\gamma=\num{-0.2139e-11}$ is the characteristic relaxation rate and $t$ is the time (black line in Fig.
-\ref{fig:crameri}). Maximum topography at the characteristic relaxation time $t_{rlx}=$ \SI{14.825}{\kilo\year} can be found to be $h_{rlx}=$ \SI{2576}{\m}.
-The results show that a viscosity of the air of \SI{e18}{\pascal\s} is needed in case of $h_{st}=$ \SI{100}{\km} to correctly describe the topography relaxation
-for this problem, as pointed out by \citet{Crameri2012} (purple line in Fig. \ref{fig:crameri}). The case with a true free surface well follows the analytical
-solution (orange line in Fig. \ref{fig:crameri}). In particular, at the relaxation time in case of the true free surface predicts a maximum topography of
-\SI{2570}{\m}, with an error of \SI{6}{\m} with respect to the analytical solution. Results are also compared with results obtained with a free surface in
-UNDERWORLD (\num{256x64} elements) and SULEC (\num{401x201} elements) (with courtesy from \citealp{Crameri2012}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Topography_relaxation}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Crameri.pdf}
-\caption{Maximum topography as function of time for the Crameri benchmark (coloured lines) in comparison with the analytical solution (black line) and results
-obtained with UNDERWORLD (yellow line) and SULEC (blue line).}
-\label{fig:crameri}
-\end{figure}
-
-\subsection{Spontaneous subduction}\label{sec:subduction}
-This experiment is performed as presented by \citet{Schmeling2008}, in case of both a sticky air and a true free surface and using different average schemes
-for the viscosities. The domain is rectangular with $L_x=$ \SI{3000}{\km} and $L_y=$ \SI{700}{\km}, a grid resolution of \num{256x64} elements with 50 markers
-per element and Courant number of 0.01. At the beginning of the simulation, a 100 km-thick lithospheric layer with $\rho_m=$ \SI{3300}{\kg\per\cubic\metre}
-and $\eta_m=$ \SI{e23}{\pascal\s} is located at the top of the domain for $x$ between 1000 and 3000 km, while the asthenospheric mantle has $\rho_m=$
-\SI{3200}{\pascal\s} and $\eta_m=$ \SI{e21}{\pascal\s}. In addition, a 200 km-depth lithospheric slab is already subducted in the mantle in order to have
-a spontaneous subduction. In case of sticky air, $L_y=750$ km and a 50 km-thick air layer with $\eta_a$ from \SIrange{e19}{e21}{\pascal\s} overlie the mantle.
-Velocity boundary conditions are set to free slip on the sides and on the bottom of the domain. In case of sticky air, velocities on the top boundary are set
-to free slip conditions as well.
-
-The maximum depth of the slab is tracked for all the simulations (Fig. \ref{fig:subduction}) and they are compared with results obtained by
-\citet{Schmeling2008} with models with regular grids and comparable grid resolutions (rectangles in Fig. \ref{fig:subduction}). Low viscous sticky air models
-(i.e., \SI{e19}{\pascal\s}) enlighten a strong dependence of the sinking velocity with the chosen average scheme (dashed lines in Fig. \ref{fig:subduction}),
-while high viscous sticky air models (i.e., \SI{e21}{\pascal\s}) show a higher resistance at the trench, underestimating the correct solution (dotted lines in
-Fig. \ref{fig:subduction}). A high resistance at the trench can be also observed in case of a true free surface for the low resolution grid used (continuous
-lines in Fig. \ref{fig:subduction}), as pointed out by \citet{Schmeling2008}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Spontaneous_subduction}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Subduction.pdf}
-\caption{Maximum depth of the slab as function of time for all the simulations of the \citet{Schmeling2008} benchmark, in case of harmonic, geometric and
-arithmetic means (red, green and blue lines, respectively) using a sticky air or a true free surface (discontinuous and continuous lines, respectively).
-Red, green and blue rectangular areas indicate the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{500}{\km} in case of sticky air,
-while the grey rectangular area indicates the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{600}{\km} in case of true free surface.}
-\label{fig:subduction}
-\end{figure}
-
-\subsection{The slab detachment}\label{sec:slab}
-The slab detachment benchmark is performed as described by \citet{Schmalholz2011} and \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{1000}{\km}
-and $L_y=$ \SI{660}{\km} and a grid resolution of \num{256x256} elements. A non-linear viscous T-shaped layer with $\rho_l=$ \SI{3300}{\kg\per\cubic\metre} is
-placed at the top of the domain and surrounded by a linear viscous fluid with $\rho_f=$ \SI{3150}{\kg\per\cubic\m}. The top layer is 80 km-thick and a 250
-km-long and 80 km-wide slab is placed at $x=L_x/2$ (Fig. \ref{fig:slab}a). The effective viscosity of the top layer is given by
-\[\eta_{eff}=\eta_0 I_2^{\frac{(1-n)}{n}}\]
-with $\eta_0=$ \SI{4.75e11}{\pascal\s} and $n=4$, while the surrounding fluid has $\eta_f=$ \SI{e21}{\pascal\s}. The effective viscosity is capped between
-$\eta_{min}=$ \SI{e21}{\pascal\s} and $\eta_{max}=$ \SI{e25}{\pascal\s}. Velocity boundary conditions are set to free slip at the top and the bottom, and to
-no slip at the sides of the domain.
-
-The sinking of the slab determines a decrease of the effective viscosity in the T-shaped layer during the first part of the evolution (Fig. \ref{fig:slab}a),
-as consequence of local strain rates, while effective viscosities increase up to $\eta_{max}$ in the top layer after approximately \SI{20}{\mega\year}, when
-necking is complete (Fig. \ref{fig:slab}b).
-
-The width of necking is tracked by means of markers position at the side of the slab. Necking width and time are normalised in function of the initial slab
-(\SI{80}{\km}) and the characteristic time ($t_c=$ \SI{7.1158e14}{\s}) \citep{Schmalholz2011,Glerum2018}. The results in case of arithmetic, geometric and
-harmonic mean of the viscosity are shown in Fig. \ref{fig:necking}, compared with results from \citet{Glerum2018} for which an infinity norm mean was used.
-The influence of the average schemes agrees with results shown in Section \ref{sec:subduction}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Slab_detachment}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Slab.pdf}
-\caption{Effective viscosity for the slab detachment benchmark at the beginning of the evolution (panel a) and when necking is complete (panel b).}
-\label{fig:slab}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Necking.pdf}
-\caption{Normalised width of the necking of the slab detachment benchmark as function of normalised time.}
-\label{fig:necking}
-\end{figure}
-
-\subsection{The indenter experiment}\label{sec:indenter}
-The indenter benchmark simulate a rigid punch on a purely plastic von Mises material, for which exists an analytical solution
-\citep{Thieulot2008,Thieulot2014,Glerum2018}. The domain is a unit square with a grid resolution of \num{256x256} elements. The gravity acceleration is set
-to $\bm{g}=\bm{0}$ and effective viscosity is capped using $\eta_{min}=10^{-4}$ and $\eta_{max}=10^3$. The tolerance for the convergence of the non-linear
-solution is $tol=10^{-9}$, with a maximum number of non-linear iterations set to 500. The medium in the domain has $\rho=0.01$, an initial viscosity
-$\eta_0=10$, cohesion $C=1$ and an angle of internal friction $\phi=0$°. Velocity boundary conditions are set to no slip at the bottom and free slip at the
-sides of the domain. The top of the domain is open with the exception of the central portion (punch area) with width $w_p=0.125$, where $v=-1.05$ and $u$ is
-fixed either to 0 (rough punch experiment) or free (smooth punch experiment).
-
-Fig. \ref{fig:indenter} show the results in terms of viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure
-(panels g and h) for both experiments. The analytical solution indicate that pressure at ($0.5;1$) and ($0.5 \pm w_p;1$)is $p=\pi +1$ and $p=1$, respectively
-(black and red lines in Fig. \ref{fig:smooth_p}). Fig. \ref{fig:smooth_p} shows a clear improvement in the pressure solution when passing from a rough to a
-smooth punch (green and blue lines in Fig. \ref{fig:smooth_p}, respectively), as previously observed by \citet{Thieulot2014} and \citet{Glerum2018}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Indenter}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Indenter.pdf}
-\caption{Viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure (panels g and h) fields for rough (left column) and
-smooth (right column) punch experiments.}
-\label{fig:indenter}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Smooth_Rough.pdf}
-\caption{Nodal pressure in function of the x coordinate at $y=1$ for rough (green line) and smooth (blue line) punch experiment. Black and red lines indicate
-the analytical solution at $x=0.5$ and $x=0.5 \pm w_p$, respectively.}
-\label{fig:smooth_p}
-\end{figure}
-
-\subsection{The brick experiment}\label{sec:brick}
-An instantaneous version of the brick benchmark is performed to verify the correctness of the pressure-dependent plasticity for different angle of internal
-friction, as proposed by \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{40}{\km} and $L_y=$ \SI{10}{\km} and a grid resolution of \num{512x128}
-elements. The gravity acceleration is set to $g_y=$ \SI{-10}{\m\square\s} and effective viscosity is capped using $\eta_{min}=$ \SI{e19}{\pascal\s} and
-$\eta_{max}=$ \SI{e26}{\pascal\s}. The tolerance for the convergence of the non-linear solution is $tol=10^{-7}$, with a maximum number of non-linear iterations
-set to 1000. A 800 m-wide and 400 m-height inclusion with $\eta_b=$ \SI{e20}{\pascal\s} and $\rho_b=$ \SI{2700}{\kg\per\cubic\m} is placed at the bottom of the
-domain at $x=L_x/2$. The inclusion is surrounded by a non-linear viscous medium with $\rho_m=$ \SI{2700}{\kg\per\cubic\m}, an initial viscosity of
-\SI{e23}{\pascal\s}, a linear viscous viscosity of \SI{e25}{\pascal\s} and a cohesion of \SI{40}{\mega\pascal}. Velocities are set to free slip conditions at
-the bottom of the domain and the top is open. Velocities on sides of the domain are fixed to $u=$ \SI{+-2e-11}{\m\per\s} and $v=0$. The experiment is performed
-in compressional and extensional contexts with an angle of internal friction $\phi$ of the non-linear medium variable between 0° and 30°.
-
-As expected, two shear bands stem from the inclusion with variable angles in relation to both the dynamics context and the internal friction angle. Shear band
-angles formed at 45° for $\phi=0$° in both extensional and compressional contexts (Fig. \ref{fig:brick_beam}a), while they are different in case of $\phi\neq0$°.
-Shear band angles in case of $\phi=20$° are shown for compression and extension in Fig. \ref{fig:brick_beam}b  and c, respectively. Values of shear band angles
-as function of internal friction angles from 0° to 30° are extracted at different depths and minimum and maximum values are plotted in Fig. \ref{fig:brick},
-in comparison with theoretical Roscoe, Arthur and Coulomb shear band angles. For all tests the tolerance for nonlinear convergence is defined $tol=10^{-7}$
-and none converges before the maximum number of iterations ($it_{max}=1000$) is reached. However, tests with internal friction angles up to 15° show a constant
-decrease of the velocity residuals, which can not be observed for for tests with higher internal friction angles (Fig. \ref{fig:convergence}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Brick_experiment}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Brick_Beam.pdf}
-\caption{Shear band angles predicted for the brick experiment for $\phi=0$° (panel a) and for $\phi=20$° in case of compression and extension (panels b and c,
-respectively).}
-\label{fig:brick_beam}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Brick.pdf}
-\caption{Shear band angles predicted for the brick experiment in case of compressional and extensional contexts as function of different internal angle of
-friction (continuous black line and red dots), compared with theoretical Roscoe, Artur and Coulomb angles (discontinuous black lines). Red lines indicate the
-range of angles calculated at different depths.}
-\label{fig:brick}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \noindent\includegraphics[width=500px]{./Figures/convergence.pdf}
-  \caption{Convergence rates for all tests. Normalised residuals (left column) are compared with velocity residuals (right column) for extensional (first row)
-  and compressional (second row) settings.}
-  \label{fig:convergence}
-  \end{figure}
-
-\subsection{Advection stabilisation}\label{sec:advection}
-The 1D advection problem proposed by \citet{Donea2003} and \citet{Thieulot2011} is performed to verify the effectiveness of the SUPG method to stabilise the
-advection term of the energy equation. The domain is a 1D segment with $L_x=1$ composed by 50 elements and a discontinuity in the thermal field placed at
-$x=0.25$. Temperature is set to 1 for $x \leq 0.25$ and to 0 for $x > 0.25$. Velocity is set to ${u}=1$ in the entire domain. The simulation is performed for
-250 time steps, with $dt=0.002$, so the thermal discontinuity should be at $x=0.75$ at the end of the simulation.
-
-Temperature profiles at the end of the simulation are shown in \ref{fig:advection} as function of the dimensionless coefficient $\gamma=\tau \bm{v}/h$. In case
-of the classic Galerkin method ($\gamma=0$, blue line in \ref{fig:advection}) the final thermal profile is characterised by strong oscillations, which are
-eliminated in case of the SUPG method ($\gamma=0.045$, orange line in \ref{fig:advection}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Advection_stabilisation}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Advection.pdf}
-\caption{Temperature profile as function of \textit{x} for the advection stabilisation benchmark. Purple line indicates the initial temperature profile; the
-green line indicates the analytical temperature profile after 250 time steps; blue line indicates the temperature profile after 250 time steps in case of the
-classic Galerkin method ($\gamma=0$); orange line indicates the temperature profile after 250 time steps in case of the SUPG method ($\gamma=0.045$).}
-\label{fig:advection}
-\end{figure}
-
-\subsection{Shear heating}\label{sec:simple_shear}
-This exercise is performed in an unit square domain composed by \num{128x128} elements. The velocity field is prescribed on the entire domain with
-$\bm{v}=(L_y-y)y\bm{e_x}$; viscosity, density and specific heat are fixed to 1, while thermal conductivity and radiogenic energy are fixed to 0. Therefore,
-the energy equation (Eq. \ref{eq:energy}) can be simplified as \[\frac{\partial T}{\partial t}=H_s\] and fixing $T(t=0)=0$, the temperature can be find as
-\[T(t)=H_s t\]
-In this case we have
-\begin{eqnarray}
-\dot{\epsilon_{xx}}&=&\frac{\partial u}{\partial x}=0 \nonumber \\
-\dot{\epsilon_{yy}}&=&\frac{\partial v}{\partial y}=0 \nonumber \\
-\dot{\epsilon_{xy}}&=&\frac{1}{2}\left(\frac{\partial v}{\partial x}+\frac{\partial u}{\partial y}\right)=\frac{1}{2}(L_y-2y)\nonumber
-\end{eqnarray}
-and, simplifying Eq. \ref{eq:shear}, shear heating can be calculated as \[H_s(x,y)=(1-2y)^2\]
-
-The solution predicted by the model in terms of velocity, temperature and shear heating match well with the analytical solutions (Fig. \ref{fig:analytical_en}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Simple_shear}.
-
-%\begin{figure}
-%\centering
-%\noindent\includegraphics[width=400px]{./Figures/Energy_Thieulot.pdf}
-%\caption{Temperature and shear heating predicted by the model after 1 time step.}
-%\label{fig:en_thieulot}
-%\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=500px]{./Figures/Analytical_Shear.pdf}
-\caption{Velocity (panel a), temperature (panel b) and shear heating (panel c) as function of the \textit{y} coordinate for the simple shear experiment. The
-solutions predicted by the model (red lines) are compared with the analytical solutions (dashed blue lines).}
-\label{fig:analytical_en}
-\end{figure}
-
-\subsection{Shear and adiabatic heating}\label{sec:shear}
-This problem is performed as presented in Exercise 9.4 in \citet{Gerya2010b}. Two materials are vertically separated in a rectangular domain with $L_x=$
-\SI{1000}{\km}, $L_y=$ \SI{1500}{\km} and a grid resolution of \num{30x20} elements. Constant thermal coefficient expansion ($\alpha=$ \SI{3e-5}{\per\kelvin}),
-temperature ($T=$ \SI{1300}{\kelvin}) and gravity acceleration ($g_y=$ \SI{-10}{\m\square\s}) are assumed in the whole domain. Fluid 1 (on the left side of the
-domain) has $\rho_1=$ \SI{3200}{\kg\per\cubic\m} and $\eta_1=$ \SI{e20}{\pascal\s}; fluid 2 (on the right side of the domain) has $\rho_2=$
-\SI{3300}{\kg\per\cubic\m} and $\eta_2=$ \SI{e22}{\pascal\s}. Velocity boundary conditions are set to free slip on all sides of the domain.
-
-As shown in Fig. \ref{fig:energy}, both shear and adiabatic heating predicted by the code (first row in Fig. \ref{fig:energy}) well recreate the results
-obtained by means of example ${Shear\_adiabatic\_heating.m}$ from \citet{Gerya2010b} (second row in Fig. \ref{fig:energy}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Adiabatic%2BShear}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Energy.pdf}
-\caption{Comparison between shear (first column), adiabatic (second column) and total (third column) energy predicted by the code (panels a, b and c) and those
-created using example ${Shear\_adiabatic\_heating.m}$ from Exercise 9.4 in \citet{Gerya2010b} (panels d, e and f).}
-\label{fig:energy}
-\end{figure}
-
-\subsection{Mantle convection}\label{sec:mantle}
-This problem is performed as presented by \citet{Blankenbach1989} (constant viscosity cases) and \citet{Thieulot2014}, in a 2D unit square domain with gravity
-acceleration $g_y=10^{10}Ra$. The experiment is performed with three different Rayleigh numbers ($Ra=10^4$, $10^5$ and $10^6$) and with different grid
-resolution (between \num{32x32} and \num{128x128} elements). The fluid has constant viscosity, initial density, heat capacity, thermal conductivity
-($\eta= \rho_0= C_p = k=1$), reference temperature ($T_0=0$) and thermal expansion coefficient ($\alpha=10^{-10}$). Temperatures are set to 0 on top and 1 on
-bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\]
-
-The solution generated by the code in terms of $v_{\textrm{rms}}$ (calculated as in Sec. \ref{sec:error}) and Nu as function of time are reported for all the
-simulations in Fig. \ref{fig:mantle}. At the steady state, $v_{\textrm{rms}}$ and Nu of all simulations converge well toward the values from \citet{Blankenbach1989},
-with lower errors for higher resolution grids (Table \ref{tab:mantle}). All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Mantle_convection}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Convection.pdf}
-\caption{$v_{\textrm{rms}}$ (panels a, c and e) and Nu (panels b, d and f) for the mantle convection benchmark as function of time for different grid resolution.
-Panels a and b show the results for $Ra=10^4$; Panels c and d show the results for $Ra=10^5$; panels e and f show the results for $Ra=10^6$. Purple lines
-indicate the convergence values for the $v_{\textrm{rms}}$ and the Nu from \citet{Blankenbach1989} at the steady state.}
-\label{fig:mantle}
-\end{figure}
-
-\begin{table}[H]
-\caption{Comparison between $v_{\textrm{rms}}$ and Nu predicted by the code for the mantle convection experiment and same values as reported in literature.}
-\centering
-\small
-\begin{tabular}{|>{\centering}m{0.09\textwidth}|>{\centering}m{0.05\textwidth}|>{\centering}m{0.24\textwidth}|>{\centering}m{0.15\textwidth}|
-  >{\centering}m{0.1\textwidth} >{\centering}m{0.1\textwidth} >{\centering\arraybackslash}m{0.1\textwidth}|}
-\toprule
-& & \multirow{2}{*}{\citet{Blankenbach1989}} & \citet{Thieulot2014} & \multicolumn{3}{c|}{FALCON}  \\
- & & & \num{200x200} & \num{32x32} & \num{64x64} & \num{128x128}  \\
-\midrule
-\multirow{2}{*}{$Ra=10^4$} & $v_{\textrm{rms}}$ & $42.864947 \pm 0.000020$ & 42.867 & 42.83226 & 42.852793 & 42.861394  \\
-           & Nu      & $4.884409 \pm 0.000010$  & 4.882  & 4.781297 & 4.857475  & 4.877573   \\
-\hline
-\multirow{2}{*}{$Ra=10^5$} & $v_{\textrm{rms}}$ & $193.21454 \pm 0.00010$  & 193.255& 193.872643 & 193.377472 & 193.252290  \\
-           & Nu      & $10.534095 \pm 0.000010$ & 10.507 & 9.602514 & 10.270735  & 10.465629   \\
-\hline
-\multirow{2}{*}{$Ra=10^6$} & $v_{\textrm{rms}}$ & $833.98977 \pm 0.00020$  & 834.712& 848.091176 & 837.767911 & 834.945793  \\
-           & Nu      & $21.972465 \pm 0.000020$ & 21.695 & 15.999266 & 19.703682  & 21.306939   \\
-\bottomrule
-\end{tabular}
-\label{tab:mantle}
-\end{table}
-
-\subsection{Viscoplastic mantle convection}\label{sec:visco_mantle}
-This benchmark is performed as Cases 1-5a described in \citet{Tosi2015} in a 2D unit square domain with two different grid resolutions of \num{32x32} 
-and \num{100x100} elements. Reference viscosity, reference density, heat capacity, thermal conductivity and thermal expansion coefficient are fixed to a constant
-value of 1 ($\eta_0= \rho_0= C_p = k=\alpha=1$), while the gravity has been chosen as $g_y=10^2$, in order to obtain a Rayleigh number $Ra=10^2$. Temperatures
-are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\].
-The viscosity field $\eta$ is calculated as the harmonic average between a linear part $\eta_{lin}$ that depends on temperature only or on temperature and 
-depth $y$ and a nonlinear-plastic part $\eta_{pl}$ that depends on the strain rate $\dot{\bm{\epsilon}}$, as follows
-\[\eta(T,y,\dot{\bm{\epsilon}})=2\left(\frac{1}{\eta_{lin}(T,y)}+\frac{1}{\eta_{pl}(\dot{\bm{\epsilon}})}\right)^{-1}\]
-The linear and the nonlinear-plastic parts of the viscosity are calculated following \citet{Tosi2015} as follows
-\[\eta_{lin}(T,y)=\exp(-\ln(\Delta\eta_T)T+\ln(\Delta\eta_y)y)\]
-\[\eta_{pl}(\dot{\bm{\epsilon}})=\eta^*+\frac{\sigma_Y}{\sqrt{\dot{\bm{\epsilon}}:\dot{\bm{\epsilon}}}}\]
-where $\Delta\eta_T$, $\Delta\eta_y$, $\eta^*$ and $\sigma_Y$ are parameters chosen as listed in Table \ref{tab:tosi_setup}.
-
-For each case, some diagnostic quantities measured at the steady state are reported in Table \ref{tab:visco_mantle} (see Section \ref{sec:error} for 
-the computation of all the quantities), showing a well fit with results described in \citet{Tosi2015}. Data and charts for all cases can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Viscoplastic_mantle_convection}.
-
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{white!100!}{gray!25!}
-%\begin{table}[H]
-%  \caption{Results of the viscoplastic mantle convection benchmark for Cases 1-4 compared with results obtained with ELEFANT and YACC (from \citealp{Tosi2015}).}
-%  \centering
-%  \resizebox{\textwidth}{!}{
-%\begin{tabular}{|c|ccc|ccc|ccc|ccc|}
-%  \toprule
-%    & \multicolumn{3}{c|}{\textbf{Case 1}} & \multicolumn{3}{c|}{\textbf{Case 2}} & \multicolumn{3}{c|}{\textbf{Case 3}} &
-%    \multicolumn{3}{c|}{\textbf{Case 4}} \\
-%   & FALCON & ELEFANT& YACC & FALCON& ELEFANT& YACC & FALCON & ELEFANT & YACC & FALCON & ELEFANT & YACC \\
-%  \midrule
-%   $\left\langle T \right\rangle$ & 0.7759 & 0.7758 & 0.7767 & 0.6043 & 0.6036 & 0.6058 & 0.7275 & 0.7275 & 0.7286 & & 0.5277 & 0.5289\\
-%    Nu$^{top}$ & 3.4280 & 3.4214 & 3.4298 & 8.5242 & 8.5115 & 8.5278 & 3.0368 & 3.0347 & 3.0374 && 6.5912 & 6.5572 \\
-%    Nu$^{bottom}$ & 3.3168 & 3.313 & 3.3143 & 8.4599 & 8.4465 & 8.3990 & 3.0086 & 2.9908 & 2.9628 && 6.5834 & 6.5243 \\
-%    $v_{\textrm{rms}}$ & 249.901 & 249.134 & 251.7997 & 141.2451 & 140.7087 & 142.2020 & 100.2776 & 100.1208 & 100.9467 && 79.1105 & 79.6202 \\
-%    $v_{\textrm{rms}}^{top}$ & 1.8907 & 1.8642 & 1.8298 & 104.7928 & 104.1228 & 105.1767 & 2.0832 & 2.0652 & 2.0374 && 74.7596 & 75.8414 \\
-%    $u_{max}^{top}$ & 2.6340 & 2.6119 & 2.5516 & 121.9212& 121.766 & 122.3238 & 2.9086 & 2.9019 & 2.8458 && 88.9146 & 89.2940 \\
-%    $\eta_{min}$ & - & - & - & $2.0303\times10^{-5}$ & $2.0600\times10^{-5}$ & $1.9853\times10^{-5}$ & $4.8129\times10^{-5}$ & $4.8080\times10^{-5}$ & $4.7907\times10^{-5}$ &
-%    & $1.9860\times10^{-4}$ & $1.9174\times10^{-4}$ \\
-%    $\eta_{max}$ & - & - & - & 1.3325 & 1.85 & 1.6095 & 0.9360 & 0.9023 & 1 && 1.52 & 1.6773 \\
-%    $\left\langle W \right\rangle$ & 2.4299 & 2.4316 & 2.4583 & 7.5648 & 7.5665 & 7.6222 & 2.0387 & 2.0384 & 2.0400 && 5.6216 & 5.6512 \\
-%    $\left\langle \Phi \right\rangle/Ra$ & 2.4311 & 2.4276 & 2.4333 & 7.5661 & 7.5585 & 7.6194 & 2.0386 & 2.0370 & 2.0335 && 5.6182 & 5.6463 \\
-%    $\delta$ & 0.05\% & 0.16\% & 1.02\% & 0.02\% & 0.11\% & 0.04\% & $<$0.01\% & 0.07\% & 0.32\% && 0.06\% & 0.09\% \\
-%  \bottomrule
-%  \end{tabular}}
-%  \label{tab:visco_mantle}
-%  \end{table}
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{2}{white!100!}{gray!25!}
-\begin{table}[H]
-  \caption{Benchmark cases and corresponding parameters.}
-  \centering
-  \small
-\begin{tabular}{|>{\centering}m{0.08\textwidth}|>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}
-  >{\centering}m{0.05\textwidth} >{\centering}m{0.05\textwidth} |>{\centering\arraybackslash}m{0.2\textwidth}|}
-  \toprule
-    Case & $Ra$ & $\Delta\eta_T$ & $\Delta\eta_y$ & $\eta^*$ & $\sigma_Y$ & Convective regime \\
-  \midrule
-  \rowcolor{gray!25!}
-    1 & $10^2$ & $10^5$ & 1 & - & - & Stagnant lid \\
-    2 & $10^2$ & $10^5$ & 1 & $10^{-3}$ & 1 & Mobile lid  \\
-    \rowcolor{gray!25!}
-    3 & $10^2$ & $10^5$ & 10 & - & - & Stagnant lid  \\
-    4 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 1 & Mobile lid \\
-    \rowcolor{gray!25!}
-    5 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 4 & Periodic \\
-  \bottomrule
-  \end{tabular}
-  \label{tab:tosi_setup}
-  \end{table}
-
-\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{gray!25!}{white!100!}
-\begin{table}
-  \caption{Results of the viscoplastic mantle convection benchmark for different grid resolutions in Cases 1-5 compared with results obtained with ELEFANT
-  and YACC ($100\times100$ elements, from \citealp{Tosi2015}) and in Stone 28 ($32\times32$ elements, from
-  \url{https://github.com/cedrict/fieldstone/tree/master/python_codes/fieldstone_28}).}
-  \centering
-  \resizebox{\textwidth}{!}{
-\begin{tabular}{|cc|ccccccccccc|}
-  \toprule
-  \multicolumn{13}{|c|}{\textbf{Case 1}} \\
-  \midrule
-  Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-  & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-  \hline
-  \rowcolor{gray!25!}
-    YACC & $100\times100$ & 0.7767 & 3.4298 & 3.3143 & 251.7997 & 1.8298 & 2.5516 & - & - & 2.4583 & 2.4333 & 1.02\% \\
-    ELEFANT & $100\times100$ & 0.7758 & 3.4214 & 3.313 & 249.134 & 1.8642 & 2.6119 & - & - & 2.4316 & 2.4276 & 0.16\% \\
-    \rowcolor{gray!25!}
-    Stone 28 & $32\times32$ & 0.7737 & 3.3987 & - & 243.872 & - & - & - & - & - & - & - \\
-%    FALCON & $32\times32$ & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & - & - & 2.4745 & 2.4864 & 0.48\% \\
-%    \rowcolor{gray!25!}
-    FALCON & $100\times100$ & 0.7759 & 3.4280 & 3.3168 & 249.901 & 1.8907 & 2.6340 & - & - & 2.4299 & 2.4311 & 0.05\% \\
-    \midrule
-    \multicolumn{13}{|c|}{\textbf{Case 2}} \\
-    \midrule
-    Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-    & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-    \hline
-    \rowcolor{gray!25!}
-     YACC & $100\times100$ & 0.6058 & 8.5278 & 8.3990 & 142.2020 & 105.1767 & 122.3238 & $1.9853\times10^{-5}$ & 1.6095 & 7.6222 & 7.6194 & 0.04\% \\
-     ELEFANT & $100\times100$ & 0.6034 & 8.5115 & 8.4465 & 140.8067 & 104.1228 & 121.766 & $2.0600\times10^{-5}$ & 1.85 & 7.5665 & 7.5585 & 0.11\% \\
-     \rowcolor{gray!25!}
-     Stone 28 & $32\times32$ & 0.6052 & 8.1715 & - & 141.518 & - & - & - & - & - & - & - \\
-%     FALCON & $32\times32$ & 0.6454 & 8.1422 & 7.4020 & 156.6908 & 95.7226 & 116.9302 & $2.2572\times10^{-5}$ & 0.7194 & 7.4289 & 7.4452 & 0.22\% \\
-%     \rowcolor{gray!25!}
-     FALCON & $100\times100$ & 0.6044 & 8.5266 & 8.4590 & 141.3684 & 104.8867 & 122.0470 & $2.0333\times10^{-5}$ & 1.2914 & 7.5707 & 7.5718 & 0.01\% \\
-     \midrule
-     \multicolumn{13}{|c|}{\textbf{Case 3}} \\
-     \midrule
-     Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-     & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-     \hline
-     \rowcolor{gray!25!} 
-     YACC & $100\times100$ & 0.7286 & 3.0374 & 2.9628 & 100.9467 & 2.0374 & 2.8458 & $4.7907\times10^{-5}$ & 1 & 2.0400 & 2.0335 & 0.32\% \\
-      ELEFANT & $100\times100$ & 0.7275 & 3.0347 & 2.9908 & 100.1208 & 2.0652 & 2.9019 & $4.8080\times10^{-5}$ & 0.9023 & 2.0384 & 2.037 & 0.07\% \\
-      \rowcolor{gray!25!}
-      Stone 28 & $32\times32$ & 0.7277 & 3.0347 & - & 100.018 & - & - & - & - & - & - & - \\
-%      FALCON & $32\times32$ & 0.7281 & 3.0618 & 2.8132 & 101.0092 & 2.1630 & 3.0178 & $4.8500\times10^{-5}$ & 0.7845 & 2.0699 & 2.0695 & 0.02\% \\
-%      \rowcolor{gray!25!}
-      FALCON & $100\times100$ & 0.7275 & 3.0368 & 3.0086 & 100.2776 & 2.0832 & 2.9086 & $4.8129\times10^{-5}$ & 0.9360 & 2.0387 & 2.0386 & $<$0.01\% \\
-      \midrule
-      \multicolumn{13}{|c|}{\textbf{Case 4}} \\
-      \midrule
-      Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-      & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-      \hline
-      \rowcolor{gray!25!} 
-      YACC & $100\times100$ & 0.5289 & 6.5572 & 6.5243 & 79.6202 & 75.4814 & 89.2940 & $1.9174\times10^{-4}$ & 1.6773 & 5.6512 & 5.6463 & 0.09\% \\
-       ELEFANT & $100\times100$ & 0.5277 & 6.5912 & 6.5834 & 79.1105 & 74.7596 & 88.9146 & $1.9860\times10^{-4}$ & 1.5200 & 5.6216 & 5.6182 & 0.06\% \\
-       \rowcolor{gray!25!}
-       Stone 28 & $32\times32$ & 0.5286 & 6.4036 & - & 79.4746 & - & - & - & - & - & - & - \\
-%       FALCON & $32\times32$ & 0.5334 & 6.4679 & 6.3435 & 80.6153 & 76.0511 & 90.2058 & $2.0393\times10^{-4}$ & 0.6453 & 5.6676 & 5.6698 & 0.04\% \\
-%       \rowcolor{gray!25!}
-       FALCON & $100\times100$ & 0.5279 & 6.5897 & 6.5830 & 79.1220 & 75.0715 & 88.8484 & $1.9693\times10^{-4}$ & 1.4133 & 5.6127 & 5.6113 & 0.03\% \\
-       \midrule
-       \multicolumn{13}{|c|}{\textbf{Case 5}} \\
-       \midrule
-       Code & Resolution & $P$ & $\left\langle T \right\rangle_{min}$ & $\left\langle T \right\rangle_{max}$ & Nu$^{top}_{min}$ & Nu$^{top}_{max}$ &
-       $v_{\textrm{rms}}^{min}$ & $v_{\textrm{rms}}^{max}$ & \multicolumn{2}{c}{$(\left\langle \Phi \right\rangle/Ra)_{min}$} & 
-       \multicolumn{2}{c|}{$(\left\langle \Phi \right\rangle/Ra)_{max}$} \\
-       \hline
-       \rowcolor{gray!25!}
-         YACC & $100\times100$ & 0.0792 & 0.6543 & 0.6722 & 2.6950 & 7.2792 & 41.9813 & 98.8230 & \multicolumn{2}{c}{1.4347} & 
-         \multicolumn{2}{c|}{9.2748} \\
-         ELEFANT & $100\times100$ & 0.076 & 0.6514 & 0.6693 & 2.6808 & 7.3506 & 41.2578 & 100.089 & \multicolumn{2}{c}{1.3324} & 
-         \multicolumn{2}{c|}{9.5289} \\
-         \rowcolor{gray!25!}
-         Stone 28 & $32\times32$ & 0.0764 & 0.6521 & 0.6697 & 2.6912 & 7.0763 & 41.61 & 98.183 & \multicolumn{2}{c}{-} & 
-         \multicolumn{2}{c|}{-} \\
-     %    FALCON & $32\times32$ & & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & \multicolumn{2}{c}{-} & 
-     %     \multicolumn{2}{c|}{-} \\
-     %    \rowcolor{gray!25!}
-         FALCON & $100\times100$ & 0.0775 & 0.6519 & 0.6697 & 2.6847 & 7.3629 & 41.5256 & 99.5466 & \multicolumn{2}{c}{1.3399} & 
-         \multicolumn{2}{c|}{9.3993} \\
-  \bottomrule
-  \end{tabular}}
-  \label{tab:visco_mantle}
-  \end{table}
-
-\subsection{Thin layer entrainment}\label{sec:thin}
-This experiment is performed as originally proposed by \citet{vanKeken1997} with two incompressible fluids in a rectangular domain with $L_x=2$ and $L_y=1$
-and gravity acceleration $g_y=10^{10}Ra$. The Rayleigh number and the compositional Rayleigh number are fixed to $Ra=300000$ and $Rc=450000$, respectively.
-Both fluids have constant viscosity, thermal conductivity, specific heat ($\eta=\rho=C_p=1$) and thermal expansion coefficient ($\alpha=10^{-10}$). Fluid 1
-has a density $\rho_1=1$, while fluid 2 is denser ($\rho_2=\rho_1+1.5 \cdot 10^{-10}$) and is placed at the bottom of the domain, for $y \leq 0.025$.
-Temperature are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides of the domain. The initial
-temperature field is given by
-\[T(x,y)=T_u(x,y)+T_l(x,y)+T_r(x,y)+T_s(x,y)-\frac{3}{2}\]
-with
-\begin{eqnarray}
-T_u(x,y)&=&\frac{1}{2}erf\left(\frac{1-y}{2}\sqrt{\frac{u_0}{x}}\right)\nonumber \\
-T_l(x,y)&=&1-\frac{1}{2}erf\left(\frac{y}{2}\sqrt{\frac{u_0}{L_x-x}}\right)\nonumber \\
-T_r(x,y)&=&\frac{1}{2}+\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{y+1}} \exp\left(-\frac{x^2u_0}{4y+4}\right)\nonumber \\
-T_s(x,y)&=&\frac{1}{2}-\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{2-y}} \exp\left(-\frac{(L_x-x)^2u_0}{8-4y}\right)\nonumber
-\end{eqnarray}
-and
-\begin{eqnarray}
-u_0&=&\frac{L_x^{7/3}}{(1+L_x^4)^{2/3}}\left(\frac{Ra}{2\sqrt{\pi}}\right)^{\frac{2}{3}}\nonumber \\
-Q&=&2\sqrt{\frac{L_x}{\pi u_0}}\nonumber
-\end{eqnarray}
-
-The experiment is performed with two grid resolution (\num{125x40} and \num{200x80} elements), with 100 markers per element and CFL of 0.25. $v_{\textrm{rms}}$
-(see Sec. \ref{sec:error}) calculated for both the simulations match well with results from \citet{vanKeken1997} and obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and ELEFANT \citep{Thieulot2014}, obtained with similar grid resolutions
-(Fig. \ref{fig:thin}). In addition, a variety of simulations with different resolutions and aspect ratios are performed to verify that the sensitivity of the
-initial velocity field. Results match well with results from \citet{vanKeken1997} (in the grey area) and \citet{Thieulot2014} (Fig. \ref{fig:thin_initial}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Thin_layer}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_thin.pdf}
-\caption{$v_{\textrm{rms}}$ for the thin layer experiment as function of time for different grid resolution. Results are compared with results obtained by
-\citet{vanKeken1997} (black lines) and with ASPECT and ELEFANT.}
-\label{fig:thin}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrmszero.pdf}
-\caption{$v_{\textrm{rms}}$ at $t=0$ for the thin layer experiment as function of the element size for different aspect ratios of the elements. Results are compared
-with results obtained by \citet{vanKeken1997} (black dots in the grey area) and with ASPECT and ELEFANT.}
-\label{fig:thin_initial}
-\end{figure}
-
-\subsection{Sinking hydrated cylinder}\label{sec:quinquis}
-This experiment is performed as explained in \citet{Quinquis2014} and simulates the sinking of a cold, hydrated cylinder into a warm, dry mantle. The domain
-is square with $L_x=L_y=$ \SI{300}{\km}, a grid resolution of \num{300x300} elements and 25 markers per element. The sinking cylinder is located at $x=$
-\SI{150}{\km} and $y=$ \SI{170}{\km} with a radius of \SI{20}{\km} and has $\rho_c=$ \SI{3250}{\kg\per\cubic\m}, $\eta_c=$ \SI{e23}{\pascal\s}, a thermal
-conductivity $k_c=$ \SI{4.5}{\watt\per\m\per\kelvin},a specific heat $C_{p_c}=$ \SI{1250}{\joule\per\kg\per\kelvin}; the surrounding mantle has $\rho_m=$
-\SI{3200}{\kg\per\cubic\m}, $\eta_m=$ \SI{e20}{\pascal\s}, a thermal conductivity $k_m=$ \SI{105}{\watt\per\m\per\kelvin} and specific heat $C_{p_m}=$
-\SI{1250}{\joule\per\kg\per\kelvin}; a 58 km-thick lithosphere overlies the mantle and has $\rho_l=$ \SI{3200}{\kg\per\cubic\m}, $\eta_l=$ \SI{e23}{\pascal\s},
-a thermal conductivity $k_l=$ \SI{4.5}{\watt\per\m\per\kelvin}, specific heat $C_{p_l}=$ \SI{750}{\joule\per\kg\per\kelvin} and a thermal diffusivity of
-\SI{e-6}{\square\m\per\s}. Velocity boundary conditions are set to free slip on all sides of the domain. Temperature increase linearly in the lithosphere,
-from 0° to 1300°C, and in the mantle, from 1300° to 1360.5°C. The initial temperature of the cylinder is fixed to 400°C. Throughout the evolution, temperature
-is fixed to 0° and 1360.5°C at the top and bottom of the domain, respectively. The initial bound water content is imposed at 2 wt.\% in the cylinder and at
-0 wt.\% in both the mantle and the lithosphere. The maximum amount of water in the mantle is fixed to 0.2 wt.\%, while maximum water content in the cylinder
-and in the lithosphere is function of pressure and temperature and is calculated using a serpentinized harzburgite with Perple\_X, as in \citet{Quinquis2014}.
-
-As shown in \citet{Quinquis2014}, the model is characterised by a progressive dehydration in the external portion of the cylinder, due to the increase of
-temperature, with a consequent vertical migration of free water that hydrates the mantle above the cylinder, up to the lithosphere. The distribution of
-bound and free water during the evolution of the experiment is shown in Fig. \ref{fig:quinquis} (left and right columns, respectively). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Hydration}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Hydration.pdf}
-\caption{Distribution of bound and free water (left and right columns, respectively) at $t=0.5$ and $t=$ \SI{1.25}{\mega\year} (first and second rows,
-respectively) for the sinking hydrated cylinder experiment. Only markers with bound or free water not equal to 0 are plotted.}
-\label{fig:quinquis}
-\end{figure}
-
-\subsection{Mantle melting curves}\label{sec:katz}
-This instantaneous experiment is performed in order to recreate isobaric melting curves, as function of temperature, bulk water and content of water in the
-melt, to compare with melting curves obtained by \citet{Katz2003}. The domain is square with $L_x=L_y=100$ km, a grid resolution of \num{100x100} elements and
-25 markers per element. Density is set in the entire domain to \SI{3300}{\kg\per\cubic\m} and temperature increases from the sides to the centre from 1000° to
-1500°C. The experiment is performed for bulk water contents from 0 to 0.3 wt.\%.
-
-Fig. \ref{fig:fraction} shows isobaric melting curves for \SI{1}{\giga\pascal} (panel a) and \SI{3}{\giga\pascal} (panel b) calculated for different bulk water
-contents. The curves well match with those obtained for the mantle by \citet{Katz2003}, showing the saturation at the solidus for bulk water of 0.2 and 0.3 wt.\%, indicating
-that the code correctly determines melt fraction as function of pressure, temperature and bulk water content. Same results can be observed in Fig.
-\ref{fig:water}a, where isobaric and isothermal melting curves are shown as function of bulk water content. In addition, Fig. \ref{fig:water}b shows that the
-code correctly determines also the percentage of water in the melt as function of pressure, temperature and melt fraction. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Melting}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Fraction.pdf}
-\caption{Isobaric melting curves for 1 GPa (panel a) and 3 GPa (panel b) as function of temperature with different bulk water contents.}
-\label{fig:fraction}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Water.pdf}
-\caption{Isobaric melting curves for 1.5 GPa as function of bulk water (panel a) and water in the melt (panel b) calculated for different temperatures.}
-\label{fig:water}
-\end{figure}
-
-\clearpage
-
-\bibliography{export}
-
-\end{document}

### 2022-09-07 tex/Benchmark/katz.tex
update main

--- 2022-03-31 FALCON_description.tex
@@
-\documentclass[hidelinks,10pt,a4paper]{article}
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{textcomp}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage{booktabs}
-\usepackage{multirow}
-\usepackage[cm]{fullpage}
-\usepackage[round]{natbib}
-\bibliographystyle{abbrvnat}
-\usepackage{doi}
-\usepackage{bm}
-\usepackage[table]{xcolor}
-\hypersetup{
-    colorlinks,
-    linkcolor={blue},
-    citecolor={blue},
-    urlcolor={blue}
-}
-\usepackage{float}
-\restylefloat{table}
-\usepackage{siunitx}
-\DeclareSIUnit\year{yr}
-
-\begin{document}
-
-\section{Introduction}\label{sec:intro}
-This document includes a description of the code FALCON, with a presentation of the features implemented followed by results of extensive benchmarks listed in
-Table \ref{tab:list}, performed to verify that the code solves correctly a wide range of problems. Description and results of all benchmarks are shown
-in Section \ref{sec:benchmark}.
-
-\begin{table}[H]
-\caption{List of benchmarks performed for each feature of the code. Data of all benchmarks can be found at 
-\url{https://github.com/aleregorda/Benchmarks}}
-\centering
-\begin{tabular}{| l | l | l |}
-\toprule
- \textbf{Feature}  & \textbf{Benchmark} & \textbf{Section} \\
-\midrule
-  Solver  & Stokes flow & \ref{sec:stokes} \\
-\hline
-  \multirow{2}{*}{Markers advection} & Zalesak disk & \ref{sec:runge} \\
-    & Conservative Interpolation Velocity & \ref{sec:CVI} \\
-\hline
-  \multirow{4}{*}{Momentum equation}  & Poiseuille flow & \ref{sec:poiseuille}\\
-    & Instantaneous 2D sphere & \ref{sec:ist_sphere} \\
-    & Rayleigh-Taylor experiment & \ref{sec:rayleigh} \\
-    & Falling block & \ref{sec:block} \\
-\hline
-    & 2D Stokes sphere & \ref{sec:sphere} \\
-    Sticky air and  & Stabilisation algorithm & \ref{sec:stab} \\
-    free surface & Topography relaxation & \ref{sec:crameri} \\
-  & Spontaneous subduction & \ref{sec:subduction} \\
-\hline
-  Erosion and sedimentation  & - & - \\
-\hline
-   Non-linear  & Slab detachment & \ref{sec:slab} \\
-   visco-plasticity & Indenter & \ref{sec:indenter} \\
-    & Brick & \ref{sec:brick} \\
-\hline
-  \multirow{3}{*}{Energy equation}  & Advection stabilisation & \ref{sec:advection} \\
-    & Simple shear heating & \ref{sec:simple_shear} \\
-    & Shear and adiabatic heating & \ref{sec:shear} \\
-\hline
-\multirow{3}{*}{Energy + momentum} & Mantle convection & \ref{sec:mantle} \\
-& Viscoplastic mantle convection & \ref{sec:visco_mantle} \\
-& Thin layer entrainment & \ref{sec:thin} \\
-\hline
-  Phase changes  & \multirow{2}{*}{Hydrated sinking cylinder} & \multirow{2}{*}{\ref{sec:quinquis}} \\
-  and hydration  &  & \\
-\hline
-  Melting & Experimental melting curves & \ref{sec:katz} \\
-\bottomrule
-\end{tabular}
-\label{tab:list}
-\end{table}
-
-The code is written in Fortran90 and uses the finite element method (FEM) with quadrilateral $Q_1 \times P_0$ elements (continuous bilinear velocity and
-discontinuous constant pressure), associated with MUMPS\footnote{\url{http://mumps.enseeiht.fr/}} \citep{Amestoy2001,Amestoy2019} that is a software package
-for solving systems of linear equations of the form $A \cdot x = b$, where \textit{A} is a square sparse matrix, by means of a direct method. Since
-$Q_1 \times P_0$ elements do not satisfy Ladyzhenskaya, Babuska and Brezzi (LBB) stability condition \citep{Donea2003}, elemental pressure is elaborated
-in post-processing to avoid spurious pressures by means of a double interpolation to smooth the pressure. Following this procedure, the elemental pressure
-is interpolated onto nodes and then back onto elements. After the smoothing procedure, the pressure field is calculated again on the nodes to be used in
-combination with the lithostatic pressure, which is calculated onto nodes. Correctness of the solution and performances of the solver in terms of time and
-memory usage are tested solving the Stokes flow with the analytical solution proposed by \citet{Donea2003} (Section \ref{sec:stokes}).
-
-The thermo-mechanics of crust-mantle systems is described by means of the conservation of mass, momentum and energy equations, expressed as follows:
-\begin{eqnarray}
-\label{eq:mass}\frac{\partial \rho}{\partial t}+\bm{\nabla} \cdot (\rho \bm{u})&=&0\\
-\label{eq:momentum}-\bm{\nabla} p + \bm{\nabla} \cdot \bm{\tau} + \rho \bm{g}&=&0\\
-\label{eq:energy}\rho C_p \left(\frac{\partial T}{\partial t} + \bm{u} \cdot \bm{\nabla} T \right)&=&\bm{\nabla} \cdot \left(k\bm{\nabla} T\right) + H_{tot}
-\end{eqnarray}
-where $\rho$ is the density, $\bm{u}$ is the velocity, \textit{p} is the pressure, $\bm{\tau}$ is the deviatoric stress, $\bm{g}$ is the gravity acceleration,
-$C_p$ is the specific heat at constant pressure, \textit{T} is the temperature, \textit{k} is the thermal conductivity and $H_{tot}$ is the total internal heat
-production. Density variations due to temperature are generally small enough to assume the density as constant ($\rho=\rho_0$) in Eq. \ref{eq:mass} and in
-Eq. \ref{eq:energy}, while it must be treated as a variable in the buoyancy term of Eq. \ref{eq:momentum}, such that \[\rho=\rho_0(1-\alpha(T-T_0))\]
-where $\rho_0$ is the density at a reference temperature $T_0$ and $\alpha$ is the coefficient of thermal expansion. Eq. \ref{eq:mass} then can be rewritten as
-\begin{equation}\label{eq:mass_0}
-\bm{\nabla} \cdot \bm{u}=0
-\end{equation}
-This simplification is known as the extended Boussinesq approximation.
-
-The time step $dt$ is chosen in according to Courant-Friedrichs-Lewy condition \citep{Anderson1995}, such that
-\[dt=C_n\; \min\left(\frac{\Delta x}{\max|\bm{u}|},\frac{(\Delta x)^2}{\kappa}\right)\]
-where $0<C_n<1$ is the Courant (CFL) number, $\Delta x$ is the minimum dimension of smallest element, $\max|\bm{u}|$ is the maximum velocity calculated on the
-entire domain and
-$\kappa=\frac{k}{\rho C_p}$ is the thermal diffusivity \citep{Thieulot2014}.
-
-\section{Lagrangian markers}\label{sec:markers}
-Elemental properties (density, viscosity, thermal conductivity, specific heat and thermal expansion) needed to solve Eqs. \ref{eq:mass_0}, \ref{eq:momentum}
-and \ref{eq:energy} are related to the composition of each element, determined by means of Lagrangian markers that are characteristic of different materials
-in the domain. Elemental properties (with the sole exception of the viscosity) are calculated using an arithmetic mean, as \[P_e=\frac{1}{n}\sum_{i=1}^n P_i\]
-where $P_e$ is the elemental property, $P_i$ is the property characteristic of the material of each marker and $n$ is the number of markers of the element.
-Differently, the average scheme for the elemental viscosity can be chosen between harmonic, geometric and arithmetic mean. Markers can be placed regularly or
-randomly at the beginning of the simulation and their advection is performed by either a 2nd-order or a 4th-order Runge-Kutta in space, interpolating the velocity 
-field on each marker by means of the shape functions (see Section \ref{sec:runge}). The interpolated velocity is then corrected by means of the Conservative
-Velocity Interpolation (CVI), which introduces a corrective term to reduce dispersion and clustering of particles in both steady state and time-dependent 
-\citep{Wang2015} (see Section \ref{sec:CVI}).
-
-The initial distribution of the markers is created using the open-source code library
-Geodynamic World Builder \citep{Fraters2019}. During the simulation, each marker carries memory of temperature, pressure and accumulated strain, which are
-determined interpolating the relative nodal parameter. In particular, pressure is interpolated onto each marker after the smoothing procedure. The number
-of markers contained in each element is maintained between $n_{min}$ and $n_{max}$. When in an element there are less markers than $n_{min}$ the code adds
-random markers to reach the $n_{min}$, while if the number is higher than $n_{max}$ some of them are randomly deleted. When new markers are added, they assume
-the properties of the nearest marker. In this way elements are never empty and maintain a number of markers inside a prefixed range.
-
-\section{The momentum equation}\label{sec:mom_eq}
-The penalty method is implemented so that Eq. \ref{eq:mass_0} can be written as:
-\begin{equation}\label{eq:mass_penalty}
-\bm{\nabla} \cdot \bm{u} + \frac{p}{\lambda}=0
-\end{equation}
-where $\lambda$ is the so-called penalty parameter, which should be 6-7 orders of magnitude larger than the shear viscosity to ensure that mass conservation
-is satisfied \citep{Donea2003,Thieulot2014}.
-
-The deviatoric stress tensor in Eq. \ref{eq:momentum} can be written in terms of the strain rate tensor as $\bm{\tau}=2\eta\dot{\bm{\epsilon}}$, with
-$\dot{\bm{\epsilon}}=\frac{1}{2}\left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)$. Therefore, Eq. \ref{eq:momentum} can be rewritten as:
-\begin{equation}\label{eq:momentum_pressure}
--\bm{\nabla} p + \bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right) + \rho \bm{g} = 0
-\end{equation}
-Finally, using pressure from Eq. \ref{eq:mass_penalty}, Eq. \ref{eq:momentum_pressure} can be rewritten as:
-\begin{equation}\label{eq:momentum_penalty}
-\lambda \bm{\nabla} \left(\bm{\nabla} \cdot \bm{v}\right)+\bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right)+\rho\bm{g}=0
-\end{equation}
-The penalty method is been associated to the iterative Uzawa method, as described in detail in \citet{Dabrowski2008} and \citet{Thieulot2014}
-(see Section \ref{sec:poiseuille}). Results of benchmarks performed to verify the correctness in the implementation of Eq. \ref{eq:momentum_penalty} are
-shown in Sections \ref{sec:ist_sphere} and \ref{sec:rayleigh}.
-
-To support large viscosities variations the penalty parameter is related to the effective elemental viscosity by means of a dimensionless coefficient, so that
-$\lambda=\lambda_e(e)\eta_{eff}(e)$ \citep{Marotta2006,Dabrowski2008,Thieulot2014}. Benchmark of the falling block \citep{Gerya2003a,Gerya2010b,Thieulot2011}
-is performed to verify that the code can correctly deal buoyancy driven flows with strong viscosity contrasts (Section \ref{sec:block}).
-
-\section{Sticky air and free surface}\label{sec:sticky}
-The Earth's surface can be treated by means of either the so-called sticky-air or a true free surface method, both of them implemented in the code. In the
-sticky-air method the surface is approximated with the introduction of a buoyant layer with a viscosity at least four orders of magnitude lower than the crust
-\citep{Schmeling2008,Crameri2012} and the interface between lithosphere and air is defined using a chain of passive markers that are advected as the Lagrangian
-markers. The correctness of the evolution of the markers chain is tested with the experiment of a 2D time-dependent Stokes sphere below a free surface and
-compared with results from ASPECT\footnote{\url{https://aspect.geodynamics.org/}} \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} (Section
-\ref{sec:sphere}). In the true free surface case the top boundary is assumed stress-free and velocities are not fixed. In this case topography variations are
-described by vertical deformations of the mesh that depend on the velocity field of the nodes that identify the free surface, while horizontal deformations are
-not taken into account. This procedure is known as the Arbitrary Lagrangian-Eulerian (ALE) method and its implementation follows the technique described in
-\citet{Thieulot2011}. However, although the implementation of a true free surface better reproduces laboratory experiment, extremely small time steps can be
-necessary to maintain stability \citep{Kaus2010a,Quinquis2011,Thieulot2014}. Therefore, the stability algorithm proposed by \citet{Kaus2010a} is implemented
-to avoid instabilities due to high density differences at the free surface when using too large time steps. The implementation of this algorithm are tested by
-performing the experiment described by \citet{Kaus2010a} (Section \ref{sec:stab}).
-
-The topography relaxation benchmark proposed by \citet{Crameri2012} is performed to verify that the code correctly recovers topography variations in case of
-both the sticky-air and the true free surface method (Section \ref{sec:crameri}). Finally, Section \ref{sec:subduction} show the results of the spontaneous
-subduction experiment described by \citet{Schmeling2008}.
-
-\section{Erosion and sedimentation}\label{sec:erosion}
-Surface processes at either the lithosphere-air interface or in correspondence of the free surface have been implemented by means of the software
-Fastscape\footnote{\url{https://fastscape.org/}} \citep{Braun2013,Cordonnier2019,Yuan2019a,Yuan2019b}. FastScapeLib is a set of routines that solve the stream
-power law (enriched by a sediment transport/deposition term) \citep{Yuan2019a}, hillslope diffusion and marine transport and deposition \citep{Yuan2019b}.
-The partial differential equation solved by FastScapeLib is:
-\begin{equation}
-\frac{\partial h}{ \partial t}=U-K_f A^{m_e}S^{n_e} + \frac{G}{A}\int_A \left( U-\frac{\partial h}{\partial t} \right) dA + K_d \bm{\nabla}^2 h 
-\end{equation}
-where $h$ is the topography, $U$ is the uplift, $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $S$ is the slope, $m_e$ and $n_e$ are
-parameters that should be chosen to constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}, $G$ is the dimensionless
-deposition/transport coefficient for the enriched stream power law and $K_d$ is the effective diffusivity. Terms $K_f A^{m_e}S^{n_e}$ and $K_d \bm{\nabla}^2 h$
-consider long-range processes, described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, and short-range processes,
-proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}.
-
-In case of sticky air, slope is determined by means of the markers chain, which is vertically corrected considering erosion and sedimentation rates. After the
-correction, continental markers above the markers chain are transformed in air markers and air markers below the markers chain become sediments. Similarly,
-slope is determined considering the nodes on the top boundary in case of true free surface and they are vertically corrected as for the markers chain. In this
-case, continental markers above the top boundary are simply deleted.
-%Surface processes at the lithosphere-air interface are implemented taking into account fluvial erosion (long-range processes), hillslope processes (short-range
-%processes) and failure of slopes exceeding a threshold angle $\theta$ \citep{Burov1997,Tucker1999,Sacek2012,Croissant2014,Ueda2015,Beucher2020}. Erosion rates
-%$\dot{E}_{lr}$ due to long-range processes can be described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{lr}=K_f A^{m_e} \bm{\nabla} h^{n_e}
-%\end{equation}
-%where $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $\bm{\nabla} h$ is the slope and $m_e$ and $n_e$ are parameters chosen in to
-%constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}. Erosion rates $\dot{E}_{sr}$ due to short-range processes are
-%implemented by assuming that transport is linearly proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{sr}=K_d \bm{\nabla}^2 h
-%\end{equation}
-%where $K_d$ is the effective diffusivity. In addition, the slope is modified if it still exceeds a threshold angle after long- and short-range processes
-%\citep{Ueda2015}.
-
-%Offshore sedimentation processes are implemented similarly to hillslope processes but with a depth-dependent coefficient \citep{Sacek2012,Beucher2020}.
-%The sedimentation rate $\dot{S}$ is calculated as
-%\begin{equation}
-%\dot{S}=K_s \bm{\nabla}^2 h
-%\end{equation}
-%with
-%\begin{equation}
-%K_s=K^*_s \exp\left(\frac{h-h_{sl}}{h_0}\right)
-%\end{equation}
-%where $K^*_s$ is a constant, $h_{sl}$ is the sea level and $h_0$ is a characteristic length. In case of sticky air, slope is determined by means of the markers
-%chain, which is vertically corrected considering erosion and sedimentation rates. After the correction, continental markers above the markers chain are
-%transformed in air markers and air markers below the markers chain become sediments. Similarly, slope is determined considering the nodes on the top boundary
-%in case of true free surface and they are vertically corrected as for the markers chain. In this case, continental markers above the top boundary are simply
-%deleted.
-
-\section{Non-linear rheologies}\label{sec:nonlinear}
-Non-linear rheologies are implemented combining viscous creep (dislocation and diffusion) and plastic yielding. For each marker, diffusion and dislocation
-viscosity ($\eta_{df|ds}$) can be determined as follows \citep{Karato1993,Warren2008b,Wang2016}:
-\begin{equation}\label{eq:viscous}
-\eta_{df|ds}=f_s \left(\frac{d^m}{A}\right)^{\frac{1}{n}} \left (\frac{2^{\frac{1-n}{n}}}{3^{\frac{n+1}{2n}}} \right)
-I_2^{\frac{(1-n)}{n}}\exp\left(\frac{Q+pV}{nRT}\right)
-\end{equation}
-where $f_s$ is a scaling factor used to represent lithologies that are stronger or weaker than the base set, $d$ is the grain size, $m$ is the grain size exponent,
-$n$ is the stress exponent, $A$ is the uniaxial pre-exponential factor, $I_2$ is the square root of the second invariant of the strain rate tensor, $Q$ is the
-activation energy, $p$ is the pressure, $V$ is the activation volume, $R$ is the gas constant and $T$ is the temperature.
-Pressures and temperatures are determined by the interpolation of the nodal parameters, while strain rates are calculated by means of the
-derivative of the velocities on the nodes. In case of diffusion creep $n=1$ and $m>0$, in case of dislocation creep $n>1$ and $m=0$. The viscous creep $\eta_{cp}$
-is then calculated as the harmonic average between $\eta_{df}$ and $\eta_{ds}$:
-\begin{equation}\label{eq:visco_creep}
-\eta_{cp}=\left(\frac{1}{\eta_{df}}+\frac{1}{\eta_{ds}}\right)^{-1}
-\end{equation}
-The implementation of the non-linear viscous creep viscosity depending on the strain rate is tested by means of the slab detachment benchmark
-(Section \ref{sec:slab}).
-
-Plastic yielding is implemented rescaling $\eta_{cp}$ in order to limit the stress below the yield stress $\sigma_y$ \citep{Thieulot2008,Thieulot2014,Glerum2018},
-obtaining
-\begin{equation}\label{eq:plastic}
-\eta_{pl}=\frac{\sigma_y}{2I_2}
-\end{equation}
-where $\eta_{pl}$ is the plastic viscosity and the yield stress is determined following the Drucker-Prager criterion, such as
-\begin{equation}\label{eq:yield}
-\sigma_y=C\;\cos(\phi)+p\;\sin(\phi)
-\end{equation}
-where $C$ is the cohesion and $\phi$ is the internal friction angle. In case of negative pressure it is imposed equal to 0, so that negative yield stress are
-excluded. The correctness of the non-linear solution in case of plastic viscosity with variable internal friction angle is verified performing indenter and
-brick experiments (Sections \ref{sec:indenter} and \ref{sec:brick}, respectively).
-
-Strain softening is taken into account for both viscous creep and plastic viscosity \citep{Huismans2003,Babeyko2005,Huismans2005,Sobolev2005,Warren2008a} by
-means of the accumulated strain $\epsilon$ memorised by each marker. Softening in the viscous creep determines a linear decrease of $\eta_{vc}$ by means of a
-viscous strain softening factor $W_S$ that increases linearly from $W_{S_0}$ to $W_{S_{\infty}}$ for $\epsilon_{S_0}<\epsilon_S<\epsilon_{S_{\infty}}$ 
-\citep{Warren2008a}. This viscous softening can be related to strain-induced grain size reduction \citep{Warren2008a}. Differently, plastic softening is 
-simulated with a linear decrease of internal friction angle $\phi(\epsilon)$ and cohesion $C(\epsilon)$ in according to
-\begin{eqnarray}
-\label{eq:friction1}\phi(\epsilon)&=&\phi_0+(\phi_{\infty}-\phi_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}\\
-\label{eq:friction2}C(\epsilon)&=&C_0+(C_{\infty}-C_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}
-\end{eqnarray}
-where $\phi_0$, $C_0$ and $\phi_{\infty}$, $C_{\infty}$ are internal friction angle and cohesion for $\epsilon_0$ and $\epsilon_{\infty}$, respectively
-\citep{Huismans2003,Huismans2005,Warren2008a,Thieulot2014}. Plastic softening approximates deformation-induced softening of faults and brittle shear zones
-\citep{Warren2008a}.
-
-Plastic yielding and viscous creep are then combined to obtain a viscoplastic viscosity $\eta_{vp}$ as follows, assuming that they are independent processes
-\citep{Karato2008,Glerum2018}:
-\begin{equation}\label{eq:viscoplastic}
-\eta_{vp}=\min(\eta_{cp},\eta_{pl})
-\end{equation}
-Finally, effective viscosity $\eta_{eff}$ is capped by the minimum and the maximum viscosity ($\eta_{min}$ and $\eta_{max}$, respectively) to avoid extremely
-low or high viscosity \citep{Glerum2018} as follows
-\begin{equation}\label{eq:effective}
-\eta_{eff}=\min(\max(\eta_{vp},\eta_{min}),\eta_{max})
-\end{equation}
-Viscous creep and plastic yielding are non-linear rheologies because of their dependence on the velocity field through pressure and strain rates. Therefore,
-the solution is determined by means of Picard-type iterations, until convergence of the velocity field \citep{Glerum2018}. The convergence is verified at each
-iteration $i$ via the nonlinear residual $\bm{\mathcal{R}}^i$ that can be determined as
-\begin{equation}\label{eq:convergence}
-\bm{\mathcal{R}}^i= \mathbb{K} (\eta_{eff}(\dot{\bm{\epsilon}}^{i-1},p^{i-1}))\cdot \bm{v}^{i-1} - \bm{f}^i
-\end{equation}
-%\begin{equation}\label{eq:convergence}
-%\chi=1-\frac{\langle (\bm{v}^i-\langle\bm{v}^i\rangle)\cdot (\bm{v}^{i+1}-\langle\bm{v}^{i+1}\rangle) \rangle}{\sqrt{\sigma^i\sigma^{i+1}}}
-%\end{equation}
-%Iterations are performed until either convergence ($\chi<tol$) or maximal number of non-linear iterations ($it_{max}$) is reached.
-where $\mathbb{K}$ is the Stiffness matrix and $\bm{f}$ is the right hand side vector \citep{Spiegelman2016,Glerum2018}.  The $L_2$-norm is extracted from
-$\bm{\mathcal{R}}^i$ and it is normalised as follows
-\begin{equation}\label{eq:convergence_res}
-\frac{||\bm{\mathcal{R}}^i||_2}{||\bm{\mathcal{R}}^0||_2}
-\end{equation}
-where $||\bm{\mathcal{R}}^0||_2$ is the $L_2$-norm of the first nonlinear iteration.
-Since often the normalised nonlinear residual drop very quickly over the first 2-3 iterations, also the $L_2$-norm of velocity and pressure residuals are
-calculated as
-\begin{equation}\label{eq:convergence_vel}
-\frac{||\Delta\bm{v}||_2}{||\bm{v}||_2}=\frac{||\bm{v}_i-\bm{v}_{i-1}||_2}{||\bm{v}_{i}||_2}
-\end{equation}
-and
-\begin{equation}\label{eq:convergence_press}
-  \frac{||\Delta p||_2}{||p||_2}=\frac{||p_i-p_{i-1}||_2}{||p_{i}||_2}
-\end{equation}
-respectively. Iterations are performed until either maximal number of non-linear iterations ($it_{max}$) is reached or all the normalised nonlinear residual, 
-the velocity residual and the pressure residual converge under a defined tolerance $tol$.
-
-\section{The energy equation}\label{sec:en_eq}
-The stabilisation of the advection term of the energy equation (Eq. \ref{eq:energy}) needed to avoid possible oscillations in the thermal solution in the case
-when advection dominates over diffusion, is implemented by means of a streamline-upwind Petrov–Galerkin (SUPG) method, for which the advection term is modified
-follows the discussion in \citet{Thieulot2011} and \citet{Thieulot2014} (see Section \ref{sec:advection} for benchmark). The total internal heat production
-$H_{tot}$ in Eq. \ref{eq:energy} includes radiogenic heating $H_r$, shear heating $H_s$ and adiabatic heating $H_a$, where
-\begin{equation}\label{eq:shear}
-H_s=2\eta \dot{\bm{\epsilon}} : \dot{\bm{\epsilon}} = 2\eta(\dot{\epsilon}_{xx}^2+\dot{\epsilon}_{yy}^2+2\dot{\epsilon}_{xy}^2)
-\end{equation}
-and
-\begin{equation}\label{eq:adiabatic}
-H_a=T \alpha \frac{Dp}{Dt} \approx -\alpha T \rho g_y v
-\end{equation}
-The correctness of the implementation and the computation of shear and adiabatic heating is verified performing an exercise with an analytical solution
-(Section \ref{sec:simple_shear}) and Exercise 9.4 in \citet{Gerya2010b} (Section \ref{sec:shear}). Furthermore, mantle convection and thin layer entrainment
-benchmarks (Sections \ref{sec:mantle} and \ref{sec:thin}, respectively) are performed to verify that the code solve correctly Eqs. \ref{eq:momentum_penalty}
-and \ref{eq:energy} in case of a temperature field characterised by an initial perturbation and with a temperature-dependent density. At last, the viscoplastic
-mantle convection benchmark is performed to check also the viscoplasticity in case of a initial perturbed temperature field (Section \ref{sec:visco_mantle}).
-
-\section{Phase transitions and hydration}\label{sec:phase}
-Variations of effective density, specific heat and coefficient of thermal expansion during the evolution of crustal and mantle materials at different
-pressure-temperature ($p$-$T$) conditions are computed by means of version 6.8.6 of Perple\_X software package \citep{Connolly2005}, similarly to the
-implementation described in \citet{Marotta2020}. By default, mineral assemblages and properties of each lithology (oceanic crust, lower and upper continental
-crust, sediments and dry or hydrated mantle) are calculated for temperatures from \SIrange{330}{1600}{\kelvin}, with increments of approximately \SI{4}{\kelvin},
-and pressures from \SIrange{0.1}{30}{\giga\pascal}, with increments of approximately \SI{0.03}{\giga\pascal}, for a total of almost 100000 points.
-
-The introduction of phase transitions produces two main effects that must be taken into account in numerical models: variations in density and release/absorption
-of latent heat \citep{Tackley2010}. Effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into
-account considering effective coefficient of thermal expansion \citep{Christensen1985,Zhong2015}. Similarly, effects of latent heat in the energy equation
-(eq \ref{eq:energy}) are taken into account by considering effective specific heat and coefficient of thermal expansion
-\citep{Christensen1985,Tackley2010,Zhong2015}. Effective density, specific heat and coefficient of thermal expansion are included in Perple\_X files and are
-assumed by each Lagrangian marker in according to its $p$-$T$ conditions.
-
-Hydration processes related to dehydration of subducting lithosphere \citep{Schmidt1998,Liu2007,Faccenda2009,Faccenda2010,Faccenda2014,Rosas2016} strongly
-influence the thermo-mechanics inside the mantle wedge, mainly because of both weakening effects on the mantle rheology and density variations in case of
-mantle serpentinization \citep{Gerya2002,Honda2003,Arcay2005,Roda2010,Regorda2017}. In the case that hydration is switched on, the amount of bound and free
-water is memorised by each marker, following the implementation of \citet{Quinquis2014}; hydration and dehydration processes are related to the amount of bound
-water of each marker and to the maximum amount of water it can transport, i.e. if the amount of bound water exceeds the maximum amount of water, the marker
-dehydrates and releases free water that can hydrates under-saturated markers. Maximum water content of each marker is determined as function of lithology and
-$p$-$T$ conditions and it is calculated using Perple\_X, in the same way than effective density, specific heat and coefficient of thermal expansion. Bound water
-is advected together with the markers, neglecting the effect of bound water diffusion, while free water simply migrates vertically and is not coupled to the
-solid-phase flow of the mantle wedge \citep{Arcay2005,Quinquis2014}. The correctness of the migration scheme of the free water and its absorption into bound
-water by under-saturated markers are verified performing the experiment described in \citet{Quinquis2014} (Sec. \ref{sec:quinquis}).
-
-The mantle viscosity weakening related to the amount of water has been extensively studied \citep{Chopra1981,Hirth2003} and it is implemented in according to
-\citet{Arcay2005} and \citet{Horiuchi2016}, as follows
-\begin{equation}\label{eq:wet_visc}
-\eta_{wet}=\eta_{dry}\left(\left[1-\frac{1}{f_v}\right]\exp\left(-\frac{[OH^-]}{[OH^-]_0}\right)+\frac{1}{f_v}\right)
-\end{equation}
-where $f_v$ is the viscosity reduction factor between dry and wet conditions, $[OH^-]$ is the water content and $[OH^-]_0$ is a reference water content set to
-620 ppm (0.062 wt.\%) \citep{Arcay2005}. By default, $f_v$ is set to 100, in according to \citet{Arcay2005} and \citet{Horiuchi2016}.
-
-\section{Mantle melting}\label{sec:melt}
-Melt fraction in the mantle depends on temperature ($T$, in \textdegree C), pressure ($p$, in \si{\giga\pascal}) and the water content in the melt ($X_{H_2O}$, in wt.\%).
-The determination of the melt fraction $M$ is implemented as explained by \citet{Katz2003}, according to the successive modification by \citet{Langmuir2006} and
-\citet{Kelley2010}, as follows
-\begin{eqnarray}\label{eq:melt_fr}
-M(p,T,X_{H_2O})&=&-T+T_{sol}+(x_t\cdot ln(p+y_t))\cdot X_{H_2O}-\nonumber\\
-&&+K_m\left(\frac{X_{H_2O}^{bulk}}{D_{H_2O}(1-X_{H_2O})+X_{H_2O}}\right)^{\gamma}
-\end{eqnarray}
-where $T_{sol}$ is the temperature of the dry solidus, $D_{H_2O}$ is the partition coefficient and $x_t\cdot log(p+y_t)$ indicates the pressure dependence of
-the melting curve \citep{Kelley2010}. $x_t$, $y_t$, $K_m$ and $\gamma$ are parameters chosen in according to \citet{Kelley2010}.
-However, $X_{H_2O}$ depends on the melt fraction as
-\begin{equation}\label{eq:melt_wt}
-X_{H_2O}(M)=\frac{X_{H_2O}^{bulk}}{D_{H_2O}+M(1-D_{H_2O})}
-\end{equation}
-and a numerical solution can be found using a root-finder method \citep{Katz2003,Wang2016}. Furthermore, the content of water in the melt is limited by the
-pressure-dependent saturation concentration of water in the melt, determined as
-\begin{equation}\label{eq:melt_max}
-X_{H_2O}^{sat}=\chi_1 p^{\lambda_m}+\chi_2 p
-\end{equation}
-where $\chi_1$, $\chi_2$ and $\lambda_m$ are parameters chosen according to \citet{Katz2003}, \citet{Langmuir2006} and \citet{Kelley2010}. The correctness in
-the determination of melt fractions and water contents in the melt are tested comparing model results with experimental curves obtained by \citet{Katz2003}
-(Sec. \ref{sec:katz}).
-
-In case of melting, viscous creep viscosity ($\eta_{vp}$) is modified according to \citet{Wang2016}, as
-\begin{equation}\label{eq:melt_visc}
-\eta_{melt}=\eta_{vp}\cdot \exp(\alpha_m M)
-\end{equation}
-where $\alpha_m$ are melt fraction factors for dislocation and diffusion creep, chosen in according to \citet{Wang2016}.
-
-Effective density ($\rho_{eff}$) for partially molten rocks is calculated as
-\begin{equation}\label{eq:melt_dens}
-\rho_{eff}=\rho_s(1-M)+\rho_m M
-\end{equation}
-where $\rho_s$ and $\rho_m$ are the densities of the solid and the molten rock, respectively \citep{Gerya2004b,Gerya2007,Wang2016}. The density of the
-solid rock is extracted by Perple\_X, while the density of the molten portion is calculated as
-\begin{equation}\label{eq:melt_dens_M}
-\rho_m(p,T)=\rho_0[1-\alpha(T-T_0)][1+\beta(p-p_0)]
-\end{equation}
-where $\rho_0$ is the density at the reference temperature ($T_0$) and pressure($p_0$), and $\alpha$ and $\beta$ are the thermal expansion and compressibility
-coefficients, respectively \citep{Gerya2004b,Gerya2007,Wang2016}.
-
-As for the phase transitions, effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into account
-considering effective coefficient of thermal expansion. Similarly, latent heat due to melting/crystallisation can be implicitly included in the energy equation
-(Eq. \ref{eq:energy}) by considering effective specific heat ($C_{p_{eff}}$) and thermal expansion ($\alpha_{eff}$) of the partially molten rocks, in according to
-\begin{eqnarray}
-\label{eq:cp_effective}C_{p_{eff}}&=&C_p+H_L\left(\frac{\partial M}{\partial T}\right)_p \\
-\label{eq:alpha_effective}\alpha_{eff}&=&\alpha+\rho\frac{H_L}{T}\left(\frac{\partial M}{\partial p}\right)_T
-\end{eqnarray}
-where $C_p$ and $\alpha$ are the specific heat and the coefficient of thermal expansion of the solid rock, respectively, and $H_L$ is the latent heat of the
-molten rock \citep{Gerya2004b,Gerya2007,Tackley2010}.
-
-\section{Benchmarks}\label{sec:benchmark}
-\subsection{Error measurements}\label{sec:error}
-In order to determine the accuracy of velocity and pressure field of the benchmarks, the $L_2$-norm is computed by numerical integration on the Gauss-Legendre
-quadrature points. $L_2$-norm for pressure and velocity errors can be evaluated as
-\begin{eqnarray}
-\textrm{err}_p&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|p^n_i(r_q)-p^a_i(r_q)|^2\right)w_q|J_q|}\nonumber \\
-\textrm{err}_v&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|u^n_i(r_q)-u^a_i(r_q)|^2+|v^n_i(r_q)-v^a_i(r_q)|^2\right)w_q|J_q|}\nonumber
-\end{eqnarray}
-respectively, where $n_e$ is the number of elements, $n_q$ is the number of quadrature points per element, $p^n_i(r_q)$ and $p^a_i(r_q)$ are the numerical and
-analytical pressure, respectively, in each quadrature point $q$, $w_q$  and $J_q$ are the weight and the Jacobian at the quadrature point
-$q$, $u^n_i(r_q)$, $v^n_i(r_q)$, $u^a_i(r_q)$ and $v^a_i(r_q)$ are the numerical and analytical velocities, respectively, in each quadrature point $q$.
-
-Other quantities used as comparison with original benchmarks are the root-mean-square velocity over the whole domain and over the surface
-\begin{eqnarray}
-v_{\textrm{rms}}&=&\sqrt{\int_{\Omega}|\bm{v}|^2 d\Omega}\nonumber \\
-v_{\textrm{rms}}^{top}&=&\sqrt{\int_0^1 u^2 \bigg|_{y=1} dx}\nonumber
-\end{eqnarray}
-respectively, top and bottom Nusselt numbers
-\[\textrm{Nu}^{top/bottom}=-\int_0^1\frac{\partial T}{\partial y} \bigg|_{y=1/y=0} dx\]
-the average rate of work done against gravity
-\[\left\langle W \right\rangle=\int_{\Omega}T u_y d\Omega\]
-the average rate of viscous dissipation
-\[\left\langle \Phi \right\rangle=\int_{\Omega}\tau_{ij}\dot{\epsilon}_{ij} d\Omega\]
-and the percentage error between $\left\langle W \right\rangle$ and $\left\langle \Phi \right\rangle/Ra$
-\[\delta=\frac{\left|\left\langle W \right\rangle -\frac{\left\langle \Phi \right\rangle}{Ra} \right|}{\max \left(\left\langle W \right\rangle,
-\frac{\left\langle \Phi \right\rangle}{Ra}\right)}\times100\]
-
-\subsection{Stokes flow}\label{sec:stokes}
-The problem consists of determining velocity field $(u,v)$ and pressure $p$ in case of a manufactured solution with prescribed body forces such as
-\begin{eqnarray}
-b_1&=&(12-24y)x^4+(-24+48y)x^3+(-48y+72y^2-48y^3+12)x^2+\nonumber \\
-&&+(-2+24y-72y^2+48y^3)x+1-4y+12y^2-8y^3\nonumber \\
-b_2&=&(8-48y+48y^2)x^3+(-12+72y-72y^2)x^2+(4-24y+48y^2-48y^3+\nonumber \\
-&&+24y^4)x-12y^2+24y^3-12y^4\nonumber
-\end{eqnarray}
-for which the exact solution is:
-\begin{eqnarray}
-u(x,y)&=&x^2(1-x)^2(2y-6y^2+4y^3)\nonumber \\
-v(x,y)&=&-y^2(1-y)^2(2x-6x^2+4x^3)\nonumber \\
-p(x,y)&=&x(1-x)-\frac{1}{6}\nonumber
-\end{eqnarray}
-The domain is a unit square a constant viscosity ($\eta=1$) and the penalty parameter is set to $\lambda=10^7$. Velocity boundary conditions are set to no slip
-$(\bm{v}=\bm{0})$ on all boundaries. The problem is performed for different grid resolution between \num{8x8} and \num{1024x1024} elements.
-
-The errors between the analytical solution and the numerical prediction of the velocity and the pressure field are calculated as in Section \ref{sec:error}.
-Fig. \ref{fig:errors}a shows that both velocity and pressure field converge to the exact solution with the decrease of the element size, following the
-theoretical convergence rate. The convergence of pressure, in contrast with the observation by \citet{Donea2003} for $Q_1 \times P_0$ elements, indicate the
-effectiveness of the smoothing procedure. Smoothed pressure field for grid resolution of \num{128x128} elements is shown in Fig. \ref{fig:errors}b (continuous
-red line), in comparison with the analytical solution (dashed blue line). Solve times and memory usage needed to generate the solution are shown in
-Fig. \ref{fig:MUMPS}. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Solver/Stokes_Flow}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/errors.pdf}
-\caption{Velocity and pressure error for the Stokes flow experiment between generated and analytical solution as a function of element size (panel a),
-and comparison between smoothed pressure and analytical solution as function of $x$ coordinate for a grid resolution of \num{128x128} elements (panel b).}
-\label{fig:errors}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/MUMPS.pdf}
-\caption{Panel a) Analysis, factorisation, solution and total solve times (panel a) and factorisation memory usage (panel b) as a function of the total number
-of degrees of freedom for the Stokes flow experiment.}
-\label{fig:MUMPS}
-\end{figure}
-
-\subsection{Markers advection}\label{sec:runge}
-The 2nd-order Runge-Kutta advection scheme in space is tested by means of the Zalesak disk test \citep{Zalesak1979,Thieulot2014}. The benchmark is performed
-in a unit square domain with a grid resolution of \num{32x32} elements and 1000000 markers and values of Courant number between 0.25 and 1.
-At $t=0$, the disk is centred at position $(0.5;0.75)$ with a radius $R=0.15$ and has a vertical fissure 0.05 wide and 0.2 high (orange in Fig. \ref{fig:runge})
-and with the velocity field prescribed in the entire domain as
-\begin{eqnarray}
-u(x,y)&=&2\pi\left(y-\frac{L_x}{2}\right)\nonumber \\
-v(x,y)&=&-2\pi\left(x-\frac{L_x}{2}\right)\nonumber
-\end{eqnarray}
-the markers are back to their initial location after a $2\pi$ rotation ($t=1$). The clockwise rotation of the disk at $t=0.25$, 0.5 and 0.75 
-is shown in Fig. \ref{fig:runge} (red, blue and green, respectively). The distance from the centre of 1 marker is calculated throughout the rotation to
-evaluate the error for different values of Courant number. As expected, the error increase with the increasing of the time step, as shown in 
-Fig. \ref{fig:runge_err}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/Zalesak_Disk}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=200px]{./Figures/Zalesak.pdf}
-\caption{Position of the Zalesak disk at $t=0$, 0.25, 0.5 and 0.75 (orange, red, blue and green, respectively) for the case with Courant number of 1.}
-\label{fig:runge}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/zal_marker.pdf}
-\caption{Distance from the centre as function of time for values of Courant number of  of 0.25, 0.3, 0.5, 0.75 and 1 (green, orange, blue, black and red,
-respectively). R indicates the distance from the centre considering a perfect circle trajectory.}
-\label{fig:runge_err}
-\end{figure}
-
-\subsection{Conservative Velocity Interpolation (CVI)}\label{sec:CVI}
-The CVI correction is checked by means of the Stokes flow experiment by \citet{Donea2003}, as presented in \ref{sec:stokes}. The advection of Lagrangian
-markers is performed using either a 2nd-order or a 4th-order Runge-Kutta scheme with CFL=0.5 and an initial random distribution of 25 markers per element.
-Fig. \ref{fig:CVI} shows the minimum and maximum values of markers per element throughout the experiment with and without the CVI correction and using
-different Runge-Kutta schemes. The introduction of the CVI correction allows the
-tests to run for times longer than $t=2000$ without show empty elements, while tests without the CVI correction stop at approximately $t=600$, when at least
-one element goes empty. Fig. \ref{fig:CVI_time} shows the different number of markers per element and the markers distribution at $t=600$ obtained without and
-with the CVI correction (panels b and c, respectively) in case of a 2nd-order Runge-Kutta scheme. All data can be found at
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/CVI}.
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/CVI.pdf}
-\caption{Maximum and minimum number of markers per element throughout the experiment, with (green and orange lines) or without (purple and blue lines)
-the CVI correction and using either a 2nd-order (blue and orange lines) or a 4th-order (purple and green lines) Runge-Kutta scheme.}
-\label{fig:CVI}
-\end{figure}
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/CVI_time.pdf}
-\caption{Number of markers per element and markers distribution through time using 2nd.order Runge-Kutta scheme.
-Panels a: number of markers per element (panel a1) and distribution of the markers (panel a2) after the first
-iteration; panels b and c: comparison between number of markers per element (panels b1 and c1) and markers distribution (panels b2 and c2) without and with
-the CVI correction at $t=600$; panels d: number of markers per element (panel d1) and markers distribution (panel d2) in case of CVI correction at $t=2000$.}
-\label{fig:CVI_time}
-\end{figure}
-
-\subsection{Poiseuille flow}\label{sec:poiseuille}
-The domain is a rectangle with $L_x=2$ and $L_y=1$ and constant density and viscosity ($\rho=1$ and $\eta=1$), gravity acceleration $\bm{g}=\bm{0}$ and penalty
-parameter $\lambda=10^8$. The grid is composed by \num{40x20} elements. Velocity boundary conditions are set to no slip $(\bm{v}=\bm{0})$ at the top and the
-bottom, and a parabolic profile is imposed on the sides, with $u=y(1-y)$ and $v=0$. The analytical solution is then given by:
-\begin{eqnarray}
-u(x,y)&=&y(1-y)\nonumber \\
-v(x,y,)&=&0\nonumber \\
-p(x,y)&=&2\eta\left(\frac{L_x}{2}-x\right)\nonumber
-\end{eqnarray}
-
-The velocity field predicted by the model follows the expected parabolic profile (Fig. \ref{fig:poiseuille}a and Fig. \ref{fig:poi_plot}a) and, in case of the
-classic penalty method with no iterations, the pressure is clearly related to the divergence by means of the penalty parameter (Fig. \ref{fig:poiseuille}b and
-c and Fig. \ref{fig:poi_plot}b). Fig. \ref{fig:poiseuille}d shows that one Uzawa iteration is sufficient to bring the divergence down to \num{e-15}, with no
-correlation with the pressure field. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Poiseuille%20Flow}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Poiseuille.pdf}
-\caption{Velocity field (panel a), pressure (panel b) and divergence velocity of a Poiseuille flow in case of the classic penalty method (no iterations) and
-after one Uzawa iteration (panel c and d, respectively).}
-\label{fig:poiseuille}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Analytical.pdf}
-\caption{Velocity (panel a) and pressure (panel b) field predicted by the model for a Poiseuille flow with respect with their analytical solutions. The
-velocity field is plotted as function of the vertical coordinate in $L_x/2$ and the pressure is plotted as function of the x coordinate in $L_y/2$.}
-\label{fig:poi_plot}
-\end{figure}
-
-\subsection{Instantaneous 2D sphere}\label{sec:ist_sphere}
-The domain is a unit square with gravity $\bm{g}=(0,-1)$. The fluid has constant density and viscosity ($\rho_f=1$ and $\eta_f=1$). The sphere is in the middle
-of the domain with a radius R = 0.123456798 and has constant density and viscosity ($\rho_s=10^{-2}+\rho_f$ and $\eta_s=10^3 \cdot \eta_f$).
-We distinguish three different types of velocity boundary conditions:
-\begin{enumerate}
-\item FS: free slip conditions on all sides.
-\item NS: no slip conditions on all sides.
-\item OT: free slip conditions on the sides and the bottom, and open boundary at the top.
-\end{enumerate}
-All three cases are performed for different grid resolutions, between \num{16x16} and \num{512x512} elements with 50 randomly distributed markers per element,
-and different average schemes for the viscosity (harmonic, geometric and arithmetic). The velocity in the centre of the sphere $\bm{v}(0.5;0.5)$), minimum and
-maximum velocities ($u_{min}$, $u_{max}$, $v_{min}$, $v_{max}$) and pressures ($p_{min}$, $p_{max}$) on the entire domain, $v_{\textrm{rms}}$ and average pressure on the
-entire domain ($p_{avg}$) are compared with solutions generated by ASPECT \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and all results
-can be found at \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere2D} and 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Instantaneous_2D_Sphere}. Results in terms of the $v_{\textrm{rms}}$ for NS model 
-are also shown in Fig. \ref{fig:inst_sphere}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_NS.pdf}
-\caption{$v_{\textrm{rms}}$ as function of element size for the instantaneous 2D sphere experiment in case of no slip boundary conditions and with different average
-schemes for the viscosity. Results are compared with results obtained by other numerical codes.}
-\label{fig:inst_sphere}
-\end{figure}
-
-\subsection{Rayleigh-Taylor instability}\label{sec:rayleigh}
-This problem has been originally presented by \citet{vanKeken1997} and here it is performed as the isoviscous case (Case 1a). The domain has $L_x=0.9142$ and
-$L_y=1$ and gravity $\bm{g}=(0,-1)$. Two fluids with same constant viscosities ($\eta_1=\eta_2=1$) and different densities ($\rho_1=1000$ and $\rho_2=1010$),
-with the lighter fluid at the bottom. The initial interface between the fluids is given by $y(x)=0.2+0.02 \cos \left(\frac{\pi x}{L_x}\right)$. The experiment
-is performed with different grid sizes (\num{50x50}, \num{80x80}, \num{100x100} and \num{256x256}). A total of 1960000 markers are randomly distributed at the
-beginning of the simulation. Velocity boundary conditions are set to no slip at the top and the bottom, and to free slip at the sides of the domain.
-
-$v_{\textrm{rms}}$ as function of time is reported in Fig. \ref{fig:RT}, matching well with results shown by \citet{vanKeken1997}, \citet{Tackley2003} and
-\citet{Thieulot2014}. Fig. \ref{fig:rayleigh} shows the evolution of the experiment at different time steps. All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Rayleigh_Taylor_experiment/ISOVISCOUS}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/RT.pdf}
-\caption{$v_{\textrm{rms}}$ of the Rayleigh-Taylor experiment as function of time for different resolution of the grid.}
-\label{fig:RT}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=300px]{./Figures/Rayleigh.pdf}
-\caption{Evolution of the Rayleigh-Taylor experiment for a grid of \num{256x256} elements at $t=0, 500, 1000$ and $2000$ (panels a, b, c and d, respectively).}
-\label{fig:rayleigh}
-\end{figure}
-
-\subsection{Falling blocks}\label{sec:block}
-This benchmark is described as proposed by \citet{Gerya2003a}, \citet{Gerya2010b} and \citet{Thieulot2011}. The domain is square with $L_x=L_y=$ \SI{500}{\km}
-and the grid is composed by \num{50x50} elements with 25 markers in each element. The block is initially centred at ($x=$ \SI{250}{\km}; $y=$ \SI{400}{\km})
-and has a size of \num{100x100} km (Fig. \ref{fig:block}a). The fluid surrounding the block has $\eta_f=$ \SI{e21}{\pascal\per\s} and $\rho_f=$
-\SI{3200}{\kg\per\cubic\m}. The benchmark tests with different viscosities of the block, with $\eta_b$ from \SIrange{e15}{e27}{\pascal\per\s}. In each
-experiment also the density $\rho_b$ of the block varies from \SIrange{3220}{9900}{\kg\per\cubic\m}. Velocity boundary conditions are set to free slip
-conditions on all sides of the domain. Density distributions at $t=$ \SI{20}{\mega\year} for $\rho_b=$ \SI{3300}{\kg\per\cubic\m} and different $\eta_b$
-are plotted in Fig. \ref{fig:block}b-f, showing that the code correctly preserve the block geometry in case of large viscosity contrast, with a block stiffer
-than the surrounding fluid (Fig. \ref{fig:block}f).
-
-The velocity in the centre of the falling block at $t=0$ is measured for all experiments and, since the velocity should increase with density contrast, the
-quantity $\bm{v}/(\rho_b-\rho_f)$ is plotted as function of the viscosity contrast. All results perfectly match with those from \citet{Gerya2010b} and they
-line up on a single curve, demonstrating that the code can correctly deal with large viscosity and density contrasts (Fig. \ref{fig:falling}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Falling%20blocks}.
-
-\begin{figure}
-\noindent\includegraphics[width=300px]{./Figures/Block.pdf}
-\centering
-\caption{Density evolution of the falling block experiment at $t=$ \SI{0}{\mega\year} (panel a) and $t=$ \SI{20}{\mega\year} for different viscosities of the
-block (panels b-f).}
-\label{fig:block}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Falling.pdf}
-\caption{Initial velocity relative to the density contrast at the centre of the falling block as function of the viscosity contrast between the block and the
-surrounding fluid.}
-\label{fig:falling}
-\end{figure}
-
-\subsection{2D time-dependent Stokes sphere with deformable free surface}\label{sec:sphere}
-This experiment is performed in a unit square domain with the gravity acceleration fixed to $g_y=-1$. The sphere has $\rho_s=2$ and $\eta_s=10^3$ and is
-initially centred at ($0.5;0.6$) with radius $R=0.123456789$. The fluid surrounding the sphere has $\rho_f=1$ and $\eta_f=1$ and occupies the domain for
-$y\leq 0.75$, while for $y>0.75$ the air has $\rho_a=0$ and $\eta_a=10^{-3}$. Velocity boundary conditions are set to free slip on all sides. The Courant
-number is set to 0.25. The experiments run for 200 s using grid resolutions from \num{150x150} to \num{512x512} elements, with 25 randomly distributed markers
-per element, and for different average schemes for the viscosity. The interface between the fluid and the air is tracked by means of the markers chain. Results
-in terms of velocity and pressure fields and topography variations are compared with results obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and they can be found at
-\url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere_fs2D} and
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Time_dependent_sphere}. Results in terms of the $v_{\textrm{rms}}$ in case of 
-an arithmetic average are also shown in Fig. \ref{fig:inst_spherefs}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_arithm.pdf}
-\caption{$v_{\textrm{rms}}$ for different grid resolutions as function of time for the 2D time-dependent Stokes sphere experiment in case of an arithmetic average.
-Results are compared with results obtained by ASPECT.}
-\label{fig:inst_spherefs}
-\end{figure}
-
-\subsection{Free surface stabilisation}\label{sec:stab}
-The benchmark is performed as discussed in \citet{Kaus2010a} and \citet{Thieulot2014}. The domain is square with $L_x=L_y=$ \SI{500}{\km} and a grid resolution
-of \num{200x200} elements, each of them containing 16 markers. A buoyant fluid with $\rho_1=$ \SI{3200}{\kg\per\cubic\metre} and $\eta_1=$ \SI{e20}{\pascal\s}
-is overlain by a denser fluid with $\rho_2=$ \SI{3300}{\kg\per\cubic\metre} and $\eta_2=$ \SI{e21}{\pascal\s}. The initial interface between the fluids has an
-initial sinusoidal shape of 5 km amplitude. Velocity boundary conditions are set to free slip conditions at the sides and no slip at the bottom, while the top
-is free surface. The experiment is performed with various fixed time steps, without the use of the Courant number. The vertical position of the free surface at
-$x=L_x$ is tracked for each simulation, with or without the stabilisation algorithm.
-
-The results show an instability (drunken sailor effect) increasing the time step in case that the stabilisation algorithm is not activated
-($dt=$ \SI{4200}{\year}, green line in Fig. \ref{fig:stab}), as already observed by \citet{Kaus2010a} and \citet{Thieulot2014}. Activating the algorithm the
-instability is fixed and the simulation remains stable using time step up to \SI{50000}{\year} (red line in Fig. \ref{fig:stab}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Stabilisation_algorithm}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Kaus_random.pdf}
-\caption{Evolution of the y coordinate in $x=L_x$ as function of time for different time steps, with and without the stabilisation algorithm.}
-\label{fig:stab}
-\end{figure}
-
-\subsection{Topography relaxation}\label{sec:crameri}
-This experiment is performed as in \citet{Crameri2012} using free surface and sticky air (Case 1). The domain is rectangular with $L_x=$ \SI{2800}{\km} and
-$L_y=$ \SI{700}{\km} in case of free surface and $L_y=$ \SI{800}{\km} in case of sticky air ($h_{st}=$ \SI{100}{\km}). The grid is composed by \num{256x64}
-elements for the free surface \num{560x320} elements for the sticky air. For both cases, there is a mantle of 600 km thickness with $\eta_m=$ \SI{e21}{\pascal\s}
-overlain by a highly viscous cosine-shaped lithosphere with thickness between 93 and 107 km and $\eta_l=$ \SI{e23}{\pascal\s}. Both mediums have $\rho=$
-\SI{3300}{\kg\per\cubic\metre}. In case of sticky air experiment, the air has $\rho_a=$ \SI{0}{\kg\per\cubic\metre} and $\eta_a=$ \SI{e18}{\pascal\s}. The
-gravity acceleration is set to $g_y=$ \SI{-10}{\metre\per\square\s}. Velocity boundary conditions are set to no slip at the bottom and free slip at the sides
-of the domain. In case of sticky air, velocities on top are set to free slip conditions.
-
-The maximum topography as function of time ($h(t)$) can be analytically derived in according to \[h(t)=h_0 \exp(\gamma t)\]
-where $h_0=7$ km is the initial topography, $\gamma=\num{-0.2139e-11}$ is the characteristic relaxation rate and $t$ is the time (black line in Fig.
-\ref{fig:crameri}). Maximum topography at the characteristic relaxation time $t_{rlx}=$ \SI{14.825}{\kilo\year} can be found to be $h_{rlx}=$ \SI{2576}{\m}.
-The results show that a viscosity of the air of \SI{e18}{\pascal\s} is needed in case of $h_{st}=$ \SI{100}{\km} to correctly describe the topography relaxation
-for this problem, as pointed out by \citet{Crameri2012} (purple line in Fig. \ref{fig:crameri}). The case with a true free surface well follows the analytical
-solution (orange line in Fig. \ref{fig:crameri}). In particular, at the relaxation time in case of the true free surface predicts a maximum topography of
-\SI{2570}{\m}, with an error of \SI{6}{\m} with respect to the analytical solution. Results are also compared with results obtained with a free surface in
-UNDERWORLD (\num{256x64} elements) and SULEC (\num{401x201} elements) (with courtesy from \citealp{Crameri2012}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Topography_relaxation}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Crameri.pdf}
-\caption{Maximum topography as function of time for the Crameri benchmark (coloured lines) in comparison with the analytical solution (black line) and results
-obtained with UNDERWORLD (yellow line) and SULEC (blue line).}
-\label{fig:crameri}
-\end{figure}
-
-\subsection{Spontaneous subduction}\label{sec:subduction}
-This experiment is performed as presented by \citet{Schmeling2008}, in case of both a sticky air and a true free surface and using different average schemes
-for the viscosities. The domain is rectangular with $L_x=$ \SI{3000}{\km} and $L_y=$ \SI{700}{\km}, a grid resolution of \num{256x64} elements with 50 markers
-per element and Courant number of 0.01. At the beginning of the simulation, a 100 km-thick lithospheric layer with $\rho_m=$ \SI{3300}{\kg\per\cubic\metre}
-and $\eta_m=$ \SI{e23}{\pascal\s} is located at the top of the domain for $x$ between 1000 and 3000 km, while the asthenospheric mantle has $\rho_m=$
-\SI{3200}{\pascal\s} and $\eta_m=$ \SI{e21}{\pascal\s}. In addition, a 200 km-depth lithospheric slab is already subducted in the mantle in order to have
-a spontaneous subduction. In case of sticky air, $L_y=750$ km and a 50 km-thick air layer with $\eta_a$ from \SIrange{e19}{e21}{\pascal\s} overlie the mantle.
-Velocity boundary conditions are set to free slip on the sides and on the bottom of the domain. In case of sticky air, velocities on the top boundary are set
-to free slip conditions as well.
-
-The maximum depth of the slab is tracked for all the simulations (Fig. \ref{fig:subduction}) and they are compared with results obtained by
-\citet{Schmeling2008} with models with regular grids and comparable grid resolutions (rectangles in Fig. \ref{fig:subduction}). Low viscous sticky air models
-(i.e., \SI{e19}{\pascal\s}) enlighten a strong dependence of the sinking velocity with the chosen average scheme (dashed lines in Fig. \ref{fig:subduction}),
-while high viscous sticky air models (i.e., \SI{e21}{\pascal\s}) show a higher resistance at the trench, underestimating the correct solution (dotted lines in
-Fig. \ref{fig:subduction}). A high resistance at the trench can be also observed in case of a true free surface for the low resolution grid used (continuous
-lines in Fig. \ref{fig:subduction}), as pointed out by \citet{Schmeling2008}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Spontaneous_subduction}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Subduction.pdf}
-\caption{Maximum depth of the slab as function of time for all the simulations of the \citet{Schmeling2008} benchmark, in case of harmonic, geometric and
-arithmetic means (red, green and blue lines, respectively) using a sticky air or a true free surface (discontinuous and continuous lines, respectively).
-Red, green and blue rectangular areas indicate the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{500}{\km} in case of sticky air,
-while the grey rectangular area indicates the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{600}{\km} in case of true free surface.}
-\label{fig:subduction}
-\end{figure}
-
-\subsection{The slab detachment}\label{sec:slab}
-The slab detachment benchmark is performed as described by \citet{Schmalholz2011} and \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{1000}{\km}
-and $L_y=$ \SI{660}{\km} and a grid resolution of \num{256x256} elements. A non-linear viscous T-shaped layer with $\rho_l=$ \SI{3300}{\kg\per\cubic\metre} is
-placed at the top of the domain and surrounded by a linear viscous fluid with $\rho_f=$ \SI{3150}{\kg\per\cubic\m}. The top layer is 80 km-thick and a 250
-km-long and 80 km-wide slab is placed at $x=L_x/2$ (Fig. \ref{fig:slab}a). The effective viscosity of the top layer is given by
-\[\eta_{eff}=\eta_0 I_2^{\frac{(1-n)}{n}}\]
-with $\eta_0=$ \SI{4.75e11}{\pascal\s} and $n=4$, while the surrounding fluid has $\eta_f=$ \SI{e21}{\pascal\s}. The effective viscosity is capped between
-$\eta_{min}=$ \SI{e21}{\pascal\s} and $\eta_{max}=$ \SI{e25}{\pascal\s}. Velocity boundary conditions are set to free slip at the top and the bottom, and to
-no slip at the sides of the domain.
-
-The sinking of the slab determines a decrease of the effective viscosity in the T-shaped layer during the first part of the evolution (Fig. \ref{fig:slab}a),
-as consequence of local strain rates, while effective viscosities increase up to $\eta_{max}$ in the top layer after approximately \SI{20}{\mega\year}, when
-necking is complete (Fig. \ref{fig:slab}b).
-
-The width of necking is tracked by means of markers position at the side of the slab. Necking width and time are normalised in function of the initial slab
-(\SI{80}{\km}) and the characteristic time ($t_c=$ \SI{7.1158e14}{\s}) \citep{Schmalholz2011,Glerum2018}. The results in case of arithmetic, geometric and
-harmonic mean of the viscosity are shown in Fig. \ref{fig:necking}, compared with results from \citet{Glerum2018} for which an infinity norm mean was used.
-The influence of the average schemes agrees with results shown in Section \ref{sec:subduction}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Slab_detachment}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Slab.pdf}
-\caption{Effective viscosity for the slab detachment benchmark at the beginning of the evolution (panel a) and when necking is complete (panel b).}
-\label{fig:slab}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Necking.pdf}
-\caption{Normalised width of the necking of the slab detachment benchmark as function of normalised time.}
-\label{fig:necking}
-\end{figure}
-
-\subsection{The indenter experiment}\label{sec:indenter}
-The indenter benchmark simulate a rigid punch on a purely plastic von Mises material, for which exists an analytical solution
-\citep{Thieulot2008,Thieulot2014,Glerum2018}. The domain is a unit square with a grid resolution of \num{256x256} elements. The gravity acceleration is set
-to $\bm{g}=\bm{0}$ and effective viscosity is capped using $\eta_{min}=10^{-4}$ and $\eta_{max}=10^3$. The tolerance for the convergence of the non-linear
-solution is $tol=10^{-9}$, with a maximum number of non-linear iterations set to 500. The medium in the domain has $\rho=0.01$, an initial viscosity
-$\eta_0=10$, cohesion $C=1$ and an angle of internal friction $\phi=0$°. Velocity boundary conditions are set to no slip at the bottom and free slip at the
-sides of the domain. The top of the domain is open with the exception of the central portion (punch area) with width $w_p=0.125$, where $v=-1.05$ and $u$ is
-fixed either to 0 (rough punch experiment) or free (smooth punch experiment).
-
-Fig. \ref{fig:indenter} show the results in terms of viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure
-(panels g and h) for both experiments. The analytical solution indicate that pressure at ($0.5;1$) and ($0.5 \pm w_p;1$)is $p=\pi +1$ and $p=1$, respectively
-(black and red lines in Fig. \ref{fig:smooth_p}). Fig. \ref{fig:smooth_p} shows a clear improvement in the pressure solution when passing from a rough to a
-smooth punch (green and blue lines in Fig. \ref{fig:smooth_p}, respectively), as previously observed by \citet{Thieulot2014} and \citet{Glerum2018}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Indenter}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Indenter.pdf}
-\caption{Viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure (panels g and h) fields for rough (left column) and
-smooth (right column) punch experiments.}
-\label{fig:indenter}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Smooth_Rough.pdf}
-\caption{Nodal pressure in function of the x coordinate at $y=1$ for rough (green line) and smooth (blue line) punch experiment. Black and red lines indicate
-the analytical solution at $x=0.5$ and $x=0.5 \pm w_p$, respectively.}
-\label{fig:smooth_p}
-\end{figure}
-
-\subsection{The brick experiment}\label{sec:brick}
-An instantaneous version of the brick benchmark is performed to verify the correctness of the pressure-dependent plasticity for different angle of internal
-friction, as proposed by \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{40}{\km} and $L_y=$ \SI{10}{\km} and a grid resolution of \num{512x128}
-elements. The gravity acceleration is set to $g_y=$ \SI{-10}{\m\square\s} and effective viscosity is capped using $\eta_{min}=$ \SI{e19}{\pascal\s} and
-$\eta_{max}=$ \SI{e26}{\pascal\s}. The tolerance for the convergence of the non-linear solution is $tol=10^{-7}$, with a maximum number of non-linear iterations
-set to 1000. A 800 m-wide and 400 m-height inclusion with $\eta_b=$ \SI{e20}{\pascal\s} and $\rho_b=$ \SI{2700}{\kg\per\cubic\m} is placed at the bottom of the
-domain at $x=L_x/2$. The inclusion is surrounded by a non-linear viscous medium with $\rho_m=$ \SI{2700}{\kg\per\cubic\m}, an initial viscosity of
-\SI{e23}{\pascal\s}, a linear viscous viscosity of \SI{e25}{\pascal\s} and a cohesion of \SI{40}{\mega\pascal}. Velocities are set to free slip conditions at
-the bottom of the domain and the top is open. Velocities on sides of the domain are fixed to $u=$ \SI{+-2e-11}{\m\per\s} and $v=0$. The experiment is performed
-in compressional and extensional contexts with an angle of internal friction $\phi$ of the non-linear medium variable between 0° and 30°.
-
-As expected, two shear bands stem from the inclusion with variable angles in relation to both the dynamics context and the internal friction angle. Shear band
-angles formed at 45° for $\phi=0$° in both extensional and compressional contexts (Fig. \ref{fig:brick_beam}a), while they are different in case of $\phi\neq0$°.
-Shear band angles in case of $\phi=20$° are shown for compression and extension in Fig. \ref{fig:brick_beam}b  and c, respectively. Values of shear band angles
-as function of internal friction angles from 0° to 30° are extracted at different depths and minimum and maximum values are plotted in Fig. \ref{fig:brick},
-in comparison with theoretical Roscoe, Arthur and Coulomb shear band angles. For all tests the tolerance for nonlinear convergence is defined $tol=10^{-7}$
-and none converges before the maximum number of iterations ($it_{max}=1000$) is reached. However, tests with internal friction angles up to 15° show a constant
-decrease of the velocity residuals, which can not be observed for for tests with higher internal friction angles (Fig. \ref{fig:convergence}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Brick_experiment}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Brick_Beam.pdf}
-\caption{Shear band angles predicted for the brick experiment for $\phi=0$° (panel a) and for $\phi=20$° in case of compression and extension (panels b and c,
-respectively).}
-\label{fig:brick_beam}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Brick.pdf}
-\caption{Shear band angles predicted for the brick experiment in case of compressional and extensional contexts as function of different internal angle of
-friction (continuous black line and red dots), compared with theoretical Roscoe, Artur and Coulomb angles (discontinuous black lines). Red lines indicate the
-range of angles calculated at different depths.}
-\label{fig:brick}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \noindent\includegraphics[width=500px]{./Figures/convergence.pdf}
-  \caption{Convergence rates for all tests. Normalised residuals (left column) are compared with velocity residuals (right column) for extensional (first row)
-  and compressional (second row) settings.}
-  \label{fig:convergence}
-  \end{figure}
-
-\subsection{Advection stabilisation}\label{sec:advection}
-The 1D advection problem proposed by \citet{Donea2003} and \citet{Thieulot2011} is performed to verify the effectiveness of the SUPG method to stabilise the
-advection term of the energy equation. The domain is a 1D segment with $L_x=1$ composed by 50 elements and a discontinuity in the thermal field placed at
-$x=0.25$. Temperature is set to 1 for $x \leq 0.25$ and to 0 for $x > 0.25$. Velocity is set to ${u}=1$ in the entire domain. The simulation is performed for
-250 time steps, with $dt=0.002$, so the thermal discontinuity should be at $x=0.75$ at the end of the simulation.
-
-Temperature profiles at the end of the simulation are shown in \ref{fig:advection} as function of the dimensionless coefficient $\gamma=\tau \bm{v}/h$. In case
-of the classic Galerkin method ($\gamma=0$, blue line in \ref{fig:advection}) the final thermal profile is characterised by strong oscillations, which are
-eliminated in case of the SUPG method ($\gamma=0.045$, orange line in \ref{fig:advection}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Advection_stabilisation}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Advection.pdf}
-\caption{Temperature profile as function of \textit{x} for the advection stabilisation benchmark. Purple line indicates the initial temperature profile; the
-green line indicates the analytical temperature profile after 250 time steps; blue line indicates the temperature profile after 250 time steps in case of the
-classic Galerkin method ($\gamma=0$); orange line indicates the temperature profile after 250 time steps in case of the SUPG method ($\gamma=0.045$).}
-\label{fig:advection}
-\end{figure}
-
-\subsection{Shear heating}\label{sec:simple_shear}
-This exercise is performed in an unit square domain composed by \num{128x128} elements. The velocity field is prescribed on the entire domain with
-$\bm{v}=(L_y-y)y\bm{e_x}$; viscosity, density and specific heat are fixed to 1, while thermal conductivity and radiogenic energy are fixed to 0. Therefore,
-the energy equation (Eq. \ref{eq:energy}) can be simplified as \[\frac{\partial T}{\partial t}=H_s\] and fixing $T(t=0)=0$, the temperature can be find as
-\[T(t)=H_s t\]
-In this case we have
-\begin{eqnarray}
-\dot{\epsilon_{xx}}&=&\frac{\partial u}{\partial x}=0 \nonumber \\
-\dot{\epsilon_{yy}}&=&\frac{\partial v}{\partial y}=0 \nonumber \\
-\dot{\epsilon_{xy}}&=&\frac{1}{2}\left(\frac{\partial v}{\partial x}+\frac{\partial u}{\partial y}\right)=\frac{1}{2}(L_y-2y)\nonumber
-\end{eqnarray}
-and, simplifying Eq. \ref{eq:shear}, shear heating can be calculated as \[H_s(x,y)=(1-2y)^2\]
-
-The solution predicted by the model in terms of velocity, temperature and shear heating match well with the analytical solutions (Fig. \ref{fig:analytical_en}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Simple_shear}.
-
-%\begin{figure}
-%\centering
-%\noindent\includegraphics[width=400px]{./Figures/Energy_Thieulot.pdf}
-%\caption{Temperature and shear heating predicted by the model after 1 time step.}
-%\label{fig:en_thieulot}
-%\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=500px]{./Figures/Analytical_Shear.pdf}
-\caption{Velocity (panel a), temperature (panel b) and shear heating (panel c) as function of the \textit{y} coordinate for the simple shear experiment. The
-solutions predicted by the model (red lines) are compared with the analytical solutions (dashed blue lines).}
-\label{fig:analytical_en}
-\end{figure}
-
-\subsection{Shear and adiabatic heating}\label{sec:shear}
-This problem is performed as presented in Exercise 9.4 in \citet{Gerya2010b}. Two materials are vertically separated in a rectangular domain with $L_x=$
-\SI{1000}{\km}, $L_y=$ \SI{1500}{\km} and a grid resolution of \num{30x20} elements. Constant thermal coefficient expansion ($\alpha=$ \SI{3e-5}{\per\kelvin}),
-temperature ($T=$ \SI{1300}{\kelvin}) and gravity acceleration ($g_y=$ \SI{-10}{\m\square\s}) are assumed in the whole domain. Fluid 1 (on the left side of the
-domain) has $\rho_1=$ \SI{3200}{\kg\per\cubic\m} and $\eta_1=$ \SI{e20}{\pascal\s}; fluid 2 (on the right side of the domain) has $\rho_2=$
-\SI{3300}{\kg\per\cubic\m} and $\eta_2=$ \SI{e22}{\pascal\s}. Velocity boundary conditions are set to free slip on all sides of the domain.
-
-As shown in Fig. \ref{fig:energy}, both shear and adiabatic heating predicted by the code (first row in Fig. \ref{fig:energy}) well recreate the results
-obtained by means of example ${Shear\_adiabatic\_heating.m}$ from \citet{Gerya2010b} (second row in Fig. \ref{fig:energy}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Adiabatic%2BShear}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Energy.pdf}
-\caption{Comparison between shear (first column), adiabatic (second column) and total (third column) energy predicted by the code (panels a, b and c) and those
-created using example ${Shear\_adiabatic\_heating.m}$ from Exercise 9.4 in \citet{Gerya2010b} (panels d, e and f).}
-\label{fig:energy}
-\end{figure}
-
-\subsection{Mantle convection}\label{sec:mantle}
-This problem is performed as presented by \citet{Blankenbach1989} (constant viscosity cases) and \citet{Thieulot2014}, in a 2D unit square domain with gravity
-acceleration $g_y=10^{10}Ra$. The experiment is performed with three different Rayleigh numbers ($Ra=10^4$, $10^5$ and $10^6$) and with different grid
-resolution (between \num{32x32} and \num{128x128} elements). The fluid has constant viscosity, initial density, heat capacity, thermal conductivity
-($\eta= \rho_0= C_p = k=1$), reference temperature ($T_0=0$) and thermal expansion coefficient ($\alpha=10^{-10}$). Temperatures are set to 0 on top and 1 on
-bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\]
-
-The solution generated by the code in terms of $v_{\textrm{rms}}$ (calculated as in Sec. \ref{sec:error}) and Nu as function of time are reported for all the
-simulations in Fig. \ref{fig:mantle}. At the steady state, $v_{\textrm{rms}}$ and Nu of all simulations converge well toward the values from \citet{Blankenbach1989},
-with lower errors for higher resolution grids (Table \ref{tab:mantle}). All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Mantle_convection}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Convection.pdf}
-\caption{$v_{\textrm{rms}}$ (panels a, c and e) and Nu (panels b, d and f) for the mantle convection benchmark as function of time for different grid resolution.
-Panels a and b show the results for $Ra=10^4$; Panels c and d show the results for $Ra=10^5$; panels e and f show the results for $Ra=10^6$. Purple lines
-indicate the convergence values for the $v_{\textrm{rms}}$ and the Nu from \citet{Blankenbach1989} at the steady state.}
-\label{fig:mantle}
-\end{figure}
-
-\begin{table}[H]
-\caption{Comparison between $v_{\textrm{rms}}$ and Nu predicted by the code for the mantle convection experiment and same values as reported in literature.}
-\centering
-\small
-\begin{tabular}{|>{\centering}m{0.09\textwidth}|>{\centering}m{0.05\textwidth}|>{\centering}m{0.24\textwidth}|>{\centering}m{0.15\textwidth}|
-  >{\centering}m{0.1\textwidth} >{\centering}m{0.1\textwidth} >{\centering\arraybackslash}m{0.1\textwidth}|}
-\toprule
-& & \multirow{2}{*}{\citet{Blankenbach1989}} & \citet{Thieulot2014} & \multicolumn{3}{c|}{FALCON}  \\
- & & & \num{200x200} & \num{32x32} & \num{64x64} & \num{128x128}  \\
-\midrule
-\multirow{2}{*}{$Ra=10^4$} & $v_{\textrm{rms}}$ & $42.864947 \pm 0.000020$ & 42.867 & 42.83226 & 42.852793 & 42.861394  \\
-           & Nu      & $4.884409 \pm 0.000010$  & 4.882  & 4.781297 & 4.857475  & 4.877573   \\
-\hline
-\multirow{2}{*}{$Ra=10^5$} & $v_{\textrm{rms}}$ & $193.21454 \pm 0.00010$  & 193.255& 193.872643 & 193.377472 & 193.252290  \\
-           & Nu      & $10.534095 \pm 0.000010$ & 10.507 & 9.602514 & 10.270735  & 10.465629   \\
-\hline
-\multirow{2}{*}{$Ra=10^6$} & $v_{\textrm{rms}}$ & $833.98977 \pm 0.00020$  & 834.712& 848.091176 & 837.767911 & 834.945793  \\
-           & Nu      & $21.972465 \pm 0.000020$ & 21.695 & 15.999266 & 19.703682  & 21.306939   \\
-\bottomrule
-\end{tabular}
-\label{tab:mantle}
-\end{table}
-
-\subsection{Viscoplastic mantle convection}\label{sec:visco_mantle}
-This benchmark is performed as Cases 1-5a described in \citet{Tosi2015} in a 2D unit square domain with two different grid resolutions of \num{32x32} 
-and \num{100x100} elements. Reference viscosity, reference density, heat capacity, thermal conductivity and thermal expansion coefficient are fixed to a constant
-value of 1 ($\eta_0= \rho_0= C_p = k=\alpha=1$), while the gravity has been chosen as $g_y=10^2$, in order to obtain a Rayleigh number $Ra=10^2$. Temperatures
-are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\].
-The viscosity field $\eta$ is calculated as the harmonic average between a linear part $\eta_{lin}$ that depends on temperature only or on temperature and 
-depth $y$ and a nonlinear-plastic part $\eta_{pl}$ that depends on the strain rate $\dot{\bm{\epsilon}}$, as follows
-\[\eta(T,y,\dot{\bm{\epsilon}})=2\left(\frac{1}{\eta_{lin}(T,y)}+\frac{1}{\eta_{pl}(\dot{\bm{\epsilon}})}\right)^{-1}\]
-The linear and the nonlinear-plastic parts of the viscosity are calculated following \citet{Tosi2015} as follows
-\[\eta_{lin}(T,y)=\exp(-\ln(\Delta\eta_T)T+\ln(\Delta\eta_y)y)\]
-\[\eta_{pl}(\dot{\bm{\epsilon}})=\eta^*+\frac{\sigma_Y}{\sqrt{\dot{\bm{\epsilon}}:\dot{\bm{\epsilon}}}}\]
-where $\Delta\eta_T$, $\Delta\eta_y$, $\eta^*$ and $\sigma_Y$ are parameters chosen as listed in Table \ref{tab:tosi_setup}.
-
-For each case, some diagnostic quantities measured at the steady state are reported in Table \ref{tab:visco_mantle} (see Section \ref{sec:error} for 
-the computation of all the quantities), showing a well fit with results described in \citet{Tosi2015}. Data and charts for all cases can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Viscoplastic_mantle_convection}.
-
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{white!100!}{gray!25!}
-%\begin{table}[H]
-%  \caption{Results of the viscoplastic mantle convection benchmark for Cases 1-4 compared with results obtained with ELEFANT and YACC (from \citealp{Tosi2015}).}
-%  \centering
-%  \resizebox{\textwidth}{!}{
-%\begin{tabular}{|c|ccc|ccc|ccc|ccc|}
-%  \toprule
-%    & \multicolumn{3}{c|}{\textbf{Case 1}} & \multicolumn{3}{c|}{\textbf{Case 2}} & \multicolumn{3}{c|}{\textbf{Case 3}} &
-%    \multicolumn{3}{c|}{\textbf{Case 4}} \\
-%   & FALCON & ELEFANT& YACC & FALCON& ELEFANT& YACC & FALCON & ELEFANT & YACC & FALCON & ELEFANT & YACC \\
-%  \midrule
-%   $\left\langle T \right\rangle$ & 0.7759 & 0.7758 & 0.7767 & 0.6043 & 0.6036 & 0.6058 & 0.7275 & 0.7275 & 0.7286 & & 0.5277 & 0.5289\\
-%    Nu$^{top}$ & 3.4280 & 3.4214 & 3.4298 & 8.5242 & 8.5115 & 8.5278 & 3.0368 & 3.0347 & 3.0374 && 6.5912 & 6.5572 \\
-%    Nu$^{bottom}$ & 3.3168 & 3.313 & 3.3143 & 8.4599 & 8.4465 & 8.3990 & 3.0086 & 2.9908 & 2.9628 && 6.5834 & 6.5243 \\
-%    $v_{\textrm{rms}}$ & 249.901 & 249.134 & 251.7997 & 141.2451 & 140.7087 & 142.2020 & 100.2776 & 100.1208 & 100.9467 && 79.1105 & 79.6202 \\
-%    $v_{\textrm{rms}}^{top}$ & 1.8907 & 1.8642 & 1.8298 & 104.7928 & 104.1228 & 105.1767 & 2.0832 & 2.0652 & 2.0374 && 74.7596 & 75.8414 \\
-%    $u_{max}^{top}$ & 2.6340 & 2.6119 & 2.5516 & 121.9212& 121.766 & 122.3238 & 2.9086 & 2.9019 & 2.8458 && 88.9146 & 89.2940 \\
-%    $\eta_{min}$ & - & - & - & $2.0303\times10^{-5}$ & $2.0600\times10^{-5}$ & $1.9853\times10^{-5}$ & $4.8129\times10^{-5}$ & $4.8080\times10^{-5}$ & $4.7907\times10^{-5}$ &
-%    & $1.9860\times10^{-4}$ & $1.9174\times10^{-4}$ \\
-%    $\eta_{max}$ & - & - & - & 1.3325 & 1.85 & 1.6095 & 0.9360 & 0.9023 & 1 && 1.52 & 1.6773 \\
-%    $\left\langle W \right\rangle$ & 2.4299 & 2.4316 & 2.4583 & 7.5648 & 7.5665 & 7.6222 & 2.0387 & 2.0384 & 2.0400 && 5.6216 & 5.6512 \\
-%    $\left\langle \Phi \right\rangle/Ra$ & 2.4311 & 2.4276 & 2.4333 & 7.5661 & 7.5585 & 7.6194 & 2.0386 & 2.0370 & 2.0335 && 5.6182 & 5.6463 \\
-%    $\delta$ & 0.05\% & 0.16\% & 1.02\% & 0.02\% & 0.11\% & 0.04\% & $<$0.01\% & 0.07\% & 0.32\% && 0.06\% & 0.09\% \\
-%  \bottomrule
-%  \end{tabular}}
-%  \label{tab:visco_mantle}
-%  \end{table}
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{2}{white!100!}{gray!25!}
-\begin{table}[H]
-  \caption{Benchmark cases and corresponding parameters.}
-  \centering
-  \small
-\begin{tabular}{|>{\centering}m{0.08\textwidth}|>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}
-  >{\centering}m{0.05\textwidth} >{\centering}m{0.05\textwidth} |>{\centering\arraybackslash}m{0.2\textwidth}|}
-  \toprule
-    Case & $Ra$ & $\Delta\eta_T$ & $\Delta\eta_y$ & $\eta^*$ & $\sigma_Y$ & Convective regime \\
-  \midrule
-  \rowcolor{gray!25!}
-    1 & $10^2$ & $10^5$ & 1 & - & - & Stagnant lid \\
-    2 & $10^2$ & $10^5$ & 1 & $10^{-3}$ & 1 & Mobile lid  \\
-    \rowcolor{gray!25!}
-    3 & $10^2$ & $10^5$ & 10 & - & - & Stagnant lid  \\
-    4 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 1 & Mobile lid \\
-    \rowcolor{gray!25!}
-    5 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 4 & Periodic \\
-  \bottomrule
-  \end{tabular}
-  \label{tab:tosi_setup}
-  \end{table}
-
-\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{gray!25!}{white!100!}
-\begin{table}
-  \caption{Results of the viscoplastic mantle convection benchmark for different grid resolutions in Cases 1-5 compared with results obtained with ELEFANT
-  and YACC ($100\times100$ elements, from \citealp{Tosi2015}) and in Stone 28 ($32\times32$ elements, from
-  \url{https://github.com/cedrict/fieldstone/tree/master/python_codes/fieldstone_28}).}
-  \centering
-  \resizebox{\textwidth}{!}{
-\begin{tabular}{|cc|ccccccccccc|}
-  \toprule
-  \multicolumn{13}{|c|}{\textbf{Case 1}} \\
-  \midrule
-  Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-  & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-  \hline
-  \rowcolor{gray!25!}
-    YACC & $100\times100$ & 0.7767 & 3.4298 & 3.3143 & 251.7997 & 1.8298 & 2.5516 & - & - & 2.4583 & 2.4333 & 1.02\% \\
-    ELEFANT & $100\times100$ & 0.7758 & 3.4214 & 3.313 & 249.134 & 1.8642 & 2.6119 & - & - & 2.4316 & 2.4276 & 0.16\% \\
-    \rowcolor{gray!25!}
-    Stone 28 & $32\times32$ & 0.7737 & 3.3987 & - & 243.872 & - & - & - & - & - & - & - \\
-%    FALCON & $32\times32$ & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & - & - & 2.4745 & 2.4864 & 0.48\% \\
-%    \rowcolor{gray!25!}
-    FALCON & $100\times100$ & 0.7759 & 3.4280 & 3.3168 & 249.901 & 1.8907 & 2.6340 & - & - & 2.4299 & 2.4311 & 0.05\% \\
-    \midrule
-    \multicolumn{13}{|c|}{\textbf{Case 2}} \\
-    \midrule
-    Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-    & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-    \hline
-    \rowcolor{gray!25!}
-     YACC & $100\times100$ & 0.6058 & 8.5278 & 8.3990 & 142.2020 & 105.1767 & 122.3238 & $1.9853\times10^{-5}$ & 1.6095 & 7.6222 & 7.6194 & 0.04\% \\
-     ELEFANT & $100\times100$ & 0.6034 & 8.5115 & 8.4465 & 140.8067 & 104.1228 & 121.766 & $2.0600\times10^{-5}$ & 1.85 & 7.5665 & 7.5585 & 0.11\% \\
-     \rowcolor{gray!25!}
-     Stone 28 & $32\times32$ & 0.6052 & 8.1715 & - & 141.518 & - & - & - & - & - & - & - \\
-%     FALCON & $32\times32$ & 0.6454 & 8.1422 & 7.4020 & 156.6908 & 95.7226 & 116.9302 & $2.2572\times10^{-5}$ & 0.7194 & 7.4289 & 7.4452 & 0.22\% \\
-%     \rowcolor{gray!25!}
-     FALCON & $100\times100$ & 0.6044 & 8.5266 & 8.4590 & 141.3684 & 104.8867 & 122.0470 & $2.0333\times10^{-5}$ & 1.2914 & 7.5707 & 7.5718 & 0.01\% \\
-     \midrule
-     \multicolumn{13}{|c|}{\textbf{Case 3}} \\
-     \midrule
-     Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-     & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-     \hline
-     \rowcolor{gray!25!} 
-     YACC & $100\times100$ & 0.7286 & 3.0374 & 2.9628 & 100.9467 & 2.0374 & 2.8458 & $4.7907\times10^{-5}$ & 1 & 2.0400 & 2.0335 & 0.32\% \\
-      ELEFANT & $100\times100$ & 0.7275 & 3.0347 & 2.9908 & 100.1208 & 2.0652 & 2.9019 & $4.8080\times10^{-5}$ & 0.9023 & 2.0384 & 2.037 & 0.07\% \\
-      \rowcolor{gray!25!}
-      Stone 28 & $32\times32$ & 0.7277 & 3.0347 & - & 100.018 & - & - & - & - & - & - & - \\
-%      FALCON & $32\times32$ & 0.7281 & 3.0618 & 2.8132 & 101.0092 & 2.1630 & 3.0178 & $4.8500\times10^{-5}$ & 0.7845 & 2.0699 & 2.0695 & 0.02\% \\
-%      \rowcolor{gray!25!}
-      FALCON & $100\times100$ & 0.7275 & 3.0368 & 3.0086 & 100.2776 & 2.0832 & 2.9086 & $4.8129\times10^{-5}$ & 0.9360 & 2.0387 & 2.0386 & $<$0.01\% \\
-      \midrule
-      \multicolumn{13}{|c|}{\textbf{Case 4}} \\
-      \midrule
-      Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-      & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-      \hline
-      \rowcolor{gray!25!} 
-      YACC & $100\times100$ & 0.5289 & 6.5572 & 6.5243 & 79.6202 & 75.4814 & 89.2940 & $1.9174\times10^{-4}$ & 1.6773 & 5.6512 & 5.6463 & 0.09\% \\
-       ELEFANT & $100\times100$ & 0.5277 & 6.5912 & 6.5834 & 79.1105 & 74.7596 & 88.9146 & $1.9860\times10^{-4}$ & 1.5200 & 5.6216 & 5.6182 & 0.06\% \\
-       \rowcolor{gray!25!}
-       Stone 28 & $32\times32$ & 0.5286 & 6.4036 & - & 79.4746 & - & - & - & - & - & - & - \\
-%       FALCON & $32\times32$ & 0.5334 & 6.4679 & 6.3435 & 80.6153 & 76.0511 & 90.2058 & $2.0393\times10^{-4}$ & 0.6453 & 5.6676 & 5.6698 & 0.04\% \\
-%       \rowcolor{gray!25!}
-       FALCON & $100\times100$ & 0.5279 & 6.5897 & 6.5830 & 79.1220 & 75.0715 & 88.8484 & $1.9693\times10^{-4}$ & 1.4133 & 5.6127 & 5.6113 & 0.03\% \\
-       \midrule
-       \multicolumn{13}{|c|}{\textbf{Case 5}} \\
-       \midrule
-       Code & Resolution & $P$ & $\left\langle T \right\rangle_{min}$ & $\left\langle T \right\rangle_{max}$ & Nu$^{top}_{min}$ & Nu$^{top}_{max}$ &
-       $v_{\textrm{rms}}^{min}$ & $v_{\textrm{rms}}^{max}$ & \multicolumn{2}{c}{$(\left\langle \Phi \right\rangle/Ra)_{min}$} & 
-       \multicolumn{2}{c|}{$(\left\langle \Phi \right\rangle/Ra)_{max}$} \\
-       \hline
-       \rowcolor{gray!25!}
-         YACC & $100\times100$ & 0.0792 & 0.6543 & 0.6722 & 2.6950 & 7.2792 & 41.9813 & 98.8230 & \multicolumn{2}{c}{1.4347} & 
-         \multicolumn{2}{c|}{9.2748} \\
-         ELEFANT & $100\times100$ & 0.076 & 0.6514 & 0.6693 & 2.6808 & 7.3506 & 41.2578 & 100.089 & \multicolumn{2}{c}{1.3324} & 
-         \multicolumn{2}{c|}{9.5289} \\
-         \rowcolor{gray!25!}
-         Stone 28 & $32\times32$ & 0.0764 & 0.6521 & 0.6697 & 2.6912 & 7.0763 & 41.61 & 98.183 & \multicolumn{2}{c}{-} & 
-         \multicolumn{2}{c|}{-} \\
-     %    FALCON & $32\times32$ & & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & \multicolumn{2}{c}{-} & 
-     %     \multicolumn{2}{c|}{-} \\
-     %    \rowcolor{gray!25!}
-         FALCON & $100\times100$ & 0.0775 & 0.6519 & 0.6697 & 2.6847 & 7.3629 & 41.5256 & 99.5466 & \multicolumn{2}{c}{1.3399} & 
-         \multicolumn{2}{c|}{9.3993} \\
-  \bottomrule
-  \end{tabular}}
-  \label{tab:visco_mantle}
-  \end{table}
-
-\subsection{Thin layer entrainment}\label{sec:thin}
-This experiment is performed as originally proposed by \citet{vanKeken1997} with two incompressible fluids in a rectangular domain with $L_x=2$ and $L_y=1$
-and gravity acceleration $g_y=10^{10}Ra$. The Rayleigh number and the compositional Rayleigh number are fixed to $Ra=300000$ and $Rc=450000$, respectively.
-Both fluids have constant viscosity, thermal conductivity, specific heat ($\eta=\rho=C_p=1$) and thermal expansion coefficient ($\alpha=10^{-10}$). Fluid 1
-has a density $\rho_1=1$, while fluid 2 is denser ($\rho_2=\rho_1+1.5 \cdot 10^{-10}$) and is placed at the bottom of the domain, for $y \leq 0.025$.
-Temperature are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides of the domain. The initial
-temperature field is given by
-\[T(x,y)=T_u(x,y)+T_l(x,y)+T_r(x,y)+T_s(x,y)-\frac{3}{2}\]
-with
-\begin{eqnarray}
-T_u(x,y)&=&\frac{1}{2}erf\left(\frac{1-y}{2}\sqrt{\frac{u_0}{x}}\right)\nonumber \\
-T_l(x,y)&=&1-\frac{1}{2}erf\left(\frac{y}{2}\sqrt{\frac{u_0}{L_x-x}}\right)\nonumber \\
-T_r(x,y)&=&\frac{1}{2}+\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{y+1}} \exp\left(-\frac{x^2u_0}{4y+4}\right)\nonumber \\
-T_s(x,y)&=&\frac{1}{2}-\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{2-y}} \exp\left(-\frac{(L_x-x)^2u_0}{8-4y}\right)\nonumber
-\end{eqnarray}
-and
-\begin{eqnarray}
-u_0&=&\frac{L_x^{7/3}}{(1+L_x^4)^{2/3}}\left(\frac{Ra}{2\sqrt{\pi}}\right)^{\frac{2}{3}}\nonumber \\
-Q&=&2\sqrt{\frac{L_x}{\pi u_0}}\nonumber
-\end{eqnarray}
-
-The experiment is performed with two grid resolution (\num{125x40} and \num{200x80} elements), with 100 markers per element and CFL of 0.25. $v_{\textrm{rms}}$
-(see Sec. \ref{sec:error}) calculated for both the simulations match well with results from \citet{vanKeken1997} and obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and ELEFANT \citep{Thieulot2014}, obtained with similar grid resolutions
-(Fig. \ref{fig:thin}). In addition, a variety of simulations with different resolutions and aspect ratios are performed to verify that the sensitivity of the
-initial velocity field. Results match well with results from \citet{vanKeken1997} (in the grey area) and \citet{Thieulot2014} (Fig. \ref{fig:thin_initial}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Thin_layer}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_thin.pdf}
-\caption{$v_{\textrm{rms}}$ for the thin layer experiment as function of time for different grid resolution. Results are compared with results obtained by
-\citet{vanKeken1997} (black lines) and with ASPECT and ELEFANT.}
-\label{fig:thin}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrmszero.pdf}
-\caption{$v_{\textrm{rms}}$ at $t=0$ for the thin layer experiment as function of the element size for different aspect ratios of the elements. Results are compared
-with results obtained by \citet{vanKeken1997} (black dots in the grey area) and with ASPECT and ELEFANT.}
-\label{fig:thin_initial}
-\end{figure}
-
-\subsection{Sinking hydrated cylinder}\label{sec:quinquis}
-This experiment is performed as explained in \citet{Quinquis2014} and simulates the sinking of a cold, hydrated cylinder into a warm, dry mantle. The domain
-is square with $L_x=L_y=$ \SI{300}{\km}, a grid resolution of \num{300x300} elements and 25 markers per element. The sinking cylinder is located at $x=$
-\SI{150}{\km} and $y=$ \SI{170}{\km} with a radius of \SI{20}{\km} and has $\rho_c=$ \SI{3250}{\kg\per\cubic\m}, $\eta_c=$ \SI{e23}{\pascal\s}, a thermal
-conductivity $k_c=$ \SI{4.5}{\watt\per\m\per\kelvin},a specific heat $C_{p_c}=$ \SI{1250}{\joule\per\kg\per\kelvin}; the surrounding mantle has $\rho_m=$
-\SI{3200}{\kg\per\cubic\m}, $\eta_m=$ \SI{e20}{\pascal\s}, a thermal conductivity $k_m=$ \SI{105}{\watt\per\m\per\kelvin} and specific heat $C_{p_m}=$
-\SI{1250}{\joule\per\kg\per\kelvin}; a 58 km-thick lithosphere overlies the mantle and has $\rho_l=$ \SI{3200}{\kg\per\cubic\m}, $\eta_l=$ \SI{e23}{\pascal\s},
-a thermal conductivity $k_l=$ \SI{4.5}{\watt\per\m\per\kelvin}, specific heat $C_{p_l}=$ \SI{750}{\joule\per\kg\per\kelvin} and a thermal diffusivity of
-\SI{e-6}{\square\m\per\s}. Velocity boundary conditions are set to free slip on all sides of the domain. Temperature increase linearly in the lithosphere,
-from 0° to 1300°C, and in the mantle, from 1300° to 1360.5°C. The initial temperature of the cylinder is fixed to 400°C. Throughout the evolution, temperature
-is fixed to 0° and 1360.5°C at the top and bottom of the domain, respectively. The initial bound water content is imposed at 2 wt.\% in the cylinder and at
-0 wt.\% in both the mantle and the lithosphere. The maximum amount of water in the mantle is fixed to 0.2 wt.\%, while maximum water content in the cylinder
-and in the lithosphere is function of pressure and temperature and is calculated using a serpentinized harzburgite with Perple\_X, as in \citet{Quinquis2014}.
-
-As shown in \citet{Quinquis2014}, the model is characterised by a progressive dehydration in the external portion of the cylinder, due to the increase of
-temperature, with a consequent vertical migration of free water that hydrates the mantle above the cylinder, up to the lithosphere. The distribution of
-bound and free water during the evolution of the experiment is shown in Fig. \ref{fig:quinquis} (left and right columns, respectively). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Hydration}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Hydration.pdf}
-\caption{Distribution of bound and free water (left and right columns, respectively) at $t=0.5$ and $t=$ \SI{1.25}{\mega\year} (first and second rows,
-respectively) for the sinking hydrated cylinder experiment. Only markers with bound or free water not equal to 0 are plotted.}
-\label{fig:quinquis}
-\end{figure}
-
 \subsection{Mantle melting curves}\label{sec:katz}
 This instantaneous experiment is performed in order to recreate isobaric melting curves, as function of temperature, bulk water and content of water in the
-melt, to compare with melting curves obtained by \citet{Katz2003}. The domain is square with $L_x=L_y=100$ km, a grid resolution of \num{100x100} elements and
+melt, to compare with melting curves obtained by \citet{Katz2003}. The domain is square with $L_x=L_y=100$ km, a grid resolution of $100\times100$ elements and
 25 markers per element. Density is set in the entire domain to \SI{3300}{\kg\per\cubic\m} and temperature increases from the sides to the centre from 1000° to
@@
 \end{figure}
-
-\clearpage
-
-\bibliography{export}
-
-\end{document}

### 2022-09-07 tex/Benchmark/mantle.tex
update main

--- 2022-07-20 FALCON_description.tex
@@
-\documentclass[hidelinks,10pt,a4paper]{article}
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{textcomp}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage{booktabs}
-\usepackage{multirow}
-\usepackage[cm]{fullpage}
-\usepackage[round]{natbib}
-\bibliographystyle{abbrvnat}
-\usepackage{doi}
-\usepackage{bm}
-\usepackage[table]{xcolor}
-\hypersetup{
-    colorlinks,
-    linkcolor={blue},
-    citecolor={blue},
-    urlcolor={blue}
-}
-\usepackage{float}
-\restylefloat{table}
-\usepackage{siunitx}
-\DeclareSIUnit\year{yr}
-
-\begin{document}
-
-\section{Introduction}\label{sec:intro}
-This document includes a description of the code FALCON, with a presentation of the features implemented followed by results of extensive benchmarks listed in
-Table \ref{tab:list}, performed to verify that the code solves correctly a wide range of problems. Description and results of all benchmarks are shown
-in Section \ref{sec:benchmark}.
-
-\begin{table}[H]
-\caption{List of benchmarks performed for each feature of the code. Data of all benchmarks can be found at 
-\url{https://github.com/aleregorda/Benchmarks}}
-\centering
-\begin{tabular}{| l | l | l |}
-\toprule
- \textbf{Feature}  & \textbf{Benchmark} & \textbf{Section} \\
-\midrule
-  Solver  & Stokes flow & \ref{sec:stokes} \\
-\hline
-  \multirow{2}{*}{Markers advection} & Zalesak disk & \ref{sec:runge} \\
-    & Conservative Interpolation Velocity & \ref{sec:CVI} \\
-\hline
-  \multirow{4}{*}{Momentum equation}  & Poiseuille flow & \ref{sec:poiseuille}\\
-    & Instantaneous 2D sphere & \ref{sec:ist_sphere} \\
-    & Rayleigh-Taylor experiment & \ref{sec:rayleigh} \\
-    & Falling block & \ref{sec:block} \\
-\hline
-    & 2D Stokes sphere & \ref{sec:sphere} \\
-    Sticky air and  & Stabilisation algorithm & \ref{sec:stab} \\
-    free surface & Topography relaxation & \ref{sec:crameri} \\
-  & Spontaneous subduction & \ref{sec:subduction} \\
-\hline
-  Erosion and sedimentation  & - & - \\
-\hline
-   Non-linear  & Slab detachment & \ref{sec:slab} \\
-   visco-plasticity & Indenter & \ref{sec:indenter} \\
-    & Brick & \ref{sec:brick} \\
-\hline
-  \multirow{3}{*}{Energy equation}  & Advection stabilisation & \ref{sec:advection} \\
-    & Simple shear heating & \ref{sec:simple_shear} \\
-    & Shear and adiabatic heating & \ref{sec:shear} \\
-\hline
-\multirow{3}{*}{Energy + momentum} & Mantle convection & \ref{sec:mantle} \\
-& Viscoplastic mantle convection & \ref{sec:visco_mantle} \\
-& Thin layer entrainment & \ref{sec:thin} \\
-\hline
-  Phase changes  & \multirow{2}{*}{Hydrated sinking cylinder} & \multirow{2}{*}{\ref{sec:quinquis}} \\
-  and hydration  &  & \\
-\hline
-  Melting & Experimental melting curves & \ref{sec:katz} \\
-\bottomrule
-\end{tabular}
-\label{tab:list}
-\end{table}
-
-The code is written in Fortran90 and uses the finite element method (FEM) with quadrilateral $Q_1 \times P_0$ elements (continuous bilinear velocity and
-discontinuous constant pressure), associated with MUMPS\footnote{\url{http://mumps.enseeiht.fr/}} \citep{Amestoy2001,Amestoy2019} that is a software package
-for solving systems of linear equations of the form $A \cdot x = b$, where \textit{A} is a square sparse matrix, by means of a direct method. Since
-$Q_1 \times P_0$ elements do not satisfy Ladyzhenskaya, Babuska and Brezzi (LBB) stability condition \citep{Donea2003}, elemental pressure is elaborated
-in post-processing to avoid spurious pressures by means of a double interpolation to smooth the pressure. Following this procedure, the elemental pressure
-is interpolated onto nodes and then back onto elements. After the smoothing procedure, the pressure field is calculated again on the nodes to be used in
-combination with the lithostatic pressure, which is calculated onto nodes. Correctness of the solution and performances of the solver in terms of time and
-memory usage are tested solving the Stokes flow with the analytical solution proposed by \citet{Donea2003} (Section \ref{sec:stokes}).
-
-The thermo-mechanics of crust-mantle systems is described by means of the conservation of mass, momentum and energy equations, expressed as follows:
-\begin{eqnarray}
-\label{eq:mass}\frac{\partial \rho}{\partial t}+\bm{\nabla} \cdot (\rho \bm{u})&=&0\\
-\label{eq:momentum}-\bm{\nabla} p + \bm{\nabla} \cdot \bm{\tau} + \rho \bm{g}&=&0\\
-\label{eq:energy}\rho C_p \left(\frac{\partial T}{\partial t} + \bm{u} \cdot \bm{\nabla} T \right)&=&\bm{\nabla} \cdot \left(k\bm{\nabla} T\right) + H_{tot}
-\end{eqnarray}
-where $\rho$ is the density, $\bm{u}$ is the velocity, \textit{p} is the pressure, $\bm{\tau}$ is the deviatoric stress, $\bm{g}$ is the gravity acceleration,
-$C_p$ is the specific heat at constant pressure, \textit{T} is the temperature, \textit{k} is the thermal conductivity and $H_{tot}$ is the total internal heat
-production. Density variations due to temperature are generally small enough to assume the density as constant ($\rho=\rho_0$) in Eq. \ref{eq:mass} and in
-Eq. \ref{eq:energy}, while it must be treated as a variable in the buoyancy term of Eq. \ref{eq:momentum}, such that \[\rho=\rho_0(1-\alpha(T-T_0))\]
-where $\rho_0$ is the density at a reference temperature $T_0$ and $\alpha$ is the coefficient of thermal expansion. Eq. \ref{eq:mass} then can be rewritten as
-\begin{equation}\label{eq:mass_0}
-\bm{\nabla} \cdot \bm{u}=0
-\end{equation}
-This simplification is known as the extended Boussinesq approximation.
-
-The time step $dt$ is chosen in according to Courant-Friedrichs-Lewy condition \citep{Anderson1995}, such that
-\[dt=C_n\; \min\left(\frac{\Delta x}{\max|\bm{u}|},\frac{(\Delta x)^2}{\kappa}\right)\]
-where $0<C_n<1$ is the Courant (CFL) number, $\Delta x$ is the minimum dimension of smallest element, $\max|\bm{u}|$ is the maximum velocity calculated on the
-entire domain and
-$\kappa=\frac{k}{\rho C_p}$ is the thermal diffusivity \citep{Thieulot2014}.
-
-\section{Lagrangian markers}\label{sec:markers}
-Elemental properties (density, viscosity, thermal conductivity, specific heat and thermal expansion) needed to solve Eqs. \ref{eq:mass_0}, \ref{eq:momentum}
-and \ref{eq:energy} are related to the composition of each element, determined by means of Lagrangian markers that are characteristic of different materials
-in the domain. Elemental properties (with the sole exception of the viscosity) are calculated using an arithmetic mean, as \[P_e=\frac{1}{n}\sum_{i=1}^n P_i\]
-where $P_e$ is the elemental property, $P_i$ is the property characteristic of the material of each marker and $n$ is the number of markers of the element.
-Differently, the average scheme for the elemental viscosity can be chosen between harmonic, geometric and arithmetic mean. Markers can be placed regularly or
-randomly at the beginning of the simulation and their advection is performed by either a 2nd-order or a 4th-order Runge-Kutta in space, interpolating the velocity 
-field on each marker by means of the shape functions (see Section \ref{sec:runge}). The interpolated velocity is then corrected by means of the Conservative
-Velocity Interpolation (CVI), which introduces a corrective term to reduce dispersion and clustering of particles in both steady state and time-dependent 
-\citep{Wang2015} (see Section \ref{sec:CVI}).
-
-The initial distribution of the markers is created using the open-source code library
-Geodynamic World Builder \citep{Fraters2019}. During the simulation, each marker carries memory of temperature, pressure and accumulated strain, which are
-determined interpolating the relative nodal parameter. In particular, pressure is interpolated onto each marker after the smoothing procedure. The number
-of markers contained in each element is maintained between $n_{min}$ and $n_{max}$. When in an element there are less markers than $n_{min}$ the code adds
-random markers to reach the $n_{min}$, while if the number is higher than $n_{max}$ some of them are randomly deleted. When new markers are added, they assume
-the properties of the nearest marker. In this way elements are never empty and maintain a number of markers inside a prefixed range.
-
-\section{The momentum equation}\label{sec:mom_eq}
-The penalty method is implemented so that Eq. \ref{eq:mass_0} can be written as:
-\begin{equation}\label{eq:mass_penalty}
-\bm{\nabla} \cdot \bm{u} + \frac{p}{\lambda}=0
-\end{equation}
-where $\lambda$ is the so-called penalty parameter, which should be 6-7 orders of magnitude larger than the shear viscosity to ensure that mass conservation
-is satisfied \citep{Donea2003,Thieulot2014}.
-
-The deviatoric stress tensor in Eq. \ref{eq:momentum} can be written in terms of the strain rate tensor as $\bm{\tau}=2\eta\dot{\bm{\epsilon}}$, with
-$\dot{\bm{\epsilon}}=\frac{1}{2}\left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)$. Therefore, Eq. \ref{eq:momentum} can be rewritten as:
-\begin{equation}\label{eq:momentum_pressure}
--\bm{\nabla} p + \bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right) + \rho \bm{g} = 0
-\end{equation}
-Finally, using pressure from Eq. \ref{eq:mass_penalty}, Eq. \ref{eq:momentum_pressure} can be rewritten as:
-\begin{equation}\label{eq:momentum_penalty}
-\lambda \bm{\nabla} \left(\bm{\nabla} \cdot \bm{v}\right)+\bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right)+\rho\bm{g}=0
-\end{equation}
-The penalty method is been associated to the iterative Uzawa method, as described in detail in \citet{Dabrowski2008} and \citet{Thieulot2014}
-(see Section \ref{sec:poiseuille}). Results of benchmarks performed to verify the correctness in the implementation of Eq. \ref{eq:momentum_penalty} are
-shown in Sections \ref{sec:ist_sphere} and \ref{sec:rayleigh}.
-
-To support large viscosities variations the penalty parameter is related to the effective elemental viscosity by means of a dimensionless coefficient, so that
-$\lambda=\lambda_e(e)\eta_{eff}(e)$ \citep{Marotta2006,Dabrowski2008,Thieulot2014}. Benchmark of the falling block \citep{Gerya2003a,Gerya2010,Thieulot2011}
-is performed to verify that the code can correctly deal buoyancy driven flows with strong viscosity contrasts (Section \ref{sec:block}).
-
-\section{Sticky air and free surface}\label{sec:sticky}
-The Earth's surface can be treated by means of either the so-called sticky-air or a true free surface method, both of them implemented in the code. In the
-sticky-air method the surface is approximated with the introduction of a buoyant layer with a viscosity at least four orders of magnitude lower than the crust
-\citep{Schmeling2008,Crameri2012} and the interface between lithosphere and air is defined using a chain of passive markers that are advected as the Lagrangian
-markers. The correctness of the evolution of the markers chain is tested with the experiment of a 2D time-dependent Stokes sphere below a free surface and
-compared with results from ASPECT\footnote{\url{https://aspect.geodynamics.org/}} \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} (Section
-\ref{sec:sphere}). In the true free surface case the top boundary is assumed stress-free and velocities are not fixed. In this case topography variations are
-described by vertical deformations of the mesh that depend on the velocity field of the nodes that identify the free surface, while horizontal deformations are
-not taken into account. This procedure is known as the Arbitrary Lagrangian-Eulerian (ALE) method and its implementation follows the technique described in
-\citet{Thieulot2011}. However, although the implementation of a true free surface better reproduces laboratory experiment, extremely small time steps can be
-necessary to maintain stability \citep{Kaus2010a,Quinquis2011,Thieulot2014}. Therefore, the stability algorithm proposed by \citet{Kaus2010a} is implemented
-to avoid instabilities due to high density differences at the free surface when using too large time steps. The implementation of this algorithm are tested by
-performing the experiment described by \citet{Kaus2010a} (Section \ref{sec:stab}).
-
-The topography relaxation benchmark proposed by \citet{Crameri2012} is performed to verify that the code correctly recovers topography variations in case of
-both the sticky-air and the true free surface method (Section \ref{sec:crameri}). Finally, Section \ref{sec:subduction} show the results of the spontaneous
-subduction experiment described by \citet{Schmeling2008}.
-
-\section{Erosion and sedimentation}\label{sec:erosion}
-Surface processes at either the lithosphere-air interface or in correspondence of the free surface have been implemented by means of the software
-Fastscape\footnote{\url{https://fastscape.org/}} \citep{Braun2013,Cordonnier2019,Yuan2019a,Yuan2019b}. FastScapeLib is a set of routines that solve the stream
-power law (enriched by a sediment transport/deposition term) \citep{Yuan2019a}, hillslope diffusion and marine transport and deposition \citep{Yuan2019b}.
-The partial differential equation solved by FastScapeLib is:
-\begin{equation}
-\frac{\partial h}{ \partial t}=U-K_f A^{m_e}S^{n_e} + \frac{G}{A}\int_A \left( U-\frac{\partial h}{\partial t} \right) dA + K_d \bm{\nabla}^2 h 
-\end{equation}
-where $h$ is the topography, $U$ is the uplift, $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $S$ is the slope, $m_e$ and $n_e$ are
-parameters that should be chosen to constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}, $G$ is the dimensionless
-deposition/transport coefficient for the enriched stream power law and $K_d$ is the effective diffusivity. Terms $K_f A^{m_e}S^{n_e}$ and $K_d \bm{\nabla}^2 h$
-consider long-range processes, described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, and short-range processes,
-proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}.
-
-In case of sticky air, slope is determined by means of the markers chain, which is vertically corrected considering erosion and sedimentation rates. After the
-correction, continental markers above the markers chain are transformed in air markers and air markers below the markers chain become sediments. Similarly,
-slope is determined considering the nodes on the top boundary in case of true free surface and they are vertically corrected as for the markers chain. In this
-case, continental markers above the top boundary are simply deleted.
-%Surface processes at the lithosphere-air interface are implemented taking into account fluvial erosion (long-range processes), hillslope processes (short-range
-%processes) and failure of slopes exceeding a threshold angle $\theta$ \citep{Burov1997,Tucker1999,Sacek2012,Croissant2014,Ueda2015,Beucher2020}. Erosion rates
-%$\dot{E}_{lr}$ due to long-range processes can be described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{lr}=K_f A^{m_e} \bm{\nabla} h^{n_e}
-%\end{equation}
-%where $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $\bm{\nabla} h$ is the slope and $m_e$ and $n_e$ are parameters chosen in to
-%constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}. Erosion rates $\dot{E}_{sr}$ due to short-range processes are
-%implemented by assuming that transport is linearly proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{sr}=K_d \bm{\nabla}^2 h
-%\end{equation}
-%where $K_d$ is the effective diffusivity. In addition, the slope is modified if it still exceeds a threshold angle after long- and short-range processes
-%\citep{Ueda2015}.
-
-%Offshore sedimentation processes are implemented similarly to hillslope processes but with a depth-dependent coefficient \citep{Sacek2012,Beucher2020}.
-%The sedimentation rate $\dot{S}$ is calculated as
-%\begin{equation}
-%\dot{S}=K_s \bm{\nabla}^2 h
-%\end{equation}
-%with
-%\begin{equation}
-%K_s=K^*_s \exp\left(\frac{h-h_{sl}}{h_0}\right)
-%\end{equation}
-%where $K^*_s$ is a constant, $h_{sl}$ is the sea level and $h_0$ is a characteristic length. In case of sticky air, slope is determined by means of the markers
-%chain, which is vertically corrected considering erosion and sedimentation rates. After the correction, continental markers above the markers chain are
-%transformed in air markers and air markers below the markers chain become sediments. Similarly, slope is determined considering the nodes on the top boundary
-%in case of true free surface and they are vertically corrected as for the markers chain. In this case, continental markers above the top boundary are simply
-%deleted.
-
-\section{Non-linear rheologies}\label{sec:nonlinear}
-Non-linear rheologies are implemented combining viscous creep (dislocation and diffusion) and plastic yielding. For each marker, diffusion and dislocation
-viscosity ($\eta_{df|ds}$) can be determined as follows \citep{Karato1993,Warren2008b,Wang2016}:
-\begin{equation}\label{eq:viscous}
-\eta_{df|ds}=f_s \left(\frac{d^m}{A}\right)^{\frac{1}{n}} \left (\frac{2^{\frac{1-n}{n}}}{3^{\frac{n+1}{2n}}} \right)
-I_2^{\frac{(1-n)}{n}}\exp\left(\frac{Q+pV}{nRT}\right)
-\end{equation}
-where $f_s$ is a scaling factor used to represent lithologies that are stronger or weaker than the base set, $d$ is the grain size, $m$ is the grain size exponent,
-$n$ is the stress exponent, $A$ is the uniaxial pre-exponential factor, $I_2$ is the square root of the second invariant of the strain rate tensor, $Q$ is the
-activation energy, $p$ is the pressure, $V$ is the activation volume, $R$ is the gas constant and $T$ is the temperature.
-Pressures and temperatures are determined by the interpolation of the nodal parameters, while strain rates are calculated by means of the
-derivative of the velocities on the nodes. In case of diffusion creep $n=1$ and $m>0$, in case of dislocation creep $n>1$ and $m=0$. The viscous creep $\eta_{cp}$
-is then calculated as the harmonic average between $\eta_{df}$ and $\eta_{ds}$:
-\begin{equation}\label{eq:visco_creep}
-\eta_{cp}=\left(\frac{1}{\eta_{df}}+\frac{1}{\eta_{ds}}\right)^{-1}
-\end{equation}
-The implementation of the non-linear viscous creep viscosity depending on the strain rate is tested by means of the slab detachment benchmark
-(Section \ref{sec:slab}).
-
-Plastic yielding is implemented rescaling $\eta_{cp}$ in order to limit the stress below the yield stress $\sigma_y$ \citep{Thieulot2008,Thieulot2014,Glerum2018},
-obtaining
-\begin{equation}\label{eq:plastic}
-\eta_{pl}=\frac{\sigma_y}{2I_2}
-\end{equation}
-where $\eta_{pl}$ is the plastic viscosity and the yield stress is determined following the Drucker-Prager criterion, such as
-\begin{equation}\label{eq:yield}
-\sigma_y=C\;\cos(\phi)+p\;\sin(\phi)
-\end{equation}
-where $C$ is the cohesion and $\phi$ is the internal friction angle. In case of negative pressure it is imposed equal to 0, so that negative yield stress are
-excluded. The correctness of the non-linear solution in case of plastic viscosity with variable internal friction angle is verified performing indenter and
-brick experiments (Sections \ref{sec:indenter} and \ref{sec:brick}, respectively).
-
-Plastic yielding and viscous creep are then combined to obtain a viscoplastic viscosity $\eta_{vp}$ as follows, assuming that they are independent processes
-\citep{Karato2008,Glerum2018}:
-\begin{equation}\label{eq:viscoplastic}
-\eta_{vp}=\min(\eta_{cp},\eta_{pl})
-\end{equation}
-Finally, effective viscosity $\eta_{eff}$ is capped by the minimum and the maximum viscosity ($\eta_{min}$ and $\eta_{max}$, respectively) to avoid extremely
-low or high viscosity \citep{Glerum2018} as follows
-\begin{equation}\label{eq:effective}
-\eta_{eff}=\min(\max(\eta_{vp},\eta_{min}),\eta_{max})
-\end{equation}
-Viscous creep and plastic yielding are non-linear rheologies because of their dependence on the velocity field through pressure and strain rates. Therefore,
-the solution is determined by means of Picard-type iterations, until convergence of the velocity field \citep{Glerum2018}. The convergence is verified at each
-iteration $i$ via the nonlinear residual $\bm{\mathcal{R}}^i$ that can be determined as
-\begin{equation}\label{eq:convergence}
-\bm{\mathcal{R}}^i= \mathbb{K} (\eta_{eff}(\dot{\bm{\epsilon}}^{i-1},p^{i-1}))\cdot \bm{v}^{i-1} - \bm{f}^i
-\end{equation}
-%\begin{equation}\label{eq:convergence}
-%\chi=1-\frac{\langle (\bm{v}^i-\langle\bm{v}^i\rangle)\cdot (\bm{v}^{i+1}-\langle\bm{v}^{i+1}\rangle) \rangle}{\sqrt{\sigma^i\sigma^{i+1}}}
-%\end{equation}
-%Iterations are performed until either convergence ($\chi<tol$) or maximal number of non-linear iterations ($it_{max}$) is reached.
-where $\mathbb{K}$ is the Stiffness matrix and $\bm{f}$ is the right hand side vector \citep{Spiegelman2016,Glerum2018}.  The $L_2$-norm is extracted from
-$\bm{\mathcal{R}}^i$ and it is normalised as follows
-\begin{equation}\label{eq:convergence_res}
-\frac{||\bm{\mathcal{R}}^i||_2}{||\bm{\mathcal{R}}^0||_2}
-\end{equation}
-where $||\bm{\mathcal{R}}^0||_2$ is the $L_2$-norm of the first nonlinear iteration.
-Since often the normalised nonlinear residual drop very quickly over the first 2-3 iterations, also the $L_2$-norm of velocity and pressure residuals are
-calculated as
-\begin{equation}\label{eq:convergence_vel}
-\frac{||\Delta\bm{v}||_2}{||\bm{v}||_2}=\frac{||\bm{v}_i-\bm{v}_{i-1}||_2}{||\bm{v}_{i}||_2}
-\end{equation}
-and
-\begin{equation}\label{eq:convergence_press}
-  \frac{||\Delta p||_2}{||p||_2}=\frac{||p_i-p_{i-1}||_2}{||p_{i}||_2}
-\end{equation}
-respectively. Iterations are performed until either maximal number of non-linear iterations ($it_{max}$) is reached or all the normalised nonlinear residual, 
-the velocity residual and the pressure residual converge under a defined tolerance $tol$.
-
-\section{Healing and weakening}
-The temporal evolution of the accumulated strain $\epsilon$ has been implemented as in \citet{Fuchs2019,Fuchs2021}, as following
-\begin{equation}\label{eq:healing}
-  \frac{d\epsilon}{dt}=I_2 - \epsilon H_{rate}(T)
-\end{equation}
-where the first term on the right-hand side is a source term given by the second invariant of the strain rate $I_2$ and the second term a
-temperature-dependent healing factor $H_{rate}$, calculated as
-\begin{equation}\label{eq:healing_rate}
-  H_{rate}(T)=B \exp \left[-\frac{\mu}{2}\left(\frac{1}{T+1}-\frac{1}{2}\right)\right]
-\end{equation}
-where $B$ is a constant describing the time scale of healing, while $\mu$ and $T$ are the non-dimensional temperature activation and temperature,
-respectively \citep{Fuchs2019,Fuchs2021}.
-
-Strain softening is taken into account for both viscous creep and plastic viscosity \citep{Huismans2003,Babeyko2005,Huismans2005,Sobolev2005,Warren2008a} by
-means of the accumulated strain $\epsilon$ memorised by each marker. Softening in the viscous creep determines a linear decrease of $\eta_{vc}$ by means of a
-viscous strain softening factor $W_S$ that increases linearly from $W_{S_0}$ to $W_{S_{\infty}}$ for $\epsilon_{S_0}<\epsilon_S<\epsilon_{S_{\infty}}$ 
-\citep{Warren2008a}. This viscous softening can be related to strain-induced grain size reduction \citep{Warren2008a}. Differently, plastic softening is 
-simulated with a linear decrease of internal friction angle $\phi(\epsilon)$ and cohesion $C(\epsilon)$ in according to
-\begin{eqnarray}
-\label{eq:friction1}\phi(\epsilon)&=&\phi_0+(\phi_{\infty}-\phi_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}\\
-\label{eq:friction2}C(\epsilon)&=&C_0+(C_{\infty}-C_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}
-\end{eqnarray}
-where $\phi_0$, $C_0$ and $\phi_{\infty}$, $C_{\infty}$ are internal friction angle and cohesion for $\epsilon_0$ and $\epsilon_{\infty}$, respectively
-\citep{Huismans2003,Huismans2005,Warren2008a,Thieulot2014}. Plastic softening approximates deformation-induced softening of faults and brittle shear zones
-\citep{Warren2008a}.
-
-\section{The energy equation}\label{sec:en_eq}
-The stabilisation of the advection term of the energy equation (Eq. \ref{eq:energy}) needed to avoid possible oscillations in the thermal solution in the case
-when advection dominates over diffusion, is implemented by means of a streamline-upwind Petrov–Galerkin (SUPG) method, for which the advection term is modified
-follows the discussion in \citet{Thieulot2011} and \citet{Thieulot2014} (see Section \ref{sec:advection} for benchmark). The total internal heat production
-$H_{tot}$ in Eq. \ref{eq:energy} includes radiogenic heating $H_r$, shear heating $H_s$ and adiabatic heating $H_a$, where
-\begin{equation}\label{eq:shear}
-H_s=2\eta \dot{\bm{\epsilon}} : \dot{\bm{\epsilon}} = 2\eta(\dot{\epsilon}_{xx}^2+\dot{\epsilon}_{yy}^2+2\dot{\epsilon}_{xy}^2)
-\end{equation}
-and
-\begin{equation}\label{eq:adiabatic}
-H_a=T \alpha \frac{Dp}{Dt} \approx -\alpha T \rho g_y v
-\end{equation}
-The correctness of the implementation and the computation of shear and adiabatic heating is verified performing an exercise with an analytical solution
-(Section \ref{sec:simple_shear}) and Exercise 9.4 in \citet{Gerya2010} (Section \ref{sec:shear}). Furthermore, mantle convection and thin layer entrainment
-benchmarks (Sections \ref{sec:mantle} and \ref{sec:thin}, respectively) are performed to verify that the code solve correctly Eqs. \ref{eq:momentum_penalty}
-and \ref{eq:energy} in case of a temperature field characterised by an initial perturbation and with a temperature-dependent density. At last, the viscoplastic
-mantle convection benchmark is performed to check also the viscoplasticity in case of a initial perturbed temperature field (Section \ref{sec:visco_mantle}).
-
-\section{Phase transitions and hydration}\label{sec:phase}
-Variations of effective density, specific heat and coefficient of thermal expansion during the evolution of crustal and mantle materials at different
-pressure-temperature ($p$-$T$) conditions are computed by means of version 6.8.6 of Perple\_X software package \citep{Connolly2005}, similarly to the
-implementation described in \citet{Marotta2020}. By default, mineral assemblages and properties of each lithology (oceanic crust, lower and upper continental
-crust, sediments and dry or hydrated mantle) are calculated for temperatures from \SIrange{330}{1600}{\kelvin}, with increments of approximately \SI{4}{\kelvin},
-and pressures from \SIrange{0.1}{30}{\giga\pascal}, with increments of approximately \SI{0.03}{\giga\pascal}, for a total of almost 100000 points.
-
-The introduction of phase transitions produces two main effects that must be taken into account in numerical models: variations in density and release/absorption
-of latent heat \citep{Tackley2010}. Effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into
-account considering effective coefficient of thermal expansion \citep{Christensen1985,Zhong2015}. Similarly, effects of latent heat in the energy equation
-(eq \ref{eq:energy}) are taken into account by considering effective specific heat and coefficient of thermal expansion
-\citep{Christensen1985,Tackley2010,Zhong2015}. Effective density, specific heat and coefficient of thermal expansion are included in Perple\_X files and are
-assumed by each Lagrangian marker in according to its $p$-$T$ conditions.
-
-Hydration processes related to dehydration of subducting lithosphere \citep{Schmidt1998,Liu2007,Faccenda2009,Faccenda2010,Faccenda2014,Rosas2016} strongly
-influence the thermo-mechanics inside the mantle wedge, mainly because of both weakening effects on the mantle rheology and density variations in case of
-mantle serpentinization \citep{Gerya2002,Honda2003,Arcay2005,Roda2010,Regorda2017}. In the case that hydration is switched on, the amount of bound and free
-water is memorised by each marker, following the implementation of \citet{Quinquis2014}; hydration and dehydration processes are related to the amount of bound
-water of each marker and to the maximum amount of water it can transport, i.e. if the amount of bound water exceeds the maximum amount of water, the marker
-dehydrates and releases free water that can hydrates under-saturated markers. Maximum water content of each marker is determined as function of lithology and
-$p$-$T$ conditions and it is calculated using Perple\_X, in the same way than effective density, specific heat and coefficient of thermal expansion. Bound water
-is advected together with the markers, neglecting the effect of bound water diffusion, while free water simply migrates vertically and is not coupled to the
-solid-phase flow of the mantle wedge \citep{Arcay2005,Quinquis2014}. The correctness of the migration scheme of the free water and its absorption into bound
-water by under-saturated markers are verified performing the experiment described in \citet{Quinquis2014} (Sec. \ref{sec:quinquis}).
-
-The mantle viscosity weakening related to the amount of water has been extensively studied \citep{Chopra1981,Hirth2003} and it is implemented in according to
-\citet{Arcay2005} and \citet{Horiuchi2016}, as follows
-\begin{equation}\label{eq:wet_visc}
-\eta_{wet}=\eta_{dry}\left(\left[1-\frac{1}{f_v}\right]\exp\left(-\frac{[OH^-]}{[OH^-]_0}\right)+\frac{1}{f_v}\right)
-\end{equation}
-where $f_v$ is the viscosity reduction factor between dry and wet conditions, $[OH^-]$ is the water content and $[OH^-]_0$ is a reference water content set to
-620 ppm (0.062 wt.\%) \citep{Arcay2005}. By default, $f_v$ is set to 100, in according to \citet{Arcay2005} and \citet{Horiuchi2016}.
-
-\section{Mantle melting}\label{sec:melt}
-Melt fraction in the mantle depends on temperature ($T$, in \textdegree C), pressure ($p$, in \si{\giga\pascal}) and the water content in the melt ($X_{H_2O}$, in wt.\%).
-The determination of the melt fraction $M$ is implemented as explained by \citet{Katz2003}, according to the successive modification by \citet{Langmuir2006} and
-\citet{Kelley2010}, as follows
-\begin{eqnarray}\label{eq:melt_fr}
-M(p,T,X_{H_2O})&=&-T+T_{sol}+(x_t\cdot ln(p+y_t))\cdot X_{H_2O}-\nonumber\\
-&&+K_m\left(\frac{X_{H_2O}^{bulk}}{D_{H_2O}(1-X_{H_2O})+X_{H_2O}}\right)^{\gamma}
-\end{eqnarray}
-where $T_{sol}$ is the temperature of the dry solidus, $D_{H_2O}$ is the partition coefficient and $x_t\cdot log(p+y_t)$ indicates the pressure dependence of
-the melting curve \citep{Kelley2010}. $x_t$, $y_t$, $K_m$ and $\gamma$ are parameters chosen in according to \citet{Kelley2010}.
-However, $X_{H_2O}$ depends on the melt fraction as
-\begin{equation}\label{eq:melt_wt}
-X_{H_2O}(M)=\frac{X_{H_2O}^{bulk}}{D_{H_2O}+M(1-D_{H_2O})}
-\end{equation}
-and a numerical solution can be found using a root-finder method \citep{Katz2003,Wang2016}. Furthermore, the content of water in the melt is limited by the
-pressure-dependent saturation concentration of water in the melt, determined as
-\begin{equation}\label{eq:melt_max}
-X_{H_2O}^{sat}=\chi_1 p^{\lambda_m}+\chi_2 p
-\end{equation}
-where $\chi_1$, $\chi_2$ and $\lambda_m$ are parameters chosen according to \citet{Katz2003}, \citet{Langmuir2006} and \citet{Kelley2010}. The correctness in
-the determination of melt fractions and water contents in the melt are tested comparing model results with experimental curves obtained by \citet{Katz2003}
-(Sec. \ref{sec:katz}).
-
-In case of melting, viscous creep viscosity ($\eta_{vp}$) is modified according to \citet{Wang2016}, as
-\begin{equation}\label{eq:melt_visc}
-\eta_{melt}=\eta_{vp}\cdot \exp(\alpha_m M)
-\end{equation}
-where $\alpha_m$ are melt fraction factors for dislocation and diffusion creep, chosen in according to \citet{Wang2016}.
-
-Effective density ($\rho_{eff}$) for partially molten rocks is calculated as
-\begin{equation}\label{eq:melt_dens}
-\rho_{eff}=\rho_s(1-M)+\rho_m M
-\end{equation}
-where $\rho_s$ and $\rho_m$ are the densities of the solid and the molten rock, respectively \citep{Gerya2004b,Gerya2007,Wang2016}. The density of the
-solid rock is extracted by Perple\_X, while the density of the molten portion is calculated as
-\begin{equation}\label{eq:melt_dens_M}
-\rho_m(p,T)=\rho_0[1-\alpha(T-T_0)][1+\beta(p-p_0)]
-\end{equation}
-where $\rho_0$ is the density at the reference temperature ($T_0$) and pressure($p_0$), and $\alpha$ and $\beta$ are the thermal expansion and compressibility
-coefficients, respectively \citep{Gerya2004b,Gerya2007,Wang2016}.
-
-As for the phase transitions, effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into account
-considering effective coefficient of thermal expansion. Similarly, latent heat due to melting/crystallisation can be implicitly included in the energy equation
-(Eq. \ref{eq:energy}) by considering effective specific heat ($C_{p_{eff}}$) and thermal expansion ($\alpha_{eff}$) of the partially molten rocks, in according to
-\begin{eqnarray}
-\label{eq:cp_effective}C_{p_{eff}}&=&C_p+H_L\left(\frac{\partial M}{\partial T}\right)_p \\
-\label{eq:alpha_effective}\alpha_{eff}&=&\alpha+\rho\frac{H_L}{T}\left(\frac{\partial M}{\partial p}\right)_T
-\end{eqnarray}
-where $C_p$ and $\alpha$ are the specific heat and the coefficient of thermal expansion of the solid rock, respectively, and $H_L$ is the latent heat of the
-molten rock \citep{Gerya2004b,Gerya2007,Tackley2010}.
-
-\section{Benchmarks}\label{sec:benchmark}
-\subsection{Error measurements}\label{sec:error}
-In order to determine the accuracy of velocity and pressure field of the benchmarks, the $L_2$-norm is computed by numerical integration on the Gauss-Legendre
-quadrature points. $L_2$-norm for pressure and velocity errors can be evaluated as
-\begin{eqnarray}
-\textrm{err}_p&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|p^n_i(r_q)-p^a_i(r_q)|^2\right)w_q|J_q|}\nonumber \\
-\textrm{err}_v&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|u^n_i(r_q)-u^a_i(r_q)|^2+|v^n_i(r_q)-v^a_i(r_q)|^2\right)w_q|J_q|}\nonumber
-\end{eqnarray}
-respectively, where $n_e$ is the number of elements, $n_q$ is the number of quadrature points per element, $p^n_i(r_q)$ and $p^a_i(r_q)$ are the numerical and
-analytical pressure, respectively, in each quadrature point $q$, $w_q$  and $J_q$ are the weight and the Jacobian at the quadrature point
-$q$, $u^n_i(r_q)$, $v^n_i(r_q)$, $u^a_i(r_q)$ and $v^a_i(r_q)$ are the numerical and analytical velocities, respectively, in each quadrature point $q$.
-
-Other quantities used as comparison with original benchmarks are the root-mean-square velocity over the whole domain and over the surface
-\begin{eqnarray}
-v_{\textrm{rms}}&=&\sqrt{\int_{\Omega}|\bm{v}|^2 d\Omega}\nonumber \\
-v_{\textrm{rms}}^{top}&=&\sqrt{\int_0^1 u^2 \bigg|_{y=1} dx}\nonumber
-\end{eqnarray}
-respectively, top and bottom Nusselt numbers
-\[\textrm{Nu}^{top/bottom}=-\int_0^1\frac{\partial T}{\partial y} \bigg|_{y=1/y=0} dx\]
-the average rate of work done against gravity
-\[\left\langle W \right\rangle=\int_{\Omega}T u_y d\Omega\]
-the average rate of viscous dissipation
-\[\left\langle \Phi \right\rangle=\int_{\Omega}\tau_{ij}\dot{\epsilon}_{ij} d\Omega\]
-and the percentage error between $\left\langle W \right\rangle$ and $\left\langle \Phi \right\rangle/Ra$
-\[\delta=\frac{\left|\left\langle W \right\rangle -\frac{\left\langle \Phi \right\rangle}{Ra} \right|}{\max \left(\left\langle W \right\rangle,
-\frac{\left\langle \Phi \right\rangle}{Ra}\right)}\times100\]
-
-\subsection{Stokes flow}\label{sec:stokes}
-The problem consists of determining velocity field $(u,v)$ and pressure $p$ in case of a manufactured solution with prescribed body forces such as
-\begin{eqnarray}
-b_1&=&(12-24y)x^4+(-24+48y)x^3+(-48y+72y^2-48y^3+12)x^2+\nonumber \\
-&&+(-2+24y-72y^2+48y^3)x+1-4y+12y^2-8y^3\nonumber \\
-b_2&=&(8-48y+48y^2)x^3+(-12+72y-72y^2)x^2+(4-24y+48y^2-48y^3+\nonumber \\
-&&+24y^4)x-12y^2+24y^3-12y^4\nonumber
-\end{eqnarray}
-for which the exact solution is:
-\begin{eqnarray}
-u(x,y)&=&x^2(1-x)^2(2y-6y^2+4y^3)\nonumber \\
-v(x,y)&=&-y^2(1-y)^2(2x-6x^2+4x^3)\nonumber \\
-p(x,y)&=&x(1-x)-\frac{1}{6}\nonumber
-\end{eqnarray}
-The domain is a unit square a constant viscosity ($\eta=1$) and the penalty parameter is set to $\lambda=10^7$. Velocity boundary conditions are set to no slip
-$(\bm{v}=\bm{0})$ on all boundaries. The problem is performed for different grid resolution between \num{8x8} and \num{1024x1024} elements.
-
-The errors between the analytical solution and the numerical prediction of the velocity and the pressure field are calculated as in Section \ref{sec:error}.
-Fig. \ref{fig:errors}a shows that both velocity and pressure field converge to the exact solution with the decrease of the element size, following the
-theoretical convergence rate. The convergence of pressure, in contrast with the observation by \citet{Donea2003} for $Q_1 \times P_0$ elements, indicate the
-effectiveness of the smoothing procedure. Smoothed pressure field for grid resolution of \num{128x128} elements is shown in Fig. \ref{fig:errors}b (continuous
-red line), in comparison with the analytical solution (dashed blue line). Solve times and memory usage needed to generate the solution are shown in
-Fig. \ref{fig:MUMPS}. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Solver/Stokes_Flow}.
-
-\begin{figure}
-\centering
-\includegraphics[width=\linewidth]{./Figures/errors.pdf}
-\caption{Velocity and pressure error for the Stokes flow experiment between generated and analytical solution as a function of element size (panel a),
-and comparison between smoothed pressure and analytical solution as function of $x$ coordinate for a grid resolution of \num{128x128} elements (panel b).}
-\label{fig:errors}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\linewidth]{./Figures/MUMPS.pdf}
-\caption{Panel a) Analysis, factorisation, solution and total solve times (panel a) and factorisation memory usage (panel b) as a function of the total number
-of degrees of freedom for the Stokes flow experiment.}
-\label{fig:MUMPS}
-\end{figure}
-
-\subsection{Markers advection}\label{sec:runge}
-The 2nd-order Runge-Kutta advection scheme in space is tested by means of the Zalesak disk test \citep{Zalesak1979,Thieulot2014}. The benchmark is performed
-in a unit square domain with a grid resolution of \num{32x32} elements and 1000000 markers and values of Courant number between 0.25 and 1.
-At $t=0$, the disk is centred at position $(0.5;0.75)$ with a radius $R=0.15$ and has a vertical fissure 0.05 wide and 0.2 high (orange in Fig. \ref{fig:runge})
-and with the velocity field prescribed in the entire domain as
-\begin{eqnarray}
-u(x,y)&=&2\pi\left(y-\frac{L_x}{2}\right)\nonumber \\
-v(x,y)&=&-2\pi\left(x-\frac{L_x}{2}\right)\nonumber
-\end{eqnarray}
-the markers are back to their initial location after a $2\pi$ rotation ($t=1$). The clockwise rotation of the disk at $t=0.25$, 0.5 and 0.75 
-is shown in Fig. \ref{fig:runge} (red, blue and green, respectively). The distance from the centre of 1 marker is calculated throughout the rotation to
-evaluate the error for different values of Courant number. As expected, the error increase with the increasing of the time step, as shown in 
-Fig. \ref{fig:runge_err}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/Zalesak_Disk}.
-
-\begin{figure}
-\centering
-\includegraphics[width=200px]{./Figures/Zalesak.pdf}
-\caption{Position of the Zalesak disk at $t=0$, 0.25, 0.5 and 0.75 (orange, red, blue and green, respectively) for the case with Courant number of 1.}
-\label{fig:runge}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/zal_marker.pdf}
-\caption{Distance from the centre as function of time for values of Courant number of  of 0.25, 0.3, 0.5, 0.75 and 1 (green, orange, blue, black and red,
-respectively). R indicates the distance from the centre considering a perfect circle trajectory.}
-\label{fig:runge_err}
-\end{figure}
-
-\subsection{Conservative Velocity Interpolation (CVI)}\label{sec:CVI}
-The CVI correction is checked by means of the Stokes flow experiment by \citet{Donea2003}, as presented in \ref{sec:stokes}. The advection of Lagrangian
-markers is performed using either a 2nd-order or a 4th-order Runge-Kutta scheme with CFL=0.5 and an initial random distribution of 25 markers per element.
-Fig. \ref{fig:CVI} shows the minimum and maximum values of markers per element throughout the experiment with and without the CVI correction and using
-different Runge-Kutta schemes. The introduction of the CVI correction allows the
-tests to run for times longer than $t=2000$ without show empty elements, while tests without the CVI correction stop at approximately $t=600$, when at least
-one element goes empty. Fig. \ref{fig:CVI_time} shows the different number of markers per element and the markers distribution at $t=600$ obtained without and
-with the CVI correction (panels b and c, respectively) in case of a 2nd-order Runge-Kutta scheme. All data can be found at
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/CVI}.
-\begin{figure}
-\centering
-\includegraphics[width=\linewidth]{./Figures/CVI.pdf}
-\caption{Maximum and minimum number of markers per element throughout the experiment, with (green and orange lines) or without (purple and blue lines)
-the CVI correction and using either a 2nd-order (blue and orange lines) or a 4th-order (purple and green lines) Runge-Kutta scheme.}
-\label{fig:CVI}
-\end{figure}
-\begin{figure}
-\centering
-\includegraphics[width=\linewidth]{./Figures/CVI_time.pdf}
-\caption{Number of markers per element and markers distribution through time using 2nd-order Runge-Kutta scheme.
-Panels a: number of markers per element (panel a1) and distribution of the markers (panel a2) after the first
-iteration; panels b and c: comparison between number of markers per element (panels b1 and c1) and markers distribution (panels b2 and c2) without and with
-the CVI correction at $t=600$; panels d: number of markers per element (panel d1) and markers distribution (panel d2) in case of CVI correction at $t=2000$.}
-\label{fig:CVI_time}
-\end{figure}
-
-\subsection{Poiseuille flow}\label{sec:poiseuille}
-The domain is a rectangle with $L_x=2$ and $L_y=1$ and constant density and viscosity ($\rho=1$ and $\eta=1$), gravity acceleration $\bm{g}=\bm{0}$ and penalty
-parameter $\lambda=10^8$. The grid is composed by \num{40x20} elements. Velocity boundary conditions are set to no slip $(\bm{v}=\bm{0})$ at the top and the
-bottom, and a parabolic profile is imposed on the sides, with $u=y(1-y)$ and $v=0$. The analytical solution is then given by:
-\begin{eqnarray}
-u(x,y)&=&y(1-y)\nonumber \\
-v(x,y,)&=&0\nonumber \\
-p(x,y)&=&2\eta\left(\frac{L_x}{2}-x\right)\nonumber
-\end{eqnarray}
-
-The velocity field predicted by the model follows the expected parabolic profile (Fig. \ref{fig:poiseuille}a and Fig. \ref{fig:poi_plot}a) and, in case of the
-classic penalty method with no iterations, the pressure is clearly related to the divergence by means of the penalty parameter (Fig. \ref{fig:poiseuille}b and
-c and Fig. \ref{fig:poi_plot}b). Fig. \ref{fig:poiseuille}d shows that one Uzawa iteration is sufficient to bring the divergence down to \num{e-15}, with no
-correlation with the pressure field. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Poiseuille%20Flow}.
-
-\begin{figure}
-\includegraphics[width=\linewidth]{./Figures/Poiseuille.pdf}
-\caption{Velocity field (panel a), pressure (panel b) and divergence velocity of a Poiseuille flow in case of the classic penalty method (no iterations) and
-after one Uzawa iteration (panel c and d, respectively).}
-\label{fig:poiseuille}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\linewidth]{./Figures/Analytical.pdf}
-\caption{Velocity (panel a) and pressure (panel b) field predicted by the model for a Poiseuille flow with respect with their analytical solutions. The
-velocity field is plotted as function of the vertical coordinate in $L_x/2$ and the pressure is plotted as function of the x coordinate in $L_y/2$.}
-\label{fig:poi_plot}
-\end{figure}
-
-\subsection{Instantaneous 2D sphere}\label{sec:ist_sphere}
-The domain is a unit square with gravity $\bm{g}=(0,-1)$. The fluid has constant density and viscosity ($\rho_f=1$ and $\eta_f=1$). The sphere is in the middle
-of the domain with a radius R = 0.123456798 and has constant density and viscosity ($\rho_s=10^{-2}+\rho_f$ and $\eta_s=10^3 \cdot \eta_f$).
-We distinguish three different types of velocity boundary conditions:
-\begin{enumerate}
-\item FS: free slip conditions on all sides.
-\item NS: no slip conditions on all sides.
-\item OT: free slip conditions on the sides and the bottom, and open boundary at the top.
-\end{enumerate}
-All three cases are performed for different grid resolutions, between \num{16x16} and \num{512x512} elements with 50 randomly distributed markers per element,
-and different average schemes for the viscosity (harmonic, geometric and arithmetic). The velocity in the centre of the sphere $\bm{v}(0.5;0.5)$), minimum and
-maximum velocities ($u_{min}$, $u_{max}$, $v_{min}$, $v_{max}$) and pressures ($p_{min}$, $p_{max}$) on the entire domain, $v_{\textrm{rms}}$ and average pressure on the
-entire domain ($p_{avg}$) are compared with solutions generated by ASPECT \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and all results
-can be found at \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere2D} and 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Instantaneous_2D_Sphere}. Results in terms of the $v_{\textrm{rms}}$ for NS model 
-are also shown in Fig. \ref{fig:inst_sphere}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/vrms_NS.pdf}
-\caption{$v_{\textrm{rms}}$ as function of element size for the instantaneous 2D sphere experiment in case of no slip boundary conditions and with different average
-schemes for the viscosity. Results are compared with results obtained by other numerical codes.}
-\label{fig:inst_sphere}
-\end{figure}
-
-\subsection{Rayleigh-Taylor instability}\label{sec:rayleigh}
-This problem has been originally presented by \citet{vanKeken1997} and here it is performed as the isoviscous case (Case 1a). The domain has $L_x=0.9142$ and
-$L_y=1$ and gravity $\bm{g}=(0,-1)$. Two fluids with same constant viscosities ($\eta_1=\eta_2=1$) and different densities ($\rho_1=1000$ and $\rho_2=1010$),
-with the lighter fluid at the bottom. The initial interface between the fluids is given by $y(x)=0.2+0.02 \cos \left(\frac{\pi x}{L_x}\right)$. The experiment
-is performed with different grid sizes (\num{50x50}, \num{80x80}, \num{100x100} and \num{256x256}). A total of 1960000 markers are randomly distributed at the
-beginning of the simulation. Velocity boundary conditions are set to no slip at the top and the bottom, and to free slip at the sides of the domain.
-
-$v_{\textrm{rms}}$ as function of time is reported in Fig. \ref{fig:RT}, matching well with results shown by \citet{vanKeken1997}, \citet{Tackley2003} and
-\citet{Thieulot2014}. Fig. \ref{fig:rayleigh} shows the evolution of the experiment at different time steps. All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Rayleigh_Taylor_experiment/ISOVISCOUS}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/RT.pdf}
-\caption{$v_{\textrm{rms}}$ of the Rayleigh-Taylor experiment as function of time for different resolution of the grid.}
-\label{fig:RT}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=300px]{./Figures/Rayleigh.pdf}
-\caption{Evolution of the Rayleigh-Taylor experiment for a grid of \num{256x256} elements at $t=0, 500, 1000$ and $2000$ (panels a, b, c and d, respectively).}
-\label{fig:rayleigh}
-\end{figure}
-
-\subsection{Falling blocks}\label{sec:block}
-This benchmark is described as proposed by \citet{Gerya2003a}, \citet{Gerya2010} and \citet{Thieulot2011}. The domain is square with $L_x=L_y=$ \SI{500}{\km}
-and the grid is composed by \num{50x50} elements with 25 markers in each element. The block is initially centred at ($x=$ \SI{250}{\km}; $y=$ \SI{400}{\km})
-and has a size of \num{100x100} km (Fig. \ref{fig:block}a). The fluid surrounding the block has $\eta_f=$ \SI{e21}{\pascal\per\s} and $\rho_f=$
-\SI{3200}{\kg\per\cubic\m}. The benchmark tests with different viscosities of the block, with $\eta_b$ from \SIrange{e15}{e27}{\pascal\per\s}. In each
-experiment also the density $\rho_b$ of the block varies from \SIrange{3220}{9900}{\kg\per\cubic\m}. Velocity boundary conditions are set to free slip
-conditions on all sides of the domain. Density distributions at $t=$ \SI{20}{\mega\year} for $\rho_b=$ \SI{3300}{\kg\per\cubic\m} and different $\eta_b$
-are plotted in Fig. \ref{fig:block}b-f, showing that the code correctly preserve the block geometry in case of large viscosity contrast, with a block stiffer
-than the surrounding fluid (Fig. \ref{fig:block}f).
-
-The velocity in the centre of the falling block at $t=0$ is measured for all experiments and, since the velocity should increase with density contrast, the
-quantity $\bm{v}/(\rho_b-\rho_f)$ is plotted as function of the viscosity contrast. All results perfectly match with those from \citet{Gerya2010} and they
-line up on a single curve, demonstrating that the code can correctly deal with large viscosity and density contrasts (Fig. \ref{fig:falling}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Falling%20blocks}.
-
-\begin{figure}
-\includegraphics[width=300px]{./Figures/Block.pdf}
-\centering
-\caption{Density evolution of the falling block experiment at $t=$ \SI{0}{\mega\year} (panel a) and $t=$ \SI{20}{\mega\year} for different viscosities of the
-block (panels b-f).}
-\label{fig:block}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Falling.pdf}
-\caption{Initial velocity relative to the density contrast at the centre of the falling block as function of the viscosity contrast between the block and the
-surrounding fluid.}
-\label{fig:falling}
-\end{figure}
-
-\subsection{2D time-dependent Stokes sphere with deformable free surface}\label{sec:sphere}
-This experiment is performed in a unit square domain with the gravity acceleration fixed to $g_y=-1$. The sphere has $\rho_s=2$ and $\eta_s=10^3$ and is
-initially centred at ($0.5;0.6$) with radius $R=0.123456789$. The fluid surrounding the sphere has $\rho_f=1$ and $\eta_f=1$ and occupies the domain for
-$y\leq 0.75$, while for $y>0.75$ the air has $\rho_a=0$ and $\eta_a=10^{-3}$. Velocity boundary conditions are set to free slip on all sides. The Courant
-number is set to 0.25. The experiments run for 200 s using grid resolutions from \num{150x150} to \num{512x512} elements, with 25 randomly distributed markers
-per element, and for different average schemes for the viscosity. The interface between the fluid and the air is tracked by means of the markers chain. Results
-in terms of velocity and pressure fields and topography variations are compared with results obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and they can be found at
-\url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere_fs2D} and
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Time_dependent_sphere}. Results in terms of the $v_{\textrm{rms}}$ in case of 
-an arithmetic average are also shown in Fig. \ref{fig:inst_spherefs}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/vrms_arithm.pdf}
-\caption{$v_{\textrm{rms}}$ for different grid resolutions as function of time for the 2D time-dependent Stokes sphere experiment in case of an arithmetic average.
-Results are compared with results obtained by ASPECT.}
-\label{fig:inst_spherefs}
-\end{figure}
-
-\subsection{Free surface stabilisation}\label{sec:stab}
-The benchmark is performed as discussed in \citet{Kaus2010a} and \citet{Thieulot2014}. The domain is square with $L_x=L_y=$ \SI{500}{\km} and a grid resolution
-of \num{200x200} elements, each of them containing 16 markers. A buoyant fluid with $\rho_1=$ \SI{3200}{\kg\per\cubic\metre} and $\eta_1=$ \SI{e20}{\pascal\s}
-is overlain by a denser fluid with $\rho_2=$ \SI{3300}{\kg\per\cubic\metre} and $\eta_2=$ \SI{e21}{\pascal\s}. The initial interface between the fluids has an
-initial sinusoidal shape of 5 km amplitude. Velocity boundary conditions are set to free slip conditions at the sides and no slip at the bottom, while the top
-is free surface. The experiment is performed with various fixed time steps, without the use of the Courant number. The vertical position of the free surface at
-$x=L_x$ is tracked for each simulation, with or without the stabilisation algorithm.
-
-The results show an instability (drunken sailor effect) increasing the time step in case that the stabilisation algorithm is not activated
-($dt=$ \SI{4200}{\year}, green line in Fig. \ref{fig:stab}), as already observed by \citet{Kaus2010a} and \citet{Thieulot2014}. Activating the algorithm the
-instability is fixed and the simulation remains stable using time step up to \SI{50000}{\year} (red line in Fig. \ref{fig:stab}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Stabilisation_algorithm}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Kaus_random.pdf}
-\caption{Evolution of the y coordinate in $x=L_x$ as function of time for different time steps, with and without the stabilisation algorithm.}
-\label{fig:stab}
-\end{figure}
-
-\subsection{Topography relaxation}\label{sec:crameri}
-This experiment is performed as in \citet{Crameri2012} using free surface and sticky air (Case 1). The domain is rectangular with $L_x=$ \SI{2800}{\km} and
-$L_y=$ \SI{700}{\km} in case of free surface and $L_y=$ \SI{800}{\km} in case of sticky air ($h_{st}=$ \SI{100}{\km}). The grid is composed by \num{256x64}
-elements for the free surface \num{560x320} elements for the sticky air. For both cases, there is a mantle of 600 km thickness with $\eta_m=$ \SI{e21}{\pascal\s}
-overlain by a highly viscous cosine-shaped lithosphere with thickness between 93 and 107 km and $\eta_l=$ \SI{e23}{\pascal\s}. Both mediums have $\rho=$
-\SI{3300}{\kg\per\cubic\metre}. In case of sticky air experiment, the air has $\rho_a=$ \SI{0}{\kg\per\cubic\metre} and $\eta_a=$ \SI{e18}{\pascal\s}. The
-gravity acceleration is set to $g_y=$ \SI{-10}{\metre\per\square\s}. Velocity boundary conditions are set to no slip at the bottom and free slip at the sides
-of the domain. In case of sticky air, velocities on top are set to free slip conditions.
-
-The maximum topography as function of time ($h(t)$) can be analytically derived in according to \[h(t)=h_0 \exp(\gamma t)\]
-where $h_0=7$ km is the initial topography, $\gamma=\num{-0.2139e-11}$ is the characteristic relaxation rate and $t$ is the time (black line in Fig.
-\ref{fig:crameri}). Maximum topography at the characteristic relaxation time $t_{rlx}=$ \SI{14.825}{\kilo\year} can be found to be $h_{rlx}=$ \SI{2576}{\m}.
-The results show that a viscosity of the air of \SI{e18}{\pascal\s} is needed in case of $h_{st}=$ \SI{100}{\km} to correctly describe the topography relaxation
-for this problem, as pointed out by \citet{Crameri2012} (purple line in Fig. \ref{fig:crameri}). The case with a true free surface well follows the analytical
-solution (orange line in Fig. \ref{fig:crameri}). In particular, at the relaxation time in case of the true free surface predicts a maximum topography of
-\SI{2570}{\m}, with an error of \SI{6}{\m} with respect to the analytical solution. Results are also compared with results obtained with a free surface in
-UNDERWORLD (\num{256x64} elements) and SULEC (\num{401x201} elements) (with courtesy from \citealp{Crameri2012}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Topography_relaxation}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Crameri.pdf}
-\caption{Maximum topography as function of time for the Crameri benchmark (coloured lines) in comparison with the analytical solution (black line) and results
-obtained with UNDERWORLD (yellow line) and SULEC (blue line).}
-\label{fig:crameri}
-\end{figure}
-
-\subsection{Spontaneous subduction}\label{sec:subduction}
-This experiment is performed as presented by \citet{Schmeling2008}, in case of both a sticky air and a true free surface and using different average schemes
-for the viscosities. The domain is rectangular with $L_x=$ \SI{3000}{\km} and $L_y=$ \SI{700}{\km}, a grid resolution of \num{256x64} elements with 50 markers
-per element and Courant number of 0.01. At the beginning of the simulation, a 100 km-thick lithospheric layer with $\rho_m=$ \SI{3300}{\kg\per\cubic\metre}
-and $\eta_m=$ \SI{e23}{\pascal\s} is located at the top of the domain for $x$ between 1000 and 3000 km, while the asthenospheric mantle has $\rho_m=$
-\SI{3200}{\pascal\s} and $\eta_m=$ \SI{e21}{\pascal\s}. In addition, a 200 km-depth lithospheric slab is already subducted in the mantle in order to have
-a spontaneous subduction. In case of sticky air, $L_y=750$ km and a 50 km-thick air layer with $\eta_a$ from \SIrange{e19}{e21}{\pascal\s} overlie the mantle.
-Velocity boundary conditions are set to free slip on the sides and on the bottom of the domain. In case of sticky air, velocities on the top boundary are set
-to free slip conditions as well.
-
-The maximum depth of the slab is tracked for all the simulations (Fig. \ref{fig:subduction}) and they are compared with results obtained by
-\citet{Schmeling2008} with models with regular grids and comparable grid resolutions (rectangles in Fig. \ref{fig:subduction}). Low viscous sticky air models
-(i.e., \SI{e19}{\pascal\s}) enlighten a strong dependence of the sinking velocity with the chosen average scheme (dashed lines in Fig. \ref{fig:subduction}),
-while high viscous sticky air models (i.e., \SI{e21}{\pascal\s}) show a higher resistance at the trench, underestimating the correct solution (dotted lines in
-Fig. \ref{fig:subduction}). A high resistance at the trench can be also observed in case of a true free surface for the low resolution grid used (continuous
-lines in Fig. \ref{fig:subduction}), as pointed out by \citet{Schmeling2008}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Spontaneous_subduction}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Subduction.pdf}
-\caption{Maximum depth of the slab as function of time for all the simulations of the \citet{Schmeling2008} benchmark, in case of harmonic, geometric and
-arithmetic means (red, green and blue lines, respectively) using a sticky air or a true free surface (discontinuous and continuous lines, respectively).
-Red, green and blue rectangular areas indicate the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{500}{\km} in case of sticky air,
-while the grey rectangular area indicates the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{600}{\km} in case of true free surface.}
-\label{fig:subduction}
-\end{figure}
-
-\subsection{The slab detachment}\label{sec:slab}
-The slab detachment benchmark is performed as described by \citet{Schmalholz2011} and \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{1000}{\km}
-and $L_y=$ \SI{660}{\km} and a grid resolution of \num{256x256} elements. A non-linear viscous T-shaped layer with $\rho_l=$ \SI{3300}{\kg\per\cubic\metre} is
-placed at the top of the domain and surrounded by a linear viscous fluid with $\rho_f=$ \SI{3150}{\kg\per\cubic\m}. The top layer is 80 km-thick and a 250
-km-long and 80 km-wide slab is placed at $x=L_x/2$ (Fig. \ref{fig:slab}a). The effective viscosity of the top layer is given by
-\[\eta_{eff}=\eta_0 I_2^{\frac{(1-n)}{n}}\]
-with $\eta_0=$ \SI{4.75e11}{\pascal\s} and $n=4$, while the surrounding fluid has $\eta_f=$ \SI{e21}{\pascal\s}. The effective viscosity is capped between
-$\eta_{min}=$ \SI{e21}{\pascal\s} and $\eta_{max}=$ \SI{e25}{\pascal\s}. Velocity boundary conditions are set to free slip at the top and the bottom, and to
-no slip at the sides of the domain.
-
-The sinking of the slab determines a decrease of the effective viscosity in the T-shaped layer during the first part of the evolution (Fig. \ref{fig:slab}a),
-as consequence of local strain rates, while effective viscosities increase up to $\eta_{max}$ in the top layer after approximately \SI{20}{\mega\year}, when
-necking is complete (Fig. \ref{fig:slab}b).
-
-The width of necking is tracked by means of markers position at the side of the slab. Necking width and time are normalised in function of the initial slab
-(\SI{80}{\km}) and the characteristic time ($t_c=$ \SI{7.1158e14}{\s}) \citep{Schmalholz2011,Glerum2018}. The results in case of arithmetic, geometric and
-harmonic mean of the viscosity are shown in Fig. \ref{fig:necking}, compared with results from \citet{Glerum2018} for which an infinity norm mean was used.
-The influence of the average schemes agrees with results shown in Section \ref{sec:subduction}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Slab_detachment}.
-
-\begin{figure}
-\includegraphics[width=\linewidth]{./Figures/Slab.pdf}
-\caption{Effective viscosity for the slab detachment benchmark at the beginning of the evolution (panel a) and when necking is complete (panel b).}
-\label{fig:slab}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Necking.pdf}
-\caption{Normalised width of the necking of the slab detachment benchmark as function of normalised time.}
-\label{fig:necking}
-\end{figure}
-
-\subsection{The indenter experiment}\label{sec:indenter}
-The indenter benchmark simulate a rigid punch on a purely plastic von Mises material, for which exists an analytical solution
-\citep{Thieulot2008,Thieulot2014,Glerum2018}. The domain is a unit square with a grid resolution of \num{256x256} elements. The gravity acceleration is set
-to $\bm{g}=\bm{0}$ and effective viscosity is capped using $\eta_{min}=10^{-4}$ and $\eta_{max}=10^3$. The tolerance for the convergence of the non-linear
-solution is $tol=10^{-9}$, with a maximum number of non-linear iterations set to 500. The medium in the domain has $\rho=0.01$, an initial viscosity
-$\eta_0=10$, cohesion $C=1$ and an angle of internal friction $\phi=0$°. Velocity boundary conditions are set to no slip at the bottom and free slip at the
-sides of the domain. The top of the domain is open with the exception of the central portion (punch area) with width $w_p=0.125$, where $v=-1.05$ and $u$ is
-fixed either to 0 (rough punch experiment) or free (smooth punch experiment).
-
-Fig. \ref{fig:indenter} show the results in terms of viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure
-(panels g and h) for both experiments. The analytical solution indicate that pressure at ($0.5;1$) and ($0.5 \pm w_p;1$)is $p=\pi +1$ and $p=1$, respectively
-(black and red lines in Fig. \ref{fig:smooth_p}). Fig. \ref{fig:smooth_p} shows a clear improvement in the pressure solution when passing from a rough to a
-smooth punch (green and blue lines in Fig. \ref{fig:smooth_p}, respectively), as previously observed by \citet{Thieulot2014} and \citet{Glerum2018}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Indenter}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Indenter.pdf}
-\caption{Viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure (panels g and h) fields for rough (left column) and
-smooth (right column) punch experiments.}
-\label{fig:indenter}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Smooth_Rough.pdf}
-\caption{Nodal pressure in function of the x coordinate at $y=1$ for rough (green line) and smooth (blue line) punch experiment. Black and red lines indicate
-the analytical solution at $x=0.5$ and $x=0.5 \pm w_p$, respectively.}
-\label{fig:smooth_p}
-\end{figure}
-
-\subsection{The brick experiment}\label{sec:brick}
-An instantaneous version of the brick benchmark is performed to verify the correctness of the pressure-dependent plasticity for different angle of internal
-friction, as proposed by \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{40}{\km} and $L_y=$ \SI{10}{\km} and a grid resolution of \num{512x128}
-elements. The gravity acceleration is set to $g_y=$ \SI{-10}{\m\square\s} and effective viscosity is capped using $\eta_{min}=$ \SI{e19}{\pascal\s} and
-$\eta_{max}=$ \SI{e26}{\pascal\s}. The tolerance for the convergence of the non-linear solution is $tol=10^{-7}$, with a maximum number of non-linear iterations
-set to 1000. A 800 m-wide and 400 m-height inclusion with $\eta_b=$ \SI{e20}{\pascal\s} and $\rho_b=$ \SI{2700}{\kg\per\cubic\m} is placed at the bottom of the
-domain at $x=L_x/2$. The inclusion is surrounded by a non-linear viscous medium with $\rho_m=$ \SI{2700}{\kg\per\cubic\m}, an initial viscosity of
-\SI{e23}{\pascal\s}, a linear viscous viscosity of \SI{e25}{\pascal\s} and a cohesion of \SI{40}{\mega\pascal}. Velocities are set to free slip conditions at
-the bottom of the domain and the top is open. Velocities on sides of the domain are fixed to $u=$ \SI{+-2e-11}{\m\per\s} and $v=0$. The experiment is performed
-in compressional and extensional contexts with an angle of internal friction $\phi$ of the non-linear medium variable between 0° and 30°.
-
-As expected, two shear bands stem from the inclusion with variable angles in relation to both the dynamics context and the internal friction angle. Shear band
-angles formed at 45° for $\phi=0$° in both extensional and compressional contexts (Fig. \ref{fig:brick_beam}a), while they are different in case of $\phi\neq0$°.
-Shear band angles in case of $\phi=20$° are shown for compression and extension in Fig. \ref{fig:brick_beam}b  and c, respectively. Values of shear band angles
-as function of internal friction angles from 0° to 30° are extracted at different depths and minimum and maximum values are plotted in Fig. \ref{fig:brick},
-in comparison with theoretical Roscoe, Arthur and Coulomb shear band angles. For all tests the tolerance for nonlinear convergence is defined $tol=10^{-7}$
-and none converges before the maximum number of iterations ($it_{max}=1000$) is reached. However, tests with internal friction angles up to 15° show a constant
-decrease of the velocity residuals, which can not be observed for for tests with higher internal friction angles (Fig. \ref{fig:convergence}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Brick_experiment}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Brick_Beam.pdf}
-\caption{Shear band angles predicted for the brick experiment for $\phi=0$° (panel a) and for $\phi=20$° in case of compression and extension (panels b and c,
-respectively).}
-\label{fig:brick_beam}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Brick.pdf}
-\caption{Shear band angles predicted for the brick experiment in case of compressional and extensional contexts as function of different internal angle of
-friction (continuous black line and red dots), compared with theoretical Roscoe, Artur and Coulomb angles (discontinuous black lines). Red lines indicate the
-range of angles calculated at different depths.}
-\label{fig:brick}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=500px]{./Figures/convergence.pdf}
-  \caption{Convergence rates for all tests. Normalised residuals (left column) are compared with velocity residuals (right column) for extensional (first row)
-  and compressional (second row) settings.}
-  \label{fig:convergence}
-  \end{figure}
-
-\subsection{Advection stabilisation}\label{sec:advection}
-The 1D advection problem proposed by \citet{Donea2003} and \citet{Thieulot2011} is performed to verify the effectiveness of the SUPG method to stabilise the
-advection term of the energy equation. The domain is a 1D segment with $L_x=1$ composed by 50 elements and a discontinuity in the thermal field placed at
-$x=0.25$. Temperature is set to 1 for $x \leq 0.25$ and to 0 for $x > 0.25$. Velocity is set to ${u}=1$ in the entire domain. The simulation is performed for
-250 time steps, with $dt=0.002$, so the thermal discontinuity should be at $x=0.75$ at the end of the simulation.
-
-Temperature profiles at the end of the simulation are shown in \ref{fig:advection} as function of the dimensionless coefficient $\gamma=\tau \bm{v}/h$. In case
-of the classic Galerkin method ($\gamma=0$, blue line in \ref{fig:advection}) the final thermal profile is characterised by strong oscillations, which are
-eliminated in case of the SUPG method ($\gamma=0.045$, orange line in \ref{fig:advection}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Advection_stabilisation}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Advection.pdf}
-\caption{Temperature profile as function of \textit{x} for the advection stabilisation benchmark. Purple line indicates the initial temperature profile; the
-green line indicates the analytical temperature profile after 250 time steps; blue line indicates the temperature profile after 250 time steps in case of the
-classic Galerkin method ($\gamma=0$); orange line indicates the temperature profile after 250 time steps in case of the SUPG method ($\gamma=0.045$).}
-\label{fig:advection}
-\end{figure}
-
-\subsection{Shear heating}\label{sec:simple_shear}
-This exercise is performed in an unit square domain composed by \num{128x128} elements. The velocity field is prescribed on the entire domain with
-$\bm{v}=(L_y-y)y\bm{e_x}$; viscosity, density and specific heat are fixed to 1, while thermal conductivity and radiogenic energy are fixed to 0. Therefore,
-the energy equation (Eq. \ref{eq:energy}) can be simplified as \[\frac{\partial T}{\partial t}=H_s\] and fixing $T(t=0)=0$, the temperature can be find as
-\[T(t)=H_s t\]
-In this case we have
-\begin{eqnarray}
-\dot{\epsilon_{xx}}&=&\frac{\partial u}{\partial x}=0 \nonumber \\
-\dot{\epsilon_{yy}}&=&\frac{\partial v}{\partial y}=0 \nonumber \\
-\dot{\epsilon_{xy}}&=&\frac{1}{2}\left(\frac{\partial v}{\partial x}+\frac{\partial u}{\partial y}\right)=\frac{1}{2}(L_y-2y)\nonumber
-\end{eqnarray}
-and, simplifying Eq. \ref{eq:shear}, shear heating can be calculated as \[H_s(x,y)=(1-2y)^2\]
-
-The solution predicted by the model in terms of velocity, temperature and shear heating match well with the analytical solutions (Fig. \ref{fig:analytical_en}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Simple_shear}.
-
-%\begin{figure}
-%\centering
-%\includegraphics[width=400px]{./Figures/Energy_Thieulot.pdf}
-%\caption{Temperature and shear heating predicted by the model after 1 time step.}
-%\label{fig:en_thieulot}
-%\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=500px]{./Figures/Analytical_Shear.pdf}
-\caption{Velocity (panel a), temperature (panel b) and shear heating (panel c) as function of the \textit{y} coordinate for the simple shear experiment. The
-solutions predicted by the model (red lines) are compared with the analytical solutions (dashed blue lines).}
-\label{fig:analytical_en}
-\end{figure}
-
-\subsection{Shear and adiabatic heating}\label{sec:shear}
-This problem is performed as presented in Exercise 9.4 in \citet{Gerya2010}. Two materials are vertically separated in a rectangular domain with $L_x=$
-\SI{1000}{\km}, $L_y=$ \SI{1500}{\km} and a grid resolution of \num{30x20} elements. Constant thermal coefficient expansion ($\alpha=$ \SI{3e-5}{\per\kelvin}),
-temperature ($T=$ \SI{1300}{\kelvin}) and gravity acceleration ($g_y=$ \SI{-10}{\m\square\s}) are assumed in the whole domain. Fluid 1 (on the left side of the
-domain) has $\rho_1=$ \SI{3200}{\kg\per\cubic\m} and $\eta_1=$ \SI{e20}{\pascal\s}; fluid 2 (on the right side of the domain) has $\rho_2=$
-\SI{3300}{\kg\per\cubic\m} and $\eta_2=$ \SI{e22}{\pascal\s}. Velocity boundary conditions are set to free slip on all sides of the domain.
-
-As shown in Fig. \ref{fig:energy}, both shear and adiabatic heating predicted by the code (first row in Fig. \ref{fig:energy}) well recreate the results
-obtained by means of example ${Shear\_adiabatic\_heating.m}$ from \citet{Gerya2010} (second row in Fig. \ref{fig:energy}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Adiabatic%2BShear}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Energy.pdf}
-\caption{Comparison between shear (first column), adiabatic (second column) and total (third column) energy predicted by the code (panels a, b and c) and those
-created using example ${Shear\_adiabatic\_heating.m}$ from Exercise 9.4 in \citet{Gerya2010} (panels d, e and f).}
-\label{fig:energy}
-\end{figure}
-
 \subsection{Mantle convection}\label{sec:mantle}
@@
 acceleration $g_y=10^{10}Ra$. The experiment is performed with three different Rayleigh numbers ($Ra=10^4$, $10^5$ and $10^6$) and with different grid
-resolution (between \num{32x32} and \num{128x128} elements). The fluid has constant viscosity, initial density, heat capacity, thermal conductivity
+resolution (between $32\times32$ and $128\times128$ elements). The fluid has constant viscosity, initial density, heat capacity, thermal conductivity
 ($\eta= \rho_0= C_p = k=1$), reference temperature ($T_0=0$) and thermal expansion coefficient ($\alpha=10^{-10}$). Temperatures are set to 0 on top and 1 on
@@
 & & \multirow{2}{*}{\citet{Blankenbach1989}} & \citet{Thieulot2014} & \multicolumn{3}{c|}{FALCON}  \\
- & & & \num{200x200} & \num{32x32} & \num{64x64} & \num{128x128}  \\
+ & & & $200\times200$ & $32\times32$ & $64\times64$ & $128\times128$  \\
 \midrule
@@
 \end{table}
-
-\subsection{Viscoplastic mantle convection}\label{sec:visco_mantle}
-This benchmark is performed as Cases 1-5a described in \citet{Tosi2015} in a 2D unit square domain with two different grid resolutions of \num{32x32} 
-and \num{100x100} elements. Reference viscosity, reference density, heat capacity, thermal conductivity and thermal expansion coefficient are fixed to a constant
-value of 1 ($\eta_0= \rho_0= C_p = k=\alpha=1$), while the gravity has been chosen as $g_y=10^2$, in order to obtain a Rayleigh number $Ra=10^2$. Temperatures
-are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\].
-The viscosity field $\eta$ is calculated as the harmonic average between a linear part $\eta_{lin}$ that depends on temperature only or on temperature and 
-depth $y$ and a nonlinear-plastic part $\eta_{pl}$ that depends on the strain rate $\dot{\bm{\epsilon}}$, as follows
-\[\eta(T,y,\dot{\bm{\epsilon}})=2\left(\frac{1}{\eta_{lin}(T,y)}+\frac{1}{\eta_{pl}(\dot{\bm{\epsilon}})}\right)^{-1}\]
-The linear and the nonlinear-plastic parts of the viscosity are calculated following \citet{Tosi2015} as follows
-\[\eta_{lin}(T,y)=\exp(-\ln(\Delta\eta_T)T+\ln(\Delta\eta_y)y)\]
-\[\eta_{pl}(\dot{\bm{\epsilon}})=\eta^*+\frac{\sigma_Y}{\sqrt{\dot{\bm{\epsilon}}:\dot{\bm{\epsilon}}}}\]
-where $\Delta\eta_T$, $\Delta\eta_y$, $\eta^*$ and $\sigma_Y$ are parameters chosen as listed in Table \ref{tab:tosi_setup}.
-
-For each case, some diagnostic quantities measured at the steady state are reported in Table \ref{tab:visco_mantle} (see Section \ref{sec:error} for 
-the computation of all the quantities), showing a well fit with results described in \citet{Tosi2015}. Data and charts for all cases can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Viscoplastic_mantle_convection}.
-
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{white!100!}{gray!25!}
-%\begin{table}[H]
-%  \caption{Results of the viscoplastic mantle convection benchmark for Cases 1-4 compared with results obtained with ELEFANT and YACC (from \citealp{Tosi2015}).}
-%  \centering
-%  \resizebox{\linewidth}{!}{
-%\begin{tabular}{|c|ccc|ccc|ccc|ccc|}
-%  \toprule
-%    & \multicolumn{3}{c|}{\textbf{Case 1}} & \multicolumn{3}{c|}{\textbf{Case 2}} & \multicolumn{3}{c|}{\textbf{Case 3}} &
-%    \multicolumn{3}{c|}{\textbf{Case 4}} \\
-%   & FALCON & ELEFANT& YACC & FALCON& ELEFANT& YACC & FALCON & ELEFANT & YACC & FALCON & ELEFANT & YACC \\
-%  \midrule
-%   $\left\langle T \right\rangle$ & 0.7759 & 0.7758 & 0.7767 & 0.6043 & 0.6036 & 0.6058 & 0.7275 & 0.7275 & 0.7286 & & 0.5277 & 0.5289\\
-%    Nu$^{top}$ & 3.4280 & 3.4214 & 3.4298 & 8.5242 & 8.5115 & 8.5278 & 3.0368 & 3.0347 & 3.0374 && 6.5912 & 6.5572 \\
-%    Nu$^{bottom}$ & 3.3168 & 3.313 & 3.3143 & 8.4599 & 8.4465 & 8.3990 & 3.0086 & 2.9908 & 2.9628 && 6.5834 & 6.5243 \\
-%    $v_{\textrm{rms}}$ & 249.901 & 249.134 & 251.7997 & 141.2451 & 140.7087 & 142.2020 & 100.2776 & 100.1208 & 100.9467 && 79.1105 & 79.6202 \\
-%    $v_{\textrm{rms}}^{top}$ & 1.8907 & 1.8642 & 1.8298 & 104.7928 & 104.1228 & 105.1767 & 2.0832 & 2.0652 & 2.0374 && 74.7596 & 75.8414 \\
-%    $u_{max}^{top}$ & 2.6340 & 2.6119 & 2.5516 & 121.9212& 121.766 & 122.3238 & 2.9086 & 2.9019 & 2.8458 && 88.9146 & 89.2940 \\
-%    $\eta_{min}$ & - & - & - & $2.0303\times10^{-5}$ & $2.0600\times10^{-5}$ & $1.9853\times10^{-5}$ & $4.8129\times10^{-5}$ & $4.8080\times10^{-5}$ & $4.7907\times10^{-5}$ &
-%    & $1.9860\times10^{-4}$ & $1.9174\times10^{-4}$ \\
-%    $\eta_{max}$ & - & - & - & 1.3325 & 1.85 & 1.6095 & 0.9360 & 0.9023 & 1 && 1.52 & 1.6773 \\
-%    $\left\langle W \right\rangle$ & 2.4299 & 2.4316 & 2.4583 & 7.5648 & 7.5665 & 7.6222 & 2.0387 & 2.0384 & 2.0400 && 5.6216 & 5.6512 \\
-%    $\left\langle \Phi \right\rangle/Ra$ & 2.4311 & 2.4276 & 2.4333 & 7.5661 & 7.5585 & 7.6194 & 2.0386 & 2.0370 & 2.0335 && 5.6182 & 5.6463 \\
-%    $\delta$ & 0.05\% & 0.16\% & 1.02\% & 0.02\% & 0.11\% & 0.04\% & $<$0.01\% & 0.07\% & 0.32\% && 0.06\% & 0.09\% \\
-%  \bottomrule
-%  \end{tabular}}
-%  \label{tab:visco_mantle}
-%  \end{table}
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{2}{white!100!}{gray!25!}
-\begin{table}[H]
-  \caption{Benchmark cases and corresponding parameters.}
-  \centering
-  \small
-\begin{tabular}{|>{\centering}m{0.08\linewidth}|>{\centering}m{0.05\linewidth}>{\centering}m{0.05\linewidth}>{\centering}m{0.05\linewidth}
-  >{\centering}m{0.05\linewidth} >{\centering}m{0.05\linewidth} |>{\centering\arraybackslash}m{0.2\linewidth}|}
-  \toprule
-    Case & $Ra$ & $\Delta\eta_T$ & $\Delta\eta_y$ & $\eta^*$ & $\sigma_Y$ & Convective regime \\
-  \midrule
-  \rowcolor{gray!25!}
-    1 & $10^2$ & $10^5$ & 1 & - & - & Stagnant lid \\
-    2 & $10^2$ & $10^5$ & 1 & $10^{-3}$ & 1 & Mobile lid  \\
-    \rowcolor{gray!25!}
-    3 & $10^2$ & $10^5$ & 10 & - & - & Stagnant lid  \\
-    4 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 1 & Mobile lid \\
-    \rowcolor{gray!25!}
-    5 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 4 & Periodic \\
-  \bottomrule
-  \end{tabular}
-  \label{tab:tosi_setup}
-  \end{table}
-
-\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{gray!25!}{white!100!}
-\begin{table}
-  \caption{Results of the viscoplastic mantle convection benchmark for different grid resolutions in Cases 1-5 compared with results obtained with ELEFANT
-  and YACC ($100\times100$ elements, from \citealp{Tosi2015}) and in Stone 28 ($32\times32$ elements, from
-  \url{https://github.com/cedrict/fieldstone/tree/master/python_codes/fieldstone_28}).}
-  \centering
-  \resizebox{\linewidth}{!}{
-\begin{tabular}{|cc|ccccccccccc|}
-  \toprule
-  \multicolumn{13}{|c|}{\textbf{Case 1}} \\
-  \midrule
-  Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-  & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-  \hline
-  \rowcolor{gray!25!}
-    YACC & $100\times100$ & 0.7767 & 3.4298 & 3.3143 & 251.7997 & 1.8298 & 2.5516 & - & - & 2.4583 & 2.4333 & 1.02\% \\
-    ELEFANT & $100\times100$ & 0.7758 & 3.4214 & 3.313 & 249.134 & 1.8642 & 2.6119 & - & - & 2.4316 & 2.4276 & 0.16\% \\
-    \rowcolor{gray!25!}
-    Stone 28 & $32\times32$ & 0.7737 & 3.3987 & - & 243.872 & - & - & - & - & - & - & - \\
-%    FALCON & $32\times32$ & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & - & - & 2.4745 & 2.4864 & 0.48\% \\
-%    \rowcolor{gray!25!}
-    FALCON & $100\times100$ & 0.7759 & 3.4280 & 3.3168 & 249.901 & 1.8907 & 2.6340 & - & - & 2.4299 & 2.4311 & 0.05\% \\
-    \midrule
-    \multicolumn{13}{|c|}{\textbf{Case 2}} \\
-    \midrule
-    Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-    & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-    \hline
-    \rowcolor{gray!25!}
-     YACC & $100\times100$ & 0.6058 & 8.5278 & 8.3990 & 142.2020 & 105.1767 & 122.3238 & $1.9853\times10^{-5}$ & 1.6095 & 7.6222 & 7.6194 & 0.04\% \\
-     ELEFANT & $100\times100$ & 0.6034 & 8.5115 & 8.4465 & 140.8067 & 104.1228 & 121.766 & $2.0600\times10^{-5}$ & 1.85 & 7.5665 & 7.5585 & 0.11\% \\
-     \rowcolor{gray!25!}
-     Stone 28 & $32\times32$ & 0.6052 & 8.1715 & - & 141.518 & - & - & - & - & - & - & - \\
-%     FALCON & $32\times32$ & 0.6454 & 8.1422 & 7.4020 & 156.6908 & 95.7226 & 116.9302 & $2.2572\times10^{-5}$ & 0.7194 & 7.4289 & 7.4452 & 0.22\% \\
-%     \rowcolor{gray!25!}
-     FALCON & $100\times100$ & 0.6044 & 8.5266 & 8.4590 & 141.3684 & 104.8867 & 122.0470 & $2.0333\times10^{-5}$ & 1.2914 & 7.5707 & 7.5718 & 0.01\% \\
-     \midrule
-     \multicolumn{13}{|c|}{\textbf{Case 3}} \\
-     \midrule
-     Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-     & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-     \hline
-     \rowcolor{gray!25!} 
-     YACC & $100\times100$ & 0.7286 & 3.0374 & 2.9628 & 100.9467 & 2.0374 & 2.8458 & $4.7907\times10^{-5}$ & 1 & 2.0400 & 2.0335 & 0.32\% \\
-      ELEFANT & $100\times100$ & 0.7275 & 3.0347 & 2.9908 & 100.1208 & 2.0652 & 2.9019 & $4.8080\times10^{-5}$ & 0.9023 & 2.0384 & 2.037 & 0.07\% \\
-      \rowcolor{gray!25!}
-      Stone 28 & $32\times32$ & 0.7277 & 3.0347 & - & 100.018 & - & - & - & - & - & - & - \\
-%      FALCON & $32\times32$ & 0.7281 & 3.0618 & 2.8132 & 101.0092 & 2.1630 & 3.0178 & $4.8500\times10^{-5}$ & 0.7845 & 2.0699 & 2.0695 & 0.02\% \\
-%      \rowcolor{gray!25!}
-      FALCON & $100\times100$ & 0.7275 & 3.0368 & 3.0086 & 100.2776 & 2.0832 & 2.9086 & $4.8129\times10^{-5}$ & 0.9360 & 2.0387 & 2.0386 & $<$0.01\% \\
-      \midrule
-      \multicolumn{13}{|c|}{\textbf{Case 4}} \\
-      \midrule
-      Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-      & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-      \hline
-      \rowcolor{gray!25!} 
-      YACC & $100\times100$ & 0.5289 & 6.5572 & 6.5243 & 79.6202 & 75.4814 & 89.2940 & $1.9174\times10^{-4}$ & 1.6773 & 5.6512 & 5.6463 & 0.09\% \\
-       ELEFANT & $100\times100$ & 0.5277 & 6.5912 & 6.5834 & 79.1105 & 74.7596 & 88.9146 & $1.9860\times10^{-4}$ & 1.5200 & 5.6216 & 5.6182 & 0.06\% \\
-       \rowcolor{gray!25!}
-       Stone 28 & $32\times32$ & 0.5286 & 6.4036 & - & 79.4746 & - & - & - & - & - & - & - \\
-%       FALCON & $32\times32$ & 0.5334 & 6.4679 & 6.3435 & 80.6153 & 76.0511 & 90.2058 & $2.0393\times10^{-4}$ & 0.6453 & 5.6676 & 5.6698 & 0.04\% \\
-%       \rowcolor{gray!25!}
-       FALCON & $100\times100$ & 0.5279 & 6.5897 & 6.5830 & 79.1220 & 75.0715 & 88.8484 & $1.9693\times10^{-4}$ & 1.4133 & 5.6127 & 5.6113 & 0.03\% \\
-       \midrule
-       \multicolumn{13}{|c|}{\textbf{Case 5}} \\
-       \midrule
-       Code & Resolution & $P$ & $\left\langle T \right\rangle_{min}$ & $\left\langle T \right\rangle_{max}$ & Nu$^{top}_{min}$ & Nu$^{top}_{max}$ &
-       $v_{\textrm{rms}}^{min}$ & $v_{\textrm{rms}}^{max}$ & \multicolumn{2}{c}{$(\left\langle \Phi \right\rangle/Ra)_{min}$} & 
-       \multicolumn{2}{c|}{$(\left\langle \Phi \right\rangle/Ra)_{max}$} \\
-       \hline
-       \rowcolor{gray!25!}
-         YACC & $100\times100$ & 0.0792 & 0.6543 & 0.6722 & 2.6950 & 7.2792 & 41.9813 & 98.8230 & \multicolumn{2}{c}{1.4347} & 
-         \multicolumn{2}{c|}{9.2748} \\
-         ELEFANT & $100\times100$ & 0.076 & 0.6514 & 0.6693 & 2.6808 & 7.3506 & 41.2578 & 100.089 & \multicolumn{2}{c}{1.3324} & 
-         \multicolumn{2}{c|}{9.5289} \\
-         \rowcolor{gray!25!}
-         Stone 28 & $32\times32$ & 0.0764 & 0.6521 & 0.6697 & 2.6912 & 7.0763 & 41.61 & 98.183 & \multicolumn{2}{c}{-} & 
-         \multicolumn{2}{c|}{-} \\
-     %    FALCON & $32\times32$ & & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & \multicolumn{2}{c}{-} & 
-     %     \multicolumn{2}{c|}{-} \\
-     %    \rowcolor{gray!25!}
-         FALCON & $100\times100$ & 0.0775 & 0.6519 & 0.6697 & 2.6847 & 7.3629 & 41.5256 & 99.5466 & \multicolumn{2}{c}{1.3399} & 
-         \multicolumn{2}{c|}{9.3993} \\
-  \bottomrule
-  \end{tabular}}
-  \label{tab:visco_mantle}
-  \end{table}
-
-\subsection{Thin layer entrainment}\label{sec:thin}
-This experiment is performed as originally proposed by \citet{vanKeken1997} with two incompressible fluids in a rectangular domain with $L_x=2$ and $L_y=1$
-and gravity acceleration $g_y=10^{10}Ra$. The Rayleigh number and the compositional Rayleigh number are fixed to $Ra=300000$ and $Rc=450000$, respectively.
-Both fluids have constant viscosity, thermal conductivity, specific heat ($\eta=\rho=C_p=1$) and thermal expansion coefficient ($\alpha=10^{-10}$). Fluid 1
-has a density $\rho_1=1$, while fluid 2 is denser ($\rho_2=\rho_1+1.5 \cdot 10^{-10}$) and is placed at the bottom of the domain, for $y \leq 0.025$.
-Temperature are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides of the domain. The initial
-temperature field is given by
-\[T(x,y)=T_u(x,y)+T_l(x,y)+T_r(x,y)+T_s(x,y)-\frac{3}{2}\]
-with
-\begin{eqnarray}
-T_u(x,y)&=&\frac{1}{2}erf\left(\frac{1-y}{2}\sqrt{\frac{u_0}{x}}\right)\nonumber \\
-T_l(x,y)&=&1-\frac{1}{2}erf\left(\frac{y}{2}\sqrt{\frac{u_0}{L_x-x}}\right)\nonumber \\
-T_r(x,y)&=&\frac{1}{2}+\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{y+1}} \exp\left(-\frac{x^2u_0}{4y+4}\right)\nonumber \\
-T_s(x,y)&=&\frac{1}{2}-\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{2-y}} \exp\left(-\frac{(L_x-x)^2u_0}{8-4y}\right)\nonumber
-\end{eqnarray}
-and
-\begin{eqnarray}
-u_0&=&\frac{L_x^{7/3}}{(1+L_x^4)^{2/3}}\left(\frac{Ra}{2\sqrt{\pi}}\right)^{\frac{2}{3}}\nonumber \\
-Q&=&2\sqrt{\frac{L_x}{\pi u_0}}\nonumber
-\end{eqnarray}
-
-The experiment is performed with two grid resolution (\num{125x40} and \num{200x80} elements), with 100 markers per element and CFL of 0.25. $v_{\textrm{rms}}$
-(see Sec. \ref{sec:error}) calculated for both the simulations match well with results from \citet{vanKeken1997} and obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and ELEFANT \citep{Thieulot2014}, obtained with similar grid resolutions
-(Fig. \ref{fig:thin}). In addition, a variety of simulations with different resolutions and aspect ratios are performed to verify that the sensitivity of the
-initial velocity field. Results match well with results from \citet{vanKeken1997} (in the grey area) and \citet{Thieulot2014} (Fig. \ref{fig:thin_initial}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Thin_layer}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/vrms_thin.pdf}
-\caption{$v_{\textrm{rms}}$ for the thin layer experiment as function of time for different grid resolution. Results are compared with results obtained by
-\citet{vanKeken1997} (black lines) and with ASPECT and ELEFANT.}
-\label{fig:thin}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/vrmszero.pdf}
-\caption{$v_{\textrm{rms}}$ at $t=0$ for the thin layer experiment as function of the element size for different aspect ratios of the elements. Results are compared
-with results obtained by \citet{vanKeken1997} (black dots in the grey area) and with ASPECT and ELEFANT.}
-\label{fig:thin_initial}
-\end{figure}
-
-\subsection{Sinking hydrated cylinder}\label{sec:quinquis}
-This experiment is performed as explained in \citet{Quinquis2014} and simulates the sinking of a cold, hydrated cylinder into a warm, dry mantle. The domain
-is square with $L_x=L_y=$ \SI{300}{\km}, a grid resolution of \num{300x300} elements and 25 markers per element. The sinking cylinder is located at $x=$
-\SI{150}{\km} and $y=$ \SI{170}{\km} with a radius of \SI{20}{\km} and has $\rho_c=$ \SI{3250}{\kg\per\cubic\m}, $\eta_c=$ \SI{e23}{\pascal\s}, a thermal
-conductivity $k_c=$ \SI{4.5}{\watt\per\m\per\kelvin},a specific heat $C_{p_c}=$ \SI{1250}{\joule\per\kg\per\kelvin}; the surrounding mantle has $\rho_m=$
-\SI{3200}{\kg\per\cubic\m}, $\eta_m=$ \SI{e20}{\pascal\s}, a thermal conductivity $k_m=$ \SI{105}{\watt\per\m\per\kelvin} and specific heat $C_{p_m}=$
-\SI{1250}{\joule\per\kg\per\kelvin}; a 58 km-thick lithosphere overlies the mantle and has $\rho_l=$ \SI{3200}{\kg\per\cubic\m}, $\eta_l=$ \SI{e23}{\pascal\s},
-a thermal conductivity $k_l=$ \SI{4.5}{\watt\per\m\per\kelvin}, specific heat $C_{p_l}=$ \SI{750}{\joule\per\kg\per\kelvin} and a thermal diffusivity of
-\SI{e-6}{\square\m\per\s}. Velocity boundary conditions are set to free slip on all sides of the domain. Temperature increase linearly in the lithosphere,
-from 0° to 1300°C, and in the mantle, from 1300° to 1360.5°C. The initial temperature of the cylinder is fixed to 400°C. Throughout the evolution, temperature
-is fixed to 0° and 1360.5°C at the top and bottom of the domain, respectively. The initial bound water content is imposed at 2 wt.\% in the cylinder and at
-0 wt.\% in both the mantle and the lithosphere. The maximum amount of water in the mantle is fixed to 0.2 wt.\%, while maximum water content in the cylinder
-and in the lithosphere is function of pressure and temperature and is calculated using a serpentinized harzburgite with Perple\_X, as in \citet{Quinquis2014}.
-
-As shown in \citet{Quinquis2014}, the model is characterised by a progressive dehydration in the external portion of the cylinder, due to the increase of
-temperature, with a consequent vertical migration of free water that hydrates the mantle above the cylinder, up to the lithosphere. The distribution of
-bound and free water during the evolution of the experiment is shown in Fig. \ref{fig:quinquis} (left and right columns, respectively). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Hydration}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Hydration.pdf}
-\caption{Distribution of bound and free water (left and right columns, respectively) at $t=0.5$ and $t=$ \SI{1.25}{\mega\year} (first and second rows,
-respectively) for the sinking hydrated cylinder experiment. Only markers with bound or free water not equal to 0 are plotted.}
-\label{fig:quinquis}
-\end{figure}
-
-\subsection{Mantle melting curves}\label{sec:katz}
-This instantaneous experiment is performed in order to recreate isobaric melting curves, as function of temperature, bulk water and content of water in the
-melt, to compare with melting curves obtained by \citet{Katz2003}. The domain is square with $L_x=L_y=100$ km, a grid resolution of \num{100x100} elements and
-25 markers per element. Density is set in the entire domain to \SI{3300}{\kg\per\cubic\m} and temperature increases from the sides to the centre from 1000° to
-1500°C. The experiment is performed for bulk water contents from 0 to 0.3 wt.\%.
-
-Fig. \ref{fig:fraction} shows isobaric melting curves for \SI{1}{\giga\pascal} (panel a) and \SI{3}{\giga\pascal} (panel b) calculated for different bulk water
-contents. The curves well match with those obtained for the mantle by \citet{Katz2003}, showing the saturation at the solidus for bulk water of 0.2 and 0.3 wt.\%, indicating
-that the code correctly determines melt fraction as function of pressure, temperature and bulk water content. Same results can be observed in Fig.
-\ref{fig:water}a, where isobaric and isothermal melting curves are shown as function of bulk water content. In addition, Fig. \ref{fig:water}b shows that the
-code correctly determines also the percentage of water in the melt as function of pressure, temperature and melt fraction. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Melting}.
-
-\begin{figure}
-\includegraphics[width=\linewidth]{./Figures/Fraction.pdf}
-\caption{Isobaric melting curves for 1 GPa (panel a) and 3 GPa (panel b) as function of temperature with different bulk water contents.}
-\label{fig:fraction}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\linewidth]{./Figures/Water.pdf}
-\caption{Isobaric melting curves for 1.5 GPa as function of bulk water (panel a) and water in the melt (panel b) calculated for different temperatures.}
-\label{fig:water}
-\end{figure}
-
-\clearpage
-
-\bibliography{export}
-
-\end{document}

### 2022-09-07 tex/Benchmark/poiseuille.tex
update main

--- 2022-03-31 FALCON_description.tex
@@
-\documentclass[hidelinks,10pt,a4paper]{article}
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{textcomp}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage{booktabs}
-\usepackage{multirow}
-\usepackage[cm]{fullpage}
-\usepackage[round]{natbib}
-\bibliographystyle{abbrvnat}
-\usepackage{doi}
-\usepackage{bm}
-\usepackage[table]{xcolor}
-\hypersetup{
-    colorlinks,
-    linkcolor={blue},
-    citecolor={blue},
-    urlcolor={blue}
-}
-\usepackage{float}
-\restylefloat{table}
-\usepackage{siunitx}
-\DeclareSIUnit\year{yr}
-
-\begin{document}
-
-\section{Introduction}\label{sec:intro}
-This document includes a description of the code FALCON, with a presentation of the features implemented followed by results of extensive benchmarks listed in
-Table \ref{tab:list}, performed to verify that the code solves correctly a wide range of problems. Description and results of all benchmarks are shown
-in Section \ref{sec:benchmark}.
-
-\begin{table}[H]
-\caption{List of benchmarks performed for each feature of the code. Data of all benchmarks can be found at 
-\url{https://github.com/aleregorda/Benchmarks}}
-\centering
-\begin{tabular}{| l | l | l |}
-\toprule
- \textbf{Feature}  & \textbf{Benchmark} & \textbf{Section} \\
-\midrule
-  Solver  & Stokes flow & \ref{sec:stokes} \\
-\hline
-  \multirow{2}{*}{Markers advection} & Zalesak disk & \ref{sec:runge} \\
-    & Conservative Interpolation Velocity & \ref{sec:CVI} \\
-\hline
-  \multirow{4}{*}{Momentum equation}  & Poiseuille flow & \ref{sec:poiseuille}\\
-    & Instantaneous 2D sphere & \ref{sec:ist_sphere} \\
-    & Rayleigh-Taylor experiment & \ref{sec:rayleigh} \\
-    & Falling block & \ref{sec:block} \\
-\hline
-    & 2D Stokes sphere & \ref{sec:sphere} \\
-    Sticky air and  & Stabilisation algorithm & \ref{sec:stab} \\
-    free surface & Topography relaxation & \ref{sec:crameri} \\
-  & Spontaneous subduction & \ref{sec:subduction} \\
-\hline
-  Erosion and sedimentation  & - & - \\
-\hline
-   Non-linear  & Slab detachment & \ref{sec:slab} \\
-   visco-plasticity & Indenter & \ref{sec:indenter} \\
-    & Brick & \ref{sec:brick} \\
-\hline
-  \multirow{3}{*}{Energy equation}  & Advection stabilisation & \ref{sec:advection} \\
-    & Simple shear heating & \ref{sec:simple_shear} \\
-    & Shear and adiabatic heating & \ref{sec:shear} \\
-\hline
-\multirow{3}{*}{Energy + momentum} & Mantle convection & \ref{sec:mantle} \\
-& Viscoplastic mantle convection & \ref{sec:visco_mantle} \\
-& Thin layer entrainment & \ref{sec:thin} \\
-\hline
-  Phase changes  & \multirow{2}{*}{Hydrated sinking cylinder} & \multirow{2}{*}{\ref{sec:quinquis}} \\
-  and hydration  &  & \\
-\hline
-  Melting & Experimental melting curves & \ref{sec:katz} \\
-\bottomrule
-\end{tabular}
-\label{tab:list}
-\end{table}
-
-The code is written in Fortran90 and uses the finite element method (FEM) with quadrilateral $Q_1 \times P_0$ elements (continuous bilinear velocity and
-discontinuous constant pressure), associated with MUMPS\footnote{\url{http://mumps.enseeiht.fr/}} \citep{Amestoy2001,Amestoy2019} that is a software package
-for solving systems of linear equations of the form $A \cdot x = b$, where \textit{A} is a square sparse matrix, by means of a direct method. Since
-$Q_1 \times P_0$ elements do not satisfy Ladyzhenskaya, Babuska and Brezzi (LBB) stability condition \citep{Donea2003}, elemental pressure is elaborated
-in post-processing to avoid spurious pressures by means of a double interpolation to smooth the pressure. Following this procedure, the elemental pressure
-is interpolated onto nodes and then back onto elements. After the smoothing procedure, the pressure field is calculated again on the nodes to be used in
-combination with the lithostatic pressure, which is calculated onto nodes. Correctness of the solution and performances of the solver in terms of time and
-memory usage are tested solving the Stokes flow with the analytical solution proposed by \citet{Donea2003} (Section \ref{sec:stokes}).
-
-The thermo-mechanics of crust-mantle systems is described by means of the conservation of mass, momentum and energy equations, expressed as follows:
-\begin{eqnarray}
-\label{eq:mass}\frac{\partial \rho}{\partial t}+\bm{\nabla} \cdot (\rho \bm{u})&=&0\\
-\label{eq:momentum}-\bm{\nabla} p + \bm{\nabla} \cdot \bm{\tau} + \rho \bm{g}&=&0\\
-\label{eq:energy}\rho C_p \left(\frac{\partial T}{\partial t} + \bm{u} \cdot \bm{\nabla} T \right)&=&\bm{\nabla} \cdot \left(k\bm{\nabla} T\right) + H_{tot}
-\end{eqnarray}
-where $\rho$ is the density, $\bm{u}$ is the velocity, \textit{p} is the pressure, $\bm{\tau}$ is the deviatoric stress, $\bm{g}$ is the gravity acceleration,
-$C_p$ is the specific heat at constant pressure, \textit{T} is the temperature, \textit{k} is the thermal conductivity and $H_{tot}$ is the total internal heat
-production. Density variations due to temperature are generally small enough to assume the density as constant ($\rho=\rho_0$) in Eq. \ref{eq:mass} and in
-Eq. \ref{eq:energy}, while it must be treated as a variable in the buoyancy term of Eq. \ref{eq:momentum}, such that \[\rho=\rho_0(1-\alpha(T-T_0))\]
-where $\rho_0$ is the density at a reference temperature $T_0$ and $\alpha$ is the coefficient of thermal expansion. Eq. \ref{eq:mass} then can be rewritten as
-\begin{equation}\label{eq:mass_0}
-\bm{\nabla} \cdot \bm{u}=0
-\end{equation}
-This simplification is known as the extended Boussinesq approximation.
-
-The time step $dt$ is chosen in according to Courant-Friedrichs-Lewy condition \citep{Anderson1995}, such that
-\[dt=C_n\; \min\left(\frac{\Delta x}{\max|\bm{u}|},\frac{(\Delta x)^2}{\kappa}\right)\]
-where $0<C_n<1$ is the Courant (CFL) number, $\Delta x$ is the minimum dimension of smallest element, $\max|\bm{u}|$ is the maximum velocity calculated on the
-entire domain and
-$\kappa=\frac{k}{\rho C_p}$ is the thermal diffusivity \citep{Thieulot2014}.
-
-\section{Lagrangian markers}\label{sec:markers}
-Elemental properties (density, viscosity, thermal conductivity, specific heat and thermal expansion) needed to solve Eqs. \ref{eq:mass_0}, \ref{eq:momentum}
-and \ref{eq:energy} are related to the composition of each element, determined by means of Lagrangian markers that are characteristic of different materials
-in the domain. Elemental properties (with the sole exception of the viscosity) are calculated using an arithmetic mean, as \[P_e=\frac{1}{n}\sum_{i=1}^n P_i\]
-where $P_e$ is the elemental property, $P_i$ is the property characteristic of the material of each marker and $n$ is the number of markers of the element.
-Differently, the average scheme for the elemental viscosity can be chosen between harmonic, geometric and arithmetic mean. Markers can be placed regularly or
-randomly at the beginning of the simulation and their advection is performed by either a 2nd-order or a 4th-order Runge-Kutta in space, interpolating the velocity 
-field on each marker by means of the shape functions (see Section \ref{sec:runge}). The interpolated velocity is then corrected by means of the Conservative
-Velocity Interpolation (CVI), which introduces a corrective term to reduce dispersion and clustering of particles in both steady state and time-dependent 
-\citep{Wang2015} (see Section \ref{sec:CVI}).
-
-The initial distribution of the markers is created using the open-source code library
-Geodynamic World Builder \citep{Fraters2019}. During the simulation, each marker carries memory of temperature, pressure and accumulated strain, which are
-determined interpolating the relative nodal parameter. In particular, pressure is interpolated onto each marker after the smoothing procedure. The number
-of markers contained in each element is maintained between $n_{min}$ and $n_{max}$. When in an element there are less markers than $n_{min}$ the code adds
-random markers to reach the $n_{min}$, while if the number is higher than $n_{max}$ some of them are randomly deleted. When new markers are added, they assume
-the properties of the nearest marker. In this way elements are never empty and maintain a number of markers inside a prefixed range.
-
-\section{The momentum equation}\label{sec:mom_eq}
-The penalty method is implemented so that Eq. \ref{eq:mass_0} can be written as:
-\begin{equation}\label{eq:mass_penalty}
-\bm{\nabla} \cdot \bm{u} + \frac{p}{\lambda}=0
-\end{equation}
-where $\lambda$ is the so-called penalty parameter, which should be 6-7 orders of magnitude larger than the shear viscosity to ensure that mass conservation
-is satisfied \citep{Donea2003,Thieulot2014}.
-
-The deviatoric stress tensor in Eq. \ref{eq:momentum} can be written in terms of the strain rate tensor as $\bm{\tau}=2\eta\dot{\bm{\epsilon}}$, with
-$\dot{\bm{\epsilon}}=\frac{1}{2}\left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)$. Therefore, Eq. \ref{eq:momentum} can be rewritten as:
-\begin{equation}\label{eq:momentum_pressure}
--\bm{\nabla} p + \bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right) + \rho \bm{g} = 0
-\end{equation}
-Finally, using pressure from Eq. \ref{eq:mass_penalty}, Eq. \ref{eq:momentum_pressure} can be rewritten as:
-\begin{equation}\label{eq:momentum_penalty}
-\lambda \bm{\nabla} \left(\bm{\nabla} \cdot \bm{v}\right)+\bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right)+\rho\bm{g}=0
-\end{equation}
-The penalty method is been associated to the iterative Uzawa method, as described in detail in \citet{Dabrowski2008} and \citet{Thieulot2014}
-(see Section \ref{sec:poiseuille}). Results of benchmarks performed to verify the correctness in the implementation of Eq. \ref{eq:momentum_penalty} are
-shown in Sections \ref{sec:ist_sphere} and \ref{sec:rayleigh}.
-
-To support large viscosities variations the penalty parameter is related to the effective elemental viscosity by means of a dimensionless coefficient, so that
-$\lambda=\lambda_e(e)\eta_{eff}(e)$ \citep{Marotta2006,Dabrowski2008,Thieulot2014}. Benchmark of the falling block \citep{Gerya2003a,Gerya2010b,Thieulot2011}
-is performed to verify that the code can correctly deal buoyancy driven flows with strong viscosity contrasts (Section \ref{sec:block}).
-
-\section{Sticky air and free surface}\label{sec:sticky}
-The Earth's surface can be treated by means of either the so-called sticky-air or a true free surface method, both of them implemented in the code. In the
-sticky-air method the surface is approximated with the introduction of a buoyant layer with a viscosity at least four orders of magnitude lower than the crust
-\citep{Schmeling2008,Crameri2012} and the interface between lithosphere and air is defined using a chain of passive markers that are advected as the Lagrangian
-markers. The correctness of the evolution of the markers chain is tested with the experiment of a 2D time-dependent Stokes sphere below a free surface and
-compared with results from ASPECT\footnote{\url{https://aspect.geodynamics.org/}} \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} (Section
-\ref{sec:sphere}). In the true free surface case the top boundary is assumed stress-free and velocities are not fixed. In this case topography variations are
-described by vertical deformations of the mesh that depend on the velocity field of the nodes that identify the free surface, while horizontal deformations are
-not taken into account. This procedure is known as the Arbitrary Lagrangian-Eulerian (ALE) method and its implementation follows the technique described in
-\citet{Thieulot2011}. However, although the implementation of a true free surface better reproduces laboratory experiment, extremely small time steps can be
-necessary to maintain stability \citep{Kaus2010a,Quinquis2011,Thieulot2014}. Therefore, the stability algorithm proposed by \citet{Kaus2010a} is implemented
-to avoid instabilities due to high density differences at the free surface when using too large time steps. The implementation of this algorithm are tested by
-performing the experiment described by \citet{Kaus2010a} (Section \ref{sec:stab}).
-
-The topography relaxation benchmark proposed by \citet{Crameri2012} is performed to verify that the code correctly recovers topography variations in case of
-both the sticky-air and the true free surface method (Section \ref{sec:crameri}). Finally, Section \ref{sec:subduction} show the results of the spontaneous
-subduction experiment described by \citet{Schmeling2008}.
-
-\section{Erosion and sedimentation}\label{sec:erosion}
-Surface processes at either the lithosphere-air interface or in correspondence of the free surface have been implemented by means of the software
-Fastscape\footnote{\url{https://fastscape.org/}} \citep{Braun2013,Cordonnier2019,Yuan2019a,Yuan2019b}. FastScapeLib is a set of routines that solve the stream
-power law (enriched by a sediment transport/deposition term) \citep{Yuan2019a}, hillslope diffusion and marine transport and deposition \citep{Yuan2019b}.
-The partial differential equation solved by FastScapeLib is:
-\begin{equation}
-\frac{\partial h}{ \partial t}=U-K_f A^{m_e}S^{n_e} + \frac{G}{A}\int_A \left( U-\frac{\partial h}{\partial t} \right) dA + K_d \bm{\nabla}^2 h 
-\end{equation}
-where $h$ is the topography, $U$ is the uplift, $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $S$ is the slope, $m_e$ and $n_e$ are
-parameters that should be chosen to constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}, $G$ is the dimensionless
-deposition/transport coefficient for the enriched stream power law and $K_d$ is the effective diffusivity. Terms $K_f A^{m_e}S^{n_e}$ and $K_d \bm{\nabla}^2 h$
-consider long-range processes, described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, and short-range processes,
-proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}.
-
-In case of sticky air, slope is determined by means of the markers chain, which is vertically corrected considering erosion and sedimentation rates. After the
-correction, continental markers above the markers chain are transformed in air markers and air markers below the markers chain become sediments. Similarly,
-slope is determined considering the nodes on the top boundary in case of true free surface and they are vertically corrected as for the markers chain. In this
-case, continental markers above the top boundary are simply deleted.
-%Surface processes at the lithosphere-air interface are implemented taking into account fluvial erosion (long-range processes), hillslope processes (short-range
-%processes) and failure of slopes exceeding a threshold angle $\theta$ \citep{Burov1997,Tucker1999,Sacek2012,Croissant2014,Ueda2015,Beucher2020}. Erosion rates
-%$\dot{E}_{lr}$ due to long-range processes can be described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{lr}=K_f A^{m_e} \bm{\nabla} h^{n_e}
-%\end{equation}
-%where $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $\bm{\nabla} h$ is the slope and $m_e$ and $n_e$ are parameters chosen in to
-%constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}. Erosion rates $\dot{E}_{sr}$ due to short-range processes are
-%implemented by assuming that transport is linearly proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{sr}=K_d \bm{\nabla}^2 h
-%\end{equation}
-%where $K_d$ is the effective diffusivity. In addition, the slope is modified if it still exceeds a threshold angle after long- and short-range processes
-%\citep{Ueda2015}.
-
-%Offshore sedimentation processes are implemented similarly to hillslope processes but with a depth-dependent coefficient \citep{Sacek2012,Beucher2020}.
-%The sedimentation rate $\dot{S}$ is calculated as
-%\begin{equation}
-%\dot{S}=K_s \bm{\nabla}^2 h
-%\end{equation}
-%with
-%\begin{equation}
-%K_s=K^*_s \exp\left(\frac{h-h_{sl}}{h_0}\right)
-%\end{equation}
-%where $K^*_s$ is a constant, $h_{sl}$ is the sea level and $h_0$ is a characteristic length. In case of sticky air, slope is determined by means of the markers
-%chain, which is vertically corrected considering erosion and sedimentation rates. After the correction, continental markers above the markers chain are
-%transformed in air markers and air markers below the markers chain become sediments. Similarly, slope is determined considering the nodes on the top boundary
-%in case of true free surface and they are vertically corrected as for the markers chain. In this case, continental markers above the top boundary are simply
-%deleted.
-
-\section{Non-linear rheologies}\label{sec:nonlinear}
-Non-linear rheologies are implemented combining viscous creep (dislocation and diffusion) and plastic yielding. For each marker, diffusion and dislocation
-viscosity ($\eta_{df|ds}$) can be determined as follows \citep{Karato1993,Warren2008b,Wang2016}:
-\begin{equation}\label{eq:viscous}
-\eta_{df|ds}=f_s \left(\frac{d^m}{A}\right)^{\frac{1}{n}} \left (\frac{2^{\frac{1-n}{n}}}{3^{\frac{n+1}{2n}}} \right)
-I_2^{\frac{(1-n)}{n}}\exp\left(\frac{Q+pV}{nRT}\right)
-\end{equation}
-where $f_s$ is a scaling factor used to represent lithologies that are stronger or weaker than the base set, $d$ is the grain size, $m$ is the grain size exponent,
-$n$ is the stress exponent, $A$ is the uniaxial pre-exponential factor, $I_2$ is the square root of the second invariant of the strain rate tensor, $Q$ is the
-activation energy, $p$ is the pressure, $V$ is the activation volume, $R$ is the gas constant and $T$ is the temperature.
-Pressures and temperatures are determined by the interpolation of the nodal parameters, while strain rates are calculated by means of the
-derivative of the velocities on the nodes. In case of diffusion creep $n=1$ and $m>0$, in case of dislocation creep $n>1$ and $m=0$. The viscous creep $\eta_{cp}$
-is then calculated as the harmonic average between $\eta_{df}$ and $\eta_{ds}$:
-\begin{equation}\label{eq:visco_creep}
-\eta_{cp}=\left(\frac{1}{\eta_{df}}+\frac{1}{\eta_{ds}}\right)^{-1}
-\end{equation}
-The implementation of the non-linear viscous creep viscosity depending on the strain rate is tested by means of the slab detachment benchmark
-(Section \ref{sec:slab}).
-
-Plastic yielding is implemented rescaling $\eta_{cp}$ in order to limit the stress below the yield stress $\sigma_y$ \citep{Thieulot2008,Thieulot2014,Glerum2018},
-obtaining
-\begin{equation}\label{eq:plastic}
-\eta_{pl}=\frac{\sigma_y}{2I_2}
-\end{equation}
-where $\eta_{pl}$ is the plastic viscosity and the yield stress is determined following the Drucker-Prager criterion, such as
-\begin{equation}\label{eq:yield}
-\sigma_y=C\;\cos(\phi)+p\;\sin(\phi)
-\end{equation}
-where $C$ is the cohesion and $\phi$ is the internal friction angle. In case of negative pressure it is imposed equal to 0, so that negative yield stress are
-excluded. The correctness of the non-linear solution in case of plastic viscosity with variable internal friction angle is verified performing indenter and
-brick experiments (Sections \ref{sec:indenter} and \ref{sec:brick}, respectively).
-
-Strain softening is taken into account for both viscous creep and plastic viscosity \citep{Huismans2003,Babeyko2005,Huismans2005,Sobolev2005,Warren2008a} by
-means of the accumulated strain $\epsilon$ memorised by each marker. Softening in the viscous creep determines a linear decrease of $\eta_{vc}$ by means of a
-viscous strain softening factor $W_S$ that increases linearly from $W_{S_0}$ to $W_{S_{\infty}}$ for $\epsilon_{S_0}<\epsilon_S<\epsilon_{S_{\infty}}$ 
-\citep{Warren2008a}. This viscous softening can be related to strain-induced grain size reduction \citep{Warren2008a}. Differently, plastic softening is 
-simulated with a linear decrease of internal friction angle $\phi(\epsilon)$ and cohesion $C(\epsilon)$ in according to
-\begin{eqnarray}
-\label{eq:friction1}\phi(\epsilon)&=&\phi_0+(\phi_{\infty}-\phi_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}\\
-\label{eq:friction2}C(\epsilon)&=&C_0+(C_{\infty}-C_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}
-\end{eqnarray}
-where $\phi_0$, $C_0$ and $\phi_{\infty}$, $C_{\infty}$ are internal friction angle and cohesion for $\epsilon_0$ and $\epsilon_{\infty}$, respectively
-\citep{Huismans2003,Huismans2005,Warren2008a,Thieulot2014}. Plastic softening approximates deformation-induced softening of faults and brittle shear zones
-\citep{Warren2008a}.
-
-Plastic yielding and viscous creep are then combined to obtain a viscoplastic viscosity $\eta_{vp}$ as follows, assuming that they are independent processes
-\citep{Karato2008,Glerum2018}:
-\begin{equation}\label{eq:viscoplastic}
-\eta_{vp}=\min(\eta_{cp},\eta_{pl})
-\end{equation}
-Finally, effective viscosity $\eta_{eff}$ is capped by the minimum and the maximum viscosity ($\eta_{min}$ and $\eta_{max}$, respectively) to avoid extremely
-low or high viscosity \citep{Glerum2018} as follows
-\begin{equation}\label{eq:effective}
-\eta_{eff}=\min(\max(\eta_{vp},\eta_{min}),\eta_{max})
-\end{equation}
-Viscous creep and plastic yielding are non-linear rheologies because of their dependence on the velocity field through pressure and strain rates. Therefore,
-the solution is determined by means of Picard-type iterations, until convergence of the velocity field \citep{Glerum2018}. The convergence is verified at each
-iteration $i$ via the nonlinear residual $\bm{\mathcal{R}}^i$ that can be determined as
-\begin{equation}\label{eq:convergence}
-\bm{\mathcal{R}}^i= \mathbb{K} (\eta_{eff}(\dot{\bm{\epsilon}}^{i-1},p^{i-1}))\cdot \bm{v}^{i-1} - \bm{f}^i
-\end{equation}
-%\begin{equation}\label{eq:convergence}
-%\chi=1-\frac{\langle (\bm{v}^i-\langle\bm{v}^i\rangle)\cdot (\bm{v}^{i+1}-\langle\bm{v}^{i+1}\rangle) \rangle}{\sqrt{\sigma^i\sigma^{i+1}}}
-%\end{equation}
-%Iterations are performed until either convergence ($\chi<tol$) or maximal number of non-linear iterations ($it_{max}$) is reached.
-where $\mathbb{K}$ is the Stiffness matrix and $\bm{f}$ is the right hand side vector \citep{Spiegelman2016,Glerum2018}.  The $L_2$-norm is extracted from
-$\bm{\mathcal{R}}^i$ and it is normalised as follows
-\begin{equation}\label{eq:convergence_res}
-\frac{||\bm{\mathcal{R}}^i||_2}{||\bm{\mathcal{R}}^0||_2}
-\end{equation}
-where $||\bm{\mathcal{R}}^0||_2$ is the $L_2$-norm of the first nonlinear iteration.
-Since often the normalised nonlinear residual drop very quickly over the first 2-3 iterations, also the $L_2$-norm of velocity and pressure residuals are
-calculated as
-\begin{equation}\label{eq:convergence_vel}
-\frac{||\Delta\bm{v}||_2}{||\bm{v}||_2}=\frac{||\bm{v}_i-\bm{v}_{i-1}||_2}{||\bm{v}_{i}||_2}
-\end{equation}
-and
-\begin{equation}\label{eq:convergence_press}
-  \frac{||\Delta p||_2}{||p||_2}=\frac{||p_i-p_{i-1}||_2}{||p_{i}||_2}
-\end{equation}
-respectively. Iterations are performed until either maximal number of non-linear iterations ($it_{max}$) is reached or all the normalised nonlinear residual, 
-the velocity residual and the pressure residual converge under a defined tolerance $tol$.
-
-\section{The energy equation}\label{sec:en_eq}
-The stabilisation of the advection term of the energy equation (Eq. \ref{eq:energy}) needed to avoid possible oscillations in the thermal solution in the case
-when advection dominates over diffusion, is implemented by means of a streamline-upwind Petrov–Galerkin (SUPG) method, for which the advection term is modified
-follows the discussion in \citet{Thieulot2011} and \citet{Thieulot2014} (see Section \ref{sec:advection} for benchmark). The total internal heat production
-$H_{tot}$ in Eq. \ref{eq:energy} includes radiogenic heating $H_r$, shear heating $H_s$ and adiabatic heating $H_a$, where
-\begin{equation}\label{eq:shear}
-H_s=2\eta \dot{\bm{\epsilon}} : \dot{\bm{\epsilon}} = 2\eta(\dot{\epsilon}_{xx}^2+\dot{\epsilon}_{yy}^2+2\dot{\epsilon}_{xy}^2)
-\end{equation}
-and
-\begin{equation}\label{eq:adiabatic}
-H_a=T \alpha \frac{Dp}{Dt} \approx -\alpha T \rho g_y v
-\end{equation}
-The correctness of the implementation and the computation of shear and adiabatic heating is verified performing an exercise with an analytical solution
-(Section \ref{sec:simple_shear}) and Exercise 9.4 in \citet{Gerya2010b} (Section \ref{sec:shear}). Furthermore, mantle convection and thin layer entrainment
-benchmarks (Sections \ref{sec:mantle} and \ref{sec:thin}, respectively) are performed to verify that the code solve correctly Eqs. \ref{eq:momentum_penalty}
-and \ref{eq:energy} in case of a temperature field characterised by an initial perturbation and with a temperature-dependent density. At last, the viscoplastic
-mantle convection benchmark is performed to check also the viscoplasticity in case of a initial perturbed temperature field (Section \ref{sec:visco_mantle}).
-
-\section{Phase transitions and hydration}\label{sec:phase}
-Variations of effective density, specific heat and coefficient of thermal expansion during the evolution of crustal and mantle materials at different
-pressure-temperature ($p$-$T$) conditions are computed by means of version 6.8.6 of Perple\_X software package \citep{Connolly2005}, similarly to the
-implementation described in \citet{Marotta2020}. By default, mineral assemblages and properties of each lithology (oceanic crust, lower and upper continental
-crust, sediments and dry or hydrated mantle) are calculated for temperatures from \SIrange{330}{1600}{\kelvin}, with increments of approximately \SI{4}{\kelvin},
-and pressures from \SIrange{0.1}{30}{\giga\pascal}, with increments of approximately \SI{0.03}{\giga\pascal}, for a total of almost 100000 points.
-
-The introduction of phase transitions produces two main effects that must be taken into account in numerical models: variations in density and release/absorption
-of latent heat \citep{Tackley2010}. Effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into
-account considering effective coefficient of thermal expansion \citep{Christensen1985,Zhong2015}. Similarly, effects of latent heat in the energy equation
-(eq \ref{eq:energy}) are taken into account by considering effective specific heat and coefficient of thermal expansion
-\citep{Christensen1985,Tackley2010,Zhong2015}. Effective density, specific heat and coefficient of thermal expansion are included in Perple\_X files and are
-assumed by each Lagrangian marker in according to its $p$-$T$ conditions.
-
-Hydration processes related to dehydration of subducting lithosphere \citep{Schmidt1998,Liu2007,Faccenda2009,Faccenda2010,Faccenda2014,Rosas2016} strongly
-influence the thermo-mechanics inside the mantle wedge, mainly because of both weakening effects on the mantle rheology and density variations in case of
-mantle serpentinization \citep{Gerya2002,Honda2003,Arcay2005,Roda2010,Regorda2017}. In the case that hydration is switched on, the amount of bound and free
-water is memorised by each marker, following the implementation of \citet{Quinquis2014}; hydration and dehydration processes are related to the amount of bound
-water of each marker and to the maximum amount of water it can transport, i.e. if the amount of bound water exceeds the maximum amount of water, the marker
-dehydrates and releases free water that can hydrates under-saturated markers. Maximum water content of each marker is determined as function of lithology and
-$p$-$T$ conditions and it is calculated using Perple\_X, in the same way than effective density, specific heat and coefficient of thermal expansion. Bound water
-is advected together with the markers, neglecting the effect of bound water diffusion, while free water simply migrates vertically and is not coupled to the
-solid-phase flow of the mantle wedge \citep{Arcay2005,Quinquis2014}. The correctness of the migration scheme of the free water and its absorption into bound
-water by under-saturated markers are verified performing the experiment described in \citet{Quinquis2014} (Sec. \ref{sec:quinquis}).
-
-The mantle viscosity weakening related to the amount of water has been extensively studied \citep{Chopra1981,Hirth2003} and it is implemented in according to
-\citet{Arcay2005} and \citet{Horiuchi2016}, as follows
-\begin{equation}\label{eq:wet_visc}
-\eta_{wet}=\eta_{dry}\left(\left[1-\frac{1}{f_v}\right]\exp\left(-\frac{[OH^-]}{[OH^-]_0}\right)+\frac{1}{f_v}\right)
-\end{equation}
-where $f_v$ is the viscosity reduction factor between dry and wet conditions, $[OH^-]$ is the water content and $[OH^-]_0$ is a reference water content set to
-620 ppm (0.062 wt.\%) \citep{Arcay2005}. By default, $f_v$ is set to 100, in according to \citet{Arcay2005} and \citet{Horiuchi2016}.
-
-\section{Mantle melting}\label{sec:melt}
-Melt fraction in the mantle depends on temperature ($T$, in \textdegree C), pressure ($p$, in \si{\giga\pascal}) and the water content in the melt ($X_{H_2O}$, in wt.\%).
-The determination of the melt fraction $M$ is implemented as explained by \citet{Katz2003}, according to the successive modification by \citet{Langmuir2006} and
-\citet{Kelley2010}, as follows
-\begin{eqnarray}\label{eq:melt_fr}
-M(p,T,X_{H_2O})&=&-T+T_{sol}+(x_t\cdot ln(p+y_t))\cdot X_{H_2O}-\nonumber\\
-&&+K_m\left(\frac{X_{H_2O}^{bulk}}{D_{H_2O}(1-X_{H_2O})+X_{H_2O}}\right)^{\gamma}
-\end{eqnarray}
-where $T_{sol}$ is the temperature of the dry solidus, $D_{H_2O}$ is the partition coefficient and $x_t\cdot log(p+y_t)$ indicates the pressure dependence of
-the melting curve \citep{Kelley2010}. $x_t$, $y_t$, $K_m$ and $\gamma$ are parameters chosen in according to \citet{Kelley2010}.
-However, $X_{H_2O}$ depends on the melt fraction as
-\begin{equation}\label{eq:melt_wt}
-X_{H_2O}(M)=\frac{X_{H_2O}^{bulk}}{D_{H_2O}+M(1-D_{H_2O})}
-\end{equation}
-and a numerical solution can be found using a root-finder method \citep{Katz2003,Wang2016}. Furthermore, the content of water in the melt is limited by the
-pressure-dependent saturation concentration of water in the melt, determined as
-\begin{equation}\label{eq:melt_max}
-X_{H_2O}^{sat}=\chi_1 p^{\lambda_m}+\chi_2 p
-\end{equation}
-where $\chi_1$, $\chi_2$ and $\lambda_m$ are parameters chosen according to \citet{Katz2003}, \citet{Langmuir2006} and \citet{Kelley2010}. The correctness in
-the determination of melt fractions and water contents in the melt are tested comparing model results with experimental curves obtained by \citet{Katz2003}
-(Sec. \ref{sec:katz}).
-
-In case of melting, viscous creep viscosity ($\eta_{vp}$) is modified according to \citet{Wang2016}, as
-\begin{equation}\label{eq:melt_visc}
-\eta_{melt}=\eta_{vp}\cdot \exp(\alpha_m M)
-\end{equation}
-where $\alpha_m$ are melt fraction factors for dislocation and diffusion creep, chosen in according to \citet{Wang2016}.
-
-Effective density ($\rho_{eff}$) for partially molten rocks is calculated as
-\begin{equation}\label{eq:melt_dens}
-\rho_{eff}=\rho_s(1-M)+\rho_m M
-\end{equation}
-where $\rho_s$ and $\rho_m$ are the densities of the solid and the molten rock, respectively \citep{Gerya2004b,Gerya2007,Wang2016}. The density of the
-solid rock is extracted by Perple\_X, while the density of the molten portion is calculated as
-\begin{equation}\label{eq:melt_dens_M}
-\rho_m(p,T)=\rho_0[1-\alpha(T-T_0)][1+\beta(p-p_0)]
-\end{equation}
-where $\rho_0$ is the density at the reference temperature ($T_0$) and pressure($p_0$), and $\alpha$ and $\beta$ are the thermal expansion and compressibility
-coefficients, respectively \citep{Gerya2004b,Gerya2007,Wang2016}.
-
-As for the phase transitions, effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into account
-considering effective coefficient of thermal expansion. Similarly, latent heat due to melting/crystallisation can be implicitly included in the energy equation
-(Eq. \ref{eq:energy}) by considering effective specific heat ($C_{p_{eff}}$) and thermal expansion ($\alpha_{eff}$) of the partially molten rocks, in according to
-\begin{eqnarray}
-\label{eq:cp_effective}C_{p_{eff}}&=&C_p+H_L\left(\frac{\partial M}{\partial T}\right)_p \\
-\label{eq:alpha_effective}\alpha_{eff}&=&\alpha+\rho\frac{H_L}{T}\left(\frac{\partial M}{\partial p}\right)_T
-\end{eqnarray}
-where $C_p$ and $\alpha$ are the specific heat and the coefficient of thermal expansion of the solid rock, respectively, and $H_L$ is the latent heat of the
-molten rock \citep{Gerya2004b,Gerya2007,Tackley2010}.
-
-\section{Benchmarks}\label{sec:benchmark}
-\subsection{Error measurements}\label{sec:error}
-In order to determine the accuracy of velocity and pressure field of the benchmarks, the $L_2$-norm is computed by numerical integration on the Gauss-Legendre
-quadrature points. $L_2$-norm for pressure and velocity errors can be evaluated as
-\begin{eqnarray}
-\textrm{err}_p&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|p^n_i(r_q)-p^a_i(r_q)|^2\right)w_q|J_q|}\nonumber \\
-\textrm{err}_v&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|u^n_i(r_q)-u^a_i(r_q)|^2+|v^n_i(r_q)-v^a_i(r_q)|^2\right)w_q|J_q|}\nonumber
-\end{eqnarray}
-respectively, where $n_e$ is the number of elements, $n_q$ is the number of quadrature points per element, $p^n_i(r_q)$ and $p^a_i(r_q)$ are the numerical and
-analytical pressure, respectively, in each quadrature point $q$, $w_q$  and $J_q$ are the weight and the Jacobian at the quadrature point
-$q$, $u^n_i(r_q)$, $v^n_i(r_q)$, $u^a_i(r_q)$ and $v^a_i(r_q)$ are the numerical and analytical velocities, respectively, in each quadrature point $q$.
-
-Other quantities used as comparison with original benchmarks are the root-mean-square velocity over the whole domain and over the surface
-\begin{eqnarray}
-v_{\textrm{rms}}&=&\sqrt{\int_{\Omega}|\bm{v}|^2 d\Omega}\nonumber \\
-v_{\textrm{rms}}^{top}&=&\sqrt{\int_0^1 u^2 \bigg|_{y=1} dx}\nonumber
-\end{eqnarray}
-respectively, top and bottom Nusselt numbers
-\[\textrm{Nu}^{top/bottom}=-\int_0^1\frac{\partial T}{\partial y} \bigg|_{y=1/y=0} dx\]
-the average rate of work done against gravity
-\[\left\langle W \right\rangle=\int_{\Omega}T u_y d\Omega\]
-the average rate of viscous dissipation
-\[\left\langle \Phi \right\rangle=\int_{\Omega}\tau_{ij}\dot{\epsilon}_{ij} d\Omega\]
-and the percentage error between $\left\langle W \right\rangle$ and $\left\langle \Phi \right\rangle/Ra$
-\[\delta=\frac{\left|\left\langle W \right\rangle -\frac{\left\langle \Phi \right\rangle}{Ra} \right|}{\max \left(\left\langle W \right\rangle,
-\frac{\left\langle \Phi \right\rangle}{Ra}\right)}\times100\]
-
-\subsection{Stokes flow}\label{sec:stokes}
-The problem consists of determining velocity field $(u,v)$ and pressure $p$ in case of a manufactured solution with prescribed body forces such as
-\begin{eqnarray}
-b_1&=&(12-24y)x^4+(-24+48y)x^3+(-48y+72y^2-48y^3+12)x^2+\nonumber \\
-&&+(-2+24y-72y^2+48y^3)x+1-4y+12y^2-8y^3\nonumber \\
-b_2&=&(8-48y+48y^2)x^3+(-12+72y-72y^2)x^2+(4-24y+48y^2-48y^3+\nonumber \\
-&&+24y^4)x-12y^2+24y^3-12y^4\nonumber
-\end{eqnarray}
-for which the exact solution is:
-\begin{eqnarray}
-u(x,y)&=&x^2(1-x)^2(2y-6y^2+4y^3)\nonumber \\
-v(x,y)&=&-y^2(1-y)^2(2x-6x^2+4x^3)\nonumber \\
-p(x,y)&=&x(1-x)-\frac{1}{6}\nonumber
-\end{eqnarray}
-The domain is a unit square a constant viscosity ($\eta=1$) and the penalty parameter is set to $\lambda=10^7$. Velocity boundary conditions are set to no slip
-$(\bm{v}=\bm{0})$ on all boundaries. The problem is performed for different grid resolution between \num{8x8} and \num{1024x1024} elements.
-
-The errors between the analytical solution and the numerical prediction of the velocity and the pressure field are calculated as in Section \ref{sec:error}.
-Fig. \ref{fig:errors}a shows that both velocity and pressure field converge to the exact solution with the decrease of the element size, following the
-theoretical convergence rate. The convergence of pressure, in contrast with the observation by \citet{Donea2003} for $Q_1 \times P_0$ elements, indicate the
-effectiveness of the smoothing procedure. Smoothed pressure field for grid resolution of \num{128x128} elements is shown in Fig. \ref{fig:errors}b (continuous
-red line), in comparison with the analytical solution (dashed blue line). Solve times and memory usage needed to generate the solution are shown in
-Fig. \ref{fig:MUMPS}. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Solver/Stokes_Flow}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/errors.pdf}
-\caption{Velocity and pressure error for the Stokes flow experiment between generated and analytical solution as a function of element size (panel a),
-and comparison between smoothed pressure and analytical solution as function of $x$ coordinate for a grid resolution of \num{128x128} elements (panel b).}
-\label{fig:errors}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/MUMPS.pdf}
-\caption{Panel a) Analysis, factorisation, solution and total solve times (panel a) and factorisation memory usage (panel b) as a function of the total number
-of degrees of freedom for the Stokes flow experiment.}
-\label{fig:MUMPS}
-\end{figure}
-
-\subsection{Markers advection}\label{sec:runge}
-The 2nd-order Runge-Kutta advection scheme in space is tested by means of the Zalesak disk test \citep{Zalesak1979,Thieulot2014}. The benchmark is performed
-in a unit square domain with a grid resolution of \num{32x32} elements and 1000000 markers and values of Courant number between 0.25 and 1.
-At $t=0$, the disk is centred at position $(0.5;0.75)$ with a radius $R=0.15$ and has a vertical fissure 0.05 wide and 0.2 high (orange in Fig. \ref{fig:runge})
-and with the velocity field prescribed in the entire domain as
-\begin{eqnarray}
-u(x,y)&=&2\pi\left(y-\frac{L_x}{2}\right)\nonumber \\
-v(x,y)&=&-2\pi\left(x-\frac{L_x}{2}\right)\nonumber
-\end{eqnarray}
-the markers are back to their initial location after a $2\pi$ rotation ($t=1$). The clockwise rotation of the disk at $t=0.25$, 0.5 and 0.75 
-is shown in Fig. \ref{fig:runge} (red, blue and green, respectively). The distance from the centre of 1 marker is calculated throughout the rotation to
-evaluate the error for different values of Courant number. As expected, the error increase with the increasing of the time step, as shown in 
-Fig. \ref{fig:runge_err}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/Zalesak_Disk}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=200px]{./Figures/Zalesak.pdf}
-\caption{Position of the Zalesak disk at $t=0$, 0.25, 0.5 and 0.75 (orange, red, blue and green, respectively) for the case with Courant number of 1.}
-\label{fig:runge}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/zal_marker.pdf}
-\caption{Distance from the centre as function of time for values of Courant number of  of 0.25, 0.3, 0.5, 0.75 and 1 (green, orange, blue, black and red,
-respectively). R indicates the distance from the centre considering a perfect circle trajectory.}
-\label{fig:runge_err}
-\end{figure}
-
-\subsection{Conservative Velocity Interpolation (CVI)}\label{sec:CVI}
-The CVI correction is checked by means of the Stokes flow experiment by \citet{Donea2003}, as presented in \ref{sec:stokes}. The advection of Lagrangian
-markers is performed using either a 2nd-order or a 4th-order Runge-Kutta scheme with CFL=0.5 and an initial random distribution of 25 markers per element.
-Fig. \ref{fig:CVI} shows the minimum and maximum values of markers per element throughout the experiment with and without the CVI correction and using
-different Runge-Kutta schemes. The introduction of the CVI correction allows the
-tests to run for times longer than $t=2000$ without show empty elements, while tests without the CVI correction stop at approximately $t=600$, when at least
-one element goes empty. Fig. \ref{fig:CVI_time} shows the different number of markers per element and the markers distribution at $t=600$ obtained without and
-with the CVI correction (panels b and c, respectively) in case of a 2nd-order Runge-Kutta scheme. All data can be found at
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/CVI}.
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/CVI.pdf}
-\caption{Maximum and minimum number of markers per element throughout the experiment, with (green and orange lines) or without (purple and blue lines)
-the CVI correction and using either a 2nd-order (blue and orange lines) or a 4th-order (purple and green lines) Runge-Kutta scheme.}
-\label{fig:CVI}
-\end{figure}
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/CVI_time.pdf}
-\caption{Number of markers per element and markers distribution through time using 2nd.order Runge-Kutta scheme.
-Panels a: number of markers per element (panel a1) and distribution of the markers (panel a2) after the first
-iteration; panels b and c: comparison between number of markers per element (panels b1 and c1) and markers distribution (panels b2 and c2) without and with
-the CVI correction at $t=600$; panels d: number of markers per element (panel d1) and markers distribution (panel d2) in case of CVI correction at $t=2000$.}
-\label{fig:CVI_time}
-\end{figure}
-
 \subsection{Poiseuille flow}\label{sec:poiseuille}
 The domain is a rectangle with $L_x=2$ and $L_y=1$ and constant density and viscosity ($\rho=1$ and $\eta=1$), gravity acceleration $\bm{g}=\bm{0}$ and penalty
-parameter $\lambda=10^8$. The grid is composed by \num{40x20} elements. Velocity boundary conditions are set to no slip $(\bm{v}=\bm{0})$ at the top and the
+parameter $\lambda=10^8$. The grid is composed by $40 \times 20$ elements. Velocity boundary conditions are set to no slip $(\bm{v}=\bm{0})$ at the top and the
 bottom, and a parabolic profile is imposed on the sides, with $u=y(1-y)$ and $v=0$. The analytical solution is then given by:
@@
 \end{figure}
-
-\subsection{Instantaneous 2D sphere}\label{sec:ist_sphere}
-The domain is a unit square with gravity $\bm{g}=(0,-1)$. The fluid has constant density and viscosity ($\rho_f=1$ and $\eta_f=1$). The sphere is in the middle
-of the domain with a radius R = 0.123456798 and has constant density and viscosity ($\rho_s=10^{-2}+\rho_f$ and $\eta_s=10^3 \cdot \eta_f$).
-We distinguish three different types of velocity boundary conditions:
-\begin{enumerate}
-\item FS: free slip conditions on all sides.
-\item NS: no slip conditions on all sides.
-\item OT: free slip conditions on the sides and the bottom, and open boundary at the top.
-\end{enumerate}
-All three cases are performed for different grid resolutions, between \num{16x16} and \num{512x512} elements with 50 randomly distributed markers per element,
-and different average schemes for the viscosity (harmonic, geometric and arithmetic). The velocity in the centre of the sphere $\bm{v}(0.5;0.5)$), minimum and
-maximum velocities ($u_{min}$, $u_{max}$, $v_{min}$, $v_{max}$) and pressures ($p_{min}$, $p_{max}$) on the entire domain, $v_{\textrm{rms}}$ and average pressure on the
-entire domain ($p_{avg}$) are compared with solutions generated by ASPECT \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and all results
-can be found at \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere2D} and 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Instantaneous_2D_Sphere}. Results in terms of the $v_{\textrm{rms}}$ for NS model 
-are also shown in Fig. \ref{fig:inst_sphere}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_NS.pdf}
-\caption{$v_{\textrm{rms}}$ as function of element size for the instantaneous 2D sphere experiment in case of no slip boundary conditions and with different average
-schemes for the viscosity. Results are compared with results obtained by other numerical codes.}
-\label{fig:inst_sphere}
-\end{figure}
-
-\subsection{Rayleigh-Taylor instability}\label{sec:rayleigh}
-This problem has been originally presented by \citet{vanKeken1997} and here it is performed as the isoviscous case (Case 1a). The domain has $L_x=0.9142$ and
-$L_y=1$ and gravity $\bm{g}=(0,-1)$. Two fluids with same constant viscosities ($\eta_1=\eta_2=1$) and different densities ($\rho_1=1000$ and $\rho_2=1010$),
-with the lighter fluid at the bottom. The initial interface between the fluids is given by $y(x)=0.2+0.02 \cos \left(\frac{\pi x}{L_x}\right)$. The experiment
-is performed with different grid sizes (\num{50x50}, \num{80x80}, \num{100x100} and \num{256x256}). A total of 1960000 markers are randomly distributed at the
-beginning of the simulation. Velocity boundary conditions are set to no slip at the top and the bottom, and to free slip at the sides of the domain.
-
-$v_{\textrm{rms}}$ as function of time is reported in Fig. \ref{fig:RT}, matching well with results shown by \citet{vanKeken1997}, \citet{Tackley2003} and
-\citet{Thieulot2014}. Fig. \ref{fig:rayleigh} shows the evolution of the experiment at different time steps. All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Rayleigh_Taylor_experiment/ISOVISCOUS}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/RT.pdf}
-\caption{$v_{\textrm{rms}}$ of the Rayleigh-Taylor experiment as function of time for different resolution of the grid.}
-\label{fig:RT}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=300px]{./Figures/Rayleigh.pdf}
-\caption{Evolution of the Rayleigh-Taylor experiment for a grid of \num{256x256} elements at $t=0, 500, 1000$ and $2000$ (panels a, b, c and d, respectively).}
-\label{fig:rayleigh}
-\end{figure}
-
-\subsection{Falling blocks}\label{sec:block}
-This benchmark is described as proposed by \citet{Gerya2003a}, \citet{Gerya2010b} and \citet{Thieulot2011}. The domain is square with $L_x=L_y=$ \SI{500}{\km}
-and the grid is composed by \num{50x50} elements with 25 markers in each element. The block is initially centred at ($x=$ \SI{250}{\km}; $y=$ \SI{400}{\km})
-and has a size of \num{100x100} km (Fig. \ref{fig:block}a). The fluid surrounding the block has $\eta_f=$ \SI{e21}{\pascal\per\s} and $\rho_f=$
-\SI{3200}{\kg\per\cubic\m}. The benchmark tests with different viscosities of the block, with $\eta_b$ from \SIrange{e15}{e27}{\pascal\per\s}. In each
-experiment also the density $\rho_b$ of the block varies from \SIrange{3220}{9900}{\kg\per\cubic\m}. Velocity boundary conditions are set to free slip
-conditions on all sides of the domain. Density distributions at $t=$ \SI{20}{\mega\year} for $\rho_b=$ \SI{3300}{\kg\per\cubic\m} and different $\eta_b$
-are plotted in Fig. \ref{fig:block}b-f, showing that the code correctly preserve the block geometry in case of large viscosity contrast, with a block stiffer
-than the surrounding fluid (Fig. \ref{fig:block}f).
-
-The velocity in the centre of the falling block at $t=0$ is measured for all experiments and, since the velocity should increase with density contrast, the
-quantity $\bm{v}/(\rho_b-\rho_f)$ is plotted as function of the viscosity contrast. All results perfectly match with those from \citet{Gerya2010b} and they
-line up on a single curve, demonstrating that the code can correctly deal with large viscosity and density contrasts (Fig. \ref{fig:falling}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Falling%20blocks}.
-
-\begin{figure}
-\noindent\includegraphics[width=300px]{./Figures/Block.pdf}
-\centering
-\caption{Density evolution of the falling block experiment at $t=$ \SI{0}{\mega\year} (panel a) and $t=$ \SI{20}{\mega\year} for different viscosities of the
-block (panels b-f).}
-\label{fig:block}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Falling.pdf}
-\caption{Initial velocity relative to the density contrast at the centre of the falling block as function of the viscosity contrast between the block and the
-surrounding fluid.}
-\label{fig:falling}
-\end{figure}
-
-\subsection{2D time-dependent Stokes sphere with deformable free surface}\label{sec:sphere}
-This experiment is performed in a unit square domain with the gravity acceleration fixed to $g_y=-1$. The sphere has $\rho_s=2$ and $\eta_s=10^3$ and is
-initially centred at ($0.5;0.6$) with radius $R=0.123456789$. The fluid surrounding the sphere has $\rho_f=1$ and $\eta_f=1$ and occupies the domain for
-$y\leq 0.75$, while for $y>0.75$ the air has $\rho_a=0$ and $\eta_a=10^{-3}$. Velocity boundary conditions are set to free slip on all sides. The Courant
-number is set to 0.25. The experiments run for 200 s using grid resolutions from \num{150x150} to \num{512x512} elements, with 25 randomly distributed markers
-per element, and for different average schemes for the viscosity. The interface between the fluid and the air is tracked by means of the markers chain. Results
-in terms of velocity and pressure fields and topography variations are compared with results obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and they can be found at
-\url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere_fs2D} and
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Time_dependent_sphere}. Results in terms of the $v_{\textrm{rms}}$ in case of 
-an arithmetic average are also shown in Fig. \ref{fig:inst_spherefs}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_arithm.pdf}
-\caption{$v_{\textrm{rms}}$ for different grid resolutions as function of time for the 2D time-dependent Stokes sphere experiment in case of an arithmetic average.
-Results are compared with results obtained by ASPECT.}
-\label{fig:inst_spherefs}
-\end{figure}
-
-\subsection{Free surface stabilisation}\label{sec:stab}
-The benchmark is performed as discussed in \citet{Kaus2010a} and \citet{Thieulot2014}. The domain is square with $L_x=L_y=$ \SI{500}{\km} and a grid resolution
-of \num{200x200} elements, each of them containing 16 markers. A buoyant fluid with $\rho_1=$ \SI{3200}{\kg\per\cubic\metre} and $\eta_1=$ \SI{e20}{\pascal\s}
-is overlain by a denser fluid with $\rho_2=$ \SI{3300}{\kg\per\cubic\metre} and $\eta_2=$ \SI{e21}{\pascal\s}. The initial interface between the fluids has an
-initial sinusoidal shape of 5 km amplitude. Velocity boundary conditions are set to free slip conditions at the sides and no slip at the bottom, while the top
-is free surface. The experiment is performed with various fixed time steps, without the use of the Courant number. The vertical position of the free surface at
-$x=L_x$ is tracked for each simulation, with or without the stabilisation algorithm.
-
-The results show an instability (drunken sailor effect) increasing the time step in case that the stabilisation algorithm is not activated
-($dt=$ \SI{4200}{\year}, green line in Fig. \ref{fig:stab}), as already observed by \citet{Kaus2010a} and \citet{Thieulot2014}. Activating the algorithm the
-instability is fixed and the simulation remains stable using time step up to \SI{50000}{\year} (red line in Fig. \ref{fig:stab}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Stabilisation_algorithm}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Kaus_random.pdf}
-\caption{Evolution of the y coordinate in $x=L_x$ as function of time for different time steps, with and without the stabilisation algorithm.}
-\label{fig:stab}
-\end{figure}
-
-\subsection{Topography relaxation}\label{sec:crameri}
-This experiment is performed as in \citet{Crameri2012} using free surface and sticky air (Case 1). The domain is rectangular with $L_x=$ \SI{2800}{\km} and
-$L_y=$ \SI{700}{\km} in case of free surface and $L_y=$ \SI{800}{\km} in case of sticky air ($h_{st}=$ \SI{100}{\km}). The grid is composed by \num{256x64}
-elements for the free surface \num{560x320} elements for the sticky air. For both cases, there is a mantle of 600 km thickness with $\eta_m=$ \SI{e21}{\pascal\s}
-overlain by a highly viscous cosine-shaped lithosphere with thickness between 93 and 107 km and $\eta_l=$ \SI{e23}{\pascal\s}. Both mediums have $\rho=$
-\SI{3300}{\kg\per\cubic\metre}. In case of sticky air experiment, the air has $\rho_a=$ \SI{0}{\kg\per\cubic\metre} and $\eta_a=$ \SI{e18}{\pascal\s}. The
-gravity acceleration is set to $g_y=$ \SI{-10}{\metre\per\square\s}. Velocity boundary conditions are set to no slip at the bottom and free slip at the sides
-of the domain. In case of sticky air, velocities on top are set to free slip conditions.
-
-The maximum topography as function of time ($h(t)$) can be analytically derived in according to \[h(t)=h_0 \exp(\gamma t)\]
-where $h_0=7$ km is the initial topography, $\gamma=\num{-0.2139e-11}$ is the characteristic relaxation rate and $t$ is the time (black line in Fig.
-\ref{fig:crameri}). Maximum topography at the characteristic relaxation time $t_{rlx}=$ \SI{14.825}{\kilo\year} can be found to be $h_{rlx}=$ \SI{2576}{\m}.
-The results show that a viscosity of the air of \SI{e18}{\pascal\s} is needed in case of $h_{st}=$ \SI{100}{\km} to correctly describe the topography relaxation
-for this problem, as pointed out by \citet{Crameri2012} (purple line in Fig. \ref{fig:crameri}). The case with a true free surface well follows the analytical
-solution (orange line in Fig. \ref{fig:crameri}). In particular, at the relaxation time in case of the true free surface predicts a maximum topography of
-\SI{2570}{\m}, with an error of \SI{6}{\m} with respect to the analytical solution. Results are also compared with results obtained with a free surface in
-UNDERWORLD (\num{256x64} elements) and SULEC (\num{401x201} elements) (with courtesy from \citealp{Crameri2012}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Topography_relaxation}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Crameri.pdf}
-\caption{Maximum topography as function of time for the Crameri benchmark (coloured lines) in comparison with the analytical solution (black line) and results
-obtained with UNDERWORLD (yellow line) and SULEC (blue line).}
-\label{fig:crameri}
-\end{figure}
-
-\subsection{Spontaneous subduction}\label{sec:subduction}
-This experiment is performed as presented by \citet{Schmeling2008}, in case of both a sticky air and a true free surface and using different average schemes
-for the viscosities. The domain is rectangular with $L_x=$ \SI{3000}{\km} and $L_y=$ \SI{700}{\km}, a grid resolution of \num{256x64} elements with 50 markers
-per element and Courant number of 0.01. At the beginning of the simulation, a 100 km-thick lithospheric layer with $\rho_m=$ \SI{3300}{\kg\per\cubic\metre}
-and $\eta_m=$ \SI{e23}{\pascal\s} is located at the top of the domain for $x$ between 1000 and 3000 km, while the asthenospheric mantle has $\rho_m=$
-\SI{3200}{\pascal\s} and $\eta_m=$ \SI{e21}{\pascal\s}. In addition, a 200 km-depth lithospheric slab is already subducted in the mantle in order to have
-a spontaneous subduction. In case of sticky air, $L_y=750$ km and a 50 km-thick air layer with $\eta_a$ from \SIrange{e19}{e21}{\pascal\s} overlie the mantle.
-Velocity boundary conditions are set to free slip on the sides and on the bottom of the domain. In case of sticky air, velocities on the top boundary are set
-to free slip conditions as well.
-
-The maximum depth of the slab is tracked for all the simulations (Fig. \ref{fig:subduction}) and they are compared with results obtained by
-\citet{Schmeling2008} with models with regular grids and comparable grid resolutions (rectangles in Fig. \ref{fig:subduction}). Low viscous sticky air models
-(i.e., \SI{e19}{\pascal\s}) enlighten a strong dependence of the sinking velocity with the chosen average scheme (dashed lines in Fig. \ref{fig:subduction}),
-while high viscous sticky air models (i.e., \SI{e21}{\pascal\s}) show a higher resistance at the trench, underestimating the correct solution (dotted lines in
-Fig. \ref{fig:subduction}). A high resistance at the trench can be also observed in case of a true free surface for the low resolution grid used (continuous
-lines in Fig. \ref{fig:subduction}), as pointed out by \citet{Schmeling2008}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Spontaneous_subduction}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Subduction.pdf}
-\caption{Maximum depth of the slab as function of time for all the simulations of the \citet{Schmeling2008} benchmark, in case of harmonic, geometric and
-arithmetic means (red, green and blue lines, respectively) using a sticky air or a true free surface (discontinuous and continuous lines, respectively).
-Red, green and blue rectangular areas indicate the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{500}{\km} in case of sticky air,
-while the grey rectangular area indicates the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{600}{\km} in case of true free surface.}
-\label{fig:subduction}
-\end{figure}
-
-\subsection{The slab detachment}\label{sec:slab}
-The slab detachment benchmark is performed as described by \citet{Schmalholz2011} and \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{1000}{\km}
-and $L_y=$ \SI{660}{\km} and a grid resolution of \num{256x256} elements. A non-linear viscous T-shaped layer with $\rho_l=$ \SI{3300}{\kg\per\cubic\metre} is
-placed at the top of the domain and surrounded by a linear viscous fluid with $\rho_f=$ \SI{3150}{\kg\per\cubic\m}. The top layer is 80 km-thick and a 250
-km-long and 80 km-wide slab is placed at $x=L_x/2$ (Fig. \ref{fig:slab}a). The effective viscosity of the top layer is given by
-\[\eta_{eff}=\eta_0 I_2^{\frac{(1-n)}{n}}\]
-with $\eta_0=$ \SI{4.75e11}{\pascal\s} and $n=4$, while the surrounding fluid has $\eta_f=$ \SI{e21}{\pascal\s}. The effective viscosity is capped between
-$\eta_{min}=$ \SI{e21}{\pascal\s} and $\eta_{max}=$ \SI{e25}{\pascal\s}. Velocity boundary conditions are set to free slip at the top and the bottom, and to
-no slip at the sides of the domain.
-
-The sinking of the slab determines a decrease of the effective viscosity in the T-shaped layer during the first part of the evolution (Fig. \ref{fig:slab}a),
-as consequence of local strain rates, while effective viscosities increase up to $\eta_{max}$ in the top layer after approximately \SI{20}{\mega\year}, when
-necking is complete (Fig. \ref{fig:slab}b).
-
-The width of necking is tracked by means of markers position at the side of the slab. Necking width and time are normalised in function of the initial slab
-(\SI{80}{\km}) and the characteristic time ($t_c=$ \SI{7.1158e14}{\s}) \citep{Schmalholz2011,Glerum2018}. The results in case of arithmetic, geometric and
-harmonic mean of the viscosity are shown in Fig. \ref{fig:necking}, compared with results from \citet{Glerum2018} for which an infinity norm mean was used.
-The influence of the average schemes agrees with results shown in Section \ref{sec:subduction}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Slab_detachment}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Slab.pdf}
-\caption{Effective viscosity for the slab detachment benchmark at the beginning of the evolution (panel a) and when necking is complete (panel b).}
-\label{fig:slab}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Necking.pdf}
-\caption{Normalised width of the necking of the slab detachment benchmark as function of normalised time.}
-\label{fig:necking}
-\end{figure}
-
-\subsection{The indenter experiment}\label{sec:indenter}
-The indenter benchmark simulate a rigid punch on a purely plastic von Mises material, for which exists an analytical solution
-\citep{Thieulot2008,Thieulot2014,Glerum2018}. The domain is a unit square with a grid resolution of \num{256x256} elements. The gravity acceleration is set
-to $\bm{g}=\bm{0}$ and effective viscosity is capped using $\eta_{min}=10^{-4}$ and $\eta_{max}=10^3$. The tolerance for the convergence of the non-linear
-solution is $tol=10^{-9}$, with a maximum number of non-linear iterations set to 500. The medium in the domain has $\rho=0.01$, an initial viscosity
-$\eta_0=10$, cohesion $C=1$ and an angle of internal friction $\phi=0$°. Velocity boundary conditions are set to no slip at the bottom and free slip at the
-sides of the domain. The top of the domain is open with the exception of the central portion (punch area) with width $w_p=0.125$, where $v=-1.05$ and $u$ is
-fixed either to 0 (rough punch experiment) or free (smooth punch experiment).
-
-Fig. \ref{fig:indenter} show the results in terms of viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure
-(panels g and h) for both experiments. The analytical solution indicate that pressure at ($0.5;1$) and ($0.5 \pm w_p;1$)is $p=\pi +1$ and $p=1$, respectively
-(black and red lines in Fig. \ref{fig:smooth_p}). Fig. \ref{fig:smooth_p} shows a clear improvement in the pressure solution when passing from a rough to a
-smooth punch (green and blue lines in Fig. \ref{fig:smooth_p}, respectively), as previously observed by \citet{Thieulot2014} and \citet{Glerum2018}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Indenter}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Indenter.pdf}
-\caption{Viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure (panels g and h) fields for rough (left column) and
-smooth (right column) punch experiments.}
-\label{fig:indenter}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Smooth_Rough.pdf}
-\caption{Nodal pressure in function of the x coordinate at $y=1$ for rough (green line) and smooth (blue line) punch experiment. Black and red lines indicate
-the analytical solution at $x=0.5$ and $x=0.5 \pm w_p$, respectively.}
-\label{fig:smooth_p}
-\end{figure}
-
-\subsection{The brick experiment}\label{sec:brick}
-An instantaneous version of the brick benchmark is performed to verify the correctness of the pressure-dependent plasticity for different angle of internal
-friction, as proposed by \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{40}{\km} and $L_y=$ \SI{10}{\km} and a grid resolution of \num{512x128}
-elements. The gravity acceleration is set to $g_y=$ \SI{-10}{\m\square\s} and effective viscosity is capped using $\eta_{min}=$ \SI{e19}{\pascal\s} and
-$\eta_{max}=$ \SI{e26}{\pascal\s}. The tolerance for the convergence of the non-linear solution is $tol=10^{-7}$, with a maximum number of non-linear iterations
-set to 1000. A 800 m-wide and 400 m-height inclusion with $\eta_b=$ \SI{e20}{\pascal\s} and $\rho_b=$ \SI{2700}{\kg\per\cubic\m} is placed at the bottom of the
-domain at $x=L_x/2$. The inclusion is surrounded by a non-linear viscous medium with $\rho_m=$ \SI{2700}{\kg\per\cubic\m}, an initial viscosity of
-\SI{e23}{\pascal\s}, a linear viscous viscosity of \SI{e25}{\pascal\s} and a cohesion of \SI{40}{\mega\pascal}. Velocities are set to free slip conditions at
-the bottom of the domain and the top is open. Velocities on sides of the domain are fixed to $u=$ \SI{+-2e-11}{\m\per\s} and $v=0$. The experiment is performed
-in compressional and extensional contexts with an angle of internal friction $\phi$ of the non-linear medium variable between 0° and 30°.
-
-As expected, two shear bands stem from the inclusion with variable angles in relation to both the dynamics context and the internal friction angle. Shear band
-angles formed at 45° for $\phi=0$° in both extensional and compressional contexts (Fig. \ref{fig:brick_beam}a), while they are different in case of $\phi\neq0$°.
-Shear band angles in case of $\phi=20$° are shown for compression and extension in Fig. \ref{fig:brick_beam}b  and c, respectively. Values of shear band angles
-as function of internal friction angles from 0° to 30° are extracted at different depths and minimum and maximum values are plotted in Fig. \ref{fig:brick},
-in comparison with theoretical Roscoe, Arthur and Coulomb shear band angles. For all tests the tolerance for nonlinear convergence is defined $tol=10^{-7}$
-and none converges before the maximum number of iterations ($it_{max}=1000$) is reached. However, tests with internal friction angles up to 15° show a constant
-decrease of the velocity residuals, which can not be observed for for tests with higher internal friction angles (Fig. \ref{fig:convergence}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Brick_experiment}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Brick_Beam.pdf}
-\caption{Shear band angles predicted for the brick experiment for $\phi=0$° (panel a) and for $\phi=20$° in case of compression and extension (panels b and c,
-respectively).}
-\label{fig:brick_beam}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Brick.pdf}
-\caption{Shear band angles predicted for the brick experiment in case of compressional and extensional contexts as function of different internal angle of
-friction (continuous black line and red dots), compared with theoretical Roscoe, Artur and Coulomb angles (discontinuous black lines). Red lines indicate the
-range of angles calculated at different depths.}
-\label{fig:brick}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \noindent\includegraphics[width=500px]{./Figures/convergence.pdf}
-  \caption{Convergence rates for all tests. Normalised residuals (left column) are compared with velocity residuals (right column) for extensional (first row)
-  and compressional (second row) settings.}
-  \label{fig:convergence}
-  \end{figure}
-
-\subsection{Advection stabilisation}\label{sec:advection}
-The 1D advection problem proposed by \citet{Donea2003} and \citet{Thieulot2011} is performed to verify the effectiveness of the SUPG method to stabilise the
-advection term of the energy equation. The domain is a 1D segment with $L_x=1$ composed by 50 elements and a discontinuity in the thermal field placed at
-$x=0.25$. Temperature is set to 1 for $x \leq 0.25$ and to 0 for $x > 0.25$. Velocity is set to ${u}=1$ in the entire domain. The simulation is performed for
-250 time steps, with $dt=0.002$, so the thermal discontinuity should be at $x=0.75$ at the end of the simulation.
-
-Temperature profiles at the end of the simulation are shown in \ref{fig:advection} as function of the dimensionless coefficient $\gamma=\tau \bm{v}/h$. In case
-of the classic Galerkin method ($\gamma=0$, blue line in \ref{fig:advection}) the final thermal profile is characterised by strong oscillations, which are
-eliminated in case of the SUPG method ($\gamma=0.045$, orange line in \ref{fig:advection}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Advection_stabilisation}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Advection.pdf}
-\caption{Temperature profile as function of \textit{x} for the advection stabilisation benchmark. Purple line indicates the initial temperature profile; the
-green line indicates the analytical temperature profile after 250 time steps; blue line indicates the temperature profile after 250 time steps in case of the
-classic Galerkin method ($\gamma=0$); orange line indicates the temperature profile after 250 time steps in case of the SUPG method ($\gamma=0.045$).}
-\label{fig:advection}
-\end{figure}
-
-\subsection{Shear heating}\label{sec:simple_shear}
-This exercise is performed in an unit square domain composed by \num{128x128} elements. The velocity field is prescribed on the entire domain with
-$\bm{v}=(L_y-y)y\bm{e_x}$; viscosity, density and specific heat are fixed to 1, while thermal conductivity and radiogenic energy are fixed to 0. Therefore,
-the energy equation (Eq. \ref{eq:energy}) can be simplified as \[\frac{\partial T}{\partial t}=H_s\] and fixing $T(t=0)=0$, the temperature can be find as
-\[T(t)=H_s t\]
-In this case we have
-\begin{eqnarray}
-\dot{\epsilon_{xx}}&=&\frac{\partial u}{\partial x}=0 \nonumber \\
-\dot{\epsilon_{yy}}&=&\frac{\partial v}{\partial y}=0 \nonumber \\
-\dot{\epsilon_{xy}}&=&\frac{1}{2}\left(\frac{\partial v}{\partial x}+\frac{\partial u}{\partial y}\right)=\frac{1}{2}(L_y-2y)\nonumber
-\end{eqnarray}
-and, simplifying Eq. \ref{eq:shear}, shear heating can be calculated as \[H_s(x,y)=(1-2y)^2\]
-
-The solution predicted by the model in terms of velocity, temperature and shear heating match well with the analytical solutions (Fig. \ref{fig:analytical_en}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Simple_shear}.
-
-%\begin{figure}
-%\centering
-%\noindent\includegraphics[width=400px]{./Figures/Energy_Thieulot.pdf}
-%\caption{Temperature and shear heating predicted by the model after 1 time step.}
-%\label{fig:en_thieulot}
-%\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=500px]{./Figures/Analytical_Shear.pdf}
-\caption{Velocity (panel a), temperature (panel b) and shear heating (panel c) as function of the \textit{y} coordinate for the simple shear experiment. The
-solutions predicted by the model (red lines) are compared with the analytical solutions (dashed blue lines).}
-\label{fig:analytical_en}
-\end{figure}
-
-\subsection{Shear and adiabatic heating}\label{sec:shear}
-This problem is performed as presented in Exercise 9.4 in \citet{Gerya2010b}. Two materials are vertically separated in a rectangular domain with $L_x=$
-\SI{1000}{\km}, $L_y=$ \SI{1500}{\km} and a grid resolution of \num{30x20} elements. Constant thermal coefficient expansion ($\alpha=$ \SI{3e-5}{\per\kelvin}),
-temperature ($T=$ \SI{1300}{\kelvin}) and gravity acceleration ($g_y=$ \SI{-10}{\m\square\s}) are assumed in the whole domain. Fluid 1 (on the left side of the
-domain) has $\rho_1=$ \SI{3200}{\kg\per\cubic\m} and $\eta_1=$ \SI{e20}{\pascal\s}; fluid 2 (on the right side of the domain) has $\rho_2=$
-\SI{3300}{\kg\per\cubic\m} and $\eta_2=$ \SI{e22}{\pascal\s}. Velocity boundary conditions are set to free slip on all sides of the domain.
-
-As shown in Fig. \ref{fig:energy}, both shear and adiabatic heating predicted by the code (first row in Fig. \ref{fig:energy}) well recreate the results
-obtained by means of example ${Shear\_adiabatic\_heating.m}$ from \citet{Gerya2010b} (second row in Fig. \ref{fig:energy}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Adiabatic%2BShear}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Energy.pdf}
-\caption{Comparison between shear (first column), adiabatic (second column) and total (third column) energy predicted by the code (panels a, b and c) and those
-created using example ${Shear\_adiabatic\_heating.m}$ from Exercise 9.4 in \citet{Gerya2010b} (panels d, e and f).}
-\label{fig:energy}
-\end{figure}
-
-\subsection{Mantle convection}\label{sec:mantle}
-This problem is performed as presented by \citet{Blankenbach1989} (constant viscosity cases) and \citet{Thieulot2014}, in a 2D unit square domain with gravity
-acceleration $g_y=10^{10}Ra$. The experiment is performed with three different Rayleigh numbers ($Ra=10^4$, $10^5$ and $10^6$) and with different grid
-resolution (between \num{32x32} and \num{128x128} elements). The fluid has constant viscosity, initial density, heat capacity, thermal conductivity
-($\eta= \rho_0= C_p = k=1$), reference temperature ($T_0=0$) and thermal expansion coefficient ($\alpha=10^{-10}$). Temperatures are set to 0 on top and 1 on
-bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\]
-
-The solution generated by the code in terms of $v_{\textrm{rms}}$ (calculated as in Sec. \ref{sec:error}) and Nu as function of time are reported for all the
-simulations in Fig. \ref{fig:mantle}. At the steady state, $v_{\textrm{rms}}$ and Nu of all simulations converge well toward the values from \citet{Blankenbach1989},
-with lower errors for higher resolution grids (Table \ref{tab:mantle}). All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Mantle_convection}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Convection.pdf}
-\caption{$v_{\textrm{rms}}$ (panels a, c and e) and Nu (panels b, d and f) for the mantle convection benchmark as function of time for different grid resolution.
-Panels a and b show the results for $Ra=10^4$; Panels c and d show the results for $Ra=10^5$; panels e and f show the results for $Ra=10^6$. Purple lines
-indicate the convergence values for the $v_{\textrm{rms}}$ and the Nu from \citet{Blankenbach1989} at the steady state.}
-\label{fig:mantle}
-\end{figure}
-
-\begin{table}[H]
-\caption{Comparison between $v_{\textrm{rms}}$ and Nu predicted by the code for the mantle convection experiment and same values as reported in literature.}
-\centering
-\small
-\begin{tabular}{|>{\centering}m{0.09\textwidth}|>{\centering}m{0.05\textwidth}|>{\centering}m{0.24\textwidth}|>{\centering}m{0.15\textwidth}|
-  >{\centering}m{0.1\textwidth} >{\centering}m{0.1\textwidth} >{\centering\arraybackslash}m{0.1\textwidth}|}
-\toprule
-& & \multirow{2}{*}{\citet{Blankenbach1989}} & \citet{Thieulot2014} & \multicolumn{3}{c|}{FALCON}  \\
- & & & \num{200x200} & \num{32x32} & \num{64x64} & \num{128x128}  \\
-\midrule
-\multirow{2}{*}{$Ra=10^4$} & $v_{\textrm{rms}}$ & $42.864947 \pm 0.000020$ & 42.867 & 42.83226 & 42.852793 & 42.861394  \\
-           & Nu      & $4.884409 \pm 0.000010$  & 4.882  & 4.781297 & 4.857475  & 4.877573   \\
-\hline
-\multirow{2}{*}{$Ra=10^5$} & $v_{\textrm{rms}}$ & $193.21454 \pm 0.00010$  & 193.255& 193.872643 & 193.377472 & 193.252290  \\
-           & Nu      & $10.534095 \pm 0.000010$ & 10.507 & 9.602514 & 10.270735  & 10.465629   \\
-\hline
-\multirow{2}{*}{$Ra=10^6$} & $v_{\textrm{rms}}$ & $833.98977 \pm 0.00020$  & 834.712& 848.091176 & 837.767911 & 834.945793  \\
-           & Nu      & $21.972465 \pm 0.000020$ & 21.695 & 15.999266 & 19.703682  & 21.306939   \\
-\bottomrule
-\end{tabular}
-\label{tab:mantle}
-\end{table}
-
-\subsection{Viscoplastic mantle convection}\label{sec:visco_mantle}
-This benchmark is performed as Cases 1-5a described in \citet{Tosi2015} in a 2D unit square domain with two different grid resolutions of \num{32x32} 
-and \num{100x100} elements. Reference viscosity, reference density, heat capacity, thermal conductivity and thermal expansion coefficient are fixed to a constant
-value of 1 ($\eta_0= \rho_0= C_p = k=\alpha=1$), while the gravity has been chosen as $g_y=10^2$, in order to obtain a Rayleigh number $Ra=10^2$. Temperatures
-are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\].
-The viscosity field $\eta$ is calculated as the harmonic average between a linear part $\eta_{lin}$ that depends on temperature only or on temperature and 
-depth $y$ and a nonlinear-plastic part $\eta_{pl}$ that depends on the strain rate $\dot{\bm{\epsilon}}$, as follows
-\[\eta(T,y,\dot{\bm{\epsilon}})=2\left(\frac{1}{\eta_{lin}(T,y)}+\frac{1}{\eta_{pl}(\dot{\bm{\epsilon}})}\right)^{-1}\]
-The linear and the nonlinear-plastic parts of the viscosity are calculated following \citet{Tosi2015} as follows
-\[\eta_{lin}(T,y)=\exp(-\ln(\Delta\eta_T)T+\ln(\Delta\eta_y)y)\]
-\[\eta_{pl}(\dot{\bm{\epsilon}})=\eta^*+\frac{\sigma_Y}{\sqrt{\dot{\bm{\epsilon}}:\dot{\bm{\epsilon}}}}\]
-where $\Delta\eta_T$, $\Delta\eta_y$, $\eta^*$ and $\sigma_Y$ are parameters chosen as listed in Table \ref{tab:tosi_setup}.
-
-For each case, some diagnostic quantities measured at the steady state are reported in Table \ref{tab:visco_mantle} (see Section \ref{sec:error} for 
-the computation of all the quantities), showing a well fit with results described in \citet{Tosi2015}. Data and charts for all cases can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Viscoplastic_mantle_convection}.
-
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{white!100!}{gray!25!}
-%\begin{table}[H]
-%  \caption{Results of the viscoplastic mantle convection benchmark for Cases 1-4 compared with results obtained with ELEFANT and YACC (from \citealp{Tosi2015}).}
-%  \centering
-%  \resizebox{\textwidth}{!}{
-%\begin{tabular}{|c|ccc|ccc|ccc|ccc|}
-%  \toprule
-%    & \multicolumn{3}{c|}{\textbf{Case 1}} & \multicolumn{3}{c|}{\textbf{Case 2}} & \multicolumn{3}{c|}{\textbf{Case 3}} &
-%    \multicolumn{3}{c|}{\textbf{Case 4}} \\
-%   & FALCON & ELEFANT& YACC & FALCON& ELEFANT& YACC & FALCON & ELEFANT & YACC & FALCON & ELEFANT & YACC \\
-%  \midrule
-%   $\left\langle T \right\rangle$ & 0.7759 & 0.7758 & 0.7767 & 0.6043 & 0.6036 & 0.6058 & 0.7275 & 0.7275 & 0.7286 & & 0.5277 & 0.5289\\
-%    Nu$^{top}$ & 3.4280 & 3.4214 & 3.4298 & 8.5242 & 8.5115 & 8.5278 & 3.0368 & 3.0347 & 3.0374 && 6.5912 & 6.5572 \\
-%    Nu$^{bottom}$ & 3.3168 & 3.313 & 3.3143 & 8.4599 & 8.4465 & 8.3990 & 3.0086 & 2.9908 & 2.9628 && 6.5834 & 6.5243 \\
-%    $v_{\textrm{rms}}$ & 249.901 & 249.134 & 251.7997 & 141.2451 & 140.7087 & 142.2020 & 100.2776 & 100.1208 & 100.9467 && 79.1105 & 79.6202 \\
-%    $v_{\textrm{rms}}^{top}$ & 1.8907 & 1.8642 & 1.8298 & 104.7928 & 104.1228 & 105.1767 & 2.0832 & 2.0652 & 2.0374 && 74.7596 & 75.8414 \\
-%    $u_{max}^{top}$ & 2.6340 & 2.6119 & 2.5516 & 121.9212& 121.766 & 122.3238 & 2.9086 & 2.9019 & 2.8458 && 88.9146 & 89.2940 \\
-%    $\eta_{min}$ & - & - & - & $2.0303\times10^{-5}$ & $2.0600\times10^{-5}$ & $1.9853\times10^{-5}$ & $4.8129\times10^{-5}$ & $4.8080\times10^{-5}$ & $4.7907\times10^{-5}$ &
-%    & $1.9860\times10^{-4}$ & $1.9174\times10^{-4}$ \\
-%    $\eta_{max}$ & - & - & - & 1.3325 & 1.85 & 1.6095 & 0.9360 & 0.9023 & 1 && 1.52 & 1.6773 \\
-%    $\left\langle W \right\rangle$ & 2.4299 & 2.4316 & 2.4583 & 7.5648 & 7.5665 & 7.6222 & 2.0387 & 2.0384 & 2.0400 && 5.6216 & 5.6512 \\
-%    $\left\langle \Phi \right\rangle/Ra$ & 2.4311 & 2.4276 & 2.4333 & 7.5661 & 7.5585 & 7.6194 & 2.0386 & 2.0370 & 2.0335 && 5.6182 & 5.6463 \\
-%    $\delta$ & 0.05\% & 0.16\% & 1.02\% & 0.02\% & 0.11\% & 0.04\% & $<$0.01\% & 0.07\% & 0.32\% && 0.06\% & 0.09\% \\
-%  \bottomrule
-%  \end{tabular}}
-%  \label{tab:visco_mantle}
-%  \end{table}
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{2}{white!100!}{gray!25!}
-\begin{table}[H]
-  \caption{Benchmark cases and corresponding parameters.}
-  \centering
-  \small
-\begin{tabular}{|>{\centering}m{0.08\textwidth}|>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}
-  >{\centering}m{0.05\textwidth} >{\centering}m{0.05\textwidth} |>{\centering\arraybackslash}m{0.2\textwidth}|}
-  \toprule
-    Case & $Ra$ & $\Delta\eta_T$ & $\Delta\eta_y$ & $\eta^*$ & $\sigma_Y$ & Convective regime \\
-  \midrule
-  \rowcolor{gray!25!}
-    1 & $10^2$ & $10^5$ & 1 & - & - & Stagnant lid \\
-    2 & $10^2$ & $10^5$ & 1 & $10^{-3}$ & 1 & Mobile lid  \\
-    \rowcolor{gray!25!}
-    3 & $10^2$ & $10^5$ & 10 & - & - & Stagnant lid  \\
-    4 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 1 & Mobile lid \\
-    \rowcolor{gray!25!}
-    5 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 4 & Periodic \\
-  \bottomrule
-  \end{tabular}
-  \label{tab:tosi_setup}
-  \end{table}
-
-\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{gray!25!}{white!100!}
-\begin{table}
-  \caption{Results of the viscoplastic mantle convection benchmark for different grid resolutions in Cases 1-5 compared with results obtained with ELEFANT
-  and YACC ($100\times100$ elements, from \citealp{Tosi2015}) and in Stone 28 ($32\times32$ elements, from
-  \url{https://github.com/cedrict/fieldstone/tree/master/python_codes/fieldstone_28}).}
-  \centering
-  \resizebox{\textwidth}{!}{
-\begin{tabular}{|cc|ccccccccccc|}
-  \toprule
-  \multicolumn{13}{|c|}{\textbf{Case 1}} \\
-  \midrule
-  Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-  & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-  \hline
-  \rowcolor{gray!25!}
-    YACC & $100\times100$ & 0.7767 & 3.4298 & 3.3143 & 251.7997 & 1.8298 & 2.5516 & - & - & 2.4583 & 2.4333 & 1.02\% \\
-    ELEFANT & $100\times100$ & 0.7758 & 3.4214 & 3.313 & 249.134 & 1.8642 & 2.6119 & - & - & 2.4316 & 2.4276 & 0.16\% \\
-    \rowcolor{gray!25!}
-    Stone 28 & $32\times32$ & 0.7737 & 3.3987 & - & 243.872 & - & - & - & - & - & - & - \\
-%    FALCON & $32\times32$ & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & - & - & 2.4745 & 2.4864 & 0.48\% \\
-%    \rowcolor{gray!25!}
-    FALCON & $100\times100$ & 0.7759 & 3.4280 & 3.3168 & 249.901 & 1.8907 & 2.6340 & - & - & 2.4299 & 2.4311 & 0.05\% \\
-    \midrule
-    \multicolumn{13}{|c|}{\textbf{Case 2}} \\
-    \midrule
-    Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-    & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-    \hline
-    \rowcolor{gray!25!}
-     YACC & $100\times100$ & 0.6058 & 8.5278 & 8.3990 & 142.2020 & 105.1767 & 122.3238 & $1.9853\times10^{-5}$ & 1.6095 & 7.6222 & 7.6194 & 0.04\% \\
-     ELEFANT & $100\times100$ & 0.6034 & 8.5115 & 8.4465 & 140.8067 & 104.1228 & 121.766 & $2.0600\times10^{-5}$ & 1.85 & 7.5665 & 7.5585 & 0.11\% \\
-     \rowcolor{gray!25!}
-     Stone 28 & $32\times32$ & 0.6052 & 8.1715 & - & 141.518 & - & - & - & - & - & - & - \\
-%     FALCON & $32\times32$ & 0.6454 & 8.1422 & 7.4020 & 156.6908 & 95.7226 & 116.9302 & $2.2572\times10^{-5}$ & 0.7194 & 7.4289 & 7.4452 & 0.22\% \\
-%     \rowcolor{gray!25!}
-     FALCON & $100\times100$ & 0.6044 & 8.5266 & 8.4590 & 141.3684 & 104.8867 & 122.0470 & $2.0333\times10^{-5}$ & 1.2914 & 7.5707 & 7.5718 & 0.01\% \\
-     \midrule
-     \multicolumn{13}{|c|}{\textbf{Case 3}} \\
-     \midrule
-     Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-     & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-     \hline
-     \rowcolor{gray!25!} 
-     YACC & $100\times100$ & 0.7286 & 3.0374 & 2.9628 & 100.9467 & 2.0374 & 2.8458 & $4.7907\times10^{-5}$ & 1 & 2.0400 & 2.0335 & 0.32\% \\
-      ELEFANT & $100\times100$ & 0.7275 & 3.0347 & 2.9908 & 100.1208 & 2.0652 & 2.9019 & $4.8080\times10^{-5}$ & 0.9023 & 2.0384 & 2.037 & 0.07\% \\
-      \rowcolor{gray!25!}
-      Stone 28 & $32\times32$ & 0.7277 & 3.0347 & - & 100.018 & - & - & - & - & - & - & - \\
-%      FALCON & $32\times32$ & 0.7281 & 3.0618 & 2.8132 & 101.0092 & 2.1630 & 3.0178 & $4.8500\times10^{-5}$ & 0.7845 & 2.0699 & 2.0695 & 0.02\% \\
-%      \rowcolor{gray!25!}
-      FALCON & $100\times100$ & 0.7275 & 3.0368 & 3.0086 & 100.2776 & 2.0832 & 2.9086 & $4.8129\times10^{-5}$ & 0.9360 & 2.0387 & 2.0386 & $<$0.01\% \\
-      \midrule
-      \multicolumn{13}{|c|}{\textbf{Case 4}} \\
-      \midrule
-      Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-      & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-      \hline
-      \rowcolor{gray!25!} 
-      YACC & $100\times100$ & 0.5289 & 6.5572 & 6.5243 & 79.6202 & 75.4814 & 89.2940 & $1.9174\times10^{-4}$ & 1.6773 & 5.6512 & 5.6463 & 0.09\% \\
-       ELEFANT & $100\times100$ & 0.5277 & 6.5912 & 6.5834 & 79.1105 & 74.7596 & 88.9146 & $1.9860\times10^{-4}$ & 1.5200 & 5.6216 & 5.6182 & 0.06\% \\
-       \rowcolor{gray!25!}
-       Stone 28 & $32\times32$ & 0.5286 & 6.4036 & - & 79.4746 & - & - & - & - & - & - & - \\
-%       FALCON & $32\times32$ & 0.5334 & 6.4679 & 6.3435 & 80.6153 & 76.0511 & 90.2058 & $2.0393\times10^{-4}$ & 0.6453 & 5.6676 & 5.6698 & 0.04\% \\
-%       \rowcolor{gray!25!}
-       FALCON & $100\times100$ & 0.5279 & 6.5897 & 6.5830 & 79.1220 & 75.0715 & 88.8484 & $1.9693\times10^{-4}$ & 1.4133 & 5.6127 & 5.6113 & 0.03\% \\
-       \midrule
-       \multicolumn{13}{|c|}{\textbf{Case 5}} \\
-       \midrule
-       Code & Resolution & $P$ & $\left\langle T \right\rangle_{min}$ & $\left\langle T \right\rangle_{max}$ & Nu$^{top}_{min}$ & Nu$^{top}_{max}$ &
-       $v_{\textrm{rms}}^{min}$ & $v_{\textrm{rms}}^{max}$ & \multicolumn{2}{c}{$(\left\langle \Phi \right\rangle/Ra)_{min}$} & 
-       \multicolumn{2}{c|}{$(\left\langle \Phi \right\rangle/Ra)_{max}$} \\
-       \hline
-       \rowcolor{gray!25!}
-         YACC & $100\times100$ & 0.0792 & 0.6543 & 0.6722 & 2.6950 & 7.2792 & 41.9813 & 98.8230 & \multicolumn{2}{c}{1.4347} & 
-         \multicolumn{2}{c|}{9.2748} \\
-         ELEFANT & $100\times100$ & 0.076 & 0.6514 & 0.6693 & 2.6808 & 7.3506 & 41.2578 & 100.089 & \multicolumn{2}{c}{1.3324} & 
-         \multicolumn{2}{c|}{9.5289} \\
-         \rowcolor{gray!25!}
-         Stone 28 & $32\times32$ & 0.0764 & 0.6521 & 0.6697 & 2.6912 & 7.0763 & 41.61 & 98.183 & \multicolumn{2}{c}{-} & 
-         \multicolumn{2}{c|}{-} \\
-     %    FALCON & $32\times32$ & & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & \multicolumn{2}{c}{-} & 
-     %     \multicolumn{2}{c|}{-} \\
-     %    \rowcolor{gray!25!}
-         FALCON & $100\times100$ & 0.0775 & 0.6519 & 0.6697 & 2.6847 & 7.3629 & 41.5256 & 99.5466 & \multicolumn{2}{c}{1.3399} & 
-         \multicolumn{2}{c|}{9.3993} \\
-  \bottomrule
-  \end{tabular}}
-  \label{tab:visco_mantle}
-  \end{table}
-
-\subsection{Thin layer entrainment}\label{sec:thin}
-This experiment is performed as originally proposed by \citet{vanKeken1997} with two incompressible fluids in a rectangular domain with $L_x=2$ and $L_y=1$
-and gravity acceleration $g_y=10^{10}Ra$. The Rayleigh number and the compositional Rayleigh number are fixed to $Ra=300000$ and $Rc=450000$, respectively.
-Both fluids have constant viscosity, thermal conductivity, specific heat ($\eta=\rho=C_p=1$) and thermal expansion coefficient ($\alpha=10^{-10}$). Fluid 1
-has a density $\rho_1=1$, while fluid 2 is denser ($\rho_2=\rho_1+1.5 \cdot 10^{-10}$) and is placed at the bottom of the domain, for $y \leq 0.025$.
-Temperature are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides of the domain. The initial
-temperature field is given by
-\[T(x,y)=T_u(x,y)+T_l(x,y)+T_r(x,y)+T_s(x,y)-\frac{3}{2}\]
-with
-\begin{eqnarray}
-T_u(x,y)&=&\frac{1}{2}erf\left(\frac{1-y}{2}\sqrt{\frac{u_0}{x}}\right)\nonumber \\
-T_l(x,y)&=&1-\frac{1}{2}erf\left(\frac{y}{2}\sqrt{\frac{u_0}{L_x-x}}\right)\nonumber \\
-T_r(x,y)&=&\frac{1}{2}+\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{y+1}} \exp\left(-\frac{x^2u_0}{4y+4}\right)\nonumber \\
-T_s(x,y)&=&\frac{1}{2}-\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{2-y}} \exp\left(-\frac{(L_x-x)^2u_0}{8-4y}\right)\nonumber
-\end{eqnarray}
-and
-\begin{eqnarray}
-u_0&=&\frac{L_x^{7/3}}{(1+L_x^4)^{2/3}}\left(\frac{Ra}{2\sqrt{\pi}}\right)^{\frac{2}{3}}\nonumber \\
-Q&=&2\sqrt{\frac{L_x}{\pi u_0}}\nonumber
-\end{eqnarray}
-
-The experiment is performed with two grid resolution (\num{125x40} and \num{200x80} elements), with 100 markers per element and CFL of 0.25. $v_{\textrm{rms}}$
-(see Sec. \ref{sec:error}) calculated for both the simulations match well with results from \citet{vanKeken1997} and obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and ELEFANT \citep{Thieulot2014}, obtained with similar grid resolutions
-(Fig. \ref{fig:thin}). In addition, a variety of simulations with different resolutions and aspect ratios are performed to verify that the sensitivity of the
-initial velocity field. Results match well with results from \citet{vanKeken1997} (in the grey area) and \citet{Thieulot2014} (Fig. \ref{fig:thin_initial}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Thin_layer}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_thin.pdf}
-\caption{$v_{\textrm{rms}}$ for the thin layer experiment as function of time for different grid resolution. Results are compared with results obtained by
-\citet{vanKeken1997} (black lines) and with ASPECT and ELEFANT.}
-\label{fig:thin}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrmszero.pdf}
-\caption{$v_{\textrm{rms}}$ at $t=0$ for the thin layer experiment as function of the element size for different aspect ratios of the elements. Results are compared
-with results obtained by \citet{vanKeken1997} (black dots in the grey area) and with ASPECT and ELEFANT.}
-\label{fig:thin_initial}
-\end{figure}
-
-\subsection{Sinking hydrated cylinder}\label{sec:quinquis}
-This experiment is performed as explained in \citet{Quinquis2014} and simulates the sinking of a cold, hydrated cylinder into a warm, dry mantle. The domain
-is square with $L_x=L_y=$ \SI{300}{\km}, a grid resolution of \num{300x300} elements and 25 markers per element. The sinking cylinder is located at $x=$
-\SI{150}{\km} and $y=$ \SI{170}{\km} with a radius of \SI{20}{\km} and has $\rho_c=$ \SI{3250}{\kg\per\cubic\m}, $\eta_c=$ \SI{e23}{\pascal\s}, a thermal
-conductivity $k_c=$ \SI{4.5}{\watt\per\m\per\kelvin},a specific heat $C_{p_c}=$ \SI{1250}{\joule\per\kg\per\kelvin}; the surrounding mantle has $\rho_m=$
-\SI{3200}{\kg\per\cubic\m}, $\eta_m=$ \SI{e20}{\pascal\s}, a thermal conductivity $k_m=$ \SI{105}{\watt\per\m\per\kelvin} and specific heat $C_{p_m}=$
-\SI{1250}{\joule\per\kg\per\kelvin}; a 58 km-thick lithosphere overlies the mantle and has $\rho_l=$ \SI{3200}{\kg\per\cubic\m}, $\eta_l=$ \SI{e23}{\pascal\s},
-a thermal conductivity $k_l=$ \SI{4.5}{\watt\per\m\per\kelvin}, specific heat $C_{p_l}=$ \SI{750}{\joule\per\kg\per\kelvin} and a thermal diffusivity of
-\SI{e-6}{\square\m\per\s}. Velocity boundary conditions are set to free slip on all sides of the domain. Temperature increase linearly in the lithosphere,
-from 0° to 1300°C, and in the mantle, from 1300° to 1360.5°C. The initial temperature of the cylinder is fixed to 400°C. Throughout the evolution, temperature
-is fixed to 0° and 1360.5°C at the top and bottom of the domain, respectively. The initial bound water content is imposed at 2 wt.\% in the cylinder and at
-0 wt.\% in both the mantle and the lithosphere. The maximum amount of water in the mantle is fixed to 0.2 wt.\%, while maximum water content in the cylinder
-and in the lithosphere is function of pressure and temperature and is calculated using a serpentinized harzburgite with Perple\_X, as in \citet{Quinquis2014}.
-
-As shown in \citet{Quinquis2014}, the model is characterised by a progressive dehydration in the external portion of the cylinder, due to the increase of
-temperature, with a consequent vertical migration of free water that hydrates the mantle above the cylinder, up to the lithosphere. The distribution of
-bound and free water during the evolution of the experiment is shown in Fig. \ref{fig:quinquis} (left and right columns, respectively). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Hydration}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Hydration.pdf}
-\caption{Distribution of bound and free water (left and right columns, respectively) at $t=0.5$ and $t=$ \SI{1.25}{\mega\year} (first and second rows,
-respectively) for the sinking hydrated cylinder experiment. Only markers with bound or free water not equal to 0 are plotted.}
-\label{fig:quinquis}
-\end{figure}
-
-\subsection{Mantle melting curves}\label{sec:katz}
-This instantaneous experiment is performed in order to recreate isobaric melting curves, as function of temperature, bulk water and content of water in the
-melt, to compare with melting curves obtained by \citet{Katz2003}. The domain is square with $L_x=L_y=100$ km, a grid resolution of \num{100x100} elements and
-25 markers per element. Density is set in the entire domain to \SI{3300}{\kg\per\cubic\m} and temperature increases from the sides to the centre from 1000° to
-1500°C. The experiment is performed for bulk water contents from 0 to 0.3 wt.\%.
-
-Fig. \ref{fig:fraction} shows isobaric melting curves for \SI{1}{\giga\pascal} (panel a) and \SI{3}{\giga\pascal} (panel b) calculated for different bulk water
-contents. The curves well match with those obtained for the mantle by \citet{Katz2003}, showing the saturation at the solidus for bulk water of 0.2 and 0.3 wt.\%, indicating
-that the code correctly determines melt fraction as function of pressure, temperature and bulk water content. Same results can be observed in Fig.
-\ref{fig:water}a, where isobaric and isothermal melting curves are shown as function of bulk water content. In addition, Fig. \ref{fig:water}b shows that the
-code correctly determines also the percentage of water in the melt as function of pressure, temperature and melt fraction. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Melting}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Fraction.pdf}
-\caption{Isobaric melting curves for 1 GPa (panel a) and 3 GPa (panel b) as function of temperature with different bulk water contents.}
-\label{fig:fraction}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Water.pdf}
-\caption{Isobaric melting curves for 1.5 GPa as function of bulk water (panel a) and water in the melt (panel b) calculated for different temperatures.}
-\label{fig:water}
-\end{figure}
-
-\clearpage
-
-\bibliography{export}
-
-\end{document}

### 2022-09-07 tex/Benchmark/quinquis.tex
update main

--- 2022-03-31 FALCON_description.tex
@@
-\documentclass[hidelinks,10pt,a4paper]{article}
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{textcomp}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage{booktabs}
-\usepackage{multirow}
-\usepackage[cm]{fullpage}
-\usepackage[round]{natbib}
-\bibliographystyle{abbrvnat}
-\usepackage{doi}
-\usepackage{bm}
-\usepackage[table]{xcolor}
-\hypersetup{
-    colorlinks,
-    linkcolor={blue},
-    citecolor={blue},
-    urlcolor={blue}
-}
-\usepackage{float}
-\restylefloat{table}
-\usepackage{siunitx}
-\DeclareSIUnit\year{yr}
-
-\begin{document}
-
-\section{Introduction}\label{sec:intro}
-This document includes a description of the code FALCON, with a presentation of the features implemented followed by results of extensive benchmarks listed in
-Table \ref{tab:list}, performed to verify that the code solves correctly a wide range of problems. Description and results of all benchmarks are shown
-in Section \ref{sec:benchmark}.
-
-\begin{table}[H]
-\caption{List of benchmarks performed for each feature of the code. Data of all benchmarks can be found at 
-\url{https://github.com/aleregorda/Benchmarks}}
-\centering
-\begin{tabular}{| l | l | l |}
-\toprule
- \textbf{Feature}  & \textbf{Benchmark} & \textbf{Section} \\
-\midrule
-  Solver  & Stokes flow & \ref{sec:stokes} \\
-\hline
-  \multirow{2}{*}{Markers advection} & Zalesak disk & \ref{sec:runge} \\
-    & Conservative Interpolation Velocity & \ref{sec:CVI} \\
-\hline
-  \multirow{4}{*}{Momentum equation}  & Poiseuille flow & \ref{sec:poiseuille}\\
-    & Instantaneous 2D sphere & \ref{sec:ist_sphere} \\
-    & Rayleigh-Taylor experiment & \ref{sec:rayleigh} \\
-    & Falling block & \ref{sec:block} \\
-\hline
-    & 2D Stokes sphere & \ref{sec:sphere} \\
-    Sticky air and  & Stabilisation algorithm & \ref{sec:stab} \\
-    free surface & Topography relaxation & \ref{sec:crameri} \\
-  & Spontaneous subduction & \ref{sec:subduction} \\
-\hline
-  Erosion and sedimentation  & - & - \\
-\hline
-   Non-linear  & Slab detachment & \ref{sec:slab} \\
-   visco-plasticity & Indenter & \ref{sec:indenter} \\
-    & Brick & \ref{sec:brick} \\
-\hline
-  \multirow{3}{*}{Energy equation}  & Advection stabilisation & \ref{sec:advection} \\
-    & Simple shear heating & \ref{sec:simple_shear} \\
-    & Shear and adiabatic heating & \ref{sec:shear} \\
-\hline
-\multirow{3}{*}{Energy + momentum} & Mantle convection & \ref{sec:mantle} \\
-& Viscoplastic mantle convection & \ref{sec:visco_mantle} \\
-& Thin layer entrainment & \ref{sec:thin} \\
-\hline
-  Phase changes  & \multirow{2}{*}{Hydrated sinking cylinder} & \multirow{2}{*}{\ref{sec:quinquis}} \\
-  and hydration  &  & \\
-\hline
-  Melting & Experimental melting curves & \ref{sec:katz} \\
-\bottomrule
-\end{tabular}
-\label{tab:list}
-\end{table}
-
-The code is written in Fortran90 and uses the finite element method (FEM) with quadrilateral $Q_1 \times P_0$ elements (continuous bilinear velocity and
-discontinuous constant pressure), associated with MUMPS\footnote{\url{http://mumps.enseeiht.fr/}} \citep{Amestoy2001,Amestoy2019} that is a software package
-for solving systems of linear equations of the form $A \cdot x = b$, where \textit{A} is a square sparse matrix, by means of a direct method. Since
-$Q_1 \times P_0$ elements do not satisfy Ladyzhenskaya, Babuska and Brezzi (LBB) stability condition \citep{Donea2003}, elemental pressure is elaborated
-in post-processing to avoid spurious pressures by means of a double interpolation to smooth the pressure. Following this procedure, the elemental pressure
-is interpolated onto nodes and then back onto elements. After the smoothing procedure, the pressure field is calculated again on the nodes to be used in
-combination with the lithostatic pressure, which is calculated onto nodes. Correctness of the solution and performances of the solver in terms of time and
-memory usage are tested solving the Stokes flow with the analytical solution proposed by \citet{Donea2003} (Section \ref{sec:stokes}).
-
-The thermo-mechanics of crust-mantle systems is described by means of the conservation of mass, momentum and energy equations, expressed as follows:
-\begin{eqnarray}
-\label{eq:mass}\frac{\partial \rho}{\partial t}+\bm{\nabla} \cdot (\rho \bm{u})&=&0\\
-\label{eq:momentum}-\bm{\nabla} p + \bm{\nabla} \cdot \bm{\tau} + \rho \bm{g}&=&0\\
-\label{eq:energy}\rho C_p \left(\frac{\partial T}{\partial t} + \bm{u} \cdot \bm{\nabla} T \right)&=&\bm{\nabla} \cdot \left(k\bm{\nabla} T\right) + H_{tot}
-\end{eqnarray}
-where $\rho$ is the density, $\bm{u}$ is the velocity, \textit{p} is the pressure, $\bm{\tau}$ is the deviatoric stress, $\bm{g}$ is the gravity acceleration,
-$C_p$ is the specific heat at constant pressure, \textit{T} is the temperature, \textit{k} is the thermal conductivity and $H_{tot}$ is the total internal heat
-production. Density variations due to temperature are generally small enough to assume the density as constant ($\rho=\rho_0$) in Eq. \ref{eq:mass} and in
-Eq. \ref{eq:energy}, while it must be treated as a variable in the buoyancy term of Eq. \ref{eq:momentum}, such that \[\rho=\rho_0(1-\alpha(T-T_0))\]
-where $\rho_0$ is the density at a reference temperature $T_0$ and $\alpha$ is the coefficient of thermal expansion. Eq. \ref{eq:mass} then can be rewritten as
-\begin{equation}\label{eq:mass_0}
-\bm{\nabla} \cdot \bm{u}=0
-\end{equation}
-This simplification is known as the extended Boussinesq approximation.
-
-The time step $dt$ is chosen in according to Courant-Friedrichs-Lewy condition \citep{Anderson1995}, such that
-\[dt=C_n\; \min\left(\frac{\Delta x}{\max|\bm{u}|},\frac{(\Delta x)^2}{\kappa}\right)\]
-where $0<C_n<1$ is the Courant (CFL) number, $\Delta x$ is the minimum dimension of smallest element, $\max|\bm{u}|$ is the maximum velocity calculated on the
-entire domain and
-$\kappa=\frac{k}{\rho C_p}$ is the thermal diffusivity \citep{Thieulot2014}.
-
-\section{Lagrangian markers}\label{sec:markers}
-Elemental properties (density, viscosity, thermal conductivity, specific heat and thermal expansion) needed to solve Eqs. \ref{eq:mass_0}, \ref{eq:momentum}
-and \ref{eq:energy} are related to the composition of each element, determined by means of Lagrangian markers that are characteristic of different materials
-in the domain. Elemental properties (with the sole exception of the viscosity) are calculated using an arithmetic mean, as \[P_e=\frac{1}{n}\sum_{i=1}^n P_i\]
-where $P_e$ is the elemental property, $P_i$ is the property characteristic of the material of each marker and $n$ is the number of markers of the element.
-Differently, the average scheme for the elemental viscosity can be chosen between harmonic, geometric and arithmetic mean. Markers can be placed regularly or
-randomly at the beginning of the simulation and their advection is performed by either a 2nd-order or a 4th-order Runge-Kutta in space, interpolating the velocity 
-field on each marker by means of the shape functions (see Section \ref{sec:runge}). The interpolated velocity is then corrected by means of the Conservative
-Velocity Interpolation (CVI), which introduces a corrective term to reduce dispersion and clustering of particles in both steady state and time-dependent 
-\citep{Wang2015} (see Section \ref{sec:CVI}).
-
-The initial distribution of the markers is created using the open-source code library
-Geodynamic World Builder \citep{Fraters2019}. During the simulation, each marker carries memory of temperature, pressure and accumulated strain, which are
-determined interpolating the relative nodal parameter. In particular, pressure is interpolated onto each marker after the smoothing procedure. The number
-of markers contained in each element is maintained between $n_{min}$ and $n_{max}$. When in an element there are less markers than $n_{min}$ the code adds
-random markers to reach the $n_{min}$, while if the number is higher than $n_{max}$ some of them are randomly deleted. When new markers are added, they assume
-the properties of the nearest marker. In this way elements are never empty and maintain a number of markers inside a prefixed range.
-
-\section{The momentum equation}\label{sec:mom_eq}
-The penalty method is implemented so that Eq. \ref{eq:mass_0} can be written as:
-\begin{equation}\label{eq:mass_penalty}
-\bm{\nabla} \cdot \bm{u} + \frac{p}{\lambda}=0
-\end{equation}
-where $\lambda$ is the so-called penalty parameter, which should be 6-7 orders of magnitude larger than the shear viscosity to ensure that mass conservation
-is satisfied \citep{Donea2003,Thieulot2014}.
-
-The deviatoric stress tensor in Eq. \ref{eq:momentum} can be written in terms of the strain rate tensor as $\bm{\tau}=2\eta\dot{\bm{\epsilon}}$, with
-$\dot{\bm{\epsilon}}=\frac{1}{2}\left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)$. Therefore, Eq. \ref{eq:momentum} can be rewritten as:
-\begin{equation}\label{eq:momentum_pressure}
--\bm{\nabla} p + \bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right) + \rho \bm{g} = 0
-\end{equation}
-Finally, using pressure from Eq. \ref{eq:mass_penalty}, Eq. \ref{eq:momentum_pressure} can be rewritten as:
-\begin{equation}\label{eq:momentum_penalty}
-\lambda \bm{\nabla} \left(\bm{\nabla} \cdot \bm{v}\right)+\bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right)+\rho\bm{g}=0
-\end{equation}
-The penalty method is been associated to the iterative Uzawa method, as described in detail in \citet{Dabrowski2008} and \citet{Thieulot2014}
-(see Section \ref{sec:poiseuille}). Results of benchmarks performed to verify the correctness in the implementation of Eq. \ref{eq:momentum_penalty} are
-shown in Sections \ref{sec:ist_sphere} and \ref{sec:rayleigh}.
-
-To support large viscosities variations the penalty parameter is related to the effective elemental viscosity by means of a dimensionless coefficient, so that
-$\lambda=\lambda_e(e)\eta_{eff}(e)$ \citep{Marotta2006,Dabrowski2008,Thieulot2014}. Benchmark of the falling block \citep{Gerya2003a,Gerya2010b,Thieulot2011}
-is performed to verify that the code can correctly deal buoyancy driven flows with strong viscosity contrasts (Section \ref{sec:block}).
-
-\section{Sticky air and free surface}\label{sec:sticky}
-The Earth's surface can be treated by means of either the so-called sticky-air or a true free surface method, both of them implemented in the code. In the
-sticky-air method the surface is approximated with the introduction of a buoyant layer with a viscosity at least four orders of magnitude lower than the crust
-\citep{Schmeling2008,Crameri2012} and the interface between lithosphere and air is defined using a chain of passive markers that are advected as the Lagrangian
-markers. The correctness of the evolution of the markers chain is tested with the experiment of a 2D time-dependent Stokes sphere below a free surface and
-compared with results from ASPECT\footnote{\url{https://aspect.geodynamics.org/}} \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} (Section
-\ref{sec:sphere}). In the true free surface case the top boundary is assumed stress-free and velocities are not fixed. In this case topography variations are
-described by vertical deformations of the mesh that depend on the velocity field of the nodes that identify the free surface, while horizontal deformations are
-not taken into account. This procedure is known as the Arbitrary Lagrangian-Eulerian (ALE) method and its implementation follows the technique described in
-\citet{Thieulot2011}. However, although the implementation of a true free surface better reproduces laboratory experiment, extremely small time steps can be
-necessary to maintain stability \citep{Kaus2010a,Quinquis2011,Thieulot2014}. Therefore, the stability algorithm proposed by \citet{Kaus2010a} is implemented
-to avoid instabilities due to high density differences at the free surface when using too large time steps. The implementation of this algorithm are tested by
-performing the experiment described by \citet{Kaus2010a} (Section \ref{sec:stab}).
-
-The topography relaxation benchmark proposed by \citet{Crameri2012} is performed to verify that the code correctly recovers topography variations in case of
-both the sticky-air and the true free surface method (Section \ref{sec:crameri}). Finally, Section \ref{sec:subduction} show the results of the spontaneous
-subduction experiment described by \citet{Schmeling2008}.
-
-\section{Erosion and sedimentation}\label{sec:erosion}
-Surface processes at either the lithosphere-air interface or in correspondence of the free surface have been implemented by means of the software
-Fastscape\footnote{\url{https://fastscape.org/}} \citep{Braun2013,Cordonnier2019,Yuan2019a,Yuan2019b}. FastScapeLib is a set of routines that solve the stream
-power law (enriched by a sediment transport/deposition term) \citep{Yuan2019a}, hillslope diffusion and marine transport and deposition \citep{Yuan2019b}.
-The partial differential equation solved by FastScapeLib is:
-\begin{equation}
-\frac{\partial h}{ \partial t}=U-K_f A^{m_e}S^{n_e} + \frac{G}{A}\int_A \left( U-\frac{\partial h}{\partial t} \right) dA + K_d \bm{\nabla}^2 h 
-\end{equation}
-where $h$ is the topography, $U$ is the uplift, $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $S$ is the slope, $m_e$ and $n_e$ are
-parameters that should be chosen to constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}, $G$ is the dimensionless
-deposition/transport coefficient for the enriched stream power law and $K_d$ is the effective diffusivity. Terms $K_f A^{m_e}S^{n_e}$ and $K_d \bm{\nabla}^2 h$
-consider long-range processes, described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, and short-range processes,
-proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}.
-
-In case of sticky air, slope is determined by means of the markers chain, which is vertically corrected considering erosion and sedimentation rates. After the
-correction, continental markers above the markers chain are transformed in air markers and air markers below the markers chain become sediments. Similarly,
-slope is determined considering the nodes on the top boundary in case of true free surface and they are vertically corrected as for the markers chain. In this
-case, continental markers above the top boundary are simply deleted.
-%Surface processes at the lithosphere-air interface are implemented taking into account fluvial erosion (long-range processes), hillslope processes (short-range
-%processes) and failure of slopes exceeding a threshold angle $\theta$ \citep{Burov1997,Tucker1999,Sacek2012,Croissant2014,Ueda2015,Beucher2020}. Erosion rates
-%$\dot{E}_{lr}$ due to long-range processes can be described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{lr}=K_f A^{m_e} \bm{\nabla} h^{n_e}
-%\end{equation}
-%where $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $\bm{\nabla} h$ is the slope and $m_e$ and $n_e$ are parameters chosen in to
-%constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}. Erosion rates $\dot{E}_{sr}$ due to short-range processes are
-%implemented by assuming that transport is linearly proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{sr}=K_d \bm{\nabla}^2 h
-%\end{equation}
-%where $K_d$ is the effective diffusivity. In addition, the slope is modified if it still exceeds a threshold angle after long- and short-range processes
-%\citep{Ueda2015}.
-
-%Offshore sedimentation processes are implemented similarly to hillslope processes but with a depth-dependent coefficient \citep{Sacek2012,Beucher2020}.
-%The sedimentation rate $\dot{S}$ is calculated as
-%\begin{equation}
-%\dot{S}=K_s \bm{\nabla}^2 h
-%\end{equation}
-%with
-%\begin{equation}
-%K_s=K^*_s \exp\left(\frac{h-h_{sl}}{h_0}\right)
-%\end{equation}
-%where $K^*_s$ is a constant, $h_{sl}$ is the sea level and $h_0$ is a characteristic length. In case of sticky air, slope is determined by means of the markers
-%chain, which is vertically corrected considering erosion and sedimentation rates. After the correction, continental markers above the markers chain are
-%transformed in air markers and air markers below the markers chain become sediments. Similarly, slope is determined considering the nodes on the top boundary
-%in case of true free surface and they are vertically corrected as for the markers chain. In this case, continental markers above the top boundary are simply
-%deleted.
-
-\section{Non-linear rheologies}\label{sec:nonlinear}
-Non-linear rheologies are implemented combining viscous creep (dislocation and diffusion) and plastic yielding. For each marker, diffusion and dislocation
-viscosity ($\eta_{df|ds}$) can be determined as follows \citep{Karato1993,Warren2008b,Wang2016}:
-\begin{equation}\label{eq:viscous}
-\eta_{df|ds}=f_s \left(\frac{d^m}{A}\right)^{\frac{1}{n}} \left (\frac{2^{\frac{1-n}{n}}}{3^{\frac{n+1}{2n}}} \right)
-I_2^{\frac{(1-n)}{n}}\exp\left(\frac{Q+pV}{nRT}\right)
-\end{equation}
-where $f_s$ is a scaling factor used to represent lithologies that are stronger or weaker than the base set, $d$ is the grain size, $m$ is the grain size exponent,
-$n$ is the stress exponent, $A$ is the uniaxial pre-exponential factor, $I_2$ is the square root of the second invariant of the strain rate tensor, $Q$ is the
-activation energy, $p$ is the pressure, $V$ is the activation volume, $R$ is the gas constant and $T$ is the temperature.
-Pressures and temperatures are determined by the interpolation of the nodal parameters, while strain rates are calculated by means of the
-derivative of the velocities on the nodes. In case of diffusion creep $n=1$ and $m>0$, in case of dislocation creep $n>1$ and $m=0$. The viscous creep $\eta_{cp}$
-is then calculated as the harmonic average between $\eta_{df}$ and $\eta_{ds}$:
-\begin{equation}\label{eq:visco_creep}
-\eta_{cp}=\left(\frac{1}{\eta_{df}}+\frac{1}{\eta_{ds}}\right)^{-1}
-\end{equation}
-The implementation of the non-linear viscous creep viscosity depending on the strain rate is tested by means of the slab detachment benchmark
-(Section \ref{sec:slab}).
-
-Plastic yielding is implemented rescaling $\eta_{cp}$ in order to limit the stress below the yield stress $\sigma_y$ \citep{Thieulot2008,Thieulot2014,Glerum2018},
-obtaining
-\begin{equation}\label{eq:plastic}
-\eta_{pl}=\frac{\sigma_y}{2I_2}
-\end{equation}
-where $\eta_{pl}$ is the plastic viscosity and the yield stress is determined following the Drucker-Prager criterion, such as
-\begin{equation}\label{eq:yield}
-\sigma_y=C\;\cos(\phi)+p\;\sin(\phi)
-\end{equation}
-where $C$ is the cohesion and $\phi$ is the internal friction angle. In case of negative pressure it is imposed equal to 0, so that negative yield stress are
-excluded. The correctness of the non-linear solution in case of plastic viscosity with variable internal friction angle is verified performing indenter and
-brick experiments (Sections \ref{sec:indenter} and \ref{sec:brick}, respectively).
-
-Strain softening is taken into account for both viscous creep and plastic viscosity \citep{Huismans2003,Babeyko2005,Huismans2005,Sobolev2005,Warren2008a} by
-means of the accumulated strain $\epsilon$ memorised by each marker. Softening in the viscous creep determines a linear decrease of $\eta_{vc}$ by means of a
-viscous strain softening factor $W_S$ that increases linearly from $W_{S_0}$ to $W_{S_{\infty}}$ for $\epsilon_{S_0}<\epsilon_S<\epsilon_{S_{\infty}}$ 
-\citep{Warren2008a}. This viscous softening can be related to strain-induced grain size reduction \citep{Warren2008a}. Differently, plastic softening is 
-simulated with a linear decrease of internal friction angle $\phi(\epsilon)$ and cohesion $C(\epsilon)$ in according to
-\begin{eqnarray}
-\label{eq:friction1}\phi(\epsilon)&=&\phi_0+(\phi_{\infty}-\phi_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}\\
-\label{eq:friction2}C(\epsilon)&=&C_0+(C_{\infty}-C_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}
-\end{eqnarray}
-where $\phi_0$, $C_0$ and $\phi_{\infty}$, $C_{\infty}$ are internal friction angle and cohesion for $\epsilon_0$ and $\epsilon_{\infty}$, respectively
-\citep{Huismans2003,Huismans2005,Warren2008a,Thieulot2014}. Plastic softening approximates deformation-induced softening of faults and brittle shear zones
-\citep{Warren2008a}.
-
-Plastic yielding and viscous creep are then combined to obtain a viscoplastic viscosity $\eta_{vp}$ as follows, assuming that they are independent processes
-\citep{Karato2008,Glerum2018}:
-\begin{equation}\label{eq:viscoplastic}
-\eta_{vp}=\min(\eta_{cp},\eta_{pl})
-\end{equation}
-Finally, effective viscosity $\eta_{eff}$ is capped by the minimum and the maximum viscosity ($\eta_{min}$ and $\eta_{max}$, respectively) to avoid extremely
-low or high viscosity \citep{Glerum2018} as follows
-\begin{equation}\label{eq:effective}
-\eta_{eff}=\min(\max(\eta_{vp},\eta_{min}),\eta_{max})
-\end{equation}
-Viscous creep and plastic yielding are non-linear rheologies because of their dependence on the velocity field through pressure and strain rates. Therefore,
-the solution is determined by means of Picard-type iterations, until convergence of the velocity field \citep{Glerum2018}. The convergence is verified at each
-iteration $i$ via the nonlinear residual $\bm{\mathcal{R}}^i$ that can be determined as
-\begin{equation}\label{eq:convergence}
-\bm{\mathcal{R}}^i= \mathbb{K} (\eta_{eff}(\dot{\bm{\epsilon}}^{i-1},p^{i-1}))\cdot \bm{v}^{i-1} - \bm{f}^i
-\end{equation}
-%\begin{equation}\label{eq:convergence}
-%\chi=1-\frac{\langle (\bm{v}^i-\langle\bm{v}^i\rangle)\cdot (\bm{v}^{i+1}-\langle\bm{v}^{i+1}\rangle) \rangle}{\sqrt{\sigma^i\sigma^{i+1}}}
-%\end{equation}
-%Iterations are performed until either convergence ($\chi<tol$) or maximal number of non-linear iterations ($it_{max}$) is reached.
-where $\mathbb{K}$ is the Stiffness matrix and $\bm{f}$ is the right hand side vector \citep{Spiegelman2016,Glerum2018}.  The $L_2$-norm is extracted from
-$\bm{\mathcal{R}}^i$ and it is normalised as follows
-\begin{equation}\label{eq:convergence_res}
-\frac{||\bm{\mathcal{R}}^i||_2}{||\bm{\mathcal{R}}^0||_2}
-\end{equation}
-where $||\bm{\mathcal{R}}^0||_2$ is the $L_2$-norm of the first nonlinear iteration.
-Since often the normalised nonlinear residual drop very quickly over the first 2-3 iterations, also the $L_2$-norm of velocity and pressure residuals are
-calculated as
-\begin{equation}\label{eq:convergence_vel}
-\frac{||\Delta\bm{v}||_2}{||\bm{v}||_2}=\frac{||\bm{v}_i-\bm{v}_{i-1}||_2}{||\bm{v}_{i}||_2}
-\end{equation}
-and
-\begin{equation}\label{eq:convergence_press}
-  \frac{||\Delta p||_2}{||p||_2}=\frac{||p_i-p_{i-1}||_2}{||p_{i}||_2}
-\end{equation}
-respectively. Iterations are performed until either maximal number of non-linear iterations ($it_{max}$) is reached or all the normalised nonlinear residual, 
-the velocity residual and the pressure residual converge under a defined tolerance $tol$.
-
-\section{The energy equation}\label{sec:en_eq}
-The stabilisation of the advection term of the energy equation (Eq. \ref{eq:energy}) needed to avoid possible oscillations in the thermal solution in the case
-when advection dominates over diffusion, is implemented by means of a streamline-upwind Petrov–Galerkin (SUPG) method, for which the advection term is modified
-follows the discussion in \citet{Thieulot2011} and \citet{Thieulot2014} (see Section \ref{sec:advection} for benchmark). The total internal heat production
-$H_{tot}$ in Eq. \ref{eq:energy} includes radiogenic heating $H_r$, shear heating $H_s$ and adiabatic heating $H_a$, where
-\begin{equation}\label{eq:shear}
-H_s=2\eta \dot{\bm{\epsilon}} : \dot{\bm{\epsilon}} = 2\eta(\dot{\epsilon}_{xx}^2+\dot{\epsilon}_{yy}^2+2\dot{\epsilon}_{xy}^2)
-\end{equation}
-and
-\begin{equation}\label{eq:adiabatic}
-H_a=T \alpha \frac{Dp}{Dt} \approx -\alpha T \rho g_y v
-\end{equation}
-The correctness of the implementation and the computation of shear and adiabatic heating is verified performing an exercise with an analytical solution
-(Section \ref{sec:simple_shear}) and Exercise 9.4 in \citet{Gerya2010b} (Section \ref{sec:shear}). Furthermore, mantle convection and thin layer entrainment
-benchmarks (Sections \ref{sec:mantle} and \ref{sec:thin}, respectively) are performed to verify that the code solve correctly Eqs. \ref{eq:momentum_penalty}
-and \ref{eq:energy} in case of a temperature field characterised by an initial perturbation and with a temperature-dependent density. At last, the viscoplastic
-mantle convection benchmark is performed to check also the viscoplasticity in case of a initial perturbed temperature field (Section \ref{sec:visco_mantle}).
-
-\section{Phase transitions and hydration}\label{sec:phase}
-Variations of effective density, specific heat and coefficient of thermal expansion during the evolution of crustal and mantle materials at different
-pressure-temperature ($p$-$T$) conditions are computed by means of version 6.8.6 of Perple\_X software package \citep{Connolly2005}, similarly to the
-implementation described in \citet{Marotta2020}. By default, mineral assemblages and properties of each lithology (oceanic crust, lower and upper continental
-crust, sediments and dry or hydrated mantle) are calculated for temperatures from \SIrange{330}{1600}{\kelvin}, with increments of approximately \SI{4}{\kelvin},
-and pressures from \SIrange{0.1}{30}{\giga\pascal}, with increments of approximately \SI{0.03}{\giga\pascal}, for a total of almost 100000 points.
-
-The introduction of phase transitions produces two main effects that must be taken into account in numerical models: variations in density and release/absorption
-of latent heat \citep{Tackley2010}. Effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into
-account considering effective coefficient of thermal expansion \citep{Christensen1985,Zhong2015}. Similarly, effects of latent heat in the energy equation
-(eq \ref{eq:energy}) are taken into account by considering effective specific heat and coefficient of thermal expansion
-\citep{Christensen1985,Tackley2010,Zhong2015}. Effective density, specific heat and coefficient of thermal expansion are included in Perple\_X files and are
-assumed by each Lagrangian marker in according to its $p$-$T$ conditions.
-
-Hydration processes related to dehydration of subducting lithosphere \citep{Schmidt1998,Liu2007,Faccenda2009,Faccenda2010,Faccenda2014,Rosas2016} strongly
-influence the thermo-mechanics inside the mantle wedge, mainly because of both weakening effects on the mantle rheology and density variations in case of
-mantle serpentinization \citep{Gerya2002,Honda2003,Arcay2005,Roda2010,Regorda2017}. In the case that hydration is switched on, the amount of bound and free
-water is memorised by each marker, following the implementation of \citet{Quinquis2014}; hydration and dehydration processes are related to the amount of bound
-water of each marker and to the maximum amount of water it can transport, i.e. if the amount of bound water exceeds the maximum amount of water, the marker
-dehydrates and releases free water that can hydrates under-saturated markers. Maximum water content of each marker is determined as function of lithology and
-$p$-$T$ conditions and it is calculated using Perple\_X, in the same way than effective density, specific heat and coefficient of thermal expansion. Bound water
-is advected together with the markers, neglecting the effect of bound water diffusion, while free water simply migrates vertically and is not coupled to the
-solid-phase flow of the mantle wedge \citep{Arcay2005,Quinquis2014}. The correctness of the migration scheme of the free water and its absorption into bound
-water by under-saturated markers are verified performing the experiment described in \citet{Quinquis2014} (Sec. \ref{sec:quinquis}).
-
-The mantle viscosity weakening related to the amount of water has been extensively studied \citep{Chopra1981,Hirth2003} and it is implemented in according to
-\citet{Arcay2005} and \citet{Horiuchi2016}, as follows
-\begin{equation}\label{eq:wet_visc}
-\eta_{wet}=\eta_{dry}\left(\left[1-\frac{1}{f_v}\right]\exp\left(-\frac{[OH^-]}{[OH^-]_0}\right)+\frac{1}{f_v}\right)
-\end{equation}
-where $f_v$ is the viscosity reduction factor between dry and wet conditions, $[OH^-]$ is the water content and $[OH^-]_0$ is a reference water content set to
-620 ppm (0.062 wt.\%) \citep{Arcay2005}. By default, $f_v$ is set to 100, in according to \citet{Arcay2005} and \citet{Horiuchi2016}.
-
-\section{Mantle melting}\label{sec:melt}
-Melt fraction in the mantle depends on temperature ($T$, in \textdegree C), pressure ($p$, in \si{\giga\pascal}) and the water content in the melt ($X_{H_2O}$, in wt.\%).
-The determination of the melt fraction $M$ is implemented as explained by \citet{Katz2003}, according to the successive modification by \citet{Langmuir2006} and
-\citet{Kelley2010}, as follows
-\begin{eqnarray}\label{eq:melt_fr}
-M(p,T,X_{H_2O})&=&-T+T_{sol}+(x_t\cdot ln(p+y_t))\cdot X_{H_2O}-\nonumber\\
-&&+K_m\left(\frac{X_{H_2O}^{bulk}}{D_{H_2O}(1-X_{H_2O})+X_{H_2O}}\right)^{\gamma}
-\end{eqnarray}
-where $T_{sol}$ is the temperature of the dry solidus, $D_{H_2O}$ is the partition coefficient and $x_t\cdot log(p+y_t)$ indicates the pressure dependence of
-the melting curve \citep{Kelley2010}. $x_t$, $y_t$, $K_m$ and $\gamma$ are parameters chosen in according to \citet{Kelley2010}.
-However, $X_{H_2O}$ depends on the melt fraction as
-\begin{equation}\label{eq:melt_wt}
-X_{H_2O}(M)=\frac{X_{H_2O}^{bulk}}{D_{H_2O}+M(1-D_{H_2O})}
-\end{equation}
-and a numerical solution can be found using a root-finder method \citep{Katz2003,Wang2016}. Furthermore, the content of water in the melt is limited by the
-pressure-dependent saturation concentration of water in the melt, determined as
-\begin{equation}\label{eq:melt_max}
-X_{H_2O}^{sat}=\chi_1 p^{\lambda_m}+\chi_2 p
-\end{equation}
-where $\chi_1$, $\chi_2$ and $\lambda_m$ are parameters chosen according to \citet{Katz2003}, \citet{Langmuir2006} and \citet{Kelley2010}. The correctness in
-the determination of melt fractions and water contents in the melt are tested comparing model results with experimental curves obtained by \citet{Katz2003}
-(Sec. \ref{sec:katz}).
-
-In case of melting, viscous creep viscosity ($\eta_{vp}$) is modified according to \citet{Wang2016}, as
-\begin{equation}\label{eq:melt_visc}
-\eta_{melt}=\eta_{vp}\cdot \exp(\alpha_m M)
-\end{equation}
-where $\alpha_m$ are melt fraction factors for dislocation and diffusion creep, chosen in according to \citet{Wang2016}.
-
-Effective density ($\rho_{eff}$) for partially molten rocks is calculated as
-\begin{equation}\label{eq:melt_dens}
-\rho_{eff}=\rho_s(1-M)+\rho_m M
-\end{equation}
-where $\rho_s$ and $\rho_m$ are the densities of the solid and the molten rock, respectively \citep{Gerya2004b,Gerya2007,Wang2016}. The density of the
-solid rock is extracted by Perple\_X, while the density of the molten portion is calculated as
-\begin{equation}\label{eq:melt_dens_M}
-\rho_m(p,T)=\rho_0[1-\alpha(T-T_0)][1+\beta(p-p_0)]
-\end{equation}
-where $\rho_0$ is the density at the reference temperature ($T_0$) and pressure($p_0$), and $\alpha$ and $\beta$ are the thermal expansion and compressibility
-coefficients, respectively \citep{Gerya2004b,Gerya2007,Wang2016}.
-
-As for the phase transitions, effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into account
-considering effective coefficient of thermal expansion. Similarly, latent heat due to melting/crystallisation can be implicitly included in the energy equation
-(Eq. \ref{eq:energy}) by considering effective specific heat ($C_{p_{eff}}$) and thermal expansion ($\alpha_{eff}$) of the partially molten rocks, in according to
-\begin{eqnarray}
-\label{eq:cp_effective}C_{p_{eff}}&=&C_p+H_L\left(\frac{\partial M}{\partial T}\right)_p \\
-\label{eq:alpha_effective}\alpha_{eff}&=&\alpha+\rho\frac{H_L}{T}\left(\frac{\partial M}{\partial p}\right)_T
-\end{eqnarray}
-where $C_p$ and $\alpha$ are the specific heat and the coefficient of thermal expansion of the solid rock, respectively, and $H_L$ is the latent heat of the
-molten rock \citep{Gerya2004b,Gerya2007,Tackley2010}.
-
-\section{Benchmarks}\label{sec:benchmark}
-\subsection{Error measurements}\label{sec:error}
-In order to determine the accuracy of velocity and pressure field of the benchmarks, the $L_2$-norm is computed by numerical integration on the Gauss-Legendre
-quadrature points. $L_2$-norm for pressure and velocity errors can be evaluated as
-\begin{eqnarray}
-\textrm{err}_p&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|p^n_i(r_q)-p^a_i(r_q)|^2\right)w_q|J_q|}\nonumber \\
-\textrm{err}_v&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|u^n_i(r_q)-u^a_i(r_q)|^2+|v^n_i(r_q)-v^a_i(r_q)|^2\right)w_q|J_q|}\nonumber
-\end{eqnarray}
-respectively, where $n_e$ is the number of elements, $n_q$ is the number of quadrature points per element, $p^n_i(r_q)$ and $p^a_i(r_q)$ are the numerical and
-analytical pressure, respectively, in each quadrature point $q$, $w_q$  and $J_q$ are the weight and the Jacobian at the quadrature point
-$q$, $u^n_i(r_q)$, $v^n_i(r_q)$, $u^a_i(r_q)$ and $v^a_i(r_q)$ are the numerical and analytical velocities, respectively, in each quadrature point $q$.
-
-Other quantities used as comparison with original benchmarks are the root-mean-square velocity over the whole domain and over the surface
-\begin{eqnarray}
-v_{\textrm{rms}}&=&\sqrt{\int_{\Omega}|\bm{v}|^2 d\Omega}\nonumber \\
-v_{\textrm{rms}}^{top}&=&\sqrt{\int_0^1 u^2 \bigg|_{y=1} dx}\nonumber
-\end{eqnarray}
-respectively, top and bottom Nusselt numbers
-\[\textrm{Nu}^{top/bottom}=-\int_0^1\frac{\partial T}{\partial y} \bigg|_{y=1/y=0} dx\]
-the average rate of work done against gravity
-\[\left\langle W \right\rangle=\int_{\Omega}T u_y d\Omega\]
-the average rate of viscous dissipation
-\[\left\langle \Phi \right\rangle=\int_{\Omega}\tau_{ij}\dot{\epsilon}_{ij} d\Omega\]
-and the percentage error between $\left\langle W \right\rangle$ and $\left\langle \Phi \right\rangle/Ra$
-\[\delta=\frac{\left|\left\langle W \right\rangle -\frac{\left\langle \Phi \right\rangle}{Ra} \right|}{\max \left(\left\langle W \right\rangle,
-\frac{\left\langle \Phi \right\rangle}{Ra}\right)}\times100\]
-
-\subsection{Stokes flow}\label{sec:stokes}
-The problem consists of determining velocity field $(u,v)$ and pressure $p$ in case of a manufactured solution with prescribed body forces such as
-\begin{eqnarray}
-b_1&=&(12-24y)x^4+(-24+48y)x^3+(-48y+72y^2-48y^3+12)x^2+\nonumber \\
-&&+(-2+24y-72y^2+48y^3)x+1-4y+12y^2-8y^3\nonumber \\
-b_2&=&(8-48y+48y^2)x^3+(-12+72y-72y^2)x^2+(4-24y+48y^2-48y^3+\nonumber \\
-&&+24y^4)x-12y^2+24y^3-12y^4\nonumber
-\end{eqnarray}
-for which the exact solution is:
-\begin{eqnarray}
-u(x,y)&=&x^2(1-x)^2(2y-6y^2+4y^3)\nonumber \\
-v(x,y)&=&-y^2(1-y)^2(2x-6x^2+4x^3)\nonumber \\
-p(x,y)&=&x(1-x)-\frac{1}{6}\nonumber
-\end{eqnarray}
-The domain is a unit square a constant viscosity ($\eta=1$) and the penalty parameter is set to $\lambda=10^7$. Velocity boundary conditions are set to no slip
-$(\bm{v}=\bm{0})$ on all boundaries. The problem is performed for different grid resolution between \num{8x8} and \num{1024x1024} elements.
-
-The errors between the analytical solution and the numerical prediction of the velocity and the pressure field are calculated as in Section \ref{sec:error}.
-Fig. \ref{fig:errors}a shows that both velocity and pressure field converge to the exact solution with the decrease of the element size, following the
-theoretical convergence rate. The convergence of pressure, in contrast with the observation by \citet{Donea2003} for $Q_1 \times P_0$ elements, indicate the
-effectiveness of the smoothing procedure. Smoothed pressure field for grid resolution of \num{128x128} elements is shown in Fig. \ref{fig:errors}b (continuous
-red line), in comparison with the analytical solution (dashed blue line). Solve times and memory usage needed to generate the solution are shown in
-Fig. \ref{fig:MUMPS}. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Solver/Stokes_Flow}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/errors.pdf}
-\caption{Velocity and pressure error for the Stokes flow experiment between generated and analytical solution as a function of element size (panel a),
-and comparison between smoothed pressure and analytical solution as function of $x$ coordinate for a grid resolution of \num{128x128} elements (panel b).}
-\label{fig:errors}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/MUMPS.pdf}
-\caption{Panel a) Analysis, factorisation, solution and total solve times (panel a) and factorisation memory usage (panel b) as a function of the total number
-of degrees of freedom for the Stokes flow experiment.}
-\label{fig:MUMPS}
-\end{figure}
-
-\subsection{Markers advection}\label{sec:runge}
-The 2nd-order Runge-Kutta advection scheme in space is tested by means of the Zalesak disk test \citep{Zalesak1979,Thieulot2014}. The benchmark is performed
-in a unit square domain with a grid resolution of \num{32x32} elements and 1000000 markers and values of Courant number between 0.25 and 1.
-At $t=0$, the disk is centred at position $(0.5;0.75)$ with a radius $R=0.15$ and has a vertical fissure 0.05 wide and 0.2 high (orange in Fig. \ref{fig:runge})
-and with the velocity field prescribed in the entire domain as
-\begin{eqnarray}
-u(x,y)&=&2\pi\left(y-\frac{L_x}{2}\right)\nonumber \\
-v(x,y)&=&-2\pi\left(x-\frac{L_x}{2}\right)\nonumber
-\end{eqnarray}
-the markers are back to their initial location after a $2\pi$ rotation ($t=1$). The clockwise rotation of the disk at $t=0.25$, 0.5 and 0.75 
-is shown in Fig. \ref{fig:runge} (red, blue and green, respectively). The distance from the centre of 1 marker is calculated throughout the rotation to
-evaluate the error for different values of Courant number. As expected, the error increase with the increasing of the time step, as shown in 
-Fig. \ref{fig:runge_err}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/Zalesak_Disk}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=200px]{./Figures/Zalesak.pdf}
-\caption{Position of the Zalesak disk at $t=0$, 0.25, 0.5 and 0.75 (orange, red, blue and green, respectively) for the case with Courant number of 1.}
-\label{fig:runge}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/zal_marker.pdf}
-\caption{Distance from the centre as function of time for values of Courant number of  of 0.25, 0.3, 0.5, 0.75 and 1 (green, orange, blue, black and red,
-respectively). R indicates the distance from the centre considering a perfect circle trajectory.}
-\label{fig:runge_err}
-\end{figure}
-
-\subsection{Conservative Velocity Interpolation (CVI)}\label{sec:CVI}
-The CVI correction is checked by means of the Stokes flow experiment by \citet{Donea2003}, as presented in \ref{sec:stokes}. The advection of Lagrangian
-markers is performed using either a 2nd-order or a 4th-order Runge-Kutta scheme with CFL=0.5 and an initial random distribution of 25 markers per element.
-Fig. \ref{fig:CVI} shows the minimum and maximum values of markers per element throughout the experiment with and without the CVI correction and using
-different Runge-Kutta schemes. The introduction of the CVI correction allows the
-tests to run for times longer than $t=2000$ without show empty elements, while tests without the CVI correction stop at approximately $t=600$, when at least
-one element goes empty. Fig. \ref{fig:CVI_time} shows the different number of markers per element and the markers distribution at $t=600$ obtained without and
-with the CVI correction (panels b and c, respectively) in case of a 2nd-order Runge-Kutta scheme. All data can be found at
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/CVI}.
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/CVI.pdf}
-\caption{Maximum and minimum number of markers per element throughout the experiment, with (green and orange lines) or without (purple and blue lines)
-the CVI correction and using either a 2nd-order (blue and orange lines) or a 4th-order (purple and green lines) Runge-Kutta scheme.}
-\label{fig:CVI}
-\end{figure}
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/CVI_time.pdf}
-\caption{Number of markers per element and markers distribution through time using 2nd.order Runge-Kutta scheme.
-Panels a: number of markers per element (panel a1) and distribution of the markers (panel a2) after the first
-iteration; panels b and c: comparison between number of markers per element (panels b1 and c1) and markers distribution (panels b2 and c2) without and with
-the CVI correction at $t=600$; panels d: number of markers per element (panel d1) and markers distribution (panel d2) in case of CVI correction at $t=2000$.}
-\label{fig:CVI_time}
-\end{figure}
-
-\subsection{Poiseuille flow}\label{sec:poiseuille}
-The domain is a rectangle with $L_x=2$ and $L_y=1$ and constant density and viscosity ($\rho=1$ and $\eta=1$), gravity acceleration $\bm{g}=\bm{0}$ and penalty
-parameter $\lambda=10^8$. The grid is composed by \num{40x20} elements. Velocity boundary conditions are set to no slip $(\bm{v}=\bm{0})$ at the top and the
-bottom, and a parabolic profile is imposed on the sides, with $u=y(1-y)$ and $v=0$. The analytical solution is then given by:
-\begin{eqnarray}
-u(x,y)&=&y(1-y)\nonumber \\
-v(x,y,)&=&0\nonumber \\
-p(x,y)&=&2\eta\left(\frac{L_x}{2}-x\right)\nonumber
-\end{eqnarray}
-
-The velocity field predicted by the model follows the expected parabolic profile (Fig. \ref{fig:poiseuille}a and Fig. \ref{fig:poi_plot}a) and, in case of the
-classic penalty method with no iterations, the pressure is clearly related to the divergence by means of the penalty parameter (Fig. \ref{fig:poiseuille}b and
-c and Fig. \ref{fig:poi_plot}b). Fig. \ref{fig:poiseuille}d shows that one Uzawa iteration is sufficient to bring the divergence down to \num{e-15}, with no
-correlation with the pressure field. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Poiseuille%20Flow}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Poiseuille.pdf}
-\caption{Velocity field (panel a), pressure (panel b) and divergence velocity of a Poiseuille flow in case of the classic penalty method (no iterations) and
-after one Uzawa iteration (panel c and d, respectively).}
-\label{fig:poiseuille}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Analytical.pdf}
-\caption{Velocity (panel a) and pressure (panel b) field predicted by the model for a Poiseuille flow with respect with their analytical solutions. The
-velocity field is plotted as function of the vertical coordinate in $L_x/2$ and the pressure is plotted as function of the x coordinate in $L_y/2$.}
-\label{fig:poi_plot}
-\end{figure}
-
-\subsection{Instantaneous 2D sphere}\label{sec:ist_sphere}
-The domain is a unit square with gravity $\bm{g}=(0,-1)$. The fluid has constant density and viscosity ($\rho_f=1$ and $\eta_f=1$). The sphere is in the middle
-of the domain with a radius R = 0.123456798 and has constant density and viscosity ($\rho_s=10^{-2}+\rho_f$ and $\eta_s=10^3 \cdot \eta_f$).
-We distinguish three different types of velocity boundary conditions:
-\begin{enumerate}
-\item FS: free slip conditions on all sides.
-\item NS: no slip conditions on all sides.
-\item OT: free slip conditions on the sides and the bottom, and open boundary at the top.
-\end{enumerate}
-All three cases are performed for different grid resolutions, between \num{16x16} and \num{512x512} elements with 50 randomly distributed markers per element,
-and different average schemes for the viscosity (harmonic, geometric and arithmetic). The velocity in the centre of the sphere $\bm{v}(0.5;0.5)$), minimum and
-maximum velocities ($u_{min}$, $u_{max}$, $v_{min}$, $v_{max}$) and pressures ($p_{min}$, $p_{max}$) on the entire domain, $v_{\textrm{rms}}$ and average pressure on the
-entire domain ($p_{avg}$) are compared with solutions generated by ASPECT \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and all results
-can be found at \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere2D} and 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Instantaneous_2D_Sphere}. Results in terms of the $v_{\textrm{rms}}$ for NS model 
-are also shown in Fig. \ref{fig:inst_sphere}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_NS.pdf}
-\caption{$v_{\textrm{rms}}$ as function of element size for the instantaneous 2D sphere experiment in case of no slip boundary conditions and with different average
-schemes for the viscosity. Results are compared with results obtained by other numerical codes.}
-\label{fig:inst_sphere}
-\end{figure}
-
-\subsection{Rayleigh-Taylor instability}\label{sec:rayleigh}
-This problem has been originally presented by \citet{vanKeken1997} and here it is performed as the isoviscous case (Case 1a). The domain has $L_x=0.9142$ and
-$L_y=1$ and gravity $\bm{g}=(0,-1)$. Two fluids with same constant viscosities ($\eta_1=\eta_2=1$) and different densities ($\rho_1=1000$ and $\rho_2=1010$),
-with the lighter fluid at the bottom. The initial interface between the fluids is given by $y(x)=0.2+0.02 \cos \left(\frac{\pi x}{L_x}\right)$. The experiment
-is performed with different grid sizes (\num{50x50}, \num{80x80}, \num{100x100} and \num{256x256}). A total of 1960000 markers are randomly distributed at the
-beginning of the simulation. Velocity boundary conditions are set to no slip at the top and the bottom, and to free slip at the sides of the domain.
-
-$v_{\textrm{rms}}$ as function of time is reported in Fig. \ref{fig:RT}, matching well with results shown by \citet{vanKeken1997}, \citet{Tackley2003} and
-\citet{Thieulot2014}. Fig. \ref{fig:rayleigh} shows the evolution of the experiment at different time steps. All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Rayleigh_Taylor_experiment/ISOVISCOUS}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/RT.pdf}
-\caption{$v_{\textrm{rms}}$ of the Rayleigh-Taylor experiment as function of time for different resolution of the grid.}
-\label{fig:RT}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=300px]{./Figures/Rayleigh.pdf}
-\caption{Evolution of the Rayleigh-Taylor experiment for a grid of \num{256x256} elements at $t=0, 500, 1000$ and $2000$ (panels a, b, c and d, respectively).}
-\label{fig:rayleigh}
-\end{figure}
-
-\subsection{Falling blocks}\label{sec:block}
-This benchmark is described as proposed by \citet{Gerya2003a}, \citet{Gerya2010b} and \citet{Thieulot2011}. The domain is square with $L_x=L_y=$ \SI{500}{\km}
-and the grid is composed by \num{50x50} elements with 25 markers in each element. The block is initially centred at ($x=$ \SI{250}{\km}; $y=$ \SI{400}{\km})
-and has a size of \num{100x100} km (Fig. \ref{fig:block}a). The fluid surrounding the block has $\eta_f=$ \SI{e21}{\pascal\per\s} and $\rho_f=$
-\SI{3200}{\kg\per\cubic\m}. The benchmark tests with different viscosities of the block, with $\eta_b$ from \SIrange{e15}{e27}{\pascal\per\s}. In each
-experiment also the density $\rho_b$ of the block varies from \SIrange{3220}{9900}{\kg\per\cubic\m}. Velocity boundary conditions are set to free slip
-conditions on all sides of the domain. Density distributions at $t=$ \SI{20}{\mega\year} for $\rho_b=$ \SI{3300}{\kg\per\cubic\m} and different $\eta_b$
-are plotted in Fig. \ref{fig:block}b-f, showing that the code correctly preserve the block geometry in case of large viscosity contrast, with a block stiffer
-than the surrounding fluid (Fig. \ref{fig:block}f).
-
-The velocity in the centre of the falling block at $t=0$ is measured for all experiments and, since the velocity should increase with density contrast, the
-quantity $\bm{v}/(\rho_b-\rho_f)$ is plotted as function of the viscosity contrast. All results perfectly match with those from \citet{Gerya2010b} and they
-line up on a single curve, demonstrating that the code can correctly deal with large viscosity and density contrasts (Fig. \ref{fig:falling}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Falling%20blocks}.
-
-\begin{figure}
-\noindent\includegraphics[width=300px]{./Figures/Block.pdf}
-\centering
-\caption{Density evolution of the falling block experiment at $t=$ \SI{0}{\mega\year} (panel a) and $t=$ \SI{20}{\mega\year} for different viscosities of the
-block (panels b-f).}
-\label{fig:block}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Falling.pdf}
-\caption{Initial velocity relative to the density contrast at the centre of the falling block as function of the viscosity contrast between the block and the
-surrounding fluid.}
-\label{fig:falling}
-\end{figure}
-
-\subsection{2D time-dependent Stokes sphere with deformable free surface}\label{sec:sphere}
-This experiment is performed in a unit square domain with the gravity acceleration fixed to $g_y=-1$. The sphere has $\rho_s=2$ and $\eta_s=10^3$ and is
-initially centred at ($0.5;0.6$) with radius $R=0.123456789$. The fluid surrounding the sphere has $\rho_f=1$ and $\eta_f=1$ and occupies the domain for
-$y\leq 0.75$, while for $y>0.75$ the air has $\rho_a=0$ and $\eta_a=10^{-3}$. Velocity boundary conditions are set to free slip on all sides. The Courant
-number is set to 0.25. The experiments run for 200 s using grid resolutions from \num{150x150} to \num{512x512} elements, with 25 randomly distributed markers
-per element, and for different average schemes for the viscosity. The interface between the fluid and the air is tracked by means of the markers chain. Results
-in terms of velocity and pressure fields and topography variations are compared with results obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and they can be found at
-\url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere_fs2D} and
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Time_dependent_sphere}. Results in terms of the $v_{\textrm{rms}}$ in case of 
-an arithmetic average are also shown in Fig. \ref{fig:inst_spherefs}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_arithm.pdf}
-\caption{$v_{\textrm{rms}}$ for different grid resolutions as function of time for the 2D time-dependent Stokes sphere experiment in case of an arithmetic average.
-Results are compared with results obtained by ASPECT.}
-\label{fig:inst_spherefs}
-\end{figure}
-
-\subsection{Free surface stabilisation}\label{sec:stab}
-The benchmark is performed as discussed in \citet{Kaus2010a} and \citet{Thieulot2014}. The domain is square with $L_x=L_y=$ \SI{500}{\km} and a grid resolution
-of \num{200x200} elements, each of them containing 16 markers. A buoyant fluid with $\rho_1=$ \SI{3200}{\kg\per\cubic\metre} and $\eta_1=$ \SI{e20}{\pascal\s}
-is overlain by a denser fluid with $\rho_2=$ \SI{3300}{\kg\per\cubic\metre} and $\eta_2=$ \SI{e21}{\pascal\s}. The initial interface between the fluids has an
-initial sinusoidal shape of 5 km amplitude. Velocity boundary conditions are set to free slip conditions at the sides and no slip at the bottom, while the top
-is free surface. The experiment is performed with various fixed time steps, without the use of the Courant number. The vertical position of the free surface at
-$x=L_x$ is tracked for each simulation, with or without the stabilisation algorithm.
-
-The results show an instability (drunken sailor effect) increasing the time step in case that the stabilisation algorithm is not activated
-($dt=$ \SI{4200}{\year}, green line in Fig. \ref{fig:stab}), as already observed by \citet{Kaus2010a} and \citet{Thieulot2014}. Activating the algorithm the
-instability is fixed and the simulation remains stable using time step up to \SI{50000}{\year} (red line in Fig. \ref{fig:stab}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Stabilisation_algorithm}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Kaus_random.pdf}
-\caption{Evolution of the y coordinate in $x=L_x$ as function of time for different time steps, with and without the stabilisation algorithm.}
-\label{fig:stab}
-\end{figure}
-
-\subsection{Topography relaxation}\label{sec:crameri}
-This experiment is performed as in \citet{Crameri2012} using free surface and sticky air (Case 1). The domain is rectangular with $L_x=$ \SI{2800}{\km} and
-$L_y=$ \SI{700}{\km} in case of free surface and $L_y=$ \SI{800}{\km} in case of sticky air ($h_{st}=$ \SI{100}{\km}). The grid is composed by \num{256x64}
-elements for the free surface \num{560x320} elements for the sticky air. For both cases, there is a mantle of 600 km thickness with $\eta_m=$ \SI{e21}{\pascal\s}
-overlain by a highly viscous cosine-shaped lithosphere with thickness between 93 and 107 km and $\eta_l=$ \SI{e23}{\pascal\s}. Both mediums have $\rho=$
-\SI{3300}{\kg\per\cubic\metre}. In case of sticky air experiment, the air has $\rho_a=$ \SI{0}{\kg\per\cubic\metre} and $\eta_a=$ \SI{e18}{\pascal\s}. The
-gravity acceleration is set to $g_y=$ \SI{-10}{\metre\per\square\s}. Velocity boundary conditions are set to no slip at the bottom and free slip at the sides
-of the domain. In case of sticky air, velocities on top are set to free slip conditions.
-
-The maximum topography as function of time ($h(t)$) can be analytically derived in according to \[h(t)=h_0 \exp(\gamma t)\]
-where $h_0=7$ km is the initial topography, $\gamma=\num{-0.2139e-11}$ is the characteristic relaxation rate and $t$ is the time (black line in Fig.
-\ref{fig:crameri}). Maximum topography at the characteristic relaxation time $t_{rlx}=$ \SI{14.825}{\kilo\year} can be found to be $h_{rlx}=$ \SI{2576}{\m}.
-The results show that a viscosity of the air of \SI{e18}{\pascal\s} is needed in case of $h_{st}=$ \SI{100}{\km} to correctly describe the topography relaxation
-for this problem, as pointed out by \citet{Crameri2012} (purple line in Fig. \ref{fig:crameri}). The case with a true free surface well follows the analytical
-solution (orange line in Fig. \ref{fig:crameri}). In particular, at the relaxation time in case of the true free surface predicts a maximum topography of
-\SI{2570}{\m}, with an error of \SI{6}{\m} with respect to the analytical solution. Results are also compared with results obtained with a free surface in
-UNDERWORLD (\num{256x64} elements) and SULEC (\num{401x201} elements) (with courtesy from \citealp{Crameri2012}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Topography_relaxation}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Crameri.pdf}
-\caption{Maximum topography as function of time for the Crameri benchmark (coloured lines) in comparison with the analytical solution (black line) and results
-obtained with UNDERWORLD (yellow line) and SULEC (blue line).}
-\label{fig:crameri}
-\end{figure}
-
-\subsection{Spontaneous subduction}\label{sec:subduction}
-This experiment is performed as presented by \citet{Schmeling2008}, in case of both a sticky air and a true free surface and using different average schemes
-for the viscosities. The domain is rectangular with $L_x=$ \SI{3000}{\km} and $L_y=$ \SI{700}{\km}, a grid resolution of \num{256x64} elements with 50 markers
-per element and Courant number of 0.01. At the beginning of the simulation, a 100 km-thick lithospheric layer with $\rho_m=$ \SI{3300}{\kg\per\cubic\metre}
-and $\eta_m=$ \SI{e23}{\pascal\s} is located at the top of the domain for $x$ between 1000 and 3000 km, while the asthenospheric mantle has $\rho_m=$
-\SI{3200}{\pascal\s} and $\eta_m=$ \SI{e21}{\pascal\s}. In addition, a 200 km-depth lithospheric slab is already subducted in the mantle in order to have
-a spontaneous subduction. In case of sticky air, $L_y=750$ km and a 50 km-thick air layer with $\eta_a$ from \SIrange{e19}{e21}{\pascal\s} overlie the mantle.
-Velocity boundary conditions are set to free slip on the sides and on the bottom of the domain. In case of sticky air, velocities on the top boundary are set
-to free slip conditions as well.
-
-The maximum depth of the slab is tracked for all the simulations (Fig. \ref{fig:subduction}) and they are compared with results obtained by
-\citet{Schmeling2008} with models with regular grids and comparable grid resolutions (rectangles in Fig. \ref{fig:subduction}). Low viscous sticky air models
-(i.e., \SI{e19}{\pascal\s}) enlighten a strong dependence of the sinking velocity with the chosen average scheme (dashed lines in Fig. \ref{fig:subduction}),
-while high viscous sticky air models (i.e., \SI{e21}{\pascal\s}) show a higher resistance at the trench, underestimating the correct solution (dotted lines in
-Fig. \ref{fig:subduction}). A high resistance at the trench can be also observed in case of a true free surface for the low resolution grid used (continuous
-lines in Fig. \ref{fig:subduction}), as pointed out by \citet{Schmeling2008}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Spontaneous_subduction}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Subduction.pdf}
-\caption{Maximum depth of the slab as function of time for all the simulations of the \citet{Schmeling2008} benchmark, in case of harmonic, geometric and
-arithmetic means (red, green and blue lines, respectively) using a sticky air or a true free surface (discontinuous and continuous lines, respectively).
-Red, green and blue rectangular areas indicate the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{500}{\km} in case of sticky air,
-while the grey rectangular area indicates the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{600}{\km} in case of true free surface.}
-\label{fig:subduction}
-\end{figure}
-
-\subsection{The slab detachment}\label{sec:slab}
-The slab detachment benchmark is performed as described by \citet{Schmalholz2011} and \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{1000}{\km}
-and $L_y=$ \SI{660}{\km} and a grid resolution of \num{256x256} elements. A non-linear viscous T-shaped layer with $\rho_l=$ \SI{3300}{\kg\per\cubic\metre} is
-placed at the top of the domain and surrounded by a linear viscous fluid with $\rho_f=$ \SI{3150}{\kg\per\cubic\m}. The top layer is 80 km-thick and a 250
-km-long and 80 km-wide slab is placed at $x=L_x/2$ (Fig. \ref{fig:slab}a). The effective viscosity of the top layer is given by
-\[\eta_{eff}=\eta_0 I_2^{\frac{(1-n)}{n}}\]
-with $\eta_0=$ \SI{4.75e11}{\pascal\s} and $n=4$, while the surrounding fluid has $\eta_f=$ \SI{e21}{\pascal\s}. The effective viscosity is capped between
-$\eta_{min}=$ \SI{e21}{\pascal\s} and $\eta_{max}=$ \SI{e25}{\pascal\s}. Velocity boundary conditions are set to free slip at the top and the bottom, and to
-no slip at the sides of the domain.
-
-The sinking of the slab determines a decrease of the effective viscosity in the T-shaped layer during the first part of the evolution (Fig. \ref{fig:slab}a),
-as consequence of local strain rates, while effective viscosities increase up to $\eta_{max}$ in the top layer after approximately \SI{20}{\mega\year}, when
-necking is complete (Fig. \ref{fig:slab}b).
-
-The width of necking is tracked by means of markers position at the side of the slab. Necking width and time are normalised in function of the initial slab
-(\SI{80}{\km}) and the characteristic time ($t_c=$ \SI{7.1158e14}{\s}) \citep{Schmalholz2011,Glerum2018}. The results in case of arithmetic, geometric and
-harmonic mean of the viscosity are shown in Fig. \ref{fig:necking}, compared with results from \citet{Glerum2018} for which an infinity norm mean was used.
-The influence of the average schemes agrees with results shown in Section \ref{sec:subduction}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Slab_detachment}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Slab.pdf}
-\caption{Effective viscosity for the slab detachment benchmark at the beginning of the evolution (panel a) and when necking is complete (panel b).}
-\label{fig:slab}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Necking.pdf}
-\caption{Normalised width of the necking of the slab detachment benchmark as function of normalised time.}
-\label{fig:necking}
-\end{figure}
-
-\subsection{The indenter experiment}\label{sec:indenter}
-The indenter benchmark simulate a rigid punch on a purely plastic von Mises material, for which exists an analytical solution
-\citep{Thieulot2008,Thieulot2014,Glerum2018}. The domain is a unit square with a grid resolution of \num{256x256} elements. The gravity acceleration is set
-to $\bm{g}=\bm{0}$ and effective viscosity is capped using $\eta_{min}=10^{-4}$ and $\eta_{max}=10^3$. The tolerance for the convergence of the non-linear
-solution is $tol=10^{-9}$, with a maximum number of non-linear iterations set to 500. The medium in the domain has $\rho=0.01$, an initial viscosity
-$\eta_0=10$, cohesion $C=1$ and an angle of internal friction $\phi=0$°. Velocity boundary conditions are set to no slip at the bottom and free slip at the
-sides of the domain. The top of the domain is open with the exception of the central portion (punch area) with width $w_p=0.125$, where $v=-1.05$ and $u$ is
-fixed either to 0 (rough punch experiment) or free (smooth punch experiment).
-
-Fig. \ref{fig:indenter} show the results in terms of viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure
-(panels g and h) for both experiments. The analytical solution indicate that pressure at ($0.5;1$) and ($0.5 \pm w_p;1$)is $p=\pi +1$ and $p=1$, respectively
-(black and red lines in Fig. \ref{fig:smooth_p}). Fig. \ref{fig:smooth_p} shows a clear improvement in the pressure solution when passing from a rough to a
-smooth punch (green and blue lines in Fig. \ref{fig:smooth_p}, respectively), as previously observed by \citet{Thieulot2014} and \citet{Glerum2018}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Indenter}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Indenter.pdf}
-\caption{Viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure (panels g and h) fields for rough (left column) and
-smooth (right column) punch experiments.}
-\label{fig:indenter}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Smooth_Rough.pdf}
-\caption{Nodal pressure in function of the x coordinate at $y=1$ for rough (green line) and smooth (blue line) punch experiment. Black and red lines indicate
-the analytical solution at $x=0.5$ and $x=0.5 \pm w_p$, respectively.}
-\label{fig:smooth_p}
-\end{figure}
-
-\subsection{The brick experiment}\label{sec:brick}
-An instantaneous version of the brick benchmark is performed to verify the correctness of the pressure-dependent plasticity for different angle of internal
-friction, as proposed by \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{40}{\km} and $L_y=$ \SI{10}{\km} and a grid resolution of \num{512x128}
-elements. The gravity acceleration is set to $g_y=$ \SI{-10}{\m\square\s} and effective viscosity is capped using $\eta_{min}=$ \SI{e19}{\pascal\s} and
-$\eta_{max}=$ \SI{e26}{\pascal\s}. The tolerance for the convergence of the non-linear solution is $tol=10^{-7}$, with a maximum number of non-linear iterations
-set to 1000. A 800 m-wide and 400 m-height inclusion with $\eta_b=$ \SI{e20}{\pascal\s} and $\rho_b=$ \SI{2700}{\kg\per\cubic\m} is placed at the bottom of the
-domain at $x=L_x/2$. The inclusion is surrounded by a non-linear viscous medium with $\rho_m=$ \SI{2700}{\kg\per\cubic\m}, an initial viscosity of
-\SI{e23}{\pascal\s}, a linear viscous viscosity of \SI{e25}{\pascal\s} and a cohesion of \SI{40}{\mega\pascal}. Velocities are set to free slip conditions at
-the bottom of the domain and the top is open. Velocities on sides of the domain are fixed to $u=$ \SI{+-2e-11}{\m\per\s} and $v=0$. The experiment is performed
-in compressional and extensional contexts with an angle of internal friction $\phi$ of the non-linear medium variable between 0° and 30°.
-
-As expected, two shear bands stem from the inclusion with variable angles in relation to both the dynamics context and the internal friction angle. Shear band
-angles formed at 45° for $\phi=0$° in both extensional and compressional contexts (Fig. \ref{fig:brick_beam}a), while they are different in case of $\phi\neq0$°.
-Shear band angles in case of $\phi=20$° are shown for compression and extension in Fig. \ref{fig:brick_beam}b  and c, respectively. Values of shear band angles
-as function of internal friction angles from 0° to 30° are extracted at different depths and minimum and maximum values are plotted in Fig. \ref{fig:brick},
-in comparison with theoretical Roscoe, Arthur and Coulomb shear band angles. For all tests the tolerance for nonlinear convergence is defined $tol=10^{-7}$
-and none converges before the maximum number of iterations ($it_{max}=1000$) is reached. However, tests with internal friction angles up to 15° show a constant
-decrease of the velocity residuals, which can not be observed for for tests with higher internal friction angles (Fig. \ref{fig:convergence}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Brick_experiment}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Brick_Beam.pdf}
-\caption{Shear band angles predicted for the brick experiment for $\phi=0$° (panel a) and for $\phi=20$° in case of compression and extension (panels b and c,
-respectively).}
-\label{fig:brick_beam}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Brick.pdf}
-\caption{Shear band angles predicted for the brick experiment in case of compressional and extensional contexts as function of different internal angle of
-friction (continuous black line and red dots), compared with theoretical Roscoe, Artur and Coulomb angles (discontinuous black lines). Red lines indicate the
-range of angles calculated at different depths.}
-\label{fig:brick}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \noindent\includegraphics[width=500px]{./Figures/convergence.pdf}
-  \caption{Convergence rates for all tests. Normalised residuals (left column) are compared with velocity residuals (right column) for extensional (first row)
-  and compressional (second row) settings.}
-  \label{fig:convergence}
-  \end{figure}
-
-\subsection{Advection stabilisation}\label{sec:advection}
-The 1D advection problem proposed by \citet{Donea2003} and \citet{Thieulot2011} is performed to verify the effectiveness of the SUPG method to stabilise the
-advection term of the energy equation. The domain is a 1D segment with $L_x=1$ composed by 50 elements and a discontinuity in the thermal field placed at
-$x=0.25$. Temperature is set to 1 for $x \leq 0.25$ and to 0 for $x > 0.25$. Velocity is set to ${u}=1$ in the entire domain. The simulation is performed for
-250 time steps, with $dt=0.002$, so the thermal discontinuity should be at $x=0.75$ at the end of the simulation.
-
-Temperature profiles at the end of the simulation are shown in \ref{fig:advection} as function of the dimensionless coefficient $\gamma=\tau \bm{v}/h$. In case
-of the classic Galerkin method ($\gamma=0$, blue line in \ref{fig:advection}) the final thermal profile is characterised by strong oscillations, which are
-eliminated in case of the SUPG method ($\gamma=0.045$, orange line in \ref{fig:advection}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Advection_stabilisation}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Advection.pdf}
-\caption{Temperature profile as function of \textit{x} for the advection stabilisation benchmark. Purple line indicates the initial temperature profile; the
-green line indicates the analytical temperature profile after 250 time steps; blue line indicates the temperature profile after 250 time steps in case of the
-classic Galerkin method ($\gamma=0$); orange line indicates the temperature profile after 250 time steps in case of the SUPG method ($\gamma=0.045$).}
-\label{fig:advection}
-\end{figure}
-
-\subsection{Shear heating}\label{sec:simple_shear}
-This exercise is performed in an unit square domain composed by \num{128x128} elements. The velocity field is prescribed on the entire domain with
-$\bm{v}=(L_y-y)y\bm{e_x}$; viscosity, density and specific heat are fixed to 1, while thermal conductivity and radiogenic energy are fixed to 0. Therefore,
-the energy equation (Eq. \ref{eq:energy}) can be simplified as \[\frac{\partial T}{\partial t}=H_s\] and fixing $T(t=0)=0$, the temperature can be find as
-\[T(t)=H_s t\]
-In this case we have
-\begin{eqnarray}
-\dot{\epsilon_{xx}}&=&\frac{\partial u}{\partial x}=0 \nonumber \\
-\dot{\epsilon_{yy}}&=&\frac{\partial v}{\partial y}=0 \nonumber \\
-\dot{\epsilon_{xy}}&=&\frac{1}{2}\left(\frac{\partial v}{\partial x}+\frac{\partial u}{\partial y}\right)=\frac{1}{2}(L_y-2y)\nonumber
-\end{eqnarray}
-and, simplifying Eq. \ref{eq:shear}, shear heating can be calculated as \[H_s(x,y)=(1-2y)^2\]
-
-The solution predicted by the model in terms of velocity, temperature and shear heating match well with the analytical solutions (Fig. \ref{fig:analytical_en}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Simple_shear}.
-
-%\begin{figure}
-%\centering
-%\noindent\includegraphics[width=400px]{./Figures/Energy_Thieulot.pdf}
-%\caption{Temperature and shear heating predicted by the model after 1 time step.}
-%\label{fig:en_thieulot}
-%\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=500px]{./Figures/Analytical_Shear.pdf}
-\caption{Velocity (panel a), temperature (panel b) and shear heating (panel c) as function of the \textit{y} coordinate for the simple shear experiment. The
-solutions predicted by the model (red lines) are compared with the analytical solutions (dashed blue lines).}
-\label{fig:analytical_en}
-\end{figure}
-
-\subsection{Shear and adiabatic heating}\label{sec:shear}
-This problem is performed as presented in Exercise 9.4 in \citet{Gerya2010b}. Two materials are vertically separated in a rectangular domain with $L_x=$
-\SI{1000}{\km}, $L_y=$ \SI{1500}{\km} and a grid resolution of \num{30x20} elements. Constant thermal coefficient expansion ($\alpha=$ \SI{3e-5}{\per\kelvin}),
-temperature ($T=$ \SI{1300}{\kelvin}) and gravity acceleration ($g_y=$ \SI{-10}{\m\square\s}) are assumed in the whole domain. Fluid 1 (on the left side of the
-domain) has $\rho_1=$ \SI{3200}{\kg\per\cubic\m} and $\eta_1=$ \SI{e20}{\pascal\s}; fluid 2 (on the right side of the domain) has $\rho_2=$
-\SI{3300}{\kg\per\cubic\m} and $\eta_2=$ \SI{e22}{\pascal\s}. Velocity boundary conditions are set to free slip on all sides of the domain.
-
-As shown in Fig. \ref{fig:energy}, both shear and adiabatic heating predicted by the code (first row in Fig. \ref{fig:energy}) well recreate the results
-obtained by means of example ${Shear\_adiabatic\_heating.m}$ from \citet{Gerya2010b} (second row in Fig. \ref{fig:energy}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Adiabatic%2BShear}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Energy.pdf}
-\caption{Comparison between shear (first column), adiabatic (second column) and total (third column) energy predicted by the code (panels a, b and c) and those
-created using example ${Shear\_adiabatic\_heating.m}$ from Exercise 9.4 in \citet{Gerya2010b} (panels d, e and f).}
-\label{fig:energy}
-\end{figure}
-
-\subsection{Mantle convection}\label{sec:mantle}
-This problem is performed as presented by \citet{Blankenbach1989} (constant viscosity cases) and \citet{Thieulot2014}, in a 2D unit square domain with gravity
-acceleration $g_y=10^{10}Ra$. The experiment is performed with three different Rayleigh numbers ($Ra=10^4$, $10^5$ and $10^6$) and with different grid
-resolution (between \num{32x32} and \num{128x128} elements). The fluid has constant viscosity, initial density, heat capacity, thermal conductivity
-($\eta= \rho_0= C_p = k=1$), reference temperature ($T_0=0$) and thermal expansion coefficient ($\alpha=10^{-10}$). Temperatures are set to 0 on top and 1 on
-bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\]
-
-The solution generated by the code in terms of $v_{\textrm{rms}}$ (calculated as in Sec. \ref{sec:error}) and Nu as function of time are reported for all the
-simulations in Fig. \ref{fig:mantle}. At the steady state, $v_{\textrm{rms}}$ and Nu of all simulations converge well toward the values from \citet{Blankenbach1989},
-with lower errors for higher resolution grids (Table \ref{tab:mantle}). All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Mantle_convection}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Convection.pdf}
-\caption{$v_{\textrm{rms}}$ (panels a, c and e) and Nu (panels b, d and f) for the mantle convection benchmark as function of time for different grid resolution.
-Panels a and b show the results for $Ra=10^4$; Panels c and d show the results for $Ra=10^5$; panels e and f show the results for $Ra=10^6$. Purple lines
-indicate the convergence values for the $v_{\textrm{rms}}$ and the Nu from \citet{Blankenbach1989} at the steady state.}
-\label{fig:mantle}
-\end{figure}
-
-\begin{table}[H]
-\caption{Comparison between $v_{\textrm{rms}}$ and Nu predicted by the code for the mantle convection experiment and same values as reported in literature.}
-\centering
-\small
-\begin{tabular}{|>{\centering}m{0.09\textwidth}|>{\centering}m{0.05\textwidth}|>{\centering}m{0.24\textwidth}|>{\centering}m{0.15\textwidth}|
-  >{\centering}m{0.1\textwidth} >{\centering}m{0.1\textwidth} >{\centering\arraybackslash}m{0.1\textwidth}|}
-\toprule
-& & \multirow{2}{*}{\citet{Blankenbach1989}} & \citet{Thieulot2014} & \multicolumn{3}{c|}{FALCON}  \\
- & & & \num{200x200} & \num{32x32} & \num{64x64} & \num{128x128}  \\
-\midrule
-\multirow{2}{*}{$Ra=10^4$} & $v_{\textrm{rms}}$ & $42.864947 \pm 0.000020$ & 42.867 & 42.83226 & 42.852793 & 42.861394  \\
-           & Nu      & $4.884409 \pm 0.000010$  & 4.882  & 4.781297 & 4.857475  & 4.877573   \\
-\hline
-\multirow{2}{*}{$Ra=10^5$} & $v_{\textrm{rms}}$ & $193.21454 \pm 0.00010$  & 193.255& 193.872643 & 193.377472 & 193.252290  \\
-           & Nu      & $10.534095 \pm 0.000010$ & 10.507 & 9.602514 & 10.270735  & 10.465629   \\
-\hline
-\multirow{2}{*}{$Ra=10^6$} & $v_{\textrm{rms}}$ & $833.98977 \pm 0.00020$  & 834.712& 848.091176 & 837.767911 & 834.945793  \\
-           & Nu      & $21.972465 \pm 0.000020$ & 21.695 & 15.999266 & 19.703682  & 21.306939   \\
-\bottomrule
-\end{tabular}
-\label{tab:mantle}
-\end{table}
-
-\subsection{Viscoplastic mantle convection}\label{sec:visco_mantle}
-This benchmark is performed as Cases 1-5a described in \citet{Tosi2015} in a 2D unit square domain with two different grid resolutions of \num{32x32} 
-and \num{100x100} elements. Reference viscosity, reference density, heat capacity, thermal conductivity and thermal expansion coefficient are fixed to a constant
-value of 1 ($\eta_0= \rho_0= C_p = k=\alpha=1$), while the gravity has been chosen as $g_y=10^2$, in order to obtain a Rayleigh number $Ra=10^2$. Temperatures
-are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\].
-The viscosity field $\eta$ is calculated as the harmonic average between a linear part $\eta_{lin}$ that depends on temperature only or on temperature and 
-depth $y$ and a nonlinear-plastic part $\eta_{pl}$ that depends on the strain rate $\dot{\bm{\epsilon}}$, as follows
-\[\eta(T,y,\dot{\bm{\epsilon}})=2\left(\frac{1}{\eta_{lin}(T,y)}+\frac{1}{\eta_{pl}(\dot{\bm{\epsilon}})}\right)^{-1}\]
-The linear and the nonlinear-plastic parts of the viscosity are calculated following \citet{Tosi2015} as follows
-\[\eta_{lin}(T,y)=\exp(-\ln(\Delta\eta_T)T+\ln(\Delta\eta_y)y)\]
-\[\eta_{pl}(\dot{\bm{\epsilon}})=\eta^*+\frac{\sigma_Y}{\sqrt{\dot{\bm{\epsilon}}:\dot{\bm{\epsilon}}}}\]
-where $\Delta\eta_T$, $\Delta\eta_y$, $\eta^*$ and $\sigma_Y$ are parameters chosen as listed in Table \ref{tab:tosi_setup}.
-
-For each case, some diagnostic quantities measured at the steady state are reported in Table \ref{tab:visco_mantle} (see Section \ref{sec:error} for 
-the computation of all the quantities), showing a well fit with results described in \citet{Tosi2015}. Data and charts for all cases can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Viscoplastic_mantle_convection}.
-
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{white!100!}{gray!25!}
-%\begin{table}[H]
-%  \caption{Results of the viscoplastic mantle convection benchmark for Cases 1-4 compared with results obtained with ELEFANT and YACC (from \citealp{Tosi2015}).}
-%  \centering
-%  \resizebox{\textwidth}{!}{
-%\begin{tabular}{|c|ccc|ccc|ccc|ccc|}
-%  \toprule
-%    & \multicolumn{3}{c|}{\textbf{Case 1}} & \multicolumn{3}{c|}{\textbf{Case 2}} & \multicolumn{3}{c|}{\textbf{Case 3}} &
-%    \multicolumn{3}{c|}{\textbf{Case 4}} \\
-%   & FALCON & ELEFANT& YACC & FALCON& ELEFANT& YACC & FALCON & ELEFANT & YACC & FALCON & ELEFANT & YACC \\
-%  \midrule
-%   $\left\langle T \right\rangle$ & 0.7759 & 0.7758 & 0.7767 & 0.6043 & 0.6036 & 0.6058 & 0.7275 & 0.7275 & 0.7286 & & 0.5277 & 0.5289\\
-%    Nu$^{top}$ & 3.4280 & 3.4214 & 3.4298 & 8.5242 & 8.5115 & 8.5278 & 3.0368 & 3.0347 & 3.0374 && 6.5912 & 6.5572 \\
-%    Nu$^{bottom}$ & 3.3168 & 3.313 & 3.3143 & 8.4599 & 8.4465 & 8.3990 & 3.0086 & 2.9908 & 2.9628 && 6.5834 & 6.5243 \\
-%    $v_{\textrm{rms}}$ & 249.901 & 249.134 & 251.7997 & 141.2451 & 140.7087 & 142.2020 & 100.2776 & 100.1208 & 100.9467 && 79.1105 & 79.6202 \\
-%    $v_{\textrm{rms}}^{top}$ & 1.8907 & 1.8642 & 1.8298 & 104.7928 & 104.1228 & 105.1767 & 2.0832 & 2.0652 & 2.0374 && 74.7596 & 75.8414 \\
-%    $u_{max}^{top}$ & 2.6340 & 2.6119 & 2.5516 & 121.9212& 121.766 & 122.3238 & 2.9086 & 2.9019 & 2.8458 && 88.9146 & 89.2940 \\
-%    $\eta_{min}$ & - & - & - & $2.0303\times10^{-5}$ & $2.0600\times10^{-5}$ & $1.9853\times10^{-5}$ & $4.8129\times10^{-5}$ & $4.8080\times10^{-5}$ & $4.7907\times10^{-5}$ &
-%    & $1.9860\times10^{-4}$ & $1.9174\times10^{-4}$ \\
-%    $\eta_{max}$ & - & - & - & 1.3325 & 1.85 & 1.6095 & 0.9360 & 0.9023 & 1 && 1.52 & 1.6773 \\
-%    $\left\langle W \right\rangle$ & 2.4299 & 2.4316 & 2.4583 & 7.5648 & 7.5665 & 7.6222 & 2.0387 & 2.0384 & 2.0400 && 5.6216 & 5.6512 \\
-%    $\left\langle \Phi \right\rangle/Ra$ & 2.4311 & 2.4276 & 2.4333 & 7.5661 & 7.5585 & 7.6194 & 2.0386 & 2.0370 & 2.0335 && 5.6182 & 5.6463 \\
-%    $\delta$ & 0.05\% & 0.16\% & 1.02\% & 0.02\% & 0.11\% & 0.04\% & $<$0.01\% & 0.07\% & 0.32\% && 0.06\% & 0.09\% \\
-%  \bottomrule
-%  \end{tabular}}
-%  \label{tab:visco_mantle}
-%  \end{table}
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{2}{white!100!}{gray!25!}
-\begin{table}[H]
-  \caption{Benchmark cases and corresponding parameters.}
-  \centering
-  \small
-\begin{tabular}{|>{\centering}m{0.08\textwidth}|>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}
-  >{\centering}m{0.05\textwidth} >{\centering}m{0.05\textwidth} |>{\centering\arraybackslash}m{0.2\textwidth}|}
-  \toprule
-    Case & $Ra$ & $\Delta\eta_T$ & $\Delta\eta_y$ & $\eta^*$ & $\sigma_Y$ & Convective regime \\
-  \midrule
-  \rowcolor{gray!25!}
-    1 & $10^2$ & $10^5$ & 1 & - & - & Stagnant lid \\
-    2 & $10^2$ & $10^5$ & 1 & $10^{-3}$ & 1 & Mobile lid  \\
-    \rowcolor{gray!25!}
-    3 & $10^2$ & $10^5$ & 10 & - & - & Stagnant lid  \\
-    4 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 1 & Mobile lid \\
-    \rowcolor{gray!25!}
-    5 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 4 & Periodic \\
-  \bottomrule
-  \end{tabular}
-  \label{tab:tosi_setup}
-  \end{table}
-
-\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{gray!25!}{white!100!}
-\begin{table}
-  \caption{Results of the viscoplastic mantle convection benchmark for different grid resolutions in Cases 1-5 compared with results obtained with ELEFANT
-  and YACC ($100\times100$ elements, from \citealp{Tosi2015}) and in Stone 28 ($32\times32$ elements, from
-  \url{https://github.com/cedrict/fieldstone/tree/master/python_codes/fieldstone_28}).}
-  \centering
-  \resizebox{\textwidth}{!}{
-\begin{tabular}{|cc|ccccccccccc|}
-  \toprule
-  \multicolumn{13}{|c|}{\textbf{Case 1}} \\
-  \midrule
-  Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-  & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-  \hline
-  \rowcolor{gray!25!}
-    YACC & $100\times100$ & 0.7767 & 3.4298 & 3.3143 & 251.7997 & 1.8298 & 2.5516 & - & - & 2.4583 & 2.4333 & 1.02\% \\
-    ELEFANT & $100\times100$ & 0.7758 & 3.4214 & 3.313 & 249.134 & 1.8642 & 2.6119 & - & - & 2.4316 & 2.4276 & 0.16\% \\
-    \rowcolor{gray!25!}
-    Stone 28 & $32\times32$ & 0.7737 & 3.3987 & - & 243.872 & - & - & - & - & - & - & - \\
-%    FALCON & $32\times32$ & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & - & - & 2.4745 & 2.4864 & 0.48\% \\
-%    \rowcolor{gray!25!}
-    FALCON & $100\times100$ & 0.7759 & 3.4280 & 3.3168 & 249.901 & 1.8907 & 2.6340 & - & - & 2.4299 & 2.4311 & 0.05\% \\
-    \midrule
-    \multicolumn{13}{|c|}{\textbf{Case 2}} \\
-    \midrule
-    Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-    & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-    \hline
-    \rowcolor{gray!25!}
-     YACC & $100\times100$ & 0.6058 & 8.5278 & 8.3990 & 142.2020 & 105.1767 & 122.3238 & $1.9853\times10^{-5}$ & 1.6095 & 7.6222 & 7.6194 & 0.04\% \\
-     ELEFANT & $100\times100$ & 0.6034 & 8.5115 & 8.4465 & 140.8067 & 104.1228 & 121.766 & $2.0600\times10^{-5}$ & 1.85 & 7.5665 & 7.5585 & 0.11\% \\
-     \rowcolor{gray!25!}
-     Stone 28 & $32\times32$ & 0.6052 & 8.1715 & - & 141.518 & - & - & - & - & - & - & - \\
-%     FALCON & $32\times32$ & 0.6454 & 8.1422 & 7.4020 & 156.6908 & 95.7226 & 116.9302 & $2.2572\times10^{-5}$ & 0.7194 & 7.4289 & 7.4452 & 0.22\% \\
-%     \rowcolor{gray!25!}
-     FALCON & $100\times100$ & 0.6044 & 8.5266 & 8.4590 & 141.3684 & 104.8867 & 122.0470 & $2.0333\times10^{-5}$ & 1.2914 & 7.5707 & 7.5718 & 0.01\% \\
-     \midrule
-     \multicolumn{13}{|c|}{\textbf{Case 3}} \\
-     \midrule
-     Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-     & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-     \hline
-     \rowcolor{gray!25!} 
-     YACC & $100\times100$ & 0.7286 & 3.0374 & 2.9628 & 100.9467 & 2.0374 & 2.8458 & $4.7907\times10^{-5}$ & 1 & 2.0400 & 2.0335 & 0.32\% \\
-      ELEFANT & $100\times100$ & 0.7275 & 3.0347 & 2.9908 & 100.1208 & 2.0652 & 2.9019 & $4.8080\times10^{-5}$ & 0.9023 & 2.0384 & 2.037 & 0.07\% \\
-      \rowcolor{gray!25!}
-      Stone 28 & $32\times32$ & 0.7277 & 3.0347 & - & 100.018 & - & - & - & - & - & - & - \\
-%      FALCON & $32\times32$ & 0.7281 & 3.0618 & 2.8132 & 101.0092 & 2.1630 & 3.0178 & $4.8500\times10^{-5}$ & 0.7845 & 2.0699 & 2.0695 & 0.02\% \\
-%      \rowcolor{gray!25!}
-      FALCON & $100\times100$ & 0.7275 & 3.0368 & 3.0086 & 100.2776 & 2.0832 & 2.9086 & $4.8129\times10^{-5}$ & 0.9360 & 2.0387 & 2.0386 & $<$0.01\% \\
-      \midrule
-      \multicolumn{13}{|c|}{\textbf{Case 4}} \\
-      \midrule
-      Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-      & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-      \hline
-      \rowcolor{gray!25!} 
-      YACC & $100\times100$ & 0.5289 & 6.5572 & 6.5243 & 79.6202 & 75.4814 & 89.2940 & $1.9174\times10^{-4}$ & 1.6773 & 5.6512 & 5.6463 & 0.09\% \\
-       ELEFANT & $100\times100$ & 0.5277 & 6.5912 & 6.5834 & 79.1105 & 74.7596 & 88.9146 & $1.9860\times10^{-4}$ & 1.5200 & 5.6216 & 5.6182 & 0.06\% \\
-       \rowcolor{gray!25!}
-       Stone 28 & $32\times32$ & 0.5286 & 6.4036 & - & 79.4746 & - & - & - & - & - & - & - \\
-%       FALCON & $32\times32$ & 0.5334 & 6.4679 & 6.3435 & 80.6153 & 76.0511 & 90.2058 & $2.0393\times10^{-4}$ & 0.6453 & 5.6676 & 5.6698 & 0.04\% \\
-%       \rowcolor{gray!25!}
-       FALCON & $100\times100$ & 0.5279 & 6.5897 & 6.5830 & 79.1220 & 75.0715 & 88.8484 & $1.9693\times10^{-4}$ & 1.4133 & 5.6127 & 5.6113 & 0.03\% \\
-       \midrule
-       \multicolumn{13}{|c|}{\textbf{Case 5}} \\
-       \midrule
-       Code & Resolution & $P$ & $\left\langle T \right\rangle_{min}$ & $\left\langle T \right\rangle_{max}$ & Nu$^{top}_{min}$ & Nu$^{top}_{max}$ &
-       $v_{\textrm{rms}}^{min}$ & $v_{\textrm{rms}}^{max}$ & \multicolumn{2}{c}{$(\left\langle \Phi \right\rangle/Ra)_{min}$} & 
-       \multicolumn{2}{c|}{$(\left\langle \Phi \right\rangle/Ra)_{max}$} \\
-       \hline
-       \rowcolor{gray!25!}
-         YACC & $100\times100$ & 0.0792 & 0.6543 & 0.6722 & 2.6950 & 7.2792 & 41.9813 & 98.8230 & \multicolumn{2}{c}{1.4347} & 
-         \multicolumn{2}{c|}{9.2748} \\
-         ELEFANT & $100\times100$ & 0.076 & 0.6514 & 0.6693 & 2.6808 & 7.3506 & 41.2578 & 100.089 & \multicolumn{2}{c}{1.3324} & 
-         \multicolumn{2}{c|}{9.5289} \\
-         \rowcolor{gray!25!}
-         Stone 28 & $32\times32$ & 0.0764 & 0.6521 & 0.6697 & 2.6912 & 7.0763 & 41.61 & 98.183 & \multicolumn{2}{c}{-} & 
-         \multicolumn{2}{c|}{-} \\
-     %    FALCON & $32\times32$ & & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & \multicolumn{2}{c}{-} & 
-     %     \multicolumn{2}{c|}{-} \\
-     %    \rowcolor{gray!25!}
-         FALCON & $100\times100$ & 0.0775 & 0.6519 & 0.6697 & 2.6847 & 7.3629 & 41.5256 & 99.5466 & \multicolumn{2}{c}{1.3399} & 
-         \multicolumn{2}{c|}{9.3993} \\
-  \bottomrule
-  \end{tabular}}
-  \label{tab:visco_mantle}
-  \end{table}
-
-\subsection{Thin layer entrainment}\label{sec:thin}
-This experiment is performed as originally proposed by \citet{vanKeken1997} with two incompressible fluids in a rectangular domain with $L_x=2$ and $L_y=1$
-and gravity acceleration $g_y=10^{10}Ra$. The Rayleigh number and the compositional Rayleigh number are fixed to $Ra=300000$ and $Rc=450000$, respectively.
-Both fluids have constant viscosity, thermal conductivity, specific heat ($\eta=\rho=C_p=1$) and thermal expansion coefficient ($\alpha=10^{-10}$). Fluid 1
-has a density $\rho_1=1$, while fluid 2 is denser ($\rho_2=\rho_1+1.5 \cdot 10^{-10}$) and is placed at the bottom of the domain, for $y \leq 0.025$.
-Temperature are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides of the domain. The initial
-temperature field is given by
-\[T(x,y)=T_u(x,y)+T_l(x,y)+T_r(x,y)+T_s(x,y)-\frac{3}{2}\]
-with
-\begin{eqnarray}
-T_u(x,y)&=&\frac{1}{2}erf\left(\frac{1-y}{2}\sqrt{\frac{u_0}{x}}\right)\nonumber \\
-T_l(x,y)&=&1-\frac{1}{2}erf\left(\frac{y}{2}\sqrt{\frac{u_0}{L_x-x}}\right)\nonumber \\
-T_r(x,y)&=&\frac{1}{2}+\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{y+1}} \exp\left(-\frac{x^2u_0}{4y+4}\right)\nonumber \\
-T_s(x,y)&=&\frac{1}{2}-\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{2-y}} \exp\left(-\frac{(L_x-x)^2u_0}{8-4y}\right)\nonumber
-\end{eqnarray}
-and
-\begin{eqnarray}
-u_0&=&\frac{L_x^{7/3}}{(1+L_x^4)^{2/3}}\left(\frac{Ra}{2\sqrt{\pi}}\right)^{\frac{2}{3}}\nonumber \\
-Q&=&2\sqrt{\frac{L_x}{\pi u_0}}\nonumber
-\end{eqnarray}
-
-The experiment is performed with two grid resolution (\num{125x40} and \num{200x80} elements), with 100 markers per element and CFL of 0.25. $v_{\textrm{rms}}$
-(see Sec. \ref{sec:error}) calculated for both the simulations match well with results from \citet{vanKeken1997} and obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and ELEFANT \citep{Thieulot2014}, obtained with similar grid resolutions
-(Fig. \ref{fig:thin}). In addition, a variety of simulations with different resolutions and aspect ratios are performed to verify that the sensitivity of the
-initial velocity field. Results match well with results from \citet{vanKeken1997} (in the grey area) and \citet{Thieulot2014} (Fig. \ref{fig:thin_initial}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Thin_layer}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_thin.pdf}
-\caption{$v_{\textrm{rms}}$ for the thin layer experiment as function of time for different grid resolution. Results are compared with results obtained by
-\citet{vanKeken1997} (black lines) and with ASPECT and ELEFANT.}
-\label{fig:thin}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrmszero.pdf}
-\caption{$v_{\textrm{rms}}$ at $t=0$ for the thin layer experiment as function of the element size for different aspect ratios of the elements. Results are compared
-with results obtained by \citet{vanKeken1997} (black dots in the grey area) and with ASPECT and ELEFANT.}
-\label{fig:thin_initial}
-\end{figure}
-
 \subsection{Sinking hydrated cylinder}\label{sec:quinquis}
 This experiment is performed as explained in \citet{Quinquis2014} and simulates the sinking of a cold, hydrated cylinder into a warm, dry mantle. The domain
-is square with $L_x=L_y=$ \SI{300}{\km}, a grid resolution of \num{300x300} elements and 25 markers per element. The sinking cylinder is located at $x=$
+is square with $L_x=L_y=$ \SI{300}{\km}, a grid resolution of $300\times300$ elements and 25 markers per element. The sinking cylinder is located at $x=$
 \SI{150}{\km} and $y=$ \SI{170}{\km} with a radius of \SI{20}{\km} and has $\rho_c=$ \SI{3250}{\kg\per\cubic\m}, $\eta_c=$ \SI{e23}{\pascal\s}, a thermal
@@
 \end{figure}
-
-\subsection{Mantle melting curves}\label{sec:katz}
-This instantaneous experiment is performed in order to recreate isobaric melting curves, as function of temperature, bulk water and content of water in the
-melt, to compare with melting curves obtained by \citet{Katz2003}. The domain is square with $L_x=L_y=100$ km, a grid resolution of \num{100x100} elements and
-25 markers per element. Density is set in the entire domain to \SI{3300}{\kg\per\cubic\m} and temperature increases from the sides to the centre from 1000° to
-1500°C. The experiment is performed for bulk water contents from 0 to 0.3 wt.\%.
-
-Fig. \ref{fig:fraction} shows isobaric melting curves for \SI{1}{\giga\pascal} (panel a) and \SI{3}{\giga\pascal} (panel b) calculated for different bulk water
-contents. The curves well match with those obtained for the mantle by \citet{Katz2003}, showing the saturation at the solidus for bulk water of 0.2 and 0.3 wt.\%, indicating
-that the code correctly determines melt fraction as function of pressure, temperature and bulk water content. Same results can be observed in Fig.
-\ref{fig:water}a, where isobaric and isothermal melting curves are shown as function of bulk water content. In addition, Fig. \ref{fig:water}b shows that the
-code correctly determines also the percentage of water in the melt as function of pressure, temperature and melt fraction. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Melting}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Fraction.pdf}
-\caption{Isobaric melting curves for 1 GPa (panel a) and 3 GPa (panel b) as function of temperature with different bulk water contents.}
-\label{fig:fraction}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Water.pdf}
-\caption{Isobaric melting curves for 1.5 GPa as function of bulk water (panel a) and water in the melt (panel b) calculated for different temperatures.}
-\label{fig:water}
-\end{figure}
-
-\clearpage
-
-\bibliography{export}
-
-\end{document}

### 2022-09-07 tex/Benchmark/rayleigh.tex
update main

--- 2022-03-31 FALCON_description.tex
@@
-\documentclass[hidelinks,10pt,a4paper]{article}
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{textcomp}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage{booktabs}
-\usepackage{multirow}
-\usepackage[cm]{fullpage}
-\usepackage[round]{natbib}
-\bibliographystyle{abbrvnat}
-\usepackage{doi}
-\usepackage{bm}
-\usepackage[table]{xcolor}
-\hypersetup{
-    colorlinks,
-    linkcolor={blue},
-    citecolor={blue},
-    urlcolor={blue}
-}
-\usepackage{float}
-\restylefloat{table}
-\usepackage{siunitx}
-\DeclareSIUnit\year{yr}
-
-\begin{document}
-
-\section{Introduction}\label{sec:intro}
-This document includes a description of the code FALCON, with a presentation of the features implemented followed by results of extensive benchmarks listed in
-Table \ref{tab:list}, performed to verify that the code solves correctly a wide range of problems. Description and results of all benchmarks are shown
-in Section \ref{sec:benchmark}.
-
-\begin{table}[H]
-\caption{List of benchmarks performed for each feature of the code. Data of all benchmarks can be found at 
-\url{https://github.com/aleregorda/Benchmarks}}
-\centering
-\begin{tabular}{| l | l | l |}
-\toprule
- \textbf{Feature}  & \textbf{Benchmark} & \textbf{Section} \\
-\midrule
-  Solver  & Stokes flow & \ref{sec:stokes} \\
-\hline
-  \multirow{2}{*}{Markers advection} & Zalesak disk & \ref{sec:runge} \\
-    & Conservative Interpolation Velocity & \ref{sec:CVI} \\
-\hline
-  \multirow{4}{*}{Momentum equation}  & Poiseuille flow & \ref{sec:poiseuille}\\
-    & Instantaneous 2D sphere & \ref{sec:ist_sphere} \\
-    & Rayleigh-Taylor experiment & \ref{sec:rayleigh} \\
-    & Falling block & \ref{sec:block} \\
-\hline
-    & 2D Stokes sphere & \ref{sec:sphere} \\
-    Sticky air and  & Stabilisation algorithm & \ref{sec:stab} \\
-    free surface & Topography relaxation & \ref{sec:crameri} \\
-  & Spontaneous subduction & \ref{sec:subduction} \\
-\hline
-  Erosion and sedimentation  & - & - \\
-\hline
-   Non-linear  & Slab detachment & \ref{sec:slab} \\
-   visco-plasticity & Indenter & \ref{sec:indenter} \\
-    & Brick & \ref{sec:brick} \\
-\hline
-  \multirow{3}{*}{Energy equation}  & Advection stabilisation & \ref{sec:advection} \\
-    & Simple shear heating & \ref{sec:simple_shear} \\
-    & Shear and adiabatic heating & \ref{sec:shear} \\
-\hline
-\multirow{3}{*}{Energy + momentum} & Mantle convection & \ref{sec:mantle} \\
-& Viscoplastic mantle convection & \ref{sec:visco_mantle} \\
-& Thin layer entrainment & \ref{sec:thin} \\
-\hline
-  Phase changes  & \multirow{2}{*}{Hydrated sinking cylinder} & \multirow{2}{*}{\ref{sec:quinquis}} \\
-  and hydration  &  & \\
-\hline
-  Melting & Experimental melting curves & \ref{sec:katz} \\
-\bottomrule
-\end{tabular}
-\label{tab:list}
-\end{table}
-
-The code is written in Fortran90 and uses the finite element method (FEM) with quadrilateral $Q_1 \times P_0$ elements (continuous bilinear velocity and
-discontinuous constant pressure), associated with MUMPS\footnote{\url{http://mumps.enseeiht.fr/}} \citep{Amestoy2001,Amestoy2019} that is a software package
-for solving systems of linear equations of the form $A \cdot x = b$, where \textit{A} is a square sparse matrix, by means of a direct method. Since
-$Q_1 \times P_0$ elements do not satisfy Ladyzhenskaya, Babuska and Brezzi (LBB) stability condition \citep{Donea2003}, elemental pressure is elaborated
-in post-processing to avoid spurious pressures by means of a double interpolation to smooth the pressure. Following this procedure, the elemental pressure
-is interpolated onto nodes and then back onto elements. After the smoothing procedure, the pressure field is calculated again on the nodes to be used in
-combination with the lithostatic pressure, which is calculated onto nodes. Correctness of the solution and performances of the solver in terms of time and
-memory usage are tested solving the Stokes flow with the analytical solution proposed by \citet{Donea2003} (Section \ref{sec:stokes}).
-
-The thermo-mechanics of crust-mantle systems is described by means of the conservation of mass, momentum and energy equations, expressed as follows:
-\begin{eqnarray}
-\label{eq:mass}\frac{\partial \rho}{\partial t}+\bm{\nabla} \cdot (\rho \bm{u})&=&0\\
-\label{eq:momentum}-\bm{\nabla} p + \bm{\nabla} \cdot \bm{\tau} + \rho \bm{g}&=&0\\
-\label{eq:energy}\rho C_p \left(\frac{\partial T}{\partial t} + \bm{u} \cdot \bm{\nabla} T \right)&=&\bm{\nabla} \cdot \left(k\bm{\nabla} T\right) + H_{tot}
-\end{eqnarray}
-where $\rho$ is the density, $\bm{u}$ is the velocity, \textit{p} is the pressure, $\bm{\tau}$ is the deviatoric stress, $\bm{g}$ is the gravity acceleration,
-$C_p$ is the specific heat at constant pressure, \textit{T} is the temperature, \textit{k} is the thermal conductivity and $H_{tot}$ is the total internal heat
-production. Density variations due to temperature are generally small enough to assume the density as constant ($\rho=\rho_0$) in Eq. \ref{eq:mass} and in
-Eq. \ref{eq:energy}, while it must be treated as a variable in the buoyancy term of Eq. \ref{eq:momentum}, such that \[\rho=\rho_0(1-\alpha(T-T_0))\]
-where $\rho_0$ is the density at a reference temperature $T_0$ and $\alpha$ is the coefficient of thermal expansion. Eq. \ref{eq:mass} then can be rewritten as
-\begin{equation}\label{eq:mass_0}
-\bm{\nabla} \cdot \bm{u}=0
-\end{equation}
-This simplification is known as the extended Boussinesq approximation.
-
-The time step $dt$ is chosen in according to Courant-Friedrichs-Lewy condition \citep{Anderson1995}, such that
-\[dt=C_n\; \min\left(\frac{\Delta x}{\max|\bm{u}|},\frac{(\Delta x)^2}{\kappa}\right)\]
-where $0<C_n<1$ is the Courant (CFL) number, $\Delta x$ is the minimum dimension of smallest element, $\max|\bm{u}|$ is the maximum velocity calculated on the
-entire domain and
-$\kappa=\frac{k}{\rho C_p}$ is the thermal diffusivity \citep{Thieulot2014}.
-
-\section{Lagrangian markers}\label{sec:markers}
-Elemental properties (density, viscosity, thermal conductivity, specific heat and thermal expansion) needed to solve Eqs. \ref{eq:mass_0}, \ref{eq:momentum}
-and \ref{eq:energy} are related to the composition of each element, determined by means of Lagrangian markers that are characteristic of different materials
-in the domain. Elemental properties (with the sole exception of the viscosity) are calculated using an arithmetic mean, as \[P_e=\frac{1}{n}\sum_{i=1}^n P_i\]
-where $P_e$ is the elemental property, $P_i$ is the property characteristic of the material of each marker and $n$ is the number of markers of the element.
-Differently, the average scheme for the elemental viscosity can be chosen between harmonic, geometric and arithmetic mean. Markers can be placed regularly or
-randomly at the beginning of the simulation and their advection is performed by either a 2nd-order or a 4th-order Runge-Kutta in space, interpolating the velocity 
-field on each marker by means of the shape functions (see Section \ref{sec:runge}). The interpolated velocity is then corrected by means of the Conservative
-Velocity Interpolation (CVI), which introduces a corrective term to reduce dispersion and clustering of particles in both steady state and time-dependent 
-\citep{Wang2015} (see Section \ref{sec:CVI}).
-
-The initial distribution of the markers is created using the open-source code library
-Geodynamic World Builder \citep{Fraters2019}. During the simulation, each marker carries memory of temperature, pressure and accumulated strain, which are
-determined interpolating the relative nodal parameter. In particular, pressure is interpolated onto each marker after the smoothing procedure. The number
-of markers contained in each element is maintained between $n_{min}$ and $n_{max}$. When in an element there are less markers than $n_{min}$ the code adds
-random markers to reach the $n_{min}$, while if the number is higher than $n_{max}$ some of them are randomly deleted. When new markers are added, they assume
-the properties of the nearest marker. In this way elements are never empty and maintain a number of markers inside a prefixed range.
-
-\section{The momentum equation}\label{sec:mom_eq}
-The penalty method is implemented so that Eq. \ref{eq:mass_0} can be written as:
-\begin{equation}\label{eq:mass_penalty}
-\bm{\nabla} \cdot \bm{u} + \frac{p}{\lambda}=0
-\end{equation}
-where $\lambda$ is the so-called penalty parameter, which should be 6-7 orders of magnitude larger than the shear viscosity to ensure that mass conservation
-is satisfied \citep{Donea2003,Thieulot2014}.
-
-The deviatoric stress tensor in Eq. \ref{eq:momentum} can be written in terms of the strain rate tensor as $\bm{\tau}=2\eta\dot{\bm{\epsilon}}$, with
-$\dot{\bm{\epsilon}}=\frac{1}{2}\left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)$. Therefore, Eq. \ref{eq:momentum} can be rewritten as:
-\begin{equation}\label{eq:momentum_pressure}
--\bm{\nabla} p + \bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right) + \rho \bm{g} = 0
-\end{equation}
-Finally, using pressure from Eq. \ref{eq:mass_penalty}, Eq. \ref{eq:momentum_pressure} can be rewritten as:
-\begin{equation}\label{eq:momentum_penalty}
-\lambda \bm{\nabla} \left(\bm{\nabla} \cdot \bm{v}\right)+\bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right)+\rho\bm{g}=0
-\end{equation}
-The penalty method is been associated to the iterative Uzawa method, as described in detail in \citet{Dabrowski2008} and \citet{Thieulot2014}
-(see Section \ref{sec:poiseuille}). Results of benchmarks performed to verify the correctness in the implementation of Eq. \ref{eq:momentum_penalty} are
-shown in Sections \ref{sec:ist_sphere} and \ref{sec:rayleigh}.
-
-To support large viscosities variations the penalty parameter is related to the effective elemental viscosity by means of a dimensionless coefficient, so that
-$\lambda=\lambda_e(e)\eta_{eff}(e)$ \citep{Marotta2006,Dabrowski2008,Thieulot2014}. Benchmark of the falling block \citep{Gerya2003a,Gerya2010b,Thieulot2011}
-is performed to verify that the code can correctly deal buoyancy driven flows with strong viscosity contrasts (Section \ref{sec:block}).
-
-\section{Sticky air and free surface}\label{sec:sticky}
-The Earth's surface can be treated by means of either the so-called sticky-air or a true free surface method, both of them implemented in the code. In the
-sticky-air method the surface is approximated with the introduction of a buoyant layer with a viscosity at least four orders of magnitude lower than the crust
-\citep{Schmeling2008,Crameri2012} and the interface between lithosphere and air is defined using a chain of passive markers that are advected as the Lagrangian
-markers. The correctness of the evolution of the markers chain is tested with the experiment of a 2D time-dependent Stokes sphere below a free surface and
-compared with results from ASPECT\footnote{\url{https://aspect.geodynamics.org/}} \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} (Section
-\ref{sec:sphere}). In the true free surface case the top boundary is assumed stress-free and velocities are not fixed. In this case topography variations are
-described by vertical deformations of the mesh that depend on the velocity field of the nodes that identify the free surface, while horizontal deformations are
-not taken into account. This procedure is known as the Arbitrary Lagrangian-Eulerian (ALE) method and its implementation follows the technique described in
-\citet{Thieulot2011}. However, although the implementation of a true free surface better reproduces laboratory experiment, extremely small time steps can be
-necessary to maintain stability \citep{Kaus2010a,Quinquis2011,Thieulot2014}. Therefore, the stability algorithm proposed by \citet{Kaus2010a} is implemented
-to avoid instabilities due to high density differences at the free surface when using too large time steps. The implementation of this algorithm are tested by
-performing the experiment described by \citet{Kaus2010a} (Section \ref{sec:stab}).
-
-The topography relaxation benchmark proposed by \citet{Crameri2012} is performed to verify that the code correctly recovers topography variations in case of
-both the sticky-air and the true free surface method (Section \ref{sec:crameri}). Finally, Section \ref{sec:subduction} show the results of the spontaneous
-subduction experiment described by \citet{Schmeling2008}.
-
-\section{Erosion and sedimentation}\label{sec:erosion}
-Surface processes at either the lithosphere-air interface or in correspondence of the free surface have been implemented by means of the software
-Fastscape\footnote{\url{https://fastscape.org/}} \citep{Braun2013,Cordonnier2019,Yuan2019a,Yuan2019b}. FastScapeLib is a set of routines that solve the stream
-power law (enriched by a sediment transport/deposition term) \citep{Yuan2019a}, hillslope diffusion and marine transport and deposition \citep{Yuan2019b}.
-The partial differential equation solved by FastScapeLib is:
-\begin{equation}
-\frac{\partial h}{ \partial t}=U-K_f A^{m_e}S^{n_e} + \frac{G}{A}\int_A \left( U-\frac{\partial h}{\partial t} \right) dA + K_d \bm{\nabla}^2 h 
-\end{equation}
-where $h$ is the topography, $U$ is the uplift, $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $S$ is the slope, $m_e$ and $n_e$ are
-parameters that should be chosen to constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}, $G$ is the dimensionless
-deposition/transport coefficient for the enriched stream power law and $K_d$ is the effective diffusivity. Terms $K_f A^{m_e}S^{n_e}$ and $K_d \bm{\nabla}^2 h$
-consider long-range processes, described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, and short-range processes,
-proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}.
-
-In case of sticky air, slope is determined by means of the markers chain, which is vertically corrected considering erosion and sedimentation rates. After the
-correction, continental markers above the markers chain are transformed in air markers and air markers below the markers chain become sediments. Similarly,
-slope is determined considering the nodes on the top boundary in case of true free surface and they are vertically corrected as for the markers chain. In this
-case, continental markers above the top boundary are simply deleted.
-%Surface processes at the lithosphere-air interface are implemented taking into account fluvial erosion (long-range processes), hillslope processes (short-range
-%processes) and failure of slopes exceeding a threshold angle $\theta$ \citep{Burov1997,Tucker1999,Sacek2012,Croissant2014,Ueda2015,Beucher2020}. Erosion rates
-%$\dot{E}_{lr}$ due to long-range processes can be described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{lr}=K_f A^{m_e} \bm{\nabla} h^{n_e}
-%\end{equation}
-%where $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $\bm{\nabla} h$ is the slope and $m_e$ and $n_e$ are parameters chosen in to
-%constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}. Erosion rates $\dot{E}_{sr}$ due to short-range processes are
-%implemented by assuming that transport is linearly proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{sr}=K_d \bm{\nabla}^2 h
-%\end{equation}
-%where $K_d$ is the effective diffusivity. In addition, the slope is modified if it still exceeds a threshold angle after long- and short-range processes
-%\citep{Ueda2015}.
-
-%Offshore sedimentation processes are implemented similarly to hillslope processes but with a depth-dependent coefficient \citep{Sacek2012,Beucher2020}.
-%The sedimentation rate $\dot{S}$ is calculated as
-%\begin{equation}
-%\dot{S}=K_s \bm{\nabla}^2 h
-%\end{equation}
-%with
-%\begin{equation}
-%K_s=K^*_s \exp\left(\frac{h-h_{sl}}{h_0}\right)
-%\end{equation}
-%where $K^*_s$ is a constant, $h_{sl}$ is the sea level and $h_0$ is a characteristic length. In case of sticky air, slope is determined by means of the markers
-%chain, which is vertically corrected considering erosion and sedimentation rates. After the correction, continental markers above the markers chain are
-%transformed in air markers and air markers below the markers chain become sediments. Similarly, slope is determined considering the nodes on the top boundary
-%in case of true free surface and they are vertically corrected as for the markers chain. In this case, continental markers above the top boundary are simply
-%deleted.
-
-\section{Non-linear rheologies}\label{sec:nonlinear}
-Non-linear rheologies are implemented combining viscous creep (dislocation and diffusion) and plastic yielding. For each marker, diffusion and dislocation
-viscosity ($\eta_{df|ds}$) can be determined as follows \citep{Karato1993,Warren2008b,Wang2016}:
-\begin{equation}\label{eq:viscous}
-\eta_{df|ds}=f_s \left(\frac{d^m}{A}\right)^{\frac{1}{n}} \left (\frac{2^{\frac{1-n}{n}}}{3^{\frac{n+1}{2n}}} \right)
-I_2^{\frac{(1-n)}{n}}\exp\left(\frac{Q+pV}{nRT}\right)
-\end{equation}
-where $f_s$ is a scaling factor used to represent lithologies that are stronger or weaker than the base set, $d$ is the grain size, $m$ is the grain size exponent,
-$n$ is the stress exponent, $A$ is the uniaxial pre-exponential factor, $I_2$ is the square root of the second invariant of the strain rate tensor, $Q$ is the
-activation energy, $p$ is the pressure, $V$ is the activation volume, $R$ is the gas constant and $T$ is the temperature.
-Pressures and temperatures are determined by the interpolation of the nodal parameters, while strain rates are calculated by means of the
-derivative of the velocities on the nodes. In case of diffusion creep $n=1$ and $m>0$, in case of dislocation creep $n>1$ and $m=0$. The viscous creep $\eta_{cp}$
-is then calculated as the harmonic average between $\eta_{df}$ and $\eta_{ds}$:
-\begin{equation}\label{eq:visco_creep}
-\eta_{cp}=\left(\frac{1}{\eta_{df}}+\frac{1}{\eta_{ds}}\right)^{-1}
-\end{equation}
-The implementation of the non-linear viscous creep viscosity depending on the strain rate is tested by means of the slab detachment benchmark
-(Section \ref{sec:slab}).
-
-Plastic yielding is implemented rescaling $\eta_{cp}$ in order to limit the stress below the yield stress $\sigma_y$ \citep{Thieulot2008,Thieulot2014,Glerum2018},
-obtaining
-\begin{equation}\label{eq:plastic}
-\eta_{pl}=\frac{\sigma_y}{2I_2}
-\end{equation}
-where $\eta_{pl}$ is the plastic viscosity and the yield stress is determined following the Drucker-Prager criterion, such as
-\begin{equation}\label{eq:yield}
-\sigma_y=C\;\cos(\phi)+p\;\sin(\phi)
-\end{equation}
-where $C$ is the cohesion and $\phi$ is the internal friction angle. In case of negative pressure it is imposed equal to 0, so that negative yield stress are
-excluded. The correctness of the non-linear solution in case of plastic viscosity with variable internal friction angle is verified performing indenter and
-brick experiments (Sections \ref{sec:indenter} and \ref{sec:brick}, respectively).
-
-Strain softening is taken into account for both viscous creep and plastic viscosity \citep{Huismans2003,Babeyko2005,Huismans2005,Sobolev2005,Warren2008a} by
-means of the accumulated strain $\epsilon$ memorised by each marker. Softening in the viscous creep determines a linear decrease of $\eta_{vc}$ by means of a
-viscous strain softening factor $W_S$ that increases linearly from $W_{S_0}$ to $W_{S_{\infty}}$ for $\epsilon_{S_0}<\epsilon_S<\epsilon_{S_{\infty}}$ 
-\citep{Warren2008a}. This viscous softening can be related to strain-induced grain size reduction \citep{Warren2008a}. Differently, plastic softening is 
-simulated with a linear decrease of internal friction angle $\phi(\epsilon)$ and cohesion $C(\epsilon)$ in according to
-\begin{eqnarray}
-\label{eq:friction1}\phi(\epsilon)&=&\phi_0+(\phi_{\infty}-\phi_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}\\
-\label{eq:friction2}C(\epsilon)&=&C_0+(C_{\infty}-C_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}
-\end{eqnarray}
-where $\phi_0$, $C_0$ and $\phi_{\infty}$, $C_{\infty}$ are internal friction angle and cohesion for $\epsilon_0$ and $\epsilon_{\infty}$, respectively
-\citep{Huismans2003,Huismans2005,Warren2008a,Thieulot2014}. Plastic softening approximates deformation-induced softening of faults and brittle shear zones
-\citep{Warren2008a}.
-
-Plastic yielding and viscous creep are then combined to obtain a viscoplastic viscosity $\eta_{vp}$ as follows, assuming that they are independent processes
-\citep{Karato2008,Glerum2018}:
-\begin{equation}\label{eq:viscoplastic}
-\eta_{vp}=\min(\eta_{cp},\eta_{pl})
-\end{equation}
-Finally, effective viscosity $\eta_{eff}$ is capped by the minimum and the maximum viscosity ($\eta_{min}$ and $\eta_{max}$, respectively) to avoid extremely
-low or high viscosity \citep{Glerum2018} as follows
-\begin{equation}\label{eq:effective}
-\eta_{eff}=\min(\max(\eta_{vp},\eta_{min}),\eta_{max})
-\end{equation}
-Viscous creep and plastic yielding are non-linear rheologies because of their dependence on the velocity field through pressure and strain rates. Therefore,
-the solution is determined by means of Picard-type iterations, until convergence of the velocity field \citep{Glerum2018}. The convergence is verified at each
-iteration $i$ via the nonlinear residual $\bm{\mathcal{R}}^i$ that can be determined as
-\begin{equation}\label{eq:convergence}
-\bm{\mathcal{R}}^i= \mathbb{K} (\eta_{eff}(\dot{\bm{\epsilon}}^{i-1},p^{i-1}))\cdot \bm{v}^{i-1} - \bm{f}^i
-\end{equation}
-%\begin{equation}\label{eq:convergence}
-%\chi=1-\frac{\langle (\bm{v}^i-\langle\bm{v}^i\rangle)\cdot (\bm{v}^{i+1}-\langle\bm{v}^{i+1}\rangle) \rangle}{\sqrt{\sigma^i\sigma^{i+1}}}
-%\end{equation}
-%Iterations are performed until either convergence ($\chi<tol$) or maximal number of non-linear iterations ($it_{max}$) is reached.
-where $\mathbb{K}$ is the Stiffness matrix and $\bm{f}$ is the right hand side vector \citep{Spiegelman2016,Glerum2018}.  The $L_2$-norm is extracted from
-$\bm{\mathcal{R}}^i$ and it is normalised as follows
-\begin{equation}\label{eq:convergence_res}
-\frac{||\bm{\mathcal{R}}^i||_2}{||\bm{\mathcal{R}}^0||_2}
-\end{equation}
-where $||\bm{\mathcal{R}}^0||_2$ is the $L_2$-norm of the first nonlinear iteration.
-Since often the normalised nonlinear residual drop very quickly over the first 2-3 iterations, also the $L_2$-norm of velocity and pressure residuals are
-calculated as
-\begin{equation}\label{eq:convergence_vel}
-\frac{||\Delta\bm{v}||_2}{||\bm{v}||_2}=\frac{||\bm{v}_i-\bm{v}_{i-1}||_2}{||\bm{v}_{i}||_2}
-\end{equation}
-and
-\begin{equation}\label{eq:convergence_press}
-  \frac{||\Delta p||_2}{||p||_2}=\frac{||p_i-p_{i-1}||_2}{||p_{i}||_2}
-\end{equation}
-respectively. Iterations are performed until either maximal number of non-linear iterations ($it_{max}$) is reached or all the normalised nonlinear residual, 
-the velocity residual and the pressure residual converge under a defined tolerance $tol$.
-
-\section{The energy equation}\label{sec:en_eq}
-The stabilisation of the advection term of the energy equation (Eq. \ref{eq:energy}) needed to avoid possible oscillations in the thermal solution in the case
-when advection dominates over diffusion, is implemented by means of a streamline-upwind Petrov–Galerkin (SUPG) method, for which the advection term is modified
-follows the discussion in \citet{Thieulot2011} and \citet{Thieulot2014} (see Section \ref{sec:advection} for benchmark). The total internal heat production
-$H_{tot}$ in Eq. \ref{eq:energy} includes radiogenic heating $H_r$, shear heating $H_s$ and adiabatic heating $H_a$, where
-\begin{equation}\label{eq:shear}
-H_s=2\eta \dot{\bm{\epsilon}} : \dot{\bm{\epsilon}} = 2\eta(\dot{\epsilon}_{xx}^2+\dot{\epsilon}_{yy}^2+2\dot{\epsilon}_{xy}^2)
-\end{equation}
-and
-\begin{equation}\label{eq:adiabatic}
-H_a=T \alpha \frac{Dp}{Dt} \approx -\alpha T \rho g_y v
-\end{equation}
-The correctness of the implementation and the computation of shear and adiabatic heating is verified performing an exercise with an analytical solution
-(Section \ref{sec:simple_shear}) and Exercise 9.4 in \citet{Gerya2010b} (Section \ref{sec:shear}). Furthermore, mantle convection and thin layer entrainment
-benchmarks (Sections \ref{sec:mantle} and \ref{sec:thin}, respectively) are performed to verify that the code solve correctly Eqs. \ref{eq:momentum_penalty}
-and \ref{eq:energy} in case of a temperature field characterised by an initial perturbation and with a temperature-dependent density. At last, the viscoplastic
-mantle convection benchmark is performed to check also the viscoplasticity in case of a initial perturbed temperature field (Section \ref{sec:visco_mantle}).
-
-\section{Phase transitions and hydration}\label{sec:phase}
-Variations of effective density, specific heat and coefficient of thermal expansion during the evolution of crustal and mantle materials at different
-pressure-temperature ($p$-$T$) conditions are computed by means of version 6.8.6 of Perple\_X software package \citep{Connolly2005}, similarly to the
-implementation described in \citet{Marotta2020}. By default, mineral assemblages and properties of each lithology (oceanic crust, lower and upper continental
-crust, sediments and dry or hydrated mantle) are calculated for temperatures from \SIrange{330}{1600}{\kelvin}, with increments of approximately \SI{4}{\kelvin},
-and pressures from \SIrange{0.1}{30}{\giga\pascal}, with increments of approximately \SI{0.03}{\giga\pascal}, for a total of almost 100000 points.
-
-The introduction of phase transitions produces two main effects that must be taken into account in numerical models: variations in density and release/absorption
-of latent heat \citep{Tackley2010}. Effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into
-account considering effective coefficient of thermal expansion \citep{Christensen1985,Zhong2015}. Similarly, effects of latent heat in the energy equation
-(eq \ref{eq:energy}) are taken into account by considering effective specific heat and coefficient of thermal expansion
-\citep{Christensen1985,Tackley2010,Zhong2015}. Effective density, specific heat and coefficient of thermal expansion are included in Perple\_X files and are
-assumed by each Lagrangian marker in according to its $p$-$T$ conditions.
-
-Hydration processes related to dehydration of subducting lithosphere \citep{Schmidt1998,Liu2007,Faccenda2009,Faccenda2010,Faccenda2014,Rosas2016} strongly
-influence the thermo-mechanics inside the mantle wedge, mainly because of both weakening effects on the mantle rheology and density variations in case of
-mantle serpentinization \citep{Gerya2002,Honda2003,Arcay2005,Roda2010,Regorda2017}. In the case that hydration is switched on, the amount of bound and free
-water is memorised by each marker, following the implementation of \citet{Quinquis2014}; hydration and dehydration processes are related to the amount of bound
-water of each marker and to the maximum amount of water it can transport, i.e. if the amount of bound water exceeds the maximum amount of water, the marker
-dehydrates and releases free water that can hydrates under-saturated markers. Maximum water content of each marker is determined as function of lithology and
-$p$-$T$ conditions and it is calculated using Perple\_X, in the same way than effective density, specific heat and coefficient of thermal expansion. Bound water
-is advected together with the markers, neglecting the effect of bound water diffusion, while free water simply migrates vertically and is not coupled to the
-solid-phase flow of the mantle wedge \citep{Arcay2005,Quinquis2014}. The correctness of the migration scheme of the free water and its absorption into bound
-water by under-saturated markers are verified performing the experiment described in \citet{Quinquis2014} (Sec. \ref{sec:quinquis}).
-
-The mantle viscosity weakening related to the amount of water has been extensively studied \citep{Chopra1981,Hirth2003} and it is implemented in according to
-\citet{Arcay2005} and \citet{Horiuchi2016}, as follows
-\begin{equation}\label{eq:wet_visc}
-\eta_{wet}=\eta_{dry}\left(\left[1-\frac{1}{f_v}\right]\exp\left(-\frac{[OH^-]}{[OH^-]_0}\right)+\frac{1}{f_v}\right)
-\end{equation}
-where $f_v$ is the viscosity reduction factor between dry and wet conditions, $[OH^-]$ is the water content and $[OH^-]_0$ is a reference water content set to
-620 ppm (0.062 wt.\%) \citep{Arcay2005}. By default, $f_v$ is set to 100, in according to \citet{Arcay2005} and \citet{Horiuchi2016}.
-
-\section{Mantle melting}\label{sec:melt}
-Melt fraction in the mantle depends on temperature ($T$, in \textdegree C), pressure ($p$, in \si{\giga\pascal}) and the water content in the melt ($X_{H_2O}$, in wt.\%).
-The determination of the melt fraction $M$ is implemented as explained by \citet{Katz2003}, according to the successive modification by \citet{Langmuir2006} and
-\citet{Kelley2010}, as follows
-\begin{eqnarray}\label{eq:melt_fr}
-M(p,T,X_{H_2O})&=&-T+T_{sol}+(x_t\cdot ln(p+y_t))\cdot X_{H_2O}-\nonumber\\
-&&+K_m\left(\frac{X_{H_2O}^{bulk}}{D_{H_2O}(1-X_{H_2O})+X_{H_2O}}\right)^{\gamma}
-\end{eqnarray}
-where $T_{sol}$ is the temperature of the dry solidus, $D_{H_2O}$ is the partition coefficient and $x_t\cdot log(p+y_t)$ indicates the pressure dependence of
-the melting curve \citep{Kelley2010}. $x_t$, $y_t$, $K_m$ and $\gamma$ are parameters chosen in according to \citet{Kelley2010}.
-However, $X_{H_2O}$ depends on the melt fraction as
-\begin{equation}\label{eq:melt_wt}
-X_{H_2O}(M)=\frac{X_{H_2O}^{bulk}}{D_{H_2O}+M(1-D_{H_2O})}
-\end{equation}
-and a numerical solution can be found using a root-finder method \citep{Katz2003,Wang2016}. Furthermore, the content of water in the melt is limited by the
-pressure-dependent saturation concentration of water in the melt, determined as
-\begin{equation}\label{eq:melt_max}
-X_{H_2O}^{sat}=\chi_1 p^{\lambda_m}+\chi_2 p
-\end{equation}
-where $\chi_1$, $\chi_2$ and $\lambda_m$ are parameters chosen according to \citet{Katz2003}, \citet{Langmuir2006} and \citet{Kelley2010}. The correctness in
-the determination of melt fractions and water contents in the melt are tested comparing model results with experimental curves obtained by \citet{Katz2003}
-(Sec. \ref{sec:katz}).
-
-In case of melting, viscous creep viscosity ($\eta_{vp}$) is modified according to \citet{Wang2016}, as
-\begin{equation}\label{eq:melt_visc}
-\eta_{melt}=\eta_{vp}\cdot \exp(\alpha_m M)
-\end{equation}
-where $\alpha_m$ are melt fraction factors for dislocation and diffusion creep, chosen in according to \citet{Wang2016}.
-
-Effective density ($\rho_{eff}$) for partially molten rocks is calculated as
-\begin{equation}\label{eq:melt_dens}
-\rho_{eff}=\rho_s(1-M)+\rho_m M
-\end{equation}
-where $\rho_s$ and $\rho_m$ are the densities of the solid and the molten rock, respectively \citep{Gerya2004b,Gerya2007,Wang2016}. The density of the
-solid rock is extracted by Perple\_X, while the density of the molten portion is calculated as
-\begin{equation}\label{eq:melt_dens_M}
-\rho_m(p,T)=\rho_0[1-\alpha(T-T_0)][1+\beta(p-p_0)]
-\end{equation}
-where $\rho_0$ is the density at the reference temperature ($T_0$) and pressure($p_0$), and $\alpha$ and $\beta$ are the thermal expansion and compressibility
-coefficients, respectively \citep{Gerya2004b,Gerya2007,Wang2016}.
-
-As for the phase transitions, effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into account
-considering effective coefficient of thermal expansion. Similarly, latent heat due to melting/crystallisation can be implicitly included in the energy equation
-(Eq. \ref{eq:energy}) by considering effective specific heat ($C_{p_{eff}}$) and thermal expansion ($\alpha_{eff}$) of the partially molten rocks, in according to
-\begin{eqnarray}
-\label{eq:cp_effective}C_{p_{eff}}&=&C_p+H_L\left(\frac{\partial M}{\partial T}\right)_p \\
-\label{eq:alpha_effective}\alpha_{eff}&=&\alpha+\rho\frac{H_L}{T}\left(\frac{\partial M}{\partial p}\right)_T
-\end{eqnarray}
-where $C_p$ and $\alpha$ are the specific heat and the coefficient of thermal expansion of the solid rock, respectively, and $H_L$ is the latent heat of the
-molten rock \citep{Gerya2004b,Gerya2007,Tackley2010}.
-
-\section{Benchmarks}\label{sec:benchmark}
-\subsection{Error measurements}\label{sec:error}
-In order to determine the accuracy of velocity and pressure field of the benchmarks, the $L_2$-norm is computed by numerical integration on the Gauss-Legendre
-quadrature points. $L_2$-norm for pressure and velocity errors can be evaluated as
-\begin{eqnarray}
-\textrm{err}_p&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|p^n_i(r_q)-p^a_i(r_q)|^2\right)w_q|J_q|}\nonumber \\
-\textrm{err}_v&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|u^n_i(r_q)-u^a_i(r_q)|^2+|v^n_i(r_q)-v^a_i(r_q)|^2\right)w_q|J_q|}\nonumber
-\end{eqnarray}
-respectively, where $n_e$ is the number of elements, $n_q$ is the number of quadrature points per element, $p^n_i(r_q)$ and $p^a_i(r_q)$ are the numerical and
-analytical pressure, respectively, in each quadrature point $q$, $w_q$  and $J_q$ are the weight and the Jacobian at the quadrature point
-$q$, $u^n_i(r_q)$, $v^n_i(r_q)$, $u^a_i(r_q)$ and $v^a_i(r_q)$ are the numerical and analytical velocities, respectively, in each quadrature point $q$.
-
-Other quantities used as comparison with original benchmarks are the root-mean-square velocity over the whole domain and over the surface
-\begin{eqnarray}
-v_{\textrm{rms}}&=&\sqrt{\int_{\Omega}|\bm{v}|^2 d\Omega}\nonumber \\
-v_{\textrm{rms}}^{top}&=&\sqrt{\int_0^1 u^2 \bigg|_{y=1} dx}\nonumber
-\end{eqnarray}
-respectively, top and bottom Nusselt numbers
-\[\textrm{Nu}^{top/bottom}=-\int_0^1\frac{\partial T}{\partial y} \bigg|_{y=1/y=0} dx\]
-the average rate of work done against gravity
-\[\left\langle W \right\rangle=\int_{\Omega}T u_y d\Omega\]
-the average rate of viscous dissipation
-\[\left\langle \Phi \right\rangle=\int_{\Omega}\tau_{ij}\dot{\epsilon}_{ij} d\Omega\]
-and the percentage error between $\left\langle W \right\rangle$ and $\left\langle \Phi \right\rangle/Ra$
-\[\delta=\frac{\left|\left\langle W \right\rangle -\frac{\left\langle \Phi \right\rangle}{Ra} \right|}{\max \left(\left\langle W \right\rangle,
-\frac{\left\langle \Phi \right\rangle}{Ra}\right)}\times100\]
-
-\subsection{Stokes flow}\label{sec:stokes}
-The problem consists of determining velocity field $(u,v)$ and pressure $p$ in case of a manufactured solution with prescribed body forces such as
-\begin{eqnarray}
-b_1&=&(12-24y)x^4+(-24+48y)x^3+(-48y+72y^2-48y^3+12)x^2+\nonumber \\
-&&+(-2+24y-72y^2+48y^3)x+1-4y+12y^2-8y^3\nonumber \\
-b_2&=&(8-48y+48y^2)x^3+(-12+72y-72y^2)x^2+(4-24y+48y^2-48y^3+\nonumber \\
-&&+24y^4)x-12y^2+24y^3-12y^4\nonumber
-\end{eqnarray}
-for which the exact solution is:
-\begin{eqnarray}
-u(x,y)&=&x^2(1-x)^2(2y-6y^2+4y^3)\nonumber \\
-v(x,y)&=&-y^2(1-y)^2(2x-6x^2+4x^3)\nonumber \\
-p(x,y)&=&x(1-x)-\frac{1}{6}\nonumber
-\end{eqnarray}
-The domain is a unit square a constant viscosity ($\eta=1$) and the penalty parameter is set to $\lambda=10^7$. Velocity boundary conditions are set to no slip
-$(\bm{v}=\bm{0})$ on all boundaries. The problem is performed for different grid resolution between \num{8x8} and \num{1024x1024} elements.
-
-The errors between the analytical solution and the numerical prediction of the velocity and the pressure field are calculated as in Section \ref{sec:error}.
-Fig. \ref{fig:errors}a shows that both velocity and pressure field converge to the exact solution with the decrease of the element size, following the
-theoretical convergence rate. The convergence of pressure, in contrast with the observation by \citet{Donea2003} for $Q_1 \times P_0$ elements, indicate the
-effectiveness of the smoothing procedure. Smoothed pressure field for grid resolution of \num{128x128} elements is shown in Fig. \ref{fig:errors}b (continuous
-red line), in comparison with the analytical solution (dashed blue line). Solve times and memory usage needed to generate the solution are shown in
-Fig. \ref{fig:MUMPS}. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Solver/Stokes_Flow}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/errors.pdf}
-\caption{Velocity and pressure error for the Stokes flow experiment between generated and analytical solution as a function of element size (panel a),
-and comparison between smoothed pressure and analytical solution as function of $x$ coordinate for a grid resolution of \num{128x128} elements (panel b).}
-\label{fig:errors}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/MUMPS.pdf}
-\caption{Panel a) Analysis, factorisation, solution and total solve times (panel a) and factorisation memory usage (panel b) as a function of the total number
-of degrees of freedom for the Stokes flow experiment.}
-\label{fig:MUMPS}
-\end{figure}
-
-\subsection{Markers advection}\label{sec:runge}
-The 2nd-order Runge-Kutta advection scheme in space is tested by means of the Zalesak disk test \citep{Zalesak1979,Thieulot2014}. The benchmark is performed
-in a unit square domain with a grid resolution of \num{32x32} elements and 1000000 markers and values of Courant number between 0.25 and 1.
-At $t=0$, the disk is centred at position $(0.5;0.75)$ with a radius $R=0.15$ and has a vertical fissure 0.05 wide and 0.2 high (orange in Fig. \ref{fig:runge})
-and with the velocity field prescribed in the entire domain as
-\begin{eqnarray}
-u(x,y)&=&2\pi\left(y-\frac{L_x}{2}\right)\nonumber \\
-v(x,y)&=&-2\pi\left(x-\frac{L_x}{2}\right)\nonumber
-\end{eqnarray}
-the markers are back to their initial location after a $2\pi$ rotation ($t=1$). The clockwise rotation of the disk at $t=0.25$, 0.5 and 0.75 
-is shown in Fig. \ref{fig:runge} (red, blue and green, respectively). The distance from the centre of 1 marker is calculated throughout the rotation to
-evaluate the error for different values of Courant number. As expected, the error increase with the increasing of the time step, as shown in 
-Fig. \ref{fig:runge_err}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/Zalesak_Disk}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=200px]{./Figures/Zalesak.pdf}
-\caption{Position of the Zalesak disk at $t=0$, 0.25, 0.5 and 0.75 (orange, red, blue and green, respectively) for the case with Courant number of 1.}
-\label{fig:runge}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/zal_marker.pdf}
-\caption{Distance from the centre as function of time for values of Courant number of  of 0.25, 0.3, 0.5, 0.75 and 1 (green, orange, blue, black and red,
-respectively). R indicates the distance from the centre considering a perfect circle trajectory.}
-\label{fig:runge_err}
-\end{figure}
-
-\subsection{Conservative Velocity Interpolation (CVI)}\label{sec:CVI}
-The CVI correction is checked by means of the Stokes flow experiment by \citet{Donea2003}, as presented in \ref{sec:stokes}. The advection of Lagrangian
-markers is performed using either a 2nd-order or a 4th-order Runge-Kutta scheme with CFL=0.5 and an initial random distribution of 25 markers per element.
-Fig. \ref{fig:CVI} shows the minimum and maximum values of markers per element throughout the experiment with and without the CVI correction and using
-different Runge-Kutta schemes. The introduction of the CVI correction allows the
-tests to run for times longer than $t=2000$ without show empty elements, while tests without the CVI correction stop at approximately $t=600$, when at least
-one element goes empty. Fig. \ref{fig:CVI_time} shows the different number of markers per element and the markers distribution at $t=600$ obtained without and
-with the CVI correction (panels b and c, respectively) in case of a 2nd-order Runge-Kutta scheme. All data can be found at
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/CVI}.
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/CVI.pdf}
-\caption{Maximum and minimum number of markers per element throughout the experiment, with (green and orange lines) or without (purple and blue lines)
-the CVI correction and using either a 2nd-order (blue and orange lines) or a 4th-order (purple and green lines) Runge-Kutta scheme.}
-\label{fig:CVI}
-\end{figure}
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/CVI_time.pdf}
-\caption{Number of markers per element and markers distribution through time using 2nd.order Runge-Kutta scheme.
-Panels a: number of markers per element (panel a1) and distribution of the markers (panel a2) after the first
-iteration; panels b and c: comparison between number of markers per element (panels b1 and c1) and markers distribution (panels b2 and c2) without and with
-the CVI correction at $t=600$; panels d: number of markers per element (panel d1) and markers distribution (panel d2) in case of CVI correction at $t=2000$.}
-\label{fig:CVI_time}
-\end{figure}
-
-\subsection{Poiseuille flow}\label{sec:poiseuille}
-The domain is a rectangle with $L_x=2$ and $L_y=1$ and constant density and viscosity ($\rho=1$ and $\eta=1$), gravity acceleration $\bm{g}=\bm{0}$ and penalty
-parameter $\lambda=10^8$. The grid is composed by \num{40x20} elements. Velocity boundary conditions are set to no slip $(\bm{v}=\bm{0})$ at the top and the
-bottom, and a parabolic profile is imposed on the sides, with $u=y(1-y)$ and $v=0$. The analytical solution is then given by:
-\begin{eqnarray}
-u(x,y)&=&y(1-y)\nonumber \\
-v(x,y,)&=&0\nonumber \\
-p(x,y)&=&2\eta\left(\frac{L_x}{2}-x\right)\nonumber
-\end{eqnarray}
-
-The velocity field predicted by the model follows the expected parabolic profile (Fig. \ref{fig:poiseuille}a and Fig. \ref{fig:poi_plot}a) and, in case of the
-classic penalty method with no iterations, the pressure is clearly related to the divergence by means of the penalty parameter (Fig. \ref{fig:poiseuille}b and
-c and Fig. \ref{fig:poi_plot}b). Fig. \ref{fig:poiseuille}d shows that one Uzawa iteration is sufficient to bring the divergence down to \num{e-15}, with no
-correlation with the pressure field. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Poiseuille%20Flow}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Poiseuille.pdf}
-\caption{Velocity field (panel a), pressure (panel b) and divergence velocity of a Poiseuille flow in case of the classic penalty method (no iterations) and
-after one Uzawa iteration (panel c and d, respectively).}
-\label{fig:poiseuille}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Analytical.pdf}
-\caption{Velocity (panel a) and pressure (panel b) field predicted by the model for a Poiseuille flow with respect with their analytical solutions. The
-velocity field is plotted as function of the vertical coordinate in $L_x/2$ and the pressure is plotted as function of the x coordinate in $L_y/2$.}
-\label{fig:poi_plot}
-\end{figure}
-
-\subsection{Instantaneous 2D sphere}\label{sec:ist_sphere}
-The domain is a unit square with gravity $\bm{g}=(0,-1)$. The fluid has constant density and viscosity ($\rho_f=1$ and $\eta_f=1$). The sphere is in the middle
-of the domain with a radius R = 0.123456798 and has constant density and viscosity ($\rho_s=10^{-2}+\rho_f$ and $\eta_s=10^3 \cdot \eta_f$).
-We distinguish three different types of velocity boundary conditions:
-\begin{enumerate}
-\item FS: free slip conditions on all sides.
-\item NS: no slip conditions on all sides.
-\item OT: free slip conditions on the sides and the bottom, and open boundary at the top.
-\end{enumerate}
-All three cases are performed for different grid resolutions, between \num{16x16} and \num{512x512} elements with 50 randomly distributed markers per element,
-and different average schemes for the viscosity (harmonic, geometric and arithmetic). The velocity in the centre of the sphere $\bm{v}(0.5;0.5)$), minimum and
-maximum velocities ($u_{min}$, $u_{max}$, $v_{min}$, $v_{max}$) and pressures ($p_{min}$, $p_{max}$) on the entire domain, $v_{\textrm{rms}}$ and average pressure on the
-entire domain ($p_{avg}$) are compared with solutions generated by ASPECT \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and all results
-can be found at \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere2D} and 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Instantaneous_2D_Sphere}. Results in terms of the $v_{\textrm{rms}}$ for NS model 
-are also shown in Fig. \ref{fig:inst_sphere}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_NS.pdf}
-\caption{$v_{\textrm{rms}}$ as function of element size for the instantaneous 2D sphere experiment in case of no slip boundary conditions and with different average
-schemes for the viscosity. Results are compared with results obtained by other numerical codes.}
-\label{fig:inst_sphere}
-\end{figure}
-
 \subsection{Rayleigh-Taylor instability}\label{sec:rayleigh}
@@
 with the lighter fluid at the bottom. The initial interface between the fluids is given by $y(x)=0.2+0.02 \cos \left(\frac{\pi x}{L_x}\right)$. The experiment
-is performed with different grid sizes (\num{50x50}, \num{80x80}, \num{100x100} and \num{256x256}). A total of 1960000 markers are randomly distributed at the
+is performed with different grid sizes ($50\times50, 80\times80, 100\times100$ and $256\times256$). A total of 1960000 markers are randomly distributed at the
 beginning of the simulation. Velocity boundary conditions are set to no slip at the top and the bottom, and to free slip at the sides of the domain.
@@
 \centering
-\noindent\includegraphics[width=300px]{./Figures/Rayleigh.pdf}
-\caption{Evolution of the Rayleigh-Taylor experiment for a grid of \num{256x256} elements at $t=0, 500, 1000$ and $2000$ (panels a, b, c and d, respectively).}
+\includegraphics[width=300px]{./Figures/Rayleigh.pdf}
+\caption{Evolution of the Rayleigh-Taylor experiment for a grid of $256\times256$ elements at $t=0, 500, 1000$ and $2000$ (panels a, b, c and d, respectively).}
 \label{fig:rayleigh}
 \end{figure}
-
-\subsection{Falling blocks}\label{sec:block}
-This benchmark is described as proposed by \citet{Gerya2003a}, \citet{Gerya2010b} and \citet{Thieulot2011}. The domain is square with $L_x=L_y=$ \SI{500}{\km}
-and the grid is composed by \num{50x50} elements with 25 markers in each element. The block is initially centred at ($x=$ \SI{250}{\km}; $y=$ \SI{400}{\km})
-and has a size of \num{100x100} km (Fig. \ref{fig:block}a). The fluid surrounding the block has $\eta_f=$ \SI{e21}{\pascal\per\s} and $\rho_f=$
-\SI{3200}{\kg\per\cubic\m}. The benchmark tests with different viscosities of the block, with $\eta_b$ from \SIrange{e15}{e27}{\pascal\per\s}. In each
-experiment also the density $\rho_b$ of the block varies from \SIrange{3220}{9900}{\kg\per\cubic\m}. Velocity boundary conditions are set to free slip
-conditions on all sides of the domain. Density distributions at $t=$ \SI{20}{\mega\year} for $\rho_b=$ \SI{3300}{\kg\per\cubic\m} and different $\eta_b$
-are plotted in Fig. \ref{fig:block}b-f, showing that the code correctly preserve the block geometry in case of large viscosity contrast, with a block stiffer
-than the surrounding fluid (Fig. \ref{fig:block}f).
-
-The velocity in the centre of the falling block at $t=0$ is measured for all experiments and, since the velocity should increase with density contrast, the
-quantity $\bm{v}/(\rho_b-\rho_f)$ is plotted as function of the viscosity contrast. All results perfectly match with those from \citet{Gerya2010b} and they
-line up on a single curve, demonstrating that the code can correctly deal with large viscosity and density contrasts (Fig. \ref{fig:falling}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Falling%20blocks}.
-
-\begin{figure}
-\noindent\includegraphics[width=300px]{./Figures/Block.pdf}
-\centering
-\caption{Density evolution of the falling block experiment at $t=$ \SI{0}{\mega\year} (panel a) and $t=$ \SI{20}{\mega\year} for different viscosities of the
-block (panels b-f).}
-\label{fig:block}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Falling.pdf}
-\caption{Initial velocity relative to the density contrast at the centre of the falling block as function of the viscosity contrast between the block and the
-surrounding fluid.}
-\label{fig:falling}
-\end{figure}
-
-\subsection{2D time-dependent Stokes sphere with deformable free surface}\label{sec:sphere}
-This experiment is performed in a unit square domain with the gravity acceleration fixed to $g_y=-1$. The sphere has $\rho_s=2$ and $\eta_s=10^3$ and is
-initially centred at ($0.5;0.6$) with radius $R=0.123456789$. The fluid surrounding the sphere has $\rho_f=1$ and $\eta_f=1$ and occupies the domain for
-$y\leq 0.75$, while for $y>0.75$ the air has $\rho_a=0$ and $\eta_a=10^{-3}$. Velocity boundary conditions are set to free slip on all sides. The Courant
-number is set to 0.25. The experiments run for 200 s using grid resolutions from \num{150x150} to \num{512x512} elements, with 25 randomly distributed markers
-per element, and for different average schemes for the viscosity. The interface between the fluid and the air is tracked by means of the markers chain. Results
-in terms of velocity and pressure fields and topography variations are compared with results obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and they can be found at
-\url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere_fs2D} and
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Time_dependent_sphere}. Results in terms of the $v_{\textrm{rms}}$ in case of 
-an arithmetic average are also shown in Fig. \ref{fig:inst_spherefs}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_arithm.pdf}
-\caption{$v_{\textrm{rms}}$ for different grid resolutions as function of time for the 2D time-dependent Stokes sphere experiment in case of an arithmetic average.
-Results are compared with results obtained by ASPECT.}
-\label{fig:inst_spherefs}
-\end{figure}
-
-\subsection{Free surface stabilisation}\label{sec:stab}
-The benchmark is performed as discussed in \citet{Kaus2010a} and \citet{Thieulot2014}. The domain is square with $L_x=L_y=$ \SI{500}{\km} and a grid resolution
-of \num{200x200} elements, each of them containing 16 markers. A buoyant fluid with $\rho_1=$ \SI{3200}{\kg\per\cubic\metre} and $\eta_1=$ \SI{e20}{\pascal\s}
-is overlain by a denser fluid with $\rho_2=$ \SI{3300}{\kg\per\cubic\metre} and $\eta_2=$ \SI{e21}{\pascal\s}. The initial interface between the fluids has an
-initial sinusoidal shape of 5 km amplitude. Velocity boundary conditions are set to free slip conditions at the sides and no slip at the bottom, while the top
-is free surface. The experiment is performed with various fixed time steps, without the use of the Courant number. The vertical position of the free surface at
-$x=L_x$ is tracked for each simulation, with or without the stabilisation algorithm.
-
-The results show an instability (drunken sailor effect) increasing the time step in case that the stabilisation algorithm is not activated
-($dt=$ \SI{4200}{\year}, green line in Fig. \ref{fig:stab}), as already observed by \citet{Kaus2010a} and \citet{Thieulot2014}. Activating the algorithm the
-instability is fixed and the simulation remains stable using time step up to \SI{50000}{\year} (red line in Fig. \ref{fig:stab}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Stabilisation_algorithm}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Kaus_random.pdf}
-\caption{Evolution of the y coordinate in $x=L_x$ as function of time for different time steps, with and without the stabilisation algorithm.}
-\label{fig:stab}
-\end{figure}
-
-\subsection{Topography relaxation}\label{sec:crameri}
-This experiment is performed as in \citet{Crameri2012} using free surface and sticky air (Case 1). The domain is rectangular with $L_x=$ \SI{2800}{\km} and
-$L_y=$ \SI{700}{\km} in case of free surface and $L_y=$ \SI{800}{\km} in case of sticky air ($h_{st}=$ \SI{100}{\km}). The grid is composed by \num{256x64}
-elements for the free surface \num{560x320} elements for the sticky air. For both cases, there is a mantle of 600 km thickness with $\eta_m=$ \SI{e21}{\pascal\s}
-overlain by a highly viscous cosine-shaped lithosphere with thickness between 93 and 107 km and $\eta_l=$ \SI{e23}{\pascal\s}. Both mediums have $\rho=$
-\SI{3300}{\kg\per\cubic\metre}. In case of sticky air experiment, the air has $\rho_a=$ \SI{0}{\kg\per\cubic\metre} and $\eta_a=$ \SI{e18}{\pascal\s}. The
-gravity acceleration is set to $g_y=$ \SI{-10}{\metre\per\square\s}. Velocity boundary conditions are set to no slip at the bottom and free slip at the sides
-of the domain. In case of sticky air, velocities on top are set to free slip conditions.
-
-The maximum topography as function of time ($h(t)$) can be analytically derived in according to \[h(t)=h_0 \exp(\gamma t)\]
-where $h_0=7$ km is the initial topography, $\gamma=\num{-0.2139e-11}$ is the characteristic relaxation rate and $t$ is the time (black line in Fig.
-\ref{fig:crameri}). Maximum topography at the characteristic relaxation time $t_{rlx}=$ \SI{14.825}{\kilo\year} can be found to be $h_{rlx}=$ \SI{2576}{\m}.
-The results show that a viscosity of the air of \SI{e18}{\pascal\s} is needed in case of $h_{st}=$ \SI{100}{\km} to correctly describe the topography relaxation
-for this problem, as pointed out by \citet{Crameri2012} (purple line in Fig. \ref{fig:crameri}). The case with a true free surface well follows the analytical
-solution (orange line in Fig. \ref{fig:crameri}). In particular, at the relaxation time in case of the true free surface predicts a maximum topography of
-\SI{2570}{\m}, with an error of \SI{6}{\m} with respect to the analytical solution. Results are also compared with results obtained with a free surface in
-UNDERWORLD (\num{256x64} elements) and SULEC (\num{401x201} elements) (with courtesy from \citealp{Crameri2012}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Topography_relaxation}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Crameri.pdf}
-\caption{Maximum topography as function of time for the Crameri benchmark (coloured lines) in comparison with the analytical solution (black line) and results
-obtained with UNDERWORLD (yellow line) and SULEC (blue line).}
-\label{fig:crameri}
-\end{figure}
-
-\subsection{Spontaneous subduction}\label{sec:subduction}
-This experiment is performed as presented by \citet{Schmeling2008}, in case of both a sticky air and a true free surface and using different average schemes
-for the viscosities. The domain is rectangular with $L_x=$ \SI{3000}{\km} and $L_y=$ \SI{700}{\km}, a grid resolution of \num{256x64} elements with 50 markers
-per element and Courant number of 0.01. At the beginning of the simulation, a 100 km-thick lithospheric layer with $\rho_m=$ \SI{3300}{\kg\per\cubic\metre}
-and $\eta_m=$ \SI{e23}{\pascal\s} is located at the top of the domain for $x$ between 1000 and 3000 km, while the asthenospheric mantle has $\rho_m=$
-\SI{3200}{\pascal\s} and $\eta_m=$ \SI{e21}{\pascal\s}. In addition, a 200 km-depth lithospheric slab is already subducted in the mantle in order to have
-a spontaneous subduction. In case of sticky air, $L_y=750$ km and a 50 km-thick air layer with $\eta_a$ from \SIrange{e19}{e21}{\pascal\s} overlie the mantle.
-Velocity boundary conditions are set to free slip on the sides and on the bottom of the domain. In case of sticky air, velocities on the top boundary are set
-to free slip conditions as well.
-
-The maximum depth of the slab is tracked for all the simulations (Fig. \ref{fig:subduction}) and they are compared with results obtained by
-\citet{Schmeling2008} with models with regular grids and comparable grid resolutions (rectangles in Fig. \ref{fig:subduction}). Low viscous sticky air models
-(i.e., \SI{e19}{\pascal\s}) enlighten a strong dependence of the sinking velocity with the chosen average scheme (dashed lines in Fig. \ref{fig:subduction}),
-while high viscous sticky air models (i.e., \SI{e21}{\pascal\s}) show a higher resistance at the trench, underestimating the correct solution (dotted lines in
-Fig. \ref{fig:subduction}). A high resistance at the trench can be also observed in case of a true free surface for the low resolution grid used (continuous
-lines in Fig. \ref{fig:subduction}), as pointed out by \citet{Schmeling2008}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Spontaneous_subduction}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Subduction.pdf}
-\caption{Maximum depth of the slab as function of time for all the simulations of the \citet{Schmeling2008} benchmark, in case of harmonic, geometric and
-arithmetic means (red, green and blue lines, respectively) using a sticky air or a true free surface (discontinuous and continuous lines, respectively).
-Red, green and blue rectangular areas indicate the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{500}{\km} in case of sticky air,
-while the grey rectangular area indicates the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{600}{\km} in case of true free surface.}
-\label{fig:subduction}
-\end{figure}
-
-\subsection{The slab detachment}\label{sec:slab}
-The slab detachment benchmark is performed as described by \citet{Schmalholz2011} and \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{1000}{\km}
-and $L_y=$ \SI{660}{\km} and a grid resolution of \num{256x256} elements. A non-linear viscous T-shaped layer with $\rho_l=$ \SI{3300}{\kg\per\cubic\metre} is
-placed at the top of the domain and surrounded by a linear viscous fluid with $\rho_f=$ \SI{3150}{\kg\per\cubic\m}. The top layer is 80 km-thick and a 250
-km-long and 80 km-wide slab is placed at $x=L_x/2$ (Fig. \ref{fig:slab}a). The effective viscosity of the top layer is given by
-\[\eta_{eff}=\eta_0 I_2^{\frac{(1-n)}{n}}\]
-with $\eta_0=$ \SI{4.75e11}{\pascal\s} and $n=4$, while the surrounding fluid has $\eta_f=$ \SI{e21}{\pascal\s}. The effective viscosity is capped between
-$\eta_{min}=$ \SI{e21}{\pascal\s} and $\eta_{max}=$ \SI{e25}{\pascal\s}. Velocity boundary conditions are set to free slip at the top and the bottom, and to
-no slip at the sides of the domain.
-
-The sinking of the slab determines a decrease of the effective viscosity in the T-shaped layer during the first part of the evolution (Fig. \ref{fig:slab}a),
-as consequence of local strain rates, while effective viscosities increase up to $\eta_{max}$ in the top layer after approximately \SI{20}{\mega\year}, when
-necking is complete (Fig. \ref{fig:slab}b).
-
-The width of necking is tracked by means of markers position at the side of the slab. Necking width and time are normalised in function of the initial slab
-(\SI{80}{\km}) and the characteristic time ($t_c=$ \SI{7.1158e14}{\s}) \citep{Schmalholz2011,Glerum2018}. The results in case of arithmetic, geometric and
-harmonic mean of the viscosity are shown in Fig. \ref{fig:necking}, compared with results from \citet{Glerum2018} for which an infinity norm mean was used.
-The influence of the average schemes agrees with results shown in Section \ref{sec:subduction}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Slab_detachment}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Slab.pdf}
-\caption{Effective viscosity for the slab detachment benchmark at the beginning of the evolution (panel a) and when necking is complete (panel b).}
-\label{fig:slab}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Necking.pdf}
-\caption{Normalised width of the necking of the slab detachment benchmark as function of normalised time.}
-\label{fig:necking}
-\end{figure}
-
-\subsection{The indenter experiment}\label{sec:indenter}
-The indenter benchmark simulate a rigid punch on a purely plastic von Mises material, for which exists an analytical solution
-\citep{Thieulot2008,Thieulot2014,Glerum2018}. The domain is a unit square with a grid resolution of \num{256x256} elements. The gravity acceleration is set
-to $\bm{g}=\bm{0}$ and effective viscosity is capped using $\eta_{min}=10^{-4}$ and $\eta_{max}=10^3$. The tolerance for the convergence of the non-linear
-solution is $tol=10^{-9}$, with a maximum number of non-linear iterations set to 500. The medium in the domain has $\rho=0.01$, an initial viscosity
-$\eta_0=10$, cohesion $C=1$ and an angle of internal friction $\phi=0$°. Velocity boundary conditions are set to no slip at the bottom and free slip at the
-sides of the domain. The top of the domain is open with the exception of the central portion (punch area) with width $w_p=0.125$, where $v=-1.05$ and $u$ is
-fixed either to 0 (rough punch experiment) or free (smooth punch experiment).
-
-Fig. \ref{fig:indenter} show the results in terms of viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure
-(panels g and h) for both experiments. The analytical solution indicate that pressure at ($0.5;1$) and ($0.5 \pm w_p;1$)is $p=\pi +1$ and $p=1$, respectively
-(black and red lines in Fig. \ref{fig:smooth_p}). Fig. \ref{fig:smooth_p} shows a clear improvement in the pressure solution when passing from a rough to a
-smooth punch (green and blue lines in Fig. \ref{fig:smooth_p}, respectively), as previously observed by \citet{Thieulot2014} and \citet{Glerum2018}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Indenter}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Indenter.pdf}
-\caption{Viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure (panels g and h) fields for rough (left column) and
-smooth (right column) punch experiments.}
-\label{fig:indenter}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Smooth_Rough.pdf}
-\caption{Nodal pressure in function of the x coordinate at $y=1$ for rough (green line) and smooth (blue line) punch experiment. Black and red lines indicate
-the analytical solution at $x=0.5$ and $x=0.5 \pm w_p$, respectively.}
-\label{fig:smooth_p}
-\end{figure}
-
-\subsection{The brick experiment}\label{sec:brick}
-An instantaneous version of the brick benchmark is performed to verify the correctness of the pressure-dependent plasticity for different angle of internal
-friction, as proposed by \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{40}{\km} and $L_y=$ \SI{10}{\km} and a grid resolution of \num{512x128}
-elements. The gravity acceleration is set to $g_y=$ \SI{-10}{\m\square\s} and effective viscosity is capped using $\eta_{min}=$ \SI{e19}{\pascal\s} and
-$\eta_{max}=$ \SI{e26}{\pascal\s}. The tolerance for the convergence of the non-linear solution is $tol=10^{-7}$, with a maximum number of non-linear iterations
-set to 1000. A 800 m-wide and 400 m-height inclusion with $\eta_b=$ \SI{e20}{\pascal\s} and $\rho_b=$ \SI{2700}{\kg\per\cubic\m} is placed at the bottom of the
-domain at $x=L_x/2$. The inclusion is surrounded by a non-linear viscous medium with $\rho_m=$ \SI{2700}{\kg\per\cubic\m}, an initial viscosity of
-\SI{e23}{\pascal\s}, a linear viscous viscosity of \SI{e25}{\pascal\s} and a cohesion of \SI{40}{\mega\pascal}. Velocities are set to free slip conditions at
-the bottom of the domain and the top is open. Velocities on sides of the domain are fixed to $u=$ \SI{+-2e-11}{\m\per\s} and $v=0$. The experiment is performed
-in compressional and extensional contexts with an angle of internal friction $\phi$ of the non-linear medium variable between 0° and 30°.
-
-As expected, two shear bands stem from the inclusion with variable angles in relation to both the dynamics context and the internal friction angle. Shear band
-angles formed at 45° for $\phi=0$° in both extensional and compressional contexts (Fig. \ref{fig:brick_beam}a), while they are different in case of $\phi\neq0$°.
-Shear band angles in case of $\phi=20$° are shown for compression and extension in Fig. \ref{fig:brick_beam}b  and c, respectively. Values of shear band angles
-as function of internal friction angles from 0° to 30° are extracted at different depths and minimum and maximum values are plotted in Fig. \ref{fig:brick},
-in comparison with theoretical Roscoe, Arthur and Coulomb shear band angles. For all tests the tolerance for nonlinear convergence is defined $tol=10^{-7}$
-and none converges before the maximum number of iterations ($it_{max}=1000$) is reached. However, tests with internal friction angles up to 15° show a constant
-decrease of the velocity residuals, which can not be observed for for tests with higher internal friction angles (Fig. \ref{fig:convergence}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Brick_experiment}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Brick_Beam.pdf}
-\caption{Shear band angles predicted for the brick experiment for $\phi=0$° (panel a) and for $\phi=20$° in case of compression and extension (panels b and c,
-respectively).}
-\label{fig:brick_beam}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Brick.pdf}
-\caption{Shear band angles predicted for the brick experiment in case of compressional and extensional contexts as function of different internal angle of
-friction (continuous black line and red dots), compared with theoretical Roscoe, Artur and Coulomb angles (discontinuous black lines). Red lines indicate the
-range of angles calculated at different depths.}
-\label{fig:brick}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \noindent\includegraphics[width=500px]{./Figures/convergence.pdf}
-  \caption{Convergence rates for all tests. Normalised residuals (left column) are compared with velocity residuals (right column) for extensional (first row)
-  and compressional (second row) settings.}
-  \label{fig:convergence}
-  \end{figure}
-
-\subsection{Advection stabilisation}\label{sec:advection}
-The 1D advection problem proposed by \citet{Donea2003} and \citet{Thieulot2011} is performed to verify the effectiveness of the SUPG method to stabilise the
-advection term of the energy equation. The domain is a 1D segment with $L_x=1$ composed by 50 elements and a discontinuity in the thermal field placed at
-$x=0.25$. Temperature is set to 1 for $x \leq 0.25$ and to 0 for $x > 0.25$. Velocity is set to ${u}=1$ in the entire domain. The simulation is performed for
-250 time steps, with $dt=0.002$, so the thermal discontinuity should be at $x=0.75$ at the end of the simulation.
-
-Temperature profiles at the end of the simulation are shown in \ref{fig:advection} as function of the dimensionless coefficient $\gamma=\tau \bm{v}/h$. In case
-of the classic Galerkin method ($\gamma=0$, blue line in \ref{fig:advection}) the final thermal profile is characterised by strong oscillations, which are
-eliminated in case of the SUPG method ($\gamma=0.045$, orange line in \ref{fig:advection}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Advection_stabilisation}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Advection.pdf}
-\caption{Temperature profile as function of \textit{x} for the advection stabilisation benchmark. Purple line indicates the initial temperature profile; the
-green line indicates the analytical temperature profile after 250 time steps; blue line indicates the temperature profile after 250 time steps in case of the
-classic Galerkin method ($\gamma=0$); orange line indicates the temperature profile after 250 time steps in case of the SUPG method ($\gamma=0.045$).}
-\label{fig:advection}
-\end{figure}
-
-\subsection{Shear heating}\label{sec:simple_shear}
-This exercise is performed in an unit square domain composed by \num{128x128} elements. The velocity field is prescribed on the entire domain with
-$\bm{v}=(L_y-y)y\bm{e_x}$; viscosity, density and specific heat are fixed to 1, while thermal conductivity and radiogenic energy are fixed to 0. Therefore,
-the energy equation (Eq. \ref{eq:energy}) can be simplified as \[\frac{\partial T}{\partial t}=H_s\] and fixing $T(t=0)=0$, the temperature can be find as
-\[T(t)=H_s t\]
-In this case we have
-\begin{eqnarray}
-\dot{\epsilon_{xx}}&=&\frac{\partial u}{\partial x}=0 \nonumber \\
-\dot{\epsilon_{yy}}&=&\frac{\partial v}{\partial y}=0 \nonumber \\
-\dot{\epsilon_{xy}}&=&\frac{1}{2}\left(\frac{\partial v}{\partial x}+\frac{\partial u}{\partial y}\right)=\frac{1}{2}(L_y-2y)\nonumber
-\end{eqnarray}
-and, simplifying Eq. \ref{eq:shear}, shear heating can be calculated as \[H_s(x,y)=(1-2y)^2\]
-
-The solution predicted by the model in terms of velocity, temperature and shear heating match well with the analytical solutions (Fig. \ref{fig:analytical_en}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Simple_shear}.
-
-%\begin{figure}
-%\centering
-%\noindent\includegraphics[width=400px]{./Figures/Energy_Thieulot.pdf}
-%\caption{Temperature and shear heating predicted by the model after 1 time step.}
-%\label{fig:en_thieulot}
-%\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=500px]{./Figures/Analytical_Shear.pdf}
-\caption{Velocity (panel a), temperature (panel b) and shear heating (panel c) as function of the \textit{y} coordinate for the simple shear experiment. The
-solutions predicted by the model (red lines) are compared with the analytical solutions (dashed blue lines).}
-\label{fig:analytical_en}
-\end{figure}
-
-\subsection{Shear and adiabatic heating}\label{sec:shear}
-This problem is performed as presented in Exercise 9.4 in \citet{Gerya2010b}. Two materials are vertically separated in a rectangular domain with $L_x=$
-\SI{1000}{\km}, $L_y=$ \SI{1500}{\km} and a grid resolution of \num{30x20} elements. Constant thermal coefficient expansion ($\alpha=$ \SI{3e-5}{\per\kelvin}),
-temperature ($T=$ \SI{1300}{\kelvin}) and gravity acceleration ($g_y=$ \SI{-10}{\m\square\s}) are assumed in the whole domain. Fluid 1 (on the left side of the
-domain) has $\rho_1=$ \SI{3200}{\kg\per\cubic\m} and $\eta_1=$ \SI{e20}{\pascal\s}; fluid 2 (on the right side of the domain) has $\rho_2=$
-\SI{3300}{\kg\per\cubic\m} and $\eta_2=$ \SI{e22}{\pascal\s}. Velocity boundary conditions are set to free slip on all sides of the domain.
-
-As shown in Fig. \ref{fig:energy}, both shear and adiabatic heating predicted by the code (first row in Fig. \ref{fig:energy}) well recreate the results
-obtained by means of example ${Shear\_adiabatic\_heating.m}$ from \citet{Gerya2010b} (second row in Fig. \ref{fig:energy}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Adiabatic%2BShear}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Energy.pdf}
-\caption{Comparison between shear (first column), adiabatic (second column) and total (third column) energy predicted by the code (panels a, b and c) and those
-created using example ${Shear\_adiabatic\_heating.m}$ from Exercise 9.4 in \citet{Gerya2010b} (panels d, e and f).}
-\label{fig:energy}
-\end{figure}
-
-\subsection{Mantle convection}\label{sec:mantle}
-This problem is performed as presented by \citet{Blankenbach1989} (constant viscosity cases) and \citet{Thieulot2014}, in a 2D unit square domain with gravity
-acceleration $g_y=10^{10}Ra$. The experiment is performed with three different Rayleigh numbers ($Ra=10^4$, $10^5$ and $10^6$) and with different grid
-resolution (between \num{32x32} and \num{128x128} elements). The fluid has constant viscosity, initial density, heat capacity, thermal conductivity
-($\eta= \rho_0= C_p = k=1$), reference temperature ($T_0=0$) and thermal expansion coefficient ($\alpha=10^{-10}$). Temperatures are set to 0 on top and 1 on
-bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\]
-
-The solution generated by the code in terms of $v_{\textrm{rms}}$ (calculated as in Sec. \ref{sec:error}) and Nu as function of time are reported for all the
-simulations in Fig. \ref{fig:mantle}. At the steady state, $v_{\textrm{rms}}$ and Nu of all simulations converge well toward the values from \citet{Blankenbach1989},
-with lower errors for higher resolution grids (Table \ref{tab:mantle}). All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Mantle_convection}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Convection.pdf}
-\caption{$v_{\textrm{rms}}$ (panels a, c and e) and Nu (panels b, d and f) for the mantle convection benchmark as function of time for different grid resolution.
-Panels a and b show the results for $Ra=10^4$; Panels c and d show the results for $Ra=10^5$; panels e and f show the results for $Ra=10^6$. Purple lines
-indicate the convergence values for the $v_{\textrm{rms}}$ and the Nu from \citet{Blankenbach1989} at the steady state.}
-\label{fig:mantle}
-\end{figure}
-
-\begin{table}[H]
-\caption{Comparison between $v_{\textrm{rms}}$ and Nu predicted by the code for the mantle convection experiment and same values as reported in literature.}
-\centering
-\small
-\begin{tabular}{|>{\centering}m{0.09\textwidth}|>{\centering}m{0.05\textwidth}|>{\centering}m{0.24\textwidth}|>{\centering}m{0.15\textwidth}|
-  >{\centering}m{0.1\textwidth} >{\centering}m{0.1\textwidth} >{\centering\arraybackslash}m{0.1\textwidth}|}
-\toprule
-& & \multirow{2}{*}{\citet{Blankenbach1989}} & \citet{Thieulot2014} & \multicolumn{3}{c|}{FALCON}  \\
- & & & \num{200x200} & \num{32x32} & \num{64x64} & \num{128x128}  \\
-\midrule
-\multirow{2}{*}{$Ra=10^4$} & $v_{\textrm{rms}}$ & $42.864947 \pm 0.000020$ & 42.867 & 42.83226 & 42.852793 & 42.861394  \\
-           & Nu      & $4.884409 \pm 0.000010$  & 4.882  & 4.781297 & 4.857475  & 4.877573   \\
-\hline
-\multirow{2}{*}{$Ra=10^5$} & $v_{\textrm{rms}}$ & $193.21454 \pm 0.00010$  & 193.255& 193.872643 & 193.377472 & 193.252290  \\
-           & Nu      & $10.534095 \pm 0.000010$ & 10.507 & 9.602514 & 10.270735  & 10.465629   \\
-\hline
-\multirow{2}{*}{$Ra=10^6$} & $v_{\textrm{rms}}$ & $833.98977 \pm 0.00020$  & 834.712& 848.091176 & 837.767911 & 834.945793  \\
-           & Nu      & $21.972465 \pm 0.000020$ & 21.695 & 15.999266 & 19.703682  & 21.306939   \\
-\bottomrule
-\end{tabular}
-\label{tab:mantle}
-\end{table}
-
-\subsection{Viscoplastic mantle convection}\label{sec:visco_mantle}
-This benchmark is performed as Cases 1-5a described in \citet{Tosi2015} in a 2D unit square domain with two different grid resolutions of \num{32x32} 
-and \num{100x100} elements. Reference viscosity, reference density, heat capacity, thermal conductivity and thermal expansion coefficient are fixed to a constant
-value of 1 ($\eta_0= \rho_0= C_p = k=\alpha=1$), while the gravity has been chosen as $g_y=10^2$, in order to obtain a Rayleigh number $Ra=10^2$. Temperatures
-are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\].
-The viscosity field $\eta$ is calculated as the harmonic average between a linear part $\eta_{lin}$ that depends on temperature only or on temperature and 
-depth $y$ and a nonlinear-plastic part $\eta_{pl}$ that depends on the strain rate $\dot{\bm{\epsilon}}$, as follows
-\[\eta(T,y,\dot{\bm{\epsilon}})=2\left(\frac{1}{\eta_{lin}(T,y)}+\frac{1}{\eta_{pl}(\dot{\bm{\epsilon}})}\right)^{-1}\]
-The linear and the nonlinear-plastic parts of the viscosity are calculated following \citet{Tosi2015} as follows
-\[\eta_{lin}(T,y)=\exp(-\ln(\Delta\eta_T)T+\ln(\Delta\eta_y)y)\]
-\[\eta_{pl}(\dot{\bm{\epsilon}})=\eta^*+\frac{\sigma_Y}{\sqrt{\dot{\bm{\epsilon}}:\dot{\bm{\epsilon}}}}\]
-where $\Delta\eta_T$, $\Delta\eta_y$, $\eta^*$ and $\sigma_Y$ are parameters chosen as listed in Table \ref{tab:tosi_setup}.
-
-For each case, some diagnostic quantities measured at the steady state are reported in Table \ref{tab:visco_mantle} (see Section \ref{sec:error} for 
-the computation of all the quantities), showing a well fit with results described in \citet{Tosi2015}. Data and charts for all cases can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Viscoplastic_mantle_convection}.
-
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{white!100!}{gray!25!}
-%\begin{table}[H]
-%  \caption{Results of the viscoplastic mantle convection benchmark for Cases 1-4 compared with results obtained with ELEFANT and YACC (from \citealp{Tosi2015}).}
-%  \centering
-%  \resizebox{\textwidth}{!}{
-%\begin{tabular}{|c|ccc|ccc|ccc|ccc|}
-%  \toprule
-%    & \multicolumn{3}{c|}{\textbf{Case 1}} & \multicolumn{3}{c|}{\textbf{Case 2}} & \multicolumn{3}{c|}{\textbf{Case 3}} &
-%    \multicolumn{3}{c|}{\textbf{Case 4}} \\
-%   & FALCON & ELEFANT& YACC & FALCON& ELEFANT& YACC & FALCON & ELEFANT & YACC & FALCON & ELEFANT & YACC \\
-%  \midrule
-%   $\left\langle T \right\rangle$ & 0.7759 & 0.7758 & 0.7767 & 0.6043 & 0.6036 & 0.6058 & 0.7275 & 0.7275 & 0.7286 & & 0.5277 & 0.5289\\
-%    Nu$^{top}$ & 3.4280 & 3.4214 & 3.4298 & 8.5242 & 8.5115 & 8.5278 & 3.0368 & 3.0347 & 3.0374 && 6.5912 & 6.5572 \\
-%    Nu$^{bottom}$ & 3.3168 & 3.313 & 3.3143 & 8.4599 & 8.4465 & 8.3990 & 3.0086 & 2.9908 & 2.9628 && 6.5834 & 6.5243 \\
-%    $v_{\textrm{rms}}$ & 249.901 & 249.134 & 251.7997 & 141.2451 & 140.7087 & 142.2020 & 100.2776 & 100.1208 & 100.9467 && 79.1105 & 79.6202 \\
-%    $v_{\textrm{rms}}^{top}$ & 1.8907 & 1.8642 & 1.8298 & 104.7928 & 104.1228 & 105.1767 & 2.0832 & 2.0652 & 2.0374 && 74.7596 & 75.8414 \\
-%    $u_{max}^{top}$ & 2.6340 & 2.6119 & 2.5516 & 121.9212& 121.766 & 122.3238 & 2.9086 & 2.9019 & 2.8458 && 88.9146 & 89.2940 \\
-%    $\eta_{min}$ & - & - & - & $2.0303\times10^{-5}$ & $2.0600\times10^{-5}$ & $1.9853\times10^{-5}$ & $4.8129\times10^{-5}$ & $4.8080\times10^{-5}$ & $4.7907\times10^{-5}$ &
-%    & $1.9860\times10^{-4}$ & $1.9174\times10^{-4}$ \\
-%    $\eta_{max}$ & - & - & - & 1.3325 & 1.85 & 1.6095 & 0.9360 & 0.9023 & 1 && 1.52 & 1.6773 \\
-%    $\left\langle W \right\rangle$ & 2.4299 & 2.4316 & 2.4583 & 7.5648 & 7.5665 & 7.6222 & 2.0387 & 2.0384 & 2.0400 && 5.6216 & 5.6512 \\
-%    $\left\langle \Phi \right\rangle/Ra$ & 2.4311 & 2.4276 & 2.4333 & 7.5661 & 7.5585 & 7.6194 & 2.0386 & 2.0370 & 2.0335 && 5.6182 & 5.6463 \\
-%    $\delta$ & 0.05\% & 0.16\% & 1.02\% & 0.02\% & 0.11\% & 0.04\% & $<$0.01\% & 0.07\% & 0.32\% && 0.06\% & 0.09\% \\
-%  \bottomrule
-%  \end{tabular}}
-%  \label{tab:visco_mantle}
-%  \end{table}
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{2}{white!100!}{gray!25!}
-\begin{table}[H]
-  \caption{Benchmark cases and corresponding parameters.}
-  \centering
-  \small
-\begin{tabular}{|>{\centering}m{0.08\textwidth}|>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}
-  >{\centering}m{0.05\textwidth} >{\centering}m{0.05\textwidth} |>{\centering\arraybackslash}m{0.2\textwidth}|}
-  \toprule
-    Case & $Ra$ & $\Delta\eta_T$ & $\Delta\eta_y$ & $\eta^*$ & $\sigma_Y$ & Convective regime \\
-  \midrule
-  \rowcolor{gray!25!}
-    1 & $10^2$ & $10^5$ & 1 & - & - & Stagnant lid \\
-    2 & $10^2$ & $10^5$ & 1 & $10^{-3}$ & 1 & Mobile lid  \\
-    \rowcolor{gray!25!}
-    3 & $10^2$ & $10^5$ & 10 & - & - & Stagnant lid  \\
-    4 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 1 & Mobile lid \\
-    \rowcolor{gray!25!}
-    5 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 4 & Periodic \\
-  \bottomrule
-  \end{tabular}
-  \label{tab:tosi_setup}
-  \end{table}
-
-\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{gray!25!}{white!100!}
-\begin{table}
-  \caption{Results of the viscoplastic mantle convection benchmark for different grid resolutions in Cases 1-5 compared with results obtained with ELEFANT
-  and YACC ($100\times100$ elements, from \citealp{Tosi2015}) and in Stone 28 ($32\times32$ elements, from
-  \url{https://github.com/cedrict/fieldstone/tree/master/python_codes/fieldstone_28}).}
-  \centering
-  \resizebox{\textwidth}{!}{
-\begin{tabular}{|cc|ccccccccccc|}
-  \toprule
-  \multicolumn{13}{|c|}{\textbf{Case 1}} \\
-  \midrule
-  Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-  & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-  \hline
-  \rowcolor{gray!25!}
-    YACC & $100\times100$ & 0.7767 & 3.4298 & 3.3143 & 251.7997 & 1.8298 & 2.5516 & - & - & 2.4583 & 2.4333 & 1.02\% \\
-    ELEFANT & $100\times100$ & 0.7758 & 3.4214 & 3.313 & 249.134 & 1.8642 & 2.6119 & - & - & 2.4316 & 2.4276 & 0.16\% \\
-    \rowcolor{gray!25!}
-    Stone 28 & $32\times32$ & 0.7737 & 3.3987 & - & 243.872 & - & - & - & - & - & - & - \\
-%    FALCON & $32\times32$ & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & - & - & 2.4745 & 2.4864 & 0.48\% \\
-%    \rowcolor{gray!25!}
-    FALCON & $100\times100$ & 0.7759 & 3.4280 & 3.3168 & 249.901 & 1.8907 & 2.6340 & - & - & 2.4299 & 2.4311 & 0.05\% \\
-    \midrule
-    \multicolumn{13}{|c|}{\textbf{Case 2}} \\
-    \midrule
-    Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-    & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-    \hline
-    \rowcolor{gray!25!}
-     YACC & $100\times100$ & 0.6058 & 8.5278 & 8.3990 & 142.2020 & 105.1767 & 122.3238 & $1.9853\times10^{-5}$ & 1.6095 & 7.6222 & 7.6194 & 0.04\% \\
-     ELEFANT & $100\times100$ & 0.6034 & 8.5115 & 8.4465 & 140.8067 & 104.1228 & 121.766 & $2.0600\times10^{-5}$ & 1.85 & 7.5665 & 7.5585 & 0.11\% \\
-     \rowcolor{gray!25!}
-     Stone 28 & $32\times32$ & 0.6052 & 8.1715 & - & 141.518 & - & - & - & - & - & - & - \\
-%     FALCON & $32\times32$ & 0.6454 & 8.1422 & 7.4020 & 156.6908 & 95.7226 & 116.9302 & $2.2572\times10^{-5}$ & 0.7194 & 7.4289 & 7.4452 & 0.22\% \\
-%     \rowcolor{gray!25!}
-     FALCON & $100\times100$ & 0.6044 & 8.5266 & 8.4590 & 141.3684 & 104.8867 & 122.0470 & $2.0333\times10^{-5}$ & 1.2914 & 7.5707 & 7.5718 & 0.01\% \\
-     \midrule
-     \multicolumn{13}{|c|}{\textbf{Case 3}} \\
-     \midrule
-     Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-     & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-     \hline
-     \rowcolor{gray!25!} 
-     YACC & $100\times100$ & 0.7286 & 3.0374 & 2.9628 & 100.9467 & 2.0374 & 2.8458 & $4.7907\times10^{-5}$ & 1 & 2.0400 & 2.0335 & 0.32\% \\
-      ELEFANT & $100\times100$ & 0.7275 & 3.0347 & 2.9908 & 100.1208 & 2.0652 & 2.9019 & $4.8080\times10^{-5}$ & 0.9023 & 2.0384 & 2.037 & 0.07\% \\
-      \rowcolor{gray!25!}
-      Stone 28 & $32\times32$ & 0.7277 & 3.0347 & - & 100.018 & - & - & - & - & - & - & - \\
-%      FALCON & $32\times32$ & 0.7281 & 3.0618 & 2.8132 & 101.0092 & 2.1630 & 3.0178 & $4.8500\times10^{-5}$ & 0.7845 & 2.0699 & 2.0695 & 0.02\% \\
-%      \rowcolor{gray!25!}
-      FALCON & $100\times100$ & 0.7275 & 3.0368 & 3.0086 & 100.2776 & 2.0832 & 2.9086 & $4.8129\times10^{-5}$ & 0.9360 & 2.0387 & 2.0386 & $<$0.01\% \\
-      \midrule
-      \multicolumn{13}{|c|}{\textbf{Case 4}} \\
-      \midrule
-      Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-      & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-      \hline
-      \rowcolor{gray!25!} 
-      YACC & $100\times100$ & 0.5289 & 6.5572 & 6.5243 & 79.6202 & 75.4814 & 89.2940 & $1.9174\times10^{-4}$ & 1.6773 & 5.6512 & 5.6463 & 0.09\% \\
-       ELEFANT & $100\times100$ & 0.5277 & 6.5912 & 6.5834 & 79.1105 & 74.7596 & 88.9146 & $1.9860\times10^{-4}$ & 1.5200 & 5.6216 & 5.6182 & 0.06\% \\
-       \rowcolor{gray!25!}
-       Stone 28 & $32\times32$ & 0.5286 & 6.4036 & - & 79.4746 & - & - & - & - & - & - & - \\
-%       FALCON & $32\times32$ & 0.5334 & 6.4679 & 6.3435 & 80.6153 & 76.0511 & 90.2058 & $2.0393\times10^{-4}$ & 0.6453 & 5.6676 & 5.6698 & 0.04\% \\
-%       \rowcolor{gray!25!}
-       FALCON & $100\times100$ & 0.5279 & 6.5897 & 6.5830 & 79.1220 & 75.0715 & 88.8484 & $1.9693\times10^{-4}$ & 1.4133 & 5.6127 & 5.6113 & 0.03\% \\
-       \midrule
-       \multicolumn{13}{|c|}{\textbf{Case 5}} \\
-       \midrule
-       Code & Resolution & $P$ & $\left\langle T \right\rangle_{min}$ & $\left\langle T \right\rangle_{max}$ & Nu$^{top}_{min}$ & Nu$^{top}_{max}$ &
-       $v_{\textrm{rms}}^{min}$ & $v_{\textrm{rms}}^{max}$ & \multicolumn{2}{c}{$(\left\langle \Phi \right\rangle/Ra)_{min}$} & 
-       \multicolumn{2}{c|}{$(\left\langle \Phi \right\rangle/Ra)_{max}$} \\
-       \hline
-       \rowcolor{gray!25!}
-         YACC & $100\times100$ & 0.0792 & 0.6543 & 0.6722 & 2.6950 & 7.2792 & 41.9813 & 98.8230 & \multicolumn{2}{c}{1.4347} & 
-         \multicolumn{2}{c|}{9.2748} \\
-         ELEFANT & $100\times100$ & 0.076 & 0.6514 & 0.6693 & 2.6808 & 7.3506 & 41.2578 & 100.089 & \multicolumn{2}{c}{1.3324} & 
-         \multicolumn{2}{c|}{9.5289} \\
-         \rowcolor{gray!25!}
-         Stone 28 & $32\times32$ & 0.0764 & 0.6521 & 0.6697 & 2.6912 & 7.0763 & 41.61 & 98.183 & \multicolumn{2}{c}{-} & 
-         \multicolumn{2}{c|}{-} \\
-     %    FALCON & $32\times32$ & & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & \multicolumn{2}{c}{-} & 
-     %     \multicolumn{2}{c|}{-} \\
-     %    \rowcolor{gray!25!}
-         FALCON & $100\times100$ & 0.0775 & 0.6519 & 0.6697 & 2.6847 & 7.3629 & 41.5256 & 99.5466 & \multicolumn{2}{c}{1.3399} & 
-         \multicolumn{2}{c|}{9.3993} \\
-  \bottomrule
-  \end{tabular}}
-  \label{tab:visco_mantle}
-  \end{table}
-
-\subsection{Thin layer entrainment}\label{sec:thin}
-This experiment is performed as originally proposed by \citet{vanKeken1997} with two incompressible fluids in a rectangular domain with $L_x=2$ and $L_y=1$
-and gravity acceleration $g_y=10^{10}Ra$. The Rayleigh number and the compositional Rayleigh number are fixed to $Ra=300000$ and $Rc=450000$, respectively.
-Both fluids have constant viscosity, thermal conductivity, specific heat ($\eta=\rho=C_p=1$) and thermal expansion coefficient ($\alpha=10^{-10}$). Fluid 1
-has a density $\rho_1=1$, while fluid 2 is denser ($\rho_2=\rho_1+1.5 \cdot 10^{-10}$) and is placed at the bottom of the domain, for $y \leq 0.025$.
-Temperature are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides of the domain. The initial
-temperature field is given by
-\[T(x,y)=T_u(x,y)+T_l(x,y)+T_r(x,y)+T_s(x,y)-\frac{3}{2}\]
-with
-\begin{eqnarray}
-T_u(x,y)&=&\frac{1}{2}erf\left(\frac{1-y}{2}\sqrt{\frac{u_0}{x}}\right)\nonumber \\
-T_l(x,y)&=&1-\frac{1}{2}erf\left(\frac{y}{2}\sqrt{\frac{u_0}{L_x-x}}\right)\nonumber \\
-T_r(x,y)&=&\frac{1}{2}+\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{y+1}} \exp\left(-\frac{x^2u_0}{4y+4}\right)\nonumber \\
-T_s(x,y)&=&\frac{1}{2}-\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{2-y}} \exp\left(-\frac{(L_x-x)^2u_0}{8-4y}\right)\nonumber
-\end{eqnarray}
-and
-\begin{eqnarray}
-u_0&=&\frac{L_x^{7/3}}{(1+L_x^4)^{2/3}}\left(\frac{Ra}{2\sqrt{\pi}}\right)^{\frac{2}{3}}\nonumber \\
-Q&=&2\sqrt{\frac{L_x}{\pi u_0}}\nonumber
-\end{eqnarray}
-
-The experiment is performed with two grid resolution (\num{125x40} and \num{200x80} elements), with 100 markers per element and CFL of 0.25. $v_{\textrm{rms}}$
-(see Sec. \ref{sec:error}) calculated for both the simulations match well with results from \citet{vanKeken1997} and obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and ELEFANT \citep{Thieulot2014}, obtained with similar grid resolutions
-(Fig. \ref{fig:thin}). In addition, a variety of simulations with different resolutions and aspect ratios are performed to verify that the sensitivity of the
-initial velocity field. Results match well with results from \citet{vanKeken1997} (in the grey area) and \citet{Thieulot2014} (Fig. \ref{fig:thin_initial}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Thin_layer}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_thin.pdf}
-\caption{$v_{\textrm{rms}}$ for the thin layer experiment as function of time for different grid resolution. Results are compared with results obtained by
-\citet{vanKeken1997} (black lines) and with ASPECT and ELEFANT.}
-\label{fig:thin}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrmszero.pdf}
-\caption{$v_{\textrm{rms}}$ at $t=0$ for the thin layer experiment as function of the element size for different aspect ratios of the elements. Results are compared
-with results obtained by \citet{vanKeken1997} (black dots in the grey area) and with ASPECT and ELEFANT.}
-\label{fig:thin_initial}
-\end{figure}
-
-\subsection{Sinking hydrated cylinder}\label{sec:quinquis}
-This experiment is performed as explained in \citet{Quinquis2014} and simulates the sinking of a cold, hydrated cylinder into a warm, dry mantle. The domain
-is square with $L_x=L_y=$ \SI{300}{\km}, a grid resolution of \num{300x300} elements and 25 markers per element. The sinking cylinder is located at $x=$
-\SI{150}{\km} and $y=$ \SI{170}{\km} with a radius of \SI{20}{\km} and has $\rho_c=$ \SI{3250}{\kg\per\cubic\m}, $\eta_c=$ \SI{e23}{\pascal\s}, a thermal
-conductivity $k_c=$ \SI{4.5}{\watt\per\m\per\kelvin},a specific heat $C_{p_c}=$ \SI{1250}{\joule\per\kg\per\kelvin}; the surrounding mantle has $\rho_m=$
-\SI{3200}{\kg\per\cubic\m}, $\eta_m=$ \SI{e20}{\pascal\s}, a thermal conductivity $k_m=$ \SI{105}{\watt\per\m\per\kelvin} and specific heat $C_{p_m}=$
-\SI{1250}{\joule\per\kg\per\kelvin}; a 58 km-thick lithosphere overlies the mantle and has $\rho_l=$ \SI{3200}{\kg\per\cubic\m}, $\eta_l=$ \SI{e23}{\pascal\s},
-a thermal conductivity $k_l=$ \SI{4.5}{\watt\per\m\per\kelvin}, specific heat $C_{p_l}=$ \SI{750}{\joule\per\kg\per\kelvin} and a thermal diffusivity of
-\SI{e-6}{\square\m\per\s}. Velocity boundary conditions are set to free slip on all sides of the domain. Temperature increase linearly in the lithosphere,
-from 0° to 1300°C, and in the mantle, from 1300° to 1360.5°C. The initial temperature of the cylinder is fixed to 400°C. Throughout the evolution, temperature
-is fixed to 0° and 1360.5°C at the top and bottom of the domain, respectively. The initial bound water content is imposed at 2 wt.\% in the cylinder and at
-0 wt.\% in both the mantle and the lithosphere. The maximum amount of water in the mantle is fixed to 0.2 wt.\%, while maximum water content in the cylinder
-and in the lithosphere is function of pressure and temperature and is calculated using a serpentinized harzburgite with Perple\_X, as in \citet{Quinquis2014}.
-
-As shown in \citet{Quinquis2014}, the model is characterised by a progressive dehydration in the external portion of the cylinder, due to the increase of
-temperature, with a consequent vertical migration of free water that hydrates the mantle above the cylinder, up to the lithosphere. The distribution of
-bound and free water during the evolution of the experiment is shown in Fig. \ref{fig:quinquis} (left and right columns, respectively). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Hydration}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Hydration.pdf}
-\caption{Distribution of bound and free water (left and right columns, respectively) at $t=0.5$ and $t=$ \SI{1.25}{\mega\year} (first and second rows,
-respectively) for the sinking hydrated cylinder experiment. Only markers with bound or free water not equal to 0 are plotted.}
-\label{fig:quinquis}
-\end{figure}
-
-\subsection{Mantle melting curves}\label{sec:katz}
-This instantaneous experiment is performed in order to recreate isobaric melting curves, as function of temperature, bulk water and content of water in the
-melt, to compare with melting curves obtained by \citet{Katz2003}. The domain is square with $L_x=L_y=100$ km, a grid resolution of \num{100x100} elements and
-25 markers per element. Density is set in the entire domain to \SI{3300}{\kg\per\cubic\m} and temperature increases from the sides to the centre from 1000° to
-1500°C. The experiment is performed for bulk water contents from 0 to 0.3 wt.\%.
-
-Fig. \ref{fig:fraction} shows isobaric melting curves for \SI{1}{\giga\pascal} (panel a) and \SI{3}{\giga\pascal} (panel b) calculated for different bulk water
-contents. The curves well match with those obtained for the mantle by \citet{Katz2003}, showing the saturation at the solidus for bulk water of 0.2 and 0.3 wt.\%, indicating
-that the code correctly determines melt fraction as function of pressure, temperature and bulk water content. Same results can be observed in Fig.
-\ref{fig:water}a, where isobaric and isothermal melting curves are shown as function of bulk water content. In addition, Fig. \ref{fig:water}b shows that the
-code correctly determines also the percentage of water in the melt as function of pressure, temperature and melt fraction. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Melting}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Fraction.pdf}
-\caption{Isobaric melting curves for 1 GPa (panel a) and 3 GPa (panel b) as function of temperature with different bulk water contents.}
-\label{fig:fraction}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Water.pdf}
-\caption{Isobaric melting curves for 1.5 GPa as function of bulk water (panel a) and water in the melt (panel b) calculated for different temperatures.}
-\label{fig:water}
-\end{figure}
-
-\clearpage
-
-\bibliography{export}
-
-\end{document}

### 2022-09-07 tex/Benchmark/runge.tex
update main

--- 2022-03-31 FALCON_description.tex
@@
-\documentclass[hidelinks,10pt,a4paper]{article}
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{textcomp}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage{booktabs}
-\usepackage{multirow}
-\usepackage[cm]{fullpage}
-\usepackage[round]{natbib}
-\bibliographystyle{abbrvnat}
-\usepackage{doi}
-\usepackage{bm}
-\usepackage[table]{xcolor}
-\hypersetup{
-    colorlinks,
-    linkcolor={blue},
-    citecolor={blue},
-    urlcolor={blue}
-}
-\usepackage{float}
-\restylefloat{table}
-\usepackage{siunitx}
-\DeclareSIUnit\year{yr}
-
-\begin{document}
-
-\section{Introduction}\label{sec:intro}
-This document includes a description of the code FALCON, with a presentation of the features implemented followed by results of extensive benchmarks listed in
-Table \ref{tab:list}, performed to verify that the code solves correctly a wide range of problems. Description and results of all benchmarks are shown
-in Section \ref{sec:benchmark}.
-
-\begin{table}[H]
-\caption{List of benchmarks performed for each feature of the code. Data of all benchmarks can be found at 
-\url{https://github.com/aleregorda/Benchmarks}}
-\centering
-\begin{tabular}{| l | l | l |}
-\toprule
- \textbf{Feature}  & \textbf{Benchmark} & \textbf{Section} \\
-\midrule
-  Solver  & Stokes flow & \ref{sec:stokes} \\
-\hline
-  \multirow{2}{*}{Markers advection} & Zalesak disk & \ref{sec:runge} \\
-    & Conservative Interpolation Velocity & \ref{sec:CVI} \\
-\hline
-  \multirow{4}{*}{Momentum equation}  & Poiseuille flow & \ref{sec:poiseuille}\\
-    & Instantaneous 2D sphere & \ref{sec:ist_sphere} \\
-    & Rayleigh-Taylor experiment & \ref{sec:rayleigh} \\
-    & Falling block & \ref{sec:block} \\
-\hline
-    & 2D Stokes sphere & \ref{sec:sphere} \\
-    Sticky air and  & Stabilisation algorithm & \ref{sec:stab} \\
-    free surface & Topography relaxation & \ref{sec:crameri} \\
-  & Spontaneous subduction & \ref{sec:subduction} \\
-\hline
-  Erosion and sedimentation  & - & - \\
-\hline
-   Non-linear  & Slab detachment & \ref{sec:slab} \\
-   visco-plasticity & Indenter & \ref{sec:indenter} \\
-    & Brick & \ref{sec:brick} \\
-\hline
-  \multirow{3}{*}{Energy equation}  & Advection stabilisation & \ref{sec:advection} \\
-    & Simple shear heating & \ref{sec:simple_shear} \\
-    & Shear and adiabatic heating & \ref{sec:shear} \\
-\hline
-\multirow{3}{*}{Energy + momentum} & Mantle convection & \ref{sec:mantle} \\
-& Viscoplastic mantle convection & \ref{sec:visco_mantle} \\
-& Thin layer entrainment & \ref{sec:thin} \\
-\hline
-  Phase changes  & \multirow{2}{*}{Hydrated sinking cylinder} & \multirow{2}{*}{\ref{sec:quinquis}} \\
-  and hydration  &  & \\
-\hline
-  Melting & Experimental melting curves & \ref{sec:katz} \\
-\bottomrule
-\end{tabular}
-\label{tab:list}
-\end{table}
-
-The code is written in Fortran90 and uses the finite element method (FEM) with quadrilateral $Q_1 \times P_0$ elements (continuous bilinear velocity and
-discontinuous constant pressure), associated with MUMPS\footnote{\url{http://mumps.enseeiht.fr/}} \citep{Amestoy2001,Amestoy2019} that is a software package
-for solving systems of linear equations of the form $A \cdot x = b$, where \textit{A} is a square sparse matrix, by means of a direct method. Since
-$Q_1 \times P_0$ elements do not satisfy Ladyzhenskaya, Babuska and Brezzi (LBB) stability condition \citep{Donea2003}, elemental pressure is elaborated
-in post-processing to avoid spurious pressures by means of a double interpolation to smooth the pressure. Following this procedure, the elemental pressure
-is interpolated onto nodes and then back onto elements. After the smoothing procedure, the pressure field is calculated again on the nodes to be used in
-combination with the lithostatic pressure, which is calculated onto nodes. Correctness of the solution and performances of the solver in terms of time and
-memory usage are tested solving the Stokes flow with the analytical solution proposed by \citet{Donea2003} (Section \ref{sec:stokes}).
-
-The thermo-mechanics of crust-mantle systems is described by means of the conservation of mass, momentum and energy equations, expressed as follows:
-\begin{eqnarray}
-\label{eq:mass}\frac{\partial \rho}{\partial t}+\bm{\nabla} \cdot (\rho \bm{u})&=&0\\
-\label{eq:momentum}-\bm{\nabla} p + \bm{\nabla} \cdot \bm{\tau} + \rho \bm{g}&=&0\\
-\label{eq:energy}\rho C_p \left(\frac{\partial T}{\partial t} + \bm{u} \cdot \bm{\nabla} T \right)&=&\bm{\nabla} \cdot \left(k\bm{\nabla} T\right) + H_{tot}
-\end{eqnarray}
-where $\rho$ is the density, $\bm{u}$ is the velocity, \textit{p} is the pressure, $\bm{\tau}$ is the deviatoric stress, $\bm{g}$ is the gravity acceleration,
-$C_p$ is the specific heat at constant pressure, \textit{T} is the temperature, \textit{k} is the thermal conductivity and $H_{tot}$ is the total internal heat
-production. Density variations due to temperature are generally small enough to assume the density as constant ($\rho=\rho_0$) in Eq. \ref{eq:mass} and in
-Eq. \ref{eq:energy}, while it must be treated as a variable in the buoyancy term of Eq. \ref{eq:momentum}, such that \[\rho=\rho_0(1-\alpha(T-T_0))\]
-where $\rho_0$ is the density at a reference temperature $T_0$ and $\alpha$ is the coefficient of thermal expansion. Eq. \ref{eq:mass} then can be rewritten as
-\begin{equation}\label{eq:mass_0}
-\bm{\nabla} \cdot \bm{u}=0
-\end{equation}
-This simplification is known as the extended Boussinesq approximation.
-
-The time step $dt$ is chosen in according to Courant-Friedrichs-Lewy condition \citep{Anderson1995}, such that
-\[dt=C_n\; \min\left(\frac{\Delta x}{\max|\bm{u}|},\frac{(\Delta x)^2}{\kappa}\right)\]
-where $0<C_n<1$ is the Courant (CFL) number, $\Delta x$ is the minimum dimension of smallest element, $\max|\bm{u}|$ is the maximum velocity calculated on the
-entire domain and
-$\kappa=\frac{k}{\rho C_p}$ is the thermal diffusivity \citep{Thieulot2014}.
-
-\section{Lagrangian markers}\label{sec:markers}
-Elemental properties (density, viscosity, thermal conductivity, specific heat and thermal expansion) needed to solve Eqs. \ref{eq:mass_0}, \ref{eq:momentum}
-and \ref{eq:energy} are related to the composition of each element, determined by means of Lagrangian markers that are characteristic of different materials
-in the domain. Elemental properties (with the sole exception of the viscosity) are calculated using an arithmetic mean, as \[P_e=\frac{1}{n}\sum_{i=1}^n P_i\]
-where $P_e$ is the elemental property, $P_i$ is the property characteristic of the material of each marker and $n$ is the number of markers of the element.
-Differently, the average scheme for the elemental viscosity can be chosen between harmonic, geometric and arithmetic mean. Markers can be placed regularly or
-randomly at the beginning of the simulation and their advection is performed by either a 2nd-order or a 4th-order Runge-Kutta in space, interpolating the velocity 
-field on each marker by means of the shape functions (see Section \ref{sec:runge}). The interpolated velocity is then corrected by means of the Conservative
-Velocity Interpolation (CVI), which introduces a corrective term to reduce dispersion and clustering of particles in both steady state and time-dependent 
-\citep{Wang2015} (see Section \ref{sec:CVI}).
-
-The initial distribution of the markers is created using the open-source code library
-Geodynamic World Builder \citep{Fraters2019}. During the simulation, each marker carries memory of temperature, pressure and accumulated strain, which are
-determined interpolating the relative nodal parameter. In particular, pressure is interpolated onto each marker after the smoothing procedure. The number
-of markers contained in each element is maintained between $n_{min}$ and $n_{max}$. When in an element there are less markers than $n_{min}$ the code adds
-random markers to reach the $n_{min}$, while if the number is higher than $n_{max}$ some of them are randomly deleted. When new markers are added, they assume
-the properties of the nearest marker. In this way elements are never empty and maintain a number of markers inside a prefixed range.
-
-\section{The momentum equation}\label{sec:mom_eq}
-The penalty method is implemented so that Eq. \ref{eq:mass_0} can be written as:
-\begin{equation}\label{eq:mass_penalty}
-\bm{\nabla} \cdot \bm{u} + \frac{p}{\lambda}=0
-\end{equation}
-where $\lambda$ is the so-called penalty parameter, which should be 6-7 orders of magnitude larger than the shear viscosity to ensure that mass conservation
-is satisfied \citep{Donea2003,Thieulot2014}.
-
-The deviatoric stress tensor in Eq. \ref{eq:momentum} can be written in terms of the strain rate tensor as $\bm{\tau}=2\eta\dot{\bm{\epsilon}}$, with
-$\dot{\bm{\epsilon}}=\frac{1}{2}\left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)$. Therefore, Eq. \ref{eq:momentum} can be rewritten as:
-\begin{equation}\label{eq:momentum_pressure}
--\bm{\nabla} p + \bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right) + \rho \bm{g} = 0
-\end{equation}
-Finally, using pressure from Eq. \ref{eq:mass_penalty}, Eq. \ref{eq:momentum_pressure} can be rewritten as:
-\begin{equation}\label{eq:momentum_penalty}
-\lambda \bm{\nabla} \left(\bm{\nabla} \cdot \bm{v}\right)+\bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right)+\rho\bm{g}=0
-\end{equation}
-The penalty method is been associated to the iterative Uzawa method, as described in detail in \citet{Dabrowski2008} and \citet{Thieulot2014}
-(see Section \ref{sec:poiseuille}). Results of benchmarks performed to verify the correctness in the implementation of Eq. \ref{eq:momentum_penalty} are
-shown in Sections \ref{sec:ist_sphere} and \ref{sec:rayleigh}.
-
-To support large viscosities variations the penalty parameter is related to the effective elemental viscosity by means of a dimensionless coefficient, so that
-$\lambda=\lambda_e(e)\eta_{eff}(e)$ \citep{Marotta2006,Dabrowski2008,Thieulot2014}. Benchmark of the falling block \citep{Gerya2003a,Gerya2010b,Thieulot2011}
-is performed to verify that the code can correctly deal buoyancy driven flows with strong viscosity contrasts (Section \ref{sec:block}).
-
-\section{Sticky air and free surface}\label{sec:sticky}
-The Earth's surface can be treated by means of either the so-called sticky-air or a true free surface method, both of them implemented in the code. In the
-sticky-air method the surface is approximated with the introduction of a buoyant layer with a viscosity at least four orders of magnitude lower than the crust
-\citep{Schmeling2008,Crameri2012} and the interface between lithosphere and air is defined using a chain of passive markers that are advected as the Lagrangian
-markers. The correctness of the evolution of the markers chain is tested with the experiment of a 2D time-dependent Stokes sphere below a free surface and
-compared with results from ASPECT\footnote{\url{https://aspect.geodynamics.org/}} \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} (Section
-\ref{sec:sphere}). In the true free surface case the top boundary is assumed stress-free and velocities are not fixed. In this case topography variations are
-described by vertical deformations of the mesh that depend on the velocity field of the nodes that identify the free surface, while horizontal deformations are
-not taken into account. This procedure is known as the Arbitrary Lagrangian-Eulerian (ALE) method and its implementation follows the technique described in
-\citet{Thieulot2011}. However, although the implementation of a true free surface better reproduces laboratory experiment, extremely small time steps can be
-necessary to maintain stability \citep{Kaus2010a,Quinquis2011,Thieulot2014}. Therefore, the stability algorithm proposed by \citet{Kaus2010a} is implemented
-to avoid instabilities due to high density differences at the free surface when using too large time steps. The implementation of this algorithm are tested by
-performing the experiment described by \citet{Kaus2010a} (Section \ref{sec:stab}).
-
-The topography relaxation benchmark proposed by \citet{Crameri2012} is performed to verify that the code correctly recovers topography variations in case of
-both the sticky-air and the true free surface method (Section \ref{sec:crameri}). Finally, Section \ref{sec:subduction} show the results of the spontaneous
-subduction experiment described by \citet{Schmeling2008}.
-
-\section{Erosion and sedimentation}\label{sec:erosion}
-Surface processes at either the lithosphere-air interface or in correspondence of the free surface have been implemented by means of the software
-Fastscape\footnote{\url{https://fastscape.org/}} \citep{Braun2013,Cordonnier2019,Yuan2019a,Yuan2019b}. FastScapeLib is a set of routines that solve the stream
-power law (enriched by a sediment transport/deposition term) \citep{Yuan2019a}, hillslope diffusion and marine transport and deposition \citep{Yuan2019b}.
-The partial differential equation solved by FastScapeLib is:
-\begin{equation}
-\frac{\partial h}{ \partial t}=U-K_f A^{m_e}S^{n_e} + \frac{G}{A}\int_A \left( U-\frac{\partial h}{\partial t} \right) dA + K_d \bm{\nabla}^2 h 
-\end{equation}
-where $h$ is the topography, $U$ is the uplift, $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $S$ is the slope, $m_e$ and $n_e$ are
-parameters that should be chosen to constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}, $G$ is the dimensionless
-deposition/transport coefficient for the enriched stream power law and $K_d$ is the effective diffusivity. Terms $K_f A^{m_e}S^{n_e}$ and $K_d \bm{\nabla}^2 h$
-consider long-range processes, described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, and short-range processes,
-proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}.
-
-In case of sticky air, slope is determined by means of the markers chain, which is vertically corrected considering erosion and sedimentation rates. After the
-correction, continental markers above the markers chain are transformed in air markers and air markers below the markers chain become sediments. Similarly,
-slope is determined considering the nodes on the top boundary in case of true free surface and they are vertically corrected as for the markers chain. In this
-case, continental markers above the top boundary are simply deleted.
-%Surface processes at the lithosphere-air interface are implemented taking into account fluvial erosion (long-range processes), hillslope processes (short-range
-%processes) and failure of slopes exceeding a threshold angle $\theta$ \citep{Burov1997,Tucker1999,Sacek2012,Croissant2014,Ueda2015,Beucher2020}. Erosion rates
-%$\dot{E}_{lr}$ due to long-range processes can be described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{lr}=K_f A^{m_e} \bm{\nabla} h^{n_e}
-%\end{equation}
-%where $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $\bm{\nabla} h$ is the slope and $m_e$ and $n_e$ are parameters chosen in to
-%constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}. Erosion rates $\dot{E}_{sr}$ due to short-range processes are
-%implemented by assuming that transport is linearly proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{sr}=K_d \bm{\nabla}^2 h
-%\end{equation}
-%where $K_d$ is the effective diffusivity. In addition, the slope is modified if it still exceeds a threshold angle after long- and short-range processes
-%\citep{Ueda2015}.
-
-%Offshore sedimentation processes are implemented similarly to hillslope processes but with a depth-dependent coefficient \citep{Sacek2012,Beucher2020}.
-%The sedimentation rate $\dot{S}$ is calculated as
-%\begin{equation}
-%\dot{S}=K_s \bm{\nabla}^2 h
-%\end{equation}
-%with
-%\begin{equation}
-%K_s=K^*_s \exp\left(\frac{h-h_{sl}}{h_0}\right)
-%\end{equation}
-%where $K^*_s$ is a constant, $h_{sl}$ is the sea level and $h_0$ is a characteristic length. In case of sticky air, slope is determined by means of the markers
-%chain, which is vertically corrected considering erosion and sedimentation rates. After the correction, continental markers above the markers chain are
-%transformed in air markers and air markers below the markers chain become sediments. Similarly, slope is determined considering the nodes on the top boundary
-%in case of true free surface and they are vertically corrected as for the markers chain. In this case, continental markers above the top boundary are simply
-%deleted.
-
-\section{Non-linear rheologies}\label{sec:nonlinear}
-Non-linear rheologies are implemented combining viscous creep (dislocation and diffusion) and plastic yielding. For each marker, diffusion and dislocation
-viscosity ($\eta_{df|ds}$) can be determined as follows \citep{Karato1993,Warren2008b,Wang2016}:
-\begin{equation}\label{eq:viscous}
-\eta_{df|ds}=f_s \left(\frac{d^m}{A}\right)^{\frac{1}{n}} \left (\frac{2^{\frac{1-n}{n}}}{3^{\frac{n+1}{2n}}} \right)
-I_2^{\frac{(1-n)}{n}}\exp\left(\frac{Q+pV}{nRT}\right)
-\end{equation}
-where $f_s$ is a scaling factor used to represent lithologies that are stronger or weaker than the base set, $d$ is the grain size, $m$ is the grain size exponent,
-$n$ is the stress exponent, $A$ is the uniaxial pre-exponential factor, $I_2$ is the square root of the second invariant of the strain rate tensor, $Q$ is the
-activation energy, $p$ is the pressure, $V$ is the activation volume, $R$ is the gas constant and $T$ is the temperature.
-Pressures and temperatures are determined by the interpolation of the nodal parameters, while strain rates are calculated by means of the
-derivative of the velocities on the nodes. In case of diffusion creep $n=1$ and $m>0$, in case of dislocation creep $n>1$ and $m=0$. The viscous creep $\eta_{cp}$
-is then calculated as the harmonic average between $\eta_{df}$ and $\eta_{ds}$:
-\begin{equation}\label{eq:visco_creep}
-\eta_{cp}=\left(\frac{1}{\eta_{df}}+\frac{1}{\eta_{ds}}\right)^{-1}
-\end{equation}
-The implementation of the non-linear viscous creep viscosity depending on the strain rate is tested by means of the slab detachment benchmark
-(Section \ref{sec:slab}).
-
-Plastic yielding is implemented rescaling $\eta_{cp}$ in order to limit the stress below the yield stress $\sigma_y$ \citep{Thieulot2008,Thieulot2014,Glerum2018},
-obtaining
-\begin{equation}\label{eq:plastic}
-\eta_{pl}=\frac{\sigma_y}{2I_2}
-\end{equation}
-where $\eta_{pl}$ is the plastic viscosity and the yield stress is determined following the Drucker-Prager criterion, such as
-\begin{equation}\label{eq:yield}
-\sigma_y=C\;\cos(\phi)+p\;\sin(\phi)
-\end{equation}
-where $C$ is the cohesion and $\phi$ is the internal friction angle. In case of negative pressure it is imposed equal to 0, so that negative yield stress are
-excluded. The correctness of the non-linear solution in case of plastic viscosity with variable internal friction angle is verified performing indenter and
-brick experiments (Sections \ref{sec:indenter} and \ref{sec:brick}, respectively).
-
-Strain softening is taken into account for both viscous creep and plastic viscosity \citep{Huismans2003,Babeyko2005,Huismans2005,Sobolev2005,Warren2008a} by
-means of the accumulated strain $\epsilon$ memorised by each marker. Softening in the viscous creep determines a linear decrease of $\eta_{vc}$ by means of a
-viscous strain softening factor $W_S$ that increases linearly from $W_{S_0}$ to $W_{S_{\infty}}$ for $\epsilon_{S_0}<\epsilon_S<\epsilon_{S_{\infty}}$ 
-\citep{Warren2008a}. This viscous softening can be related to strain-induced grain size reduction \citep{Warren2008a}. Differently, plastic softening is 
-simulated with a linear decrease of internal friction angle $\phi(\epsilon)$ and cohesion $C(\epsilon)$ in according to
-\begin{eqnarray}
-\label{eq:friction1}\phi(\epsilon)&=&\phi_0+(\phi_{\infty}-\phi_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}\\
-\label{eq:friction2}C(\epsilon)&=&C_0+(C_{\infty}-C_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}
-\end{eqnarray}
-where $\phi_0$, $C_0$ and $\phi_{\infty}$, $C_{\infty}$ are internal friction angle and cohesion for $\epsilon_0$ and $\epsilon_{\infty}$, respectively
-\citep{Huismans2003,Huismans2005,Warren2008a,Thieulot2014}. Plastic softening approximates deformation-induced softening of faults and brittle shear zones
-\citep{Warren2008a}.
-
-Plastic yielding and viscous creep are then combined to obtain a viscoplastic viscosity $\eta_{vp}$ as follows, assuming that they are independent processes
-\citep{Karato2008,Glerum2018}:
-\begin{equation}\label{eq:viscoplastic}
-\eta_{vp}=\min(\eta_{cp},\eta_{pl})
-\end{equation}
-Finally, effective viscosity $\eta_{eff}$ is capped by the minimum and the maximum viscosity ($\eta_{min}$ and $\eta_{max}$, respectively) to avoid extremely
-low or high viscosity \citep{Glerum2018} as follows
-\begin{equation}\label{eq:effective}
-\eta_{eff}=\min(\max(\eta_{vp},\eta_{min}),\eta_{max})
-\end{equation}
-Viscous creep and plastic yielding are non-linear rheologies because of their dependence on the velocity field through pressure and strain rates. Therefore,
-the solution is determined by means of Picard-type iterations, until convergence of the velocity field \citep{Glerum2018}. The convergence is verified at each
-iteration $i$ via the nonlinear residual $\bm{\mathcal{R}}^i$ that can be determined as
-\begin{equation}\label{eq:convergence}
-\bm{\mathcal{R}}^i= \mathbb{K} (\eta_{eff}(\dot{\bm{\epsilon}}^{i-1},p^{i-1}))\cdot \bm{v}^{i-1} - \bm{f}^i
-\end{equation}
-%\begin{equation}\label{eq:convergence}
-%\chi=1-\frac{\langle (\bm{v}^i-\langle\bm{v}^i\rangle)\cdot (\bm{v}^{i+1}-\langle\bm{v}^{i+1}\rangle) \rangle}{\sqrt{\sigma^i\sigma^{i+1}}}
-%\end{equation}
-%Iterations are performed until either convergence ($\chi<tol$) or maximal number of non-linear iterations ($it_{max}$) is reached.
-where $\mathbb{K}$ is the Stiffness matrix and $\bm{f}$ is the right hand side vector \citep{Spiegelman2016,Glerum2018}.  The $L_2$-norm is extracted from
-$\bm{\mathcal{R}}^i$ and it is normalised as follows
-\begin{equation}\label{eq:convergence_res}
-\frac{||\bm{\mathcal{R}}^i||_2}{||\bm{\mathcal{R}}^0||_2}
-\end{equation}
-where $||\bm{\mathcal{R}}^0||_2$ is the $L_2$-norm of the first nonlinear iteration.
-Since often the normalised nonlinear residual drop very quickly over the first 2-3 iterations, also the $L_2$-norm of velocity and pressure residuals are
-calculated as
-\begin{equation}\label{eq:convergence_vel}
-\frac{||\Delta\bm{v}||_2}{||\bm{v}||_2}=\frac{||\bm{v}_i-\bm{v}_{i-1}||_2}{||\bm{v}_{i}||_2}
-\end{equation}
-and
-\begin{equation}\label{eq:convergence_press}
-  \frac{||\Delta p||_2}{||p||_2}=\frac{||p_i-p_{i-1}||_2}{||p_{i}||_2}
-\end{equation}
-respectively. Iterations are performed until either maximal number of non-linear iterations ($it_{max}$) is reached or all the normalised nonlinear residual, 
-the velocity residual and the pressure residual converge under a defined tolerance $tol$.
-
-\section{The energy equation}\label{sec:en_eq}
-The stabilisation of the advection term of the energy equation (Eq. \ref{eq:energy}) needed to avoid possible oscillations in the thermal solution in the case
-when advection dominates over diffusion, is implemented by means of a streamline-upwind Petrov–Galerkin (SUPG) method, for which the advection term is modified
-follows the discussion in \citet{Thieulot2011} and \citet{Thieulot2014} (see Section \ref{sec:advection} for benchmark). The total internal heat production
-$H_{tot}$ in Eq. \ref{eq:energy} includes radiogenic heating $H_r$, shear heating $H_s$ and adiabatic heating $H_a$, where
-\begin{equation}\label{eq:shear}
-H_s=2\eta \dot{\bm{\epsilon}} : \dot{\bm{\epsilon}} = 2\eta(\dot{\epsilon}_{xx}^2+\dot{\epsilon}_{yy}^2+2\dot{\epsilon}_{xy}^2)
-\end{equation}
-and
-\begin{equation}\label{eq:adiabatic}
-H_a=T \alpha \frac{Dp}{Dt} \approx -\alpha T \rho g_y v
-\end{equation}
-The correctness of the implementation and the computation of shear and adiabatic heating is verified performing an exercise with an analytical solution
-(Section \ref{sec:simple_shear}) and Exercise 9.4 in \citet{Gerya2010b} (Section \ref{sec:shear}). Furthermore, mantle convection and thin layer entrainment
-benchmarks (Sections \ref{sec:mantle} and \ref{sec:thin}, respectively) are performed to verify that the code solve correctly Eqs. \ref{eq:momentum_penalty}
-and \ref{eq:energy} in case of a temperature field characterised by an initial perturbation and with a temperature-dependent density. At last, the viscoplastic
-mantle convection benchmark is performed to check also the viscoplasticity in case of a initial perturbed temperature field (Section \ref{sec:visco_mantle}).
-
-\section{Phase transitions and hydration}\label{sec:phase}
-Variations of effective density, specific heat and coefficient of thermal expansion during the evolution of crustal and mantle materials at different
-pressure-temperature ($p$-$T$) conditions are computed by means of version 6.8.6 of Perple\_X software package \citep{Connolly2005}, similarly to the
-implementation described in \citet{Marotta2020}. By default, mineral assemblages and properties of each lithology (oceanic crust, lower and upper continental
-crust, sediments and dry or hydrated mantle) are calculated for temperatures from \SIrange{330}{1600}{\kelvin}, with increments of approximately \SI{4}{\kelvin},
-and pressures from \SIrange{0.1}{30}{\giga\pascal}, with increments of approximately \SI{0.03}{\giga\pascal}, for a total of almost 100000 points.
-
-The introduction of phase transitions produces two main effects that must be taken into account in numerical models: variations in density and release/absorption
-of latent heat \citep{Tackley2010}. Effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into
-account considering effective coefficient of thermal expansion \citep{Christensen1985,Zhong2015}. Similarly, effects of latent heat in the energy equation
-(eq \ref{eq:energy}) are taken into account by considering effective specific heat and coefficient of thermal expansion
-\citep{Christensen1985,Tackley2010,Zhong2015}. Effective density, specific heat and coefficient of thermal expansion are included in Perple\_X files and are
-assumed by each Lagrangian marker in according to its $p$-$T$ conditions.
-
-Hydration processes related to dehydration of subducting lithosphere \citep{Schmidt1998,Liu2007,Faccenda2009,Faccenda2010,Faccenda2014,Rosas2016} strongly
-influence the thermo-mechanics inside the mantle wedge, mainly because of both weakening effects on the mantle rheology and density variations in case of
-mantle serpentinization \citep{Gerya2002,Honda2003,Arcay2005,Roda2010,Regorda2017}. In the case that hydration is switched on, the amount of bound and free
-water is memorised by each marker, following the implementation of \citet{Quinquis2014}; hydration and dehydration processes are related to the amount of bound
-water of each marker and to the maximum amount of water it can transport, i.e. if the amount of bound water exceeds the maximum amount of water, the marker
-dehydrates and releases free water that can hydrates under-saturated markers. Maximum water content of each marker is determined as function of lithology and
-$p$-$T$ conditions and it is calculated using Perple\_X, in the same way than effective density, specific heat and coefficient of thermal expansion. Bound water
-is advected together with the markers, neglecting the effect of bound water diffusion, while free water simply migrates vertically and is not coupled to the
-solid-phase flow of the mantle wedge \citep{Arcay2005,Quinquis2014}. The correctness of the migration scheme of the free water and its absorption into bound
-water by under-saturated markers are verified performing the experiment described in \citet{Quinquis2014} (Sec. \ref{sec:quinquis}).
-
-The mantle viscosity weakening related to the amount of water has been extensively studied \citep{Chopra1981,Hirth2003} and it is implemented in according to
-\citet{Arcay2005} and \citet{Horiuchi2016}, as follows
-\begin{equation}\label{eq:wet_visc}
-\eta_{wet}=\eta_{dry}\left(\left[1-\frac{1}{f_v}\right]\exp\left(-\frac{[OH^-]}{[OH^-]_0}\right)+\frac{1}{f_v}\right)
-\end{equation}
-where $f_v$ is the viscosity reduction factor between dry and wet conditions, $[OH^-]$ is the water content and $[OH^-]_0$ is a reference water content set to
-620 ppm (0.062 wt.\%) \citep{Arcay2005}. By default, $f_v$ is set to 100, in according to \citet{Arcay2005} and \citet{Horiuchi2016}.
-
-\section{Mantle melting}\label{sec:melt}
-Melt fraction in the mantle depends on temperature ($T$, in \textdegree C), pressure ($p$, in \si{\giga\pascal}) and the water content in the melt ($X_{H_2O}$, in wt.\%).
-The determination of the melt fraction $M$ is implemented as explained by \citet{Katz2003}, according to the successive modification by \citet{Langmuir2006} and
-\citet{Kelley2010}, as follows
-\begin{eqnarray}\label{eq:melt_fr}
-M(p,T,X_{H_2O})&=&-T+T_{sol}+(x_t\cdot ln(p+y_t))\cdot X_{H_2O}-\nonumber\\
-&&+K_m\left(\frac{X_{H_2O}^{bulk}}{D_{H_2O}(1-X_{H_2O})+X_{H_2O}}\right)^{\gamma}
-\end{eqnarray}
-where $T_{sol}$ is the temperature of the dry solidus, $D_{H_2O}$ is the partition coefficient and $x_t\cdot log(p+y_t)$ indicates the pressure dependence of
-the melting curve \citep{Kelley2010}. $x_t$, $y_t$, $K_m$ and $\gamma$ are parameters chosen in according to \citet{Kelley2010}.
-However, $X_{H_2O}$ depends on the melt fraction as
-\begin{equation}\label{eq:melt_wt}
-X_{H_2O}(M)=\frac{X_{H_2O}^{bulk}}{D_{H_2O}+M(1-D_{H_2O})}
-\end{equation}
-and a numerical solution can be found using a root-finder method \citep{Katz2003,Wang2016}. Furthermore, the content of water in the melt is limited by the
-pressure-dependent saturation concentration of water in the melt, determined as
-\begin{equation}\label{eq:melt_max}
-X_{H_2O}^{sat}=\chi_1 p^{\lambda_m}+\chi_2 p
-\end{equation}
-where $\chi_1$, $\chi_2$ and $\lambda_m$ are parameters chosen according to \citet{Katz2003}, \citet{Langmuir2006} and \citet{Kelley2010}. The correctness in
-the determination of melt fractions and water contents in the melt are tested comparing model results with experimental curves obtained by \citet{Katz2003}
-(Sec. \ref{sec:katz}).
-
-In case of melting, viscous creep viscosity ($\eta_{vp}$) is modified according to \citet{Wang2016}, as
-\begin{equation}\label{eq:melt_visc}
-\eta_{melt}=\eta_{vp}\cdot \exp(\alpha_m M)
-\end{equation}
-where $\alpha_m$ are melt fraction factors for dislocation and diffusion creep, chosen in according to \citet{Wang2016}.
-
-Effective density ($\rho_{eff}$) for partially molten rocks is calculated as
-\begin{equation}\label{eq:melt_dens}
-\rho_{eff}=\rho_s(1-M)+\rho_m M
-\end{equation}
-where $\rho_s$ and $\rho_m$ are the densities of the solid and the molten rock, respectively \citep{Gerya2004b,Gerya2007,Wang2016}. The density of the
-solid rock is extracted by Perple\_X, while the density of the molten portion is calculated as
-\begin{equation}\label{eq:melt_dens_M}
-\rho_m(p,T)=\rho_0[1-\alpha(T-T_0)][1+\beta(p-p_0)]
-\end{equation}
-where $\rho_0$ is the density at the reference temperature ($T_0$) and pressure($p_0$), and $\alpha$ and $\beta$ are the thermal expansion and compressibility
-coefficients, respectively \citep{Gerya2004b,Gerya2007,Wang2016}.
-
-As for the phase transitions, effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into account
-considering effective coefficient of thermal expansion. Similarly, latent heat due to melting/crystallisation can be implicitly included in the energy equation
-(Eq. \ref{eq:energy}) by considering effective specific heat ($C_{p_{eff}}$) and thermal expansion ($\alpha_{eff}$) of the partially molten rocks, in according to
-\begin{eqnarray}
-\label{eq:cp_effective}C_{p_{eff}}&=&C_p+H_L\left(\frac{\partial M}{\partial T}\right)_p \\
-\label{eq:alpha_effective}\alpha_{eff}&=&\alpha+\rho\frac{H_L}{T}\left(\frac{\partial M}{\partial p}\right)_T
-\end{eqnarray}
-where $C_p$ and $\alpha$ are the specific heat and the coefficient of thermal expansion of the solid rock, respectively, and $H_L$ is the latent heat of the
-molten rock \citep{Gerya2004b,Gerya2007,Tackley2010}.
-
-\section{Benchmarks}\label{sec:benchmark}
-\subsection{Error measurements}\label{sec:error}
-In order to determine the accuracy of velocity and pressure field of the benchmarks, the $L_2$-norm is computed by numerical integration on the Gauss-Legendre
-quadrature points. $L_2$-norm for pressure and velocity errors can be evaluated as
-\begin{eqnarray}
-\textrm{err}_p&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|p^n_i(r_q)-p^a_i(r_q)|^2\right)w_q|J_q|}\nonumber \\
-\textrm{err}_v&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|u^n_i(r_q)-u^a_i(r_q)|^2+|v^n_i(r_q)-v^a_i(r_q)|^2\right)w_q|J_q|}\nonumber
-\end{eqnarray}
-respectively, where $n_e$ is the number of elements, $n_q$ is the number of quadrature points per element, $p^n_i(r_q)$ and $p^a_i(r_q)$ are the numerical and
-analytical pressure, respectively, in each quadrature point $q$, $w_q$  and $J_q$ are the weight and the Jacobian at the quadrature point
-$q$, $u^n_i(r_q)$, $v^n_i(r_q)$, $u^a_i(r_q)$ and $v^a_i(r_q)$ are the numerical and analytical velocities, respectively, in each quadrature point $q$.
-
-Other quantities used as comparison with original benchmarks are the root-mean-square velocity over the whole domain and over the surface
-\begin{eqnarray}
-v_{\textrm{rms}}&=&\sqrt{\int_{\Omega}|\bm{v}|^2 d\Omega}\nonumber \\
-v_{\textrm{rms}}^{top}&=&\sqrt{\int_0^1 u^2 \bigg|_{y=1} dx}\nonumber
-\end{eqnarray}
-respectively, top and bottom Nusselt numbers
-\[\textrm{Nu}^{top/bottom}=-\int_0^1\frac{\partial T}{\partial y} \bigg|_{y=1/y=0} dx\]
-the average rate of work done against gravity
-\[\left\langle W \right\rangle=\int_{\Omega}T u_y d\Omega\]
-the average rate of viscous dissipation
-\[\left\langle \Phi \right\rangle=\int_{\Omega}\tau_{ij}\dot{\epsilon}_{ij} d\Omega\]
-and the percentage error between $\left\langle W \right\rangle$ and $\left\langle \Phi \right\rangle/Ra$
-\[\delta=\frac{\left|\left\langle W \right\rangle -\frac{\left\langle \Phi \right\rangle}{Ra} \right|}{\max \left(\left\langle W \right\rangle,
-\frac{\left\langle \Phi \right\rangle}{Ra}\right)}\times100\]
-
-\subsection{Stokes flow}\label{sec:stokes}
-The problem consists of determining velocity field $(u,v)$ and pressure $p$ in case of a manufactured solution with prescribed body forces such as
-\begin{eqnarray}
-b_1&=&(12-24y)x^4+(-24+48y)x^3+(-48y+72y^2-48y^3+12)x^2+\nonumber \\
-&&+(-2+24y-72y^2+48y^3)x+1-4y+12y^2-8y^3\nonumber \\
-b_2&=&(8-48y+48y^2)x^3+(-12+72y-72y^2)x^2+(4-24y+48y^2-48y^3+\nonumber \\
-&&+24y^4)x-12y^2+24y^3-12y^4\nonumber
-\end{eqnarray}
-for which the exact solution is:
-\begin{eqnarray}
-u(x,y)&=&x^2(1-x)^2(2y-6y^2+4y^3)\nonumber \\
-v(x,y)&=&-y^2(1-y)^2(2x-6x^2+4x^3)\nonumber \\
-p(x,y)&=&x(1-x)-\frac{1}{6}\nonumber
-\end{eqnarray}
-The domain is a unit square a constant viscosity ($\eta=1$) and the penalty parameter is set to $\lambda=10^7$. Velocity boundary conditions are set to no slip
-$(\bm{v}=\bm{0})$ on all boundaries. The problem is performed for different grid resolution between \num{8x8} and \num{1024x1024} elements.
-
-The errors between the analytical solution and the numerical prediction of the velocity and the pressure field are calculated as in Section \ref{sec:error}.
-Fig. \ref{fig:errors}a shows that both velocity and pressure field converge to the exact solution with the decrease of the element size, following the
-theoretical convergence rate. The convergence of pressure, in contrast with the observation by \citet{Donea2003} for $Q_1 \times P_0$ elements, indicate the
-effectiveness of the smoothing procedure. Smoothed pressure field for grid resolution of \num{128x128} elements is shown in Fig. \ref{fig:errors}b (continuous
-red line), in comparison with the analytical solution (dashed blue line). Solve times and memory usage needed to generate the solution are shown in
-Fig. \ref{fig:MUMPS}. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Solver/Stokes_Flow}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/errors.pdf}
-\caption{Velocity and pressure error for the Stokes flow experiment between generated and analytical solution as a function of element size (panel a),
-and comparison between smoothed pressure and analytical solution as function of $x$ coordinate for a grid resolution of \num{128x128} elements (panel b).}
-\label{fig:errors}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/MUMPS.pdf}
-\caption{Panel a) Analysis, factorisation, solution and total solve times (panel a) and factorisation memory usage (panel b) as a function of the total number
-of degrees of freedom for the Stokes flow experiment.}
-\label{fig:MUMPS}
-\end{figure}
-
 \subsection{Markers advection}\label{sec:runge}
 The 2nd-order Runge-Kutta advection scheme in space is tested by means of the Zalesak disk test \citep{Zalesak1979,Thieulot2014}. The benchmark is performed
-in a unit square domain with a grid resolution of \num{32x32} elements and 1000000 markers and values of Courant number between 0.25 and 1.
+in a unit square domain with a grid resolution of $32\times32$ elements and 1000000 markers and values of Courant number between 0.25 and 1.
 At $t=0$, the disk is centred at position $(0.5;0.75)$ with a radius $R=0.15$ and has a vertical fissure 0.05 wide and 0.2 high (orange in Fig. \ref{fig:runge})
@@
 \end{figure}
-
-\subsection{Conservative Velocity Interpolation (CVI)}\label{sec:CVI}
-The CVI correction is checked by means of the Stokes flow experiment by \citet{Donea2003}, as presented in \ref{sec:stokes}. The advection of Lagrangian
-markers is performed using either a 2nd-order or a 4th-order Runge-Kutta scheme with CFL=0.5 and an initial random distribution of 25 markers per element.
-Fig. \ref{fig:CVI} shows the minimum and maximum values of markers per element throughout the experiment with and without the CVI correction and using
-different Runge-Kutta schemes. The introduction of the CVI correction allows the
-tests to run for times longer than $t=2000$ without show empty elements, while tests without the CVI correction stop at approximately $t=600$, when at least
-one element goes empty. Fig. \ref{fig:CVI_time} shows the different number of markers per element and the markers distribution at $t=600$ obtained without and
-with the CVI correction (panels b and c, respectively) in case of a 2nd-order Runge-Kutta scheme. All data can be found at
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/CVI}.
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/CVI.pdf}
-\caption{Maximum and minimum number of markers per element throughout the experiment, with (green and orange lines) or without (purple and blue lines)
-the CVI correction and using either a 2nd-order (blue and orange lines) or a 4th-order (purple and green lines) Runge-Kutta scheme.}
-\label{fig:CVI}
-\end{figure}
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/CVI_time.pdf}
-\caption{Number of markers per element and markers distribution through time using 2nd.order Runge-Kutta scheme.
-Panels a: number of markers per element (panel a1) and distribution of the markers (panel a2) after the first
-iteration; panels b and c: comparison between number of markers per element (panels b1 and c1) and markers distribution (panels b2 and c2) without and with
-the CVI correction at $t=600$; panels d: number of markers per element (panel d1) and markers distribution (panel d2) in case of CVI correction at $t=2000$.}
-\label{fig:CVI_time}
-\end{figure}
-
-\subsection{Poiseuille flow}\label{sec:poiseuille}
-The domain is a rectangle with $L_x=2$ and $L_y=1$ and constant density and viscosity ($\rho=1$ and $\eta=1$), gravity acceleration $\bm{g}=\bm{0}$ and penalty
-parameter $\lambda=10^8$. The grid is composed by \num{40x20} elements. Velocity boundary conditions are set to no slip $(\bm{v}=\bm{0})$ at the top and the
-bottom, and a parabolic profile is imposed on the sides, with $u=y(1-y)$ and $v=0$. The analytical solution is then given by:
-\begin{eqnarray}
-u(x,y)&=&y(1-y)\nonumber \\
-v(x,y,)&=&0\nonumber \\
-p(x,y)&=&2\eta\left(\frac{L_x}{2}-x\right)\nonumber
-\end{eqnarray}
-
-The velocity field predicted by the model follows the expected parabolic profile (Fig. \ref{fig:poiseuille}a and Fig. \ref{fig:poi_plot}a) and, in case of the
-classic penalty method with no iterations, the pressure is clearly related to the divergence by means of the penalty parameter (Fig. \ref{fig:poiseuille}b and
-c and Fig. \ref{fig:poi_plot}b). Fig. \ref{fig:poiseuille}d shows that one Uzawa iteration is sufficient to bring the divergence down to \num{e-15}, with no
-correlation with the pressure field. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Poiseuille%20Flow}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Poiseuille.pdf}
-\caption{Velocity field (panel a), pressure (panel b) and divergence velocity of a Poiseuille flow in case of the classic penalty method (no iterations) and
-after one Uzawa iteration (panel c and d, respectively).}
-\label{fig:poiseuille}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Analytical.pdf}
-\caption{Velocity (panel a) and pressure (panel b) field predicted by the model for a Poiseuille flow with respect with their analytical solutions. The
-velocity field is plotted as function of the vertical coordinate in $L_x/2$ and the pressure is plotted as function of the x coordinate in $L_y/2$.}
-\label{fig:poi_plot}
-\end{figure}
-
-\subsection{Instantaneous 2D sphere}\label{sec:ist_sphere}
-The domain is a unit square with gravity $\bm{g}=(0,-1)$. The fluid has constant density and viscosity ($\rho_f=1$ and $\eta_f=1$). The sphere is in the middle
-of the domain with a radius R = 0.123456798 and has constant density and viscosity ($\rho_s=10^{-2}+\rho_f$ and $\eta_s=10^3 \cdot \eta_f$).
-We distinguish three different types of velocity boundary conditions:
-\begin{enumerate}
-\item FS: free slip conditions on all sides.
-\item NS: no slip conditions on all sides.
-\item OT: free slip conditions on the sides and the bottom, and open boundary at the top.
-\end{enumerate}
-All three cases are performed for different grid resolutions, between \num{16x16} and \num{512x512} elements with 50 randomly distributed markers per element,
-and different average schemes for the viscosity (harmonic, geometric and arithmetic). The velocity in the centre of the sphere $\bm{v}(0.5;0.5)$), minimum and
-maximum velocities ($u_{min}$, $u_{max}$, $v_{min}$, $v_{max}$) and pressures ($p_{min}$, $p_{max}$) on the entire domain, $v_{\textrm{rms}}$ and average pressure on the
-entire domain ($p_{avg}$) are compared with solutions generated by ASPECT \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and all results
-can be found at \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere2D} and 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Instantaneous_2D_Sphere}. Results in terms of the $v_{\textrm{rms}}$ for NS model 
-are also shown in Fig. \ref{fig:inst_sphere}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_NS.pdf}
-\caption{$v_{\textrm{rms}}$ as function of element size for the instantaneous 2D sphere experiment in case of no slip boundary conditions and with different average
-schemes for the viscosity. Results are compared with results obtained by other numerical codes.}
-\label{fig:inst_sphere}
-\end{figure}
-
-\subsection{Rayleigh-Taylor instability}\label{sec:rayleigh}
-This problem has been originally presented by \citet{vanKeken1997} and here it is performed as the isoviscous case (Case 1a). The domain has $L_x=0.9142$ and
-$L_y=1$ and gravity $\bm{g}=(0,-1)$. Two fluids with same constant viscosities ($\eta_1=\eta_2=1$) and different densities ($\rho_1=1000$ and $\rho_2=1010$),
-with the lighter fluid at the bottom. The initial interface between the fluids is given by $y(x)=0.2+0.02 \cos \left(\frac{\pi x}{L_x}\right)$. The experiment
-is performed with different grid sizes (\num{50x50}, \num{80x80}, \num{100x100} and \num{256x256}). A total of 1960000 markers are randomly distributed at the
-beginning of the simulation. Velocity boundary conditions are set to no slip at the top and the bottom, and to free slip at the sides of the domain.
-
-$v_{\textrm{rms}}$ as function of time is reported in Fig. \ref{fig:RT}, matching well with results shown by \citet{vanKeken1997}, \citet{Tackley2003} and
-\citet{Thieulot2014}. Fig. \ref{fig:rayleigh} shows the evolution of the experiment at different time steps. All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Rayleigh_Taylor_experiment/ISOVISCOUS}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/RT.pdf}
-\caption{$v_{\textrm{rms}}$ of the Rayleigh-Taylor experiment as function of time for different resolution of the grid.}
-\label{fig:RT}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=300px]{./Figures/Rayleigh.pdf}
-\caption{Evolution of the Rayleigh-Taylor experiment for a grid of \num{256x256} elements at $t=0, 500, 1000$ and $2000$ (panels a, b, c and d, respectively).}
-\label{fig:rayleigh}
-\end{figure}
-
-\subsection{Falling blocks}\label{sec:block}
-This benchmark is described as proposed by \citet{Gerya2003a}, \citet{Gerya2010b} and \citet{Thieulot2011}. The domain is square with $L_x=L_y=$ \SI{500}{\km}
-and the grid is composed by \num{50x50} elements with 25 markers in each element. The block is initially centred at ($x=$ \SI{250}{\km}; $y=$ \SI{400}{\km})
-and has a size of \num{100x100} km (Fig. \ref{fig:block}a). The fluid surrounding the block has $\eta_f=$ \SI{e21}{\pascal\per\s} and $\rho_f=$
-\SI{3200}{\kg\per\cubic\m}. The benchmark tests with different viscosities of the block, with $\eta_b$ from \SIrange{e15}{e27}{\pascal\per\s}. In each
-experiment also the density $\rho_b$ of the block varies from \SIrange{3220}{9900}{\kg\per\cubic\m}. Velocity boundary conditions are set to free slip
-conditions on all sides of the domain. Density distributions at $t=$ \SI{20}{\mega\year} for $\rho_b=$ \SI{3300}{\kg\per\cubic\m} and different $\eta_b$
-are plotted in Fig. \ref{fig:block}b-f, showing that the code correctly preserve the block geometry in case of large viscosity contrast, with a block stiffer
-than the surrounding fluid (Fig. \ref{fig:block}f).
-
-The velocity in the centre of the falling block at $t=0$ is measured for all experiments and, since the velocity should increase with density contrast, the
-quantity $\bm{v}/(\rho_b-\rho_f)$ is plotted as function of the viscosity contrast. All results perfectly match with those from \citet{Gerya2010b} and they
-line up on a single curve, demonstrating that the code can correctly deal with large viscosity and density contrasts (Fig. \ref{fig:falling}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Falling%20blocks}.
-
-\begin{figure}
-\noindent\includegraphics[width=300px]{./Figures/Block.pdf}
-\centering
-\caption{Density evolution of the falling block experiment at $t=$ \SI{0}{\mega\year} (panel a) and $t=$ \SI{20}{\mega\year} for different viscosities of the
-block (panels b-f).}
-\label{fig:block}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Falling.pdf}
-\caption{Initial velocity relative to the density contrast at the centre of the falling block as function of the viscosity contrast between the block and the
-surrounding fluid.}
-\label{fig:falling}
-\end{figure}
-
-\subsection{2D time-dependent Stokes sphere with deformable free surface}\label{sec:sphere}
-This experiment is performed in a unit square domain with the gravity acceleration fixed to $g_y=-1$. The sphere has $\rho_s=2$ and $\eta_s=10^3$ and is
-initially centred at ($0.5;0.6$) with radius $R=0.123456789$. The fluid surrounding the sphere has $\rho_f=1$ and $\eta_f=1$ and occupies the domain for
-$y\leq 0.75$, while for $y>0.75$ the air has $\rho_a=0$ and $\eta_a=10^{-3}$. Velocity boundary conditions are set to free slip on all sides. The Courant
-number is set to 0.25. The experiments run for 200 s using grid resolutions from \num{150x150} to \num{512x512} elements, with 25 randomly distributed markers
-per element, and for different average schemes for the viscosity. The interface between the fluid and the air is tracked by means of the markers chain. Results
-in terms of velocity and pressure fields and topography variations are compared with results obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and they can be found at
-\url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere_fs2D} and
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Time_dependent_sphere}. Results in terms of the $v_{\textrm{rms}}$ in case of 
-an arithmetic average are also shown in Fig. \ref{fig:inst_spherefs}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_arithm.pdf}
-\caption{$v_{\textrm{rms}}$ for different grid resolutions as function of time for the 2D time-dependent Stokes sphere experiment in case of an arithmetic average.
-Results are compared with results obtained by ASPECT.}
-\label{fig:inst_spherefs}
-\end{figure}
-
-\subsection{Free surface stabilisation}\label{sec:stab}
-The benchmark is performed as discussed in \citet{Kaus2010a} and \citet{Thieulot2014}. The domain is square with $L_x=L_y=$ \SI{500}{\km} and a grid resolution
-of \num{200x200} elements, each of them containing 16 markers. A buoyant fluid with $\rho_1=$ \SI{3200}{\kg\per\cubic\metre} and $\eta_1=$ \SI{e20}{\pascal\s}
-is overlain by a denser fluid with $\rho_2=$ \SI{3300}{\kg\per\cubic\metre} and $\eta_2=$ \SI{e21}{\pascal\s}. The initial interface between the fluids has an
-initial sinusoidal shape of 5 km amplitude. Velocity boundary conditions are set to free slip conditions at the sides and no slip at the bottom, while the top
-is free surface. The experiment is performed with various fixed time steps, without the use of the Courant number. The vertical position of the free surface at
-$x=L_x$ is tracked for each simulation, with or without the stabilisation algorithm.
-
-The results show an instability (drunken sailor effect) increasing the time step in case that the stabilisation algorithm is not activated
-($dt=$ \SI{4200}{\year}, green line in Fig. \ref{fig:stab}), as already observed by \citet{Kaus2010a} and \citet{Thieulot2014}. Activating the algorithm the
-instability is fixed and the simulation remains stable using time step up to \SI{50000}{\year} (red line in Fig. \ref{fig:stab}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Stabilisation_algorithm}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Kaus_random.pdf}
-\caption{Evolution of the y coordinate in $x=L_x$ as function of time for different time steps, with and without the stabilisation algorithm.}
-\label{fig:stab}
-\end{figure}
-
-\subsection{Topography relaxation}\label{sec:crameri}
-This experiment is performed as in \citet{Crameri2012} using free surface and sticky air (Case 1). The domain is rectangular with $L_x=$ \SI{2800}{\km} and
-$L_y=$ \SI{700}{\km} in case of free surface and $L_y=$ \SI{800}{\km} in case of sticky air ($h_{st}=$ \SI{100}{\km}). The grid is composed by \num{256x64}
-elements for the free surface \num{560x320} elements for the sticky air. For both cases, there is a mantle of 600 km thickness with $\eta_m=$ \SI{e21}{\pascal\s}
-overlain by a highly viscous cosine-shaped lithosphere with thickness between 93 and 107 km and $\eta_l=$ \SI{e23}{\pascal\s}. Both mediums have $\rho=$
-\SI{3300}{\kg\per\cubic\metre}. In case of sticky air experiment, the air has $\rho_a=$ \SI{0}{\kg\per\cubic\metre} and $\eta_a=$ \SI{e18}{\pascal\s}. The
-gravity acceleration is set to $g_y=$ \SI{-10}{\metre\per\square\s}. Velocity boundary conditions are set to no slip at the bottom and free slip at the sides
-of the domain. In case of sticky air, velocities on top are set to free slip conditions.
-
-The maximum topography as function of time ($h(t)$) can be analytically derived in according to \[h(t)=h_0 \exp(\gamma t)\]
-where $h_0=7$ km is the initial topography, $\gamma=\num{-0.2139e-11}$ is the characteristic relaxation rate and $t$ is the time (black line in Fig.
-\ref{fig:crameri}). Maximum topography at the characteristic relaxation time $t_{rlx}=$ \SI{14.825}{\kilo\year} can be found to be $h_{rlx}=$ \SI{2576}{\m}.
-The results show that a viscosity of the air of \SI{e18}{\pascal\s} is needed in case of $h_{st}=$ \SI{100}{\km} to correctly describe the topography relaxation
-for this problem, as pointed out by \citet{Crameri2012} (purple line in Fig. \ref{fig:crameri}). The case with a true free surface well follows the analytical
-solution (orange line in Fig. \ref{fig:crameri}). In particular, at the relaxation time in case of the true free surface predicts a maximum topography of
-\SI{2570}{\m}, with an error of \SI{6}{\m} with respect to the analytical solution. Results are also compared with results obtained with a free surface in
-UNDERWORLD (\num{256x64} elements) and SULEC (\num{401x201} elements) (with courtesy from \citealp{Crameri2012}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Topography_relaxation}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Crameri.pdf}
-\caption{Maximum topography as function of time for the Crameri benchmark (coloured lines) in comparison with the analytical solution (black line) and results
-obtained with UNDERWORLD (yellow line) and SULEC (blue line).}
-\label{fig:crameri}
-\end{figure}
-
-\subsection{Spontaneous subduction}\label{sec:subduction}
-This experiment is performed as presented by \citet{Schmeling2008}, in case of both a sticky air and a true free surface and using different average schemes
-for the viscosities. The domain is rectangular with $L_x=$ \SI{3000}{\km} and $L_y=$ \SI{700}{\km}, a grid resolution of \num{256x64} elements with 50 markers
-per element and Courant number of 0.01. At the beginning of the simulation, a 100 km-thick lithospheric layer with $\rho_m=$ \SI{3300}{\kg\per\cubic\metre}
-and $\eta_m=$ \SI{e23}{\pascal\s} is located at the top of the domain for $x$ between 1000 and 3000 km, while the asthenospheric mantle has $\rho_m=$
-\SI{3200}{\pascal\s} and $\eta_m=$ \SI{e21}{\pascal\s}. In addition, a 200 km-depth lithospheric slab is already subducted in the mantle in order to have
-a spontaneous subduction. In case of sticky air, $L_y=750$ km and a 50 km-thick air layer with $\eta_a$ from \SIrange{e19}{e21}{\pascal\s} overlie the mantle.
-Velocity boundary conditions are set to free slip on the sides and on the bottom of the domain. In case of sticky air, velocities on the top boundary are set
-to free slip conditions as well.
-
-The maximum depth of the slab is tracked for all the simulations (Fig. \ref{fig:subduction}) and they are compared with results obtained by
-\citet{Schmeling2008} with models with regular grids and comparable grid resolutions (rectangles in Fig. \ref{fig:subduction}). Low viscous sticky air models
-(i.e., \SI{e19}{\pascal\s}) enlighten a strong dependence of the sinking velocity with the chosen average scheme (dashed lines in Fig. \ref{fig:subduction}),
-while high viscous sticky air models (i.e., \SI{e21}{\pascal\s}) show a higher resistance at the trench, underestimating the correct solution (dotted lines in
-Fig. \ref{fig:subduction}). A high resistance at the trench can be also observed in case of a true free surface for the low resolution grid used (continuous
-lines in Fig. \ref{fig:subduction}), as pointed out by \citet{Schmeling2008}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Spontaneous_subduction}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Subduction.pdf}
-\caption{Maximum depth of the slab as function of time for all the simulations of the \citet{Schmeling2008} benchmark, in case of harmonic, geometric and
-arithmetic means (red, green and blue lines, respectively) using a sticky air or a true free surface (discontinuous and continuous lines, respectively).
-Red, green and blue rectangular areas indicate the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{500}{\km} in case of sticky air,
-while the grey rectangular area indicates the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{600}{\km} in case of true free surface.}
-\label{fig:subduction}
-\end{figure}
-
-\subsection{The slab detachment}\label{sec:slab}
-The slab detachment benchmark is performed as described by \citet{Schmalholz2011} and \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{1000}{\km}
-and $L_y=$ \SI{660}{\km} and a grid resolution of \num{256x256} elements. A non-linear viscous T-shaped layer with $\rho_l=$ \SI{3300}{\kg\per\cubic\metre} is
-placed at the top of the domain and surrounded by a linear viscous fluid with $\rho_f=$ \SI{3150}{\kg\per\cubic\m}. The top layer is 80 km-thick and a 250
-km-long and 80 km-wide slab is placed at $x=L_x/2$ (Fig. \ref{fig:slab}a). The effective viscosity of the top layer is given by
-\[\eta_{eff}=\eta_0 I_2^{\frac{(1-n)}{n}}\]
-with $\eta_0=$ \SI{4.75e11}{\pascal\s} and $n=4$, while the surrounding fluid has $\eta_f=$ \SI{e21}{\pascal\s}. The effective viscosity is capped between
-$\eta_{min}=$ \SI{e21}{\pascal\s} and $\eta_{max}=$ \SI{e25}{\pascal\s}. Velocity boundary conditions are set to free slip at the top and the bottom, and to
-no slip at the sides of the domain.
-
-The sinking of the slab determines a decrease of the effective viscosity in the T-shaped layer during the first part of the evolution (Fig. \ref{fig:slab}a),
-as consequence of local strain rates, while effective viscosities increase up to $\eta_{max}$ in the top layer after approximately \SI{20}{\mega\year}, when
-necking is complete (Fig. \ref{fig:slab}b).
-
-The width of necking is tracked by means of markers position at the side of the slab. Necking width and time are normalised in function of the initial slab
-(\SI{80}{\km}) and the characteristic time ($t_c=$ \SI{7.1158e14}{\s}) \citep{Schmalholz2011,Glerum2018}. The results in case of arithmetic, geometric and
-harmonic mean of the viscosity are shown in Fig. \ref{fig:necking}, compared with results from \citet{Glerum2018} for which an infinity norm mean was used.
-The influence of the average schemes agrees with results shown in Section \ref{sec:subduction}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Slab_detachment}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Slab.pdf}
-\caption{Effective viscosity for the slab detachment benchmark at the beginning of the evolution (panel a) and when necking is complete (panel b).}
-\label{fig:slab}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Necking.pdf}
-\caption{Normalised width of the necking of the slab detachment benchmark as function of normalised time.}
-\label{fig:necking}
-\end{figure}
-
-\subsection{The indenter experiment}\label{sec:indenter}
-The indenter benchmark simulate a rigid punch on a purely plastic von Mises material, for which exists an analytical solution
-\citep{Thieulot2008,Thieulot2014,Glerum2018}. The domain is a unit square with a grid resolution of \num{256x256} elements. The gravity acceleration is set
-to $\bm{g}=\bm{0}$ and effective viscosity is capped using $\eta_{min}=10^{-4}$ and $\eta_{max}=10^3$. The tolerance for the convergence of the non-linear
-solution is $tol=10^{-9}$, with a maximum number of non-linear iterations set to 500. The medium in the domain has $\rho=0.01$, an initial viscosity
-$\eta_0=10$, cohesion $C=1$ and an angle of internal friction $\phi=0$°. Velocity boundary conditions are set to no slip at the bottom and free slip at the
-sides of the domain. The top of the domain is open with the exception of the central portion (punch area) with width $w_p=0.125$, where $v=-1.05$ and $u$ is
-fixed either to 0 (rough punch experiment) or free (smooth punch experiment).
-
-Fig. \ref{fig:indenter} show the results in terms of viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure
-(panels g and h) for both experiments. The analytical solution indicate that pressure at ($0.5;1$) and ($0.5 \pm w_p;1$)is $p=\pi +1$ and $p=1$, respectively
-(black and red lines in Fig. \ref{fig:smooth_p}). Fig. \ref{fig:smooth_p} shows a clear improvement in the pressure solution when passing from a rough to a
-smooth punch (green and blue lines in Fig. \ref{fig:smooth_p}, respectively), as previously observed by \citet{Thieulot2014} and \citet{Glerum2018}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Indenter}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Indenter.pdf}
-\caption{Viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure (panels g and h) fields for rough (left column) and
-smooth (right column) punch experiments.}
-\label{fig:indenter}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Smooth_Rough.pdf}
-\caption{Nodal pressure in function of the x coordinate at $y=1$ for rough (green line) and smooth (blue line) punch experiment. Black and red lines indicate
-the analytical solution at $x=0.5$ and $x=0.5 \pm w_p$, respectively.}
-\label{fig:smooth_p}
-\end{figure}
-
-\subsection{The brick experiment}\label{sec:brick}
-An instantaneous version of the brick benchmark is performed to verify the correctness of the pressure-dependent plasticity for different angle of internal
-friction, as proposed by \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{40}{\km} and $L_y=$ \SI{10}{\km} and a grid resolution of \num{512x128}
-elements. The gravity acceleration is set to $g_y=$ \SI{-10}{\m\square\s} and effective viscosity is capped using $\eta_{min}=$ \SI{e19}{\pascal\s} and
-$\eta_{max}=$ \SI{e26}{\pascal\s}. The tolerance for the convergence of the non-linear solution is $tol=10^{-7}$, with a maximum number of non-linear iterations
-set to 1000. A 800 m-wide and 400 m-height inclusion with $\eta_b=$ \SI{e20}{\pascal\s} and $\rho_b=$ \SI{2700}{\kg\per\cubic\m} is placed at the bottom of the
-domain at $x=L_x/2$. The inclusion is surrounded by a non-linear viscous medium with $\rho_m=$ \SI{2700}{\kg\per\cubic\m}, an initial viscosity of
-\SI{e23}{\pascal\s}, a linear viscous viscosity of \SI{e25}{\pascal\s} and a cohesion of \SI{40}{\mega\pascal}. Velocities are set to free slip conditions at
-the bottom of the domain and the top is open. Velocities on sides of the domain are fixed to $u=$ \SI{+-2e-11}{\m\per\s} and $v=0$. The experiment is performed
-in compressional and extensional contexts with an angle of internal friction $\phi$ of the non-linear medium variable between 0° and 30°.
-
-As expected, two shear bands stem from the inclusion with variable angles in relation to both the dynamics context and the internal friction angle. Shear band
-angles formed at 45° for $\phi=0$° in both extensional and compressional contexts (Fig. \ref{fig:brick_beam}a), while they are different in case of $\phi\neq0$°.
-Shear band angles in case of $\phi=20$° are shown for compression and extension in Fig. \ref{fig:brick_beam}b  and c, respectively. Values of shear band angles
-as function of internal friction angles from 0° to 30° are extracted at different depths and minimum and maximum values are plotted in Fig. \ref{fig:brick},
-in comparison with theoretical Roscoe, Arthur and Coulomb shear band angles. For all tests the tolerance for nonlinear convergence is defined $tol=10^{-7}$
-and none converges before the maximum number of iterations ($it_{max}=1000$) is reached. However, tests with internal friction angles up to 15° show a constant
-decrease of the velocity residuals, which can not be observed for for tests with higher internal friction angles (Fig. \ref{fig:convergence}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Brick_experiment}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Brick_Beam.pdf}
-\caption{Shear band angles predicted for the brick experiment for $\phi=0$° (panel a) and for $\phi=20$° in case of compression and extension (panels b and c,
-respectively).}
-\label{fig:brick_beam}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Brick.pdf}
-\caption{Shear band angles predicted for the brick experiment in case of compressional and extensional contexts as function of different internal angle of
-friction (continuous black line and red dots), compared with theoretical Roscoe, Artur and Coulomb angles (discontinuous black lines). Red lines indicate the
-range of angles calculated at different depths.}
-\label{fig:brick}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \noindent\includegraphics[width=500px]{./Figures/convergence.pdf}
-  \caption{Convergence rates for all tests. Normalised residuals (left column) are compared with velocity residuals (right column) for extensional (first row)
-  and compressional (second row) settings.}
-  \label{fig:convergence}
-  \end{figure}
-
-\subsection{Advection stabilisation}\label{sec:advection}
-The 1D advection problem proposed by \citet{Donea2003} and \citet{Thieulot2011} is performed to verify the effectiveness of the SUPG method to stabilise the
-advection term of the energy equation. The domain is a 1D segment with $L_x=1$ composed by 50 elements and a discontinuity in the thermal field placed at
-$x=0.25$. Temperature is set to 1 for $x \leq 0.25$ and to 0 for $x > 0.25$. Velocity is set to ${u}=1$ in the entire domain. The simulation is performed for
-250 time steps, with $dt=0.002$, so the thermal discontinuity should be at $x=0.75$ at the end of the simulation.
-
-Temperature profiles at the end of the simulation are shown in \ref{fig:advection} as function of the dimensionless coefficient $\gamma=\tau \bm{v}/h$. In case
-of the classic Galerkin method ($\gamma=0$, blue line in \ref{fig:advection}) the final thermal profile is characterised by strong oscillations, which are
-eliminated in case of the SUPG method ($\gamma=0.045$, orange line in \ref{fig:advection}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Advection_stabilisation}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Advection.pdf}
-\caption{Temperature profile as function of \textit{x} for the advection stabilisation benchmark. Purple line indicates the initial temperature profile; the
-green line indicates the analytical temperature profile after 250 time steps; blue line indicates the temperature profile after 250 time steps in case of the
-classic Galerkin method ($\gamma=0$); orange line indicates the temperature profile after 250 time steps in case of the SUPG method ($\gamma=0.045$).}
-\label{fig:advection}
-\end{figure}
-
-\subsection{Shear heating}\label{sec:simple_shear}
-This exercise is performed in an unit square domain composed by \num{128x128} elements. The velocity field is prescribed on the entire domain with
-$\bm{v}=(L_y-y)y\bm{e_x}$; viscosity, density and specific heat are fixed to 1, while thermal conductivity and radiogenic energy are fixed to 0. Therefore,
-the energy equation (Eq. \ref{eq:energy}) can be simplified as \[\frac{\partial T}{\partial t}=H_s\] and fixing $T(t=0)=0$, the temperature can be find as
-\[T(t)=H_s t\]
-In this case we have
-\begin{eqnarray}
-\dot{\epsilon_{xx}}&=&\frac{\partial u}{\partial x}=0 \nonumber \\
-\dot{\epsilon_{yy}}&=&\frac{\partial v}{\partial y}=0 \nonumber \\
-\dot{\epsilon_{xy}}&=&\frac{1}{2}\left(\frac{\partial v}{\partial x}+\frac{\partial u}{\partial y}\right)=\frac{1}{2}(L_y-2y)\nonumber
-\end{eqnarray}
-and, simplifying Eq. \ref{eq:shear}, shear heating can be calculated as \[H_s(x,y)=(1-2y)^2\]
-
-The solution predicted by the model in terms of velocity, temperature and shear heating match well with the analytical solutions (Fig. \ref{fig:analytical_en}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Simple_shear}.
-
-%\begin{figure}
-%\centering
-%\noindent\includegraphics[width=400px]{./Figures/Energy_Thieulot.pdf}
-%\caption{Temperature and shear heating predicted by the model after 1 time step.}
-%\label{fig:en_thieulot}
-%\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=500px]{./Figures/Analytical_Shear.pdf}
-\caption{Velocity (panel a), temperature (panel b) and shear heating (panel c) as function of the \textit{y} coordinate for the simple shear experiment. The
-solutions predicted by the model (red lines) are compared with the analytical solutions (dashed blue lines).}
-\label{fig:analytical_en}
-\end{figure}
-
-\subsection{Shear and adiabatic heating}\label{sec:shear}
-This problem is performed as presented in Exercise 9.4 in \citet{Gerya2010b}. Two materials are vertically separated in a rectangular domain with $L_x=$
-\SI{1000}{\km}, $L_y=$ \SI{1500}{\km} and a grid resolution of \num{30x20} elements. Constant thermal coefficient expansion ($\alpha=$ \SI{3e-5}{\per\kelvin}),
-temperature ($T=$ \SI{1300}{\kelvin}) and gravity acceleration ($g_y=$ \SI{-10}{\m\square\s}) are assumed in the whole domain. Fluid 1 (on the left side of the
-domain) has $\rho_1=$ \SI{3200}{\kg\per\cubic\m} and $\eta_1=$ \SI{e20}{\pascal\s}; fluid 2 (on the right side of the domain) has $\rho_2=$
-\SI{3300}{\kg\per\cubic\m} and $\eta_2=$ \SI{e22}{\pascal\s}. Velocity boundary conditions are set to free slip on all sides of the domain.
-
-As shown in Fig. \ref{fig:energy}, both shear and adiabatic heating predicted by the code (first row in Fig. \ref{fig:energy}) well recreate the results
-obtained by means of example ${Shear\_adiabatic\_heating.m}$ from \citet{Gerya2010b} (second row in Fig. \ref{fig:energy}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Adiabatic%2BShear}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Energy.pdf}
-\caption{Comparison between shear (first column), adiabatic (second column) and total (third column) energy predicted by the code (panels a, b and c) and those
-created using example ${Shear\_adiabatic\_heating.m}$ from Exercise 9.4 in \citet{Gerya2010b} (panels d, e and f).}
-\label{fig:energy}
-\end{figure}
-
-\subsection{Mantle convection}\label{sec:mantle}
-This problem is performed as presented by \citet{Blankenbach1989} (constant viscosity cases) and \citet{Thieulot2014}, in a 2D unit square domain with gravity
-acceleration $g_y=10^{10}Ra$. The experiment is performed with three different Rayleigh numbers ($Ra=10^4$, $10^5$ and $10^6$) and with different grid
-resolution (between \num{32x32} and \num{128x128} elements). The fluid has constant viscosity, initial density, heat capacity, thermal conductivity
-($\eta= \rho_0= C_p = k=1$), reference temperature ($T_0=0$) and thermal expansion coefficient ($\alpha=10^{-10}$). Temperatures are set to 0 on top and 1 on
-bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\]
-
-The solution generated by the code in terms of $v_{\textrm{rms}}$ (calculated as in Sec. \ref{sec:error}) and Nu as function of time are reported for all the
-simulations in Fig. \ref{fig:mantle}. At the steady state, $v_{\textrm{rms}}$ and Nu of all simulations converge well toward the values from \citet{Blankenbach1989},
-with lower errors for higher resolution grids (Table \ref{tab:mantle}). All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Mantle_convection}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Convection.pdf}
-\caption{$v_{\textrm{rms}}$ (panels a, c and e) and Nu (panels b, d and f) for the mantle convection benchmark as function of time for different grid resolution.
-Panels a and b show the results for $Ra=10^4$; Panels c and d show the results for $Ra=10^5$; panels e and f show the results for $Ra=10^6$. Purple lines
-indicate the convergence values for the $v_{\textrm{rms}}$ and the Nu from \citet{Blankenbach1989} at the steady state.}
-\label{fig:mantle}
-\end{figure}
-
-\begin{table}[H]
-\caption{Comparison between $v_{\textrm{rms}}$ and Nu predicted by the code for the mantle convection experiment and same values as reported in literature.}
-\centering
-\small
-\begin{tabular}{|>{\centering}m{0.09\textwidth}|>{\centering}m{0.05\textwidth}|>{\centering}m{0.24\textwidth}|>{\centering}m{0.15\textwidth}|
-  >{\centering}m{0.1\textwidth} >{\centering}m{0.1\textwidth} >{\centering\arraybackslash}m{0.1\textwidth}|}
-\toprule
-& & \multirow{2}{*}{\citet{Blankenbach1989}} & \citet{Thieulot2014} & \multicolumn{3}{c|}{FALCON}  \\
- & & & \num{200x200} & \num{32x32} & \num{64x64} & \num{128x128}  \\
-\midrule
-\multirow{2}{*}{$Ra=10^4$} & $v_{\textrm{rms}}$ & $42.864947 \pm 0.000020$ & 42.867 & 42.83226 & 42.852793 & 42.861394  \\
-           & Nu      & $4.884409 \pm 0.000010$  & 4.882  & 4.781297 & 4.857475  & 4.877573   \\
-\hline
-\multirow{2}{*}{$Ra=10^5$} & $v_{\textrm{rms}}$ & $193.21454 \pm 0.00010$  & 193.255& 193.872643 & 193.377472 & 193.252290  \\
-           & Nu      & $10.534095 \pm 0.000010$ & 10.507 & 9.602514 & 10.270735  & 10.465629   \\
-\hline
-\multirow{2}{*}{$Ra=10^6$} & $v_{\textrm{rms}}$ & $833.98977 \pm 0.00020$  & 834.712& 848.091176 & 837.767911 & 834.945793  \\
-           & Nu      & $21.972465 \pm 0.000020$ & 21.695 & 15.999266 & 19.703682  & 21.306939   \\
-\bottomrule
-\end{tabular}
-\label{tab:mantle}
-\end{table}
-
-\subsection{Viscoplastic mantle convection}\label{sec:visco_mantle}
-This benchmark is performed as Cases 1-5a described in \citet{Tosi2015} in a 2D unit square domain with two different grid resolutions of \num{32x32} 
-and \num{100x100} elements. Reference viscosity, reference density, heat capacity, thermal conductivity and thermal expansion coefficient are fixed to a constant
-value of 1 ($\eta_0= \rho_0= C_p = k=\alpha=1$), while the gravity has been chosen as $g_y=10^2$, in order to obtain a Rayleigh number $Ra=10^2$. Temperatures
-are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\].
-The viscosity field $\eta$ is calculated as the harmonic average between a linear part $\eta_{lin}$ that depends on temperature only or on temperature and 
-depth $y$ and a nonlinear-plastic part $\eta_{pl}$ that depends on the strain rate $\dot{\bm{\epsilon}}$, as follows
-\[\eta(T,y,\dot{\bm{\epsilon}})=2\left(\frac{1}{\eta_{lin}(T,y)}+\frac{1}{\eta_{pl}(\dot{\bm{\epsilon}})}\right)^{-1}\]
-The linear and the nonlinear-plastic parts of the viscosity are calculated following \citet{Tosi2015} as follows
-\[\eta_{lin}(T,y)=\exp(-\ln(\Delta\eta_T)T+\ln(\Delta\eta_y)y)\]
-\[\eta_{pl}(\dot{\bm{\epsilon}})=\eta^*+\frac{\sigma_Y}{\sqrt{\dot{\bm{\epsilon}}:\dot{\bm{\epsilon}}}}\]
-where $\Delta\eta_T$, $\Delta\eta_y$, $\eta^*$ and $\sigma_Y$ are parameters chosen as listed in Table \ref{tab:tosi_setup}.
-
-For each case, some diagnostic quantities measured at the steady state are reported in Table \ref{tab:visco_mantle} (see Section \ref{sec:error} for 
-the computation of all the quantities), showing a well fit with results described in \citet{Tosi2015}. Data and charts for all cases can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Viscoplastic_mantle_convection}.
-
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{white!100!}{gray!25!}
-%\begin{table}[H]
-%  \caption{Results of the viscoplastic mantle convection benchmark for Cases 1-4 compared with results obtained with ELEFANT and YACC (from \citealp{Tosi2015}).}
-%  \centering
-%  \resizebox{\textwidth}{!}{
-%\begin{tabular}{|c|ccc|ccc|ccc|ccc|}
-%  \toprule
-%    & \multicolumn{3}{c|}{\textbf{Case 1}} & \multicolumn{3}{c|}{\textbf{Case 2}} & \multicolumn{3}{c|}{\textbf{Case 3}} &
-%    \multicolumn{3}{c|}{\textbf{Case 4}} \\
-%   & FALCON & ELEFANT& YACC & FALCON& ELEFANT& YACC & FALCON & ELEFANT & YACC & FALCON & ELEFANT & YACC \\
-%  \midrule
-%   $\left\langle T \right\rangle$ & 0.7759 & 0.7758 & 0.7767 & 0.6043 & 0.6036 & 0.6058 & 0.7275 & 0.7275 & 0.7286 & & 0.5277 & 0.5289\\
-%    Nu$^{top}$ & 3.4280 & 3.4214 & 3.4298 & 8.5242 & 8.5115 & 8.5278 & 3.0368 & 3.0347 & 3.0374 && 6.5912 & 6.5572 \\
-%    Nu$^{bottom}$ & 3.3168 & 3.313 & 3.3143 & 8.4599 & 8.4465 & 8.3990 & 3.0086 & 2.9908 & 2.9628 && 6.5834 & 6.5243 \\
-%    $v_{\textrm{rms}}$ & 249.901 & 249.134 & 251.7997 & 141.2451 & 140.7087 & 142.2020 & 100.2776 & 100.1208 & 100.9467 && 79.1105 & 79.6202 \\
-%    $v_{\textrm{rms}}^{top}$ & 1.8907 & 1.8642 & 1.8298 & 104.7928 & 104.1228 & 105.1767 & 2.0832 & 2.0652 & 2.0374 && 74.7596 & 75.8414 \\
-%    $u_{max}^{top}$ & 2.6340 & 2.6119 & 2.5516 & 121.9212& 121.766 & 122.3238 & 2.9086 & 2.9019 & 2.8458 && 88.9146 & 89.2940 \\
-%    $\eta_{min}$ & - & - & - & $2.0303\times10^{-5}$ & $2.0600\times10^{-5}$ & $1.9853\times10^{-5}$ & $4.8129\times10^{-5}$ & $4.8080\times10^{-5}$ & $4.7907\times10^{-5}$ &
-%    & $1.9860\times10^{-4}$ & $1.9174\times10^{-4}$ \\
-%    $\eta_{max}$ & - & - & - & 1.3325 & 1.85 & 1.6095 & 0.9360 & 0.9023 & 1 && 1.52 & 1.6773 \\
-%    $\left\langle W \right\rangle$ & 2.4299 & 2.4316 & 2.4583 & 7.5648 & 7.5665 & 7.6222 & 2.0387 & 2.0384 & 2.0400 && 5.6216 & 5.6512 \\
-%    $\left\langle \Phi \right\rangle/Ra$ & 2.4311 & 2.4276 & 2.4333 & 7.5661 & 7.5585 & 7.6194 & 2.0386 & 2.0370 & 2.0335 && 5.6182 & 5.6463 \\
-%    $\delta$ & 0.05\% & 0.16\% & 1.02\% & 0.02\% & 0.11\% & 0.04\% & $<$0.01\% & 0.07\% & 0.32\% && 0.06\% & 0.09\% \\
-%  \bottomrule
-%  \end{tabular}}
-%  \label{tab:visco_mantle}
-%  \end{table}
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{2}{white!100!}{gray!25!}
-\begin{table}[H]
-  \caption{Benchmark cases and corresponding parameters.}
-  \centering
-  \small
-\begin{tabular}{|>{\centering}m{0.08\textwidth}|>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}
-  >{\centering}m{0.05\textwidth} >{\centering}m{0.05\textwidth} |>{\centering\arraybackslash}m{0.2\textwidth}|}
-  \toprule
-    Case & $Ra$ & $\Delta\eta_T$ & $\Delta\eta_y$ & $\eta^*$ & $\sigma_Y$ & Convective regime \\
-  \midrule
-  \rowcolor{gray!25!}
-    1 & $10^2$ & $10^5$ & 1 & - & - & Stagnant lid \\
-    2 & $10^2$ & $10^5$ & 1 & $10^{-3}$ & 1 & Mobile lid  \\
-    \rowcolor{gray!25!}
-    3 & $10^2$ & $10^5$ & 10 & - & - & Stagnant lid  \\
-    4 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 1 & Mobile lid \\
-    \rowcolor{gray!25!}
-    5 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 4 & Periodic \\
-  \bottomrule
-  \end{tabular}
-  \label{tab:tosi_setup}
-  \end{table}
-
-\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{gray!25!}{white!100!}
-\begin{table}
-  \caption{Results of the viscoplastic mantle convection benchmark for different grid resolutions in Cases 1-5 compared with results obtained with ELEFANT
-  and YACC ($100\times100$ elements, from \citealp{Tosi2015}) and in Stone 28 ($32\times32$ elements, from
-  \url{https://github.com/cedrict/fieldstone/tree/master/python_codes/fieldstone_28}).}
-  \centering
-  \resizebox{\textwidth}{!}{
-\begin{tabular}{|cc|ccccccccccc|}
-  \toprule
-  \multicolumn{13}{|c|}{\textbf{Case 1}} \\
-  \midrule
-  Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-  & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-  \hline
-  \rowcolor{gray!25!}
-    YACC & $100\times100$ & 0.7767 & 3.4298 & 3.3143 & 251.7997 & 1.8298 & 2.5516 & - & - & 2.4583 & 2.4333 & 1.02\% \\
-    ELEFANT & $100\times100$ & 0.7758 & 3.4214 & 3.313 & 249.134 & 1.8642 & 2.6119 & - & - & 2.4316 & 2.4276 & 0.16\% \\
-    \rowcolor{gray!25!}
-    Stone 28 & $32\times32$ & 0.7737 & 3.3987 & - & 243.872 & - & - & - & - & - & - & - \\
-%    FALCON & $32\times32$ & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & - & - & 2.4745 & 2.4864 & 0.48\% \\
-%    \rowcolor{gray!25!}
-    FALCON & $100\times100$ & 0.7759 & 3.4280 & 3.3168 & 249.901 & 1.8907 & 2.6340 & - & - & 2.4299 & 2.4311 & 0.05\% \\
-    \midrule
-    \multicolumn{13}{|c|}{\textbf{Case 2}} \\
-    \midrule
-    Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-    & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-    \hline
-    \rowcolor{gray!25!}
-     YACC & $100\times100$ & 0.6058 & 8.5278 & 8.3990 & 142.2020 & 105.1767 & 122.3238 & $1.9853\times10^{-5}$ & 1.6095 & 7.6222 & 7.6194 & 0.04\% \\
-     ELEFANT & $100\times100$ & 0.6034 & 8.5115 & 8.4465 & 140.8067 & 104.1228 & 121.766 & $2.0600\times10^{-5}$ & 1.85 & 7.5665 & 7.5585 & 0.11\% \\
-     \rowcolor{gray!25!}
-     Stone 28 & $32\times32$ & 0.6052 & 8.1715 & - & 141.518 & - & - & - & - & - & - & - \\
-%     FALCON & $32\times32$ & 0.6454 & 8.1422 & 7.4020 & 156.6908 & 95.7226 & 116.9302 & $2.2572\times10^{-5}$ & 0.7194 & 7.4289 & 7.4452 & 0.22\% \\
-%     \rowcolor{gray!25!}
-     FALCON & $100\times100$ & 0.6044 & 8.5266 & 8.4590 & 141.3684 & 104.8867 & 122.0470 & $2.0333\times10^{-5}$ & 1.2914 & 7.5707 & 7.5718 & 0.01\% \\
-     \midrule
-     \multicolumn{13}{|c|}{\textbf{Case 3}} \\
-     \midrule
-     Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-     & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-     \hline
-     \rowcolor{gray!25!} 
-     YACC & $100\times100$ & 0.7286 & 3.0374 & 2.9628 & 100.9467 & 2.0374 & 2.8458 & $4.7907\times10^{-5}$ & 1 & 2.0400 & 2.0335 & 0.32\% \\
-      ELEFANT & $100\times100$ & 0.7275 & 3.0347 & 2.9908 & 100.1208 & 2.0652 & 2.9019 & $4.8080\times10^{-5}$ & 0.9023 & 2.0384 & 2.037 & 0.07\% \\
-      \rowcolor{gray!25!}
-      Stone 28 & $32\times32$ & 0.7277 & 3.0347 & - & 100.018 & - & - & - & - & - & - & - \\
-%      FALCON & $32\times32$ & 0.7281 & 3.0618 & 2.8132 & 101.0092 & 2.1630 & 3.0178 & $4.8500\times10^{-5}$ & 0.7845 & 2.0699 & 2.0695 & 0.02\% \\
-%      \rowcolor{gray!25!}
-      FALCON & $100\times100$ & 0.7275 & 3.0368 & 3.0086 & 100.2776 & 2.0832 & 2.9086 & $4.8129\times10^{-5}$ & 0.9360 & 2.0387 & 2.0386 & $<$0.01\% \\
-      \midrule
-      \multicolumn{13}{|c|}{\textbf{Case 4}} \\
-      \midrule
-      Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-      & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-      \hline
-      \rowcolor{gray!25!} 
-      YACC & $100\times100$ & 0.5289 & 6.5572 & 6.5243 & 79.6202 & 75.4814 & 89.2940 & $1.9174\times10^{-4}$ & 1.6773 & 5.6512 & 5.6463 & 0.09\% \\
-       ELEFANT & $100\times100$ & 0.5277 & 6.5912 & 6.5834 & 79.1105 & 74.7596 & 88.9146 & $1.9860\times10^{-4}$ & 1.5200 & 5.6216 & 5.6182 & 0.06\% \\
-       \rowcolor{gray!25!}
-       Stone 28 & $32\times32$ & 0.5286 & 6.4036 & - & 79.4746 & - & - & - & - & - & - & - \\
-%       FALCON & $32\times32$ & 0.5334 & 6.4679 & 6.3435 & 80.6153 & 76.0511 & 90.2058 & $2.0393\times10^{-4}$ & 0.6453 & 5.6676 & 5.6698 & 0.04\% \\
-%       \rowcolor{gray!25!}
-       FALCON & $100\times100$ & 0.5279 & 6.5897 & 6.5830 & 79.1220 & 75.0715 & 88.8484 & $1.9693\times10^{-4}$ & 1.4133 & 5.6127 & 5.6113 & 0.03\% \\
-       \midrule
-       \multicolumn{13}{|c|}{\textbf{Case 5}} \\
-       \midrule
-       Code & Resolution & $P$ & $\left\langle T \right\rangle_{min}$ & $\left\langle T \right\rangle_{max}$ & Nu$^{top}_{min}$ & Nu$^{top}_{max}$ &
-       $v_{\textrm{rms}}^{min}$ & $v_{\textrm{rms}}^{max}$ & \multicolumn{2}{c}{$(\left\langle \Phi \right\rangle/Ra)_{min}$} & 
-       \multicolumn{2}{c|}{$(\left\langle \Phi \right\rangle/Ra)_{max}$} \\
-       \hline
-       \rowcolor{gray!25!}
-         YACC & $100\times100$ & 0.0792 & 0.6543 & 0.6722 & 2.6950 & 7.2792 & 41.9813 & 98.8230 & \multicolumn{2}{c}{1.4347} & 
-         \multicolumn{2}{c|}{9.2748} \\
-         ELEFANT & $100\times100$ & 0.076 & 0.6514 & 0.6693 & 2.6808 & 7.3506 & 41.2578 & 100.089 & \multicolumn{2}{c}{1.3324} & 
-         \multicolumn{2}{c|}{9.5289} \\
-         \rowcolor{gray!25!}
-         Stone 28 & $32\times32$ & 0.0764 & 0.6521 & 0.6697 & 2.6912 & 7.0763 & 41.61 & 98.183 & \multicolumn{2}{c}{-} & 
-         \multicolumn{2}{c|}{-} \\
-     %    FALCON & $32\times32$ & & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & \multicolumn{2}{c}{-} & 
-     %     \multicolumn{2}{c|}{-} \\
-     %    \rowcolor{gray!25!}
-         FALCON & $100\times100$ & 0.0775 & 0.6519 & 0.6697 & 2.6847 & 7.3629 & 41.5256 & 99.5466 & \multicolumn{2}{c}{1.3399} & 
-         \multicolumn{2}{c|}{9.3993} \\
-  \bottomrule
-  \end{tabular}}
-  \label{tab:visco_mantle}
-  \end{table}
-
-\subsection{Thin layer entrainment}\label{sec:thin}
-This experiment is performed as originally proposed by \citet{vanKeken1997} with two incompressible fluids in a rectangular domain with $L_x=2$ and $L_y=1$
-and gravity acceleration $g_y=10^{10}Ra$. The Rayleigh number and the compositional Rayleigh number are fixed to $Ra=300000$ and $Rc=450000$, respectively.
-Both fluids have constant viscosity, thermal conductivity, specific heat ($\eta=\rho=C_p=1$) and thermal expansion coefficient ($\alpha=10^{-10}$). Fluid 1
-has a density $\rho_1=1$, while fluid 2 is denser ($\rho_2=\rho_1+1.5 \cdot 10^{-10}$) and is placed at the bottom of the domain, for $y \leq 0.025$.
-Temperature are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides of the domain. The initial
-temperature field is given by
-\[T(x,y)=T_u(x,y)+T_l(x,y)+T_r(x,y)+T_s(x,y)-\frac{3}{2}\]
-with
-\begin{eqnarray}
-T_u(x,y)&=&\frac{1}{2}erf\left(\frac{1-y}{2}\sqrt{\frac{u_0}{x}}\right)\nonumber \\
-T_l(x,y)&=&1-\frac{1}{2}erf\left(\frac{y}{2}\sqrt{\frac{u_0}{L_x-x}}\right)\nonumber \\
-T_r(x,y)&=&\frac{1}{2}+\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{y+1}} \exp\left(-\frac{x^2u_0}{4y+4}\right)\nonumber \\
-T_s(x,y)&=&\frac{1}{2}-\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{2-y}} \exp\left(-\frac{(L_x-x)^2u_0}{8-4y}\right)\nonumber
-\end{eqnarray}
-and
-\begin{eqnarray}
-u_0&=&\frac{L_x^{7/3}}{(1+L_x^4)^{2/3}}\left(\frac{Ra}{2\sqrt{\pi}}\right)^{\frac{2}{3}}\nonumber \\
-Q&=&2\sqrt{\frac{L_x}{\pi u_0}}\nonumber
-\end{eqnarray}
-
-The experiment is performed with two grid resolution (\num{125x40} and \num{200x80} elements), with 100 markers per element and CFL of 0.25. $v_{\textrm{rms}}$
-(see Sec. \ref{sec:error}) calculated for both the simulations match well with results from \citet{vanKeken1997} and obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and ELEFANT \citep{Thieulot2014}, obtained with similar grid resolutions
-(Fig. \ref{fig:thin}). In addition, a variety of simulations with different resolutions and aspect ratios are performed to verify that the sensitivity of the
-initial velocity field. Results match well with results from \citet{vanKeken1997} (in the grey area) and \citet{Thieulot2014} (Fig. \ref{fig:thin_initial}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Thin_layer}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_thin.pdf}
-\caption{$v_{\textrm{rms}}$ for the thin layer experiment as function of time for different grid resolution. Results are compared with results obtained by
-\citet{vanKeken1997} (black lines) and with ASPECT and ELEFANT.}
-\label{fig:thin}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrmszero.pdf}
-\caption{$v_{\textrm{rms}}$ at $t=0$ for the thin layer experiment as function of the element size for different aspect ratios of the elements. Results are compared
-with results obtained by \citet{vanKeken1997} (black dots in the grey area) and with ASPECT and ELEFANT.}
-\label{fig:thin_initial}
-\end{figure}
-
-\subsection{Sinking hydrated cylinder}\label{sec:quinquis}
-This experiment is performed as explained in \citet{Quinquis2014} and simulates the sinking of a cold, hydrated cylinder into a warm, dry mantle. The domain
-is square with $L_x=L_y=$ \SI{300}{\km}, a grid resolution of \num{300x300} elements and 25 markers per element. The sinking cylinder is located at $x=$
-\SI{150}{\km} and $y=$ \SI{170}{\km} with a radius of \SI{20}{\km} and has $\rho_c=$ \SI{3250}{\kg\per\cubic\m}, $\eta_c=$ \SI{e23}{\pascal\s}, a thermal
-conductivity $k_c=$ \SI{4.5}{\watt\per\m\per\kelvin},a specific heat $C_{p_c}=$ \SI{1250}{\joule\per\kg\per\kelvin}; the surrounding mantle has $\rho_m=$
-\SI{3200}{\kg\per\cubic\m}, $\eta_m=$ \SI{e20}{\pascal\s}, a thermal conductivity $k_m=$ \SI{105}{\watt\per\m\per\kelvin} and specific heat $C_{p_m}=$
-\SI{1250}{\joule\per\kg\per\kelvin}; a 58 km-thick lithosphere overlies the mantle and has $\rho_l=$ \SI{3200}{\kg\per\cubic\m}, $\eta_l=$ \SI{e23}{\pascal\s},
-a thermal conductivity $k_l=$ \SI{4.5}{\watt\per\m\per\kelvin}, specific heat $C_{p_l}=$ \SI{750}{\joule\per\kg\per\kelvin} and a thermal diffusivity of
-\SI{e-6}{\square\m\per\s}. Velocity boundary conditions are set to free slip on all sides of the domain. Temperature increase linearly in the lithosphere,
-from 0° to 1300°C, and in the mantle, from 1300° to 1360.5°C. The initial temperature of the cylinder is fixed to 400°C. Throughout the evolution, temperature
-is fixed to 0° and 1360.5°C at the top and bottom of the domain, respectively. The initial bound water content is imposed at 2 wt.\% in the cylinder and at
-0 wt.\% in both the mantle and the lithosphere. The maximum amount of water in the mantle is fixed to 0.2 wt.\%, while maximum water content in the cylinder
-and in the lithosphere is function of pressure and temperature and is calculated using a serpentinized harzburgite with Perple\_X, as in \citet{Quinquis2014}.
-
-As shown in \citet{Quinquis2014}, the model is characterised by a progressive dehydration in the external portion of the cylinder, due to the increase of
-temperature, with a consequent vertical migration of free water that hydrates the mantle above the cylinder, up to the lithosphere. The distribution of
-bound and free water during the evolution of the experiment is shown in Fig. \ref{fig:quinquis} (left and right columns, respectively). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Hydration}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Hydration.pdf}
-\caption{Distribution of bound and free water (left and right columns, respectively) at $t=0.5$ and $t=$ \SI{1.25}{\mega\year} (first and second rows,
-respectively) for the sinking hydrated cylinder experiment. Only markers with bound or free water not equal to 0 are plotted.}
-\label{fig:quinquis}
-\end{figure}
-
-\subsection{Mantle melting curves}\label{sec:katz}
-This instantaneous experiment is performed in order to recreate isobaric melting curves, as function of temperature, bulk water and content of water in the
-melt, to compare with melting curves obtained by \citet{Katz2003}. The domain is square with $L_x=L_y=100$ km, a grid resolution of \num{100x100} elements and
-25 markers per element. Density is set in the entire domain to \SI{3300}{\kg\per\cubic\m} and temperature increases from the sides to the centre from 1000° to
-1500°C. The experiment is performed for bulk water contents from 0 to 0.3 wt.\%.
-
-Fig. \ref{fig:fraction} shows isobaric melting curves for \SI{1}{\giga\pascal} (panel a) and \SI{3}{\giga\pascal} (panel b) calculated for different bulk water
-contents. The curves well match with those obtained for the mantle by \citet{Katz2003}, showing the saturation at the solidus for bulk water of 0.2 and 0.3 wt.\%, indicating
-that the code correctly determines melt fraction as function of pressure, temperature and bulk water content. Same results can be observed in Fig.
-\ref{fig:water}a, where isobaric and isothermal melting curves are shown as function of bulk water content. In addition, Fig. \ref{fig:water}b shows that the
-code correctly determines also the percentage of water in the melt as function of pressure, temperature and melt fraction. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Melting}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Fraction.pdf}
-\caption{Isobaric melting curves for 1 GPa (panel a) and 3 GPa (panel b) as function of temperature with different bulk water contents.}
-\label{fig:fraction}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Water.pdf}
-\caption{Isobaric melting curves for 1.5 GPa as function of bulk water (panel a) and water in the melt (panel b) calculated for different temperatures.}
-\label{fig:water}
-\end{figure}
-
-\clearpage
-
-\bibliography{export}
-
-\end{document}

### 2022-09-07 tex/Benchmark/shear.tex
update main

--- 2022-07-20 FALCON_description.tex
@@
-\documentclass[hidelinks,10pt,a4paper]{article}
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{textcomp}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage{booktabs}
-\usepackage{multirow}
-\usepackage[cm]{fullpage}
-\usepackage[round]{natbib}
-\bibliographystyle{abbrvnat}
-\usepackage{doi}
-\usepackage{bm}
-\usepackage[table]{xcolor}
-\hypersetup{
-    colorlinks,
-    linkcolor={blue},
-    citecolor={blue},
-    urlcolor={blue}
-}
-\usepackage{float}
-\restylefloat{table}
-\usepackage{siunitx}
-\DeclareSIUnit\year{yr}
-
-\begin{document}
-
-\section{Introduction}\label{sec:intro}
-This document includes a description of the code FALCON, with a presentation of the features implemented followed by results of extensive benchmarks listed in
-Table \ref{tab:list}, performed to verify that the code solves correctly a wide range of problems. Description and results of all benchmarks are shown
-in Section \ref{sec:benchmark}.
-
-\begin{table}[H]
-\caption{List of benchmarks performed for each feature of the code. Data of all benchmarks can be found at 
-\url{https://github.com/aleregorda/Benchmarks}}
-\centering
-\begin{tabular}{| l | l | l |}
-\toprule
- \textbf{Feature}  & \textbf{Benchmark} & \textbf{Section} \\
-\midrule
-  Solver  & Stokes flow & \ref{sec:stokes} \\
-\hline
-  \multirow{2}{*}{Markers advection} & Zalesak disk & \ref{sec:runge} \\
-    & Conservative Interpolation Velocity & \ref{sec:CVI} \\
-\hline
-  \multirow{4}{*}{Momentum equation}  & Poiseuille flow & \ref{sec:poiseuille}\\
-    & Instantaneous 2D sphere & \ref{sec:ist_sphere} \\
-    & Rayleigh-Taylor experiment & \ref{sec:rayleigh} \\
-    & Falling block & \ref{sec:block} \\
-\hline
-    & 2D Stokes sphere & \ref{sec:sphere} \\
-    Sticky air and  & Stabilisation algorithm & \ref{sec:stab} \\
-    free surface & Topography relaxation & \ref{sec:crameri} \\
-  & Spontaneous subduction & \ref{sec:subduction} \\
-\hline
-  Erosion and sedimentation  & - & - \\
-\hline
-   Non-linear  & Slab detachment & \ref{sec:slab} \\
-   visco-plasticity & Indenter & \ref{sec:indenter} \\
-    & Brick & \ref{sec:brick} \\
-\hline
-  \multirow{3}{*}{Energy equation}  & Advection stabilisation & \ref{sec:advection} \\
-    & Simple shear heating & \ref{sec:simple_shear} \\
-    & Shear and adiabatic heating & \ref{sec:shear} \\
-\hline
-\multirow{3}{*}{Energy + momentum} & Mantle convection & \ref{sec:mantle} \\
-& Viscoplastic mantle convection & \ref{sec:visco_mantle} \\
-& Thin layer entrainment & \ref{sec:thin} \\
-\hline
-  Phase changes  & \multirow{2}{*}{Hydrated sinking cylinder} & \multirow{2}{*}{\ref{sec:quinquis}} \\
-  and hydration  &  & \\
-\hline
-  Melting & Experimental melting curves & \ref{sec:katz} \\
-\bottomrule
-\end{tabular}
-\label{tab:list}
-\end{table}
-
-The code is written in Fortran90 and uses the finite element method (FEM) with quadrilateral $Q_1 \times P_0$ elements (continuous bilinear velocity and
-discontinuous constant pressure), associated with MUMPS\footnote{\url{http://mumps.enseeiht.fr/}} \citep{Amestoy2001,Amestoy2019} that is a software package
-for solving systems of linear equations of the form $A \cdot x = b$, where \textit{A} is a square sparse matrix, by means of a direct method. Since
-$Q_1 \times P_0$ elements do not satisfy Ladyzhenskaya, Babuska and Brezzi (LBB) stability condition \citep{Donea2003}, elemental pressure is elaborated
-in post-processing to avoid spurious pressures by means of a double interpolation to smooth the pressure. Following this procedure, the elemental pressure
-is interpolated onto nodes and then back onto elements. After the smoothing procedure, the pressure field is calculated again on the nodes to be used in
-combination with the lithostatic pressure, which is calculated onto nodes. Correctness of the solution and performances of the solver in terms of time and
-memory usage are tested solving the Stokes flow with the analytical solution proposed by \citet{Donea2003} (Section \ref{sec:stokes}).
-
-The thermo-mechanics of crust-mantle systems is described by means of the conservation of mass, momentum and energy equations, expressed as follows:
-\begin{eqnarray}
-\label{eq:mass}\frac{\partial \rho}{\partial t}+\bm{\nabla} \cdot (\rho \bm{u})&=&0\\
-\label{eq:momentum}-\bm{\nabla} p + \bm{\nabla} \cdot \bm{\tau} + \rho \bm{g}&=&0\\
-\label{eq:energy}\rho C_p \left(\frac{\partial T}{\partial t} + \bm{u} \cdot \bm{\nabla} T \right)&=&\bm{\nabla} \cdot \left(k\bm{\nabla} T\right) + H_{tot}
-\end{eqnarray}
-where $\rho$ is the density, $\bm{u}$ is the velocity, \textit{p} is the pressure, $\bm{\tau}$ is the deviatoric stress, $\bm{g}$ is the gravity acceleration,
-$C_p$ is the specific heat at constant pressure, \textit{T} is the temperature, \textit{k} is the thermal conductivity and $H_{tot}$ is the total internal heat
-production. Density variations due to temperature are generally small enough to assume the density as constant ($\rho=\rho_0$) in Eq. \ref{eq:mass} and in
-Eq. \ref{eq:energy}, while it must be treated as a variable in the buoyancy term of Eq. \ref{eq:momentum}, such that \[\rho=\rho_0(1-\alpha(T-T_0))\]
-where $\rho_0$ is the density at a reference temperature $T_0$ and $\alpha$ is the coefficient of thermal expansion. Eq. \ref{eq:mass} then can be rewritten as
-\begin{equation}\label{eq:mass_0}
-\bm{\nabla} \cdot \bm{u}=0
-\end{equation}
-This simplification is known as the extended Boussinesq approximation.
-
-The time step $dt$ is chosen in according to Courant-Friedrichs-Lewy condition \citep{Anderson1995}, such that
-\[dt=C_n\; \min\left(\frac{\Delta x}{\max|\bm{u}|},\frac{(\Delta x)^2}{\kappa}\right)\]
-where $0<C_n<1$ is the Courant (CFL) number, $\Delta x$ is the minimum dimension of smallest element, $\max|\bm{u}|$ is the maximum velocity calculated on the
-entire domain and
-$\kappa=\frac{k}{\rho C_p}$ is the thermal diffusivity \citep{Thieulot2014}.
-
-\section{Lagrangian markers}\label{sec:markers}
-Elemental properties (density, viscosity, thermal conductivity, specific heat and thermal expansion) needed to solve Eqs. \ref{eq:mass_0}, \ref{eq:momentum}
-and \ref{eq:energy} are related to the composition of each element, determined by means of Lagrangian markers that are characteristic of different materials
-in the domain. Elemental properties (with the sole exception of the viscosity) are calculated using an arithmetic mean, as \[P_e=\frac{1}{n}\sum_{i=1}^n P_i\]
-where $P_e$ is the elemental property, $P_i$ is the property characteristic of the material of each marker and $n$ is the number of markers of the element.
-Differently, the average scheme for the elemental viscosity can be chosen between harmonic, geometric and arithmetic mean. Markers can be placed regularly or
-randomly at the beginning of the simulation and their advection is performed by either a 2nd-order or a 4th-order Runge-Kutta in space, interpolating the velocity 
-field on each marker by means of the shape functions (see Section \ref{sec:runge}). The interpolated velocity is then corrected by means of the Conservative
-Velocity Interpolation (CVI), which introduces a corrective term to reduce dispersion and clustering of particles in both steady state and time-dependent 
-\citep{Wang2015} (see Section \ref{sec:CVI}).
-
-The initial distribution of the markers is created using the open-source code library
-Geodynamic World Builder \citep{Fraters2019}. During the simulation, each marker carries memory of temperature, pressure and accumulated strain, which are
-determined interpolating the relative nodal parameter. In particular, pressure is interpolated onto each marker after the smoothing procedure. The number
-of markers contained in each element is maintained between $n_{min}$ and $n_{max}$. When in an element there are less markers than $n_{min}$ the code adds
-random markers to reach the $n_{min}$, while if the number is higher than $n_{max}$ some of them are randomly deleted. When new markers are added, they assume
-the properties of the nearest marker. In this way elements are never empty and maintain a number of markers inside a prefixed range.
-
-\section{The momentum equation}\label{sec:mom_eq}
-The penalty method is implemented so that Eq. \ref{eq:mass_0} can be written as:
-\begin{equation}\label{eq:mass_penalty}
-\bm{\nabla} \cdot \bm{u} + \frac{p}{\lambda}=0
-\end{equation}
-where $\lambda$ is the so-called penalty parameter, which should be 6-7 orders of magnitude larger than the shear viscosity to ensure that mass conservation
-is satisfied \citep{Donea2003,Thieulot2014}.
-
-The deviatoric stress tensor in Eq. \ref{eq:momentum} can be written in terms of the strain rate tensor as $\bm{\tau}=2\eta\dot{\bm{\epsilon}}$, with
-$\dot{\bm{\epsilon}}=\frac{1}{2}\left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)$. Therefore, Eq. \ref{eq:momentum} can be rewritten as:
-\begin{equation}\label{eq:momentum_pressure}
--\bm{\nabla} p + \bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right) + \rho \bm{g} = 0
-\end{equation}
-Finally, using pressure from Eq. \ref{eq:mass_penalty}, Eq. \ref{eq:momentum_pressure} can be rewritten as:
-\begin{equation}\label{eq:momentum_penalty}
-\lambda \bm{\nabla} \left(\bm{\nabla} \cdot \bm{v}\right)+\bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right)+\rho\bm{g}=0
-\end{equation}
-The penalty method is been associated to the iterative Uzawa method, as described in detail in \citet{Dabrowski2008} and \citet{Thieulot2014}
-(see Section \ref{sec:poiseuille}). Results of benchmarks performed to verify the correctness in the implementation of Eq. \ref{eq:momentum_penalty} are
-shown in Sections \ref{sec:ist_sphere} and \ref{sec:rayleigh}.
-
-To support large viscosities variations the penalty parameter is related to the effective elemental viscosity by means of a dimensionless coefficient, so that
-$\lambda=\lambda_e(e)\eta_{eff}(e)$ \citep{Marotta2006,Dabrowski2008,Thieulot2014}. Benchmark of the falling block \citep{Gerya2003a,Gerya2010,Thieulot2011}
-is performed to verify that the code can correctly deal buoyancy driven flows with strong viscosity contrasts (Section \ref{sec:block}).
-
-\section{Sticky air and free surface}\label{sec:sticky}
-The Earth's surface can be treated by means of either the so-called sticky-air or a true free surface method, both of them implemented in the code. In the
-sticky-air method the surface is approximated with the introduction of a buoyant layer with a viscosity at least four orders of magnitude lower than the crust
-\citep{Schmeling2008,Crameri2012} and the interface between lithosphere and air is defined using a chain of passive markers that are advected as the Lagrangian
-markers. The correctness of the evolution of the markers chain is tested with the experiment of a 2D time-dependent Stokes sphere below a free surface and
-compared with results from ASPECT\footnote{\url{https://aspect.geodynamics.org/}} \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} (Section
-\ref{sec:sphere}). In the true free surface case the top boundary is assumed stress-free and velocities are not fixed. In this case topography variations are
-described by vertical deformations of the mesh that depend on the velocity field of the nodes that identify the free surface, while horizontal deformations are
-not taken into account. This procedure is known as the Arbitrary Lagrangian-Eulerian (ALE) method and its implementation follows the technique described in
-\citet{Thieulot2011}. However, although the implementation of a true free surface better reproduces laboratory experiment, extremely small time steps can be
-necessary to maintain stability \citep{Kaus2010a,Quinquis2011,Thieulot2014}. Therefore, the stability algorithm proposed by \citet{Kaus2010a} is implemented
-to avoid instabilities due to high density differences at the free surface when using too large time steps. The implementation of this algorithm are tested by
-performing the experiment described by \citet{Kaus2010a} (Section \ref{sec:stab}).
-
-The topography relaxation benchmark proposed by \citet{Crameri2012} is performed to verify that the code correctly recovers topography variations in case of
-both the sticky-air and the true free surface method (Section \ref{sec:crameri}). Finally, Section \ref{sec:subduction} show the results of the spontaneous
-subduction experiment described by \citet{Schmeling2008}.
-
-\section{Erosion and sedimentation}\label{sec:erosion}
-Surface processes at either the lithosphere-air interface or in correspondence of the free surface have been implemented by means of the software
-Fastscape\footnote{\url{https://fastscape.org/}} \citep{Braun2013,Cordonnier2019,Yuan2019a,Yuan2019b}. FastScapeLib is a set of routines that solve the stream
-power law (enriched by a sediment transport/deposition term) \citep{Yuan2019a}, hillslope diffusion and marine transport and deposition \citep{Yuan2019b}.
-The partial differential equation solved by FastScapeLib is:
-\begin{equation}
-\frac{\partial h}{ \partial t}=U-K_f A^{m_e}S^{n_e} + \frac{G}{A}\int_A \left( U-\frac{\partial h}{\partial t} \right) dA + K_d \bm{\nabla}^2 h 
-\end{equation}
-where $h$ is the topography, $U$ is the uplift, $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $S$ is the slope, $m_e$ and $n_e$ are
-parameters that should be chosen to constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}, $G$ is the dimensionless
-deposition/transport coefficient for the enriched stream power law and $K_d$ is the effective diffusivity. Terms $K_f A^{m_e}S^{n_e}$ and $K_d \bm{\nabla}^2 h$
-consider long-range processes, described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, and short-range processes,
-proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}.
-
-In case of sticky air, slope is determined by means of the markers chain, which is vertically corrected considering erosion and sedimentation rates. After the
-correction, continental markers above the markers chain are transformed in air markers and air markers below the markers chain become sediments. Similarly,
-slope is determined considering the nodes on the top boundary in case of true free surface and they are vertically corrected as for the markers chain. In this
-case, continental markers above the top boundary are simply deleted.
-%Surface processes at the lithosphere-air interface are implemented taking into account fluvial erosion (long-range processes), hillslope processes (short-range
-%processes) and failure of slopes exceeding a threshold angle $\theta$ \citep{Burov1997,Tucker1999,Sacek2012,Croissant2014,Ueda2015,Beucher2020}. Erosion rates
-%$\dot{E}_{lr}$ due to long-range processes can be described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{lr}=K_f A^{m_e} \bm{\nabla} h^{n_e}
-%\end{equation}
-%where $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $\bm{\nabla} h$ is the slope and $m_e$ and $n_e$ are parameters chosen in to
-%constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}. Erosion rates $\dot{E}_{sr}$ due to short-range processes are
-%implemented by assuming that transport is linearly proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{sr}=K_d \bm{\nabla}^2 h
-%\end{equation}
-%where $K_d$ is the effective diffusivity. In addition, the slope is modified if it still exceeds a threshold angle after long- and short-range processes
-%\citep{Ueda2015}.
-
-%Offshore sedimentation processes are implemented similarly to hillslope processes but with a depth-dependent coefficient \citep{Sacek2012,Beucher2020}.
-%The sedimentation rate $\dot{S}$ is calculated as
-%\begin{equation}
-%\dot{S}=K_s \bm{\nabla}^2 h
-%\end{equation}
-%with
-%\begin{equation}
-%K_s=K^*_s \exp\left(\frac{h-h_{sl}}{h_0}\right)
-%\end{equation}
-%where $K^*_s$ is a constant, $h_{sl}$ is the sea level and $h_0$ is a characteristic length. In case of sticky air, slope is determined by means of the markers
-%chain, which is vertically corrected considering erosion and sedimentation rates. After the correction, continental markers above the markers chain are
-%transformed in air markers and air markers below the markers chain become sediments. Similarly, slope is determined considering the nodes on the top boundary
-%in case of true free surface and they are vertically corrected as for the markers chain. In this case, continental markers above the top boundary are simply
-%deleted.
-
-\section{Non-linear rheologies}\label{sec:nonlinear}
-Non-linear rheologies are implemented combining viscous creep (dislocation and diffusion) and plastic yielding. For each marker, diffusion and dislocation
-viscosity ($\eta_{df|ds}$) can be determined as follows \citep{Karato1993,Warren2008b,Wang2016}:
-\begin{equation}\label{eq:viscous}
-\eta_{df|ds}=f_s \left(\frac{d^m}{A}\right)^{\frac{1}{n}} \left (\frac{2^{\frac{1-n}{n}}}{3^{\frac{n+1}{2n}}} \right)
-I_2^{\frac{(1-n)}{n}}\exp\left(\frac{Q+pV}{nRT}\right)
-\end{equation}
-where $f_s$ is a scaling factor used to represent lithologies that are stronger or weaker than the base set, $d$ is the grain size, $m$ is the grain size exponent,
-$n$ is the stress exponent, $A$ is the uniaxial pre-exponential factor, $I_2$ is the square root of the second invariant of the strain rate tensor, $Q$ is the
-activation energy, $p$ is the pressure, $V$ is the activation volume, $R$ is the gas constant and $T$ is the temperature.
-Pressures and temperatures are determined by the interpolation of the nodal parameters, while strain rates are calculated by means of the
-derivative of the velocities on the nodes. In case of diffusion creep $n=1$ and $m>0$, in case of dislocation creep $n>1$ and $m=0$. The viscous creep $\eta_{cp}$
-is then calculated as the harmonic average between $\eta_{df}$ and $\eta_{ds}$:
-\begin{equation}\label{eq:visco_creep}
-\eta_{cp}=\left(\frac{1}{\eta_{df}}+\frac{1}{\eta_{ds}}\right)^{-1}
-\end{equation}
-The implementation of the non-linear viscous creep viscosity depending on the strain rate is tested by means of the slab detachment benchmark
-(Section \ref{sec:slab}).
-
-Plastic yielding is implemented rescaling $\eta_{cp}$ in order to limit the stress below the yield stress $\sigma_y$ \citep{Thieulot2008,Thieulot2014,Glerum2018},
-obtaining
-\begin{equation}\label{eq:plastic}
-\eta_{pl}=\frac{\sigma_y}{2I_2}
-\end{equation}
-where $\eta_{pl}$ is the plastic viscosity and the yield stress is determined following the Drucker-Prager criterion, such as
-\begin{equation}\label{eq:yield}
-\sigma_y=C\;\cos(\phi)+p\;\sin(\phi)
-\end{equation}
-where $C$ is the cohesion and $\phi$ is the internal friction angle. In case of negative pressure it is imposed equal to 0, so that negative yield stress are
-excluded. The correctness of the non-linear solution in case of plastic viscosity with variable internal friction angle is verified performing indenter and
-brick experiments (Sections \ref{sec:indenter} and \ref{sec:brick}, respectively).
-
-Plastic yielding and viscous creep are then combined to obtain a viscoplastic viscosity $\eta_{vp}$ as follows, assuming that they are independent processes
-\citep{Karato2008,Glerum2018}:
-\begin{equation}\label{eq:viscoplastic}
-\eta_{vp}=\min(\eta_{cp},\eta_{pl})
-\end{equation}
-Finally, effective viscosity $\eta_{eff}$ is capped by the minimum and the maximum viscosity ($\eta_{min}$ and $\eta_{max}$, respectively) to avoid extremely
-low or high viscosity \citep{Glerum2018} as follows
-\begin{equation}\label{eq:effective}
-\eta_{eff}=\min(\max(\eta_{vp},\eta_{min}),\eta_{max})
-\end{equation}
-Viscous creep and plastic yielding are non-linear rheologies because of their dependence on the velocity field through pressure and strain rates. Therefore,
-the solution is determined by means of Picard-type iterations, until convergence of the velocity field \citep{Glerum2018}. The convergence is verified at each
-iteration $i$ via the nonlinear residual $\bm{\mathcal{R}}^i$ that can be determined as
-\begin{equation}\label{eq:convergence}
-\bm{\mathcal{R}}^i= \mathbb{K} (\eta_{eff}(\dot{\bm{\epsilon}}^{i-1},p^{i-1}))\cdot \bm{v}^{i-1} - \bm{f}^i
-\end{equation}
-%\begin{equation}\label{eq:convergence}
-%\chi=1-\frac{\langle (\bm{v}^i-\langle\bm{v}^i\rangle)\cdot (\bm{v}^{i+1}-\langle\bm{v}^{i+1}\rangle) \rangle}{\sqrt{\sigma^i\sigma^{i+1}}}
-%\end{equation}
-%Iterations are performed until either convergence ($\chi<tol$) or maximal number of non-linear iterations ($it_{max}$) is reached.
-where $\mathbb{K}$ is the Stiffness matrix and $\bm{f}$ is the right hand side vector \citep{Spiegelman2016,Glerum2018}.  The $L_2$-norm is extracted from
-$\bm{\mathcal{R}}^i$ and it is normalised as follows
-\begin{equation}\label{eq:convergence_res}
-\frac{||\bm{\mathcal{R}}^i||_2}{||\bm{\mathcal{R}}^0||_2}
-\end{equation}
-where $||\bm{\mathcal{R}}^0||_2$ is the $L_2$-norm of the first nonlinear iteration.
-Since often the normalised nonlinear residual drop very quickly over the first 2-3 iterations, also the $L_2$-norm of velocity and pressure residuals are
-calculated as
-\begin{equation}\label{eq:convergence_vel}
-\frac{||\Delta\bm{v}||_2}{||\bm{v}||_2}=\frac{||\bm{v}_i-\bm{v}_{i-1}||_2}{||\bm{v}_{i}||_2}
-\end{equation}
-and
-\begin{equation}\label{eq:convergence_press}
-  \frac{||\Delta p||_2}{||p||_2}=\frac{||p_i-p_{i-1}||_2}{||p_{i}||_2}
-\end{equation}
-respectively. Iterations are performed until either maximal number of non-linear iterations ($it_{max}$) is reached or all the normalised nonlinear residual, 
-the velocity residual and the pressure residual converge under a defined tolerance $tol$.
-
-\section{Healing and weakening}
-The temporal evolution of the accumulated strain $\epsilon$ has been implemented as in \citet{Fuchs2019,Fuchs2021}, as following
-\begin{equation}\label{eq:healing}
-  \frac{d\epsilon}{dt}=I_2 - \epsilon H_{rate}(T)
-\end{equation}
-where the first term on the right-hand side is a source term given by the second invariant of the strain rate $I_2$ and the second term a
-temperature-dependent healing factor $H_{rate}$, calculated as
-\begin{equation}\label{eq:healing_rate}
-  H_{rate}(T)=B \exp \left[-\frac{\mu}{2}\left(\frac{1}{T+1}-\frac{1}{2}\right)\right]
-\end{equation}
-where $B$ is a constant describing the time scale of healing, while $\mu$ and $T$ are the non-dimensional temperature activation and temperature,
-respectively \citep{Fuchs2019,Fuchs2021}.
-
-Strain softening is taken into account for both viscous creep and plastic viscosity \citep{Huismans2003,Babeyko2005,Huismans2005,Sobolev2005,Warren2008a} by
-means of the accumulated strain $\epsilon$ memorised by each marker. Softening in the viscous creep determines a linear decrease of $\eta_{vc}$ by means of a
-viscous strain softening factor $W_S$ that increases linearly from $W_{S_0}$ to $W_{S_{\infty}}$ for $\epsilon_{S_0}<\epsilon_S<\epsilon_{S_{\infty}}$ 
-\citep{Warren2008a}. This viscous softening can be related to strain-induced grain size reduction \citep{Warren2008a}. Differently, plastic softening is 
-simulated with a linear decrease of internal friction angle $\phi(\epsilon)$ and cohesion $C(\epsilon)$ in according to
-\begin{eqnarray}
-\label{eq:friction1}\phi(\epsilon)&=&\phi_0+(\phi_{\infty}-\phi_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}\\
-\label{eq:friction2}C(\epsilon)&=&C_0+(C_{\infty}-C_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}
-\end{eqnarray}
-where $\phi_0$, $C_0$ and $\phi_{\infty}$, $C_{\infty}$ are internal friction angle and cohesion for $\epsilon_0$ and $\epsilon_{\infty}$, respectively
-\citep{Huismans2003,Huismans2005,Warren2008a,Thieulot2014}. Plastic softening approximates deformation-induced softening of faults and brittle shear zones
-\citep{Warren2008a}.
-
-\section{The energy equation}\label{sec:en_eq}
-The stabilisation of the advection term of the energy equation (Eq. \ref{eq:energy}) needed to avoid possible oscillations in the thermal solution in the case
-when advection dominates over diffusion, is implemented by means of a streamline-upwind Petrov–Galerkin (SUPG) method, for which the advection term is modified
-follows the discussion in \citet{Thieulot2011} and \citet{Thieulot2014} (see Section \ref{sec:advection} for benchmark). The total internal heat production
-$H_{tot}$ in Eq. \ref{eq:energy} includes radiogenic heating $H_r$, shear heating $H_s$ and adiabatic heating $H_a$, where
-\begin{equation}\label{eq:shear}
-H_s=2\eta \dot{\bm{\epsilon}} : \dot{\bm{\epsilon}} = 2\eta(\dot{\epsilon}_{xx}^2+\dot{\epsilon}_{yy}^2+2\dot{\epsilon}_{xy}^2)
-\end{equation}
-and
-\begin{equation}\label{eq:adiabatic}
-H_a=T \alpha \frac{Dp}{Dt} \approx -\alpha T \rho g_y v
-\end{equation}
-The correctness of the implementation and the computation of shear and adiabatic heating is verified performing an exercise with an analytical solution
-(Section \ref{sec:simple_shear}) and Exercise 9.4 in \citet{Gerya2010} (Section \ref{sec:shear}). Furthermore, mantle convection and thin layer entrainment
-benchmarks (Sections \ref{sec:mantle} and \ref{sec:thin}, respectively) are performed to verify that the code solve correctly Eqs. \ref{eq:momentum_penalty}
-and \ref{eq:energy} in case of a temperature field characterised by an initial perturbation and with a temperature-dependent density. At last, the viscoplastic
-mantle convection benchmark is performed to check also the viscoplasticity in case of a initial perturbed temperature field (Section \ref{sec:visco_mantle}).
-
-\section{Phase transitions and hydration}\label{sec:phase}
-Variations of effective density, specific heat and coefficient of thermal expansion during the evolution of crustal and mantle materials at different
-pressure-temperature ($p$-$T$) conditions are computed by means of version 6.8.6 of Perple\_X software package \citep{Connolly2005}, similarly to the
-implementation described in \citet{Marotta2020}. By default, mineral assemblages and properties of each lithology (oceanic crust, lower and upper continental
-crust, sediments and dry or hydrated mantle) are calculated for temperatures from \SIrange{330}{1600}{\kelvin}, with increments of approximately \SI{4}{\kelvin},
-and pressures from \SIrange{0.1}{30}{\giga\pascal}, with increments of approximately \SI{0.03}{\giga\pascal}, for a total of almost 100000 points.
-
-The introduction of phase transitions produces two main effects that must be taken into account in numerical models: variations in density and release/absorption
-of latent heat \citep{Tackley2010}. Effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into
-account considering effective coefficient of thermal expansion \citep{Christensen1985,Zhong2015}. Similarly, effects of latent heat in the energy equation
-(eq \ref{eq:energy}) are taken into account by considering effective specific heat and coefficient of thermal expansion
-\citep{Christensen1985,Tackley2010,Zhong2015}. Effective density, specific heat and coefficient of thermal expansion are included in Perple\_X files and are
-assumed by each Lagrangian marker in according to its $p$-$T$ conditions.
-
-Hydration processes related to dehydration of subducting lithosphere \citep{Schmidt1998,Liu2007,Faccenda2009,Faccenda2010,Faccenda2014,Rosas2016} strongly
-influence the thermo-mechanics inside the mantle wedge, mainly because of both weakening effects on the mantle rheology and density variations in case of
-mantle serpentinization \citep{Gerya2002,Honda2003,Arcay2005,Roda2010,Regorda2017}. In the case that hydration is switched on, the amount of bound and free
-water is memorised by each marker, following the implementation of \citet{Quinquis2014}; hydration and dehydration processes are related to the amount of bound
-water of each marker and to the maximum amount of water it can transport, i.e. if the amount of bound water exceeds the maximum amount of water, the marker
-dehydrates and releases free water that can hydrates under-saturated markers. Maximum water content of each marker is determined as function of lithology and
-$p$-$T$ conditions and it is calculated using Perple\_X, in the same way than effective density, specific heat and coefficient of thermal expansion. Bound water
-is advected together with the markers, neglecting the effect of bound water diffusion, while free water simply migrates vertically and is not coupled to the
-solid-phase flow of the mantle wedge \citep{Arcay2005,Quinquis2014}. The correctness of the migration scheme of the free water and its absorption into bound
-water by under-saturated markers are verified performing the experiment described in \citet{Quinquis2014} (Sec. \ref{sec:quinquis}).
-
-The mantle viscosity weakening related to the amount of water has been extensively studied \citep{Chopra1981,Hirth2003} and it is implemented in according to
-\citet{Arcay2005} and \citet{Horiuchi2016}, as follows
-\begin{equation}\label{eq:wet_visc}
-\eta_{wet}=\eta_{dry}\left(\left[1-\frac{1}{f_v}\right]\exp\left(-\frac{[OH^-]}{[OH^-]_0}\right)+\frac{1}{f_v}\right)
-\end{equation}
-where $f_v$ is the viscosity reduction factor between dry and wet conditions, $[OH^-]$ is the water content and $[OH^-]_0$ is a reference water content set to
-620 ppm (0.062 wt.\%) \citep{Arcay2005}. By default, $f_v$ is set to 100, in according to \citet{Arcay2005} and \citet{Horiuchi2016}.
-
-\section{Mantle melting}\label{sec:melt}
-Melt fraction in the mantle depends on temperature ($T$, in \textdegree C), pressure ($p$, in \si{\giga\pascal}) and the water content in the melt ($X_{H_2O}$, in wt.\%).
-The determination of the melt fraction $M$ is implemented as explained by \citet{Katz2003}, according to the successive modification by \citet{Langmuir2006} and
-\citet{Kelley2010}, as follows
-\begin{eqnarray}\label{eq:melt_fr}
-M(p,T,X_{H_2O})&=&-T+T_{sol}+(x_t\cdot ln(p+y_t))\cdot X_{H_2O}-\nonumber\\
-&&+K_m\left(\frac{X_{H_2O}^{bulk}}{D_{H_2O}(1-X_{H_2O})+X_{H_2O}}\right)^{\gamma}
-\end{eqnarray}
-where $T_{sol}$ is the temperature of the dry solidus, $D_{H_2O}$ is the partition coefficient and $x_t\cdot log(p+y_t)$ indicates the pressure dependence of
-the melting curve \citep{Kelley2010}. $x_t$, $y_t$, $K_m$ and $\gamma$ are parameters chosen in according to \citet{Kelley2010}.
-However, $X_{H_2O}$ depends on the melt fraction as
-\begin{equation}\label{eq:melt_wt}
-X_{H_2O}(M)=\frac{X_{H_2O}^{bulk}}{D_{H_2O}+M(1-D_{H_2O})}
-\end{equation}
-and a numerical solution can be found using a root-finder method \citep{Katz2003,Wang2016}. Furthermore, the content of water in the melt is limited by the
-pressure-dependent saturation concentration of water in the melt, determined as
-\begin{equation}\label{eq:melt_max}
-X_{H_2O}^{sat}=\chi_1 p^{\lambda_m}+\chi_2 p
-\end{equation}
-where $\chi_1$, $\chi_2$ and $\lambda_m$ are parameters chosen according to \citet{Katz2003}, \citet{Langmuir2006} and \citet{Kelley2010}. The correctness in
-the determination of melt fractions and water contents in the melt are tested comparing model results with experimental curves obtained by \citet{Katz2003}
-(Sec. \ref{sec:katz}).
-
-In case of melting, viscous creep viscosity ($\eta_{vp}$) is modified according to \citet{Wang2016}, as
-\begin{equation}\label{eq:melt_visc}
-\eta_{melt}=\eta_{vp}\cdot \exp(\alpha_m M)
-\end{equation}
-where $\alpha_m$ are melt fraction factors for dislocation and diffusion creep, chosen in according to \citet{Wang2016}.
-
-Effective density ($\rho_{eff}$) for partially molten rocks is calculated as
-\begin{equation}\label{eq:melt_dens}
-\rho_{eff}=\rho_s(1-M)+\rho_m M
-\end{equation}
-where $\rho_s$ and $\rho_m$ are the densities of the solid and the molten rock, respectively \citep{Gerya2004b,Gerya2007,Wang2016}. The density of the
-solid rock is extracted by Perple\_X, while the density of the molten portion is calculated as
-\begin{equation}\label{eq:melt_dens_M}
-\rho_m(p,T)=\rho_0[1-\alpha(T-T_0)][1+\beta(p-p_0)]
-\end{equation}
-where $\rho_0$ is the density at the reference temperature ($T_0$) and pressure($p_0$), and $\alpha$ and $\beta$ are the thermal expansion and compressibility
-coefficients, respectively \citep{Gerya2004b,Gerya2007,Wang2016}.
-
-As for the phase transitions, effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into account
-considering effective coefficient of thermal expansion. Similarly, latent heat due to melting/crystallisation can be implicitly included in the energy equation
-(Eq. \ref{eq:energy}) by considering effective specific heat ($C_{p_{eff}}$) and thermal expansion ($\alpha_{eff}$) of the partially molten rocks, in according to
-\begin{eqnarray}
-\label{eq:cp_effective}C_{p_{eff}}&=&C_p+H_L\left(\frac{\partial M}{\partial T}\right)_p \\
-\label{eq:alpha_effective}\alpha_{eff}&=&\alpha+\rho\frac{H_L}{T}\left(\frac{\partial M}{\partial p}\right)_T
-\end{eqnarray}
-where $C_p$ and $\alpha$ are the specific heat and the coefficient of thermal expansion of the solid rock, respectively, and $H_L$ is the latent heat of the
-molten rock \citep{Gerya2004b,Gerya2007,Tackley2010}.
-
-\section{Benchmarks}\label{sec:benchmark}
-\subsection{Error measurements}\label{sec:error}
-In order to determine the accuracy of velocity and pressure field of the benchmarks, the $L_2$-norm is computed by numerical integration on the Gauss-Legendre
-quadrature points. $L_2$-norm for pressure and velocity errors can be evaluated as
-\begin{eqnarray}
-\textrm{err}_p&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|p^n_i(r_q)-p^a_i(r_q)|^2\right)w_q|J_q|}\nonumber \\
-\textrm{err}_v&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|u^n_i(r_q)-u^a_i(r_q)|^2+|v^n_i(r_q)-v^a_i(r_q)|^2\right)w_q|J_q|}\nonumber
-\end{eqnarray}
-respectively, where $n_e$ is the number of elements, $n_q$ is the number of quadrature points per element, $p^n_i(r_q)$ and $p^a_i(r_q)$ are the numerical and
-analytical pressure, respectively, in each quadrature point $q$, $w_q$  and $J_q$ are the weight and the Jacobian at the quadrature point
-$q$, $u^n_i(r_q)$, $v^n_i(r_q)$, $u^a_i(r_q)$ and $v^a_i(r_q)$ are the numerical and analytical velocities, respectively, in each quadrature point $q$.
-
-Other quantities used as comparison with original benchmarks are the root-mean-square velocity over the whole domain and over the surface
-\begin{eqnarray}
-v_{\textrm{rms}}&=&\sqrt{\int_{\Omega}|\bm{v}|^2 d\Omega}\nonumber \\
-v_{\textrm{rms}}^{top}&=&\sqrt{\int_0^1 u^2 \bigg|_{y=1} dx}\nonumber
-\end{eqnarray}
-respectively, top and bottom Nusselt numbers
-\[\textrm{Nu}^{top/bottom}=-\int_0^1\frac{\partial T}{\partial y} \bigg|_{y=1/y=0} dx\]
-the average rate of work done against gravity
-\[\left\langle W \right\rangle=\int_{\Omega}T u_y d\Omega\]
-the average rate of viscous dissipation
-\[\left\langle \Phi \right\rangle=\int_{\Omega}\tau_{ij}\dot{\epsilon}_{ij} d\Omega\]
-and the percentage error between $\left\langle W \right\rangle$ and $\left\langle \Phi \right\rangle/Ra$
-\[\delta=\frac{\left|\left\langle W \right\rangle -\frac{\left\langle \Phi \right\rangle}{Ra} \right|}{\max \left(\left\langle W \right\rangle,
-\frac{\left\langle \Phi \right\rangle}{Ra}\right)}\times100\]
-
-\subsection{Stokes flow}\label{sec:stokes}
-The problem consists of determining velocity field $(u,v)$ and pressure $p$ in case of a manufactured solution with prescribed body forces such as
-\begin{eqnarray}
-b_1&=&(12-24y)x^4+(-24+48y)x^3+(-48y+72y^2-48y^3+12)x^2+\nonumber \\
-&&+(-2+24y-72y^2+48y^3)x+1-4y+12y^2-8y^3\nonumber \\
-b_2&=&(8-48y+48y^2)x^3+(-12+72y-72y^2)x^2+(4-24y+48y^2-48y^3+\nonumber \\
-&&+24y^4)x-12y^2+24y^3-12y^4\nonumber
-\end{eqnarray}
-for which the exact solution is:
-\begin{eqnarray}
-u(x,y)&=&x^2(1-x)^2(2y-6y^2+4y^3)\nonumber \\
-v(x,y)&=&-y^2(1-y)^2(2x-6x^2+4x^3)\nonumber \\
-p(x,y)&=&x(1-x)-\frac{1}{6}\nonumber
-\end{eqnarray}
-The domain is a unit square a constant viscosity ($\eta=1$) and the penalty parameter is set to $\lambda=10^7$. Velocity boundary conditions are set to no slip
-$(\bm{v}=\bm{0})$ on all boundaries. The problem is performed for different grid resolution between \num{8x8} and \num{1024x1024} elements.
-
-The errors between the analytical solution and the numerical prediction of the velocity and the pressure field are calculated as in Section \ref{sec:error}.
-Fig. \ref{fig:errors}a shows that both velocity and pressure field converge to the exact solution with the decrease of the element size, following the
-theoretical convergence rate. The convergence of pressure, in contrast with the observation by \citet{Donea2003} for $Q_1 \times P_0$ elements, indicate the
-effectiveness of the smoothing procedure. Smoothed pressure field for grid resolution of \num{128x128} elements is shown in Fig. \ref{fig:errors}b (continuous
-red line), in comparison with the analytical solution (dashed blue line). Solve times and memory usage needed to generate the solution are shown in
-Fig. \ref{fig:MUMPS}. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Solver/Stokes_Flow}.
-
-\begin{figure}
-\centering
-\includegraphics[width=\linewidth]{./Figures/errors.pdf}
-\caption{Velocity and pressure error for the Stokes flow experiment between generated and analytical solution as a function of element size (panel a),
-and comparison between smoothed pressure and analytical solution as function of $x$ coordinate for a grid resolution of \num{128x128} elements (panel b).}
-\label{fig:errors}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\linewidth]{./Figures/MUMPS.pdf}
-\caption{Panel a) Analysis, factorisation, solution and total solve times (panel a) and factorisation memory usage (panel b) as a function of the total number
-of degrees of freedom for the Stokes flow experiment.}
-\label{fig:MUMPS}
-\end{figure}
-
-\subsection{Markers advection}\label{sec:runge}
-The 2nd-order Runge-Kutta advection scheme in space is tested by means of the Zalesak disk test \citep{Zalesak1979,Thieulot2014}. The benchmark is performed
-in a unit square domain with a grid resolution of \num{32x32} elements and 1000000 markers and values of Courant number between 0.25 and 1.
-At $t=0$, the disk is centred at position $(0.5;0.75)$ with a radius $R=0.15$ and has a vertical fissure 0.05 wide and 0.2 high (orange in Fig. \ref{fig:runge})
-and with the velocity field prescribed in the entire domain as
-\begin{eqnarray}
-u(x,y)&=&2\pi\left(y-\frac{L_x}{2}\right)\nonumber \\
-v(x,y)&=&-2\pi\left(x-\frac{L_x}{2}\right)\nonumber
-\end{eqnarray}
-the markers are back to their initial location after a $2\pi$ rotation ($t=1$). The clockwise rotation of the disk at $t=0.25$, 0.5 and 0.75 
-is shown in Fig. \ref{fig:runge} (red, blue and green, respectively). The distance from the centre of 1 marker is calculated throughout the rotation to
-evaluate the error for different values of Courant number. As expected, the error increase with the increasing of the time step, as shown in 
-Fig. \ref{fig:runge_err}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/Zalesak_Disk}.
-
-\begin{figure}
-\centering
-\includegraphics[width=200px]{./Figures/Zalesak.pdf}
-\caption{Position of the Zalesak disk at $t=0$, 0.25, 0.5 and 0.75 (orange, red, blue and green, respectively) for the case with Courant number of 1.}
-\label{fig:runge}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/zal_marker.pdf}
-\caption{Distance from the centre as function of time for values of Courant number of  of 0.25, 0.3, 0.5, 0.75 and 1 (green, orange, blue, black and red,
-respectively). R indicates the distance from the centre considering a perfect circle trajectory.}
-\label{fig:runge_err}
-\end{figure}
-
-\subsection{Conservative Velocity Interpolation (CVI)}\label{sec:CVI}
-The CVI correction is checked by means of the Stokes flow experiment by \citet{Donea2003}, as presented in \ref{sec:stokes}. The advection of Lagrangian
-markers is performed using either a 2nd-order or a 4th-order Runge-Kutta scheme with CFL=0.5 and an initial random distribution of 25 markers per element.
-Fig. \ref{fig:CVI} shows the minimum and maximum values of markers per element throughout the experiment with and without the CVI correction and using
-different Runge-Kutta schemes. The introduction of the CVI correction allows the
-tests to run for times longer than $t=2000$ without show empty elements, while tests without the CVI correction stop at approximately $t=600$, when at least
-one element goes empty. Fig. \ref{fig:CVI_time} shows the different number of markers per element and the markers distribution at $t=600$ obtained without and
-with the CVI correction (panels b and c, respectively) in case of a 2nd-order Runge-Kutta scheme. All data can be found at
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/CVI}.
-\begin{figure}
-\centering
-\includegraphics[width=\linewidth]{./Figures/CVI.pdf}
-\caption{Maximum and minimum number of markers per element throughout the experiment, with (green and orange lines) or without (purple and blue lines)
-the CVI correction and using either a 2nd-order (blue and orange lines) or a 4th-order (purple and green lines) Runge-Kutta scheme.}
-\label{fig:CVI}
-\end{figure}
-\begin{figure}
-\centering
-\includegraphics[width=\linewidth]{./Figures/CVI_time.pdf}
-\caption{Number of markers per element and markers distribution through time using 2nd-order Runge-Kutta scheme.
-Panels a: number of markers per element (panel a1) and distribution of the markers (panel a2) after the first
-iteration; panels b and c: comparison between number of markers per element (panels b1 and c1) and markers distribution (panels b2 and c2) without and with
-the CVI correction at $t=600$; panels d: number of markers per element (panel d1) and markers distribution (panel d2) in case of CVI correction at $t=2000$.}
-\label{fig:CVI_time}
-\end{figure}
-
-\subsection{Poiseuille flow}\label{sec:poiseuille}
-The domain is a rectangle with $L_x=2$ and $L_y=1$ and constant density and viscosity ($\rho=1$ and $\eta=1$), gravity acceleration $\bm{g}=\bm{0}$ and penalty
-parameter $\lambda=10^8$. The grid is composed by \num{40x20} elements. Velocity boundary conditions are set to no slip $(\bm{v}=\bm{0})$ at the top and the
-bottom, and a parabolic profile is imposed on the sides, with $u=y(1-y)$ and $v=0$. The analytical solution is then given by:
-\begin{eqnarray}
-u(x,y)&=&y(1-y)\nonumber \\
-v(x,y,)&=&0\nonumber \\
-p(x,y)&=&2\eta\left(\frac{L_x}{2}-x\right)\nonumber
-\end{eqnarray}
-
-The velocity field predicted by the model follows the expected parabolic profile (Fig. \ref{fig:poiseuille}a and Fig. \ref{fig:poi_plot}a) and, in case of the
-classic penalty method with no iterations, the pressure is clearly related to the divergence by means of the penalty parameter (Fig. \ref{fig:poiseuille}b and
-c and Fig. \ref{fig:poi_plot}b). Fig. \ref{fig:poiseuille}d shows that one Uzawa iteration is sufficient to bring the divergence down to \num{e-15}, with no
-correlation with the pressure field. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Poiseuille%20Flow}.
-
-\begin{figure}
-\includegraphics[width=\linewidth]{./Figures/Poiseuille.pdf}
-\caption{Velocity field (panel a), pressure (panel b) and divergence velocity of a Poiseuille flow in case of the classic penalty method (no iterations) and
-after one Uzawa iteration (panel c and d, respectively).}
-\label{fig:poiseuille}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\linewidth]{./Figures/Analytical.pdf}
-\caption{Velocity (panel a) and pressure (panel b) field predicted by the model for a Poiseuille flow with respect with their analytical solutions. The
-velocity field is plotted as function of the vertical coordinate in $L_x/2$ and the pressure is plotted as function of the x coordinate in $L_y/2$.}
-\label{fig:poi_plot}
-\end{figure}
-
-\subsection{Instantaneous 2D sphere}\label{sec:ist_sphere}
-The domain is a unit square with gravity $\bm{g}=(0,-1)$. The fluid has constant density and viscosity ($\rho_f=1$ and $\eta_f=1$). The sphere is in the middle
-of the domain with a radius R = 0.123456798 and has constant density and viscosity ($\rho_s=10^{-2}+\rho_f$ and $\eta_s=10^3 \cdot \eta_f$).
-We distinguish three different types of velocity boundary conditions:
-\begin{enumerate}
-\item FS: free slip conditions on all sides.
-\item NS: no slip conditions on all sides.
-\item OT: free slip conditions on the sides and the bottom, and open boundary at the top.
-\end{enumerate}
-All three cases are performed for different grid resolutions, between \num{16x16} and \num{512x512} elements with 50 randomly distributed markers per element,
-and different average schemes for the viscosity (harmonic, geometric and arithmetic). The velocity in the centre of the sphere $\bm{v}(0.5;0.5)$), minimum and
-maximum velocities ($u_{min}$, $u_{max}$, $v_{min}$, $v_{max}$) and pressures ($p_{min}$, $p_{max}$) on the entire domain, $v_{\textrm{rms}}$ and average pressure on the
-entire domain ($p_{avg}$) are compared with solutions generated by ASPECT \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and all results
-can be found at \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere2D} and 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Instantaneous_2D_Sphere}. Results in terms of the $v_{\textrm{rms}}$ for NS model 
-are also shown in Fig. \ref{fig:inst_sphere}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/vrms_NS.pdf}
-\caption{$v_{\textrm{rms}}$ as function of element size for the instantaneous 2D sphere experiment in case of no slip boundary conditions and with different average
-schemes for the viscosity. Results are compared with results obtained by other numerical codes.}
-\label{fig:inst_sphere}
-\end{figure}
-
-\subsection{Rayleigh-Taylor instability}\label{sec:rayleigh}
-This problem has been originally presented by \citet{vanKeken1997} and here it is performed as the isoviscous case (Case 1a). The domain has $L_x=0.9142$ and
-$L_y=1$ and gravity $\bm{g}=(0,-1)$. Two fluids with same constant viscosities ($\eta_1=\eta_2=1$) and different densities ($\rho_1=1000$ and $\rho_2=1010$),
-with the lighter fluid at the bottom. The initial interface between the fluids is given by $y(x)=0.2+0.02 \cos \left(\frac{\pi x}{L_x}\right)$. The experiment
-is performed with different grid sizes (\num{50x50}, \num{80x80}, \num{100x100} and \num{256x256}). A total of 1960000 markers are randomly distributed at the
-beginning of the simulation. Velocity boundary conditions are set to no slip at the top and the bottom, and to free slip at the sides of the domain.
-
-$v_{\textrm{rms}}$ as function of time is reported in Fig. \ref{fig:RT}, matching well with results shown by \citet{vanKeken1997}, \citet{Tackley2003} and
-\citet{Thieulot2014}. Fig. \ref{fig:rayleigh} shows the evolution of the experiment at different time steps. All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Rayleigh_Taylor_experiment/ISOVISCOUS}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/RT.pdf}
-\caption{$v_{\textrm{rms}}$ of the Rayleigh-Taylor experiment as function of time for different resolution of the grid.}
-\label{fig:RT}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=300px]{./Figures/Rayleigh.pdf}
-\caption{Evolution of the Rayleigh-Taylor experiment for a grid of \num{256x256} elements at $t=0, 500, 1000$ and $2000$ (panels a, b, c and d, respectively).}
-\label{fig:rayleigh}
-\end{figure}
-
-\subsection{Falling blocks}\label{sec:block}
-This benchmark is described as proposed by \citet{Gerya2003a}, \citet{Gerya2010} and \citet{Thieulot2011}. The domain is square with $L_x=L_y=$ \SI{500}{\km}
-and the grid is composed by \num{50x50} elements with 25 markers in each element. The block is initially centred at ($x=$ \SI{250}{\km}; $y=$ \SI{400}{\km})
-and has a size of \num{100x100} km (Fig. \ref{fig:block}a). The fluid surrounding the block has $\eta_f=$ \SI{e21}{\pascal\per\s} and $\rho_f=$
-\SI{3200}{\kg\per\cubic\m}. The benchmark tests with different viscosities of the block, with $\eta_b$ from \SIrange{e15}{e27}{\pascal\per\s}. In each
-experiment also the density $\rho_b$ of the block varies from \SIrange{3220}{9900}{\kg\per\cubic\m}. Velocity boundary conditions are set to free slip
-conditions on all sides of the domain. Density distributions at $t=$ \SI{20}{\mega\year} for $\rho_b=$ \SI{3300}{\kg\per\cubic\m} and different $\eta_b$
-are plotted in Fig. \ref{fig:block}b-f, showing that the code correctly preserve the block geometry in case of large viscosity contrast, with a block stiffer
-than the surrounding fluid (Fig. \ref{fig:block}f).
-
-The velocity in the centre of the falling block at $t=0$ is measured for all experiments and, since the velocity should increase with density contrast, the
-quantity $\bm{v}/(\rho_b-\rho_f)$ is plotted as function of the viscosity contrast. All results perfectly match with those from \citet{Gerya2010} and they
-line up on a single curve, demonstrating that the code can correctly deal with large viscosity and density contrasts (Fig. \ref{fig:falling}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Falling%20blocks}.
-
-\begin{figure}
-\includegraphics[width=300px]{./Figures/Block.pdf}
-\centering
-\caption{Density evolution of the falling block experiment at $t=$ \SI{0}{\mega\year} (panel a) and $t=$ \SI{20}{\mega\year} for different viscosities of the
-block (panels b-f).}
-\label{fig:block}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Falling.pdf}
-\caption{Initial velocity relative to the density contrast at the centre of the falling block as function of the viscosity contrast between the block and the
-surrounding fluid.}
-\label{fig:falling}
-\end{figure}
-
-\subsection{2D time-dependent Stokes sphere with deformable free surface}\label{sec:sphere}
-This experiment is performed in a unit square domain with the gravity acceleration fixed to $g_y=-1$. The sphere has $\rho_s=2$ and $\eta_s=10^3$ and is
-initially centred at ($0.5;0.6$) with radius $R=0.123456789$. The fluid surrounding the sphere has $\rho_f=1$ and $\eta_f=1$ and occupies the domain for
-$y\leq 0.75$, while for $y>0.75$ the air has $\rho_a=0$ and $\eta_a=10^{-3}$. Velocity boundary conditions are set to free slip on all sides. The Courant
-number is set to 0.25. The experiments run for 200 s using grid resolutions from \num{150x150} to \num{512x512} elements, with 25 randomly distributed markers
-per element, and for different average schemes for the viscosity. The interface between the fluid and the air is tracked by means of the markers chain. Results
-in terms of velocity and pressure fields and topography variations are compared with results obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and they can be found at
-\url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere_fs2D} and
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Time_dependent_sphere}. Results in terms of the $v_{\textrm{rms}}$ in case of 
-an arithmetic average are also shown in Fig. \ref{fig:inst_spherefs}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/vrms_arithm.pdf}
-\caption{$v_{\textrm{rms}}$ for different grid resolutions as function of time for the 2D time-dependent Stokes sphere experiment in case of an arithmetic average.
-Results are compared with results obtained by ASPECT.}
-\label{fig:inst_spherefs}
-\end{figure}
-
-\subsection{Free surface stabilisation}\label{sec:stab}
-The benchmark is performed as discussed in \citet{Kaus2010a} and \citet{Thieulot2014}. The domain is square with $L_x=L_y=$ \SI{500}{\km} and a grid resolution
-of \num{200x200} elements, each of them containing 16 markers. A buoyant fluid with $\rho_1=$ \SI{3200}{\kg\per\cubic\metre} and $\eta_1=$ \SI{e20}{\pascal\s}
-is overlain by a denser fluid with $\rho_2=$ \SI{3300}{\kg\per\cubic\metre} and $\eta_2=$ \SI{e21}{\pascal\s}. The initial interface between the fluids has an
-initial sinusoidal shape of 5 km amplitude. Velocity boundary conditions are set to free slip conditions at the sides and no slip at the bottom, while the top
-is free surface. The experiment is performed with various fixed time steps, without the use of the Courant number. The vertical position of the free surface at
-$x=L_x$ is tracked for each simulation, with or without the stabilisation algorithm.
-
-The results show an instability (drunken sailor effect) increasing the time step in case that the stabilisation algorithm is not activated
-($dt=$ \SI{4200}{\year}, green line in Fig. \ref{fig:stab}), as already observed by \citet{Kaus2010a} and \citet{Thieulot2014}. Activating the algorithm the
-instability is fixed and the simulation remains stable using time step up to \SI{50000}{\year} (red line in Fig. \ref{fig:stab}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Stabilisation_algorithm}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Kaus_random.pdf}
-\caption{Evolution of the y coordinate in $x=L_x$ as function of time for different time steps, with and without the stabilisation algorithm.}
-\label{fig:stab}
-\end{figure}
-
-\subsection{Topography relaxation}\label{sec:crameri}
-This experiment is performed as in \citet{Crameri2012} using free surface and sticky air (Case 1). The domain is rectangular with $L_x=$ \SI{2800}{\km} and
-$L_y=$ \SI{700}{\km} in case of free surface and $L_y=$ \SI{800}{\km} in case of sticky air ($h_{st}=$ \SI{100}{\km}). The grid is composed by \num{256x64}
-elements for the free surface \num{560x320} elements for the sticky air. For both cases, there is a mantle of 600 km thickness with $\eta_m=$ \SI{e21}{\pascal\s}
-overlain by a highly viscous cosine-shaped lithosphere with thickness between 93 and 107 km and $\eta_l=$ \SI{e23}{\pascal\s}. Both mediums have $\rho=$
-\SI{3300}{\kg\per\cubic\metre}. In case of sticky air experiment, the air has $\rho_a=$ \SI{0}{\kg\per\cubic\metre} and $\eta_a=$ \SI{e18}{\pascal\s}. The
-gravity acceleration is set to $g_y=$ \SI{-10}{\metre\per\square\s}. Velocity boundary conditions are set to no slip at the bottom and free slip at the sides
-of the domain. In case of sticky air, velocities on top are set to free slip conditions.
-
-The maximum topography as function of time ($h(t)$) can be analytically derived in according to \[h(t)=h_0 \exp(\gamma t)\]
-where $h_0=7$ km is the initial topography, $\gamma=\num{-0.2139e-11}$ is the characteristic relaxation rate and $t$ is the time (black line in Fig.
-\ref{fig:crameri}). Maximum topography at the characteristic relaxation time $t_{rlx}=$ \SI{14.825}{\kilo\year} can be found to be $h_{rlx}=$ \SI{2576}{\m}.
-The results show that a viscosity of the air of \SI{e18}{\pascal\s} is needed in case of $h_{st}=$ \SI{100}{\km} to correctly describe the topography relaxation
-for this problem, as pointed out by \citet{Crameri2012} (purple line in Fig. \ref{fig:crameri}). The case with a true free surface well follows the analytical
-solution (orange line in Fig. \ref{fig:crameri}). In particular, at the relaxation time in case of the true free surface predicts a maximum topography of
-\SI{2570}{\m}, with an error of \SI{6}{\m} with respect to the analytical solution. Results are also compared with results obtained with a free surface in
-UNDERWORLD (\num{256x64} elements) and SULEC (\num{401x201} elements) (with courtesy from \citealp{Crameri2012}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Topography_relaxation}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Crameri.pdf}
-\caption{Maximum topography as function of time for the Crameri benchmark (coloured lines) in comparison with the analytical solution (black line) and results
-obtained with UNDERWORLD (yellow line) and SULEC (blue line).}
-\label{fig:crameri}
-\end{figure}
-
-\subsection{Spontaneous subduction}\label{sec:subduction}
-This experiment is performed as presented by \citet{Schmeling2008}, in case of both a sticky air and a true free surface and using different average schemes
-for the viscosities. The domain is rectangular with $L_x=$ \SI{3000}{\km} and $L_y=$ \SI{700}{\km}, a grid resolution of \num{256x64} elements with 50 markers
-per element and Courant number of 0.01. At the beginning of the simulation, a 100 km-thick lithospheric layer with $\rho_m=$ \SI{3300}{\kg\per\cubic\metre}
-and $\eta_m=$ \SI{e23}{\pascal\s} is located at the top of the domain for $x$ between 1000 and 3000 km, while the asthenospheric mantle has $\rho_m=$
-\SI{3200}{\pascal\s} and $\eta_m=$ \SI{e21}{\pascal\s}. In addition, a 200 km-depth lithospheric slab is already subducted in the mantle in order to have
-a spontaneous subduction. In case of sticky air, $L_y=750$ km and a 50 km-thick air layer with $\eta_a$ from \SIrange{e19}{e21}{\pascal\s} overlie the mantle.
-Velocity boundary conditions are set to free slip on the sides and on the bottom of the domain. In case of sticky air, velocities on the top boundary are set
-to free slip conditions as well.
-
-The maximum depth of the slab is tracked for all the simulations (Fig. \ref{fig:subduction}) and they are compared with results obtained by
-\citet{Schmeling2008} with models with regular grids and comparable grid resolutions (rectangles in Fig. \ref{fig:subduction}). Low viscous sticky air models
-(i.e., \SI{e19}{\pascal\s}) enlighten a strong dependence of the sinking velocity with the chosen average scheme (dashed lines in Fig. \ref{fig:subduction}),
-while high viscous sticky air models (i.e., \SI{e21}{\pascal\s}) show a higher resistance at the trench, underestimating the correct solution (dotted lines in
-Fig. \ref{fig:subduction}). A high resistance at the trench can be also observed in case of a true free surface for the low resolution grid used (continuous
-lines in Fig. \ref{fig:subduction}), as pointed out by \citet{Schmeling2008}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Spontaneous_subduction}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Subduction.pdf}
-\caption{Maximum depth of the slab as function of time for all the simulations of the \citet{Schmeling2008} benchmark, in case of harmonic, geometric and
-arithmetic means (red, green and blue lines, respectively) using a sticky air or a true free surface (discontinuous and continuous lines, respectively).
-Red, green and blue rectangular areas indicate the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{500}{\km} in case of sticky air,
-while the grey rectangular area indicates the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{600}{\km} in case of true free surface.}
-\label{fig:subduction}
-\end{figure}
-
-\subsection{The slab detachment}\label{sec:slab}
-The slab detachment benchmark is performed as described by \citet{Schmalholz2011} and \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{1000}{\km}
-and $L_y=$ \SI{660}{\km} and a grid resolution of \num{256x256} elements. A non-linear viscous T-shaped layer with $\rho_l=$ \SI{3300}{\kg\per\cubic\metre} is
-placed at the top of the domain and surrounded by a linear viscous fluid with $\rho_f=$ \SI{3150}{\kg\per\cubic\m}. The top layer is 80 km-thick and a 250
-km-long and 80 km-wide slab is placed at $x=L_x/2$ (Fig. \ref{fig:slab}a). The effective viscosity of the top layer is given by
-\[\eta_{eff}=\eta_0 I_2^{\frac{(1-n)}{n}}\]
-with $\eta_0=$ \SI{4.75e11}{\pascal\s} and $n=4$, while the surrounding fluid has $\eta_f=$ \SI{e21}{\pascal\s}. The effective viscosity is capped between
-$\eta_{min}=$ \SI{e21}{\pascal\s} and $\eta_{max}=$ \SI{e25}{\pascal\s}. Velocity boundary conditions are set to free slip at the top and the bottom, and to
-no slip at the sides of the domain.
-
-The sinking of the slab determines a decrease of the effective viscosity in the T-shaped layer during the first part of the evolution (Fig. \ref{fig:slab}a),
-as consequence of local strain rates, while effective viscosities increase up to $\eta_{max}$ in the top layer after approximately \SI{20}{\mega\year}, when
-necking is complete (Fig. \ref{fig:slab}b).
-
-The width of necking is tracked by means of markers position at the side of the slab. Necking width and time are normalised in function of the initial slab
-(\SI{80}{\km}) and the characteristic time ($t_c=$ \SI{7.1158e14}{\s}) \citep{Schmalholz2011,Glerum2018}. The results in case of arithmetic, geometric and
-harmonic mean of the viscosity are shown in Fig. \ref{fig:necking}, compared with results from \citet{Glerum2018} for which an infinity norm mean was used.
-The influence of the average schemes agrees with results shown in Section \ref{sec:subduction}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Slab_detachment}.
-
-\begin{figure}
-\includegraphics[width=\linewidth]{./Figures/Slab.pdf}
-\caption{Effective viscosity for the slab detachment benchmark at the beginning of the evolution (panel a) and when necking is complete (panel b).}
-\label{fig:slab}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Necking.pdf}
-\caption{Normalised width of the necking of the slab detachment benchmark as function of normalised time.}
-\label{fig:necking}
-\end{figure}
-
-\subsection{The indenter experiment}\label{sec:indenter}
-The indenter benchmark simulate a rigid punch on a purely plastic von Mises material, for which exists an analytical solution
-\citep{Thieulot2008,Thieulot2014,Glerum2018}. The domain is a unit square with a grid resolution of \num{256x256} elements. The gravity acceleration is set
-to $\bm{g}=\bm{0}$ and effective viscosity is capped using $\eta_{min}=10^{-4}$ and $\eta_{max}=10^3$. The tolerance for the convergence of the non-linear
-solution is $tol=10^{-9}$, with a maximum number of non-linear iterations set to 500. The medium in the domain has $\rho=0.01$, an initial viscosity
-$\eta_0=10$, cohesion $C=1$ and an angle of internal friction $\phi=0$°. Velocity boundary conditions are set to no slip at the bottom and free slip at the
-sides of the domain. The top of the domain is open with the exception of the central portion (punch area) with width $w_p=0.125$, where $v=-1.05$ and $u$ is
-fixed either to 0 (rough punch experiment) or free (smooth punch experiment).
-
-Fig. \ref{fig:indenter} show the results in terms of viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure
-(panels g and h) for both experiments. The analytical solution indicate that pressure at ($0.5;1$) and ($0.5 \pm w_p;1$)is $p=\pi +1$ and $p=1$, respectively
-(black and red lines in Fig. \ref{fig:smooth_p}). Fig. \ref{fig:smooth_p} shows a clear improvement in the pressure solution when passing from a rough to a
-smooth punch (green and blue lines in Fig. \ref{fig:smooth_p}, respectively), as previously observed by \citet{Thieulot2014} and \citet{Glerum2018}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Indenter}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Indenter.pdf}
-\caption{Viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure (panels g and h) fields for rough (left column) and
-smooth (right column) punch experiments.}
-\label{fig:indenter}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Smooth_Rough.pdf}
-\caption{Nodal pressure in function of the x coordinate at $y=1$ for rough (green line) and smooth (blue line) punch experiment. Black and red lines indicate
-the analytical solution at $x=0.5$ and $x=0.5 \pm w_p$, respectively.}
-\label{fig:smooth_p}
-\end{figure}
-
-\subsection{The brick experiment}\label{sec:brick}
-An instantaneous version of the brick benchmark is performed to verify the correctness of the pressure-dependent plasticity for different angle of internal
-friction, as proposed by \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{40}{\km} and $L_y=$ \SI{10}{\km} and a grid resolution of \num{512x128}
-elements. The gravity acceleration is set to $g_y=$ \SI{-10}{\m\square\s} and effective viscosity is capped using $\eta_{min}=$ \SI{e19}{\pascal\s} and
-$\eta_{max}=$ \SI{e26}{\pascal\s}. The tolerance for the convergence of the non-linear solution is $tol=10^{-7}$, with a maximum number of non-linear iterations
-set to 1000. A 800 m-wide and 400 m-height inclusion with $\eta_b=$ \SI{e20}{\pascal\s} and $\rho_b=$ \SI{2700}{\kg\per\cubic\m} is placed at the bottom of the
-domain at $x=L_x/2$. The inclusion is surrounded by a non-linear viscous medium with $\rho_m=$ \SI{2700}{\kg\per\cubic\m}, an initial viscosity of
-\SI{e23}{\pascal\s}, a linear viscous viscosity of \SI{e25}{\pascal\s} and a cohesion of \SI{40}{\mega\pascal}. Velocities are set to free slip conditions at
-the bottom of the domain and the top is open. Velocities on sides of the domain are fixed to $u=$ \SI{+-2e-11}{\m\per\s} and $v=0$. The experiment is performed
-in compressional and extensional contexts with an angle of internal friction $\phi$ of the non-linear medium variable between 0° and 30°.
-
-As expected, two shear bands stem from the inclusion with variable angles in relation to both the dynamics context and the internal friction angle. Shear band
-angles formed at 45° for $\phi=0$° in both extensional and compressional contexts (Fig. \ref{fig:brick_beam}a), while they are different in case of $\phi\neq0$°.
-Shear band angles in case of $\phi=20$° are shown for compression and extension in Fig. \ref{fig:brick_beam}b  and c, respectively. Values of shear band angles
-as function of internal friction angles from 0° to 30° are extracted at different depths and minimum and maximum values are plotted in Fig. \ref{fig:brick},
-in comparison with theoretical Roscoe, Arthur and Coulomb shear band angles. For all tests the tolerance for nonlinear convergence is defined $tol=10^{-7}$
-and none converges before the maximum number of iterations ($it_{max}=1000$) is reached. However, tests with internal friction angles up to 15° show a constant
-decrease of the velocity residuals, which can not be observed for for tests with higher internal friction angles (Fig. \ref{fig:convergence}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Brick_experiment}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Brick_Beam.pdf}
-\caption{Shear band angles predicted for the brick experiment for $\phi=0$° (panel a) and for $\phi=20$° in case of compression and extension (panels b and c,
-respectively).}
-\label{fig:brick_beam}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Brick.pdf}
-\caption{Shear band angles predicted for the brick experiment in case of compressional and extensional contexts as function of different internal angle of
-friction (continuous black line and red dots), compared with theoretical Roscoe, Artur and Coulomb angles (discontinuous black lines). Red lines indicate the
-range of angles calculated at different depths.}
-\label{fig:brick}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=500px]{./Figures/convergence.pdf}
-  \caption{Convergence rates for all tests. Normalised residuals (left column) are compared with velocity residuals (right column) for extensional (first row)
-  and compressional (second row) settings.}
-  \label{fig:convergence}
-  \end{figure}
-
-\subsection{Advection stabilisation}\label{sec:advection}
-The 1D advection problem proposed by \citet{Donea2003} and \citet{Thieulot2011} is performed to verify the effectiveness of the SUPG method to stabilise the
-advection term of the energy equation. The domain is a 1D segment with $L_x=1$ composed by 50 elements and a discontinuity in the thermal field placed at
-$x=0.25$. Temperature is set to 1 for $x \leq 0.25$ and to 0 for $x > 0.25$. Velocity is set to ${u}=1$ in the entire domain. The simulation is performed for
-250 time steps, with $dt=0.002$, so the thermal discontinuity should be at $x=0.75$ at the end of the simulation.
-
-Temperature profiles at the end of the simulation are shown in \ref{fig:advection} as function of the dimensionless coefficient $\gamma=\tau \bm{v}/h$. In case
-of the classic Galerkin method ($\gamma=0$, blue line in \ref{fig:advection}) the final thermal profile is characterised by strong oscillations, which are
-eliminated in case of the SUPG method ($\gamma=0.045$, orange line in \ref{fig:advection}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Advection_stabilisation}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Advection.pdf}
-\caption{Temperature profile as function of \textit{x} for the advection stabilisation benchmark. Purple line indicates the initial temperature profile; the
-green line indicates the analytical temperature profile after 250 time steps; blue line indicates the temperature profile after 250 time steps in case of the
-classic Galerkin method ($\gamma=0$); orange line indicates the temperature profile after 250 time steps in case of the SUPG method ($\gamma=0.045$).}
-\label{fig:advection}
-\end{figure}
-
-\subsection{Shear heating}\label{sec:simple_shear}
-This exercise is performed in an unit square domain composed by \num{128x128} elements. The velocity field is prescribed on the entire domain with
-$\bm{v}=(L_y-y)y\bm{e_x}$; viscosity, density and specific heat are fixed to 1, while thermal conductivity and radiogenic energy are fixed to 0. Therefore,
-the energy equation (Eq. \ref{eq:energy}) can be simplified as \[\frac{\partial T}{\partial t}=H_s\] and fixing $T(t=0)=0$, the temperature can be find as
-\[T(t)=H_s t\]
-In this case we have
-\begin{eqnarray}
-\dot{\epsilon_{xx}}&=&\frac{\partial u}{\partial x}=0 \nonumber \\
-\dot{\epsilon_{yy}}&=&\frac{\partial v}{\partial y}=0 \nonumber \\
-\dot{\epsilon_{xy}}&=&\frac{1}{2}\left(\frac{\partial v}{\partial x}+\frac{\partial u}{\partial y}\right)=\frac{1}{2}(L_y-2y)\nonumber
-\end{eqnarray}
-and, simplifying Eq. \ref{eq:shear}, shear heating can be calculated as \[H_s(x,y)=(1-2y)^2\]
-
-The solution predicted by the model in terms of velocity, temperature and shear heating match well with the analytical solutions (Fig. \ref{fig:analytical_en}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Simple_shear}.
-
-%\begin{figure}
-%\centering
-%\includegraphics[width=400px]{./Figures/Energy_Thieulot.pdf}
-%\caption{Temperature and shear heating predicted by the model after 1 time step.}
-%\label{fig:en_thieulot}
-%\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=500px]{./Figures/Analytical_Shear.pdf}
-\caption{Velocity (panel a), temperature (panel b) and shear heating (panel c) as function of the \textit{y} coordinate for the simple shear experiment. The
-solutions predicted by the model (red lines) are compared with the analytical solutions (dashed blue lines).}
-\label{fig:analytical_en}
-\end{figure}
-
 \subsection{Shear and adiabatic heating}\label{sec:shear}
 This problem is performed as presented in Exercise 9.4 in \citet{Gerya2010}. Two materials are vertically separated in a rectangular domain with $L_x=$
-\SI{1000}{\km}, $L_y=$ \SI{1500}{\km} and a grid resolution of \num{30x20} elements. Constant thermal coefficient expansion ($\alpha=$ \SI{3e-5}{\per\kelvin}),
+\SI{1000}{\km}, $L_y=$ \SI{1500}{\km} and a grid resolution of $30\times20$ elements. Constant thermal coefficient expansion ($\alpha=$ \SI{3e-5}{\per\kelvin}),
 temperature ($T=$ \SI{1300}{\kelvin}) and gravity acceleration ($g_y=$ \SI{-10}{\m\square\s}) are assumed in the whole domain. Fluid 1 (on the left side of the
@@
 \end{figure}
-
-\subsection{Mantle convection}\label{sec:mantle}
-This problem is performed as presented by \citet{Blankenbach1989} (constant viscosity cases) and \citet{Thieulot2014}, in a 2D unit square domain with gravity
-acceleration $g_y=10^{10}Ra$. The experiment is performed with three different Rayleigh numbers ($Ra=10^4$, $10^5$ and $10^6$) and with different grid
-resolution (between \num{32x32} and \num{128x128} elements). The fluid has constant viscosity, initial density, heat capacity, thermal conductivity
-($\eta= \rho_0= C_p = k=1$), reference temperature ($T_0=0$) and thermal expansion coefficient ($\alpha=10^{-10}$). Temperatures are set to 0 on top and 1 on
-bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\]
-
-The solution generated by the code in terms of $v_{\textrm{rms}}$ (calculated as in Sec. \ref{sec:error}) and Nu as function of time are reported for all the
-simulations in Fig. \ref{fig:mantle}. At the steady state, $v_{\textrm{rms}}$ and Nu of all simulations converge well toward the values from \citet{Blankenbach1989},
-with lower errors for higher resolution grids (Table \ref{tab:mantle}). All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Mantle_convection}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Convection.pdf}
-\caption{$v_{\textrm{rms}}$ (panels a, c and e) and Nu (panels b, d and f) for the mantle convection benchmark as function of time for different grid resolution.
-Panels a and b show the results for $Ra=10^4$; Panels c and d show the results for $Ra=10^5$; panels e and f show the results for $Ra=10^6$. Purple lines
-indicate the convergence values for the $v_{\textrm{rms}}$ and the Nu from \citet{Blankenbach1989} at the steady state.}
-\label{fig:mantle}
-\end{figure}
-
-\begin{table}[H]
-\caption{Comparison between $v_{\textrm{rms}}$ and Nu predicted by the code for the mantle convection experiment and same values as reported in literature.}
-\centering
-\small
-\begin{tabular}{|>{\centering}m{0.09\linewidth}|>{\centering}m{0.05\linewidth}|>{\centering}m{0.24\linewidth}|>{\centering}m{0.15\linewidth}|
-  >{\centering}m{0.1\linewidth} >{\centering}m{0.1\linewidth} >{\centering\arraybackslash}m{0.1\linewidth}|}
-\toprule
-& & \multirow{2}{*}{\citet{Blankenbach1989}} & \citet{Thieulot2014} & \multicolumn{3}{c|}{FALCON}  \\
- & & & \num{200x200} & \num{32x32} & \num{64x64} & \num{128x128}  \\
-\midrule
-\multirow{2}{*}{$Ra=10^4$} & $v_{\textrm{rms}}$ & $42.864947 \pm 0.000020$ & 42.867 & 42.83226 & 42.852793 & 42.861394  \\
-           & Nu      & $4.884409 \pm 0.000010$  & 4.882  & 4.781297 & 4.857475  & 4.877573   \\
-\hline
-\multirow{2}{*}{$Ra=10^5$} & $v_{\textrm{rms}}$ & $193.21454 \pm 0.00010$  & 193.255& 193.872643 & 193.377472 & 193.252290  \\
-           & Nu      & $10.534095 \pm 0.000010$ & 10.507 & 9.602514 & 10.270735  & 10.465629   \\
-\hline
-\multirow{2}{*}{$Ra=10^6$} & $v_{\textrm{rms}}$ & $833.98977 \pm 0.00020$  & 834.712& 848.091176 & 837.767911 & 834.945793  \\
-           & Nu      & $21.972465 \pm 0.000020$ & 21.695 & 15.999266 & 19.703682  & 21.306939   \\
-\bottomrule
-\end{tabular}
-\label{tab:mantle}
-\end{table}
-
-\subsection{Viscoplastic mantle convection}\label{sec:visco_mantle}
-This benchmark is performed as Cases 1-5a described in \citet{Tosi2015} in a 2D unit square domain with two different grid resolutions of \num{32x32} 
-and \num{100x100} elements. Reference viscosity, reference density, heat capacity, thermal conductivity and thermal expansion coefficient are fixed to a constant
-value of 1 ($\eta_0= \rho_0= C_p = k=\alpha=1$), while the gravity has been chosen as $g_y=10^2$, in order to obtain a Rayleigh number $Ra=10^2$. Temperatures
-are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\].
-The viscosity field $\eta$ is calculated as the harmonic average between a linear part $\eta_{lin}$ that depends on temperature only or on temperature and 
-depth $y$ and a nonlinear-plastic part $\eta_{pl}$ that depends on the strain rate $\dot{\bm{\epsilon}}$, as follows
-\[\eta(T,y,\dot{\bm{\epsilon}})=2\left(\frac{1}{\eta_{lin}(T,y)}+\frac{1}{\eta_{pl}(\dot{\bm{\epsilon}})}\right)^{-1}\]
-The linear and the nonlinear-plastic parts of the viscosity are calculated following \citet{Tosi2015} as follows
-\[\eta_{lin}(T,y)=\exp(-\ln(\Delta\eta_T)T+\ln(\Delta\eta_y)y)\]
-\[\eta_{pl}(\dot{\bm{\epsilon}})=\eta^*+\frac{\sigma_Y}{\sqrt{\dot{\bm{\epsilon}}:\dot{\bm{\epsilon}}}}\]
-where $\Delta\eta_T$, $\Delta\eta_y$, $\eta^*$ and $\sigma_Y$ are parameters chosen as listed in Table \ref{tab:tosi_setup}.
-
-For each case, some diagnostic quantities measured at the steady state are reported in Table \ref{tab:visco_mantle} (see Section \ref{sec:error} for 
-the computation of all the quantities), showing a well fit with results described in \citet{Tosi2015}. Data and charts for all cases can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Viscoplastic_mantle_convection}.
-
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{white!100!}{gray!25!}
-%\begin{table}[H]
-%  \caption{Results of the viscoplastic mantle convection benchmark for Cases 1-4 compared with results obtained with ELEFANT and YACC (from \citealp{Tosi2015}).}
-%  \centering
-%  \resizebox{\linewidth}{!}{
-%\begin{tabular}{|c|ccc|ccc|ccc|ccc|}
-%  \toprule
-%    & \multicolumn{3}{c|}{\textbf{Case 1}} & \multicolumn{3}{c|}{\textbf{Case 2}} & \multicolumn{3}{c|}{\textbf{Case 3}} &
-%    \multicolumn{3}{c|}{\textbf{Case 4}} \\
-%   & FALCON & ELEFANT& YACC & FALCON& ELEFANT& YACC & FALCON & ELEFANT & YACC & FALCON & ELEFANT & YACC \\
-%  \midrule
-%   $\left\langle T \right\rangle$ & 0.7759 & 0.7758 & 0.7767 & 0.6043 & 0.6036 & 0.6058 & 0.7275 & 0.7275 & 0.7286 & & 0.5277 & 0.5289\\
-%    Nu$^{top}$ & 3.4280 & 3.4214 & 3.4298 & 8.5242 & 8.5115 & 8.5278 & 3.0368 & 3.0347 & 3.0374 && 6.5912 & 6.5572 \\
-%    Nu$^{bottom}$ & 3.3168 & 3.313 & 3.3143 & 8.4599 & 8.4465 & 8.3990 & 3.0086 & 2.9908 & 2.9628 && 6.5834 & 6.5243 \\
-%    $v_{\textrm{rms}}$ & 249.901 & 249.134 & 251.7997 & 141.2451 & 140.7087 & 142.2020 & 100.2776 & 100.1208 & 100.9467 && 79.1105 & 79.6202 \\
-%    $v_{\textrm{rms}}^{top}$ & 1.8907 & 1.8642 & 1.8298 & 104.7928 & 104.1228 & 105.1767 & 2.0832 & 2.0652 & 2.0374 && 74.7596 & 75.8414 \\
-%    $u_{max}^{top}$ & 2.6340 & 2.6119 & 2.5516 & 121.9212& 121.766 & 122.3238 & 2.9086 & 2.9019 & 2.8458 && 88.9146 & 89.2940 \\
-%    $\eta_{min}$ & - & - & - & $2.0303\times10^{-5}$ & $2.0600\times10^{-5}$ & $1.9853\times10^{-5}$ & $4.8129\times10^{-5}$ & $4.8080\times10^{-5}$ & $4.7907\times10^{-5}$ &
-%    & $1.9860\times10^{-4}$ & $1.9174\times10^{-4}$ \\
-%    $\eta_{max}$ & - & - & - & 1.3325 & 1.85 & 1.6095 & 0.9360 & 0.9023 & 1 && 1.52 & 1.6773 \\
-%    $\left\langle W \right\rangle$ & 2.4299 & 2.4316 & 2.4583 & 7.5648 & 7.5665 & 7.6222 & 2.0387 & 2.0384 & 2.0400 && 5.6216 & 5.6512 \\
-%    $\left\langle \Phi \right\rangle/Ra$ & 2.4311 & 2.4276 & 2.4333 & 7.5661 & 7.5585 & 7.6194 & 2.0386 & 2.0370 & 2.0335 && 5.6182 & 5.6463 \\
-%    $\delta$ & 0.05\% & 0.16\% & 1.02\% & 0.02\% & 0.11\% & 0.04\% & $<$0.01\% & 0.07\% & 0.32\% && 0.06\% & 0.09\% \\
-%  \bottomrule
-%  \end{tabular}}
-%  \label{tab:visco_mantle}
-%  \end{table}
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{2}{white!100!}{gray!25!}
-\begin{table}[H]
-  \caption{Benchmark cases and corresponding parameters.}
-  \centering
-  \small
-\begin{tabular}{|>{\centering}m{0.08\linewidth}|>{\centering}m{0.05\linewidth}>{\centering}m{0.05\linewidth}>{\centering}m{0.05\linewidth}
-  >{\centering}m{0.05\linewidth} >{\centering}m{0.05\linewidth} |>{\centering\arraybackslash}m{0.2\linewidth}|}
-  \toprule
-    Case & $Ra$ & $\Delta\eta_T$ & $\Delta\eta_y$ & $\eta^*$ & $\sigma_Y$ & Convective regime \\
-  \midrule
-  \rowcolor{gray!25!}
-    1 & $10^2$ & $10^5$ & 1 & - & - & Stagnant lid \\
-    2 & $10^2$ & $10^5$ & 1 & $10^{-3}$ & 1 & Mobile lid  \\
-    \rowcolor{gray!25!}
-    3 & $10^2$ & $10^5$ & 10 & - & - & Stagnant lid  \\
-    4 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 1 & Mobile lid \\
-    \rowcolor{gray!25!}
-    5 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 4 & Periodic \\
-  \bottomrule
-  \end{tabular}
-  \label{tab:tosi_setup}
-  \end{table}
-
-\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{gray!25!}{white!100!}
-\begin{table}
-  \caption{Results of the viscoplastic mantle convection benchmark for different grid resolutions in Cases 1-5 compared with results obtained with ELEFANT
-  and YACC ($100\times100$ elements, from \citealp{Tosi2015}) and in Stone 28 ($32\times32$ elements, from
-  \url{https://github.com/cedrict/fieldstone/tree/master/python_codes/fieldstone_28}).}
-  \centering
-  \resizebox{\linewidth}{!}{
-\begin{tabular}{|cc|ccccccccccc|}
-  \toprule
-  \multicolumn{13}{|c|}{\textbf{Case 1}} \\
-  \midrule
-  Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-  & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-  \hline
-  \rowcolor{gray!25!}
-    YACC & $100\times100$ & 0.7767 & 3.4298 & 3.3143 & 251.7997 & 1.8298 & 2.5516 & - & - & 2.4583 & 2.4333 & 1.02\% \\
-    ELEFANT & $100\times100$ & 0.7758 & 3.4214 & 3.313 & 249.134 & 1.8642 & 2.6119 & - & - & 2.4316 & 2.4276 & 0.16\% \\
-    \rowcolor{gray!25!}
-    Stone 28 & $32\times32$ & 0.7737 & 3.3987 & - & 243.872 & - & - & - & - & - & - & - \\
-%    FALCON & $32\times32$ & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & - & - & 2.4745 & 2.4864 & 0.48\% \\
-%    \rowcolor{gray!25!}
-    FALCON & $100\times100$ & 0.7759 & 3.4280 & 3.3168 & 249.901 & 1.8907 & 2.6340 & - & - & 2.4299 & 2.4311 & 0.05\% \\
-    \midrule
-    \multicolumn{13}{|c|}{\textbf{Case 2}} \\
-    \midrule
-    Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-    & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-    \hline
-    \rowcolor{gray!25!}
-     YACC & $100\times100$ & 0.6058 & 8.5278 & 8.3990 & 142.2020 & 105.1767 & 122.3238 & $1.9853\times10^{-5}$ & 1.6095 & 7.6222 & 7.6194 & 0.04\% \\
-     ELEFANT & $100\times100$ & 0.6034 & 8.5115 & 8.4465 & 140.8067 & 104.1228 & 121.766 & $2.0600\times10^{-5}$ & 1.85 & 7.5665 & 7.5585 & 0.11\% \\
-     \rowcolor{gray!25!}
-     Stone 28 & $32\times32$ & 0.6052 & 8.1715 & - & 141.518 & - & - & - & - & - & - & - \\
-%     FALCON & $32\times32$ & 0.6454 & 8.1422 & 7.4020 & 156.6908 & 95.7226 & 116.9302 & $2.2572\times10^{-5}$ & 0.7194 & 7.4289 & 7.4452 & 0.22\% \\
-%     \rowcolor{gray!25!}
-     FALCON & $100\times100$ & 0.6044 & 8.5266 & 8.4590 & 141.3684 & 104.8867 & 122.0470 & $2.0333\times10^{-5}$ & 1.2914 & 7.5707 & 7.5718 & 0.01\% \\
-     \midrule
-     \multicolumn{13}{|c|}{\textbf{Case 3}} \\
-     \midrule
-     Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-     & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-     \hline
-     \rowcolor{gray!25!} 
-     YACC & $100\times100$ & 0.7286 & 3.0374 & 2.9628 & 100.9467 & 2.0374 & 2.8458 & $4.7907\times10^{-5}$ & 1 & 2.0400 & 2.0335 & 0.32\% \\
-      ELEFANT & $100\times100$ & 0.7275 & 3.0347 & 2.9908 & 100.1208 & 2.0652 & 2.9019 & $4.8080\times10^{-5}$ & 0.9023 & 2.0384 & 2.037 & 0.07\% \\
-      \rowcolor{gray!25!}
-      Stone 28 & $32\times32$ & 0.7277 & 3.0347 & - & 100.018 & - & - & - & - & - & - & - \\
-%      FALCON & $32\times32$ & 0.7281 & 3.0618 & 2.8132 & 101.0092 & 2.1630 & 3.0178 & $4.8500\times10^{-5}$ & 0.7845 & 2.0699 & 2.0695 & 0.02\% \\
-%      \rowcolor{gray!25!}
-      FALCON & $100\times100$ & 0.7275 & 3.0368 & 3.0086 & 100.2776 & 2.0832 & 2.9086 & $4.8129\times10^{-5}$ & 0.9360 & 2.0387 & 2.0386 & $<$0.01\% \\
-      \midrule
-      \multicolumn{13}{|c|}{\textbf{Case 4}} \\
-      \midrule
-      Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-      & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-      \hline
-      \rowcolor{gray!25!} 
-      YACC & $100\times100$ & 0.5289 & 6.5572 & 6.5243 & 79.6202 & 75.4814 & 89.2940 & $1.9174\times10^{-4}$ & 1.6773 & 5.6512 & 5.6463 & 0.09\% \\
-       ELEFANT & $100\times100$ & 0.5277 & 6.5912 & 6.5834 & 79.1105 & 74.7596 & 88.9146 & $1.9860\times10^{-4}$ & 1.5200 & 5.6216 & 5.6182 & 0.06\% \\
-       \rowcolor{gray!25!}
-       Stone 28 & $32\times32$ & 0.5286 & 6.4036 & - & 79.4746 & - & - & - & - & - & - & - \\
-%       FALCON & $32\times32$ & 0.5334 & 6.4679 & 6.3435 & 80.6153 & 76.0511 & 90.2058 & $2.0393\times10^{-4}$ & 0.6453 & 5.6676 & 5.6698 & 0.04\% \\
-%       \rowcolor{gray!25!}
-       FALCON & $100\times100$ & 0.5279 & 6.5897 & 6.5830 & 79.1220 & 75.0715 & 88.8484 & $1.9693\times10^{-4}$ & 1.4133 & 5.6127 & 5.6113 & 0.03\% \\
-       \midrule
-       \multicolumn{13}{|c|}{\textbf{Case 5}} \\
-       \midrule
-       Code & Resolution & $P$ & $\left\langle T \right\rangle_{min}$ & $\left\langle T \right\rangle_{max}$ & Nu$^{top}_{min}$ & Nu$^{top}_{max}$ &
-       $v_{\textrm{rms}}^{min}$ & $v_{\textrm{rms}}^{max}$ & \multicolumn{2}{c}{$(\left\langle \Phi \right\rangle/Ra)_{min}$} & 
-       \multicolumn{2}{c|}{$(\left\langle \Phi \right\rangle/Ra)_{max}$} \\
-       \hline
-       \rowcolor{gray!25!}
-         YACC & $100\times100$ & 0.0792 & 0.6543 & 0.6722 & 2.6950 & 7.2792 & 41.9813 & 98.8230 & \multicolumn{2}{c}{1.4347} & 
-         \multicolumn{2}{c|}{9.2748} \\
-         ELEFANT & $100\times100$ & 0.076 & 0.6514 & 0.6693 & 2.6808 & 7.3506 & 41.2578 & 100.089 & \multicolumn{2}{c}{1.3324} & 
-         \multicolumn{2}{c|}{9.5289} \\
-         \rowcolor{gray!25!}
-         Stone 28 & $32\times32$ & 0.0764 & 0.6521 & 0.6697 & 2.6912 & 7.0763 & 41.61 & 98.183 & \multicolumn{2}{c}{-} & 
-         \multicolumn{2}{c|}{-} \\
-     %    FALCON & $32\times32$ & & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & \multicolumn{2}{c}{-} & 
-     %     \multicolumn{2}{c|}{-} \\
-     %    \rowcolor{gray!25!}
-         FALCON & $100\times100$ & 0.0775 & 0.6519 & 0.6697 & 2.6847 & 7.3629 & 41.5256 & 99.5466 & \multicolumn{2}{c}{1.3399} & 
-         \multicolumn{2}{c|}{9.3993} \\
-  \bottomrule
-  \end{tabular}}
-  \label{tab:visco_mantle}
-  \end{table}
-
-\subsection{Thin layer entrainment}\label{sec:thin}
-This experiment is performed as originally proposed by \citet{vanKeken1997} with two incompressible fluids in a rectangular domain with $L_x=2$ and $L_y=1$
-and gravity acceleration $g_y=10^{10}Ra$. The Rayleigh number and the compositional Rayleigh number are fixed to $Ra=300000$ and $Rc=450000$, respectively.
-Both fluids have constant viscosity, thermal conductivity, specific heat ($\eta=\rho=C_p=1$) and thermal expansion coefficient ($\alpha=10^{-10}$). Fluid 1
-has a density $\rho_1=1$, while fluid 2 is denser ($\rho_2=\rho_1+1.5 \cdot 10^{-10}$) and is placed at the bottom of the domain, for $y \leq 0.025$.
-Temperature are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides of the domain. The initial
-temperature field is given by
-\[T(x,y)=T_u(x,y)+T_l(x,y)+T_r(x,y)+T_s(x,y)-\frac{3}{2}\]
-with
-\begin{eqnarray}
-T_u(x,y)&=&\frac{1}{2}erf\left(\frac{1-y}{2}\sqrt{\frac{u_0}{x}}\right)\nonumber \\
-T_l(x,y)&=&1-\frac{1}{2}erf\left(\frac{y}{2}\sqrt{\frac{u_0}{L_x-x}}\right)\nonumber \\
-T_r(x,y)&=&\frac{1}{2}+\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{y+1}} \exp\left(-\frac{x^2u_0}{4y+4}\right)\nonumber \\
-T_s(x,y)&=&\frac{1}{2}-\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{2-y}} \exp\left(-\frac{(L_x-x)^2u_0}{8-4y}\right)\nonumber
-\end{eqnarray}
-and
-\begin{eqnarray}
-u_0&=&\frac{L_x^{7/3}}{(1+L_x^4)^{2/3}}\left(\frac{Ra}{2\sqrt{\pi}}\right)^{\frac{2}{3}}\nonumber \\
-Q&=&2\sqrt{\frac{L_x}{\pi u_0}}\nonumber
-\end{eqnarray}
-
-The experiment is performed with two grid resolution (\num{125x40} and \num{200x80} elements), with 100 markers per element and CFL of 0.25. $v_{\textrm{rms}}$
-(see Sec. \ref{sec:error}) calculated for both the simulations match well with results from \citet{vanKeken1997} and obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and ELEFANT \citep{Thieulot2014}, obtained with similar grid resolutions
-(Fig. \ref{fig:thin}). In addition, a variety of simulations with different resolutions and aspect ratios are performed to verify that the sensitivity of the
-initial velocity field. Results match well with results from \citet{vanKeken1997} (in the grey area) and \citet{Thieulot2014} (Fig. \ref{fig:thin_initial}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Thin_layer}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/vrms_thin.pdf}
-\caption{$v_{\textrm{rms}}$ for the thin layer experiment as function of time for different grid resolution. Results are compared with results obtained by
-\citet{vanKeken1997} (black lines) and with ASPECT and ELEFANT.}
-\label{fig:thin}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/vrmszero.pdf}
-\caption{$v_{\textrm{rms}}$ at $t=0$ for the thin layer experiment as function of the element size for different aspect ratios of the elements. Results are compared
-with results obtained by \citet{vanKeken1997} (black dots in the grey area) and with ASPECT and ELEFANT.}
-\label{fig:thin_initial}
-\end{figure}
-
-\subsection{Sinking hydrated cylinder}\label{sec:quinquis}
-This experiment is performed as explained in \citet{Quinquis2014} and simulates the sinking of a cold, hydrated cylinder into a warm, dry mantle. The domain
-is square with $L_x=L_y=$ \SI{300}{\km}, a grid resolution of \num{300x300} elements and 25 markers per element. The sinking cylinder is located at $x=$
-\SI{150}{\km} and $y=$ \SI{170}{\km} with a radius of \SI{20}{\km} and has $\rho_c=$ \SI{3250}{\kg\per\cubic\m}, $\eta_c=$ \SI{e23}{\pascal\s}, a thermal
-conductivity $k_c=$ \SI{4.5}{\watt\per\m\per\kelvin},a specific heat $C_{p_c}=$ \SI{1250}{\joule\per\kg\per\kelvin}; the surrounding mantle has $\rho_m=$
-\SI{3200}{\kg\per\cubic\m}, $\eta_m=$ \SI{e20}{\pascal\s}, a thermal conductivity $k_m=$ \SI{105}{\watt\per\m\per\kelvin} and specific heat $C_{p_m}=$
-\SI{1250}{\joule\per\kg\per\kelvin}; a 58 km-thick lithosphere overlies the mantle and has $\rho_l=$ \SI{3200}{\kg\per\cubic\m}, $\eta_l=$ \SI{e23}{\pascal\s},
-a thermal conductivity $k_l=$ \SI{4.5}{\watt\per\m\per\kelvin}, specific heat $C_{p_l}=$ \SI{750}{\joule\per\kg\per\kelvin} and a thermal diffusivity of
-\SI{e-6}{\square\m\per\s}. Velocity boundary conditions are set to free slip on all sides of the domain. Temperature increase linearly in the lithosphere,
-from 0° to 1300°C, and in the mantle, from 1300° to 1360.5°C. The initial temperature of the cylinder is fixed to 400°C. Throughout the evolution, temperature
-is fixed to 0° and 1360.5°C at the top and bottom of the domain, respectively. The initial bound water content is imposed at 2 wt.\% in the cylinder and at
-0 wt.\% in both the mantle and the lithosphere. The maximum amount of water in the mantle is fixed to 0.2 wt.\%, while maximum water content in the cylinder
-and in the lithosphere is function of pressure and temperature and is calculated using a serpentinized harzburgite with Perple\_X, as in \citet{Quinquis2014}.
-
-As shown in \citet{Quinquis2014}, the model is characterised by a progressive dehydration in the external portion of the cylinder, due to the increase of
-temperature, with a consequent vertical migration of free water that hydrates the mantle above the cylinder, up to the lithosphere. The distribution of
-bound and free water during the evolution of the experiment is shown in Fig. \ref{fig:quinquis} (left and right columns, respectively). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Hydration}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Hydration.pdf}
-\caption{Distribution of bound and free water (left and right columns, respectively) at $t=0.5$ and $t=$ \SI{1.25}{\mega\year} (first and second rows,
-respectively) for the sinking hydrated cylinder experiment. Only markers with bound or free water not equal to 0 are plotted.}
-\label{fig:quinquis}
-\end{figure}
-
-\subsection{Mantle melting curves}\label{sec:katz}
-This instantaneous experiment is performed in order to recreate isobaric melting curves, as function of temperature, bulk water and content of water in the
-melt, to compare with melting curves obtained by \citet{Katz2003}. The domain is square with $L_x=L_y=100$ km, a grid resolution of \num{100x100} elements and
-25 markers per element. Density is set in the entire domain to \SI{3300}{\kg\per\cubic\m} and temperature increases from the sides to the centre from 1000° to
-1500°C. The experiment is performed for bulk water contents from 0 to 0.3 wt.\%.
-
-Fig. \ref{fig:fraction} shows isobaric melting curves for \SI{1}{\giga\pascal} (panel a) and \SI{3}{\giga\pascal} (panel b) calculated for different bulk water
-contents. The curves well match with those obtained for the mantle by \citet{Katz2003}, showing the saturation at the solidus for bulk water of 0.2 and 0.3 wt.\%, indicating
-that the code correctly determines melt fraction as function of pressure, temperature and bulk water content. Same results can be observed in Fig.
-\ref{fig:water}a, where isobaric and isothermal melting curves are shown as function of bulk water content. In addition, Fig. \ref{fig:water}b shows that the
-code correctly determines also the percentage of water in the melt as function of pressure, temperature and melt fraction. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Melting}.
-
-\begin{figure}
-\includegraphics[width=\linewidth]{./Figures/Fraction.pdf}
-\caption{Isobaric melting curves for 1 GPa (panel a) and 3 GPa (panel b) as function of temperature with different bulk water contents.}
-\label{fig:fraction}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\linewidth]{./Figures/Water.pdf}
-\caption{Isobaric melting curves for 1.5 GPa as function of bulk water (panel a) and water in the melt (panel b) calculated for different temperatures.}
-\label{fig:water}
-\end{figure}
-
-\clearpage
-
-\bibliography{export}
-
-\end{document}

### 2022-09-07 tex/Benchmark/simple_shear.tex
update main

--- 2022-03-31 FALCON_description.tex
@@
-\documentclass[hidelinks,10pt,a4paper]{article}
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{textcomp}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage{booktabs}
-\usepackage{multirow}
-\usepackage[cm]{fullpage}
-\usepackage[round]{natbib}
-\bibliographystyle{abbrvnat}
-\usepackage{doi}
-\usepackage{bm}
-\usepackage[table]{xcolor}
-\hypersetup{
-    colorlinks,
-    linkcolor={blue},
-    citecolor={blue},
-    urlcolor={blue}
-}
-\usepackage{float}
-\restylefloat{table}
-\usepackage{siunitx}
-\DeclareSIUnit\year{yr}
-
-\begin{document}
-
-\section{Introduction}\label{sec:intro}
-This document includes a description of the code FALCON, with a presentation of the features implemented followed by results of extensive benchmarks listed in
-Table \ref{tab:list}, performed to verify that the code solves correctly a wide range of problems. Description and results of all benchmarks are shown
-in Section \ref{sec:benchmark}.
-
-\begin{table}[H]
-\caption{List of benchmarks performed for each feature of the code. Data of all benchmarks can be found at 
-\url{https://github.com/aleregorda/Benchmarks}}
-\centering
-\begin{tabular}{| l | l | l |}
-\toprule
- \textbf{Feature}  & \textbf{Benchmark} & \textbf{Section} \\
-\midrule
-  Solver  & Stokes flow & \ref{sec:stokes} \\
-\hline
-  \multirow{2}{*}{Markers advection} & Zalesak disk & \ref{sec:runge} \\
-    & Conservative Interpolation Velocity & \ref{sec:CVI} \\
-\hline
-  \multirow{4}{*}{Momentum equation}  & Poiseuille flow & \ref{sec:poiseuille}\\
-    & Instantaneous 2D sphere & \ref{sec:ist_sphere} \\
-    & Rayleigh-Taylor experiment & \ref{sec:rayleigh} \\
-    & Falling block & \ref{sec:block} \\
-\hline
-    & 2D Stokes sphere & \ref{sec:sphere} \\
-    Sticky air and  & Stabilisation algorithm & \ref{sec:stab} \\
-    free surface & Topography relaxation & \ref{sec:crameri} \\
-  & Spontaneous subduction & \ref{sec:subduction} \\
-\hline
-  Erosion and sedimentation  & - & - \\
-\hline
-   Non-linear  & Slab detachment & \ref{sec:slab} \\
-   visco-plasticity & Indenter & \ref{sec:indenter} \\
-    & Brick & \ref{sec:brick} \\
-\hline
-  \multirow{3}{*}{Energy equation}  & Advection stabilisation & \ref{sec:advection} \\
-    & Simple shear heating & \ref{sec:simple_shear} \\
-    & Shear and adiabatic heating & \ref{sec:shear} \\
-\hline
-\multirow{3}{*}{Energy + momentum} & Mantle convection & \ref{sec:mantle} \\
-& Viscoplastic mantle convection & \ref{sec:visco_mantle} \\
-& Thin layer entrainment & \ref{sec:thin} \\
-\hline
-  Phase changes  & \multirow{2}{*}{Hydrated sinking cylinder} & \multirow{2}{*}{\ref{sec:quinquis}} \\
-  and hydration  &  & \\
-\hline
-  Melting & Experimental melting curves & \ref{sec:katz} \\
-\bottomrule
-\end{tabular}
-\label{tab:list}
-\end{table}
-
-The code is written in Fortran90 and uses the finite element method (FEM) with quadrilateral $Q_1 \times P_0$ elements (continuous bilinear velocity and
-discontinuous constant pressure), associated with MUMPS\footnote{\url{http://mumps.enseeiht.fr/}} \citep{Amestoy2001,Amestoy2019} that is a software package
-for solving systems of linear equations of the form $A \cdot x = b$, where \textit{A} is a square sparse matrix, by means of a direct method. Since
-$Q_1 \times P_0$ elements do not satisfy Ladyzhenskaya, Babuska and Brezzi (LBB) stability condition \citep{Donea2003}, elemental pressure is elaborated
-in post-processing to avoid spurious pressures by means of a double interpolation to smooth the pressure. Following this procedure, the elemental pressure
-is interpolated onto nodes and then back onto elements. After the smoothing procedure, the pressure field is calculated again on the nodes to be used in
-combination with the lithostatic pressure, which is calculated onto nodes. Correctness of the solution and performances of the solver in terms of time and
-memory usage are tested solving the Stokes flow with the analytical solution proposed by \citet{Donea2003} (Section \ref{sec:stokes}).
-
-The thermo-mechanics of crust-mantle systems is described by means of the conservation of mass, momentum and energy equations, expressed as follows:
-\begin{eqnarray}
-\label{eq:mass}\frac{\partial \rho}{\partial t}+\bm{\nabla} \cdot (\rho \bm{u})&=&0\\
-\label{eq:momentum}-\bm{\nabla} p + \bm{\nabla} \cdot \bm{\tau} + \rho \bm{g}&=&0\\
-\label{eq:energy}\rho C_p \left(\frac{\partial T}{\partial t} + \bm{u} \cdot \bm{\nabla} T \right)&=&\bm{\nabla} \cdot \left(k\bm{\nabla} T\right) + H_{tot}
-\end{eqnarray}
-where $\rho$ is the density, $\bm{u}$ is the velocity, \textit{p} is the pressure, $\bm{\tau}$ is the deviatoric stress, $\bm{g}$ is the gravity acceleration,
-$C_p$ is the specific heat at constant pressure, \textit{T} is the temperature, \textit{k} is the thermal conductivity and $H_{tot}$ is the total internal heat
-production. Density variations due to temperature are generally small enough to assume the density as constant ($\rho=\rho_0$) in Eq. \ref{eq:mass} and in
-Eq. \ref{eq:energy}, while it must be treated as a variable in the buoyancy term of Eq. \ref{eq:momentum}, such that \[\rho=\rho_0(1-\alpha(T-T_0))\]
-where $\rho_0$ is the density at a reference temperature $T_0$ and $\alpha$ is the coefficient of thermal expansion. Eq. \ref{eq:mass} then can be rewritten as
-\begin{equation}\label{eq:mass_0}
-\bm{\nabla} \cdot \bm{u}=0
-\end{equation}
-This simplification is known as the extended Boussinesq approximation.
-
-The time step $dt$ is chosen in according to Courant-Friedrichs-Lewy condition \citep{Anderson1995}, such that
-\[dt=C_n\; \min\left(\frac{\Delta x}{\max|\bm{u}|},\frac{(\Delta x)^2}{\kappa}\right)\]
-where $0<C_n<1$ is the Courant (CFL) number, $\Delta x$ is the minimum dimension of smallest element, $\max|\bm{u}|$ is the maximum velocity calculated on the
-entire domain and
-$\kappa=\frac{k}{\rho C_p}$ is the thermal diffusivity \citep{Thieulot2014}.
-
-\section{Lagrangian markers}\label{sec:markers}
-Elemental properties (density, viscosity, thermal conductivity, specific heat and thermal expansion) needed to solve Eqs. \ref{eq:mass_0}, \ref{eq:momentum}
-and \ref{eq:energy} are related to the composition of each element, determined by means of Lagrangian markers that are characteristic of different materials
-in the domain. Elemental properties (with the sole exception of the viscosity) are calculated using an arithmetic mean, as \[P_e=\frac{1}{n}\sum_{i=1}^n P_i\]
-where $P_e$ is the elemental property, $P_i$ is the property characteristic of the material of each marker and $n$ is the number of markers of the element.
-Differently, the average scheme for the elemental viscosity can be chosen between harmonic, geometric and arithmetic mean. Markers can be placed regularly or
-randomly at the beginning of the simulation and their advection is performed by either a 2nd-order or a 4th-order Runge-Kutta in space, interpolating the velocity 
-field on each marker by means of the shape functions (see Section \ref{sec:runge}). The interpolated velocity is then corrected by means of the Conservative
-Velocity Interpolation (CVI), which introduces a corrective term to reduce dispersion and clustering of particles in both steady state and time-dependent 
-\citep{Wang2015} (see Section \ref{sec:CVI}).
-
-The initial distribution of the markers is created using the open-source code library
-Geodynamic World Builder \citep{Fraters2019}. During the simulation, each marker carries memory of temperature, pressure and accumulated strain, which are
-determined interpolating the relative nodal parameter. In particular, pressure is interpolated onto each marker after the smoothing procedure. The number
-of markers contained in each element is maintained between $n_{min}$ and $n_{max}$. When in an element there are less markers than $n_{min}$ the code adds
-random markers to reach the $n_{min}$, while if the number is higher than $n_{max}$ some of them are randomly deleted. When new markers are added, they assume
-the properties of the nearest marker. In this way elements are never empty and maintain a number of markers inside a prefixed range.
-
-\section{The momentum equation}\label{sec:mom_eq}
-The penalty method is implemented so that Eq. \ref{eq:mass_0} can be written as:
-\begin{equation}\label{eq:mass_penalty}
-\bm{\nabla} \cdot \bm{u} + \frac{p}{\lambda}=0
-\end{equation}
-where $\lambda$ is the so-called penalty parameter, which should be 6-7 orders of magnitude larger than the shear viscosity to ensure that mass conservation
-is satisfied \citep{Donea2003,Thieulot2014}.
-
-The deviatoric stress tensor in Eq. \ref{eq:momentum} can be written in terms of the strain rate tensor as $\bm{\tau}=2\eta\dot{\bm{\epsilon}}$, with
-$\dot{\bm{\epsilon}}=\frac{1}{2}\left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)$. Therefore, Eq. \ref{eq:momentum} can be rewritten as:
-\begin{equation}\label{eq:momentum_pressure}
--\bm{\nabla} p + \bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right) + \rho \bm{g} = 0
-\end{equation}
-Finally, using pressure from Eq. \ref{eq:mass_penalty}, Eq. \ref{eq:momentum_pressure} can be rewritten as:
-\begin{equation}\label{eq:momentum_penalty}
-\lambda \bm{\nabla} \left(\bm{\nabla} \cdot \bm{v}\right)+\bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right)+\rho\bm{g}=0
-\end{equation}
-The penalty method is been associated to the iterative Uzawa method, as described in detail in \citet{Dabrowski2008} and \citet{Thieulot2014}
-(see Section \ref{sec:poiseuille}). Results of benchmarks performed to verify the correctness in the implementation of Eq. \ref{eq:momentum_penalty} are
-shown in Sections \ref{sec:ist_sphere} and \ref{sec:rayleigh}.
-
-To support large viscosities variations the penalty parameter is related to the effective elemental viscosity by means of a dimensionless coefficient, so that
-$\lambda=\lambda_e(e)\eta_{eff}(e)$ \citep{Marotta2006,Dabrowski2008,Thieulot2014}. Benchmark of the falling block \citep{Gerya2003a,Gerya2010b,Thieulot2011}
-is performed to verify that the code can correctly deal buoyancy driven flows with strong viscosity contrasts (Section \ref{sec:block}).
-
-\section{Sticky air and free surface}\label{sec:sticky}
-The Earth's surface can be treated by means of either the so-called sticky-air or a true free surface method, both of them implemented in the code. In the
-sticky-air method the surface is approximated with the introduction of a buoyant layer with a viscosity at least four orders of magnitude lower than the crust
-\citep{Schmeling2008,Crameri2012} and the interface between lithosphere and air is defined using a chain of passive markers that are advected as the Lagrangian
-markers. The correctness of the evolution of the markers chain is tested with the experiment of a 2D time-dependent Stokes sphere below a free surface and
-compared with results from ASPECT\footnote{\url{https://aspect.geodynamics.org/}} \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} (Section
-\ref{sec:sphere}). In the true free surface case the top boundary is assumed stress-free and velocities are not fixed. In this case topography variations are
-described by vertical deformations of the mesh that depend on the velocity field of the nodes that identify the free surface, while horizontal deformations are
-not taken into account. This procedure is known as the Arbitrary Lagrangian-Eulerian (ALE) method and its implementation follows the technique described in
-\citet{Thieulot2011}. However, although the implementation of a true free surface better reproduces laboratory experiment, extremely small time steps can be
-necessary to maintain stability \citep{Kaus2010a,Quinquis2011,Thieulot2014}. Therefore, the stability algorithm proposed by \citet{Kaus2010a} is implemented
-to avoid instabilities due to high density differences at the free surface when using too large time steps. The implementation of this algorithm are tested by
-performing the experiment described by \citet{Kaus2010a} (Section \ref{sec:stab}).
-
-The topography relaxation benchmark proposed by \citet{Crameri2012} is performed to verify that the code correctly recovers topography variations in case of
-both the sticky-air and the true free surface method (Section \ref{sec:crameri}). Finally, Section \ref{sec:subduction} show the results of the spontaneous
-subduction experiment described by \citet{Schmeling2008}.
-
-\section{Erosion and sedimentation}\label{sec:erosion}
-Surface processes at either the lithosphere-air interface or in correspondence of the free surface have been implemented by means of the software
-Fastscape\footnote{\url{https://fastscape.org/}} \citep{Braun2013,Cordonnier2019,Yuan2019a,Yuan2019b}. FastScapeLib is a set of routines that solve the stream
-power law (enriched by a sediment transport/deposition term) \citep{Yuan2019a}, hillslope diffusion and marine transport and deposition \citep{Yuan2019b}.
-The partial differential equation solved by FastScapeLib is:
-\begin{equation}
-\frac{\partial h}{ \partial t}=U-K_f A^{m_e}S^{n_e} + \frac{G}{A}\int_A \left( U-\frac{\partial h}{\partial t} \right) dA + K_d \bm{\nabla}^2 h 
-\end{equation}
-where $h$ is the topography, $U$ is the uplift, $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $S$ is the slope, $m_e$ and $n_e$ are
-parameters that should be chosen to constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}, $G$ is the dimensionless
-deposition/transport coefficient for the enriched stream power law and $K_d$ is the effective diffusivity. Terms $K_f A^{m_e}S^{n_e}$ and $K_d \bm{\nabla}^2 h$
-consider long-range processes, described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, and short-range processes,
-proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}.
-
-In case of sticky air, slope is determined by means of the markers chain, which is vertically corrected considering erosion and sedimentation rates. After the
-correction, continental markers above the markers chain are transformed in air markers and air markers below the markers chain become sediments. Similarly,
-slope is determined considering the nodes on the top boundary in case of true free surface and they are vertically corrected as for the markers chain. In this
-case, continental markers above the top boundary are simply deleted.
-%Surface processes at the lithosphere-air interface are implemented taking into account fluvial erosion (long-range processes), hillslope processes (short-range
-%processes) and failure of slopes exceeding a threshold angle $\theta$ \citep{Burov1997,Tucker1999,Sacek2012,Croissant2014,Ueda2015,Beucher2020}. Erosion rates
-%$\dot{E}_{lr}$ due to long-range processes can be described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{lr}=K_f A^{m_e} \bm{\nabla} h^{n_e}
-%\end{equation}
-%where $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $\bm{\nabla} h$ is the slope and $m_e$ and $n_e$ are parameters chosen in to
-%constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}. Erosion rates $\dot{E}_{sr}$ due to short-range processes are
-%implemented by assuming that transport is linearly proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{sr}=K_d \bm{\nabla}^2 h
-%\end{equation}
-%where $K_d$ is the effective diffusivity. In addition, the slope is modified if it still exceeds a threshold angle after long- and short-range processes
-%\citep{Ueda2015}.
-
-%Offshore sedimentation processes are implemented similarly to hillslope processes but with a depth-dependent coefficient \citep{Sacek2012,Beucher2020}.
-%The sedimentation rate $\dot{S}$ is calculated as
-%\begin{equation}
-%\dot{S}=K_s \bm{\nabla}^2 h
-%\end{equation}
-%with
-%\begin{equation}
-%K_s=K^*_s \exp\left(\frac{h-h_{sl}}{h_0}\right)
-%\end{equation}
-%where $K^*_s$ is a constant, $h_{sl}$ is the sea level and $h_0$ is a characteristic length. In case of sticky air, slope is determined by means of the markers
-%chain, which is vertically corrected considering erosion and sedimentation rates. After the correction, continental markers above the markers chain are
-%transformed in air markers and air markers below the markers chain become sediments. Similarly, slope is determined considering the nodes on the top boundary
-%in case of true free surface and they are vertically corrected as for the markers chain. In this case, continental markers above the top boundary are simply
-%deleted.
-
-\section{Non-linear rheologies}\label{sec:nonlinear}
-Non-linear rheologies are implemented combining viscous creep (dislocation and diffusion) and plastic yielding. For each marker, diffusion and dislocation
-viscosity ($\eta_{df|ds}$) can be determined as follows \citep{Karato1993,Warren2008b,Wang2016}:
-\begin{equation}\label{eq:viscous}
-\eta_{df|ds}=f_s \left(\frac{d^m}{A}\right)^{\frac{1}{n}} \left (\frac{2^{\frac{1-n}{n}}}{3^{\frac{n+1}{2n}}} \right)
-I_2^{\frac{(1-n)}{n}}\exp\left(\frac{Q+pV}{nRT}\right)
-\end{equation}
-where $f_s$ is a scaling factor used to represent lithologies that are stronger or weaker than the base set, $d$ is the grain size, $m$ is the grain size exponent,
-$n$ is the stress exponent, $A$ is the uniaxial pre-exponential factor, $I_2$ is the square root of the second invariant of the strain rate tensor, $Q$ is the
-activation energy, $p$ is the pressure, $V$ is the activation volume, $R$ is the gas constant and $T$ is the temperature.
-Pressures and temperatures are determined by the interpolation of the nodal parameters, while strain rates are calculated by means of the
-derivative of the velocities on the nodes. In case of diffusion creep $n=1$ and $m>0$, in case of dislocation creep $n>1$ and $m=0$. The viscous creep $\eta_{cp}$
-is then calculated as the harmonic average between $\eta_{df}$ and $\eta_{ds}$:
-\begin{equation}\label{eq:visco_creep}
-\eta_{cp}=\left(\frac{1}{\eta_{df}}+\frac{1}{\eta_{ds}}\right)^{-1}
-\end{equation}
-The implementation of the non-linear viscous creep viscosity depending on the strain rate is tested by means of the slab detachment benchmark
-(Section \ref{sec:slab}).
-
-Plastic yielding is implemented rescaling $\eta_{cp}$ in order to limit the stress below the yield stress $\sigma_y$ \citep{Thieulot2008,Thieulot2014,Glerum2018},
-obtaining
-\begin{equation}\label{eq:plastic}
-\eta_{pl}=\frac{\sigma_y}{2I_2}
-\end{equation}
-where $\eta_{pl}$ is the plastic viscosity and the yield stress is determined following the Drucker-Prager criterion, such as
-\begin{equation}\label{eq:yield}
-\sigma_y=C\;\cos(\phi)+p\;\sin(\phi)
-\end{equation}
-where $C$ is the cohesion and $\phi$ is the internal friction angle. In case of negative pressure it is imposed equal to 0, so that negative yield stress are
-excluded. The correctness of the non-linear solution in case of plastic viscosity with variable internal friction angle is verified performing indenter and
-brick experiments (Sections \ref{sec:indenter} and \ref{sec:brick}, respectively).
-
-Strain softening is taken into account for both viscous creep and plastic viscosity \citep{Huismans2003,Babeyko2005,Huismans2005,Sobolev2005,Warren2008a} by
-means of the accumulated strain $\epsilon$ memorised by each marker. Softening in the viscous creep determines a linear decrease of $\eta_{vc}$ by means of a
-viscous strain softening factor $W_S$ that increases linearly from $W_{S_0}$ to $W_{S_{\infty}}$ for $\epsilon_{S_0}<\epsilon_S<\epsilon_{S_{\infty}}$ 
-\citep{Warren2008a}. This viscous softening can be related to strain-induced grain size reduction \citep{Warren2008a}. Differently, plastic softening is 
-simulated with a linear decrease of internal friction angle $\phi(\epsilon)$ and cohesion $C(\epsilon)$ in according to
-\begin{eqnarray}
-\label{eq:friction1}\phi(\epsilon)&=&\phi_0+(\phi_{\infty}-\phi_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}\\
-\label{eq:friction2}C(\epsilon)&=&C_0+(C_{\infty}-C_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}
-\end{eqnarray}
-where $\phi_0$, $C_0$ and $\phi_{\infty}$, $C_{\infty}$ are internal friction angle and cohesion for $\epsilon_0$ and $\epsilon_{\infty}$, respectively
-\citep{Huismans2003,Huismans2005,Warren2008a,Thieulot2014}. Plastic softening approximates deformation-induced softening of faults and brittle shear zones
-\citep{Warren2008a}.
-
-Plastic yielding and viscous creep are then combined to obtain a viscoplastic viscosity $\eta_{vp}$ as follows, assuming that they are independent processes
-\citep{Karato2008,Glerum2018}:
-\begin{equation}\label{eq:viscoplastic}
-\eta_{vp}=\min(\eta_{cp},\eta_{pl})
-\end{equation}
-Finally, effective viscosity $\eta_{eff}$ is capped by the minimum and the maximum viscosity ($\eta_{min}$ and $\eta_{max}$, respectively) to avoid extremely
-low or high viscosity \citep{Glerum2018} as follows
-\begin{equation}\label{eq:effective}
-\eta_{eff}=\min(\max(\eta_{vp},\eta_{min}),\eta_{max})
-\end{equation}
-Viscous creep and plastic yielding are non-linear rheologies because of their dependence on the velocity field through pressure and strain rates. Therefore,
-the solution is determined by means of Picard-type iterations, until convergence of the velocity field \citep{Glerum2018}. The convergence is verified at each
-iteration $i$ via the nonlinear residual $\bm{\mathcal{R}}^i$ that can be determined as
-\begin{equation}\label{eq:convergence}
-\bm{\mathcal{R}}^i= \mathbb{K} (\eta_{eff}(\dot{\bm{\epsilon}}^{i-1},p^{i-1}))\cdot \bm{v}^{i-1} - \bm{f}^i
-\end{equation}
-%\begin{equation}\label{eq:convergence}
-%\chi=1-\frac{\langle (\bm{v}^i-\langle\bm{v}^i\rangle)\cdot (\bm{v}^{i+1}-\langle\bm{v}^{i+1}\rangle) \rangle}{\sqrt{\sigma^i\sigma^{i+1}}}
-%\end{equation}
-%Iterations are performed until either convergence ($\chi<tol$) or maximal number of non-linear iterations ($it_{max}$) is reached.
-where $\mathbb{K}$ is the Stiffness matrix and $\bm{f}$ is the right hand side vector \citep{Spiegelman2016,Glerum2018}.  The $L_2$-norm is extracted from
-$\bm{\mathcal{R}}^i$ and it is normalised as follows
-\begin{equation}\label{eq:convergence_res}
-\frac{||\bm{\mathcal{R}}^i||_2}{||\bm{\mathcal{R}}^0||_2}
-\end{equation}
-where $||\bm{\mathcal{R}}^0||_2$ is the $L_2$-norm of the first nonlinear iteration.
-Since often the normalised nonlinear residual drop very quickly over the first 2-3 iterations, also the $L_2$-norm of velocity and pressure residuals are
-calculated as
-\begin{equation}\label{eq:convergence_vel}
-\frac{||\Delta\bm{v}||_2}{||\bm{v}||_2}=\frac{||\bm{v}_i-\bm{v}_{i-1}||_2}{||\bm{v}_{i}||_2}
-\end{equation}
-and
-\begin{equation}\label{eq:convergence_press}
-  \frac{||\Delta p||_2}{||p||_2}=\frac{||p_i-p_{i-1}||_2}{||p_{i}||_2}
-\end{equation}
-respectively. Iterations are performed until either maximal number of non-linear iterations ($it_{max}$) is reached or all the normalised nonlinear residual, 
-the velocity residual and the pressure residual converge under a defined tolerance $tol$.
-
-\section{The energy equation}\label{sec:en_eq}
-The stabilisation of the advection term of the energy equation (Eq. \ref{eq:energy}) needed to avoid possible oscillations in the thermal solution in the case
-when advection dominates over diffusion, is implemented by means of a streamline-upwind Petrov–Galerkin (SUPG) method, for which the advection term is modified
-follows the discussion in \citet{Thieulot2011} and \citet{Thieulot2014} (see Section \ref{sec:advection} for benchmark). The total internal heat production
-$H_{tot}$ in Eq. \ref{eq:energy} includes radiogenic heating $H_r$, shear heating $H_s$ and adiabatic heating $H_a$, where
-\begin{equation}\label{eq:shear}
-H_s=2\eta \dot{\bm{\epsilon}} : \dot{\bm{\epsilon}} = 2\eta(\dot{\epsilon}_{xx}^2+\dot{\epsilon}_{yy}^2+2\dot{\epsilon}_{xy}^2)
-\end{equation}
-and
-\begin{equation}\label{eq:adiabatic}
-H_a=T \alpha \frac{Dp}{Dt} \approx -\alpha T \rho g_y v
-\end{equation}
-The correctness of the implementation and the computation of shear and adiabatic heating is verified performing an exercise with an analytical solution
-(Section \ref{sec:simple_shear}) and Exercise 9.4 in \citet{Gerya2010b} (Section \ref{sec:shear}). Furthermore, mantle convection and thin layer entrainment
-benchmarks (Sections \ref{sec:mantle} and \ref{sec:thin}, respectively) are performed to verify that the code solve correctly Eqs. \ref{eq:momentum_penalty}
-and \ref{eq:energy} in case of a temperature field characterised by an initial perturbation and with a temperature-dependent density. At last, the viscoplastic
-mantle convection benchmark is performed to check also the viscoplasticity in case of a initial perturbed temperature field (Section \ref{sec:visco_mantle}).
-
-\section{Phase transitions and hydration}\label{sec:phase}
-Variations of effective density, specific heat and coefficient of thermal expansion during the evolution of crustal and mantle materials at different
-pressure-temperature ($p$-$T$) conditions are computed by means of version 6.8.6 of Perple\_X software package \citep{Connolly2005}, similarly to the
-implementation described in \citet{Marotta2020}. By default, mineral assemblages and properties of each lithology (oceanic crust, lower and upper continental
-crust, sediments and dry or hydrated mantle) are calculated for temperatures from \SIrange{330}{1600}{\kelvin}, with increments of approximately \SI{4}{\kelvin},
-and pressures from \SIrange{0.1}{30}{\giga\pascal}, with increments of approximately \SI{0.03}{\giga\pascal}, for a total of almost 100000 points.
-
-The introduction of phase transitions produces two main effects that must be taken into account in numerical models: variations in density and release/absorption
-of latent heat \citep{Tackley2010}. Effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into
-account considering effective coefficient of thermal expansion \citep{Christensen1985,Zhong2015}. Similarly, effects of latent heat in the energy equation
-(eq \ref{eq:energy}) are taken into account by considering effective specific heat and coefficient of thermal expansion
-\citep{Christensen1985,Tackley2010,Zhong2015}. Effective density, specific heat and coefficient of thermal expansion are included in Perple\_X files and are
-assumed by each Lagrangian marker in according to its $p$-$T$ conditions.
-
-Hydration processes related to dehydration of subducting lithosphere \citep{Schmidt1998,Liu2007,Faccenda2009,Faccenda2010,Faccenda2014,Rosas2016} strongly
-influence the thermo-mechanics inside the mantle wedge, mainly because of both weakening effects on the mantle rheology and density variations in case of
-mantle serpentinization \citep{Gerya2002,Honda2003,Arcay2005,Roda2010,Regorda2017}. In the case that hydration is switched on, the amount of bound and free
-water is memorised by each marker, following the implementation of \citet{Quinquis2014}; hydration and dehydration processes are related to the amount of bound
-water of each marker and to the maximum amount of water it can transport, i.e. if the amount of bound water exceeds the maximum amount of water, the marker
-dehydrates and releases free water that can hydrates under-saturated markers. Maximum water content of each marker is determined as function of lithology and
-$p$-$T$ conditions and it is calculated using Perple\_X, in the same way than effective density, specific heat and coefficient of thermal expansion. Bound water
-is advected together with the markers, neglecting the effect of bound water diffusion, while free water simply migrates vertically and is not coupled to the
-solid-phase flow of the mantle wedge \citep{Arcay2005,Quinquis2014}. The correctness of the migration scheme of the free water and its absorption into bound
-water by under-saturated markers are verified performing the experiment described in \citet{Quinquis2014} (Sec. \ref{sec:quinquis}).
-
-The mantle viscosity weakening related to the amount of water has been extensively studied \citep{Chopra1981,Hirth2003} and it is implemented in according to
-\citet{Arcay2005} and \citet{Horiuchi2016}, as follows
-\begin{equation}\label{eq:wet_visc}
-\eta_{wet}=\eta_{dry}\left(\left[1-\frac{1}{f_v}\right]\exp\left(-\frac{[OH^-]}{[OH^-]_0}\right)+\frac{1}{f_v}\right)
-\end{equation}
-where $f_v$ is the viscosity reduction factor between dry and wet conditions, $[OH^-]$ is the water content and $[OH^-]_0$ is a reference water content set to
-620 ppm (0.062 wt.\%) \citep{Arcay2005}. By default, $f_v$ is set to 100, in according to \citet{Arcay2005} and \citet{Horiuchi2016}.
-
-\section{Mantle melting}\label{sec:melt}
-Melt fraction in the mantle depends on temperature ($T$, in \textdegree C), pressure ($p$, in \si{\giga\pascal}) and the water content in the melt ($X_{H_2O}$, in wt.\%).
-The determination of the melt fraction $M$ is implemented as explained by \citet{Katz2003}, according to the successive modification by \citet{Langmuir2006} and
-\citet{Kelley2010}, as follows
-\begin{eqnarray}\label{eq:melt_fr}
-M(p,T,X_{H_2O})&=&-T+T_{sol}+(x_t\cdot ln(p+y_t))\cdot X_{H_2O}-\nonumber\\
-&&+K_m\left(\frac{X_{H_2O}^{bulk}}{D_{H_2O}(1-X_{H_2O})+X_{H_2O}}\right)^{\gamma}
-\end{eqnarray}
-where $T_{sol}$ is the temperature of the dry solidus, $D_{H_2O}$ is the partition coefficient and $x_t\cdot log(p+y_t)$ indicates the pressure dependence of
-the melting curve \citep{Kelley2010}. $x_t$, $y_t$, $K_m$ and $\gamma$ are parameters chosen in according to \citet{Kelley2010}.
-However, $X_{H_2O}$ depends on the melt fraction as
-\begin{equation}\label{eq:melt_wt}
-X_{H_2O}(M)=\frac{X_{H_2O}^{bulk}}{D_{H_2O}+M(1-D_{H_2O})}
-\end{equation}
-and a numerical solution can be found using a root-finder method \citep{Katz2003,Wang2016}. Furthermore, the content of water in the melt is limited by the
-pressure-dependent saturation concentration of water in the melt, determined as
-\begin{equation}\label{eq:melt_max}
-X_{H_2O}^{sat}=\chi_1 p^{\lambda_m}+\chi_2 p
-\end{equation}
-where $\chi_1$, $\chi_2$ and $\lambda_m$ are parameters chosen according to \citet{Katz2003}, \citet{Langmuir2006} and \citet{Kelley2010}. The correctness in
-the determination of melt fractions and water contents in the melt are tested comparing model results with experimental curves obtained by \citet{Katz2003}
-(Sec. \ref{sec:katz}).
-
-In case of melting, viscous creep viscosity ($\eta_{vp}$) is modified according to \citet{Wang2016}, as
-\begin{equation}\label{eq:melt_visc}
-\eta_{melt}=\eta_{vp}\cdot \exp(\alpha_m M)
-\end{equation}
-where $\alpha_m$ are melt fraction factors for dislocation and diffusion creep, chosen in according to \citet{Wang2016}.
-
-Effective density ($\rho_{eff}$) for partially molten rocks is calculated as
-\begin{equation}\label{eq:melt_dens}
-\rho_{eff}=\rho_s(1-M)+\rho_m M
-\end{equation}
-where $\rho_s$ and $\rho_m$ are the densities of the solid and the molten rock, respectively \citep{Gerya2004b,Gerya2007,Wang2016}. The density of the
-solid rock is extracted by Perple\_X, while the density of the molten portion is calculated as
-\begin{equation}\label{eq:melt_dens_M}
-\rho_m(p,T)=\rho_0[1-\alpha(T-T_0)][1+\beta(p-p_0)]
-\end{equation}
-where $\rho_0$ is the density at the reference temperature ($T_0$) and pressure($p_0$), and $\alpha$ and $\beta$ are the thermal expansion and compressibility
-coefficients, respectively \citep{Gerya2004b,Gerya2007,Wang2016}.
-
-As for the phase transitions, effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into account
-considering effective coefficient of thermal expansion. Similarly, latent heat due to melting/crystallisation can be implicitly included in the energy equation
-(Eq. \ref{eq:energy}) by considering effective specific heat ($C_{p_{eff}}$) and thermal expansion ($\alpha_{eff}$) of the partially molten rocks, in according to
-\begin{eqnarray}
-\label{eq:cp_effective}C_{p_{eff}}&=&C_p+H_L\left(\frac{\partial M}{\partial T}\right)_p \\
-\label{eq:alpha_effective}\alpha_{eff}&=&\alpha+\rho\frac{H_L}{T}\left(\frac{\partial M}{\partial p}\right)_T
-\end{eqnarray}
-where $C_p$ and $\alpha$ are the specific heat and the coefficient of thermal expansion of the solid rock, respectively, and $H_L$ is the latent heat of the
-molten rock \citep{Gerya2004b,Gerya2007,Tackley2010}.
-
-\section{Benchmarks}\label{sec:benchmark}
-\subsection{Error measurements}\label{sec:error}
-In order to determine the accuracy of velocity and pressure field of the benchmarks, the $L_2$-norm is computed by numerical integration on the Gauss-Legendre
-quadrature points. $L_2$-norm for pressure and velocity errors can be evaluated as
-\begin{eqnarray}
-\textrm{err}_p&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|p^n_i(r_q)-p^a_i(r_q)|^2\right)w_q|J_q|}\nonumber \\
-\textrm{err}_v&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|u^n_i(r_q)-u^a_i(r_q)|^2+|v^n_i(r_q)-v^a_i(r_q)|^2\right)w_q|J_q|}\nonumber
-\end{eqnarray}
-respectively, where $n_e$ is the number of elements, $n_q$ is the number of quadrature points per element, $p^n_i(r_q)$ and $p^a_i(r_q)$ are the numerical and
-analytical pressure, respectively, in each quadrature point $q$, $w_q$  and $J_q$ are the weight and the Jacobian at the quadrature point
-$q$, $u^n_i(r_q)$, $v^n_i(r_q)$, $u^a_i(r_q)$ and $v^a_i(r_q)$ are the numerical and analytical velocities, respectively, in each quadrature point $q$.
-
-Other quantities used as comparison with original benchmarks are the root-mean-square velocity over the whole domain and over the surface
-\begin{eqnarray}
-v_{\textrm{rms}}&=&\sqrt{\int_{\Omega}|\bm{v}|^2 d\Omega}\nonumber \\
-v_{\textrm{rms}}^{top}&=&\sqrt{\int_0^1 u^2 \bigg|_{y=1} dx}\nonumber
-\end{eqnarray}
-respectively, top and bottom Nusselt numbers
-\[\textrm{Nu}^{top/bottom}=-\int_0^1\frac{\partial T}{\partial y} \bigg|_{y=1/y=0} dx\]
-the average rate of work done against gravity
-\[\left\langle W \right\rangle=\int_{\Omega}T u_y d\Omega\]
-the average rate of viscous dissipation
-\[\left\langle \Phi \right\rangle=\int_{\Omega}\tau_{ij}\dot{\epsilon}_{ij} d\Omega\]
-and the percentage error between $\left\langle W \right\rangle$ and $\left\langle \Phi \right\rangle/Ra$
-\[\delta=\frac{\left|\left\langle W \right\rangle -\frac{\left\langle \Phi \right\rangle}{Ra} \right|}{\max \left(\left\langle W \right\rangle,
-\frac{\left\langle \Phi \right\rangle}{Ra}\right)}\times100\]
-
-\subsection{Stokes flow}\label{sec:stokes}
-The problem consists of determining velocity field $(u,v)$ and pressure $p$ in case of a manufactured solution with prescribed body forces such as
-\begin{eqnarray}
-b_1&=&(12-24y)x^4+(-24+48y)x^3+(-48y+72y^2-48y^3+12)x^2+\nonumber \\
-&&+(-2+24y-72y^2+48y^3)x+1-4y+12y^2-8y^3\nonumber \\
-b_2&=&(8-48y+48y^2)x^3+(-12+72y-72y^2)x^2+(4-24y+48y^2-48y^3+\nonumber \\
-&&+24y^4)x-12y^2+24y^3-12y^4\nonumber
-\end{eqnarray}
-for which the exact solution is:
-\begin{eqnarray}
-u(x,y)&=&x^2(1-x)^2(2y-6y^2+4y^3)\nonumber \\
-v(x,y)&=&-y^2(1-y)^2(2x-6x^2+4x^3)\nonumber \\
-p(x,y)&=&x(1-x)-\frac{1}{6}\nonumber
-\end{eqnarray}
-The domain is a unit square a constant viscosity ($\eta=1$) and the penalty parameter is set to $\lambda=10^7$. Velocity boundary conditions are set to no slip
-$(\bm{v}=\bm{0})$ on all boundaries. The problem is performed for different grid resolution between \num{8x8} and \num{1024x1024} elements.
-
-The errors between the analytical solution and the numerical prediction of the velocity and the pressure field are calculated as in Section \ref{sec:error}.
-Fig. \ref{fig:errors}a shows that both velocity and pressure field converge to the exact solution with the decrease of the element size, following the
-theoretical convergence rate. The convergence of pressure, in contrast with the observation by \citet{Donea2003} for $Q_1 \times P_0$ elements, indicate the
-effectiveness of the smoothing procedure. Smoothed pressure field for grid resolution of \num{128x128} elements is shown in Fig. \ref{fig:errors}b (continuous
-red line), in comparison with the analytical solution (dashed blue line). Solve times and memory usage needed to generate the solution are shown in
-Fig. \ref{fig:MUMPS}. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Solver/Stokes_Flow}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/errors.pdf}
-\caption{Velocity and pressure error for the Stokes flow experiment between generated and analytical solution as a function of element size (panel a),
-and comparison between smoothed pressure and analytical solution as function of $x$ coordinate for a grid resolution of \num{128x128} elements (panel b).}
-\label{fig:errors}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/MUMPS.pdf}
-\caption{Panel a) Analysis, factorisation, solution and total solve times (panel a) and factorisation memory usage (panel b) as a function of the total number
-of degrees of freedom for the Stokes flow experiment.}
-\label{fig:MUMPS}
-\end{figure}
-
-\subsection{Markers advection}\label{sec:runge}
-The 2nd-order Runge-Kutta advection scheme in space is tested by means of the Zalesak disk test \citep{Zalesak1979,Thieulot2014}. The benchmark is performed
-in a unit square domain with a grid resolution of \num{32x32} elements and 1000000 markers and values of Courant number between 0.25 and 1.
-At $t=0$, the disk is centred at position $(0.5;0.75)$ with a radius $R=0.15$ and has a vertical fissure 0.05 wide and 0.2 high (orange in Fig. \ref{fig:runge})
-and with the velocity field prescribed in the entire domain as
-\begin{eqnarray}
-u(x,y)&=&2\pi\left(y-\frac{L_x}{2}\right)\nonumber \\
-v(x,y)&=&-2\pi\left(x-\frac{L_x}{2}\right)\nonumber
-\end{eqnarray}
-the markers are back to their initial location after a $2\pi$ rotation ($t=1$). The clockwise rotation of the disk at $t=0.25$, 0.5 and 0.75 
-is shown in Fig. \ref{fig:runge} (red, blue and green, respectively). The distance from the centre of 1 marker is calculated throughout the rotation to
-evaluate the error for different values of Courant number. As expected, the error increase with the increasing of the time step, as shown in 
-Fig. \ref{fig:runge_err}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/Zalesak_Disk}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=200px]{./Figures/Zalesak.pdf}
-\caption{Position of the Zalesak disk at $t=0$, 0.25, 0.5 and 0.75 (orange, red, blue and green, respectively) for the case with Courant number of 1.}
-\label{fig:runge}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/zal_marker.pdf}
-\caption{Distance from the centre as function of time for values of Courant number of  of 0.25, 0.3, 0.5, 0.75 and 1 (green, orange, blue, black and red,
-respectively). R indicates the distance from the centre considering a perfect circle trajectory.}
-\label{fig:runge_err}
-\end{figure}
-
-\subsection{Conservative Velocity Interpolation (CVI)}\label{sec:CVI}
-The CVI correction is checked by means of the Stokes flow experiment by \citet{Donea2003}, as presented in \ref{sec:stokes}. The advection of Lagrangian
-markers is performed using either a 2nd-order or a 4th-order Runge-Kutta scheme with CFL=0.5 and an initial random distribution of 25 markers per element.
-Fig. \ref{fig:CVI} shows the minimum and maximum values of markers per element throughout the experiment with and without the CVI correction and using
-different Runge-Kutta schemes. The introduction of the CVI correction allows the
-tests to run for times longer than $t=2000$ without show empty elements, while tests without the CVI correction stop at approximately $t=600$, when at least
-one element goes empty. Fig. \ref{fig:CVI_time} shows the different number of markers per element and the markers distribution at $t=600$ obtained without and
-with the CVI correction (panels b and c, respectively) in case of a 2nd-order Runge-Kutta scheme. All data can be found at
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/CVI}.
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/CVI.pdf}
-\caption{Maximum and minimum number of markers per element throughout the experiment, with (green and orange lines) or without (purple and blue lines)
-the CVI correction and using either a 2nd-order (blue and orange lines) or a 4th-order (purple and green lines) Runge-Kutta scheme.}
-\label{fig:CVI}
-\end{figure}
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/CVI_time.pdf}
-\caption{Number of markers per element and markers distribution through time using 2nd.order Runge-Kutta scheme.
-Panels a: number of markers per element (panel a1) and distribution of the markers (panel a2) after the first
-iteration; panels b and c: comparison between number of markers per element (panels b1 and c1) and markers distribution (panels b2 and c2) without and with
-the CVI correction at $t=600$; panels d: number of markers per element (panel d1) and markers distribution (panel d2) in case of CVI correction at $t=2000$.}
-\label{fig:CVI_time}
-\end{figure}
-
-\subsection{Poiseuille flow}\label{sec:poiseuille}
-The domain is a rectangle with $L_x=2$ and $L_y=1$ and constant density and viscosity ($\rho=1$ and $\eta=1$), gravity acceleration $\bm{g}=\bm{0}$ and penalty
-parameter $\lambda=10^8$. The grid is composed by \num{40x20} elements. Velocity boundary conditions are set to no slip $(\bm{v}=\bm{0})$ at the top and the
-bottom, and a parabolic profile is imposed on the sides, with $u=y(1-y)$ and $v=0$. The analytical solution is then given by:
-\begin{eqnarray}
-u(x,y)&=&y(1-y)\nonumber \\
-v(x,y,)&=&0\nonumber \\
-p(x,y)&=&2\eta\left(\frac{L_x}{2}-x\right)\nonumber
-\end{eqnarray}
-
-The velocity field predicted by the model follows the expected parabolic profile (Fig. \ref{fig:poiseuille}a and Fig. \ref{fig:poi_plot}a) and, in case of the
-classic penalty method with no iterations, the pressure is clearly related to the divergence by means of the penalty parameter (Fig. \ref{fig:poiseuille}b and
-c and Fig. \ref{fig:poi_plot}b). Fig. \ref{fig:poiseuille}d shows that one Uzawa iteration is sufficient to bring the divergence down to \num{e-15}, with no
-correlation with the pressure field. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Poiseuille%20Flow}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Poiseuille.pdf}
-\caption{Velocity field (panel a), pressure (panel b) and divergence velocity of a Poiseuille flow in case of the classic penalty method (no iterations) and
-after one Uzawa iteration (panel c and d, respectively).}
-\label{fig:poiseuille}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Analytical.pdf}
-\caption{Velocity (panel a) and pressure (panel b) field predicted by the model for a Poiseuille flow with respect with their analytical solutions. The
-velocity field is plotted as function of the vertical coordinate in $L_x/2$ and the pressure is plotted as function of the x coordinate in $L_y/2$.}
-\label{fig:poi_plot}
-\end{figure}
-
-\subsection{Instantaneous 2D sphere}\label{sec:ist_sphere}
-The domain is a unit square with gravity $\bm{g}=(0,-1)$. The fluid has constant density and viscosity ($\rho_f=1$ and $\eta_f=1$). The sphere is in the middle
-of the domain with a radius R = 0.123456798 and has constant density and viscosity ($\rho_s=10^{-2}+\rho_f$ and $\eta_s=10^3 \cdot \eta_f$).
-We distinguish three different types of velocity boundary conditions:
-\begin{enumerate}
-\item FS: free slip conditions on all sides.
-\item NS: no slip conditions on all sides.
-\item OT: free slip conditions on the sides and the bottom, and open boundary at the top.
-\end{enumerate}
-All three cases are performed for different grid resolutions, between \num{16x16} and \num{512x512} elements with 50 randomly distributed markers per element,
-and different average schemes for the viscosity (harmonic, geometric and arithmetic). The velocity in the centre of the sphere $\bm{v}(0.5;0.5)$), minimum and
-maximum velocities ($u_{min}$, $u_{max}$, $v_{min}$, $v_{max}$) and pressures ($p_{min}$, $p_{max}$) on the entire domain, $v_{\textrm{rms}}$ and average pressure on the
-entire domain ($p_{avg}$) are compared with solutions generated by ASPECT \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and all results
-can be found at \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere2D} and 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Instantaneous_2D_Sphere}. Results in terms of the $v_{\textrm{rms}}$ for NS model 
-are also shown in Fig. \ref{fig:inst_sphere}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_NS.pdf}
-\caption{$v_{\textrm{rms}}$ as function of element size for the instantaneous 2D sphere experiment in case of no slip boundary conditions and with different average
-schemes for the viscosity. Results are compared with results obtained by other numerical codes.}
-\label{fig:inst_sphere}
-\end{figure}
-
-\subsection{Rayleigh-Taylor instability}\label{sec:rayleigh}
-This problem has been originally presented by \citet{vanKeken1997} and here it is performed as the isoviscous case (Case 1a). The domain has $L_x=0.9142$ and
-$L_y=1$ and gravity $\bm{g}=(0,-1)$. Two fluids with same constant viscosities ($\eta_1=\eta_2=1$) and different densities ($\rho_1=1000$ and $\rho_2=1010$),
-with the lighter fluid at the bottom. The initial interface between the fluids is given by $y(x)=0.2+0.02 \cos \left(\frac{\pi x}{L_x}\right)$. The experiment
-is performed with different grid sizes (\num{50x50}, \num{80x80}, \num{100x100} and \num{256x256}). A total of 1960000 markers are randomly distributed at the
-beginning of the simulation. Velocity boundary conditions are set to no slip at the top and the bottom, and to free slip at the sides of the domain.
-
-$v_{\textrm{rms}}$ as function of time is reported in Fig. \ref{fig:RT}, matching well with results shown by \citet{vanKeken1997}, \citet{Tackley2003} and
-\citet{Thieulot2014}. Fig. \ref{fig:rayleigh} shows the evolution of the experiment at different time steps. All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Rayleigh_Taylor_experiment/ISOVISCOUS}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/RT.pdf}
-\caption{$v_{\textrm{rms}}$ of the Rayleigh-Taylor experiment as function of time for different resolution of the grid.}
-\label{fig:RT}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=300px]{./Figures/Rayleigh.pdf}
-\caption{Evolution of the Rayleigh-Taylor experiment for a grid of \num{256x256} elements at $t=0, 500, 1000$ and $2000$ (panels a, b, c and d, respectively).}
-\label{fig:rayleigh}
-\end{figure}
-
-\subsection{Falling blocks}\label{sec:block}
-This benchmark is described as proposed by \citet{Gerya2003a}, \citet{Gerya2010b} and \citet{Thieulot2011}. The domain is square with $L_x=L_y=$ \SI{500}{\km}
-and the grid is composed by \num{50x50} elements with 25 markers in each element. The block is initially centred at ($x=$ \SI{250}{\km}; $y=$ \SI{400}{\km})
-and has a size of \num{100x100} km (Fig. \ref{fig:block}a). The fluid surrounding the block has $\eta_f=$ \SI{e21}{\pascal\per\s} and $\rho_f=$
-\SI{3200}{\kg\per\cubic\m}. The benchmark tests with different viscosities of the block, with $\eta_b$ from \SIrange{e15}{e27}{\pascal\per\s}. In each
-experiment also the density $\rho_b$ of the block varies from \SIrange{3220}{9900}{\kg\per\cubic\m}. Velocity boundary conditions are set to free slip
-conditions on all sides of the domain. Density distributions at $t=$ \SI{20}{\mega\year} for $\rho_b=$ \SI{3300}{\kg\per\cubic\m} and different $\eta_b$
-are plotted in Fig. \ref{fig:block}b-f, showing that the code correctly preserve the block geometry in case of large viscosity contrast, with a block stiffer
-than the surrounding fluid (Fig. \ref{fig:block}f).
-
-The velocity in the centre of the falling block at $t=0$ is measured for all experiments and, since the velocity should increase with density contrast, the
-quantity $\bm{v}/(\rho_b-\rho_f)$ is plotted as function of the viscosity contrast. All results perfectly match with those from \citet{Gerya2010b} and they
-line up on a single curve, demonstrating that the code can correctly deal with large viscosity and density contrasts (Fig. \ref{fig:falling}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Falling%20blocks}.
-
-\begin{figure}
-\noindent\includegraphics[width=300px]{./Figures/Block.pdf}
-\centering
-\caption{Density evolution of the falling block experiment at $t=$ \SI{0}{\mega\year} (panel a) and $t=$ \SI{20}{\mega\year} for different viscosities of the
-block (panels b-f).}
-\label{fig:block}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Falling.pdf}
-\caption{Initial velocity relative to the density contrast at the centre of the falling block as function of the viscosity contrast between the block and the
-surrounding fluid.}
-\label{fig:falling}
-\end{figure}
-
-\subsection{2D time-dependent Stokes sphere with deformable free surface}\label{sec:sphere}
-This experiment is performed in a unit square domain with the gravity acceleration fixed to $g_y=-1$. The sphere has $\rho_s=2$ and $\eta_s=10^3$ and is
-initially centred at ($0.5;0.6$) with radius $R=0.123456789$. The fluid surrounding the sphere has $\rho_f=1$ and $\eta_f=1$ and occupies the domain for
-$y\leq 0.75$, while for $y>0.75$ the air has $\rho_a=0$ and $\eta_a=10^{-3}$. Velocity boundary conditions are set to free slip on all sides. The Courant
-number is set to 0.25. The experiments run for 200 s using grid resolutions from \num{150x150} to \num{512x512} elements, with 25 randomly distributed markers
-per element, and for different average schemes for the viscosity. The interface between the fluid and the air is tracked by means of the markers chain. Results
-in terms of velocity and pressure fields and topography variations are compared with results obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and they can be found at
-\url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere_fs2D} and
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Time_dependent_sphere}. Results in terms of the $v_{\textrm{rms}}$ in case of 
-an arithmetic average are also shown in Fig. \ref{fig:inst_spherefs}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_arithm.pdf}
-\caption{$v_{\textrm{rms}}$ for different grid resolutions as function of time for the 2D time-dependent Stokes sphere experiment in case of an arithmetic average.
-Results are compared with results obtained by ASPECT.}
-\label{fig:inst_spherefs}
-\end{figure}
-
-\subsection{Free surface stabilisation}\label{sec:stab}
-The benchmark is performed as discussed in \citet{Kaus2010a} and \citet{Thieulot2014}. The domain is square with $L_x=L_y=$ \SI{500}{\km} and a grid resolution
-of \num{200x200} elements, each of them containing 16 markers. A buoyant fluid with $\rho_1=$ \SI{3200}{\kg\per\cubic\metre} and $\eta_1=$ \SI{e20}{\pascal\s}
-is overlain by a denser fluid with $\rho_2=$ \SI{3300}{\kg\per\cubic\metre} and $\eta_2=$ \SI{e21}{\pascal\s}. The initial interface between the fluids has an
-initial sinusoidal shape of 5 km amplitude. Velocity boundary conditions are set to free slip conditions at the sides and no slip at the bottom, while the top
-is free surface. The experiment is performed with various fixed time steps, without the use of the Courant number. The vertical position of the free surface at
-$x=L_x$ is tracked for each simulation, with or without the stabilisation algorithm.
-
-The results show an instability (drunken sailor effect) increasing the time step in case that the stabilisation algorithm is not activated
-($dt=$ \SI{4200}{\year}, green line in Fig. \ref{fig:stab}), as already observed by \citet{Kaus2010a} and \citet{Thieulot2014}. Activating the algorithm the
-instability is fixed and the simulation remains stable using time step up to \SI{50000}{\year} (red line in Fig. \ref{fig:stab}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Stabilisation_algorithm}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Kaus_random.pdf}
-\caption{Evolution of the y coordinate in $x=L_x$ as function of time for different time steps, with and without the stabilisation algorithm.}
-\label{fig:stab}
-\end{figure}
-
-\subsection{Topography relaxation}\label{sec:crameri}
-This experiment is performed as in \citet{Crameri2012} using free surface and sticky air (Case 1). The domain is rectangular with $L_x=$ \SI{2800}{\km} and
-$L_y=$ \SI{700}{\km} in case of free surface and $L_y=$ \SI{800}{\km} in case of sticky air ($h_{st}=$ \SI{100}{\km}). The grid is composed by \num{256x64}
-elements for the free surface \num{560x320} elements for the sticky air. For both cases, there is a mantle of 600 km thickness with $\eta_m=$ \SI{e21}{\pascal\s}
-overlain by a highly viscous cosine-shaped lithosphere with thickness between 93 and 107 km and $\eta_l=$ \SI{e23}{\pascal\s}. Both mediums have $\rho=$
-\SI{3300}{\kg\per\cubic\metre}. In case of sticky air experiment, the air has $\rho_a=$ \SI{0}{\kg\per\cubic\metre} and $\eta_a=$ \SI{e18}{\pascal\s}. The
-gravity acceleration is set to $g_y=$ \SI{-10}{\metre\per\square\s}. Velocity boundary conditions are set to no slip at the bottom and free slip at the sides
-of the domain. In case of sticky air, velocities on top are set to free slip conditions.
-
-The maximum topography as function of time ($h(t)$) can be analytically derived in according to \[h(t)=h_0 \exp(\gamma t)\]
-where $h_0=7$ km is the initial topography, $\gamma=\num{-0.2139e-11}$ is the characteristic relaxation rate and $t$ is the time (black line in Fig.
-\ref{fig:crameri}). Maximum topography at the characteristic relaxation time $t_{rlx}=$ \SI{14.825}{\kilo\year} can be found to be $h_{rlx}=$ \SI{2576}{\m}.
-The results show that a viscosity of the air of \SI{e18}{\pascal\s} is needed in case of $h_{st}=$ \SI{100}{\km} to correctly describe the topography relaxation
-for this problem, as pointed out by \citet{Crameri2012} (purple line in Fig. \ref{fig:crameri}). The case with a true free surface well follows the analytical
-solution (orange line in Fig. \ref{fig:crameri}). In particular, at the relaxation time in case of the true free surface predicts a maximum topography of
-\SI{2570}{\m}, with an error of \SI{6}{\m} with respect to the analytical solution. Results are also compared with results obtained with a free surface in
-UNDERWORLD (\num{256x64} elements) and SULEC (\num{401x201} elements) (with courtesy from \citealp{Crameri2012}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Topography_relaxation}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Crameri.pdf}
-\caption{Maximum topography as function of time for the Crameri benchmark (coloured lines) in comparison with the analytical solution (black line) and results
-obtained with UNDERWORLD (yellow line) and SULEC (blue line).}
-\label{fig:crameri}
-\end{figure}
-
-\subsection{Spontaneous subduction}\label{sec:subduction}
-This experiment is performed as presented by \citet{Schmeling2008}, in case of both a sticky air and a true free surface and using different average schemes
-for the viscosities. The domain is rectangular with $L_x=$ \SI{3000}{\km} and $L_y=$ \SI{700}{\km}, a grid resolution of \num{256x64} elements with 50 markers
-per element and Courant number of 0.01. At the beginning of the simulation, a 100 km-thick lithospheric layer with $\rho_m=$ \SI{3300}{\kg\per\cubic\metre}
-and $\eta_m=$ \SI{e23}{\pascal\s} is located at the top of the domain for $x$ between 1000 and 3000 km, while the asthenospheric mantle has $\rho_m=$
-\SI{3200}{\pascal\s} and $\eta_m=$ \SI{e21}{\pascal\s}. In addition, a 200 km-depth lithospheric slab is already subducted in the mantle in order to have
-a spontaneous subduction. In case of sticky air, $L_y=750$ km and a 50 km-thick air layer with $\eta_a$ from \SIrange{e19}{e21}{\pascal\s} overlie the mantle.
-Velocity boundary conditions are set to free slip on the sides and on the bottom of the domain. In case of sticky air, velocities on the top boundary are set
-to free slip conditions as well.
-
-The maximum depth of the slab is tracked for all the simulations (Fig. \ref{fig:subduction}) and they are compared with results obtained by
-\citet{Schmeling2008} with models with regular grids and comparable grid resolutions (rectangles in Fig. \ref{fig:subduction}). Low viscous sticky air models
-(i.e., \SI{e19}{\pascal\s}) enlighten a strong dependence of the sinking velocity with the chosen average scheme (dashed lines in Fig. \ref{fig:subduction}),
-while high viscous sticky air models (i.e., \SI{e21}{\pascal\s}) show a higher resistance at the trench, underestimating the correct solution (dotted lines in
-Fig. \ref{fig:subduction}). A high resistance at the trench can be also observed in case of a true free surface for the low resolution grid used (continuous
-lines in Fig. \ref{fig:subduction}), as pointed out by \citet{Schmeling2008}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Spontaneous_subduction}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Subduction.pdf}
-\caption{Maximum depth of the slab as function of time for all the simulations of the \citet{Schmeling2008} benchmark, in case of harmonic, geometric and
-arithmetic means (red, green and blue lines, respectively) using a sticky air or a true free surface (discontinuous and continuous lines, respectively).
-Red, green and blue rectangular areas indicate the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{500}{\km} in case of sticky air,
-while the grey rectangular area indicates the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{600}{\km} in case of true free surface.}
-\label{fig:subduction}
-\end{figure}
-
-\subsection{The slab detachment}\label{sec:slab}
-The slab detachment benchmark is performed as described by \citet{Schmalholz2011} and \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{1000}{\km}
-and $L_y=$ \SI{660}{\km} and a grid resolution of \num{256x256} elements. A non-linear viscous T-shaped layer with $\rho_l=$ \SI{3300}{\kg\per\cubic\metre} is
-placed at the top of the domain and surrounded by a linear viscous fluid with $\rho_f=$ \SI{3150}{\kg\per\cubic\m}. The top layer is 80 km-thick and a 250
-km-long and 80 km-wide slab is placed at $x=L_x/2$ (Fig. \ref{fig:slab}a). The effective viscosity of the top layer is given by
-\[\eta_{eff}=\eta_0 I_2^{\frac{(1-n)}{n}}\]
-with $\eta_0=$ \SI{4.75e11}{\pascal\s} and $n=4$, while the surrounding fluid has $\eta_f=$ \SI{e21}{\pascal\s}. The effective viscosity is capped between
-$\eta_{min}=$ \SI{e21}{\pascal\s} and $\eta_{max}=$ \SI{e25}{\pascal\s}. Velocity boundary conditions are set to free slip at the top and the bottom, and to
-no slip at the sides of the domain.
-
-The sinking of the slab determines a decrease of the effective viscosity in the T-shaped layer during the first part of the evolution (Fig. \ref{fig:slab}a),
-as consequence of local strain rates, while effective viscosities increase up to $\eta_{max}$ in the top layer after approximately \SI{20}{\mega\year}, when
-necking is complete (Fig. \ref{fig:slab}b).
-
-The width of necking is tracked by means of markers position at the side of the slab. Necking width and time are normalised in function of the initial slab
-(\SI{80}{\km}) and the characteristic time ($t_c=$ \SI{7.1158e14}{\s}) \citep{Schmalholz2011,Glerum2018}. The results in case of arithmetic, geometric and
-harmonic mean of the viscosity are shown in Fig. \ref{fig:necking}, compared with results from \citet{Glerum2018} for which an infinity norm mean was used.
-The influence of the average schemes agrees with results shown in Section \ref{sec:subduction}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Slab_detachment}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Slab.pdf}
-\caption{Effective viscosity for the slab detachment benchmark at the beginning of the evolution (panel a) and when necking is complete (panel b).}
-\label{fig:slab}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Necking.pdf}
-\caption{Normalised width of the necking of the slab detachment benchmark as function of normalised time.}
-\label{fig:necking}
-\end{figure}
-
-\subsection{The indenter experiment}\label{sec:indenter}
-The indenter benchmark simulate a rigid punch on a purely plastic von Mises material, for which exists an analytical solution
-\citep{Thieulot2008,Thieulot2014,Glerum2018}. The domain is a unit square with a grid resolution of \num{256x256} elements. The gravity acceleration is set
-to $\bm{g}=\bm{0}$ and effective viscosity is capped using $\eta_{min}=10^{-4}$ and $\eta_{max}=10^3$. The tolerance for the convergence of the non-linear
-solution is $tol=10^{-9}$, with a maximum number of non-linear iterations set to 500. The medium in the domain has $\rho=0.01$, an initial viscosity
-$\eta_0=10$, cohesion $C=1$ and an angle of internal friction $\phi=0$°. Velocity boundary conditions are set to no slip at the bottom and free slip at the
-sides of the domain. The top of the domain is open with the exception of the central portion (punch area) with width $w_p=0.125$, where $v=-1.05$ and $u$ is
-fixed either to 0 (rough punch experiment) or free (smooth punch experiment).
-
-Fig. \ref{fig:indenter} show the results in terms of viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure
-(panels g and h) for both experiments. The analytical solution indicate that pressure at ($0.5;1$) and ($0.5 \pm w_p;1$)is $p=\pi +1$ and $p=1$, respectively
-(black and red lines in Fig. \ref{fig:smooth_p}). Fig. \ref{fig:smooth_p} shows a clear improvement in the pressure solution when passing from a rough to a
-smooth punch (green and blue lines in Fig. \ref{fig:smooth_p}, respectively), as previously observed by \citet{Thieulot2014} and \citet{Glerum2018}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Indenter}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Indenter.pdf}
-\caption{Viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure (panels g and h) fields for rough (left column) and
-smooth (right column) punch experiments.}
-\label{fig:indenter}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Smooth_Rough.pdf}
-\caption{Nodal pressure in function of the x coordinate at $y=1$ for rough (green line) and smooth (blue line) punch experiment. Black and red lines indicate
-the analytical solution at $x=0.5$ and $x=0.5 \pm w_p$, respectively.}
-\label{fig:smooth_p}
-\end{figure}
-
-\subsection{The brick experiment}\label{sec:brick}
-An instantaneous version of the brick benchmark is performed to verify the correctness of the pressure-dependent plasticity for different angle of internal
-friction, as proposed by \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{40}{\km} and $L_y=$ \SI{10}{\km} and a grid resolution of \num{512x128}
-elements. The gravity acceleration is set to $g_y=$ \SI{-10}{\m\square\s} and effective viscosity is capped using $\eta_{min}=$ \SI{e19}{\pascal\s} and
-$\eta_{max}=$ \SI{e26}{\pascal\s}. The tolerance for the convergence of the non-linear solution is $tol=10^{-7}$, with a maximum number of non-linear iterations
-set to 1000. A 800 m-wide and 400 m-height inclusion with $\eta_b=$ \SI{e20}{\pascal\s} and $\rho_b=$ \SI{2700}{\kg\per\cubic\m} is placed at the bottom of the
-domain at $x=L_x/2$. The inclusion is surrounded by a non-linear viscous medium with $\rho_m=$ \SI{2700}{\kg\per\cubic\m}, an initial viscosity of
-\SI{e23}{\pascal\s}, a linear viscous viscosity of \SI{e25}{\pascal\s} and a cohesion of \SI{40}{\mega\pascal}. Velocities are set to free slip conditions at
-the bottom of the domain and the top is open. Velocities on sides of the domain are fixed to $u=$ \SI{+-2e-11}{\m\per\s} and $v=0$. The experiment is performed
-in compressional and extensional contexts with an angle of internal friction $\phi$ of the non-linear medium variable between 0° and 30°.
-
-As expected, two shear bands stem from the inclusion with variable angles in relation to both the dynamics context and the internal friction angle. Shear band
-angles formed at 45° for $\phi=0$° in both extensional and compressional contexts (Fig. \ref{fig:brick_beam}a), while they are different in case of $\phi\neq0$°.
-Shear band angles in case of $\phi=20$° are shown for compression and extension in Fig. \ref{fig:brick_beam}b  and c, respectively. Values of shear band angles
-as function of internal friction angles from 0° to 30° are extracted at different depths and minimum and maximum values are plotted in Fig. \ref{fig:brick},
-in comparison with theoretical Roscoe, Arthur and Coulomb shear band angles. For all tests the tolerance for nonlinear convergence is defined $tol=10^{-7}$
-and none converges before the maximum number of iterations ($it_{max}=1000$) is reached. However, tests with internal friction angles up to 15° show a constant
-decrease of the velocity residuals, which can not be observed for for tests with higher internal friction angles (Fig. \ref{fig:convergence}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Brick_experiment}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Brick_Beam.pdf}
-\caption{Shear band angles predicted for the brick experiment for $\phi=0$° (panel a) and for $\phi=20$° in case of compression and extension (panels b and c,
-respectively).}
-\label{fig:brick_beam}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Brick.pdf}
-\caption{Shear band angles predicted for the brick experiment in case of compressional and extensional contexts as function of different internal angle of
-friction (continuous black line and red dots), compared with theoretical Roscoe, Artur and Coulomb angles (discontinuous black lines). Red lines indicate the
-range of angles calculated at different depths.}
-\label{fig:brick}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \noindent\includegraphics[width=500px]{./Figures/convergence.pdf}
-  \caption{Convergence rates for all tests. Normalised residuals (left column) are compared with velocity residuals (right column) for extensional (first row)
-  and compressional (second row) settings.}
-  \label{fig:convergence}
-  \end{figure}
-
-\subsection{Advection stabilisation}\label{sec:advection}
-The 1D advection problem proposed by \citet{Donea2003} and \citet{Thieulot2011} is performed to verify the effectiveness of the SUPG method to stabilise the
-advection term of the energy equation. The domain is a 1D segment with $L_x=1$ composed by 50 elements and a discontinuity in the thermal field placed at
-$x=0.25$. Temperature is set to 1 for $x \leq 0.25$ and to 0 for $x > 0.25$. Velocity is set to ${u}=1$ in the entire domain. The simulation is performed for
-250 time steps, with $dt=0.002$, so the thermal discontinuity should be at $x=0.75$ at the end of the simulation.
-
-Temperature profiles at the end of the simulation are shown in \ref{fig:advection} as function of the dimensionless coefficient $\gamma=\tau \bm{v}/h$. In case
-of the classic Galerkin method ($\gamma=0$, blue line in \ref{fig:advection}) the final thermal profile is characterised by strong oscillations, which are
-eliminated in case of the SUPG method ($\gamma=0.045$, orange line in \ref{fig:advection}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Advection_stabilisation}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Advection.pdf}
-\caption{Temperature profile as function of \textit{x} for the advection stabilisation benchmark. Purple line indicates the initial temperature profile; the
-green line indicates the analytical temperature profile after 250 time steps; blue line indicates the temperature profile after 250 time steps in case of the
-classic Galerkin method ($\gamma=0$); orange line indicates the temperature profile after 250 time steps in case of the SUPG method ($\gamma=0.045$).}
-\label{fig:advection}
-\end{figure}
-
 \subsection{Shear heating}\label{sec:simple_shear}
-This exercise is performed in an unit square domain composed by \num{128x128} elements. The velocity field is prescribed on the entire domain with
+This exercise is performed in an unit square domain composed by $128\times128$ elements. The velocity field is prescribed on the entire domain with
 $\bm{v}=(L_y-y)y\bm{e_x}$; viscosity, density and specific heat are fixed to 1, while thermal conductivity and radiogenic energy are fixed to 0. Therefore,
@@
 \end{figure}
-
-\subsection{Shear and adiabatic heating}\label{sec:shear}
-This problem is performed as presented in Exercise 9.4 in \citet{Gerya2010b}. Two materials are vertically separated in a rectangular domain with $L_x=$
-\SI{1000}{\km}, $L_y=$ \SI{1500}{\km} and a grid resolution of \num{30x20} elements. Constant thermal coefficient expansion ($\alpha=$ \SI{3e-5}{\per\kelvin}),
-temperature ($T=$ \SI{1300}{\kelvin}) and gravity acceleration ($g_y=$ \SI{-10}{\m\square\s}) are assumed in the whole domain. Fluid 1 (on the left side of the
-domain) has $\rho_1=$ \SI{3200}{\kg\per\cubic\m} and $\eta_1=$ \SI{e20}{\pascal\s}; fluid 2 (on the right side of the domain) has $\rho_2=$
-\SI{3300}{\kg\per\cubic\m} and $\eta_2=$ \SI{e22}{\pascal\s}. Velocity boundary conditions are set to free slip on all sides of the domain.
-
-As shown in Fig. \ref{fig:energy}, both shear and adiabatic heating predicted by the code (first row in Fig. \ref{fig:energy}) well recreate the results
-obtained by means of example ${Shear\_adiabatic\_heating.m}$ from \citet{Gerya2010b} (second row in Fig. \ref{fig:energy}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Adiabatic%2BShear}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Energy.pdf}
-\caption{Comparison between shear (first column), adiabatic (second column) and total (third column) energy predicted by the code (panels a, b and c) and those
-created using example ${Shear\_adiabatic\_heating.m}$ from Exercise 9.4 in \citet{Gerya2010b} (panels d, e and f).}
-\label{fig:energy}
-\end{figure}
-
-\subsection{Mantle convection}\label{sec:mantle}
-This problem is performed as presented by \citet{Blankenbach1989} (constant viscosity cases) and \citet{Thieulot2014}, in a 2D unit square domain with gravity
-acceleration $g_y=10^{10}Ra$. The experiment is performed with three different Rayleigh numbers ($Ra=10^4$, $10^5$ and $10^6$) and with different grid
-resolution (between \num{32x32} and \num{128x128} elements). The fluid has constant viscosity, initial density, heat capacity, thermal conductivity
-($\eta= \rho_0= C_p = k=1$), reference temperature ($T_0=0$) and thermal expansion coefficient ($\alpha=10^{-10}$). Temperatures are set to 0 on top and 1 on
-bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\]
-
-The solution generated by the code in terms of $v_{\textrm{rms}}$ (calculated as in Sec. \ref{sec:error}) and Nu as function of time are reported for all the
-simulations in Fig. \ref{fig:mantle}. At the steady state, $v_{\textrm{rms}}$ and Nu of all simulations converge well toward the values from \citet{Blankenbach1989},
-with lower errors for higher resolution grids (Table \ref{tab:mantle}). All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Mantle_convection}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Convection.pdf}
-\caption{$v_{\textrm{rms}}$ (panels a, c and e) and Nu (panels b, d and f) for the mantle convection benchmark as function of time for different grid resolution.
-Panels a and b show the results for $Ra=10^4$; Panels c and d show the results for $Ra=10^5$; panels e and f show the results for $Ra=10^6$. Purple lines
-indicate the convergence values for the $v_{\textrm{rms}}$ and the Nu from \citet{Blankenbach1989} at the steady state.}
-\label{fig:mantle}
-\end{figure}
-
-\begin{table}[H]
-\caption{Comparison between $v_{\textrm{rms}}$ and Nu predicted by the code for the mantle convection experiment and same values as reported in literature.}
-\centering
-\small
-\begin{tabular}{|>{\centering}m{0.09\textwidth}|>{\centering}m{0.05\textwidth}|>{\centering}m{0.24\textwidth}|>{\centering}m{0.15\textwidth}|
-  >{\centering}m{0.1\textwidth} >{\centering}m{0.1\textwidth} >{\centering\arraybackslash}m{0.1\textwidth}|}
-\toprule
-& & \multirow{2}{*}{\citet{Blankenbach1989}} & \citet{Thieulot2014} & \multicolumn{3}{c|}{FALCON}  \\
- & & & \num{200x200} & \num{32x32} & \num{64x64} & \num{128x128}  \\
-\midrule
-\multirow{2}{*}{$Ra=10^4$} & $v_{\textrm{rms}}$ & $42.864947 \pm 0.000020$ & 42.867 & 42.83226 & 42.852793 & 42.861394  \\
-           & Nu      & $4.884409 \pm 0.000010$  & 4.882  & 4.781297 & 4.857475  & 4.877573   \\
-\hline
-\multirow{2}{*}{$Ra=10^5$} & $v_{\textrm{rms}}$ & $193.21454 \pm 0.00010$  & 193.255& 193.872643 & 193.377472 & 193.252290  \\
-           & Nu      & $10.534095 \pm 0.000010$ & 10.507 & 9.602514 & 10.270735  & 10.465629   \\
-\hline
-\multirow{2}{*}{$Ra=10^6$} & $v_{\textrm{rms}}$ & $833.98977 \pm 0.00020$  & 834.712& 848.091176 & 837.767911 & 834.945793  \\
-           & Nu      & $21.972465 \pm 0.000020$ & 21.695 & 15.999266 & 19.703682  & 21.306939   \\
-\bottomrule
-\end{tabular}
-\label{tab:mantle}
-\end{table}
-
-\subsection{Viscoplastic mantle convection}\label{sec:visco_mantle}
-This benchmark is performed as Cases 1-5a described in \citet{Tosi2015} in a 2D unit square domain with two different grid resolutions of \num{32x32} 
-and \num{100x100} elements. Reference viscosity, reference density, heat capacity, thermal conductivity and thermal expansion coefficient are fixed to a constant
-value of 1 ($\eta_0= \rho_0= C_p = k=\alpha=1$), while the gravity has been chosen as $g_y=10^2$, in order to obtain a Rayleigh number $Ra=10^2$. Temperatures
-are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\].
-The viscosity field $\eta$ is calculated as the harmonic average between a linear part $\eta_{lin}$ that depends on temperature only or on temperature and 
-depth $y$ and a nonlinear-plastic part $\eta_{pl}$ that depends on the strain rate $\dot{\bm{\epsilon}}$, as follows
-\[\eta(T,y,\dot{\bm{\epsilon}})=2\left(\frac{1}{\eta_{lin}(T,y)}+\frac{1}{\eta_{pl}(\dot{\bm{\epsilon}})}\right)^{-1}\]
-The linear and the nonlinear-plastic parts of the viscosity are calculated following \citet{Tosi2015} as follows
-\[\eta_{lin}(T,y)=\exp(-\ln(\Delta\eta_T)T+\ln(\Delta\eta_y)y)\]
-\[\eta_{pl}(\dot{\bm{\epsilon}})=\eta^*+\frac{\sigma_Y}{\sqrt{\dot{\bm{\epsilon}}:\dot{\bm{\epsilon}}}}\]
-where $\Delta\eta_T$, $\Delta\eta_y$, $\eta^*$ and $\sigma_Y$ are parameters chosen as listed in Table \ref{tab:tosi_setup}.
-
-For each case, some diagnostic quantities measured at the steady state are reported in Table \ref{tab:visco_mantle} (see Section \ref{sec:error} for 
-the computation of all the quantities), showing a well fit with results described in \citet{Tosi2015}. Data and charts for all cases can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Viscoplastic_mantle_convection}.
-
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{white!100!}{gray!25!}
-%\begin{table}[H]
-%  \caption{Results of the viscoplastic mantle convection benchmark for Cases 1-4 compared with results obtained with ELEFANT and YACC (from \citealp{Tosi2015}).}
-%  \centering
-%  \resizebox{\textwidth}{!}{
-%\begin{tabular}{|c|ccc|ccc|ccc|ccc|}
-%  \toprule
-%    & \multicolumn{3}{c|}{\textbf{Case 1}} & \multicolumn{3}{c|}{\textbf{Case 2}} & \multicolumn{3}{c|}{\textbf{Case 3}} &
-%    \multicolumn{3}{c|}{\textbf{Case 4}} \\
-%   & FALCON & ELEFANT& YACC & FALCON& ELEFANT& YACC & FALCON & ELEFANT & YACC & FALCON & ELEFANT & YACC \\
-%  \midrule
-%   $\left\langle T \right\rangle$ & 0.7759 & 0.7758 & 0.7767 & 0.6043 & 0.6036 & 0.6058 & 0.7275 & 0.7275 & 0.7286 & & 0.5277 & 0.5289\\
-%    Nu$^{top}$ & 3.4280 & 3.4214 & 3.4298 & 8.5242 & 8.5115 & 8.5278 & 3.0368 & 3.0347 & 3.0374 && 6.5912 & 6.5572 \\
-%    Nu$^{bottom}$ & 3.3168 & 3.313 & 3.3143 & 8.4599 & 8.4465 & 8.3990 & 3.0086 & 2.9908 & 2.9628 && 6.5834 & 6.5243 \\
-%    $v_{\textrm{rms}}$ & 249.901 & 249.134 & 251.7997 & 141.2451 & 140.7087 & 142.2020 & 100.2776 & 100.1208 & 100.9467 && 79.1105 & 79.6202 \\
-%    $v_{\textrm{rms}}^{top}$ & 1.8907 & 1.8642 & 1.8298 & 104.7928 & 104.1228 & 105.1767 & 2.0832 & 2.0652 & 2.0374 && 74.7596 & 75.8414 \\
-%    $u_{max}^{top}$ & 2.6340 & 2.6119 & 2.5516 & 121.9212& 121.766 & 122.3238 & 2.9086 & 2.9019 & 2.8458 && 88.9146 & 89.2940 \\
-%    $\eta_{min}$ & - & - & - & $2.0303\times10^{-5}$ & $2.0600\times10^{-5}$ & $1.9853\times10^{-5}$ & $4.8129\times10^{-5}$ & $4.8080\times10^{-5}$ & $4.7907\times10^{-5}$ &
-%    & $1.9860\times10^{-4}$ & $1.9174\times10^{-4}$ \\
-%    $\eta_{max}$ & - & - & - & 1.3325 & 1.85 & 1.6095 & 0.9360 & 0.9023 & 1 && 1.52 & 1.6773 \\
-%    $\left\langle W \right\rangle$ & 2.4299 & 2.4316 & 2.4583 & 7.5648 & 7.5665 & 7.6222 & 2.0387 & 2.0384 & 2.0400 && 5.6216 & 5.6512 \\
-%    $\left\langle \Phi \right\rangle/Ra$ & 2.4311 & 2.4276 & 2.4333 & 7.5661 & 7.5585 & 7.6194 & 2.0386 & 2.0370 & 2.0335 && 5.6182 & 5.6463 \\
-%    $\delta$ & 0.05\% & 0.16\% & 1.02\% & 0.02\% & 0.11\% & 0.04\% & $<$0.01\% & 0.07\% & 0.32\% && 0.06\% & 0.09\% \\
-%  \bottomrule
-%  \end{tabular}}
-%  \label{tab:visco_mantle}
-%  \end{table}
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{2}{white!100!}{gray!25!}
-\begin{table}[H]
-  \caption{Benchmark cases and corresponding parameters.}
-  \centering
-  \small
-\begin{tabular}{|>{\centering}m{0.08\textwidth}|>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}
-  >{\centering}m{0.05\textwidth} >{\centering}m{0.05\textwidth} |>{\centering\arraybackslash}m{0.2\textwidth}|}
-  \toprule
-    Case & $Ra$ & $\Delta\eta_T$ & $\Delta\eta_y$ & $\eta^*$ & $\sigma_Y$ & Convective regime \\
-  \midrule
-  \rowcolor{gray!25!}
-    1 & $10^2$ & $10^5$ & 1 & - & - & Stagnant lid \\
-    2 & $10^2$ & $10^5$ & 1 & $10^{-3}$ & 1 & Mobile lid  \\
-    \rowcolor{gray!25!}
-    3 & $10^2$ & $10^5$ & 10 & - & - & Stagnant lid  \\
-    4 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 1 & Mobile lid \\
-    \rowcolor{gray!25!}
-    5 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 4 & Periodic \\
-  \bottomrule
-  \end{tabular}
-  \label{tab:tosi_setup}
-  \end{table}
-
-\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{gray!25!}{white!100!}
-\begin{table}
-  \caption{Results of the viscoplastic mantle convection benchmark for different grid resolutions in Cases 1-5 compared with results obtained with ELEFANT
-  and YACC ($100\times100$ elements, from \citealp{Tosi2015}) and in Stone 28 ($32\times32$ elements, from
-  \url{https://github.com/cedrict/fieldstone/tree/master/python_codes/fieldstone_28}).}
-  \centering
-  \resizebox{\textwidth}{!}{
-\begin{tabular}{|cc|ccccccccccc|}
-  \toprule
-  \multicolumn{13}{|c|}{\textbf{Case 1}} \\
-  \midrule
-  Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-  & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-  \hline
-  \rowcolor{gray!25!}
-    YACC & $100\times100$ & 0.7767 & 3.4298 & 3.3143 & 251.7997 & 1.8298 & 2.5516 & - & - & 2.4583 & 2.4333 & 1.02\% \\
-    ELEFANT & $100\times100$ & 0.7758 & 3.4214 & 3.313 & 249.134 & 1.8642 & 2.6119 & - & - & 2.4316 & 2.4276 & 0.16\% \\
-    \rowcolor{gray!25!}
-    Stone 28 & $32\times32$ & 0.7737 & 3.3987 & - & 243.872 & - & - & - & - & - & - & - \\
-%    FALCON & $32\times32$ & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & - & - & 2.4745 & 2.4864 & 0.48\% \\
-%    \rowcolor{gray!25!}
-    FALCON & $100\times100$ & 0.7759 & 3.4280 & 3.3168 & 249.901 & 1.8907 & 2.6340 & - & - & 2.4299 & 2.4311 & 0.05\% \\
-    \midrule
-    \multicolumn{13}{|c|}{\textbf{Case 2}} \\
-    \midrule
-    Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-    & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-    \hline
-    \rowcolor{gray!25!}
-     YACC & $100\times100$ & 0.6058 & 8.5278 & 8.3990 & 142.2020 & 105.1767 & 122.3238 & $1.9853\times10^{-5}$ & 1.6095 & 7.6222 & 7.6194 & 0.04\% \\
-     ELEFANT & $100\times100$ & 0.6034 & 8.5115 & 8.4465 & 140.8067 & 104.1228 & 121.766 & $2.0600\times10^{-5}$ & 1.85 & 7.5665 & 7.5585 & 0.11\% \\
-     \rowcolor{gray!25!}
-     Stone 28 & $32\times32$ & 0.6052 & 8.1715 & - & 141.518 & - & - & - & - & - & - & - \\
-%     FALCON & $32\times32$ & 0.6454 & 8.1422 & 7.4020 & 156.6908 & 95.7226 & 116.9302 & $2.2572\times10^{-5}$ & 0.7194 & 7.4289 & 7.4452 & 0.22\% \\
-%     \rowcolor{gray!25!}
-     FALCON & $100\times100$ & 0.6044 & 8.5266 & 8.4590 & 141.3684 & 104.8867 & 122.0470 & $2.0333\times10^{-5}$ & 1.2914 & 7.5707 & 7.5718 & 0.01\% \\
-     \midrule
-     \multicolumn{13}{|c|}{\textbf{Case 3}} \\
-     \midrule
-     Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-     & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-     \hline
-     \rowcolor{gray!25!} 
-     YACC & $100\times100$ & 0.7286 & 3.0374 & 2.9628 & 100.9467 & 2.0374 & 2.8458 & $4.7907\times10^{-5}$ & 1 & 2.0400 & 2.0335 & 0.32\% \\
-      ELEFANT & $100\times100$ & 0.7275 & 3.0347 & 2.9908 & 100.1208 & 2.0652 & 2.9019 & $4.8080\times10^{-5}$ & 0.9023 & 2.0384 & 2.037 & 0.07\% \\
-      \rowcolor{gray!25!}
-      Stone 28 & $32\times32$ & 0.7277 & 3.0347 & - & 100.018 & - & - & - & - & - & - & - \\
-%      FALCON & $32\times32$ & 0.7281 & 3.0618 & 2.8132 & 101.0092 & 2.1630 & 3.0178 & $4.8500\times10^{-5}$ & 0.7845 & 2.0699 & 2.0695 & 0.02\% \\
-%      \rowcolor{gray!25!}
-      FALCON & $100\times100$ & 0.7275 & 3.0368 & 3.0086 & 100.2776 & 2.0832 & 2.9086 & $4.8129\times10^{-5}$ & 0.9360 & 2.0387 & 2.0386 & $<$0.01\% \\
-      \midrule
-      \multicolumn{13}{|c|}{\textbf{Case 4}} \\
-      \midrule
-      Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-      & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-      \hline
-      \rowcolor{gray!25!} 
-      YACC & $100\times100$ & 0.5289 & 6.5572 & 6.5243 & 79.6202 & 75.4814 & 89.2940 & $1.9174\times10^{-4}$ & 1.6773 & 5.6512 & 5.6463 & 0.09\% \\
-       ELEFANT & $100\times100$ & 0.5277 & 6.5912 & 6.5834 & 79.1105 & 74.7596 & 88.9146 & $1.9860\times10^{-4}$ & 1.5200 & 5.6216 & 5.6182 & 0.06\% \\
-       \rowcolor{gray!25!}
-       Stone 28 & $32\times32$ & 0.5286 & 6.4036 & - & 79.4746 & - & - & - & - & - & - & - \\
-%       FALCON & $32\times32$ & 0.5334 & 6.4679 & 6.3435 & 80.6153 & 76.0511 & 90.2058 & $2.0393\times10^{-4}$ & 0.6453 & 5.6676 & 5.6698 & 0.04\% \\
-%       \rowcolor{gray!25!}
-       FALCON & $100\times100$ & 0.5279 & 6.5897 & 6.5830 & 79.1220 & 75.0715 & 88.8484 & $1.9693\times10^{-4}$ & 1.4133 & 5.6127 & 5.6113 & 0.03\% \\
-       \midrule
-       \multicolumn{13}{|c|}{\textbf{Case 5}} \\
-       \midrule
-       Code & Resolution & $P$ & $\left\langle T \right\rangle_{min}$ & $\left\langle T \right\rangle_{max}$ & Nu$^{top}_{min}$ & Nu$^{top}_{max}$ &
-       $v_{\textrm{rms}}^{min}$ & $v_{\textrm{rms}}^{max}$ & \multicolumn{2}{c}{$(\left\langle \Phi \right\rangle/Ra)_{min}$} & 
-       \multicolumn{2}{c|}{$(\left\langle \Phi \right\rangle/Ra)_{max}$} \\
-       \hline
-       \rowcolor{gray!25!}
-         YACC & $100\times100$ & 0.0792 & 0.6543 & 0.6722 & 2.6950 & 7.2792 & 41.9813 & 98.8230 & \multicolumn{2}{c}{1.4347} & 
-         \multicolumn{2}{c|}{9.2748} \\
-         ELEFANT & $100\times100$ & 0.076 & 0.6514 & 0.6693 & 2.6808 & 7.3506 & 41.2578 & 100.089 & \multicolumn{2}{c}{1.3324} & 
-         \multicolumn{2}{c|}{9.5289} \\
-         \rowcolor{gray!25!}
-         Stone 28 & $32\times32$ & 0.0764 & 0.6521 & 0.6697 & 2.6912 & 7.0763 & 41.61 & 98.183 & \multicolumn{2}{c}{-} & 
-         \multicolumn{2}{c|}{-} \\
-     %    FALCON & $32\times32$ & & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & \multicolumn{2}{c}{-} & 
-     %     \multicolumn{2}{c|}{-} \\
-     %    \rowcolor{gray!25!}
-         FALCON & $100\times100$ & 0.0775 & 0.6519 & 0.6697 & 2.6847 & 7.3629 & 41.5256 & 99.5466 & \multicolumn{2}{c}{1.3399} & 
-         \multicolumn{2}{c|}{9.3993} \\
-  \bottomrule
-  \end{tabular}}
-  \label{tab:visco_mantle}
-  \end{table}
-
-\subsection{Thin layer entrainment}\label{sec:thin}
-This experiment is performed as originally proposed by \citet{vanKeken1997} with two incompressible fluids in a rectangular domain with $L_x=2$ and $L_y=1$
-and gravity acceleration $g_y=10^{10}Ra$. The Rayleigh number and the compositional Rayleigh number are fixed to $Ra=300000$ and $Rc=450000$, respectively.
-Both fluids have constant viscosity, thermal conductivity, specific heat ($\eta=\rho=C_p=1$) and thermal expansion coefficient ($\alpha=10^{-10}$). Fluid 1
-has a density $\rho_1=1$, while fluid 2 is denser ($\rho_2=\rho_1+1.5 \cdot 10^{-10}$) and is placed at the bottom of the domain, for $y \leq 0.025$.
-Temperature are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides of the domain. The initial
-temperature field is given by
-\[T(x,y)=T_u(x,y)+T_l(x,y)+T_r(x,y)+T_s(x,y)-\frac{3}{2}\]
-with
-\begin{eqnarray}
-T_u(x,y)&=&\frac{1}{2}erf\left(\frac{1-y}{2}\sqrt{\frac{u_0}{x}}\right)\nonumber \\
-T_l(x,y)&=&1-\frac{1}{2}erf\left(\frac{y}{2}\sqrt{\frac{u_0}{L_x-x}}\right)\nonumber \\
-T_r(x,y)&=&\frac{1}{2}+\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{y+1}} \exp\left(-\frac{x^2u_0}{4y+4}\right)\nonumber \\
-T_s(x,y)&=&\frac{1}{2}-\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{2-y}} \exp\left(-\frac{(L_x-x)^2u_0}{8-4y}\right)\nonumber
-\end{eqnarray}
-and
-\begin{eqnarray}
-u_0&=&\frac{L_x^{7/3}}{(1+L_x^4)^{2/3}}\left(\frac{Ra}{2\sqrt{\pi}}\right)^{\frac{2}{3}}\nonumber \\
-Q&=&2\sqrt{\frac{L_x}{\pi u_0}}\nonumber
-\end{eqnarray}
-
-The experiment is performed with two grid resolution (\num{125x40} and \num{200x80} elements), with 100 markers per element and CFL of 0.25. $v_{\textrm{rms}}$
-(see Sec. \ref{sec:error}) calculated for both the simulations match well with results from \citet{vanKeken1997} and obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and ELEFANT \citep{Thieulot2014}, obtained with similar grid resolutions
-(Fig. \ref{fig:thin}). In addition, a variety of simulations with different resolutions and aspect ratios are performed to verify that the sensitivity of the
-initial velocity field. Results match well with results from \citet{vanKeken1997} (in the grey area) and \citet{Thieulot2014} (Fig. \ref{fig:thin_initial}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Thin_layer}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_thin.pdf}
-\caption{$v_{\textrm{rms}}$ for the thin layer experiment as function of time for different grid resolution. Results are compared with results obtained by
-\citet{vanKeken1997} (black lines) and with ASPECT and ELEFANT.}
-\label{fig:thin}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrmszero.pdf}
-\caption{$v_{\textrm{rms}}$ at $t=0$ for the thin layer experiment as function of the element size for different aspect ratios of the elements. Results are compared
-with results obtained by \citet{vanKeken1997} (black dots in the grey area) and with ASPECT and ELEFANT.}
-\label{fig:thin_initial}
-\end{figure}
-
-\subsection{Sinking hydrated cylinder}\label{sec:quinquis}
-This experiment is performed as explained in \citet{Quinquis2014} and simulates the sinking of a cold, hydrated cylinder into a warm, dry mantle. The domain
-is square with $L_x=L_y=$ \SI{300}{\km}, a grid resolution of \num{300x300} elements and 25 markers per element. The sinking cylinder is located at $x=$
-\SI{150}{\km} and $y=$ \SI{170}{\km} with a radius of \SI{20}{\km} and has $\rho_c=$ \SI{3250}{\kg\per\cubic\m}, $\eta_c=$ \SI{e23}{\pascal\s}, a thermal
-conductivity $k_c=$ \SI{4.5}{\watt\per\m\per\kelvin},a specific heat $C_{p_c}=$ \SI{1250}{\joule\per\kg\per\kelvin}; the surrounding mantle has $\rho_m=$
-\SI{3200}{\kg\per\cubic\m}, $\eta_m=$ \SI{e20}{\pascal\s}, a thermal conductivity $k_m=$ \SI{105}{\watt\per\m\per\kelvin} and specific heat $C_{p_m}=$
-\SI{1250}{\joule\per\kg\per\kelvin}; a 58 km-thick lithosphere overlies the mantle and has $\rho_l=$ \SI{3200}{\kg\per\cubic\m}, $\eta_l=$ \SI{e23}{\pascal\s},
-a thermal conductivity $k_l=$ \SI{4.5}{\watt\per\m\per\kelvin}, specific heat $C_{p_l}=$ \SI{750}{\joule\per\kg\per\kelvin} and a thermal diffusivity of
-\SI{e-6}{\square\m\per\s}. Velocity boundary conditions are set to free slip on all sides of the domain. Temperature increase linearly in the lithosphere,
-from 0° to 1300°C, and in the mantle, from 1300° to 1360.5°C. The initial temperature of the cylinder is fixed to 400°C. Throughout the evolution, temperature
-is fixed to 0° and 1360.5°C at the top and bottom of the domain, respectively. The initial bound water content is imposed at 2 wt.\% in the cylinder and at
-0 wt.\% in both the mantle and the lithosphere. The maximum amount of water in the mantle is fixed to 0.2 wt.\%, while maximum water content in the cylinder
-and in the lithosphere is function of pressure and temperature and is calculated using a serpentinized harzburgite with Perple\_X, as in \citet{Quinquis2014}.
-
-As shown in \citet{Quinquis2014}, the model is characterised by a progressive dehydration in the external portion of the cylinder, due to the increase of
-temperature, with a consequent vertical migration of free water that hydrates the mantle above the cylinder, up to the lithosphere. The distribution of
-bound and free water during the evolution of the experiment is shown in Fig. \ref{fig:quinquis} (left and right columns, respectively). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Hydration}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Hydration.pdf}
-\caption{Distribution of bound and free water (left and right columns, respectively) at $t=0.5$ and $t=$ \SI{1.25}{\mega\year} (first and second rows,
-respectively) for the sinking hydrated cylinder experiment. Only markers with bound or free water not equal to 0 are plotted.}
-\label{fig:quinquis}
-\end{figure}
-
-\subsection{Mantle melting curves}\label{sec:katz}
-This instantaneous experiment is performed in order to recreate isobaric melting curves, as function of temperature, bulk water and content of water in the
-melt, to compare with melting curves obtained by \citet{Katz2003}. The domain is square with $L_x=L_y=100$ km, a grid resolution of \num{100x100} elements and
-25 markers per element. Density is set in the entire domain to \SI{3300}{\kg\per\cubic\m} and temperature increases from the sides to the centre from 1000° to
-1500°C. The experiment is performed for bulk water contents from 0 to 0.3 wt.\%.
-
-Fig. \ref{fig:fraction} shows isobaric melting curves for \SI{1}{\giga\pascal} (panel a) and \SI{3}{\giga\pascal} (panel b) calculated for different bulk water
-contents. The curves well match with those obtained for the mantle by \citet{Katz2003}, showing the saturation at the solidus for bulk water of 0.2 and 0.3 wt.\%, indicating
-that the code correctly determines melt fraction as function of pressure, temperature and bulk water content. Same results can be observed in Fig.
-\ref{fig:water}a, where isobaric and isothermal melting curves are shown as function of bulk water content. In addition, Fig. \ref{fig:water}b shows that the
-code correctly determines also the percentage of water in the melt as function of pressure, temperature and melt fraction. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Melting}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Fraction.pdf}
-\caption{Isobaric melting curves for 1 GPa (panel a) and 3 GPa (panel b) as function of temperature with different bulk water contents.}
-\label{fig:fraction}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Water.pdf}
-\caption{Isobaric melting curves for 1.5 GPa as function of bulk water (panel a) and water in the melt (panel b) calculated for different temperatures.}
-\label{fig:water}
-\end{figure}
-
-\clearpage
-
-\bibliography{export}
-
-\end{document}

### 2022-09-07 tex/Benchmark/slab.tex
update main

--- 2022-03-31 FALCON_description.tex
@@
-\documentclass[hidelinks,10pt,a4paper]{article}
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{textcomp}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage{booktabs}
-\usepackage{multirow}
-\usepackage[cm]{fullpage}
-\usepackage[round]{natbib}
-\bibliographystyle{abbrvnat}
-\usepackage{doi}
-\usepackage{bm}
-\usepackage[table]{xcolor}
-\hypersetup{
-    colorlinks,
-    linkcolor={blue},
-    citecolor={blue},
-    urlcolor={blue}
-}
-\usepackage{float}
-\restylefloat{table}
-\usepackage{siunitx}
-\DeclareSIUnit\year{yr}
-
-\begin{document}
-
-\section{Introduction}\label{sec:intro}
-This document includes a description of the code FALCON, with a presentation of the features implemented followed by results of extensive benchmarks listed in
-Table \ref{tab:list}, performed to verify that the code solves correctly a wide range of problems. Description and results of all benchmarks are shown
-in Section \ref{sec:benchmark}.
-
-\begin{table}[H]
-\caption{List of benchmarks performed for each feature of the code. Data of all benchmarks can be found at 
-\url{https://github.com/aleregorda/Benchmarks}}
-\centering
-\begin{tabular}{| l | l | l |}
-\toprule
- \textbf{Feature}  & \textbf{Benchmark} & \textbf{Section} \\
-\midrule
-  Solver  & Stokes flow & \ref{sec:stokes} \\
-\hline
-  \multirow{2}{*}{Markers advection} & Zalesak disk & \ref{sec:runge} \\
-    & Conservative Interpolation Velocity & \ref{sec:CVI} \\
-\hline
-  \multirow{4}{*}{Momentum equation}  & Poiseuille flow & \ref{sec:poiseuille}\\
-    & Instantaneous 2D sphere & \ref{sec:ist_sphere} \\
-    & Rayleigh-Taylor experiment & \ref{sec:rayleigh} \\
-    & Falling block & \ref{sec:block} \\
-\hline
-    & 2D Stokes sphere & \ref{sec:sphere} \\
-    Sticky air and  & Stabilisation algorithm & \ref{sec:stab} \\
-    free surface & Topography relaxation & \ref{sec:crameri} \\
-  & Spontaneous subduction & \ref{sec:subduction} \\
-\hline
-  Erosion and sedimentation  & - & - \\
-\hline
-   Non-linear  & Slab detachment & \ref{sec:slab} \\
-   visco-plasticity & Indenter & \ref{sec:indenter} \\
-    & Brick & \ref{sec:brick} \\
-\hline
-  \multirow{3}{*}{Energy equation}  & Advection stabilisation & \ref{sec:advection} \\
-    & Simple shear heating & \ref{sec:simple_shear} \\
-    & Shear and adiabatic heating & \ref{sec:shear} \\
-\hline
-\multirow{3}{*}{Energy + momentum} & Mantle convection & \ref{sec:mantle} \\
-& Viscoplastic mantle convection & \ref{sec:visco_mantle} \\
-& Thin layer entrainment & \ref{sec:thin} \\
-\hline
-  Phase changes  & \multirow{2}{*}{Hydrated sinking cylinder} & \multirow{2}{*}{\ref{sec:quinquis}} \\
-  and hydration  &  & \\
-\hline
-  Melting & Experimental melting curves & \ref{sec:katz} \\
-\bottomrule
-\end{tabular}
-\label{tab:list}
-\end{table}
-
-The code is written in Fortran90 and uses the finite element method (FEM) with quadrilateral $Q_1 \times P_0$ elements (continuous bilinear velocity and
-discontinuous constant pressure), associated with MUMPS\footnote{\url{http://mumps.enseeiht.fr/}} \citep{Amestoy2001,Amestoy2019} that is a software package
-for solving systems of linear equations of the form $A \cdot x = b$, where \textit{A} is a square sparse matrix, by means of a direct method. Since
-$Q_1 \times P_0$ elements do not satisfy Ladyzhenskaya, Babuska and Brezzi (LBB) stability condition \citep{Donea2003}, elemental pressure is elaborated
-in post-processing to avoid spurious pressures by means of a double interpolation to smooth the pressure. Following this procedure, the elemental pressure
-is interpolated onto nodes and then back onto elements. After the smoothing procedure, the pressure field is calculated again on the nodes to be used in
-combination with the lithostatic pressure, which is calculated onto nodes. Correctness of the solution and performances of the solver in terms of time and
-memory usage are tested solving the Stokes flow with the analytical solution proposed by \citet{Donea2003} (Section \ref{sec:stokes}).
-
-The thermo-mechanics of crust-mantle systems is described by means of the conservation of mass, momentum and energy equations, expressed as follows:
-\begin{eqnarray}
-\label{eq:mass}\frac{\partial \rho}{\partial t}+\bm{\nabla} \cdot (\rho \bm{u})&=&0\\
-\label{eq:momentum}-\bm{\nabla} p + \bm{\nabla} \cdot \bm{\tau} + \rho \bm{g}&=&0\\
-\label{eq:energy}\rho C_p \left(\frac{\partial T}{\partial t} + \bm{u} \cdot \bm{\nabla} T \right)&=&\bm{\nabla} \cdot \left(k\bm{\nabla} T\right) + H_{tot}
-\end{eqnarray}
-where $\rho$ is the density, $\bm{u}$ is the velocity, \textit{p} is the pressure, $\bm{\tau}$ is the deviatoric stress, $\bm{g}$ is the gravity acceleration,
-$C_p$ is the specific heat at constant pressure, \textit{T} is the temperature, \textit{k} is the thermal conductivity and $H_{tot}$ is the total internal heat
-production. Density variations due to temperature are generally small enough to assume the density as constant ($\rho=\rho_0$) in Eq. \ref{eq:mass} and in
-Eq. \ref{eq:energy}, while it must be treated as a variable in the buoyancy term of Eq. \ref{eq:momentum}, such that \[\rho=\rho_0(1-\alpha(T-T_0))\]
-where $\rho_0$ is the density at a reference temperature $T_0$ and $\alpha$ is the coefficient of thermal expansion. Eq. \ref{eq:mass} then can be rewritten as
-\begin{equation}\label{eq:mass_0}
-\bm{\nabla} \cdot \bm{u}=0
-\end{equation}
-This simplification is known as the extended Boussinesq approximation.
-
-The time step $dt$ is chosen in according to Courant-Friedrichs-Lewy condition \citep{Anderson1995}, such that
-\[dt=C_n\; \min\left(\frac{\Delta x}{\max|\bm{u}|},\frac{(\Delta x)^2}{\kappa}\right)\]
-where $0<C_n<1$ is the Courant (CFL) number, $\Delta x$ is the minimum dimension of smallest element, $\max|\bm{u}|$ is the maximum velocity calculated on the
-entire domain and
-$\kappa=\frac{k}{\rho C_p}$ is the thermal diffusivity \citep{Thieulot2014}.
-
-\section{Lagrangian markers}\label{sec:markers}
-Elemental properties (density, viscosity, thermal conductivity, specific heat and thermal expansion) needed to solve Eqs. \ref{eq:mass_0}, \ref{eq:momentum}
-and \ref{eq:energy} are related to the composition of each element, determined by means of Lagrangian markers that are characteristic of different materials
-in the domain. Elemental properties (with the sole exception of the viscosity) are calculated using an arithmetic mean, as \[P_e=\frac{1}{n}\sum_{i=1}^n P_i\]
-where $P_e$ is the elemental property, $P_i$ is the property characteristic of the material of each marker and $n$ is the number of markers of the element.
-Differently, the average scheme for the elemental viscosity can be chosen between harmonic, geometric and arithmetic mean. Markers can be placed regularly or
-randomly at the beginning of the simulation and their advection is performed by either a 2nd-order or a 4th-order Runge-Kutta in space, interpolating the velocity 
-field on each marker by means of the shape functions (see Section \ref{sec:runge}). The interpolated velocity is then corrected by means of the Conservative
-Velocity Interpolation (CVI), which introduces a corrective term to reduce dispersion and clustering of particles in both steady state and time-dependent 
-\citep{Wang2015} (see Section \ref{sec:CVI}).
-
-The initial distribution of the markers is created using the open-source code library
-Geodynamic World Builder \citep{Fraters2019}. During the simulation, each marker carries memory of temperature, pressure and accumulated strain, which are
-determined interpolating the relative nodal parameter. In particular, pressure is interpolated onto each marker after the smoothing procedure. The number
-of markers contained in each element is maintained between $n_{min}$ and $n_{max}$. When in an element there are less markers than $n_{min}$ the code adds
-random markers to reach the $n_{min}$, while if the number is higher than $n_{max}$ some of them are randomly deleted. When new markers are added, they assume
-the properties of the nearest marker. In this way elements are never empty and maintain a number of markers inside a prefixed range.
-
-\section{The momentum equation}\label{sec:mom_eq}
-The penalty method is implemented so that Eq. \ref{eq:mass_0} can be written as:
-\begin{equation}\label{eq:mass_penalty}
-\bm{\nabla} \cdot \bm{u} + \frac{p}{\lambda}=0
-\end{equation}
-where $\lambda$ is the so-called penalty parameter, which should be 6-7 orders of magnitude larger than the shear viscosity to ensure that mass conservation
-is satisfied \citep{Donea2003,Thieulot2014}.
-
-The deviatoric stress tensor in Eq. \ref{eq:momentum} can be written in terms of the strain rate tensor as $\bm{\tau}=2\eta\dot{\bm{\epsilon}}$, with
-$\dot{\bm{\epsilon}}=\frac{1}{2}\left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)$. Therefore, Eq. \ref{eq:momentum} can be rewritten as:
-\begin{equation}\label{eq:momentum_pressure}
--\bm{\nabla} p + \bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right) + \rho \bm{g} = 0
-\end{equation}
-Finally, using pressure from Eq. \ref{eq:mass_penalty}, Eq. \ref{eq:momentum_pressure} can be rewritten as:
-\begin{equation}\label{eq:momentum_penalty}
-\lambda \bm{\nabla} \left(\bm{\nabla} \cdot \bm{v}\right)+\bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right)+\rho\bm{g}=0
-\end{equation}
-The penalty method is been associated to the iterative Uzawa method, as described in detail in \citet{Dabrowski2008} and \citet{Thieulot2014}
-(see Section \ref{sec:poiseuille}). Results of benchmarks performed to verify the correctness in the implementation of Eq. \ref{eq:momentum_penalty} are
-shown in Sections \ref{sec:ist_sphere} and \ref{sec:rayleigh}.
-
-To support large viscosities variations the penalty parameter is related to the effective elemental viscosity by means of a dimensionless coefficient, so that
-$\lambda=\lambda_e(e)\eta_{eff}(e)$ \citep{Marotta2006,Dabrowski2008,Thieulot2014}. Benchmark of the falling block \citep{Gerya2003a,Gerya2010b,Thieulot2011}
-is performed to verify that the code can correctly deal buoyancy driven flows with strong viscosity contrasts (Section \ref{sec:block}).
-
-\section{Sticky air and free surface}\label{sec:sticky}
-The Earth's surface can be treated by means of either the so-called sticky-air or a true free surface method, both of them implemented in the code. In the
-sticky-air method the surface is approximated with the introduction of a buoyant layer with a viscosity at least four orders of magnitude lower than the crust
-\citep{Schmeling2008,Crameri2012} and the interface between lithosphere and air is defined using a chain of passive markers that are advected as the Lagrangian
-markers. The correctness of the evolution of the markers chain is tested with the experiment of a 2D time-dependent Stokes sphere below a free surface and
-compared with results from ASPECT\footnote{\url{https://aspect.geodynamics.org/}} \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} (Section
-\ref{sec:sphere}). In the true free surface case the top boundary is assumed stress-free and velocities are not fixed. In this case topography variations are
-described by vertical deformations of the mesh that depend on the velocity field of the nodes that identify the free surface, while horizontal deformations are
-not taken into account. This procedure is known as the Arbitrary Lagrangian-Eulerian (ALE) method and its implementation follows the technique described in
-\citet{Thieulot2011}. However, although the implementation of a true free surface better reproduces laboratory experiment, extremely small time steps can be
-necessary to maintain stability \citep{Kaus2010a,Quinquis2011,Thieulot2014}. Therefore, the stability algorithm proposed by \citet{Kaus2010a} is implemented
-to avoid instabilities due to high density differences at the free surface when using too large time steps. The implementation of this algorithm are tested by
-performing the experiment described by \citet{Kaus2010a} (Section \ref{sec:stab}).
-
-The topography relaxation benchmark proposed by \citet{Crameri2012} is performed to verify that the code correctly recovers topography variations in case of
-both the sticky-air and the true free surface method (Section \ref{sec:crameri}). Finally, Section \ref{sec:subduction} show the results of the spontaneous
-subduction experiment described by \citet{Schmeling2008}.
-
-\section{Erosion and sedimentation}\label{sec:erosion}
-Surface processes at either the lithosphere-air interface or in correspondence of the free surface have been implemented by means of the software
-Fastscape\footnote{\url{https://fastscape.org/}} \citep{Braun2013,Cordonnier2019,Yuan2019a,Yuan2019b}. FastScapeLib is a set of routines that solve the stream
-power law (enriched by a sediment transport/deposition term) \citep{Yuan2019a}, hillslope diffusion and marine transport and deposition \citep{Yuan2019b}.
-The partial differential equation solved by FastScapeLib is:
-\begin{equation}
-\frac{\partial h}{ \partial t}=U-K_f A^{m_e}S^{n_e} + \frac{G}{A}\int_A \left( U-\frac{\partial h}{\partial t} \right) dA + K_d \bm{\nabla}^2 h 
-\end{equation}
-where $h$ is the topography, $U$ is the uplift, $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $S$ is the slope, $m_e$ and $n_e$ are
-parameters that should be chosen to constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}, $G$ is the dimensionless
-deposition/transport coefficient for the enriched stream power law and $K_d$ is the effective diffusivity. Terms $K_f A^{m_e}S^{n_e}$ and $K_d \bm{\nabla}^2 h$
-consider long-range processes, described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, and short-range processes,
-proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}.
-
-In case of sticky air, slope is determined by means of the markers chain, which is vertically corrected considering erosion and sedimentation rates. After the
-correction, continental markers above the markers chain are transformed in air markers and air markers below the markers chain become sediments. Similarly,
-slope is determined considering the nodes on the top boundary in case of true free surface and they are vertically corrected as for the markers chain. In this
-case, continental markers above the top boundary are simply deleted.
-%Surface processes at the lithosphere-air interface are implemented taking into account fluvial erosion (long-range processes), hillslope processes (short-range
-%processes) and failure of slopes exceeding a threshold angle $\theta$ \citep{Burov1997,Tucker1999,Sacek2012,Croissant2014,Ueda2015,Beucher2020}. Erosion rates
-%$\dot{E}_{lr}$ due to long-range processes can be described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{lr}=K_f A^{m_e} \bm{\nabla} h^{n_e}
-%\end{equation}
-%where $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $\bm{\nabla} h$ is the slope and $m_e$ and $n_e$ are parameters chosen in to
-%constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}. Erosion rates $\dot{E}_{sr}$ due to short-range processes are
-%implemented by assuming that transport is linearly proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{sr}=K_d \bm{\nabla}^2 h
-%\end{equation}
-%where $K_d$ is the effective diffusivity. In addition, the slope is modified if it still exceeds a threshold angle after long- and short-range processes
-%\citep{Ueda2015}.
-
-%Offshore sedimentation processes are implemented similarly to hillslope processes but with a depth-dependent coefficient \citep{Sacek2012,Beucher2020}.
-%The sedimentation rate $\dot{S}$ is calculated as
-%\begin{equation}
-%\dot{S}=K_s \bm{\nabla}^2 h
-%\end{equation}
-%with
-%\begin{equation}
-%K_s=K^*_s \exp\left(\frac{h-h_{sl}}{h_0}\right)
-%\end{equation}
-%where $K^*_s$ is a constant, $h_{sl}$ is the sea level and $h_0$ is a characteristic length. In case of sticky air, slope is determined by means of the markers
-%chain, which is vertically corrected considering erosion and sedimentation rates. After the correction, continental markers above the markers chain are
-%transformed in air markers and air markers below the markers chain become sediments. Similarly, slope is determined considering the nodes on the top boundary
-%in case of true free surface and they are vertically corrected as for the markers chain. In this case, continental markers above the top boundary are simply
-%deleted.
-
-\section{Non-linear rheologies}\label{sec:nonlinear}
-Non-linear rheologies are implemented combining viscous creep (dislocation and diffusion) and plastic yielding. For each marker, diffusion and dislocation
-viscosity ($\eta_{df|ds}$) can be determined as follows \citep{Karato1993,Warren2008b,Wang2016}:
-\begin{equation}\label{eq:viscous}
-\eta_{df|ds}=f_s \left(\frac{d^m}{A}\right)^{\frac{1}{n}} \left (\frac{2^{\frac{1-n}{n}}}{3^{\frac{n+1}{2n}}} \right)
-I_2^{\frac{(1-n)}{n}}\exp\left(\frac{Q+pV}{nRT}\right)
-\end{equation}
-where $f_s$ is a scaling factor used to represent lithologies that are stronger or weaker than the base set, $d$ is the grain size, $m$ is the grain size exponent,
-$n$ is the stress exponent, $A$ is the uniaxial pre-exponential factor, $I_2$ is the square root of the second invariant of the strain rate tensor, $Q$ is the
-activation energy, $p$ is the pressure, $V$ is the activation volume, $R$ is the gas constant and $T$ is the temperature.
-Pressures and temperatures are determined by the interpolation of the nodal parameters, while strain rates are calculated by means of the
-derivative of the velocities on the nodes. In case of diffusion creep $n=1$ and $m>0$, in case of dislocation creep $n>1$ and $m=0$. The viscous creep $\eta_{cp}$
-is then calculated as the harmonic average between $\eta_{df}$ and $\eta_{ds}$:
-\begin{equation}\label{eq:visco_creep}
-\eta_{cp}=\left(\frac{1}{\eta_{df}}+\frac{1}{\eta_{ds}}\right)^{-1}
-\end{equation}
-The implementation of the non-linear viscous creep viscosity depending on the strain rate is tested by means of the slab detachment benchmark
-(Section \ref{sec:slab}).
-
-Plastic yielding is implemented rescaling $\eta_{cp}$ in order to limit the stress below the yield stress $\sigma_y$ \citep{Thieulot2008,Thieulot2014,Glerum2018},
-obtaining
-\begin{equation}\label{eq:plastic}
-\eta_{pl}=\frac{\sigma_y}{2I_2}
-\end{equation}
-where $\eta_{pl}$ is the plastic viscosity and the yield stress is determined following the Drucker-Prager criterion, such as
-\begin{equation}\label{eq:yield}
-\sigma_y=C\;\cos(\phi)+p\;\sin(\phi)
-\end{equation}
-where $C$ is the cohesion and $\phi$ is the internal friction angle. In case of negative pressure it is imposed equal to 0, so that negative yield stress are
-excluded. The correctness of the non-linear solution in case of plastic viscosity with variable internal friction angle is verified performing indenter and
-brick experiments (Sections \ref{sec:indenter} and \ref{sec:brick}, respectively).
-
-Strain softening is taken into account for both viscous creep and plastic viscosity \citep{Huismans2003,Babeyko2005,Huismans2005,Sobolev2005,Warren2008a} by
-means of the accumulated strain $\epsilon$ memorised by each marker. Softening in the viscous creep determines a linear decrease of $\eta_{vc}$ by means of a
-viscous strain softening factor $W_S$ that increases linearly from $W_{S_0}$ to $W_{S_{\infty}}$ for $\epsilon_{S_0}<\epsilon_S<\epsilon_{S_{\infty}}$ 
-\citep{Warren2008a}. This viscous softening can be related to strain-induced grain size reduction \citep{Warren2008a}. Differently, plastic softening is 
-simulated with a linear decrease of internal friction angle $\phi(\epsilon)$ and cohesion $C(\epsilon)$ in according to
-\begin{eqnarray}
-\label{eq:friction1}\phi(\epsilon)&=&\phi_0+(\phi_{\infty}-\phi_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}\\
-\label{eq:friction2}C(\epsilon)&=&C_0+(C_{\infty}-C_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}
-\end{eqnarray}
-where $\phi_0$, $C_0$ and $\phi_{\infty}$, $C_{\infty}$ are internal friction angle and cohesion for $\epsilon_0$ and $\epsilon_{\infty}$, respectively
-\citep{Huismans2003,Huismans2005,Warren2008a,Thieulot2014}. Plastic softening approximates deformation-induced softening of faults and brittle shear zones
-\citep{Warren2008a}.
-
-Plastic yielding and viscous creep are then combined to obtain a viscoplastic viscosity $\eta_{vp}$ as follows, assuming that they are independent processes
-\citep{Karato2008,Glerum2018}:
-\begin{equation}\label{eq:viscoplastic}
-\eta_{vp}=\min(\eta_{cp},\eta_{pl})
-\end{equation}
-Finally, effective viscosity $\eta_{eff}$ is capped by the minimum and the maximum viscosity ($\eta_{min}$ and $\eta_{max}$, respectively) to avoid extremely
-low or high viscosity \citep{Glerum2018} as follows
-\begin{equation}\label{eq:effective}
-\eta_{eff}=\min(\max(\eta_{vp},\eta_{min}),\eta_{max})
-\end{equation}
-Viscous creep and plastic yielding are non-linear rheologies because of their dependence on the velocity field through pressure and strain rates. Therefore,
-the solution is determined by means of Picard-type iterations, until convergence of the velocity field \citep{Glerum2018}. The convergence is verified at each
-iteration $i$ via the nonlinear residual $\bm{\mathcal{R}}^i$ that can be determined as
-\begin{equation}\label{eq:convergence}
-\bm{\mathcal{R}}^i= \mathbb{K} (\eta_{eff}(\dot{\bm{\epsilon}}^{i-1},p^{i-1}))\cdot \bm{v}^{i-1} - \bm{f}^i
-\end{equation}
-%\begin{equation}\label{eq:convergence}
-%\chi=1-\frac{\langle (\bm{v}^i-\langle\bm{v}^i\rangle)\cdot (\bm{v}^{i+1}-\langle\bm{v}^{i+1}\rangle) \rangle}{\sqrt{\sigma^i\sigma^{i+1}}}
-%\end{equation}
-%Iterations are performed until either convergence ($\chi<tol$) or maximal number of non-linear iterations ($it_{max}$) is reached.
-where $\mathbb{K}$ is the Stiffness matrix and $\bm{f}$ is the right hand side vector \citep{Spiegelman2016,Glerum2018}.  The $L_2$-norm is extracted from
-$\bm{\mathcal{R}}^i$ and it is normalised as follows
-\begin{equation}\label{eq:convergence_res}
-\frac{||\bm{\mathcal{R}}^i||_2}{||\bm{\mathcal{R}}^0||_2}
-\end{equation}
-where $||\bm{\mathcal{R}}^0||_2$ is the $L_2$-norm of the first nonlinear iteration.
-Since often the normalised nonlinear residual drop very quickly over the first 2-3 iterations, also the $L_2$-norm of velocity and pressure residuals are
-calculated as
-\begin{equation}\label{eq:convergence_vel}
-\frac{||\Delta\bm{v}||_2}{||\bm{v}||_2}=\frac{||\bm{v}_i-\bm{v}_{i-1}||_2}{||\bm{v}_{i}||_2}
-\end{equation}
-and
-\begin{equation}\label{eq:convergence_press}
-  \frac{||\Delta p||_2}{||p||_2}=\frac{||p_i-p_{i-1}||_2}{||p_{i}||_2}
-\end{equation}
-respectively. Iterations are performed until either maximal number of non-linear iterations ($it_{max}$) is reached or all the normalised nonlinear residual, 
-the velocity residual and the pressure residual converge under a defined tolerance $tol$.
-
-\section{The energy equation}\label{sec:en_eq}
-The stabilisation of the advection term of the energy equation (Eq. \ref{eq:energy}) needed to avoid possible oscillations in the thermal solution in the case
-when advection dominates over diffusion, is implemented by means of a streamline-upwind Petrov–Galerkin (SUPG) method, for which the advection term is modified
-follows the discussion in \citet{Thieulot2011} and \citet{Thieulot2014} (see Section \ref{sec:advection} for benchmark). The total internal heat production
-$H_{tot}$ in Eq. \ref{eq:energy} includes radiogenic heating $H_r$, shear heating $H_s$ and adiabatic heating $H_a$, where
-\begin{equation}\label{eq:shear}
-H_s=2\eta \dot{\bm{\epsilon}} : \dot{\bm{\epsilon}} = 2\eta(\dot{\epsilon}_{xx}^2+\dot{\epsilon}_{yy}^2+2\dot{\epsilon}_{xy}^2)
-\end{equation}
-and
-\begin{equation}\label{eq:adiabatic}
-H_a=T \alpha \frac{Dp}{Dt} \approx -\alpha T \rho g_y v
-\end{equation}
-The correctness of the implementation and the computation of shear and adiabatic heating is verified performing an exercise with an analytical solution
-(Section \ref{sec:simple_shear}) and Exercise 9.4 in \citet{Gerya2010b} (Section \ref{sec:shear}). Furthermore, mantle convection and thin layer entrainment
-benchmarks (Sections \ref{sec:mantle} and \ref{sec:thin}, respectively) are performed to verify that the code solve correctly Eqs. \ref{eq:momentum_penalty}
-and \ref{eq:energy} in case of a temperature field characterised by an initial perturbation and with a temperature-dependent density. At last, the viscoplastic
-mantle convection benchmark is performed to check also the viscoplasticity in case of a initial perturbed temperature field (Section \ref{sec:visco_mantle}).
-
-\section{Phase transitions and hydration}\label{sec:phase}
-Variations of effective density, specific heat and coefficient of thermal expansion during the evolution of crustal and mantle materials at different
-pressure-temperature ($p$-$T$) conditions are computed by means of version 6.8.6 of Perple\_X software package \citep{Connolly2005}, similarly to the
-implementation described in \citet{Marotta2020}. By default, mineral assemblages and properties of each lithology (oceanic crust, lower and upper continental
-crust, sediments and dry or hydrated mantle) are calculated for temperatures from \SIrange{330}{1600}{\kelvin}, with increments of approximately \SI{4}{\kelvin},
-and pressures from \SIrange{0.1}{30}{\giga\pascal}, with increments of approximately \SI{0.03}{\giga\pascal}, for a total of almost 100000 points.
-
-The introduction of phase transitions produces two main effects that must be taken into account in numerical models: variations in density and release/absorption
-of latent heat \citep{Tackley2010}. Effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into
-account considering effective coefficient of thermal expansion \citep{Christensen1985,Zhong2015}. Similarly, effects of latent heat in the energy equation
-(eq \ref{eq:energy}) are taken into account by considering effective specific heat and coefficient of thermal expansion
-\citep{Christensen1985,Tackley2010,Zhong2015}. Effective density, specific heat and coefficient of thermal expansion are included in Perple\_X files and are
-assumed by each Lagrangian marker in according to its $p$-$T$ conditions.
-
-Hydration processes related to dehydration of subducting lithosphere \citep{Schmidt1998,Liu2007,Faccenda2009,Faccenda2010,Faccenda2014,Rosas2016} strongly
-influence the thermo-mechanics inside the mantle wedge, mainly because of both weakening effects on the mantle rheology and density variations in case of
-mantle serpentinization \citep{Gerya2002,Honda2003,Arcay2005,Roda2010,Regorda2017}. In the case that hydration is switched on, the amount of bound and free
-water is memorised by each marker, following the implementation of \citet{Quinquis2014}; hydration and dehydration processes are related to the amount of bound
-water of each marker and to the maximum amount of water it can transport, i.e. if the amount of bound water exceeds the maximum amount of water, the marker
-dehydrates and releases free water that can hydrates under-saturated markers. Maximum water content of each marker is determined as function of lithology and
-$p$-$T$ conditions and it is calculated using Perple\_X, in the same way than effective density, specific heat and coefficient of thermal expansion. Bound water
-is advected together with the markers, neglecting the effect of bound water diffusion, while free water simply migrates vertically and is not coupled to the
-solid-phase flow of the mantle wedge \citep{Arcay2005,Quinquis2014}. The correctness of the migration scheme of the free water and its absorption into bound
-water by under-saturated markers are verified performing the experiment described in \citet{Quinquis2014} (Sec. \ref{sec:quinquis}).
-
-The mantle viscosity weakening related to the amount of water has been extensively studied \citep{Chopra1981,Hirth2003} and it is implemented in according to
-\citet{Arcay2005} and \citet{Horiuchi2016}, as follows
-\begin{equation}\label{eq:wet_visc}
-\eta_{wet}=\eta_{dry}\left(\left[1-\frac{1}{f_v}\right]\exp\left(-\frac{[OH^-]}{[OH^-]_0}\right)+\frac{1}{f_v}\right)
-\end{equation}
-where $f_v$ is the viscosity reduction factor between dry and wet conditions, $[OH^-]$ is the water content and $[OH^-]_0$ is a reference water content set to
-620 ppm (0.062 wt.\%) \citep{Arcay2005}. By default, $f_v$ is set to 100, in according to \citet{Arcay2005} and \citet{Horiuchi2016}.
-
-\section{Mantle melting}\label{sec:melt}
-Melt fraction in the mantle depends on temperature ($T$, in \textdegree C), pressure ($p$, in \si{\giga\pascal}) and the water content in the melt ($X_{H_2O}$, in wt.\%).
-The determination of the melt fraction $M$ is implemented as explained by \citet{Katz2003}, according to the successive modification by \citet{Langmuir2006} and
-\citet{Kelley2010}, as follows
-\begin{eqnarray}\label{eq:melt_fr}
-M(p,T,X_{H_2O})&=&-T+T_{sol}+(x_t\cdot ln(p+y_t))\cdot X_{H_2O}-\nonumber\\
-&&+K_m\left(\frac{X_{H_2O}^{bulk}}{D_{H_2O}(1-X_{H_2O})+X_{H_2O}}\right)^{\gamma}
-\end{eqnarray}
-where $T_{sol}$ is the temperature of the dry solidus, $D_{H_2O}$ is the partition coefficient and $x_t\cdot log(p+y_t)$ indicates the pressure dependence of
-the melting curve \citep{Kelley2010}. $x_t$, $y_t$, $K_m$ and $\gamma$ are parameters chosen in according to \citet{Kelley2010}.
-However, $X_{H_2O}$ depends on the melt fraction as
-\begin{equation}\label{eq:melt_wt}
-X_{H_2O}(M)=\frac{X_{H_2O}^{bulk}}{D_{H_2O}+M(1-D_{H_2O})}
-\end{equation}
-and a numerical solution can be found using a root-finder method \citep{Katz2003,Wang2016}. Furthermore, the content of water in the melt is limited by the
-pressure-dependent saturation concentration of water in the melt, determined as
-\begin{equation}\label{eq:melt_max}
-X_{H_2O}^{sat}=\chi_1 p^{\lambda_m}+\chi_2 p
-\end{equation}
-where $\chi_1$, $\chi_2$ and $\lambda_m$ are parameters chosen according to \citet{Katz2003}, \citet{Langmuir2006} and \citet{Kelley2010}. The correctness in
-the determination of melt fractions and water contents in the melt are tested comparing model results with experimental curves obtained by \citet{Katz2003}
-(Sec. \ref{sec:katz}).
-
-In case of melting, viscous creep viscosity ($\eta_{vp}$) is modified according to \citet{Wang2016}, as
-\begin{equation}\label{eq:melt_visc}
-\eta_{melt}=\eta_{vp}\cdot \exp(\alpha_m M)
-\end{equation}
-where $\alpha_m$ are melt fraction factors for dislocation and diffusion creep, chosen in according to \citet{Wang2016}.
-
-Effective density ($\rho_{eff}$) for partially molten rocks is calculated as
-\begin{equation}\label{eq:melt_dens}
-\rho_{eff}=\rho_s(1-M)+\rho_m M
-\end{equation}
-where $\rho_s$ and $\rho_m$ are the densities of the solid and the molten rock, respectively \citep{Gerya2004b,Gerya2007,Wang2016}. The density of the
-solid rock is extracted by Perple\_X, while the density of the molten portion is calculated as
-\begin{equation}\label{eq:melt_dens_M}
-\rho_m(p,T)=\rho_0[1-\alpha(T-T_0)][1+\beta(p-p_0)]
-\end{equation}
-where $\rho_0$ is the density at the reference temperature ($T_0$) and pressure($p_0$), and $\alpha$ and $\beta$ are the thermal expansion and compressibility
-coefficients, respectively \citep{Gerya2004b,Gerya2007,Wang2016}.
-
-As for the phase transitions, effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into account
-considering effective coefficient of thermal expansion. Similarly, latent heat due to melting/crystallisation can be implicitly included in the energy equation
-(Eq. \ref{eq:energy}) by considering effective specific heat ($C_{p_{eff}}$) and thermal expansion ($\alpha_{eff}$) of the partially molten rocks, in according to
-\begin{eqnarray}
-\label{eq:cp_effective}C_{p_{eff}}&=&C_p+H_L\left(\frac{\partial M}{\partial T}\right)_p \\
-\label{eq:alpha_effective}\alpha_{eff}&=&\alpha+\rho\frac{H_L}{T}\left(\frac{\partial M}{\partial p}\right)_T
-\end{eqnarray}
-where $C_p$ and $\alpha$ are the specific heat and the coefficient of thermal expansion of the solid rock, respectively, and $H_L$ is the latent heat of the
-molten rock \citep{Gerya2004b,Gerya2007,Tackley2010}.
-
-\section{Benchmarks}\label{sec:benchmark}
-\subsection{Error measurements}\label{sec:error}
-In order to determine the accuracy of velocity and pressure field of the benchmarks, the $L_2$-norm is computed by numerical integration on the Gauss-Legendre
-quadrature points. $L_2$-norm for pressure and velocity errors can be evaluated as
-\begin{eqnarray}
-\textrm{err}_p&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|p^n_i(r_q)-p^a_i(r_q)|^2\right)w_q|J_q|}\nonumber \\
-\textrm{err}_v&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|u^n_i(r_q)-u^a_i(r_q)|^2+|v^n_i(r_q)-v^a_i(r_q)|^2\right)w_q|J_q|}\nonumber
-\end{eqnarray}
-respectively, where $n_e$ is the number of elements, $n_q$ is the number of quadrature points per element, $p^n_i(r_q)$ and $p^a_i(r_q)$ are the numerical and
-analytical pressure, respectively, in each quadrature point $q$, $w_q$  and $J_q$ are the weight and the Jacobian at the quadrature point
-$q$, $u^n_i(r_q)$, $v^n_i(r_q)$, $u^a_i(r_q)$ and $v^a_i(r_q)$ are the numerical and analytical velocities, respectively, in each quadrature point $q$.
-
-Other quantities used as comparison with original benchmarks are the root-mean-square velocity over the whole domain and over the surface
-\begin{eqnarray}
-v_{\textrm{rms}}&=&\sqrt{\int_{\Omega}|\bm{v}|^2 d\Omega}\nonumber \\
-v_{\textrm{rms}}^{top}&=&\sqrt{\int_0^1 u^2 \bigg|_{y=1} dx}\nonumber
-\end{eqnarray}
-respectively, top and bottom Nusselt numbers
-\[\textrm{Nu}^{top/bottom}=-\int_0^1\frac{\partial T}{\partial y} \bigg|_{y=1/y=0} dx\]
-the average rate of work done against gravity
-\[\left\langle W \right\rangle=\int_{\Omega}T u_y d\Omega\]
-the average rate of viscous dissipation
-\[\left\langle \Phi \right\rangle=\int_{\Omega}\tau_{ij}\dot{\epsilon}_{ij} d\Omega\]
-and the percentage error between $\left\langle W \right\rangle$ and $\left\langle \Phi \right\rangle/Ra$
-\[\delta=\frac{\left|\left\langle W \right\rangle -\frac{\left\langle \Phi \right\rangle}{Ra} \right|}{\max \left(\left\langle W \right\rangle,
-\frac{\left\langle \Phi \right\rangle}{Ra}\right)}\times100\]
-
-\subsection{Stokes flow}\label{sec:stokes}
-The problem consists of determining velocity field $(u,v)$ and pressure $p$ in case of a manufactured solution with prescribed body forces such as
-\begin{eqnarray}
-b_1&=&(12-24y)x^4+(-24+48y)x^3+(-48y+72y^2-48y^3+12)x^2+\nonumber \\
-&&+(-2+24y-72y^2+48y^3)x+1-4y+12y^2-8y^3\nonumber \\
-b_2&=&(8-48y+48y^2)x^3+(-12+72y-72y^2)x^2+(4-24y+48y^2-48y^3+\nonumber \\
-&&+24y^4)x-12y^2+24y^3-12y^4\nonumber
-\end{eqnarray}
-for which the exact solution is:
-\begin{eqnarray}
-u(x,y)&=&x^2(1-x)^2(2y-6y^2+4y^3)\nonumber \\
-v(x,y)&=&-y^2(1-y)^2(2x-6x^2+4x^3)\nonumber \\
-p(x,y)&=&x(1-x)-\frac{1}{6}\nonumber
-\end{eqnarray}
-The domain is a unit square a constant viscosity ($\eta=1$) and the penalty parameter is set to $\lambda=10^7$. Velocity boundary conditions are set to no slip
-$(\bm{v}=\bm{0})$ on all boundaries. The problem is performed for different grid resolution between \num{8x8} and \num{1024x1024} elements.
-
-The errors between the analytical solution and the numerical prediction of the velocity and the pressure field are calculated as in Section \ref{sec:error}.
-Fig. \ref{fig:errors}a shows that both velocity and pressure field converge to the exact solution with the decrease of the element size, following the
-theoretical convergence rate. The convergence of pressure, in contrast with the observation by \citet{Donea2003} for $Q_1 \times P_0$ elements, indicate the
-effectiveness of the smoothing procedure. Smoothed pressure field for grid resolution of \num{128x128} elements is shown in Fig. \ref{fig:errors}b (continuous
-red line), in comparison with the analytical solution (dashed blue line). Solve times and memory usage needed to generate the solution are shown in
-Fig. \ref{fig:MUMPS}. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Solver/Stokes_Flow}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/errors.pdf}
-\caption{Velocity and pressure error for the Stokes flow experiment between generated and analytical solution as a function of element size (panel a),
-and comparison between smoothed pressure and analytical solution as function of $x$ coordinate for a grid resolution of \num{128x128} elements (panel b).}
-\label{fig:errors}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/MUMPS.pdf}
-\caption{Panel a) Analysis, factorisation, solution and total solve times (panel a) and factorisation memory usage (panel b) as a function of the total number
-of degrees of freedom for the Stokes flow experiment.}
-\label{fig:MUMPS}
-\end{figure}
-
-\subsection{Markers advection}\label{sec:runge}
-The 2nd-order Runge-Kutta advection scheme in space is tested by means of the Zalesak disk test \citep{Zalesak1979,Thieulot2014}. The benchmark is performed
-in a unit square domain with a grid resolution of \num{32x32} elements and 1000000 markers and values of Courant number between 0.25 and 1.
-At $t=0$, the disk is centred at position $(0.5;0.75)$ with a radius $R=0.15$ and has a vertical fissure 0.05 wide and 0.2 high (orange in Fig. \ref{fig:runge})
-and with the velocity field prescribed in the entire domain as
-\begin{eqnarray}
-u(x,y)&=&2\pi\left(y-\frac{L_x}{2}\right)\nonumber \\
-v(x,y)&=&-2\pi\left(x-\frac{L_x}{2}\right)\nonumber
-\end{eqnarray}
-the markers are back to their initial location after a $2\pi$ rotation ($t=1$). The clockwise rotation of the disk at $t=0.25$, 0.5 and 0.75 
-is shown in Fig. \ref{fig:runge} (red, blue and green, respectively). The distance from the centre of 1 marker is calculated throughout the rotation to
-evaluate the error for different values of Courant number. As expected, the error increase with the increasing of the time step, as shown in 
-Fig. \ref{fig:runge_err}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/Zalesak_Disk}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=200px]{./Figures/Zalesak.pdf}
-\caption{Position of the Zalesak disk at $t=0$, 0.25, 0.5 and 0.75 (orange, red, blue and green, respectively) for the case with Courant number of 1.}
-\label{fig:runge}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/zal_marker.pdf}
-\caption{Distance from the centre as function of time for values of Courant number of  of 0.25, 0.3, 0.5, 0.75 and 1 (green, orange, blue, black and red,
-respectively). R indicates the distance from the centre considering a perfect circle trajectory.}
-\label{fig:runge_err}
-\end{figure}
-
-\subsection{Conservative Velocity Interpolation (CVI)}\label{sec:CVI}
-The CVI correction is checked by means of the Stokes flow experiment by \citet{Donea2003}, as presented in \ref{sec:stokes}. The advection of Lagrangian
-markers is performed using either a 2nd-order or a 4th-order Runge-Kutta scheme with CFL=0.5 and an initial random distribution of 25 markers per element.
-Fig. \ref{fig:CVI} shows the minimum and maximum values of markers per element throughout the experiment with and without the CVI correction and using
-different Runge-Kutta schemes. The introduction of the CVI correction allows the
-tests to run for times longer than $t=2000$ without show empty elements, while tests without the CVI correction stop at approximately $t=600$, when at least
-one element goes empty. Fig. \ref{fig:CVI_time} shows the different number of markers per element and the markers distribution at $t=600$ obtained without and
-with the CVI correction (panels b and c, respectively) in case of a 2nd-order Runge-Kutta scheme. All data can be found at
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/CVI}.
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/CVI.pdf}
-\caption{Maximum and minimum number of markers per element throughout the experiment, with (green and orange lines) or without (purple and blue lines)
-the CVI correction and using either a 2nd-order (blue and orange lines) or a 4th-order (purple and green lines) Runge-Kutta scheme.}
-\label{fig:CVI}
-\end{figure}
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/CVI_time.pdf}
-\caption{Number of markers per element and markers distribution through time using 2nd.order Runge-Kutta scheme.
-Panels a: number of markers per element (panel a1) and distribution of the markers (panel a2) after the first
-iteration; panels b and c: comparison between number of markers per element (panels b1 and c1) and markers distribution (panels b2 and c2) without and with
-the CVI correction at $t=600$; panels d: number of markers per element (panel d1) and markers distribution (panel d2) in case of CVI correction at $t=2000$.}
-\label{fig:CVI_time}
-\end{figure}
-
-\subsection{Poiseuille flow}\label{sec:poiseuille}
-The domain is a rectangle with $L_x=2$ and $L_y=1$ and constant density and viscosity ($\rho=1$ and $\eta=1$), gravity acceleration $\bm{g}=\bm{0}$ and penalty
-parameter $\lambda=10^8$. The grid is composed by \num{40x20} elements. Velocity boundary conditions are set to no slip $(\bm{v}=\bm{0})$ at the top and the
-bottom, and a parabolic profile is imposed on the sides, with $u=y(1-y)$ and $v=0$. The analytical solution is then given by:
-\begin{eqnarray}
-u(x,y)&=&y(1-y)\nonumber \\
-v(x,y,)&=&0\nonumber \\
-p(x,y)&=&2\eta\left(\frac{L_x}{2}-x\right)\nonumber
-\end{eqnarray}
-
-The velocity field predicted by the model follows the expected parabolic profile (Fig. \ref{fig:poiseuille}a and Fig. \ref{fig:poi_plot}a) and, in case of the
-classic penalty method with no iterations, the pressure is clearly related to the divergence by means of the penalty parameter (Fig. \ref{fig:poiseuille}b and
-c and Fig. \ref{fig:poi_plot}b). Fig. \ref{fig:poiseuille}d shows that one Uzawa iteration is sufficient to bring the divergence down to \num{e-15}, with no
-correlation with the pressure field. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Poiseuille%20Flow}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Poiseuille.pdf}
-\caption{Velocity field (panel a), pressure (panel b) and divergence velocity of a Poiseuille flow in case of the classic penalty method (no iterations) and
-after one Uzawa iteration (panel c and d, respectively).}
-\label{fig:poiseuille}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Analytical.pdf}
-\caption{Velocity (panel a) and pressure (panel b) field predicted by the model for a Poiseuille flow with respect with their analytical solutions. The
-velocity field is plotted as function of the vertical coordinate in $L_x/2$ and the pressure is plotted as function of the x coordinate in $L_y/2$.}
-\label{fig:poi_plot}
-\end{figure}
-
-\subsection{Instantaneous 2D sphere}\label{sec:ist_sphere}
-The domain is a unit square with gravity $\bm{g}=(0,-1)$. The fluid has constant density and viscosity ($\rho_f=1$ and $\eta_f=1$). The sphere is in the middle
-of the domain with a radius R = 0.123456798 and has constant density and viscosity ($\rho_s=10^{-2}+\rho_f$ and $\eta_s=10^3 \cdot \eta_f$).
-We distinguish three different types of velocity boundary conditions:
-\begin{enumerate}
-\item FS: free slip conditions on all sides.
-\item NS: no slip conditions on all sides.
-\item OT: free slip conditions on the sides and the bottom, and open boundary at the top.
-\end{enumerate}
-All three cases are performed for different grid resolutions, between \num{16x16} and \num{512x512} elements with 50 randomly distributed markers per element,
-and different average schemes for the viscosity (harmonic, geometric and arithmetic). The velocity in the centre of the sphere $\bm{v}(0.5;0.5)$), minimum and
-maximum velocities ($u_{min}$, $u_{max}$, $v_{min}$, $v_{max}$) and pressures ($p_{min}$, $p_{max}$) on the entire domain, $v_{\textrm{rms}}$ and average pressure on the
-entire domain ($p_{avg}$) are compared with solutions generated by ASPECT \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and all results
-can be found at \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere2D} and 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Instantaneous_2D_Sphere}. Results in terms of the $v_{\textrm{rms}}$ for NS model 
-are also shown in Fig. \ref{fig:inst_sphere}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_NS.pdf}
-\caption{$v_{\textrm{rms}}$ as function of element size for the instantaneous 2D sphere experiment in case of no slip boundary conditions and with different average
-schemes for the viscosity. Results are compared with results obtained by other numerical codes.}
-\label{fig:inst_sphere}
-\end{figure}
-
-\subsection{Rayleigh-Taylor instability}\label{sec:rayleigh}
-This problem has been originally presented by \citet{vanKeken1997} and here it is performed as the isoviscous case (Case 1a). The domain has $L_x=0.9142$ and
-$L_y=1$ and gravity $\bm{g}=(0,-1)$. Two fluids with same constant viscosities ($\eta_1=\eta_2=1$) and different densities ($\rho_1=1000$ and $\rho_2=1010$),
-with the lighter fluid at the bottom. The initial interface between the fluids is given by $y(x)=0.2+0.02 \cos \left(\frac{\pi x}{L_x}\right)$. The experiment
-is performed with different grid sizes (\num{50x50}, \num{80x80}, \num{100x100} and \num{256x256}). A total of 1960000 markers are randomly distributed at the
-beginning of the simulation. Velocity boundary conditions are set to no slip at the top and the bottom, and to free slip at the sides of the domain.
-
-$v_{\textrm{rms}}$ as function of time is reported in Fig. \ref{fig:RT}, matching well with results shown by \citet{vanKeken1997}, \citet{Tackley2003} and
-\citet{Thieulot2014}. Fig. \ref{fig:rayleigh} shows the evolution of the experiment at different time steps. All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Rayleigh_Taylor_experiment/ISOVISCOUS}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/RT.pdf}
-\caption{$v_{\textrm{rms}}$ of the Rayleigh-Taylor experiment as function of time for different resolution of the grid.}
-\label{fig:RT}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=300px]{./Figures/Rayleigh.pdf}
-\caption{Evolution of the Rayleigh-Taylor experiment for a grid of \num{256x256} elements at $t=0, 500, 1000$ and $2000$ (panels a, b, c and d, respectively).}
-\label{fig:rayleigh}
-\end{figure}
-
-\subsection{Falling blocks}\label{sec:block}
-This benchmark is described as proposed by \citet{Gerya2003a}, \citet{Gerya2010b} and \citet{Thieulot2011}. The domain is square with $L_x=L_y=$ \SI{500}{\km}
-and the grid is composed by \num{50x50} elements with 25 markers in each element. The block is initially centred at ($x=$ \SI{250}{\km}; $y=$ \SI{400}{\km})
-and has a size of \num{100x100} km (Fig. \ref{fig:block}a). The fluid surrounding the block has $\eta_f=$ \SI{e21}{\pascal\per\s} and $\rho_f=$
-\SI{3200}{\kg\per\cubic\m}. The benchmark tests with different viscosities of the block, with $\eta_b$ from \SIrange{e15}{e27}{\pascal\per\s}. In each
-experiment also the density $\rho_b$ of the block varies from \SIrange{3220}{9900}{\kg\per\cubic\m}. Velocity boundary conditions are set to free slip
-conditions on all sides of the domain. Density distributions at $t=$ \SI{20}{\mega\year} for $\rho_b=$ \SI{3300}{\kg\per\cubic\m} and different $\eta_b$
-are plotted in Fig. \ref{fig:block}b-f, showing that the code correctly preserve the block geometry in case of large viscosity contrast, with a block stiffer
-than the surrounding fluid (Fig. \ref{fig:block}f).
-
-The velocity in the centre of the falling block at $t=0$ is measured for all experiments and, since the velocity should increase with density contrast, the
-quantity $\bm{v}/(\rho_b-\rho_f)$ is plotted as function of the viscosity contrast. All results perfectly match with those from \citet{Gerya2010b} and they
-line up on a single curve, demonstrating that the code can correctly deal with large viscosity and density contrasts (Fig. \ref{fig:falling}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Falling%20blocks}.
-
-\begin{figure}
-\noindent\includegraphics[width=300px]{./Figures/Block.pdf}
-\centering
-\caption{Density evolution of the falling block experiment at $t=$ \SI{0}{\mega\year} (panel a) and $t=$ \SI{20}{\mega\year} for different viscosities of the
-block (panels b-f).}
-\label{fig:block}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Falling.pdf}
-\caption{Initial velocity relative to the density contrast at the centre of the falling block as function of the viscosity contrast between the block and the
-surrounding fluid.}
-\label{fig:falling}
-\end{figure}
-
-\subsection{2D time-dependent Stokes sphere with deformable free surface}\label{sec:sphere}
-This experiment is performed in a unit square domain with the gravity acceleration fixed to $g_y=-1$. The sphere has $\rho_s=2$ and $\eta_s=10^3$ and is
-initially centred at ($0.5;0.6$) with radius $R=0.123456789$. The fluid surrounding the sphere has $\rho_f=1$ and $\eta_f=1$ and occupies the domain for
-$y\leq 0.75$, while for $y>0.75$ the air has $\rho_a=0$ and $\eta_a=10^{-3}$. Velocity boundary conditions are set to free slip on all sides. The Courant
-number is set to 0.25. The experiments run for 200 s using grid resolutions from \num{150x150} to \num{512x512} elements, with 25 randomly distributed markers
-per element, and for different average schemes for the viscosity. The interface between the fluid and the air is tracked by means of the markers chain. Results
-in terms of velocity and pressure fields and topography variations are compared with results obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and they can be found at
-\url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere_fs2D} and
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Time_dependent_sphere}. Results in terms of the $v_{\textrm{rms}}$ in case of 
-an arithmetic average are also shown in Fig. \ref{fig:inst_spherefs}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_arithm.pdf}
-\caption{$v_{\textrm{rms}}$ for different grid resolutions as function of time for the 2D time-dependent Stokes sphere experiment in case of an arithmetic average.
-Results are compared with results obtained by ASPECT.}
-\label{fig:inst_spherefs}
-\end{figure}
-
-\subsection{Free surface stabilisation}\label{sec:stab}
-The benchmark is performed as discussed in \citet{Kaus2010a} and \citet{Thieulot2014}. The domain is square with $L_x=L_y=$ \SI{500}{\km} and a grid resolution
-of \num{200x200} elements, each of them containing 16 markers. A buoyant fluid with $\rho_1=$ \SI{3200}{\kg\per\cubic\metre} and $\eta_1=$ \SI{e20}{\pascal\s}
-is overlain by a denser fluid with $\rho_2=$ \SI{3300}{\kg\per\cubic\metre} and $\eta_2=$ \SI{e21}{\pascal\s}. The initial interface between the fluids has an
-initial sinusoidal shape of 5 km amplitude. Velocity boundary conditions are set to free slip conditions at the sides and no slip at the bottom, while the top
-is free surface. The experiment is performed with various fixed time steps, without the use of the Courant number. The vertical position of the free surface at
-$x=L_x$ is tracked for each simulation, with or without the stabilisation algorithm.
-
-The results show an instability (drunken sailor effect) increasing the time step in case that the stabilisation algorithm is not activated
-($dt=$ \SI{4200}{\year}, green line in Fig. \ref{fig:stab}), as already observed by \citet{Kaus2010a} and \citet{Thieulot2014}. Activating the algorithm the
-instability is fixed and the simulation remains stable using time step up to \SI{50000}{\year} (red line in Fig. \ref{fig:stab}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Stabilisation_algorithm}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Kaus_random.pdf}
-\caption{Evolution of the y coordinate in $x=L_x$ as function of time for different time steps, with and without the stabilisation algorithm.}
-\label{fig:stab}
-\end{figure}
-
-\subsection{Topography relaxation}\label{sec:crameri}
-This experiment is performed as in \citet{Crameri2012} using free surface and sticky air (Case 1). The domain is rectangular with $L_x=$ \SI{2800}{\km} and
-$L_y=$ \SI{700}{\km} in case of free surface and $L_y=$ \SI{800}{\km} in case of sticky air ($h_{st}=$ \SI{100}{\km}). The grid is composed by \num{256x64}
-elements for the free surface \num{560x320} elements for the sticky air. For both cases, there is a mantle of 600 km thickness with $\eta_m=$ \SI{e21}{\pascal\s}
-overlain by a highly viscous cosine-shaped lithosphere with thickness between 93 and 107 km and $\eta_l=$ \SI{e23}{\pascal\s}. Both mediums have $\rho=$
-\SI{3300}{\kg\per\cubic\metre}. In case of sticky air experiment, the air has $\rho_a=$ \SI{0}{\kg\per\cubic\metre} and $\eta_a=$ \SI{e18}{\pascal\s}. The
-gravity acceleration is set to $g_y=$ \SI{-10}{\metre\per\square\s}. Velocity boundary conditions are set to no slip at the bottom and free slip at the sides
-of the domain. In case of sticky air, velocities on top are set to free slip conditions.
-
-The maximum topography as function of time ($h(t)$) can be analytically derived in according to \[h(t)=h_0 \exp(\gamma t)\]
-where $h_0=7$ km is the initial topography, $\gamma=\num{-0.2139e-11}$ is the characteristic relaxation rate and $t$ is the time (black line in Fig.
-\ref{fig:crameri}). Maximum topography at the characteristic relaxation time $t_{rlx}=$ \SI{14.825}{\kilo\year} can be found to be $h_{rlx}=$ \SI{2576}{\m}.
-The results show that a viscosity of the air of \SI{e18}{\pascal\s} is needed in case of $h_{st}=$ \SI{100}{\km} to correctly describe the topography relaxation
-for this problem, as pointed out by \citet{Crameri2012} (purple line in Fig. \ref{fig:crameri}). The case with a true free surface well follows the analytical
-solution (orange line in Fig. \ref{fig:crameri}). In particular, at the relaxation time in case of the true free surface predicts a maximum topography of
-\SI{2570}{\m}, with an error of \SI{6}{\m} with respect to the analytical solution. Results are also compared with results obtained with a free surface in
-UNDERWORLD (\num{256x64} elements) and SULEC (\num{401x201} elements) (with courtesy from \citealp{Crameri2012}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Topography_relaxation}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Crameri.pdf}
-\caption{Maximum topography as function of time for the Crameri benchmark (coloured lines) in comparison with the analytical solution (black line) and results
-obtained with UNDERWORLD (yellow line) and SULEC (blue line).}
-\label{fig:crameri}
-\end{figure}
-
-\subsection{Spontaneous subduction}\label{sec:subduction}
-This experiment is performed as presented by \citet{Schmeling2008}, in case of both a sticky air and a true free surface and using different average schemes
-for the viscosities. The domain is rectangular with $L_x=$ \SI{3000}{\km} and $L_y=$ \SI{700}{\km}, a grid resolution of \num{256x64} elements with 50 markers
-per element and Courant number of 0.01. At the beginning of the simulation, a 100 km-thick lithospheric layer with $\rho_m=$ \SI{3300}{\kg\per\cubic\metre}
-and $\eta_m=$ \SI{e23}{\pascal\s} is located at the top of the domain for $x$ between 1000 and 3000 km, while the asthenospheric mantle has $\rho_m=$
-\SI{3200}{\pascal\s} and $\eta_m=$ \SI{e21}{\pascal\s}. In addition, a 200 km-depth lithospheric slab is already subducted in the mantle in order to have
-a spontaneous subduction. In case of sticky air, $L_y=750$ km and a 50 km-thick air layer with $\eta_a$ from \SIrange{e19}{e21}{\pascal\s} overlie the mantle.
-Velocity boundary conditions are set to free slip on the sides and on the bottom of the domain. In case of sticky air, velocities on the top boundary are set
-to free slip conditions as well.
-
-The maximum depth of the slab is tracked for all the simulations (Fig. \ref{fig:subduction}) and they are compared with results obtained by
-\citet{Schmeling2008} with models with regular grids and comparable grid resolutions (rectangles in Fig. \ref{fig:subduction}). Low viscous sticky air models
-(i.e., \SI{e19}{\pascal\s}) enlighten a strong dependence of the sinking velocity with the chosen average scheme (dashed lines in Fig. \ref{fig:subduction}),
-while high viscous sticky air models (i.e., \SI{e21}{\pascal\s}) show a higher resistance at the trench, underestimating the correct solution (dotted lines in
-Fig. \ref{fig:subduction}). A high resistance at the trench can be also observed in case of a true free surface for the low resolution grid used (continuous
-lines in Fig. \ref{fig:subduction}), as pointed out by \citet{Schmeling2008}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Spontaneous_subduction}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Subduction.pdf}
-\caption{Maximum depth of the slab as function of time for all the simulations of the \citet{Schmeling2008} benchmark, in case of harmonic, geometric and
-arithmetic means (red, green and blue lines, respectively) using a sticky air or a true free surface (discontinuous and continuous lines, respectively).
-Red, green and blue rectangular areas indicate the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{500}{\km} in case of sticky air,
-while the grey rectangular area indicates the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{600}{\km} in case of true free surface.}
-\label{fig:subduction}
-\end{figure}
-
 \subsection{The slab detachment}\label{sec:slab}
 The slab detachment benchmark is performed as described by \citet{Schmalholz2011} and \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{1000}{\km}
-and $L_y=$ \SI{660}{\km} and a grid resolution of \num{256x256} elements. A non-linear viscous T-shaped layer with $\rho_l=$ \SI{3300}{\kg\per\cubic\metre} is
+and $L_y=$ \SI{660}{\km} and a grid resolution of $256\times256$ elements. A non-linear viscous T-shaped layer with $\rho_l=$ \SI{3300}{\kg\per\cubic\metre} is
 placed at the top of the domain and surrounded by a linear viscous fluid with $\rho_f=$ \SI{3150}{\kg\per\cubic\m}. The top layer is 80 km-thick and a 250
@@
 \end{figure}
-
-\subsection{The indenter experiment}\label{sec:indenter}
-The indenter benchmark simulate a rigid punch on a purely plastic von Mises material, for which exists an analytical solution
-\citep{Thieulot2008,Thieulot2014,Glerum2018}. The domain is a unit square with a grid resolution of \num{256x256} elements. The gravity acceleration is set
-to $\bm{g}=\bm{0}$ and effective viscosity is capped using $\eta_{min}=10^{-4}$ and $\eta_{max}=10^3$. The tolerance for the convergence of the non-linear
-solution is $tol=10^{-9}$, with a maximum number of non-linear iterations set to 500. The medium in the domain has $\rho=0.01$, an initial viscosity
-$\eta_0=10$, cohesion $C=1$ and an angle of internal friction $\phi=0$°. Velocity boundary conditions are set to no slip at the bottom and free slip at the
-sides of the domain. The top of the domain is open with the exception of the central portion (punch area) with width $w_p=0.125$, where $v=-1.05$ and $u$ is
-fixed either to 0 (rough punch experiment) or free (smooth punch experiment).
-
-Fig. \ref{fig:indenter} show the results in terms of viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure
-(panels g and h) for both experiments. The analytical solution indicate that pressure at ($0.5;1$) and ($0.5 \pm w_p;1$)is $p=\pi +1$ and $p=1$, respectively
-(black and red lines in Fig. \ref{fig:smooth_p}). Fig. \ref{fig:smooth_p} shows a clear improvement in the pressure solution when passing from a rough to a
-smooth punch (green and blue lines in Fig. \ref{fig:smooth_p}, respectively), as previously observed by \citet{Thieulot2014} and \citet{Glerum2018}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Indenter}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Indenter.pdf}
-\caption{Viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure (panels g and h) fields for rough (left column) and
-smooth (right column) punch experiments.}
-\label{fig:indenter}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Smooth_Rough.pdf}
-\caption{Nodal pressure in function of the x coordinate at $y=1$ for rough (green line) and smooth (blue line) punch experiment. Black and red lines indicate
-the analytical solution at $x=0.5$ and $x=0.5 \pm w_p$, respectively.}
-\label{fig:smooth_p}
-\end{figure}
-
-\subsection{The brick experiment}\label{sec:brick}
-An instantaneous version of the brick benchmark is performed to verify the correctness of the pressure-dependent plasticity for different angle of internal
-friction, as proposed by \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{40}{\km} and $L_y=$ \SI{10}{\km} and a grid resolution of \num{512x128}
-elements. The gravity acceleration is set to $g_y=$ \SI{-10}{\m\square\s} and effective viscosity is capped using $\eta_{min}=$ \SI{e19}{\pascal\s} and
-$\eta_{max}=$ \SI{e26}{\pascal\s}. The tolerance for the convergence of the non-linear solution is $tol=10^{-7}$, with a maximum number of non-linear iterations
-set to 1000. A 800 m-wide and 400 m-height inclusion with $\eta_b=$ \SI{e20}{\pascal\s} and $\rho_b=$ \SI{2700}{\kg\per\cubic\m} is placed at the bottom of the
-domain at $x=L_x/2$. The inclusion is surrounded by a non-linear viscous medium with $\rho_m=$ \SI{2700}{\kg\per\cubic\m}, an initial viscosity of
-\SI{e23}{\pascal\s}, a linear viscous viscosity of \SI{e25}{\pascal\s} and a cohesion of \SI{40}{\mega\pascal}. Velocities are set to free slip conditions at
-the bottom of the domain and the top is open. Velocities on sides of the domain are fixed to $u=$ \SI{+-2e-11}{\m\per\s} and $v=0$. The experiment is performed
-in compressional and extensional contexts with an angle of internal friction $\phi$ of the non-linear medium variable between 0° and 30°.
-
-As expected, two shear bands stem from the inclusion with variable angles in relation to both the dynamics context and the internal friction angle. Shear band
-angles formed at 45° for $\phi=0$° in both extensional and compressional contexts (Fig. \ref{fig:brick_beam}a), while they are different in case of $\phi\neq0$°.
-Shear band angles in case of $\phi=20$° are shown for compression and extension in Fig. \ref{fig:brick_beam}b  and c, respectively. Values of shear band angles
-as function of internal friction angles from 0° to 30° are extracted at different depths and minimum and maximum values are plotted in Fig. \ref{fig:brick},
-in comparison with theoretical Roscoe, Arthur and Coulomb shear band angles. For all tests the tolerance for nonlinear convergence is defined $tol=10^{-7}$
-and none converges before the maximum number of iterations ($it_{max}=1000$) is reached. However, tests with internal friction angles up to 15° show a constant
-decrease of the velocity residuals, which can not be observed for for tests with higher internal friction angles (Fig. \ref{fig:convergence}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Brick_experiment}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Brick_Beam.pdf}
-\caption{Shear band angles predicted for the brick experiment for $\phi=0$° (panel a) and for $\phi=20$° in case of compression and extension (panels b and c,
-respectively).}
-\label{fig:brick_beam}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Brick.pdf}
-\caption{Shear band angles predicted for the brick experiment in case of compressional and extensional contexts as function of different internal angle of
-friction (continuous black line and red dots), compared with theoretical Roscoe, Artur and Coulomb angles (discontinuous black lines). Red lines indicate the
-range of angles calculated at different depths.}
-\label{fig:brick}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \noindent\includegraphics[width=500px]{./Figures/convergence.pdf}
-  \caption{Convergence rates for all tests. Normalised residuals (left column) are compared with velocity residuals (right column) for extensional (first row)
-  and compressional (second row) settings.}
-  \label{fig:convergence}
-  \end{figure}
-
-\subsection{Advection stabilisation}\label{sec:advection}
-The 1D advection problem proposed by \citet{Donea2003} and \citet{Thieulot2011} is performed to verify the effectiveness of the SUPG method to stabilise the
-advection term of the energy equation. The domain is a 1D segment with $L_x=1$ composed by 50 elements and a discontinuity in the thermal field placed at
-$x=0.25$. Temperature is set to 1 for $x \leq 0.25$ and to 0 for $x > 0.25$. Velocity is set to ${u}=1$ in the entire domain. The simulation is performed for
-250 time steps, with $dt=0.002$, so the thermal discontinuity should be at $x=0.75$ at the end of the simulation.
-
-Temperature profiles at the end of the simulation are shown in \ref{fig:advection} as function of the dimensionless coefficient $\gamma=\tau \bm{v}/h$. In case
-of the classic Galerkin method ($\gamma=0$, blue line in \ref{fig:advection}) the final thermal profile is characterised by strong oscillations, which are
-eliminated in case of the SUPG method ($\gamma=0.045$, orange line in \ref{fig:advection}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Advection_stabilisation}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Advection.pdf}
-\caption{Temperature profile as function of \textit{x} for the advection stabilisation benchmark. Purple line indicates the initial temperature profile; the
-green line indicates the analytical temperature profile after 250 time steps; blue line indicates the temperature profile after 250 time steps in case of the
-classic Galerkin method ($\gamma=0$); orange line indicates the temperature profile after 250 time steps in case of the SUPG method ($\gamma=0.045$).}
-\label{fig:advection}
-\end{figure}
-
-\subsection{Shear heating}\label{sec:simple_shear}
-This exercise is performed in an unit square domain composed by \num{128x128} elements. The velocity field is prescribed on the entire domain with
-$\bm{v}=(L_y-y)y\bm{e_x}$; viscosity, density and specific heat are fixed to 1, while thermal conductivity and radiogenic energy are fixed to 0. Therefore,
-the energy equation (Eq. \ref{eq:energy}) can be simplified as \[\frac{\partial T}{\partial t}=H_s\] and fixing $T(t=0)=0$, the temperature can be find as
-\[T(t)=H_s t\]
-In this case we have
-\begin{eqnarray}
-\dot{\epsilon_{xx}}&=&\frac{\partial u}{\partial x}=0 \nonumber \\
-\dot{\epsilon_{yy}}&=&\frac{\partial v}{\partial y}=0 \nonumber \\
-\dot{\epsilon_{xy}}&=&\frac{1}{2}\left(\frac{\partial v}{\partial x}+\frac{\partial u}{\partial y}\right)=\frac{1}{2}(L_y-2y)\nonumber
-\end{eqnarray}
-and, simplifying Eq. \ref{eq:shear}, shear heating can be calculated as \[H_s(x,y)=(1-2y)^2\]
-
-The solution predicted by the model in terms of velocity, temperature and shear heating match well with the analytical solutions (Fig. \ref{fig:analytical_en}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Simple_shear}.
-
-%\begin{figure}
-%\centering
-%\noindent\includegraphics[width=400px]{./Figures/Energy_Thieulot.pdf}
-%\caption{Temperature and shear heating predicted by the model after 1 time step.}
-%\label{fig:en_thieulot}
-%\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=500px]{./Figures/Analytical_Shear.pdf}
-\caption{Velocity (panel a), temperature (panel b) and shear heating (panel c) as function of the \textit{y} coordinate for the simple shear experiment. The
-solutions predicted by the model (red lines) are compared with the analytical solutions (dashed blue lines).}
-\label{fig:analytical_en}
-\end{figure}
-
-\subsection{Shear and adiabatic heating}\label{sec:shear}
-This problem is performed as presented in Exercise 9.4 in \citet{Gerya2010b}. Two materials are vertically separated in a rectangular domain with $L_x=$
-\SI{1000}{\km}, $L_y=$ \SI{1500}{\km} and a grid resolution of \num{30x20} elements. Constant thermal coefficient expansion ($\alpha=$ \SI{3e-5}{\per\kelvin}),
-temperature ($T=$ \SI{1300}{\kelvin}) and gravity acceleration ($g_y=$ \SI{-10}{\m\square\s}) are assumed in the whole domain. Fluid 1 (on the left side of the
-domain) has $\rho_1=$ \SI{3200}{\kg\per\cubic\m} and $\eta_1=$ \SI{e20}{\pascal\s}; fluid 2 (on the right side of the domain) has $\rho_2=$
-\SI{3300}{\kg\per\cubic\m} and $\eta_2=$ \SI{e22}{\pascal\s}. Velocity boundary conditions are set to free slip on all sides of the domain.
-
-As shown in Fig. \ref{fig:energy}, both shear and adiabatic heating predicted by the code (first row in Fig. \ref{fig:energy}) well recreate the results
-obtained by means of example ${Shear\_adiabatic\_heating.m}$ from \citet{Gerya2010b} (second row in Fig. \ref{fig:energy}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Adiabatic%2BShear}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Energy.pdf}
-\caption{Comparison between shear (first column), adiabatic (second column) and total (third column) energy predicted by the code (panels a, b and c) and those
-created using example ${Shear\_adiabatic\_heating.m}$ from Exercise 9.4 in \citet{Gerya2010b} (panels d, e and f).}
-\label{fig:energy}
-\end{figure}
-
-\subsection{Mantle convection}\label{sec:mantle}
-This problem is performed as presented by \citet{Blankenbach1989} (constant viscosity cases) and \citet{Thieulot2014}, in a 2D unit square domain with gravity
-acceleration $g_y=10^{10}Ra$. The experiment is performed with three different Rayleigh numbers ($Ra=10^4$, $10^5$ and $10^6$) and with different grid
-resolution (between \num{32x32} and \num{128x128} elements). The fluid has constant viscosity, initial density, heat capacity, thermal conductivity
-($\eta= \rho_0= C_p = k=1$), reference temperature ($T_0=0$) and thermal expansion coefficient ($\alpha=10^{-10}$). Temperatures are set to 0 on top and 1 on
-bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\]
-
-The solution generated by the code in terms of $v_{\textrm{rms}}$ (calculated as in Sec. \ref{sec:error}) and Nu as function of time are reported for all the
-simulations in Fig. \ref{fig:mantle}. At the steady state, $v_{\textrm{rms}}$ and Nu of all simulations converge well toward the values from \citet{Blankenbach1989},
-with lower errors for higher resolution grids (Table \ref{tab:mantle}). All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Mantle_convection}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Convection.pdf}
-\caption{$v_{\textrm{rms}}$ (panels a, c and e) and Nu (panels b, d and f) for the mantle convection benchmark as function of time for different grid resolution.
-Panels a and b show the results for $Ra=10^4$; Panels c and d show the results for $Ra=10^5$; panels e and f show the results for $Ra=10^6$. Purple lines
-indicate the convergence values for the $v_{\textrm{rms}}$ and the Nu from \citet{Blankenbach1989} at the steady state.}
-\label{fig:mantle}
-\end{figure}
-
-\begin{table}[H]
-\caption{Comparison between $v_{\textrm{rms}}$ and Nu predicted by the code for the mantle convection experiment and same values as reported in literature.}
-\centering
-\small
-\begin{tabular}{|>{\centering}m{0.09\textwidth}|>{\centering}m{0.05\textwidth}|>{\centering}m{0.24\textwidth}|>{\centering}m{0.15\textwidth}|
-  >{\centering}m{0.1\textwidth} >{\centering}m{0.1\textwidth} >{\centering\arraybackslash}m{0.1\textwidth}|}
-\toprule
-& & \multirow{2}{*}{\citet{Blankenbach1989}} & \citet{Thieulot2014} & \multicolumn{3}{c|}{FALCON}  \\
- & & & \num{200x200} & \num{32x32} & \num{64x64} & \num{128x128}  \\
-\midrule
-\multirow{2}{*}{$Ra=10^4$} & $v_{\textrm{rms}}$ & $42.864947 \pm 0.000020$ & 42.867 & 42.83226 & 42.852793 & 42.861394  \\
-           & Nu      & $4.884409 \pm 0.000010$  & 4.882  & 4.781297 & 4.857475  & 4.877573   \\
-\hline
-\multirow{2}{*}{$Ra=10^5$} & $v_{\textrm{rms}}$ & $193.21454 \pm 0.00010$  & 193.255& 193.872643 & 193.377472 & 193.252290  \\
-           & Nu      & $10.534095 \pm 0.000010$ & 10.507 & 9.602514 & 10.270735  & 10.465629   \\
-\hline
-\multirow{2}{*}{$Ra=10^6$} & $v_{\textrm{rms}}$ & $833.98977 \pm 0.00020$  & 834.712& 848.091176 & 837.767911 & 834.945793  \\
-           & Nu      & $21.972465 \pm 0.000020$ & 21.695 & 15.999266 & 19.703682  & 21.306939   \\
-\bottomrule
-\end{tabular}
-\label{tab:mantle}
-\end{table}
-
-\subsection{Viscoplastic mantle convection}\label{sec:visco_mantle}
-This benchmark is performed as Cases 1-5a described in \citet{Tosi2015} in a 2D unit square domain with two different grid resolutions of \num{32x32} 
-and \num{100x100} elements. Reference viscosity, reference density, heat capacity, thermal conductivity and thermal expansion coefficient are fixed to a constant
-value of 1 ($\eta_0= \rho_0= C_p = k=\alpha=1$), while the gravity has been chosen as $g_y=10^2$, in order to obtain a Rayleigh number $Ra=10^2$. Temperatures
-are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\].
-The viscosity field $\eta$ is calculated as the harmonic average between a linear part $\eta_{lin}$ that depends on temperature only or on temperature and 
-depth $y$ and a nonlinear-plastic part $\eta_{pl}$ that depends on the strain rate $\dot{\bm{\epsilon}}$, as follows
-\[\eta(T,y,\dot{\bm{\epsilon}})=2\left(\frac{1}{\eta_{lin}(T,y)}+\frac{1}{\eta_{pl}(\dot{\bm{\epsilon}})}\right)^{-1}\]
-The linear and the nonlinear-plastic parts of the viscosity are calculated following \citet{Tosi2015} as follows
-\[\eta_{lin}(T,y)=\exp(-\ln(\Delta\eta_T)T+\ln(\Delta\eta_y)y)\]
-\[\eta_{pl}(\dot{\bm{\epsilon}})=\eta^*+\frac{\sigma_Y}{\sqrt{\dot{\bm{\epsilon}}:\dot{\bm{\epsilon}}}}\]
-where $\Delta\eta_T$, $\Delta\eta_y$, $\eta^*$ and $\sigma_Y$ are parameters chosen as listed in Table \ref{tab:tosi_setup}.
-
-For each case, some diagnostic quantities measured at the steady state are reported in Table \ref{tab:visco_mantle} (see Section \ref{sec:error} for 
-the computation of all the quantities), showing a well fit with results described in \citet{Tosi2015}. Data and charts for all cases can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Viscoplastic_mantle_convection}.
-
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{white!100!}{gray!25!}
-%\begin{table}[H]
-%  \caption{Results of the viscoplastic mantle convection benchmark for Cases 1-4 compared with results obtained with ELEFANT and YACC (from \citealp{Tosi2015}).}
-%  \centering
-%  \resizebox{\textwidth}{!}{
-%\begin{tabular}{|c|ccc|ccc|ccc|ccc|}
-%  \toprule
-%    & \multicolumn{3}{c|}{\textbf{Case 1}} & \multicolumn{3}{c|}{\textbf{Case 2}} & \multicolumn{3}{c|}{\textbf{Case 3}} &
-%    \multicolumn{3}{c|}{\textbf{Case 4}} \\
-%   & FALCON & ELEFANT& YACC & FALCON& ELEFANT& YACC & FALCON & ELEFANT & YACC & FALCON & ELEFANT & YACC \\
-%  \midrule
-%   $\left\langle T \right\rangle$ & 0.7759 & 0.7758 & 0.7767 & 0.6043 & 0.6036 & 0.6058 & 0.7275 & 0.7275 & 0.7286 & & 0.5277 & 0.5289\\
-%    Nu$^{top}$ & 3.4280 & 3.4214 & 3.4298 & 8.5242 & 8.5115 & 8.5278 & 3.0368 & 3.0347 & 3.0374 && 6.5912 & 6.5572 \\
-%    Nu$^{bottom}$ & 3.3168 & 3.313 & 3.3143 & 8.4599 & 8.4465 & 8.3990 & 3.0086 & 2.9908 & 2.9628 && 6.5834 & 6.5243 \\
-%    $v_{\textrm{rms}}$ & 249.901 & 249.134 & 251.7997 & 141.2451 & 140.7087 & 142.2020 & 100.2776 & 100.1208 & 100.9467 && 79.1105 & 79.6202 \\
-%    $v_{\textrm{rms}}^{top}$ & 1.8907 & 1.8642 & 1.8298 & 104.7928 & 104.1228 & 105.1767 & 2.0832 & 2.0652 & 2.0374 && 74.7596 & 75.8414 \\
-%    $u_{max}^{top}$ & 2.6340 & 2.6119 & 2.5516 & 121.9212& 121.766 & 122.3238 & 2.9086 & 2.9019 & 2.8458 && 88.9146 & 89.2940 \\
-%    $\eta_{min}$ & - & - & - & $2.0303\times10^{-5}$ & $2.0600\times10^{-5}$ & $1.9853\times10^{-5}$ & $4.8129\times10^{-5}$ & $4.8080\times10^{-5}$ & $4.7907\times10^{-5}$ &
-%    & $1.9860\times10^{-4}$ & $1.9174\times10^{-4}$ \\
-%    $\eta_{max}$ & - & - & - & 1.3325 & 1.85 & 1.6095 & 0.9360 & 0.9023 & 1 && 1.52 & 1.6773 \\
-%    $\left\langle W \right\rangle$ & 2.4299 & 2.4316 & 2.4583 & 7.5648 & 7.5665 & 7.6222 & 2.0387 & 2.0384 & 2.0400 && 5.6216 & 5.6512 \\
-%    $\left\langle \Phi \right\rangle/Ra$ & 2.4311 & 2.4276 & 2.4333 & 7.5661 & 7.5585 & 7.6194 & 2.0386 & 2.0370 & 2.0335 && 5.6182 & 5.6463 \\
-%    $\delta$ & 0.05\% & 0.16\% & 1.02\% & 0.02\% & 0.11\% & 0.04\% & $<$0.01\% & 0.07\% & 0.32\% && 0.06\% & 0.09\% \\
-%  \bottomrule
-%  \end{tabular}}
-%  \label{tab:visco_mantle}
-%  \end{table}
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{2}{white!100!}{gray!25!}
-\begin{table}[H]
-  \caption{Benchmark cases and corresponding parameters.}
-  \centering
-  \small
-\begin{tabular}{|>{\centering}m{0.08\textwidth}|>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}
-  >{\centering}m{0.05\textwidth} >{\centering}m{0.05\textwidth} |>{\centering\arraybackslash}m{0.2\textwidth}|}
-  \toprule
-    Case & $Ra$ & $\Delta\eta_T$ & $\Delta\eta_y$ & $\eta^*$ & $\sigma_Y$ & Convective regime \\
-  \midrule
-  \rowcolor{gray!25!}
-    1 & $10^2$ & $10^5$ & 1 & - & - & Stagnant lid \\
-    2 & $10^2$ & $10^5$ & 1 & $10^{-3}$ & 1 & Mobile lid  \\
-    \rowcolor{gray!25!}
-    3 & $10^2$ & $10^5$ & 10 & - & - & Stagnant lid  \\
-    4 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 1 & Mobile lid \\
-    \rowcolor{gray!25!}
-    5 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 4 & Periodic \\
-  \bottomrule
-  \end{tabular}
-  \label{tab:tosi_setup}
-  \end{table}
-
-\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{gray!25!}{white!100!}
-\begin{table}
-  \caption{Results of the viscoplastic mantle convection benchmark for different grid resolutions in Cases 1-5 compared with results obtained with ELEFANT
-  and YACC ($100\times100$ elements, from \citealp{Tosi2015}) and in Stone 28 ($32\times32$ elements, from
-  \url{https://github.com/cedrict/fieldstone/tree/master/python_codes/fieldstone_28}).}
-  \centering
-  \resizebox{\textwidth}{!}{
-\begin{tabular}{|cc|ccccccccccc|}
-  \toprule
-  \multicolumn{13}{|c|}{\textbf{Case 1}} \\
-  \midrule
-  Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-  & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-  \hline
-  \rowcolor{gray!25!}
-    YACC & $100\times100$ & 0.7767 & 3.4298 & 3.3143 & 251.7997 & 1.8298 & 2.5516 & - & - & 2.4583 & 2.4333 & 1.02\% \\
-    ELEFANT & $100\times100$ & 0.7758 & 3.4214 & 3.313 & 249.134 & 1.8642 & 2.6119 & - & - & 2.4316 & 2.4276 & 0.16\% \\
-    \rowcolor{gray!25!}
-    Stone 28 & $32\times32$ & 0.7737 & 3.3987 & - & 243.872 & - & - & - & - & - & - & - \\
-%    FALCON & $32\times32$ & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & - & - & 2.4745 & 2.4864 & 0.48\% \\
-%    \rowcolor{gray!25!}
-    FALCON & $100\times100$ & 0.7759 & 3.4280 & 3.3168 & 249.901 & 1.8907 & 2.6340 & - & - & 2.4299 & 2.4311 & 0.05\% \\
-    \midrule
-    \multicolumn{13}{|c|}{\textbf{Case 2}} \\
-    \midrule
-    Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-    & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-    \hline
-    \rowcolor{gray!25!}
-     YACC & $100\times100$ & 0.6058 & 8.5278 & 8.3990 & 142.2020 & 105.1767 & 122.3238 & $1.9853\times10^{-5}$ & 1.6095 & 7.6222 & 7.6194 & 0.04\% \\
-     ELEFANT & $100\times100$ & 0.6034 & 8.5115 & 8.4465 & 140.8067 & 104.1228 & 121.766 & $2.0600\times10^{-5}$ & 1.85 & 7.5665 & 7.5585 & 0.11\% \\
-     \rowcolor{gray!25!}
-     Stone 28 & $32\times32$ & 0.6052 & 8.1715 & - & 141.518 & - & - & - & - & - & - & - \\
-%     FALCON & $32\times32$ & 0.6454 & 8.1422 & 7.4020 & 156.6908 & 95.7226 & 116.9302 & $2.2572\times10^{-5}$ & 0.7194 & 7.4289 & 7.4452 & 0.22\% \\
-%     \rowcolor{gray!25!}
-     FALCON & $100\times100$ & 0.6044 & 8.5266 & 8.4590 & 141.3684 & 104.8867 & 122.0470 & $2.0333\times10^{-5}$ & 1.2914 & 7.5707 & 7.5718 & 0.01\% \\
-     \midrule
-     \multicolumn{13}{|c|}{\textbf{Case 3}} \\
-     \midrule
-     Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-     & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-     \hline
-     \rowcolor{gray!25!} 
-     YACC & $100\times100$ & 0.7286 & 3.0374 & 2.9628 & 100.9467 & 2.0374 & 2.8458 & $4.7907\times10^{-5}$ & 1 & 2.0400 & 2.0335 & 0.32\% \\
-      ELEFANT & $100\times100$ & 0.7275 & 3.0347 & 2.9908 & 100.1208 & 2.0652 & 2.9019 & $4.8080\times10^{-5}$ & 0.9023 & 2.0384 & 2.037 & 0.07\% \\
-      \rowcolor{gray!25!}
-      Stone 28 & $32\times32$ & 0.7277 & 3.0347 & - & 100.018 & - & - & - & - & - & - & - \\
-%      FALCON & $32\times32$ & 0.7281 & 3.0618 & 2.8132 & 101.0092 & 2.1630 & 3.0178 & $4.8500\times10^{-5}$ & 0.7845 & 2.0699 & 2.0695 & 0.02\% \\
-%      \rowcolor{gray!25!}
-      FALCON & $100\times100$ & 0.7275 & 3.0368 & 3.0086 & 100.2776 & 2.0832 & 2.9086 & $4.8129\times10^{-5}$ & 0.9360 & 2.0387 & 2.0386 & $<$0.01\% \\
-      \midrule
-      \multicolumn{13}{|c|}{\textbf{Case 4}} \\
-      \midrule
-      Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-      & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-      \hline
-      \rowcolor{gray!25!} 
-      YACC & $100\times100$ & 0.5289 & 6.5572 & 6.5243 & 79.6202 & 75.4814 & 89.2940 & $1.9174\times10^{-4}$ & 1.6773 & 5.6512 & 5.6463 & 0.09\% \\
-       ELEFANT & $100\times100$ & 0.5277 & 6.5912 & 6.5834 & 79.1105 & 74.7596 & 88.9146 & $1.9860\times10^{-4}$ & 1.5200 & 5.6216 & 5.6182 & 0.06\% \\
-       \rowcolor{gray!25!}
-       Stone 28 & $32\times32$ & 0.5286 & 6.4036 & - & 79.4746 & - & - & - & - & - & - & - \\
-%       FALCON & $32\times32$ & 0.5334 & 6.4679 & 6.3435 & 80.6153 & 76.0511 & 90.2058 & $2.0393\times10^{-4}$ & 0.6453 & 5.6676 & 5.6698 & 0.04\% \\
-%       \rowcolor{gray!25!}
-       FALCON & $100\times100$ & 0.5279 & 6.5897 & 6.5830 & 79.1220 & 75.0715 & 88.8484 & $1.9693\times10^{-4}$ & 1.4133 & 5.6127 & 5.6113 & 0.03\% \\
-       \midrule
-       \multicolumn{13}{|c|}{\textbf{Case 5}} \\
-       \midrule
-       Code & Resolution & $P$ & $\left\langle T \right\rangle_{min}$ & $\left\langle T \right\rangle_{max}$ & Nu$^{top}_{min}$ & Nu$^{top}_{max}$ &
-       $v_{\textrm{rms}}^{min}$ & $v_{\textrm{rms}}^{max}$ & \multicolumn{2}{c}{$(\left\langle \Phi \right\rangle/Ra)_{min}$} & 
-       \multicolumn{2}{c|}{$(\left\langle \Phi \right\rangle/Ra)_{max}$} \\
-       \hline
-       \rowcolor{gray!25!}
-         YACC & $100\times100$ & 0.0792 & 0.6543 & 0.6722 & 2.6950 & 7.2792 & 41.9813 & 98.8230 & \multicolumn{2}{c}{1.4347} & 
-         \multicolumn{2}{c|}{9.2748} \\
-         ELEFANT & $100\times100$ & 0.076 & 0.6514 & 0.6693 & 2.6808 & 7.3506 & 41.2578 & 100.089 & \multicolumn{2}{c}{1.3324} & 
-         \multicolumn{2}{c|}{9.5289} \\
-         \rowcolor{gray!25!}
-         Stone 28 & $32\times32$ & 0.0764 & 0.6521 & 0.6697 & 2.6912 & 7.0763 & 41.61 & 98.183 & \multicolumn{2}{c}{-} & 
-         \multicolumn{2}{c|}{-} \\
-     %    FALCON & $32\times32$ & & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & \multicolumn{2}{c}{-} & 
-     %     \multicolumn{2}{c|}{-} \\
-     %    \rowcolor{gray!25!}
-         FALCON & $100\times100$ & 0.0775 & 0.6519 & 0.6697 & 2.6847 & 7.3629 & 41.5256 & 99.5466 & \multicolumn{2}{c}{1.3399} & 
-         \multicolumn{2}{c|}{9.3993} \\
-  \bottomrule
-  \end{tabular}}
-  \label{tab:visco_mantle}
-  \end{table}
-
-\subsection{Thin layer entrainment}\label{sec:thin}
-This experiment is performed as originally proposed by \citet{vanKeken1997} with two incompressible fluids in a rectangular domain with $L_x=2$ and $L_y=1$
-and gravity acceleration $g_y=10^{10}Ra$. The Rayleigh number and the compositional Rayleigh number are fixed to $Ra=300000$ and $Rc=450000$, respectively.
-Both fluids have constant viscosity, thermal conductivity, specific heat ($\eta=\rho=C_p=1$) and thermal expansion coefficient ($\alpha=10^{-10}$). Fluid 1
-has a density $\rho_1=1$, while fluid 2 is denser ($\rho_2=\rho_1+1.5 \cdot 10^{-10}$) and is placed at the bottom of the domain, for $y \leq 0.025$.
-Temperature are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides of the domain. The initial
-temperature field is given by
-\[T(x,y)=T_u(x,y)+T_l(x,y)+T_r(x,y)+T_s(x,y)-\frac{3}{2}\]
-with
-\begin{eqnarray}
-T_u(x,y)&=&\frac{1}{2}erf\left(\frac{1-y}{2}\sqrt{\frac{u_0}{x}}\right)\nonumber \\
-T_l(x,y)&=&1-\frac{1}{2}erf\left(\frac{y}{2}\sqrt{\frac{u_0}{L_x-x}}\right)\nonumber \\
-T_r(x,y)&=&\frac{1}{2}+\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{y+1}} \exp\left(-\frac{x^2u_0}{4y+4}\right)\nonumber \\
-T_s(x,y)&=&\frac{1}{2}-\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{2-y}} \exp\left(-\frac{(L_x-x)^2u_0}{8-4y}\right)\nonumber
-\end{eqnarray}
-and
-\begin{eqnarray}
-u_0&=&\frac{L_x^{7/3}}{(1+L_x^4)^{2/3}}\left(\frac{Ra}{2\sqrt{\pi}}\right)^{\frac{2}{3}}\nonumber \\
-Q&=&2\sqrt{\frac{L_x}{\pi u_0}}\nonumber
-\end{eqnarray}
-
-The experiment is performed with two grid resolution (\num{125x40} and \num{200x80} elements), with 100 markers per element and CFL of 0.25. $v_{\textrm{rms}}$
-(see Sec. \ref{sec:error}) calculated for both the simulations match well with results from \citet{vanKeken1997} and obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and ELEFANT \citep{Thieulot2014}, obtained with similar grid resolutions
-(Fig. \ref{fig:thin}). In addition, a variety of simulations with different resolutions and aspect ratios are performed to verify that the sensitivity of the
-initial velocity field. Results match well with results from \citet{vanKeken1997} (in the grey area) and \citet{Thieulot2014} (Fig. \ref{fig:thin_initial}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Thin_layer}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_thin.pdf}
-\caption{$v_{\textrm{rms}}$ for the thin layer experiment as function of time for different grid resolution. Results are compared with results obtained by
-\citet{vanKeken1997} (black lines) and with ASPECT and ELEFANT.}
-\label{fig:thin}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrmszero.pdf}
-\caption{$v_{\textrm{rms}}$ at $t=0$ for the thin layer experiment as function of the element size for different aspect ratios of the elements. Results are compared
-with results obtained by \citet{vanKeken1997} (black dots in the grey area) and with ASPECT and ELEFANT.}
-\label{fig:thin_initial}
-\end{figure}
-
-\subsection{Sinking hydrated cylinder}\label{sec:quinquis}
-This experiment is performed as explained in \citet{Quinquis2014} and simulates the sinking of a cold, hydrated cylinder into a warm, dry mantle. The domain
-is square with $L_x=L_y=$ \SI{300}{\km}, a grid resolution of \num{300x300} elements and 25 markers per element. The sinking cylinder is located at $x=$
-\SI{150}{\km} and $y=$ \SI{170}{\km} with a radius of \SI{20}{\km} and has $\rho_c=$ \SI{3250}{\kg\per\cubic\m}, $\eta_c=$ \SI{e23}{\pascal\s}, a thermal
-conductivity $k_c=$ \SI{4.5}{\watt\per\m\per\kelvin},a specific heat $C_{p_c}=$ \SI{1250}{\joule\per\kg\per\kelvin}; the surrounding mantle has $\rho_m=$
-\SI{3200}{\kg\per\cubic\m}, $\eta_m=$ \SI{e20}{\pascal\s}, a thermal conductivity $k_m=$ \SI{105}{\watt\per\m\per\kelvin} and specific heat $C_{p_m}=$
-\SI{1250}{\joule\per\kg\per\kelvin}; a 58 km-thick lithosphere overlies the mantle and has $\rho_l=$ \SI{3200}{\kg\per\cubic\m}, $\eta_l=$ \SI{e23}{\pascal\s},
-a thermal conductivity $k_l=$ \SI{4.5}{\watt\per\m\per\kelvin}, specific heat $C_{p_l}=$ \SI{750}{\joule\per\kg\per\kelvin} and a thermal diffusivity of
-\SI{e-6}{\square\m\per\s}. Velocity boundary conditions are set to free slip on all sides of the domain. Temperature increase linearly in the lithosphere,
-from 0° to 1300°C, and in the mantle, from 1300° to 1360.5°C. The initial temperature of the cylinder is fixed to 400°C. Throughout the evolution, temperature
-is fixed to 0° and 1360.5°C at the top and bottom of the domain, respectively. The initial bound water content is imposed at 2 wt.\% in the cylinder and at
-0 wt.\% in both the mantle and the lithosphere. The maximum amount of water in the mantle is fixed to 0.2 wt.\%, while maximum water content in the cylinder
-and in the lithosphere is function of pressure and temperature and is calculated using a serpentinized harzburgite with Perple\_X, as in \citet{Quinquis2014}.
-
-As shown in \citet{Quinquis2014}, the model is characterised by a progressive dehydration in the external portion of the cylinder, due to the increase of
-temperature, with a consequent vertical migration of free water that hydrates the mantle above the cylinder, up to the lithosphere. The distribution of
-bound and free water during the evolution of the experiment is shown in Fig. \ref{fig:quinquis} (left and right columns, respectively). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Hydration}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Hydration.pdf}
-\caption{Distribution of bound and free water (left and right columns, respectively) at $t=0.5$ and $t=$ \SI{1.25}{\mega\year} (first and second rows,
-respectively) for the sinking hydrated cylinder experiment. Only markers with bound or free water not equal to 0 are plotted.}
-\label{fig:quinquis}
-\end{figure}
-
-\subsection{Mantle melting curves}\label{sec:katz}
-This instantaneous experiment is performed in order to recreate isobaric melting curves, as function of temperature, bulk water and content of water in the
-melt, to compare with melting curves obtained by \citet{Katz2003}. The domain is square with $L_x=L_y=100$ km, a grid resolution of \num{100x100} elements and
-25 markers per element. Density is set in the entire domain to \SI{3300}{\kg\per\cubic\m} and temperature increases from the sides to the centre from 1000° to
-1500°C. The experiment is performed for bulk water contents from 0 to 0.3 wt.\%.
-
-Fig. \ref{fig:fraction} shows isobaric melting curves for \SI{1}{\giga\pascal} (panel a) and \SI{3}{\giga\pascal} (panel b) calculated for different bulk water
-contents. The curves well match with those obtained for the mantle by \citet{Katz2003}, showing the saturation at the solidus for bulk water of 0.2 and 0.3 wt.\%, indicating
-that the code correctly determines melt fraction as function of pressure, temperature and bulk water content. Same results can be observed in Fig.
-\ref{fig:water}a, where isobaric and isothermal melting curves are shown as function of bulk water content. In addition, Fig. \ref{fig:water}b shows that the
-code correctly determines also the percentage of water in the melt as function of pressure, temperature and melt fraction. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Melting}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Fraction.pdf}
-\caption{Isobaric melting curves for 1 GPa (panel a) and 3 GPa (panel b) as function of temperature with different bulk water contents.}
-\label{fig:fraction}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Water.pdf}
-\caption{Isobaric melting curves for 1.5 GPa as function of bulk water (panel a) and water in the melt (panel b) calculated for different temperatures.}
-\label{fig:water}
-\end{figure}
-
-\clearpage
-
-\bibliography{export}
-
-\end{document}

### 2022-09-07 tex/Benchmark/sphere.tex
update main

--- 2022-03-31 FALCON_description.tex
@@
-\documentclass[hidelinks,10pt,a4paper]{article}
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{textcomp}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage{booktabs}
-\usepackage{multirow}
-\usepackage[cm]{fullpage}
-\usepackage[round]{natbib}
-\bibliographystyle{abbrvnat}
-\usepackage{doi}
-\usepackage{bm}
-\usepackage[table]{xcolor}
-\hypersetup{
-    colorlinks,
-    linkcolor={blue},
-    citecolor={blue},
-    urlcolor={blue}
-}
-\usepackage{float}
-\restylefloat{table}
-\usepackage{siunitx}
-\DeclareSIUnit\year{yr}
-
-\begin{document}
-
-\section{Introduction}\label{sec:intro}
-This document includes a description of the code FALCON, with a presentation of the features implemented followed by results of extensive benchmarks listed in
-Table \ref{tab:list}, performed to verify that the code solves correctly a wide range of problems. Description and results of all benchmarks are shown
-in Section \ref{sec:benchmark}.
-
-\begin{table}[H]
-\caption{List of benchmarks performed for each feature of the code. Data of all benchmarks can be found at 
-\url{https://github.com/aleregorda/Benchmarks}}
-\centering
-\begin{tabular}{| l | l | l |}
-\toprule
- \textbf{Feature}  & \textbf{Benchmark} & \textbf{Section} \\
-\midrule
-  Solver  & Stokes flow & \ref{sec:stokes} \\
-\hline
-  \multirow{2}{*}{Markers advection} & Zalesak disk & \ref{sec:runge} \\
-    & Conservative Interpolation Velocity & \ref{sec:CVI} \\
-\hline
-  \multirow{4}{*}{Momentum equation}  & Poiseuille flow & \ref{sec:poiseuille}\\
-    & Instantaneous 2D sphere & \ref{sec:ist_sphere} \\
-    & Rayleigh-Taylor experiment & \ref{sec:rayleigh} \\
-    & Falling block & \ref{sec:block} \\
-\hline
-    & 2D Stokes sphere & \ref{sec:sphere} \\
-    Sticky air and  & Stabilisation algorithm & \ref{sec:stab} \\
-    free surface & Topography relaxation & \ref{sec:crameri} \\
-  & Spontaneous subduction & \ref{sec:subduction} \\
-\hline
-  Erosion and sedimentation  & - & - \\
-\hline
-   Non-linear  & Slab detachment & \ref{sec:slab} \\
-   visco-plasticity & Indenter & \ref{sec:indenter} \\
-    & Brick & \ref{sec:brick} \\
-\hline
-  \multirow{3}{*}{Energy equation}  & Advection stabilisation & \ref{sec:advection} \\
-    & Simple shear heating & \ref{sec:simple_shear} \\
-    & Shear and adiabatic heating & \ref{sec:shear} \\
-\hline
-\multirow{3}{*}{Energy + momentum} & Mantle convection & \ref{sec:mantle} \\
-& Viscoplastic mantle convection & \ref{sec:visco_mantle} \\
-& Thin layer entrainment & \ref{sec:thin} \\
-\hline
-  Phase changes  & \multirow{2}{*}{Hydrated sinking cylinder} & \multirow{2}{*}{\ref{sec:quinquis}} \\
-  and hydration  &  & \\
-\hline
-  Melting & Experimental melting curves & \ref{sec:katz} \\
-\bottomrule
-\end{tabular}
-\label{tab:list}
-\end{table}
-
-The code is written in Fortran90 and uses the finite element method (FEM) with quadrilateral $Q_1 \times P_0$ elements (continuous bilinear velocity and
-discontinuous constant pressure), associated with MUMPS\footnote{\url{http://mumps.enseeiht.fr/}} \citep{Amestoy2001,Amestoy2019} that is a software package
-for solving systems of linear equations of the form $A \cdot x = b$, where \textit{A} is a square sparse matrix, by means of a direct method. Since
-$Q_1 \times P_0$ elements do not satisfy Ladyzhenskaya, Babuska and Brezzi (LBB) stability condition \citep{Donea2003}, elemental pressure is elaborated
-in post-processing to avoid spurious pressures by means of a double interpolation to smooth the pressure. Following this procedure, the elemental pressure
-is interpolated onto nodes and then back onto elements. After the smoothing procedure, the pressure field is calculated again on the nodes to be used in
-combination with the lithostatic pressure, which is calculated onto nodes. Correctness of the solution and performances of the solver in terms of time and
-memory usage are tested solving the Stokes flow with the analytical solution proposed by \citet{Donea2003} (Section \ref{sec:stokes}).
-
-The thermo-mechanics of crust-mantle systems is described by means of the conservation of mass, momentum and energy equations, expressed as follows:
-\begin{eqnarray}
-\label{eq:mass}\frac{\partial \rho}{\partial t}+\bm{\nabla} \cdot (\rho \bm{u})&=&0\\
-\label{eq:momentum}-\bm{\nabla} p + \bm{\nabla} \cdot \bm{\tau} + \rho \bm{g}&=&0\\
-\label{eq:energy}\rho C_p \left(\frac{\partial T}{\partial t} + \bm{u} \cdot \bm{\nabla} T \right)&=&\bm{\nabla} \cdot \left(k\bm{\nabla} T\right) + H_{tot}
-\end{eqnarray}
-where $\rho$ is the density, $\bm{u}$ is the velocity, \textit{p} is the pressure, $\bm{\tau}$ is the deviatoric stress, $\bm{g}$ is the gravity acceleration,
-$C_p$ is the specific heat at constant pressure, \textit{T} is the temperature, \textit{k} is the thermal conductivity and $H_{tot}$ is the total internal heat
-production. Density variations due to temperature are generally small enough to assume the density as constant ($\rho=\rho_0$) in Eq. \ref{eq:mass} and in
-Eq. \ref{eq:energy}, while it must be treated as a variable in the buoyancy term of Eq. \ref{eq:momentum}, such that \[\rho=\rho_0(1-\alpha(T-T_0))\]
-where $\rho_0$ is the density at a reference temperature $T_0$ and $\alpha$ is the coefficient of thermal expansion. Eq. \ref{eq:mass} then can be rewritten as
-\begin{equation}\label{eq:mass_0}
-\bm{\nabla} \cdot \bm{u}=0
-\end{equation}
-This simplification is known as the extended Boussinesq approximation.
-
-The time step $dt$ is chosen in according to Courant-Friedrichs-Lewy condition \citep{Anderson1995}, such that
-\[dt=C_n\; \min\left(\frac{\Delta x}{\max|\bm{u}|},\frac{(\Delta x)^2}{\kappa}\right)\]
-where $0<C_n<1$ is the Courant (CFL) number, $\Delta x$ is the minimum dimension of smallest element, $\max|\bm{u}|$ is the maximum velocity calculated on the
-entire domain and
-$\kappa=\frac{k}{\rho C_p}$ is the thermal diffusivity \citep{Thieulot2014}.
-
-\section{Lagrangian markers}\label{sec:markers}
-Elemental properties (density, viscosity, thermal conductivity, specific heat and thermal expansion) needed to solve Eqs. \ref{eq:mass_0}, \ref{eq:momentum}
-and \ref{eq:energy} are related to the composition of each element, determined by means of Lagrangian markers that are characteristic of different materials
-in the domain. Elemental properties (with the sole exception of the viscosity) are calculated using an arithmetic mean, as \[P_e=\frac{1}{n}\sum_{i=1}^n P_i\]
-where $P_e$ is the elemental property, $P_i$ is the property characteristic of the material of each marker and $n$ is the number of markers of the element.
-Differently, the average scheme for the elemental viscosity can be chosen between harmonic, geometric and arithmetic mean. Markers can be placed regularly or
-randomly at the beginning of the simulation and their advection is performed by either a 2nd-order or a 4th-order Runge-Kutta in space, interpolating the velocity 
-field on each marker by means of the shape functions (see Section \ref{sec:runge}). The interpolated velocity is then corrected by means of the Conservative
-Velocity Interpolation (CVI), which introduces a corrective term to reduce dispersion and clustering of particles in both steady state and time-dependent 
-\citep{Wang2015} (see Section \ref{sec:CVI}).
-
-The initial distribution of the markers is created using the open-source code library
-Geodynamic World Builder \citep{Fraters2019}. During the simulation, each marker carries memory of temperature, pressure and accumulated strain, which are
-determined interpolating the relative nodal parameter. In particular, pressure is interpolated onto each marker after the smoothing procedure. The number
-of markers contained in each element is maintained between $n_{min}$ and $n_{max}$. When in an element there are less markers than $n_{min}$ the code adds
-random markers to reach the $n_{min}$, while if the number is higher than $n_{max}$ some of them are randomly deleted. When new markers are added, they assume
-the properties of the nearest marker. In this way elements are never empty and maintain a number of markers inside a prefixed range.
-
-\section{The momentum equation}\label{sec:mom_eq}
-The penalty method is implemented so that Eq. \ref{eq:mass_0} can be written as:
-\begin{equation}\label{eq:mass_penalty}
-\bm{\nabla} \cdot \bm{u} + \frac{p}{\lambda}=0
-\end{equation}
-where $\lambda$ is the so-called penalty parameter, which should be 6-7 orders of magnitude larger than the shear viscosity to ensure that mass conservation
-is satisfied \citep{Donea2003,Thieulot2014}.
-
-The deviatoric stress tensor in Eq. \ref{eq:momentum} can be written in terms of the strain rate tensor as $\bm{\tau}=2\eta\dot{\bm{\epsilon}}$, with
-$\dot{\bm{\epsilon}}=\frac{1}{2}\left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)$. Therefore, Eq. \ref{eq:momentum} can be rewritten as:
-\begin{equation}\label{eq:momentum_pressure}
--\bm{\nabla} p + \bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right) + \rho \bm{g} = 0
-\end{equation}
-Finally, using pressure from Eq. \ref{eq:mass_penalty}, Eq. \ref{eq:momentum_pressure} can be rewritten as:
-\begin{equation}\label{eq:momentum_penalty}
-\lambda \bm{\nabla} \left(\bm{\nabla} \cdot \bm{v}\right)+\bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right)+\rho\bm{g}=0
-\end{equation}
-The penalty method is been associated to the iterative Uzawa method, as described in detail in \citet{Dabrowski2008} and \citet{Thieulot2014}
-(see Section \ref{sec:poiseuille}). Results of benchmarks performed to verify the correctness in the implementation of Eq. \ref{eq:momentum_penalty} are
-shown in Sections \ref{sec:ist_sphere} and \ref{sec:rayleigh}.
-
-To support large viscosities variations the penalty parameter is related to the effective elemental viscosity by means of a dimensionless coefficient, so that
-$\lambda=\lambda_e(e)\eta_{eff}(e)$ \citep{Marotta2006,Dabrowski2008,Thieulot2014}. Benchmark of the falling block \citep{Gerya2003a,Gerya2010b,Thieulot2011}
-is performed to verify that the code can correctly deal buoyancy driven flows with strong viscosity contrasts (Section \ref{sec:block}).
-
-\section{Sticky air and free surface}\label{sec:sticky}
-The Earth's surface can be treated by means of either the so-called sticky-air or a true free surface method, both of them implemented in the code. In the
-sticky-air method the surface is approximated with the introduction of a buoyant layer with a viscosity at least four orders of magnitude lower than the crust
-\citep{Schmeling2008,Crameri2012} and the interface between lithosphere and air is defined using a chain of passive markers that are advected as the Lagrangian
-markers. The correctness of the evolution of the markers chain is tested with the experiment of a 2D time-dependent Stokes sphere below a free surface and
-compared with results from ASPECT\footnote{\url{https://aspect.geodynamics.org/}} \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} (Section
-\ref{sec:sphere}). In the true free surface case the top boundary is assumed stress-free and velocities are not fixed. In this case topography variations are
-described by vertical deformations of the mesh that depend on the velocity field of the nodes that identify the free surface, while horizontal deformations are
-not taken into account. This procedure is known as the Arbitrary Lagrangian-Eulerian (ALE) method and its implementation follows the technique described in
-\citet{Thieulot2011}. However, although the implementation of a true free surface better reproduces laboratory experiment, extremely small time steps can be
-necessary to maintain stability \citep{Kaus2010a,Quinquis2011,Thieulot2014}. Therefore, the stability algorithm proposed by \citet{Kaus2010a} is implemented
-to avoid instabilities due to high density differences at the free surface when using too large time steps. The implementation of this algorithm are tested by
-performing the experiment described by \citet{Kaus2010a} (Section \ref{sec:stab}).
-
-The topography relaxation benchmark proposed by \citet{Crameri2012} is performed to verify that the code correctly recovers topography variations in case of
-both the sticky-air and the true free surface method (Section \ref{sec:crameri}). Finally, Section \ref{sec:subduction} show the results of the spontaneous
-subduction experiment described by \citet{Schmeling2008}.
-
-\section{Erosion and sedimentation}\label{sec:erosion}
-Surface processes at either the lithosphere-air interface or in correspondence of the free surface have been implemented by means of the software
-Fastscape\footnote{\url{https://fastscape.org/}} \citep{Braun2013,Cordonnier2019,Yuan2019a,Yuan2019b}. FastScapeLib is a set of routines that solve the stream
-power law (enriched by a sediment transport/deposition term) \citep{Yuan2019a}, hillslope diffusion and marine transport and deposition \citep{Yuan2019b}.
-The partial differential equation solved by FastScapeLib is:
-\begin{equation}
-\frac{\partial h}{ \partial t}=U-K_f A^{m_e}S^{n_e} + \frac{G}{A}\int_A \left( U-\frac{\partial h}{\partial t} \right) dA + K_d \bm{\nabla}^2 h 
-\end{equation}
-where $h$ is the topography, $U$ is the uplift, $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $S$ is the slope, $m_e$ and $n_e$ are
-parameters that should be chosen to constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}, $G$ is the dimensionless
-deposition/transport coefficient for the enriched stream power law and $K_d$ is the effective diffusivity. Terms $K_f A^{m_e}S^{n_e}$ and $K_d \bm{\nabla}^2 h$
-consider long-range processes, described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, and short-range processes,
-proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}.
-
-In case of sticky air, slope is determined by means of the markers chain, which is vertically corrected considering erosion and sedimentation rates. After the
-correction, continental markers above the markers chain are transformed in air markers and air markers below the markers chain become sediments. Similarly,
-slope is determined considering the nodes on the top boundary in case of true free surface and they are vertically corrected as for the markers chain. In this
-case, continental markers above the top boundary are simply deleted.
-%Surface processes at the lithosphere-air interface are implemented taking into account fluvial erosion (long-range processes), hillslope processes (short-range
-%processes) and failure of slopes exceeding a threshold angle $\theta$ \citep{Burov1997,Tucker1999,Sacek2012,Croissant2014,Ueda2015,Beucher2020}. Erosion rates
-%$\dot{E}_{lr}$ due to long-range processes can be described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{lr}=K_f A^{m_e} \bm{\nabla} h^{n_e}
-%\end{equation}
-%where $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $\bm{\nabla} h$ is the slope and $m_e$ and $n_e$ are parameters chosen in to
-%constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}. Erosion rates $\dot{E}_{sr}$ due to short-range processes are
-%implemented by assuming that transport is linearly proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{sr}=K_d \bm{\nabla}^2 h
-%\end{equation}
-%where $K_d$ is the effective diffusivity. In addition, the slope is modified if it still exceeds a threshold angle after long- and short-range processes
-%\citep{Ueda2015}.
-
-%Offshore sedimentation processes are implemented similarly to hillslope processes but with a depth-dependent coefficient \citep{Sacek2012,Beucher2020}.
-%The sedimentation rate $\dot{S}$ is calculated as
-%\begin{equation}
-%\dot{S}=K_s \bm{\nabla}^2 h
-%\end{equation}
-%with
-%\begin{equation}
-%K_s=K^*_s \exp\left(\frac{h-h_{sl}}{h_0}\right)
-%\end{equation}
-%where $K^*_s$ is a constant, $h_{sl}$ is the sea level and $h_0$ is a characteristic length. In case of sticky air, slope is determined by means of the markers
-%chain, which is vertically corrected considering erosion and sedimentation rates. After the correction, continental markers above the markers chain are
-%transformed in air markers and air markers below the markers chain become sediments. Similarly, slope is determined considering the nodes on the top boundary
-%in case of true free surface and they are vertically corrected as for the markers chain. In this case, continental markers above the top boundary are simply
-%deleted.
-
-\section{Non-linear rheologies}\label{sec:nonlinear}
-Non-linear rheologies are implemented combining viscous creep (dislocation and diffusion) and plastic yielding. For each marker, diffusion and dislocation
-viscosity ($\eta_{df|ds}$) can be determined as follows \citep{Karato1993,Warren2008b,Wang2016}:
-\begin{equation}\label{eq:viscous}
-\eta_{df|ds}=f_s \left(\frac{d^m}{A}\right)^{\frac{1}{n}} \left (\frac{2^{\frac{1-n}{n}}}{3^{\frac{n+1}{2n}}} \right)
-I_2^{\frac{(1-n)}{n}}\exp\left(\frac{Q+pV}{nRT}\right)
-\end{equation}
-where $f_s$ is a scaling factor used to represent lithologies that are stronger or weaker than the base set, $d$ is the grain size, $m$ is the grain size exponent,
-$n$ is the stress exponent, $A$ is the uniaxial pre-exponential factor, $I_2$ is the square root of the second invariant of the strain rate tensor, $Q$ is the
-activation energy, $p$ is the pressure, $V$ is the activation volume, $R$ is the gas constant and $T$ is the temperature.
-Pressures and temperatures are determined by the interpolation of the nodal parameters, while strain rates are calculated by means of the
-derivative of the velocities on the nodes. In case of diffusion creep $n=1$ and $m>0$, in case of dislocation creep $n>1$ and $m=0$. The viscous creep $\eta_{cp}$
-is then calculated as the harmonic average between $\eta_{df}$ and $\eta_{ds}$:
-\begin{equation}\label{eq:visco_creep}
-\eta_{cp}=\left(\frac{1}{\eta_{df}}+\frac{1}{\eta_{ds}}\right)^{-1}
-\end{equation}
-The implementation of the non-linear viscous creep viscosity depending on the strain rate is tested by means of the slab detachment benchmark
-(Section \ref{sec:slab}).
-
-Plastic yielding is implemented rescaling $\eta_{cp}$ in order to limit the stress below the yield stress $\sigma_y$ \citep{Thieulot2008,Thieulot2014,Glerum2018},
-obtaining
-\begin{equation}\label{eq:plastic}
-\eta_{pl}=\frac{\sigma_y}{2I_2}
-\end{equation}
-where $\eta_{pl}$ is the plastic viscosity and the yield stress is determined following the Drucker-Prager criterion, such as
-\begin{equation}\label{eq:yield}
-\sigma_y=C\;\cos(\phi)+p\;\sin(\phi)
-\end{equation}
-where $C$ is the cohesion and $\phi$ is the internal friction angle. In case of negative pressure it is imposed equal to 0, so that negative yield stress are
-excluded. The correctness of the non-linear solution in case of plastic viscosity with variable internal friction angle is verified performing indenter and
-brick experiments (Sections \ref{sec:indenter} and \ref{sec:brick}, respectively).
-
-Strain softening is taken into account for both viscous creep and plastic viscosity \citep{Huismans2003,Babeyko2005,Huismans2005,Sobolev2005,Warren2008a} by
-means of the accumulated strain $\epsilon$ memorised by each marker. Softening in the viscous creep determines a linear decrease of $\eta_{vc}$ by means of a
-viscous strain softening factor $W_S$ that increases linearly from $W_{S_0}$ to $W_{S_{\infty}}$ for $\epsilon_{S_0}<\epsilon_S<\epsilon_{S_{\infty}}$ 
-\citep{Warren2008a}. This viscous softening can be related to strain-induced grain size reduction \citep{Warren2008a}. Differently, plastic softening is 
-simulated with a linear decrease of internal friction angle $\phi(\epsilon)$ and cohesion $C(\epsilon)$ in according to
-\begin{eqnarray}
-\label{eq:friction1}\phi(\epsilon)&=&\phi_0+(\phi_{\infty}-\phi_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}\\
-\label{eq:friction2}C(\epsilon)&=&C_0+(C_{\infty}-C_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}
-\end{eqnarray}
-where $\phi_0$, $C_0$ and $\phi_{\infty}$, $C_{\infty}$ are internal friction angle and cohesion for $\epsilon_0$ and $\epsilon_{\infty}$, respectively
-\citep{Huismans2003,Huismans2005,Warren2008a,Thieulot2014}. Plastic softening approximates deformation-induced softening of faults and brittle shear zones
-\citep{Warren2008a}.
-
-Plastic yielding and viscous creep are then combined to obtain a viscoplastic viscosity $\eta_{vp}$ as follows, assuming that they are independent processes
-\citep{Karato2008,Glerum2018}:
-\begin{equation}\label{eq:viscoplastic}
-\eta_{vp}=\min(\eta_{cp},\eta_{pl})
-\end{equation}
-Finally, effective viscosity $\eta_{eff}$ is capped by the minimum and the maximum viscosity ($\eta_{min}$ and $\eta_{max}$, respectively) to avoid extremely
-low or high viscosity \citep{Glerum2018} as follows
-\begin{equation}\label{eq:effective}
-\eta_{eff}=\min(\max(\eta_{vp},\eta_{min}),\eta_{max})
-\end{equation}
-Viscous creep and plastic yielding are non-linear rheologies because of their dependence on the velocity field through pressure and strain rates. Therefore,
-the solution is determined by means of Picard-type iterations, until convergence of the velocity field \citep{Glerum2018}. The convergence is verified at each
-iteration $i$ via the nonlinear residual $\bm{\mathcal{R}}^i$ that can be determined as
-\begin{equation}\label{eq:convergence}
-\bm{\mathcal{R}}^i= \mathbb{K} (\eta_{eff}(\dot{\bm{\epsilon}}^{i-1},p^{i-1}))\cdot \bm{v}^{i-1} - \bm{f}^i
-\end{equation}
-%\begin{equation}\label{eq:convergence}
-%\chi=1-\frac{\langle (\bm{v}^i-\langle\bm{v}^i\rangle)\cdot (\bm{v}^{i+1}-\langle\bm{v}^{i+1}\rangle) \rangle}{\sqrt{\sigma^i\sigma^{i+1}}}
-%\end{equation}
-%Iterations are performed until either convergence ($\chi<tol$) or maximal number of non-linear iterations ($it_{max}$) is reached.
-where $\mathbb{K}$ is the Stiffness matrix and $\bm{f}$ is the right hand side vector \citep{Spiegelman2016,Glerum2018}.  The $L_2$-norm is extracted from
-$\bm{\mathcal{R}}^i$ and it is normalised as follows
-\begin{equation}\label{eq:convergence_res}
-\frac{||\bm{\mathcal{R}}^i||_2}{||\bm{\mathcal{R}}^0||_2}
-\end{equation}
-where $||\bm{\mathcal{R}}^0||_2$ is the $L_2$-norm of the first nonlinear iteration.
-Since often the normalised nonlinear residual drop very quickly over the first 2-3 iterations, also the $L_2$-norm of velocity and pressure residuals are
-calculated as
-\begin{equation}\label{eq:convergence_vel}
-\frac{||\Delta\bm{v}||_2}{||\bm{v}||_2}=\frac{||\bm{v}_i-\bm{v}_{i-1}||_2}{||\bm{v}_{i}||_2}
-\end{equation}
-and
-\begin{equation}\label{eq:convergence_press}
-  \frac{||\Delta p||_2}{||p||_2}=\frac{||p_i-p_{i-1}||_2}{||p_{i}||_2}
-\end{equation}
-respectively. Iterations are performed until either maximal number of non-linear iterations ($it_{max}$) is reached or all the normalised nonlinear residual, 
-the velocity residual and the pressure residual converge under a defined tolerance $tol$.
-
-\section{The energy equation}\label{sec:en_eq}
-The stabilisation of the advection term of the energy equation (Eq. \ref{eq:energy}) needed to avoid possible oscillations in the thermal solution in the case
-when advection dominates over diffusion, is implemented by means of a streamline-upwind Petrov–Galerkin (SUPG) method, for which the advection term is modified
-follows the discussion in \citet{Thieulot2011} and \citet{Thieulot2014} (see Section \ref{sec:advection} for benchmark). The total internal heat production
-$H_{tot}$ in Eq. \ref{eq:energy} includes radiogenic heating $H_r$, shear heating $H_s$ and adiabatic heating $H_a$, where
-\begin{equation}\label{eq:shear}
-H_s=2\eta \dot{\bm{\epsilon}} : \dot{\bm{\epsilon}} = 2\eta(\dot{\epsilon}_{xx}^2+\dot{\epsilon}_{yy}^2+2\dot{\epsilon}_{xy}^2)
-\end{equation}
-and
-\begin{equation}\label{eq:adiabatic}
-H_a=T \alpha \frac{Dp}{Dt} \approx -\alpha T \rho g_y v
-\end{equation}
-The correctness of the implementation and the computation of shear and adiabatic heating is verified performing an exercise with an analytical solution
-(Section \ref{sec:simple_shear}) and Exercise 9.4 in \citet{Gerya2010b} (Section \ref{sec:shear}). Furthermore, mantle convection and thin layer entrainment
-benchmarks (Sections \ref{sec:mantle} and \ref{sec:thin}, respectively) are performed to verify that the code solve correctly Eqs. \ref{eq:momentum_penalty}
-and \ref{eq:energy} in case of a temperature field characterised by an initial perturbation and with a temperature-dependent density. At last, the viscoplastic
-mantle convection benchmark is performed to check also the viscoplasticity in case of a initial perturbed temperature field (Section \ref{sec:visco_mantle}).
-
-\section{Phase transitions and hydration}\label{sec:phase}
-Variations of effective density, specific heat and coefficient of thermal expansion during the evolution of crustal and mantle materials at different
-pressure-temperature ($p$-$T$) conditions are computed by means of version 6.8.6 of Perple\_X software package \citep{Connolly2005}, similarly to the
-implementation described in \citet{Marotta2020}. By default, mineral assemblages and properties of each lithology (oceanic crust, lower and upper continental
-crust, sediments and dry or hydrated mantle) are calculated for temperatures from \SIrange{330}{1600}{\kelvin}, with increments of approximately \SI{4}{\kelvin},
-and pressures from \SIrange{0.1}{30}{\giga\pascal}, with increments of approximately \SI{0.03}{\giga\pascal}, for a total of almost 100000 points.
-
-The introduction of phase transitions produces two main effects that must be taken into account in numerical models: variations in density and release/absorption
-of latent heat \citep{Tackley2010}. Effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into
-account considering effective coefficient of thermal expansion \citep{Christensen1985,Zhong2015}. Similarly, effects of latent heat in the energy equation
-(eq \ref{eq:energy}) are taken into account by considering effective specific heat and coefficient of thermal expansion
-\citep{Christensen1985,Tackley2010,Zhong2015}. Effective density, specific heat and coefficient of thermal expansion are included in Perple\_X files and are
-assumed by each Lagrangian marker in according to its $p$-$T$ conditions.
-
-Hydration processes related to dehydration of subducting lithosphere \citep{Schmidt1998,Liu2007,Faccenda2009,Faccenda2010,Faccenda2014,Rosas2016} strongly
-influence the thermo-mechanics inside the mantle wedge, mainly because of both weakening effects on the mantle rheology and density variations in case of
-mantle serpentinization \citep{Gerya2002,Honda2003,Arcay2005,Roda2010,Regorda2017}. In the case that hydration is switched on, the amount of bound and free
-water is memorised by each marker, following the implementation of \citet{Quinquis2014}; hydration and dehydration processes are related to the amount of bound
-water of each marker and to the maximum amount of water it can transport, i.e. if the amount of bound water exceeds the maximum amount of water, the marker
-dehydrates and releases free water that can hydrates under-saturated markers. Maximum water content of each marker is determined as function of lithology and
-$p$-$T$ conditions and it is calculated using Perple\_X, in the same way than effective density, specific heat and coefficient of thermal expansion. Bound water
-is advected together with the markers, neglecting the effect of bound water diffusion, while free water simply migrates vertically and is not coupled to the
-solid-phase flow of the mantle wedge \citep{Arcay2005,Quinquis2014}. The correctness of the migration scheme of the free water and its absorption into bound
-water by under-saturated markers are verified performing the experiment described in \citet{Quinquis2014} (Sec. \ref{sec:quinquis}).
-
-The mantle viscosity weakening related to the amount of water has been extensively studied \citep{Chopra1981,Hirth2003} and it is implemented in according to
-\citet{Arcay2005} and \citet{Horiuchi2016}, as follows
-\begin{equation}\label{eq:wet_visc}
-\eta_{wet}=\eta_{dry}\left(\left[1-\frac{1}{f_v}\right]\exp\left(-\frac{[OH^-]}{[OH^-]_0}\right)+\frac{1}{f_v}\right)
-\end{equation}
-where $f_v$ is the viscosity reduction factor between dry and wet conditions, $[OH^-]$ is the water content and $[OH^-]_0$ is a reference water content set to
-620 ppm (0.062 wt.\%) \citep{Arcay2005}. By default, $f_v$ is set to 100, in according to \citet{Arcay2005} and \citet{Horiuchi2016}.
-
-\section{Mantle melting}\label{sec:melt}
-Melt fraction in the mantle depends on temperature ($T$, in \textdegree C), pressure ($p$, in \si{\giga\pascal}) and the water content in the melt ($X_{H_2O}$, in wt.\%).
-The determination of the melt fraction $M$ is implemented as explained by \citet{Katz2003}, according to the successive modification by \citet{Langmuir2006} and
-\citet{Kelley2010}, as follows
-\begin{eqnarray}\label{eq:melt_fr}
-M(p,T,X_{H_2O})&=&-T+T_{sol}+(x_t\cdot ln(p+y_t))\cdot X_{H_2O}-\nonumber\\
-&&+K_m\left(\frac{X_{H_2O}^{bulk}}{D_{H_2O}(1-X_{H_2O})+X_{H_2O}}\right)^{\gamma}
-\end{eqnarray}
-where $T_{sol}$ is the temperature of the dry solidus, $D_{H_2O}$ is the partition coefficient and $x_t\cdot log(p+y_t)$ indicates the pressure dependence of
-the melting curve \citep{Kelley2010}. $x_t$, $y_t$, $K_m$ and $\gamma$ are parameters chosen in according to \citet{Kelley2010}.
-However, $X_{H_2O}$ depends on the melt fraction as
-\begin{equation}\label{eq:melt_wt}
-X_{H_2O}(M)=\frac{X_{H_2O}^{bulk}}{D_{H_2O}+M(1-D_{H_2O})}
-\end{equation}
-and a numerical solution can be found using a root-finder method \citep{Katz2003,Wang2016}. Furthermore, the content of water in the melt is limited by the
-pressure-dependent saturation concentration of water in the melt, determined as
-\begin{equation}\label{eq:melt_max}
-X_{H_2O}^{sat}=\chi_1 p^{\lambda_m}+\chi_2 p
-\end{equation}
-where $\chi_1$, $\chi_2$ and $\lambda_m$ are parameters chosen according to \citet{Katz2003}, \citet{Langmuir2006} and \citet{Kelley2010}. The correctness in
-the determination of melt fractions and water contents in the melt are tested comparing model results with experimental curves obtained by \citet{Katz2003}
-(Sec. \ref{sec:katz}).
-
-In case of melting, viscous creep viscosity ($\eta_{vp}$) is modified according to \citet{Wang2016}, as
-\begin{equation}\label{eq:melt_visc}
-\eta_{melt}=\eta_{vp}\cdot \exp(\alpha_m M)
-\end{equation}
-where $\alpha_m$ are melt fraction factors for dislocation and diffusion creep, chosen in according to \citet{Wang2016}.
-
-Effective density ($\rho_{eff}$) for partially molten rocks is calculated as
-\begin{equation}\label{eq:melt_dens}
-\rho_{eff}=\rho_s(1-M)+\rho_m M
-\end{equation}
-where $\rho_s$ and $\rho_m$ are the densities of the solid and the molten rock, respectively \citep{Gerya2004b,Gerya2007,Wang2016}. The density of the
-solid rock is extracted by Perple\_X, while the density of the molten portion is calculated as
-\begin{equation}\label{eq:melt_dens_M}
-\rho_m(p,T)=\rho_0[1-\alpha(T-T_0)][1+\beta(p-p_0)]
-\end{equation}
-where $\rho_0$ is the density at the reference temperature ($T_0$) and pressure($p_0$), and $\alpha$ and $\beta$ are the thermal expansion and compressibility
-coefficients, respectively \citep{Gerya2004b,Gerya2007,Wang2016}.
-
-As for the phase transitions, effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into account
-considering effective coefficient of thermal expansion. Similarly, latent heat due to melting/crystallisation can be implicitly included in the energy equation
-(Eq. \ref{eq:energy}) by considering effective specific heat ($C_{p_{eff}}$) and thermal expansion ($\alpha_{eff}$) of the partially molten rocks, in according to
-\begin{eqnarray}
-\label{eq:cp_effective}C_{p_{eff}}&=&C_p+H_L\left(\frac{\partial M}{\partial T}\right)_p \\
-\label{eq:alpha_effective}\alpha_{eff}&=&\alpha+\rho\frac{H_L}{T}\left(\frac{\partial M}{\partial p}\right)_T
-\end{eqnarray}
-where $C_p$ and $\alpha$ are the specific heat and the coefficient of thermal expansion of the solid rock, respectively, and $H_L$ is the latent heat of the
-molten rock \citep{Gerya2004b,Gerya2007,Tackley2010}.
-
-\section{Benchmarks}\label{sec:benchmark}
-\subsection{Error measurements}\label{sec:error}
-In order to determine the accuracy of velocity and pressure field of the benchmarks, the $L_2$-norm is computed by numerical integration on the Gauss-Legendre
-quadrature points. $L_2$-norm for pressure and velocity errors can be evaluated as
-\begin{eqnarray}
-\textrm{err}_p&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|p^n_i(r_q)-p^a_i(r_q)|^2\right)w_q|J_q|}\nonumber \\
-\textrm{err}_v&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|u^n_i(r_q)-u^a_i(r_q)|^2+|v^n_i(r_q)-v^a_i(r_q)|^2\right)w_q|J_q|}\nonumber
-\end{eqnarray}
-respectively, where $n_e$ is the number of elements, $n_q$ is the number of quadrature points per element, $p^n_i(r_q)$ and $p^a_i(r_q)$ are the numerical and
-analytical pressure, respectively, in each quadrature point $q$, $w_q$  and $J_q$ are the weight and the Jacobian at the quadrature point
-$q$, $u^n_i(r_q)$, $v^n_i(r_q)$, $u^a_i(r_q)$ and $v^a_i(r_q)$ are the numerical and analytical velocities, respectively, in each quadrature point $q$.
-
-Other quantities used as comparison with original benchmarks are the root-mean-square velocity over the whole domain and over the surface
-\begin{eqnarray}
-v_{\textrm{rms}}&=&\sqrt{\int_{\Omega}|\bm{v}|^2 d\Omega}\nonumber \\
-v_{\textrm{rms}}^{top}&=&\sqrt{\int_0^1 u^2 \bigg|_{y=1} dx}\nonumber
-\end{eqnarray}
-respectively, top and bottom Nusselt numbers
-\[\textrm{Nu}^{top/bottom}=-\int_0^1\frac{\partial T}{\partial y} \bigg|_{y=1/y=0} dx\]
-the average rate of work done against gravity
-\[\left\langle W \right\rangle=\int_{\Omega}T u_y d\Omega\]
-the average rate of viscous dissipation
-\[\left\langle \Phi \right\rangle=\int_{\Omega}\tau_{ij}\dot{\epsilon}_{ij} d\Omega\]
-and the percentage error between $\left\langle W \right\rangle$ and $\left\langle \Phi \right\rangle/Ra$
-\[\delta=\frac{\left|\left\langle W \right\rangle -\frac{\left\langle \Phi \right\rangle}{Ra} \right|}{\max \left(\left\langle W \right\rangle,
-\frac{\left\langle \Phi \right\rangle}{Ra}\right)}\times100\]
-
-\subsection{Stokes flow}\label{sec:stokes}
-The problem consists of determining velocity field $(u,v)$ and pressure $p$ in case of a manufactured solution with prescribed body forces such as
-\begin{eqnarray}
-b_1&=&(12-24y)x^4+(-24+48y)x^3+(-48y+72y^2-48y^3+12)x^2+\nonumber \\
-&&+(-2+24y-72y^2+48y^3)x+1-4y+12y^2-8y^3\nonumber \\
-b_2&=&(8-48y+48y^2)x^3+(-12+72y-72y^2)x^2+(4-24y+48y^2-48y^3+\nonumber \\
-&&+24y^4)x-12y^2+24y^3-12y^4\nonumber
-\end{eqnarray}
-for which the exact solution is:
-\begin{eqnarray}
-u(x,y)&=&x^2(1-x)^2(2y-6y^2+4y^3)\nonumber \\
-v(x,y)&=&-y^2(1-y)^2(2x-6x^2+4x^3)\nonumber \\
-p(x,y)&=&x(1-x)-\frac{1}{6}\nonumber
-\end{eqnarray}
-The domain is a unit square a constant viscosity ($\eta=1$) and the penalty parameter is set to $\lambda=10^7$. Velocity boundary conditions are set to no slip
-$(\bm{v}=\bm{0})$ on all boundaries. The problem is performed for different grid resolution between \num{8x8} and \num{1024x1024} elements.
-
-The errors between the analytical solution and the numerical prediction of the velocity and the pressure field are calculated as in Section \ref{sec:error}.
-Fig. \ref{fig:errors}a shows that both velocity and pressure field converge to the exact solution with the decrease of the element size, following the
-theoretical convergence rate. The convergence of pressure, in contrast with the observation by \citet{Donea2003} for $Q_1 \times P_0$ elements, indicate the
-effectiveness of the smoothing procedure. Smoothed pressure field for grid resolution of \num{128x128} elements is shown in Fig. \ref{fig:errors}b (continuous
-red line), in comparison with the analytical solution (dashed blue line). Solve times and memory usage needed to generate the solution are shown in
-Fig. \ref{fig:MUMPS}. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Solver/Stokes_Flow}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/errors.pdf}
-\caption{Velocity and pressure error for the Stokes flow experiment between generated and analytical solution as a function of element size (panel a),
-and comparison between smoothed pressure and analytical solution as function of $x$ coordinate for a grid resolution of \num{128x128} elements (panel b).}
-\label{fig:errors}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/MUMPS.pdf}
-\caption{Panel a) Analysis, factorisation, solution and total solve times (panel a) and factorisation memory usage (panel b) as a function of the total number
-of degrees of freedom for the Stokes flow experiment.}
-\label{fig:MUMPS}
-\end{figure}
-
-\subsection{Markers advection}\label{sec:runge}
-The 2nd-order Runge-Kutta advection scheme in space is tested by means of the Zalesak disk test \citep{Zalesak1979,Thieulot2014}. The benchmark is performed
-in a unit square domain with a grid resolution of \num{32x32} elements and 1000000 markers and values of Courant number between 0.25 and 1.
-At $t=0$, the disk is centred at position $(0.5;0.75)$ with a radius $R=0.15$ and has a vertical fissure 0.05 wide and 0.2 high (orange in Fig. \ref{fig:runge})
-and with the velocity field prescribed in the entire domain as
-\begin{eqnarray}
-u(x,y)&=&2\pi\left(y-\frac{L_x}{2}\right)\nonumber \\
-v(x,y)&=&-2\pi\left(x-\frac{L_x}{2}\right)\nonumber
-\end{eqnarray}
-the markers are back to their initial location after a $2\pi$ rotation ($t=1$). The clockwise rotation of the disk at $t=0.25$, 0.5 and 0.75 
-is shown in Fig. \ref{fig:runge} (red, blue and green, respectively). The distance from the centre of 1 marker is calculated throughout the rotation to
-evaluate the error for different values of Courant number. As expected, the error increase with the increasing of the time step, as shown in 
-Fig. \ref{fig:runge_err}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/Zalesak_Disk}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=200px]{./Figures/Zalesak.pdf}
-\caption{Position of the Zalesak disk at $t=0$, 0.25, 0.5 and 0.75 (orange, red, blue and green, respectively) for the case with Courant number of 1.}
-\label{fig:runge}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/zal_marker.pdf}
-\caption{Distance from the centre as function of time for values of Courant number of  of 0.25, 0.3, 0.5, 0.75 and 1 (green, orange, blue, black and red,
-respectively). R indicates the distance from the centre considering a perfect circle trajectory.}
-\label{fig:runge_err}
-\end{figure}
-
-\subsection{Conservative Velocity Interpolation (CVI)}\label{sec:CVI}
-The CVI correction is checked by means of the Stokes flow experiment by \citet{Donea2003}, as presented in \ref{sec:stokes}. The advection of Lagrangian
-markers is performed using either a 2nd-order or a 4th-order Runge-Kutta scheme with CFL=0.5 and an initial random distribution of 25 markers per element.
-Fig. \ref{fig:CVI} shows the minimum and maximum values of markers per element throughout the experiment with and without the CVI correction and using
-different Runge-Kutta schemes. The introduction of the CVI correction allows the
-tests to run for times longer than $t=2000$ without show empty elements, while tests without the CVI correction stop at approximately $t=600$, when at least
-one element goes empty. Fig. \ref{fig:CVI_time} shows the different number of markers per element and the markers distribution at $t=600$ obtained without and
-with the CVI correction (panels b and c, respectively) in case of a 2nd-order Runge-Kutta scheme. All data can be found at
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/CVI}.
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/CVI.pdf}
-\caption{Maximum and minimum number of markers per element throughout the experiment, with (green and orange lines) or without (purple and blue lines)
-the CVI correction and using either a 2nd-order (blue and orange lines) or a 4th-order (purple and green lines) Runge-Kutta scheme.}
-\label{fig:CVI}
-\end{figure}
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/CVI_time.pdf}
-\caption{Number of markers per element and markers distribution through time using 2nd.order Runge-Kutta scheme.
-Panels a: number of markers per element (panel a1) and distribution of the markers (panel a2) after the first
-iteration; panels b and c: comparison between number of markers per element (panels b1 and c1) and markers distribution (panels b2 and c2) without and with
-the CVI correction at $t=600$; panels d: number of markers per element (panel d1) and markers distribution (panel d2) in case of CVI correction at $t=2000$.}
-\label{fig:CVI_time}
-\end{figure}
-
-\subsection{Poiseuille flow}\label{sec:poiseuille}
-The domain is a rectangle with $L_x=2$ and $L_y=1$ and constant density and viscosity ($\rho=1$ and $\eta=1$), gravity acceleration $\bm{g}=\bm{0}$ and penalty
-parameter $\lambda=10^8$. The grid is composed by \num{40x20} elements. Velocity boundary conditions are set to no slip $(\bm{v}=\bm{0})$ at the top and the
-bottom, and a parabolic profile is imposed on the sides, with $u=y(1-y)$ and $v=0$. The analytical solution is then given by:
-\begin{eqnarray}
-u(x,y)&=&y(1-y)\nonumber \\
-v(x,y,)&=&0\nonumber \\
-p(x,y)&=&2\eta\left(\frac{L_x}{2}-x\right)\nonumber
-\end{eqnarray}
-
-The velocity field predicted by the model follows the expected parabolic profile (Fig. \ref{fig:poiseuille}a and Fig. \ref{fig:poi_plot}a) and, in case of the
-classic penalty method with no iterations, the pressure is clearly related to the divergence by means of the penalty parameter (Fig. \ref{fig:poiseuille}b and
-c and Fig. \ref{fig:poi_plot}b). Fig. \ref{fig:poiseuille}d shows that one Uzawa iteration is sufficient to bring the divergence down to \num{e-15}, with no
-correlation with the pressure field. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Poiseuille%20Flow}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Poiseuille.pdf}
-\caption{Velocity field (panel a), pressure (panel b) and divergence velocity of a Poiseuille flow in case of the classic penalty method (no iterations) and
-after one Uzawa iteration (panel c and d, respectively).}
-\label{fig:poiseuille}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Analytical.pdf}
-\caption{Velocity (panel a) and pressure (panel b) field predicted by the model for a Poiseuille flow with respect with their analytical solutions. The
-velocity field is plotted as function of the vertical coordinate in $L_x/2$ and the pressure is plotted as function of the x coordinate in $L_y/2$.}
-\label{fig:poi_plot}
-\end{figure}
-
-\subsection{Instantaneous 2D sphere}\label{sec:ist_sphere}
-The domain is a unit square with gravity $\bm{g}=(0,-1)$. The fluid has constant density and viscosity ($\rho_f=1$ and $\eta_f=1$). The sphere is in the middle
-of the domain with a radius R = 0.123456798 and has constant density and viscosity ($\rho_s=10^{-2}+\rho_f$ and $\eta_s=10^3 \cdot \eta_f$).
-We distinguish three different types of velocity boundary conditions:
-\begin{enumerate}
-\item FS: free slip conditions on all sides.
-\item NS: no slip conditions on all sides.
-\item OT: free slip conditions on the sides and the bottom, and open boundary at the top.
-\end{enumerate}
-All three cases are performed for different grid resolutions, between \num{16x16} and \num{512x512} elements with 50 randomly distributed markers per element,
-and different average schemes for the viscosity (harmonic, geometric and arithmetic). The velocity in the centre of the sphere $\bm{v}(0.5;0.5)$), minimum and
-maximum velocities ($u_{min}$, $u_{max}$, $v_{min}$, $v_{max}$) and pressures ($p_{min}$, $p_{max}$) on the entire domain, $v_{\textrm{rms}}$ and average pressure on the
-entire domain ($p_{avg}$) are compared with solutions generated by ASPECT \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and all results
-can be found at \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere2D} and 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Instantaneous_2D_Sphere}. Results in terms of the $v_{\textrm{rms}}$ for NS model 
-are also shown in Fig. \ref{fig:inst_sphere}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_NS.pdf}
-\caption{$v_{\textrm{rms}}$ as function of element size for the instantaneous 2D sphere experiment in case of no slip boundary conditions and with different average
-schemes for the viscosity. Results are compared with results obtained by other numerical codes.}
-\label{fig:inst_sphere}
-\end{figure}
-
-\subsection{Rayleigh-Taylor instability}\label{sec:rayleigh}
-This problem has been originally presented by \citet{vanKeken1997} and here it is performed as the isoviscous case (Case 1a). The domain has $L_x=0.9142$ and
-$L_y=1$ and gravity $\bm{g}=(0,-1)$. Two fluids with same constant viscosities ($\eta_1=\eta_2=1$) and different densities ($\rho_1=1000$ and $\rho_2=1010$),
-with the lighter fluid at the bottom. The initial interface between the fluids is given by $y(x)=0.2+0.02 \cos \left(\frac{\pi x}{L_x}\right)$. The experiment
-is performed with different grid sizes (\num{50x50}, \num{80x80}, \num{100x100} and \num{256x256}). A total of 1960000 markers are randomly distributed at the
-beginning of the simulation. Velocity boundary conditions are set to no slip at the top and the bottom, and to free slip at the sides of the domain.
-
-$v_{\textrm{rms}}$ as function of time is reported in Fig. \ref{fig:RT}, matching well with results shown by \citet{vanKeken1997}, \citet{Tackley2003} and
-\citet{Thieulot2014}. Fig. \ref{fig:rayleigh} shows the evolution of the experiment at different time steps. All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Rayleigh_Taylor_experiment/ISOVISCOUS}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/RT.pdf}
-\caption{$v_{\textrm{rms}}$ of the Rayleigh-Taylor experiment as function of time for different resolution of the grid.}
-\label{fig:RT}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=300px]{./Figures/Rayleigh.pdf}
-\caption{Evolution of the Rayleigh-Taylor experiment for a grid of \num{256x256} elements at $t=0, 500, 1000$ and $2000$ (panels a, b, c and d, respectively).}
-\label{fig:rayleigh}
-\end{figure}
-
-\subsection{Falling blocks}\label{sec:block}
-This benchmark is described as proposed by \citet{Gerya2003a}, \citet{Gerya2010b} and \citet{Thieulot2011}. The domain is square with $L_x=L_y=$ \SI{500}{\km}
-and the grid is composed by \num{50x50} elements with 25 markers in each element. The block is initially centred at ($x=$ \SI{250}{\km}; $y=$ \SI{400}{\km})
-and has a size of \num{100x100} km (Fig. \ref{fig:block}a). The fluid surrounding the block has $\eta_f=$ \SI{e21}{\pascal\per\s} and $\rho_f=$
-\SI{3200}{\kg\per\cubic\m}. The benchmark tests with different viscosities of the block, with $\eta_b$ from \SIrange{e15}{e27}{\pascal\per\s}. In each
-experiment also the density $\rho_b$ of the block varies from \SIrange{3220}{9900}{\kg\per\cubic\m}. Velocity boundary conditions are set to free slip
-conditions on all sides of the domain. Density distributions at $t=$ \SI{20}{\mega\year} for $\rho_b=$ \SI{3300}{\kg\per\cubic\m} and different $\eta_b$
-are plotted in Fig. \ref{fig:block}b-f, showing that the code correctly preserve the block geometry in case of large viscosity contrast, with a block stiffer
-than the surrounding fluid (Fig. \ref{fig:block}f).
-
-The velocity in the centre of the falling block at $t=0$ is measured for all experiments and, since the velocity should increase with density contrast, the
-quantity $\bm{v}/(\rho_b-\rho_f)$ is plotted as function of the viscosity contrast. All results perfectly match with those from \citet{Gerya2010b} and they
-line up on a single curve, demonstrating that the code can correctly deal with large viscosity and density contrasts (Fig. \ref{fig:falling}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Falling%20blocks}.
-
-\begin{figure}
-\noindent\includegraphics[width=300px]{./Figures/Block.pdf}
-\centering
-\caption{Density evolution of the falling block experiment at $t=$ \SI{0}{\mega\year} (panel a) and $t=$ \SI{20}{\mega\year} for different viscosities of the
-block (panels b-f).}
-\label{fig:block}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Falling.pdf}
-\caption{Initial velocity relative to the density contrast at the centre of the falling block as function of the viscosity contrast between the block and the
-surrounding fluid.}
-\label{fig:falling}
-\end{figure}
-
 \subsection{2D time-dependent Stokes sphere with deformable free surface}\label{sec:sphere}
@@
 $y\leq 0.75$, while for $y>0.75$ the air has $\rho_a=0$ and $\eta_a=10^{-3}$. Velocity boundary conditions are set to free slip on all sides. The Courant
-number is set to 0.25. The experiments run for 200 s using grid resolutions from \num{150x150} to \num{512x512} elements, with 25 randomly distributed markers
+number is set to 0.25. The experiments run for 200 s using grid resolutions from $150\times150$ to $512\times512$ elements, with 25 randomly distributed markers
 per element, and for different average schemes for the viscosity. The interface between the fluid and the air is tracked by means of the markers chain. Results
@@
 \end{figure}
-
-\subsection{Free surface stabilisation}\label{sec:stab}
-The benchmark is performed as discussed in \citet{Kaus2010a} and \citet{Thieulot2014}. The domain is square with $L_x=L_y=$ \SI{500}{\km} and a grid resolution
-of \num{200x200} elements, each of them containing 16 markers. A buoyant fluid with $\rho_1=$ \SI{3200}{\kg\per\cubic\metre} and $\eta_1=$ \SI{e20}{\pascal\s}
-is overlain by a denser fluid with $\rho_2=$ \SI{3300}{\kg\per\cubic\metre} and $\eta_2=$ \SI{e21}{\pascal\s}. The initial interface between the fluids has an
-initial sinusoidal shape of 5 km amplitude. Velocity boundary conditions are set to free slip conditions at the sides and no slip at the bottom, while the top
-is free surface. The experiment is performed with various fixed time steps, without the use of the Courant number. The vertical position of the free surface at
-$x=L_x$ is tracked for each simulation, with or without the stabilisation algorithm.
-
-The results show an instability (drunken sailor effect) increasing the time step in case that the stabilisation algorithm is not activated
-($dt=$ \SI{4200}{\year}, green line in Fig. \ref{fig:stab}), as already observed by \citet{Kaus2010a} and \citet{Thieulot2014}. Activating the algorithm the
-instability is fixed and the simulation remains stable using time step up to \SI{50000}{\year} (red line in Fig. \ref{fig:stab}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Stabilisation_algorithm}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Kaus_random.pdf}
-\caption{Evolution of the y coordinate in $x=L_x$ as function of time for different time steps, with and without the stabilisation algorithm.}
-\label{fig:stab}
-\end{figure}
-
-\subsection{Topography relaxation}\label{sec:crameri}
-This experiment is performed as in \citet{Crameri2012} using free surface and sticky air (Case 1). The domain is rectangular with $L_x=$ \SI{2800}{\km} and
-$L_y=$ \SI{700}{\km} in case of free surface and $L_y=$ \SI{800}{\km} in case of sticky air ($h_{st}=$ \SI{100}{\km}). The grid is composed by \num{256x64}
-elements for the free surface \num{560x320} elements for the sticky air. For both cases, there is a mantle of 600 km thickness with $\eta_m=$ \SI{e21}{\pascal\s}
-overlain by a highly viscous cosine-shaped lithosphere with thickness between 93 and 107 km and $\eta_l=$ \SI{e23}{\pascal\s}. Both mediums have $\rho=$
-\SI{3300}{\kg\per\cubic\metre}. In case of sticky air experiment, the air has $\rho_a=$ \SI{0}{\kg\per\cubic\metre} and $\eta_a=$ \SI{e18}{\pascal\s}. The
-gravity acceleration is set to $g_y=$ \SI{-10}{\metre\per\square\s}. Velocity boundary conditions are set to no slip at the bottom and free slip at the sides
-of the domain. In case of sticky air, velocities on top are set to free slip conditions.
-
-The maximum topography as function of time ($h(t)$) can be analytically derived in according to \[h(t)=h_0 \exp(\gamma t)\]
-where $h_0=7$ km is the initial topography, $\gamma=\num{-0.2139e-11}$ is the characteristic relaxation rate and $t$ is the time (black line in Fig.
-\ref{fig:crameri}). Maximum topography at the characteristic relaxation time $t_{rlx}=$ \SI{14.825}{\kilo\year} can be found to be $h_{rlx}=$ \SI{2576}{\m}.
-The results show that a viscosity of the air of \SI{e18}{\pascal\s} is needed in case of $h_{st}=$ \SI{100}{\km} to correctly describe the topography relaxation
-for this problem, as pointed out by \citet{Crameri2012} (purple line in Fig. \ref{fig:crameri}). The case with a true free surface well follows the analytical
-solution (orange line in Fig. \ref{fig:crameri}). In particular, at the relaxation time in case of the true free surface predicts a maximum topography of
-\SI{2570}{\m}, with an error of \SI{6}{\m} with respect to the analytical solution. Results are also compared with results obtained with a free surface in
-UNDERWORLD (\num{256x64} elements) and SULEC (\num{401x201} elements) (with courtesy from \citealp{Crameri2012}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Topography_relaxation}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Crameri.pdf}
-\caption{Maximum topography as function of time for the Crameri benchmark (coloured lines) in comparison with the analytical solution (black line) and results
-obtained with UNDERWORLD (yellow line) and SULEC (blue line).}
-\label{fig:crameri}
-\end{figure}
-
-\subsection{Spontaneous subduction}\label{sec:subduction}
-This experiment is performed as presented by \citet{Schmeling2008}, in case of both a sticky air and a true free surface and using different average schemes
-for the viscosities. The domain is rectangular with $L_x=$ \SI{3000}{\km} and $L_y=$ \SI{700}{\km}, a grid resolution of \num{256x64} elements with 50 markers
-per element and Courant number of 0.01. At the beginning of the simulation, a 100 km-thick lithospheric layer with $\rho_m=$ \SI{3300}{\kg\per\cubic\metre}
-and $\eta_m=$ \SI{e23}{\pascal\s} is located at the top of the domain for $x$ between 1000 and 3000 km, while the asthenospheric mantle has $\rho_m=$
-\SI{3200}{\pascal\s} and $\eta_m=$ \SI{e21}{\pascal\s}. In addition, a 200 km-depth lithospheric slab is already subducted in the mantle in order to have
-a spontaneous subduction. In case of sticky air, $L_y=750$ km and a 50 km-thick air layer with $\eta_a$ from \SIrange{e19}{e21}{\pascal\s} overlie the mantle.
-Velocity boundary conditions are set to free slip on the sides and on the bottom of the domain. In case of sticky air, velocities on the top boundary are set
-to free slip conditions as well.
-
-The maximum depth of the slab is tracked for all the simulations (Fig. \ref{fig:subduction}) and they are compared with results obtained by
-\citet{Schmeling2008} with models with regular grids and comparable grid resolutions (rectangles in Fig. \ref{fig:subduction}). Low viscous sticky air models
-(i.e., \SI{e19}{\pascal\s}) enlighten a strong dependence of the sinking velocity with the chosen average scheme (dashed lines in Fig. \ref{fig:subduction}),
-while high viscous sticky air models (i.e., \SI{e21}{\pascal\s}) show a higher resistance at the trench, underestimating the correct solution (dotted lines in
-Fig. \ref{fig:subduction}). A high resistance at the trench can be also observed in case of a true free surface for the low resolution grid used (continuous
-lines in Fig. \ref{fig:subduction}), as pointed out by \citet{Schmeling2008}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Spontaneous_subduction}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Subduction.pdf}
-\caption{Maximum depth of the slab as function of time for all the simulations of the \citet{Schmeling2008} benchmark, in case of harmonic, geometric and
-arithmetic means (red, green and blue lines, respectively) using a sticky air or a true free surface (discontinuous and continuous lines, respectively).
-Red, green and blue rectangular areas indicate the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{500}{\km} in case of sticky air,
-while the grey rectangular area indicates the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{600}{\km} in case of true free surface.}
-\label{fig:subduction}
-\end{figure}
-
-\subsection{The slab detachment}\label{sec:slab}
-The slab detachment benchmark is performed as described by \citet{Schmalholz2011} and \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{1000}{\km}
-and $L_y=$ \SI{660}{\km} and a grid resolution of \num{256x256} elements. A non-linear viscous T-shaped layer with $\rho_l=$ \SI{3300}{\kg\per\cubic\metre} is
-placed at the top of the domain and surrounded by a linear viscous fluid with $\rho_f=$ \SI{3150}{\kg\per\cubic\m}. The top layer is 80 km-thick and a 250
-km-long and 80 km-wide slab is placed at $x=L_x/2$ (Fig. \ref{fig:slab}a). The effective viscosity of the top layer is given by
-\[\eta_{eff}=\eta_0 I_2^{\frac{(1-n)}{n}}\]
-with $\eta_0=$ \SI{4.75e11}{\pascal\s} and $n=4$, while the surrounding fluid has $\eta_f=$ \SI{e21}{\pascal\s}. The effective viscosity is capped between
-$\eta_{min}=$ \SI{e21}{\pascal\s} and $\eta_{max}=$ \SI{e25}{\pascal\s}. Velocity boundary conditions are set to free slip at the top and the bottom, and to
-no slip at the sides of the domain.
-
-The sinking of the slab determines a decrease of the effective viscosity in the T-shaped layer during the first part of the evolution (Fig. \ref{fig:slab}a),
-as consequence of local strain rates, while effective viscosities increase up to $\eta_{max}$ in the top layer after approximately \SI{20}{\mega\year}, when
-necking is complete (Fig. \ref{fig:slab}b).
-
-The width of necking is tracked by means of markers position at the side of the slab. Necking width and time are normalised in function of the initial slab
-(\SI{80}{\km}) and the characteristic time ($t_c=$ \SI{7.1158e14}{\s}) \citep{Schmalholz2011,Glerum2018}. The results in case of arithmetic, geometric and
-harmonic mean of the viscosity are shown in Fig. \ref{fig:necking}, compared with results from \citet{Glerum2018} for which an infinity norm mean was used.
-The influence of the average schemes agrees with results shown in Section \ref{sec:subduction}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Slab_detachment}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Slab.pdf}
-\caption{Effective viscosity for the slab detachment benchmark at the beginning of the evolution (panel a) and when necking is complete (panel b).}
-\label{fig:slab}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Necking.pdf}
-\caption{Normalised width of the necking of the slab detachment benchmark as function of normalised time.}
-\label{fig:necking}
-\end{figure}
-
-\subsection{The indenter experiment}\label{sec:indenter}
-The indenter benchmark simulate a rigid punch on a purely plastic von Mises material, for which exists an analytical solution
-\citep{Thieulot2008,Thieulot2014,Glerum2018}. The domain is a unit square with a grid resolution of \num{256x256} elements. The gravity acceleration is set
-to $\bm{g}=\bm{0}$ and effective viscosity is capped using $\eta_{min}=10^{-4}$ and $\eta_{max}=10^3$. The tolerance for the convergence of the non-linear
-solution is $tol=10^{-9}$, with a maximum number of non-linear iterations set to 500. The medium in the domain has $\rho=0.01$, an initial viscosity
-$\eta_0=10$, cohesion $C=1$ and an angle of internal friction $\phi=0$°. Velocity boundary conditions are set to no slip at the bottom and free slip at the
-sides of the domain. The top of the domain is open with the exception of the central portion (punch area) with width $w_p=0.125$, where $v=-1.05$ and $u$ is
-fixed either to 0 (rough punch experiment) or free (smooth punch experiment).
-
-Fig. \ref{fig:indenter} show the results in terms of viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure
-(panels g and h) for both experiments. The analytical solution indicate that pressure at ($0.5;1$) and ($0.5 \pm w_p;1$)is $p=\pi +1$ and $p=1$, respectively
-(black and red lines in Fig. \ref{fig:smooth_p}). Fig. \ref{fig:smooth_p} shows a clear improvement in the pressure solution when passing from a rough to a
-smooth punch (green and blue lines in Fig. \ref{fig:smooth_p}, respectively), as previously observed by \citet{Thieulot2014} and \citet{Glerum2018}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Indenter}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Indenter.pdf}
-\caption{Viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure (panels g and h) fields for rough (left column) and
-smooth (right column) punch experiments.}
-\label{fig:indenter}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Smooth_Rough.pdf}
-\caption{Nodal pressure in function of the x coordinate at $y=1$ for rough (green line) and smooth (blue line) punch experiment. Black and red lines indicate
-the analytical solution at $x=0.5$ and $x=0.5 \pm w_p$, respectively.}
-\label{fig:smooth_p}
-\end{figure}
-
-\subsection{The brick experiment}\label{sec:brick}
-An instantaneous version of the brick benchmark is performed to verify the correctness of the pressure-dependent plasticity for different angle of internal
-friction, as proposed by \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{40}{\km} and $L_y=$ \SI{10}{\km} and a grid resolution of \num{512x128}
-elements. The gravity acceleration is set to $g_y=$ \SI{-10}{\m\square\s} and effective viscosity is capped using $\eta_{min}=$ \SI{e19}{\pascal\s} and
-$\eta_{max}=$ \SI{e26}{\pascal\s}. The tolerance for the convergence of the non-linear solution is $tol=10^{-7}$, with a maximum number of non-linear iterations
-set to 1000. A 800 m-wide and 400 m-height inclusion with $\eta_b=$ \SI{e20}{\pascal\s} and $\rho_b=$ \SI{2700}{\kg\per\cubic\m} is placed at the bottom of the
-domain at $x=L_x/2$. The inclusion is surrounded by a non-linear viscous medium with $\rho_m=$ \SI{2700}{\kg\per\cubic\m}, an initial viscosity of
-\SI{e23}{\pascal\s}, a linear viscous viscosity of \SI{e25}{\pascal\s} and a cohesion of \SI{40}{\mega\pascal}. Velocities are set to free slip conditions at
-the bottom of the domain and the top is open. Velocities on sides of the domain are fixed to $u=$ \SI{+-2e-11}{\m\per\s} and $v=0$. The experiment is performed
-in compressional and extensional contexts with an angle of internal friction $\phi$ of the non-linear medium variable between 0° and 30°.
-
-As expected, two shear bands stem from the inclusion with variable angles in relation to both the dynamics context and the internal friction angle. Shear band
-angles formed at 45° for $\phi=0$° in both extensional and compressional contexts (Fig. \ref{fig:brick_beam}a), while they are different in case of $\phi\neq0$°.
-Shear band angles in case of $\phi=20$° are shown for compression and extension in Fig. \ref{fig:brick_beam}b  and c, respectively. Values of shear band angles
-as function of internal friction angles from 0° to 30° are extracted at different depths and minimum and maximum values are plotted in Fig. \ref{fig:brick},
-in comparison with theoretical Roscoe, Arthur and Coulomb shear band angles. For all tests the tolerance for nonlinear convergence is defined $tol=10^{-7}$
-and none converges before the maximum number of iterations ($it_{max}=1000$) is reached. However, tests with internal friction angles up to 15° show a constant
-decrease of the velocity residuals, which can not be observed for for tests with higher internal friction angles (Fig. \ref{fig:convergence}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Brick_experiment}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Brick_Beam.pdf}
-\caption{Shear band angles predicted for the brick experiment for $\phi=0$° (panel a) and for $\phi=20$° in case of compression and extension (panels b and c,
-respectively).}
-\label{fig:brick_beam}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Brick.pdf}
-\caption{Shear band angles predicted for the brick experiment in case of compressional and extensional contexts as function of different internal angle of
-friction (continuous black line and red dots), compared with theoretical Roscoe, Artur and Coulomb angles (discontinuous black lines). Red lines indicate the
-range of angles calculated at different depths.}
-\label{fig:brick}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \noindent\includegraphics[width=500px]{./Figures/convergence.pdf}
-  \caption{Convergence rates for all tests. Normalised residuals (left column) are compared with velocity residuals (right column) for extensional (first row)
-  and compressional (second row) settings.}
-  \label{fig:convergence}
-  \end{figure}
-
-\subsection{Advection stabilisation}\label{sec:advection}
-The 1D advection problem proposed by \citet{Donea2003} and \citet{Thieulot2011} is performed to verify the effectiveness of the SUPG method to stabilise the
-advection term of the energy equation. The domain is a 1D segment with $L_x=1$ composed by 50 elements and a discontinuity in the thermal field placed at
-$x=0.25$. Temperature is set to 1 for $x \leq 0.25$ and to 0 for $x > 0.25$. Velocity is set to ${u}=1$ in the entire domain. The simulation is performed for
-250 time steps, with $dt=0.002$, so the thermal discontinuity should be at $x=0.75$ at the end of the simulation.
-
-Temperature profiles at the end of the simulation are shown in \ref{fig:advection} as function of the dimensionless coefficient $\gamma=\tau \bm{v}/h$. In case
-of the classic Galerkin method ($\gamma=0$, blue line in \ref{fig:advection}) the final thermal profile is characterised by strong oscillations, which are
-eliminated in case of the SUPG method ($\gamma=0.045$, orange line in \ref{fig:advection}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Advection_stabilisation}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Advection.pdf}
-\caption{Temperature profile as function of \textit{x} for the advection stabilisation benchmark. Purple line indicates the initial temperature profile; the
-green line indicates the analytical temperature profile after 250 time steps; blue line indicates the temperature profile after 250 time steps in case of the
-classic Galerkin method ($\gamma=0$); orange line indicates the temperature profile after 250 time steps in case of the SUPG method ($\gamma=0.045$).}
-\label{fig:advection}
-\end{figure}
-
-\subsection{Shear heating}\label{sec:simple_shear}
-This exercise is performed in an unit square domain composed by \num{128x128} elements. The velocity field is prescribed on the entire domain with
-$\bm{v}=(L_y-y)y\bm{e_x}$; viscosity, density and specific heat are fixed to 1, while thermal conductivity and radiogenic energy are fixed to 0. Therefore,
-the energy equation (Eq. \ref{eq:energy}) can be simplified as \[\frac{\partial T}{\partial t}=H_s\] and fixing $T(t=0)=0$, the temperature can be find as
-\[T(t)=H_s t\]
-In this case we have
-\begin{eqnarray}
-\dot{\epsilon_{xx}}&=&\frac{\partial u}{\partial x}=0 \nonumber \\
-\dot{\epsilon_{yy}}&=&\frac{\partial v}{\partial y}=0 \nonumber \\
-\dot{\epsilon_{xy}}&=&\frac{1}{2}\left(\frac{\partial v}{\partial x}+\frac{\partial u}{\partial y}\right)=\frac{1}{2}(L_y-2y)\nonumber
-\end{eqnarray}
-and, simplifying Eq. \ref{eq:shear}, shear heating can be calculated as \[H_s(x,y)=(1-2y)^2\]
-
-The solution predicted by the model in terms of velocity, temperature and shear heating match well with the analytical solutions (Fig. \ref{fig:analytical_en}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Simple_shear}.
-
-%\begin{figure}
-%\centering
-%\noindent\includegraphics[width=400px]{./Figures/Energy_Thieulot.pdf}
-%\caption{Temperature and shear heating predicted by the model after 1 time step.}
-%\label{fig:en_thieulot}
-%\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=500px]{./Figures/Analytical_Shear.pdf}
-\caption{Velocity (panel a), temperature (panel b) and shear heating (panel c) as function of the \textit{y} coordinate for the simple shear experiment. The
-solutions predicted by the model (red lines) are compared with the analytical solutions (dashed blue lines).}
-\label{fig:analytical_en}
-\end{figure}
-
-\subsection{Shear and adiabatic heating}\label{sec:shear}
-This problem is performed as presented in Exercise 9.4 in \citet{Gerya2010b}. Two materials are vertically separated in a rectangular domain with $L_x=$
-\SI{1000}{\km}, $L_y=$ \SI{1500}{\km} and a grid resolution of \num{30x20} elements. Constant thermal coefficient expansion ($\alpha=$ \SI{3e-5}{\per\kelvin}),
-temperature ($T=$ \SI{1300}{\kelvin}) and gravity acceleration ($g_y=$ \SI{-10}{\m\square\s}) are assumed in the whole domain. Fluid 1 (on the left side of the
-domain) has $\rho_1=$ \SI{3200}{\kg\per\cubic\m} and $\eta_1=$ \SI{e20}{\pascal\s}; fluid 2 (on the right side of the domain) has $\rho_2=$
-\SI{3300}{\kg\per\cubic\m} and $\eta_2=$ \SI{e22}{\pascal\s}. Velocity boundary conditions are set to free slip on all sides of the domain.
-
-As shown in Fig. \ref{fig:energy}, both shear and adiabatic heating predicted by the code (first row in Fig. \ref{fig:energy}) well recreate the results
-obtained by means of example ${Shear\_adiabatic\_heating.m}$ from \citet{Gerya2010b} (second row in Fig. \ref{fig:energy}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Adiabatic%2BShear}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Energy.pdf}
-\caption{Comparison between shear (first column), adiabatic (second column) and total (third column) energy predicted by the code (panels a, b and c) and those
-created using example ${Shear\_adiabatic\_heating.m}$ from Exercise 9.4 in \citet{Gerya2010b} (panels d, e and f).}
-\label{fig:energy}
-\end{figure}
-
-\subsection{Mantle convection}\label{sec:mantle}
-This problem is performed as presented by \citet{Blankenbach1989} (constant viscosity cases) and \citet{Thieulot2014}, in a 2D unit square domain with gravity
-acceleration $g_y=10^{10}Ra$. The experiment is performed with three different Rayleigh numbers ($Ra=10^4$, $10^5$ and $10^6$) and with different grid
-resolution (between \num{32x32} and \num{128x128} elements). The fluid has constant viscosity, initial density, heat capacity, thermal conductivity
-($\eta= \rho_0= C_p = k=1$), reference temperature ($T_0=0$) and thermal expansion coefficient ($\alpha=10^{-10}$). Temperatures are set to 0 on top and 1 on
-bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\]
-
-The solution generated by the code in terms of $v_{\textrm{rms}}$ (calculated as in Sec. \ref{sec:error}) and Nu as function of time are reported for all the
-simulations in Fig. \ref{fig:mantle}. At the steady state, $v_{\textrm{rms}}$ and Nu of all simulations converge well toward the values from \citet{Blankenbach1989},
-with lower errors for higher resolution grids (Table \ref{tab:mantle}). All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Mantle_convection}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Convection.pdf}
-\caption{$v_{\textrm{rms}}$ (panels a, c and e) and Nu (panels b, d and f) for the mantle convection benchmark as function of time for different grid resolution.
-Panels a and b show the results for $Ra=10^4$; Panels c and d show the results for $Ra=10^5$; panels e and f show the results for $Ra=10^6$. Purple lines
-indicate the convergence values for the $v_{\textrm{rms}}$ and the Nu from \citet{Blankenbach1989} at the steady state.}
-\label{fig:mantle}
-\end{figure}
-
-\begin{table}[H]
-\caption{Comparison between $v_{\textrm{rms}}$ and Nu predicted by the code for the mantle convection experiment and same values as reported in literature.}
-\centering
-\small
-\begin{tabular}{|>{\centering}m{0.09\textwidth}|>{\centering}m{0.05\textwidth}|>{\centering}m{0.24\textwidth}|>{\centering}m{0.15\textwidth}|
-  >{\centering}m{0.1\textwidth} >{\centering}m{0.1\textwidth} >{\centering\arraybackslash}m{0.1\textwidth}|}
-\toprule
-& & \multirow{2}{*}{\citet{Blankenbach1989}} & \citet{Thieulot2014} & \multicolumn{3}{c|}{FALCON}  \\
- & & & \num{200x200} & \num{32x32} & \num{64x64} & \num{128x128}  \\
-\midrule
-\multirow{2}{*}{$Ra=10^4$} & $v_{\textrm{rms}}$ & $42.864947 \pm 0.000020$ & 42.867 & 42.83226 & 42.852793 & 42.861394  \\
-           & Nu      & $4.884409 \pm 0.000010$  & 4.882  & 4.781297 & 4.857475  & 4.877573   \\
-\hline
-\multirow{2}{*}{$Ra=10^5$} & $v_{\textrm{rms}}$ & $193.21454 \pm 0.00010$  & 193.255& 193.872643 & 193.377472 & 193.252290  \\
-           & Nu      & $10.534095 \pm 0.000010$ & 10.507 & 9.602514 & 10.270735  & 10.465629   \\
-\hline
-\multirow{2}{*}{$Ra=10^6$} & $v_{\textrm{rms}}$ & $833.98977 \pm 0.00020$  & 834.712& 848.091176 & 837.767911 & 834.945793  \\
-           & Nu      & $21.972465 \pm 0.000020$ & 21.695 & 15.999266 & 19.703682  & 21.306939   \\
-\bottomrule
-\end{tabular}
-\label{tab:mantle}
-\end{table}
-
-\subsection{Viscoplastic mantle convection}\label{sec:visco_mantle}
-This benchmark is performed as Cases 1-5a described in \citet{Tosi2015} in a 2D unit square domain with two different grid resolutions of \num{32x32} 
-and \num{100x100} elements. Reference viscosity, reference density, heat capacity, thermal conductivity and thermal expansion coefficient are fixed to a constant
-value of 1 ($\eta_0= \rho_0= C_p = k=\alpha=1$), while the gravity has been chosen as $g_y=10^2$, in order to obtain a Rayleigh number $Ra=10^2$. Temperatures
-are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\].
-The viscosity field $\eta$ is calculated as the harmonic average between a linear part $\eta_{lin}$ that depends on temperature only or on temperature and 
-depth $y$ and a nonlinear-plastic part $\eta_{pl}$ that depends on the strain rate $\dot{\bm{\epsilon}}$, as follows
-\[\eta(T,y,\dot{\bm{\epsilon}})=2\left(\frac{1}{\eta_{lin}(T,y)}+\frac{1}{\eta_{pl}(\dot{\bm{\epsilon}})}\right)^{-1}\]
-The linear and the nonlinear-plastic parts of the viscosity are calculated following \citet{Tosi2015} as follows
-\[\eta_{lin}(T,y)=\exp(-\ln(\Delta\eta_T)T+\ln(\Delta\eta_y)y)\]
-\[\eta_{pl}(\dot{\bm{\epsilon}})=\eta^*+\frac{\sigma_Y}{\sqrt{\dot{\bm{\epsilon}}:\dot{\bm{\epsilon}}}}\]
-where $\Delta\eta_T$, $\Delta\eta_y$, $\eta^*$ and $\sigma_Y$ are parameters chosen as listed in Table \ref{tab:tosi_setup}.
-
-For each case, some diagnostic quantities measured at the steady state are reported in Table \ref{tab:visco_mantle} (see Section \ref{sec:error} for 
-the computation of all the quantities), showing a well fit with results described in \citet{Tosi2015}. Data and charts for all cases can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Viscoplastic_mantle_convection}.
-
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{white!100!}{gray!25!}
-%\begin{table}[H]
-%  \caption{Results of the viscoplastic mantle convection benchmark for Cases 1-4 compared with results obtained with ELEFANT and YACC (from \citealp{Tosi2015}).}
-%  \centering
-%  \resizebox{\textwidth}{!}{
-%\begin{tabular}{|c|ccc|ccc|ccc|ccc|}
-%  \toprule
-%    & \multicolumn{3}{c|}{\textbf{Case 1}} & \multicolumn{3}{c|}{\textbf{Case 2}} & \multicolumn{3}{c|}{\textbf{Case 3}} &
-%    \multicolumn{3}{c|}{\textbf{Case 4}} \\
-%   & FALCON & ELEFANT& YACC & FALCON& ELEFANT& YACC & FALCON & ELEFANT & YACC & FALCON & ELEFANT & YACC \\
-%  \midrule
-%   $\left\langle T \right\rangle$ & 0.7759 & 0.7758 & 0.7767 & 0.6043 & 0.6036 & 0.6058 & 0.7275 & 0.7275 & 0.7286 & & 0.5277 & 0.5289\\
-%    Nu$^{top}$ & 3.4280 & 3.4214 & 3.4298 & 8.5242 & 8.5115 & 8.5278 & 3.0368 & 3.0347 & 3.0374 && 6.5912 & 6.5572 \\
-%    Nu$^{bottom}$ & 3.3168 & 3.313 & 3.3143 & 8.4599 & 8.4465 & 8.3990 & 3.0086 & 2.9908 & 2.9628 && 6.5834 & 6.5243 \\
-%    $v_{\textrm{rms}}$ & 249.901 & 249.134 & 251.7997 & 141.2451 & 140.7087 & 142.2020 & 100.2776 & 100.1208 & 100.9467 && 79.1105 & 79.6202 \\
-%    $v_{\textrm{rms}}^{top}$ & 1.8907 & 1.8642 & 1.8298 & 104.7928 & 104.1228 & 105.1767 & 2.0832 & 2.0652 & 2.0374 && 74.7596 & 75.8414 \\
-%    $u_{max}^{top}$ & 2.6340 & 2.6119 & 2.5516 & 121.9212& 121.766 & 122.3238 & 2.9086 & 2.9019 & 2.8458 && 88.9146 & 89.2940 \\
-%    $\eta_{min}$ & - & - & - & $2.0303\times10^{-5}$ & $2.0600\times10^{-5}$ & $1.9853\times10^{-5}$ & $4.8129\times10^{-5}$ & $4.8080\times10^{-5}$ & $4.7907\times10^{-5}$ &
-%    & $1.9860\times10^{-4}$ & $1.9174\times10^{-4}$ \\
-%    $\eta_{max}$ & - & - & - & 1.3325 & 1.85 & 1.6095 & 0.9360 & 0.9023 & 1 && 1.52 & 1.6773 \\
-%    $\left\langle W \right\rangle$ & 2.4299 & 2.4316 & 2.4583 & 7.5648 & 7.5665 & 7.6222 & 2.0387 & 2.0384 & 2.0400 && 5.6216 & 5.6512 \\
-%    $\left\langle \Phi \right\rangle/Ra$ & 2.4311 & 2.4276 & 2.4333 & 7.5661 & 7.5585 & 7.6194 & 2.0386 & 2.0370 & 2.0335 && 5.6182 & 5.6463 \\
-%    $\delta$ & 0.05\% & 0.16\% & 1.02\% & 0.02\% & 0.11\% & 0.04\% & $<$0.01\% & 0.07\% & 0.32\% && 0.06\% & 0.09\% \\
-%  \bottomrule
-%  \end{tabular}}
-%  \label{tab:visco_mantle}
-%  \end{table}
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{2}{white!100!}{gray!25!}
-\begin{table}[H]
-  \caption{Benchmark cases and corresponding parameters.}
-  \centering
-  \small
-\begin{tabular}{|>{\centering}m{0.08\textwidth}|>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}
-  >{\centering}m{0.05\textwidth} >{\centering}m{0.05\textwidth} |>{\centering\arraybackslash}m{0.2\textwidth}|}
-  \toprule
-    Case & $Ra$ & $\Delta\eta_T$ & $\Delta\eta_y$ & $\eta^*$ & $\sigma_Y$ & Convective regime \\
-  \midrule
-  \rowcolor{gray!25!}
-    1 & $10^2$ & $10^5$ & 1 & - & - & Stagnant lid \\
-    2 & $10^2$ & $10^5$ & 1 & $10^{-3}$ & 1 & Mobile lid  \\
-    \rowcolor{gray!25!}
-    3 & $10^2$ & $10^5$ & 10 & - & - & Stagnant lid  \\
-    4 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 1 & Mobile lid \\
-    \rowcolor{gray!25!}
-    5 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 4 & Periodic \\
-  \bottomrule
-  \end{tabular}
-  \label{tab:tosi_setup}
-  \end{table}
-
-\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{gray!25!}{white!100!}
-\begin{table}
-  \caption{Results of the viscoplastic mantle convection benchmark for different grid resolutions in Cases 1-5 compared with results obtained with ELEFANT
-  and YACC ($100\times100$ elements, from \citealp{Tosi2015}) and in Stone 28 ($32\times32$ elements, from
-  \url{https://github.com/cedrict/fieldstone/tree/master/python_codes/fieldstone_28}).}
-  \centering
-  \resizebox{\textwidth}{!}{
-\begin{tabular}{|cc|ccccccccccc|}
-  \toprule
-  \multicolumn{13}{|c|}{\textbf{Case 1}} \\
-  \midrule
-  Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-  & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-  \hline
-  \rowcolor{gray!25!}
-    YACC & $100\times100$ & 0.7767 & 3.4298 & 3.3143 & 251.7997 & 1.8298 & 2.5516 & - & - & 2.4583 & 2.4333 & 1.02\% \\
-    ELEFANT & $100\times100$ & 0.7758 & 3.4214 & 3.313 & 249.134 & 1.8642 & 2.6119 & - & - & 2.4316 & 2.4276 & 0.16\% \\
-    \rowcolor{gray!25!}
-    Stone 28 & $32\times32$ & 0.7737 & 3.3987 & - & 243.872 & - & - & - & - & - & - & - \\
-%    FALCON & $32\times32$ & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & - & - & 2.4745 & 2.4864 & 0.48\% \\
-%    \rowcolor{gray!25!}
-    FALCON & $100\times100$ & 0.7759 & 3.4280 & 3.3168 & 249.901 & 1.8907 & 2.6340 & - & - & 2.4299 & 2.4311 & 0.05\% \\
-    \midrule
-    \multicolumn{13}{|c|}{\textbf{Case 2}} \\
-    \midrule
-    Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-    & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-    \hline
-    \rowcolor{gray!25!}
-     YACC & $100\times100$ & 0.6058 & 8.5278 & 8.3990 & 142.2020 & 105.1767 & 122.3238 & $1.9853\times10^{-5}$ & 1.6095 & 7.6222 & 7.6194 & 0.04\% \\
-     ELEFANT & $100\times100$ & 0.6034 & 8.5115 & 8.4465 & 140.8067 & 104.1228 & 121.766 & $2.0600\times10^{-5}$ & 1.85 & 7.5665 & 7.5585 & 0.11\% \\
-     \rowcolor{gray!25!}
-     Stone 28 & $32\times32$ & 0.6052 & 8.1715 & - & 141.518 & - & - & - & - & - & - & - \\
-%     FALCON & $32\times32$ & 0.6454 & 8.1422 & 7.4020 & 156.6908 & 95.7226 & 116.9302 & $2.2572\times10^{-5}$ & 0.7194 & 7.4289 & 7.4452 & 0.22\% \\
-%     \rowcolor{gray!25!}
-     FALCON & $100\times100$ & 0.6044 & 8.5266 & 8.4590 & 141.3684 & 104.8867 & 122.0470 & $2.0333\times10^{-5}$ & 1.2914 & 7.5707 & 7.5718 & 0.01\% \\
-     \midrule
-     \multicolumn{13}{|c|}{\textbf{Case 3}} \\
-     \midrule
-     Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-     & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-     \hline
-     \rowcolor{gray!25!} 
-     YACC & $100\times100$ & 0.7286 & 3.0374 & 2.9628 & 100.9467 & 2.0374 & 2.8458 & $4.7907\times10^{-5}$ & 1 & 2.0400 & 2.0335 & 0.32\% \\
-      ELEFANT & $100\times100$ & 0.7275 & 3.0347 & 2.9908 & 100.1208 & 2.0652 & 2.9019 & $4.8080\times10^{-5}$ & 0.9023 & 2.0384 & 2.037 & 0.07\% \\
-      \rowcolor{gray!25!}
-      Stone 28 & $32\times32$ & 0.7277 & 3.0347 & - & 100.018 & - & - & - & - & - & - & - \\
-%      FALCON & $32\times32$ & 0.7281 & 3.0618 & 2.8132 & 101.0092 & 2.1630 & 3.0178 & $4.8500\times10^{-5}$ & 0.7845 & 2.0699 & 2.0695 & 0.02\% \\
-%      \rowcolor{gray!25!}
-      FALCON & $100\times100$ & 0.7275 & 3.0368 & 3.0086 & 100.2776 & 2.0832 & 2.9086 & $4.8129\times10^{-5}$ & 0.9360 & 2.0387 & 2.0386 & $<$0.01\% \\
-      \midrule
-      \multicolumn{13}{|c|}{\textbf{Case 4}} \\
-      \midrule
-      Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-      & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-      \hline
-      \rowcolor{gray!25!} 
-      YACC & $100\times100$ & 0.5289 & 6.5572 & 6.5243 & 79.6202 & 75.4814 & 89.2940 & $1.9174\times10^{-4}$ & 1.6773 & 5.6512 & 5.6463 & 0.09\% \\
-       ELEFANT & $100\times100$ & 0.5277 & 6.5912 & 6.5834 & 79.1105 & 74.7596 & 88.9146 & $1.9860\times10^{-4}$ & 1.5200 & 5.6216 & 5.6182 & 0.06\% \\
-       \rowcolor{gray!25!}
-       Stone 28 & $32\times32$ & 0.5286 & 6.4036 & - & 79.4746 & - & - & - & - & - & - & - \\
-%       FALCON & $32\times32$ & 0.5334 & 6.4679 & 6.3435 & 80.6153 & 76.0511 & 90.2058 & $2.0393\times10^{-4}$ & 0.6453 & 5.6676 & 5.6698 & 0.04\% \\
-%       \rowcolor{gray!25!}
-       FALCON & $100\times100$ & 0.5279 & 6.5897 & 6.5830 & 79.1220 & 75.0715 & 88.8484 & $1.9693\times10^{-4}$ & 1.4133 & 5.6127 & 5.6113 & 0.03\% \\
-       \midrule
-       \multicolumn{13}{|c|}{\textbf{Case 5}} \\
-       \midrule
-       Code & Resolution & $P$ & $\left\langle T \right\rangle_{min}$ & $\left\langle T \right\rangle_{max}$ & Nu$^{top}_{min}$ & Nu$^{top}_{max}$ &
-       $v_{\textrm{rms}}^{min}$ & $v_{\textrm{rms}}^{max}$ & \multicolumn{2}{c}{$(\left\langle \Phi \right\rangle/Ra)_{min}$} & 
-       \multicolumn{2}{c|}{$(\left\langle \Phi \right\rangle/Ra)_{max}$} \\
-       \hline
-       \rowcolor{gray!25!}
-         YACC & $100\times100$ & 0.0792 & 0.6543 & 0.6722 & 2.6950 & 7.2792 & 41.9813 & 98.8230 & \multicolumn{2}{c}{1.4347} & 
-         \multicolumn{2}{c|}{9.2748} \\
-         ELEFANT & $100\times100$ & 0.076 & 0.6514 & 0.6693 & 2.6808 & 7.3506 & 41.2578 & 100.089 & \multicolumn{2}{c}{1.3324} & 
-         \multicolumn{2}{c|}{9.5289} \\
-         \rowcolor{gray!25!}
-         Stone 28 & $32\times32$ & 0.0764 & 0.6521 & 0.6697 & 2.6912 & 7.0763 & 41.61 & 98.183 & \multicolumn{2}{c}{-} & 
-         \multicolumn{2}{c|}{-} \\
-     %    FALCON & $32\times32$ & & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & \multicolumn{2}{c}{-} & 
-     %     \multicolumn{2}{c|}{-} \\
-     %    \rowcolor{gray!25!}
-         FALCON & $100\times100$ & 0.0775 & 0.6519 & 0.6697 & 2.6847 & 7.3629 & 41.5256 & 99.5466 & \multicolumn{2}{c}{1.3399} & 
-         \multicolumn{2}{c|}{9.3993} \\
-  \bottomrule
-  \end{tabular}}
-  \label{tab:visco_mantle}
-  \end{table}
-
-\subsection{Thin layer entrainment}\label{sec:thin}
-This experiment is performed as originally proposed by \citet{vanKeken1997} with two incompressible fluids in a rectangular domain with $L_x=2$ and $L_y=1$
-and gravity acceleration $g_y=10^{10}Ra$. The Rayleigh number and the compositional Rayleigh number are fixed to $Ra=300000$ and $Rc=450000$, respectively.
-Both fluids have constant viscosity, thermal conductivity, specific heat ($\eta=\rho=C_p=1$) and thermal expansion coefficient ($\alpha=10^{-10}$). Fluid 1
-has a density $\rho_1=1$, while fluid 2 is denser ($\rho_2=\rho_1+1.5 \cdot 10^{-10}$) and is placed at the bottom of the domain, for $y \leq 0.025$.
-Temperature are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides of the domain. The initial
-temperature field is given by
-\[T(x,y)=T_u(x,y)+T_l(x,y)+T_r(x,y)+T_s(x,y)-\frac{3}{2}\]
-with
-\begin{eqnarray}
-T_u(x,y)&=&\frac{1}{2}erf\left(\frac{1-y}{2}\sqrt{\frac{u_0}{x}}\right)\nonumber \\
-T_l(x,y)&=&1-\frac{1}{2}erf\left(\frac{y}{2}\sqrt{\frac{u_0}{L_x-x}}\right)\nonumber \\
-T_r(x,y)&=&\frac{1}{2}+\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{y+1}} \exp\left(-\frac{x^2u_0}{4y+4}\right)\nonumber \\
-T_s(x,y)&=&\frac{1}{2}-\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{2-y}} \exp\left(-\frac{(L_x-x)^2u_0}{8-4y}\right)\nonumber
-\end{eqnarray}
-and
-\begin{eqnarray}
-u_0&=&\frac{L_x^{7/3}}{(1+L_x^4)^{2/3}}\left(\frac{Ra}{2\sqrt{\pi}}\right)^{\frac{2}{3}}\nonumber \\
-Q&=&2\sqrt{\frac{L_x}{\pi u_0}}\nonumber
-\end{eqnarray}
-
-The experiment is performed with two grid resolution (\num{125x40} and \num{200x80} elements), with 100 markers per element and CFL of 0.25. $v_{\textrm{rms}}$
-(see Sec. \ref{sec:error}) calculated for both the simulations match well with results from \citet{vanKeken1997} and obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and ELEFANT \citep{Thieulot2014}, obtained with similar grid resolutions
-(Fig. \ref{fig:thin}). In addition, a variety of simulations with different resolutions and aspect ratios are performed to verify that the sensitivity of the
-initial velocity field. Results match well with results from \citet{vanKeken1997} (in the grey area) and \citet{Thieulot2014} (Fig. \ref{fig:thin_initial}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Thin_layer}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_thin.pdf}
-\caption{$v_{\textrm{rms}}$ for the thin layer experiment as function of time for different grid resolution. Results are compared with results obtained by
-\citet{vanKeken1997} (black lines) and with ASPECT and ELEFANT.}
-\label{fig:thin}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrmszero.pdf}
-\caption{$v_{\textrm{rms}}$ at $t=0$ for the thin layer experiment as function of the element size for different aspect ratios of the elements. Results are compared
-with results obtained by \citet{vanKeken1997} (black dots in the grey area) and with ASPECT and ELEFANT.}
-\label{fig:thin_initial}
-\end{figure}
-
-\subsection{Sinking hydrated cylinder}\label{sec:quinquis}
-This experiment is performed as explained in \citet{Quinquis2014} and simulates the sinking of a cold, hydrated cylinder into a warm, dry mantle. The domain
-is square with $L_x=L_y=$ \SI{300}{\km}, a grid resolution of \num{300x300} elements and 25 markers per element. The sinking cylinder is located at $x=$
-\SI{150}{\km} and $y=$ \SI{170}{\km} with a radius of \SI{20}{\km} and has $\rho_c=$ \SI{3250}{\kg\per\cubic\m}, $\eta_c=$ \SI{e23}{\pascal\s}, a thermal
-conductivity $k_c=$ \SI{4.5}{\watt\per\m\per\kelvin},a specific heat $C_{p_c}=$ \SI{1250}{\joule\per\kg\per\kelvin}; the surrounding mantle has $\rho_m=$
-\SI{3200}{\kg\per\cubic\m}, $\eta_m=$ \SI{e20}{\pascal\s}, a thermal conductivity $k_m=$ \SI{105}{\watt\per\m\per\kelvin} and specific heat $C_{p_m}=$
-\SI{1250}{\joule\per\kg\per\kelvin}; a 58 km-thick lithosphere overlies the mantle and has $\rho_l=$ \SI{3200}{\kg\per\cubic\m}, $\eta_l=$ \SI{e23}{\pascal\s},
-a thermal conductivity $k_l=$ \SI{4.5}{\watt\per\m\per\kelvin}, specific heat $C_{p_l}=$ \SI{750}{\joule\per\kg\per\kelvin} and a thermal diffusivity of
-\SI{e-6}{\square\m\per\s}. Velocity boundary conditions are set to free slip on all sides of the domain. Temperature increase linearly in the lithosphere,
-from 0° to 1300°C, and in the mantle, from 1300° to 1360.5°C. The initial temperature of the cylinder is fixed to 400°C. Throughout the evolution, temperature
-is fixed to 0° and 1360.5°C at the top and bottom of the domain, respectively. The initial bound water content is imposed at 2 wt.\% in the cylinder and at
-0 wt.\% in both the mantle and the lithosphere. The maximum amount of water in the mantle is fixed to 0.2 wt.\%, while maximum water content in the cylinder
-and in the lithosphere is function of pressure and temperature and is calculated using a serpentinized harzburgite with Perple\_X, as in \citet{Quinquis2014}.
-
-As shown in \citet{Quinquis2014}, the model is characterised by a progressive dehydration in the external portion of the cylinder, due to the increase of
-temperature, with a consequent vertical migration of free water that hydrates the mantle above the cylinder, up to the lithosphere. The distribution of
-bound and free water during the evolution of the experiment is shown in Fig. \ref{fig:quinquis} (left and right columns, respectively). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Hydration}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Hydration.pdf}
-\caption{Distribution of bound and free water (left and right columns, respectively) at $t=0.5$ and $t=$ \SI{1.25}{\mega\year} (first and second rows,
-respectively) for the sinking hydrated cylinder experiment. Only markers with bound or free water not equal to 0 are plotted.}
-\label{fig:quinquis}
-\end{figure}
-
-\subsection{Mantle melting curves}\label{sec:katz}
-This instantaneous experiment is performed in order to recreate isobaric melting curves, as function of temperature, bulk water and content of water in the
-melt, to compare with melting curves obtained by \citet{Katz2003}. The domain is square with $L_x=L_y=100$ km, a grid resolution of \num{100x100} elements and
-25 markers per element. Density is set in the entire domain to \SI{3300}{\kg\per\cubic\m} and temperature increases from the sides to the centre from 1000° to
-1500°C. The experiment is performed for bulk water contents from 0 to 0.3 wt.\%.
-
-Fig. \ref{fig:fraction} shows isobaric melting curves for \SI{1}{\giga\pascal} (panel a) and \SI{3}{\giga\pascal} (panel b) calculated for different bulk water
-contents. The curves well match with those obtained for the mantle by \citet{Katz2003}, showing the saturation at the solidus for bulk water of 0.2 and 0.3 wt.\%, indicating
-that the code correctly determines melt fraction as function of pressure, temperature and bulk water content. Same results can be observed in Fig.
-\ref{fig:water}a, where isobaric and isothermal melting curves are shown as function of bulk water content. In addition, Fig. \ref{fig:water}b shows that the
-code correctly determines also the percentage of water in the melt as function of pressure, temperature and melt fraction. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Melting}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Fraction.pdf}
-\caption{Isobaric melting curves for 1 GPa (panel a) and 3 GPa (panel b) as function of temperature with different bulk water contents.}
-\label{fig:fraction}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Water.pdf}
-\caption{Isobaric melting curves for 1.5 GPa as function of bulk water (panel a) and water in the melt (panel b) calculated for different temperatures.}
-\label{fig:water}
-\end{figure}
-
-\clearpage
-
-\bibliography{export}
-
-\end{document}

### 2022-09-07 tex/Benchmark/stab.tex
update main

--- 2022-03-31 FALCON_description.tex
@@
-\documentclass[hidelinks,10pt,a4paper]{article}
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{textcomp}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage{booktabs}
-\usepackage{multirow}
-\usepackage[cm]{fullpage}
-\usepackage[round]{natbib}
-\bibliographystyle{abbrvnat}
-\usepackage{doi}
-\usepackage{bm}
-\usepackage[table]{xcolor}
-\hypersetup{
-    colorlinks,
-    linkcolor={blue},
-    citecolor={blue},
-    urlcolor={blue}
-}
-\usepackage{float}
-\restylefloat{table}
-\usepackage{siunitx}
-\DeclareSIUnit\year{yr}
-
-\begin{document}
-
-\section{Introduction}\label{sec:intro}
-This document includes a description of the code FALCON, with a presentation of the features implemented followed by results of extensive benchmarks listed in
-Table \ref{tab:list}, performed to verify that the code solves correctly a wide range of problems. Description and results of all benchmarks are shown
-in Section \ref{sec:benchmark}.
-
-\begin{table}[H]
-\caption{List of benchmarks performed for each feature of the code. Data of all benchmarks can be found at 
-\url{https://github.com/aleregorda/Benchmarks}}
-\centering
-\begin{tabular}{| l | l | l |}
-\toprule
- \textbf{Feature}  & \textbf{Benchmark} & \textbf{Section} \\
-\midrule
-  Solver  & Stokes flow & \ref{sec:stokes} \\
-\hline
-  \multirow{2}{*}{Markers advection} & Zalesak disk & \ref{sec:runge} \\
-    & Conservative Interpolation Velocity & \ref{sec:CVI} \\
-\hline
-  \multirow{4}{*}{Momentum equation}  & Poiseuille flow & \ref{sec:poiseuille}\\
-    & Instantaneous 2D sphere & \ref{sec:ist_sphere} \\
-    & Rayleigh-Taylor experiment & \ref{sec:rayleigh} \\
-    & Falling block & \ref{sec:block} \\
-\hline
-    & 2D Stokes sphere & \ref{sec:sphere} \\
-    Sticky air and  & Stabilisation algorithm & \ref{sec:stab} \\
-    free surface & Topography relaxation & \ref{sec:crameri} \\
-  & Spontaneous subduction & \ref{sec:subduction} \\
-\hline
-  Erosion and sedimentation  & - & - \\
-\hline
-   Non-linear  & Slab detachment & \ref{sec:slab} \\
-   visco-plasticity & Indenter & \ref{sec:indenter} \\
-    & Brick & \ref{sec:brick} \\
-\hline
-  \multirow{3}{*}{Energy equation}  & Advection stabilisation & \ref{sec:advection} \\
-    & Simple shear heating & \ref{sec:simple_shear} \\
-    & Shear and adiabatic heating & \ref{sec:shear} \\
-\hline
-\multirow{3}{*}{Energy + momentum} & Mantle convection & \ref{sec:mantle} \\
-& Viscoplastic mantle convection & \ref{sec:visco_mantle} \\
-& Thin layer entrainment & \ref{sec:thin} \\
-\hline
-  Phase changes  & \multirow{2}{*}{Hydrated sinking cylinder} & \multirow{2}{*}{\ref{sec:quinquis}} \\
-  and hydration  &  & \\
-\hline
-  Melting & Experimental melting curves & \ref{sec:katz} \\
-\bottomrule
-\end{tabular}
-\label{tab:list}
-\end{table}
-
-The code is written in Fortran90 and uses the finite element method (FEM) with quadrilateral $Q_1 \times P_0$ elements (continuous bilinear velocity and
-discontinuous constant pressure), associated with MUMPS\footnote{\url{http://mumps.enseeiht.fr/}} \citep{Amestoy2001,Amestoy2019} that is a software package
-for solving systems of linear equations of the form $A \cdot x = b$, where \textit{A} is a square sparse matrix, by means of a direct method. Since
-$Q_1 \times P_0$ elements do not satisfy Ladyzhenskaya, Babuska and Brezzi (LBB) stability condition \citep{Donea2003}, elemental pressure is elaborated
-in post-processing to avoid spurious pressures by means of a double interpolation to smooth the pressure. Following this procedure, the elemental pressure
-is interpolated onto nodes and then back onto elements. After the smoothing procedure, the pressure field is calculated again on the nodes to be used in
-combination with the lithostatic pressure, which is calculated onto nodes. Correctness of the solution and performances of the solver in terms of time and
-memory usage are tested solving the Stokes flow with the analytical solution proposed by \citet{Donea2003} (Section \ref{sec:stokes}).
-
-The thermo-mechanics of crust-mantle systems is described by means of the conservation of mass, momentum and energy equations, expressed as follows:
-\begin{eqnarray}
-\label{eq:mass}\frac{\partial \rho}{\partial t}+\bm{\nabla} \cdot (\rho \bm{u})&=&0\\
-\label{eq:momentum}-\bm{\nabla} p + \bm{\nabla} \cdot \bm{\tau} + \rho \bm{g}&=&0\\
-\label{eq:energy}\rho C_p \left(\frac{\partial T}{\partial t} + \bm{u} \cdot \bm{\nabla} T \right)&=&\bm{\nabla} \cdot \left(k\bm{\nabla} T\right) + H_{tot}
-\end{eqnarray}
-where $\rho$ is the density, $\bm{u}$ is the velocity, \textit{p} is the pressure, $\bm{\tau}$ is the deviatoric stress, $\bm{g}$ is the gravity acceleration,
-$C_p$ is the specific heat at constant pressure, \textit{T} is the temperature, \textit{k} is the thermal conductivity and $H_{tot}$ is the total internal heat
-production. Density variations due to temperature are generally small enough to assume the density as constant ($\rho=\rho_0$) in Eq. \ref{eq:mass} and in
-Eq. \ref{eq:energy}, while it must be treated as a variable in the buoyancy term of Eq. \ref{eq:momentum}, such that \[\rho=\rho_0(1-\alpha(T-T_0))\]
-where $\rho_0$ is the density at a reference temperature $T_0$ and $\alpha$ is the coefficient of thermal expansion. Eq. \ref{eq:mass} then can be rewritten as
-\begin{equation}\label{eq:mass_0}
-\bm{\nabla} \cdot \bm{u}=0
-\end{equation}
-This simplification is known as the extended Boussinesq approximation.
-
-The time step $dt$ is chosen in according to Courant-Friedrichs-Lewy condition \citep{Anderson1995}, such that
-\[dt=C_n\; \min\left(\frac{\Delta x}{\max|\bm{u}|},\frac{(\Delta x)^2}{\kappa}\right)\]
-where $0<C_n<1$ is the Courant (CFL) number, $\Delta x$ is the minimum dimension of smallest element, $\max|\bm{u}|$ is the maximum velocity calculated on the
-entire domain and
-$\kappa=\frac{k}{\rho C_p}$ is the thermal diffusivity \citep{Thieulot2014}.
-
-\section{Lagrangian markers}\label{sec:markers}
-Elemental properties (density, viscosity, thermal conductivity, specific heat and thermal expansion) needed to solve Eqs. \ref{eq:mass_0}, \ref{eq:momentum}
-and \ref{eq:energy} are related to the composition of each element, determined by means of Lagrangian markers that are characteristic of different materials
-in the domain. Elemental properties (with the sole exception of the viscosity) are calculated using an arithmetic mean, as \[P_e=\frac{1}{n}\sum_{i=1}^n P_i\]
-where $P_e$ is the elemental property, $P_i$ is the property characteristic of the material of each marker and $n$ is the number of markers of the element.
-Differently, the average scheme for the elemental viscosity can be chosen between harmonic, geometric and arithmetic mean. Markers can be placed regularly or
-randomly at the beginning of the simulation and their advection is performed by either a 2nd-order or a 4th-order Runge-Kutta in space, interpolating the velocity 
-field on each marker by means of the shape functions (see Section \ref{sec:runge}). The interpolated velocity is then corrected by means of the Conservative
-Velocity Interpolation (CVI), which introduces a corrective term to reduce dispersion and clustering of particles in both steady state and time-dependent 
-\citep{Wang2015} (see Section \ref{sec:CVI}).
-
-The initial distribution of the markers is created using the open-source code library
-Geodynamic World Builder \citep{Fraters2019}. During the simulation, each marker carries memory of temperature, pressure and accumulated strain, which are
-determined interpolating the relative nodal parameter. In particular, pressure is interpolated onto each marker after the smoothing procedure. The number
-of markers contained in each element is maintained between $n_{min}$ and $n_{max}$. When in an element there are less markers than $n_{min}$ the code adds
-random markers to reach the $n_{min}$, while if the number is higher than $n_{max}$ some of them are randomly deleted. When new markers are added, they assume
-the properties of the nearest marker. In this way elements are never empty and maintain a number of markers inside a prefixed range.
-
-\section{The momentum equation}\label{sec:mom_eq}
-The penalty method is implemented so that Eq. \ref{eq:mass_0} can be written as:
-\begin{equation}\label{eq:mass_penalty}
-\bm{\nabla} \cdot \bm{u} + \frac{p}{\lambda}=0
-\end{equation}
-where $\lambda$ is the so-called penalty parameter, which should be 6-7 orders of magnitude larger than the shear viscosity to ensure that mass conservation
-is satisfied \citep{Donea2003,Thieulot2014}.
-
-The deviatoric stress tensor in Eq. \ref{eq:momentum} can be written in terms of the strain rate tensor as $\bm{\tau}=2\eta\dot{\bm{\epsilon}}$, with
-$\dot{\bm{\epsilon}}=\frac{1}{2}\left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)$. Therefore, Eq. \ref{eq:momentum} can be rewritten as:
-\begin{equation}\label{eq:momentum_pressure}
--\bm{\nabla} p + \bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right) + \rho \bm{g} = 0
-\end{equation}
-Finally, using pressure from Eq. \ref{eq:mass_penalty}, Eq. \ref{eq:momentum_pressure} can be rewritten as:
-\begin{equation}\label{eq:momentum_penalty}
-\lambda \bm{\nabla} \left(\bm{\nabla} \cdot \bm{v}\right)+\bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right)+\rho\bm{g}=0
-\end{equation}
-The penalty method is been associated to the iterative Uzawa method, as described in detail in \citet{Dabrowski2008} and \citet{Thieulot2014}
-(see Section \ref{sec:poiseuille}). Results of benchmarks performed to verify the correctness in the implementation of Eq. \ref{eq:momentum_penalty} are
-shown in Sections \ref{sec:ist_sphere} and \ref{sec:rayleigh}.
-
-To support large viscosities variations the penalty parameter is related to the effective elemental viscosity by means of a dimensionless coefficient, so that
-$\lambda=\lambda_e(e)\eta_{eff}(e)$ \citep{Marotta2006,Dabrowski2008,Thieulot2014}. Benchmark of the falling block \citep{Gerya2003a,Gerya2010b,Thieulot2011}
-is performed to verify that the code can correctly deal buoyancy driven flows with strong viscosity contrasts (Section \ref{sec:block}).
-
-\section{Sticky air and free surface}\label{sec:sticky}
-The Earth's surface can be treated by means of either the so-called sticky-air or a true free surface method, both of them implemented in the code. In the
-sticky-air method the surface is approximated with the introduction of a buoyant layer with a viscosity at least four orders of magnitude lower than the crust
-\citep{Schmeling2008,Crameri2012} and the interface between lithosphere and air is defined using a chain of passive markers that are advected as the Lagrangian
-markers. The correctness of the evolution of the markers chain is tested with the experiment of a 2D time-dependent Stokes sphere below a free surface and
-compared with results from ASPECT\footnote{\url{https://aspect.geodynamics.org/}} \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} (Section
-\ref{sec:sphere}). In the true free surface case the top boundary is assumed stress-free and velocities are not fixed. In this case topography variations are
-described by vertical deformations of the mesh that depend on the velocity field of the nodes that identify the free surface, while horizontal deformations are
-not taken into account. This procedure is known as the Arbitrary Lagrangian-Eulerian (ALE) method and its implementation follows the technique described in
-\citet{Thieulot2011}. However, although the implementation of a true free surface better reproduces laboratory experiment, extremely small time steps can be
-necessary to maintain stability \citep{Kaus2010a,Quinquis2011,Thieulot2014}. Therefore, the stability algorithm proposed by \citet{Kaus2010a} is implemented
-to avoid instabilities due to high density differences at the free surface when using too large time steps. The implementation of this algorithm are tested by
-performing the experiment described by \citet{Kaus2010a} (Section \ref{sec:stab}).
-
-The topography relaxation benchmark proposed by \citet{Crameri2012} is performed to verify that the code correctly recovers topography variations in case of
-both the sticky-air and the true free surface method (Section \ref{sec:crameri}). Finally, Section \ref{sec:subduction} show the results of the spontaneous
-subduction experiment described by \citet{Schmeling2008}.
-
-\section{Erosion and sedimentation}\label{sec:erosion}
-Surface processes at either the lithosphere-air interface or in correspondence of the free surface have been implemented by means of the software
-Fastscape\footnote{\url{https://fastscape.org/}} \citep{Braun2013,Cordonnier2019,Yuan2019a,Yuan2019b}. FastScapeLib is a set of routines that solve the stream
-power law (enriched by a sediment transport/deposition term) \citep{Yuan2019a}, hillslope diffusion and marine transport and deposition \citep{Yuan2019b}.
-The partial differential equation solved by FastScapeLib is:
-\begin{equation}
-\frac{\partial h}{ \partial t}=U-K_f A^{m_e}S^{n_e} + \frac{G}{A}\int_A \left( U-\frac{\partial h}{\partial t} \right) dA + K_d \bm{\nabla}^2 h 
-\end{equation}
-where $h$ is the topography, $U$ is the uplift, $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $S$ is the slope, $m_e$ and $n_e$ are
-parameters that should be chosen to constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}, $G$ is the dimensionless
-deposition/transport coefficient for the enriched stream power law and $K_d$ is the effective diffusivity. Terms $K_f A^{m_e}S^{n_e}$ and $K_d \bm{\nabla}^2 h$
-consider long-range processes, described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, and short-range processes,
-proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}.
-
-In case of sticky air, slope is determined by means of the markers chain, which is vertically corrected considering erosion and sedimentation rates. After the
-correction, continental markers above the markers chain are transformed in air markers and air markers below the markers chain become sediments. Similarly,
-slope is determined considering the nodes on the top boundary in case of true free surface and they are vertically corrected as for the markers chain. In this
-case, continental markers above the top boundary are simply deleted.
-%Surface processes at the lithosphere-air interface are implemented taking into account fluvial erosion (long-range processes), hillslope processes (short-range
-%processes) and failure of slopes exceeding a threshold angle $\theta$ \citep{Burov1997,Tucker1999,Sacek2012,Croissant2014,Ueda2015,Beucher2020}. Erosion rates
-%$\dot{E}_{lr}$ due to long-range processes can be described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{lr}=K_f A^{m_e} \bm{\nabla} h^{n_e}
-%\end{equation}
-%where $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $\bm{\nabla} h$ is the slope and $m_e$ and $n_e$ are parameters chosen in to
-%constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}. Erosion rates $\dot{E}_{sr}$ due to short-range processes are
-%implemented by assuming that transport is linearly proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{sr}=K_d \bm{\nabla}^2 h
-%\end{equation}
-%where $K_d$ is the effective diffusivity. In addition, the slope is modified if it still exceeds a threshold angle after long- and short-range processes
-%\citep{Ueda2015}.
-
-%Offshore sedimentation processes are implemented similarly to hillslope processes but with a depth-dependent coefficient \citep{Sacek2012,Beucher2020}.
-%The sedimentation rate $\dot{S}$ is calculated as
-%\begin{equation}
-%\dot{S}=K_s \bm{\nabla}^2 h
-%\end{equation}
-%with
-%\begin{equation}
-%K_s=K^*_s \exp\left(\frac{h-h_{sl}}{h_0}\right)
-%\end{equation}
-%where $K^*_s$ is a constant, $h_{sl}$ is the sea level and $h_0$ is a characteristic length. In case of sticky air, slope is determined by means of the markers
-%chain, which is vertically corrected considering erosion and sedimentation rates. After the correction, continental markers above the markers chain are
-%transformed in air markers and air markers below the markers chain become sediments. Similarly, slope is determined considering the nodes on the top boundary
-%in case of true free surface and they are vertically corrected as for the markers chain. In this case, continental markers above the top boundary are simply
-%deleted.
-
-\section{Non-linear rheologies}\label{sec:nonlinear}
-Non-linear rheologies are implemented combining viscous creep (dislocation and diffusion) and plastic yielding. For each marker, diffusion and dislocation
-viscosity ($\eta_{df|ds}$) can be determined as follows \citep{Karato1993,Warren2008b,Wang2016}:
-\begin{equation}\label{eq:viscous}
-\eta_{df|ds}=f_s \left(\frac{d^m}{A}\right)^{\frac{1}{n}} \left (\frac{2^{\frac{1-n}{n}}}{3^{\frac{n+1}{2n}}} \right)
-I_2^{\frac{(1-n)}{n}}\exp\left(\frac{Q+pV}{nRT}\right)
-\end{equation}
-where $f_s$ is a scaling factor used to represent lithologies that are stronger or weaker than the base set, $d$ is the grain size, $m$ is the grain size exponent,
-$n$ is the stress exponent, $A$ is the uniaxial pre-exponential factor, $I_2$ is the square root of the second invariant of the strain rate tensor, $Q$ is the
-activation energy, $p$ is the pressure, $V$ is the activation volume, $R$ is the gas constant and $T$ is the temperature.
-Pressures and temperatures are determined by the interpolation of the nodal parameters, while strain rates are calculated by means of the
-derivative of the velocities on the nodes. In case of diffusion creep $n=1$ and $m>0$, in case of dislocation creep $n>1$ and $m=0$. The viscous creep $\eta_{cp}$
-is then calculated as the harmonic average between $\eta_{df}$ and $\eta_{ds}$:
-\begin{equation}\label{eq:visco_creep}
-\eta_{cp}=\left(\frac{1}{\eta_{df}}+\frac{1}{\eta_{ds}}\right)^{-1}
-\end{equation}
-The implementation of the non-linear viscous creep viscosity depending on the strain rate is tested by means of the slab detachment benchmark
-(Section \ref{sec:slab}).
-
-Plastic yielding is implemented rescaling $\eta_{cp}$ in order to limit the stress below the yield stress $\sigma_y$ \citep{Thieulot2008,Thieulot2014,Glerum2018},
-obtaining
-\begin{equation}\label{eq:plastic}
-\eta_{pl}=\frac{\sigma_y}{2I_2}
-\end{equation}
-where $\eta_{pl}$ is the plastic viscosity and the yield stress is determined following the Drucker-Prager criterion, such as
-\begin{equation}\label{eq:yield}
-\sigma_y=C\;\cos(\phi)+p\;\sin(\phi)
-\end{equation}
-where $C$ is the cohesion and $\phi$ is the internal friction angle. In case of negative pressure it is imposed equal to 0, so that negative yield stress are
-excluded. The correctness of the non-linear solution in case of plastic viscosity with variable internal friction angle is verified performing indenter and
-brick experiments (Sections \ref{sec:indenter} and \ref{sec:brick}, respectively).
-
-Strain softening is taken into account for both viscous creep and plastic viscosity \citep{Huismans2003,Babeyko2005,Huismans2005,Sobolev2005,Warren2008a} by
-means of the accumulated strain $\epsilon$ memorised by each marker. Softening in the viscous creep determines a linear decrease of $\eta_{vc}$ by means of a
-viscous strain softening factor $W_S$ that increases linearly from $W_{S_0}$ to $W_{S_{\infty}}$ for $\epsilon_{S_0}<\epsilon_S<\epsilon_{S_{\infty}}$ 
-\citep{Warren2008a}. This viscous softening can be related to strain-induced grain size reduction \citep{Warren2008a}. Differently, plastic softening is 
-simulated with a linear decrease of internal friction angle $\phi(\epsilon)$ and cohesion $C(\epsilon)$ in according to
-\begin{eqnarray}
-\label{eq:friction1}\phi(\epsilon)&=&\phi_0+(\phi_{\infty}-\phi_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}\\
-\label{eq:friction2}C(\epsilon)&=&C_0+(C_{\infty}-C_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}
-\end{eqnarray}
-where $\phi_0$, $C_0$ and $\phi_{\infty}$, $C_{\infty}$ are internal friction angle and cohesion for $\epsilon_0$ and $\epsilon_{\infty}$, respectively
-\citep{Huismans2003,Huismans2005,Warren2008a,Thieulot2014}. Plastic softening approximates deformation-induced softening of faults and brittle shear zones
-\citep{Warren2008a}.
-
-Plastic yielding and viscous creep are then combined to obtain a viscoplastic viscosity $\eta_{vp}$ as follows, assuming that they are independent processes
-\citep{Karato2008,Glerum2018}:
-\begin{equation}\label{eq:viscoplastic}
-\eta_{vp}=\min(\eta_{cp},\eta_{pl})
-\end{equation}
-Finally, effective viscosity $\eta_{eff}$ is capped by the minimum and the maximum viscosity ($\eta_{min}$ and $\eta_{max}$, respectively) to avoid extremely
-low or high viscosity \citep{Glerum2018} as follows
-\begin{equation}\label{eq:effective}
-\eta_{eff}=\min(\max(\eta_{vp},\eta_{min}),\eta_{max})
-\end{equation}
-Viscous creep and plastic yielding are non-linear rheologies because of their dependence on the velocity field through pressure and strain rates. Therefore,
-the solution is determined by means of Picard-type iterations, until convergence of the velocity field \citep{Glerum2018}. The convergence is verified at each
-iteration $i$ via the nonlinear residual $\bm{\mathcal{R}}^i$ that can be determined as
-\begin{equation}\label{eq:convergence}
-\bm{\mathcal{R}}^i= \mathbb{K} (\eta_{eff}(\dot{\bm{\epsilon}}^{i-1},p^{i-1}))\cdot \bm{v}^{i-1} - \bm{f}^i
-\end{equation}
-%\begin{equation}\label{eq:convergence}
-%\chi=1-\frac{\langle (\bm{v}^i-\langle\bm{v}^i\rangle)\cdot (\bm{v}^{i+1}-\langle\bm{v}^{i+1}\rangle) \rangle}{\sqrt{\sigma^i\sigma^{i+1}}}
-%\end{equation}
-%Iterations are performed until either convergence ($\chi<tol$) or maximal number of non-linear iterations ($it_{max}$) is reached.
-where $\mathbb{K}$ is the Stiffness matrix and $\bm{f}$ is the right hand side vector \citep{Spiegelman2016,Glerum2018}.  The $L_2$-norm is extracted from
-$\bm{\mathcal{R}}^i$ and it is normalised as follows
-\begin{equation}\label{eq:convergence_res}
-\frac{||\bm{\mathcal{R}}^i||_2}{||\bm{\mathcal{R}}^0||_2}
-\end{equation}
-where $||\bm{\mathcal{R}}^0||_2$ is the $L_2$-norm of the first nonlinear iteration.
-Since often the normalised nonlinear residual drop very quickly over the first 2-3 iterations, also the $L_2$-norm of velocity and pressure residuals are
-calculated as
-\begin{equation}\label{eq:convergence_vel}
-\frac{||\Delta\bm{v}||_2}{||\bm{v}||_2}=\frac{||\bm{v}_i-\bm{v}_{i-1}||_2}{||\bm{v}_{i}||_2}
-\end{equation}
-and
-\begin{equation}\label{eq:convergence_press}
-  \frac{||\Delta p||_2}{||p||_2}=\frac{||p_i-p_{i-1}||_2}{||p_{i}||_2}
-\end{equation}
-respectively. Iterations are performed until either maximal number of non-linear iterations ($it_{max}$) is reached or all the normalised nonlinear residual, 
-the velocity residual and the pressure residual converge under a defined tolerance $tol$.
-
-\section{The energy equation}\label{sec:en_eq}
-The stabilisation of the advection term of the energy equation (Eq. \ref{eq:energy}) needed to avoid possible oscillations in the thermal solution in the case
-when advection dominates over diffusion, is implemented by means of a streamline-upwind Petrov–Galerkin (SUPG) method, for which the advection term is modified
-follows the discussion in \citet{Thieulot2011} and \citet{Thieulot2014} (see Section \ref{sec:advection} for benchmark). The total internal heat production
-$H_{tot}$ in Eq. \ref{eq:energy} includes radiogenic heating $H_r$, shear heating $H_s$ and adiabatic heating $H_a$, where
-\begin{equation}\label{eq:shear}
-H_s=2\eta \dot{\bm{\epsilon}} : \dot{\bm{\epsilon}} = 2\eta(\dot{\epsilon}_{xx}^2+\dot{\epsilon}_{yy}^2+2\dot{\epsilon}_{xy}^2)
-\end{equation}
-and
-\begin{equation}\label{eq:adiabatic}
-H_a=T \alpha \frac{Dp}{Dt} \approx -\alpha T \rho g_y v
-\end{equation}
-The correctness of the implementation and the computation of shear and adiabatic heating is verified performing an exercise with an analytical solution
-(Section \ref{sec:simple_shear}) and Exercise 9.4 in \citet{Gerya2010b} (Section \ref{sec:shear}). Furthermore, mantle convection and thin layer entrainment
-benchmarks (Sections \ref{sec:mantle} and \ref{sec:thin}, respectively) are performed to verify that the code solve correctly Eqs. \ref{eq:momentum_penalty}
-and \ref{eq:energy} in case of a temperature field characterised by an initial perturbation and with a temperature-dependent density. At last, the viscoplastic
-mantle convection benchmark is performed to check also the viscoplasticity in case of a initial perturbed temperature field (Section \ref{sec:visco_mantle}).
-
-\section{Phase transitions and hydration}\label{sec:phase}
-Variations of effective density, specific heat and coefficient of thermal expansion during the evolution of crustal and mantle materials at different
-pressure-temperature ($p$-$T$) conditions are computed by means of version 6.8.6 of Perple\_X software package \citep{Connolly2005}, similarly to the
-implementation described in \citet{Marotta2020}. By default, mineral assemblages and properties of each lithology (oceanic crust, lower and upper continental
-crust, sediments and dry or hydrated mantle) are calculated for temperatures from \SIrange{330}{1600}{\kelvin}, with increments of approximately \SI{4}{\kelvin},
-and pressures from \SIrange{0.1}{30}{\giga\pascal}, with increments of approximately \SI{0.03}{\giga\pascal}, for a total of almost 100000 points.
-
-The introduction of phase transitions produces two main effects that must be taken into account in numerical models: variations in density and release/absorption
-of latent heat \citep{Tackley2010}. Effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into
-account considering effective coefficient of thermal expansion \citep{Christensen1985,Zhong2015}. Similarly, effects of latent heat in the energy equation
-(eq \ref{eq:energy}) are taken into account by considering effective specific heat and coefficient of thermal expansion
-\citep{Christensen1985,Tackley2010,Zhong2015}. Effective density, specific heat and coefficient of thermal expansion are included in Perple\_X files and are
-assumed by each Lagrangian marker in according to its $p$-$T$ conditions.
-
-Hydration processes related to dehydration of subducting lithosphere \citep{Schmidt1998,Liu2007,Faccenda2009,Faccenda2010,Faccenda2014,Rosas2016} strongly
-influence the thermo-mechanics inside the mantle wedge, mainly because of both weakening effects on the mantle rheology and density variations in case of
-mantle serpentinization \citep{Gerya2002,Honda2003,Arcay2005,Roda2010,Regorda2017}. In the case that hydration is switched on, the amount of bound and free
-water is memorised by each marker, following the implementation of \citet{Quinquis2014}; hydration and dehydration processes are related to the amount of bound
-water of each marker and to the maximum amount of water it can transport, i.e. if the amount of bound water exceeds the maximum amount of water, the marker
-dehydrates and releases free water that can hydrates under-saturated markers. Maximum water content of each marker is determined as function of lithology and
-$p$-$T$ conditions and it is calculated using Perple\_X, in the same way than effective density, specific heat and coefficient of thermal expansion. Bound water
-is advected together with the markers, neglecting the effect of bound water diffusion, while free water simply migrates vertically and is not coupled to the
-solid-phase flow of the mantle wedge \citep{Arcay2005,Quinquis2014}. The correctness of the migration scheme of the free water and its absorption into bound
-water by under-saturated markers are verified performing the experiment described in \citet{Quinquis2014} (Sec. \ref{sec:quinquis}).
-
-The mantle viscosity weakening related to the amount of water has been extensively studied \citep{Chopra1981,Hirth2003} and it is implemented in according to
-\citet{Arcay2005} and \citet{Horiuchi2016}, as follows
-\begin{equation}\label{eq:wet_visc}
-\eta_{wet}=\eta_{dry}\left(\left[1-\frac{1}{f_v}\right]\exp\left(-\frac{[OH^-]}{[OH^-]_0}\right)+\frac{1}{f_v}\right)
-\end{equation}
-where $f_v$ is the viscosity reduction factor between dry and wet conditions, $[OH^-]$ is the water content and $[OH^-]_0$ is a reference water content set to
-620 ppm (0.062 wt.\%) \citep{Arcay2005}. By default, $f_v$ is set to 100, in according to \citet{Arcay2005} and \citet{Horiuchi2016}.
-
-\section{Mantle melting}\label{sec:melt}
-Melt fraction in the mantle depends on temperature ($T$, in \textdegree C), pressure ($p$, in \si{\giga\pascal}) and the water content in the melt ($X_{H_2O}$, in wt.\%).
-The determination of the melt fraction $M$ is implemented as explained by \citet{Katz2003}, according to the successive modification by \citet{Langmuir2006} and
-\citet{Kelley2010}, as follows
-\begin{eqnarray}\label{eq:melt_fr}
-M(p,T,X_{H_2O})&=&-T+T_{sol}+(x_t\cdot ln(p+y_t))\cdot X_{H_2O}-\nonumber\\
-&&+K_m\left(\frac{X_{H_2O}^{bulk}}{D_{H_2O}(1-X_{H_2O})+X_{H_2O}}\right)^{\gamma}
-\end{eqnarray}
-where $T_{sol}$ is the temperature of the dry solidus, $D_{H_2O}$ is the partition coefficient and $x_t\cdot log(p+y_t)$ indicates the pressure dependence of
-the melting curve \citep{Kelley2010}. $x_t$, $y_t$, $K_m$ and $\gamma$ are parameters chosen in according to \citet{Kelley2010}.
-However, $X_{H_2O}$ depends on the melt fraction as
-\begin{equation}\label{eq:melt_wt}
-X_{H_2O}(M)=\frac{X_{H_2O}^{bulk}}{D_{H_2O}+M(1-D_{H_2O})}
-\end{equation}
-and a numerical solution can be found using a root-finder method \citep{Katz2003,Wang2016}. Furthermore, the content of water in the melt is limited by the
-pressure-dependent saturation concentration of water in the melt, determined as
-\begin{equation}\label{eq:melt_max}
-X_{H_2O}^{sat}=\chi_1 p^{\lambda_m}+\chi_2 p
-\end{equation}
-where $\chi_1$, $\chi_2$ and $\lambda_m$ are parameters chosen according to \citet{Katz2003}, \citet{Langmuir2006} and \citet{Kelley2010}. The correctness in
-the determination of melt fractions and water contents in the melt are tested comparing model results with experimental curves obtained by \citet{Katz2003}
-(Sec. \ref{sec:katz}).
-
-In case of melting, viscous creep viscosity ($\eta_{vp}$) is modified according to \citet{Wang2016}, as
-\begin{equation}\label{eq:melt_visc}
-\eta_{melt}=\eta_{vp}\cdot \exp(\alpha_m M)
-\end{equation}
-where $\alpha_m$ are melt fraction factors for dislocation and diffusion creep, chosen in according to \citet{Wang2016}.
-
-Effective density ($\rho_{eff}$) for partially molten rocks is calculated as
-\begin{equation}\label{eq:melt_dens}
-\rho_{eff}=\rho_s(1-M)+\rho_m M
-\end{equation}
-where $\rho_s$ and $\rho_m$ are the densities of the solid and the molten rock, respectively \citep{Gerya2004b,Gerya2007,Wang2016}. The density of the
-solid rock is extracted by Perple\_X, while the density of the molten portion is calculated as
-\begin{equation}\label{eq:melt_dens_M}
-\rho_m(p,T)=\rho_0[1-\alpha(T-T_0)][1+\beta(p-p_0)]
-\end{equation}
-where $\rho_0$ is the density at the reference temperature ($T_0$) and pressure($p_0$), and $\alpha$ and $\beta$ are the thermal expansion and compressibility
-coefficients, respectively \citep{Gerya2004b,Gerya2007,Wang2016}.
-
-As for the phase transitions, effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into account
-considering effective coefficient of thermal expansion. Similarly, latent heat due to melting/crystallisation can be implicitly included in the energy equation
-(Eq. \ref{eq:energy}) by considering effective specific heat ($C_{p_{eff}}$) and thermal expansion ($\alpha_{eff}$) of the partially molten rocks, in according to
-\begin{eqnarray}
-\label{eq:cp_effective}C_{p_{eff}}&=&C_p+H_L\left(\frac{\partial M}{\partial T}\right)_p \\
-\label{eq:alpha_effective}\alpha_{eff}&=&\alpha+\rho\frac{H_L}{T}\left(\frac{\partial M}{\partial p}\right)_T
-\end{eqnarray}
-where $C_p$ and $\alpha$ are the specific heat and the coefficient of thermal expansion of the solid rock, respectively, and $H_L$ is the latent heat of the
-molten rock \citep{Gerya2004b,Gerya2007,Tackley2010}.
-
-\section{Benchmarks}\label{sec:benchmark}
-\subsection{Error measurements}\label{sec:error}
-In order to determine the accuracy of velocity and pressure field of the benchmarks, the $L_2$-norm is computed by numerical integration on the Gauss-Legendre
-quadrature points. $L_2$-norm for pressure and velocity errors can be evaluated as
-\begin{eqnarray}
-\textrm{err}_p&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|p^n_i(r_q)-p^a_i(r_q)|^2\right)w_q|J_q|}\nonumber \\
-\textrm{err}_v&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|u^n_i(r_q)-u^a_i(r_q)|^2+|v^n_i(r_q)-v^a_i(r_q)|^2\right)w_q|J_q|}\nonumber
-\end{eqnarray}
-respectively, where $n_e$ is the number of elements, $n_q$ is the number of quadrature points per element, $p^n_i(r_q)$ and $p^a_i(r_q)$ are the numerical and
-analytical pressure, respectively, in each quadrature point $q$, $w_q$  and $J_q$ are the weight and the Jacobian at the quadrature point
-$q$, $u^n_i(r_q)$, $v^n_i(r_q)$, $u^a_i(r_q)$ and $v^a_i(r_q)$ are the numerical and analytical velocities, respectively, in each quadrature point $q$.
-
-Other quantities used as comparison with original benchmarks are the root-mean-square velocity over the whole domain and over the surface
-\begin{eqnarray}
-v_{\textrm{rms}}&=&\sqrt{\int_{\Omega}|\bm{v}|^2 d\Omega}\nonumber \\
-v_{\textrm{rms}}^{top}&=&\sqrt{\int_0^1 u^2 \bigg|_{y=1} dx}\nonumber
-\end{eqnarray}
-respectively, top and bottom Nusselt numbers
-\[\textrm{Nu}^{top/bottom}=-\int_0^1\frac{\partial T}{\partial y} \bigg|_{y=1/y=0} dx\]
-the average rate of work done against gravity
-\[\left\langle W \right\rangle=\int_{\Omega}T u_y d\Omega\]
-the average rate of viscous dissipation
-\[\left\langle \Phi \right\rangle=\int_{\Omega}\tau_{ij}\dot{\epsilon}_{ij} d\Omega\]
-and the percentage error between $\left\langle W \right\rangle$ and $\left\langle \Phi \right\rangle/Ra$
-\[\delta=\frac{\left|\left\langle W \right\rangle -\frac{\left\langle \Phi \right\rangle}{Ra} \right|}{\max \left(\left\langle W \right\rangle,
-\frac{\left\langle \Phi \right\rangle}{Ra}\right)}\times100\]
-
-\subsection{Stokes flow}\label{sec:stokes}
-The problem consists of determining velocity field $(u,v)$ and pressure $p$ in case of a manufactured solution with prescribed body forces such as
-\begin{eqnarray}
-b_1&=&(12-24y)x^4+(-24+48y)x^3+(-48y+72y^2-48y^3+12)x^2+\nonumber \\
-&&+(-2+24y-72y^2+48y^3)x+1-4y+12y^2-8y^3\nonumber \\
-b_2&=&(8-48y+48y^2)x^3+(-12+72y-72y^2)x^2+(4-24y+48y^2-48y^3+\nonumber \\
-&&+24y^4)x-12y^2+24y^3-12y^4\nonumber
-\end{eqnarray}
-for which the exact solution is:
-\begin{eqnarray}
-u(x,y)&=&x^2(1-x)^2(2y-6y^2+4y^3)\nonumber \\
-v(x,y)&=&-y^2(1-y)^2(2x-6x^2+4x^3)\nonumber \\
-p(x,y)&=&x(1-x)-\frac{1}{6}\nonumber
-\end{eqnarray}
-The domain is a unit square a constant viscosity ($\eta=1$) and the penalty parameter is set to $\lambda=10^7$. Velocity boundary conditions are set to no slip
-$(\bm{v}=\bm{0})$ on all boundaries. The problem is performed for different grid resolution between \num{8x8} and \num{1024x1024} elements.
-
-The errors between the analytical solution and the numerical prediction of the velocity and the pressure field are calculated as in Section \ref{sec:error}.
-Fig. \ref{fig:errors}a shows that both velocity and pressure field converge to the exact solution with the decrease of the element size, following the
-theoretical convergence rate. The convergence of pressure, in contrast with the observation by \citet{Donea2003} for $Q_1 \times P_0$ elements, indicate the
-effectiveness of the smoothing procedure. Smoothed pressure field for grid resolution of \num{128x128} elements is shown in Fig. \ref{fig:errors}b (continuous
-red line), in comparison with the analytical solution (dashed blue line). Solve times and memory usage needed to generate the solution are shown in
-Fig. \ref{fig:MUMPS}. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Solver/Stokes_Flow}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/errors.pdf}
-\caption{Velocity and pressure error for the Stokes flow experiment between generated and analytical solution as a function of element size (panel a),
-and comparison between smoothed pressure and analytical solution as function of $x$ coordinate for a grid resolution of \num{128x128} elements (panel b).}
-\label{fig:errors}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/MUMPS.pdf}
-\caption{Panel a) Analysis, factorisation, solution and total solve times (panel a) and factorisation memory usage (panel b) as a function of the total number
-of degrees of freedom for the Stokes flow experiment.}
-\label{fig:MUMPS}
-\end{figure}
-
-\subsection{Markers advection}\label{sec:runge}
-The 2nd-order Runge-Kutta advection scheme in space is tested by means of the Zalesak disk test \citep{Zalesak1979,Thieulot2014}. The benchmark is performed
-in a unit square domain with a grid resolution of \num{32x32} elements and 1000000 markers and values of Courant number between 0.25 and 1.
-At $t=0$, the disk is centred at position $(0.5;0.75)$ with a radius $R=0.15$ and has a vertical fissure 0.05 wide and 0.2 high (orange in Fig. \ref{fig:runge})
-and with the velocity field prescribed in the entire domain as
-\begin{eqnarray}
-u(x,y)&=&2\pi\left(y-\frac{L_x}{2}\right)\nonumber \\
-v(x,y)&=&-2\pi\left(x-\frac{L_x}{2}\right)\nonumber
-\end{eqnarray}
-the markers are back to their initial location after a $2\pi$ rotation ($t=1$). The clockwise rotation of the disk at $t=0.25$, 0.5 and 0.75 
-is shown in Fig. \ref{fig:runge} (red, blue and green, respectively). The distance from the centre of 1 marker is calculated throughout the rotation to
-evaluate the error for different values of Courant number. As expected, the error increase with the increasing of the time step, as shown in 
-Fig. \ref{fig:runge_err}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/Zalesak_Disk}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=200px]{./Figures/Zalesak.pdf}
-\caption{Position of the Zalesak disk at $t=0$, 0.25, 0.5 and 0.75 (orange, red, blue and green, respectively) for the case with Courant number of 1.}
-\label{fig:runge}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/zal_marker.pdf}
-\caption{Distance from the centre as function of time for values of Courant number of  of 0.25, 0.3, 0.5, 0.75 and 1 (green, orange, blue, black and red,
-respectively). R indicates the distance from the centre considering a perfect circle trajectory.}
-\label{fig:runge_err}
-\end{figure}
-
-\subsection{Conservative Velocity Interpolation (CVI)}\label{sec:CVI}
-The CVI correction is checked by means of the Stokes flow experiment by \citet{Donea2003}, as presented in \ref{sec:stokes}. The advection of Lagrangian
-markers is performed using either a 2nd-order or a 4th-order Runge-Kutta scheme with CFL=0.5 and an initial random distribution of 25 markers per element.
-Fig. \ref{fig:CVI} shows the minimum and maximum values of markers per element throughout the experiment with and without the CVI correction and using
-different Runge-Kutta schemes. The introduction of the CVI correction allows the
-tests to run for times longer than $t=2000$ without show empty elements, while tests without the CVI correction stop at approximately $t=600$, when at least
-one element goes empty. Fig. \ref{fig:CVI_time} shows the different number of markers per element and the markers distribution at $t=600$ obtained without and
-with the CVI correction (panels b and c, respectively) in case of a 2nd-order Runge-Kutta scheme. All data can be found at
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/CVI}.
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/CVI.pdf}
-\caption{Maximum and minimum number of markers per element throughout the experiment, with (green and orange lines) or without (purple and blue lines)
-the CVI correction and using either a 2nd-order (blue and orange lines) or a 4th-order (purple and green lines) Runge-Kutta scheme.}
-\label{fig:CVI}
-\end{figure}
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/CVI_time.pdf}
-\caption{Number of markers per element and markers distribution through time using 2nd.order Runge-Kutta scheme.
-Panels a: number of markers per element (panel a1) and distribution of the markers (panel a2) after the first
-iteration; panels b and c: comparison between number of markers per element (panels b1 and c1) and markers distribution (panels b2 and c2) without and with
-the CVI correction at $t=600$; panels d: number of markers per element (panel d1) and markers distribution (panel d2) in case of CVI correction at $t=2000$.}
-\label{fig:CVI_time}
-\end{figure}
-
-\subsection{Poiseuille flow}\label{sec:poiseuille}
-The domain is a rectangle with $L_x=2$ and $L_y=1$ and constant density and viscosity ($\rho=1$ and $\eta=1$), gravity acceleration $\bm{g}=\bm{0}$ and penalty
-parameter $\lambda=10^8$. The grid is composed by \num{40x20} elements. Velocity boundary conditions are set to no slip $(\bm{v}=\bm{0})$ at the top and the
-bottom, and a parabolic profile is imposed on the sides, with $u=y(1-y)$ and $v=0$. The analytical solution is then given by:
-\begin{eqnarray}
-u(x,y)&=&y(1-y)\nonumber \\
-v(x,y,)&=&0\nonumber \\
-p(x,y)&=&2\eta\left(\frac{L_x}{2}-x\right)\nonumber
-\end{eqnarray}
-
-The velocity field predicted by the model follows the expected parabolic profile (Fig. \ref{fig:poiseuille}a and Fig. \ref{fig:poi_plot}a) and, in case of the
-classic penalty method with no iterations, the pressure is clearly related to the divergence by means of the penalty parameter (Fig. \ref{fig:poiseuille}b and
-c and Fig. \ref{fig:poi_plot}b). Fig. \ref{fig:poiseuille}d shows that one Uzawa iteration is sufficient to bring the divergence down to \num{e-15}, with no
-correlation with the pressure field. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Poiseuille%20Flow}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Poiseuille.pdf}
-\caption{Velocity field (panel a), pressure (panel b) and divergence velocity of a Poiseuille flow in case of the classic penalty method (no iterations) and
-after one Uzawa iteration (panel c and d, respectively).}
-\label{fig:poiseuille}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Analytical.pdf}
-\caption{Velocity (panel a) and pressure (panel b) field predicted by the model for a Poiseuille flow with respect with their analytical solutions. The
-velocity field is plotted as function of the vertical coordinate in $L_x/2$ and the pressure is plotted as function of the x coordinate in $L_y/2$.}
-\label{fig:poi_plot}
-\end{figure}
-
-\subsection{Instantaneous 2D sphere}\label{sec:ist_sphere}
-The domain is a unit square with gravity $\bm{g}=(0,-1)$. The fluid has constant density and viscosity ($\rho_f=1$ and $\eta_f=1$). The sphere is in the middle
-of the domain with a radius R = 0.123456798 and has constant density and viscosity ($\rho_s=10^{-2}+\rho_f$ and $\eta_s=10^3 \cdot \eta_f$).
-We distinguish three different types of velocity boundary conditions:
-\begin{enumerate}
-\item FS: free slip conditions on all sides.
-\item NS: no slip conditions on all sides.
-\item OT: free slip conditions on the sides and the bottom, and open boundary at the top.
-\end{enumerate}
-All three cases are performed for different grid resolutions, between \num{16x16} and \num{512x512} elements with 50 randomly distributed markers per element,
-and different average schemes for the viscosity (harmonic, geometric and arithmetic). The velocity in the centre of the sphere $\bm{v}(0.5;0.5)$), minimum and
-maximum velocities ($u_{min}$, $u_{max}$, $v_{min}$, $v_{max}$) and pressures ($p_{min}$, $p_{max}$) on the entire domain, $v_{\textrm{rms}}$ and average pressure on the
-entire domain ($p_{avg}$) are compared with solutions generated by ASPECT \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and all results
-can be found at \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere2D} and 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Instantaneous_2D_Sphere}. Results in terms of the $v_{\textrm{rms}}$ for NS model 
-are also shown in Fig. \ref{fig:inst_sphere}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_NS.pdf}
-\caption{$v_{\textrm{rms}}$ as function of element size for the instantaneous 2D sphere experiment in case of no slip boundary conditions and with different average
-schemes for the viscosity. Results are compared with results obtained by other numerical codes.}
-\label{fig:inst_sphere}
-\end{figure}
-
-\subsection{Rayleigh-Taylor instability}\label{sec:rayleigh}
-This problem has been originally presented by \citet{vanKeken1997} and here it is performed as the isoviscous case (Case 1a). The domain has $L_x=0.9142$ and
-$L_y=1$ and gravity $\bm{g}=(0,-1)$. Two fluids with same constant viscosities ($\eta_1=\eta_2=1$) and different densities ($\rho_1=1000$ and $\rho_2=1010$),
-with the lighter fluid at the bottom. The initial interface between the fluids is given by $y(x)=0.2+0.02 \cos \left(\frac{\pi x}{L_x}\right)$. The experiment
-is performed with different grid sizes (\num{50x50}, \num{80x80}, \num{100x100} and \num{256x256}). A total of 1960000 markers are randomly distributed at the
-beginning of the simulation. Velocity boundary conditions are set to no slip at the top and the bottom, and to free slip at the sides of the domain.
-
-$v_{\textrm{rms}}$ as function of time is reported in Fig. \ref{fig:RT}, matching well with results shown by \citet{vanKeken1997}, \citet{Tackley2003} and
-\citet{Thieulot2014}. Fig. \ref{fig:rayleigh} shows the evolution of the experiment at different time steps. All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Rayleigh_Taylor_experiment/ISOVISCOUS}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/RT.pdf}
-\caption{$v_{\textrm{rms}}$ of the Rayleigh-Taylor experiment as function of time for different resolution of the grid.}
-\label{fig:RT}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=300px]{./Figures/Rayleigh.pdf}
-\caption{Evolution of the Rayleigh-Taylor experiment for a grid of \num{256x256} elements at $t=0, 500, 1000$ and $2000$ (panels a, b, c and d, respectively).}
-\label{fig:rayleigh}
-\end{figure}
-
-\subsection{Falling blocks}\label{sec:block}
-This benchmark is described as proposed by \citet{Gerya2003a}, \citet{Gerya2010b} and \citet{Thieulot2011}. The domain is square with $L_x=L_y=$ \SI{500}{\km}
-and the grid is composed by \num{50x50} elements with 25 markers in each element. The block is initially centred at ($x=$ \SI{250}{\km}; $y=$ \SI{400}{\km})
-and has a size of \num{100x100} km (Fig. \ref{fig:block}a). The fluid surrounding the block has $\eta_f=$ \SI{e21}{\pascal\per\s} and $\rho_f=$
-\SI{3200}{\kg\per\cubic\m}. The benchmark tests with different viscosities of the block, with $\eta_b$ from \SIrange{e15}{e27}{\pascal\per\s}. In each
-experiment also the density $\rho_b$ of the block varies from \SIrange{3220}{9900}{\kg\per\cubic\m}. Velocity boundary conditions are set to free slip
-conditions on all sides of the domain. Density distributions at $t=$ \SI{20}{\mega\year} for $\rho_b=$ \SI{3300}{\kg\per\cubic\m} and different $\eta_b$
-are plotted in Fig. \ref{fig:block}b-f, showing that the code correctly preserve the block geometry in case of large viscosity contrast, with a block stiffer
-than the surrounding fluid (Fig. \ref{fig:block}f).
-
-The velocity in the centre of the falling block at $t=0$ is measured for all experiments and, since the velocity should increase with density contrast, the
-quantity $\bm{v}/(\rho_b-\rho_f)$ is plotted as function of the viscosity contrast. All results perfectly match with those from \citet{Gerya2010b} and they
-line up on a single curve, demonstrating that the code can correctly deal with large viscosity and density contrasts (Fig. \ref{fig:falling}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Falling%20blocks}.
-
-\begin{figure}
-\noindent\includegraphics[width=300px]{./Figures/Block.pdf}
-\centering
-\caption{Density evolution of the falling block experiment at $t=$ \SI{0}{\mega\year} (panel a) and $t=$ \SI{20}{\mega\year} for different viscosities of the
-block (panels b-f).}
-\label{fig:block}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Falling.pdf}
-\caption{Initial velocity relative to the density contrast at the centre of the falling block as function of the viscosity contrast between the block and the
-surrounding fluid.}
-\label{fig:falling}
-\end{figure}
-
-\subsection{2D time-dependent Stokes sphere with deformable free surface}\label{sec:sphere}
-This experiment is performed in a unit square domain with the gravity acceleration fixed to $g_y=-1$. The sphere has $\rho_s=2$ and $\eta_s=10^3$ and is
-initially centred at ($0.5;0.6$) with radius $R=0.123456789$. The fluid surrounding the sphere has $\rho_f=1$ and $\eta_f=1$ and occupies the domain for
-$y\leq 0.75$, while for $y>0.75$ the air has $\rho_a=0$ and $\eta_a=10^{-3}$. Velocity boundary conditions are set to free slip on all sides. The Courant
-number is set to 0.25. The experiments run for 200 s using grid resolutions from \num{150x150} to \num{512x512} elements, with 25 randomly distributed markers
-per element, and for different average schemes for the viscosity. The interface between the fluid and the air is tracked by means of the markers chain. Results
-in terms of velocity and pressure fields and topography variations are compared with results obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and they can be found at
-\url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere_fs2D} and
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Time_dependent_sphere}. Results in terms of the $v_{\textrm{rms}}$ in case of 
-an arithmetic average are also shown in Fig. \ref{fig:inst_spherefs}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_arithm.pdf}
-\caption{$v_{\textrm{rms}}$ for different grid resolutions as function of time for the 2D time-dependent Stokes sphere experiment in case of an arithmetic average.
-Results are compared with results obtained by ASPECT.}
-\label{fig:inst_spherefs}
-\end{figure}
-
 \subsection{Free surface stabilisation}\label{sec:stab}
 The benchmark is performed as discussed in \citet{Kaus2010a} and \citet{Thieulot2014}. The domain is square with $L_x=L_y=$ \SI{500}{\km} and a grid resolution
-of \num{200x200} elements, each of them containing 16 markers. A buoyant fluid with $\rho_1=$ \SI{3200}{\kg\per\cubic\metre} and $\eta_1=$ \SI{e20}{\pascal\s}
+of $200\times200$ elements, each of them containing 16 markers. A buoyant fluid with $\rho_1=$ \SI{3200}{\kg\per\cubic\metre} and $\eta_1=$ \SI{e20}{\pascal\s}
 is overlain by a denser fluid with $\rho_2=$ \SI{3300}{\kg\per\cubic\metre} and $\eta_2=$ \SI{e21}{\pascal\s}. The initial interface between the fluids has an
@@
 \end{figure}
-
-\subsection{Topography relaxation}\label{sec:crameri}
-This experiment is performed as in \citet{Crameri2012} using free surface and sticky air (Case 1). The domain is rectangular with $L_x=$ \SI{2800}{\km} and
-$L_y=$ \SI{700}{\km} in case of free surface and $L_y=$ \SI{800}{\km} in case of sticky air ($h_{st}=$ \SI{100}{\km}). The grid is composed by \num{256x64}
-elements for the free surface \num{560x320} elements for the sticky air. For both cases, there is a mantle of 600 km thickness with $\eta_m=$ \SI{e21}{\pascal\s}
-overlain by a highly viscous cosine-shaped lithosphere with thickness between 93 and 107 km and $\eta_l=$ \SI{e23}{\pascal\s}. Both mediums have $\rho=$
-\SI{3300}{\kg\per\cubic\metre}. In case of sticky air experiment, the air has $\rho_a=$ \SI{0}{\kg\per\cubic\metre} and $\eta_a=$ \SI{e18}{\pascal\s}. The
-gravity acceleration is set to $g_y=$ \SI{-10}{\metre\per\square\s}. Velocity boundary conditions are set to no slip at the bottom and free slip at the sides
-of the domain. In case of sticky air, velocities on top are set to free slip conditions.
-
-The maximum topography as function of time ($h(t)$) can be analytically derived in according to \[h(t)=h_0 \exp(\gamma t)\]
-where $h_0=7$ km is the initial topography, $\gamma=\num{-0.2139e-11}$ is the characteristic relaxation rate and $t$ is the time (black line in Fig.
-\ref{fig:crameri}). Maximum topography at the characteristic relaxation time $t_{rlx}=$ \SI{14.825}{\kilo\year} can be found to be $h_{rlx}=$ \SI{2576}{\m}.
-The results show that a viscosity of the air of \SI{e18}{\pascal\s} is needed in case of $h_{st}=$ \SI{100}{\km} to correctly describe the topography relaxation
-for this problem, as pointed out by \citet{Crameri2012} (purple line in Fig. \ref{fig:crameri}). The case with a true free surface well follows the analytical
-solution (orange line in Fig. \ref{fig:crameri}). In particular, at the relaxation time in case of the true free surface predicts a maximum topography of
-\SI{2570}{\m}, with an error of \SI{6}{\m} with respect to the analytical solution. Results are also compared with results obtained with a free surface in
-UNDERWORLD (\num{256x64} elements) and SULEC (\num{401x201} elements) (with courtesy from \citealp{Crameri2012}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Topography_relaxation}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Crameri.pdf}
-\caption{Maximum topography as function of time for the Crameri benchmark (coloured lines) in comparison with the analytical solution (black line) and results
-obtained with UNDERWORLD (yellow line) and SULEC (blue line).}
-\label{fig:crameri}
-\end{figure}
-
-\subsection{Spontaneous subduction}\label{sec:subduction}
-This experiment is performed as presented by \citet{Schmeling2008}, in case of both a sticky air and a true free surface and using different average schemes
-for the viscosities. The domain is rectangular with $L_x=$ \SI{3000}{\km} and $L_y=$ \SI{700}{\km}, a grid resolution of \num{256x64} elements with 50 markers
-per element and Courant number of 0.01. At the beginning of the simulation, a 100 km-thick lithospheric layer with $\rho_m=$ \SI{3300}{\kg\per\cubic\metre}
-and $\eta_m=$ \SI{e23}{\pascal\s} is located at the top of the domain for $x$ between 1000 and 3000 km, while the asthenospheric mantle has $\rho_m=$
-\SI{3200}{\pascal\s} and $\eta_m=$ \SI{e21}{\pascal\s}. In addition, a 200 km-depth lithospheric slab is already subducted in the mantle in order to have
-a spontaneous subduction. In case of sticky air, $L_y=750$ km and a 50 km-thick air layer with $\eta_a$ from \SIrange{e19}{e21}{\pascal\s} overlie the mantle.
-Velocity boundary conditions are set to free slip on the sides and on the bottom of the domain. In case of sticky air, velocities on the top boundary are set
-to free slip conditions as well.
-
-The maximum depth of the slab is tracked for all the simulations (Fig. \ref{fig:subduction}) and they are compared with results obtained by
-\citet{Schmeling2008} with models with regular grids and comparable grid resolutions (rectangles in Fig. \ref{fig:subduction}). Low viscous sticky air models
-(i.e., \SI{e19}{\pascal\s}) enlighten a strong dependence of the sinking velocity with the chosen average scheme (dashed lines in Fig. \ref{fig:subduction}),
-while high viscous sticky air models (i.e., \SI{e21}{\pascal\s}) show a higher resistance at the trench, underestimating the correct solution (dotted lines in
-Fig. \ref{fig:subduction}). A high resistance at the trench can be also observed in case of a true free surface for the low resolution grid used (continuous
-lines in Fig. \ref{fig:subduction}), as pointed out by \citet{Schmeling2008}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Spontaneous_subduction}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Subduction.pdf}
-\caption{Maximum depth of the slab as function of time for all the simulations of the \citet{Schmeling2008} benchmark, in case of harmonic, geometric and
-arithmetic means (red, green and blue lines, respectively) using a sticky air or a true free surface (discontinuous and continuous lines, respectively).
-Red, green and blue rectangular areas indicate the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{500}{\km} in case of sticky air,
-while the grey rectangular area indicates the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{600}{\km} in case of true free surface.}
-\label{fig:subduction}
-\end{figure}
-
-\subsection{The slab detachment}\label{sec:slab}
-The slab detachment benchmark is performed as described by \citet{Schmalholz2011} and \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{1000}{\km}
-and $L_y=$ \SI{660}{\km} and a grid resolution of \num{256x256} elements. A non-linear viscous T-shaped layer with $\rho_l=$ \SI{3300}{\kg\per\cubic\metre} is
-placed at the top of the domain and surrounded by a linear viscous fluid with $\rho_f=$ \SI{3150}{\kg\per\cubic\m}. The top layer is 80 km-thick and a 250
-km-long and 80 km-wide slab is placed at $x=L_x/2$ (Fig. \ref{fig:slab}a). The effective viscosity of the top layer is given by
-\[\eta_{eff}=\eta_0 I_2^{\frac{(1-n)}{n}}\]
-with $\eta_0=$ \SI{4.75e11}{\pascal\s} and $n=4$, while the surrounding fluid has $\eta_f=$ \SI{e21}{\pascal\s}. The effective viscosity is capped between
-$\eta_{min}=$ \SI{e21}{\pascal\s} and $\eta_{max}=$ \SI{e25}{\pascal\s}. Velocity boundary conditions are set to free slip at the top and the bottom, and to
-no slip at the sides of the domain.
-
-The sinking of the slab determines a decrease of the effective viscosity in the T-shaped layer during the first part of the evolution (Fig. \ref{fig:slab}a),
-as consequence of local strain rates, while effective viscosities increase up to $\eta_{max}$ in the top layer after approximately \SI{20}{\mega\year}, when
-necking is complete (Fig. \ref{fig:slab}b).
-
-The width of necking is tracked by means of markers position at the side of the slab. Necking width and time are normalised in function of the initial slab
-(\SI{80}{\km}) and the characteristic time ($t_c=$ \SI{7.1158e14}{\s}) \citep{Schmalholz2011,Glerum2018}. The results in case of arithmetic, geometric and
-harmonic mean of the viscosity are shown in Fig. \ref{fig:necking}, compared with results from \citet{Glerum2018} for which an infinity norm mean was used.
-The influence of the average schemes agrees with results shown in Section \ref{sec:subduction}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Slab_detachment}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Slab.pdf}
-\caption{Effective viscosity for the slab detachment benchmark at the beginning of the evolution (panel a) and when necking is complete (panel b).}
-\label{fig:slab}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Necking.pdf}
-\caption{Normalised width of the necking of the slab detachment benchmark as function of normalised time.}
-\label{fig:necking}
-\end{figure}
-
-\subsection{The indenter experiment}\label{sec:indenter}
-The indenter benchmark simulate a rigid punch on a purely plastic von Mises material, for which exists an analytical solution
-\citep{Thieulot2008,Thieulot2014,Glerum2018}. The domain is a unit square with a grid resolution of \num{256x256} elements. The gravity acceleration is set
-to $\bm{g}=\bm{0}$ and effective viscosity is capped using $\eta_{min}=10^{-4}$ and $\eta_{max}=10^3$. The tolerance for the convergence of the non-linear
-solution is $tol=10^{-9}$, with a maximum number of non-linear iterations set to 500. The medium in the domain has $\rho=0.01$, an initial viscosity
-$\eta_0=10$, cohesion $C=1$ and an angle of internal friction $\phi=0$°. Velocity boundary conditions are set to no slip at the bottom and free slip at the
-sides of the domain. The top of the domain is open with the exception of the central portion (punch area) with width $w_p=0.125$, where $v=-1.05$ and $u$ is
-fixed either to 0 (rough punch experiment) or free (smooth punch experiment).
-
-Fig. \ref{fig:indenter} show the results in terms of viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure
-(panels g and h) for both experiments. The analytical solution indicate that pressure at ($0.5;1$) and ($0.5 \pm w_p;1$)is $p=\pi +1$ and $p=1$, respectively
-(black and red lines in Fig. \ref{fig:smooth_p}). Fig. \ref{fig:smooth_p} shows a clear improvement in the pressure solution when passing from a rough to a
-smooth punch (green and blue lines in Fig. \ref{fig:smooth_p}, respectively), as previously observed by \citet{Thieulot2014} and \citet{Glerum2018}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Indenter}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Indenter.pdf}
-\caption{Viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure (panels g and h) fields for rough (left column) and
-smooth (right column) punch experiments.}
-\label{fig:indenter}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Smooth_Rough.pdf}
-\caption{Nodal pressure in function of the x coordinate at $y=1$ for rough (green line) and smooth (blue line) punch experiment. Black and red lines indicate
-the analytical solution at $x=0.5$ and $x=0.5 \pm w_p$, respectively.}
-\label{fig:smooth_p}
-\end{figure}
-
-\subsection{The brick experiment}\label{sec:brick}
-An instantaneous version of the brick benchmark is performed to verify the correctness of the pressure-dependent plasticity for different angle of internal
-friction, as proposed by \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{40}{\km} and $L_y=$ \SI{10}{\km} and a grid resolution of \num{512x128}
-elements. The gravity acceleration is set to $g_y=$ \SI{-10}{\m\square\s} and effective viscosity is capped using $\eta_{min}=$ \SI{e19}{\pascal\s} and
-$\eta_{max}=$ \SI{e26}{\pascal\s}. The tolerance for the convergence of the non-linear solution is $tol=10^{-7}$, with a maximum number of non-linear iterations
-set to 1000. A 800 m-wide and 400 m-height inclusion with $\eta_b=$ \SI{e20}{\pascal\s} and $\rho_b=$ \SI{2700}{\kg\per\cubic\m} is placed at the bottom of the
-domain at $x=L_x/2$. The inclusion is surrounded by a non-linear viscous medium with $\rho_m=$ \SI{2700}{\kg\per\cubic\m}, an initial viscosity of
-\SI{e23}{\pascal\s}, a linear viscous viscosity of \SI{e25}{\pascal\s} and a cohesion of \SI{40}{\mega\pascal}. Velocities are set to free slip conditions at
-the bottom of the domain and the top is open. Velocities on sides of the domain are fixed to $u=$ \SI{+-2e-11}{\m\per\s} and $v=0$. The experiment is performed
-in compressional and extensional contexts with an angle of internal friction $\phi$ of the non-linear medium variable between 0° and 30°.
-
-As expected, two shear bands stem from the inclusion with variable angles in relation to both the dynamics context and the internal friction angle. Shear band
-angles formed at 45° for $\phi=0$° in both extensional and compressional contexts (Fig. \ref{fig:brick_beam}a), while they are different in case of $\phi\neq0$°.
-Shear band angles in case of $\phi=20$° are shown for compression and extension in Fig. \ref{fig:brick_beam}b  and c, respectively. Values of shear band angles
-as function of internal friction angles from 0° to 30° are extracted at different depths and minimum and maximum values are plotted in Fig. \ref{fig:brick},
-in comparison with theoretical Roscoe, Arthur and Coulomb shear band angles. For all tests the tolerance for nonlinear convergence is defined $tol=10^{-7}$
-and none converges before the maximum number of iterations ($it_{max}=1000$) is reached. However, tests with internal friction angles up to 15° show a constant
-decrease of the velocity residuals, which can not be observed for for tests with higher internal friction angles (Fig. \ref{fig:convergence}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Brick_experiment}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Brick_Beam.pdf}
-\caption{Shear band angles predicted for the brick experiment for $\phi=0$° (panel a) and for $\phi=20$° in case of compression and extension (panels b and c,
-respectively).}
-\label{fig:brick_beam}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Brick.pdf}
-\caption{Shear band angles predicted for the brick experiment in case of compressional and extensional contexts as function of different internal angle of
-friction (continuous black line and red dots), compared with theoretical Roscoe, Artur and Coulomb angles (discontinuous black lines). Red lines indicate the
-range of angles calculated at different depths.}
-\label{fig:brick}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \noindent\includegraphics[width=500px]{./Figures/convergence.pdf}
-  \caption{Convergence rates for all tests. Normalised residuals (left column) are compared with velocity residuals (right column) for extensional (first row)
-  and compressional (second row) settings.}
-  \label{fig:convergence}
-  \end{figure}
-
-\subsection{Advection stabilisation}\label{sec:advection}
-The 1D advection problem proposed by \citet{Donea2003} and \citet{Thieulot2011} is performed to verify the effectiveness of the SUPG method to stabilise the
-advection term of the energy equation. The domain is a 1D segment with $L_x=1$ composed by 50 elements and a discontinuity in the thermal field placed at
-$x=0.25$. Temperature is set to 1 for $x \leq 0.25$ and to 0 for $x > 0.25$. Velocity is set to ${u}=1$ in the entire domain. The simulation is performed for
-250 time steps, with $dt=0.002$, so the thermal discontinuity should be at $x=0.75$ at the end of the simulation.
-
-Temperature profiles at the end of the simulation are shown in \ref{fig:advection} as function of the dimensionless coefficient $\gamma=\tau \bm{v}/h$. In case
-of the classic Galerkin method ($\gamma=0$, blue line in \ref{fig:advection}) the final thermal profile is characterised by strong oscillations, which are
-eliminated in case of the SUPG method ($\gamma=0.045$, orange line in \ref{fig:advection}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Advection_stabilisation}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Advection.pdf}
-\caption{Temperature profile as function of \textit{x} for the advection stabilisation benchmark. Purple line indicates the initial temperature profile; the
-green line indicates the analytical temperature profile after 250 time steps; blue line indicates the temperature profile after 250 time steps in case of the
-classic Galerkin method ($\gamma=0$); orange line indicates the temperature profile after 250 time steps in case of the SUPG method ($\gamma=0.045$).}
-\label{fig:advection}
-\end{figure}
-
-\subsection{Shear heating}\label{sec:simple_shear}
-This exercise is performed in an unit square domain composed by \num{128x128} elements. The velocity field is prescribed on the entire domain with
-$\bm{v}=(L_y-y)y\bm{e_x}$; viscosity, density and specific heat are fixed to 1, while thermal conductivity and radiogenic energy are fixed to 0. Therefore,
-the energy equation (Eq. \ref{eq:energy}) can be simplified as \[\frac{\partial T}{\partial t}=H_s\] and fixing $T(t=0)=0$, the temperature can be find as
-\[T(t)=H_s t\]
-In this case we have
-\begin{eqnarray}
-\dot{\epsilon_{xx}}&=&\frac{\partial u}{\partial x}=0 \nonumber \\
-\dot{\epsilon_{yy}}&=&\frac{\partial v}{\partial y}=0 \nonumber \\
-\dot{\epsilon_{xy}}&=&\frac{1}{2}\left(\frac{\partial v}{\partial x}+\frac{\partial u}{\partial y}\right)=\frac{1}{2}(L_y-2y)\nonumber
-\end{eqnarray}
-and, simplifying Eq. \ref{eq:shear}, shear heating can be calculated as \[H_s(x,y)=(1-2y)^2\]
-
-The solution predicted by the model in terms of velocity, temperature and shear heating match well with the analytical solutions (Fig. \ref{fig:analytical_en}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Simple_shear}.
-
-%\begin{figure}
-%\centering
-%\noindent\includegraphics[width=400px]{./Figures/Energy_Thieulot.pdf}
-%\caption{Temperature and shear heating predicted by the model after 1 time step.}
-%\label{fig:en_thieulot}
-%\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=500px]{./Figures/Analytical_Shear.pdf}
-\caption{Velocity (panel a), temperature (panel b) and shear heating (panel c) as function of the \textit{y} coordinate for the simple shear experiment. The
-solutions predicted by the model (red lines) are compared with the analytical solutions (dashed blue lines).}
-\label{fig:analytical_en}
-\end{figure}
-
-\subsection{Shear and adiabatic heating}\label{sec:shear}
-This problem is performed as presented in Exercise 9.4 in \citet{Gerya2010b}. Two materials are vertically separated in a rectangular domain with $L_x=$
-\SI{1000}{\km}, $L_y=$ \SI{1500}{\km} and a grid resolution of \num{30x20} elements. Constant thermal coefficient expansion ($\alpha=$ \SI{3e-5}{\per\kelvin}),
-temperature ($T=$ \SI{1300}{\kelvin}) and gravity acceleration ($g_y=$ \SI{-10}{\m\square\s}) are assumed in the whole domain. Fluid 1 (on the left side of the
-domain) has $\rho_1=$ \SI{3200}{\kg\per\cubic\m} and $\eta_1=$ \SI{e20}{\pascal\s}; fluid 2 (on the right side of the domain) has $\rho_2=$
-\SI{3300}{\kg\per\cubic\m} and $\eta_2=$ \SI{e22}{\pascal\s}. Velocity boundary conditions are set to free slip on all sides of the domain.
-
-As shown in Fig. \ref{fig:energy}, both shear and adiabatic heating predicted by the code (first row in Fig. \ref{fig:energy}) well recreate the results
-obtained by means of example ${Shear\_adiabatic\_heating.m}$ from \citet{Gerya2010b} (second row in Fig. \ref{fig:energy}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Adiabatic%2BShear}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Energy.pdf}
-\caption{Comparison between shear (first column), adiabatic (second column) and total (third column) energy predicted by the code (panels a, b and c) and those
-created using example ${Shear\_adiabatic\_heating.m}$ from Exercise 9.4 in \citet{Gerya2010b} (panels d, e and f).}
-\label{fig:energy}
-\end{figure}
-
-\subsection{Mantle convection}\label{sec:mantle}
-This problem is performed as presented by \citet{Blankenbach1989} (constant viscosity cases) and \citet{Thieulot2014}, in a 2D unit square domain with gravity
-acceleration $g_y=10^{10}Ra$. The experiment is performed with three different Rayleigh numbers ($Ra=10^4$, $10^5$ and $10^6$) and with different grid
-resolution (between \num{32x32} and \num{128x128} elements). The fluid has constant viscosity, initial density, heat capacity, thermal conductivity
-($\eta= \rho_0= C_p = k=1$), reference temperature ($T_0=0$) and thermal expansion coefficient ($\alpha=10^{-10}$). Temperatures are set to 0 on top and 1 on
-bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\]
-
-The solution generated by the code in terms of $v_{\textrm{rms}}$ (calculated as in Sec. \ref{sec:error}) and Nu as function of time are reported for all the
-simulations in Fig. \ref{fig:mantle}. At the steady state, $v_{\textrm{rms}}$ and Nu of all simulations converge well toward the values from \citet{Blankenbach1989},
-with lower errors for higher resolution grids (Table \ref{tab:mantle}). All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Mantle_convection}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Convection.pdf}
-\caption{$v_{\textrm{rms}}$ (panels a, c and e) and Nu (panels b, d and f) for the mantle convection benchmark as function of time for different grid resolution.
-Panels a and b show the results for $Ra=10^4$; Panels c and d show the results for $Ra=10^5$; panels e and f show the results for $Ra=10^6$. Purple lines
-indicate the convergence values for the $v_{\textrm{rms}}$ and the Nu from \citet{Blankenbach1989} at the steady state.}
-\label{fig:mantle}
-\end{figure}
-
-\begin{table}[H]
-\caption{Comparison between $v_{\textrm{rms}}$ and Nu predicted by the code for the mantle convection experiment and same values as reported in literature.}
-\centering
-\small
-\begin{tabular}{|>{\centering}m{0.09\textwidth}|>{\centering}m{0.05\textwidth}|>{\centering}m{0.24\textwidth}|>{\centering}m{0.15\textwidth}|
-  >{\centering}m{0.1\textwidth} >{\centering}m{0.1\textwidth} >{\centering\arraybackslash}m{0.1\textwidth}|}
-\toprule
-& & \multirow{2}{*}{\citet{Blankenbach1989}} & \citet{Thieulot2014} & \multicolumn{3}{c|}{FALCON}  \\
- & & & \num{200x200} & \num{32x32} & \num{64x64} & \num{128x128}  \\
-\midrule
-\multirow{2}{*}{$Ra=10^4$} & $v_{\textrm{rms}}$ & $42.864947 \pm 0.000020$ & 42.867 & 42.83226 & 42.852793 & 42.861394  \\
-           & Nu      & $4.884409 \pm 0.000010$  & 4.882  & 4.781297 & 4.857475  & 4.877573   \\
-\hline
-\multirow{2}{*}{$Ra=10^5$} & $v_{\textrm{rms}}$ & $193.21454 \pm 0.00010$  & 193.255& 193.872643 & 193.377472 & 193.252290  \\
-           & Nu      & $10.534095 \pm 0.000010$ & 10.507 & 9.602514 & 10.270735  & 10.465629   \\
-\hline
-\multirow{2}{*}{$Ra=10^6$} & $v_{\textrm{rms}}$ & $833.98977 \pm 0.00020$  & 834.712& 848.091176 & 837.767911 & 834.945793  \\
-           & Nu      & $21.972465 \pm 0.000020$ & 21.695 & 15.999266 & 19.703682  & 21.306939   \\
-\bottomrule
-\end{tabular}
-\label{tab:mantle}
-\end{table}
-
-\subsection{Viscoplastic mantle convection}\label{sec:visco_mantle}
-This benchmark is performed as Cases 1-5a described in \citet{Tosi2015} in a 2D unit square domain with two different grid resolutions of \num{32x32} 
-and \num{100x100} elements. Reference viscosity, reference density, heat capacity, thermal conductivity and thermal expansion coefficient are fixed to a constant
-value of 1 ($\eta_0= \rho_0= C_p = k=\alpha=1$), while the gravity has been chosen as $g_y=10^2$, in order to obtain a Rayleigh number $Ra=10^2$. Temperatures
-are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\].
-The viscosity field $\eta$ is calculated as the harmonic average between a linear part $\eta_{lin}$ that depends on temperature only or on temperature and 
-depth $y$ and a nonlinear-plastic part $\eta_{pl}$ that depends on the strain rate $\dot{\bm{\epsilon}}$, as follows
-\[\eta(T,y,\dot{\bm{\epsilon}})=2\left(\frac{1}{\eta_{lin}(T,y)}+\frac{1}{\eta_{pl}(\dot{\bm{\epsilon}})}\right)^{-1}\]
-The linear and the nonlinear-plastic parts of the viscosity are calculated following \citet{Tosi2015} as follows
-\[\eta_{lin}(T,y)=\exp(-\ln(\Delta\eta_T)T+\ln(\Delta\eta_y)y)\]
-\[\eta_{pl}(\dot{\bm{\epsilon}})=\eta^*+\frac{\sigma_Y}{\sqrt{\dot{\bm{\epsilon}}:\dot{\bm{\epsilon}}}}\]
-where $\Delta\eta_T$, $\Delta\eta_y$, $\eta^*$ and $\sigma_Y$ are parameters chosen as listed in Table \ref{tab:tosi_setup}.
-
-For each case, some diagnostic quantities measured at the steady state are reported in Table \ref{tab:visco_mantle} (see Section \ref{sec:error} for 
-the computation of all the quantities), showing a well fit with results described in \citet{Tosi2015}. Data and charts for all cases can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Viscoplastic_mantle_convection}.
-
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{white!100!}{gray!25!}
-%\begin{table}[H]
-%  \caption{Results of the viscoplastic mantle convection benchmark for Cases 1-4 compared with results obtained with ELEFANT and YACC (from \citealp{Tosi2015}).}
-%  \centering
-%  \resizebox{\textwidth}{!}{
-%\begin{tabular}{|c|ccc|ccc|ccc|ccc|}
-%  \toprule
-%    & \multicolumn{3}{c|}{\textbf{Case 1}} & \multicolumn{3}{c|}{\textbf{Case 2}} & \multicolumn{3}{c|}{\textbf{Case 3}} &
-%    \multicolumn{3}{c|}{\textbf{Case 4}} \\
-%   & FALCON & ELEFANT& YACC & FALCON& ELEFANT& YACC & FALCON & ELEFANT & YACC & FALCON & ELEFANT & YACC \\
-%  \midrule
-%   $\left\langle T \right\rangle$ & 0.7759 & 0.7758 & 0.7767 & 0.6043 & 0.6036 & 0.6058 & 0.7275 & 0.7275 & 0.7286 & & 0.5277 & 0.5289\\
-%    Nu$^{top}$ & 3.4280 & 3.4214 & 3.4298 & 8.5242 & 8.5115 & 8.5278 & 3.0368 & 3.0347 & 3.0374 && 6.5912 & 6.5572 \\
-%    Nu$^{bottom}$ & 3.3168 & 3.313 & 3.3143 & 8.4599 & 8.4465 & 8.3990 & 3.0086 & 2.9908 & 2.9628 && 6.5834 & 6.5243 \\
-%    $v_{\textrm{rms}}$ & 249.901 & 249.134 & 251.7997 & 141.2451 & 140.7087 & 142.2020 & 100.2776 & 100.1208 & 100.9467 && 79.1105 & 79.6202 \\
-%    $v_{\textrm{rms}}^{top}$ & 1.8907 & 1.8642 & 1.8298 & 104.7928 & 104.1228 & 105.1767 & 2.0832 & 2.0652 & 2.0374 && 74.7596 & 75.8414 \\
-%    $u_{max}^{top}$ & 2.6340 & 2.6119 & 2.5516 & 121.9212& 121.766 & 122.3238 & 2.9086 & 2.9019 & 2.8458 && 88.9146 & 89.2940 \\
-%    $\eta_{min}$ & - & - & - & $2.0303\times10^{-5}$ & $2.0600\times10^{-5}$ & $1.9853\times10^{-5}$ & $4.8129\times10^{-5}$ & $4.8080\times10^{-5}$ & $4.7907\times10^{-5}$ &
-%    & $1.9860\times10^{-4}$ & $1.9174\times10^{-4}$ \\
-%    $\eta_{max}$ & - & - & - & 1.3325 & 1.85 & 1.6095 & 0.9360 & 0.9023 & 1 && 1.52 & 1.6773 \\
-%    $\left\langle W \right\rangle$ & 2.4299 & 2.4316 & 2.4583 & 7.5648 & 7.5665 & 7.6222 & 2.0387 & 2.0384 & 2.0400 && 5.6216 & 5.6512 \\
-%    $\left\langle \Phi \right\rangle/Ra$ & 2.4311 & 2.4276 & 2.4333 & 7.5661 & 7.5585 & 7.6194 & 2.0386 & 2.0370 & 2.0335 && 5.6182 & 5.6463 \\
-%    $\delta$ & 0.05\% & 0.16\% & 1.02\% & 0.02\% & 0.11\% & 0.04\% & $<$0.01\% & 0.07\% & 0.32\% && 0.06\% & 0.09\% \\
-%  \bottomrule
-%  \end{tabular}}
-%  \label{tab:visco_mantle}
-%  \end{table}
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{2}{white!100!}{gray!25!}
-\begin{table}[H]
-  \caption{Benchmark cases and corresponding parameters.}
-  \centering
-  \small
-\begin{tabular}{|>{\centering}m{0.08\textwidth}|>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}
-  >{\centering}m{0.05\textwidth} >{\centering}m{0.05\textwidth} |>{\centering\arraybackslash}m{0.2\textwidth}|}
-  \toprule
-    Case & $Ra$ & $\Delta\eta_T$ & $\Delta\eta_y$ & $\eta^*$ & $\sigma_Y$ & Convective regime \\
-  \midrule
-  \rowcolor{gray!25!}
-    1 & $10^2$ & $10^5$ & 1 & - & - & Stagnant lid \\
-    2 & $10^2$ & $10^5$ & 1 & $10^{-3}$ & 1 & Mobile lid  \\
-    \rowcolor{gray!25!}
-    3 & $10^2$ & $10^5$ & 10 & - & - & Stagnant lid  \\
-    4 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 1 & Mobile lid \\
-    \rowcolor{gray!25!}
-    5 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 4 & Periodic \\
-  \bottomrule
-  \end{tabular}
-  \label{tab:tosi_setup}
-  \end{table}
-
-\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{gray!25!}{white!100!}
-\begin{table}
-  \caption{Results of the viscoplastic mantle convection benchmark for different grid resolutions in Cases 1-5 compared with results obtained with ELEFANT
-  and YACC ($100\times100$ elements, from \citealp{Tosi2015}) and in Stone 28 ($32\times32$ elements, from
-  \url{https://github.com/cedrict/fieldstone/tree/master/python_codes/fieldstone_28}).}
-  \centering
-  \resizebox{\textwidth}{!}{
-\begin{tabular}{|cc|ccccccccccc|}
-  \toprule
-  \multicolumn{13}{|c|}{\textbf{Case 1}} \\
-  \midrule
-  Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-  & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-  \hline
-  \rowcolor{gray!25!}
-    YACC & $100\times100$ & 0.7767 & 3.4298 & 3.3143 & 251.7997 & 1.8298 & 2.5516 & - & - & 2.4583 & 2.4333 & 1.02\% \\
-    ELEFANT & $100\times100$ & 0.7758 & 3.4214 & 3.313 & 249.134 & 1.8642 & 2.6119 & - & - & 2.4316 & 2.4276 & 0.16\% \\
-    \rowcolor{gray!25!}
-    Stone 28 & $32\times32$ & 0.7737 & 3.3987 & - & 243.872 & - & - & - & - & - & - & - \\
-%    FALCON & $32\times32$ & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & - & - & 2.4745 & 2.4864 & 0.48\% \\
-%    \rowcolor{gray!25!}
-    FALCON & $100\times100$ & 0.7759 & 3.4280 & 3.3168 & 249.901 & 1.8907 & 2.6340 & - & - & 2.4299 & 2.4311 & 0.05\% \\
-    \midrule
-    \multicolumn{13}{|c|}{\textbf{Case 2}} \\
-    \midrule
-    Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-    & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-    \hline
-    \rowcolor{gray!25!}
-     YACC & $100\times100$ & 0.6058 & 8.5278 & 8.3990 & 142.2020 & 105.1767 & 122.3238 & $1.9853\times10^{-5}$ & 1.6095 & 7.6222 & 7.6194 & 0.04\% \\
-     ELEFANT & $100\times100$ & 0.6034 & 8.5115 & 8.4465 & 140.8067 & 104.1228 & 121.766 & $2.0600\times10^{-5}$ & 1.85 & 7.5665 & 7.5585 & 0.11\% \\
-     \rowcolor{gray!25!}
-     Stone 28 & $32\times32$ & 0.6052 & 8.1715 & - & 141.518 & - & - & - & - & - & - & - \\
-%     FALCON & $32\times32$ & 0.6454 & 8.1422 & 7.4020 & 156.6908 & 95.7226 & 116.9302 & $2.2572\times10^{-5}$ & 0.7194 & 7.4289 & 7.4452 & 0.22\% \\
-%     \rowcolor{gray!25!}
-     FALCON & $100\times100$ & 0.6044 & 8.5266 & 8.4590 & 141.3684 & 104.8867 & 122.0470 & $2.0333\times10^{-5}$ & 1.2914 & 7.5707 & 7.5718 & 0.01\% \\
-     \midrule
-     \multicolumn{13}{|c|}{\textbf{Case 3}} \\
-     \midrule
-     Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-     & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-     \hline
-     \rowcolor{gray!25!} 
-     YACC & $100\times100$ & 0.7286 & 3.0374 & 2.9628 & 100.9467 & 2.0374 & 2.8458 & $4.7907\times10^{-5}$ & 1 & 2.0400 & 2.0335 & 0.32\% \\
-      ELEFANT & $100\times100$ & 0.7275 & 3.0347 & 2.9908 & 100.1208 & 2.0652 & 2.9019 & $4.8080\times10^{-5}$ & 0.9023 & 2.0384 & 2.037 & 0.07\% \\
-      \rowcolor{gray!25!}
-      Stone 28 & $32\times32$ & 0.7277 & 3.0347 & - & 100.018 & - & - & - & - & - & - & - \\
-%      FALCON & $32\times32$ & 0.7281 & 3.0618 & 2.8132 & 101.0092 & 2.1630 & 3.0178 & $4.8500\times10^{-5}$ & 0.7845 & 2.0699 & 2.0695 & 0.02\% \\
-%      \rowcolor{gray!25!}
-      FALCON & $100\times100$ & 0.7275 & 3.0368 & 3.0086 & 100.2776 & 2.0832 & 2.9086 & $4.8129\times10^{-5}$ & 0.9360 & 2.0387 & 2.0386 & $<$0.01\% \\
-      \midrule
-      \multicolumn{13}{|c|}{\textbf{Case 4}} \\
-      \midrule
-      Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-      & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-      \hline
-      \rowcolor{gray!25!} 
-      YACC & $100\times100$ & 0.5289 & 6.5572 & 6.5243 & 79.6202 & 75.4814 & 89.2940 & $1.9174\times10^{-4}$ & 1.6773 & 5.6512 & 5.6463 & 0.09\% \\
-       ELEFANT & $100\times100$ & 0.5277 & 6.5912 & 6.5834 & 79.1105 & 74.7596 & 88.9146 & $1.9860\times10^{-4}$ & 1.5200 & 5.6216 & 5.6182 & 0.06\% \\
-       \rowcolor{gray!25!}
-       Stone 28 & $32\times32$ & 0.5286 & 6.4036 & - & 79.4746 & - & - & - & - & - & - & - \\
-%       FALCON & $32\times32$ & 0.5334 & 6.4679 & 6.3435 & 80.6153 & 76.0511 & 90.2058 & $2.0393\times10^{-4}$ & 0.6453 & 5.6676 & 5.6698 & 0.04\% \\
-%       \rowcolor{gray!25!}
-       FALCON & $100\times100$ & 0.5279 & 6.5897 & 6.5830 & 79.1220 & 75.0715 & 88.8484 & $1.9693\times10^{-4}$ & 1.4133 & 5.6127 & 5.6113 & 0.03\% \\
-       \midrule
-       \multicolumn{13}{|c|}{\textbf{Case 5}} \\
-       \midrule
-       Code & Resolution & $P$ & $\left\langle T \right\rangle_{min}$ & $\left\langle T \right\rangle_{max}$ & Nu$^{top}_{min}$ & Nu$^{top}_{max}$ &
-       $v_{\textrm{rms}}^{min}$ & $v_{\textrm{rms}}^{max}$ & \multicolumn{2}{c}{$(\left\langle \Phi \right\rangle/Ra)_{min}$} & 
-       \multicolumn{2}{c|}{$(\left\langle \Phi \right\rangle/Ra)_{max}$} \\
-       \hline
-       \rowcolor{gray!25!}
-         YACC & $100\times100$ & 0.0792 & 0.6543 & 0.6722 & 2.6950 & 7.2792 & 41.9813 & 98.8230 & \multicolumn{2}{c}{1.4347} & 
-         \multicolumn{2}{c|}{9.2748} \\
-         ELEFANT & $100\times100$ & 0.076 & 0.6514 & 0.6693 & 2.6808 & 7.3506 & 41.2578 & 100.089 & \multicolumn{2}{c}{1.3324} & 
-         \multicolumn{2}{c|}{9.5289} \\
-         \rowcolor{gray!25!}
-         Stone 28 & $32\times32$ & 0.0764 & 0.6521 & 0.6697 & 2.6912 & 7.0763 & 41.61 & 98.183 & \multicolumn{2}{c}{-} & 
-         \multicolumn{2}{c|}{-} \\
-     %    FALCON & $32\times32$ & & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & \multicolumn{2}{c}{-} & 
-     %     \multicolumn{2}{c|}{-} \\
-     %    \rowcolor{gray!25!}
-         FALCON & $100\times100$ & 0.0775 & 0.6519 & 0.6697 & 2.6847 & 7.3629 & 41.5256 & 99.5466 & \multicolumn{2}{c}{1.3399} & 
-         \multicolumn{2}{c|}{9.3993} \\
-  \bottomrule
-  \end{tabular}}
-  \label{tab:visco_mantle}
-  \end{table}
-
-\subsection{Thin layer entrainment}\label{sec:thin}
-This experiment is performed as originally proposed by \citet{vanKeken1997} with two incompressible fluids in a rectangular domain with $L_x=2$ and $L_y=1$
-and gravity acceleration $g_y=10^{10}Ra$. The Rayleigh number and the compositional Rayleigh number are fixed to $Ra=300000$ and $Rc=450000$, respectively.
-Both fluids have constant viscosity, thermal conductivity, specific heat ($\eta=\rho=C_p=1$) and thermal expansion coefficient ($\alpha=10^{-10}$). Fluid 1
-has a density $\rho_1=1$, while fluid 2 is denser ($\rho_2=\rho_1+1.5 \cdot 10^{-10}$) and is placed at the bottom of the domain, for $y \leq 0.025$.
-Temperature are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides of the domain. The initial
-temperature field is given by
-\[T(x,y)=T_u(x,y)+T_l(x,y)+T_r(x,y)+T_s(x,y)-\frac{3}{2}\]
-with
-\begin{eqnarray}
-T_u(x,y)&=&\frac{1}{2}erf\left(\frac{1-y}{2}\sqrt{\frac{u_0}{x}}\right)\nonumber \\
-T_l(x,y)&=&1-\frac{1}{2}erf\left(\frac{y}{2}\sqrt{\frac{u_0}{L_x-x}}\right)\nonumber \\
-T_r(x,y)&=&\frac{1}{2}+\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{y+1}} \exp\left(-\frac{x^2u_0}{4y+4}\right)\nonumber \\
-T_s(x,y)&=&\frac{1}{2}-\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{2-y}} \exp\left(-\frac{(L_x-x)^2u_0}{8-4y}\right)\nonumber
-\end{eqnarray}
-and
-\begin{eqnarray}
-u_0&=&\frac{L_x^{7/3}}{(1+L_x^4)^{2/3}}\left(\frac{Ra}{2\sqrt{\pi}}\right)^{\frac{2}{3}}\nonumber \\
-Q&=&2\sqrt{\frac{L_x}{\pi u_0}}\nonumber
-\end{eqnarray}
-
-The experiment is performed with two grid resolution (\num{125x40} and \num{200x80} elements), with 100 markers per element and CFL of 0.25. $v_{\textrm{rms}}$
-(see Sec. \ref{sec:error}) calculated for both the simulations match well with results from \citet{vanKeken1997} and obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and ELEFANT \citep{Thieulot2014}, obtained with similar grid resolutions
-(Fig. \ref{fig:thin}). In addition, a variety of simulations with different resolutions and aspect ratios are performed to verify that the sensitivity of the
-initial velocity field. Results match well with results from \citet{vanKeken1997} (in the grey area) and \citet{Thieulot2014} (Fig. \ref{fig:thin_initial}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Thin_layer}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_thin.pdf}
-\caption{$v_{\textrm{rms}}$ for the thin layer experiment as function of time for different grid resolution. Results are compared with results obtained by
-\citet{vanKeken1997} (black lines) and with ASPECT and ELEFANT.}
-\label{fig:thin}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrmszero.pdf}
-\caption{$v_{\textrm{rms}}$ at $t=0$ for the thin layer experiment as function of the element size for different aspect ratios of the elements. Results are compared
-with results obtained by \citet{vanKeken1997} (black dots in the grey area) and with ASPECT and ELEFANT.}
-\label{fig:thin_initial}
-\end{figure}
-
-\subsection{Sinking hydrated cylinder}\label{sec:quinquis}
-This experiment is performed as explained in \citet{Quinquis2014} and simulates the sinking of a cold, hydrated cylinder into a warm, dry mantle. The domain
-is square with $L_x=L_y=$ \SI{300}{\km}, a grid resolution of \num{300x300} elements and 25 markers per element. The sinking cylinder is located at $x=$
-\SI{150}{\km} and $y=$ \SI{170}{\km} with a radius of \SI{20}{\km} and has $\rho_c=$ \SI{3250}{\kg\per\cubic\m}, $\eta_c=$ \SI{e23}{\pascal\s}, a thermal
-conductivity $k_c=$ \SI{4.5}{\watt\per\m\per\kelvin},a specific heat $C_{p_c}=$ \SI{1250}{\joule\per\kg\per\kelvin}; the surrounding mantle has $\rho_m=$
-\SI{3200}{\kg\per\cubic\m}, $\eta_m=$ \SI{e20}{\pascal\s}, a thermal conductivity $k_m=$ \SI{105}{\watt\per\m\per\kelvin} and specific heat $C_{p_m}=$
-\SI{1250}{\joule\per\kg\per\kelvin}; a 58 km-thick lithosphere overlies the mantle and has $\rho_l=$ \SI{3200}{\kg\per\cubic\m}, $\eta_l=$ \SI{e23}{\pascal\s},
-a thermal conductivity $k_l=$ \SI{4.5}{\watt\per\m\per\kelvin}, specific heat $C_{p_l}=$ \SI{750}{\joule\per\kg\per\kelvin} and a thermal diffusivity of
-\SI{e-6}{\square\m\per\s}. Velocity boundary conditions are set to free slip on all sides of the domain. Temperature increase linearly in the lithosphere,
-from 0° to 1300°C, and in the mantle, from 1300° to 1360.5°C. The initial temperature of the cylinder is fixed to 400°C. Throughout the evolution, temperature
-is fixed to 0° and 1360.5°C at the top and bottom of the domain, respectively. The initial bound water content is imposed at 2 wt.\% in the cylinder and at
-0 wt.\% in both the mantle and the lithosphere. The maximum amount of water in the mantle is fixed to 0.2 wt.\%, while maximum water content in the cylinder
-and in the lithosphere is function of pressure and temperature and is calculated using a serpentinized harzburgite with Perple\_X, as in \citet{Quinquis2014}.
-
-As shown in \citet{Quinquis2014}, the model is characterised by a progressive dehydration in the external portion of the cylinder, due to the increase of
-temperature, with a consequent vertical migration of free water that hydrates the mantle above the cylinder, up to the lithosphere. The distribution of
-bound and free water during the evolution of the experiment is shown in Fig. \ref{fig:quinquis} (left and right columns, respectively). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Hydration}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Hydration.pdf}
-\caption{Distribution of bound and free water (left and right columns, respectively) at $t=0.5$ and $t=$ \SI{1.25}{\mega\year} (first and second rows,
-respectively) for the sinking hydrated cylinder experiment. Only markers with bound or free water not equal to 0 are plotted.}
-\label{fig:quinquis}
-\end{figure}
-
-\subsection{Mantle melting curves}\label{sec:katz}
-This instantaneous experiment is performed in order to recreate isobaric melting curves, as function of temperature, bulk water and content of water in the
-melt, to compare with melting curves obtained by \citet{Katz2003}. The domain is square with $L_x=L_y=100$ km, a grid resolution of \num{100x100} elements and
-25 markers per element. Density is set in the entire domain to \SI{3300}{\kg\per\cubic\m} and temperature increases from the sides to the centre from 1000° to
-1500°C. The experiment is performed for bulk water contents from 0 to 0.3 wt.\%.
-
-Fig. \ref{fig:fraction} shows isobaric melting curves for \SI{1}{\giga\pascal} (panel a) and \SI{3}{\giga\pascal} (panel b) calculated for different bulk water
-contents. The curves well match with those obtained for the mantle by \citet{Katz2003}, showing the saturation at the solidus for bulk water of 0.2 and 0.3 wt.\%, indicating
-that the code correctly determines melt fraction as function of pressure, temperature and bulk water content. Same results can be observed in Fig.
-\ref{fig:water}a, where isobaric and isothermal melting curves are shown as function of bulk water content. In addition, Fig. \ref{fig:water}b shows that the
-code correctly determines also the percentage of water in the melt as function of pressure, temperature and melt fraction. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Melting}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Fraction.pdf}
-\caption{Isobaric melting curves for 1 GPa (panel a) and 3 GPa (panel b) as function of temperature with different bulk water contents.}
-\label{fig:fraction}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Water.pdf}
-\caption{Isobaric melting curves for 1.5 GPa as function of bulk water (panel a) and water in the melt (panel b) calculated for different temperatures.}
-\label{fig:water}
-\end{figure}
-
-\clearpage
-
-\bibliography{export}
-
-\end{document}

### 2022-09-07 tex/Benchmark/stokes.tex
update main

--- 2022-03-31 FALCON_description.tex
@@
-\documentclass[hidelinks,10pt,a4paper]{article}
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{textcomp}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage{booktabs}
-\usepackage{multirow}
-\usepackage[cm]{fullpage}
-\usepackage[round]{natbib}
-\bibliographystyle{abbrvnat}
-\usepackage{doi}
-\usepackage{bm}
-\usepackage[table]{xcolor}
-\hypersetup{
-    colorlinks,
-    linkcolor={blue},
-    citecolor={blue},
-    urlcolor={blue}
-}
-\usepackage{float}
-\restylefloat{table}
-\usepackage{siunitx}
-\DeclareSIUnit\year{yr}
-
-\begin{document}
-
-\section{Introduction}\label{sec:intro}
-This document includes a description of the code FALCON, with a presentation of the features implemented followed by results of extensive benchmarks listed in
-Table \ref{tab:list}, performed to verify that the code solves correctly a wide range of problems. Description and results of all benchmarks are shown
-in Section \ref{sec:benchmark}.
-
-\begin{table}[H]
-\caption{List of benchmarks performed for each feature of the code. Data of all benchmarks can be found at 
-\url{https://github.com/aleregorda/Benchmarks}}
-\centering
-\begin{tabular}{| l | l | l |}
-\toprule
- \textbf{Feature}  & \textbf{Benchmark} & \textbf{Section} \\
-\midrule
-  Solver  & Stokes flow & \ref{sec:stokes} \\
-\hline
-  \multirow{2}{*}{Markers advection} & Zalesak disk & \ref{sec:runge} \\
-    & Conservative Interpolation Velocity & \ref{sec:CVI} \\
-\hline
-  \multirow{4}{*}{Momentum equation}  & Poiseuille flow & \ref{sec:poiseuille}\\
-    & Instantaneous 2D sphere & \ref{sec:ist_sphere} \\
-    & Rayleigh-Taylor experiment & \ref{sec:rayleigh} \\
-    & Falling block & \ref{sec:block} \\
-\hline
-    & 2D Stokes sphere & \ref{sec:sphere} \\
-    Sticky air and  & Stabilisation algorithm & \ref{sec:stab} \\
-    free surface & Topography relaxation & \ref{sec:crameri} \\
-  & Spontaneous subduction & \ref{sec:subduction} \\
-\hline
-  Erosion and sedimentation  & - & - \\
-\hline
-   Non-linear  & Slab detachment & \ref{sec:slab} \\
-   visco-plasticity & Indenter & \ref{sec:indenter} \\
-    & Brick & \ref{sec:brick} \\
-\hline
-  \multirow{3}{*}{Energy equation}  & Advection stabilisation & \ref{sec:advection} \\
-    & Simple shear heating & \ref{sec:simple_shear} \\
-    & Shear and adiabatic heating & \ref{sec:shear} \\
-\hline
-\multirow{3}{*}{Energy + momentum} & Mantle convection & \ref{sec:mantle} \\
-& Viscoplastic mantle convection & \ref{sec:visco_mantle} \\
-& Thin layer entrainment & \ref{sec:thin} \\
-\hline
-  Phase changes  & \multirow{2}{*}{Hydrated sinking cylinder} & \multirow{2}{*}{\ref{sec:quinquis}} \\
-  and hydration  &  & \\
-\hline
-  Melting & Experimental melting curves & \ref{sec:katz} \\
-\bottomrule
-\end{tabular}
-\label{tab:list}
-\end{table}
-
-The code is written in Fortran90 and uses the finite element method (FEM) with quadrilateral $Q_1 \times P_0$ elements (continuous bilinear velocity and
-discontinuous constant pressure), associated with MUMPS\footnote{\url{http://mumps.enseeiht.fr/}} \citep{Amestoy2001,Amestoy2019} that is a software package
-for solving systems of linear equations of the form $A \cdot x = b$, where \textit{A} is a square sparse matrix, by means of a direct method. Since
-$Q_1 \times P_0$ elements do not satisfy Ladyzhenskaya, Babuska and Brezzi (LBB) stability condition \citep{Donea2003}, elemental pressure is elaborated
-in post-processing to avoid spurious pressures by means of a double interpolation to smooth the pressure. Following this procedure, the elemental pressure
-is interpolated onto nodes and then back onto elements. After the smoothing procedure, the pressure field is calculated again on the nodes to be used in
-combination with the lithostatic pressure, which is calculated onto nodes. Correctness of the solution and performances of the solver in terms of time and
-memory usage are tested solving the Stokes flow with the analytical solution proposed by \citet{Donea2003} (Section \ref{sec:stokes}).
-
-The thermo-mechanics of crust-mantle systems is described by means of the conservation of mass, momentum and energy equations, expressed as follows:
-\begin{eqnarray}
-\label{eq:mass}\frac{\partial \rho}{\partial t}+\bm{\nabla} \cdot (\rho \bm{u})&=&0\\
-\label{eq:momentum}-\bm{\nabla} p + \bm{\nabla} \cdot \bm{\tau} + \rho \bm{g}&=&0\\
-\label{eq:energy}\rho C_p \left(\frac{\partial T}{\partial t} + \bm{u} \cdot \bm{\nabla} T \right)&=&\bm{\nabla} \cdot \left(k\bm{\nabla} T\right) + H_{tot}
-\end{eqnarray}
-where $\rho$ is the density, $\bm{u}$ is the velocity, \textit{p} is the pressure, $\bm{\tau}$ is the deviatoric stress, $\bm{g}$ is the gravity acceleration,
-$C_p$ is the specific heat at constant pressure, \textit{T} is the temperature, \textit{k} is the thermal conductivity and $H_{tot}$ is the total internal heat
-production. Density variations due to temperature are generally small enough to assume the density as constant ($\rho=\rho_0$) in Eq. \ref{eq:mass} and in
-Eq. \ref{eq:energy}, while it must be treated as a variable in the buoyancy term of Eq. \ref{eq:momentum}, such that \[\rho=\rho_0(1-\alpha(T-T_0))\]
-where $\rho_0$ is the density at a reference temperature $T_0$ and $\alpha$ is the coefficient of thermal expansion. Eq. \ref{eq:mass} then can be rewritten as
-\begin{equation}\label{eq:mass_0}
-\bm{\nabla} \cdot \bm{u}=0
-\end{equation}
-This simplification is known as the extended Boussinesq approximation.
-
-The time step $dt$ is chosen in according to Courant-Friedrichs-Lewy condition \citep{Anderson1995}, such that
-\[dt=C_n\; \min\left(\frac{\Delta x}{\max|\bm{u}|},\frac{(\Delta x)^2}{\kappa}\right)\]
-where $0<C_n<1$ is the Courant (CFL) number, $\Delta x$ is the minimum dimension of smallest element, $\max|\bm{u}|$ is the maximum velocity calculated on the
-entire domain and
-$\kappa=\frac{k}{\rho C_p}$ is the thermal diffusivity \citep{Thieulot2014}.
-
-\section{Lagrangian markers}\label{sec:markers}
-Elemental properties (density, viscosity, thermal conductivity, specific heat and thermal expansion) needed to solve Eqs. \ref{eq:mass_0}, \ref{eq:momentum}
-and \ref{eq:energy} are related to the composition of each element, determined by means of Lagrangian markers that are characteristic of different materials
-in the domain. Elemental properties (with the sole exception of the viscosity) are calculated using an arithmetic mean, as \[P_e=\frac{1}{n}\sum_{i=1}^n P_i\]
-where $P_e$ is the elemental property, $P_i$ is the property characteristic of the material of each marker and $n$ is the number of markers of the element.
-Differently, the average scheme for the elemental viscosity can be chosen between harmonic, geometric and arithmetic mean. Markers can be placed regularly or
-randomly at the beginning of the simulation and their advection is performed by either a 2nd-order or a 4th-order Runge-Kutta in space, interpolating the velocity 
-field on each marker by means of the shape functions (see Section \ref{sec:runge}). The interpolated velocity is then corrected by means of the Conservative
-Velocity Interpolation (CVI), which introduces a corrective term to reduce dispersion and clustering of particles in both steady state and time-dependent 
-\citep{Wang2015} (see Section \ref{sec:CVI}).
-
-The initial distribution of the markers is created using the open-source code library
-Geodynamic World Builder \citep{Fraters2019}. During the simulation, each marker carries memory of temperature, pressure and accumulated strain, which are
-determined interpolating the relative nodal parameter. In particular, pressure is interpolated onto each marker after the smoothing procedure. The number
-of markers contained in each element is maintained between $n_{min}$ and $n_{max}$. When in an element there are less markers than $n_{min}$ the code adds
-random markers to reach the $n_{min}$, while if the number is higher than $n_{max}$ some of them are randomly deleted. When new markers are added, they assume
-the properties of the nearest marker. In this way elements are never empty and maintain a number of markers inside a prefixed range.
-
-\section{The momentum equation}\label{sec:mom_eq}
-The penalty method is implemented so that Eq. \ref{eq:mass_0} can be written as:
-\begin{equation}\label{eq:mass_penalty}
-\bm{\nabla} \cdot \bm{u} + \frac{p}{\lambda}=0
-\end{equation}
-where $\lambda$ is the so-called penalty parameter, which should be 6-7 orders of magnitude larger than the shear viscosity to ensure that mass conservation
-is satisfied \citep{Donea2003,Thieulot2014}.
-
-The deviatoric stress tensor in Eq. \ref{eq:momentum} can be written in terms of the strain rate tensor as $\bm{\tau}=2\eta\dot{\bm{\epsilon}}$, with
-$\dot{\bm{\epsilon}}=\frac{1}{2}\left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)$. Therefore, Eq. \ref{eq:momentum} can be rewritten as:
-\begin{equation}\label{eq:momentum_pressure}
--\bm{\nabla} p + \bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right) + \rho \bm{g} = 0
-\end{equation}
-Finally, using pressure from Eq. \ref{eq:mass_penalty}, Eq. \ref{eq:momentum_pressure} can be rewritten as:
-\begin{equation}\label{eq:momentum_penalty}
-\lambda \bm{\nabla} \left(\bm{\nabla} \cdot \bm{v}\right)+\bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right)+\rho\bm{g}=0
-\end{equation}
-The penalty method is been associated to the iterative Uzawa method, as described in detail in \citet{Dabrowski2008} and \citet{Thieulot2014}
-(see Section \ref{sec:poiseuille}). Results of benchmarks performed to verify the correctness in the implementation of Eq. \ref{eq:momentum_penalty} are
-shown in Sections \ref{sec:ist_sphere} and \ref{sec:rayleigh}.
-
-To support large viscosities variations the penalty parameter is related to the effective elemental viscosity by means of a dimensionless coefficient, so that
-$\lambda=\lambda_e(e)\eta_{eff}(e)$ \citep{Marotta2006,Dabrowski2008,Thieulot2014}. Benchmark of the falling block \citep{Gerya2003a,Gerya2010b,Thieulot2011}
-is performed to verify that the code can correctly deal buoyancy driven flows with strong viscosity contrasts (Section \ref{sec:block}).
-
-\section{Sticky air and free surface}\label{sec:sticky}
-The Earth's surface can be treated by means of either the so-called sticky-air or a true free surface method, both of them implemented in the code. In the
-sticky-air method the surface is approximated with the introduction of a buoyant layer with a viscosity at least four orders of magnitude lower than the crust
-\citep{Schmeling2008,Crameri2012} and the interface between lithosphere and air is defined using a chain of passive markers that are advected as the Lagrangian
-markers. The correctness of the evolution of the markers chain is tested with the experiment of a 2D time-dependent Stokes sphere below a free surface and
-compared with results from ASPECT\footnote{\url{https://aspect.geodynamics.org/}} \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} (Section
-\ref{sec:sphere}). In the true free surface case the top boundary is assumed stress-free and velocities are not fixed. In this case topography variations are
-described by vertical deformations of the mesh that depend on the velocity field of the nodes that identify the free surface, while horizontal deformations are
-not taken into account. This procedure is known as the Arbitrary Lagrangian-Eulerian (ALE) method and its implementation follows the technique described in
-\citet{Thieulot2011}. However, although the implementation of a true free surface better reproduces laboratory experiment, extremely small time steps can be
-necessary to maintain stability \citep{Kaus2010a,Quinquis2011,Thieulot2014}. Therefore, the stability algorithm proposed by \citet{Kaus2010a} is implemented
-to avoid instabilities due to high density differences at the free surface when using too large time steps. The implementation of this algorithm are tested by
-performing the experiment described by \citet{Kaus2010a} (Section \ref{sec:stab}).
-
-The topography relaxation benchmark proposed by \citet{Crameri2012} is performed to verify that the code correctly recovers topography variations in case of
-both the sticky-air and the true free surface method (Section \ref{sec:crameri}). Finally, Section \ref{sec:subduction} show the results of the spontaneous
-subduction experiment described by \citet{Schmeling2008}.
-
-\section{Erosion and sedimentation}\label{sec:erosion}
-Surface processes at either the lithosphere-air interface or in correspondence of the free surface have been implemented by means of the software
-Fastscape\footnote{\url{https://fastscape.org/}} \citep{Braun2013,Cordonnier2019,Yuan2019a,Yuan2019b}. FastScapeLib is a set of routines that solve the stream
-power law (enriched by a sediment transport/deposition term) \citep{Yuan2019a}, hillslope diffusion and marine transport and deposition \citep{Yuan2019b}.
-The partial differential equation solved by FastScapeLib is:
-\begin{equation}
-\frac{\partial h}{ \partial t}=U-K_f A^{m_e}S^{n_e} + \frac{G}{A}\int_A \left( U-\frac{\partial h}{\partial t} \right) dA + K_d \bm{\nabla}^2 h 
-\end{equation}
-where $h$ is the topography, $U$ is the uplift, $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $S$ is the slope, $m_e$ and $n_e$ are
-parameters that should be chosen to constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}, $G$ is the dimensionless
-deposition/transport coefficient for the enriched stream power law and $K_d$ is the effective diffusivity. Terms $K_f A^{m_e}S^{n_e}$ and $K_d \bm{\nabla}^2 h$
-consider long-range processes, described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, and short-range processes,
-proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}.
-
-In case of sticky air, slope is determined by means of the markers chain, which is vertically corrected considering erosion and sedimentation rates. After the
-correction, continental markers above the markers chain are transformed in air markers and air markers below the markers chain become sediments. Similarly,
-slope is determined considering the nodes on the top boundary in case of true free surface and they are vertically corrected as for the markers chain. In this
-case, continental markers above the top boundary are simply deleted.
-%Surface processes at the lithosphere-air interface are implemented taking into account fluvial erosion (long-range processes), hillslope processes (short-range
-%processes) and failure of slopes exceeding a threshold angle $\theta$ \citep{Burov1997,Tucker1999,Sacek2012,Croissant2014,Ueda2015,Beucher2020}. Erosion rates
-%$\dot{E}_{lr}$ due to long-range processes can be described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{lr}=K_f A^{m_e} \bm{\nabla} h^{n_e}
-%\end{equation}
-%where $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $\bm{\nabla} h$ is the slope and $m_e$ and $n_e$ are parameters chosen in to
-%constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}. Erosion rates $\dot{E}_{sr}$ due to short-range processes are
-%implemented by assuming that transport is linearly proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{sr}=K_d \bm{\nabla}^2 h
-%\end{equation}
-%where $K_d$ is the effective diffusivity. In addition, the slope is modified if it still exceeds a threshold angle after long- and short-range processes
-%\citep{Ueda2015}.
-
-%Offshore sedimentation processes are implemented similarly to hillslope processes but with a depth-dependent coefficient \citep{Sacek2012,Beucher2020}.
-%The sedimentation rate $\dot{S}$ is calculated as
-%\begin{equation}
-%\dot{S}=K_s \bm{\nabla}^2 h
-%\end{equation}
-%with
-%\begin{equation}
-%K_s=K^*_s \exp\left(\frac{h-h_{sl}}{h_0}\right)
-%\end{equation}
-%where $K^*_s$ is a constant, $h_{sl}$ is the sea level and $h_0$ is a characteristic length. In case of sticky air, slope is determined by means of the markers
-%chain, which is vertically corrected considering erosion and sedimentation rates. After the correction, continental markers above the markers chain are
-%transformed in air markers and air markers below the markers chain become sediments. Similarly, slope is determined considering the nodes on the top boundary
-%in case of true free surface and they are vertically corrected as for the markers chain. In this case, continental markers above the top boundary are simply
-%deleted.
-
-\section{Non-linear rheologies}\label{sec:nonlinear}
-Non-linear rheologies are implemented combining viscous creep (dislocation and diffusion) and plastic yielding. For each marker, diffusion and dislocation
-viscosity ($\eta_{df|ds}$) can be determined as follows \citep{Karato1993,Warren2008b,Wang2016}:
-\begin{equation}\label{eq:viscous}
-\eta_{df|ds}=f_s \left(\frac{d^m}{A}\right)^{\frac{1}{n}} \left (\frac{2^{\frac{1-n}{n}}}{3^{\frac{n+1}{2n}}} \right)
-I_2^{\frac{(1-n)}{n}}\exp\left(\frac{Q+pV}{nRT}\right)
-\end{equation}
-where $f_s$ is a scaling factor used to represent lithologies that are stronger or weaker than the base set, $d$ is the grain size, $m$ is the grain size exponent,
-$n$ is the stress exponent, $A$ is the uniaxial pre-exponential factor, $I_2$ is the square root of the second invariant of the strain rate tensor, $Q$ is the
-activation energy, $p$ is the pressure, $V$ is the activation volume, $R$ is the gas constant and $T$ is the temperature.
-Pressures and temperatures are determined by the interpolation of the nodal parameters, while strain rates are calculated by means of the
-derivative of the velocities on the nodes. In case of diffusion creep $n=1$ and $m>0$, in case of dislocation creep $n>1$ and $m=0$. The viscous creep $\eta_{cp}$
-is then calculated as the harmonic average between $\eta_{df}$ and $\eta_{ds}$:
-\begin{equation}\label{eq:visco_creep}
-\eta_{cp}=\left(\frac{1}{\eta_{df}}+\frac{1}{\eta_{ds}}\right)^{-1}
-\end{equation}
-The implementation of the non-linear viscous creep viscosity depending on the strain rate is tested by means of the slab detachment benchmark
-(Section \ref{sec:slab}).
-
-Plastic yielding is implemented rescaling $\eta_{cp}$ in order to limit the stress below the yield stress $\sigma_y$ \citep{Thieulot2008,Thieulot2014,Glerum2018},
-obtaining
-\begin{equation}\label{eq:plastic}
-\eta_{pl}=\frac{\sigma_y}{2I_2}
-\end{equation}
-where $\eta_{pl}$ is the plastic viscosity and the yield stress is determined following the Drucker-Prager criterion, such as
-\begin{equation}\label{eq:yield}
-\sigma_y=C\;\cos(\phi)+p\;\sin(\phi)
-\end{equation}
-where $C$ is the cohesion and $\phi$ is the internal friction angle. In case of negative pressure it is imposed equal to 0, so that negative yield stress are
-excluded. The correctness of the non-linear solution in case of plastic viscosity with variable internal friction angle is verified performing indenter and
-brick experiments (Sections \ref{sec:indenter} and \ref{sec:brick}, respectively).
-
-Strain softening is taken into account for both viscous creep and plastic viscosity \citep{Huismans2003,Babeyko2005,Huismans2005,Sobolev2005,Warren2008a} by
-means of the accumulated strain $\epsilon$ memorised by each marker. Softening in the viscous creep determines a linear decrease of $\eta_{vc}$ by means of a
-viscous strain softening factor $W_S$ that increases linearly from $W_{S_0}$ to $W_{S_{\infty}}$ for $\epsilon_{S_0}<\epsilon_S<\epsilon_{S_{\infty}}$ 
-\citep{Warren2008a}. This viscous softening can be related to strain-induced grain size reduction \citep{Warren2008a}. Differently, plastic softening is 
-simulated with a linear decrease of internal friction angle $\phi(\epsilon)$ and cohesion $C(\epsilon)$ in according to
-\begin{eqnarray}
-\label{eq:friction1}\phi(\epsilon)&=&\phi_0+(\phi_{\infty}-\phi_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}\\
-\label{eq:friction2}C(\epsilon)&=&C_0+(C_{\infty}-C_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}
-\end{eqnarray}
-where $\phi_0$, $C_0$ and $\phi_{\infty}$, $C_{\infty}$ are internal friction angle and cohesion for $\epsilon_0$ and $\epsilon_{\infty}$, respectively
-\citep{Huismans2003,Huismans2005,Warren2008a,Thieulot2014}. Plastic softening approximates deformation-induced softening of faults and brittle shear zones
-\citep{Warren2008a}.
-
-Plastic yielding and viscous creep are then combined to obtain a viscoplastic viscosity $\eta_{vp}$ as follows, assuming that they are independent processes
-\citep{Karato2008,Glerum2018}:
-\begin{equation}\label{eq:viscoplastic}
-\eta_{vp}=\min(\eta_{cp},\eta_{pl})
-\end{equation}
-Finally, effective viscosity $\eta_{eff}$ is capped by the minimum and the maximum viscosity ($\eta_{min}$ and $\eta_{max}$, respectively) to avoid extremely
-low or high viscosity \citep{Glerum2018} as follows
-\begin{equation}\label{eq:effective}
-\eta_{eff}=\min(\max(\eta_{vp},\eta_{min}),\eta_{max})
-\end{equation}
-Viscous creep and plastic yielding are non-linear rheologies because of their dependence on the velocity field through pressure and strain rates. Therefore,
-the solution is determined by means of Picard-type iterations, until convergence of the velocity field \citep{Glerum2018}. The convergence is verified at each
-iteration $i$ via the nonlinear residual $\bm{\mathcal{R}}^i$ that can be determined as
-\begin{equation}\label{eq:convergence}
-\bm{\mathcal{R}}^i= \mathbb{K} (\eta_{eff}(\dot{\bm{\epsilon}}^{i-1},p^{i-1}))\cdot \bm{v}^{i-1} - \bm{f}^i
-\end{equation}
-%\begin{equation}\label{eq:convergence}
-%\chi=1-\frac{\langle (\bm{v}^i-\langle\bm{v}^i\rangle)\cdot (\bm{v}^{i+1}-\langle\bm{v}^{i+1}\rangle) \rangle}{\sqrt{\sigma^i\sigma^{i+1}}}
-%\end{equation}
-%Iterations are performed until either convergence ($\chi<tol$) or maximal number of non-linear iterations ($it_{max}$) is reached.
-where $\mathbb{K}$ is the Stiffness matrix and $\bm{f}$ is the right hand side vector \citep{Spiegelman2016,Glerum2018}.  The $L_2$-norm is extracted from
-$\bm{\mathcal{R}}^i$ and it is normalised as follows
-\begin{equation}\label{eq:convergence_res}
-\frac{||\bm{\mathcal{R}}^i||_2}{||\bm{\mathcal{R}}^0||_2}
-\end{equation}
-where $||\bm{\mathcal{R}}^0||_2$ is the $L_2$-norm of the first nonlinear iteration.
-Since often the normalised nonlinear residual drop very quickly over the first 2-3 iterations, also the $L_2$-norm of velocity and pressure residuals are
-calculated as
-\begin{equation}\label{eq:convergence_vel}
-\frac{||\Delta\bm{v}||_2}{||\bm{v}||_2}=\frac{||\bm{v}_i-\bm{v}_{i-1}||_2}{||\bm{v}_{i}||_2}
-\end{equation}
-and
-\begin{equation}\label{eq:convergence_press}
-  \frac{||\Delta p||_2}{||p||_2}=\frac{||p_i-p_{i-1}||_2}{||p_{i}||_2}
-\end{equation}
-respectively. Iterations are performed until either maximal number of non-linear iterations ($it_{max}$) is reached or all the normalised nonlinear residual, 
-the velocity residual and the pressure residual converge under a defined tolerance $tol$.
-
-\section{The energy equation}\label{sec:en_eq}
-The stabilisation of the advection term of the energy equation (Eq. \ref{eq:energy}) needed to avoid possible oscillations in the thermal solution in the case
-when advection dominates over diffusion, is implemented by means of a streamline-upwind Petrov–Galerkin (SUPG) method, for which the advection term is modified
-follows the discussion in \citet{Thieulot2011} and \citet{Thieulot2014} (see Section \ref{sec:advection} for benchmark). The total internal heat production
-$H_{tot}$ in Eq. \ref{eq:energy} includes radiogenic heating $H_r$, shear heating $H_s$ and adiabatic heating $H_a$, where
-\begin{equation}\label{eq:shear}
-H_s=2\eta \dot{\bm{\epsilon}} : \dot{\bm{\epsilon}} = 2\eta(\dot{\epsilon}_{xx}^2+\dot{\epsilon}_{yy}^2+2\dot{\epsilon}_{xy}^2)
-\end{equation}
-and
-\begin{equation}\label{eq:adiabatic}
-H_a=T \alpha \frac{Dp}{Dt} \approx -\alpha T \rho g_y v
-\end{equation}
-The correctness of the implementation and the computation of shear and adiabatic heating is verified performing an exercise with an analytical solution
-(Section \ref{sec:simple_shear}) and Exercise 9.4 in \citet{Gerya2010b} (Section \ref{sec:shear}). Furthermore, mantle convection and thin layer entrainment
-benchmarks (Sections \ref{sec:mantle} and \ref{sec:thin}, respectively) are performed to verify that the code solve correctly Eqs. \ref{eq:momentum_penalty}
-and \ref{eq:energy} in case of a temperature field characterised by an initial perturbation and with a temperature-dependent density. At last, the viscoplastic
-mantle convection benchmark is performed to check also the viscoplasticity in case of a initial perturbed temperature field (Section \ref{sec:visco_mantle}).
-
-\section{Phase transitions and hydration}\label{sec:phase}
-Variations of effective density, specific heat and coefficient of thermal expansion during the evolution of crustal and mantle materials at different
-pressure-temperature ($p$-$T$) conditions are computed by means of version 6.8.6 of Perple\_X software package \citep{Connolly2005}, similarly to the
-implementation described in \citet{Marotta2020}. By default, mineral assemblages and properties of each lithology (oceanic crust, lower and upper continental
-crust, sediments and dry or hydrated mantle) are calculated for temperatures from \SIrange{330}{1600}{\kelvin}, with increments of approximately \SI{4}{\kelvin},
-and pressures from \SIrange{0.1}{30}{\giga\pascal}, with increments of approximately \SI{0.03}{\giga\pascal}, for a total of almost 100000 points.
-
-The introduction of phase transitions produces two main effects that must be taken into account in numerical models: variations in density and release/absorption
-of latent heat \citep{Tackley2010}. Effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into
-account considering effective coefficient of thermal expansion \citep{Christensen1985,Zhong2015}. Similarly, effects of latent heat in the energy equation
-(eq \ref{eq:energy}) are taken into account by considering effective specific heat and coefficient of thermal expansion
-\citep{Christensen1985,Tackley2010,Zhong2015}. Effective density, specific heat and coefficient of thermal expansion are included in Perple\_X files and are
-assumed by each Lagrangian marker in according to its $p$-$T$ conditions.
-
-Hydration processes related to dehydration of subducting lithosphere \citep{Schmidt1998,Liu2007,Faccenda2009,Faccenda2010,Faccenda2014,Rosas2016} strongly
-influence the thermo-mechanics inside the mantle wedge, mainly because of both weakening effects on the mantle rheology and density variations in case of
-mantle serpentinization \citep{Gerya2002,Honda2003,Arcay2005,Roda2010,Regorda2017}. In the case that hydration is switched on, the amount of bound and free
-water is memorised by each marker, following the implementation of \citet{Quinquis2014}; hydration and dehydration processes are related to the amount of bound
-water of each marker and to the maximum amount of water it can transport, i.e. if the amount of bound water exceeds the maximum amount of water, the marker
-dehydrates and releases free water that can hydrates under-saturated markers. Maximum water content of each marker is determined as function of lithology and
-$p$-$T$ conditions and it is calculated using Perple\_X, in the same way than effective density, specific heat and coefficient of thermal expansion. Bound water
-is advected together with the markers, neglecting the effect of bound water diffusion, while free water simply migrates vertically and is not coupled to the
-solid-phase flow of the mantle wedge \citep{Arcay2005,Quinquis2014}. The correctness of the migration scheme of the free water and its absorption into bound
-water by under-saturated markers are verified performing the experiment described in \citet{Quinquis2014} (Sec. \ref{sec:quinquis}).
-
-The mantle viscosity weakening related to the amount of water has been extensively studied \citep{Chopra1981,Hirth2003} and it is implemented in according to
-\citet{Arcay2005} and \citet{Horiuchi2016}, as follows
-\begin{equation}\label{eq:wet_visc}
-\eta_{wet}=\eta_{dry}\left(\left[1-\frac{1}{f_v}\right]\exp\left(-\frac{[OH^-]}{[OH^-]_0}\right)+\frac{1}{f_v}\right)
-\end{equation}
-where $f_v$ is the viscosity reduction factor between dry and wet conditions, $[OH^-]$ is the water content and $[OH^-]_0$ is a reference water content set to
-620 ppm (0.062 wt.\%) \citep{Arcay2005}. By default, $f_v$ is set to 100, in according to \citet{Arcay2005} and \citet{Horiuchi2016}.
-
-\section{Mantle melting}\label{sec:melt}
-Melt fraction in the mantle depends on temperature ($T$, in \textdegree C), pressure ($p$, in \si{\giga\pascal}) and the water content in the melt ($X_{H_2O}$, in wt.\%).
-The determination of the melt fraction $M$ is implemented as explained by \citet{Katz2003}, according to the successive modification by \citet{Langmuir2006} and
-\citet{Kelley2010}, as follows
-\begin{eqnarray}\label{eq:melt_fr}
-M(p,T,X_{H_2O})&=&-T+T_{sol}+(x_t\cdot ln(p+y_t))\cdot X_{H_2O}-\nonumber\\
-&&+K_m\left(\frac{X_{H_2O}^{bulk}}{D_{H_2O}(1-X_{H_2O})+X_{H_2O}}\right)^{\gamma}
-\end{eqnarray}
-where $T_{sol}$ is the temperature of the dry solidus, $D_{H_2O}$ is the partition coefficient and $x_t\cdot log(p+y_t)$ indicates the pressure dependence of
-the melting curve \citep{Kelley2010}. $x_t$, $y_t$, $K_m$ and $\gamma$ are parameters chosen in according to \citet{Kelley2010}.
-However, $X_{H_2O}$ depends on the melt fraction as
-\begin{equation}\label{eq:melt_wt}
-X_{H_2O}(M)=\frac{X_{H_2O}^{bulk}}{D_{H_2O}+M(1-D_{H_2O})}
-\end{equation}
-and a numerical solution can be found using a root-finder method \citep{Katz2003,Wang2016}. Furthermore, the content of water in the melt is limited by the
-pressure-dependent saturation concentration of water in the melt, determined as
-\begin{equation}\label{eq:melt_max}
-X_{H_2O}^{sat}=\chi_1 p^{\lambda_m}+\chi_2 p
-\end{equation}
-where $\chi_1$, $\chi_2$ and $\lambda_m$ are parameters chosen according to \citet{Katz2003}, \citet{Langmuir2006} and \citet{Kelley2010}. The correctness in
-the determination of melt fractions and water contents in the melt are tested comparing model results with experimental curves obtained by \citet{Katz2003}
-(Sec. \ref{sec:katz}).
-
-In case of melting, viscous creep viscosity ($\eta_{vp}$) is modified according to \citet{Wang2016}, as
-\begin{equation}\label{eq:melt_visc}
-\eta_{melt}=\eta_{vp}\cdot \exp(\alpha_m M)
-\end{equation}
-where $\alpha_m$ are melt fraction factors for dislocation and diffusion creep, chosen in according to \citet{Wang2016}.
-
-Effective density ($\rho_{eff}$) for partially molten rocks is calculated as
-\begin{equation}\label{eq:melt_dens}
-\rho_{eff}=\rho_s(1-M)+\rho_m M
-\end{equation}
-where $\rho_s$ and $\rho_m$ are the densities of the solid and the molten rock, respectively \citep{Gerya2004b,Gerya2007,Wang2016}. The density of the
-solid rock is extracted by Perple\_X, while the density of the molten portion is calculated as
-\begin{equation}\label{eq:melt_dens_M}
-\rho_m(p,T)=\rho_0[1-\alpha(T-T_0)][1+\beta(p-p_0)]
-\end{equation}
-where $\rho_0$ is the density at the reference temperature ($T_0$) and pressure($p_0$), and $\alpha$ and $\beta$ are the thermal expansion and compressibility
-coefficients, respectively \citep{Gerya2004b,Gerya2007,Wang2016}.
-
-As for the phase transitions, effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into account
-considering effective coefficient of thermal expansion. Similarly, latent heat due to melting/crystallisation can be implicitly included in the energy equation
-(Eq. \ref{eq:energy}) by considering effective specific heat ($C_{p_{eff}}$) and thermal expansion ($\alpha_{eff}$) of the partially molten rocks, in according to
-\begin{eqnarray}
-\label{eq:cp_effective}C_{p_{eff}}&=&C_p+H_L\left(\frac{\partial M}{\partial T}\right)_p \\
-\label{eq:alpha_effective}\alpha_{eff}&=&\alpha+\rho\frac{H_L}{T}\left(\frac{\partial M}{\partial p}\right)_T
-\end{eqnarray}
-where $C_p$ and $\alpha$ are the specific heat and the coefficient of thermal expansion of the solid rock, respectively, and $H_L$ is the latent heat of the
-molten rock \citep{Gerya2004b,Gerya2007,Tackley2010}.
-
-\section{Benchmarks}\label{sec:benchmark}
-\subsection{Error measurements}\label{sec:error}
-In order to determine the accuracy of velocity and pressure field of the benchmarks, the $L_2$-norm is computed by numerical integration on the Gauss-Legendre
-quadrature points. $L_2$-norm for pressure and velocity errors can be evaluated as
-\begin{eqnarray}
-\textrm{err}_p&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|p^n_i(r_q)-p^a_i(r_q)|^2\right)w_q|J_q|}\nonumber \\
-\textrm{err}_v&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|u^n_i(r_q)-u^a_i(r_q)|^2+|v^n_i(r_q)-v^a_i(r_q)|^2\right)w_q|J_q|}\nonumber
-\end{eqnarray}
-respectively, where $n_e$ is the number of elements, $n_q$ is the number of quadrature points per element, $p^n_i(r_q)$ and $p^a_i(r_q)$ are the numerical and
-analytical pressure, respectively, in each quadrature point $q$, $w_q$  and $J_q$ are the weight and the Jacobian at the quadrature point
-$q$, $u^n_i(r_q)$, $v^n_i(r_q)$, $u^a_i(r_q)$ and $v^a_i(r_q)$ are the numerical and analytical velocities, respectively, in each quadrature point $q$.
-
-Other quantities used as comparison with original benchmarks are the root-mean-square velocity over the whole domain and over the surface
-\begin{eqnarray}
-v_{\textrm{rms}}&=&\sqrt{\int_{\Omega}|\bm{v}|^2 d\Omega}\nonumber \\
-v_{\textrm{rms}}^{top}&=&\sqrt{\int_0^1 u^2 \bigg|_{y=1} dx}\nonumber
-\end{eqnarray}
-respectively, top and bottom Nusselt numbers
-\[\textrm{Nu}^{top/bottom}=-\int_0^1\frac{\partial T}{\partial y} \bigg|_{y=1/y=0} dx\]
-the average rate of work done against gravity
-\[\left\langle W \right\rangle=\int_{\Omega}T u_y d\Omega\]
-the average rate of viscous dissipation
-\[\left\langle \Phi \right\rangle=\int_{\Omega}\tau_{ij}\dot{\epsilon}_{ij} d\Omega\]
-and the percentage error between $\left\langle W \right\rangle$ and $\left\langle \Phi \right\rangle/Ra$
-\[\delta=\frac{\left|\left\langle W \right\rangle -\frac{\left\langle \Phi \right\rangle}{Ra} \right|}{\max \left(\left\langle W \right\rangle,
-\frac{\left\langle \Phi \right\rangle}{Ra}\right)}\times100\]
-
 \subsection{Stokes flow}\label{sec:stokes}
@@
 The domain is a unit square a constant viscosity ($\eta=1$) and the penalty parameter is set to $\lambda=10^7$. Velocity boundary conditions are set to no slip
-$(\bm{v}=\bm{0})$ on all boundaries. The problem is performed for different grid resolution between \num{8x8} and \num{1024x1024} elements.
+$(\bm{v}=\bm{0})$ on all boundaries. The problem is performed for different grid resolution between $8\times8$ and $1024\times1024$ elements.
 
@@
 theoretical convergence rate. The convergence of pressure, in contrast with the observation by \citet{Donea2003} for $Q_1 \times P_0$ elements, indicate the
-effectiveness of the smoothing procedure. Smoothed pressure field for grid resolution of \num{128x128} elements is shown in Fig. \ref{fig:errors}b (continuous
+effectiveness of the smoothing procedure. Smoothed pressure field for grid resolution of $128\times128$ elements is shown in Fig. \ref{fig:errors}b (continuous
 red line), in comparison with the analytical solution (dashed blue line). Solve times and memory usage needed to generate the solution are shown in
@@
 \centering
-\noindent\includegraphics[width=\textwidth]{./Figures/errors.pdf}
+\includegraphics[width=\linewidth]{./Figures/errors.pdf}
 \caption{Velocity and pressure error for the Stokes flow experiment between generated and analytical solution as a function of element size (panel a),
-and comparison between smoothed pressure and analytical solution as function of $x$ coordinate for a grid resolution of \num{128x128} elements (panel b).}
+and comparison between smoothed pressure and analytical solution as function of $x$ coordinate for a grid resolution of $128\times128$ elements (panel b).}
 \label{fig:errors}
@@
 \end{figure}
-
-\subsection{Markers advection}\label{sec:runge}
-The 2nd-order Runge-Kutta advection scheme in space is tested by means of the Zalesak disk test \citep{Zalesak1979,Thieulot2014}. The benchmark is performed
-in a unit square domain with a grid resolution of \num{32x32} elements and 1000000 markers and values of Courant number between 0.25 and 1.
-At $t=0$, the disk is centred at position $(0.5;0.75)$ with a radius $R=0.15$ and has a vertical fissure 0.05 wide and 0.2 high (orange in Fig. \ref{fig:runge})
-and with the velocity field prescribed in the entire domain as
-\begin{eqnarray}
-u(x,y)&=&2\pi\left(y-\frac{L_x}{2}\right)\nonumber \\
-v(x,y)&=&-2\pi\left(x-\frac{L_x}{2}\right)\nonumber
-\end{eqnarray}
-the markers are back to their initial location after a $2\pi$ rotation ($t=1$). The clockwise rotation of the disk at $t=0.25$, 0.5 and 0.75 
-is shown in Fig. \ref{fig:runge} (red, blue and green, respectively). The distance from the centre of 1 marker is calculated throughout the rotation to
-evaluate the error for different values of Courant number. As expected, the error increase with the increasing of the time step, as shown in 
-Fig. \ref{fig:runge_err}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/Zalesak_Disk}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=200px]{./Figures/Zalesak.pdf}
-\caption{Position of the Zalesak disk at $t=0$, 0.25, 0.5 and 0.75 (orange, red, blue and green, respectively) for the case with Courant number of 1.}
-\label{fig:runge}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/zal_marker.pdf}
-\caption{Distance from the centre as function of time for values of Courant number of  of 0.25, 0.3, 0.5, 0.75 and 1 (green, orange, blue, black and red,
-respectively). R indicates the distance from the centre considering a perfect circle trajectory.}
-\label{fig:runge_err}
-\end{figure}
-
-\subsection{Conservative Velocity Interpolation (CVI)}\label{sec:CVI}
-The CVI correction is checked by means of the Stokes flow experiment by \citet{Donea2003}, as presented in \ref{sec:stokes}. The advection of Lagrangian
-markers is performed using either a 2nd-order or a 4th-order Runge-Kutta scheme with CFL=0.5 and an initial random distribution of 25 markers per element.
-Fig. \ref{fig:CVI} shows the minimum and maximum values of markers per element throughout the experiment with and without the CVI correction and using
-different Runge-Kutta schemes. The introduction of the CVI correction allows the
-tests to run for times longer than $t=2000$ without show empty elements, while tests without the CVI correction stop at approximately $t=600$, when at least
-one element goes empty. Fig. \ref{fig:CVI_time} shows the different number of markers per element and the markers distribution at $t=600$ obtained without and
-with the CVI correction (panels b and c, respectively) in case of a 2nd-order Runge-Kutta scheme. All data can be found at
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/CVI}.
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/CVI.pdf}
-\caption{Maximum and minimum number of markers per element throughout the experiment, with (green and orange lines) or without (purple and blue lines)
-the CVI correction and using either a 2nd-order (blue and orange lines) or a 4th-order (purple and green lines) Runge-Kutta scheme.}
-\label{fig:CVI}
-\end{figure}
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/CVI_time.pdf}
-\caption{Number of markers per element and markers distribution through time using 2nd.order Runge-Kutta scheme.
-Panels a: number of markers per element (panel a1) and distribution of the markers (panel a2) after the first
-iteration; panels b and c: comparison between number of markers per element (panels b1 and c1) and markers distribution (panels b2 and c2) without and with
-the CVI correction at $t=600$; panels d: number of markers per element (panel d1) and markers distribution (panel d2) in case of CVI correction at $t=2000$.}
-\label{fig:CVI_time}
-\end{figure}
-
-\subsection{Poiseuille flow}\label{sec:poiseuille}
-The domain is a rectangle with $L_x=2$ and $L_y=1$ and constant density and viscosity ($\rho=1$ and $\eta=1$), gravity acceleration $\bm{g}=\bm{0}$ and penalty
-parameter $\lambda=10^8$. The grid is composed by \num{40x20} elements. Velocity boundary conditions are set to no slip $(\bm{v}=\bm{0})$ at the top and the
-bottom, and a parabolic profile is imposed on the sides, with $u=y(1-y)$ and $v=0$. The analytical solution is then given by:
-\begin{eqnarray}
-u(x,y)&=&y(1-y)\nonumber \\
-v(x,y,)&=&0\nonumber \\
-p(x,y)&=&2\eta\left(\frac{L_x}{2}-x\right)\nonumber
-\end{eqnarray}
-
-The velocity field predicted by the model follows the expected parabolic profile (Fig. \ref{fig:poiseuille}a and Fig. \ref{fig:poi_plot}a) and, in case of the
-classic penalty method with no iterations, the pressure is clearly related to the divergence by means of the penalty parameter (Fig. \ref{fig:poiseuille}b and
-c and Fig. \ref{fig:poi_plot}b). Fig. \ref{fig:poiseuille}d shows that one Uzawa iteration is sufficient to bring the divergence down to \num{e-15}, with no
-correlation with the pressure field. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Poiseuille%20Flow}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Poiseuille.pdf}
-\caption{Velocity field (panel a), pressure (panel b) and divergence velocity of a Poiseuille flow in case of the classic penalty method (no iterations) and
-after one Uzawa iteration (panel c and d, respectively).}
-\label{fig:poiseuille}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Analytical.pdf}
-\caption{Velocity (panel a) and pressure (panel b) field predicted by the model for a Poiseuille flow with respect with their analytical solutions. The
-velocity field is plotted as function of the vertical coordinate in $L_x/2$ and the pressure is plotted as function of the x coordinate in $L_y/2$.}
-\label{fig:poi_plot}
-\end{figure}
-
-\subsection{Instantaneous 2D sphere}\label{sec:ist_sphere}
-The domain is a unit square with gravity $\bm{g}=(0,-1)$. The fluid has constant density and viscosity ($\rho_f=1$ and $\eta_f=1$). The sphere is in the middle
-of the domain with a radius R = 0.123456798 and has constant density and viscosity ($\rho_s=10^{-2}+\rho_f$ and $\eta_s=10^3 \cdot \eta_f$).
-We distinguish three different types of velocity boundary conditions:
-\begin{enumerate}
-\item FS: free slip conditions on all sides.
-\item NS: no slip conditions on all sides.
-\item OT: free slip conditions on the sides and the bottom, and open boundary at the top.
-\end{enumerate}
-All three cases are performed for different grid resolutions, between \num{16x16} and \num{512x512} elements with 50 randomly distributed markers per element,
-and different average schemes for the viscosity (harmonic, geometric and arithmetic). The velocity in the centre of the sphere $\bm{v}(0.5;0.5)$), minimum and
-maximum velocities ($u_{min}$, $u_{max}$, $v_{min}$, $v_{max}$) and pressures ($p_{min}$, $p_{max}$) on the entire domain, $v_{\textrm{rms}}$ and average pressure on the
-entire domain ($p_{avg}$) are compared with solutions generated by ASPECT \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and all results
-can be found at \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere2D} and 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Instantaneous_2D_Sphere}. Results in terms of the $v_{\textrm{rms}}$ for NS model 
-are also shown in Fig. \ref{fig:inst_sphere}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_NS.pdf}
-\caption{$v_{\textrm{rms}}$ as function of element size for the instantaneous 2D sphere experiment in case of no slip boundary conditions and with different average
-schemes for the viscosity. Results are compared with results obtained by other numerical codes.}
-\label{fig:inst_sphere}
-\end{figure}
-
-\subsection{Rayleigh-Taylor instability}\label{sec:rayleigh}
-This problem has been originally presented by \citet{vanKeken1997} and here it is performed as the isoviscous case (Case 1a). The domain has $L_x=0.9142$ and
-$L_y=1$ and gravity $\bm{g}=(0,-1)$. Two fluids with same constant viscosities ($\eta_1=\eta_2=1$) and different densities ($\rho_1=1000$ and $\rho_2=1010$),
-with the lighter fluid at the bottom. The initial interface between the fluids is given by $y(x)=0.2+0.02 \cos \left(\frac{\pi x}{L_x}\right)$. The experiment
-is performed with different grid sizes (\num{50x50}, \num{80x80}, \num{100x100} and \num{256x256}). A total of 1960000 markers are randomly distributed at the
-beginning of the simulation. Velocity boundary conditions are set to no slip at the top and the bottom, and to free slip at the sides of the domain.
-
-$v_{\textrm{rms}}$ as function of time is reported in Fig. \ref{fig:RT}, matching well with results shown by \citet{vanKeken1997}, \citet{Tackley2003} and
-\citet{Thieulot2014}. Fig. \ref{fig:rayleigh} shows the evolution of the experiment at different time steps. All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Rayleigh_Taylor_experiment/ISOVISCOUS}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/RT.pdf}
-\caption{$v_{\textrm{rms}}$ of the Rayleigh-Taylor experiment as function of time for different resolution of the grid.}
-\label{fig:RT}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=300px]{./Figures/Rayleigh.pdf}
-\caption{Evolution of the Rayleigh-Taylor experiment for a grid of \num{256x256} elements at $t=0, 500, 1000$ and $2000$ (panels a, b, c and d, respectively).}
-\label{fig:rayleigh}
-\end{figure}
-
-\subsection{Falling blocks}\label{sec:block}
-This benchmark is described as proposed by \citet{Gerya2003a}, \citet{Gerya2010b} and \citet{Thieulot2011}. The domain is square with $L_x=L_y=$ \SI{500}{\km}
-and the grid is composed by \num{50x50} elements with 25 markers in each element. The block is initially centred at ($x=$ \SI{250}{\km}; $y=$ \SI{400}{\km})
-and has a size of \num{100x100} km (Fig. \ref{fig:block}a). The fluid surrounding the block has $\eta_f=$ \SI{e21}{\pascal\per\s} and $\rho_f=$
-\SI{3200}{\kg\per\cubic\m}. The benchmark tests with different viscosities of the block, with $\eta_b$ from \SIrange{e15}{e27}{\pascal\per\s}. In each
-experiment also the density $\rho_b$ of the block varies from \SIrange{3220}{9900}{\kg\per\cubic\m}. Velocity boundary conditions are set to free slip
-conditions on all sides of the domain. Density distributions at $t=$ \SI{20}{\mega\year} for $\rho_b=$ \SI{3300}{\kg\per\cubic\m} and different $\eta_b$
-are plotted in Fig. \ref{fig:block}b-f, showing that the code correctly preserve the block geometry in case of large viscosity contrast, with a block stiffer
-than the surrounding fluid (Fig. \ref{fig:block}f).
-
-The velocity in the centre of the falling block at $t=0$ is measured for all experiments and, since the velocity should increase with density contrast, the
-quantity $\bm{v}/(\rho_b-\rho_f)$ is plotted as function of the viscosity contrast. All results perfectly match with those from \citet{Gerya2010b} and they
-line up on a single curve, demonstrating that the code can correctly deal with large viscosity and density contrasts (Fig. \ref{fig:falling}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Falling%20blocks}.
-
-\begin{figure}
-\noindent\includegraphics[width=300px]{./Figures/Block.pdf}
-\centering
-\caption{Density evolution of the falling block experiment at $t=$ \SI{0}{\mega\year} (panel a) and $t=$ \SI{20}{\mega\year} for different viscosities of the
-block (panels b-f).}
-\label{fig:block}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Falling.pdf}
-\caption{Initial velocity relative to the density contrast at the centre of the falling block as function of the viscosity contrast between the block and the
-surrounding fluid.}
-\label{fig:falling}
-\end{figure}
-
-\subsection{2D time-dependent Stokes sphere with deformable free surface}\label{sec:sphere}
-This experiment is performed in a unit square domain with the gravity acceleration fixed to $g_y=-1$. The sphere has $\rho_s=2$ and $\eta_s=10^3$ and is
-initially centred at ($0.5;0.6$) with radius $R=0.123456789$. The fluid surrounding the sphere has $\rho_f=1$ and $\eta_f=1$ and occupies the domain for
-$y\leq 0.75$, while for $y>0.75$ the air has $\rho_a=0$ and $\eta_a=10^{-3}$. Velocity boundary conditions are set to free slip on all sides. The Courant
-number is set to 0.25. The experiments run for 200 s using grid resolutions from \num{150x150} to \num{512x512} elements, with 25 randomly distributed markers
-per element, and for different average schemes for the viscosity. The interface between the fluid and the air is tracked by means of the markers chain. Results
-in terms of velocity and pressure fields and topography variations are compared with results obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and they can be found at
-\url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere_fs2D} and
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Time_dependent_sphere}. Results in terms of the $v_{\textrm{rms}}$ in case of 
-an arithmetic average are also shown in Fig. \ref{fig:inst_spherefs}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_arithm.pdf}
-\caption{$v_{\textrm{rms}}$ for different grid resolutions as function of time for the 2D time-dependent Stokes sphere experiment in case of an arithmetic average.
-Results are compared with results obtained by ASPECT.}
-\label{fig:inst_spherefs}
-\end{figure}
-
-\subsection{Free surface stabilisation}\label{sec:stab}
-The benchmark is performed as discussed in \citet{Kaus2010a} and \citet{Thieulot2014}. The domain is square with $L_x=L_y=$ \SI{500}{\km} and a grid resolution
-of \num{200x200} elements, each of them containing 16 markers. A buoyant fluid with $\rho_1=$ \SI{3200}{\kg\per\cubic\metre} and $\eta_1=$ \SI{e20}{\pascal\s}
-is overlain by a denser fluid with $\rho_2=$ \SI{3300}{\kg\per\cubic\metre} and $\eta_2=$ \SI{e21}{\pascal\s}. The initial interface between the fluids has an
-initial sinusoidal shape of 5 km amplitude. Velocity boundary conditions are set to free slip conditions at the sides and no slip at the bottom, while the top
-is free surface. The experiment is performed with various fixed time steps, without the use of the Courant number. The vertical position of the free surface at
-$x=L_x$ is tracked for each simulation, with or without the stabilisation algorithm.
-
-The results show an instability (drunken sailor effect) increasing the time step in case that the stabilisation algorithm is not activated
-($dt=$ \SI{4200}{\year}, green line in Fig. \ref{fig:stab}), as already observed by \citet{Kaus2010a} and \citet{Thieulot2014}. Activating the algorithm the
-instability is fixed and the simulation remains stable using time step up to \SI{50000}{\year} (red line in Fig. \ref{fig:stab}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Stabilisation_algorithm}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Kaus_random.pdf}
-\caption{Evolution of the y coordinate in $x=L_x$ as function of time for different time steps, with and without the stabilisation algorithm.}
-\label{fig:stab}
-\end{figure}
-
-\subsection{Topography relaxation}\label{sec:crameri}
-This experiment is performed as in \citet{Crameri2012} using free surface and sticky air (Case 1). The domain is rectangular with $L_x=$ \SI{2800}{\km} and
-$L_y=$ \SI{700}{\km} in case of free surface and $L_y=$ \SI{800}{\km} in case of sticky air ($h_{st}=$ \SI{100}{\km}). The grid is composed by \num{256x64}
-elements for the free surface \num{560x320} elements for the sticky air. For both cases, there is a mantle of 600 km thickness with $\eta_m=$ \SI{e21}{\pascal\s}
-overlain by a highly viscous cosine-shaped lithosphere with thickness between 93 and 107 km and $\eta_l=$ \SI{e23}{\pascal\s}. Both mediums have $\rho=$
-\SI{3300}{\kg\per\cubic\metre}. In case of sticky air experiment, the air has $\rho_a=$ \SI{0}{\kg\per\cubic\metre} and $\eta_a=$ \SI{e18}{\pascal\s}. The
-gravity acceleration is set to $g_y=$ \SI{-10}{\metre\per\square\s}. Velocity boundary conditions are set to no slip at the bottom and free slip at the sides
-of the domain. In case of sticky air, velocities on top are set to free slip conditions.
-
-The maximum topography as function of time ($h(t)$) can be analytically derived in according to \[h(t)=h_0 \exp(\gamma t)\]
-where $h_0=7$ km is the initial topography, $\gamma=\num{-0.2139e-11}$ is the characteristic relaxation rate and $t$ is the time (black line in Fig.
-\ref{fig:crameri}). Maximum topography at the characteristic relaxation time $t_{rlx}=$ \SI{14.825}{\kilo\year} can be found to be $h_{rlx}=$ \SI{2576}{\m}.
-The results show that a viscosity of the air of \SI{e18}{\pascal\s} is needed in case of $h_{st}=$ \SI{100}{\km} to correctly describe the topography relaxation
-for this problem, as pointed out by \citet{Crameri2012} (purple line in Fig. \ref{fig:crameri}). The case with a true free surface well follows the analytical
-solution (orange line in Fig. \ref{fig:crameri}). In particular, at the relaxation time in case of the true free surface predicts a maximum topography of
-\SI{2570}{\m}, with an error of \SI{6}{\m} with respect to the analytical solution. Results are also compared with results obtained with a free surface in
-UNDERWORLD (\num{256x64} elements) and SULEC (\num{401x201} elements) (with courtesy from \citealp{Crameri2012}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Topography_relaxation}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Crameri.pdf}
-\caption{Maximum topography as function of time for the Crameri benchmark (coloured lines) in comparison with the analytical solution (black line) and results
-obtained with UNDERWORLD (yellow line) and SULEC (blue line).}
-\label{fig:crameri}
-\end{figure}
-
-\subsection{Spontaneous subduction}\label{sec:subduction}
-This experiment is performed as presented by \citet{Schmeling2008}, in case of both a sticky air and a true free surface and using different average schemes
-for the viscosities. The domain is rectangular with $L_x=$ \SI{3000}{\km} and $L_y=$ \SI{700}{\km}, a grid resolution of \num{256x64} elements with 50 markers
-per element and Courant number of 0.01. At the beginning of the simulation, a 100 km-thick lithospheric layer with $\rho_m=$ \SI{3300}{\kg\per\cubic\metre}
-and $\eta_m=$ \SI{e23}{\pascal\s} is located at the top of the domain for $x$ between 1000 and 3000 km, while the asthenospheric mantle has $\rho_m=$
-\SI{3200}{\pascal\s} and $\eta_m=$ \SI{e21}{\pascal\s}. In addition, a 200 km-depth lithospheric slab is already subducted in the mantle in order to have
-a spontaneous subduction. In case of sticky air, $L_y=750$ km and a 50 km-thick air layer with $\eta_a$ from \SIrange{e19}{e21}{\pascal\s} overlie the mantle.
-Velocity boundary conditions are set to free slip on the sides and on the bottom of the domain. In case of sticky air, velocities on the top boundary are set
-to free slip conditions as well.
-
-The maximum depth of the slab is tracked for all the simulations (Fig. \ref{fig:subduction}) and they are compared with results obtained by
-\citet{Schmeling2008} with models with regular grids and comparable grid resolutions (rectangles in Fig. \ref{fig:subduction}). Low viscous sticky air models
-(i.e., \SI{e19}{\pascal\s}) enlighten a strong dependence of the sinking velocity with the chosen average scheme (dashed lines in Fig. \ref{fig:subduction}),
-while high viscous sticky air models (i.e., \SI{e21}{\pascal\s}) show a higher resistance at the trench, underestimating the correct solution (dotted lines in
-Fig. \ref{fig:subduction}). A high resistance at the trench can be also observed in case of a true free surface for the low resolution grid used (continuous
-lines in Fig. \ref{fig:subduction}), as pointed out by \citet{Schmeling2008}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Spontaneous_subduction}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Subduction.pdf}
-\caption{Maximum depth of the slab as function of time for all the simulations of the \citet{Schmeling2008} benchmark, in case of harmonic, geometric and
-arithmetic means (red, green and blue lines, respectively) using a sticky air or a true free surface (discontinuous and continuous lines, respectively).
-Red, green and blue rectangular areas indicate the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{500}{\km} in case of sticky air,
-while the grey rectangular area indicates the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{600}{\km} in case of true free surface.}
-\label{fig:subduction}
-\end{figure}
-
-\subsection{The slab detachment}\label{sec:slab}
-The slab detachment benchmark is performed as described by \citet{Schmalholz2011} and \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{1000}{\km}
-and $L_y=$ \SI{660}{\km} and a grid resolution of \num{256x256} elements. A non-linear viscous T-shaped layer with $\rho_l=$ \SI{3300}{\kg\per\cubic\metre} is
-placed at the top of the domain and surrounded by a linear viscous fluid with $\rho_f=$ \SI{3150}{\kg\per\cubic\m}. The top layer is 80 km-thick and a 250
-km-long and 80 km-wide slab is placed at $x=L_x/2$ (Fig. \ref{fig:slab}a). The effective viscosity of the top layer is given by
-\[\eta_{eff}=\eta_0 I_2^{\frac{(1-n)}{n}}\]
-with $\eta_0=$ \SI{4.75e11}{\pascal\s} and $n=4$, while the surrounding fluid has $\eta_f=$ \SI{e21}{\pascal\s}. The effective viscosity is capped between
-$\eta_{min}=$ \SI{e21}{\pascal\s} and $\eta_{max}=$ \SI{e25}{\pascal\s}. Velocity boundary conditions are set to free slip at the top and the bottom, and to
-no slip at the sides of the domain.
-
-The sinking of the slab determines a decrease of the effective viscosity in the T-shaped layer during the first part of the evolution (Fig. \ref{fig:slab}a),
-as consequence of local strain rates, while effective viscosities increase up to $\eta_{max}$ in the top layer after approximately \SI{20}{\mega\year}, when
-necking is complete (Fig. \ref{fig:slab}b).
-
-The width of necking is tracked by means of markers position at the side of the slab. Necking width and time are normalised in function of the initial slab
-(\SI{80}{\km}) and the characteristic time ($t_c=$ \SI{7.1158e14}{\s}) \citep{Schmalholz2011,Glerum2018}. The results in case of arithmetic, geometric and
-harmonic mean of the viscosity are shown in Fig. \ref{fig:necking}, compared with results from \citet{Glerum2018} for which an infinity norm mean was used.
-The influence of the average schemes agrees with results shown in Section \ref{sec:subduction}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Slab_detachment}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Slab.pdf}
-\caption{Effective viscosity for the slab detachment benchmark at the beginning of the evolution (panel a) and when necking is complete (panel b).}
-\label{fig:slab}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Necking.pdf}
-\caption{Normalised width of the necking of the slab detachment benchmark as function of normalised time.}
-\label{fig:necking}
-\end{figure}
-
-\subsection{The indenter experiment}\label{sec:indenter}
-The indenter benchmark simulate a rigid punch on a purely plastic von Mises material, for which exists an analytical solution
-\citep{Thieulot2008,Thieulot2014,Glerum2018}. The domain is a unit square with a grid resolution of \num{256x256} elements. The gravity acceleration is set
-to $\bm{g}=\bm{0}$ and effective viscosity is capped using $\eta_{min}=10^{-4}$ and $\eta_{max}=10^3$. The tolerance for the convergence of the non-linear
-solution is $tol=10^{-9}$, with a maximum number of non-linear iterations set to 500. The medium in the domain has $\rho=0.01$, an initial viscosity
-$\eta_0=10$, cohesion $C=1$ and an angle of internal friction $\phi=0$°. Velocity boundary conditions are set to no slip at the bottom and free slip at the
-sides of the domain. The top of the domain is open with the exception of the central portion (punch area) with width $w_p=0.125$, where $v=-1.05$ and $u$ is
-fixed either to 0 (rough punch experiment) or free (smooth punch experiment).
-
-Fig. \ref{fig:indenter} show the results in terms of viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure
-(panels g and h) for both experiments. The analytical solution indicate that pressure at ($0.5;1$) and ($0.5 \pm w_p;1$)is $p=\pi +1$ and $p=1$, respectively
-(black and red lines in Fig. \ref{fig:smooth_p}). Fig. \ref{fig:smooth_p} shows a clear improvement in the pressure solution when passing from a rough to a
-smooth punch (green and blue lines in Fig. \ref{fig:smooth_p}, respectively), as previously observed by \citet{Thieulot2014} and \citet{Glerum2018}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Indenter}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Indenter.pdf}
-\caption{Viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure (panels g and h) fields for rough (left column) and
-smooth (right column) punch experiments.}
-\label{fig:indenter}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Smooth_Rough.pdf}
-\caption{Nodal pressure in function of the x coordinate at $y=1$ for rough (green line) and smooth (blue line) punch experiment. Black and red lines indicate
-the analytical solution at $x=0.5$ and $x=0.5 \pm w_p$, respectively.}
-\label{fig:smooth_p}
-\end{figure}
-
-\subsection{The brick experiment}\label{sec:brick}
-An instantaneous version of the brick benchmark is performed to verify the correctness of the pressure-dependent plasticity for different angle of internal
-friction, as proposed by \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{40}{\km} and $L_y=$ \SI{10}{\km} and a grid resolution of \num{512x128}
-elements. The gravity acceleration is set to $g_y=$ \SI{-10}{\m\square\s} and effective viscosity is capped using $\eta_{min}=$ \SI{e19}{\pascal\s} and
-$\eta_{max}=$ \SI{e26}{\pascal\s}. The tolerance for the convergence of the non-linear solution is $tol=10^{-7}$, with a maximum number of non-linear iterations
-set to 1000. A 800 m-wide and 400 m-height inclusion with $\eta_b=$ \SI{e20}{\pascal\s} and $\rho_b=$ \SI{2700}{\kg\per\cubic\m} is placed at the bottom of the
-domain at $x=L_x/2$. The inclusion is surrounded by a non-linear viscous medium with $\rho_m=$ \SI{2700}{\kg\per\cubic\m}, an initial viscosity of
-\SI{e23}{\pascal\s}, a linear viscous viscosity of \SI{e25}{\pascal\s} and a cohesion of \SI{40}{\mega\pascal}. Velocities are set to free slip conditions at
-the bottom of the domain and the top is open. Velocities on sides of the domain are fixed to $u=$ \SI{+-2e-11}{\m\per\s} and $v=0$. The experiment is performed
-in compressional and extensional contexts with an angle of internal friction $\phi$ of the non-linear medium variable between 0° and 30°.
-
-As expected, two shear bands stem from the inclusion with variable angles in relation to both the dynamics context and the internal friction angle. Shear band
-angles formed at 45° for $\phi=0$° in both extensional and compressional contexts (Fig. \ref{fig:brick_beam}a), while they are different in case of $\phi\neq0$°.
-Shear band angles in case of $\phi=20$° are shown for compression and extension in Fig. \ref{fig:brick_beam}b  and c, respectively. Values of shear band angles
-as function of internal friction angles from 0° to 30° are extracted at different depths and minimum and maximum values are plotted in Fig. \ref{fig:brick},
-in comparison with theoretical Roscoe, Arthur and Coulomb shear band angles. For all tests the tolerance for nonlinear convergence is defined $tol=10^{-7}$
-and none converges before the maximum number of iterations ($it_{max}=1000$) is reached. However, tests with internal friction angles up to 15° show a constant
-decrease of the velocity residuals, which can not be observed for for tests with higher internal friction angles (Fig. \ref{fig:convergence}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Brick_experiment}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Brick_Beam.pdf}
-\caption{Shear band angles predicted for the brick experiment for $\phi=0$° (panel a) and for $\phi=20$° in case of compression and extension (panels b and c,
-respectively).}
-\label{fig:brick_beam}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Brick.pdf}
-\caption{Shear band angles predicted for the brick experiment in case of compressional and extensional contexts as function of different internal angle of
-friction (continuous black line and red dots), compared with theoretical Roscoe, Artur and Coulomb angles (discontinuous black lines). Red lines indicate the
-range of angles calculated at different depths.}
-\label{fig:brick}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \noindent\includegraphics[width=500px]{./Figures/convergence.pdf}
-  \caption{Convergence rates for all tests. Normalised residuals (left column) are compared with velocity residuals (right column) for extensional (first row)
-  and compressional (second row) settings.}
-  \label{fig:convergence}
-  \end{figure}
-
-\subsection{Advection stabilisation}\label{sec:advection}
-The 1D advection problem proposed by \citet{Donea2003} and \citet{Thieulot2011} is performed to verify the effectiveness of the SUPG method to stabilise the
-advection term of the energy equation. The domain is a 1D segment with $L_x=1$ composed by 50 elements and a discontinuity in the thermal field placed at
-$x=0.25$. Temperature is set to 1 for $x \leq 0.25$ and to 0 for $x > 0.25$. Velocity is set to ${u}=1$ in the entire domain. The simulation is performed for
-250 time steps, with $dt=0.002$, so the thermal discontinuity should be at $x=0.75$ at the end of the simulation.
-
-Temperature profiles at the end of the simulation are shown in \ref{fig:advection} as function of the dimensionless coefficient $\gamma=\tau \bm{v}/h$. In case
-of the classic Galerkin method ($\gamma=0$, blue line in \ref{fig:advection}) the final thermal profile is characterised by strong oscillations, which are
-eliminated in case of the SUPG method ($\gamma=0.045$, orange line in \ref{fig:advection}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Advection_stabilisation}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Advection.pdf}
-\caption{Temperature profile as function of \textit{x} for the advection stabilisation benchmark. Purple line indicates the initial temperature profile; the
-green line indicates the analytical temperature profile after 250 time steps; blue line indicates the temperature profile after 250 time steps in case of the
-classic Galerkin method ($\gamma=0$); orange line indicates the temperature profile after 250 time steps in case of the SUPG method ($\gamma=0.045$).}
-\label{fig:advection}
-\end{figure}
-
-\subsection{Shear heating}\label{sec:simple_shear}
-This exercise is performed in an unit square domain composed by \num{128x128} elements. The velocity field is prescribed on the entire domain with
-$\bm{v}=(L_y-y)y\bm{e_x}$; viscosity, density and specific heat are fixed to 1, while thermal conductivity and radiogenic energy are fixed to 0. Therefore,
-the energy equation (Eq. \ref{eq:energy}) can be simplified as \[\frac{\partial T}{\partial t}=H_s\] and fixing $T(t=0)=0$, the temperature can be find as
-\[T(t)=H_s t\]
-In this case we have
-\begin{eqnarray}
-\dot{\epsilon_{xx}}&=&\frac{\partial u}{\partial x}=0 \nonumber \\
-\dot{\epsilon_{yy}}&=&\frac{\partial v}{\partial y}=0 \nonumber \\
-\dot{\epsilon_{xy}}&=&\frac{1}{2}\left(\frac{\partial v}{\partial x}+\frac{\partial u}{\partial y}\right)=\frac{1}{2}(L_y-2y)\nonumber
-\end{eqnarray}
-and, simplifying Eq. \ref{eq:shear}, shear heating can be calculated as \[H_s(x,y)=(1-2y)^2\]
-
-The solution predicted by the model in terms of velocity, temperature and shear heating match well with the analytical solutions (Fig. \ref{fig:analytical_en}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Simple_shear}.
-
-%\begin{figure}
-%\centering
-%\noindent\includegraphics[width=400px]{./Figures/Energy_Thieulot.pdf}
-%\caption{Temperature and shear heating predicted by the model after 1 time step.}
-%\label{fig:en_thieulot}
-%\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=500px]{./Figures/Analytical_Shear.pdf}
-\caption{Velocity (panel a), temperature (panel b) and shear heating (panel c) as function of the \textit{y} coordinate for the simple shear experiment. The
-solutions predicted by the model (red lines) are compared with the analytical solutions (dashed blue lines).}
-\label{fig:analytical_en}
-\end{figure}
-
-\subsection{Shear and adiabatic heating}\label{sec:shear}
-This problem is performed as presented in Exercise 9.4 in \citet{Gerya2010b}. Two materials are vertically separated in a rectangular domain with $L_x=$
-\SI{1000}{\km}, $L_y=$ \SI{1500}{\km} and a grid resolution of \num{30x20} elements. Constant thermal coefficient expansion ($\alpha=$ \SI{3e-5}{\per\kelvin}),
-temperature ($T=$ \SI{1300}{\kelvin}) and gravity acceleration ($g_y=$ \SI{-10}{\m\square\s}) are assumed in the whole domain. Fluid 1 (on the left side of the
-domain) has $\rho_1=$ \SI{3200}{\kg\per\cubic\m} and $\eta_1=$ \SI{e20}{\pascal\s}; fluid 2 (on the right side of the domain) has $\rho_2=$
-\SI{3300}{\kg\per\cubic\m} and $\eta_2=$ \SI{e22}{\pascal\s}. Velocity boundary conditions are set to free slip on all sides of the domain.
-
-As shown in Fig. \ref{fig:energy}, both shear and adiabatic heating predicted by the code (first row in Fig. \ref{fig:energy}) well recreate the results
-obtained by means of example ${Shear\_adiabatic\_heating.m}$ from \citet{Gerya2010b} (second row in Fig. \ref{fig:energy}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Adiabatic%2BShear}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Energy.pdf}
-\caption{Comparison between shear (first column), adiabatic (second column) and total (third column) energy predicted by the code (panels a, b and c) and those
-created using example ${Shear\_adiabatic\_heating.m}$ from Exercise 9.4 in \citet{Gerya2010b} (panels d, e and f).}
-\label{fig:energy}
-\end{figure}
-
-\subsection{Mantle convection}\label{sec:mantle}
-This problem is performed as presented by \citet{Blankenbach1989} (constant viscosity cases) and \citet{Thieulot2014}, in a 2D unit square domain with gravity
-acceleration $g_y=10^{10}Ra$. The experiment is performed with three different Rayleigh numbers ($Ra=10^4$, $10^5$ and $10^6$) and with different grid
-resolution (between \num{32x32} and \num{128x128} elements). The fluid has constant viscosity, initial density, heat capacity, thermal conductivity
-($\eta= \rho_0= C_p = k=1$), reference temperature ($T_0=0$) and thermal expansion coefficient ($\alpha=10^{-10}$). Temperatures are set to 0 on top and 1 on
-bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\]
-
-The solution generated by the code in terms of $v_{\textrm{rms}}$ (calculated as in Sec. \ref{sec:error}) and Nu as function of time are reported for all the
-simulations in Fig. \ref{fig:mantle}. At the steady state, $v_{\textrm{rms}}$ and Nu of all simulations converge well toward the values from \citet{Blankenbach1989},
-with lower errors for higher resolution grids (Table \ref{tab:mantle}). All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Mantle_convection}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Convection.pdf}
-\caption{$v_{\textrm{rms}}$ (panels a, c and e) and Nu (panels b, d and f) for the mantle convection benchmark as function of time for different grid resolution.
-Panels a and b show the results for $Ra=10^4$; Panels c and d show the results for $Ra=10^5$; panels e and f show the results for $Ra=10^6$. Purple lines
-indicate the convergence values for the $v_{\textrm{rms}}$ and the Nu from \citet{Blankenbach1989} at the steady state.}
-\label{fig:mantle}
-\end{figure}
-
-\begin{table}[H]
-\caption{Comparison between $v_{\textrm{rms}}$ and Nu predicted by the code for the mantle convection experiment and same values as reported in literature.}
-\centering
-\small
-\begin{tabular}{|>{\centering}m{0.09\textwidth}|>{\centering}m{0.05\textwidth}|>{\centering}m{0.24\textwidth}|>{\centering}m{0.15\textwidth}|
-  >{\centering}m{0.1\textwidth} >{\centering}m{0.1\textwidth} >{\centering\arraybackslash}m{0.1\textwidth}|}
-\toprule
-& & \multirow{2}{*}{\citet{Blankenbach1989}} & \citet{Thieulot2014} & \multicolumn{3}{c|}{FALCON}  \\
- & & & \num{200x200} & \num{32x32} & \num{64x64} & \num{128x128}  \\
-\midrule
-\multirow{2}{*}{$Ra=10^4$} & $v_{\textrm{rms}}$ & $42.864947 \pm 0.000020$ & 42.867 & 42.83226 & 42.852793 & 42.861394  \\
-           & Nu      & $4.884409 \pm 0.000010$  & 4.882  & 4.781297 & 4.857475  & 4.877573   \\
-\hline
-\multirow{2}{*}{$Ra=10^5$} & $v_{\textrm{rms}}$ & $193.21454 \pm 0.00010$  & 193.255& 193.872643 & 193.377472 & 193.252290  \\
-           & Nu      & $10.534095 \pm 0.000010$ & 10.507 & 9.602514 & 10.270735  & 10.465629   \\
-\hline
-\multirow{2}{*}{$Ra=10^6$} & $v_{\textrm{rms}}$ & $833.98977 \pm 0.00020$  & 834.712& 848.091176 & 837.767911 & 834.945793  \\
-           & Nu      & $21.972465 \pm 0.000020$ & 21.695 & 15.999266 & 19.703682  & 21.306939   \\
-\bottomrule
-\end{tabular}
-\label{tab:mantle}
-\end{table}
-
-\subsection{Viscoplastic mantle convection}\label{sec:visco_mantle}
-This benchmark is performed as Cases 1-5a described in \citet{Tosi2015} in a 2D unit square domain with two different grid resolutions of \num{32x32} 
-and \num{100x100} elements. Reference viscosity, reference density, heat capacity, thermal conductivity and thermal expansion coefficient are fixed to a constant
-value of 1 ($\eta_0= \rho_0= C_p = k=\alpha=1$), while the gravity has been chosen as $g_y=10^2$, in order to obtain a Rayleigh number $Ra=10^2$. Temperatures
-are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\].
-The viscosity field $\eta$ is calculated as the harmonic average between a linear part $\eta_{lin}$ that depends on temperature only or on temperature and 
-depth $y$ and a nonlinear-plastic part $\eta_{pl}$ that depends on the strain rate $\dot{\bm{\epsilon}}$, as follows
-\[\eta(T,y,\dot{\bm{\epsilon}})=2\left(\frac{1}{\eta_{lin}(T,y)}+\frac{1}{\eta_{pl}(\dot{\bm{\epsilon}})}\right)^{-1}\]
-The linear and the nonlinear-plastic parts of the viscosity are calculated following \citet{Tosi2015} as follows
-\[\eta_{lin}(T,y)=\exp(-\ln(\Delta\eta_T)T+\ln(\Delta\eta_y)y)\]
-\[\eta_{pl}(\dot{\bm{\epsilon}})=\eta^*+\frac{\sigma_Y}{\sqrt{\dot{\bm{\epsilon}}:\dot{\bm{\epsilon}}}}\]
-where $\Delta\eta_T$, $\Delta\eta_y$, $\eta^*$ and $\sigma_Y$ are parameters chosen as listed in Table \ref{tab:tosi_setup}.
-
-For each case, some diagnostic quantities measured at the steady state are reported in Table \ref{tab:visco_mantle} (see Section \ref{sec:error} for 
-the computation of all the quantities), showing a well fit with results described in \citet{Tosi2015}. Data and charts for all cases can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Viscoplastic_mantle_convection}.
-
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{white!100!}{gray!25!}
-%\begin{table}[H]
-%  \caption{Results of the viscoplastic mantle convection benchmark for Cases 1-4 compared with results obtained with ELEFANT and YACC (from \citealp{Tosi2015}).}
-%  \centering
-%  \resizebox{\textwidth}{!}{
-%\begin{tabular}{|c|ccc|ccc|ccc|ccc|}
-%  \toprule
-%    & \multicolumn{3}{c|}{\textbf{Case 1}} & \multicolumn{3}{c|}{\textbf{Case 2}} & \multicolumn{3}{c|}{\textbf{Case 3}} &
-%    \multicolumn{3}{c|}{\textbf{Case 4}} \\
-%   & FALCON & ELEFANT& YACC & FALCON& ELEFANT& YACC & FALCON & ELEFANT & YACC & FALCON & ELEFANT & YACC \\
-%  \midrule
-%   $\left\langle T \right\rangle$ & 0.7759 & 0.7758 & 0.7767 & 0.6043 & 0.6036 & 0.6058 & 0.7275 & 0.7275 & 0.7286 & & 0.5277 & 0.5289\\
-%    Nu$^{top}$ & 3.4280 & 3.4214 & 3.4298 & 8.5242 & 8.5115 & 8.5278 & 3.0368 & 3.0347 & 3.0374 && 6.5912 & 6.5572 \\
-%    Nu$^{bottom}$ & 3.3168 & 3.313 & 3.3143 & 8.4599 & 8.4465 & 8.3990 & 3.0086 & 2.9908 & 2.9628 && 6.5834 & 6.5243 \\
-%    $v_{\textrm{rms}}$ & 249.901 & 249.134 & 251.7997 & 141.2451 & 140.7087 & 142.2020 & 100.2776 & 100.1208 & 100.9467 && 79.1105 & 79.6202 \\
-%    $v_{\textrm{rms}}^{top}$ & 1.8907 & 1.8642 & 1.8298 & 104.7928 & 104.1228 & 105.1767 & 2.0832 & 2.0652 & 2.0374 && 74.7596 & 75.8414 \\
-%    $u_{max}^{top}$ & 2.6340 & 2.6119 & 2.5516 & 121.9212& 121.766 & 122.3238 & 2.9086 & 2.9019 & 2.8458 && 88.9146 & 89.2940 \\
-%    $\eta_{min}$ & - & - & - & $2.0303\times10^{-5}$ & $2.0600\times10^{-5}$ & $1.9853\times10^{-5}$ & $4.8129\times10^{-5}$ & $4.8080\times10^{-5}$ & $4.7907\times10^{-5}$ &
-%    & $1.9860\times10^{-4}$ & $1.9174\times10^{-4}$ \\
-%    $\eta_{max}$ & - & - & - & 1.3325 & 1.85 & 1.6095 & 0.9360 & 0.9023 & 1 && 1.52 & 1.6773 \\
-%    $\left\langle W \right\rangle$ & 2.4299 & 2.4316 & 2.4583 & 7.5648 & 7.5665 & 7.6222 & 2.0387 & 2.0384 & 2.0400 && 5.6216 & 5.6512 \\
-%    $\left\langle \Phi \right\rangle/Ra$ & 2.4311 & 2.4276 & 2.4333 & 7.5661 & 7.5585 & 7.6194 & 2.0386 & 2.0370 & 2.0335 && 5.6182 & 5.6463 \\
-%    $\delta$ & 0.05\% & 0.16\% & 1.02\% & 0.02\% & 0.11\% & 0.04\% & $<$0.01\% & 0.07\% & 0.32\% && 0.06\% & 0.09\% \\
-%  \bottomrule
-%  \end{tabular}}
-%  \label{tab:visco_mantle}
-%  \end{table}
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{2}{white!100!}{gray!25!}
-\begin{table}[H]
-  \caption{Benchmark cases and corresponding parameters.}
-  \centering
-  \small
-\begin{tabular}{|>{\centering}m{0.08\textwidth}|>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}
-  >{\centering}m{0.05\textwidth} >{\centering}m{0.05\textwidth} |>{\centering\arraybackslash}m{0.2\textwidth}|}
-  \toprule
-    Case & $Ra$ & $\Delta\eta_T$ & $\Delta\eta_y$ & $\eta^*$ & $\sigma_Y$ & Convective regime \\
-  \midrule
-  \rowcolor{gray!25!}
-    1 & $10^2$ & $10^5$ & 1 & - & - & Stagnant lid \\
-    2 & $10^2$ & $10^5$ & 1 & $10^{-3}$ & 1 & Mobile lid  \\
-    \rowcolor{gray!25!}
-    3 & $10^2$ & $10^5$ & 10 & - & - & Stagnant lid  \\
-    4 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 1 & Mobile lid \\
-    \rowcolor{gray!25!}
-    5 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 4 & Periodic \\
-  \bottomrule
-  \end{tabular}
-  \label{tab:tosi_setup}
-  \end{table}
-
-\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{gray!25!}{white!100!}
-\begin{table}
-  \caption{Results of the viscoplastic mantle convection benchmark for different grid resolutions in Cases 1-5 compared with results obtained with ELEFANT
-  and YACC ($100\times100$ elements, from \citealp{Tosi2015}) and in Stone 28 ($32\times32$ elements, from
-  \url{https://github.com/cedrict/fieldstone/tree/master/python_codes/fieldstone_28}).}
-  \centering
-  \resizebox{\textwidth}{!}{
-\begin{tabular}{|cc|ccccccccccc|}
-  \toprule
-  \multicolumn{13}{|c|}{\textbf{Case 1}} \\
-  \midrule
-  Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-  & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-  \hline
-  \rowcolor{gray!25!}
-    YACC & $100\times100$ & 0.7767 & 3.4298 & 3.3143 & 251.7997 & 1.8298 & 2.5516 & - & - & 2.4583 & 2.4333 & 1.02\% \\
-    ELEFANT & $100\times100$ & 0.7758 & 3.4214 & 3.313 & 249.134 & 1.8642 & 2.6119 & - & - & 2.4316 & 2.4276 & 0.16\% \\
-    \rowcolor{gray!25!}
-    Stone 28 & $32\times32$ & 0.7737 & 3.3987 & - & 243.872 & - & - & - & - & - & - & - \\
-%    FALCON & $32\times32$ & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & - & - & 2.4745 & 2.4864 & 0.48\% \\
-%    \rowcolor{gray!25!}
-    FALCON & $100\times100$ & 0.7759 & 3.4280 & 3.3168 & 249.901 & 1.8907 & 2.6340 & - & - & 2.4299 & 2.4311 & 0.05\% \\
-    \midrule
-    \multicolumn{13}{|c|}{\textbf{Case 2}} \\
-    \midrule
-    Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-    & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-    \hline
-    \rowcolor{gray!25!}
-     YACC & $100\times100$ & 0.6058 & 8.5278 & 8.3990 & 142.2020 & 105.1767 & 122.3238 & $1.9853\times10^{-5}$ & 1.6095 & 7.6222 & 7.6194 & 0.04\% \\
-     ELEFANT & $100\times100$ & 0.6034 & 8.5115 & 8.4465 & 140.8067 & 104.1228 & 121.766 & $2.0600\times10^{-5}$ & 1.85 & 7.5665 & 7.5585 & 0.11\% \\
-     \rowcolor{gray!25!}
-     Stone 28 & $32\times32$ & 0.6052 & 8.1715 & - & 141.518 & - & - & - & - & - & - & - \\
-%     FALCON & $32\times32$ & 0.6454 & 8.1422 & 7.4020 & 156.6908 & 95.7226 & 116.9302 & $2.2572\times10^{-5}$ & 0.7194 & 7.4289 & 7.4452 & 0.22\% \\
-%     \rowcolor{gray!25!}
-     FALCON & $100\times100$ & 0.6044 & 8.5266 & 8.4590 & 141.3684 & 104.8867 & 122.0470 & $2.0333\times10^{-5}$ & 1.2914 & 7.5707 & 7.5718 & 0.01\% \\
-     \midrule
-     \multicolumn{13}{|c|}{\textbf{Case 3}} \\
-     \midrule
-     Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-     & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-     \hline
-     \rowcolor{gray!25!} 
-     YACC & $100\times100$ & 0.7286 & 3.0374 & 2.9628 & 100.9467 & 2.0374 & 2.8458 & $4.7907\times10^{-5}$ & 1 & 2.0400 & 2.0335 & 0.32\% \\
-      ELEFANT & $100\times100$ & 0.7275 & 3.0347 & 2.9908 & 100.1208 & 2.0652 & 2.9019 & $4.8080\times10^{-5}$ & 0.9023 & 2.0384 & 2.037 & 0.07\% \\
-      \rowcolor{gray!25!}
-      Stone 28 & $32\times32$ & 0.7277 & 3.0347 & - & 100.018 & - & - & - & - & - & - & - \\
-%      FALCON & $32\times32$ & 0.7281 & 3.0618 & 2.8132 & 101.0092 & 2.1630 & 3.0178 & $4.8500\times10^{-5}$ & 0.7845 & 2.0699 & 2.0695 & 0.02\% \\
-%      \rowcolor{gray!25!}
-      FALCON & $100\times100$ & 0.7275 & 3.0368 & 3.0086 & 100.2776 & 2.0832 & 2.9086 & $4.8129\times10^{-5}$ & 0.9360 & 2.0387 & 2.0386 & $<$0.01\% \\
-      \midrule
-      \multicolumn{13}{|c|}{\textbf{Case 4}} \\
-      \midrule
-      Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-      & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-      \hline
-      \rowcolor{gray!25!} 
-      YACC & $100\times100$ & 0.5289 & 6.5572 & 6.5243 & 79.6202 & 75.4814 & 89.2940 & $1.9174\times10^{-4}$ & 1.6773 & 5.6512 & 5.6463 & 0.09\% \\
-       ELEFANT & $100\times100$ & 0.5277 & 6.5912 & 6.5834 & 79.1105 & 74.7596 & 88.9146 & $1.9860\times10^{-4}$ & 1.5200 & 5.6216 & 5.6182 & 0.06\% \\
-       \rowcolor{gray!25!}
-       Stone 28 & $32\times32$ & 0.5286 & 6.4036 & - & 79.4746 & - & - & - & - & - & - & - \\
-%       FALCON & $32\times32$ & 0.5334 & 6.4679 & 6.3435 & 80.6153 & 76.0511 & 90.2058 & $2.0393\times10^{-4}$ & 0.6453 & 5.6676 & 5.6698 & 0.04\% \\
-%       \rowcolor{gray!25!}
-       FALCON & $100\times100$ & 0.5279 & 6.5897 & 6.5830 & 79.1220 & 75.0715 & 88.8484 & $1.9693\times10^{-4}$ & 1.4133 & 5.6127 & 5.6113 & 0.03\% \\
-       \midrule
-       \multicolumn{13}{|c|}{\textbf{Case 5}} \\
-       \midrule
-       Code & Resolution & $P$ & $\left\langle T \right\rangle_{min}$ & $\left\langle T \right\rangle_{max}$ & Nu$^{top}_{min}$ & Nu$^{top}_{max}$ &
-       $v_{\textrm{rms}}^{min}$ & $v_{\textrm{rms}}^{max}$ & \multicolumn{2}{c}{$(\left\langle \Phi \right\rangle/Ra)_{min}$} & 
-       \multicolumn{2}{c|}{$(\left\langle \Phi \right\rangle/Ra)_{max}$} \\
-       \hline
-       \rowcolor{gray!25!}
-         YACC & $100\times100$ & 0.0792 & 0.6543 & 0.6722 & 2.6950 & 7.2792 & 41.9813 & 98.8230 & \multicolumn{2}{c}{1.4347} & 
-         \multicolumn{2}{c|}{9.2748} \\
-         ELEFANT & $100\times100$ & 0.076 & 0.6514 & 0.6693 & 2.6808 & 7.3506 & 41.2578 & 100.089 & \multicolumn{2}{c}{1.3324} & 
-         \multicolumn{2}{c|}{9.5289} \\
-         \rowcolor{gray!25!}
-         Stone 28 & $32\times32$ & 0.0764 & 0.6521 & 0.6697 & 2.6912 & 7.0763 & 41.61 & 98.183 & \multicolumn{2}{c}{-} & 
-         \multicolumn{2}{c|}{-} \\
-     %    FALCON & $32\times32$ & & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & \multicolumn{2}{c}{-} & 
-     %     \multicolumn{2}{c|}{-} \\
-     %    \rowcolor{gray!25!}
-         FALCON & $100\times100$ & 0.0775 & 0.6519 & 0.6697 & 2.6847 & 7.3629 & 41.5256 & 99.5466 & \multicolumn{2}{c}{1.3399} & 
-         \multicolumn{2}{c|}{9.3993} \\
-  \bottomrule
-  \end{tabular}}
-  \label{tab:visco_mantle}
-  \end{table}
-
-\subsection{Thin layer entrainment}\label{sec:thin}
-This experiment is performed as originally proposed by \citet{vanKeken1997} with two incompressible fluids in a rectangular domain with $L_x=2$ and $L_y=1$
-and gravity acceleration $g_y=10^{10}Ra$. The Rayleigh number and the compositional Rayleigh number are fixed to $Ra=300000$ and $Rc=450000$, respectively.
-Both fluids have constant viscosity, thermal conductivity, specific heat ($\eta=\rho=C_p=1$) and thermal expansion coefficient ($\alpha=10^{-10}$). Fluid 1
-has a density $\rho_1=1$, while fluid 2 is denser ($\rho_2=\rho_1+1.5 \cdot 10^{-10}$) and is placed at the bottom of the domain, for $y \leq 0.025$.
-Temperature are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides of the domain. The initial
-temperature field is given by
-\[T(x,y)=T_u(x,y)+T_l(x,y)+T_r(x,y)+T_s(x,y)-\frac{3}{2}\]
-with
-\begin{eqnarray}
-T_u(x,y)&=&\frac{1}{2}erf\left(\frac{1-y}{2}\sqrt{\frac{u_0}{x}}\right)\nonumber \\
-T_l(x,y)&=&1-\frac{1}{2}erf\left(\frac{y}{2}\sqrt{\frac{u_0}{L_x-x}}\right)\nonumber \\
-T_r(x,y)&=&\frac{1}{2}+\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{y+1}} \exp\left(-\frac{x^2u_0}{4y+4}\right)\nonumber \\
-T_s(x,y)&=&\frac{1}{2}-\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{2-y}} \exp\left(-\frac{(L_x-x)^2u_0}{8-4y}\right)\nonumber
-\end{eqnarray}
-and
-\begin{eqnarray}
-u_0&=&\frac{L_x^{7/3}}{(1+L_x^4)^{2/3}}\left(\frac{Ra}{2\sqrt{\pi}}\right)^{\frac{2}{3}}\nonumber \\
-Q&=&2\sqrt{\frac{L_x}{\pi u_0}}\nonumber
-\end{eqnarray}
-
-The experiment is performed with two grid resolution (\num{125x40} and \num{200x80} elements), with 100 markers per element and CFL of 0.25. $v_{\textrm{rms}}$
-(see Sec. \ref{sec:error}) calculated for both the simulations match well with results from \citet{vanKeken1997} and obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and ELEFANT \citep{Thieulot2014}, obtained with similar grid resolutions
-(Fig. \ref{fig:thin}). In addition, a variety of simulations with different resolutions and aspect ratios are performed to verify that the sensitivity of the
-initial velocity field. Results match well with results from \citet{vanKeken1997} (in the grey area) and \citet{Thieulot2014} (Fig. \ref{fig:thin_initial}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Thin_layer}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_thin.pdf}
-\caption{$v_{\textrm{rms}}$ for the thin layer experiment as function of time for different grid resolution. Results are compared with results obtained by
-\citet{vanKeken1997} (black lines) and with ASPECT and ELEFANT.}
-\label{fig:thin}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrmszero.pdf}
-\caption{$v_{\textrm{rms}}$ at $t=0$ for the thin layer experiment as function of the element size for different aspect ratios of the elements. Results are compared
-with results obtained by \citet{vanKeken1997} (black dots in the grey area) and with ASPECT and ELEFANT.}
-\label{fig:thin_initial}
-\end{figure}
-
-\subsection{Sinking hydrated cylinder}\label{sec:quinquis}
-This experiment is performed as explained in \citet{Quinquis2014} and simulates the sinking of a cold, hydrated cylinder into a warm, dry mantle. The domain
-is square with $L_x=L_y=$ \SI{300}{\km}, a grid resolution of \num{300x300} elements and 25 markers per element. The sinking cylinder is located at $x=$
-\SI{150}{\km} and $y=$ \SI{170}{\km} with a radius of \SI{20}{\km} and has $\rho_c=$ \SI{3250}{\kg\per\cubic\m}, $\eta_c=$ \SI{e23}{\pascal\s}, a thermal
-conductivity $k_c=$ \SI{4.5}{\watt\per\m\per\kelvin},a specific heat $C_{p_c}=$ \SI{1250}{\joule\per\kg\per\kelvin}; the surrounding mantle has $\rho_m=$
-\SI{3200}{\kg\per\cubic\m}, $\eta_m=$ \SI{e20}{\pascal\s}, a thermal conductivity $k_m=$ \SI{105}{\watt\per\m\per\kelvin} and specific heat $C_{p_m}=$
-\SI{1250}{\joule\per\kg\per\kelvin}; a 58 km-thick lithosphere overlies the mantle and has $\rho_l=$ \SI{3200}{\kg\per\cubic\m}, $\eta_l=$ \SI{e23}{\pascal\s},
-a thermal conductivity $k_l=$ \SI{4.5}{\watt\per\m\per\kelvin}, specific heat $C_{p_l}=$ \SI{750}{\joule\per\kg\per\kelvin} and a thermal diffusivity of
-\SI{e-6}{\square\m\per\s}. Velocity boundary conditions are set to free slip on all sides of the domain. Temperature increase linearly in the lithosphere,
-from 0° to 1300°C, and in the mantle, from 1300° to 1360.5°C. The initial temperature of the cylinder is fixed to 400°C. Throughout the evolution, temperature
-is fixed to 0° and 1360.5°C at the top and bottom of the domain, respectively. The initial bound water content is imposed at 2 wt.\% in the cylinder and at
-0 wt.\% in both the mantle and the lithosphere. The maximum amount of water in the mantle is fixed to 0.2 wt.\%, while maximum water content in the cylinder
-and in the lithosphere is function of pressure and temperature and is calculated using a serpentinized harzburgite with Perple\_X, as in \citet{Quinquis2014}.
-
-As shown in \citet{Quinquis2014}, the model is characterised by a progressive dehydration in the external portion of the cylinder, due to the increase of
-temperature, with a consequent vertical migration of free water that hydrates the mantle above the cylinder, up to the lithosphere. The distribution of
-bound and free water during the evolution of the experiment is shown in Fig. \ref{fig:quinquis} (left and right columns, respectively). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Hydration}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Hydration.pdf}
-\caption{Distribution of bound and free water (left and right columns, respectively) at $t=0.5$ and $t=$ \SI{1.25}{\mega\year} (first and second rows,
-respectively) for the sinking hydrated cylinder experiment. Only markers with bound or free water not equal to 0 are plotted.}
-\label{fig:quinquis}
-\end{figure}
-
-\subsection{Mantle melting curves}\label{sec:katz}
-This instantaneous experiment is performed in order to recreate isobaric melting curves, as function of temperature, bulk water and content of water in the
-melt, to compare with melting curves obtained by \citet{Katz2003}. The domain is square with $L_x=L_y=100$ km, a grid resolution of \num{100x100} elements and
-25 markers per element. Density is set in the entire domain to \SI{3300}{\kg\per\cubic\m} and temperature increases from the sides to the centre from 1000° to
-1500°C. The experiment is performed for bulk water contents from 0 to 0.3 wt.\%.
-
-Fig. \ref{fig:fraction} shows isobaric melting curves for \SI{1}{\giga\pascal} (panel a) and \SI{3}{\giga\pascal} (panel b) calculated for different bulk water
-contents. The curves well match with those obtained for the mantle by \citet{Katz2003}, showing the saturation at the solidus for bulk water of 0.2 and 0.3 wt.\%, indicating
-that the code correctly determines melt fraction as function of pressure, temperature and bulk water content. Same results can be observed in Fig.
-\ref{fig:water}a, where isobaric and isothermal melting curves are shown as function of bulk water content. In addition, Fig. \ref{fig:water}b shows that the
-code correctly determines also the percentage of water in the melt as function of pressure, temperature and melt fraction. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Melting}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Fraction.pdf}
-\caption{Isobaric melting curves for 1 GPa (panel a) and 3 GPa (panel b) as function of temperature with different bulk water contents.}
-\label{fig:fraction}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Water.pdf}
-\caption{Isobaric melting curves for 1.5 GPa as function of bulk water (panel a) and water in the melt (panel b) calculated for different temperatures.}
-\label{fig:water}
-\end{figure}
-
-\clearpage
-
-\bibliography{export}
-
-\end{document}

### 2022-09-07 tex/Benchmark/subduction.tex
update main

--- 2022-03-31 FALCON_description.tex
@@
-\documentclass[hidelinks,10pt,a4paper]{article}
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{textcomp}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage{booktabs}
-\usepackage{multirow}
-\usepackage[cm]{fullpage}
-\usepackage[round]{natbib}
-\bibliographystyle{abbrvnat}
-\usepackage{doi}
-\usepackage{bm}
-\usepackage[table]{xcolor}
-\hypersetup{
-    colorlinks,
-    linkcolor={blue},
-    citecolor={blue},
-    urlcolor={blue}
-}
-\usepackage{float}
-\restylefloat{table}
-\usepackage{siunitx}
-\DeclareSIUnit\year{yr}
-
-\begin{document}
-
-\section{Introduction}\label{sec:intro}
-This document includes a description of the code FALCON, with a presentation of the features implemented followed by results of extensive benchmarks listed in
-Table \ref{tab:list}, performed to verify that the code solves correctly a wide range of problems. Description and results of all benchmarks are shown
-in Section \ref{sec:benchmark}.
-
-\begin{table}[H]
-\caption{List of benchmarks performed for each feature of the code. Data of all benchmarks can be found at 
-\url{https://github.com/aleregorda/Benchmarks}}
-\centering
-\begin{tabular}{| l | l | l |}
-\toprule
- \textbf{Feature}  & \textbf{Benchmark} & \textbf{Section} \\
-\midrule
-  Solver  & Stokes flow & \ref{sec:stokes} \\
-\hline
-  \multirow{2}{*}{Markers advection} & Zalesak disk & \ref{sec:runge} \\
-    & Conservative Interpolation Velocity & \ref{sec:CVI} \\
-\hline
-  \multirow{4}{*}{Momentum equation}  & Poiseuille flow & \ref{sec:poiseuille}\\
-    & Instantaneous 2D sphere & \ref{sec:ist_sphere} \\
-    & Rayleigh-Taylor experiment & \ref{sec:rayleigh} \\
-    & Falling block & \ref{sec:block} \\
-\hline
-    & 2D Stokes sphere & \ref{sec:sphere} \\
-    Sticky air and  & Stabilisation algorithm & \ref{sec:stab} \\
-    free surface & Topography relaxation & \ref{sec:crameri} \\
-  & Spontaneous subduction & \ref{sec:subduction} \\
-\hline
-  Erosion and sedimentation  & - & - \\
-\hline
-   Non-linear  & Slab detachment & \ref{sec:slab} \\
-   visco-plasticity & Indenter & \ref{sec:indenter} \\
-    & Brick & \ref{sec:brick} \\
-\hline
-  \multirow{3}{*}{Energy equation}  & Advection stabilisation & \ref{sec:advection} \\
-    & Simple shear heating & \ref{sec:simple_shear} \\
-    & Shear and adiabatic heating & \ref{sec:shear} \\
-\hline
-\multirow{3}{*}{Energy + momentum} & Mantle convection & \ref{sec:mantle} \\
-& Viscoplastic mantle convection & \ref{sec:visco_mantle} \\
-& Thin layer entrainment & \ref{sec:thin} \\
-\hline
-  Phase changes  & \multirow{2}{*}{Hydrated sinking cylinder} & \multirow{2}{*}{\ref{sec:quinquis}} \\
-  and hydration  &  & \\
-\hline
-  Melting & Experimental melting curves & \ref{sec:katz} \\
-\bottomrule
-\end{tabular}
-\label{tab:list}
-\end{table}
-
-The code is written in Fortran90 and uses the finite element method (FEM) with quadrilateral $Q_1 \times P_0$ elements (continuous bilinear velocity and
-discontinuous constant pressure), associated with MUMPS\footnote{\url{http://mumps.enseeiht.fr/}} \citep{Amestoy2001,Amestoy2019} that is a software package
-for solving systems of linear equations of the form $A \cdot x = b$, where \textit{A} is a square sparse matrix, by means of a direct method. Since
-$Q_1 \times P_0$ elements do not satisfy Ladyzhenskaya, Babuska and Brezzi (LBB) stability condition \citep{Donea2003}, elemental pressure is elaborated
-in post-processing to avoid spurious pressures by means of a double interpolation to smooth the pressure. Following this procedure, the elemental pressure
-is interpolated onto nodes and then back onto elements. After the smoothing procedure, the pressure field is calculated again on the nodes to be used in
-combination with the lithostatic pressure, which is calculated onto nodes. Correctness of the solution and performances of the solver in terms of time and
-memory usage are tested solving the Stokes flow with the analytical solution proposed by \citet{Donea2003} (Section \ref{sec:stokes}).
-
-The thermo-mechanics of crust-mantle systems is described by means of the conservation of mass, momentum and energy equations, expressed as follows:
-\begin{eqnarray}
-\label{eq:mass}\frac{\partial \rho}{\partial t}+\bm{\nabla} \cdot (\rho \bm{u})&=&0\\
-\label{eq:momentum}-\bm{\nabla} p + \bm{\nabla} \cdot \bm{\tau} + \rho \bm{g}&=&0\\
-\label{eq:energy}\rho C_p \left(\frac{\partial T}{\partial t} + \bm{u} \cdot \bm{\nabla} T \right)&=&\bm{\nabla} \cdot \left(k\bm{\nabla} T\right) + H_{tot}
-\end{eqnarray}
-where $\rho$ is the density, $\bm{u}$ is the velocity, \textit{p} is the pressure, $\bm{\tau}$ is the deviatoric stress, $\bm{g}$ is the gravity acceleration,
-$C_p$ is the specific heat at constant pressure, \textit{T} is the temperature, \textit{k} is the thermal conductivity and $H_{tot}$ is the total internal heat
-production. Density variations due to temperature are generally small enough to assume the density as constant ($\rho=\rho_0$) in Eq. \ref{eq:mass} and in
-Eq. \ref{eq:energy}, while it must be treated as a variable in the buoyancy term of Eq. \ref{eq:momentum}, such that \[\rho=\rho_0(1-\alpha(T-T_0))\]
-where $\rho_0$ is the density at a reference temperature $T_0$ and $\alpha$ is the coefficient of thermal expansion. Eq. \ref{eq:mass} then can be rewritten as
-\begin{equation}\label{eq:mass_0}
-\bm{\nabla} \cdot \bm{u}=0
-\end{equation}
-This simplification is known as the extended Boussinesq approximation.
-
-The time step $dt$ is chosen in according to Courant-Friedrichs-Lewy condition \citep{Anderson1995}, such that
-\[dt=C_n\; \min\left(\frac{\Delta x}{\max|\bm{u}|},\frac{(\Delta x)^2}{\kappa}\right)\]
-where $0<C_n<1$ is the Courant (CFL) number, $\Delta x$ is the minimum dimension of smallest element, $\max|\bm{u}|$ is the maximum velocity calculated on the
-entire domain and
-$\kappa=\frac{k}{\rho C_p}$ is the thermal diffusivity \citep{Thieulot2014}.
-
-\section{Lagrangian markers}\label{sec:markers}
-Elemental properties (density, viscosity, thermal conductivity, specific heat and thermal expansion) needed to solve Eqs. \ref{eq:mass_0}, \ref{eq:momentum}
-and \ref{eq:energy} are related to the composition of each element, determined by means of Lagrangian markers that are characteristic of different materials
-in the domain. Elemental properties (with the sole exception of the viscosity) are calculated using an arithmetic mean, as \[P_e=\frac{1}{n}\sum_{i=1}^n P_i\]
-where $P_e$ is the elemental property, $P_i$ is the property characteristic of the material of each marker and $n$ is the number of markers of the element.
-Differently, the average scheme for the elemental viscosity can be chosen between harmonic, geometric and arithmetic mean. Markers can be placed regularly or
-randomly at the beginning of the simulation and their advection is performed by either a 2nd-order or a 4th-order Runge-Kutta in space, interpolating the velocity 
-field on each marker by means of the shape functions (see Section \ref{sec:runge}). The interpolated velocity is then corrected by means of the Conservative
-Velocity Interpolation (CVI), which introduces a corrective term to reduce dispersion and clustering of particles in both steady state and time-dependent 
-\citep{Wang2015} (see Section \ref{sec:CVI}).
-
-The initial distribution of the markers is created using the open-source code library
-Geodynamic World Builder \citep{Fraters2019}. During the simulation, each marker carries memory of temperature, pressure and accumulated strain, which are
-determined interpolating the relative nodal parameter. In particular, pressure is interpolated onto each marker after the smoothing procedure. The number
-of markers contained in each element is maintained between $n_{min}$ and $n_{max}$. When in an element there are less markers than $n_{min}$ the code adds
-random markers to reach the $n_{min}$, while if the number is higher than $n_{max}$ some of them are randomly deleted. When new markers are added, they assume
-the properties of the nearest marker. In this way elements are never empty and maintain a number of markers inside a prefixed range.
-
-\section{The momentum equation}\label{sec:mom_eq}
-The penalty method is implemented so that Eq. \ref{eq:mass_0} can be written as:
-\begin{equation}\label{eq:mass_penalty}
-\bm{\nabla} \cdot \bm{u} + \frac{p}{\lambda}=0
-\end{equation}
-where $\lambda$ is the so-called penalty parameter, which should be 6-7 orders of magnitude larger than the shear viscosity to ensure that mass conservation
-is satisfied \citep{Donea2003,Thieulot2014}.
-
-The deviatoric stress tensor in Eq. \ref{eq:momentum} can be written in terms of the strain rate tensor as $\bm{\tau}=2\eta\dot{\bm{\epsilon}}$, with
-$\dot{\bm{\epsilon}}=\frac{1}{2}\left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)$. Therefore, Eq. \ref{eq:momentum} can be rewritten as:
-\begin{equation}\label{eq:momentum_pressure}
--\bm{\nabla} p + \bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right) + \rho \bm{g} = 0
-\end{equation}
-Finally, using pressure from Eq. \ref{eq:mass_penalty}, Eq. \ref{eq:momentum_pressure} can be rewritten as:
-\begin{equation}\label{eq:momentum_penalty}
-\lambda \bm{\nabla} \left(\bm{\nabla} \cdot \bm{v}\right)+\bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right)+\rho\bm{g}=0
-\end{equation}
-The penalty method is been associated to the iterative Uzawa method, as described in detail in \citet{Dabrowski2008} and \citet{Thieulot2014}
-(see Section \ref{sec:poiseuille}). Results of benchmarks performed to verify the correctness in the implementation of Eq. \ref{eq:momentum_penalty} are
-shown in Sections \ref{sec:ist_sphere} and \ref{sec:rayleigh}.
-
-To support large viscosities variations the penalty parameter is related to the effective elemental viscosity by means of a dimensionless coefficient, so that
-$\lambda=\lambda_e(e)\eta_{eff}(e)$ \citep{Marotta2006,Dabrowski2008,Thieulot2014}. Benchmark of the falling block \citep{Gerya2003a,Gerya2010b,Thieulot2011}
-is performed to verify that the code can correctly deal buoyancy driven flows with strong viscosity contrasts (Section \ref{sec:block}).
-
-\section{Sticky air and free surface}\label{sec:sticky}
-The Earth's surface can be treated by means of either the so-called sticky-air or a true free surface method, both of them implemented in the code. In the
-sticky-air method the surface is approximated with the introduction of a buoyant layer with a viscosity at least four orders of magnitude lower than the crust
-\citep{Schmeling2008,Crameri2012} and the interface between lithosphere and air is defined using a chain of passive markers that are advected as the Lagrangian
-markers. The correctness of the evolution of the markers chain is tested with the experiment of a 2D time-dependent Stokes sphere below a free surface and
-compared with results from ASPECT\footnote{\url{https://aspect.geodynamics.org/}} \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} (Section
-\ref{sec:sphere}). In the true free surface case the top boundary is assumed stress-free and velocities are not fixed. In this case topography variations are
-described by vertical deformations of the mesh that depend on the velocity field of the nodes that identify the free surface, while horizontal deformations are
-not taken into account. This procedure is known as the Arbitrary Lagrangian-Eulerian (ALE) method and its implementation follows the technique described in
-\citet{Thieulot2011}. However, although the implementation of a true free surface better reproduces laboratory experiment, extremely small time steps can be
-necessary to maintain stability \citep{Kaus2010a,Quinquis2011,Thieulot2014}. Therefore, the stability algorithm proposed by \citet{Kaus2010a} is implemented
-to avoid instabilities due to high density differences at the free surface when using too large time steps. The implementation of this algorithm are tested by
-performing the experiment described by \citet{Kaus2010a} (Section \ref{sec:stab}).
-
-The topography relaxation benchmark proposed by \citet{Crameri2012} is performed to verify that the code correctly recovers topography variations in case of
-both the sticky-air and the true free surface method (Section \ref{sec:crameri}). Finally, Section \ref{sec:subduction} show the results of the spontaneous
-subduction experiment described by \citet{Schmeling2008}.
-
-\section{Erosion and sedimentation}\label{sec:erosion}
-Surface processes at either the lithosphere-air interface or in correspondence of the free surface have been implemented by means of the software
-Fastscape\footnote{\url{https://fastscape.org/}} \citep{Braun2013,Cordonnier2019,Yuan2019a,Yuan2019b}. FastScapeLib is a set of routines that solve the stream
-power law (enriched by a sediment transport/deposition term) \citep{Yuan2019a}, hillslope diffusion and marine transport and deposition \citep{Yuan2019b}.
-The partial differential equation solved by FastScapeLib is:
-\begin{equation}
-\frac{\partial h}{ \partial t}=U-K_f A^{m_e}S^{n_e} + \frac{G}{A}\int_A \left( U-\frac{\partial h}{\partial t} \right) dA + K_d \bm{\nabla}^2 h 
-\end{equation}
-where $h$ is the topography, $U$ is the uplift, $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $S$ is the slope, $m_e$ and $n_e$ are
-parameters that should be chosen to constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}, $G$ is the dimensionless
-deposition/transport coefficient for the enriched stream power law and $K_d$ is the effective diffusivity. Terms $K_f A^{m_e}S^{n_e}$ and $K_d \bm{\nabla}^2 h$
-consider long-range processes, described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, and short-range processes,
-proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}.
-
-In case of sticky air, slope is determined by means of the markers chain, which is vertically corrected considering erosion and sedimentation rates. After the
-correction, continental markers above the markers chain are transformed in air markers and air markers below the markers chain become sediments. Similarly,
-slope is determined considering the nodes on the top boundary in case of true free surface and they are vertically corrected as for the markers chain. In this
-case, continental markers above the top boundary are simply deleted.
-%Surface processes at the lithosphere-air interface are implemented taking into account fluvial erosion (long-range processes), hillslope processes (short-range
-%processes) and failure of slopes exceeding a threshold angle $\theta$ \citep{Burov1997,Tucker1999,Sacek2012,Croissant2014,Ueda2015,Beucher2020}. Erosion rates
-%$\dot{E}_{lr}$ due to long-range processes can be described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{lr}=K_f A^{m_e} \bm{\nabla} h^{n_e}
-%\end{equation}
-%where $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $\bm{\nabla} h$ is the slope and $m_e$ and $n_e$ are parameters chosen in to
-%constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}. Erosion rates $\dot{E}_{sr}$ due to short-range processes are
-%implemented by assuming that transport is linearly proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{sr}=K_d \bm{\nabla}^2 h
-%\end{equation}
-%where $K_d$ is the effective diffusivity. In addition, the slope is modified if it still exceeds a threshold angle after long- and short-range processes
-%\citep{Ueda2015}.
-
-%Offshore sedimentation processes are implemented similarly to hillslope processes but with a depth-dependent coefficient \citep{Sacek2012,Beucher2020}.
-%The sedimentation rate $\dot{S}$ is calculated as
-%\begin{equation}
-%\dot{S}=K_s \bm{\nabla}^2 h
-%\end{equation}
-%with
-%\begin{equation}
-%K_s=K^*_s \exp\left(\frac{h-h_{sl}}{h_0}\right)
-%\end{equation}
-%where $K^*_s$ is a constant, $h_{sl}$ is the sea level and $h_0$ is a characteristic length. In case of sticky air, slope is determined by means of the markers
-%chain, which is vertically corrected considering erosion and sedimentation rates. After the correction, continental markers above the markers chain are
-%transformed in air markers and air markers below the markers chain become sediments. Similarly, slope is determined considering the nodes on the top boundary
-%in case of true free surface and they are vertically corrected as for the markers chain. In this case, continental markers above the top boundary are simply
-%deleted.
-
-\section{Non-linear rheologies}\label{sec:nonlinear}
-Non-linear rheologies are implemented combining viscous creep (dislocation and diffusion) and plastic yielding. For each marker, diffusion and dislocation
-viscosity ($\eta_{df|ds}$) can be determined as follows \citep{Karato1993,Warren2008b,Wang2016}:
-\begin{equation}\label{eq:viscous}
-\eta_{df|ds}=f_s \left(\frac{d^m}{A}\right)^{\frac{1}{n}} \left (\frac{2^{\frac{1-n}{n}}}{3^{\frac{n+1}{2n}}} \right)
-I_2^{\frac{(1-n)}{n}}\exp\left(\frac{Q+pV}{nRT}\right)
-\end{equation}
-where $f_s$ is a scaling factor used to represent lithologies that are stronger or weaker than the base set, $d$ is the grain size, $m$ is the grain size exponent,
-$n$ is the stress exponent, $A$ is the uniaxial pre-exponential factor, $I_2$ is the square root of the second invariant of the strain rate tensor, $Q$ is the
-activation energy, $p$ is the pressure, $V$ is the activation volume, $R$ is the gas constant and $T$ is the temperature.
-Pressures and temperatures are determined by the interpolation of the nodal parameters, while strain rates are calculated by means of the
-derivative of the velocities on the nodes. In case of diffusion creep $n=1$ and $m>0$, in case of dislocation creep $n>1$ and $m=0$. The viscous creep $\eta_{cp}$
-is then calculated as the harmonic average between $\eta_{df}$ and $\eta_{ds}$:
-\begin{equation}\label{eq:visco_creep}
-\eta_{cp}=\left(\frac{1}{\eta_{df}}+\frac{1}{\eta_{ds}}\right)^{-1}
-\end{equation}
-The implementation of the non-linear viscous creep viscosity depending on the strain rate is tested by means of the slab detachment benchmark
-(Section \ref{sec:slab}).
-
-Plastic yielding is implemented rescaling $\eta_{cp}$ in order to limit the stress below the yield stress $\sigma_y$ \citep{Thieulot2008,Thieulot2014,Glerum2018},
-obtaining
-\begin{equation}\label{eq:plastic}
-\eta_{pl}=\frac{\sigma_y}{2I_2}
-\end{equation}
-where $\eta_{pl}$ is the plastic viscosity and the yield stress is determined following the Drucker-Prager criterion, such as
-\begin{equation}\label{eq:yield}
-\sigma_y=C\;\cos(\phi)+p\;\sin(\phi)
-\end{equation}
-where $C$ is the cohesion and $\phi$ is the internal friction angle. In case of negative pressure it is imposed equal to 0, so that negative yield stress are
-excluded. The correctness of the non-linear solution in case of plastic viscosity with variable internal friction angle is verified performing indenter and
-brick experiments (Sections \ref{sec:indenter} and \ref{sec:brick}, respectively).
-
-Strain softening is taken into account for both viscous creep and plastic viscosity \citep{Huismans2003,Babeyko2005,Huismans2005,Sobolev2005,Warren2008a} by
-means of the accumulated strain $\epsilon$ memorised by each marker. Softening in the viscous creep determines a linear decrease of $\eta_{vc}$ by means of a
-viscous strain softening factor $W_S$ that increases linearly from $W_{S_0}$ to $W_{S_{\infty}}$ for $\epsilon_{S_0}<\epsilon_S<\epsilon_{S_{\infty}}$ 
-\citep{Warren2008a}. This viscous softening can be related to strain-induced grain size reduction \citep{Warren2008a}. Differently, plastic softening is 
-simulated with a linear decrease of internal friction angle $\phi(\epsilon)$ and cohesion $C(\epsilon)$ in according to
-\begin{eqnarray}
-\label{eq:friction1}\phi(\epsilon)&=&\phi_0+(\phi_{\infty}-\phi_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}\\
-\label{eq:friction2}C(\epsilon)&=&C_0+(C_{\infty}-C_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}
-\end{eqnarray}
-where $\phi_0$, $C_0$ and $\phi_{\infty}$, $C_{\infty}$ are internal friction angle and cohesion for $\epsilon_0$ and $\epsilon_{\infty}$, respectively
-\citep{Huismans2003,Huismans2005,Warren2008a,Thieulot2014}. Plastic softening approximates deformation-induced softening of faults and brittle shear zones
-\citep{Warren2008a}.
-
-Plastic yielding and viscous creep are then combined to obtain a viscoplastic viscosity $\eta_{vp}$ as follows, assuming that they are independent processes
-\citep{Karato2008,Glerum2018}:
-\begin{equation}\label{eq:viscoplastic}
-\eta_{vp}=\min(\eta_{cp},\eta_{pl})
-\end{equation}
-Finally, effective viscosity $\eta_{eff}$ is capped by the minimum and the maximum viscosity ($\eta_{min}$ and $\eta_{max}$, respectively) to avoid extremely
-low or high viscosity \citep{Glerum2018} as follows
-\begin{equation}\label{eq:effective}
-\eta_{eff}=\min(\max(\eta_{vp},\eta_{min}),\eta_{max})
-\end{equation}
-Viscous creep and plastic yielding are non-linear rheologies because of their dependence on the velocity field through pressure and strain rates. Therefore,
-the solution is determined by means of Picard-type iterations, until convergence of the velocity field \citep{Glerum2018}. The convergence is verified at each
-iteration $i$ via the nonlinear residual $\bm{\mathcal{R}}^i$ that can be determined as
-\begin{equation}\label{eq:convergence}
-\bm{\mathcal{R}}^i= \mathbb{K} (\eta_{eff}(\dot{\bm{\epsilon}}^{i-1},p^{i-1}))\cdot \bm{v}^{i-1} - \bm{f}^i
-\end{equation}
-%\begin{equation}\label{eq:convergence}
-%\chi=1-\frac{\langle (\bm{v}^i-\langle\bm{v}^i\rangle)\cdot (\bm{v}^{i+1}-\langle\bm{v}^{i+1}\rangle) \rangle}{\sqrt{\sigma^i\sigma^{i+1}}}
-%\end{equation}
-%Iterations are performed until either convergence ($\chi<tol$) or maximal number of non-linear iterations ($it_{max}$) is reached.
-where $\mathbb{K}$ is the Stiffness matrix and $\bm{f}$ is the right hand side vector \citep{Spiegelman2016,Glerum2018}.  The $L_2$-norm is extracted from
-$\bm{\mathcal{R}}^i$ and it is normalised as follows
-\begin{equation}\label{eq:convergence_res}
-\frac{||\bm{\mathcal{R}}^i||_2}{||\bm{\mathcal{R}}^0||_2}
-\end{equation}
-where $||\bm{\mathcal{R}}^0||_2$ is the $L_2$-norm of the first nonlinear iteration.
-Since often the normalised nonlinear residual drop very quickly over the first 2-3 iterations, also the $L_2$-norm of velocity and pressure residuals are
-calculated as
-\begin{equation}\label{eq:convergence_vel}
-\frac{||\Delta\bm{v}||_2}{||\bm{v}||_2}=\frac{||\bm{v}_i-\bm{v}_{i-1}||_2}{||\bm{v}_{i}||_2}
-\end{equation}
-and
-\begin{equation}\label{eq:convergence_press}
-  \frac{||\Delta p||_2}{||p||_2}=\frac{||p_i-p_{i-1}||_2}{||p_{i}||_2}
-\end{equation}
-respectively. Iterations are performed until either maximal number of non-linear iterations ($it_{max}$) is reached or all the normalised nonlinear residual, 
-the velocity residual and the pressure residual converge under a defined tolerance $tol$.
-
-\section{The energy equation}\label{sec:en_eq}
-The stabilisation of the advection term of the energy equation (Eq. \ref{eq:energy}) needed to avoid possible oscillations in the thermal solution in the case
-when advection dominates over diffusion, is implemented by means of a streamline-upwind Petrov–Galerkin (SUPG) method, for which the advection term is modified
-follows the discussion in \citet{Thieulot2011} and \citet{Thieulot2014} (see Section \ref{sec:advection} for benchmark). The total internal heat production
-$H_{tot}$ in Eq. \ref{eq:energy} includes radiogenic heating $H_r$, shear heating $H_s$ and adiabatic heating $H_a$, where
-\begin{equation}\label{eq:shear}
-H_s=2\eta \dot{\bm{\epsilon}} : \dot{\bm{\epsilon}} = 2\eta(\dot{\epsilon}_{xx}^2+\dot{\epsilon}_{yy}^2+2\dot{\epsilon}_{xy}^2)
-\end{equation}
-and
-\begin{equation}\label{eq:adiabatic}
-H_a=T \alpha \frac{Dp}{Dt} \approx -\alpha T \rho g_y v
-\end{equation}
-The correctness of the implementation and the computation of shear and adiabatic heating is verified performing an exercise with an analytical solution
-(Section \ref{sec:simple_shear}) and Exercise 9.4 in \citet{Gerya2010b} (Section \ref{sec:shear}). Furthermore, mantle convection and thin layer entrainment
-benchmarks (Sections \ref{sec:mantle} and \ref{sec:thin}, respectively) are performed to verify that the code solve correctly Eqs. \ref{eq:momentum_penalty}
-and \ref{eq:energy} in case of a temperature field characterised by an initial perturbation and with a temperature-dependent density. At last, the viscoplastic
-mantle convection benchmark is performed to check also the viscoplasticity in case of a initial perturbed temperature field (Section \ref{sec:visco_mantle}).
-
-\section{Phase transitions and hydration}\label{sec:phase}
-Variations of effective density, specific heat and coefficient of thermal expansion during the evolution of crustal and mantle materials at different
-pressure-temperature ($p$-$T$) conditions are computed by means of version 6.8.6 of Perple\_X software package \citep{Connolly2005}, similarly to the
-implementation described in \citet{Marotta2020}. By default, mineral assemblages and properties of each lithology (oceanic crust, lower and upper continental
-crust, sediments and dry or hydrated mantle) are calculated for temperatures from \SIrange{330}{1600}{\kelvin}, with increments of approximately \SI{4}{\kelvin},
-and pressures from \SIrange{0.1}{30}{\giga\pascal}, with increments of approximately \SI{0.03}{\giga\pascal}, for a total of almost 100000 points.
-
-The introduction of phase transitions produces two main effects that must be taken into account in numerical models: variations in density and release/absorption
-of latent heat \citep{Tackley2010}. Effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into
-account considering effective coefficient of thermal expansion \citep{Christensen1985,Zhong2015}. Similarly, effects of latent heat in the energy equation
-(eq \ref{eq:energy}) are taken into account by considering effective specific heat and coefficient of thermal expansion
-\citep{Christensen1985,Tackley2010,Zhong2015}. Effective density, specific heat and coefficient of thermal expansion are included in Perple\_X files and are
-assumed by each Lagrangian marker in according to its $p$-$T$ conditions.
-
-Hydration processes related to dehydration of subducting lithosphere \citep{Schmidt1998,Liu2007,Faccenda2009,Faccenda2010,Faccenda2014,Rosas2016} strongly
-influence the thermo-mechanics inside the mantle wedge, mainly because of both weakening effects on the mantle rheology and density variations in case of
-mantle serpentinization \citep{Gerya2002,Honda2003,Arcay2005,Roda2010,Regorda2017}. In the case that hydration is switched on, the amount of bound and free
-water is memorised by each marker, following the implementation of \citet{Quinquis2014}; hydration and dehydration processes are related to the amount of bound
-water of each marker and to the maximum amount of water it can transport, i.e. if the amount of bound water exceeds the maximum amount of water, the marker
-dehydrates and releases free water that can hydrates under-saturated markers. Maximum water content of each marker is determined as function of lithology and
-$p$-$T$ conditions and it is calculated using Perple\_X, in the same way than effective density, specific heat and coefficient of thermal expansion. Bound water
-is advected together with the markers, neglecting the effect of bound water diffusion, while free water simply migrates vertically and is not coupled to the
-solid-phase flow of the mantle wedge \citep{Arcay2005,Quinquis2014}. The correctness of the migration scheme of the free water and its absorption into bound
-water by under-saturated markers are verified performing the experiment described in \citet{Quinquis2014} (Sec. \ref{sec:quinquis}).
-
-The mantle viscosity weakening related to the amount of water has been extensively studied \citep{Chopra1981,Hirth2003} and it is implemented in according to
-\citet{Arcay2005} and \citet{Horiuchi2016}, as follows
-\begin{equation}\label{eq:wet_visc}
-\eta_{wet}=\eta_{dry}\left(\left[1-\frac{1}{f_v}\right]\exp\left(-\frac{[OH^-]}{[OH^-]_0}\right)+\frac{1}{f_v}\right)
-\end{equation}
-where $f_v$ is the viscosity reduction factor between dry and wet conditions, $[OH^-]$ is the water content and $[OH^-]_0$ is a reference water content set to
-620 ppm (0.062 wt.\%) \citep{Arcay2005}. By default, $f_v$ is set to 100, in according to \citet{Arcay2005} and \citet{Horiuchi2016}.
-
-\section{Mantle melting}\label{sec:melt}
-Melt fraction in the mantle depends on temperature ($T$, in \textdegree C), pressure ($p$, in \si{\giga\pascal}) and the water content in the melt ($X_{H_2O}$, in wt.\%).
-The determination of the melt fraction $M$ is implemented as explained by \citet{Katz2003}, according to the successive modification by \citet{Langmuir2006} and
-\citet{Kelley2010}, as follows
-\begin{eqnarray}\label{eq:melt_fr}
-M(p,T,X_{H_2O})&=&-T+T_{sol}+(x_t\cdot ln(p+y_t))\cdot X_{H_2O}-\nonumber\\
-&&+K_m\left(\frac{X_{H_2O}^{bulk}}{D_{H_2O}(1-X_{H_2O})+X_{H_2O}}\right)^{\gamma}
-\end{eqnarray}
-where $T_{sol}$ is the temperature of the dry solidus, $D_{H_2O}$ is the partition coefficient and $x_t\cdot log(p+y_t)$ indicates the pressure dependence of
-the melting curve \citep{Kelley2010}. $x_t$, $y_t$, $K_m$ and $\gamma$ are parameters chosen in according to \citet{Kelley2010}.
-However, $X_{H_2O}$ depends on the melt fraction as
-\begin{equation}\label{eq:melt_wt}
-X_{H_2O}(M)=\frac{X_{H_2O}^{bulk}}{D_{H_2O}+M(1-D_{H_2O})}
-\end{equation}
-and a numerical solution can be found using a root-finder method \citep{Katz2003,Wang2016}. Furthermore, the content of water in the melt is limited by the
-pressure-dependent saturation concentration of water in the melt, determined as
-\begin{equation}\label{eq:melt_max}
-X_{H_2O}^{sat}=\chi_1 p^{\lambda_m}+\chi_2 p
-\end{equation}
-where $\chi_1$, $\chi_2$ and $\lambda_m$ are parameters chosen according to \citet{Katz2003}, \citet{Langmuir2006} and \citet{Kelley2010}. The correctness in
-the determination of melt fractions and water contents in the melt are tested comparing model results with experimental curves obtained by \citet{Katz2003}
-(Sec. \ref{sec:katz}).
-
-In case of melting, viscous creep viscosity ($\eta_{vp}$) is modified according to \citet{Wang2016}, as
-\begin{equation}\label{eq:melt_visc}
-\eta_{melt}=\eta_{vp}\cdot \exp(\alpha_m M)
-\end{equation}
-where $\alpha_m$ are melt fraction factors for dislocation and diffusion creep, chosen in according to \citet{Wang2016}.
-
-Effective density ($\rho_{eff}$) for partially molten rocks is calculated as
-\begin{equation}\label{eq:melt_dens}
-\rho_{eff}=\rho_s(1-M)+\rho_m M
-\end{equation}
-where $\rho_s$ and $\rho_m$ are the densities of the solid and the molten rock, respectively \citep{Gerya2004b,Gerya2007,Wang2016}. The density of the
-solid rock is extracted by Perple\_X, while the density of the molten portion is calculated as
-\begin{equation}\label{eq:melt_dens_M}
-\rho_m(p,T)=\rho_0[1-\alpha(T-T_0)][1+\beta(p-p_0)]
-\end{equation}
-where $\rho_0$ is the density at the reference temperature ($T_0$) and pressure($p_0$), and $\alpha$ and $\beta$ are the thermal expansion and compressibility
-coefficients, respectively \citep{Gerya2004b,Gerya2007,Wang2016}.
-
-As for the phase transitions, effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into account
-considering effective coefficient of thermal expansion. Similarly, latent heat due to melting/crystallisation can be implicitly included in the energy equation
-(Eq. \ref{eq:energy}) by considering effective specific heat ($C_{p_{eff}}$) and thermal expansion ($\alpha_{eff}$) of the partially molten rocks, in according to
-\begin{eqnarray}
-\label{eq:cp_effective}C_{p_{eff}}&=&C_p+H_L\left(\frac{\partial M}{\partial T}\right)_p \\
-\label{eq:alpha_effective}\alpha_{eff}&=&\alpha+\rho\frac{H_L}{T}\left(\frac{\partial M}{\partial p}\right)_T
-\end{eqnarray}
-where $C_p$ and $\alpha$ are the specific heat and the coefficient of thermal expansion of the solid rock, respectively, and $H_L$ is the latent heat of the
-molten rock \citep{Gerya2004b,Gerya2007,Tackley2010}.
-
-\section{Benchmarks}\label{sec:benchmark}
-\subsection{Error measurements}\label{sec:error}
-In order to determine the accuracy of velocity and pressure field of the benchmarks, the $L_2$-norm is computed by numerical integration on the Gauss-Legendre
-quadrature points. $L_2$-norm for pressure and velocity errors can be evaluated as
-\begin{eqnarray}
-\textrm{err}_p&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|p^n_i(r_q)-p^a_i(r_q)|^2\right)w_q|J_q|}\nonumber \\
-\textrm{err}_v&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|u^n_i(r_q)-u^a_i(r_q)|^2+|v^n_i(r_q)-v^a_i(r_q)|^2\right)w_q|J_q|}\nonumber
-\end{eqnarray}
-respectively, where $n_e$ is the number of elements, $n_q$ is the number of quadrature points per element, $p^n_i(r_q)$ and $p^a_i(r_q)$ are the numerical and
-analytical pressure, respectively, in each quadrature point $q$, $w_q$  and $J_q$ are the weight and the Jacobian at the quadrature point
-$q$, $u^n_i(r_q)$, $v^n_i(r_q)$, $u^a_i(r_q)$ and $v^a_i(r_q)$ are the numerical and analytical velocities, respectively, in each quadrature point $q$.
-
-Other quantities used as comparison with original benchmarks are the root-mean-square velocity over the whole domain and over the surface
-\begin{eqnarray}
-v_{\textrm{rms}}&=&\sqrt{\int_{\Omega}|\bm{v}|^2 d\Omega}\nonumber \\
-v_{\textrm{rms}}^{top}&=&\sqrt{\int_0^1 u^2 \bigg|_{y=1} dx}\nonumber
-\end{eqnarray}
-respectively, top and bottom Nusselt numbers
-\[\textrm{Nu}^{top/bottom}=-\int_0^1\frac{\partial T}{\partial y} \bigg|_{y=1/y=0} dx\]
-the average rate of work done against gravity
-\[\left\langle W \right\rangle=\int_{\Omega}T u_y d\Omega\]
-the average rate of viscous dissipation
-\[\left\langle \Phi \right\rangle=\int_{\Omega}\tau_{ij}\dot{\epsilon}_{ij} d\Omega\]
-and the percentage error between $\left\langle W \right\rangle$ and $\left\langle \Phi \right\rangle/Ra$
-\[\delta=\frac{\left|\left\langle W \right\rangle -\frac{\left\langle \Phi \right\rangle}{Ra} \right|}{\max \left(\left\langle W \right\rangle,
-\frac{\left\langle \Phi \right\rangle}{Ra}\right)}\times100\]
-
-\subsection{Stokes flow}\label{sec:stokes}
-The problem consists of determining velocity field $(u,v)$ and pressure $p$ in case of a manufactured solution with prescribed body forces such as
-\begin{eqnarray}
-b_1&=&(12-24y)x^4+(-24+48y)x^3+(-48y+72y^2-48y^3+12)x^2+\nonumber \\
-&&+(-2+24y-72y^2+48y^3)x+1-4y+12y^2-8y^3\nonumber \\
-b_2&=&(8-48y+48y^2)x^3+(-12+72y-72y^2)x^2+(4-24y+48y^2-48y^3+\nonumber \\
-&&+24y^4)x-12y^2+24y^3-12y^4\nonumber
-\end{eqnarray}
-for which the exact solution is:
-\begin{eqnarray}
-u(x,y)&=&x^2(1-x)^2(2y-6y^2+4y^3)\nonumber \\
-v(x,y)&=&-y^2(1-y)^2(2x-6x^2+4x^3)\nonumber \\
-p(x,y)&=&x(1-x)-\frac{1}{6}\nonumber
-\end{eqnarray}
-The domain is a unit square a constant viscosity ($\eta=1$) and the penalty parameter is set to $\lambda=10^7$. Velocity boundary conditions are set to no slip
-$(\bm{v}=\bm{0})$ on all boundaries. The problem is performed for different grid resolution between \num{8x8} and \num{1024x1024} elements.
-
-The errors between the analytical solution and the numerical prediction of the velocity and the pressure field are calculated as in Section \ref{sec:error}.
-Fig. \ref{fig:errors}a shows that both velocity and pressure field converge to the exact solution with the decrease of the element size, following the
-theoretical convergence rate. The convergence of pressure, in contrast with the observation by \citet{Donea2003} for $Q_1 \times P_0$ elements, indicate the
-effectiveness of the smoothing procedure. Smoothed pressure field for grid resolution of \num{128x128} elements is shown in Fig. \ref{fig:errors}b (continuous
-red line), in comparison with the analytical solution (dashed blue line). Solve times and memory usage needed to generate the solution are shown in
-Fig. \ref{fig:MUMPS}. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Solver/Stokes_Flow}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/errors.pdf}
-\caption{Velocity and pressure error for the Stokes flow experiment between generated and analytical solution as a function of element size (panel a),
-and comparison between smoothed pressure and analytical solution as function of $x$ coordinate for a grid resolution of \num{128x128} elements (panel b).}
-\label{fig:errors}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/MUMPS.pdf}
-\caption{Panel a) Analysis, factorisation, solution and total solve times (panel a) and factorisation memory usage (panel b) as a function of the total number
-of degrees of freedom for the Stokes flow experiment.}
-\label{fig:MUMPS}
-\end{figure}
-
-\subsection{Markers advection}\label{sec:runge}
-The 2nd-order Runge-Kutta advection scheme in space is tested by means of the Zalesak disk test \citep{Zalesak1979,Thieulot2014}. The benchmark is performed
-in a unit square domain with a grid resolution of \num{32x32} elements and 1000000 markers and values of Courant number between 0.25 and 1.
-At $t=0$, the disk is centred at position $(0.5;0.75)$ with a radius $R=0.15$ and has a vertical fissure 0.05 wide and 0.2 high (orange in Fig. \ref{fig:runge})
-and with the velocity field prescribed in the entire domain as
-\begin{eqnarray}
-u(x,y)&=&2\pi\left(y-\frac{L_x}{2}\right)\nonumber \\
-v(x,y)&=&-2\pi\left(x-\frac{L_x}{2}\right)\nonumber
-\end{eqnarray}
-the markers are back to their initial location after a $2\pi$ rotation ($t=1$). The clockwise rotation of the disk at $t=0.25$, 0.5 and 0.75 
-is shown in Fig. \ref{fig:runge} (red, blue and green, respectively). The distance from the centre of 1 marker is calculated throughout the rotation to
-evaluate the error for different values of Courant number. As expected, the error increase with the increasing of the time step, as shown in 
-Fig. \ref{fig:runge_err}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/Zalesak_Disk}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=200px]{./Figures/Zalesak.pdf}
-\caption{Position of the Zalesak disk at $t=0$, 0.25, 0.5 and 0.75 (orange, red, blue and green, respectively) for the case with Courant number of 1.}
-\label{fig:runge}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/zal_marker.pdf}
-\caption{Distance from the centre as function of time for values of Courant number of  of 0.25, 0.3, 0.5, 0.75 and 1 (green, orange, blue, black and red,
-respectively). R indicates the distance from the centre considering a perfect circle trajectory.}
-\label{fig:runge_err}
-\end{figure}
-
-\subsection{Conservative Velocity Interpolation (CVI)}\label{sec:CVI}
-The CVI correction is checked by means of the Stokes flow experiment by \citet{Donea2003}, as presented in \ref{sec:stokes}. The advection of Lagrangian
-markers is performed using either a 2nd-order or a 4th-order Runge-Kutta scheme with CFL=0.5 and an initial random distribution of 25 markers per element.
-Fig. \ref{fig:CVI} shows the minimum and maximum values of markers per element throughout the experiment with and without the CVI correction and using
-different Runge-Kutta schemes. The introduction of the CVI correction allows the
-tests to run for times longer than $t=2000$ without show empty elements, while tests without the CVI correction stop at approximately $t=600$, when at least
-one element goes empty. Fig. \ref{fig:CVI_time} shows the different number of markers per element and the markers distribution at $t=600$ obtained without and
-with the CVI correction (panels b and c, respectively) in case of a 2nd-order Runge-Kutta scheme. All data can be found at
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/CVI}.
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/CVI.pdf}
-\caption{Maximum and minimum number of markers per element throughout the experiment, with (green and orange lines) or without (purple and blue lines)
-the CVI correction and using either a 2nd-order (blue and orange lines) or a 4th-order (purple and green lines) Runge-Kutta scheme.}
-\label{fig:CVI}
-\end{figure}
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/CVI_time.pdf}
-\caption{Number of markers per element and markers distribution through time using 2nd.order Runge-Kutta scheme.
-Panels a: number of markers per element (panel a1) and distribution of the markers (panel a2) after the first
-iteration; panels b and c: comparison between number of markers per element (panels b1 and c1) and markers distribution (panels b2 and c2) without and with
-the CVI correction at $t=600$; panels d: number of markers per element (panel d1) and markers distribution (panel d2) in case of CVI correction at $t=2000$.}
-\label{fig:CVI_time}
-\end{figure}
-
-\subsection{Poiseuille flow}\label{sec:poiseuille}
-The domain is a rectangle with $L_x=2$ and $L_y=1$ and constant density and viscosity ($\rho=1$ and $\eta=1$), gravity acceleration $\bm{g}=\bm{0}$ and penalty
-parameter $\lambda=10^8$. The grid is composed by \num{40x20} elements. Velocity boundary conditions are set to no slip $(\bm{v}=\bm{0})$ at the top and the
-bottom, and a parabolic profile is imposed on the sides, with $u=y(1-y)$ and $v=0$. The analytical solution is then given by:
-\begin{eqnarray}
-u(x,y)&=&y(1-y)\nonumber \\
-v(x,y,)&=&0\nonumber \\
-p(x,y)&=&2\eta\left(\frac{L_x}{2}-x\right)\nonumber
-\end{eqnarray}
-
-The velocity field predicted by the model follows the expected parabolic profile (Fig. \ref{fig:poiseuille}a and Fig. \ref{fig:poi_plot}a) and, in case of the
-classic penalty method with no iterations, the pressure is clearly related to the divergence by means of the penalty parameter (Fig. \ref{fig:poiseuille}b and
-c and Fig. \ref{fig:poi_plot}b). Fig. \ref{fig:poiseuille}d shows that one Uzawa iteration is sufficient to bring the divergence down to \num{e-15}, with no
-correlation with the pressure field. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Poiseuille%20Flow}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Poiseuille.pdf}
-\caption{Velocity field (panel a), pressure (panel b) and divergence velocity of a Poiseuille flow in case of the classic penalty method (no iterations) and
-after one Uzawa iteration (panel c and d, respectively).}
-\label{fig:poiseuille}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Analytical.pdf}
-\caption{Velocity (panel a) and pressure (panel b) field predicted by the model for a Poiseuille flow with respect with their analytical solutions. The
-velocity field is plotted as function of the vertical coordinate in $L_x/2$ and the pressure is plotted as function of the x coordinate in $L_y/2$.}
-\label{fig:poi_plot}
-\end{figure}
-
-\subsection{Instantaneous 2D sphere}\label{sec:ist_sphere}
-The domain is a unit square with gravity $\bm{g}=(0,-1)$. The fluid has constant density and viscosity ($\rho_f=1$ and $\eta_f=1$). The sphere is in the middle
-of the domain with a radius R = 0.123456798 and has constant density and viscosity ($\rho_s=10^{-2}+\rho_f$ and $\eta_s=10^3 \cdot \eta_f$).
-We distinguish three different types of velocity boundary conditions:
-\begin{enumerate}
-\item FS: free slip conditions on all sides.
-\item NS: no slip conditions on all sides.
-\item OT: free slip conditions on the sides and the bottom, and open boundary at the top.
-\end{enumerate}
-All three cases are performed for different grid resolutions, between \num{16x16} and \num{512x512} elements with 50 randomly distributed markers per element,
-and different average schemes for the viscosity (harmonic, geometric and arithmetic). The velocity in the centre of the sphere $\bm{v}(0.5;0.5)$), minimum and
-maximum velocities ($u_{min}$, $u_{max}$, $v_{min}$, $v_{max}$) and pressures ($p_{min}$, $p_{max}$) on the entire domain, $v_{\textrm{rms}}$ and average pressure on the
-entire domain ($p_{avg}$) are compared with solutions generated by ASPECT \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and all results
-can be found at \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere2D} and 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Instantaneous_2D_Sphere}. Results in terms of the $v_{\textrm{rms}}$ for NS model 
-are also shown in Fig. \ref{fig:inst_sphere}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_NS.pdf}
-\caption{$v_{\textrm{rms}}$ as function of element size for the instantaneous 2D sphere experiment in case of no slip boundary conditions and with different average
-schemes for the viscosity. Results are compared with results obtained by other numerical codes.}
-\label{fig:inst_sphere}
-\end{figure}
-
-\subsection{Rayleigh-Taylor instability}\label{sec:rayleigh}
-This problem has been originally presented by \citet{vanKeken1997} and here it is performed as the isoviscous case (Case 1a). The domain has $L_x=0.9142$ and
-$L_y=1$ and gravity $\bm{g}=(0,-1)$. Two fluids with same constant viscosities ($\eta_1=\eta_2=1$) and different densities ($\rho_1=1000$ and $\rho_2=1010$),
-with the lighter fluid at the bottom. The initial interface between the fluids is given by $y(x)=0.2+0.02 \cos \left(\frac{\pi x}{L_x}\right)$. The experiment
-is performed with different grid sizes (\num{50x50}, \num{80x80}, \num{100x100} and \num{256x256}). A total of 1960000 markers are randomly distributed at the
-beginning of the simulation. Velocity boundary conditions are set to no slip at the top and the bottom, and to free slip at the sides of the domain.
-
-$v_{\textrm{rms}}$ as function of time is reported in Fig. \ref{fig:RT}, matching well with results shown by \citet{vanKeken1997}, \citet{Tackley2003} and
-\citet{Thieulot2014}. Fig. \ref{fig:rayleigh} shows the evolution of the experiment at different time steps. All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Rayleigh_Taylor_experiment/ISOVISCOUS}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/RT.pdf}
-\caption{$v_{\textrm{rms}}$ of the Rayleigh-Taylor experiment as function of time for different resolution of the grid.}
-\label{fig:RT}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=300px]{./Figures/Rayleigh.pdf}
-\caption{Evolution of the Rayleigh-Taylor experiment for a grid of \num{256x256} elements at $t=0, 500, 1000$ and $2000$ (panels a, b, c and d, respectively).}
-\label{fig:rayleigh}
-\end{figure}
-
-\subsection{Falling blocks}\label{sec:block}
-This benchmark is described as proposed by \citet{Gerya2003a}, \citet{Gerya2010b} and \citet{Thieulot2011}. The domain is square with $L_x=L_y=$ \SI{500}{\km}
-and the grid is composed by \num{50x50} elements with 25 markers in each element. The block is initially centred at ($x=$ \SI{250}{\km}; $y=$ \SI{400}{\km})
-and has a size of \num{100x100} km (Fig. \ref{fig:block}a). The fluid surrounding the block has $\eta_f=$ \SI{e21}{\pascal\per\s} and $\rho_f=$
-\SI{3200}{\kg\per\cubic\m}. The benchmark tests with different viscosities of the block, with $\eta_b$ from \SIrange{e15}{e27}{\pascal\per\s}. In each
-experiment also the density $\rho_b$ of the block varies from \SIrange{3220}{9900}{\kg\per\cubic\m}. Velocity boundary conditions are set to free slip
-conditions on all sides of the domain. Density distributions at $t=$ \SI{20}{\mega\year} for $\rho_b=$ \SI{3300}{\kg\per\cubic\m} and different $\eta_b$
-are plotted in Fig. \ref{fig:block}b-f, showing that the code correctly preserve the block geometry in case of large viscosity contrast, with a block stiffer
-than the surrounding fluid (Fig. \ref{fig:block}f).
-
-The velocity in the centre of the falling block at $t=0$ is measured for all experiments and, since the velocity should increase with density contrast, the
-quantity $\bm{v}/(\rho_b-\rho_f)$ is plotted as function of the viscosity contrast. All results perfectly match with those from \citet{Gerya2010b} and they
-line up on a single curve, demonstrating that the code can correctly deal with large viscosity and density contrasts (Fig. \ref{fig:falling}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Falling%20blocks}.
-
-\begin{figure}
-\noindent\includegraphics[width=300px]{./Figures/Block.pdf}
-\centering
-\caption{Density evolution of the falling block experiment at $t=$ \SI{0}{\mega\year} (panel a) and $t=$ \SI{20}{\mega\year} for different viscosities of the
-block (panels b-f).}
-\label{fig:block}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Falling.pdf}
-\caption{Initial velocity relative to the density contrast at the centre of the falling block as function of the viscosity contrast between the block and the
-surrounding fluid.}
-\label{fig:falling}
-\end{figure}
-
-\subsection{2D time-dependent Stokes sphere with deformable free surface}\label{sec:sphere}
-This experiment is performed in a unit square domain with the gravity acceleration fixed to $g_y=-1$. The sphere has $\rho_s=2$ and $\eta_s=10^3$ and is
-initially centred at ($0.5;0.6$) with radius $R=0.123456789$. The fluid surrounding the sphere has $\rho_f=1$ and $\eta_f=1$ and occupies the domain for
-$y\leq 0.75$, while for $y>0.75$ the air has $\rho_a=0$ and $\eta_a=10^{-3}$. Velocity boundary conditions are set to free slip on all sides. The Courant
-number is set to 0.25. The experiments run for 200 s using grid resolutions from \num{150x150} to \num{512x512} elements, with 25 randomly distributed markers
-per element, and for different average schemes for the viscosity. The interface between the fluid and the air is tracked by means of the markers chain. Results
-in terms of velocity and pressure fields and topography variations are compared with results obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and they can be found at
-\url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere_fs2D} and
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Time_dependent_sphere}. Results in terms of the $v_{\textrm{rms}}$ in case of 
-an arithmetic average are also shown in Fig. \ref{fig:inst_spherefs}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_arithm.pdf}
-\caption{$v_{\textrm{rms}}$ for different grid resolutions as function of time for the 2D time-dependent Stokes sphere experiment in case of an arithmetic average.
-Results are compared with results obtained by ASPECT.}
-\label{fig:inst_spherefs}
-\end{figure}
-
-\subsection{Free surface stabilisation}\label{sec:stab}
-The benchmark is performed as discussed in \citet{Kaus2010a} and \citet{Thieulot2014}. The domain is square with $L_x=L_y=$ \SI{500}{\km} and a grid resolution
-of \num{200x200} elements, each of them containing 16 markers. A buoyant fluid with $\rho_1=$ \SI{3200}{\kg\per\cubic\metre} and $\eta_1=$ \SI{e20}{\pascal\s}
-is overlain by a denser fluid with $\rho_2=$ \SI{3300}{\kg\per\cubic\metre} and $\eta_2=$ \SI{e21}{\pascal\s}. The initial interface between the fluids has an
-initial sinusoidal shape of 5 km amplitude. Velocity boundary conditions are set to free slip conditions at the sides and no slip at the bottom, while the top
-is free surface. The experiment is performed with various fixed time steps, without the use of the Courant number. The vertical position of the free surface at
-$x=L_x$ is tracked for each simulation, with or without the stabilisation algorithm.
-
-The results show an instability (drunken sailor effect) increasing the time step in case that the stabilisation algorithm is not activated
-($dt=$ \SI{4200}{\year}, green line in Fig. \ref{fig:stab}), as already observed by \citet{Kaus2010a} and \citet{Thieulot2014}. Activating the algorithm the
-instability is fixed and the simulation remains stable using time step up to \SI{50000}{\year} (red line in Fig. \ref{fig:stab}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Stabilisation_algorithm}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Kaus_random.pdf}
-\caption{Evolution of the y coordinate in $x=L_x$ as function of time for different time steps, with and without the stabilisation algorithm.}
-\label{fig:stab}
-\end{figure}
-
-\subsection{Topography relaxation}\label{sec:crameri}
-This experiment is performed as in \citet{Crameri2012} using free surface and sticky air (Case 1). The domain is rectangular with $L_x=$ \SI{2800}{\km} and
-$L_y=$ \SI{700}{\km} in case of free surface and $L_y=$ \SI{800}{\km} in case of sticky air ($h_{st}=$ \SI{100}{\km}). The grid is composed by \num{256x64}
-elements for the free surface \num{560x320} elements for the sticky air. For both cases, there is a mantle of 600 km thickness with $\eta_m=$ \SI{e21}{\pascal\s}
-overlain by a highly viscous cosine-shaped lithosphere with thickness between 93 and 107 km and $\eta_l=$ \SI{e23}{\pascal\s}. Both mediums have $\rho=$
-\SI{3300}{\kg\per\cubic\metre}. In case of sticky air experiment, the air has $\rho_a=$ \SI{0}{\kg\per\cubic\metre} and $\eta_a=$ \SI{e18}{\pascal\s}. The
-gravity acceleration is set to $g_y=$ \SI{-10}{\metre\per\square\s}. Velocity boundary conditions are set to no slip at the bottom and free slip at the sides
-of the domain. In case of sticky air, velocities on top are set to free slip conditions.
-
-The maximum topography as function of time ($h(t)$) can be analytically derived in according to \[h(t)=h_0 \exp(\gamma t)\]
-where $h_0=7$ km is the initial topography, $\gamma=\num{-0.2139e-11}$ is the characteristic relaxation rate and $t$ is the time (black line in Fig.
-\ref{fig:crameri}). Maximum topography at the characteristic relaxation time $t_{rlx}=$ \SI{14.825}{\kilo\year} can be found to be $h_{rlx}=$ \SI{2576}{\m}.
-The results show that a viscosity of the air of \SI{e18}{\pascal\s} is needed in case of $h_{st}=$ \SI{100}{\km} to correctly describe the topography relaxation
-for this problem, as pointed out by \citet{Crameri2012} (purple line in Fig. \ref{fig:crameri}). The case with a true free surface well follows the analytical
-solution (orange line in Fig. \ref{fig:crameri}). In particular, at the relaxation time in case of the true free surface predicts a maximum topography of
-\SI{2570}{\m}, with an error of \SI{6}{\m} with respect to the analytical solution. Results are also compared with results obtained with a free surface in
-UNDERWORLD (\num{256x64} elements) and SULEC (\num{401x201} elements) (with courtesy from \citealp{Crameri2012}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Topography_relaxation}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Crameri.pdf}
-\caption{Maximum topography as function of time for the Crameri benchmark (coloured lines) in comparison with the analytical solution (black line) and results
-obtained with UNDERWORLD (yellow line) and SULEC (blue line).}
-\label{fig:crameri}
-\end{figure}
-
 \subsection{Spontaneous subduction}\label{sec:subduction}
 This experiment is performed as presented by \citet{Schmeling2008}, in case of both a sticky air and a true free surface and using different average schemes
-for the viscosities. The domain is rectangular with $L_x=$ \SI{3000}{\km} and $L_y=$ \SI{700}{\km}, a grid resolution of \num{256x64} elements with 50 markers
+for the viscosities. The domain is rectangular with $L_x=$ \SI{3000}{\km} and $L_y=$ \SI{700}{\km}, a grid resolution of $256\times64$ elements with 50 markers
 per element and Courant number of 0.01. At the beginning of the simulation, a 100 km-thick lithospheric layer with $\rho_m=$ \SI{3300}{\kg\per\cubic\metre}
@@
 \end{figure}
-
-\subsection{The slab detachment}\label{sec:slab}
-The slab detachment benchmark is performed as described by \citet{Schmalholz2011} and \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{1000}{\km}
-and $L_y=$ \SI{660}{\km} and a grid resolution of \num{256x256} elements. A non-linear viscous T-shaped layer with $\rho_l=$ \SI{3300}{\kg\per\cubic\metre} is
-placed at the top of the domain and surrounded by a linear viscous fluid with $\rho_f=$ \SI{3150}{\kg\per\cubic\m}. The top layer is 80 km-thick and a 250
-km-long and 80 km-wide slab is placed at $x=L_x/2$ (Fig. \ref{fig:slab}a). The effective viscosity of the top layer is given by
-\[\eta_{eff}=\eta_0 I_2^{\frac{(1-n)}{n}}\]
-with $\eta_0=$ \SI{4.75e11}{\pascal\s} and $n=4$, while the surrounding fluid has $\eta_f=$ \SI{e21}{\pascal\s}. The effective viscosity is capped between
-$\eta_{min}=$ \SI{e21}{\pascal\s} and $\eta_{max}=$ \SI{e25}{\pascal\s}. Velocity boundary conditions are set to free slip at the top and the bottom, and to
-no slip at the sides of the domain.
-
-The sinking of the slab determines a decrease of the effective viscosity in the T-shaped layer during the first part of the evolution (Fig. \ref{fig:slab}a),
-as consequence of local strain rates, while effective viscosities increase up to $\eta_{max}$ in the top layer after approximately \SI{20}{\mega\year}, when
-necking is complete (Fig. \ref{fig:slab}b).
-
-The width of necking is tracked by means of markers position at the side of the slab. Necking width and time are normalised in function of the initial slab
-(\SI{80}{\km}) and the characteristic time ($t_c=$ \SI{7.1158e14}{\s}) \citep{Schmalholz2011,Glerum2018}. The results in case of arithmetic, geometric and
-harmonic mean of the viscosity are shown in Fig. \ref{fig:necking}, compared with results from \citet{Glerum2018} for which an infinity norm mean was used.
-The influence of the average schemes agrees with results shown in Section \ref{sec:subduction}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Slab_detachment}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Slab.pdf}
-\caption{Effective viscosity for the slab detachment benchmark at the beginning of the evolution (panel a) and when necking is complete (panel b).}
-\label{fig:slab}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Necking.pdf}
-\caption{Normalised width of the necking of the slab detachment benchmark as function of normalised time.}
-\label{fig:necking}
-\end{figure}
-
-\subsection{The indenter experiment}\label{sec:indenter}
-The indenter benchmark simulate a rigid punch on a purely plastic von Mises material, for which exists an analytical solution
-\citep{Thieulot2008,Thieulot2014,Glerum2018}. The domain is a unit square with a grid resolution of \num{256x256} elements. The gravity acceleration is set
-to $\bm{g}=\bm{0}$ and effective viscosity is capped using $\eta_{min}=10^{-4}$ and $\eta_{max}=10^3$. The tolerance for the convergence of the non-linear
-solution is $tol=10^{-9}$, with a maximum number of non-linear iterations set to 500. The medium in the domain has $\rho=0.01$, an initial viscosity
-$\eta_0=10$, cohesion $C=1$ and an angle of internal friction $\phi=0$°. Velocity boundary conditions are set to no slip at the bottom and free slip at the
-sides of the domain. The top of the domain is open with the exception of the central portion (punch area) with width $w_p=0.125$, where $v=-1.05$ and $u$ is
-fixed either to 0 (rough punch experiment) or free (smooth punch experiment).
-
-Fig. \ref{fig:indenter} show the results in terms of viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure
-(panels g and h) for both experiments. The analytical solution indicate that pressure at ($0.5;1$) and ($0.5 \pm w_p;1$)is $p=\pi +1$ and $p=1$, respectively
-(black and red lines in Fig. \ref{fig:smooth_p}). Fig. \ref{fig:smooth_p} shows a clear improvement in the pressure solution when passing from a rough to a
-smooth punch (green and blue lines in Fig. \ref{fig:smooth_p}, respectively), as previously observed by \citet{Thieulot2014} and \citet{Glerum2018}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Indenter}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Indenter.pdf}
-\caption{Viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure (panels g and h) fields for rough (left column) and
-smooth (right column) punch experiments.}
-\label{fig:indenter}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Smooth_Rough.pdf}
-\caption{Nodal pressure in function of the x coordinate at $y=1$ for rough (green line) and smooth (blue line) punch experiment. Black and red lines indicate
-the analytical solution at $x=0.5$ and $x=0.5 \pm w_p$, respectively.}
-\label{fig:smooth_p}
-\end{figure}
-
-\subsection{The brick experiment}\label{sec:brick}
-An instantaneous version of the brick benchmark is performed to verify the correctness of the pressure-dependent plasticity for different angle of internal
-friction, as proposed by \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{40}{\km} and $L_y=$ \SI{10}{\km} and a grid resolution of \num{512x128}
-elements. The gravity acceleration is set to $g_y=$ \SI{-10}{\m\square\s} and effective viscosity is capped using $\eta_{min}=$ \SI{e19}{\pascal\s} and
-$\eta_{max}=$ \SI{e26}{\pascal\s}. The tolerance for the convergence of the non-linear solution is $tol=10^{-7}$, with a maximum number of non-linear iterations
-set to 1000. A 800 m-wide and 400 m-height inclusion with $\eta_b=$ \SI{e20}{\pascal\s} and $\rho_b=$ \SI{2700}{\kg\per\cubic\m} is placed at the bottom of the
-domain at $x=L_x/2$. The inclusion is surrounded by a non-linear viscous medium with $\rho_m=$ \SI{2700}{\kg\per\cubic\m}, an initial viscosity of
-\SI{e23}{\pascal\s}, a linear viscous viscosity of \SI{e25}{\pascal\s} and a cohesion of \SI{40}{\mega\pascal}. Velocities are set to free slip conditions at
-the bottom of the domain and the top is open. Velocities on sides of the domain are fixed to $u=$ \SI{+-2e-11}{\m\per\s} and $v=0$. The experiment is performed
-in compressional and extensional contexts with an angle of internal friction $\phi$ of the non-linear medium variable between 0° and 30°.
-
-As expected, two shear bands stem from the inclusion with variable angles in relation to both the dynamics context and the internal friction angle. Shear band
-angles formed at 45° for $\phi=0$° in both extensional and compressional contexts (Fig. \ref{fig:brick_beam}a), while they are different in case of $\phi\neq0$°.
-Shear band angles in case of $\phi=20$° are shown for compression and extension in Fig. \ref{fig:brick_beam}b  and c, respectively. Values of shear band angles
-as function of internal friction angles from 0° to 30° are extracted at different depths and minimum and maximum values are plotted in Fig. \ref{fig:brick},
-in comparison with theoretical Roscoe, Arthur and Coulomb shear band angles. For all tests the tolerance for nonlinear convergence is defined $tol=10^{-7}$
-and none converges before the maximum number of iterations ($it_{max}=1000$) is reached. However, tests with internal friction angles up to 15° show a constant
-decrease of the velocity residuals, which can not be observed for for tests with higher internal friction angles (Fig. \ref{fig:convergence}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Brick_experiment}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Brick_Beam.pdf}
-\caption{Shear band angles predicted for the brick experiment for $\phi=0$° (panel a) and for $\phi=20$° in case of compression and extension (panels b and c,
-respectively).}
-\label{fig:brick_beam}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Brick.pdf}
-\caption{Shear band angles predicted for the brick experiment in case of compressional and extensional contexts as function of different internal angle of
-friction (continuous black line and red dots), compared with theoretical Roscoe, Artur and Coulomb angles (discontinuous black lines). Red lines indicate the
-range of angles calculated at different depths.}
-\label{fig:brick}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \noindent\includegraphics[width=500px]{./Figures/convergence.pdf}
-  \caption{Convergence rates for all tests. Normalised residuals (left column) are compared with velocity residuals (right column) for extensional (first row)
-  and compressional (second row) settings.}
-  \label{fig:convergence}
-  \end{figure}
-
-\subsection{Advection stabilisation}\label{sec:advection}
-The 1D advection problem proposed by \citet{Donea2003} and \citet{Thieulot2011} is performed to verify the effectiveness of the SUPG method to stabilise the
-advection term of the energy equation. The domain is a 1D segment with $L_x=1$ composed by 50 elements and a discontinuity in the thermal field placed at
-$x=0.25$. Temperature is set to 1 for $x \leq 0.25$ and to 0 for $x > 0.25$. Velocity is set to ${u}=1$ in the entire domain. The simulation is performed for
-250 time steps, with $dt=0.002$, so the thermal discontinuity should be at $x=0.75$ at the end of the simulation.
-
-Temperature profiles at the end of the simulation are shown in \ref{fig:advection} as function of the dimensionless coefficient $\gamma=\tau \bm{v}/h$. In case
-of the classic Galerkin method ($\gamma=0$, blue line in \ref{fig:advection}) the final thermal profile is characterised by strong oscillations, which are
-eliminated in case of the SUPG method ($\gamma=0.045$, orange line in \ref{fig:advection}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Advection_stabilisation}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Advection.pdf}
-\caption{Temperature profile as function of \textit{x} for the advection stabilisation benchmark. Purple line indicates the initial temperature profile; the
-green line indicates the analytical temperature profile after 250 time steps; blue line indicates the temperature profile after 250 time steps in case of the
-classic Galerkin method ($\gamma=0$); orange line indicates the temperature profile after 250 time steps in case of the SUPG method ($\gamma=0.045$).}
-\label{fig:advection}
-\end{figure}
-
-\subsection{Shear heating}\label{sec:simple_shear}
-This exercise is performed in an unit square domain composed by \num{128x128} elements. The velocity field is prescribed on the entire domain with
-$\bm{v}=(L_y-y)y\bm{e_x}$; viscosity, density and specific heat are fixed to 1, while thermal conductivity and radiogenic energy are fixed to 0. Therefore,
-the energy equation (Eq. \ref{eq:energy}) can be simplified as \[\frac{\partial T}{\partial t}=H_s\] and fixing $T(t=0)=0$, the temperature can be find as
-\[T(t)=H_s t\]
-In this case we have
-\begin{eqnarray}
-\dot{\epsilon_{xx}}&=&\frac{\partial u}{\partial x}=0 \nonumber \\
-\dot{\epsilon_{yy}}&=&\frac{\partial v}{\partial y}=0 \nonumber \\
-\dot{\epsilon_{xy}}&=&\frac{1}{2}\left(\frac{\partial v}{\partial x}+\frac{\partial u}{\partial y}\right)=\frac{1}{2}(L_y-2y)\nonumber
-\end{eqnarray}
-and, simplifying Eq. \ref{eq:shear}, shear heating can be calculated as \[H_s(x,y)=(1-2y)^2\]
-
-The solution predicted by the model in terms of velocity, temperature and shear heating match well with the analytical solutions (Fig. \ref{fig:analytical_en}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Simple_shear}.
-
-%\begin{figure}
-%\centering
-%\noindent\includegraphics[width=400px]{./Figures/Energy_Thieulot.pdf}
-%\caption{Temperature and shear heating predicted by the model after 1 time step.}
-%\label{fig:en_thieulot}
-%\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=500px]{./Figures/Analytical_Shear.pdf}
-\caption{Velocity (panel a), temperature (panel b) and shear heating (panel c) as function of the \textit{y} coordinate for the simple shear experiment. The
-solutions predicted by the model (red lines) are compared with the analytical solutions (dashed blue lines).}
-\label{fig:analytical_en}
-\end{figure}
-
-\subsection{Shear and adiabatic heating}\label{sec:shear}
-This problem is performed as presented in Exercise 9.4 in \citet{Gerya2010b}. Two materials are vertically separated in a rectangular domain with $L_x=$
-\SI{1000}{\km}, $L_y=$ \SI{1500}{\km} and a grid resolution of \num{30x20} elements. Constant thermal coefficient expansion ($\alpha=$ \SI{3e-5}{\per\kelvin}),
-temperature ($T=$ \SI{1300}{\kelvin}) and gravity acceleration ($g_y=$ \SI{-10}{\m\square\s}) are assumed in the whole domain. Fluid 1 (on the left side of the
-domain) has $\rho_1=$ \SI{3200}{\kg\per\cubic\m} and $\eta_1=$ \SI{e20}{\pascal\s}; fluid 2 (on the right side of the domain) has $\rho_2=$
-\SI{3300}{\kg\per\cubic\m} and $\eta_2=$ \SI{e22}{\pascal\s}. Velocity boundary conditions are set to free slip on all sides of the domain.
-
-As shown in Fig. \ref{fig:energy}, both shear and adiabatic heating predicted by the code (first row in Fig. \ref{fig:energy}) well recreate the results
-obtained by means of example ${Shear\_adiabatic\_heating.m}$ from \citet{Gerya2010b} (second row in Fig. \ref{fig:energy}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Adiabatic%2BShear}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Energy.pdf}
-\caption{Comparison between shear (first column), adiabatic (second column) and total (third column) energy predicted by the code (panels a, b and c) and those
-created using example ${Shear\_adiabatic\_heating.m}$ from Exercise 9.4 in \citet{Gerya2010b} (panels d, e and f).}
-\label{fig:energy}
-\end{figure}
-
-\subsection{Mantle convection}\label{sec:mantle}
-This problem is performed as presented by \citet{Blankenbach1989} (constant viscosity cases) and \citet{Thieulot2014}, in a 2D unit square domain with gravity
-acceleration $g_y=10^{10}Ra$. The experiment is performed with three different Rayleigh numbers ($Ra=10^4$, $10^5$ and $10^6$) and with different grid
-resolution (between \num{32x32} and \num{128x128} elements). The fluid has constant viscosity, initial density, heat capacity, thermal conductivity
-($\eta= \rho_0= C_p = k=1$), reference temperature ($T_0=0$) and thermal expansion coefficient ($\alpha=10^{-10}$). Temperatures are set to 0 on top and 1 on
-bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\]
-
-The solution generated by the code in terms of $v_{\textrm{rms}}$ (calculated as in Sec. \ref{sec:error}) and Nu as function of time are reported for all the
-simulations in Fig. \ref{fig:mantle}. At the steady state, $v_{\textrm{rms}}$ and Nu of all simulations converge well toward the values from \citet{Blankenbach1989},
-with lower errors for higher resolution grids (Table \ref{tab:mantle}). All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Mantle_convection}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Convection.pdf}
-\caption{$v_{\textrm{rms}}$ (panels a, c and e) and Nu (panels b, d and f) for the mantle convection benchmark as function of time for different grid resolution.
-Panels a and b show the results for $Ra=10^4$; Panels c and d show the results for $Ra=10^5$; panels e and f show the results for $Ra=10^6$. Purple lines
-indicate the convergence values for the $v_{\textrm{rms}}$ and the Nu from \citet{Blankenbach1989} at the steady state.}
-\label{fig:mantle}
-\end{figure}
-
-\begin{table}[H]
-\caption{Comparison between $v_{\textrm{rms}}$ and Nu predicted by the code for the mantle convection experiment and same values as reported in literature.}
-\centering
-\small
-\begin{tabular}{|>{\centering}m{0.09\textwidth}|>{\centering}m{0.05\textwidth}|>{\centering}m{0.24\textwidth}|>{\centering}m{0.15\textwidth}|
-  >{\centering}m{0.1\textwidth} >{\centering}m{0.1\textwidth} >{\centering\arraybackslash}m{0.1\textwidth}|}
-\toprule
-& & \multirow{2}{*}{\citet{Blankenbach1989}} & \citet{Thieulot2014} & \multicolumn{3}{c|}{FALCON}  \\
- & & & \num{200x200} & \num{32x32} & \num{64x64} & \num{128x128}  \\
-\midrule
-\multirow{2}{*}{$Ra=10^4$} & $v_{\textrm{rms}}$ & $42.864947 \pm 0.000020$ & 42.867 & 42.83226 & 42.852793 & 42.861394  \\
-           & Nu      & $4.884409 \pm 0.000010$  & 4.882  & 4.781297 & 4.857475  & 4.877573   \\
-\hline
-\multirow{2}{*}{$Ra=10^5$} & $v_{\textrm{rms}}$ & $193.21454 \pm 0.00010$  & 193.255& 193.872643 & 193.377472 & 193.252290  \\
-           & Nu      & $10.534095 \pm 0.000010$ & 10.507 & 9.602514 & 10.270735  & 10.465629   \\
-\hline
-\multirow{2}{*}{$Ra=10^6$} & $v_{\textrm{rms}}$ & $833.98977 \pm 0.00020$  & 834.712& 848.091176 & 837.767911 & 834.945793  \\
-           & Nu      & $21.972465 \pm 0.000020$ & 21.695 & 15.999266 & 19.703682  & 21.306939   \\
-\bottomrule
-\end{tabular}
-\label{tab:mantle}
-\end{table}
-
-\subsection{Viscoplastic mantle convection}\label{sec:visco_mantle}
-This benchmark is performed as Cases 1-5a described in \citet{Tosi2015} in a 2D unit square domain with two different grid resolutions of \num{32x32} 
-and \num{100x100} elements. Reference viscosity, reference density, heat capacity, thermal conductivity and thermal expansion coefficient are fixed to a constant
-value of 1 ($\eta_0= \rho_0= C_p = k=\alpha=1$), while the gravity has been chosen as $g_y=10^2$, in order to obtain a Rayleigh number $Ra=10^2$. Temperatures
-are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\].
-The viscosity field $\eta$ is calculated as the harmonic average between a linear part $\eta_{lin}$ that depends on temperature only or on temperature and 
-depth $y$ and a nonlinear-plastic part $\eta_{pl}$ that depends on the strain rate $\dot{\bm{\epsilon}}$, as follows
-\[\eta(T,y,\dot{\bm{\epsilon}})=2\left(\frac{1}{\eta_{lin}(T,y)}+\frac{1}{\eta_{pl}(\dot{\bm{\epsilon}})}\right)^{-1}\]
-The linear and the nonlinear-plastic parts of the viscosity are calculated following \citet{Tosi2015} as follows
-\[\eta_{lin}(T,y)=\exp(-\ln(\Delta\eta_T)T+\ln(\Delta\eta_y)y)\]
-\[\eta_{pl}(\dot{\bm{\epsilon}})=\eta^*+\frac{\sigma_Y}{\sqrt{\dot{\bm{\epsilon}}:\dot{\bm{\epsilon}}}}\]
-where $\Delta\eta_T$, $\Delta\eta_y$, $\eta^*$ and $\sigma_Y$ are parameters chosen as listed in Table \ref{tab:tosi_setup}.
-
-For each case, some diagnostic quantities measured at the steady state are reported in Table \ref{tab:visco_mantle} (see Section \ref{sec:error} for 
-the computation of all the quantities), showing a well fit with results described in \citet{Tosi2015}. Data and charts for all cases can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Viscoplastic_mantle_convection}.
-
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{white!100!}{gray!25!}
-%\begin{table}[H]
-%  \caption{Results of the viscoplastic mantle convection benchmark for Cases 1-4 compared with results obtained with ELEFANT and YACC (from \citealp{Tosi2015}).}
-%  \centering
-%  \resizebox{\textwidth}{!}{
-%\begin{tabular}{|c|ccc|ccc|ccc|ccc|}
-%  \toprule
-%    & \multicolumn{3}{c|}{\textbf{Case 1}} & \multicolumn{3}{c|}{\textbf{Case 2}} & \multicolumn{3}{c|}{\textbf{Case 3}} &
-%    \multicolumn{3}{c|}{\textbf{Case 4}} \\
-%   & FALCON & ELEFANT& YACC & FALCON& ELEFANT& YACC & FALCON & ELEFANT & YACC & FALCON & ELEFANT & YACC \\
-%  \midrule
-%   $\left\langle T \right\rangle$ & 0.7759 & 0.7758 & 0.7767 & 0.6043 & 0.6036 & 0.6058 & 0.7275 & 0.7275 & 0.7286 & & 0.5277 & 0.5289\\
-%    Nu$^{top}$ & 3.4280 & 3.4214 & 3.4298 & 8.5242 & 8.5115 & 8.5278 & 3.0368 & 3.0347 & 3.0374 && 6.5912 & 6.5572 \\
-%    Nu$^{bottom}$ & 3.3168 & 3.313 & 3.3143 & 8.4599 & 8.4465 & 8.3990 & 3.0086 & 2.9908 & 2.9628 && 6.5834 & 6.5243 \\
-%    $v_{\textrm{rms}}$ & 249.901 & 249.134 & 251.7997 & 141.2451 & 140.7087 & 142.2020 & 100.2776 & 100.1208 & 100.9467 && 79.1105 & 79.6202 \\
-%    $v_{\textrm{rms}}^{top}$ & 1.8907 & 1.8642 & 1.8298 & 104.7928 & 104.1228 & 105.1767 & 2.0832 & 2.0652 & 2.0374 && 74.7596 & 75.8414 \\
-%    $u_{max}^{top}$ & 2.6340 & 2.6119 & 2.5516 & 121.9212& 121.766 & 122.3238 & 2.9086 & 2.9019 & 2.8458 && 88.9146 & 89.2940 \\
-%    $\eta_{min}$ & - & - & - & $2.0303\times10^{-5}$ & $2.0600\times10^{-5}$ & $1.9853\times10^{-5}$ & $4.8129\times10^{-5}$ & $4.8080\times10^{-5}$ & $4.7907\times10^{-5}$ &
-%    & $1.9860\times10^{-4}$ & $1.9174\times10^{-4}$ \\
-%    $\eta_{max}$ & - & - & - & 1.3325 & 1.85 & 1.6095 & 0.9360 & 0.9023 & 1 && 1.52 & 1.6773 \\
-%    $\left\langle W \right\rangle$ & 2.4299 & 2.4316 & 2.4583 & 7.5648 & 7.5665 & 7.6222 & 2.0387 & 2.0384 & 2.0400 && 5.6216 & 5.6512 \\
-%    $\left\langle \Phi \right\rangle/Ra$ & 2.4311 & 2.4276 & 2.4333 & 7.5661 & 7.5585 & 7.6194 & 2.0386 & 2.0370 & 2.0335 && 5.6182 & 5.6463 \\
-%    $\delta$ & 0.05\% & 0.16\% & 1.02\% & 0.02\% & 0.11\% & 0.04\% & $<$0.01\% & 0.07\% & 0.32\% && 0.06\% & 0.09\% \\
-%  \bottomrule
-%  \end{tabular}}
-%  \label{tab:visco_mantle}
-%  \end{table}
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{2}{white!100!}{gray!25!}
-\begin{table}[H]
-  \caption{Benchmark cases and corresponding parameters.}
-  \centering
-  \small
-\begin{tabular}{|>{\centering}m{0.08\textwidth}|>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}
-  >{\centering}m{0.05\textwidth} >{\centering}m{0.05\textwidth} |>{\centering\arraybackslash}m{0.2\textwidth}|}
-  \toprule
-    Case & $Ra$ & $\Delta\eta_T$ & $\Delta\eta_y$ & $\eta^*$ & $\sigma_Y$ & Convective regime \\
-  \midrule
-  \rowcolor{gray!25!}
-    1 & $10^2$ & $10^5$ & 1 & - & - & Stagnant lid \\
-    2 & $10^2$ & $10^5$ & 1 & $10^{-3}$ & 1 & Mobile lid  \\
-    \rowcolor{gray!25!}
-    3 & $10^2$ & $10^5$ & 10 & - & - & Stagnant lid  \\
-    4 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 1 & Mobile lid \\
-    \rowcolor{gray!25!}
-    5 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 4 & Periodic \\
-  \bottomrule
-  \end{tabular}
-  \label{tab:tosi_setup}
-  \end{table}
-
-\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{gray!25!}{white!100!}
-\begin{table}
-  \caption{Results of the viscoplastic mantle convection benchmark for different grid resolutions in Cases 1-5 compared with results obtained with ELEFANT
-  and YACC ($100\times100$ elements, from \citealp{Tosi2015}) and in Stone 28 ($32\times32$ elements, from
-  \url{https://github.com/cedrict/fieldstone/tree/master/python_codes/fieldstone_28}).}
-  \centering
-  \resizebox{\textwidth}{!}{
-\begin{tabular}{|cc|ccccccccccc|}
-  \toprule
-  \multicolumn{13}{|c|}{\textbf{Case 1}} \\
-  \midrule
-  Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-  & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-  \hline
-  \rowcolor{gray!25!}
-    YACC & $100\times100$ & 0.7767 & 3.4298 & 3.3143 & 251.7997 & 1.8298 & 2.5516 & - & - & 2.4583 & 2.4333 & 1.02\% \\
-    ELEFANT & $100\times100$ & 0.7758 & 3.4214 & 3.313 & 249.134 & 1.8642 & 2.6119 & - & - & 2.4316 & 2.4276 & 0.16\% \\
-    \rowcolor{gray!25!}
-    Stone 28 & $32\times32$ & 0.7737 & 3.3987 & - & 243.872 & - & - & - & - & - & - & - \\
-%    FALCON & $32\times32$ & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & - & - & 2.4745 & 2.4864 & 0.48\% \\
-%    \rowcolor{gray!25!}
-    FALCON & $100\times100$ & 0.7759 & 3.4280 & 3.3168 & 249.901 & 1.8907 & 2.6340 & - & - & 2.4299 & 2.4311 & 0.05\% \\
-    \midrule
-    \multicolumn{13}{|c|}{\textbf{Case 2}} \\
-    \midrule
-    Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-    & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-    \hline
-    \rowcolor{gray!25!}
-     YACC & $100\times100$ & 0.6058 & 8.5278 & 8.3990 & 142.2020 & 105.1767 & 122.3238 & $1.9853\times10^{-5}$ & 1.6095 & 7.6222 & 7.6194 & 0.04\% \\
-     ELEFANT & $100\times100$ & 0.6034 & 8.5115 & 8.4465 & 140.8067 & 104.1228 & 121.766 & $2.0600\times10^{-5}$ & 1.85 & 7.5665 & 7.5585 & 0.11\% \\
-     \rowcolor{gray!25!}
-     Stone 28 & $32\times32$ & 0.6052 & 8.1715 & - & 141.518 & - & - & - & - & - & - & - \\
-%     FALCON & $32\times32$ & 0.6454 & 8.1422 & 7.4020 & 156.6908 & 95.7226 & 116.9302 & $2.2572\times10^{-5}$ & 0.7194 & 7.4289 & 7.4452 & 0.22\% \\
-%     \rowcolor{gray!25!}
-     FALCON & $100\times100$ & 0.6044 & 8.5266 & 8.4590 & 141.3684 & 104.8867 & 122.0470 & $2.0333\times10^{-5}$ & 1.2914 & 7.5707 & 7.5718 & 0.01\% \\
-     \midrule
-     \multicolumn{13}{|c|}{\textbf{Case 3}} \\
-     \midrule
-     Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-     & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-     \hline
-     \rowcolor{gray!25!} 
-     YACC & $100\times100$ & 0.7286 & 3.0374 & 2.9628 & 100.9467 & 2.0374 & 2.8458 & $4.7907\times10^{-5}$ & 1 & 2.0400 & 2.0335 & 0.32\% \\
-      ELEFANT & $100\times100$ & 0.7275 & 3.0347 & 2.9908 & 100.1208 & 2.0652 & 2.9019 & $4.8080\times10^{-5}$ & 0.9023 & 2.0384 & 2.037 & 0.07\% \\
-      \rowcolor{gray!25!}
-      Stone 28 & $32\times32$ & 0.7277 & 3.0347 & - & 100.018 & - & - & - & - & - & - & - \\
-%      FALCON & $32\times32$ & 0.7281 & 3.0618 & 2.8132 & 101.0092 & 2.1630 & 3.0178 & $4.8500\times10^{-5}$ & 0.7845 & 2.0699 & 2.0695 & 0.02\% \\
-%      \rowcolor{gray!25!}
-      FALCON & $100\times100$ & 0.7275 & 3.0368 & 3.0086 & 100.2776 & 2.0832 & 2.9086 & $4.8129\times10^{-5}$ & 0.9360 & 2.0387 & 2.0386 & $<$0.01\% \\
-      \midrule
-      \multicolumn{13}{|c|}{\textbf{Case 4}} \\
-      \midrule
-      Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-      & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-      \hline
-      \rowcolor{gray!25!} 
-      YACC & $100\times100$ & 0.5289 & 6.5572 & 6.5243 & 79.6202 & 75.4814 & 89.2940 & $1.9174\times10^{-4}$ & 1.6773 & 5.6512 & 5.6463 & 0.09\% \\
-       ELEFANT & $100\times100$ & 0.5277 & 6.5912 & 6.5834 & 79.1105 & 74.7596 & 88.9146 & $1.9860\times10^{-4}$ & 1.5200 & 5.6216 & 5.6182 & 0.06\% \\
-       \rowcolor{gray!25!}
-       Stone 28 & $32\times32$ & 0.5286 & 6.4036 & - & 79.4746 & - & - & - & - & - & - & - \\
-%       FALCON & $32\times32$ & 0.5334 & 6.4679 & 6.3435 & 80.6153 & 76.0511 & 90.2058 & $2.0393\times10^{-4}$ & 0.6453 & 5.6676 & 5.6698 & 0.04\% \\
-%       \rowcolor{gray!25!}
-       FALCON & $100\times100$ & 0.5279 & 6.5897 & 6.5830 & 79.1220 & 75.0715 & 88.8484 & $1.9693\times10^{-4}$ & 1.4133 & 5.6127 & 5.6113 & 0.03\% \\
-       \midrule
-       \multicolumn{13}{|c|}{\textbf{Case 5}} \\
-       \midrule
-       Code & Resolution & $P$ & $\left\langle T \right\rangle_{min}$ & $\left\langle T \right\rangle_{max}$ & Nu$^{top}_{min}$ & Nu$^{top}_{max}$ &
-       $v_{\textrm{rms}}^{min}$ & $v_{\textrm{rms}}^{max}$ & \multicolumn{2}{c}{$(\left\langle \Phi \right\rangle/Ra)_{min}$} & 
-       \multicolumn{2}{c|}{$(\left\langle \Phi \right\rangle/Ra)_{max}$} \\
-       \hline
-       \rowcolor{gray!25!}
-         YACC & $100\times100$ & 0.0792 & 0.6543 & 0.6722 & 2.6950 & 7.2792 & 41.9813 & 98.8230 & \multicolumn{2}{c}{1.4347} & 
-         \multicolumn{2}{c|}{9.2748} \\
-         ELEFANT & $100\times100$ & 0.076 & 0.6514 & 0.6693 & 2.6808 & 7.3506 & 41.2578 & 100.089 & \multicolumn{2}{c}{1.3324} & 
-         \multicolumn{2}{c|}{9.5289} \\
-         \rowcolor{gray!25!}
-         Stone 28 & $32\times32$ & 0.0764 & 0.6521 & 0.6697 & 2.6912 & 7.0763 & 41.61 & 98.183 & \multicolumn{2}{c}{-} & 
-         \multicolumn{2}{c|}{-} \\
-     %    FALCON & $32\times32$ & & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & \multicolumn{2}{c}{-} & 
-     %     \multicolumn{2}{c|}{-} \\
-     %    \rowcolor{gray!25!}
-         FALCON & $100\times100$ & 0.0775 & 0.6519 & 0.6697 & 2.6847 & 7.3629 & 41.5256 & 99.5466 & \multicolumn{2}{c}{1.3399} & 
-         \multicolumn{2}{c|}{9.3993} \\
-  \bottomrule
-  \end{tabular}}
-  \label{tab:visco_mantle}
-  \end{table}
-
-\subsection{Thin layer entrainment}\label{sec:thin}
-This experiment is performed as originally proposed by \citet{vanKeken1997} with two incompressible fluids in a rectangular domain with $L_x=2$ and $L_y=1$
-and gravity acceleration $g_y=10^{10}Ra$. The Rayleigh number and the compositional Rayleigh number are fixed to $Ra=300000$ and $Rc=450000$, respectively.
-Both fluids have constant viscosity, thermal conductivity, specific heat ($\eta=\rho=C_p=1$) and thermal expansion coefficient ($\alpha=10^{-10}$). Fluid 1
-has a density $\rho_1=1$, while fluid 2 is denser ($\rho_2=\rho_1+1.5 \cdot 10^{-10}$) and is placed at the bottom of the domain, for $y \leq 0.025$.
-Temperature are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides of the domain. The initial
-temperature field is given by
-\[T(x,y)=T_u(x,y)+T_l(x,y)+T_r(x,y)+T_s(x,y)-\frac{3}{2}\]
-with
-\begin{eqnarray}
-T_u(x,y)&=&\frac{1}{2}erf\left(\frac{1-y}{2}\sqrt{\frac{u_0}{x}}\right)\nonumber \\
-T_l(x,y)&=&1-\frac{1}{2}erf\left(\frac{y}{2}\sqrt{\frac{u_0}{L_x-x}}\right)\nonumber \\
-T_r(x,y)&=&\frac{1}{2}+\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{y+1}} \exp\left(-\frac{x^2u_0}{4y+4}\right)\nonumber \\
-T_s(x,y)&=&\frac{1}{2}-\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{2-y}} \exp\left(-\frac{(L_x-x)^2u_0}{8-4y}\right)\nonumber
-\end{eqnarray}
-and
-\begin{eqnarray}
-u_0&=&\frac{L_x^{7/3}}{(1+L_x^4)^{2/3}}\left(\frac{Ra}{2\sqrt{\pi}}\right)^{\frac{2}{3}}\nonumber \\
-Q&=&2\sqrt{\frac{L_x}{\pi u_0}}\nonumber
-\end{eqnarray}
-
-The experiment is performed with two grid resolution (\num{125x40} and \num{200x80} elements), with 100 markers per element and CFL of 0.25. $v_{\textrm{rms}}$
-(see Sec. \ref{sec:error}) calculated for both the simulations match well with results from \citet{vanKeken1997} and obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and ELEFANT \citep{Thieulot2014}, obtained with similar grid resolutions
-(Fig. \ref{fig:thin}). In addition, a variety of simulations with different resolutions and aspect ratios are performed to verify that the sensitivity of the
-initial velocity field. Results match well with results from \citet{vanKeken1997} (in the grey area) and \citet{Thieulot2014} (Fig. \ref{fig:thin_initial}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Thin_layer}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_thin.pdf}
-\caption{$v_{\textrm{rms}}$ for the thin layer experiment as function of time for different grid resolution. Results are compared with results obtained by
-\citet{vanKeken1997} (black lines) and with ASPECT and ELEFANT.}
-\label{fig:thin}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrmszero.pdf}
-\caption{$v_{\textrm{rms}}$ at $t=0$ for the thin layer experiment as function of the element size for different aspect ratios of the elements. Results are compared
-with results obtained by \citet{vanKeken1997} (black dots in the grey area) and with ASPECT and ELEFANT.}
-\label{fig:thin_initial}
-\end{figure}
-
-\subsection{Sinking hydrated cylinder}\label{sec:quinquis}
-This experiment is performed as explained in \citet{Quinquis2014} and simulates the sinking of a cold, hydrated cylinder into a warm, dry mantle. The domain
-is square with $L_x=L_y=$ \SI{300}{\km}, a grid resolution of \num{300x300} elements and 25 markers per element. The sinking cylinder is located at $x=$
-\SI{150}{\km} and $y=$ \SI{170}{\km} with a radius of \SI{20}{\km} and has $\rho_c=$ \SI{3250}{\kg\per\cubic\m}, $\eta_c=$ \SI{e23}{\pascal\s}, a thermal
-conductivity $k_c=$ \SI{4.5}{\watt\per\m\per\kelvin},a specific heat $C_{p_c}=$ \SI{1250}{\joule\per\kg\per\kelvin}; the surrounding mantle has $\rho_m=$
-\SI{3200}{\kg\per\cubic\m}, $\eta_m=$ \SI{e20}{\pascal\s}, a thermal conductivity $k_m=$ \SI{105}{\watt\per\m\per\kelvin} and specific heat $C_{p_m}=$
-\SI{1250}{\joule\per\kg\per\kelvin}; a 58 km-thick lithosphere overlies the mantle and has $\rho_l=$ \SI{3200}{\kg\per\cubic\m}, $\eta_l=$ \SI{e23}{\pascal\s},
-a thermal conductivity $k_l=$ \SI{4.5}{\watt\per\m\per\kelvin}, specific heat $C_{p_l}=$ \SI{750}{\joule\per\kg\per\kelvin} and a thermal diffusivity of
-\SI{e-6}{\square\m\per\s}. Velocity boundary conditions are set to free slip on all sides of the domain. Temperature increase linearly in the lithosphere,
-from 0° to 1300°C, and in the mantle, from 1300° to 1360.5°C. The initial temperature of the cylinder is fixed to 400°C. Throughout the evolution, temperature
-is fixed to 0° and 1360.5°C at the top and bottom of the domain, respectively. The initial bound water content is imposed at 2 wt.\% in the cylinder and at
-0 wt.\% in both the mantle and the lithosphere. The maximum amount of water in the mantle is fixed to 0.2 wt.\%, while maximum water content in the cylinder
-and in the lithosphere is function of pressure and temperature and is calculated using a serpentinized harzburgite with Perple\_X, as in \citet{Quinquis2014}.
-
-As shown in \citet{Quinquis2014}, the model is characterised by a progressive dehydration in the external portion of the cylinder, due to the increase of
-temperature, with a consequent vertical migration of free water that hydrates the mantle above the cylinder, up to the lithosphere. The distribution of
-bound and free water during the evolution of the experiment is shown in Fig. \ref{fig:quinquis} (left and right columns, respectively). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Hydration}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Hydration.pdf}
-\caption{Distribution of bound and free water (left and right columns, respectively) at $t=0.5$ and $t=$ \SI{1.25}{\mega\year} (first and second rows,
-respectively) for the sinking hydrated cylinder experiment. Only markers with bound or free water not equal to 0 are plotted.}
-\label{fig:quinquis}
-\end{figure}
-
-\subsection{Mantle melting curves}\label{sec:katz}
-This instantaneous experiment is performed in order to recreate isobaric melting curves, as function of temperature, bulk water and content of water in the
-melt, to compare with melting curves obtained by \citet{Katz2003}. The domain is square with $L_x=L_y=100$ km, a grid resolution of \num{100x100} elements and
-25 markers per element. Density is set in the entire domain to \SI{3300}{\kg\per\cubic\m} and temperature increases from the sides to the centre from 1000° to
-1500°C. The experiment is performed for bulk water contents from 0 to 0.3 wt.\%.
-
-Fig. \ref{fig:fraction} shows isobaric melting curves for \SI{1}{\giga\pascal} (panel a) and \SI{3}{\giga\pascal} (panel b) calculated for different bulk water
-contents. The curves well match with those obtained for the mantle by \citet{Katz2003}, showing the saturation at the solidus for bulk water of 0.2 and 0.3 wt.\%, indicating
-that the code correctly determines melt fraction as function of pressure, temperature and bulk water content. Same results can be observed in Fig.
-\ref{fig:water}a, where isobaric and isothermal melting curves are shown as function of bulk water content. In addition, Fig. \ref{fig:water}b shows that the
-code correctly determines also the percentage of water in the melt as function of pressure, temperature and melt fraction. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Melting}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Fraction.pdf}
-\caption{Isobaric melting curves for 1 GPa (panel a) and 3 GPa (panel b) as function of temperature with different bulk water contents.}
-\label{fig:fraction}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Water.pdf}
-\caption{Isobaric melting curves for 1.5 GPa as function of bulk water (panel a) and water in the melt (panel b) calculated for different temperatures.}
-\label{fig:water}
-\end{figure}
-
-\clearpage
-
-\bibliography{export}
-
-\end{document}

### 2022-09-07 tex/Benchmark/thin.tex
update main

--- 2022-03-31 FALCON_description.tex
@@
-\documentclass[hidelinks,10pt,a4paper]{article}
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{textcomp}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage{booktabs}
-\usepackage{multirow}
-\usepackage[cm]{fullpage}
-\usepackage[round]{natbib}
-\bibliographystyle{abbrvnat}
-\usepackage{doi}
-\usepackage{bm}
-\usepackage[table]{xcolor}
-\hypersetup{
-    colorlinks,
-    linkcolor={blue},
-    citecolor={blue},
-    urlcolor={blue}
-}
-\usepackage{float}
-\restylefloat{table}
-\usepackage{siunitx}
-\DeclareSIUnit\year{yr}
-
-\begin{document}
-
-\section{Introduction}\label{sec:intro}
-This document includes a description of the code FALCON, with a presentation of the features implemented followed by results of extensive benchmarks listed in
-Table \ref{tab:list}, performed to verify that the code solves correctly a wide range of problems. Description and results of all benchmarks are shown
-in Section \ref{sec:benchmark}.
-
-\begin{table}[H]
-\caption{List of benchmarks performed for each feature of the code. Data of all benchmarks can be found at 
-\url{https://github.com/aleregorda/Benchmarks}}
-\centering
-\begin{tabular}{| l | l | l |}
-\toprule
- \textbf{Feature}  & \textbf{Benchmark} & \textbf{Section} \\
-\midrule
-  Solver  & Stokes flow & \ref{sec:stokes} \\
-\hline
-  \multirow{2}{*}{Markers advection} & Zalesak disk & \ref{sec:runge} \\
-    & Conservative Interpolation Velocity & \ref{sec:CVI} \\
-\hline
-  \multirow{4}{*}{Momentum equation}  & Poiseuille flow & \ref{sec:poiseuille}\\
-    & Instantaneous 2D sphere & \ref{sec:ist_sphere} \\
-    & Rayleigh-Taylor experiment & \ref{sec:rayleigh} \\
-    & Falling block & \ref{sec:block} \\
-\hline
-    & 2D Stokes sphere & \ref{sec:sphere} \\
-    Sticky air and  & Stabilisation algorithm & \ref{sec:stab} \\
-    free surface & Topography relaxation & \ref{sec:crameri} \\
-  & Spontaneous subduction & \ref{sec:subduction} \\
-\hline
-  Erosion and sedimentation  & - & - \\
-\hline
-   Non-linear  & Slab detachment & \ref{sec:slab} \\
-   visco-plasticity & Indenter & \ref{sec:indenter} \\
-    & Brick & \ref{sec:brick} \\
-\hline
-  \multirow{3}{*}{Energy equation}  & Advection stabilisation & \ref{sec:advection} \\
-    & Simple shear heating & \ref{sec:simple_shear} \\
-    & Shear and adiabatic heating & \ref{sec:shear} \\
-\hline
-\multirow{3}{*}{Energy + momentum} & Mantle convection & \ref{sec:mantle} \\
-& Viscoplastic mantle convection & \ref{sec:visco_mantle} \\
-& Thin layer entrainment & \ref{sec:thin} \\
-\hline
-  Phase changes  & \multirow{2}{*}{Hydrated sinking cylinder} & \multirow{2}{*}{\ref{sec:quinquis}} \\
-  and hydration  &  & \\
-\hline
-  Melting & Experimental melting curves & \ref{sec:katz} \\
-\bottomrule
-\end{tabular}
-\label{tab:list}
-\end{table}
-
-The code is written in Fortran90 and uses the finite element method (FEM) with quadrilateral $Q_1 \times P_0$ elements (continuous bilinear velocity and
-discontinuous constant pressure), associated with MUMPS\footnote{\url{http://mumps.enseeiht.fr/}} \citep{Amestoy2001,Amestoy2019} that is a software package
-for solving systems of linear equations of the form $A \cdot x = b$, where \textit{A} is a square sparse matrix, by means of a direct method. Since
-$Q_1 \times P_0$ elements do not satisfy Ladyzhenskaya, Babuska and Brezzi (LBB) stability condition \citep{Donea2003}, elemental pressure is elaborated
-in post-processing to avoid spurious pressures by means of a double interpolation to smooth the pressure. Following this procedure, the elemental pressure
-is interpolated onto nodes and then back onto elements. After the smoothing procedure, the pressure field is calculated again on the nodes to be used in
-combination with the lithostatic pressure, which is calculated onto nodes. Correctness of the solution and performances of the solver in terms of time and
-memory usage are tested solving the Stokes flow with the analytical solution proposed by \citet{Donea2003} (Section \ref{sec:stokes}).
-
-The thermo-mechanics of crust-mantle systems is described by means of the conservation of mass, momentum and energy equations, expressed as follows:
-\begin{eqnarray}
-\label{eq:mass}\frac{\partial \rho}{\partial t}+\bm{\nabla} \cdot (\rho \bm{u})&=&0\\
-\label{eq:momentum}-\bm{\nabla} p + \bm{\nabla} \cdot \bm{\tau} + \rho \bm{g}&=&0\\
-\label{eq:energy}\rho C_p \left(\frac{\partial T}{\partial t} + \bm{u} \cdot \bm{\nabla} T \right)&=&\bm{\nabla} \cdot \left(k\bm{\nabla} T\right) + H_{tot}
-\end{eqnarray}
-where $\rho$ is the density, $\bm{u}$ is the velocity, \textit{p} is the pressure, $\bm{\tau}$ is the deviatoric stress, $\bm{g}$ is the gravity acceleration,
-$C_p$ is the specific heat at constant pressure, \textit{T} is the temperature, \textit{k} is the thermal conductivity and $H_{tot}$ is the total internal heat
-production. Density variations due to temperature are generally small enough to assume the density as constant ($\rho=\rho_0$) in Eq. \ref{eq:mass} and in
-Eq. \ref{eq:energy}, while it must be treated as a variable in the buoyancy term of Eq. \ref{eq:momentum}, such that \[\rho=\rho_0(1-\alpha(T-T_0))\]
-where $\rho_0$ is the density at a reference temperature $T_0$ and $\alpha$ is the coefficient of thermal expansion. Eq. \ref{eq:mass} then can be rewritten as
-\begin{equation}\label{eq:mass_0}
-\bm{\nabla} \cdot \bm{u}=0
-\end{equation}
-This simplification is known as the extended Boussinesq approximation.
-
-The time step $dt$ is chosen in according to Courant-Friedrichs-Lewy condition \citep{Anderson1995}, such that
-\[dt=C_n\; \min\left(\frac{\Delta x}{\max|\bm{u}|},\frac{(\Delta x)^2}{\kappa}\right)\]
-where $0<C_n<1$ is the Courant (CFL) number, $\Delta x$ is the minimum dimension of smallest element, $\max|\bm{u}|$ is the maximum velocity calculated on the
-entire domain and
-$\kappa=\frac{k}{\rho C_p}$ is the thermal diffusivity \citep{Thieulot2014}.
-
-\section{Lagrangian markers}\label{sec:markers}
-Elemental properties (density, viscosity, thermal conductivity, specific heat and thermal expansion) needed to solve Eqs. \ref{eq:mass_0}, \ref{eq:momentum}
-and \ref{eq:energy} are related to the composition of each element, determined by means of Lagrangian markers that are characteristic of different materials
-in the domain. Elemental properties (with the sole exception of the viscosity) are calculated using an arithmetic mean, as \[P_e=\frac{1}{n}\sum_{i=1}^n P_i\]
-where $P_e$ is the elemental property, $P_i$ is the property characteristic of the material of each marker and $n$ is the number of markers of the element.
-Differently, the average scheme for the elemental viscosity can be chosen between harmonic, geometric and arithmetic mean. Markers can be placed regularly or
-randomly at the beginning of the simulation and their advection is performed by either a 2nd-order or a 4th-order Runge-Kutta in space, interpolating the velocity 
-field on each marker by means of the shape functions (see Section \ref{sec:runge}). The interpolated velocity is then corrected by means of the Conservative
-Velocity Interpolation (CVI), which introduces a corrective term to reduce dispersion and clustering of particles in both steady state and time-dependent 
-\citep{Wang2015} (see Section \ref{sec:CVI}).
-
-The initial distribution of the markers is created using the open-source code library
-Geodynamic World Builder \citep{Fraters2019}. During the simulation, each marker carries memory of temperature, pressure and accumulated strain, which are
-determined interpolating the relative nodal parameter. In particular, pressure is interpolated onto each marker after the smoothing procedure. The number
-of markers contained in each element is maintained between $n_{min}$ and $n_{max}$. When in an element there are less markers than $n_{min}$ the code adds
-random markers to reach the $n_{min}$, while if the number is higher than $n_{max}$ some of them are randomly deleted. When new markers are added, they assume
-the properties of the nearest marker. In this way elements are never empty and maintain a number of markers inside a prefixed range.
-
-\section{The momentum equation}\label{sec:mom_eq}
-The penalty method is implemented so that Eq. \ref{eq:mass_0} can be written as:
-\begin{equation}\label{eq:mass_penalty}
-\bm{\nabla} \cdot \bm{u} + \frac{p}{\lambda}=0
-\end{equation}
-where $\lambda$ is the so-called penalty parameter, which should be 6-7 orders of magnitude larger than the shear viscosity to ensure that mass conservation
-is satisfied \citep{Donea2003,Thieulot2014}.
-
-The deviatoric stress tensor in Eq. \ref{eq:momentum} can be written in terms of the strain rate tensor as $\bm{\tau}=2\eta\dot{\bm{\epsilon}}$, with
-$\dot{\bm{\epsilon}}=\frac{1}{2}\left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)$. Therefore, Eq. \ref{eq:momentum} can be rewritten as:
-\begin{equation}\label{eq:momentum_pressure}
--\bm{\nabla} p + \bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right) + \rho \bm{g} = 0
-\end{equation}
-Finally, using pressure from Eq. \ref{eq:mass_penalty}, Eq. \ref{eq:momentum_pressure} can be rewritten as:
-\begin{equation}\label{eq:momentum_penalty}
-\lambda \bm{\nabla} \left(\bm{\nabla} \cdot \bm{v}\right)+\bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right)+\rho\bm{g}=0
-\end{equation}
-The penalty method is been associated to the iterative Uzawa method, as described in detail in \citet{Dabrowski2008} and \citet{Thieulot2014}
-(see Section \ref{sec:poiseuille}). Results of benchmarks performed to verify the correctness in the implementation of Eq. \ref{eq:momentum_penalty} are
-shown in Sections \ref{sec:ist_sphere} and \ref{sec:rayleigh}.
-
-To support large viscosities variations the penalty parameter is related to the effective elemental viscosity by means of a dimensionless coefficient, so that
-$\lambda=\lambda_e(e)\eta_{eff}(e)$ \citep{Marotta2006,Dabrowski2008,Thieulot2014}. Benchmark of the falling block \citep{Gerya2003a,Gerya2010b,Thieulot2011}
-is performed to verify that the code can correctly deal buoyancy driven flows with strong viscosity contrasts (Section \ref{sec:block}).
-
-\section{Sticky air and free surface}\label{sec:sticky}
-The Earth's surface can be treated by means of either the so-called sticky-air or a true free surface method, both of them implemented in the code. In the
-sticky-air method the surface is approximated with the introduction of a buoyant layer with a viscosity at least four orders of magnitude lower than the crust
-\citep{Schmeling2008,Crameri2012} and the interface between lithosphere and air is defined using a chain of passive markers that are advected as the Lagrangian
-markers. The correctness of the evolution of the markers chain is tested with the experiment of a 2D time-dependent Stokes sphere below a free surface and
-compared with results from ASPECT\footnote{\url{https://aspect.geodynamics.org/}} \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} (Section
-\ref{sec:sphere}). In the true free surface case the top boundary is assumed stress-free and velocities are not fixed. In this case topography variations are
-described by vertical deformations of the mesh that depend on the velocity field of the nodes that identify the free surface, while horizontal deformations are
-not taken into account. This procedure is known as the Arbitrary Lagrangian-Eulerian (ALE) method and its implementation follows the technique described in
-\citet{Thieulot2011}. However, although the implementation of a true free surface better reproduces laboratory experiment, extremely small time steps can be
-necessary to maintain stability \citep{Kaus2010a,Quinquis2011,Thieulot2014}. Therefore, the stability algorithm proposed by \citet{Kaus2010a} is implemented
-to avoid instabilities due to high density differences at the free surface when using too large time steps. The implementation of this algorithm are tested by
-performing the experiment described by \citet{Kaus2010a} (Section \ref{sec:stab}).
-
-The topography relaxation benchmark proposed by \citet{Crameri2012} is performed to verify that the code correctly recovers topography variations in case of
-both the sticky-air and the true free surface method (Section \ref{sec:crameri}). Finally, Section \ref{sec:subduction} show the results of the spontaneous
-subduction experiment described by \citet{Schmeling2008}.
-
-\section{Erosion and sedimentation}\label{sec:erosion}
-Surface processes at either the lithosphere-air interface or in correspondence of the free surface have been implemented by means of the software
-Fastscape\footnote{\url{https://fastscape.org/}} \citep{Braun2013,Cordonnier2019,Yuan2019a,Yuan2019b}. FastScapeLib is a set of routines that solve the stream
-power law (enriched by a sediment transport/deposition term) \citep{Yuan2019a}, hillslope diffusion and marine transport and deposition \citep{Yuan2019b}.
-The partial differential equation solved by FastScapeLib is:
-\begin{equation}
-\frac{\partial h}{ \partial t}=U-K_f A^{m_e}S^{n_e} + \frac{G}{A}\int_A \left( U-\frac{\partial h}{\partial t} \right) dA + K_d \bm{\nabla}^2 h 
-\end{equation}
-where $h$ is the topography, $U$ is the uplift, $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $S$ is the slope, $m_e$ and $n_e$ are
-parameters that should be chosen to constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}, $G$ is the dimensionless
-deposition/transport coefficient for the enriched stream power law and $K_d$ is the effective diffusivity. Terms $K_f A^{m_e}S^{n_e}$ and $K_d \bm{\nabla}^2 h$
-consider long-range processes, described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, and short-range processes,
-proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}.
-
-In case of sticky air, slope is determined by means of the markers chain, which is vertically corrected considering erosion and sedimentation rates. After the
-correction, continental markers above the markers chain are transformed in air markers and air markers below the markers chain become sediments. Similarly,
-slope is determined considering the nodes on the top boundary in case of true free surface and they are vertically corrected as for the markers chain. In this
-case, continental markers above the top boundary are simply deleted.
-%Surface processes at the lithosphere-air interface are implemented taking into account fluvial erosion (long-range processes), hillslope processes (short-range
-%processes) and failure of slopes exceeding a threshold angle $\theta$ \citep{Burov1997,Tucker1999,Sacek2012,Croissant2014,Ueda2015,Beucher2020}. Erosion rates
-%$\dot{E}_{lr}$ due to long-range processes can be described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{lr}=K_f A^{m_e} \bm{\nabla} h^{n_e}
-%\end{equation}
-%where $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $\bm{\nabla} h$ is the slope and $m_e$ and $n_e$ are parameters chosen in to
-%constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}. Erosion rates $\dot{E}_{sr}$ due to short-range processes are
-%implemented by assuming that transport is linearly proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{sr}=K_d \bm{\nabla}^2 h
-%\end{equation}
-%where $K_d$ is the effective diffusivity. In addition, the slope is modified if it still exceeds a threshold angle after long- and short-range processes
-%\citep{Ueda2015}.
-
-%Offshore sedimentation processes are implemented similarly to hillslope processes but with a depth-dependent coefficient \citep{Sacek2012,Beucher2020}.
-%The sedimentation rate $\dot{S}$ is calculated as
-%\begin{equation}
-%\dot{S}=K_s \bm{\nabla}^2 h
-%\end{equation}
-%with
-%\begin{equation}
-%K_s=K^*_s \exp\left(\frac{h-h_{sl}}{h_0}\right)
-%\end{equation}
-%where $K^*_s$ is a constant, $h_{sl}$ is the sea level and $h_0$ is a characteristic length. In case of sticky air, slope is determined by means of the markers
-%chain, which is vertically corrected considering erosion and sedimentation rates. After the correction, continental markers above the markers chain are
-%transformed in air markers and air markers below the markers chain become sediments. Similarly, slope is determined considering the nodes on the top boundary
-%in case of true free surface and they are vertically corrected as for the markers chain. In this case, continental markers above the top boundary are simply
-%deleted.
-
-\section{Non-linear rheologies}\label{sec:nonlinear}
-Non-linear rheologies are implemented combining viscous creep (dislocation and diffusion) and plastic yielding. For each marker, diffusion and dislocation
-viscosity ($\eta_{df|ds}$) can be determined as follows \citep{Karato1993,Warren2008b,Wang2016}:
-\begin{equation}\label{eq:viscous}
-\eta_{df|ds}=f_s \left(\frac{d^m}{A}\right)^{\frac{1}{n}} \left (\frac{2^{\frac{1-n}{n}}}{3^{\frac{n+1}{2n}}} \right)
-I_2^{\frac{(1-n)}{n}}\exp\left(\frac{Q+pV}{nRT}\right)
-\end{equation}
-where $f_s$ is a scaling factor used to represent lithologies that are stronger or weaker than the base set, $d$ is the grain size, $m$ is the grain size exponent,
-$n$ is the stress exponent, $A$ is the uniaxial pre-exponential factor, $I_2$ is the square root of the second invariant of the strain rate tensor, $Q$ is the
-activation energy, $p$ is the pressure, $V$ is the activation volume, $R$ is the gas constant and $T$ is the temperature.
-Pressures and temperatures are determined by the interpolation of the nodal parameters, while strain rates are calculated by means of the
-derivative of the velocities on the nodes. In case of diffusion creep $n=1$ and $m>0$, in case of dislocation creep $n>1$ and $m=0$. The viscous creep $\eta_{cp}$
-is then calculated as the harmonic average between $\eta_{df}$ and $\eta_{ds}$:
-\begin{equation}\label{eq:visco_creep}
-\eta_{cp}=\left(\frac{1}{\eta_{df}}+\frac{1}{\eta_{ds}}\right)^{-1}
-\end{equation}
-The implementation of the non-linear viscous creep viscosity depending on the strain rate is tested by means of the slab detachment benchmark
-(Section \ref{sec:slab}).
-
-Plastic yielding is implemented rescaling $\eta_{cp}$ in order to limit the stress below the yield stress $\sigma_y$ \citep{Thieulot2008,Thieulot2014,Glerum2018},
-obtaining
-\begin{equation}\label{eq:plastic}
-\eta_{pl}=\frac{\sigma_y}{2I_2}
-\end{equation}
-where $\eta_{pl}$ is the plastic viscosity and the yield stress is determined following the Drucker-Prager criterion, such as
-\begin{equation}\label{eq:yield}
-\sigma_y=C\;\cos(\phi)+p\;\sin(\phi)
-\end{equation}
-where $C$ is the cohesion and $\phi$ is the internal friction angle. In case of negative pressure it is imposed equal to 0, so that negative yield stress are
-excluded. The correctness of the non-linear solution in case of plastic viscosity with variable internal friction angle is verified performing indenter and
-brick experiments (Sections \ref{sec:indenter} and \ref{sec:brick}, respectively).
-
-Strain softening is taken into account for both viscous creep and plastic viscosity \citep{Huismans2003,Babeyko2005,Huismans2005,Sobolev2005,Warren2008a} by
-means of the accumulated strain $\epsilon$ memorised by each marker. Softening in the viscous creep determines a linear decrease of $\eta_{vc}$ by means of a
-viscous strain softening factor $W_S$ that increases linearly from $W_{S_0}$ to $W_{S_{\infty}}$ for $\epsilon_{S_0}<\epsilon_S<\epsilon_{S_{\infty}}$ 
-\citep{Warren2008a}. This viscous softening can be related to strain-induced grain size reduction \citep{Warren2008a}. Differently, plastic softening is 
-simulated with a linear decrease of internal friction angle $\phi(\epsilon)$ and cohesion $C(\epsilon)$ in according to
-\begin{eqnarray}
-\label{eq:friction1}\phi(\epsilon)&=&\phi_0+(\phi_{\infty}-\phi_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}\\
-\label{eq:friction2}C(\epsilon)&=&C_0+(C_{\infty}-C_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}
-\end{eqnarray}
-where $\phi_0$, $C_0$ and $\phi_{\infty}$, $C_{\infty}$ are internal friction angle and cohesion for $\epsilon_0$ and $\epsilon_{\infty}$, respectively
-\citep{Huismans2003,Huismans2005,Warren2008a,Thieulot2014}. Plastic softening approximates deformation-induced softening of faults and brittle shear zones
-\citep{Warren2008a}.
-
-Plastic yielding and viscous creep are then combined to obtain a viscoplastic viscosity $\eta_{vp}$ as follows, assuming that they are independent processes
-\citep{Karato2008,Glerum2018}:
-\begin{equation}\label{eq:viscoplastic}
-\eta_{vp}=\min(\eta_{cp},\eta_{pl})
-\end{equation}
-Finally, effective viscosity $\eta_{eff}$ is capped by the minimum and the maximum viscosity ($\eta_{min}$ and $\eta_{max}$, respectively) to avoid extremely
-low or high viscosity \citep{Glerum2018} as follows
-\begin{equation}\label{eq:effective}
-\eta_{eff}=\min(\max(\eta_{vp},\eta_{min}),\eta_{max})
-\end{equation}
-Viscous creep and plastic yielding are non-linear rheologies because of their dependence on the velocity field through pressure and strain rates. Therefore,
-the solution is determined by means of Picard-type iterations, until convergence of the velocity field \citep{Glerum2018}. The convergence is verified at each
-iteration $i$ via the nonlinear residual $\bm{\mathcal{R}}^i$ that can be determined as
-\begin{equation}\label{eq:convergence}
-\bm{\mathcal{R}}^i= \mathbb{K} (\eta_{eff}(\dot{\bm{\epsilon}}^{i-1},p^{i-1}))\cdot \bm{v}^{i-1} - \bm{f}^i
-\end{equation}
-%\begin{equation}\label{eq:convergence}
-%\chi=1-\frac{\langle (\bm{v}^i-\langle\bm{v}^i\rangle)\cdot (\bm{v}^{i+1}-\langle\bm{v}^{i+1}\rangle) \rangle}{\sqrt{\sigma^i\sigma^{i+1}}}
-%\end{equation}
-%Iterations are performed until either convergence ($\chi<tol$) or maximal number of non-linear iterations ($it_{max}$) is reached.
-where $\mathbb{K}$ is the Stiffness matrix and $\bm{f}$ is the right hand side vector \citep{Spiegelman2016,Glerum2018}.  The $L_2$-norm is extracted from
-$\bm{\mathcal{R}}^i$ and it is normalised as follows
-\begin{equation}\label{eq:convergence_res}
-\frac{||\bm{\mathcal{R}}^i||_2}{||\bm{\mathcal{R}}^0||_2}
-\end{equation}
-where $||\bm{\mathcal{R}}^0||_2$ is the $L_2$-norm of the first nonlinear iteration.
-Since often the normalised nonlinear residual drop very quickly over the first 2-3 iterations, also the $L_2$-norm of velocity and pressure residuals are
-calculated as
-\begin{equation}\label{eq:convergence_vel}
-\frac{||\Delta\bm{v}||_2}{||\bm{v}||_2}=\frac{||\bm{v}_i-\bm{v}_{i-1}||_2}{||\bm{v}_{i}||_2}
-\end{equation}
-and
-\begin{equation}\label{eq:convergence_press}
-  \frac{||\Delta p||_2}{||p||_2}=\frac{||p_i-p_{i-1}||_2}{||p_{i}||_2}
-\end{equation}
-respectively. Iterations are performed until either maximal number of non-linear iterations ($it_{max}$) is reached or all the normalised nonlinear residual, 
-the velocity residual and the pressure residual converge under a defined tolerance $tol$.
-
-\section{The energy equation}\label{sec:en_eq}
-The stabilisation of the advection term of the energy equation (Eq. \ref{eq:energy}) needed to avoid possible oscillations in the thermal solution in the case
-when advection dominates over diffusion, is implemented by means of a streamline-upwind Petrov–Galerkin (SUPG) method, for which the advection term is modified
-follows the discussion in \citet{Thieulot2011} and \citet{Thieulot2014} (see Section \ref{sec:advection} for benchmark). The total internal heat production
-$H_{tot}$ in Eq. \ref{eq:energy} includes radiogenic heating $H_r$, shear heating $H_s$ and adiabatic heating $H_a$, where
-\begin{equation}\label{eq:shear}
-H_s=2\eta \dot{\bm{\epsilon}} : \dot{\bm{\epsilon}} = 2\eta(\dot{\epsilon}_{xx}^2+\dot{\epsilon}_{yy}^2+2\dot{\epsilon}_{xy}^2)
-\end{equation}
-and
-\begin{equation}\label{eq:adiabatic}
-H_a=T \alpha \frac{Dp}{Dt} \approx -\alpha T \rho g_y v
-\end{equation}
-The correctness of the implementation and the computation of shear and adiabatic heating is verified performing an exercise with an analytical solution
-(Section \ref{sec:simple_shear}) and Exercise 9.4 in \citet{Gerya2010b} (Section \ref{sec:shear}). Furthermore, mantle convection and thin layer entrainment
-benchmarks (Sections \ref{sec:mantle} and \ref{sec:thin}, respectively) are performed to verify that the code solve correctly Eqs. \ref{eq:momentum_penalty}
-and \ref{eq:energy} in case of a temperature field characterised by an initial perturbation and with a temperature-dependent density. At last, the viscoplastic
-mantle convection benchmark is performed to check also the viscoplasticity in case of a initial perturbed temperature field (Section \ref{sec:visco_mantle}).
-
-\section{Phase transitions and hydration}\label{sec:phase}
-Variations of effective density, specific heat and coefficient of thermal expansion during the evolution of crustal and mantle materials at different
-pressure-temperature ($p$-$T$) conditions are computed by means of version 6.8.6 of Perple\_X software package \citep{Connolly2005}, similarly to the
-implementation described in \citet{Marotta2020}. By default, mineral assemblages and properties of each lithology (oceanic crust, lower and upper continental
-crust, sediments and dry or hydrated mantle) are calculated for temperatures from \SIrange{330}{1600}{\kelvin}, with increments of approximately \SI{4}{\kelvin},
-and pressures from \SIrange{0.1}{30}{\giga\pascal}, with increments of approximately \SI{0.03}{\giga\pascal}, for a total of almost 100000 points.
-
-The introduction of phase transitions produces two main effects that must be taken into account in numerical models: variations in density and release/absorption
-of latent heat \citep{Tackley2010}. Effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into
-account considering effective coefficient of thermal expansion \citep{Christensen1985,Zhong2015}. Similarly, effects of latent heat in the energy equation
-(eq \ref{eq:energy}) are taken into account by considering effective specific heat and coefficient of thermal expansion
-\citep{Christensen1985,Tackley2010,Zhong2015}. Effective density, specific heat and coefficient of thermal expansion are included in Perple\_X files and are
-assumed by each Lagrangian marker in according to its $p$-$T$ conditions.
-
-Hydration processes related to dehydration of subducting lithosphere \citep{Schmidt1998,Liu2007,Faccenda2009,Faccenda2010,Faccenda2014,Rosas2016} strongly
-influence the thermo-mechanics inside the mantle wedge, mainly because of both weakening effects on the mantle rheology and density variations in case of
-mantle serpentinization \citep{Gerya2002,Honda2003,Arcay2005,Roda2010,Regorda2017}. In the case that hydration is switched on, the amount of bound and free
-water is memorised by each marker, following the implementation of \citet{Quinquis2014}; hydration and dehydration processes are related to the amount of bound
-water of each marker and to the maximum amount of water it can transport, i.e. if the amount of bound water exceeds the maximum amount of water, the marker
-dehydrates and releases free water that can hydrates under-saturated markers. Maximum water content of each marker is determined as function of lithology and
-$p$-$T$ conditions and it is calculated using Perple\_X, in the same way than effective density, specific heat and coefficient of thermal expansion. Bound water
-is advected together with the markers, neglecting the effect of bound water diffusion, while free water simply migrates vertically and is not coupled to the
-solid-phase flow of the mantle wedge \citep{Arcay2005,Quinquis2014}. The correctness of the migration scheme of the free water and its absorption into bound
-water by under-saturated markers are verified performing the experiment described in \citet{Quinquis2014} (Sec. \ref{sec:quinquis}).
-
-The mantle viscosity weakening related to the amount of water has been extensively studied \citep{Chopra1981,Hirth2003} and it is implemented in according to
-\citet{Arcay2005} and \citet{Horiuchi2016}, as follows
-\begin{equation}\label{eq:wet_visc}
-\eta_{wet}=\eta_{dry}\left(\left[1-\frac{1}{f_v}\right]\exp\left(-\frac{[OH^-]}{[OH^-]_0}\right)+\frac{1}{f_v}\right)
-\end{equation}
-where $f_v$ is the viscosity reduction factor between dry and wet conditions, $[OH^-]$ is the water content and $[OH^-]_0$ is a reference water content set to
-620 ppm (0.062 wt.\%) \citep{Arcay2005}. By default, $f_v$ is set to 100, in according to \citet{Arcay2005} and \citet{Horiuchi2016}.
-
-\section{Mantle melting}\label{sec:melt}
-Melt fraction in the mantle depends on temperature ($T$, in \textdegree C), pressure ($p$, in \si{\giga\pascal}) and the water content in the melt ($X_{H_2O}$, in wt.\%).
-The determination of the melt fraction $M$ is implemented as explained by \citet{Katz2003}, according to the successive modification by \citet{Langmuir2006} and
-\citet{Kelley2010}, as follows
-\begin{eqnarray}\label{eq:melt_fr}
-M(p,T,X_{H_2O})&=&-T+T_{sol}+(x_t\cdot ln(p+y_t))\cdot X_{H_2O}-\nonumber\\
-&&+K_m\left(\frac{X_{H_2O}^{bulk}}{D_{H_2O}(1-X_{H_2O})+X_{H_2O}}\right)^{\gamma}
-\end{eqnarray}
-where $T_{sol}$ is the temperature of the dry solidus, $D_{H_2O}$ is the partition coefficient and $x_t\cdot log(p+y_t)$ indicates the pressure dependence of
-the melting curve \citep{Kelley2010}. $x_t$, $y_t$, $K_m$ and $\gamma$ are parameters chosen in according to \citet{Kelley2010}.
-However, $X_{H_2O}$ depends on the melt fraction as
-\begin{equation}\label{eq:melt_wt}
-X_{H_2O}(M)=\frac{X_{H_2O}^{bulk}}{D_{H_2O}+M(1-D_{H_2O})}
-\end{equation}
-and a numerical solution can be found using a root-finder method \citep{Katz2003,Wang2016}. Furthermore, the content of water in the melt is limited by the
-pressure-dependent saturation concentration of water in the melt, determined as
-\begin{equation}\label{eq:melt_max}
-X_{H_2O}^{sat}=\chi_1 p^{\lambda_m}+\chi_2 p
-\end{equation}
-where $\chi_1$, $\chi_2$ and $\lambda_m$ are parameters chosen according to \citet{Katz2003}, \citet{Langmuir2006} and \citet{Kelley2010}. The correctness in
-the determination of melt fractions and water contents in the melt are tested comparing model results with experimental curves obtained by \citet{Katz2003}
-(Sec. \ref{sec:katz}).
-
-In case of melting, viscous creep viscosity ($\eta_{vp}$) is modified according to \citet{Wang2016}, as
-\begin{equation}\label{eq:melt_visc}
-\eta_{melt}=\eta_{vp}\cdot \exp(\alpha_m M)
-\end{equation}
-where $\alpha_m$ are melt fraction factors for dislocation and diffusion creep, chosen in according to \citet{Wang2016}.
-
-Effective density ($\rho_{eff}$) for partially molten rocks is calculated as
-\begin{equation}\label{eq:melt_dens}
-\rho_{eff}=\rho_s(1-M)+\rho_m M
-\end{equation}
-where $\rho_s$ and $\rho_m$ are the densities of the solid and the molten rock, respectively \citep{Gerya2004b,Gerya2007,Wang2016}. The density of the
-solid rock is extracted by Perple\_X, while the density of the molten portion is calculated as
-\begin{equation}\label{eq:melt_dens_M}
-\rho_m(p,T)=\rho_0[1-\alpha(T-T_0)][1+\beta(p-p_0)]
-\end{equation}
-where $\rho_0$ is the density at the reference temperature ($T_0$) and pressure($p_0$), and $\alpha$ and $\beta$ are the thermal expansion and compressibility
-coefficients, respectively \citep{Gerya2004b,Gerya2007,Wang2016}.
-
-As for the phase transitions, effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into account
-considering effective coefficient of thermal expansion. Similarly, latent heat due to melting/crystallisation can be implicitly included in the energy equation
-(Eq. \ref{eq:energy}) by considering effective specific heat ($C_{p_{eff}}$) and thermal expansion ($\alpha_{eff}$) of the partially molten rocks, in according to
-\begin{eqnarray}
-\label{eq:cp_effective}C_{p_{eff}}&=&C_p+H_L\left(\frac{\partial M}{\partial T}\right)_p \\
-\label{eq:alpha_effective}\alpha_{eff}&=&\alpha+\rho\frac{H_L}{T}\left(\frac{\partial M}{\partial p}\right)_T
-\end{eqnarray}
-where $C_p$ and $\alpha$ are the specific heat and the coefficient of thermal expansion of the solid rock, respectively, and $H_L$ is the latent heat of the
-molten rock \citep{Gerya2004b,Gerya2007,Tackley2010}.
-
-\section{Benchmarks}\label{sec:benchmark}
-\subsection{Error measurements}\label{sec:error}
-In order to determine the accuracy of velocity and pressure field of the benchmarks, the $L_2$-norm is computed by numerical integration on the Gauss-Legendre
-quadrature points. $L_2$-norm for pressure and velocity errors can be evaluated as
-\begin{eqnarray}
-\textrm{err}_p&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|p^n_i(r_q)-p^a_i(r_q)|^2\right)w_q|J_q|}\nonumber \\
-\textrm{err}_v&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|u^n_i(r_q)-u^a_i(r_q)|^2+|v^n_i(r_q)-v^a_i(r_q)|^2\right)w_q|J_q|}\nonumber
-\end{eqnarray}
-respectively, where $n_e$ is the number of elements, $n_q$ is the number of quadrature points per element, $p^n_i(r_q)$ and $p^a_i(r_q)$ are the numerical and
-analytical pressure, respectively, in each quadrature point $q$, $w_q$  and $J_q$ are the weight and the Jacobian at the quadrature point
-$q$, $u^n_i(r_q)$, $v^n_i(r_q)$, $u^a_i(r_q)$ and $v^a_i(r_q)$ are the numerical and analytical velocities, respectively, in each quadrature point $q$.
-
-Other quantities used as comparison with original benchmarks are the root-mean-square velocity over the whole domain and over the surface
-\begin{eqnarray}
-v_{\textrm{rms}}&=&\sqrt{\int_{\Omega}|\bm{v}|^2 d\Omega}\nonumber \\
-v_{\textrm{rms}}^{top}&=&\sqrt{\int_0^1 u^2 \bigg|_{y=1} dx}\nonumber
-\end{eqnarray}
-respectively, top and bottom Nusselt numbers
-\[\textrm{Nu}^{top/bottom}=-\int_0^1\frac{\partial T}{\partial y} \bigg|_{y=1/y=0} dx\]
-the average rate of work done against gravity
-\[\left\langle W \right\rangle=\int_{\Omega}T u_y d\Omega\]
-the average rate of viscous dissipation
-\[\left\langle \Phi \right\rangle=\int_{\Omega}\tau_{ij}\dot{\epsilon}_{ij} d\Omega\]
-and the percentage error between $\left\langle W \right\rangle$ and $\left\langle \Phi \right\rangle/Ra$
-\[\delta=\frac{\left|\left\langle W \right\rangle -\frac{\left\langle \Phi \right\rangle}{Ra} \right|}{\max \left(\left\langle W \right\rangle,
-\frac{\left\langle \Phi \right\rangle}{Ra}\right)}\times100\]
-
-\subsection{Stokes flow}\label{sec:stokes}
-The problem consists of determining velocity field $(u,v)$ and pressure $p$ in case of a manufactured solution with prescribed body forces such as
-\begin{eqnarray}
-b_1&=&(12-24y)x^4+(-24+48y)x^3+(-48y+72y^2-48y^3+12)x^2+\nonumber \\
-&&+(-2+24y-72y^2+48y^3)x+1-4y+12y^2-8y^3\nonumber \\
-b_2&=&(8-48y+48y^2)x^3+(-12+72y-72y^2)x^2+(4-24y+48y^2-48y^3+\nonumber \\
-&&+24y^4)x-12y^2+24y^3-12y^4\nonumber
-\end{eqnarray}
-for which the exact solution is:
-\begin{eqnarray}
-u(x,y)&=&x^2(1-x)^2(2y-6y^2+4y^3)\nonumber \\
-v(x,y)&=&-y^2(1-y)^2(2x-6x^2+4x^3)\nonumber \\
-p(x,y)&=&x(1-x)-\frac{1}{6}\nonumber
-\end{eqnarray}
-The domain is a unit square a constant viscosity ($\eta=1$) and the penalty parameter is set to $\lambda=10^7$. Velocity boundary conditions are set to no slip
-$(\bm{v}=\bm{0})$ on all boundaries. The problem is performed for different grid resolution between \num{8x8} and \num{1024x1024} elements.
-
-The errors between the analytical solution and the numerical prediction of the velocity and the pressure field are calculated as in Section \ref{sec:error}.
-Fig. \ref{fig:errors}a shows that both velocity and pressure field converge to the exact solution with the decrease of the element size, following the
-theoretical convergence rate. The convergence of pressure, in contrast with the observation by \citet{Donea2003} for $Q_1 \times P_0$ elements, indicate the
-effectiveness of the smoothing procedure. Smoothed pressure field for grid resolution of \num{128x128} elements is shown in Fig. \ref{fig:errors}b (continuous
-red line), in comparison with the analytical solution (dashed blue line). Solve times and memory usage needed to generate the solution are shown in
-Fig. \ref{fig:MUMPS}. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Solver/Stokes_Flow}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/errors.pdf}
-\caption{Velocity and pressure error for the Stokes flow experiment between generated and analytical solution as a function of element size (panel a),
-and comparison between smoothed pressure and analytical solution as function of $x$ coordinate for a grid resolution of \num{128x128} elements (panel b).}
-\label{fig:errors}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/MUMPS.pdf}
-\caption{Panel a) Analysis, factorisation, solution and total solve times (panel a) and factorisation memory usage (panel b) as a function of the total number
-of degrees of freedom for the Stokes flow experiment.}
-\label{fig:MUMPS}
-\end{figure}
-
-\subsection{Markers advection}\label{sec:runge}
-The 2nd-order Runge-Kutta advection scheme in space is tested by means of the Zalesak disk test \citep{Zalesak1979,Thieulot2014}. The benchmark is performed
-in a unit square domain with a grid resolution of \num{32x32} elements and 1000000 markers and values of Courant number between 0.25 and 1.
-At $t=0$, the disk is centred at position $(0.5;0.75)$ with a radius $R=0.15$ and has a vertical fissure 0.05 wide and 0.2 high (orange in Fig. \ref{fig:runge})
-and with the velocity field prescribed in the entire domain as
-\begin{eqnarray}
-u(x,y)&=&2\pi\left(y-\frac{L_x}{2}\right)\nonumber \\
-v(x,y)&=&-2\pi\left(x-\frac{L_x}{2}\right)\nonumber
-\end{eqnarray}
-the markers are back to their initial location after a $2\pi$ rotation ($t=1$). The clockwise rotation of the disk at $t=0.25$, 0.5 and 0.75 
-is shown in Fig. \ref{fig:runge} (red, blue and green, respectively). The distance from the centre of 1 marker is calculated throughout the rotation to
-evaluate the error for different values of Courant number. As expected, the error increase with the increasing of the time step, as shown in 
-Fig. \ref{fig:runge_err}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/Zalesak_Disk}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=200px]{./Figures/Zalesak.pdf}
-\caption{Position of the Zalesak disk at $t=0$, 0.25, 0.5 and 0.75 (orange, red, blue and green, respectively) for the case with Courant number of 1.}
-\label{fig:runge}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/zal_marker.pdf}
-\caption{Distance from the centre as function of time for values of Courant number of  of 0.25, 0.3, 0.5, 0.75 and 1 (green, orange, blue, black and red,
-respectively). R indicates the distance from the centre considering a perfect circle trajectory.}
-\label{fig:runge_err}
-\end{figure}
-
-\subsection{Conservative Velocity Interpolation (CVI)}\label{sec:CVI}
-The CVI correction is checked by means of the Stokes flow experiment by \citet{Donea2003}, as presented in \ref{sec:stokes}. The advection of Lagrangian
-markers is performed using either a 2nd-order or a 4th-order Runge-Kutta scheme with CFL=0.5 and an initial random distribution of 25 markers per element.
-Fig. \ref{fig:CVI} shows the minimum and maximum values of markers per element throughout the experiment with and without the CVI correction and using
-different Runge-Kutta schemes. The introduction of the CVI correction allows the
-tests to run for times longer than $t=2000$ without show empty elements, while tests without the CVI correction stop at approximately $t=600$, when at least
-one element goes empty. Fig. \ref{fig:CVI_time} shows the different number of markers per element and the markers distribution at $t=600$ obtained without and
-with the CVI correction (panels b and c, respectively) in case of a 2nd-order Runge-Kutta scheme. All data can be found at
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/CVI}.
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/CVI.pdf}
-\caption{Maximum and minimum number of markers per element throughout the experiment, with (green and orange lines) or without (purple and blue lines)
-the CVI correction and using either a 2nd-order (blue and orange lines) or a 4th-order (purple and green lines) Runge-Kutta scheme.}
-\label{fig:CVI}
-\end{figure}
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/CVI_time.pdf}
-\caption{Number of markers per element and markers distribution through time using 2nd.order Runge-Kutta scheme.
-Panels a: number of markers per element (panel a1) and distribution of the markers (panel a2) after the first
-iteration; panels b and c: comparison between number of markers per element (panels b1 and c1) and markers distribution (panels b2 and c2) without and with
-the CVI correction at $t=600$; panels d: number of markers per element (panel d1) and markers distribution (panel d2) in case of CVI correction at $t=2000$.}
-\label{fig:CVI_time}
-\end{figure}
-
-\subsection{Poiseuille flow}\label{sec:poiseuille}
-The domain is a rectangle with $L_x=2$ and $L_y=1$ and constant density and viscosity ($\rho=1$ and $\eta=1$), gravity acceleration $\bm{g}=\bm{0}$ and penalty
-parameter $\lambda=10^8$. The grid is composed by \num{40x20} elements. Velocity boundary conditions are set to no slip $(\bm{v}=\bm{0})$ at the top and the
-bottom, and a parabolic profile is imposed on the sides, with $u=y(1-y)$ and $v=0$. The analytical solution is then given by:
-\begin{eqnarray}
-u(x,y)&=&y(1-y)\nonumber \\
-v(x,y,)&=&0\nonumber \\
-p(x,y)&=&2\eta\left(\frac{L_x}{2}-x\right)\nonumber
-\end{eqnarray}
-
-The velocity field predicted by the model follows the expected parabolic profile (Fig. \ref{fig:poiseuille}a and Fig. \ref{fig:poi_plot}a) and, in case of the
-classic penalty method with no iterations, the pressure is clearly related to the divergence by means of the penalty parameter (Fig. \ref{fig:poiseuille}b and
-c and Fig. \ref{fig:poi_plot}b). Fig. \ref{fig:poiseuille}d shows that one Uzawa iteration is sufficient to bring the divergence down to \num{e-15}, with no
-correlation with the pressure field. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Poiseuille%20Flow}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Poiseuille.pdf}
-\caption{Velocity field (panel a), pressure (panel b) and divergence velocity of a Poiseuille flow in case of the classic penalty method (no iterations) and
-after one Uzawa iteration (panel c and d, respectively).}
-\label{fig:poiseuille}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Analytical.pdf}
-\caption{Velocity (panel a) and pressure (panel b) field predicted by the model for a Poiseuille flow with respect with their analytical solutions. The
-velocity field is plotted as function of the vertical coordinate in $L_x/2$ and the pressure is plotted as function of the x coordinate in $L_y/2$.}
-\label{fig:poi_plot}
-\end{figure}
-
-\subsection{Instantaneous 2D sphere}\label{sec:ist_sphere}
-The domain is a unit square with gravity $\bm{g}=(0,-1)$. The fluid has constant density and viscosity ($\rho_f=1$ and $\eta_f=1$). The sphere is in the middle
-of the domain with a radius R = 0.123456798 and has constant density and viscosity ($\rho_s=10^{-2}+\rho_f$ and $\eta_s=10^3 \cdot \eta_f$).
-We distinguish three different types of velocity boundary conditions:
-\begin{enumerate}
-\item FS: free slip conditions on all sides.
-\item NS: no slip conditions on all sides.
-\item OT: free slip conditions on the sides and the bottom, and open boundary at the top.
-\end{enumerate}
-All three cases are performed for different grid resolutions, between \num{16x16} and \num{512x512} elements with 50 randomly distributed markers per element,
-and different average schemes for the viscosity (harmonic, geometric and arithmetic). The velocity in the centre of the sphere $\bm{v}(0.5;0.5)$), minimum and
-maximum velocities ($u_{min}$, $u_{max}$, $v_{min}$, $v_{max}$) and pressures ($p_{min}$, $p_{max}$) on the entire domain, $v_{\textrm{rms}}$ and average pressure on the
-entire domain ($p_{avg}$) are compared with solutions generated by ASPECT \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and all results
-can be found at \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere2D} and 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Instantaneous_2D_Sphere}. Results in terms of the $v_{\textrm{rms}}$ for NS model 
-are also shown in Fig. \ref{fig:inst_sphere}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_NS.pdf}
-\caption{$v_{\textrm{rms}}$ as function of element size for the instantaneous 2D sphere experiment in case of no slip boundary conditions and with different average
-schemes for the viscosity. Results are compared with results obtained by other numerical codes.}
-\label{fig:inst_sphere}
-\end{figure}
-
-\subsection{Rayleigh-Taylor instability}\label{sec:rayleigh}
-This problem has been originally presented by \citet{vanKeken1997} and here it is performed as the isoviscous case (Case 1a). The domain has $L_x=0.9142$ and
-$L_y=1$ and gravity $\bm{g}=(0,-1)$. Two fluids with same constant viscosities ($\eta_1=\eta_2=1$) and different densities ($\rho_1=1000$ and $\rho_2=1010$),
-with the lighter fluid at the bottom. The initial interface between the fluids is given by $y(x)=0.2+0.02 \cos \left(\frac{\pi x}{L_x}\right)$. The experiment
-is performed with different grid sizes (\num{50x50}, \num{80x80}, \num{100x100} and \num{256x256}). A total of 1960000 markers are randomly distributed at the
-beginning of the simulation. Velocity boundary conditions are set to no slip at the top and the bottom, and to free slip at the sides of the domain.
-
-$v_{\textrm{rms}}$ as function of time is reported in Fig. \ref{fig:RT}, matching well with results shown by \citet{vanKeken1997}, \citet{Tackley2003} and
-\citet{Thieulot2014}. Fig. \ref{fig:rayleigh} shows the evolution of the experiment at different time steps. All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Rayleigh_Taylor_experiment/ISOVISCOUS}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/RT.pdf}
-\caption{$v_{\textrm{rms}}$ of the Rayleigh-Taylor experiment as function of time for different resolution of the grid.}
-\label{fig:RT}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=300px]{./Figures/Rayleigh.pdf}
-\caption{Evolution of the Rayleigh-Taylor experiment for a grid of \num{256x256} elements at $t=0, 500, 1000$ and $2000$ (panels a, b, c and d, respectively).}
-\label{fig:rayleigh}
-\end{figure}
-
-\subsection{Falling blocks}\label{sec:block}
-This benchmark is described as proposed by \citet{Gerya2003a}, \citet{Gerya2010b} and \citet{Thieulot2011}. The domain is square with $L_x=L_y=$ \SI{500}{\km}
-and the grid is composed by \num{50x50} elements with 25 markers in each element. The block is initially centred at ($x=$ \SI{250}{\km}; $y=$ \SI{400}{\km})
-and has a size of \num{100x100} km (Fig. \ref{fig:block}a). The fluid surrounding the block has $\eta_f=$ \SI{e21}{\pascal\per\s} and $\rho_f=$
-\SI{3200}{\kg\per\cubic\m}. The benchmark tests with different viscosities of the block, with $\eta_b$ from \SIrange{e15}{e27}{\pascal\per\s}. In each
-experiment also the density $\rho_b$ of the block varies from \SIrange{3220}{9900}{\kg\per\cubic\m}. Velocity boundary conditions are set to free slip
-conditions on all sides of the domain. Density distributions at $t=$ \SI{20}{\mega\year} for $\rho_b=$ \SI{3300}{\kg\per\cubic\m} and different $\eta_b$
-are plotted in Fig. \ref{fig:block}b-f, showing that the code correctly preserve the block geometry in case of large viscosity contrast, with a block stiffer
-than the surrounding fluid (Fig. \ref{fig:block}f).
-
-The velocity in the centre of the falling block at $t=0$ is measured for all experiments and, since the velocity should increase with density contrast, the
-quantity $\bm{v}/(\rho_b-\rho_f)$ is plotted as function of the viscosity contrast. All results perfectly match with those from \citet{Gerya2010b} and they
-line up on a single curve, demonstrating that the code can correctly deal with large viscosity and density contrasts (Fig. \ref{fig:falling}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Falling%20blocks}.
-
-\begin{figure}
-\noindent\includegraphics[width=300px]{./Figures/Block.pdf}
-\centering
-\caption{Density evolution of the falling block experiment at $t=$ \SI{0}{\mega\year} (panel a) and $t=$ \SI{20}{\mega\year} for different viscosities of the
-block (panels b-f).}
-\label{fig:block}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Falling.pdf}
-\caption{Initial velocity relative to the density contrast at the centre of the falling block as function of the viscosity contrast between the block and the
-surrounding fluid.}
-\label{fig:falling}
-\end{figure}
-
-\subsection{2D time-dependent Stokes sphere with deformable free surface}\label{sec:sphere}
-This experiment is performed in a unit square domain with the gravity acceleration fixed to $g_y=-1$. The sphere has $\rho_s=2$ and $\eta_s=10^3$ and is
-initially centred at ($0.5;0.6$) with radius $R=0.123456789$. The fluid surrounding the sphere has $\rho_f=1$ and $\eta_f=1$ and occupies the domain for
-$y\leq 0.75$, while for $y>0.75$ the air has $\rho_a=0$ and $\eta_a=10^{-3}$. Velocity boundary conditions are set to free slip on all sides. The Courant
-number is set to 0.25. The experiments run for 200 s using grid resolutions from \num{150x150} to \num{512x512} elements, with 25 randomly distributed markers
-per element, and for different average schemes for the viscosity. The interface between the fluid and the air is tracked by means of the markers chain. Results
-in terms of velocity and pressure fields and topography variations are compared with results obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and they can be found at
-\url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere_fs2D} and
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Time_dependent_sphere}. Results in terms of the $v_{\textrm{rms}}$ in case of 
-an arithmetic average are also shown in Fig. \ref{fig:inst_spherefs}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_arithm.pdf}
-\caption{$v_{\textrm{rms}}$ for different grid resolutions as function of time for the 2D time-dependent Stokes sphere experiment in case of an arithmetic average.
-Results are compared with results obtained by ASPECT.}
-\label{fig:inst_spherefs}
-\end{figure}
-
-\subsection{Free surface stabilisation}\label{sec:stab}
-The benchmark is performed as discussed in \citet{Kaus2010a} and \citet{Thieulot2014}. The domain is square with $L_x=L_y=$ \SI{500}{\km} and a grid resolution
-of \num{200x200} elements, each of them containing 16 markers. A buoyant fluid with $\rho_1=$ \SI{3200}{\kg\per\cubic\metre} and $\eta_1=$ \SI{e20}{\pascal\s}
-is overlain by a denser fluid with $\rho_2=$ \SI{3300}{\kg\per\cubic\metre} and $\eta_2=$ \SI{e21}{\pascal\s}. The initial interface between the fluids has an
-initial sinusoidal shape of 5 km amplitude. Velocity boundary conditions are set to free slip conditions at the sides and no slip at the bottom, while the top
-is free surface. The experiment is performed with various fixed time steps, without the use of the Courant number. The vertical position of the free surface at
-$x=L_x$ is tracked for each simulation, with or without the stabilisation algorithm.
-
-The results show an instability (drunken sailor effect) increasing the time step in case that the stabilisation algorithm is not activated
-($dt=$ \SI{4200}{\year}, green line in Fig. \ref{fig:stab}), as already observed by \citet{Kaus2010a} and \citet{Thieulot2014}. Activating the algorithm the
-instability is fixed and the simulation remains stable using time step up to \SI{50000}{\year} (red line in Fig. \ref{fig:stab}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Stabilisation_algorithm}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Kaus_random.pdf}
-\caption{Evolution of the y coordinate in $x=L_x$ as function of time for different time steps, with and without the stabilisation algorithm.}
-\label{fig:stab}
-\end{figure}
-
-\subsection{Topography relaxation}\label{sec:crameri}
-This experiment is performed as in \citet{Crameri2012} using free surface and sticky air (Case 1). The domain is rectangular with $L_x=$ \SI{2800}{\km} and
-$L_y=$ \SI{700}{\km} in case of free surface and $L_y=$ \SI{800}{\km} in case of sticky air ($h_{st}=$ \SI{100}{\km}). The grid is composed by \num{256x64}
-elements for the free surface \num{560x320} elements for the sticky air. For both cases, there is a mantle of 600 km thickness with $\eta_m=$ \SI{e21}{\pascal\s}
-overlain by a highly viscous cosine-shaped lithosphere with thickness between 93 and 107 km and $\eta_l=$ \SI{e23}{\pascal\s}. Both mediums have $\rho=$
-\SI{3300}{\kg\per\cubic\metre}. In case of sticky air experiment, the air has $\rho_a=$ \SI{0}{\kg\per\cubic\metre} and $\eta_a=$ \SI{e18}{\pascal\s}. The
-gravity acceleration is set to $g_y=$ \SI{-10}{\metre\per\square\s}. Velocity boundary conditions are set to no slip at the bottom and free slip at the sides
-of the domain. In case of sticky air, velocities on top are set to free slip conditions.
-
-The maximum topography as function of time ($h(t)$) can be analytically derived in according to \[h(t)=h_0 \exp(\gamma t)\]
-where $h_0=7$ km is the initial topography, $\gamma=\num{-0.2139e-11}$ is the characteristic relaxation rate and $t$ is the time (black line in Fig.
-\ref{fig:crameri}). Maximum topography at the characteristic relaxation time $t_{rlx}=$ \SI{14.825}{\kilo\year} can be found to be $h_{rlx}=$ \SI{2576}{\m}.
-The results show that a viscosity of the air of \SI{e18}{\pascal\s} is needed in case of $h_{st}=$ \SI{100}{\km} to correctly describe the topography relaxation
-for this problem, as pointed out by \citet{Crameri2012} (purple line in Fig. \ref{fig:crameri}). The case with a true free surface well follows the analytical
-solution (orange line in Fig. \ref{fig:crameri}). In particular, at the relaxation time in case of the true free surface predicts a maximum topography of
-\SI{2570}{\m}, with an error of \SI{6}{\m} with respect to the analytical solution. Results are also compared with results obtained with a free surface in
-UNDERWORLD (\num{256x64} elements) and SULEC (\num{401x201} elements) (with courtesy from \citealp{Crameri2012}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Topography_relaxation}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Crameri.pdf}
-\caption{Maximum topography as function of time for the Crameri benchmark (coloured lines) in comparison with the analytical solution (black line) and results
-obtained with UNDERWORLD (yellow line) and SULEC (blue line).}
-\label{fig:crameri}
-\end{figure}
-
-\subsection{Spontaneous subduction}\label{sec:subduction}
-This experiment is performed as presented by \citet{Schmeling2008}, in case of both a sticky air and a true free surface and using different average schemes
-for the viscosities. The domain is rectangular with $L_x=$ \SI{3000}{\km} and $L_y=$ \SI{700}{\km}, a grid resolution of \num{256x64} elements with 50 markers
-per element and Courant number of 0.01. At the beginning of the simulation, a 100 km-thick lithospheric layer with $\rho_m=$ \SI{3300}{\kg\per\cubic\metre}
-and $\eta_m=$ \SI{e23}{\pascal\s} is located at the top of the domain for $x$ between 1000 and 3000 km, while the asthenospheric mantle has $\rho_m=$
-\SI{3200}{\pascal\s} and $\eta_m=$ \SI{e21}{\pascal\s}. In addition, a 200 km-depth lithospheric slab is already subducted in the mantle in order to have
-a spontaneous subduction. In case of sticky air, $L_y=750$ km and a 50 km-thick air layer with $\eta_a$ from \SIrange{e19}{e21}{\pascal\s} overlie the mantle.
-Velocity boundary conditions are set to free slip on the sides and on the bottom of the domain. In case of sticky air, velocities on the top boundary are set
-to free slip conditions as well.
-
-The maximum depth of the slab is tracked for all the simulations (Fig. \ref{fig:subduction}) and they are compared with results obtained by
-\citet{Schmeling2008} with models with regular grids and comparable grid resolutions (rectangles in Fig. \ref{fig:subduction}). Low viscous sticky air models
-(i.e., \SI{e19}{\pascal\s}) enlighten a strong dependence of the sinking velocity with the chosen average scheme (dashed lines in Fig. \ref{fig:subduction}),
-while high viscous sticky air models (i.e., \SI{e21}{\pascal\s}) show a higher resistance at the trench, underestimating the correct solution (dotted lines in
-Fig. \ref{fig:subduction}). A high resistance at the trench can be also observed in case of a true free surface for the low resolution grid used (continuous
-lines in Fig. \ref{fig:subduction}), as pointed out by \citet{Schmeling2008}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Spontaneous_subduction}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Subduction.pdf}
-\caption{Maximum depth of the slab as function of time for all the simulations of the \citet{Schmeling2008} benchmark, in case of harmonic, geometric and
-arithmetic means (red, green and blue lines, respectively) using a sticky air or a true free surface (discontinuous and continuous lines, respectively).
-Red, green and blue rectangular areas indicate the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{500}{\km} in case of sticky air,
-while the grey rectangular area indicates the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{600}{\km} in case of true free surface.}
-\label{fig:subduction}
-\end{figure}
-
-\subsection{The slab detachment}\label{sec:slab}
-The slab detachment benchmark is performed as described by \citet{Schmalholz2011} and \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{1000}{\km}
-and $L_y=$ \SI{660}{\km} and a grid resolution of \num{256x256} elements. A non-linear viscous T-shaped layer with $\rho_l=$ \SI{3300}{\kg\per\cubic\metre} is
-placed at the top of the domain and surrounded by a linear viscous fluid with $\rho_f=$ \SI{3150}{\kg\per\cubic\m}. The top layer is 80 km-thick and a 250
-km-long and 80 km-wide slab is placed at $x=L_x/2$ (Fig. \ref{fig:slab}a). The effective viscosity of the top layer is given by
-\[\eta_{eff}=\eta_0 I_2^{\frac{(1-n)}{n}}\]
-with $\eta_0=$ \SI{4.75e11}{\pascal\s} and $n=4$, while the surrounding fluid has $\eta_f=$ \SI{e21}{\pascal\s}. The effective viscosity is capped between
-$\eta_{min}=$ \SI{e21}{\pascal\s} and $\eta_{max}=$ \SI{e25}{\pascal\s}. Velocity boundary conditions are set to free slip at the top and the bottom, and to
-no slip at the sides of the domain.
-
-The sinking of the slab determines a decrease of the effective viscosity in the T-shaped layer during the first part of the evolution (Fig. \ref{fig:slab}a),
-as consequence of local strain rates, while effective viscosities increase up to $\eta_{max}$ in the top layer after approximately \SI{20}{\mega\year}, when
-necking is complete (Fig. \ref{fig:slab}b).
-
-The width of necking is tracked by means of markers position at the side of the slab. Necking width and time are normalised in function of the initial slab
-(\SI{80}{\km}) and the characteristic time ($t_c=$ \SI{7.1158e14}{\s}) \citep{Schmalholz2011,Glerum2018}. The results in case of arithmetic, geometric and
-harmonic mean of the viscosity are shown in Fig. \ref{fig:necking}, compared with results from \citet{Glerum2018} for which an infinity norm mean was used.
-The influence of the average schemes agrees with results shown in Section \ref{sec:subduction}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Slab_detachment}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Slab.pdf}
-\caption{Effective viscosity for the slab detachment benchmark at the beginning of the evolution (panel a) and when necking is complete (panel b).}
-\label{fig:slab}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Necking.pdf}
-\caption{Normalised width of the necking of the slab detachment benchmark as function of normalised time.}
-\label{fig:necking}
-\end{figure}
-
-\subsection{The indenter experiment}\label{sec:indenter}
-The indenter benchmark simulate a rigid punch on a purely plastic von Mises material, for which exists an analytical solution
-\citep{Thieulot2008,Thieulot2014,Glerum2018}. The domain is a unit square with a grid resolution of \num{256x256} elements. The gravity acceleration is set
-to $\bm{g}=\bm{0}$ and effective viscosity is capped using $\eta_{min}=10^{-4}$ and $\eta_{max}=10^3$. The tolerance for the convergence of the non-linear
-solution is $tol=10^{-9}$, with a maximum number of non-linear iterations set to 500. The medium in the domain has $\rho=0.01$, an initial viscosity
-$\eta_0=10$, cohesion $C=1$ and an angle of internal friction $\phi=0$°. Velocity boundary conditions are set to no slip at the bottom and free slip at the
-sides of the domain. The top of the domain is open with the exception of the central portion (punch area) with width $w_p=0.125$, where $v=-1.05$ and $u$ is
-fixed either to 0 (rough punch experiment) or free (smooth punch experiment).
-
-Fig. \ref{fig:indenter} show the results in terms of viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure
-(panels g and h) for both experiments. The analytical solution indicate that pressure at ($0.5;1$) and ($0.5 \pm w_p;1$)is $p=\pi +1$ and $p=1$, respectively
-(black and red lines in Fig. \ref{fig:smooth_p}). Fig. \ref{fig:smooth_p} shows a clear improvement in the pressure solution when passing from a rough to a
-smooth punch (green and blue lines in Fig. \ref{fig:smooth_p}, respectively), as previously observed by \citet{Thieulot2014} and \citet{Glerum2018}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Indenter}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Indenter.pdf}
-\caption{Viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure (panels g and h) fields for rough (left column) and
-smooth (right column) punch experiments.}
-\label{fig:indenter}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Smooth_Rough.pdf}
-\caption{Nodal pressure in function of the x coordinate at $y=1$ for rough (green line) and smooth (blue line) punch experiment. Black and red lines indicate
-the analytical solution at $x=0.5$ and $x=0.5 \pm w_p$, respectively.}
-\label{fig:smooth_p}
-\end{figure}
-
-\subsection{The brick experiment}\label{sec:brick}
-An instantaneous version of the brick benchmark is performed to verify the correctness of the pressure-dependent plasticity for different angle of internal
-friction, as proposed by \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{40}{\km} and $L_y=$ \SI{10}{\km} and a grid resolution of \num{512x128}
-elements. The gravity acceleration is set to $g_y=$ \SI{-10}{\m\square\s} and effective viscosity is capped using $\eta_{min}=$ \SI{e19}{\pascal\s} and
-$\eta_{max}=$ \SI{e26}{\pascal\s}. The tolerance for the convergence of the non-linear solution is $tol=10^{-7}$, with a maximum number of non-linear iterations
-set to 1000. A 800 m-wide and 400 m-height inclusion with $\eta_b=$ \SI{e20}{\pascal\s} and $\rho_b=$ \SI{2700}{\kg\per\cubic\m} is placed at the bottom of the
-domain at $x=L_x/2$. The inclusion is surrounded by a non-linear viscous medium with $\rho_m=$ \SI{2700}{\kg\per\cubic\m}, an initial viscosity of
-\SI{e23}{\pascal\s}, a linear viscous viscosity of \SI{e25}{\pascal\s} and a cohesion of \SI{40}{\mega\pascal}. Velocities are set to free slip conditions at
-the bottom of the domain and the top is open. Velocities on sides of the domain are fixed to $u=$ \SI{+-2e-11}{\m\per\s} and $v=0$. The experiment is performed
-in compressional and extensional contexts with an angle of internal friction $\phi$ of the non-linear medium variable between 0° and 30°.
-
-As expected, two shear bands stem from the inclusion with variable angles in relation to both the dynamics context and the internal friction angle. Shear band
-angles formed at 45° for $\phi=0$° in both extensional and compressional contexts (Fig. \ref{fig:brick_beam}a), while they are different in case of $\phi\neq0$°.
-Shear band angles in case of $\phi=20$° are shown for compression and extension in Fig. \ref{fig:brick_beam}b  and c, respectively. Values of shear band angles
-as function of internal friction angles from 0° to 30° are extracted at different depths and minimum and maximum values are plotted in Fig. \ref{fig:brick},
-in comparison with theoretical Roscoe, Arthur and Coulomb shear band angles. For all tests the tolerance for nonlinear convergence is defined $tol=10^{-7}$
-and none converges before the maximum number of iterations ($it_{max}=1000$) is reached. However, tests with internal friction angles up to 15° show a constant
-decrease of the velocity residuals, which can not be observed for for tests with higher internal friction angles (Fig. \ref{fig:convergence}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Brick_experiment}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Brick_Beam.pdf}
-\caption{Shear band angles predicted for the brick experiment for $\phi=0$° (panel a) and for $\phi=20$° in case of compression and extension (panels b and c,
-respectively).}
-\label{fig:brick_beam}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Brick.pdf}
-\caption{Shear band angles predicted for the brick experiment in case of compressional and extensional contexts as function of different internal angle of
-friction (continuous black line and red dots), compared with theoretical Roscoe, Artur and Coulomb angles (discontinuous black lines). Red lines indicate the
-range of angles calculated at different depths.}
-\label{fig:brick}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \noindent\includegraphics[width=500px]{./Figures/convergence.pdf}
-  \caption{Convergence rates for all tests. Normalised residuals (left column) are compared with velocity residuals (right column) for extensional (first row)
-  and compressional (second row) settings.}
-  \label{fig:convergence}
-  \end{figure}
-
-\subsection{Advection stabilisation}\label{sec:advection}
-The 1D advection problem proposed by \citet{Donea2003} and \citet{Thieulot2011} is performed to verify the effectiveness of the SUPG method to stabilise the
-advection term of the energy equation. The domain is a 1D segment with $L_x=1$ composed by 50 elements and a discontinuity in the thermal field placed at
-$x=0.25$. Temperature is set to 1 for $x \leq 0.25$ and to 0 for $x > 0.25$. Velocity is set to ${u}=1$ in the entire domain. The simulation is performed for
-250 time steps, with $dt=0.002$, so the thermal discontinuity should be at $x=0.75$ at the end of the simulation.
-
-Temperature profiles at the end of the simulation are shown in \ref{fig:advection} as function of the dimensionless coefficient $\gamma=\tau \bm{v}/h$. In case
-of the classic Galerkin method ($\gamma=0$, blue line in \ref{fig:advection}) the final thermal profile is characterised by strong oscillations, which are
-eliminated in case of the SUPG method ($\gamma=0.045$, orange line in \ref{fig:advection}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Advection_stabilisation}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Advection.pdf}
-\caption{Temperature profile as function of \textit{x} for the advection stabilisation benchmark. Purple line indicates the initial temperature profile; the
-green line indicates the analytical temperature profile after 250 time steps; blue line indicates the temperature profile after 250 time steps in case of the
-classic Galerkin method ($\gamma=0$); orange line indicates the temperature profile after 250 time steps in case of the SUPG method ($\gamma=0.045$).}
-\label{fig:advection}
-\end{figure}
-
-\subsection{Shear heating}\label{sec:simple_shear}
-This exercise is performed in an unit square domain composed by \num{128x128} elements. The velocity field is prescribed on the entire domain with
-$\bm{v}=(L_y-y)y\bm{e_x}$; viscosity, density and specific heat are fixed to 1, while thermal conductivity and radiogenic energy are fixed to 0. Therefore,
-the energy equation (Eq. \ref{eq:energy}) can be simplified as \[\frac{\partial T}{\partial t}=H_s\] and fixing $T(t=0)=0$, the temperature can be find as
-\[T(t)=H_s t\]
-In this case we have
-\begin{eqnarray}
-\dot{\epsilon_{xx}}&=&\frac{\partial u}{\partial x}=0 \nonumber \\
-\dot{\epsilon_{yy}}&=&\frac{\partial v}{\partial y}=0 \nonumber \\
-\dot{\epsilon_{xy}}&=&\frac{1}{2}\left(\frac{\partial v}{\partial x}+\frac{\partial u}{\partial y}\right)=\frac{1}{2}(L_y-2y)\nonumber
-\end{eqnarray}
-and, simplifying Eq. \ref{eq:shear}, shear heating can be calculated as \[H_s(x,y)=(1-2y)^2\]
-
-The solution predicted by the model in terms of velocity, temperature and shear heating match well with the analytical solutions (Fig. \ref{fig:analytical_en}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Simple_shear}.
-
-%\begin{figure}
-%\centering
-%\noindent\includegraphics[width=400px]{./Figures/Energy_Thieulot.pdf}
-%\caption{Temperature and shear heating predicted by the model after 1 time step.}
-%\label{fig:en_thieulot}
-%\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=500px]{./Figures/Analytical_Shear.pdf}
-\caption{Velocity (panel a), temperature (panel b) and shear heating (panel c) as function of the \textit{y} coordinate for the simple shear experiment. The
-solutions predicted by the model (red lines) are compared with the analytical solutions (dashed blue lines).}
-\label{fig:analytical_en}
-\end{figure}
-
-\subsection{Shear and adiabatic heating}\label{sec:shear}
-This problem is performed as presented in Exercise 9.4 in \citet{Gerya2010b}. Two materials are vertically separated in a rectangular domain with $L_x=$
-\SI{1000}{\km}, $L_y=$ \SI{1500}{\km} and a grid resolution of \num{30x20} elements. Constant thermal coefficient expansion ($\alpha=$ \SI{3e-5}{\per\kelvin}),
-temperature ($T=$ \SI{1300}{\kelvin}) and gravity acceleration ($g_y=$ \SI{-10}{\m\square\s}) are assumed in the whole domain. Fluid 1 (on the left side of the
-domain) has $\rho_1=$ \SI{3200}{\kg\per\cubic\m} and $\eta_1=$ \SI{e20}{\pascal\s}; fluid 2 (on the right side of the domain) has $\rho_2=$
-\SI{3300}{\kg\per\cubic\m} and $\eta_2=$ \SI{e22}{\pascal\s}. Velocity boundary conditions are set to free slip on all sides of the domain.
-
-As shown in Fig. \ref{fig:energy}, both shear and adiabatic heating predicted by the code (first row in Fig. \ref{fig:energy}) well recreate the results
-obtained by means of example ${Shear\_adiabatic\_heating.m}$ from \citet{Gerya2010b} (second row in Fig. \ref{fig:energy}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Adiabatic%2BShear}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Energy.pdf}
-\caption{Comparison between shear (first column), adiabatic (second column) and total (third column) energy predicted by the code (panels a, b and c) and those
-created using example ${Shear\_adiabatic\_heating.m}$ from Exercise 9.4 in \citet{Gerya2010b} (panels d, e and f).}
-\label{fig:energy}
-\end{figure}
-
-\subsection{Mantle convection}\label{sec:mantle}
-This problem is performed as presented by \citet{Blankenbach1989} (constant viscosity cases) and \citet{Thieulot2014}, in a 2D unit square domain with gravity
-acceleration $g_y=10^{10}Ra$. The experiment is performed with three different Rayleigh numbers ($Ra=10^4$, $10^5$ and $10^6$) and with different grid
-resolution (between \num{32x32} and \num{128x128} elements). The fluid has constant viscosity, initial density, heat capacity, thermal conductivity
-($\eta= \rho_0= C_p = k=1$), reference temperature ($T_0=0$) and thermal expansion coefficient ($\alpha=10^{-10}$). Temperatures are set to 0 on top and 1 on
-bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\]
-
-The solution generated by the code in terms of $v_{\textrm{rms}}$ (calculated as in Sec. \ref{sec:error}) and Nu as function of time are reported for all the
-simulations in Fig. \ref{fig:mantle}. At the steady state, $v_{\textrm{rms}}$ and Nu of all simulations converge well toward the values from \citet{Blankenbach1989},
-with lower errors for higher resolution grids (Table \ref{tab:mantle}). All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Mantle_convection}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Convection.pdf}
-\caption{$v_{\textrm{rms}}$ (panels a, c and e) and Nu (panels b, d and f) for the mantle convection benchmark as function of time for different grid resolution.
-Panels a and b show the results for $Ra=10^4$; Panels c and d show the results for $Ra=10^5$; panels e and f show the results for $Ra=10^6$. Purple lines
-indicate the convergence values for the $v_{\textrm{rms}}$ and the Nu from \citet{Blankenbach1989} at the steady state.}
-\label{fig:mantle}
-\end{figure}
-
-\begin{table}[H]
-\caption{Comparison between $v_{\textrm{rms}}$ and Nu predicted by the code for the mantle convection experiment and same values as reported in literature.}
-\centering
-\small
-\begin{tabular}{|>{\centering}m{0.09\textwidth}|>{\centering}m{0.05\textwidth}|>{\centering}m{0.24\textwidth}|>{\centering}m{0.15\textwidth}|
-  >{\centering}m{0.1\textwidth} >{\centering}m{0.1\textwidth} >{\centering\arraybackslash}m{0.1\textwidth}|}
-\toprule
-& & \multirow{2}{*}{\citet{Blankenbach1989}} & \citet{Thieulot2014} & \multicolumn{3}{c|}{FALCON}  \\
- & & & \num{200x200} & \num{32x32} & \num{64x64} & \num{128x128}  \\
-\midrule
-\multirow{2}{*}{$Ra=10^4$} & $v_{\textrm{rms}}$ & $42.864947 \pm 0.000020$ & 42.867 & 42.83226 & 42.852793 & 42.861394  \\
-           & Nu      & $4.884409 \pm 0.000010$  & 4.882  & 4.781297 & 4.857475  & 4.877573   \\
-\hline
-\multirow{2}{*}{$Ra=10^5$} & $v_{\textrm{rms}}$ & $193.21454 \pm 0.00010$  & 193.255& 193.872643 & 193.377472 & 193.252290  \\
-           & Nu      & $10.534095 \pm 0.000010$ & 10.507 & 9.602514 & 10.270735  & 10.465629   \\
-\hline
-\multirow{2}{*}{$Ra=10^6$} & $v_{\textrm{rms}}$ & $833.98977 \pm 0.00020$  & 834.712& 848.091176 & 837.767911 & 834.945793  \\
-           & Nu      & $21.972465 \pm 0.000020$ & 21.695 & 15.999266 & 19.703682  & 21.306939   \\
-\bottomrule
-\end{tabular}
-\label{tab:mantle}
-\end{table}
-
-\subsection{Viscoplastic mantle convection}\label{sec:visco_mantle}
-This benchmark is performed as Cases 1-5a described in \citet{Tosi2015} in a 2D unit square domain with two different grid resolutions of \num{32x32} 
-and \num{100x100} elements. Reference viscosity, reference density, heat capacity, thermal conductivity and thermal expansion coefficient are fixed to a constant
-value of 1 ($\eta_0= \rho_0= C_p = k=\alpha=1$), while the gravity has been chosen as $g_y=10^2$, in order to obtain a Rayleigh number $Ra=10^2$. Temperatures
-are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\].
-The viscosity field $\eta$ is calculated as the harmonic average between a linear part $\eta_{lin}$ that depends on temperature only or on temperature and 
-depth $y$ and a nonlinear-plastic part $\eta_{pl}$ that depends on the strain rate $\dot{\bm{\epsilon}}$, as follows
-\[\eta(T,y,\dot{\bm{\epsilon}})=2\left(\frac{1}{\eta_{lin}(T,y)}+\frac{1}{\eta_{pl}(\dot{\bm{\epsilon}})}\right)^{-1}\]
-The linear and the nonlinear-plastic parts of the viscosity are calculated following \citet{Tosi2015} as follows
-\[\eta_{lin}(T,y)=\exp(-\ln(\Delta\eta_T)T+\ln(\Delta\eta_y)y)\]
-\[\eta_{pl}(\dot{\bm{\epsilon}})=\eta^*+\frac{\sigma_Y}{\sqrt{\dot{\bm{\epsilon}}:\dot{\bm{\epsilon}}}}\]
-where $\Delta\eta_T$, $\Delta\eta_y$, $\eta^*$ and $\sigma_Y$ are parameters chosen as listed in Table \ref{tab:tosi_setup}.
-
-For each case, some diagnostic quantities measured at the steady state are reported in Table \ref{tab:visco_mantle} (see Section \ref{sec:error} for 
-the computation of all the quantities), showing a well fit with results described in \citet{Tosi2015}. Data and charts for all cases can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Viscoplastic_mantle_convection}.
-
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{white!100!}{gray!25!}
-%\begin{table}[H]
-%  \caption{Results of the viscoplastic mantle convection benchmark for Cases 1-4 compared with results obtained with ELEFANT and YACC (from \citealp{Tosi2015}).}
-%  \centering
-%  \resizebox{\textwidth}{!}{
-%\begin{tabular}{|c|ccc|ccc|ccc|ccc|}
-%  \toprule
-%    & \multicolumn{3}{c|}{\textbf{Case 1}} & \multicolumn{3}{c|}{\textbf{Case 2}} & \multicolumn{3}{c|}{\textbf{Case 3}} &
-%    \multicolumn{3}{c|}{\textbf{Case 4}} \\
-%   & FALCON & ELEFANT& YACC & FALCON& ELEFANT& YACC & FALCON & ELEFANT & YACC & FALCON & ELEFANT & YACC \\
-%  \midrule
-%   $\left\langle T \right\rangle$ & 0.7759 & 0.7758 & 0.7767 & 0.6043 & 0.6036 & 0.6058 & 0.7275 & 0.7275 & 0.7286 & & 0.5277 & 0.5289\\
-%    Nu$^{top}$ & 3.4280 & 3.4214 & 3.4298 & 8.5242 & 8.5115 & 8.5278 & 3.0368 & 3.0347 & 3.0374 && 6.5912 & 6.5572 \\
-%    Nu$^{bottom}$ & 3.3168 & 3.313 & 3.3143 & 8.4599 & 8.4465 & 8.3990 & 3.0086 & 2.9908 & 2.9628 && 6.5834 & 6.5243 \\
-%    $v_{\textrm{rms}}$ & 249.901 & 249.134 & 251.7997 & 141.2451 & 140.7087 & 142.2020 & 100.2776 & 100.1208 & 100.9467 && 79.1105 & 79.6202 \\
-%    $v_{\textrm{rms}}^{top}$ & 1.8907 & 1.8642 & 1.8298 & 104.7928 & 104.1228 & 105.1767 & 2.0832 & 2.0652 & 2.0374 && 74.7596 & 75.8414 \\
-%    $u_{max}^{top}$ & 2.6340 & 2.6119 & 2.5516 & 121.9212& 121.766 & 122.3238 & 2.9086 & 2.9019 & 2.8458 && 88.9146 & 89.2940 \\
-%    $\eta_{min}$ & - & - & - & $2.0303\times10^{-5}$ & $2.0600\times10^{-5}$ & $1.9853\times10^{-5}$ & $4.8129\times10^{-5}$ & $4.8080\times10^{-5}$ & $4.7907\times10^{-5}$ &
-%    & $1.9860\times10^{-4}$ & $1.9174\times10^{-4}$ \\
-%    $\eta_{max}$ & - & - & - & 1.3325 & 1.85 & 1.6095 & 0.9360 & 0.9023 & 1 && 1.52 & 1.6773 \\
-%    $\left\langle W \right\rangle$ & 2.4299 & 2.4316 & 2.4583 & 7.5648 & 7.5665 & 7.6222 & 2.0387 & 2.0384 & 2.0400 && 5.6216 & 5.6512 \\
-%    $\left\langle \Phi \right\rangle/Ra$ & 2.4311 & 2.4276 & 2.4333 & 7.5661 & 7.5585 & 7.6194 & 2.0386 & 2.0370 & 2.0335 && 5.6182 & 5.6463 \\
-%    $\delta$ & 0.05\% & 0.16\% & 1.02\% & 0.02\% & 0.11\% & 0.04\% & $<$0.01\% & 0.07\% & 0.32\% && 0.06\% & 0.09\% \\
-%  \bottomrule
-%  \end{tabular}}
-%  \label{tab:visco_mantle}
-%  \end{table}
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{2}{white!100!}{gray!25!}
-\begin{table}[H]
-  \caption{Benchmark cases and corresponding parameters.}
-  \centering
-  \small
-\begin{tabular}{|>{\centering}m{0.08\textwidth}|>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}
-  >{\centering}m{0.05\textwidth} >{\centering}m{0.05\textwidth} |>{\centering\arraybackslash}m{0.2\textwidth}|}
-  \toprule
-    Case & $Ra$ & $\Delta\eta_T$ & $\Delta\eta_y$ & $\eta^*$ & $\sigma_Y$ & Convective regime \\
-  \midrule
-  \rowcolor{gray!25!}
-    1 & $10^2$ & $10^5$ & 1 & - & - & Stagnant lid \\
-    2 & $10^2$ & $10^5$ & 1 & $10^{-3}$ & 1 & Mobile lid  \\
-    \rowcolor{gray!25!}
-    3 & $10^2$ & $10^5$ & 10 & - & - & Stagnant lid  \\
-    4 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 1 & Mobile lid \\
-    \rowcolor{gray!25!}
-    5 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 4 & Periodic \\
-  \bottomrule
-  \end{tabular}
-  \label{tab:tosi_setup}
-  \end{table}
-
-\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{gray!25!}{white!100!}
-\begin{table}
-  \caption{Results of the viscoplastic mantle convection benchmark for different grid resolutions in Cases 1-5 compared with results obtained with ELEFANT
-  and YACC ($100\times100$ elements, from \citealp{Tosi2015}) and in Stone 28 ($32\times32$ elements, from
-  \url{https://github.com/cedrict/fieldstone/tree/master/python_codes/fieldstone_28}).}
-  \centering
-  \resizebox{\textwidth}{!}{
-\begin{tabular}{|cc|ccccccccccc|}
-  \toprule
-  \multicolumn{13}{|c|}{\textbf{Case 1}} \\
-  \midrule
-  Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-  & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-  \hline
-  \rowcolor{gray!25!}
-    YACC & $100\times100$ & 0.7767 & 3.4298 & 3.3143 & 251.7997 & 1.8298 & 2.5516 & - & - & 2.4583 & 2.4333 & 1.02\% \\
-    ELEFANT & $100\times100$ & 0.7758 & 3.4214 & 3.313 & 249.134 & 1.8642 & 2.6119 & - & - & 2.4316 & 2.4276 & 0.16\% \\
-    \rowcolor{gray!25!}
-    Stone 28 & $32\times32$ & 0.7737 & 3.3987 & - & 243.872 & - & - & - & - & - & - & - \\
-%    FALCON & $32\times32$ & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & - & - & 2.4745 & 2.4864 & 0.48\% \\
-%    \rowcolor{gray!25!}
-    FALCON & $100\times100$ & 0.7759 & 3.4280 & 3.3168 & 249.901 & 1.8907 & 2.6340 & - & - & 2.4299 & 2.4311 & 0.05\% \\
-    \midrule
-    \multicolumn{13}{|c|}{\textbf{Case 2}} \\
-    \midrule
-    Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-    & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-    \hline
-    \rowcolor{gray!25!}
-     YACC & $100\times100$ & 0.6058 & 8.5278 & 8.3990 & 142.2020 & 105.1767 & 122.3238 & $1.9853\times10^{-5}$ & 1.6095 & 7.6222 & 7.6194 & 0.04\% \\
-     ELEFANT & $100\times100$ & 0.6034 & 8.5115 & 8.4465 & 140.8067 & 104.1228 & 121.766 & $2.0600\times10^{-5}$ & 1.85 & 7.5665 & 7.5585 & 0.11\% \\
-     \rowcolor{gray!25!}
-     Stone 28 & $32\times32$ & 0.6052 & 8.1715 & - & 141.518 & - & - & - & - & - & - & - \\
-%     FALCON & $32\times32$ & 0.6454 & 8.1422 & 7.4020 & 156.6908 & 95.7226 & 116.9302 & $2.2572\times10^{-5}$ & 0.7194 & 7.4289 & 7.4452 & 0.22\% \\
-%     \rowcolor{gray!25!}
-     FALCON & $100\times100$ & 0.6044 & 8.5266 & 8.4590 & 141.3684 & 104.8867 & 122.0470 & $2.0333\times10^{-5}$ & 1.2914 & 7.5707 & 7.5718 & 0.01\% \\
-     \midrule
-     \multicolumn{13}{|c|}{\textbf{Case 3}} \\
-     \midrule
-     Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-     & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-     \hline
-     \rowcolor{gray!25!} 
-     YACC & $100\times100$ & 0.7286 & 3.0374 & 2.9628 & 100.9467 & 2.0374 & 2.8458 & $4.7907\times10^{-5}$ & 1 & 2.0400 & 2.0335 & 0.32\% \\
-      ELEFANT & $100\times100$ & 0.7275 & 3.0347 & 2.9908 & 100.1208 & 2.0652 & 2.9019 & $4.8080\times10^{-5}$ & 0.9023 & 2.0384 & 2.037 & 0.07\% \\
-      \rowcolor{gray!25!}
-      Stone 28 & $32\times32$ & 0.7277 & 3.0347 & - & 100.018 & - & - & - & - & - & - & - \\
-%      FALCON & $32\times32$ & 0.7281 & 3.0618 & 2.8132 & 101.0092 & 2.1630 & 3.0178 & $4.8500\times10^{-5}$ & 0.7845 & 2.0699 & 2.0695 & 0.02\% \\
-%      \rowcolor{gray!25!}
-      FALCON & $100\times100$ & 0.7275 & 3.0368 & 3.0086 & 100.2776 & 2.0832 & 2.9086 & $4.8129\times10^{-5}$ & 0.9360 & 2.0387 & 2.0386 & $<$0.01\% \\
-      \midrule
-      \multicolumn{13}{|c|}{\textbf{Case 4}} \\
-      \midrule
-      Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-      & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-      \hline
-      \rowcolor{gray!25!} 
-      YACC & $100\times100$ & 0.5289 & 6.5572 & 6.5243 & 79.6202 & 75.4814 & 89.2940 & $1.9174\times10^{-4}$ & 1.6773 & 5.6512 & 5.6463 & 0.09\% \\
-       ELEFANT & $100\times100$ & 0.5277 & 6.5912 & 6.5834 & 79.1105 & 74.7596 & 88.9146 & $1.9860\times10^{-4}$ & 1.5200 & 5.6216 & 5.6182 & 0.06\% \\
-       \rowcolor{gray!25!}
-       Stone 28 & $32\times32$ & 0.5286 & 6.4036 & - & 79.4746 & - & - & - & - & - & - & - \\
-%       FALCON & $32\times32$ & 0.5334 & 6.4679 & 6.3435 & 80.6153 & 76.0511 & 90.2058 & $2.0393\times10^{-4}$ & 0.6453 & 5.6676 & 5.6698 & 0.04\% \\
-%       \rowcolor{gray!25!}
-       FALCON & $100\times100$ & 0.5279 & 6.5897 & 6.5830 & 79.1220 & 75.0715 & 88.8484 & $1.9693\times10^{-4}$ & 1.4133 & 5.6127 & 5.6113 & 0.03\% \\
-       \midrule
-       \multicolumn{13}{|c|}{\textbf{Case 5}} \\
-       \midrule
-       Code & Resolution & $P$ & $\left\langle T \right\rangle_{min}$ & $\left\langle T \right\rangle_{max}$ & Nu$^{top}_{min}$ & Nu$^{top}_{max}$ &
-       $v_{\textrm{rms}}^{min}$ & $v_{\textrm{rms}}^{max}$ & \multicolumn{2}{c}{$(\left\langle \Phi \right\rangle/Ra)_{min}$} & 
-       \multicolumn{2}{c|}{$(\left\langle \Phi \right\rangle/Ra)_{max}$} \\
-       \hline
-       \rowcolor{gray!25!}
-         YACC & $100\times100$ & 0.0792 & 0.6543 & 0.6722 & 2.6950 & 7.2792 & 41.9813 & 98.8230 & \multicolumn{2}{c}{1.4347} & 
-         \multicolumn{2}{c|}{9.2748} \\
-         ELEFANT & $100\times100$ & 0.076 & 0.6514 & 0.6693 & 2.6808 & 7.3506 & 41.2578 & 100.089 & \multicolumn{2}{c}{1.3324} & 
-         \multicolumn{2}{c|}{9.5289} \\
-         \rowcolor{gray!25!}
-         Stone 28 & $32\times32$ & 0.0764 & 0.6521 & 0.6697 & 2.6912 & 7.0763 & 41.61 & 98.183 & \multicolumn{2}{c}{-} & 
-         \multicolumn{2}{c|}{-} \\
-     %    FALCON & $32\times32$ & & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & \multicolumn{2}{c}{-} & 
-     %     \multicolumn{2}{c|}{-} \\
-     %    \rowcolor{gray!25!}
-         FALCON & $100\times100$ & 0.0775 & 0.6519 & 0.6697 & 2.6847 & 7.3629 & 41.5256 & 99.5466 & \multicolumn{2}{c}{1.3399} & 
-         \multicolumn{2}{c|}{9.3993} \\
-  \bottomrule
-  \end{tabular}}
-  \label{tab:visco_mantle}
-  \end{table}
-
 \subsection{Thin layer entrainment}\label{sec:thin}
@@
 
-The experiment is performed with two grid resolution (\num{125x40} and \num{200x80} elements), with 100 markers per element and CFL of 0.25. $v_{\textrm{rms}}$
+The experiment is performed with two grid resolution ($125\times40$ and $200\times80$ elements), with 100 markers per element and CFL of 0.25. $v_{\textrm{rms}}$
 (see Sec. \ref{sec:error}) calculated for both the simulations match well with results from \citet{vanKeken1997} and obtained with ASPECT
@@
 \end{figure}
-
-\subsection{Sinking hydrated cylinder}\label{sec:quinquis}
-This experiment is performed as explained in \citet{Quinquis2014} and simulates the sinking of a cold, hydrated cylinder into a warm, dry mantle. The domain
-is square with $L_x=L_y=$ \SI{300}{\km}, a grid resolution of \num{300x300} elements and 25 markers per element. The sinking cylinder is located at $x=$
-\SI{150}{\km} and $y=$ \SI{170}{\km} with a radius of \SI{20}{\km} and has $\rho_c=$ \SI{3250}{\kg\per\cubic\m}, $\eta_c=$ \SI{e23}{\pascal\s}, a thermal
-conductivity $k_c=$ \SI{4.5}{\watt\per\m\per\kelvin},a specific heat $C_{p_c}=$ \SI{1250}{\joule\per\kg\per\kelvin}; the surrounding mantle has $\rho_m=$
-\SI{3200}{\kg\per\cubic\m}, $\eta_m=$ \SI{e20}{\pascal\s}, a thermal conductivity $k_m=$ \SI{105}{\watt\per\m\per\kelvin} and specific heat $C_{p_m}=$
-\SI{1250}{\joule\per\kg\per\kelvin}; a 58 km-thick lithosphere overlies the mantle and has $\rho_l=$ \SI{3200}{\kg\per\cubic\m}, $\eta_l=$ \SI{e23}{\pascal\s},
-a thermal conductivity $k_l=$ \SI{4.5}{\watt\per\m\per\kelvin}, specific heat $C_{p_l}=$ \SI{750}{\joule\per\kg\per\kelvin} and a thermal diffusivity of
-\SI{e-6}{\square\m\per\s}. Velocity boundary conditions are set to free slip on all sides of the domain. Temperature increase linearly in the lithosphere,
-from 0° to 1300°C, and in the mantle, from 1300° to 1360.5°C. The initial temperature of the cylinder is fixed to 400°C. Throughout the evolution, temperature
-is fixed to 0° and 1360.5°C at the top and bottom of the domain, respectively. The initial bound water content is imposed at 2 wt.\% in the cylinder and at
-0 wt.\% in both the mantle and the lithosphere. The maximum amount of water in the mantle is fixed to 0.2 wt.\%, while maximum water content in the cylinder
-and in the lithosphere is function of pressure and temperature and is calculated using a serpentinized harzburgite with Perple\_X, as in \citet{Quinquis2014}.
-
-As shown in \citet{Quinquis2014}, the model is characterised by a progressive dehydration in the external portion of the cylinder, due to the increase of
-temperature, with a consequent vertical migration of free water that hydrates the mantle above the cylinder, up to the lithosphere. The distribution of
-bound and free water during the evolution of the experiment is shown in Fig. \ref{fig:quinquis} (left and right columns, respectively). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Hydration}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Hydration.pdf}
-\caption{Distribution of bound and free water (left and right columns, respectively) at $t=0.5$ and $t=$ \SI{1.25}{\mega\year} (first and second rows,
-respectively) for the sinking hydrated cylinder experiment. Only markers with bound or free water not equal to 0 are plotted.}
-\label{fig:quinquis}
-\end{figure}
-
-\subsection{Mantle melting curves}\label{sec:katz}
-This instantaneous experiment is performed in order to recreate isobaric melting curves, as function of temperature, bulk water and content of water in the
-melt, to compare with melting curves obtained by \citet{Katz2003}. The domain is square with $L_x=L_y=100$ km, a grid resolution of \num{100x100} elements and
-25 markers per element. Density is set in the entire domain to \SI{3300}{\kg\per\cubic\m} and temperature increases from the sides to the centre from 1000° to
-1500°C. The experiment is performed for bulk water contents from 0 to 0.3 wt.\%.
-
-Fig. \ref{fig:fraction} shows isobaric melting curves for \SI{1}{\giga\pascal} (panel a) and \SI{3}{\giga\pascal} (panel b) calculated for different bulk water
-contents. The curves well match with those obtained for the mantle by \citet{Katz2003}, showing the saturation at the solidus for bulk water of 0.2 and 0.3 wt.\%, indicating
-that the code correctly determines melt fraction as function of pressure, temperature and bulk water content. Same results can be observed in Fig.
-\ref{fig:water}a, where isobaric and isothermal melting curves are shown as function of bulk water content. In addition, Fig. \ref{fig:water}b shows that the
-code correctly determines also the percentage of water in the melt as function of pressure, temperature and melt fraction. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Melting}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Fraction.pdf}
-\caption{Isobaric melting curves for 1 GPa (panel a) and 3 GPa (panel b) as function of temperature with different bulk water contents.}
-\label{fig:fraction}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Water.pdf}
-\caption{Isobaric melting curves for 1.5 GPa as function of bulk water (panel a) and water in the melt (panel b) calculated for different temperatures.}
-\label{fig:water}
-\end{figure}
-
-\clearpage
-
-\bibliography{export}
-
-\end{document}

### 2022-09-07 tex/en_eq.tex
update main

--- 2022-07-20 FALCON_description.tex
@@
-\documentclass[hidelinks,10pt,a4paper]{article}
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{textcomp}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage{booktabs}
-\usepackage{multirow}
-\usepackage[cm]{fullpage}
-\usepackage[round]{natbib}
-\bibliographystyle{abbrvnat}
-\usepackage{doi}
-\usepackage{bm}
-\usepackage[table]{xcolor}
-\hypersetup{
-    colorlinks,
-    linkcolor={blue},
-    citecolor={blue},
-    urlcolor={blue}
-}
-\usepackage{float}
-\restylefloat{table}
-\usepackage{siunitx}
-\DeclareSIUnit\year{yr}
-
-\begin{document}
-
-\section{Introduction}\label{sec:intro}
-This document includes a description of the code FALCON, with a presentation of the features implemented followed by results of extensive benchmarks listed in
-Table \ref{tab:list}, performed to verify that the code solves correctly a wide range of problems. Description and results of all benchmarks are shown
-in Section \ref{sec:benchmark}.
-
-\begin{table}[H]
-\caption{List of benchmarks performed for each feature of the code. Data of all benchmarks can be found at 
-\url{https://github.com/aleregorda/Benchmarks}}
-\centering
-\begin{tabular}{| l | l | l |}
-\toprule
- \textbf{Feature}  & \textbf{Benchmark} & \textbf{Section} \\
-\midrule
-  Solver  & Stokes flow & \ref{sec:stokes} \\
-\hline
-  \multirow{2}{*}{Markers advection} & Zalesak disk & \ref{sec:runge} \\
-    & Conservative Interpolation Velocity & \ref{sec:CVI} \\
-\hline
-  \multirow{4}{*}{Momentum equation}  & Poiseuille flow & \ref{sec:poiseuille}\\
-    & Instantaneous 2D sphere & \ref{sec:ist_sphere} \\
-    & Rayleigh-Taylor experiment & \ref{sec:rayleigh} \\
-    & Falling block & \ref{sec:block} \\
-\hline
-    & 2D Stokes sphere & \ref{sec:sphere} \\
-    Sticky air and  & Stabilisation algorithm & \ref{sec:stab} \\
-    free surface & Topography relaxation & \ref{sec:crameri} \\
-  & Spontaneous subduction & \ref{sec:subduction} \\
-\hline
-  Erosion and sedimentation  & - & - \\
-\hline
-   Non-linear  & Slab detachment & \ref{sec:slab} \\
-   visco-plasticity & Indenter & \ref{sec:indenter} \\
-    & Brick & \ref{sec:brick} \\
-\hline
-  \multirow{3}{*}{Energy equation}  & Advection stabilisation & \ref{sec:advection} \\
-    & Simple shear heating & \ref{sec:simple_shear} \\
-    & Shear and adiabatic heating & \ref{sec:shear} \\
-\hline
-\multirow{3}{*}{Energy + momentum} & Mantle convection & \ref{sec:mantle} \\
-& Viscoplastic mantle convection & \ref{sec:visco_mantle} \\
-& Thin layer entrainment & \ref{sec:thin} \\
-\hline
-  Phase changes  & \multirow{2}{*}{Hydrated sinking cylinder} & \multirow{2}{*}{\ref{sec:quinquis}} \\
-  and hydration  &  & \\
-\hline
-  Melting & Experimental melting curves & \ref{sec:katz} \\
-\bottomrule
-\end{tabular}
-\label{tab:list}
-\end{table}
-
-The code is written in Fortran90 and uses the finite element method (FEM) with quadrilateral $Q_1 \times P_0$ elements (continuous bilinear velocity and
-discontinuous constant pressure), associated with MUMPS\footnote{\url{http://mumps.enseeiht.fr/}} \citep{Amestoy2001,Amestoy2019} that is a software package
-for solving systems of linear equations of the form $A \cdot x = b$, where \textit{A} is a square sparse matrix, by means of a direct method. Since
-$Q_1 \times P_0$ elements do not satisfy Ladyzhenskaya, Babuska and Brezzi (LBB) stability condition \citep{Donea2003}, elemental pressure is elaborated
-in post-processing to avoid spurious pressures by means of a double interpolation to smooth the pressure. Following this procedure, the elemental pressure
-is interpolated onto nodes and then back onto elements. After the smoothing procedure, the pressure field is calculated again on the nodes to be used in
-combination with the lithostatic pressure, which is calculated onto nodes. Correctness of the solution and performances of the solver in terms of time and
-memory usage are tested solving the Stokes flow with the analytical solution proposed by \citet{Donea2003} (Section \ref{sec:stokes}).
-
-The thermo-mechanics of crust-mantle systems is described by means of the conservation of mass, momentum and energy equations, expressed as follows:
-\begin{eqnarray}
-\label{eq:mass}\frac{\partial \rho}{\partial t}+\bm{\nabla} \cdot (\rho \bm{u})&=&0\\
-\label{eq:momentum}-\bm{\nabla} p + \bm{\nabla} \cdot \bm{\tau} + \rho \bm{g}&=&0\\
-\label{eq:energy}\rho C_p \left(\frac{\partial T}{\partial t} + \bm{u} \cdot \bm{\nabla} T \right)&=&\bm{\nabla} \cdot \left(k\bm{\nabla} T\right) + H_{tot}
-\end{eqnarray}
-where $\rho$ is the density, $\bm{u}$ is the velocity, \textit{p} is the pressure, $\bm{\tau}$ is the deviatoric stress, $\bm{g}$ is the gravity acceleration,
-$C_p$ is the specific heat at constant pressure, \textit{T} is the temperature, \textit{k} is the thermal conductivity and $H_{tot}$ is the total internal heat
-production. Density variations due to temperature are generally small enough to assume the density as constant ($\rho=\rho_0$) in Eq. \ref{eq:mass} and in
-Eq. \ref{eq:energy}, while it must be treated as a variable in the buoyancy term of Eq. \ref{eq:momentum}, such that \[\rho=\rho_0(1-\alpha(T-T_0))\]
-where $\rho_0$ is the density at a reference temperature $T_0$ and $\alpha$ is the coefficient of thermal expansion. Eq. \ref{eq:mass} then can be rewritten as
-\begin{equation}\label{eq:mass_0}
-\bm{\nabla} \cdot \bm{u}=0
-\end{equation}
-This simplification is known as the extended Boussinesq approximation.
-
-The time step $dt$ is chosen in according to Courant-Friedrichs-Lewy condition \citep{Anderson1995}, such that
-\[dt=C_n\; \min\left(\frac{\Delta x}{\max|\bm{u}|},\frac{(\Delta x)^2}{\kappa}\right)\]
-where $0<C_n<1$ is the Courant (CFL) number, $\Delta x$ is the minimum dimension of smallest element, $\max|\bm{u}|$ is the maximum velocity calculated on the
-entire domain and
-$\kappa=\frac{k}{\rho C_p}$ is the thermal diffusivity \citep{Thieulot2014}.
-
-\section{Lagrangian markers}\label{sec:markers}
-Elemental properties (density, viscosity, thermal conductivity, specific heat and thermal expansion) needed to solve Eqs. \ref{eq:mass_0}, \ref{eq:momentum}
-and \ref{eq:energy} are related to the composition of each element, determined by means of Lagrangian markers that are characteristic of different materials
-in the domain. Elemental properties (with the sole exception of the viscosity) are calculated using an arithmetic mean, as \[P_e=\frac{1}{n}\sum_{i=1}^n P_i\]
-where $P_e$ is the elemental property, $P_i$ is the property characteristic of the material of each marker and $n$ is the number of markers of the element.
-Differently, the average scheme for the elemental viscosity can be chosen between harmonic, geometric and arithmetic mean. Markers can be placed regularly or
-randomly at the beginning of the simulation and their advection is performed by either a 2nd-order or a 4th-order Runge-Kutta in space, interpolating the velocity 
-field on each marker by means of the shape functions (see Section \ref{sec:runge}). The interpolated velocity is then corrected by means of the Conservative
-Velocity Interpolation (CVI), which introduces a corrective term to reduce dispersion and clustering of particles in both steady state and time-dependent 
-\citep{Wang2015} (see Section \ref{sec:CVI}).
-
-The initial distribution of the markers is created using the open-source code library
-Geodynamic World Builder \citep{Fraters2019}. During the simulation, each marker carries memory of temperature, pressure and accumulated strain, which are
-determined interpolating the relative nodal parameter. In particular, pressure is interpolated onto each marker after the smoothing procedure. The number
-of markers contained in each element is maintained between $n_{min}$ and $n_{max}$. When in an element there are less markers than $n_{min}$ the code adds
-random markers to reach the $n_{min}$, while if the number is higher than $n_{max}$ some of them are randomly deleted. When new markers are added, they assume
-the properties of the nearest marker. In this way elements are never empty and maintain a number of markers inside a prefixed range.
-
-\section{The momentum equation}\label{sec:mom_eq}
-The penalty method is implemented so that Eq. \ref{eq:mass_0} can be written as:
-\begin{equation}\label{eq:mass_penalty}
-\bm{\nabla} \cdot \bm{u} + \frac{p}{\lambda}=0
-\end{equation}
-where $\lambda$ is the so-called penalty parameter, which should be 6-7 orders of magnitude larger than the shear viscosity to ensure that mass conservation
-is satisfied \citep{Donea2003,Thieulot2014}.
-
-The deviatoric stress tensor in Eq. \ref{eq:momentum} can be written in terms of the strain rate tensor as $\bm{\tau}=2\eta\dot{\bm{\epsilon}}$, with
-$\dot{\bm{\epsilon}}=\frac{1}{2}\left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)$. Therefore, Eq. \ref{eq:momentum} can be rewritten as:
-\begin{equation}\label{eq:momentum_pressure}
--\bm{\nabla} p + \bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right) + \rho \bm{g} = 0
-\end{equation}
-Finally, using pressure from Eq. \ref{eq:mass_penalty}, Eq. \ref{eq:momentum_pressure} can be rewritten as:
-\begin{equation}\label{eq:momentum_penalty}
-\lambda \bm{\nabla} \left(\bm{\nabla} \cdot \bm{v}\right)+\bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right)+\rho\bm{g}=0
-\end{equation}
-The penalty method is been associated to the iterative Uzawa method, as described in detail in \citet{Dabrowski2008} and \citet{Thieulot2014}
-(see Section \ref{sec:poiseuille}). Results of benchmarks performed to verify the correctness in the implementation of Eq. \ref{eq:momentum_penalty} are
-shown in Sections \ref{sec:ist_sphere} and \ref{sec:rayleigh}.
-
-To support large viscosities variations the penalty parameter is related to the effective elemental viscosity by means of a dimensionless coefficient, so that
-$\lambda=\lambda_e(e)\eta_{eff}(e)$ \citep{Marotta2006,Dabrowski2008,Thieulot2014}. Benchmark of the falling block \citep{Gerya2003a,Gerya2010,Thieulot2011}
-is performed to verify that the code can correctly deal buoyancy driven flows with strong viscosity contrasts (Section \ref{sec:block}).
-
-\section{Sticky air and free surface}\label{sec:sticky}
-The Earth's surface can be treated by means of either the so-called sticky-air or a true free surface method, both of them implemented in the code. In the
-sticky-air method the surface is approximated with the introduction of a buoyant layer with a viscosity at least four orders of magnitude lower than the crust
-\citep{Schmeling2008,Crameri2012} and the interface between lithosphere and air is defined using a chain of passive markers that are advected as the Lagrangian
-markers. The correctness of the evolution of the markers chain is tested with the experiment of a 2D time-dependent Stokes sphere below a free surface and
-compared with results from ASPECT\footnote{\url{https://aspect.geodynamics.org/}} \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} (Section
-\ref{sec:sphere}). In the true free surface case the top boundary is assumed stress-free and velocities are not fixed. In this case topography variations are
-described by vertical deformations of the mesh that depend on the velocity field of the nodes that identify the free surface, while horizontal deformations are
-not taken into account. This procedure is known as the Arbitrary Lagrangian-Eulerian (ALE) method and its implementation follows the technique described in
-\citet{Thieulot2011}. However, although the implementation of a true free surface better reproduces laboratory experiment, extremely small time steps can be
-necessary to maintain stability \citep{Kaus2010a,Quinquis2011,Thieulot2014}. Therefore, the stability algorithm proposed by \citet{Kaus2010a} is implemented
-to avoid instabilities due to high density differences at the free surface when using too large time steps. The implementation of this algorithm are tested by
-performing the experiment described by \citet{Kaus2010a} (Section \ref{sec:stab}).
-
-The topography relaxation benchmark proposed by \citet{Crameri2012} is performed to verify that the code correctly recovers topography variations in case of
-both the sticky-air and the true free surface method (Section \ref{sec:crameri}). Finally, Section \ref{sec:subduction} show the results of the spontaneous
-subduction experiment described by \citet{Schmeling2008}.
-
-\section{Erosion and sedimentation}\label{sec:erosion}
-Surface processes at either the lithosphere-air interface or in correspondence of the free surface have been implemented by means of the software
-Fastscape\footnote{\url{https://fastscape.org/}} \citep{Braun2013,Cordonnier2019,Yuan2019a,Yuan2019b}. FastScapeLib is a set of routines that solve the stream
-power law (enriched by a sediment transport/deposition term) \citep{Yuan2019a}, hillslope diffusion and marine transport and deposition \citep{Yuan2019b}.
-The partial differential equation solved by FastScapeLib is:
-\begin{equation}
-\frac{\partial h}{ \partial t}=U-K_f A^{m_e}S^{n_e} + \frac{G}{A}\int_A \left( U-\frac{\partial h}{\partial t} \right) dA + K_d \bm{\nabla}^2 h 
-\end{equation}
-where $h$ is the topography, $U$ is the uplift, $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $S$ is the slope, $m_e$ and $n_e$ are
-parameters that should be chosen to constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}, $G$ is the dimensionless
-deposition/transport coefficient for the enriched stream power law and $K_d$ is the effective diffusivity. Terms $K_f A^{m_e}S^{n_e}$ and $K_d \bm{\nabla}^2 h$
-consider long-range processes, described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, and short-range processes,
-proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}.
-
-In case of sticky air, slope is determined by means of the markers chain, which is vertically corrected considering erosion and sedimentation rates. After the
-correction, continental markers above the markers chain are transformed in air markers and air markers below the markers chain become sediments. Similarly,
-slope is determined considering the nodes on the top boundary in case of true free surface and they are vertically corrected as for the markers chain. In this
-case, continental markers above the top boundary are simply deleted.
-%Surface processes at the lithosphere-air interface are implemented taking into account fluvial erosion (long-range processes), hillslope processes (short-range
-%processes) and failure of slopes exceeding a threshold angle $\theta$ \citep{Burov1997,Tucker1999,Sacek2012,Croissant2014,Ueda2015,Beucher2020}. Erosion rates
-%$\dot{E}_{lr}$ due to long-range processes can be described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{lr}=K_f A^{m_e} \bm{\nabla} h^{n_e}
-%\end{equation}
-%where $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $\bm{\nabla} h$ is the slope and $m_e$ and $n_e$ are parameters chosen in to
-%constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}. Erosion rates $\dot{E}_{sr}$ due to short-range processes are
-%implemented by assuming that transport is linearly proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{sr}=K_d \bm{\nabla}^2 h
-%\end{equation}
-%where $K_d$ is the effective diffusivity. In addition, the slope is modified if it still exceeds a threshold angle after long- and short-range processes
-%\citep{Ueda2015}.
-
-%Offshore sedimentation processes are implemented similarly to hillslope processes but with a depth-dependent coefficient \citep{Sacek2012,Beucher2020}.
-%The sedimentation rate $\dot{S}$ is calculated as
-%\begin{equation}
-%\dot{S}=K_s \bm{\nabla}^2 h
-%\end{equation}
-%with
-%\begin{equation}
-%K_s=K^*_s \exp\left(\frac{h-h_{sl}}{h_0}\right)
-%\end{equation}
-%where $K^*_s$ is a constant, $h_{sl}$ is the sea level and $h_0$ is a characteristic length. In case of sticky air, slope is determined by means of the markers
-%chain, which is vertically corrected considering erosion and sedimentation rates. After the correction, continental markers above the markers chain are
-%transformed in air markers and air markers below the markers chain become sediments. Similarly, slope is determined considering the nodes on the top boundary
-%in case of true free surface and they are vertically corrected as for the markers chain. In this case, continental markers above the top boundary are simply
-%deleted.
-
-\section{Non-linear rheologies}\label{sec:nonlinear}
-Non-linear rheologies are implemented combining viscous creep (dislocation and diffusion) and plastic yielding. For each marker, diffusion and dislocation
-viscosity ($\eta_{df|ds}$) can be determined as follows \citep{Karato1993,Warren2008b,Wang2016}:
-\begin{equation}\label{eq:viscous}
-\eta_{df|ds}=f_s \left(\frac{d^m}{A}\right)^{\frac{1}{n}} \left (\frac{2^{\frac{1-n}{n}}}{3^{\frac{n+1}{2n}}} \right)
-I_2^{\frac{(1-n)}{n}}\exp\left(\frac{Q+pV}{nRT}\right)
-\end{equation}
-where $f_s$ is a scaling factor used to represent lithologies that are stronger or weaker than the base set, $d$ is the grain size, $m$ is the grain size exponent,
-$n$ is the stress exponent, $A$ is the uniaxial pre-exponential factor, $I_2$ is the square root of the second invariant of the strain rate tensor, $Q$ is the
-activation energy, $p$ is the pressure, $V$ is the activation volume, $R$ is the gas constant and $T$ is the temperature.
-Pressures and temperatures are determined by the interpolation of the nodal parameters, while strain rates are calculated by means of the
-derivative of the velocities on the nodes. In case of diffusion creep $n=1$ and $m>0$, in case of dislocation creep $n>1$ and $m=0$. The viscous creep $\eta_{cp}$
-is then calculated as the harmonic average between $\eta_{df}$ and $\eta_{ds}$:
-\begin{equation}\label{eq:visco_creep}
-\eta_{cp}=\left(\frac{1}{\eta_{df}}+\frac{1}{\eta_{ds}}\right)^{-1}
-\end{equation}
-The implementation of the non-linear viscous creep viscosity depending on the strain rate is tested by means of the slab detachment benchmark
-(Section \ref{sec:slab}).
-
-Plastic yielding is implemented rescaling $\eta_{cp}$ in order to limit the stress below the yield stress $\sigma_y$ \citep{Thieulot2008,Thieulot2014,Glerum2018},
-obtaining
-\begin{equation}\label{eq:plastic}
-\eta_{pl}=\frac{\sigma_y}{2I_2}
-\end{equation}
-where $\eta_{pl}$ is the plastic viscosity and the yield stress is determined following the Drucker-Prager criterion, such as
-\begin{equation}\label{eq:yield}
-\sigma_y=C\;\cos(\phi)+p\;\sin(\phi)
-\end{equation}
-where $C$ is the cohesion and $\phi$ is the internal friction angle. In case of negative pressure it is imposed equal to 0, so that negative yield stress are
-excluded. The correctness of the non-linear solution in case of plastic viscosity with variable internal friction angle is verified performing indenter and
-brick experiments (Sections \ref{sec:indenter} and \ref{sec:brick}, respectively).
-
-Plastic yielding and viscous creep are then combined to obtain a viscoplastic viscosity $\eta_{vp}$ as follows, assuming that they are independent processes
-\citep{Karato2008,Glerum2018}:
-\begin{equation}\label{eq:viscoplastic}
-\eta_{vp}=\min(\eta_{cp},\eta_{pl})
-\end{equation}
-Finally, effective viscosity $\eta_{eff}$ is capped by the minimum and the maximum viscosity ($\eta_{min}$ and $\eta_{max}$, respectively) to avoid extremely
-low or high viscosity \citep{Glerum2018} as follows
-\begin{equation}\label{eq:effective}
-\eta_{eff}=\min(\max(\eta_{vp},\eta_{min}),\eta_{max})
-\end{equation}
-Viscous creep and plastic yielding are non-linear rheologies because of their dependence on the velocity field through pressure and strain rates. Therefore,
-the solution is determined by means of Picard-type iterations, until convergence of the velocity field \citep{Glerum2018}. The convergence is verified at each
-iteration $i$ via the nonlinear residual $\bm{\mathcal{R}}^i$ that can be determined as
-\begin{equation}\label{eq:convergence}
-\bm{\mathcal{R}}^i= \mathbb{K} (\eta_{eff}(\dot{\bm{\epsilon}}^{i-1},p^{i-1}))\cdot \bm{v}^{i-1} - \bm{f}^i
-\end{equation}
-%\begin{equation}\label{eq:convergence}
-%\chi=1-\frac{\langle (\bm{v}^i-\langle\bm{v}^i\rangle)\cdot (\bm{v}^{i+1}-\langle\bm{v}^{i+1}\rangle) \rangle}{\sqrt{\sigma^i\sigma^{i+1}}}
-%\end{equation}
-%Iterations are performed until either convergence ($\chi<tol$) or maximal number of non-linear iterations ($it_{max}$) is reached.
-where $\mathbb{K}$ is the Stiffness matrix and $\bm{f}$ is the right hand side vector \citep{Spiegelman2016,Glerum2018}.  The $L_2$-norm is extracted from
-$\bm{\mathcal{R}}^i$ and it is normalised as follows
-\begin{equation}\label{eq:convergence_res}
-\frac{||\bm{\mathcal{R}}^i||_2}{||\bm{\mathcal{R}}^0||_2}
-\end{equation}
-where $||\bm{\mathcal{R}}^0||_2$ is the $L_2$-norm of the first nonlinear iteration.
-Since often the normalised nonlinear residual drop very quickly over the first 2-3 iterations, also the $L_2$-norm of velocity and pressure residuals are
-calculated as
-\begin{equation}\label{eq:convergence_vel}
-\frac{||\Delta\bm{v}||_2}{||\bm{v}||_2}=\frac{||\bm{v}_i-\bm{v}_{i-1}||_2}{||\bm{v}_{i}||_2}
-\end{equation}
-and
-\begin{equation}\label{eq:convergence_press}
-  \frac{||\Delta p||_2}{||p||_2}=\frac{||p_i-p_{i-1}||_2}{||p_{i}||_2}
-\end{equation}
-respectively. Iterations are performed until either maximal number of non-linear iterations ($it_{max}$) is reached or all the normalised nonlinear residual, 
-the velocity residual and the pressure residual converge under a defined tolerance $tol$.
-
-\section{Healing and weakening}
-The temporal evolution of the accumulated strain $\epsilon$ has been implemented as in \citet{Fuchs2019,Fuchs2021}, as following
-\begin{equation}\label{eq:healing}
-  \frac{d\epsilon}{dt}=I_2 - \epsilon H_{rate}(T)
-\end{equation}
-where the first term on the right-hand side is a source term given by the second invariant of the strain rate $I_2$ and the second term a
-temperature-dependent healing factor $H_{rate}$, calculated as
-\begin{equation}\label{eq:healing_rate}
-  H_{rate}(T)=B \exp \left[-\frac{\mu}{2}\left(\frac{1}{T+1}-\frac{1}{2}\right)\right]
-\end{equation}
-where $B$ is a constant describing the time scale of healing, while $\mu$ and $T$ are the non-dimensional temperature activation and temperature,
-respectively \citep{Fuchs2019,Fuchs2021}.
-
-Strain softening is taken into account for both viscous creep and plastic viscosity \citep{Huismans2003,Babeyko2005,Huismans2005,Sobolev2005,Warren2008a} by
-means of the accumulated strain $\epsilon$ memorised by each marker. Softening in the viscous creep determines a linear decrease of $\eta_{vc}$ by means of a
-viscous strain softening factor $W_S$ that increases linearly from $W_{S_0}$ to $W_{S_{\infty}}$ for $\epsilon_{S_0}<\epsilon_S<\epsilon_{S_{\infty}}$ 
-\citep{Warren2008a}. This viscous softening can be related to strain-induced grain size reduction \citep{Warren2008a}. Differently, plastic softening is 
-simulated with a linear decrease of internal friction angle $\phi(\epsilon)$ and cohesion $C(\epsilon)$ in according to
-\begin{eqnarray}
-\label{eq:friction1}\phi(\epsilon)&=&\phi_0+(\phi_{\infty}-\phi_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}\\
-\label{eq:friction2}C(\epsilon)&=&C_0+(C_{\infty}-C_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}
-\end{eqnarray}
-where $\phi_0$, $C_0$ and $\phi_{\infty}$, $C_{\infty}$ are internal friction angle and cohesion for $\epsilon_0$ and $\epsilon_{\infty}$, respectively
-\citep{Huismans2003,Huismans2005,Warren2008a,Thieulot2014}. Plastic softening approximates deformation-induced softening of faults and brittle shear zones
-\citep{Warren2008a}.
-
 \section{The energy equation}\label{sec:en_eq}
@@
 mantle convection benchmark is performed to check also the viscoplasticity in case of a initial perturbed temperature field (Section \ref{sec:visco_mantle}).
-
-\section{Phase transitions and hydration}\label{sec:phase}
-Variations of effective density, specific heat and coefficient of thermal expansion during the evolution of crustal and mantle materials at different
-pressure-temperature ($p$-$T$) conditions are computed by means of version 6.8.6 of Perple\_X software package \citep{Connolly2005}, similarly to the
-implementation described in \citet{Marotta2020}. By default, mineral assemblages and properties of each lithology (oceanic crust, lower and upper continental
-crust, sediments and dry or hydrated mantle) are calculated for temperatures from \SIrange{330}{1600}{\kelvin}, with increments of approximately \SI{4}{\kelvin},
-and pressures from \SIrange{0.1}{30}{\giga\pascal}, with increments of approximately \SI{0.03}{\giga\pascal}, for a total of almost 100000 points.
-
-The introduction of phase transitions produces two main effects that must be taken into account in numerical models: variations in density and release/absorption
-of latent heat \citep{Tackley2010}. Effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into
-account considering effective coefficient of thermal expansion \citep{Christensen1985,Zhong2015}. Similarly, effects of latent heat in the energy equation
-(eq \ref{eq:energy}) are taken into account by considering effective specific heat and coefficient of thermal expansion
-\citep{Christensen1985,Tackley2010,Zhong2015}. Effective density, specific heat and coefficient of thermal expansion are included in Perple\_X files and are
-assumed by each Lagrangian marker in according to its $p$-$T$ conditions.
-
-Hydration processes related to dehydration of subducting lithosphere \citep{Schmidt1998,Liu2007,Faccenda2009,Faccenda2010,Faccenda2014,Rosas2016} strongly
-influence the thermo-mechanics inside the mantle wedge, mainly because of both weakening effects on the mantle rheology and density variations in case of
-mantle serpentinization \citep{Gerya2002,Honda2003,Arcay2005,Roda2010,Regorda2017}. In the case that hydration is switched on, the amount of bound and free
-water is memorised by each marker, following the implementation of \citet{Quinquis2014}; hydration and dehydration processes are related to the amount of bound
-water of each marker and to the maximum amount of water it can transport, i.e. if the amount of bound water exceeds the maximum amount of water, the marker
-dehydrates and releases free water that can hydrates under-saturated markers. Maximum water content of each marker is determined as function of lithology and
-$p$-$T$ conditions and it is calculated using Perple\_X, in the same way than effective density, specific heat and coefficient of thermal expansion. Bound water
-is advected together with the markers, neglecting the effect of bound water diffusion, while free water simply migrates vertically and is not coupled to the
-solid-phase flow of the mantle wedge \citep{Arcay2005,Quinquis2014}. The correctness of the migration scheme of the free water and its absorption into bound
-water by under-saturated markers are verified performing the experiment described in \citet{Quinquis2014} (Sec. \ref{sec:quinquis}).
-
-The mantle viscosity weakening related to the amount of water has been extensively studied \citep{Chopra1981,Hirth2003} and it is implemented in according to
-\citet{Arcay2005} and \citet{Horiuchi2016}, as follows
-\begin{equation}\label{eq:wet_visc}
-\eta_{wet}=\eta_{dry}\left(\left[1-\frac{1}{f_v}\right]\exp\left(-\frac{[OH^-]}{[OH^-]_0}\right)+\frac{1}{f_v}\right)
-\end{equation}
-where $f_v$ is the viscosity reduction factor between dry and wet conditions, $[OH^-]$ is the water content and $[OH^-]_0$ is a reference water content set to
-620 ppm (0.062 wt.\%) \citep{Arcay2005}. By default, $f_v$ is set to 100, in according to \citet{Arcay2005} and \citet{Horiuchi2016}.
-
-\section{Mantle melting}\label{sec:melt}
-Melt fraction in the mantle depends on temperature ($T$, in \textdegree C), pressure ($p$, in \si{\giga\pascal}) and the water content in the melt ($X_{H_2O}$, in wt.\%).
-The determination of the melt fraction $M$ is implemented as explained by \citet{Katz2003}, according to the successive modification by \citet{Langmuir2006} and
-\citet{Kelley2010}, as follows
-\begin{eqnarray}\label{eq:melt_fr}
-M(p,T,X_{H_2O})&=&-T+T_{sol}+(x_t\cdot ln(p+y_t))\cdot X_{H_2O}-\nonumber\\
-&&+K_m\left(\frac{X_{H_2O}^{bulk}}{D_{H_2O}(1-X_{H_2O})+X_{H_2O}}\right)^{\gamma}
-\end{eqnarray}
-where $T_{sol}$ is the temperature of the dry solidus, $D_{H_2O}$ is the partition coefficient and $x_t\cdot log(p+y_t)$ indicates the pressure dependence of
-the melting curve \citep{Kelley2010}. $x_t$, $y_t$, $K_m$ and $\gamma$ are parameters chosen in according to \citet{Kelley2010}.
-However, $X_{H_2O}$ depends on the melt fraction as
-\begin{equation}\label{eq:melt_wt}
-X_{H_2O}(M)=\frac{X_{H_2O}^{bulk}}{D_{H_2O}+M(1-D_{H_2O})}
-\end{equation}
-and a numerical solution can be found using a root-finder method \citep{Katz2003,Wang2016}. Furthermore, the content of water in the melt is limited by the
-pressure-dependent saturation concentration of water in the melt, determined as
-\begin{equation}\label{eq:melt_max}
-X_{H_2O}^{sat}=\chi_1 p^{\lambda_m}+\chi_2 p
-\end{equation}
-where $\chi_1$, $\chi_2$ and $\lambda_m$ are parameters chosen according to \citet{Katz2003}, \citet{Langmuir2006} and \citet{Kelley2010}. The correctness in
-the determination of melt fractions and water contents in the melt are tested comparing model results with experimental curves obtained by \citet{Katz2003}
-(Sec. \ref{sec:katz}).
-
-In case of melting, viscous creep viscosity ($\eta_{vp}$) is modified according to \citet{Wang2016}, as
-\begin{equation}\label{eq:melt_visc}
-\eta_{melt}=\eta_{vp}\cdot \exp(\alpha_m M)
-\end{equation}
-where $\alpha_m$ are melt fraction factors for dislocation and diffusion creep, chosen in according to \citet{Wang2016}.
-
-Effective density ($\rho_{eff}$) for partially molten rocks is calculated as
-\begin{equation}\label{eq:melt_dens}
-\rho_{eff}=\rho_s(1-M)+\rho_m M
-\end{equation}
-where $\rho_s$ and $\rho_m$ are the densities of the solid and the molten rock, respectively \citep{Gerya2004b,Gerya2007,Wang2016}. The density of the
-solid rock is extracted by Perple\_X, while the density of the molten portion is calculated as
-\begin{equation}\label{eq:melt_dens_M}
-\rho_m(p,T)=\rho_0[1-\alpha(T-T_0)][1+\beta(p-p_0)]
-\end{equation}
-where $\rho_0$ is the density at the reference temperature ($T_0$) and pressure($p_0$), and $\alpha$ and $\beta$ are the thermal expansion and compressibility
-coefficients, respectively \citep{Gerya2004b,Gerya2007,Wang2016}.
-
-As for the phase transitions, effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into account
-considering effective coefficient of thermal expansion. Similarly, latent heat due to melting/crystallisation can be implicitly included in the energy equation
-(Eq. \ref{eq:energy}) by considering effective specific heat ($C_{p_{eff}}$) and thermal expansion ($\alpha_{eff}$) of the partially molten rocks, in according to
-\begin{eqnarray}
-\label{eq:cp_effective}C_{p_{eff}}&=&C_p+H_L\left(\frac{\partial M}{\partial T}\right)_p \\
-\label{eq:alpha_effective}\alpha_{eff}&=&\alpha+\rho\frac{H_L}{T}\left(\frac{\partial M}{\partial p}\right)_T
-\end{eqnarray}
-where $C_p$ and $\alpha$ are the specific heat and the coefficient of thermal expansion of the solid rock, respectively, and $H_L$ is the latent heat of the
-molten rock \citep{Gerya2004b,Gerya2007,Tackley2010}.
-
-\section{Benchmarks}\label{sec:benchmark}
-\subsection{Error measurements}\label{sec:error}
-In order to determine the accuracy of velocity and pressure field of the benchmarks, the $L_2$-norm is computed by numerical integration on the Gauss-Legendre
-quadrature points. $L_2$-norm for pressure and velocity errors can be evaluated as
-\begin{eqnarray}
-\textrm{err}_p&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|p^n_i(r_q)-p^a_i(r_q)|^2\right)w_q|J_q|}\nonumber \\
-\textrm{err}_v&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|u^n_i(r_q)-u^a_i(r_q)|^2+|v^n_i(r_q)-v^a_i(r_q)|^2\right)w_q|J_q|}\nonumber
-\end{eqnarray}
-respectively, where $n_e$ is the number of elements, $n_q$ is the number of quadrature points per element, $p^n_i(r_q)$ and $p^a_i(r_q)$ are the numerical and
-analytical pressure, respectively, in each quadrature point $q$, $w_q$  and $J_q$ are the weight and the Jacobian at the quadrature point
-$q$, $u^n_i(r_q)$, $v^n_i(r_q)$, $u^a_i(r_q)$ and $v^a_i(r_q)$ are the numerical and analytical velocities, respectively, in each quadrature point $q$.
-
-Other quantities used as comparison with original benchmarks are the root-mean-square velocity over the whole domain and over the surface
-\begin{eqnarray}
-v_{\textrm{rms}}&=&\sqrt{\int_{\Omega}|\bm{v}|^2 d\Omega}\nonumber \\
-v_{\textrm{rms}}^{top}&=&\sqrt{\int_0^1 u^2 \bigg|_{y=1} dx}\nonumber
-\end{eqnarray}
-respectively, top and bottom Nusselt numbers
-\[\textrm{Nu}^{top/bottom}=-\int_0^1\frac{\partial T}{\partial y} \bigg|_{y=1/y=0} dx\]
-the average rate of work done against gravity
-\[\left\langle W \right\rangle=\int_{\Omega}T u_y d\Omega\]
-the average rate of viscous dissipation
-\[\left\langle \Phi \right\rangle=\int_{\Omega}\tau_{ij}\dot{\epsilon}_{ij} d\Omega\]
-and the percentage error between $\left\langle W \right\rangle$ and $\left\langle \Phi \right\rangle/Ra$
-\[\delta=\frac{\left|\left\langle W \right\rangle -\frac{\left\langle \Phi \right\rangle}{Ra} \right|}{\max \left(\left\langle W \right\rangle,
-\frac{\left\langle \Phi \right\rangle}{Ra}\right)}\times100\]
-
-\subsection{Stokes flow}\label{sec:stokes}
-The problem consists of determining velocity field $(u,v)$ and pressure $p$ in case of a manufactured solution with prescribed body forces such as
-\begin{eqnarray}
-b_1&=&(12-24y)x^4+(-24+48y)x^3+(-48y+72y^2-48y^3+12)x^2+\nonumber \\
-&&+(-2+24y-72y^2+48y^3)x+1-4y+12y^2-8y^3\nonumber \\
-b_2&=&(8-48y+48y^2)x^3+(-12+72y-72y^2)x^2+(4-24y+48y^2-48y^3+\nonumber \\
-&&+24y^4)x-12y^2+24y^3-12y^4\nonumber
-\end{eqnarray}
-for which the exact solution is:
-\begin{eqnarray}
-u(x,y)&=&x^2(1-x)^2(2y-6y^2+4y^3)\nonumber \\
-v(x,y)&=&-y^2(1-y)^2(2x-6x^2+4x^3)\nonumber \\
-p(x,y)&=&x(1-x)-\frac{1}{6}\nonumber
-\end{eqnarray}
-The domain is a unit square a constant viscosity ($\eta=1$) and the penalty parameter is set to $\lambda=10^7$. Velocity boundary conditions are set to no slip
-$(\bm{v}=\bm{0})$ on all boundaries. The problem is performed for different grid resolution between \num{8x8} and \num{1024x1024} elements.
-
-The errors between the analytical solution and the numerical prediction of the velocity and the pressure field are calculated as in Section \ref{sec:error}.
-Fig. \ref{fig:errors}a shows that both velocity and pressure field converge to the exact solution with the decrease of the element size, following the
-theoretical convergence rate. The convergence of pressure, in contrast with the observation by \citet{Donea2003} for $Q_1 \times P_0$ elements, indicate the
-effectiveness of the smoothing procedure. Smoothed pressure field for grid resolution of \num{128x128} elements is shown in Fig. \ref{fig:errors}b (continuous
-red line), in comparison with the analytical solution (dashed blue line). Solve times and memory usage needed to generate the solution are shown in
-Fig. \ref{fig:MUMPS}. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Solver/Stokes_Flow}.
-
-\begin{figure}
-\centering
-\includegraphics[width=\linewidth]{./Figures/errors.pdf}
-\caption{Velocity and pressure error for the Stokes flow experiment between generated and analytical solution as a function of element size (panel a),
-and comparison between smoothed pressure and analytical solution as function of $x$ coordinate for a grid resolution of \num{128x128} elements (panel b).}
-\label{fig:errors}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\linewidth]{./Figures/MUMPS.pdf}
-\caption{Panel a) Analysis, factorisation, solution and total solve times (panel a) and factorisation memory usage (panel b) as a function of the total number
-of degrees of freedom for the Stokes flow experiment.}
-\label{fig:MUMPS}
-\end{figure}
-
-\subsection{Markers advection}\label{sec:runge}
-The 2nd-order Runge-Kutta advection scheme in space is tested by means of the Zalesak disk test \citep{Zalesak1979,Thieulot2014}. The benchmark is performed
-in a unit square domain with a grid resolution of \num{32x32} elements and 1000000 markers and values of Courant number between 0.25 and 1.
-At $t=0$, the disk is centred at position $(0.5;0.75)$ with a radius $R=0.15$ and has a vertical fissure 0.05 wide and 0.2 high (orange in Fig. \ref{fig:runge})
-and with the velocity field prescribed in the entire domain as
-\begin{eqnarray}
-u(x,y)&=&2\pi\left(y-\frac{L_x}{2}\right)\nonumber \\
-v(x,y)&=&-2\pi\left(x-\frac{L_x}{2}\right)\nonumber
-\end{eqnarray}
-the markers are back to their initial location after a $2\pi$ rotation ($t=1$). The clockwise rotation of the disk at $t=0.25$, 0.5 and 0.75 
-is shown in Fig. \ref{fig:runge} (red, blue and green, respectively). The distance from the centre of 1 marker is calculated throughout the rotation to
-evaluate the error for different values of Courant number. As expected, the error increase with the increasing of the time step, as shown in 
-Fig. \ref{fig:runge_err}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/Zalesak_Disk}.
-
-\begin{figure}
-\centering
-\includegraphics[width=200px]{./Figures/Zalesak.pdf}
-\caption{Position of the Zalesak disk at $t=0$, 0.25, 0.5 and 0.75 (orange, red, blue and green, respectively) for the case with Courant number of 1.}
-\label{fig:runge}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/zal_marker.pdf}
-\caption{Distance from the centre as function of time for values of Courant number of  of 0.25, 0.3, 0.5, 0.75 and 1 (green, orange, blue, black and red,
-respectively). R indicates the distance from the centre considering a perfect circle trajectory.}
-\label{fig:runge_err}
-\end{figure}
-
-\subsection{Conservative Velocity Interpolation (CVI)}\label{sec:CVI}
-The CVI correction is checked by means of the Stokes flow experiment by \citet{Donea2003}, as presented in \ref{sec:stokes}. The advection of Lagrangian
-markers is performed using either a 2nd-order or a 4th-order Runge-Kutta scheme with CFL=0.5 and an initial random distribution of 25 markers per element.
-Fig. \ref{fig:CVI} shows the minimum and maximum values of markers per element throughout the experiment with and without the CVI correction and using
-different Runge-Kutta schemes. The introduction of the CVI correction allows the
-tests to run for times longer than $t=2000$ without show empty elements, while tests without the CVI correction stop at approximately $t=600$, when at least
-one element goes empty. Fig. \ref{fig:CVI_time} shows the different number of markers per element and the markers distribution at $t=600$ obtained without and
-with the CVI correction (panels b and c, respectively) in case of a 2nd-order Runge-Kutta scheme. All data can be found at
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/CVI}.
-\begin{figure}
-\centering
-\includegraphics[width=\linewidth]{./Figures/CVI.pdf}
-\caption{Maximum and minimum number of markers per element throughout the experiment, with (green and orange lines) or without (purple and blue lines)
-the CVI correction and using either a 2nd-order (blue and orange lines) or a 4th-order (purple and green lines) Runge-Kutta scheme.}
-\label{fig:CVI}
-\end{figure}
-\begin{figure}
-\centering
-\includegraphics[width=\linewidth]{./Figures/CVI_time.pdf}
-\caption{Number of markers per element and markers distribution through time using 2nd-order Runge-Kutta scheme.
-Panels a: number of markers per element (panel a1) and distribution of the markers (panel a2) after the first
-iteration; panels b and c: comparison between number of markers per element (panels b1 and c1) and markers distribution (panels b2 and c2) without and with
-the CVI correction at $t=600$; panels d: number of markers per element (panel d1) and markers distribution (panel d2) in case of CVI correction at $t=2000$.}
-\label{fig:CVI_time}
-\end{figure}
-
-\subsection{Poiseuille flow}\label{sec:poiseuille}
-The domain is a rectangle with $L_x=2$ and $L_y=1$ and constant density and viscosity ($\rho=1$ and $\eta=1$), gravity acceleration $\bm{g}=\bm{0}$ and penalty
-parameter $\lambda=10^8$. The grid is composed by \num{40x20} elements. Velocity boundary conditions are set to no slip $(\bm{v}=\bm{0})$ at the top and the
-bottom, and a parabolic profile is imposed on the sides, with $u=y(1-y)$ and $v=0$. The analytical solution is then given by:
-\begin{eqnarray}
-u(x,y)&=&y(1-y)\nonumber \\
-v(x,y,)&=&0\nonumber \\
-p(x,y)&=&2\eta\left(\frac{L_x}{2}-x\right)\nonumber
-\end{eqnarray}
-
-The velocity field predicted by the model follows the expected parabolic profile (Fig. \ref{fig:poiseuille}a and Fig. \ref{fig:poi_plot}a) and, in case of the
-classic penalty method with no iterations, the pressure is clearly related to the divergence by means of the penalty parameter (Fig. \ref{fig:poiseuille}b and
-c and Fig. \ref{fig:poi_plot}b). Fig. \ref{fig:poiseuille}d shows that one Uzawa iteration is sufficient to bring the divergence down to \num{e-15}, with no
-correlation with the pressure field. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Poiseuille%20Flow}.
-
-\begin{figure}
-\includegraphics[width=\linewidth]{./Figures/Poiseuille.pdf}
-\caption{Velocity field (panel a), pressure (panel b) and divergence velocity of a Poiseuille flow in case of the classic penalty method (no iterations) and
-after one Uzawa iteration (panel c and d, respectively).}
-\label{fig:poiseuille}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\linewidth]{./Figures/Analytical.pdf}
-\caption{Velocity (panel a) and pressure (panel b) field predicted by the model for a Poiseuille flow with respect with their analytical solutions. The
-velocity field is plotted as function of the vertical coordinate in $L_x/2$ and the pressure is plotted as function of the x coordinate in $L_y/2$.}
-\label{fig:poi_plot}
-\end{figure}
-
-\subsection{Instantaneous 2D sphere}\label{sec:ist_sphere}
-The domain is a unit square with gravity $\bm{g}=(0,-1)$. The fluid has constant density and viscosity ($\rho_f=1$ and $\eta_f=1$). The sphere is in the middle
-of the domain with a radius R = 0.123456798 and has constant density and viscosity ($\rho_s=10^{-2}+\rho_f$ and $\eta_s=10^3 \cdot \eta_f$).
-We distinguish three different types of velocity boundary conditions:
-\begin{enumerate}
-\item FS: free slip conditions on all sides.
-\item NS: no slip conditions on all sides.
-\item OT: free slip conditions on the sides and the bottom, and open boundary at the top.
-\end{enumerate}
-All three cases are performed for different grid resolutions, between \num{16x16} and \num{512x512} elements with 50 randomly distributed markers per element,
-and different average schemes for the viscosity (harmonic, geometric and arithmetic). The velocity in the centre of the sphere $\bm{v}(0.5;0.5)$), minimum and
-maximum velocities ($u_{min}$, $u_{max}$, $v_{min}$, $v_{max}$) and pressures ($p_{min}$, $p_{max}$) on the entire domain, $v_{\textrm{rms}}$ and average pressure on the
-entire domain ($p_{avg}$) are compared with solutions generated by ASPECT \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and all results
-can be found at \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere2D} and 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Instantaneous_2D_Sphere}. Results in terms of the $v_{\textrm{rms}}$ for NS model 
-are also shown in Fig. \ref{fig:inst_sphere}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/vrms_NS.pdf}
-\caption{$v_{\textrm{rms}}$ as function of element size for the instantaneous 2D sphere experiment in case of no slip boundary conditions and with different average
-schemes for the viscosity. Results are compared with results obtained by other numerical codes.}
-\label{fig:inst_sphere}
-\end{figure}
-
-\subsection{Rayleigh-Taylor instability}\label{sec:rayleigh}
-This problem has been originally presented by \citet{vanKeken1997} and here it is performed as the isoviscous case (Case 1a). The domain has $L_x=0.9142$ and
-$L_y=1$ and gravity $\bm{g}=(0,-1)$. Two fluids with same constant viscosities ($\eta_1=\eta_2=1$) and different densities ($\rho_1=1000$ and $\rho_2=1010$),
-with the lighter fluid at the bottom. The initial interface between the fluids is given by $y(x)=0.2+0.02 \cos \left(\frac{\pi x}{L_x}\right)$. The experiment
-is performed with different grid sizes (\num{50x50}, \num{80x80}, \num{100x100} and \num{256x256}). A total of 1960000 markers are randomly distributed at the
-beginning of the simulation. Velocity boundary conditions are set to no slip at the top and the bottom, and to free slip at the sides of the domain.
-
-$v_{\textrm{rms}}$ as function of time is reported in Fig. \ref{fig:RT}, matching well with results shown by \citet{vanKeken1997}, \citet{Tackley2003} and
-\citet{Thieulot2014}. Fig. \ref{fig:rayleigh} shows the evolution of the experiment at different time steps. All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Rayleigh_Taylor_experiment/ISOVISCOUS}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/RT.pdf}
-\caption{$v_{\textrm{rms}}$ of the Rayleigh-Taylor experiment as function of time for different resolution of the grid.}
-\label{fig:RT}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=300px]{./Figures/Rayleigh.pdf}
-\caption{Evolution of the Rayleigh-Taylor experiment for a grid of \num{256x256} elements at $t=0, 500, 1000$ and $2000$ (panels a, b, c and d, respectively).}
-\label{fig:rayleigh}
-\end{figure}
-
-\subsection{Falling blocks}\label{sec:block}
-This benchmark is described as proposed by \citet{Gerya2003a}, \citet{Gerya2010} and \citet{Thieulot2011}. The domain is square with $L_x=L_y=$ \SI{500}{\km}
-and the grid is composed by \num{50x50} elements with 25 markers in each element. The block is initially centred at ($x=$ \SI{250}{\km}; $y=$ \SI{400}{\km})
-and has a size of \num{100x100} km (Fig. \ref{fig:block}a). The fluid surrounding the block has $\eta_f=$ \SI{e21}{\pascal\per\s} and $\rho_f=$
-\SI{3200}{\kg\per\cubic\m}. The benchmark tests with different viscosities of the block, with $\eta_b$ from \SIrange{e15}{e27}{\pascal\per\s}. In each
-experiment also the density $\rho_b$ of the block varies from \SIrange{3220}{9900}{\kg\per\cubic\m}. Velocity boundary conditions are set to free slip
-conditions on all sides of the domain. Density distributions at $t=$ \SI{20}{\mega\year} for $\rho_b=$ \SI{3300}{\kg\per\cubic\m} and different $\eta_b$
-are plotted in Fig. \ref{fig:block}b-f, showing that the code correctly preserve the block geometry in case of large viscosity contrast, with a block stiffer
-than the surrounding fluid (Fig. \ref{fig:block}f).
-
-The velocity in the centre of the falling block at $t=0$ is measured for all experiments and, since the velocity should increase with density contrast, the
-quantity $\bm{v}/(\rho_b-\rho_f)$ is plotted as function of the viscosity contrast. All results perfectly match with those from \citet{Gerya2010} and they
-line up on a single curve, demonstrating that the code can correctly deal with large viscosity and density contrasts (Fig. \ref{fig:falling}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Falling%20blocks}.
-
-\begin{figure}
-\includegraphics[width=300px]{./Figures/Block.pdf}
-\centering
-\caption{Density evolution of the falling block experiment at $t=$ \SI{0}{\mega\year} (panel a) and $t=$ \SI{20}{\mega\year} for different viscosities of the
-block (panels b-f).}
-\label{fig:block}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Falling.pdf}
-\caption{Initial velocity relative to the density contrast at the centre of the falling block as function of the viscosity contrast between the block and the
-surrounding fluid.}
-\label{fig:falling}
-\end{figure}
-
-\subsection{2D time-dependent Stokes sphere with deformable free surface}\label{sec:sphere}
-This experiment is performed in a unit square domain with the gravity acceleration fixed to $g_y=-1$. The sphere has $\rho_s=2$ and $\eta_s=10^3$ and is
-initially centred at ($0.5;0.6$) with radius $R=0.123456789$. The fluid surrounding the sphere has $\rho_f=1$ and $\eta_f=1$ and occupies the domain for
-$y\leq 0.75$, while for $y>0.75$ the air has $\rho_a=0$ and $\eta_a=10^{-3}$. Velocity boundary conditions are set to free slip on all sides. The Courant
-number is set to 0.25. The experiments run for 200 s using grid resolutions from \num{150x150} to \num{512x512} elements, with 25 randomly distributed markers
-per element, and for different average schemes for the viscosity. The interface between the fluid and the air is tracked by means of the markers chain. Results
-in terms of velocity and pressure fields and topography variations are compared with results obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and they can be found at
-\url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere_fs2D} and
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Time_dependent_sphere}. Results in terms of the $v_{\textrm{rms}}$ in case of 
-an arithmetic average are also shown in Fig. \ref{fig:inst_spherefs}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/vrms_arithm.pdf}
-\caption{$v_{\textrm{rms}}$ for different grid resolutions as function of time for the 2D time-dependent Stokes sphere experiment in case of an arithmetic average.
-Results are compared with results obtained by ASPECT.}
-\label{fig:inst_spherefs}
-\end{figure}
-
-\subsection{Free surface stabilisation}\label{sec:stab}
-The benchmark is performed as discussed in \citet{Kaus2010a} and \citet{Thieulot2014}. The domain is square with $L_x=L_y=$ \SI{500}{\km} and a grid resolution
-of \num{200x200} elements, each of them containing 16 markers. A buoyant fluid with $\rho_1=$ \SI{3200}{\kg\per\cubic\metre} and $\eta_1=$ \SI{e20}{\pascal\s}
-is overlain by a denser fluid with $\rho_2=$ \SI{3300}{\kg\per\cubic\metre} and $\eta_2=$ \SI{e21}{\pascal\s}. The initial interface between the fluids has an
-initial sinusoidal shape of 5 km amplitude. Velocity boundary conditions are set to free slip conditions at the sides and no slip at the bottom, while the top
-is free surface. The experiment is performed with various fixed time steps, without the use of the Courant number. The vertical position of the free surface at
-$x=L_x$ is tracked for each simulation, with or without the stabilisation algorithm.
-
-The results show an instability (drunken sailor effect) increasing the time step in case that the stabilisation algorithm is not activated
-($dt=$ \SI{4200}{\year}, green line in Fig. \ref{fig:stab}), as already observed by \citet{Kaus2010a} and \citet{Thieulot2014}. Activating the algorithm the
-instability is fixed and the simulation remains stable using time step up to \SI{50000}{\year} (red line in Fig. \ref{fig:stab}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Stabilisation_algorithm}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Kaus_random.pdf}
-\caption{Evolution of the y coordinate in $x=L_x$ as function of time for different time steps, with and without the stabilisation algorithm.}
-\label{fig:stab}
-\end{figure}
-
-\subsection{Topography relaxation}\label{sec:crameri}
-This experiment is performed as in \citet{Crameri2012} using free surface and sticky air (Case 1). The domain is rectangular with $L_x=$ \SI{2800}{\km} and
-$L_y=$ \SI{700}{\km} in case of free surface and $L_y=$ \SI{800}{\km} in case of sticky air ($h_{st}=$ \SI{100}{\km}). The grid is composed by \num{256x64}
-elements for the free surface \num{560x320} elements for the sticky air. For both cases, there is a mantle of 600 km thickness with $\eta_m=$ \SI{e21}{\pascal\s}
-overlain by a highly viscous cosine-shaped lithosphere with thickness between 93 and 107 km and $\eta_l=$ \SI{e23}{\pascal\s}. Both mediums have $\rho=$
-\SI{3300}{\kg\per\cubic\metre}. In case of sticky air experiment, the air has $\rho_a=$ \SI{0}{\kg\per\cubic\metre} and $\eta_a=$ \SI{e18}{\pascal\s}. The
-gravity acceleration is set to $g_y=$ \SI{-10}{\metre\per\square\s}. Velocity boundary conditions are set to no slip at the bottom and free slip at the sides
-of the domain. In case of sticky air, velocities on top are set to free slip conditions.
-
-The maximum topography as function of time ($h(t)$) can be analytically derived in according to \[h(t)=h_0 \exp(\gamma t)\]
-where $h_0=7$ km is the initial topography, $\gamma=\num{-0.2139e-11}$ is the characteristic relaxation rate and $t$ is the time (black line in Fig.
-\ref{fig:crameri}). Maximum topography at the characteristic relaxation time $t_{rlx}=$ \SI{14.825}{\kilo\year} can be found to be $h_{rlx}=$ \SI{2576}{\m}.
-The results show that a viscosity of the air of \SI{e18}{\pascal\s} is needed in case of $h_{st}=$ \SI{100}{\km} to correctly describe the topography relaxation
-for this problem, as pointed out by \citet{Crameri2012} (purple line in Fig. \ref{fig:crameri}). The case with a true free surface well follows the analytical
-solution (orange line in Fig. \ref{fig:crameri}). In particular, at the relaxation time in case of the true free surface predicts a maximum topography of
-\SI{2570}{\m}, with an error of \SI{6}{\m} with respect to the analytical solution. Results are also compared with results obtained with a free surface in
-UNDERWORLD (\num{256x64} elements) and SULEC (\num{401x201} elements) (with courtesy from \citealp{Crameri2012}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Topography_relaxation}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Crameri.pdf}
-\caption{Maximum topography as function of time for the Crameri benchmark (coloured lines) in comparison with the analytical solution (black line) and results
-obtained with UNDERWORLD (yellow line) and SULEC (blue line).}
-\label{fig:crameri}
-\end{figure}
-
-\subsection{Spontaneous subduction}\label{sec:subduction}
-This experiment is performed as presented by \citet{Schmeling2008}, in case of both a sticky air and a true free surface and using different average schemes
-for the viscosities. The domain is rectangular with $L_x=$ \SI{3000}{\km} and $L_y=$ \SI{700}{\km}, a grid resolution of \num{256x64} elements with 50 markers
-per element and Courant number of 0.01. At the beginning of the simulation, a 100 km-thick lithospheric layer with $\rho_m=$ \SI{3300}{\kg\per\cubic\metre}
-and $\eta_m=$ \SI{e23}{\pascal\s} is located at the top of the domain for $x$ between 1000 and 3000 km, while the asthenospheric mantle has $\rho_m=$
-\SI{3200}{\pascal\s} and $\eta_m=$ \SI{e21}{\pascal\s}. In addition, a 200 km-depth lithospheric slab is already subducted in the mantle in order to have
-a spontaneous subduction. In case of sticky air, $L_y=750$ km and a 50 km-thick air layer with $\eta_a$ from \SIrange{e19}{e21}{\pascal\s} overlie the mantle.
-Velocity boundary conditions are set to free slip on the sides and on the bottom of the domain. In case of sticky air, velocities on the top boundary are set
-to free slip conditions as well.
-
-The maximum depth of the slab is tracked for all the simulations (Fig. \ref{fig:subduction}) and they are compared with results obtained by
-\citet{Schmeling2008} with models with regular grids and comparable grid resolutions (rectangles in Fig. \ref{fig:subduction}). Low viscous sticky air models
-(i.e., \SI{e19}{\pascal\s}) enlighten a strong dependence of the sinking velocity with the chosen average scheme (dashed lines in Fig. \ref{fig:subduction}),
-while high viscous sticky air models (i.e., \SI{e21}{\pascal\s}) show a higher resistance at the trench, underestimating the correct solution (dotted lines in
-Fig. \ref{fig:subduction}). A high resistance at the trench can be also observed in case of a true free surface for the low resolution grid used (continuous
-lines in Fig. \ref{fig:subduction}), as pointed out by \citet{Schmeling2008}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Spontaneous_subduction}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Subduction.pdf}
-\caption{Maximum depth of the slab as function of time for all the simulations of the \citet{Schmeling2008} benchmark, in case of harmonic, geometric and
-arithmetic means (red, green and blue lines, respectively) using a sticky air or a true free surface (discontinuous and continuous lines, respectively).
-Red, green and blue rectangular areas indicate the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{500}{\km} in case of sticky air,
-while the grey rectangular area indicates the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{600}{\km} in case of true free surface.}
-\label{fig:subduction}
-\end{figure}
-
-\subsection{The slab detachment}\label{sec:slab}
-The slab detachment benchmark is performed as described by \citet{Schmalholz2011} and \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{1000}{\km}
-and $L_y=$ \SI{660}{\km} and a grid resolution of \num{256x256} elements. A non-linear viscous T-shaped layer with $\rho_l=$ \SI{3300}{\kg\per\cubic\metre} is
-placed at the top of the domain and surrounded by a linear viscous fluid with $\rho_f=$ \SI{3150}{\kg\per\cubic\m}. The top layer is 80 km-thick and a 250
-km-long and 80 km-wide slab is placed at $x=L_x/2$ (Fig. \ref{fig:slab}a). The effective viscosity of the top layer is given by
-\[\eta_{eff}=\eta_0 I_2^{\frac{(1-n)}{n}}\]
-with $\eta_0=$ \SI{4.75e11}{\pascal\s} and $n=4$, while the surrounding fluid has $\eta_f=$ \SI{e21}{\pascal\s}. The effective viscosity is capped between
-$\eta_{min}=$ \SI{e21}{\pascal\s} and $\eta_{max}=$ \SI{e25}{\pascal\s}. Velocity boundary conditions are set to free slip at the top and the bottom, and to
-no slip at the sides of the domain.
-
-The sinking of the slab determines a decrease of the effective viscosity in the T-shaped layer during the first part of the evolution (Fig. \ref{fig:slab}a),
-as consequence of local strain rates, while effective viscosities increase up to $\eta_{max}$ in the top layer after approximately \SI{20}{\mega\year}, when
-necking is complete (Fig. \ref{fig:slab}b).
-
-The width of necking is tracked by means of markers position at the side of the slab. Necking width and time are normalised in function of the initial slab
-(\SI{80}{\km}) and the characteristic time ($t_c=$ \SI{7.1158e14}{\s}) \citep{Schmalholz2011,Glerum2018}. The results in case of arithmetic, geometric and
-harmonic mean of the viscosity are shown in Fig. \ref{fig:necking}, compared with results from \citet{Glerum2018} for which an infinity norm mean was used.
-The influence of the average schemes agrees with results shown in Section \ref{sec:subduction}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Slab_detachment}.
-
-\begin{figure}
-\includegraphics[width=\linewidth]{./Figures/Slab.pdf}
-\caption{Effective viscosity for the slab detachment benchmark at the beginning of the evolution (panel a) and when necking is complete (panel b).}
-\label{fig:slab}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Necking.pdf}
-\caption{Normalised width of the necking of the slab detachment benchmark as function of normalised time.}
-\label{fig:necking}
-\end{figure}
-
-\subsection{The indenter experiment}\label{sec:indenter}
-The indenter benchmark simulate a rigid punch on a purely plastic von Mises material, for which exists an analytical solution
-\citep{Thieulot2008,Thieulot2014,Glerum2018}. The domain is a unit square with a grid resolution of \num{256x256} elements. The gravity acceleration is set
-to $\bm{g}=\bm{0}$ and effective viscosity is capped using $\eta_{min}=10^{-4}$ and $\eta_{max}=10^3$. The tolerance for the convergence of the non-linear
-solution is $tol=10^{-9}$, with a maximum number of non-linear iterations set to 500. The medium in the domain has $\rho=0.01$, an initial viscosity
-$\eta_0=10$, cohesion $C=1$ and an angle of internal friction $\phi=0$°. Velocity boundary conditions are set to no slip at the bottom and free slip at the
-sides of the domain. The top of the domain is open with the exception of the central portion (punch area) with width $w_p=0.125$, where $v=-1.05$ and $u$ is
-fixed either to 0 (rough punch experiment) or free (smooth punch experiment).
-
-Fig. \ref{fig:indenter} show the results in terms of viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure
-(panels g and h) for both experiments. The analytical solution indicate that pressure at ($0.5;1$) and ($0.5 \pm w_p;1$)is $p=\pi +1$ and $p=1$, respectively
-(black and red lines in Fig. \ref{fig:smooth_p}). Fig. \ref{fig:smooth_p} shows a clear improvement in the pressure solution when passing from a rough to a
-smooth punch (green and blue lines in Fig. \ref{fig:smooth_p}, respectively), as previously observed by \citet{Thieulot2014} and \citet{Glerum2018}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Indenter}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Indenter.pdf}
-\caption{Viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure (panels g and h) fields for rough (left column) and
-smooth (right column) punch experiments.}
-\label{fig:indenter}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Smooth_Rough.pdf}
-\caption{Nodal pressure in function of the x coordinate at $y=1$ for rough (green line) and smooth (blue line) punch experiment. Black and red lines indicate
-the analytical solution at $x=0.5$ and $x=0.5 \pm w_p$, respectively.}
-\label{fig:smooth_p}
-\end{figure}
-
-\subsection{The brick experiment}\label{sec:brick}
-An instantaneous version of the brick benchmark is performed to verify the correctness of the pressure-dependent plasticity for different angle of internal
-friction, as proposed by \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{40}{\km} and $L_y=$ \SI{10}{\km} and a grid resolution of \num{512x128}
-elements. The gravity acceleration is set to $g_y=$ \SI{-10}{\m\square\s} and effective viscosity is capped using $\eta_{min}=$ \SI{e19}{\pascal\s} and
-$\eta_{max}=$ \SI{e26}{\pascal\s}. The tolerance for the convergence of the non-linear solution is $tol=10^{-7}$, with a maximum number of non-linear iterations
-set to 1000. A 800 m-wide and 400 m-height inclusion with $\eta_b=$ \SI{e20}{\pascal\s} and $\rho_b=$ \SI{2700}{\kg\per\cubic\m} is placed at the bottom of the
-domain at $x=L_x/2$. The inclusion is surrounded by a non-linear viscous medium with $\rho_m=$ \SI{2700}{\kg\per\cubic\m}, an initial viscosity of
-\SI{e23}{\pascal\s}, a linear viscous viscosity of \SI{e25}{\pascal\s} and a cohesion of \SI{40}{\mega\pascal}. Velocities are set to free slip conditions at
-the bottom of the domain and the top is open. Velocities on sides of the domain are fixed to $u=$ \SI{+-2e-11}{\m\per\s} and $v=0$. The experiment is performed
-in compressional and extensional contexts with an angle of internal friction $\phi$ of the non-linear medium variable between 0° and 30°.
-
-As expected, two shear bands stem from the inclusion with variable angles in relation to both the dynamics context and the internal friction angle. Shear band
-angles formed at 45° for $\phi=0$° in both extensional and compressional contexts (Fig. \ref{fig:brick_beam}a), while they are different in case of $\phi\neq0$°.
-Shear band angles in case of $\phi=20$° are shown for compression and extension in Fig. \ref{fig:brick_beam}b  and c, respectively. Values of shear band angles
-as function of internal friction angles from 0° to 30° are extracted at different depths and minimum and maximum values are plotted in Fig. \ref{fig:brick},
-in comparison with theoretical Roscoe, Arthur and Coulomb shear band angles. For all tests the tolerance for nonlinear convergence is defined $tol=10^{-7}$
-and none converges before the maximum number of iterations ($it_{max}=1000$) is reached. However, tests with internal friction angles up to 15° show a constant
-decrease of the velocity residuals, which can not be observed for for tests with higher internal friction angles (Fig. \ref{fig:convergence}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Brick_experiment}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Brick_Beam.pdf}
-\caption{Shear band angles predicted for the brick experiment for $\phi=0$° (panel a) and for $\phi=20$° in case of compression and extension (panels b and c,
-respectively).}
-\label{fig:brick_beam}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Brick.pdf}
-\caption{Shear band angles predicted for the brick experiment in case of compressional and extensional contexts as function of different internal angle of
-friction (continuous black line and red dots), compared with theoretical Roscoe, Artur and Coulomb angles (discontinuous black lines). Red lines indicate the
-range of angles calculated at different depths.}
-\label{fig:brick}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=500px]{./Figures/convergence.pdf}
-  \caption{Convergence rates for all tests. Normalised residuals (left column) are compared with velocity residuals (right column) for extensional (first row)
-  and compressional (second row) settings.}
-  \label{fig:convergence}
-  \end{figure}
-
-\subsection{Advection stabilisation}\label{sec:advection}
-The 1D advection problem proposed by \citet{Donea2003} and \citet{Thieulot2011} is performed to verify the effectiveness of the SUPG method to stabilise the
-advection term of the energy equation. The domain is a 1D segment with $L_x=1$ composed by 50 elements and a discontinuity in the thermal field placed at
-$x=0.25$. Temperature is set to 1 for $x \leq 0.25$ and to 0 for $x > 0.25$. Velocity is set to ${u}=1$ in the entire domain. The simulation is performed for
-250 time steps, with $dt=0.002$, so the thermal discontinuity should be at $x=0.75$ at the end of the simulation.
-
-Temperature profiles at the end of the simulation are shown in \ref{fig:advection} as function of the dimensionless coefficient $\gamma=\tau \bm{v}/h$. In case
-of the classic Galerkin method ($\gamma=0$, blue line in \ref{fig:advection}) the final thermal profile is characterised by strong oscillations, which are
-eliminated in case of the SUPG method ($\gamma=0.045$, orange line in \ref{fig:advection}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Advection_stabilisation}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Advection.pdf}
-\caption{Temperature profile as function of \textit{x} for the advection stabilisation benchmark. Purple line indicates the initial temperature profile; the
-green line indicates the analytical temperature profile after 250 time steps; blue line indicates the temperature profile after 250 time steps in case of the
-classic Galerkin method ($\gamma=0$); orange line indicates the temperature profile after 250 time steps in case of the SUPG method ($\gamma=0.045$).}
-\label{fig:advection}
-\end{figure}
-
-\subsection{Shear heating}\label{sec:simple_shear}
-This exercise is performed in an unit square domain composed by \num{128x128} elements. The velocity field is prescribed on the entire domain with
-$\bm{v}=(L_y-y)y\bm{e_x}$; viscosity, density and specific heat are fixed to 1, while thermal conductivity and radiogenic energy are fixed to 0. Therefore,
-the energy equation (Eq. \ref{eq:energy}) can be simplified as \[\frac{\partial T}{\partial t}=H_s\] and fixing $T(t=0)=0$, the temperature can be find as
-\[T(t)=H_s t\]
-In this case we have
-\begin{eqnarray}
-\dot{\epsilon_{xx}}&=&\frac{\partial u}{\partial x}=0 \nonumber \\
-\dot{\epsilon_{yy}}&=&\frac{\partial v}{\partial y}=0 \nonumber \\
-\dot{\epsilon_{xy}}&=&\frac{1}{2}\left(\frac{\partial v}{\partial x}+\frac{\partial u}{\partial y}\right)=\frac{1}{2}(L_y-2y)\nonumber
-\end{eqnarray}
-and, simplifying Eq. \ref{eq:shear}, shear heating can be calculated as \[H_s(x,y)=(1-2y)^2\]
-
-The solution predicted by the model in terms of velocity, temperature and shear heating match well with the analytical solutions (Fig. \ref{fig:analytical_en}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Simple_shear}.
-
-%\begin{figure}
-%\centering
-%\includegraphics[width=400px]{./Figures/Energy_Thieulot.pdf}
-%\caption{Temperature and shear heating predicted by the model after 1 time step.}
-%\label{fig:en_thieulot}
-%\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=500px]{./Figures/Analytical_Shear.pdf}
-\caption{Velocity (panel a), temperature (panel b) and shear heating (panel c) as function of the \textit{y} coordinate for the simple shear experiment. The
-solutions predicted by the model (red lines) are compared with the analytical solutions (dashed blue lines).}
-\label{fig:analytical_en}
-\end{figure}
-
-\subsection{Shear and adiabatic heating}\label{sec:shear}
-This problem is performed as presented in Exercise 9.4 in \citet{Gerya2010}. Two materials are vertically separated in a rectangular domain with $L_x=$
-\SI{1000}{\km}, $L_y=$ \SI{1500}{\km} and a grid resolution of \num{30x20} elements. Constant thermal coefficient expansion ($\alpha=$ \SI{3e-5}{\per\kelvin}),
-temperature ($T=$ \SI{1300}{\kelvin}) and gravity acceleration ($g_y=$ \SI{-10}{\m\square\s}) are assumed in the whole domain. Fluid 1 (on the left side of the
-domain) has $\rho_1=$ \SI{3200}{\kg\per\cubic\m} and $\eta_1=$ \SI{e20}{\pascal\s}; fluid 2 (on the right side of the domain) has $\rho_2=$
-\SI{3300}{\kg\per\cubic\m} and $\eta_2=$ \SI{e22}{\pascal\s}. Velocity boundary conditions are set to free slip on all sides of the domain.
-
-As shown in Fig. \ref{fig:energy}, both shear and adiabatic heating predicted by the code (first row in Fig. \ref{fig:energy}) well recreate the results
-obtained by means of example ${Shear\_adiabatic\_heating.m}$ from \citet{Gerya2010} (second row in Fig. \ref{fig:energy}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Adiabatic%2BShear}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Energy.pdf}
-\caption{Comparison between shear (first column), adiabatic (second column) and total (third column) energy predicted by the code (panels a, b and c) and those
-created using example ${Shear\_adiabatic\_heating.m}$ from Exercise 9.4 in \citet{Gerya2010} (panels d, e and f).}
-\label{fig:energy}
-\end{figure}
-
-\subsection{Mantle convection}\label{sec:mantle}
-This problem is performed as presented by \citet{Blankenbach1989} (constant viscosity cases) and \citet{Thieulot2014}, in a 2D unit square domain with gravity
-acceleration $g_y=10^{10}Ra$. The experiment is performed with three different Rayleigh numbers ($Ra=10^4$, $10^5$ and $10^6$) and with different grid
-resolution (between \num{32x32} and \num{128x128} elements). The fluid has constant viscosity, initial density, heat capacity, thermal conductivity
-($\eta= \rho_0= C_p = k=1$), reference temperature ($T_0=0$) and thermal expansion coefficient ($\alpha=10^{-10}$). Temperatures are set to 0 on top and 1 on
-bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\]
-
-The solution generated by the code in terms of $v_{\textrm{rms}}$ (calculated as in Sec. \ref{sec:error}) and Nu as function of time are reported for all the
-simulations in Fig. \ref{fig:mantle}. At the steady state, $v_{\textrm{rms}}$ and Nu of all simulations converge well toward the values from \citet{Blankenbach1989},
-with lower errors for higher resolution grids (Table \ref{tab:mantle}). All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Mantle_convection}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Convection.pdf}
-\caption{$v_{\textrm{rms}}$ (panels a, c and e) and Nu (panels b, d and f) for the mantle convection benchmark as function of time for different grid resolution.
-Panels a and b show the results for $Ra=10^4$; Panels c and d show the results for $Ra=10^5$; panels e and f show the results for $Ra=10^6$. Purple lines
-indicate the convergence values for the $v_{\textrm{rms}}$ and the Nu from \citet{Blankenbach1989} at the steady state.}
-\label{fig:mantle}
-\end{figure}
-
-\begin{table}[H]
-\caption{Comparison between $v_{\textrm{rms}}$ and Nu predicted by the code for the mantle convection experiment and same values as reported in literature.}
-\centering
-\small
-\begin{tabular}{|>{\centering}m{0.09\linewidth}|>{\centering}m{0.05\linewidth}|>{\centering}m{0.24\linewidth}|>{\centering}m{0.15\linewidth}|
-  >{\centering}m{0.1\linewidth} >{\centering}m{0.1\linewidth} >{\centering\arraybackslash}m{0.1\linewidth}|}
-\toprule
-& & \multirow{2}{*}{\citet{Blankenbach1989}} & \citet{Thieulot2014} & \multicolumn{3}{c|}{FALCON}  \\
- & & & \num{200x200} & \num{32x32} & \num{64x64} & \num{128x128}  \\
-\midrule
-\multirow{2}{*}{$Ra=10^4$} & $v_{\textrm{rms}}$ & $42.864947 \pm 0.000020$ & 42.867 & 42.83226 & 42.852793 & 42.861394  \\
-           & Nu      & $4.884409 \pm 0.000010$  & 4.882  & 4.781297 & 4.857475  & 4.877573   \\
-\hline
-\multirow{2}{*}{$Ra=10^5$} & $v_{\textrm{rms}}$ & $193.21454 \pm 0.00010$  & 193.255& 193.872643 & 193.377472 & 193.252290  \\
-           & Nu      & $10.534095 \pm 0.000010$ & 10.507 & 9.602514 & 10.270735  & 10.465629   \\
-\hline
-\multirow{2}{*}{$Ra=10^6$} & $v_{\textrm{rms}}$ & $833.98977 \pm 0.00020$  & 834.712& 848.091176 & 837.767911 & 834.945793  \\
-           & Nu      & $21.972465 \pm 0.000020$ & 21.695 & 15.999266 & 19.703682  & 21.306939   \\
-\bottomrule
-\end{tabular}
-\label{tab:mantle}
-\end{table}
-
-\subsection{Viscoplastic mantle convection}\label{sec:visco_mantle}
-This benchmark is performed as Cases 1-5a described in \citet{Tosi2015} in a 2D unit square domain with two different grid resolutions of \num{32x32} 
-and \num{100x100} elements. Reference viscosity, reference density, heat capacity, thermal conductivity and thermal expansion coefficient are fixed to a constant
-value of 1 ($\eta_0= \rho_0= C_p = k=\alpha=1$), while the gravity has been chosen as $g_y=10^2$, in order to obtain a Rayleigh number $Ra=10^2$. Temperatures
-are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\].
-The viscosity field $\eta$ is calculated as the harmonic average between a linear part $\eta_{lin}$ that depends on temperature only or on temperature and 
-depth $y$ and a nonlinear-plastic part $\eta_{pl}$ that depends on the strain rate $\dot{\bm{\epsilon}}$, as follows
-\[\eta(T,y,\dot{\bm{\epsilon}})=2\left(\frac{1}{\eta_{lin}(T,y)}+\frac{1}{\eta_{pl}(\dot{\bm{\epsilon}})}\right)^{-1}\]
-The linear and the nonlinear-plastic parts of the viscosity are calculated following \citet{Tosi2015} as follows
-\[\eta_{lin}(T,y)=\exp(-\ln(\Delta\eta_T)T+\ln(\Delta\eta_y)y)\]
-\[\eta_{pl}(\dot{\bm{\epsilon}})=\eta^*+\frac{\sigma_Y}{\sqrt{\dot{\bm{\epsilon}}:\dot{\bm{\epsilon}}}}\]
-where $\Delta\eta_T$, $\Delta\eta_y$, $\eta^*$ and $\sigma_Y$ are parameters chosen as listed in Table \ref{tab:tosi_setup}.
-
-For each case, some diagnostic quantities measured at the steady state are reported in Table \ref{tab:visco_mantle} (see Section \ref{sec:error} for 
-the computation of all the quantities), showing a well fit with results described in \citet{Tosi2015}. Data and charts for all cases can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Viscoplastic_mantle_convection}.
-
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{white!100!}{gray!25!}
-%\begin{table}[H]
-%  \caption{Results of the viscoplastic mantle convection benchmark for Cases 1-4 compared with results obtained with ELEFANT and YACC (from \citealp{Tosi2015}).}
-%  \centering
-%  \resizebox{\linewidth}{!}{
-%\begin{tabular}{|c|ccc|ccc|ccc|ccc|}
-%  \toprule
-%    & \multicolumn{3}{c|}{\textbf{Case 1}} & \multicolumn{3}{c|}{\textbf{Case 2}} & \multicolumn{3}{c|}{\textbf{Case 3}} &
-%    \multicolumn{3}{c|}{\textbf{Case 4}} \\
-%   & FALCON & ELEFANT& YACC & FALCON& ELEFANT& YACC & FALCON & ELEFANT & YACC & FALCON & ELEFANT & YACC \\
-%  \midrule
-%   $\left\langle T \right\rangle$ & 0.7759 & 0.7758 & 0.7767 & 0.6043 & 0.6036 & 0.6058 & 0.7275 & 0.7275 & 0.7286 & & 0.5277 & 0.5289\\
-%    Nu$^{top}$ & 3.4280 & 3.4214 & 3.4298 & 8.5242 & 8.5115 & 8.5278 & 3.0368 & 3.0347 & 3.0374 && 6.5912 & 6.5572 \\
-%    Nu$^{bottom}$ & 3.3168 & 3.313 & 3.3143 & 8.4599 & 8.4465 & 8.3990 & 3.0086 & 2.9908 & 2.9628 && 6.5834 & 6.5243 \\
-%    $v_{\textrm{rms}}$ & 249.901 & 249.134 & 251.7997 & 141.2451 & 140.7087 & 142.2020 & 100.2776 & 100.1208 & 100.9467 && 79.1105 & 79.6202 \\
-%    $v_{\textrm{rms}}^{top}$ & 1.8907 & 1.8642 & 1.8298 & 104.7928 & 104.1228 & 105.1767 & 2.0832 & 2.0652 & 2.0374 && 74.7596 & 75.8414 \\
-%    $u_{max}^{top}$ & 2.6340 & 2.6119 & 2.5516 & 121.9212& 121.766 & 122.3238 & 2.9086 & 2.9019 & 2.8458 && 88.9146 & 89.2940 \\
-%    $\eta_{min}$ & - & - & - & $2.0303\times10^{-5}$ & $2.0600\times10^{-5}$ & $1.9853\times10^{-5}$ & $4.8129\times10^{-5}$ & $4.8080\times10^{-5}$ & $4.7907\times10^{-5}$ &
-%    & $1.9860\times10^{-4}$ & $1.9174\times10^{-4}$ \\
-%    $\eta_{max}$ & - & - & - & 1.3325 & 1.85 & 1.6095 & 0.9360 & 0.9023 & 1 && 1.52 & 1.6773 \\
-%    $\left\langle W \right\rangle$ & 2.4299 & 2.4316 & 2.4583 & 7.5648 & 7.5665 & 7.6222 & 2.0387 & 2.0384 & 2.0400 && 5.6216 & 5.6512 \\
-%    $\left\langle \Phi \right\rangle/Ra$ & 2.4311 & 2.4276 & 2.4333 & 7.5661 & 7.5585 & 7.6194 & 2.0386 & 2.0370 & 2.0335 && 5.6182 & 5.6463 \\
-%    $\delta$ & 0.05\% & 0.16\% & 1.02\% & 0.02\% & 0.11\% & 0.04\% & $<$0.01\% & 0.07\% & 0.32\% && 0.06\% & 0.09\% \\
-%  \bottomrule
-%  \end{tabular}}
-%  \label{tab:visco_mantle}
-%  \end{table}
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{2}{white!100!}{gray!25!}
-\begin{table}[H]
-  \caption{Benchmark cases and corresponding parameters.}
-  \centering
-  \small
-\begin{tabular}{|>{\centering}m{0.08\linewidth}|>{\centering}m{0.05\linewidth}>{\centering}m{0.05\linewidth}>{\centering}m{0.05\linewidth}
-  >{\centering}m{0.05\linewidth} >{\centering}m{0.05\linewidth} |>{\centering\arraybackslash}m{0.2\linewidth}|}
-  \toprule
-    Case & $Ra$ & $\Delta\eta_T$ & $\Delta\eta_y$ & $\eta^*$ & $\sigma_Y$ & Convective regime \\
-  \midrule
-  \rowcolor{gray!25!}
-    1 & $10^2$ & $10^5$ & 1 & - & - & Stagnant lid \\
-    2 & $10^2$ & $10^5$ & 1 & $10^{-3}$ & 1 & Mobile lid  \\
-    \rowcolor{gray!25!}
-    3 & $10^2$ & $10^5$ & 10 & - & - & Stagnant lid  \\
-    4 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 1 & Mobile lid \\
-    \rowcolor{gray!25!}
-    5 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 4 & Periodic \\
-  \bottomrule
-  \end{tabular}
-  \label{tab:tosi_setup}
-  \end{table}
-
-\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{gray!25!}{white!100!}
-\begin{table}
-  \caption{Results of the viscoplastic mantle convection benchmark for different grid resolutions in Cases 1-5 compared with results obtained with ELEFANT
-  and YACC ($100\times100$ elements, from \citealp{Tosi2015}) and in Stone 28 ($32\times32$ elements, from
-  \url{https://github.com/cedrict/fieldstone/tree/master/python_codes/fieldstone_28}).}
-  \centering
-  \resizebox{\linewidth}{!}{
-\begin{tabular}{|cc|ccccccccccc|}
-  \toprule
-  \multicolumn{13}{|c|}{\textbf{Case 1}} \\
-  \midrule
-  Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-  & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-  \hline
-  \rowcolor{gray!25!}
-    YACC & $100\times100$ & 0.7767 & 3.4298 & 3.3143 & 251.7997 & 1.8298 & 2.5516 & - & - & 2.4583 & 2.4333 & 1.02\% \\
-    ELEFANT & $100\times100$ & 0.7758 & 3.4214 & 3.313 & 249.134 & 1.8642 & 2.6119 & - & - & 2.4316 & 2.4276 & 0.16\% \\
-    \rowcolor{gray!25!}
-    Stone 28 & $32\times32$ & 0.7737 & 3.3987 & - & 243.872 & - & - & - & - & - & - & - \\
-%    FALCON & $32\times32$ & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & - & - & 2.4745 & 2.4864 & 0.48\% \\
-%    \rowcolor{gray!25!}
-    FALCON & $100\times100$ & 0.7759 & 3.4280 & 3.3168 & 249.901 & 1.8907 & 2.6340 & - & - & 2.4299 & 2.4311 & 0.05\% \\
-    \midrule
-    \multicolumn{13}{|c|}{\textbf{Case 2}} \\
-    \midrule
-    Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-    & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-    \hline
-    \rowcolor{gray!25!}
-     YACC & $100\times100$ & 0.6058 & 8.5278 & 8.3990 & 142.2020 & 105.1767 & 122.3238 & $1.9853\times10^{-5}$ & 1.6095 & 7.6222 & 7.6194 & 0.04\% \\
-     ELEFANT & $100\times100$ & 0.6034 & 8.5115 & 8.4465 & 140.8067 & 104.1228 & 121.766 & $2.0600\times10^{-5}$ & 1.85 & 7.5665 & 7.5585 & 0.11\% \\
-     \rowcolor{gray!25!}
-     Stone 28 & $32\times32$ & 0.6052 & 8.1715 & - & 141.518 & - & - & - & - & - & - & - \\
-%     FALCON & $32\times32$ & 0.6454 & 8.1422 & 7.4020 & 156.6908 & 95.7226 & 116.9302 & $2.2572\times10^{-5}$ & 0.7194 & 7.4289 & 7.4452 & 0.22\% \\
-%     \rowcolor{gray!25!}
-     FALCON & $100\times100$ & 0.6044 & 8.5266 & 8.4590 & 141.3684 & 104.8867 & 122.0470 & $2.0333\times10^{-5}$ & 1.2914 & 7.5707 & 7.5718 & 0.01\% \\
-     \midrule
-     \multicolumn{13}{|c|}{\textbf{Case 3}} \\
-     \midrule
-     Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-     & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-     \hline
-     \rowcolor{gray!25!} 
-     YACC & $100\times100$ & 0.7286 & 3.0374 & 2.9628 & 100.9467 & 2.0374 & 2.8458 & $4.7907\times10^{-5}$ & 1 & 2.0400 & 2.0335 & 0.32\% \\
-      ELEFANT & $100\times100$ & 0.7275 & 3.0347 & 2.9908 & 100.1208 & 2.0652 & 2.9019 & $4.8080\times10^{-5}$ & 0.9023 & 2.0384 & 2.037 & 0.07\% \\
-      \rowcolor{gray!25!}
-      Stone 28 & $32\times32$ & 0.7277 & 3.0347 & - & 100.018 & - & - & - & - & - & - & - \\
-%      FALCON & $32\times32$ & 0.7281 & 3.0618 & 2.8132 & 101.0092 & 2.1630 & 3.0178 & $4.8500\times10^{-5}$ & 0.7845 & 2.0699 & 2.0695 & 0.02\% \\
-%      \rowcolor{gray!25!}
-      FALCON & $100\times100$ & 0.7275 & 3.0368 & 3.0086 & 100.2776 & 2.0832 & 2.9086 & $4.8129\times10^{-5}$ & 0.9360 & 2.0387 & 2.0386 & $<$0.01\% \\
-      \midrule
-      \multicolumn{13}{|c|}{\textbf{Case 4}} \\
-      \midrule
-      Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-      & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-      \hline
-      \rowcolor{gray!25!} 
-      YACC & $100\times100$ & 0.5289 & 6.5572 & 6.5243 & 79.6202 & 75.4814 & 89.2940 & $1.9174\times10^{-4}$ & 1.6773 & 5.6512 & 5.6463 & 0.09\% \\
-       ELEFANT & $100\times100$ & 0.5277 & 6.5912 & 6.5834 & 79.1105 & 74.7596 & 88.9146 & $1.9860\times10^{-4}$ & 1.5200 & 5.6216 & 5.6182 & 0.06\% \\
-       \rowcolor{gray!25!}
-       Stone 28 & $32\times32$ & 0.5286 & 6.4036 & - & 79.4746 & - & - & - & - & - & - & - \\
-%       FALCON & $32\times32$ & 0.5334 & 6.4679 & 6.3435 & 80.6153 & 76.0511 & 90.2058 & $2.0393\times10^{-4}$ & 0.6453 & 5.6676 & 5.6698 & 0.04\% \\
-%       \rowcolor{gray!25!}
-       FALCON & $100\times100$ & 0.5279 & 6.5897 & 6.5830 & 79.1220 & 75.0715 & 88.8484 & $1.9693\times10^{-4}$ & 1.4133 & 5.6127 & 5.6113 & 0.03\% \\
-       \midrule
-       \multicolumn{13}{|c|}{\textbf{Case 5}} \\
-       \midrule
-       Code & Resolution & $P$ & $\left\langle T \right\rangle_{min}$ & $\left\langle T \right\rangle_{max}$ & Nu$^{top}_{min}$ & Nu$^{top}_{max}$ &
-       $v_{\textrm{rms}}^{min}$ & $v_{\textrm{rms}}^{max}$ & \multicolumn{2}{c}{$(\left\langle \Phi \right\rangle/Ra)_{min}$} & 
-       \multicolumn{2}{c|}{$(\left\langle \Phi \right\rangle/Ra)_{max}$} \\
-       \hline
-       \rowcolor{gray!25!}
-         YACC & $100\times100$ & 0.0792 & 0.6543 & 0.6722 & 2.6950 & 7.2792 & 41.9813 & 98.8230 & \multicolumn{2}{c}{1.4347} & 
-         \multicolumn{2}{c|}{9.2748} \\
-         ELEFANT & $100\times100$ & 0.076 & 0.6514 & 0.6693 & 2.6808 & 7.3506 & 41.2578 & 100.089 & \multicolumn{2}{c}{1.3324} & 
-         \multicolumn{2}{c|}{9.5289} \\
-         \rowcolor{gray!25!}
-         Stone 28 & $32\times32$ & 0.0764 & 0.6521 & 0.6697 & 2.6912 & 7.0763 & 41.61 & 98.183 & \multicolumn{2}{c}{-} & 
-         \multicolumn{2}{c|}{-} \\
-     %    FALCON & $32\times32$ & & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & \multicolumn{2}{c}{-} & 
-     %     \multicolumn{2}{c|}{-} \\
-     %    \rowcolor{gray!25!}
-         FALCON & $100\times100$ & 0.0775 & 0.6519 & 0.6697 & 2.6847 & 7.3629 & 41.5256 & 99.5466 & \multicolumn{2}{c}{1.3399} & 
-         \multicolumn{2}{c|}{9.3993} \\
-  \bottomrule
-  \end{tabular}}
-  \label{tab:visco_mantle}
-  \end{table}
-
-\subsection{Thin layer entrainment}\label{sec:thin}
-This experiment is performed as originally proposed by \citet{vanKeken1997} with two incompressible fluids in a rectangular domain with $L_x=2$ and $L_y=1$
-and gravity acceleration $g_y=10^{10}Ra$. The Rayleigh number and the compositional Rayleigh number are fixed to $Ra=300000$ and $Rc=450000$, respectively.
-Both fluids have constant viscosity, thermal conductivity, specific heat ($\eta=\rho=C_p=1$) and thermal expansion coefficient ($\alpha=10^{-10}$). Fluid 1
-has a density $\rho_1=1$, while fluid 2 is denser ($\rho_2=\rho_1+1.5 \cdot 10^{-10}$) and is placed at the bottom of the domain, for $y \leq 0.025$.
-Temperature are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides of the domain. The initial
-temperature field is given by
-\[T(x,y)=T_u(x,y)+T_l(x,y)+T_r(x,y)+T_s(x,y)-\frac{3}{2}\]
-with
-\begin{eqnarray}
-T_u(x,y)&=&\frac{1}{2}erf\left(\frac{1-y}{2}\sqrt{\frac{u_0}{x}}\right)\nonumber \\
-T_l(x,y)&=&1-\frac{1}{2}erf\left(\frac{y}{2}\sqrt{\frac{u_0}{L_x-x}}\right)\nonumber \\
-T_r(x,y)&=&\frac{1}{2}+\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{y+1}} \exp\left(-\frac{x^2u_0}{4y+4}\right)\nonumber \\
-T_s(x,y)&=&\frac{1}{2}-\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{2-y}} \exp\left(-\frac{(L_x-x)^2u_0}{8-4y}\right)\nonumber
-\end{eqnarray}
-and
-\begin{eqnarray}
-u_0&=&\frac{L_x^{7/3}}{(1+L_x^4)^{2/3}}\left(\frac{Ra}{2\sqrt{\pi}}\right)^{\frac{2}{3}}\nonumber \\
-Q&=&2\sqrt{\frac{L_x}{\pi u_0}}\nonumber
-\end{eqnarray}
-
-The experiment is performed with two grid resolution (\num{125x40} and \num{200x80} elements), with 100 markers per element and CFL of 0.25. $v_{\textrm{rms}}$
-(see Sec. \ref{sec:error}) calculated for both the simulations match well with results from \citet{vanKeken1997} and obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and ELEFANT \citep{Thieulot2014}, obtained with similar grid resolutions
-(Fig. \ref{fig:thin}). In addition, a variety of simulations with different resolutions and aspect ratios are performed to verify that the sensitivity of the
-initial velocity field. Results match well with results from \citet{vanKeken1997} (in the grey area) and \citet{Thieulot2014} (Fig. \ref{fig:thin_initial}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Thin_layer}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/vrms_thin.pdf}
-\caption{$v_{\textrm{rms}}$ for the thin layer experiment as function of time for different grid resolution. Results are compared with results obtained by
-\citet{vanKeken1997} (black lines) and with ASPECT and ELEFANT.}
-\label{fig:thin}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/vrmszero.pdf}
-\caption{$v_{\textrm{rms}}$ at $t=0$ for the thin layer experiment as function of the element size for different aspect ratios of the elements. Results are compared
-with results obtained by \citet{vanKeken1997} (black dots in the grey area) and with ASPECT and ELEFANT.}
-\label{fig:thin_initial}
-\end{figure}
-
-\subsection{Sinking hydrated cylinder}\label{sec:quinquis}
-This experiment is performed as explained in \citet{Quinquis2014} and simulates the sinking of a cold, hydrated cylinder into a warm, dry mantle. The domain
-is square with $L_x=L_y=$ \SI{300}{\km}, a grid resolution of \num{300x300} elements and 25 markers per element. The sinking cylinder is located at $x=$
-\SI{150}{\km} and $y=$ \SI{170}{\km} with a radius of \SI{20}{\km} and has $\rho_c=$ \SI{3250}{\kg\per\cubic\m}, $\eta_c=$ \SI{e23}{\pascal\s}, a thermal
-conductivity $k_c=$ \SI{4.5}{\watt\per\m\per\kelvin},a specific heat $C_{p_c}=$ \SI{1250}{\joule\per\kg\per\kelvin}; the surrounding mantle has $\rho_m=$
-\SI{3200}{\kg\per\cubic\m}, $\eta_m=$ \SI{e20}{\pascal\s}, a thermal conductivity $k_m=$ \SI{105}{\watt\per\m\per\kelvin} and specific heat $C_{p_m}=$
-\SI{1250}{\joule\per\kg\per\kelvin}; a 58 km-thick lithosphere overlies the mantle and has $\rho_l=$ \SI{3200}{\kg\per\cubic\m}, $\eta_l=$ \SI{e23}{\pascal\s},
-a thermal conductivity $k_l=$ \SI{4.5}{\watt\per\m\per\kelvin}, specific heat $C_{p_l}=$ \SI{750}{\joule\per\kg\per\kelvin} and a thermal diffusivity of
-\SI{e-6}{\square\m\per\s}. Velocity boundary conditions are set to free slip on all sides of the domain. Temperature increase linearly in the lithosphere,
-from 0° to 1300°C, and in the mantle, from 1300° to 1360.5°C. The initial temperature of the cylinder is fixed to 400°C. Throughout the evolution, temperature
-is fixed to 0° and 1360.5°C at the top and bottom of the domain, respectively. The initial bound water content is imposed at 2 wt.\% in the cylinder and at
-0 wt.\% in both the mantle and the lithosphere. The maximum amount of water in the mantle is fixed to 0.2 wt.\%, while maximum water content in the cylinder
-and in the lithosphere is function of pressure and temperature and is calculated using a serpentinized harzburgite with Perple\_X, as in \citet{Quinquis2014}.
-
-As shown in \citet{Quinquis2014}, the model is characterised by a progressive dehydration in the external portion of the cylinder, due to the increase of
-temperature, with a consequent vertical migration of free water that hydrates the mantle above the cylinder, up to the lithosphere. The distribution of
-bound and free water during the evolution of the experiment is shown in Fig. \ref{fig:quinquis} (left and right columns, respectively). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Hydration}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Hydration.pdf}
-\caption{Distribution of bound and free water (left and right columns, respectively) at $t=0.5$ and $t=$ \SI{1.25}{\mega\year} (first and second rows,
-respectively) for the sinking hydrated cylinder experiment. Only markers with bound or free water not equal to 0 are plotted.}
-\label{fig:quinquis}
-\end{figure}
-
-\subsection{Mantle melting curves}\label{sec:katz}
-This instantaneous experiment is performed in order to recreate isobaric melting curves, as function of temperature, bulk water and content of water in the
-melt, to compare with melting curves obtained by \citet{Katz2003}. The domain is square with $L_x=L_y=100$ km, a grid resolution of \num{100x100} elements and
-25 markers per element. Density is set in the entire domain to \SI{3300}{\kg\per\cubic\m} and temperature increases from the sides to the centre from 1000° to
-1500°C. The experiment is performed for bulk water contents from 0 to 0.3 wt.\%.
-
-Fig. \ref{fig:fraction} shows isobaric melting curves for \SI{1}{\giga\pascal} (panel a) and \SI{3}{\giga\pascal} (panel b) calculated for different bulk water
-contents. The curves well match with those obtained for the mantle by \citet{Katz2003}, showing the saturation at the solidus for bulk water of 0.2 and 0.3 wt.\%, indicating
-that the code correctly determines melt fraction as function of pressure, temperature and bulk water content. Same results can be observed in Fig.
-\ref{fig:water}a, where isobaric and isothermal melting curves are shown as function of bulk water content. In addition, Fig. \ref{fig:water}b shows that the
-code correctly determines also the percentage of water in the melt as function of pressure, temperature and melt fraction. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Melting}.
-
-\begin{figure}
-\includegraphics[width=\linewidth]{./Figures/Fraction.pdf}
-\caption{Isobaric melting curves for 1 GPa (panel a) and 3 GPa (panel b) as function of temperature with different bulk water contents.}
-\label{fig:fraction}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\linewidth]{./Figures/Water.pdf}
-\caption{Isobaric melting curves for 1.5 GPa as function of bulk water (panel a) and water in the melt (panel b) calculated for different temperatures.}
-\label{fig:water}
-\end{figure}
-
-\clearpage
-
-\bibliography{export}
-
-\end{document}

### 2022-09-07 tex/erosion.tex
update main

--- 2021-05-26 Paper.tex
@@
-\documentclass[hidelinks,10pt,a4paper]{article}
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage{booktabs}
-\usepackage{multirow}
-\usepackage[cm]{fullpage}
-\usepackage[round]{natbib}
-\bibliographystyle{abbrvnat}
-\usepackage{doi}
-\usepackage{bm}
-\usepackage[table]{xcolor}
-\hypersetup{
-    colorlinks,
-    linkcolor={blue},
-    citecolor={blue},
-    urlcolor={blue}
-}
-\usepackage{float}
-\restylefloat{table}
-\usepackage{siunitx}
-\DeclareSIUnit\year{yr}
+\section{Erosion and sedimentation}\label{sec:erosion}
+Surface processes at either the lithosphere-air interface or in correspondence of the free surface have been implemented by means of the software
+Fastscape\footnote{\url{https://fastscape.org/}} \citep{Braun2013,Cordonnier2019,Yuan2019a,Yuan2019b}. FastScapeLib is a set of routines that solve the stream
+power law (enriched by a sediment transport/deposition term) \citep{Yuan2019a}, hillslope diffusion and marine transport and deposition \citep{Yuan2019b}.
+The partial differential equation solved by FastScapeLib is:
+\begin{equation}
+\frac{\partial h}{ \partial t}=U-K_f A^{m_e}S^{n_e} + \frac{G}{A}\int_A \left( U-\frac{\partial h}{\partial t} \right) dA + K_d \bm{\nabla}^2 h 
+\end{equation}
+where $h$ is the topography, $U$ is the uplift, $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $S$ is the slope, $m_e$ and $n_e$ are
+parameters that should be chosen to constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}, $G$ is the dimensionless
+deposition/transport coefficient for the enriched stream power law and $K_d$ is the effective diffusivity. Terms $K_f A^{m_e}S^{n_e}$ and $K_d \bm{\nabla}^2 h$
+consider long-range processes, described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, and short-range processes,
+proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}.
 
-\begin{document}
-
-\section{Introduction}\label{sec:intro}
-This supplement includes a description of the code, with a presentation of the features implemented followed by results of extensive benchmarks listed in
-Table \ref{tab:list}, implemented to verify that the code solves correctly a large range of problems. Description and results of all benchmarks are shown
-in Section \ref{sec:benchmark}.
-
-\begin{table}[H]
-\caption{List of benchmarks implemented for each feature of the code. Data of all benchmarks can be found at 
-\url{https://github.com/aleregorda/Benchmarks}}
-\centering
-\begin{tabular}{| l | l | l |}
-\toprule
- \textbf{Feature}  & \textbf{Benchmark} & \textbf{Section} \\
-\midrule
-  Solver  & Stokes flow & \ref{sec:stokes} \\
-\hline
-  4th-order Runge-Kutta  & \multirow{2}{*}{Zalesak disk} & \multirow{2}{*}{\ref{sec:runge}} \\
-  markers advection  &  & \\
-\hline
-  \multirow{4}{*}{Momentum equation}  & Poiseuille flow & \ref{sec:poiseuille}\\
-    & Instantaneous 2D sphere & \ref{sec:ist_sphere} \\
-    & Rayleigh-Taylor experiment & \ref{sec:rayleigh} \\
-    & Falling block & \ref{sec:block} \\
-\hline
-    & 2D Stokes sphere & \ref{sec:sphere} \\
-    Sticky air and  & Stabilisation algorithm & \ref{sec:stab} \\
-    free surface & Topography relaxation & \ref{sec:crameri} \\
-  & Spontaneous subduction & \ref{sec:subduction} \\
-\hline
-  Erosion and sedimentation  & - & - \\
-\hline
-   Non-linear  & Slab detachment & \ref{sec:slab} \\
-   visco-plasticity & Indenter & \ref{sec:indenter} \\
-    & Brick & \ref{sec:brick} \\
-\hline
-  \multirow{3}{*}{Energy equation}  & Advection stabilisation & \ref{sec:advection} \\
-    & Simple shear heating & \ref{sec:simple_shear} \\
-    & Shear and adiabatic heating & \ref{sec:shear} \\
-\hline
-\multirow{3}{*}{Energy + momentum} & Mantle convection & \ref{sec:mantle} \\
-& Viscoplastic mantle convection & \ref{sec:visco_mantle} \\
-& Thin layer entrainment & \ref{sec:thin} \\
-\hline
-  Phase changes  & \multirow{2}{*}{Hydrated sinking cylinder} & \multirow{2}{*}{\ref{sec:quinquis}} \\
-  and hydration  &  & \\
-\hline
-  Melting & Experimental melting curves & \ref{sec:katz} \\
-\bottomrule
-\end{tabular}
-\label{tab:list}
-\end{table}
-
-The code is written in Fortran90 and uses the finite element method (FEM) with quadrilateral $Q_1 \times P_0$ elements (continuous bilinear velocity and
-discontinuous constant pressure), associated with MUMPS\footnote{\url{http://mumps.enseeiht.fr/}} \citep{Amestoy2001,Amestoy2019} that is a software package
-for solving systems of linear equations of the form $A \cdot x = b$, where \textit{A} is a square sparse matrix, by means of a direct method. Since
-$Q_1 \times P_0$ elements do not satisfy Ladyzhenskaya, Babuska and Brezzi (LBB) stability condition \citep{Donea2003}, elemental pressure is elaborated
-in post-processing to avoid spurious pressures by means of a double interpolation to smooth the pressure. Following this procedure, the elemental pressure
-is interpolated onto nodes and then back onto elements. After the smoothing procedure, the pressure field is calculated again on the nodes to be used in
-combination with the lithostatic pressure, which is calculated onto nodes. Correctness of the solution and performances of the solver in terms of time and
-memory usage are tested solving the Stokes flow with the analytical solution proposed by \citet{Donea2003} (Section \ref{sec:stokes}).
-
-The thermo-mechanics of crust-mantle systems is described by means of the conservation of mass, momentum and energy equations, expressed as follows:
-\begin{eqnarray}
-\label{eq:mass}\frac{\partial \rho}{\partial t}+\bm{\nabla} \cdot (\rho \bm{u})&=&0\\
-\label{eq:momentum}-\bm{\nabla} p + \bm{\nabla} \cdot \bm{\tau} + \rho \bm{g}&=&0\\
-\label{eq:energy}\rho C_p \left(\frac{\partial T}{\partial t} + \bm{u} \cdot \bm{\nabla} T \right)&=&\bm{\nabla} \cdot \left(k\bm{\nabla} T\right) + H_{tot}
-\end{eqnarray}
-where $\rho$ is the density, $\bm{u}$ is the velocity, \textit{p} is the pressure, $\bm{\tau}$ is the deviatoric stress, $\bm{g}$ is the gravity acceleration,
-$C_p$ is the specific heat at constant pressure, \textit{T} is the temperature, \textit{k} is the thermal conductivity and $H_{tot}$ is the total internal heat
-production. Density variations due to temperature are generally small enough to assume the density as constant ($\rho=\rho_0$) in Eq. \ref{eq:mass} and in
-Eq. \ref{eq:energy}, while it must be treated as a variable in the buoyancy term of Eq. \ref{eq:momentum}, such that \[\rho=\rho_0(1-\alpha(T-T_0))\]
-where $\rho_0$ is the density at a reference temperature $T_0$ and $\alpha$ is the coefficient of thermal expansion. Eq. \ref{eq:mass} then can be rewritten as
-\begin{equation}\label{eq:mass_0}
-\bm{\nabla} \cdot \bm{u}=0
-\end{equation}
-This simplification is known as the extended Boussinesq approximation.
-
-The time step $dt$ is chosen in according to Courant-Friedrichs-Lewy condition \citep{Anderson}, such that
-\[dt=C_n\; \min\left(\frac{\Delta x}{\max|\bm{u}|},\frac{(\Delta x)^2}{\kappa}\right)\]
-where $0<C_n<1$ is the Courant (CFL) number, $\Delta x$ is the minimum dimension of smallest element, $\max|\bm{u}|$ is the maximum velocity calculated on the
-entire domain and
-$\kappa=\frac{k}{\rho C_p}$ is the thermal diffusivity \citep{Thieulot2014}.
-
-\section{Lagrangian markers}\label{sec:markers}
-Elemental properties (density, viscosity, thermal conductivity, specific heat and thermal expansion) needed to solve Eqs. \ref{eq:mass_0}, \ref{eq:momentum}
-and \ref{eq:energy} are related to the composition of each element, determined by means of Lagrangian markers that are characteristic of different materials
-in the domain. Elemental properties (with the sole exception of the viscosity) are calculated using an arithmetic mean, as \[P_e=\frac{1}{n}\sum_{i=1}^n P_i\]
-where $P_e$ is the elemental property, $P_i$ is the property characteristic of the material of each marker and $n$ is the number of markers of the element.
-Differently, the average scheme for the elemental viscosity can be chosen between harmonic, geometric and arithmetic mean. Markers can be placed regularly or
-randomly at the beginning of the simulation and their advection is performed by a 4th-order Runge-Kutta in space, interpolating the velocity field on each
-marker by means of the shape functions (see Section \ref{sec:runge}). The initial distribution of the markers is created using the open-source code library
-Geodynamic World Builder \citep{Fraters2019}. During the simulation, each marker carries memory of temperature, pressure and accumulated strain, which are
-determined interpolating the relative nodal parameter. In particular, pressure is interpolated onto each marker after the smoothing procedure. The number
-of markers contained in each element is maintained between $n_{min}$ and $n_{max}$. When in an element there are less markers than $n_{min}$ the code adds
-random markers to reach the $n_{min}$, while if the number is higher than $n_{max}$ some of them are randomly deleted. When new markers are added, they assume
-the properties of the nearest marker. In this way elements are never empty and maintain a number of markers inside a prefixed range.
-
-\section{The momentum equation}\label{sec:mom_eq}
-The penalty method is implemented so that Eq. \ref{eq:mass_0} can be written as:
-\begin{equation}\label{eq:mass_penalty}
-\bm{\nabla} \cdot \bm{u} + \frac{p}{\lambda}=0
-\end{equation}
-where $\lambda$ is the so-called penalty parameter, which should be 6-7 orders of magnitude larger than the shear viscosity to ensure that mass conservation
-is satisfied \citep{Donea2003,Thieulot2014}.
-
-The deviatoric stress tensor in Eq. \ref{eq:momentum} can be written in terms of the strain rate tensor as $\bm{\tau}=2\eta\dot{\bm{\epsilon}}$, with
-$\dot{\bm{\epsilon}}=\frac{1}{2}\left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)$. Therefore, Eq. \ref{eq:momentum} can be rewritten as:
-\begin{equation}\label{eq:momentum_pressure}
--\bm{\nabla} p + \bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right) + \rho \bm{g} = 0
-\end{equation}
-Finally, using pressure from Eq. \ref{eq:mass_penalty}, Eq. \ref{eq:momentum_pressure} can be rewritten as:
-\begin{equation}\label{eq:momentum_penalty}
-\lambda \bm{\nabla} \left(\bm{\nabla} \cdot \bm{v}\right)+\bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right)+\rho\bm{g}=0
-\end{equation}
-The penalty method is been associated to the iterative Uzawa method, as described in detail in \citet{Dabrowski2008} and \citet{Thieulot2014}
-(see Section \ref{sec:poiseuille}). Results of benchmarks performed to verify the correctness in the implementation of Eq. \ref{eq:momentum_penalty} are
-shown in Sections \ref{sec:ist_sphere} and \ref{sec:rayleigh}.
-
-To support large viscosities variations the penalty parameter is related to the effective elemental viscosity by means of a dimensionless coefficient, so that
-$\lambda=\lambda_e(e)\eta_{eff}(e)$ \citep{Marotta2006,Dabrowski2008,Thieulot2014}. Benchmark of the falling block \citep{Gerya2003a,Gerya2010b,Thieulot2011}
-is performed to verify that the code can correctly deal buoyancy driven flows with strong viscosity contrasts (Section \ref{sec:block}).
-
-\section{Sticky air and free surface}\label{sec:sticky}
-The Earth's surface can be treated by means of either the so-called sticky-air or a true free surface method, both of them implemented in the code. In the
-sticky-air method the surface is approximated with the introduction of a buoyant layer with a viscosity at least four orders of magnitude lower than the crust
-\citep{Schmeling2008,Crameri2012} and the interface between lithosphere and air is defined using a chain of passive markers that are advected as the Lagrangian
-markers. The correctness of the evolution of the markers chain is tested with the experiment of a 2D time-dependent Stokes sphere below a free surface and
-compared with results from ASPECT\footnote{\url{https://aspect.geodynamics.org/}} \citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} (Section
-\ref{sec:sphere}). In the true free surface case the top boundary is assumed stress-free and velocities are not fixed. In this case topography variations are
-described by vertical deformations of the mesh that depend on the velocity field of the nodes that identify the free surface, while horizontal deformations are
-not taken into account. This procedure is known as the Arbitrary Lagrangian-Eulerian (ALE) method and its implementation follows the technique described in
-\citet{Thieulot2011}. However, although the implementation of a true free surface better reproduces laboratory experiment, extremely small time steps can be
-necessary to maintain stability \citep{Kaus2010a,Quinquis2011,Thieulot2014}. Therefore, the stability algorithm proposed by \citet{Kaus2010a} is implemented
-to avoid instabilities due to high density differences at the free surface when using too large time steps. The implementation of this algorithm are tested by
-performing the experiment described by \citet{Kaus2010a} (Section \ref{sec:stab}).
-
-The topography relaxation benchmark proposed by \citet{Crameri2012} is performed to verify that the code correctly recovers topography variations in case of
-both the sticky-air and the true free surface method (Section \ref{sec:crameri}). Finally, Section \ref{sec:subduction} show the results of the spontaneous
-subduction experiment described by \citet{Schmeling2008}.
-
-\section{Erosion and sedimentation}\label{sec:erosion}
-Surface processes at the lithosphere-air interface are implemented taking into account fluvial erosion (long-range processes), hillslope processes (short-range
-processes) and failure of slopes exceeding a threshold angle $\theta$ \citep{Burov1997,Tucker1999,Sacek2012,Croissant2014,Ueda2015,Beucher2020}. Erosion rates
-$\dot{E}_{lr}$ due to long-range processes can be described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, in according to
-\begin{equation}
-\dot{E}_{lr}=K_f A^{m_e} \bm{\nabla} h^{n_e}
-\end{equation}
-where $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $\bm{\nabla} h$ is the slope and $m_e$ and $n_e$ are parameters chosen in to
-constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}. Erosion rates $\dot{E}_{sr}$ due to short-range processes are
-implemented by assuming that transport is linearly proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}, in according to
-\begin{equation}
-\dot{E}_{sr}=K_d \bm{\nabla}^2 h
-\end{equation}
-where $K_d$ is the effective diffusivity. In addition, the slope is modified if it still exceeds a threshold angle after long- and short-range processes
-\citep{Ueda2015}.
-
-Offshore sedimentation processes are implemented similarly to hillslope processes but with a depth-dependent coefficient \citep{Sacek2012,Beucher2020}.
-The sedimentation rate $\dot{S}$ is calculated as
-\begin{equation}
-\dot{S}=K_s \bm{\nabla}^2 h
-\end{equation}
-with
-\begin{equation}
-K_s=K^*_s \exp\left(\frac{h-h_{sl}}{h_0}\right)
-\end{equation}
-where $K^*_s$ is a constant, $h_{sl}$ is the sea level and $h_0$ is a characteristic length. In case of sticky air, slope is determined by means of the markers
-chain, which is vertically corrected considering erosion and sedimentation rates. After the correction, continental markers above the markers chain are
-transformed in air markers and air markers below the markers chain become sediments. Similarly, slope is determined considering the nodes on the top boundary
-in case of true free surface and they are vertically corrected as for the markers chain. In this case, continental markers above the top boundary are simply
-deleted.
-
-\section{Non-linear rheologies}\label{sec:nonlinear}
-Non-linear rheologies are implemented combining viscous creep (dislocation and diffusion) and plastic yielding. For each marker, diffusion and dislocation
-viscosity ($\eta_{df|ds}$) can be determined as follows \citep{Karato1993,Wang2016,Glerum2018}:
-\begin{equation}\label{eq:viscous}
-\eta_{df|ds}=\left(\frac{d^m}{A}\right)^{\frac{1}{n}}I_2^{\frac{(1-n)}{n}}\exp\left(\frac{Q+pV}{nRT}\right)
-\end{equation}
-where $d$ is the grain size, $m$ is the grain size exponent, $n$ is the stress exponent, $A$ is the pre-exponential factor, $I_2$ is the square root of the
-second invariant of the strain rate tensor, $Q$ is the activation energy, $p$ is the pressure, $V$ is the activation volume, $R$ is the gas constant and $T$
-is the temperature. Pressures and temperatures are determined by the interpolation of the nodal parameters, while strain rates are calculated by means of the
-derivative of the velocities on the nodes. In case of diffusion creep $n=1$ and $m>0$, in case of dislocation creep $n>1$ and $m=0$. The implementation of the
-non-linear viscous creep viscosity depending on the strain rate is tested by means of the slab detachment benchmark (Section \ref{sec:slab}).
-
-Plastic yielding is implemented rescaling $\eta_{vc}$ in order to limit the stress below the yield stress $\sigma_y$ \citep{Thieulot2008,Thieulot2014,Glerum2018},
-obtaining
-\begin{equation}\label{eq:plastic}
-\eta_{pl}=\frac{\sigma_y}{2I_2}
-\end{equation}
-where $\eta_{pl}$ is the plastic viscosity and the yield stress is determined following the Drucker-Prager criterion, such as
-\begin{equation}\label{eq:yield}
-\sigma_y=C\;\cos(\phi)+p\;\sin(\phi)
-\end{equation}
-where $C$ is the cohesion and $\phi$ is the internal friction angle. In case of negative pressure it is imposed equal to 0, so that negative yield stress are
-excluded. The correctness of the non-linear solution in case of plastic viscosity with variable internal friction angle is verified performing indenter and
-brick experiments (Sections \ref{sec:indenter} and \ref{sec:brick}, respectively).
-
-Strain softening is taken into account for both viscous creep and plastic viscosity \citep{Huismans2003,Babeyko2005,Huismans2005,Sobolev2005,Warren2008a} by
-means of the accumulated strain $\epsilon$ memorised by each marker. Softening in the viscous creep determines a linear decrease of $\eta_{vc}$ by means of a
-viscous strain softening factor $W_S$ that increases linearly from 1 to 10 for $5<\epsilon<10$ \citep{Warren2008a}. This viscous softening can be related to
-strain-induced grain size reduction \citep{Warren2008a}. Differently, plastic softening is simulated with a linear decrease of internal friction angle
-$\phi(\epsilon)$ and cohesion $C(\epsilon)$ in according to
-\begin{eqnarray}
-\label{eq:friction1}\phi(\epsilon)&=&\phi_0+(\phi_{\infty}-\phi_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}\\
-\label{eq:friction2}C(\epsilon)&=&C_0+(C_{\infty}-C_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}
-\end{eqnarray}
-where $\phi_0$, $C_0$ and $\phi_{\infty}$, $C_{\infty}$ are internal friction angle and cohesion for $\epsilon=0.5$ and $\epsilon=1.5$, respectively
-\citep{Huismans2003,Huismans2005,Warren2008a,Thieulot2014}. Plastic softening approximates deformation-induced softening of faults and brittle shear zones
-\citep{Warren2008a}.
-
-Viscous creep and plastic yielding are then combined to obtain a viscoplastic viscosity $\eta_{vp}$, which is computed with an harmonic average
-\citep{Tackley2010,Glerum2018} as follows
-\begin{equation}\label{eq:viscoplastic}
-\eta_{vp}=\left(\frac{1}{\eta_{df}}+\frac{1}{\eta_{ds}}+\frac{1}{\eta_{pl}}\right)^{-1}
-\end{equation}
-Finally, effective viscosity $\eta_{eff}$ is capped by the minimum and the maximum viscosity ($\eta_{min}$ and $\eta_{max}$, respectively) to avoid extremely
-low or high viscosity \citep{Glerum2018} as follows
-\begin{equation}\label{eq:effective}
-\eta_{eff}=\min(\max(\eta_{vp},\eta_{min}),\eta_{max})
-\end{equation}
-Viscous creep and plastic yielding are non-linear rheologies because of their dependence on the velocity field through pressure and strain rates. Therefore,
-the solution is determined by means of Picard-type iterations, until convergence of the velocity field \citep{Glerum2018}. The convergence is verified at each
-iteration $i$ via the nonlinear residual $\bm{\mathcal{R}}^i$ that can be determined as
-\begin{equation}\label{eq:convergence}
-\bm{\mathcal{R}}^i= \mathbb{K} (\eta_{eff}(\dot{\bm{\epsilon}}^{i-1},p^{i-1}))\cdot \bm{v}^{i-1} - \bm{f}^i
-\end{equation}
-%\begin{equation}\label{eq:convergence}
-%\chi=1-\frac{\langle (\bm{v}^i-\langle\bm{v}^i\rangle)\cdot (\bm{v}^{i+1}-\langle\bm{v}^{i+1}\rangle) \rangle}{\sqrt{\sigma^i\sigma^{i+1}}}
-%\end{equation}
-%Iterations are performed until either convergence ($\chi<tol$) or maximal number of non-linear iterations ($it_{max}$) is reached.
-where $\mathbb{K}$ is the Stiffness matrix and $\bm{f}$ is the right hand side vector \citep{Spiegelman2016,Glerum2018}.  The $L_2$-norm is extracted from
-$\bm{\mathcal{R}}^i$ and it is normalised as follows
-\begin{equation}\label{eq:convergence_res}
-\frac{||\bm{\mathcal{R}}^i||_2}{||\bm{\mathcal{R}}^0||_2}
-\end{equation}
-where $||\bm{\mathcal{R}}^0||_2$ is the $L_2$-norm of the first nonlinear iteration.
-Since often the normalised nonlinear residual drop very quickly over the first 2-3 iterations, also the $L_2$-norm of the velocity residual is calculated as
-\begin{equation}\label{eq:convergence_vel}
-\frac{||\bm{v}_1-\bm{v}_{1-1}||_2}{||\bm{v}_{i-1}||_2}
-\end{equation}
-Iterations are performed until either maximal number of non-linear iterations ($it_{max}$) is reached or both the normalised nonlinear residual and the
-velocity residual converge under a defined tolerance $tol$.
-
-\section{The energy equation}\label{sec:en_eq}
-The stabilisation of the advection term of the energy equation (Eq. \ref{eq:energy}) needed to avoid possible oscillations in the thermal solution in the case
-when advection dominates over diffusion, is implemented by means of a streamline-upwind Petrov–Galerkin (SUPG) method, for which the advection term is modified
-follows the discussion in \citet{Thieulot2011} and \citet{Thieulot2014} (see Section \ref{sec:advection} for benchmark). The total internal heat production
-$H_{tot}$ in Eq. \ref{eq:energy} includes radiogenic heating $H_r$, shear heating $H_s$ and adiabatic heating $H_a$, where
-\begin{equation}\label{eq:shear}
-H_s=2\eta \dot{\bm{\epsilon}} : \dot{\bm{\epsilon}} = 2\eta(\dot{\epsilon}_{xx}^2+\dot{\epsilon}_{yy}^2+2\dot{\epsilon}_{xy}^2)
-\end{equation}
-and
-\begin{equation}\label{eq:adiabatic}
-H_a=T \alpha \frac{Dp}{Dt} \approx -\alpha T \rho g_y v
-\end{equation}
-The correctness of the implementation and the computation of shear and adiabatic heating is verified performing an exercise with an analytical solution
-(Section \ref{sec:simple_shear}) and Exercise 9.4 in \citet{Gerya2010b} (Section \ref{sec:shear}). Furthermore, mantle convection and thin layer entrainment
-benchmarks (Sections \ref{sec:mantle} and \ref{sec:thin}, respectively) are performed to verify that the code solve correctly Eqs. \ref{eq:momentum_penalty}
-and \ref{eq:energy} in case of a temperature field characterised by an initial perturbation and with a temperature-dependent density. At last, the viscoplastic
-mantle convection benchmark is performed to check also the viscoplasticity in case of a initial perturbed temperature field (Section \ref{sec:visco_mantle}).
-
-\section{Phase transitions and hydration}\label{sec:phase}
-Variations of effective density, specific heat and coefficient of thermal expansion during the evolution of crustal and mantle materials at different
-pressure-temperature ($p$-$T$) conditions are computed by means of version 6.8.6 of Perple\_X software package \citep{Connolly2005}, similarly to the
-implementation described in \citet{Marotta2020}. By default, mineral assemblages and properties of each lithology (oceanic crust, lower and upper continental
-crust, sediments and dry or hydrated mantle) are calculated for temperatures from \SIrange{330}{1600}{\kelvin}, with increments of approximately \SI{4}{\kelvin},
-and pressures from \SIrange{0.1}{30}{\giga\pascal}, with increments of approximately \SI{0.03}{\giga\pascal}, for a total of almost 100000 points.
-
-The introduction of phase transitions produces two main effects that must be taken into account in numerical models: variations in density and release/absorption
-of latent heat \citep{Tackley2010}. Effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into
-account considering effective coefficient of thermal expansion \citep{Christensen1985,Zhong2015}. Similarly, effects of latent heat in the energy equation
-(eq \ref{eq:energy}) are taken into account by considering effective specific heat and coefficient of thermal expansion
-\citep{Christensen1985,Tackley2010,Zhong2015}. Effective density, specific heat and coefficient of thermal expansion are included in Perple\_X files and are
-assumed by each Lagrangian marker in according to its $p$-$T$ conditions.
-
-Hydration processes related to dehydration of subducting lithosphere \citep{Schmidt1998,Liu2007,Faccenda2009,Faccenda2010,Faccenda2014,Rosas2016} strongly
-influence the thermo-mechanics inside the mantle wedge, mainly because of both weakening effects on the mantle rheology and density variations in case of
-mantle serpentinization \citep{Gerya2002,Honda2003,Arcay2005,Roda2010,Regorda2017}. In the case that hydration is switched on, the amount of bound and free
-water is memorised by each marker, following the implementation of \citet{Quinquis2014}; hydration and dehydration processes are related to the amount of bound
-water of each marker and to the maximum amount of water it can transport, i.e. if the amount of bound water exceeds the maximum amount of water, the marker
-dehydrates and releases free water that can hydrates under-saturated markers. Maximum water content of each marker is determined as function of lithology and
-$p$-$T$ conditions and it is calculated using Perple\_X, in the same way than effective density, specific heat and coefficient of thermal expansion. Bound water
-is advected together with the markers, neglecting the effect of bound water diffusion, while free water simply migrates vertically and is not coupled to the
-solid-phase flow of the mantle wedge \citep{Arcay2005,Quinquis2014}. The correctness of the migration scheme of the free water and its absorption into bound
-water by under-saturated markers are verified performing the experiment described in \citet{Quinquis2014} (Sec. \ref{sec:quinquis}).
-
-The mantle viscosity weakening related to the amount of water has been extensively studied \citep{Chopra1981,Hirth2003} and it is implemented in according to
-\citet{Arcay2005} and \citet{Horiuchi2016}, as follows
-\begin{equation}\label{eq:wet_visc}
-\eta_{wet}=\eta_{dry}\left(\left[1-\frac{1}{f_v}\right]\exp\left(-\frac{[OH^-]}{[OH^-]_0}\right)+\frac{1}{f_v}\right)
-\end{equation}
-where $f_v$ is the viscosity reduction factor between dry and wet conditions, $[OH^-]$ is the water content and $[OH^-]_0$ is a reference water content set to
-620 ppm (0.062 wt.\%) \citep{Arcay2005}. By default, $f_v$ is set to 100, in according to \citet{Arcay2005} and \citet{Horiuchi2016}.
-
-\section{Mantle melting}\label{sec:melt}
-Melt fraction in the mantle depends on temperature ($T$, in °C), pressure ($p$, in \si{\giga\pascal}) and the water content in the melt ($X_{H_2O}$, in wt.\%).
-The determination of the melt fraction $M$ is implemented as explained by \citet{Katz2003}, according to the successive modification by \citet{Langmuir2006} and
-\citet{Kelley2010}, as follows
-\begin{eqnarray}\label{eq:melt_fr}
-M(p,T,X_{H_2O})&=&-T+T_{sol}+(x_t\cdot ln(p+y_t))\cdot X_{H_2O}-\nonumber\\
-&&+K_m\left(\frac{X_{H_2O}^{bulk}}{D_{H_2O}(1-X_{H_2O})+X_{H_2O}}\right)^{\gamma}
-\end{eqnarray}
-where $T_{sol}$ is the temperature of the dry solidus, $D_{H_2O}$ is the partition coefficient and $x_t\cdot log(p+y_t)$ indicates the pressure dependence of
-the melting curve \citep{Kelley2010}. $x_t$, $y_t$, $K_m$ and $\gamma$ are parameters chosen in according to \citet{Kelley2010}.
-However, $X_{H_2O}$ depends on the melt fraction as
-\begin{equation}\label{eq:melt_wt}
-X_{H_2O}(M)=\frac{X_{H_2O}^{bulk}}{D_{H_2O}+M(1-D_{H_2O})}
-\end{equation}
-and a numerical solution can be found using a root-finder method \citep{Katz2003,Wang2016}. Furthermore, the content of water in the melt is limited by the
-pressure-dependent saturation concentration of water in the melt, determined as
-\begin{equation}\label{eq:melt_max}
-X_{H_2O}^{sat}=\chi_1 p^{\lambda_m}+\chi_2 p
-\end{equation}
-where $\chi_1$, $\chi_2$ and $\lambda_m$ are parameters chosen according to \citet{Katz2003}, \citet{Langmuir2006} and \citet{Kelley2010}. The correctness in
-the determination of melt fractions and water contents in the melt are tested comparing model results with experimental curves obtained by \citet{Katz2003}
-(Sec. \ref{sec:katz}).
-
-In case of melting, viscous creep viscosity ($\eta_{vp}$) is modified according to \citet{Wang2016}, as
-\begin{equation}\label{eq:melt_visc}
-\eta_{melt}=\eta_{vp}\cdot \exp(\alpha_m M)
-\end{equation}
-where $\alpha_m$ are melt fraction factors for dislocation and diffusion creep, chosen in according to \citet{Wang2016}.
-
-Effective density ($\rho_{eff}$) for partially molten rocks is calculated as
-\begin{equation}\label{eq:melt_dens}
-\rho_{eff}=\rho_s(1-M)+\rho_m M
-\end{equation}
-where $\rho_s$ and $\rho_m$ are the densities of the solid and the molten rock, respectively \citep{Gerya2003a,Wang2016}. The density of the solid rock is
-extracted by Perple\_X, while the density of the molten portion is calculated as
-\begin{equation}\label{eq:melt_dens_M}
-\rho_m(p,T)=\rho_0[1-\alpha(T-T_0)][1+\beta(p-p_0)]
-\end{equation}
-where $\rho_0$ is the density at the reference temperature ($T_0$) and pressure($p_0$), and $\alpha$ and $\beta$ are the thermal expansion and compressibility
-coefficients, respectively \citep{Gerya2003a,Wang2016}.
-
-As for the phase transitions, effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into account
-considering effective coefficient of thermal expansion. Similarly, latent heat due to melting/crystallisation can be implicitly included in the energy equation
-(Eq. \ref{eq:energy}) by considering effective specific heat ($C_{p_{eff}}$) and thermal expansion ($\alpha_{eff}$) of the partially molten rocks, in according to
-\begin{eqnarray}
-\label{eq:cp_effective}C_{p_{eff}}&=&C_p+H_L\left(\frac{\partial M}{\partial T}\right)_p \\
-\label{eq:alpha_effective}\alpha_{eff}&=&\alpha+\rho\frac{H_L}{T}\left(\frac{\partial M}{\partial p}\right)_T
-\end{eqnarray}
-where $C_p$ and $\alpha$ are the specific heat and the coefficient of thermal expansion of the solid rock, respectively, and $H_L$ is the latent heat of the
-molten rock \citep{Gerya2003a,Gerya2007,Tackley2010}.
-
-\section{Benchmarks}\label{sec:benchmark}
-\subsection{Error measurements}\label{sec:error}
-In order to determine the accuracy of velocity and pressure field of the benchmarks, the $L_2$-norm is computed by numerical integration on the Gauss-Legendre
-quadrature points. $L_2$-norm for pressure and velocity errors can be evaluated as
-\begin{eqnarray}
-\textrm{err}_p&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|p^n_i(r_q)-p^a_i(r_q)|^2\right)w_q|J_q|}\nonumber \\
-\textrm{err}_v&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|u^n_i(r_q)-u^a_i(r_q)|^2+|v^n_i(r_q)-v^a_i(r_q)|^2\right)w_q|J_q|}\nonumber
-\end{eqnarray}
-respectively, where $n_e$ is the number of elements, $n_q$ is the number of quadrature points per element, $p^n_i(r_q)$ and $p^a_i(r_q)$ are the numerical and
-analytical pressure, respectively, in each quadrature point $q$, $w_q$  and $J_q$ are the weight and the Jacobian at the quadrature point
-$q$, $u^n_i(r_q)$, $v^n_i(r_q)$, $u^a_i(r_q)$ and $v^a_i(r_q)$ are the numerical and analytical velocities, respectively, in each quadrature point $q$.
-
-Other quantities used as comparison with original benchmarks are the root-mean-square velocity over the whole domain and over the surface
-\begin{eqnarray}
-v_{\textrm{rms}}&=&\sqrt{\int_{\Omega}|\bm{v}|^2 d\Omega}\nonumber \\
-v_{\textrm{rms}}^{top}&=&\sqrt{\int_0^1 u^2 \bigg|_{y=1} dx}\nonumber
-\end{eqnarray}
-respectively, top and bottom Nusselt numbers
-\[\textrm{Nu}^{top/bottom}=-\int_0^1\frac{\partial T}{\partial y} \bigg|_{y=1/y=0} dx\]
-the average rate of work done against gravity
-\[\left\langle W \right\rangle=\int_{\Omega}T u_y d\Omega\]
-the average rate of viscous dissipation
-\[\left\langle \Phi \right\rangle=\int_{\Omega}\tau_{ij}\dot{\epsilon}_{ij} d\Omega\]
-and the percentage error between $\left\langle W \right\rangle$ and $\left\langle \Phi \right\rangle/Ra$
-\[\delta=\frac{\left|\left\langle W \right\rangle -\frac{\left\langle \Phi \right\rangle}{Ra} \right|}{\max \left(\left\langle W \right\rangle,
-\frac{\left\langle \Phi \right\rangle}{Ra}\right)}\times100\]
-
-\subsection{Stokes flow}\label{sec:stokes}
-The problem consists of determining velocity field $(u,v)$ and pressure $p$ in case of a manufactured solution with prescribed body forces such as
-\begin{eqnarray}
-b_1&=&(12-24y)x^4+(-24+48y)x^3+(-48y+72y^2-48y^3+12)x^2+\nonumber \\
-&&+(-2+24y-72y^2+48y^3)x+1-4y+12y^2-8y^3\nonumber \\
-b_2&=&(8-48y+48y^2)x^3+(-12y+72y-72y^2)x^2+(4-24y+48y^2-48y^3+\nonumber \\
-&&+24y^4)x-12y^2+24y^3-12y^4\nonumber
-\end{eqnarray}
-for which the exact solution is:
-\begin{eqnarray}
-u(x,y)&=&x^2(1-x)^2(2y-6y^2+4y^3)\nonumber \\
-v(x,y)&=&-y^2(1-y)^2(2x-6x^2+4x^3)\nonumber \\
-p(x,y)&=&x(1-x)-\frac{1}{6}\nonumber
-\end{eqnarray}
-The domain is a unit square a constant viscosity ($\eta=1$) and the penalty parameter is set to $\lambda=10^7$. Velocity boundary conditions are set to no slip
-$(\bm{v}=\bm{0})$ on all boundaries. The problem is performed for different grid resolution between \num{8x8} and \num{1024x1024} elements.
-
-The errors between the analytical solution and the numerical prediction of the velocity and the pressure field are calculated as in Section \ref{sec:error}.
-Fig. \ref{fig:errors}a shows that both velocity and pressure field converge to the exact solution with the decrease of the element size, following the
-theoretical convergence rate. The convergence of pressure, in contrast with the observation by \citet{Donea2003} for $Q_1 \times P_0$ elements, indicate the
-effectiveness of the smoothing procedure. Smoothed pressure field for grid resolution of \num{128x128} elements is shown in Fig. \ref{fig:errors}b (continuous
-red line), in comparison with the analytical solution (dashed blue line). Solve times and memory usage needed to generate the solution are shown in
-Fig. \ref{fig:MUMPS}. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Solver/Stokes_Flow}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/errors.pdf}
-\caption{Velocity and pressure error for the Stokes flow experiment between generated and analytical solution as a function of element size (panel a),
-and comparison between smoothed pressure and analytical solution as function of $x$ coordinate for a grid resolution of \num{128x128} elements (panel b).}
-\label{fig:errors}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/MUMPS.pdf}
-\caption{Panel a) Analysis, factorisation, solution and total solve times (panel a) and factorisation memory usage (panel b) as a function of the total number
-of degrees of freedom for the Stokes flow experiment.}
-\label{fig:MUMPS}
-\end{figure}
-
-\subsection{Markers advection}\label{sec:runge}
-The 4th-order Runge-Kutta advection scheme in space is tested by means of the Zalesak disk test \citep{Zalesak1979,Thieulot2014}. The benchmark is performed
-in a unit square domain with different grid resolution (\num{100x100} and \num{200x200} elements) and 1000000 markers. The Courant number varies between 0.1
-and 0.5 and the velocity field is prescribed in the entire domain as
-\begin{eqnarray}
-u(x,y)&=&2\pi\left(y-\frac{L_x}{2}\right)\nonumber \\
-v(x,y)&=&-2\pi\left(x-\frac{L_x}{2}\right)\nonumber
-\end{eqnarray}
-At $t=0$, the disk is centred at position $(0.5;0.75)$ with a radius $R=0.15$ and has a vertical fissure 0.05 wide and 0.2 high (orange in Fig. \ref{fig:runge}).
-As we expect, markers are back to their initial location after a $2\pi$ rotation ($t=1$, red, green and blue in Fig. \ref{fig:runge}), with an error of the
-final position that increase moving away from the centre, more evident in case of higher time steps or lower grid resolution (Fig. \ref{fig:runge_err}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Zalesak_Disk}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=200px]{./Figures/Zalesak.pdf}
-\caption{Comparison between the initial position of markers of the Zalesak disk experiment (orange) and their final position for different CFL in case of grid
-resolution of \num{100x100} elements. Blue, green and red markers indicate the final position of the disk in case of Courant number of 0.1, 0.25 and 0.5,
-respectively.}
-\label{fig:runge}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/zal_error.pdf}
-\caption{Error between final and initial position of markers of the disk as function of their distance from the centre of the domain. Results for different
-time steps are shown in blue, red and green for a grid resolution of \num{100x100} elements (for CFL of 0.1, 0.25 and 0.5, respectively) and in light blue,
-dark green and purple in case of \num{200x200} elements (for CFL of 0.1, 0.25 and 0.5, respectively).}
-\label{fig:runge_err}
-\end{figure}
-
-\subsection{Poiseuille flow}\label{sec:poiseuille}
-The domain is a rectangle with $L_x=2$ and $L_y=1$ and constant density and viscosity ($\rho=1$ and $\eta=1$), gravity acceleration $\bm{g}=\bm{0}$ and penalty
-parameter $\lambda=10^8$. The grid is composed by \num{40x20} elements. Velocity boundary conditions are set to no slip $(\bm{v}=\bm{0})$ at the top and the
-bottom, and a parabolic profile is imposed on the sides, with $u=y(1-y)$ and $v=0$. The analytical solution is then given by:
-\begin{eqnarray}
-u(x,y)&=&y(1-y)\nonumber \\
-v(x,y,)&=&0\nonumber \\
-p(x,y)&=&2\eta\left(\frac{L_x}{2}-x\right)\nonumber
-\end{eqnarray}
-
-The velocity field predicted by the model follows the expected parabolic profile (Fig. \ref{fig:poiseuille}a and Fig. \ref{fig:poi_plot}a) and, in case of the
-classic penalty method with no iterations, the pressure is clearly related to the divergence by means of the penalty parameter (Fig. \ref{fig:poiseuille}b and
-c and Fig. \ref{fig:poi_plot}b). Fig. \ref{fig:poiseuille}d shows that one Uzawa iteration is sufficient to bring the divergence down to \num{e-15}, with no
-correlation with the pressure field. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Poiseuille%20Flow}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Poiseuille.pdf}
-\caption{Velocity field (panel a), pressure (panel b) and divergence velocity of a Poiseuille flow in case of the classic penalty method (no iterations) and
-after one Uzawa iteration (panel c and d, respectively).}
-\label{fig:poiseuille}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Analytical.pdf}
-\caption{Velocity (panel a) and pressure (panel b) field predicted by the model for a Poiseuille flow with respect with their analytical solutions. The
-velocity field is plotted as function of the vertical coordinate in $L_x/2$ and the pressure is plotted as function of the x coordinate in $L_y/2$.}
-\label{fig:poi_plot}
-\end{figure}
-
-\subsection{Instantaneous 2D sphere}\label{sec:ist_sphere}
-The domain is a unit square with gravity $\bm{g}=(0,-1)$. The fluid has constant density and viscosity ($\rho_f=1$ and $\eta_f=1$). The sphere is in the middle
-of the domain with a radius R = 0.123456798 and has constant density and viscosity ($\rho_s=10^{-2}+\rho_f$ and $\eta_s=10^3 \cdot \eta_f$).
-We distinguish three different types of velocity boundary conditions:
-\begin{enumerate}
-\item FS: free slip conditions on all sides.
-\item NS: no slip conditions on all sides.
-\item OT: free slip conditions on the sides and the bottom, and open boundary at the top.
-\end{enumerate}
-All three cases are performed for different grid resolutions, between \num{16x16} and \num{512x512} elements with 50 randomly distributed markers per element,
-and different average schemes for the viscosity (harmonic, geometric and arithmetic). The velocity in the centre of the sphere $\bm{v}(0.5;0.5)$), minimum and
-maximum velocities ($u_{min}$, $u_{max}$, $v_{min}$, $v_{max}$) and pressures ($p_{min}$, $p_{max}$) on the entire domain, $v_{\textrm{rms}}$ and average pressure on the
-entire domain ($p_{avg}$) are compared with solutions generated by ASPECT \citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} and all results
-can be found at \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere2D} and 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Instantaneous_2D_Sphere}. Results in terms of the $v_{\textrm{rms}}$ for NS model 
-are also shown in Fig. \ref{fig:inst_sphere}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_NS.pdf}
-\caption{$v_{\textrm{rms}}$ as function of element size for the instantaneous 2D sphere experiment in case of no slip boundary conditions and with different average
-schemes for the viscosity. Results are compared with results obtained by other numerical codes.}
-\label{fig:inst_sphere}
-\end{figure}
-
-\subsection{Rayleigh-Taylor instability}\label{sec:rayleigh}
-This problem has been originally presented by \citet{vanKeken1997} and here it is performed as the isoviscous case (Case 1a). The domain has $L_x=0.9142$ and
-$L_y=1$ and gravity $\bm{g}=(0,-1)$. Two fluids with same constant viscosities ($\eta_1=\eta_2=1$) and different densities ($\rho_1=1000$ and $\rho_2=1010$),
-with the lighter fluid at the bottom. The initial interface between the fluids is given by $y(x)=0.2+0.02 \cos \left(\frac{\pi x}{L_x}\right)$. The experiment
-is performed with different grid sizes (\num{50x50}, \num{80x80}, \num{100x100} and \num{256x256}). A total of 1960000 markers are randomly distributed at the
-beginning of the simulation. Velocity boundary conditions are set to no slip at the top and the bottom, and to free slip at the sides of the domain.
-
-$v_{\textrm{rms}}$ as function of time is reported in Fig. \ref{fig:RT}, matching well with results shown by \citet{vanKeken1997}, \citet{Tackley2003} and
-\citet{Thieulot2014}. Fig. \ref{fig:rayleigh} shows the evolution of the experiment at different time steps. All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Rayleigh_Taylor_experiment/ISOVISCOUS}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/RT.pdf}
-\caption{$v_{\textrm{rms}}$ of the Rayleigh-Taylor experiment as function of time for different resolution of the grid.}
-\label{fig:RT}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=300px]{./Figures/Rayleigh.pdf}
-\caption{Evolution of the Rayleigh-Taylor experiment for a grid of \num{256x256} elements at $t=0, 500, 1000$ and $2000$ (panels a, b, c and d, respectively).}
-\label{fig:rayleigh}
-\end{figure}
-
-\subsection{Falling blocks}\label{sec:block}
-This benchmark is described as proposed by \citet{Gerya2003a}, \citet{Gerya2010b} and \citet{Thieulot2011}. The domain is square with $L_x=L_y=$ \SI{500}{\km}
-and the grid is composed by \num{50x50} elements with 25 markers in each element. The block is initially centred at ($x=$ \SI{250}{\km}; $y=$ \SI{400}{\km})
-and has a size of \num{100x100} km (Fig. \ref{fig:block}a). The fluid surrounding the block has $\eta_f=$ \SI{e21}{\pascal\per\s} and $\rho_f=$
-\SI{3200}{\kg\per\cubic\m}. The benchmark tests with different viscosities of the block, with $\eta_b$ from \SIrange{e15}{e27}{\pascal\per\s}. In each
-experiment also the density $\rho_b$ of the block varies from \SIrange{3220}{9900}{\kg\per\cubic\m}. Velocity boundary conditions are set to free slip
-conditions on all sides of the domain. Density distributions at $t=$ \SI{20}{\mega\year} for $\rho_b=$ \SI{3300}{\kg\per\cubic\m} and different $\eta_b$
-are plotted in Fig. \ref{fig:block}b-f, showing that the code correctly preserve the block geometry in case of large viscosity contrast, with a block stiffer
-than the surrounding fluid (Fig. \ref{fig:block}f).
-
-The velocity in the centre of the falling block at $t=0$ is measured for all experiments and, since the velocity should increase with density contrast, the
-quantity $\bm{v}/(\rho_b-\rho_f)$ is plotted as function of the viscosity contrast. All results perfectly match with those from \citet{Gerya2010b} and they
-line up on a single curve, demonstrating that the code can correctly deal with large viscosity and density contrasts (Fig. \ref{fig:falling}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Falling%20blocks}.
-
-\begin{figure}
-\noindent\includegraphics[width=300px]{./Figures/Block.pdf}
-\centering
-\caption{Density evolution of the falling block experiment at $t=$ \SI{0}{\mega\year} (panel a) and $t=$ \SI{20}{\mega\year} for different viscosities of the
-block (panels b-f).}
-\label{fig:block}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Falling.pdf}
-\caption{Initial velocity relative to the density contrast at the centre of the falling block as function of the viscosity contrast between the block and the
-surrounding fluid.}
-\label{fig:falling}
-\end{figure}
-
-\subsection{2D time-dependent Stokes sphere with deformable free surface}\label{sec:sphere}
-This experiment is performed in a unit square domain with the gravity acceleration fixed to $g_y=-1$. The sphere has $\rho_s=2$ and $\eta_s=10^3$ and is
-initially centred at ($0.5;0.6$) with radius $R=0.123456789$. The fluid surrounding the sphere has $\rho_f=1$ and $\eta_f=1$ and occupies the domain for
-$y\leq 0.75$, while for $y>0.75$ the air has $\rho_a=0$ and $\eta_a=10^{-3}$. Velocity boundary conditions are set to free slip on all sides. The Courant
-number is set to 0.25. The experiments run for 200 s using grid resolutions from \num{150x150} to \num{512x512} elements, with 25 randomly distributed markers
-per element, and for different average schemes for the viscosity. The interface between the fluid and the air is tracked by means of the markers chain. Results
-in terms of velocity and pressure fields and topography variations are compared with results obtained with ASPECT
-\citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} and they can be found at
-\url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere_fs2D} and
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Time_dependent_sphere}. Results in terms of the $v_{\textrm{rms}}$ in case of 
-an arithmetic average are also shown in Fig. \ref{fig:inst_spherefs}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_arithm.pdf}
-\caption{$v_{\textrm{rms}}$ for different grid resolutions as function of time for the 2D time-dependent Stokes sphere experiment in case of an arithmetic average.
-Results are compared with results obtained by ASPECT.}
-\label{fig:inst_spherefs}
-\end{figure}
-
-\subsection{Free surface stabilisation}\label{sec:stab}
-The benchmark is performed as discussed in \citet{Kaus2010a} and \citet{Thieulot2014}. The domain is square with $L_x=L_y=$ \SI{500}{\km} and a grid resolution
-of \num{200x200} elements, each of them containing 16 markers. A buoyant fluid with $\rho_1=$ \SI{3200}{\kg\per\cubic\metre} and $\eta_1=$ \SI{e20}{\pascal\s}
-is overlain by a denser fluid with $\rho_2=$ \SI{3300}{\kg\per\cubic\metre} and $\eta_2=$ \SI{e21}{\pascal\s}. The initial interface between the fluids has an
-initial sinusoidal shape of 5 km amplitude. Velocity boundary conditions are set to free slip conditions at the sides and no slip at the bottom, while the top
-is free surface. The experiment is performed with various fixed time steps, without the use of the Courant number. The vertical position of the free surface at
-$x=L_x$ is tracked for each simulation, with or without the stabilisation algorithm.
-
-The results show an instability (drunken sailor effect) increasing the time step in case that the stabilisation algorithm is not activated
-($dt=$ \SI{4200}{\year}, green line in Fig. \ref{fig:stab}), as already observed by \citet{Kaus2010a} and \citet{Thieulot2014}. Activating the algorithm the
-instability is fixed and the simulation remains stable using time step up to \SI{50000}{\year} (red line in Fig. \ref{fig:stab}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Stabilisation_algorithm}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Kaus_random.pdf}
-\caption{Evolution of the y coordinate in $x=L_x$ as function of time for different time steps, with and without the stabilisation algorithm.}
-\label{fig:stab}
-\end{figure}
-
-\subsection{Topography relaxation}\label{sec:crameri}
-This experiment is performed as in \citet{Crameri2012} using free surface and sticky air (Case 1). The domain is rectangular with $L_x=$ \SI{2800}{\km} and
-$L_y=$ \SI{700}{\km} in case of free surface and $L_y=$ \SI{800}{\km} in case of sticky air ($h_{st}=$ \SI{100}{\km}). The grid is composed by \num{256x64}
-elements for the free surface \num{560x320} elements for the sticky air. For both cases, there is a mantle of 600 km thickness with $\eta_m=$ \SI{e21}{\pascal\s}
-overlain by a highly viscous cosine-shaped lithosphere with thickness between 93 and 107 km and $\eta_l=$ \SI{e23}{\pascal\s}. Both mediums have $\rho=$
-\SI{3300}{\kg\per\cubic\metre}. In case of sticky air experiment, the air has $\rho_a=$ \SI{0}{\kg\per\cubic\metre} and $\eta_a=$ \SI{e18}{\pascal\s}. The
-gravity acceleration is set to $g_y=$ \SI{-10}{\metre\per\square\s}. Velocity boundary conditions are set to no slip at the bottom and free slip at the sides
-of the domain. In case of sticky air, velocities on top are set to free slip conditions.
-
-The maximum topography as function of time ($h(t)$) can be analytically derived in according to \[h(t)=h_0 \exp(\gamma t)\]
-where $h_0=7$ km is the initial topography, $\gamma=\num{-0.2139e-11}$ is the characteristic relaxation rate and $t$ is the time (black line in Fig.
-\ref{fig:crameri}). Maximum topography at the characteristic relaxation time $t_{rlx}=$ \SI{14.825}{\kilo\year} can be found to be $h_{rlx}=$ \SI{2576}{\m}.
-The results show that a viscosity of the air of \SI{e18}{\pascal\s} is needed in case of $h_{st}=$ \SI{100}{\km} to correctly describe the topography relaxation
-for this problem, as pointed out by \citet{Crameri2012} (purple line in Fig. \ref{fig:crameri}). The case with a true free surface well follows the analytical
-solution (orange line in Fig. \ref{fig:crameri}). In particular, at the relaxation time in case of the true free surface predicts a maximum topography of
-\SI{2570}{\m}, with an error of \SI{6}{\m} with respect to the analytical solution. Results are also compared with results obtained with a free surface in
-UNDERWORLD (\num{256x64} elements) and SULEC (\num{401x201} elements) (with courtesy from \citealp{Crameri2012}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Topography_relaxation}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Crameri.pdf}
-\caption{Maximum topography as function of time for the Crameri benchmark (coloured lines) in comparison with the analytical solution (black line) and results
-obtained with UNDERWORLD (yellow line) and SULEC (blue line).}
-\label{fig:crameri}
-\end{figure}
-
-\subsection{Spontaneous subduction}\label{sec:subduction}
-This experiment is performed as presented by \citet{Schmeling2008}, in case of both a sticky air and a true free surface and using different average schemes
-for the viscosities. The domain is rectangular with $L_x=$ \SI{3000}{\km} and $L_y=$ \SI{700}{\km}, a grid resolution of \num{256x64} elements with 50 markers
-per element and Courant number of 0.01. At the beginning of the simulation, a 100 km-thick lithospheric layer with $\rho_m=$ \SI{3300}{\kg\per\cubic\metre}
-and $\eta_m=$ \SI{e23}{\pascal\s} is located at the top of the domain for $x$ between 1000 and 3000 km, while the asthenospheric mantle has $\rho_m=$
-\SI{3200}{\pascal\s} and $\eta_m=$ \SI{e21}{\pascal\s}. In addition, a 200 km-depth lithospheric slab is already subducted in the mantle in order to have
-a spontaneous subduction. In case of sticky air, $L_y=750$ km and a 50 km-thick air layer with $\eta_a$ from \SIrange{e19}{e21}{\pascal\s} overlie the mantle.
-Velocity boundary conditions are set to free slip on the sides and on the bottom of the domain. In case of sticky air, velocities on the top boundary are set
-to free slip conditions as well.
-
-The maximum depth of the slab is tracked for all the simulations (Fig. \ref{fig:subduction}) and they are compared with results obtained by
-\citet{Schmeling2008} with models with regular grids and comparable grid resolutions (rectangles in Fig. \ref{fig:subduction}). Low viscous sticky air models
-(i.e., \SI{e19}{\pascal\s}) enlighten a strong dependence of the sinking velocity with the chosen average scheme (dashed lines in Fig. \ref{fig:subduction}),
-while high viscous sticky air models (i.e., \SI{e21}{\pascal\s}) show a higher resistance at the trench, underestimating the correct solution (dotted lines in
-Fig. \ref{fig:subduction}). A high resistance at the trench can be also observed in case of a true free surface for the low resolution grid used (continuous
-lines in Fig. \ref{fig:subduction}), as pointed out by \citet{Schmeling2008}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Spontaneous_subduction}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Subduction.pdf}
-\caption{Maximum depth of the slab as function of time for all the simulations of the \citet{Schmeling2008} benchmark, in case of harmonic, geometric and
-arithmetic means (red, green and blue lines, respectively) using a sticky air or a true free surface (discontinuous and continuous lines, respectively).
-Red, green and blue rectangular areas indicate the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{500}{\km} in case of sticky air,
-while the grey rectangular area indicates the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{600}{\km} in case of true free surface.}
-\label{fig:subduction}
-\end{figure}
-
-\subsection{The slab detachment}\label{sec:slab}
-The slab detachment benchmark is performed as described by \citet{Schmalholz2011} and \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{1000}{\km}
-and $L_y=$ \SI{660}{\km} and a grid resolution of \num{256x256} elements. A non-linear viscous T-shaped layer with $\rho_l=$ \SI{3300}{\kg\per\cubic\metre} is
-placed at the top of the domain and surrounded by a linear viscous fluid with $\rho_f=$ \SI{3150}{\kg\per\cubic\m}. The top layer is 80 km-thick and a 250
-km-long and 80 km-wide slab is placed at $x=L_x/2$ (Fig. \ref{fig:slab}a). The effective viscosity of the top layer is given by
-\[\eta_{eff}=\eta_0 I_2^{\frac{(1-n)}{n}}\]
-with $\eta_0=$ \SI{4.75e11}{\pascal\s} and $n=4$, while the surrounding fluid has $\eta_f=$ \SI{e21}{\pascal\s}. The effective viscosity is capped between
-$\eta_{min}=$ \SI{e21}{\pascal\s} and $\eta_{max}=$ \SI{e25}{\pascal\s}. Velocity boundary conditions are set to free slip at the top and the bottom, and to
-no slip at the sides of the domain.
-
-The sinking of the slab determines a decrease of the effective viscosity in the T-shaped layer during the first part of the evolution (Fig. \ref{fig:slab}a),
-as consequence of local strain rates, while effective viscosities increase up to $\eta_{max}$ in the top layer after approximately \SI{20}{\mega\year}, when
-necking is complete (Fig. \ref{fig:slab}b).
-
-The width of necking is tracked by means of markers position at the side of the slab. Necking width and time are normalised in function of the initial slab
-(\SI{80}{\km}) and the characteristic time ($t_c=$ \SI{7.1158e14}{\s}) \citep{Schmalholz2011,Glerum2018}. The results in case of arithmetic, geometric and
-harmonic mean of the viscosity are shown in Fig. \ref{fig:necking}, compared with results from \citet{Glerum2018} for which an infinity norm mean was used.
-The influence of the average schemes agrees with results shown in Section \ref{sec:subduction}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Slab_detachment}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Slab.pdf}
-\caption{Effective viscosity for the slab detachment benchmark at the beginning of the evolution (panel a) and when necking is complete (panel b).}
-\label{fig:slab}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Necking.pdf}
-\caption{Normalised width of the necking of the slab detachment benchmark as function of normalised time.}
-\label{fig:necking}
-\end{figure}
-
-\subsection{The indenter experiment}\label{sec:indenter}
-The indenter benchmark simulate a rigid punch on a purely plastic von Mises material, for which exists an analytical solution
-\citep{Thieulot2008,Thieulot2014,Glerum2018}. The domain is a unit square with a grid resolution of \num{256x256} elements. The gravity acceleration is set
-to $\bm{g}=\bm{0}$ and effective viscosity is capped using $\eta_{min}=10^{-4}$ and $\eta_{max}=10^3$. The tolerance for the convergence of the non-linear
-solution is $tol=10^{-9}$, with a maximum number of non-linear iterations set to 500. The medium in the domain has $\rho=0.01$, an initial viscosity
-$\eta_0=10$, cohesion $C=1$ and an angle of internal friction $\phi=0$°. Velocity boundary conditions are set to no slip at the bottom and free slip at the
-sides of the domain. The top of the domain is open with the exception of the central portion (punch area) with width $w_p=0.125$, where $v=-1.05$ and $u$ is
-fixed either to 0 (rough punch experiment) or free (smooth punch experiment).
-
-Fig. \ref{fig:indenter} show the results in terms of viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure
-(panels g and h) for both experiments. The analytical solution indicate that pressure at ($0.5;1$) and ($0.5 \pm w_p;1$)is $p=\pi +1$ and $p=1$, respectively
-(black and red lines in Fig. \ref{fig:smooth_p}). Fig. \ref{fig:smooth_p} shows a clear improvement in the pressure solution when passing from a rough to a
-smooth punch (green and blue lines in Fig. \ref{fig:smooth_p}, respectively), as previously observed by \citet{Thieulot2014} and \citet{Glerum2018}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Indenter}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Indenter.pdf}
-\caption{Viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure (panels g and h) fields for rough (left column) and
-smooth (right column) punch experiments.}
-\label{fig:indenter}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Smooth_Rough.pdf}
-\caption{Nodal pressure in function of the x coordinate at $y=1$ for rough (green line) and smooth (blue line) punch experiment. Black and red lines indicate
-the analytical solution at $x=0.5$ and $x=0.5 \pm w_p$, respectively.}
-\label{fig:smooth_p}
-\end{figure}
-
-\subsection{The brick experiment}\label{sec:brick}
-An instantaneous version of the brick benchmark is performed to verify the correctness of the pressure-dependent plasticity for different angle of internal
-friction, as proposed by \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{40}{\km} and $L_y=$ \SI{10}{\km} and a grid resolution of \num{512x128}
-elements. The gravity acceleration is set to $g_y=$ \SI{-10}{\m\square\s} and effective viscosity is capped using $\eta_{min}=$ \SI{e19}{\pascal\s} and
-$\eta_{max}=$ \SI{e26}{\pascal\s}. The tolerance for the convergence of the non-linear solution is $tol=10^{-7}$, with a maximum number of non-linear iterations
-set to 1000. A 800 m-wide and 400 m-height inclusion with $\eta_b=$ \SI{e20}{\pascal\s} and $\rho_b=$ \SI{2700}{\kg\per\cubic\m} is placed at the bottom of the
-domain at $x=L_x/2$. The inclusion is surrounded by a non-linear viscous medium with $\rho_m=$ \SI{2700}{\kg\per\cubic\m}, an initial viscosity of
-\SI{e23}{\pascal\s}, a linear viscous viscosity of \SI{e25}{\pascal\s} and a cohesion of \SI{40}{\mega\pascal}. Velocities are set to free slip conditions at
-the bottom of the domain and the top is open. Velocities on sides of the domain are fixed to $u=$ \SI{+-2e-11}{\m\per\s} and $v=0$. The experiment is performed
-in compressional and extensional contexts with an angle of internal friction $\phi$ of the non-linear medium variable between 0° and 30°.
-
-As expected, two shear bands stem from the inclusion with variable angles in relation to both the dynamics context and the internal friction angle. Shear band
-angles formed at 45° for $\phi=0$° in both extensional and compressional contexts (Fig. \ref{fig:brick_beam}a), while they are different in case of $\phi\neq0$°.
-Shear band angles in case of $\phi=20$° are shown for compression and extension in Fig. \ref{fig:brick_beam}b  and c, respectively. Values of shear band angles
-as function of internal friction angles from 0° to 30° are extracted at different depths and minimum and maximum values are plotted in Fig. \ref{fig:brick},
-in comparison with theoretical Roscoe, Arthur and Coulomb shear band angles. For all tests the tolerance for nonlinear convergence is defined $tol=10^{-7}$
-and none converges before the maximum number of iterations ($it_{max}=1000$) is reached. However, tests with internal friction angles up to 15° show a constant
-decrease of the velocity residuals, which can not be observed for for tests with higher internal friction angles (Fig. \ref{fig:convergence}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Brick_experiment}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Brick_Beam.pdf}
-\caption{Shear band angles predicted for the brick experiment for $\phi=0$° (panel a) and for $\phi=20$° in case of compression and extension (panels b and c,
-respectively).}
-\label{fig:brick_beam}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Brick.pdf}
-\caption{Shear band angles predicted for the brick experiment in case of compressional and extensional contexts as function of different internal angle of
-friction (continuous black line and red dots), compared with theoretical Roscoe, Artur and Coulomb angles (discontinuous black lines). Red lines indicate the
-range of angles calculated at different depths.}
-\label{fig:brick}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \noindent\includegraphics[width=500px]{./Figures/convergence.pdf}
-  \caption{Convergence rates for all tests. Normalised residuals (left column) are compared with velocity residuals (right column) for extensional (first row)
-  and compressional (second row) settings.}
-  \label{fig:convergence}
-  \end{figure}
-
-\subsection{Advection stabilisation}\label{sec:advection}
-The 1D advection problem proposed by \citet{Donea2003} and \citet{Thieulot2011} is performed to verify the effectiveness of the SUPG method to stabilise the
-advection term of the energy equation. The domain is a 1D segment with $L_x=1$ composed by 50 elements and a discontinuity in the thermal field placed at
-$x=0.25$. Temperature is set to 1 for $x \leq 0.25$ and to 0 for $x > 0.25$. Velocity is set to ${u}=1$ in the entire domain. The simulation is performed for
-250 time steps, with $dt=0.002$, so the thermal discontinuity should be at $x=0.75$ at the end of the simulation.
-
-Temperature profiles at the end of the simulation are shown in \ref{fig:advection} as function of the dimensionless coefficient $\gamma=\tau \bm{v}/h$. In case
-of the classic Galerkin method ($\gamma=0$, blue line in \ref{fig:advection}) the final thermal profile is characterised by strong oscillations, which are
-eliminated in case of the SUPG method ($\gamma=0.045$, orange line in \ref{fig:advection}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Advection_stabilisation}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Advection.pdf}
-\caption{Temperature profile as function of \textit{x} for the advection stabilisation benchmark. Purple line indicates the initial temperature profile; the
-green line indicates the analytical temperature profile after 250 time steps; blue line indicates the temperature profile after 250 time steps in case of the
-classic Galerkin method ($\gamma=0$); orange line indicates the temperature profile after 250 time steps in case of the SUPG method ($\gamma=0.045$).}
-\label{fig:advection}
-\end{figure}
-
-\subsection{Shear heating}\label{sec:simple_shear}
-This exercise is performed in an unit square domain composed by \num{128x128} elements. The velocity field is prescribed on the entire domain with
-$\bm{v}=(L_y-y)y\bm{e_x}$; viscosity, density and specific heat are fixed to 1, while thermal conductivity and radiogenic energy are fixed to 0. Therefore,
-the energy equation (Eq. \ref{eq:energy}) can be simplified as \[\frac{\partial T}{\partial t}=H_s\] and fixing $T(t=0)=0$, the temperature can be find as
-\[T(t)=H_s t\]
-In this case we have
-\begin{eqnarray}
-\dot{\epsilon_{xx}}&=&\frac{\partial u}{\partial x}=0 \nonumber \\
-\dot{\epsilon_{yy}}&=&\frac{\partial v}{\partial y}=0 \nonumber \\
-\dot{\epsilon_{xy}}&=&\frac{1}{2}\left(\frac{\partial v}{\partial x}+\frac{\partial u}{\partial y}\right)=\frac{1}{2}(L_y-2y)\nonumber
-\end{eqnarray}
-and, simplifying Eq. \ref{eq:shear}, shear heating can be calculated as \[H_s(x,y)=(1-2y)^2\]
-
-The solution predicted by the model in terms of velocity, temperature and shear heating match well with the analytical solutions (Fig. \ref{fig:analytical_en}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Simple_shear}.
-
-%\begin{figure}
-%\centering
-%\noindent\includegraphics[width=400px]{./Figures/Energy_Thieulot.pdf}
-%\caption{Temperature and shear heating predicted by the model after 1 time step.}
-%\label{fig:en_thieulot}
-%\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=500px]{./Figures/Analytical_Shear.pdf}
-\caption{Velocity (panel a), temperature (panel b) and shear heating (panel c) as function of the \textit{y} coordinate for the simple shear experiment. The
-solutions predicted by the model (red lines) are compared with the analytical solutions (dashed blue lines).}
-\label{fig:analytical_en}
-\end{figure}
-
-\subsection{Shear and adiabatic heating}\label{sec:shear}
-This problem is performed as presented in Exercise 9.4 in \citet{Gerya2010b}. Two materials are vertically separated in a rectangular domain with $L_x=$
-\SI{1000}{\km}, $L_y=$ \SI{1500}{\km} and a grid resolution of \num{30x20} elements. Constant thermal coefficient expansion ($\alpha=$ \SI{3e-5}{\per\kelvin}),
-temperature ($T=$ \SI{1300}{\kelvin}) and gravity acceleration ($g_y=$ \SI{-10}{\m\square\s}) are assumed in the whole domain. Fluid 1 (on the left side of the
-domain) has $\rho_1=$ \SI{3200}{\kg\per\cubic\m} and $\eta_1=$ \SI{e20}{\pascal\s}; fluid 2 (on the right side of the domain) has $\rho_2=$
-\SI{3300}{\kg\per\cubic\m} and $\eta_2=$ \SI{e22}{\pascal\s}. Velocity boundary conditions are set to free slip on all sides of the domain.
-
-As shown in Fig. \ref{fig:energy}, both shear and adiabatic heating predicted by the code (first row in Fig. \ref{fig:energy}) well recreate the results
-obtained by means of example ${Shear\_adiabatic\_heating.m}$ from \citet{Gerya2010b} (second row in Fig. \ref{fig:energy}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Adiabatic%2BShear}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Energy.pdf}
-\caption{Comparison between shear (first column), adiabatic (second column) and total (third column) energy predicted by the code (panels a, b and c) and those
-created using example ${Shear\_adiabatic\_heating.m}$ from Exercise 9.4 in \citet{Gerya2010b} (panels d, e and f).}
-\label{fig:energy}
-\end{figure}
-
-\subsection{Mantle convection}\label{sec:mantle}
-This problem is performed as presented by \citet{Blankenbach1989} (constant viscosity cases) and \citet{Thieulot2014}, in a 2D unit square domain with gravity
-acceleration $g_y=10^{10}Ra$. The experiment is performed with three different Rayleigh numbers ($Ra=10^4$, $10^5$ and $10^6$) and with different grid
-resolution (between \num{32x32} and \num{128x128} elements). The fluid has constant viscosity, initial density, heat capacity, thermal conductivity
-($\eta= \rho_0= C_p = k=1$), reference temperature ($T_0=0$) and thermal expansion coefficient ($\alpha=10^{-10}$). Temperatures are set to 0 on top and 1 on
-bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\]
-
-The solution generated by the code in terms of $v_{\textrm{rms}}$ (calculated as in Sec. \ref{sec:error}) and Nu as function of time are reported for all the
-simulations in Fig. \ref{fig:mantle}. At the steady state, $v_{\textrm{rms}}$ and Nu of all simulations converge well toward the values from \citet{Blankenbach1989},
-with lower errors for higher resolution grids (Table \ref{tab:mantle}). All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Mantle_convection}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Convection.pdf}
-\caption{$v_{\textrm{rms}}$ (panels a, c and e) and Nu (panels b, d and f) for the mantle convection benchmark as function of time for different grid resolution.
-Panels a and b show the results for $Ra=10^4$; Panels c and d show the results for $Ra=10^5$; panels e and f show the results for $Ra=10^6$. Purple lines
-indicate the convergence values for the $v_{\textrm{rms}}$ and the Nu from \citet{Blankenbach1989} at the steady state.}
-\label{fig:mantle}
-\end{figure}
-
-\begin{table}[H]
-\caption{Comparison between $v_{\textrm{rms}}$ and Nu predicted by the code for the mantle convection experiment and same values as reported in literature.}
-\centering
-\small
-\begin{tabular}{|>{\centering}m{0.09\textwidth}|>{\centering}m{0.05\textwidth}|>{\centering}m{0.24\textwidth}|>{\centering}m{0.15\textwidth}|
-  >{\centering}m{0.1\textwidth} >{\centering}m{0.1\textwidth} >{\centering\arraybackslash}m{0.1\textwidth}|}
-\toprule
-& & \multirow{2}{*}{\citet{Blankenbach1989}} & \citet{Thieulot2014} & \multicolumn{3}{c|}{FALCON}  \\
- & & & \num{200x200} & \num{32x32} & \num{64x64} & \num{128x128}  \\
-\midrule
-\multirow{2}{*}{$Ra=10^4$} & $v_{\textrm{rms}}$ & $42.864947 \pm 0.000020$ & 42.867 & 42.83226 & 42.852793 & 42.861394  \\
-           & Nu      & $4.884409 \pm 0.000010$  & 4.882  & 4.781297 & 4.857475  & 4.877573   \\
-\hline
-\multirow{2}{*}{$Ra=10^5$} & $v_{\textrm{rms}}$ & $193.21454 \pm 0.00010$  & 193.255& 193.872643 & 193.377472 & 193.252290  \\
-           & Nu      & $10.534095 \pm 0.000010$ & 10.507 & 9.602514 & 10.270735  & 10.465629   \\
-\hline
-\multirow{2}{*}{$Ra=10^6$} & $v_{\textrm{rms}}$ & $833.98977 \pm 0.00020$  & 834.712& 848.091176 & 837.767911 & 834.945793  \\
-           & Nu      & $21.972465 \pm 0.000020$ & 21.695 & 15.999266 & 19.703682  & 21.306939   \\
-\bottomrule
-\end{tabular}
-\label{tab:mantle}
-\end{table}
-
-\subsection{Viscoplastic mantle convection}\label{sec:visco_mantle}
-This benchmark is performed as Cases 1-5a described in \citet{Tosi2015} in a 2D unit square domain with two different grid resolutions of \num{32x32} 
-and \num{100x100} elements. Reference viscosity, reference density, heat capacity, thermal conductivity and thermal expansion coefficient are fixed to a constant
-value of 1 ($\eta_0= \rho_0= C_p = k=\alpha=1$), while the gravity has been chosen as $g_y=10^2$, in order to obtain a Rayleigh number $Ra=10^2$. Temperatures
-are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\].
-The viscosity ﬁeld $\eta$ is calculated as the harmonic average between a linear part $\eta_{lin}$ that depends on temperature only or on temperature and 
-depth $y$ and a nonlinear-plastic part $\eta_{pl}$ that depends on the strain rate $\dot{\bm{\epsilon}}$, as follows
-\[\eta(T,y,\dot{\bm{\epsilon}})=2\left(\frac{1}{\eta_{lin}(T,y)}+\frac{1}{\eta_{pl}(\dot{\bm{\epsilon}})}\right)^{-1}\]
-The linear and the nonlinear-plastic parts of the viscosity are calculated following \citet{Tosi2015} as follows
-\[\eta_{lin}(T,y)=\exp(-\ln(\Delta\eta_T)T+\ln(\Delta\eta_y)y)\]
-\[\eta_{pl}(\dot{\bm{\epsilon}})=\eta^*+\frac{\sigma_Y}{\sqrt{\dot{\bm{\epsilon}}:\dot{\bm{\epsilon}}}}\]
-where $\Delta\eta_T$, $\Delta\eta_y$, $\eta^*$ and $\sigma_Y$ are parameters chosen as listed in Table \ref{tab:tosi_setup}.
-
-For each case, some diagnostic quantities measured at the steady state are reported in Table \ref{tab:visco_mantle} (see Section \ref{sec:error} for 
-the computation of all the quantities), showing a well fit with results described in \citet{Tosi2015}. Data and charts for all cases can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Viscoplastic_mantle_convection}.
-
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{white!100!}{gray!25!}
-%\begin{table}[H]
-%  \caption{Results of the viscoplastic mantle convection benchmark for Cases 1-4 compared with results obtained with ELEFANT and YACC (from \citealp{Tosi2015}).}
-%  \centering
-%  \resizebox{\textwidth}{!}{
-%\begin{tabular}{|c|ccc|ccc|ccc|ccc|}
-%  \toprule
-%    & \multicolumn{3}{c|}{\textbf{Case 1}} & \multicolumn{3}{c|}{\textbf{Case 2}} & \multicolumn{3}{c|}{\textbf{Case 3}} &
-%    \multicolumn{3}{c|}{\textbf{Case 4}} \\
-%   & FALCON & ELEFANT& YACC & FALCON& ELEFANT& YACC & FALCON & ELEFANT & YACC & FALCON & ELEFANT & YACC \\
-%  \midrule
-%   $\left\langle T \right\rangle$ & 0.7759 & 0.7758 & 0.7767 & 0.6043 & 0.6036 & 0.6058 & 0.7275 & 0.7275 & 0.7286 & & 0.5277 & 0.5289\\
-%    Nu$^{top}$ & 3.4280 & 3.4214 & 3.4298 & 8.5242 & 8.5115 & 8.5278 & 3.0368 & 3.0347 & 3.0374 && 6.5912 & 6.5572 \\
-%    Nu$^{bottom}$ & 3.3168 & 3.313 & 3.3143 & 8.4599 & 8.4465 & 8.3990 & 3.0086 & 2.9908 & 2.9628 && 6.5834 & 6.5243 \\
-%    $v_{\textrm{rms}}$ & 249.901 & 249.134 & 251.7997 & 141.2451 & 140.7087 & 142.2020 & 100.2776 & 100.1208 & 100.9467 && 79.1105 & 79.6202 \\
-%    $v_{\textrm{rms}}^{top}$ & 1.8907 & 1.8642 & 1.8298 & 104.7928 & 104.1228 & 105.1767 & 2.0832 & 2.0652 & 2.0374 && 74.7596 & 75.8414 \\
-%    $u_{max}^{top}$ & 2.6340 & 2.6119 & 2.5516 & 121.9212& 121.766 & 122.3238 & 2.9086 & 2.9019 & 2.8458 && 88.9146 & 89.2940 \\
-%    $\eta_{min}$ & - & - & - & $2.0303\times10^{-5}$ & $2.0600\times10^{-5}$ & $1.9853\times10^{-5}$ & $4.8129\times10^{-5}$ & $4.8080\times10^{-5}$ & $4.7907\times10^{-5}$ &
-%    & $1.9860\times10^{-4}$ & $1.9174\times10^{-4}$ \\
-%    $\eta_{max}$ & - & - & - & 1.3325 & 1.85 & 1.6095 & 0.9360 & 0.9023 & 1 && 1.52 & 1.6773 \\
-%    $\left\langle W \right\rangle$ & 2.4299 & 2.4316 & 2.4583 & 7.5648 & 7.5665 & 7.6222 & 2.0387 & 2.0384 & 2.0400 && 5.6216 & 5.6512 \\
-%    $\left\langle \Phi \right\rangle/Ra$ & 2.4311 & 2.4276 & 2.4333 & 7.5661 & 7.5585 & 7.6194 & 2.0386 & 2.0370 & 2.0335 && 5.6182 & 5.6463 \\
-%    $\delta$ & 0.05\% & 0.16\% & 1.02\% & 0.02\% & 0.11\% & 0.04\% & $<$0.01\% & 0.07\% & 0.32\% && 0.06\% & 0.09\% \\
-%  \bottomrule
-%  \end{tabular}}
-%  \label{tab:visco_mantle}
-%  \end{table}
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{2}{white!100!}{gray!25!}
-\begin{table}[H]
-  \caption{Benchmark cases and corresponding parameters.}
-  \centering
-  \small
-\begin{tabular}{|>{\centering}m{0.08\textwidth}|>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}
-  >{\centering}m{0.05\textwidth} >{\centering}m{0.05\textwidth} |>{\centering\arraybackslash}m{0.2\textwidth}|}
-  \toprule
-    Case & $Ra$ & $\Delta\eta_T$ & $\Delta\eta_y$ & $\eta^*$ & $\sigma_Y$ & Convective regime \\
-  \midrule
-  \rowcolor{gray!25!}
-    1 & $10^2$ & $10^5$ & 1 & - & - & Stagnant lid \\
-    2 & $10^2$ & $10^5$ & 1 & $10^{-3}$ & 1 & Mobile lid  \\
-    \rowcolor{gray!25!}
-    3 & $10^2$ & $10^5$ & 10 & - & - & Stagnant lid  \\
-    4 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 1 & Mobile lid \\
-    \rowcolor{gray!25!}
-    5 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 4 & Periodic \\
-  \bottomrule
-  \end{tabular}
-  \label{tab:tosi_setup}
-  \end{table}
-
-\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{gray!25!}{white!100!}
-\begin{table}
-  \caption{Results of the viscoplastic mantle convection benchmark for different grid resolutions in Cases 1-5 compared with results obtained with ELEFANT
-  and YACC ($100\times100$ elements, from \citealp{Tosi2015}) and in Stone 28 ($32\times32$ elements, from
-  \url{https://github.com/cedrict/fieldstone/tree/master/python_codes/fieldstone_28}).}
-  \centering
-  \resizebox{\textwidth}{!}{
-\begin{tabular}{|cc|ccccccccccc|}
-  \toprule
-  \multicolumn{13}{|c|}{\textbf{Case 1}} \\
-  \midrule
-  Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-  & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-  \hline
-  \rowcolor{gray!25!}
-    YACC & $100\times100$ & 0.7767 & 3.4298 & 3.3143 & 251.7997 & 1.8298 & 2.5516 & - & - & 2.4583 & 2.4333 & 1.02\% \\
-    ELEFANT & $100\times100$ & 0.7758 & 3.4214 & 3.313 & 249.134 & 1.8642 & 2.6119 & - & - & 2.4316 & 2.4276 & 0.16\% \\
-    \rowcolor{gray!25!}
-    Stone 28 & $32\times32$ & 0.7737 & 3.3987 & - & 243.872 & - & - & - & - & - & - & - \\
-%    FALCON & $32\times32$ & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & - & - & 2.4745 & 2.4864 & 0.48\% \\
-%    \rowcolor{gray!25!}
-    FALCON & $100\times100$ & 0.7759 & 3.4280 & 3.3168 & 249.901 & 1.8907 & 2.6340 & - & - & 2.4299 & 2.4311 & 0.05\% \\
-    \midrule
-    \multicolumn{13}{|c|}{\textbf{Case 2}} \\
-    \midrule
-    Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-    & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-    \hline
-    \rowcolor{gray!25!}
-     YACC & $100\times100$ & 0.6058 & 8.5278 & 8.3990 & 142.2020 & 105.1767 & 122.3238 & $1.9853\times10^{-5}$ & 1.6095 & 7.6222 & 7.6194 & 0.04\% \\
-     ELEFANT & $100\times100$ & 0.6034 & 8.5115 & 8.4465 & 140.8067 & 104.1228 & 121.766 & $2.0600\times10^{-5}$ & 1.85 & 7.5665 & 7.5585 & 0.11\% \\
-     \rowcolor{gray!25!}
-     Stone 28 & $32\times32$ & 0.6052 & 8.1715 & - & 141.518 & - & - & - & - & - & - & - \\
-%     FALCON & $32\times32$ & 0.6454 & 8.1422 & 7.4020 & 156.6908 & 95.7226 & 116.9302 & $2.2572\times10^{-5}$ & 0.7194 & 7.4289 & 7.4452 & 0.22\% \\
-%     \rowcolor{gray!25!}
-     FALCON & $100\times100$ & 0.6044 & 8.5266 & 8.4590 & 141.3684 & 104.8867 & 122.0470 & $2.0333\times10^{-5}$ & 1.2914 & 7.5707 & 7.5718 & 0.01\% \\
-     \midrule
-     \multicolumn{13}{|c|}{\textbf{Case 3}} \\
-     \midrule
-     Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-     & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-     \hline
-     \rowcolor{gray!25!} 
-     YACC & $100\times100$ & 0.7286 & 3.0374 & 2.9628 & 100.9467 & 2.0374 & 2.8458 & $4.7907\times10^{-5}$ & 1 & 2.0400 & 2.0335 & 0.32\% \\
-      ELEFANT & $100\times100$ & 0.7275 & 3.0347 & 2.9908 & 100.1208 & 2.0652 & 2.9019 & $4.8080\times10^{-5}$ & 0.9023 & 2.0384 & 2.037 & 0.07\% \\
-      \rowcolor{gray!25!}
-      Stone 28 & $32\times32$ & 0.7277 & 3.0347 & - & 100.018 & - & - & - & - & - & - & - \\
-%      FALCON & $32\times32$ & 0.7281 & 3.0618 & 2.8132 & 101.0092 & 2.1630 & 3.0178 & $4.8500\times10^{-5}$ & 0.7845 & 2.0699 & 2.0695 & 0.02\% \\
-%      \rowcolor{gray!25!}
-      FALCON & $100\times100$ & 0.7275 & 3.0368 & 3.0086 & 100.2776 & 2.0832 & 2.9086 & $4.8129\times10^{-5}$ & 0.9360 & 2.0387 & 2.0386 & $<$0.01\% \\
-      \midrule
-      \multicolumn{13}{|c|}{\textbf{Case 4}} \\
-      \midrule
-      Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-      & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-      \hline
-      \rowcolor{gray!25!} 
-      YACC & $100\times100$ & 0.5289 & 6.5572 & 6.5243 & 79.6202 & 75.4814 & 89.2940 & $1.9174\times10^{-4}$ & 1.6773 & 5.6512 & 5.6463 & 0.09\% \\
-       ELEFANT & $100\times100$ & 0.5277 & 6.5912 & 6.5834 & 79.1105 & 74.7596 & 88.9146 & $1.9860\times10^{-4}$ & 1.5200 & 5.6216 & 5.6182 & 0.06\% \\
-       \rowcolor{gray!25!}
-       Stone 28 & $32\times32$ & 0.5286 & 6.4036 & - & 79.4746 & - & - & - & - & - & - & - \\
-%       FALCON & $32\times32$ & 0.5334 & 6.4679 & 6.3435 & 80.6153 & 76.0511 & 90.2058 & $2.0393\times10^{-4}$ & 0.6453 & 5.6676 & 5.6698 & 0.04\% \\
-%       \rowcolor{gray!25!}
-       FALCON & $100\times100$ & 0.5279 & 6.5897 & 6.5830 & 79.1220 & 75.0715 & 88.8484 & $1.9693\times10^{-4}$ & 1.4133 & 5.6127 & 5.6113 & 0.03\% \\
-       \midrule
-       \multicolumn{13}{|c|}{\textbf{Case 5}} \\
-       \midrule
-       Code & Resolution & $P$ & $\left\langle T \right\rangle_{min}$ & $\left\langle T \right\rangle_{max}$ & Nu$^{top}_{min}$ & Nu$^{top}_{max}$ &
-       $v_{\textrm{rms}}^{min}$ & $v_{\textrm{rms}}^{max}$ & \multicolumn{2}{c}{$(\left\langle \Phi \right\rangle/Ra)_{min}$} & 
-       \multicolumn{2}{c|}{$(\left\langle \Phi \right\rangle/Ra)_{max}$} \\
-       \hline
-       \rowcolor{gray!25!}
-         YACC & $100\times100$ & 0.0792 & 0.6543 & 0.6722 & 2.6950 & 7.2792 & 41.9813 & 98.8230 & \multicolumn{2}{c}{1.4347} & 
-         \multicolumn{2}{c|}{9.2748} \\
-         ELEFANT & $100\times100$ & 0.076 & 0.6514 & 0.6693 & 2.6808 & 7.3506 & 41.2578 & 100.089 & \multicolumn{2}{c}{1.3324} & 
-         \multicolumn{2}{c|}{9.5289} \\
-         \rowcolor{gray!25!}
-         Stone 28 & $32\times32$ & 0.0764 & 0.6521 & 0.6697 & 2.6912 & 7.0763 & 41.61 & 98.183 & \multicolumn{2}{c}{-} & 
-         \multicolumn{2}{c|}{-} \\
-     %    FALCON & $32\times32$ & & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & \multicolumn{2}{c}{-} & 
-     %     \multicolumn{2}{c|}{-} \\
-     %    \rowcolor{gray!25!}
-         FALCON & $100\times100$ & 0.0775 & 0.6519 & 0.6697 & 2.6847 & 7.3629 & 41.5256 & 99.5466 & \multicolumn{2}{c}{1.3399} & 
-         \multicolumn{2}{c|}{9.3993} \\
-  \bottomrule
-  \end{tabular}}
-  \label{tab:visco_mantle}
-  \end{table}
-
-\subsection{Thin layer entrainment}\label{sec:thin}
-This experiment is performed as originally proposed by \citet{vanKeken1997} with two incompressible fluids in a rectangular domain with $L_x=2$ and $L_y=1$
-and gravity acceleration $g_y=10^{10}Ra$. The Rayleigh number and the compositional Rayleigh number are fixed to $Ra=300000$ and $Rc=450000$, respectively.
-Both fluids have constant viscosity, thermal conductivity, specific heat ($\eta=\rho=C_p=1$) and thermal expansion coefficient ($\alpha=10^{-10}$). Fluid 1
-has a density $\rho_1=1$, while fluid 2 is denser ($\rho_2=\rho_1+1.5 \cdot 10^{-10}$) and is placed at the bottom of the domain, for $y \leq 0.025$.
-Temperature are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides of the domain. The initial
-temperature field is given by
-\[T(x,y)=T_u(x,y)+T_l(x,y)+T_r(x,y)+T_s(x,y)-\frac{3}{2}\]
-with
-\begin{eqnarray}
-T_u(x,y)&=&\frac{1}{2}erf\left(\frac{1-y}{2}\sqrt{\frac{u_0}{x}}\right)\nonumber \\
-T_l(x,y)&=&1-\frac{1}{2}erf\left(\frac{y}{2}\sqrt{\frac{u_0}{L_x-x}}\right)\nonumber \\
-T_r(x,y)&=&\frac{1}{2}+\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{y+1}} \exp\left(-\frac{x^2u_0}{4y+4}\right)\nonumber \\
-T_s(x,y)&=&\frac{1}{2}-\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{2-y}} \exp\left(-\frac{(L_x-x)^2u_0}{8-4y}\right)\nonumber
-\end{eqnarray}
-and
-\begin{eqnarray}
-u_0&=&\frac{L_x^{7/3}}{(1+L_x^4)^{2/3}}\left(\frac{Ra}{2\sqrt{\pi}}\right)^{\frac{2}{3}}\nonumber \\
-Q&=&2\sqrt{\frac{L_x}{\pi u_0}}\nonumber
-\end{eqnarray}
-
-The experiment is performed with two grid resolution (\num{125x40} and \num{200x80} elements), with 100 markers per element and CFL of 0.25. $v_{\textrm{rms}}$
-(see Sec. \ref{sec:error}) calculated for both the simulations match well with results from \citet{vanKeken1997} and obtained with ASPECT
-\citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} and ELEFANT \citep{Thieulot2014}, obtained with similar grid resolutions
-(Fig. \ref{fig:thin}). In addition, a variety of simulations with different resolutions and aspect ratios are performed to verify that the sensitivity of the
-initial velocity field. Results match well with results from \citet{vanKeken1997} (in the grey area) and \citet{Thieulot2014} (Fig. \ref{fig:thin_initial}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Thin_layer}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_thin.pdf}
-\caption{$v_{\textrm{rms}}$ for the thin layer experiment as function of time for different grid resolution. Results are compared with results obtained by
-\citet{vanKeken1997} (black lines) and with ASPECT and ELEFANT.}
-\label{fig:thin}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrmszero.pdf}
-\caption{$v_{\textrm{rms}}$ at $t=0$ for the thin layer experiment as function of the element size for different aspect ratios of the elements. Results are compared
-with results obtained by \citet{vanKeken1997} (black dots in the grey area) and with ASPECT and ELEFANT.}
-\label{fig:thin_initial}
-\end{figure}
-
-\subsection{Sinking hydrated cylinder}\label{sec:quinquis}
-This experiment is performed as explained in \citet{Quinquis2014} and simulates the sinking of a cold, hydrated cylinder into a warm, dry mantle. The domain
-is square with $L_x=L_y=$ \SI{300}{\km}, a grid resolution of \num{300x300} elements and 25 markers per element. The sinking cylinder is located at $x=$
-\SI{150}{\km} and $y=$ \SI{170}{\km} with a radius of \SI{20}{\km} and has $\rho_c=$ \SI{3250}{\kg\per\cubic\m}, $\eta_c=$ \SI{e23}{\pascal\s}, a thermal
-conductivity $k_c=$ \SI{4.5}{\watt\per\m\per\kelvin},a specific heat $C_{p_c}=$ \SI{1250}{\joule\per\kg\per\kelvin}; the surrounding mantle has $\rho_m=$
-\SI{3200}{\kg\per\cubic\m}, $\eta_m=$ \SI{e20}{\pascal\s}, a thermal conductivity $k_m=$ \SI{105}{\watt\per\m\per\kelvin} and specific heat $C_{p_m}=$
-\SI{1250}{\joule\per\kg\per\kelvin}; a 58 km-thick lithosphere overlies the mantle and has $\rho_l=$ \SI{3200}{\kg\per\cubic\m}, $\eta_l=$ \SI{e23}{\pascal\s},
-a thermal conductivity $k_l=$ \SI{4.5}{\watt\per\m\per\kelvin}, specific heat $C_{p_l}=$ \SI{750}{\joule\per\kg\per\kelvin} and a thermal diffusivity of
-\SI{e-6}{\square\m\per\s}. Velocity boundary conditions are set to free slip on all sides of the domain. Temperature increase linearly in the lithosphere,
-from 0° to 1300°C, and in the mantle, from 1300° to 1360.5°C. The initial temperature of the cylinder is fixed to 400°C. Throughout the evolution, temperature
-is fixed to 0° and 1360.5°C at the top and bottom of the domain, respectively. The initial bound water content is imposed at 2 wt.\% in the cylinder and at
-0 wt.\% in both the mantle and the lithosphere. The maximum amount of water in the mantle is fixed to 0.2 wt.\%, while maximum water content in the cylinder
-and in the lithosphere is function of pressure and temperature and is calculated using a serpentinized harzburgite with Perple\_X, as in \citet{Quinquis2014}.
-
-As shown in \citet{Quinquis2014}, the model is characterised by a progressive dehydration in the external portion of the cylinder, due to the increase of
-temperature, with a consequent vertical migration of free water that hydrates the mantle above the cylinder, up to the lithosphere. The distribution of
-bound and free water during the evolution of the experiment is shown in Fig. \ref{fig:quinquis} (left and right columns, respectively). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Hydration}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Hydration.pdf}
-\caption{Distribution of bound and free water (left and right columns, respectively) at $t=0.5$ and $t=$ \SI{1.25}{\mega\year} (first and second rows,
-respectively) for the sinking hydrated cylinder experiment. Only markers with bound or free water not equal to 0 are plotted.}
-\label{fig:quinquis}
-\end{figure}
-
-\subsection{Mantle melting curves}\label{sec:katz}
-This instantaneous experiment is performed in order to recreate isobaric melting curves, as function of temperature, bulk water and content of water in the
-melt, to compare with melting curves obtained by \citet{Katz2003}. The domain is square with $L_x=L_y=100$ km, a grid resolution of \num{100x100} elements and
-25 markers per element. Density is set in the entire domain to \SI{3300}{\kg\per\cubic\m} and temperature increases from the sides to the centre from 1000° to
-1500°C. The experiment is performed for bulk water contents from 0 to 0.3 wt.\%.
-
-Fig. \ref{fig:fraction} shows isobaric melting curves for \SI{1}{\giga\pascal} (panel a) and \SI{3}{\giga\pascal} (panel b) calculated for different bulk water
-contents. The curves well match with those obtained for the mantle by \citet{Katz2003}, showing the saturation at the solidus for bulk water of 0.2 and 0.3 wt.\%, indicating
-that the code correctly determines melt fraction as function of pressure, temperature and bulk water content. Same results can be observed in Fig.
-\ref{fig:water}a, where isobaric and isothermal melting curves are shown as function of bulk water content. In addition, Fig. \ref{fig:water}b shows that the
-code correctly determines also the percentage of water in the melt as function of pressure, temperature and melt fraction. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Melting}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Fraction.pdf}
-\caption{Isobaric melting curves for 1 GPa (panel a) and 3 GPa (panel b) as function of temperature with different bulk water contents.}
-\label{fig:fraction}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Water.pdf}
-\caption{Isobaric melting curves for 1.5 GPa as function of bulk water (panel a) and water in the melt (panel b) calculated for different temperatures.}
-\label{fig:water}
-\end{figure}
-
-\clearpage
-
-\bibliography{export}
-
-\end{document}
+In case of sticky air, slope is determined by means of the markers chain, which is vertically corrected considering erosion and sedimentation rates. After the
+correction, continental markers above the markers chain are transformed in air markers and air markers below the markers chain become sediments. Similarly,
+slope is determined considering the nodes on the top boundary in case of true free surface and they are vertically corrected as for the markers chain. In this
+case, continental markers above the top boundary are simply deleted.

### 2022-09-07 tex/healing.tex
update main

--- 2022-07-20 FALCON_description.tex
@@
-\documentclass[hidelinks,10pt,a4paper]{article}
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{textcomp}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage{booktabs}
-\usepackage{multirow}
-\usepackage[cm]{fullpage}
-\usepackage[round]{natbib}
-\bibliographystyle{abbrvnat}
-\usepackage{doi}
-\usepackage{bm}
-\usepackage[table]{xcolor}
-\hypersetup{
-    colorlinks,
-    linkcolor={blue},
-    citecolor={blue},
-    urlcolor={blue}
-}
-\usepackage{float}
-\restylefloat{table}
-\usepackage{siunitx}
-\DeclareSIUnit\year{yr}
-
-\begin{document}
-
-\section{Introduction}\label{sec:intro}
-This document includes a description of the code FALCON, with a presentation of the features implemented followed by results of extensive benchmarks listed in
-Table \ref{tab:list}, performed to verify that the code solves correctly a wide range of problems. Description and results of all benchmarks are shown
-in Section \ref{sec:benchmark}.
-
-\begin{table}[H]
-\caption{List of benchmarks performed for each feature of the code. Data of all benchmarks can be found at 
-\url{https://github.com/aleregorda/Benchmarks}}
-\centering
-\begin{tabular}{| l | l | l |}
-\toprule
- \textbf{Feature}  & \textbf{Benchmark} & \textbf{Section} \\
-\midrule
-  Solver  & Stokes flow & \ref{sec:stokes} \\
-\hline
-  \multirow{2}{*}{Markers advection} & Zalesak disk & \ref{sec:runge} \\
-    & Conservative Interpolation Velocity & \ref{sec:CVI} \\
-\hline
-  \multirow{4}{*}{Momentum equation}  & Poiseuille flow & \ref{sec:poiseuille}\\
-    & Instantaneous 2D sphere & \ref{sec:ist_sphere} \\
-    & Rayleigh-Taylor experiment & \ref{sec:rayleigh} \\
-    & Falling block & \ref{sec:block} \\
-\hline
-    & 2D Stokes sphere & \ref{sec:sphere} \\
-    Sticky air and  & Stabilisation algorithm & \ref{sec:stab} \\
-    free surface & Topography relaxation & \ref{sec:crameri} \\
-  & Spontaneous subduction & \ref{sec:subduction} \\
-\hline
-  Erosion and sedimentation  & - & - \\
-\hline
-   Non-linear  & Slab detachment & \ref{sec:slab} \\
-   visco-plasticity & Indenter & \ref{sec:indenter} \\
-    & Brick & \ref{sec:brick} \\
-\hline
-  \multirow{3}{*}{Energy equation}  & Advection stabilisation & \ref{sec:advection} \\
-    & Simple shear heating & \ref{sec:simple_shear} \\
-    & Shear and adiabatic heating & \ref{sec:shear} \\
-\hline
-\multirow{3}{*}{Energy + momentum} & Mantle convection & \ref{sec:mantle} \\
-& Viscoplastic mantle convection & \ref{sec:visco_mantle} \\
-& Thin layer entrainment & \ref{sec:thin} \\
-\hline
-  Phase changes  & \multirow{2}{*}{Hydrated sinking cylinder} & \multirow{2}{*}{\ref{sec:quinquis}} \\
-  and hydration  &  & \\
-\hline
-  Melting & Experimental melting curves & \ref{sec:katz} \\
-\bottomrule
-\end{tabular}
-\label{tab:list}
-\end{table}
-
-The code is written in Fortran90 and uses the finite element method (FEM) with quadrilateral $Q_1 \times P_0$ elements (continuous bilinear velocity and
-discontinuous constant pressure), associated with MUMPS\footnote{\url{http://mumps.enseeiht.fr/}} \citep{Amestoy2001,Amestoy2019} that is a software package
-for solving systems of linear equations of the form $A \cdot x = b$, where \textit{A} is a square sparse matrix, by means of a direct method. Since
-$Q_1 \times P_0$ elements do not satisfy Ladyzhenskaya, Babuska and Brezzi (LBB) stability condition \citep{Donea2003}, elemental pressure is elaborated
-in post-processing to avoid spurious pressures by means of a double interpolation to smooth the pressure. Following this procedure, the elemental pressure
-is interpolated onto nodes and then back onto elements. After the smoothing procedure, the pressure field is calculated again on the nodes to be used in
-combination with the lithostatic pressure, which is calculated onto nodes. Correctness of the solution and performances of the solver in terms of time and
-memory usage are tested solving the Stokes flow with the analytical solution proposed by \citet{Donea2003} (Section \ref{sec:stokes}).
-
-The thermo-mechanics of crust-mantle systems is described by means of the conservation of mass, momentum and energy equations, expressed as follows:
-\begin{eqnarray}
-\label{eq:mass}\frac{\partial \rho}{\partial t}+\bm{\nabla} \cdot (\rho \bm{u})&=&0\\
-\label{eq:momentum}-\bm{\nabla} p + \bm{\nabla} \cdot \bm{\tau} + \rho \bm{g}&=&0\\
-\label{eq:energy}\rho C_p \left(\frac{\partial T}{\partial t} + \bm{u} \cdot \bm{\nabla} T \right)&=&\bm{\nabla} \cdot \left(k\bm{\nabla} T\right) + H_{tot}
-\end{eqnarray}
-where $\rho$ is the density, $\bm{u}$ is the velocity, \textit{p} is the pressure, $\bm{\tau}$ is the deviatoric stress, $\bm{g}$ is the gravity acceleration,
-$C_p$ is the specific heat at constant pressure, \textit{T} is the temperature, \textit{k} is the thermal conductivity and $H_{tot}$ is the total internal heat
-production. Density variations due to temperature are generally small enough to assume the density as constant ($\rho=\rho_0$) in Eq. \ref{eq:mass} and in
-Eq. \ref{eq:energy}, while it must be treated as a variable in the buoyancy term of Eq. \ref{eq:momentum}, such that \[\rho=\rho_0(1-\alpha(T-T_0))\]
-where $\rho_0$ is the density at a reference temperature $T_0$ and $\alpha$ is the coefficient of thermal expansion. Eq. \ref{eq:mass} then can be rewritten as
-\begin{equation}\label{eq:mass_0}
-\bm{\nabla} \cdot \bm{u}=0
-\end{equation}
-This simplification is known as the extended Boussinesq approximation.
-
-The time step $dt$ is chosen in according to Courant-Friedrichs-Lewy condition \citep{Anderson1995}, such that
-\[dt=C_n\; \min\left(\frac{\Delta x}{\max|\bm{u}|},\frac{(\Delta x)^2}{\kappa}\right)\]
-where $0<C_n<1$ is the Courant (CFL) number, $\Delta x$ is the minimum dimension of smallest element, $\max|\bm{u}|$ is the maximum velocity calculated on the
-entire domain and
-$\kappa=\frac{k}{\rho C_p}$ is the thermal diffusivity \citep{Thieulot2014}.
-
-\section{Lagrangian markers}\label{sec:markers}
-Elemental properties (density, viscosity, thermal conductivity, specific heat and thermal expansion) needed to solve Eqs. \ref{eq:mass_0}, \ref{eq:momentum}
-and \ref{eq:energy} are related to the composition of each element, determined by means of Lagrangian markers that are characteristic of different materials
-in the domain. Elemental properties (with the sole exception of the viscosity) are calculated using an arithmetic mean, as \[P_e=\frac{1}{n}\sum_{i=1}^n P_i\]
-where $P_e$ is the elemental property, $P_i$ is the property characteristic of the material of each marker and $n$ is the number of markers of the element.
-Differently, the average scheme for the elemental viscosity can be chosen between harmonic, geometric and arithmetic mean. Markers can be placed regularly or
-randomly at the beginning of the simulation and their advection is performed by either a 2nd-order or a 4th-order Runge-Kutta in space, interpolating the velocity 
-field on each marker by means of the shape functions (see Section \ref{sec:runge}). The interpolated velocity is then corrected by means of the Conservative
-Velocity Interpolation (CVI), which introduces a corrective term to reduce dispersion and clustering of particles in both steady state and time-dependent 
-\citep{Wang2015} (see Section \ref{sec:CVI}).
-
-The initial distribution of the markers is created using the open-source code library
-Geodynamic World Builder \citep{Fraters2019}. During the simulation, each marker carries memory of temperature, pressure and accumulated strain, which are
-determined interpolating the relative nodal parameter. In particular, pressure is interpolated onto each marker after the smoothing procedure. The number
-of markers contained in each element is maintained between $n_{min}$ and $n_{max}$. When in an element there are less markers than $n_{min}$ the code adds
-random markers to reach the $n_{min}$, while if the number is higher than $n_{max}$ some of them are randomly deleted. When new markers are added, they assume
-the properties of the nearest marker. In this way elements are never empty and maintain a number of markers inside a prefixed range.
-
-\section{The momentum equation}\label{sec:mom_eq}
-The penalty method is implemented so that Eq. \ref{eq:mass_0} can be written as:
-\begin{equation}\label{eq:mass_penalty}
-\bm{\nabla} \cdot \bm{u} + \frac{p}{\lambda}=0
-\end{equation}
-where $\lambda$ is the so-called penalty parameter, which should be 6-7 orders of magnitude larger than the shear viscosity to ensure that mass conservation
-is satisfied \citep{Donea2003,Thieulot2014}.
-
-The deviatoric stress tensor in Eq. \ref{eq:momentum} can be written in terms of the strain rate tensor as $\bm{\tau}=2\eta\dot{\bm{\epsilon}}$, with
-$\dot{\bm{\epsilon}}=\frac{1}{2}\left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)$. Therefore, Eq. \ref{eq:momentum} can be rewritten as:
-\begin{equation}\label{eq:momentum_pressure}
--\bm{\nabla} p + \bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right) + \rho \bm{g} = 0
-\end{equation}
-Finally, using pressure from Eq. \ref{eq:mass_penalty}, Eq. \ref{eq:momentum_pressure} can be rewritten as:
-\begin{equation}\label{eq:momentum_penalty}
-\lambda \bm{\nabla} \left(\bm{\nabla} \cdot \bm{v}\right)+\bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right)+\rho\bm{g}=0
-\end{equation}
-The penalty method is been associated to the iterative Uzawa method, as described in detail in \citet{Dabrowski2008} and \citet{Thieulot2014}
-(see Section \ref{sec:poiseuille}). Results of benchmarks performed to verify the correctness in the implementation of Eq. \ref{eq:momentum_penalty} are
-shown in Sections \ref{sec:ist_sphere} and \ref{sec:rayleigh}.
-
-To support large viscosities variations the penalty parameter is related to the effective elemental viscosity by means of a dimensionless coefficient, so that
-$\lambda=\lambda_e(e)\eta_{eff}(e)$ \citep{Marotta2006,Dabrowski2008,Thieulot2014}. Benchmark of the falling block \citep{Gerya2003a,Gerya2010,Thieulot2011}
-is performed to verify that the code can correctly deal buoyancy driven flows with strong viscosity contrasts (Section \ref{sec:block}).
-
-\section{Sticky air and free surface}\label{sec:sticky}
-The Earth's surface can be treated by means of either the so-called sticky-air or a true free surface method, both of them implemented in the code. In the
-sticky-air method the surface is approximated with the introduction of a buoyant layer with a viscosity at least four orders of magnitude lower than the crust
-\citep{Schmeling2008,Crameri2012} and the interface between lithosphere and air is defined using a chain of passive markers that are advected as the Lagrangian
-markers. The correctness of the evolution of the markers chain is tested with the experiment of a 2D time-dependent Stokes sphere below a free surface and
-compared with results from ASPECT\footnote{\url{https://aspect.geodynamics.org/}} \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} (Section
-\ref{sec:sphere}). In the true free surface case the top boundary is assumed stress-free and velocities are not fixed. In this case topography variations are
-described by vertical deformations of the mesh that depend on the velocity field of the nodes that identify the free surface, while horizontal deformations are
-not taken into account. This procedure is known as the Arbitrary Lagrangian-Eulerian (ALE) method and its implementation follows the technique described in
-\citet{Thieulot2011}. However, although the implementation of a true free surface better reproduces laboratory experiment, extremely small time steps can be
-necessary to maintain stability \citep{Kaus2010a,Quinquis2011,Thieulot2014}. Therefore, the stability algorithm proposed by \citet{Kaus2010a} is implemented
-to avoid instabilities due to high density differences at the free surface when using too large time steps. The implementation of this algorithm are tested by
-performing the experiment described by \citet{Kaus2010a} (Section \ref{sec:stab}).
-
-The topography relaxation benchmark proposed by \citet{Crameri2012} is performed to verify that the code correctly recovers topography variations in case of
-both the sticky-air and the true free surface method (Section \ref{sec:crameri}). Finally, Section \ref{sec:subduction} show the results of the spontaneous
-subduction experiment described by \citet{Schmeling2008}.
-
-\section{Erosion and sedimentation}\label{sec:erosion}
-Surface processes at either the lithosphere-air interface or in correspondence of the free surface have been implemented by means of the software
-Fastscape\footnote{\url{https://fastscape.org/}} \citep{Braun2013,Cordonnier2019,Yuan2019a,Yuan2019b}. FastScapeLib is a set of routines that solve the stream
-power law (enriched by a sediment transport/deposition term) \citep{Yuan2019a}, hillslope diffusion and marine transport and deposition \citep{Yuan2019b}.
-The partial differential equation solved by FastScapeLib is:
-\begin{equation}
-\frac{\partial h}{ \partial t}=U-K_f A^{m_e}S^{n_e} + \frac{G}{A}\int_A \left( U-\frac{\partial h}{\partial t} \right) dA + K_d \bm{\nabla}^2 h 
-\end{equation}
-where $h$ is the topography, $U$ is the uplift, $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $S$ is the slope, $m_e$ and $n_e$ are
-parameters that should be chosen to constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}, $G$ is the dimensionless
-deposition/transport coefficient for the enriched stream power law and $K_d$ is the effective diffusivity. Terms $K_f A^{m_e}S^{n_e}$ and $K_d \bm{\nabla}^2 h$
-consider long-range processes, described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, and short-range processes,
-proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}.
-
-In case of sticky air, slope is determined by means of the markers chain, which is vertically corrected considering erosion and sedimentation rates. After the
-correction, continental markers above the markers chain are transformed in air markers and air markers below the markers chain become sediments. Similarly,
-slope is determined considering the nodes on the top boundary in case of true free surface and they are vertically corrected as for the markers chain. In this
-case, continental markers above the top boundary are simply deleted.
-%Surface processes at the lithosphere-air interface are implemented taking into account fluvial erosion (long-range processes), hillslope processes (short-range
-%processes) and failure of slopes exceeding a threshold angle $\theta$ \citep{Burov1997,Tucker1999,Sacek2012,Croissant2014,Ueda2015,Beucher2020}. Erosion rates
-%$\dot{E}_{lr}$ due to long-range processes can be described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{lr}=K_f A^{m_e} \bm{\nabla} h^{n_e}
-%\end{equation}
-%where $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $\bm{\nabla} h$ is the slope and $m_e$ and $n_e$ are parameters chosen in to
-%constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}. Erosion rates $\dot{E}_{sr}$ due to short-range processes are
-%implemented by assuming that transport is linearly proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{sr}=K_d \bm{\nabla}^2 h
-%\end{equation}
-%where $K_d$ is the effective diffusivity. In addition, the slope is modified if it still exceeds a threshold angle after long- and short-range processes
-%\citep{Ueda2015}.
-
-%Offshore sedimentation processes are implemented similarly to hillslope processes but with a depth-dependent coefficient \citep{Sacek2012,Beucher2020}.
-%The sedimentation rate $\dot{S}$ is calculated as
-%\begin{equation}
-%\dot{S}=K_s \bm{\nabla}^2 h
-%\end{equation}
-%with
-%\begin{equation}
-%K_s=K^*_s \exp\left(\frac{h-h_{sl}}{h_0}\right)
-%\end{equation}
-%where $K^*_s$ is a constant, $h_{sl}$ is the sea level and $h_0$ is a characteristic length. In case of sticky air, slope is determined by means of the markers
-%chain, which is vertically corrected considering erosion and sedimentation rates. After the correction, continental markers above the markers chain are
-%transformed in air markers and air markers below the markers chain become sediments. Similarly, slope is determined considering the nodes on the top boundary
-%in case of true free surface and they are vertically corrected as for the markers chain. In this case, continental markers above the top boundary are simply
-%deleted.
-
-\section{Non-linear rheologies}\label{sec:nonlinear}
-Non-linear rheologies are implemented combining viscous creep (dislocation and diffusion) and plastic yielding. For each marker, diffusion and dislocation
-viscosity ($\eta_{df|ds}$) can be determined as follows \citep{Karato1993,Warren2008b,Wang2016}:
-\begin{equation}\label{eq:viscous}
-\eta_{df|ds}=f_s \left(\frac{d^m}{A}\right)^{\frac{1}{n}} \left (\frac{2^{\frac{1-n}{n}}}{3^{\frac{n+1}{2n}}} \right)
-I_2^{\frac{(1-n)}{n}}\exp\left(\frac{Q+pV}{nRT}\right)
-\end{equation}
-where $f_s$ is a scaling factor used to represent lithologies that are stronger or weaker than the base set, $d$ is the grain size, $m$ is the grain size exponent,
-$n$ is the stress exponent, $A$ is the uniaxial pre-exponential factor, $I_2$ is the square root of the second invariant of the strain rate tensor, $Q$ is the
-activation energy, $p$ is the pressure, $V$ is the activation volume, $R$ is the gas constant and $T$ is the temperature.
-Pressures and temperatures are determined by the interpolation of the nodal parameters, while strain rates are calculated by means of the
-derivative of the velocities on the nodes. In case of diffusion creep $n=1$ and $m>0$, in case of dislocation creep $n>1$ and $m=0$. The viscous creep $\eta_{cp}$
-is then calculated as the harmonic average between $\eta_{df}$ and $\eta_{ds}$:
-\begin{equation}\label{eq:visco_creep}
-\eta_{cp}=\left(\frac{1}{\eta_{df}}+\frac{1}{\eta_{ds}}\right)^{-1}
-\end{equation}
-The implementation of the non-linear viscous creep viscosity depending on the strain rate is tested by means of the slab detachment benchmark
-(Section \ref{sec:slab}).
-
-Plastic yielding is implemented rescaling $\eta_{cp}$ in order to limit the stress below the yield stress $\sigma_y$ \citep{Thieulot2008,Thieulot2014,Glerum2018},
-obtaining
-\begin{equation}\label{eq:plastic}
-\eta_{pl}=\frac{\sigma_y}{2I_2}
-\end{equation}
-where $\eta_{pl}$ is the plastic viscosity and the yield stress is determined following the Drucker-Prager criterion, such as
-\begin{equation}\label{eq:yield}
-\sigma_y=C\;\cos(\phi)+p\;\sin(\phi)
-\end{equation}
-where $C$ is the cohesion and $\phi$ is the internal friction angle. In case of negative pressure it is imposed equal to 0, so that negative yield stress are
-excluded. The correctness of the non-linear solution in case of plastic viscosity with variable internal friction angle is verified performing indenter and
-brick experiments (Sections \ref{sec:indenter} and \ref{sec:brick}, respectively).
-
-Plastic yielding and viscous creep are then combined to obtain a viscoplastic viscosity $\eta_{vp}$ as follows, assuming that they are independent processes
-\citep{Karato2008,Glerum2018}:
-\begin{equation}\label{eq:viscoplastic}
-\eta_{vp}=\min(\eta_{cp},\eta_{pl})
-\end{equation}
-Finally, effective viscosity $\eta_{eff}$ is capped by the minimum and the maximum viscosity ($\eta_{min}$ and $\eta_{max}$, respectively) to avoid extremely
-low or high viscosity \citep{Glerum2018} as follows
-\begin{equation}\label{eq:effective}
-\eta_{eff}=\min(\max(\eta_{vp},\eta_{min}),\eta_{max})
-\end{equation}
-Viscous creep and plastic yielding are non-linear rheologies because of their dependence on the velocity field through pressure and strain rates. Therefore,
-the solution is determined by means of Picard-type iterations, until convergence of the velocity field \citep{Glerum2018}. The convergence is verified at each
-iteration $i$ via the nonlinear residual $\bm{\mathcal{R}}^i$ that can be determined as
-\begin{equation}\label{eq:convergence}
-\bm{\mathcal{R}}^i= \mathbb{K} (\eta_{eff}(\dot{\bm{\epsilon}}^{i-1},p^{i-1}))\cdot \bm{v}^{i-1} - \bm{f}^i
-\end{equation}
-%\begin{equation}\label{eq:convergence}
-%\chi=1-\frac{\langle (\bm{v}^i-\langle\bm{v}^i\rangle)\cdot (\bm{v}^{i+1}-\langle\bm{v}^{i+1}\rangle) \rangle}{\sqrt{\sigma^i\sigma^{i+1}}}
-%\end{equation}
-%Iterations are performed until either convergence ($\chi<tol$) or maximal number of non-linear iterations ($it_{max}$) is reached.
-where $\mathbb{K}$ is the Stiffness matrix and $\bm{f}$ is the right hand side vector \citep{Spiegelman2016,Glerum2018}.  The $L_2$-norm is extracted from
-$\bm{\mathcal{R}}^i$ and it is normalised as follows
-\begin{equation}\label{eq:convergence_res}
-\frac{||\bm{\mathcal{R}}^i||_2}{||\bm{\mathcal{R}}^0||_2}
-\end{equation}
-where $||\bm{\mathcal{R}}^0||_2$ is the $L_2$-norm of the first nonlinear iteration.
-Since often the normalised nonlinear residual drop very quickly over the first 2-3 iterations, also the $L_2$-norm of velocity and pressure residuals are
-calculated as
-\begin{equation}\label{eq:convergence_vel}
-\frac{||\Delta\bm{v}||_2}{||\bm{v}||_2}=\frac{||\bm{v}_i-\bm{v}_{i-1}||_2}{||\bm{v}_{i}||_2}
-\end{equation}
-and
-\begin{equation}\label{eq:convergence_press}
-  \frac{||\Delta p||_2}{||p||_2}=\frac{||p_i-p_{i-1}||_2}{||p_{i}||_2}
-\end{equation}
-respectively. Iterations are performed until either maximal number of non-linear iterations ($it_{max}$) is reached or all the normalised nonlinear residual, 
-the velocity residual and the pressure residual converge under a defined tolerance $tol$.
-
-\section{Healing and weakening}
+\section{Healing and weakening}\label{sec:healing}
 The temporal evolution of the accumulated strain $\epsilon$ has been implemented as in \citet{Fuchs2019,Fuchs2021}, as following
@@
 \citep{Warren2008a}.
-
-\section{The energy equation}\label{sec:en_eq}
-The stabilisation of the advection term of the energy equation (Eq. \ref{eq:energy}) needed to avoid possible oscillations in the thermal solution in the case
-when advection dominates over diffusion, is implemented by means of a streamline-upwind Petrov–Galerkin (SUPG) method, for which the advection term is modified
-follows the discussion in \citet{Thieulot2011} and \citet{Thieulot2014} (see Section \ref{sec:advection} for benchmark). The total internal heat production
-$H_{tot}$ in Eq. \ref{eq:energy} includes radiogenic heating $H_r$, shear heating $H_s$ and adiabatic heating $H_a$, where
-\begin{equation}\label{eq:shear}
-H_s=2\eta \dot{\bm{\epsilon}} : \dot{\bm{\epsilon}} = 2\eta(\dot{\epsilon}_{xx}^2+\dot{\epsilon}_{yy}^2+2\dot{\epsilon}_{xy}^2)
-\end{equation}
-and
-\begin{equation}\label{eq:adiabatic}
-H_a=T \alpha \frac{Dp}{Dt} \approx -\alpha T \rho g_y v
-\end{equation}
-The correctness of the implementation and the computation of shear and adiabatic heating is verified performing an exercise with an analytical solution
-(Section \ref{sec:simple_shear}) and Exercise 9.4 in \citet{Gerya2010} (Section \ref{sec:shear}). Furthermore, mantle convection and thin layer entrainment
-benchmarks (Sections \ref{sec:mantle} and \ref{sec:thin}, respectively) are performed to verify that the code solve correctly Eqs. \ref{eq:momentum_penalty}
-and \ref{eq:energy} in case of a temperature field characterised by an initial perturbation and with a temperature-dependent density. At last, the viscoplastic
-mantle convection benchmark is performed to check also the viscoplasticity in case of a initial perturbed temperature field (Section \ref{sec:visco_mantle}).
-
-\section{Phase transitions and hydration}\label{sec:phase}
-Variations of effective density, specific heat and coefficient of thermal expansion during the evolution of crustal and mantle materials at different
-pressure-temperature ($p$-$T$) conditions are computed by means of version 6.8.6 of Perple\_X software package \citep{Connolly2005}, similarly to the
-implementation described in \citet{Marotta2020}. By default, mineral assemblages and properties of each lithology (oceanic crust, lower and upper continental
-crust, sediments and dry or hydrated mantle) are calculated for temperatures from \SIrange{330}{1600}{\kelvin}, with increments of approximately \SI{4}{\kelvin},
-and pressures from \SIrange{0.1}{30}{\giga\pascal}, with increments of approximately \SI{0.03}{\giga\pascal}, for a total of almost 100000 points.
-
-The introduction of phase transitions produces two main effects that must be taken into account in numerical models: variations in density and release/absorption
-of latent heat \citep{Tackley2010}. Effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into
-account considering effective coefficient of thermal expansion \citep{Christensen1985,Zhong2015}. Similarly, effects of latent heat in the energy equation
-(eq \ref{eq:energy}) are taken into account by considering effective specific heat and coefficient of thermal expansion
-\citep{Christensen1985,Tackley2010,Zhong2015}. Effective density, specific heat and coefficient of thermal expansion are included in Perple\_X files and are
-assumed by each Lagrangian marker in according to its $p$-$T$ conditions.
-
-Hydration processes related to dehydration of subducting lithosphere \citep{Schmidt1998,Liu2007,Faccenda2009,Faccenda2010,Faccenda2014,Rosas2016} strongly
-influence the thermo-mechanics inside the mantle wedge, mainly because of both weakening effects on the mantle rheology and density variations in case of
-mantle serpentinization \citep{Gerya2002,Honda2003,Arcay2005,Roda2010,Regorda2017}. In the case that hydration is switched on, the amount of bound and free
-water is memorised by each marker, following the implementation of \citet{Quinquis2014}; hydration and dehydration processes are related to the amount of bound
-water of each marker and to the maximum amount of water it can transport, i.e. if the amount of bound water exceeds the maximum amount of water, the marker
-dehydrates and releases free water that can hydrates under-saturated markers. Maximum water content of each marker is determined as function of lithology and
-$p$-$T$ conditions and it is calculated using Perple\_X, in the same way than effective density, specific heat and coefficient of thermal expansion. Bound water
-is advected together with the markers, neglecting the effect of bound water diffusion, while free water simply migrates vertically and is not coupled to the
-solid-phase flow of the mantle wedge \citep{Arcay2005,Quinquis2014}. The correctness of the migration scheme of the free water and its absorption into bound
-water by under-saturated markers are verified performing the experiment described in \citet{Quinquis2014} (Sec. \ref{sec:quinquis}).
-
-The mantle viscosity weakening related to the amount of water has been extensively studied \citep{Chopra1981,Hirth2003} and it is implemented in according to
-\citet{Arcay2005} and \citet{Horiuchi2016}, as follows
-\begin{equation}\label{eq:wet_visc}
-\eta_{wet}=\eta_{dry}\left(\left[1-\frac{1}{f_v}\right]\exp\left(-\frac{[OH^-]}{[OH^-]_0}\right)+\frac{1}{f_v}\right)
-\end{equation}
-where $f_v$ is the viscosity reduction factor between dry and wet conditions, $[OH^-]$ is the water content and $[OH^-]_0$ is a reference water content set to
-620 ppm (0.062 wt.\%) \citep{Arcay2005}. By default, $f_v$ is set to 100, in according to \citet{Arcay2005} and \citet{Horiuchi2016}.
-
-\section{Mantle melting}\label{sec:melt}
-Melt fraction in the mantle depends on temperature ($T$, in \textdegree C), pressure ($p$, in \si{\giga\pascal}) and the water content in the melt ($X_{H_2O}$, in wt.\%).
-The determination of the melt fraction $M$ is implemented as explained by \citet{Katz2003}, according to the successive modification by \citet{Langmuir2006} and
-\citet{Kelley2010}, as follows
-\begin{eqnarray}\label{eq:melt_fr}
-M(p,T,X_{H_2O})&=&-T+T_{sol}+(x_t\cdot ln(p+y_t))\cdot X_{H_2O}-\nonumber\\
-&&+K_m\left(\frac{X_{H_2O}^{bulk}}{D_{H_2O}(1-X_{H_2O})+X_{H_2O}}\right)^{\gamma}
-\end{eqnarray}
-where $T_{sol}$ is the temperature of the dry solidus, $D_{H_2O}$ is the partition coefficient and $x_t\cdot log(p+y_t)$ indicates the pressure dependence of
-the melting curve \citep{Kelley2010}. $x_t$, $y_t$, $K_m$ and $\gamma$ are parameters chosen in according to \citet{Kelley2010}.
-However, $X_{H_2O}$ depends on the melt fraction as
-\begin{equation}\label{eq:melt_wt}
-X_{H_2O}(M)=\frac{X_{H_2O}^{bulk}}{D_{H_2O}+M(1-D_{H_2O})}
-\end{equation}
-and a numerical solution can be found using a root-finder method \citep{Katz2003,Wang2016}. Furthermore, the content of water in the melt is limited by the
-pressure-dependent saturation concentration of water in the melt, determined as
-\begin{equation}\label{eq:melt_max}
-X_{H_2O}^{sat}=\chi_1 p^{\lambda_m}+\chi_2 p
-\end{equation}
-where $\chi_1$, $\chi_2$ and $\lambda_m$ are parameters chosen according to \citet{Katz2003}, \citet{Langmuir2006} and \citet{Kelley2010}. The correctness in
-the determination of melt fractions and water contents in the melt are tested comparing model results with experimental curves obtained by \citet{Katz2003}
-(Sec. \ref{sec:katz}).
-
-In case of melting, viscous creep viscosity ($\eta_{vp}$) is modified according to \citet{Wang2016}, as
-\begin{equation}\label{eq:melt_visc}
-\eta_{melt}=\eta_{vp}\cdot \exp(\alpha_m M)
-\end{equation}
-where $\alpha_m$ are melt fraction factors for dislocation and diffusion creep, chosen in according to \citet{Wang2016}.
-
-Effective density ($\rho_{eff}$) for partially molten rocks is calculated as
-\begin{equation}\label{eq:melt_dens}
-\rho_{eff}=\rho_s(1-M)+\rho_m M
-\end{equation}
-where $\rho_s$ and $\rho_m$ are the densities of the solid and the molten rock, respectively \citep{Gerya2004b,Gerya2007,Wang2016}. The density of the
-solid rock is extracted by Perple\_X, while the density of the molten portion is calculated as
-\begin{equation}\label{eq:melt_dens_M}
-\rho_m(p,T)=\rho_0[1-\alpha(T-T_0)][1+\beta(p-p_0)]
-\end{equation}
-where $\rho_0$ is the density at the reference temperature ($T_0$) and pressure($p_0$), and $\alpha$ and $\beta$ are the thermal expansion and compressibility
-coefficients, respectively \citep{Gerya2004b,Gerya2007,Wang2016}.
-
-As for the phase transitions, effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into account
-considering effective coefficient of thermal expansion. Similarly, latent heat due to melting/crystallisation can be implicitly included in the energy equation
-(Eq. \ref{eq:energy}) by considering effective specific heat ($C_{p_{eff}}$) and thermal expansion ($\alpha_{eff}$) of the partially molten rocks, in according to
-\begin{eqnarray}
-\label{eq:cp_effective}C_{p_{eff}}&=&C_p+H_L\left(\frac{\partial M}{\partial T}\right)_p \\
-\label{eq:alpha_effective}\alpha_{eff}&=&\alpha+\rho\frac{H_L}{T}\left(\frac{\partial M}{\partial p}\right)_T
-\end{eqnarray}
-where $C_p$ and $\alpha$ are the specific heat and the coefficient of thermal expansion of the solid rock, respectively, and $H_L$ is the latent heat of the
-molten rock \citep{Gerya2004b,Gerya2007,Tackley2010}.
-
-\section{Benchmarks}\label{sec:benchmark}
-\subsection{Error measurements}\label{sec:error}
-In order to determine the accuracy of velocity and pressure field of the benchmarks, the $L_2$-norm is computed by numerical integration on the Gauss-Legendre
-quadrature points. $L_2$-norm for pressure and velocity errors can be evaluated as
-\begin{eqnarray}
-\textrm{err}_p&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|p^n_i(r_q)-p^a_i(r_q)|^2\right)w_q|J_q|}\nonumber \\
-\textrm{err}_v&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|u^n_i(r_q)-u^a_i(r_q)|^2+|v^n_i(r_q)-v^a_i(r_q)|^2\right)w_q|J_q|}\nonumber
-\end{eqnarray}
-respectively, where $n_e$ is the number of elements, $n_q$ is the number of quadrature points per element, $p^n_i(r_q)$ and $p^a_i(r_q)$ are the numerical and
-analytical pressure, respectively, in each quadrature point $q$, $w_q$  and $J_q$ are the weight and the Jacobian at the quadrature point
-$q$, $u^n_i(r_q)$, $v^n_i(r_q)$, $u^a_i(r_q)$ and $v^a_i(r_q)$ are the numerical and analytical velocities, respectively, in each quadrature point $q$.
-
-Other quantities used as comparison with original benchmarks are the root-mean-square velocity over the whole domain and over the surface
-\begin{eqnarray}
-v_{\textrm{rms}}&=&\sqrt{\int_{\Omega}|\bm{v}|^2 d\Omega}\nonumber \\
-v_{\textrm{rms}}^{top}&=&\sqrt{\int_0^1 u^2 \bigg|_{y=1} dx}\nonumber
-\end{eqnarray}
-respectively, top and bottom Nusselt numbers
-\[\textrm{Nu}^{top/bottom}=-\int_0^1\frac{\partial T}{\partial y} \bigg|_{y=1/y=0} dx\]
-the average rate of work done against gravity
-\[\left\langle W \right\rangle=\int_{\Omega}T u_y d\Omega\]
-the average rate of viscous dissipation
-\[\left\langle \Phi \right\rangle=\int_{\Omega}\tau_{ij}\dot{\epsilon}_{ij} d\Omega\]
-and the percentage error between $\left\langle W \right\rangle$ and $\left\langle \Phi \right\rangle/Ra$
-\[\delta=\frac{\left|\left\langle W \right\rangle -\frac{\left\langle \Phi \right\rangle}{Ra} \right|}{\max \left(\left\langle W \right\rangle,
-\frac{\left\langle \Phi \right\rangle}{Ra}\right)}\times100\]
-
-\subsection{Stokes flow}\label{sec:stokes}
-The problem consists of determining velocity field $(u,v)$ and pressure $p$ in case of a manufactured solution with prescribed body forces such as
-\begin{eqnarray}
-b_1&=&(12-24y)x^4+(-24+48y)x^3+(-48y+72y^2-48y^3+12)x^2+\nonumber \\
-&&+(-2+24y-72y^2+48y^3)x+1-4y+12y^2-8y^3\nonumber \\
-b_2&=&(8-48y+48y^2)x^3+(-12+72y-72y^2)x^2+(4-24y+48y^2-48y^3+\nonumber \\
-&&+24y^4)x-12y^2+24y^3-12y^4\nonumber
-\end{eqnarray}
-for which the exact solution is:
-\begin{eqnarray}
-u(x,y)&=&x^2(1-x)^2(2y-6y^2+4y^3)\nonumber \\
-v(x,y)&=&-y^2(1-y)^2(2x-6x^2+4x^3)\nonumber \\
-p(x,y)&=&x(1-x)-\frac{1}{6}\nonumber
-\end{eqnarray}
-The domain is a unit square a constant viscosity ($\eta=1$) and the penalty parameter is set to $\lambda=10^7$. Velocity boundary conditions are set to no slip
-$(\bm{v}=\bm{0})$ on all boundaries. The problem is performed for different grid resolution between \num{8x8} and \num{1024x1024} elements.
-
-The errors between the analytical solution and the numerical prediction of the velocity and the pressure field are calculated as in Section \ref{sec:error}.
-Fig. \ref{fig:errors}a shows that both velocity and pressure field converge to the exact solution with the decrease of the element size, following the
-theoretical convergence rate. The convergence of pressure, in contrast with the observation by \citet{Donea2003} for $Q_1 \times P_0$ elements, indicate the
-effectiveness of the smoothing procedure. Smoothed pressure field for grid resolution of \num{128x128} elements is shown in Fig. \ref{fig:errors}b (continuous
-red line), in comparison with the analytical solution (dashed blue line). Solve times and memory usage needed to generate the solution are shown in
-Fig. \ref{fig:MUMPS}. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Solver/Stokes_Flow}.
-
-\begin{figure}
-\centering
-\includegraphics[width=\linewidth]{./Figures/errors.pdf}
-\caption{Velocity and pressure error for the Stokes flow experiment between generated and analytical solution as a function of element size (panel a),
-and comparison between smoothed pressure and analytical solution as function of $x$ coordinate for a grid resolution of \num{128x128} elements (panel b).}
-\label{fig:errors}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\linewidth]{./Figures/MUMPS.pdf}
-\caption{Panel a) Analysis, factorisation, solution and total solve times (panel a) and factorisation memory usage (panel b) as a function of the total number
-of degrees of freedom for the Stokes flow experiment.}
-\label{fig:MUMPS}
-\end{figure}
-
-\subsection{Markers advection}\label{sec:runge}
-The 2nd-order Runge-Kutta advection scheme in space is tested by means of the Zalesak disk test \citep{Zalesak1979,Thieulot2014}. The benchmark is performed
-in a unit square domain with a grid resolution of \num{32x32} elements and 1000000 markers and values of Courant number between 0.25 and 1.
-At $t=0$, the disk is centred at position $(0.5;0.75)$ with a radius $R=0.15$ and has a vertical fissure 0.05 wide and 0.2 high (orange in Fig. \ref{fig:runge})
-and with the velocity field prescribed in the entire domain as
-\begin{eqnarray}
-u(x,y)&=&2\pi\left(y-\frac{L_x}{2}\right)\nonumber \\
-v(x,y)&=&-2\pi\left(x-\frac{L_x}{2}\right)\nonumber
-\end{eqnarray}
-the markers are back to their initial location after a $2\pi$ rotation ($t=1$). The clockwise rotation of the disk at $t=0.25$, 0.5 and 0.75 
-is shown in Fig. \ref{fig:runge} (red, blue and green, respectively). The distance from the centre of 1 marker is calculated throughout the rotation to
-evaluate the error for different values of Courant number. As expected, the error increase with the increasing of the time step, as shown in 
-Fig. \ref{fig:runge_err}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/Zalesak_Disk}.
-
-\begin{figure}
-\centering
-\includegraphics[width=200px]{./Figures/Zalesak.pdf}
-\caption{Position of the Zalesak disk at $t=0$, 0.25, 0.5 and 0.75 (orange, red, blue and green, respectively) for the case with Courant number of 1.}
-\label{fig:runge}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/zal_marker.pdf}
-\caption{Distance from the centre as function of time for values of Courant number of  of 0.25, 0.3, 0.5, 0.75 and 1 (green, orange, blue, black and red,
-respectively). R indicates the distance from the centre considering a perfect circle trajectory.}
-\label{fig:runge_err}
-\end{figure}
-
-\subsection{Conservative Velocity Interpolation (CVI)}\label{sec:CVI}
-The CVI correction is checked by means of the Stokes flow experiment by \citet{Donea2003}, as presented in \ref{sec:stokes}. The advection of Lagrangian
-markers is performed using either a 2nd-order or a 4th-order Runge-Kutta scheme with CFL=0.5 and an initial random distribution of 25 markers per element.
-Fig. \ref{fig:CVI} shows the minimum and maximum values of markers per element throughout the experiment with and without the CVI correction and using
-different Runge-Kutta schemes. The introduction of the CVI correction allows the
-tests to run for times longer than $t=2000$ without show empty elements, while tests without the CVI correction stop at approximately $t=600$, when at least
-one element goes empty. Fig. \ref{fig:CVI_time} shows the different number of markers per element and the markers distribution at $t=600$ obtained without and
-with the CVI correction (panels b and c, respectively) in case of a 2nd-order Runge-Kutta scheme. All data can be found at
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/CVI}.
-\begin{figure}
-\centering
-\includegraphics[width=\linewidth]{./Figures/CVI.pdf}
-\caption{Maximum and minimum number of markers per element throughout the experiment, with (green and orange lines) or without (purple and blue lines)
-the CVI correction and using either a 2nd-order (blue and orange lines) or a 4th-order (purple and green lines) Runge-Kutta scheme.}
-\label{fig:CVI}
-\end{figure}
-\begin{figure}
-\centering
-\includegraphics[width=\linewidth]{./Figures/CVI_time.pdf}
-\caption{Number of markers per element and markers distribution through time using 2nd-order Runge-Kutta scheme.
-Panels a: number of markers per element (panel a1) and distribution of the markers (panel a2) after the first
-iteration; panels b and c: comparison between number of markers per element (panels b1 and c1) and markers distribution (panels b2 and c2) without and with
-the CVI correction at $t=600$; panels d: number of markers per element (panel d1) and markers distribution (panel d2) in case of CVI correction at $t=2000$.}
-\label{fig:CVI_time}
-\end{figure}
-
-\subsection{Poiseuille flow}\label{sec:poiseuille}
-The domain is a rectangle with $L_x=2$ and $L_y=1$ and constant density and viscosity ($\rho=1$ and $\eta=1$), gravity acceleration $\bm{g}=\bm{0}$ and penalty
-parameter $\lambda=10^8$. The grid is composed by \num{40x20} elements. Velocity boundary conditions are set to no slip $(\bm{v}=\bm{0})$ at the top and the
-bottom, and a parabolic profile is imposed on the sides, with $u=y(1-y)$ and $v=0$. The analytical solution is then given by:
-\begin{eqnarray}
-u(x,y)&=&y(1-y)\nonumber \\
-v(x,y,)&=&0\nonumber \\
-p(x,y)&=&2\eta\left(\frac{L_x}{2}-x\right)\nonumber
-\end{eqnarray}
-
-The velocity field predicted by the model follows the expected parabolic profile (Fig. \ref{fig:poiseuille}a and Fig. \ref{fig:poi_plot}a) and, in case of the
-classic penalty method with no iterations, the pressure is clearly related to the divergence by means of the penalty parameter (Fig. \ref{fig:poiseuille}b and
-c and Fig. \ref{fig:poi_plot}b). Fig. \ref{fig:poiseuille}d shows that one Uzawa iteration is sufficient to bring the divergence down to \num{e-15}, with no
-correlation with the pressure field. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Poiseuille%20Flow}.
-
-\begin{figure}
-\includegraphics[width=\linewidth]{./Figures/Poiseuille.pdf}
-\caption{Velocity field (panel a), pressure (panel b) and divergence velocity of a Poiseuille flow in case of the classic penalty method (no iterations) and
-after one Uzawa iteration (panel c and d, respectively).}
-\label{fig:poiseuille}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\linewidth]{./Figures/Analytical.pdf}
-\caption{Velocity (panel a) and pressure (panel b) field predicted by the model for a Poiseuille flow with respect with their analytical solutions. The
-velocity field is plotted as function of the vertical coordinate in $L_x/2$ and the pressure is plotted as function of the x coordinate in $L_y/2$.}
-\label{fig:poi_plot}
-\end{figure}
-
-\subsection{Instantaneous 2D sphere}\label{sec:ist_sphere}
-The domain is a unit square with gravity $\bm{g}=(0,-1)$. The fluid has constant density and viscosity ($\rho_f=1$ and $\eta_f=1$). The sphere is in the middle
-of the domain with a radius R = 0.123456798 and has constant density and viscosity ($\rho_s=10^{-2}+\rho_f$ and $\eta_s=10^3 \cdot \eta_f$).
-We distinguish three different types of velocity boundary conditions:
-\begin{enumerate}
-\item FS: free slip conditions on all sides.
-\item NS: no slip conditions on all sides.
-\item OT: free slip conditions on the sides and the bottom, and open boundary at the top.
-\end{enumerate}
-All three cases are performed for different grid resolutions, between \num{16x16} and \num{512x512} elements with 50 randomly distributed markers per element,
-and different average schemes for the viscosity (harmonic, geometric and arithmetic). The velocity in the centre of the sphere $\bm{v}(0.5;0.5)$), minimum and
-maximum velocities ($u_{min}$, $u_{max}$, $v_{min}$, $v_{max}$) and pressures ($p_{min}$, $p_{max}$) on the entire domain, $v_{\textrm{rms}}$ and average pressure on the
-entire domain ($p_{avg}$) are compared with solutions generated by ASPECT \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and all results
-can be found at \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere2D} and 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Instantaneous_2D_Sphere}. Results in terms of the $v_{\textrm{rms}}$ for NS model 
-are also shown in Fig. \ref{fig:inst_sphere}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/vrms_NS.pdf}
-\caption{$v_{\textrm{rms}}$ as function of element size for the instantaneous 2D sphere experiment in case of no slip boundary conditions and with different average
-schemes for the viscosity. Results are compared with results obtained by other numerical codes.}
-\label{fig:inst_sphere}
-\end{figure}
-
-\subsection{Rayleigh-Taylor instability}\label{sec:rayleigh}
-This problem has been originally presented by \citet{vanKeken1997} and here it is performed as the isoviscous case (Case 1a). The domain has $L_x=0.9142$ and
-$L_y=1$ and gravity $\bm{g}=(0,-1)$. Two fluids with same constant viscosities ($\eta_1=\eta_2=1$) and different densities ($\rho_1=1000$ and $\rho_2=1010$),
-with the lighter fluid at the bottom. The initial interface between the fluids is given by $y(x)=0.2+0.02 \cos \left(\frac{\pi x}{L_x}\right)$. The experiment
-is performed with different grid sizes (\num{50x50}, \num{80x80}, \num{100x100} and \num{256x256}). A total of 1960000 markers are randomly distributed at the
-beginning of the simulation. Velocity boundary conditions are set to no slip at the top and the bottom, and to free slip at the sides of the domain.
-
-$v_{\textrm{rms}}$ as function of time is reported in Fig. \ref{fig:RT}, matching well with results shown by \citet{vanKeken1997}, \citet{Tackley2003} and
-\citet{Thieulot2014}. Fig. \ref{fig:rayleigh} shows the evolution of the experiment at different time steps. All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Rayleigh_Taylor_experiment/ISOVISCOUS}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/RT.pdf}
-\caption{$v_{\textrm{rms}}$ of the Rayleigh-Taylor experiment as function of time for different resolution of the grid.}
-\label{fig:RT}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=300px]{./Figures/Rayleigh.pdf}
-\caption{Evolution of the Rayleigh-Taylor experiment for a grid of \num{256x256} elements at $t=0, 500, 1000$ and $2000$ (panels a, b, c and d, respectively).}
-\label{fig:rayleigh}
-\end{figure}
-
-\subsection{Falling blocks}\label{sec:block}
-This benchmark is described as proposed by \citet{Gerya2003a}, \citet{Gerya2010} and \citet{Thieulot2011}. The domain is square with $L_x=L_y=$ \SI{500}{\km}
-and the grid is composed by \num{50x50} elements with 25 markers in each element. The block is initially centred at ($x=$ \SI{250}{\km}; $y=$ \SI{400}{\km})
-and has a size of \num{100x100} km (Fig. \ref{fig:block}a). The fluid surrounding the block has $\eta_f=$ \SI{e21}{\pascal\per\s} and $\rho_f=$
-\SI{3200}{\kg\per\cubic\m}. The benchmark tests with different viscosities of the block, with $\eta_b$ from \SIrange{e15}{e27}{\pascal\per\s}. In each
-experiment also the density $\rho_b$ of the block varies from \SIrange{3220}{9900}{\kg\per\cubic\m}. Velocity boundary conditions are set to free slip
-conditions on all sides of the domain. Density distributions at $t=$ \SI{20}{\mega\year} for $\rho_b=$ \SI{3300}{\kg\per\cubic\m} and different $\eta_b$
-are plotted in Fig. \ref{fig:block}b-f, showing that the code correctly preserve the block geometry in case of large viscosity contrast, with a block stiffer
-than the surrounding fluid (Fig. \ref{fig:block}f).
-
-The velocity in the centre of the falling block at $t=0$ is measured for all experiments and, since the velocity should increase with density contrast, the
-quantity $\bm{v}/(\rho_b-\rho_f)$ is plotted as function of the viscosity contrast. All results perfectly match with those from \citet{Gerya2010} and they
-line up on a single curve, demonstrating that the code can correctly deal with large viscosity and density contrasts (Fig. \ref{fig:falling}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Falling%20blocks}.
-
-\begin{figure}
-\includegraphics[width=300px]{./Figures/Block.pdf}
-\centering
-\caption{Density evolution of the falling block experiment at $t=$ \SI{0}{\mega\year} (panel a) and $t=$ \SI{20}{\mega\year} for different viscosities of the
-block (panels b-f).}
-\label{fig:block}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Falling.pdf}
-\caption{Initial velocity relative to the density contrast at the centre of the falling block as function of the viscosity contrast between the block and the
-surrounding fluid.}
-\label{fig:falling}
-\end{figure}
-
-\subsection{2D time-dependent Stokes sphere with deformable free surface}\label{sec:sphere}
-This experiment is performed in a unit square domain with the gravity acceleration fixed to $g_y=-1$. The sphere has $\rho_s=2$ and $\eta_s=10^3$ and is
-initially centred at ($0.5;0.6$) with radius $R=0.123456789$. The fluid surrounding the sphere has $\rho_f=1$ and $\eta_f=1$ and occupies the domain for
-$y\leq 0.75$, while for $y>0.75$ the air has $\rho_a=0$ and $\eta_a=10^{-3}$. Velocity boundary conditions are set to free slip on all sides. The Courant
-number is set to 0.25. The experiments run for 200 s using grid resolutions from \num{150x150} to \num{512x512} elements, with 25 randomly distributed markers
-per element, and for different average schemes for the viscosity. The interface between the fluid and the air is tracked by means of the markers chain. Results
-in terms of velocity and pressure fields and topography variations are compared with results obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and they can be found at
-\url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere_fs2D} and
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Time_dependent_sphere}. Results in terms of the $v_{\textrm{rms}}$ in case of 
-an arithmetic average are also shown in Fig. \ref{fig:inst_spherefs}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/vrms_arithm.pdf}
-\caption{$v_{\textrm{rms}}$ for different grid resolutions as function of time for the 2D time-dependent Stokes sphere experiment in case of an arithmetic average.
-Results are compared with results obtained by ASPECT.}
-\label{fig:inst_spherefs}
-\end{figure}
-
-\subsection{Free surface stabilisation}\label{sec:stab}
-The benchmark is performed as discussed in \citet{Kaus2010a} and \citet{Thieulot2014}. The domain is square with $L_x=L_y=$ \SI{500}{\km} and a grid resolution
-of \num{200x200} elements, each of them containing 16 markers. A buoyant fluid with $\rho_1=$ \SI{3200}{\kg\per\cubic\metre} and $\eta_1=$ \SI{e20}{\pascal\s}
-is overlain by a denser fluid with $\rho_2=$ \SI{3300}{\kg\per\cubic\metre} and $\eta_2=$ \SI{e21}{\pascal\s}. The initial interface between the fluids has an
-initial sinusoidal shape of 5 km amplitude. Velocity boundary conditions are set to free slip conditions at the sides and no slip at the bottom, while the top
-is free surface. The experiment is performed with various fixed time steps, without the use of the Courant number. The vertical position of the free surface at
-$x=L_x$ is tracked for each simulation, with or without the stabilisation algorithm.
-
-The results show an instability (drunken sailor effect) increasing the time step in case that the stabilisation algorithm is not activated
-($dt=$ \SI{4200}{\year}, green line in Fig. \ref{fig:stab}), as already observed by \citet{Kaus2010a} and \citet{Thieulot2014}. Activating the algorithm the
-instability is fixed and the simulation remains stable using time step up to \SI{50000}{\year} (red line in Fig. \ref{fig:stab}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Stabilisation_algorithm}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Kaus_random.pdf}
-\caption{Evolution of the y coordinate in $x=L_x$ as function of time for different time steps, with and without the stabilisation algorithm.}
-\label{fig:stab}
-\end{figure}
-
-\subsection{Topography relaxation}\label{sec:crameri}
-This experiment is performed as in \citet{Crameri2012} using free surface and sticky air (Case 1). The domain is rectangular with $L_x=$ \SI{2800}{\km} and
-$L_y=$ \SI{700}{\km} in case of free surface and $L_y=$ \SI{800}{\km} in case of sticky air ($h_{st}=$ \SI{100}{\km}). The grid is composed by \num{256x64}
-elements for the free surface \num{560x320} elements for the sticky air. For both cases, there is a mantle of 600 km thickness with $\eta_m=$ \SI{e21}{\pascal\s}
-overlain by a highly viscous cosine-shaped lithosphere with thickness between 93 and 107 km and $\eta_l=$ \SI{e23}{\pascal\s}. Both mediums have $\rho=$
-\SI{3300}{\kg\per\cubic\metre}. In case of sticky air experiment, the air has $\rho_a=$ \SI{0}{\kg\per\cubic\metre} and $\eta_a=$ \SI{e18}{\pascal\s}. The
-gravity acceleration is set to $g_y=$ \SI{-10}{\metre\per\square\s}. Velocity boundary conditions are set to no slip at the bottom and free slip at the sides
-of the domain. In case of sticky air, velocities on top are set to free slip conditions.
-
-The maximum topography as function of time ($h(t)$) can be analytically derived in according to \[h(t)=h_0 \exp(\gamma t)\]
-where $h_0=7$ km is the initial topography, $\gamma=\num{-0.2139e-11}$ is the characteristic relaxation rate and $t$ is the time (black line in Fig.
-\ref{fig:crameri}). Maximum topography at the characteristic relaxation time $t_{rlx}=$ \SI{14.825}{\kilo\year} can be found to be $h_{rlx}=$ \SI{2576}{\m}.
-The results show that a viscosity of the air of \SI{e18}{\pascal\s} is needed in case of $h_{st}=$ \SI{100}{\km} to correctly describe the topography relaxation
-for this problem, as pointed out by \citet{Crameri2012} (purple line in Fig. \ref{fig:crameri}). The case with a true free surface well follows the analytical
-solution (orange line in Fig. \ref{fig:crameri}). In particular, at the relaxation time in case of the true free surface predicts a maximum topography of
-\SI{2570}{\m}, with an error of \SI{6}{\m} with respect to the analytical solution. Results are also compared with results obtained with a free surface in
-UNDERWORLD (\num{256x64} elements) and SULEC (\num{401x201} elements) (with courtesy from \citealp{Crameri2012}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Topography_relaxation}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Crameri.pdf}
-\caption{Maximum topography as function of time for the Crameri benchmark (coloured lines) in comparison with the analytical solution (black line) and results
-obtained with UNDERWORLD (yellow line) and SULEC (blue line).}
-\label{fig:crameri}
-\end{figure}
-
-\subsection{Spontaneous subduction}\label{sec:subduction}
-This experiment is performed as presented by \citet{Schmeling2008}, in case of both a sticky air and a true free surface and using different average schemes
-for the viscosities. The domain is rectangular with $L_x=$ \SI{3000}{\km} and $L_y=$ \SI{700}{\km}, a grid resolution of \num{256x64} elements with 50 markers
-per element and Courant number of 0.01. At the beginning of the simulation, a 100 km-thick lithospheric layer with $\rho_m=$ \SI{3300}{\kg\per\cubic\metre}
-and $\eta_m=$ \SI{e23}{\pascal\s} is located at the top of the domain for $x$ between 1000 and 3000 km, while the asthenospheric mantle has $\rho_m=$
-\SI{3200}{\pascal\s} and $\eta_m=$ \SI{e21}{\pascal\s}. In addition, a 200 km-depth lithospheric slab is already subducted in the mantle in order to have
-a spontaneous subduction. In case of sticky air, $L_y=750$ km and a 50 km-thick air layer with $\eta_a$ from \SIrange{e19}{e21}{\pascal\s} overlie the mantle.
-Velocity boundary conditions are set to free slip on the sides and on the bottom of the domain. In case of sticky air, velocities on the top boundary are set
-to free slip conditions as well.
-
-The maximum depth of the slab is tracked for all the simulations (Fig. \ref{fig:subduction}) and they are compared with results obtained by
-\citet{Schmeling2008} with models with regular grids and comparable grid resolutions (rectangles in Fig. \ref{fig:subduction}). Low viscous sticky air models
-(i.e., \SI{e19}{\pascal\s}) enlighten a strong dependence of the sinking velocity with the chosen average scheme (dashed lines in Fig. \ref{fig:subduction}),
-while high viscous sticky air models (i.e., \SI{e21}{\pascal\s}) show a higher resistance at the trench, underestimating the correct solution (dotted lines in
-Fig. \ref{fig:subduction}). A high resistance at the trench can be also observed in case of a true free surface for the low resolution grid used (continuous
-lines in Fig. \ref{fig:subduction}), as pointed out by \citet{Schmeling2008}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Spontaneous_subduction}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Subduction.pdf}
-\caption{Maximum depth of the slab as function of time for all the simulations of the \citet{Schmeling2008} benchmark, in case of harmonic, geometric and
-arithmetic means (red, green and blue lines, respectively) using a sticky air or a true free surface (discontinuous and continuous lines, respectively).
-Red, green and blue rectangular areas indicate the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{500}{\km} in case of sticky air,
-while the grey rectangular area indicates the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{600}{\km} in case of true free surface.}
-\label{fig:subduction}
-\end{figure}
-
-\subsection{The slab detachment}\label{sec:slab}
-The slab detachment benchmark is performed as described by \citet{Schmalholz2011} and \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{1000}{\km}
-and $L_y=$ \SI{660}{\km} and a grid resolution of \num{256x256} elements. A non-linear viscous T-shaped layer with $\rho_l=$ \SI{3300}{\kg\per\cubic\metre} is
-placed at the top of the domain and surrounded by a linear viscous fluid with $\rho_f=$ \SI{3150}{\kg\per\cubic\m}. The top layer is 80 km-thick and a 250
-km-long and 80 km-wide slab is placed at $x=L_x/2$ (Fig. \ref{fig:slab}a). The effective viscosity of the top layer is given by
-\[\eta_{eff}=\eta_0 I_2^{\frac{(1-n)}{n}}\]
-with $\eta_0=$ \SI{4.75e11}{\pascal\s} and $n=4$, while the surrounding fluid has $\eta_f=$ \SI{e21}{\pascal\s}. The effective viscosity is capped between
-$\eta_{min}=$ \SI{e21}{\pascal\s} and $\eta_{max}=$ \SI{e25}{\pascal\s}. Velocity boundary conditions are set to free slip at the top and the bottom, and to
-no slip at the sides of the domain.
-
-The sinking of the slab determines a decrease of the effective viscosity in the T-shaped layer during the first part of the evolution (Fig. \ref{fig:slab}a),
-as consequence of local strain rates, while effective viscosities increase up to $\eta_{max}$ in the top layer after approximately \SI{20}{\mega\year}, when
-necking is complete (Fig. \ref{fig:slab}b).
-
-The width of necking is tracked by means of markers position at the side of the slab. Necking width and time are normalised in function of the initial slab
-(\SI{80}{\km}) and the characteristic time ($t_c=$ \SI{7.1158e14}{\s}) \citep{Schmalholz2011,Glerum2018}. The results in case of arithmetic, geometric and
-harmonic mean of the viscosity are shown in Fig. \ref{fig:necking}, compared with results from \citet{Glerum2018} for which an infinity norm mean was used.
-The influence of the average schemes agrees with results shown in Section \ref{sec:subduction}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Slab_detachment}.
-
-\begin{figure}
-\includegraphics[width=\linewidth]{./Figures/Slab.pdf}
-\caption{Effective viscosity for the slab detachment benchmark at the beginning of the evolution (panel a) and when necking is complete (panel b).}
-\label{fig:slab}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Necking.pdf}
-\caption{Normalised width of the necking of the slab detachment benchmark as function of normalised time.}
-\label{fig:necking}
-\end{figure}
-
-\subsection{The indenter experiment}\label{sec:indenter}
-The indenter benchmark simulate a rigid punch on a purely plastic von Mises material, for which exists an analytical solution
-\citep{Thieulot2008,Thieulot2014,Glerum2018}. The domain is a unit square with a grid resolution of \num{256x256} elements. The gravity acceleration is set
-to $\bm{g}=\bm{0}$ and effective viscosity is capped using $\eta_{min}=10^{-4}$ and $\eta_{max}=10^3$. The tolerance for the convergence of the non-linear
-solution is $tol=10^{-9}$, with a maximum number of non-linear iterations set to 500. The medium in the domain has $\rho=0.01$, an initial viscosity
-$\eta_0=10$, cohesion $C=1$ and an angle of internal friction $\phi=0$°. Velocity boundary conditions are set to no slip at the bottom and free slip at the
-sides of the domain. The top of the domain is open with the exception of the central portion (punch area) with width $w_p=0.125$, where $v=-1.05$ and $u$ is
-fixed either to 0 (rough punch experiment) or free (smooth punch experiment).
-
-Fig. \ref{fig:indenter} show the results in terms of viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure
-(panels g and h) for both experiments. The analytical solution indicate that pressure at ($0.5;1$) and ($0.5 \pm w_p;1$)is $p=\pi +1$ and $p=1$, respectively
-(black and red lines in Fig. \ref{fig:smooth_p}). Fig. \ref{fig:smooth_p} shows a clear improvement in the pressure solution when passing from a rough to a
-smooth punch (green and blue lines in Fig. \ref{fig:smooth_p}, respectively), as previously observed by \citet{Thieulot2014} and \citet{Glerum2018}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Indenter}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Indenter.pdf}
-\caption{Viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure (panels g and h) fields for rough (left column) and
-smooth (right column) punch experiments.}
-\label{fig:indenter}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Smooth_Rough.pdf}
-\caption{Nodal pressure in function of the x coordinate at $y=1$ for rough (green line) and smooth (blue line) punch experiment. Black and red lines indicate
-the analytical solution at $x=0.5$ and $x=0.5 \pm w_p$, respectively.}
-\label{fig:smooth_p}
-\end{figure}
-
-\subsection{The brick experiment}\label{sec:brick}
-An instantaneous version of the brick benchmark is performed to verify the correctness of the pressure-dependent plasticity for different angle of internal
-friction, as proposed by \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{40}{\km} and $L_y=$ \SI{10}{\km} and a grid resolution of \num{512x128}
-elements. The gravity acceleration is set to $g_y=$ \SI{-10}{\m\square\s} and effective viscosity is capped using $\eta_{min}=$ \SI{e19}{\pascal\s} and
-$\eta_{max}=$ \SI{e26}{\pascal\s}. The tolerance for the convergence of the non-linear solution is $tol=10^{-7}$, with a maximum number of non-linear iterations
-set to 1000. A 800 m-wide and 400 m-height inclusion with $\eta_b=$ \SI{e20}{\pascal\s} and $\rho_b=$ \SI{2700}{\kg\per\cubic\m} is placed at the bottom of the
-domain at $x=L_x/2$. The inclusion is surrounded by a non-linear viscous medium with $\rho_m=$ \SI{2700}{\kg\per\cubic\m}, an initial viscosity of
-\SI{e23}{\pascal\s}, a linear viscous viscosity of \SI{e25}{\pascal\s} and a cohesion of \SI{40}{\mega\pascal}. Velocities are set to free slip conditions at
-the bottom of the domain and the top is open. Velocities on sides of the domain are fixed to $u=$ \SI{+-2e-11}{\m\per\s} and $v=0$. The experiment is performed
-in compressional and extensional contexts with an angle of internal friction $\phi$ of the non-linear medium variable between 0° and 30°.
-
-As expected, two shear bands stem from the inclusion with variable angles in relation to both the dynamics context and the internal friction angle. Shear band
-angles formed at 45° for $\phi=0$° in both extensional and compressional contexts (Fig. \ref{fig:brick_beam}a), while they are different in case of $\phi\neq0$°.
-Shear band angles in case of $\phi=20$° are shown for compression and extension in Fig. \ref{fig:brick_beam}b  and c, respectively. Values of shear band angles
-as function of internal friction angles from 0° to 30° are extracted at different depths and minimum and maximum values are plotted in Fig. \ref{fig:brick},
-in comparison with theoretical Roscoe, Arthur and Coulomb shear band angles. For all tests the tolerance for nonlinear convergence is defined $tol=10^{-7}$
-and none converges before the maximum number of iterations ($it_{max}=1000$) is reached. However, tests with internal friction angles up to 15° show a constant
-decrease of the velocity residuals, which can not be observed for for tests with higher internal friction angles (Fig. \ref{fig:convergence}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Brick_experiment}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Brick_Beam.pdf}
-\caption{Shear band angles predicted for the brick experiment for $\phi=0$° (panel a) and for $\phi=20$° in case of compression and extension (panels b and c,
-respectively).}
-\label{fig:brick_beam}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Brick.pdf}
-\caption{Shear band angles predicted for the brick experiment in case of compressional and extensional contexts as function of different internal angle of
-friction (continuous black line and red dots), compared with theoretical Roscoe, Artur and Coulomb angles (discontinuous black lines). Red lines indicate the
-range of angles calculated at different depths.}
-\label{fig:brick}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=500px]{./Figures/convergence.pdf}
-  \caption{Convergence rates for all tests. Normalised residuals (left column) are compared with velocity residuals (right column) for extensional (first row)
-  and compressional (second row) settings.}
-  \label{fig:convergence}
-  \end{figure}
-
-\subsection{Advection stabilisation}\label{sec:advection}
-The 1D advection problem proposed by \citet{Donea2003} and \citet{Thieulot2011} is performed to verify the effectiveness of the SUPG method to stabilise the
-advection term of the energy equation. The domain is a 1D segment with $L_x=1$ composed by 50 elements and a discontinuity in the thermal field placed at
-$x=0.25$. Temperature is set to 1 for $x \leq 0.25$ and to 0 for $x > 0.25$. Velocity is set to ${u}=1$ in the entire domain. The simulation is performed for
-250 time steps, with $dt=0.002$, so the thermal discontinuity should be at $x=0.75$ at the end of the simulation.
-
-Temperature profiles at the end of the simulation are shown in \ref{fig:advection} as function of the dimensionless coefficient $\gamma=\tau \bm{v}/h$. In case
-of the classic Galerkin method ($\gamma=0$, blue line in \ref{fig:advection}) the final thermal profile is characterised by strong oscillations, which are
-eliminated in case of the SUPG method ($\gamma=0.045$, orange line in \ref{fig:advection}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Advection_stabilisation}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Advection.pdf}
-\caption{Temperature profile as function of \textit{x} for the advection stabilisation benchmark. Purple line indicates the initial temperature profile; the
-green line indicates the analytical temperature profile after 250 time steps; blue line indicates the temperature profile after 250 time steps in case of the
-classic Galerkin method ($\gamma=0$); orange line indicates the temperature profile after 250 time steps in case of the SUPG method ($\gamma=0.045$).}
-\label{fig:advection}
-\end{figure}
-
-\subsection{Shear heating}\label{sec:simple_shear}
-This exercise is performed in an unit square domain composed by \num{128x128} elements. The velocity field is prescribed on the entire domain with
-$\bm{v}=(L_y-y)y\bm{e_x}$; viscosity, density and specific heat are fixed to 1, while thermal conductivity and radiogenic energy are fixed to 0. Therefore,
-the energy equation (Eq. \ref{eq:energy}) can be simplified as \[\frac{\partial T}{\partial t}=H_s\] and fixing $T(t=0)=0$, the temperature can be find as
-\[T(t)=H_s t\]
-In this case we have
-\begin{eqnarray}
-\dot{\epsilon_{xx}}&=&\frac{\partial u}{\partial x}=0 \nonumber \\
-\dot{\epsilon_{yy}}&=&\frac{\partial v}{\partial y}=0 \nonumber \\
-\dot{\epsilon_{xy}}&=&\frac{1}{2}\left(\frac{\partial v}{\partial x}+\frac{\partial u}{\partial y}\right)=\frac{1}{2}(L_y-2y)\nonumber
-\end{eqnarray}
-and, simplifying Eq. \ref{eq:shear}, shear heating can be calculated as \[H_s(x,y)=(1-2y)^2\]
-
-The solution predicted by the model in terms of velocity, temperature and shear heating match well with the analytical solutions (Fig. \ref{fig:analytical_en}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Simple_shear}.
-
-%\begin{figure}
-%\centering
-%\includegraphics[width=400px]{./Figures/Energy_Thieulot.pdf}
-%\caption{Temperature and shear heating predicted by the model after 1 time step.}
-%\label{fig:en_thieulot}
-%\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=500px]{./Figures/Analytical_Shear.pdf}
-\caption{Velocity (panel a), temperature (panel b) and shear heating (panel c) as function of the \textit{y} coordinate for the simple shear experiment. The
-solutions predicted by the model (red lines) are compared with the analytical solutions (dashed blue lines).}
-\label{fig:analytical_en}
-\end{figure}
-
-\subsection{Shear and adiabatic heating}\label{sec:shear}
-This problem is performed as presented in Exercise 9.4 in \citet{Gerya2010}. Two materials are vertically separated in a rectangular domain with $L_x=$
-\SI{1000}{\km}, $L_y=$ \SI{1500}{\km} and a grid resolution of \num{30x20} elements. Constant thermal coefficient expansion ($\alpha=$ \SI{3e-5}{\per\kelvin}),
-temperature ($T=$ \SI{1300}{\kelvin}) and gravity acceleration ($g_y=$ \SI{-10}{\m\square\s}) are assumed in the whole domain. Fluid 1 (on the left side of the
-domain) has $\rho_1=$ \SI{3200}{\kg\per\cubic\m} and $\eta_1=$ \SI{e20}{\pascal\s}; fluid 2 (on the right side of the domain) has $\rho_2=$
-\SI{3300}{\kg\per\cubic\m} and $\eta_2=$ \SI{e22}{\pascal\s}. Velocity boundary conditions are set to free slip on all sides of the domain.
-
-As shown in Fig. \ref{fig:energy}, both shear and adiabatic heating predicted by the code (first row in Fig. \ref{fig:energy}) well recreate the results
-obtained by means of example ${Shear\_adiabatic\_heating.m}$ from \citet{Gerya2010} (second row in Fig. \ref{fig:energy}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Adiabatic%2BShear}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Energy.pdf}
-\caption{Comparison between shear (first column), adiabatic (second column) and total (third column) energy predicted by the code (panels a, b and c) and those
-created using example ${Shear\_adiabatic\_heating.m}$ from Exercise 9.4 in \citet{Gerya2010} (panels d, e and f).}
-\label{fig:energy}
-\end{figure}
-
-\subsection{Mantle convection}\label{sec:mantle}
-This problem is performed as presented by \citet{Blankenbach1989} (constant viscosity cases) and \citet{Thieulot2014}, in a 2D unit square domain with gravity
-acceleration $g_y=10^{10}Ra$. The experiment is performed with three different Rayleigh numbers ($Ra=10^4$, $10^5$ and $10^6$) and with different grid
-resolution (between \num{32x32} and \num{128x128} elements). The fluid has constant viscosity, initial density, heat capacity, thermal conductivity
-($\eta= \rho_0= C_p = k=1$), reference temperature ($T_0=0$) and thermal expansion coefficient ($\alpha=10^{-10}$). Temperatures are set to 0 on top and 1 on
-bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\]
-
-The solution generated by the code in terms of $v_{\textrm{rms}}$ (calculated as in Sec. \ref{sec:error}) and Nu as function of time are reported for all the
-simulations in Fig. \ref{fig:mantle}. At the steady state, $v_{\textrm{rms}}$ and Nu of all simulations converge well toward the values from \citet{Blankenbach1989},
-with lower errors for higher resolution grids (Table \ref{tab:mantle}). All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Mantle_convection}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Convection.pdf}
-\caption{$v_{\textrm{rms}}$ (panels a, c and e) and Nu (panels b, d and f) for the mantle convection benchmark as function of time for different grid resolution.
-Panels a and b show the results for $Ra=10^4$; Panels c and d show the results for $Ra=10^5$; panels e and f show the results for $Ra=10^6$. Purple lines
-indicate the convergence values for the $v_{\textrm{rms}}$ and the Nu from \citet{Blankenbach1989} at the steady state.}
-\label{fig:mantle}
-\end{figure}
-
-\begin{table}[H]
-\caption{Comparison between $v_{\textrm{rms}}$ and Nu predicted by the code for the mantle convection experiment and same values as reported in literature.}
-\centering
-\small
-\begin{tabular}{|>{\centering}m{0.09\linewidth}|>{\centering}m{0.05\linewidth}|>{\centering}m{0.24\linewidth}|>{\centering}m{0.15\linewidth}|
-  >{\centering}m{0.1\linewidth} >{\centering}m{0.1\linewidth} >{\centering\arraybackslash}m{0.1\linewidth}|}
-\toprule
-& & \multirow{2}{*}{\citet{Blankenbach1989}} & \citet{Thieulot2014} & \multicolumn{3}{c|}{FALCON}  \\
- & & & \num{200x200} & \num{32x32} & \num{64x64} & \num{128x128}  \\
-\midrule
-\multirow{2}{*}{$Ra=10^4$} & $v_{\textrm{rms}}$ & $42.864947 \pm 0.000020$ & 42.867 & 42.83226 & 42.852793 & 42.861394  \\
-           & Nu      & $4.884409 \pm 0.000010$  & 4.882  & 4.781297 & 4.857475  & 4.877573   \\
-\hline
-\multirow{2}{*}{$Ra=10^5$} & $v_{\textrm{rms}}$ & $193.21454 \pm 0.00010$  & 193.255& 193.872643 & 193.377472 & 193.252290  \\
-           & Nu      & $10.534095 \pm 0.000010$ & 10.507 & 9.602514 & 10.270735  & 10.465629   \\
-\hline
-\multirow{2}{*}{$Ra=10^6$} & $v_{\textrm{rms}}$ & $833.98977 \pm 0.00020$  & 834.712& 848.091176 & 837.767911 & 834.945793  \\
-           & Nu      & $21.972465 \pm 0.000020$ & 21.695 & 15.999266 & 19.703682  & 21.306939   \\
-\bottomrule
-\end{tabular}
-\label{tab:mantle}
-\end{table}
-
-\subsection{Viscoplastic mantle convection}\label{sec:visco_mantle}
-This benchmark is performed as Cases 1-5a described in \citet{Tosi2015} in a 2D unit square domain with two different grid resolutions of \num{32x32} 
-and \num{100x100} elements. Reference viscosity, reference density, heat capacity, thermal conductivity and thermal expansion coefficient are fixed to a constant
-value of 1 ($\eta_0= \rho_0= C_p = k=\alpha=1$), while the gravity has been chosen as $g_y=10^2$, in order to obtain a Rayleigh number $Ra=10^2$. Temperatures
-are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\].
-The viscosity field $\eta$ is calculated as the harmonic average between a linear part $\eta_{lin}$ that depends on temperature only or on temperature and 
-depth $y$ and a nonlinear-plastic part $\eta_{pl}$ that depends on the strain rate $\dot{\bm{\epsilon}}$, as follows
-\[\eta(T,y,\dot{\bm{\epsilon}})=2\left(\frac{1}{\eta_{lin}(T,y)}+\frac{1}{\eta_{pl}(\dot{\bm{\epsilon}})}\right)^{-1}\]
-The linear and the nonlinear-plastic parts of the viscosity are calculated following \citet{Tosi2015} as follows
-\[\eta_{lin}(T,y)=\exp(-\ln(\Delta\eta_T)T+\ln(\Delta\eta_y)y)\]
-\[\eta_{pl}(\dot{\bm{\epsilon}})=\eta^*+\frac{\sigma_Y}{\sqrt{\dot{\bm{\epsilon}}:\dot{\bm{\epsilon}}}}\]
-where $\Delta\eta_T$, $\Delta\eta_y$, $\eta^*$ and $\sigma_Y$ are parameters chosen as listed in Table \ref{tab:tosi_setup}.
-
-For each case, some diagnostic quantities measured at the steady state are reported in Table \ref{tab:visco_mantle} (see Section \ref{sec:error} for 
-the computation of all the quantities), showing a well fit with results described in \citet{Tosi2015}. Data and charts for all cases can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Viscoplastic_mantle_convection}.
-
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{white!100!}{gray!25!}
-%\begin{table}[H]
-%  \caption{Results of the viscoplastic mantle convection benchmark for Cases 1-4 compared with results obtained with ELEFANT and YACC (from \citealp{Tosi2015}).}
-%  \centering
-%  \resizebox{\linewidth}{!}{
-%\begin{tabular}{|c|ccc|ccc|ccc|ccc|}
-%  \toprule
-%    & \multicolumn{3}{c|}{\textbf{Case 1}} & \multicolumn{3}{c|}{\textbf{Case 2}} & \multicolumn{3}{c|}{\textbf{Case 3}} &
-%    \multicolumn{3}{c|}{\textbf{Case 4}} \\
-%   & FALCON & ELEFANT& YACC & FALCON& ELEFANT& YACC & FALCON & ELEFANT & YACC & FALCON & ELEFANT & YACC \\
-%  \midrule
-%   $\left\langle T \right\rangle$ & 0.7759 & 0.7758 & 0.7767 & 0.6043 & 0.6036 & 0.6058 & 0.7275 & 0.7275 & 0.7286 & & 0.5277 & 0.5289\\
-%    Nu$^{top}$ & 3.4280 & 3.4214 & 3.4298 & 8.5242 & 8.5115 & 8.5278 & 3.0368 & 3.0347 & 3.0374 && 6.5912 & 6.5572 \\
-%    Nu$^{bottom}$ & 3.3168 & 3.313 & 3.3143 & 8.4599 & 8.4465 & 8.3990 & 3.0086 & 2.9908 & 2.9628 && 6.5834 & 6.5243 \\
-%    $v_{\textrm{rms}}$ & 249.901 & 249.134 & 251.7997 & 141.2451 & 140.7087 & 142.2020 & 100.2776 & 100.1208 & 100.9467 && 79.1105 & 79.6202 \\
-%    $v_{\textrm{rms}}^{top}$ & 1.8907 & 1.8642 & 1.8298 & 104.7928 & 104.1228 & 105.1767 & 2.0832 & 2.0652 & 2.0374 && 74.7596 & 75.8414 \\
-%    $u_{max}^{top}$ & 2.6340 & 2.6119 & 2.5516 & 121.9212& 121.766 & 122.3238 & 2.9086 & 2.9019 & 2.8458 && 88.9146 & 89.2940 \\
-%    $\eta_{min}$ & - & - & - & $2.0303\times10^{-5}$ & $2.0600\times10^{-5}$ & $1.9853\times10^{-5}$ & $4.8129\times10^{-5}$ & $4.8080\times10^{-5}$ & $4.7907\times10^{-5}$ &
-%    & $1.9860\times10^{-4}$ & $1.9174\times10^{-4}$ \\
-%    $\eta_{max}$ & - & - & - & 1.3325 & 1.85 & 1.6095 & 0.9360 & 0.9023 & 1 && 1.52 & 1.6773 \\
-%    $\left\langle W \right\rangle$ & 2.4299 & 2.4316 & 2.4583 & 7.5648 & 7.5665 & 7.6222 & 2.0387 & 2.0384 & 2.0400 && 5.6216 & 5.6512 \\
-%    $\left\langle \Phi \right\rangle/Ra$ & 2.4311 & 2.4276 & 2.4333 & 7.5661 & 7.5585 & 7.6194 & 2.0386 & 2.0370 & 2.0335 && 5.6182 & 5.6463 \\
-%    $\delta$ & 0.05\% & 0.16\% & 1.02\% & 0.02\% & 0.11\% & 0.04\% & $<$0.01\% & 0.07\% & 0.32\% && 0.06\% & 0.09\% \\
-%  \bottomrule
-%  \end{tabular}}
-%  \label{tab:visco_mantle}
-%  \end{table}
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{2}{white!100!}{gray!25!}
-\begin{table}[H]
-  \caption{Benchmark cases and corresponding parameters.}
-  \centering
-  \small
-\begin{tabular}{|>{\centering}m{0.08\linewidth}|>{\centering}m{0.05\linewidth}>{\centering}m{0.05\linewidth}>{\centering}m{0.05\linewidth}
-  >{\centering}m{0.05\linewidth} >{\centering}m{0.05\linewidth} |>{\centering\arraybackslash}m{0.2\linewidth}|}
-  \toprule
-    Case & $Ra$ & $\Delta\eta_T$ & $\Delta\eta_y$ & $\eta^*$ & $\sigma_Y$ & Convective regime \\
-  \midrule
-  \rowcolor{gray!25!}
-    1 & $10^2$ & $10^5$ & 1 & - & - & Stagnant lid \\
-    2 & $10^2$ & $10^5$ & 1 & $10^{-3}$ & 1 & Mobile lid  \\
-    \rowcolor{gray!25!}
-    3 & $10^2$ & $10^5$ & 10 & - & - & Stagnant lid  \\
-    4 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 1 & Mobile lid \\
-    \rowcolor{gray!25!}
-    5 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 4 & Periodic \\
-  \bottomrule
-  \end{tabular}
-  \label{tab:tosi_setup}
-  \end{table}
-
-\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{gray!25!}{white!100!}
-\begin{table}
-  \caption{Results of the viscoplastic mantle convection benchmark for different grid resolutions in Cases 1-5 compared with results obtained with ELEFANT
-  and YACC ($100\times100$ elements, from \citealp{Tosi2015}) and in Stone 28 ($32\times32$ elements, from
-  \url{https://github.com/cedrict/fieldstone/tree/master/python_codes/fieldstone_28}).}
-  \centering
-  \resizebox{\linewidth}{!}{
-\begin{tabular}{|cc|ccccccccccc|}
-  \toprule
-  \multicolumn{13}{|c|}{\textbf{Case 1}} \\
-  \midrule
-  Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-  & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-  \hline
-  \rowcolor{gray!25!}
-    YACC & $100\times100$ & 0.7767 & 3.4298 & 3.3143 & 251.7997 & 1.8298 & 2.5516 & - & - & 2.4583 & 2.4333 & 1.02\% \\
-    ELEFANT & $100\times100$ & 0.7758 & 3.4214 & 3.313 & 249.134 & 1.8642 & 2.6119 & - & - & 2.4316 & 2.4276 & 0.16\% \\
-    \rowcolor{gray!25!}
-    Stone 28 & $32\times32$ & 0.7737 & 3.3987 & - & 243.872 & - & - & - & - & - & - & - \\
-%    FALCON & $32\times32$ & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & - & - & 2.4745 & 2.4864 & 0.48\% \\
-%    \rowcolor{gray!25!}
-    FALCON & $100\times100$ & 0.7759 & 3.4280 & 3.3168 & 249.901 & 1.8907 & 2.6340 & - & - & 2.4299 & 2.4311 & 0.05\% \\
-    \midrule
-    \multicolumn{13}{|c|}{\textbf{Case 2}} \\
-    \midrule
-    Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-    & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-    \hline
-    \rowcolor{gray!25!}
-     YACC & $100\times100$ & 0.6058 & 8.5278 & 8.3990 & 142.2020 & 105.1767 & 122.3238 & $1.9853\times10^{-5}$ & 1.6095 & 7.6222 & 7.6194 & 0.04\% \\
-     ELEFANT & $100\times100$ & 0.6034 & 8.5115 & 8.4465 & 140.8067 & 104.1228 & 121.766 & $2.0600\times10^{-5}$ & 1.85 & 7.5665 & 7.5585 & 0.11\% \\
-     \rowcolor{gray!25!}
-     Stone 28 & $32\times32$ & 0.6052 & 8.1715 & - & 141.518 & - & - & - & - & - & - & - \\
-%     FALCON & $32\times32$ & 0.6454 & 8.1422 & 7.4020 & 156.6908 & 95.7226 & 116.9302 & $2.2572\times10^{-5}$ & 0.7194 & 7.4289 & 7.4452 & 0.22\% \\
-%     \rowcolor{gray!25!}
-     FALCON & $100\times100$ & 0.6044 & 8.5266 & 8.4590 & 141.3684 & 104.8867 & 122.0470 & $2.0333\times10^{-5}$ & 1.2914 & 7.5707 & 7.5718 & 0.01\% \\
-     \midrule
-     \multicolumn{13}{|c|}{\textbf{Case 3}} \\
-     \midrule
-     Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-     & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-     \hline
-     \rowcolor{gray!25!} 
-     YACC & $100\times100$ & 0.7286 & 3.0374 & 2.9628 & 100.9467 & 2.0374 & 2.8458 & $4.7907\times10^{-5}$ & 1 & 2.0400 & 2.0335 & 0.32\% \\
-      ELEFANT & $100\times100$ & 0.7275 & 3.0347 & 2.9908 & 100.1208 & 2.0652 & 2.9019 & $4.8080\times10^{-5}$ & 0.9023 & 2.0384 & 2.037 & 0.07\% \\
-      \rowcolor{gray!25!}
-      Stone 28 & $32\times32$ & 0.7277 & 3.0347 & - & 100.018 & - & - & - & - & - & - & - \\
-%      FALCON & $32\times32$ & 0.7281 & 3.0618 & 2.8132 & 101.0092 & 2.1630 & 3.0178 & $4.8500\times10^{-5}$ & 0.7845 & 2.0699 & 2.0695 & 0.02\% \\
-%      \rowcolor{gray!25!}
-      FALCON & $100\times100$ & 0.7275 & 3.0368 & 3.0086 & 100.2776 & 2.0832 & 2.9086 & $4.8129\times10^{-5}$ & 0.9360 & 2.0387 & 2.0386 & $<$0.01\% \\
-      \midrule
-      \multicolumn{13}{|c|}{\textbf{Case 4}} \\
-      \midrule
-      Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-      & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-      \hline
-      \rowcolor{gray!25!} 
-      YACC & $100\times100$ & 0.5289 & 6.5572 & 6.5243 & 79.6202 & 75.4814 & 89.2940 & $1.9174\times10^{-4}$ & 1.6773 & 5.6512 & 5.6463 & 0.09\% \\
-       ELEFANT & $100\times100$ & 0.5277 & 6.5912 & 6.5834 & 79.1105 & 74.7596 & 88.9146 & $1.9860\times10^{-4}$ & 1.5200 & 5.6216 & 5.6182 & 0.06\% \\
-       \rowcolor{gray!25!}
-       Stone 28 & $32\times32$ & 0.5286 & 6.4036 & - & 79.4746 & - & - & - & - & - & - & - \\
-%       FALCON & $32\times32$ & 0.5334 & 6.4679 & 6.3435 & 80.6153 & 76.0511 & 90.2058 & $2.0393\times10^{-4}$ & 0.6453 & 5.6676 & 5.6698 & 0.04\% \\
-%       \rowcolor{gray!25!}
-       FALCON & $100\times100$ & 0.5279 & 6.5897 & 6.5830 & 79.1220 & 75.0715 & 88.8484 & $1.9693\times10^{-4}$ & 1.4133 & 5.6127 & 5.6113 & 0.03\% \\
-       \midrule
-       \multicolumn{13}{|c|}{\textbf{Case 5}} \\
-       \midrule
-       Code & Resolution & $P$ & $\left\langle T \right\rangle_{min}$ & $\left\langle T \right\rangle_{max}$ & Nu$^{top}_{min}$ & Nu$^{top}_{max}$ &
-       $v_{\textrm{rms}}^{min}$ & $v_{\textrm{rms}}^{max}$ & \multicolumn{2}{c}{$(\left\langle \Phi \right\rangle/Ra)_{min}$} & 
-       \multicolumn{2}{c|}{$(\left\langle \Phi \right\rangle/Ra)_{max}$} \\
-       \hline
-       \rowcolor{gray!25!}
-         YACC & $100\times100$ & 0.0792 & 0.6543 & 0.6722 & 2.6950 & 7.2792 & 41.9813 & 98.8230 & \multicolumn{2}{c}{1.4347} & 
-         \multicolumn{2}{c|}{9.2748} \\
-         ELEFANT & $100\times100$ & 0.076 & 0.6514 & 0.6693 & 2.6808 & 7.3506 & 41.2578 & 100.089 & \multicolumn{2}{c}{1.3324} & 
-         \multicolumn{2}{c|}{9.5289} \\
-         \rowcolor{gray!25!}
-         Stone 28 & $32\times32$ & 0.0764 & 0.6521 & 0.6697 & 2.6912 & 7.0763 & 41.61 & 98.183 & \multicolumn{2}{c}{-} & 
-         \multicolumn{2}{c|}{-} \\
-     %    FALCON & $32\times32$ & & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & \multicolumn{2}{c}{-} & 
-     %     \multicolumn{2}{c|}{-} \\
-     %    \rowcolor{gray!25!}
-         FALCON & $100\times100$ & 0.0775 & 0.6519 & 0.6697 & 2.6847 & 7.3629 & 41.5256 & 99.5466 & \multicolumn{2}{c}{1.3399} & 
-         \multicolumn{2}{c|}{9.3993} \\
-  \bottomrule
-  \end{tabular}}
-  \label{tab:visco_mantle}
-  \end{table}
-
-\subsection{Thin layer entrainment}\label{sec:thin}
-This experiment is performed as originally proposed by \citet{vanKeken1997} with two incompressible fluids in a rectangular domain with $L_x=2$ and $L_y=1$
-and gravity acceleration $g_y=10^{10}Ra$. The Rayleigh number and the compositional Rayleigh number are fixed to $Ra=300000$ and $Rc=450000$, respectively.
-Both fluids have constant viscosity, thermal conductivity, specific heat ($\eta=\rho=C_p=1$) and thermal expansion coefficient ($\alpha=10^{-10}$). Fluid 1
-has a density $\rho_1=1$, while fluid 2 is denser ($\rho_2=\rho_1+1.5 \cdot 10^{-10}$) and is placed at the bottom of the domain, for $y \leq 0.025$.
-Temperature are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides of the domain. The initial
-temperature field is given by
-\[T(x,y)=T_u(x,y)+T_l(x,y)+T_r(x,y)+T_s(x,y)-\frac{3}{2}\]
-with
-\begin{eqnarray}
-T_u(x,y)&=&\frac{1}{2}erf\left(\frac{1-y}{2}\sqrt{\frac{u_0}{x}}\right)\nonumber \\
-T_l(x,y)&=&1-\frac{1}{2}erf\left(\frac{y}{2}\sqrt{\frac{u_0}{L_x-x}}\right)\nonumber \\
-T_r(x,y)&=&\frac{1}{2}+\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{y+1}} \exp\left(-\frac{x^2u_0}{4y+4}\right)\nonumber \\
-T_s(x,y)&=&\frac{1}{2}-\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{2-y}} \exp\left(-\frac{(L_x-x)^2u_0}{8-4y}\right)\nonumber
-\end{eqnarray}
-and
-\begin{eqnarray}
-u_0&=&\frac{L_x^{7/3}}{(1+L_x^4)^{2/3}}\left(\frac{Ra}{2\sqrt{\pi}}\right)^{\frac{2}{3}}\nonumber \\
-Q&=&2\sqrt{\frac{L_x}{\pi u_0}}\nonumber
-\end{eqnarray}
-
-The experiment is performed with two grid resolution (\num{125x40} and \num{200x80} elements), with 100 markers per element and CFL of 0.25. $v_{\textrm{rms}}$
-(see Sec. \ref{sec:error}) calculated for both the simulations match well with results from \citet{vanKeken1997} and obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and ELEFANT \citep{Thieulot2014}, obtained with similar grid resolutions
-(Fig. \ref{fig:thin}). In addition, a variety of simulations with different resolutions and aspect ratios are performed to verify that the sensitivity of the
-initial velocity field. Results match well with results from \citet{vanKeken1997} (in the grey area) and \citet{Thieulot2014} (Fig. \ref{fig:thin_initial}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Thin_layer}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/vrms_thin.pdf}
-\caption{$v_{\textrm{rms}}$ for the thin layer experiment as function of time for different grid resolution. Results are compared with results obtained by
-\citet{vanKeken1997} (black lines) and with ASPECT and ELEFANT.}
-\label{fig:thin}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/vrmszero.pdf}
-\caption{$v_{\textrm{rms}}$ at $t=0$ for the thin layer experiment as function of the element size for different aspect ratios of the elements. Results are compared
-with results obtained by \citet{vanKeken1997} (black dots in the grey area) and with ASPECT and ELEFANT.}
-\label{fig:thin_initial}
-\end{figure}
-
-\subsection{Sinking hydrated cylinder}\label{sec:quinquis}
-This experiment is performed as explained in \citet{Quinquis2014} and simulates the sinking of a cold, hydrated cylinder into a warm, dry mantle. The domain
-is square with $L_x=L_y=$ \SI{300}{\km}, a grid resolution of \num{300x300} elements and 25 markers per element. The sinking cylinder is located at $x=$
-\SI{150}{\km} and $y=$ \SI{170}{\km} with a radius of \SI{20}{\km} and has $\rho_c=$ \SI{3250}{\kg\per\cubic\m}, $\eta_c=$ \SI{e23}{\pascal\s}, a thermal
-conductivity $k_c=$ \SI{4.5}{\watt\per\m\per\kelvin},a specific heat $C_{p_c}=$ \SI{1250}{\joule\per\kg\per\kelvin}; the surrounding mantle has $\rho_m=$
-\SI{3200}{\kg\per\cubic\m}, $\eta_m=$ \SI{e20}{\pascal\s}, a thermal conductivity $k_m=$ \SI{105}{\watt\per\m\per\kelvin} and specific heat $C_{p_m}=$
-\SI{1250}{\joule\per\kg\per\kelvin}; a 58 km-thick lithosphere overlies the mantle and has $\rho_l=$ \SI{3200}{\kg\per\cubic\m}, $\eta_l=$ \SI{e23}{\pascal\s},
-a thermal conductivity $k_l=$ \SI{4.5}{\watt\per\m\per\kelvin}, specific heat $C_{p_l}=$ \SI{750}{\joule\per\kg\per\kelvin} and a thermal diffusivity of
-\SI{e-6}{\square\m\per\s}. Velocity boundary conditions are set to free slip on all sides of the domain. Temperature increase linearly in the lithosphere,
-from 0° to 1300°C, and in the mantle, from 1300° to 1360.5°C. The initial temperature of the cylinder is fixed to 400°C. Throughout the evolution, temperature
-is fixed to 0° and 1360.5°C at the top and bottom of the domain, respectively. The initial bound water content is imposed at 2 wt.\% in the cylinder and at
-0 wt.\% in both the mantle and the lithosphere. The maximum amount of water in the mantle is fixed to 0.2 wt.\%, while maximum water content in the cylinder
-and in the lithosphere is function of pressure and temperature and is calculated using a serpentinized harzburgite with Perple\_X, as in \citet{Quinquis2014}.
-
-As shown in \citet{Quinquis2014}, the model is characterised by a progressive dehydration in the external portion of the cylinder, due to the increase of
-temperature, with a consequent vertical migration of free water that hydrates the mantle above the cylinder, up to the lithosphere. The distribution of
-bound and free water during the evolution of the experiment is shown in Fig. \ref{fig:quinquis} (left and right columns, respectively). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Hydration}.
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Hydration.pdf}
-\caption{Distribution of bound and free water (left and right columns, respectively) at $t=0.5$ and $t=$ \SI{1.25}{\mega\year} (first and second rows,
-respectively) for the sinking hydrated cylinder experiment. Only markers with bound or free water not equal to 0 are plotted.}
-\label{fig:quinquis}
-\end{figure}
-
-\subsection{Mantle melting curves}\label{sec:katz}
-This instantaneous experiment is performed in order to recreate isobaric melting curves, as function of temperature, bulk water and content of water in the
-melt, to compare with melting curves obtained by \citet{Katz2003}. The domain is square with $L_x=L_y=100$ km, a grid resolution of \num{100x100} elements and
-25 markers per element. Density is set in the entire domain to \SI{3300}{\kg\per\cubic\m} and temperature increases from the sides to the centre from 1000° to
-1500°C. The experiment is performed for bulk water contents from 0 to 0.3 wt.\%.
-
-Fig. \ref{fig:fraction} shows isobaric melting curves for \SI{1}{\giga\pascal} (panel a) and \SI{3}{\giga\pascal} (panel b) calculated for different bulk water
-contents. The curves well match with those obtained for the mantle by \citet{Katz2003}, showing the saturation at the solidus for bulk water of 0.2 and 0.3 wt.\%, indicating
-that the code correctly determines melt fraction as function of pressure, temperature and bulk water content. Same results can be observed in Fig.
-\ref{fig:water}a, where isobaric and isothermal melting curves are shown as function of bulk water content. In addition, Fig. \ref{fig:water}b shows that the
-code correctly determines also the percentage of water in the melt as function of pressure, temperature and melt fraction. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Melting}.
-
-\begin{figure}
-\includegraphics[width=\linewidth]{./Figures/Fraction.pdf}
-\caption{Isobaric melting curves for 1 GPa (panel a) and 3 GPa (panel b) as function of temperature with different bulk water contents.}
-\label{fig:fraction}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\linewidth]{./Figures/Water.pdf}
-\caption{Isobaric melting curves for 1.5 GPa as function of bulk water (panel a) and water in the melt (panel b) calculated for different temperatures.}
-\label{fig:water}
-\end{figure}
-
-\clearpage
-
-\bibliography{export}
-
-\end{document}

### 2022-09-07 tex/intro.tex
update main

--- 2022-03-31 FALCON_description.tex
@@
-\documentclass[hidelinks,10pt,a4paper]{article}
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{textcomp}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage{booktabs}
-\usepackage{multirow}
-\usepackage[cm]{fullpage}
-\usepackage[round]{natbib}
-\bibliographystyle{abbrvnat}
-\usepackage{doi}
-\usepackage{bm}
-\usepackage[table]{xcolor}
-\hypersetup{
-    colorlinks,
-    linkcolor={blue},
-    citecolor={blue},
-    urlcolor={blue}
-}
-\usepackage{float}
-\restylefloat{table}
-\usepackage{siunitx}
-\DeclareSIUnit\year{yr}
-
-\begin{document}
-
 \section{Introduction}\label{sec:intro}
@@
 \end{equation}
-This simplification is known as the extended Boussinesq approximation.
+This simplification is known as the Boussinesq approximation.
 
@@
 $\kappa=\frac{k}{\rho C_p}$ is the thermal diffusivity \citep{Thieulot2014}.
-
-\section{Lagrangian markers}\label{sec:markers}
-Elemental properties (density, viscosity, thermal conductivity, specific heat and thermal expansion) needed to solve Eqs. \ref{eq:mass_0}, \ref{eq:momentum}
-and \ref{eq:energy} are related to the composition of each element, determined by means of Lagrangian markers that are characteristic of different materials
-in the domain. Elemental properties (with the sole exception of the viscosity) are calculated using an arithmetic mean, as \[P_e=\frac{1}{n}\sum_{i=1}^n P_i\]
-where $P_e$ is the elemental property, $P_i$ is the property characteristic of the material of each marker and $n$ is the number of markers of the element.
-Differently, the average scheme for the elemental viscosity can be chosen between harmonic, geometric and arithmetic mean. Markers can be placed regularly or
-randomly at the beginning of the simulation and their advection is performed by either a 2nd-order or a 4th-order Runge-Kutta in space, interpolating the velocity 
-field on each marker by means of the shape functions (see Section \ref{sec:runge}). The interpolated velocity is then corrected by means of the Conservative
-Velocity Interpolation (CVI), which introduces a corrective term to reduce dispersion and clustering of particles in both steady state and time-dependent 
-\citep{Wang2015} (see Section \ref{sec:CVI}).
-
-The initial distribution of the markers is created using the open-source code library
-Geodynamic World Builder \citep{Fraters2019}. During the simulation, each marker carries memory of temperature, pressure and accumulated strain, which are
-determined interpolating the relative nodal parameter. In particular, pressure is interpolated onto each marker after the smoothing procedure. The number
-of markers contained in each element is maintained between $n_{min}$ and $n_{max}$. When in an element there are less markers than $n_{min}$ the code adds
-random markers to reach the $n_{min}$, while if the number is higher than $n_{max}$ some of them are randomly deleted. When new markers are added, they assume
-the properties of the nearest marker. In this way elements are never empty and maintain a number of markers inside a prefixed range.
-
-\section{The momentum equation}\label{sec:mom_eq}
-The penalty method is implemented so that Eq. \ref{eq:mass_0} can be written as:
-\begin{equation}\label{eq:mass_penalty}
-\bm{\nabla} \cdot \bm{u} + \frac{p}{\lambda}=0
-\end{equation}
-where $\lambda$ is the so-called penalty parameter, which should be 6-7 orders of magnitude larger than the shear viscosity to ensure that mass conservation
-is satisfied \citep{Donea2003,Thieulot2014}.
-
-The deviatoric stress tensor in Eq. \ref{eq:momentum} can be written in terms of the strain rate tensor as $\bm{\tau}=2\eta\dot{\bm{\epsilon}}$, with
-$\dot{\bm{\epsilon}}=\frac{1}{2}\left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)$. Therefore, Eq. \ref{eq:momentum} can be rewritten as:
-\begin{equation}\label{eq:momentum_pressure}
--\bm{\nabla} p + \bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right) + \rho \bm{g} = 0
-\end{equation}
-Finally, using pressure from Eq. \ref{eq:mass_penalty}, Eq. \ref{eq:momentum_pressure} can be rewritten as:
-\begin{equation}\label{eq:momentum_penalty}
-\lambda \bm{\nabla} \left(\bm{\nabla} \cdot \bm{v}\right)+\bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right)+\rho\bm{g}=0
-\end{equation}
-The penalty method is been associated to the iterative Uzawa method, as described in detail in \citet{Dabrowski2008} and \citet{Thieulot2014}
-(see Section \ref{sec:poiseuille}). Results of benchmarks performed to verify the correctness in the implementation of Eq. \ref{eq:momentum_penalty} are
-shown in Sections \ref{sec:ist_sphere} and \ref{sec:rayleigh}.
-
-To support large viscosities variations the penalty parameter is related to the effective elemental viscosity by means of a dimensionless coefficient, so that
-$\lambda=\lambda_e(e)\eta_{eff}(e)$ \citep{Marotta2006,Dabrowski2008,Thieulot2014}. Benchmark of the falling block \citep{Gerya2003a,Gerya2010b,Thieulot2011}
-is performed to verify that the code can correctly deal buoyancy driven flows with strong viscosity contrasts (Section \ref{sec:block}).
-
-\section{Sticky air and free surface}\label{sec:sticky}
-The Earth's surface can be treated by means of either the so-called sticky-air or a true free surface method, both of them implemented in the code. In the
-sticky-air method the surface is approximated with the introduction of a buoyant layer with a viscosity at least four orders of magnitude lower than the crust
-\citep{Schmeling2008,Crameri2012} and the interface between lithosphere and air is defined using a chain of passive markers that are advected as the Lagrangian
-markers. The correctness of the evolution of the markers chain is tested with the experiment of a 2D time-dependent Stokes sphere below a free surface and
-compared with results from ASPECT\footnote{\url{https://aspect.geodynamics.org/}} \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} (Section
-\ref{sec:sphere}). In the true free surface case the top boundary is assumed stress-free and velocities are not fixed. In this case topography variations are
-described by vertical deformations of the mesh that depend on the velocity field of the nodes that identify the free surface, while horizontal deformations are
-not taken into account. This procedure is known as the Arbitrary Lagrangian-Eulerian (ALE) method and its implementation follows the technique described in
-\citet{Thieulot2011}. However, although the implementation of a true free surface better reproduces laboratory experiment, extremely small time steps can be
-necessary to maintain stability \citep{Kaus2010a,Quinquis2011,Thieulot2014}. Therefore, the stability algorithm proposed by \citet{Kaus2010a} is implemented
-to avoid instabilities due to high density differences at the free surface when using too large time steps. The implementation of this algorithm are tested by
-performing the experiment described by \citet{Kaus2010a} (Section \ref{sec:stab}).
-
-The topography relaxation benchmark proposed by \citet{Crameri2012} is performed to verify that the code correctly recovers topography variations in case of
-both the sticky-air and the true free surface method (Section \ref{sec:crameri}). Finally, Section \ref{sec:subduction} show the results of the spontaneous
-subduction experiment described by \citet{Schmeling2008}.
-
-\section{Erosion and sedimentation}\label{sec:erosion}
-Surface processes at either the lithosphere-air interface or in correspondence of the free surface have been implemented by means of the software
-Fastscape\footnote{\url{https://fastscape.org/}} \citep{Braun2013,Cordonnier2019,Yuan2019a,Yuan2019b}. FastScapeLib is a set of routines that solve the stream
-power law (enriched by a sediment transport/deposition term) \citep{Yuan2019a}, hillslope diffusion and marine transport and deposition \citep{Yuan2019b}.
-The partial differential equation solved by FastScapeLib is:
-\begin{equation}
-\frac{\partial h}{ \partial t}=U-K_f A^{m_e}S^{n_e} + \frac{G}{A}\int_A \left( U-\frac{\partial h}{\partial t} \right) dA + K_d \bm{\nabla}^2 h 
-\end{equation}
-where $h$ is the topography, $U$ is the uplift, $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $S$ is the slope, $m_e$ and $n_e$ are
-parameters that should be chosen to constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}, $G$ is the dimensionless
-deposition/transport coefficient for the enriched stream power law and $K_d$ is the effective diffusivity. Terms $K_f A^{m_e}S^{n_e}$ and $K_d \bm{\nabla}^2 h$
-consider long-range processes, described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, and short-range processes,
-proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}.
-
-In case of sticky air, slope is determined by means of the markers chain, which is vertically corrected considering erosion and sedimentation rates. After the
-correction, continental markers above the markers chain are transformed in air markers and air markers below the markers chain become sediments. Similarly,
-slope is determined considering the nodes on the top boundary in case of true free surface and they are vertically corrected as for the markers chain. In this
-case, continental markers above the top boundary are simply deleted.
-%Surface processes at the lithosphere-air interface are implemented taking into account fluvial erosion (long-range processes), hillslope processes (short-range
-%processes) and failure of slopes exceeding a threshold angle $\theta$ \citep{Burov1997,Tucker1999,Sacek2012,Croissant2014,Ueda2015,Beucher2020}. Erosion rates
-%$\dot{E}_{lr}$ due to long-range processes can be described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{lr}=K_f A^{m_e} \bm{\nabla} h^{n_e}
-%\end{equation}
-%where $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $\bm{\nabla} h$ is the slope and $m_e$ and $n_e$ are parameters chosen in to
-%constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}. Erosion rates $\dot{E}_{sr}$ due to short-range processes are
-%implemented by assuming that transport is linearly proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{sr}=K_d \bm{\nabla}^2 h
-%\end{equation}
-%where $K_d$ is the effective diffusivity. In addition, the slope is modified if it still exceeds a threshold angle after long- and short-range processes
-%\citep{Ueda2015}.
-
-%Offshore sedimentation processes are implemented similarly to hillslope processes but with a depth-dependent coefficient \citep{Sacek2012,Beucher2020}.
-%The sedimentation rate $\dot{S}$ is calculated as
-%\begin{equation}
-%\dot{S}=K_s \bm{\nabla}^2 h
-%\end{equation}
-%with
-%\begin{equation}
-%K_s=K^*_s \exp\left(\frac{h-h_{sl}}{h_0}\right)
-%\end{equation}
-%where $K^*_s$ is a constant, $h_{sl}$ is the sea level and $h_0$ is a characteristic length. In case of sticky air, slope is determined by means of the markers
-%chain, which is vertically corrected considering erosion and sedimentation rates. After the correction, continental markers above the markers chain are
-%transformed in air markers and air markers below the markers chain become sediments. Similarly, slope is determined considering the nodes on the top boundary
-%in case of true free surface and they are vertically corrected as for the markers chain. In this case, continental markers above the top boundary are simply
-%deleted.
-
-\section{Non-linear rheologies}\label{sec:nonlinear}
-Non-linear rheologies are implemented combining viscous creep (dislocation and diffusion) and plastic yielding. For each marker, diffusion and dislocation
-viscosity ($\eta_{df|ds}$) can be determined as follows \citep{Karato1993,Warren2008b,Wang2016}:
-\begin{equation}\label{eq:viscous}
-\eta_{df|ds}=f_s \left(\frac{d^m}{A}\right)^{\frac{1}{n}} \left (\frac{2^{\frac{1-n}{n}}}{3^{\frac{n+1}{2n}}} \right)
-I_2^{\frac{(1-n)}{n}}\exp\left(\frac{Q+pV}{nRT}\right)
-\end{equation}
-where $f_s$ is a scaling factor used to represent lithologies that are stronger or weaker than the base set, $d$ is the grain size, $m$ is the grain size exponent,
-$n$ is the stress exponent, $A$ is the uniaxial pre-exponential factor, $I_2$ is the square root of the second invariant of the strain rate tensor, $Q$ is the
-activation energy, $p$ is the pressure, $V$ is the activation volume, $R$ is the gas constant and $T$ is the temperature.
-Pressures and temperatures are determined by the interpolation of the nodal parameters, while strain rates are calculated by means of the
-derivative of the velocities on the nodes. In case of diffusion creep $n=1$ and $m>0$, in case of dislocation creep $n>1$ and $m=0$. The viscous creep $\eta_{cp}$
-is then calculated as the harmonic average between $\eta_{df}$ and $\eta_{ds}$:
-\begin{equation}\label{eq:visco_creep}
-\eta_{cp}=\left(\frac{1}{\eta_{df}}+\frac{1}{\eta_{ds}}\right)^{-1}
-\end{equation}
-The implementation of the non-linear viscous creep viscosity depending on the strain rate is tested by means of the slab detachment benchmark
-(Section \ref{sec:slab}).
-
-Plastic yielding is implemented rescaling $\eta_{cp}$ in order to limit the stress below the yield stress $\sigma_y$ \citep{Thieulot2008,Thieulot2014,Glerum2018},
-obtaining
-\begin{equation}\label{eq:plastic}
-\eta_{pl}=\frac{\sigma_y}{2I_2}
-\end{equation}
-where $\eta_{pl}$ is the plastic viscosity and the yield stress is determined following the Drucker-Prager criterion, such as
-\begin{equation}\label{eq:yield}
-\sigma_y=C\;\cos(\phi)+p\;\sin(\phi)
-\end{equation}
-where $C$ is the cohesion and $\phi$ is the internal friction angle. In case of negative pressure it is imposed equal to 0, so that negative yield stress are
-excluded. The correctness of the non-linear solution in case of plastic viscosity with variable internal friction angle is verified performing indenter and
-brick experiments (Sections \ref{sec:indenter} and \ref{sec:brick}, respectively).
-
-Strain softening is taken into account for both viscous creep and plastic viscosity \citep{Huismans2003,Babeyko2005,Huismans2005,Sobolev2005,Warren2008a} by
-means of the accumulated strain $\epsilon$ memorised by each marker. Softening in the viscous creep determines a linear decrease of $\eta_{vc}$ by means of a
-viscous strain softening factor $W_S$ that increases linearly from $W_{S_0}$ to $W_{S_{\infty}}$ for $\epsilon_{S_0}<\epsilon_S<\epsilon_{S_{\infty}}$ 
-\citep{Warren2008a}. This viscous softening can be related to strain-induced grain size reduction \citep{Warren2008a}. Differently, plastic softening is 
-simulated with a linear decrease of internal friction angle $\phi(\epsilon)$ and cohesion $C(\epsilon)$ in according to
-\begin{eqnarray}
-\label{eq:friction1}\phi(\epsilon)&=&\phi_0+(\phi_{\infty}-\phi_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}\\
-\label{eq:friction2}C(\epsilon)&=&C_0+(C_{\infty}-C_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}
-\end{eqnarray}
-where $\phi_0$, $C_0$ and $\phi_{\infty}$, $C_{\infty}$ are internal friction angle and cohesion for $\epsilon_0$ and $\epsilon_{\infty}$, respectively
-\citep{Huismans2003,Huismans2005,Warren2008a,Thieulot2014}. Plastic softening approximates deformation-induced softening of faults and brittle shear zones
-\citep{Warren2008a}.
-
-Plastic yielding and viscous creep are then combined to obtain a viscoplastic viscosity $\eta_{vp}$ as follows, assuming that they are independent processes
-\citep{Karato2008,Glerum2018}:
-\begin{equation}\label{eq:viscoplastic}
-\eta_{vp}=\min(\eta_{cp},\eta_{pl})
-\end{equation}
-Finally, effective viscosity $\eta_{eff}$ is capped by the minimum and the maximum viscosity ($\eta_{min}$ and $\eta_{max}$, respectively) to avoid extremely
-low or high viscosity \citep{Glerum2018} as follows
-\begin{equation}\label{eq:effective}
-\eta_{eff}=\min(\max(\eta_{vp},\eta_{min}),\eta_{max})
-\end{equation}
-Viscous creep and plastic yielding are non-linear rheologies because of their dependence on the velocity field through pressure and strain rates. Therefore,
-the solution is determined by means of Picard-type iterations, until convergence of the velocity field \citep{Glerum2018}. The convergence is verified at each
-iteration $i$ via the nonlinear residual $\bm{\mathcal{R}}^i$ that can be determined as
-\begin{equation}\label{eq:convergence}
-\bm{\mathcal{R}}^i= \mathbb{K} (\eta_{eff}(\dot{\bm{\epsilon}}^{i-1},p^{i-1}))\cdot \bm{v}^{i-1} - \bm{f}^i
-\end{equation}
-%\begin{equation}\label{eq:convergence}
-%\chi=1-\frac{\langle (\bm{v}^i-\langle\bm{v}^i\rangle)\cdot (\bm{v}^{i+1}-\langle\bm{v}^{i+1}\rangle) \rangle}{\sqrt{\sigma^i\sigma^{i+1}}}
-%\end{equation}
-%Iterations are performed until either convergence ($\chi<tol$) or maximal number of non-linear iterations ($it_{max}$) is reached.
-where $\mathbb{K}$ is the Stiffness matrix and $\bm{f}$ is the right hand side vector \citep{Spiegelman2016,Glerum2018}.  The $L_2$-norm is extracted from
-$\bm{\mathcal{R}}^i$ and it is normalised as follows
-\begin{equation}\label{eq:convergence_res}
-\frac{||\bm{\mathcal{R}}^i||_2}{||\bm{\mathcal{R}}^0||_2}
-\end{equation}
-where $||\bm{\mathcal{R}}^0||_2$ is the $L_2$-norm of the first nonlinear iteration.
-Since often the normalised nonlinear residual drop very quickly over the first 2-3 iterations, also the $L_2$-norm of velocity and pressure residuals are
-calculated as
-\begin{equation}\label{eq:convergence_vel}
-\frac{||\Delta\bm{v}||_2}{||\bm{v}||_2}=\frac{||\bm{v}_i-\bm{v}_{i-1}||_2}{||\bm{v}_{i}||_2}
-\end{equation}
-and
-\begin{equation}\label{eq:convergence_press}
-  \frac{||\Delta p||_2}{||p||_2}=\frac{||p_i-p_{i-1}||_2}{||p_{i}||_2}
-\end{equation}
-respectively. Iterations are performed until either maximal number of non-linear iterations ($it_{max}$) is reached or all the normalised nonlinear residual, 
-the velocity residual and the pressure residual converge under a defined tolerance $tol$.
-
-\section{The energy equation}\label{sec:en_eq}
-The stabilisation of the advection term of the energy equation (Eq. \ref{eq:energy}) needed to avoid possible oscillations in the thermal solution in the case
-when advection dominates over diffusion, is implemented by means of a streamline-upwind Petrov–Galerkin (SUPG) method, for which the advection term is modified
-follows the discussion in \citet{Thieulot2011} and \citet{Thieulot2014} (see Section \ref{sec:advection} for benchmark). The total internal heat production
-$H_{tot}$ in Eq. \ref{eq:energy} includes radiogenic heating $H_r$, shear heating $H_s$ and adiabatic heating $H_a$, where
-\begin{equation}\label{eq:shear}
-H_s=2\eta \dot{\bm{\epsilon}} : \dot{\bm{\epsilon}} = 2\eta(\dot{\epsilon}_{xx}^2+\dot{\epsilon}_{yy}^2+2\dot{\epsilon}_{xy}^2)
-\end{equation}
-and
-\begin{equation}\label{eq:adiabatic}
-H_a=T \alpha \frac{Dp}{Dt} \approx -\alpha T \rho g_y v
-\end{equation}
-The correctness of the implementation and the computation of shear and adiabatic heating is verified performing an exercise with an analytical solution
-(Section \ref{sec:simple_shear}) and Exercise 9.4 in \citet{Gerya2010b} (Section \ref{sec:shear}). Furthermore, mantle convection and thin layer entrainment
-benchmarks (Sections \ref{sec:mantle} and \ref{sec:thin}, respectively) are performed to verify that the code solve correctly Eqs. \ref{eq:momentum_penalty}
-and \ref{eq:energy} in case of a temperature field characterised by an initial perturbation and with a temperature-dependent density. At last, the viscoplastic
-mantle convection benchmark is performed to check also the viscoplasticity in case of a initial perturbed temperature field (Section \ref{sec:visco_mantle}).
-
-\section{Phase transitions and hydration}\label{sec:phase}
-Variations of effective density, specific heat and coefficient of thermal expansion during the evolution of crustal and mantle materials at different
-pressure-temperature ($p$-$T$) conditions are computed by means of version 6.8.6 of Perple\_X software package \citep{Connolly2005}, similarly to the
-implementation described in \citet{Marotta2020}. By default, mineral assemblages and properties of each lithology (oceanic crust, lower and upper continental
-crust, sediments and dry or hydrated mantle) are calculated for temperatures from \SIrange{330}{1600}{\kelvin}, with increments of approximately \SI{4}{\kelvin},
-and pressures from \SIrange{0.1}{30}{\giga\pascal}, with increments of approximately \SI{0.03}{\giga\pascal}, for a total of almost 100000 points.
-
-The introduction of phase transitions produces two main effects that must be taken into account in numerical models: variations in density and release/absorption
-of latent heat \citep{Tackley2010}. Effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into
-account considering effective coefficient of thermal expansion \citep{Christensen1985,Zhong2015}. Similarly, effects of latent heat in the energy equation
-(eq \ref{eq:energy}) are taken into account by considering effective specific heat and coefficient of thermal expansion
-\citep{Christensen1985,Tackley2010,Zhong2015}. Effective density, specific heat and coefficient of thermal expansion are included in Perple\_X files and are
-assumed by each Lagrangian marker in according to its $p$-$T$ conditions.
-
-Hydration processes related to dehydration of subducting lithosphere \citep{Schmidt1998,Liu2007,Faccenda2009,Faccenda2010,Faccenda2014,Rosas2016} strongly
-influence the thermo-mechanics inside the mantle wedge, mainly because of both weakening effects on the mantle rheology and density variations in case of
-mantle serpentinization \citep{Gerya2002,Honda2003,Arcay2005,Roda2010,Regorda2017}. In the case that hydration is switched on, the amount of bound and free
-water is memorised by each marker, following the implementation of \citet{Quinquis2014}; hydration and dehydration processes are related to the amount of bound
-water of each marker and to the maximum amount of water it can transport, i.e. if the amount of bound water exceeds the maximum amount of water, the marker
-dehydrates and releases free water that can hydrates under-saturated markers. Maximum water content of each marker is determined as function of lithology and
-$p$-$T$ conditions and it is calculated using Perple\_X, in the same way than effective density, specific heat and coefficient of thermal expansion. Bound water
-is advected together with the markers, neglecting the effect of bound water diffusion, while free water simply migrates vertically and is not coupled to the
-solid-phase flow of the mantle wedge \citep{Arcay2005,Quinquis2014}. The correctness of the migration scheme of the free water and its absorption into bound
-water by under-saturated markers are verified performing the experiment described in \citet{Quinquis2014} (Sec. \ref{sec:quinquis}).
-
-The mantle viscosity weakening related to the amount of water has been extensively studied \citep{Chopra1981,Hirth2003} and it is implemented in according to
-\citet{Arcay2005} and \citet{Horiuchi2016}, as follows
-\begin{equation}\label{eq:wet_visc}
-\eta_{wet}=\eta_{dry}\left(\left[1-\frac{1}{f_v}\right]\exp\left(-\frac{[OH^-]}{[OH^-]_0}\right)+\frac{1}{f_v}\right)
-\end{equation}
-where $f_v$ is the viscosity reduction factor between dry and wet conditions, $[OH^-]$ is the water content and $[OH^-]_0$ is a reference water content set to
-620 ppm (0.062 wt.\%) \citep{Arcay2005}. By default, $f_v$ is set to 100, in according to \citet{Arcay2005} and \citet{Horiuchi2016}.
-
-\section{Mantle melting}\label{sec:melt}
-Melt fraction in the mantle depends on temperature ($T$, in \textdegree C), pressure ($p$, in \si{\giga\pascal}) and the water content in the melt ($X_{H_2O}$, in wt.\%).
-The determination of the melt fraction $M$ is implemented as explained by \citet{Katz2003}, according to the successive modification by \citet{Langmuir2006} and
-\citet{Kelley2010}, as follows
-\begin{eqnarray}\label{eq:melt_fr}
-M(p,T,X_{H_2O})&=&-T+T_{sol}+(x_t\cdot ln(p+y_t))\cdot X_{H_2O}-\nonumber\\
-&&+K_m\left(\frac{X_{H_2O}^{bulk}}{D_{H_2O}(1-X_{H_2O})+X_{H_2O}}\right)^{\gamma}
-\end{eqnarray}
-where $T_{sol}$ is the temperature of the dry solidus, $D_{H_2O}$ is the partition coefficient and $x_t\cdot log(p+y_t)$ indicates the pressure dependence of
-the melting curve \citep{Kelley2010}. $x_t$, $y_t$, $K_m$ and $\gamma$ are parameters chosen in according to \citet{Kelley2010}.
-However, $X_{H_2O}$ depends on the melt fraction as
-\begin{equation}\label{eq:melt_wt}
-X_{H_2O}(M)=\frac{X_{H_2O}^{bulk}}{D_{H_2O}+M(1-D_{H_2O})}
-\end{equation}
-and a numerical solution can be found using a root-finder method \citep{Katz2003,Wang2016}. Furthermore, the content of water in the melt is limited by the
-pressure-dependent saturation concentration of water in the melt, determined as
-\begin{equation}\label{eq:melt_max}
-X_{H_2O}^{sat}=\chi_1 p^{\lambda_m}+\chi_2 p
-\end{equation}
-where $\chi_1$, $\chi_2$ and $\lambda_m$ are parameters chosen according to \citet{Katz2003}, \citet{Langmuir2006} and \citet{Kelley2010}. The correctness in
-the determination of melt fractions and water contents in the melt are tested comparing model results with experimental curves obtained by \citet{Katz2003}
-(Sec. \ref{sec:katz}).
-
-In case of melting, viscous creep viscosity ($\eta_{vp}$) is modified according to \citet{Wang2016}, as
-\begin{equation}\label{eq:melt_visc}
-\eta_{melt}=\eta_{vp}\cdot \exp(\alpha_m M)
-\end{equation}
-where $\alpha_m$ are melt fraction factors for dislocation and diffusion creep, chosen in according to \citet{Wang2016}.
-
-Effective density ($\rho_{eff}$) for partially molten rocks is calculated as
-\begin{equation}\label{eq:melt_dens}
-\rho_{eff}=\rho_s(1-M)+\rho_m M
-\end{equation}
-where $\rho_s$ and $\rho_m$ are the densities of the solid and the molten rock, respectively \citep{Gerya2004b,Gerya2007,Wang2016}. The density of the
-solid rock is extracted by Perple\_X, while the density of the molten portion is calculated as
-\begin{equation}\label{eq:melt_dens_M}
-\rho_m(p,T)=\rho_0[1-\alpha(T-T_0)][1+\beta(p-p_0)]
-\end{equation}
-where $\rho_0$ is the density at the reference temperature ($T_0$) and pressure($p_0$), and $\alpha$ and $\beta$ are the thermal expansion and compressibility
-coefficients, respectively \citep{Gerya2004b,Gerya2007,Wang2016}.
-
-As for the phase transitions, effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into account
-considering effective coefficient of thermal expansion. Similarly, latent heat due to melting/crystallisation can be implicitly included in the energy equation
-(Eq. \ref{eq:energy}) by considering effective specific heat ($C_{p_{eff}}$) and thermal expansion ($\alpha_{eff}$) of the partially molten rocks, in according to
-\begin{eqnarray}
-\label{eq:cp_effective}C_{p_{eff}}&=&C_p+H_L\left(\frac{\partial M}{\partial T}\right)_p \\
-\label{eq:alpha_effective}\alpha_{eff}&=&\alpha+\rho\frac{H_L}{T}\left(\frac{\partial M}{\partial p}\right)_T
-\end{eqnarray}
-where $C_p$ and $\alpha$ are the specific heat and the coefficient of thermal expansion of the solid rock, respectively, and $H_L$ is the latent heat of the
-molten rock \citep{Gerya2004b,Gerya2007,Tackley2010}.
-
-\section{Benchmarks}\label{sec:benchmark}
-\subsection{Error measurements}\label{sec:error}
-In order to determine the accuracy of velocity and pressure field of the benchmarks, the $L_2$-norm is computed by numerical integration on the Gauss-Legendre
-quadrature points. $L_2$-norm for pressure and velocity errors can be evaluated as
-\begin{eqnarray}
-\textrm{err}_p&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|p^n_i(r_q)-p^a_i(r_q)|^2\right)w_q|J_q|}\nonumber \\
-\textrm{err}_v&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|u^n_i(r_q)-u^a_i(r_q)|^2+|v^n_i(r_q)-v^a_i(r_q)|^2\right)w_q|J_q|}\nonumber
-\end{eqnarray}
-respectively, where $n_e$ is the number of elements, $n_q$ is the number of quadrature points per element, $p^n_i(r_q)$ and $p^a_i(r_q)$ are the numerical and
-analytical pressure, respectively, in each quadrature point $q$, $w_q$  and $J_q$ are the weight and the Jacobian at the quadrature point
-$q$, $u^n_i(r_q)$, $v^n_i(r_q)$, $u^a_i(r_q)$ and $v^a_i(r_q)$ are the numerical and analytical velocities, respectively, in each quadrature point $q$.
-
-Other quantities used as comparison with original benchmarks are the root-mean-square velocity over the whole domain and over the surface
-\begin{eqnarray}
-v_{\textrm{rms}}&=&\sqrt{\int_{\Omega}|\bm{v}|^2 d\Omega}\nonumber \\
-v_{\textrm{rms}}^{top}&=&\sqrt{\int_0^1 u^2 \bigg|_{y=1} dx}\nonumber
-\end{eqnarray}
-respectively, top and bottom Nusselt numbers
-\[\textrm{Nu}^{top/bottom}=-\int_0^1\frac{\partial T}{\partial y} \bigg|_{y=1/y=0} dx\]
-the average rate of work done against gravity
-\[\left\langle W \right\rangle=\int_{\Omega}T u_y d\Omega\]
-the average rate of viscous dissipation
-\[\left\langle \Phi \right\rangle=\int_{\Omega}\tau_{ij}\dot{\epsilon}_{ij} d\Omega\]
-and the percentage error between $\left\langle W \right\rangle$ and $\left\langle \Phi \right\rangle/Ra$
-\[\delta=\frac{\left|\left\langle W \right\rangle -\frac{\left\langle \Phi \right\rangle}{Ra} \right|}{\max \left(\left\langle W \right\rangle,
-\frac{\left\langle \Phi \right\rangle}{Ra}\right)}\times100\]
-
-\subsection{Stokes flow}\label{sec:stokes}
-The problem consists of determining velocity field $(u,v)$ and pressure $p$ in case of a manufactured solution with prescribed body forces such as
-\begin{eqnarray}
-b_1&=&(12-24y)x^4+(-24+48y)x^3+(-48y+72y^2-48y^3+12)x^2+\nonumber \\
-&&+(-2+24y-72y^2+48y^3)x+1-4y+12y^2-8y^3\nonumber \\
-b_2&=&(8-48y+48y^2)x^3+(-12+72y-72y^2)x^2+(4-24y+48y^2-48y^3+\nonumber \\
-&&+24y^4)x-12y^2+24y^3-12y^4\nonumber
-\end{eqnarray}
-for which the exact solution is:
-\begin{eqnarray}
-u(x,y)&=&x^2(1-x)^2(2y-6y^2+4y^3)\nonumber \\
-v(x,y)&=&-y^2(1-y)^2(2x-6x^2+4x^3)\nonumber \\
-p(x,y)&=&x(1-x)-\frac{1}{6}\nonumber
-\end{eqnarray}
-The domain is a unit square a constant viscosity ($\eta=1$) and the penalty parameter is set to $\lambda=10^7$. Velocity boundary conditions are set to no slip
-$(\bm{v}=\bm{0})$ on all boundaries. The problem is performed for different grid resolution between \num{8x8} and \num{1024x1024} elements.
-
-The errors between the analytical solution and the numerical prediction of the velocity and the pressure field are calculated as in Section \ref{sec:error}.
-Fig. \ref{fig:errors}a shows that both velocity and pressure field converge to the exact solution with the decrease of the element size, following the
-theoretical convergence rate. The convergence of pressure, in contrast with the observation by \citet{Donea2003} for $Q_1 \times P_0$ elements, indicate the
-effectiveness of the smoothing procedure. Smoothed pressure field for grid resolution of \num{128x128} elements is shown in Fig. \ref{fig:errors}b (continuous
-red line), in comparison with the analytical solution (dashed blue line). Solve times and memory usage needed to generate the solution are shown in
-Fig. \ref{fig:MUMPS}. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Solver/Stokes_Flow}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/errors.pdf}
-\caption{Velocity and pressure error for the Stokes flow experiment between generated and analytical solution as a function of element size (panel a),
-and comparison between smoothed pressure and analytical solution as function of $x$ coordinate for a grid resolution of \num{128x128} elements (panel b).}
-\label{fig:errors}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/MUMPS.pdf}
-\caption{Panel a) Analysis, factorisation, solution and total solve times (panel a) and factorisation memory usage (panel b) as a function of the total number
-of degrees of freedom for the Stokes flow experiment.}
-\label{fig:MUMPS}
-\end{figure}
-
-\subsection{Markers advection}\label{sec:runge}
-The 2nd-order Runge-Kutta advection scheme in space is tested by means of the Zalesak disk test \citep{Zalesak1979,Thieulot2014}. The benchmark is performed
-in a unit square domain with a grid resolution of \num{32x32} elements and 1000000 markers and values of Courant number between 0.25 and 1.
-At $t=0$, the disk is centred at position $(0.5;0.75)$ with a radius $R=0.15$ and has a vertical fissure 0.05 wide and 0.2 high (orange in Fig. \ref{fig:runge})
-and with the velocity field prescribed in the entire domain as
-\begin{eqnarray}
-u(x,y)&=&2\pi\left(y-\frac{L_x}{2}\right)\nonumber \\
-v(x,y)&=&-2\pi\left(x-\frac{L_x}{2}\right)\nonumber
-\end{eqnarray}
-the markers are back to their initial location after a $2\pi$ rotation ($t=1$). The clockwise rotation of the disk at $t=0.25$, 0.5 and 0.75 
-is shown in Fig. \ref{fig:runge} (red, blue and green, respectively). The distance from the centre of 1 marker is calculated throughout the rotation to
-evaluate the error for different values of Courant number. As expected, the error increase with the increasing of the time step, as shown in 
-Fig. \ref{fig:runge_err}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/Zalesak_Disk}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=200px]{./Figures/Zalesak.pdf}
-\caption{Position of the Zalesak disk at $t=0$, 0.25, 0.5 and 0.75 (orange, red, blue and green, respectively) for the case with Courant number of 1.}
-\label{fig:runge}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/zal_marker.pdf}
-\caption{Distance from the centre as function of time for values of Courant number of  of 0.25, 0.3, 0.5, 0.75 and 1 (green, orange, blue, black and red,
-respectively). R indicates the distance from the centre considering a perfect circle trajectory.}
-\label{fig:runge_err}
-\end{figure}
-
-\subsection{Conservative Velocity Interpolation (CVI)}\label{sec:CVI}
-The CVI correction is checked by means of the Stokes flow experiment by \citet{Donea2003}, as presented in \ref{sec:stokes}. The advection of Lagrangian
-markers is performed using either a 2nd-order or a 4th-order Runge-Kutta scheme with CFL=0.5 and an initial random distribution of 25 markers per element.
-Fig. \ref{fig:CVI} shows the minimum and maximum values of markers per element throughout the experiment with and without the CVI correction and using
-different Runge-Kutta schemes. The introduction of the CVI correction allows the
-tests to run for times longer than $t=2000$ without show empty elements, while tests without the CVI correction stop at approximately $t=600$, when at least
-one element goes empty. Fig. \ref{fig:CVI_time} shows the different number of markers per element and the markers distribution at $t=600$ obtained without and
-with the CVI correction (panels b and c, respectively) in case of a 2nd-order Runge-Kutta scheme. All data can be found at
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/CVI}.
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/CVI.pdf}
-\caption{Maximum and minimum number of markers per element throughout the experiment, with (green and orange lines) or without (purple and blue lines)
-the CVI correction and using either a 2nd-order (blue and orange lines) or a 4th-order (purple and green lines) Runge-Kutta scheme.}
-\label{fig:CVI}
-\end{figure}
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/CVI_time.pdf}
-\caption{Number of markers per element and markers distribution through time using 2nd.order Runge-Kutta scheme.
-Panels a: number of markers per element (panel a1) and distribution of the markers (panel a2) after the first
-iteration; panels b and c: comparison between number of markers per element (panels b1 and c1) and markers distribution (panels b2 and c2) without and with
-the CVI correction at $t=600$; panels d: number of markers per element (panel d1) and markers distribution (panel d2) in case of CVI correction at $t=2000$.}
-\label{fig:CVI_time}
-\end{figure}
-
-\subsection{Poiseuille flow}\label{sec:poiseuille}
-The domain is a rectangle with $L_x=2$ and $L_y=1$ and constant density and viscosity ($\rho=1$ and $\eta=1$), gravity acceleration $\bm{g}=\bm{0}$ and penalty
-parameter $\lambda=10^8$. The grid is composed by \num{40x20} elements. Velocity boundary conditions are set to no slip $(\bm{v}=\bm{0})$ at the top and the
-bottom, and a parabolic profile is imposed on the sides, with $u=y(1-y)$ and $v=0$. The analytical solution is then given by:
-\begin{eqnarray}
-u(x,y)&=&y(1-y)\nonumber \\
-v(x,y,)&=&0\nonumber \\
-p(x,y)&=&2\eta\left(\frac{L_x}{2}-x\right)\nonumber
-\end{eqnarray}
-
-The velocity field predicted by the model follows the expected parabolic profile (Fig. \ref{fig:poiseuille}a and Fig. \ref{fig:poi_plot}a) and, in case of the
-classic penalty method with no iterations, the pressure is clearly related to the divergence by means of the penalty parameter (Fig. \ref{fig:poiseuille}b and
-c and Fig. \ref{fig:poi_plot}b). Fig. \ref{fig:poiseuille}d shows that one Uzawa iteration is sufficient to bring the divergence down to \num{e-15}, with no
-correlation with the pressure field. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Poiseuille%20Flow}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Poiseuille.pdf}
-\caption{Velocity field (panel a), pressure (panel b) and divergence velocity of a Poiseuille flow in case of the classic penalty method (no iterations) and
-after one Uzawa iteration (panel c and d, respectively).}
-\label{fig:poiseuille}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Analytical.pdf}
-\caption{Velocity (panel a) and pressure (panel b) field predicted by the model for a Poiseuille flow with respect with their analytical solutions. The
-velocity field is plotted as function of the vertical coordinate in $L_x/2$ and the pressure is plotted as function of the x coordinate in $L_y/2$.}
-\label{fig:poi_plot}
-\end{figure}
-
-\subsection{Instantaneous 2D sphere}\label{sec:ist_sphere}
-The domain is a unit square with gravity $\bm{g}=(0,-1)$. The fluid has constant density and viscosity ($\rho_f=1$ and $\eta_f=1$). The sphere is in the middle
-of the domain with a radius R = 0.123456798 and has constant density and viscosity ($\rho_s=10^{-2}+\rho_f$ and $\eta_s=10^3 \cdot \eta_f$).
-We distinguish three different types of velocity boundary conditions:
-\begin{enumerate}
-\item FS: free slip conditions on all sides.
-\item NS: no slip conditions on all sides.
-\item OT: free slip conditions on the sides and the bottom, and open boundary at the top.
-\end{enumerate}
-All three cases are performed for different grid resolutions, between \num{16x16} and \num{512x512} elements with 50 randomly distributed markers per element,
-and different average schemes for the viscosity (harmonic, geometric and arithmetic). The velocity in the centre of the sphere $\bm{v}(0.5;0.5)$), minimum and
-maximum velocities ($u_{min}$, $u_{max}$, $v_{min}$, $v_{max}$) and pressures ($p_{min}$, $p_{max}$) on the entire domain, $v_{\textrm{rms}}$ and average pressure on the
-entire domain ($p_{avg}$) are compared with solutions generated by ASPECT \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and all results
-can be found at \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere2D} and 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Instantaneous_2D_Sphere}. Results in terms of the $v_{\textrm{rms}}$ for NS model 
-are also shown in Fig. \ref{fig:inst_sphere}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_NS.pdf}
-\caption{$v_{\textrm{rms}}$ as function of element size for the instantaneous 2D sphere experiment in case of no slip boundary conditions and with different average
-schemes for the viscosity. Results are compared with results obtained by other numerical codes.}
-\label{fig:inst_sphere}
-\end{figure}
-
-\subsection{Rayleigh-Taylor instability}\label{sec:rayleigh}
-This problem has been originally presented by \citet{vanKeken1997} and here it is performed as the isoviscous case (Case 1a). The domain has $L_x=0.9142$ and
-$L_y=1$ and gravity $\bm{g}=(0,-1)$. Two fluids with same constant viscosities ($\eta_1=\eta_2=1$) and different densities ($\rho_1=1000$ and $\rho_2=1010$),
-with the lighter fluid at the bottom. The initial interface between the fluids is given by $y(x)=0.2+0.02 \cos \left(\frac{\pi x}{L_x}\right)$. The experiment
-is performed with different grid sizes (\num{50x50}, \num{80x80}, \num{100x100} and \num{256x256}). A total of 1960000 markers are randomly distributed at the
-beginning of the simulation. Velocity boundary conditions are set to no slip at the top and the bottom, and to free slip at the sides of the domain.
-
-$v_{\textrm{rms}}$ as function of time is reported in Fig. \ref{fig:RT}, matching well with results shown by \citet{vanKeken1997}, \citet{Tackley2003} and
-\citet{Thieulot2014}. Fig. \ref{fig:rayleigh} shows the evolution of the experiment at different time steps. All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Rayleigh_Taylor_experiment/ISOVISCOUS}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/RT.pdf}
-\caption{$v_{\textrm{rms}}$ of the Rayleigh-Taylor experiment as function of time for different resolution of the grid.}
-\label{fig:RT}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=300px]{./Figures/Rayleigh.pdf}
-\caption{Evolution of the Rayleigh-Taylor experiment for a grid of \num{256x256} elements at $t=0, 500, 1000$ and $2000$ (panels a, b, c and d, respectively).}
-\label{fig:rayleigh}
-\end{figure}
-
-\subsection{Falling blocks}\label{sec:block}
-This benchmark is described as proposed by \citet{Gerya2003a}, \citet{Gerya2010b} and \citet{Thieulot2011}. The domain is square with $L_x=L_y=$ \SI{500}{\km}
-and the grid is composed by \num{50x50} elements with 25 markers in each element. The block is initially centred at ($x=$ \SI{250}{\km}; $y=$ \SI{400}{\km})
-and has a size of \num{100x100} km (Fig. \ref{fig:block}a). The fluid surrounding the block has $\eta_f=$ \SI{e21}{\pascal\per\s} and $\rho_f=$
-\SI{3200}{\kg\per\cubic\m}. The benchmark tests with different viscosities of the block, with $\eta_b$ from \SIrange{e15}{e27}{\pascal\per\s}. In each
-experiment also the density $\rho_b$ of the block varies from \SIrange{3220}{9900}{\kg\per\cubic\m}. Velocity boundary conditions are set to free slip
-conditions on all sides of the domain. Density distributions at $t=$ \SI{20}{\mega\year} for $\rho_b=$ \SI{3300}{\kg\per\cubic\m} and different $\eta_b$
-are plotted in Fig. \ref{fig:block}b-f, showing that the code correctly preserve the block geometry in case of large viscosity contrast, with a block stiffer
-than the surrounding fluid (Fig. \ref{fig:block}f).
-
-The velocity in the centre of the falling block at $t=0$ is measured for all experiments and, since the velocity should increase with density contrast, the
-quantity $\bm{v}/(\rho_b-\rho_f)$ is plotted as function of the viscosity contrast. All results perfectly match with those from \citet{Gerya2010b} and they
-line up on a single curve, demonstrating that the code can correctly deal with large viscosity and density contrasts (Fig. \ref{fig:falling}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Falling%20blocks}.
-
-\begin{figure}
-\noindent\includegraphics[width=300px]{./Figures/Block.pdf}
-\centering
-\caption{Density evolution of the falling block experiment at $t=$ \SI{0}{\mega\year} (panel a) and $t=$ \SI{20}{\mega\year} for different viscosities of the
-block (panels b-f).}
-\label{fig:block}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Falling.pdf}
-\caption{Initial velocity relative to the density contrast at the centre of the falling block as function of the viscosity contrast between the block and the
-surrounding fluid.}
-\label{fig:falling}
-\end{figure}
-
-\subsection{2D time-dependent Stokes sphere with deformable free surface}\label{sec:sphere}
-This experiment is performed in a unit square domain with the gravity acceleration fixed to $g_y=-1$. The sphere has $\rho_s=2$ and $\eta_s=10^3$ and is
-initially centred at ($0.5;0.6$) with radius $R=0.123456789$. The fluid surrounding the sphere has $\rho_f=1$ and $\eta_f=1$ and occupies the domain for
-$y\leq 0.75$, while for $y>0.75$ the air has $\rho_a=0$ and $\eta_a=10^{-3}$. Velocity boundary conditions are set to free slip on all sides. The Courant
-number is set to 0.25. The experiments run for 200 s using grid resolutions from \num{150x150} to \num{512x512} elements, with 25 randomly distributed markers
-per element, and for different average schemes for the viscosity. The interface between the fluid and the air is tracked by means of the markers chain. Results
-in terms of velocity and pressure fields and topography variations are compared with results obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and they can be found at
-\url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere_fs2D} and
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Time_dependent_sphere}. Results in terms of the $v_{\textrm{rms}}$ in case of 
-an arithmetic average are also shown in Fig. \ref{fig:inst_spherefs}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_arithm.pdf}
-\caption{$v_{\textrm{rms}}$ for different grid resolutions as function of time for the 2D time-dependent Stokes sphere experiment in case of an arithmetic average.
-Results are compared with results obtained by ASPECT.}
-\label{fig:inst_spherefs}
-\end{figure}
-
-\subsection{Free surface stabilisation}\label{sec:stab}
-The benchmark is performed as discussed in \citet{Kaus2010a} and \citet{Thieulot2014}. The domain is square with $L_x=L_y=$ \SI{500}{\km} and a grid resolution
-of \num{200x200} elements, each of them containing 16 markers. A buoyant fluid with $\rho_1=$ \SI{3200}{\kg\per\cubic\metre} and $\eta_1=$ \SI{e20}{\pascal\s}
-is overlain by a denser fluid with $\rho_2=$ \SI{3300}{\kg\per\cubic\metre} and $\eta_2=$ \SI{e21}{\pascal\s}. The initial interface between the fluids has an
-initial sinusoidal shape of 5 km amplitude. Velocity boundary conditions are set to free slip conditions at the sides and no slip at the bottom, while the top
-is free surface. The experiment is performed with various fixed time steps, without the use of the Courant number. The vertical position of the free surface at
-$x=L_x$ is tracked for each simulation, with or without the stabilisation algorithm.
-
-The results show an instability (drunken sailor effect) increasing the time step in case that the stabilisation algorithm is not activated
-($dt=$ \SI{4200}{\year}, green line in Fig. \ref{fig:stab}), as already observed by \citet{Kaus2010a} and \citet{Thieulot2014}. Activating the algorithm the
-instability is fixed and the simulation remains stable using time step up to \SI{50000}{\year} (red line in Fig. \ref{fig:stab}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Stabilisation_algorithm}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Kaus_random.pdf}
-\caption{Evolution of the y coordinate in $x=L_x$ as function of time for different time steps, with and without the stabilisation algorithm.}
-\label{fig:stab}
-\end{figure}
-
-\subsection{Topography relaxation}\label{sec:crameri}
-This experiment is performed as in \citet{Crameri2012} using free surface and sticky air (Case 1). The domain is rectangular with $L_x=$ \SI{2800}{\km} and
-$L_y=$ \SI{700}{\km} in case of free surface and $L_y=$ \SI{800}{\km} in case of sticky air ($h_{st}=$ \SI{100}{\km}). The grid is composed by \num{256x64}
-elements for the free surface \num{560x320} elements for the sticky air. For both cases, there is a mantle of 600 km thickness with $\eta_m=$ \SI{e21}{\pascal\s}
-overlain by a highly viscous cosine-shaped lithosphere with thickness between 93 and 107 km and $\eta_l=$ \SI{e23}{\pascal\s}. Both mediums have $\rho=$
-\SI{3300}{\kg\per\cubic\metre}. In case of sticky air experiment, the air has $\rho_a=$ \SI{0}{\kg\per\cubic\metre} and $\eta_a=$ \SI{e18}{\pascal\s}. The
-gravity acceleration is set to $g_y=$ \SI{-10}{\metre\per\square\s}. Velocity boundary conditions are set to no slip at the bottom and free slip at the sides
-of the domain. In case of sticky air, velocities on top are set to free slip conditions.
-
-The maximum topography as function of time ($h(t)$) can be analytically derived in according to \[h(t)=h_0 \exp(\gamma t)\]
-where $h_0=7$ km is the initial topography, $\gamma=\num{-0.2139e-11}$ is the characteristic relaxation rate and $t$ is the time (black line in Fig.
-\ref{fig:crameri}). Maximum topography at the characteristic relaxation time $t_{rlx}=$ \SI{14.825}{\kilo\year} can be found to be $h_{rlx}=$ \SI{2576}{\m}.
-The results show that a viscosity of the air of \SI{e18}{\pascal\s} is needed in case of $h_{st}=$ \SI{100}{\km} to correctly describe the topography relaxation
-for this problem, as pointed out by \citet{Crameri2012} (purple line in Fig. \ref{fig:crameri}). The case with a true free surface well follows the analytical
-solution (orange line in Fig. \ref{fig:crameri}). In particular, at the relaxation time in case of the true free surface predicts a maximum topography of
-\SI{2570}{\m}, with an error of \SI{6}{\m} with respect to the analytical solution. Results are also compared with results obtained with a free surface in
-UNDERWORLD (\num{256x64} elements) and SULEC (\num{401x201} elements) (with courtesy from \citealp{Crameri2012}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Topography_relaxation}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Crameri.pdf}
-\caption{Maximum topography as function of time for the Crameri benchmark (coloured lines) in comparison with the analytical solution (black line) and results
-obtained with UNDERWORLD (yellow line) and SULEC (blue line).}
-\label{fig:crameri}
-\end{figure}
-
-\subsection{Spontaneous subduction}\label{sec:subduction}
-This experiment is performed as presented by \citet{Schmeling2008}, in case of both a sticky air and a true free surface and using different average schemes
-for the viscosities. The domain is rectangular with $L_x=$ \SI{3000}{\km} and $L_y=$ \SI{700}{\km}, a grid resolution of \num{256x64} elements with 50 markers
-per element and Courant number of 0.01. At the beginning of the simulation, a 100 km-thick lithospheric layer with $\rho_m=$ \SI{3300}{\kg\per\cubic\metre}
-and $\eta_m=$ \SI{e23}{\pascal\s} is located at the top of the domain for $x$ between 1000 and 3000 km, while the asthenospheric mantle has $\rho_m=$
-\SI{3200}{\pascal\s} and $\eta_m=$ \SI{e21}{\pascal\s}. In addition, a 200 km-depth lithospheric slab is already subducted in the mantle in order to have
-a spontaneous subduction. In case of sticky air, $L_y=750$ km and a 50 km-thick air layer with $\eta_a$ from \SIrange{e19}{e21}{\pascal\s} overlie the mantle.
-Velocity boundary conditions are set to free slip on the sides and on the bottom of the domain. In case of sticky air, velocities on the top boundary are set
-to free slip conditions as well.
-
-The maximum depth of the slab is tracked for all the simulations (Fig. \ref{fig:subduction}) and they are compared with results obtained by
-\citet{Schmeling2008} with models with regular grids and comparable grid resolutions (rectangles in Fig. \ref{fig:subduction}). Low viscous sticky air models
-(i.e., \SI{e19}{\pascal\s}) enlighten a strong dependence of the sinking velocity with the chosen average scheme (dashed lines in Fig. \ref{fig:subduction}),
-while high viscous sticky air models (i.e., \SI{e21}{\pascal\s}) show a higher resistance at the trench, underestimating the correct solution (dotted lines in
-Fig. \ref{fig:subduction}). A high resistance at the trench can be also observed in case of a true free surface for the low resolution grid used (continuous
-lines in Fig. \ref{fig:subduction}), as pointed out by \citet{Schmeling2008}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Spontaneous_subduction}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Subduction.pdf}
-\caption{Maximum depth of the slab as function of time for all the simulations of the \citet{Schmeling2008} benchmark, in case of harmonic, geometric and
-arithmetic means (red, green and blue lines, respectively) using a sticky air or a true free surface (discontinuous and continuous lines, respectively).
-Red, green and blue rectangular areas indicate the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{500}{\km} in case of sticky air,
-while the grey rectangular area indicates the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{600}{\km} in case of true free surface.}
-\label{fig:subduction}
-\end{figure}
-
-\subsection{The slab detachment}\label{sec:slab}
-The slab detachment benchmark is performed as described by \citet{Schmalholz2011} and \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{1000}{\km}
-and $L_y=$ \SI{660}{\km} and a grid resolution of \num{256x256} elements. A non-linear viscous T-shaped layer with $\rho_l=$ \SI{3300}{\kg\per\cubic\metre} is
-placed at the top of the domain and surrounded by a linear viscous fluid with $\rho_f=$ \SI{3150}{\kg\per\cubic\m}. The top layer is 80 km-thick and a 250
-km-long and 80 km-wide slab is placed at $x=L_x/2$ (Fig. \ref{fig:slab}a). The effective viscosity of the top layer is given by
-\[\eta_{eff}=\eta_0 I_2^{\frac{(1-n)}{n}}\]
-with $\eta_0=$ \SI{4.75e11}{\pascal\s} and $n=4$, while the surrounding fluid has $\eta_f=$ \SI{e21}{\pascal\s}. The effective viscosity is capped between
-$\eta_{min}=$ \SI{e21}{\pascal\s} and $\eta_{max}=$ \SI{e25}{\pascal\s}. Velocity boundary conditions are set to free slip at the top and the bottom, and to
-no slip at the sides of the domain.
-
-The sinking of the slab determines a decrease of the effective viscosity in the T-shaped layer during the first part of the evolution (Fig. \ref{fig:slab}a),
-as consequence of local strain rates, while effective viscosities increase up to $\eta_{max}$ in the top layer after approximately \SI{20}{\mega\year}, when
-necking is complete (Fig. \ref{fig:slab}b).
-
-The width of necking is tracked by means of markers position at the side of the slab. Necking width and time are normalised in function of the initial slab
-(\SI{80}{\km}) and the characteristic time ($t_c=$ \SI{7.1158e14}{\s}) \citep{Schmalholz2011,Glerum2018}. The results in case of arithmetic, geometric and
-harmonic mean of the viscosity are shown in Fig. \ref{fig:necking}, compared with results from \citet{Glerum2018} for which an infinity norm mean was used.
-The influence of the average schemes agrees with results shown in Section \ref{sec:subduction}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Slab_detachment}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Slab.pdf}
-\caption{Effective viscosity for the slab detachment benchmark at the beginning of the evolution (panel a) and when necking is complete (panel b).}
-\label{fig:slab}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Necking.pdf}
-\caption{Normalised width of the necking of the slab detachment benchmark as function of normalised time.}
-\label{fig:necking}
-\end{figure}
-
-\subsection{The indenter experiment}\label{sec:indenter}
-The indenter benchmark simulate a rigid punch on a purely plastic von Mises material, for which exists an analytical solution
-\citep{Thieulot2008,Thieulot2014,Glerum2018}. The domain is a unit square with a grid resolution of \num{256x256} elements. The gravity acceleration is set
-to $\bm{g}=\bm{0}$ and effective viscosity is capped using $\eta_{min}=10^{-4}$ and $\eta_{max}=10^3$. The tolerance for the convergence of the non-linear
-solution is $tol=10^{-9}$, with a maximum number of non-linear iterations set to 500. The medium in the domain has $\rho=0.01$, an initial viscosity
-$\eta_0=10$, cohesion $C=1$ and an angle of internal friction $\phi=0$°. Velocity boundary conditions are set to no slip at the bottom and free slip at the
-sides of the domain. The top of the domain is open with the exception of the central portion (punch area) with width $w_p=0.125$, where $v=-1.05$ and $u$ is
-fixed either to 0 (rough punch experiment) or free (smooth punch experiment).
-
-Fig. \ref{fig:indenter} show the results in terms of viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure
-(panels g and h) for both experiments. The analytical solution indicate that pressure at ($0.5;1$) and ($0.5 \pm w_p;1$)is $p=\pi +1$ and $p=1$, respectively
-(black and red lines in Fig. \ref{fig:smooth_p}). Fig. \ref{fig:smooth_p} shows a clear improvement in the pressure solution when passing from a rough to a
-smooth punch (green and blue lines in Fig. \ref{fig:smooth_p}, respectively), as previously observed by \citet{Thieulot2014} and \citet{Glerum2018}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Indenter}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Indenter.pdf}
-\caption{Viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure (panels g and h) fields for rough (left column) and
-smooth (right column) punch experiments.}
-\label{fig:indenter}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Smooth_Rough.pdf}
-\caption{Nodal pressure in function of the x coordinate at $y=1$ for rough (green line) and smooth (blue line) punch experiment. Black and red lines indicate
-the analytical solution at $x=0.5$ and $x=0.5 \pm w_p$, respectively.}
-\label{fig:smooth_p}
-\end{figure}
-
-\subsection{The brick experiment}\label{sec:brick}
-An instantaneous version of the brick benchmark is performed to verify the correctness of the pressure-dependent plasticity for different angle of internal
-friction, as proposed by \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{40}{\km} and $L_y=$ \SI{10}{\km} and a grid resolution of \num{512x128}
-elements. The gravity acceleration is set to $g_y=$ \SI{-10}{\m\square\s} and effective viscosity is capped using $\eta_{min}=$ \SI{e19}{\pascal\s} and
-$\eta_{max}=$ \SI{e26}{\pascal\s}. The tolerance for the convergence of the non-linear solution is $tol=10^{-7}$, with a maximum number of non-linear iterations
-set to 1000. A 800 m-wide and 400 m-height inclusion with $\eta_b=$ \SI{e20}{\pascal\s} and $\rho_b=$ \SI{2700}{\kg\per\cubic\m} is placed at the bottom of the
-domain at $x=L_x/2$. The inclusion is surrounded by a non-linear viscous medium with $\rho_m=$ \SI{2700}{\kg\per\cubic\m}, an initial viscosity of
-\SI{e23}{\pascal\s}, a linear viscous viscosity of \SI{e25}{\pascal\s} and a cohesion of \SI{40}{\mega\pascal}. Velocities are set to free slip conditions at
-the bottom of the domain and the top is open. Velocities on sides of the domain are fixed to $u=$ \SI{+-2e-11}{\m\per\s} and $v=0$. The experiment is performed
-in compressional and extensional contexts with an angle of internal friction $\phi$ of the non-linear medium variable between 0° and 30°.
-
-As expected, two shear bands stem from the inclusion with variable angles in relation to both the dynamics context and the internal friction angle. Shear band
-angles formed at 45° for $\phi=0$° in both extensional and compressional contexts (Fig. \ref{fig:brick_beam}a), while they are different in case of $\phi\neq0$°.
-Shear band angles in case of $\phi=20$° are shown for compression and extension in Fig. \ref{fig:brick_beam}b  and c, respectively. Values of shear band angles
-as function of internal friction angles from 0° to 30° are extracted at different depths and minimum and maximum values are plotted in Fig. \ref{fig:brick},
-in comparison with theoretical Roscoe, Arthur and Coulomb shear band angles. For all tests the tolerance for nonlinear convergence is defined $tol=10^{-7}$
-and none converges before the maximum number of iterations ($it_{max}=1000$) is reached. However, tests with internal friction angles up to 15° show a constant
-decrease of the velocity residuals, which can not be observed for for tests with higher internal friction angles (Fig. \ref{fig:convergence}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Brick_experiment}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Brick_Beam.pdf}
-\caption{Shear band angles predicted for the brick experiment for $\phi=0$° (panel a) and for $\phi=20$° in case of compression and extension (panels b and c,
-respectively).}
-\label{fig:brick_beam}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Brick.pdf}
-\caption{Shear band angles predicted for the brick experiment in case of compressional and extensional contexts as function of different internal angle of
-friction (continuous black line and red dots), compared with theoretical Roscoe, Artur and Coulomb angles (discontinuous black lines). Red lines indicate the
-range of angles calculated at different depths.}
-\label{fig:brick}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \noindent\includegraphics[width=500px]{./Figures/convergence.pdf}
-  \caption{Convergence rates for all tests. Normalised residuals (left column) are compared with velocity residuals (right column) for extensional (first row)
-  and compressional (second row) settings.}
-  \label{fig:convergence}
-  \end{figure}
-
-\subsection{Advection stabilisation}\label{sec:advection}
-The 1D advection problem proposed by \citet{Donea2003} and \citet{Thieulot2011} is performed to verify the effectiveness of the SUPG method to stabilise the
-advection term of the energy equation. The domain is a 1D segment with $L_x=1$ composed by 50 elements and a discontinuity in the thermal field placed at
-$x=0.25$. Temperature is set to 1 for $x \leq 0.25$ and to 0 for $x > 0.25$. Velocity is set to ${u}=1$ in the entire domain. The simulation is performed for
-250 time steps, with $dt=0.002$, so the thermal discontinuity should be at $x=0.75$ at the end of the simulation.
-
-Temperature profiles at the end of the simulation are shown in \ref{fig:advection} as function of the dimensionless coefficient $\gamma=\tau \bm{v}/h$. In case
-of the classic Galerkin method ($\gamma=0$, blue line in \ref{fig:advection}) the final thermal profile is characterised by strong oscillations, which are
-eliminated in case of the SUPG method ($\gamma=0.045$, orange line in \ref{fig:advection}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Advection_stabilisation}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Advection.pdf}
-\caption{Temperature profile as function of \textit{x} for the advection stabilisation benchmark. Purple line indicates the initial temperature profile; the
-green line indicates the analytical temperature profile after 250 time steps; blue line indicates the temperature profile after 250 time steps in case of the
-classic Galerkin method ($\gamma=0$); orange line indicates the temperature profile after 250 time steps in case of the SUPG method ($\gamma=0.045$).}
-\label{fig:advection}
-\end{figure}
-
-\subsection{Shear heating}\label{sec:simple_shear}
-This exercise is performed in an unit square domain composed by \num{128x128} elements. The velocity field is prescribed on the entire domain with
-$\bm{v}=(L_y-y)y\bm{e_x}$; viscosity, density and specific heat are fixed to 1, while thermal conductivity and radiogenic energy are fixed to 0. Therefore,
-the energy equation (Eq. \ref{eq:energy}) can be simplified as \[\frac{\partial T}{\partial t}=H_s\] and fixing $T(t=0)=0$, the temperature can be find as
-\[T(t)=H_s t\]
-In this case we have
-\begin{eqnarray}
-\dot{\epsilon_{xx}}&=&\frac{\partial u}{\partial x}=0 \nonumber \\
-\dot{\epsilon_{yy}}&=&\frac{\partial v}{\partial y}=0 \nonumber \\
-\dot{\epsilon_{xy}}&=&\frac{1}{2}\left(\frac{\partial v}{\partial x}+\frac{\partial u}{\partial y}\right)=\frac{1}{2}(L_y-2y)\nonumber
-\end{eqnarray}
-and, simplifying Eq. \ref{eq:shear}, shear heating can be calculated as \[H_s(x,y)=(1-2y)^2\]
-
-The solution predicted by the model in terms of velocity, temperature and shear heating match well with the analytical solutions (Fig. \ref{fig:analytical_en}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Simple_shear}.
-
-%\begin{figure}
-%\centering
-%\noindent\includegraphics[width=400px]{./Figures/Energy_Thieulot.pdf}
-%\caption{Temperature and shear heating predicted by the model after 1 time step.}
-%\label{fig:en_thieulot}
-%\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=500px]{./Figures/Analytical_Shear.pdf}
-\caption{Velocity (panel a), temperature (panel b) and shear heating (panel c) as function of the \textit{y} coordinate for the simple shear experiment. The
-solutions predicted by the model (red lines) are compared with the analytical solutions (dashed blue lines).}
-\label{fig:analytical_en}
-\end{figure}
-
-\subsection{Shear and adiabatic heating}\label{sec:shear}
-This problem is performed as presented in Exercise 9.4 in \citet{Gerya2010b}. Two materials are vertically separated in a rectangular domain with $L_x=$
-\SI{1000}{\km}, $L_y=$ \SI{1500}{\km} and a grid resolution of \num{30x20} elements. Constant thermal coefficient expansion ($\alpha=$ \SI{3e-5}{\per\kelvin}),
-temperature ($T=$ \SI{1300}{\kelvin}) and gravity acceleration ($g_y=$ \SI{-10}{\m\square\s}) are assumed in the whole domain. Fluid 1 (on the left side of the
-domain) has $\rho_1=$ \SI{3200}{\kg\per\cubic\m} and $\eta_1=$ \SI{e20}{\pascal\s}; fluid 2 (on the right side of the domain) has $\rho_2=$
-\SI{3300}{\kg\per\cubic\m} and $\eta_2=$ \SI{e22}{\pascal\s}. Velocity boundary conditions are set to free slip on all sides of the domain.
-
-As shown in Fig. \ref{fig:energy}, both shear and adiabatic heating predicted by the code (first row in Fig. \ref{fig:energy}) well recreate the results
-obtained by means of example ${Shear\_adiabatic\_heating.m}$ from \citet{Gerya2010b} (second row in Fig. \ref{fig:energy}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Adiabatic%2BShear}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Energy.pdf}
-\caption{Comparison between shear (first column), adiabatic (second column) and total (third column) energy predicted by the code (panels a, b and c) and those
-created using example ${Shear\_adiabatic\_heating.m}$ from Exercise 9.4 in \citet{Gerya2010b} (panels d, e and f).}
-\label{fig:energy}
-\end{figure}
-
-\subsection{Mantle convection}\label{sec:mantle}
-This problem is performed as presented by \citet{Blankenbach1989} (constant viscosity cases) and \citet{Thieulot2014}, in a 2D unit square domain with gravity
-acceleration $g_y=10^{10}Ra$. The experiment is performed with three different Rayleigh numbers ($Ra=10^4$, $10^5$ and $10^6$) and with different grid
-resolution (between \num{32x32} and \num{128x128} elements). The fluid has constant viscosity, initial density, heat capacity, thermal conductivity
-($\eta= \rho_0= C_p = k=1$), reference temperature ($T_0=0$) and thermal expansion coefficient ($\alpha=10^{-10}$). Temperatures are set to 0 on top and 1 on
-bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\]
-
-The solution generated by the code in terms of $v_{\textrm{rms}}$ (calculated as in Sec. \ref{sec:error}) and Nu as function of time are reported for all the
-simulations in Fig. \ref{fig:mantle}. At the steady state, $v_{\textrm{rms}}$ and Nu of all simulations converge well toward the values from \citet{Blankenbach1989},
-with lower errors for higher resolution grids (Table \ref{tab:mantle}). All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Mantle_convection}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Convection.pdf}
-\caption{$v_{\textrm{rms}}$ (panels a, c and e) and Nu (panels b, d and f) for the mantle convection benchmark as function of time for different grid resolution.
-Panels a and b show the results for $Ra=10^4$; Panels c and d show the results for $Ra=10^5$; panels e and f show the results for $Ra=10^6$. Purple lines
-indicate the convergence values for the $v_{\textrm{rms}}$ and the Nu from \citet{Blankenbach1989} at the steady state.}
-\label{fig:mantle}
-\end{figure}
-
-\begin{table}[H]
-\caption{Comparison between $v_{\textrm{rms}}$ and Nu predicted by the code for the mantle convection experiment and same values as reported in literature.}
-\centering
-\small
-\begin{tabular}{|>{\centering}m{0.09\textwidth}|>{\centering}m{0.05\textwidth}|>{\centering}m{0.24\textwidth}|>{\centering}m{0.15\textwidth}|
-  >{\centering}m{0.1\textwidth} >{\centering}m{0.1\textwidth} >{\centering\arraybackslash}m{0.1\textwidth}|}
-\toprule
-& & \multirow{2}{*}{\citet{Blankenbach1989}} & \citet{Thieulot2014} & \multicolumn{3}{c|}{FALCON}  \\
- & & & \num{200x200} & \num{32x32} & \num{64x64} & \num{128x128}  \\
-\midrule
-\multirow{2}{*}{$Ra=10^4$} & $v_{\textrm{rms}}$ & $42.864947 \pm 0.000020$ & 42.867 & 42.83226 & 42.852793 & 42.861394  \\
-           & Nu      & $4.884409 \pm 0.000010$  & 4.882  & 4.781297 & 4.857475  & 4.877573   \\
-\hline
-\multirow{2}{*}{$Ra=10^5$} & $v_{\textrm{rms}}$ & $193.21454 \pm 0.00010$  & 193.255& 193.872643 & 193.377472 & 193.252290  \\
-           & Nu      & $10.534095 \pm 0.000010$ & 10.507 & 9.602514 & 10.270735  & 10.465629   \\
-\hline
-\multirow{2}{*}{$Ra=10^6$} & $v_{\textrm{rms}}$ & $833.98977 \pm 0.00020$  & 834.712& 848.091176 & 837.767911 & 834.945793  \\
-           & Nu      & $21.972465 \pm 0.000020$ & 21.695 & 15.999266 & 19.703682  & 21.306939   \\
-\bottomrule
-\end{tabular}
-\label{tab:mantle}
-\end{table}
-
-\subsection{Viscoplastic mantle convection}\label{sec:visco_mantle}
-This benchmark is performed as Cases 1-5a described in \citet{Tosi2015} in a 2D unit square domain with two different grid resolutions of \num{32x32} 
-and \num{100x100} elements. Reference viscosity, reference density, heat capacity, thermal conductivity and thermal expansion coefficient are fixed to a constant
-value of 1 ($\eta_0= \rho_0= C_p = k=\alpha=1$), while the gravity has been chosen as $g_y=10^2$, in order to obtain a Rayleigh number $Ra=10^2$. Temperatures
-are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\].
-The viscosity field $\eta$ is calculated as the harmonic average between a linear part $\eta_{lin}$ that depends on temperature only or on temperature and 
-depth $y$ and a nonlinear-plastic part $\eta_{pl}$ that depends on the strain rate $\dot{\bm{\epsilon}}$, as follows
-\[\eta(T,y,\dot{\bm{\epsilon}})=2\left(\frac{1}{\eta_{lin}(T,y)}+\frac{1}{\eta_{pl}(\dot{\bm{\epsilon}})}\right)^{-1}\]
-The linear and the nonlinear-plastic parts of the viscosity are calculated following \citet{Tosi2015} as follows
-\[\eta_{lin}(T,y)=\exp(-\ln(\Delta\eta_T)T+\ln(\Delta\eta_y)y)\]
-\[\eta_{pl}(\dot{\bm{\epsilon}})=\eta^*+\frac{\sigma_Y}{\sqrt{\dot{\bm{\epsilon}}:\dot{\bm{\epsilon}}}}\]
-where $\Delta\eta_T$, $\Delta\eta_y$, $\eta^*$ and $\sigma_Y$ are parameters chosen as listed in Table \ref{tab:tosi_setup}.
-
-For each case, some diagnostic quantities measured at the steady state are reported in Table \ref{tab:visco_mantle} (see Section \ref{sec:error} for 
-the computation of all the quantities), showing a well fit with results described in \citet{Tosi2015}. Data and charts for all cases can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Viscoplastic_mantle_convection}.
-
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{white!100!}{gray!25!}
-%\begin{table}[H]
-%  \caption{Results of the viscoplastic mantle convection benchmark for Cases 1-4 compared with results obtained with ELEFANT and YACC (from \citealp{Tosi2015}).}
-%  \centering
-%  \resizebox{\textwidth}{!}{
-%\begin{tabular}{|c|ccc|ccc|ccc|ccc|}
-%  \toprule
-%    & \multicolumn{3}{c|}{\textbf{Case 1}} & \multicolumn{3}{c|}{\textbf{Case 2}} & \multicolumn{3}{c|}{\textbf{Case 3}} &
-%    \multicolumn{3}{c|}{\textbf{Case 4}} \\
-%   & FALCON & ELEFANT& YACC & FALCON& ELEFANT& YACC & FALCON & ELEFANT & YACC & FALCON & ELEFANT & YACC \\
-%  \midrule
-%   $\left\langle T \right\rangle$ & 0.7759 & 0.7758 & 0.7767 & 0.6043 & 0.6036 & 0.6058 & 0.7275 & 0.7275 & 0.7286 & & 0.5277 & 0.5289\\
-%    Nu$^{top}$ & 3.4280 & 3.4214 & 3.4298 & 8.5242 & 8.5115 & 8.5278 & 3.0368 & 3.0347 & 3.0374 && 6.5912 & 6.5572 \\
-%    Nu$^{bottom}$ & 3.3168 & 3.313 & 3.3143 & 8.4599 & 8.4465 & 8.3990 & 3.0086 & 2.9908 & 2.9628 && 6.5834 & 6.5243 \\
-%    $v_{\textrm{rms}}$ & 249.901 & 249.134 & 251.7997 & 141.2451 & 140.7087 & 142.2020 & 100.2776 & 100.1208 & 100.9467 && 79.1105 & 79.6202 \\
-%    $v_{\textrm{rms}}^{top}$ & 1.8907 & 1.8642 & 1.8298 & 104.7928 & 104.1228 & 105.1767 & 2.0832 & 2.0652 & 2.0374 && 74.7596 & 75.8414 \\
-%    $u_{max}^{top}$ & 2.6340 & 2.6119 & 2.5516 & 121.9212& 121.766 & 122.3238 & 2.9086 & 2.9019 & 2.8458 && 88.9146 & 89.2940 \\
-%    $\eta_{min}$ & - & - & - & $2.0303\times10^{-5}$ & $2.0600\times10^{-5}$ & $1.9853\times10^{-5}$ & $4.8129\times10^{-5}$ & $4.8080\times10^{-5}$ & $4.7907\times10^{-5}$ &
-%    & $1.9860\times10^{-4}$ & $1.9174\times10^{-4}$ \\
-%    $\eta_{max}$ & - & - & - & 1.3325 & 1.85 & 1.6095 & 0.9360 & 0.9023 & 1 && 1.52 & 1.6773 \\
-%    $\left\langle W \right\rangle$ & 2.4299 & 2.4316 & 2.4583 & 7.5648 & 7.5665 & 7.6222 & 2.0387 & 2.0384 & 2.0400 && 5.6216 & 5.6512 \\
-%    $\left\langle \Phi \right\rangle/Ra$ & 2.4311 & 2.4276 & 2.4333 & 7.5661 & 7.5585 & 7.6194 & 2.0386 & 2.0370 & 2.0335 && 5.6182 & 5.6463 \\
-%    $\delta$ & 0.05\% & 0.16\% & 1.02\% & 0.02\% & 0.11\% & 0.04\% & $<$0.01\% & 0.07\% & 0.32\% && 0.06\% & 0.09\% \\
-%  \bottomrule
-%  \end{tabular}}
-%  \label{tab:visco_mantle}
-%  \end{table}
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{2}{white!100!}{gray!25!}
-\begin{table}[H]
-  \caption{Benchmark cases and corresponding parameters.}
-  \centering
-  \small
-\begin{tabular}{|>{\centering}m{0.08\textwidth}|>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}
-  >{\centering}m{0.05\textwidth} >{\centering}m{0.05\textwidth} |>{\centering\arraybackslash}m{0.2\textwidth}|}
-  \toprule
-    Case & $Ra$ & $\Delta\eta_T$ & $\Delta\eta_y$ & $\eta^*$ & $\sigma_Y$ & Convective regime \\
-  \midrule
-  \rowcolor{gray!25!}
-    1 & $10^2$ & $10^5$ & 1 & - & - & Stagnant lid \\
-    2 & $10^2$ & $10^5$ & 1 & $10^{-3}$ & 1 & Mobile lid  \\
-    \rowcolor{gray!25!}
-    3 & $10^2$ & $10^5$ & 10 & - & - & Stagnant lid  \\
-    4 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 1 & Mobile lid \\
-    \rowcolor{gray!25!}
-    5 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 4 & Periodic \\
-  \bottomrule
-  \end{tabular}
-  \label{tab:tosi_setup}
-  \end{table}
-
-\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{gray!25!}{white!100!}
-\begin{table}
-  \caption{Results of the viscoplastic mantle convection benchmark for different grid resolutions in Cases 1-5 compared with results obtained with ELEFANT
-  and YACC ($100\times100$ elements, from \citealp{Tosi2015}) and in Stone 28 ($32\times32$ elements, from
-  \url{https://github.com/cedrict/fieldstone/tree/master/python_codes/fieldstone_28}).}
-  \centering
-  \resizebox{\textwidth}{!}{
-\begin{tabular}{|cc|ccccccccccc|}
-  \toprule
-  \multicolumn{13}{|c|}{\textbf{Case 1}} \\
-  \midrule
-  Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-  & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-  \hline
-  \rowcolor{gray!25!}
-    YACC & $100\times100$ & 0.7767 & 3.4298 & 3.3143 & 251.7997 & 1.8298 & 2.5516 & - & - & 2.4583 & 2.4333 & 1.02\% \\
-    ELEFANT & $100\times100$ & 0.7758 & 3.4214 & 3.313 & 249.134 & 1.8642 & 2.6119 & - & - & 2.4316 & 2.4276 & 0.16\% \\
-    \rowcolor{gray!25!}
-    Stone 28 & $32\times32$ & 0.7737 & 3.3987 & - & 243.872 & - & - & - & - & - & - & - \\
-%    FALCON & $32\times32$ & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & - & - & 2.4745 & 2.4864 & 0.48\% \\
-%    \rowcolor{gray!25!}
-    FALCON & $100\times100$ & 0.7759 & 3.4280 & 3.3168 & 249.901 & 1.8907 & 2.6340 & - & - & 2.4299 & 2.4311 & 0.05\% \\
-    \midrule
-    \multicolumn{13}{|c|}{\textbf{Case 2}} \\
-    \midrule
-    Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-    & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-    \hline
-    \rowcolor{gray!25!}
-     YACC & $100\times100$ & 0.6058 & 8.5278 & 8.3990 & 142.2020 & 105.1767 & 122.3238 & $1.9853\times10^{-5}$ & 1.6095 & 7.6222 & 7.6194 & 0.04\% \\
-     ELEFANT & $100\times100$ & 0.6034 & 8.5115 & 8.4465 & 140.8067 & 104.1228 & 121.766 & $2.0600\times10^{-5}$ & 1.85 & 7.5665 & 7.5585 & 0.11\% \\
-     \rowcolor{gray!25!}
-     Stone 28 & $32\times32$ & 0.6052 & 8.1715 & - & 141.518 & - & - & - & - & - & - & - \\
-%     FALCON & $32\times32$ & 0.6454 & 8.1422 & 7.4020 & 156.6908 & 95.7226 & 116.9302 & $2.2572\times10^{-5}$ & 0.7194 & 7.4289 & 7.4452 & 0.22\% \\
-%     \rowcolor{gray!25!}
-     FALCON & $100\times100$ & 0.6044 & 8.5266 & 8.4590 & 141.3684 & 104.8867 & 122.0470 & $2.0333\times10^{-5}$ & 1.2914 & 7.5707 & 7.5718 & 0.01\% \\
-     \midrule
-     \multicolumn{13}{|c|}{\textbf{Case 3}} \\
-     \midrule
-     Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-     & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-     \hline
-     \rowcolor{gray!25!} 
-     YACC & $100\times100$ & 0.7286 & 3.0374 & 2.9628 & 100.9467 & 2.0374 & 2.8458 & $4.7907\times10^{-5}$ & 1 & 2.0400 & 2.0335 & 0.32\% \\
-      ELEFANT & $100\times100$ & 0.7275 & 3.0347 & 2.9908 & 100.1208 & 2.0652 & 2.9019 & $4.8080\times10^{-5}$ & 0.9023 & 2.0384 & 2.037 & 0.07\% \\
-      \rowcolor{gray!25!}
-      Stone 28 & $32\times32$ & 0.7277 & 3.0347 & - & 100.018 & - & - & - & - & - & - & - \\
-%      FALCON & $32\times32$ & 0.7281 & 3.0618 & 2.8132 & 101.0092 & 2.1630 & 3.0178 & $4.8500\times10^{-5}$ & 0.7845 & 2.0699 & 2.0695 & 0.02\% \\
-%      \rowcolor{gray!25!}
-      FALCON & $100\times100$ & 0.7275 & 3.0368 & 3.0086 & 100.2776 & 2.0832 & 2.9086 & $4.8129\times10^{-5}$ & 0.9360 & 2.0387 & 2.0386 & $<$0.01\% \\
-      \midrule
-      \multicolumn{13}{|c|}{\textbf{Case 4}} \\
-      \midrule
-      Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-      & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-      \hline
-      \rowcolor{gray!25!} 
-      YACC & $100\times100$ & 0.5289 & 6.5572 & 6.5243 & 79.6202 & 75.4814 & 89.2940 & $1.9174\times10^{-4}$ & 1.6773 & 5.6512 & 5.6463 & 0.09\% \\
-       ELEFANT & $100\times100$ & 0.5277 & 6.5912 & 6.5834 & 79.1105 & 74.7596 & 88.9146 & $1.9860\times10^{-4}$ & 1.5200 & 5.6216 & 5.6182 & 0.06\% \\
-       \rowcolor{gray!25!}
-       Stone 28 & $32\times32$ & 0.5286 & 6.4036 & - & 79.4746 & - & - & - & - & - & - & - \\
-%       FALCON & $32\times32$ & 0.5334 & 6.4679 & 6.3435 & 80.6153 & 76.0511 & 90.2058 & $2.0393\times10^{-4}$ & 0.6453 & 5.6676 & 5.6698 & 0.04\% \\
-%       \rowcolor{gray!25!}
-       FALCON & $100\times100$ & 0.5279 & 6.5897 & 6.5830 & 79.1220 & 75.0715 & 88.8484 & $1.9693\times10^{-4}$ & 1.4133 & 5.6127 & 5.6113 & 0.03\% \\
-       \midrule
-       \multicolumn{13}{|c|}{\textbf{Case 5}} \\
-       \midrule
-       Code & Resolution & $P$ & $\left\langle T \right\rangle_{min}$ & $\left\langle T \right\rangle_{max}$ & Nu$^{top}_{min}$ & Nu$^{top}_{max}$ &
-       $v_{\textrm{rms}}^{min}$ & $v_{\textrm{rms}}^{max}$ & \multicolumn{2}{c}{$(\left\langle \Phi \right\rangle/Ra)_{min}$} & 
-       \multicolumn{2}{c|}{$(\left\langle \Phi \right\rangle/Ra)_{max}$} \\
-       \hline
-       \rowcolor{gray!25!}
-         YACC & $100\times100$ & 0.0792 & 0.6543 & 0.6722 & 2.6950 & 7.2792 & 41.9813 & 98.8230 & \multicolumn{2}{c}{1.4347} & 
-         \multicolumn{2}{c|}{9.2748} \\
-         ELEFANT & $100\times100$ & 0.076 & 0.6514 & 0.6693 & 2.6808 & 7.3506 & 41.2578 & 100.089 & \multicolumn{2}{c}{1.3324} & 
-         \multicolumn{2}{c|}{9.5289} \\
-         \rowcolor{gray!25!}
-         Stone 28 & $32\times32$ & 0.0764 & 0.6521 & 0.6697 & 2.6912 & 7.0763 & 41.61 & 98.183 & \multicolumn{2}{c}{-} & 
-         \multicolumn{2}{c|}{-} \\
-     %    FALCON & $32\times32$ & & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & \multicolumn{2}{c}{-} & 
-     %     \multicolumn{2}{c|}{-} \\
-     %    \rowcolor{gray!25!}
-         FALCON & $100\times100$ & 0.0775 & 0.6519 & 0.6697 & 2.6847 & 7.3629 & 41.5256 & 99.5466 & \multicolumn{2}{c}{1.3399} & 
-         \multicolumn{2}{c|}{9.3993} \\
-  \bottomrule
-  \end{tabular}}
-  \label{tab:visco_mantle}
-  \end{table}
-
-\subsection{Thin layer entrainment}\label{sec:thin}
-This experiment is performed as originally proposed by \citet{vanKeken1997} with two incompressible fluids in a rectangular domain with $L_x=2$ and $L_y=1$
-and gravity acceleration $g_y=10^{10}Ra$. The Rayleigh number and the compositional Rayleigh number are fixed to $Ra=300000$ and $Rc=450000$, respectively.
-Both fluids have constant viscosity, thermal conductivity, specific heat ($\eta=\rho=C_p=1$) and thermal expansion coefficient ($\alpha=10^{-10}$). Fluid 1
-has a density $\rho_1=1$, while fluid 2 is denser ($\rho_2=\rho_1+1.5 \cdot 10^{-10}$) and is placed at the bottom of the domain, for $y \leq 0.025$.
-Temperature are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides of the domain. The initial
-temperature field is given by
-\[T(x,y)=T_u(x,y)+T_l(x,y)+T_r(x,y)+T_s(x,y)-\frac{3}{2}\]
-with
-\begin{eqnarray}
-T_u(x,y)&=&\frac{1}{2}erf\left(\frac{1-y}{2}\sqrt{\frac{u_0}{x}}\right)\nonumber \\
-T_l(x,y)&=&1-\frac{1}{2}erf\left(\frac{y}{2}\sqrt{\frac{u_0}{L_x-x}}\right)\nonumber \\
-T_r(x,y)&=&\frac{1}{2}+\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{y+1}} \exp\left(-\frac{x^2u_0}{4y+4}\right)\nonumber \\
-T_s(x,y)&=&\frac{1}{2}-\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{2-y}} \exp\left(-\frac{(L_x-x)^2u_0}{8-4y}\right)\nonumber
-\end{eqnarray}
-and
-\begin{eqnarray}
-u_0&=&\frac{L_x^{7/3}}{(1+L_x^4)^{2/3}}\left(\frac{Ra}{2\sqrt{\pi}}\right)^{\frac{2}{3}}\nonumber \\
-Q&=&2\sqrt{\frac{L_x}{\pi u_0}}\nonumber
-\end{eqnarray}
-
-The experiment is performed with two grid resolution (\num{125x40} and \num{200x80} elements), with 100 markers per element and CFL of 0.25. $v_{\textrm{rms}}$
-(see Sec. \ref{sec:error}) calculated for both the simulations match well with results from \citet{vanKeken1997} and obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and ELEFANT \citep{Thieulot2014}, obtained with similar grid resolutions
-(Fig. \ref{fig:thin}). In addition, a variety of simulations with different resolutions and aspect ratios are performed to verify that the sensitivity of the
-initial velocity field. Results match well with results from \citet{vanKeken1997} (in the grey area) and \citet{Thieulot2014} (Fig. \ref{fig:thin_initial}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Thin_layer}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_thin.pdf}
-\caption{$v_{\textrm{rms}}$ for the thin layer experiment as function of time for different grid resolution. Results are compared with results obtained by
-\citet{vanKeken1997} (black lines) and with ASPECT and ELEFANT.}
-\label{fig:thin}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrmszero.pdf}
-\caption{$v_{\textrm{rms}}$ at $t=0$ for the thin layer experiment as function of the element size for different aspect ratios of the elements. Results are compared
-with results obtained by \citet{vanKeken1997} (black dots in the grey area) and with ASPECT and ELEFANT.}
-\label{fig:thin_initial}
-\end{figure}
-
-\subsection{Sinking hydrated cylinder}\label{sec:quinquis}
-This experiment is performed as explained in \citet{Quinquis2014} and simulates the sinking of a cold, hydrated cylinder into a warm, dry mantle. The domain
-is square with $L_x=L_y=$ \SI{300}{\km}, a grid resolution of \num{300x300} elements and 25 markers per element. The sinking cylinder is located at $x=$
-\SI{150}{\km} and $y=$ \SI{170}{\km} with a radius of \SI{20}{\km} and has $\rho_c=$ \SI{3250}{\kg\per\cubic\m}, $\eta_c=$ \SI{e23}{\pascal\s}, a thermal
-conductivity $k_c=$ \SI{4.5}{\watt\per\m\per\kelvin},a specific heat $C_{p_c}=$ \SI{1250}{\joule\per\kg\per\kelvin}; the surrounding mantle has $\rho_m=$
-\SI{3200}{\kg\per\cubic\m}, $\eta_m=$ \SI{e20}{\pascal\s}, a thermal conductivity $k_m=$ \SI{105}{\watt\per\m\per\kelvin} and specific heat $C_{p_m}=$
-\SI{1250}{\joule\per\kg\per\kelvin}; a 58 km-thick lithosphere overlies the mantle and has $\rho_l=$ \SI{3200}{\kg\per\cubic\m}, $\eta_l=$ \SI{e23}{\pascal\s},
-a thermal conductivity $k_l=$ \SI{4.5}{\watt\per\m\per\kelvin}, specific heat $C_{p_l}=$ \SI{750}{\joule\per\kg\per\kelvin} and a thermal diffusivity of
-\SI{e-6}{\square\m\per\s}. Velocity boundary conditions are set to free slip on all sides of the domain. Temperature increase linearly in the lithosphere,
-from 0° to 1300°C, and in the mantle, from 1300° to 1360.5°C. The initial temperature of the cylinder is fixed to 400°C. Throughout the evolution, temperature
-is fixed to 0° and 1360.5°C at the top and bottom of the domain, respectively. The initial bound water content is imposed at 2 wt.\% in the cylinder and at
-0 wt.\% in both the mantle and the lithosphere. The maximum amount of water in the mantle is fixed to 0.2 wt.\%, while maximum water content in the cylinder
-and in the lithosphere is function of pressure and temperature and is calculated using a serpentinized harzburgite with Perple\_X, as in \citet{Quinquis2014}.
-
-As shown in \citet{Quinquis2014}, the model is characterised by a progressive dehydration in the external portion of the cylinder, due to the increase of
-temperature, with a consequent vertical migration of free water that hydrates the mantle above the cylinder, up to the lithosphere. The distribution of
-bound and free water during the evolution of the experiment is shown in Fig. \ref{fig:quinquis} (left and right columns, respectively). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Hydration}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Hydration.pdf}
-\caption{Distribution of bound and free water (left and right columns, respectively) at $t=0.5$ and $t=$ \SI{1.25}{\mega\year} (first and second rows,
-respectively) for the sinking hydrated cylinder experiment. Only markers with bound or free water not equal to 0 are plotted.}
-\label{fig:quinquis}
-\end{figure}
-
-\subsection{Mantle melting curves}\label{sec:katz}
-This instantaneous experiment is performed in order to recreate isobaric melting curves, as function of temperature, bulk water and content of water in the
-melt, to compare with melting curves obtained by \citet{Katz2003}. The domain is square with $L_x=L_y=100$ km, a grid resolution of \num{100x100} elements and
-25 markers per element. Density is set in the entire domain to \SI{3300}{\kg\per\cubic\m} and temperature increases from the sides to the centre from 1000° to
-1500°C. The experiment is performed for bulk water contents from 0 to 0.3 wt.\%.
-
-Fig. \ref{fig:fraction} shows isobaric melting curves for \SI{1}{\giga\pascal} (panel a) and \SI{3}{\giga\pascal} (panel b) calculated for different bulk water
-contents. The curves well match with those obtained for the mantle by \citet{Katz2003}, showing the saturation at the solidus for bulk water of 0.2 and 0.3 wt.\%, indicating
-that the code correctly determines melt fraction as function of pressure, temperature and bulk water content. Same results can be observed in Fig.
-\ref{fig:water}a, where isobaric and isothermal melting curves are shown as function of bulk water content. In addition, Fig. \ref{fig:water}b shows that the
-code correctly determines also the percentage of water in the melt as function of pressure, temperature and melt fraction. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Melting}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Fraction.pdf}
-\caption{Isobaric melting curves for 1 GPa (panel a) and 3 GPa (panel b) as function of temperature with different bulk water contents.}
-\label{fig:fraction}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Water.pdf}
-\caption{Isobaric melting curves for 1.5 GPa as function of bulk water (panel a) and water in the melt (panel b) calculated for different temperatures.}
-\label{fig:water}
-\end{figure}
-
-\clearpage
-
-\bibliography{export}
-
-\end{document}

### 2022-09-07 tex/markers.tex
update main

--- 2022-01-28 .debris/2022-01-27/Paper.tex
@@
-\documentclass[hidelinks,10pt,a4paper]{article}
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage{booktabs}
-\usepackage{multirow}
-\usepackage[cm]{fullpage}
-\usepackage[round]{natbib}
-\bibliographystyle{abbrvnat}
-\usepackage{doi}
-\usepackage{bm}
-\usepackage[table]{xcolor}
-\hypersetup{
-    colorlinks,
-    linkcolor={blue},
-    citecolor={blue},
-    urlcolor={blue}
-}
-\usepackage{float}
-\restylefloat{table}
-\usepackage{siunitx}
-\DeclareSIUnit\year{yr}
-
-\begin{document}
-
-\section{Introduction}\label{sec:intro}
-This supplement includes a description of the code, with a presentation of the features implemented followed by results of extensive benchmarks listed in
-Table \ref{tab:list}, implemented to verify that the code solves correctly a large range of problems. Description and results of all benchmarks are shown
-in Section \ref{sec:benchmark}.
-
-\begin{table}[H]
-\caption{List of benchmarks implemented for each feature of the code. Data of all benchmarks can be found at 
-\url{https://github.com/aleregorda/Benchmarks}}
-\centering
-\begin{tabular}{| l | l | l |}
-\toprule
- \textbf{Feature}  & \textbf{Benchmark} & \textbf{Section} \\
-\midrule
-  Solver  & Stokes flow & \ref{sec:stokes} \\
-\hline
-  \multirow{2}{*}{Markers advection} & Zalesak disk & \ref{sec:runge} \\
-    & Conservative Interpolation Velocity & \ref{sec:CVI} \\
-\hline
-  \multirow{4}{*}{Momentum equation}  & Poiseuille flow & \ref{sec:poiseuille}\\
-    & Instantaneous 2D sphere & \ref{sec:ist_sphere} \\
-    & Rayleigh-Taylor experiment & \ref{sec:rayleigh} \\
-    & Falling block & \ref{sec:block} \\
-\hline
-    & 2D Stokes sphere & \ref{sec:sphere} \\
-    Sticky air and  & Stabilisation algorithm & \ref{sec:stab} \\
-    free surface & Topography relaxation & \ref{sec:crameri} \\
-  & Spontaneous subduction & \ref{sec:subduction} \\
-\hline
-  Erosion and sedimentation  & - & - \\
-\hline
-   Non-linear  & Slab detachment & \ref{sec:slab} \\
-   visco-plasticity & Indenter & \ref{sec:indenter} \\
-    & Brick & \ref{sec:brick} \\
-\hline
-  \multirow{3}{*}{Energy equation}  & Advection stabilisation & \ref{sec:advection} \\
-    & Simple shear heating & \ref{sec:simple_shear} \\
-    & Shear and adiabatic heating & \ref{sec:shear} \\
-\hline
-\multirow{3}{*}{Energy + momentum} & Mantle convection & \ref{sec:mantle} \\
-& Viscoplastic mantle convection & \ref{sec:visco_mantle} \\
-& Thin layer entrainment & \ref{sec:thin} \\
-\hline
-  Phase changes  & \multirow{2}{*}{Hydrated sinking cylinder} & \multirow{2}{*}{\ref{sec:quinquis}} \\
-  and hydration  &  & \\
-\hline
-  Melting & Experimental melting curves & \ref{sec:katz} \\
-\bottomrule
-\end{tabular}
-\label{tab:list}
-\end{table}
-
-The code is written in Fortran90 and uses the finite element method (FEM) with quadrilateral $Q_1 \times P_0$ elements (continuous bilinear velocity and
-discontinuous constant pressure), associated with MUMPS\footnote{\url{http://mumps.enseeiht.fr/}} \citep{Amestoy2001,Amestoy2019} that is a software package
-for solving systems of linear equations of the form $A \cdot x = b$, where \textit{A} is a square sparse matrix, by means of a direct method. Since
-$Q_1 \times P_0$ elements do not satisfy Ladyzhenskaya, Babuska and Brezzi (LBB) stability condition \citep{Donea2003}, elemental pressure is elaborated
-in post-processing to avoid spurious pressures by means of a double interpolation to smooth the pressure. Following this procedure, the elemental pressure
-is interpolated onto nodes and then back onto elements. After the smoothing procedure, the pressure field is calculated again on the nodes to be used in
-combination with the lithostatic pressure, which is calculated onto nodes. Correctness of the solution and performances of the solver in terms of time and
-memory usage are tested solving the Stokes flow with the analytical solution proposed by \citet{Donea2003} (Section \ref{sec:stokes}).
-
-The thermo-mechanics of crust-mantle systems is described by means of the conservation of mass, momentum and energy equations, expressed as follows:
-\begin{eqnarray}
-\label{eq:mass}\frac{\partial \rho}{\partial t}+\bm{\nabla} \cdot (\rho \bm{u})&=&0\\
-\label{eq:momentum}-\bm{\nabla} p + \bm{\nabla} \cdot \bm{\tau} + \rho \bm{g}&=&0\\
-\label{eq:energy}\rho C_p \left(\frac{\partial T}{\partial t} + \bm{u} \cdot \bm{\nabla} T \right)&=&\bm{\nabla} \cdot \left(k\bm{\nabla} T\right) + H_{tot}
-\end{eqnarray}
-where $\rho$ is the density, $\bm{u}$ is the velocity, \textit{p} is the pressure, $\bm{\tau}$ is the deviatoric stress, $\bm{g}$ is the gravity acceleration,
-$C_p$ is the specific heat at constant pressure, \textit{T} is the temperature, \textit{k} is the thermal conductivity and $H_{tot}$ is the total internal heat
-production. Density variations due to temperature are generally small enough to assume the density as constant ($\rho=\rho_0$) in Eq. \ref{eq:mass} and in
-Eq. \ref{eq:energy}, while it must be treated as a variable in the buoyancy term of Eq. \ref{eq:momentum}, such that \[\rho=\rho_0(1-\alpha(T-T_0))\]
-where $\rho_0$ is the density at a reference temperature $T_0$ and $\alpha$ is the coefficient of thermal expansion. Eq. \ref{eq:mass} then can be rewritten as
-\begin{equation}\label{eq:mass_0}
-\bm{\nabla} \cdot \bm{u}=0
-\end{equation}
-This simplification is known as the extended Boussinesq approximation.
-
-The time step $dt$ is chosen in according to Courant-Friedrichs-Lewy condition \citep{Anderson}, such that
-\[dt=C_n\; \min\left(\frac{\Delta x}{\max|\bm{u}|},\frac{(\Delta x)^2}{\kappa}\right)\]
-where $0<C_n<1$ is the Courant (CFL) number, $\Delta x$ is the minimum dimension of smallest element, $\max|\bm{u}|$ is the maximum velocity calculated on the
-entire domain and
-$\kappa=\frac{k}{\rho C_p}$ is the thermal diffusivity \citep{Thieulot2014}.
-
 \section{Lagrangian markers}\label{sec:markers}
@@
 the properties of the nearest marker. In this way elements are never empty and maintain a number of markers inside a prefixed range.
-
-\section{The momentum equation}\label{sec:mom_eq}
-The penalty method is implemented so that Eq. \ref{eq:mass_0} can be written as:
-\begin{equation}\label{eq:mass_penalty}
-\bm{\nabla} \cdot \bm{u} + \frac{p}{\lambda}=0
-\end{equation}
-where $\lambda$ is the so-called penalty parameter, which should be 6-7 orders of magnitude larger than the shear viscosity to ensure that mass conservation
-is satisfied \citep{Donea2003,Thieulot2014}.
-
-The deviatoric stress tensor in Eq. \ref{eq:momentum} can be written in terms of the strain rate tensor as $\bm{\tau}=2\eta\dot{\bm{\epsilon}}$, with
-$\dot{\bm{\epsilon}}=\frac{1}{2}\left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)$. Therefore, Eq. \ref{eq:momentum} can be rewritten as:
-\begin{equation}\label{eq:momentum_pressure}
--\bm{\nabla} p + \bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right) + \rho \bm{g} = 0
-\end{equation}
-Finally, using pressure from Eq. \ref{eq:mass_penalty}, Eq. \ref{eq:momentum_pressure} can be rewritten as:
-\begin{equation}\label{eq:momentum_penalty}
-\lambda \bm{\nabla} \left(\bm{\nabla} \cdot \bm{v}\right)+\bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right)+\rho\bm{g}=0
-\end{equation}
-The penalty method is been associated to the iterative Uzawa method, as described in detail in \citet{Dabrowski2008} and \citet{Thieulot2014}
-(see Section \ref{sec:poiseuille}). Results of benchmarks performed to verify the correctness in the implementation of Eq. \ref{eq:momentum_penalty} are
-shown in Sections \ref{sec:ist_sphere} and \ref{sec:rayleigh}.
-
-To support large viscosities variations the penalty parameter is related to the effective elemental viscosity by means of a dimensionless coefficient, so that
-$\lambda=\lambda_e(e)\eta_{eff}(e)$ \citep{Marotta2006,Dabrowski2008,Thieulot2014}. Benchmark of the falling block \citep{Gerya2003a,Gerya2010b,Thieulot2011}
-is performed to verify that the code can correctly deal buoyancy driven flows with strong viscosity contrasts (Section \ref{sec:block}).
-
-\section{Sticky air and free surface}\label{sec:sticky}
-The Earth's surface can be treated by means of either the so-called sticky-air or a true free surface method, both of them implemented in the code. In the
-sticky-air method the surface is approximated with the introduction of a buoyant layer with a viscosity at least four orders of magnitude lower than the crust
-\citep{Schmeling2008,Crameri2012} and the interface between lithosphere and air is defined using a chain of passive markers that are advected as the Lagrangian
-markers. The correctness of the evolution of the markers chain is tested with the experiment of a 2D time-dependent Stokes sphere below a free surface and
-compared with results from ASPECT\footnote{\url{https://aspect.geodynamics.org/}} \citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} (Section
-\ref{sec:sphere}). In the true free surface case the top boundary is assumed stress-free and velocities are not fixed. In this case topography variations are
-described by vertical deformations of the mesh that depend on the velocity field of the nodes that identify the free surface, while horizontal deformations are
-not taken into account. This procedure is known as the Arbitrary Lagrangian-Eulerian (ALE) method and its implementation follows the technique described in
-\citet{Thieulot2011}. However, although the implementation of a true free surface better reproduces laboratory experiment, extremely small time steps can be
-necessary to maintain stability \citep{Kaus2010a,Quinquis2011,Thieulot2014}. Therefore, the stability algorithm proposed by \citet{Kaus2010a} is implemented
-to avoid instabilities due to high density differences at the free surface when using too large time steps. The implementation of this algorithm are tested by
-performing the experiment described by \citet{Kaus2010a} (Section \ref{sec:stab}).
-
-The topography relaxation benchmark proposed by \citet{Crameri2012} is performed to verify that the code correctly recovers topography variations in case of
-both the sticky-air and the true free surface method (Section \ref{sec:crameri}). Finally, Section \ref{sec:subduction} show the results of the spontaneous
-subduction experiment described by \citet{Schmeling2008}.
-
-\section{Erosion and sedimentation}\label{sec:erosion}
-Surface processes at the lithosphere-air interface are implemented taking into account fluvial erosion (long-range processes), hillslope processes (short-range
-processes) and failure of slopes exceeding a threshold angle $\theta$ \citep{Burov1997,Tucker1999,Sacek2012,Croissant2014,Ueda2015,Beucher2020}. Erosion rates
-$\dot{E}_{lr}$ due to long-range processes can be described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, in according to
-\begin{equation}
-\dot{E}_{lr}=K_f A^{m_e} \bm{\nabla} h^{n_e}
-\end{equation}
-where $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $\bm{\nabla} h$ is the slope and $m_e$ and $n_e$ are parameters chosen in to
-constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}. Erosion rates $\dot{E}_{sr}$ due to short-range processes are
-implemented by assuming that transport is linearly proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}, in according to
-\begin{equation}
-\dot{E}_{sr}=K_d \bm{\nabla}^2 h
-\end{equation}
-where $K_d$ is the effective diffusivity. In addition, the slope is modified if it still exceeds a threshold angle after long- and short-range processes
-\citep{Ueda2015}.
-
-Offshore sedimentation processes are implemented similarly to hillslope processes but with a depth-dependent coefficient \citep{Sacek2012,Beucher2020}.
-The sedimentation rate $\dot{S}$ is calculated as
-\begin{equation}
-\dot{S}=K_s \bm{\nabla}^2 h
-\end{equation}
-with
-\begin{equation}
-K_s=K^*_s \exp\left(\frac{h-h_{sl}}{h_0}\right)
-\end{equation}
-where $K^*_s$ is a constant, $h_{sl}$ is the sea level and $h_0$ is a characteristic length. In case of sticky air, slope is determined by means of the markers
-chain, which is vertically corrected considering erosion and sedimentation rates. After the correction, continental markers above the markers chain are
-transformed in air markers and air markers below the markers chain become sediments. Similarly, slope is determined considering the nodes on the top boundary
-in case of true free surface and they are vertically corrected as for the markers chain. In this case, continental markers above the top boundary are simply
-deleted.
-
-\section{Non-linear rheologies}\label{sec:nonlinear}
-Non-linear rheologies are implemented combining viscous creep (dislocation and diffusion) and plastic yielding. For each marker, diffusion and dislocation
-viscosity ($\eta_{df|ds}$) can be determined as follows \citep{Karato1993,Wang2016}:
-\begin{equation}\label{eq:viscous}
-\eta_{df|ds}=\left(\frac{d^m}{A}\right)^{\frac{1}{n}}I_2^{\frac{(1-n)}{n}}\exp\left(\frac{Q+pV}{nRT}\right)
-\end{equation}
-where $d$ is the grain size, $m$ is the grain size exponent, $n$ is the stress exponent, $A$ is the pre-exponential factor, $I_2$ is the square root of the
-second invariant of the strain rate tensor, $Q$ is the activation energy, $p$ is the pressure, $V$ is the activation volume, $R$ is the gas constant and $T$
-is the temperature. Pressures and temperatures are determined by the interpolation of the nodal parameters, while strain rates are calculated by means of the
-derivative of the velocities on the nodes. In case of diffusion creep $n=1$ and $m>0$, in case of dislocation creep $n>1$ and $m=0$. The viscous creep $\eta_{cp}$
-is then calculated as the harmonic average between $\eta_{df}$ and $\eta_{ds}$:
-\begin{equation}\label{eq:visco_creep}
-\eta_{cp}=\left(\frac{1}{\eta_{df}}+\frac{1}{\eta_{ds}}\right)^{-1}
-\end{equation}
-The implementation of the non-linear viscous creep viscosity depending on the strain rate is tested by means of the slab detachment benchmark
-(Section \ref{sec:slab}).
-
-Plastic yielding is implemented rescaling $\eta_{cp}$ in order to limit the stress below the yield stress $\sigma_y$ \citep{Thieulot2008,Thieulot2014,Glerum2018},
-obtaining
-\begin{equation}\label{eq:plastic}
-\eta_{pl}=\frac{\sigma_y}{2I_2}
-\end{equation}
-where $\eta_{pl}$ is the plastic viscosity and the yield stress is determined following the Drucker-Prager criterion, such as
-\begin{equation}\label{eq:yield}
-\sigma_y=C\;\cos(\phi)+p\;\sin(\phi)
-\end{equation}
-where $C$ is the cohesion and $\phi$ is the internal friction angle. In case of negative pressure it is imposed equal to 0, so that negative yield stress are
-excluded. The correctness of the non-linear solution in case of plastic viscosity with variable internal friction angle is verified performing indenter and
-brick experiments (Sections \ref{sec:indenter} and \ref{sec:brick}, respectively).
-
-Strain softening is taken into account for both viscous creep and plastic viscosity \citep{Huismans2003,Babeyko2005,Huismans2005,Sobolev2005,Warren2008a} by
-means of the accumulated strain $\epsilon$ memorised by each marker. Softening in the viscous creep determines a linear decrease of $\eta_{vc}$ by means of a
-viscous strain softening factor $W_S$ that increases linearly from $W_{S_0}$ to $W_{S_{\infty}}$ for $\epsilon_{S_0}<\epsilon_S<\epsilon_{S_{\infty}}$ 
-\citep{Warren2008a}. This viscous softening can be related to strain-induced grain size reduction \citep{Warren2008a}. Differently, plastic softening is 
-simulated with a linear decrease of internal friction angle $\phi(\epsilon)$ and cohesion $C(\epsilon)$ in according to
-\begin{eqnarray}
-\label{eq:friction1}\phi(\epsilon)&=&\phi_0+(\phi_{\infty}-\phi_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}\\
-\label{eq:friction2}C(\epsilon)&=&C_0+(C_{\infty}-C_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}
-\end{eqnarray}
-where $\phi_0$, $C_0$ and $\phi_{\infty}$, $C_{\infty}$ are internal friction angle and cohesion for $\epsilon_0$ and $\epsilon_{\infty}$, respectively
-\citep{Huismans2003,Huismans2005,Warren2008a,Thieulot2014}. Plastic softening approximates deformation-induced softening of faults and brittle shear zones
-\citep{Warren2008a}.
-
-Plastic yielding and viscous creep are then combined to obtain a viscoplastic viscosity $\eta_{vp}$ as follows, assuming that they are independent processes
-\citep{Karato2008,Glerum2018}:
-\begin{equation}\label{eq:viscoplastic}
-\eta_{vp}=\min(\eta_{cp},\eta_{pl})
-\end{equation}
-Finally, effective viscosity $\eta_{eff}$ is capped by the minimum and the maximum viscosity ($\eta_{min}$ and $\eta_{max}$, respectively) to avoid extremely
-low or high viscosity \citep{Glerum2018} as follows
-\begin{equation}\label{eq:effective}
-\eta_{eff}=\min(\max(\eta_{vp},\eta_{min}),\eta_{max})
-\end{equation}
-Viscous creep and plastic yielding are non-linear rheologies because of their dependence on the velocity field through pressure and strain rates. Therefore,
-the solution is determined by means of Picard-type iterations, until convergence of the velocity field \citep{Glerum2018}. The convergence is verified at each
-iteration $i$ via the nonlinear residual $\bm{\mathcal{R}}^i$ that can be determined as
-\begin{equation}\label{eq:convergence}
-\bm{\mathcal{R}}^i= \mathbb{K} (\eta_{eff}(\dot{\bm{\epsilon}}^{i-1},p^{i-1}))\cdot \bm{v}^{i-1} - \bm{f}^i
-\end{equation}
-%\begin{equation}\label{eq:convergence}
-%\chi=1-\frac{\langle (\bm{v}^i-\langle\bm{v}^i\rangle)\cdot (\bm{v}^{i+1}-\langle\bm{v}^{i+1}\rangle) \rangle}{\sqrt{\sigma^i\sigma^{i+1}}}
-%\end{equation}
-%Iterations are performed until either convergence ($\chi<tol$) or maximal number of non-linear iterations ($it_{max}$) is reached.
-where $\mathbb{K}$ is the Stiffness matrix and $\bm{f}$ is the right hand side vector \citep{Spiegelman2016,Glerum2018}.  The $L_2$-norm is extracted from
-$\bm{\mathcal{R}}^i$ and it is normalised as follows
-\begin{equation}\label{eq:convergence_res}
-\frac{||\bm{\mathcal{R}}^i||_2}{||\bm{\mathcal{R}}^0||_2}
-\end{equation}
-where $||\bm{\mathcal{R}}^0||_2$ is the $L_2$-norm of the first nonlinear iteration.
-Since often the normalised nonlinear residual drop very quickly over the first 2-3 iterations, also the $L_2$-norm of velocity and pressure residuals are
-calculated as
-\begin{equation}\label{eq:convergence_vel}
-\frac{||\Delta\bm{v}||_2}{||\bm{v}||_2}=\frac{||\bm{v}_i-\bm{v}_{i-1}||_2}{||\bm{v}_{i}||_2}
-\end{equation}
-and
-\begin{equation}\label{eq:convergence_press}
-  \frac{||\Delta p||_2}{||p||_2}=\frac{||p_i-p_{i-1}||_2}{||p_{i}||_2}
-\end{equation}
-respectively. Iterations are performed until either maximal number of non-linear iterations ($it_{max}$) is reached or all the normalised nonlinear residual, 
-the velocity residual and the pressure residual converge under a defined tolerance $tol$.
-
-\section{The energy equation}\label{sec:en_eq}
-The stabilisation of the advection term of the energy equation (Eq. \ref{eq:energy}) needed to avoid possible oscillations in the thermal solution in the case
-when advection dominates over diffusion, is implemented by means of a streamline-upwind Petrov–Galerkin (SUPG) method, for which the advection term is modified
-follows the discussion in \citet{Thieulot2011} and \citet{Thieulot2014} (see Section \ref{sec:advection} for benchmark). The total internal heat production
-$H_{tot}$ in Eq. \ref{eq:energy} includes radiogenic heating $H_r$, shear heating $H_s$ and adiabatic heating $H_a$, where
-\begin{equation}\label{eq:shear}
-H_s=2\eta \dot{\bm{\epsilon}} : \dot{\bm{\epsilon}} = 2\eta(\dot{\epsilon}_{xx}^2+\dot{\epsilon}_{yy}^2+2\dot{\epsilon}_{xy}^2)
-\end{equation}
-and
-\begin{equation}\label{eq:adiabatic}
-H_a=T \alpha \frac{Dp}{Dt} \approx -\alpha T \rho g_y v
-\end{equation}
-The correctness of the implementation and the computation of shear and adiabatic heating is verified performing an exercise with an analytical solution
-(Section \ref{sec:simple_shear}) and Exercise 9.4 in \citet{Gerya2010b} (Section \ref{sec:shear}). Furthermore, mantle convection and thin layer entrainment
-benchmarks (Sections \ref{sec:mantle} and \ref{sec:thin}, respectively) are performed to verify that the code solve correctly Eqs. \ref{eq:momentum_penalty}
-and \ref{eq:energy} in case of a temperature field characterised by an initial perturbation and with a temperature-dependent density. At last, the viscoplastic
-mantle convection benchmark is performed to check also the viscoplasticity in case of a initial perturbed temperature field (Section \ref{sec:visco_mantle}).
-
-\section{Phase transitions and hydration}\label{sec:phase}
-Variations of effective density, specific heat and coefficient of thermal expansion during the evolution of crustal and mantle materials at different
-pressure-temperature ($p$-$T$) conditions are computed by means of version 6.8.6 of Perple\_X software package \citep{Connolly2005}, similarly to the
-implementation described in \citet{Marotta2020}. By default, mineral assemblages and properties of each lithology (oceanic crust, lower and upper continental
-crust, sediments and dry or hydrated mantle) are calculated for temperatures from \SIrange{330}{1600}{\kelvin}, with increments of approximately \SI{4}{\kelvin},
-and pressures from \SIrange{0.1}{30}{\giga\pascal}, with increments of approximately \SI{0.03}{\giga\pascal}, for a total of almost 100000 points.
-
-The introduction of phase transitions produces two main effects that must be taken into account in numerical models: variations in density and release/absorption
-of latent heat \citep{Tackley2010}. Effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into
-account considering effective coefficient of thermal expansion \citep{Christensen1985,Zhong2015}. Similarly, effects of latent heat in the energy equation
-(eq \ref{eq:energy}) are taken into account by considering effective specific heat and coefficient of thermal expansion
-\citep{Christensen1985,Tackley2010,Zhong2015}. Effective density, specific heat and coefficient of thermal expansion are included in Perple\_X files and are
-assumed by each Lagrangian marker in according to its $p$-$T$ conditions.
-
-Hydration processes related to dehydration of subducting lithosphere \citep{Schmidt1998,Liu2007,Faccenda2009,Faccenda2010,Faccenda2014,Rosas2016} strongly
-influence the thermo-mechanics inside the mantle wedge, mainly because of both weakening effects on the mantle rheology and density variations in case of
-mantle serpentinization \citep{Gerya2002,Honda2003,Arcay2005,Roda2010,Regorda2017}. In the case that hydration is switched on, the amount of bound and free
-water is memorised by each marker, following the implementation of \citet{Quinquis2014}; hydration and dehydration processes are related to the amount of bound
-water of each marker and to the maximum amount of water it can transport, i.e. if the amount of bound water exceeds the maximum amount of water, the marker
-dehydrates and releases free water that can hydrates under-saturated markers. Maximum water content of each marker is determined as function of lithology and
-$p$-$T$ conditions and it is calculated using Perple\_X, in the same way than effective density, specific heat and coefficient of thermal expansion. Bound water
-is advected together with the markers, neglecting the effect of bound water diffusion, while free water simply migrates vertically and is not coupled to the
-solid-phase flow of the mantle wedge \citep{Arcay2005,Quinquis2014}. The correctness of the migration scheme of the free water and its absorption into bound
-water by under-saturated markers are verified performing the experiment described in \citet{Quinquis2014} (Sec. \ref{sec:quinquis}).
-
-The mantle viscosity weakening related to the amount of water has been extensively studied \citep{Chopra1981,Hirth2003} and it is implemented in according to
-\citet{Arcay2005} and \citet{Horiuchi2016}, as follows
-\begin{equation}\label{eq:wet_visc}
-\eta_{wet}=\eta_{dry}\left(\left[1-\frac{1}{f_v}\right]\exp\left(-\frac{[OH^-]}{[OH^-]_0}\right)+\frac{1}{f_v}\right)
-\end{equation}
-where $f_v$ is the viscosity reduction factor between dry and wet conditions, $[OH^-]$ is the water content and $[OH^-]_0$ is a reference water content set to
-620 ppm (0.062 wt.\%) \citep{Arcay2005}. By default, $f_v$ is set to 100, in according to \citet{Arcay2005} and \citet{Horiuchi2016}.
-
-\section{Mantle melting}\label{sec:melt}
-Melt fraction in the mantle depends on temperature ($T$, in °C), pressure ($p$, in \si{\giga\pascal}) and the water content in the melt ($X_{H_2O}$, in wt.\%).
-The determination of the melt fraction $M$ is implemented as explained by \citet{Katz2003}, according to the successive modification by \citet{Langmuir2006} and
-\citet{Kelley2010}, as follows
-\begin{eqnarray}\label{eq:melt_fr}
-M(p,T,X_{H_2O})&=&-T+T_{sol}+(x_t\cdot ln(p+y_t))\cdot X_{H_2O}-\nonumber\\
-&&+K_m\left(\frac{X_{H_2O}^{bulk}}{D_{H_2O}(1-X_{H_2O})+X_{H_2O}}\right)^{\gamma}
-\end{eqnarray}
-where $T_{sol}$ is the temperature of the dry solidus, $D_{H_2O}$ is the partition coefficient and $x_t\cdot log(p+y_t)$ indicates the pressure dependence of
-the melting curve \citep{Kelley2010}. $x_t$, $y_t$, $K_m$ and $\gamma$ are parameters chosen in according to \citet{Kelley2010}.
-However, $X_{H_2O}$ depends on the melt fraction as
-\begin{equation}\label{eq:melt_wt}
-X_{H_2O}(M)=\frac{X_{H_2O}^{bulk}}{D_{H_2O}+M(1-D_{H_2O})}
-\end{equation}
-and a numerical solution can be found using a root-finder method \citep{Katz2003,Wang2016}. Furthermore, the content of water in the melt is limited by the
-pressure-dependent saturation concentration of water in the melt, determined as
-\begin{equation}\label{eq:melt_max}
-X_{H_2O}^{sat}=\chi_1 p^{\lambda_m}+\chi_2 p
-\end{equation}
-where $\chi_1$, $\chi_2$ and $\lambda_m$ are parameters chosen according to \citet{Katz2003}, \citet{Langmuir2006} and \citet{Kelley2010}. The correctness in
-the determination of melt fractions and water contents in the melt are tested comparing model results with experimental curves obtained by \citet{Katz2003}
-(Sec. \ref{sec:katz}).
-
-In case of melting, viscous creep viscosity ($\eta_{vp}$) is modified according to \citet{Wang2016}, as
-\begin{equation}\label{eq:melt_visc}
-\eta_{melt}=\eta_{vp}\cdot \exp(\alpha_m M)
-\end{equation}
-where $\alpha_m$ are melt fraction factors for dislocation and diffusion creep, chosen in according to \citet{Wang2016}.
-
-Effective density ($\rho_{eff}$) for partially molten rocks is calculated as
-\begin{equation}\label{eq:melt_dens}
-\rho_{eff}=\rho_s(1-M)+\rho_m M
-\end{equation}
-where $\rho_s$ and $\rho_m$ are the densities of the solid and the molten rock, respectively \citep{Gerya2003a,Wang2016}. The density of the solid rock is
-extracted by Perple\_X, while the density of the molten portion is calculated as
-\begin{equation}\label{eq:melt_dens_M}
-\rho_m(p,T)=\rho_0[1-\alpha(T-T_0)][1+\beta(p-p_0)]
-\end{equation}
-where $\rho_0$ is the density at the reference temperature ($T_0$) and pressure($p_0$), and $\alpha$ and $\beta$ are the thermal expansion and compressibility
-coefficients, respectively \citep{Gerya2003a,Wang2016}.
-
-As for the phase transitions, effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into account
-considering effective coefficient of thermal expansion. Similarly, latent heat due to melting/crystallisation can be implicitly included in the energy equation
-(Eq. \ref{eq:energy}) by considering effective specific heat ($C_{p_{eff}}$) and thermal expansion ($\alpha_{eff}$) of the partially molten rocks, in according to
-\begin{eqnarray}
-\label{eq:cp_effective}C_{p_{eff}}&=&C_p+H_L\left(\frac{\partial M}{\partial T}\right)_p \\
-\label{eq:alpha_effective}\alpha_{eff}&=&\alpha+\rho\frac{H_L}{T}\left(\frac{\partial M}{\partial p}\right)_T
-\end{eqnarray}
-where $C_p$ and $\alpha$ are the specific heat and the coefficient of thermal expansion of the solid rock, respectively, and $H_L$ is the latent heat of the
-molten rock \citep{Gerya2003a,Gerya2007,Tackley2010}.
-
-\section{Benchmarks}\label{sec:benchmark}
-\subsection{Error measurements}\label{sec:error}
-In order to determine the accuracy of velocity and pressure field of the benchmarks, the $L_2$-norm is computed by numerical integration on the Gauss-Legendre
-quadrature points. $L_2$-norm for pressure and velocity errors can be evaluated as
-\begin{eqnarray}
-\textrm{err}_p&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|p^n_i(r_q)-p^a_i(r_q)|^2\right)w_q|J_q|}\nonumber \\
-\textrm{err}_v&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|u^n_i(r_q)-u^a_i(r_q)|^2+|v^n_i(r_q)-v^a_i(r_q)|^2\right)w_q|J_q|}\nonumber
-\end{eqnarray}
-respectively, where $n_e$ is the number of elements, $n_q$ is the number of quadrature points per element, $p^n_i(r_q)$ and $p^a_i(r_q)$ are the numerical and
-analytical pressure, respectively, in each quadrature point $q$, $w_q$  and $J_q$ are the weight and the Jacobian at the quadrature point
-$q$, $u^n_i(r_q)$, $v^n_i(r_q)$, $u^a_i(r_q)$ and $v^a_i(r_q)$ are the numerical and analytical velocities, respectively, in each quadrature point $q$.
-
-Other quantities used as comparison with original benchmarks are the root-mean-square velocity over the whole domain and over the surface
-\begin{eqnarray}
-v_{\textrm{rms}}&=&\sqrt{\int_{\Omega}|\bm{v}|^2 d\Omega}\nonumber \\
-v_{\textrm{rms}}^{top}&=&\sqrt{\int_0^1 u^2 \bigg|_{y=1} dx}\nonumber
-\end{eqnarray}
-respectively, top and bottom Nusselt numbers
-\[\textrm{Nu}^{top/bottom}=-\int_0^1\frac{\partial T}{\partial y} \bigg|_{y=1/y=0} dx\]
-the average rate of work done against gravity
-\[\left\langle W \right\rangle=\int_{\Omega}T u_y d\Omega\]
-the average rate of viscous dissipation
-\[\left\langle \Phi \right\rangle=\int_{\Omega}\tau_{ij}\dot{\epsilon}_{ij} d\Omega\]
-and the percentage error between $\left\langle W \right\rangle$ and $\left\langle \Phi \right\rangle/Ra$
-\[\delta=\frac{\left|\left\langle W \right\rangle -\frac{\left\langle \Phi \right\rangle}{Ra} \right|}{\max \left(\left\langle W \right\rangle,
-\frac{\left\langle \Phi \right\rangle}{Ra}\right)}\times100\]
-
-\subsection{Stokes flow}\label{sec:stokes}
-The problem consists of determining velocity field $(u,v)$ and pressure $p$ in case of a manufactured solution with prescribed body forces such as
-\begin{eqnarray}
-b_1&=&(12-24y)x^4+(-24+48y)x^3+(-48y+72y^2-48y^3+12)x^2+\nonumber \\
-&&+(-2+24y-72y^2+48y^3)x+1-4y+12y^2-8y^3\nonumber \\
-b_2&=&(8-48y+48y^2)x^3+(-12+72y-72y^2)x^2+(4-24y+48y^2-48y^3+\nonumber \\
-&&+24y^4)x-12y^2+24y^3-12y^4\nonumber
-\end{eqnarray}
-for which the exact solution is:
-\begin{eqnarray}
-u(x,y)&=&x^2(1-x)^2(2y-6y^2+4y^3)\nonumber \\
-v(x,y)&=&-y^2(1-y)^2(2x-6x^2+4x^3)\nonumber \\
-p(x,y)&=&x(1-x)-\frac{1}{6}\nonumber
-\end{eqnarray}
-The domain is a unit square a constant viscosity ($\eta=1$) and the penalty parameter is set to $\lambda=10^7$. Velocity boundary conditions are set to no slip
-$(\bm{v}=\bm{0})$ on all boundaries. The problem is performed for different grid resolution between \num{8x8} and \num{1024x1024} elements.
-
-The errors between the analytical solution and the numerical prediction of the velocity and the pressure field are calculated as in Section \ref{sec:error}.
-Fig. \ref{fig:errors}a shows that both velocity and pressure field converge to the exact solution with the decrease of the element size, following the
-theoretical convergence rate. The convergence of pressure, in contrast with the observation by \citet{Donea2003} for $Q_1 \times P_0$ elements, indicate the
-effectiveness of the smoothing procedure. Smoothed pressure field for grid resolution of \num{128x128} elements is shown in Fig. \ref{fig:errors}b (continuous
-red line), in comparison with the analytical solution (dashed blue line). Solve times and memory usage needed to generate the solution are shown in
-Fig. \ref{fig:MUMPS}. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Solver/Stokes_Flow}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/errors.pdf}
-\caption{Velocity and pressure error for the Stokes flow experiment between generated and analytical solution as a function of element size (panel a),
-and comparison between smoothed pressure and analytical solution as function of $x$ coordinate for a grid resolution of \num{128x128} elements (panel b).}
-\label{fig:errors}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/MUMPS.pdf}
-\caption{Panel a) Analysis, factorisation, solution and total solve times (panel a) and factorisation memory usage (panel b) as a function of the total number
-of degrees of freedom for the Stokes flow experiment.}
-\label{fig:MUMPS}
-\end{figure}
-
-\subsection{Markers advection}\label{sec:runge}
-The 2nd-order Runge-Kutta advection scheme in space is tested by means of the Zalesak disk test \citep{Zalesak1979,Thieulot2014}. The benchmark is performed
-in a unit square domain with a grid resolution of \num{32x32} elements and 1000000 markers and values of Courant number between 0.25 and 1.
-At $t=0$, the disk is centred at position $(0.5;0.75)$ with a radius $R=0.15$ and has a vertical fissure 0.05 wide and 0.2 high (orange in Fig. \ref{fig:runge})
-and with the velocity field prescribed in the entire domain as
-\begin{eqnarray}
-u(x,y)&=&2\pi\left(y-\frac{L_x}{2}\right)\nonumber \\
-v(x,y)&=&-2\pi\left(x-\frac{L_x}{2}\right)\nonumber
-\end{eqnarray}
-the markers are back to their initial location after a $2\pi$ rotation ($t=1$). The clockwise rotation of the disk at $t=0.25$, 0.5 and 0.75 
-is shown in Fig. \ref{fig:runge} (red, blue and green, respectively). The distance from the centre of 1 marker is calculated throughout the rotation to
-evaluate the error for different values of Courant number. As expected, the error increase with the increasing of the time step, as shown in 
-Fig. \ref{fig:runge_err}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/Zalesak_Disk}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=200px]{./Figures/Zalesak.pdf}
-\caption{Position of the Zalesak disk at $t=0$, 0.25, 0.5 and 0.75 (orange, red, blue and green, respectively) for the case with Courant number of 1.}
-\label{fig:runge}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/zal_marker.pdf}
-\caption{Distance from the centre as function of time for values of Courant number of  of 0.25, 0.3, 0.5, 0.75 and 1 (green, orange, blue, black and red,
-respectively). R indicates the distance from the centre considering a perfect circle trajectory.}
-\label{fig:runge_err}
-\end{figure}
-
-\subsection{Conservative Velocity Interpolation (CVI)}\label{sec:CVI}
-The CVI correction is checked by means of the Stokes flow experiment by \citet{Donea2003}, as presented in \ref{sec:stokes}. The advection of Lagrangian
-markers is performed using either a 2nd-order or a 4th-order Runge-Kutta scheme with CFL=0.5 and an initial random distribution of 25 markers per element.
-Fig. \ref{fig:CVI} shows the minimum and maximum values of markers per element throughout the experiment with and without the CVI correction and using
-different Runge-Kutta schemes. The introduction of the CVI correction allows the
-tests to run for times longer than $t=2000$ without show empty elements, while tests without the CVI correction stop at approximately $t=600$, when at least
-one element goes empty. Fig. \ref{fig:CVI_time} shows the different number of markers per element and the markers distribution at $t=600$ obtained without and
-with the CVI correction (panels b and c, respectively) in case of a 2nd-order Runge-Kutta scheme. All data can be found at
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/CVI}.
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/CVI.pdf}
-\caption{Maximum and minimum number of markers per element throughout the experiment, with (green and orange lines) or without (purple and blue lines)
-the CVI correction and using either a 2nd-order (blue and orange lines) or a 4th-order (purple and green lines) Runge-Kutta scheme.}
-\label{fig:CVI}
-\end{figure}
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/CVI_time.pdf}
-\caption{Number of markers per element and markers distribution through time using 2nd.order Runge-Kutta scheme.
-Panels a: number of markers per element (panel a1) and distribution of the markers (panel a2) after the first
-iteration; panels b and c: comparison between number of markers per element (panels b1 and c1) and markers distribution (panels b2 and c2) without and with
-the CVI correction at $t=600$; panels d: number of markers per element (panel d1) and markers distribution (panel d2) in case of CVI correction at $t=2000$.}
-\label{fig:CVI_time}
-\end{figure}
-
-\subsection{Poiseuille flow}\label{sec:poiseuille}
-The domain is a rectangle with $L_x=2$ and $L_y=1$ and constant density and viscosity ($\rho=1$ and $\eta=1$), gravity acceleration $\bm{g}=\bm{0}$ and penalty
-parameter $\lambda=10^8$. The grid is composed by \num{40x20} elements. Velocity boundary conditions are set to no slip $(\bm{v}=\bm{0})$ at the top and the
-bottom, and a parabolic profile is imposed on the sides, with $u=y(1-y)$ and $v=0$. The analytical solution is then given by:
-\begin{eqnarray}
-u(x,y)&=&y(1-y)\nonumber \\
-v(x,y,)&=&0\nonumber \\
-p(x,y)&=&2\eta\left(\frac{L_x}{2}-x\right)\nonumber
-\end{eqnarray}
-
-The velocity field predicted by the model follows the expected parabolic profile (Fig. \ref{fig:poiseuille}a and Fig. \ref{fig:poi_plot}a) and, in case of the
-classic penalty method with no iterations, the pressure is clearly related to the divergence by means of the penalty parameter (Fig. \ref{fig:poiseuille}b and
-c and Fig. \ref{fig:poi_plot}b). Fig. \ref{fig:poiseuille}d shows that one Uzawa iteration is sufficient to bring the divergence down to \num{e-15}, with no
-correlation with the pressure field. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Poiseuille%20Flow}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Poiseuille.pdf}
-\caption{Velocity field (panel a), pressure (panel b) and divergence velocity of a Poiseuille flow in case of the classic penalty method (no iterations) and
-after one Uzawa iteration (panel c and d, respectively).}
-\label{fig:poiseuille}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Analytical.pdf}
-\caption{Velocity (panel a) and pressure (panel b) field predicted by the model for a Poiseuille flow with respect with their analytical solutions. The
-velocity field is plotted as function of the vertical coordinate in $L_x/2$ and the pressure is plotted as function of the x coordinate in $L_y/2$.}
-\label{fig:poi_plot}
-\end{figure}
-
-\subsection{Instantaneous 2D sphere}\label{sec:ist_sphere}
-The domain is a unit square with gravity $\bm{g}=(0,-1)$. The fluid has constant density and viscosity ($\rho_f=1$ and $\eta_f=1$). The sphere is in the middle
-of the domain with a radius R = 0.123456798 and has constant density and viscosity ($\rho_s=10^{-2}+\rho_f$ and $\eta_s=10^3 \cdot \eta_f$).
-We distinguish three different types of velocity boundary conditions:
-\begin{enumerate}
-\item FS: free slip conditions on all sides.
-\item NS: no slip conditions on all sides.
-\item OT: free slip conditions on the sides and the bottom, and open boundary at the top.
-\end{enumerate}
-All three cases are performed for different grid resolutions, between \num{16x16} and \num{512x512} elements with 50 randomly distributed markers per element,
-and different average schemes for the viscosity (harmonic, geometric and arithmetic). The velocity in the centre of the sphere $\bm{v}(0.5;0.5)$), minimum and
-maximum velocities ($u_{min}$, $u_{max}$, $v_{min}$, $v_{max}$) and pressures ($p_{min}$, $p_{max}$) on the entire domain, $v_{\textrm{rms}}$ and average pressure on the
-entire domain ($p_{avg}$) are compared with solutions generated by ASPECT \citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} and all results
-can be found at \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere2D} and 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Instantaneous_2D_Sphere}. Results in terms of the $v_{\textrm{rms}}$ for NS model 
-are also shown in Fig. \ref{fig:inst_sphere}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_NS.pdf}
-\caption{$v_{\textrm{rms}}$ as function of element size for the instantaneous 2D sphere experiment in case of no slip boundary conditions and with different average
-schemes for the viscosity. Results are compared with results obtained by other numerical codes.}
-\label{fig:inst_sphere}
-\end{figure}
-
-\subsection{Rayleigh-Taylor instability}\label{sec:rayleigh}
-This problem has been originally presented by \citet{vanKeken1997} and here it is performed as the isoviscous case (Case 1a). The domain has $L_x=0.9142$ and
-$L_y=1$ and gravity $\bm{g}=(0,-1)$. Two fluids with same constant viscosities ($\eta_1=\eta_2=1$) and different densities ($\rho_1=1000$ and $\rho_2=1010$),
-with the lighter fluid at the bottom. The initial interface between the fluids is given by $y(x)=0.2+0.02 \cos \left(\frac{\pi x}{L_x}\right)$. The experiment
-is performed with different grid sizes (\num{50x50}, \num{80x80}, \num{100x100} and \num{256x256}). A total of 1960000 markers are randomly distributed at the
-beginning of the simulation. Velocity boundary conditions are set to no slip at the top and the bottom, and to free slip at the sides of the domain.
-
-$v_{\textrm{rms}}$ as function of time is reported in Fig. \ref{fig:RT}, matching well with results shown by \citet{vanKeken1997}, \citet{Tackley2003} and
-\citet{Thieulot2014}. Fig. \ref{fig:rayleigh} shows the evolution of the experiment at different time steps. All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Rayleigh_Taylor_experiment/ISOVISCOUS}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/RT.pdf}
-\caption{$v_{\textrm{rms}}$ of the Rayleigh-Taylor experiment as function of time for different resolution of the grid.}
-\label{fig:RT}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=300px]{./Figures/Rayleigh.pdf}
-\caption{Evolution of the Rayleigh-Taylor experiment for a grid of \num{256x256} elements at $t=0, 500, 1000$ and $2000$ (panels a, b, c and d, respectively).}
-\label{fig:rayleigh}
-\end{figure}
-
-\subsection{Falling blocks}\label{sec:block}
-This benchmark is described as proposed by \citet{Gerya2003a}, \citet{Gerya2010b} and \citet{Thieulot2011}. The domain is square with $L_x=L_y=$ \SI{500}{\km}
-and the grid is composed by \num{50x50} elements with 25 markers in each element. The block is initially centred at ($x=$ \SI{250}{\km}; $y=$ \SI{400}{\km})
-and has a size of \num{100x100} km (Fig. \ref{fig:block}a). The fluid surrounding the block has $\eta_f=$ \SI{e21}{\pascal\per\s} and $\rho_f=$
-\SI{3200}{\kg\per\cubic\m}. The benchmark tests with different viscosities of the block, with $\eta_b$ from \SIrange{e15}{e27}{\pascal\per\s}. In each
-experiment also the density $\rho_b$ of the block varies from \SIrange{3220}{9900}{\kg\per\cubic\m}. Velocity boundary conditions are set to free slip
-conditions on all sides of the domain. Density distributions at $t=$ \SI{20}{\mega\year} for $\rho_b=$ \SI{3300}{\kg\per\cubic\m} and different $\eta_b$
-are plotted in Fig. \ref{fig:block}b-f, showing that the code correctly preserve the block geometry in case of large viscosity contrast, with a block stiffer
-than the surrounding fluid (Fig. \ref{fig:block}f).
-
-The velocity in the centre of the falling block at $t=0$ is measured for all experiments and, since the velocity should increase with density contrast, the
-quantity $\bm{v}/(\rho_b-\rho_f)$ is plotted as function of the viscosity contrast. All results perfectly match with those from \citet{Gerya2010b} and they
-line up on a single curve, demonstrating that the code can correctly deal with large viscosity and density contrasts (Fig. \ref{fig:falling}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Falling%20blocks}.
-
-\begin{figure}
-\noindent\includegraphics[width=300px]{./Figures/Block.pdf}
-\centering
-\caption{Density evolution of the falling block experiment at $t=$ \SI{0}{\mega\year} (panel a) and $t=$ \SI{20}{\mega\year} for different viscosities of the
-block (panels b-f).}
-\label{fig:block}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Falling.pdf}
-\caption{Initial velocity relative to the density contrast at the centre of the falling block as function of the viscosity contrast between the block and the
-surrounding fluid.}
-\label{fig:falling}
-\end{figure}
-
-\subsection{2D time-dependent Stokes sphere with deformable free surface}\label{sec:sphere}
-This experiment is performed in a unit square domain with the gravity acceleration fixed to $g_y=-1$. The sphere has $\rho_s=2$ and $\eta_s=10^3$ and is
-initially centred at ($0.5;0.6$) with radius $R=0.123456789$. The fluid surrounding the sphere has $\rho_f=1$ and $\eta_f=1$ and occupies the domain for
-$y\leq 0.75$, while for $y>0.75$ the air has $\rho_a=0$ and $\eta_a=10^{-3}$. Velocity boundary conditions are set to free slip on all sides. The Courant
-number is set to 0.25. The experiments run for 200 s using grid resolutions from \num{150x150} to \num{512x512} elements, with 25 randomly distributed markers
-per element, and for different average schemes for the viscosity. The interface between the fluid and the air is tracked by means of the markers chain. Results
-in terms of velocity and pressure fields and topography variations are compared with results obtained with ASPECT
-\citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} and they can be found at
-\url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere_fs2D} and
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Time_dependent_sphere}. Results in terms of the $v_{\textrm{rms}}$ in case of 
-an arithmetic average are also shown in Fig. \ref{fig:inst_spherefs}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_arithm.pdf}
-\caption{$v_{\textrm{rms}}$ for different grid resolutions as function of time for the 2D time-dependent Stokes sphere experiment in case of an arithmetic average.
-Results are compared with results obtained by ASPECT.}
-\label{fig:inst_spherefs}
-\end{figure}
-
-\subsection{Free surface stabilisation}\label{sec:stab}
-The benchmark is performed as discussed in \citet{Kaus2010a} and \citet{Thieulot2014}. The domain is square with $L_x=L_y=$ \SI{500}{\km} and a grid resolution
-of \num{200x200} elements, each of them containing 16 markers. A buoyant fluid with $\rho_1=$ \SI{3200}{\kg\per\cubic\metre} and $\eta_1=$ \SI{e20}{\pascal\s}
-is overlain by a denser fluid with $\rho_2=$ \SI{3300}{\kg\per\cubic\metre} and $\eta_2=$ \SI{e21}{\pascal\s}. The initial interface between the fluids has an
-initial sinusoidal shape of 5 km amplitude. Velocity boundary conditions are set to free slip conditions at the sides and no slip at the bottom, while the top
-is free surface. The experiment is performed with various fixed time steps, without the use of the Courant number. The vertical position of the free surface at
-$x=L_x$ is tracked for each simulation, with or without the stabilisation algorithm.
-
-The results show an instability (drunken sailor effect) increasing the time step in case that the stabilisation algorithm is not activated
-($dt=$ \SI{4200}{\year}, green line in Fig. \ref{fig:stab}), as already observed by \citet{Kaus2010a} and \citet{Thieulot2014}. Activating the algorithm the
-instability is fixed and the simulation remains stable using time step up to \SI{50000}{\year} (red line in Fig. \ref{fig:stab}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Stabilisation_algorithm}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Kaus_random.pdf}
-\caption{Evolution of the y coordinate in $x=L_x$ as function of time for different time steps, with and without the stabilisation algorithm.}
-\label{fig:stab}
-\end{figure}
-
-\subsection{Topography relaxation}\label{sec:crameri}
-This experiment is performed as in \citet{Crameri2012} using free surface and sticky air (Case 1). The domain is rectangular with $L_x=$ \SI{2800}{\km} and
-$L_y=$ \SI{700}{\km} in case of free surface and $L_y=$ \SI{800}{\km} in case of sticky air ($h_{st}=$ \SI{100}{\km}). The grid is composed by \num{256x64}
-elements for the free surface \num{560x320} elements for the sticky air. For both cases, there is a mantle of 600 km thickness with $\eta_m=$ \SI{e21}{\pascal\s}
-overlain by a highly viscous cosine-shaped lithosphere with thickness between 93 and 107 km and $\eta_l=$ \SI{e23}{\pascal\s}. Both mediums have $\rho=$
-\SI{3300}{\kg\per\cubic\metre}. In case of sticky air experiment, the air has $\rho_a=$ \SI{0}{\kg\per\cubic\metre} and $\eta_a=$ \SI{e18}{\pascal\s}. The
-gravity acceleration is set to $g_y=$ \SI{-10}{\metre\per\square\s}. Velocity boundary conditions are set to no slip at the bottom and free slip at the sides
-of the domain. In case of sticky air, velocities on top are set to free slip conditions.
-
-The maximum topography as function of time ($h(t)$) can be analytically derived in according to \[h(t)=h_0 \exp(\gamma t)\]
-where $h_0=7$ km is the initial topography, $\gamma=\num{-0.2139e-11}$ is the characteristic relaxation rate and $t$ is the time (black line in Fig.
-\ref{fig:crameri}). Maximum topography at the characteristic relaxation time $t_{rlx}=$ \SI{14.825}{\kilo\year} can be found to be $h_{rlx}=$ \SI{2576}{\m}.
-The results show that a viscosity of the air of \SI{e18}{\pascal\s} is needed in case of $h_{st}=$ \SI{100}{\km} to correctly describe the topography relaxation
-for this problem, as pointed out by \citet{Crameri2012} (purple line in Fig. \ref{fig:crameri}). The case with a true free surface well follows the analytical
-solution (orange line in Fig. \ref{fig:crameri}). In particular, at the relaxation time in case of the true free surface predicts a maximum topography of
-\SI{2570}{\m}, with an error of \SI{6}{\m} with respect to the analytical solution. Results are also compared with results obtained with a free surface in
-UNDERWORLD (\num{256x64} elements) and SULEC (\num{401x201} elements) (with courtesy from \citealp{Crameri2012}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Topography_relaxation}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Crameri.pdf}
-\caption{Maximum topography as function of time for the Crameri benchmark (coloured lines) in comparison with the analytical solution (black line) and results
-obtained with UNDERWORLD (yellow line) and SULEC (blue line).}
-\label{fig:crameri}
-\end{figure}
-
-\subsection{Spontaneous subduction}\label{sec:subduction}
-This experiment is performed as presented by \citet{Schmeling2008}, in case of both a sticky air and a true free surface and using different average schemes
-for the viscosities. The domain is rectangular with $L_x=$ \SI{3000}{\km} and $L_y=$ \SI{700}{\km}, a grid resolution of \num{256x64} elements with 50 markers
-per element and Courant number of 0.01. At the beginning of the simulation, a 100 km-thick lithospheric layer with $\rho_m=$ \SI{3300}{\kg\per\cubic\metre}
-and $\eta_m=$ \SI{e23}{\pascal\s} is located at the top of the domain for $x$ between 1000 and 3000 km, while the asthenospheric mantle has $\rho_m=$
-\SI{3200}{\pascal\s} and $\eta_m=$ \SI{e21}{\pascal\s}. In addition, a 200 km-depth lithospheric slab is already subducted in the mantle in order to have
-a spontaneous subduction. In case of sticky air, $L_y=750$ km and a 50 km-thick air layer with $\eta_a$ from \SIrange{e19}{e21}{\pascal\s} overlie the mantle.
-Velocity boundary conditions are set to free slip on the sides and on the bottom of the domain. In case of sticky air, velocities on the top boundary are set
-to free slip conditions as well.
-
-The maximum depth of the slab is tracked for all the simulations (Fig. \ref{fig:subduction}) and they are compared with results obtained by
-\citet{Schmeling2008} with models with regular grids and comparable grid resolutions (rectangles in Fig. \ref{fig:subduction}). Low viscous sticky air models
-(i.e., \SI{e19}{\pascal\s}) enlighten a strong dependence of the sinking velocity with the chosen average scheme (dashed lines in Fig. \ref{fig:subduction}),
-while high viscous sticky air models (i.e., \SI{e21}{\pascal\s}) show a higher resistance at the trench, underestimating the correct solution (dotted lines in
-Fig. \ref{fig:subduction}). A high resistance at the trench can be also observed in case of a true free surface for the low resolution grid used (continuous
-lines in Fig. \ref{fig:subduction}), as pointed out by \citet{Schmeling2008}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Spontaneous_subduction}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Subduction.pdf}
-\caption{Maximum depth of the slab as function of time for all the simulations of the \citet{Schmeling2008} benchmark, in case of harmonic, geometric and
-arithmetic means (red, green and blue lines, respectively) using a sticky air or a true free surface (discontinuous and continuous lines, respectively).
-Red, green and blue rectangular areas indicate the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{500}{\km} in case of sticky air,
-while the grey rectangular area indicates the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{600}{\km} in case of true free surface.}
-\label{fig:subduction}
-\end{figure}
-
-\subsection{The slab detachment}\label{sec:slab}
-The slab detachment benchmark is performed as described by \citet{Schmalholz2011} and \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{1000}{\km}
-and $L_y=$ \SI{660}{\km} and a grid resolution of \num{256x256} elements. A non-linear viscous T-shaped layer with $\rho_l=$ \SI{3300}{\kg\per\cubic\metre} is
-placed at the top of the domain and surrounded by a linear viscous fluid with $\rho_f=$ \SI{3150}{\kg\per\cubic\m}. The top layer is 80 km-thick and a 250
-km-long and 80 km-wide slab is placed at $x=L_x/2$ (Fig. \ref{fig:slab}a). The effective viscosity of the top layer is given by
-\[\eta_{eff}=\eta_0 I_2^{\frac{(1-n)}{n}}\]
-with $\eta_0=$ \SI{4.75e11}{\pascal\s} and $n=4$, while the surrounding fluid has $\eta_f=$ \SI{e21}{\pascal\s}. The effective viscosity is capped between
-$\eta_{min}=$ \SI{e21}{\pascal\s} and $\eta_{max}=$ \SI{e25}{\pascal\s}. Velocity boundary conditions are set to free slip at the top and the bottom, and to
-no slip at the sides of the domain.
-
-The sinking of the slab determines a decrease of the effective viscosity in the T-shaped layer during the first part of the evolution (Fig. \ref{fig:slab}a),
-as consequence of local strain rates, while effective viscosities increase up to $\eta_{max}$ in the top layer after approximately \SI{20}{\mega\year}, when
-necking is complete (Fig. \ref{fig:slab}b).
-
-The width of necking is tracked by means of markers position at the side of the slab. Necking width and time are normalised in function of the initial slab
-(\SI{80}{\km}) and the characteristic time ($t_c=$ \SI{7.1158e14}{\s}) \citep{Schmalholz2011,Glerum2018}. The results in case of arithmetic, geometric and
-harmonic mean of the viscosity are shown in Fig. \ref{fig:necking}, compared with results from \citet{Glerum2018} for which an infinity norm mean was used.
-The influence of the average schemes agrees with results shown in Section \ref{sec:subduction}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Slab_detachment}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Slab.pdf}
-\caption{Effective viscosity for the slab detachment benchmark at the beginning of the evolution (panel a) and when necking is complete (panel b).}
-\label{fig:slab}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Necking.pdf}
-\caption{Normalised width of the necking of the slab detachment benchmark as function of normalised time.}
-\label{fig:necking}
-\end{figure}
-
-\subsection{The indenter experiment}\label{sec:indenter}
-The indenter benchmark simulate a rigid punch on a purely plastic von Mises material, for which exists an analytical solution
-\citep{Thieulot2008,Thieulot2014,Glerum2018}. The domain is a unit square with a grid resolution of \num{256x256} elements. The gravity acceleration is set
-to $\bm{g}=\bm{0}$ and effective viscosity is capped using $\eta_{min}=10^{-4}$ and $\eta_{max}=10^3$. The tolerance for the convergence of the non-linear
-solution is $tol=10^{-9}$, with a maximum number of non-linear iterations set to 500. The medium in the domain has $\rho=0.01$, an initial viscosity
-$\eta_0=10$, cohesion $C=1$ and an angle of internal friction $\phi=0$°. Velocity boundary conditions are set to no slip at the bottom and free slip at the
-sides of the domain. The top of the domain is open with the exception of the central portion (punch area) with width $w_p=0.125$, where $v=-1.05$ and $u$ is
-fixed either to 0 (rough punch experiment) or free (smooth punch experiment).
-
-Fig. \ref{fig:indenter} show the results in terms of viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure
-(panels g and h) for both experiments. The analytical solution indicate that pressure at ($0.5;1$) and ($0.5 \pm w_p;1$)is $p=\pi +1$ and $p=1$, respectively
-(black and red lines in Fig. \ref{fig:smooth_p}). Fig. \ref{fig:smooth_p} shows a clear improvement in the pressure solution when passing from a rough to a
-smooth punch (green and blue lines in Fig. \ref{fig:smooth_p}, respectively), as previously observed by \citet{Thieulot2014} and \citet{Glerum2018}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Indenter}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Indenter.pdf}
-\caption{Viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure (panels g and h) fields for rough (left column) and
-smooth (right column) punch experiments.}
-\label{fig:indenter}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Smooth_Rough.pdf}
-\caption{Nodal pressure in function of the x coordinate at $y=1$ for rough (green line) and smooth (blue line) punch experiment. Black and red lines indicate
-the analytical solution at $x=0.5$ and $x=0.5 \pm w_p$, respectively.}
-\label{fig:smooth_p}
-\end{figure}
-
-\subsection{The brick experiment}\label{sec:brick}
-An instantaneous version of the brick benchmark is performed to verify the correctness of the pressure-dependent plasticity for different angle of internal
-friction, as proposed by \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{40}{\km} and $L_y=$ \SI{10}{\km} and a grid resolution of \num{512x128}
-elements. The gravity acceleration is set to $g_y=$ \SI{-10}{\m\square\s} and effective viscosity is capped using $\eta_{min}=$ \SI{e19}{\pascal\s} and
-$\eta_{max}=$ \SI{e26}{\pascal\s}. The tolerance for the convergence of the non-linear solution is $tol=10^{-7}$, with a maximum number of non-linear iterations
-set to 1000. A 800 m-wide and 400 m-height inclusion with $\eta_b=$ \SI{e20}{\pascal\s} and $\rho_b=$ \SI{2700}{\kg\per\cubic\m} is placed at the bottom of the
-domain at $x=L_x/2$. The inclusion is surrounded by a non-linear viscous medium with $\rho_m=$ \SI{2700}{\kg\per\cubic\m}, an initial viscosity of
-\SI{e23}{\pascal\s}, a linear viscous viscosity of \SI{e25}{\pascal\s} and a cohesion of \SI{40}{\mega\pascal}. Velocities are set to free slip conditions at
-the bottom of the domain and the top is open. Velocities on sides of the domain are fixed to $u=$ \SI{+-2e-11}{\m\per\s} and $v=0$. The experiment is performed
-in compressional and extensional contexts with an angle of internal friction $\phi$ of the non-linear medium variable between 0° and 30°.
-
-As expected, two shear bands stem from the inclusion with variable angles in relation to both the dynamics context and the internal friction angle. Shear band
-angles formed at 45° for $\phi=0$° in both extensional and compressional contexts (Fig. \ref{fig:brick_beam}a), while they are different in case of $\phi\neq0$°.
-Shear band angles in case of $\phi=20$° are shown for compression and extension in Fig. \ref{fig:brick_beam}b  and c, respectively. Values of shear band angles
-as function of internal friction angles from 0° to 30° are extracted at different depths and minimum and maximum values are plotted in Fig. \ref{fig:brick},
-in comparison with theoretical Roscoe, Arthur and Coulomb shear band angles. For all tests the tolerance for nonlinear convergence is defined $tol=10^{-7}$
-and none converges before the maximum number of iterations ($it_{max}=1000$) is reached. However, tests with internal friction angles up to 15° show a constant
-decrease of the velocity residuals, which can not be observed for for tests with higher internal friction angles (Fig. \ref{fig:convergence}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Brick_experiment}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Brick_Beam.pdf}
-\caption{Shear band angles predicted for the brick experiment for $\phi=0$° (panel a) and for $\phi=20$° in case of compression and extension (panels b and c,
-respectively).}
-\label{fig:brick_beam}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Brick.pdf}
-\caption{Shear band angles predicted for the brick experiment in case of compressional and extensional contexts as function of different internal angle of
-friction (continuous black line and red dots), compared with theoretical Roscoe, Artur and Coulomb angles (discontinuous black lines). Red lines indicate the
-range of angles calculated at different depths.}
-\label{fig:brick}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \noindent\includegraphics[width=500px]{./Figures/convergence.pdf}
-  \caption{Convergence rates for all tests. Normalised residuals (left column) are compared with velocity residuals (right column) for extensional (first row)
-  and compressional (second row) settings.}
-  \label{fig:convergence}
-  \end{figure}
-
-\subsection{Advection stabilisation}\label{sec:advection}
-The 1D advection problem proposed by \citet{Donea2003} and \citet{Thieulot2011} is performed to verify the effectiveness of the SUPG method to stabilise the
-advection term of the energy equation. The domain is a 1D segment with $L_x=1$ composed by 50 elements and a discontinuity in the thermal field placed at
-$x=0.25$. Temperature is set to 1 for $x \leq 0.25$ and to 0 for $x > 0.25$. Velocity is set to ${u}=1$ in the entire domain. The simulation is performed for
-250 time steps, with $dt=0.002$, so the thermal discontinuity should be at $x=0.75$ at the end of the simulation.
-
-Temperature profiles at the end of the simulation are shown in \ref{fig:advection} as function of the dimensionless coefficient $\gamma=\tau \bm{v}/h$. In case
-of the classic Galerkin method ($\gamma=0$, blue line in \ref{fig:advection}) the final thermal profile is characterised by strong oscillations, which are
-eliminated in case of the SUPG method ($\gamma=0.045$, orange line in \ref{fig:advection}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Advection_stabilisation}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Advection.pdf}
-\caption{Temperature profile as function of \textit{x} for the advection stabilisation benchmark. Purple line indicates the initial temperature profile; the
-green line indicates the analytical temperature profile after 250 time steps; blue line indicates the temperature profile after 250 time steps in case of the
-classic Galerkin method ($\gamma=0$); orange line indicates the temperature profile after 250 time steps in case of the SUPG method ($\gamma=0.045$).}
-\label{fig:advection}
-\end{figure}
-
-\subsection{Shear heating}\label{sec:simple_shear}
-This exercise is performed in an unit square domain composed by \num{128x128} elements. The velocity field is prescribed on the entire domain with
-$\bm{v}=(L_y-y)y\bm{e_x}$; viscosity, density and specific heat are fixed to 1, while thermal conductivity and radiogenic energy are fixed to 0. Therefore,
-the energy equation (Eq. \ref{eq:energy}) can be simplified as \[\frac{\partial T}{\partial t}=H_s\] and fixing $T(t=0)=0$, the temperature can be find as
-\[T(t)=H_s t\]
-In this case we have
-\begin{eqnarray}
-\dot{\epsilon_{xx}}&=&\frac{\partial u}{\partial x}=0 \nonumber \\
-\dot{\epsilon_{yy}}&=&\frac{\partial v}{\partial y}=0 \nonumber \\
-\dot{\epsilon_{xy}}&=&\frac{1}{2}\left(\frac{\partial v}{\partial x}+\frac{\partial u}{\partial y}\right)=\frac{1}{2}(L_y-2y)\nonumber
-\end{eqnarray}
-and, simplifying Eq. \ref{eq:shear}, shear heating can be calculated as \[H_s(x,y)=(1-2y)^2\]
-
-The solution predicted by the model in terms of velocity, temperature and shear heating match well with the analytical solutions (Fig. \ref{fig:analytical_en}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Simple_shear}.
-
-%\begin{figure}
-%\centering
-%\noindent\includegraphics[width=400px]{./Figures/Energy_Thieulot.pdf}
-%\caption{Temperature and shear heating predicted by the model after 1 time step.}
-%\label{fig:en_thieulot}
-%\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=500px]{./Figures/Analytical_Shear.pdf}
-\caption{Velocity (panel a), temperature (panel b) and shear heating (panel c) as function of the \textit{y} coordinate for the simple shear experiment. The
-solutions predicted by the model (red lines) are compared with the analytical solutions (dashed blue lines).}
-\label{fig:analytical_en}
-\end{figure}
-
-\subsection{Shear and adiabatic heating}\label{sec:shear}
-This problem is performed as presented in Exercise 9.4 in \citet{Gerya2010b}. Two materials are vertically separated in a rectangular domain with $L_x=$
-\SI{1000}{\km}, $L_y=$ \SI{1500}{\km} and a grid resolution of \num{30x20} elements. Constant thermal coefficient expansion ($\alpha=$ \SI{3e-5}{\per\kelvin}),
-temperature ($T=$ \SI{1300}{\kelvin}) and gravity acceleration ($g_y=$ \SI{-10}{\m\square\s}) are assumed in the whole domain. Fluid 1 (on the left side of the
-domain) has $\rho_1=$ \SI{3200}{\kg\per\cubic\m} and $\eta_1=$ \SI{e20}{\pascal\s}; fluid 2 (on the right side of the domain) has $\rho_2=$
-\SI{3300}{\kg\per\cubic\m} and $\eta_2=$ \SI{e22}{\pascal\s}. Velocity boundary conditions are set to free slip on all sides of the domain.
-
-As shown in Fig. \ref{fig:energy}, both shear and adiabatic heating predicted by the code (first row in Fig. \ref{fig:energy}) well recreate the results
-obtained by means of example ${Shear\_adiabatic\_heating.m}$ from \citet{Gerya2010b} (second row in Fig. \ref{fig:energy}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Adiabatic%2BShear}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Energy.pdf}
-\caption{Comparison between shear (first column), adiabatic (second column) and total (third column) energy predicted by the code (panels a, b and c) and those
-created using example ${Shear\_adiabatic\_heating.m}$ from Exercise 9.4 in \citet{Gerya2010b} (panels d, e and f).}
-\label{fig:energy}
-\end{figure}
-
-\subsection{Mantle convection}\label{sec:mantle}
-This problem is performed as presented by \citet{Blankenbach1989} (constant viscosity cases) and \citet{Thieulot2014}, in a 2D unit square domain with gravity
-acceleration $g_y=10^{10}Ra$. The experiment is performed with three different Rayleigh numbers ($Ra=10^4$, $10^5$ and $10^6$) and with different grid
-resolution (between \num{32x32} and \num{128x128} elements). The fluid has constant viscosity, initial density, heat capacity, thermal conductivity
-($\eta= \rho_0= C_p = k=1$), reference temperature ($T_0=0$) and thermal expansion coefficient ($\alpha=10^{-10}$). Temperatures are set to 0 on top and 1 on
-bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\]
-
-The solution generated by the code in terms of $v_{\textrm{rms}}$ (calculated as in Sec. \ref{sec:error}) and Nu as function of time are reported for all the
-simulations in Fig. \ref{fig:mantle}. At the steady state, $v_{\textrm{rms}}$ and Nu of all simulations converge well toward the values from \citet{Blankenbach1989},
-with lower errors for higher resolution grids (Table \ref{tab:mantle}). All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Mantle_convection}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Convection.pdf}
-\caption{$v_{\textrm{rms}}$ (panels a, c and e) and Nu (panels b, d and f) for the mantle convection benchmark as function of time for different grid resolution.
-Panels a and b show the results for $Ra=10^4$; Panels c and d show the results for $Ra=10^5$; panels e and f show the results for $Ra=10^6$. Purple lines
-indicate the convergence values for the $v_{\textrm{rms}}$ and the Nu from \citet{Blankenbach1989} at the steady state.}
-\label{fig:mantle}
-\end{figure}
-
-\begin{table}[H]
-\caption{Comparison between $v_{\textrm{rms}}$ and Nu predicted by the code for the mantle convection experiment and same values as reported in literature.}
-\centering
-\small
-\begin{tabular}{|>{\centering}m{0.09\textwidth}|>{\centering}m{0.05\textwidth}|>{\centering}m{0.24\textwidth}|>{\centering}m{0.15\textwidth}|
-  >{\centering}m{0.1\textwidth} >{\centering}m{0.1\textwidth} >{\centering\arraybackslash}m{0.1\textwidth}|}
-\toprule
-& & \multirow{2}{*}{\citet{Blankenbach1989}} & \citet{Thieulot2014} & \multicolumn{3}{c|}{FALCON}  \\
- & & & \num{200x200} & \num{32x32} & \num{64x64} & \num{128x128}  \\
-\midrule
-\multirow{2}{*}{$Ra=10^4$} & $v_{\textrm{rms}}$ & $42.864947 \pm 0.000020$ & 42.867 & 42.83226 & 42.852793 & 42.861394  \\
-           & Nu      & $4.884409 \pm 0.000010$  & 4.882  & 4.781297 & 4.857475  & 4.877573   \\
-\hline
-\multirow{2}{*}{$Ra=10^5$} & $v_{\textrm{rms}}$ & $193.21454 \pm 0.00010$  & 193.255& 193.872643 & 193.377472 & 193.252290  \\
-           & Nu      & $10.534095 \pm 0.000010$ & 10.507 & 9.602514 & 10.270735  & 10.465629   \\
-\hline
-\multirow{2}{*}{$Ra=10^6$} & $v_{\textrm{rms}}$ & $833.98977 \pm 0.00020$  & 834.712& 848.091176 & 837.767911 & 834.945793  \\
-           & Nu      & $21.972465 \pm 0.000020$ & 21.695 & 15.999266 & 19.703682  & 21.306939   \\
-\bottomrule
-\end{tabular}
-\label{tab:mantle}
-\end{table}
-
-\subsection{Viscoplastic mantle convection}\label{sec:visco_mantle}
-This benchmark is performed as Cases 1-5a described in \citet{Tosi2015} in a 2D unit square domain with two different grid resolutions of \num{32x32} 
-and \num{100x100} elements. Reference viscosity, reference density, heat capacity, thermal conductivity and thermal expansion coefficient are fixed to a constant
-value of 1 ($\eta_0= \rho_0= C_p = k=\alpha=1$), while the gravity has been chosen as $g_y=10^2$, in order to obtain a Rayleigh number $Ra=10^2$. Temperatures
-are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\].
-The viscosity ﬁeld $\eta$ is calculated as the harmonic average between a linear part $\eta_{lin}$ that depends on temperature only or on temperature and 
-depth $y$ and a nonlinear-plastic part $\eta_{pl}$ that depends on the strain rate $\dot{\bm{\epsilon}}$, as follows
-\[\eta(T,y,\dot{\bm{\epsilon}})=2\left(\frac{1}{\eta_{lin}(T,y)}+\frac{1}{\eta_{pl}(\dot{\bm{\epsilon}})}\right)^{-1}\]
-The linear and the nonlinear-plastic parts of the viscosity are calculated following \citet{Tosi2015} as follows
-\[\eta_{lin}(T,y)=\exp(-\ln(\Delta\eta_T)T+\ln(\Delta\eta_y)y)\]
-\[\eta_{pl}(\dot{\bm{\epsilon}})=\eta^*+\frac{\sigma_Y}{\sqrt{\dot{\bm{\epsilon}}:\dot{\bm{\epsilon}}}}\]
-where $\Delta\eta_T$, $\Delta\eta_y$, $\eta^*$ and $\sigma_Y$ are parameters chosen as listed in Table \ref{tab:tosi_setup}.
-
-For each case, some diagnostic quantities measured at the steady state are reported in Table \ref{tab:visco_mantle} (see Section \ref{sec:error} for 
-the computation of all the quantities), showing a well fit with results described in \citet{Tosi2015}. Data and charts for all cases can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Viscoplastic_mantle_convection}.
-
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{white!100!}{gray!25!}
-%\begin{table}[H]
-%  \caption{Results of the viscoplastic mantle convection benchmark for Cases 1-4 compared with results obtained with ELEFANT and YACC (from \citealp{Tosi2015}).}
-%  \centering
-%  \resizebox{\textwidth}{!}{
-%\begin{tabular}{|c|ccc|ccc|ccc|ccc|}
-%  \toprule
-%    & \multicolumn{3}{c|}{\textbf{Case 1}} & \multicolumn{3}{c|}{\textbf{Case 2}} & \multicolumn{3}{c|}{\textbf{Case 3}} &
-%    \multicolumn{3}{c|}{\textbf{Case 4}} \\
-%   & FALCON & ELEFANT& YACC & FALCON& ELEFANT& YACC & FALCON & ELEFANT & YACC & FALCON & ELEFANT & YACC \\
-%  \midrule
-%   $\left\langle T \right\rangle$ & 0.7759 & 0.7758 & 0.7767 & 0.6043 & 0.6036 & 0.6058 & 0.7275 & 0.7275 & 0.7286 & & 0.5277 & 0.5289\\
-%    Nu$^{top}$ & 3.4280 & 3.4214 & 3.4298 & 8.5242 & 8.5115 & 8.5278 & 3.0368 & 3.0347 & 3.0374 && 6.5912 & 6.5572 \\
-%    Nu$^{bottom}$ & 3.3168 & 3.313 & 3.3143 & 8.4599 & 8.4465 & 8.3990 & 3.0086 & 2.9908 & 2.9628 && 6.5834 & 6.5243 \\
-%    $v_{\textrm{rms}}$ & 249.901 & 249.134 & 251.7997 & 141.2451 & 140.7087 & 142.2020 & 100.2776 & 100.1208 & 100.9467 && 79.1105 & 79.6202 \\
-%    $v_{\textrm{rms}}^{top}$ & 1.8907 & 1.8642 & 1.8298 & 104.7928 & 104.1228 & 105.1767 & 2.0832 & 2.0652 & 2.0374 && 74.7596 & 75.8414 \\
-%    $u_{max}^{top}$ & 2.6340 & 2.6119 & 2.5516 & 121.9212& 121.766 & 122.3238 & 2.9086 & 2.9019 & 2.8458 && 88.9146 & 89.2940 \\
-%    $\eta_{min}$ & - & - & - & $2.0303\times10^{-5}$ & $2.0600\times10^{-5}$ & $1.9853\times10^{-5}$ & $4.8129\times10^{-5}$ & $4.8080\times10^{-5}$ & $4.7907\times10^{-5}$ &
-%    & $1.9860\times10^{-4}$ & $1.9174\times10^{-4}$ \\
-%    $\eta_{max}$ & - & - & - & 1.3325 & 1.85 & 1.6095 & 0.9360 & 0.9023 & 1 && 1.52 & 1.6773 \\
-%    $\left\langle W \right\rangle$ & 2.4299 & 2.4316 & 2.4583 & 7.5648 & 7.5665 & 7.6222 & 2.0387 & 2.0384 & 2.0400 && 5.6216 & 5.6512 \\
-%    $\left\langle \Phi \right\rangle/Ra$ & 2.4311 & 2.4276 & 2.4333 & 7.5661 & 7.5585 & 7.6194 & 2.0386 & 2.0370 & 2.0335 && 5.6182 & 5.6463 \\
-%    $\delta$ & 0.05\% & 0.16\% & 1.02\% & 0.02\% & 0.11\% & 0.04\% & $<$0.01\% & 0.07\% & 0.32\% && 0.06\% & 0.09\% \\
-%  \bottomrule
-%  \end{tabular}}
-%  \label{tab:visco_mantle}
-%  \end{table}
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{2}{white!100!}{gray!25!}
-\begin{table}[H]
-  \caption{Benchmark cases and corresponding parameters.}
-  \centering
-  \small
-\begin{tabular}{|>{\centering}m{0.08\textwidth}|>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}
-  >{\centering}m{0.05\textwidth} >{\centering}m{0.05\textwidth} |>{\centering\arraybackslash}m{0.2\textwidth}|}
-  \toprule
-    Case & $Ra$ & $\Delta\eta_T$ & $\Delta\eta_y$ & $\eta^*$ & $\sigma_Y$ & Convective regime \\
-  \midrule
-  \rowcolor{gray!25!}
-    1 & $10^2$ & $10^5$ & 1 & - & - & Stagnant lid \\
-    2 & $10^2$ & $10^5$ & 1 & $10^{-3}$ & 1 & Mobile lid  \\
-    \rowcolor{gray!25!}
-    3 & $10^2$ & $10^5$ & 10 & - & - & Stagnant lid  \\
-    4 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 1 & Mobile lid \\
-    \rowcolor{gray!25!}
-    5 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 4 & Periodic \\
-  \bottomrule
-  \end{tabular}
-  \label{tab:tosi_setup}
-  \end{table}
-
-\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{gray!25!}{white!100!}
-\begin{table}
-  \caption{Results of the viscoplastic mantle convection benchmark for different grid resolutions in Cases 1-5 compared with results obtained with ELEFANT
-  and YACC ($100\times100$ elements, from \citealp{Tosi2015}) and in Stone 28 ($32\times32$ elements, from
-  \url{https://github.com/cedrict/fieldstone/tree/master/python_codes/fieldstone_28}).}
-  \centering
-  \resizebox{\textwidth}{!}{
-\begin{tabular}{|cc|ccccccccccc|}
-  \toprule
-  \multicolumn{13}{|c|}{\textbf{Case 1}} \\
-  \midrule
-  Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-  & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-  \hline
-  \rowcolor{gray!25!}
-    YACC & $100\times100$ & 0.7767 & 3.4298 & 3.3143 & 251.7997 & 1.8298 & 2.5516 & - & - & 2.4583 & 2.4333 & 1.02\% \\
-    ELEFANT & $100\times100$ & 0.7758 & 3.4214 & 3.313 & 249.134 & 1.8642 & 2.6119 & - & - & 2.4316 & 2.4276 & 0.16\% \\
-    \rowcolor{gray!25!}
-    Stone 28 & $32\times32$ & 0.7737 & 3.3987 & - & 243.872 & - & - & - & - & - & - & - \\
-%    FALCON & $32\times32$ & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & - & - & 2.4745 & 2.4864 & 0.48\% \\
-%    \rowcolor{gray!25!}
-    FALCON & $100\times100$ & 0.7759 & 3.4280 & 3.3168 & 249.901 & 1.8907 & 2.6340 & - & - & 2.4299 & 2.4311 & 0.05\% \\
-    \midrule
-    \multicolumn{13}{|c|}{\textbf{Case 2}} \\
-    \midrule
-    Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-    & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-    \hline
-    \rowcolor{gray!25!}
-     YACC & $100\times100$ & 0.6058 & 8.5278 & 8.3990 & 142.2020 & 105.1767 & 122.3238 & $1.9853\times10^{-5}$ & 1.6095 & 7.6222 & 7.6194 & 0.04\% \\
-     ELEFANT & $100\times100$ & 0.6034 & 8.5115 & 8.4465 & 140.8067 & 104.1228 & 121.766 & $2.0600\times10^{-5}$ & 1.85 & 7.5665 & 7.5585 & 0.11\% \\
-     \rowcolor{gray!25!}
-     Stone 28 & $32\times32$ & 0.6052 & 8.1715 & - & 141.518 & - & - & - & - & - & - & - \\
-%     FALCON & $32\times32$ & 0.6454 & 8.1422 & 7.4020 & 156.6908 & 95.7226 & 116.9302 & $2.2572\times10^{-5}$ & 0.7194 & 7.4289 & 7.4452 & 0.22\% \\
-%     \rowcolor{gray!25!}
-     FALCON & $100\times100$ & 0.6044 & 8.5266 & 8.4590 & 141.3684 & 104.8867 & 122.0470 & $2.0333\times10^{-5}$ & 1.2914 & 7.5707 & 7.5718 & 0.01\% \\
-     \midrule
-     \multicolumn{13}{|c|}{\textbf{Case 3}} \\
-     \midrule
-     Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-     & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-     \hline
-     \rowcolor{gray!25!} 
-     YACC & $100\times100$ & 0.7286 & 3.0374 & 2.9628 & 100.9467 & 2.0374 & 2.8458 & $4.7907\times10^{-5}$ & 1 & 2.0400 & 2.0335 & 0.32\% \\
-      ELEFANT & $100\times100$ & 0.7275 & 3.0347 & 2.9908 & 100.1208 & 2.0652 & 2.9019 & $4.8080\times10^{-5}$ & 0.9023 & 2.0384 & 2.037 & 0.07\% \\
-      \rowcolor{gray!25!}
-      Stone 28 & $32\times32$ & 0.7277 & 3.0347 & - & 100.018 & - & - & - & - & - & - & - \\
-%      FALCON & $32\times32$ & 0.7281 & 3.0618 & 2.8132 & 101.0092 & 2.1630 & 3.0178 & $4.8500\times10^{-5}$ & 0.7845 & 2.0699 & 2.0695 & 0.02\% \\
-%      \rowcolor{gray!25!}
-      FALCON & $100\times100$ & 0.7275 & 3.0368 & 3.0086 & 100.2776 & 2.0832 & 2.9086 & $4.8129\times10^{-5}$ & 0.9360 & 2.0387 & 2.0386 & $<$0.01\% \\
-      \midrule
-      \multicolumn{13}{|c|}{\textbf{Case 4}} \\
-      \midrule
-      Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-      & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-      \hline
-      \rowcolor{gray!25!} 
-      YACC & $100\times100$ & 0.5289 & 6.5572 & 6.5243 & 79.6202 & 75.4814 & 89.2940 & $1.9174\times10^{-4}$ & 1.6773 & 5.6512 & 5.6463 & 0.09\% \\
-       ELEFANT & $100\times100$ & 0.5277 & 6.5912 & 6.5834 & 79.1105 & 74.7596 & 88.9146 & $1.9860\times10^{-4}$ & 1.5200 & 5.6216 & 5.6182 & 0.06\% \\
-       \rowcolor{gray!25!}
-       Stone 28 & $32\times32$ & 0.5286 & 6.4036 & - & 79.4746 & - & - & - & - & - & - & - \\
-%       FALCON & $32\times32$ & 0.5334 & 6.4679 & 6.3435 & 80.6153 & 76.0511 & 90.2058 & $2.0393\times10^{-4}$ & 0.6453 & 5.6676 & 5.6698 & 0.04\% \\
-%       \rowcolor{gray!25!}
-       FALCON & $100\times100$ & 0.5279 & 6.5897 & 6.5830 & 79.1220 & 75.0715 & 88.8484 & $1.9693\times10^{-4}$ & 1.4133 & 5.6127 & 5.6113 & 0.03\% \\
-       \midrule
-       \multicolumn{13}{|c|}{\textbf{Case 5}} \\
-       \midrule
-       Code & Resolution & $P$ & $\left\langle T \right\rangle_{min}$ & $\left\langle T \right\rangle_{max}$ & Nu$^{top}_{min}$ & Nu$^{top}_{max}$ &
-       $v_{\textrm{rms}}^{min}$ & $v_{\textrm{rms}}^{max}$ & \multicolumn{2}{c}{$(\left\langle \Phi \right\rangle/Ra)_{min}$} & 
-       \multicolumn{2}{c|}{$(\left\langle \Phi \right\rangle/Ra)_{max}$} \\
-       \hline
-       \rowcolor{gray!25!}
-         YACC & $100\times100$ & 0.0792 & 0.6543 & 0.6722 & 2.6950 & 7.2792 & 41.9813 & 98.8230 & \multicolumn{2}{c}{1.4347} & 
-         \multicolumn{2}{c|}{9.2748} \\
-         ELEFANT & $100\times100$ & 0.076 & 0.6514 & 0.6693 & 2.6808 & 7.3506 & 41.2578 & 100.089 & \multicolumn{2}{c}{1.3324} & 
-         \multicolumn{2}{c|}{9.5289} \\
-         \rowcolor{gray!25!}
-         Stone 28 & $32\times32$ & 0.0764 & 0.6521 & 0.6697 & 2.6912 & 7.0763 & 41.61 & 98.183 & \multicolumn{2}{c}{-} & 
-         \multicolumn{2}{c|}{-} \\
-     %    FALCON & $32\times32$ & & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & \multicolumn{2}{c}{-} & 
-     %     \multicolumn{2}{c|}{-} \\
-     %    \rowcolor{gray!25!}
-         FALCON & $100\times100$ & 0.0775 & 0.6519 & 0.6697 & 2.6847 & 7.3629 & 41.5256 & 99.5466 & \multicolumn{2}{c}{1.3399} & 
-         \multicolumn{2}{c|}{9.3993} \\
-  \bottomrule
-  \end{tabular}}
-  \label{tab:visco_mantle}
-  \end{table}
-
-\subsection{Thin layer entrainment}\label{sec:thin}
-This experiment is performed as originally proposed by \citet{vanKeken1997} with two incompressible fluids in a rectangular domain with $L_x=2$ and $L_y=1$
-and gravity acceleration $g_y=10^{10}Ra$. The Rayleigh number and the compositional Rayleigh number are fixed to $Ra=300000$ and $Rc=450000$, respectively.
-Both fluids have constant viscosity, thermal conductivity, specific heat ($\eta=\rho=C_p=1$) and thermal expansion coefficient ($\alpha=10^{-10}$). Fluid 1
-has a density $\rho_1=1$, while fluid 2 is denser ($\rho_2=\rho_1+1.5 \cdot 10^{-10}$) and is placed at the bottom of the domain, for $y \leq 0.025$.
-Temperature are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides of the domain. The initial
-temperature field is given by
-\[T(x,y)=T_u(x,y)+T_l(x,y)+T_r(x,y)+T_s(x,y)-\frac{3}{2}\]
-with
-\begin{eqnarray}
-T_u(x,y)&=&\frac{1}{2}erf\left(\frac{1-y}{2}\sqrt{\frac{u_0}{x}}\right)\nonumber \\
-T_l(x,y)&=&1-\frac{1}{2}erf\left(\frac{y}{2}\sqrt{\frac{u_0}{L_x-x}}\right)\nonumber \\
-T_r(x,y)&=&\frac{1}{2}+\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{y+1}} \exp\left(-\frac{x^2u_0}{4y+4}\right)\nonumber \\
-T_s(x,y)&=&\frac{1}{2}-\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{2-y}} \exp\left(-\frac{(L_x-x)^2u_0}{8-4y}\right)\nonumber
-\end{eqnarray}
-and
-\begin{eqnarray}
-u_0&=&\frac{L_x^{7/3}}{(1+L_x^4)^{2/3}}\left(\frac{Ra}{2\sqrt{\pi}}\right)^{\frac{2}{3}}\nonumber \\
-Q&=&2\sqrt{\frac{L_x}{\pi u_0}}\nonumber
-\end{eqnarray}
-
-The experiment is performed with two grid resolution (\num{125x40} and \num{200x80} elements), with 100 markers per element and CFL of 0.25. $v_{\textrm{rms}}$
-(see Sec. \ref{sec:error}) calculated for both the simulations match well with results from \citet{vanKeken1997} and obtained with ASPECT
-\citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} and ELEFANT \citep{Thieulot2014}, obtained with similar grid resolutions
-(Fig. \ref{fig:thin}). In addition, a variety of simulations with different resolutions and aspect ratios are performed to verify that the sensitivity of the
-initial velocity field. Results match well with results from \citet{vanKeken1997} (in the grey area) and \citet{Thieulot2014} (Fig. \ref{fig:thin_initial}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Thin_layer}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_thin.pdf}
-\caption{$v_{\textrm{rms}}$ for the thin layer experiment as function of time for different grid resolution. Results are compared with results obtained by
-\citet{vanKeken1997} (black lines) and with ASPECT and ELEFANT.}
-\label{fig:thin}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrmszero.pdf}
-\caption{$v_{\textrm{rms}}$ at $t=0$ for the thin layer experiment as function of the element size for different aspect ratios of the elements. Results are compared
-with results obtained by \citet{vanKeken1997} (black dots in the grey area) and with ASPECT and ELEFANT.}
-\label{fig:thin_initial}
-\end{figure}
-
-\subsection{Sinking hydrated cylinder}\label{sec:quinquis}
-This experiment is performed as explained in \citet{Quinquis2014} and simulates the sinking of a cold, hydrated cylinder into a warm, dry mantle. The domain
-is square with $L_x=L_y=$ \SI{300}{\km}, a grid resolution of \num{300x300} elements and 25 markers per element. The sinking cylinder is located at $x=$
-\SI{150}{\km} and $y=$ \SI{170}{\km} with a radius of \SI{20}{\km} and has $\rho_c=$ \SI{3250}{\kg\per\cubic\m}, $\eta_c=$ \SI{e23}{\pascal\s}, a thermal
-conductivity $k_c=$ \SI{4.5}{\watt\per\m\per\kelvin},a specific heat $C_{p_c}=$ \SI{1250}{\joule\per\kg\per\kelvin}; the surrounding mantle has $\rho_m=$
-\SI{3200}{\kg\per\cubic\m}, $\eta_m=$ \SI{e20}{\pascal\s}, a thermal conductivity $k_m=$ \SI{105}{\watt\per\m\per\kelvin} and specific heat $C_{p_m}=$
-\SI{1250}{\joule\per\kg\per\kelvin}; a 58 km-thick lithosphere overlies the mantle and has $\rho_l=$ \SI{3200}{\kg\per\cubic\m}, $\eta_l=$ \SI{e23}{\pascal\s},
-a thermal conductivity $k_l=$ \SI{4.5}{\watt\per\m\per\kelvin}, specific heat $C_{p_l}=$ \SI{750}{\joule\per\kg\per\kelvin} and a thermal diffusivity of
-\SI{e-6}{\square\m\per\s}. Velocity boundary conditions are set to free slip on all sides of the domain. Temperature increase linearly in the lithosphere,
-from 0° to 1300°C, and in the mantle, from 1300° to 1360.5°C. The initial temperature of the cylinder is fixed to 400°C. Throughout the evolution, temperature
-is fixed to 0° and 1360.5°C at the top and bottom of the domain, respectively. The initial bound water content is imposed at 2 wt.\% in the cylinder and at
-0 wt.\% in both the mantle and the lithosphere. The maximum amount of water in the mantle is fixed to 0.2 wt.\%, while maximum water content in the cylinder
-and in the lithosphere is function of pressure and temperature and is calculated using a serpentinized harzburgite with Perple\_X, as in \citet{Quinquis2014}.
-
-As shown in \citet{Quinquis2014}, the model is characterised by a progressive dehydration in the external portion of the cylinder, due to the increase of
-temperature, with a consequent vertical migration of free water that hydrates the mantle above the cylinder, up to the lithosphere. The distribution of
-bound and free water during the evolution of the experiment is shown in Fig. \ref{fig:quinquis} (left and right columns, respectively). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Hydration}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Hydration.pdf}
-\caption{Distribution of bound and free water (left and right columns, respectively) at $t=0.5$ and $t=$ \SI{1.25}{\mega\year} (first and second rows,
-respectively) for the sinking hydrated cylinder experiment. Only markers with bound or free water not equal to 0 are plotted.}
-\label{fig:quinquis}
-\end{figure}
-
-\subsection{Mantle melting curves}\label{sec:katz}
-This instantaneous experiment is performed in order to recreate isobaric melting curves, as function of temperature, bulk water and content of water in the
-melt, to compare with melting curves obtained by \citet{Katz2003}. The domain is square with $L_x=L_y=100$ km, a grid resolution of \num{100x100} elements and
-25 markers per element. Density is set in the entire domain to \SI{3300}{\kg\per\cubic\m} and temperature increases from the sides to the centre from 1000° to
-1500°C. The experiment is performed for bulk water contents from 0 to 0.3 wt.\%.
-
-Fig. \ref{fig:fraction} shows isobaric melting curves for \SI{1}{\giga\pascal} (panel a) and \SI{3}{\giga\pascal} (panel b) calculated for different bulk water
-contents. The curves well match with those obtained for the mantle by \citet{Katz2003}, showing the saturation at the solidus for bulk water of 0.2 and 0.3 wt.\%, indicating
-that the code correctly determines melt fraction as function of pressure, temperature and bulk water content. Same results can be observed in Fig.
-\ref{fig:water}a, where isobaric and isothermal melting curves are shown as function of bulk water content. In addition, Fig. \ref{fig:water}b shows that the
-code correctly determines also the percentage of water in the melt as function of pressure, temperature and melt fraction. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Melting}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Fraction.pdf}
-\caption{Isobaric melting curves for 1 GPa (panel a) and 3 GPa (panel b) as function of temperature with different bulk water contents.}
-\label{fig:fraction}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Water.pdf}
-\caption{Isobaric melting curves for 1.5 GPa as function of bulk water (panel a) and water in the melt (panel b) calculated for different temperatures.}
-\label{fig:water}
-\end{figure}
-
-\clearpage
-
-\bibliography{export}
-
-\end{document}

### 2022-09-07 tex/melt.tex
update main

--- 2022-03-31 FALCON_description.tex
@@
-\documentclass[hidelinks,10pt,a4paper]{article}
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{textcomp}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage{booktabs}
-\usepackage{multirow}
-\usepackage[cm]{fullpage}
-\usepackage[round]{natbib}
-\bibliographystyle{abbrvnat}
-\usepackage{doi}
-\usepackage{bm}
-\usepackage[table]{xcolor}
-\hypersetup{
-    colorlinks,
-    linkcolor={blue},
-    citecolor={blue},
-    urlcolor={blue}
-}
-\usepackage{float}
-\restylefloat{table}
-\usepackage{siunitx}
-\DeclareSIUnit\year{yr}
-
-\begin{document}
-
-\section{Introduction}\label{sec:intro}
-This document includes a description of the code FALCON, with a presentation of the features implemented followed by results of extensive benchmarks listed in
-Table \ref{tab:list}, performed to verify that the code solves correctly a wide range of problems. Description and results of all benchmarks are shown
-in Section \ref{sec:benchmark}.
-
-\begin{table}[H]
-\caption{List of benchmarks performed for each feature of the code. Data of all benchmarks can be found at 
-\url{https://github.com/aleregorda/Benchmarks}}
-\centering
-\begin{tabular}{| l | l | l |}
-\toprule
- \textbf{Feature}  & \textbf{Benchmark} & \textbf{Section} \\
-\midrule
-  Solver  & Stokes flow & \ref{sec:stokes} \\
-\hline
-  \multirow{2}{*}{Markers advection} & Zalesak disk & \ref{sec:runge} \\
-    & Conservative Interpolation Velocity & \ref{sec:CVI} \\
-\hline
-  \multirow{4}{*}{Momentum equation}  & Poiseuille flow & \ref{sec:poiseuille}\\
-    & Instantaneous 2D sphere & \ref{sec:ist_sphere} \\
-    & Rayleigh-Taylor experiment & \ref{sec:rayleigh} \\
-    & Falling block & \ref{sec:block} \\
-\hline
-    & 2D Stokes sphere & \ref{sec:sphere} \\
-    Sticky air and  & Stabilisation algorithm & \ref{sec:stab} \\
-    free surface & Topography relaxation & \ref{sec:crameri} \\
-  & Spontaneous subduction & \ref{sec:subduction} \\
-\hline
-  Erosion and sedimentation  & - & - \\
-\hline
-   Non-linear  & Slab detachment & \ref{sec:slab} \\
-   visco-plasticity & Indenter & \ref{sec:indenter} \\
-    & Brick & \ref{sec:brick} \\
-\hline
-  \multirow{3}{*}{Energy equation}  & Advection stabilisation & \ref{sec:advection} \\
-    & Simple shear heating & \ref{sec:simple_shear} \\
-    & Shear and adiabatic heating & \ref{sec:shear} \\
-\hline
-\multirow{3}{*}{Energy + momentum} & Mantle convection & \ref{sec:mantle} \\
-& Viscoplastic mantle convection & \ref{sec:visco_mantle} \\
-& Thin layer entrainment & \ref{sec:thin} \\
-\hline
-  Phase changes  & \multirow{2}{*}{Hydrated sinking cylinder} & \multirow{2}{*}{\ref{sec:quinquis}} \\
-  and hydration  &  & \\
-\hline
-  Melting & Experimental melting curves & \ref{sec:katz} \\
-\bottomrule
-\end{tabular}
-\label{tab:list}
-\end{table}
-
-The code is written in Fortran90 and uses the finite element method (FEM) with quadrilateral $Q_1 \times P_0$ elements (continuous bilinear velocity and
-discontinuous constant pressure), associated with MUMPS\footnote{\url{http://mumps.enseeiht.fr/}} \citep{Amestoy2001,Amestoy2019} that is a software package
-for solving systems of linear equations of the form $A \cdot x = b$, where \textit{A} is a square sparse matrix, by means of a direct method. Since
-$Q_1 \times P_0$ elements do not satisfy Ladyzhenskaya, Babuska and Brezzi (LBB) stability condition \citep{Donea2003}, elemental pressure is elaborated
-in post-processing to avoid spurious pressures by means of a double interpolation to smooth the pressure. Following this procedure, the elemental pressure
-is interpolated onto nodes and then back onto elements. After the smoothing procedure, the pressure field is calculated again on the nodes to be used in
-combination with the lithostatic pressure, which is calculated onto nodes. Correctness of the solution and performances of the solver in terms of time and
-memory usage are tested solving the Stokes flow with the analytical solution proposed by \citet{Donea2003} (Section \ref{sec:stokes}).
-
-The thermo-mechanics of crust-mantle systems is described by means of the conservation of mass, momentum and energy equations, expressed as follows:
-\begin{eqnarray}
-\label{eq:mass}\frac{\partial \rho}{\partial t}+\bm{\nabla} \cdot (\rho \bm{u})&=&0\\
-\label{eq:momentum}-\bm{\nabla} p + \bm{\nabla} \cdot \bm{\tau} + \rho \bm{g}&=&0\\
-\label{eq:energy}\rho C_p \left(\frac{\partial T}{\partial t} + \bm{u} \cdot \bm{\nabla} T \right)&=&\bm{\nabla} \cdot \left(k\bm{\nabla} T\right) + H_{tot}
-\end{eqnarray}
-where $\rho$ is the density, $\bm{u}$ is the velocity, \textit{p} is the pressure, $\bm{\tau}$ is the deviatoric stress, $\bm{g}$ is the gravity acceleration,
-$C_p$ is the specific heat at constant pressure, \textit{T} is the temperature, \textit{k} is the thermal conductivity and $H_{tot}$ is the total internal heat
-production. Density variations due to temperature are generally small enough to assume the density as constant ($\rho=\rho_0$) in Eq. \ref{eq:mass} and in
-Eq. \ref{eq:energy}, while it must be treated as a variable in the buoyancy term of Eq. \ref{eq:momentum}, such that \[\rho=\rho_0(1-\alpha(T-T_0))\]
-where $\rho_0$ is the density at a reference temperature $T_0$ and $\alpha$ is the coefficient of thermal expansion. Eq. \ref{eq:mass} then can be rewritten as
-\begin{equation}\label{eq:mass_0}
-\bm{\nabla} \cdot \bm{u}=0
-\end{equation}
-This simplification is known as the extended Boussinesq approximation.
-
-The time step $dt$ is chosen in according to Courant-Friedrichs-Lewy condition \citep{Anderson1995}, such that
-\[dt=C_n\; \min\left(\frac{\Delta x}{\max|\bm{u}|},\frac{(\Delta x)^2}{\kappa}\right)\]
-where $0<C_n<1$ is the Courant (CFL) number, $\Delta x$ is the minimum dimension of smallest element, $\max|\bm{u}|$ is the maximum velocity calculated on the
-entire domain and
-$\kappa=\frac{k}{\rho C_p}$ is the thermal diffusivity \citep{Thieulot2014}.
-
-\section{Lagrangian markers}\label{sec:markers}
-Elemental properties (density, viscosity, thermal conductivity, specific heat and thermal expansion) needed to solve Eqs. \ref{eq:mass_0}, \ref{eq:momentum}
-and \ref{eq:energy} are related to the composition of each element, determined by means of Lagrangian markers that are characteristic of different materials
-in the domain. Elemental properties (with the sole exception of the viscosity) are calculated using an arithmetic mean, as \[P_e=\frac{1}{n}\sum_{i=1}^n P_i\]
-where $P_e$ is the elemental property, $P_i$ is the property characteristic of the material of each marker and $n$ is the number of markers of the element.
-Differently, the average scheme for the elemental viscosity can be chosen between harmonic, geometric and arithmetic mean. Markers can be placed regularly or
-randomly at the beginning of the simulation and their advection is performed by either a 2nd-order or a 4th-order Runge-Kutta in space, interpolating the velocity 
-field on each marker by means of the shape functions (see Section \ref{sec:runge}). The interpolated velocity is then corrected by means of the Conservative
-Velocity Interpolation (CVI), which introduces a corrective term to reduce dispersion and clustering of particles in both steady state and time-dependent 
-\citep{Wang2015} (see Section \ref{sec:CVI}).
-
-The initial distribution of the markers is created using the open-source code library
-Geodynamic World Builder \citep{Fraters2019}. During the simulation, each marker carries memory of temperature, pressure and accumulated strain, which are
-determined interpolating the relative nodal parameter. In particular, pressure is interpolated onto each marker after the smoothing procedure. The number
-of markers contained in each element is maintained between $n_{min}$ and $n_{max}$. When in an element there are less markers than $n_{min}$ the code adds
-random markers to reach the $n_{min}$, while if the number is higher than $n_{max}$ some of them are randomly deleted. When new markers are added, they assume
-the properties of the nearest marker. In this way elements are never empty and maintain a number of markers inside a prefixed range.
-
-\section{The momentum equation}\label{sec:mom_eq}
-The penalty method is implemented so that Eq. \ref{eq:mass_0} can be written as:
-\begin{equation}\label{eq:mass_penalty}
-\bm{\nabla} \cdot \bm{u} + \frac{p}{\lambda}=0
-\end{equation}
-where $\lambda$ is the so-called penalty parameter, which should be 6-7 orders of magnitude larger than the shear viscosity to ensure that mass conservation
-is satisfied \citep{Donea2003,Thieulot2014}.
-
-The deviatoric stress tensor in Eq. \ref{eq:momentum} can be written in terms of the strain rate tensor as $\bm{\tau}=2\eta\dot{\bm{\epsilon}}$, with
-$\dot{\bm{\epsilon}}=\frac{1}{2}\left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)$. Therefore, Eq. \ref{eq:momentum} can be rewritten as:
-\begin{equation}\label{eq:momentum_pressure}
--\bm{\nabla} p + \bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right) + \rho \bm{g} = 0
-\end{equation}
-Finally, using pressure from Eq. \ref{eq:mass_penalty}, Eq. \ref{eq:momentum_pressure} can be rewritten as:
-\begin{equation}\label{eq:momentum_penalty}
-\lambda \bm{\nabla} \left(\bm{\nabla} \cdot \bm{v}\right)+\bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right)+\rho\bm{g}=0
-\end{equation}
-The penalty method is been associated to the iterative Uzawa method, as described in detail in \citet{Dabrowski2008} and \citet{Thieulot2014}
-(see Section \ref{sec:poiseuille}). Results of benchmarks performed to verify the correctness in the implementation of Eq. \ref{eq:momentum_penalty} are
-shown in Sections \ref{sec:ist_sphere} and \ref{sec:rayleigh}.
-
-To support large viscosities variations the penalty parameter is related to the effective elemental viscosity by means of a dimensionless coefficient, so that
-$\lambda=\lambda_e(e)\eta_{eff}(e)$ \citep{Marotta2006,Dabrowski2008,Thieulot2014}. Benchmark of the falling block \citep{Gerya2003a,Gerya2010b,Thieulot2011}
-is performed to verify that the code can correctly deal buoyancy driven flows with strong viscosity contrasts (Section \ref{sec:block}).
-
-\section{Sticky air and free surface}\label{sec:sticky}
-The Earth's surface can be treated by means of either the so-called sticky-air or a true free surface method, both of them implemented in the code. In the
-sticky-air method the surface is approximated with the introduction of a buoyant layer with a viscosity at least four orders of magnitude lower than the crust
-\citep{Schmeling2008,Crameri2012} and the interface between lithosphere and air is defined using a chain of passive markers that are advected as the Lagrangian
-markers. The correctness of the evolution of the markers chain is tested with the experiment of a 2D time-dependent Stokes sphere below a free surface and
-compared with results from ASPECT\footnote{\url{https://aspect.geodynamics.org/}} \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} (Section
-\ref{sec:sphere}). In the true free surface case the top boundary is assumed stress-free and velocities are not fixed. In this case topography variations are
-described by vertical deformations of the mesh that depend on the velocity field of the nodes that identify the free surface, while horizontal deformations are
-not taken into account. This procedure is known as the Arbitrary Lagrangian-Eulerian (ALE) method and its implementation follows the technique described in
-\citet{Thieulot2011}. However, although the implementation of a true free surface better reproduces laboratory experiment, extremely small time steps can be
-necessary to maintain stability \citep{Kaus2010a,Quinquis2011,Thieulot2014}. Therefore, the stability algorithm proposed by \citet{Kaus2010a} is implemented
-to avoid instabilities due to high density differences at the free surface when using too large time steps. The implementation of this algorithm are tested by
-performing the experiment described by \citet{Kaus2010a} (Section \ref{sec:stab}).
-
-The topography relaxation benchmark proposed by \citet{Crameri2012} is performed to verify that the code correctly recovers topography variations in case of
-both the sticky-air and the true free surface method (Section \ref{sec:crameri}). Finally, Section \ref{sec:subduction} show the results of the spontaneous
-subduction experiment described by \citet{Schmeling2008}.
-
-\section{Erosion and sedimentation}\label{sec:erosion}
-Surface processes at either the lithosphere-air interface or in correspondence of the free surface have been implemented by means of the software
-Fastscape\footnote{\url{https://fastscape.org/}} \citep{Braun2013,Cordonnier2019,Yuan2019a,Yuan2019b}. FastScapeLib is a set of routines that solve the stream
-power law (enriched by a sediment transport/deposition term) \citep{Yuan2019a}, hillslope diffusion and marine transport and deposition \citep{Yuan2019b}.
-The partial differential equation solved by FastScapeLib is:
-\begin{equation}
-\frac{\partial h}{ \partial t}=U-K_f A^{m_e}S^{n_e} + \frac{G}{A}\int_A \left( U-\frac{\partial h}{\partial t} \right) dA + K_d \bm{\nabla}^2 h 
-\end{equation}
-where $h$ is the topography, $U$ is the uplift, $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $S$ is the slope, $m_e$ and $n_e$ are
-parameters that should be chosen to constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}, $G$ is the dimensionless
-deposition/transport coefficient for the enriched stream power law and $K_d$ is the effective diffusivity. Terms $K_f A^{m_e}S^{n_e}$ and $K_d \bm{\nabla}^2 h$
-consider long-range processes, described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, and short-range processes,
-proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}.
-
-In case of sticky air, slope is determined by means of the markers chain, which is vertically corrected considering erosion and sedimentation rates. After the
-correction, continental markers above the markers chain are transformed in air markers and air markers below the markers chain become sediments. Similarly,
-slope is determined considering the nodes on the top boundary in case of true free surface and they are vertically corrected as for the markers chain. In this
-case, continental markers above the top boundary are simply deleted.
-%Surface processes at the lithosphere-air interface are implemented taking into account fluvial erosion (long-range processes), hillslope processes (short-range
-%processes) and failure of slopes exceeding a threshold angle $\theta$ \citep{Burov1997,Tucker1999,Sacek2012,Croissant2014,Ueda2015,Beucher2020}. Erosion rates
-%$\dot{E}_{lr}$ due to long-range processes can be described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{lr}=K_f A^{m_e} \bm{\nabla} h^{n_e}
-%\end{equation}
-%where $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $\bm{\nabla} h$ is the slope and $m_e$ and $n_e$ are parameters chosen in to
-%constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}. Erosion rates $\dot{E}_{sr}$ due to short-range processes are
-%implemented by assuming that transport is linearly proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{sr}=K_d \bm{\nabla}^2 h
-%\end{equation}
-%where $K_d$ is the effective diffusivity. In addition, the slope is modified if it still exceeds a threshold angle after long- and short-range processes
-%\citep{Ueda2015}.
-
-%Offshore sedimentation processes are implemented similarly to hillslope processes but with a depth-dependent coefficient \citep{Sacek2012,Beucher2020}.
-%The sedimentation rate $\dot{S}$ is calculated as
-%\begin{equation}
-%\dot{S}=K_s \bm{\nabla}^2 h
-%\end{equation}
-%with
-%\begin{equation}
-%K_s=K^*_s \exp\left(\frac{h-h_{sl}}{h_0}\right)
-%\end{equation}
-%where $K^*_s$ is a constant, $h_{sl}$ is the sea level and $h_0$ is a characteristic length. In case of sticky air, slope is determined by means of the markers
-%chain, which is vertically corrected considering erosion and sedimentation rates. After the correction, continental markers above the markers chain are
-%transformed in air markers and air markers below the markers chain become sediments. Similarly, slope is determined considering the nodes on the top boundary
-%in case of true free surface and they are vertically corrected as for the markers chain. In this case, continental markers above the top boundary are simply
-%deleted.
-
-\section{Non-linear rheologies}\label{sec:nonlinear}
-Non-linear rheologies are implemented combining viscous creep (dislocation and diffusion) and plastic yielding. For each marker, diffusion and dislocation
-viscosity ($\eta_{df|ds}$) can be determined as follows \citep{Karato1993,Warren2008b,Wang2016}:
-\begin{equation}\label{eq:viscous}
-\eta_{df|ds}=f_s \left(\frac{d^m}{A}\right)^{\frac{1}{n}} \left (\frac{2^{\frac{1-n}{n}}}{3^{\frac{n+1}{2n}}} \right)
-I_2^{\frac{(1-n)}{n}}\exp\left(\frac{Q+pV}{nRT}\right)
-\end{equation}
-where $f_s$ is a scaling factor used to represent lithologies that are stronger or weaker than the base set, $d$ is the grain size, $m$ is the grain size exponent,
-$n$ is the stress exponent, $A$ is the uniaxial pre-exponential factor, $I_2$ is the square root of the second invariant of the strain rate tensor, $Q$ is the
-activation energy, $p$ is the pressure, $V$ is the activation volume, $R$ is the gas constant and $T$ is the temperature.
-Pressures and temperatures are determined by the interpolation of the nodal parameters, while strain rates are calculated by means of the
-derivative of the velocities on the nodes. In case of diffusion creep $n=1$ and $m>0$, in case of dislocation creep $n>1$ and $m=0$. The viscous creep $\eta_{cp}$
-is then calculated as the harmonic average between $\eta_{df}$ and $\eta_{ds}$:
-\begin{equation}\label{eq:visco_creep}
-\eta_{cp}=\left(\frac{1}{\eta_{df}}+\frac{1}{\eta_{ds}}\right)^{-1}
-\end{equation}
-The implementation of the non-linear viscous creep viscosity depending on the strain rate is tested by means of the slab detachment benchmark
-(Section \ref{sec:slab}).
-
-Plastic yielding is implemented rescaling $\eta_{cp}$ in order to limit the stress below the yield stress $\sigma_y$ \citep{Thieulot2008,Thieulot2014,Glerum2018},
-obtaining
-\begin{equation}\label{eq:plastic}
-\eta_{pl}=\frac{\sigma_y}{2I_2}
-\end{equation}
-where $\eta_{pl}$ is the plastic viscosity and the yield stress is determined following the Drucker-Prager criterion, such as
-\begin{equation}\label{eq:yield}
-\sigma_y=C\;\cos(\phi)+p\;\sin(\phi)
-\end{equation}
-where $C$ is the cohesion and $\phi$ is the internal friction angle. In case of negative pressure it is imposed equal to 0, so that negative yield stress are
-excluded. The correctness of the non-linear solution in case of plastic viscosity with variable internal friction angle is verified performing indenter and
-brick experiments (Sections \ref{sec:indenter} and \ref{sec:brick}, respectively).
-
-Strain softening is taken into account for both viscous creep and plastic viscosity \citep{Huismans2003,Babeyko2005,Huismans2005,Sobolev2005,Warren2008a} by
-means of the accumulated strain $\epsilon$ memorised by each marker. Softening in the viscous creep determines a linear decrease of $\eta_{vc}$ by means of a
-viscous strain softening factor $W_S$ that increases linearly from $W_{S_0}$ to $W_{S_{\infty}}$ for $\epsilon_{S_0}<\epsilon_S<\epsilon_{S_{\infty}}$ 
-\citep{Warren2008a}. This viscous softening can be related to strain-induced grain size reduction \citep{Warren2008a}. Differently, plastic softening is 
-simulated with a linear decrease of internal friction angle $\phi(\epsilon)$ and cohesion $C(\epsilon)$ in according to
-\begin{eqnarray}
-\label{eq:friction1}\phi(\epsilon)&=&\phi_0+(\phi_{\infty}-\phi_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}\\
-\label{eq:friction2}C(\epsilon)&=&C_0+(C_{\infty}-C_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}
-\end{eqnarray}
-where $\phi_0$, $C_0$ and $\phi_{\infty}$, $C_{\infty}$ are internal friction angle and cohesion for $\epsilon_0$ and $\epsilon_{\infty}$, respectively
-\citep{Huismans2003,Huismans2005,Warren2008a,Thieulot2014}. Plastic softening approximates deformation-induced softening of faults and brittle shear zones
-\citep{Warren2008a}.
-
-Plastic yielding and viscous creep are then combined to obtain a viscoplastic viscosity $\eta_{vp}$ as follows, assuming that they are independent processes
-\citep{Karato2008,Glerum2018}:
-\begin{equation}\label{eq:viscoplastic}
-\eta_{vp}=\min(\eta_{cp},\eta_{pl})
-\end{equation}
-Finally, effective viscosity $\eta_{eff}$ is capped by the minimum and the maximum viscosity ($\eta_{min}$ and $\eta_{max}$, respectively) to avoid extremely
-low or high viscosity \citep{Glerum2018} as follows
-\begin{equation}\label{eq:effective}
-\eta_{eff}=\min(\max(\eta_{vp},\eta_{min}),\eta_{max})
-\end{equation}
-Viscous creep and plastic yielding are non-linear rheologies because of their dependence on the velocity field through pressure and strain rates. Therefore,
-the solution is determined by means of Picard-type iterations, until convergence of the velocity field \citep{Glerum2018}. The convergence is verified at each
-iteration $i$ via the nonlinear residual $\bm{\mathcal{R}}^i$ that can be determined as
-\begin{equation}\label{eq:convergence}
-\bm{\mathcal{R}}^i= \mathbb{K} (\eta_{eff}(\dot{\bm{\epsilon}}^{i-1},p^{i-1}))\cdot \bm{v}^{i-1} - \bm{f}^i
-\end{equation}
-%\begin{equation}\label{eq:convergence}
-%\chi=1-\frac{\langle (\bm{v}^i-\langle\bm{v}^i\rangle)\cdot (\bm{v}^{i+1}-\langle\bm{v}^{i+1}\rangle) \rangle}{\sqrt{\sigma^i\sigma^{i+1}}}
-%\end{equation}
-%Iterations are performed until either convergence ($\chi<tol$) or maximal number of non-linear iterations ($it_{max}$) is reached.
-where $\mathbb{K}$ is the Stiffness matrix and $\bm{f}$ is the right hand side vector \citep{Spiegelman2016,Glerum2018}.  The $L_2$-norm is extracted from
-$\bm{\mathcal{R}}^i$ and it is normalised as follows
-\begin{equation}\label{eq:convergence_res}
-\frac{||\bm{\mathcal{R}}^i||_2}{||\bm{\mathcal{R}}^0||_2}
-\end{equation}
-where $||\bm{\mathcal{R}}^0||_2$ is the $L_2$-norm of the first nonlinear iteration.
-Since often the normalised nonlinear residual drop very quickly over the first 2-3 iterations, also the $L_2$-norm of velocity and pressure residuals are
-calculated as
-\begin{equation}\label{eq:convergence_vel}
-\frac{||\Delta\bm{v}||_2}{||\bm{v}||_2}=\frac{||\bm{v}_i-\bm{v}_{i-1}||_2}{||\bm{v}_{i}||_2}
-\end{equation}
-and
-\begin{equation}\label{eq:convergence_press}
-  \frac{||\Delta p||_2}{||p||_2}=\frac{||p_i-p_{i-1}||_2}{||p_{i}||_2}
-\end{equation}
-respectively. Iterations are performed until either maximal number of non-linear iterations ($it_{max}$) is reached or all the normalised nonlinear residual, 
-the velocity residual and the pressure residual converge under a defined tolerance $tol$.
-
-\section{The energy equation}\label{sec:en_eq}
-The stabilisation of the advection term of the energy equation (Eq. \ref{eq:energy}) needed to avoid possible oscillations in the thermal solution in the case
-when advection dominates over diffusion, is implemented by means of a streamline-upwind Petrov–Galerkin (SUPG) method, for which the advection term is modified
-follows the discussion in \citet{Thieulot2011} and \citet{Thieulot2014} (see Section \ref{sec:advection} for benchmark). The total internal heat production
-$H_{tot}$ in Eq. \ref{eq:energy} includes radiogenic heating $H_r$, shear heating $H_s$ and adiabatic heating $H_a$, where
-\begin{equation}\label{eq:shear}
-H_s=2\eta \dot{\bm{\epsilon}} : \dot{\bm{\epsilon}} = 2\eta(\dot{\epsilon}_{xx}^2+\dot{\epsilon}_{yy}^2+2\dot{\epsilon}_{xy}^2)
-\end{equation}
-and
-\begin{equation}\label{eq:adiabatic}
-H_a=T \alpha \frac{Dp}{Dt} \approx -\alpha T \rho g_y v
-\end{equation}
-The correctness of the implementation and the computation of shear and adiabatic heating is verified performing an exercise with an analytical solution
-(Section \ref{sec:simple_shear}) and Exercise 9.4 in \citet{Gerya2010b} (Section \ref{sec:shear}). Furthermore, mantle convection and thin layer entrainment
-benchmarks (Sections \ref{sec:mantle} and \ref{sec:thin}, respectively) are performed to verify that the code solve correctly Eqs. \ref{eq:momentum_penalty}
-and \ref{eq:energy} in case of a temperature field characterised by an initial perturbation and with a temperature-dependent density. At last, the viscoplastic
-mantle convection benchmark is performed to check also the viscoplasticity in case of a initial perturbed temperature field (Section \ref{sec:visco_mantle}).
-
-\section{Phase transitions and hydration}\label{sec:phase}
-Variations of effective density, specific heat and coefficient of thermal expansion during the evolution of crustal and mantle materials at different
-pressure-temperature ($p$-$T$) conditions are computed by means of version 6.8.6 of Perple\_X software package \citep{Connolly2005}, similarly to the
-implementation described in \citet{Marotta2020}. By default, mineral assemblages and properties of each lithology (oceanic crust, lower and upper continental
-crust, sediments and dry or hydrated mantle) are calculated for temperatures from \SIrange{330}{1600}{\kelvin}, with increments of approximately \SI{4}{\kelvin},
-and pressures from \SIrange{0.1}{30}{\giga\pascal}, with increments of approximately \SI{0.03}{\giga\pascal}, for a total of almost 100000 points.
-
-The introduction of phase transitions produces two main effects that must be taken into account in numerical models: variations in density and release/absorption
-of latent heat \citep{Tackley2010}. Effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into
-account considering effective coefficient of thermal expansion \citep{Christensen1985,Zhong2015}. Similarly, effects of latent heat in the energy equation
-(eq \ref{eq:energy}) are taken into account by considering effective specific heat and coefficient of thermal expansion
-\citep{Christensen1985,Tackley2010,Zhong2015}. Effective density, specific heat and coefficient of thermal expansion are included in Perple\_X files and are
-assumed by each Lagrangian marker in according to its $p$-$T$ conditions.
-
-Hydration processes related to dehydration of subducting lithosphere \citep{Schmidt1998,Liu2007,Faccenda2009,Faccenda2010,Faccenda2014,Rosas2016} strongly
-influence the thermo-mechanics inside the mantle wedge, mainly because of both weakening effects on the mantle rheology and density variations in case of
-mantle serpentinization \citep{Gerya2002,Honda2003,Arcay2005,Roda2010,Regorda2017}. In the case that hydration is switched on, the amount of bound and free
-water is memorised by each marker, following the implementation of \citet{Quinquis2014}; hydration and dehydration processes are related to the amount of bound
-water of each marker and to the maximum amount of water it can transport, i.e. if the amount of bound water exceeds the maximum amount of water, the marker
-dehydrates and releases free water that can hydrates under-saturated markers. Maximum water content of each marker is determined as function of lithology and
-$p$-$T$ conditions and it is calculated using Perple\_X, in the same way than effective density, specific heat and coefficient of thermal expansion. Bound water
-is advected together with the markers, neglecting the effect of bound water diffusion, while free water simply migrates vertically and is not coupled to the
-solid-phase flow of the mantle wedge \citep{Arcay2005,Quinquis2014}. The correctness of the migration scheme of the free water and its absorption into bound
-water by under-saturated markers are verified performing the experiment described in \citet{Quinquis2014} (Sec. \ref{sec:quinquis}).
-
-The mantle viscosity weakening related to the amount of water has been extensively studied \citep{Chopra1981,Hirth2003} and it is implemented in according to
-\citet{Arcay2005} and \citet{Horiuchi2016}, as follows
-\begin{equation}\label{eq:wet_visc}
-\eta_{wet}=\eta_{dry}\left(\left[1-\frac{1}{f_v}\right]\exp\left(-\frac{[OH^-]}{[OH^-]_0}\right)+\frac{1}{f_v}\right)
-\end{equation}
-where $f_v$ is the viscosity reduction factor between dry and wet conditions, $[OH^-]$ is the water content and $[OH^-]_0$ is a reference water content set to
-620 ppm (0.062 wt.\%) \citep{Arcay2005}. By default, $f_v$ is set to 100, in according to \citet{Arcay2005} and \citet{Horiuchi2016}.
-
 \section{Mantle melting}\label{sec:melt}
@@
 molten rock \citep{Gerya2004b,Gerya2007,Tackley2010}.
-
-\section{Benchmarks}\label{sec:benchmark}
-\subsection{Error measurements}\label{sec:error}
-In order to determine the accuracy of velocity and pressure field of the benchmarks, the $L_2$-norm is computed by numerical integration on the Gauss-Legendre
-quadrature points. $L_2$-norm for pressure and velocity errors can be evaluated as
-\begin{eqnarray}
-\textrm{err}_p&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|p^n_i(r_q)-p^a_i(r_q)|^2\right)w_q|J_q|}\nonumber \\
-\textrm{err}_v&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|u^n_i(r_q)-u^a_i(r_q)|^2+|v^n_i(r_q)-v^a_i(r_q)|^2\right)w_q|J_q|}\nonumber
-\end{eqnarray}
-respectively, where $n_e$ is the number of elements, $n_q$ is the number of quadrature points per element, $p^n_i(r_q)$ and $p^a_i(r_q)$ are the numerical and
-analytical pressure, respectively, in each quadrature point $q$, $w_q$  and $J_q$ are the weight and the Jacobian at the quadrature point
-$q$, $u^n_i(r_q)$, $v^n_i(r_q)$, $u^a_i(r_q)$ and $v^a_i(r_q)$ are the numerical and analytical velocities, respectively, in each quadrature point $q$.
-
-Other quantities used as comparison with original benchmarks are the root-mean-square velocity over the whole domain and over the surface
-\begin{eqnarray}
-v_{\textrm{rms}}&=&\sqrt{\int_{\Omega}|\bm{v}|^2 d\Omega}\nonumber \\
-v_{\textrm{rms}}^{top}&=&\sqrt{\int_0^1 u^2 \bigg|_{y=1} dx}\nonumber
-\end{eqnarray}
-respectively, top and bottom Nusselt numbers
-\[\textrm{Nu}^{top/bottom}=-\int_0^1\frac{\partial T}{\partial y} \bigg|_{y=1/y=0} dx\]
-the average rate of work done against gravity
-\[\left\langle W \right\rangle=\int_{\Omega}T u_y d\Omega\]
-the average rate of viscous dissipation
-\[\left\langle \Phi \right\rangle=\int_{\Omega}\tau_{ij}\dot{\epsilon}_{ij} d\Omega\]
-and the percentage error between $\left\langle W \right\rangle$ and $\left\langle \Phi \right\rangle/Ra$
-\[\delta=\frac{\left|\left\langle W \right\rangle -\frac{\left\langle \Phi \right\rangle}{Ra} \right|}{\max \left(\left\langle W \right\rangle,
-\frac{\left\langle \Phi \right\rangle}{Ra}\right)}\times100\]
-
-\subsection{Stokes flow}\label{sec:stokes}
-The problem consists of determining velocity field $(u,v)$ and pressure $p$ in case of a manufactured solution with prescribed body forces such as
-\begin{eqnarray}
-b_1&=&(12-24y)x^4+(-24+48y)x^3+(-48y+72y^2-48y^3+12)x^2+\nonumber \\
-&&+(-2+24y-72y^2+48y^3)x+1-4y+12y^2-8y^3\nonumber \\
-b_2&=&(8-48y+48y^2)x^3+(-12+72y-72y^2)x^2+(4-24y+48y^2-48y^3+\nonumber \\
-&&+24y^4)x-12y^2+24y^3-12y^4\nonumber
-\end{eqnarray}
-for which the exact solution is:
-\begin{eqnarray}
-u(x,y)&=&x^2(1-x)^2(2y-6y^2+4y^3)\nonumber \\
-v(x,y)&=&-y^2(1-y)^2(2x-6x^2+4x^3)\nonumber \\
-p(x,y)&=&x(1-x)-\frac{1}{6}\nonumber
-\end{eqnarray}
-The domain is a unit square a constant viscosity ($\eta=1$) and the penalty parameter is set to $\lambda=10^7$. Velocity boundary conditions are set to no slip
-$(\bm{v}=\bm{0})$ on all boundaries. The problem is performed for different grid resolution between \num{8x8} and \num{1024x1024} elements.
-
-The errors between the analytical solution and the numerical prediction of the velocity and the pressure field are calculated as in Section \ref{sec:error}.
-Fig. \ref{fig:errors}a shows that both velocity and pressure field converge to the exact solution with the decrease of the element size, following the
-theoretical convergence rate. The convergence of pressure, in contrast with the observation by \citet{Donea2003} for $Q_1 \times P_0$ elements, indicate the
-effectiveness of the smoothing procedure. Smoothed pressure field for grid resolution of \num{128x128} elements is shown in Fig. \ref{fig:errors}b (continuous
-red line), in comparison with the analytical solution (dashed blue line). Solve times and memory usage needed to generate the solution are shown in
-Fig. \ref{fig:MUMPS}. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Solver/Stokes_Flow}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/errors.pdf}
-\caption{Velocity and pressure error for the Stokes flow experiment between generated and analytical solution as a function of element size (panel a),
-and comparison between smoothed pressure and analytical solution as function of $x$ coordinate for a grid resolution of \num{128x128} elements (panel b).}
-\label{fig:errors}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/MUMPS.pdf}
-\caption{Panel a) Analysis, factorisation, solution and total solve times (panel a) and factorisation memory usage (panel b) as a function of the total number
-of degrees of freedom for the Stokes flow experiment.}
-\label{fig:MUMPS}
-\end{figure}
-
-\subsection{Markers advection}\label{sec:runge}
-The 2nd-order Runge-Kutta advection scheme in space is tested by means of the Zalesak disk test \citep{Zalesak1979,Thieulot2014}. The benchmark is performed
-in a unit square domain with a grid resolution of \num{32x32} elements and 1000000 markers and values of Courant number between 0.25 and 1.
-At $t=0$, the disk is centred at position $(0.5;0.75)$ with a radius $R=0.15$ and has a vertical fissure 0.05 wide and 0.2 high (orange in Fig. \ref{fig:runge})
-and with the velocity field prescribed in the entire domain as
-\begin{eqnarray}
-u(x,y)&=&2\pi\left(y-\frac{L_x}{2}\right)\nonumber \\
-v(x,y)&=&-2\pi\left(x-\frac{L_x}{2}\right)\nonumber
-\end{eqnarray}
-the markers are back to their initial location after a $2\pi$ rotation ($t=1$). The clockwise rotation of the disk at $t=0.25$, 0.5 and 0.75 
-is shown in Fig. \ref{fig:runge} (red, blue and green, respectively). The distance from the centre of 1 marker is calculated throughout the rotation to
-evaluate the error for different values of Courant number. As expected, the error increase with the increasing of the time step, as shown in 
-Fig. \ref{fig:runge_err}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/Zalesak_Disk}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=200px]{./Figures/Zalesak.pdf}
-\caption{Position of the Zalesak disk at $t=0$, 0.25, 0.5 and 0.75 (orange, red, blue and green, respectively) for the case with Courant number of 1.}
-\label{fig:runge}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/zal_marker.pdf}
-\caption{Distance from the centre as function of time for values of Courant number of  of 0.25, 0.3, 0.5, 0.75 and 1 (green, orange, blue, black and red,
-respectively). R indicates the distance from the centre considering a perfect circle trajectory.}
-\label{fig:runge_err}
-\end{figure}
-
-\subsection{Conservative Velocity Interpolation (CVI)}\label{sec:CVI}
-The CVI correction is checked by means of the Stokes flow experiment by \citet{Donea2003}, as presented in \ref{sec:stokes}. The advection of Lagrangian
-markers is performed using either a 2nd-order or a 4th-order Runge-Kutta scheme with CFL=0.5 and an initial random distribution of 25 markers per element.
-Fig. \ref{fig:CVI} shows the minimum and maximum values of markers per element throughout the experiment with and without the CVI correction and using
-different Runge-Kutta schemes. The introduction of the CVI correction allows the
-tests to run for times longer than $t=2000$ without show empty elements, while tests without the CVI correction stop at approximately $t=600$, when at least
-one element goes empty. Fig. \ref{fig:CVI_time} shows the different number of markers per element and the markers distribution at $t=600$ obtained without and
-with the CVI correction (panels b and c, respectively) in case of a 2nd-order Runge-Kutta scheme. All data can be found at
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/CVI}.
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/CVI.pdf}
-\caption{Maximum and minimum number of markers per element throughout the experiment, with (green and orange lines) or without (purple and blue lines)
-the CVI correction and using either a 2nd-order (blue and orange lines) or a 4th-order (purple and green lines) Runge-Kutta scheme.}
-\label{fig:CVI}
-\end{figure}
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/CVI_time.pdf}
-\caption{Number of markers per element and markers distribution through time using 2nd.order Runge-Kutta scheme.
-Panels a: number of markers per element (panel a1) and distribution of the markers (panel a2) after the first
-iteration; panels b and c: comparison between number of markers per element (panels b1 and c1) and markers distribution (panels b2 and c2) without and with
-the CVI correction at $t=600$; panels d: number of markers per element (panel d1) and markers distribution (panel d2) in case of CVI correction at $t=2000$.}
-\label{fig:CVI_time}
-\end{figure}
-
-\subsection{Poiseuille flow}\label{sec:poiseuille}
-The domain is a rectangle with $L_x=2$ and $L_y=1$ and constant density and viscosity ($\rho=1$ and $\eta=1$), gravity acceleration $\bm{g}=\bm{0}$ and penalty
-parameter $\lambda=10^8$. The grid is composed by \num{40x20} elements. Velocity boundary conditions are set to no slip $(\bm{v}=\bm{0})$ at the top and the
-bottom, and a parabolic profile is imposed on the sides, with $u=y(1-y)$ and $v=0$. The analytical solution is then given by:
-\begin{eqnarray}
-u(x,y)&=&y(1-y)\nonumber \\
-v(x,y,)&=&0\nonumber \\
-p(x,y)&=&2\eta\left(\frac{L_x}{2}-x\right)\nonumber
-\end{eqnarray}
-
-The velocity field predicted by the model follows the expected parabolic profile (Fig. \ref{fig:poiseuille}a and Fig. \ref{fig:poi_plot}a) and, in case of the
-classic penalty method with no iterations, the pressure is clearly related to the divergence by means of the penalty parameter (Fig. \ref{fig:poiseuille}b and
-c and Fig. \ref{fig:poi_plot}b). Fig. \ref{fig:poiseuille}d shows that one Uzawa iteration is sufficient to bring the divergence down to \num{e-15}, with no
-correlation with the pressure field. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Poiseuille%20Flow}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Poiseuille.pdf}
-\caption{Velocity field (panel a), pressure (panel b) and divergence velocity of a Poiseuille flow in case of the classic penalty method (no iterations) and
-after one Uzawa iteration (panel c and d, respectively).}
-\label{fig:poiseuille}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Analytical.pdf}
-\caption{Velocity (panel a) and pressure (panel b) field predicted by the model for a Poiseuille flow with respect with their analytical solutions. The
-velocity field is plotted as function of the vertical coordinate in $L_x/2$ and the pressure is plotted as function of the x coordinate in $L_y/2$.}
-\label{fig:poi_plot}
-\end{figure}
-
-\subsection{Instantaneous 2D sphere}\label{sec:ist_sphere}
-The domain is a unit square with gravity $\bm{g}=(0,-1)$. The fluid has constant density and viscosity ($\rho_f=1$ and $\eta_f=1$). The sphere is in the middle
-of the domain with a radius R = 0.123456798 and has constant density and viscosity ($\rho_s=10^{-2}+\rho_f$ and $\eta_s=10^3 \cdot \eta_f$).
-We distinguish three different types of velocity boundary conditions:
-\begin{enumerate}
-\item FS: free slip conditions on all sides.
-\item NS: no slip conditions on all sides.
-\item OT: free slip conditions on the sides and the bottom, and open boundary at the top.
-\end{enumerate}
-All three cases are performed for different grid resolutions, between \num{16x16} and \num{512x512} elements with 50 randomly distributed markers per element,
-and different average schemes for the viscosity (harmonic, geometric and arithmetic). The velocity in the centre of the sphere $\bm{v}(0.5;0.5)$), minimum and
-maximum velocities ($u_{min}$, $u_{max}$, $v_{min}$, $v_{max}$) and pressures ($p_{min}$, $p_{max}$) on the entire domain, $v_{\textrm{rms}}$ and average pressure on the
-entire domain ($p_{avg}$) are compared with solutions generated by ASPECT \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and all results
-can be found at \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere2D} and 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Instantaneous_2D_Sphere}. Results in terms of the $v_{\textrm{rms}}$ for NS model 
-are also shown in Fig. \ref{fig:inst_sphere}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_NS.pdf}
-\caption{$v_{\textrm{rms}}$ as function of element size for the instantaneous 2D sphere experiment in case of no slip boundary conditions and with different average
-schemes for the viscosity. Results are compared with results obtained by other numerical codes.}
-\label{fig:inst_sphere}
-\end{figure}
-
-\subsection{Rayleigh-Taylor instability}\label{sec:rayleigh}
-This problem has been originally presented by \citet{vanKeken1997} and here it is performed as the isoviscous case (Case 1a). The domain has $L_x=0.9142$ and
-$L_y=1$ and gravity $\bm{g}=(0,-1)$. Two fluids with same constant viscosities ($\eta_1=\eta_2=1$) and different densities ($\rho_1=1000$ and $\rho_2=1010$),
-with the lighter fluid at the bottom. The initial interface between the fluids is given by $y(x)=0.2+0.02 \cos \left(\frac{\pi x}{L_x}\right)$. The experiment
-is performed with different grid sizes (\num{50x50}, \num{80x80}, \num{100x100} and \num{256x256}). A total of 1960000 markers are randomly distributed at the
-beginning of the simulation. Velocity boundary conditions are set to no slip at the top and the bottom, and to free slip at the sides of the domain.
-
-$v_{\textrm{rms}}$ as function of time is reported in Fig. \ref{fig:RT}, matching well with results shown by \citet{vanKeken1997}, \citet{Tackley2003} and
-\citet{Thieulot2014}. Fig. \ref{fig:rayleigh} shows the evolution of the experiment at different time steps. All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Rayleigh_Taylor_experiment/ISOVISCOUS}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/RT.pdf}
-\caption{$v_{\textrm{rms}}$ of the Rayleigh-Taylor experiment as function of time for different resolution of the grid.}
-\label{fig:RT}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=300px]{./Figures/Rayleigh.pdf}
-\caption{Evolution of the Rayleigh-Taylor experiment for a grid of \num{256x256} elements at $t=0, 500, 1000$ and $2000$ (panels a, b, c and d, respectively).}
-\label{fig:rayleigh}
-\end{figure}
-
-\subsection{Falling blocks}\label{sec:block}
-This benchmark is described as proposed by \citet{Gerya2003a}, \citet{Gerya2010b} and \citet{Thieulot2011}. The domain is square with $L_x=L_y=$ \SI{500}{\km}
-and the grid is composed by \num{50x50} elements with 25 markers in each element. The block is initially centred at ($x=$ \SI{250}{\km}; $y=$ \SI{400}{\km})
-and has a size of \num{100x100} km (Fig. \ref{fig:block}a). The fluid surrounding the block has $\eta_f=$ \SI{e21}{\pascal\per\s} and $\rho_f=$
-\SI{3200}{\kg\per\cubic\m}. The benchmark tests with different viscosities of the block, with $\eta_b$ from \SIrange{e15}{e27}{\pascal\per\s}. In each
-experiment also the density $\rho_b$ of the block varies from \SIrange{3220}{9900}{\kg\per\cubic\m}. Velocity boundary conditions are set to free slip
-conditions on all sides of the domain. Density distributions at $t=$ \SI{20}{\mega\year} for $\rho_b=$ \SI{3300}{\kg\per\cubic\m} and different $\eta_b$
-are plotted in Fig. \ref{fig:block}b-f, showing that the code correctly preserve the block geometry in case of large viscosity contrast, with a block stiffer
-than the surrounding fluid (Fig. \ref{fig:block}f).
-
-The velocity in the centre of the falling block at $t=0$ is measured for all experiments and, since the velocity should increase with density contrast, the
-quantity $\bm{v}/(\rho_b-\rho_f)$ is plotted as function of the viscosity contrast. All results perfectly match with those from \citet{Gerya2010b} and they
-line up on a single curve, demonstrating that the code can correctly deal with large viscosity and density contrasts (Fig. \ref{fig:falling}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Falling%20blocks}.
-
-\begin{figure}
-\noindent\includegraphics[width=300px]{./Figures/Block.pdf}
-\centering
-\caption{Density evolution of the falling block experiment at $t=$ \SI{0}{\mega\year} (panel a) and $t=$ \SI{20}{\mega\year} for different viscosities of the
-block (panels b-f).}
-\label{fig:block}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Falling.pdf}
-\caption{Initial velocity relative to the density contrast at the centre of the falling block as function of the viscosity contrast between the block and the
-surrounding fluid.}
-\label{fig:falling}
-\end{figure}
-
-\subsection{2D time-dependent Stokes sphere with deformable free surface}\label{sec:sphere}
-This experiment is performed in a unit square domain with the gravity acceleration fixed to $g_y=-1$. The sphere has $\rho_s=2$ and $\eta_s=10^3$ and is
-initially centred at ($0.5;0.6$) with radius $R=0.123456789$. The fluid surrounding the sphere has $\rho_f=1$ and $\eta_f=1$ and occupies the domain for
-$y\leq 0.75$, while for $y>0.75$ the air has $\rho_a=0$ and $\eta_a=10^{-3}$. Velocity boundary conditions are set to free slip on all sides. The Courant
-number is set to 0.25. The experiments run for 200 s using grid resolutions from \num{150x150} to \num{512x512} elements, with 25 randomly distributed markers
-per element, and for different average schemes for the viscosity. The interface between the fluid and the air is tracked by means of the markers chain. Results
-in terms of velocity and pressure fields and topography variations are compared with results obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and they can be found at
-\url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere_fs2D} and
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Time_dependent_sphere}. Results in terms of the $v_{\textrm{rms}}$ in case of 
-an arithmetic average are also shown in Fig. \ref{fig:inst_spherefs}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_arithm.pdf}
-\caption{$v_{\textrm{rms}}$ for different grid resolutions as function of time for the 2D time-dependent Stokes sphere experiment in case of an arithmetic average.
-Results are compared with results obtained by ASPECT.}
-\label{fig:inst_spherefs}
-\end{figure}
-
-\subsection{Free surface stabilisation}\label{sec:stab}
-The benchmark is performed as discussed in \citet{Kaus2010a} and \citet{Thieulot2014}. The domain is square with $L_x=L_y=$ \SI{500}{\km} and a grid resolution
-of \num{200x200} elements, each of them containing 16 markers. A buoyant fluid with $\rho_1=$ \SI{3200}{\kg\per\cubic\metre} and $\eta_1=$ \SI{e20}{\pascal\s}
-is overlain by a denser fluid with $\rho_2=$ \SI{3300}{\kg\per\cubic\metre} and $\eta_2=$ \SI{e21}{\pascal\s}. The initial interface between the fluids has an
-initial sinusoidal shape of 5 km amplitude. Velocity boundary conditions are set to free slip conditions at the sides and no slip at the bottom, while the top
-is free surface. The experiment is performed with various fixed time steps, without the use of the Courant number. The vertical position of the free surface at
-$x=L_x$ is tracked for each simulation, with or without the stabilisation algorithm.
-
-The results show an instability (drunken sailor effect) increasing the time step in case that the stabilisation algorithm is not activated
-($dt=$ \SI{4200}{\year}, green line in Fig. \ref{fig:stab}), as already observed by \citet{Kaus2010a} and \citet{Thieulot2014}. Activating the algorithm the
-instability is fixed and the simulation remains stable using time step up to \SI{50000}{\year} (red line in Fig. \ref{fig:stab}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Stabilisation_algorithm}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Kaus_random.pdf}
-\caption{Evolution of the y coordinate in $x=L_x$ as function of time for different time steps, with and without the stabilisation algorithm.}
-\label{fig:stab}
-\end{figure}
-
-\subsection{Topography relaxation}\label{sec:crameri}
-This experiment is performed as in \citet{Crameri2012} using free surface and sticky air (Case 1). The domain is rectangular with $L_x=$ \SI{2800}{\km} and
-$L_y=$ \SI{700}{\km} in case of free surface and $L_y=$ \SI{800}{\km} in case of sticky air ($h_{st}=$ \SI{100}{\km}). The grid is composed by \num{256x64}
-elements for the free surface \num{560x320} elements for the sticky air. For both cases, there is a mantle of 600 km thickness with $\eta_m=$ \SI{e21}{\pascal\s}
-overlain by a highly viscous cosine-shaped lithosphere with thickness between 93 and 107 km and $\eta_l=$ \SI{e23}{\pascal\s}. Both mediums have $\rho=$
-\SI{3300}{\kg\per\cubic\metre}. In case of sticky air experiment, the air has $\rho_a=$ \SI{0}{\kg\per\cubic\metre} and $\eta_a=$ \SI{e18}{\pascal\s}. The
-gravity acceleration is set to $g_y=$ \SI{-10}{\metre\per\square\s}. Velocity boundary conditions are set to no slip at the bottom and free slip at the sides
-of the domain. In case of sticky air, velocities on top are set to free slip conditions.
-
-The maximum topography as function of time ($h(t)$) can be analytically derived in according to \[h(t)=h_0 \exp(\gamma t)\]
-where $h_0=7$ km is the initial topography, $\gamma=\num{-0.2139e-11}$ is the characteristic relaxation rate and $t$ is the time (black line in Fig.
-\ref{fig:crameri}). Maximum topography at the characteristic relaxation time $t_{rlx}=$ \SI{14.825}{\kilo\year} can be found to be $h_{rlx}=$ \SI{2576}{\m}.
-The results show that a viscosity of the air of \SI{e18}{\pascal\s} is needed in case of $h_{st}=$ \SI{100}{\km} to correctly describe the topography relaxation
-for this problem, as pointed out by \citet{Crameri2012} (purple line in Fig. \ref{fig:crameri}). The case with a true free surface well follows the analytical
-solution (orange line in Fig. \ref{fig:crameri}). In particular, at the relaxation time in case of the true free surface predicts a maximum topography of
-\SI{2570}{\m}, with an error of \SI{6}{\m} with respect to the analytical solution. Results are also compared with results obtained with a free surface in
-UNDERWORLD (\num{256x64} elements) and SULEC (\num{401x201} elements) (with courtesy from \citealp{Crameri2012}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Topography_relaxation}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Crameri.pdf}
-\caption{Maximum topography as function of time for the Crameri benchmark (coloured lines) in comparison with the analytical solution (black line) and results
-obtained with UNDERWORLD (yellow line) and SULEC (blue line).}
-\label{fig:crameri}
-\end{figure}
-
-\subsection{Spontaneous subduction}\label{sec:subduction}
-This experiment is performed as presented by \citet{Schmeling2008}, in case of both a sticky air and a true free surface and using different average schemes
-for the viscosities. The domain is rectangular with $L_x=$ \SI{3000}{\km} and $L_y=$ \SI{700}{\km}, a grid resolution of \num{256x64} elements with 50 markers
-per element and Courant number of 0.01. At the beginning of the simulation, a 100 km-thick lithospheric layer with $\rho_m=$ \SI{3300}{\kg\per\cubic\metre}
-and $\eta_m=$ \SI{e23}{\pascal\s} is located at the top of the domain for $x$ between 1000 and 3000 km, while the asthenospheric mantle has $\rho_m=$
-\SI{3200}{\pascal\s} and $\eta_m=$ \SI{e21}{\pascal\s}. In addition, a 200 km-depth lithospheric slab is already subducted in the mantle in order to have
-a spontaneous subduction. In case of sticky air, $L_y=750$ km and a 50 km-thick air layer with $\eta_a$ from \SIrange{e19}{e21}{\pascal\s} overlie the mantle.
-Velocity boundary conditions are set to free slip on the sides and on the bottom of the domain. In case of sticky air, velocities on the top boundary are set
-to free slip conditions as well.
-
-The maximum depth of the slab is tracked for all the simulations (Fig. \ref{fig:subduction}) and they are compared with results obtained by
-\citet{Schmeling2008} with models with regular grids and comparable grid resolutions (rectangles in Fig. \ref{fig:subduction}). Low viscous sticky air models
-(i.e., \SI{e19}{\pascal\s}) enlighten a strong dependence of the sinking velocity with the chosen average scheme (dashed lines in Fig. \ref{fig:subduction}),
-while high viscous sticky air models (i.e., \SI{e21}{\pascal\s}) show a higher resistance at the trench, underestimating the correct solution (dotted lines in
-Fig. \ref{fig:subduction}). A high resistance at the trench can be also observed in case of a true free surface for the low resolution grid used (continuous
-lines in Fig. \ref{fig:subduction}), as pointed out by \citet{Schmeling2008}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Spontaneous_subduction}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Subduction.pdf}
-\caption{Maximum depth of the slab as function of time for all the simulations of the \citet{Schmeling2008} benchmark, in case of harmonic, geometric and
-arithmetic means (red, green and blue lines, respectively) using a sticky air or a true free surface (discontinuous and continuous lines, respectively).
-Red, green and blue rectangular areas indicate the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{500}{\km} in case of sticky air,
-while the grey rectangular area indicates the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{600}{\km} in case of true free surface.}
-\label{fig:subduction}
-\end{figure}
-
-\subsection{The slab detachment}\label{sec:slab}
-The slab detachment benchmark is performed as described by \citet{Schmalholz2011} and \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{1000}{\km}
-and $L_y=$ \SI{660}{\km} and a grid resolution of \num{256x256} elements. A non-linear viscous T-shaped layer with $\rho_l=$ \SI{3300}{\kg\per\cubic\metre} is
-placed at the top of the domain and surrounded by a linear viscous fluid with $\rho_f=$ \SI{3150}{\kg\per\cubic\m}. The top layer is 80 km-thick and a 250
-km-long and 80 km-wide slab is placed at $x=L_x/2$ (Fig. \ref{fig:slab}a). The effective viscosity of the top layer is given by
-\[\eta_{eff}=\eta_0 I_2^{\frac{(1-n)}{n}}\]
-with $\eta_0=$ \SI{4.75e11}{\pascal\s} and $n=4$, while the surrounding fluid has $\eta_f=$ \SI{e21}{\pascal\s}. The effective viscosity is capped between
-$\eta_{min}=$ \SI{e21}{\pascal\s} and $\eta_{max}=$ \SI{e25}{\pascal\s}. Velocity boundary conditions are set to free slip at the top and the bottom, and to
-no slip at the sides of the domain.
-
-The sinking of the slab determines a decrease of the effective viscosity in the T-shaped layer during the first part of the evolution (Fig. \ref{fig:slab}a),
-as consequence of local strain rates, while effective viscosities increase up to $\eta_{max}$ in the top layer after approximately \SI{20}{\mega\year}, when
-necking is complete (Fig. \ref{fig:slab}b).
-
-The width of necking is tracked by means of markers position at the side of the slab. Necking width and time are normalised in function of the initial slab
-(\SI{80}{\km}) and the characteristic time ($t_c=$ \SI{7.1158e14}{\s}) \citep{Schmalholz2011,Glerum2018}. The results in case of arithmetic, geometric and
-harmonic mean of the viscosity are shown in Fig. \ref{fig:necking}, compared with results from \citet{Glerum2018} for which an infinity norm mean was used.
-The influence of the average schemes agrees with results shown in Section \ref{sec:subduction}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Slab_detachment}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Slab.pdf}
-\caption{Effective viscosity for the slab detachment benchmark at the beginning of the evolution (panel a) and when necking is complete (panel b).}
-\label{fig:slab}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Necking.pdf}
-\caption{Normalised width of the necking of the slab detachment benchmark as function of normalised time.}
-\label{fig:necking}
-\end{figure}
-
-\subsection{The indenter experiment}\label{sec:indenter}
-The indenter benchmark simulate a rigid punch on a purely plastic von Mises material, for which exists an analytical solution
-\citep{Thieulot2008,Thieulot2014,Glerum2018}. The domain is a unit square with a grid resolution of \num{256x256} elements. The gravity acceleration is set
-to $\bm{g}=\bm{0}$ and effective viscosity is capped using $\eta_{min}=10^{-4}$ and $\eta_{max}=10^3$. The tolerance for the convergence of the non-linear
-solution is $tol=10^{-9}$, with a maximum number of non-linear iterations set to 500. The medium in the domain has $\rho=0.01$, an initial viscosity
-$\eta_0=10$, cohesion $C=1$ and an angle of internal friction $\phi=0$°. Velocity boundary conditions are set to no slip at the bottom and free slip at the
-sides of the domain. The top of the domain is open with the exception of the central portion (punch area) with width $w_p=0.125$, where $v=-1.05$ and $u$ is
-fixed either to 0 (rough punch experiment) or free (smooth punch experiment).
-
-Fig. \ref{fig:indenter} show the results in terms of viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure
-(panels g and h) for both experiments. The analytical solution indicate that pressure at ($0.5;1$) and ($0.5 \pm w_p;1$)is $p=\pi +1$ and $p=1$, respectively
-(black and red lines in Fig. \ref{fig:smooth_p}). Fig. \ref{fig:smooth_p} shows a clear improvement in the pressure solution when passing from a rough to a
-smooth punch (green and blue lines in Fig. \ref{fig:smooth_p}, respectively), as previously observed by \citet{Thieulot2014} and \citet{Glerum2018}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Indenter}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Indenter.pdf}
-\caption{Viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure (panels g and h) fields for rough (left column) and
-smooth (right column) punch experiments.}
-\label{fig:indenter}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Smooth_Rough.pdf}
-\caption{Nodal pressure in function of the x coordinate at $y=1$ for rough (green line) and smooth (blue line) punch experiment. Black and red lines indicate
-the analytical solution at $x=0.5$ and $x=0.5 \pm w_p$, respectively.}
-\label{fig:smooth_p}
-\end{figure}
-
-\subsection{The brick experiment}\label{sec:brick}
-An instantaneous version of the brick benchmark is performed to verify the correctness of the pressure-dependent plasticity for different angle of internal
-friction, as proposed by \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{40}{\km} and $L_y=$ \SI{10}{\km} and a grid resolution of \num{512x128}
-elements. The gravity acceleration is set to $g_y=$ \SI{-10}{\m\square\s} and effective viscosity is capped using $\eta_{min}=$ \SI{e19}{\pascal\s} and
-$\eta_{max}=$ \SI{e26}{\pascal\s}. The tolerance for the convergence of the non-linear solution is $tol=10^{-7}$, with a maximum number of non-linear iterations
-set to 1000. A 800 m-wide and 400 m-height inclusion with $\eta_b=$ \SI{e20}{\pascal\s} and $\rho_b=$ \SI{2700}{\kg\per\cubic\m} is placed at the bottom of the
-domain at $x=L_x/2$. The inclusion is surrounded by a non-linear viscous medium with $\rho_m=$ \SI{2700}{\kg\per\cubic\m}, an initial viscosity of
-\SI{e23}{\pascal\s}, a linear viscous viscosity of \SI{e25}{\pascal\s} and a cohesion of \SI{40}{\mega\pascal}. Velocities are set to free slip conditions at
-the bottom of the domain and the top is open. Velocities on sides of the domain are fixed to $u=$ \SI{+-2e-11}{\m\per\s} and $v=0$. The experiment is performed
-in compressional and extensional contexts with an angle of internal friction $\phi$ of the non-linear medium variable between 0° and 30°.
-
-As expected, two shear bands stem from the inclusion with variable angles in relation to both the dynamics context and the internal friction angle. Shear band
-angles formed at 45° for $\phi=0$° in both extensional and compressional contexts (Fig. \ref{fig:brick_beam}a), while they are different in case of $\phi\neq0$°.
-Shear band angles in case of $\phi=20$° are shown for compression and extension in Fig. \ref{fig:brick_beam}b  and c, respectively. Values of shear band angles
-as function of internal friction angles from 0° to 30° are extracted at different depths and minimum and maximum values are plotted in Fig. \ref{fig:brick},
-in comparison with theoretical Roscoe, Arthur and Coulomb shear band angles. For all tests the tolerance for nonlinear convergence is defined $tol=10^{-7}$
-and none converges before the maximum number of iterations ($it_{max}=1000$) is reached. However, tests with internal friction angles up to 15° show a constant
-decrease of the velocity residuals, which can not be observed for for tests with higher internal friction angles (Fig. \ref{fig:convergence}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Brick_experiment}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Brick_Beam.pdf}
-\caption{Shear band angles predicted for the brick experiment for $\phi=0$° (panel a) and for $\phi=20$° in case of compression and extension (panels b and c,
-respectively).}
-\label{fig:brick_beam}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Brick.pdf}
-\caption{Shear band angles predicted for the brick experiment in case of compressional and extensional contexts as function of different internal angle of
-friction (continuous black line and red dots), compared with theoretical Roscoe, Artur and Coulomb angles (discontinuous black lines). Red lines indicate the
-range of angles calculated at different depths.}
-\label{fig:brick}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \noindent\includegraphics[width=500px]{./Figures/convergence.pdf}
-  \caption{Convergence rates for all tests. Normalised residuals (left column) are compared with velocity residuals (right column) for extensional (first row)
-  and compressional (second row) settings.}
-  \label{fig:convergence}
-  \end{figure}
-
-\subsection{Advection stabilisation}\label{sec:advection}
-The 1D advection problem proposed by \citet{Donea2003} and \citet{Thieulot2011} is performed to verify the effectiveness of the SUPG method to stabilise the
-advection term of the energy equation. The domain is a 1D segment with $L_x=1$ composed by 50 elements and a discontinuity in the thermal field placed at
-$x=0.25$. Temperature is set to 1 for $x \leq 0.25$ and to 0 for $x > 0.25$. Velocity is set to ${u}=1$ in the entire domain. The simulation is performed for
-250 time steps, with $dt=0.002$, so the thermal discontinuity should be at $x=0.75$ at the end of the simulation.
-
-Temperature profiles at the end of the simulation are shown in \ref{fig:advection} as function of the dimensionless coefficient $\gamma=\tau \bm{v}/h$. In case
-of the classic Galerkin method ($\gamma=0$, blue line in \ref{fig:advection}) the final thermal profile is characterised by strong oscillations, which are
-eliminated in case of the SUPG method ($\gamma=0.045$, orange line in \ref{fig:advection}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Advection_stabilisation}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Advection.pdf}
-\caption{Temperature profile as function of \textit{x} for the advection stabilisation benchmark. Purple line indicates the initial temperature profile; the
-green line indicates the analytical temperature profile after 250 time steps; blue line indicates the temperature profile after 250 time steps in case of the
-classic Galerkin method ($\gamma=0$); orange line indicates the temperature profile after 250 time steps in case of the SUPG method ($\gamma=0.045$).}
-\label{fig:advection}
-\end{figure}
-
-\subsection{Shear heating}\label{sec:simple_shear}
-This exercise is performed in an unit square domain composed by \num{128x128} elements. The velocity field is prescribed on the entire domain with
-$\bm{v}=(L_y-y)y\bm{e_x}$; viscosity, density and specific heat are fixed to 1, while thermal conductivity and radiogenic energy are fixed to 0. Therefore,
-the energy equation (Eq. \ref{eq:energy}) can be simplified as \[\frac{\partial T}{\partial t}=H_s\] and fixing $T(t=0)=0$, the temperature can be find as
-\[T(t)=H_s t\]
-In this case we have
-\begin{eqnarray}
-\dot{\epsilon_{xx}}&=&\frac{\partial u}{\partial x}=0 \nonumber \\
-\dot{\epsilon_{yy}}&=&\frac{\partial v}{\partial y}=0 \nonumber \\
-\dot{\epsilon_{xy}}&=&\frac{1}{2}\left(\frac{\partial v}{\partial x}+\frac{\partial u}{\partial y}\right)=\frac{1}{2}(L_y-2y)\nonumber
-\end{eqnarray}
-and, simplifying Eq. \ref{eq:shear}, shear heating can be calculated as \[H_s(x,y)=(1-2y)^2\]
-
-The solution predicted by the model in terms of velocity, temperature and shear heating match well with the analytical solutions (Fig. \ref{fig:analytical_en}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Simple_shear}.
-
-%\begin{figure}
-%\centering
-%\noindent\includegraphics[width=400px]{./Figures/Energy_Thieulot.pdf}
-%\caption{Temperature and shear heating predicted by the model after 1 time step.}
-%\label{fig:en_thieulot}
-%\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=500px]{./Figures/Analytical_Shear.pdf}
-\caption{Velocity (panel a), temperature (panel b) and shear heating (panel c) as function of the \textit{y} coordinate for the simple shear experiment. The
-solutions predicted by the model (red lines) are compared with the analytical solutions (dashed blue lines).}
-\label{fig:analytical_en}
-\end{figure}
-
-\subsection{Shear and adiabatic heating}\label{sec:shear}
-This problem is performed as presented in Exercise 9.4 in \citet{Gerya2010b}. Two materials are vertically separated in a rectangular domain with $L_x=$
-\SI{1000}{\km}, $L_y=$ \SI{1500}{\km} and a grid resolution of \num{30x20} elements. Constant thermal coefficient expansion ($\alpha=$ \SI{3e-5}{\per\kelvin}),
-temperature ($T=$ \SI{1300}{\kelvin}) and gravity acceleration ($g_y=$ \SI{-10}{\m\square\s}) are assumed in the whole domain. Fluid 1 (on the left side of the
-domain) has $\rho_1=$ \SI{3200}{\kg\per\cubic\m} and $\eta_1=$ \SI{e20}{\pascal\s}; fluid 2 (on the right side of the domain) has $\rho_2=$
-\SI{3300}{\kg\per\cubic\m} and $\eta_2=$ \SI{e22}{\pascal\s}. Velocity boundary conditions are set to free slip on all sides of the domain.
-
-As shown in Fig. \ref{fig:energy}, both shear and adiabatic heating predicted by the code (first row in Fig. \ref{fig:energy}) well recreate the results
-obtained by means of example ${Shear\_adiabatic\_heating.m}$ from \citet{Gerya2010b} (second row in Fig. \ref{fig:energy}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Adiabatic%2BShear}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Energy.pdf}
-\caption{Comparison between shear (first column), adiabatic (second column) and total (third column) energy predicted by the code (panels a, b and c) and those
-created using example ${Shear\_adiabatic\_heating.m}$ from Exercise 9.4 in \citet{Gerya2010b} (panels d, e and f).}
-\label{fig:energy}
-\end{figure}
-
-\subsection{Mantle convection}\label{sec:mantle}
-This problem is performed as presented by \citet{Blankenbach1989} (constant viscosity cases) and \citet{Thieulot2014}, in a 2D unit square domain with gravity
-acceleration $g_y=10^{10}Ra$. The experiment is performed with three different Rayleigh numbers ($Ra=10^4$, $10^5$ and $10^6$) and with different grid
-resolution (between \num{32x32} and \num{128x128} elements). The fluid has constant viscosity, initial density, heat capacity, thermal conductivity
-($\eta= \rho_0= C_p = k=1$), reference temperature ($T_0=0$) and thermal expansion coefficient ($\alpha=10^{-10}$). Temperatures are set to 0 on top and 1 on
-bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\]
-
-The solution generated by the code in terms of $v_{\textrm{rms}}$ (calculated as in Sec. \ref{sec:error}) and Nu as function of time are reported for all the
-simulations in Fig. \ref{fig:mantle}. At the steady state, $v_{\textrm{rms}}$ and Nu of all simulations converge well toward the values from \citet{Blankenbach1989},
-with lower errors for higher resolution grids (Table \ref{tab:mantle}). All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Mantle_convection}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Convection.pdf}
-\caption{$v_{\textrm{rms}}$ (panels a, c and e) and Nu (panels b, d and f) for the mantle convection benchmark as function of time for different grid resolution.
-Panels a and b show the results for $Ra=10^4$; Panels c and d show the results for $Ra=10^5$; panels e and f show the results for $Ra=10^6$. Purple lines
-indicate the convergence values for the $v_{\textrm{rms}}$ and the Nu from \citet{Blankenbach1989} at the steady state.}
-\label{fig:mantle}
-\end{figure}
-
-\begin{table}[H]
-\caption{Comparison between $v_{\textrm{rms}}$ and Nu predicted by the code for the mantle convection experiment and same values as reported in literature.}
-\centering
-\small
-\begin{tabular}{|>{\centering}m{0.09\textwidth}|>{\centering}m{0.05\textwidth}|>{\centering}m{0.24\textwidth}|>{\centering}m{0.15\textwidth}|
-  >{\centering}m{0.1\textwidth} >{\centering}m{0.1\textwidth} >{\centering\arraybackslash}m{0.1\textwidth}|}
-\toprule
-& & \multirow{2}{*}{\citet{Blankenbach1989}} & \citet{Thieulot2014} & \multicolumn{3}{c|}{FALCON}  \\
- & & & \num{200x200} & \num{32x32} & \num{64x64} & \num{128x128}  \\
-\midrule
-\multirow{2}{*}{$Ra=10^4$} & $v_{\textrm{rms}}$ & $42.864947 \pm 0.000020$ & 42.867 & 42.83226 & 42.852793 & 42.861394  \\
-           & Nu      & $4.884409 \pm 0.000010$  & 4.882  & 4.781297 & 4.857475  & 4.877573   \\
-\hline
-\multirow{2}{*}{$Ra=10^5$} & $v_{\textrm{rms}}$ & $193.21454 \pm 0.00010$  & 193.255& 193.872643 & 193.377472 & 193.252290  \\
-           & Nu      & $10.534095 \pm 0.000010$ & 10.507 & 9.602514 & 10.270735  & 10.465629   \\
-\hline
-\multirow{2}{*}{$Ra=10^6$} & $v_{\textrm{rms}}$ & $833.98977 \pm 0.00020$  & 834.712& 848.091176 & 837.767911 & 834.945793  \\
-           & Nu      & $21.972465 \pm 0.000020$ & 21.695 & 15.999266 & 19.703682  & 21.306939   \\
-\bottomrule
-\end{tabular}
-\label{tab:mantle}
-\end{table}
-
-\subsection{Viscoplastic mantle convection}\label{sec:visco_mantle}
-This benchmark is performed as Cases 1-5a described in \citet{Tosi2015} in a 2D unit square domain with two different grid resolutions of \num{32x32} 
-and \num{100x100} elements. Reference viscosity, reference density, heat capacity, thermal conductivity and thermal expansion coefficient are fixed to a constant
-value of 1 ($\eta_0= \rho_0= C_p = k=\alpha=1$), while the gravity has been chosen as $g_y=10^2$, in order to obtain a Rayleigh number $Ra=10^2$. Temperatures
-are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\].
-The viscosity field $\eta$ is calculated as the harmonic average between a linear part $\eta_{lin}$ that depends on temperature only or on temperature and 
-depth $y$ and a nonlinear-plastic part $\eta_{pl}$ that depends on the strain rate $\dot{\bm{\epsilon}}$, as follows
-\[\eta(T,y,\dot{\bm{\epsilon}})=2\left(\frac{1}{\eta_{lin}(T,y)}+\frac{1}{\eta_{pl}(\dot{\bm{\epsilon}})}\right)^{-1}\]
-The linear and the nonlinear-plastic parts of the viscosity are calculated following \citet{Tosi2015} as follows
-\[\eta_{lin}(T,y)=\exp(-\ln(\Delta\eta_T)T+\ln(\Delta\eta_y)y)\]
-\[\eta_{pl}(\dot{\bm{\epsilon}})=\eta^*+\frac{\sigma_Y}{\sqrt{\dot{\bm{\epsilon}}:\dot{\bm{\epsilon}}}}\]
-where $\Delta\eta_T$, $\Delta\eta_y$, $\eta^*$ and $\sigma_Y$ are parameters chosen as listed in Table \ref{tab:tosi_setup}.
-
-For each case, some diagnostic quantities measured at the steady state are reported in Table \ref{tab:visco_mantle} (see Section \ref{sec:error} for 
-the computation of all the quantities), showing a well fit with results described in \citet{Tosi2015}. Data and charts for all cases can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Viscoplastic_mantle_convection}.
-
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{white!100!}{gray!25!}
-%\begin{table}[H]
-%  \caption{Results of the viscoplastic mantle convection benchmark for Cases 1-4 compared with results obtained with ELEFANT and YACC (from \citealp{Tosi2015}).}
-%  \centering
-%  \resizebox{\textwidth}{!}{
-%\begin{tabular}{|c|ccc|ccc|ccc|ccc|}
-%  \toprule
-%    & \multicolumn{3}{c|}{\textbf{Case 1}} & \multicolumn{3}{c|}{\textbf{Case 2}} & \multicolumn{3}{c|}{\textbf{Case 3}} &
-%    \multicolumn{3}{c|}{\textbf{Case 4}} \\
-%   & FALCON & ELEFANT& YACC & FALCON& ELEFANT& YACC & FALCON & ELEFANT & YACC & FALCON & ELEFANT & YACC \\
-%  \midrule
-%   $\left\langle T \right\rangle$ & 0.7759 & 0.7758 & 0.7767 & 0.6043 & 0.6036 & 0.6058 & 0.7275 & 0.7275 & 0.7286 & & 0.5277 & 0.5289\\
-%    Nu$^{top}$ & 3.4280 & 3.4214 & 3.4298 & 8.5242 & 8.5115 & 8.5278 & 3.0368 & 3.0347 & 3.0374 && 6.5912 & 6.5572 \\
-%    Nu$^{bottom}$ & 3.3168 & 3.313 & 3.3143 & 8.4599 & 8.4465 & 8.3990 & 3.0086 & 2.9908 & 2.9628 && 6.5834 & 6.5243 \\
-%    $v_{\textrm{rms}}$ & 249.901 & 249.134 & 251.7997 & 141.2451 & 140.7087 & 142.2020 & 100.2776 & 100.1208 & 100.9467 && 79.1105 & 79.6202 \\
-%    $v_{\textrm{rms}}^{top}$ & 1.8907 & 1.8642 & 1.8298 & 104.7928 & 104.1228 & 105.1767 & 2.0832 & 2.0652 & 2.0374 && 74.7596 & 75.8414 \\
-%    $u_{max}^{top}$ & 2.6340 & 2.6119 & 2.5516 & 121.9212& 121.766 & 122.3238 & 2.9086 & 2.9019 & 2.8458 && 88.9146 & 89.2940 \\
-%    $\eta_{min}$ & - & - & - & $2.0303\times10^{-5}$ & $2.0600\times10^{-5}$ & $1.9853\times10^{-5}$ & $4.8129\times10^{-5}$ & $4.8080\times10^{-5}$ & $4.7907\times10^{-5}$ &
-%    & $1.9860\times10^{-4}$ & $1.9174\times10^{-4}$ \\
-%    $\eta_{max}$ & - & - & - & 1.3325 & 1.85 & 1.6095 & 0.9360 & 0.9023 & 1 && 1.52 & 1.6773 \\
-%    $\left\langle W \right\rangle$ & 2.4299 & 2.4316 & 2.4583 & 7.5648 & 7.5665 & 7.6222 & 2.0387 & 2.0384 & 2.0400 && 5.6216 & 5.6512 \\
-%    $\left\langle \Phi \right\rangle/Ra$ & 2.4311 & 2.4276 & 2.4333 & 7.5661 & 7.5585 & 7.6194 & 2.0386 & 2.0370 & 2.0335 && 5.6182 & 5.6463 \\
-%    $\delta$ & 0.05\% & 0.16\% & 1.02\% & 0.02\% & 0.11\% & 0.04\% & $<$0.01\% & 0.07\% & 0.32\% && 0.06\% & 0.09\% \\
-%  \bottomrule
-%  \end{tabular}}
-%  \label{tab:visco_mantle}
-%  \end{table}
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{2}{white!100!}{gray!25!}
-\begin{table}[H]
-  \caption{Benchmark cases and corresponding parameters.}
-  \centering
-  \small
-\begin{tabular}{|>{\centering}m{0.08\textwidth}|>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}
-  >{\centering}m{0.05\textwidth} >{\centering}m{0.05\textwidth} |>{\centering\arraybackslash}m{0.2\textwidth}|}
-  \toprule
-    Case & $Ra$ & $\Delta\eta_T$ & $\Delta\eta_y$ & $\eta^*$ & $\sigma_Y$ & Convective regime \\
-  \midrule
-  \rowcolor{gray!25!}
-    1 & $10^2$ & $10^5$ & 1 & - & - & Stagnant lid \\
-    2 & $10^2$ & $10^5$ & 1 & $10^{-3}$ & 1 & Mobile lid  \\
-    \rowcolor{gray!25!}
-    3 & $10^2$ & $10^5$ & 10 & - & - & Stagnant lid  \\
-    4 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 1 & Mobile lid \\
-    \rowcolor{gray!25!}
-    5 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 4 & Periodic \\
-  \bottomrule
-  \end{tabular}
-  \label{tab:tosi_setup}
-  \end{table}
-
-\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{gray!25!}{white!100!}
-\begin{table}
-  \caption{Results of the viscoplastic mantle convection benchmark for different grid resolutions in Cases 1-5 compared with results obtained with ELEFANT
-  and YACC ($100\times100$ elements, from \citealp{Tosi2015}) and in Stone 28 ($32\times32$ elements, from
-  \url{https://github.com/cedrict/fieldstone/tree/master/python_codes/fieldstone_28}).}
-  \centering
-  \resizebox{\textwidth}{!}{
-\begin{tabular}{|cc|ccccccccccc|}
-  \toprule
-  \multicolumn{13}{|c|}{\textbf{Case 1}} \\
-  \midrule
-  Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-  & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-  \hline
-  \rowcolor{gray!25!}
-    YACC & $100\times100$ & 0.7767 & 3.4298 & 3.3143 & 251.7997 & 1.8298 & 2.5516 & - & - & 2.4583 & 2.4333 & 1.02\% \\
-    ELEFANT & $100\times100$ & 0.7758 & 3.4214 & 3.313 & 249.134 & 1.8642 & 2.6119 & - & - & 2.4316 & 2.4276 & 0.16\% \\
-    \rowcolor{gray!25!}
-    Stone 28 & $32\times32$ & 0.7737 & 3.3987 & - & 243.872 & - & - & - & - & - & - & - \\
-%    FALCON & $32\times32$ & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & - & - & 2.4745 & 2.4864 & 0.48\% \\
-%    \rowcolor{gray!25!}
-    FALCON & $100\times100$ & 0.7759 & 3.4280 & 3.3168 & 249.901 & 1.8907 & 2.6340 & - & - & 2.4299 & 2.4311 & 0.05\% \\
-    \midrule
-    \multicolumn{13}{|c|}{\textbf{Case 2}} \\
-    \midrule
-    Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-    & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-    \hline
-    \rowcolor{gray!25!}
-     YACC & $100\times100$ & 0.6058 & 8.5278 & 8.3990 & 142.2020 & 105.1767 & 122.3238 & $1.9853\times10^{-5}$ & 1.6095 & 7.6222 & 7.6194 & 0.04\% \\
-     ELEFANT & $100\times100$ & 0.6034 & 8.5115 & 8.4465 & 140.8067 & 104.1228 & 121.766 & $2.0600\times10^{-5}$ & 1.85 & 7.5665 & 7.5585 & 0.11\% \\
-     \rowcolor{gray!25!}
-     Stone 28 & $32\times32$ & 0.6052 & 8.1715 & - & 141.518 & - & - & - & - & - & - & - \\
-%     FALCON & $32\times32$ & 0.6454 & 8.1422 & 7.4020 & 156.6908 & 95.7226 & 116.9302 & $2.2572\times10^{-5}$ & 0.7194 & 7.4289 & 7.4452 & 0.22\% \\
-%     \rowcolor{gray!25!}
-     FALCON & $100\times100$ & 0.6044 & 8.5266 & 8.4590 & 141.3684 & 104.8867 & 122.0470 & $2.0333\times10^{-5}$ & 1.2914 & 7.5707 & 7.5718 & 0.01\% \\
-     \midrule
-     \multicolumn{13}{|c|}{\textbf{Case 3}} \\
-     \midrule
-     Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-     & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-     \hline
-     \rowcolor{gray!25!} 
-     YACC & $100\times100$ & 0.7286 & 3.0374 & 2.9628 & 100.9467 & 2.0374 & 2.8458 & $4.7907\times10^{-5}$ & 1 & 2.0400 & 2.0335 & 0.32\% \\
-      ELEFANT & $100\times100$ & 0.7275 & 3.0347 & 2.9908 & 100.1208 & 2.0652 & 2.9019 & $4.8080\times10^{-5}$ & 0.9023 & 2.0384 & 2.037 & 0.07\% \\
-      \rowcolor{gray!25!}
-      Stone 28 & $32\times32$ & 0.7277 & 3.0347 & - & 100.018 & - & - & - & - & - & - & - \\
-%      FALCON & $32\times32$ & 0.7281 & 3.0618 & 2.8132 & 101.0092 & 2.1630 & 3.0178 & $4.8500\times10^{-5}$ & 0.7845 & 2.0699 & 2.0695 & 0.02\% \\
-%      \rowcolor{gray!25!}
-      FALCON & $100\times100$ & 0.7275 & 3.0368 & 3.0086 & 100.2776 & 2.0832 & 2.9086 & $4.8129\times10^{-5}$ & 0.9360 & 2.0387 & 2.0386 & $<$0.01\% \\
-      \midrule
-      \multicolumn{13}{|c|}{\textbf{Case 4}} \\
-      \midrule
-      Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-      & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-      \hline
-      \rowcolor{gray!25!} 
-      YACC & $100\times100$ & 0.5289 & 6.5572 & 6.5243 & 79.6202 & 75.4814 & 89.2940 & $1.9174\times10^{-4}$ & 1.6773 & 5.6512 & 5.6463 & 0.09\% \\
-       ELEFANT & $100\times100$ & 0.5277 & 6.5912 & 6.5834 & 79.1105 & 74.7596 & 88.9146 & $1.9860\times10^{-4}$ & 1.5200 & 5.6216 & 5.6182 & 0.06\% \\
-       \rowcolor{gray!25!}
-       Stone 28 & $32\times32$ & 0.5286 & 6.4036 & - & 79.4746 & - & - & - & - & - & - & - \\
-%       FALCON & $32\times32$ & 0.5334 & 6.4679 & 6.3435 & 80.6153 & 76.0511 & 90.2058 & $2.0393\times10^{-4}$ & 0.6453 & 5.6676 & 5.6698 & 0.04\% \\
-%       \rowcolor{gray!25!}
-       FALCON & $100\times100$ & 0.5279 & 6.5897 & 6.5830 & 79.1220 & 75.0715 & 88.8484 & $1.9693\times10^{-4}$ & 1.4133 & 5.6127 & 5.6113 & 0.03\% \\
-       \midrule
-       \multicolumn{13}{|c|}{\textbf{Case 5}} \\
-       \midrule
-       Code & Resolution & $P$ & $\left\langle T \right\rangle_{min}$ & $\left\langle T \right\rangle_{max}$ & Nu$^{top}_{min}$ & Nu$^{top}_{max}$ &
-       $v_{\textrm{rms}}^{min}$ & $v_{\textrm{rms}}^{max}$ & \multicolumn{2}{c}{$(\left\langle \Phi \right\rangle/Ra)_{min}$} & 
-       \multicolumn{2}{c|}{$(\left\langle \Phi \right\rangle/Ra)_{max}$} \\
-       \hline
-       \rowcolor{gray!25!}
-         YACC & $100\times100$ & 0.0792 & 0.6543 & 0.6722 & 2.6950 & 7.2792 & 41.9813 & 98.8230 & \multicolumn{2}{c}{1.4347} & 
-         \multicolumn{2}{c|}{9.2748} \\
-         ELEFANT & $100\times100$ & 0.076 & 0.6514 & 0.6693 & 2.6808 & 7.3506 & 41.2578 & 100.089 & \multicolumn{2}{c}{1.3324} & 
-         \multicolumn{2}{c|}{9.5289} \\
-         \rowcolor{gray!25!}
-         Stone 28 & $32\times32$ & 0.0764 & 0.6521 & 0.6697 & 2.6912 & 7.0763 & 41.61 & 98.183 & \multicolumn{2}{c}{-} & 
-         \multicolumn{2}{c|}{-} \\
-     %    FALCON & $32\times32$ & & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & \multicolumn{2}{c}{-} & 
-     %     \multicolumn{2}{c|}{-} \\
-     %    \rowcolor{gray!25!}
-         FALCON & $100\times100$ & 0.0775 & 0.6519 & 0.6697 & 2.6847 & 7.3629 & 41.5256 & 99.5466 & \multicolumn{2}{c}{1.3399} & 
-         \multicolumn{2}{c|}{9.3993} \\
-  \bottomrule
-  \end{tabular}}
-  \label{tab:visco_mantle}
-  \end{table}
-
-\subsection{Thin layer entrainment}\label{sec:thin}
-This experiment is performed as originally proposed by \citet{vanKeken1997} with two incompressible fluids in a rectangular domain with $L_x=2$ and $L_y=1$
-and gravity acceleration $g_y=10^{10}Ra$. The Rayleigh number and the compositional Rayleigh number are fixed to $Ra=300000$ and $Rc=450000$, respectively.
-Both fluids have constant viscosity, thermal conductivity, specific heat ($\eta=\rho=C_p=1$) and thermal expansion coefficient ($\alpha=10^{-10}$). Fluid 1
-has a density $\rho_1=1$, while fluid 2 is denser ($\rho_2=\rho_1+1.5 \cdot 10^{-10}$) and is placed at the bottom of the domain, for $y \leq 0.025$.
-Temperature are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides of the domain. The initial
-temperature field is given by
-\[T(x,y)=T_u(x,y)+T_l(x,y)+T_r(x,y)+T_s(x,y)-\frac{3}{2}\]
-with
-\begin{eqnarray}
-T_u(x,y)&=&\frac{1}{2}erf\left(\frac{1-y}{2}\sqrt{\frac{u_0}{x}}\right)\nonumber \\
-T_l(x,y)&=&1-\frac{1}{2}erf\left(\frac{y}{2}\sqrt{\frac{u_0}{L_x-x}}\right)\nonumber \\
-T_r(x,y)&=&\frac{1}{2}+\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{y+1}} \exp\left(-\frac{x^2u_0}{4y+4}\right)\nonumber \\
-T_s(x,y)&=&\frac{1}{2}-\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{2-y}} \exp\left(-\frac{(L_x-x)^2u_0}{8-4y}\right)\nonumber
-\end{eqnarray}
-and
-\begin{eqnarray}
-u_0&=&\frac{L_x^{7/3}}{(1+L_x^4)^{2/3}}\left(\frac{Ra}{2\sqrt{\pi}}\right)^{\frac{2}{3}}\nonumber \\
-Q&=&2\sqrt{\frac{L_x}{\pi u_0}}\nonumber
-\end{eqnarray}
-
-The experiment is performed with two grid resolution (\num{125x40} and \num{200x80} elements), with 100 markers per element and CFL of 0.25. $v_{\textrm{rms}}$
-(see Sec. \ref{sec:error}) calculated for both the simulations match well with results from \citet{vanKeken1997} and obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and ELEFANT \citep{Thieulot2014}, obtained with similar grid resolutions
-(Fig. \ref{fig:thin}). In addition, a variety of simulations with different resolutions and aspect ratios are performed to verify that the sensitivity of the
-initial velocity field. Results match well with results from \citet{vanKeken1997} (in the grey area) and \citet{Thieulot2014} (Fig. \ref{fig:thin_initial}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Thin_layer}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_thin.pdf}
-\caption{$v_{\textrm{rms}}$ for the thin layer experiment as function of time for different grid resolution. Results are compared with results obtained by
-\citet{vanKeken1997} (black lines) and with ASPECT and ELEFANT.}
-\label{fig:thin}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrmszero.pdf}
-\caption{$v_{\textrm{rms}}$ at $t=0$ for the thin layer experiment as function of the element size for different aspect ratios of the elements. Results are compared
-with results obtained by \citet{vanKeken1997} (black dots in the grey area) and with ASPECT and ELEFANT.}
-\label{fig:thin_initial}
-\end{figure}
-
-\subsection{Sinking hydrated cylinder}\label{sec:quinquis}
-This experiment is performed as explained in \citet{Quinquis2014} and simulates the sinking of a cold, hydrated cylinder into a warm, dry mantle. The domain
-is square with $L_x=L_y=$ \SI{300}{\km}, a grid resolution of \num{300x300} elements and 25 markers per element. The sinking cylinder is located at $x=$
-\SI{150}{\km} and $y=$ \SI{170}{\km} with a radius of \SI{20}{\km} and has $\rho_c=$ \SI{3250}{\kg\per\cubic\m}, $\eta_c=$ \SI{e23}{\pascal\s}, a thermal
-conductivity $k_c=$ \SI{4.5}{\watt\per\m\per\kelvin},a specific heat $C_{p_c}=$ \SI{1250}{\joule\per\kg\per\kelvin}; the surrounding mantle has $\rho_m=$
-\SI{3200}{\kg\per\cubic\m}, $\eta_m=$ \SI{e20}{\pascal\s}, a thermal conductivity $k_m=$ \SI{105}{\watt\per\m\per\kelvin} and specific heat $C_{p_m}=$
-\SI{1250}{\joule\per\kg\per\kelvin}; a 58 km-thick lithosphere overlies the mantle and has $\rho_l=$ \SI{3200}{\kg\per\cubic\m}, $\eta_l=$ \SI{e23}{\pascal\s},
-a thermal conductivity $k_l=$ \SI{4.5}{\watt\per\m\per\kelvin}, specific heat $C_{p_l}=$ \SI{750}{\joule\per\kg\per\kelvin} and a thermal diffusivity of
-\SI{e-6}{\square\m\per\s}. Velocity boundary conditions are set to free slip on all sides of the domain. Temperature increase linearly in the lithosphere,
-from 0° to 1300°C, and in the mantle, from 1300° to 1360.5°C. The initial temperature of the cylinder is fixed to 400°C. Throughout the evolution, temperature
-is fixed to 0° and 1360.5°C at the top and bottom of the domain, respectively. The initial bound water content is imposed at 2 wt.\% in the cylinder and at
-0 wt.\% in both the mantle and the lithosphere. The maximum amount of water in the mantle is fixed to 0.2 wt.\%, while maximum water content in the cylinder
-and in the lithosphere is function of pressure and temperature and is calculated using a serpentinized harzburgite with Perple\_X, as in \citet{Quinquis2014}.
-
-As shown in \citet{Quinquis2014}, the model is characterised by a progressive dehydration in the external portion of the cylinder, due to the increase of
-temperature, with a consequent vertical migration of free water that hydrates the mantle above the cylinder, up to the lithosphere. The distribution of
-bound and free water during the evolution of the experiment is shown in Fig. \ref{fig:quinquis} (left and right columns, respectively). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Hydration}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Hydration.pdf}
-\caption{Distribution of bound and free water (left and right columns, respectively) at $t=0.5$ and $t=$ \SI{1.25}{\mega\year} (first and second rows,
-respectively) for the sinking hydrated cylinder experiment. Only markers with bound or free water not equal to 0 are plotted.}
-\label{fig:quinquis}
-\end{figure}
-
-\subsection{Mantle melting curves}\label{sec:katz}
-This instantaneous experiment is performed in order to recreate isobaric melting curves, as function of temperature, bulk water and content of water in the
-melt, to compare with melting curves obtained by \citet{Katz2003}. The domain is square with $L_x=L_y=100$ km, a grid resolution of \num{100x100} elements and
-25 markers per element. Density is set in the entire domain to \SI{3300}{\kg\per\cubic\m} and temperature increases from the sides to the centre from 1000° to
-1500°C. The experiment is performed for bulk water contents from 0 to 0.3 wt.\%.
-
-Fig. \ref{fig:fraction} shows isobaric melting curves for \SI{1}{\giga\pascal} (panel a) and \SI{3}{\giga\pascal} (panel b) calculated for different bulk water
-contents. The curves well match with those obtained for the mantle by \citet{Katz2003}, showing the saturation at the solidus for bulk water of 0.2 and 0.3 wt.\%, indicating
-that the code correctly determines melt fraction as function of pressure, temperature and bulk water content. Same results can be observed in Fig.
-\ref{fig:water}a, where isobaric and isothermal melting curves are shown as function of bulk water content. In addition, Fig. \ref{fig:water}b shows that the
-code correctly determines also the percentage of water in the melt as function of pressure, temperature and melt fraction. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Melting}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Fraction.pdf}
-\caption{Isobaric melting curves for 1 GPa (panel a) and 3 GPa (panel b) as function of temperature with different bulk water contents.}
-\label{fig:fraction}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Water.pdf}
-\caption{Isobaric melting curves for 1.5 GPa as function of bulk water (panel a) and water in the melt (panel b) calculated for different temperatures.}
-\label{fig:water}
-\end{figure}
-
-\clearpage
-
-\bibliography{export}
-
-\end{document}

### 2022-09-07 tex/mom_eq.tex
update main

--- 2022-03-31 FALCON_description.tex
@@
-\documentclass[hidelinks,10pt,a4paper]{article}
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{textcomp}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage{booktabs}
-\usepackage{multirow}
-\usepackage[cm]{fullpage}
-\usepackage[round]{natbib}
-\bibliographystyle{abbrvnat}
-\usepackage{doi}
-\usepackage{bm}
-\usepackage[table]{xcolor}
-\hypersetup{
-    colorlinks,
-    linkcolor={blue},
-    citecolor={blue},
-    urlcolor={blue}
-}
-\usepackage{float}
-\restylefloat{table}
-\usepackage{siunitx}
-\DeclareSIUnit\year{yr}
-
-\begin{document}
-
-\section{Introduction}\label{sec:intro}
-This document includes a description of the code FALCON, with a presentation of the features implemented followed by results of extensive benchmarks listed in
-Table \ref{tab:list}, performed to verify that the code solves correctly a wide range of problems. Description and results of all benchmarks are shown
-in Section \ref{sec:benchmark}.
-
-\begin{table}[H]
-\caption{List of benchmarks performed for each feature of the code. Data of all benchmarks can be found at 
-\url{https://github.com/aleregorda/Benchmarks}}
-\centering
-\begin{tabular}{| l | l | l |}
-\toprule
- \textbf{Feature}  & \textbf{Benchmark} & \textbf{Section} \\
-\midrule
-  Solver  & Stokes flow & \ref{sec:stokes} \\
-\hline
-  \multirow{2}{*}{Markers advection} & Zalesak disk & \ref{sec:runge} \\
-    & Conservative Interpolation Velocity & \ref{sec:CVI} \\
-\hline
-  \multirow{4}{*}{Momentum equation}  & Poiseuille flow & \ref{sec:poiseuille}\\
-    & Instantaneous 2D sphere & \ref{sec:ist_sphere} \\
-    & Rayleigh-Taylor experiment & \ref{sec:rayleigh} \\
-    & Falling block & \ref{sec:block} \\
-\hline
-    & 2D Stokes sphere & \ref{sec:sphere} \\
-    Sticky air and  & Stabilisation algorithm & \ref{sec:stab} \\
-    free surface & Topography relaxation & \ref{sec:crameri} \\
-  & Spontaneous subduction & \ref{sec:subduction} \\
-\hline
-  Erosion and sedimentation  & - & - \\
-\hline
-   Non-linear  & Slab detachment & \ref{sec:slab} \\
-   visco-plasticity & Indenter & \ref{sec:indenter} \\
-    & Brick & \ref{sec:brick} \\
-\hline
-  \multirow{3}{*}{Energy equation}  & Advection stabilisation & \ref{sec:advection} \\
-    & Simple shear heating & \ref{sec:simple_shear} \\
-    & Shear and adiabatic heating & \ref{sec:shear} \\
-\hline
-\multirow{3}{*}{Energy + momentum} & Mantle convection & \ref{sec:mantle} \\
-& Viscoplastic mantle convection & \ref{sec:visco_mantle} \\
-& Thin layer entrainment & \ref{sec:thin} \\
-\hline
-  Phase changes  & \multirow{2}{*}{Hydrated sinking cylinder} & \multirow{2}{*}{\ref{sec:quinquis}} \\
-  and hydration  &  & \\
-\hline
-  Melting & Experimental melting curves & \ref{sec:katz} \\
-\bottomrule
-\end{tabular}
-\label{tab:list}
-\end{table}
-
-The code is written in Fortran90 and uses the finite element method (FEM) with quadrilateral $Q_1 \times P_0$ elements (continuous bilinear velocity and
-discontinuous constant pressure), associated with MUMPS\footnote{\url{http://mumps.enseeiht.fr/}} \citep{Amestoy2001,Amestoy2019} that is a software package
-for solving systems of linear equations of the form $A \cdot x = b$, where \textit{A} is a square sparse matrix, by means of a direct method. Since
-$Q_1 \times P_0$ elements do not satisfy Ladyzhenskaya, Babuska and Brezzi (LBB) stability condition \citep{Donea2003}, elemental pressure is elaborated
-in post-processing to avoid spurious pressures by means of a double interpolation to smooth the pressure. Following this procedure, the elemental pressure
-is interpolated onto nodes and then back onto elements. After the smoothing procedure, the pressure field is calculated again on the nodes to be used in
-combination with the lithostatic pressure, which is calculated onto nodes. Correctness of the solution and performances of the solver in terms of time and
-memory usage are tested solving the Stokes flow with the analytical solution proposed by \citet{Donea2003} (Section \ref{sec:stokes}).
-
-The thermo-mechanics of crust-mantle systems is described by means of the conservation of mass, momentum and energy equations, expressed as follows:
-\begin{eqnarray}
-\label{eq:mass}\frac{\partial \rho}{\partial t}+\bm{\nabla} \cdot (\rho \bm{u})&=&0\\
-\label{eq:momentum}-\bm{\nabla} p + \bm{\nabla} \cdot \bm{\tau} + \rho \bm{g}&=&0\\
-\label{eq:energy}\rho C_p \left(\frac{\partial T}{\partial t} + \bm{u} \cdot \bm{\nabla} T \right)&=&\bm{\nabla} \cdot \left(k\bm{\nabla} T\right) + H_{tot}
-\end{eqnarray}
-where $\rho$ is the density, $\bm{u}$ is the velocity, \textit{p} is the pressure, $\bm{\tau}$ is the deviatoric stress, $\bm{g}$ is the gravity acceleration,
-$C_p$ is the specific heat at constant pressure, \textit{T} is the temperature, \textit{k} is the thermal conductivity and $H_{tot}$ is the total internal heat
-production. Density variations due to temperature are generally small enough to assume the density as constant ($\rho=\rho_0$) in Eq. \ref{eq:mass} and in
-Eq. \ref{eq:energy}, while it must be treated as a variable in the buoyancy term of Eq. \ref{eq:momentum}, such that \[\rho=\rho_0(1-\alpha(T-T_0))\]
-where $\rho_0$ is the density at a reference temperature $T_0$ and $\alpha$ is the coefficient of thermal expansion. Eq. \ref{eq:mass} then can be rewritten as
-\begin{equation}\label{eq:mass_0}
-\bm{\nabla} \cdot \bm{u}=0
-\end{equation}
-This simplification is known as the extended Boussinesq approximation.
-
-The time step $dt$ is chosen in according to Courant-Friedrichs-Lewy condition \citep{Anderson1995}, such that
-\[dt=C_n\; \min\left(\frac{\Delta x}{\max|\bm{u}|},\frac{(\Delta x)^2}{\kappa}\right)\]
-where $0<C_n<1$ is the Courant (CFL) number, $\Delta x$ is the minimum dimension of smallest element, $\max|\bm{u}|$ is the maximum velocity calculated on the
-entire domain and
-$\kappa=\frac{k}{\rho C_p}$ is the thermal diffusivity \citep{Thieulot2014}.
-
-\section{Lagrangian markers}\label{sec:markers}
-Elemental properties (density, viscosity, thermal conductivity, specific heat and thermal expansion) needed to solve Eqs. \ref{eq:mass_0}, \ref{eq:momentum}
-and \ref{eq:energy} are related to the composition of each element, determined by means of Lagrangian markers that are characteristic of different materials
-in the domain. Elemental properties (with the sole exception of the viscosity) are calculated using an arithmetic mean, as \[P_e=\frac{1}{n}\sum_{i=1}^n P_i\]
-where $P_e$ is the elemental property, $P_i$ is the property characteristic of the material of each marker and $n$ is the number of markers of the element.
-Differently, the average scheme for the elemental viscosity can be chosen between harmonic, geometric and arithmetic mean. Markers can be placed regularly or
-randomly at the beginning of the simulation and their advection is performed by either a 2nd-order or a 4th-order Runge-Kutta in space, interpolating the velocity 
-field on each marker by means of the shape functions (see Section \ref{sec:runge}). The interpolated velocity is then corrected by means of the Conservative
-Velocity Interpolation (CVI), which introduces a corrective term to reduce dispersion and clustering of particles in both steady state and time-dependent 
-\citep{Wang2015} (see Section \ref{sec:CVI}).
-
-The initial distribution of the markers is created using the open-source code library
-Geodynamic World Builder \citep{Fraters2019}. During the simulation, each marker carries memory of temperature, pressure and accumulated strain, which are
-determined interpolating the relative nodal parameter. In particular, pressure is interpolated onto each marker after the smoothing procedure. The number
-of markers contained in each element is maintained between $n_{min}$ and $n_{max}$. When in an element there are less markers than $n_{min}$ the code adds
-random markers to reach the $n_{min}$, while if the number is higher than $n_{max}$ some of them are randomly deleted. When new markers are added, they assume
-the properties of the nearest marker. In this way elements are never empty and maintain a number of markers inside a prefixed range.
-
 \section{The momentum equation}\label{sec:mom_eq}
@@
 To support large viscosities variations the penalty parameter is related to the effective elemental viscosity by means of a dimensionless coefficient, so that
-$\lambda=\lambda_e(e)\eta_{eff}(e)$ \citep{Marotta2006,Dabrowski2008,Thieulot2014}. Benchmark of the falling block \citep{Gerya2003a,Gerya2010b,Thieulot2011}
+$\lambda=\lambda_e(e)\eta_{eff}(e)$ \citep{Marotta2006,Dabrowski2008,Thieulot2014}. Benchmark of the falling block \citep{Gerya2003a,Gerya2010,Thieulot2011}
 is performed to verify that the code can correctly deal buoyancy driven flows with strong viscosity contrasts (Section \ref{sec:block}).
-
-\section{Sticky air and free surface}\label{sec:sticky}
-The Earth's surface can be treated by means of either the so-called sticky-air or a true free surface method, both of them implemented in the code. In the
-sticky-air method the surface is approximated with the introduction of a buoyant layer with a viscosity at least four orders of magnitude lower than the crust
-\citep{Schmeling2008,Crameri2012} and the interface between lithosphere and air is defined using a chain of passive markers that are advected as the Lagrangian
-markers. The correctness of the evolution of the markers chain is tested with the experiment of a 2D time-dependent Stokes sphere below a free surface and
-compared with results from ASPECT\footnote{\url{https://aspect.geodynamics.org/}} \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} (Section
-\ref{sec:sphere}). In the true free surface case the top boundary is assumed stress-free and velocities are not fixed. In this case topography variations are
-described by vertical deformations of the mesh that depend on the velocity field of the nodes that identify the free surface, while horizontal deformations are
-not taken into account. This procedure is known as the Arbitrary Lagrangian-Eulerian (ALE) method and its implementation follows the technique described in
-\citet{Thieulot2011}. However, although the implementation of a true free surface better reproduces laboratory experiment, extremely small time steps can be
-necessary to maintain stability \citep{Kaus2010a,Quinquis2011,Thieulot2014}. Therefore, the stability algorithm proposed by \citet{Kaus2010a} is implemented
-to avoid instabilities due to high density differences at the free surface when using too large time steps. The implementation of this algorithm are tested by
-performing the experiment described by \citet{Kaus2010a} (Section \ref{sec:stab}).
-
-The topography relaxation benchmark proposed by \citet{Crameri2012} is performed to verify that the code correctly recovers topography variations in case of
-both the sticky-air and the true free surface method (Section \ref{sec:crameri}). Finally, Section \ref{sec:subduction} show the results of the spontaneous
-subduction experiment described by \citet{Schmeling2008}.
-
-\section{Erosion and sedimentation}\label{sec:erosion}
-Surface processes at either the lithosphere-air interface or in correspondence of the free surface have been implemented by means of the software
-Fastscape\footnote{\url{https://fastscape.org/}} \citep{Braun2013,Cordonnier2019,Yuan2019a,Yuan2019b}. FastScapeLib is a set of routines that solve the stream
-power law (enriched by a sediment transport/deposition term) \citep{Yuan2019a}, hillslope diffusion and marine transport and deposition \citep{Yuan2019b}.
-The partial differential equation solved by FastScapeLib is:
-\begin{equation}
-\frac{\partial h}{ \partial t}=U-K_f A^{m_e}S^{n_e} + \frac{G}{A}\int_A \left( U-\frac{\partial h}{\partial t} \right) dA + K_d \bm{\nabla}^2 h 
-\end{equation}
-where $h$ is the topography, $U$ is the uplift, $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $S$ is the slope, $m_e$ and $n_e$ are
-parameters that should be chosen to constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}, $G$ is the dimensionless
-deposition/transport coefficient for the enriched stream power law and $K_d$ is the effective diffusivity. Terms $K_f A^{m_e}S^{n_e}$ and $K_d \bm{\nabla}^2 h$
-consider long-range processes, described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, and short-range processes,
-proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}.
-
-In case of sticky air, slope is determined by means of the markers chain, which is vertically corrected considering erosion and sedimentation rates. After the
-correction, continental markers above the markers chain are transformed in air markers and air markers below the markers chain become sediments. Similarly,
-slope is determined considering the nodes on the top boundary in case of true free surface and they are vertically corrected as for the markers chain. In this
-case, continental markers above the top boundary are simply deleted.
-%Surface processes at the lithosphere-air interface are implemented taking into account fluvial erosion (long-range processes), hillslope processes (short-range
-%processes) and failure of slopes exceeding a threshold angle $\theta$ \citep{Burov1997,Tucker1999,Sacek2012,Croissant2014,Ueda2015,Beucher2020}. Erosion rates
-%$\dot{E}_{lr}$ due to long-range processes can be described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{lr}=K_f A^{m_e} \bm{\nabla} h^{n_e}
-%\end{equation}
-%where $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $\bm{\nabla} h$ is the slope and $m_e$ and $n_e$ are parameters chosen in to
-%constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}. Erosion rates $\dot{E}_{sr}$ due to short-range processes are
-%implemented by assuming that transport is linearly proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{sr}=K_d \bm{\nabla}^2 h
-%\end{equation}
-%where $K_d$ is the effective diffusivity. In addition, the slope is modified if it still exceeds a threshold angle after long- and short-range processes
-%\citep{Ueda2015}.
-
-%Offshore sedimentation processes are implemented similarly to hillslope processes but with a depth-dependent coefficient \citep{Sacek2012,Beucher2020}.
-%The sedimentation rate $\dot{S}$ is calculated as
-%\begin{equation}
-%\dot{S}=K_s \bm{\nabla}^2 h
-%\end{equation}
-%with
-%\begin{equation}
-%K_s=K^*_s \exp\left(\frac{h-h_{sl}}{h_0}\right)
-%\end{equation}
-%where $K^*_s$ is a constant, $h_{sl}$ is the sea level and $h_0$ is a characteristic length. In case of sticky air, slope is determined by means of the markers
-%chain, which is vertically corrected considering erosion and sedimentation rates. After the correction, continental markers above the markers chain are
-%transformed in air markers and air markers below the markers chain become sediments. Similarly, slope is determined considering the nodes on the top boundary
-%in case of true free surface and they are vertically corrected as for the markers chain. In this case, continental markers above the top boundary are simply
-%deleted.
-
-\section{Non-linear rheologies}\label{sec:nonlinear}
-Non-linear rheologies are implemented combining viscous creep (dislocation and diffusion) and plastic yielding. For each marker, diffusion and dislocation
-viscosity ($\eta_{df|ds}$) can be determined as follows \citep{Karato1993,Warren2008b,Wang2016}:
-\begin{equation}\label{eq:viscous}
-\eta_{df|ds}=f_s \left(\frac{d^m}{A}\right)^{\frac{1}{n}} \left (\frac{2^{\frac{1-n}{n}}}{3^{\frac{n+1}{2n}}} \right)
-I_2^{\frac{(1-n)}{n}}\exp\left(\frac{Q+pV}{nRT}\right)
-\end{equation}
-where $f_s$ is a scaling factor used to represent lithologies that are stronger or weaker than the base set, $d$ is the grain size, $m$ is the grain size exponent,
-$n$ is the stress exponent, $A$ is the uniaxial pre-exponential factor, $I_2$ is the square root of the second invariant of the strain rate tensor, $Q$ is the
-activation energy, $p$ is the pressure, $V$ is the activation volume, $R$ is the gas constant and $T$ is the temperature.
-Pressures and temperatures are determined by the interpolation of the nodal parameters, while strain rates are calculated by means of the
-derivative of the velocities on the nodes. In case of diffusion creep $n=1$ and $m>0$, in case of dislocation creep $n>1$ and $m=0$. The viscous creep $\eta_{cp}$
-is then calculated as the harmonic average between $\eta_{df}$ and $\eta_{ds}$:
-\begin{equation}\label{eq:visco_creep}
-\eta_{cp}=\left(\frac{1}{\eta_{df}}+\frac{1}{\eta_{ds}}\right)^{-1}
-\end{equation}
-The implementation of the non-linear viscous creep viscosity depending on the strain rate is tested by means of the slab detachment benchmark
-(Section \ref{sec:slab}).
-
-Plastic yielding is implemented rescaling $\eta_{cp}$ in order to limit the stress below the yield stress $\sigma_y$ \citep{Thieulot2008,Thieulot2014,Glerum2018},
-obtaining
-\begin{equation}\label{eq:plastic}
-\eta_{pl}=\frac{\sigma_y}{2I_2}
-\end{equation}
-where $\eta_{pl}$ is the plastic viscosity and the yield stress is determined following the Drucker-Prager criterion, such as
-\begin{equation}\label{eq:yield}
-\sigma_y=C\;\cos(\phi)+p\;\sin(\phi)
-\end{equation}
-where $C$ is the cohesion and $\phi$ is the internal friction angle. In case of negative pressure it is imposed equal to 0, so that negative yield stress are
-excluded. The correctness of the non-linear solution in case of plastic viscosity with variable internal friction angle is verified performing indenter and
-brick experiments (Sections \ref{sec:indenter} and \ref{sec:brick}, respectively).
-
-Strain softening is taken into account for both viscous creep and plastic viscosity \citep{Huismans2003,Babeyko2005,Huismans2005,Sobolev2005,Warren2008a} by
-means of the accumulated strain $\epsilon$ memorised by each marker. Softening in the viscous creep determines a linear decrease of $\eta_{vc}$ by means of a
-viscous strain softening factor $W_S$ that increases linearly from $W_{S_0}$ to $W_{S_{\infty}}$ for $\epsilon_{S_0}<\epsilon_S<\epsilon_{S_{\infty}}$ 
-\citep{Warren2008a}. This viscous softening can be related to strain-induced grain size reduction \citep{Warren2008a}. Differently, plastic softening is 
-simulated with a linear decrease of internal friction angle $\phi(\epsilon)$ and cohesion $C(\epsilon)$ in according to
-\begin{eqnarray}
-\label{eq:friction1}\phi(\epsilon)&=&\phi_0+(\phi_{\infty}-\phi_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}\\
-\label{eq:friction2}C(\epsilon)&=&C_0+(C_{\infty}-C_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}
-\end{eqnarray}
-where $\phi_0$, $C_0$ and $\phi_{\infty}$, $C_{\infty}$ are internal friction angle and cohesion for $\epsilon_0$ and $\epsilon_{\infty}$, respectively
-\citep{Huismans2003,Huismans2005,Warren2008a,Thieulot2014}. Plastic softening approximates deformation-induced softening of faults and brittle shear zones
-\citep{Warren2008a}.
-
-Plastic yielding and viscous creep are then combined to obtain a viscoplastic viscosity $\eta_{vp}$ as follows, assuming that they are independent processes
-\citep{Karato2008,Glerum2018}:
-\begin{equation}\label{eq:viscoplastic}
-\eta_{vp}=\min(\eta_{cp},\eta_{pl})
-\end{equation}
-Finally, effective viscosity $\eta_{eff}$ is capped by the minimum and the maximum viscosity ($\eta_{min}$ and $\eta_{max}$, respectively) to avoid extremely
-low or high viscosity \citep{Glerum2018} as follows
-\begin{equation}\label{eq:effective}
-\eta_{eff}=\min(\max(\eta_{vp},\eta_{min}),\eta_{max})
-\end{equation}
-Viscous creep and plastic yielding are non-linear rheologies because of their dependence on the velocity field through pressure and strain rates. Therefore,
-the solution is determined by means of Picard-type iterations, until convergence of the velocity field \citep{Glerum2018}. The convergence is verified at each
-iteration $i$ via the nonlinear residual $\bm{\mathcal{R}}^i$ that can be determined as
-\begin{equation}\label{eq:convergence}
-\bm{\mathcal{R}}^i= \mathbb{K} (\eta_{eff}(\dot{\bm{\epsilon}}^{i-1},p^{i-1}))\cdot \bm{v}^{i-1} - \bm{f}^i
-\end{equation}
-%\begin{equation}\label{eq:convergence}
-%\chi=1-\frac{\langle (\bm{v}^i-\langle\bm{v}^i\rangle)\cdot (\bm{v}^{i+1}-\langle\bm{v}^{i+1}\rangle) \rangle}{\sqrt{\sigma^i\sigma^{i+1}}}
-%\end{equation}
-%Iterations are performed until either convergence ($\chi<tol$) or maximal number of non-linear iterations ($it_{max}$) is reached.
-where $\mathbb{K}$ is the Stiffness matrix and $\bm{f}$ is the right hand side vector \citep{Spiegelman2016,Glerum2018}.  The $L_2$-norm is extracted from
-$\bm{\mathcal{R}}^i$ and it is normalised as follows
-\begin{equation}\label{eq:convergence_res}
-\frac{||\bm{\mathcal{R}}^i||_2}{||\bm{\mathcal{R}}^0||_2}
-\end{equation}
-where $||\bm{\mathcal{R}}^0||_2$ is the $L_2$-norm of the first nonlinear iteration.
-Since often the normalised nonlinear residual drop very quickly over the first 2-3 iterations, also the $L_2$-norm of velocity and pressure residuals are
-calculated as
-\begin{equation}\label{eq:convergence_vel}
-\frac{||\Delta\bm{v}||_2}{||\bm{v}||_2}=\frac{||\bm{v}_i-\bm{v}_{i-1}||_2}{||\bm{v}_{i}||_2}
-\end{equation}
-and
-\begin{equation}\label{eq:convergence_press}
-  \frac{||\Delta p||_2}{||p||_2}=\frac{||p_i-p_{i-1}||_2}{||p_{i}||_2}
-\end{equation}
-respectively. Iterations are performed until either maximal number of non-linear iterations ($it_{max}$) is reached or all the normalised nonlinear residual, 
-the velocity residual and the pressure residual converge under a defined tolerance $tol$.
-
-\section{The energy equation}\label{sec:en_eq}
-The stabilisation of the advection term of the energy equation (Eq. \ref{eq:energy}) needed to avoid possible oscillations in the thermal solution in the case
-when advection dominates over diffusion, is implemented by means of a streamline-upwind Petrov–Galerkin (SUPG) method, for which the advection term is modified
-follows the discussion in \citet{Thieulot2011} and \citet{Thieulot2014} (see Section \ref{sec:advection} for benchmark). The total internal heat production
-$H_{tot}$ in Eq. \ref{eq:energy} includes radiogenic heating $H_r$, shear heating $H_s$ and adiabatic heating $H_a$, where
-\begin{equation}\label{eq:shear}
-H_s=2\eta \dot{\bm{\epsilon}} : \dot{\bm{\epsilon}} = 2\eta(\dot{\epsilon}_{xx}^2+\dot{\epsilon}_{yy}^2+2\dot{\epsilon}_{xy}^2)
-\end{equation}
-and
-\begin{equation}\label{eq:adiabatic}
-H_a=T \alpha \frac{Dp}{Dt} \approx -\alpha T \rho g_y v
-\end{equation}
-The correctness of the implementation and the computation of shear and adiabatic heating is verified performing an exercise with an analytical solution
-(Section \ref{sec:simple_shear}) and Exercise 9.4 in \citet{Gerya2010b} (Section \ref{sec:shear}). Furthermore, mantle convection and thin layer entrainment
-benchmarks (Sections \ref{sec:mantle} and \ref{sec:thin}, respectively) are performed to verify that the code solve correctly Eqs. \ref{eq:momentum_penalty}
-and \ref{eq:energy} in case of a temperature field characterised by an initial perturbation and with a temperature-dependent density. At last, the viscoplastic
-mantle convection benchmark is performed to check also the viscoplasticity in case of a initial perturbed temperature field (Section \ref{sec:visco_mantle}).
-
-\section{Phase transitions and hydration}\label{sec:phase}
-Variations of effective density, specific heat and coefficient of thermal expansion during the evolution of crustal and mantle materials at different
-pressure-temperature ($p$-$T$) conditions are computed by means of version 6.8.6 of Perple\_X software package \citep{Connolly2005}, similarly to the
-implementation described in \citet{Marotta2020}. By default, mineral assemblages and properties of each lithology (oceanic crust, lower and upper continental
-crust, sediments and dry or hydrated mantle) are calculated for temperatures from \SIrange{330}{1600}{\kelvin}, with increments of approximately \SI{4}{\kelvin},
-and pressures from \SIrange{0.1}{30}{\giga\pascal}, with increments of approximately \SI{0.03}{\giga\pascal}, for a total of almost 100000 points.
-
-The introduction of phase transitions produces two main effects that must be taken into account in numerical models: variations in density and release/absorption
-of latent heat \citep{Tackley2010}. Effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into
-account considering effective coefficient of thermal expansion \citep{Christensen1985,Zhong2015}. Similarly, effects of latent heat in the energy equation
-(eq \ref{eq:energy}) are taken into account by considering effective specific heat and coefficient of thermal expansion
-\citep{Christensen1985,Tackley2010,Zhong2015}. Effective density, specific heat and coefficient of thermal expansion are included in Perple\_X files and are
-assumed by each Lagrangian marker in according to its $p$-$T$ conditions.
-
-Hydration processes related to dehydration of subducting lithosphere \citep{Schmidt1998,Liu2007,Faccenda2009,Faccenda2010,Faccenda2014,Rosas2016} strongly
-influence the thermo-mechanics inside the mantle wedge, mainly because of both weakening effects on the mantle rheology and density variations in case of
-mantle serpentinization \citep{Gerya2002,Honda2003,Arcay2005,Roda2010,Regorda2017}. In the case that hydration is switched on, the amount of bound and free
-water is memorised by each marker, following the implementation of \citet{Quinquis2014}; hydration and dehydration processes are related to the amount of bound
-water of each marker and to the maximum amount of water it can transport, i.e. if the amount of bound water exceeds the maximum amount of water, the marker
-dehydrates and releases free water that can hydrates under-saturated markers. Maximum water content of each marker is determined as function of lithology and
-$p$-$T$ conditions and it is calculated using Perple\_X, in the same way than effective density, specific heat and coefficient of thermal expansion. Bound water
-is advected together with the markers, neglecting the effect of bound water diffusion, while free water simply migrates vertically and is not coupled to the
-solid-phase flow of the mantle wedge \citep{Arcay2005,Quinquis2014}. The correctness of the migration scheme of the free water and its absorption into bound
-water by under-saturated markers are verified performing the experiment described in \citet{Quinquis2014} (Sec. \ref{sec:quinquis}).
-
-The mantle viscosity weakening related to the amount of water has been extensively studied \citep{Chopra1981,Hirth2003} and it is implemented in according to
-\citet{Arcay2005} and \citet{Horiuchi2016}, as follows
-\begin{equation}\label{eq:wet_visc}
-\eta_{wet}=\eta_{dry}\left(\left[1-\frac{1}{f_v}\right]\exp\left(-\frac{[OH^-]}{[OH^-]_0}\right)+\frac{1}{f_v}\right)
-\end{equation}
-where $f_v$ is the viscosity reduction factor between dry and wet conditions, $[OH^-]$ is the water content and $[OH^-]_0$ is a reference water content set to
-620 ppm (0.062 wt.\%) \citep{Arcay2005}. By default, $f_v$ is set to 100, in according to \citet{Arcay2005} and \citet{Horiuchi2016}.
-
-\section{Mantle melting}\label{sec:melt}
-Melt fraction in the mantle depends on temperature ($T$, in \textdegree C), pressure ($p$, in \si{\giga\pascal}) and the water content in the melt ($X_{H_2O}$, in wt.\%).
-The determination of the melt fraction $M$ is implemented as explained by \citet{Katz2003}, according to the successive modification by \citet{Langmuir2006} and
-\citet{Kelley2010}, as follows
-\begin{eqnarray}\label{eq:melt_fr}
-M(p,T,X_{H_2O})&=&-T+T_{sol}+(x_t\cdot ln(p+y_t))\cdot X_{H_2O}-\nonumber\\
-&&+K_m\left(\frac{X_{H_2O}^{bulk}}{D_{H_2O}(1-X_{H_2O})+X_{H_2O}}\right)^{\gamma}
-\end{eqnarray}
-where $T_{sol}$ is the temperature of the dry solidus, $D_{H_2O}$ is the partition coefficient and $x_t\cdot log(p+y_t)$ indicates the pressure dependence of
-the melting curve \citep{Kelley2010}. $x_t$, $y_t$, $K_m$ and $\gamma$ are parameters chosen in according to \citet{Kelley2010}.
-However, $X_{H_2O}$ depends on the melt fraction as
-\begin{equation}\label{eq:melt_wt}
-X_{H_2O}(M)=\frac{X_{H_2O}^{bulk}}{D_{H_2O}+M(1-D_{H_2O})}
-\end{equation}
-and a numerical solution can be found using a root-finder method \citep{Katz2003,Wang2016}. Furthermore, the content of water in the melt is limited by the
-pressure-dependent saturation concentration of water in the melt, determined as
-\begin{equation}\label{eq:melt_max}
-X_{H_2O}^{sat}=\chi_1 p^{\lambda_m}+\chi_2 p
-\end{equation}
-where $\chi_1$, $\chi_2$ and $\lambda_m$ are parameters chosen according to \citet{Katz2003}, \citet{Langmuir2006} and \citet{Kelley2010}. The correctness in
-the determination of melt fractions and water contents in the melt are tested comparing model results with experimental curves obtained by \citet{Katz2003}
-(Sec. \ref{sec:katz}).
-
-In case of melting, viscous creep viscosity ($\eta_{vp}$) is modified according to \citet{Wang2016}, as
-\begin{equation}\label{eq:melt_visc}
-\eta_{melt}=\eta_{vp}\cdot \exp(\alpha_m M)
-\end{equation}
-where $\alpha_m$ are melt fraction factors for dislocation and diffusion creep, chosen in according to \citet{Wang2016}.
-
-Effective density ($\rho_{eff}$) for partially molten rocks is calculated as
-\begin{equation}\label{eq:melt_dens}
-\rho_{eff}=\rho_s(1-M)+\rho_m M
-\end{equation}
-where $\rho_s$ and $\rho_m$ are the densities of the solid and the molten rock, respectively \citep{Gerya2004b,Gerya2007,Wang2016}. The density of the
-solid rock is extracted by Perple\_X, while the density of the molten portion is calculated as
-\begin{equation}\label{eq:melt_dens_M}
-\rho_m(p,T)=\rho_0[1-\alpha(T-T_0)][1+\beta(p-p_0)]
-\end{equation}
-where $\rho_0$ is the density at the reference temperature ($T_0$) and pressure($p_0$), and $\alpha$ and $\beta$ are the thermal expansion and compressibility
-coefficients, respectively \citep{Gerya2004b,Gerya2007,Wang2016}.
-
-As for the phase transitions, effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into account
-considering effective coefficient of thermal expansion. Similarly, latent heat due to melting/crystallisation can be implicitly included in the energy equation
-(Eq. \ref{eq:energy}) by considering effective specific heat ($C_{p_{eff}}$) and thermal expansion ($\alpha_{eff}$) of the partially molten rocks, in according to
-\begin{eqnarray}
-\label{eq:cp_effective}C_{p_{eff}}&=&C_p+H_L\left(\frac{\partial M}{\partial T}\right)_p \\
-\label{eq:alpha_effective}\alpha_{eff}&=&\alpha+\rho\frac{H_L}{T}\left(\frac{\partial M}{\partial p}\right)_T
-\end{eqnarray}
-where $C_p$ and $\alpha$ are the specific heat and the coefficient of thermal expansion of the solid rock, respectively, and $H_L$ is the latent heat of the
-molten rock \citep{Gerya2004b,Gerya2007,Tackley2010}.
-
-\section{Benchmarks}\label{sec:benchmark}
-\subsection{Error measurements}\label{sec:error}
-In order to determine the accuracy of velocity and pressure field of the benchmarks, the $L_2$-norm is computed by numerical integration on the Gauss-Legendre
-quadrature points. $L_2$-norm for pressure and velocity errors can be evaluated as
-\begin{eqnarray}
-\textrm{err}_p&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|p^n_i(r_q)-p^a_i(r_q)|^2\right)w_q|J_q|}\nonumber \\
-\textrm{err}_v&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|u^n_i(r_q)-u^a_i(r_q)|^2+|v^n_i(r_q)-v^a_i(r_q)|^2\right)w_q|J_q|}\nonumber
-\end{eqnarray}
-respectively, where $n_e$ is the number of elements, $n_q$ is the number of quadrature points per element, $p^n_i(r_q)$ and $p^a_i(r_q)$ are the numerical and
-analytical pressure, respectively, in each quadrature point $q$, $w_q$  and $J_q$ are the weight and the Jacobian at the quadrature point
-$q$, $u^n_i(r_q)$, $v^n_i(r_q)$, $u^a_i(r_q)$ and $v^a_i(r_q)$ are the numerical and analytical velocities, respectively, in each quadrature point $q$.
-
-Other quantities used as comparison with original benchmarks are the root-mean-square velocity over the whole domain and over the surface
-\begin{eqnarray}
-v_{\textrm{rms}}&=&\sqrt{\int_{\Omega}|\bm{v}|^2 d\Omega}\nonumber \\
-v_{\textrm{rms}}^{top}&=&\sqrt{\int_0^1 u^2 \bigg|_{y=1} dx}\nonumber
-\end{eqnarray}
-respectively, top and bottom Nusselt numbers
-\[\textrm{Nu}^{top/bottom}=-\int_0^1\frac{\partial T}{\partial y} \bigg|_{y=1/y=0} dx\]
-the average rate of work done against gravity
-\[\left\langle W \right\rangle=\int_{\Omega}T u_y d\Omega\]
-the average rate of viscous dissipation
-\[\left\langle \Phi \right\rangle=\int_{\Omega}\tau_{ij}\dot{\epsilon}_{ij} d\Omega\]
-and the percentage error between $\left\langle W \right\rangle$ and $\left\langle \Phi \right\rangle/Ra$
-\[\delta=\frac{\left|\left\langle W \right\rangle -\frac{\left\langle \Phi \right\rangle}{Ra} \right|}{\max \left(\left\langle W \right\rangle,
-\frac{\left\langle \Phi \right\rangle}{Ra}\right)}\times100\]
-
-\subsection{Stokes flow}\label{sec:stokes}
-The problem consists of determining velocity field $(u,v)$ and pressure $p$ in case of a manufactured solution with prescribed body forces such as
-\begin{eqnarray}
-b_1&=&(12-24y)x^4+(-24+48y)x^3+(-48y+72y^2-48y^3+12)x^2+\nonumber \\
-&&+(-2+24y-72y^2+48y^3)x+1-4y+12y^2-8y^3\nonumber \\
-b_2&=&(8-48y+48y^2)x^3+(-12+72y-72y^2)x^2+(4-24y+48y^2-48y^3+\nonumber \\
-&&+24y^4)x-12y^2+24y^3-12y^4\nonumber
-\end{eqnarray}
-for which the exact solution is:
-\begin{eqnarray}
-u(x,y)&=&x^2(1-x)^2(2y-6y^2+4y^3)\nonumber \\
-v(x,y)&=&-y^2(1-y)^2(2x-6x^2+4x^3)\nonumber \\
-p(x,y)&=&x(1-x)-\frac{1}{6}\nonumber
-\end{eqnarray}
-The domain is a unit square a constant viscosity ($\eta=1$) and the penalty parameter is set to $\lambda=10^7$. Velocity boundary conditions are set to no slip
-$(\bm{v}=\bm{0})$ on all boundaries. The problem is performed for different grid resolution between \num{8x8} and \num{1024x1024} elements.
-
-The errors between the analytical solution and the numerical prediction of the velocity and the pressure field are calculated as in Section \ref{sec:error}.
-Fig. \ref{fig:errors}a shows that both velocity and pressure field converge to the exact solution with the decrease of the element size, following the
-theoretical convergence rate. The convergence of pressure, in contrast with the observation by \citet{Donea2003} for $Q_1 \times P_0$ elements, indicate the
-effectiveness of the smoothing procedure. Smoothed pressure field for grid resolution of \num{128x128} elements is shown in Fig. \ref{fig:errors}b (continuous
-red line), in comparison with the analytical solution (dashed blue line). Solve times and memory usage needed to generate the solution are shown in
-Fig. \ref{fig:MUMPS}. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Solver/Stokes_Flow}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/errors.pdf}
-\caption{Velocity and pressure error for the Stokes flow experiment between generated and analytical solution as a function of element size (panel a),
-and comparison between smoothed pressure and analytical solution as function of $x$ coordinate for a grid resolution of \num{128x128} elements (panel b).}
-\label{fig:errors}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/MUMPS.pdf}
-\caption{Panel a) Analysis, factorisation, solution and total solve times (panel a) and factorisation memory usage (panel b) as a function of the total number
-of degrees of freedom for the Stokes flow experiment.}
-\label{fig:MUMPS}
-\end{figure}
-
-\subsection{Markers advection}\label{sec:runge}
-The 2nd-order Runge-Kutta advection scheme in space is tested by means of the Zalesak disk test \citep{Zalesak1979,Thieulot2014}. The benchmark is performed
-in a unit square domain with a grid resolution of \num{32x32} elements and 1000000 markers and values of Courant number between 0.25 and 1.
-At $t=0$, the disk is centred at position $(0.5;0.75)$ with a radius $R=0.15$ and has a vertical fissure 0.05 wide and 0.2 high (orange in Fig. \ref{fig:runge})
-and with the velocity field prescribed in the entire domain as
-\begin{eqnarray}
-u(x,y)&=&2\pi\left(y-\frac{L_x}{2}\right)\nonumber \\
-v(x,y)&=&-2\pi\left(x-\frac{L_x}{2}\right)\nonumber
-\end{eqnarray}
-the markers are back to their initial location after a $2\pi$ rotation ($t=1$). The clockwise rotation of the disk at $t=0.25$, 0.5 and 0.75 
-is shown in Fig. \ref{fig:runge} (red, blue and green, respectively). The distance from the centre of 1 marker is calculated throughout the rotation to
-evaluate the error for different values of Courant number. As expected, the error increase with the increasing of the time step, as shown in 
-Fig. \ref{fig:runge_err}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/Zalesak_Disk}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=200px]{./Figures/Zalesak.pdf}
-\caption{Position of the Zalesak disk at $t=0$, 0.25, 0.5 and 0.75 (orange, red, blue and green, respectively) for the case with Courant number of 1.}
-\label{fig:runge}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/zal_marker.pdf}
-\caption{Distance from the centre as function of time for values of Courant number of  of 0.25, 0.3, 0.5, 0.75 and 1 (green, orange, blue, black and red,
-respectively). R indicates the distance from the centre considering a perfect circle trajectory.}
-\label{fig:runge_err}
-\end{figure}
-
-\subsection{Conservative Velocity Interpolation (CVI)}\label{sec:CVI}
-The CVI correction is checked by means of the Stokes flow experiment by \citet{Donea2003}, as presented in \ref{sec:stokes}. The advection of Lagrangian
-markers is performed using either a 2nd-order or a 4th-order Runge-Kutta scheme with CFL=0.5 and an initial random distribution of 25 markers per element.
-Fig. \ref{fig:CVI} shows the minimum and maximum values of markers per element throughout the experiment with and without the CVI correction and using
-different Runge-Kutta schemes. The introduction of the CVI correction allows the
-tests to run for times longer than $t=2000$ without show empty elements, while tests without the CVI correction stop at approximately $t=600$, when at least
-one element goes empty. Fig. \ref{fig:CVI_time} shows the different number of markers per element and the markers distribution at $t=600$ obtained without and
-with the CVI correction (panels b and c, respectively) in case of a 2nd-order Runge-Kutta scheme. All data can be found at
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/CVI}.
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/CVI.pdf}
-\caption{Maximum and minimum number of markers per element throughout the experiment, with (green and orange lines) or without (purple and blue lines)
-the CVI correction and using either a 2nd-order (blue and orange lines) or a 4th-order (purple and green lines) Runge-Kutta scheme.}
-\label{fig:CVI}
-\end{figure}
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/CVI_time.pdf}
-\caption{Number of markers per element and markers distribution through time using 2nd.order Runge-Kutta scheme.
-Panels a: number of markers per element (panel a1) and distribution of the markers (panel a2) after the first
-iteration; panels b and c: comparison between number of markers per element (panels b1 and c1) and markers distribution (panels b2 and c2) without and with
-the CVI correction at $t=600$; panels d: number of markers per element (panel d1) and markers distribution (panel d2) in case of CVI correction at $t=2000$.}
-\label{fig:CVI_time}
-\end{figure}
-
-\subsection{Poiseuille flow}\label{sec:poiseuille}
-The domain is a rectangle with $L_x=2$ and $L_y=1$ and constant density and viscosity ($\rho=1$ and $\eta=1$), gravity acceleration $\bm{g}=\bm{0}$ and penalty
-parameter $\lambda=10^8$. The grid is composed by \num{40x20} elements. Velocity boundary conditions are set to no slip $(\bm{v}=\bm{0})$ at the top and the
-bottom, and a parabolic profile is imposed on the sides, with $u=y(1-y)$ and $v=0$. The analytical solution is then given by:
-\begin{eqnarray}
-u(x,y)&=&y(1-y)\nonumber \\
-v(x,y,)&=&0\nonumber \\
-p(x,y)&=&2\eta\left(\frac{L_x}{2}-x\right)\nonumber
-\end{eqnarray}
-
-The velocity field predicted by the model follows the expected parabolic profile (Fig. \ref{fig:poiseuille}a and Fig. \ref{fig:poi_plot}a) and, in case of the
-classic penalty method with no iterations, the pressure is clearly related to the divergence by means of the penalty parameter (Fig. \ref{fig:poiseuille}b and
-c and Fig. \ref{fig:poi_plot}b). Fig. \ref{fig:poiseuille}d shows that one Uzawa iteration is sufficient to bring the divergence down to \num{e-15}, with no
-correlation with the pressure field. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Poiseuille%20Flow}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Poiseuille.pdf}
-\caption{Velocity field (panel a), pressure (panel b) and divergence velocity of a Poiseuille flow in case of the classic penalty method (no iterations) and
-after one Uzawa iteration (panel c and d, respectively).}
-\label{fig:poiseuille}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Analytical.pdf}
-\caption{Velocity (panel a) and pressure (panel b) field predicted by the model for a Poiseuille flow with respect with their analytical solutions. The
-velocity field is plotted as function of the vertical coordinate in $L_x/2$ and the pressure is plotted as function of the x coordinate in $L_y/2$.}
-\label{fig:poi_plot}
-\end{figure}
-
-\subsection{Instantaneous 2D sphere}\label{sec:ist_sphere}
-The domain is a unit square with gravity $\bm{g}=(0,-1)$. The fluid has constant density and viscosity ($\rho_f=1$ and $\eta_f=1$). The sphere is in the middle
-of the domain with a radius R = 0.123456798 and has constant density and viscosity ($\rho_s=10^{-2}+\rho_f$ and $\eta_s=10^3 \cdot \eta_f$).
-We distinguish three different types of velocity boundary conditions:
-\begin{enumerate}
-\item FS: free slip conditions on all sides.
-\item NS: no slip conditions on all sides.
-\item OT: free slip conditions on the sides and the bottom, and open boundary at the top.
-\end{enumerate}
-All three cases are performed for different grid resolutions, between \num{16x16} and \num{512x512} elements with 50 randomly distributed markers per element,
-and different average schemes for the viscosity (harmonic, geometric and arithmetic). The velocity in the centre of the sphere $\bm{v}(0.5;0.5)$), minimum and
-maximum velocities ($u_{min}$, $u_{max}$, $v_{min}$, $v_{max}$) and pressures ($p_{min}$, $p_{max}$) on the entire domain, $v_{\textrm{rms}}$ and average pressure on the
-entire domain ($p_{avg}$) are compared with solutions generated by ASPECT \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and all results
-can be found at \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere2D} and 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Instantaneous_2D_Sphere}. Results in terms of the $v_{\textrm{rms}}$ for NS model 
-are also shown in Fig. \ref{fig:inst_sphere}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_NS.pdf}
-\caption{$v_{\textrm{rms}}$ as function of element size for the instantaneous 2D sphere experiment in case of no slip boundary conditions and with different average
-schemes for the viscosity. Results are compared with results obtained by other numerical codes.}
-\label{fig:inst_sphere}
-\end{figure}
-
-\subsection{Rayleigh-Taylor instability}\label{sec:rayleigh}
-This problem has been originally presented by \citet{vanKeken1997} and here it is performed as the isoviscous case (Case 1a). The domain has $L_x=0.9142$ and
-$L_y=1$ and gravity $\bm{g}=(0,-1)$. Two fluids with same constant viscosities ($\eta_1=\eta_2=1$) and different densities ($\rho_1=1000$ and $\rho_2=1010$),
-with the lighter fluid at the bottom. The initial interface between the fluids is given by $y(x)=0.2+0.02 \cos \left(\frac{\pi x}{L_x}\right)$. The experiment
-is performed with different grid sizes (\num{50x50}, \num{80x80}, \num{100x100} and \num{256x256}). A total of 1960000 markers are randomly distributed at the
-beginning of the simulation. Velocity boundary conditions are set to no slip at the top and the bottom, and to free slip at the sides of the domain.
-
-$v_{\textrm{rms}}$ as function of time is reported in Fig. \ref{fig:RT}, matching well with results shown by \citet{vanKeken1997}, \citet{Tackley2003} and
-\citet{Thieulot2014}. Fig. \ref{fig:rayleigh} shows the evolution of the experiment at different time steps. All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Rayleigh_Taylor_experiment/ISOVISCOUS}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/RT.pdf}
-\caption{$v_{\textrm{rms}}$ of the Rayleigh-Taylor experiment as function of time for different resolution of the grid.}
-\label{fig:RT}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=300px]{./Figures/Rayleigh.pdf}
-\caption{Evolution of the Rayleigh-Taylor experiment for a grid of \num{256x256} elements at $t=0, 500, 1000$ and $2000$ (panels a, b, c and d, respectively).}
-\label{fig:rayleigh}
-\end{figure}
-
-\subsection{Falling blocks}\label{sec:block}
-This benchmark is described as proposed by \citet{Gerya2003a}, \citet{Gerya2010b} and \citet{Thieulot2011}. The domain is square with $L_x=L_y=$ \SI{500}{\km}
-and the grid is composed by \num{50x50} elements with 25 markers in each element. The block is initially centred at ($x=$ \SI{250}{\km}; $y=$ \SI{400}{\km})
-and has a size of \num{100x100} km (Fig. \ref{fig:block}a). The fluid surrounding the block has $\eta_f=$ \SI{e21}{\pascal\per\s} and $\rho_f=$
-\SI{3200}{\kg\per\cubic\m}. The benchmark tests with different viscosities of the block, with $\eta_b$ from \SIrange{e15}{e27}{\pascal\per\s}. In each
-experiment also the density $\rho_b$ of the block varies from \SIrange{3220}{9900}{\kg\per\cubic\m}. Velocity boundary conditions are set to free slip
-conditions on all sides of the domain. Density distributions at $t=$ \SI{20}{\mega\year} for $\rho_b=$ \SI{3300}{\kg\per\cubic\m} and different $\eta_b$
-are plotted in Fig. \ref{fig:block}b-f, showing that the code correctly preserve the block geometry in case of large viscosity contrast, with a block stiffer
-than the surrounding fluid (Fig. \ref{fig:block}f).
-
-The velocity in the centre of the falling block at $t=0$ is measured for all experiments and, since the velocity should increase with density contrast, the
-quantity $\bm{v}/(\rho_b-\rho_f)$ is plotted as function of the viscosity contrast. All results perfectly match with those from \citet{Gerya2010b} and they
-line up on a single curve, demonstrating that the code can correctly deal with large viscosity and density contrasts (Fig. \ref{fig:falling}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Falling%20blocks}.
-
-\begin{figure}
-\noindent\includegraphics[width=300px]{./Figures/Block.pdf}
-\centering
-\caption{Density evolution of the falling block experiment at $t=$ \SI{0}{\mega\year} (panel a) and $t=$ \SI{20}{\mega\year} for different viscosities of the
-block (panels b-f).}
-\label{fig:block}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Falling.pdf}
-\caption{Initial velocity relative to the density contrast at the centre of the falling block as function of the viscosity contrast between the block and the
-surrounding fluid.}
-\label{fig:falling}
-\end{figure}
-
-\subsection{2D time-dependent Stokes sphere with deformable free surface}\label{sec:sphere}
-This experiment is performed in a unit square domain with the gravity acceleration fixed to $g_y=-1$. The sphere has $\rho_s=2$ and $\eta_s=10^3$ and is
-initially centred at ($0.5;0.6$) with radius $R=0.123456789$. The fluid surrounding the sphere has $\rho_f=1$ and $\eta_f=1$ and occupies the domain for
-$y\leq 0.75$, while for $y>0.75$ the air has $\rho_a=0$ and $\eta_a=10^{-3}$. Velocity boundary conditions are set to free slip on all sides. The Courant
-number is set to 0.25. The experiments run for 200 s using grid resolutions from \num{150x150} to \num{512x512} elements, with 25 randomly distributed markers
-per element, and for different average schemes for the viscosity. The interface between the fluid and the air is tracked by means of the markers chain. Results
-in terms of velocity and pressure fields and topography variations are compared with results obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and they can be found at
-\url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere_fs2D} and
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Time_dependent_sphere}. Results in terms of the $v_{\textrm{rms}}$ in case of 
-an arithmetic average are also shown in Fig. \ref{fig:inst_spherefs}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_arithm.pdf}
-\caption{$v_{\textrm{rms}}$ for different grid resolutions as function of time for the 2D time-dependent Stokes sphere experiment in case of an arithmetic average.
-Results are compared with results obtained by ASPECT.}
-\label{fig:inst_spherefs}
-\end{figure}
-
-\subsection{Free surface stabilisation}\label{sec:stab}
-The benchmark is performed as discussed in \citet{Kaus2010a} and \citet{Thieulot2014}. The domain is square with $L_x=L_y=$ \SI{500}{\km} and a grid resolution
-of \num{200x200} elements, each of them containing 16 markers. A buoyant fluid with $\rho_1=$ \SI{3200}{\kg\per\cubic\metre} and $\eta_1=$ \SI{e20}{\pascal\s}
-is overlain by a denser fluid with $\rho_2=$ \SI{3300}{\kg\per\cubic\metre} and $\eta_2=$ \SI{e21}{\pascal\s}. The initial interface between the fluids has an
-initial sinusoidal shape of 5 km amplitude. Velocity boundary conditions are set to free slip conditions at the sides and no slip at the bottom, while the top
-is free surface. The experiment is performed with various fixed time steps, without the use of the Courant number. The vertical position of the free surface at
-$x=L_x$ is tracked for each simulation, with or without the stabilisation algorithm.
-
-The results show an instability (drunken sailor effect) increasing the time step in case that the stabilisation algorithm is not activated
-($dt=$ \SI{4200}{\year}, green line in Fig. \ref{fig:stab}), as already observed by \citet{Kaus2010a} and \citet{Thieulot2014}. Activating the algorithm the
-instability is fixed and the simulation remains stable using time step up to \SI{50000}{\year} (red line in Fig. \ref{fig:stab}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Stabilisation_algorithm}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Kaus_random.pdf}
-\caption{Evolution of the y coordinate in $x=L_x$ as function of time for different time steps, with and without the stabilisation algorithm.}
-\label{fig:stab}
-\end{figure}
-
-\subsection{Topography relaxation}\label{sec:crameri}
-This experiment is performed as in \citet{Crameri2012} using free surface and sticky air (Case 1). The domain is rectangular with $L_x=$ \SI{2800}{\km} and
-$L_y=$ \SI{700}{\km} in case of free surface and $L_y=$ \SI{800}{\km} in case of sticky air ($h_{st}=$ \SI{100}{\km}). The grid is composed by \num{256x64}
-elements for the free surface \num{560x320} elements for the sticky air. For both cases, there is a mantle of 600 km thickness with $\eta_m=$ \SI{e21}{\pascal\s}
-overlain by a highly viscous cosine-shaped lithosphere with thickness between 93 and 107 km and $\eta_l=$ \SI{e23}{\pascal\s}. Both mediums have $\rho=$
-\SI{3300}{\kg\per\cubic\metre}. In case of sticky air experiment, the air has $\rho_a=$ \SI{0}{\kg\per\cubic\metre} and $\eta_a=$ \SI{e18}{\pascal\s}. The
-gravity acceleration is set to $g_y=$ \SI{-10}{\metre\per\square\s}. Velocity boundary conditions are set to no slip at the bottom and free slip at the sides
-of the domain. In case of sticky air, velocities on top are set to free slip conditions.
-
-The maximum topography as function of time ($h(t)$) can be analytically derived in according to \[h(t)=h_0 \exp(\gamma t)\]
-where $h_0=7$ km is the initial topography, $\gamma=\num{-0.2139e-11}$ is the characteristic relaxation rate and $t$ is the time (black line in Fig.
-\ref{fig:crameri}). Maximum topography at the characteristic relaxation time $t_{rlx}=$ \SI{14.825}{\kilo\year} can be found to be $h_{rlx}=$ \SI{2576}{\m}.
-The results show that a viscosity of the air of \SI{e18}{\pascal\s} is needed in case of $h_{st}=$ \SI{100}{\km} to correctly describe the topography relaxation
-for this problem, as pointed out by \citet{Crameri2012} (purple line in Fig. \ref{fig:crameri}). The case with a true free surface well follows the analytical
-solution (orange line in Fig. \ref{fig:crameri}). In particular, at the relaxation time in case of the true free surface predicts a maximum topography of
-\SI{2570}{\m}, with an error of \SI{6}{\m} with respect to the analytical solution. Results are also compared with results obtained with a free surface in
-UNDERWORLD (\num{256x64} elements) and SULEC (\num{401x201} elements) (with courtesy from \citealp{Crameri2012}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Topography_relaxation}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Crameri.pdf}
-\caption{Maximum topography as function of time for the Crameri benchmark (coloured lines) in comparison with the analytical solution (black line) and results
-obtained with UNDERWORLD (yellow line) and SULEC (blue line).}
-\label{fig:crameri}
-\end{figure}
-
-\subsection{Spontaneous subduction}\label{sec:subduction}
-This experiment is performed as presented by \citet{Schmeling2008}, in case of both a sticky air and a true free surface and using different average schemes
-for the viscosities. The domain is rectangular with $L_x=$ \SI{3000}{\km} and $L_y=$ \SI{700}{\km}, a grid resolution of \num{256x64} elements with 50 markers
-per element and Courant number of 0.01. At the beginning of the simulation, a 100 km-thick lithospheric layer with $\rho_m=$ \SI{3300}{\kg\per\cubic\metre}
-and $\eta_m=$ \SI{e23}{\pascal\s} is located at the top of the domain for $x$ between 1000 and 3000 km, while the asthenospheric mantle has $\rho_m=$
-\SI{3200}{\pascal\s} and $\eta_m=$ \SI{e21}{\pascal\s}. In addition, a 200 km-depth lithospheric slab is already subducted in the mantle in order to have
-a spontaneous subduction. In case of sticky air, $L_y=750$ km and a 50 km-thick air layer with $\eta_a$ from \SIrange{e19}{e21}{\pascal\s} overlie the mantle.
-Velocity boundary conditions are set to free slip on the sides and on the bottom of the domain. In case of sticky air, velocities on the top boundary are set
-to free slip conditions as well.
-
-The maximum depth of the slab is tracked for all the simulations (Fig. \ref{fig:subduction}) and they are compared with results obtained by
-\citet{Schmeling2008} with models with regular grids and comparable grid resolutions (rectangles in Fig. \ref{fig:subduction}). Low viscous sticky air models
-(i.e., \SI{e19}{\pascal\s}) enlighten a strong dependence of the sinking velocity with the chosen average scheme (dashed lines in Fig. \ref{fig:subduction}),
-while high viscous sticky air models (i.e., \SI{e21}{\pascal\s}) show a higher resistance at the trench, underestimating the correct solution (dotted lines in
-Fig. \ref{fig:subduction}). A high resistance at the trench can be also observed in case of a true free surface for the low resolution grid used (continuous
-lines in Fig. \ref{fig:subduction}), as pointed out by \citet{Schmeling2008}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Spontaneous_subduction}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Subduction.pdf}
-\caption{Maximum depth of the slab as function of time for all the simulations of the \citet{Schmeling2008} benchmark, in case of harmonic, geometric and
-arithmetic means (red, green and blue lines, respectively) using a sticky air or a true free surface (discontinuous and continuous lines, respectively).
-Red, green and blue rectangular areas indicate the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{500}{\km} in case of sticky air,
-while the grey rectangular area indicates the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{600}{\km} in case of true free surface.}
-\label{fig:subduction}
-\end{figure}
-
-\subsection{The slab detachment}\label{sec:slab}
-The slab detachment benchmark is performed as described by \citet{Schmalholz2011} and \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{1000}{\km}
-and $L_y=$ \SI{660}{\km} and a grid resolution of \num{256x256} elements. A non-linear viscous T-shaped layer with $\rho_l=$ \SI{3300}{\kg\per\cubic\metre} is
-placed at the top of the domain and surrounded by a linear viscous fluid with $\rho_f=$ \SI{3150}{\kg\per\cubic\m}. The top layer is 80 km-thick and a 250
-km-long and 80 km-wide slab is placed at $x=L_x/2$ (Fig. \ref{fig:slab}a). The effective viscosity of the top layer is given by
-\[\eta_{eff}=\eta_0 I_2^{\frac{(1-n)}{n}}\]
-with $\eta_0=$ \SI{4.75e11}{\pascal\s} and $n=4$, while the surrounding fluid has $\eta_f=$ \SI{e21}{\pascal\s}. The effective viscosity is capped between
-$\eta_{min}=$ \SI{e21}{\pascal\s} and $\eta_{max}=$ \SI{e25}{\pascal\s}. Velocity boundary conditions are set to free slip at the top and the bottom, and to
-no slip at the sides of the domain.
-
-The sinking of the slab determines a decrease of the effective viscosity in the T-shaped layer during the first part of the evolution (Fig. \ref{fig:slab}a),
-as consequence of local strain rates, while effective viscosities increase up to $\eta_{max}$ in the top layer after approximately \SI{20}{\mega\year}, when
-necking is complete (Fig. \ref{fig:slab}b).
-
-The width of necking is tracked by means of markers position at the side of the slab. Necking width and time are normalised in function of the initial slab
-(\SI{80}{\km}) and the characteristic time ($t_c=$ \SI{7.1158e14}{\s}) \citep{Schmalholz2011,Glerum2018}. The results in case of arithmetic, geometric and
-harmonic mean of the viscosity are shown in Fig. \ref{fig:necking}, compared with results from \citet{Glerum2018} for which an infinity norm mean was used.
-The influence of the average schemes agrees with results shown in Section \ref{sec:subduction}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Slab_detachment}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Slab.pdf}
-\caption{Effective viscosity for the slab detachment benchmark at the beginning of the evolution (panel a) and when necking is complete (panel b).}
-\label{fig:slab}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Necking.pdf}
-\caption{Normalised width of the necking of the slab detachment benchmark as function of normalised time.}
-\label{fig:necking}
-\end{figure}
-
-\subsection{The indenter experiment}\label{sec:indenter}
-The indenter benchmark simulate a rigid punch on a purely plastic von Mises material, for which exists an analytical solution
-\citep{Thieulot2008,Thieulot2014,Glerum2018}. The domain is a unit square with a grid resolution of \num{256x256} elements. The gravity acceleration is set
-to $\bm{g}=\bm{0}$ and effective viscosity is capped using $\eta_{min}=10^{-4}$ and $\eta_{max}=10^3$. The tolerance for the convergence of the non-linear
-solution is $tol=10^{-9}$, with a maximum number of non-linear iterations set to 500. The medium in the domain has $\rho=0.01$, an initial viscosity
-$\eta_0=10$, cohesion $C=1$ and an angle of internal friction $\phi=0$°. Velocity boundary conditions are set to no slip at the bottom and free slip at the
-sides of the domain. The top of the domain is open with the exception of the central portion (punch area) with width $w_p=0.125$, where $v=-1.05$ and $u$ is
-fixed either to 0 (rough punch experiment) or free (smooth punch experiment).
-
-Fig. \ref{fig:indenter} show the results in terms of viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure
-(panels g and h) for both experiments. The analytical solution indicate that pressure at ($0.5;1$) and ($0.5 \pm w_p;1$)is $p=\pi +1$ and $p=1$, respectively
-(black and red lines in Fig. \ref{fig:smooth_p}). Fig. \ref{fig:smooth_p} shows a clear improvement in the pressure solution when passing from a rough to a
-smooth punch (green and blue lines in Fig. \ref{fig:smooth_p}, respectively), as previously observed by \citet{Thieulot2014} and \citet{Glerum2018}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Indenter}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Indenter.pdf}
-\caption{Viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure (panels g and h) fields for rough (left column) and
-smooth (right column) punch experiments.}
-\label{fig:indenter}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Smooth_Rough.pdf}
-\caption{Nodal pressure in function of the x coordinate at $y=1$ for rough (green line) and smooth (blue line) punch experiment. Black and red lines indicate
-the analytical solution at $x=0.5$ and $x=0.5 \pm w_p$, respectively.}
-\label{fig:smooth_p}
-\end{figure}
-
-\subsection{The brick experiment}\label{sec:brick}
-An instantaneous version of the brick benchmark is performed to verify the correctness of the pressure-dependent plasticity for different angle of internal
-friction, as proposed by \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{40}{\km} and $L_y=$ \SI{10}{\km} and a grid resolution of \num{512x128}
-elements. The gravity acceleration is set to $g_y=$ \SI{-10}{\m\square\s} and effective viscosity is capped using $\eta_{min}=$ \SI{e19}{\pascal\s} and
-$\eta_{max}=$ \SI{e26}{\pascal\s}. The tolerance for the convergence of the non-linear solution is $tol=10^{-7}$, with a maximum number of non-linear iterations
-set to 1000. A 800 m-wide and 400 m-height inclusion with $\eta_b=$ \SI{e20}{\pascal\s} and $\rho_b=$ \SI{2700}{\kg\per\cubic\m} is placed at the bottom of the
-domain at $x=L_x/2$. The inclusion is surrounded by a non-linear viscous medium with $\rho_m=$ \SI{2700}{\kg\per\cubic\m}, an initial viscosity of
-\SI{e23}{\pascal\s}, a linear viscous viscosity of \SI{e25}{\pascal\s} and a cohesion of \SI{40}{\mega\pascal}. Velocities are set to free slip conditions at
-the bottom of the domain and the top is open. Velocities on sides of the domain are fixed to $u=$ \SI{+-2e-11}{\m\per\s} and $v=0$. The experiment is performed
-in compressional and extensional contexts with an angle of internal friction $\phi$ of the non-linear medium variable between 0° and 30°.
-
-As expected, two shear bands stem from the inclusion with variable angles in relation to both the dynamics context and the internal friction angle. Shear band
-angles formed at 45° for $\phi=0$° in both extensional and compressional contexts (Fig. \ref{fig:brick_beam}a), while they are different in case of $\phi\neq0$°.
-Shear band angles in case of $\phi=20$° are shown for compression and extension in Fig. \ref{fig:brick_beam}b  and c, respectively. Values of shear band angles
-as function of internal friction angles from 0° to 30° are extracted at different depths and minimum and maximum values are plotted in Fig. \ref{fig:brick},
-in comparison with theoretical Roscoe, Arthur and Coulomb shear band angles. For all tests the tolerance for nonlinear convergence is defined $tol=10^{-7}$
-and none converges before the maximum number of iterations ($it_{max}=1000$) is reached. However, tests with internal friction angles up to 15° show a constant
-decrease of the velocity residuals, which can not be observed for for tests with higher internal friction angles (Fig. \ref{fig:convergence}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Brick_experiment}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Brick_Beam.pdf}
-\caption{Shear band angles predicted for the brick experiment for $\phi=0$° (panel a) and for $\phi=20$° in case of compression and extension (panels b and c,
-respectively).}
-\label{fig:brick_beam}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Brick.pdf}
-\caption{Shear band angles predicted for the brick experiment in case of compressional and extensional contexts as function of different internal angle of
-friction (continuous black line and red dots), compared with theoretical Roscoe, Artur and Coulomb angles (discontinuous black lines). Red lines indicate the
-range of angles calculated at different depths.}
-\label{fig:brick}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \noindent\includegraphics[width=500px]{./Figures/convergence.pdf}
-  \caption{Convergence rates for all tests. Normalised residuals (left column) are compared with velocity residuals (right column) for extensional (first row)
-  and compressional (second row) settings.}
-  \label{fig:convergence}
-  \end{figure}
-
-\subsection{Advection stabilisation}\label{sec:advection}
-The 1D advection problem proposed by \citet{Donea2003} and \citet{Thieulot2011} is performed to verify the effectiveness of the SUPG method to stabilise the
-advection term of the energy equation. The domain is a 1D segment with $L_x=1$ composed by 50 elements and a discontinuity in the thermal field placed at
-$x=0.25$. Temperature is set to 1 for $x \leq 0.25$ and to 0 for $x > 0.25$. Velocity is set to ${u}=1$ in the entire domain. The simulation is performed for
-250 time steps, with $dt=0.002$, so the thermal discontinuity should be at $x=0.75$ at the end of the simulation.
-
-Temperature profiles at the end of the simulation are shown in \ref{fig:advection} as function of the dimensionless coefficient $\gamma=\tau \bm{v}/h$. In case
-of the classic Galerkin method ($\gamma=0$, blue line in \ref{fig:advection}) the final thermal profile is characterised by strong oscillations, which are
-eliminated in case of the SUPG method ($\gamma=0.045$, orange line in \ref{fig:advection}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Advection_stabilisation}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Advection.pdf}
-\caption{Temperature profile as function of \textit{x} for the advection stabilisation benchmark. Purple line indicates the initial temperature profile; the
-green line indicates the analytical temperature profile after 250 time steps; blue line indicates the temperature profile after 250 time steps in case of the
-classic Galerkin method ($\gamma=0$); orange line indicates the temperature profile after 250 time steps in case of the SUPG method ($\gamma=0.045$).}
-\label{fig:advection}
-\end{figure}
-
-\subsection{Shear heating}\label{sec:simple_shear}
-This exercise is performed in an unit square domain composed by \num{128x128} elements. The velocity field is prescribed on the entire domain with
-$\bm{v}=(L_y-y)y\bm{e_x}$; viscosity, density and specific heat are fixed to 1, while thermal conductivity and radiogenic energy are fixed to 0. Therefore,
-the energy equation (Eq. \ref{eq:energy}) can be simplified as \[\frac{\partial T}{\partial t}=H_s\] and fixing $T(t=0)=0$, the temperature can be find as
-\[T(t)=H_s t\]
-In this case we have
-\begin{eqnarray}
-\dot{\epsilon_{xx}}&=&\frac{\partial u}{\partial x}=0 \nonumber \\
-\dot{\epsilon_{yy}}&=&\frac{\partial v}{\partial y}=0 \nonumber \\
-\dot{\epsilon_{xy}}&=&\frac{1}{2}\left(\frac{\partial v}{\partial x}+\frac{\partial u}{\partial y}\right)=\frac{1}{2}(L_y-2y)\nonumber
-\end{eqnarray}
-and, simplifying Eq. \ref{eq:shear}, shear heating can be calculated as \[H_s(x,y)=(1-2y)^2\]
-
-The solution predicted by the model in terms of velocity, temperature and shear heating match well with the analytical solutions (Fig. \ref{fig:analytical_en}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Simple_shear}.
-
-%\begin{figure}
-%\centering
-%\noindent\includegraphics[width=400px]{./Figures/Energy_Thieulot.pdf}
-%\caption{Temperature and shear heating predicted by the model after 1 time step.}
-%\label{fig:en_thieulot}
-%\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=500px]{./Figures/Analytical_Shear.pdf}
-\caption{Velocity (panel a), temperature (panel b) and shear heating (panel c) as function of the \textit{y} coordinate for the simple shear experiment. The
-solutions predicted by the model (red lines) are compared with the analytical solutions (dashed blue lines).}
-\label{fig:analytical_en}
-\end{figure}
-
-\subsection{Shear and adiabatic heating}\label{sec:shear}
-This problem is performed as presented in Exercise 9.4 in \citet{Gerya2010b}. Two materials are vertically separated in a rectangular domain with $L_x=$
-\SI{1000}{\km}, $L_y=$ \SI{1500}{\km} and a grid resolution of \num{30x20} elements. Constant thermal coefficient expansion ($\alpha=$ \SI{3e-5}{\per\kelvin}),
-temperature ($T=$ \SI{1300}{\kelvin}) and gravity acceleration ($g_y=$ \SI{-10}{\m\square\s}) are assumed in the whole domain. Fluid 1 (on the left side of the
-domain) has $\rho_1=$ \SI{3200}{\kg\per\cubic\m} and $\eta_1=$ \SI{e20}{\pascal\s}; fluid 2 (on the right side of the domain) has $\rho_2=$
-\SI{3300}{\kg\per\cubic\m} and $\eta_2=$ \SI{e22}{\pascal\s}. Velocity boundary conditions are set to free slip on all sides of the domain.
-
-As shown in Fig. \ref{fig:energy}, both shear and adiabatic heating predicted by the code (first row in Fig. \ref{fig:energy}) well recreate the results
-obtained by means of example ${Shear\_adiabatic\_heating.m}$ from \citet{Gerya2010b} (second row in Fig. \ref{fig:energy}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Adiabatic%2BShear}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Energy.pdf}
-\caption{Comparison between shear (first column), adiabatic (second column) and total (third column) energy predicted by the code (panels a, b and c) and those
-created using example ${Shear\_adiabatic\_heating.m}$ from Exercise 9.4 in \citet{Gerya2010b} (panels d, e and f).}
-\label{fig:energy}
-\end{figure}
-
-\subsection{Mantle convection}\label{sec:mantle}
-This problem is performed as presented by \citet{Blankenbach1989} (constant viscosity cases) and \citet{Thieulot2014}, in a 2D unit square domain with gravity
-acceleration $g_y=10^{10}Ra$. The experiment is performed with three different Rayleigh numbers ($Ra=10^4$, $10^5$ and $10^6$) and with different grid
-resolution (between \num{32x32} and \num{128x128} elements). The fluid has constant viscosity, initial density, heat capacity, thermal conductivity
-($\eta= \rho_0= C_p = k=1$), reference temperature ($T_0=0$) and thermal expansion coefficient ($\alpha=10^{-10}$). Temperatures are set to 0 on top and 1 on
-bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\]
-
-The solution generated by the code in terms of $v_{\textrm{rms}}$ (calculated as in Sec. \ref{sec:error}) and Nu as function of time are reported for all the
-simulations in Fig. \ref{fig:mantle}. At the steady state, $v_{\textrm{rms}}$ and Nu of all simulations converge well toward the values from \citet{Blankenbach1989},
-with lower errors for higher resolution grids (Table \ref{tab:mantle}). All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Mantle_convection}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Convection.pdf}
-\caption{$v_{\textrm{rms}}$ (panels a, c and e) and Nu (panels b, d and f) for the mantle convection benchmark as function of time for different grid resolution.
-Panels a and b show the results for $Ra=10^4$; Panels c and d show the results for $Ra=10^5$; panels e and f show the results for $Ra=10^6$. Purple lines
-indicate the convergence values for the $v_{\textrm{rms}}$ and the Nu from \citet{Blankenbach1989} at the steady state.}
-\label{fig:mantle}
-\end{figure}
-
-\begin{table}[H]
-\caption{Comparison between $v_{\textrm{rms}}$ and Nu predicted by the code for the mantle convection experiment and same values as reported in literature.}
-\centering
-\small
-\begin{tabular}{|>{\centering}m{0.09\textwidth}|>{\centering}m{0.05\textwidth}|>{\centering}m{0.24\textwidth}|>{\centering}m{0.15\textwidth}|
-  >{\centering}m{0.1\textwidth} >{\centering}m{0.1\textwidth} >{\centering\arraybackslash}m{0.1\textwidth}|}
-\toprule
-& & \multirow{2}{*}{\citet{Blankenbach1989}} & \citet{Thieulot2014} & \multicolumn{3}{c|}{FALCON}  \\
- & & & \num{200x200} & \num{32x32} & \num{64x64} & \num{128x128}  \\
-\midrule
-\multirow{2}{*}{$Ra=10^4$} & $v_{\textrm{rms}}$ & $42.864947 \pm 0.000020$ & 42.867 & 42.83226 & 42.852793 & 42.861394  \\
-           & Nu      & $4.884409 \pm 0.000010$  & 4.882  & 4.781297 & 4.857475  & 4.877573   \\
-\hline
-\multirow{2}{*}{$Ra=10^5$} & $v_{\textrm{rms}}$ & $193.21454 \pm 0.00010$  & 193.255& 193.872643 & 193.377472 & 193.252290  \\
-           & Nu      & $10.534095 \pm 0.000010$ & 10.507 & 9.602514 & 10.270735  & 10.465629   \\
-\hline
-\multirow{2}{*}{$Ra=10^6$} & $v_{\textrm{rms}}$ & $833.98977 \pm 0.00020$  & 834.712& 848.091176 & 837.767911 & 834.945793  \\
-           & Nu      & $21.972465 \pm 0.000020$ & 21.695 & 15.999266 & 19.703682  & 21.306939   \\
-\bottomrule
-\end{tabular}
-\label{tab:mantle}
-\end{table}
-
-\subsection{Viscoplastic mantle convection}\label{sec:visco_mantle}
-This benchmark is performed as Cases 1-5a described in \citet{Tosi2015} in a 2D unit square domain with two different grid resolutions of \num{32x32} 
-and \num{100x100} elements. Reference viscosity, reference density, heat capacity, thermal conductivity and thermal expansion coefficient are fixed to a constant
-value of 1 ($\eta_0= \rho_0= C_p = k=\alpha=1$), while the gravity has been chosen as $g_y=10^2$, in order to obtain a Rayleigh number $Ra=10^2$. Temperatures
-are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\].
-The viscosity field $\eta$ is calculated as the harmonic average between a linear part $\eta_{lin}$ that depends on temperature only or on temperature and 
-depth $y$ and a nonlinear-plastic part $\eta_{pl}$ that depends on the strain rate $\dot{\bm{\epsilon}}$, as follows
-\[\eta(T,y,\dot{\bm{\epsilon}})=2\left(\frac{1}{\eta_{lin}(T,y)}+\frac{1}{\eta_{pl}(\dot{\bm{\epsilon}})}\right)^{-1}\]
-The linear and the nonlinear-plastic parts of the viscosity are calculated following \citet{Tosi2015} as follows
-\[\eta_{lin}(T,y)=\exp(-\ln(\Delta\eta_T)T+\ln(\Delta\eta_y)y)\]
-\[\eta_{pl}(\dot{\bm{\epsilon}})=\eta^*+\frac{\sigma_Y}{\sqrt{\dot{\bm{\epsilon}}:\dot{\bm{\epsilon}}}}\]
-where $\Delta\eta_T$, $\Delta\eta_y$, $\eta^*$ and $\sigma_Y$ are parameters chosen as listed in Table \ref{tab:tosi_setup}.
-
-For each case, some diagnostic quantities measured at the steady state are reported in Table \ref{tab:visco_mantle} (see Section \ref{sec:error} for 
-the computation of all the quantities), showing a well fit with results described in \citet{Tosi2015}. Data and charts for all cases can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Viscoplastic_mantle_convection}.
-
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{white!100!}{gray!25!}
-%\begin{table}[H]
-%  \caption{Results of the viscoplastic mantle convection benchmark for Cases 1-4 compared with results obtained with ELEFANT and YACC (from \citealp{Tosi2015}).}
-%  \centering
-%  \resizebox{\textwidth}{!}{
-%\begin{tabular}{|c|ccc|ccc|ccc|ccc|}
-%  \toprule
-%    & \multicolumn{3}{c|}{\textbf{Case 1}} & \multicolumn{3}{c|}{\textbf{Case 2}} & \multicolumn{3}{c|}{\textbf{Case 3}} &
-%    \multicolumn{3}{c|}{\textbf{Case 4}} \\
-%   & FALCON & ELEFANT& YACC & FALCON& ELEFANT& YACC & FALCON & ELEFANT & YACC & FALCON & ELEFANT & YACC \\
-%  \midrule
-%   $\left\langle T \right\rangle$ & 0.7759 & 0.7758 & 0.7767 & 0.6043 & 0.6036 & 0.6058 & 0.7275 & 0.7275 & 0.7286 & & 0.5277 & 0.5289\\
-%    Nu$^{top}$ & 3.4280 & 3.4214 & 3.4298 & 8.5242 & 8.5115 & 8.5278 & 3.0368 & 3.0347 & 3.0374 && 6.5912 & 6.5572 \\
-%    Nu$^{bottom}$ & 3.3168 & 3.313 & 3.3143 & 8.4599 & 8.4465 & 8.3990 & 3.0086 & 2.9908 & 2.9628 && 6.5834 & 6.5243 \\
-%    $v_{\textrm{rms}}$ & 249.901 & 249.134 & 251.7997 & 141.2451 & 140.7087 & 142.2020 & 100.2776 & 100.1208 & 100.9467 && 79.1105 & 79.6202 \\
-%    $v_{\textrm{rms}}^{top}$ & 1.8907 & 1.8642 & 1.8298 & 104.7928 & 104.1228 & 105.1767 & 2.0832 & 2.0652 & 2.0374 && 74.7596 & 75.8414 \\
-%    $u_{max}^{top}$ & 2.6340 & 2.6119 & 2.5516 & 121.9212& 121.766 & 122.3238 & 2.9086 & 2.9019 & 2.8458 && 88.9146 & 89.2940 \\
-%    $\eta_{min}$ & - & - & - & $2.0303\times10^{-5}$ & $2.0600\times10^{-5}$ & $1.9853\times10^{-5}$ & $4.8129\times10^{-5}$ & $4.8080\times10^{-5}$ & $4.7907\times10^{-5}$ &
-%    & $1.9860\times10^{-4}$ & $1.9174\times10^{-4}$ \\
-%    $\eta_{max}$ & - & - & - & 1.3325 & 1.85 & 1.6095 & 0.9360 & 0.9023 & 1 && 1.52 & 1.6773 \\
-%    $\left\langle W \right\rangle$ & 2.4299 & 2.4316 & 2.4583 & 7.5648 & 7.5665 & 7.6222 & 2.0387 & 2.0384 & 2.0400 && 5.6216 & 5.6512 \\
-%    $\left\langle \Phi \right\rangle/Ra$ & 2.4311 & 2.4276 & 2.4333 & 7.5661 & 7.5585 & 7.6194 & 2.0386 & 2.0370 & 2.0335 && 5.6182 & 5.6463 \\
-%    $\delta$ & 0.05\% & 0.16\% & 1.02\% & 0.02\% & 0.11\% & 0.04\% & $<$0.01\% & 0.07\% & 0.32\% && 0.06\% & 0.09\% \\
-%  \bottomrule
-%  \end{tabular}}
-%  \label{tab:visco_mantle}
-%  \end{table}
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{2}{white!100!}{gray!25!}
-\begin{table}[H]
-  \caption{Benchmark cases and corresponding parameters.}
-  \centering
-  \small
-\begin{tabular}{|>{\centering}m{0.08\textwidth}|>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}
-  >{\centering}m{0.05\textwidth} >{\centering}m{0.05\textwidth} |>{\centering\arraybackslash}m{0.2\textwidth}|}
-  \toprule
-    Case & $Ra$ & $\Delta\eta_T$ & $\Delta\eta_y$ & $\eta^*$ & $\sigma_Y$ & Convective regime \\
-  \midrule
-  \rowcolor{gray!25!}
-    1 & $10^2$ & $10^5$ & 1 & - & - & Stagnant lid \\
-    2 & $10^2$ & $10^5$ & 1 & $10^{-3}$ & 1 & Mobile lid  \\
-    \rowcolor{gray!25!}
-    3 & $10^2$ & $10^5$ & 10 & - & - & Stagnant lid  \\
-    4 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 1 & Mobile lid \\
-    \rowcolor{gray!25!}
-    5 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 4 & Periodic \\
-  \bottomrule
-  \end{tabular}
-  \label{tab:tosi_setup}
-  \end{table}
-
-\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{gray!25!}{white!100!}
-\begin{table}
-  \caption{Results of the viscoplastic mantle convection benchmark for different grid resolutions in Cases 1-5 compared with results obtained with ELEFANT
-  and YACC ($100\times100$ elements, from \citealp{Tosi2015}) and in Stone 28 ($32\times32$ elements, from
-  \url{https://github.com/cedrict/fieldstone/tree/master/python_codes/fieldstone_28}).}
-  \centering
-  \resizebox{\textwidth}{!}{
-\begin{tabular}{|cc|ccccccccccc|}
-  \toprule
-  \multicolumn{13}{|c|}{\textbf{Case 1}} \\
-  \midrule
-  Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-  & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-  \hline
-  \rowcolor{gray!25!}
-    YACC & $100\times100$ & 0.7767 & 3.4298 & 3.3143 & 251.7997 & 1.8298 & 2.5516 & - & - & 2.4583 & 2.4333 & 1.02\% \\
-    ELEFANT & $100\times100$ & 0.7758 & 3.4214 & 3.313 & 249.134 & 1.8642 & 2.6119 & - & - & 2.4316 & 2.4276 & 0.16\% \\
-    \rowcolor{gray!25!}
-    Stone 28 & $32\times32$ & 0.7737 & 3.3987 & - & 243.872 & - & - & - & - & - & - & - \\
-%    FALCON & $32\times32$ & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & - & - & 2.4745 & 2.4864 & 0.48\% \\
-%    \rowcolor{gray!25!}
-    FALCON & $100\times100$ & 0.7759 & 3.4280 & 3.3168 & 249.901 & 1.8907 & 2.6340 & - & - & 2.4299 & 2.4311 & 0.05\% \\
-    \midrule
-    \multicolumn{13}{|c|}{\textbf{Case 2}} \\
-    \midrule
-    Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-    & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-    \hline
-    \rowcolor{gray!25!}
-     YACC & $100\times100$ & 0.6058 & 8.5278 & 8.3990 & 142.2020 & 105.1767 & 122.3238 & $1.9853\times10^{-5}$ & 1.6095 & 7.6222 & 7.6194 & 0.04\% \\
-     ELEFANT & $100\times100$ & 0.6034 & 8.5115 & 8.4465 & 140.8067 & 104.1228 & 121.766 & $2.0600\times10^{-5}$ & 1.85 & 7.5665 & 7.5585 & 0.11\% \\
-     \rowcolor{gray!25!}
-     Stone 28 & $32\times32$ & 0.6052 & 8.1715 & - & 141.518 & - & - & - & - & - & - & - \\
-%     FALCON & $32\times32$ & 0.6454 & 8.1422 & 7.4020 & 156.6908 & 95.7226 & 116.9302 & $2.2572\times10^{-5}$ & 0.7194 & 7.4289 & 7.4452 & 0.22\% \\
-%     \rowcolor{gray!25!}
-     FALCON & $100\times100$ & 0.6044 & 8.5266 & 8.4590 & 141.3684 & 104.8867 & 122.0470 & $2.0333\times10^{-5}$ & 1.2914 & 7.5707 & 7.5718 & 0.01\% \\
-     \midrule
-     \multicolumn{13}{|c|}{\textbf{Case 3}} \\
-     \midrule
-     Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-     & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-     \hline
-     \rowcolor{gray!25!} 
-     YACC & $100\times100$ & 0.7286 & 3.0374 & 2.9628 & 100.9467 & 2.0374 & 2.8458 & $4.7907\times10^{-5}$ & 1 & 2.0400 & 2.0335 & 0.32\% \\
-      ELEFANT & $100\times100$ & 0.7275 & 3.0347 & 2.9908 & 100.1208 & 2.0652 & 2.9019 & $4.8080\times10^{-5}$ & 0.9023 & 2.0384 & 2.037 & 0.07\% \\
-      \rowcolor{gray!25!}
-      Stone 28 & $32\times32$ & 0.7277 & 3.0347 & - & 100.018 & - & - & - & - & - & - & - \\
-%      FALCON & $32\times32$ & 0.7281 & 3.0618 & 2.8132 & 101.0092 & 2.1630 & 3.0178 & $4.8500\times10^{-5}$ & 0.7845 & 2.0699 & 2.0695 & 0.02\% \\
-%      \rowcolor{gray!25!}
-      FALCON & $100\times100$ & 0.7275 & 3.0368 & 3.0086 & 100.2776 & 2.0832 & 2.9086 & $4.8129\times10^{-5}$ & 0.9360 & 2.0387 & 2.0386 & $<$0.01\% \\
-      \midrule
-      \multicolumn{13}{|c|}{\textbf{Case 4}} \\
-      \midrule
-      Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-      & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-      \hline
-      \rowcolor{gray!25!} 
-      YACC & $100\times100$ & 0.5289 & 6.5572 & 6.5243 & 79.6202 & 75.4814 & 89.2940 & $1.9174\times10^{-4}$ & 1.6773 & 5.6512 & 5.6463 & 0.09\% \\
-       ELEFANT & $100\times100$ & 0.5277 & 6.5912 & 6.5834 & 79.1105 & 74.7596 & 88.9146 & $1.9860\times10^{-4}$ & 1.5200 & 5.6216 & 5.6182 & 0.06\% \\
-       \rowcolor{gray!25!}
-       Stone 28 & $32\times32$ & 0.5286 & 6.4036 & - & 79.4746 & - & - & - & - & - & - & - \\
-%       FALCON & $32\times32$ & 0.5334 & 6.4679 & 6.3435 & 80.6153 & 76.0511 & 90.2058 & $2.0393\times10^{-4}$ & 0.6453 & 5.6676 & 5.6698 & 0.04\% \\
-%       \rowcolor{gray!25!}
-       FALCON & $100\times100$ & 0.5279 & 6.5897 & 6.5830 & 79.1220 & 75.0715 & 88.8484 & $1.9693\times10^{-4}$ & 1.4133 & 5.6127 & 5.6113 & 0.03\% \\
-       \midrule
-       \multicolumn{13}{|c|}{\textbf{Case 5}} \\
-       \midrule
-       Code & Resolution & $P$ & $\left\langle T \right\rangle_{min}$ & $\left\langle T \right\rangle_{max}$ & Nu$^{top}_{min}$ & Nu$^{top}_{max}$ &
-       $v_{\textrm{rms}}^{min}$ & $v_{\textrm{rms}}^{max}$ & \multicolumn{2}{c}{$(\left\langle \Phi \right\rangle/Ra)_{min}$} & 
-       \multicolumn{2}{c|}{$(\left\langle \Phi \right\rangle/Ra)_{max}$} \\
-       \hline
-       \rowcolor{gray!25!}
-         YACC & $100\times100$ & 0.0792 & 0.6543 & 0.6722 & 2.6950 & 7.2792 & 41.9813 & 98.8230 & \multicolumn{2}{c}{1.4347} & 
-         \multicolumn{2}{c|}{9.2748} \\
-         ELEFANT & $100\times100$ & 0.076 & 0.6514 & 0.6693 & 2.6808 & 7.3506 & 41.2578 & 100.089 & \multicolumn{2}{c}{1.3324} & 
-         \multicolumn{2}{c|}{9.5289} \\
-         \rowcolor{gray!25!}
-         Stone 28 & $32\times32$ & 0.0764 & 0.6521 & 0.6697 & 2.6912 & 7.0763 & 41.61 & 98.183 & \multicolumn{2}{c}{-} & 
-         \multicolumn{2}{c|}{-} \\
-     %    FALCON & $32\times32$ & & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & \multicolumn{2}{c}{-} & 
-     %     \multicolumn{2}{c|}{-} \\
-     %    \rowcolor{gray!25!}
-         FALCON & $100\times100$ & 0.0775 & 0.6519 & 0.6697 & 2.6847 & 7.3629 & 41.5256 & 99.5466 & \multicolumn{2}{c}{1.3399} & 
-         \multicolumn{2}{c|}{9.3993} \\
-  \bottomrule
-  \end{tabular}}
-  \label{tab:visco_mantle}
-  \end{table}
-
-\subsection{Thin layer entrainment}\label{sec:thin}
-This experiment is performed as originally proposed by \citet{vanKeken1997} with two incompressible fluids in a rectangular domain with $L_x=2$ and $L_y=1$
-and gravity acceleration $g_y=10^{10}Ra$. The Rayleigh number and the compositional Rayleigh number are fixed to $Ra=300000$ and $Rc=450000$, respectively.
-Both fluids have constant viscosity, thermal conductivity, specific heat ($\eta=\rho=C_p=1$) and thermal expansion coefficient ($\alpha=10^{-10}$). Fluid 1
-has a density $\rho_1=1$, while fluid 2 is denser ($\rho_2=\rho_1+1.5 \cdot 10^{-10}$) and is placed at the bottom of the domain, for $y \leq 0.025$.
-Temperature are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides of the domain. The initial
-temperature field is given by
-\[T(x,y)=T_u(x,y)+T_l(x,y)+T_r(x,y)+T_s(x,y)-\frac{3}{2}\]
-with
-\begin{eqnarray}
-T_u(x,y)&=&\frac{1}{2}erf\left(\frac{1-y}{2}\sqrt{\frac{u_0}{x}}\right)\nonumber \\
-T_l(x,y)&=&1-\frac{1}{2}erf\left(\frac{y}{2}\sqrt{\frac{u_0}{L_x-x}}\right)\nonumber \\
-T_r(x,y)&=&\frac{1}{2}+\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{y+1}} \exp\left(-\frac{x^2u_0}{4y+4}\right)\nonumber \\
-T_s(x,y)&=&\frac{1}{2}-\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{2-y}} \exp\left(-\frac{(L_x-x)^2u_0}{8-4y}\right)\nonumber
-\end{eqnarray}
-and
-\begin{eqnarray}
-u_0&=&\frac{L_x^{7/3}}{(1+L_x^4)^{2/3}}\left(\frac{Ra}{2\sqrt{\pi}}\right)^{\frac{2}{3}}\nonumber \\
-Q&=&2\sqrt{\frac{L_x}{\pi u_0}}\nonumber
-\end{eqnarray}
-
-The experiment is performed with two grid resolution (\num{125x40} and \num{200x80} elements), with 100 markers per element and CFL of 0.25. $v_{\textrm{rms}}$
-(see Sec. \ref{sec:error}) calculated for both the simulations match well with results from \citet{vanKeken1997} and obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and ELEFANT \citep{Thieulot2014}, obtained with similar grid resolutions
-(Fig. \ref{fig:thin}). In addition, a variety of simulations with different resolutions and aspect ratios are performed to verify that the sensitivity of the
-initial velocity field. Results match well with results from \citet{vanKeken1997} (in the grey area) and \citet{Thieulot2014} (Fig. \ref{fig:thin_initial}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Thin_layer}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_thin.pdf}
-\caption{$v_{\textrm{rms}}$ for the thin layer experiment as function of time for different grid resolution. Results are compared with results obtained by
-\citet{vanKeken1997} (black lines) and with ASPECT and ELEFANT.}
-\label{fig:thin}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrmszero.pdf}
-\caption{$v_{\textrm{rms}}$ at $t=0$ for the thin layer experiment as function of the element size for different aspect ratios of the elements. Results are compared
-with results obtained by \citet{vanKeken1997} (black dots in the grey area) and with ASPECT and ELEFANT.}
-\label{fig:thin_initial}
-\end{figure}
-
-\subsection{Sinking hydrated cylinder}\label{sec:quinquis}
-This experiment is performed as explained in \citet{Quinquis2014} and simulates the sinking of a cold, hydrated cylinder into a warm, dry mantle. The domain
-is square with $L_x=L_y=$ \SI{300}{\km}, a grid resolution of \num{300x300} elements and 25 markers per element. The sinking cylinder is located at $x=$
-\SI{150}{\km} and $y=$ \SI{170}{\km} with a radius of \SI{20}{\km} and has $\rho_c=$ \SI{3250}{\kg\per\cubic\m}, $\eta_c=$ \SI{e23}{\pascal\s}, a thermal
-conductivity $k_c=$ \SI{4.5}{\watt\per\m\per\kelvin},a specific heat $C_{p_c}=$ \SI{1250}{\joule\per\kg\per\kelvin}; the surrounding mantle has $\rho_m=$
-\SI{3200}{\kg\per\cubic\m}, $\eta_m=$ \SI{e20}{\pascal\s}, a thermal conductivity $k_m=$ \SI{105}{\watt\per\m\per\kelvin} and specific heat $C_{p_m}=$
-\SI{1250}{\joule\per\kg\per\kelvin}; a 58 km-thick lithosphere overlies the mantle and has $\rho_l=$ \SI{3200}{\kg\per\cubic\m}, $\eta_l=$ \SI{e23}{\pascal\s},
-a thermal conductivity $k_l=$ \SI{4.5}{\watt\per\m\per\kelvin}, specific heat $C_{p_l}=$ \SI{750}{\joule\per\kg\per\kelvin} and a thermal diffusivity of
-\SI{e-6}{\square\m\per\s}. Velocity boundary conditions are set to free slip on all sides of the domain. Temperature increase linearly in the lithosphere,
-from 0° to 1300°C, and in the mantle, from 1300° to 1360.5°C. The initial temperature of the cylinder is fixed to 400°C. Throughout the evolution, temperature
-is fixed to 0° and 1360.5°C at the top and bottom of the domain, respectively. The initial bound water content is imposed at 2 wt.\% in the cylinder and at
-0 wt.\% in both the mantle and the lithosphere. The maximum amount of water in the mantle is fixed to 0.2 wt.\%, while maximum water content in the cylinder
-and in the lithosphere is function of pressure and temperature and is calculated using a serpentinized harzburgite with Perple\_X, as in \citet{Quinquis2014}.
-
-As shown in \citet{Quinquis2014}, the model is characterised by a progressive dehydration in the external portion of the cylinder, due to the increase of
-temperature, with a consequent vertical migration of free water that hydrates the mantle above the cylinder, up to the lithosphere. The distribution of
-bound and free water during the evolution of the experiment is shown in Fig. \ref{fig:quinquis} (left and right columns, respectively). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Hydration}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Hydration.pdf}
-\caption{Distribution of bound and free water (left and right columns, respectively) at $t=0.5$ and $t=$ \SI{1.25}{\mega\year} (first and second rows,
-respectively) for the sinking hydrated cylinder experiment. Only markers with bound or free water not equal to 0 are plotted.}
-\label{fig:quinquis}
-\end{figure}
-
-\subsection{Mantle melting curves}\label{sec:katz}
-This instantaneous experiment is performed in order to recreate isobaric melting curves, as function of temperature, bulk water and content of water in the
-melt, to compare with melting curves obtained by \citet{Katz2003}. The domain is square with $L_x=L_y=100$ km, a grid resolution of \num{100x100} elements and
-25 markers per element. Density is set in the entire domain to \SI{3300}{\kg\per\cubic\m} and temperature increases from the sides to the centre from 1000° to
-1500°C. The experiment is performed for bulk water contents from 0 to 0.3 wt.\%.
-
-Fig. \ref{fig:fraction} shows isobaric melting curves for \SI{1}{\giga\pascal} (panel a) and \SI{3}{\giga\pascal} (panel b) calculated for different bulk water
-contents. The curves well match with those obtained for the mantle by \citet{Katz2003}, showing the saturation at the solidus for bulk water of 0.2 and 0.3 wt.\%, indicating
-that the code correctly determines melt fraction as function of pressure, temperature and bulk water content. Same results can be observed in Fig.
-\ref{fig:water}a, where isobaric and isothermal melting curves are shown as function of bulk water content. In addition, Fig. \ref{fig:water}b shows that the
-code correctly determines also the percentage of water in the melt as function of pressure, temperature and melt fraction. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Melting}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Fraction.pdf}
-\caption{Isobaric melting curves for 1 GPa (panel a) and 3 GPa (panel b) as function of temperature with different bulk water contents.}
-\label{fig:fraction}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Water.pdf}
-\caption{Isobaric melting curves for 1.5 GPa as function of bulk water (panel a) and water in the melt (panel b) calculated for different temperatures.}
-\label{fig:water}
-\end{figure}
-
-\clearpage
-
-\bibliography{export}
-
-\end{document}

### 2022-09-07 tex/nonlinear.tex
update main

--- 2022-03-31 FALCON_description.tex
@@
-\documentclass[hidelinks,10pt,a4paper]{article}
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{textcomp}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage{booktabs}
-\usepackage{multirow}
-\usepackage[cm]{fullpage}
-\usepackage[round]{natbib}
-\bibliographystyle{abbrvnat}
-\usepackage{doi}
-\usepackage{bm}
-\usepackage[table]{xcolor}
-\hypersetup{
-    colorlinks,
-    linkcolor={blue},
-    citecolor={blue},
-    urlcolor={blue}
-}
-\usepackage{float}
-\restylefloat{table}
-\usepackage{siunitx}
-\DeclareSIUnit\year{yr}
-
-\begin{document}
-
-\section{Introduction}\label{sec:intro}
-This document includes a description of the code FALCON, with a presentation of the features implemented followed by results of extensive benchmarks listed in
-Table \ref{tab:list}, performed to verify that the code solves correctly a wide range of problems. Description and results of all benchmarks are shown
-in Section \ref{sec:benchmark}.
-
-\begin{table}[H]
-\caption{List of benchmarks performed for each feature of the code. Data of all benchmarks can be found at 
-\url{https://github.com/aleregorda/Benchmarks}}
-\centering
-\begin{tabular}{| l | l | l |}
-\toprule
- \textbf{Feature}  & \textbf{Benchmark} & \textbf{Section} \\
-\midrule
-  Solver  & Stokes flow & \ref{sec:stokes} \\
-\hline
-  \multirow{2}{*}{Markers advection} & Zalesak disk & \ref{sec:runge} \\
-    & Conservative Interpolation Velocity & \ref{sec:CVI} \\
-\hline
-  \multirow{4}{*}{Momentum equation}  & Poiseuille flow & \ref{sec:poiseuille}\\
-    & Instantaneous 2D sphere & \ref{sec:ist_sphere} \\
-    & Rayleigh-Taylor experiment & \ref{sec:rayleigh} \\
-    & Falling block & \ref{sec:block} \\
-\hline
-    & 2D Stokes sphere & \ref{sec:sphere} \\
-    Sticky air and  & Stabilisation algorithm & \ref{sec:stab} \\
-    free surface & Topography relaxation & \ref{sec:crameri} \\
-  & Spontaneous subduction & \ref{sec:subduction} \\
-\hline
-  Erosion and sedimentation  & - & - \\
-\hline
-   Non-linear  & Slab detachment & \ref{sec:slab} \\
-   visco-plasticity & Indenter & \ref{sec:indenter} \\
-    & Brick & \ref{sec:brick} \\
-\hline
-  \multirow{3}{*}{Energy equation}  & Advection stabilisation & \ref{sec:advection} \\
-    & Simple shear heating & \ref{sec:simple_shear} \\
-    & Shear and adiabatic heating & \ref{sec:shear} \\
-\hline
-\multirow{3}{*}{Energy + momentum} & Mantle convection & \ref{sec:mantle} \\
-& Viscoplastic mantle convection & \ref{sec:visco_mantle} \\
-& Thin layer entrainment & \ref{sec:thin} \\
-\hline
-  Phase changes  & \multirow{2}{*}{Hydrated sinking cylinder} & \multirow{2}{*}{\ref{sec:quinquis}} \\
-  and hydration  &  & \\
-\hline
-  Melting & Experimental melting curves & \ref{sec:katz} \\
-\bottomrule
-\end{tabular}
-\label{tab:list}
-\end{table}
-
-The code is written in Fortran90 and uses the finite element method (FEM) with quadrilateral $Q_1 \times P_0$ elements (continuous bilinear velocity and
-discontinuous constant pressure), associated with MUMPS\footnote{\url{http://mumps.enseeiht.fr/}} \citep{Amestoy2001,Amestoy2019} that is a software package
-for solving systems of linear equations of the form $A \cdot x = b$, where \textit{A} is a square sparse matrix, by means of a direct method. Since
-$Q_1 \times P_0$ elements do not satisfy Ladyzhenskaya, Babuska and Brezzi (LBB) stability condition \citep{Donea2003}, elemental pressure is elaborated
-in post-processing to avoid spurious pressures by means of a double interpolation to smooth the pressure. Following this procedure, the elemental pressure
-is interpolated onto nodes and then back onto elements. After the smoothing procedure, the pressure field is calculated again on the nodes to be used in
-combination with the lithostatic pressure, which is calculated onto nodes. Correctness of the solution and performances of the solver in terms of time and
-memory usage are tested solving the Stokes flow with the analytical solution proposed by \citet{Donea2003} (Section \ref{sec:stokes}).
-
-The thermo-mechanics of crust-mantle systems is described by means of the conservation of mass, momentum and energy equations, expressed as follows:
-\begin{eqnarray}
-\label{eq:mass}\frac{\partial \rho}{\partial t}+\bm{\nabla} \cdot (\rho \bm{u})&=&0\\
-\label{eq:momentum}-\bm{\nabla} p + \bm{\nabla} \cdot \bm{\tau} + \rho \bm{g}&=&0\\
-\label{eq:energy}\rho C_p \left(\frac{\partial T}{\partial t} + \bm{u} \cdot \bm{\nabla} T \right)&=&\bm{\nabla} \cdot \left(k\bm{\nabla} T\right) + H_{tot}
-\end{eqnarray}
-where $\rho$ is the density, $\bm{u}$ is the velocity, \textit{p} is the pressure, $\bm{\tau}$ is the deviatoric stress, $\bm{g}$ is the gravity acceleration,
-$C_p$ is the specific heat at constant pressure, \textit{T} is the temperature, \textit{k} is the thermal conductivity and $H_{tot}$ is the total internal heat
-production. Density variations due to temperature are generally small enough to assume the density as constant ($\rho=\rho_0$) in Eq. \ref{eq:mass} and in
-Eq. \ref{eq:energy}, while it must be treated as a variable in the buoyancy term of Eq. \ref{eq:momentum}, such that \[\rho=\rho_0(1-\alpha(T-T_0))\]
-where $\rho_0$ is the density at a reference temperature $T_0$ and $\alpha$ is the coefficient of thermal expansion. Eq. \ref{eq:mass} then can be rewritten as
-\begin{equation}\label{eq:mass_0}
-\bm{\nabla} \cdot \bm{u}=0
-\end{equation}
-This simplification is known as the extended Boussinesq approximation.
-
-The time step $dt$ is chosen in according to Courant-Friedrichs-Lewy condition \citep{Anderson1995}, such that
-\[dt=C_n\; \min\left(\frac{\Delta x}{\max|\bm{u}|},\frac{(\Delta x)^2}{\kappa}\right)\]
-where $0<C_n<1$ is the Courant (CFL) number, $\Delta x$ is the minimum dimension of smallest element, $\max|\bm{u}|$ is the maximum velocity calculated on the
-entire domain and
-$\kappa=\frac{k}{\rho C_p}$ is the thermal diffusivity \citep{Thieulot2014}.
-
-\section{Lagrangian markers}\label{sec:markers}
-Elemental properties (density, viscosity, thermal conductivity, specific heat and thermal expansion) needed to solve Eqs. \ref{eq:mass_0}, \ref{eq:momentum}
-and \ref{eq:energy} are related to the composition of each element, determined by means of Lagrangian markers that are characteristic of different materials
-in the domain. Elemental properties (with the sole exception of the viscosity) are calculated using an arithmetic mean, as \[P_e=\frac{1}{n}\sum_{i=1}^n P_i\]
-where $P_e$ is the elemental property, $P_i$ is the property characteristic of the material of each marker and $n$ is the number of markers of the element.
-Differently, the average scheme for the elemental viscosity can be chosen between harmonic, geometric and arithmetic mean. Markers can be placed regularly or
-randomly at the beginning of the simulation and their advection is performed by either a 2nd-order or a 4th-order Runge-Kutta in space, interpolating the velocity 
-field on each marker by means of the shape functions (see Section \ref{sec:runge}). The interpolated velocity is then corrected by means of the Conservative
-Velocity Interpolation (CVI), which introduces a corrective term to reduce dispersion and clustering of particles in both steady state and time-dependent 
-\citep{Wang2015} (see Section \ref{sec:CVI}).
-
-The initial distribution of the markers is created using the open-source code library
-Geodynamic World Builder \citep{Fraters2019}. During the simulation, each marker carries memory of temperature, pressure and accumulated strain, which are
-determined interpolating the relative nodal parameter. In particular, pressure is interpolated onto each marker after the smoothing procedure. The number
-of markers contained in each element is maintained between $n_{min}$ and $n_{max}$. When in an element there are less markers than $n_{min}$ the code adds
-random markers to reach the $n_{min}$, while if the number is higher than $n_{max}$ some of them are randomly deleted. When new markers are added, they assume
-the properties of the nearest marker. In this way elements are never empty and maintain a number of markers inside a prefixed range.
-
-\section{The momentum equation}\label{sec:mom_eq}
-The penalty method is implemented so that Eq. \ref{eq:mass_0} can be written as:
-\begin{equation}\label{eq:mass_penalty}
-\bm{\nabla} \cdot \bm{u} + \frac{p}{\lambda}=0
-\end{equation}
-where $\lambda$ is the so-called penalty parameter, which should be 6-7 orders of magnitude larger than the shear viscosity to ensure that mass conservation
-is satisfied \citep{Donea2003,Thieulot2014}.
-
-The deviatoric stress tensor in Eq. \ref{eq:momentum} can be written in terms of the strain rate tensor as $\bm{\tau}=2\eta\dot{\bm{\epsilon}}$, with
-$\dot{\bm{\epsilon}}=\frac{1}{2}\left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)$. Therefore, Eq. \ref{eq:momentum} can be rewritten as:
-\begin{equation}\label{eq:momentum_pressure}
--\bm{\nabla} p + \bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right) + \rho \bm{g} = 0
-\end{equation}
-Finally, using pressure from Eq. \ref{eq:mass_penalty}, Eq. \ref{eq:momentum_pressure} can be rewritten as:
-\begin{equation}\label{eq:momentum_penalty}
-\lambda \bm{\nabla} \left(\bm{\nabla} \cdot \bm{v}\right)+\bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right)+\rho\bm{g}=0
-\end{equation}
-The penalty method is been associated to the iterative Uzawa method, as described in detail in \citet{Dabrowski2008} and \citet{Thieulot2014}
-(see Section \ref{sec:poiseuille}). Results of benchmarks performed to verify the correctness in the implementation of Eq. \ref{eq:momentum_penalty} are
-shown in Sections \ref{sec:ist_sphere} and \ref{sec:rayleigh}.
-
-To support large viscosities variations the penalty parameter is related to the effective elemental viscosity by means of a dimensionless coefficient, so that
-$\lambda=\lambda_e(e)\eta_{eff}(e)$ \citep{Marotta2006,Dabrowski2008,Thieulot2014}. Benchmark of the falling block \citep{Gerya2003a,Gerya2010b,Thieulot2011}
-is performed to verify that the code can correctly deal buoyancy driven flows with strong viscosity contrasts (Section \ref{sec:block}).
-
-\section{Sticky air and free surface}\label{sec:sticky}
-The Earth's surface can be treated by means of either the so-called sticky-air or a true free surface method, both of them implemented in the code. In the
-sticky-air method the surface is approximated with the introduction of a buoyant layer with a viscosity at least four orders of magnitude lower than the crust
-\citep{Schmeling2008,Crameri2012} and the interface between lithosphere and air is defined using a chain of passive markers that are advected as the Lagrangian
-markers. The correctness of the evolution of the markers chain is tested with the experiment of a 2D time-dependent Stokes sphere below a free surface and
-compared with results from ASPECT\footnote{\url{https://aspect.geodynamics.org/}} \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} (Section
-\ref{sec:sphere}). In the true free surface case the top boundary is assumed stress-free and velocities are not fixed. In this case topography variations are
-described by vertical deformations of the mesh that depend on the velocity field of the nodes that identify the free surface, while horizontal deformations are
-not taken into account. This procedure is known as the Arbitrary Lagrangian-Eulerian (ALE) method and its implementation follows the technique described in
-\citet{Thieulot2011}. However, although the implementation of a true free surface better reproduces laboratory experiment, extremely small time steps can be
-necessary to maintain stability \citep{Kaus2010a,Quinquis2011,Thieulot2014}. Therefore, the stability algorithm proposed by \citet{Kaus2010a} is implemented
-to avoid instabilities due to high density differences at the free surface when using too large time steps. The implementation of this algorithm are tested by
-performing the experiment described by \citet{Kaus2010a} (Section \ref{sec:stab}).
-
-The topography relaxation benchmark proposed by \citet{Crameri2012} is performed to verify that the code correctly recovers topography variations in case of
-both the sticky-air and the true free surface method (Section \ref{sec:crameri}). Finally, Section \ref{sec:subduction} show the results of the spontaneous
-subduction experiment described by \citet{Schmeling2008}.
-
-\section{Erosion and sedimentation}\label{sec:erosion}
-Surface processes at either the lithosphere-air interface or in correspondence of the free surface have been implemented by means of the software
-Fastscape\footnote{\url{https://fastscape.org/}} \citep{Braun2013,Cordonnier2019,Yuan2019a,Yuan2019b}. FastScapeLib is a set of routines that solve the stream
-power law (enriched by a sediment transport/deposition term) \citep{Yuan2019a}, hillslope diffusion and marine transport and deposition \citep{Yuan2019b}.
-The partial differential equation solved by FastScapeLib is:
-\begin{equation}
-\frac{\partial h}{ \partial t}=U-K_f A^{m_e}S^{n_e} + \frac{G}{A}\int_A \left( U-\frac{\partial h}{\partial t} \right) dA + K_d \bm{\nabla}^2 h 
-\end{equation}
-where $h$ is the topography, $U$ is the uplift, $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $S$ is the slope, $m_e$ and $n_e$ are
-parameters that should be chosen to constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}, $G$ is the dimensionless
-deposition/transport coefficient for the enriched stream power law and $K_d$ is the effective diffusivity. Terms $K_f A^{m_e}S^{n_e}$ and $K_d \bm{\nabla}^2 h$
-consider long-range processes, described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, and short-range processes,
-proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}.
-
-In case of sticky air, slope is determined by means of the markers chain, which is vertically corrected considering erosion and sedimentation rates. After the
-correction, continental markers above the markers chain are transformed in air markers and air markers below the markers chain become sediments. Similarly,
-slope is determined considering the nodes on the top boundary in case of true free surface and they are vertically corrected as for the markers chain. In this
-case, continental markers above the top boundary are simply deleted.
-%Surface processes at the lithosphere-air interface are implemented taking into account fluvial erosion (long-range processes), hillslope processes (short-range
-%processes) and failure of slopes exceeding a threshold angle $\theta$ \citep{Burov1997,Tucker1999,Sacek2012,Croissant2014,Ueda2015,Beucher2020}. Erosion rates
-%$\dot{E}_{lr}$ due to long-range processes can be described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{lr}=K_f A^{m_e} \bm{\nabla} h^{n_e}
-%\end{equation}
-%where $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $\bm{\nabla} h$ is the slope and $m_e$ and $n_e$ are parameters chosen in to
-%constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}. Erosion rates $\dot{E}_{sr}$ due to short-range processes are
-%implemented by assuming that transport is linearly proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}, in according to
-%\begin{equation}
-%\dot{E}_{sr}=K_d \bm{\nabla}^2 h
-%\end{equation}
-%where $K_d$ is the effective diffusivity. In addition, the slope is modified if it still exceeds a threshold angle after long- and short-range processes
-%\citep{Ueda2015}.
-
-%Offshore sedimentation processes are implemented similarly to hillslope processes but with a depth-dependent coefficient \citep{Sacek2012,Beucher2020}.
-%The sedimentation rate $\dot{S}$ is calculated as
-%\begin{equation}
-%\dot{S}=K_s \bm{\nabla}^2 h
-%\end{equation}
-%with
-%\begin{equation}
-%K_s=K^*_s \exp\left(\frac{h-h_{sl}}{h_0}\right)
-%\end{equation}
-%where $K^*_s$ is a constant, $h_{sl}$ is the sea level and $h_0$ is a characteristic length. In case of sticky air, slope is determined by means of the markers
-%chain, which is vertically corrected considering erosion and sedimentation rates. After the correction, continental markers above the markers chain are
-%transformed in air markers and air markers below the markers chain become sediments. Similarly, slope is determined considering the nodes on the top boundary
-%in case of true free surface and they are vertically corrected as for the markers chain. In this case, continental markers above the top boundary are simply
-%deleted.
-
 \section{Non-linear rheologies}\label{sec:nonlinear}
@@
 brick experiments (Sections \ref{sec:indenter} and \ref{sec:brick}, respectively).
-
-Strain softening is taken into account for both viscous creep and plastic viscosity \citep{Huismans2003,Babeyko2005,Huismans2005,Sobolev2005,Warren2008a} by
-means of the accumulated strain $\epsilon$ memorised by each marker. Softening in the viscous creep determines a linear decrease of $\eta_{vc}$ by means of a
-viscous strain softening factor $W_S$ that increases linearly from $W_{S_0}$ to $W_{S_{\infty}}$ for $\epsilon_{S_0}<\epsilon_S<\epsilon_{S_{\infty}}$ 
-\citep{Warren2008a}. This viscous softening can be related to strain-induced grain size reduction \citep{Warren2008a}. Differently, plastic softening is 
-simulated with a linear decrease of internal friction angle $\phi(\epsilon)$ and cohesion $C(\epsilon)$ in according to
-\begin{eqnarray}
-\label{eq:friction1}\phi(\epsilon)&=&\phi_0+(\phi_{\infty}-\phi_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}\\
-\label{eq:friction2}C(\epsilon)&=&C_0+(C_{\infty}-C_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}
-\end{eqnarray}
-where $\phi_0$, $C_0$ and $\phi_{\infty}$, $C_{\infty}$ are internal friction angle and cohesion for $\epsilon_0$ and $\epsilon_{\infty}$, respectively
-\citep{Huismans2003,Huismans2005,Warren2008a,Thieulot2014}. Plastic softening approximates deformation-induced softening of faults and brittle shear zones
-\citep{Warren2008a}.
 
@@
 the velocity residual and the pressure residual converge under a defined tolerance $tol$.
-
-\section{The energy equation}\label{sec:en_eq}
-The stabilisation of the advection term of the energy equation (Eq. \ref{eq:energy}) needed to avoid possible oscillations in the thermal solution in the case
-when advection dominates over diffusion, is implemented by means of a streamline-upwind Petrov–Galerkin (SUPG) method, for which the advection term is modified
-follows the discussion in \citet{Thieulot2011} and \citet{Thieulot2014} (see Section \ref{sec:advection} for benchmark). The total internal heat production
-$H_{tot}$ in Eq. \ref{eq:energy} includes radiogenic heating $H_r$, shear heating $H_s$ and adiabatic heating $H_a$, where
-\begin{equation}\label{eq:shear}
-H_s=2\eta \dot{\bm{\epsilon}} : \dot{\bm{\epsilon}} = 2\eta(\dot{\epsilon}_{xx}^2+\dot{\epsilon}_{yy}^2+2\dot{\epsilon}_{xy}^2)
-\end{equation}
-and
-\begin{equation}\label{eq:adiabatic}
-H_a=T \alpha \frac{Dp}{Dt} \approx -\alpha T \rho g_y v
-\end{equation}
-The correctness of the implementation and the computation of shear and adiabatic heating is verified performing an exercise with an analytical solution
-(Section \ref{sec:simple_shear}) and Exercise 9.4 in \citet{Gerya2010b} (Section \ref{sec:shear}). Furthermore, mantle convection and thin layer entrainment
-benchmarks (Sections \ref{sec:mantle} and \ref{sec:thin}, respectively) are performed to verify that the code solve correctly Eqs. \ref{eq:momentum_penalty}
-and \ref{eq:energy} in case of a temperature field characterised by an initial perturbation and with a temperature-dependent density. At last, the viscoplastic
-mantle convection benchmark is performed to check also the viscoplasticity in case of a initial perturbed temperature field (Section \ref{sec:visco_mantle}).
-
-\section{Phase transitions and hydration}\label{sec:phase}
-Variations of effective density, specific heat and coefficient of thermal expansion during the evolution of crustal and mantle materials at different
-pressure-temperature ($p$-$T$) conditions are computed by means of version 6.8.6 of Perple\_X software package \citep{Connolly2005}, similarly to the
-implementation described in \citet{Marotta2020}. By default, mineral assemblages and properties of each lithology (oceanic crust, lower and upper continental
-crust, sediments and dry or hydrated mantle) are calculated for temperatures from \SIrange{330}{1600}{\kelvin}, with increments of approximately \SI{4}{\kelvin},
-and pressures from \SIrange{0.1}{30}{\giga\pascal}, with increments of approximately \SI{0.03}{\giga\pascal}, for a total of almost 100000 points.
-
-The introduction of phase transitions produces two main effects that must be taken into account in numerical models: variations in density and release/absorption
-of latent heat \citep{Tackley2010}. Effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into
-account considering effective coefficient of thermal expansion \citep{Christensen1985,Zhong2015}. Similarly, effects of latent heat in the energy equation
-(eq \ref{eq:energy}) are taken into account by considering effective specific heat and coefficient of thermal expansion
-\citep{Christensen1985,Tackley2010,Zhong2015}. Effective density, specific heat and coefficient of thermal expansion are included in Perple\_X files and are
-assumed by each Lagrangian marker in according to its $p$-$T$ conditions.
-
-Hydration processes related to dehydration of subducting lithosphere \citep{Schmidt1998,Liu2007,Faccenda2009,Faccenda2010,Faccenda2014,Rosas2016} strongly
-influence the thermo-mechanics inside the mantle wedge, mainly because of both weakening effects on the mantle rheology and density variations in case of
-mantle serpentinization \citep{Gerya2002,Honda2003,Arcay2005,Roda2010,Regorda2017}. In the case that hydration is switched on, the amount of bound and free
-water is memorised by each marker, following the implementation of \citet{Quinquis2014}; hydration and dehydration processes are related to the amount of bound
-water of each marker and to the maximum amount of water it can transport, i.e. if the amount of bound water exceeds the maximum amount of water, the marker
-dehydrates and releases free water that can hydrates under-saturated markers. Maximum water content of each marker is determined as function of lithology and
-$p$-$T$ conditions and it is calculated using Perple\_X, in the same way than effective density, specific heat and coefficient of thermal expansion. Bound water
-is advected together with the markers, neglecting the effect of bound water diffusion, while free water simply migrates vertically and is not coupled to the
-solid-phase flow of the mantle wedge \citep{Arcay2005,Quinquis2014}. The correctness of the migration scheme of the free water and its absorption into bound
-water by under-saturated markers are verified performing the experiment described in \citet{Quinquis2014} (Sec. \ref{sec:quinquis}).
-
-The mantle viscosity weakening related to the amount of water has been extensively studied \citep{Chopra1981,Hirth2003} and it is implemented in according to
-\citet{Arcay2005} and \citet{Horiuchi2016}, as follows
-\begin{equation}\label{eq:wet_visc}
-\eta_{wet}=\eta_{dry}\left(\left[1-\frac{1}{f_v}\right]\exp\left(-\frac{[OH^-]}{[OH^-]_0}\right)+\frac{1}{f_v}\right)
-\end{equation}
-where $f_v$ is the viscosity reduction factor between dry and wet conditions, $[OH^-]$ is the water content and $[OH^-]_0$ is a reference water content set to
-620 ppm (0.062 wt.\%) \citep{Arcay2005}. By default, $f_v$ is set to 100, in according to \citet{Arcay2005} and \citet{Horiuchi2016}.
-
-\section{Mantle melting}\label{sec:melt}
-Melt fraction in the mantle depends on temperature ($T$, in \textdegree C), pressure ($p$, in \si{\giga\pascal}) and the water content in the melt ($X_{H_2O}$, in wt.\%).
-The determination of the melt fraction $M$ is implemented as explained by \citet{Katz2003}, according to the successive modification by \citet{Langmuir2006} and
-\citet{Kelley2010}, as follows
-\begin{eqnarray}\label{eq:melt_fr}
-M(p,T,X_{H_2O})&=&-T+T_{sol}+(x_t\cdot ln(p+y_t))\cdot X_{H_2O}-\nonumber\\
-&&+K_m\left(\frac{X_{H_2O}^{bulk}}{D_{H_2O}(1-X_{H_2O})+X_{H_2O}}\right)^{\gamma}
-\end{eqnarray}
-where $T_{sol}$ is the temperature of the dry solidus, $D_{H_2O}$ is the partition coefficient and $x_t\cdot log(p+y_t)$ indicates the pressure dependence of
-the melting curve \citep{Kelley2010}. $x_t$, $y_t$, $K_m$ and $\gamma$ are parameters chosen in according to \citet{Kelley2010}.
-However, $X_{H_2O}$ depends on the melt fraction as
-\begin{equation}\label{eq:melt_wt}
-X_{H_2O}(M)=\frac{X_{H_2O}^{bulk}}{D_{H_2O}+M(1-D_{H_2O})}
-\end{equation}
-and a numerical solution can be found using a root-finder method \citep{Katz2003,Wang2016}. Furthermore, the content of water in the melt is limited by the
-pressure-dependent saturation concentration of water in the melt, determined as
-\begin{equation}\label{eq:melt_max}
-X_{H_2O}^{sat}=\chi_1 p^{\lambda_m}+\chi_2 p
-\end{equation}
-where $\chi_1$, $\chi_2$ and $\lambda_m$ are parameters chosen according to \citet{Katz2003}, \citet{Langmuir2006} and \citet{Kelley2010}. The correctness in
-the determination of melt fractions and water contents in the melt are tested comparing model results with experimental curves obtained by \citet{Katz2003}
-(Sec. \ref{sec:katz}).
-
-In case of melting, viscous creep viscosity ($\eta_{vp}$) is modified according to \citet{Wang2016}, as
-\begin{equation}\label{eq:melt_visc}
-\eta_{melt}=\eta_{vp}\cdot \exp(\alpha_m M)
-\end{equation}
-where $\alpha_m$ are melt fraction factors for dislocation and diffusion creep, chosen in according to \citet{Wang2016}.
-
-Effective density ($\rho_{eff}$) for partially molten rocks is calculated as
-\begin{equation}\label{eq:melt_dens}
-\rho_{eff}=\rho_s(1-M)+\rho_m M
-\end{equation}
-where $\rho_s$ and $\rho_m$ are the densities of the solid and the molten rock, respectively \citep{Gerya2004b,Gerya2007,Wang2016}. The density of the
-solid rock is extracted by Perple\_X, while the density of the molten portion is calculated as
-\begin{equation}\label{eq:melt_dens_M}
-\rho_m(p,T)=\rho_0[1-\alpha(T-T_0)][1+\beta(p-p_0)]
-\end{equation}
-where $\rho_0$ is the density at the reference temperature ($T_0$) and pressure($p_0$), and $\alpha$ and $\beta$ are the thermal expansion and compressibility
-coefficients, respectively \citep{Gerya2004b,Gerya2007,Wang2016}.
-
-As for the phase transitions, effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into account
-considering effective coefficient of thermal expansion. Similarly, latent heat due to melting/crystallisation can be implicitly included in the energy equation
-(Eq. \ref{eq:energy}) by considering effective specific heat ($C_{p_{eff}}$) and thermal expansion ($\alpha_{eff}$) of the partially molten rocks, in according to
-\begin{eqnarray}
-\label{eq:cp_effective}C_{p_{eff}}&=&C_p+H_L\left(\frac{\partial M}{\partial T}\right)_p \\
-\label{eq:alpha_effective}\alpha_{eff}&=&\alpha+\rho\frac{H_L}{T}\left(\frac{\partial M}{\partial p}\right)_T
-\end{eqnarray}
-where $C_p$ and $\alpha$ are the specific heat and the coefficient of thermal expansion of the solid rock, respectively, and $H_L$ is the latent heat of the
-molten rock \citep{Gerya2004b,Gerya2007,Tackley2010}.
-
-\section{Benchmarks}\label{sec:benchmark}
-\subsection{Error measurements}\label{sec:error}
-In order to determine the accuracy of velocity and pressure field of the benchmarks, the $L_2$-norm is computed by numerical integration on the Gauss-Legendre
-quadrature points. $L_2$-norm for pressure and velocity errors can be evaluated as
-\begin{eqnarray}
-\textrm{err}_p&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|p^n_i(r_q)-p^a_i(r_q)|^2\right)w_q|J_q|}\nonumber \\
-\textrm{err}_v&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|u^n_i(r_q)-u^a_i(r_q)|^2+|v^n_i(r_q)-v^a_i(r_q)|^2\right)w_q|J_q|}\nonumber
-\end{eqnarray}
-respectively, where $n_e$ is the number of elements, $n_q$ is the number of quadrature points per element, $p^n_i(r_q)$ and $p^a_i(r_q)$ are the numerical and
-analytical pressure, respectively, in each quadrature point $q$, $w_q$  and $J_q$ are the weight and the Jacobian at the quadrature point
-$q$, $u^n_i(r_q)$, $v^n_i(r_q)$, $u^a_i(r_q)$ and $v^a_i(r_q)$ are the numerical and analytical velocities, respectively, in each quadrature point $q$.
-
-Other quantities used as comparison with original benchmarks are the root-mean-square velocity over the whole domain and over the surface
-\begin{eqnarray}
-v_{\textrm{rms}}&=&\sqrt{\int_{\Omega}|\bm{v}|^2 d\Omega}\nonumber \\
-v_{\textrm{rms}}^{top}&=&\sqrt{\int_0^1 u^2 \bigg|_{y=1} dx}\nonumber
-\end{eqnarray}
-respectively, top and bottom Nusselt numbers
-\[\textrm{Nu}^{top/bottom}=-\int_0^1\frac{\partial T}{\partial y} \bigg|_{y=1/y=0} dx\]
-the average rate of work done against gravity
-\[\left\langle W \right\rangle=\int_{\Omega}T u_y d\Omega\]
-the average rate of viscous dissipation
-\[\left\langle \Phi \right\rangle=\int_{\Omega}\tau_{ij}\dot{\epsilon}_{ij} d\Omega\]
-and the percentage error between $\left\langle W \right\rangle$ and $\left\langle \Phi \right\rangle/Ra$
-\[\delta=\frac{\left|\left\langle W \right\rangle -\frac{\left\langle \Phi \right\rangle}{Ra} \right|}{\max \left(\left\langle W \right\rangle,
-\frac{\left\langle \Phi \right\rangle}{Ra}\right)}\times100\]
-
-\subsection{Stokes flow}\label{sec:stokes}
-The problem consists of determining velocity field $(u,v)$ and pressure $p$ in case of a manufactured solution with prescribed body forces such as
-\begin{eqnarray}
-b_1&=&(12-24y)x^4+(-24+48y)x^3+(-48y+72y^2-48y^3+12)x^2+\nonumber \\
-&&+(-2+24y-72y^2+48y^3)x+1-4y+12y^2-8y^3\nonumber \\
-b_2&=&(8-48y+48y^2)x^3+(-12+72y-72y^2)x^2+(4-24y+48y^2-48y^3+\nonumber \\
-&&+24y^4)x-12y^2+24y^3-12y^4\nonumber
-\end{eqnarray}
-for which the exact solution is:
-\begin{eqnarray}
-u(x,y)&=&x^2(1-x)^2(2y-6y^2+4y^3)\nonumber \\
-v(x,y)&=&-y^2(1-y)^2(2x-6x^2+4x^3)\nonumber \\
-p(x,y)&=&x(1-x)-\frac{1}{6}\nonumber
-\end{eqnarray}
-The domain is a unit square a constant viscosity ($\eta=1$) and the penalty parameter is set to $\lambda=10^7$. Velocity boundary conditions are set to no slip
-$(\bm{v}=\bm{0})$ on all boundaries. The problem is performed for different grid resolution between \num{8x8} and \num{1024x1024} elements.
-
-The errors between the analytical solution and the numerical prediction of the velocity and the pressure field are calculated as in Section \ref{sec:error}.
-Fig. \ref{fig:errors}a shows that both velocity and pressure field converge to the exact solution with the decrease of the element size, following the
-theoretical convergence rate. The convergence of pressure, in contrast with the observation by \citet{Donea2003} for $Q_1 \times P_0$ elements, indicate the
-effectiveness of the smoothing procedure. Smoothed pressure field for grid resolution of \num{128x128} elements is shown in Fig. \ref{fig:errors}b (continuous
-red line), in comparison with the analytical solution (dashed blue line). Solve times and memory usage needed to generate the solution are shown in
-Fig. \ref{fig:MUMPS}. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Solver/Stokes_Flow}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/errors.pdf}
-\caption{Velocity and pressure error for the Stokes flow experiment between generated and analytical solution as a function of element size (panel a),
-and comparison between smoothed pressure and analytical solution as function of $x$ coordinate for a grid resolution of \num{128x128} elements (panel b).}
-\label{fig:errors}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/MUMPS.pdf}
-\caption{Panel a) Analysis, factorisation, solution and total solve times (panel a) and factorisation memory usage (panel b) as a function of the total number
-of degrees of freedom for the Stokes flow experiment.}
-\label{fig:MUMPS}
-\end{figure}
-
-\subsection{Markers advection}\label{sec:runge}
-The 2nd-order Runge-Kutta advection scheme in space is tested by means of the Zalesak disk test \citep{Zalesak1979,Thieulot2014}. The benchmark is performed
-in a unit square domain with a grid resolution of \num{32x32} elements and 1000000 markers and values of Courant number between 0.25 and 1.
-At $t=0$, the disk is centred at position $(0.5;0.75)$ with a radius $R=0.15$ and has a vertical fissure 0.05 wide and 0.2 high (orange in Fig. \ref{fig:runge})
-and with the velocity field prescribed in the entire domain as
-\begin{eqnarray}
-u(x,y)&=&2\pi\left(y-\frac{L_x}{2}\right)\nonumber \\
-v(x,y)&=&-2\pi\left(x-\frac{L_x}{2}\right)\nonumber
-\end{eqnarray}
-the markers are back to their initial location after a $2\pi$ rotation ($t=1$). The clockwise rotation of the disk at $t=0.25$, 0.5 and 0.75 
-is shown in Fig. \ref{fig:runge} (red, blue and green, respectively). The distance from the centre of 1 marker is calculated throughout the rotation to
-evaluate the error for different values of Courant number. As expected, the error increase with the increasing of the time step, as shown in 
-Fig. \ref{fig:runge_err}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/Zalesak_Disk}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=200px]{./Figures/Zalesak.pdf}
-\caption{Position of the Zalesak disk at $t=0$, 0.25, 0.5 and 0.75 (orange, red, blue and green, respectively) for the case with Courant number of 1.}
-\label{fig:runge}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/zal_marker.pdf}
-\caption{Distance from the centre as function of time for values of Courant number of  of 0.25, 0.3, 0.5, 0.75 and 1 (green, orange, blue, black and red,
-respectively). R indicates the distance from the centre considering a perfect circle trajectory.}
-\label{fig:runge_err}
-\end{figure}
-
-\subsection{Conservative Velocity Interpolation (CVI)}\label{sec:CVI}
-The CVI correction is checked by means of the Stokes flow experiment by \citet{Donea2003}, as presented in \ref{sec:stokes}. The advection of Lagrangian
-markers is performed using either a 2nd-order or a 4th-order Runge-Kutta scheme with CFL=0.5 and an initial random distribution of 25 markers per element.
-Fig. \ref{fig:CVI} shows the minimum and maximum values of markers per element throughout the experiment with and without the CVI correction and using
-different Runge-Kutta schemes. The introduction of the CVI correction allows the
-tests to run for times longer than $t=2000$ without show empty elements, while tests without the CVI correction stop at approximately $t=600$, when at least
-one element goes empty. Fig. \ref{fig:CVI_time} shows the different number of markers per element and the markers distribution at $t=600$ obtained without and
-with the CVI correction (panels b and c, respectively) in case of a 2nd-order Runge-Kutta scheme. All data can be found at
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/CVI}.
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/CVI.pdf}
-\caption{Maximum and minimum number of markers per element throughout the experiment, with (green and orange lines) or without (purple and blue lines)
-the CVI correction and using either a 2nd-order (blue and orange lines) or a 4th-order (purple and green lines) Runge-Kutta scheme.}
-\label{fig:CVI}
-\end{figure}
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/CVI_time.pdf}
-\caption{Number of markers per element and markers distribution through time using 2nd.order Runge-Kutta scheme.
-Panels a: number of markers per element (panel a1) and distribution of the markers (panel a2) after the first
-iteration; panels b and c: comparison between number of markers per element (panels b1 and c1) and markers distribution (panels b2 and c2) without and with
-the CVI correction at $t=600$; panels d: number of markers per element (panel d1) and markers distribution (panel d2) in case of CVI correction at $t=2000$.}
-\label{fig:CVI_time}
-\end{figure}
-
-\subsection{Poiseuille flow}\label{sec:poiseuille}
-The domain is a rectangle with $L_x=2$ and $L_y=1$ and constant density and viscosity ($\rho=1$ and $\eta=1$), gravity acceleration $\bm{g}=\bm{0}$ and penalty
-parameter $\lambda=10^8$. The grid is composed by \num{40x20} elements. Velocity boundary conditions are set to no slip $(\bm{v}=\bm{0})$ at the top and the
-bottom, and a parabolic profile is imposed on the sides, with $u=y(1-y)$ and $v=0$. The analytical solution is then given by:
-\begin{eqnarray}
-u(x,y)&=&y(1-y)\nonumber \\
-v(x,y,)&=&0\nonumber \\
-p(x,y)&=&2\eta\left(\frac{L_x}{2}-x\right)\nonumber
-\end{eqnarray}
-
-The velocity field predicted by the model follows the expected parabolic profile (Fig. \ref{fig:poiseuille}a and Fig. \ref{fig:poi_plot}a) and, in case of the
-classic penalty method with no iterations, the pressure is clearly related to the divergence by means of the penalty parameter (Fig. \ref{fig:poiseuille}b and
-c and Fig. \ref{fig:poi_plot}b). Fig. \ref{fig:poiseuille}d shows that one Uzawa iteration is sufficient to bring the divergence down to \num{e-15}, with no
-correlation with the pressure field. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Poiseuille%20Flow}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Poiseuille.pdf}
-\caption{Velocity field (panel a), pressure (panel b) and divergence velocity of a Poiseuille flow in case of the classic penalty method (no iterations) and
-after one Uzawa iteration (panel c and d, respectively).}
-\label{fig:poiseuille}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Analytical.pdf}
-\caption{Velocity (panel a) and pressure (panel b) field predicted by the model for a Poiseuille flow with respect with their analytical solutions. The
-velocity field is plotted as function of the vertical coordinate in $L_x/2$ and the pressure is plotted as function of the x coordinate in $L_y/2$.}
-\label{fig:poi_plot}
-\end{figure}
-
-\subsection{Instantaneous 2D sphere}\label{sec:ist_sphere}
-The domain is a unit square with gravity $\bm{g}=(0,-1)$. The fluid has constant density and viscosity ($\rho_f=1$ and $\eta_f=1$). The sphere is in the middle
-of the domain with a radius R = 0.123456798 and has constant density and viscosity ($\rho_s=10^{-2}+\rho_f$ and $\eta_s=10^3 \cdot \eta_f$).
-We distinguish three different types of velocity boundary conditions:
-\begin{enumerate}
-\item FS: free slip conditions on all sides.
-\item NS: no slip conditions on all sides.
-\item OT: free slip conditions on the sides and the bottom, and open boundary at the top.
-\end{enumerate}
-All three cases are performed for different grid resolutions, between \num{16x16} and \num{512x512} elements with 50 randomly distributed markers per element,
-and different average schemes for the viscosity (harmonic, geometric and arithmetic). The velocity in the centre of the sphere $\bm{v}(0.5;0.5)$), minimum and
-maximum velocities ($u_{min}$, $u_{max}$, $v_{min}$, $v_{max}$) and pressures ($p_{min}$, $p_{max}$) on the entire domain, $v_{\textrm{rms}}$ and average pressure on the
-entire domain ($p_{avg}$) are compared with solutions generated by ASPECT \citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and all results
-can be found at \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere2D} and 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Instantaneous_2D_Sphere}. Results in terms of the $v_{\textrm{rms}}$ for NS model 
-are also shown in Fig. \ref{fig:inst_sphere}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_NS.pdf}
-\caption{$v_{\textrm{rms}}$ as function of element size for the instantaneous 2D sphere experiment in case of no slip boundary conditions and with different average
-schemes for the viscosity. Results are compared with results obtained by other numerical codes.}
-\label{fig:inst_sphere}
-\end{figure}
-
-\subsection{Rayleigh-Taylor instability}\label{sec:rayleigh}
-This problem has been originally presented by \citet{vanKeken1997} and here it is performed as the isoviscous case (Case 1a). The domain has $L_x=0.9142$ and
-$L_y=1$ and gravity $\bm{g}=(0,-1)$. Two fluids with same constant viscosities ($\eta_1=\eta_2=1$) and different densities ($\rho_1=1000$ and $\rho_2=1010$),
-with the lighter fluid at the bottom. The initial interface between the fluids is given by $y(x)=0.2+0.02 \cos \left(\frac{\pi x}{L_x}\right)$. The experiment
-is performed with different grid sizes (\num{50x50}, \num{80x80}, \num{100x100} and \num{256x256}). A total of 1960000 markers are randomly distributed at the
-beginning of the simulation. Velocity boundary conditions are set to no slip at the top and the bottom, and to free slip at the sides of the domain.
-
-$v_{\textrm{rms}}$ as function of time is reported in Fig. \ref{fig:RT}, matching well with results shown by \citet{vanKeken1997}, \citet{Tackley2003} and
-\citet{Thieulot2014}. Fig. \ref{fig:rayleigh} shows the evolution of the experiment at different time steps. All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Rayleigh_Taylor_experiment/ISOVISCOUS}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/RT.pdf}
-\caption{$v_{\textrm{rms}}$ of the Rayleigh-Taylor experiment as function of time for different resolution of the grid.}
-\label{fig:RT}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=300px]{./Figures/Rayleigh.pdf}
-\caption{Evolution of the Rayleigh-Taylor experiment for a grid of \num{256x256} elements at $t=0, 500, 1000$ and $2000$ (panels a, b, c and d, respectively).}
-\label{fig:rayleigh}
-\end{figure}
-
-\subsection{Falling blocks}\label{sec:block}
-This benchmark is described as proposed by \citet{Gerya2003a}, \citet{Gerya2010b} and \citet{Thieulot2011}. The domain is square with $L_x=L_y=$ \SI{500}{\km}
-and the grid is composed by \num{50x50} elements with 25 markers in each element. The block is initially centred at ($x=$ \SI{250}{\km}; $y=$ \SI{400}{\km})
-and has a size of \num{100x100} km (Fig. \ref{fig:block}a). The fluid surrounding the block has $\eta_f=$ \SI{e21}{\pascal\per\s} and $\rho_f=$
-\SI{3200}{\kg\per\cubic\m}. The benchmark tests with different viscosities of the block, with $\eta_b$ from \SIrange{e15}{e27}{\pascal\per\s}. In each
-experiment also the density $\rho_b$ of the block varies from \SIrange{3220}{9900}{\kg\per\cubic\m}. Velocity boundary conditions are set to free slip
-conditions on all sides of the domain. Density distributions at $t=$ \SI{20}{\mega\year} for $\rho_b=$ \SI{3300}{\kg\per\cubic\m} and different $\eta_b$
-are plotted in Fig. \ref{fig:block}b-f, showing that the code correctly preserve the block geometry in case of large viscosity contrast, with a block stiffer
-than the surrounding fluid (Fig. \ref{fig:block}f).
-
-The velocity in the centre of the falling block at $t=0$ is measured for all experiments and, since the velocity should increase with density contrast, the
-quantity $\bm{v}/(\rho_b-\rho_f)$ is plotted as function of the viscosity contrast. All results perfectly match with those from \citet{Gerya2010b} and they
-line up on a single curve, demonstrating that the code can correctly deal with large viscosity and density contrasts (Fig. \ref{fig:falling}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Falling%20blocks}.
-
-\begin{figure}
-\noindent\includegraphics[width=300px]{./Figures/Block.pdf}
-\centering
-\caption{Density evolution of the falling block experiment at $t=$ \SI{0}{\mega\year} (panel a) and $t=$ \SI{20}{\mega\year} for different viscosities of the
-block (panels b-f).}
-\label{fig:block}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Falling.pdf}
-\caption{Initial velocity relative to the density contrast at the centre of the falling block as function of the viscosity contrast between the block and the
-surrounding fluid.}
-\label{fig:falling}
-\end{figure}
-
-\subsection{2D time-dependent Stokes sphere with deformable free surface}\label{sec:sphere}
-This experiment is performed in a unit square domain with the gravity acceleration fixed to $g_y=-1$. The sphere has $\rho_s=2$ and $\eta_s=10^3$ and is
-initially centred at ($0.5;0.6$) with radius $R=0.123456789$. The fluid surrounding the sphere has $\rho_f=1$ and $\eta_f=1$ and occupies the domain for
-$y\leq 0.75$, while for $y>0.75$ the air has $\rho_a=0$ and $\eta_a=10^{-3}$. Velocity boundary conditions are set to free slip on all sides. The Courant
-number is set to 0.25. The experiments run for 200 s using grid resolutions from \num{150x150} to \num{512x512} elements, with 25 randomly distributed markers
-per element, and for different average schemes for the viscosity. The interface between the fluid and the air is tracked by means of the markers chain. Results
-in terms of velocity and pressure fields and topography variations are compared with results obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and they can be found at
-\url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere_fs2D} and
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Time_dependent_sphere}. Results in terms of the $v_{\textrm{rms}}$ in case of 
-an arithmetic average are also shown in Fig. \ref{fig:inst_spherefs}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_arithm.pdf}
-\caption{$v_{\textrm{rms}}$ for different grid resolutions as function of time for the 2D time-dependent Stokes sphere experiment in case of an arithmetic average.
-Results are compared with results obtained by ASPECT.}
-\label{fig:inst_spherefs}
-\end{figure}
-
-\subsection{Free surface stabilisation}\label{sec:stab}
-The benchmark is performed as discussed in \citet{Kaus2010a} and \citet{Thieulot2014}. The domain is square with $L_x=L_y=$ \SI{500}{\km} and a grid resolution
-of \num{200x200} elements, each of them containing 16 markers. A buoyant fluid with $\rho_1=$ \SI{3200}{\kg\per\cubic\metre} and $\eta_1=$ \SI{e20}{\pascal\s}
-is overlain by a denser fluid with $\rho_2=$ \SI{3300}{\kg\per\cubic\metre} and $\eta_2=$ \SI{e21}{\pascal\s}. The initial interface between the fluids has an
-initial sinusoidal shape of 5 km amplitude. Velocity boundary conditions are set to free slip conditions at the sides and no slip at the bottom, while the top
-is free surface. The experiment is performed with various fixed time steps, without the use of the Courant number. The vertical position of the free surface at
-$x=L_x$ is tracked for each simulation, with or without the stabilisation algorithm.
-
-The results show an instability (drunken sailor effect) increasing the time step in case that the stabilisation algorithm is not activated
-($dt=$ \SI{4200}{\year}, green line in Fig. \ref{fig:stab}), as already observed by \citet{Kaus2010a} and \citet{Thieulot2014}. Activating the algorithm the
-instability is fixed and the simulation remains stable using time step up to \SI{50000}{\year} (red line in Fig. \ref{fig:stab}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Stabilisation_algorithm}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Kaus_random.pdf}
-\caption{Evolution of the y coordinate in $x=L_x$ as function of time for different time steps, with and without the stabilisation algorithm.}
-\label{fig:stab}
-\end{figure}
-
-\subsection{Topography relaxation}\label{sec:crameri}
-This experiment is performed as in \citet{Crameri2012} using free surface and sticky air (Case 1). The domain is rectangular with $L_x=$ \SI{2800}{\km} and
-$L_y=$ \SI{700}{\km} in case of free surface and $L_y=$ \SI{800}{\km} in case of sticky air ($h_{st}=$ \SI{100}{\km}). The grid is composed by \num{256x64}
-elements for the free surface \num{560x320} elements for the sticky air. For both cases, there is a mantle of 600 km thickness with $\eta_m=$ \SI{e21}{\pascal\s}
-overlain by a highly viscous cosine-shaped lithosphere with thickness between 93 and 107 km and $\eta_l=$ \SI{e23}{\pascal\s}. Both mediums have $\rho=$
-\SI{3300}{\kg\per\cubic\metre}. In case of sticky air experiment, the air has $\rho_a=$ \SI{0}{\kg\per\cubic\metre} and $\eta_a=$ \SI{e18}{\pascal\s}. The
-gravity acceleration is set to $g_y=$ \SI{-10}{\metre\per\square\s}. Velocity boundary conditions are set to no slip at the bottom and free slip at the sides
-of the domain. In case of sticky air, velocities on top are set to free slip conditions.
-
-The maximum topography as function of time ($h(t)$) can be analytically derived in according to \[h(t)=h_0 \exp(\gamma t)\]
-where $h_0=7$ km is the initial topography, $\gamma=\num{-0.2139e-11}$ is the characteristic relaxation rate and $t$ is the time (black line in Fig.
-\ref{fig:crameri}). Maximum topography at the characteristic relaxation time $t_{rlx}=$ \SI{14.825}{\kilo\year} can be found to be $h_{rlx}=$ \SI{2576}{\m}.
-The results show that a viscosity of the air of \SI{e18}{\pascal\s} is needed in case of $h_{st}=$ \SI{100}{\km} to correctly describe the topography relaxation
-for this problem, as pointed out by \citet{Crameri2012} (purple line in Fig. \ref{fig:crameri}). The case with a true free surface well follows the analytical
-solution (orange line in Fig. \ref{fig:crameri}). In particular, at the relaxation time in case of the true free surface predicts a maximum topography of
-\SI{2570}{\m}, with an error of \SI{6}{\m} with respect to the analytical solution. Results are also compared with results obtained with a free surface in
-UNDERWORLD (\num{256x64} elements) and SULEC (\num{401x201} elements) (with courtesy from \citealp{Crameri2012}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Topography_relaxation}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Crameri.pdf}
-\caption{Maximum topography as function of time for the Crameri benchmark (coloured lines) in comparison with the analytical solution (black line) and results
-obtained with UNDERWORLD (yellow line) and SULEC (blue line).}
-\label{fig:crameri}
-\end{figure}
-
-\subsection{Spontaneous subduction}\label{sec:subduction}
-This experiment is performed as presented by \citet{Schmeling2008}, in case of both a sticky air and a true free surface and using different average schemes
-for the viscosities. The domain is rectangular with $L_x=$ \SI{3000}{\km} and $L_y=$ \SI{700}{\km}, a grid resolution of \num{256x64} elements with 50 markers
-per element and Courant number of 0.01. At the beginning of the simulation, a 100 km-thick lithospheric layer with $\rho_m=$ \SI{3300}{\kg\per\cubic\metre}
-and $\eta_m=$ \SI{e23}{\pascal\s} is located at the top of the domain for $x$ between 1000 and 3000 km, while the asthenospheric mantle has $\rho_m=$
-\SI{3200}{\pascal\s} and $\eta_m=$ \SI{e21}{\pascal\s}. In addition, a 200 km-depth lithospheric slab is already subducted in the mantle in order to have
-a spontaneous subduction. In case of sticky air, $L_y=750$ km and a 50 km-thick air layer with $\eta_a$ from \SIrange{e19}{e21}{\pascal\s} overlie the mantle.
-Velocity boundary conditions are set to free slip on the sides and on the bottom of the domain. In case of sticky air, velocities on the top boundary are set
-to free slip conditions as well.
-
-The maximum depth of the slab is tracked for all the simulations (Fig. \ref{fig:subduction}) and they are compared with results obtained by
-\citet{Schmeling2008} with models with regular grids and comparable grid resolutions (rectangles in Fig. \ref{fig:subduction}). Low viscous sticky air models
-(i.e., \SI{e19}{\pascal\s}) enlighten a strong dependence of the sinking velocity with the chosen average scheme (dashed lines in Fig. \ref{fig:subduction}),
-while high viscous sticky air models (i.e., \SI{e21}{\pascal\s}) show a higher resistance at the trench, underestimating the correct solution (dotted lines in
-Fig. \ref{fig:subduction}). A high resistance at the trench can be also observed in case of a true free surface for the low resolution grid used (continuous
-lines in Fig. \ref{fig:subduction}), as pointed out by \citet{Schmeling2008}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Spontaneous_subduction}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Subduction.pdf}
-\caption{Maximum depth of the slab as function of time for all the simulations of the \citet{Schmeling2008} benchmark, in case of harmonic, geometric and
-arithmetic means (red, green and blue lines, respectively) using a sticky air or a true free surface (discontinuous and continuous lines, respectively).
-Red, green and blue rectangular areas indicate the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{500}{\km} in case of sticky air,
-while the grey rectangular area indicates the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{600}{\km} in case of true free surface.}
-\label{fig:subduction}
-\end{figure}
-
-\subsection{The slab detachment}\label{sec:slab}
-The slab detachment benchmark is performed as described by \citet{Schmalholz2011} and \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{1000}{\km}
-and $L_y=$ \SI{660}{\km} and a grid resolution of \num{256x256} elements. A non-linear viscous T-shaped layer with $\rho_l=$ \SI{3300}{\kg\per\cubic\metre} is
-placed at the top of the domain and surrounded by a linear viscous fluid with $\rho_f=$ \SI{3150}{\kg\per\cubic\m}. The top layer is 80 km-thick and a 250
-km-long and 80 km-wide slab is placed at $x=L_x/2$ (Fig. \ref{fig:slab}a). The effective viscosity of the top layer is given by
-\[\eta_{eff}=\eta_0 I_2^{\frac{(1-n)}{n}}\]
-with $\eta_0=$ \SI{4.75e11}{\pascal\s} and $n=4$, while the surrounding fluid has $\eta_f=$ \SI{e21}{\pascal\s}. The effective viscosity is capped between
-$\eta_{min}=$ \SI{e21}{\pascal\s} and $\eta_{max}=$ \SI{e25}{\pascal\s}. Velocity boundary conditions are set to free slip at the top and the bottom, and to
-no slip at the sides of the domain.
-
-The sinking of the slab determines a decrease of the effective viscosity in the T-shaped layer during the first part of the evolution (Fig. \ref{fig:slab}a),
-as consequence of local strain rates, while effective viscosities increase up to $\eta_{max}$ in the top layer after approximately \SI{20}{\mega\year}, when
-necking is complete (Fig. \ref{fig:slab}b).
-
-The width of necking is tracked by means of markers position at the side of the slab. Necking width and time are normalised in function of the initial slab
-(\SI{80}{\km}) and the characteristic time ($t_c=$ \SI{7.1158e14}{\s}) \citep{Schmalholz2011,Glerum2018}. The results in case of arithmetic, geometric and
-harmonic mean of the viscosity are shown in Fig. \ref{fig:necking}, compared with results from \citet{Glerum2018} for which an infinity norm mean was used.
-The influence of the average schemes agrees with results shown in Section \ref{sec:subduction}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Slab_detachment}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Slab.pdf}
-\caption{Effective viscosity for the slab detachment benchmark at the beginning of the evolution (panel a) and when necking is complete (panel b).}
-\label{fig:slab}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Necking.pdf}
-\caption{Normalised width of the necking of the slab detachment benchmark as function of normalised time.}
-\label{fig:necking}
-\end{figure}
-
-\subsection{The indenter experiment}\label{sec:indenter}
-The indenter benchmark simulate a rigid punch on a purely plastic von Mises material, for which exists an analytical solution
-\citep{Thieulot2008,Thieulot2014,Glerum2018}. The domain is a unit square with a grid resolution of \num{256x256} elements. The gravity acceleration is set
-to $\bm{g}=\bm{0}$ and effective viscosity is capped using $\eta_{min}=10^{-4}$ and $\eta_{max}=10^3$. The tolerance for the convergence of the non-linear
-solution is $tol=10^{-9}$, with a maximum number of non-linear iterations set to 500. The medium in the domain has $\rho=0.01$, an initial viscosity
-$\eta_0=10$, cohesion $C=1$ and an angle of internal friction $\phi=0$°. Velocity boundary conditions are set to no slip at the bottom and free slip at the
-sides of the domain. The top of the domain is open with the exception of the central portion (punch area) with width $w_p=0.125$, where $v=-1.05$ and $u$ is
-fixed either to 0 (rough punch experiment) or free (smooth punch experiment).
-
-Fig. \ref{fig:indenter} show the results in terms of viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure
-(panels g and h) for both experiments. The analytical solution indicate that pressure at ($0.5;1$) and ($0.5 \pm w_p;1$)is $p=\pi +1$ and $p=1$, respectively
-(black and red lines in Fig. \ref{fig:smooth_p}). Fig. \ref{fig:smooth_p} shows a clear improvement in the pressure solution when passing from a rough to a
-smooth punch (green and blue lines in Fig. \ref{fig:smooth_p}, respectively), as previously observed by \citet{Thieulot2014} and \citet{Glerum2018}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Indenter}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Indenter.pdf}
-\caption{Viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure (panels g and h) fields for rough (left column) and
-smooth (right column) punch experiments.}
-\label{fig:indenter}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Smooth_Rough.pdf}
-\caption{Nodal pressure in function of the x coordinate at $y=1$ for rough (green line) and smooth (blue line) punch experiment. Black and red lines indicate
-the analytical solution at $x=0.5$ and $x=0.5 \pm w_p$, respectively.}
-\label{fig:smooth_p}
-\end{figure}
-
-\subsection{The brick experiment}\label{sec:brick}
-An instantaneous version of the brick benchmark is performed to verify the correctness of the pressure-dependent plasticity for different angle of internal
-friction, as proposed by \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{40}{\km} and $L_y=$ \SI{10}{\km} and a grid resolution of \num{512x128}
-elements. The gravity acceleration is set to $g_y=$ \SI{-10}{\m\square\s} and effective viscosity is capped using $\eta_{min}=$ \SI{e19}{\pascal\s} and
-$\eta_{max}=$ \SI{e26}{\pascal\s}. The tolerance for the convergence of the non-linear solution is $tol=10^{-7}$, with a maximum number of non-linear iterations
-set to 1000. A 800 m-wide and 400 m-height inclusion with $\eta_b=$ \SI{e20}{\pascal\s} and $\rho_b=$ \SI{2700}{\kg\per\cubic\m} is placed at the bottom of the
-domain at $x=L_x/2$. The inclusion is surrounded by a non-linear viscous medium with $\rho_m=$ \SI{2700}{\kg\per\cubic\m}, an initial viscosity of
-\SI{e23}{\pascal\s}, a linear viscous viscosity of \SI{e25}{\pascal\s} and a cohesion of \SI{40}{\mega\pascal}. Velocities are set to free slip conditions at
-the bottom of the domain and the top is open. Velocities on sides of the domain are fixed to $u=$ \SI{+-2e-11}{\m\per\s} and $v=0$. The experiment is performed
-in compressional and extensional contexts with an angle of internal friction $\phi$ of the non-linear medium variable between 0° and 30°.
-
-As expected, two shear bands stem from the inclusion with variable angles in relation to both the dynamics context and the internal friction angle. Shear band
-angles formed at 45° for $\phi=0$° in both extensional and compressional contexts (Fig. \ref{fig:brick_beam}a), while they are different in case of $\phi\neq0$°.
-Shear band angles in case of $\phi=20$° are shown for compression and extension in Fig. \ref{fig:brick_beam}b  and c, respectively. Values of shear band angles
-as function of internal friction angles from 0° to 30° are extracted at different depths and minimum and maximum values are plotted in Fig. \ref{fig:brick},
-in comparison with theoretical Roscoe, Arthur and Coulomb shear band angles. For all tests the tolerance for nonlinear convergence is defined $tol=10^{-7}$
-and none converges before the maximum number of iterations ($it_{max}=1000$) is reached. However, tests with internal friction angles up to 15° show a constant
-decrease of the velocity residuals, which can not be observed for for tests with higher internal friction angles (Fig. \ref{fig:convergence}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Brick_experiment}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Brick_Beam.pdf}
-\caption{Shear band angles predicted for the brick experiment for $\phi=0$° (panel a) and for $\phi=20$° in case of compression and extension (panels b and c,
-respectively).}
-\label{fig:brick_beam}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Brick.pdf}
-\caption{Shear band angles predicted for the brick experiment in case of compressional and extensional contexts as function of different internal angle of
-friction (continuous black line and red dots), compared with theoretical Roscoe, Artur and Coulomb angles (discontinuous black lines). Red lines indicate the
-range of angles calculated at different depths.}
-\label{fig:brick}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \noindent\includegraphics[width=500px]{./Figures/convergence.pdf}
-  \caption{Convergence rates for all tests. Normalised residuals (left column) are compared with velocity residuals (right column) for extensional (first row)
-  and compressional (second row) settings.}
-  \label{fig:convergence}
-  \end{figure}
-
-\subsection{Advection stabilisation}\label{sec:advection}
-The 1D advection problem proposed by \citet{Donea2003} and \citet{Thieulot2011} is performed to verify the effectiveness of the SUPG method to stabilise the
-advection term of the energy equation. The domain is a 1D segment with $L_x=1$ composed by 50 elements and a discontinuity in the thermal field placed at
-$x=0.25$. Temperature is set to 1 for $x \leq 0.25$ and to 0 for $x > 0.25$. Velocity is set to ${u}=1$ in the entire domain. The simulation is performed for
-250 time steps, with $dt=0.002$, so the thermal discontinuity should be at $x=0.75$ at the end of the simulation.
-
-Temperature profiles at the end of the simulation are shown in \ref{fig:advection} as function of the dimensionless coefficient $\gamma=\tau \bm{v}/h$. In case
-of the classic Galerkin method ($\gamma=0$, blue line in \ref{fig:advection}) the final thermal profile is characterised by strong oscillations, which are
-eliminated in case of the SUPG method ($\gamma=0.045$, orange line in \ref{fig:advection}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Advection_stabilisation}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Advection.pdf}
-\caption{Temperature profile as function of \textit{x} for the advection stabilisation benchmark. Purple line indicates the initial temperature profile; the
-green line indicates the analytical temperature profile after 250 time steps; blue line indicates the temperature profile after 250 time steps in case of the
-classic Galerkin method ($\gamma=0$); orange line indicates the temperature profile after 250 time steps in case of the SUPG method ($\gamma=0.045$).}
-\label{fig:advection}
-\end{figure}
-
-\subsection{Shear heating}\label{sec:simple_shear}
-This exercise is performed in an unit square domain composed by \num{128x128} elements. The velocity field is prescribed on the entire domain with
-$\bm{v}=(L_y-y)y\bm{e_x}$; viscosity, density and specific heat are fixed to 1, while thermal conductivity and radiogenic energy are fixed to 0. Therefore,
-the energy equation (Eq. \ref{eq:energy}) can be simplified as \[\frac{\partial T}{\partial t}=H_s\] and fixing $T(t=0)=0$, the temperature can be find as
-\[T(t)=H_s t\]
-In this case we have
-\begin{eqnarray}
-\dot{\epsilon_{xx}}&=&\frac{\partial u}{\partial x}=0 \nonumber \\
-\dot{\epsilon_{yy}}&=&\frac{\partial v}{\partial y}=0 \nonumber \\
-\dot{\epsilon_{xy}}&=&\frac{1}{2}\left(\frac{\partial v}{\partial x}+\frac{\partial u}{\partial y}\right)=\frac{1}{2}(L_y-2y)\nonumber
-\end{eqnarray}
-and, simplifying Eq. \ref{eq:shear}, shear heating can be calculated as \[H_s(x,y)=(1-2y)^2\]
-
-The solution predicted by the model in terms of velocity, temperature and shear heating match well with the analytical solutions (Fig. \ref{fig:analytical_en}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Simple_shear}.
-
-%\begin{figure}
-%\centering
-%\noindent\includegraphics[width=400px]{./Figures/Energy_Thieulot.pdf}
-%\caption{Temperature and shear heating predicted by the model after 1 time step.}
-%\label{fig:en_thieulot}
-%\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=500px]{./Figures/Analytical_Shear.pdf}
-\caption{Velocity (panel a), temperature (panel b) and shear heating (panel c) as function of the \textit{y} coordinate for the simple shear experiment. The
-solutions predicted by the model (red lines) are compared with the analytical solutions (dashed blue lines).}
-\label{fig:analytical_en}
-\end{figure}
-
-\subsection{Shear and adiabatic heating}\label{sec:shear}
-This problem is performed as presented in Exercise 9.4 in \citet{Gerya2010b}. Two materials are vertically separated in a rectangular domain with $L_x=$
-\SI{1000}{\km}, $L_y=$ \SI{1500}{\km} and a grid resolution of \num{30x20} elements. Constant thermal coefficient expansion ($\alpha=$ \SI{3e-5}{\per\kelvin}),
-temperature ($T=$ \SI{1300}{\kelvin}) and gravity acceleration ($g_y=$ \SI{-10}{\m\square\s}) are assumed in the whole domain. Fluid 1 (on the left side of the
-domain) has $\rho_1=$ \SI{3200}{\kg\per\cubic\m} and $\eta_1=$ \SI{e20}{\pascal\s}; fluid 2 (on the right side of the domain) has $\rho_2=$
-\SI{3300}{\kg\per\cubic\m} and $\eta_2=$ \SI{e22}{\pascal\s}. Velocity boundary conditions are set to free slip on all sides of the domain.
-
-As shown in Fig. \ref{fig:energy}, both shear and adiabatic heating predicted by the code (first row in Fig. \ref{fig:energy}) well recreate the results
-obtained by means of example ${Shear\_adiabatic\_heating.m}$ from \citet{Gerya2010b} (second row in Fig. \ref{fig:energy}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Adiabatic%2BShear}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Energy.pdf}
-\caption{Comparison between shear (first column), adiabatic (second column) and total (third column) energy predicted by the code (panels a, b and c) and those
-created using example ${Shear\_adiabatic\_heating.m}$ from Exercise 9.4 in \citet{Gerya2010b} (panels d, e and f).}
-\label{fig:energy}
-\end{figure}
-
-\subsection{Mantle convection}\label{sec:mantle}
-This problem is performed as presented by \citet{Blankenbach1989} (constant viscosity cases) and \citet{Thieulot2014}, in a 2D unit square domain with gravity
-acceleration $g_y=10^{10}Ra$. The experiment is performed with three different Rayleigh numbers ($Ra=10^4$, $10^5$ and $10^6$) and with different grid
-resolution (between \num{32x32} and \num{128x128} elements). The fluid has constant viscosity, initial density, heat capacity, thermal conductivity
-($\eta= \rho_0= C_p = k=1$), reference temperature ($T_0=0$) and thermal expansion coefficient ($\alpha=10^{-10}$). Temperatures are set to 0 on top and 1 on
-bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\]
-
-The solution generated by the code in terms of $v_{\textrm{rms}}$ (calculated as in Sec. \ref{sec:error}) and Nu as function of time are reported for all the
-simulations in Fig. \ref{fig:mantle}. At the steady state, $v_{\textrm{rms}}$ and Nu of all simulations converge well toward the values from \citet{Blankenbach1989},
-with lower errors for higher resolution grids (Table \ref{tab:mantle}). All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Mantle_convection}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Convection.pdf}
-\caption{$v_{\textrm{rms}}$ (panels a, c and e) and Nu (panels b, d and f) for the mantle convection benchmark as function of time for different grid resolution.
-Panels a and b show the results for $Ra=10^4$; Panels c and d show the results for $Ra=10^5$; panels e and f show the results for $Ra=10^6$. Purple lines
-indicate the convergence values for the $v_{\textrm{rms}}$ and the Nu from \citet{Blankenbach1989} at the steady state.}
-\label{fig:mantle}
-\end{figure}
-
-\begin{table}[H]
-\caption{Comparison between $v_{\textrm{rms}}$ and Nu predicted by the code for the mantle convection experiment and same values as reported in literature.}
-\centering
-\small
-\begin{tabular}{|>{\centering}m{0.09\textwidth}|>{\centering}m{0.05\textwidth}|>{\centering}m{0.24\textwidth}|>{\centering}m{0.15\textwidth}|
-  >{\centering}m{0.1\textwidth} >{\centering}m{0.1\textwidth} >{\centering\arraybackslash}m{0.1\textwidth}|}
-\toprule
-& & \multirow{2}{*}{\citet{Blankenbach1989}} & \citet{Thieulot2014} & \multicolumn{3}{c|}{FALCON}  \\
- & & & \num{200x200} & \num{32x32} & \num{64x64} & \num{128x128}  \\
-\midrule
-\multirow{2}{*}{$Ra=10^4$} & $v_{\textrm{rms}}$ & $42.864947 \pm 0.000020$ & 42.867 & 42.83226 & 42.852793 & 42.861394  \\
-           & Nu      & $4.884409 \pm 0.000010$  & 4.882  & 4.781297 & 4.857475  & 4.877573   \\
-\hline
-\multirow{2}{*}{$Ra=10^5$} & $v_{\textrm{rms}}$ & $193.21454 \pm 0.00010$  & 193.255& 193.872643 & 193.377472 & 193.252290  \\
-           & Nu      & $10.534095 \pm 0.000010$ & 10.507 & 9.602514 & 10.270735  & 10.465629   \\
-\hline
-\multirow{2}{*}{$Ra=10^6$} & $v_{\textrm{rms}}$ & $833.98977 \pm 0.00020$  & 834.712& 848.091176 & 837.767911 & 834.945793  \\
-           & Nu      & $21.972465 \pm 0.000020$ & 21.695 & 15.999266 & 19.703682  & 21.306939   \\
-\bottomrule
-\end{tabular}
-\label{tab:mantle}
-\end{table}
-
-\subsection{Viscoplastic mantle convection}\label{sec:visco_mantle}
-This benchmark is performed as Cases 1-5a described in \citet{Tosi2015} in a 2D unit square domain with two different grid resolutions of \num{32x32} 
-and \num{100x100} elements. Reference viscosity, reference density, heat capacity, thermal conductivity and thermal expansion coefficient are fixed to a constant
-value of 1 ($\eta_0= \rho_0= C_p = k=\alpha=1$), while the gravity has been chosen as $g_y=10^2$, in order to obtain a Rayleigh number $Ra=10^2$. Temperatures
-are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\].
-The viscosity field $\eta$ is calculated as the harmonic average between a linear part $\eta_{lin}$ that depends on temperature only or on temperature and 
-depth $y$ and a nonlinear-plastic part $\eta_{pl}$ that depends on the strain rate $\dot{\bm{\epsilon}}$, as follows
-\[\eta(T,y,\dot{\bm{\epsilon}})=2\left(\frac{1}{\eta_{lin}(T,y)}+\frac{1}{\eta_{pl}(\dot{\bm{\epsilon}})}\right)^{-1}\]
-The linear and the nonlinear-plastic parts of the viscosity are calculated following \citet{Tosi2015} as follows
-\[\eta_{lin}(T,y)=\exp(-\ln(\Delta\eta_T)T+\ln(\Delta\eta_y)y)\]
-\[\eta_{pl}(\dot{\bm{\epsilon}})=\eta^*+\frac{\sigma_Y}{\sqrt{\dot{\bm{\epsilon}}:\dot{\bm{\epsilon}}}}\]
-where $\Delta\eta_T$, $\Delta\eta_y$, $\eta^*$ and $\sigma_Y$ are parameters chosen as listed in Table \ref{tab:tosi_setup}.
-
-For each case, some diagnostic quantities measured at the steady state are reported in Table \ref{tab:visco_mantle} (see Section \ref{sec:error} for 
-the computation of all the quantities), showing a well fit with results described in \citet{Tosi2015}. Data and charts for all cases can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Viscoplastic_mantle_convection}.
-
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{white!100!}{gray!25!}
-%\begin{table}[H]
-%  \caption{Results of the viscoplastic mantle convection benchmark for Cases 1-4 compared with results obtained with ELEFANT and YACC (from \citealp{Tosi2015}).}
-%  \centering
-%  \resizebox{\textwidth}{!}{
-%\begin{tabular}{|c|ccc|ccc|ccc|ccc|}
-%  \toprule
-%    & \multicolumn{3}{c|}{\textbf{Case 1}} & \multicolumn{3}{c|}{\textbf{Case 2}} & \multicolumn{3}{c|}{\textbf{Case 3}} &
-%    \multicolumn{3}{c|}{\textbf{Case 4}} \\
-%   & FALCON & ELEFANT& YACC & FALCON& ELEFANT& YACC & FALCON & ELEFANT & YACC & FALCON & ELEFANT & YACC \\
-%  \midrule
-%   $\left\langle T \right\rangle$ & 0.7759 & 0.7758 & 0.7767 & 0.6043 & 0.6036 & 0.6058 & 0.7275 & 0.7275 & 0.7286 & & 0.5277 & 0.5289\\
-%    Nu$^{top}$ & 3.4280 & 3.4214 & 3.4298 & 8.5242 & 8.5115 & 8.5278 & 3.0368 & 3.0347 & 3.0374 && 6.5912 & 6.5572 \\
-%    Nu$^{bottom}$ & 3.3168 & 3.313 & 3.3143 & 8.4599 & 8.4465 & 8.3990 & 3.0086 & 2.9908 & 2.9628 && 6.5834 & 6.5243 \\
-%    $v_{\textrm{rms}}$ & 249.901 & 249.134 & 251.7997 & 141.2451 & 140.7087 & 142.2020 & 100.2776 & 100.1208 & 100.9467 && 79.1105 & 79.6202 \\
-%    $v_{\textrm{rms}}^{top}$ & 1.8907 & 1.8642 & 1.8298 & 104.7928 & 104.1228 & 105.1767 & 2.0832 & 2.0652 & 2.0374 && 74.7596 & 75.8414 \\
-%    $u_{max}^{top}$ & 2.6340 & 2.6119 & 2.5516 & 121.9212& 121.766 & 122.3238 & 2.9086 & 2.9019 & 2.8458 && 88.9146 & 89.2940 \\
-%    $\eta_{min}$ & - & - & - & $2.0303\times10^{-5}$ & $2.0600\times10^{-5}$ & $1.9853\times10^{-5}$ & $4.8129\times10^{-5}$ & $4.8080\times10^{-5}$ & $4.7907\times10^{-5}$ &
-%    & $1.9860\times10^{-4}$ & $1.9174\times10^{-4}$ \\
-%    $\eta_{max}$ & - & - & - & 1.3325 & 1.85 & 1.6095 & 0.9360 & 0.9023 & 1 && 1.52 & 1.6773 \\
-%    $\left\langle W \right\rangle$ & 2.4299 & 2.4316 & 2.4583 & 7.5648 & 7.5665 & 7.6222 & 2.0387 & 2.0384 & 2.0400 && 5.6216 & 5.6512 \\
-%    $\left\langle \Phi \right\rangle/Ra$ & 2.4311 & 2.4276 & 2.4333 & 7.5661 & 7.5585 & 7.6194 & 2.0386 & 2.0370 & 2.0335 && 5.6182 & 5.6463 \\
-%    $\delta$ & 0.05\% & 0.16\% & 1.02\% & 0.02\% & 0.11\% & 0.04\% & $<$0.01\% & 0.07\% & 0.32\% && 0.06\% & 0.09\% \\
-%  \bottomrule
-%  \end{tabular}}
-%  \label{tab:visco_mantle}
-%  \end{table}
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{2}{white!100!}{gray!25!}
-\begin{table}[H]
-  \caption{Benchmark cases and corresponding parameters.}
-  \centering
-  \small
-\begin{tabular}{|>{\centering}m{0.08\textwidth}|>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}
-  >{\centering}m{0.05\textwidth} >{\centering}m{0.05\textwidth} |>{\centering\arraybackslash}m{0.2\textwidth}|}
-  \toprule
-    Case & $Ra$ & $\Delta\eta_T$ & $\Delta\eta_y$ & $\eta^*$ & $\sigma_Y$ & Convective regime \\
-  \midrule
-  \rowcolor{gray!25!}
-    1 & $10^2$ & $10^5$ & 1 & - & - & Stagnant lid \\
-    2 & $10^2$ & $10^5$ & 1 & $10^{-3}$ & 1 & Mobile lid  \\
-    \rowcolor{gray!25!}
-    3 & $10^2$ & $10^5$ & 10 & - & - & Stagnant lid  \\
-    4 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 1 & Mobile lid \\
-    \rowcolor{gray!25!}
-    5 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 4 & Periodic \\
-  \bottomrule
-  \end{tabular}
-  \label{tab:tosi_setup}
-  \end{table}
-
-\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{gray!25!}{white!100!}
-\begin{table}
-  \caption{Results of the viscoplastic mantle convection benchmark for different grid resolutions in Cases 1-5 compared with results obtained with ELEFANT
-  and YACC ($100\times100$ elements, from \citealp{Tosi2015}) and in Stone 28 ($32\times32$ elements, from
-  \url{https://github.com/cedrict/fieldstone/tree/master/python_codes/fieldstone_28}).}
-  \centering
-  \resizebox{\textwidth}{!}{
-\begin{tabular}{|cc|ccccccccccc|}
-  \toprule
-  \multicolumn{13}{|c|}{\textbf{Case 1}} \\
-  \midrule
-  Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-  & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-  \hline
-  \rowcolor{gray!25!}
-    YACC & $100\times100$ & 0.7767 & 3.4298 & 3.3143 & 251.7997 & 1.8298 & 2.5516 & - & - & 2.4583 & 2.4333 & 1.02\% \\
-    ELEFANT & $100\times100$ & 0.7758 & 3.4214 & 3.313 & 249.134 & 1.8642 & 2.6119 & - & - & 2.4316 & 2.4276 & 0.16\% \\
-    \rowcolor{gray!25!}
-    Stone 28 & $32\times32$ & 0.7737 & 3.3987 & - & 243.872 & - & - & - & - & - & - & - \\
-%    FALCON & $32\times32$ & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & - & - & 2.4745 & 2.4864 & 0.48\% \\
-%    \rowcolor{gray!25!}
-    FALCON & $100\times100$ & 0.7759 & 3.4280 & 3.3168 & 249.901 & 1.8907 & 2.6340 & - & - & 2.4299 & 2.4311 & 0.05\% \\
-    \midrule
-    \multicolumn{13}{|c|}{\textbf{Case 2}} \\
-    \midrule
-    Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-    & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-    \hline
-    \rowcolor{gray!25!}
-     YACC & $100\times100$ & 0.6058 & 8.5278 & 8.3990 & 142.2020 & 105.1767 & 122.3238 & $1.9853\times10^{-5}$ & 1.6095 & 7.6222 & 7.6194 & 0.04\% \\
-     ELEFANT & $100\times100$ & 0.6034 & 8.5115 & 8.4465 & 140.8067 & 104.1228 & 121.766 & $2.0600\times10^{-5}$ & 1.85 & 7.5665 & 7.5585 & 0.11\% \\
-     \rowcolor{gray!25!}
-     Stone 28 & $32\times32$ & 0.6052 & 8.1715 & - & 141.518 & - & - & - & - & - & - & - \\
-%     FALCON & $32\times32$ & 0.6454 & 8.1422 & 7.4020 & 156.6908 & 95.7226 & 116.9302 & $2.2572\times10^{-5}$ & 0.7194 & 7.4289 & 7.4452 & 0.22\% \\
-%     \rowcolor{gray!25!}
-     FALCON & $100\times100$ & 0.6044 & 8.5266 & 8.4590 & 141.3684 & 104.8867 & 122.0470 & $2.0333\times10^{-5}$ & 1.2914 & 7.5707 & 7.5718 & 0.01\% \\
-     \midrule
-     \multicolumn{13}{|c|}{\textbf{Case 3}} \\
-     \midrule
-     Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-     & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-     \hline
-     \rowcolor{gray!25!} 
-     YACC & $100\times100$ & 0.7286 & 3.0374 & 2.9628 & 100.9467 & 2.0374 & 2.8458 & $4.7907\times10^{-5}$ & 1 & 2.0400 & 2.0335 & 0.32\% \\
-      ELEFANT & $100\times100$ & 0.7275 & 3.0347 & 2.9908 & 100.1208 & 2.0652 & 2.9019 & $4.8080\times10^{-5}$ & 0.9023 & 2.0384 & 2.037 & 0.07\% \\
-      \rowcolor{gray!25!}
-      Stone 28 & $32\times32$ & 0.7277 & 3.0347 & - & 100.018 & - & - & - & - & - & - & - \\
-%      FALCON & $32\times32$ & 0.7281 & 3.0618 & 2.8132 & 101.0092 & 2.1630 & 3.0178 & $4.8500\times10^{-5}$ & 0.7845 & 2.0699 & 2.0695 & 0.02\% \\
-%      \rowcolor{gray!25!}
-      FALCON & $100\times100$ & 0.7275 & 3.0368 & 3.0086 & 100.2776 & 2.0832 & 2.9086 & $4.8129\times10^{-5}$ & 0.9360 & 2.0387 & 2.0386 & $<$0.01\% \\
-      \midrule
-      \multicolumn{13}{|c|}{\textbf{Case 4}} \\
-      \midrule
-      Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-      & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-      \hline
-      \rowcolor{gray!25!} 
-      YACC & $100\times100$ & 0.5289 & 6.5572 & 6.5243 & 79.6202 & 75.4814 & 89.2940 & $1.9174\times10^{-4}$ & 1.6773 & 5.6512 & 5.6463 & 0.09\% \\
-       ELEFANT & $100\times100$ & 0.5277 & 6.5912 & 6.5834 & 79.1105 & 74.7596 & 88.9146 & $1.9860\times10^{-4}$ & 1.5200 & 5.6216 & 5.6182 & 0.06\% \\
-       \rowcolor{gray!25!}
-       Stone 28 & $32\times32$ & 0.5286 & 6.4036 & - & 79.4746 & - & - & - & - & - & - & - \\
-%       FALCON & $32\times32$ & 0.5334 & 6.4679 & 6.3435 & 80.6153 & 76.0511 & 90.2058 & $2.0393\times10^{-4}$ & 0.6453 & 5.6676 & 5.6698 & 0.04\% \\
-%       \rowcolor{gray!25!}
-       FALCON & $100\times100$ & 0.5279 & 6.5897 & 6.5830 & 79.1220 & 75.0715 & 88.8484 & $1.9693\times10^{-4}$ & 1.4133 & 5.6127 & 5.6113 & 0.03\% \\
-       \midrule
-       \multicolumn{13}{|c|}{\textbf{Case 5}} \\
-       \midrule
-       Code & Resolution & $P$ & $\left\langle T \right\rangle_{min}$ & $\left\langle T \right\rangle_{max}$ & Nu$^{top}_{min}$ & Nu$^{top}_{max}$ &
-       $v_{\textrm{rms}}^{min}$ & $v_{\textrm{rms}}^{max}$ & \multicolumn{2}{c}{$(\left\langle \Phi \right\rangle/Ra)_{min}$} & 
-       \multicolumn{2}{c|}{$(\left\langle \Phi \right\rangle/Ra)_{max}$} \\
-       \hline
-       \rowcolor{gray!25!}
-         YACC & $100\times100$ & 0.0792 & 0.6543 & 0.6722 & 2.6950 & 7.2792 & 41.9813 & 98.8230 & \multicolumn{2}{c}{1.4347} & 
-         \multicolumn{2}{c|}{9.2748} \\
-         ELEFANT & $100\times100$ & 0.076 & 0.6514 & 0.6693 & 2.6808 & 7.3506 & 41.2578 & 100.089 & \multicolumn{2}{c}{1.3324} & 
-         \multicolumn{2}{c|}{9.5289} \\
-         \rowcolor{gray!25!}
-         Stone 28 & $32\times32$ & 0.0764 & 0.6521 & 0.6697 & 2.6912 & 7.0763 & 41.61 & 98.183 & \multicolumn{2}{c}{-} & 
-         \multicolumn{2}{c|}{-} \\
-     %    FALCON & $32\times32$ & & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & \multicolumn{2}{c}{-} & 
-     %     \multicolumn{2}{c|}{-} \\
-     %    \rowcolor{gray!25!}
-         FALCON & $100\times100$ & 0.0775 & 0.6519 & 0.6697 & 2.6847 & 7.3629 & 41.5256 & 99.5466 & \multicolumn{2}{c}{1.3399} & 
-         \multicolumn{2}{c|}{9.3993} \\
-  \bottomrule
-  \end{tabular}}
-  \label{tab:visco_mantle}
-  \end{table}
-
-\subsection{Thin layer entrainment}\label{sec:thin}
-This experiment is performed as originally proposed by \citet{vanKeken1997} with two incompressible fluids in a rectangular domain with $L_x=2$ and $L_y=1$
-and gravity acceleration $g_y=10^{10}Ra$. The Rayleigh number and the compositional Rayleigh number are fixed to $Ra=300000$ and $Rc=450000$, respectively.
-Both fluids have constant viscosity, thermal conductivity, specific heat ($\eta=\rho=C_p=1$) and thermal expansion coefficient ($\alpha=10^{-10}$). Fluid 1
-has a density $\rho_1=1$, while fluid 2 is denser ($\rho_2=\rho_1+1.5 \cdot 10^{-10}$) and is placed at the bottom of the domain, for $y \leq 0.025$.
-Temperature are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides of the domain. The initial
-temperature field is given by
-\[T(x,y)=T_u(x,y)+T_l(x,y)+T_r(x,y)+T_s(x,y)-\frac{3}{2}\]
-with
-\begin{eqnarray}
-T_u(x,y)&=&\frac{1}{2}erf\left(\frac{1-y}{2}\sqrt{\frac{u_0}{x}}\right)\nonumber \\
-T_l(x,y)&=&1-\frac{1}{2}erf\left(\frac{y}{2}\sqrt{\frac{u_0}{L_x-x}}\right)\nonumber \\
-T_r(x,y)&=&\frac{1}{2}+\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{y+1}} \exp\left(-\frac{x^2u_0}{4y+4}\right)\nonumber \\
-T_s(x,y)&=&\frac{1}{2}-\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{2-y}} \exp\left(-\frac{(L_x-x)^2u_0}{8-4y}\right)\nonumber
-\end{eqnarray}
-and
-\begin{eqnarray}
-u_0&=&\frac{L_x^{7/3}}{(1+L_x^4)^{2/3}}\left(\frac{Ra}{2\sqrt{\pi}}\right)^{\frac{2}{3}}\nonumber \\
-Q&=&2\sqrt{\frac{L_x}{\pi u_0}}\nonumber
-\end{eqnarray}
-
-The experiment is performed with two grid resolution (\num{125x40} and \num{200x80} elements), with 100 markers per element and CFL of 0.25. $v_{\textrm{rms}}$
-(see Sec. \ref{sec:error}) calculated for both the simulations match well with results from \citet{vanKeken1997} and obtained with ASPECT
-\citep{Kronbichler2012,Heister2017,Bangerth2020,Bangerth2020a} and ELEFANT \citep{Thieulot2014}, obtained with similar grid resolutions
-(Fig. \ref{fig:thin}). In addition, a variety of simulations with different resolutions and aspect ratios are performed to verify that the sensitivity of the
-initial velocity field. Results match well with results from \citet{vanKeken1997} (in the grey area) and \citet{Thieulot2014} (Fig. \ref{fig:thin_initial}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Thin_layer}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_thin.pdf}
-\caption{$v_{\textrm{rms}}$ for the thin layer experiment as function of time for different grid resolution. Results are compared with results obtained by
-\citet{vanKeken1997} (black lines) and with ASPECT and ELEFANT.}
-\label{fig:thin}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrmszero.pdf}
-\caption{$v_{\textrm{rms}}$ at $t=0$ for the thin layer experiment as function of the element size for different aspect ratios of the elements. Results are compared
-with results obtained by \citet{vanKeken1997} (black dots in the grey area) and with ASPECT and ELEFANT.}
-\label{fig:thin_initial}
-\end{figure}
-
-\subsection{Sinking hydrated cylinder}\label{sec:quinquis}
-This experiment is performed as explained in \citet{Quinquis2014} and simulates the sinking of a cold, hydrated cylinder into a warm, dry mantle. The domain
-is square with $L_x=L_y=$ \SI{300}{\km}, a grid resolution of \num{300x300} elements and 25 markers per element. The sinking cylinder is located at $x=$
-\SI{150}{\km} and $y=$ \SI{170}{\km} with a radius of \SI{20}{\km} and has $\rho_c=$ \SI{3250}{\kg\per\cubic\m}, $\eta_c=$ \SI{e23}{\pascal\s}, a thermal
-conductivity $k_c=$ \SI{4.5}{\watt\per\m\per\kelvin},a specific heat $C_{p_c}=$ \SI{1250}{\joule\per\kg\per\kelvin}; the surrounding mantle has $\rho_m=$
-\SI{3200}{\kg\per\cubic\m}, $\eta_m=$ \SI{e20}{\pascal\s}, a thermal conductivity $k_m=$ \SI{105}{\watt\per\m\per\kelvin} and specific heat $C_{p_m}=$
-\SI{1250}{\joule\per\kg\per\kelvin}; a 58 km-thick lithosphere overlies the mantle and has $\rho_l=$ \SI{3200}{\kg\per\cubic\m}, $\eta_l=$ \SI{e23}{\pascal\s},
-a thermal conductivity $k_l=$ \SI{4.5}{\watt\per\m\per\kelvin}, specific heat $C_{p_l}=$ \SI{750}{\joule\per\kg\per\kelvin} and a thermal diffusivity of
-\SI{e-6}{\square\m\per\s}. Velocity boundary conditions are set to free slip on all sides of the domain. Temperature increase linearly in the lithosphere,
-from 0° to 1300°C, and in the mantle, from 1300° to 1360.5°C. The initial temperature of the cylinder is fixed to 400°C. Throughout the evolution, temperature
-is fixed to 0° and 1360.5°C at the top and bottom of the domain, respectively. The initial bound water content is imposed at 2 wt.\% in the cylinder and at
-0 wt.\% in both the mantle and the lithosphere. The maximum amount of water in the mantle is fixed to 0.2 wt.\%, while maximum water content in the cylinder
-and in the lithosphere is function of pressure and temperature and is calculated using a serpentinized harzburgite with Perple\_X, as in \citet{Quinquis2014}.
-
-As shown in \citet{Quinquis2014}, the model is characterised by a progressive dehydration in the external portion of the cylinder, due to the increase of
-temperature, with a consequent vertical migration of free water that hydrates the mantle above the cylinder, up to the lithosphere. The distribution of
-bound and free water during the evolution of the experiment is shown in Fig. \ref{fig:quinquis} (left and right columns, respectively). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Hydration}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Hydration.pdf}
-\caption{Distribution of bound and free water (left and right columns, respectively) at $t=0.5$ and $t=$ \SI{1.25}{\mega\year} (first and second rows,
-respectively) for the sinking hydrated cylinder experiment. Only markers with bound or free water not equal to 0 are plotted.}
-\label{fig:quinquis}
-\end{figure}
-
-\subsection{Mantle melting curves}\label{sec:katz}
-This instantaneous experiment is performed in order to recreate isobaric melting curves, as function of temperature, bulk water and content of water in the
-melt, to compare with melting curves obtained by \citet{Katz2003}. The domain is square with $L_x=L_y=100$ km, a grid resolution of \num{100x100} elements and
-25 markers per element. Density is set in the entire domain to \SI{3300}{\kg\per\cubic\m} and temperature increases from the sides to the centre from 1000° to
-1500°C. The experiment is performed for bulk water contents from 0 to 0.3 wt.\%.
-
-Fig. \ref{fig:fraction} shows isobaric melting curves for \SI{1}{\giga\pascal} (panel a) and \SI{3}{\giga\pascal} (panel b) calculated for different bulk water
-contents. The curves well match with those obtained for the mantle by \citet{Katz2003}, showing the saturation at the solidus for bulk water of 0.2 and 0.3 wt.\%, indicating
-that the code correctly determines melt fraction as function of pressure, temperature and bulk water content. Same results can be observed in Fig.
-\ref{fig:water}a, where isobaric and isothermal melting curves are shown as function of bulk water content. In addition, Fig. \ref{fig:water}b shows that the
-code correctly determines also the percentage of water in the melt as function of pressure, temperature and melt fraction. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Melting}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Fraction.pdf}
-\caption{Isobaric melting curves for 1 GPa (panel a) and 3 GPa (panel b) as function of temperature with different bulk water contents.}
-\label{fig:fraction}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Water.pdf}
-\caption{Isobaric melting curves for 1.5 GPa as function of bulk water (panel a) and water in the melt (panel b) calculated for different temperatures.}
-\label{fig:water}
-\end{figure}
-
-\clearpage
-
-\bibliography{export}
-
-\end{document}

### 2022-09-07 tex/phase.tex
update main

--- 2022-01-28 .debris/2022-01-27/Paper.tex
@@
-\documentclass[hidelinks,10pt,a4paper]{article}
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage{booktabs}
-\usepackage{multirow}
-\usepackage[cm]{fullpage}
-\usepackage[round]{natbib}
-\bibliographystyle{abbrvnat}
-\usepackage{doi}
-\usepackage{bm}
-\usepackage[table]{xcolor}
-\hypersetup{
-    colorlinks,
-    linkcolor={blue},
-    citecolor={blue},
-    urlcolor={blue}
-}
-\usepackage{float}
-\restylefloat{table}
-\usepackage{siunitx}
-\DeclareSIUnit\year{yr}
+\section{Phase transitions and hydration}\label{sec:phase}
+Variations of effective density, specific heat and coefficient of thermal expansion during the evolution of crustal and mantle materials at different pressure-temperature ($p$-$T$) conditions are computed by means of version 6.8.6 of Perple\_X software package \citep{Connolly2005}, similarly to the implementation described in \citet{Marotta2020}.
+By default, mineral assemblages and properties of each lithology (oceanic crust, lower and upper continental crust, sediments and dry or hydrated mantle) are calculated for temperatures from \SIrange{330}{1600}{\kelvin}, with increments of approximately \SI{4}{\kelvin}, and pressures from \SIrange{0.1}{30}{\giga\pascal}, with increments of approximately \SI{0.03}{\giga\pascal}, for a total of almost 100000 points.
 
-\begin{document}
-
-\section{Introduction}\label{sec:intro}
-This supplement includes a description of the code, with a presentation of the features implemented followed by results of extensive benchmarks listed in
-Table \ref{tab:list}, implemented to verify that the code solves correctly a large range of problems. Description and results of all benchmarks are shown
-in Section \ref{sec:benchmark}.
-
-\begin{table}[H]
-\caption{List of benchmarks implemented for each feature of the code. Data of all benchmarks can be found at 
-\url{https://github.com/aleregorda/Benchmarks}}
-\centering
-\begin{tabular}{| l | l | l |}
-\toprule
- \textbf{Feature}  & \textbf{Benchmark} & \textbf{Section} \\
-\midrule
-  Solver  & Stokes flow & \ref{sec:stokes} \\
-\hline
-  \multirow{2}{*}{Markers advection} & Zalesak disk & \ref{sec:runge} \\
-    & Conservative Interpolation Velocity & \ref{sec:CVI} \\
-\hline
-  \multirow{4}{*}{Momentum equation}  & Poiseuille flow & \ref{sec:poiseuille}\\
-    & Instantaneous 2D sphere & \ref{sec:ist_sphere} \\
-    & Rayleigh-Taylor experiment & \ref{sec:rayleigh} \\
-    & Falling block & \ref{sec:block} \\
-\hline
-    & 2D Stokes sphere & \ref{sec:sphere} \\
-    Sticky air and  & Stabilisation algorithm & \ref{sec:stab} \\
-    free surface & Topography relaxation & \ref{sec:crameri} \\
-  & Spontaneous subduction & \ref{sec:subduction} \\
-\hline
-  Erosion and sedimentation  & - & - \\
-\hline
-   Non-linear  & Slab detachment & \ref{sec:slab} \\
-   visco-plasticity & Indenter & \ref{sec:indenter} \\
-    & Brick & \ref{sec:brick} \\
-\hline
-  \multirow{3}{*}{Energy equation}  & Advection stabilisation & \ref{sec:advection} \\
-    & Simple shear heating & \ref{sec:simple_shear} \\
-    & Shear and adiabatic heating & \ref{sec:shear} \\
-\hline
-\multirow{3}{*}{Energy + momentum} & Mantle convection & \ref{sec:mantle} \\
-& Viscoplastic mantle convection & \ref{sec:visco_mantle} \\
-& Thin layer entrainment & \ref{sec:thin} \\
-\hline
-  Phase changes  & \multirow{2}{*}{Hydrated sinking cylinder} & \multirow{2}{*}{\ref{sec:quinquis}} \\
-  and hydration  &  & \\
-\hline
-  Melting & Experimental melting curves & \ref{sec:katz} \\
-\bottomrule
-\end{tabular}
-\label{tab:list}
-\end{table}
-
-The code is written in Fortran90 and uses the finite element method (FEM) with quadrilateral $Q_1 \times P_0$ elements (continuous bilinear velocity and
-discontinuous constant pressure), associated with MUMPS\footnote{\url{http://mumps.enseeiht.fr/}} \citep{Amestoy2001,Amestoy2019} that is a software package
-for solving systems of linear equations of the form $A \cdot x = b$, where \textit{A} is a square sparse matrix, by means of a direct method. Since
-$Q_1 \times P_0$ elements do not satisfy Ladyzhenskaya, Babuska and Brezzi (LBB) stability condition \citep{Donea2003}, elemental pressure is elaborated
-in post-processing to avoid spurious pressures by means of a double interpolation to smooth the pressure. Following this procedure, the elemental pressure
-is interpolated onto nodes and then back onto elements. After the smoothing procedure, the pressure field is calculated again on the nodes to be used in
-combination with the lithostatic pressure, which is calculated onto nodes. Correctness of the solution and performances of the solver in terms of time and
-memory usage are tested solving the Stokes flow with the analytical solution proposed by \citet{Donea2003} (Section \ref{sec:stokes}).
-
-The thermo-mechanics of crust-mantle systems is described by means of the conservation of mass, momentum and energy equations, expressed as follows:
-\begin{eqnarray}
-\label{eq:mass}\frac{\partial \rho}{\partial t}+\bm{\nabla} \cdot (\rho \bm{u})&=&0\\
-\label{eq:momentum}-\bm{\nabla} p + \bm{\nabla} \cdot \bm{\tau} + \rho \bm{g}&=&0\\
-\label{eq:energy}\rho C_p \left(\frac{\partial T}{\partial t} + \bm{u} \cdot \bm{\nabla} T \right)&=&\bm{\nabla} \cdot \left(k\bm{\nabla} T\right) + H_{tot}
-\end{eqnarray}
-where $\rho$ is the density, $\bm{u}$ is the velocity, \textit{p} is the pressure, $\bm{\tau}$ is the deviatoric stress, $\bm{g}$ is the gravity acceleration,
-$C_p$ is the specific heat at constant pressure, \textit{T} is the temperature, \textit{k} is the thermal conductivity and $H_{tot}$ is the total internal heat
-production. Density variations due to temperature are generally small enough to assume the density as constant ($\rho=\rho_0$) in Eq. \ref{eq:mass} and in
-Eq. \ref{eq:energy}, while it must be treated as a variable in the buoyancy term of Eq. \ref{eq:momentum}, such that \[\rho=\rho_0(1-\alpha(T-T_0))\]
-where $\rho_0$ is the density at a reference temperature $T_0$ and $\alpha$ is the coefficient of thermal expansion. Eq. \ref{eq:mass} then can be rewritten as
-\begin{equation}\label{eq:mass_0}
-\bm{\nabla} \cdot \bm{u}=0
-\end{equation}
-This simplification is known as the extended Boussinesq approximation.
-
-The time step $dt$ is chosen in according to Courant-Friedrichs-Lewy condition \citep{Anderson}, such that
-\[dt=C_n\; \min\left(\frac{\Delta x}{\max|\bm{u}|},\frac{(\Delta x)^2}{\kappa}\right)\]
-where $0<C_n<1$ is the Courant (CFL) number, $\Delta x$ is the minimum dimension of smallest element, $\max|\bm{u}|$ is the maximum velocity calculated on the
-entire domain and
-$\kappa=\frac{k}{\rho C_p}$ is the thermal diffusivity \citep{Thieulot2014}.
-
-\section{Lagrangian markers}\label{sec:markers}
-Elemental properties (density, viscosity, thermal conductivity, specific heat and thermal expansion) needed to solve Eqs. \ref{eq:mass_0}, \ref{eq:momentum}
-and \ref{eq:energy} are related to the composition of each element, determined by means of Lagrangian markers that are characteristic of different materials
-in the domain. Elemental properties (with the sole exception of the viscosity) are calculated using an arithmetic mean, as \[P_e=\frac{1}{n}\sum_{i=1}^n P_i\]
-where $P_e$ is the elemental property, $P_i$ is the property characteristic of the material of each marker and $n$ is the number of markers of the element.
-Differently, the average scheme for the elemental viscosity can be chosen between harmonic, geometric and arithmetic mean. Markers can be placed regularly or
-randomly at the beginning of the simulation and their advection is performed by either a 2nd-order or a 4th-order Runge-Kutta in space, interpolating the velocity 
-field on each marker by means of the shape functions (see Section \ref{sec:runge}). The interpolated velocity is then corrected by means of the Conservative
-Velocity Interpolation (CVI), which introduces a corrective term to reduce dispersion and clustering of particles in both steady state and time-dependent 
-\citep{Wang2015} (see Section \ref{sec:CVI}).
-
-The initial distribution of the markers is created using the open-source code library
-Geodynamic World Builder \citep{Fraters2019}. During the simulation, each marker carries memory of temperature, pressure and accumulated strain, which are
-determined interpolating the relative nodal parameter. In particular, pressure is interpolated onto each marker after the smoothing procedure. The number
-of markers contained in each element is maintained between $n_{min}$ and $n_{max}$. When in an element there are less markers than $n_{min}$ the code adds
-random markers to reach the $n_{min}$, while if the number is higher than $n_{max}$ some of them are randomly deleted. When new markers are added, they assume
-the properties of the nearest marker. In this way elements are never empty and maintain a number of markers inside a prefixed range.
-
-\section{The momentum equation}\label{sec:mom_eq}
-The penalty method is implemented so that Eq. \ref{eq:mass_0} can be written as:
-\begin{equation}\label{eq:mass_penalty}
-\bm{\nabla} \cdot \bm{u} + \frac{p}{\lambda}=0
-\end{equation}
-where $\lambda$ is the so-called penalty parameter, which should be 6-7 orders of magnitude larger than the shear viscosity to ensure that mass conservation
-is satisfied \citep{Donea2003,Thieulot2014}.
-
-The deviatoric stress tensor in Eq. \ref{eq:momentum} can be written in terms of the strain rate tensor as $\bm{\tau}=2\eta\dot{\bm{\epsilon}}$, with
-$\dot{\bm{\epsilon}}=\frac{1}{2}\left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)$. Therefore, Eq. \ref{eq:momentum} can be rewritten as:
-\begin{equation}\label{eq:momentum_pressure}
--\bm{\nabla} p + \bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right) + \rho \bm{g} = 0
-\end{equation}
-Finally, using pressure from Eq. \ref{eq:mass_penalty}, Eq. \ref{eq:momentum_pressure} can be rewritten as:
-\begin{equation}\label{eq:momentum_penalty}
-\lambda \bm{\nabla} \left(\bm{\nabla} \cdot \bm{v}\right)+\bm{\nabla} \cdot \left(\eta \left(\bm{\nabla}\bm{v}+(\bm{\nabla}\bm{v})^T\right)\right)+\rho\bm{g}=0
-\end{equation}
-The penalty method is been associated to the iterative Uzawa method, as described in detail in \citet{Dabrowski2008} and \citet{Thieulot2014}
-(see Section \ref{sec:poiseuille}). Results of benchmarks performed to verify the correctness in the implementation of Eq. \ref{eq:momentum_penalty} are
-shown in Sections \ref{sec:ist_sphere} and \ref{sec:rayleigh}.
-
-To support large viscosities variations the penalty parameter is related to the effective elemental viscosity by means of a dimensionless coefficient, so that
-$\lambda=\lambda_e(e)\eta_{eff}(e)$ \citep{Marotta2006,Dabrowski2008,Thieulot2014}. Benchmark of the falling block \citep{Gerya2003a,Gerya2010b,Thieulot2011}
-is performed to verify that the code can correctly deal buoyancy driven flows with strong viscosity contrasts (Section \ref{sec:block}).
-
-\section{Sticky air and free surface}\label{sec:sticky}
-The Earth's surface can be treated by means of either the so-called sticky-air or a true free surface method, both of them implemented in the code. In the
-sticky-air method the surface is approximated with the introduction of a buoyant layer with a viscosity at least four orders of magnitude lower than the crust
-\citep{Schmeling2008,Crameri2012} and the interface between lithosphere and air is defined using a chain of passive markers that are advected as the Lagrangian
-markers. The correctness of the evolution of the markers chain is tested with the experiment of a 2D time-dependent Stokes sphere below a free surface and
-compared with results from ASPECT\footnote{\url{https://aspect.geodynamics.org/}} \citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} (Section
-\ref{sec:sphere}). In the true free surface case the top boundary is assumed stress-free and velocities are not fixed. In this case topography variations are
-described by vertical deformations of the mesh that depend on the velocity field of the nodes that identify the free surface, while horizontal deformations are
-not taken into account. This procedure is known as the Arbitrary Lagrangian-Eulerian (ALE) method and its implementation follows the technique described in
-\citet{Thieulot2011}. However, although the implementation of a true free surface better reproduces laboratory experiment, extremely small time steps can be
-necessary to maintain stability \citep{Kaus2010a,Quinquis2011,Thieulot2014}. Therefore, the stability algorithm proposed by \citet{Kaus2010a} is implemented
-to avoid instabilities due to high density differences at the free surface when using too large time steps. The implementation of this algorithm are tested by
-performing the experiment described by \citet{Kaus2010a} (Section \ref{sec:stab}).
-
-The topography relaxation benchmark proposed by \citet{Crameri2012} is performed to verify that the code correctly recovers topography variations in case of
-both the sticky-air and the true free surface method (Section \ref{sec:crameri}). Finally, Section \ref{sec:subduction} show the results of the spontaneous
-subduction experiment described by \citet{Schmeling2008}.
-
-\section{Erosion and sedimentation}\label{sec:erosion}
-Surface processes at the lithosphere-air interface are implemented taking into account fluvial erosion (long-range processes), hillslope processes (short-range
-processes) and failure of slopes exceeding a threshold angle $\theta$ \citep{Burov1997,Tucker1999,Sacek2012,Croissant2014,Ueda2015,Beucher2020}. Erosion rates
-$\dot{E}_{lr}$ due to long-range processes can be described by a stream power law equation \citep{Tucker1999,Ueda2015,Beucher2020}, in according to
-\begin{equation}
-\dot{E}_{lr}=K_f A^{m_e} \bm{\nabla} h^{n_e}
-\end{equation}
-where $K_f$ is the erodibility coefficient, $A$ is the upstream drainage area, $\bm{\nabla} h$ is the slope and $m_e$ and $n_e$ are parameters chosen in to
-constrain the ratio $m_e/n_e$ between 0.35 and 0.8 \citep{Croissant2014,Ueda2015,Beucher2020}. Erosion rates $\dot{E}_{sr}$ due to short-range processes are
-implemented by assuming that transport is linearly proportional to slope \citep{Burov1997,Ueda2015,Beucher2020}, in according to
-\begin{equation}
-\dot{E}_{sr}=K_d \bm{\nabla}^2 h
-\end{equation}
-where $K_d$ is the effective diffusivity. In addition, the slope is modified if it still exceeds a threshold angle after long- and short-range processes
-\citep{Ueda2015}.
-
-Offshore sedimentation processes are implemented similarly to hillslope processes but with a depth-dependent coefficient \citep{Sacek2012,Beucher2020}.
-The sedimentation rate $\dot{S}$ is calculated as
-\begin{equation}
-\dot{S}=K_s \bm{\nabla}^2 h
-\end{equation}
-with
-\begin{equation}
-K_s=K^*_s \exp\left(\frac{h-h_{sl}}{h_0}\right)
-\end{equation}
-where $K^*_s$ is a constant, $h_{sl}$ is the sea level and $h_0$ is a characteristic length. In case of sticky air, slope is determined by means of the markers
-chain, which is vertically corrected considering erosion and sedimentation rates. After the correction, continental markers above the markers chain are
-transformed in air markers and air markers below the markers chain become sediments. Similarly, slope is determined considering the nodes on the top boundary
-in case of true free surface and they are vertically corrected as for the markers chain. In this case, continental markers above the top boundary are simply
-deleted.
-
-\section{Non-linear rheologies}\label{sec:nonlinear}
-Non-linear rheologies are implemented combining viscous creep (dislocation and diffusion) and plastic yielding. For each marker, diffusion and dislocation
-viscosity ($\eta_{df|ds}$) can be determined as follows \citep{Karato1993,Wang2016}:
-\begin{equation}\label{eq:viscous}
-\eta_{df|ds}=\left(\frac{d^m}{A}\right)^{\frac{1}{n}}I_2^{\frac{(1-n)}{n}}\exp\left(\frac{Q+pV}{nRT}\right)
-\end{equation}
-where $d$ is the grain size, $m$ is the grain size exponent, $n$ is the stress exponent, $A$ is the pre-exponential factor, $I_2$ is the square root of the
-second invariant of the strain rate tensor, $Q$ is the activation energy, $p$ is the pressure, $V$ is the activation volume, $R$ is the gas constant and $T$
-is the temperature. Pressures and temperatures are determined by the interpolation of the nodal parameters, while strain rates are calculated by means of the
-derivative of the velocities on the nodes. In case of diffusion creep $n=1$ and $m>0$, in case of dislocation creep $n>1$ and $m=0$. The viscous creep $\eta_{cp}$
-is then calculated as the harmonic average between $\eta_{df}$ and $\eta_{ds}$:
-\begin{equation}\label{eq:visco_creep}
-\eta_{cp}=\left(\frac{1}{\eta_{df}}+\frac{1}{\eta_{ds}}\right)^{-1}
-\end{equation}
-The implementation of the non-linear viscous creep viscosity depending on the strain rate is tested by means of the slab detachment benchmark
-(Section \ref{sec:slab}).
-
-Plastic yielding is implemented rescaling $\eta_{cp}$ in order to limit the stress below the yield stress $\sigma_y$ \citep{Thieulot2008,Thieulot2014,Glerum2018},
-obtaining
-\begin{equation}\label{eq:plastic}
-\eta_{pl}=\frac{\sigma_y}{2I_2}
-\end{equation}
-where $\eta_{pl}$ is the plastic viscosity and the yield stress is determined following the Drucker-Prager criterion, such as
-\begin{equation}\label{eq:yield}
-\sigma_y=C\;\cos(\phi)+p\;\sin(\phi)
-\end{equation}
-where $C$ is the cohesion and $\phi$ is the internal friction angle. In case of negative pressure it is imposed equal to 0, so that negative yield stress are
-excluded. The correctness of the non-linear solution in case of plastic viscosity with variable internal friction angle is verified performing indenter and
-brick experiments (Sections \ref{sec:indenter} and \ref{sec:brick}, respectively).
-
-Strain softening is taken into account for both viscous creep and plastic viscosity \citep{Huismans2003,Babeyko2005,Huismans2005,Sobolev2005,Warren2008a} by
-means of the accumulated strain $\epsilon$ memorised by each marker. Softening in the viscous creep determines a linear decrease of $\eta_{vc}$ by means of a
-viscous strain softening factor $W_S$ that increases linearly from $W_{S_0}$ to $W_{S_{\infty}}$ for $\epsilon_{S_0}<\epsilon_S<\epsilon_{S_{\infty}}$ 
-\citep{Warren2008a}. This viscous softening can be related to strain-induced grain size reduction \citep{Warren2008a}. Differently, plastic softening is 
-simulated with a linear decrease of internal friction angle $\phi(\epsilon)$ and cohesion $C(\epsilon)$ in according to
-\begin{eqnarray}
-\label{eq:friction1}\phi(\epsilon)&=&\phi_0+(\phi_{\infty}-\phi_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}\\
-\label{eq:friction2}C(\epsilon)&=&C_0+(C_{\infty}-C_0)\frac{\epsilon-\epsilon_0}{\epsilon_{\infty}-\epsilon_0}
-\end{eqnarray}
-where $\phi_0$, $C_0$ and $\phi_{\infty}$, $C_{\infty}$ are internal friction angle and cohesion for $\epsilon_0$ and $\epsilon_{\infty}$, respectively
-\citep{Huismans2003,Huismans2005,Warren2008a,Thieulot2014}. Plastic softening approximates deformation-induced softening of faults and brittle shear zones
-\citep{Warren2008a}.
-
-Plastic yielding and viscous creep are then combined to obtain a viscoplastic viscosity $\eta_{vp}$ as follows, assuming that they are independent processes
-\citep{Karato2008,Glerum2018}:
-\begin{equation}\label{eq:viscoplastic}
-\eta_{vp}=\min(\eta_{cp},\eta_{pl})
-\end{equation}
-Finally, effective viscosity $\eta_{eff}$ is capped by the minimum and the maximum viscosity ($\eta_{min}$ and $\eta_{max}$, respectively) to avoid extremely
-low or high viscosity \citep{Glerum2018} as follows
-\begin{equation}\label{eq:effective}
-\eta_{eff}=\min(\max(\eta_{vp},\eta_{min}),\eta_{max})
-\end{equation}
-Viscous creep and plastic yielding are non-linear rheologies because of their dependence on the velocity field through pressure and strain rates. Therefore,
-the solution is determined by means of Picard-type iterations, until convergence of the velocity field \citep{Glerum2018}. The convergence is verified at each
-iteration $i$ via the nonlinear residual $\bm{\mathcal{R}}^i$ that can be determined as
-\begin{equation}\label{eq:convergence}
-\bm{\mathcal{R}}^i= \mathbb{K} (\eta_{eff}(\dot{\bm{\epsilon}}^{i-1},p^{i-1}))\cdot \bm{v}^{i-1} - \bm{f}^i
-\end{equation}
-%\begin{equation}\label{eq:convergence}
-%\chi=1-\frac{\langle (\bm{v}^i-\langle\bm{v}^i\rangle)\cdot (\bm{v}^{i+1}-\langle\bm{v}^{i+1}\rangle) \rangle}{\sqrt{\sigma^i\sigma^{i+1}}}
-%\end{equation}
-%Iterations are performed until either convergence ($\chi<tol$) or maximal number of non-linear iterations ($it_{max}$) is reached.
-where $\mathbb{K}$ is the Stiffness matrix and $\bm{f}$ is the right hand side vector \citep{Spiegelman2016,Glerum2018}.  The $L_2$-norm is extracted from
-$\bm{\mathcal{R}}^i$ and it is normalised as follows
-\begin{equation}\label{eq:convergence_res}
-\frac{||\bm{\mathcal{R}}^i||_2}{||\bm{\mathcal{R}}^0||_2}
-\end{equation}
-where $||\bm{\mathcal{R}}^0||_2$ is the $L_2$-norm of the first nonlinear iteration.
-Since often the normalised nonlinear residual drop very quickly over the first 2-3 iterations, also the $L_2$-norm of velocity and pressure residuals are
-calculated as
-\begin{equation}\label{eq:convergence_vel}
-\frac{||\Delta\bm{v}||_2}{||\bm{v}||_2}=\frac{||\bm{v}_i-\bm{v}_{i-1}||_2}{||\bm{v}_{i}||_2}
-\end{equation}
-and
-\begin{equation}\label{eq:convergence_press}
-  \frac{||\Delta p||_2}{||p||_2}=\frac{||p_i-p_{i-1}||_2}{||p_{i}||_2}
-\end{equation}
-respectively. Iterations are performed until either maximal number of non-linear iterations ($it_{max}$) is reached or all the normalised nonlinear residual, 
-the velocity residual and the pressure residual converge under a defined tolerance $tol$.
-
-\section{The energy equation}\label{sec:en_eq}
-The stabilisation of the advection term of the energy equation (Eq. \ref{eq:energy}) needed to avoid possible oscillations in the thermal solution in the case
-when advection dominates over diffusion, is implemented by means of a streamline-upwind Petrov–Galerkin (SUPG) method, for which the advection term is modified
-follows the discussion in \citet{Thieulot2011} and \citet{Thieulot2014} (see Section \ref{sec:advection} for benchmark). The total internal heat production
-$H_{tot}$ in Eq. \ref{eq:energy} includes radiogenic heating $H_r$, shear heating $H_s$ and adiabatic heating $H_a$, where
-\begin{equation}\label{eq:shear}
-H_s=2\eta \dot{\bm{\epsilon}} : \dot{\bm{\epsilon}} = 2\eta(\dot{\epsilon}_{xx}^2+\dot{\epsilon}_{yy}^2+2\dot{\epsilon}_{xy}^2)
-\end{equation}
-and
-\begin{equation}\label{eq:adiabatic}
-H_a=T \alpha \frac{Dp}{Dt} \approx -\alpha T \rho g_y v
-\end{equation}
-The correctness of the implementation and the computation of shear and adiabatic heating is verified performing an exercise with an analytical solution
-(Section \ref{sec:simple_shear}) and Exercise 9.4 in \citet{Gerya2010b} (Section \ref{sec:shear}). Furthermore, mantle convection and thin layer entrainment
-benchmarks (Sections \ref{sec:mantle} and \ref{sec:thin}, respectively) are performed to verify that the code solve correctly Eqs. \ref{eq:momentum_penalty}
-and \ref{eq:energy} in case of a temperature field characterised by an initial perturbation and with a temperature-dependent density. At last, the viscoplastic
-mantle convection benchmark is performed to check also the viscoplasticity in case of a initial perturbed temperature field (Section \ref{sec:visco_mantle}).
-
-\section{Phase transitions and hydration}\label{sec:phase}
-Variations of effective density, specific heat and coefficient of thermal expansion during the evolution of crustal and mantle materials at different
-pressure-temperature ($p$-$T$) conditions are computed by means of version 6.8.6 of Perple\_X software package \citep{Connolly2005}, similarly to the
-implementation described in \citet{Marotta2020}. By default, mineral assemblages and properties of each lithology (oceanic crust, lower and upper continental
-crust, sediments and dry or hydrated mantle) are calculated for temperatures from \SIrange{330}{1600}{\kelvin}, with increments of approximately \SI{4}{\kelvin},
-and pressures from \SIrange{0.1}{30}{\giga\pascal}, with increments of approximately \SI{0.03}{\giga\pascal}, for a total of almost 100000 points.
-
-The introduction of phase transitions produces two main effects that must be taken into account in numerical models: variations in density and release/absorption
-of latent heat \citep{Tackley2010}. Effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into
-account considering effective coefficient of thermal expansion \citep{Christensen1985,Zhong2015}. Similarly, effects of latent heat in the energy equation
-(eq \ref{eq:energy}) are taken into account by considering effective specific heat and coefficient of thermal expansion
-\citep{Christensen1985,Tackley2010,Zhong2015}. Effective density, specific heat and coefficient of thermal expansion are included in Perple\_X files and are
-assumed by each Lagrangian marker in according to its $p$-$T$ conditions.
+The introduction of phase transitions produces two main effects that must be taken into account in numerical models: variations in density and release/absorption of latent heat that is required by the extended Boussinesq approximation \citep{Tackley2010}.
+Effects of density variations on buoyancy force in the momentum equation (Eq.
+\ref{eq:momentum_penalty}) are taken into account considering effective coefficient of thermal expansion \citep{Christensen1985,Zhong2015}.
+Similarly, effects of latent heat in the energy equation (eq \ref{eq:energy}) are taken into account by considering effective specific heat and coefficient of thermal expansion
+\citep{Christensen1985,Tackley2010,Zhong2015}.
+Effective density, specific heat and coefficient of thermal expansion are included in Perple\_X files and are assumed by each Lagrangian marker in according to its $p$-$T$ conditions.
 
 Hydration processes related to dehydration of subducting lithosphere \citep{Schmidt1998,Liu2007,Faccenda2009,Faccenda2010,Faccenda2014,Rosas2016} strongly
-influence the thermo-mechanics inside the mantle wedge, mainly because of both weakening effects on the mantle rheology and density variations in case of
-mantle serpentinization \citep{Gerya2002,Honda2003,Arcay2005,Roda2010,Regorda2017}. In the case that hydration is switched on, the amount of bound and free
-water is memorised by each marker, following the implementation of \citet{Quinquis2014}; hydration and dehydration processes are related to the amount of bound
-water of each marker and to the maximum amount of water it can transport, i.e. if the amount of bound water exceeds the maximum amount of water, the marker
-dehydrates and releases free water that can hydrates under-saturated markers. Maximum water content of each marker is determined as function of lithology and
-$p$-$T$ conditions and it is calculated using Perple\_X, in the same way than effective density, specific heat and coefficient of thermal expansion. Bound water
-is advected together with the markers, neglecting the effect of bound water diffusion, while free water simply migrates vertically and is not coupled to the
-solid-phase flow of the mantle wedge \citep{Arcay2005,Quinquis2014}. The correctness of the migration scheme of the free water and its absorption into bound
+influence the thermo-mechanics inside the mantle wedge, mainly because of both weakening effects on the mantle rheology and density variations in case of mantle serpentinization \citep{Gerya2002,Honda2003,Arcay2005,Roda2010,Regorda2017}.
+In the case that hydration is switched on, the amount of bound and free water is memorised by each marker, following the implementation of \citet{Quinquis2014}; hydration and dehydration processes are related to the amount of bound water of each marker and to the maximum amount of water it can transport, i.e.
+if the amount of bound water exceeds the maximum amount of water, the marker dehydrates and releases free water that can hydrates under-saturated markers.
+Maximum water content of each marker is determined as function of lithology and $p$-$T$ conditions and it is calculated using Perple\_X, in the same way than effective density, specific heat and coefficient of thermal expansion.
+Bound water is advected together with the markers, neglecting the effect of bound water diffusion, while free water simply migrates vertically and is not coupled to the solid-phase flow of the mantle wedge \citep{Arcay2005,Quinquis2014}.
+The correctness of the migration scheme of the free water and its absorption into bound
 water by under-saturated markers are verified performing the experiment described in \citet{Quinquis2014} (Sec. \ref{sec:quinquis}).
 
-The mantle viscosity weakening related to the amount of water has been extensively studied \citep{Chopra1981,Hirth2003} and it is implemented in according to
-\citet{Arcay2005} and \citet{Horiuchi2016}, as follows
+The mantle viscosity weakening related to the amount of water has been extensively studied \citep{Chopra1981,Hirth2003} and it is implemented in according to \citet{Arcay2005} and \citet{Horiuchi2016}, as follows
 \begin{equation}\label{eq:wet_visc}
@@
 \end{equation}
-where $f_v$ is the viscosity reduction factor between dry and wet conditions, $[OH^-]$ is the water content and $[OH^-]_0$ is a reference water content set to
-620 ppm (0.062 wt.\%) \citep{Arcay2005}. By default, $f_v$ is set to 100, in according to \citet{Arcay2005} and \citet{Horiuchi2016}.
-
-\section{Mantle melting}\label{sec:melt}
-Melt fraction in the mantle depends on temperature ($T$, in °C), pressure ($p$, in \si{\giga\pascal}) and the water content in the melt ($X_{H_2O}$, in wt.\%).
-The determination of the melt fraction $M$ is implemented as explained by \citet{Katz2003}, according to the successive modification by \citet{Langmuir2006} and
-\citet{Kelley2010}, as follows
-\begin{eqnarray}\label{eq:melt_fr}
-M(p,T,X_{H_2O})&=&-T+T_{sol}+(x_t\cdot ln(p+y_t))\cdot X_{H_2O}-\nonumber\\
-&&+K_m\left(\frac{X_{H_2O}^{bulk}}{D_{H_2O}(1-X_{H_2O})+X_{H_2O}}\right)^{\gamma}
-\end{eqnarray}
-where $T_{sol}$ is the temperature of the dry solidus, $D_{H_2O}$ is the partition coefficient and $x_t\cdot log(p+y_t)$ indicates the pressure dependence of
-the melting curve \citep{Kelley2010}. $x_t$, $y_t$, $K_m$ and $\gamma$ are parameters chosen in according to \citet{Kelley2010}.
-However, $X_{H_2O}$ depends on the melt fraction as
-\begin{equation}\label{eq:melt_wt}
-X_{H_2O}(M)=\frac{X_{H_2O}^{bulk}}{D_{H_2O}+M(1-D_{H_2O})}
-\end{equation}
-and a numerical solution can be found using a root-finder method \citep{Katz2003,Wang2016}. Furthermore, the content of water in the melt is limited by the
-pressure-dependent saturation concentration of water in the melt, determined as
-\begin{equation}\label{eq:melt_max}
-X_{H_2O}^{sat}=\chi_1 p^{\lambda_m}+\chi_2 p
-\end{equation}
-where $\chi_1$, $\chi_2$ and $\lambda_m$ are parameters chosen according to \citet{Katz2003}, \citet{Langmuir2006} and \citet{Kelley2010}. The correctness in
-the determination of melt fractions and water contents in the melt are tested comparing model results with experimental curves obtained by \citet{Katz2003}
-(Sec. \ref{sec:katz}).
-
-In case of melting, viscous creep viscosity ($\eta_{vp}$) is modified according to \citet{Wang2016}, as
-\begin{equation}\label{eq:melt_visc}
-\eta_{melt}=\eta_{vp}\cdot \exp(\alpha_m M)
-\end{equation}
-where $\alpha_m$ are melt fraction factors for dislocation and diffusion creep, chosen in according to \citet{Wang2016}.
-
-Effective density ($\rho_{eff}$) for partially molten rocks is calculated as
-\begin{equation}\label{eq:melt_dens}
-\rho_{eff}=\rho_s(1-M)+\rho_m M
-\end{equation}
-where $\rho_s$ and $\rho_m$ are the densities of the solid and the molten rock, respectively \citep{Gerya2003a,Wang2016}. The density of the solid rock is
-extracted by Perple\_X, while the density of the molten portion is calculated as
-\begin{equation}\label{eq:melt_dens_M}
-\rho_m(p,T)=\rho_0[1-\alpha(T-T_0)][1+\beta(p-p_0)]
-\end{equation}
-where $\rho_0$ is the density at the reference temperature ($T_0$) and pressure($p_0$), and $\alpha$ and $\beta$ are the thermal expansion and compressibility
-coefficients, respectively \citep{Gerya2003a,Wang2016}.
-
-As for the phase transitions, effects of density variations on buoyancy force in the momentum equation (Eq. \ref{eq:momentum_penalty}) are taken into account
-considering effective coefficient of thermal expansion. Similarly, latent heat due to melting/crystallisation can be implicitly included in the energy equation
-(Eq. \ref{eq:energy}) by considering effective specific heat ($C_{p_{eff}}$) and thermal expansion ($\alpha_{eff}$) of the partially molten rocks, in according to
-\begin{eqnarray}
-\label{eq:cp_effective}C_{p_{eff}}&=&C_p+H_L\left(\frac{\partial M}{\partial T}\right)_p \\
-\label{eq:alpha_effective}\alpha_{eff}&=&\alpha+\rho\frac{H_L}{T}\left(\frac{\partial M}{\partial p}\right)_T
-\end{eqnarray}
-where $C_p$ and $\alpha$ are the specific heat and the coefficient of thermal expansion of the solid rock, respectively, and $H_L$ is the latent heat of the
-molten rock \citep{Gerya2003a,Gerya2007,Tackley2010}.
-
-\section{Benchmarks}\label{sec:benchmark}
-\subsection{Error measurements}\label{sec:error}
-In order to determine the accuracy of velocity and pressure field of the benchmarks, the $L_2$-norm is computed by numerical integration on the Gauss-Legendre
-quadrature points. $L_2$-norm for pressure and velocity errors can be evaluated as
-\begin{eqnarray}
-\textrm{err}_p&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|p^n_i(r_q)-p^a_i(r_q)|^2\right)w_q|J_q|}\nonumber \\
-\textrm{err}_v&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|u^n_i(r_q)-u^a_i(r_q)|^2+|v^n_i(r_q)-v^a_i(r_q)|^2\right)w_q|J_q|}\nonumber
-\end{eqnarray}
-respectively, where $n_e$ is the number of elements, $n_q$ is the number of quadrature points per element, $p^n_i(r_q)$ and $p^a_i(r_q)$ are the numerical and
-analytical pressure, respectively, in each quadrature point $q$, $w_q$  and $J_q$ are the weight and the Jacobian at the quadrature point
-$q$, $u^n_i(r_q)$, $v^n_i(r_q)$, $u^a_i(r_q)$ and $v^a_i(r_q)$ are the numerical and analytical velocities, respectively, in each quadrature point $q$.
-
-Other quantities used as comparison with original benchmarks are the root-mean-square velocity over the whole domain and over the surface
-\begin{eqnarray}
-v_{\textrm{rms}}&=&\sqrt{\int_{\Omega}|\bm{v}|^2 d\Omega}\nonumber \\
-v_{\textrm{rms}}^{top}&=&\sqrt{\int_0^1 u^2 \bigg|_{y=1} dx}\nonumber
-\end{eqnarray}
-respectively, top and bottom Nusselt numbers
-\[\textrm{Nu}^{top/bottom}=-\int_0^1\frac{\partial T}{\partial y} \bigg|_{y=1/y=0} dx\]
-the average rate of work done against gravity
-\[\left\langle W \right\rangle=\int_{\Omega}T u_y d\Omega\]
-the average rate of viscous dissipation
-\[\left\langle \Phi \right\rangle=\int_{\Omega}\tau_{ij}\dot{\epsilon}_{ij} d\Omega\]
-and the percentage error between $\left\langle W \right\rangle$ and $\left\langle \Phi \right\rangle/Ra$
-\[\delta=\frac{\left|\left\langle W \right\rangle -\frac{\left\langle \Phi \right\rangle}{Ra} \right|}{\max \left(\left\langle W \right\rangle,
-\frac{\left\langle \Phi \right\rangle}{Ra}\right)}\times100\]
-
-\subsection{Stokes flow}\label{sec:stokes}
-The problem consists of determining velocity field $(u,v)$ and pressure $p$ in case of a manufactured solution with prescribed body forces such as
-\begin{eqnarray}
-b_1&=&(12-24y)x^4+(-24+48y)x^3+(-48y+72y^2-48y^3+12)x^2+\nonumber \\
-&&+(-2+24y-72y^2+48y^3)x+1-4y+12y^2-8y^3\nonumber \\
-b_2&=&(8-48y+48y^2)x^3+(-12+72y-72y^2)x^2+(4-24y+48y^2-48y^3+\nonumber \\
-&&+24y^4)x-12y^2+24y^3-12y^4\nonumber
-\end{eqnarray}
-for which the exact solution is:
-\begin{eqnarray}
-u(x,y)&=&x^2(1-x)^2(2y-6y^2+4y^3)\nonumber \\
-v(x,y)&=&-y^2(1-y)^2(2x-6x^2+4x^3)\nonumber \\
-p(x,y)&=&x(1-x)-\frac{1}{6}\nonumber
-\end{eqnarray}
-The domain is a unit square a constant viscosity ($\eta=1$) and the penalty parameter is set to $\lambda=10^7$. Velocity boundary conditions are set to no slip
-$(\bm{v}=\bm{0})$ on all boundaries. The problem is performed for different grid resolution between \num{8x8} and \num{1024x1024} elements.
-
-The errors between the analytical solution and the numerical prediction of the velocity and the pressure field are calculated as in Section \ref{sec:error}.
-Fig. \ref{fig:errors}a shows that both velocity and pressure field converge to the exact solution with the decrease of the element size, following the
-theoretical convergence rate. The convergence of pressure, in contrast with the observation by \citet{Donea2003} for $Q_1 \times P_0$ elements, indicate the
-effectiveness of the smoothing procedure. Smoothed pressure field for grid resolution of \num{128x128} elements is shown in Fig. \ref{fig:errors}b (continuous
-red line), in comparison with the analytical solution (dashed blue line). Solve times and memory usage needed to generate the solution are shown in
-Fig. \ref{fig:MUMPS}. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Solver/Stokes_Flow}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/errors.pdf}
-\caption{Velocity and pressure error for the Stokes flow experiment between generated and analytical solution as a function of element size (panel a),
-and comparison between smoothed pressure and analytical solution as function of $x$ coordinate for a grid resolution of \num{128x128} elements (panel b).}
-\label{fig:errors}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/MUMPS.pdf}
-\caption{Panel a) Analysis, factorisation, solution and total solve times (panel a) and factorisation memory usage (panel b) as a function of the total number
-of degrees of freedom for the Stokes flow experiment.}
-\label{fig:MUMPS}
-\end{figure}
-
-\subsection{Markers advection}\label{sec:runge}
-The 2nd-order Runge-Kutta advection scheme in space is tested by means of the Zalesak disk test \citep{Zalesak1979,Thieulot2014}. The benchmark is performed
-in a unit square domain with a grid resolution of \num{32x32} elements and 1000000 markers and values of Courant number between 0.25 and 1.
-At $t=0$, the disk is centred at position $(0.5;0.75)$ with a radius $R=0.15$ and has a vertical fissure 0.05 wide and 0.2 high (orange in Fig. \ref{fig:runge})
-and with the velocity field prescribed in the entire domain as
-\begin{eqnarray}
-u(x,y)&=&2\pi\left(y-\frac{L_x}{2}\right)\nonumber \\
-v(x,y)&=&-2\pi\left(x-\frac{L_x}{2}\right)\nonumber
-\end{eqnarray}
-the markers are back to their initial location after a $2\pi$ rotation ($t=1$). The clockwise rotation of the disk at $t=0.25$, 0.5 and 0.75 
-is shown in Fig. \ref{fig:runge} (red, blue and green, respectively). The distance from the centre of 1 marker is calculated throughout the rotation to
-evaluate the error for different values of Courant number. As expected, the error increase with the increasing of the time step, as shown in 
-Fig. \ref{fig:runge_err}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/Zalesak_Disk}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=200px]{./Figures/Zalesak.pdf}
-\caption{Position of the Zalesak disk at $t=0$, 0.25, 0.5 and 0.75 (orange, red, blue and green, respectively) for the case with Courant number of 1.}
-\label{fig:runge}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/zal_marker.pdf}
-\caption{Distance from the centre as function of time for values of Courant number of  of 0.25, 0.3, 0.5, 0.75 and 1 (green, orange, blue, black and red,
-respectively). R indicates the distance from the centre considering a perfect circle trajectory.}
-\label{fig:runge_err}
-\end{figure}
-
-\subsection{Conservative Velocity Interpolation (CVI)}\label{sec:CVI}
-The CVI correction is checked by means of the Stokes flow experiment by \citet{Donea2003}, as presented in \ref{sec:stokes}. The advection of Lagrangian
-markers is performed using either a 2nd-order or a 4th-order Runge-Kutta scheme with CFL=0.5 and an initial random distribution of 25 markers per element.
-Fig. \ref{fig:CVI} shows the minimum and maximum values of markers per element throughout the experiment with and without the CVI correction and using
-different Runge-Kutta schemes. The introduction of the CVI correction allows the
-tests to run for times longer than $t=2000$ without show empty elements, while tests without the CVI correction stop at approximately $t=600$, when at least
-one element goes empty. Fig. \ref{fig:CVI_time} shows the different number of markers per element and the markers distribution at $t=600$ obtained without and
-with the CVI correction (panels b and c, respectively) in case of a 2nd-order Runge-Kutta scheme. All data can be found at
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Advection/CVI}.
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/CVI.pdf}
-\caption{Maximum and minimum number of markers per element throughout the experiment, with (green and orange lines) or without (purple and blue lines)
-the CVI correction and using either a 2nd-order (blue and orange lines) or a 4th-order (purple and green lines) Runge-Kutta scheme.}
-\label{fig:CVI}
-\end{figure}
-\begin{figure}
-\centering
-\noindent\includegraphics[width=\textwidth]{./Figures/CVI_time.pdf}
-\caption{Number of markers per element and markers distribution through time using 2nd.order Runge-Kutta scheme.
-Panels a: number of markers per element (panel a1) and distribution of the markers (panel a2) after the first
-iteration; panels b and c: comparison between number of markers per element (panels b1 and c1) and markers distribution (panels b2 and c2) without and with
-the CVI correction at $t=600$; panels d: number of markers per element (panel d1) and markers distribution (panel d2) in case of CVI correction at $t=2000$.}
-\label{fig:CVI_time}
-\end{figure}
-
-\subsection{Poiseuille flow}\label{sec:poiseuille}
-The domain is a rectangle with $L_x=2$ and $L_y=1$ and constant density and viscosity ($\rho=1$ and $\eta=1$), gravity acceleration $\bm{g}=\bm{0}$ and penalty
-parameter $\lambda=10^8$. The grid is composed by \num{40x20} elements. Velocity boundary conditions are set to no slip $(\bm{v}=\bm{0})$ at the top and the
-bottom, and a parabolic profile is imposed on the sides, with $u=y(1-y)$ and $v=0$. The analytical solution is then given by:
-\begin{eqnarray}
-u(x,y)&=&y(1-y)\nonumber \\
-v(x,y,)&=&0\nonumber \\
-p(x,y)&=&2\eta\left(\frac{L_x}{2}-x\right)\nonumber
-\end{eqnarray}
-
-The velocity field predicted by the model follows the expected parabolic profile (Fig. \ref{fig:poiseuille}a and Fig. \ref{fig:poi_plot}a) and, in case of the
-classic penalty method with no iterations, the pressure is clearly related to the divergence by means of the penalty parameter (Fig. \ref{fig:poiseuille}b and
-c and Fig. \ref{fig:poi_plot}b). Fig. \ref{fig:poiseuille}d shows that one Uzawa iteration is sufficient to bring the divergence down to \num{e-15}, with no
-correlation with the pressure field. All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Poiseuille%20Flow}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Poiseuille.pdf}
-\caption{Velocity field (panel a), pressure (panel b) and divergence velocity of a Poiseuille flow in case of the classic penalty method (no iterations) and
-after one Uzawa iteration (panel c and d, respectively).}
-\label{fig:poiseuille}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Analytical.pdf}
-\caption{Velocity (panel a) and pressure (panel b) field predicted by the model for a Poiseuille flow with respect with their analytical solutions. The
-velocity field is plotted as function of the vertical coordinate in $L_x/2$ and the pressure is plotted as function of the x coordinate in $L_y/2$.}
-\label{fig:poi_plot}
-\end{figure}
-
-\subsection{Instantaneous 2D sphere}\label{sec:ist_sphere}
-The domain is a unit square with gravity $\bm{g}=(0,-1)$. The fluid has constant density and viscosity ($\rho_f=1$ and $\eta_f=1$). The sphere is in the middle
-of the domain with a radius R = 0.123456798 and has constant density and viscosity ($\rho_s=10^{-2}+\rho_f$ and $\eta_s=10^3 \cdot \eta_f$).
-We distinguish three different types of velocity boundary conditions:
-\begin{enumerate}
-\item FS: free slip conditions on all sides.
-\item NS: no slip conditions on all sides.
-\item OT: free slip conditions on the sides and the bottom, and open boundary at the top.
-\end{enumerate}
-All three cases are performed for different grid resolutions, between \num{16x16} and \num{512x512} elements with 50 randomly distributed markers per element,
-and different average schemes for the viscosity (harmonic, geometric and arithmetic). The velocity in the centre of the sphere $\bm{v}(0.5;0.5)$), minimum and
-maximum velocities ($u_{min}$, $u_{max}$, $v_{min}$, $v_{max}$) and pressures ($p_{min}$, $p_{max}$) on the entire domain, $v_{\textrm{rms}}$ and average pressure on the
-entire domain ($p_{avg}$) are compared with solutions generated by ASPECT \citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} and all results
-can be found at \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere2D} and 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Instantaneous_2D_Sphere}. Results in terms of the $v_{\textrm{rms}}$ for NS model 
-are also shown in Fig. \ref{fig:inst_sphere}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_NS.pdf}
-\caption{$v_{\textrm{rms}}$ as function of element size for the instantaneous 2D sphere experiment in case of no slip boundary conditions and with different average
-schemes for the viscosity. Results are compared with results obtained by other numerical codes.}
-\label{fig:inst_sphere}
-\end{figure}
-
-\subsection{Rayleigh-Taylor instability}\label{sec:rayleigh}
-This problem has been originally presented by \citet{vanKeken1997} and here it is performed as the isoviscous case (Case 1a). The domain has $L_x=0.9142$ and
-$L_y=1$ and gravity $\bm{g}=(0,-1)$. Two fluids with same constant viscosities ($\eta_1=\eta_2=1$) and different densities ($\rho_1=1000$ and $\rho_2=1010$),
-with the lighter fluid at the bottom. The initial interface between the fluids is given by $y(x)=0.2+0.02 \cos \left(\frac{\pi x}{L_x}\right)$. The experiment
-is performed with different grid sizes (\num{50x50}, \num{80x80}, \num{100x100} and \num{256x256}). A total of 1960000 markers are randomly distributed at the
-beginning of the simulation. Velocity boundary conditions are set to no slip at the top and the bottom, and to free slip at the sides of the domain.
-
-$v_{\textrm{rms}}$ as function of time is reported in Fig. \ref{fig:RT}, matching well with results shown by \citet{vanKeken1997}, \citet{Tackley2003} and
-\citet{Thieulot2014}. Fig. \ref{fig:rayleigh} shows the evolution of the experiment at different time steps. All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Rayleigh_Taylor_experiment/ISOVISCOUS}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/RT.pdf}
-\caption{$v_{\textrm{rms}}$ of the Rayleigh-Taylor experiment as function of time for different resolution of the grid.}
-\label{fig:RT}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=300px]{./Figures/Rayleigh.pdf}
-\caption{Evolution of the Rayleigh-Taylor experiment for a grid of \num{256x256} elements at $t=0, 500, 1000$ and $2000$ (panels a, b, c and d, respectively).}
-\label{fig:rayleigh}
-\end{figure}
-
-\subsection{Falling blocks}\label{sec:block}
-This benchmark is described as proposed by \citet{Gerya2003a}, \citet{Gerya2010b} and \citet{Thieulot2011}. The domain is square with $L_x=L_y=$ \SI{500}{\km}
-and the grid is composed by \num{50x50} elements with 25 markers in each element. The block is initially centred at ($x=$ \SI{250}{\km}; $y=$ \SI{400}{\km})
-and has a size of \num{100x100} km (Fig. \ref{fig:block}a). The fluid surrounding the block has $\eta_f=$ \SI{e21}{\pascal\per\s} and $\rho_f=$
-\SI{3200}{\kg\per\cubic\m}. The benchmark tests with different viscosities of the block, with $\eta_b$ from \SIrange{e15}{e27}{\pascal\per\s}. In each
-experiment also the density $\rho_b$ of the block varies from \SIrange{3220}{9900}{\kg\per\cubic\m}. Velocity boundary conditions are set to free slip
-conditions on all sides of the domain. Density distributions at $t=$ \SI{20}{\mega\year} for $\rho_b=$ \SI{3300}{\kg\per\cubic\m} and different $\eta_b$
-are plotted in Fig. \ref{fig:block}b-f, showing that the code correctly preserve the block geometry in case of large viscosity contrast, with a block stiffer
-than the surrounding fluid (Fig. \ref{fig:block}f).
-
-The velocity in the centre of the falling block at $t=0$ is measured for all experiments and, since the velocity should increase with density contrast, the
-quantity $\bm{v}/(\rho_b-\rho_f)$ is plotted as function of the viscosity contrast. All results perfectly match with those from \citet{Gerya2010b} and they
-line up on a single curve, demonstrating that the code can correctly deal with large viscosity and density contrasts (Fig. \ref{fig:falling}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Falling%20blocks}.
-
-\begin{figure}
-\noindent\includegraphics[width=300px]{./Figures/Block.pdf}
-\centering
-\caption{Density evolution of the falling block experiment at $t=$ \SI{0}{\mega\year} (panel a) and $t=$ \SI{20}{\mega\year} for different viscosities of the
-block (panels b-f).}
-\label{fig:block}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Falling.pdf}
-\caption{Initial velocity relative to the density contrast at the centre of the falling block as function of the viscosity contrast between the block and the
-surrounding fluid.}
-\label{fig:falling}
-\end{figure}
-
-\subsection{2D time-dependent Stokes sphere with deformable free surface}\label{sec:sphere}
-This experiment is performed in a unit square domain with the gravity acceleration fixed to $g_y=-1$. The sphere has $\rho_s=2$ and $\eta_s=10^3$ and is
-initially centred at ($0.5;0.6$) with radius $R=0.123456789$. The fluid surrounding the sphere has $\rho_f=1$ and $\eta_f=1$ and occupies the domain for
-$y\leq 0.75$, while for $y>0.75$ the air has $\rho_a=0$ and $\eta_a=10^{-3}$. Velocity boundary conditions are set to free slip on all sides. The Courant
-number is set to 0.25. The experiments run for 200 s using grid resolutions from \num{150x150} to \num{512x512} elements, with 25 randomly distributed markers
-per element, and for different average schemes for the viscosity. The interface between the fluid and the air is tracked by means of the markers chain. Results
-in terms of velocity and pressure fields and topography variations are compared with results obtained with ASPECT
-\citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} and they can be found at
-\url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere_fs2D} and
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Time_dependent_sphere}. Results in terms of the $v_{\textrm{rms}}$ in case of 
-an arithmetic average are also shown in Fig. \ref{fig:inst_spherefs}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_arithm.pdf}
-\caption{$v_{\textrm{rms}}$ for different grid resolutions as function of time for the 2D time-dependent Stokes sphere experiment in case of an arithmetic average.
-Results are compared with results obtained by ASPECT.}
-\label{fig:inst_spherefs}
-\end{figure}
-
-\subsection{Free surface stabilisation}\label{sec:stab}
-The benchmark is performed as discussed in \citet{Kaus2010a} and \citet{Thieulot2014}. The domain is square with $L_x=L_y=$ \SI{500}{\km} and a grid resolution
-of \num{200x200} elements, each of them containing 16 markers. A buoyant fluid with $\rho_1=$ \SI{3200}{\kg\per\cubic\metre} and $\eta_1=$ \SI{e20}{\pascal\s}
-is overlain by a denser fluid with $\rho_2=$ \SI{3300}{\kg\per\cubic\metre} and $\eta_2=$ \SI{e21}{\pascal\s}. The initial interface between the fluids has an
-initial sinusoidal shape of 5 km amplitude. Velocity boundary conditions are set to free slip conditions at the sides and no slip at the bottom, while the top
-is free surface. The experiment is performed with various fixed time steps, without the use of the Courant number. The vertical position of the free surface at
-$x=L_x$ is tracked for each simulation, with or without the stabilisation algorithm.
-
-The results show an instability (drunken sailor effect) increasing the time step in case that the stabilisation algorithm is not activated
-($dt=$ \SI{4200}{\year}, green line in Fig. \ref{fig:stab}), as already observed by \citet{Kaus2010a} and \citet{Thieulot2014}. Activating the algorithm the
-instability is fixed and the simulation remains stable using time step up to \SI{50000}{\year} (red line in Fig. \ref{fig:stab}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Stabilisation_algorithm}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Kaus_random.pdf}
-\caption{Evolution of the y coordinate in $x=L_x$ as function of time for different time steps, with and without the stabilisation algorithm.}
-\label{fig:stab}
-\end{figure}
-
-\subsection{Topography relaxation}\label{sec:crameri}
-This experiment is performed as in \citet{Crameri2012} using free surface and sticky air (Case 1). The domain is rectangular with $L_x=$ \SI{2800}{\km} and
-$L_y=$ \SI{700}{\km} in case of free surface and $L_y=$ \SI{800}{\km} in case of sticky air ($h_{st}=$ \SI{100}{\km}). The grid is composed by \num{256x64}
-elements for the free surface \num{560x320} elements for the sticky air. For both cases, there is a mantle of 600 km thickness with $\eta_m=$ \SI{e21}{\pascal\s}
-overlain by a highly viscous cosine-shaped lithosphere with thickness between 93 and 107 km and $\eta_l=$ \SI{e23}{\pascal\s}. Both mediums have $\rho=$
-\SI{3300}{\kg\per\cubic\metre}. In case of sticky air experiment, the air has $\rho_a=$ \SI{0}{\kg\per\cubic\metre} and $\eta_a=$ \SI{e18}{\pascal\s}. The
-gravity acceleration is set to $g_y=$ \SI{-10}{\metre\per\square\s}. Velocity boundary conditions are set to no slip at the bottom and free slip at the sides
-of the domain. In case of sticky air, velocities on top are set to free slip conditions.
-
-The maximum topography as function of time ($h(t)$) can be analytically derived in according to \[h(t)=h_0 \exp(\gamma t)\]
-where $h_0=7$ km is the initial topography, $\gamma=\num{-0.2139e-11}$ is the characteristic relaxation rate and $t$ is the time (black line in Fig.
-\ref{fig:crameri}). Maximum topography at the characteristic relaxation time $t_{rlx}=$ \SI{14.825}{\kilo\year} can be found to be $h_{rlx}=$ \SI{2576}{\m}.
-The results show that a viscosity of the air of \SI{e18}{\pascal\s} is needed in case of $h_{st}=$ \SI{100}{\km} to correctly describe the topography relaxation
-for this problem, as pointed out by \citet{Crameri2012} (purple line in Fig. \ref{fig:crameri}). The case with a true free surface well follows the analytical
-solution (orange line in Fig. \ref{fig:crameri}). In particular, at the relaxation time in case of the true free surface predicts a maximum topography of
-\SI{2570}{\m}, with an error of \SI{6}{\m} with respect to the analytical solution. Results are also compared with results obtained with a free surface in
-UNDERWORLD (\num{256x64} elements) and SULEC (\num{401x201} elements) (with courtesy from \citealp{Crameri2012}). All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Topography_relaxation}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Crameri.pdf}
-\caption{Maximum topography as function of time for the Crameri benchmark (coloured lines) in comparison with the analytical solution (black line) and results
-obtained with UNDERWORLD (yellow line) and SULEC (blue line).}
-\label{fig:crameri}
-\end{figure}
-
-\subsection{Spontaneous subduction}\label{sec:subduction}
-This experiment is performed as presented by \citet{Schmeling2008}, in case of both a sticky air and a true free surface and using different average schemes
-for the viscosities. The domain is rectangular with $L_x=$ \SI{3000}{\km} and $L_y=$ \SI{700}{\km}, a grid resolution of \num{256x64} elements with 50 markers
-per element and Courant number of 0.01. At the beginning of the simulation, a 100 km-thick lithospheric layer with $\rho_m=$ \SI{3300}{\kg\per\cubic\metre}
-and $\eta_m=$ \SI{e23}{\pascal\s} is located at the top of the domain for $x$ between 1000 and 3000 km, while the asthenospheric mantle has $\rho_m=$
-\SI{3200}{\pascal\s} and $\eta_m=$ \SI{e21}{\pascal\s}. In addition, a 200 km-depth lithospheric slab is already subducted in the mantle in order to have
-a spontaneous subduction. In case of sticky air, $L_y=750$ km and a 50 km-thick air layer with $\eta_a$ from \SIrange{e19}{e21}{\pascal\s} overlie the mantle.
-Velocity boundary conditions are set to free slip on the sides and on the bottom of the domain. In case of sticky air, velocities on the top boundary are set
-to free slip conditions as well.
-
-The maximum depth of the slab is tracked for all the simulations (Fig. \ref{fig:subduction}) and they are compared with results obtained by
-\citet{Schmeling2008} with models with regular grids and comparable grid resolutions (rectangles in Fig. \ref{fig:subduction}). Low viscous sticky air models
-(i.e., \SI{e19}{\pascal\s}) enlighten a strong dependence of the sinking velocity with the chosen average scheme (dashed lines in Fig. \ref{fig:subduction}),
-while high viscous sticky air models (i.e., \SI{e21}{\pascal\s}) show a higher resistance at the trench, underestimating the correct solution (dotted lines in
-Fig. \ref{fig:subduction}). A high resistance at the trench can be also observed in case of a true free surface for the low resolution grid used (continuous
-lines in Fig. \ref{fig:subduction}), as pointed out by \citet{Schmeling2008}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Surface_processes/Spontaneous_subduction}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Subduction.pdf}
-\caption{Maximum depth of the slab as function of time for all the simulations of the \citet{Schmeling2008} benchmark, in case of harmonic, geometric and
-arithmetic means (red, green and blue lines, respectively) using a sticky air or a true free surface (discontinuous and continuous lines, respectively).
-Red, green and blue rectangular areas indicate the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{500}{\km} in case of sticky air,
-while the grey rectangular area indicates the range of times from \citet{Schmeling2008} when the slab tip reaches \SI{600}{\km} in case of true free surface.}
-\label{fig:subduction}
-\end{figure}
-
-\subsection{The slab detachment}\label{sec:slab}
-The slab detachment benchmark is performed as described by \citet{Schmalholz2011} and \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{1000}{\km}
-and $L_y=$ \SI{660}{\km} and a grid resolution of \num{256x256} elements. A non-linear viscous T-shaped layer with $\rho_l=$ \SI{3300}{\kg\per\cubic\metre} is
-placed at the top of the domain and surrounded by a linear viscous fluid with $\rho_f=$ \SI{3150}{\kg\per\cubic\m}. The top layer is 80 km-thick and a 250
-km-long and 80 km-wide slab is placed at $x=L_x/2$ (Fig. \ref{fig:slab}a). The effective viscosity of the top layer is given by
-\[\eta_{eff}=\eta_0 I_2^{\frac{(1-n)}{n}}\]
-with $\eta_0=$ \SI{4.75e11}{\pascal\s} and $n=4$, while the surrounding fluid has $\eta_f=$ \SI{e21}{\pascal\s}. The effective viscosity is capped between
-$\eta_{min}=$ \SI{e21}{\pascal\s} and $\eta_{max}=$ \SI{e25}{\pascal\s}. Velocity boundary conditions are set to free slip at the top and the bottom, and to
-no slip at the sides of the domain.
-
-The sinking of the slab determines a decrease of the effective viscosity in the T-shaped layer during the first part of the evolution (Fig. \ref{fig:slab}a),
-as consequence of local strain rates, while effective viscosities increase up to $\eta_{max}$ in the top layer after approximately \SI{20}{\mega\year}, when
-necking is complete (Fig. \ref{fig:slab}b).
-
-The width of necking is tracked by means of markers position at the side of the slab. Necking width and time are normalised in function of the initial slab
-(\SI{80}{\km}) and the characteristic time ($t_c=$ \SI{7.1158e14}{\s}) \citep{Schmalholz2011,Glerum2018}. The results in case of arithmetic, geometric and
-harmonic mean of the viscosity are shown in Fig. \ref{fig:necking}, compared with results from \citet{Glerum2018} for which an infinity norm mean was used.
-The influence of the average schemes agrees with results shown in Section \ref{sec:subduction}. All data can be 
-found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Slab_detachment}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Slab.pdf}
-\caption{Effective viscosity for the slab detachment benchmark at the beginning of the evolution (panel a) and when necking is complete (panel b).}
-\label{fig:slab}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Necking.pdf}
-\caption{Normalised width of the necking of the slab detachment benchmark as function of normalised time.}
-\label{fig:necking}
-\end{figure}
-
-\subsection{The indenter experiment}\label{sec:indenter}
-The indenter benchmark simulate a rigid punch on a purely plastic von Mises material, for which exists an analytical solution
-\citep{Thieulot2008,Thieulot2014,Glerum2018}. The domain is a unit square with a grid resolution of \num{256x256} elements. The gravity acceleration is set
-to $\bm{g}=\bm{0}$ and effective viscosity is capped using $\eta_{min}=10^{-4}$ and $\eta_{max}=10^3$. The tolerance for the convergence of the non-linear
-solution is $tol=10^{-9}$, with a maximum number of non-linear iterations set to 500. The medium in the domain has $\rho=0.01$, an initial viscosity
-$\eta_0=10$, cohesion $C=1$ and an angle of internal friction $\phi=0$°. Velocity boundary conditions are set to no slip at the bottom and free slip at the
-sides of the domain. The top of the domain is open with the exception of the central portion (punch area) with width $w_p=0.125$, where $v=-1.05$ and $u$ is
-fixed either to 0 (rough punch experiment) or free (smooth punch experiment).
-
-Fig. \ref{fig:indenter} show the results in terms of viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure
-(panels g and h) for both experiments. The analytical solution indicate that pressure at ($0.5;1$) and ($0.5 \pm w_p;1$)is $p=\pi +1$ and $p=1$, respectively
-(black and red lines in Fig. \ref{fig:smooth_p}). Fig. \ref{fig:smooth_p} shows a clear improvement in the pressure solution when passing from a rough to a
-smooth punch (green and blue lines in Fig. \ref{fig:smooth_p}, respectively), as previously observed by \citet{Thieulot2014} and \citet{Glerum2018}. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Indenter}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Indenter.pdf}
-\caption{Viscosity (panels a and b), strain rate (panels c and d), velocity (panels e and f) and pressure (panels g and h) fields for rough (left column) and
-smooth (right column) punch experiments.}
-\label{fig:indenter}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Smooth_Rough.pdf}
-\caption{Nodal pressure in function of the x coordinate at $y=1$ for rough (green line) and smooth (blue line) punch experiment. Black and red lines indicate
-the analytical solution at $x=0.5$ and $x=0.5 \pm w_p$, respectively.}
-\label{fig:smooth_p}
-\end{figure}
-
-\subsection{The brick experiment}\label{sec:brick}
-An instantaneous version of the brick benchmark is performed to verify the correctness of the pressure-dependent plasticity for different angle of internal
-friction, as proposed by \citet{Glerum2018}. The domain is rectangular with $L_x=$ \SI{40}{\km} and $L_y=$ \SI{10}{\km} and a grid resolution of \num{512x128}
-elements. The gravity acceleration is set to $g_y=$ \SI{-10}{\m\square\s} and effective viscosity is capped using $\eta_{min}=$ \SI{e19}{\pascal\s} and
-$\eta_{max}=$ \SI{e26}{\pascal\s}. The tolerance for the convergence of the non-linear solution is $tol=10^{-7}$, with a maximum number of non-linear iterations
-set to 1000. A 800 m-wide and 400 m-height inclusion with $\eta_b=$ \SI{e20}{\pascal\s} and $\rho_b=$ \SI{2700}{\kg\per\cubic\m} is placed at the bottom of the
-domain at $x=L_x/2$. The inclusion is surrounded by a non-linear viscous medium with $\rho_m=$ \SI{2700}{\kg\per\cubic\m}, an initial viscosity of
-\SI{e23}{\pascal\s}, a linear viscous viscosity of \SI{e25}{\pascal\s} and a cohesion of \SI{40}{\mega\pascal}. Velocities are set to free slip conditions at
-the bottom of the domain and the top is open. Velocities on sides of the domain are fixed to $u=$ \SI{+-2e-11}{\m\per\s} and $v=0$. The experiment is performed
-in compressional and extensional contexts with an angle of internal friction $\phi$ of the non-linear medium variable between 0° and 30°.
-
-As expected, two shear bands stem from the inclusion with variable angles in relation to both the dynamics context and the internal friction angle. Shear band
-angles formed at 45° for $\phi=0$° in both extensional and compressional contexts (Fig. \ref{fig:brick_beam}a), while they are different in case of $\phi\neq0$°.
-Shear band angles in case of $\phi=20$° are shown for compression and extension in Fig. \ref{fig:brick_beam}b  and c, respectively. Values of shear band angles
-as function of internal friction angles from 0° to 30° are extracted at different depths and minimum and maximum values are plotted in Fig. \ref{fig:brick},
-in comparison with theoretical Roscoe, Arthur and Coulomb shear band angles. For all tests the tolerance for nonlinear convergence is defined $tol=10^{-7}$
-and none converges before the maximum number of iterations ($it_{max}=1000$) is reached. However, tests with internal friction angles up to 15° show a constant
-decrease of the velocity residuals, which can not be observed for for tests with higher internal friction angles (Fig. \ref{fig:convergence}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Nonlinear_visco_plasticity/Brick_experiment}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Brick_Beam.pdf}
-\caption{Shear band angles predicted for the brick experiment for $\phi=0$° (panel a) and for $\phi=20$° in case of compression and extension (panels b and c,
-respectively).}
-\label{fig:brick_beam}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Brick.pdf}
-\caption{Shear band angles predicted for the brick experiment in case of compressional and extensional contexts as function of different internal angle of
-friction (continuous black line and red dots), compared with theoretical Roscoe, Artur and Coulomb angles (discontinuous black lines). Red lines indicate the
-range of angles calculated at different depths.}
-\label{fig:brick}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \noindent\includegraphics[width=500px]{./Figures/convergence.pdf}
-  \caption{Convergence rates for all tests. Normalised residuals (left column) are compared with velocity residuals (right column) for extensional (first row)
-  and compressional (second row) settings.}
-  \label{fig:convergence}
-  \end{figure}
-
-\subsection{Advection stabilisation}\label{sec:advection}
-The 1D advection problem proposed by \citet{Donea2003} and \citet{Thieulot2011} is performed to verify the effectiveness of the SUPG method to stabilise the
-advection term of the energy equation. The domain is a 1D segment with $L_x=1$ composed by 50 elements and a discontinuity in the thermal field placed at
-$x=0.25$. Temperature is set to 1 for $x \leq 0.25$ and to 0 for $x > 0.25$. Velocity is set to ${u}=1$ in the entire domain. The simulation is performed for
-250 time steps, with $dt=0.002$, so the thermal discontinuity should be at $x=0.75$ at the end of the simulation.
-
-Temperature profiles at the end of the simulation are shown in \ref{fig:advection} as function of the dimensionless coefficient $\gamma=\tau \bm{v}/h$. In case
-of the classic Galerkin method ($\gamma=0$, blue line in \ref{fig:advection}) the final thermal profile is characterised by strong oscillations, which are
-eliminated in case of the SUPG method ($\gamma=0.045$, orange line in \ref{fig:advection}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Advection_stabilisation}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Advection.pdf}
-\caption{Temperature profile as function of \textit{x} for the advection stabilisation benchmark. Purple line indicates the initial temperature profile; the
-green line indicates the analytical temperature profile after 250 time steps; blue line indicates the temperature profile after 250 time steps in case of the
-classic Galerkin method ($\gamma=0$); orange line indicates the temperature profile after 250 time steps in case of the SUPG method ($\gamma=0.045$).}
-\label{fig:advection}
-\end{figure}
-
-\subsection{Shear heating}\label{sec:simple_shear}
-This exercise is performed in an unit square domain composed by \num{128x128} elements. The velocity field is prescribed on the entire domain with
-$\bm{v}=(L_y-y)y\bm{e_x}$; viscosity, density and specific heat are fixed to 1, while thermal conductivity and radiogenic energy are fixed to 0. Therefore,
-the energy equation (Eq. \ref{eq:energy}) can be simplified as \[\frac{\partial T}{\partial t}=H_s\] and fixing $T(t=0)=0$, the temperature can be find as
-\[T(t)=H_s t\]
-In this case we have
-\begin{eqnarray}
-\dot{\epsilon_{xx}}&=&\frac{\partial u}{\partial x}=0 \nonumber \\
-\dot{\epsilon_{yy}}&=&\frac{\partial v}{\partial y}=0 \nonumber \\
-\dot{\epsilon_{xy}}&=&\frac{1}{2}\left(\frac{\partial v}{\partial x}+\frac{\partial u}{\partial y}\right)=\frac{1}{2}(L_y-2y)\nonumber
-\end{eqnarray}
-and, simplifying Eq. \ref{eq:shear}, shear heating can be calculated as \[H_s(x,y)=(1-2y)^2\]
-
-The solution predicted by the model in terms of velocity, temperature and shear heating match well with the analytical solutions (Fig. \ref{fig:analytical_en}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Simple_shear}.
-
-%\begin{figure}
-%\centering
-%\noindent\includegraphics[width=400px]{./Figures/Energy_Thieulot.pdf}
-%\caption{Temperature and shear heating predicted by the model after 1 time step.}
-%\label{fig:en_thieulot}
-%\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=500px]{./Figures/Analytical_Shear.pdf}
-\caption{Velocity (panel a), temperature (panel b) and shear heating (panel c) as function of the \textit{y} coordinate for the simple shear experiment. The
-solutions predicted by the model (red lines) are compared with the analytical solutions (dashed blue lines).}
-\label{fig:analytical_en}
-\end{figure}
-
-\subsection{Shear and adiabatic heating}\label{sec:shear}
-This problem is performed as presented in Exercise 9.4 in \citet{Gerya2010b}. Two materials are vertically separated in a rectangular domain with $L_x=$
-\SI{1000}{\km}, $L_y=$ \SI{1500}{\km} and a grid resolution of \num{30x20} elements. Constant thermal coefficient expansion ($\alpha=$ \SI{3e-5}{\per\kelvin}),
-temperature ($T=$ \SI{1300}{\kelvin}) and gravity acceleration ($g_y=$ \SI{-10}{\m\square\s}) are assumed in the whole domain. Fluid 1 (on the left side of the
-domain) has $\rho_1=$ \SI{3200}{\kg\per\cubic\m} and $\eta_1=$ \SI{e20}{\pascal\s}; fluid 2 (on the right side of the domain) has $\rho_2=$
-\SI{3300}{\kg\per\cubic\m} and $\eta_2=$ \SI{e22}{\pascal\s}. Velocity boundary conditions are set to free slip on all sides of the domain.
-
-As shown in Fig. \ref{fig:energy}, both shear and adiabatic heating predicted by the code (first row in Fig. \ref{fig:energy}) well recreate the results
-obtained by means of example ${Shear\_adiabatic\_heating.m}$ from \citet{Gerya2010b} (second row in Fig. \ref{fig:energy}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Adiabatic%2BShear}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Energy.pdf}
-\caption{Comparison between shear (first column), adiabatic (second column) and total (third column) energy predicted by the code (panels a, b and c) and those
-created using example ${Shear\_adiabatic\_heating.m}$ from Exercise 9.4 in \citet{Gerya2010b} (panels d, e and f).}
-\label{fig:energy}
-\end{figure}
-
-\subsection{Mantle convection}\label{sec:mantle}
-This problem is performed as presented by \citet{Blankenbach1989} (constant viscosity cases) and \citet{Thieulot2014}, in a 2D unit square domain with gravity
-acceleration $g_y=10^{10}Ra$. The experiment is performed with three different Rayleigh numbers ($Ra=10^4$, $10^5$ and $10^6$) and with different grid
-resolution (between \num{32x32} and \num{128x128} elements). The fluid has constant viscosity, initial density, heat capacity, thermal conductivity
-($\eta= \rho_0= C_p = k=1$), reference temperature ($T_0=0$) and thermal expansion coefficient ($\alpha=10^{-10}$). Temperatures are set to 0 on top and 1 on
-bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\]
-
-The solution generated by the code in terms of $v_{\textrm{rms}}$ (calculated as in Sec. \ref{sec:error}) and Nu as function of time are reported for all the
-simulations in Fig. \ref{fig:mantle}. At the steady state, $v_{\textrm{rms}}$ and Nu of all simulations converge well toward the values from \citet{Blankenbach1989},
-with lower errors for higher resolution grids (Table \ref{tab:mantle}). All data can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Mantle_convection}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Convection.pdf}
-\caption{$v_{\textrm{rms}}$ (panels a, c and e) and Nu (panels b, d and f) for the mantle convection benchmark as function of time for different grid resolution.
-Panels a and b show the results for $Ra=10^4$; Panels c and d show the results for $Ra=10^5$; panels e and f show the results for $Ra=10^6$. Purple lines
-indicate the convergence values for the $v_{\textrm{rms}}$ and the Nu from \citet{Blankenbach1989} at the steady state.}
-\label{fig:mantle}
-\end{figure}
-
-\begin{table}[H]
-\caption{Comparison between $v_{\textrm{rms}}$ and Nu predicted by the code for the mantle convection experiment and same values as reported in literature.}
-\centering
-\small
-\begin{tabular}{|>{\centering}m{0.09\textwidth}|>{\centering}m{0.05\textwidth}|>{\centering}m{0.24\textwidth}|>{\centering}m{0.15\textwidth}|
-  >{\centering}m{0.1\textwidth} >{\centering}m{0.1\textwidth} >{\centering\arraybackslash}m{0.1\textwidth}|}
-\toprule
-& & \multirow{2}{*}{\citet{Blankenbach1989}} & \citet{Thieulot2014} & \multicolumn{3}{c|}{FALCON}  \\
- & & & \num{200x200} & \num{32x32} & \num{64x64} & \num{128x128}  \\
-\midrule
-\multirow{2}{*}{$Ra=10^4$} & $v_{\textrm{rms}}$ & $42.864947 \pm 0.000020$ & 42.867 & 42.83226 & 42.852793 & 42.861394  \\
-           & Nu      & $4.884409 \pm 0.000010$  & 4.882  & 4.781297 & 4.857475  & 4.877573   \\
-\hline
-\multirow{2}{*}{$Ra=10^5$} & $v_{\textrm{rms}}$ & $193.21454 \pm 0.00010$  & 193.255& 193.872643 & 193.377472 & 193.252290  \\
-           & Nu      & $10.534095 \pm 0.000010$ & 10.507 & 9.602514 & 10.270735  & 10.465629   \\
-\hline
-\multirow{2}{*}{$Ra=10^6$} & $v_{\textrm{rms}}$ & $833.98977 \pm 0.00020$  & 834.712& 848.091176 & 837.767911 & 834.945793  \\
-           & Nu      & $21.972465 \pm 0.000020$ & 21.695 & 15.999266 & 19.703682  & 21.306939   \\
-\bottomrule
-\end{tabular}
-\label{tab:mantle}
-\end{table}
-
-\subsection{Viscoplastic mantle convection}\label{sec:visco_mantle}
-This benchmark is performed as Cases 1-5a described in \citet{Tosi2015} in a 2D unit square domain with two different grid resolutions of \num{32x32} 
-and \num{100x100} elements. Reference viscosity, reference density, heat capacity, thermal conductivity and thermal expansion coefficient are fixed to a constant
-value of 1 ($\eta_0= \rho_0= C_p = k=\alpha=1$), while the gravity has been chosen as $g_y=10^2$, in order to obtain a Rayleigh number $Ra=10^2$. Temperatures
-are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
-\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\].
-The viscosity ﬁeld $\eta$ is calculated as the harmonic average between a linear part $\eta_{lin}$ that depends on temperature only or on temperature and 
-depth $y$ and a nonlinear-plastic part $\eta_{pl}$ that depends on the strain rate $\dot{\bm{\epsilon}}$, as follows
-\[\eta(T,y,\dot{\bm{\epsilon}})=2\left(\frac{1}{\eta_{lin}(T,y)}+\frac{1}{\eta_{pl}(\dot{\bm{\epsilon}})}\right)^{-1}\]
-The linear and the nonlinear-plastic parts of the viscosity are calculated following \citet{Tosi2015} as follows
-\[\eta_{lin}(T,y)=\exp(-\ln(\Delta\eta_T)T+\ln(\Delta\eta_y)y)\]
-\[\eta_{pl}(\dot{\bm{\epsilon}})=\eta^*+\frac{\sigma_Y}{\sqrt{\dot{\bm{\epsilon}}:\dot{\bm{\epsilon}}}}\]
-where $\Delta\eta_T$, $\Delta\eta_y$, $\eta^*$ and $\sigma_Y$ are parameters chosen as listed in Table \ref{tab:tosi_setup}.
-
-For each case, some diagnostic quantities measured at the steady state are reported in Table \ref{tab:visco_mantle} (see Section \ref{sec:error} for 
-the computation of all the quantities), showing a well fit with results described in \citet{Tosi2015}. Data and charts for all cases can be found at 
-\url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Viscoplastic_mantle_convection}.
-
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{white!100!}{gray!25!}
-%\begin{table}[H]
-%  \caption{Results of the viscoplastic mantle convection benchmark for Cases 1-4 compared with results obtained with ELEFANT and YACC (from \citealp{Tosi2015}).}
-%  \centering
-%  \resizebox{\textwidth}{!}{
-%\begin{tabular}{|c|ccc|ccc|ccc|ccc|}
-%  \toprule
-%    & \multicolumn{3}{c|}{\textbf{Case 1}} & \multicolumn{3}{c|}{\textbf{Case 2}} & \multicolumn{3}{c|}{\textbf{Case 3}} &
-%    \multicolumn{3}{c|}{\textbf{Case 4}} \\
-%   & FALCON & ELEFANT& YACC & FALCON& ELEFANT& YACC & FALCON & ELEFANT & YACC & FALCON & ELEFANT & YACC \\
-%  \midrule
-%   $\left\langle T \right\rangle$ & 0.7759 & 0.7758 & 0.7767 & 0.6043 & 0.6036 & 0.6058 & 0.7275 & 0.7275 & 0.7286 & & 0.5277 & 0.5289\\
-%    Nu$^{top}$ & 3.4280 & 3.4214 & 3.4298 & 8.5242 & 8.5115 & 8.5278 & 3.0368 & 3.0347 & 3.0374 && 6.5912 & 6.5572 \\
-%    Nu$^{bottom}$ & 3.3168 & 3.313 & 3.3143 & 8.4599 & 8.4465 & 8.3990 & 3.0086 & 2.9908 & 2.9628 && 6.5834 & 6.5243 \\
-%    $v_{\textrm{rms}}$ & 249.901 & 249.134 & 251.7997 & 141.2451 & 140.7087 & 142.2020 & 100.2776 & 100.1208 & 100.9467 && 79.1105 & 79.6202 \\
-%    $v_{\textrm{rms}}^{top}$ & 1.8907 & 1.8642 & 1.8298 & 104.7928 & 104.1228 & 105.1767 & 2.0832 & 2.0652 & 2.0374 && 74.7596 & 75.8414 \\
-%    $u_{max}^{top}$ & 2.6340 & 2.6119 & 2.5516 & 121.9212& 121.766 & 122.3238 & 2.9086 & 2.9019 & 2.8458 && 88.9146 & 89.2940 \\
-%    $\eta_{min}$ & - & - & - & $2.0303\times10^{-5}$ & $2.0600\times10^{-5}$ & $1.9853\times10^{-5}$ & $4.8129\times10^{-5}$ & $4.8080\times10^{-5}$ & $4.7907\times10^{-5}$ &
-%    & $1.9860\times10^{-4}$ & $1.9174\times10^{-4}$ \\
-%    $\eta_{max}$ & - & - & - & 1.3325 & 1.85 & 1.6095 & 0.9360 & 0.9023 & 1 && 1.52 & 1.6773 \\
-%    $\left\langle W \right\rangle$ & 2.4299 & 2.4316 & 2.4583 & 7.5648 & 7.5665 & 7.6222 & 2.0387 & 2.0384 & 2.0400 && 5.6216 & 5.6512 \\
-%    $\left\langle \Phi \right\rangle/Ra$ & 2.4311 & 2.4276 & 2.4333 & 7.5661 & 7.5585 & 7.6194 & 2.0386 & 2.0370 & 2.0335 && 5.6182 & 5.6463 \\
-%    $\delta$ & 0.05\% & 0.16\% & 1.02\% & 0.02\% & 0.11\% & 0.04\% & $<$0.01\% & 0.07\% & 0.32\% && 0.06\% & 0.09\% \\
-%  \bottomrule
-%  \end{tabular}}
-%  \label{tab:visco_mantle}
-%  \end{table}
-%\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{2}{white!100!}{gray!25!}
-\begin{table}[H]
-  \caption{Benchmark cases and corresponding parameters.}
-  \centering
-  \small
-\begin{tabular}{|>{\centering}m{0.08\textwidth}|>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}>{\centering}m{0.05\textwidth}
-  >{\centering}m{0.05\textwidth} >{\centering}m{0.05\textwidth} |>{\centering\arraybackslash}m{0.2\textwidth}|}
-  \toprule
-    Case & $Ra$ & $\Delta\eta_T$ & $\Delta\eta_y$ & $\eta^*$ & $\sigma_Y$ & Convective regime \\
-  \midrule
-  \rowcolor{gray!25!}
-    1 & $10^2$ & $10^5$ & 1 & - & - & Stagnant lid \\
-    2 & $10^2$ & $10^5$ & 1 & $10^{-3}$ & 1 & Mobile lid  \\
-    \rowcolor{gray!25!}
-    3 & $10^2$ & $10^5$ & 10 & - & - & Stagnant lid  \\
-    4 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 1 & Mobile lid \\
-    \rowcolor{gray!25!}
-    5 & $10^2$ & $10^5$ & 10 & $10^{-3}$ & 4 & Periodic \\
-  \bottomrule
-  \end{tabular}
-  \label{tab:tosi_setup}
-  \end{table}
-
-\renewcommand{\arraystretch}{1.5}
-%{\rowcolors{3}{gray!25!}{white!100!}
-\begin{table}
-  \caption{Results of the viscoplastic mantle convection benchmark for different grid resolutions in Cases 1-5 compared with results obtained with ELEFANT
-  and YACC ($100\times100$ elements, from \citealp{Tosi2015}) and in Stone 28 ($32\times32$ elements, from
-  \url{https://github.com/cedrict/fieldstone/tree/master/python_codes/fieldstone_28}).}
-  \centering
-  \resizebox{\textwidth}{!}{
-\begin{tabular}{|cc|ccccccccccc|}
-  \toprule
-  \multicolumn{13}{|c|}{\textbf{Case 1}} \\
-  \midrule
-  Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-  & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-  \hline
-  \rowcolor{gray!25!}
-    YACC & $100\times100$ & 0.7767 & 3.4298 & 3.3143 & 251.7997 & 1.8298 & 2.5516 & - & - & 2.4583 & 2.4333 & 1.02\% \\
-    ELEFANT & $100\times100$ & 0.7758 & 3.4214 & 3.313 & 249.134 & 1.8642 & 2.6119 & - & - & 2.4316 & 2.4276 & 0.16\% \\
-    \rowcolor{gray!25!}
-    Stone 28 & $32\times32$ & 0.7737 & 3.3987 & - & 243.872 & - & - & - & - & - & - & - \\
-%    FALCON & $32\times32$ & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & - & - & 2.4745 & 2.4864 & 0.48\% \\
-%    \rowcolor{gray!25!}
-    FALCON & $100\times100$ & 0.7759 & 3.4280 & 3.3168 & 249.901 & 1.8907 & 2.6340 & - & - & 2.4299 & 2.4311 & 0.05\% \\
-    \midrule
-    \multicolumn{13}{|c|}{\textbf{Case 2}} \\
-    \midrule
-    Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-    & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-    \hline
-    \rowcolor{gray!25!}
-     YACC & $100\times100$ & 0.6058 & 8.5278 & 8.3990 & 142.2020 & 105.1767 & 122.3238 & $1.9853\times10^{-5}$ & 1.6095 & 7.6222 & 7.6194 & 0.04\% \\
-     ELEFANT & $100\times100$ & 0.6034 & 8.5115 & 8.4465 & 140.8067 & 104.1228 & 121.766 & $2.0600\times10^{-5}$ & 1.85 & 7.5665 & 7.5585 & 0.11\% \\
-     \rowcolor{gray!25!}
-     Stone 28 & $32\times32$ & 0.6052 & 8.1715 & - & 141.518 & - & - & - & - & - & - & - \\
-%     FALCON & $32\times32$ & 0.6454 & 8.1422 & 7.4020 & 156.6908 & 95.7226 & 116.9302 & $2.2572\times10^{-5}$ & 0.7194 & 7.4289 & 7.4452 & 0.22\% \\
-%     \rowcolor{gray!25!}
-     FALCON & $100\times100$ & 0.6044 & 8.5266 & 8.4590 & 141.3684 & 104.8867 & 122.0470 & $2.0333\times10^{-5}$ & 1.2914 & 7.5707 & 7.5718 & 0.01\% \\
-     \midrule
-     \multicolumn{13}{|c|}{\textbf{Case 3}} \\
-     \midrule
-     Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-     & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-     \hline
-     \rowcolor{gray!25!} 
-     YACC & $100\times100$ & 0.7286 & 3.0374 & 2.9628 & 100.9467 & 2.0374 & 2.8458 & $4.7907\times10^{-5}$ & 1 & 2.0400 & 2.0335 & 0.32\% \\
-      ELEFANT & $100\times100$ & 0.7275 & 3.0347 & 2.9908 & 100.1208 & 2.0652 & 2.9019 & $4.8080\times10^{-5}$ & 0.9023 & 2.0384 & 2.037 & 0.07\% \\
-      \rowcolor{gray!25!}
-      Stone 28 & $32\times32$ & 0.7277 & 3.0347 & - & 100.018 & - & - & - & - & - & - & - \\
-%      FALCON & $32\times32$ & 0.7281 & 3.0618 & 2.8132 & 101.0092 & 2.1630 & 3.0178 & $4.8500\times10^{-5}$ & 0.7845 & 2.0699 & 2.0695 & 0.02\% \\
-%      \rowcolor{gray!25!}
-      FALCON & $100\times100$ & 0.7275 & 3.0368 & 3.0086 & 100.2776 & 2.0832 & 2.9086 & $4.8129\times10^{-5}$ & 0.9360 & 2.0387 & 2.0386 & $<$0.01\% \\
-      \midrule
-      \multicolumn{13}{|c|}{\textbf{Case 4}} \\
-      \midrule
-      Code & Resolution & $\left\langle T \right\rangle$ & Nu$^{top}$ & Nu$^{bottom}$ & $v_{\textrm{rms}}$ & $v_{\textrm{rms}}^{top}$& $u_{max}^{top}$& $\eta_{min}$
-      & $\eta_{max}$ & $\left\langle W \right\rangle$ & $\left\langle \Phi \right\rangle/Ra$ & $\delta$ \\
-      \hline
-      \rowcolor{gray!25!} 
-      YACC & $100\times100$ & 0.5289 & 6.5572 & 6.5243 & 79.6202 & 75.4814 & 89.2940 & $1.9174\times10^{-4}$ & 1.6773 & 5.6512 & 5.6463 & 0.09\% \\
-       ELEFANT & $100\times100$ & 0.5277 & 6.5912 & 6.5834 & 79.1105 & 74.7596 & 88.9146 & $1.9860\times10^{-4}$ & 1.5200 & 5.6216 & 5.6182 & 0.06\% \\
-       \rowcolor{gray!25!}
-       Stone 28 & $32\times32$ & 0.5286 & 6.4036 & - & 79.4746 & - & - & - & - & - & - & - \\
-%       FALCON & $32\times32$ & 0.5334 & 6.4679 & 6.3435 & 80.6153 & 76.0511 & 90.2058 & $2.0393\times10^{-4}$ & 0.6453 & 5.6676 & 5.6698 & 0.04\% \\
-%       \rowcolor{gray!25!}
-       FALCON & $100\times100$ & 0.5279 & 6.5897 & 6.5830 & 79.1220 & 75.0715 & 88.8484 & $1.9693\times10^{-4}$ & 1.4133 & 5.6127 & 5.6113 & 0.03\% \\
-       \midrule
-       \multicolumn{13}{|c|}{\textbf{Case 5}} \\
-       \midrule
-       Code & Resolution & $P$ & $\left\langle T \right\rangle_{min}$ & $\left\langle T \right\rangle_{max}$ & Nu$^{top}_{min}$ & Nu$^{top}_{max}$ &
-       $v_{\textrm{rms}}^{min}$ & $v_{\textrm{rms}}^{max}$ & \multicolumn{2}{c}{$(\left\langle \Phi \right\rangle/Ra)_{min}$} & 
-       \multicolumn{2}{c|}{$(\left\langle \Phi \right\rangle/Ra)_{max}$} \\
-       \hline
-       \rowcolor{gray!25!}
-         YACC & $100\times100$ & 0.0792 & 0.6543 & 0.6722 & 2.6950 & 7.2792 & 41.9813 & 98.8230 & \multicolumn{2}{c}{1.4347} & 
-         \multicolumn{2}{c|}{9.2748} \\
-         ELEFANT & $100\times100$ & 0.076 & 0.6514 & 0.6693 & 2.6808 & 7.3506 & 41.2578 & 100.089 & \multicolumn{2}{c}{1.3324} & 
-         \multicolumn{2}{c|}{9.5289} \\
-         \rowcolor{gray!25!}
-         Stone 28 & $32\times32$ & 0.0764 & 0.6521 & 0.6697 & 2.6912 & 7.0763 & 41.61 & 98.183 & \multicolumn{2}{c}{-} & 
-         \multicolumn{2}{c|}{-} \\
-     %    FALCON & $32\times32$ & & 0.7750 & 3.4381 & 2.7363 & 248.3337 & 2.0022 & 2.7813 & \multicolumn{2}{c}{-} & 
-     %     \multicolumn{2}{c|}{-} \\
-     %    \rowcolor{gray!25!}
-         FALCON & $100\times100$ & 0.0775 & 0.6519 & 0.6697 & 2.6847 & 7.3629 & 41.5256 & 99.5466 & \multicolumn{2}{c}{1.3399} & 
-         \multicolumn{2}{c|}{9.3993} \\
-  \bottomrule
-  \end{tabular}}
-  \label{tab:visco_mantle}
-  \end{table}
-
-\subsection{Thin layer entrainment}\label{sec:thin}
-This experiment is performed as originally proposed by \citet{vanKeken1997} with two incompressible fluids in a rectangular domain with $L_x=2$ and $L_y=1$
-and gravity acceleration $g_y=10^{10}Ra$. The Rayleigh number and the compositional Rayleigh number are fixed to $Ra=300000$ and $Rc=450000$, respectively.
-Both fluids have constant viscosity, thermal conductivity, specific heat ($\eta=\rho=C_p=1$) and thermal expansion coefficient ($\alpha=10^{-10}$). Fluid 1
-has a density $\rho_1=1$, while fluid 2 is denser ($\rho_2=\rho_1+1.5 \cdot 10^{-10}$) and is placed at the bottom of the domain, for $y \leq 0.025$.
-Temperature are set to 0 on top and 1 on bottom of the domain. Velocity boundary conditions are set to free slip on all sides of the domain. The initial
-temperature field is given by
-\[T(x,y)=T_u(x,y)+T_l(x,y)+T_r(x,y)+T_s(x,y)-\frac{3}{2}\]
-with
-\begin{eqnarray}
-T_u(x,y)&=&\frac{1}{2}erf\left(\frac{1-y}{2}\sqrt{\frac{u_0}{x}}\right)\nonumber \\
-T_l(x,y)&=&1-\frac{1}{2}erf\left(\frac{y}{2}\sqrt{\frac{u_0}{L_x-x}}\right)\nonumber \\
-T_r(x,y)&=&\frac{1}{2}+\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{y+1}} \exp\left(-\frac{x^2u_0}{4y+4}\right)\nonumber \\
-T_s(x,y)&=&\frac{1}{2}-\frac{Q}{2\sqrt{\pi}}\sqrt{\frac{u_0}{2-y}} \exp\left(-\frac{(L_x-x)^2u_0}{8-4y}\right)\nonumber
-\end{eqnarray}
-and
-\begin{eqnarray}
-u_0&=&\frac{L_x^{7/3}}{(1+L_x^4)^{2/3}}\left(\frac{Ra}{2\sqrt{\pi}}\right)^{\frac{2}{3}}\nonumber \\
-Q&=&2\sqrt{\frac{L_x}{\pi u_0}}\nonumber
-\end{eqnarray}
-
-The experiment is performed with two grid resolution (\num{125x40} and \num{200x80} elements), with 100 markers per element and CFL of 0.25. $v_{\textrm{rms}}$
-(see Sec. \ref{sec:error}) calculated for both the simulations match well with results from \citet{vanKeken1997} and obtained with ASPECT
-\citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} and ELEFANT \citep{Thieulot2014}, obtained with similar grid resolutions
-(Fig. \ref{fig:thin}). In addition, a variety of simulations with different resolutions and aspect ratios are performed to verify that the sensitivity of the
-initial velocity field. Results match well with results from \citet{vanKeken1997} (in the grey area) and \citet{Thieulot2014} (Fig. \ref{fig:thin_initial}). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum%2BEnergy/Thin_layer}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrms_thin.pdf}
-\caption{$v_{\textrm{rms}}$ for the thin layer experiment as function of time for different grid resolution. Results are compared with results obtained by
-\citet{vanKeken1997} (black lines) and with ASPECT and ELEFANT.}
-\label{fig:thin}
-\end{figure}
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/vrmszero.pdf}
-\caption{$v_{\textrm{rms}}$ at $t=0$ for the thin layer experiment as function of the element size for different aspect ratios of the elements. Results are compared
-with results obtained by \citet{vanKeken1997} (black dots in the grey area) and with ASPECT and ELEFANT.}
-\label{fig:thin_initial}
-\end{figure}
-
-\subsection{Sinking hydrated cylinder}\label{sec:quinquis}
-This experiment is performed as explained in \citet{Quinquis2014} and simulates the sinking of a cold, hydrated cylinder into a warm, dry mantle. The domain
-is square with $L_x=L_y=$ \SI{300}{\km}, a grid resolution of \num{300x300} elements and 25 markers per element. The sinking cylinder is located at $x=$
-\SI{150}{\km} and $y=$ \SI{170}{\km} with a radius of \SI{20}{\km} and has $\rho_c=$ \SI{3250}{\kg\per\cubic\m}, $\eta_c=$ \SI{e23}{\pascal\s}, a thermal
-conductivity $k_c=$ \SI{4.5}{\watt\per\m\per\kelvin},a specific heat $C_{p_c}=$ \SI{1250}{\joule\per\kg\per\kelvin}; the surrounding mantle has $\rho_m=$
-\SI{3200}{\kg\per\cubic\m}, $\eta_m=$ \SI{e20}{\pascal\s}, a thermal conductivity $k_m=$ \SI{105}{\watt\per\m\per\kelvin} and specific heat $C_{p_m}=$
-\SI{1250}{\joule\per\kg\per\kelvin}; a 58 km-thick lithosphere overlies the mantle and has $\rho_l=$ \SI{3200}{\kg\per\cubic\m}, $\eta_l=$ \SI{e23}{\pascal\s},
-a thermal conductivity $k_l=$ \SI{4.5}{\watt\per\m\per\kelvin}, specific heat $C_{p_l}=$ \SI{750}{\joule\per\kg\per\kelvin} and a thermal diffusivity of
-\SI{e-6}{\square\m\per\s}. Velocity boundary conditions are set to free slip on all sides of the domain. Temperature increase linearly in the lithosphere,
-from 0° to 1300°C, and in the mantle, from 1300° to 1360.5°C. The initial temperature of the cylinder is fixed to 400°C. Throughout the evolution, temperature
-is fixed to 0° and 1360.5°C at the top and bottom of the domain, respectively. The initial bound water content is imposed at 2 wt.\% in the cylinder and at
-0 wt.\% in both the mantle and the lithosphere. The maximum amount of water in the mantle is fixed to 0.2 wt.\%, while maximum water content in the cylinder
-and in the lithosphere is function of pressure and temperature and is calculated using a serpentinized harzburgite with Perple\_X, as in \citet{Quinquis2014}.
-
-As shown in \citet{Quinquis2014}, the model is characterised by a progressive dehydration in the external portion of the cylinder, due to the increase of
-temperature, with a consequent vertical migration of free water that hydrates the mantle above the cylinder, up to the lithosphere. The distribution of
-bound and free water during the evolution of the experiment is shown in Fig. \ref{fig:quinquis} (left and right columns, respectively). 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Hydration}.
-
-\begin{figure}
-\centering
-\noindent\includegraphics[width=400px]{./Figures/Hydration.pdf}
-\caption{Distribution of bound and free water (left and right columns, respectively) at $t=0.5$ and $t=$ \SI{1.25}{\mega\year} (first and second rows,
-respectively) for the sinking hydrated cylinder experiment. Only markers with bound or free water not equal to 0 are plotted.}
-\label{fig:quinquis}
-\end{figure}
-
-\subsection{Mantle melting curves}\label{sec:katz}
-This instantaneous experiment is performed in order to recreate isobaric melting curves, as function of temperature, bulk water and content of water in the
-melt, to compare with melting curves obtained by \citet{Katz2003}. The domain is square with $L_x=L_y=100$ km, a grid resolution of \num{100x100} elements and
-25 markers per element. Density is set in the entire domain to \SI{3300}{\kg\per\cubic\m} and temperature increases from the sides to the centre from 1000° to
-1500°C. The experiment is performed for bulk water contents from 0 to 0.3 wt.\%.
-
-Fig. \ref{fig:fraction} shows isobaric melting curves for \SI{1}{\giga\pascal} (panel a) and \SI{3}{\giga\pascal} (panel b) calculated for different bulk water
-contents. The curves well match with those obtained for the mantle by \citet{Katz2003}, showing the saturation at the solidus for bulk water of 0.2 and 0.3 wt.\%, indicating
-that the code correctly determines melt fraction as function of pressure, temperature and bulk water content. Same results can be observed in Fig.
-\ref{fig:water}a, where isobaric and isothermal melting curves are shown as function of bulk water content. In addition, Fig. \ref{fig:water}b shows that the
-code correctly determines also the percentage of water in the melt as function of pressure, temperature and melt fraction. 
-All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Melting}.
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Fraction.pdf}
-\caption{Isobaric melting curves for 1 GPa (panel a) and 3 GPa (panel b) as function of temperature with different bulk water contents.}
-\label{fig:fraction}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{./Figures/Water.pdf}
-\caption{Isobaric melting curves for 1.5 GPa as function of bulk water (panel a) and water in the melt (panel b) calculated for different temperatures.}
-\label{fig:water}
-\end{figure}
-
-\clearpage
-
-\bibliography{export}
-
-\end{document}
+where $f_v$ is the viscosity reduction factor between dry and wet conditions, $[OH^-]$ is the water content and $[OH^-]_0$ is a reference water content set to 620 ppm (0.062 wt.\%) \citep{Arcay2005}.
+By default, $f_v$ is set to 100, in according to \citet{Arcay2005} and \citet{Horiuchi2016}.

### 2022-09-07 tex/intro.tex
table

@@
   \multirow{2}{*}{Markers advection} & Zalesak disk & \ref{sec:runge} \\
-    & Conservative Interpolation Velocity & \ref{sec:CVI} \\
+    & Conservative Velocity Interpolation & \ref{sec:CVI} \\
 \hline
@@
 \hline
-   Non-linear  & Slab detachment & \ref{sec:slab} \\
-   visco-plasticity & Indenter & \ref{sec:indenter} \\
+    \multirow{3}{*}{Nonlinear rheology} & Slab detachment & \ref{sec:slab} \\
+    & Indenter & \ref{sec:indenter} \\
     & Brick & \ref{sec:brick} \\

### 2022-09-07 tex/Benchmark/stokes.tex
stokes

@@
 \includegraphics[width=\linewidth]{./Figures/MUMPS.pdf}
-\caption{Panel a) Analysis, factorisation, solution and total solve times (panel a) and factorisation memory usage (panel b) as a function of the total number
+\caption{Analysis, factorisation, solution and total solve times (panel a) and factorisation memory usage (panel b) as a function of the total number
 of degrees of freedom for the Stokes flow experiment.}

### 2022-09-09 tex/Benchmark/runge.tex
zalesak

@@
 \caption{Distance from the centre as function of time for values of Courant number of  of 0.25, 0.3, 0.5, 0.75 and 1 (green, orange, blue, black and red,
-respectively). R indicates the distance from the centre considering a perfect circle trajectory.}
+respectively).}
 \label{fig:runge_err}

### 2022-09-09 tex/Benchmark/ist_sphere.tex
inst sphere

@@
 All three cases are performed for different grid resolutions, between $16 \times16$ and $512\times 512$ elements with 50 randomly distributed markers per element,
-and different average schemes for the viscosity (harmonic, geometric and arithmetic). The velocity in the centre of the sphere $\bm{v}(0.5;0.5)$), minimum and
+and different average schemes for the viscosity (harmonic, geometric and arithmetic). The velocity in the centre of the sphere $\bm{v}(0.5;0.5)$, minimum and
 maximum velocities ($u_{min}$, $u_{max}$, $v_{min}$, $v_{max}$) and pressures ($p_{min}$, $p_{max}$) on the entire domain, $v_{\textrm{rms}}$ and average pressure on the

### 2022-09-16 tex/Benchmark/block.tex
zenodo

@@
 \subsection{Falling blocks}\label{sec:block}
-This benchmark is described as proposed by \citet{Gerya2003a}, \citet{Gerya2010} and \citet{Thieulot2011}. The domain is square with $L_x=L_y=$ \SI{500}{\km}
-and the grid is composed by $50\times50$ elements with 25 markers in each element. The block is initially centred at ($x=$ \SI{250}{\km}; $y=$ \SI{400}{\km})
-and has a size of $100\times100$ km (Fig. \ref{fig:block}a). The fluid surrounding the block has $\eta_f=$ \SI{e21}{\pascal\per\s} and $\rho_f=$
-\SI{3200}{\kg\per\cubic\m}. The benchmark tests with different viscosities of the block, with $\eta_b$ from \SIrange{e15}{e27}{\pascal\per\s}. In each
-experiment also the density $\rho_b$ of the block varies from \SIrange{3220}{9900}{\kg\per\cubic\m}. Velocity boundary conditions are set to free slip
-conditions on all sides of the domain. Density distributions at $t=$ \SI{20}{\mega\year} for $\rho_b=$ \SI{3300}{\kg\per\cubic\m} and different $\eta_b$
-are plotted in Fig. \ref{fig:block}b-f, showing that the code correctly preserve the block geometry in case of large viscosity contrast, with a block stiffer
-than the surrounding fluid (Fig. \ref{fig:block}f).
 
+\begin{wrapfigure}{r}{8cm}
+\includegraphics[width=8cm]{./Figures/Block.pdf}
+\caption{Density evolution of the falling block experiment at $t=$ \SI{0}{\mega\year} (panel a) and $t=$ \SI{20}{\mega\year} for different viscosities of the
+block (panels b-f).}
+\label{fig:block}
+\end{wrapfigure}
+This benchmark is described as proposed by \citet{Gerya2003a}, \citet{Gerya2010} and \citet{Thieulot2011}.
+The domain is square with $L_x=L_y=$ \SI{500}{\km} and the grid is composed by $50\times50$ elements with 25 markers in each element.
+The block is initially centred at ($x=$ \SI{250}{\km}; $y=$ \SI{400}{\km}) and has a size of $100\times100$ km (Fig.
+\ref{fig:block}a).
+The fluid surrounding the block has $\eta_f=$ \SI{e21}{\pascal\per\s} and $\rho_f=$ \SI{3200}{\kg\per\cubic\m}.
+The benchmark tests with different viscosities of the block, with $\eta_b$ from \SIrange{e15}{e27}{\pascal\per\s}.
+In each experiment also the density $\rho_b$ of the block varies from \SIrange{3220}{9900}{\kg\per\cubic\m}.
+Velocity boundary conditions are set to free slip conditions on all sides of the domain.
+Density distributions at $t=$ \SI{20}{\mega\year} for $\rho_b=$ \SI{3300}{\kg\per\cubic\m} and different $\eta_b$ are plotted in Fig.
+\ref{fig:block}b-f, showing that the code correctly preserve the block geometry in case of large viscosity contrast, with a block stiffer than the surrounding fluid (Fig.
+\ref{fig:block}f).
 The velocity in the centre of the falling block at $t=0$ is measured for all experiments and, since the velocity should increase with density contrast, the
@@
 found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Momentum_equation/Falling%20blocks}.
-
-\begin{figure}
-\includegraphics[width=300px]{./Figures/Block.pdf}
+\begin{figure}[h!]
 \centering
-\caption{Density evolution of the falling block experiment at $t=$ \SI{0}{\mega\year} (panel a) and $t=$ \SI{20}{\mega\year} for different viscosities of the
-block (panels b-f).}
-\label{fig:block}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=400px]{./Figures/Falling.pdf}
+\includegraphics[width=13cm]{./Figures/Falling.pdf}
 \caption{Initial velocity relative to the density contrast at the centre of the falling block as function of the viscosity contrast between the block and the

### 2022-10-14 tex/Benchmark/advection.tex
zenodo
energy

@@
 
-Temperature profiles at the end of the simulation are shown in \ref{fig:advection} as function of the dimensionless coefficient $\gamma=\tau \bm{v}/h$. In case
-of the classic Galerkin method ($\gamma=0$, blue line in \ref{fig:advection}) the final thermal profile is characterised by strong oscillations, which are
-eliminated in case of the SUPG method ($\gamma=0.045$, orange line in \ref{fig:advection}). 
+Temperature profiles at the end of the simulation are shown in Fig. \ref{fig:advection} as function of the dimensionless coefficient $\gamma=\tau \bm{v}/h$. In case
+of the classic Galerkin method ($\gamma=0$, blue line in Fig. \ref{fig:advection}) the final thermal profile is characterised by strong oscillations, which are
+eliminated in case of the SUPG method ($\gamma=0.045$, orange line in Fig. \ref{fig:advection}).
 All data can be found at \url{https://github.com/aleregorda/Benchmarks/tree/main/Energy_equation/Advection_stabilisation}.
 
-\begin{figure}
+\begin{figure}[h!]
 \centering
-\includegraphics[width=400px]{./Figures/Advection.pdf}
+\includegraphics[width=9.5cm]{./Figures/Advection.pdf}
 \caption{Temperature profile as function of \textit{x} for the advection stabilisation benchmark. Purple line indicates the initial temperature profile; the
